---
title: "The Analytic Layer for the MD/Harm Pathway: Verification and Prediction"
subtitle: "Certified Identities, the Approximation Ladder, and Semi-Analytic Operating Characteristics (Continuous Outcomes)"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>")
```

# Scope and provenance

This document consolidates the analytic materials for the continuous/MD harm
pathway into a single reference. It combines, and is intended to supersede as
the working reference, two committed documents in this directory:
`analytic_verification_math_details.qmd` (the verification tiers: what closed
forms certify exactly) and
`analytic_derivations_mr_operating_characteristics.qmd` (the prediction
layer: which MR operating characteristics are derivable before a grid runs).
Both parents are left unchanged in the repository. Beyond consolidation, this
document adds two things: the connective material between verification and
prediction — the explicit ladder from exact identities to each approximation
actually carried out, including the population-level covariance decomposition
previously deferred and now derived and verified — and a detailed worked
example (@sec-worked) that carries the machinery end to end against a
specific committed simulation design, stating for every predicted quantity
what is computed, from which inputs, and which column of the design's own
report it should be compared against.

Supporting artifacts: `dev/detection-theory-continuous/`
(`NOTE_declaration_gaussian_md.md`; `check_declaration_gaussian_md.R`,
on-record 16 PASS / 0 FAIL, cross-stack identical output on
R 4.6.1/mvtnorm 1.4.2 and R 4.3.3/mvtnorm 1.2.4; and
`check_additive_covariance_md.R`, the additive-decomposition verification
cited in @sec-ladder).

Estimand language is strict throughout. $\beta(\hat H)$ is the conditional
estimand — the exact treatment effect at the realized, data-selected rule,
computed on the fixed super-population with zero Monte Carlo error.
$\theta^\dagger(H)$ is the marginal true-subgroup effect at the designed
region $Q$. MR conditions on the realized candidate family, targets
$\beta(\hat H)$, and is scored against $\beta(\hat H)$, never
$\theta^\dagger$. Effects are reported on the oriented scale
(positive = harm) under `adverse_outcome = FALSE`; exact references are
stored on the raw scale and the orientation bridge applies the sign map once.

# Fixture, estimator, and event conventions

The committed harm-pathway fixture (`df_super`, $N = 5000$;
`actg175_continuous_simulations.qmd` lines 65, 100, 118–119): designed
region $Q$ with $|Q| = 1723$, prevalence $\pi = 0.3446$; raw-scale effects
$\tau_Q = -40$ (calibration target), $\tau_{Q^c} = -26.255235876$,
$\beta_{\mathrm{int}} = \tau_Q - \tau_{Q^c} = -13.744764124$; oriented
values $+40$, $+26.255235876$, $+13.744764124$. The CATE is two-valued.
Screening floor $c_1 = 30$ and consistency floor $c_2 = 10$, oriented.

Three conventions, verified against package source, govern everything below:
the estimator is the within-subgroup OLS treatment coefficient of
`lm(y ~ treat)` — the difference in arm means — on the oriented scale
(`R/glm_effect_estimators.R:814-821`); the screening and consistency events
compare **point estimates** to the floors
(`R/subgroup_search.R:626`; `R/subgroup_consistency_main.R:528-531`), so
estimated variance never enters an event definition; and thresholds are
one-sided floors on the oriented scale, never direction switches
(`dev/glm-continuous-sims/NOTE_threshold_sign_md.md`).

One translation is used repeatedly and is worth stating once in words. By
the mixture identity of Part I, a region's true oriented effect exceeds the
screening floor exactly when its purity does:
$$
\beta(g) \ge c_1
\quad\Longleftrightarrow\quad
P(Q \mid g) \;\ge\; \frac{c_1 - \tau_{Q^c}^{\mathrm{or}}}{\beta_{\mathrm{int}}^{\mathrm{or}}}
= \frac{30 - 26.2552}{13.7448} = 0.2724.
$$
The floor $c_1 = 30$ is therefore a **27.2% purity requirement** — and since
the marginal prevalence of $Q$ is 34.5%, even a rule carrying no information
about $Q$ has a true effect above the floor. This single fact organizes much
of what the simulations observe on this pathway, and the worked example
returns to it.

# Part I — The certified layer

## Tier A: algebraic closed forms (exactly verifiable)

**Proposition 1 (mixture identity; the content of CHECK A).** For every
$H \subseteq \mathcal{S}$ with $|H| > 0$,

$$
\beta(H)
\;=\; \frac{|H \cap Q|\,\tau_Q + |H \cap Q^c|\,\tau_{Q^c}}{|H|}
\;=\; \tau_{Q^c} + \beta_{\mathrm{int}}\, P(Q \mid H).
$$

*Proof.* Identity link and randomization give collapsibility —
$\beta(H)$ is the average CATE over $H$ — and a two-valued $\tau$ makes
that average a two-point mixture determined by one overlap count. $\square$

The force of Proposition 1 is its quantifier: it holds for **every** rule,
including every data-selected $\hat H$, so the scoring target behind every
bias and coverage figure reduces to counting on a fixed finite set —
machine-exact. In the committed runs, CHECK A recomputed $\beta(\hat H)$
this way for each realized rule and held at $1.421\times 10^{-14}$ across
601–912 distinct rules per cell.

**Corollaries.** (i) $\beta(H^c)$ by the same mixture on the complement.
(ii) The ITT chain: $\beta(\mathcal{S}) = \tau_{Q^c} +
\beta_{\mathrm{int}}\pi = -30.991681593$ raw, which the bundles reproduce
as the complement effect on undetected rows. (iii) Disjointness:
$H \cap Q = \varnothing \Rightarrow \beta(H) = \tau_{Q^c}$ — observed
verbatim when a replicate selected a region disjoint from $Q$ and scored
exactly $26.255$ oriented. (iv) Calibration (CHECK B):
$\theta^\dagger(Q) = \beta(Q) = \tau_Q$ identically. (v) Classification
identities: with $s = |\hat H \cap Q|$, $\mathrm{sens} = s/|Q|$,
$\mathrm{PPV} = s/|\hat H|$, and on the complement
$\mathrm{spec} = |\hat H^c \cap Q^c|/|Q^c|$,
$\mathrm{NPV} = |\hat H^c \cap Q^c|/|\hat H^c|$; hence $\mathrm{PPV} \equiv \mathrm{sens}\cdot
|Q|/|\hat H|$ — the reported metrics carry one degree of freedom given the
sizes, an exact internal consistency relation.

```{r itt-live}
pfile <- "mr_sweep_md_harm/md_harm_s50_pilot/fs_md_harm_n1000_res.rds"
if (file.exists(pfile)) {
  b <- readRDS(pfile); tr <- b$truth; r <- b$results
  pred_itt <- tr$effect_Qc + tr$beta_inter * tr$prevalence_Q
  obs_itt  <- unique(round(r$betaHhat_Hc[r$detected %in% 0L |
                                         is.na(r$detected)], 9))
  stopifnot(length(obs_itt) == 1L, abs(pred_itt - obs_itt) < 1e-8)
  cat(sprintf("ITT chain: tau_Qc + beta_int*pi = %.9f; bundle records %.9f\n",
              pred_itt, obs_itt))
}
```

(The chunk is guarded: the `mr_sweep_md_harm` bundles are summary-era and
machine-local; where absent, the chunk skips silently and the committed
record above stands.)

## Tier B: exact by computation, not by formula

Quantities defined on $\mathcal{S}$ for a realized object are computable
with zero Monte Carlo error by finite summation even when no algebraic form
exists. First, in the binary/OR port the mixture identity fails — for a
link $g$, $g^{-1}$ of a mixture is not the mixture of $g^{-1}$'s — yet the
exact OR over $H$ remains a deterministic function of the DGM's probability
surfaces evaluated on $\mathcal{S}$: Tier A degrades to Tier B, the
reference stays exact (the risk-difference scale retains full Tier A).
Second, the covariance structure of the candidate estimators: conditional
on a realized family and sample, the difference-in-means estimators over
regions $H_j, H_k$ with arm-wise counts $n_{j,a}$ satisfy, under
homoscedastic $\varepsilon$,

$$
\operatorname{Cov}\!\big(\hat\beta_j, \hat\beta_k\big)
\;=\; \sigma^2 \sum_{a\in\{0,1\}}
\frac{\,n_{jk,a}\,}{\,n_{j,a}\, n_{k,a}\,},
\qquad n_{jk,a} = |H_j \cap H_k \cap \{A=a\}|,
$$

an exact overlap computation; under balanced arms the correlation is the
overlap fraction $P(H_j \cap H_k)/\sqrt{P(H_j)P(H_k)}$, free of $\sigma$.
The caveat is stated rather than hidden: family membership and quantile
cuts are data-dependent, so this exactness is conditional on the realized
family.

## Tier C: computational identities

Dual-path equalities (two independent computations of one quantity must
agree exactly — CHECK C), the partition $|H| + |H^c| = N$ on every row, and
bit-level reproducibility (a replicate is a deterministic function of
(seed, RNG kind), so the standard is equality, not tolerance). These verify
the *implementation*, not the estimand, and their exactness makes failures
self-diagnosing: CHECK C's two quantities disagreeing by
$1.98\times 10^{2}$ while summing to $2.7\times 10^{-13}$ is uniquely the
signature of a sign flip — a tolerance-based check could only have said
"off."

# Part II — From identities to approximations: the ladder {#sec-ladder}

The connective structure of the analytic layer is a ladder of six levels.
At each level, exactly one new source of error enters; everything below it
remains exact. This ordering is what converts "Gaussian approximation" from
a label into an audited chain: when a computed rate and a measured rate
disagree, the ladder localizes which rung the disagreement lives on.

## Level 0–1: identities and exact computation

Tier A supplies the mean structure exactly (mixture values for every rule);
Tier B supplies the covariance structure exactly (overlap counts) and exact
references where formulas fail. Error: none.

## Level 2: the design-conditional law is an identity, not an approximation

Fix a candidate family $\mathcal G = \{g_1, \dots, g_M\}$ and a design
$D = \{(X_i, A_i)\}$. The estimator vector is linear in $Y$,
$\hat\beta = C Y$ with $C$ the $M \times n$ coefficient matrix
($c_i(g) = \mathbb 1\{i \in g\}(A_i/n_{g,1} - (1-A_i)/n_{g,0})$). Under
Gaussian noise with known $\sigma^2$,

$$
\hat\beta \mid D \;\sim\; N_M\!\big(m(D),\ \sigma^2 C C^\top\big)
$$

**exactly** — a distributional identity, not a limit theorem. The
conditional mean decomposes linearly in the DGM effect parameters,
$m(D) = a_0 + \tau_{Q^c}\, a_1 + \beta_{\mathrm{int}}\, a_2$ (prognostic
imbalance, treatment, and realized-$P(Q \mid g, \text{arm } 1)$
components), so null, borderline, and alternative scenarios share one
design geometry; and $\sigma^2 C C^\top$ coincides with the Tier B overlap
formula (verified identical at $1.9\times 10^{-15}$). Declaration and
selection events are then exact probabilities of a known law, and the only
error at this level is numerical quadrature: `mvtnorm::pmvnorm`'s
Genz–Bretz quasi–Monte Carlo, with a reported error estimate
($3\times 10^{-8}$ observed at correlation $0.98$) and a stated requirement
that tail rates use `abseps` $\le 10^{-5}$ (the default $10^{-3}$ is 40%+
relative error at $P \approx 0.002$). Exact duplicates make the law
singular; de-duplicate the family first.

## Level 3: CLT and plug-in — where estimation error enters, and where it cannot

Two departures from Level 2, each measured. **Non-Gaussian noise:** the
same MVN law becomes a CLT approximation for the weighted sums $CY$;
accuracy is governed by the smallest arm counts and the noise shape, and
was measured at $0.016$ absolute rate error with minimum arm size 10 and
noise skewness $\approx 6.2$ — the linear estimator's averaging does the
work. **Estimated variance:** because the procedure's events are defined on
point estimates, $\widehat\Sigma$ enters only the *computed probabilities*,
never the events. The single-dataset plug-in (per-candidate `lm` standard
errors with design-known overlap correlations) displaced the family rate by
$0.012$. One structural observation connects this level to the next: the
per-subgroup `lm` residual variance absorbs the within-region prognostic
and CATE spread (measured inflation factors $0.94$–$1.09$ over
$\sigma$-only truth), so the analyst's plug-in is a nearly unbiased
estimate of the *population-level* effective variance of Level 4 rather
than the conditional $\sigma^2$ — plug-in computations approximate the
unconditional law, known-$\sigma$ computations the design-conditional one.
This has a practical corollary used in the worked example: **anchoring the
scale to a fitted standard error already incorporates the Level 4 variance
correction**, leaving only the covariance refinement below.

## Level 4: population averaging — the additive covariance decomposition {#sec-additive}

Averaging over designs, the means converge to the mixture values. The
variances and covariances require care, and this is where a purely
multiplicative repair fails. Write, for subject $i$ in region $g$, the
centered prognostic and CATE contributions

$$
u^{(g)}_i = \mu_0(X_i) - \mathbb E[\mu_0 \mid g], \qquad
t^{(g)}_i = \beta_{\mathrm{int}}\big(\mathbb 1\{X_i \in Q\} - P(Q \mid g)\big),
$$

so that, to leading order, the estimator's deviation from its target is a
sum of independent per-subject influence contributions
$\hat\beta(g) - \beta(g) \approx \sum_i \psi_i(g)$ with

$$
\psi_i(g) = \frac{2}{nP(g)}\,\mathbb 1\{i \in g\}
\Big[A_i\big(u^{(g)}_i + t^{(g)}_i + \varepsilon_i\big)
   - (1 - A_i)\big(u^{(g)}_i + \varepsilon_i\big)\Big].
$$

Taking expectations over $(A, \varepsilon)$ with $A \sim
\mathrm{Bern}(1/2)$ independent of $X$ yields the **additive population
covariance**: for any pair $(g, g')$,

$$
\Sigma_{\mathrm{pop}}(g, g')
= \frac{4}{n}\cdot\frac{P(g \cap g')}{P(g)P(g')}
\left(\sigma^2 \;+\;
\mathbb E\!\left[\, u^{(g)} u^{(g')}
+ \tfrac{1}{2}\big(u^{(g)} t^{(g')} + t^{(g)} u^{(g')} + t^{(g)} t^{(g')}\big)
\;\middle|\; g \cap g'\right]\right),
$$

with $\Sigma_{\mathrm{pop}}(g, g') = 0$ exactly for disjoint pairs. Every
term is enumerable on `df_super`. Three structural facts follow. (i) The
diagonal reproduces the variance correction
$\sigma^2 + \operatorname{Var}(\mu_0 \mid g) +
\operatorname{Var}(\tau \mid g)/2 + \operatorname{Cov}(\mu_0, \tau \mid g)$
— verified identical at $5.9\times 10^{-16}$. (ii) $\Sigma_{\mathrm{pop}}$
is a Gram matrix of the influence functions $\psi(g)$ in $L^2$, hence
positive semi-definite; near-duplicate candidates place it close to
singular, so numerical evaluation uses an eigenvalue floor. (iii) The
off-diagonal is **not** the $\sigma$-overlap correlation applied to
corrected variances: the imbalance components covary through their own
intersection moments. This is precisely the mechanism behind the borderline
residual — for the engineered near-duplicate (no prognostic variation along
its differing coordinate), the additive law reduces
$\operatorname{corr}(Q, \mathrm{dup})$ from the $\sigma$-overlap value
$0.9747$ to $0.9385$, decorrelating exactly where both candidates sit at
the threshold.

**Verification** (`dev/detection-theory-continuous/check_additive_covariance_md.R`,
sandbox R 4.3.3/mvtnorm 1.2.4, seed 46, fresh population Monte Carlo,
$n_{\mathrm{rep}} = 6000$): at the adversarial borderline, MC $0.5580$
(SE $0.0064$) against additive computation $0.5568$ — gap $-0.0012$,
$0.2$ MC standard errors, where the naive and multiplicatively corrected
computations both sat at $0.536$ ($\approx 3.4$ SEs off); at the
alternative, MC $0.8507$ against additive $0.8511$ ($0.1$ SEs). The
population layer is therefore exact to leading order with all terms
enumerable; what remains outside it is second order — the $O(1/n)$
ratio-estimator terms from random arm counts and the non-normality of
small binomial pieces.

## Level 5: idealization — the gaps between the computed procedure and forestsearch

The computed functionals idealize the procedure in four named ways, each
with a diagnostic signature rather than a hidden discrepancy. **Fixed
versus regenerated family:** the computation conditions on an enumerated
family; the real identifier regenerates candidates per replicate from
data-dependent quantile cuts under `n.min` viability. Growth of a fitted
$L_{\mathrm{eff}}$ with $N$, and a measured sensitivity decline against a
fixed-family law that concentrates on maximal-mean candidates, are
regeneration signatures. **Single split versus
$\hat p_{\mathrm{cons}}$:** the computed consistency rung is the idealized
two-way split; the real screen aggregates hundreds of random splits into
$\hat p_{\mathrm{cons}}$ and gates on $\hat p_{\mathrm{cons}} \ge$ a
threshold. The single-split event and the multi-split gate bracket one
another in a definite direction discussed in the worked example; the exact
composition uses the stability closed form and remains a deferred rung.
**Argmax scope:** the computed selection is argmax-over-all; the procedure
takes the maximum over *admitted* candidates under `sg_focus` — the same
difference transform on a subset. **Nomination:** GRF/LASSO screening into
the family is unmodeled, exactly as in the April $L_{\mathrm{eff}}$
document.

## Level 6: MR-specific — the fidelity benchmark and the empirical core

MR's per-draw re-selection ranks by `maxcons` on unrounded `zcons`
(`R/fs_mr_inference.R:169`) — a different statistic than the realized
selection. Both rankings are functionals of the same Gaussian object:
effect-ranking uses the identity transform; consistency-ranking uses, to
first order with design-known SEs, the diagonal rescaling
$D = \mathrm{diag}(1/\mathrm{SE}_m)$, so the zcons-ranked argmax is the
argmax of $D\hat\beta$ under the same difference-transform recipe.
Consistency-ranking weights small-SE (large) regions — the computed
counterpart of the measured $2.46\times$ ($n = 1000$) to $10.29\times$
($n = 4000$) winner-size inflation — and the two computed selection laws
give the semi-analytic benchmark for the fidelity fix, including the
falsifiable prediction that simulated winner sizes collapse to the realized
range, before any grid re-run. What stays empirical, by construction: MR's
*sufficiency* — the residual plateau near 37 and the coverage decay
$0.957 \to 0.731$ are measurements against certified references, and the
resolved mechanism (`REPORT_mr_mechanism.md`, `5b4efb4`) is the fidelity
gap, not linearization failure. The correction *scale* is derivable
($O(n^{-1/2})$; measured decay $n^{-0.60}$).

## The error ordering at a glance

| Level | New error source | Enters | Measured magnitude | Control |
|---|---|---|---|---|
| 0–1 | none | — | $10^{-14}$–$10^{-15}$ | machine |
| 2 | quadrature | computed probabilities | $3\times 10^{-8}$–$10^{-5}$ | `abseps` |
| 3 | CLT; plug-in $\widehat\Sigma$ | law shape; scale | $0.016$; $0.012$ | arm sizes; pooling |
| 4 | leading-order influence | population law | $0.001$ (additive); second order remains | enumerated moments |
| 5 | idealization | procedure fidelity | regeneration signatures on record | conditioning; deferred rungs |
| 6 | ranking statistic; MR sufficiency | re-selection; correction | winner-size $2.46$–$10.29\times$; residual $\sim 37$ | benchmark; empirical |

# Part III — Prediction recipes {#sec-recipes}

## What exactly is being predicted {#sec-what}

Every recipe below computes a probability or expectation of the **candidate
effect process** — the joint Gaussian law of the within-region estimators
over a fixed family — and each corresponds to a specific quantity a
simulation report tabulates. Before the individual recipes, the full map:

| Recipe | Input required | Output (the predicted quantity) | Compares to (simulation column) | Exactness |
|---|---|---|---|---|
| Marginal clearance | $\beta(g)$, $\sigma_g(n)$, $c_1$ | $P(\hat\beta(g) \ge c_1)$ for one fixed $g$ | per-region clearance counts (e.g., how often $Q$'s refit exceeds the floor) | Level 2–4 |
| Family declaration (Rung A) | means, $\Sigma$, $c_1$ | $P(\max_m \hat\beta_m \ge c_1)$ | detection rate if the procedure screened only | Level 2–4 |
| Split-half consistency (Rung B) | means, $\Sigma$, $c_1$, $c_2$ | per-candidate $P_1$ and $P(\text{any pass})$ | `DETECTED` share of the status table | Level 2–4 + single-split idealization |
| Selection distribution (Rung C) | means, $\Sigma$, gates, ranking rule | $P(\hat H = g_m)$ for each $m$ | realized-rule frequency table (`sg_def`); covariate-involvement shares | Levels 2–5 |
| Winner size | selection distribution, $P(g)$ | $\mathbb E[\,|\hat H|\,]$ | `n_harm` mean (identified-subgroup-sizes callout) | Levels 2–5 |
| Classification | selection distribution, overlaps with $Q$ | expected sens/spec/PPV/NPV | identification table | Levels 2–5 |
| Conditional estimand | selection distribution, mixture means | $\mathbb E[\beta(\hat H)]$ | `betaHhat_H` mean (oriented) | Levels 2–5 |
| Naive selection bias | selection distribution, winner noise | $\mathbb E[\hat\beta(\hat H) - \beta(\hat H)]$ | naive row, "vs $\beta(\hat H)$" bias column | Levels 2–5 |
| Coverage (conditional) | $(b, w, \sigma)$ | interval coverage given bias and width | coverage columns, given measured $(b, w)$ | closed form |
| Implied $L_{\mathrm{eff}}$ | null-family law | effective independent-candidate count | null-cell detection rate | Level 2–4 |
| Design calculators | any recipe, inverted | $n$ (or floors) achieving a target rate | — (planning, not comparison) | as the recipe |

(The Rung A/B/C labels in the table are the names these functionals carry
in the verification artifacts under `dev/detection-theory-continuous/`,
kept for traceability.) Two global caveats attach to every row. First, the family is
**conditioned on**: the recipes take $\mathcal G$ as given, while
forestsearch regenerates it per replicate (Level 5) — so fixed-family
predictions are structural benchmarks for the real design, exact for the
scenario they define, and become replicate-faithful when fed a realized
family (@sec-realized). Second, the consistency gate is idealized as a
single split unless the $\hat p_{\mathrm{cons}}$ composition rung is added.

## Marginal floor clearance of a fixed region

In words: for one pre-specified region, the probability that its fitted
effect clears the screening floor is a single normal tail probability,
because the mean is the mixture value and the scale is the Tier B standard
error. $P\big(\hat\beta(H) \ge c_1\big) = \bar\Phi\big((c_1 -
\beta_o(H))/\sigma_H(n)\big)$. The measured anchor (O1: the committed
50-replicate overnight anchor run):

```{r clearance-example}
z <- (40 - 30) / 13.6786          # oriented truth, floor, measured SE at n=1000
pred <- pnorm(z)
cat(sprintf("Marginal clearance of Q at n = 1000: Phi(%.3f) = %.3f;\n", z, pred),
    "measured (overnight O1): 36/50 = 0.720, MC SE 0.064 -- consistent.\n")
```

The marginal recipe alone under-predicts *family* detection, and the gap is
the point: at $n = 500$, $\sigma_Q$ scales to $13.6786\sqrt 2 = 19.34$,
giving marginal clearance $\Phi(10/19.34) = 0.697$ for $Q$ — against a
measured family detection rate of $997/1000$. The difference is candidate
competition (Level 2, computed exactly by the next recipe) plus family
regeneration (Level 5, isolated).

## Family declaration (screening exceedance)

In words: the probability that *some* candidate's fitted effect clears the
floor. Because the joint law is known, this is one rectangle probability:
$P(\max_m \hat\beta_m \ge c_1) = 1 - \Phi_M(c_1\mathbf 1;\ m, \Sigma)$,
with $\Phi_M$ the $M$-variate normal distribution function.
Verified design-conditionally at $0.8297/0.5281/0.0262$ computed against
data-level Monte Carlo $0.8243/0.5400/0.0288$ (alternative / borderline /
null).

## Split-half consistency

In words: the idealized consistency gate splits the sample in two and asks
that the candidate clear the floor on average and a lower floor on each
half, with $W_1, W_2$ the candidate's fitted effects on the two halves. Per
candidate the passing probability is a one-dimensional integral,
$$
P_1 \;=\; \int_{c_2}^{\infty} \phi_{m_1, s_1}(w)\,
\bar\Phi_{m_2, s_2}\!\big(\max(c_2,\ 2c_1 - w)\big)\, dw,
$$
the Leon et al. (2024, SIM §2.1) object with $(m_j, s_j)$ that candidate's
mean and standard deviation on half $j$ — verified to
$7.5\times 10^{-4}$; the family rate ("some candidate
passes") is Gaussian-level Monte Carlo on the exact $2M$-block law,
verified $0.8206$ vs $0.8187$. The real gate replaces the single split by
$\hat p_{\mathrm{cons}} \ge$ threshold over hundreds of splits; the
relation between the two is taken up concretely in the worked example.

## Selection distribution and its consequences

In words: not just *whether* something is declared but *which* candidate
wins. For the effect-ranked argmax,
$$
P\big(\hat H = g_m,\ \text{declared}\big)
= P\big(\hat\beta_m \ge c_1,\ \hat\beta_m - \hat\beta_j \ge 0\ \forall j\big),
$$
a rectangle in the difference transform — one `pmvnorm` call per candidate,
with the internal identity
$\sum_m P(\hat H = g_m, \text{declared}) = P(\text{family declaration})$
verified to $3\times 10^{-7}$. Gated rules (such as maxeffCons: argmax of
effect among consistency-passers) are the same functional with the gate
event intersected, evaluated by Gaussian-level Monte Carlo. Selection
probabilities are what per-candidate clearances are not, and the
distinction drives every downstream prediction: winner size, expected
$\beta(\hat H)$, classification metrics, and the naive-bias forecast (the
selection-noise functional $\mathbb E[\text{noise of the argmax}]$ — the
quantity `design-checks/oracle_design.R` already evaluates on the same
law). The zcons-ranked variant of Level 6 uses the same recipe under the
diagonal rescaling.

## Coverage, composed

Conditionally on $(b, w, \sigma)$, coverage is
$\Phi\big((w/2 - b)/\sigma\big) - \Phi\big((-w/2 - b)/\sigma\big)$,
demonstrated against the grid in the simulation summary. Unconditionally it
composes over the selection law,
$\sum_m P(\hat H = g_m)\,\mathrm{cov}_m$ — semi-analytic given MR's
realized $(b_m, w_m)$, which remain empirical inputs (Level 6).

## Implied $L_{\mathrm{eff}}$ as an output

The family null rate from the joint law yields an implied count directly;
with prevalence-heterogeneous candidate variances a single-$P_1$
normalization is ill-defined in the deep tail, and the stable readout is
against $\max_m P_1$,
$L_{\mathrm{eff}} = \log(1 - P_{\mathrm{fam}})/\log(1 - \max_m P_1)$:
$2.33, 1.80, 1.08$ at $N = 300, 600, 1200$ in the
verification family, with the family/union-bound ratio ($0.671 \to 0.989$)
exhibiting the correlation discount. Growth of a fitted
$L_{\mathrm{eff}}$ with $N$ is a Level 5 regeneration phenomenon.

## A three-candidate warm-up, read line by line {#sec-warmup}

Before the full worked example, a minimal geometry that shows what the
recipes return and how to read them. Three candidates at the fixture's
means and measured scale: $Q$ itself; an overlapping proxy $P$
(prevalence $0.45$, sharing probability mass $0.28$ with $Q$, hence purity
$P(Q \mid P) = 0.28/0.45 = 0.622$); and a near-duplicate $D \subset Q$
(prevalence $0.31$, purity $1$). SEs are anchored to the measured
$\sigma_Q(n{=}1000) = 13.6786$ and scaled by the balanced-arms prevalence
rule $\sigma_H = \sigma_Q\sqrt{\pi_Q/\pi_H}$; correlations are the
$\sigma$-overlap block (per @sec-ladder Level 3, the fitted-SE anchor
already carries the Level 4 variance component; the covariance refinement
of @sec-additive applies when DGM moments are in hand).

```{r worked-prediction}
set.seed(20260824)   # pins pmvnorm's quasi-MC randomization for the gap readout
library(mvtnorm)
tauQc <- 26.255235876; bint <- 13.744764124   # oriented
c1 <- 30; seQ <- 13.6786; piQ <- 0.3446
cand <- data.frame(
  name = c("Q", "P (overlap proxy)", "D (near-dup, subset of Q)"),
  pi   = c(piQ, 0.45, 0.31),
  piQg = c(1.00, 0.28/0.45, 1.00))          # P(Q | g)
cand$beta <- tauQc + bint * cand$piQg
cand$se   <- seQ * sqrt(piQ / cand$pi)
ov <- matrix(c(piQ, 0.28, 0.31,
               0.28, 0.45, 0.28*0.31/piQ,   # |P∩D| via D ⊂ Q composition
               0.31, 0.28*0.31/piQ, 0.31), 3, 3)
R  <- ov / sqrt(outer(cand$pi, cand$pi))
Sg <- R * outer(cand$se, cand$se)
p_marg <- pnorm(c1, cand$beta, cand$se, lower.tail = FALSE)
p_fam  <- 1 - pmvnorm(upper = rep(c1, 3), mean = cand$beta, sigma = Sg,
                      algorithm = GenzBretz(abseps = 1e-6))
sel <- vapply(1:3, function(m) {
  A <- rbind(diag(3)[m, ],
             do.call(rbind, lapply(setdiff(1:3, m),
                                   function(j) diag(3)[m, ] - diag(3)[j, ])))
  as.numeric(pmvnorm(lower = c(c1, 0, 0), upper = rep(Inf, 3),
                     mean = as.numeric(A %*% cand$beta),
                     sigma = A %*% Sg %*% t(A),
                     algorithm = GenzBretz(abseps = 1e-6)))
}, numeric(1))
out <- data.frame(candidate = cand$name,
                  beta_oriented = round(cand$beta, 3),
                  se = round(cand$se, 3),
                  marginal_clearance = round(p_marg, 4),
                  selection_prob = round(sel, 4))
knitr::kable(out, caption = sprintf(
  "Warm-up at n = 1000: family declaration %.4f; selection identity gap %.1e.",
  as.numeric(p_fam), abs(sum(sel) - as.numeric(p_fam))))
```

Reading the table. **Column `beta_oriented`** is the mixture mean: $Q$ and
$D$ both sit at $40$ (both have purity $1$), $P$ at $34.81$. **Column
`se`** is the prevalence-scaled measured scale: smaller regions are
noisier, so $D$ ($\pi = 0.31$) carries the largest SE. **Column
`marginal_clearance`** treats each candidate alone: $Q$ clears the floor
$76.8\%$ of the time, $D$ $75.6\%$, $P$ $65.6\%$. **Column
`selection_prob`** is the joint answer, and it is *not* proportional to the
marginals: $D$ takes the largest share ($0.360$) despite the smaller
marginal clearance, because at equal means the *noisier* candidate wins the
argmax more often — the Gaussian-level mechanism behind the observed
tendency of the identifier to return small, noise-favored regions. The
three selection probabilities sum to the family declaration rate ($0.8474$)
up to the printed quadrature gap — an internal consistency identity, not a
new assumption. Finally, family declaration ($0.847$) exceeds the best
marginal ($0.768$): competition helps *declaration* even as it dilutes any
single candidate's *selection*.

## Design calculators

The recipes invert. Sample size for a target marginal clearance of $Q$:

```{r design-calculator}
n_for <- function(target, se_n1000 = 13.6786, mu = 40, c1 = 30)
  1000 * (se_n1000 * qnorm(target) / (mu - c1))^2
cat(sprintf("n for %d%% marginal clearance of Q at floor 30: %s\n",
            c(80L, 90L, 95L),
            format(round(n_for(c(.80, .90, .95))), big.mark = ",")), sep = "")
```

Family-level and selection-level counterparts invert the corresponding
functionals by the same one-dimensional search; Monte Carlo error budgets
for grid design follow from computed rates rather than convention.

# A detailed worked example: the maxeffCons MR design at n = 500 {#sec-worked}

This section carries the machinery end to end against a specific committed
design:
`sim_fs_maxeffCons_mr_md40_knoise0_n500_batch_1_1000.qmd`. Every input is
taken from that document's own setup chunks; every output is stated
together with the report column it should be compared against.

## Step 1 — the design under study

From the driver's setup, verbatim: $n = 500$ per replicate; the committed
alternative cell (`target_md_harm = -40` raw, oriented $+40$ in $Q$);
floors `md_threshold = 30`, `md_consistency = 10`; consistency gate
$\hat p_{\mathrm{cons}} \ge 0.90$ over `fs.splits = 400` random splits;
`n.min = 60` (a $60/500 = 12\%$ prevalence floor on candidates);
`maxk = 2` (conjunctions of at most two conditions);
`sg_focus = "maxeffCons"` — in the driver's own words, "highest effect
meeting the consistency rate"; MR with 5000 draws; no added noise
covariates. The candidate cuts are the deliberate part of the design: J
$= 10$ quantile grids on `age` and `preanti` **replace** the default cuts,
so the true boundaries ($34$ and $744.5$) are *not representable*; the
driver documents the best achievable rule as
`age > 33 & preanti <= 677.9` at Jaccard $0.882$ against $Q$, and gives
the grid quantile positions (the $33$ cut at the $5/11 = 45.5\%$ age
quantile; the preanti grid $0, 0, 0, 0, 55.8, 214.1, 432, 677.9, 872.8,
1070.5$ at the $k/11$ quantiles, so $677.9$ sits at $72.7\%$). Here and
throughout, the **Jaccard index** of a rule $g$ against $Q$ is
$$
J(g, Q) \;=\; \frac{P(g \cap Q)}{P(g \cup Q)},
$$
the overlap as a fraction of the union — equal to $1$ exactly when
$g = Q$, and tied to the classification identities of Part I,
Corollary (v), by the exact relation
$1/J = 1/\mathrm{sens} + 1/\mathrm{PPV} - 1$: the driver's rule-recovery
metric is the harmonic composite of recovery and purity.

Its report tabulates, per batch: the status trichotomy
(`DETECTED / NO-DETECTION / CONFIG-ERROR`); identified-subgroup sizes
(`n_harm`); the realized-rule frequency table (`sg_def`) and
covariate-involvement shares; classification (sens/spec/PPV/NPV); and the
estimation table scoring oracle/naive/MR against the exact
$\beta(\hat H)$, the oracle refit, and the structural and fitted targets.
Those are exactly the quantities the recipes predict.

## Step 2 — what must be specified, and what the driver already supplies

The recipes need five inputs. The driver supplies four directly: the
oriented effects ($\tau_{Q^c} = +26.2552$, $\beta_{\mathrm{int}} =
+13.7448$, hence purity translation $P(Q \mid g) \ge 0.2724$ for true
clearance — the driver's own "27.2%" remark); the sample size and floors;
the selection rule (maxeffCons); and the scale anchor (the committed
measured $\sigma_Q(n{=}1000) = 13.6786$, which the fitted-SE observation
of Level 3 licenses as a population-effective scale, rescaled to $n = 500$
by $\sqrt 2$). The fifth input — **the candidate family and the joint
covariate law that determines every $P(g)$, $P(Q \mid g)$, and overlap** —
is the one forestsearch regenerates per replicate, so a prediction must
either condition on a realized family (@sec-realized) or construct a
scenario satisfying the fixed-family requirements. This example does the
latter, using only quantities the driver itself documents.

## Step 3 — constructing the scenario from documented anchors

Cross the two true-region axes at their *representable* resolutions. Age
takes three categories $\{\le 33,\ (33, 34],\ > 34\}$ — the family can cut
at $33$ (a grid cut) but not at $34$ (the true boundary, absent from the
grid). Preanti takes three categories
$\{\le 677.9,\ (677.9, 744.5],\ > 744.5\}$ — cut at $677.9$ representable,
true boundary $744.5$ not. Marginals come from the documented quantile
positions ($P(\text{age} > 33) = 1 - 5/11 = 0.545$;
$P(\text{age} > 34) = 0.50$, the median;
$P(\text{preanti} \le 677.9) = 8/11 = 0.727$;
$P(\text{preanti} \le 744.5) = 0.75$, the q75 default). The $3 \times 3$
joint below satisfies **all** the documented anchors simultaneously — the
four marginals, $P(Q) = 0.3446$, and Jaccard of the best representable
rule against $Q$ equal to $0.8820$ — and the chunk asserts each one. A
third, stylized covariate $V$ (prevalence $0.42$, independent) stands in
for the distractor pool; the driver's real distractors include `str2`,
which nests inside the true partner condition and is *not* independent —
the stylization is stated, not hidden. Candidates are all single
conditions and cross-variable pairs over the three representable cuts
(both sides), pruned at the $12\%$ `n.min` floor.

```{r worked-scenario}
library(mvtnorm)
set.seed(20260825)
tauQc <- 26.255235876; bint <- 13.744764124        # oriented
c1 <- 30; c2 <- 10; n_rep <- 500; pcons_gate <- 0.90
seQ1000 <- 13.6786; piQ_fix <- 0.3446
# 3x3 joint: rows age (<=33, (33,34], >34); cols preanti (<=677.9,
# (677.9,744.5], >744.5). Hand-solved to hit every documented anchor.
J3 <- matrix(c(0.3553, 0.0121, 0.0876,
               0.0360, 0.0020, 0.0070,
               0.3357, 0.0089, 0.1554), 3, 3, byrow = TRUE)
stopifnot(abs(sum(J3) - 1) < 1e-12,
          abs(rowSums(J3) - c(0.455, 0.045, 0.500)) < 1e-12,
          abs(colSums(J3) - c(0.727, 0.023, 0.250)) < 1e-12)
PQ <- J3[3, 1] + J3[3, 2]                    # Q = age>34 & pre<=744.5
stopifnot(abs(PQ - piQ_fix) < 1e-4)
gstar_P <- J3[2, 1] + J3[3, 1]               # g* = age>33 & pre<=677.9
jac <- J3[3, 1] / (gstar_P + PQ - J3[3, 1])
stopifnot(abs(jac - 0.882) < 1e-3)
cat(sprintf("Scenario anchors: P(Q) = %.4f; Jaccard(g*, Q) = %.4f (driver: 0.882)\n",
            PQ, jac))
# 18 cells: 9 (age x pre) x V in {1, 0}, V independent, P(V=1) = .42
pV <- 0.42
cells <- expand.grid(a = 1:3, p = 1:3, v = 1:0)
cells$prob <- J3[cbind(cells$a, cells$p)] * ifelse(cells$v == 1, pV, 1 - pV)
cells$inQ  <- cells$a == 3 & cells$p <= 2
# Candidate conditions on representable cuts (both sides) + V
conds <- list(`age>33` = cells$a >= 2, `age<=33` = cells$a == 1,
              `pre<=677.9` = cells$p == 1, `pre>677.9` = cells$p >= 2,
              `V=1` = cells$v == 1, `V=0` = cells$v == 0)
fam <- c(lapply(names(conds), function(nm) list(nm = nm, m = conds[[nm]])),
         unlist(lapply(list(c(1,3), c(1,4), c(2,3), c(2,4),
                            c(1,5), c(1,6), c(2,5), c(2,6),
                            c(3,5), c(3,6), c(4,5), c(4,6)),
   function(ix) list(list(nm = paste(names(conds)[ix], collapse = " & "),
                          m = conds[[ix[1]]] & conds[[ix[2]]]))),
   recursive = FALSE))
Pg  <- vapply(fam, function(f) sum(cells$prob[f$m]), numeric(1))
keep <- Pg >= 0.12                                  # n.min = 60 of 500
fam <- fam[keep]; Pg <- Pg[keep]; M <- length(fam)
lab <- vapply(fam, `[[`, character(1), "nm")
PQg <- vapply(fam, function(f) sum(cells$prob[f$m & cells$inQ]), numeric(1)) / Pg
beta_g <- tauQc + bint * PQg
se_g   <- seQ1000 * sqrt(2) * sqrt(piQ_fix / Pg)    # n = 500 anchor scaling
sens_g <- vapply(fam, function(f) sum(cells$prob[f$m & cells$inQ]), numeric(1)) / PQ
spec_g <- vapply(fam, function(f) 1 - sum(cells$prob[f$m & !cells$inQ]) / (1 - PQ),
                 numeric(1))
ovl <- outer(seq_len(M), seq_len(M), Vectorize(function(i, j)
  sum(cells$prob[fam[[i]]$m & fam[[j]]$m])))
Rho <- ovl / sqrt(outer(Pg, Pg))
Sg  <- Rho * outer(se_g, se_g)
tab3 <- data.frame(candidate = lab, prevalence = round(Pg, 4),
                   purity_PQg = round(PQg, 4), beta = round(beta_g, 2),
                   se_n500 = round(se_g, 1),
                   clears_floor = ifelse(beta_g >= c1, "yes", "no"))
knitr::kable(tab3[order(-tab3$beta), ], row.names = FALSE, caption = sprintf(
  "The scenario family (M = %d after the 12%% floor): mixture means, anchored scales, and the purity translation of the floor.", M))
```

Reading the family table. The purity column *is* the mean column in
disguise (Proposition 1), and the floor line falls at purity $0.2724$ —
below the marginal prevalence $0.3446$, so **nine of the sixteen
candidates have true effects above the floor, including the pure
distractors** $V{=}1$ and $V{=}0$ at $\beta = 30.99$. On this pathway the
screening floor does not separate signal from noise regions; it separates
regions *depleted* of $Q$ from everything else. The best representable
rule $g^\ast$ leads at $\beta = 38.67$ (purity $0.903$), $1.33$ units
below the unrepresentable truth — the price of the J-grid design,
quantified before any simulation.

## Step 4 — predictions, each mapped to its report column

The maxeffCons rule is: among candidates passing the consistency gate,
select the one with the largest fitted effect; declare `NO-DETECTION` if
none passes. The idealized computation below replaces the
$\hat p_{\mathrm{cons}} \ge 0.90$ multi-split gate with the single-split
event $\{W_1 + W_2 \ge 2c_1,\ \min(W_1, W_2) \ge c_2\}$ — $W_1, W_2$
again the half-sample fitted effects — and evaluates
everything on the known Gaussian law by Monte Carlo (Level 2 numerics;
$2 \times 10^5$ draws, so rate SEs are $\approx 0.001$). Direction of the
idealization, stated: with $c_2 = 10$ sitting far below every candidate
mean (the driver's open question D2 — the consistency floor is
non-discriminating on this cell), the per-split floor events are nearly
always satisfied and both gates reduce to variants of "the effect
replicates above threshold"; a demanding $\hat p_{\mathrm{cons}} \ge 0.90$
over 400 splits is *stricter* than one split, so at fixed family the
computed detection rate is an upper benchmark and the computed selection
law shifts mass toward stable (larger, higher-mean) candidates relative to
the real gate; the far larger real family pushes detection the other way
and dominates it here.

```{r worked-predictions}
Rdraw <- 2e5
# Complement pairs make the population law EXACTLY rank-deficient: for any
# cut, P(g+)c(g+) + P(g-)c(g-) is the full-sample coefficient vector, so the
# candidate covariance is a singular Gram matrix (the Level 4 PSD fact, made
# concrete). Sample by eigendecomposition; degenerate directions get zero
# variance, which enforces the linear constraints exactly.
eS <- eigen((Sg + t(Sg)) / 2, symmetric = TRUE)
root_half <- eS$vectors %*% diag(sqrt(pmax(2 * eS$values, 0)))  # cov = 2*Sg
W1 <- matrix(rnorm(Rdraw * M), Rdraw, M) %*% t(root_half) +
      rep(beta_g, each = Rdraw)
W2 <- matrix(rnorm(Rdraw * M), Rdraw, M) %*% t(root_half) +
      rep(beta_g, each = Rdraw)
Bhat <- (W1 + W2) / 2                   # full-sample effects, exact law N(beta, Sg)
pass <- (W1 + W2 >= 2 * c1) & (W1 >= c2) & (W2 >= c2)
det  <- rowSums(pass) > 0
Bmask <- Bhat; Bmask[!pass] <- -Inf
winner <- max.col(Bmask, ties.method = "first"); winner[!det] <- NA_integer_
P1 <- colMeans(pass)
p_sel <- tabulate(winner[det], M) / Rdraw
det_rate <- mean(det)
sel_c <- p_sel / det_rate                       # selection | detection
EnH   <- n_rep * sum(sel_c * Pg)                # expected |Hhat| | detection
Esens <- sum(sel_c * sens_g); Espec <- sum(sel_c * spec_g)
Eppv  <- sum(sel_c * PQg)
Enpv  <- sum(sel_c * vapply(seq_len(M), function(m) {
  pHc <- 1 - Pg[m]; (pHc - (PQ - PQg[m] * Pg[m])) / pHc }, numeric(1)))
EbetaH <- sum(sel_c * beta_g)
noise  <- Bhat - rep(beta_g, each = Rdraw)
Enaive_bias <- mean(noise[cbind(which(det), winner[det])])
mass_below <- sum(sel_c[beta_g < c1])          # selections on true-mean-below-floor rules
top <- order(-p_sel)[1:6]
knitr::kable(data.frame(candidate = lab[top],
                        P1_single_split = round(P1[top], 3),
                        P_selected = round(p_sel[top], 3),
                        sel_given_det = round(sel_c[top], 3)),
             row.names = FALSE, caption = sprintf(
  "Top of the predicted selection distribution (detection rate %.3f). Compare P_selected against the driver's realized-rule (sg_def) frequency table.",
  det_rate))
knitr::kable(data.frame(
  quantity = c("detection rate (DETECTED share)",
               "E[|Hhat|] given detection (n_harm mean)",
               "E[sens]", "E[spec]", "E[PPV]", "E[NPV]",
               "E[beta(Hhat)] oriented (betaHhat_H mean)",
               "E[naive - beta(Hhat)] (naive bias vs beta(Hhat))",
               "selection mass on rules with true mean below the floor"),
  predicted = round(c(det_rate, EnH, Esens, Espec, Eppv, Enpv,
                      EbetaH, Enaive_bias, mass_below), 3)),
  row.names = FALSE, caption =
  "Scenario predictions with their driver-report comparison targets in parentheses.")
```

Reading the predictions, line by line.

**Detection.** The computed rate is $0.965$: with nine candidate means
above the floor and half-sample SEs of $19$–$40$, *some* candidate clears
on nearly every replicate. The committed real-family measurement at
$n = 500$ is $997/1000$ — higher still, in exactly the direction the
family-size caveat below predicts — and the driver's own status table is
this prediction's comparison target. Detection on this cell is not
evidence of finding $Q$; the interesting prediction is *what* gets
selected.

**The selection distribution** is the substantive output, and it inverts
the marginal ordering. $g^\ast$ leads the gate ($P_1 = 0.63$, the largest)
yet takes only the second-largest selection share: maxeffCons is an
effect-ranked argmax over gate-passers, and *conditional on passing*, a
high-variance candidate's fitted effect is inflated far above its mean —
so noisy low-purity rules out-bid the leader whenever their noise draws
large. The aggregate diagnostic in the second table makes this exact: a
substantial fraction of all selections lands on rules whose **true** mean
sits below the floor — rule-level false-harm declarations, predicted from
the law, not discovered in a run. Each row of the selection table is
comparable rule for rule against the driver's `sg_def` frequency table,
and every share taken by a low-purity rule drags the classification and
estimand predictions below their $g^\ast$ ceilings.

**Winner size and classification.** $\mathbb E[\,|\hat H|\,]$ is the
selection-weighted prevalence times $n$ — compare to the `n_harm` mean in
the identified-subgroup-sizes callout — and expected sens/spec/PPV/NPV are
selection-weighted overlaps with $Q$, comparable to the identification
table. Because $g^\ast$ carries sens $= 0.974$ and PPV $= 0.903$, the
predictions sit high exactly to the extent selection concentrates on it;
the shortfall from those ceilings measures the noise-selection share.

**The conditional estimand and the naive bias.** $\mathbb E[\beta(\hat
H)]$ is the selection-weighted mixture mean (compare: oriented
`betaHhat_H` mean over detected replicates), and the naive-bias forecast
is the mean noise of the winner (compare: the naive row's "vs
$\beta(\hat H)$" column). The latter is the winner's-curse functional of
`oracle_design.R`, now with this design's own means and covariances as
inputs.

**What this scenario cannot predict** — stated as sharply as what it can.
The real family is far larger (six continuous covariates at ten cuts each,
both sides, plus seven binaries, under `maxk = 2` and the viability
floors), so real selection mass is diluted across many more low-purity
competitors: relative to this $M = 16$ scenario, expect *lower*
concentration on $g^\ast$, *larger* naive bias, and classification
metrics *below* these predictions — with the direction certain and the
magnitude requiring either the realized family (@sec-realized) or a larger
constructed one. And the multi-split gate strengthens the consistency
requirement relative to the single split, filtering marginal passers; both
idealizations are Level 5 rungs with signs attached, not silent gaps.

## Step 5 — the null-cell twin, predicted before it runs

The driver's `null_cell = TRUE` switch sets $\beta_{\mathrm{int}} = 0$, so
every region's oriented mean collapses to the common treatment effect
($\approx +26.26$, the complement value). The same machinery then predicts
the null cell's operating characteristics — and the prediction is the
useful kind of surprising: the floor sits only $(30 - 26.26)/\mathrm{SE}
\approx 0.13$–$0.28$ SEs above every candidate mean, so this "null" is a
**near-threshold null**, not a zero null.

```{r worked-null}
beta_null <- rep(tauQc, M)
W1n <- matrix(rnorm(Rdraw * M), Rdraw, M) %*% t(root_half) +
       rep(beta_null, each = Rdraw)
W2n <- matrix(rnorm(Rdraw * M), Rdraw, M) %*% t(root_half) +
       rep(beta_null, each = Rdraw)
pass_n <- (W1n + W2n >= 2 * c1) & (W1n >= c2) & (W2n >= c2)
cat(sprintf("Null-cell predicted detection (single-split gate): %.3f\n",
            mean(rowSums(pass_n) > 0)))
p1n <- colMeans(pass_n)
cat(sprintf("  per-candidate P1 range: %.3f-%.3f; implied L_eff vs max(P1): %.2f\n",
            min(p1n), max(p1n),
            log(1 - mean(rowSums(pass_n) > 0)) / log(1 - max(p1n))))
```

The predicted null-cell detection rate is far above any conventional
false-positive intuition, for a structural reason the mixture identity
makes exact: with a nonzero common effect near the floor, declaration is a
coin-flip per candidate and near-certain per family. This is the
borderline-null regime of the manuscript's §7 anchors, derived rather than
discovered — and it quantifies the driver's open question D2 from the
other side: on the null cell it is the *screening* floor, not the
consistency floor, that is barely discriminating.

## Step 6 — inverting for design

The same objects answer planning questions. For example: at what $n$ would
the marginal clearance of the best representable rule $g^\ast$ reach
$90\%$, and at what screening floor would the null-cell family detection
fall to $10\%$? Both are one-dimensional inversions of computed rates —
the first is the closed-form calculator of @sec-recipes applied at
$\beta = 38.67$; the second solves $c_1$ in the null family functional.
Floors chosen this way are calibrated to computed operating
characteristics rather than convention — with the family-size caveat of
Step 4 carried along.

## Step 7 — doing it with the realized family {#sec-realized}

For replicate-faithful numbers, replace the constructed scenario by a
realized family: from a fitted `forestsearch` object, the evaluated
candidate set with per-candidate membership (equivalently the arm-wise
counts $n_{g,a}$ and pairwise intersections) determines the exact
design-conditional law of Level 2 — means from the mixture identity
evaluated on `df_super` at each candidate's rule, covariance from the
overlap formula with no balanced-arms approximation. Every prediction of
Step 4 then re-runs unchanged on that law, now exact-conditional rather
than scenario-structural, with only the gate idealization and the
quadrature error remaining between computation and the replicate's truth.
That extraction is the natural next artifact on this thread; it is a
consumer of existing objects and requires no package changes.

# Part IV — Consolidated verification record {#sec-accuracy}

**Certified layer, exercised in the committed grid:** CHECK A at
$1.421\times 10^{-14}$ across 601–912 realized rules per cell; CHECK B
reproduced the calibration to machine precision on every run; CHECK C
caught a real scale-convention defect on first firing and now agrees to
$0.000$; the bit-level standard isolated an RNG-kind dependence and the
post-fix re-render reproduced the pilot on 35 of 35 scientific columns.
Every defect of the campaign was caught by an exact expectation; none by a
tolerance.

**Prediction layer:** the on-record suite
(`dev/detection-theory-continuous/check_declaration_gaussian_md.R`)
verified 16 PASS / 0 FAIL on the primary machine (R 4.6.1, mvtnorm 1.4.2,
2.4 s), numerically identical line for line to the drafting-stage sandbox
run (R 4.3.3, mvtnorm 1.2.4) — cross-stack bit-level reproduction. The
additive population covariance of @sec-additive verified at $0.2$ / $0.1$
MC standard errors (borderline / alternative). The document's own rendered
chunks reconcile against the same record; the selection-identity gap
readout fluctuates at the quadrature tolerance ($10^{-6}$) between
platforms unless seeded, and the warm-up chunk pins the seed for that
reason. The worked example's Gaussian-level Monte Carlo uses $2 \times
10^5$ draws (rate SEs $\approx 10^{-3}$) and a fixed seed.

The verification family of the check suite is a synthetic $M = 14$
stressor whose complement effect is beneficial (oriented $-26.26$) —
deliberately harder competition geometry than the fixture's harmful
complement ($+26.2552$); the machinery checks are parameter-point-agnostic,
and the fixture values used throughout this document are the committed ones
of §2.

# Part V — The boundary, restated {#sec-boundary}

| Quantity | Status |
|---|---|
| $\beta(H)$ any rule; $\theta^\dagger(Q)$; ITT/complement/disjoint; classification identities | Tier A, exact |
| $\beta(H)$, binary/OR | Tier B, exact computation (mixture identity fails; RD retains Tier A) |
| Candidate covariance given family | Tier B, exact overlaps |
| Design-conditional joint law; family declaration; split-half consistency; selection distribution; winner-size law; naive-bias forecast | Distributional identity + quadrature; **verified** on fixed families |
| Population-level law | **Exact to leading order** — additive covariance decomposition derived, enumerable, verified ($0.2$ SEs at the adversarial point); second-order $O(1/n)$ terms remain |
| Marginal clearance; design calculators | Closed form; measured anchors consistent |
| Implied $L_{\mathrm{eff}}$ | Derived output; single-$P_1$ form requires variance homogeneity |
| Coverage | Conditionally closed-form; unconditional composes over the selection law given MR's $(b, w)$ |
| MR sufficiency | Empirical (mechanism resolved as the re-selection fidelity gap; the Level 6 benchmark specifies the fix's falsifiable prediction) |
| Family regeneration; nomination; multi-split $\hat p_{\mathrm{cons}}$ composition | Beyond the fixed-family computation (conditioning caveat stated; regeneration signatures identified; gate direction stated in @sec-worked) |
| Identifier's $n \to \infty$ limit | Property of the algorithm on a data-dependent cut lattice; carried as measurement |
| Survival, forward | AFT log-time: Tier A transfers wholesale for the uncensored contrast; observed-data Cox log-HR non-collapsible and censoring-dependent — Tier B at best, estimand distinctions maintained |

The division of labor, in one sentence: closed forms certify targets and
implementations exactly; the selection layer is predictable on fixed
families from certified inputs through an audited approximation ladder
whose every rung carries a measured error; and family regeneration,
nomination, and MR's sufficiency remain the empirical core the simulations
exist to measure — now against both certified references and computed
benchmarks.
