---
title: "The Analytic Layer for the MD/Harm Pathway: Verification and Prediction"
subtitle: "Certified Identities, the Approximation Ladder, and Semi-Analytic Operating Characteristics (Continuous Outcomes)"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>")
```

# Scope and provenance

This document consolidates the analytic materials for the continuous/MD harm
pathway into a single reference. It combines, and is intended to supersede as
the working reference, two committed documents in this directory:
`analytic_verification_math_details.qmd` (the verification tiers: what closed
forms certify exactly) and
`analytic_derivations_mr_operating_characteristics.qmd` (the prediction
layer: which MR operating characteristics are derivable before a grid runs).
Both parents are left unchanged in the repository. Beyond consolidation, this
document adds two things: the connective material between verification and
prediction — the explicit ladder from exact identities to each approximation
actually carried out, including the population-level covariance decomposition
previously deferred and now derived and verified — and a detailed worked
example (@sec-worked) that carries the machinery end to end against a
specific committed simulation design, stating for every predicted quantity
what is computed, from which inputs, and which column of the design's own
report it should be compared against.

Supporting artifacts: `dev/detection-theory-continuous/`
(`NOTE_declaration_gaussian_md.md`; `check_declaration_gaussian_md.R`,
on-record 16 PASS / 0 FAIL, cross-stack identical output on
R 4.6.1/mvtnorm 1.4.2 and R 4.3.3/mvtnorm 1.2.4; and
`check_additive_covariance_md.R`, the additive-decomposition verification
cited in @sec-ladder).

Estimand language is strict throughout. $\beta(\hat H)$ is the conditional
estimand — the exact treatment effect at the realized, data-selected rule,
computed on the fixed super-population with zero Monte Carlo error.
$\theta^\dagger(H)$ is the marginal true-subgroup effect at the designed
region $Q$. MR conditions on the realized candidate family, targets
$\beta(\hat H)$, and is scored against $\beta(\hat H)$, never
$\theta^\dagger$. Effects are reported on the oriented scale
(positive = harm) under `adverse_outcome = FALSE`; exact references are
stored on the raw scale and the orientation bridge applies the sign map once.

# Fixture, estimator, and event conventions

The committed harm-pathway fixture (`df_super`, $N = 5000$;
`actg175_continuous_simulations.qmd` lines 65, 100, 118–119): the designed
harm region is the covariate rule
$Q = \{\text{age} > 34 \ \wedge\ \text{preanti} \le 744.5\}$, with
$|Q| = 1723$ members and prevalence $\pi = P(Q) = |Q|/N = 0.3446$.
Raw-scale (stored) effects: $\tau_Q = -40$ (calibration target),
$\tau_{Q^c} = -26.255$,
$\beta_{\mathrm{int}} = \tau_Q - \tau_{Q^c} = -13.745$. The CATE is
two-valued. Screening floor $c_1 = 30$ and consistency floor $c_2 = 10$,
both on the oriented scale defined next.

**Notation used throughout.** A *region* (equivalently a *rule*) $H$ or
$g$ is a subset of the covariate space — in practice a conjunction of
covariate conditions — identified with the subset of `df_super` rows it
selects; $\hat H$ is the data-selected region. Its *prevalence* is
$P(g) = |g|/N$, and its **purity** is
$$
P(Q \mid g) \;=\; \frac{|g \cap Q|}{|g|},
$$
the fraction of $g$'s members lying inside the designed region $Q$ — so
$P(Q \mid Q) = 1$, and $P(Q \mid g) = \pi$ for any rule carrying no
information about $Q$. The *effect* of a region,
$$
\beta(g) \;=\; \mathbb E\big[\,Y(1) - Y(0) \mid X \in g\,\big]
\ \text{on the oriented scale},
$$
is its average treatment effect, computed exactly on `df_super`. **All
effect symbols in this document are oriented by default** (positive =
harm — the scale the estimator operates on under
`adverse_outcome = FALSE`): $\tau_{Q^c} = +26.255$,
$\beta_{\mathrm{int}} = +13.745$, $\tau_Q = \beta(Q) = +40$. Raw
(stored) values appear only when quoting stored references, and the word
*oriented* is never abbreviated to a super- or subscript — in this
project the prefix `or_` belongs to the **oracle** estimator, and the two
must not collide. Likewise the letter $g$ is reserved for regions: where a
GLM link function is needed it is written $\kappa$, never $g$. Display convention: numeric constants are shown to
three decimals (four where a stored anchor is itself a four-decimal
quantity, e.g., $\pi = 0.3446$ and Jaccard $0.882$); stored references
retain full precision, and every exactness check compares at machine
precision regardless of display.

Three conventions, verified against package source, govern everything below:
the estimator is the within-subgroup OLS treatment coefficient of
`lm(y ~ treat)` — the difference in arm means — on the oriented scale
(`R/glm_effect_estimators.R:814-821`), i.e., the difference-in-means
estimator of Wager ([W] §1.1, eq. 1.4; see Appendix B, @sec-refwager)
restricted to a region; the screening and consistency events
compare **point estimates** to the floors
(`R/subgroup_search.R:626`; `R/subgroup_consistency_main.R:528-531`), so
estimated variance never enters an event definition; and thresholds are
one-sided floors on the oriented scale, never direction switches
(`dev/glm-continuous-sims/NOTE_threshold_sign_md.md`).

One translation is used repeatedly and is worth stating once in words. By
the mixture identity of Part I (Proposition 1:
$\beta(g) = \tau_{Q^c} + \beta_{\mathrm{int}}\, P(Q \mid g)$), a
region's true effect exceeds the screening floor exactly when its purity
does:
$$
\beta(g) \ge c_1
\quad\Longleftrightarrow\quad
P(Q \mid g) \;\ge\; \frac{c_1 - \tau_{Q^c}}{\beta_{\mathrm{int}}}
= \frac{30 - 26.255}{13.745} = 0.272.
$$
The floor $c_1 = 30$ is therefore a **27.2% purity requirement** — and since
the marginal prevalence of $Q$ is $\pi = 0.3446$ (34.5%), even a rule
carrying no information
about $Q$ has a true effect above the floor. This single fact organizes much
of what the simulations observe on this pathway, and the worked example
returns to it.

# Part I — The certified layer

## Tier A: algebraic closed forms (exactly verifiable)

**Proposition 1 (mixture identity; the content of CHECK A).** For every
$H \subseteq \mathcal{S}$ with $|H| > 0$,

$$
\beta(H)
\;=\; \frac{|H \cap Q|\,\tau_Q + |H \cap Q^c|\,\tau_{Q^c}}{|H|}
\;=\; \tau_{Q^c} + \beta_{\mathrm{int}}\, P(Q \mid H).
$$

*Proof, in four steps, naming the assumption each step consumes.*

**Step 1 (linearity of the effect scale $\Rightarrow$ collapsibility).**
Write $\mu_w(x) = \mathbb E[Y(w) \mid X = x]$ for the DGM's stored arm
surfaces and $\tau(x) = \mu_1(x) - \mu_0(x)$ for the row-level effect. By
definition the region effect is a difference of region-averaged arm means,
and because a difference is linear, averaging over $H$ commutes with
forming the effect:
$$
\beta(H)
= \frac{1}{|H|}\sum_{i \in H} \mu_1(X_i)
- \frac{1}{|H|}\sum_{i \in H} \mu_0(X_i)
= \frac{1}{|H|}\sum_{i \in H} \tau(X_i)
= \mathbb E\big[\tau(X) \mid X \in H\big].
$$
This is collapsibility. The only property used is that the effect measure
is a *difference of means* — the identity (collapsible) scale — and the
step therefore holds for **every** region: no rectangle shape, no
covariate independence, no restriction on how $H$ was chosen. It is
exactly this step that fails on nonlinear scales (log-odds, log-hazard),
which is why Tier A degrades to Tier B there.

**Step 2 (randomization $\Rightarrow$ the estimand is what the estimator
targets).** Under randomization, $W \perp (Y(0), Y(1), X)$, and since
$\{X \in H\}$ is an event determined by $X$ alone, the independence
survives the conditioning: $W \perp (Y(0), Y(1)) \mid X \in H$. Hence
$\mathbb E[Y \mid X \in H, W = w] = \mathbb E[\mu_w(X) \mid X \in H]$
for $w = 0, 1$, so the within-region difference in arm means — the
estimator of §2 — targets $\beta(H)$, up to the finite-sample imbalance
terms made exact in Part II, Level 2. This step is a design property,
agnostic to the outcome type, and it is consumed here only to connect the
algebraic estimand to the fitted quantity; the identity of Step 1 itself
is pure arithmetic on `df_super`. It is the region-restricted form of
the randomization assumption and unbiasedness theorem of [W] §1.1
(eq. 1.6, Theorem 1.1); see @sec-refwager-rct.

**Step 3 (two-valued CATE $\Rightarrow$ two-point mixture).** The fixture
sets $\tau(x) = \tau_{Q^c} + \beta_{\mathrm{int}}\,\mathbb 1\{x \in Q\}$,
so $\tau$ takes the value $\tau_Q = \tau_{Q^c} + \beta_{\mathrm{int}}$ on
$H \cap Q$ and $\tau_{Q^c}$ on $H \cap Q^c$. The Step 1 sum then splits
on one overlap count:
$$
\sum_{i \in H} \tau(X_i)
= |H \cap Q|\,\tau_Q + |H \cap Q^c|\,\tau_{Q^c},
$$
which is the first display; dividing by $|H|$ and writing
$P(Q \mid H) = |H \cap Q|/|H|$ (the purity of §2, with
$|H \cap Q^c|/|H| = 1 - P(Q \mid H)$) gives the second,
$\beta(H) = \tau_{Q^c} + \beta_{\mathrm{int}}\, P(Q \mid H)$.

**Step 4 (ledger).** Linearity bought Step 1 and is the piece lost on the
OR/HR scales; randomization bought Step 2 and transfers to every outcome
type; two-valuedness bought Step 3 and is a fixture choice — under a
general $\tau$, $\beta(H) = \mathbb E[\tau(X) \mid H]$ still holds
exactly, and only the reduction to a single overlap count is lost.
$\square$

The force of Proposition 1 is its quantifier: it holds for **every** rule,
including every data-selected $\hat H$, so the scoring target behind every
bias and coverage figure reduces to counting on a fixed finite set —
machine-exact. In the committed runs, CHECK A recomputed $\beta(\hat H)$
this way for each realized rule and held at $1.421\times 10^{-14}$ across
601–912 distinct rules per cell.

**Corollaries.** (i) $\beta(H^c)$ by the same mixture on the complement.
(ii) The ITT chain: $\beta(\mathcal{S}) = \tau_{Q^c} +
\beta_{\mathrm{int}}\pi = -30.992$ raw, which the bundles reproduce
as the complement effect on undetected rows. (iii) Disjointness:
$H \cap Q = \varnothing \Rightarrow \beta(H) = \tau_{Q^c}$ — observed
verbatim when a replicate selected a region disjoint from $Q$ and scored
exactly $26.255$ oriented. (iv) Calibration (CHECK B):
$\theta^\dagger(Q) = \beta(Q) = \tau_Q$ identically. (v) Classification
identities: with $s = |\hat H \cap Q|$, $\mathrm{sens} = s/|Q|$,
$\mathrm{PPV} = s/|\hat H|$, and on the complement
$\mathrm{spec} = |\hat H^c \cap Q^c|/|Q^c|$,
$\mathrm{NPV} = |\hat H^c \cap Q^c|/|\hat H^c|$; hence $\mathrm{PPV} \equiv \mathrm{sens}\cdot
|Q|/|\hat H|$ — the reported metrics carry one degree of freedom given the
sizes, an exact internal consistency relation.

```{r itt-live}
pfile <- "mr_sweep_md_harm/md_harm_s50_pilot/fs_md_harm_n1000_res.rds"
if (file.exists(pfile)) {
  b <- readRDS(pfile); tr <- b$truth; r <- b$results
  pred_itt <- tr$effect_Qc + tr$beta_inter * tr$prevalence_Q
  obs_itt  <- unique(round(r$betaHhat_Hc[r$detected %in% 0L |
                                         is.na(r$detected)], 9))
  stopifnot(length(obs_itt) == 1L, abs(pred_itt - obs_itt) < 1e-8)
  cat(sprintf("ITT chain: tau_Qc + beta_int*pi = %.3f; bundle records %.3f; |difference| = %.1e\n",
              pred_itt, obs_itt, abs(pred_itt - obs_itt)))
}
```

(The chunk is guarded: the `mr_sweep_md_harm` bundles are summary-era and
machine-local; where absent, the chunk skips silently and the committed
record above stands.)

## Tier B: exact by computation, not by formula

Quantities defined on $\mathcal{S}$ for a realized object are computable
with zero Monte Carlo error by finite summation even when no algebraic form
exists. First, in the binary/OR port the mixture identity fails because
averaging does not commute with a nonlinear scale: the log-odds of an
average risk is not the average of the log-odds (for a GLM link $\kappa$,
$\kappa^{-1}$ of a mixture is not the mixture of $\kappa^{-1}$'s) — yet the
exact OR over $H$ remains a deterministic function of the DGM's probability
surfaces evaluated on $\mathcal{S}$: Tier A degrades to Tier B, the
reference stays exact (the risk-difference scale retains full Tier A).
Second, the covariance structure of the candidate estimators: conditional
on a realized family and sample, the difference-in-means estimators over
regions $H_j, H_k$ with arm-wise counts $n_{j,a}$ satisfy, under
homoscedastic $\varepsilon$,

$$
\operatorname{Cov}\!\big(\hat\beta_j, \hat\beta_k\big)
\;=\; \sigma^2 \sum_{a\in\{0,1\}}
\frac{\,n_{jk,a}\,}{\,n_{j,a}\, n_{k,a}\,},
\qquad n_{jk,a} = |H_j \cap H_k \cap \{A=a\}|,
$$

an exact overlap computation; under balanced arms the correlation is the
overlap fraction $P(H_j \cap H_k)/\sqrt{P(H_j)P(H_k)}$, free of $\sigma$;
its diagonal is the classical Neyman/difference-in-means sampling
variance of [W] §1.1, restricted to a region (@sec-refwager-rct).
The caveat is stated rather than hidden: family membership and quantile
cuts are data-dependent, so this exactness is conditional on the realized
family.

## Tier C: computational identities

Dual-path equalities (two independent computations of one quantity must
agree exactly — CHECK C), the partition $|H| + |H^c| = N$ on every row, and
bit-level reproducibility (a replicate is a deterministic function of
(seed, RNG kind), so the standard is equality, not tolerance). These verify
the *implementation*, not the estimand, and their exactness makes failures
self-diagnosing: CHECK C's two quantities disagreeing by
$1.98\times 10^{2}$ while summing to $2.7\times 10^{-13}$ is uniquely the
signature of a sign flip — a tolerance-based check could only have said
"off."

# Part II — From identities to approximations: the ladder {#sec-ladder}

The connective structure of the analytic layer is a ladder of six levels.
At each level, exactly one new source of error enters; everything below it
remains exact. This ordering is what converts "Gaussian approximation" from
a label into an audited chain: when a computed rate and a measured rate
disagree, the ladder localizes which rung the disagreement lives on.

## Level 0–1: identities and exact computation

Tier A supplies the mean structure exactly (mixture values for every rule);
Tier B supplies the covariance structure exactly (overlap counts) and exact
references where formulas fail. Error: none.

## Level 2: the design-conditional law is an identity, not an approximation

Fix a candidate family $\mathcal G = \{g_1, \dots, g_M\}$ and a design
$D = \{(X_i, A_i)\}$. The estimator vector is linear in $Y$,
$\hat\beta = C Y$ with $C$ the $M \times n$ coefficient matrix
($c_i(g) = \mathbb 1\{i \in g\}(A_i/n_{g,1} - (1-A_i)/n_{g,0})$). Under
Gaussian noise with known $\sigma^2$,

$$
\hat\beta \mid D \;\sim\; N_M\!\big(m(D),\ \sigma^2 C C^\top\big)
$$

**exactly** — a distributional identity, not a limit theorem. The
conditional mean decomposes linearly in the DGM effect parameters,
$m(D) = a_0 + \tau_{Q^c}\, a_1 + \beta_{\mathrm{int}}\, a_2$ (prognostic
imbalance, treatment, and realized-$P(Q \mid g, \text{arm } 1)$
components), so null, borderline, and alternative scenarios share one
design geometry; and $\sigma^2 C C^\top$ coincides with the Tier B overlap
formula (verified identical at $1.9\times 10^{-15}$). Declaration and
selection events are then exact probabilities of a known law, and the only
error at this level is numerical quadrature: `mvtnorm::pmvnorm`'s
Genz–Bretz quasi–Monte Carlo, with a reported error estimate
($3\times 10^{-8}$ observed at correlation $0.98$) and a stated requirement
that tail rates use `abseps` $\le 10^{-5}$ (the default $10^{-3}$ is 40%+
relative error at $P \approx 0.002$). Exact duplicates make the law
singular; de-duplicate the family first.

## Level 3: CLT and plug-in — where estimation error enters, and where it cannot

Two departures from Level 2, each measured. **Non-Gaussian noise:** the
same MVN law becomes a CLT approximation for the weighted sums $CY$;
accuracy is governed by the smallest arm counts and the noise shape, and
was measured at $0.016$ absolute rate error with minimum arm size 10 and
noise skewness $\approx 6.2$ — the linear estimator's averaging does the
work. **Estimated variance:** because the procedure's events are defined on
point estimates, $\widehat\Sigma$ enters only the *computed probabilities*,
never the events. The single-dataset plug-in (per-candidate `lm` standard
errors with design-known overlap correlations) displaced the family rate by
$0.012$. One structural observation connects this level to the next: the
per-subgroup `lm` residual variance absorbs the within-region prognostic
and CATE spread (measured inflation factors $0.94$–$1.09$ over
$\sigma$-only truth), so the analyst's plug-in is a nearly unbiased
estimate of the *population-level* effective variance of Level 4 rather
than the conditional $\sigma^2$ — plug-in computations approximate the
unconditional law, known-$\sigma$ computations the design-conditional one
(the finite-population face of the variance-absorption phenomenon that
motivates regression adjustment, [W] §1.2; @sec-refwager-adj).
This has a practical corollary used in the worked example: **anchoring the
scale to a fitted standard error already incorporates the Level 4 variance
correction**, leaving only the covariance refinement below.

## Level 4: population averaging — the additive covariance decomposition {#sec-additive}

Averaging over designs, the means converge to the mixture values. The
variances and covariances require care, and this is where a purely
multiplicative repair fails. Write, for subject $i$ in region $g$, the
centered prognostic and CATE contributions

$$
u^{(g)}_i = \mu_0(X_i) - \mathbb E[\mu_0 \mid g], \qquad
t^{(g)}_i = \beta_{\mathrm{int}}\big(\mathbb 1\{X_i \in Q\} - P(Q \mid g)\big),
$$

so that, to leading order, the estimator's deviation from its target is a
sum of independent per-subject influence contributions
$\hat\beta(g) - \beta(g) \approx \sum_i \psi_i(g)$ with

$$
\psi_i(g) = \frac{2}{nP(g)}\,\mathbb 1\{i \in g\}
\Big[A_i\big(u^{(g)}_i + t^{(g)}_i + \varepsilon_i\big)
   - (1 - A_i)\big(u^{(g)}_i + \varepsilon_i\big)\Big].
$$

Taking expectations over $(A, \varepsilon)$ with $A \sim
\mathrm{Bern}(1/2)$ independent of $X$ yields the **additive population
covariance**: for any pair $(g, g')$,

$$
\Sigma_{\mathrm{pop}}(g, g')
= \frac{4}{n}\cdot\frac{P(g \cap g')}{P(g)P(g')}
\left(\sigma^2 \;+\;
\mathbb E\!\left[\, u^{(g)} u^{(g')}
+ \tfrac{1}{2}\big(u^{(g)} t^{(g')} + t^{(g)} u^{(g')} + t^{(g)} t^{(g')}\big)
\;\middle|\; g \cap g'\right]\right),
$$

with $\Sigma_{\mathrm{pop}}(g, g') = 0$ exactly for disjoint pairs. Every
term is enumerable on `df_super`. Three structural facts follow. (i) The
diagonal reproduces the variance correction
$\sigma^2 + \operatorname{Var}(\mu_0 \mid g) +
\operatorname{Var}(\tau \mid g)/2 + \operatorname{Cov}(\mu_0, \tau \mid g)$
— verified identical at $5.9\times 10^{-16}$. (ii) $\Sigma_{\mathrm{pop}}$
is a Gram matrix of the influence functions $\psi(g)$ in $L^2$, hence
positive semi-definite; near-duplicate candidates place it close to
singular, so numerical evaluation uses an eigenvalue floor. (iii) The
off-diagonal is **not** the $\sigma$-overlap correlation applied to
corrected variances: the imbalance components covary through their own
intersection moments. This is precisely the mechanism behind the borderline
residual — for the engineered near-duplicate (no prognostic variation along
its differing coordinate), the additive law reduces
$\operatorname{corr}(Q, \mathrm{dup})$ from the $\sigma$-overlap value
$0.9747$ to $0.9385$, decorrelating exactly where both candidates sit at
the threshold.

**The classical identity behind the diagonal, and what efficiency would
buy.** The diagonal of $\Sigma_{\mathrm{pop}}$ is not merely analogous to
classical theory — it *is* the Neyman superpopulation variance of [W]
§1.1, restricted to the region at $e = 1/2$: with
$\mathrm{Var}[Y(1)\mid g] = \sigma^2 + \mathrm{Var}(\mu_0+\tau\mid g)$
and $\mathrm{Var}[Y(0)\mid g] = \sigma^2 + \mathrm{Var}(\mu_0\mid g)$,
$$
V_{\mathrm{DM}}(g)
= 2\,\mathrm{Var}[Y(1)\mid g] + 2\,\mathrm{Var}[Y(0)\mid g]
= 4\Big(\sigma^2 + \mathrm{Var}(\mu_0\mid g)
+ \tfrac12\mathrm{Var}(\tau\mid g) + \mathrm{Cov}(\mu_0,\tau\mid g)\Big),
$$
exactly the bracket above. The semiparametric efficiency bound of [W]
§3.2 (eq. 3.20; Theorems 3.2 and 3.4), specialized to $e = 1/2$ and
homoscedastic noise within $g$, is
$V^\ast(g) = 4\sigma^2 + \mathrm{Var}(\tau\mid g)$, and the excess of
the unadjusted estimator over the bound has an exact closed form:
$$
V_{\mathrm{DM}}(g) - V^\ast(g)
= 4\,\mathrm{Var}(\mu_0\mid g) + \mathrm{Var}(\tau\mid g)
+ 4\,\mathrm{Cov}(\mu_0,\tau\mid g)
= \mathrm{Var}\big(2\mu_0 + \tau \,\big|\, g\big) \;\ge\; 0,
$$
the adjustable prognostic component expressed as a single variance. The
package's unadjusted within-region fit sits at $V_{\mathrm{DM}}$;
Appendix B (@sec-refwager-eff) states the bound and the AIPW estimator
that attains it.

**Verification** (`dev/detection-theory-continuous/check_additive_covariance_md.R`,
sandbox R 4.3.3/mvtnorm 1.2.4, seed 46, fresh population Monte Carlo,
$n_{\mathrm{rep}} = 6000$): at the adversarial borderline, MC $0.5580$
(SE $0.0064$) against additive computation $0.5568$ — gap $-0.0012$,
$0.2$ MC standard errors, where the naive and multiplicatively corrected
computations both sat at $0.536$ ($\approx 3.4$ SEs off); at the
alternative, MC $0.8507$ against additive $0.8511$ ($0.1$ SEs). The
population layer is therefore exact to leading order with all terms
enumerable; what remains outside it is second order — the $O(1/n)$
ratio-estimator terms from random arm counts and the non-normality of
small binomial pieces.

## Level 5: idealization — the gaps between the computed procedure and forestsearch

The computed functionals idealize the procedure in four named ways, each
with a diagnostic signature rather than a hidden discrepancy. **Fixed
versus regenerated family:** the computation conditions on an enumerated
family; the real identifier regenerates candidates per replicate from
data-dependent quantile cuts under `n.min` viability. Growth of a fitted
$L_{\mathrm{eff}}$ with $N$, and a measured sensitivity decline against a
fixed-family law that concentrates on maximal-mean candidates, are
regeneration signatures. **Single split versus
$\hat p_{\mathrm{cons}}$:** the computed consistency rung is the idealized
two-way split; the real screen aggregates hundreds of random splits into
$\hat p_{\mathrm{cons}}$ and gates on $\hat p_{\mathrm{cons}} \ge$ a
threshold. The single-split event and the multi-split gate bracket one
another in a definite direction discussed in the worked example; the exact
composition uses the stability closed form and remains a deferred rung.
**Argmax scope:** the computed selection is argmax-over-all; the procedure
takes the maximum over *admitted* candidates under `sg_focus` — the same
difference transform on a subset. **Nomination:** GRF/LASSO screening into
the family is unmodeled — its target is the CATE in the R-learner frame
of [W] §4.1–4.2, instantiated by the causal forest
(@sec-refwager-cate) — exactly as in the April $L_{\mathrm{eff}}$
document.

## Level 6: MR-specific — the fidelity benchmark and the empirical core

MR's per-draw re-selection ranks by `maxcons` on unrounded `zcons`
(`R/fs_mr_inference.R:169`) — a different statistic than the realized
selection. Both rankings are functionals of the same Gaussian object:
effect-ranking uses the identity transform; consistency-ranking uses, to
first order with design-known SEs, the diagonal rescaling
$D = \mathrm{diag}(1/\mathrm{SE}_m)$, so the zcons-ranked argmax is the
argmax of $D\hat\beta$ under the same difference-transform recipe.
Consistency-ranking weights small-SE (large) regions — the computed
counterpart of the measured $2.46\times$ ($n = 1000$) to $10.29\times$
($n = 4000$) winner-size inflation — and the two computed selection laws
give the semi-analytic benchmark for the fidelity fix, including the
falsifiable prediction that simulated winner sizes collapse to the realized
range, before any grid re-run. What stays empirical, by construction: MR's
*sufficiency* — the residual plateau near 37 and the coverage decay
$0.957 \to 0.731$ are measurements against certified references, and the
resolved mechanism (`REPORT_mr_mechanism.md`, `5b4efb4`) is the fidelity
gap, not linearization failure. The correction *scale* is derivable
($O(n^{-1/2})$; measured decay $n^{-0.60}$).

## The error ordering at a glance

| Level | New error source | Enters | Measured magnitude | Control |
|---|---|---|---|---|
| 0–1 | none | — | $10^{-14}$–$10^{-15}$ | machine |
| 2 | quadrature | computed probabilities | $3\times 10^{-8}$–$10^{-5}$ | `abseps` |
| 3 | CLT; plug-in $\widehat\Sigma$ | law shape; scale | $0.016$; $0.012$ | arm sizes; pooling |
| 4 | leading-order influence | population law | $0.001$ (additive); second order remains | enumerated moments |
| 5 | idealization | procedure fidelity | regeneration signatures on record | conditioning; deferred rungs |
| 6 | ranking statistic; MR sufficiency | re-selection; correction | winner-size $2.46$–$10.29\times$; residual $\sim 37$ | benchmark; empirical |

# Part III — Prediction recipes {#sec-recipes}

## What exactly is being predicted {#sec-what}

Every recipe below computes a probability or expectation of the **candidate
effect process** — the joint Gaussian law of the within-region estimators
over a fixed family — and each corresponds to a specific quantity a
simulation report tabulates. Before the individual recipes, the full map:

| Recipe | Input required | Output (the predicted quantity) | Compares to (simulation column) | Exactness |
|---|---|---|---|---|
| Marginal clearance | $\beta(g)$, $\sigma_g(n)$, $c_1$ | $P(\hat\beta(g) \ge c_1)$ for one fixed $g$ | per-region clearance counts (e.g., how often $Q$'s refit exceeds the floor) | Level 2–4 |
| Family declaration (Rung A) | means, $\Sigma$, $c_1$ | $P(\max_m \hat\beta_m \ge c_1)$ | detection rate if the procedure screened only | Level 2–4 |
| Split-half consistency (Rung B) | means, $\Sigma$, $c_1$, $c_2$ | per-candidate $P_1$ and $P(\text{any pass})$ | `DETECTED` share of the status table | Level 2–4 + single-split idealization |
| Selection distribution (Rung C) | means, $\Sigma$, gates, ranking rule | $P(\hat H = g_m)$ for each $m$ | realized-rule frequency table (`sg_def`); covariate-involvement shares | Levels 2–5 |
| Winner size | selection distribution, $P(g)$ | $\mathbb E[\,|\hat H|\,]$ | `n_harm` mean (identified-subgroup-sizes callout) | Levels 2–5 |
| Classification | selection distribution, overlaps with $Q$ | expected sens/spec/PPV/NPV | identification table | Levels 2–5 |
| Conditional estimand | selection distribution, mixture means | $\mathbb E[\beta(\hat H)]$ | `betaHhat_H` mean (oriented) | Levels 2–5 |
| Naive selection bias | selection distribution, winner noise | $\mathbb E[\hat\beta(\hat H) - \beta(\hat H)]$ | naive row, "vs $\beta(\hat H)$" bias column | Levels 2–5 |
| Coverage (conditional) | $(b, w, \sigma)$ | interval coverage given bias and width | coverage columns, given measured $(b, w)$ | closed form |
| Implied $L_{\mathrm{eff}}$ | null-family law | effective independent-candidate count | null-cell detection rate | Level 2–4 |
| Design calculators | any recipe, inverted | $n$ (or floors) achieving a target rate | — (planning, not comparison) | as the recipe |

(The Rung A/B/C labels in the table are the names these functionals carry
in the verification artifacts under `dev/detection-theory-continuous/`,
kept for traceability.) Two global caveats attach to every row. First, the family is
**conditioned on**: the recipes take $\mathcal G$ as given, while
forestsearch regenerates it per replicate (Level 5) — so fixed-family
predictions are structural benchmarks for the real design, exact for the
scenario they define, and become replicate-faithful when fed a realized
family (@sec-realized). Second, the consistency gate is idealized as a
single split unless the $\hat p_{\mathrm{cons}}$ composition rung is added.

## Marginal floor clearance of a fixed region

In words: for one pre-specified region, the probability that its fitted
effect clears the screening floor is a single normal tail probability,
because the mean is the mixture value and the scale is the Tier B standard
error. $P\big(\hat\beta(H) \ge c_1\big) = \bar\Phi\big((c_1 -
\beta(H))/\sigma_H(n)\big)$, $\beta(H)$ the oriented mixture value of
§2. The measured anchor (O1: the committed
50-replicate overnight anchor run):

```{r clearance-example}
z <- (40 - 30) / 13.6786          # oriented truth, floor, measured SE at n=1000
pred <- pnorm(z)
cat(sprintf("Marginal clearance of Q at n = 1000: Phi(%.3f) = %.3f;\n", z, pred),
    "measured (overnight O1): 36/50 = 0.720, MC SE 0.064 -- consistent.\n")
```

The marginal recipe alone under-predicts *family* detection, and the gap is
the point: at $n = 500$, $\sigma_Q$ scales to $13.679\sqrt 2 = 19.34$,
giving marginal clearance $\Phi(10/19.34) = 0.697$ for $Q$ — against a
measured family detection rate of $997/1000$. The difference is candidate
competition (Level 2, computed exactly by the next recipe) plus family
regeneration (Level 5, isolated).

## Family declaration (screening exceedance)

In words: the probability that *some* candidate's fitted effect clears the
floor. Because the joint law is known, this is one rectangle probability:
$P(\max_m \hat\beta_m \ge c_1) = 1 - \Phi_M(c_1\mathbf 1;\ m, \Sigma)$,
with $\Phi_M$ the $M$-variate normal distribution function.
Verified design-conditionally at $0.8297/0.5281/0.0262$ computed against
data-level Monte Carlo $0.8243/0.5400/0.0288$ (alternative / borderline /
null).

## Split-half consistency

In words: the idealized consistency gate splits the sample in two and asks
that the candidate clear the floor on average and a lower floor on each
half, with $W_1, W_2$ the candidate's fitted effects on the two halves. Per
candidate the passing probability is a one-dimensional integral,
$$
P_1 \;=\; \int_{c_2}^{\infty} \phi_{m_1, s_1}(w)\,
\bar\Phi_{m_2, s_2}\!\big(\max(c_2,\ 2c_1 - w)\big)\, dw,
$$
the Leon et al. (2024, SIM §2.1) object with $(m_j, s_j)$ that candidate's
mean and standard deviation on half $j$ — verified to
$7.5\times 10^{-4}$; the family rate ("some candidate
passes") is Gaussian-level Monte Carlo on the exact $2M$-block law,
verified $0.8206$ vs $0.8187$. The real gate replaces the single split by
$\hat p_{\mathrm{cons}} \ge$ threshold over hundreds of splits; the
relation between the two is taken up concretely in the worked example.

## Selection distribution and its consequences

In words: not just *whether* something is declared but *which* candidate
wins. For the effect-ranked argmax,
$$
P\big(\hat H = g_m,\ \text{declared}\big)
= P\big(\hat\beta_m \ge c_1,\ \hat\beta_m - \hat\beta_j \ge 0\ \forall j\big),
$$
a rectangle in the difference transform — one `pmvnorm` call per candidate,
with the internal identity
$\sum_m P(\hat H = g_m, \text{declared}) = P(\text{family declaration})$
verified to $3\times 10^{-7}$. Gated rules (such as maxeffCons: argmax of
effect among consistency-passers) are the same functional with the gate
event intersected, evaluated by Gaussian-level Monte Carlo. Selection
probabilities are what per-candidate clearances are not, and the
distinction drives every downstream prediction: winner size, expected
$\beta(\hat H)$, classification metrics, and the naive-bias forecast (the
selection-noise functional $\mathbb E[\text{noise of the argmax}]$ — the
quantity `design-checks/oracle_design.R` already evaluates on the same
law). The zcons-ranked variant of Level 6 uses the same recipe under the
diagonal rescaling. The formal relative of this layer in the causal
inference literature is policy learning ([W] Ch. 5); the estimand
contrast and the complexity/regret kinship are set out in
@sec-refwager-policy.

## Coverage, composed

Conditionally on $(b, w, \sigma)$, coverage is
$\Phi\big((w/2 - b)/\sigma\big) - \Phi\big((-w/2 - b)/\sigma\big)$,
demonstrated against the grid in the simulation summary. Unconditionally it
composes over the selection law,
$\sum_m P(\hat H = g_m)\,\mathrm{cov}_m$ — semi-analytic given MR's
realized $(b_m, w_m)$, which remain empirical inputs (Level 6).

## Implied $L_{\mathrm{eff}}$ as an output

The family null rate from the joint law yields an implied count directly;
with prevalence-heterogeneous candidate variances a single-$P_1$
normalization is ill-defined in the deep tail, and the stable readout is
against $\max_m P_1$,
$L_{\mathrm{eff}} = \log(1 - P_{\mathrm{fam}})/\log(1 - \max_m P_1)$:
$2.33, 1.80, 1.08$ at $N = 300, 600, 1200$ in the
verification family, with the family/union-bound ratio ($0.671 \to 0.989$)
exhibiting the correlation discount. Growth of a fitted
$L_{\mathrm{eff}}$ with $N$ is a Level 5 regeneration phenomenon.
$L_{\mathrm{eff}}$ plays the role the policy-class complexity term
plays in empirical-welfare-maximization regret bounds ([W] §5.2;
@sec-refwager-policy) — the same object seen from the exact-law side.

## The floors and family declaration: the single-candidate regime {#sec-floors}

Before any competition enters, the cleanest question is the one a design
must answer first: **with a single candidate — the designed region $Q$
itself, well separated, no family — how do the screening floor $c_1$ and
the consistency floor $c_2$ set the detection rate, and how does that
rate depend on the ITT, $Q$, and $Q^c$ effects?** Everything here is
closed-form or a one-dimensional integral, and the three-candidate
warm-up below is this section plus competition.

**How the triple enters.** The design's effect geometry is pinned by
three numbers — the prevalence $\pi$ and any two of
$(\tau_Q, \tau_{Q^c}, \mathrm{ITT})$ — through the mixture algebra of
Proposition 1:
$$
\mathrm{ITT} = \tau_{Q^c} + \beta_{\mathrm{int}}\,\pi,
\qquad
\tau_Q = \tau_{Q^c} + \frac{\mathrm{ITT} - \tau_{Q^c}}{\pi}.
$$
At the fixture triple ($\pi = 0.3446$, $\tau_{Q^c} = 26.255$,
$\beta_{\mathrm{int}} = 13.745$): $\tau_Q = 40$ and
$\mathrm{ITT} = 30.992$. Three *channels* follow, one per ideal
candidate, each with its mean fixed by the triple and its scale by the
prevalence rule $\sigma_g(n) = \sigma_Q(n)\sqrt{\pi_Q/\pi_g}$: the $Q$
channel (mean $\tau_Q$, the power channel); the **ITT channel** (the
whole-sample rule: mean $\mathrm{ITT}$, the smallest scale,
$\sigma = \sigma_Q\sqrt{\pi}$) — the canonical *uninformative* clearance
route, since any rule carrying no information about $Q$ has this mean;
and the **$Q^c$ channel** (mean $\tau_{Q^c}$, the mean *floor* over all
rules, attained at zero purity). The complement therefore matters even
when it is not a candidate: it sets both the lower end of every
candidate's possible mean and, through the identity, how much of the ITT
is "explained" by $Q$.

**The exact single-candidate rate.** With half-sample effects
$W_1, W_2 \stackrel{iid}{\sim} N\!\big(m,\ 2\sigma^2(n)\big)$ and the
idealized gate $\{\bar W \ge c_1,\ \min(W_1, W_2) \ge c_2\}$,
$$
P_1(c_1, c_2; m, \sigma)
= \int_{c_2}^{\infty} \phi_{m,\sigma\sqrt2}(w)\,
  \bar\Phi_{m,\sigma\sqrt2}\!\big(\max(c_2,\ 2c_1 - w)\big)\, dw,
$$
monotone decreasing in both floors, with three instructive limits:
$c_2 \to -\infty$ gives the screening-only rate
$\Phi\!\big((m - c_1)/\sigma\big)$; $c_1 \to -\infty$ gives the pure
replication rate $\bar\Phi\!\big((c_2 - m)/(\sigma\sqrt2)\big)^2$ —
which is also the **attainable ceiling at fixed $c_2$**: no screening
floor can push $P_1$ above it, and at the fixture ($c_2 = 10$, $m = 40$)
that ceiling is $0.883$, so a $95\%$ single-candidate target is
infeasible at these floors without lowering $c_2$; and
$c_2 \ge c_1$ makes screening redundant (the minimum constraint implies
the mean one). Between the limits, $c_2$ bites through a mechanism worth
naming: the event lost relative to screening alone is
$\{\bar W \ge c_1,\ \min W_j < c_2\}$ — **one low half forces the other
half to carry the sum**, so even a $c_2$ far below $m$ removes the
anti-correlated corner. At the fixture ($m = 40$,
$\sigma_Q(1000) = 13.679$, floors $30/10$) this costs about two
points: screening-only $0.768$ versus $P_1 = 0.745$ — the
quantitative content behind the "non-discriminating $c_2$" reading of
open question D2: non-discriminating *between candidates*, not free.

**Can $(c_1, c_2)$ be tuned to a target rate? Yes — with a feasibility
boundary once the ITT channel is watched.** For a single channel the
inversion is immediate (closed form at non-binding $c_2$,
$c_1 = m - z_{\rho}\,\sigma(n)$; one `uniroot` in general, done in the
chunk). The design question with teeth is *joint*: power $\ge \rho$ at
$Q$ while the uninformative ITT channel clears at most $\alpha$. To
first order (screening-dominant regime) that requires
$$
c_1 \le \tau_Q - z_\rho\,\sigma_Q(n)
\quad\text{and}\quad
c_1 \ge \mathrm{ITT} + z_{1-\alpha}\,\sigma_{\mathrm{ITT}}(n),
$$
which is feasible iff the **separation inequality** holds:
$$
\tau_Q - \mathrm{ITT} \;=\; \beta_{\mathrm{int}}\,(1 - \pi)
\;\ge\; z_\rho\,\sigma_Q(n) + z_{1-\alpha}\,\sigma_{\mathrm{ITT}}(n).
$$
The left side is pure geometry — the interaction times the complement
share — and the right side shrinks as $\sqrt{1000/n}$, giving the sample
size at which a $(\rho, \alpha)$ pair first becomes achievable. At the
fixture triple the numbers are sobering and worth stating plainly:
$(\rho, \alpha) = (0.90, 0.10)$ first becomes feasible near
$n \approx 9{,}500$; at $n = 1000$ the floors can *trade* the two
channels along the $c_1$ axis but cannot meet both — the borderline-null
regime of the manuscript's §7 anchors, derived rather than observed. The
$Q^c$ channel sits below both (mean $26.255$) and is the mild one at
these floors; $c_2$ moves all three channels together (multiplicatively,
through the replication factor) and is the secondary knob.

```{r floors-single}
#| fig-width: 7
#| fig-height: 5
# Exact single-candidate machinery, fixture triple, oriented scale.
detect1 <- function(c1, c2, m, s) {
  s2 <- s * sqrt(2)
  integrate(function(w) dnorm(w, m, s2) *
              pnorm(pmax(c2, 2*c1 - w), m, s2, lower.tail = FALSE),
            lower = c2, upper = Inf, rel.tol = 1e-10)$value
}
c1_for <- function(rho, c2, m, s) {        # exact inversion in c1
  ceiling <- pnorm(c2, m, s*sqrt(2), lower.tail = FALSE)^2
  if (rho >= ceiling) return(NA_real_)     # target above the c2 ceiling
  uniroot(function(c1) detect1(c1, c2, m, s) - rho,
          interval = c(m - 8*s, m + 8*s), tol = 1e-8)$root
}
piQ <- 0.3446; tQc <- 26.255235876; bint <- 13.744764124
tQ  <- tQc + bint; ITT <- tQc + bint*piQ
sQ1 <- 13.6786
chan <- data.frame(channel = c("Q (power)", "ITT (uninformative)",
                               "Qc (complement)"),
                   mean = c(tQ, ITT, tQc),
                   prevalence = c(piQ, 1, 1 - piQ))
chan$sigma_n1000 <- sQ1 * sqrt(piQ / chan$prevalence)
res <- do.call(rbind, lapply(c(500, 700, 1000), function(n) {
  sc <- sqrt(1000/n)
  data.frame(n = n, channel = chan$channel,
    screen_only = pnorm(30, chan$mean, chan$sigma_n1000*sc, lower.tail = FALSE),
    P1_c2_10 = vapply(seq_len(3), function(i)
      detect1(30, 10, chan$mean[i], chan$sigma_n1000[i]*sc), numeric(1)),
    P1_c2_25 = vapply(seq_len(3), function(i)
      detect1(30, 25, chan$mean[i], chan$sigma_n1000[i]*sc), numeric(1)))
}))
res[, 3:5] <- round(res[, 3:5], 3)
knitr::kable(res, row.names = FALSE, caption = paste(
  "The three ideal channels at floors (30, c2): screening-only vs the exact",
  "single-split rate. The c2 = 10 column quantifies the sum-forcing penalty;",
  "c2 = 25 shows the floor beginning to bind on its own."))
zr <- function(x) qnorm(x)
nstar <- function(rho, alpha)
  1000 * ((zr(rho)*sQ1 + zr(1-alpha)*sQ1*sqrt(piQ)) / (bint*(1-piQ)))^2
cat(sprintf("Separation: tau_Q - ITT = %.3f.  n* for (rho, alpha): (.90,.10) %s | (.80,.20) %s | (.95,.05) %s\n",
    bint*(1-piQ),
    format(round(nstar(.90,.10)), big.mark=","),
    format(round(nstar(.80,.20)), big.mark=","),
    format(round(nstar(.95,.05)), big.mark=",")))
tr <- data.frame(c1 = seq(28, 42, 2))
tr$power_Q   <- vapply(tr$c1, detect1, numeric(1), c2 = 10, m = tQ,  s = sQ1)
tr$ITT_chan  <- vapply(tr$c1, detect1, numeric(1), c2 = 10, m = ITT, s = sQ1*sqrt(piQ))
tr$Qc_chan   <- vapply(tr$c1, detect1, numeric(1), c2 = 10, m = tQc, s = sQ1*sqrt(piQ/(1-piQ)))
tr[, 2:4] <- round(tr[, 2:4], 3)
knitr::kable(tr, row.names = FALSE, caption = paste(
  "The trade-off at n = 1000, c2 = 10: sliding c1 exchanges power at Q",
  "against the ITT-channel clearance rate; no c1 meets (0.90, 0.10) jointly."))
inv <- vapply(c(.80, .90, .95), c1_for, numeric(1), c2 = 10, m = tQ, s = sQ1)
cat(sprintf("Exact inversions at n = 1000, c2 = 10: c1 for 80/90%% power at Q: %.2f / %.2f; 95%% infeasible (c2 ceiling %.3f)\n",
    inv[1], inv[2], pnorm(10, tQ, sQ1*sqrt(2), lower.tail = FALSE)^2))
g1 <- seq(22, 44, by = 0.5); g2 <- seq(0, 40, by = 0.5)
Z <- outer(g1, g2, Vectorize(function(a, b)
  if (b > a) NA_real_ else detect1(a, b, tQ, sQ1)))
contour(g1, g2, Z, levels = c(.5, .6, .7, .743, .8, .9, .95),
        xlab = "screening floor c1", ylab = "consistency floor c2",
        main = "Single-candidate detection P1(c1, c2) for Q at n = 1000")
points(30, 10, pch = 19); text(30, 10, "  fixture (30, 10)", adj = 0)
abline(0, 1, lty = 3)
```

Reading the outputs. The channel table is the section's core: at the
fixture floors the $Q$ channel detects at $0.745$ ($n = 1000$) while the
uninformative ITT channel *also* clears at $0.55$ — the two rates a
$(c_1, c_2)$ choice must trade, with $Q^c$ mild below both. The
trade-off table walks $c_1$ across the gap and shows the exchange rate
explicitly; the separation printout gives the $n$ at which chosen
$(\rho, \alpha)$ pairs become jointly feasible; and the contour displays
the full level sets of $P_1(c_1, c_2)$ with the fixture marked — flat in
$c_2$ well below the diagonal, bending as $c_2$ approaches $c_1$ where
replication takes over. Family declaration for a real family is this
section composed with competition: each member is one channel with its
purity-determined mean, and the family rate is the joint exceedance the
next recipes compute.

## A three-candidate warm-up, read line by line {#sec-warmup}

Before the full worked example, a minimal geometry that shows what the
recipes return and how to read them. Three candidates at the fixture's
means and measured scale: $Q$ itself; an overlapping proxy $P$
(prevalence $0.45$, sharing probability mass $0.28$ with $Q$, hence purity
$P(Q \mid P) = 0.28/0.45 = 0.622$); and a near-duplicate $D \subset Q$
(prevalence $0.31$, purity $1$). SEs are anchored to the measured
$\sigma_Q(n{=}1000) = 13.679$ and scaled by the balanced-arms prevalence
rule $\sigma_H = \sigma_Q\sqrt{\pi_Q/\pi_H}$; correlations are the
$\sigma$-overlap block (per @sec-ladder Level 3, the fitted-SE anchor
already carries the Level 4 variance component; the covariance refinement
of @sec-additive applies when DGM moments are in hand).

```{r worked-prediction}
set.seed(20260824)   # pins pmvnorm's quasi-MC randomization for the gap readout
library(mvtnorm)
tauQc <- 26.255235876; bint <- 13.744764124   # oriented
c1 <- 30; seQ <- 13.6786; piQ <- 0.3446
cand <- data.frame(
  name = c("Q", "P (overlap proxy)", "D (near-dup, subset of Q)"),
  pi   = c(piQ, 0.45, 0.31),
  piQg = c(1.00, 0.28/0.45, 1.00))          # P(Q | g)
cand$beta <- tauQc + bint * cand$piQg
cand$se   <- seQ * sqrt(piQ / cand$pi)
ov <- matrix(c(piQ, 0.28, 0.31,
               0.28, 0.45, 0.28*0.31/piQ,   # |P∩D| via D ⊂ Q composition
               0.31, 0.28*0.31/piQ, 0.31), 3, 3)
R  <- ov / sqrt(outer(cand$pi, cand$pi))
Sg <- R * outer(cand$se, cand$se)
p_marg <- pnorm(c1, cand$beta, cand$se, lower.tail = FALSE)
p_fam  <- 1 - pmvnorm(upper = rep(c1, 3), mean = cand$beta, sigma = Sg,
                      algorithm = GenzBretz(abseps = 1e-6))
sel <- vapply(1:3, function(m) {
  A <- rbind(diag(3)[m, ],
             do.call(rbind, lapply(setdiff(1:3, m),
                                   function(j) diag(3)[m, ] - diag(3)[j, ])))
  as.numeric(pmvnorm(lower = c(c1, 0, 0), upper = rep(Inf, 3),
                     mean = as.numeric(A %*% cand$beta),
                     sigma = A %*% Sg %*% t(A),
                     algorithm = GenzBretz(abseps = 1e-6)))
}, numeric(1))
out <- data.frame(candidate = cand$name,
                  beta_oriented = round(cand$beta, 3),
                  se = round(cand$se, 3),
                  marginal_clearance = round(p_marg, 4),
                  selection_prob = round(sel, 4))
knitr::kable(out, caption = sprintf(
  "Warm-up at n = 1000: family declaration %.4f; selection identity gap %.1e.",
  as.numeric(p_fam), abs(sum(sel) - as.numeric(p_fam))))
```

Reading the table. **Column `beta_oriented`** is the mixture mean: $Q$ and
$D$ both sit at $40$ (both have purity $1$), $P$ at $34.81$. **Column
`se`** is the prevalence-scaled measured scale: smaller regions are
noisier, so $D$ ($\pi = 0.31$) carries the largest SE. **Column
`marginal_clearance`** treats each candidate alone: $Q$ clears the floor
$76.8\%$ of the time, $D$ $75.6\%$, $P$ $65.6\%$. **Column
`selection_prob`** is the joint answer, and it is *not* proportional to the
marginals: $D$ takes the largest share ($0.360$) despite the smaller
marginal clearance, because at equal means the *noisier* candidate wins the
argmax more often — the Gaussian-level mechanism behind the observed
tendency of the identifier to return small, noise-favored regions. The
three selection probabilities sum to the family declaration rate ($0.8474$)
up to the printed quadrature gap — an internal consistency identity, not a
new assumption. Finally, family declaration ($0.847$) exceeds the best
marginal ($0.768$): competition helps *declaration* even as it dilutes any
single candidate's *selection*.

### Adding the consistency floor, and the ITT and $Q^c$ channels {#sec-warmup-floors}

The warm-up table above is screening-only. Two additions complete it in
the direction of @sec-floors: the consistency floor, and the two ideal
channels as explicit family members.

```{r warmup-floors}
set.seed(20260826)
P1_3 <- vapply(seq_len(3), function(m)
  detect1(c1, 10, cand$beta[m], cand$se[m]), numeric(1))
e3 <- eigen(2 * Sg, symmetric = TRUE)
rt3 <- e3$vectors %*% diag(sqrt(pmax(e3$values, 0)))
Rg <- 2e5
fam_split <- function(mu, root, Mloc, c2v) {
  V1 <- matrix(rnorm(Rg*Mloc), Rg, Mloc) %*% t(root) + rep(mu, each = Rg)
  V2 <- matrix(rnorm(Rg*Mloc), Rg, Mloc) %*% t(root) + rep(mu, each = Rg)
  mean(rowSums((V1 + V2 >= 2*c1) & (V1 >= c2v) & (V2 >= c2v)) > 0)
}
fam10 <- fam_split(cand$beta, rt3, 3, 10)
fam25 <- fam_split(cand$beta, rt3, 3, 25)
# Augment with the ITT rule S (prevalence 1) and the complement Qc
piQ <- 0.3446; sQ1 <- 13.6786
beta5 <- c(cand$beta, 30.992, 26.255)
prev5 <- c(cand$pi, 1, 1 - piQ)
se5   <- c(cand$se, sQ1*sqrt(piQ), sQ1*sqrt(piQ/(1-piQ)))
ov5 <- matrix(0, 5, 5)
ov5[1:3, 1:3] <- ov
ov5[4, ] <- ov5[, 4] <- prev5            # S overlaps everything at P(g)
ov5[5, 1:3] <- ov5[1:3, 5] <- c(0, 0.45 - 0.28, 0)   # Qc: none with Q, D
ov5[5, 4] <- ov5[4, 5] <- 1 - piQ; ov5[5, 5] <- 1 - piQ
diag(ov5)[1:3] <- cand$pi
R5 <- ov5 / sqrt(outer(prev5, prev5))
S5 <- R5 * outer(se5, se5)
e5 <- eigen(2 * S5, symmetric = TRUE)
rt5 <- e5$vectors %*% diag(sqrt(pmax(e5$values, 0)))
fam5_10 <- fam_split(beta5, rt5, 5, 10)
scr5 <- pnorm(c1, beta5, se5, lower.tail = FALSE)
P1_5 <- vapply(seq_len(5), function(m) detect1(c1, 10, beta5[m], se5[m]),
               numeric(1))
knitr::kable(data.frame(
  member = c(as.character(cand$name), "S (ITT rule)", "Qc (complement)"),
  beta = round(beta5, 3), se = round(se5, 3),
  screen_only = round(scr5, 3), P1_c1_30_c2_10 = round(P1_5, 3)),
  row.names = FALSE, caption =
  "The warm-up family with the two ideal channels added, at floors (30, 10).")
cat(sprintf("Family single-split declaration: 3-candidate %.4f (c2=10) / %.4f (c2=25); screening-only %.4f.\n",
            fam10, fam25, as.numeric(p_fam)))
cat(sprintf("5-member family with ITT and Qc channels, c2=10: %.4f.\n", fam5_10))
```

Reading the additions. The consistency floor at $c_2 = 10$ trims the
three-candidate family from the screening-only $0.847$ by the
sum-forcing mechanism of @sec-floors, and $c_2 = 25$ shows the floor
biting in earnest. Adding the two ideal channels raises family
declaration further — the ITT rule alone contributes a $0.55$-per-family
clearance route at these floors — which is the warm-up-scale
demonstration of the section's central point: **"overall detection" is a
property of the channel set, and the $(c_1, c_2)$ pair prices every
channel simultaneously through its purity-determined mean.** The
three-class structure (truth, proxy, near-duplicate) governs *which*
member is selected; the ITT and $Q^c$ channels govern how often
*something* clears when the informative members are absent or weak.

## Design calculators

The recipes invert. Sample size for a target marginal clearance of $Q$:

```{r design-calculator}
n_for <- function(target, se_n1000 = 13.6786, mu = 40, c1 = 30)
  1000 * (se_n1000 * qnorm(target) / (mu - c1))^2
cat(sprintf("n for %d%% marginal clearance of Q at floor 30: %s\n",
            c(80L, 90L, 95L),
            format(round(n_for(c(.80, .90, .95))), big.mark = ",")), sep = "")
```

Family-level and selection-level counterparts invert the corresponding
functionals by the same one-dimensional search; Monte Carlo error budgets
for grid design follow from computed rates rather than convention.

# A detailed worked example: the maxeffCons MR design at n = 500 {#sec-worked}

This section carries the machinery end to end against a specific committed
design:
`sim_fs_maxeffCons_mr_md40_knoise0_n500_batch_1_1000.qmd`. Every input is
taken from that document's own setup chunks; every output is stated
together with the report column it should be compared against.

## Step 1 — the design under study

From the driver's setup, verbatim: $n = 500$ per replicate; the committed
alternative cell (`target_md_harm = -40` raw, oriented $+40$ in $Q$);
floors `md_threshold = 30`, `md_consistency = 10`; consistency gate
$\hat p_{\mathrm{cons}} \ge 0.90$ over `fs.splits = 400` random splits;
`n.min = 60` (a $60/500 = 12\%$ prevalence floor on candidates);
`maxk = 2` (conjunctions of at most two conditions);
`sg_focus = "maxeffCons"` — in the driver's own words, "highest effect
meeting the consistency rate"; MR with 5000 draws; no added noise
covariates. The candidate cuts are the deliberate part of the design: J
$= 10$ quantile grids on `age` and `preanti` **replace** the default cuts,
so the true boundaries ($34$ and $744.5$) are *not representable*; the
driver documents the best achievable rule as
`age > 33 & preanti <= 677.9` at Jaccard $0.882$ against $Q$, and gives
the grid quantile positions (the $33$ cut at the $5/11 = 45.5\%$ age
quantile; the preanti grid $0, 0, 0, 0, 55.8, 214.1, 432, 677.9, 872.8,
1070.5$ at the $k/11$ quantiles, so $677.9$ sits at $72.7\%$). Here and
throughout, the **Jaccard index** of a rule $g$ against $Q$ is
$$
J(g, Q) \;=\; \frac{P(g \cap Q)}{P(g \cup Q)},
$$
the overlap as a fraction of the union — equal to $1$ exactly when
$g = Q$, and tied to the classification identities of Part I,
Corollary (v), by the exact relation
$1/J = 1/\mathrm{sens} + 1/\mathrm{PPV} - 1$: the driver's rule-recovery
metric is the harmonic composite of recovery and purity.

Its report tabulates, per batch: the status trichotomy
(`DETECTED / NO-DETECTION / CONFIG-ERROR`); identified-subgroup sizes
(`n_harm`); the realized-rule frequency table (`sg_def`) and
covariate-involvement shares; classification (sens/spec/PPV/NPV); and the
estimation table scoring oracle/naive/MR against the exact
$\beta(\hat H)$, the oracle refit, and the structural and fitted targets.
Those are exactly the quantities the recipes predict.

## Step 2 — what must be specified, and what the driver already supplies

The recipes need five inputs. The driver supplies four directly: the
oriented effects ($\tau_{Q^c} = +26.255$, $\beta_{\mathrm{int}} =
+13.745$, hence purity translation $P(Q \mid g) \ge 0.272$ for true
clearance — the driver's own "27.2%" remark); the sample size and floors;
the selection rule (maxeffCons); and the scale anchor (the committed
measured $\sigma_Q(n{=}1000) = 13.679$, which the fitted-SE observation
of Level 3 licenses as a population-effective scale, rescaled to $n = 500$
by $\sqrt 2$). The fifth input — **the candidate family and the joint
covariate law that determines every $P(g)$, $P(Q \mid g)$, and overlap** —
is the one forestsearch regenerates per replicate, so a prediction must
either condition on a realized family (@sec-realized) or construct a
scenario satisfying the fixed-family requirements. This example does the
latter, using only quantities the driver itself documents.

## Step 3 — constructing the scenario from documented anchors

Cross the two true-region axes at their *representable* resolutions. Age
takes three categories $\{\le 33,\ (33, 34],\ > 34\}$ — the family can cut
at $33$ (a grid cut) but not at $34$ (the true boundary, absent from the
grid). Preanti takes three categories
$\{\le 677.9,\ (677.9, 744.5],\ > 744.5\}$ — cut at $677.9$ representable,
true boundary $744.5$ not. Marginals come from the documented quantile
positions ($P(\text{age} > 33) = 1 - 5/11 = 0.545$;
$P(\text{age} > 34) = 0.50$, the median;
$P(\text{preanti} \le 677.9) = 8/11 = 0.727$;
$P(\text{preanti} \le 744.5) = 0.75$, the q75 default). The $3 \times 3$
joint below satisfies **all** the documented anchors simultaneously — the
four marginals, $P(Q) = 0.3446$, and Jaccard of the best representable
rule against $Q$ equal to $0.8820$ — and the chunk asserts each one. A
third, stylized covariate $V$ (prevalence $0.42$, independent) stands in
for the distractor pool; the driver's real distractors include `str2`,
which nests inside the true partner condition and is *not* independent —
the stylization is stated, not hidden. Candidates are all single
conditions and cross-variable pairs over the three representable cuts
(both sides), pruned at the $12\%$ `n.min` floor.

```{r worked-scenario}
library(mvtnorm)
set.seed(20260825)
tauQc <- 26.255235876; bint <- 13.744764124        # oriented
c1 <- 30; c2 <- 10; n_rep <- 500; pcons_gate <- 0.90
seQ1000 <- 13.6786; piQ_fix <- 0.3446
# 3x3 joint: rows age (<=33, (33,34], >34); cols preanti (<=677.9,
# (677.9,744.5], >744.5). Hand-solved to hit every documented anchor.
J3 <- matrix(c(0.3553, 0.0121, 0.0876,
               0.0360, 0.0020, 0.0070,
               0.3357, 0.0089, 0.1554), 3, 3, byrow = TRUE)
stopifnot(abs(sum(J3) - 1) < 1e-12,
          abs(rowSums(J3) - c(0.455, 0.045, 0.500)) < 1e-12,
          abs(colSums(J3) - c(0.727, 0.023, 0.250)) < 1e-12)
PQ <- J3[3, 1] + J3[3, 2]                    # Q = age>34 & pre<=744.5
stopifnot(abs(PQ - piQ_fix) < 1e-4)
gstar_P <- J3[2, 1] + J3[3, 1]               # g* = age>33 & pre<=677.9
jac <- J3[3, 1] / (gstar_P + PQ - J3[3, 1])
stopifnot(abs(jac - 0.882) < 1e-3)
cat(sprintf("Scenario anchors: P(Q) = %.4f; Jaccard(g*, Q) = %.4f (driver: 0.882)\n",
            PQ, jac))
# 18 cells: 9 (age x pre) x V in {1, 0}, V independent, P(V=1) = .42
pV <- 0.42
cells <- expand.grid(a = 1:3, p = 1:3, v = 1:0)
cells$prob <- J3[cbind(cells$a, cells$p)] * ifelse(cells$v == 1, pV, 1 - pV)
cells$inQ  <- cells$a == 3 & cells$p <= 2
# Candidate conditions on representable cuts (both sides) + V
conds <- list(`age>33` = cells$a >= 2, `age<=33` = cells$a == 1,
              `pre<=677.9` = cells$p == 1, `pre>677.9` = cells$p >= 2,
              `V=1` = cells$v == 1, `V=0` = cells$v == 0)
fam <- c(lapply(names(conds), function(nm) list(nm = nm, m = conds[[nm]])),
         unlist(lapply(list(c(1,3), c(1,4), c(2,3), c(2,4),
                            c(1,5), c(1,6), c(2,5), c(2,6),
                            c(3,5), c(3,6), c(4,5), c(4,6)),
   function(ix) list(list(nm = paste(names(conds)[ix], collapse = " & "),
                          m = conds[[ix[1]]] & conds[[ix[2]]]))),
   recursive = FALSE))
Pg  <- vapply(fam, function(f) sum(cells$prob[f$m]), numeric(1))
keep <- Pg >= 0.12                                  # n.min = 60 of 500
fam <- fam[keep]; Pg <- Pg[keep]; M <- length(fam)
lab <- vapply(fam, `[[`, character(1), "nm")
PQg <- vapply(fam, function(f) sum(cells$prob[f$m & cells$inQ]), numeric(1)) / Pg
beta_g <- tauQc + bint * PQg
se_g   <- seQ1000 * sqrt(2) * sqrt(piQ_fix / Pg)    # n = 500 anchor scaling
sens_g <- vapply(fam, function(f) sum(cells$prob[f$m & cells$inQ]), numeric(1)) / PQ
spec_g <- vapply(fam, function(f) 1 - sum(cells$prob[f$m & !cells$inQ]) / (1 - PQ),
                 numeric(1))
ovl <- outer(seq_len(M), seq_len(M), Vectorize(function(i, j)
  sum(cells$prob[fam[[i]]$m & fam[[j]]$m])))
Rho <- ovl / sqrt(outer(Pg, Pg))
Sg  <- Rho * outer(se_g, se_g)
tab3 <- data.frame(candidate = lab, prevalence = round(Pg, 4),
                   purity_PQg = round(PQg, 4), beta = round(beta_g, 2),
                   se_n500 = round(se_g, 1),
                   clears_floor = ifelse(beta_g >= c1, "yes", "no"))
knitr::kable(tab3[order(-tab3$beta), ], row.names = FALSE, caption = sprintf(
  "The scenario family (M = %d after the 12%% floor): mixture means, anchored scales, and the purity translation of the floor.", M))
```

Reading the family table. The purity column *is* the mean column in
disguise (Proposition 1), and the floor line falls at purity $0.2724$ —
below the marginal prevalence $0.3446$, so **nine of the sixteen
candidates have true effects above the floor, including the pure
distractors** $V{=}1$ and $V{=}0$ at $\beta = 30.99$. On this pathway the
screening floor does not separate signal from noise regions; it separates
regions *depleted* of $Q$ from everything else. The best representable
rule $g^\ast$ leads at $\beta = 38.67$ (purity $0.903$), $1.33$ units
below the unrepresentable truth — the price of the J-grid design,
quantified before any simulation.

## Step 4 — predictions, each mapped to its report column

The maxeffCons rule is: among candidates passing the consistency gate,
select the one with the largest fitted effect; declare `NO-DETECTION` if
none passes. The idealized computation below replaces the
$\hat p_{\mathrm{cons}} \ge 0.90$ multi-split gate with the single-split
event $\{W_1 + W_2 \ge 2c_1,\ \min(W_1, W_2) \ge c_2\}$ — $W_1, W_2$
again the half-sample fitted effects — and evaluates
everything on the known Gaussian law by Monte Carlo (Level 2 numerics;
$2 \times 10^5$ draws, so rate SEs are $\approx 0.001$). Direction of the
idealization, stated: with $c_2 = 10$ sitting far below every candidate
mean (the driver's open question D2 — the consistency floor is
non-discriminating on this cell), the per-split floor events are nearly
always satisfied and both gates reduce to variants of "the effect
replicates above threshold"; a demanding $\hat p_{\mathrm{cons}} \ge 0.90$
over 400 splits is *stricter* than one split, so at fixed family the
computed detection rate is an upper benchmark and the computed selection
law shifts mass toward stable (larger, higher-mean) candidates relative to
the real gate; the far larger real family pushes detection the other way
and dominates it here.

```{r worked-predictions}
Rdraw <- 2e5
# Complement pairs make the population law EXACTLY rank-deficient: for any
# cut, P(g+)c(g+) + P(g-)c(g-) is the full-sample coefficient vector, so the
# candidate covariance is a singular Gram matrix (the Level 4 PSD fact, made
# concrete). Sample by eigendecomposition; degenerate directions get zero
# variance, which enforces the linear constraints exactly.
eS <- eigen((Sg + t(Sg)) / 2, symmetric = TRUE)
root_half <- eS$vectors %*% diag(sqrt(pmax(2 * eS$values, 0)))  # cov = 2*Sg
W1 <- matrix(rnorm(Rdraw * M), Rdraw, M) %*% t(root_half) +
      rep(beta_g, each = Rdraw)
W2 <- matrix(rnorm(Rdraw * M), Rdraw, M) %*% t(root_half) +
      rep(beta_g, each = Rdraw)
Bhat <- (W1 + W2) / 2                   # full-sample effects, exact law N(beta, Sg)
pass <- (W1 + W2 >= 2 * c1) & (W1 >= c2) & (W2 >= c2)
det  <- rowSums(pass) > 0
Bmask <- Bhat; Bmask[!pass] <- -Inf
winner <- max.col(Bmask, ties.method = "first"); winner[!det] <- NA_integer_
P1 <- colMeans(pass)
p_sel <- tabulate(winner[det], M) / Rdraw
det_rate <- mean(det)
sel_c <- p_sel / det_rate                       # selection | detection
EnH   <- n_rep * sum(sel_c * Pg)                # expected |Hhat| | detection
Esens <- sum(sel_c * sens_g); Espec <- sum(sel_c * spec_g)
Eppv  <- sum(sel_c * PQg)
Enpv  <- sum(sel_c * vapply(seq_len(M), function(m) {
  pHc <- 1 - Pg[m]; (pHc - (PQ - PQg[m] * Pg[m])) / pHc }, numeric(1)))
EbetaH <- sum(sel_c * beta_g)
noise  <- Bhat - rep(beta_g, each = Rdraw)
Enaive_bias <- mean(noise[cbind(which(det), winner[det])])
mass_below <- sum(sel_c[beta_g < c1])          # selections on true-mean-below-floor rules
top <- order(-p_sel)[1:6]
knitr::kable(data.frame(candidate = lab[top],
                        P1_single_split = round(P1[top], 3),
                        P_selected = round(p_sel[top], 3),
                        sel_given_det = round(sel_c[top], 3)),
             row.names = FALSE, caption = sprintf(
  "Top of the predicted selection distribution (detection rate %.3f). Compare P_selected against the driver's realized-rule (sg_def) frequency table.",
  det_rate))
knitr::kable(data.frame(
  quantity = c("detection rate (DETECTED share)",
               "E[|Hhat|] given detection (n_harm mean)",
               "E[sens]", "E[spec]", "E[PPV]", "E[NPV]",
               "E[beta(Hhat)] oriented (betaHhat_H mean)",
               "E[naive - beta(Hhat)] (naive bias vs beta(Hhat))",
               "selection mass on rules with true mean below the floor"),
  predicted = round(c(det_rate, EnH, Esens, Espec, Eppv, Enpv,
                      EbetaH, Enaive_bias, mass_below), 3)),
  row.names = FALSE, caption =
  "Scenario predictions with their driver-report comparison targets in parentheses.")
```

Reading the predictions, line by line.

**Detection.** The computed rate is $0.965$: with nine candidate means
above the floor and half-sample SEs of $19$–$40$, *some* candidate clears
on nearly every replicate. The committed real-family measurement at
$n = 500$ is $997/1000$ — higher still, in exactly the direction the
family-size caveat below predicts — and the driver's own status table is
this prediction's comparison target. Detection on this cell is not
evidence of finding $Q$; the interesting prediction is *what* gets
selected.

**The selection distribution** is the substantive output, and it inverts
the marginal ordering. $g^\ast$ leads the gate ($P_1 = 0.63$, the largest)
yet takes only the second-largest selection share: maxeffCons is an
effect-ranked argmax over gate-passers, and *conditional on passing*, a
high-variance candidate's fitted effect is inflated far above its mean —
so noisy low-purity rules out-bid the leader whenever their noise draws
large. The aggregate diagnostic in the second table makes this exact: a
substantial fraction of all selections lands on rules whose **true** mean
sits below the floor — rule-level false-harm declarations, predicted from
the law, not discovered in a run. Each row of the selection table is
comparable rule for rule against the driver's `sg_def` frequency table,
and every share taken by a low-purity rule drags the classification and
estimand predictions below their $g^\ast$ ceilings.

**Winner size and classification.** $\mathbb E[\,|\hat H|\,]$ is the
selection-weighted prevalence times $n$ — compare to the `n_harm` mean in
the identified-subgroup-sizes callout — and expected sens/spec/PPV/NPV are
selection-weighted overlaps with $Q$, comparable to the identification
table. Because $g^\ast$ carries sens $= 0.974$ and PPV $= 0.903$, the
predictions sit high exactly to the extent selection concentrates on it;
the shortfall from those ceilings measures the noise-selection share.

**The conditional estimand and the naive bias.** $\mathbb E[\beta(\hat
H)]$ is the selection-weighted mixture mean (compare: oriented
`betaHhat_H` mean over detected replicates), and the naive-bias forecast
is the mean noise of the winner (compare: the naive row's "vs
$\beta(\hat H)$" column). The latter is the winner's-curse functional of
`oracle_design.R`, now with this design's own means and covariances as
inputs.

**What this scenario cannot predict** — stated as sharply as what it can.
The real family is far larger (six continuous covariates at ten cuts each,
both sides, plus seven binaries, under `maxk = 2` and the viability
floors), so real selection mass is diluted across many more low-purity
competitors: relative to this $M = 16$ scenario, expect *lower*
concentration on $g^\ast$, *larger* naive bias, and classification
metrics *below* these predictions — with the direction certain and the
magnitude requiring either the realized family (@sec-realized) or a larger
constructed one. And the multi-split gate strengthens the consistency
requirement relative to the single split, filtering marginal passers; both
idealizations are Level 5 rungs with signs attached, not silent gaps.

## Step 5 — the null-cell twin, predicted before it runs

The driver's `null_cell = TRUE` switch sets $\beta_{\mathrm{int}} = 0$, so
every region's oriented mean collapses to the common treatment effect
($\approx +26.26$, the complement value). The same machinery then predicts
the null cell's operating characteristics — and the prediction is the
useful kind of surprising: the floor sits only $(30 - 26.26)/\mathrm{SE}
\approx 0.13$–$0.28$ SEs above every candidate mean, so this "null" is a
**near-threshold null**, not a zero null.

```{r worked-null}
beta_null <- rep(tauQc, M)
W1n <- matrix(rnorm(Rdraw * M), Rdraw, M) %*% t(root_half) +
       rep(beta_null, each = Rdraw)
W2n <- matrix(rnorm(Rdraw * M), Rdraw, M) %*% t(root_half) +
       rep(beta_null, each = Rdraw)
pass_n <- (W1n + W2n >= 2 * c1) & (W1n >= c2) & (W2n >= c2)
cat(sprintf("Null-cell predicted detection (single-split gate): %.3f\n",
            mean(rowSums(pass_n) > 0)))
p1n <- colMeans(pass_n)
cat(sprintf("  per-candidate P1 range: %.3f-%.3f; implied L_eff vs max(P1): %.2f\n",
            min(p1n), max(p1n),
            log(1 - mean(rowSums(pass_n) > 0)) / log(1 - max(p1n))))
```

The predicted null-cell detection rate is far above any conventional
false-positive intuition, for a structural reason the mixture identity
makes exact: with a nonzero common effect near the floor, declaration is a
coin-flip per candidate and near-certain per family. This is the
borderline-null regime of the manuscript's §7 anchors, derived rather than
discovered — and it quantifies the driver's open question D2 from the
other side: on the null cell it is the *screening* floor, not the
consistency floor, that is barely discriminating.

## Step 6 — inverting for design

The same objects answer planning questions. For example: at what $n$ would
the marginal clearance of the best representable rule $g^\ast$ reach
$90\%$, and at what screening floor would the null-cell family detection
fall to $10\%$? Both are one-dimensional inversions of computed rates —
the first is the closed-form calculator of @sec-recipes applied at
$\beta = 38.67$; the second solves $c_1$ in the null family functional.
Floors chosen this way are calibrated to computed operating
characteristics rather than convention — with the family-size caveat of
Step 4 carried along.

## Step 7 — doing it with the realized family {#sec-realized}

For replicate-faithful numbers, replace the constructed scenario by a
realized family: from a fitted `forestsearch` object, the evaluated
candidate set with per-candidate membership (equivalently the arm-wise
counts $n_{g,a}$ and pairwise intersections) determines the exact
design-conditional law of Level 2 — means from the mixture identity
evaluated on `df_super` at each candidate's rule, covariance from the
overlap formula with no balanced-arms approximation. Every prediction of
Step 4 then re-runs unchanged on that law, now exact-conditional rather
than scenario-structural, with only the gate idealization and the
quadrature error remaining between computation and the replicate's truth.
That extraction is the natural next artifact on this thread; it is a
consumer of existing objects and requires no package changes. A completed
$s = 1000$ run of this design at $n = 700$ is set against these recipes,
tier by tier, in the Appendix (@sec-appendix).

# Part IV — Consolidated verification record {#sec-accuracy}

**Certified layer, exercised in the committed grid:** CHECK A at
$1.421\times 10^{-14}$ across 601–912 realized rules per cell; CHECK B
reproduced the calibration to machine precision on every run; CHECK C
caught a real scale-convention defect on first firing and now agrees to
$0.000$; the bit-level standard isolated an RNG-kind dependence and the
post-fix re-render reproduced the pilot on 35 of 35 scientific columns.
Every defect of the campaign was caught by an exact expectation; none by a
tolerance.

**Prediction layer:** the on-record suite
(`dev/detection-theory-continuous/check_declaration_gaussian_md.R`)
verified 16 PASS / 0 FAIL on the primary machine (R 4.6.1, mvtnorm 1.4.2,
2.4 s), numerically identical line for line to the drafting-stage sandbox
run (R 4.3.3, mvtnorm 1.2.4) — cross-stack bit-level reproduction. The
additive population covariance of @sec-additive verified at $0.2$ / $0.1$
MC standard errors (borderline / alternative). The document's own rendered
chunks reconcile against the same record; the selection-identity gap
readout fluctuates at the quadrature tolerance ($10^{-6}$) between
platforms unless seeded, and the warm-up chunk pins the seed for that
reason. The worked example's Gaussian-level Monte Carlo uses $2 \times
10^5$ draws (rate SEs $\approx 10^{-3}$) and a fixed seed.

The verification family of the check suite is a synthetic $M = 14$
stressor whose complement effect is beneficial (oriented $-26.26$) —
deliberately harder competition geometry than the fixture's harmful
complement ($+26.255$); the machinery checks are parameter-point-agnostic,
and the fixture values used throughout this document are the committed ones
of §2.

# Part V — The boundary, restated {#sec-boundary}

| Quantity | Status |
|---|---|
| $\beta(H)$ any rule; $\theta^\dagger(Q)$; ITT/complement/disjoint; classification identities | Tier A, exact |
| $\beta(H)$, binary/OR | Tier B, exact computation (mixture identity fails; RD retains Tier A) |
| Candidate covariance given family | Tier B, exact overlaps |
| Design-conditional joint law; family declaration; split-half consistency; selection distribution; winner-size law; naive-bias forecast | Distributional identity + quadrature; **verified** on fixed families |
| Population-level law | **Exact to leading order** — additive covariance decomposition derived, enumerable, verified ($0.2$ SEs at the adversarial point); second-order $O(1/n)$ terms remain |
| Marginal clearance; design calculators | Closed form; measured anchors consistent |
| Implied $L_{\mathrm{eff}}$ | Derived output; single-$P_1$ form requires variance homogeneity |
| Coverage | Conditionally closed-form; unconditional composes over the selection law given MR's $(b, w)$ |
| MR sufficiency | Empirical (mechanism resolved as the re-selection fidelity gap; the Level 6 benchmark specifies the fix's falsifiable prediction) |
| Family regeneration; nomination; multi-split $\hat p_{\mathrm{cons}}$ composition | Beyond the fixed-family computation (conditioning caveat stated; regeneration signatures identified; gate direction stated in @sec-worked) |
| Identifier's $n \to \infty$ limit | Property of the algorithm on a data-dependent cut lattice; carried as measurement |
| Survival, forward | AFT log-time: Tier A transfers wholesale for the uncensored contrast; observed-data Cox log-HR non-collapsible and censoring-dependent — Tier B at best, estimand distinctions maintained |

The division of labor, in one sentence: closed forms certify targets and
implementations exactly; the selection layer is predictable on fixed
families from certified inputs through an audited approximation ladder
whose every rung carries a measured error; and family regeneration,
nomination, and MR's sufficiency remain the empirical core the simulations
exist to measure — now against both certified references and computed
benchmarks.

# Appendix A — Analytic calculations versus the simulations: the maxeffCons design at n = 700 {#sec-appendix}

This appendix evaluates the document's recipes against a **completed**
simulation of the Step 1 design: the rendered report
`sim_fs_maxeffCons_mr_md40_knoise0_n700_batch_1_1000` ($n = 700$,
$s = 1000$ replicates, MR draws $= 5000$; the report's quick-run banner is
a stale comment — its tables are the full run, with $999$ of $1000$
replicates detected). The comparison is organized by the three tiers of
predictability the document defines: exact identities (zero error),
fixed-family structural benchmarks (quantitative on their scenario,
directional for the regenerated real family), and the replicate-faithful
route (specified, not yet executed). Every measured number below is cited
to its report section.

## The measured record

From the report: detection $999/1000$ (identification line: "detection
99.9% of 1000; anchor recovered 45.3%"); mean $|\hat H| = 74$ and mean
$|\hat H^c| = 626$ (estimation-table block headers, §6.3); classification
over detected replicates sens $0.123$, spec $0.904$, PPV $0.402$, NPV
$0.662$ (§6.7 kable, means); estimation table (§6.3, harm block,
$n = 999$): oracle mean $40.66$ with bias $8.90$ vs $\beta(\hat H)$;
naive mean $108.85$, SD $17.57$, bias $77.09$ vs $\beta(\hat H)$, mean CI
width $121.40$, coverage of $\beta(\hat H)$ $0.18$; MR mean $53.94$, SD
$22.37$, bias $22.17$, width $144.49$, coverage $0.97$.

## Tier 1 — identity locks: exact relations the report must satisfy

Three certified relations connect independently tabulated report columns.
None is a forecast; each is a zero-error constraint, and checking them is
an integrity audit on the run.

```{r appendix-identity-locks}
# Measured values, cited: report sections 6.3 (estimation table, harm
# block) and 6.7 (classification kable); sizes from the 6.3 block header.
meas <- list(sens = 0.123, ppv = 0.402, nH = 74, n = 700,
             est  = c(oracle = 40.66, naive = 108.85, mr = 53.94),
             bias = c(oracle =  8.90, naive =  77.09, mr = 22.17),
             sd_naive = 17.57, w_naive = 121.40, cov_naive = 0.18,
             sd_mr    = 22.37, w_mr    = 144.49, cov_mr    = 0.97)
# (i) Proposition 1 at the summary level: E[beta(Hhat)] = tauQc + bint*E[PPV]
backout <- meas$est - meas$bias            # three independent E[beta(Hhat)]
lock1   <- tauQc + bint * meas$ppv
# (ii) classification identity: PPV = sens * |Q| / |Hhat| (per replicate);
#      summary-level application with E|Q-sample| = n * pi
lock2 <- meas$sens * (meas$n * piQ_fix) / meas$nH
# (iii) conditional coverage from measured (bias, width, SD)
cov_formula <- function(b, w, s) pnorm((w/2 - b)/s) - pnorm((-w/2 - b)/s)
lock3n <- cov_formula(meas$bias["naive"], meas$w_naive, meas$sd_naive)
lock3m <- cov_formula(meas$bias["mr"],    meas$w_mr,    meas$sd_mr)
knitr::kable(data.frame(
  relation = c("E[beta(Hhat)]: Prop. 1 vs oracle-row back-out",
               "E[beta(Hhat)]: Prop. 1 vs naive-row back-out",
               "E[beta(Hhat)]: Prop. 1 vs MR-row back-out",
               "PPV from sens (classification identity)",
               "naive coverage from (bias, width, SD)",
               "MR coverage from (bias, width, SD)"),
  computed = round(c(lock1, lock1, lock1, lock2, lock3n, lock3m), 3),
  measured = round(c(backout, meas$ppv, meas$cov_naive, meas$cov_mr), 3)),
  caption = paste("Identity locks: the computed value of each relation",
                  "against the report's own columns. The one Prop.-1",
                  "prediction faces three mutually consistent measured",
                  "back-outs of E[beta(Hhat)], one per estimator row."))
```

Reading the locks. The mixture identity holds **per replicate** with the
realized rule's purity, so its summary form ties mean $\beta(\hat H)$ to
mean PPV exactly up to averaging; the three estimator rows back out the
same $31.76$–$31.77$, against the computed $31.78$. The classification
lock applies the per-replicate identity at the summary level ($n\pi$ for
the expected in-sample $|Q|$), agreeing to $0.001$. The coverage formula
reproduces the naive row's measured $0.18$ from its bias, width, and
spread alone, and places MR near $0.99$ against the measured $0.97$ — the
small gap is the mean-width plug-in (per-replicate widths vary), a
first-order composition exactly as §Part III's coverage recipe states.
What the formula does **not** supply is MR's bias and width themselves:
those are the empirical core.

## Tier 2 — the fixed-family benchmark at n = 700, against the real family

The scenario of the worked example, re-evaluated at this design's sample
size and viability floor ($n.min = 60$ at $n = 700$ is an $8.6\%$
prevalence floor, admitting $M = 18$ candidates), with the same gate and
selection rule. The chunk reuses the scenario's covariate law and
recomputes everything at $n = 700$; the measured column carries the
report's values.

```{r appendix-tier2}
set.seed(20260825)
Pg7  <- vapply(fam0 <- c(lapply(names(conds), function(nm)
          list(nm = nm, m = conds[[nm]])),
        unlist(lapply(list(c(1,3), c(1,4), c(2,3), c(2,4),
                           c(1,5), c(1,6), c(2,5), c(2,6),
                           c(3,5), c(3,6), c(4,5), c(4,6)),
          function(ix) list(list(nm = paste(names(conds)[ix], collapse = " & "),
                                 m = conds[[ix[1]]] & conds[[ix[2]]]))),
          recursive = FALSE)),
        function(f) sum(cells$prob[f$m]), numeric(1))
keep7 <- Pg7 >= 60/700
fam7  <- fam0[keep7]; Pg7 <- Pg7[keep7]; M7 <- length(fam7)
lab7  <- vapply(fam7, `[[`, character(1), "nm")
PQg7  <- vapply(fam7, function(f) sum(cells$prob[f$m & cells$inQ]),
                numeric(1)) / Pg7
beta7 <- tauQc + bint * PQg7
se7   <- seQ1000 * sqrt(1000/700) * sqrt(piQ_fix / Pg7)
sens7 <- vapply(fam7, function(f) sum(cells$prob[f$m & cells$inQ]),
                numeric(1)) / piQ_fix
spec7 <- vapply(fam7, function(f)
  1 - sum(cells$prob[f$m & !cells$inQ]) / (1 - piQ_fix), numeric(1))
ovl7 <- outer(seq_len(M7), seq_len(M7), Vectorize(function(i, j)
  sum(cells$prob[fam7[[i]]$m & fam7[[j]]$m])))
Sg7 <- (ovl7 / sqrt(outer(Pg7, Pg7))) * outer(se7, se7)
e7  <- eigen((Sg7 + t(Sg7))/2, symmetric = TRUE)
rt7 <- e7$vectors %*% diag(sqrt(pmax(2 * e7$values, 0)))
R7  <- 2e5
W1 <- matrix(rnorm(R7 * M7), R7, M7) %*% t(rt7) + rep(beta7, each = R7)
W2 <- matrix(rnorm(R7 * M7), R7, M7) %*% t(rt7) + rep(beta7, each = R7)
Bh <- (W1 + W2)/2
pass <- (W1 + W2 >= 2 * c1) & (W1 >= c2) & (W2 >= c2)
det  <- rowSums(pass) > 0
Bm <- Bh; Bm[!pass] <- -Inf
win <- max.col(Bm, ties.method = "first"); win[!det] <- NA
sel <- tabulate(win[det], M7) / R7; selc <- sel / mean(det)
npv7 <- vapply(seq_len(M7), function(m) {
  pc <- 1 - Pg7[m]; (pc - (piQ_fix - PQg7[m] * Pg7[m])) / pc }, numeric(1))
pred <- c(detection = mean(det),
          EnH  = 700 * sum(selc * Pg7),
          sens = sum(selc * sens7), spec = sum(selc * spec7),
          ppv  = sum(selc * PQg7),  npv  = sum(selc * npv7),
          EbH  = sum(selc * beta7),
          naive = mean((Bh - rep(beta7, each = R7))[cbind(which(det), win[det])]),
          below = sum(selc[beta7 < c1]))
measd <- c(sprintf("%.3f", c(0.999, 74, 0.123, 0.904, 0.402, 0.662,
                              31.76, 77.09)),
           "not tabulated (bundle-computable)")
measd[2] <- "74"
knitr::kable(data.frame(
  quantity = c("detection rate", "E|Hhat| given detection",
               "E[sens]", "E[spec]", "E[PPV]", "E[NPV]",
               "E[beta(Hhat)] oriented", "naive bias vs beta(Hhat)",
               "selection mass, true mean below floor"),
  predicted_M18 = round(pred, 3), measured = measd),
  caption = sprintf(
    "Fixed-family benchmark (M = %d, floor 60/700) vs the measured n = 700 record.", M7))
top7 <- order(-sel)[1:3]
cat(sprintf("Top predicted selections: %s\n",
    paste(sprintf("%s %.3f", lab7[top7], sel[top7]), collapse = " | ")))
```

Reading the comparison, row by row, with each residual assigned to its
rung. **Detection** ($0.978$ predicted vs $0.999$): the real family is
far larger; more candidates raise the family exceedance — the stated
Level 5 direction. **Winner size and sensitivity** ($163$ vs $74$;
$0.287$ vs $0.123$): the regenerated family's fine cuts supply many small
candidates, and the noise-seeking argmax takes them — the scenario's own
signature of this regime appears in its top predicted selections, where
at the $8.6\%$ floor a **zero-purity** rule leads. **PPV** ($0.312$
predicted vs $0.402$ measured) is the honest miss in direction: the real
family's fine cuts also supply small *high-purity* slices of $Q$ that the
coarse $M = 18$ scenario cannot represent, so measured purity beats the
scenario ceiling — a regeneration effect with the opposite sign to the
sensitivity one, and a caution against reading the scenario's
classification rows as bounds. **E[$\beta(\hat H)$]** ($30.55$ vs
$31.76$) is near-quantitative, for a reason the locks make plain: the
quantity is pinned by the mixture identity to mean purity, and selection
concentrates just above the floor in both worlds. **Naive bias** ($+29.7$
vs $+77.1$): regeneration multiplies the winner's curse roughly
$2.6\times$ here — the direction stated in Step 4, its magnitude the
clearest single measure of what the fixed family cannot supply. The
predicted below-floor selection mass ($0.544$) has no directly tabulated
counterpart; its per-replicate analogue (share of detected replicates
with $\beta(\hat H) < c_1$) is computable from the bundle and is a
natural addition to the report.

## Tier 3 and the boundary

The replicate-faithful route of the worked example's final step — feeding
each replicate's realized family (candidate memberships and arm counts
from the fitted objects) into the design-conditional law — would replace
every Tier 2 attribution above with a measured decomposition: computed
detection, selection, and naive-bias per replicate, exact up to the gate
idealization and quadrature. It has not yet been executed; it is a
consumer of existing bundle objects. What no tier predicts, by
construction, is the MR correction's own residual: the measured $+22.17$
bias and $144.49$ mean width are inputs to the coverage composition
above, not outputs of the analytic layer — the empirical core the
simulations exist to measure, scored against references the certified
layer supplies exactly.

# Appendix B — Causal-inference reference results (Wager) {#sec-refwager}

This appendix makes the document self-contained on its causal-inference
foundations by restating, in this document's notation and with precise
citations, the results it uses or parallels from Stefan Wager's *Causal
Inference: A Statistical Learning Approach* (Stanford; draft dated
November 26, 2025), cited throughout as **[W]** with the draft's equation
and theorem numbers, and its ancestor, the STATS 361 lecture notes
(Spring 2020), cited as **[W-notes]**. Because [W] is a dated draft,
numbering may shift in print; the cited section titles are stable. Each
subsection ends with the role the result plays here.

## Potential outcomes, randomization, and difference in means {#sec-refwager-rct}

**Setup ([W] §1.1).** Each unit carries potential outcomes
$(Y_i(0), Y_i(1))$. SUTVA (eq. 1.5) states $Y_i = Y_i(W_i)$: the
observed outcome is the potential outcome of the assigned arm, with no
interference across units. Randomization (eq. 1.6) states that,
conditionally on all potential outcomes and on the treated count $n_1$,
every unit is treated with the same probability $n_1/n$. The
difference-in-means estimator (eq. 1.4) is
$\hat\tau_{\mathrm{DM}} = n_1^{-1}\sum_{W_i=1} Y_i -
n_0^{-1}\sum_{W_i=0} Y_i$.

**Theorem 1.1 ([W]).** Under SUTVA and randomization,
$\hat\tau_{\mathrm{DM}}$ is finite-sample unbiased for the sample
average treatment effect, conditionally on the potential outcomes and on
both arms being non-empty. The proof is iterated expectation:
assignment independence pulls through the sum.

**Superpopulation CLT and the Neyman variance ([W] §1.1, around
eqs. 1.9–1.10).** With units i.i.d. from a population and treated
fraction $\pi_{\mathrm{tr}}$, define residuals
$\varepsilon_i(w) = Y_i(w) - \mathbb E[Y_i(w)]$. The estimation error
decomposes as
$$
\hat\tau_{\mathrm{DM}} - \tau
= \frac{n}{n_1}\cdot\frac1n\sum_i W_i\,\varepsilon_i(1)
- \frac{n}{n_0}\cdot\frac1n\sum_i (1-W_i)\,\varepsilon_i(0);
$$
the two per-arm sums are uncorrelated with variances
$\pi_{\mathrm{tr}}\,\mathrm{Var}[\varepsilon(1)]$ and
$(1-\pi_{\mathrm{tr}})\,\mathrm{Var}[\varepsilon(0)]$, and the CLT gives
$\sqrt n\,(\hat\tau_{\mathrm{DM}} - \tau) \Rightarrow
N(0, V_{\mathrm{DM}})$ with
$$
V_{\mathrm{DM}}
= \frac{\mathrm{Var}[\varepsilon(1)]}{\pi_{\mathrm{tr}}}
+ \frac{\mathrm{Var}[\varepsilon(0)]}{1-\pi_{\mathrm{tr}}},
$$
consistently estimable by the arm-wise plug-in
$\widehat V_{\mathrm{DM}}$.

**Role here.** The §2 estimator is exactly this object restricted to a
region ($\pi_{\mathrm{tr}} = 1/2$); Proposition 1's Step 2 is eq. 1.6
surviving the conditioning on $\{X \in H\}$; the residual decomposition
is the seed of the influence representation $\psi_i(g)$ of
@sec-additive; the Neyman variance restricted to $g$ **is** the Level 4
diagonal (the identity displayed in @sec-ladder, Level 4); and the
consistent variance estimator is the ancestor of the Level 3 plug-in
observation.

## Regression adjustment in randomized trials {#sec-refwager-adj}

**[W] §1.2.** Adjusting the difference in means by fitted per-arm
outcome models leaves the estimator consistent under randomization
regardless of model quality, and reduces the asymptotic variance by the
portion of $\mathrm{Var}[\varepsilon(w)]$ explained by the covariates —
the prognostic component. **Role here.** This is the population
statement of two facts the ladder measures: the per-region `lm` residual
variance absorbing within-region prognostic and CATE spread (inflation
factors $0.94$–$1.09$, Level 3), and the $u^{(g)}$ prognostic terms of
the additive decomposition (Level 4) being exactly the component an
adjusted estimator could remove — quantified as
$\mathrm{Var}(2\mu_0 + \tau \mid g)$ in the Level 4 identity.

## AIPW, the efficient influence function, and the efficiency bound {#sec-refwager-eff}

**AIPW ([W] §3.1–3.2).** Under unconfoundedness with propensity $e(x)$
and outcome regressions $\mu_{(w)}(x)$, the augmented IPW estimator
averages the scores
$$
\Gamma_i = \mu_{(1)}(X_i) - \mu_{(0)}(X_i)
+ \frac{W_i\,\big(Y_i - \mu_{(1)}(X_i)\big)}{e(X_i)}
- \frac{(1-W_i)\,\big(Y_i - \mu_{(0)}(X_i)\big)}{1 - e(X_i)}.
$$
**Theorem 3.2 ([W]), with cross-fitting:** with nuisances estimated at
$o(n^{-1/4})$ rates on held-out folds,
$\sqrt n\,(\hat\tau_{\mathrm{AIPW}} - \tau) \Rightarrow N(0, V^\ast)$
and the plug-in variance is consistent. **The bound (eq. 3.20) and
Theorem 3.4 ([W]):**
$$
V^\ast = \mathrm{Var}[\tau(X)]
+ \mathbb E\!\left[\frac{\sigma^2_{(1)}(X)}{e(X)}\right]
+ \mathbb E\!\left[\frac{\sigma^2_{(0)}(X)}{1-e(X)}\right]
$$
is the *efficient* variance in the local asymptotic minimax sense of
Hájek [1972], via the argument of Chamberlain [1992]: no estimator
improves on it uniformly over a neighborhood of any $P$.

**Role here.** Three uses. (i) Specialized to the fixture ($e = 1/2$,
homoscedastic noise) within a region,
$V^\ast(g) = 4\sigma^2 + \mathrm{Var}(\tau \mid g)$, and the Level 4
identity gives the exact excess of the package's unadjusted estimator
over the bound — $\mathrm{Var}(2\mu_0 + \tau \mid g)$, the adjustable
prognostic component (display in @sec-ladder, Level 4). (ii) The score
$\Gamma_i$, restricted to a region with known $e = 1/2$ and the
regression pieces absorbed into the centering, reduces to the
$\psi_i(g)$ of @sec-additive up to the $2/(nP(g))$ normalization; the
Gram/PSD structure of $\Sigma_{\mathrm{pop}}$ is the second-moment
structure of these scores. (iii) The Crump machinery's
`sandwich_var_glm` (the B2 and survival lanes of the detection-theory
NOTEs) is the empirical second moment of the same estimating-equation
scores at the basis level — the Level 3 plug-in layer, pre-built.

## The CATE, the R-learner frame, and the GRF identifier's target {#sec-refwager-cate}

**[W] §4.1–4.2.** The estimand is the conditional average treatment
effect $\tau(x) = \mu_{(1)}(x) - \mu_{(0)}(x)$. Estimating the two arm
surfaces separately and differencing incurs *regularization bias* — the
fits regularize differently and the difference inherits artifacts ([W]
Fig. 4.1: an oscillatory control fit against a flat treated fit
manufactures a spurious oscillatory CATE). The remedy targets $\tau$
directly: the semiparametric specification $\tau(x) = \psi(x)\cdot\beta$
with residual-on-residual regression, generalized in §4.2 to a loss
function for treatment heterogeneity whose minimization over flexible
classes defines the R-learner; the causal forest of Athey, Tibshirani,
and Wager [2019] instantiates this framework with random forests.

**Role here.** The GRF identifier's nominations target exactly this
$\tau(\cdot)$; Proposition 1 is the identity-scale collapsibility that
turns region means into linear functionals of $\tau$; and the fixture's
two-valued $\tau$ is the piecewise-constant case in which thresholding
rules are the natural optimal policies — the hinge to the next
subsection. Nomination itself remains unmodeled in the ladder (Level 5);
this subsection supplies its formal reference, not its law.

## Policy learning, empirical welfare maximization, and regret {#sec-refwager-policy}

**[W] §5.1–5.2.** A policy $\pi: x \mapsto \{0, 1\}$ has value
$V(\pi)$; with a test set under unconfoundedness, $V(\pi)$ is estimable
by IPW or AIPW value estimators (§5.1). Learning constrains $\pi$ to a
class $\Pi$ (linear thresholds, fixed-depth trees, …); the optimum is
$\pi^\ast \in \arg\max\{V(\pi'): \pi' \in \Pi\}$ (eq. 5.14), the regret
of any acceptable $\pi$ is
$R(\pi) = \sup_{\pi'} V(\pi') - V(\pi)$ (eq. 5.15), and
empirical-welfare maximization [Kitagawa and Tetenov, 2018] learns
$\hat\pi = \arg\max\{\widehat V(\pi): \pi \in \Pi\}$ (eq. 5.16). The
regret of EWM scales with the square root of the policy-class
complexity over $n$ ([W] §5.2, regret-bounds sketch), and the learning
problem reduces to weighted classification.

**Role here — kinship and contrast, with estimand discipline.** The
selection layer of this document and EWM are two lenses on one object:
an argmax over a rule class driven by noisy per-rule evaluations. The
kinships are exact in kind: the policy-class complexity term is the
bound-side face of what $L_{\mathrm{eff}}$ and the family joint law
measure on the exact-law side, and the winner's-curse functional (the
computed naive bias) is the finite-$n$ distributional content that
regret bounds envelope in expectation. The contrasts matter equally:
EWM targets the *value* $V(\pi)$ — a size-weighted welfare — with
guarantees over $\Pi$; this document's procedure ranks by the
*per-rule effect* under screening and consistency gates (maxeffCons is
argmax of effect among gate-passers, not argmax of welfare), and its
analytic layer computes the exact finite-$n$ law of that specific
procedure, targeting $\beta(\hat H)$. Neither subsumes the other: the
regret frame answers "how good a policy can learning guarantee"; the
present frame answers "what will this identifier do, with what
probabilities, on this design."

## Source map and scope

[W-notes] chapters 1–4 (RCTs; unconfoundedness; AIPW; heterogeneity)
are the ancestors of [W] chapters 1–4, and [W] Ch. 5 extends the
corresponding policy-learning lecture. Chapters of [W] beyond these —
balancing estimators, regression discontinuity, instruments and LATE,
interference, event studies, dynamic policies — are not used by this
document and are named only to delimit scope. Full bibliographic
entries for the works cited inside the restatements (Hájek 1972;
Chamberlain 1992; Kitagawa and Tetenov 2018; Athey, Tibshirani, and
Wager 2019) appear in [W]'s bibliography.
