---
title: "MR coverage sweep — continuous / MD, harm direction"
subtitle: "Unbiasedness and interval coverage of the MR-corrected estimate against the EXACT beta(Hhat)"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: true
  message: true
---

# What this document is

The GLM/continuous counterpart of the binary MR coverage sweeps, in the **harm**
direction. It measures two things from one set of replicates:

1. **Unbiasedness** — mean (estimate − β(Ĥ)) for the naive in-trial MD and for
   the MR-corrected estimate, side by side.
2. **Interval coverage** — of the MR interval for β(Ĥ), with `n_eff` reported.

## Why the continuous pathway makes this a clean measurement

The mean difference is **collapsible**, so β(Ĥ) is an exact finite mean over
`dgm$df_super` — no evaluation frame, no model fit, **zero Monte Carlo error in
the target**. Every deviation of an estimate from β(Ĥ) is therefore the method,
not the target. On the survival and binary pathways the target is itself a
fitted quantity and the two sources are entangled.

`fs_build_eval_frame(dgm, "continuous")` returns `df_super` unchanged for this
reason: the scoring frame *is* the population.

## Closed-form cross-checks

Where a readout has a closed-form counterpart, the counterpart is computed
independently and printed beside it, then asserted. The point is to confirm the
machinery against **theory**, not only against itself. A failing assertion is a
machinery defect by definition and stops the phase.

- **A** — the two-valued CATE identity on the harness's own realized rules:
  β(Ĥ) must equal δ + β_inter · P(Q | Ĥ).
- **B** — θ† must reproduce the DGM's own `effect_Q` / `effect_Qc`, exactly.
- **C** — the naive mean bias and the mean selection effect are the same
  quantity computed two ways and must agree.

## Decisions (closed — recorded, not revisited)

`focus = "harm"`; harm-calibrated DGM at `cal_target_md = -40`;
`adverse_outcome = FALSE` with positive thresholds per
`dev/glm-continuous-sims/NOTE_threshold_sign_md.md`; FS/consistency identifier
only. The selection rate at these settings is ≈ 1.0 and that is the
phenomenon, on the record — it is not tuned away.

```{r setup, message=FALSE, warning=FALSE}
.t0_doc <- Sys.time()

# forestsearch must be INSTALLED (devtools::install()), not load_all(): the
# doFuture workers are separate R processes that see only the installed package.
library(forestsearch)
library(speff2trial)
library(doFuture)
library(foreach)
library(future)

`%||%` <- function(a, b) if (is.null(a)) b else a

# ── Run identity ─────────────────────────────────────────────────────────────
run_sweep   <- TRUE
run_tag     <- "md_harm_s50_pilot"      # pilot; full grid bumps this
results_dir <- file.path("mr_sweep_md_harm", run_tag)
grid_path   <- file.path(results_dir, sprintf("mr_coverage_grid_%s.rds", run_tag))

# ── Cells ────────────────────────────────────────────────────────────────────
methods_run <- "consistency"            # v1 scope: FS/consistency only
n_grid_run  <- 1000L                    # STOP A pilot: one cell
n_sims      <- 50L                      # STOP A pilot: 50 replicates
mr_draws    <- 5000L
nb_boots    <- NULL                     # MR-only; FB dormant

# ── Seeds: PRE-GENERATED, indexed by GLOBAL replicate id ─────────────────────
# Results must be invariant to worker count and to batch position, so the seed
# for replicate i is looked up, never derived from loop position.
seed_base  <- 8316951L
MAX_SIMS   <- 5000L
set.seed(seed_base)
SEED_TABLE <- sample.int(.Machine$integer.max - 1L, MAX_SIMS)
seed_for <- function(sim_id) {
  if (sim_id < 1L || sim_id > MAX_SIMS)
    stop("sim_id ", sim_id, " outside the pre-generated seed table.")
  SEED_TABLE[sim_id]
}

# ── DGM: the verified harm calibration ───────────────────────────────────────
# Do NOT recalibrate.  cal_target_md = -40 converges at machine precision.
actg_arms        <- c(1L, 3L)
actg_treat_arm   <- 1L
actg_age_cut     <- 34
actg_preanti_cut <- 744.5
dgm_factor_vars   <- paste0("z", 1:12)
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)
dgm_n_super       <- 5000L
cal_target_md     <- -40                # HARM
cal_k_grid_range  <- c(0, 120)
cal_grid_step     <- 2

# ── Search: the corrected harm orientation ───────────────────────────────────
# adverse_outcome = FALSE negates Y internally so MD > 0 means harm; the
# thresholds are therefore POSITIVE on that oriented scale.  See
# dev/glm-continuous-sims/NOTE_threshold_sign_md.md -- a negative threshold here
# is a near-vacuous floor, not a harm specifier.
md_threshold    <- 30
md_consistency  <- 10
adverse_outcome <- FALSE
pconsistency    <- 0.90
fs_splits <- 400L; maxk <- 2L; n_min <- 60L; d0_min <- 12L; d1_min <- 12L
vi_grf_min <- -0.2
use_twostage <- TRUE; is_rct <- TRUE
# MR-COMPATIBLE: no model-based front end.  use_lasso / use_grf / use_dina are
# omitted entirely and take their forestsearch() formal defaults, all FALSE
# (R/forestsearch_main.R:1096-1100).  With any of them TRUE the MR alignment
# guard rejects the call before fitting -- see REPORT_blockers_1-3_ADDENDUM.md
# section D.1, where that guard produced the pilot's detected 0 of 50.

analysis_continuous_vars <- c("age","preanti","wtkg","karnof","cd40","cd80")
analysis_binary_vars     <- c("hemo","homo","drugs","race","gender","symptom")

outcome_name <- "y_sim"; treat_name <- "treat_sim"
id_name <- "id"; harm_col <- "flag_harm"

# ── Parallel ─────────────────────────────────────────────────────────────────
parallel_mode  <- "sims"
n_workers      <- max(1L, ceiling(0.90 * (parallel::detectCores(logical = FALSE) - 1L)))
inner_parallel <- list(plan = "sequential")

cat(sprintf("run_tag=%s  cells: %s x n in {%s}  n_sims=%d  mr_draws=%d  workers=%d\n",
            run_tag, paste(methods_run, collapse=","),
            paste(n_grid_run, collapse=","), n_sims, mr_draws, n_workers))
```

# The DGM, the scoring frame, and theta-dagger {#sec-dgm}

```{r build-dgm}
dgm <- NULL; eval_df <- NULL; truth <- NULL
if (run_sweep) {
  actg_df <- subset(speff2trial::ACTG175, arms %in% actg_arms)
  actg_df$id <- seq_len(nrow(actg_df))
  actg_df$treat_orig <- ifelse(actg_df$arms == actg_treat_arm, 1L, 0L)
  actg_df$treat      <- 1L - actg_df$treat_orig     # SWITCHED: ddI = 1
  actg_df$cd4_change <- actg_df$cd420 - actg_df$cd40
  actg_df <- actg_df[!is.na(actg_df$cd420), ]
  actg_df$z1  <- as.factor(ifelse(actg_df$age > actg_age_cut, 1L, 0L))
  actg_df$z2  <- as.factor(ifelse(actg_df$preanti <= actg_preanti_cut, 1L, 0L))
  actg_df$z3  <- as.factor(ifelse(actg_df$wtkg <= 75, 1L, 0L))
  actg_df$z4  <- as.factor(ifelse(actg_df$karnof <= median(actg_df$karnof), 1L, 0L))
  actg_df$z5  <- as.factor(ifelse(actg_df$cd40 <= median(actg_df$cd40), 1L, 0L))
  actg_df$z6  <- as.factor(ifelse(actg_df$cd80 <= median(actg_df$cd80), 1L, 0L))
  actg_df$z7  <- as.factor(actg_df$hemo);  actg_df$z8  <- as.factor(actg_df$homo)
  actg_df$z9  <- as.factor(actg_df$drugs); actg_df$z10 <- as.factor(actg_df$race)
  actg_df$z11 <- as.factor(actg_df$gender);actg_df$z12 <- as.factor(actg_df$symptom)
  for (v in analysis_binary_vars) actg_df[[v]] <- as.factor(actg_df[[v]])
  confounders_analysis <- c(analysis_continuous_vars, analysis_binary_vars)

  dgm <- calibrate_glm_interaction(
    data = actg_df, factor_vars = dgm_factor_vars,
    outcome_var = "cd4_change", treatment_var = "treat",
    target_effect = cal_target_md, outcome_type = "continuous",
    effect_measure = "MD", subgroup_vars = dgm_subgroup_vars,
    subgroup_cuts = dgm_subgroup_cuts, k_inter_range = cal_k_grid_range,
    grid_step = cal_grid_step, n_super = dgm_n_super, seed = seed_base,
    verbose = FALSE)

  # The scoring frame.  For continuous this RETURNS df_super unchanged: the MD
  # is collapsible, so the target is an exact finite mean over the population
  # and no simulation occurs.
  eval_df <- fs_build_eval_frame(dgm, outcome_type = "continuous")

  inQ        <- eval_df[[harm_col]] == 1L
  delta      <- compute_aor(eval_df, as.integer(!inQ), "MD")   # complement effect
  beta_inter <- dgm$model_params$beta_inter
  effect_Q   <- dgm$hazard_ratios$harm_subgroup
  effect_Qc  <- dgm$hazard_ratios$no_harm_subgroup

  td <- fs_betaHhat_theta_dagger_check(eval_df, outcome_type = "continuous",
                                       harm.name = harm_col,
                                       effect_measure = "MD")
  truth <- list(marg_H = unname(td[["thetaDagger_H"]]),
                marg_Hc = unname(td[["thetaDagger_Hc"]]),
                effect_Q = effect_Q, effect_Qc = effect_Qc,
                delta = delta, beta_inter = beta_inter,
                prevalence_Q = mean(inQ))

  cat(sprintf("df_super n=%d   P(Q)=%.4f\n", nrow(eval_df), mean(inQ)))
  cat(sprintf("MD(Q)=%.6f  MD(Qc)=%.6f  delta=%.6f  beta_inter=%.6f\n",
              effect_Q, effect_Qc, delta, beta_inter))
  cat(sprintf("theta-dagger  H=%.10f  Hc=%.10f\n", truth$marg_H, truth$marg_Hc))
  cat(sprintf("scoring frame IS df_super: %s\n",
              identical(eval_df, dgm$df_super)))
}
```

## Closed-form check B — θ† against the DGM's own effects {#sec-checkB}

**Provenance:** criterion 4 of `acceptance_betaHhat_md.qmd`, ported to the
package entry point `fs_betaHhat_theta_dagger_check()`.

θ† and the DGM's `effect_Q`/`effect_Qc` are the same population quantity
computed by the same `compute_aor()` dispatch on the same columns. This is an
**exact identity**, not agreement to Monte Carlo error; a tolerance here would
hide a real defect.

Both sides are read off the calibrated object: the harness value from
`fs_betaHhat_theta_dagger_check(eval_df, ...)`, the reference from
`dgm$hazard_ratios$harm_subgroup` / `$no_harm_subgroup`.

```{r check-B}
if (run_sweep) {
  chk <- data.frame(
    quantity = c("theta_dagger_H", "theta_dagger_Hc"),
    harness  = c(truth$marg_H, truth$marg_Hc),
    dgm      = c(truth$effect_Q, truth$effect_Qc))
  chk$difference <- chk$harness - chk$dgm
  print(chk, row.names = FALSE, digits = 12)
  ok_B <- identical(chk$harness, chk$dgm)
  cat(sprintf("\nCHECK B: identical() = %s   max |difference| = %.3e\n",
              ok_B, max(abs(chk$difference))))
  if (!ok_B) stop("CHECK B FAILED: theta-dagger does not reproduce the DGM's own effects.")
}
```

# The recorder {#sec-recorder}

```{r recorder}
.na_record <- function(sim_id) data.frame(
  sim_id = sim_id, seed = NA_integer_, detected = 0L, mr_ok = 0L,
  n_sel = NA_integer_, n_harm = NA_integer_, n_true = NA_integer_,
  label = NA_character_, sg_def = NA_character_,
  md_trial_H = NA_real_,                       # naive in-trial MD on Hhat
  nv_H_est = NA_real_, nv_H_lo = NA_real_, nv_H_hi = NA_real_,
  t2_H_est = NA_real_, t2_H_lo = NA_real_, t2_H_hi = NA_real_,
  t2_H_se_ij = NA_real_,
  nv_Hc_est = NA_real_, t2_Hc_est = NA_real_,
  t2_Hc_lo = NA_real_, t2_Hc_hi = NA_real_,
  t2_harm_flag = NA_integer_, ij_source = NA_character_,
  sens = NA_real_, spec = NA_real_, ppv = NA_real_, npv = NA_real_,
  t2_secs = NA_real_,
  # Diagnostics, so a configuration failure can never masquerade as a
  # scientific null result (addendum D.1/D.4).  status is one of
  # "CONFIG-ERROR" / "NO-DETECTION" / "DETECTED"; on CONFIG-ERROR the
  # detected and mr_ok fields are NA, not 0.
  status = NA_character_, err_msg = NA_character_, mr_msg = NA_character_,
  stringsAsFactors = FALSE)

.classify <- function(sg_hat, truth_flag) {
  out <- c(sens = NA_real_, spec = NA_real_, ppv = NA_real_, npv = NA_real_)
  if (is.null(sg_hat) || length(sg_hat) != length(truth_flag)) return(out)
  sg_hat <- as.integer(sg_hat); truth_flag <- as.integer(truth_flag)
  tp <- sum(sg_hat == 1 & truth_flag == 1); fp <- sum(sg_hat == 1 & truth_flag == 0)
  fn <- sum(sg_hat == 0 & truth_flag == 1); tn <- sum(sg_hat == 0 & truth_flag == 0)
  out["sens"] <- if (tp + fn > 0) tp / (tp + fn) else NA_real_
  out["spec"] <- if (tn + fp > 0) tn / (tn + fp) else NA_real_
  out["ppv"]  <- if (tp + fp > 0) tp / (tp + fp) else NA_real_
  out["npv"]  <- if (tn + fn > 0) tn / (tn + fn) else NA_real_
  out
}

record_replicate <- function(sim_id, method, n_sample, nb_boots) {
  rec <- .na_record(sim_id)
  sd_i <- seed_for(sim_id)                     # global-id lookup, not position
  rec$seed <- sd_i

  # Pin the RNG kind.  simulate_from_glm_dgm() calls set.seed(seed) with
  # kind = NULL, which KEEPS the ambient kind, so the realized draw depends on
  # whichever kind the backend installed -- Mersenne-Twister sequentially,
  # L'Ecuyer-CMRG under %dofuture% with seed = TRUE.  Pinning it here makes a
  # replicate the same replicate however it is run.  See
  # REPORT_blockers_1-3_ADDENDUM.md section D.2, where CMRG reproduces the
  # pilot bundle's n_true = 341 and Mersenne-Twister gives 334.
  RNGkind("L'Ecuyer-CMRG")
  set.seed(sd_i)

  df <- tryCatch(simulate_from_glm_dgm(dgm, n = n_sample, seed = sd_i),
                 error = function(e) NULL)
  if (is.null(df)) return(rec)
  df[[id_name]] <- seq_len(nrow(df))
  rec$n_true <- sum(df[[harm_col]] == 1L, na.rm = TRUE)

  # Diagnostics are CAPTURED, never swallowed.  The pilot's tryCatch(error =
  # function(e) NULL) turned a hard configuration error into detected = 0 on
  # all 50 replicates, and quiet = TRUE silenced the MR skip announcement --
  # see REPORT_blockers_1-3_ADDENDUM.md sections D.1 and D.4.  A configuration
  # failure must never again read as a scientific null result.
  t0 <- proc.time()[3]
  fs.est  <- NULL
  err_msg <- NA_character_
  # quiet = FALSE so the .mr_skip announcement is emitted; captured as messages
  # rather than printed, so a 50-replicate run stays readable.
  msgs <- tryCatch(
    utils::capture.output(
      fs.est <- suppressWarnings(forestsearch(
        df.analysis = df, confounders.name = confounders_analysis,
        outcome.name = outcome_name, treat.name = treat_name, id.name = id_name,
        outcome_type = "continuous", effect_measure = "MD",
        effect.threshold = md_threshold, consistency.threshold = md_consistency,
        pconsistency.threshold = pconsistency, fs.splits = fs_splits,
        n.min = n_min, d0.min = d0_min, d1.min = d1_min, maxk = maxk,
        vi.grf.min = vi_grf_min,
        consistency_method = "resample",
        use_twostage = use_twostage, is.RCT = is_rct,
        adverse_outcome = adverse_outcome,
        details = FALSE, quiet = FALSE, seedit = sd_i,
        parallel_args = inner_parallel,
        mr_inference = TRUE,
        mr_inference_args = list(ci_method = "ij", draws = mr_draws,
                                 include_complement = TRUE))),
      type = "message"),
    error = function(e) { err_msg <<- conditionMessage(e); character(0) })
  rec$t2_secs <- proc.time()[3] - t0

  # Keep the MR-relevant announcements; the rest of the console is noise.
  mr_lines <- grep("multiplier resampling|mr_inference", msgs, value = TRUE)
  if (length(mr_lines)) rec$mr_msg <- paste(mr_lines, collapse = " | ")

  # CONFIG-ERROR: forestsearch() refused the call.  Distinct from NO-DETECTION,
  # and recorded as NA rather than 0 -- nothing was ever searched.
  if (!is.na(err_msg) || is.null(fs.est)) {
    rec$status   <- "CONFIG-ERROR"
    rec$err_msg  <- err_msg
    rec$detected <- NA_integer_
    rec$mr_ok    <- NA_integer_
    return(rec)
  }

  found <- !is.null(fs.est$sg.harm) && length(fs.est$sg.harm) > 0 &&
           !all(is.na(fs.est$sg.harm))
  rec$detected <- as.integer(found)
  if (!found) { rec$status <- "NO-DETECTION"; return(rec) }
  rec$status <- "DETECTED"
  rec$sg_def <- paste(fs.est$sg.harm, collapse = " & ")

  # naive in-trial MD on the realized region, computed HERE from the trial --
  # an independent path to the same quantity the gate reports as `naive`.
  pr <- tryCatch(get_dfpred(df, fs.est$sg.harm), error = function(e) NULL)
  if (!is.null(pr) && !is.null(pr$treat.recommend)) {
    inH <- pr$treat.recommend == 0L
    rec$n_harm <- sum(inH, na.rm = TRUE)
    d <- df[which(inH), , drop = FALSE]
    if (sum(d[[treat_name]] == 1L) >= 2L && sum(d[[treat_name]] == 0L) >= 2L)
      rec$md_trial_H <- mean(d[[outcome_name]][d[[treat_name]] == 1L]) -
                        mean(d[[outcome_name]][d[[treat_name]] == 0L])
    cl <- .classify(as.integer(inH), df[[harm_col]])
    rec$sens <- cl[["sens"]]; rec$spec <- cl[["spec"]]
    rec$ppv  <- cl[["ppv"]];  rec$npv  <- cl[["npv"]]
  }

  g <- fs.est$mr_inference
  rec$mr_ok <- as.integer(!is.null(g))
  if (!is.null(g)) {
    rec$label <- g$selected_label %||% NA_character_
    rec$n_sel <- g$n_selected %||% NA_integer_
    rec$t2_harm_flag <- as.integer(isTRUE(g$harm_flag))
    rec$ij_source <- g$debiased$ij_source %||% NA_character_
    rec$nv_H_est <- g$naive$est; rec$nv_H_lo <- g$naive$lower; rec$nv_H_hi <- g$naive$upper
    rec$t2_H_est <- g$debiased$est; rec$t2_H_lo <- g$debiased$lower
    rec$t2_H_hi <- g$debiased$upper
    rec$t2_H_se_ij <- g$debiased$se_ij %||% NA_real_
    if (is.list(g$complement) && !is.null(g$complement$debiased)) {
      rec$nv_Hc_est <- g$complement$naive$est
      rec$t2_Hc_est <- g$complement$debiased$est
      rec$t2_Hc_lo  <- g$complement$debiased$lower
      rec$t2_Hc_hi  <- g$complement$debiased$upper
    }
  }
  rec
}
```

# The cell runner {#sec-runcell}

```{r run-cell}
run_cell <- function(method, n_sample, nb_boots = NULL, out = NULL,
                     sim_id_start = 1L) {
  sim_ids <- sim_id_start:(sim_id_start + n_sims - 1L)
  t0 <- proc.time()[3]
  if (identical(parallel_mode, "sims")) {
    plan("sequential"); gc(); plan("multisession", workers = n_workers)
    results <- foreach(
      s = sim_ids, .combine = rbind, .errorhandling = "remove",
      .options.future = list(packages = c("forestsearch"), seed = TRUE)
    ) %dofuture% record_replicate(s, method, n_sample, nb_boots)
    plan("sequential"); gc()
  } else {
    results <- do.call(rbind, lapply(sim_ids, function(s)
      tryCatch(record_replicate(s, method, n_sample, nb_boots),
               error = function(e) NULL)))
  }
  elapsed <- proc.time()[3] - t0
  if (is.null(results) || !nrow(results)) results <- .na_record(0L)[0, , drop = FALSE]

  # beta(Hhat) at the realized rules, on the EXACT frame.  Undetected rows take
  # the ITT complement (0 + N), so they are part of the Hhat^c denominator.
  results <- fs_attach_betaHhat(
    results, eval_df, focus = "harm", outcome_type = "continuous",
    effect_measure = "MD")

  bundle <- list(
    results = results, truth = truth,
    meta = list(n_sample = n_sample, n_sims = n_sims, nb_boots = nb_boots,
                mr_draws = mr_draws, subgroup_method = method,
                sim_id_start = sim_id_start, sim_id_end = max(sim_ids),
                seed_base = seed_base, parallel_mode = parallel_mode,
                n_workers = n_workers, elapsed_sec = elapsed,
                # --- provenance, RECORD ONLY ------------------------------
                run_tag = run_tag,
                direction = "harm",
                focus = "harm",
                outcome_type = "continuous", effect_measure = "MD",
                cal_target_md = cal_target_md,
                adverse_outcome = adverse_outcome,
                effect_threshold = md_threshold,
                consistency_threshold = md_consistency,
                threshold_orientation = "positive on the adverse_outcome=FALSE oriented scale (NOTE_threshold_sign_md.md)",
                n_super = dgm_n_super,
                seed_scheme = "pre-generated table indexed by global sim_id",
                pkg_version = as.character(utils::packageVersion("forestsearch")),
                pkg_commit = tryCatch(
                  system("git rev-parse --short HEAD", intern = TRUE)[1],
                  error = function(e) NA_character_),
                r_version = as.character(getRversion()),
                built_at = Sys.time()))
  if (!is.null(out)) saveRDS(bundle, out)
  bundle
}
```

# Run the cell {#sec-run}

```{r run-sweep}
dir.create(results_dir, showWarnings = FALSE, recursive = TRUE)
cell_path <- function(method, n)
  file.path(results_dir, sprintf("%s_md_harm_n%d_res.rds",
            if (identical(method, "consistency")) "fs" else method, n))

bundles <- list()
if (run_sweep) {
  for (m in methods_run) for (n in n_grid_run) {
    out <- cell_path(m, n)
    b <- run_cell(m, n, nb_boots = nb_boots, out = out)
    bundles[[basename(out)]] <- b
    cat(sprintf("[%s n=%d] %.0f s total, %.1f s/replicate, detected %d/%d\n",
                m, n, b$meta$elapsed_sec, b$meta$elapsed_sec / n_sims,
                sum(b$results$detected == 1L), nrow(b$results)))
  }
}
```

# Readouts {#sec-readouts}

```{r readouts}
summarise_cell <- function(b) {
  r <- b$results; tr <- b$truth
  det <- r$detected == 1L
  rd  <- r[det, , drop = FALSE]
  N   <- nrow(eval_df)

  cat(sprintf("\n================ %s, n = %d, %d replicates ================\n",
              b$meta$subgroup_method, b$meta$n_sample, b$meta$n_sims))

  ## ---- 5. partition, from the bundle -------------------------------------
  s <- r$nH_eval + r$nHc_eval
  cat("\n--- 5. PARTITION (from the bundle, all rows incl. undetected) ---\n")
  cat(sprintf("  rows: %d;  distinct nH+nHc: %s;  all == N (%d): %s\n",
              nrow(r), paste(unique(s), collapse = ", "), N, all(s == N)))
  cat(sprintf("  betaHhat_status: %s\n",
              paste(names(table(r$betaHhat_status, useNA = "ifany")),
                    table(r$betaHhat_status, useNA = "ifany"),
                    sep = "=", collapse = "  ")))
  if (!all(s == N)) stop("PARTITION FAILED — machinery defect, phase stops.")

  ## ---- 3. selection ------------------------------------------------------
  sel_rate <- mean(det)
  seff <- rd$md_trial_H - rd$betaHhat_H
  cat("\n--- 3. SELECTION ---\n")
  cat(sprintf("  selection_rate: %d/%d = %.3f\n", sum(det), nrow(r), sel_rate))
  q <- quantile(seff, c(0, .25, .5, .75, 1), na.rm = TRUE)
  cat(sprintf("  selection effect (in-trial MD - beta(Hhat)):\n"))
  cat(sprintf("    mean %.4f   SD %.4f   min %.4f   Q1 %.4f   median %.4f   Q3 %.4f   max %.4f\n",
              mean(seff, na.rm=TRUE), sd(seff, na.rm=TRUE),
              q[1], q[2], q[3], q[4], q[5]))

  ## ---- 1. bias, naive vs MR ---------------------------------------------
  b_nv <- rd$nv_H_est - rd$betaHhat_H
  b_mr <- rd$t2_H_est - rd$betaHhat_H
  se <- function(x) sd(x, na.rm = TRUE) / sqrt(sum(is.finite(x)))
  cat("\n--- 1. BIAS against the EXACT beta(Hhat) (zero MC error in the target) ---\n")
  bias <- data.frame(
    estimator = c("naive (in-trial)", "MR-corrected"),
    n         = c(sum(is.finite(b_nv)), sum(is.finite(b_mr))),
    mean_bias = c(mean(b_nv, na.rm=TRUE), mean(b_mr, na.rm=TRUE)),
    SE        = c(se(b_nv), se(b_mr)),
    SD        = c(sd(b_nv, na.rm=TRUE), sd(b_mr, na.rm=TRUE)))
  print(bias, row.names = FALSE, digits = 6)

  ## ---- 2. coverage -------------------------------------------------------
  cov1 <- function(lo, hi, tgt) {
    ok <- is.finite(lo) & is.finite(hi) & is.finite(tgt)
    c(n_eff = sum(ok),
      coverage = if (sum(ok)) mean(tgt[ok] >= lo[ok] & tgt[ok] <= hi[ok]) else NA_real_)
  }
  cH  <- cov1(rd$t2_H_lo,  rd$t2_H_hi,  rd$betaHhat_H)
  cHn <- cov1(rd$nv_H_lo,  rd$nv_H_hi,  rd$betaHhat_H)
  rc  <- r[is.finite(r$t2_Hc_lo), , drop = FALSE]
  cHc <- cov1(rc$t2_Hc_lo, rc$t2_Hc_hi, rc$betaHhat_Hc)
  cat("\n--- 2. COVERAGE of beta(Hhat) ---\n")
  covt <- data.frame(
    block = c("H", "H", "Hc"),
    interval = c("MR (IJ)", "naive", "MR (IJ)"),
    n_eff = c(cH[["n_eff"]], cHn[["n_eff"]], cHc[["n_eff"]]),
    coverage = c(cH[["coverage"]], cHn[["coverage"]], cHc[["coverage"]]))
  covt$mc_se <- sqrt(covt$coverage * (1 - covt$coverage) / pmax(covt$n_eff, 1))
  print(covt, row.names = FALSE, digits = 4)

  ## ---- 4. identification vs Q -------------------------------------------
  cat("\n--- 4. IDENTIFICATION vs Q (NPV, project convention) ---\n")
  idt <- data.frame(
    metric = c("sensitivity","specificity","PPV","NPV"),
    mean = c(mean(rd$sens,na.rm=TRUE), mean(rd$spec,na.rm=TRUE),
             mean(rd$ppv,na.rm=TRUE),  mean(rd$npv,na.rm=TRUE)),
    SD   = c(sd(rd$sens,na.rm=TRUE), sd(rd$spec,na.rm=TRUE),
             sd(rd$ppv,na.rm=TRUE),  sd(rd$npv,na.rm=TRUE)),
    min  = c(min(rd$sens,na.rm=TRUE), min(rd$spec,na.rm=TRUE),
             min(rd$ppv,na.rm=TRUE),  min(rd$npv,na.rm=TRUE)),
    max  = c(max(rd$sens,na.rm=TRUE), max(rd$spec,na.rm=TRUE),
             max(rd$ppv,na.rm=TRUE),  max(rd$npv,na.rm=TRUE)))
  print(idt, row.names = FALSE, digits = 4)

  ## ---- 6. theta-dagger reference ----------------------------------------
  cat("\n--- 6. THETA-DAGGER reference (a DISTINCT estimand from beta(Hhat)) ---\n")
  cat(sprintf("  theta_dagger_H  = %.10f   mean beta(Hhat)_H  = %.10f\n",
              tr$marg_H, mean(rd$betaHhat_H, na.rm = TRUE)))
  cat(sprintf("  theta_dagger_Hc = %.10f   mean beta(Hhat)_Hc = %.10f\n",
              tr$marg_Hc, mean(r$betaHhat_Hc, na.rm = TRUE)))
  cat("  theta-dagger is the marginal effect at the TRUE Q; beta(Hhat) is the\n")
  cat("  population effect at the REALIZED region.  They are not comparable\n")
  cat("  targets and are printed side by side only for reference.\n")

  invisible(list(bias = bias, coverage = covt, ident = idt,
                 sel_rate = sel_rate, seff = seff))
}

cell_summaries <- lapply(bundles, summarise_cell)
```

# Closed-form assertions {#sec-assertions}

Every readout with a closed-form counterpart gets it computed independently and
asserted. A failure is a machinery defect and stops the phase.

## Check A — the two-valued CATE identity on realized rules

**Provenance:** `dev/glm-continuous-sims/NOTE_target_is_collapsibility.md` §3,
and criterion 3 of `dev/glm-continuous-sims/verification/acceptance_betaHhat_md.qmd`,
whose ingredient discipline is reused verbatim.

The continuous DGM gives a two-valued subject effect: `delta` outside Q and
`delta + beta_inter` inside. For **any** region R it follows that
β(R) = δ + β_inter · P(Q | R). This is evaluated on the harness's own realized
rules, not on hand-constructed ones.

All three ingredients are **derived, never literal**: `delta` from
`compute_aor()` on Qᶜ, `beta_inter` from `dgm$model_params$beta_inter`, and
`P(Q | R)` counted on `df_super`. Nothing here is a typed constant, so the
check cannot pass by agreeing with a number someone copied in.

```{r check-A}
if (length(bundles)) {
  b <- bundles[[1]]; r <- b$results
  rules <- unique(r$sg_def[!is.na(r$sg_def) & nzchar(r$sg_def)])
  # Empty-rule guard.  With zero detections `rules` is character(0), so the
  # do.call(rbind, lapply(...)) below returns NULL and `abs(A$diff)` raises
  # "non-numeric argument to mathematical function" -- which is precisely how
  # the first pilot render died (overnight report O2.4).  An empty rule set is
  # an absence of input, not an identity violation: report it and skip.
  if (!length(rules)) {
    cat("CHECK A: SKIPPED -- 0 distinct realized rules",
        sprintf("(detected %d of %d replicates).\n",
                sum(r$detected == 1L, na.rm = TRUE), nrow(r)))
    cat("  The identity has no rules to evaluate; this is not a failure.\n")
  } else {
  inQ <- eval_df[[harm_col]] == 1L
  resolve <- function(g) {
    if (length(g) == 1L && grepl("|", g, fixed = TRUE)) g
    else if (length(g) == 1L && grepl(" & ", g, fixed = TRUE))
      strsplit(g, " & ", fixed = TRUE)[[1L]] else g
  }
  A <- do.call(rbind, lapply(rules, function(g) {
    idx <- get_dfpred(eval_df, resolve(g))$treat.recommend == 0L
    p   <- mean(inQ[idx])
    ex  <- r$betaHhat_H[match(g, r$sg_def)]
    cf  <- truth$delta + truth$beta_inter * p
    data.frame(rule = g, nH = sum(idx), P_Q = p,
               exact = ex, closed_form = cf, diff = ex - cf,
               stringsAsFactors = FALSE)
  }))
  print(A[, c("nH","P_Q","exact","closed_form","diff")], row.names = FALSE,
        digits = 12)
  mx <- max(abs(A$diff), na.rm = TRUE)
  cat(sprintf("\nCHECK A: distinct rules = %d;  max |exact - closed form| = %.3e\n",
              nrow(A), mx))
  cat(sprintf("  exactly 0 on %d of %d rules\n", sum(A$diff == 0), nrow(A)))
  # The identity is exact in real arithmetic; in floating point the two orders
  # of summation differ in the last bits.  Anything above ~1e-10 is a defect.
  if (!(mx == 0 || mx < 1e-10))
    stop(sprintf("CHECK A FAILED: max |diff| = %.3e — machinery defect.", mx))
  cat("  PASS (exact, or within floating-point reassociation)\n")
  }
}
```

## Check C — naive bias and mean selection effect are one quantity

**Provenance:** defined in the task kickoff — same quantity, two computations.

The gate's `naive` estimate and the in-trial MD computed independently in the
recorder are the same number, so their mean deviations from β(Ĥ) must agree.
The recorder's version is computed from the trial frame in `record_replicate()`
without consulting the gate, so this is a genuine second path rather than the
same value read twice.

```{r check-C}
if (length(bundles)) {
  b <- bundles[[1]]
  rd <- b$results[which(b$results$detected == 1L), , drop = FALSE]
  # Empty-detection guard, same class as CHECK A's.  With no detected rows nb
  # and sm are both NaN and `if (!(abs(nb - sm) < 1e-8))` tests NA, which is
  # itself an error.  Report the count and skip rather than erroring the
  # render (overnight report O2.5).
  if (!nrow(rd)) {
    cat("CHECK C: SKIPPED -- 0 detected replicates of", nrow(b$results), "\n")
    cat("  The two computations have nothing to compare; this is not a failure.\n")
  } else {
  nb <- mean(rd$nv_H_est  - rd$betaHhat_H, na.rm = TRUE)
  sm <- mean(rd$md_trial_H - rd$betaHhat_H, na.rm = TRUE)
  cat(sprintf("  naive mean bias        (gate)      : %.10f\n", nb))
  cat(sprintf("  mean selection effect  (recorder)  : %.10f\n", sm))
  cat(sprintf("  difference                         : %.3e\n", nb - sm))
  pd <- max(abs(rd$nv_H_est - rd$md_trial_H), na.rm = TRUE)
  cat(sprintf("  max per-replicate |gate naive - recorder MD| : %.3e\n", pd))
  if (!(abs(nb - sm) < 1e-8))
    stop(sprintf("CHECK C FAILED: %.3e — the two computations disagree.", nb - sm))
  cat("  PASS\n")
  }
}
```

# Timing and the proposed full grid {#sec-schedule}

```{r schedule}
if (length(bundles)) {
  b <- bundles[[1]]
  per <- b$meta$elapsed_sec / b$meta$n_sims
  cat(sprintf("pilot cell: n=%d, %d replicates, %.0f s total, %.2f s/replicate (%d workers)\n",
              b$meta$n_sample, b$meta$n_sims, b$meta$elapsed_sec, per, b$meta$n_workers))
  grid_n <- c(500L, 1000L, 2000L, 4000L); s_full <- 1000L
  # cost per replicate is assumed to grow with n; the pilot measures ONE n, so
  # the other three are scaled linearly and flagged as unverified.
  est <- data.frame(n = grid_n,
                    scale = grid_n / b$meta$n_sample,
                    est_s_per_rep = per * (grid_n / b$meta$n_sample))
  est$cell_hours <- est$est_s_per_rep * s_full / 3600
  print(est, row.names = FALSE, digits = 4)
  cat(sprintf("\nproposed full grid: n in {500,1000,2000,4000}, s = %d\n", s_full))
  cat(sprintf("estimated total: %.1f h  (LINEAR-IN-n ASSUMPTION, measured at n=%d only)\n",
              sum(est$cell_hours), b$meta$n_sample))
}
```

# Provenance {#sec-prov}

```{r provenance}
if (length(bundles)) str(bundles[[1]]$meta, max.level = 1)
cat(sprintf("\ndocument elapsed: %.1f min\n",
            as.numeric(difftime(Sys.time(), .t0_doc, units = "mins"))))
```

```{r session}
sessionInfo()
```
