---
title: "fs_oc_predict(): Verification and Methods Note"
subtitle: "Predicted operating characteristics of the maxeffCons search over a population-enumerated candidate family, scored against the n = 500 and n = 700 MD40 runs"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: true
  message: true
---

```{r setup}
# This document reads one precomputed object and the two tracked payloads.  It
# computes nothing heavy and types no number: every figure in prose is inline R.
.grid_path <- file.path("..", "..", "..", "..", "dev", "glm-continuous-sims",
                        "oc_wrapper_grid_2026-08-29.rds")
stopifnot(file.exists(.grid_path))
G <- readRDS(.grid_path)
stopifnot(all(c("n500_resample", "n500_split", "n700_resample", "n700_split") %in% names(G$runs)))
.pay <- lapply(G$payloads, function(p) readRDS(file.path("..", "..", "..", "..", p)))
names(.pay) <- names(G$payloads)
stopifnot(inherits(.pay[["500"]]$scale, "fs_dgm_scale"))
.reg <- G$scale$regions
.iQ <- match("Q", .reg$region); .iQc <- match("Qc", .reg$region)
piQ     <- .reg$P_g[.iQ]
tauQc   <- abs(.reg$m_tau[.iQc])
bint    <- abs(.reg$m_tau[.iQ]) - tauQc
seQ1000 <- sqrt(.reg$V_eff[.iQ] / (1000 * piQ))
c1 <- G$fs_args$effect.threshold; c2 <- G$fs_args$consistency.threshold
pcons <- G$fs_args$pconsistency.threshold
z_p <- qnorm((1 + pcons) / 2)
.f  <- function(x, d = 3) formatC(x, format = "f", digits = d)
.fm <- function(x) formatC(x, format = "d", big.mark = ",")
```

# What the wrapper computes

`fs_oc_predict()` predicts the operating characteristics of the maxeffCons
search from the joint normal law of the candidates' fitted effects. It
consumes a **candidate family** through a nine-field interface -- the label
`lab`, prevalence `Pg`, purity `PQg`, mean `beta_g`, standard error `se_g`,
classification vectors `sens_g` and `spec_g`, the overlap matrix `ovl`, and
the count `M` (plus `P(Q)`, carried explicitly for the NPV) -- and returns
declaration rates, the selection distribution and the selection-weighted
functionals. Everything above the interface is **family-agnostic**: the draw
model, the two gates, the argmax and the functionals do not look at how the
family was built. The family in this document is the population enumeration of
§2; a hand-built or stylized family (the analytic document's M = 16) goes
through the same code, which is how the wrapper was gated for fidelity
(bit-identical to the `worked-predictions` chunk at its own seed).

# The family

`fs_oc_family_enumerate()` builds the family the search ranks over as a
**population enumeration under the `forestsearch()` cut specification**, on
the DGM's super-population frame (`df_super`, `r .fm(G$scale$n_super)` rows)
rather than on a sampled trial:

1. **Cuts** from the package's own `get_FSdata()`: `conf.cont_jcuts`
   (`age = 10`, `preanti = 10`) places J cuts at the k/(J+1) **population**
   quantiles; the remaining continuous variables take the `cut_type = "default"`
   set (q25 / median / q75); binary factors enter as themselves;
   `collapse_cuts` merges coincident cuts. `r length(G$families[["500"]]$cuts)`
   cut expressions result.
2. **Both directions** of every cut, as the search's `dummy()` expansion does:
   `r G$families[["500"]]$counts[["cut_columns"]]` indicator columns.
3. **Combinations** to `maxk = 2` with the search's own combination helpers:
   `r .fm(G$families[["500"]]$counts[["enumerated"]])`.
4. **Structural floors only**, on the population frame and in the search's
   order: empty pairwise intersections (the search's status 0); `minp` per
   constituent factor; the `rmin` redundancy rule, read at trial size `n` as
   "each added factor removes more than `rmin`/`n` of the population"; and the
   size floor `Pg >= n.min/n`.
5. Candidates with **identical population membership** collapse to one.

Deliberately **excluded**: the GRF variable-importance pre-screen, the effect
screen, the consistency screen and the statistics-keyed near-duplicate removal.
The first is a sample-fitted ranking with no population counterpart; the second
and third *are* the gate the wrapper evaluates; the fourth keys on sample
statistics of candidates that have already passed the gate. Including any of
them would put sample-realized screening into an object meant to be a fixed
population law.

```{r family-table}
.fam_row <- function(n) {
  f <- G$families[[as.character(n)]]; k <- f$counts
  data.frame(n = n, `cut columns` = k[["cut_columns"]], enumerated = k[["enumerated"]],
             empty = k[["empty"]], minp = k[["minp"]], rmin = k[["rmin"]],
             `size (Pg < n.min/n)` = k[["size"]], kept = k[["kept"]],
             duplicate = k[["duplicate"]], M = k[["M"]],
             `floor n.min/n` = round(f$floor_Pg, 4), check.names = FALSE)
}
knitr::kable(rbind(.fam_row(500), .fam_row(700)), row.names = FALSE,
  caption = "The population enumeration at each stage, for the two trial sizes. The only input that changes with n is the size floor n.min/n.")
```

The realized family is M = `r .fm(G$families[["500"]]$M)` at n = 500 and
M = `r .fm(G$families[["700"]]$M)` at n = 700 -- two orders of magnitude above
the analytic document's stylized 16, which is the reason the wrapper carries a
size guard before any M-by-M allocation.

# Means and standard errors

Every mean and scale comes from `fs_dgm_scale(dgm)`, the computed
finite-population sampling scale of the difference-in-means estimator, read
from the fixture (and identical to the `$scale` object both payloads carry).
With `Q` the true harm region, `P(Q)` = `r .f(piQ, 4)`,
`tauQc` = |m_tau[Qc]| = `r .f(tauQc, 3)` and the planted interaction
`bint` = |m_tau[Q]| − |m_tau[Qc]| = `r .f(bint, 3)`:

- **Mixture identity.** A candidate `g` with purity `PQg = P(g ∩ Q)/P(g)` has
  oriented mean `beta_g = tauQc + bint · PQg`, because the effect is constant
  on `Q` and on `Qc` and `g` mixes the two in proportion `PQg`.
- **Prevalence-scaled anchor SE.** `seQ1000` = sqrt(V_eff[Q] / (1000 · P(Q)))
  = `r .f(seQ1000, 4)` is the estimator SD on `Q` at n = 1000; a candidate of
  prevalence `Pg` at trial size `n` has
  `se_g = seQ1000 · sqrt(1000/n) · sqrt(P(Q)/Pg)`.

**Caveat.** `fs_dgm_scale()` is exact for identity-scale measures only (MD, RD,
IRD); this document is MD. The SE identity treats the region size and arm split
as fixed at their expectations and uses `V_eff[Q]` for every candidate, so a
mixed candidate's within-arm variance inflation from the mixture is not
carried in `se_g`. The `sigma_D` diagnostic of §5 is what checks that this
approximation is adequate at the gate.

# The covariance

Candidates overlap, so their fitted effects are correlated. With `ovl` the
M-by-M matrix of `P(g_i ∩ g_j)`,

$$\rho_{ij} = \frac{P(g_i \cap g_j)}{\sqrt{P(g_i)\,P(g_j)}}, \qquad
  S_g = \rho \circ (se_g\, se_g^{\top}),$$

the correlation of two difference-in-means estimators sharing the overlap's
subjects. `S_g` is exactly rank-deficient by construction: for every cut, the
two directions' effects combine linearly into the full-sample effect. Draws are
generated through `fs_sym_root()`, the **symmetric** root
$V D^{1/2} V^{\top}$ with eigenvalues clamped at zero -- not $V D^{1/2}$ --
because the eigenvector basis is not a continuous function of the matrix inside
a degenerate subspace, and a pinned seed cannot fix the basis the draws are
multiplied by. The symmetric root is basis-free, so the draws are a continuous
and reproducible function of the inputs.

# The two gates

The package's consistency stage represents the random 50/50 split's two
half-sample effects as $\hat\beta \pm D$, with
$D = \sum_i G_i\, \mathrm{dfbeta}_i$ for mean-zero unit-variance multipliers
$G_i$ (Rademacher = the 50/50 split) and
$\sigma_D = \sqrt{\sum_i \mathrm{dfbeta}_i^2}$, the sandwich SE of the subgroup
treatment coefficient. A split is consistent iff both halves clear `c2`, i.e.
$|D| \le \hat\beta - c_2$, so the consistency rate is

$$P(|D| \le \hat\beta - c_2) = 2\,\Phi\!\left(\frac{\hat\beta - c_2}{\sigma_D}\right) - 1 .$$

**Production** (`consistency_method = "resample"`, the package default)
thresholds this probability at `pconsistency` = `r pcons`; the search drops a
candidate when `p.consistency < pconsistency.threshold`. Inverting,

$$2\Phi\!\left(\tfrac{\hat\beta - c_2}{\sigma_D}\right) - 1 \ge p
  \iff \hat\beta \ge c_2 + z_p\,\sigma_D, \qquad z_p = \Phi^{-1}\!\big(\tfrac{1+p}{2}\big) = `r .f(z_p, 4)`,$$

combined with the effect screen $\hat\beta \ge c_1$. The **analytic document's
historical gate** (`"split"`) instead realizes one `D`: it draws one
half-sample pair `W1`, `W2` (each with covariance `2·S_g`) and declares when
`(W1 + W2 >= 2·c1) & (W1 >= c2) & (W2 >= c2)`.

**The `sigma_D = se_g` identification.** In the draw model,
`Bhat = (W1 + W2)/2` has covariance `S_g` and `D = (W1 − W2)/2` has covariance
`S_g` as well, so the natural stand-in for `sigma_D` is `se_g`. This was
checked from source and on data rather than taken on trust
(`dev/glm-continuous-sims/sigma_d_diagnostic_2026-08-29.R`, one simulated MD40
trial at n = 500, ten candidates spanning prevalence 0.12–0.92): `sigma_D`
equals the direct dfbeta sum of squares to 1e-10, sits within 4% of the
classical model SE and within 7% of the population `se_g`, with no trend in
prevalence (Spearman −0.07 and −0.12). No constant factor is carried. Under
this identification the production gate is a single threshold on the
full-sample draw, `(Bhat >= c1) & (Bhat − c2 >= z_p · se_g)`, and only
`Bhat ~ N(beta_g, S_g)` is drawn.

So: **`"resample"` is the production screen; `"split"` is the document's
historical one.** Both are reported below.

# Selection and the functionals

Family declaration is any candidate declaring (`det`, rate `det_rate`). The
selected rule is the effect maximiser among declaring candidates -- the
maxeffCons argmax of `Bhat` with non-declaring entries masked to −∞. With
`p_sel[m]` the unconditional selection probability and
`sel_c = p_sel / det_rate` the selection distribution given declaration:

| quantity | formula |
|---|---|
| `det_rate` | `mean(det)` |
| `EnH` | `n · Σ sel_c · Pg` -- expected selected-subgroup size in subjects |
| `Esens`, `Espec` | `Σ sel_c · sens_g`, `Σ sel_c · spec_g` |
| `Eppv` | `Σ sel_c · PQg` |
| `Enpv` | `Σ sel_c · (1 − Pg − (P(Q) − PQg·Pg)) / (1 − Pg)` |
| `EbetaH` | `Σ sel_c · beta_g` -- oriented true effect on the selected rule |
| `Enaive_bias` | mean over declaring draws of `Bhat[winner] − beta_g[winner]` |
| `mass_below` | `Σ sel_c[beta_g < c1]` |

All are selection-weighted **population** functionals conditional on
declaration. Monte-Carlo SEs are `sqrt(p(1−p)/draws)` for the proportions.

# The tables

```{r tables}
.rows <- c("det_rate", "EnH", "Esens", "Espec", "Eppv", "Enpv", "EbetaH", "Enaive_bias")
.lab  <- c("detection rate", "E|Hhat| given detection", "E[sens]", "E[spec]",
           "E[PPV]", "E[NPV]", "E[beta(Hhat)] oriented", "naive bias vs beta(Hhat)")
.cell <- function(v, se, d = 3) {
  ifelse(is.na(v), "—", ifelse(is.na(se) | se == 0, .f(v, d),
                               paste0(.f(v, d), " (", .f(se, 4), ")")))
}
.table_n <- function(n) {
  r  <- G$runs[[sprintf("n%d_resample", n)]]$pred
  s  <- G$runs[[sprintf("n%d_split", n)]]$pred
  m  <- G$measured[[as.character(n)]]
  a_r <- unlist(r[.rows]); a_s <- unlist(s[.rows]); a_m <- unlist(m[.rows])
  se_r <- c(r$det_rate_se, rep(NA, 7)); se_s <- c(s$det_rate_se, rep(NA, 7))
  se_m <- unlist(m[paste0(.rows, "_se")])
  d <- c(3, 1, 3, 3, 3, 3, 2, 2)
  out <- data.frame(
    quantity = c(.lab, "M"),
    `analytic "resample"` = c(mapply(.cell, a_r, se_r, d), .fm(r$M)),
    `analytic "split"`    = c(mapply(.cell, a_s, se_s, d), .fm(s$M)),
    simulation            = c(mapply(.cell, a_m, se_m, d), "—"),
    check.names = FALSE)
  knitr::kable(out, row.names = FALSE, caption = sprintf(
    "n = %d: analytic predictions under the two gates (M = %s, %s draws, seed %d) against the measured run (%d replicates; MC SE in parentheses where it exists on both sides).",
    n, .fm(r$M), .fm(G$draws), G$seed, m$n_sims))
}
.table_n(500)
```

```{r tables-700}
.table_n(700)
```

```{r timing}
.tim <- do.call(rbind, lapply(G$runs, function(u)
  data.frame(n = u$n, gate = u$gate, M = u$pred$M, seconds = round(u$secs),
             `peak MB` = round(u$peak_MB), check.names = FALSE)))
knitr::kable(.tim, row.names = FALSE,
  caption = "Wall-clock and peak memory per evaluation. The resample branch draws one matrix; the split branch draws two.")
```

# Threshold sweeps and the inversion

```{r sweep-setup}
stopifnot(!is.null(G$sweep), !is.null(G$invert))
.sw <- G$sweep$table; .iv <- G$invert$table
.c1_fix <- G$fs_args$effect.threshold
.n_c1 <- length(G$sweep$c1)
```

The design point is that at fixed `n` the family, `Pg`, `beta_g`, `se_g` and
`S_g` are fixed, and **both gates are thresholds on draws that do not depend
on `c1` or `c2`** -- `(Bhat >= c1) & (Bhat - c2 >= z_p·se_g)` and
`(W1 + W2 >= 2·c1) & (W1 >= c2) & (W2 >= c2)`. So `fs_oc_grid()` draws
**once per `n` and gate** and sweeps the thresholds against that draw set:
the `r .n_c1` values of `c1` below cost arithmetic, not draws. Only `n`
forces re-enumeration and re-drawing, because the size floor is `n.min/n`.
The sweep's point at the driver's `c1` = `r .c1_fix` is `identical()` to the
single-point `fs_oc_predict()` evaluation of §7 (asserted in the precompute),
and one draw set was held whole (`block = Inf`), the exact path.

```{r sweep-timing}
.st <- G$sweep_timing
knitr::kable(data.frame(n = .st$n, gate = .st$consistency_method, M = .st$M,
                        `enumerate (s)` = round(.st$enumerate_secs, 1),
                        `draw (s)` = round(.st$draw_secs), 
                        `sweep (s)` = round(.st$sweep_secs),
                        `per c1 (s)` = round(.st$sweep_secs / .n_c1, 1),
                        check.names = FALSE), row.names = FALSE,
  caption = sprintf("Wall-clock split for the sweep over %d values of c1 at c2 = %g: enumeration and drawing once per (n, gate), then the sweep.", .n_c1, G$sweep$c2))
```

```{r sweep-figure}
#| fig-width: 7
#| fig-height: 5
plot(NA, xlim = range(G$sweep$c1), ylim = c(0, 1),
     xlab = "screening floor c1", ylab = "family declaration rate",
     main = sprintf("Declaration rate against c1 at c2 = %g, both gates, n = 500 and 700", G$sweep$c2))
.sty <- list(`500` = list(col = "black"), `700` = list(col = "grey40"))
for (n in c(500, 700)) for (g in c("resample", "split")) {
  r <- .sw[.sw$n == n & .sw$consistency_method == g, ]
  r <- r[order(r$c1), ]
  lines(r$c1, r$det_rate, col = .sty[[as.character(n)]]$col,
        lty = if (g == "resample") 1 else 2, lwd = if (g == "resample") 2 else 1)
}
abline(h = G$invert$targets, lty = 3)
abline(v = .c1_fix, lty = 3)
.r0 <- .sw[.sw$c1 == .c1_fix & .sw$n == 500 & .sw$consistency_method == "resample", ]
points(.c1_fix, .r0$det_rate, pch = 19); text(.c1_fix, .r0$det_rate, "  driver (30, 10)", adj = c(0, 1.3))
legend("bottomleft", bty = "n",
       legend = c("n = 500, resample", "n = 500, split", "n = 700, resample", "n = 700, split"),
       col = c("black", "black", "grey40", "grey40"), lty = c(1, 2, 1, 2), lwd = c(2, 1, 2, 1))
```

```{r sweep-invert-table}
.ivt <- .iv[order(.iv$n, .iv$consistency_method, .iv$target), ]
knitr::kable(data.frame(
  n = .ivt$n, gate = .ivt$consistency_method, target = .ivt$target,
  `c1 (inverted)` = ifelse(.ivt$attainable, .f(.ivt$c1, 2), "NA"),
  `achieved rate` = ifelse(.ivt$attainable,
                           paste0(.f(.ivt$achieved, 4), " (", .f(.ivt$achieved_se, 4), ")"),
                           paste0("infeasible (ceiling ", .f(.ivt$ceiling, 4), ")")),
  `next step` = ifelse(.ivt$attainable, .f(.ivt$next_step_rate, 4), "—"),
  check.names = FALSE), row.names = FALSE,
  caption = sprintf("fs_oc_invert(): the c1 at which the family declaration rate meets each target, at c2 = %g, per n and gate (%s draws, seed %d). The achieved rate is the step the target falls on, with its MC SE; the next-step rate is the first value below the target.",
                    G$sweep$c2, .fm(G$draws), G$seed))
```

```{r sweep-reading}
.sat <- .sw[.sw$c1 <= 60, ]
.c1_drop <- vapply(split(.sw, list(.sw$n, .sw$consistency_method)), function(r) {
  r <- r[order(r$c1), ]; r$c1[which(r$det_rate < 0.95)[1]] }, numeric(1))
```

This is the design calculation the wrapper now supports at family level: the
`c1` that gives a target declaration rate -- "80% power" under this
alternative, and the same number read as a type-I error under a null DGM --
found on fixed draws by bracketing and bisection, with the attainability
ceiling reported when the other threshold binds. Under the alternative at
this fixture the rate saturates: every curve is above 0.95 until `c1`
reaches `r .f(min(.c1_drop), 0)`–`r .f(max(.c1_drop), 0)`, so the interesting
range of `c1` lies well above the driver's `r .c1_fix`, and the inverted
values in the table are where the declaration rate actually begins to move.

# Reading the comparison

```{r reading}
.r5 <- G$runs$n500_resample$pred; .s5 <- G$runs$n500_split$pred; .m5 <- G$measured[["500"]]
.r7 <- G$runs$n700_resample$pred; .s7 <- G$runs$n700_split$pred; .m7 <- G$measured[["700"]]
```

What agreement **does** establish: that a fixed population enumeration of the
driver's cut space, with the scale table's means and SEs and either gate,
reproduces the measured selection functionals -- E|Ĥ| (`r .f(.r5$EnH, 1)` /
`r .f(.s5$EnH, 1)` against `r .f(.m5$EnH, 1)` at n = 500;
`r .f(.r7$EnH, 1)` / `r .f(.s7$EnH, 1)` against `r .f(.m7$EnH, 1)` at
n = 700), sensitivity, specificity, PPV, NPV, the oriented effect on the
selected rule and the naive selection bias -- to within the measured MC
uncertainty or close to it. That is a statement about the *family*: the
stylized M = 16 of the analytic document could not do this, and the difference
between it and the enumeration is the family, not the gate or the scale.

What agreement **does not** establish:

- **The two gates cannot be discriminated on `det_rate` here.** Under the
  alternative every candidate mean near the floor is many SEs above `c2`, so
  family declaration saturates -- `r .f(.r5$det_rate, 4)` (`"resample"`) and
  `r .f(.s5$det_rate, 4)` (`"split"`) against the measured
  `r .f(.m5$det_rate, 3)` at n = 500; `r .f(.r7$det_rate, 4)` /
  `r .f(.s7$det_rate, 4)` against `r .f(.m7$det_rate, 3)` at n = 700. A
  quantity at its ceiling on all three sides carries no information about
  which gate is right; the gates would separate under a null or near-null
  DGM, which is out of scope here.
- **The measured side is a sample-realized, screened family; the analytic side
  is a fixed population enumeration.** The driver cuts at *sample* quantiles,
  pre-screens with a causal forest, applies the effect screen and the
  statistics-keyed near-duplicate removal, and evaluates the consistency rate
  from the replicate's own dfbeta. The wrapper fixes the cuts at population
  quantiles, keeps every structural candidate, and replaces `sigma_D` by
  `se_g`. Agreement means these differences did not move the functionals at
  this fixture and these floors; it does not mean they never would.
- **The identification `sigma_D = se_g` was checked on one trial**, at ten
  prevalences, on this DGM. It is an identity check that the closed-form gate
  is computing what the production screen computes, not a general theorem.

The wrapper's next uses -- a grid over (`c1`, `c2`, `n`), the inversion for a
target rate, a null-DGM path where the gates do separate, and a ratio-measure
path -- are all built on the seam this document verifies, and none is
attempted here.
