---
title: "FS/maxeffCons on the MD/harm DGM — MR vs full-bootstrap, batchable"
subtitle: "Estimation/coverage head-to-head against the full bootstrap (FB), continuous twin of the GBSG Table-2 design"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: true
  message: true
---

# What this evaluates {#sec-intro}

Per replicate, identify $\hat H$ with forest search (`sg_focus =
"maxeffCons"`) and compute four estimates of the within-subgroup mean
difference on the same data and selected subgroup:

* **oracle** $\hat\theta(Q)$ — refit on the true region (the analysis one
  would run had the subgroup been pre-specified);
* **naive** $\hat\theta(\hat H)$ — the plug-in on the selected region;
* **FB** $\hat\theta^{*}(\hat H)$ — full nonparametric bootstrap
  bias-correction (`forestsearch_bootstrap_dofuture()`, León Eq. 7, H2)
  with its IJ interval;
* **MR** $\tilde\theta(\hat H)$ — refit-free multiplier resampling on the
  fixed fitted family, with its IJ interval.

For the consistency engine the candidate family is fixed, so FB and MR
target the same estimand and should agree to leading order — the
dfbeta-linearisation check; an FB↔MR gap here is a finding about the MR
re-selection, adjudicated by FB. Scoring uses the exact conditional
estimand β(Ĥ) computed on the DGM super-population (zero Monte Carlo
error) alongside the θ† references, kept strictly distinct.

**Swap ledger versus the survival template**
(`sim_fs_maxeffCons_fb_mr_m1_h10_knoise0_n500_batch_1_50.qmd`), every
deviation named: (1) DGM = `calibrate_glm_interaction()` /
`simulate_from_glm_dgm()` (package API; ACTG175 CD4 change, harm
direction) in place of the AFT/GBSG builders; (2) the scenario knob
`target_md_harm` defaults to −40 — the committed, verified alternative
cell — where the template runs its null (h10 = HR 1.0); the MD null
analogue is the same document with this knob changed (alignment D3);
(3) thresholds are the committed MD values (30 / 10 / 0.90); the
consistency floor sits below both CATE values — a documented open design
question (alignment D2) — and is retained unchanged here so this run is
comparable with the committed grid; (4) truth attachment uses the exported
eval API (`fs_build_eval_frame()`, exact `betaHhat_H`/`betaHhat_Hc`)
instead of a sourced script — the continuous path's package-resident
improvement; (5) the per-replicate error handling keeps the CONFIG-ERROR /
NO-DETECTION / DETECTED trichotomy and the RNG-kind pin from the committed
harness (a configuration failure must never read as a scientific null);
everything else — batching, seeding, parallel topology, stems, the
estimator set, the record schema — mirrors the template.

# Setup — all knobs {#sec-setup}

```{r setup-knobs}
#| message: false
# forestsearch must be INSTALLED (devtools::install()), not load_all(): the
# doFuture workers spawn separate R processes that only see the installed
# package.
library(forestsearch)
suppressPackageStartupMessages({
  library(doFuture); library(foreach); library(data.table)
})

# ── Parallelization topology ────────────────────────────────────────────────
#   "sims"  -> replicates fan out across workers (search + bootstrap sequential
#              within each). Best when n_sims >= n_workers.
#   "boots" -> replicates sequential; the search AND the nb_boots bootstrap fan
#              out across workers WITHIN each replicate. Best for FB runs.
# Bootstrap resample indices are seeded on the main process, so the two modes
# differ in wall-clock only, not in detection/estimates.
parallel_mode <- "sims"
n_workers     <- max(1L, parallel::detectCores() - 8L)

# ── Smoke-test counts (the only knobs to change for production) ─────────────
n_sims       <- 50L     # replicates in THIS batch          (prod: 500 -> 1000)
diag_mode    <- FALSE   # TRUE caps to 5 reps and relays worker errors verbatim
nb_boots     <- 300L    # FB bootstraps (manuscript Section 7). 0L/NULL => MR-only.
mr_draws     <- 5000L   # MR multiplier draws
k_random_noise <- 0     # add k standard-normal noise confounders (0 = none)

# ── Batching across renders (seed-disjoint pieces; pool via combine) ────────
# Every per-replicate seed is seed_base + sim_id (DGM, search, bootstrap, and
# noise), so disjoint sim_id ranges across renders pool to exactly one big run.
run_mode     <- "batch"   # "batch" (run + save one piece) | "combine" (merge)
sim_id_start <- 1L
seed_base    <- 8316951L

# ── The identifier: engine + focus (these ARE the estimator; they feed the stem)
subgroup_method <- "consistency"
sg_focus        <- "maxeffCons"  # highest effect meeting the consistency rate

# ── Scenario knob and sample size ───────────────────────────────────────────
target_md_harm <- -40     # calibrate k_inter to this MD in the harm region
                          # (raw scale; oriented +40). NULL-ANALOGUE CELL (D3):
                          # set this to the calibrator's no-interaction effect.
n_sample       <- 500L

method_tag <- if (identical(subgroup_method, "consistency")) "fs" else subgroup_method
focus_tag  <- switch(sg_focus,
                     eff = , hr = , maxcons = "maxcons",
                     effMaxSG = , hrMaxSG = "effMaxSG",
                     effMinSG = , hrMinSG = "effMinSG",
                     sg_focus)
rds_stem <- sprintf("%s_%s_fb_mr_md%02d_knoise%d_n%d",
                    method_tag, focus_tag, abs(target_md_harm),
                    k_random_noise, n_sample)
results_dir <- file.path("fb_mr_md_harm", sprintf("%s_s%d_d%d",
                         rds_stem, n_sims, mr_draws))
dir.create(results_dir, recursive = TRUE, showWarnings = FALSE)

# ── Search configuration (committed MD values; resample identification) ─────
consistency_method <- "resample"
# MR's de-biasing rule is derived by forestsearch() from sg_focus — not set
# here. Override only via mr_inference_args(reselection=) (advanced use).
selection_rule      <- "neighborhood"  # band rule for effMaxSG/effMinSG arms;
effect_neighborhood <- 0.10            # inert under maxcons-family foci; pinned
stop_threshold      <- NULL            # pinned so the resolved setting is visible
md_threshold   <- 30    # effect.threshold, oriented scale (true clearance
                        # <=> P(Q|H) >= 27.2% — alignment doc section 2)
md_consistency <- 10    # consistency.threshold. OPEN QUESTION D2: sits below
                        # both CATE values, hence non-discriminating.
pconsistency   <- 0.90
fs_splits <- 400L; maxk <- 2L; n_min <- 60L; d0_min <- 12L; d1_min <- 12L
use_lasso <- FALSE; use_dina <- FALSE; use_grf <- FALSE; use_twostage <- TRUE
is_rct <- TRUE; vi_grf_min <- -0.2
adverse_outcome <- FALSE   # higher CD4 change = better; harm = negative MD

# ── DGM knobs (NOW transplanted verbatim from the committed harness) ────────
# mr_coverage_sweep_md_harm.qmd:102-114.  Do NOT recalibrate: cal_target_md =
# -40 converges at machine precision with exactly these values.
#
# The as-received document had the SURVIVAL template's DGM parameterisation
# here, which cannot calibrate on this DGM -- see the commit message.  The
# harm region is pre-binarised into z1/z2 and the cuts are plain factor
# levels; `type = "less_eq"` is not a cutpoint type the package implements
# (generate_aft_dgm_helpers.R:750-785 accepts "greater", "multiple", "custom"
# only), and the k_inter search interval is [0, 120], not [-3, 0].
actg_arms        <- c(1L, 3L)
actg_treat_arm   <- 1L
actg_age_cut     <- 34
actg_preanti_cut <- 744.5
dgm_n_super <- 5000L
cal_k_grid_range <- c(0, 120); cal_grid_step <- 2
dgm_factor_vars   <- paste0("z", 1:12)
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)
# The ANALYSIS covariates stay the raw ACTG175 variables -- they are NOT the
# DGM's z-factors.  age and preanti enter continuous, so the search has to
# recover the z1/z2 region through its own quartile cuts; that is the design,
# and tying analysis_binary_vars to dgm_factor_vars (as received) would have
# handed the search the answer pre-binarised.
analysis_continuous_vars <- c("age","preanti","wtkg","karnof","cd40","cd80")
analysis_binary_vars     <- c("hemo","homo","drugs","race","gender","symptom")
outcome_name <- "y_sim"; treat_name <- "treat_sim"
id_name <- "id"; harm_col <- "flag_harm"
`%||%` <- function(a, b) if (is.null(a) || length(a) == 0 || all(is.na(a))) b else a
run_fb <- !is.null(nb_boots) && nb_boots >= 1L
cat(sprintf("Batch %d..%d | n=%d | FB=%s(nb=%s) | MR draws=%d | mode=%s x %d workers\n",
            sim_id_start, sim_id_start + n_sims - 1L, n_sample,
            run_fb, nb_boots %||% 0L, mr_draws, parallel_mode, n_workers))
```

# Build the DGM once (n-invariant truth) {#sec-dgm}

```{r build-dgm}
actg_df <- subset(speff2trial::ACTG175, arms %in% actg_arms)
actg_df$id <- seq_len(nrow(actg_df))
# TREATMENT IS SWITCHED (ddI = 1), verbatim from the committed harness
# (mr_coverage_sweep_md_harm.qmd:157-158).  The -40 calibration is defined
# against THIS coding; the as-received `treat <- (arms == 1)` is its negation
# and would have calibrated a different cell under the same name.
actg_df$treat_orig <- ifelse(actg_df$arms == actg_treat_arm, 1L, 0L)
actg_df$treat      <- 1L - actg_df$treat_orig
actg_df$cd4_change <- actg_df$cd420 - actg_df$cd40
actg_df <- actg_df[!is.na(actg_df$cd420), ]
actg_df$z1  <- as.factor(ifelse(actg_df$age > actg_age_cut, 1L, 0L))
actg_df$z2  <- as.factor(ifelse(actg_df$preanti <= actg_preanti_cut, 1L, 0L))
actg_df$z3  <- as.factor(ifelse(actg_df$wtkg <= 75, 1L, 0L))
actg_df$z4  <- as.factor(ifelse(actg_df$karnof <= median(actg_df$karnof), 1L, 0L))
actg_df$z5  <- as.factor(ifelse(actg_df$cd40 <= median(actg_df$cd40), 1L, 0L))
actg_df$z6  <- as.factor(ifelse(actg_df$cd80 <= median(actg_df$cd80), 1L, 0L))
actg_df$z7  <- as.factor(actg_df$hemo);  actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs); actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender);actg_df$z12 <- as.factor(actg_df$symptom)
for (v in analysis_binary_vars) actg_df[[v]] <- as.factor(actg_df[[v]])
confounders_analysis <- c(analysis_continuous_vars, analysis_binary_vars)

dgm <- calibrate_glm_interaction(
  data = actg_df, factor_vars = dgm_factor_vars,
  outcome_var = "cd4_change", treatment_var = "treat",
  target_effect = target_md_harm, outcome_type = "continuous",
  effect_measure = "MD", subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts, k_inter_range = cal_k_grid_range,
  grid_step = cal_grid_step, n_super = dgm_n_super, seed = seed_base,
  verbose = FALSE)

# Exact scoring frame: for continuous, df_super passes through unchanged —
# beta(Hhat) is an exact finite-population mean, zero Monte Carlo error.
eval_df <- fs_build_eval_frame(dgm, outcome_type = "continuous")

# truth is ASSEMBLED from the calibrated object, as in the committed harness
# (mr_coverage_sweep_md_harm.qmd:187-200).  There is no `dgm$truth` element, so
# the as-received `dgm$truth %||% dgm` fell through to dgm itself, whose
# $effect_Q is also absent -- which made the stopifnot below read
# target_md_harm on both sides and assert -40 == -40.  A vacuous guard in the
# exact position where the calibration is supposed to be verified.
inQ        <- eval_df[[harm_col]] == 1L
truth <- list(effect_Q     = dgm$hazard_ratios$harm_subgroup,
              effect_Qc    = dgm$hazard_ratios$no_harm_subgroup,
              beta_inter   = dgm$model_params$beta_inter,
              prevalence_Q = mean(inQ))
stopifnot(abs(truth$effect_Q - target_md_harm) < 1e-8)
cat(sprintf("DGM: effect_Q = %.10f | effect_Qc = %.10f | prevalence(Q) = %.4f | N_super = %d\n",
            truth$effect_Q, truth$effect_Qc, truth$prevalence_Q, nrow(eval_df)))
```

# Per-replicate machinery {#sec-machinery}

```{r machinery}
.classify <- function(sg_hat, tru) {
  out <- c(sens = NA_real_, spec = NA_real_, ppv = NA_real_, npv = NA_real_)
  if (is.null(sg_hat) || length(sg_hat) != length(tru)) return(out)
  sg_hat <- as.integer(sg_hat); tru <- as.integer(tru)
  tp <- sum(sg_hat==1 & tru==1); fp <- sum(sg_hat==1 & tru==0)
  fn <- sum(sg_hat==0 & tru==1); tn <- sum(sg_hat==0 & tru==0)
  out["sens"] <- if (tp+fn>0) tp/(tp+fn) else NA_real_
  out["spec"] <- if (tn+fp>0) tn/(tn+fp) else NA_real_
  out["ppv"]  <- if (tp+fp>0) tp/(tp+fp) else NA_real_
  out["npv"]  <- if (tn+fn>0) tn/(tn+fn) else NA_real_
  out
}

# One record; status distinguishes CONFIG-ERROR / NO-DETECTION / DETECTED.
.na_record <- function(sim_id) data.frame(
  sim_id = sim_id, status = NA_character_, detected = NA_integer_,
  mr_ok = NA_integer_, err_msg = NA_character_, mr_msg = NA_character_,
  n_sel = NA_integer_, n_harm = NA_integer_, n_true = NA_integer_,
  sg_def = NA_character_, covs = NA_character_,
  betaHhat_H = NA_real_, betaHhat_Hc = NA_real_,
  fb_secs = NA_real_, fit_mr_secs = NA_real_, fb_err = NA_character_,
  fb_src1 = NA_real_, fb_src2 = NA_real_, fb_nres = NA_integer_,
  or_H_est = NA_real_, or_H_lo = NA_real_, or_H_hi = NA_real_, or_H_se = NA_real_,
  nv_H_est = NA_real_, nv_H_lo = NA_real_, nv_H_hi = NA_real_, nv_H_se = NA_real_,
  fb_H_est = NA_real_, fb_H_lo = NA_real_, fb_H_hi = NA_real_, fb_H_se = NA_real_,
  mr_H_est = NA_real_, mr_H_lo = NA_real_, mr_H_hi = NA_real_, mr_H_se_ij = NA_real_,
  mr_Hc_est = NA_real_, mr_Hc_lo = NA_real_, mr_Hc_hi = NA_real_, mr_Hc_se_ij = NA_real_,
  ij_source = NA_character_,
  sens = NA_real_, spec = NA_real_, ppv = NA_real_, npv = NA_real_,
  stringsAsFactors = FALSE)

# Oracle: difference in means refit on the TRUE region (identity link).
.oracle_md <- function(df) {
  h <- df[[harm_col]] == 1L
  if (!any(h)) return(c(est=NA_real_, lo=NA_real_, hi=NA_real_, se=NA_real_))
  fit <- stats::lm(stats::reformulate(treat_name, response = outcome_name),
                   data = df[h, , drop = FALSE])
  co <- summary(fit)$coefficients[treat_name, ]
  est <- unname(co["Estimate"]); se <- unname(co["Std. Error"])
  s <- if (isTRUE(adverse_outcome)) 1 else -1        # oriented: + = harm
  c(est = s*est, lo = s*est - 1.96*se, hi = s*est + 1.96*se, se = se)[c(1,3,2,4)] -> v
  if (s < 0) v[c("lo","hi")] <- sort(unname(v[c("lo","hi")]))
  v
}

# Inner (within-replicate) plan.  MUST be sequential under parallel_mode =
# "sims": the FB call at the bottom of record_replicate() passes inner_parallel
# UNCONDITIONALLY, so with replicates already fanned out across n_workers an
# unguarded multisession inner plan would nest n_workers x n_workers processes.
# Under "boots" the replicate loop is sequential, so the inner plan is where all
# the parallelism lives and multisession is correct. Same pattern as the
# committed sweep (mr_coverage_sweep_md_harm.qmd:141-143).
inner_parallel <- if (identical(parallel_mode, "sims")) {
  list(plan = "sequential")
} else {
  list(plan = "multisession", workers = n_workers)
}

record_replicate <- function(sim_id) {
  rec  <- .na_record(sim_id)
  sd_i <- seed_base + sim_id
  # Pin the RNG kind: simulate_from_glm_dgm() calls set.seed(seed) with
  # kind = NULL, inheriting the ambient generator (addendum D.2).
  RNGkind("L'Ecuyer-CMRG"); set.seed(sd_i)

  df <- tryCatch(simulate_from_glm_dgm(dgm, n = n_sample, seed = sd_i),
                 error = function(e) { rec$err_msg <<- conditionMessage(e); NULL })
  if (is.null(df)) { rec$status <- "CONFIG-ERROR"; return(rec) }
  df[[id_name]] <- df[[id_name]] %||% seq_len(nrow(df))
  if (k_random_noise > 0) {
    set.seed(sd_i + 10^7)
    for (j in seq_len(k_random_noise))
      df[[paste0("noise", j)]] <- stats::rnorm(nrow(df))
  }
  confs <- c(confounders_analysis,
             if (k_random_noise > 0) paste0("noise", seq_len(k_random_noise)))
  rec$n_true <- sum(df[[harm_col]] == 1L)
  rec[c("or_H_est","or_H_lo","or_H_hi","or_H_se")] <- as.list(.oracle_md(df))

  t0 <- proc.time()[3]; fs.est <- NULL
  msgs <- tryCatch(utils::capture.output(
    fs.est <- suppressWarnings(forestsearch(
      df.analysis = df, confounders.name = confs,
      outcome.name = outcome_name, treat.name = treat_name, id.name = id_name,
      outcome_type = "continuous", effect_measure = "MD",
      effect.threshold = md_threshold, consistency.threshold = md_consistency,
      pconsistency.threshold = pconsistency, fs.splits = fs_splits,
      n.min = n_min, d0.min = d0_min, d1.min = d1_min, maxk = maxk,
      vi.grf.min = vi_grf_min, sg_focus = sg_focus,
      selection_rule = selection_rule,
      effect_neighborhood = effect_neighborhood,
      stop_threshold = stop_threshold,
      consistency_method = consistency_method,
      use_lasso = use_lasso, use_dina = use_dina, use_grf = use_grf,
      use_twostage = use_twostage, is.RCT = is_rct,
      adverse_outcome = adverse_outcome,
      details = FALSE, quiet = FALSE, seedit = sd_i,
      parallel_args = if (identical(parallel_mode, "boots")) inner_parallel else NULL,
      mr_inference = TRUE,
      mr_inference_args = list(ci_method = "ij", draws = mr_draws,
                               include_complement = TRUE))),
    type = "message"),
    error = function(e) {
      if (isTRUE(diag_mode)) warning(sprintf("FS-ERR sim %d: %s",
                                             sim_id, conditionMessage(e)))
      rec$err_msg <<- conditionMessage(e); character(0) })
  rec$fit_mr_secs <- proc.time()[3] - t0
  mr_lines <- grep("multiplier resampling|mr_inference", msgs, value = TRUE)
  if (length(mr_lines)) rec$mr_msg <- paste(mr_lines, collapse = " | ")
  if (!is.na(rec$err_msg)) { rec$status <- "CONFIG-ERROR"; return(rec) }
  if (is.null(fs.est) || is.null(fs.est$sg.harm)) {
    rec$status <- "NO-DETECTION"; rec$detected <- 0L; return(rec)
  }
  rec$status <- "DETECTED"; rec$detected <- 1L

  # The realized rule, as a single string.  fs.est$sg.harm is a CHARACTER
  # VECTOR of rule terms, not a list with a $definition element, so the
  # as-received `att$sg_def %||% fs.est$sg.harm$definition` gave NA_character_.
  # sg_def is the key fs_attach_betaHhat() joins on, so an NA here silently
  # routes every detected replicate into the undetected branch and hands it the
  # ITT complement as its target.  Collapse verbatim as both references do.
  rec$sg_def <- paste(fs.est$sg.harm, collapse = " & ")
  # `covs` (variable names only, for the identification figures) stays NA as
  # received: the survival template fills it with rule_covs(), which is not a
  # package export -- it is sourced from a helper that must sit in the render
  # working directory, and there is no such file here.  Putting the full rule
  # string in a column that downstream code reads as a variable list would be
  # worse than leaving it empty, so it is left empty and noted instead.

  # Classification of Hhat against the true harm flag, on the ANALYSIS sample.
  # fs.est$sg.harm$flag and fs.est$df_flag do not exist either; the per-subject
  # 0/1 membership vector is grp.consistency$sg.harm.id (survival
  # template:529-532), aligned to the rows of df.  Left as-received this block
  # would have written NA to n_harm/sens/spec/ppv/npv on every replicate --
  # readout 4 of the batch, gone, and gone quietly.
  sgv <- fs.est$grp.consistency$sg.harm.id
  if (!is.null(sgv) && length(sgv) == nrow(df)) {
    rec$n_harm <- sum(sgv == 1L, na.rm = TRUE)
    rec[c("sens","spec","ppv","npv")] <- as.list(.classify(sgv, df[[harm_col]]))
  }
  # Naive and MR columns live under fs.est$mr_inference, in the naive/debiased/
  # complement sub-lists -- NOT at the top level of fs.est and not under an
  # `fs.est$mr` element (there is none).  The as-received `%||%` chain over
  # fs.est[[nm]] therefore resolved to the NA seed on every replicate: a silent
  # all-NA MR arm that mr_ok, keyed on is.na(mr_H_est), would have reported as
  # a genuine MR failure.  Field paths taken verbatim from
  # mr_coverage_sweep_md_harm.qmd:365-382 and the survival template:506-526.
  #
  # mr_ok is keyed on the PRESENCE of the gate, not on a non-NA estimate, so an
  # MR that ran and legitimately returned NA is not confused with one that
  # never ran.  Detection is already recorded above and is independent of it.
  g <- fs.est$mr_inference
  rec$mr_ok <- as.integer(!is.null(g))
  if (!is.null(g)) {
    rec$n_sel     <- g$n_selected %||% NA_integer_
    rec$ij_source <- g$debiased$ij_source %||% NA_character_
    rec$nv_H_est <- g$naive$est    %||% NA_real_
    rec$nv_H_lo  <- g$naive$lower  %||% NA_real_
    rec$nv_H_hi  <- g$naive$upper  %||% NA_real_
    rec$nv_H_se  <- g$debiased$se_wald %||% NA_real_
    rec$mr_H_est <- g$debiased$est   %||% NA_real_
    rec$mr_H_lo  <- g$debiased$lower %||% NA_real_
    rec$mr_H_hi  <- g$debiased$upper %||% NA_real_
    rec$mr_H_se_ij <- g$debiased$se_ij %||% NA_real_
    if (is.list(g$complement) && !is.null(g$complement$debiased)) {
      rec$mr_Hc_est   <- g$complement$debiased$est   %||% NA_real_
      rec$mr_Hc_lo    <- g$complement$debiased$lower %||% NA_real_
      rec$mr_Hc_hi    <- g$complement$debiased$upper %||% NA_real_
      rec$mr_Hc_se_ij <- g$complement$debiased$se_ij %||% NA_real_
    }
  }

  # FB: full bootstrap of the identical fitted procedure (Eq. 7, H2).
  if (run_fb) {
    tb <- proc.time()[3]
    fb <- tryCatch(
      forestsearch_bootstrap_dofuture(
        fs.est, nb_boots = nb_boots, seed = sd_i,
        parallel_args = inner_parallel),
      error = function(e) { rec$fb_err <<- conditionMessage(e); NULL })
    rec$fb_secs <- proc.time()[3] - tb
    # forestsearch_bootstrap_dofuture() returns the bias-corrected estimates in
    # `H_estimates`, a one-row data.table with columns H0/H1/H2 and, per
    # correction order, sdH*/H*_lower/H*_upper (bootstrap_calculations_helpers.R
    # get_dfRes()).  H2 is Leon Eq.7 -- both bias sources -- which is the FB
    # comparator this document is built around.  There is no top-level fb_H_est
    # and no fb$H2 list, so the as-received chain left all four columns NA.
    #
    # SCALE.  get_dfRes() is invoked with est.loghr = FALSE for identity-scale
    # measures (bootstrap_dofuture_main.R:557-563 lists MD among them), so H2
    # and sdH2 are already on the MD scale.  The survival template's
    # .se_beta_hr(sdH2, H2) = sdH2/H2 is the delta-method step from an HR SD to
    # a log-HR SE and is deliberately NOT carried over: on an identity scale
    # sdH2 IS the standard error, and dividing by H2 would be meaningless.
    #
    # The per-column %||% NA_real_ is retained from the template for the reason
    # given there: get_dfRes() omits H2 when the second-order adjustment is
    # unavailable, and a bare NULL assignment would DELETE the column from rec.
    # rbind() then errors on the short record and one malformed replicate takes
    # down the whole batch -- neither run path survives it.
    #
    # ORIENTATION.  FB returns the RAW outcome scale (negative = harm); the MR
    # gate returns the ORIENTED scale (positive = harm).  Verified directly on
    # replicate 1: the FB uncorrected estimate and the gate's naive estimate
    # are EXACT negatives,
    #
    #   H0            = -103.1366726933
    #   g$naive$est   =  103.1366726933
    #   |H0 + naive|  =  0.000e+00
    #
    # so this is a sign flip, not a re-estimation -- the same relationship, and
    # the same hazard, that commit 1830ca92 documents between the estimator
    # columns and betaHhat/md_trial_H.  fb_* is recorded ORIENTED here so that
    # it matches its siblings or_H_*, nv_H_* and mr_H_* in this schema; left
    # raw it would read as a sign error against them in every downstream table.
    # Interval bounds are negated AND swapped (a lower bound on -X is minus the
    # upper bound on X); sdH2 is a standard deviation and is unaffected.
    if (!is.null(fb) && !is.null(fb$H_estimates)) {
      he <- fb$H_estimates
      .orient <- if (isTRUE(adverse_outcome)) 1 else -1
      rec$fb_H_est <- .orient * (he$H2 %||% NA_real_)
      fb_ci <- sort(.orient * c(he$H2_lower %||% NA_real_,
                                he$H2_upper %||% NA_real_), na.last = TRUE)
      rec$fb_H_lo  <- fb_ci[1]
      rec$fb_H_hi  <- fb_ci[2]
      rec$fb_H_se  <- he$sdH2 %||% NA_real_   # scale-free, NOT negated

      # ---- Two-source decomposition of the Eq.7 correction ------------------
      # RAW scale (not oriented), so these compose additively with H0/H2 as
      # written in the source rather than needing a sign map.
      #
      #   fb_src1 = H_star     - H_obs      (resampling noise on the ORIGINAL
      #                                      subgroup: same rule, bootstrap data)
      #   fb_src2 = Hstar_star - Hstar_obs  (RE-SELECTION: the resample's own
      #                                      subgroup, scored on its own data
      #                                      vs on the original data)
      #
      # forestsearch_bootstrap_dofuture() does NOT return the four constituents
      # -- H_obs, H_star, Hstar_star, Hstar_obs are locals inside the resample
      # loop (bootstrap_analysis_dofuture.R:488, :697, :712) and only the two
      # combinations survive into boot$results. They are nonetheless EXACTLY
      # recoverable, because the two combinations are linearly independent in
      # the two unknowns:
      #
      #   H_biasadj_1 = H_obs - (Hstar_star - Hstar_obs)
      #   H_biasadj_2 = 2*H_obs - (H_star + Hstar_star - Hstar_obs)
      #
      #   => fb_src2 = Hstar_star - Hstar_obs = H_obs - H_biasadj_1
      #   => fb_src1 = H_star     - H_obs     = H_biasadj_1 - H_biasadj_2
      #
      # H_obs is the original-sample fit on the original subgroup, which is
      # exactly H0 on the identity scale (est.loghr = FALSE for MD, so ci_est()
      # is a pass-through; verified in V_fb_adjudication.md section V1, where
      # H0 = -103.1366726933 is the exact negative of the gate's naive
      # estimate). Averaged over resamples, na.rm because a failed draw sets
      # both adjustments NA (bootstrap_analysis_dofuture.R:729-731).
      br <- fb$results
      if (!is.null(br) && all(c("H_biasadj_1","H_biasadj_2") %in% names(br))) {
        b1 <- br$H_biasadj_1; b2 <- br$H_biasadj_2
        ok <- is.finite(b1) & is.finite(b2)
        rec$fb_nres <- sum(ok)
        if (any(ok)) {
          rec$fb_src1 <- mean(b1[ok] - b2[ok])
          rec$fb_src2 <- mean((he$H0 %||% NA_real_) - b1[ok])
        }
      }
    }
  }
  rec
}
```

# Run the batch {#sec-run}

```{r run-batch}
sim_ids <- seq(sim_id_start, length.out = if (isTRUE(diag_mode)) 5L else n_sims)
if (identical(parallel_mode, "sims")) {
  registerDoFuture(); future::plan(future::multisession, workers = n_workers)
  results <- foreach(i = sim_ids, .combine = rbind,
                     .options.future = list(seed = TRUE)) %dofuture%
    record_replicate(i)
  future::plan(future::sequential)
} else {
  results <- do.call(rbind, lapply(sim_ids, record_replicate))
}

# Exact beta(Hhat) at the realized rules, on the fixed evaluation frame.
# fs_attach_betaHhat() is a POST-LOOP, whole-frame call: its signature is
# (results, frame, focus, ...) -- a results data.frame carrying an sg_def
# column, keyed to the rules it finds there -- not a per-replicate call on an
# fs.est object.  The as-received document called it inside record_replicate()
# as fs_attach_betaHhat(fs.est, eval_df = eval_df, outcome_type = ...), which
# fails three ways at once: `eval_df` is not a formal (the function takes no
# ..., so R raises "unused argument" and the tryCatch-free call aborts the
# replicate), `focus` has no default and was not supplied, and `fs.est` is not
# a results frame.  Call signature and argument values follow the committed
# mr_coverage_sweep_md_harm.qmd:411-413 exactly -- same DGM, same outcome type,
# same focus.
#
# Undetected and CONFIG-ERROR rows carry sg_def = NA and take the no-subgroup
# record (nH_eval = 0, betaHhat_Hc = the ITT effect), so they remain part of
# the Hhat^c denominator instead of dropping out.
results <- fs_attach_betaHhat(
  results, eval_df, focus = "harm", outcome_type = "continuous",
  effect_measure = "MD")

out <- file.path(results_dir, sprintf("%s_res_%d_%d.rds", rds_stem,
                 min(sim_ids), max(sim_ids)))
saveRDS(list(results = results, truth = truth,
             meta = list(n_sample = n_sample, n_sims = length(sim_ids),
                         nb_boots = nb_boots, mr_draws = mr_draws,
                         sg_focus = sg_focus, selection_rule = selection_rule,
                         consistency_method = consistency_method,
                         subgroup_method = subgroup_method,
                         target_md_harm = target_md_harm,
                         effect_threshold = md_threshold,
                         consistency_threshold = md_consistency,
                         adverse_outcome = adverse_outcome,
                         seed_base = seed_base, sim_id_start = sim_id_start,
                         parallel_mode = parallel_mode, n_workers = n_workers,
                         pkg_version = as.character(utils::packageVersion("forestsearch")),
                         built_at = Sys.time())), out)
cat("Saved:", out, "\n")
knitr::kable(within(as.data.frame(t(colMeans(
  results[results$status == "DETECTED",
          c("nv_H_est","or_H_est","fb_H_est","mr_H_est","betaHhat_H")],
  na.rm = TRUE))), {}), digits = 3,
  caption = sprintf("Batch means, detected replicates (%d/%d detected; %d mr_ok; %d FB errors).",
                    sum(results$detected %in% 1L), length(sim_ids),
                    sum(results$mr_ok %in% 1L), sum(!is.na(results$fb_err))))
```

Combine across batches with the template's `run_mode = "combine"` pattern
once multiple pieces exist; the combine-and-report twin (of
`..._combine_1_500.qmd`) follows as the next document.
