---
title: "Three-way comparison: t1_t2_legacy vs maxeffCons vs maxcons"
subtitle: "M1, target HR harm $= 1.0$, $n = 500$, 500 replicates"
date: today
format:
  pdf:
    documentclass: scrartcl
    papersize: letter
    geometry: margin=1in
    toc: false
    number-sections: true
bibliography: []
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup

data_dir <- "."

arm_files <- c(
  t1_t2_legacy = "fs_t1_t2_m1_h10_knoise0_n500_combined_1_500.rds",
  maxeffCons   = "fs_maxeffCons_fb_mr_m1_h10_knoise0_n500_combined_1_500.rds",
  maxcons      = "fs_maxcons_fb_mr_m1_h10_knoise0_n500_combined_1_500.rds"
)

#' Load simulation bundles and resolve the legacy/new column naming.
#'
#' Legacy runs record the full bootstrap as `t1_*` and multiplier resampling as
#' `t2_*`; the revised codebase records them as `fb_*` and `mr_*`.
load_arms <- function(data_dir, files = arm_files) {
  arms <- lapply(names(files), function(nm) {
    bundle <- readRDS(file.path(data_dir, files[[nm]]))
    legacy <- "t1_H_est" %in% names(bundle$results)
    list(
      name      = nm,
      results   = bundle$results,
      truth     = bundle$truth,
      meta      = bundle$meta,
      fb        = if (legacy) "t1" else "fb",
      mr        = if (legacy) "t2" else "mr",
      harm_flag = if (legacy) "t2_harm_flag" else "mr_harm_flag",
      mr_ok     = if (legacy) "gate_ok" else "mr_ok",
      fb_secs   = if (legacy) "t1_secs" else "fb_secs",
      mr_secs   = if (legacy) "t2_secs" else "fit_mr_secs"
    )
  })
  names(arms) <- names(files)
  arms
}

detected_of <- function(arm) {
  arm$results[arm$results$detected == 1, , drop = FALSE]
}

num <- function(x, digits = 3) formatC(x, format = "f", digits = digits)

arm_table <- function(arms, rows) {
  body <- lapply(rows, function(f) vapply(arms, f, character(1)))
  df <- as.data.frame(do.call(rbind, body), stringsAsFactors = FALSE)
  cbind(Metric = names(rows), df, row.names = NULL)
}

show_table <- function(df, caption) {
  knitr::kable(df, booktabs = TRUE, linesep = "", align = "l",
               escape = FALSE, caption = caption)
}

#' Align two arms on `sim_id` so replicates can be compared pairwise.
align_on_sim_id <- function(arm_a, arm_b) {
  key <- arm_a$results$sim_id
  list(a = arm_a$results,
       b = arm_b$results[match(key, arm_b$results$sim_id), , drop = FALSE])
}

arms <- load_arms(data_dir)
```

All three bundles share the design cell: `n_sample` 500, `n_sims` 500,
`nb_boots` 300, draws 5000, `seed_base` 8316951,
`consistency_method = "resample"`, `forestsearch` 0.2.0, and identical `truth`.
The arms differ only in `sg_focus`.

```{r}
#| label: tbl-design

design <- data.frame(
  arm       = names(arms),
  sg_focus  = vapply(arms, function(a) {
    if (is.null(a$meta$sg_focus)) "eff (legacy)" else a$meta$sg_focus
  }, character(1)),
  n_sample  = vapply(arms, function(a) a$meta$n_sample, integer(1)),
  nb_boots  = vapply(arms, function(a) a$meta$nb_boots, integer(1)),
  seed_base = vapply(arms, function(a) a$meta$seed_base, integer(1)),
  batches   = vapply(arms, function(a) a$meta$n_batches, integer(1)),
  row.names = NULL
)
show_table(design, "Design constants and provenance.")

stopifnot(
  all.equal(arms$t1_t2_legacy$truth, arms$maxeffCons$truth),
  all.equal(arms$t1_t2_legacy$truth, arms$maxcons$truth)
)
```

# Detection

Detection agrees replicate by replicate across all three arms.

```{r}
#| label: tbl-detection-agreement

detection_cross <- function(arm_a, arm_b) {
  al <- align_on_sim_id(arm_a, arm_b)
  sum(al$a$detected == al$b$detected)
}

pairs <- list(
  c("t1_t2_legacy", "maxeffCons"),
  c("t1_t2_legacy", "maxcons"),
  c("maxeffCons",   "maxcons")
)

detection <- data.frame(
  comparison = vapply(pairs, function(p) paste(p, collapse = " vs "), character(1)),
  agree      = vapply(pairs, function(p) {
    detection_cross(arms[[p[1]]], arms[[p[2]]])
  }, integer(1)),
  of         = nrow(arms$t1_t2_legacy$results),
  row.names  = NULL
)
show_table(detection, "Replicate-level detection agreement.")

vapply(arms, function(a) sum(a$results$detected), integer(1))
```

# Selection

```{r}
#| label: tbl-selection-agreement

#' Share of jointly detected replicates selecting the identical rule.
selection_agreement <- function(arm_a, arm_b) {
  al   <- align_on_sim_id(arm_a, arm_b)
  det  <- al$a$detected == 1 & al$b$detected == 1
  same <- det & al$a$label == al$b$label
  c(detected = sum(det, na.rm = TRUE),
    same     = sum(same, na.rm = TRUE),
    pct      = 100 * sum(same, na.rm = TRUE) / sum(det, na.rm = TRUE))
}

agreement <- do.call(rbind, lapply(pairs, function(p) {
  s <- selection_agreement(arms[[p[1]]], arms[[p[2]]])
  data.frame(comparison = paste(p, collapse = " vs "),
             detected   = s[["detected"]],
             same_rule  = s[["same"]],
             pct        = num(s[["pct"]], 1),
             row.names  = NULL)
}))
show_table(agreement, "Selection agreement on jointly detected replicates.")
```

```{r}
#| label: tbl-subgroup-size

rows_size <- list(
  "mean $\\lvert\\widehat{H}\\rvert$" = function(a) num(mean(detected_of(a)$n_harm), 1),
  "SD $\\lvert\\widehat{H}\\rvert$"   = function(a) num(sd(detected_of(a)$n_harm), 1),
  "median $\\lvert\\widehat{H}\\rvert$" = function(a) num(median(detected_of(a)$n_harm), 0),
  "min $\\lvert\\widehat{H}\\rvert$"  = function(a) num(min(detected_of(a)$n_harm), 0),
  "max $\\lvert\\widehat{H}\\rvert$"  = function(a) num(max(detected_of(a)$n_harm), 0),
  "Sensitivity" = function(a) num(mean(detected_of(a)$sens)),
  "Specificity" = function(a) num(mean(detected_of(a)$spec)),
  "PPV"         = function(a) num(mean(detected_of(a)$ppv)),
  "NPV"         = function(a) num(mean(detected_of(a)$npv)),
  "mean $\\beta(\\widehat{H})$" = function(a) num(mean(detected_of(a)$betaHhat_H))
)
show_table(arm_table(arms, rows_size),
           "Subgroup size and classification, detected replicates.")
```

# MR at coincident selection

The selection map $\mathcal{S}$ enters the correction through the per-draw
re-selected winner $\widehat{H}^{*}_{b}$, so arms with different $\mathcal{S}$
are expected to differ in MR even where the observed winner coincides. The
table below restricts to replicates on which the two arms selected the identical
rule.

```{r}
#| label: tbl-coincident

#' Compare MR and FB output on replicates where both arms selected the same
#' rule, so that any difference is attributable to the procedure rather than to
#' the region being estimated.
coincident_selection_diff <- function(arm_a, arm_b) {
  al   <- align_on_sim_id(arm_a, arm_b)
  same <- al$a$detected == 1 & al$b$detected == 1 & al$a$label == al$b$label
  same[is.na(same)] <- FALSE

  pull <- function(d, prefix, suffix) d[[paste0(prefix, "_H_", suffix)]][same]
  mr_d <- pull(al$a, arm_a$mr, "est") - pull(al$b, arm_b$mr, "est")
  hi_d <- pull(al$a, arm_a$mr, "hi")  - pull(al$b, arm_b$mr, "hi")
  fb_d <- pull(al$a, arm_a$fb, "est") - pull(al$b, arm_b$fb, "est")
  se_a <- al$a[[paste0(arm_a$mr, "_H_se_ij")]][same]
  se_b <- al$b[[paste0(arm_b$mr, "_H_se_ij")]][same]

  data.frame(
    n              = sum(same),
    mr_max_abs     = max(abs(mr_d), na.rm = TRUE),
    mr_hi_max_abs  = max(abs(hi_d), na.rm = TRUE),
    mr_mean_diff   = mean(mr_d, na.rm = TRUE),
    mr_n_exact     = sum(abs(mr_d) < 1e-10, na.rm = TRUE),
    mr_n_gt_0.01   = sum(abs(mr_d) > 0.01, na.rm = TRUE),
    mr_n_same_sign = sum(mr_d > 0, na.rm = TRUE),
    se_mean_a      = mean(se_a, na.rm = TRUE),
    se_mean_b      = mean(se_b, na.rm = TRUE),
    fb_mean_abs    = mean(abs(fb_d), na.rm = TRUE),
    row.names      = NULL
  )
}

coincident <- do.call(rbind, lapply(c("maxcons", "maxeffCons"), function(nm) {
  cbind(comparison = paste("legacy vs", nm),
        coincident_selection_diff(arms$t1_t2_legacy, arms[[nm]]))
}))

coincident_fmt <- coincident
for (j in c("mr_max_abs", "mr_hi_max_abs", "mr_mean_diff")) {
  coincident_fmt[[j]] <- formatC(coincident[[j]], format = "g", digits = 3)
}
for (j in c("se_mean_a", "se_mean_b", "fb_mean_abs")) {
  coincident_fmt[[j]] <- num(coincident[[j]], 4)
}
show_table(coincident_fmt,
           "MR and FB agreement restricted to identical-selection replicates.")
```

# Estimation and coverage

```{r}
#| label: tbl-estimation

est_summary <- function(arm, prefix, block = "H") {
  d   <- detected_of(arm)
  est <- d[[paste0(prefix, "_", block, "_est")]]
  lo  <- d[[paste0(prefix, "_", block, "_lo")]]
  hi  <- d[[paste0(prefix, "_", block, "_hi")]]
  tgt <- d[[paste0("betaHhat_", block)]]
  ok  <- is.finite(est) & is.finite(lo) & is.finite(hi) & is.finite(tgt)
  c(
    avg  = mean(est[ok]),
    len  = median((hi - lo)[ok]),
    bias = 100 * mean((est[ok] - tgt[ok]) / tgt[ok]),
    cov  = mean(tgt[ok] >= lo[ok] & tgt[ok] <= hi[ok])
  )
}

prefix_for <- function(arm, estimator) {
  switch(estimator, oracle = "or", naive = "nv", FB = arm$fb, MR = arm$mr)
}

estimation_table <- function(arms, block = "H",
                             estimators = c("oracle", "naive", "FB", "MR")) {
  metrics <- c(avg = "Avg", len = "Len", bias = "Bias (\\%)", cov = "Cov")
  parts <- lapply(estimators, function(e) {
    vals <- lapply(arms, function(a) est_summary(a, prefix_for(a, e), block))
    tab <- vapply(names(metrics), function(m) {
      digits <- if (m == "bias") 1 else 3
      vapply(vals, function(v) num(v[[m]], digits), character(1))
    }, character(length(arms)))
    data.frame(
      Estimator = c(e, rep("", length(metrics) - 1L)),
      Metric    = unname(metrics),
      t(tab),
      check.names = FALSE, row.names = NULL, stringsAsFactors = FALSE
    )
  })
  do.call(rbind, parts)
}

show_table(estimation_table(arms, "H"),
           "Harm block $\\widehat{H}$, against $\\beta(\\widehat{H})$.")
show_table(estimation_table(arms, "Hc"),
           "Complement block $\\widehat{H}^{\\mathrm{c}}$.")
```

The oracle refits on the true region $H$, so its Bias% and Cov entries are
computed against a different region from the one it estimates.

# Timings

```{r}
#| label: tbl-timings

rows_timing <- list(
  "FB mean (s)"       = function(a) num(mean(detected_of(a)[[a$fb_secs]], na.rm = TRUE), 1),
  "FB median (s)"     = function(a) num(median(detected_of(a)[[a$fb_secs]], na.rm = TRUE), 1),
  "FB min (s)"        = function(a) num(min(detected_of(a)[[a$fb_secs]], na.rm = TRUE), 1),
  "FB max (s)"        = function(a) num(max(detected_of(a)[[a$fb_secs]], na.rm = TRUE), 1),
  "fit+MR mean (s)"   = function(a) num(mean(a$results[[a$mr_secs]], na.rm = TRUE), 2),
  "fit+MR median (s)" = function(a) num(median(a$results[[a$mr_secs]], na.rm = TRUE), 2),
  "fit+MR min (s)"    = function(a) num(min(a$results[[a$mr_secs]], na.rm = TRUE), 2),
  "fit+MR max (s)"    = function(a) num(max(a$results[[a$mr_secs]], na.rm = TRUE), 2)
)
show_table(arm_table(arms, rows_timing),
           "Wall-clock seconds per replicate under 115 workers.")
```

Per-replicate cost varies by batch within each arm, so cross-arm differences
reflect render conditions as well as any code effect.

```{r}
#| label: tbl-timing-batches

#' Mean seconds per replicate within `sim_id` segments. Batch boundaries are not
#' stored numerically in `meta`, so they are read off `meta$source_files`.
timing_by_batch <- function(arm, breaks, secs = c("fb", "mr")) {
  secs <- match.arg(secs)
  d    <- arm$results
  keep <- if (secs == "fb") d$detected == 1 else rep(TRUE, nrow(d))
  col  <- if (secs == "fb") arm$fb_secs else arm$mr_secs
  seg  <- cut(d$sim_id, breaks = breaks)
  round(tapply(d[[col]][keep], seg[keep], mean, na.rm = TRUE), 2)
}

batch_breaks <- list(
  t1_t2_legacy = c(0, 20, 100, 200, 500),
  maxeffCons   = c(0, 50, 300, 500),
  maxcons      = c(0, 20, 300, 500)
)

batch_timing <- do.call(rbind, lapply(names(arms), function(nm) {
  data.frame(
    arm    = nm,
    FB     = paste(timing_by_batch(arms[[nm]], batch_breaks[[nm]], "fb"),
                   collapse = " / "),
    fit_MR = paste(timing_by_batch(arms[[nm]], batch_breaks[[nm]], "mr"),
                   collapse = " / "),
    row.names = NULL
  )
}))
show_table(batch_timing, "Mean seconds per replicate, by batch.")

#' Correlation of per-replicate cost with selected subgroup size.
vapply(arms, function(a) {
  d <- detected_of(a)
  cor(d[[a$fb_secs]], d$n_harm, use = "complete.obs")
}, numeric(1))
```

# Session information

```{r}
#| label: session
sessionInfo()
```
