---
title: "Three-way summary: t1_t2_legacy vs maxeffCons vs maxcons"
subtitle: "M1, target HR harm $= 1.0$, $n = 500$, 500 replicates"
date: today
format:
  pdf:
    documentclass: scrartcl
    papersize: letter
    geometry: margin=1in
    toc: false
    number-sections: true
bibliography: []
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup

data_dir <- "."

arm_files <- c(
  t1_t2_legacy = "fs_t1_t2_m1_h10_knoise0_n500_combined_1_500.rds",
  maxeffCons   = "fs_maxeffCons_fb_mr_m1_h10_knoise0_n500_combined_1_500.rds",
  maxcons      = "fs_maxcons_fb_mr_m1_h10_knoise0_n500_combined_1_500.rds"
)

#' Load simulation bundles and resolve the legacy/new column naming.
#'
#' Legacy runs record the full bootstrap as `t1_*` and multiplier resampling as
#' `t2_*`; the revised codebase records them as `fb_*` and `mr_*`. Detection of
#' `t1_H_est` decides which convention applies.
load_arms <- function(data_dir, files = arm_files) {
  arms <- lapply(names(files), function(nm) {
    bundle <- readRDS(file.path(data_dir, files[[nm]]))
    legacy <- "t1_H_est" %in% names(bundle$results)
    list(
      name      = nm,
      results   = bundle$results,
      truth     = bundle$truth,
      meta      = bundle$meta,
      fb        = if (legacy) "t1" else "fb",
      mr        = if (legacy) "t2" else "mr",
      harm_flag = if (legacy) "t2_harm_flag" else "mr_harm_flag",
      mr_ok     = if (legacy) "gate_ok" else "mr_ok",
      fb_secs   = if (legacy) "t1_secs" else "fb_secs",
      mr_secs   = if (legacy) "t2_secs" else "fit_mr_secs"
    )
  })
  names(arms) <- names(files)
  arms
}

detected_of <- function(arm) {
  arm$results[arm$results$detected == 1, , drop = FALSE]
}

num <- function(x, digits = 3) formatC(x, format = "f", digits = digits)

#' Evaluate a named list of functions over every arm and bind into a table.
arm_table <- function(arms, rows) {
  body <- lapply(rows, function(f) vapply(arms, f, character(1)))
  df <- as.data.frame(do.call(rbind, body), stringsAsFactors = FALSE)
  cbind(Metric = names(rows), df, row.names = NULL)
}

#' Point estimate, interval length, bias and coverage for one estimator,
#' evaluated against the conditional target $\beta(\widehat H)$.
est_summary <- function(arm, prefix, block = "H") {
  d   <- detected_of(arm)
  est <- d[[paste0(prefix, "_", block, "_est")]]
  lo  <- d[[paste0(prefix, "_", block, "_lo")]]
  hi  <- d[[paste0(prefix, "_", block, "_hi")]]
  tgt <- d[[paste0("betaHhat_", block)]]
  ok  <- is.finite(est) & is.finite(lo) & is.finite(hi) & is.finite(tgt)
  c(
    avg  = mean(est[ok]),
    len  = median((hi - lo)[ok]),
    bias = 100 * mean((est[ok] - tgt[ok]) / tgt[ok]),
    cov  = mean(tgt[ok] >= lo[ok] & tgt[ok] <= hi[ok])
  )
}

prefix_for <- function(arm, estimator) {
  switch(estimator, oracle = "or", naive = "nv", FB = arm$fb, MR = arm$mr)
}

estimation_table <- function(arms, block = "H",
                             estimators = c("oracle", "naive", "FB", "MR")) {
  metrics <- c(avg = "Avg", len = "Len", bias = "Bias (\\%)", cov = "Cov")
  parts <- lapply(estimators, function(e) {
    vals <- lapply(arms, function(a) est_summary(a, prefix_for(a, e), block))
    tab <- vapply(names(metrics), function(m) {
      digits <- if (m == "bias") 1 else 3
      vapply(vals, function(v) num(v[[m]], digits), character(1))
    }, character(length(arms)))
    data.frame(
      Estimator = c(e, rep("", length(metrics) - 1L)),
      Metric    = unname(metrics),
      t(tab),
      check.names = FALSE, row.names = NULL, stringsAsFactors = FALSE
    )
  })
  do.call(rbind, parts)
}

show_table <- function(df, caption) {
  knitr::kable(df, booktabs = TRUE, linesep = "", align = "l",
               escape = FALSE, caption = caption)
}

arms <- load_arms(data_dir)
```

All three bundles share the design cell: `n_sample` 500, `n_sims` 500,
`nb_boots` 300, draws 5000, `seed_base` 8316951,
`consistency_method = "resample"`, `forestsearch` 0.2.0, and identical `truth`.
The arms differ only in `sg_focus`: `"eff"`, `"maxeffCons"`, `"maxcons"`.

```{r}
#| label: design-check

design <- data.frame(
  arm       = names(arms),
  sg_focus  = vapply(arms, function(a) {
    if (is.null(a$meta$sg_focus)) "eff (legacy)" else a$meta$sg_focus
  }, character(1)),
  n_sample  = vapply(arms, function(a) a$meta$n_sample, integer(1)),
  n_sims    = vapply(arms, function(a) a$meta$n_sims, integer(1)),
  nb_boots  = vapply(arms, function(a) a$meta$nb_boots, integer(1)),
  seed_base = vapply(arms, function(a) a$meta$seed_base, integer(1)),
  batches   = vapply(arms, function(a) a$meta$n_batches, integer(1)),
  row.names = NULL
)
show_table(design, "Design constants and provenance.")

# truth is identical across arms
stopifnot(
  all.equal(arms$t1_t2_legacy$truth, arms$maxeffCons$truth),
  all.equal(arms$t1_t2_legacy$truth, arms$maxcons$truth)
)
```

# Detection and subgroup identification

Rows below the detection count are over the detected replicates.

```{r}
#| label: tbl-detection

rows_detection <- list(
  "Detections / 500"     = function(a) as.character(sum(a$results$detected)),
  "Detection rate (\\%)"   = function(a) num(100 * mean(a$results$detected), 1),
  "mean $\\lvert\\widehat{H}\\rvert$"          = function(a) num(mean(detected_of(a)$n_harm), 1),
  "SD $\\lvert\\widehat{H}\\rvert$"            = function(a) num(sd(detected_of(a)$n_harm), 1),
  "median $\\lvert\\widehat{H}\\rvert$"        = function(a) num(median(detected_of(a)$n_harm), 0),
  "min $\\lvert\\widehat{H}\\rvert$"           = function(a) num(min(detected_of(a)$n_harm), 0),
  "max $\\lvert\\widehat{H}\\rvert$"           = function(a) num(max(detected_of(a)$n_harm), 0),
  "mean $\\lvert\\widehat{H}\\rvert$ (\\% ITT)"  = function(a) {
    num(100 * mean(detected_of(a)$n_harm) / a$meta$n_sample, 1)
  },
  "mean true Q"          = function(a) num(mean(detected_of(a)$n_true), 1),
  "Sensitivity"          = function(a) num(mean(detected_of(a)$sens)),
  "Specificity"          = function(a) num(mean(detected_of(a)$spec)),
  "PPV"                  = function(a) num(mean(detected_of(a)$ppv)),
  "NPV"                  = function(a) num(mean(detected_of(a)$npv)),
  "MR harm flags"        = function(a) {
    as.character(sum(detected_of(a)[[a$harm_flag]], na.rm = TRUE))
  },
  "MR returned estimate" = function(a) {
    d <- detected_of(a)
    sprintf("%d / %d", sum(d[[a$mr_ok]], na.rm = TRUE), nrow(d))
  },
  "mean $\\beta(\\widehat{H})$"      = function(a) num(mean(detected_of(a)$betaHhat_H))
)

show_table(arm_table(arms, rows_detection),
           "Detection and subgroup identification.")
```

# Estimation and coverage --- harm block $\widehat{H}$

Avg is the mean point estimate, Len the median interval length; Bias% and Cov
are computed against $\beta(\widehat{H})$.

```{r}
#| label: tbl-harm

show_table(estimation_table(arms, block = "H"),
           "Harm block: estimation and coverage against $\\beta(\\widehat{H})$.")
```

The oracle refits on the true region $H$, so its Bias% and Cov entries are
computed against a different region from the one it estimates.

# Estimation and coverage --- complement block $\widehat{H}^{\mathrm{c}}$

```{r}
#| label: tbl-complement

show_table(estimation_table(arms, block = "Hc"),
           "Complement block: estimation and coverage against $\\beta(\\widehat{H}^{\\mathrm{c}})$.")
```

# Timings

Legacy records `t1_secs` and `t2_secs`; the revised arms record `fb_secs` and
`fit_mr_secs`. These are the same measured quantities. FB rows are over detected
replicates, fit+MR rows over all 500.

```{r}
#| label: tbl-timings

rows_timing <- list(
  "FB mean (s)"     = function(a) num(mean(detected_of(a)[[a$fb_secs]], na.rm = TRUE), 1),
  "FB median (s)"   = function(a) num(median(detected_of(a)[[a$fb_secs]], na.rm = TRUE), 1),
  "FB min (s)"      = function(a) num(min(detected_of(a)[[a$fb_secs]], na.rm = TRUE), 1),
  "FB max (s)"      = function(a) num(max(detected_of(a)[[a$fb_secs]], na.rm = TRUE), 1),
  "fit+MR mean (s)" = function(a) num(mean(a$results[[a$mr_secs]], na.rm = TRUE), 2),
  "fit+MR median (s)" = function(a) num(median(a$results[[a$mr_secs]], na.rm = TRUE), 2),
  "fit+MR min (s)"  = function(a) num(min(a$results[[a$mr_secs]], na.rm = TRUE), 2),
  "fit+MR max (s)"  = function(a) num(max(a$results[[a$mr_secs]], na.rm = TRUE), 2),
  "batches pooled"  = function(a) as.character(a$meta$n_batches)
)

show_table(arm_table(arms, rows_timing),
           "Wall-clock seconds per replicate under 115 workers.")
```

Timings vary by batch within each arm, so cross-arm differences reflect render
conditions as well as any code effect.

```{r}
#| label: tbl-timing-batches

#' Mean seconds per replicate within `sim_id` segments. Batch boundaries are
#' not stored numerically in `meta`, so they are supplied from
#' `meta$source_files`.
timing_by_batch <- function(arm, breaks, secs = c("fb", "mr")) {
  secs <- match.arg(secs)
  d    <- arm$results
  keep <- if (secs == "fb") d$detected == 1 else rep(TRUE, nrow(d))
  col  <- if (secs == "fb") arm$fb_secs else arm$mr_secs
  seg  <- cut(d$sim_id, breaks = breaks)
  round(tapply(d[[col]][keep], seg[keep], mean, na.rm = TRUE), 2)
}

batch_breaks <- list(
  t1_t2_legacy = c(0, 20, 100, 200, 500),
  maxeffCons   = c(0, 50, 300, 500),
  maxcons      = c(0, 20, 300, 500)
)

for (nm in names(arms)) {
  cat("\n", nm, " --- FB: ",
      paste(timing_by_batch(arms[[nm]], batch_breaks[[nm]], "fb"), collapse = " / "),
      "  |  fit+MR: ",
      paste(timing_by_batch(arms[[nm]], batch_breaks[[nm]], "mr"), collapse = " / "),
      "\n", sep = "")
}
```

# Session information

```{r}
#| label: session
sessionInfo()
```
