---
title: "Simulation Studies Based on ACTG175"
subtitle: "Adapting the ForestSearch GBSG Simulation Framework to HIV Clinical Trial Data"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ── Simulation control ──────────────────────────────────────────
# Demonstration run; increase to >= 500 for stable operating
# characteristics.  Monte Carlo SE for a 50% detection rate is
# approximately 2.2% at 500 sims and 1.0% at 2500 sims.
nsims_alt   <- 44L
nsims_null  <- 44L
sim_seed    <- 8316951L
n_workers   <- min(
  parallel::detectCores(logical = FALSE) - 1L, 120L
)

# ── Dataset preparation: GBSG ───────────────────────────────────
gbsg_er_quantile <- 0.25          # ER percentile cutpoint for ER-low definition

# ── Dataset preparation: ACTG175 ───────────────────────────────
#
# Cutpoint rationale (see Section 1.4 for full discussion):
#   z1 (age):  age <= 34  — GRF-identified split from actg175_analysis_v1.Rnw
#   z2 (wt):   wtkg <= 75 — GRF-identified prognostic split
#   z5 (CD8):  cd80 > 1292 — GRF-identified split (high CD8)
#   z3, z4:    data-driven medians (Karnofsky, CD4 baseline)
#   z6–z12:    binary indicators passed through as-is
#
actg_arms        <- c(1L, 2L)    # Trial arms to retain (1 = ddI, 2 = ZDV)
actg_treat_arm   <- 1L           # Arm coded as treatment = 1 (didanosine)
actg_age_cut     <- 34           # Age cutpoint (years): age <= cut -> z1 = 1
actg_wtkg_cut    <- 75           # Weight cutpoint (kg): wtkg <= cut -> z2 = 1
actg_cd80_cut    <- 1292         # CD8 cutpoint (cells/mm³): cd80 > cut -> z5 = 1
# z3 (Karnofsky) and z4 (CD4) use data-driven medians (see prepare-datasets)

# ── DGM construction ────────────────────────────────────────────
dgm_factor_vars   <- paste0("z", 1:12) # All binary confounder names
dgm_continuous_vars <- character(0)    # No continuous covariates
dgm_subgroup_vars <- c("z1", "z5")    # Variables defining harm subgroup H
# Within-level values that define membership in H.
#   z1 = 1 means age <= actg_age_cut   (young)
#   z5 = 1 means CD8 > actg_cd80_cut   (high CD8)
# These must match the factor coding in the data preparation chunk:
# ifelse(condition, 1L, 0L) means 1 = condition satisfied.
dgm_subgroup_cuts <- list(z1 = 1L, z5 = 1L)
dgm_k_treat       <- 1.0             # Main treatment effect multiplier (k_treat)
dgm_n_super       <- 5000L           # Super-population size for DGM
dgm_k_inter_init  <- 2.0             # Starting k_inter before calibration

# ── Calibration ─────────────────────────────────────────────────
# Target HR(H) = 2.0: treatment doubles the hazard in the harm subgroup.
# This represents a clinically meaningful level of harm and matches
# the simulation scenarios in Leon et al. (2024).  Users should choose
# their target based on the effect size they want to detect.
cal_target_hr  <- 2.0
cal_k_grid     <- seq(1.0, 5.0, by = 0.25) # k_inter grid for grid search

# ── Single trial simulation ──────────────────────────────────────
single_n          <- 1000L  # Sample size for single-trial demonstration
single_seed       <- 1L     # Seed for single-trial demonstration
single_rand_ratio <- 1      # Randomisation ratio (1:1)

# ── Simulation study ─────────────────────────────────────────────
sim_n_sample      <- 1000L  # Per-simulation sample size
# Administrative censoring time on the DGM's time scale.
# CRITICAL: must use the SAME units as the outcome variable in the
# source data.  ACTG175 uses days, so analysis_time is in days.
# Rule of thumb: set close to the observed median follow-up time,
# or the planned analysis calendar time of the trial.
sim_analysis_time <- 1200   # days (~3.3 years), near ACTG175 median follow-up
sim_cens_adjust   <- 0      # Additional censoring adjustment (log-scale)
sim_n_noise       <- 0L     # Number of noise variables added per replicate

# ── Analysis flags ───────────────────────────────────────────────
run_fs      <- TRUE   # Run ForestSearch
run_fs_grf  <- FALSE  # Run ForestSearch with GRF pre-screening only
run_grf     <- TRUE   # Run GRF-only subgroup analysis

# ── ForestSearch parameters ──────────────────────────────────────
# See Section 5.2 for detailed guidance on when to change each parameter.
fs_hr_threshold        <- 1.25   # c1: minimum HR for subgroup detection
fs_hr_consistency      <- 1.0    # c2: consistency HR threshold
fs_pconsistency_thresh <- 0.90   # Bootstrap consistency proportion threshold
fs_splits              <- 400L   # Number of ForestSearch bootstrap splits
fs_n_min               <- 60L    # Minimum subgroup size (total)
fs_d0_min              <- 12L    # Minimum events in control arm
fs_d1_min              <- 12L    # Minimum events in treatment arm
fs_maxk                <- 2L     # Maximum number of subgroup-defining variables
fs_by_risk             <- 12L    # Risk-table interval (days)
# Minimum GRF variable importance for screening.
# A negative value effectively disables GRF screening (all variables
# pass).  Increase to 0.0 or higher for stricter screening in
# datasets with many candidate variables.
fs_vi_grf_min          <- -0.2

# ── Adjusted Cox formula (used in run_simulation_analysis) ───────
# We include only the subgroup-defining variable (z1) and two strong
# prognostic confounders (z2, z3) to mimic a parsimonious clinical
# analysis.  The full set z1-z12 could be used for a saturated model
# but may over-adjust in small simulated samples.
#
# IMPORTANT: This formula must reference the simulated column names
# (y_sim, event_sim, treat_sim) and confounder names that exist in
# the DGM super-population (z1, z2, ...).  If you change the
# confounders in your dataset, update this formula to match.
cox_formula_adj <- survival::Surv(y_sim, event_sim) ~
  treat_sim + z1 + z2 + z3

# ── Detection curve ──────────────────────────────────────────────
det_theta_range <- c(0.5, 3.0)  # HR range for detection curve
det_n_points    <- 50L          # Number of points on detection curve
```


# Introduction {#sec-intro}

## Motivation

Leon et al. (2024) demonstrated the ForestSearch methodology using
the German Breast Cancer Study Group (GBSG) dataset
(`survival::gbsg`) as both the applied analysis target and the
covariate template for simulation studies.  The `setup_gbsg_dgm()`
convenience function encodes all GBSG-specific preparation choices
--- variable discretisation, subgroup definition, Weibull
calibration --- into a single reproducible call.

This document adapts the same simulation framework to the **AIDS
Clinical Trials Group Study 175** (ACTG175) dataset from the
`speff2trial` package.  ACTG175 is a randomised trial of four
antiretroviral regimens in HIV-infected subjects (Hammer et al.,
*NEJM* 1996).  The adaptation serves two purposes:

1. **Methodological validation**: confirming that the ForestSearch
   operating characteristics (detection rate, sensitivity,
   specificity, PPV) generalise beyond the GBSG breast cancer
   setting to a different disease, population, endpoint
   structure, and covariate set.

2. **Template for new-dataset simulations**: demonstrating the
   recommended workflow when `setup_gbsg_dgm()` is not
   appropriate and the analyst must use
   `generate_aft_dgm_flex()` directly.


## Adaptation Checklist {#sec-checklist}

To substitute your own trial data for ACTG175, you must address
each of the following steps.  The sections that follow implement
each one for ACTG175; cross-references are provided so you can
see the concrete implementation alongside each checklist item.

| Step | What to do | Section |
|------|-----------|---------|
| **1. Load your data** | Create `id`, `y` (observed time), `event` (0/1 indicator), and `treat` (0 = control, 1 = treatment) columns. Time units for `y` set the time scale for the entire simulation. | @sec-comparison |
| **2. Choose confounders** | Identify the candidate variables for the subgroup search. These become `factor_vars` in the DGM and `confounders_base` in the simulation. | @sec-comparison |
| **3. Discretise continuous variables** | Create binary `z1`, `z2`, ... columns from continuous covariates. Document your cutpoint rationale (see @sec-cutpoints). | @sec-comparison |
| **4. Define the harm subgroup** | Set `dgm_subgroup_vars` and `dgm_subgroup_cuts` consistently with your factor coding. Verify with an explicit `stopifnot()` check. | @sec-dgm |
| **5. Set time-scale parameters** | Ensure `sim_analysis_time` is in the same units as your outcome variable. Run the time-scale diagnostic after DGM creation. | @sec-dgm |
| **6. Update `cox_formula_adj`** | Include confounders appropriate for your study. The formula must use simulated column names (`y_sim`, `event_sim`, `treat_sim`). | Setup chunk |
| **7. Update `confounders_base`** | This must list all `z*` variable names passed to ForestSearch and GRF. It must match `factor_vars` in the DGM call. | @sec-sims |
| **8. Calibrate `k_inter`** | Run the grid search targeting a clinically meaningful HR(H). | @sec-calibrate |
| **9. Run time-scale diagnostic** | Verify `exp(dgm$model_params$mu)` is close to your observed median event time. | @sec-timescale-check |
| **10. Run censoring diagnostic** | Use `check_censoring_dgm()` to compare simulated vs observed censoring. | @sec-cens-check |
| **11. Set simulation count** | Use $\geq 500$ replicates for stable operating characteristics. | @sec-sims |


## Choosing Discretisation Cutpoints {#sec-cutpoints}

ForestSearch operates on binary (factor) covariates.  Continuous
variables must be discretised before building the DGM.  There are
three common strategies:

1. **Clinical cutpoints** --- use thresholds defined by domain
   knowledge (e.g., CD4 $< 200$ cells/mm³ for AIDS-defining
   immunosuppression).  Preferred when available.

2. **Data-driven medians** --- split at the sample median;
   appropriate when no clinical threshold exists.  Used here
   for `z3` (Karnofsky performance score) and `z4` (CD4
   baseline count).

3. **GRF-guided splits** --- run a preliminary GRF analysis on the
   observed data (`grf::causal_forest()`) to identify the
   variable-importance-weighted split points that maximise
   treatment effect heterogeneity.  This was the approach for
   `z1` (age $\leq 34$) and `z5` (CD8 $> 1292$), derived from
   the analysis in `actg175_analysis_v1.Rnw`.

For your own dataset, we recommend starting with strategy (1)
where available, falling back to (2), and optionally using (3) as
a sensitivity analysis.  Whichever approach you choose, document
the rationale alongside the cutpoint value in your setup chunk.


## Steps for Swapping GBSG with ACTG175 {#sec-swap-steps}

Replacing the GBSG template with ACTG175 requires changes at every
stage of the simulation pipeline.  The table below summarises each
change; subsequent sections implement them in detail.

| Step | GBSG (original) | ACTG175 (this document) |
|------|-----------------|-------------------------|
| **Data source** | `survival::gbsg` (686 patients) | `speff2trial::ACTG175`, arms 1 vs 2 (~1,046 patients) |
| **Outcome** | Recurrence-free survival in months (`rfstime / 30.4375`) | Time to first event in days (`days`) |
| **Event indicator** | `status` (1 = recurrence/death) | `cens` (1 = event) |
| **Treatment** | `hormon` (0/1) | `arms`: 1 = didanosine, 2 = zidovudine; coded as `treat = ifelse(arms == 1, 1, 0)` |
| **Confounders** | 7 binary variables: ER, age, meno, PgR, nodes, size, grade | 12 variables: age, weight, Karnofsky, CD4 baseline, CD8 baseline, hemophilia, homosexual, IV drugs, race, gender, prior antiretroviral, symptoms |
| **Discretisation** | ER $\leq$ Q`r gbsg_er_quantile * 100`, age $\leq$ median, meno = premenopausal, etc. | Age $\leq$ `r actg_age_cut`, weight $\leq$ `r actg_wtkg_cut` kg, CD4 $\leq$ median, CD8 $>$ `r actg_cd80_cut`, etc. (guided by GRF screening in the `.Rnw` analysis) |
| **Subgroup definition** | $H = \{\text{ER-low} \cap \text{premenopausal}\}$ ($\sim$13% prevalence) | $H = \{\text{age} \leq `r actg_age_cut` \cap \text{CD8} > `r actg_cd80_cut`\}$ ($\sim$15--20% prevalence) |
| **DGM builder** | `setup_gbsg_dgm()` | `generate_aft_dgm_flex()` |
| **Censoring** | Weibull; `analysis_time = 84` months | Weibull; `analysis_time =` `r sim_analysis_time` days ($\sim$3.3 years) |
| **ForestSearch parameters** | `n.min =` `r fs_n_min`, `maxk =` `r fs_maxk`, `hr.threshold =` `r fs_hr_threshold` | Same defaults; adjusted if needed after initial runs |

The subgroup definition for ACTG175 is derived from the GRF
screening analysis in `actg175_analysis_v1.Rnw`, which identified
age $\leq `r actg_age_cut`$ and CD8 count $> `r actg_cd80_cut`$
as the primary splits delineating a subgroup recommended for
control (i.e., potential harm from the experimental treatment).


# Side-by-Side Dataset Comparison {#sec-comparison}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(survival)
library(speff2trial)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)
```

```{r timing-init}
#| include: false
timings <- list()
t_vignette_start <- proc.time()
```


## Loading and Preparing Both Datasets

The code below prepares both datasets for the simulation framework.
For each dataset, the key steps are:

1. Create a standardised outcome variable (`y`) in the dataset's
   natural time units.
2. Create a binary treatment indicator (`treat`: 0 = control,
   1 = treatment).
3. Discretise continuous covariates into binary factors
   (`z1`, `z2`, ...) using the cutpoints defined in the setup
   chunk.
4. Construct the true harm-subgroup indicator (`flag_harm`)
   using the same variables and levels specified in
   `dgm_subgroup_vars` and `dgm_subgroup_cuts`.

```{r prepare-datasets}
#| code-fold: false

# ── GBSG ──────────────────────────────────────────────────────
gbsg_df        <- survival::gbsg
gbsg_df$y      <- gbsg_df$rfstime / 30.4375
gbsg_df$treat  <- gbsg_df$hormon

er_cut         <- quantile(gbsg_df$er, probs = gbsg_er_quantile)
gbsg_df$z1     <- as.factor(ifelse(gbsg_df$er <= er_cut, 1L, 0L))
gbsg_df$z3     <- as.factor(ifelse(gbsg_df$meno == 0, 1L, 0L))
gbsg_df$flag_harm <- as.integer(gbsg_df$z1 == 1 & gbsg_df$z3 == 1)

# ── ACTG175 ───────────────────────────────────────────────────
actg_df        <- subset(speff2trial::ACTG175, arms %in% actg_arms)
actg_df$id     <- seq_len(nrow(actg_df))
actg_df$time_days <- actg_df$days
actg_df$treat  <- ifelse(actg_df$arms == actg_treat_arm, 1L, 0L)

# ── Discretise continuous variables into binary factors ───────
# Each variable is coded so that 1 = condition satisfied.
# For example, z1 = 1 means the patient is young (age <= cutpoint).
#
# Subgroup-defining variables (from GRF screening):
# z1: age <= actg_age_cut  (subgroup factor 1)
actg_df$z1  <- as.factor(ifelse(actg_df$age  <= actg_age_cut,  1L, 0L))
# z2: weight <= actg_wtkg_cut (prognostic)
actg_df$z2  <- as.factor(ifelse(actg_df$wtkg <= actg_wtkg_cut, 1L, 0L))
# z3: Karnofsky <= median (data-driven cutpoint)
actg_df$z3  <- as.factor(ifelse(
  actg_df$karnof <= median(actg_df$karnof), 1L, 0L
))
# z4: CD4 baseline <= median (data-driven cutpoint)
actg_df$z4  <- as.factor(ifelse(
  actg_df$cd40 <= median(actg_df$cd40), 1L, 0L
))
# z5: CD8 > actg_cd80_cut  (subgroup factor 2)
actg_df$z5  <- as.factor(ifelse(actg_df$cd80 > actg_cd80_cut,  1L, 0L))
# z6-z12: binary indicators (passed through as-is from the dataset)
actg_df$z6  <- as.factor(actg_df$hemo)
actg_df$z7  <- as.factor(actg_df$homo)
actg_df$z8  <- as.factor(actg_df$drugs)
actg_df$z9  <- as.factor(actg_df$race)
actg_df$z10 <- as.factor(actg_df$gender)
actg_df$z11 <- as.factor(actg_df$oprior)
actg_df$z12 <- as.factor(actg_df$symptom)

# ── Factor coding summary ──────────────────────────────────────
# z1  = 1 => age <= 34         (young)
# z2  = 1 => weight <= 75 kg   (low weight)
# z3  = 1 => Karnofsky <= median (lower performance)
# z4  = 1 => CD4 <= median     (lower baseline CD4)
# z5  = 1 => CD8 > 1292        (high CD8)
# z6  = 1 => hemophilia yes
# z7  = 1 => homosexual activity yes
# z8  = 1 => IV drug use yes
# z9  = 1 => non-white
# z10 = 1 => male
# z11 = 1 => prior antiretroviral yes
# z12 = 1 => symptomatic yes

# ── Harm subgroup: z1 = 1 AND z5 = 1 ──────────────────────────
# This must match dgm_subgroup_vars / dgm_subgroup_cuts exactly.
actg_df$flag_harm <- as.integer(
  actg_df[[dgm_subgroup_vars[1]]] == dgm_subgroup_cuts[[dgm_subgroup_vars[1]]] &
  actg_df[[dgm_subgroup_vars[2]]] == dgm_subgroup_cuts[[dgm_subgroup_vars[2]]]
)

# Verify subgroup definition matches expectations
stopifnot(
  identical(dgm_subgroup_vars, c("z1", "z5")),
  identical(dgm_subgroup_cuts, list(z1 = 1L, z5 = 1L))
)
cat(sprintf(
  "Harm subgroup H: z1 == 1 (age <= %d) AND z5 == 1 (CD8 > %d)\n",
  actg_age_cut, actg_cd80_cut
))
cat(sprintf(
  "  H prevalence: %d / %d (%.1f%%)\n",
  sum(actg_df$flag_harm), nrow(actg_df),
  100 * mean(actg_df$flag_harm)
))
```


## Structural Comparison

```{r dataset-comparison-table}
#| tbl-cap: "Side-by-side comparison of GBSG and ACTG175 datasets"

comp <- data.frame(
  Feature = c(
    "Disease area",
    "Trial name",
    "R package",
    "Total N (used)",
    "Treatment arms",
    "Outcome",
    "Time unit",
    "Median follow-up",
    "Event rate",
    "Treatment effect (ITT HR)",
    "Subgroup definition (H)",
    "H prevalence",
    "N confounders (binary)",
    "Confounder types"
  ),
  GBSG = c(
    "Breast cancer",
    "German Breast Cancer Study Group",
    "survival::gbsg",
    sprintf("%d", nrow(gbsg_df)),
    "Hormonal therapy vs. control",
    "Recurrence-free survival",
    "Months",
    sprintf("%.1f months", median(gbsg_df$y)),
    sprintf("%.0f%%", 100 * mean(gbsg_df$status)),
    sprintf("%.3f", exp(coef(coxph(
      Surv(y, status) ~ treat, data = gbsg_df
    )))),
    sprintf("ER-low (Q%d) AND premenopausal",
            as.integer(gbsg_er_quantile * 100)),
    sprintf("%.0f%%", 100 * mean(gbsg_df$flag_harm)),
    "7",
    "All pre-discretised from continuous/categorical"
  ),
  ACTG175 = c(
    "HIV / AIDS",
    "AIDS Clinical Trials Group Study 175",
    "speff2trial::ACTG175",
    sprintf("%d", nrow(actg_df)),
    "Didanosine (ddI) vs. zidovudine (ZDV)",
    "Time to >= 50% CD4 decline, AIDS, or death",
    "Days",
    sprintf("%.0f days", median(actg_df$time_days)),
    sprintf("%.0f%%", 100 * mean(actg_df$cens)),
    sprintf("%.3f", exp(coef(coxph(
      Surv(time_days, cens) ~ treat, data = actg_df
    )))),
    sprintf("Age <= %d AND CD8 > %d", actg_age_cut, actg_cd80_cut),
    sprintf("%.0f%%", 100 * mean(actg_df$flag_harm)),
    "12",
    "Mix of binary indicators and discretised continuous"
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt(comp) |>
  tab_header(
    title = "Dataset Comparison: GBSG vs ACTG175"
  ) |>
  cols_label(Feature = "") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Feature)
  ) |>
  tab_style(
    style = cell_fill(color = "#F5F5F5"),
    locations = cells_body(rows = seq(1, nrow(comp), by = 2))
  )
```


## Confounder Mapping

```{r confounder-mapping}
#| tbl-cap: "Mapping of confounder variables between datasets"

conf_map <- data.frame(
  Role = c(
    "Subgroup factor 1",
    "Subgroup factor 2",
    "Prognostic 1",
    "Prognostic 2",
    "Prognostic 3",
    "Prognostic 4",
    "Prognostic 5",
    "Demographic 1",
    "Demographic 2",
    "Demographic 3",
    "Demographic 4",
    "Demographic 5"
  ),
  `GBSG variable` = c(
    sprintf("z1: ER <= Q%d", as.integer(gbsg_er_quantile * 100)),
    "z3: premenopausal",
    "z2: age <= median",
    "z4: PgR <= median",
    "z5: nodes <= median",
    "v6: tumour size <= median",
    "v7: grade = 3",
    rep("---", 5)
  ),
  `ACTG175 variable` = c(
    sprintf("z1: age <= %d", actg_age_cut),
    sprintf("z5: CD8 > %d", actg_cd80_cut),
    sprintf("z2: weight <= %d kg", actg_wtkg_cut),
    "z3: Karnofsky <= median",
    "z4: CD4 baseline <= median",
    "z6: hemophilia",
    "z7: homosexual activity",
    "z8: IV drug use",
    "z9: race (non-white)",
    "z10: gender (male)",
    "z11: prior antiretroviral",
    "z12: symptomatic"
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt(conf_map) |>
  tab_header(
    title = "Confounder Variable Mapping"
  ) |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(rows = 1:2)
  ) |>
  tab_source_note(
    "Highlighted rows define the harm subgroup H."
  )
```


## Survival Curves (Observed Data)

```{r km-comparison, fig.height=5, fig.width=10}
#| fig-cap: "Kaplan-Meier curves in the observed datasets. Left: GBSG (hormonal therapy). Right: ACTG175 (didanosine vs zidovudine)."

# ── GBSG KM ──────────────────────────────────────────────────
km_gbsg    <- survfit(Surv(y, status) ~ treat, data = gbsg_df)
km_gbsg_df <- data.frame(
  time = c(km_gbsg$time[1:km_gbsg$strata[1]],
           km_gbsg$time[(km_gbsg$strata[1] + 1):sum(km_gbsg$strata)]),
  surv = c(km_gbsg$surv[1:km_gbsg$strata[1]],
           km_gbsg$surv[(km_gbsg$strata[1] + 1):sum(km_gbsg$strata)]),
  arm = c(
    rep("Control", km_gbsg$strata[1]),
    rep("Hormonal", km_gbsg$strata[2])
  )
)

p_gbsg <- ggplot(km_gbsg_df, aes(x = time, y = surv, colour = arm)) +
  geom_step(linewidth = 0.8) +
  labs(
    title = "GBSG",
    x = "Months", y = "Survival Probability",
    colour = "Treatment"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# ── ACTG175 KM ───────────────────────────────────────────────
km_actg <- survfit(Surv(time_days, cens) ~ treat, data = actg_df)
km_actg_df <- data.frame(
  time = c(km_actg$time[1:km_actg$strata[1]],
           km_actg$time[(km_actg$strata[1] + 1):sum(km_actg$strata)]),
  surv = c(km_actg$surv[1:km_actg$strata[1]],
           km_actg$surv[(km_actg$strata[1] + 1):sum(km_actg$strata)]),
  arm = c(
    rep("ZDV (control)", km_actg$strata[1]),
    rep("ddI (treatment)", km_actg$strata[2])
  )
)

p_actg <- ggplot(km_actg_df, aes(x = time, y = surv, colour = arm)) +
  geom_step(linewidth = 0.8) +
  labs(
    title = "ACTG175",
    x = "Days", y = "Survival Probability",
    colour = "Treatment"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

p_gbsg + p_actg +
  plot_annotation(
    title = "Observed Survival in Source Datasets",
    theme = theme(plot.title = element_text(face = "bold"))
  )
```


## Subgroup Prevalence and Outcome

```{r subgroup-preview}
#| code-fold: false

cat("=== GBSG: Harm subgroup H = {ER-low AND premenopausal} ===\n")
cat("  N(H):", sum(gbsg_df$flag_harm), "/", nrow(gbsg_df),
    sprintf("(%.1f%%)\n", 100 * mean(gbsg_df$flag_harm)))
fit_gbsg_h <- coxph(
  Surv(y, status) ~ treat,
  data = gbsg_df[gbsg_df$flag_harm == 1, ]
)
cat("  Observed HR(H):",
    round(exp(coef(fit_gbsg_h)), 3), "\n\n")

cat(sprintf(
  "=== ACTG175: Harm subgroup H = {age<=%d AND CD8>%d} ===\n",
  actg_age_cut, actg_cd80_cut
))
cat("  N(H):", sum(actg_df$flag_harm), "/", nrow(actg_df),
    sprintf("(%.1f%%)\n", 100 * mean(actg_df$flag_harm)))
fit_actg_h <- coxph(
  Surv(time_days, cens) ~ treat,
  data = actg_df[actg_df$flag_harm == 1, ]
)
cat("  Observed HR(H):",
    round(exp(coef(fit_actg_h)), 3), "\n")
fit_actg_hc <- coxph(
  Surv(time_days, cens) ~ treat,
  data = actg_df[actg_df$flag_harm == 0, ]
)
cat("  Observed HR(Hc):",
    round(exp(coef(fit_actg_hc)), 3), "\n")
```


# Creating the ACTG175 DGM {#sec-dgm}

Because `setup_gbsg_dgm()` is specific to the GBSG dataset, we use
`generate_aft_dgm_flex()` directly.  This requires the analyst to
supply the prepared dataset, specify all modelling choices, and
define the subgroup explicitly.

The key function arguments are:

- `data`: the prepared analysis dataset with `y`, `event`,
  `treat`, and all `z*` factor columns.
- `factor_vars`: character vector of **all** binary confounder
  names (here, `z1`--`z12`).  These are used as covariates in
  the AFT model fitted to the observed data.
- `subgroup_vars` / `subgroup_cuts`: define the harm subgroup
  $H$.  `subgroup_cuts` specifies the factor level (here, `1L`)
  that indicates membership.
- `model`: `"alt"` embeds a treatment-by-subgroup interaction;
  `"null"` fits a uniform treatment effect (no HTE).
- `k_inter`: multiplier on the interaction coefficient (AFT
  log-time scale).  Higher values produce larger HR(H).
- `k_treat`: multiplier on the main treatment coefficient.
  Usually `1.0` (retain the observed ITT effect).
- `n_super`: size of the super-population drawn from the fitted
  AFT model.  $5000$ is sufficient for stable HR estimates.


## Step 1: Prepare the Analysis Dataset

The DGM function expects columns named by `outcome_var`,
`event_var`, and `treatment_var`.  We create simple aliases.

```{r prepare-actg-dgm}
#| code-fold: false

# Prepare the ACTG175 data for generate_aft_dgm_flex()
# Column names here must match the *_var arguments below.
dfa        <- actg_df
dfa$y      <- dfa$time_days   # outcome in days
dfa$event  <- dfa$cens        # 1 = event, 0 = censored
# 'treat' already exists from the preparation step above
```


## Step 2: Build the DGM (Alternative Hypothesis)

```{r create-dgm-alt}
t0 <- proc.time()[3]

dgm_alt <- generate_aft_dgm_flex(
  data            = dfa,
  outcome_var     = "y",
  event_var       = "event",
  treatment_var   = "treat",
  continuous_vars = dgm_continuous_vars,
  factor_vars     = dgm_factor_vars,
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  model           = "alt",
  k_inter         = dgm_k_inter_init,
  k_treat         = dgm_k_treat,
  n_super         = dgm_n_super,
  seed            = sim_seed,
  verbose         = TRUE
)

# Enrich the DGM with controlled direct effect (CDE) quantities.
# compute_dgm_cde() adds individual-level potential outcomes
# (theta_0, theta_1, loghr_po) to the super-population.  These are
# required by build_estimation_table() and interpret_estimation_table()
# to compute CDE bias and RMSE metrics.  Always call this after
# creating any alternative-hypothesis DGM.
dgm_alt <- compute_dgm_cde(dgm_alt)

cat("\n=== ACTG175 Alternative DGM ===\n")
cat("  HR(H):",   round(dgm_alt$hr_H_true,   3), "\n")
cat("  HR(Hc):",  round(dgm_alt$hr_Hc_true,  3), "\n")
cat("  HR(ITT):", round(dgm_alt$hr_causal,    3), "\n")
cat("  AHR(H):",  round(dgm_alt$AHR_H_true,  3), "\n")
cat("  AHR(Hc):", round(dgm_alt$AHR_Hc_true, 3), "\n")
cat("  H prevalence:",
    round(100 * mean(dgm_alt$df_super$flag_harm), 1), "%\n")

timings$dgm_alt <- proc.time()[3] - t0
```

### Time-Scale Sanity Check {#sec-timescale-check}

The single most common simulation failure is building the DGM on one
time scale (e.g., days) and then passing `analysis_time` in a different
scale (e.g., months).  This results in a follow-up window far shorter
or longer than the DGM event-time distribution, producing either
universal censoring or no censoring at all.

The diagnostic below verifies that the DGM's implied median event time
(`exp(mu)`) is in the same ballpark as the observed data and that
`analysis_time` is on the correct scale.

```{r timescale-check}
#| code-fold: false

observed_median <- median(dfa$y[dfa$event == 1])
dgm_median      <- exp(dgm_alt$model_params$mu)

cat(sprintf("Observed median event time:  %.0f days\n", observed_median))
cat(sprintf("DGM implied median (exp(mu)): %.0f days\n", dgm_median))
cat(sprintf("analysis_time:                %.0f days\n", sim_analysis_time))

if (sim_analysis_time < dgm_median * 0.5) {
  warning("analysis_time is much shorter than DGM median -- ",
          "expect heavy administrative censoring!")
}
if (sim_analysis_time > dgm_median * 10) {
  warning("analysis_time is much longer than DGM median -- ",
          "nearly all events will be observed (minimal censoring).")
}
```


## Step 3: Calibrate for Target HR {#sec-calibrate}

### Why calibrate?

The DGM's `k_inter` parameter controls the strength of the
treatment-by-subgroup interaction on the AFT (log-time) scale.
Because the mapping from `k_inter` to the Cox-model hazard ratio
HR(H) is non-linear and depends on the specific covariate
distribution, censoring structure, and baseline hazard, we cannot
set `k_inter` analytically to achieve a desired HR(H).  Instead,
we search over a grid of candidate values and select the one that
yields an HR(H) closest to the target.

A target HR(H) of `r cal_target_hr` is used here because it
represents a clinically meaningful level of harm --- treatment
doubles the hazard in the subgroup --- and matches the simulation
scenarios in León et al. (2024).  Users should choose their target
based on the effect size they want their simulation to be powered
to detect.

**Note:** `calibrate_k_inter()` automates this process for GBSG-based
DGMs.  For non-GBSG datasets, the manual grid search shown here is
required.  This is a known limitation and will be addressed in a
future release.

```{r calibrate-k-inter}
t0 <- proc.time()[3]

cal_results <- data.frame(
  k_inter = cal_k_grid,
  hr_H    = NA_real_,
  hr_Hc   = NA_real_,
  stringsAsFactors = FALSE
)

for (i in seq_along(cal_k_grid)) {
  dgm_k <- tryCatch(
    generate_aft_dgm_flex(
      data            = dfa,
      outcome_var     = "y",
      event_var       = "event",
      treatment_var   = "treat",
      continuous_vars = dgm_continuous_vars,
      factor_vars     = dgm_factor_vars,
      subgroup_vars   = dgm_subgroup_vars,
      subgroup_cuts   = dgm_subgroup_cuts,
      model           = "alt",
      k_inter         = cal_k_grid[i],
      k_treat         = dgm_k_treat,
      n_super         = dgm_n_super,
      seed            = sim_seed,
      verbose         = FALSE
    ),
    error = function(e) NULL
  )

  if (!is.null(dgm_k)) {
    cal_results$hr_H[i]  <- dgm_k$hr_H_true
    cal_results$hr_Hc[i] <- dgm_k$hr_Hc_true
  }
}

# Find k_inter closest to target
cal_valid  <- cal_results[!is.na(cal_results$hr_H), ]
best_idx   <- which.min(abs(cal_valid$hr_H - cal_target_hr))
k_inter_best <- cal_valid$k_inter[best_idx]

cat(sprintf(
  "Calibration: k_inter = %.2f gives HR(H) = %.3f (target: %.1f)\n",
  k_inter_best, cal_valid$hr_H[best_idx], cal_target_hr
))

timings$calibrate <- proc.time()[3] - t0
```


```{r calibration-plot, fig.height=5}
#| fig-cap: "Calibration of k_inter: HR(H) as a function of the interaction parameter."

ggplot(cal_valid, aes(x = k_inter)) +
  geom_line(aes(y = hr_H,  colour = "HR(H)"),  linewidth = 1) +
  geom_line(aes(y = hr_Hc, colour = "HR(Hc)"), linewidth = 1) +
  geom_hline(yintercept = cal_target_hr, linetype = "dashed",
             colour = "grey50") +
  geom_hline(yintercept = 1.0, linetype = "dotted",
             colour = "grey70") +
  geom_point(
    data = cal_valid[best_idx, ],
    aes(x = k_inter, y = hr_H),
    colour = "red", size = 4, shape = 16
  ) +
  scale_colour_manual(
    values = c("HR(H)" = "firebrick", "HR(Hc)" = "steelblue")
  ) +
  labs(
    x = "k_inter",
    y = "Hazard Ratio",
    title = "ACTG175 DGM Calibration",
    subtitle = sprintf(
      "Best: k_inter = %.2f -> HR(H) = %.3f (target: %.1f)",
      k_inter_best, cal_valid$hr_H[best_idx], cal_target_hr
    ),
    colour = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Step 4: Create Calibrated DGM

```{r dgm-calibrated}
t0 <- proc.time()[3]

dgm_calibrated <- generate_aft_dgm_flex(
  data            = dfa,
  outcome_var     = "y",
  event_var       = "event",
  treatment_var   = "treat",
  continuous_vars = dgm_continuous_vars,
  factor_vars     = dgm_factor_vars,
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  model           = "alt",
  k_inter         = k_inter_best,
  k_treat         = dgm_k_treat,
  n_super         = dgm_n_super,
  seed            = sim_seed,
  verbose         = TRUE
)
# Enrich with CDE potential outcomes (required for estimation tables)
dgm_calibrated <- compute_dgm_cde(dgm_calibrated)

cat("\n=== Calibrated ACTG175 DGM ===\n")
cat("  HR(H):",   round(dgm_calibrated$hr_H_true,  3), "\n")
cat("  HR(Hc):",  round(dgm_calibrated$hr_Hc_true, 3), "\n")
cat("  HR(ITT):", round(dgm_calibrated$hr_causal,   3), "\n")

timings$dgm_calibrated <- proc.time()[3] - t0
```


## Step 5: Create Null DGM

Under `model = "null"`, the treatment effect is uniform across
all patients (no treatment effect heterogeneity).  The `k_inter`
argument is irrelevant under the null and defaults to `1`.  Any
subgroup detected under the null is a false positive.

```{r dgm-null}
t0 <- proc.time()[3]

dgm_null <- generate_aft_dgm_flex(
  data            = dfa,
  outcome_var     = "y",
  event_var       = "event",
  treatment_var   = "treat",
  continuous_vars = dgm_continuous_vars,
  factor_vars     = dgm_factor_vars,
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  model           = "null",
  k_treat         = dgm_k_treat,
  n_super         = dgm_n_super,
  seed            = sim_seed,
  verbose         = TRUE
)

cat("\n=== ACTG175 Null DGM ===\n")
cat("  HR(ITT):", round(dgm_null$hr_causal, 3), "\n")

timings$dgm_null <- proc.time()[3] - t0
```


# Single Trial Simulation {#sec-single-trial}

Before running the full simulation study, we generate a single
simulated trial to verify that the pipeline produces sensible data.
This is a useful sanity check: the simulated event rate, subgroup
prevalence, and ITT HR should be broadly consistent with the DGM
parameters.

```{r simulate-single}
sim_data <- simulate_from_dgm(
  dgm           = dgm_calibrated,
  n             = single_n,
  rand_ratio    = single_rand_ratio,
  seed          = single_seed,
  analysis_time = sim_analysis_time,
  cens_adjust   = sim_cens_adjust
)

cat("Simulated ACTG175 trial:\n")
cat("  N =", nrow(sim_data), "\n")
cat("  Events =", sum(sim_data$event_sim),
    sprintf("(%.1f%%)\n", 100 * mean(sim_data$event_sim)))
cat("  Harm subgroup size =", sum(sim_data$flag_harm),
    sprintf("(%.1f%%)\n", 100 * mean(sim_data$flag_harm)))

fit_sim <- coxph(Surv(y_sim, event_sim) ~ treat_sim, data = sim_data)
cat("  Estimated ITT HR =", round(exp(coef(fit_sim)), 3), "\n")
```

## Censoring Diagnostic {#sec-cens-check}

The `check_censoring_dgm()` function compares the simulated censoring
distribution against the observed data used to build the DGM.  Large
discrepancies suggest a time-scale mismatch or a poorly calibrated
censoring model.

```{r censoring-diagnostic}
check_censoring_dgm(
  dgm           = dgm_calibrated,
  n             = single_n,
  analysis_time = sim_analysis_time,
  seed          = single_seed
)
```


# Running Simulation Studies {#sec-sims}

## Setup

```{r sim-setup}
#| code-fold: false

plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")

# ── Candidate variable set ────────────────────────────────────
# confounders_base tells run_simulation_analysis() which columns in
# the simulated data to pass to ForestSearch and GRF as candidate
# subgroup-defining variables.  These must:
#   1. Be present in the super-population (dgm$df_super)
#   2. Be binary factors (0/1) for ForestSearch
#   3. Match the names used in factor_vars when building the DGM
#
# If your dataset uses different variable names (e.g. "age_bin",
# "cd4_low"), update BOTH factor_vars (in the DGM call) and
# confounders_base (here) to match.
confounders_actg <- dgm_factor_vars

# ── ForestSearch parameters ───────────────────────────────────
# These are passed to run_simulation_analysis() and forwarded to
# forestsearch().  The annotations below explain each parameter
# and when you might change it for your own study.
fs_params <- list(
  # Column name mappings — must match simulate_from_dgm() output
  outcome.name           = "y_sim",
  event.name             = "event_sim",
  treat.name             = "treat_sim",
  id.name                = "id",

  # Variable screening flags
  use_lasso              = TRUE,    # LASSO pre-screening (recommended)
  use_grf                = TRUE,    # GRF pre-screening (recommended)

  # Subgroup detection thresholds (see Leon et al. 2024, Section 3)
  # c1: a candidate subgroup must have HR > hr.threshold to be flagged.
  # Increase for more conservative detection; decrease for more power.
  hr.threshold           = fs_hr_threshold,
  # c2: in each bootstrap replicate, the subgroup HR must exceed this
  # value.  Typically set to 1.0 (any harm signal).
  hr.consistency         = fs_hr_consistency,
  # The proportion of bootstrap replicates in which the subgroup HR
  # exceeds c2.  The paper default is 0.90.
  pconsistency.threshold = fs_pconsistency_thresh,

  # Computational and sample-size constraints
  fs.splits              = fs_splits,    # Bootstrap replicates (400 is default)
  n.min                  = fs_n_min,     # Minimum subgroup size (both arms)
  d0.min                 = fs_d0_min,    # Minimum events, control arm
  d1.min                 = fs_d1_min,    # Minimum events, treatment arm
  maxk                   = fs_maxk,      # Max subgroup-defining variables
  by.risk                = fs_by_risk,   # Risk-table granularity (time units)
  vi.grf.min             = fs_vi_grf_min, # GRF variable importance threshold

  # Two-stage estimation: after identifying a subgroup, re-estimates
  # the treatment effect using an independent hold-out sample.
  # Recommended for bias correction.
  use_twostage           = TRUE,
  twostage_args          = list(),

  # Inner parallelism: ForestSearch runs sequentially within each
  # worker.  The outer loop (foreach) parallelises across replicates.
  parallel_args          = list(
    plan = "sequential", workers = 1, show_message = FALSE
  )
)

# ── Simulation configuration ─────────────────────────────────
sim_config <- list(
  n_sample      = sim_n_sample,
  analysis_time = sim_analysis_time,
  cens_adjust   = sim_cens_adjust
)
```


## Alternative Hypothesis Simulations

```{r run-alt}
cat("Running", nsims_alt, "simulations under H1...\n")
t0 <- proc.time()[3]

results_alt <- foreach(
  sim = seq_len(nsims_alt),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  # Reset plan inside workers to prevent nested parallelism.
  # Each worker runs ForestSearch sequentially; the outer foreach
  # loop parallelises across simulation replicates.
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_calibrated,
    n_sample         = sim_config$n_sample,
    analysis_time    = sim_config$analysis_time,
    cens_adjust      = sim_config$cens_adjust,
    confounders_base = confounders_actg,
    cox_formula_adj  = cox_formula_adj,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    verbose          = FALSE
  )
}

timings$sims_alt <- proc.time()[3] - t0
cat("Completed in", round(timings$sims_alt / 60, 1), "minutes\n")
cat("Results:", nrow(results_alt), "rows\n")
```


## Null Hypothesis Simulations

```{r run-null}
cat("Running", nsims_null, "simulations under H0...\n")
t0 <- proc.time()[3]

results_null <- foreach(
  sim = seq_len(nsims_null),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  # Reset plan inside workers to prevent nested parallelism.
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_null,
    n_sample         = sim_config$n_sample,
    analysis_time    = sim_config$analysis_time,
    cens_adjust      = sim_config$cens_adjust,
    confounders_base = confounders_actg,
    cox_formula_adj  = cox_formula_adj,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    verbose          = FALSE
  )
}

timings$sims_null <- proc.time()[3] - t0
cat("Completed in", round(timings$sims_null / 60, 1), "minutes\n")
```


# Results {#sec-results}

## Metrics Glossary {#sec-glossary}

The operating-characteristics tables below report three groups of
metrics:

- **detection**: the proportion of simulations where ForestSearch
  (or GRF) identifies *any* subgroup ($\texttt{any.H} = 1$).
  Under the alternative, this is detection power; under the null,
  it is the false-positive rate (type I error).

- **classification**: among simulations where a subgroup was
  detected, how well the identified subgroup $\hat{H}$ overlaps
  with the true subgroup $H$.  Includes sensitivity (proportion
  of true $H$ members captured), specificity (proportion of
  $H^c$ members correctly excluded), and positive predictive value
  (PPV, proportion of $\hat{H}$ that truly belongs to $H$).

- **cde_estimates**: bias and root-mean-square error (RMSE) of
  the controlled direct effect (CDE) estimator for the treatment
  effect within $\hat{H}$ and $\hat{H}^c$, relative to the true
  DGM values.  The CDE is defined at the individual level via
  potential outcomes and represents the causal estimand
  (see León et al. 2024, Section 2.2).


## Operating Characteristics Under H1

```{r oc-alt}
t0 <- proc.time()[3]

format_oc_results(
  results_alt,
  metrics   = c("detection", "classification", "cde_estimates"),
  digits    = 2,
  digits_hr = 2,
  title     = "ACTG175: Identification Properties (HTEs)",
  subtitle  = sprintf(
    "n = %d, %d sims, HR(H) = %.2f, HR(Hc) = %.2f",
    sim_config$n_sample, nsims_alt,
    dgm_calibrated$hr_H_true, dgm_calibrated$hr_Hc_true
  )
)

build_estimation_table(
  results         = results_alt,
  dgm             = dgm_calibrated,
  analysis_method = "FS",
  subtitle = sprintf(
    "ACTG175 DGM | n = %d, %d sims",
    sim_config$n_sample, nsims_alt
  ),
  font_size = 16
)

timings$summarize_alt <- proc.time()[3] - t0
```

### Interpretation

```{r interpret-alt, results="asis"}
interpret_estimation_table(
  results_alt, dgm_calibrated,
  analysis_method = "FS",
  scenario = "alt"
)
```


## Operating Characteristics Under H0

```{r oc-null}
t0 <- proc.time()[3]

format_oc_results(
  results_null,
  metrics  = c("detection", "classification"),
  digits   = 2,
  title    = "ACTG175: Type I Error (Null)",
  subtitle = sprintf(
    "n = %d, %d sims, HR(ITT) = %.2f",
    sim_config$n_sample, nsims_null,
    dgm_null$hr_causal
  )
)

timings$summarize_null <- proc.time()[3] - t0
```


## Classification Table

```{r classification}
scenario_list <- list(
  null = list(
    results    = results_null,
    label      = "ACTG",
    n_sample   = sim_config$n_sample,
    dgm        = dgm_null,
    hypothesis = "null"
  ),
  alt = list(
    results    = results_alt,
    label      = "ACTG",
    n_sample   = sim_config$n_sample,
    dgm        = dgm_calibrated,
    hypothesis = "alt"
  )
)

build_classification_table(
  scenario_results = scenario_list,
  analyses = sort(unique(c(
    unique(results_null$analysis),
    unique(results_alt$analysis)
  ))),
  digits    = 2,
  font_size = 16,
  title     = "ACTG175: Subgroup Identification and Classification",
  n_sims    = nsims_alt
)
```


# Theoretical Detection Probability {#sec-detection}

The detection probability is the theoretical power of ForestSearch's
bootstrap consistency rule to flag a subgroup with a given HR, given
the expected subgroup size and censoring proportion.  This curve
provides a reference for interpreting the empirical detection rate
observed in the simulation.

The `method = "cubature"` argument uses numerical integration
(via the `cubature` package) to compute the joint probability
that the subgroup HR exceeds `hr.threshold` (c1) and that the
proportion of bootstrap replicates exceeding `hr.consistency`
(c2) is at least `pconsistency.threshold`.  An alternative
`method = "normal"` uses a normal approximation.

```{r detection-curve, fig.height=6}
n_sg_expected <- sim_config$n_sample *
  mean(dgm_calibrated$df_super$flag_harm)
prop_cens <- mean(results_alt$p.cens)

cat("Expected subgroup size:", round(n_sg_expected), "\n")
cat("Censoring proportion:",   round(prop_cens, 3), "\n")
cat("True HR(H):",             round(dgm_calibrated$hr_H_true, 3), "\n")

prob_detect <- compute_detection_probability(
  theta          = dgm_calibrated$hr_H_true,
  n_sg           = round(n_sg_expected),
  prop_cens      = prop_cens,
  hr_threshold   = fs_params$hr.threshold,
  hr_consistency = fs_params$hr.consistency,
  method         = "cubature"
)

cat("\nTheoretical P(detect):", round(prob_detect, 3), "\n")
cat("Empirical FS:",
    round(mean(results_alt[
      results_alt$analysis == "FS", "any.H"
    ]), 3), "\n")

# Generate curve
detection_curve <- generate_detection_curve(
  theta_range       = det_theta_range,
  n_points          = det_n_points,
  n_sg              = round(n_sg_expected),
  prop_cens         = prop_cens,
  hr_threshold      = fs_params$hr.threshold,
  hr_consistency    = fs_params$hr.consistency,
  include_reference = TRUE,
  verbose           = FALSE
)

plot_detection_curve(
  detection_curve,
  add_reference_lines = TRUE,
  add_threshold_line  = TRUE,
  title = sprintf(
    "Detection Curve (n_sg=%d, cens=%.0f%%, c1=%.2f)",
    round(n_sg_expected), 100 * prop_cens,
    fs_params$hr.threshold
  )
)

abline(v = dgm_calibrated$hr_H_true, lty = 2, col = "blue")
```


# Timing Summary {#sec-timing}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Stage = c(
    "DGM creation (H1)",
    "k_inter calibration",
    "DGM calibrated",
    "DGM creation (H0)",
    "Simulations H1",
    "Simulations H0",
    "Summarize H1",
    "Summarize H0",
    "Total"
  ),
  Seconds = c(
    timings$dgm_alt,
    timings$calibrate,
    timings$dgm_calibrated,
    timings$dgm_null,
    timings$sims_alt,
    timings$sims_null,
    timings$summarize_alt,
    timings$summarize_null,
    timings$total
  ),
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(
    title    = "Computational Timing Summary",
    subtitle = sprintf(
      "%d H1 + %d H0 sims | %d workers",
      nsims_alt, nsims_null, n_workers
    )
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  cols_label(
    Stage   = "Stage",
    Seconds = "Time (sec)",
    Minutes = "Time (min)"
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total")
  )
```


# Discussion {#sec-discussion}

## Key Differences from the GBSG Simulation

The ACTG175 dataset introduces several features that test
ForestSearch under different conditions than the GBSG template:

1. **Larger sample size** ($N \approx 1{,}000$ vs $686$): more
   candidate subgroups survive the `n.min` filter, potentially
   increasing $L_{\text{eff}}$ and the procedure-level FPR.

2. **More confounders** ($12$ vs $7$): the combinatorial search
   space is larger, testing GRF and LASSO screening effectiveness.

3. **Different event rate**: ACTG175 has a lower event rate than
   GBSG, which reduces $d_{\text{eff}}$ per subgroup and may
   affect detection power.

4. **Different subgroup structure**: the age $\times$ CD8 subgroup
   is defined by continuous biomarkers discretised at clinically
   motivated cutpoints, rather than the GBSG's receptor-status
   $\times$ menopausal-status interaction.

5. **Days vs months**: the time scale is different, affecting the
   AFT model parametrisation but not the HR-scale operating
   characteristics.

## Limitations

- The ACTG175 subgroup definition (age $\leq `r actg_age_cut`$
  AND CD8 $> `r actg_cd80_cut`$) is exploratory, derived from a
  single GRF analysis. The true treatment effect heterogeneity
  pattern in ACTG175 is unknown.

- `generate_aft_dgm_flex()` requires more analyst judgment than
  `setup_gbsg_dgm()`. The variable discretisation, subgroup
  definition, and censoring calibration must all be specified
  manually.

- The simulation count (`r nsims_alt` sims) is for demonstration.
  Production studies should use $\geq 500$ simulations for stable
  operating characteristics (Monte Carlo SE for a 50% rate $\approx
  2.2\%$ at 500 sims).

- The adjusted Cox formula (`cox_formula_adj`) includes only 3 of
  12 confounders.  This is a deliberate choice for parsimony, but
  users should select confounders appropriate for their study
  design.

- `calibrate_k_inter()` does not yet support non-GBSG datasets;
  the manual grid search in @sec-calibrate is the current
  workaround.


# Session Info {#sec-session}

```{r session-info}
plan(sequential)
sessionInfo()
```


# References {#sec-references}

- Leon, L.F., Jemielita, T., Guo, Z., Marceau West, R.,
  Anderson, K.M. (2024). Exploratory subgroup identification
  in the heterogeneous Cox model. *Statistics in Medicine*,
  43, 4253--4275.

- Hammer, S.M., Katzenstein, D.A., Hughes, M.D., et al. (1996).
  A trial comparing nucleoside monotherapy with combination
  therapy in HIV-infected adults with CD4 cell counts from 200
  to 500 per cubic millimeter. *NEJM*, 335, 1081--1090.
