---
title: "ForestSearch Consistency: Split vs. Resample"
subtitle: "GBSG Cox analysis (effMinSG) — same engine, two consistency methods"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")
```

# Purpose

A minimal comparison of the two consistency-calculation methods now exposed by
`forestsearch()` via `consistency_method`, holding everything else fixed:

* **`"split"`** — the original repeated 50/50 split-and-refit consistency rate.
* **`"resample"`** — the score-residual multiplier approximation
  (`consistency_resample()`), which returns the rate from a single subgroup fit.

This is the GBSG Cox (survival) analysis trimmed to only the parts needed for
the comparison: data preparation, the `forestsearch()` call, and the
post-selection bootstrap bias correction. The two runs
share the same `seedit`, screening engines, and thresholds, so the candidate
set is identical — only the consistency step differs. Both use
`use_twostage = TRUE`; under `"resample"` the two-stage split screening is
bypassed (the rate is computed directly), so the flag is exercised in the
configuration where the question of "does resample fit within two-stage?"
actually arises.

```{r config}
#| code-fold: false
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(grf)
library(policytree)
library(future)
library(doFuture)

n_cores <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))

# Identification parameters (mirror the effMinSG analysis)
fs_sg_focus             <- "effMinSG"
fs_hr_threshold         <- 1.0
fs_hr_consistency       <- 0.90
fs_max_subgroups_search <- 30
```

# Data

```{r data-setup}
#| code-fold: false
df.analysis <- gbsg
df.analysis <- within(df.analysis, {
  id          <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3      <- ifelse(grade == "3", 1, 0)
  treat       <- hormon
})

confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name   <- "status"
id.name      <- "id"
treat.name   <- "hormon"

cat("N =", nrow(df.analysis),
    "| Events =", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
```

# Run forestsearch under each consistency method

The identical call is wrapped in a function parameterised only by
`consistency_method`. Display/verbosity arguments are quieted; everything that
affects candidate generation and selection is held fixed.

```{r run-both}
#| code-fold: false
run_fs <- function(method) {
  t0 <- proc.time()
  fs <- forestsearch(
    df.analysis,
    outcome.name              = outcome.name,
    event.name                = event.name,
    treat.name                = treat.name,
    id.name                   = id.name,
    confounders.name          = confounders.name,
    is.RCT                    = TRUE,
    seedit                    = 8316951,
    est.scale                 = "hr",
    # screening engines (shared -> identical candidate set across methods)
    use_lasso                 = TRUE,
    use_grf                   = TRUE,
    grf_depth                 = 2,
    dmin.grf                  = 4,
    frac.tau                  = 0.80,
    return_selected_cuts_only = TRUE,
    # selection engine
    subgroup_method           = "consistency",
    sg_focus                  = fs_sg_focus,
    selection_rule            = "neighborhood",
    effect_neighborhood       = 0.10,
    # subgroup constraints
    n.min                     = 60,
    d0.min                    = 10,
    d1.min                    = 10,
    maxk                      = 2,
    max_subgroups_search      = fs_max_subgroups_search,
    # thresholds
    hr.threshold              = fs_hr_threshold,
    hr.consistency            = fs_hr_consistency,
    pconsistency.threshold    = 0.90,
    m1.threshold              = Inf,
    minp                      = 0.025,
    # consistency search
    fs.splits                 = 1000,
    stop_threshold            = 0.95,
    use_twostage              = TRUE,
    # >>> the only thing that varies between the two runs <<<
    consistency_method        = method,
    # output (quieted for the comparison)
    show_candidate_summary    = FALSE,
    details                   = FALSE,
    quiet                     = TRUE,
    plot.sg                   = FALSE,
    plot.grf                  = FALSE,
    parallel_args = list(plan = "multisession", workers = n_cores,
                         show_message = FALSE)
  )
  plan("sequential")
  list(fs = fs, elapsed = unname((proc.time() - t0)["elapsed"]))
}

res_split    <- run_fs("split")
res_resample <- run_fs("resample")
```

# Comparison

```{r compare}
#| code-fold: false
summ <- function(tag, r) {
  fs <- r$fs
  sg   <- if (is.null(fs$sg.harm)) "(none)" else paste(fs$sg.harm, collapse = " & ")
  size <- if (!is.null(fs$df.est)) sum(fs$df.est$treat.recommend == 0) else NA_integer_
  data.frame(method = tag, subgroup = sg, size = size,
             elapsed_sec = round(r$elapsed, 2))
}

cmp <- rbind(summ("split", res_split), summ("resample", res_resample))
cmp |>
  gt() |>
  tab_header(title = "Selected subgroup: split vs. resample") |>
  tab_source_note("Same seedit, screening, and thresholds; only consistency_method differs.")

cat("\nIdentical selected subgroup? ",
    identical(res_split$fs$sg.harm, res_resample$fs$sg.harm), "\n")
```

## Consistency rates per candidate

The per-candidate consistency table for each method. The selected subgroup's
`Pcons` should agree to within the approximation error documented for the
resampling method (≈0.01 on GBSG).

```{r tables}
#| code-fold: false
cat("=== split ===\n")
sg_tables(res_split$fs,    ndecimals = 3, which_df = "est")$sg10_out

cat("=== resample ===\n")
sg_tables(res_resample$fs, ndecimals = 3, which_df = "est")$sg10_out
```

## Treatment-effect estimates

```{r estimates}
#| code-fold: false
cat("=== split ===\n")
sg_tables(res_split$fs,    ndecimals = 3, which_df = "est")$tab_estimates

cat("=== resample ===\n")
sg_tables(res_resample$fs, ndecimals = 3, which_df = "est")$tab_estimates
```

# Bootstrap post-selection (bias correction)

Cox estimates from a *selected* subgroup are upwardly biased: the subgroup was
chosen precisely because it showed an extreme effect. The bootstrap re-runs the
entire selection on each resample and applies bias correction with
infinitesimal-jackknife variance.

The key point for this comparison: `forestsearch_bootstrap_dofuture()` re-runs
`forestsearch()` from the fitted object's stored call (`fs$args_call_all`), and
`consistency_method` is part of that call. So it propagates into every
resample automatically — the split-bootstrap re-selects with literal splitting,
the resample-bootstrap with the approximation. This is where the per-subgroup
resample saving compounds: the consistency step runs `nb_boots` times, once
inside each bootstrap re-fit.

```{r bootstrap-run}
#| code-fold: false
NB <- 1000   # production: 500-2000

run_boot <- function(fs_obj) {
  t0 <- proc.time()
  bc <- forestsearch_bootstrap_dofuture(
    fs.est        = fs_obj,
    nb_boots      = NB,
    show_three    = FALSE,
    details       = FALSE,
    parallel_args = list(plan = "multisession", workers = n_cores,
                         show_message = FALSE)
  )
  plan("sequential")
  list(bc = bc, elapsed = unname((proc.time() - t0)["elapsed"]))
}

boot_split    <- run_boot(res_split$fs)
boot_resample <- run_boot(res_resample$fs)
```

## Bias-corrected estimates

The bias-corrected estimate table for the selected subgroup (Ĥ) and its
complement (Ĥᶜ) under each consistency method. Because both methods selected the
same subgroup, the corrected estimates should agree up to bootstrap Monte-Carlo
error (and the small consistency-rate approximation, which only affects which
subgroup each *resample* re-selects).

```{r bootstrap-summary}
#| code-fold: false
sum_split <- summarize_bootstrap_results(
  sgharm = res_split$fs$sg.harm, boot_results = boot_split$bc,
  create_plots = FALSE, est.scale = "hr")
sum_resample <- summarize_bootstrap_results(
  sgharm = res_resample$fs$sg.harm, boot_results = boot_resample$bc,
  create_plots = FALSE, est.scale = "hr")

cat("=== split: bias-corrected ===\n")
sum_split$table

cat("=== resample: bias-corrected ===\n")
sum_resample$table
```

## Bootstrap timing

```{r bootstrap-timing}
#| code-fold: false
data.frame(
  method        = c("split", "resample"),
  nb_boots      = NB,
  bootstrap_sec = round(c(boot_split$elapsed, boot_resample$elapsed), 1)
) |>
  gt() |>
  tab_header(title = "Bootstrap wall-clock") |>
  tab_source_note(paste(
    "Each resample re-runs the whole forestsearch() (candidate generation +",
    "consistency); only the consistency step differs between the two methods."))
```

# Notes

* **`use_twostage` interaction.** Both runs set `use_twostage = TRUE`. Under
  `"split"` the two-stage Stage-1 screen + Stage-2 Wilson-CI early stopping
  governs how many of the up-to-`fs.splits` splits are drawn per candidate.
  Under `"resample"` there are no splits to stop early, so the two-stage path
  is bypassed and the rate is returned from a single fit — `use_twostage` and
  `twostage_args` have no effect on that path.
* **Timing.** The elapsed times above are for the *whole* `forestsearch()`
  call, which is dominated by the shared candidate generation (GRF/LASSO) and,
  under split, already trimmed by two-stage early stopping. The consistency
  step itself is ~3 orders of magnitude cheaper under resample (measured
  separately on the Ĥ subgroup); that per-subgroup saving compounds inside the
  bootstrap and cross-validation layers rather than in this single call.
* **Scope.** `consistency_method = "resample"` currently applies to the Cox
  (survival) path; GLM/binary outcomes fall back to `"split"`.
* **Bootstrap propagation.** The post-selection bootstrap inherits
  `consistency_method` from `fs$args_call_all`, so no extra wiring is needed:
  passing `res_split$fs` vs `res_resample$fs` is the only difference between the
  two bootstrap runs. The consistency step executes `nb_boots` times, so this is
  the layer where the resample saving accumulates.
```
