---
title: "GBSG: splitting-consistency vs. resampling approximation"
subtitle: "Forest Search consistency step, original (direct split) and new (resampling)"
format:
  html:
    toc: true
    code-fold: false
    df-print: paged
execute:
  warning: false
  message: false
---

```{r setup}
# --- knobs (all in one place) ---------------------------------------------
R_true        <- 2000      # literal random splits for the "direct" rate
draws         <- 8000      # multiplier draws for the resampling approximation
hr.consistency<- 1.0       # consistency threshold c2 (HR scale)
hr.screen     <- 1.25      # screening threshold c1 (context only here)
seed          <- 7

library(survival)          # gbsg data + coxph
# forestsearch provides run_single_consistency_split(); consistency_resample()
# is the new file. Load whichever way you have them:
if (!exists("run_single_consistency_split")) suppressMessages(library(forestsearch))
if (!exists("consistency_resample")) source("consistency_resample.R")
library(gt)
```

## Data

`survival::gbsg` (German Breast Cancer Study Group). Outcome = recurrence-free
survival in months; treatment = hormonal therapy (`hormon`). Columns are mapped
to the `Y` / `Event` / `Treat` names the forestsearch consistency engine expects.

```{r data}
df <- within(gbsg, {
  Y     <- rfstime / 30.4375
  Event <- status
  Treat <- hormon
})
```

## The paper's Ĥ subgroup: Estrogen (`er`) ≤ 0

León et al. (2024) identify `Ĥ = {er <= 0}` with $\hat\theta(\hat H)=1.95$ and a
**reported consistency rate of 95.1%**. We recompute that rate two ways:

* **direct** — `run_single_consistency_split()` looped `R_true` times (the
  original 50/50-split-and-refit calculation);
* **resampling** — `consistency_resample()` (one Cox fit + multiplier
  representation of the split pair).

```{r hhat}
sub_H <- df[df$er <= 0, c("Y", "Event", "Treat")]

cmp_H <- consistency_resample_compare(
  sub_H, hr.consistency = hr.consistency,
  R_true = R_true, draws = draws, seed = seed
)
cmp_H
# Expected (base-R Breslow check): HR ~ 1.95, n = 82, d = 45,
#   rate_true ~ 0.95 (paper: 0.951),  rate_closed ~ 0.96,  rate_mc ~ 0.96.
# coxph (Efron ties) will land within rounding of these.
```

The direct rate reproduces the paper's 95.1%; the closed-form and multiplier
approximations agree to within ~1 percentage point.

## Candidate sweep

The selection step only cares whether a candidate clears the 90% consistency
bar, so what matters is that the approximation *ranks* candidates like the
direct rate. Below, a handful of single-factor GBSG subgroups (one harm-direction
candidate, the rest benefit-direction).

```{r sweep}
df$age_hi   <- as.integer(df$age   > median(df$age))
df$nodes_hi <- as.integer(df$nodes > median(df$nodes))
df$pgr_lo   <- as.integer(df$pgr  <= median(df$pgr))
df$size_hi  <- as.integer(df$size  > median(df$size))

cands <- list(
  "er<=0"     = df$er <= 0,
  "grade==3"  = df$grade == 3,
  "meno==1"   = df$meno == 1,
  "nodes>med" = df$nodes_hi == 1,
  "pgr<=med"  = df$pgr_lo == 1,
  "size>med"  = df$size_hi == 1,
  "age>med"   = df$age_hi == 1
)

sweep <- do.call(rbind, lapply(names(cands), function(nm) {
  s <- df[cands[[nm]], c("Y", "Event", "Treat")]
  out <- consistency_resample_compare(
    s, hr.consistency = hr.consistency,
    R_true = R_true, draws = draws, seed = seed
  )
  data.frame(subgroup = nm, out)
}))

sweep |>
  gt() |>
  fmt_number(c(beta_hat, sigma_D, rate_true, rate_closed, rate_mc,
               err_closed, err_mc), decimals = 3) |>
  tab_header(title = "GBSG consistency: direct split vs. resampling approximation")
```

Benefit-direction candidates (HR < 1) read ~0 under both methods; only the harm
subgroup `er<=0` is consistent. Selection ranking is preserved.

## A `method` flag (application level)

A single entry point that dispatches between the original direct splitting and
the new approximation. This is the pattern that would live inside the engine
(see note below); here it wraps the two paths for one subgroup.

```{r flag}
consistency_rate <- function(df.x, hr.consistency = 1.0,
                             method = c("split", "resample"),
                             R = 2000, draws = 8000, seed = 7,
                             adjust_covariates = NULL) {
  method <- match.arg(method)
  if (method == "resample") {
    consistency_resample(
      df.x, hr.consistency = hr.consistency, method = "closed",
      draws = draws, adjust_covariates = adjust_covariates, seed = seed
    )$rate_closed
  } else {
    set.seed(seed)
    res <- vapply(seq_len(R), function(r) run_single_consistency_split(
      df.x = df.x, N.x = nrow(df.x), hr.consistency = hr.consistency,
      adjust_covariates = adjust_covariates), numeric(1))
    mean(res, na.rm = TRUE)
  }
}

c(split    = consistency_rate(sub_H, method = "split",    R = R_true, seed = seed),
  resample = consistency_rate(sub_H, method = "resample", draws = draws, seed = seed))
```

## Timing

```{r timing}
t_split <- system.time(
  consistency_rate(sub_H, method = "split", R = R_true, seed = seed)
)["elapsed"]
t_resamp <- system.time(
  consistency_rate(sub_H, method = "resample", draws = draws, seed = seed)
)["elapsed"]
data.frame(method = c("split", "resample"),
           elapsed_sec = c(unname(t_split), unname(t_resamp)),
           speedup = c(1, round(t_split / t_resamp, 1)))
```

## Notes

* This doc isolates the **consistency step**; it is not the full FS candidate
  generation / GRF-lasso screening / selection pipeline.
* Adjusted version: pass `adjust_covariates = "age"` (etc.) to both
  `consistency_resample_compare()` and `consistency_rate()`.
* **Core-engine flag (not yet added).** The natural home is a
  `consistency_method = c("split","resample")` argument on
  `subgroup.consistency()` that flows into `evaluate_subgroup_consistency()`,
  swapping the per-candidate `run_single_consistency_split()` loop for a single
  `consistency_resample()` call. That is a change to the core consistency files
  and is intentionally left out of this prototype.
```
