---
title: "Smoke test: Count GLM end-to-end pipeline"
subtitle: "Poisson outcome with offset, IRR effect measure"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-tools: true
    code-fold: true
    theme: cosmo
    fontsize: 11pt
    number-sections: true
params:
  quick_mode: true
  force_rerun: false
---

# Purpose {#sec-purpose}

Smoke test exercising the **count outcome path** end-to-end:
primary `forestsearch()` call → CV via `forestsearch_tenfold()` →
bootstrap via `forestsearch_bootstrap_dofuture()` → summary tables.

If any stage fails to render or produces implausible numbers, the
count-GLM path has regressed.  The DGM uses a known harm signal
(IRR = 2.0 within `age <= 50 & biomarker_hi == 1`) with follow-up
time as an offset.

This document is a smoke test, not a unit-test suite.  Success
criteria are (1) clean render, (2) pipeline identifies the planted
harm signal on the primary fit, (3) outputs look reasonable.


# Setup {#sec-setup}

```{r load-packages}
#| message: false
#| warning: false
library(forestsearch)
library(data.table)
library(gt)
library(future)
library(foreach)
library(doFuture)
```

```{r config}
if (isTRUE(params$quick_mode)) {
  N         <- 800L
  Ksims_cv  <- 5L
  Kfolds_cv <- 10L
  nb_boots  <- 20L
  ckpt_tag  <- "quick"
} else {
  N         <- 1500L
  Ksims_cv  <- 20L
  Kfolds_cv <- 10L
  nb_boots  <- 100L
  ckpt_tag  <- "full"
}
cv_ckpt   <- sprintf("smoke_count_cv_%s.rds",   ckpt_tag)
boot_ckpt <- sprintf("smoke_count_boot_%s.rds", ckpt_tag)
n_workers <- min(4L, floor(0.95 * max(1L, parallel::detectCores() - 1L)))
cat(sprintf("N=%d, Ksims=%d, nb_boots=%d\n", N, Ksims_cv, nb_boots))
```


# Synthetic count DGM {#sec-dgm}

```{r dgm}
set.seed(20260421)
IRR_harm <- 2.0
df_syn <- data.frame(
  id           = seq_len(N),
  treat        = rbinom(N, 1L, 0.5),
  age          = round(rnorm(N, 55, 12)),
  biomarker    = rnorm(N),
  wtkg         = round(rnorm(N, 75, 12), 1),
  biomarker_hi = factor(rbinom(N, 1L, 0.5), levels = c(0L, 1L)),
  sex          = factor(rbinom(N, 1L, 0.5), levels = c(0L, 1L)),
  ftime        = round(runif(N, 1, 5), 2)
)
in_harm  <- df_syn$age <= 50 & df_syn$biomarker_hi == "1"
base_rate <- 0.5
rate     <- base_rate * exp(df_syn$treat * ifelse(in_harm, log(IRR_harm), 0))
df_syn$y <- rpois(N, rate * df_syn$ftime)

cat(sprintf("N = %d\n", nrow(df_syn)))
cat(sprintf("Overall event count: mean=%.2f, median=%d, max=%d\n",
            mean(df_syn$y), median(df_syn$y), max(df_syn$y)))
cat(sprintf("Harm subgroup prevalence: %.2f\n", mean(in_harm)))
cat(sprintf("Mean rate | treat=0, in harm: %.3f\n",
            mean(df_syn$y[df_syn$treat == 0L & in_harm] /
                 df_syn$ftime[df_syn$treat == 0L & in_harm])))
cat(sprintf("Mean rate | treat=1, in harm: %.3f\n",
            mean(df_syn$y[df_syn$treat == 1L & in_harm] /
                 df_syn$ftime[df_syn$treat == 1L & in_harm])))
```


# Primary ForestSearch call {#sec-fs}

```{r fs-run}
#| message: false
fs_args <- list(
  confounders.name = c("age", "biomarker", "wtkg", "biomarker_hi", "sex"),
  outcome.name     = "y",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "count",
  effect_measure   = "IRR",
  offset.name      = "ftime",
  hr.threshold           = 1.25,
  hr.consistency         = 1.00,
  pconsistency.threshold = 0.80,
  sg_focus               = "maxSG",
  max_subgroups_search   = 10L,
  use_grf                = TRUE,
  use_lasso              = FALSE,
  is.RCT                 = TRUE,
  vi.grf.min             = -0.2,
  cont.cutoff            = 4L,
  cut_type               = "default",
  seedit                 = 8316951L,
  maxk                   = 2L,
  n.min                  = 60L,
  d0.min                 = 10L,
  d1.min                 = 10L,
  fs.splits              = 300L,
  details                = FALSE,
  quiet                  = TRUE,
  plot.sg                = FALSE,
  plot.grf               = FALSE,
  parallel_args          = list(plan = "multisession",
                                workers = n_workers,
                                show_message = FALSE)
)

t0 <- proc.time()
fs_full <- do.call(forestsearch,
                   c(list(df.analysis = df_syn), fs_args))
plan("sequential")
t_fs <- (proc.time() - t0)["elapsed"]

stopifnot(!is.null(fs_full$args_call_all))

cat(sprintf("forestsearch() completed in %.1fs\n", t_fs))
cat(sprintf("Identified subgroup: %s\n",
            if (is.null(fs_full$sg.harm)) "(none)"
            else paste(fs_full$sg.harm, collapse = " & ")))
cat(sprintf("GRF cuts:          %s\n",
            paste(fs_full$grf_cuts, collapse = " | ")))
cat(sprintf("outcome_type:      %s\n", fs_full$outcome_type))
cat(sprintf("effect_measure:    %s\n", fs_full$effect_measure))
```


# Cross-validation {#sec-cv}

```{r cv-run}
#| message: false
.run_cv <- function() {
  forestsearch_tenfold(
    fs.est        = fs_full,
    sims          = Ksims_cv,
    Kfolds        = Kfolds_cv,
    details       = FALSE,
    parallel_args = list(plan    = "multisession",
                         workers = n_workers,
                         show_message = FALSE)
  )
}

if (file.exists(cv_ckpt) && !isTRUE(params$force_rerun)) {
  message(sprintf("Loading CV from %s", cv_ckpt))
  fs_cv <- readRDS(cv_ckpt)
} else {
  message(sprintf("Running CV (sims=%d, Kfolds=%d)...",
                  Ksims_cv, Kfolds_cv))
  fs_cv <- .run_cv()
  saveRDS(fs_cv, cv_ckpt)
}
plan("sequential")

cat(sprintf("CV: %d sims x %d folds = %d rows\n",
            fs_cv$sims, fs_cv$Kfolds, nrow(fs_cv$fold_summary)))
cat(sprintf("Folds identifying a subgroup: %d / %d\n",
            sum(fs_cv$fold_summary$any_found == 1L),
            nrow(fs_cv$fold_summary)))
```

## Fold-level summary

```{r cv-summary}
cv_diag <- summarize_cv_results(
  cv_output         = fs_cv,
  original_sg       = fs_full$sg.harm,
  original_grf_cuts = fs_full$grf_cuts,
  top_n             = 10L
)

cv_diag$identification_summary
```

## Fold numeric diagnostics

Key check for count outcome: the "Training-fold subgroup effect"
median should be > 1 on the **natural IRR scale** (not log-IRR).

```{r cv-numeric}
cv_diag$fold_numeric_summary
```

## GRF cut stability across folds

```{r cv-grf-cuts}
cv_diag$grf_cut_summary
```


# Bootstrap {#sec-bootstrap}

```{r boot-run}
#| message: false
.run_boot <- function() {
  forestsearch_bootstrap_dofuture(
    fs.est        = fs_full,
    nb_boots      = nb_boots,
    parallel_args = list(plan    = "multisession",
                         workers = n_workers,
                         show_message = FALSE)
  )
}

if (file.exists(boot_ckpt) && !isTRUE(params$force_rerun)) {
  message(sprintf("Loading bootstrap from %s", boot_ckpt))
  fs_bc <- readRDS(boot_ckpt)
} else {
  message(sprintf("Running bootstrap (nb_boots=%d)...", nb_boots))
  fs_bc <- .run_boot()
  saveRDS(fs_bc, boot_ckpt)
}
plan("sequential")

cat(sprintf("Bootstrap: %d iterations\n", nrow(fs_bc$results)))
cat(sprintf("Iterations identifying a subgroup: %d / %d\n",
            sum(!is.na(fs_bc$results$sg1_b)),
            nrow(fs_bc$results)))
```

## GRF cut frequency across bootstraps

```{r boot-grf}
subg_summary <- summarize_bootstrap_subgroups(
  results     = fs_bc$results,
  nb_boots    = nb_boots,
  original_sg = fs_full$sg.harm,
  maxk        = 2L
)

if (!is.null(subg_summary$grf_cut_freq)) {
  subg_summary$grf_cut_freq |>
    utils::head(10) |>
    gt::gt() |>
    gt::tab_header(title = gt::md("**Top GRF cuts across bootstraps**")) |>
    gt::fmt_number(columns = "Percent", decimals = 2) |>
    gt::tab_options(table.font.size = gt::px(13))
}
```


# Session info {#sec-session}

```{r session}
sessionInfo()
```
