---
title: "Living test: Phase A/B/C GLM compatibility"
subtitle: "Synthetic binary outcome + OR effect measure"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-tools: true
    code-fold: true
    theme: cosmo
    fontsize: 11pt
    number-sections: true
params:
  quick_mode: true       # TRUE: smaller grid (fast). FALSE: larger.
  force_rerun: false
---

# Purpose {#sec-purpose}

Validates that the Phase A/B/C diagnostic features
(`fold_summary$grf_cuts`, `fold_summary$pconsistency`,
`fold_summary$training_fs_hr`, `fold_summary$n_candidates_evaluated`,
`summarize_cv_results()` slots, and the bootstrap `grf_cut_freq`
slot) work correctly when the outcome is **binary** rather than
survival.

The key GLM-specific concern is `training_fs_hr` on log-scale effect
measures (OR, IRR, RR): `forestsearch()`'s internal result stores
these values on log scale, but Phase B's capture applies `exp()` so
the column is always on natural scale regardless of outcome type.
This document exercises the full pipeline on synthetic binary data
with `effect_measure = "OR"` to confirm that contract.

The synthetic DGM mirrors the canonical ACTG175 binary-benefit
configuration: a mix of **continuous** covariates (where GRF
discovers cuts) and **binary factor** covariates (passed through
without cut discovery), with `sg_focus = "maxSG"` to recover the
largest qualifying subgroup.

Coverage:

1. **Synthetic data generation** — harm subgroup driven by
   `age <= 50 & biomarker_hi == 1`.  Ground truth OR ≈ 2.0 within
   subgroup, ≈ 1.0 outside.  Mix of continuous and binary factor
   covariates exercises both cut-discovery paths.
2. **Primary ForestSearch call** — confirms the GLM routing
   succeeds, returns a populated object with `df.est` and
   `args_call_all`, and surfaces `outcome_type` / `effect_measure`
   metadata correctly.
3. **CV pipeline (Phase A/B)** — runs `forestsearch_tenfold()` at
   small sims, confirms every new `fold_summary` column populates
   correctly on GLM data, confirms `summarize_cv_results()` produces
   each diagnostic slot, and specifically confirms `training_fs_hr`
   values are on OR scale (i.e., > 1 for harm, not log-OR values
   centered near 0.7).
4. **Bootstrap pipeline (Phase C)** — runs
   `forestsearch_bootstrap_dofuture()` at small `nb_boots`, confirms
   `results$grf_cuts_b` populates and `summarize_bootstrap_subgroups()`
   surfaces `grf_cut_freq`.
5. **Cross-check** — captured `training_fs_hr` on an identifying
   fold agrees within order-of-magnitude with an independent `glm()`
   refit on the same subgroup.

A cumulative pass/fail table renders in @sec-summary.


# Setup {#sec-setup}

```{r load-packages}
#| message: false
#| warning: false
library(forestsearch)
library(data.table)
library(gt)
library(future)
library(foreach)
library(doFuture)
```

```{r config}
if (isTRUE(params$quick_mode)) {
  N         <- 800L
  Ksims_cv  <- 5L
  Kfolds_cv <- 10L
  nb_boots  <- 20L
  ckpt_tag  <- "quick"
} else {
  N         <- 1500L
  Ksims_cv  <- 20L
  Kfolds_cv <- 10L
  nb_boots  <- 100L
  ckpt_tag  <- "full"
}
ckpt_path      <- sprintf("test_glm_compatibility_checkpoint_%s.rds", ckpt_tag)
boot_ckpt_path <- sprintf("test_glm_bootstrap_checkpoint_%s.rds",    ckpt_tag)
cat(sprintf("Mode: %s (N=%d, Ksims=%d, nb_boots=%d)\n",
            ckpt_tag, N, Ksims_cv, nb_boots))

n_workers <- min(4L, floor(0.95 * max(1L, parallel::detectCores() - 1L)))
```

```{r test-harness}
test_results <- data.frame(
  Section = character(0),
  Test    = character(0),
  Status  = character(0),
  Detail  = character(0),
  stringsAsFactors = FALSE
)

.tc <- function(section, test, pass, detail = "") {
  status <- if (isTRUE(pass)) "PASS" else "FAIL"
  test_results <<- rbind(test_results, data.frame(
    Section = section, Test = test, Status = status,
    Detail  = substr(detail, 1L, 140L),
    stringsAsFactors = FALSE
  ))
  invisible(pass)
}
```


# Synthetic binary-outcome DGM {#sec-dgm}

Mirrors the ACTG175 binary-benefit configuration: a mix of
continuous and binary factor covariates, with a harm signal defined
jointly by one continuous cut (`age <= 50`) and one binary factor
(`biomarker_hi == 1`).  Outcome is Bernoulli; natural effect measure
is odds ratio.

```{r generate-data}
set.seed(20260421)

OR_harm    <- 2.0     # true OR in harm subgroup (treatment worsens outcome)
OR_benefit <- 1.0     # null treatment effect outside harm subgroup

df_syn <- data.frame(
  id           = seq_len(N),
  treat        = rbinom(N, 1L, 0.5),
  age          = round(rnorm(N, mean = 55, sd = 12)),
  wtkg         = round(rnorm(N, mean = 75, sd = 12), 1),
  biomarker    = rnorm(N),
  karnof       = sample(70:100, N, replace = TRUE),
  biomarker_hi = factor(rbinom(N, 1L, 0.5), levels = c(0L, 1L)),
  sex          = factor(rbinom(N, 1L, 0.5), levels = c(0L, 1L)),
  site         = factor(sample(0:1, N, replace = TRUE),
                        levels = c(0L, 1L))
)

# True harm subgroup: age <= 50 AND biomarker_hi == 1
true_harm    <- df_syn$age <= 50 & df_syn$biomarker_hi == "1"
treat_effect <- ifelse(true_harm, log(OR_harm), log(OR_benefit))

linpred <- 0 + df_syn$treat * treat_effect
prob_y  <- plogis(linpred)
df_syn$y <- rbinom(N, 1L, prob_y)

cat(sprintf("N = %d, overall event rate = %.3f\n",
            nrow(df_syn), mean(df_syn$y)))
cat(sprintf("True harm prevalence: %.2f\n", mean(true_harm)))
cat(sprintf("Event rate | treat = 0: %.3f\n",
            mean(df_syn$y[df_syn$treat == 0])))
cat(sprintf("Event rate | treat = 1: %.3f\n",
            mean(df_syn$y[df_syn$treat == 1])))

overall_or <- (mean(df_syn$y[df_syn$treat == 1]) /
                (1 - mean(df_syn$y[df_syn$treat == 1]))) /
              (mean(df_syn$y[df_syn$treat == 0]) /
                (1 - mean(df_syn$y[df_syn$treat == 0])))
cat(sprintf("Overall empirical OR: %.2f\n", overall_or))
```

Empirical check that the harm subgroup is recoverable:

```{r dgm-sanity}
har <- df_syn[true_harm, ]
or_harm_emp <- (mean(har$y[har$treat == 1]) /
                  (1 - mean(har$y[har$treat == 1]))) /
               (mean(har$y[har$treat == 0]) /
                  (1 - mean(har$y[har$treat == 0])))
cat(sprintf("Empirical OR in true harm subgroup: %.2f (DGM = %.1f)\n",
            or_harm_emp, OR_harm))

.tc("DGM", "empirical harm OR > 1.3",
    or_harm_emp > 1.3,
    detail = sprintf("empirical OR = %.2f", or_harm_emp))

.tc("DGM", "harm subgroup prevalence between 15% and 40%",
    mean(true_harm) > 0.15 && mean(true_harm) < 0.40,
    detail = sprintf("prev = %.2f", mean(true_harm)))
```


# Primary ForestSearch call {#sec-fs-full}

Configuration mirrors the ACTG175 pattern:

- `outcome_type = "binary"`, `effect_measure = "OR"`
- `sg_focus = "maxSG"` (largest qualifying subgroup)
- `is.RCT = TRUE`, `vi.grf.min = -0.2`, `cont.cutoff = 4L`
- Mixed covariates: continuous (cuts discovered by GRF) + factor
  (passed through without cut attempt)
- `use_grf = TRUE`, `use_lasso = FALSE`

```{r fs-full}
#| message: false
fs_args <- list(
  confounders.name = c("age", "wtkg", "biomarker", "karnof",
                       "biomarker_hi", "sex", "site"),
  outcome.name     = "y",
  event.name       = "y",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "OR",
  hr.threshold             = 1.25,
  hr.consistency           = 1.00,
  pconsistency.threshold   = 0.80,
  sg_focus                 = "maxSG",
  max_subgroups_search     = 10L,
  use_grf                  = TRUE,
  use_lasso                = FALSE,
  is.RCT                   = TRUE,
  vi.grf.min               = -0.2,
  cont.cutoff              = 4L,
  cut_type                 = "default",
  seedit                   = 8316951L,
  maxk                     = 2L,
  n.min                    = 60L,
  d0.min                   = 10L,
  d1.min                   = 10L,
  fs.splits                = 300L,
  details                  = FALSE,
  quiet                    = TRUE,
  plot.sg                  = FALSE,
  plot.grf                 = FALSE,
  parallel_args            = list(plan         = "multisession",
                                  workers      = n_workers,
                                  show_message = FALSE)
)

fs_full <- do.call(forestsearch,
                   c(list(df.analysis = df_syn), fs_args))
plan("sequential")

# Hard check: if the primary call collapsed to a minimal
# early-return, fail LOUDLY here rather than letting downstream
# chunks error with cryptic messages.
if (is.null(fs_full$args_call_all) || is.null(fs_full$df.est)) {
  stop(sprintf(paste0(
    "Primary ForestSearch call collapsed to an early-return path ",
    "(args_call_all=%s, df.est=%s, sg.harm=%s).  Check for upstream ",
    "FSdata / find.grps / consistency failures."),
    !is.null(fs_full$args_call_all),
    !is.null(fs_full$df.est),
    !is.null(fs_full$sg.harm)))
}

cat(sprintf("Identified subgroup: %s\n",
            paste(fs_full$sg.harm, collapse = " & ")))
cat(sprintf("outcome_type on return: %s\n", fs_full$outcome_type))
cat(sprintf("effect_measure on return: %s\n", fs_full$effect_measure))
cat(sprintf("GRF cuts: %s\n",
            paste(fs_full$grf_cuts, collapse = " | ")))

.tc("Primary FS", "args_call_all present on return",
    !is.null(fs_full$args_call_all))
.tc("Primary FS", "df.est present on return",
    !is.null(fs_full$df.est))
.tc("Primary FS", "outcome_type == 'binary' preserved on return",
    identical(fs_full$outcome_type, "binary"))
.tc("Primary FS", "effect_measure == 'OR' preserved on return",
    identical(fs_full$effect_measure, "OR"))
.tc("Primary FS", "a subgroup was identified",
    !is.null(fs_full$sg.harm))

if (!is.null(fs_full$sg.harm)) {
  sg_str <- paste(fs_full$sg.harm, collapse = " & ")
  # Identified subgroup should involve at least one of the true
  # signal variables (age or biomarker_hi)
  .tc("Primary FS",
      "identified subgroup involves true-signal variable",
      grepl("age|biomarker_hi", sg_str),
      detail = sprintf("sg = '%s'", sg_str))
}

# The top-ranked result row's 'hr' column holds log-OR for binary+OR.
# Exponentiating should give a value roughly matching OR_harm (2.0).
if (!is.null(fs_full$grp.consistency$out_sg$result) &&
    nrow(fs_full$grp.consistency$out_sg$result) > 0L) {
  stored_hr <- as.numeric(
    fs_full$grp.consistency$out_sg$result[1L, "hr"]
  )
  or_nat <- exp(stored_hr)
  .tc("Primary FS",
      "stored 'hr' is on log scale (exp() yields plausible OR)",
      or_nat > 1.1 && or_nat < 8,
      detail = sprintf("stored='hr'=%.3f, exp()=%.2f (DGM OR = %.1f)",
                       stored_hr, or_nat, OR_harm))
}
```


# CV pipeline on GLM data {#sec-cv}

## Run CV

```{r cv-run}
#| message: false
.run_cv <- function() {
  forestsearch_tenfold(
    fs.est        = fs_full,
    sims          = Ksims_cv,
    Kfolds        = Kfolds_cv,
    details       = FALSE,
    parallel_args = list(plan    = "multisession",
                         workers = n_workers,
                         show_message = FALSE)
  )
}

if (file.exists(ckpt_path) && !isTRUE(params$force_rerun)) {
  message(sprintf("Loading CV from %s", ckpt_path))
  fs_cv <- readRDS(ckpt_path)

  required <- c("pconsistency", "training_fs_hr",
                "n_candidates_evaluated", "grf_cuts")
  if (!all(required %in% names(fs_cv$fold_summary))) {
    message("Stale checkpoint; rerunning CV...")
    fs_cv <- .run_cv()
    saveRDS(fs_cv, ckpt_path)
  }
} else {
  message(sprintf("Running CV (sims=%d, Kfolds=%d) ...",
                  Ksims_cv, Kfolds_cv))
  fs_cv <- .run_cv()
  saveRDS(fs_cv, ckpt_path)
}
plan("sequential")

cat(sprintf("CV complete: %d sims x %d folds = %d rows\n",
            fs_cv$sims, fs_cv$Kfolds, nrow(fs_cv$fold_summary)))
cat(sprintf("fold_summary columns: %s\n",
            paste(names(fs_cv$fold_summary), collapse = ", ")))
```

## Column-level checks on GLM data

```{r cv-columns}
fs <- fs_cv$fold_summary

.tc("CV columns", "fold_summary$grf_cuts present",
    "grf_cuts" %in% names(fs))
.tc("CV columns", "fold_summary$pconsistency present",
    "pconsistency" %in% names(fs))
.tc("CV columns", "fold_summary$training_fs_hr present",
    "training_fs_hr" %in% names(fs))
.tc("CV columns", "fold_summary$n_candidates_evaluated present",
    "n_candidates_evaluated" %in% names(fs))
```

## KEY GLM check: `training_fs_hr` is on natural scale

```{r cv-glm-scale}
# Under our DGM (OR_harm = 2.0), identifying folds should yield
# training_fs_hr values > 1 (typically 1.3 to 5 range on the OR
# scale).  If Phase B's exp() transform is missing, the values would
# cluster near log(2.0) ~ 0.69 instead.
hr_vals  <- fs$training_fs_hr
hr_nonna <- hr_vals[!is.na(hr_vals)]

if (length(hr_nonna) > 0L) {
  med_hr <- median(hr_nonna)
  min_hr <- min(hr_nonna)
  max_hr <- max(hr_nonna)

  cat(sprintf("training_fs_hr across %d identifying folds:\n",
              length(hr_nonna)))
  cat(sprintf("  median=%.3f, min=%.3f, max=%.3f\n",
              med_hr, min_hr, max_hr))
  cat(sprintf("  (on log scale, median would be %.3f)\n", log(med_hr)))

  # Natural-scale values should be > 1 for a harm search.  Log-scale
  # values would be near log(2) = 0.69 (below hr.threshold = 1.25).
  .tc("CV GLM scale",
      "training_fs_hr values > 1 (natural scale, not log)",
      med_hr > 1,
      detail = sprintf("median = %.3f (would be %.3f if log-scale)",
                       med_hr, log(med_hr)))

  # Given hr.threshold = 1.25 was enforced during search, identifying
  # folds must have training_fs_hr >= 1.25 on natural scale.
  .tc("CV GLM scale",
      sprintf("training_fs_hr >= hr.threshold (%.2f) for all id-ing folds",
              fs_args$hr.threshold),
      all(hr_nonna >= fs_args$hr.threshold),
      detail = sprintf("min = %.3f; threshold = %.2f",
                       min_hr, fs_args$hr.threshold))

  .tc("CV GLM scale", "training_fs_hr values finite and plausible",
      all(is.finite(hr_nonna)) && max_hr < 100,
      detail = sprintf("max = %.3f", max_hr))
}
```

## Pcons and candidate-count column checks

```{r cv-pconsistency}
pv          <- fs$pconsistency
identifying <- !is.na(fs$sg1)
.tc("CV Pcons", "pconsistency NA iff any_found == 0",
    all(is.na(pv) == (fs$any_found == 0L)))
if (sum(identifying) > 0L) {
  .tc("CV Pcons", "non-NA pconsistency in [0, 1]",
      all(pv[identifying] >= 0 & pv[identifying] <= 1))
  .tc("CV Pcons",
      sprintf("pconsistency >= threshold (%.2f) for identifying folds",
              fs_args$pconsistency.threshold),
      all(pv[identifying] >= fs_args$pconsistency.threshold))
}
```

```{r cv-candidates}
nc <- fs$n_candidates_evaluated
.tc("CV candidates", "n_candidates_evaluated is integer",
    is.integer(nc))
.tc("CV candidates", "n_candidates_evaluated non-negative",
    all(is.na(nc) | nc >= 0L))
id_idx <- which(fs$any_found == 1L)
if (length(id_idx) > 0L) {
  .tc("CV candidates",
      "identifying folds evaluated >= 1 candidate",
      all(!is.na(nc[id_idx]) & nc[id_idx] >= 1L))
}
```

## `summarize_cv_results()` on GLM data

```{r cv-summary}
cv_diag <- summarize_cv_results(
  cv_output         = fs_cv,
  original_sg       = fs_full$sg.harm,
  original_grf_cuts = fs_full$grf_cuts,
  top_n             = 10L
)

.tc("CV summary", "identification_summary slot populated",
    !is.null(cv_diag$identification_summary))
.tc("CV summary", "grf_cut_summary slot populated",
    !is.null(cv_diag$grf_cut_summary))
.tc("CV summary", "no_subgroup_decomposition slot populated",
    !is.null(cv_diag$no_subgroup_decomposition))
.tc("CV summary", "pconsistency_distribution slot populated",
    !is.null(cv_diag$pconsistency_distribution))
.tc("CV summary", "fold_numeric_summary slot populated",
    !is.null(cv_diag$fold_numeric_summary))
.tc("CV summary", "original_agreement slot populated",
    !is.null(cv_diag$original_agreement))
.tc("CV summary", "has_pconsistency metadata TRUE",
    isTRUE(cv_diag$has_pconsistency))
```

### Rendered `identification_summary`

```{r cv-ident-display}
cv_diag$identification_summary
```

### Rendered `grf_cut_summary`

```{r cv-grf-display}
cv_diag$grf_cut_summary
```

### Rendered `fold_numeric_summary`

Critical GLM check: the "Training-fold subgroup effect" row's median
should be on natural OR scale (value > 1), not log-OR scale (would be
near 0.5–1.0).

```{r cv-fnum-display}
cv_diag$fold_numeric_summary
```

```{r cv-fnum-values}
fns <- cv_diag$data$fold_numeric_summary
eff_row <- fns[
  fns$Metric == "Training-fold subgroup effect (identifying folds)",
  , drop = FALSE
]
if (nrow(eff_row) == 1L) {
  .tc("CV summary",
      "fold_numeric_summary effect-row label is 'subgroup effect' (outcome-agnostic)",
      grepl("subgroup effect", eff_row$Metric, fixed = TRUE))
  .tc("CV summary",
      "fold_numeric_summary effect-row context mentions 'natural scale'",
      grepl("natural scale", eff_row$Context, ignore.case = TRUE))
  .tc("CV summary",
      "fold_numeric_summary effect-row median > 1 (natural scale verified)",
      !is.na(eff_row$median) && eff_row$median > 1,
      detail = sprintf("median = %.3f", eff_row$median))
}
```


# Bootstrap pipeline on GLM data {#sec-boot}

## Run bootstrap

```{r boot-run}
#| message: false
.run_bootstrap <- function() {
  forestsearch_bootstrap_dofuture(
    fs.est        = fs_full,
    nb_boots      = nb_boots,
    parallel_args = list(plan    = "multisession",
                         workers = n_workers,
                         show_message = FALSE)
  )
}

if (file.exists(boot_ckpt_path) && !isTRUE(params$force_rerun)) {
  message(sprintf("Loading bootstrap from %s", boot_ckpt_path))
  fs_bc <- readRDS(boot_ckpt_path)
  if (!"grf_cuts_b" %in% names(fs_bc$results)) {
    message("Stale bootstrap checkpoint; rerunning...")
    fs_bc <- .run_bootstrap()
    saveRDS(fs_bc, boot_ckpt_path)
  }
} else {
  message(sprintf("Running bootstrap (nb_boots=%d) ...", nb_boots))
  fs_bc <- .run_bootstrap()
  saveRDS(fs_bc, boot_ckpt_path)
}
plan("sequential")

cat(sprintf("Bootstrap complete: %d iterations, columns:\n",
            nrow(fs_bc$results)))
cat(paste(names(fs_bc$results), collapse = ", "), "\n")
```

## Column and slot checks

```{r boot-column}
.tc("Bootstrap column", "results$grf_cuts_b present",
    "grf_cuts_b" %in% names(fs_bc$results))
.tc("Bootstrap column", "grf_cuts_b is character",
    is.character(fs_bc$results$grf_cuts_b))
.tc("Bootstrap column", "nrow(results) == nb_boots",
    nrow(fs_bc$results) == nb_boots)
```

```{r boot-slot}
subg_summary <- summarize_bootstrap_subgroups(
  results     = fs_bc$results,
  nb_boots    = nb_boots,
  original_sg = fs_full$sg.harm,
  maxk        = 2L
)

.tc("Bootstrap slot", "grf_cut_freq slot populated",
    !is.null(subg_summary$grf_cut_freq))

gcf <- subg_summary$grf_cut_freq
if (!is.null(gcf)) {
  .tc("Bootstrap slot", "grf_cut_freq N sums to nb_boots",
      sum(gcf$N) == nb_boots,
      detail = sprintf("sum=%d exp=%d", sum(gcf$N), nb_boots))
  .tc("Bootstrap slot", "grf_cut_freq sorted by descending N",
      all(diff(gcf$N) <= 0))
}
```

### Rendered GRF-cut frequency

```{r boot-display}
if (!is.null(subg_summary$grf_cut_freq)) {
  subg_summary$grf_cut_freq |>
    utils::head(10) |>
    gt::gt() |>
    gt::tab_header(
      title = gt::md("**GRF cuts across bootstrap iterations (GLM binary outcome)**"),
      subtitle = sprintf("Top 10 of %d unique cuts across %d iterations",
                         nrow(subg_summary$grf_cut_freq), nb_boots)
    ) |>
    gt::cols_label(
      Rank = "Rank", grf_cut = "GRF cuts",
      N = "N",       Percent = "% of bootstraps"
    ) |>
    gt::fmt_number(columns = "Percent", decimals = 2) |>
    gt::tab_options(table.font.size = gt::px(13))
}
```


# Cross-check: `training_fs_hr` vs. independent GLM refit {#sec-crosscheck}

Confirm numerical correctness of Phase B's `exp()` transform by
re-fitting the GLM directly on an identifying fold's subgroup (using
full data as a plausibility proxy, since exact per-fold splits
aren't preserved).

```{r crosscheck}
id_folds <- fs_cv$fold_summary[
  !is.na(fs_cv$fold_summary$training_fs_hr), , drop = FALSE
]

if (nrow(id_folds) > 0L) {
  # Pick the median fold by training_fs_hr
  pick <- id_folds[order(abs(id_folds$training_fs_hr -
                              median(id_folds$training_fs_hr)))[1], ]
  captured_hr <- pick$training_fs_hr

  cat(sprintf("Sampled identifying fold: sim=%d, fold=%d\n",
              pick$sim, pick$fold))
  cat(sprintf("Subgroup definition: %s%s\n",
              pick$sg1,
              if (!is.na(pick$sg2)) paste0(" & ", pick$sg2) else ""))
  cat(sprintf("Captured training_fs_hr: %.3f\n", captured_hr))

  # Evaluate the subgroup expression(s) to get a membership flag.
  # forestsearch() returns R-evaluable expressions like "age <= 50"
  # or "biomarker_hi == 1".
  in_sg <- tryCatch({
    sg1_flag <- eval(parse(text = pick$sg1), envir = df_syn)
    if (!is.na(pick$sg2)) {
      sg2_flag <- eval(parse(text = pick$sg2), envir = df_syn)
      sg1_flag & sg2_flag
    } else {
      sg1_flag
    }
  }, error = function(e) NULL)

  if (!is.null(in_sg) && sum(in_sg) > 10L) {
    df_sg <- df_syn[in_sg, ]
    m <- tryCatch(
      glm(y ~ treat, family = binomial, data = df_sg),
      error = function(e) NULL
    )
    if (!is.null(m)) {
      or_refit <- exp(coef(m)["treat"])
      cat(sprintf("Re-fit OR on full-data subgroup: %.3f\n", or_refit))

      .tc("Cross-check",
          "captured training_fs_hr and re-fit OR both > 1",
          captured_hr > 1 && or_refit > 1,
          detail = sprintf("captured = %.2f, refit = %.2f",
                           captured_hr, or_refit))

      .tc("Cross-check",
          "captured and refit agree within 3x",
          (captured_hr / or_refit > 1/3) &&
          (captured_hr / or_refit < 3),
          detail = sprintf("ratio = %.2f",
                           captured_hr / or_refit))
    }
  }
}
```


# Test summary {#sec-summary}

```{r summary-table}
n_total <- nrow(test_results)
n_pass  <- sum(test_results$Status == "PASS")
n_fail  <- n_total - n_pass

gt::gt(test_results,
       groupname_col = "Section",
       row_group_as_column = TRUE) |>
  gt::tab_header(
    title    = gt::md(sprintf(
      "**GLM compatibility test results: %d / %d passing**",
      n_pass, n_total)),
    subtitle = sprintf(
      "Binary outcome (OR), N=%d, CV sims=%d, nb_boots=%d",
      N, Ksims_cv, nb_boots)
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#d4edda"),
    locations = gt::cells_body(
      columns = "Status", rows = Status == "PASS"
    )
  ) |>
  gt::tab_style(
    style     = list(
      gt::cell_fill(color = "#f8d7da"),
      gt::cell_text(weight = "bold")
    ),
    locations = gt::cells_body(
      columns = "Status", rows = Status == "FAIL"
    )
  ) |>
  gt::cols_label(
    Test   = "Test case",
    Status = "Status",
    Detail = "Detail"
  ) |>
  gt::tab_options(table.font.size = gt::px(12))

if (n_fail == 0L) {
  message(sprintf("ALL TESTS PASSED: %d / %d", n_pass, n_total))
} else {
  message(sprintf("TESTS FAILED: %d / %d failing", n_fail, n_total))
  print(test_results[test_results$Status == "FAIL",
                     c("Section", "Test", "Detail")])
}
```


# Session info {#sec-session}

```{r session}
sessionInfo()
```
