---
title: "Living test: `summarize_cv_results()`"
subtitle: "Validates Phase A CV diagnostic expansion using GBSG"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-tools: true
    code-fold: true
    theme: cosmo
    fontsize: 11pt
    number-sections: true
params:
  quick_mode: true       # TRUE: sims=10 (~1-2 min). FALSE: sims=100 (~15-20 min).
---

# Purpose {#sec-purpose}

This document exercises every branch of
`summarize_cv_results()` against a `forestsearch_tenfold()` run
on the GBSG dataset.  It is a "living" test: re-running it with
an updated function should produce identical pass/fail results.
A cumulative pass/fail table is rendered in
@sec-summary.

Coverage:

1. **Minimal invocation** — `summarize_cv_results(tf)` with no
   optional arguments.  Inspects structure of the returned
   object and confirms every component slot is populated.
2. **Full invocation** — with `original_sg`,
   `original_grf_cuts`, and `create_plots = TRUE`.  Validates
   `original_agreement` and plot components.
3. **Input validation** — confirms the documented error
   conditions fire with expected messages.
4. **Backward compatibility** — simulates a pre-feature-branch
   object by dropping `grf_cuts` from `fold_summary`; confirms
   graceful NULL returns for the GRF-dependent slots.
5. **Cross-reference against manual aggregation** — reproduces
   `grf_cut_summary` using the `.clean_grf` + `.grf_dist`
   helpers lifted directly from `gbsg_10fold_dmin_sensitivity.qmd`,
   then asserts row-for-row equality with the function's output.
6. **Print method** — confirms `print.fs_cv_summary()` emits the
   expected metadata and section headers.
7. **Phase B diagnostic columns** — confirms the new
   `fold_summary` columns (`pconsistency`, `training_fs_hr`,
   `n_candidates_evaluated`) are present, typed correctly, `NA`
   exactly when their semantics require it, and in valid domains.
   Confirms the `pconsistency_distribution` and
   `fold_numeric_summary` slots are populated from the columns
   and produce the expected row counts.  Confirms
   backward-compatible behaviour when individual columns or the
   entire Phase B set are stripped.

The document uses `dmin.grf = 12, frac.tau = 0.8` (the strict
baseline from `gbsg_10fold_dmin_sensitivity.qmd` config A).  This
setting produces a diverse mix of `{er <= 0}`, `(no subgroup)`,
and non-`er` subgroups across folds, which exercises all
diagnostic branches even at small simulation counts.

**Checkpoint behaviour.** This document does not cache CV results;
the run executes every render.  At quick-mode sims, this takes on
the order of 1–2 minutes.


# Setup {#sec-setup}

```{r load-packages}
#| message: false
#| warning: false
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(ggplot2)
library(future)
library(foreach)
library(doFuture)
```

```{r config}
# --- Simulation size -------------------------------------------------------
if (isTRUE(params$quick_mode)) {
  Ksims_cv  <- 10L
  Kfolds_cv <- 10L
  cat("Mode: quick (sims = 10).  Set quick_mode = FALSE for sims = 100.\n")
} else {
  Ksims_cv  <- 100L
  Kfolds_cv <- 10L
  cat("Mode: full (sims = 100).  Expect 15-20 min runtime.\n")
}

# --- GBSG-specific parameters ---------------------------------------------
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name     <- "time_months"
event.name       <- "status"
treat.name       <- "hormon"
id.name          <- "id"

# --- forestsearch() base arguments (mirrors sensitivity QMD config A) -----
fs_args_base <- list(
  hr.threshold              = 1.25,
  hr.consistency            = 1.00,
  pconsistency.threshold    = 0.90,
  stop_threshold            = 0.95,
  sg_focus                  = "maxSG",
  max_subgroups_search      = 10L,
  use_twostage              = FALSE,
  use_grf                   = TRUE,
  return_selected_cuts_only = TRUE,
  use_lasso                 = TRUE,
  cut_type                  = "default",
  frac.tau                  = 0.8,
  dmin.grf                  = 12,
  maxk                      = 2L,
  n.min                     = 60L,
  d0.min                    = 10L,
  d1.min                    = 10L,
  fs.splits                 = 400L,
  details                   = FALSE,
  quiet                     = TRUE,
  plot.sg                   = FALSE,
  plot.grf                  = FALSE
)

# --- Parallel configuration ------------------------------------------------
n_workers <- floor(0.95 * max(1L, parallel::detectCores() - 1L))
```

```{r data-setup}
df              <- survival::gbsg
df$grade3       <- as.integer(df$grade == "3")
df$id           <- seq_len(nrow(df))
df$time_months  <- df$rfstime / 30.4375
```

```{r test-harness}
# Test-result collector.  Every test case appends a row; the final
# section renders these into a gt summary table.
test_results <- data.frame(
  Section = character(0),
  Test    = character(0),
  Status  = character(0),
  Detail  = character(0),
  stringsAsFactors = FALSE
)

.tc <- function(section, test, pass, detail = "") {
  status <- if (isTRUE(pass)) "PASS" else "FAIL"
  new_row <- data.frame(
    Section = section,
    Test    = test,
    Status  = status,
    Detail  = substr(detail, 1L, 120L),
    stringsAsFactors = FALSE
  )
  # Append to parent scope
  test_results <<- rbind(test_results, new_row)
  invisible(pass)
}

.expect_error <- function(section, test, expr, pattern) {
  err_msg <- NULL
  tryCatch(
    eval(expr),
    error = function(e) { err_msg <<- conditionMessage(e) }
  )
  pass   <- !is.null(err_msg) && grepl(pattern, err_msg, fixed = FALSE)
  detail <- if (!is.null(err_msg)) err_msg else "<no error thrown>"
  .tc(section, test, pass, detail)
}
```


# Baseline runs {#sec-baseline}

## Full-data ForestSearch (always runs)

```{r fs-full}
#| message: false
fs_full <- do.call(forestsearch, c(list(
  df.analysis      = df,
  confounders.name = confounders.name,
  outcome.name     = outcome.name,
  event.name       = event.name,
  treat.name       = treat.name,
  id.name          = id.name,
  parallel_args    = list(plan         = "multisession",
                          workers      = n_workers,
                          show_message = FALSE)
), fs_args_base))
plan("sequential")

cat("Full-data identified subgroup:",
    paste(fs_full$sg.harm, collapse = " & "), "\n")
cat("Full-data GRF cuts:         ",
    paste(fs_full$grf_cuts, collapse = " | "), "\n")
```

## 10-fold CV

```{r cv-run}
#| message: false
.run_cv <- function() {
  t0 <- proc.time()
  out <- forestsearch_tenfold(
    fs.est        = fs_full,
    sims          = Ksims_cv,
    Kfolds        = Kfolds_cv,
    details       = FALSE,
    keep_resCV    = FALSE,
    parallel_args = list(plan         = "multisession",
                         workers      = n_workers,
                         show_message = FALSE)
  )
  plan("sequential")
  attr(out, "runtime_seconds") <-
    as.numeric((proc.time() - t0)["elapsed"])
  out
}

message(sprintf("Running %d-sim x %d-fold CV ...",
                Ksims_cv, Kfolds_cv))
tf <- .run_cv()

cat(sprintf("CV completed: %d sims x %d folds = %d sim x fold pairs\n",
            tf$sims, tf$Kfolds, nrow(tf$fold_summary)))
cat(sprintf("fold_summary columns: %s\n",
            paste(names(tf$fold_summary), collapse = ", ")))

.tc("Baseline", "fold_summary has required columns",
    all(c("sim", "fold", "sg1", "sg2", "any_found") %in%
          names(tf$fold_summary)),
    detail = paste(names(tf$fold_summary), collapse = ", "))

.tc("Baseline", "fold_summary$grf_cuts present",
    "grf_cuts" %in% names(tf$fold_summary),
    detail = "Required for feature-branch diagnostic slots")

.tc("Baseline", "fold_summary$pconsistency present",
    "pconsistency" %in% names(tf$fold_summary),
    detail = "Required for Phase B pconsistency_distribution slot")

.tc("Baseline", "fold_summary$training_fs_hr present",
    "training_fs_hr" %in% names(tf$fold_summary),
    detail = "Phase B diagnostic column")

.tc("Baseline", "fold_summary$n_candidates_evaluated present",
    "n_candidates_evaluated" %in% names(tf$fold_summary),
    detail = "Phase B diagnostic column")
```


# Test 1: Minimal invocation {#sec-test1}

Calls `summarize_cv_results(tf)` with no optional arguments.
Every slot should be populated except `original_agreement` and
`plots` (which require opt-in).

```{r test1-run}
cv_diag <- summarize_cv_results(tf)

# Top-level slots
slot_status <- list(
  identification_summary    = !is.null(cv_diag$identification_summary),
  grf_cut_summary           = !is.null(cv_diag$grf_cut_summary),
  cut_vs_subgroup_xtab      = !is.null(cv_diag$cut_vs_subgroup_xtab),
  no_subgroup_decomposition = !is.null(cv_diag$no_subgroup_decomposition),
  original_agreement        = is.null(cv_diag$original_agreement),
  metrics_table             = !is.null(cv_diag$metrics_table),
  plots                     = is.null(cv_diag$plots)
)

for (nm in names(slot_status)) {
  expected_populated <- !grepl("^(original_agreement|plots)$", nm)
  observed_populated <- if (grepl("^(original_agreement|plots)$", nm)) {
    !slot_status[[nm]]
  } else {
    slot_status[[nm]]
  }
  .tc("Test 1: minimal",
      sprintf("slot %s matches expectation", nm),
      isTRUE(slot_status[[nm]]),
      detail = sprintf("populated = %s", observed_populated))
}

# Metadata
.tc("Test 1: minimal", "n_sims matches tf$sims",
    cv_diag$n_sims == tf$sims,
    detail = sprintf("obs=%d exp=%d", cv_diag$n_sims, tf$sims))
.tc("Test 1: minimal", "n_folds matches tf$Kfolds",
    cv_diag$n_folds == tf$Kfolds,
    detail = sprintf("obs=%d exp=%d", cv_diag$n_folds, tf$Kfolds))
.tc("Test 1: minimal", "total_pairs matches nrow(fold_summary)",
    cv_diag$total_pairs == nrow(tf$fold_summary),
    detail = sprintf("obs=%d exp=%d",
                     cv_diag$total_pairs, nrow(tf$fold_summary)))
.tc("Test 1: minimal", "has_grf_cuts == TRUE (feature branch)",
    isTRUE(cv_diag$has_grf_cuts))
.tc("Test 1: minimal", "class == 'fs_cv_summary'",
    inherits(cv_diag, "fs_cv_summary"))

# Raw data structure
.tc("Test 1: minimal", "$data$identification is data.frame",
    is.data.frame(cv_diag$data$identification))
.tc("Test 1: minimal", "$data$grf_cuts is data.frame",
    is.data.frame(cv_diag$data$grf_cuts))
.tc("Test 1: minimal", "identification df has expected columns",
    all(c("Rank", "Subgroup", "n_folds", "pct_folds") %in%
          names(cv_diag$data$identification)))
.tc("Test 1: minimal", "grf_cuts df has expected columns",
    all(c("Rank", "grf_cut", "n_folds", "pct_folds") %in%
          names(cv_diag$data$grf_cuts)))

# Row-count sanity.  When top_n truncates, an "(N others)" row
# aggregates the tail; the n_folds column must still sum to total_pairs.
total_expected <- nrow(tf$fold_summary)
total_id       <- sum(cv_diag$data$identification$n_folds)
.tc("Test 1: minimal",
    "identification n_folds sums to total pairs (including (N others))",
    total_id == total_expected,
    detail = sprintf("obs=%d exp=%d", total_id, total_expected))

total_grf <- sum(cv_diag$data$grf_cuts$n_folds)
.tc("Test 1: minimal",
    "grf_cuts n_folds sums to total pairs (including (N others))",
    total_grf == total_expected,
    detail = sprintf("obs=%d exp=%d", total_grf, total_expected))

# If the "(N others)" row is present, it should carry the tail count.
id_has_others  <- any(grepl("^\\(\\d+ others\\)$",
                            cv_diag$data$identification$Subgroup))
grf_has_others <- any(grepl("^\\(\\d+ others\\)$",
                            cv_diag$data$grf_cuts$grf_cut))
.tc("Test 1: minimal", "identification_summary row count <= top_n + 1",
    nrow(cv_diag$data$identification) <= 16L,
    detail = sprintf("rows=%d; has (N others) row=%s",
                     nrow(cv_diag$data$identification), id_has_others))
.tc("Test 1: minimal", "grf_cut_summary row count <= top_n + 1",
    nrow(cv_diag$data$grf_cuts) <= 16L,
    detail = sprintf("rows=%d; has (N others) row=%s",
                     nrow(cv_diag$data$grf_cuts), grf_has_others))
```

## Rendered gt tables from minimal invocation

### Identification summary

```{r test1-display-id}
cv_diag$identification_summary
```

### GRF cut summary

```{r test1-display-grf}
cv_diag$grf_cut_summary
```

### Cut × subgroup cross-tab

```{r test1-display-xtab}
cv_diag$cut_vs_subgroup_xtab
```

### No-subgroup decomposition

```{r test1-display-nosg}
cv_diag$no_subgroup_decomposition
```

### Metrics table (delegated to `cv_metrics_tables()`)

```{r test1-display-metrics}
cv_diag$metrics_table
```


# Test 2: Full invocation with originals and plots {#sec-test2}

```{r test2-run}
cv_diag_full <- summarize_cv_results(
  cv_output         = tf,
  original_sg       = fs_full$sg.harm,
  original_grf_cuts = fs_full$grf_cuts,
  create_plots      = TRUE
)

.tc("Test 2: full", "$original_agreement populated",
    !is.null(cv_diag_full$original_agreement))
.tc("Test 2: full", "$plots populated",
    !is.null(cv_diag_full$plots))
.tc("Test 2: full", "$plots$identification is ggplot",
    inherits(cv_diag_full$plots$identification, "ggplot"))
.tc("Test 2: full", "$plots$grf_cut is ggplot (has_grf_cuts TRUE)",
    inherits(cv_diag_full$plots$grf_cut, "ggplot"))

# original_agreement raw data structure
oa_df <- cv_diag_full$data$original_agreement
.tc("Test 2: full", "original_agreement df has Metric, Value",
    all(c("Metric", "Value") %in% names(oa_df)))
.tc("Test 2: full", "original_agreement includes Exact subgroup match row",
    any(grepl("Exact subgroup match", oa_df$Metric)))
.tc("Test 2: full", "original_agreement includes Exact GRF-cut match row",
    any(grepl("Exact GRF-cut match", oa_df$Metric)))
.tc("Test 2: full", "original_agreement includes Both match row",
    any(grepl("Both subgroup and GRF cut match", oa_df$Metric)))
```

## Rendered original agreement table

```{r test2-display-original}
cv_diag_full$original_agreement
```

## Rendered plots

```{r test2-display-plot-id, fig.height=4}
cv_diag_full$plots$identification
```

```{r test2-display-plot-grf, fig.height=4}
cv_diag_full$plots$grf_cut
```


# Test 3: Input validation {#sec-test3}

Confirms each documented error condition fires with the
expected message pattern.

```{r test3-run}
# Case 3.1: non-list input
.expect_error(
  "Test 3: errors", "non-list input rejected",
  expr    = quote(summarize_cv_results(cv_output = "not a list")),
  pattern = "must be a list"
)

# Case 3.2: empty fold_summary
tf_empty <- tf
tf_empty$fold_summary <- tf$fold_summary[0, , drop = FALSE]
.expect_error(
  "Test 3: errors", "empty fold_summary rejected",
  expr    = quote(summarize_cv_results(cv_output = tf_empty)),
  pattern = "missing or empty"
)

# Case 3.3: missing required column
tf_bad <- tf
tf_bad$fold_summary$sg1 <- NULL
.expect_error(
  "Test 3: errors", "missing required column rejected",
  expr    = quote(summarize_cv_results(cv_output = tf_bad)),
  pattern = "missing required column"
)

# Case 3.4: bad top_n
.expect_error(
  "Test 3: errors", "non-positive top_n rejected",
  expr    = quote(summarize_cv_results(cv_output = tf, top_n = 0)),
  pattern = "positive integer"
)

# Case 3.5: fs_kfold redirection
# We synthesise a minimal fs_kfold object with the right class; it must
# emit the pointer to cv_summary_tables()/cv_metrics_tables().
tf_kfold_fake <- structure(
  list(fold_summary = tf$fold_summary),
  class = c("fs_kfold", "list")
)
.expect_error(
  "Test 3: errors", "fs_kfold object redirected",
  expr    = quote(summarize_cv_results(cv_output = tf_kfold_fake)),
  pattern = "cv_summary_tables|cv_metrics_tables"
)
```


# Test 4: Backward compatibility {#sec-test4}

Drops `grf_cuts` from `fold_summary` to simulate a
pre-feature-branch object.  Expects graceful NULL for the
GRF-dependent slots and a populated `identification_summary`.

```{r test4-run}
tf_nogrf <- tf
tf_nogrf$fold_summary <- tf$fold_summary[
  , setdiff(names(tf$fold_summary), "grf_cuts"),
  drop = FALSE
]

cv_diag_nogrf <- summarize_cv_results(tf_nogrf)

.tc("Test 4: backward-compat", "has_grf_cuts FALSE",
    isFALSE(cv_diag_nogrf$has_grf_cuts))
.tc("Test 4: backward-compat", "identification_summary populated",
    !is.null(cv_diag_nogrf$identification_summary))
.tc("Test 4: backward-compat", "grf_cut_summary NULL",
    is.null(cv_diag_nogrf$grf_cut_summary))
.tc("Test 4: backward-compat", "cut_vs_subgroup_xtab NULL",
    is.null(cv_diag_nogrf$cut_vs_subgroup_xtab))
.tc("Test 4: backward-compat", "no_subgroup_decomposition NULL",
    is.null(cv_diag_nogrf$no_subgroup_decomposition))
.tc("Test 4: backward-compat", "$data$grf_cuts NULL",
    is.null(cv_diag_nogrf$data$grf_cuts))
.tc("Test 4: backward-compat", "metrics_table still populated",
    !is.null(cv_diag_nogrf$metrics_table))
```


# Test 5: Cross-reference against manual aggregation {#sec-test5}

Reproduces `grf_cut_summary` and `identification_summary` using
the `.clean_grf` + `.grf_dist` + `.build_sg_full` helpers lifted
directly from `gbsg_10fold_dmin_sensitivity.qmd`, then asserts
row-for-row equality with the function output.  This is the
refactor-safety check: the function should produce identical
results to the manual one-off aggregation the sensitivity QMD
performed.

To isolate the tabulation logic from the top-n truncation layer,
this test calls `summarize_cv_results(..., top_n = Inf)` so the
function returns the full frequency distribution without an
`"(N others)"` aggregation row.  The row-sum invariant under
truncation is covered separately in Test 1.

```{r test5-untruncated}
# Untruncated variant for byte-exact comparison
cv_diag_full_topn <- summarize_cv_results(tf, top_n = Inf)
```

```{r test5-manual-helpers}
# Lifted verbatim from gbsg_10fold_dmin_sensitivity.qmd
.clean_grf_manual <- function(x) {
  if (is.na(x) || x == "") return("(no GRF cut)")
  x
}

.grf_dist_manual <- function(fsum, top_n = Inf) {
  cuts <- vapply(fsum$grf_cuts, .clean_grf_manual, character(1))
  tab  <- as.data.frame(table(cuts = cuts), stringsAsFactors = FALSE)
  tab  <- tab[order(-tab$Freq), , drop = FALSE]
  if (is.finite(top_n) && nrow(tab) > top_n) tab <- utils::head(tab, top_n)
  tab
}

.build_sg_full_manual <- function(fsum) {
  ifelse(
    is.na(fsum$sg1),
    "(no subgroup)",
    ifelse(
      is.na(fsum$sg2),
      fsum$sg1,
      paste(fsum$sg1, fsum$sg2, sep = " & ")
    )
  )
}
```

```{r test5-compare-grf}
# Manual (no truncation)
manual_grf <- .grf_dist_manual(tf$fold_summary)
manual_grf_aligned <- data.frame(
  grf_cut = as.character(manual_grf$cuts),
  n_folds = as.integer(manual_grf$Freq),
  stringsAsFactors = FALSE
)
rownames(manual_grf_aligned) <- NULL

# Function (untruncated)
fn_grf <- cv_diag_full_topn$data$grf_cuts
fn_grf_aligned <- data.frame(
  grf_cut = fn_grf$grf_cut,
  n_folds = as.integer(fn_grf$n_folds),
  stringsAsFactors = FALSE
)
rownames(fn_grf_aligned) <- NULL

# Same number of rows?
.tc("Test 5: cross-ref", "grf_cut_summary row count matches manual",
    nrow(manual_grf_aligned) == nrow(fn_grf_aligned),
    detail = sprintf("manual=%d fn=%d",
                     nrow(manual_grf_aligned),
                     nrow(fn_grf_aligned)))

# Row-for-row equality
eq_ok <- isTRUE(all.equal(manual_grf_aligned, fn_grf_aligned))
.tc("Test 5: cross-ref", "grf_cut_summary rows match manual exactly",
    eq_ok,
    detail = if (eq_ok) "all.equal TRUE" else
      paste(utils::capture.output(
        all.equal(manual_grf_aligned, fn_grf_aligned)
      ), collapse = "; "))
```

```{r test5-compare-id}
# Manual identification tabulation (no truncation)
manual_sg <- .build_sg_full_manual(tf$fold_summary)
tab <- as.data.frame(table(Subgroup = manual_sg),
                     stringsAsFactors = FALSE)
tab <- tab[order(-tab$Freq), , drop = FALSE]

manual_id_aligned <- data.frame(
  Subgroup = as.character(tab$Subgroup),
  n_folds  = as.integer(tab$Freq),
  stringsAsFactors = FALSE
)
rownames(manual_id_aligned) <- NULL

fn_id <- cv_diag_full_topn$data$identification
fn_id_aligned <- data.frame(
  Subgroup = fn_id$Subgroup,
  n_folds  = as.integer(fn_id$n_folds),
  stringsAsFactors = FALSE
)
rownames(fn_id_aligned) <- NULL

.tc("Test 5: cross-ref", "identification_summary row count matches manual",
    nrow(manual_id_aligned) == nrow(fn_id_aligned),
    detail = sprintf("manual=%d fn=%d",
                     nrow(manual_id_aligned),
                     nrow(fn_id_aligned)))

eq_ok <- isTRUE(all.equal(manual_id_aligned, fn_id_aligned))
.tc("Test 5: cross-ref", "identification_summary rows match manual exactly",
    eq_ok,
    detail = if (eq_ok) "all.equal TRUE" else
      paste(utils::capture.output(
        all.equal(manual_id_aligned, fn_id_aligned)
      ), collapse = "; "))
```

```{r test5-compare-nosg-total}
# Manual no-subgroup count
manual_no_sg <- sum(manual_sg == "(no subgroup)")
# Function
fn_no_sg <- cv_diag$data$no_subgroup$n_folds[
  cv_diag$data$no_subgroup$Category == "Folds with no subgroup identified"
]

.tc("Test 5: cross-ref", "no_subgroup total matches manual",
    isTRUE(manual_no_sg == fn_no_sg),
    detail = sprintf("manual=%d fn=%d", manual_no_sg, fn_no_sg))

# Manual GRF-none-among-no-sg count
manual_grf_none_in_no_sg <- sum(
  (manual_sg == "(no subgroup)") &
    vapply(tf$fold_summary$grf_cuts, .clean_grf_manual,
           character(1)) == "(no GRF cut)"
)
fn_grf_none_in_no_sg <- cv_diag$data$no_subgroup$n_folds[
  cv_diag$data$no_subgroup$Category == "  GRF returned no cut"
]

.tc("Test 5: cross-ref", "decomposition 'GRF none' matches manual",
    isTRUE(manual_grf_none_in_no_sg == fn_grf_none_in_no_sg),
    detail = sprintf("manual=%d fn=%d",
                     manual_grf_none_in_no_sg, fn_grf_none_in_no_sg))
```


# Test 6: Print method {#sec-test6}

Confirms `print.fs_cv_summary()` emits the expected header
lines and section markers.

```{r test6-run}
out <- utils::capture.output(print(cv_diag))

.tc("Test 6: print", "print output non-empty",
    length(out) > 0L)
.tc("Test 6: print", "print output contains 'Simulations'",
    any(grepl("Simulations:", out)))
.tc("Test 6: print", "print output contains 'Total sim x fold pairs'",
    any(grepl("Total sim x fold pairs", out)))
.tc("Test 6: print", "print output contains 'Top identified subgroups'",
    any(grepl("Top identified subgroups", out)))
.tc("Test 6: print", "print output contains 'Top GRF cuts'",
    any(grepl("Top GRF cuts", out)))
.tc("Test 6: print", "print output contains 'No-subgroup decomposition'",
    any(grepl("No-subgroup decomposition", out)))
```

Full captured output:

```{r test6-display, results='asis'}
cat("```\n")
cat(out, sep = "\n")
cat("\n```\n")
```


# Test 7: Phase B diagnostic columns {#sec-test7}

Validates the three new `fold_summary` columns populated by
`forestsearch_tenfold()` on the feature branch
(`pconsistency`, `training_fs_hr`, `n_candidates_evaluated`)
and the two new summary slots exposed by
`summarize_cv_results()` (`pconsistency_distribution`,
`fold_numeric_summary`).

## 7.1 Column presence and value domain

```{r test7-column}
has_pcons <- "pconsistency" %in% names(tf$fold_summary)
.tc("Test 7: pconsistency", "fold_summary$pconsistency column present",
    has_pcons,
    detail = if (has_pcons) "yes" else
      "NO -- rerun forestsearch_tenfold() with Phase B branch")

if (has_pcons) {
  pv <- tf$fold_summary$pconsistency

  .tc("Test 7: pconsistency", "pconsistency is numeric",
      is.numeric(pv),
      detail = sprintf("class=%s", class(pv)[1]))

  # NA iff no subgroup identified
  na_aligned <- all(
    is.na(pv) == (tf$fold_summary$any_found == 0L)
  )
  .tc("Test 7: pconsistency",
      "NA pconsistency iff any_found == 0",
      na_aligned,
      detail = sprintf(
        "NA count=%d; any_found==0 count=%d",
        sum(is.na(pv)),
        sum(tf$fold_summary$any_found == 0L)))

  # Domain check: non-NA values should be in [0, 1]
  pv_nona <- pv[!is.na(pv)]
  in_range <- length(pv_nona) == 0L ||
    (all(pv_nona >= 0) && all(pv_nona <= 1))
  .tc("Test 7: pconsistency",
      "non-NA pconsistency values in [0, 1]",
      in_range,
      detail = if (length(pv_nona) > 0L)
        sprintf("range = [%.3f, %.3f], median = %.3f",
                min(pv_nona), max(pv_nona), median(pv_nona))
      else "no identifying folds")

  # Given pconsistency.threshold = 0.90 in config, all identifying
  # Pcons should be >= that threshold (by construction of
  # forestsearch()'s selection rule).
  thresh <- fs_args_base$pconsistency.threshold
  above_thresh <- length(pv_nona) == 0L || all(pv_nona >= thresh)
  .tc("Test 7: pconsistency",
      sprintf("identifying Pcons >= threshold (%.2f)", thresh),
      above_thresh,
      detail = if (length(pv_nona) > 0L)
        sprintf("min = %.3f; threshold = %.2f", min(pv_nona), thresh)
      else "no identifying folds")
}
```

## 7.2 Slot populated when column present

```{r test7-slot-present}
if (has_pcons) {
  .tc("Test 7: pconsistency",
      "has_pconsistency metadata TRUE",
      isTRUE(cv_diag$has_pconsistency))
  .tc("Test 7: pconsistency",
      "pconsistency_distribution slot populated",
      !is.null(cv_diag$pconsistency_distribution))
  .tc("Test 7: pconsistency",
      "$data$pconsistency is data.frame",
      is.data.frame(cv_diag$data$pconsistency))

  pcd <- cv_diag$data$pconsistency
  .tc("Test 7: pconsistency",
      "pconsistency df has expected columns",
      all(c("Pcons_range", "n_folds", "pct_of_identifying") %in%
            names(pcd)))

  # Summary rows carry median/IQR/total — last three rows
  last_rows <- pcd$Pcons_range[(nrow(pcd) - 2L):nrow(pcd)]
  .tc("Test 7: pconsistency",
      "last three rows include Median/IQR/total summary",
      any(grepl("Median", last_rows)) &&
      any(grepl("IQR",    last_rows)) &&
      any(grepl("Identifying", last_rows)),
      detail = paste(last_rows, collapse = " | "))

  # Bin row counts should sum to number of identifying folds
  n_identifying <- sum(!is.na(tf$fold_summary$pconsistency))
  bin_rows <- pcd[1:6, , drop = FALSE]  # 6 bins from 0 to 1
  bin_sum <- sum(bin_rows$n_folds, na.rm = TRUE)
  .tc("Test 7: pconsistency",
      "bin n_folds sums to identifying-fold count",
      bin_sum == n_identifying,
      detail = sprintf("sum=%d exp=%d", bin_sum, n_identifying))
}
```

## 7.3 Backward compatibility (column stripped)

```{r test7-backcompat}
if (has_pcons) {
  tf_nopcons <- tf
  tf_nopcons$fold_summary <- tf$fold_summary[
    , setdiff(names(tf$fold_summary), "pconsistency"),
    drop = FALSE
  ]
  cv_diag_nopcons <- summarize_cv_results(tf_nopcons)

  .tc("Test 7: pconsistency",
      "has_pconsistency FALSE when column absent",
      isFALSE(cv_diag_nopcons$has_pconsistency))
  .tc("Test 7: pconsistency",
      "pconsistency_distribution NULL when column absent",
      is.null(cv_diag_nopcons$pconsistency_distribution))
  .tc("Test 7: pconsistency",
      "$data$pconsistency NULL when column absent",
      is.null(cv_diag_nopcons$data$pconsistency))
  .tc("Test 7: pconsistency",
      "other slots still populated when pconsistency absent",
      !is.null(cv_diag_nopcons$identification_summary) &&
      !is.null(cv_diag_nopcons$grf_cut_summary))
}
```

## 7.4 Rendered pconsistency_distribution table

```{r test7-display}
if (has_pcons) cv_diag$pconsistency_distribution
```

## 7.5 `training_fs_hr` column

```{r test7-hr}
has_hr <- "training_fs_hr" %in% names(tf$fold_summary)
.tc("Test 7: pconsistency",
    "fold_summary$training_fs_hr column present",
    has_hr,
    detail = if (has_hr) "yes" else
      "NO -- rerun forestsearch_tenfold() with Phase B branch")

if (has_hr) {
  hr <- tf$fold_summary$training_fs_hr
  .tc("Test 7: pconsistency", "training_fs_hr is numeric",
      is.numeric(hr),
      detail = sprintf("class=%s", class(hr)[1]))

  # NA iff no subgroup identified (same gate as pconsistency)
  na_aligned <- all(is.na(hr) == (tf$fold_summary$any_found == 0L))
  .tc("Test 7: pconsistency",
      "NA training_fs_hr iff any_found == 0",
      na_aligned,
      detail = sprintf(
        "NA count=%d; any_found==0 count=%d",
        sum(is.na(hr)),
        sum(tf$fold_summary$any_found == 0L)))

  # Domain: non-NA HRs should be positive and finite
  hr_nona <- hr[!is.na(hr)]
  in_domain <- length(hr_nona) == 0L ||
    (all(hr_nona > 0) && all(is.finite(hr_nona)))
  .tc("Test 7: pconsistency",
      "non-NA training_fs_hr values positive and finite",
      in_domain,
      detail = if (length(hr_nona) > 0L)
        sprintf("range = [%.3f, %.3f], median = %.3f",
                min(hr_nona), max(hr_nona), median(hr_nona))
      else "no identifying folds")

  # GBSG at sg_focus = "maxSG" with hr.threshold = 1.25 forces
  # identifying HRs above the threshold.  Sanity check.
  hr_thresh <- fs_args_base$hr.threshold
  above_hr_thresh <- length(hr_nona) == 0L || all(hr_nona >= hr_thresh)
  .tc("Test 7: pconsistency",
      sprintf("identifying HRs >= hr.threshold (%.2f)", hr_thresh),
      above_hr_thresh,
      detail = if (length(hr_nona) > 0L)
        sprintf("min = %.3f; threshold = %.2f", min(hr_nona), hr_thresh)
      else "no identifying folds")
}
```

## 7.6 `n_candidates_evaluated` column

```{r test7-cands}
has_cands <- "n_candidates_evaluated" %in% names(tf$fold_summary)
.tc("Test 7: pconsistency",
    "fold_summary$n_candidates_evaluated column present",
    has_cands,
    detail = if (has_cands) "yes" else
      "NO -- rerun forestsearch_tenfold() with Phase B branch")

if (has_cands) {
  nc <- tf$fold_summary$n_candidates_evaluated
  .tc("Test 7: pconsistency", "n_candidates_evaluated is integer",
      is.integer(nc),
      detail = sprintf("class=%s", class(nc)[1]))

  # Every fold that identified a subgroup must have evaluated
  # at least 1 candidate.  The converse is NOT true: a fold can
  # have evaluated many candidates and still identify none.
  identifying_idx <- which(tf$fold_summary$any_found == 1L)
  if (length(identifying_idx) > 0L) {
    nc_identifying <- nc[identifying_idx]
    nc_identifying_ok <- all(
      !is.na(nc_identifying) & nc_identifying >= 1L
    )
    .tc("Test 7: pconsistency",
        "identifying folds have n_candidates_evaluated >= 1",
        nc_identifying_ok,
        detail = sprintf(
          "identifying folds: n=%d, min candidates=%s",
          length(identifying_idx),
          if (all(is.na(nc_identifying))) "all NA"
          else as.character(min(nc_identifying, na.rm = TRUE))))
  }

  # Domain: non-NA values should be non-negative
  nc_nona <- nc[!is.na(nc)]
  in_domain <- length(nc_nona) == 0L ||
    (all(nc_nona >= 0L) && all(is.finite(nc_nona)))
  .tc("Test 7: pconsistency",
      "non-NA n_candidates_evaluated values non-negative",
      in_domain,
      detail = if (length(nc_nona) > 0L)
        sprintf("range = [%d, %d], median = %d",
                as.integer(min(nc_nona)),
                as.integer(max(nc_nona)),
                as.integer(median(nc_nona)))
      else "no populated values")
}
```

## 7.7 `fold_numeric_summary` slot

```{r test7-numsum}
.tc("Test 7: pconsistency",
    "fold_numeric_summary slot populated",
    !is.null(cv_diag$fold_numeric_summary))
.tc("Test 7: pconsistency",
    "$data$fold_numeric_summary is data.frame",
    is.data.frame(cv_diag$data$fold_numeric_summary))

fns <- cv_diag$data$fold_numeric_summary
expected_cols <- c("Metric", "n", "median", "Q1", "Q3", "min",
                   "max", "n_na", "Context")
.tc("Test 7: pconsistency",
    "fold_numeric_summary has expected columns",
    all(expected_cols %in% names(fns)),
    detail = paste(setdiff(expected_cols, names(fns)),
                   collapse = ", "))

# Given GBSG + Phase B, all four metric rows should appear.
expected_metrics <- c(
  "Test-fold size (n_test)",
  "Pcons (identifying folds)",
  "Training-fold subgroup effect (identifying folds)",
  "Candidates evaluated per fold"
)
.tc("Test 7: pconsistency",
    "fold_numeric_summary includes all four Phase B metric rows",
    all(expected_metrics %in% fns$Metric),
    detail = paste(setdiff(expected_metrics, fns$Metric),
                   collapse = ", "))

# Row counts must match the per-column non-NA counts.
for (spec in list(
  list(metric = "Pcons (identifying folds)",
       col    = "pconsistency",
       filt   = function(x) !is.na(x)),
  list(metric = "Training-fold subgroup effect (identifying folds)",
       col    = "training_fs_hr",
       filt   = function(x) !is.na(x)),
  list(metric = "Candidates evaluated per fold",
       col    = "n_candidates_evaluated",
       filt   = function(x) !is.na(x)),
  list(metric = "Test-fold size (n_test)",
       col    = "n_test",
       filt   = function(x) !is.na(x))
)) {
  if (!spec$col %in% names(tf$fold_summary)) next
  row    <- fns[fns$Metric == spec$metric, , drop = FALSE]
  actual <- sum(spec$filt(tf$fold_summary[[spec$col]]))
  .tc("Test 7: pconsistency",
      sprintf("%s: n matches non-NA count", spec$metric),
      nrow(row) == 1L && row$n == actual,
      detail = sprintf("row$n=%d actual=%d",
                       if (nrow(row) == 1L) row$n else NA_integer_,
                       actual))
}
```

## 7.8 Rendered fold_numeric_summary table

```{r test7-numsum-display}
cv_diag$fold_numeric_summary
```

## 7.9 Backward compat: strip ALL Phase B columns

Also confirms that stripping Phase B columns wholesale doesn't
disturb the non-Phase-B slots.

```{r test7-backcompat-all}
tf_noB <- tf
tf_noB$fold_summary <- tf$fold_summary[
  , setdiff(names(tf$fold_summary),
            c("pconsistency", "training_fs_hr",
              "n_candidates_evaluated")),
  drop = FALSE
]
cv_diag_noB <- summarize_cv_results(tf_noB)

.tc("Test 7: pconsistency",
    "pconsistency_distribution NULL when all Phase B stripped",
    is.null(cv_diag_noB$pconsistency_distribution))
.tc("Test 7: pconsistency",
    "fold_numeric_summary still populated (n_test always present)",
    !is.null(cv_diag_noB$fold_numeric_summary))
.tc("Test 7: pconsistency",
    "non-Phase-B slots still populated when Phase B stripped",
    !is.null(cv_diag_noB$identification_summary) &&
      !is.null(cv_diag_noB$grf_cut_summary) &&
      !is.null(cv_diag_noB$no_subgroup_decomposition))

# fold_numeric_summary should have exactly 1 row (n_test only)
fns_noB <- cv_diag_noB$data$fold_numeric_summary
.tc("Test 7: pconsistency",
    "fold_numeric_summary has only n_test row when Phase B stripped",
    nrow(fns_noB) == 1L && fns_noB$Metric[1] == "Test-fold size (n_test)",
    detail = sprintf("rows=%d, metrics=%s",
                     nrow(fns_noB),
                     paste(fns_noB$Metric, collapse = ", ")))
```


# Test summary {#sec-summary}

```{r summary-table}
n_total <- nrow(test_results)
n_pass  <- sum(test_results$Status == "PASS")
n_fail  <- n_total - n_pass

gt::gt(test_results,
       groupname_col = "Section",
       row_group_as_column = TRUE) |>
  gt::tab_header(
    title    = gt::md(sprintf(
      "**Living test results: %d / %d passing**", n_pass, n_total)),
    subtitle = sprintf(
      "GBSG, %d sims x %d folds, dmin.grf = %g, frac.tau = %g",
      Ksims_cv, Kfolds_cv,
      fs_args_base$dmin.grf, fs_args_base$frac.tau)
  ) |>
  gt::tab_style(
    style    = gt::cell_fill(color = "#d4edda"),
    locations = gt::cells_body(
      columns = "Status", rows = Status == "PASS"
    )
  ) |>
  gt::tab_style(
    style    = list(
      gt::cell_fill(color = "#f8d7da"),
      gt::cell_text(weight = "bold")
    ),
    locations = gt::cells_body(
      columns = "Status", rows = Status == "FAIL"
    )
  ) |>
  gt::cols_label(
    Test   = "Test case",
    Status = "Status",
    Detail = "Detail"
  ) |>
  gt::tab_options(table.font.size = gt::px(12))

# Also emit a one-line headline suitable for grep-based CI.
if (n_fail == 0L) {
  message(sprintf("ALL TESTS PASSED: %d / %d", n_pass, n_total))
} else {
  message(sprintf("TESTS FAILED: %d / %d failing", n_fail, n_total))
  print(test_results[test_results$Status == "FAIL",
                     c("Section", "Test", "Detail")])
}
```


# Session info {#sec-session}

```{r session}
sessionInfo()
```
