---
title: "Living test: Phase A/B/C GLM compatibility (continuous / MD)"
subtitle: "Synthetic continuous outcome + mean difference on identity scale"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-tools: true
    code-fold: true
    theme: cosmo
    fontsize: 11pt
    number-sections: true
params:
  quick_mode: true       # TRUE: smaller grid (fast). FALSE: larger.
---

# Purpose {#sec-purpose}

Validates that the Phase A/B/C diagnostic features
(`fold_summary$grf_cuts`, `fold_summary$pconsistency`,
`fold_summary$training_fs_hr`, `fold_summary$n_candidates_evaluated`,
`summarize_cv_results()` slots, and the bootstrap `grf_cut_freq`
slot) work correctly when the outcome is **continuous** (Gaussian
GLM with identity link) rather than survival, binary, or count.

Parallel in intent to `test_glm_compatibility.qmd`, but specialised
to the continuous / MD path.  The continuous-specific concerns are
*inverse* to those of the log-link (binary / count) paths:

1. `training_fs_hr` on an identity-link measure (MD): the effect is
   *already* on the natural (additive) scale.  Phase B must NOT
   apply `exp()` here — values should be raw mean differences, not
   multiplicatively transformed.  A regression where `exp()` leaks
   into the identity-scale path would produce values near
   `exp(MD_harm)` instead of `MD_harm`.
2. The `hr.threshold` convention for identity-scale searches: for
   MD the threshold is a raw outcome-scale increment, not a
   multiplicative ratio.  A default of 0.5 means "at least 0.5
   outcome-scale units worse".

The synthetic DGM plants a harm signal driven by
`age <= 50 & biomarker_hi == 1`, with true MD ≈ 3.0 inside the
subgroup and ≈ 0 outside.  Covariates mix continuous (GRF cuts
discoverable) and binary factor (passed through without cut
attempt), with `sg_focus = "maxSG"` to recover the largest
qualifying subgroup.

Coverage:

1. **Synthetic data generation** — harm subgroup driven by
   `age <= 50 & biomarker_hi == 1`, true MD ≈ 3.0, mix of continuous
   and factor covariates.
2. **Primary ForestSearch call** — confirms the continuous GLM
   routing succeeds, outcome-type / effect-measure metadata is
   preserved, and the stored effect value is on identity scale
   (not log-transformed).
3. **CV pipeline (Phase A/B)** — confirms every new `fold_summary`
   column populates on continuous data, confirms
   `summarize_cv_results()` produces each diagnostic slot, and
   confirms `training_fs_hr` values are on MD / identity scale
   (i.e., values in the sensible range around `MD_harm`, not
   exponentiated to e.g. `exp(3) ~ 20`).
4. **Bootstrap pipeline (Phase C)** — confirms `results$grf_cuts_b`
   populates and `summarize_bootstrap_subgroups()` surfaces
   `grf_cut_freq`.
5. **v0.2.0 bootstrap return-value contract** — `any_found` column,
   `fs_bootstrap` S3 class tag, metadata (nb_boots, outcome_type,
   effect_measure, est.scale), and `print.fs_bootstrap` dispatch
   work on the continuous path.
6. **Cross-check** — captured `training_fs_hr` on an identifying
   fold agrees within a small additive tolerance with an independent
   `lm()` refit on the same subgroup.

A cumulative pass/fail table renders in @sec-summary.


# Setup {#sec-setup}

```{r load-packages}
#| message: false
#| warning: false
library(forestsearch)
library(data.table)
library(gt)
library(future)
library(foreach)
library(doFuture)
```

```{r config}
if (isTRUE(params$quick_mode)) {
  N         <- 800L
  Ksims_cv  <- 5L
  Kfolds_cv <- 10L
  nb_boots  <- 20L
  mode_tag  <- "quick"
} else {
  N         <- 1500L
  Ksims_cv  <- 20L
  Kfolds_cv <- 10L
  nb_boots  <- 100L
  mode_tag  <- "full"
}
cat(sprintf("Mode: %s (N=%d, Ksims=%d, nb_boots=%d)\n",
            mode_tag, N, Ksims_cv, nb_boots))

n_workers <- min(4L, floor(0.95 * max(1L, parallel::detectCores() - 1L)))
```

```{r test-harness}
test_results <- data.frame(
  Section = character(0),
  Test    = character(0),
  Status  = character(0),
  Detail  = character(0),
  stringsAsFactors = FALSE
)

.tc <- function(section, test, pass, detail = "") {
  status <- if (isTRUE(pass)) "PASS" else "FAIL"
  test_results <<- rbind(test_results, data.frame(
    Section = section, Test = test, Status = status,
    Detail  = substr(detail, 1L, 140L),
    stringsAsFactors = FALSE
  ))
  invisible(pass)
}
```


# Synthetic continuous-outcome DGM {#sec-dgm}

Gaussian outcome with treatment-by-subgroup interaction on the
identity scale: within the harm subgroup the treatment arm has
mean shifted by `MD_harm`; outside the subgroup treatment has no
effect.  Harm signal jointly defined by `age <= 50` and
`biomarker_hi == 1`.

```{r generate-data}
set.seed(20260421)

MD_harm    <- 3.0     # true MD in harm subgroup (treatment worsens mean)
MD_benefit <- 0.0     # null treatment effect outside harm subgroup

df_syn <- data.frame(
  id           = seq_len(N),
  treat        = rbinom(N, 1L, 0.5),
  age          = round(rnorm(N, mean = 55, sd = 12)),
  wtkg         = round(rnorm(N, mean = 75, sd = 12), 1),
  biomarker    = rnorm(N),
  karnof       = sample(70:100, N, replace = TRUE),
  biomarker_hi = factor(rbinom(N, 1L, 0.5), levels = c(0L, 1L)),
  sex          = factor(rbinom(N, 1L, 0.5), levels = c(0L, 1L)),
  site         = factor(sample(0:1, N, replace = TRUE),
                        levels = c(0L, 1L))
)

# True harm subgroup: age <= 50 AND biomarker_hi == 1
true_harm    <- df_syn$age <= 50 & df_syn$biomarker_hi == "1"
treat_effect <- ifelse(true_harm, MD_harm, MD_benefit)

# Linear predictor is the mean; add Gaussian noise with sd = 1 so
# MD_harm = 3 corresponds to ~3-sigma signal inside the subgroup.
mu       <- df_syn$treat * treat_effect
df_syn$y <- rnorm(N, mean = mu, sd = 1)

cat(sprintf("N = %d, mean(y) = %.3f, sd(y) = %.3f\n",
            nrow(df_syn), mean(df_syn$y), sd(df_syn$y)))
cat(sprintf("True harm prevalence: %.2f\n", mean(true_harm)))
cat(sprintf("Mean y | treat = 0: %.3f\n",
            mean(df_syn$y[df_syn$treat == 0])))
cat(sprintf("Mean y | treat = 1: %.3f\n",
            mean(df_syn$y[df_syn$treat == 1])))

overall_md <- mean(df_syn$y[df_syn$treat == 1]) -
              mean(df_syn$y[df_syn$treat == 0])
cat(sprintf("Overall empirical MD: %.3f\n", overall_md))
```

Empirical check that the harm subgroup is recoverable:

```{r dgm-sanity}
har <- df_syn[true_harm, ]
md_harm_emp <- mean(har$y[har$treat == 1]) -
               mean(har$y[har$treat == 0])
cat(sprintf("Empirical MD in true harm subgroup: %.2f (DGM = %.1f)\n",
            md_harm_emp, MD_harm))

.tc("DGM", "empirical harm MD > 1.5",
    md_harm_emp > 1.5,
    detail = sprintf("empirical MD = %.2f", md_harm_emp))

.tc("DGM", "harm subgroup prevalence between 15% and 40%",
    mean(true_harm) > 0.15 && mean(true_harm) < 0.40,
    detail = sprintf("prev = %.2f", mean(true_harm)))
```


# Primary ForestSearch call {#sec-fs-full}

Configuration exercises the continuous / MD pathway on identity
scale:

- `outcome_type = "continuous"`, `effect_measure = "MD"`
- `sg_focus = "maxSG"` (largest qualifying subgroup)
- `is.RCT = TRUE`, `vi.grf.min = -0.2`, `cont.cutoff = 4L`
- `hr.threshold = 0.5` — identity-scale increment (not multiplicative)
- `hr.consistency = 0.0` — identity-scale baseline (not 1.0 as for log links)
- Mixed covariates: continuous (cuts discovered by GRF) + factor
  (passed through without cut attempt)
- `use_grf = TRUE`, `use_lasso = FALSE`

```{r fs-full}
#| message: false
fs_args <- list(
  confounders.name = c("age", "wtkg", "biomarker", "karnof",
                       "biomarker_hi", "sex", "site"),
  outcome.name     = "y",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "continuous",
  effect_measure   = "MD",
  # MD is additive -- hr.threshold is an outcome-scale increment,
  # hr.consistency is the null-effect baseline (0 instead of 1).
  hr.threshold             = 0.5,
  hr.consistency           = 0.0,
  pconsistency.threshold   = 0.80,
  sg_focus                 = "maxSG",
  max_subgroups_search     = 10L,
  use_grf                  = TRUE,
  use_lasso                = FALSE,
  is.RCT                   = TRUE,
  vi.grf.min               = -0.2,
  cont.cutoff              = 4L,
  cut_type                 = "default",
  seedit                   = 8316951L,
  maxk                     = 2L,
  n.min                    = 60L,
  d0.min                   = 10L,
  d1.min                   = 10L,
  fs.splits                = 300L,
  details                  = FALSE,
  quiet                    = TRUE,
  plot.sg                  = FALSE,
  plot.grf                 = FALSE,
  parallel_args            = list(plan         = "multisession",
                                  workers      = n_workers,
                                  show_message = FALSE)
)

fs_full <- do.call(forestsearch,
                   c(list(df.analysis = df_syn), fs_args))
plan("sequential")

# Hard check: if the primary call collapsed to a minimal
# early-return, fail LOUDLY here rather than letting downstream
# chunks error with cryptic messages.
if (is.null(fs_full$args_call_all) || is.null(fs_full$df.est)) {
  stop(sprintf(paste0(
    "Primary ForestSearch call collapsed to an early-return path ",
    "(args_call_all=%s, df.est=%s, sg.harm=%s).  Check for upstream ",
    "FSdata / find.grps / consistency failures."),
    !is.null(fs_full$args_call_all),
    !is.null(fs_full$df.est),
    !is.null(fs_full$sg.harm)))
}

cat(sprintf("Identified subgroup: %s\n",
            paste(fs_full$sg.harm, collapse = " & ")))
cat(sprintf("outcome_type on return: %s\n", fs_full$outcome_type))
cat(sprintf("effect_measure on return: %s\n", fs_full$effect_measure))
cat(sprintf("GRF cuts: %s\n",
            paste(fs_full$grf_cuts, collapse = " | ")))

.tc("Primary FS", "args_call_all present on return",
    !is.null(fs_full$args_call_all))
.tc("Primary FS", "df.est present on return",
    !is.null(fs_full$df.est))
.tc("Primary FS", "outcome_type == 'continuous' preserved on return",
    identical(fs_full$outcome_type, "continuous"))
.tc("Primary FS", "effect_measure == 'MD' preserved on return",
    identical(fs_full$effect_measure, "MD"))
.tc("Primary FS", "a subgroup was identified",
    !is.null(fs_full$sg.harm))

if (!is.null(fs_full$sg.harm)) {
  sg_str <- paste(fs_full$sg.harm, collapse = " & ")
  # Identified subgroup should involve at least one of the true
  # signal variables (age or biomarker_hi)
  .tc("Primary FS",
      "identified subgroup involves true-signal variable",
      grepl("age|biomarker_hi", sg_str),
      detail = sprintf("sg = '%s'", sg_str))
}

# For MD / identity link, the top-ranked result row's 'hr' column
# should hold the MD directly (no log transform applied).  A
# plausible value for DGM MD_harm = 3.0 is in the range [1, 6].  If
# someone had applied exp() in error, the value would be around
# exp(3) ~ 20.
if (!is.null(fs_full$grp.consistency$out_sg$result) &&
    nrow(fs_full$grp.consistency$out_sg$result) > 0L) {
  stored_hr <- as.numeric(
    fs_full$grp.consistency$out_sg$result[1L, "hr"]
  )
  .tc("Primary FS",
      "stored 'hr' is on identity scale (MD directly, not exp())",
      stored_hr > 0.5 && stored_hr < 10,
      detail = sprintf("stored='hr'=%.3f (DGM MD = %.1f; exp() would give %.2f)",
                       stored_hr, MD_harm, exp(stored_hr)))
}
```


# CV pipeline on GLM data {#sec-cv}

## Run CV

```{r cv-run}
#| message: false
.run_cv <- function() {
  forestsearch_tenfold(
    fs.est        = fs_full,
    sims          = Ksims_cv,
    Kfolds        = Kfolds_cv,
    details       = FALSE,
    parallel_args = list(plan    = "multisession",
                         workers = n_workers,
                         show_message = FALSE)
  )
}

message(sprintf("Running CV (sims=%d, Kfolds=%d) ...",
                Ksims_cv, Kfolds_cv))
fs_cv <- .run_cv()
plan("sequential")

cat(sprintf("CV complete: %d sims x %d folds = %d rows\n",
            fs_cv$sims, fs_cv$Kfolds, nrow(fs_cv$fold_summary)))
cat(sprintf("fold_summary columns: %s\n",
            paste(names(fs_cv$fold_summary), collapse = ", ")))
```

## Column-level checks on GLM data

```{r cv-columns}
fs <- fs_cv$fold_summary

.tc("CV columns", "fold_summary$grf_cuts present",
    "grf_cuts" %in% names(fs))
.tc("CV columns", "fold_summary$pconsistency present",
    "pconsistency" %in% names(fs))
.tc("CV columns", "fold_summary$training_fs_hr present",
    "training_fs_hr" %in% names(fs))
.tc("CV columns", "fold_summary$n_candidates_evaluated present",
    "n_candidates_evaluated" %in% names(fs))
```

## KEY GLM check: `training_fs_hr` is on identity (MD) scale

```{r cv-glm-scale}
# Under our DGM (MD_harm = 3.0), identifying folds should yield
# training_fs_hr values in a sensible additive range around MD_harm
# (typically 1 to 6 on the MD scale).  The failure mode here is the
# INVERSE of the log-link tests in the binary/count compat docs: if
# exp() accidentally leaks into the identity-scale path, values
# would cluster near exp(MD) ~ 20, not near MD itself.
hr_vals  <- fs$training_fs_hr
hr_nonna <- hr_vals[!is.na(hr_vals)]

if (length(hr_nonna) > 0L) {
  med_hr <- median(hr_nonna)
  min_hr <- min(hr_nonna)
  max_hr <- max(hr_nonna)

  cat(sprintf("training_fs_hr across %d identifying folds:\n",
              length(hr_nonna)))
  cat(sprintf("  median=%.3f, min=%.3f, max=%.3f\n",
              med_hr, min_hr, max_hr))
  cat(sprintf("  (if exp() had leaked in, median would be ~ %.2f)\n",
              exp(med_hr)))

  # Identity-scale values for a harm search should be positive and
  # within a plausible additive band around MD_harm.  If values were
  # on log scale (i.e. pre-exp()), they'd be near log(3.0) ~ 1.1 which
  # still passes > 0 but fails the "within 3x MD_harm" check.  If
  # exp() leaked in, values would be near exp(3.0) ~ 20 and fail the
  # upper bound.
  .tc("CV GLM scale",
      "training_fs_hr values > 0 (harm direction preserved)",
      med_hr > 0,
      detail = sprintf("median = %.3f", med_hr))

  # Given hr.threshold = 0.5 was enforced during search, identifying
  # folds must have training_fs_hr >= 0.5 on identity scale.
  .tc("CV GLM scale",
      sprintf("training_fs_hr >= hr.threshold (%.2f) for all id-ing folds",
              fs_args$hr.threshold),
      all(hr_nonna >= fs_args$hr.threshold),
      detail = sprintf("min = %.3f; threshold = %.2f",
                       min_hr, fs_args$hr.threshold))

  # Upper-bound check: exp(MD_harm) ~ 20 would cluster well above 10,
  # so a cap at 10 catches the exp()-leak failure mode while still
  # permitting legitimate tail folds with MD estimates up to ~3x DGM.
  .tc("CV GLM scale",
      "training_fs_hr values are on identity scale (not exp-transformed)",
      max_hr < 10,
      detail = sprintf("max = %.3f (exp(MD_harm) would be %.2f)",
                       max_hr, exp(MD_harm)))

  .tc("CV GLM scale", "training_fs_hr values all finite",
      all(is.finite(hr_nonna)),
      detail = sprintf("finite=%d/%d",
                       sum(is.finite(hr_nonna)), length(hr_nonna)))
}
```

## Pcons and candidate-count column checks

```{r cv-pconsistency}
pv          <- fs$pconsistency
identifying <- !is.na(fs$sg1)
.tc("CV Pcons", "pconsistency NA iff any_found == 0",
    all(is.na(pv) == (fs$any_found == 0L)))
if (sum(identifying) > 0L) {
  .tc("CV Pcons", "non-NA pconsistency in [0, 1]",
      all(pv[identifying] >= 0 & pv[identifying] <= 1))
  .tc("CV Pcons",
      sprintf("pconsistency >= threshold (%.2f) for identifying folds",
              fs_args$pconsistency.threshold),
      all(pv[identifying] >= fs_args$pconsistency.threshold))
}
```

```{r cv-candidates}
nc <- fs$n_candidates_evaluated
.tc("CV candidates", "n_candidates_evaluated is integer",
    is.integer(nc))
.tc("CV candidates", "n_candidates_evaluated non-negative",
    all(is.na(nc) | nc >= 0L))
id_idx <- which(fs$any_found == 1L)
if (length(id_idx) > 0L) {
  .tc("CV candidates",
      "identifying folds evaluated >= 1 candidate",
      all(!is.na(nc[id_idx]) & nc[id_idx] >= 1L))
}
```

## `summarize_cv_results()` on GLM data

```{r cv-summary}
cv_diag <- summarize_cv_results(
  cv_output         = fs_cv,
  original_sg       = fs_full$sg.harm,
  original_grf_cuts = fs_full$grf_cuts,
  top_n             = 10L
)

.tc("CV summary", "identification_summary slot populated",
    !is.null(cv_diag$identification_summary))
.tc("CV summary", "grf_cut_summary slot populated",
    !is.null(cv_diag$grf_cut_summary))
.tc("CV summary", "no_subgroup_decomposition slot populated",
    !is.null(cv_diag$no_subgroup_decomposition))
.tc("CV summary", "pconsistency_distribution slot populated",
    !is.null(cv_diag$pconsistency_distribution))
.tc("CV summary", "fold_numeric_summary slot populated",
    !is.null(cv_diag$fold_numeric_summary))
.tc("CV summary", "original_agreement slot populated",
    !is.null(cv_diag$original_agreement))
.tc("CV summary", "has_pconsistency metadata TRUE",
    isTRUE(cv_diag$has_pconsistency))
```

### Rendered `identification_summary`

```{r cv-ident-display}
cv_diag$identification_summary
```

### Rendered `grf_cut_summary`

```{r cv-grf-display}
cv_diag$grf_cut_summary
```

### Rendered `fold_numeric_summary`

Critical GLM check: the "Training-fold subgroup effect" row's median
should be on identity (MD) scale — a value in the plausible additive
range around `MD_harm`, *not* an exponentially transformed value.

```{r cv-fnum-display}
cv_diag$fold_numeric_summary
```

```{r cv-fnum-values}
fns <- cv_diag$data$fold_numeric_summary
eff_row <- fns[
  fns$Metric == "Training-fold subgroup effect (identifying folds)",
  , drop = FALSE
]
if (nrow(eff_row) == 1L) {
  .tc("CV summary",
      "fold_numeric_summary effect-row label is 'subgroup effect' (outcome-agnostic)",
      grepl("subgroup effect", eff_row$Metric, fixed = TRUE))
  # Context string may say "natural scale" or "identity scale" depending
  # on how the formatter describes the continuous / MD path.  Accept
  # either; failure here would mean the context column is missing any
  # scale annotation at all.
  .tc("CV summary",
      "fold_numeric_summary effect-row context mentions a scale",
      grepl("natural scale|identity scale|MD|mean difference",
            eff_row$Context, ignore.case = TRUE),
      detail = sprintf("Context = '%s'", eff_row$Context))
  # For MD harm, median should be positive on identity scale.  If
  # exp() had leaked in, it would be near exp(MD_harm) ~ 20 and this
  # loose bound wouldn't catch it; the upper-bound check in
  # cv-glm-scale is where that failure is caught.
  .tc("CV summary",
      "fold_numeric_summary effect-row median > 0 (identity-scale harm)",
      !is.na(eff_row$median) && eff_row$median > 0,
      detail = sprintf("median = %.3f", eff_row$median))
}
```


# Bootstrap pipeline on GLM data {#sec-boot}

## Run bootstrap

```{r boot-run}
#| message: false
.run_bootstrap <- function() {
  forestsearch_bootstrap_dofuture(
    fs.est        = fs_full,
    nb_boots      = nb_boots,
    parallel_args = list(plan    = "multisession",
                         workers = n_workers,
                         show_message = FALSE)
  )
}

message(sprintf("Running bootstrap (nb_boots=%d) ...", nb_boots))
fs_bc <- .run_bootstrap()
plan("sequential")

cat(sprintf("Bootstrap complete: %d iterations, columns:\n",
            nrow(fs_bc$results)))
cat(paste(names(fs_bc$results), collapse = ", "), "\n")
```

## Column and slot checks

```{r boot-column}
.tc("Bootstrap column", "results$grf_cuts_b present",
    "grf_cuts_b" %in% names(fs_bc$results))
.tc("Bootstrap column", "grf_cuts_b is character",
    is.character(fs_bc$results$grf_cuts_b))
.tc("Bootstrap column", "nrow(results) == nb_boots",
    nrow(fs_bc$results) == nb_boots)
```

```{r boot-slot}
subg_summary <- summarize_bootstrap_subgroups(
  results     = fs_bc$results,
  nb_boots    = nb_boots,
  original_sg = fs_full$sg.harm,
  maxk        = 2L
)

.tc("Bootstrap slot", "grf_cut_freq slot populated",
    !is.null(subg_summary$grf_cut_freq))

gcf <- subg_summary$grf_cut_freq
if (!is.null(gcf)) {
  .tc("Bootstrap slot", "grf_cut_freq N sums to nb_boots",
      sum(gcf$N) == nb_boots,
      detail = sprintf("sum=%d exp=%d", sum(gcf$N), nb_boots))
  .tc("Bootstrap slot", "grf_cut_freq sorted by descending N",
      all(diff(gcf$N) <= 0))
}
```

### Rendered GRF-cut frequency

```{r boot-display}
if (!is.null(subg_summary$grf_cut_freq)) {
  subg_summary$grf_cut_freq |>
    utils::head(10) |>
    gt::gt() |>
    gt::tab_header(
      title = gt::md("**GRF cuts across bootstrap iterations (GLM continuous outcome)**"),
      subtitle = sprintf("Top 10 of %d unique cuts across %d iterations",
                         nrow(subg_summary$grf_cut_freq), nb_boots)
    ) |>
    gt::cols_label(
      Rank = "Rank", grf_cut = "GRF cuts",
      N = "N",       Percent = "% of bootstraps"
    ) |>
    gt::fmt_number(columns = "Percent", decimals = 2) |>
    gt::tab_options(table.font.size = gt::px(13))
}
```


# Cross-check: `training_fs_hr` vs. independent GLM refit {#sec-crosscheck}

Confirm numerical correctness of the identity-scale path by
re-fitting `lm(y ~ treat)` directly on an identifying fold's subgroup
(using full data as a plausibility proxy, since exact per-fold splits
aren't preserved).  Because MD is additive, the agreement check
uses an additive tolerance rather than the multiplicative 3x ratio
used for log-link compatibility tests.

```{r crosscheck}
id_folds <- fs_cv$fold_summary[
  !is.na(fs_cv$fold_summary$training_fs_hr), , drop = FALSE
]

if (nrow(id_folds) > 0L) {
  # Pick the median fold by training_fs_hr
  pick <- id_folds[order(abs(id_folds$training_fs_hr -
                              median(id_folds$training_fs_hr)))[1], ]
  captured_hr <- pick$training_fs_hr

  cat(sprintf("Sampled identifying fold: sim=%d, fold=%d\n",
              pick$sim, pick$fold))
  cat(sprintf("Subgroup definition: %s%s\n",
              pick$sg1,
              if (!is.na(pick$sg2)) paste0(" & ", pick$sg2) else ""))
  cat(sprintf("Captured training_fs_hr: %.3f\n", captured_hr))

  # Evaluate the subgroup expression(s) to get a membership flag.
  in_sg <- tryCatch({
    sg1_flag <- eval(parse(text = pick$sg1), envir = df_syn)
    if (!is.na(pick$sg2)) {
      sg2_flag <- eval(parse(text = pick$sg2), envir = df_syn)
      sg1_flag & sg2_flag
    } else {
      sg1_flag
    }
  }, error = function(e) NULL)

  if (!is.null(in_sg) && sum(in_sg) > 10L) {
    df_sg <- df_syn[in_sg, ]
    m <- tryCatch(
      lm(y ~ treat, data = df_sg),
      error = function(e) NULL
    )
    if (!is.null(m)) {
      # For Gaussian / identity link, the treat coefficient IS the
      # MD directly -- no exp() transform.
      md_refit <- as.numeric(coef(m)["treat"])
      cat(sprintf("Re-fit MD on full-data subgroup: %.3f\n", md_refit))

      .tc("Cross-check",
          "captured training_fs_hr and re-fit MD both > 0",
          captured_hr > 0 && md_refit > 0,
          detail = sprintf("captured = %.3f, refit = %.3f",
                           captured_hr, md_refit))

      # Additive tolerance: captured MD and refit MD should agree
      # within +/- 1.0 outcome units.  This is a loose bound because
      # the fold and full-data subgroups differ; it catches
      # order-of-magnitude errors (e.g. exp() leaking in) without
      # being tripped by ordinary sampling variation.
      .tc("Cross-check",
          "captured and refit MD agree within +/- 1.0 units",
          abs(captured_hr - md_refit) < 1.0,
          detail = sprintf("diff = %.3f",
                           captured_hr - md_refit))
    }
  }
}
```


# v0.2.0 bootstrap return-value contract {#sec-v020}

Validates the v0.2.0 additions to the bootstrap return object that
downstream consumers (`print.fs_bootstrap`, CV-style fold
identification diagnostics, cross-API count comparisons) depend on.
Orthogonal to Phase C; included here so each outcome-type
compatibility doc exercises the contract independently.

```{r v020-helper}
`%||%` <- function(a, b) if (is.null(a)) b else a
```

## S3 class and core schema

```{r v020-class}
.tc("v0.2.0", "fs_bc inherits 'fs_bootstrap' class",
    inherits(fs_bc, "fs_bootstrap"),
    detail = sprintf("classes = %s",
                     paste(class(fs_bc), collapse = "/")))

.tc("v0.2.0", "results has any_found column",
    "any_found" %in% names(fs_bc$results))

.tc("v0.2.0", "any_found is integer",
    is.integer(fs_bc$results$any_found),
    detail = sprintf("class = %s",
                     class(fs_bc$results$any_found)[1]))

.tc("v0.2.0", "any_found values are all 0L or 1L",
    all(fs_bc$results$any_found %in% c(0L, 1L)),
    detail = sprintf("distinct = %s",
                     paste(sort(unique(fs_bc$results$any_found)),
                           collapse = ",")))
```

## `any_found` agrees with `!is.na(Pcons)`

```{r v020-equiv}
by_pcons    <- sum(!is.na(fs_bc$results$Pcons))
by_anyfound <- sum(fs_bc$results$any_found == 1L)

.tc("v0.2.0",
    "any_found count matches !is.na(Pcons) count",
    by_pcons == by_anyfound,
    detail = sprintf("Pcons=%d, any_found=%d", by_pcons, by_anyfound))
```

## Metadata fields populated on the continuous path

```{r v020-metadata}
.tc("v0.2.0", "nb_boots metadata non-NULL",
    !is.null(fs_bc$nb_boots),
    detail = sprintf("nb_boots = %s", fs_bc$nb_boots %||% "NULL"))

.tc("v0.2.0", "original_sg slot exists on return",
    "original_sg" %in% names(fs_bc))

.tc("v0.2.0", "outcome_type metadata non-NULL",
    !is.null(fs_bc$outcome_type),
    detail = sprintf("outcome_type = %s",
                     fs_bc$outcome_type %||% "NULL"))

.tc("v0.2.0", "effect_measure metadata non-NULL",
    !is.null(fs_bc$effect_measure),
    detail = sprintf("effect_measure = %s",
                     fs_bc$effect_measure %||% "NULL"))

.tc("v0.2.0", "est.scale metadata non-NULL",
    !is.null(fs_bc$est.scale),
    detail = sprintf("est.scale = %s",
                     fs_bc$est.scale %||% "NULL"))
```

## Metadata values match the continuous call

```{r v020-metadata-values}
.tc("v0.2.0", "nb_boots echoes the call argument",
    identical(fs_bc$nb_boots, nb_boots),
    detail = sprintf("returned=%s, expected=%d",
                     fs_bc$nb_boots %||% "NULL", nb_boots))

.tc("v0.2.0", "outcome_type == 'continuous'",
    identical(fs_bc$outcome_type, "continuous"),
    detail = sprintf("got %s", fs_bc$outcome_type %||% "NULL"))

.tc("v0.2.0", "effect_measure == 'MD'",
    identical(fs_bc$effect_measure, "MD"),
    detail = sprintf("got %s", fs_bc$effect_measure %||% "NULL"))
```

## `print.fs_bootstrap` dispatches without error

```{r v020-print-default}
print_err <- tryCatch(
  { capture.output(print(fs_bc)); NULL },
  error = function(e) conditionMessage(e)
)
.tc("v0.2.0",
    "print(fs_bc) dispatches without error",
    is.null(print_err),
    detail = if (is.null(print_err)) "OK" else print_err)
```

## `top_n` parameter honoured

```{r v020-print-topn}
topn_err <- tryCatch(
  { capture.output(print(fs_bc, top_n = 5L)); NULL },
  error = function(e) conditionMessage(e)
)
.tc("v0.2.0",
    "print(fs_bc, top_n = 5L) dispatches without error",
    is.null(topn_err),
    detail = if (is.null(topn_err)) "OK" else topn_err)
```


# Test summary {#sec-summary}

```{r summary-table}
n_total <- nrow(test_results)
n_pass  <- sum(test_results$Status == "PASS")
n_fail  <- n_total - n_pass

gt::gt(test_results,
       groupname_col = "Section",
       row_group_as_column = TRUE) |>
  gt::tab_header(
    title    = gt::md(sprintf(
      "**GLM compatibility (continuous / MD) test results: %d / %d passing**",
      n_pass, n_total)),
    subtitle = sprintf(
      "Continuous outcome (MD, identity scale), N=%d, CV sims=%d, nb_boots=%d",
      N, Ksims_cv, nb_boots)
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#d4edda"),
    locations = gt::cells_body(
      columns = "Status", rows = Status == "PASS"
    )
  ) |>
  gt::tab_style(
    style     = list(
      gt::cell_fill(color = "#f8d7da"),
      gt::cell_text(weight = "bold")
    ),
    locations = gt::cells_body(
      columns = "Status", rows = Status == "FAIL"
    )
  ) |>
  gt::cols_label(
    Test   = "Test case",
    Status = "Status",
    Detail = "Detail"
  ) |>
  gt::tab_options(table.font.size = gt::px(12))

if (n_fail == 0L) {
  message(sprintf("ALL TESTS PASSED: %d / %d", n_pass, n_total))
} else {
  message(sprintf("TESTS FAILED: %d / %d failing", n_fail, n_total))
  print(test_results[test_results$Status == "FAIL",
                     c("Section", "Test", "Detail")])
}
```


# Session info {#sec-session}

```{r session}
sessionInfo()
```
