---
title: "Living test: Phase A/B/C GLM compatibility (count / IRR)"
subtitle: "Synthetic count outcome + IRR effect measure + person-time offset"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-tools: true
    code-fold: true
    theme: cosmo
    fontsize: 11pt
    number-sections: true
params:
  quick_mode: true       # TRUE: smaller grid (fast). FALSE: larger.
---

# Purpose {#sec-purpose}

Validates that the Phase A/B/C diagnostic features
(`fold_summary$grf_cuts`, `fold_summary$pconsistency`,
`fold_summary$training_fs_hr`, `fold_summary$n_candidates_evaluated`,
`summarize_cv_results()` slots, and the bootstrap `grf_cut_freq`
slot) work correctly when the outcome is a **count** (Poisson GLM
with log link and optional person-time offset) rather than survival
or binary.

Parallel in intent to `test_glm_compatibility.qmd`, but specialised
to the count / IRR path.  The count-specific concerns are:

1. `training_fs_hr` on log-link measures (IRR): `forestsearch()`
   stores internal effect values on the log scale, but Phase B's
   capture must apply `exp()` so the column is on the natural
   (multiplicative) IRR scale regardless of outcome type.  Values
   near `log(IRR_harm)` in the fold-summary would indicate a missing
   transform.
2. The `offset.name` pathway: forestsearch() must propagate the
   person-time offset through all internal GLM fits; failure mode is
   an effect estimate that treats the count as raw and drops to
   near-null even when the rate ratio is large.

The synthetic DGM plants a harm signal driven by
`age <= 50 & biomarker_hi == 1`, with true IRR ≈ 2.0 inside the
subgroup and ≈ 1.0 outside, with follow-up time (`ftime`) supplied
as an offset.  Covariates mix continuous (GRF cuts discoverable) and
binary factor (passed through without cut attempt), with
`sg_focus = "maxSG"` to recover the largest qualifying subgroup.

Coverage:

1. **Synthetic data generation** — harm subgroup driven by
   `age <= 50 & biomarker_hi == 1`, true IRR ≈ 2.0, mix of continuous
   and factor covariates, follow-up time as offset.
2. **Primary ForestSearch call** — confirms the count GLM routing
   succeeds, the offset is honoured, outcome-type / effect-measure
   metadata is preserved through the return.
3. **CV pipeline (Phase A/B)** — confirms every new `fold_summary`
   column populates on count data, confirms `summarize_cv_results()`
   produces each diagnostic slot, and confirms `training_fs_hr`
   values are on IRR scale (i.e., > 1 for harm, not log-IRR values
   centred near 0.7).
4. **Bootstrap pipeline (Phase C)** — confirms `results$grf_cuts_b`
   populates and `summarize_bootstrap_subgroups()` surfaces
   `grf_cut_freq`.
5. **v0.2.0 bootstrap return-value contract** — `any_found` column,
   `fs_bootstrap` S3 class tag, metadata (nb_boots, outcome_type,
   effect_measure, est.scale), and `print.fs_bootstrap` dispatch
   work on the count path.
6. **Cross-check** — captured `training_fs_hr` on an identifying
   fold agrees within order-of-magnitude with an independent
   Poisson `glm()` refit (with offset) on the same subgroup.

A cumulative pass/fail table renders in @sec-summary.


# Setup {#sec-setup}

```{r load-packages}
#| message: false
#| warning: false
library(forestsearch)
library(data.table)
library(gt)
library(future)
library(foreach)
library(doFuture)
```

```{r config}
if (isTRUE(params$quick_mode)) {
  N         <- 800L
  Ksims_cv  <- 5L
  Kfolds_cv <- 10L
  nb_boots  <- 20L
  mode_tag  <- "quick"
} else {
  N         <- 1500L
  Ksims_cv  <- 20L
  Kfolds_cv <- 10L
  nb_boots  <- 100L
  mode_tag  <- "full"
}
cat(sprintf("Mode: %s (N=%d, Ksims=%d, nb_boots=%d)\n",
            mode_tag, N, Ksims_cv, nb_boots))

n_workers <- min(4L, floor(0.95 * max(1L, parallel::detectCores() - 1L)))
```

```{r test-harness}
test_results <- data.frame(
  Section = character(0),
  Test    = character(0),
  Status  = character(0),
  Detail  = character(0),
  stringsAsFactors = FALSE
)

.tc <- function(section, test, pass, detail = "") {
  status <- if (isTRUE(pass)) "PASS" else "FAIL"
  test_results <<- rbind(test_results, data.frame(
    Section = section, Test = test, Status = status,
    Detail  = substr(detail, 1L, 140L),
    stringsAsFactors = FALSE
  ))
  invisible(pass)
}
```


# Synthetic count-outcome DGM {#sec-dgm}

Mirrors the binary compatibility pattern but with a count response:
event counts per subject are Poisson distributed with rate governed
by treatment and subgroup membership, scaled by individual follow-up
time (`ftime`).  The natural effect measure is the incidence rate
ratio (IRR) reached via a log link.  Harm signal is jointly defined
by a continuous cut (`age <= 50`) and a binary factor
(`biomarker_hi == 1`).

```{r generate-data}
set.seed(20260421)

IRR_harm    <- 2.0     # true IRR in harm subgroup (treatment worsens rate)
IRR_benefit <- 1.0     # null rate ratio outside harm subgroup

df_syn <- data.frame(
  id           = seq_len(N),
  treat        = rbinom(N, 1L, 0.5),
  age          = round(rnorm(N, mean = 55, sd = 12)),
  wtkg         = round(rnorm(N, mean = 75, sd = 12), 1),
  biomarker    = rnorm(N),
  karnof       = sample(70:100, N, replace = TRUE),
  biomarker_hi = factor(rbinom(N, 1L, 0.5), levels = c(0L, 1L)),
  sex          = factor(rbinom(N, 1L, 0.5), levels = c(0L, 1L)),
  site         = factor(sample(0:1, N, replace = TRUE),
                        levels = c(0L, 1L)),
  ftime        = round(runif(N, 1, 5), 2)   # person-time (arbitrary units)
)

# True harm subgroup: age <= 50 AND biomarker_hi == 1
true_harm    <- df_syn$age <= 50 & df_syn$biomarker_hi == "1"
treat_effect <- ifelse(true_harm, log(IRR_harm), log(IRR_benefit))

# Baseline rate = 0.5 events per unit ftime; apply treatment-by-subgroup
# interaction on the log scale, then scale by ftime to get expected count.
base_rate <- 0.5
rate      <- base_rate * exp(df_syn$treat * treat_effect)
df_syn$y  <- rpois(N, lambda = rate * df_syn$ftime)

cat(sprintf("N = %d, mean count = %.2f (median %d, max %d)\n",
            nrow(df_syn), mean(df_syn$y),
            median(df_syn$y), max(df_syn$y)))
cat(sprintf("True harm prevalence: %.2f\n", mean(true_harm)))

rate_0 <- sum(df_syn$y[df_syn$treat == 0L]) /
          sum(df_syn$ftime[df_syn$treat == 0L])
rate_1 <- sum(df_syn$y[df_syn$treat == 1L]) /
          sum(df_syn$ftime[df_syn$treat == 1L])
cat(sprintf("Event rate per unit ftime | treat = 0: %.3f\n", rate_0))
cat(sprintf("Event rate per unit ftime | treat = 1: %.3f\n", rate_1))
cat(sprintf("Overall empirical IRR: %.2f\n", rate_1 / rate_0))
```

Empirical check that the harm subgroup is recoverable:

```{r dgm-sanity}
har <- df_syn[true_harm, ]
r0 <- sum(har$y[har$treat == 0L]) / sum(har$ftime[har$treat == 0L])
r1 <- sum(har$y[har$treat == 1L]) / sum(har$ftime[har$treat == 1L])
irr_harm_emp <- r1 / r0
cat(sprintf("Empirical IRR in true harm subgroup: %.2f (DGM = %.1f)\n",
            irr_harm_emp, IRR_harm))

.tc("DGM", "empirical harm IRR > 1.3",
    irr_harm_emp > 1.3,
    detail = sprintf("empirical IRR = %.2f", irr_harm_emp))

.tc("DGM", "harm subgroup prevalence between 15% and 40%",
    mean(true_harm) > 0.15 && mean(true_harm) < 0.40,
    detail = sprintf("prev = %.2f", mean(true_harm)))
```


# Primary ForestSearch call {#sec-fs-full}

Configuration exercises the count / IRR pathway with a person-time
offset:

- `outcome_type = "count"`, `effect_measure = "IRR"`,
  `offset.name = "ftime"`
- `sg_focus = "maxSG"` (largest qualifying subgroup)
- `is.RCT = TRUE`, `vi.grf.min = -0.2`, `cont.cutoff = 4L`
- Mixed covariates: continuous (cuts discovered by GRF) + factor
  (passed through without cut attempt)
- `use_grf = TRUE`, `use_lasso = FALSE`

```{r fs-full}
#| message: false
fs_args <- list(
  confounders.name = c("age", "wtkg", "biomarker", "karnof",
                       "biomarker_hi", "sex", "site"),
  outcome.name     = "y",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "count",
  effect_measure   = "IRR",
  offset.name      = "ftime",
  hr.threshold             = 1.25,
  hr.consistency           = 1.00,
  pconsistency.threshold   = 0.80,
  sg_focus                 = "maxSG",
  max_subgroups_search     = 10L,
  use_grf                  = TRUE,
  use_lasso                = FALSE,
  is.RCT                   = TRUE,
  vi.grf.min               = -0.2,
  cont.cutoff              = 4L,
  cut_type                 = "default",
  seedit                   = 8316951L,
  maxk                     = 2L,
  n.min                    = 60L,
  d0.min                   = 10L,
  d1.min                   = 10L,
  fs.splits                = 300L,
  details                  = FALSE,
  quiet                    = TRUE,
  plot.sg                  = FALSE,
  plot.grf                 = FALSE,
  parallel_args            = list(plan         = "multisession",
                                  workers      = n_workers,
                                  show_message = FALSE)
)

fs_full <- do.call(forestsearch,
                   c(list(df.analysis = df_syn), fs_args))
plan("sequential")

# Hard check: if the primary call collapsed to a minimal
# early-return, fail LOUDLY here rather than letting downstream
# chunks error with cryptic messages.
if (is.null(fs_full$args_call_all) || is.null(fs_full$df.est)) {
  stop(sprintf(paste0(
    "Primary ForestSearch call collapsed to an early-return path ",
    "(args_call_all=%s, df.est=%s, sg.harm=%s).  Check for upstream ",
    "FSdata / find.grps / consistency failures."),
    !is.null(fs_full$args_call_all),
    !is.null(fs_full$df.est),
    !is.null(fs_full$sg.harm)))
}

cat(sprintf("Identified subgroup: %s\n",
            paste(fs_full$sg.harm, collapse = " & ")))
cat(sprintf("outcome_type on return: %s\n", fs_full$outcome_type))
cat(sprintf("effect_measure on return: %s\n", fs_full$effect_measure))
cat(sprintf("GRF cuts: %s\n",
            paste(fs_full$grf_cuts, collapse = " | ")))

.tc("Primary FS", "args_call_all present on return",
    !is.null(fs_full$args_call_all))
.tc("Primary FS", "df.est present on return",
    !is.null(fs_full$df.est))
.tc("Primary FS", "outcome_type == 'count' preserved on return",
    identical(fs_full$outcome_type, "count"))
.tc("Primary FS", "effect_measure == 'IRR' preserved on return",
    identical(fs_full$effect_measure, "IRR"))
.tc("Primary FS", "a subgroup was identified",
    !is.null(fs_full$sg.harm))

if (!is.null(fs_full$sg.harm)) {
  sg_str <- paste(fs_full$sg.harm, collapse = " & ")
  # Identified subgroup should involve at least one of the true
  # signal variables (age or biomarker_hi)
  .tc("Primary FS",
      "identified subgroup involves true-signal variable",
      grepl("age|biomarker_hi", sg_str),
      detail = sprintf("sg = '%s'", sg_str))
}

# The top-ranked result row's 'hr' column holds log-IRR for count+IRR.
# Exponentiating should give a value roughly matching IRR_harm (2.0).
if (!is.null(fs_full$grp.consistency$out_sg$result) &&
    nrow(fs_full$grp.consistency$out_sg$result) > 0L) {
  stored_hr <- as.numeric(
    fs_full$grp.consistency$out_sg$result[1L, "hr"]
  )
  irr_nat <- exp(stored_hr)
  .tc("Primary FS",
      "stored 'hr' is on log scale (exp() yields plausible IRR)",
      irr_nat > 1.1 && irr_nat < 8,
      detail = sprintf("stored='hr'=%.3f, exp()=%.2f (DGM IRR = %.1f)",
                       stored_hr, irr_nat, IRR_harm))
}
```


# CV pipeline on GLM data {#sec-cv}

## Run CV

```{r cv-run}
#| message: false
.run_cv <- function() {
  forestsearch_tenfold(
    fs.est        = fs_full,
    sims          = Ksims_cv,
    Kfolds        = Kfolds_cv,
    details       = FALSE,
    parallel_args = list(plan    = "multisession",
                         workers = n_workers,
                         show_message = FALSE)
  )
}

message(sprintf("Running CV (sims=%d, Kfolds=%d) ...",
                Ksims_cv, Kfolds_cv))
fs_cv <- .run_cv()
plan("sequential")

cat(sprintf("CV complete: %d sims x %d folds = %d rows\n",
            fs_cv$sims, fs_cv$Kfolds, nrow(fs_cv$fold_summary)))
cat(sprintf("fold_summary columns: %s\n",
            paste(names(fs_cv$fold_summary), collapse = ", ")))
```

## Column-level checks on GLM data

```{r cv-columns}
fs <- fs_cv$fold_summary

.tc("CV columns", "fold_summary$grf_cuts present",
    "grf_cuts" %in% names(fs))
.tc("CV columns", "fold_summary$pconsistency present",
    "pconsistency" %in% names(fs))
.tc("CV columns", "fold_summary$training_fs_hr present",
    "training_fs_hr" %in% names(fs))
.tc("CV columns", "fold_summary$n_candidates_evaluated present",
    "n_candidates_evaluated" %in% names(fs))
```

## KEY GLM check: `training_fs_hr` is on natural scale

```{r cv-glm-scale}
# Under our DGM (IRR_harm = 2.0), identifying folds should yield
# training_fs_hr values > 1 (typically 1.3 to 5 range on the IRR
# scale).  If Phase B's exp() transform is missing, the values would
# cluster near log(2.0) ~ 0.69 instead.
hr_vals  <- fs$training_fs_hr
hr_nonna <- hr_vals[!is.na(hr_vals)]

if (length(hr_nonna) > 0L) {
  med_hr <- median(hr_nonna)
  min_hr <- min(hr_nonna)
  max_hr <- max(hr_nonna)

  cat(sprintf("training_fs_hr across %d identifying folds:\n",
              length(hr_nonna)))
  cat(sprintf("  median=%.3f, min=%.3f, max=%.3f\n",
              med_hr, min_hr, max_hr))
  cat(sprintf("  (on log scale, median would be %.3f)\n", log(med_hr)))

  # Natural-scale values should be > 1 for a harm search.  Log-scale
  # values would be near log(2) = 0.69 (below hr.threshold = 1.25).
  .tc("CV GLM scale",
      "training_fs_hr values > 1 (natural scale, not log)",
      med_hr > 1,
      detail = sprintf("median = %.3f (would be %.3f if log-scale)",
                       med_hr, log(med_hr)))

  # Given hr.threshold = 1.25 was enforced during search, identifying
  # folds must have training_fs_hr >= 1.25 on natural scale.
  .tc("CV GLM scale",
      sprintf("training_fs_hr >= hr.threshold (%.2f) for all id-ing folds",
              fs_args$hr.threshold),
      all(hr_nonna >= fs_args$hr.threshold),
      detail = sprintf("min = %.3f; threshold = %.2f",
                       min_hr, fs_args$hr.threshold))

  .tc("CV GLM scale", "training_fs_hr values finite and plausible",
      all(is.finite(hr_nonna)) && max_hr < 100,
      detail = sprintf("max = %.3f", max_hr))
}
```

## Pcons and candidate-count column checks

```{r cv-pconsistency}
pv          <- fs$pconsistency
identifying <- !is.na(fs$sg1)
.tc("CV Pcons", "pconsistency NA iff any_found == 0",
    all(is.na(pv) == (fs$any_found == 0L)))
if (sum(identifying) > 0L) {
  .tc("CV Pcons", "non-NA pconsistency in [0, 1]",
      all(pv[identifying] >= 0 & pv[identifying] <= 1))
  .tc("CV Pcons",
      sprintf("pconsistency >= threshold (%.2f) for identifying folds",
              fs_args$pconsistency.threshold),
      all(pv[identifying] >= fs_args$pconsistency.threshold))
}
```

```{r cv-candidates}
nc <- fs$n_candidates_evaluated
.tc("CV candidates", "n_candidates_evaluated is integer",
    is.integer(nc))
.tc("CV candidates", "n_candidates_evaluated non-negative",
    all(is.na(nc) | nc >= 0L))
id_idx <- which(fs$any_found == 1L)
if (length(id_idx) > 0L) {
  .tc("CV candidates",
      "identifying folds evaluated >= 1 candidate",
      all(!is.na(nc[id_idx]) & nc[id_idx] >= 1L))
}
```

## `summarize_cv_results()` on GLM data

```{r cv-summary}
cv_diag <- summarize_cv_results(
  cv_output         = fs_cv,
  original_sg       = fs_full$sg.harm,
  original_grf_cuts = fs_full$grf_cuts,
  top_n             = 10L
)

.tc("CV summary", "identification_summary slot populated",
    !is.null(cv_diag$identification_summary))
.tc("CV summary", "grf_cut_summary slot populated",
    !is.null(cv_diag$grf_cut_summary))
.tc("CV summary", "no_subgroup_decomposition slot populated",
    !is.null(cv_diag$no_subgroup_decomposition))
.tc("CV summary", "pconsistency_distribution slot populated",
    !is.null(cv_diag$pconsistency_distribution))
.tc("CV summary", "fold_numeric_summary slot populated",
    !is.null(cv_diag$fold_numeric_summary))
.tc("CV summary", "original_agreement slot populated",
    !is.null(cv_diag$original_agreement))
.tc("CV summary", "has_pconsistency metadata TRUE",
    isTRUE(cv_diag$has_pconsistency))
```

### Rendered `identification_summary`

```{r cv-ident-display}
cv_diag$identification_summary
```

### Rendered `grf_cut_summary`

```{r cv-grf-display}
cv_diag$grf_cut_summary
```

### Rendered `fold_numeric_summary`

Critical GLM check: the "Training-fold subgroup effect" row's median
should be on natural IRR scale (value > 1), not log-IRR scale (would
be near 0.5–1.0).

```{r cv-fnum-display}
cv_diag$fold_numeric_summary
```

```{r cv-fnum-values}
fns <- cv_diag$data$fold_numeric_summary
eff_row <- fns[
  fns$Metric == "Training-fold subgroup effect (identifying folds)",
  , drop = FALSE
]
if (nrow(eff_row) == 1L) {
  .tc("CV summary",
      "fold_numeric_summary effect-row label is 'subgroup effect' (outcome-agnostic)",
      grepl("subgroup effect", eff_row$Metric, fixed = TRUE))
  .tc("CV summary",
      "fold_numeric_summary effect-row context mentions 'natural scale'",
      grepl("natural scale", eff_row$Context, ignore.case = TRUE))
  .tc("CV summary",
      "fold_numeric_summary effect-row median > 1 (natural scale verified)",
      !is.na(eff_row$median) && eff_row$median > 1,
      detail = sprintf("median = %.3f", eff_row$median))
}
```


# Bootstrap pipeline on GLM data {#sec-boot}

## Run bootstrap

```{r boot-run}
#| message: false
.run_bootstrap <- function() {
  forestsearch_bootstrap_dofuture(
    fs.est        = fs_full,
    nb_boots      = nb_boots,
    parallel_args = list(plan    = "multisession",
                         workers = n_workers,
                         show_message = FALSE)
  )
}

message(sprintf("Running bootstrap (nb_boots=%d) ...", nb_boots))
fs_bc <- .run_bootstrap()
plan("sequential")

cat(sprintf("Bootstrap complete: %d iterations, columns:\n",
            nrow(fs_bc$results)))
cat(paste(names(fs_bc$results), collapse = ", "), "\n")
```

## Column and slot checks

```{r boot-column}
.tc("Bootstrap column", "results$grf_cuts_b present",
    "grf_cuts_b" %in% names(fs_bc$results))
.tc("Bootstrap column", "grf_cuts_b is character",
    is.character(fs_bc$results$grf_cuts_b))
.tc("Bootstrap column", "nrow(results) == nb_boots",
    nrow(fs_bc$results) == nb_boots)
```

```{r boot-slot}
subg_summary <- summarize_bootstrap_subgroups(
  results     = fs_bc$results,
  nb_boots    = nb_boots,
  original_sg = fs_full$sg.harm,
  maxk        = 2L
)

.tc("Bootstrap slot", "grf_cut_freq slot populated",
    !is.null(subg_summary$grf_cut_freq))

gcf <- subg_summary$grf_cut_freq
if (!is.null(gcf)) {
  .tc("Bootstrap slot", "grf_cut_freq N sums to nb_boots",
      sum(gcf$N) == nb_boots,
      detail = sprintf("sum=%d exp=%d", sum(gcf$N), nb_boots))
  .tc("Bootstrap slot", "grf_cut_freq sorted by descending N",
      all(diff(gcf$N) <= 0))
}
```

### Rendered GRF-cut frequency

```{r boot-display}
if (!is.null(subg_summary$grf_cut_freq)) {
  subg_summary$grf_cut_freq |>
    utils::head(10) |>
    gt::gt() |>
    gt::tab_header(
      title = gt::md("**GRF cuts across bootstrap iterations (GLM count outcome)**"),
      subtitle = sprintf("Top 10 of %d unique cuts across %d iterations",
                         nrow(subg_summary$grf_cut_freq), nb_boots)
    ) |>
    gt::cols_label(
      Rank = "Rank", grf_cut = "GRF cuts",
      N = "N",       Percent = "% of bootstraps"
    ) |>
    gt::fmt_number(columns = "Percent", decimals = 2) |>
    gt::tab_options(table.font.size = gt::px(13))
}
```


# Cross-check: `training_fs_hr` vs. independent GLM refit {#sec-crosscheck}

Confirm numerical correctness of Phase B's `exp()` transform by
re-fitting the Poisson GLM directly on an identifying fold's subgroup
(using full data as a plausibility proxy, since exact per-fold splits
aren't preserved).  The independent refit includes `log(ftime)` as an
offset so its coefficient is comparable to forestsearch's stored IRR.

```{r crosscheck}
id_folds <- fs_cv$fold_summary[
  !is.na(fs_cv$fold_summary$training_fs_hr), , drop = FALSE
]

if (nrow(id_folds) > 0L) {
  # Pick the median fold by training_fs_hr
  pick <- id_folds[order(abs(id_folds$training_fs_hr -
                              median(id_folds$training_fs_hr)))[1], ]
  captured_hr <- pick$training_fs_hr

  cat(sprintf("Sampled identifying fold: sim=%d, fold=%d\n",
              pick$sim, pick$fold))
  cat(sprintf("Subgroup definition: %s%s\n",
              pick$sg1,
              if (!is.na(pick$sg2)) paste0(" & ", pick$sg2) else ""))
  cat(sprintf("Captured training_fs_hr: %.3f\n", captured_hr))

  # Evaluate the subgroup expression(s) to get a membership flag.
  in_sg <- tryCatch({
    sg1_flag <- eval(parse(text = pick$sg1), envir = df_syn)
    if (!is.na(pick$sg2)) {
      sg2_flag <- eval(parse(text = pick$sg2), envir = df_syn)
      sg1_flag & sg2_flag
    } else {
      sg1_flag
    }
  }, error = function(e) NULL)

  if (!is.null(in_sg) && sum(in_sg) > 10L) {
    df_sg <- df_syn[in_sg, ]
    m <- tryCatch(
      glm(y ~ treat + offset(log(ftime)),
          family = poisson, data = df_sg),
      error = function(e) NULL
    )
    if (!is.null(m)) {
      irr_refit <- exp(coef(m)["treat"])
      cat(sprintf("Re-fit IRR on full-data subgroup: %.3f\n", irr_refit))

      .tc("Cross-check",
          "captured training_fs_hr and re-fit IRR both > 1",
          captured_hr > 1 && irr_refit > 1,
          detail = sprintf("captured = %.2f, refit = %.2f",
                           captured_hr, irr_refit))

      .tc("Cross-check",
          "captured and refit agree within 3x",
          (captured_hr / irr_refit > 1/3) &&
          (captured_hr / irr_refit < 3),
          detail = sprintf("ratio = %.2f",
                           captured_hr / irr_refit))
    }
  }
}
```


# v0.2.0 bootstrap return-value contract {#sec-v020}

Validates the v0.2.0 additions to the bootstrap return object that
downstream consumers (`print.fs_bootstrap`, CV-style fold
identification diagnostics, cross-API count comparisons) depend on.
Orthogonal to Phase C; included here so each outcome-type
compatibility doc exercises the contract independently.

```{r v020-helper}
`%||%` <- function(a, b) if (is.null(a)) b else a
```

## S3 class and core schema

```{r v020-class}
.tc("v0.2.0", "fs_bc inherits 'fs_bootstrap' class",
    inherits(fs_bc, "fs_bootstrap"),
    detail = sprintf("classes = %s",
                     paste(class(fs_bc), collapse = "/")))

.tc("v0.2.0", "results has any_found column",
    "any_found" %in% names(fs_bc$results))

.tc("v0.2.0", "any_found is integer",
    is.integer(fs_bc$results$any_found),
    detail = sprintf("class = %s",
                     class(fs_bc$results$any_found)[1]))

.tc("v0.2.0", "any_found values are all 0L or 1L",
    all(fs_bc$results$any_found %in% c(0L, 1L)),
    detail = sprintf("distinct = %s",
                     paste(sort(unique(fs_bc$results$any_found)),
                           collapse = ",")))
```

## `any_found` agrees with `!is.na(Pcons)`

```{r v020-equiv}
by_pcons    <- sum(!is.na(fs_bc$results$Pcons))
by_anyfound <- sum(fs_bc$results$any_found == 1L)

.tc("v0.2.0",
    "any_found count matches !is.na(Pcons) count",
    by_pcons == by_anyfound,
    detail = sprintf("Pcons=%d, any_found=%d", by_pcons, by_anyfound))
```

## Metadata fields populated on the count path

```{r v020-metadata}
.tc("v0.2.0", "nb_boots metadata non-NULL",
    !is.null(fs_bc$nb_boots),
    detail = sprintf("nb_boots = %s", fs_bc$nb_boots %||% "NULL"))

.tc("v0.2.0", "original_sg slot exists on return",
    "original_sg" %in% names(fs_bc))

.tc("v0.2.0", "outcome_type metadata non-NULL",
    !is.null(fs_bc$outcome_type),
    detail = sprintf("outcome_type = %s",
                     fs_bc$outcome_type %||% "NULL"))

.tc("v0.2.0", "effect_measure metadata non-NULL",
    !is.null(fs_bc$effect_measure),
    detail = sprintf("effect_measure = %s",
                     fs_bc$effect_measure %||% "NULL"))

.tc("v0.2.0", "est.scale metadata non-NULL",
    !is.null(fs_bc$est.scale),
    detail = sprintf("est.scale = %s",
                     fs_bc$est.scale %||% "NULL"))
```

## Metadata values match the count call

```{r v020-metadata-values}
.tc("v0.2.0", "nb_boots echoes the call argument",
    identical(fs_bc$nb_boots, nb_boots),
    detail = sprintf("returned=%s, expected=%d",
                     fs_bc$nb_boots %||% "NULL", nb_boots))

.tc("v0.2.0", "outcome_type == 'count'",
    identical(fs_bc$outcome_type, "count"),
    detail = sprintf("got %s", fs_bc$outcome_type %||% "NULL"))

.tc("v0.2.0", "effect_measure == 'IRR'",
    identical(fs_bc$effect_measure, "IRR"),
    detail = sprintf("got %s", fs_bc$effect_measure %||% "NULL"))
```

## `print.fs_bootstrap` dispatches without error

```{r v020-print-default}
print_err <- tryCatch(
  { capture.output(print(fs_bc)); NULL },
  error = function(e) conditionMessage(e)
)
.tc("v0.2.0",
    "print(fs_bc) dispatches without error",
    is.null(print_err),
    detail = if (is.null(print_err)) "OK" else print_err)
```

## `top_n` parameter honoured

```{r v020-print-topn}
topn_err <- tryCatch(
  { capture.output(print(fs_bc, top_n = 5L)); NULL },
  error = function(e) conditionMessage(e)
)
.tc("v0.2.0",
    "print(fs_bc, top_n = 5L) dispatches without error",
    is.null(topn_err),
    detail = if (is.null(topn_err)) "OK" else topn_err)
```


# Test summary {#sec-summary}

```{r summary-table}
n_total <- nrow(test_results)
n_pass  <- sum(test_results$Status == "PASS")
n_fail  <- n_total - n_pass

gt::gt(test_results,
       groupname_col = "Section",
       row_group_as_column = TRUE) |>
  gt::tab_header(
    title    = gt::md(sprintf(
      "**GLM compatibility (count / IRR) test results: %d / %d passing**",
      n_pass, n_total)),
    subtitle = sprintf(
      "Count outcome (IRR, offset=ftime), N=%d, CV sims=%d, nb_boots=%d",
      N, Ksims_cv, nb_boots)
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#d4edda"),
    locations = gt::cells_body(
      columns = "Status", rows = Status == "PASS"
    )
  ) |>
  gt::tab_style(
    style     = list(
      gt::cell_fill(color = "#f8d7da"),
      gt::cell_text(weight = "bold")
    ),
    locations = gt::cells_body(
      columns = "Status", rows = Status == "FAIL"
    )
  ) |>
  gt::cols_label(
    Test   = "Test case",
    Status = "Status",
    Detail = "Detail"
  ) |>
  gt::tab_options(table.font.size = gt::px(12))

if (n_fail == 0L) {
  message(sprintf("ALL TESTS PASSED: %d / %d", n_pass, n_total))
} else {
  message(sprintf("TESTS FAILED: %d / %d failing", n_fail, n_total))
  print(test_results[test_results$Status == "FAIL",
                     c("Section", "Test", "Detail")])
}
```


# Session info {#sec-session}

```{r session}
sessionInfo()
```
