---
title: "Living test: Phase C bootstrap `grf_cuts_b`"
subtitle: "Validates bootstrap-side GRF-cut capture and summary"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-tools: true
    code-fold: true
    theme: cosmo
    fontsize: 11pt
    number-sections: true
params:
  quick_mode: true       # TRUE: nb_boots=30 (~30 sec). FALSE: nb_boots=200 (~3-5 min).
  force_rerun: false     # TRUE: ignore checkpoint and re-run bootstrap.
---

# Purpose {#sec-purpose}

Validates Phase C: the `grf_cuts_b` column captured per bootstrap
iteration in `bootstrap_results()` and surfaced through the
`summarize_bootstrap_subgroups()` -> `format_subgroup_summary_tables()`
pipeline.  Parallel in spirit to the CV-side
`test_summarize_cv_results.qmd` but narrower in scope: Phase C is a
single new column and a single new summary slot.

Coverage:

1. **Column presence and domain** — `results$grf_cuts_b` is a
   character column, populated whenever the bootstrap call succeeded
   regardless of subgroup identification, `NA_character_` otherwise.
2. **`summarize_bootstrap_subgroups()` `grf_cut_freq` slot** —
   populated with a tidy `data.table(Rank, grf_cut, N, Percent)` that
   spans all bootstraps (not just the successful ones).  `N` sums to
   `nb_boots`.
3. **Formatter integration** — `format_subgroup_summary_tables()`
   produces a `grf_cut_freq` gt table with a `"(N others)"` tail row
   when distinct cuts exceed the display cap.
4. **End-to-end via `summarize_bootstrap_results()`** — exposes the
   new slot through the existing pipeline without disturbing other
   slots.
5. **Backward compatibility** — stripping `grf_cuts_b` from
   `results` degrades to `NULL` for the new slots without breaking
   anything else.

A cumulative pass/fail table renders in @sec-summary.


# Setup {#sec-setup}

```{r load-packages}
#| message: false
#| warning: false
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(future)
library(foreach)
library(doFuture)
```

```{r config}
if (isTRUE(params$quick_mode)) {
  nb_boots  <- 30L
  ckpt_path <- "test_bootstrap_grf_cuts_checkpoint_quick.rds"
  cat("Mode: quick (nb_boots = 30).  Set quick_mode = FALSE for 200.\n")
} else {
  nb_boots  <- 200L
  ckpt_path <- "test_bootstrap_grf_cuts_checkpoint_full.rds"
  cat("Mode: full (nb_boots = 200).  Expect 3-5 min runtime.\n")
}

confounders.name <- c("age", "meno", "size", "grade3",
                      "nodes", "pgr", "er")
outcome.name     <- "time_months"
event.name       <- "status"
treat.name       <- "hormon"
id.name          <- "id"

fs_args_base <- list(
  hr.threshold              = 1.25,
  hr.consistency            = 1.00,
  pconsistency.threshold    = 0.90,
  stop_threshold            = 0.95,
  sg_focus                  = "maxSG",
  max_subgroups_search      = 10L,
  use_twostage              = FALSE,
  use_grf                   = TRUE,
  return_selected_cuts_only = TRUE,
  use_lasso                 = TRUE,
  cut_type                  = "default",
  frac.tau                  = 0.8,
  dmin.grf                  = 12,
  maxk                      = 2L,
  n.min                     = 60L,
  d0.min                    = 10L,
  d1.min                    = 10L,
  fs.splits                 = 400L,
  details                   = FALSE,
  quiet                     = TRUE,
  plot.sg                   = FALSE,
  plot.grf                  = FALSE
)

n_workers <- floor(0.95 * max(1L, parallel::detectCores() - 1L))
```

```{r data-setup}
df             <- survival::gbsg
df$grade3      <- as.integer(df$grade == "3")
df$id          <- seq_len(nrow(df))
df$time_months <- df$rfstime / 30.4375
```

```{r test-harness}
test_results <- data.frame(
  Section = character(0),
  Test    = character(0),
  Status  = character(0),
  Detail  = character(0),
  stringsAsFactors = FALSE
)

.tc <- function(section, test, pass, detail = "") {
  status <- if (isTRUE(pass)) "PASS" else "FAIL"
  test_results <<- rbind(test_results, data.frame(
    Section = section, Test = test, Status = status,
    Detail  = substr(detail, 1L, 120L),
    stringsAsFactors = FALSE
  ))
  invisible(pass)
}
```


# Baseline run {#sec-baseline}

Full-data ForestSearch followed by a small bootstrap.  Checkpoint
preserves the `fs_bc` object; stale checkpoints are detected via the
`grf_cuts_b` column.

```{r fs-full}
#| message: false
fs_full <- do.call(forestsearch, c(list(
  df.analysis      = df,
  confounders.name = confounders.name,
  outcome.name     = outcome.name,
  event.name       = event.name,
  treat.name       = treat.name,
  id.name          = id.name,
  parallel_args    = list(plan         = "multisession",
                          workers      = n_workers,
                          show_message = FALSE)
), fs_args_base))
plan("sequential")

cat("Full-data subgroup:",
    paste(fs_full$sg.harm, collapse = " & "), "\n")
cat("Full-data GRF cuts:",
    paste(fs_full$grf_cuts, collapse = " | "), "\n")
```

```{r bootstrap-run}
#| message: false
.run_bootstrap <- function() {
  t0 <- proc.time()
  out <- forestsearch_bootstrap_dofuture(
    fs.est        = fs_full,
    nb_boots      = nb_boots,
    parallel_args = list(plan         = "multisession",
                         workers      = n_workers,
                         show_message = FALSE)
  )
  plan("sequential")
  attr(out, "runtime_seconds") <-
    as.numeric((proc.time() - t0)["elapsed"])
  out
}

if (file.exists(ckpt_path) && !isTRUE(params$force_rerun)) {
  message(sprintf("Loading bootstrap from checkpoint: %s", ckpt_path))
  fs_bc <- readRDS(ckpt_path)

  # Phase C staleness check.
  if (!"grf_cuts_b" %in% names(fs_bc$results)) {
    message(sprintf(paste0(
      "Checkpoint %s predates Phase C (no grf_cuts_b column). ",
      "Re-running bootstrap..."), ckpt_path))
    fs_bc <- .run_bootstrap()
    saveRDS(fs_bc, ckpt_path)
    message(sprintf("Wrote refreshed checkpoint to %s", ckpt_path))
  }
} else {
  message(sprintf("Running bootstrap (nb_boots = %d) ...", nb_boots))
  fs_bc <- .run_bootstrap()
  saveRDS(fs_bc, ckpt_path)
  message(sprintf("Wrote checkpoint to %s", ckpt_path))
}

cat(sprintf("Bootstrap iterations: %d\n", nrow(fs_bc$results)))
cat(sprintf("results columns: %s\n",
            paste(names(fs_bc$results), collapse = ", ")))

.tc("Baseline", "results$grf_cuts_b column present",
    "grf_cuts_b" %in% names(fs_bc$results),
    detail = "Required for Phase C summary slots")

.tc("Baseline", "nrow(results) == nb_boots",
    nrow(fs_bc$results) == nb_boots,
    detail = sprintf("obs=%d exp=%d",
                     nrow(fs_bc$results), nb_boots))
```


# Test 1: Column domain {#sec-test1}

```{r test1}
gcb <- fs_bc$results$grf_cuts_b

.tc("Test 1: column", "grf_cuts_b is character",
    is.character(gcb),
    detail = sprintf("class=%s", class(gcb)[1]))

# Non-empty strings should look like GRF cut expressions or be NA.
# Empty strings are not expected (capture logic either populates with
# a real cut or defaults to NA_character_).
.tc("Test 1: column", "no empty-string grf_cuts_b values",
    !any(!is.na(gcb) & gcb == ""),
    detail = sprintf("n empty=%d",
                     sum(!is.na(gcb) & gcb == "")))

# grf_cuts_b NA implies the bootstrap either errored or produced no
# cut.  On a successful GBSG bootstrap at these settings, we expect
# most iterations to produce some cut.  Not a strict assertion --
# warn rather than fail if the rate is high.
na_rate <- mean(is.na(gcb))
.tc("Test 1: column", "NA rate for grf_cuts_b is reported",
    TRUE,  # always pass; informational only
    detail = sprintf("NA rate = %.1f%%", 100 * na_rate))

# Values that ARE populated should contain at least one "<=" or ">="
# (GRF cut expressions have the pattern "variable op threshold").
nonna <- gcb[!is.na(gcb) & gcb != ""]
has_op <- length(nonna) == 0L ||
  all(grepl("<=|>=|<|>|=", nonna))
.tc("Test 1: column", "populated cuts contain comparison operator",
    has_op,
    detail = sprintf("populated=%d; all look like cuts=%s",
                     length(nonna), has_op))
```


# Test 2: `summarize_bootstrap_subgroups()` `grf_cut_freq` slot {#sec-test2}

Exercises the exported `summarize_bootstrap_subgroups()` tabulation
helper directly, validating its contract independent of the
`summarize_bootstrap_results()` orchestrator (which is validated
end-to-end in @sec-test4).

```{r test2}
subg_summary <- summarize_bootstrap_subgroups(
  results     = fs_bc$results,
  nb_boots    = nb_boots,
  original_sg = fs_full$sg.harm,
  maxk        = 2L
)

.tc("Test 2: subgroups slot", "grf_cut_freq slot present",
    "grf_cut_freq" %in% names(subg_summary))
.tc("Test 2: subgroups slot", "grf_cut_freq is non-NULL",
    !is.null(subg_summary$grf_cut_freq))
.tc("Test 2: subgroups slot", "grf_cut_freq inherits data.table",
    inherits(subg_summary$grf_cut_freq, "data.table"))

gcf <- subg_summary$grf_cut_freq
expected_cols <- c("Rank", "grf_cut", "N", "Percent")
.tc("Test 2: subgroups slot", "grf_cut_freq has expected columns",
    all(expected_cols %in% names(gcf)),
    detail = paste(setdiff(expected_cols, names(gcf)), collapse = ", "))

# Spans all bootstraps, not just successful ones.  N must sum to
# nb_boots exactly.
.tc("Test 2: subgroups slot", "grf_cut_freq N sums to nb_boots",
    sum(gcf$N) == nb_boots,
    detail = sprintf("sum=%d exp=%d", sum(gcf$N), nb_boots))

# Percent must sum to ~100 (modulo rounding)
.tc("Test 2: subgroups slot", "grf_cut_freq Percent sums to ~100",
    abs(sum(gcf$Percent) - 100) < 0.5,
    detail = sprintf("sum=%.2f", sum(gcf$Percent)))

# Rank is sequential 1..nrow
.tc("Test 2: subgroups slot", "grf_cut_freq Rank is sequential",
    identical(gcf$Rank, seq_len(nrow(gcf))),
    detail = sprintf("rank range=[%d, %d], nrow=%d",
                     min(gcf$Rank), max(gcf$Rank), nrow(gcf)))

# Sort invariant: N non-increasing
.tc("Test 2: subgroups slot", "grf_cut_freq sorted by descending N",
    all(diff(gcf$N) <= 0),
    detail = sprintf("first N=%d, last N=%d",
                     gcf$N[1], gcf$N[nrow(gcf)]))

# NA values from grf_cuts_b must map to a single "(no GRF cut)" row
# (not be silently dropped).
na_count_in_col  <- sum(is.na(fs_bc$results$grf_cuts_b))
no_grf_cut_row_n <- if ("(no GRF cut)" %in% gcf$grf_cut) {
  gcf$N[gcf$grf_cut == "(no GRF cut)"]
} else {
  0L
}
.tc("Test 2: subgroups slot", "NA values map to '(no GRF cut)' row",
    na_count_in_col == no_grf_cut_row_n,
    detail = sprintf("NAs in col=%d, '(no GRF cut)' row N=%d",
                     na_count_in_col, no_grf_cut_row_n))
```


# Test 3: `format_subgroup_summary_tables()` integration {#sec-test3}

```{r test3}
fmt_tables <- forestsearch:::format_subgroup_summary_tables(
  subg_summary, nb_boots
)

.tc("Test 3: format", "grf_cut_freq gt table produced",
    "grf_cut_freq" %in% names(fmt_tables))
.tc("Test 3: format", "grf_cut_freq is a gt_tbl",
    inherits(fmt_tables$grf_cut_freq, "gt_tbl"))

# Previously-existing tables still present (no regression)
prior_tables <- c("basic_stats", "agreement")
for (tn in prior_tables) {
  if (tn %in% names(fmt_tables)) {
    .tc("Test 3: format",
        sprintf("prior table '%s' still produced", tn),
        inherits(fmt_tables[[tn]], "gt_tbl"),
        detail = "no regression")
  }
}
```

## Rendered grf_cut_freq table

```{r test3-display}
fmt_tables$grf_cut_freq
```


# Test 4: End-to-end via `summarize_bootstrap_results()` {#sec-test4}

```{r test4}
summaries <- summarize_bootstrap_results(
  sgharm       = fs_full$sg.harm,
  boot_results = fs_bc,
  create_plots = FALSE
)

.tc("Test 4: end-to-end",
    "subgroup_summary$grf_cut_freq reachable via top-level summaries",
    !is.null(summaries$subgroup_summary$grf_cut_freq))
.tc("Test 4: end-to-end",
    "top-level table slot still produced",
    inherits(summaries$table, "gt_tbl"))
.tc("Test 4: end-to-end",
    "top-level diagnostics_table_gt still produced",
    inherits(summaries$diagnostics_table_gt, "gt_tbl"))

# Same grf_cut_freq content regardless of whether we called
# summarize_bootstrap_subgroups() directly or via summarize_bootstrap_results()
.tc("Test 4: end-to-end",
    "grf_cut_freq identical through both access paths",
    isTRUE(all.equal(
      summaries$subgroup_summary$grf_cut_freq,
      subg_summary$grf_cut_freq
    )))
```


# Test 5: Backward compatibility {#sec-test5}

Strip `grf_cuts_b` from `results` to simulate a pre-Phase-C object.
Expects graceful `NULL` for the new slot and no disturbance to the
rest of the summary pipeline.

```{r test5}
fs_bc_nogrf <- fs_bc
fs_bc_nogrf$results <- fs_bc$results[
  , setdiff(names(fs_bc$results), "grf_cuts_b"), with = FALSE
]

subg_nogrf <- summarize_bootstrap_subgroups(
  results     = fs_bc_nogrf$results,
  nb_boots    = nb_boots,
  original_sg = fs_full$sg.harm,
  maxk        = 2L
)

.tc("Test 5: backward-compat",
    "grf_cut_freq NULL when column absent",
    is.null(subg_nogrf$grf_cut_freq))
.tc("Test 5: backward-compat",
    "basic_stats still produced when Phase C absent",
    !is.null(subg_nogrf$basic_stats))
.tc("Test 5: backward-compat",
    "factor_freq still produced when Phase C absent",
    !is.null(subg_nogrf$factor_freq))

fmt_nogrf <- forestsearch:::format_subgroup_summary_tables(
  subg_nogrf, nb_boots
)
.tc("Test 5: backward-compat",
    "formatter omits grf_cut_freq gt when source NULL",
    !"grf_cut_freq" %in% names(fmt_nogrf))
.tc("Test 5: backward-compat",
    "other formatter outputs still produced",
    "basic_stats" %in% names(fmt_nogrf))
```


# Test 6: Empty-results edge case {#sec-test6}

Confirms the early-return path in `summarize_bootstrap_subgroups()`
(no successful iterations) still populates `grf_cut_freq` when
`grf_cuts_b` is present -- this is the most diagnostically valuable
case.

```{r test6}
# Synthesise a results object where no iteration identified a
# subgroup (all Pcons NA) but grf_cuts_b is populated.
res_no_subg <- data.table::copy(fs_bc$results)
res_no_subg[, Pcons := NA_real_]  # force n_found = 0

subg_no <- suppressWarnings(summarize_bootstrap_subgroups(
  results     = res_no_subg,
  nb_boots    = nb_boots,
  original_sg = NULL,
  maxk        = 2L
))

.tc("Test 6: empty-results", "early-return populates grf_cut_freq",
    !is.null(subg_no$grf_cut_freq),
    detail = "most diagnostic case: GRF output when no sg found")
.tc("Test 6: empty-results", "n_found is 0 in early-return",
    subg_no$n_found == 0L)

# When early-return triggers, grf_cut_freq should still sum to nb_boots
if (!is.null(subg_no$grf_cut_freq)) {
  .tc("Test 6: empty-results",
      "grf_cut_freq N sums to nb_boots in early-return",
      sum(subg_no$grf_cut_freq$N) == nb_boots,
      detail = sprintf("sum=%d exp=%d",
                       sum(subg_no$grf_cut_freq$N), nb_boots))
}
```


# Test summary {#sec-summary}

```{r summary-table}
n_total <- nrow(test_results)
n_pass  <- sum(test_results$Status == "PASS")
n_fail  <- n_total - n_pass

gt::gt(test_results,
       groupname_col = "Section",
       row_group_as_column = TRUE) |>
  gt::tab_header(
    title    = gt::md(sprintf(
      "**Phase C test results: %d / %d passing**", n_pass, n_total)),
    subtitle = sprintf(
      "GBSG, nb_boots = %d, dmin.grf = %g, frac.tau = %g",
      nb_boots, fs_args_base$dmin.grf, fs_args_base$frac.tau)
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#d4edda"),
    locations = gt::cells_body(
      columns = "Status", rows = Status == "PASS"
    )
  ) |>
  gt::tab_style(
    style     = list(
      gt::cell_fill(color = "#f8d7da"),
      gt::cell_text(weight = "bold")
    ),
    locations = gt::cells_body(
      columns = "Status", rows = Status == "FAIL"
    )
  ) |>
  gt::cols_label(
    Test   = "Test case",
    Status = "Status",
    Detail = "Detail"
  ) |>
  gt::tab_options(table.font.size = gt::px(12))

if (n_fail == 0L) {
  message(sprintf("ALL TESTS PASSED: %d / %d", n_pass, n_total))
} else {
  message(sprintf("TESTS FAILED: %d / %d failing", n_fail, n_total))
  print(test_results[test_results$Status == "FAIL",
                     c("Section", "Test", "Detail")])
}
```


# Session info {#sec-session}

```{r session}
sessionInfo()
```
