---
title: "ACTG175 (Continuous): Comparing Selection Rules"
subtitle: "hrMaxSG with selection_rule ∈ {pareto, both}"
format: html
---

```{r setup}
#| code-fold: true
#| code-summary: "Setup (click to expand)"

# Initialize timing
t_doc_start <- proc.time()
timings <- list()
```


# Introduction

This vignette compares the selection diagnostic side-by-side across two
choices of `selection_rule`, holding all other ForestSearch parameters
fixed.  It uses the package's
[`compare_selection_rules()`](../R/compare_selection_rules.R) wrapper to
run `forestsearch()` once per combination and bundle the per-combo
diagnostics for inspection.

The structure is:

1. **Data setup + ITT** — identical to the single-rule analyses.
2. **Comparison run** — one call to `compare_selection_rules()` that
   handles the loop, captures each run's
   `CANDIDATE EVALUATION PREVIEW` + `CANDIDATE EVALUATION SUMMARY`
   banners, builds the Pareto-frontier plots, and computes split-derived
   CIs.
3. **Per-combo diagnostics** — replay of captured console output,
   individual frontier plot/table/narrative for each combination.
4. **Side-by-side composition** — patchwork-composed plot panel
   for direct visual comparison.

Bootstrap, cross-validation, and the combined forest plot are
**deliberately omitted** so the document stays focused on the
*selection mechanism* and the comparison reads as a tight diagnostic
of the rule axis alone.


# Setup

## Load Required Packages

```{r load-packages}
suppressPackageStartupMessages({
  library(forestsearch)
  library(speff2trial)       # ACTG175 dataset
  library(data.table)
  library(ggplot2)
  library(gt)
  library(future)
  library(patchwork)         # for side-by-side plot composition
})

# Source the comparison wrapper.  When this graduates into the package
# (copy compare_selection_rules.R into R/ and run devtools::document()),
# this source() line can be removed.
#source("compare_selection_rules.R")

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```


## Analysis Parameters

All ForestSearch parameters are defined here once and reused for both
combinations.  The two axes of variation are `sg_focus` and
`selection_rule`, which the wrapper handles directly.

```{r analysis-params}
#| code-fold: false

# ─── Parallel processing ────────────────────────────────────────────────
n_cores <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))

# ─── Reproducibility ────────────────────────────────────────────────────
analysis_seed <- 8316951L
set.seed(analysis_seed)

# ─── Confounders ───────────────────────────────────────────────────────
cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
bin_vars  <- c("hemo", "homo", "drugs", "race", "gender", "symptom")
confounders.name <- c(cont_vars, bin_vars)
id.name <- "id"

# ─── Variables (dual-scale framing) ────────────────────────────────────
treat.name         <- "treat"
adverse_outcome    <- "y_decline"   # FS adverse-scale outcome
beneficial_outcome <- "y_improve"   # GRF beneficial-scale outcome

# ─── ForestSearch parameters (continuous, MD scale) ────────────────────
fs_outcome_type    <- "continuous"
fs_effect_measure  <- "MD"

# NOTE: fs_sg_focus and fs_selection_rule are NOT defined here.
# They live in the `comparison-axes` chunk below as the two axes the
# wrapper varies.  All OTHER ForestSearch parameters are fixed across
# the comparison.

# Thresholds in CD4 count units (adverse scale: MD > c1 = harm)
fs_md_threshold    <- 10    # screening: MD >= 10 indicates harm
fs_md_consistency  <- 5     # consistency: MD >= 5

fs_pconsistency    <- 0.90
fs_n_min           <- 60L
fs_d0_min          <- 10L
fs_d1_min          <- 10L
fs_splits          <- 500L
fs_max_subgroups   <- 20
fs_maxk            <- 2L
fs_use_lasso       <- TRUE
fs_use_grf         <- TRUE
fs_is_RCT          <- TRUE
fs_vi_grf_min      <- -0.2
fs_dmin_grf        <- 0
fs_cont_cutoff     <- 4L

# ─── Pareto-frontier diagnostic parameters ──────────────────────────────
frontier_params <- list(
  n_splits      = 1000L,
  seed          = 1L,
  xlim          = NULL,
  digits        = 2L
)
```


## Comparison Axes

The two axes varied by the wrapper.  Tuple semantics: paired
element-by-element.

```{r comparison-axes}
sg_focus_vec       <- c("hrMinSG", "hrMinSG")
selection_rule_vec <- c("pareto",  "both")

# Pretty labels used in plot titles, list names, and prose
combo_labels       <- paste(sg_focus_vec, selection_rule_vec, sep = " / ")
n_combos           <- length(sg_focus_vec)

cat("Combinations to be run:\n")
print(data.frame(
  combo          = seq_len(n_combos),
  sg_focus       = sg_focus_vec,
  selection_rule = selection_rule_vec,
  label          = combo_labels
))
```


# Data Preparation

## Treatment Assignment

```{r data-treat}
# ACTG175: arm 1 = ZDV+ddI (experimental), arm 3 = ddI (control)
actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$id <- seq_len(nrow(actg_df))

# Treatment: ZDV+ddI = 1, ddI = 0 (NO switching)
actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)
```


## Outcome Construction

```{r data-outcomes}
# Drop subjects missing the week-20 CD4 measurement
actg_df <- actg_df[!is.na(actg_df$cd420), ]

# Adverse scale: CD4 decline (higher = worse)
actg_df$y_decline <- actg_df$cd40 - actg_df$cd420

outcome_vars <- c("y_decline")
```


## Covariates

```{r data-covariates}
# Coerce binary indicators to factors (analyst does not know true cuts
# for continuous covariates; ForestSearch handles those internally)
for (v in bin_vars) actg_df[[v]] <- as.factor(actg_df[[v]])

cat("Continuous candidates:", paste(cont_vars, collapse = ", "), "\n")
cat("Binary candidates:    ", paste(bin_vars, collapse = ", "), "\n")
cat("Analysis N:", nrow(actg_df), "\n")
```


## Baseline Characteristics

```{r baseline-table}
create_summary_table(
  data             = actg_df,
  treat_var        = treat.name,
  table_title      = "ACTG175 Baseline Characteristics and Outcomes by Treatment Arm",
  vars_continuous  = c(outcome_vars, cont_vars),
  vars_categorical = bin_vars,
  font_size        = 12
)
```


## ITT Analysis

```{r itt-analysis}
itt_decline <- lm(y_decline ~ treat, data = actg_df)
cat("=== ITT Mean Difference, adverse scale (y_decline) ===\n")
print(summary(itt_decline)$coefficients["treat", , drop = FALSE])
```


# Selection-Rule Comparison

## Running `compare_selection_rules()`

One call to the wrapper runs `forestsearch()` for each
`(sg_focus, selection_rule)` combination with all other parameters
held fixed, and bundles the per-combo diagnostics into a single
object.

```{r run-comparison}
#| message: false
#| warning: false

t0 <- proc.time()

comparison <- compare_selection_rules(
  df.analysis      = actg_df,
  sg_focus         = sg_focus_vec,
  selection_rule   = selection_rule_vec,
  combo_labels     = combo_labels,

  # Wrapper-specific controls
  compute_cis      = TRUE,
  n_splits         = frontier_params$n_splits,
  ci_seed          = frontier_params$seed,
  plot_xlim        = frontier_params$xlim,
  show_band        = TRUE,
  verbose          = TRUE,

  # All other ForestSearch arguments — pass-through via ... .
  # These match the single-rule analysis chunks 1-for-1 except for
  # sg_focus and selection_rule (which the wrapper supplies).
  confounders.name = confounders.name,
  outcome.name     = adverse_outcome,
  treat.name       = treat.name,
  id.name          = id.name,
  outcome_type     = fs_outcome_type,
  effect_measure   = fs_effect_measure,
  adverse_outcome  = TRUE,
  seedit           = analysis_seed,
  hr.threshold     = fs_md_threshold,
  hr.consistency   = fs_md_consistency,
  pconsistency.threshold = fs_pconsistency,
  max_subgroups_search = fs_max_subgroups,
  use_twostage     = TRUE,
  conf.cont_jcuts  = list(cd40 = 10, wtkg = 10),
  use_grf          = fs_use_grf,
  return_selected_cuts_only = TRUE,
  use_lasso        = fs_use_lasso,
  cut_type         = "default",
  maxk             = fs_maxk,
  n.min            = fs_n_min,
  d0.min           = fs_d0_min,
  d1.min           = fs_d1_min,
  fs.splits        = fs_splits,
  parallel_args    = list(
    plan         = "multisession",
    workers      = n_cores,
    show_message = TRUE
  ),
  show_candidate_summary = TRUE,
  details          = TRUE,
  plot.sg          = FALSE,
  is.RCT           = fs_is_RCT,
  vi.grf.min       = fs_vi_grf_min,
  dmin.grf         = fs_dmin_grf,
  cont.cutoff      = fs_cont_cutoff,
  tune_grf         = FALSE,
  quiet            = FALSE
)

plan("sequential"); invisible(gc(verbose = FALSE))
timings$compare <- (proc.time() - t0)["elapsed"]

cat("\nComparison completed in",
    round(timings$compare / 60, 1), "minutes\n")
```


## Comparison Summary

```{r comparison-summary}
print(comparison)
```


# Per-Combo Diagnostics

## Captured Console Output

The wrapper captured the in-flight diagnostic output from each
`forestsearch()` call.  Each combination's two diagnostic banners --
the `CANDIDATE EVALUATION PREVIEW` (pre-consistency; all candidates
entering evaluation) and the `CANDIDATE EVALUATION SUMMARY`
(post-consistency; passing candidates with
Pcons / Frontier / InBand / Selected flags) -- are shown below as the
primary view.  The full captured output (GRF preview, Cox-LASSO
selection, filtering summary, parallel-config messages, post-run
timings, etc.) is folded into a collapsible block for readers who want
the complete context.

### `r combo_labels[1]`

```{r console-1, results = "asis"}
diag <- comparison$diagnostics[[1]]

cat("**Pre-consistency preview:**\n\n")
if (!is.na(diag$preview)) {
  cat("```\n", diag$preview, "\n```\n\n", sep = "")
} else {
  cat("_(not available)_\n\n")
}

cat("**Post-consistency summary:**\n\n")
if (!is.na(diag$summary)) {
  cat("```\n", diag$summary, "\n```\n\n", sep = "")
} else {
  cat("_(not available)_\n\n")
}

cat("<details><summary>Full forestsearch output (expand)</summary>\n\n```\n")
cat(comparison$console[[1]], sep = "\n")
cat("\n```\n\n</details>\n")
```


### `r combo_labels[2]`

```{r console-2, results = "asis"}
diag <- comparison$diagnostics[[2]]

cat("**Pre-consistency preview:**\n\n")
if (!is.na(diag$preview)) {
  cat("```\n", diag$preview, "\n```\n\n", sep = "")
} else {
  cat("_(not available)_\n\n")
}

cat("**Post-consistency summary:**\n\n")
if (!is.na(diag$summary)) {
  cat("```\n", diag$summary, "\n```\n\n", sep = "")
} else {
  cat("_(not available)_\n\n")
}

cat("<details><summary>Full forestsearch output (expand)</summary>\n\n```\n")
cat(comparison$console[[2]], sep = "\n")
cat("\n```\n\n</details>\n")
```


## Pareto-Frontier Plots

Each combination produces its own Pareto frontier on (MD, N).  The
plot for each combo is shown individually below, then composed
side-by-side via `patchwork` for direct visual comparison.

### Individual Panels

```{r plot-combo-1}
#| fig-width: 8
#| fig-height: 5
#| fig-cap: !expr sprintf("Pareto frontier — %s", combo_labels[1])

print(comparison$plots[[1]])
```


```{r plot-combo-2}
#| fig-width: 8
#| fig-height: 5
#| fig-cap: !expr sprintf("Pareto frontier — %s", combo_labels[2])

print(comparison$plots[[2]])
```


### Side-by-Side Composition

```{r plot-side-by-side}
#| fig-width: 14
#| fig-height: 6
#| fig-cap: "Pareto frontier across selection rules (side-by-side)"

print(comparison$plot_grid)
```


## Combined Plot (when passing sets match)

When the two configurations produce the same passing set, the combined
plot below labels each selected subgroup with one or more `S<i>:`
markers naming the configuration(s) that picked it. When the sets
differ, this returns `NULL` and the side-by-side view above is the
appropriate alternative.

```{r plot-combined}
#| fig-width: 10
#| fig-height: 6
#| fig-cap: "Combined Pareto frontier with S1/S2 winner annotations"

if (!is.null(comparison$plot_combined)) {
  print(comparison$plot_combined)
} else {
  cat("No combined plot: passing sets differ across configurations.\n")
  cat("See the side-by-side composition above instead.\n")
}
```


## Frontier Tables

Per-combination Pareto-frontier table with split-derived and FSBC
CIs.  Selected subgroup is highlighted with ★.

### `r combo_labels[1]`

```{r table-combo-1}
pareto_frontier_table(
  comparison$fs[[1]],
  ci_table = comparison$ci_tab[[1]],
  digits   = frontier_params$digits
)
```


### `r combo_labels[2]`

```{r table-combo-2}
pareto_frontier_table(
  comparison$fs[[2]],
  ci_table = comparison$ci_tab[[2]],
  digits   = frontier_params$digits
)
```


## Selection Narratives

`explain_pareto_selection()` produces a verbal account of why the
selected subgroup wins on the configured lexicographic criterion
relative to the other non-dominated candidates.  Reading the two
narratives together highlights which axis of the lex key the two
rules tie-broke on differently.

### `r combo_labels[1]`

```{r explain-combo-1, results = "asis"}
cat(explain_pareto_selection(
  comparison$fs[[1]],
  ci_table = comparison$ci_tab[[1]],
  digits   = frontier_params$digits
))
```


### `r combo_labels[2]`

```{r explain-combo-2, results = "asis"}
cat(explain_pareto_selection(
  comparison$fs[[2]],
  ci_table = comparison$ci_tab[[2]],
  digits   = frontier_params$digits
))
```


# Rich Post-Consistency View

The `include_dominated = TRUE` argument to `pareto_frontier_table()`
shows **all** passing candidates (not just the frontier), with
`on_frontier` and `in_band` flag columns added.  This is the
post-hoc analog of the `CANDIDATE EVALUATION SUMMARY` banner printed
during the run.

### `r combo_labels[1]`

```{r dominated-1}
pareto_frontier_table(
  comparison$fs[[1]],
  ci_table          = comparison$ci_tab[[1]],
  include_dominated = TRUE,
  digits            = frontier_params$digits
)
```


### `r combo_labels[2]`

```{r dominated-2}
pareto_frontier_table(
  comparison$fs[[2]],
  ci_table          = comparison$ci_tab[[2]],
  include_dominated = TRUE,
  digits            = frontier_params$digits
)
```


# Programmatic Access (Reference)

```{r programmatic, eval = FALSE}
# Each combination's full fs object is available for downstream use:
comparison$fs[[1]]       # selection_rule = "pareto"
comparison$fs[[2]]       # selection_rule = "both"

# CI tables (data.tables):
comparison$ci_tab[[1]]
comparison$ci_tab[[2]]

# Plot objects (ggplot):
comparison$plots[[1]]
comparison$plots[[2]]

# Console output (full captured stdout, one character per line):
comparison$console[[1]]
comparison$console[[2]]

# Diagnostics (filtered slices: preview, summary, full):
comparison$diagnostics[[1]]$preview     # PREVIEW block as single string
comparison$diagnostics[[1]]$summary     # SUMMARY block as single string
comparison$diagnostics[[1]]$full        # entire captured stdout (single string)

# Errors (NULL if combo succeeded):
comparison$errors[[1]]
comparison$errors[[2]]

# The combos data.frame:
comparison$combos
```


# Computational Timing

```{r timing-summary}
timings$total <- (proc.time() - t_doc_start)["elapsed"]

timing_df <- data.frame(
  Component = c("Comparison run (both combos)", "Total document"),
  Seconds   = c(timings$compare, timings$total)
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(title = "Computational Timing") |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1) |>
  cols_label(
    Component = "Component",
    Seconds   = "Time (sec)",
    Minutes   = "Time (min)"
  )
```


# References

León LA, Ye J, Zhao H. *A Bootstrap Framework for Subgroup
Identification in Randomized Clinical Trials with Bias Correction.*
Statistics in Medicine. 2024.  [doi:10.1002/sim.10163](https://doi.org/10.1002/sim.10163)
