---
title: "DINA Table 2 Reproduction: Gaussian, n = 1024"
author: "forestsearch contributors"
date: today
format:
  html:
    code-fold: false
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
vignette: >
  %\VignetteIndexEntry{DINA Table 2 Reproduction: Gaussian, n = 1024}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

# Goal {#sec-goal}

Reproduce the **first block of Table 2** in the Supplementary Materials of
Gao and Hastie (2025) for the **DINA** method only:

> Coverage and width of 95% confidence intervals for Gaussian responses,
> sample size 1024, propensity score estimated by logistic regression,
> baseline natural-parameter functions by linear regression, 100 bootstrap
> CI samples per trial, averaged over 100 trials.

Paper reference values (DINA row, n = 1024):

| Coefficient | β₁ | β₂ | β₃ | β₄ | β₅ | β₀ |
|---|---|---|---|---|---|---|
| **Coverage** | 0.97 | 0.96 | 0.93 | 0.95 | 0.93 | 0.98 |
| **Width** | 1.48 | 1.49 | 1.38 | 1.35 | 1.34 | 0.789 |

The user-facing `dina()` already implements Algorithm 1 with the right
defaults for this row (`propensity_method = "logistic"`,
`baseline_method = NULL`, `cross_fitting = TRUE`, `n_folds = 2L`). This
document supplies the data-generating mechanism, the bootstrap CI wrapper,
and the simulation driver, then summarises results against the paper.

# Setup {#sec-setup}

```{r setup}
#| message: false
#| warning: false

library(forestsearch)
library(data.table)
library(doFuture)
library(future)
library(gt)
```

# Parameters {#sec-params}

All tunable parameters live in this single chunk. The DGM hyperparameters
are taken verbatim from `simulation_setting.xlsx` (Gao & Hastie supplement),
sheet `GLM`, column `Gaussian`.

```{r params}
# --- Data-generating mechanism (Gao & Hastie supp., sheet 'GLM', Gaussian)
DGM <- list(
  d         = 5L,
  alpha     = c(0.5, 0.5, 0.5, 0,   0),     # nuisance natural-parameter
  beta      = c(0.8, 0.7, 0.6, 0,   0),     # the DINA target
  theta     = c(1,  -1,   0,   0,   0),     # propensity coefficients
  delta     = 2,                            # x_1 * x_2 interaction strength
  sigma2    = 4                             # outcome residual variance
)

# True target for coverage assessment: tau(x) = beta_0 + sum_j beta_j * x_j.
# The DGM has no intercept in tau and no x_1*x_2 interaction in tau (it
# cancels between eta_1 and eta_0).
TRUE_BETA <- c("(Intercept)" = 0,
               X1 = 0.8, X2 = 0.7, X3 = 0.6, X4 = 0, X5 = 0)

# --- Simulation configuration
SIM <- list(
  n_sample  = 1024L,         # paper Table 2 first block
  n_trials  = 100L,          # paper averages over 100 trials
  n_boot    = 100L,          # paper uses 100 bootstrap samples
  level     = 0.95,
  n_workers = 20L,           # adjust to local machine
  seed_base = 20260528L
)

# Paper reference values (DINA, n = 1024) for the comparison table
PAPER_REF <- data.table::data.table(
  coef     = c("X1", "X2", "X3", "X4", "X5", "(Intercept)"),
  cov_ref  = c(0.97, 0.96, 0.93, 0.95, 0.93, 0.98),
  wid_ref  = c(1.48, 1.49, 1.38, 1.35, 1.34, 0.789)
)
```

# Helper: data-generating mechanism {#sec-dgm}

A small inline DGM for the Gao & Hastie Gaussian setting. Kept local to the
document; if it proves useful elsewhere it can later be promoted to a
package function with full roxygen documentation.

```{r dgm-helper}
#' Generate one Gaussian dataset from the Gao & Hastie (2025) DGM
#'
#' @param n integer; sample size.
#' @param dgm list with components `d`, `alpha`, `beta`, `theta`, `delta`,
#'   `sigma2`.
#' @param seed optional integer seed; passed to `set.seed()` if supplied.
#'
#' @return `data.table` with columns `X1, ..., Xd, treat, Y`.
#'
#' @keywords internal
generate_dina_gaussian <- function(n, dgm, seed = NULL) {

  if (!is.null(seed)) set.seed(seed)

  d <- dgm$d
  X <- matrix(stats::runif(n * d, min = -1, max = 1), nrow = n, ncol = d)
  colnames(X) <- paste0("X", seq_len(d))

  # Propensity: e(x) = sigmoid(x' theta), W | X ~ Bernoulli(e(x))
  lin_e <- as.numeric(X %*% dgm$theta)
  e_x   <- 1 / (1 + exp(-lin_e))
  W     <- stats::rbinom(n, size = 1L, prob = e_x)

  # Natural parameter functions: eta_w(x) = x' alpha_w + delta * x_1 * x_2,
  # where alpha_1 = alpha + beta.  The interaction is present in BOTH
  # arms, so the true DINA tau(x) = x' beta (no intercept, no interaction).
  inter <- dgm$delta * X[, 1] * X[, 2]
  eta0  <- as.numeric(X %*% dgm$alpha) + inter
  eta1  <- as.numeric(X %*% (dgm$alpha + dgm$beta)) + inter
  eta_W <- ifelse(W == 1L, eta1, eta0)

  # Gaussian outcome with sigma^2 = 4
  eps <- stats::rnorm(n, mean = 0, sd = sqrt(dgm$sigma2))
  Y   <- eta_W + eps

  # Build the data.table explicitly to avoid any matrix-expansion
  # ambiguity in `data.table::data.table(X, ...)` with a matrix X.
  dt <- data.table::as.data.table(X)
  dt$treat <- W
  dt$Y     <- Y
  dt
}
```

# Helper: bootstrap CIs from a single dataset {#sec-bootstrap}

```{r boot-helper}
#' Nonparametric bootstrap CIs for a single DINA fit
#'
#' Resamples rows with replacement `B` times, refits `dina()` on each
#' resample, and returns both percentile and normal-approximation
#' `level`-CIs for every coefficient.
#'
#' @param df data.frame containing the outcome, treatment and covariate
#'   columns named consistently with `outcome`, `treatment`, `covariates`.
#' @param outcome,treatment,covariates character; column names.
#' @param B integer number of bootstrap samples.
#' @param level confidence level (default 0.95).
#'
#' @return a list with elements
#'   * `boot_betas`: B-by-(d+1) matrix of bootstrap coefficient estimates;
#'   * `percentile_ci`: (d+1)-by-2 matrix of percentile CIs;
#'   * `normal_ci`: (d+1)-by-2 matrix of mean +/- z * sd CIs.
#'
#' @keywords internal
dina_bootstrap_cis <- function(df,
                               outcome    = "Y",
                               treatment  = "treat",
                               covariates,
                               B          = 100L,
                               level      = 0.95) {

  # `forestsearch::dina()` -> `.dina_extract_from_df()` uses base-R
  # data.frame indexing idioms (`df[, cols]`, `df[, cols, drop = FALSE]`)
  # that misbehave on data.table inputs (the third arg is interpreted as
  # `by =`).  Coerce to a plain data.frame up-front so all internal row
  # resamples below are also data.frames.
  df <- as.data.frame(df)
  n  <- nrow(df)
  p  <- length(covariates) + 1L                    # +1 for intercept
  out_names <- c("(Intercept)", covariates)

  boot_betas <- matrix(NA_real_, nrow = B, ncol = p,
                       dimnames = list(NULL, out_names))

  # Track failures so the bootstrap is resilient to single bad resamples
  # (e.g., a resample with insufficient treatment-arm balance) while
  # still surfacing the cause loudly if many or all fits fail.
  n_err   <- 0L
  err_msg <- NULL

  for (b in seq_len(B)) {
    idx  <- sample.int(n, size = n, replace = TRUE)
    df_b <- df[idx, , drop = FALSE]
    fit_b <- tryCatch(
      forestsearch::dina(
        df         = df_b,
        outcome    = outcome,
        treatment  = treatment,
        covariates = covariates,
        family     = "gaussian"
      ),
      error = function(e) {
        n_err   <<- n_err + 1L
        if (is.null(err_msg)) err_msg <<- conditionMessage(e)
        NULL
      }
    )
    if (!is.null(fit_b)) {
      cf <- stats::coef(fit_b)
      boot_betas[b, names(cf)] <- as.numeric(cf)
    }
  }

  if (n_err == B) {
    stop(sprintf(
      "dina_bootstrap_cis(): all %d bootstrap fits failed. First error: %s",
      B, err_msg), call. = FALSE)
  } else if (n_err > 0L) {
    warning(sprintf(
      "dina_bootstrap_cis(): %d of %d bootstrap fits failed. First error: %s",
      n_err, B, err_msg), call. = FALSE)
  }

  alpha <- 1 - level
  qs    <- c(alpha / 2, 1 - alpha / 2)

  percentile_ci <- t(apply(boot_betas, 2L, stats::quantile,
                           probs = qs, na.rm = TRUE))
  colnames(percentile_ci) <- c("lower", "upper")

  boot_means <- colMeans(boot_betas, na.rm = TRUE)
  boot_sds   <- apply(boot_betas, 2L, stats::sd, na.rm = TRUE)
  z          <- stats::qnorm(1 - alpha / 2)
  normal_ci  <- cbind(
    lower = boot_means - z * boot_sds,
    upper = boot_means + z * boot_sds
  )

  list(
    boot_betas    = boot_betas,
    percentile_ci = percentile_ci,
    normal_ci     = normal_ci
  )
}
```

# Helper: one simulation trial {#sec-trial}

```{r trial-helper}
#' Run one trial of the Table-2 simulation
#'
#' Generates a fresh dataset, fits `dina()` on the full sample (for the
#' point estimate), runs `B` bootstrap CIs, and returns a long-format
#' data.table with one row per coefficient and per CI method.
#'
#' @keywords internal
run_one_trial <- function(trial_id, dgm, sim, true_beta, x_names) {

  df <- generate_dina_gaussian(n = sim$n_sample, dgm = dgm, seed = NULL)

  # Full-sample point estimate (the beta-hat that the bootstrap CIs are
  # built around; needed for the empirical-SE diagnostic).  Coerce to
  # data.frame because forestsearch::dina() uses base-R data.frame
  # indexing internally; see comment in dina_bootstrap_cis().
  fit_full      <- forestsearch::dina(
    df         = as.data.frame(df),
    outcome    = "Y",
    treatment  = "treat",
    covariates = x_names,
    family     = "gaussian"
  )
  beta_hat_full <- stats::coef(fit_full)

  boot <- dina_bootstrap_cis(
    df         = df,
    outcome    = "Y",
    treatment  = "treat",
    covariates = x_names,
    B          = sim$n_boot,
    level      = sim$level
  )

  # Per-coefficient within-trial bootstrap SE (the SD of the B bootstrap
  # beta-hats around their mean).  This is what an honest bootstrap
  # variance estimator returns for each coefficient.
  boot_se <- apply(boot$boot_betas, 2L, stats::sd, na.rm = TRUE)

  # Build a long-format result: one row per (coefficient, CI method).
  make_rows <- function(ci_mat, method) {
    coef_names <- rownames(ci_mat)
    data.table::data.table(
      trial    = trial_id,
      coef     = coef_names,
      method   = method,
      beta_hat = beta_hat_full[coef_names],
      boot_se  = boot_se[coef_names],
      lower    = ci_mat[, "lower"],
      upper    = ci_mat[, "upper"],
      width    = ci_mat[, "upper"] - ci_mat[, "lower"],
      truth    = true_beta[coef_names],
      covered  = (true_beta[coef_names] >= ci_mat[, "lower"]) &
                 (true_beta[coef_names] <= ci_mat[, "upper"])
    )
  }

  rbind(
    make_rows(boot$percentile_ci, "percentile"),
    make_rows(boot$normal_ci,     "normal")
  )
}
```

# Sanity check: single trial {#sec-sanity}

Before launching the full simulation, run one trial and verify the
machinery returns sensible numbers.

```{r sanity}
x_names <- paste0("X", seq_len(DGM$d))

t0 <- proc.time()
sanity_res <- run_one_trial(
  trial_id  = 0L,
  dgm       = DGM,
  sim       = SIM,
  true_beta = TRUE_BETA,
  x_names   = x_names
)
t_sanity <- (proc.time() - t0)["elapsed"]

cat(sprintf("Sanity trial completed in %.2f seconds.\n", t_sanity))
print(sanity_res)
```

A typical trial should take a few seconds. Total expected runtime for
`SIM$n_trials` trials with `SIM$n_workers` parallel workers is roughly
`SIM$n_trials / SIM$n_workers * t_sanity` seconds.

# Simulation driver {#sec-driver}

Parallelize over trials. Each worker generates its own data and runs `B`
bootstrap reps internally. Errors are captured and surfaced by
`forestsearch::collect_results()`.

```{r run-sim}
#| message: false

future::plan(future::multisession, workers = SIM$n_workers)

t_run_start <- proc.time()

raw_results <- foreach::foreach(
  trial = seq_len(SIM$n_trials),
  .options.future = list(seed = SIM$seed_base)
) %dofuture% {
  tryCatch(
    run_one_trial(
      trial_id  = trial,
      dgm       = DGM,
      sim       = SIM,
      true_beta = TRUE_BETA,
      x_names   = x_names
    ),
    error = function(e) e
  )
}

# Return parallel resources before downstream sequential work.
future::plan("sequential")
invisible(gc())

results <- forestsearch::collect_results(
  raw_list = raw_results,
  label    = "DINA Table 2 sim"
)

# collect_results() returns a data.frame; coerce to data.table so the
# downstream `[, .(...), by = .()]` idioms dispatch correctly.
results <- data.table::as.data.table(results)

t_run <- (proc.time() - t_run_start)["elapsed"]
cat(sprintf("Completed %d trials in %.1f minutes (%.1f s / trial avg).\n",
            SIM$n_trials, t_run / 60, t_run / SIM$n_trials))
```

# Results summary {#sec-summary}

Compute mean coverage and mean width per coefficient and per CI method,
then assemble the comparison table against the paper.

```{r summarise}
summary_dt <- results[
  , .(coverage = mean(covered, na.rm = TRUE),
      width    = mean(width,   na.rm = TRUE)),
  by = .(coef, method)
]

# Reshape to wide format: one row per coefficient, columns for each method.
summary_wide <- data.table::dcast(
  summary_dt,
  coef ~ method,
  value.var = c("coverage", "width")
)

# Join the paper reference values
summary_wide <- merge(summary_wide, PAPER_REF, by = "coef", all.x = TRUE)

# Order rows as in the paper: X1, X2, X3, X4, X5, (Intercept)
row_order <- c("X1", "X2", "X3", "X4", "X5", "(Intercept)")
summary_wide <- summary_wide[match(row_order, coef)]

# Add the true beta column
summary_wide[, truth := TRUE_BETA[coef]]

# Reorder columns for display
summary_wide <- summary_wide[
  , .(coef, truth,
      coverage_percentile, width_percentile,
      coverage_normal,     width_normal,
      cov_ref, wid_ref)
]
```

```{r summary-table}
gt::gt(summary_wide) |>
  gt::tab_header(
    title    = gt::md("**DINA Table 2 Reproduction: Gaussian, n = 1024**"),
    subtitle = gt::md(
      sprintf("Averaged over %d trials, %d bootstrap reps per trial",
              SIM$n_trials, SIM$n_boot)
    )
  ) |>
  gt::cols_label(
    coef                = "Coefficient",
    truth               = "True \u03b2",
    coverage_percentile = "Coverage",
    width_percentile    = "Width",
    coverage_normal     = "Coverage",
    width_normal        = "Width",
    cov_ref             = "Coverage",
    wid_ref             = "Width"
  ) |>
  gt::tab_spanner(label = "Percentile CI (this run)",
                  columns = c(coverage_percentile, width_percentile)) |>
  gt::tab_spanner(label = "Normal CI (this run)",
                  columns = c(coverage_normal, width_normal)) |>
  gt::tab_spanner(label = "Paper Table 2 (DINA)",
                  columns = c(cov_ref, wid_ref)) |>
  gt::fmt_number(columns = c(truth,
                             coverage_percentile, coverage_normal, cov_ref),
                 decimals = 2) |>
  gt::fmt_number(columns = c(width_percentile, width_normal, wid_ref),
                 decimals = 3) |>
  gt::tab_source_note(gt::md(paste0(
    "Paper reference: Gao and Hastie (2025), *Biometrics* 81(4): ujaf162, ",
    "Supplementary Materials Table 2 (Gaussian responses, n = 1024)."
  )))
```

# SE diagnostic: empirical vs. bootstrap SE {#sec-se-diag}

If the bootstrap is honest, the mean within-trial bootstrap SE should
match the empirical SE (the SD of the per-trial point estimates across
the 100 trials). The paper-implied SE comes from inverting Table 2's
width: $\text{SE} \approx \text{width} / (2 \cdot 1.96)$.

```{r se-diagnostic}
# Use percentile rows only.  beta_hat and boot_se are identical across
# CI methods within a trial; deduplicating avoids double-counting.
se_compare <- results[
  method == "percentile",
  .(empirical_se   = stats::sd(beta_hat,  na.rm = TRUE),
    mean_boot_se   = mean(boot_se,        na.rm = TRUE),
    mean_point_est = mean(beta_hat,       na.rm = TRUE),
    truth          = truth[1]),
  by = coef
]

# Paper-implied SE: invert the normal-CI width formula
z <- stats::qnorm(0.975)
se_compare <- merge(se_compare, PAPER_REF, by = "coef", all.x = TRUE)
se_compare[, paper_implied_se := wid_ref / (2 * z)]

# Order rows as in the paper
row_order <- c("X1", "X2", "X3", "X4", "X5", "(Intercept)")
se_compare <- se_compare[match(row_order, coef)]

se_compare <- se_compare[
  , .(coef, truth, mean_point_est,
      empirical_se, mean_boot_se, paper_implied_se)
]

gt::gt(se_compare) |>
  gt::tab_header(
    title    = gt::md("**SE diagnostic: empirical vs. bootstrap vs. paper-implied**"),
    subtitle = gt::md(paste0(
      "Empirical SE = sd(\u03b2-hat) across ", SIM$n_trials,
      " trials. Mean bootstrap SE = mean across trials of the ",
      "within-trial sd of ", SIM$n_boot, " bootstrap \u03b2-hats. ",
      "Paper-implied SE = Table 2 width / (2 \u00d7 1.96)."
    ))
  ) |>
  gt::cols_label(
    coef             = "Coefficient",
    truth            = "True \u03b2",
    mean_point_est   = "Mean \u03b2\u0302",
    empirical_se     = "Empirical SE",
    mean_boot_se     = "Mean bootstrap SE",
    paper_implied_se = "Paper-implied SE"
  ) |>
  gt::fmt_number(columns = c(truth, mean_point_est,
                             empirical_se, mean_boot_se,
                             paper_implied_se),
                 decimals = 3) |>
  gt::tab_source_note(gt::md(paste0(
    "If **empirical \u2248 mean bootstrap**, the bootstrap variance ",
    "estimator is unbiased for the estimator's true sampling variance. ",
    "If **both \u003c paper-implied**, the estimator itself is less ",
    "variable in this implementation than in the paper's runs."
  )))
```



Distribution of coverage and width across trials, to surface any
trial-level outliers that the averaged table conceals.

```{r diagnostics}
diag_dt <- results[
  , .(coverage_per_trial = mean(covered, na.rm = TRUE),
      width_per_trial    = mean(width,   na.rm = TRUE)),
  by = .(trial, method)
]

diag_summary <- diag_dt[
  , .(min_cov  = min(coverage_per_trial),
      max_cov  = max(coverage_per_trial),
      sd_cov   = stats::sd(coverage_per_trial),
      mean_wid = mean(width_per_trial),
      sd_wid   = stats::sd(width_per_trial)),
  by = method
]

gt::gt(diag_summary) |>
  gt::tab_header(
    title = gt::md("**Per-trial coverage and width spread**")
  ) |>
  gt::fmt_number(columns = c(min_cov, max_cov, sd_cov,
                             mean_wid, sd_wid),
                 decimals = 3)
```

# Session info {#sec-session}

```{r session-info}
sessionInfo()
```

# References {#sec-references}

Gao Z, Hastie T (2025). "Estimating heterogeneous treatment effects for
general responses." *Biometrics* 81(4): ujaf162.
[doi:10.1093/biomtc/ujaf162](https://doi.org/10.1093/biomtc/ujaf162).
