---
title: "DINA Within-Subgroup Standard-Model Smoke Test"
subtitle: "End-to-end check of dina_subgroup_refit() and the within-subgroup bootstrap CI across all four families"
author: "forestsearch v0.2.0"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

# Purpose

Smoke-test the *standard within-subgroup treatment-effect model* that
complements DINA's BLP-analog effect:

* `forestsearch::dina_subgroup_refit()` --- fits a plain Cox model (Cox
  family) or GLM (otherwise) comparing the treatment arms inside a
  subgroup discovered by `dina_subgroup()`, with optional confounders
  and Cox stratification.
* `forestsearch::dina_subgroup_bootstrap(refit = TRUE)` --- reports that
  standard model on the original data and bootstraps it within the
  *fixed* discovered signature, giving a within-signature percentile CI
  directly comparable to the DINA `effect_ci`.

The DGM uses a deliberately strong treatment-effect contrast on `x1`
($\beta_1 = 1.2$) so a harm subgroup at high `x1` genuinely exists and
the within-subgroup treatment effect is reliably positive (harm) on the
natural-parameter scale (log-HR / log-OR / log-IRR / MD).

This is a structural + math-sanity smoke test on a single seed per
family --- not a Monte Carlo coverage study. Assertions accumulate via
the standard `.tc()` helper; a gt pass/fail summary renders at the
bottom.

Key checks per family:

1. `dina_subgroup_refit()` runs, returns a well-formed
   `"dina_subgroup_refit"` object, with a positive SE, ordered CI, and
   the correct effect scale / ratio flag.
2. The three `confounders` modes resolve correctly: `"none"`
   (unadjusted), `NULL` (DINA covariates minus the subgroup factor),
   explicit vector.
3. Cox stratification works (Cox family); `strata` errors for non-Cox
   families (negative control).
4. Supplying the subgroup covariate as an explicit confounder warns.
5. The bootstrap reports `refit`, `refit_effect_ci`, `refit_effect_dist`
   consistently, the point refit matches a standalone
   `dina_subgroup_refit()` call, and `refit = FALSE` suppresses them.

---

# Setup

```{r setup}
library(forestsearch)
library(survival)
library(data.table)
library(gt)

set.seed(20260529)

SMOKE <- list(
  n               = 800L,
  d               = 5L,
  n_boot          = 60L,        # small; this is a smoke test, not a study
  n_folds         = 2L,
  parallel        = "boots",    # "boots" or "none"
  n_workers_frac  = 0.9,
  n_workers       = NULL,
  n_min           = 60L,
  alpha           = 0.05,
  cox_T0          = 1.0
)

# Strong heterogeneity on x1 only; harm subgroup lives at high x1.
BETA_TRUE <- c(0.40, 1.20, 0.00, 0.00, 0.00, 0.00)
stopifnot(length(BETA_TRUE) == SMOKE$d + 1L)

# m_diff positioned in the right tail so a right-direction subgroup on x1
# is the expected discovery (same convention as the other DINA smoke
# tests).
M_DIFF_BY_FAMILY <- c(
  gaussian = 0.80,
  binomial = 0.80,
  poisson  = 0.80,
  cox      = 0.80
)
```

## `.tc()` accumulating helper

```{r tc-helper}
.tc_env <- new.env(parent = emptyenv())
.tc_env$rows <- list()

.tc <- function(label, condition, family, details = "") {
  passed <- isTRUE(tryCatch(condition, error = function(e) FALSE))
  msg <- if (!passed && nzchar(details)) details
         else if (!passed) "condition FALSE or errored"
         else ""
  .tc_env$rows[[length(.tc_env$rows) + 1L]] <- data.table(
    family  = family,
    label   = label,
    status  = if (passed) "PASS" else "FAIL",
    details = msg
  )
  invisible(passed)
}

.tc_table <- function() {
  if (length(.tc_env$rows) == 0L) {
    return(data.table(family = character(), label = character(),
                      status = character(), details = character()))
  }
  rbindlist(.tc_env$rows)
}

`%||%` <- function(a, b) if (is.null(a)) b else a

# Did evaluating `expr` raise a warning matching `pattern`?
gave_warning <- function(expr, pattern) {
  w <- FALSE
  withCallingHandlers(
    tryCatch(force(expr), error = function(e) NULL),
    warning = function(cnd) {
      if (grepl(pattern, conditionMessage(cnd))) w <<- TRUE
      invokeRestart("muffleWarning")
    }
  )
  w
}

# Did evaluating `expr` raise an error?
gave_error <- function(expr) {
  inherits(tryCatch(force(expr), error = function(e) e), "error")
}
```

## Parallel plan setup

```{r plan-setup}
if (SMOKE$parallel == "boots") {
  total_cores <- parallelly::availableCores()
  n_workers <- if (is.null(SMOKE$n_workers)) {
    max(1L, as.integer(floor(SMOKE$n_workers_frac * total_cores)))
  } else {
    as.integer(SMOKE$n_workers)
  }
  future::plan(future::multisession, workers = n_workers)
  message(sprintf(
    "Parallel plan: %d workers (%.0f%% of %d available cores)",
    n_workers, 100 * n_workers / total_cores, total_cores
  ))
} else {
  future::plan("sequential")
  message("Parallel plan: sequential")
}
```

# Data simulation

Same DGM as the other DINA smoke tests, with a discrete `site` column
added so Cox stratification can be exercised.

```{r data-helpers}
simulate_dina_data <- function(family, n, d, beta_true, cox_T0 = 1.0) {
  X <- matrix(stats::runif(n * d, -1, 1), n, d)
  colnames(X) <- paste0("x", seq_len(d))

  e_x  <- stats::plogis(0.3 * X[, 1])
  W    <- stats::rbinom(n, 1L, e_x)

  mu_x  <- 0.2 * X[, 1] - 0.1 * X[, 2]
  tau_x <- as.numeric(beta_true[1L] + X %*% beta_true[-1L])
  eta_w <- mu_x + W * tau_x

  out <- list(X = X, W = W, beta_true = beta_true, eta_w = eta_w,
              tau_x_true = tau_x)

  out$Y <- switch(
    family,
    gaussian = as.numeric(eta_w + stats::rnorm(n)),
    binomial = stats::rbinom(n, 1L, stats::plogis(eta_w)),
    poisson  = stats::rpois(n, exp(pmin(eta_w, 5))),
    cox = {
      t_event  <- stats::rexp(n, rate = exp(eta_w))
      t_censor <- stats::runif(n, 0, cox_T0)
      time   <- pmin(t_event, t_censor)
      status <- as.integer(t_event <= t_censor)
      data.frame(time = time, status = status)
    }
  )
  out
}

build_df <- function(dat, family) {
  site <- factor(sample(c("A", "B"), nrow(dat$X), replace = TRUE))
  if (family == "cox") {
    cbind(
      data.frame(time = dat$Y$time, status = dat$Y$status,
                 w = dat$W, site = site),
      as.data.frame(dat$X)
    )
  } else {
    cbind(
      data.frame(y = dat$Y, w = dat$W, site = site),
      as.data.frame(dat$X)
    )
  }
}
```

## Per-family runner

```{r runner}
run_refit_family_tests <- function(family) {
  m_diff <- M_DIFF_BY_FAMILY[[family]]

  dat <- simulate_dina_data(
    family = family, n = SMOKE$n, d = SMOKE$d,
    beta_true = BETA_TRUE, cox_T0 = SMOKE$cox_T0
  )
  df_fam       <- build_df(dat, family)
  outcome_col  <- if (family == "cox") "time" else "y"
  status_col   <- if (family == "cox") "status" else NULL
  covariates_v <- colnames(dat$X)

  cens_type   <- if (family == "cox") "uniform" else NULL
  cens_params <- if (family == "cox") list(T0 = SMOKE$cox_T0) else list()

  # ---- Upstream DINA fit + subgroup ---------------------------------------
  fit <- tryCatch(
    dina(df = df_fam, outcome = outcome_col, treatment = "w",
         covariates = covariates_v, family = family, status = status_col,
         n_folds = SMOKE$n_folds, cens_type = cens_type,
         cens_params = cens_params, seed = 20260529L),
    error = function(e) NULL
  )
  .tc("dina() completes", !is.null(fit), family)

  sg <- if (!is.null(fit)) {
    tryCatch(
      dina_subgroup(fit = fit, df = df_fam, covariates = covariates_v,
                    m_diff = m_diff, n_min = SMOKE$n_min, alpha = SMOKE$alpha),
      error = function(e) NULL
    )
  } else NULL
  .tc("dina_subgroup() completes", !is.null(sg), family)
  found <- !is.null(sg) && isTRUE(sg$found)
  .tc("dina_subgroup() found a subgroup at m_diff", found, family,
      details = sprintf("found = %s", if (is.null(sg)) "NULL"
                        else as.character(sg$found)))

  scale_expected <- switch(family,
    cox = "log-HR", binomial = "log-OR",
    poisson = "log-IRR", gaussian = "MD")
  ratio_expected <- family != "gaussian"

  rf_none <- NULL

  if (found) {
    # ---- (1) Unadjusted refit (default) -----------------------------------
    rf_none <- tryCatch(
      dina_subgroup_refit(
        sg = sg, df = df_fam, treatment = "w", outcome = outcome_col,
        covariates = covariates_v, status = status_col,
        confounders = "none", alpha = SMOKE$alpha
      ),
      error = function(e) {
        attr(e, "details") <- conditionMessage(e); e
      }
    )
    ok <- !inherits(rf_none, "error")
    .tc("dina_subgroup_refit() (unadjusted) completes", ok, family,
        details = if (!ok) attr(rf_none, "details") else "")

    if (ok) {
      .tc("returns object of class 'dina_subgroup_refit'",
          inherits(rf_none, "dina_subgroup_refit"), family)
      .tc("effect is finite", is.finite(rf_none$effect), family)
      .tc("SE is positive and finite",
          is.finite(rf_none$se) && rf_none$se > 0, family)
      .tc("Wald CI ordered (lower <= effect <= upper)",
          rf_none$ci[["lower"]] <= rf_none$effect &&
            rf_none$effect <= rf_none$ci[["upper"]], family)
      .tc("effect_scale matches family",
          identical(rf_none$effect_scale, scale_expected), family,
          details = sprintf("got %s, expected %s",
                            rf_none$effect_scale, scale_expected))
      .tc("ratio_scale matches family",
          identical(rf_none$ratio_scale, ratio_expected), family)
      .tc("unadjusted: confounders_used is empty",
          length(rf_none$confounders_used) == 0L, family)
      .tc("both arms non-empty in subgroup",
          rf_none$n_treated > 0L && rf_none$n_control > 0L, family,
          details = sprintf("trt %d, ctrl %d",
                            rf_none$n_treated, rf_none$n_control))
      .tc("n_subgroup matches sg$n_subgroup",
          identical(as.integer(rf_none$n_subgroup),
                    as.integer(sg$n_subgroup)), family)
      if (family == "cox") {
        .tc("Cox: event counts present and consistent",
            is.finite(rf_none$n_events) &&
              rf_none$n_events ==
                rf_none$n_events_treated + rf_none$n_events_control,
            family)
      } else {
        .tc("non-Cox: event counts are NA",
            is.na(rf_none$n_events), family)
      }
      # Directional sanity: harm subgroup at high x1 => positive effect.
      .tc("effect has expected (harm) sign: > 0",
          rf_none$effect > 0, family,
          details = sprintf("effect = %.3f", rf_none$effect))
      .tc("print() runs without error",
          !inherits(tryCatch(utils::capture.output(print(rf_none)),
                             error = function(e) e), "error"), family)
    }

    # ---- (2) Auto confounders (NULL) --------------------------------------
    rf_auto <- tryCatch(
      dina_subgroup_refit(
        sg = sg, df = df_fam, treatment = "w", outcome = outcome_col,
        covariates = covariates_v, status = status_col,
        confounders = NULL, alpha = SMOKE$alpha
      ),
      error = function(e) NULL
    )
    .tc("auto-confounders (NULL) completes", !is.null(rf_auto), family)
    if (!is.null(rf_auto)) {
      expected_auto <- setdiff(covariates_v, sg$covariate)
      .tc("auto confounders == DINA covariates minus subgroup factor",
          setequal(rf_auto$confounders_used, expected_auto), family,
          details = sprintf("used {%s}; expected {%s}",
                            paste(rf_auto$confounders_used, collapse = ","),
                            paste(expected_auto, collapse = ",")))
    }

    # ---- (3) Explicit confounders -----------------------------------------
    expl <- setdiff(covariates_v, sg$covariate)[1:2]
    rf_expl <- tryCatch(
      dina_subgroup_refit(
        sg = sg, df = df_fam, treatment = "w", outcome = outcome_col,
        covariates = covariates_v, status = status_col,
        confounders = expl, alpha = SMOKE$alpha
      ),
      error = function(e) NULL
    )
    .tc("explicit confounders completes", !is.null(rf_expl), family)
    if (!is.null(rf_expl)) {
      .tc("explicit confounders_used == specified set",
          setequal(rf_expl$confounders_used, expl), family)
    }

    # ---- (4) Warning when subgroup covariate is an explicit confounder ----
    .tc("warns when explicit confounders include the subgroup covariate",
        gave_warning(
          dina_subgroup_refit(
            sg = sg, df = df_fam, treatment = "w", outcome = outcome_col,
            covariates = covariates_v, status = status_col,
            confounders = c(sg$covariate, expl[1]), alpha = SMOKE$alpha
          ),
          pattern = "subgroup-defining covariate"
        ), family)

    # ---- (5) Strata: works for Cox, errors otherwise ----------------------
    if (family == "cox") {
      rf_strata <- tryCatch(
        dina_subgroup_refit(
          sg = sg, df = df_fam, treatment = "w", outcome = outcome_col,
          covariates = covariates_v, status = status_col,
          strata = "site", confounders = "none", alpha = SMOKE$alpha
        ),
        error = function(e) NULL
      )
      .tc("Cox: strata = 'site' completes", !is.null(rf_strata), family)
      if (!is.null(rf_strata)) {
        .tc("Cox: strata_used records the strata column",
            setequal(rf_strata$strata_used, "site"), family)
        .tc("Cox: stratified effect finite and SE positive",
            is.finite(rf_strata$effect) && rf_strata$se > 0, family)
      }
    } else {
      .tc("non-Cox: strata raises an error (negative control)",
          gave_error(
            dina_subgroup_refit(
              sg = sg, df = df_fam, treatment = "w", outcome = outcome_col,
              covariates = covariates_v, status = status_col,
              strata = "site", confounders = "none", alpha = SMOKE$alpha
            )
          ), family)
    }
  }

  # ---- (6) Negative control: refit on a not-found subgroup errors ---------
  sg_null <- if (!is.null(fit)) {
    tryCatch(
      dina_subgroup(fit = fit, df = df_fam, covariates = covariates_v,
                    m_diff = 100, n_min = SMOKE$n_min, alpha = SMOKE$alpha),
      error = function(e) NULL
    )
  } else NULL
  if (!is.null(sg_null)) {
    .tc("refit on a not-found subgroup raises an error (negative control)",
        gave_error(
          dina_subgroup_refit(
            sg = sg_null, df = df_fam, treatment = "w", outcome = outcome_col,
            covariates = covariates_v, status = status_col
          )
        ), family)
  }

  # ---- (7) Bootstrap integration ----------------------------------------
  bs <- if (found) {
    tryCatch(
      dina_subgroup_bootstrap(
        df = df_fam, outcome = outcome_col, treatment = "w",
        covariates = covariates_v, family = family, status = status_col,
        m_diff = m_diff, n_min = SMOKE$n_min, alpha = SMOKE$alpha,
        n_boot = SMOKE$n_boot, parallel = SMOKE$parallel,
        n_folds = SMOKE$n_folds, cens_type = cens_type,
        cens_params = cens_params, seed = 20260529L,
        fit = fit, sg = sg,
        refit = TRUE, refit_confounders = "none", refit_strata = NULL
      ),
      error = function(e) {
        attr(e, "details") <- conditionMessage(e); e
      }
    )
  } else NULL
  bs_ok <- !is.null(bs) && !inherits(bs, "error")
  .tc("dina_subgroup_bootstrap(refit = TRUE) completes", bs_ok, family,
      details = if (!is.null(bs) && inherits(bs, "error"))
        attr(bs, "details") else "")

  if (bs_ok) {
    .tc("bs$refit is a 'dina_subgroup_refit' object",
        inherits(bs$refit, "dina_subgroup_refit"), family)
    .tc("bs$refit_effect_ci is named length-2 numeric (lower, upper)",
        is.numeric(bs$refit_effect_ci) &&
          identical(names(bs$refit_effect_ci), c("lower", "upper")), family)
    .tc("bs$refit_effect_dist has length n_boot",
        length(bs$refit_effect_dist) == SMOKE$n_boot, family)

    # Point refit inside the bootstrap matches a standalone refit call.
    if (!is.null(rf_none) && !inherits(rf_none, "error") &&
        inherits(bs$refit, "dina_subgroup_refit")) {
      .tc("bootstrap point refit == standalone dina_subgroup_refit()",
          isTRUE(all.equal(bs$refit$effect, rf_none$effect)), family,
          details = sprintf("boot %.5f vs standalone %.5f",
                            bs$refit$effect, rf_none$effect))
    }

    refit_finite <- bs$refit_effect_dist[is.finite(bs$refit_effect_dist)]
    if (length(refit_finite) >= 2L) {
      .tc("refit bootstrap CI ordered and finite",
          all(is.finite(bs$refit_effect_ci)) &&
            bs$refit_effect_ci[["lower"]] <= bs$refit_effect_ci[["upper"]],
          family,
          details = sprintf("[%.3f, %.3f]",
                            bs$refit_effect_ci[["lower"]],
                            bs$refit_effect_ci[["upper"]]))
    }
  }

  # ---- (8) refit = FALSE suppresses the within-subgroup outputs ----------
  bs_no <- if (found) {
    tryCatch(
      dina_subgroup_bootstrap(
        df = df_fam, outcome = outcome_col, treatment = "w",
        covariates = covariates_v, family = family, status = status_col,
        m_diff = m_diff, n_min = SMOKE$n_min, alpha = SMOKE$alpha,
        n_boot = 20L, parallel = SMOKE$parallel,
        n_folds = SMOKE$n_folds, cens_type = cens_type,
        cens_params = cens_params, seed = 20260529L,
        fit = fit, sg = sg, refit = FALSE
      ),
      error = function(e) NULL
    )
  } else NULL
  if (!is.null(bs_no)) {
    .tc("refit = FALSE: bs$refit is NULL",
        is.null(bs_no$refit), family)
    .tc("refit = FALSE: refit_effect_ci is (NA, NA)",
        all(is.na(bs_no$refit_effect_ci)), family)
  }

  list(family = family, m_diff = m_diff, sg = sg, rf_none = rf_none, bs = bs)
}

# Per-family gt: DINA effect vs within-subgroup standard-model effect
family_refit_gt <- function(res, family) {
  sg <- res$sg
  rf <- res$rf_none
  bs <- res$bs

  fmt_ci <- function(ci) {
    if (is.null(ci) || any(!is.finite(ci))) return("—")
    sprintf("[%.3f, %.3f]", ci[["lower"]], ci[["upper"]])
  }

  sig <- if (!is.null(sg) && isTRUE(sg$found)) {
    cmp <- if (sg$direction == "left") "<=" else ">="
    sprintf("%s %s %.3f", sg$covariate, cmp, sg$threshold)
  } else "—"

  dina_eff   <- if (!is.null(sg) && isTRUE(sg$found))
    sprintf("%.3f", sg$mean_tau_hat) else "—"
  refit_eff  <- if (!is.null(rf) && !inherits(rf, "error"))
    sprintf("%.3f", rf$effect) else "—"
  refit_wald <- if (!is.null(rf) && !inherits(rf, "error"))
    fmt_ci(rf$ci) else "—"
  refit_boot <- if (!is.null(bs) && !inherits(bs, "error"))
    fmt_ci(bs$refit_effect_ci) else "—"

  scale_lab <- switch(family, cox = "log-HR", binomial = "log-OR",
                      poisson = "log-IRR", gaussian = "MD")

  rows <- data.table(
    Field = c(
      "Signature",
      sprintf("DINA effect (%s)", scale_lab),
      sprintf("Within-subgroup standard effect (%s)", scale_lab),
      "Within-subgroup Wald CI",
      "Within-subgroup bootstrap CI"
    ),
    Value = c(sig, dina_eff, refit_eff, refit_wald, refit_boot)
  )

  rows |>
    gt::gt() |>
    gt::tab_header(
      title    = sprintf("%s family — DINA vs within-subgroup standard model",
                         tools::toTitleCase(family)),
      subtitle = sprintf("m_diff = %.2f; n_boot = %d", res$m_diff, SMOKE$n_boot)
    ) |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_source_note(
      source_note = gt::md(
        "*Both effects are on the natural-parameter scale. The DINA effect
         is the BLP-analog; the within-subgroup standard effect is a plain
         Cox/GLM contrast inside the signature. Both CIs are conditional on
         the discovered signature.*"
      )
    )
}
```

---

# Gaussian family

```{r family-gaussian}
res_gauss <- run_refit_family_tests("gaussian")
family_refit_gt(res_gauss, "gaussian")
```

---

# Binomial family

```{r family-binomial}
res_binom <- run_refit_family_tests("binomial")
family_refit_gt(res_binom, "binomial")
```

---

# Poisson family

```{r family-poisson}
res_pois <- run_refit_family_tests("poisson")
family_refit_gt(res_pois, "poisson")
```

---

# Cox family

```{r family-cox}
res_cox <- run_refit_family_tests("cox")
family_refit_gt(res_cox, "cox")
```

```{r plan-teardown}
future::plan("sequential")
gc(verbose = FALSE)
```

---

# Summary

```{r summary}
results_tab <- .tc_table()
n_total <- nrow(results_tab)
n_pass  <- results_tab[status == "PASS", .N]
n_fail  <- n_total - n_pass

subtitle_text <- sprintf(
  "%d / %d assertions passed (%d failures)", n_pass, n_total, n_fail
)

results_tab |>
  gt::gt(groupname_col = "family") |>
  gt::tab_header(
    title    = gt::md("**DINA within-subgroup standard-model smoke-test summary**"),
    subtitle = subtitle_text
  ) |>
  gt::cols_label(
    label   = "Assertion",
    status  = "Status",
    details = "Details"
  ) |>
  gt::tab_style(
    style    = gt::cell_fill(color = "#d4f4dd"),
    locations = gt::cells_body(rows = status == "PASS")
  ) |>
  gt::tab_style(
    style    = gt::cell_fill(color = "#f9d4d4"),
    locations = gt::cells_body(rows = status == "FAIL")
  ) |>
  gt::tab_style(
    style    = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |>
  gt::tab_options(table.font.size = 12)
```

---

# Session information

```{r session-info}
sessionInfo()
```
