---
title: "forestsearch() vs dina_subgroup(): Signature-Identification Comparison"
subtitle: "Matched harm-subgroup recovery on a single-factor threshold DGM"
author: "forestsearch v0.2.0"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
---

# Purpose

A head-to-head comparison of two harm-subgroup identifiers on a DGM with
a **known single-factor signature**:

* `forestsearch(sg_focus = "effMaxSG")` --- among candidates whose effect
  is within `effect_neighborhood` of the maximum, selects the one with
  the **largest** sample size. This is the forestsearch analog of DINA's
  "largest subgroup with contrast $\ge$ threshold."
* `dina_subgroup()` --- a univariate threshold search over the qualifying
  subgroups (mean $\hat\tau \ge$ `m_diff`). Run here under **two**
  selection rules: `maxSG` (the legacy largest-qualifying rule) and
  `effMaxSG` (the new effect-neighborhood band, mirroring forestsearch's
  `sg_focus`). Comparing the two isolates what the selection rule --- not
  the search --- contributes to signature recovery.

Both are run on the **same simulated data** each replication and scored
against the **same** ground-truth subgroup. The DGM is the signature-
simulation setup: covariates $\sim U[-1,1]$, logistic propensity, Cox
outcome via `rexp(rate = exp(eta))` with uniform censoring, and a true
harm subgroup defined by a single-covariate threshold
$S = \{x_1 \ge 0.30\}$ (treatment increases the hazard inside $S$).

## Matched thresholds (harm study)

The methods use the **same harm cutoff on the natural (HR) scale**:

| Method | Threshold argument | Value |
|---|---|---|
| `dina_subgroup()` | `m_diff` (log-HR scale) | `log(1.25) ~ 0.223` |
| `forestsearch()` | `effect.threshold` (HR scale) | `1.25` |

These satisfy `exp(m_diff) = effect.threshold = 1.25`, so "largest
subgroup with effect $\ge$ 1.25 (HR)" means the same number for both.

## Fairness: the `maxk` knob

The true signature is a **single factor**. `dina_subgroup()` searches
only univariate thresholds, so it is structurally single-factor.
`forestsearch()` can combine up to `maxk` factors. To compare like with
like, the default below sets **`fs_maxk = 1L`** (both methods restricted
to single factors); set `fs_maxk = 2L` to study what the extra
two-factor flexibility costs or buys against the single-factor truth.
The setting is recorded in every results table so each run is
self-documenting.

> This is a Monte Carlo study. `forestsearch()` is far heavier per
> replication than `dina_subgroup()` (GRF screening + bootstrap
> consistency), so the default grid and `nsims` are modest; scale up on
> a multi-core machine.

---

# Configuration

```{r setup}
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(doFuture)
library(foreach)

CONFIG <- list(
  # Monte Carlo
  nsims       = 1000L,         # replications per cell (scale up for production)
  base_seed   = 20260530L,

  # design grid (forestsearch is expensive; keep modest)
  d           = 5L,
  n_levels    = c(500L, 1000L),
  tau_in      = c(moderate = 0.8, strong = 1.3),  # harm log-HR inside S
  tau_out     = 0.0,
  true_cov    = "x1",
  true_dir    = "right",      # S = { x1 >= cut }
  true_cut    = 0.30,
  cox_T0      = 1.5,          # uniform-censoring upper bound (DGM and fits match)

  # ---- matched harm thresholds (natural scale) ----
  effect_threshold = 1.25,            # forestsearch HR-scale harm threshold
  m_diff           = log(1.25),       # dina_subgroup log-HR threshold = log(effect_threshold)

  # ---- forestsearch knobs ----
  fs_maxk            = 1L,    # 1 = single-factor (matched to DINA); 2 = allow 2-factor
  fs_sg_focus        = "effMaxSG",
  fs_eff_neighbor    = 0.10,
  fs_selection_rule  = "neighborhood",
  fs_consistency     = 1.0,   # hr.consistency
  fs_pconsistency    = 0.90,
  fs_splits          = 500L,  # lowered for tractability (paper uses 1000)
  fs_n_min           = 60L,
  fs_d0_min          = 10L,
  fs_d1_min          = 10L,
  fs_vi_grf_min      = -0.2,
  fs_use_grf         = TRUE,
  fs_use_lasso       = FALSE,
  fs_use_dina        = TRUE,
  fs_seedit          = 8316951L,

  # ---- dina_subgroup knobs ----
  dina_n_folds = 2L,
  dina_n_min   = 60L,
  dina_alpha   = 0.05,
  dina_foci         = c("effMaxSG", "maxSG"),  # legacy largest-qualifying vs new effect band
  dina_eff_neighbor = 0.10,                    # effect-neighborhood band (matches fs_eff_neighbor)

  # threshold-recovery tolerance for dina (signature axis is well-defined there)
  thr_tol      = 0.10,

  # parallel
  parallel       = "sims",    # "sims" or "none"
  n_workers_frac = 0.9,
  n_workers      = NULL
)

stopifnot(abs(exp(CONFIG$m_diff) - CONFIG$effect_threshold) < 1e-8)
stopifnot(CONFIG$fs_maxk %in% c(1L, 2L))
stopifnot(all(CONFIG$dina_foci %in%
              c("maxSG", "minSG", "eff", "effMaxSG", "effMinSG")))
# harm thresholds must exceed the no-effect level and sit below tau_in so S
# is detectable.
stopifnot(CONFIG$m_diff > CONFIG$tau_out, all(CONFIG$m_diff < CONFIG$tau_in))

`%||%` <- function(a, b) if (is.null(a)) b else a
```

## Parallel plan

```{r plan-setup}
if (CONFIG$parallel == "sims") {
  total_cores <- parallelly::availableCores()
  n_workers <- if (is.null(CONFIG$n_workers))
    max(1L, as.integer(floor(CONFIG$n_workers_frac * total_cores)))
  else as.integer(CONFIG$n_workers)
  future::plan(future::multisession, workers = n_workers)
  message(sprintf("Parallel plan: %d workers of %d cores", n_workers, total_cores))
} else {
  future::plan("sequential"); n_workers <- 1L
  message("Parallel plan: sequential")
}

timings <- list()
```

---

# Data-generating mechanism

A genuine single-factor harm subgroup via a step in the treatment
effect: $\tau(x) = \tau_{\text{in}}$ (harm) inside $S=\{x_1 \ge c\}$,
$\tau_{\text{out}} \approx 0$ outside. For the null arm both are zero
(no heterogeneity). The data frame carries the true membership column
`flag_harm` for scoring but does **not** expose the cut to either
method.

```{r dgm}
simulate_threshold_cox <- function(n, d, tau_in, tau_out,
                                    true_cov, true_dir, true_cut, cox_T0) {
  X <- matrix(stats::runif(n * d, -1, 1), n, d)
  colnames(X) <- paste0("x", seq_len(d))

  e_x <- stats::plogis(0.3 * X[, 1])
  W   <- stats::rbinom(n, 1L, e_x)

  in_S <- if (true_dir == "left") X[, true_cov] <= true_cut
          else X[, true_cov] >= true_cut
  tau_x <- tau_out + (tau_in - tau_out) * as.numeric(in_S)

  mu_x  <- 0.2 * X[, 1] - 0.1 * X[, 2]
  eta_w <- mu_x + W * tau_x

  t_event  <- stats::rexp(n, rate = exp(eta_w))
  t_censor <- stats::runif(n, 0, cox_T0)

  data.table(
    id        = seq_len(n),
    y_sim     = pmin(t_event, t_censor),
    event_sim = as.integer(t_event <= t_censor),
    treat_sim = W,
    flag_harm = as.integer(in_S),
    as.data.frame(X)
  )
}
```

---

# Scoring (common, method-agnostic)

Because `forestsearch()` may return a multi-factor rule while
`dina_subgroup()` returns a univariate threshold, both are scored on
**method-agnostic** quantities: detection, whether the discovered
subgroup involves the true covariate, and membership recovery vs the
true mask (sensitivity / specificity / PPV / Jaccard). Threshold
recovery (bias vs the true cut) is reported for `dina_subgroup()`, where
it is well defined.

```{r scoring}
mask_recovery <- function(disc_mask, true_mask) {
  tp <- sum(disc_mask & true_mask); fp <- sum(disc_mask & !true_mask)
  fn <- sum(!disc_mask & true_mask); tn <- sum(!disc_mask & !true_mask)
  safe <- function(a, b) if (b > 0) a / b else NA_real_
  list(sensitivity = safe(tp, tp + fn), specificity = safe(tn, tn + fp),
       ppv = safe(tp, tp + fp), npv = safe(tn, tn + fn),
       jaccard = safe(tp, tp + fp + fn))
}

# Empty/standard per-method result row (keeps rbindlist columns aligned).
empty_row <- function(method) data.table(
  method = method, found = FALSE,
  covariate_involved = NA, n_subgroup = NA_integer_, n_factors = NA_integer_,
  threshold = NA_real_, threshold_error = NA_real_,
  sensitivity = NA_real_, specificity = NA_real_, ppv = NA_real_,
  npv = NA_real_, jaccard = NA_real_, sg_def = NA_character_
)

# Score forestsearch: discovered membership is treat.recommend == 0 (-> H).
score_fs <- function(fs, df, truth) {
  out <- empty_row("FS")
  sg <- tryCatch(fs$grp.consistency$sg.harm, error = function(e) NULL)
  found_flag <- !is.null(sg) && length(sg) > 0 && !all(is.na(sg))
  out[, found := found_flag]
  if (!found_flag) {
    if (!truth$is_null) {
      # Detected nothing: everyone predicted Hc. sens=0, ppv=0 by
      # convention (no positives predicted); spec=1; npv=|Hc|/N; jaccard=0.
      true_H <- df[[truth$mask_col]] == 1L
      npv_none <- if (length(true_H)) mean(!true_H) else NA_real_
      out[, c("sensitivity","specificity","ppv","npv","jaccard") :=
        .(0, 1, 0, npv_none, 0)]
    }
    return(out)
  }
  sg_txt <- paste(sg[!is.na(sg)], collapse = " & ")
  out[, sg_def := sg_txt]
  out[, n_factors := sum(!is.na(sg))]
  out[, covariate_involved := grepl(truth$covariate, sg_txt, fixed = TRUE)]
  df_est <- tryCatch(fs$df.est, error = function(e) NULL)
  if (!is.null(df_est) && "treat.recommend" %in% names(df_est)) {
    disc <- df_est$treat.recommend == 0
    if (length(disc) == nrow(df) && !truth$is_null) {
      true_mask <- df[[truth$mask_col]] == 1L
      rec <- mask_recovery(disc, true_mask)
      out[, c("sensitivity","specificity","ppv","npv","jaccard") :=
            .(rec$sensitivity, rec$specificity, rec$ppv, rec$npv, rec$jaccard)]
      out[, n_subgroup := sum(disc)]
    }
  }
  out
}

# Score dina_subgroup.  `method_label` lets the same scorer tag rows for
# different `sg_focus` runs (e.g. "DINA-maxSG", "DINA-effMaxSG").
score_dina <- function(sg, df, truth, thr_tol, method_label = "DINA") {
  out <- empty_row(method_label)
  found_flag <- !is.null(sg) && isTRUE(sg$found)
  out[, found := found_flag]
  if (!found_flag) {
    if (!truth$is_null) {
      true_H <- df[[truth$mask_col]] == 1L
      npv_none <- if (length(true_H)) mean(!true_H) else NA_real_
      out[, c("sensitivity","specificity","ppv","npv","jaccard") :=
        .(0, 1, 0, npv_none, 0)]
    }
    return(out)
  }
  out[, n_subgroup := as.integer(sg$n_subgroup)]
  out[, n_factors := 1L]
  out[, covariate_involved := identical(sg$covariate, truth$covariate)]
  cmp <- if (sg$direction == "left") "<=" else ">="
  out[, sg_def := sprintf("%s %s %.3f", sg$covariate, cmp, sg$threshold)]
  out[, threshold := as.numeric(sg$threshold)]
  if (!truth$is_null) {
    if (identical(sg$covariate, truth$covariate) &&
        identical(sg$direction, truth$direction)) {
      out[, threshold_error := as.numeric(sg$threshold) - truth$threshold]
    }
    true_mask <- df[[truth$mask_col]] == 1L
    rec <- mask_recovery(as.logical(sg$mask), true_mask)
    out[, c("sensitivity","specificity","ppv","npv","jaccard") :=
          .(rec$sensitivity, rec$specificity, rec$ppv, rec$npv, rec$jaccard)]
  }
  out
}
```

## Scoring unit tests

```{r scoring-tests}
.ok <- function(lbl, cond) cat(if (isTRUE(cond)) "PASS" else "FAIL", "-", lbl, "\n")

tm <- c(rep(TRUE, 30), rep(FALSE, 70))
.ok("perfect mask sens=1", isTRUE(all.equal(mask_recovery(tm, tm)$sensitivity, 1)))
.ok("perfect mask spec=1", isTRUE(all.equal(mask_recovery(tm, tm)$specificity, 1)))

mock_df <- data.table(x1 = seq(-1, 1, length.out = 100),
                      flag_harm = as.integer(seq(-1, 1, length.out = 100) >= 0.30))
truth1 <- list(is_null = FALSE, covariate = "x1", direction = "right",
               threshold = 0.30, mask_col = "flag_harm")

mock_sg <- list(found = TRUE, covariate = "x1", direction = "right",
                threshold = 0.30, n_subgroup = sum(mock_df$flag_harm),
                mask = mock_df$x1 >= 0.30)
sd1 <- score_dina(mock_sg, mock_df, truth1, 0.10)
.ok("dina correct: covariate_involved TRUE", isTRUE(sd1$covariate_involved))
.ok("dina correct: threshold_error ~ 0", isTRUE(abs(sd1$threshold_error) < 1e-9))
.ok("dina correct: sensitivity = 1", isTRUE(all.equal(sd1$sensitivity, 1)))

sd0 <- score_dina(list(found = FALSE), mock_df, truth1, 0.10)
.ok("dina not-found: found FALSE", isFALSE(sd0$found))
.ok("dina not-found: specificity = 1 (all to Hc)", isTRUE(all.equal(sd0$specificity, 1)))
.ok("dina method label flows through",
    identical(score_dina(mock_sg, mock_df, truth1, 0.10, "DINA-effMaxSG")$method,
              "DINA-effMaxSG"))

mock_fs <- list(grp.consistency = list(sg.harm = c("x1.high")),
                df.est = data.frame(treat.recommend =
                  as.integer(!(mock_df$x1 >= 0.30))))
sf1 <- score_fs(mock_fs, mock_df, truth1)
.ok("fs correct: found TRUE", isTRUE(sf1$found))
.ok("fs correct: covariate_involved TRUE (x1 in sg)", isTRUE(sf1$covariate_involved))
.ok("fs correct: sensitivity = 1", isTRUE(all.equal(sf1$sensitivity, 1)))
.ok("fs correct: n_factors = 1", identical(as.integer(sf1$n_factors), 1L))

sf0 <- score_fs(list(grp.consistency = list(sg.harm = NA)), mock_df, truth1)
.ok("fs not-found: found FALSE", isFALSE(sf0$found))
```

---

# One replication: both methods on the same data

```{r one-rep}
run_one_rep <- function(scenario, rep_id, cfg) {
  seed <- cfg$base_seed + scenario$scenario_id * 100000L + rep_id
  set.seed(seed)

  df <- simulate_threshold_cox(
    n = scenario$n, d = cfg$d,
    tau_in = scenario$tau_in, tau_out = cfg$tau_out,
    true_cov = cfg$true_cov, true_dir = cfg$true_dir,
    true_cut = cfg$true_cut, cox_T0 = cfg$cox_T0)
  df <- as.data.frame(df)
  covariates_v <- paste0("x", seq_len(cfg$d))
  truth <- list(is_null = isTRUE(scenario$is_null),
                covariate = cfg$true_cov, direction = cfg$true_dir,
                threshold = cfg$true_cut, mask_col = "flag_harm")

  # ---- forestsearch (sg_focus = effMaxSG; maxk per config) ----
  fs <- tryCatch(
    suppressWarnings(forestsearch(
      df.analysis      = df,
      confounders.name = covariates_v,
      outcome.name     = "y_sim",
      event.name       = "event_sim",
      treat.name       = "treat_sim",
      id.name          = "id",
      est.scale        = "hr",
      effect.threshold = cfg$effect_threshold,
      hr.consistency   = cfg$fs_consistency,
      sg_focus         = cfg$fs_sg_focus,
      selection_rule        = cfg$fs_selection_rule,
      effect_neighborhood    = cfg$fs_eff_neighbor,
      pconsistency.threshold = cfg$fs_pconsistency,
      fs.splits        = cfg$fs_splits,
      n.min            = cfg$fs_n_min,
      d0.min           = cfg$fs_d0_min,
      d1.min           = cfg$fs_d1_min,
      maxk             = cfg$fs_maxk,
      vi.grf.min       = cfg$fs_vi_grf_min,
      use_grf          = cfg$fs_use_grf,
      use_lasso        = cfg$fs_use_lasso,
      use_dina         = cfg$fs_use_dina,
      use_twostage     = TRUE,
      is.RCT           = TRUE,
      details          = FALSE,
      quiet            = TRUE,
      seedit           = cfg$fs_seedit,
      parallel_args    = list(plan = "sequential", workers = 1,
                              show_message = FALSE)
    )),
    error = function(e) NULL)
  fs_row <- score_fs(fs, df, truth)

  # ---- dina_subgroup (one DINA fit; scored under each sg_focus) ----
  # The expensive step is the cross-fitted dina() fit; the threshold
  # search under each focus is cheap, so all foci reuse the same fit.
  # `maxSG` is the legacy largest-qualifying rule; `effMaxSG` applies the
  # new effect-neighborhood band (the selection logic forestsearch uses),
  # anchoring selection to the maximum qualifying effect rather than the
  # m_diff floor.
  dfit <- tryCatch(
    dina(df = df, outcome = "y_sim", treatment = "treat_sim",
         covariates = covariates_v, family = "cox", status = "event_sim",
         n_folds = cfg$dina_n_folds, cens_type = "uniform",
         cens_params = list(T0 = cfg$cox_T0), seed = seed),
    error = function(e) NULL)

  dina_rows <- lapply(cfg$dina_foci, function(focus) {
    dsg <- if (!is.null(dfit)) tryCatch(
      dina_subgroup(fit = dfit, df = df, covariates = covariates_v,
                    m_diff = cfg$m_diff, n_min = cfg$dina_n_min,
                    direction = "both", sg_focus = focus,
                    effect_neighborhood = cfg$dina_eff_neighbor,
                    alpha = cfg$dina_alpha),
      error = function(e) NULL) else NULL
    score_dina(dsg, df, truth, cfg$thr_tol,
               method_label = sprintf("DINA-%s", focus))
  })

  scen_cols <- data.table(
    scenario_id = scenario$scenario_id, label = scenario$label,
    n = scenario$n, tau_in = scenario$tau_in, is_null = isTRUE(scenario$is_null),
    maxk = cfg$fs_maxk, rep_id = rep_id,
    n_true_S = sum(df[["flag_harm"]] == 1L))
  method_rows <- c(list(fs_row), dina_rows)
  rbindlist(lapply(method_rows, function(r) cbind(scen_cols, r)))
}
```

---

# Run the study

```{r build-scenarios}
main <- CJ(n = CONFIG$n_levels, eff = names(CONFIG$tau_in))
main[, tau_in := CONFIG$tau_in[eff]]
main[, label := sprintf("n=%d, %s (tau_in=%.1f)", n, eff, tau_in)]
main[, is_null := FALSE]

nullarm <- data.table(n = CONFIG$n_levels, eff = "null", tau_in = 0.0,
  label = sprintf("n=%d, NULL", CONFIG$n_levels), is_null = TRUE)

scenarios <- rbind(main[, .(n, eff, tau_in, label, is_null)], nullarm)
scenarios[, scenario_id := .I]
scenarios[]
```

```{r run-grid}
run_grid <- function(scenarios, cfg) {
  grid <- scenarios[rep(seq_len(.N), each = cfg$nsims)]
  grid[, rep_id := rep(seq_len(cfg$nsims), times = nrow(scenarios))]

  if (cfg$parallel == "sims") {
    res <- foreach::foreach(
      i = seq_len(nrow(grid)), .errorhandling = "pass",
      .options.future = list(
        packages = c("forestsearch", "survival", "data.table", "future"),
        seed = TRUE)
    ) %dofuture% {
      future::plan("sequential")
      run_one_rep(as.list(grid[i]), grid$rep_id[i], cfg)
    }
  } else {
    res <- lapply(seq_len(nrow(grid)), function(i)
      run_one_rep(as.list(grid[i]), grid$rep_id[i], cfg))
  }
  ok <- vapply(res, is.data.table, logical(1))
  if (any(!ok)) message(sprintf("%d/%d reps errored and were dropped.",
                                sum(!ok), length(res)))
  rbindlist(res[ok], fill = TRUE)
}

t0 <- proc.time()[3]
results <- run_grid(scenarios, CONFIG)
timings$study <- proc.time()[3] - t0
message(sprintf("Study completed in %.1f min (%d method-rows).",
                timings$study / 60, nrow(results)))

# Reset workers between the main study and the m_diff/maxk follow-ups.
if (CONFIG$parallel == "sims") {
  forestsearch::reset_workers(workers = n_workers)
}
```

---

# Results

## True subgroup size

The true subgroup is $S = \{x_1 \ge 0.30\}$ with $x_1 \sim U[-1,1]$, so
membership is random across replications: each subject is in $S$ with
probability $P(x_1 \ge 0.30) = (1 - 0.30)/2 = 0.35$, and
$|S| \sim \text{Binomial}(n, 0.35)$. The table reports the expected and
realized sizes; the discovered-size columns in the next table should be
read against these.

```{r true-S-size}
true_p <- (1 - CONFIG$true_cut) / 2   # P(x1 >= cut) for x1 ~ U[-1,1]

trueS <- unique(results[is_null == FALSE,
  .(scenario_id, n, label, rep_id, n_true_S)])

trueS_summary <- trueS[, .(
  expected_p     = true_p,
  expected_n     = true_p * n[1],
  mean_n_true_S  = mean(n_true_S),
  sd_n_true_S    = sd(n_true_S),
  min_n_true_S   = min(n_true_S),
  max_n_true_S   = max(n_true_S),
  mean_frac      = mean(n_true_S / n)
), by = .(n)][order(n)]

trueS_summary |>
  gt::gt() |>
  gt::tab_header(
    title = gt::md("**True subgroup size:** $S = \\{x_1 \\ge 0.30\\}$"),
    subtitle = sprintf("x1 ~ U[-1,1]; |S| ~ Binomial(n, %.2f)", true_p)) |>
  gt::fmt_number(columns = c(expected_p, mean_frac), decimals = 3) |>
  gt::fmt_number(columns = c(expected_n, mean_n_true_S, sd_n_true_S),
                 decimals = 1) |>
  gt::cols_label(n = "n", expected_p = "E[p]", expected_n = "E|S|",
                 mean_n_true_S = "mean |S|", sd_n_true_S = "sd |S|",
                 min_n_true_S = "min |S|", max_n_true_S = "max |S|",
                 mean_frac = "mean |S|/n") |>
  gt::tab_source_note(gt::md(
    "*`|S|` is random across replications (Binomial). A method that
     recovers `S` well should report a discovered size near `mean |S|`;
     a discovered size near `n` (fraction near 1) indicates the subgroup
     was extended well beyond the truth --- which drives specificity
     down even when sensitivity is high.*")) |>
  gt::tab_options(table.font.size = 12)
```

## Detection and recovery, by method (non-null cells)

```{r agg-main}
agg <- results[is_null == FALSE, .(
  detection_rate = mean(found),
  covariate_hit  = mean(covariate_involved[found], na.rm = TRUE),
  sensitivity    = mean(sensitivity, na.rm = TRUE),
  specificity    = mean(specificity, na.rm = TRUE),
  ppv            = mean(ppv, na.rm = TRUE),
  npv            = mean(npv, na.rm = TRUE),
  jaccard        = mean(jaccard, na.rm = TRUE),
  mean_sg_n      = mean(n_subgroup[found], na.rm = TRUE),
  mean_sg_frac   = mean((n_subgroup / n)[found], na.rm = TRUE),
  mean_n_factors = mean(n_factors[found], na.rm = TRUE)
), by = .(label, n, tau_in, method)][order(n, tau_in, method)]

agg |>
  gt::gt(groupname_col = "label") |>
  gt::tab_header(
    title = gt::md("**Signature identification: forestsearch vs dina_subgroup**"),
    subtitle = sprintf(
      "Cox; true S = {%s >= %.2f}; effect.threshold = %.2f (m_diff = %.3f); maxk = %d; %d sims/cell",
      CONFIG$true_cov, CONFIG$true_cut, CONFIG$effect_threshold,
      CONFIG$m_diff, CONFIG$fs_maxk, CONFIG$nsims)) |>
  gt::fmt_number(columns = c(detection_rate, covariate_hit, sensitivity,
                             specificity, ppv, npv, jaccard, mean_sg_frac,
                             mean_n_factors), decimals = 3) |>
  gt::fmt_number(columns = mean_sg_n, decimals = 1) |>
  gt::cols_label(method = "method", detection_rate = "detected",
                 covariate_hit = "cov hit", sensitivity = "sens",
                 specificity = "spec", ppv = "ppv", npv = "npv",
                 jaccard = "Jaccard",
                 mean_sg_n = "mean |S-hat|", mean_sg_frac = "|S-hat|/n",
                 mean_n_factors = "mean #factors", n = "n", tau_in = "tau_in") |>
  gt::tab_source_note(gt::md(
    "*`detected` = P(any subgroup found). `cov hit` = P(true covariate
     appears in the discovered signature | found). `sens`/`spec`/`ppv`/
     `Jaccard` compare discovered membership to the true mask. `mean
     |S-hat|` and `|S-hat|/n` are the discovered subgroup size (count and
     fraction of the sample), to be read against the true-S size summary
     below. `mean #factors` is 1 for dina_subgroup by construction; for
     forestsearch it reflects how often a multi-factor rule was returned
     (only > 1 when maxk = 2).*")) |>
  gt::tab_options(table.font.size = 12)
```

## Threshold recovery (dina_subgroup; signature-correct reps)

```{r agg-threshold}
thr <- results[startsWith(method, "DINA") & is_null == FALSE &
                 !is.na(threshold_error), .(
  n_reps   = .N,
  thr_mean = mean(threshold, na.rm = TRUE),
  thr_bias = mean(threshold_error, na.rm = TRUE),
  thr_rmse = sqrt(mean(threshold_error^2, na.rm = TRUE))
), by = .(label, n, tau_in, method)][order(n, tau_in, method)]

if (nrow(thr)) {
  thr |>
    gt::gt() |>
    gt::tab_header(
      title = gt::md("**dina_subgroup threshold recovery, by selection rule**"),
      subtitle = sprintf("true cut = %.2f; m_diff = %.3f",
                         CONFIG$true_cut, CONFIG$m_diff)) |>
    gt::fmt_number(columns = c(thr_mean, thr_bias, thr_rmse), decimals = 3) |>
    gt::cols_label(label = "Scenario", n = "n", tau_in = "tau_in",
                   method = "focus", n_reps = "reps", thr_mean = "mean thr",
                   thr_bias = "bias", thr_rmse = "RMSE") |>
    gt::tab_source_note(gt::md(
      "*Negative bias = discovered cut below the true cut. Under `maxSG`
       the largest-qualifying rule extends the subgroup down-tail until
       the running mean effect falls to `m_diff`, so the cut sits well
       below 0.30. Under `effMaxSG` the effect-neighborhood band anchors
       to the maximum effect, pulling the cut back toward the truth ---
       smaller bias and RMSE.*")) |>
    gt::tab_options(table.font.size = 12)
} else cat("No signature-correct DINA reps to summarise.\n")
```

## False-discovery rate (null arm), by method

```{r agg-null}
fdr <- results[is_null == TRUE, .(
  false_discovery_rate = mean(found)
), by = .(n, method)][order(n, method)]

fdr |>
  gt::gt(groupname_col = "n") |>
  gt::tab_header(
    title = gt::md("**False-discovery rate (null: no heterogeneity)**"),
    subtitle = sprintf("effect.threshold = %.2f; m_diff = %.3f; maxk = %d; %d sims/cell",
                       CONFIG$effect_threshold, CONFIG$m_diff,
                       CONFIG$fs_maxk, CONFIG$nsims)) |>
  gt::fmt_number(columns = false_discovery_rate, decimals = 3) |>
  gt::cols_label(method = "method",
                 false_discovery_rate = "false-discovery rate") |>
  gt::tab_source_note(gt::md(
    "*Under no treatment-effect heterogeneity, any discovered subgroup is
     a false positive. Both methods' harm thresholds (effect.threshold /
     m_diff) act as the guard.*")) |>
  gt::tab_options(table.font.size = 12)
```

```{r plan-teardown}
future::plan("sequential")
gc(verbose = FALSE)
```

---

# Interpretation

```{r interpretation, echo=FALSE, results='asis'}
fmt_pct <- function(x) if (length(x) && is.finite(x)) sprintf("%.0f%%", 100 * x) else "NA"
getm <- function(dt, m) { r <- dt[method == m]; if (nrow(r)) r else dt[0] }
cell <- agg[n == max(n) & tau_in == CONFIG$tau_in[["strong"]]]
fs_c  <- getm(cell, "FS")
dn_mx <- getm(cell, "DINA-maxSG")
dn_ef <- getm(cell, "DINA-effMaxSG")

cat(sprintf("
At the strongest signal (n = %d, tau_in = %.1f, maxk = %d):

* **forestsearch (%s)** detected in %s of replications, involved the true
  covariate in %s of detections, sensitivity %s, specificity %s.
* **dina_subgroup (maxSG, legacy)** detected in %s, sensitivity %s,
  specificity %s, mean |S-hat|/n %s.
* **dina_subgroup (effMaxSG, new band)** detected in %s, sensitivity %s,
  specificity %s, mean |S-hat|/n %s.

Same fit and same threshold search; only the selection rule differs. The
effect-neighborhood band moves dina specificity from %s (maxSG) to %s
(effMaxSG) by anchoring selection to the maximum qualifying effect rather
than the m_diff floor --- the same mechanism forestsearch's effMaxSG uses.
",
  max(CONFIG$n_levels), CONFIG$tau_in[["strong"]], CONFIG$fs_maxk,
  CONFIG$fs_sg_focus,
  fmt_pct(fs_c$detection_rate), fmt_pct(fs_c$covariate_hit),
  fmt_pct(fs_c$sensitivity), fmt_pct(fs_c$specificity),
  fmt_pct(dn_mx$detection_rate), fmt_pct(dn_mx$sensitivity),
  fmt_pct(dn_mx$specificity), fmt_pct(dn_mx$mean_sg_frac),
  fmt_pct(dn_ef$detection_rate), fmt_pct(dn_ef$sensitivity),
  fmt_pct(dn_ef$specificity), fmt_pct(dn_ef$mean_sg_frac),
  fmt_pct(dn_mx$specificity), fmt_pct(dn_ef$specificity)))
```

Points the tables support:

* **`maxk` is the fairness lever.** With `fs_maxk = 1` both methods are
  single-factor and the comparison is like-for-like against the
  single-factor truth. Re-running with `fs_maxk = 2` shows the cost of
  the larger hypothesis space: forestsearch may return a two-factor rule
  (`mean #factors` > 1), which can lower specificity / covariate-hit when
  the truth is genuinely one factor.
* **Matched thresholds.** Both methods use the same natural-scale harm
  cutoff (`exp(m_diff) = effect.threshold = 1.25`), so differences are
  attributable to the search and selection rule, not to different
  cutoffs.
* **Selection rule, not search, drives the size.** `dina_subgroup()` now
  exposes the same `sg_focus` vocabulary as forestsearch. Under `maxSG`
  it keeps the legacy largest-qualifying behaviour (extends membership
  outward: higher sensitivity, lower specificity). Under `effMaxSG` it
  applies the effect-neighborhood band --- anchoring to the maximum
  effect over qualifying candidates --- which tightens the subgroup
  toward the true cut. The two `DINA-*` rows isolate this exactly: same
  fit, same threshold search, different selection rule.
* **Why dina specificity is low here, and why that is correct.** With
  `tau_in` (e.g. 0.8 or 1.3) well above the threshold `m_diff = log(1.25)
  = 0.223`, the mean effect over `{x1 >= q}` stays above `m_diff` even for
  very small `q`: the subgroup-mean is `tau_in * P(x1 >= 0.30 | x1 >= q)`,
  which only has to clear 0.223. So the largest-qualifying subgroup
  extends far down-tail (often the whole sample), giving sensitivity near
  1 and specificity near 0 against the true `{x1 >= 0.30}`. This is the
  permissiveness of a low `m_diff` relative to the in-subgroup effect, not
  a scoring error --- the discovered-size column shows `|S-hat|/n`
  approaching 1, and the threshold-bias column shows the cut sitting far
  below 0.30. Two remedies tighten the subgroup and lift specificity:
  raising `m_diff` toward `tau_in` (which requires guessing the
  in-subgroup effect), or --- new in this comparison --- selecting under
  `effMaxSG`, whose effect-neighborhood band adapts to the realized
  maximum effect without `tau_in` specified in advance. The
  `DINA-effMaxSG` rows show the latter directly.

This study isolates **discovery** behaviour. It does not address
post-selection inference (the conditional-on-signature CIs and the
deferred selection-adjusted bootstrap), which are separate questions.

# Timing

```{r timing}
timings$total <- Reduce(`+`, timings)
data.frame(Step = names(timings),
           Minutes = round(unlist(timings) / 60, 2)) |>
  gt::gt() |>
  gt::tab_header(title = "Computation timing") |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_body(rows = Step == "total"))
```

# Session information

```{r session-info}
sessionInfo()
```
