---
title: "DINA Subgroup Bootstrap Smoke Test — Selection-Adjusted Inference Across All Four Families"
subtitle: "End-to-end pipeline check for dina_subgroup_bootstrap() with structural and math-sanity assertions"
author: "forestsearch v0.2.0"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

# Purpose

Smoke-test `forestsearch::dina_subgroup_bootstrap()` end-to-end across
all four response families.  The function provides three complementary
inference outputs for the harm subgroup discovered by `dina_subgroup()`:

* **Effect CI** on the fixed-subgroup linear functional
  $a^{*\top}\hat\beta_b$, where $a^* = (1, \bar x_{S^*})$ is held
  *fixed* from the original-data subgroup and $\hat\beta_b$ is each
  bootstrap iteration's DINA coefficient vector.  This is the
  bootstrap analog of `dina_subgroup()`'s sandwich Wald CI, with
  width reflecting sampling variance of $\hat\beta$.  See
  `dina_inference_design_notes.md`, Section 4, for the rationale
  for this choice over a naive bootstrap of `mean_tau_hat`.
* **Structural stability CIs** on $n_{\text{subgroup}}$ and the
  threshold $q$, restricted to bootstrap iterations that selected the
  same `(covariate, direction)` as the original-data subgroup.
* **Selection-frequency diagnostic** counting how often each
  `(covariate, direction)` was selected.

This document validates that the function:

1. Runs without error across all four families with `parallel = "boots"`.
2. Returns the expected `"dina_subgroup_bootstrap"` object with all
   new fields (`effect_ci`, `effect_dist`, `n_subgroup_ci`,
   `threshold_ci`, `n_modal_match`) present and well-typed.
3. Produces an effect CI with non-degenerate width — the headline
   correctness check that the previous `mean_tau_ci` would have
   failed by clustering tightly at `m_diff`.
4. Populates the selection-frequency and stability diagnostics
   correctly.
5. Handles the "no subgroup found" path cleanly via a negative-control
   call with an impossibly large `m_diff`.

Same DGM as the subgroup smoke test (strong β₁ = 1.20 signal on `x1`)
so a real harm subgroup exists for the bootstrap to discover.  This is
not a Monte Carlo simulation — single seed per family, single resampling
scheme, single set of parameters.  Coverage and bias studies belong in
a separate document.

---

# Setup

```{r setup}
library(forestsearch)
library(data.table)
library(gt)

set.seed(20260528)

SMOKE <- list(
  n                = 800L,
  d                = 5L,
  n_boot           = 100L,     # bootstrap iterations
  n_folds          = 2L,       # cross-fitting folds inside each DINA fit
  parallel         = "boots",  # "boots" or "none"
  n_workers_frac   = 0.9,
  n_workers        = NULL,
  n_min            = 60L,
  alpha            = 0.05,
  cox_T0           = 1.0
)

# Strong heterogeneity on x1 only, others zero.
BETA_TRUE <- c(0.40, 1.20, 0.00, 0.00, 0.00, 0.00)
stopifnot(length(BETA_TRUE) == SMOKE$d + 1L)

# Family-specific m_diff values: positioned to detect the right-tail
# harm subgroup on x1.
M_DIFF_BY_FAMILY <- c(
  gaussian = 0.80,
  binomial = 0.80,
  poisson  = 0.80,
  cox      = 0.80
)
```

## `.tc()` accumulating helper

```{r tc-helper}
.tc_env <- new.env(parent = emptyenv())
.tc_env$rows <- list()

.tc <- function(label, condition, family, details = "") {
  passed <- isTRUE(tryCatch(condition, error = function(e) FALSE))
  msg <- if (!passed && nzchar(details)) details
         else if (!passed) "condition FALSE or errored"
         else ""
  .tc_env$rows[[length(.tc_env$rows) + 1L]] <- data.table(
    family  = family,
    label   = label,
    status  = if (passed) "PASS" else "FAIL",
    details = msg
  )
  invisible(passed)
}

.tc_table <- function() {
  if (length(.tc_env$rows) == 0L) {
    return(data.table(family = character(), label = character(),
                      status = character(), details = character()))
  }
  rbindlist(.tc_env$rows)
}

`%||%` <- function(a, b) if (is.null(a)) b else a
```

## Parallel plan setup

```{r plan-setup}
if (SMOKE$parallel == "boots") {
  total_cores <- parallelly::availableCores()
  n_workers <- if (is.null(SMOKE$n_workers)) {
    max(1L, as.integer(floor(SMOKE$n_workers_frac * total_cores)))
  } else {
    as.integer(SMOKE$n_workers)
  }
  future::plan(future::multisession, workers = n_workers)
  message(sprintf(
    "Parallel plan: %d workers (%.0f%% of %d available cores)",
    n_workers, 100 * n_workers / total_cores, total_cores
  ))
} else {
  future::plan("sequential")
  message("Parallel plan: sequential")
}
```

# Data simulation

Same DGM as `dina_subgroup_smoke_test.qmd`: $X \sim \mathrm{Unif}[-1, 1]^d$,
mild propensity confounding via $e(x) = \mathrm{logit}^{-1}(0.3 x_1)$,
prognostic effect $\mu(x) = 0.2 x_1 - 0.1 x_2$, and a deliberately
strong treatment-effect heterogeneity $\tau(x) = 0.40 + 1.20 x_1$ so a
right-tail subgroup on `x1` is the expected discovery.

```{r data-helpers}
simulate_dina_data <- function(family, n, d, beta_true, cox_T0 = 1.0) {
  X <- matrix(stats::runif(n * d, -1, 1), n, d)
  colnames(X) <- paste0("x", seq_len(d))

  e_x  <- stats::plogis(0.3 * X[, 1])
  W    <- stats::rbinom(n, 1L, e_x)

  mu_x  <- 0.2 * X[, 1] - 0.1 * X[, 2]
  tau_x <- as.numeric(beta_true[1L] + X %*% beta_true[-1L])
  eta_w <- mu_x + W * tau_x

  out <- list(X = X, W = W, beta_true = beta_true, eta_w = eta_w,
              tau_x_true = tau_x)

  out$Y <- switch(
    family,
    gaussian = as.numeric(eta_w + stats::rnorm(n)),
    binomial = stats::rbinom(n, 1L, stats::plogis(eta_w)),
    poisson  = stats::rpois(n, exp(pmin(eta_w, 5))),
    cox = {
      t_event  <- stats::rexp(n, rate = exp(eta_w))
      t_censor <- stats::runif(n, 0, cox_T0)
      time   <- pmin(t_event, t_censor)
      status <- as.integer(t_event <= t_censor)
      data.frame(time = time, status = status)
    }
  )
  out
}

build_df <- function(dat, family) {
  if (family == "cox") {
    cbind(
      data.frame(time = dat$Y$time, status = dat$Y$status, w = dat$W),
      as.data.frame(dat$X)
    )
  } else {
    cbind(
      data.frame(y = dat$Y, w = dat$W),
      as.data.frame(dat$X)
    )
  }
}
```

## Per-family test runner

```{r runner}
run_bootstrap_family_tests <- function(family) {
  m_diff <- M_DIFF_BY_FAMILY[[family]]

  dat <- simulate_dina_data(
    family    = family,
    n         = SMOKE$n,
    d         = SMOKE$d,
    beta_true = BETA_TRUE,
    cox_T0    = SMOKE$cox_T0
  )
  df_fam       <- build_df(dat, family)
  outcome_col  <- if (family == "cox") "time" else "y"
  status_col   <- if (family == "cox") "status" else NULL
  covariates_v <- colnames(dat$X)

  cens_type   <- if (family == "cox") "uniform" else NULL
  cens_params <- if (family == "cox") list(T0 = SMOKE$cox_T0) else list()

  # ---- Main bootstrap call ------------------------------------------------
  bs <- tryCatch(
    dina_subgroup_bootstrap(
      df          = df_fam,
      outcome     = outcome_col,
      treatment   = "w",
      covariates  = covariates_v,
      family      = family,
      status      = status_col,
      m_diff      = m_diff,
      n_min       = SMOKE$n_min,
      alpha       = SMOKE$alpha,
      n_boot      = SMOKE$n_boot,
      parallel    = SMOKE$parallel,
      n_folds     = SMOKE$n_folds,
      cens_type   = cens_type,
      cens_params = cens_params,
      seed        = 20260528L,
      verbose     = FALSE
    ),
    error = function(e) {
      attr(e, "details") <- conditionMessage(e)
      e
    }
  )
  bs_ok <- !inherits(bs, "error")
  .tc("dina_subgroup_bootstrap() completes without error", bs_ok, family,
      details = if (!bs_ok) attr(bs, "details") else "")
  if (!bs_ok) bs <- structure(list(), class = "dina_subgroup_bootstrap_failed")

  # ---- Structural assertions ---------------------------------------------
  .tc("returns object of class 'dina_subgroup_bootstrap'",
      inherits(bs, "dina_subgroup_bootstrap"), family)
  .tc("point is a 'dina_subgroup' object",
      inherits(bs$point, "dina_subgroup"), family)
  .tc("boot_results is a data.frame with n_boot rows",
      is.data.frame(bs$boot_results) &&
        nrow(bs$boot_results) == SMOKE$n_boot, family,
      details = if (is.data.frame(bs$boot_results))
        sprintf("nrow = %d, n_boot = %d",
                nrow(bs$boot_results), SMOKE$n_boot) else "")
  .tc("boot_results has expected columns",
      all(c("found", "covariate", "direction", "threshold",
            "n_subgroup", "mean_tau_hat", "failed") %in%
          names(bs$boot_results)), family)
  .tc("effect_ci is named length-2 numeric (lower, upper)",
      is.numeric(bs$effect_ci) && length(bs$effect_ci) == 2L &&
        identical(names(bs$effect_ci), c("lower", "upper")), family)
  .tc("n_subgroup_ci is named length-2 numeric (lower, upper)",
      is.numeric(bs$n_subgroup_ci) && length(bs$n_subgroup_ci) == 2L &&
        identical(names(bs$n_subgroup_ci), c("lower", "upper")), family)
  .tc("threshold_ci is named length-2 numeric (lower, upper)",
      is.numeric(bs$threshold_ci) && length(bs$threshold_ci) == 2L &&
        identical(names(bs$threshold_ci), c("lower", "upper")), family)
  .tc("effect_dist has length n_boot",
      length(bs$effect_dist) == SMOKE$n_boot, family,
      details = sprintf("len = %d, n_boot = %d",
                        length(bs$effect_dist), SMOKE$n_boot))
  .tc("n_boot_found + n_boot_not_found + n_boot_failed equals n_boot",
      {
        n_nf <- sum(!is.na(bs$boot_results$found) &
                    !bs$boot_results$found)
        bs$n_boot_found + n_nf + bs$n_boot_failed == SMOKE$n_boot
      },
      family,
      details = sprintf("found = %d, failed = %d, total = %d",
                        bs$n_boot_found, bs$n_boot_failed, SMOKE$n_boot))
  .tc("print() runs without error",
      !inherits(tryCatch(utils::capture.output(print(bs)),
                         error = function(e) e), "error"), family)

  # ---- Math sanity (when point estimate found something) -----------------
  if (isTRUE(bs$point$found)) {
    n_fit <- SMOKE$n_boot - bs$n_boot_failed
    if (n_fit >= 2L) {
      # Structural assertions on effect_ci
      .tc("effect_ci lower endpoint <= upper endpoint",
          bs$effect_ci[["lower"]] <= bs$effect_ci[["upper"]], family,
          details = sprintf("[%.3f, %.3f]",
                            bs$effect_ci[["lower"]],
                            bs$effect_ci[["upper"]]))
      .tc("effect_ci endpoints finite",
          all(is.finite(bs$effect_ci)), family)
      .tc("effect_dist has at least 2 finite values",
          sum(is.finite(bs$effect_dist)) >= 2L, family)
      # The percentile CI must contain the median of effect_dist
      .tc("effect_ci contains median(effect_dist)",
          {
            med <- stats::median(bs$effect_dist, na.rm = TRUE)
            bs$effect_ci[["lower"]] <= med &&
              med <= bs$effect_ci[["upper"]]
          }, family)
      # CI width should be meaningful (this is the headline test the
      # previous mean_tau_ci would have failed by clustering at m_diff).
      # 0.05 is a conservative floor; widths in the smoke test are
      # typically O(0.1-0.5).
      .tc("effect_ci width is non-degenerate (> 0.05)",
          (bs$effect_ci[["upper"]] - bs$effect_ci[["lower"]]) > 0.05,
          family,
          details = sprintf("width = %.4f",
                            bs$effect_ci[["upper"]] - bs$effect_ci[["lower"]]))
    }

    if (!is.na(bs$n_modal_match) && bs$n_modal_match >= 2L) {
      .tc("n_subgroup_ci endpoints finite when n_modal_match >= 2",
          all(is.finite(bs$n_subgroup_ci)), family)
      .tc("threshold_ci endpoints finite when n_modal_match >= 2",
          all(is.finite(bs$threshold_ci)), family)
      .tc("n_subgroup_ci lower endpoint <= upper endpoint",
          bs$n_subgroup_ci[["lower"]] <= bs$n_subgroup_ci[["upper"]],
          family)
      .tc("threshold_ci lower endpoint <= upper endpoint",
          bs$threshold_ci[["lower"]] <= bs$threshold_ci[["upper"]],
          family)
    }
  } else {
    .tc(sprintf("Point: NOT found at m_diff = %.2f (informational)", m_diff),
        TRUE, family,
        details = "No qualifying subgroup on original data.")
  }

  # ---- Selection-frequency diagnostic ------------------------------------
  if (bs$n_boot_found > 0L) {
    .tc("selection_frequency is non-NULL when at least one iter found",
        !is.null(bs$selection_frequency), family)
    .tc("selection_frequency entries sum to n_boot_found",
        sum(bs$selection_frequency) == bs$n_boot_found, family,
        details = sprintf("sum = %d, n_boot_found = %d",
                          sum(bs$selection_frequency), bs$n_boot_found))
  }

  # ---- Negative control: impossibly large m_diff -> no iters found -------
  bs_null <- tryCatch(
    dina_subgroup_bootstrap(
      df = df_fam, outcome = outcome_col, treatment = "w",
      covariates = covariates_v, family = family, status = status_col,
      m_diff = 100, n_min = SMOKE$n_min,
      n_boot = 20L,                        # smaller for speed; same negative result
      parallel = SMOKE$parallel,
      n_folds = SMOKE$n_folds,
      cens_type = cens_type, cens_params = cens_params,
      seed = 20260528L
    ),
    error = function(e) NULL
  )
  .tc("Negative control: large m_diff -> n_boot_found == 0",
      !is.null(bs_null) && bs_null$n_boot_found == 0L, family,
      details = if (!is.null(bs_null))
        sprintf("n_boot_found = %d", bs_null$n_boot_found)
      else "errored")
  .tc("Negative control: effect_ci is (NA, NA) when point not found",
      !is.null(bs_null) && all(is.na(bs_null$effect_ci)), family)
  .tc("Negative control: selection_frequency is NULL when nothing found",
      !is.null(bs_null) && is.null(bs_null$selection_frequency), family)

  list(family = family, m_diff = m_diff, bs = bs, bs_null = bs_null)
}

# Per-family gt table: point estimate, effect CI (fixed-subgroup),
# stability CIs (n_subgroup, threshold), and top selection frequencies
family_bootstrap_gt <- function(res, family) {
  bs <- res$bs

  pt_signature <- if (isTRUE(bs$point$found)) {
    cmp <- if (bs$point$direction == "left") "<=" else ">="
    sprintf("%s %s %.3f", bs$point$covariate, cmp, bs$point$threshold)
  } else "—"

  effect_str <- if (all(is.finite(bs$effect_ci))) {
    sprintf("[%.3f, %.3f]",
            bs$effect_ci[["lower"]], bs$effect_ci[["upper"]])
  } else "—"

  n_sg_str <- if (all(is.finite(bs$n_subgroup_ci))) {
    sprintf("[%.0f, %.0f]",
            bs$n_subgroup_ci[["lower"]], bs$n_subgroup_ci[["upper"]])
  } else "—"

  thr_str <- if (all(is.finite(bs$threshold_ci))) {
    sprintf("[%.3f, %.3f]",
            bs$threshold_ci[["lower"]], bs$threshold_ci[["upper"]])
  } else "—"

  modal_match_str <- if (!is.na(bs$n_modal_match))
    sprintf("%d / %d", bs$n_modal_match, bs$n_boot_found)
  else "—"

  # Top 3 selection frequencies, if any
  sel_str <- if (!is.null(bs$selection_frequency) && bs$n_boot_found > 0L) {
    sel_long <- as.data.frame(bs$selection_frequency,
                              stringsAsFactors = FALSE)
    sel_long <- sel_long[sel_long$Freq > 0L, , drop = FALSE]
    sel_long <- sel_long[order(-sel_long$Freq), , drop = FALSE]
    top3 <- head(sel_long, 3L)
    paste(
      sprintf("%s (%s): %d/%d (%.0f%%)",
              top3$covariate, top3$direction, top3$Freq,
              bs$n_boot_found,
              100 * top3$Freq / bs$n_boot_found),
      collapse = "; "
    )
  } else "—"

  rows <- data.table(
    Field = c("m_diff",
              "Point signature",
              "Point mean_tau_hat",
              "Bootstrap found / total",
              "Bootstrap failed",
              "Effect CI (fixed-subgroup)",
              "n_subgroup CI (modal-match)",
              "Threshold CI (modal-match)",
              "Modal-match iters / found",
              "Top selections (cov, dir)"),
    Value = c(
      sprintf("%.2f", res$m_diff),
      pt_signature,
      sprintf("%.3f", bs$point$mean_tau_hat %||% NA_real_),
      sprintf("%d / %d", bs$n_boot_found, bs$n_boot),
      sprintf("%d", bs$n_boot_failed),
      effect_str,
      n_sg_str,
      thr_str,
      modal_match_str,
      sel_str
    )
  )

  rows |>
    gt::gt() |>
    gt::tab_header(
      title    = sprintf("%s family — dina_subgroup_bootstrap() results",
                         tools::toTitleCase(family)),
      subtitle = sprintf("n_boot = %d; %d-fold CV per per-iteration DINA fit",
                         SMOKE$n_boot, SMOKE$n_folds)
    ) |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_source_note(
      source_note = gt::md(
        "*True heterogeneity is on `x1` only (β₁ = 1.20).  Effect CI is the
         bootstrap percentile of `a*'beta_b` with `a*` fixed from the
         original-data subgroup; stability CIs are over bootstrap-selected
         subgroups matching the modal (covariate, direction).*"
      )
    )
}
```

---

# Gaussian family

```{r family-gaussian}
res_gauss <- run_bootstrap_family_tests("gaussian")
family_bootstrap_gt(res_gauss, "gaussian")
```

---

# Binomial family

```{r family-binomial}
res_binom <- run_bootstrap_family_tests("binomial")
family_bootstrap_gt(res_binom, "binomial")
```

---

# Poisson family

```{r family-poisson}
res_pois <- run_bootstrap_family_tests("poisson")
family_bootstrap_gt(res_pois, "poisson")
```

---

# Cox family

```{r family-cox}
res_cox <- run_bootstrap_family_tests("cox")
family_bootstrap_gt(res_cox, "cox")
```

```{r plan-teardown}
future::plan("sequential")
gc(verbose = FALSE)
```

---

# Summary

```{r summary}
results_tab <- .tc_table()
n_total <- nrow(results_tab)
n_pass  <- results_tab[status == "PASS", .N]
n_fail  <- n_total - n_pass

subtitle_text <- sprintf(
  "%d / %d assertions passed (%d failures)", n_pass, n_total, n_fail
)

results_tab |>
  gt::gt(groupname_col = "family") |>
  gt::tab_header(
    title    = gt::md("**DINA subgroup bootstrap smoke-test summary**"),
    subtitle = subtitle_text
  ) |>
  gt::cols_label(
    label   = "Assertion",
    status  = "Status",
    details = "Details"
  ) |>
  gt::tab_style(
    style    = gt::cell_fill(color = "#d4f4dd"),
    locations = gt::cells_body(rows = status == "PASS")
  ) |>
  gt::tab_style(
    style    = gt::cell_fill(color = "#f9d4d4"),
    locations = gt::cells_body(rows = status == "FAIL")
  ) |>
  gt::tab_style(
    style    = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |>
  gt::tab_options(table.font.size = 12)
```

---

# Session information

```{r session-info}
sessionInfo()
```
