---
title: "Embedded vs. standalone equivalence: DINA and GRF selection"
subtitle: "Verifying that forestsearch(subgroup_method = ...) reproduces the standalone selection functions across sg_focus / selection_rule / frontier_rule"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    df-print: paged
    embed-resources: true
execute:
  warning: false
  message: false
---

## Purpose

`forestsearch()` can delegate subgroup *selection* to either `dina_subgroup()`
(`subgroup_method = "dina"`) or the GRF routines (`subgroup_method = "grf"`).
This document verifies, on synthetic data, that the **embedded** path reproduces
the **standalone** function exactly --- the discovered subgroup membership is
byte-identical --- across every selection-rule combination, for DINA and for the
GRF frontier (survival and binary GLM).

The check is membership identity: the standalone subgroup mask versus the
embedded `treat.recommend == 0` (harm) indicator, on the same data and the same
fit/seed. Two practical alignment requirements emerged and are baked into the
calls below:

* **DINA:** `forestsearch()` takes `effect.threshold` (a hazard ratio / ratio)
  and derives the link-scale floor `m_diff = log(effect.threshold)` for ratio
  families (identity for gaussian). A standalone `dina_subgroup()` takes `m_diff`
  directly, so to match you pass `m_diff = log(effect.threshold)`. Both paths must
  share the same `dina()` fit.
* **GRF (binary/count):** `forestsearch()` defaults `adverse_outcome = NULL`,
  which resolves to **`TRUE`** for binary and count outcomes (higher outcome =
  adverse). A standalone `grf.subg.harm.glm()` comparison must therefore pass
  `adverse_outcome = TRUE` to match the embedded orientation; passing `FALSE`
  flips the harm direction and the masks will (correctly) disagree.

Valid `selection_rule` values are `neighborhood`, `pareto`, `both`; `both` is
only meaningful for `effMaxSG` / `effMinSG`. Both paths enforce this identically.

```{r setup}
library(forestsearch)
library(survival)
library(data.table)

SEED  <- 8316951L
NMIN  <- 60L

# PASS/FAIL accumulator (smoke-test style, no testthat).
.results <- list()
.tc <- function(pass, msg) {
  .results[[length(.results) + 1L]] <<-
    data.table(check = msg, status = if (isTRUE(pass)) "PASS" else "FAIL")
  invisible(NULL)
}
mask_int <- function(x) as.integer(as.logical(x))
```

## DINA: embedded vs. standalone, all valid (sg_focus x selection_rule)

A synthetic Cox DGM with a genuine harm subgroup `S = {x1 > 0}`. One shared
`dina()` fit is passed to both paths (via `dina_res`) so the cross-fitted
tau-hat is identical; the embedded path receives `effect.threshold = HR` and the
standalone path the matched `m_diff = log(HR)`.

```{r dina-check}
set.seed(5)
n <- 1200L
X <- matrix(runif(n * 5, -1, 1), n, 5); colnames(X) <- paste0("x", 1:5)
W <- rbinom(n, 1, 0.5)
tau <- ifelse(X[, 1] > 0, 1.5, -0.3)
te  <- rexp(n, exp(0.1 * X[, 2] + W * tau)); tc <- runif(n, 0, 1.6)
d_dina <- data.frame(tm = pmin(te, tc), status = as.integer(te <= tc),
                     hormon = W, as.data.frame(X), id = seq_len(n))
covs <- paste0("x", 1:5)
HR <- 1.10

dfit <- dina(df = d_dina, outcome = "tm", treatment = "hormon",
             covariates = covs, family = "cox", status = "status",
             n_folds = 2L, cens_type = "uniform",
             cens_params = list(T0 = 60), seed = SEED)

dina_args <- list(family = "cox", seed = SEED, n_folds = 2L,
                  cens_type = "uniform", cens_params = list(T0 = 60),
                  max_depth = 1L, grid_probs = seq(0.1, 0.9, by = 0.1))

# Valid combinations: neighborhood for all; both only for effMaxSG / effMinSG.
dina_combos <- list(
  c("eff", "neighborhood"),
  c("effMaxSG", "neighborhood"), c("effMaxSG", "both"),
  c("effMinSG", "neighborhood"), c("effMinSG", "both"),
  c("maxSG", "neighborhood"),
  c("minSG", "neighborhood"))

dina_tbl <- rbindlist(lapply(dina_combos, function(cb) {
  foc <- cb[1]; rule <- cb[2]
  sg <- dina_subgroup(
    fit = dfit, df = d_dina, covariates = covs, m_diff = log(HR), n_min = NMIN,
    max_depth = 1L, grid_probs = seq(0.1, 0.9, by = 0.1),
    sg_focus = foc, selection_rule = rule, effect_neighborhood = 0.10)
  fs <- forestsearch(
    df.analysis = d_dina, outcome.name = "tm", event.name = "status",
    treat.name = "hormon", id.name = "id", confounders.name = covs,
    df.predict = d_dina, subgroup_method = "dina", sg_focus = foc,
    selection_rule = rule, effect_neighborhood = 0.10, effect.threshold = HR,
    n.min = NMIN, dina_args = dina_args, dina_res = dfit, use_grf = FALSE,
    use_lasso = FALSE, is.RCT = TRUE, seedit = SEED, details = FALSE,
    quiet = TRUE,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE))
  sm <- mask_int(sg$mask); em <- mask_int(fs$df.est$treat.recommend == 0)
  identical_masks <- identical(sm, em)
  .tc(identical_masks, sprintf("DINA %s / %s", foc, rule))
  data.table(sg_focus = foc, selection_rule = rule,
             standalone_n = sum(sm), embedded_n = sum(em),
             identical = identical_masks)
}))
knitr::kable(dina_tbl, caption = "DINA: standalone dina_subgroup() vs embedded forestsearch(); membership identity.")
```

## GRF survival: embedded vs. standalone, all five frontier rules

The GRF frontier maps `sg_focus -> frontier_rule` inside `forestsearch()`; the
standalone `grf.subg.harm.survival()` takes `frontier_rule` directly. Both fit
the causal survival forest with the same `seedit`, so the forests are identical.

```{r grf-surv-check}
set.seed(21)
n <- 900L
X <- matrix(runif(n * 5, -1, 1), n, 5); colnames(X) <- paste0("x", 1:5)
W <- rbinom(n, 1, 0.5)
tau <- ifelse(X[, 1] > 0.2, 1.4, -0.2)
te  <- rexp(n, exp(0.1 * X[, 2] + W * tau)); tc <- runif(n, 0, 1.6)
d_surv <- data.frame(time = pmin(te, tc), status = as.integer(te <= tc),
                     trt = W, as.data.frame(X), id = seq_len(n))
DEPTH <- 2L; NB <- 0.20
rules <- c("eff", "effMaxSG", "effMinSG", "maxSG", "minSG")

grf_surv_tbl <- rbindlist(lapply(rules, function(r) {
  a <- forestsearch:::.build_grf_survival_args(
    data = d_surv, confounders.name = covs, outcome.name = "time",
    event.name = "status", id.name = "id", treat.name = "trt",
    frac.tau = 0.80, n.min = NMIN, dmin.grf = 0.0, is.RCT = TRUE,
    grf_depth = DEPTH, seedit = SEED, return_selected_cuts_only = TRUE,
    grf_selection = "frontier", frontier_rule = r, effect_neighborhood = NB)
  std <- do.call(grf.subg.harm.survival, a)
  sm  <- mask_int(std$data$treat.recommend == 0)
  fs <- forestsearch(
    df.analysis = d_surv, outcome.name = "time", event.name = "status",
    treat.name = "trt", id.name = "id", confounders.name = covs,
    df.predict = d_surv, subgroup_method = "grf", grf_selection = "frontier",
    sg_focus = r, effect_neighborhood = NB, grf_depth = DEPTH, dmin.grf = 0.0,
    frac.tau = 0.80, n.min = NMIN, use_grf = FALSE, use_lasso = FALSE,
    is.RCT = TRUE, seedit = SEED, details = FALSE, quiet = TRUE,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE))
  em <- mask_int(fs$df.est$treat.recommend == 0)
  identical_masks <- identical(sm, em)
  .tc(identical_masks, sprintf("GRF survival %s", r))
  data.table(frontier_rule = r, standalone_n = sum(sm), embedded_n = sum(em),
             identical = identical_masks)
}))
knitr::kable(grf_surv_tbl, caption = "GRF survival: standalone grf.subg.harm.survival() vs embedded forestsearch().")
```

## GRF binary GLM: embedded vs. standalone (adverse_outcome matched)

For binary outcomes the embedded default `adverse_outcome = NULL` resolves to
`TRUE`, so the standalone call sets `adverse_outcome = TRUE` to match.

```{r grf-glm-check}
set.seed(23)
n <- 1600L
X <- matrix(runif(n * 5, -1, 1), n, 5); colnames(X) <- paste0("x", 1:5)
W <- rbinom(n, 1, 0.5)
inH <- (X[, 1] > 0 & X[, 2] > 0)
Yb  <- rbinom(n, 1, plogis(-0.2 + 0.1 * X[, 1] + W * ifelse(inH, 1.8, -0.5)))
d_bin <- data.frame(Y = Yb, trt = W, as.data.frame(X), id = seq_len(n))

grf_glm_tbl <- rbindlist(lapply(rules, function(r) {
  a <- forestsearch:::.build_grf_glm_args(
    data = d_bin, confounders.name = covs, outcome.name = "Y",
    treat.name = "trt", id.name = "id", outcome_type = "binary", n.min = NMIN,
    dmin.grf = 0.0, is.RCT = TRUE, grf_depth = DEPTH, seedit = SEED,
    return_selected_cuts_only = TRUE, adverse_outcome = TRUE,
    grf_selection = "frontier", frontier_rule = r, effect_neighborhood = NB)
  std <- do.call(grf.subg.harm.glm, a)
  sm  <- mask_int(std$data$treat.recommend == 0)
  fs <- forestsearch(
    df.analysis = d_bin, outcome.name = "Y", treat.name = "trt", id.name = "id",
    confounders.name = covs, df.predict = d_bin, outcome_type = "binary",
    effect_measure = "OR", subgroup_method = "grf", grf_selection = "frontier",
    sg_focus = r, effect_neighborhood = NB, grf_depth = DEPTH, dmin.grf = 0.0,
    n.min = NMIN, use_grf = FALSE, use_lasso = FALSE, is.RCT = TRUE,
    seedit = SEED, details = FALSE, quiet = TRUE,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE))
  em <- mask_int(fs$df.est$treat.recommend == 0)
  identical_masks <- identical(sm, em)
  .tc(identical_masks, sprintf("GRF binary %s", r))
  data.table(frontier_rule = r, standalone_n = sum(sm), embedded_n = sum(em),
             identical = identical_masks)
}))
knitr::kable(grf_glm_tbl, caption = "GRF binary GLM: standalone grf.subg.harm.glm() vs embedded forestsearch() (adverse_outcome = TRUE).")
```

## Summary

```{r summary}
summary_tbl <- rbindlist(.results)
n_pass <- sum(summary_tbl$status == "PASS")
n_fail <- sum(summary_tbl$status == "FAIL")
knitr::kable(summary_tbl,
  caption = sprintf("All equivalence checks: %d PASS, %d FAIL.", n_pass, n_fail))
```

`r if (exists("summary_tbl")) sprintf("**%d of %d checks passed.**", sum(summary_tbl$status == "PASS"), nrow(summary_tbl))`

## Notes

* **Equivalence is exact, not approximate.** With matched fit/seed and the two
  alignment requirements above (DINA `m_diff = log(effect.threshold)`; GRF
  binary/count `adverse_outcome = TRUE`), the embedded and standalone paths
  return byte-identical membership.
* **Threshold label precision.** The embedded DINA path formats the selected
  threshold into its `sg.harm` label at 17 significant digits, so membership
  recomputed from the label round-trips exactly to the standalone `sg$mask`. A
  lower-precision format would flip subjects sitting exactly on a selected cut.
* **GRF tree mode** ignores `sg_focus` (it selects via `dmin.grf` / `grf_depth`),
  so there is no `(sg_focus x rule)` grid to align there; only the threading of
  `dmin.grf` / `grf_depth` matters, and the tree path is unchanged from the
  pre-frontier implementation.
