---
title: "Subgroup-Method Signature Recovery (Single-Factor)"
subtitle: "Matched harm-subgroup recovery for FS-consistency, FS-grf (tree and frontier), and standalone DINA on a single-factor DGM (depth-1 / maxk = 1)"
author: "forestsearch v0.2.0"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
---

# Purpose

This document studies a DGM whose true signature is a **single covariate
threshold** -- the ideal case for the univariate `dina_subgroup()` search and
for `forestsearch(maxk = 1)`. (The companion document
`subgroup_method_signature_recovery_2factor.qmd` studies the two-factor
analog, where the truth is an AND-conjunction of two covariates.)  Each
identifier is given the matching single-factor hypothesis space:

* `forestsearch(maxk = 1)` --- a single-factor subgroup.
* `dina_subgroup(max_depth = 1)` --- a univariate threshold.
* `forestsearch(subgroup_method = "grf", grf_depth = 1)` --- a single
  policy-tree split.

The consistency and DINA arms are run under **two selection rules** so the
selection rule --- not the search --- is isolated:

* **`effMaxSG`**: among candidates whose effect is within
  `effect_neighborhood` of the maximum qualifying effect, take the largest.
* **`eff`**: the max-effect rule (largest qualifying effect).

All methods are run on the **same simulated data** each replication and scored
against the **same** ground-truth subgroup. The DGM is: covariates
$\sim U[-1,1]$, logistic propensity, Cox outcome via `rexp(rate = exp(eta))`
with uniform censoring, and a true harm subgroup

$$S = \{x_1 \ge c_1\}$$

(treatment increases the hazard inside $S$).

## Matched thresholds (harm study)

The methods use the **same harm cutoff on the natural (HR) scale**:

| Method | Threshold argument | Value |
|---|---|---|
| `dina_subgroup()` | `m_diff` (log-HR scale) | `log(1.25) ~ 0.223` |
| `forestsearch()` | `effect.threshold` (HR scale) | `1.25` |

`exp(m_diff) = effect.threshold = 1.25`, so "largest subgroup with effect
$\ge$ 1.25 (HR)" means the same number for both.

## Fairness: the `maxk` / `max_depth` knob

The true signature is **one factor**. Each method is given the single-factor
hypothesis space (`fs_maxk = 1L`, `dina_max_depth = 1L`, `grf_depth = 1L`), so
the comparison is like-for-like against the univariate truth. This is the
setting most favorable to DINA's univariate search; the two-factor companion
study shows what happens when the truth is a genuine conjunction.

## Multiple selection foci for the consistency and DINA arms

`dina_subgroup()` separates an expensive cross-fitted `dina()` fit from a cheap
threshold search, so `dina_foci = c("eff", "effMaxSG")` reuses one fit and
scores it under each rule. `forestsearch()` cannot reuse a single fit this way
--- its GRF/LASSO screening and consistency search are not separable from
selection through the public API --- so `fs_foci = c("eff", "effMaxSG")`
**re-runs `forestsearch()` once per focus**. (The packaged interactive analog is
`compare_selection_rules(sg_focus = c("eff","effMaxSG"),
selection_rule = c("neighborhood","neighborhood"))`, which likewise loops the
search; `sort_subgroups_preview()` is the internal selection sorter that a future
re-selection helper could build on.)  The FS-grf arm does not use `sg_focus`, so
it is a single run, not a per-focus loop.

> This is a Monte Carlo study. `forestsearch()` is far heavier per replication
> than `dina_subgroup()`, and running it under two foci plus the FS-grf arm
> multiplies that cost. Keep `nsims` modest and scale up on a multi-core
> machine; the per-method timing tables report where the time goes.

## GRF arms: `forestsearch(subgroup_method = "grf")` (tree and frontier)

Alongside the consistency-search FS arm and standalone DINA, the study includes
**two GRF-selection arms**, both via `forestsearch(subgroup_method = "grf")` on
the same causal-survival-forest doubly-robust scores, differing only in how the
subgroup is selected from those scores:

* **`FS-grf-tree`** (`grf_selection = "tree"`, the default) selects the harm
  leaf from the GRF policy tree under `dmin.grf` / `grf_depth`.
* **`FS-grf-frontier`** (`grf_selection = "frontier"`, experimental) selects
  from the Pareto frontier of (harm-effect, size) computed on the same scores,
  under the **effMaxSG** rule (largest candidate within `effect_neighborhood` of
  the maximum harm-effect). Only the effMaxSG variant is run here.

Two things make each a single arm rather than a per-focus family:

* **The tree arm does not use `sg_focus`** at all; the frontier arm uses it only
  to fix a single rule (effMaxSG). Neither varies a rule across rows, so each
  contributes one row per replication.
* **`grf_depth = 1` matches the single-factor truth.** A depth-1 policy tree is
  a single split and the frontier here enumerates single-covariate thresholds,
  so both GRF subgroups are one box `{x op c}` -- directly comparable to a
  depth-1 DINA cut and a `maxk = 1` consistency subgroup. (The `forestsearch()`
  default is `grf_depth = 2`; we set it to 1 here to match the univariate
  truth.)

Comparing `FS-grf-tree` with `FS-grf-frontier` isolates the selection step alone
(policy tree vs Pareto frontier) holding the forest fit fixed. Because GRF
membership is taken from the selected subgroup, the membership-based recovery
metrics (sensitivity, specificity, PPV, NPV) are always well defined; at depth 1
both subgroups are single conjunctions, so the text-parsing metrics
(`cov-set hit`, `sig correct`, `mean #factors`) apply directly.

> **Scale note (important).** For survival outcomes the GRF arms operate on the
> **RMST-difference** scale, not hazard ratios. The causal survival forest's
> doubly-robust scores estimate a difference in restricted mean survival time at
> the horizon `frac.tau * tau`, so the GRF harm-eligibility threshold
> `dmin.grf` and the frontier's `effect_neighborhood` band are in **time
> units**, whereas the FS-consistency and DINA arms threshold on **HR**
> (`effect.threshold = 1.25`, `m_diff = log(1.25)`). These are different scales.
> With `dmin.grf = 0` (used here) the GRF eligibility floor degenerates to "any
> positive RMST harm," which is scale-free and so comparable across arms; a
> *positive* `dmin.grf` would **not** be directly comparable to the HR
> threshold. Likewise, because RMST-difference and HR do not rank subgroups
> identically (especially under non-proportional hazards), the GRF arms may
> select a different "most harmful" subgroup than the HR-based arms even on
> identical data.

---

# Configuration

```{r setup}
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(ggplot2)
library(doFuture)
library(foreach)

# Exploratory graphics utilities (not part of the package).  Adjust the
# relative depth if this document is rendered from a different location.
.util_dir <- "../_utils"
if (!dir.exists(.util_dir)) .util_dir <- file.path("..", "..", "_utils")
source(file.path(.util_dir, "signature_recovery_display.R"))
source(file.path(.util_dir, "plot_oc_metrics.R"))
source(file.path(.util_dir, "compute_oc_pareto_flag.R"))

CONFIG <- list(
  # Monte Carlo
  nsims       = 200L,          # replications per cell (scale up for production)
  base_seed   = 20260531L,

  # design grid (forestsearch is expensive; keep modest)
  d           = 5L,
  n_levels    = c(500L, 750L, 1000L),
  tau_in      = c(moderate = 0.9, strong = 1.4),  # harm log-HR inside S
  tau_out     = 0.0,

  # ---- true single-factor signature ----
  #   S = { x1 >= 0 }  (single covariate threshold; the ideal DINA case)
  true_covs = c("x1"),
  true_dirs = c("right"),
  true_cuts = c(0.0),

  cox_T0      = 1.5,           # uniform-censoring upper bound (DGM and fits match)

  # ---- matched harm thresholds (natural scale) ----
  effect_threshold = 1.75,
  m_diff           = log(1.75),

  # ---- forestsearch knobs ----
  fs_maxk            = 1L,     # single-factor hypothesis space (matched to truth)
  fs_foci            = c("eff", "effMaxSG"),  # NEW: both selection rules
  fs_eff_neighbor    = 0.10,
  fs_selection_rule  = "neighborhood",
  fs_consistency     = 1.5,
  fs_pconsistency    = 0.90,
  fs_subgroup_method  = "consistency",
  fs_splits          = 500L,
  fs_n_min           = NULL,
  fs_d0_min          = 10L,
  fs_d1_min          = 10L,
  fs_vi_grf_min      = -0.2,
  fs_use_grf         = TRUE,
  fs_use_lasso       = TRUE,
  fs_use_dina        = TRUE,
  fs_seedit          = 8316951L,
  # fs_dina_args = list(
  #          seed        = seed,
  #          n_folds     = cfg$dina_n_folds,
  #          cens_type   = "uniform",
  #          cens_params = list(T0 = cfg$cox_T0),
  #          max_depth   = cfg$dina_max_depth,
  #          grid_probs  = cfg$dina_grid_probs
  #        ),
  # use defaults
  fs_dina_args = list(),

  # ---- FS-grf knobs (subgroup_method = "grf"): GRF policy-tree selection.
  #   GRF selects via the policy tree / dmin.grf, NOT the sg_focus
  #   effect-neighborhood rule, so FS-grf is a SINGLE run per replication
  #   (one row, method "FS-grf"), not a per-focus loop.  grf_depth = 2 is the
  #   matched to the single-factor truth (grf_depth = 1).
  fs_grf_enable   = TRUE,
  fs_grf_frontier = TRUE,      # also run GRF frontier-selection (effMaxSG)
  fs_grf_depth    = 1L,        # GRF tree depth (single-split, matched to truth)
  # Three independent dmin.grf knobs, one per GRF use (all on the RMST /
  # time-unit scale for survival -- NOT the HR scale; see the scale note in
  # the GRF-arms section).  All default to 0.0, reproducing prior behavior;
  # raise a single one to calibrate that use (e.g. the frontier's FDR)
  # without moving the others.
  fs_grf_screen_dmin   = 0.0,  # screening cut-generation (FS-eff / FS-effMaxSG)
  fs_grf_tree_dmin     = 0.4,  # GRF policy-tree selection  (FS-grf-tree)
  fs_grf_frontier_dmin = 0.4,  # GRF frontier selection     (FS-grf-frontier)
  fs_grf_fractau  = 0.80,      # GRF time-horizon multiplier (survival)

  # ---- dina_subgroup knobs ----
  dina_n_folds   = 2L,
  dina_n_min     = NULL,
  dina_alpha     = 0.05,
  dina_max_depth = 1L,                          # depth-1 single-covariate selection
  dina_grid_probs = seq(0.1, 0.9, by = 0.1),    # per-covariate threshold grid
  dina_foci      = c("eff", "effMaxSG"),
  dina_eff_neighbor = 0.10,

  # parallel
  parallel       = "sims",     # "sims" or "none"
  n_workers_frac = 0.9,
  n_workers      = NULL
)

stopifnot(abs(exp(CONFIG$m_diff) - CONFIG$effect_threshold) < 1e-8)
stopifnot(CONFIG$fs_maxk %in% c(1L, 2L), CONFIG$dina_max_depth %in% c(1L, 2L))
stopifnot(length(CONFIG$true_covs) == 1L,
          length(CONFIG$true_dirs) == 1L,
          length(CONFIG$true_cuts) == 1L)
stopifnot(all(CONFIG$fs_foci %in%
              c("maxSG","minSG","eff","effMaxSG","effMinSG")))
stopifnot(all(CONFIG$dina_foci %in%
              c("maxSG","minSG","eff","effMaxSG","effMinSG")))
stopifnot(CONFIG$m_diff > CONFIG$tau_out, all(CONFIG$m_diff < CONFIG$tau_in))

`%||%` <- function(a, b) if (is.null(a)) b else a
```

## Parallel plan

```{r plan-setup}
if (CONFIG$parallel == "sims") {
  total_cores <- parallelly::availableCores()
  n_workers <- if (is.null(CONFIG$n_workers))
    max(1L, as.integer(floor(CONFIG$n_workers_frac * total_cores)))
  else as.integer(CONFIG$n_workers)
  future::plan(future::multisession, workers = n_workers)
  message(sprintf("Parallel plan: %d workers of %d cores", n_workers, total_cores))
} else {
  future::plan("sequential"); n_workers <- 1L
  message("Parallel plan: sequential")
}

timings <- list()
```

---

# Data-generating mechanism

A single-factor harm subgroup via a step in the treatment effect:
$\tau(x) = \tau_{\text{in}}$ (harm) inside $S = \{x_1 \ge c_1\}$, and
$\tau_{\text{out}} = 0$ outside. For the null arm both are zero. The data frame
carries the true membership column `flag_harm` for scoring but does not expose
the cut to any method. (The `simulate_conjunction_cox()` helper is general over
the number of true factors; a length-1 truth produces this single-factor
region.)

```{r dgm}
simulate_conjunction_cox <- function(n, d, tau_in, tau_out,
                                      true_covs, true_dirs, true_cuts, cox_T0) {
  X <- matrix(stats::runif(n * d, -1, 1), n, d)
  colnames(X) <- paste0("x", seq_len(d))

  e_x <- stats::plogis(0.3 * X[, 1])
  W   <- stats::rbinom(n, 1L, e_x)

  in_S <- rep(TRUE, n)
  for (j in seq_along(true_covs)) {
    xj <- X[, true_covs[j]]
    in_S <- in_S & (if (true_dirs[j] == "left") xj <= true_cuts[j]
                    else xj >= true_cuts[j])
  }
  tau_x <- tau_out + (tau_in - tau_out) * as.numeric(in_S)

  mu_x  <- 0.2 * X[, 1] - 0.1 * X[, 2]
  eta_w <- mu_x + W * tau_x

  t_event  <- stats::rexp(n, rate = exp(eta_w))
  t_censor <- stats::runif(n, 0, cox_T0)

  data.table(
    id        = seq_len(n),
    y_sim     = pmin(t_event, t_censor),
    event_sim = as.integer(t_event <= t_censor),
    treat_sim = W,
    flag_harm = as.integer(in_S),
    as.data.frame(X)
  )
}
```

---

# Scoring (common, method-agnostic)

Both methods may return a multi-factor rule, so both are scored on
**method-agnostic** quantities: detection, membership recovery vs the true mask
(sensitivity / specificity / PPV / NPV), the number of factors returned, and
two signature-identity flags --- whether the discovered rule uses exactly the
true **covariate set** (`covariate_set_hit`) and exactly the true
**(covariate, direction)** axes (`signature_correct`). For `dina_subgroup()`,
where the axes are explicit, per-component threshold bias is also reported when
the axes are correct.

```{r scoring}
mask_recovery <- function(disc_mask, true_mask) {
  tp <- sum(disc_mask & true_mask); fp <- sum(disc_mask & !true_mask)
  fn <- sum(!disc_mask & true_mask); tn <- sum(!disc_mask & !true_mask)
  safe <- function(a, b) if (b > 0) a / b else NA_real_
  list(sensitivity = safe(tp, tp + fn), specificity = safe(tn, tn + fp),
       ppv = safe(tp, tp + fp), npv = safe(tn, tn + fn),
       jaccard = safe(tp, tp + fp + fn))
}

# Empty/standard per-method result row (keeps rbindlist columns aligned).
# te_1 / te_2 are per-component threshold errors aligned to true_covs order.
empty_row <- function(method) data.table(
  method = method, found = FALSE,
  covariate_set_hit = NA, signature_correct = NA,
  n_subgroup = NA_integer_, n_factors = NA_integer_,
  te_1 = NA_real_, te_2 = NA_real_,
  sensitivity = NA_real_, specificity = NA_real_, ppv = NA_real_,
  npv = NA_real_, jaccard = NA_real_, sg_def = NA_character_,
  fit_seconds = NA_real_
)

# When nothing is found and the cell is non-null: everyone predicted Hc.
.fill_none <- function(out, df, truth) {
  if (!truth$is_null) {
    true_H <- df[[truth$mask_col]] == 1L
    npv_none <- if (length(true_H)) mean(!true_H) else NA_real_
    out[, c("sensitivity","specificity","ppv","npv","jaccard") :=
          .(0, 1, 0, npv_none, 0)]
  }
  out
}

# Does the discovered text use every true covariate (whole-word)?
.covset_in_text <- function(sg_txt, true_covs) {
  all(vapply(true_covs, function(cv)
    grepl(paste0("\\b", cv, "\\b"), sg_txt), logical(1)))
}

# Parse a forestsearch sg.harm definition into the set of signed axes it uses,
# as "covariate direction" strings.  Handles negation: "!{x1 <= 0.5}" is
# NOT(x1 <= 0.5) = (x1 > 0.5), i.e. the "right" direction on x1.  Returns
# NA_character_ if any factor cannot be parsed (e.g. a categorical/level cut),
# so signature_correct falls back to NA rather than a wrong value.
.parse_fs_axes <- function(sg_txt) {
  toks  <- strsplit(sg_txt, " & ", fixed = TRUE)[[1]]
  pairs <- character(0)
  for (t in toks) {
    neg  <- startsWith(trimws(t), "!")
    body <- gsub("[!{} ]", "", t)
    m <- regmatches(body,
                    regexec("^([A-Za-z0-9_.]+)(<=|>=|<|>)(.*)$", body))[[1]]
    if (length(m) != 4L) return(NA_character_)
    dir <- if (m[3] %in% c("<=", "<")) "left" else "right"
    if (neg) dir <- if (dir == "left") "right" else "left"
    pairs <- c(pairs, paste(m[2], dir))
  }
  pairs
}

# Score forestsearch: discovered membership is treat.recommend == 0 (-> H).
score_fs <- function(fs, df, truth, method_label = "FS") {
  out <- empty_row(method_label)
  sg <- tryCatch(fs$grp.consistency$sg.harm, error = function(e) NULL)
  found_flag <- !is.null(sg) && length(sg) > 0 && !all(is.na(sg))
  out[, found := found_flag]
  if (!found_flag) return(.fill_none(out, df, truth))

  sg_txt <- paste(sg[!is.na(sg)], collapse = " & ")
  out[, sg_def := sg_txt]

  # A GRF union of leaves is carried as a single disjunctive definition string
  # "(a & b) | (c & d)".  Membership-based metrics (sens/spec/ppv/npv) are
  # always valid because they read treat.recommend, but the text-parsing
  # metrics (n_factors, covariate_set_hit, signature_correct) do not have a
  # clean single-conjunction meaning for a union, so flag them NA rather than
  # mis-parse the disjunction.
  is_disjunction <- length(sg) == 1L && grepl("|", sg, fixed = TRUE)

  if (is_disjunction) {
    out[, n_factors := NA_integer_]
    if (!truth$is_null) {
      out[, covariate_set_hit := NA]
      out[, signature_correct := NA]
    }
  } else {
    out[, n_factors := sum(!is.na(sg))]
    if (!truth$is_null) {
      out[, covariate_set_hit := .covset_in_text(sg_txt, truth$covariates)]
      # Signed-axis recovery for forestsearch, parsed from the cut labels
      # (negations flip the direction).  NA if any label is unparseable.
      fs_axes <- .parse_fs_axes(sg_txt)
      out[, signature_correct := if (length(fs_axes) == 1L && is.na(fs_axes[1]))
            NA else setequal(fs_axes, paste(truth$covariates, truth$directions))]
    }
  }

  df_est <- tryCatch(fs$df.est, error = function(e) NULL)
  if (!is.null(df_est) && "treat.recommend" %in% names(df_est)) {
    disc <- df_est$treat.recommend == 0
    if (length(disc) == nrow(df) && !truth$is_null) {
      true_mask <- df[[truth$mask_col]] == 1L
      rec <- mask_recovery(disc, true_mask)
      out[, c("sensitivity","specificity","ppv","npv","jaccard") :=
            .(rec$sensitivity, rec$specificity, rec$ppv, rec$npv, rec$jaccard)]
      out[, n_subgroup := sum(disc)]
    }
  }
  out
}

# Score dina_subgroup (depth-1 or depth-2; covariate/direction/threshold may
# be length-1 or length-2 vectors).  Membership recovery uses sg$mask, which is
# depth-agnostic.  `method_label` tags rows per sg_focus run.
score_dina <- function(sg, df, truth, method_label = "DINA") {
  out <- empty_row(method_label)
  found_flag <- !is.null(sg) && isTRUE(sg$found)
  out[, found := found_flag]
  if (!found_flag) return(.fill_none(out, df, truth))

  covs <- as.character(sg$covariate)
  dirs <- as.character(sg$direction)
  thr  <- as.numeric(sg$threshold)
  out[, n_factors := length(covs)]
  out[, n_subgroup := as.integer(sg$n_subgroup)]
  out[, sg_def := sg$label %||%
        paste(sprintf("%s %s %.3f", covs,
                      ifelse(dirs == "left", "<=", ">="), thr),
              collapse = " & ")]

  if (!truth$is_null) {
    out[, covariate_set_hit := setequal(covs, truth$covariates)]
    sig_disc <- paste(covs, dirs)
    sig_true <- paste(truth$covariates, truth$directions)
    sig_ok   <- setequal(sig_disc, sig_true)
    out[, signature_correct := sig_ok]
    if (sig_ok) {
      idx <- match(truth$covariates, covs)        # align components by name
      te  <- thr[idx] - truth$cuts
      out[, te_1 := te[1]]; out[, te_2 := te[2]]
    }
    true_mask <- df[[truth$mask_col]] == 1L
    rec <- mask_recovery(as.logical(sg$mask), true_mask)
    out[, c("sensitivity","specificity","ppv","npv","jaccard") :=
          .(rec$sensitivity, rec$specificity, rec$ppv, rec$npv, rec$jaccard)]
  }
  out
}
```

## Scoring unit tests

```{r scoring-tests}
.ok <- function(lbl, cond) cat(if (isTRUE(cond)) "PASS" else "FAIL", "-", lbl, "\n")

tm <- c(rep(TRUE, 30), rep(FALSE, 70))
.ok("perfect mask sens=1", isTRUE(all.equal(mask_recovery(tm, tm)$sensitivity, 1)))
.ok("perfect mask spec=1", isTRUE(all.equal(mask_recovery(tm, tm)$specificity, 1)))

# mock data with a true conjunction S = {x1 >= 0 & x2 <= 0}
set.seed(1)
mock <- data.table(x1 = runif(200, -1, 1), x2 = runif(200, -1, 1))
mock[, flag_harm := as.integer(x1 >= 0 & x2 <= 0)]
truthC <- list(is_null = FALSE, covariates = c("x1", "x2"),
               directions = c("right", "left"), cuts = c(0, 0),
               mask_col = "flag_harm")

# correct depth-2 DINA selection (axes + cuts exactly right)
mock_sg2 <- list(found = TRUE, depth = 2L,
                 covariate = c("x1", "x2"), direction = c("right", "left"),
                 threshold = c(0, 0), n_subgroup = sum(mock$flag_harm),
                 label = "{x1 >= 0 & x2 <= 0}",
                 mask = mock$x1 >= 0 & mock$x2 <= 0)
sd2 <- score_dina(mock_sg2, mock, truthC, "DINA-effMaxSG")
.ok("dina depth-2: covariate_set_hit TRUE", isTRUE(sd2$covariate_set_hit))
.ok("dina depth-2: signature_correct TRUE", isTRUE(sd2$signature_correct))
.ok("dina depth-2: te aligned ~ 0", isTRUE(abs(sd2$te_1) < 1e-9 && abs(sd2$te_2) < 1e-9))
.ok("dina depth-2: sensitivity = 1", isTRUE(all.equal(sd2$sensitivity, 1)))
.ok("dina depth-2: n_factors = 2", identical(as.integer(sd2$n_factors), 2L))

# depth-1 DINA on a two-factor truth: axes incomplete
mock_sg1 <- list(found = TRUE, depth = 1L, covariate = "x1",
                 direction = "right", threshold = 0,
                 n_subgroup = sum(mock$x1 >= 0), label = "{x1 >= 0}",
                 mask = mock$x1 >= 0)
sd1 <- score_dina(mock_sg1, mock, truthC, "DINA-maxSG")
.ok("dina depth-1 vs 2-factor truth: covariate_set_hit FALSE",
    isFALSE(sd1$covariate_set_hit))
.ok("dina depth-1: n_factors = 1", identical(as.integer(sd1$n_factors), 1L))

# forestsearch with the right two covariates present
mock_fs <- list(grp.consistency = list(sg.harm = c("{x1 >= 0}", "{x2 <= 0}")),
                df.est = data.frame(treat.recommend =
                  as.integer(!(mock$x1 >= 0 & mock$x2 <= 0))))
sf2 <- score_fs(mock_fs, mock, truthC, "FS-effMaxSG")
.ok("fs: found TRUE", isTRUE(sf2$found))
.ok("fs: covariate_set_hit TRUE (x1 & x2)", isTRUE(sf2$covariate_set_hit))
.ok("fs: signature_correct TRUE (x1 right, x2 left)", isTRUE(sf2$signature_correct))
.ok("fs: sensitivity = 1", isTRUE(all.equal(sf2$sensitivity, 1)))
.ok("fs: n_factors = 2", identical(as.integer(sf2$n_factors), 2L))

# forestsearch with a NEGATED label: !{x1 <= 0} == (x1 > 0) == x1 right
mock_fs_neg <- list(grp.consistency = list(sg.harm = c("!{x1 <= 0}", "{x2 <= 0}")),
                    df.est = data.frame(treat.recommend =
                      as.integer(!(mock$x1 > 0 & mock$x2 <= 0))))
sf3 <- score_fs(mock_fs_neg, mock, truthC, "FS-effMaxSG")
.ok("fs negation: signature_correct TRUE (!<= flips to right)",
    isTRUE(sf3$signature_correct))

# forestsearch with wrong axes: signature_correct FALSE (not NA)
mock_fs_bad <- list(grp.consistency = list(sg.harm = c("{x3 <= 0.5}", "{x4 >= 0}")),
                    df.est = data.frame(treat.recommend = rep(1L, nrow(mock))))
sf4 <- score_fs(mock_fs_bad, mock, truthC, "FS-maxSG")
.ok("fs wrong axes: signature_correct FALSE", isFALSE(sf4$signature_correct))
```

---

# One replication: both methods, both foci, on the same data

```{r one-rep}
run_one_rep <- function(scenario, rep_id, cfg) {
  seed <- cfg$base_seed + scenario$scenario_id * 100000L + rep_id
  set.seed(seed)

  df <- simulate_conjunction_cox(
    n = scenario$n, d = cfg$d,
    tau_in = scenario$tau_in, tau_out = cfg$tau_out,
    true_covs = cfg$true_covs, true_dirs = cfg$true_dirs,
    true_cuts = cfg$true_cuts, cox_T0 = cfg$cox_T0)
  df <- as.data.frame(df)
  covariates_v <- paste0("x", seq_len(cfg$d))
  truth <- list(is_null = isTRUE(scenario$is_null),
                covariates = cfg$true_covs, directions = cfg$true_dirs,
                cuts = cfg$true_cuts, mask_col = "flag_harm")

  # ---- forestsearch: re-run once per focus (maxk = 2) --------------------
  # Unlike DINA, the screening + consistency search is not cached across
  # foci, so each focus is a separate forestsearch() call.
  fs_rows <- lapply(cfg$fs_foci, function(focus) {
    .t0 <- proc.time()[3]
    fs <- tryCatch(
      suppressWarnings(forestsearch(
        df.analysis      = df,
        confounders.name = covariates_v,
        outcome.name     = "y_sim",
        event.name       = "event_sim",
        treat.name       = "treat_sim",
        id.name          = "id",
        est.scale        = "hr",
        subgroup_method  = cfg$fs_subgroup_method,
        effect.threshold = cfg$effect_threshold,
        hr.consistency   = cfg$fs_consistency,
        sg_focus               = focus,
        selection_rule         = cfg$fs_selection_rule,
        effect_neighborhood    = cfg$fs_eff_neighbor,
        pconsistency.threshold = cfg$fs_pconsistency,
        fs.splits        = cfg$fs_splits,
        n.min            = cfg$fs_n_min,
        d0.min           = cfg$fs_d0_min,
        d1.min           = cfg$fs_d1_min,
        maxk             = cfg$fs_maxk,
        vi.grf.min       = cfg$fs_vi_grf_min,
        dmin.grf         = cfg$fs_grf_screen_dmin,
        use_grf          = cfg$fs_use_grf,
        use_lasso        = cfg$fs_use_lasso,
        use_dina         = cfg$fs_use_dina,
        dina_args        = cfg$fs_dina_args,
        use_twostage     = TRUE,
        is.RCT           = TRUE,
        details          = FALSE,
        quiet            = TRUE,
        seedit           = cfg$fs_seedit,
        parallel_args    = list(plan = "sequential", workers = 1,
                                show_message = FALSE)
      )),
      error = function(e) NULL)
    .sec <- proc.time()[3] - .t0
    r <- score_fs(fs, df, truth, method_label = sprintf("FS-%s", focus))
    r[, fit_seconds := .sec]
    r
  })

  # ---- forestsearch GRF-selection mode (subgroup_method = "grf") ----------
  # GRF selects the subgroup from its policy tree (depth-2 by default), NOT via
  # the sg_focus effect-neighborhood rule, so this is a SINGLE run per
  # replication labelled "FS-grf" -- not a per-focus loop.  Configured by the
  # call's GRF arguments (grf_depth / dmin.grf / frac.tau); use_lasso/use_dina
  # are irrelevant in this mode (selection bypasses screening).
  grf_rows <- if (isTRUE(cfg$fs_grf_enable)) {
    .t0 <- proc.time()[3]
    fs_grf <- tryCatch(
      suppressWarnings(forestsearch(
        df.analysis      = df,
        confounders.name = covariates_v,
        outcome.name     = "y_sim",
        event.name       = "event_sim",
        treat.name       = "treat_sim",
        id.name          = "id",
        est.scale        = "hr",
        subgroup_method  = "grf",
        effect.threshold = cfg$effect_threshold,
        grf_depth        = cfg$fs_grf_depth,
        dmin.grf         = cfg$fs_grf_tree_dmin,
        frac.tau         = cfg$fs_grf_fractau,
        n.min            = cfg$fs_n_min,
        use_grf          = FALSE,   # selection IS GRF; no separate screening
        use_lasso        = FALSE,
        is.RCT           = TRUE,
        details          = FALSE,
        quiet            = TRUE,
        seedit           = cfg$fs_seedit,
        parallel_args    = list(plan = "sequential", workers = 1,
                                show_message = FALSE)
      )),
      error = function(e) NULL)
    .sec <- proc.time()[3] - .t0
    r <- score_fs(fs_grf, df, truth, method_label = "FS-grf-tree")
    r[, fit_seconds := .sec]
    list(r)
  } else list()

  # ---- forestsearch GRF FRONTIER selection (grf_selection = "frontier") ----
  # Same GRF doubly-robust scores, but selection bypasses the policy tree: the
  # Pareto frontier of (harm-effect, size) is taken and one subgroup is chosen
  # under the effMaxSG rule (sg_focus = "effMaxSG", relative effect_neighborhood).
  # Single run per replication labelled "FS-grf-frontier" (experimental mode).
  grf_fr_rows <- if (isTRUE(cfg$fs_grf_enable) && isTRUE(cfg$fs_grf_frontier)) {
    .t0 <- proc.time()[3]
    fs_grf_fr <- tryCatch(
      suppressWarnings(forestsearch(
        df.analysis         = df,
        confounders.name    = covariates_v,
        outcome.name        = "y_sim",
        event.name          = "event_sim",
        treat.name          = "treat_sim",
        id.name             = "id",
        est.scale           = "hr",
        subgroup_method     = "grf",
        grf_selection       = "frontier",
        sg_focus            = "effMaxSG",
        effect_neighborhood = cfg$fs_eff_neighbor,
        effect.threshold    = cfg$effect_threshold,
        grf_depth           = cfg$fs_grf_depth,
        dmin.grf            = cfg$fs_grf_frontier_dmin,
        frac.tau            = cfg$fs_grf_fractau,
        n.min               = cfg$fs_n_min,
        use_grf             = FALSE,
        use_lasso           = FALSE,
        is.RCT              = TRUE,
        details             = FALSE,
        quiet               = TRUE,
        seedit              = cfg$fs_seedit,
        parallel_args       = list(plan = "sequential", workers = 1,
                                   show_message = FALSE)
      )),
      error = function(e) NULL)
    .sec <- proc.time()[3] - .t0
    r <- score_fs(fs_grf_fr, df, truth, method_label = "FS-grf-frontier")
    r[, fit_seconds := .sec]
    list(r)
  } else list()

  # ---- dina_subgroup: one fit, scored under each focus (depth-2) ---------
  # Time the shared dina() fit once and the per-focus selection separately.
  # The per-row fit_seconds is (shared fit / n_foci) + that focus's selection,
  # so summing the DINA rows recovers the true one-fit-many-selections cost.
  .t_fit <- proc.time()[3]
  dfit <- tryCatch(
    dina(df = df, outcome = "y_sim", treatment = "treat_sim",
         covariates = covariates_v, family = "cox", status = "event_sim",
         n_folds = cfg$dina_n_folds, cens_type = "uniform",
         cens_params = list(T0 = cfg$cox_T0), seed = seed),
    error = function(e) NULL)
  .fit_sec   <- proc.time()[3] - .t_fit
  .fit_share <- .fit_sec / max(1L, length(cfg$dina_foci))

  dina_rows <- lapply(cfg$dina_foci, function(focus) {
    .t0 <- proc.time()[3]
    dsg <- if (!is.null(dfit)) tryCatch(
      dina_subgroup(fit = dfit, df = df, covariates = covariates_v,
                    m_diff = cfg$m_diff, n_min = cfg$dina_n_min,
                    direction = "both",
                    max_depth = cfg$dina_max_depth,
                    grid_probs = cfg$dina_grid_probs,
                    sg_focus = focus,
                    effect_neighborhood = cfg$dina_eff_neighbor,
                    alpha = cfg$dina_alpha),
      error = function(e) NULL) else NULL
    .sec <- (proc.time()[3] - .t0) + .fit_share
    r <- score_dina(dsg, df, truth, method_label = sprintf("DINA-%s", focus))
    r[, fit_seconds := .sec]
    r
  })

  scen_cols <- data.table(
    scenario_id = scenario$scenario_id, label = scenario$label,
    n = scenario$n, tau_in = scenario$tau_in, is_null = isTRUE(scenario$is_null),
    fs_maxk = cfg$fs_maxk, dina_max_depth = cfg$dina_max_depth, rep_id = rep_id,
    n_true_S = sum(df[["flag_harm"]] == 1L))
  method_rows <- c(fs_rows, grf_rows, grf_fr_rows, dina_rows)
  rbindlist(lapply(method_rows, function(r) cbind(scen_cols, r)))
}
```

---

# Run the study

```{r build-scenarios}
main <- CJ(n = CONFIG$n_levels, eff = names(CONFIG$tau_in))
main[, tau_in := CONFIG$tau_in[eff]]
main[, label := sprintf("n=%d, %s (tau_in=%.1f)", n, eff, tau_in)]
main[, is_null := FALSE]

nullarm <- data.table(n = CONFIG$n_levels, eff = "null", tau_in = 0.0,
  label = sprintf("n=%d, NULL", CONFIG$n_levels), is_null = TRUE)

scenarios <- rbind(main[, .(n, eff, tau_in, label, is_null)], nullarm)
scenarios[, scenario_id := .I]
scenarios[]
```

```{r run-grid}
run_grid <- function(scenarios, cfg) {
  grid <- scenarios[rep(seq_len(.N), each = cfg$nsims)]
  grid[, rep_id := rep(seq_len(cfg$nsims), times = nrow(scenarios))]

  if (cfg$parallel == "sims") {
    res <- foreach::foreach(
      i = seq_len(nrow(grid)), .errorhandling = "pass",
      .options.future = list(
        packages = c("forestsearch", "survival", "data.table", "future"),
        seed = TRUE)
    ) %dofuture% {
      future::plan("sequential")
      run_one_rep(as.list(grid[i]), grid$rep_id[i], cfg)
    }
  } else {
    res <- lapply(seq_len(nrow(grid)), function(i)
      run_one_rep(as.list(grid[i]), grid$rep_id[i], cfg))
  }
  ok <- vapply(res, is.data.table, logical(1))
  if (any(!ok)) message(sprintf("%d/%d reps errored and were dropped.",
                                sum(!ok), length(res)))
  rbindlist(res[ok], fill = TRUE)
}

t0 <- proc.time()[3]
results <- run_grid(scenarios, CONFIG)
timings$study <- proc.time()[3] - t0
message(sprintf("Study completed in %.1f min (%d method-rows).",
                timings$study / 60, nrow(results)))

if (CONFIG$parallel == "sims") {
  forestsearch::reset_workers(workers = n_workers)
}
```

---

# Results

## True subgroup size

The true subgroup is $S = \{x_1 \ge 0\}$ with $x_1 \sim U[-1,1]$, so each
subject is in $S$ with probability $P(x_1 \ge 0) = 0.5$ and
$|S| \sim \text{Binomial}(n, 0.5)$.

```{r true-S-size}
marg_p <- function(dir, cut) if (dir == "left") (cut - (-1)) / 2 else (1 - cut) / 2
true_p <- prod(mapply(marg_p, CONFIG$true_dirs, CONFIG$true_cuts))

trueS <- unique(results[is_null == FALSE,
  .(scenario_id, n, label, rep_id, n_true_S)])

trueS_summary <- trueS[, .(
  expected_p     = true_p,
  expected_n     = true_p * n[1],
  mean_n_true_S  = mean(n_true_S),
  sd_n_true_S    = sd(n_true_S),
  min_n_true_S   = min(n_true_S),
  max_n_true_S   = max(n_true_S),
  mean_frac      = mean(n_true_S / n)
), by = .(n)][order(n)]

trueS_summary |>
  gt::gt() |>
  gt::tab_header(
    title = gt::md("**True subgroup size:** $S = \\{x_1 \\ge 0\\}$"),
    subtitle = sprintf("x1 ~ U[-1,1]; |S| ~ Binomial(n, %.2f)", true_p)) |>
  gt::fmt_number(columns = c(expected_p, mean_frac), decimals = 3) |>
  gt::fmt_number(columns = c(expected_n, mean_n_true_S, sd_n_true_S),
                 decimals = 1) |>
  gt::cols_label(n = "n", expected_p = "E[p]", expected_n = "E|S|",
                 mean_n_true_S = "mean |S|", sd_n_true_S = "sd |S|",
                 min_n_true_S = "min |S|", max_n_true_S = "max |S|",
                 mean_frac = "mean |S|/n") |>
  gt::tab_source_note(gt::md(
    "*`|S|` is random across replications. A method recovering the true
     subgroup should report a discovered size near `mean |S|`; a size near `n`
     indicates the subgroup was extended well beyond the truth.*")) |>
  gt::tab_options(table.font.size = 12)
```

## Detection and recovery, by method (non-null cells)

```{r agg-main}
# NA-safe mean: an all-NA slice returns NA (not NaN, which gt prints literally).
mean_or_na <- function(x) {
  x <- x[!is.na(x)]
  if (length(x)) mean(x) else NA_real_
}

agg <- results[is_null == FALSE, .(
  detection_rate    = mean(found),
  covariate_set_hit = mean_or_na(covariate_set_hit[found]),
  signature_correct = mean_or_na(signature_correct[found]),
  sensitivity       = mean(sensitivity, na.rm = TRUE),
  specificity       = mean(specificity, na.rm = TRUE),
  ppv               = mean(ppv, na.rm = TRUE),
  npv               = mean(npv, na.rm = TRUE),
  mean_sg_frac      = mean((n_subgroup / n)[found], na.rm = TRUE),
  mean_n_factors    = mean(n_factors[found], na.rm = TRUE)
), by = .(label, n, tau_in, method)][order(n, tau_in, method)]

agg |>
  gt::gt(groupname_col = "label") |>
  gt::tab_header(
    title = gt::md("**Two-factor signature identification: forestsearch vs dina_subgroup**"),
    subtitle = sprintf(
      "Cox; true S = {x1 >= 0}; HR threshold = %.2f; maxk = %d / max_depth = %d; %d sims/cell",
      CONFIG$effect_threshold, CONFIG$fs_maxk, CONFIG$dina_max_depth, CONFIG$nsims)) |>
  gt::fmt_number(columns = c(detection_rate, covariate_set_hit, signature_correct,
                             sensitivity, specificity, ppv, npv,
                             mean_sg_frac, mean_n_factors), decimals = 3) |>
  gt::sub_missing(columns = gt::everything(), missing_text = "\u2014") |>
  gt::cols_label(method = "method", detection_rate = "detected",
                 covariate_set_hit = "cov-set hit", signature_correct = "sig correct",
                 sensitivity = "sens", specificity = "spec", ppv = "ppv",
                 npv = "npv", mean_sg_frac = "|S-hat|/n",
                 mean_n_factors = "mean #factors", n = "n", tau_in = "tau_in") |>
  gt::tab_source_note(gt::md(
    "*`cov-set hit` = P(discovered rule uses exactly {x1, x2} | found).
     `sig correct` = P(discovered (covariate, direction) axes equal the truth |
     found). It is computed for **both** methods: `dina_subgroup` exposes the
     axes directly, and forestsearch's cut labels are parsed (negations flip
     the direction); an em-dash means no found replications (or an unparseable
     label). `sens`/`spec`/`ppv`/`npv` compare discovered membership to the
     true subgroup mask. `mean #factors` should sit near 1 when the
     single-factor truth is recovered.*")) |>
  gt::tab_options(table.font.size = 12)
```

## Signature axis recovery and threshold bias (dina_subgroup; axis-correct reps)

For `dina_subgroup()` the axis is explicit, so among replications that recover
the exact `(covariate, direction)` axis we can report threshold bias against the
true cut $c_1 = 0$.

```{r agg-threshold}
thr <- results[startsWith(method, "DINA") & is_null == FALSE &
                 signature_correct == TRUE, .(
  n_axis_correct = .N,
  bias_x1 = mean(te_1, na.rm = TRUE), rmse_x1 = sqrt(mean(te_1^2, na.rm = TRUE))
), by = .(label, n, tau_in, method)][order(n, tau_in, method)]

if (nrow(thr)) {
  thr |>
    gt::gt() |>
    gt::tab_header(
      title = gt::md("**dina_subgroup threshold recovery, by selection rule**"),
      subtitle = sprintf("true cut x1 = %.2f; m_diff = %.3f",
                         CONFIG$true_cuts[1], CONFIG$m_diff)) |>
    gt::fmt_number(columns = c(bias_x1, rmse_x1), decimals = 3) |>
    gt::cols_label(label = "Scenario", n = "n", tau_in = "tau_in",
                   method = "focus", n_axis_correct = "axis-correct reps",
                   bias_x1 = "bias x1", rmse_x1 = "RMSE x1") |>
    gt::tab_source_note(gt::md(
      "*Among reps recovering the axis, `effMaxSG`'s effect-neighborhood band
       anchors the cut near the truth (small bias), while the max-effect rule
       (`eff`) can push the cut outward as it chases the largest effect.*")) |>
    gt::tab_options(table.font.size = 12)
} else cat("No axis-correct DINA reps to summarise.\n")
```

## False-discovery rate (null arm), by method

```{r agg-null}
fdr <- results[is_null == TRUE, .(
  false_discovery_rate = mean(found)
), by = .(n, method)][order(n, method)]

fdr |>
  gt::gt(groupname_col = "n") |>
  gt::tab_header(
    title = gt::md("**False-discovery rate (null: no heterogeneity)**"),
    subtitle = sprintf("HR threshold = %.2f; maxk = %d / max_depth = %d; %d sims/cell",
                       CONFIG$effect_threshold, CONFIG$fs_maxk,
                       CONFIG$dina_max_depth, CONFIG$nsims)) |>
  gt::fmt_number(columns = false_discovery_rate, decimals = 3) |>
  gt::cols_label(method = "method",
                 false_discovery_rate = "false-discovery rate") |>
  gt::tab_source_note(gt::md(
    "*Under no heterogeneity any discovered subgroup is a false positive. The
     single-factor hypothesis space (`maxk = 1` / `max_depth = 1`) keeps the
     candidate space small; the harm thresholds remain the guard.*")) |>
  gt::tab_options(table.font.size = 12)
```

# Graphical summary

The displays below reuse the operating-characteristic bubble vocabulary from
the `binary_m1_harm` summary (`plot_oc_metrics()` / `compute_oc_pareto_flag()`
in `quarto/_utils/`). The signature-recovery aggregates are reshaped onto that
convention by `sigrec_to_display()` (and `sigrec_fdr_to_display()` for the null
arm), so color = method (`group`), shape = method family (`Analysis`:
FS-consistency / FS-grf / DINA), and the chosen metrics map onto the x / y / size
axes. All metrics are higher-is-better except `|S-hat|/n`, which is read against
the true subgroup fraction (0.5 here), not maximised.

```{r gfx-reshape}
display_dt <- sigrec_to_display(agg)
# Pareto frontier over the three recovery axes that matter here, all
# higher-is-better.  (|S-hat|/n is excluded: it is a target-matching metric,
# not higher-better, so it does not belong in a dominance calculation.)
display_dt <- compute_oc_pareto_flag(
  display_dt, metrics = c("Sen", "Spec", "CovHit"))
fdr_dt <- sigrec_fdr_to_display(fdr)
```

## Sensitivity × specificity (bubble size = covariate-set hit)

The single most informative view: where each method lands in
sensitivity/specificity space, with bubble size showing how often it recovered
the true covariate set. A method that recovers the signature *correctly* sits
toward the top-right (high sensitivity **and** high specificity) with a large
bubble; a method that over-extends sits low-right (high sensitivity, low
specificity).

```{r gfx-sens-spec, fig.width = 9, fig.height = 4.8}
plot_oc_metrics(
  display_dt[display_dt$tau_in == max(display_dt$tau_in), ],
  x = "Sen", y = "Spec", z = "CovHit",
  size_limits = c(0, 1),
  pareto_col  = "pareto_frontier",
  title    = "Signature recovery: sensitivity \u00d7 specificity (bubble = cov-set hit)",
  subtitle = sprintf("Strong signal (tau_in = %.1f), n = %d; frontier on {Sen, Spec, CovHit} in solid",
                     max(display_dt$tau_in), max(display_dt$n)))
```

## Subgroup-size accuracy (`|S-hat|/n` vs the truth)

The size view exposes over-extension that the sensitivity/specificity bubble can
hide. The dashed line is the true subgroup fraction; points to its right have
discovered a subgroup *larger* than the truth (membership extended beyond the
true region), points on the line have matched its size.

```{r gfx-size, fig.width = 9, fig.height = 5}
true_frac <- 0.5  # true |S|/n for S = {x1 >= 0}
ggplot(display_dt,
       aes(x = Shat_frac, y = group, colour = group, shape = Analysis)) +
  geom_vline(xintercept = true_frac, linetype = "dashed",
             colour = "grey40") +
  geom_point(size = 3) +
  facet_grid(. ~ paste0("n = ", n) + paste0("tau = ", tau_in)) +
  scale_x_continuous(limits = c(0, 1)) +
  labs(x = "|S-hat| / n  (dashed = true 0.5)", y = NULL,
       title = "Discovered subgroup size vs. the truth",
       subtitle = "Points right of the dashed line over-extend beyond the true subgroup") +
  theme_bw() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

## False-discovery rate (null arm)

The FDR story is categorical -- a couple of methods near 1, the rest near 0 --
so a bar panel reads better than a bubble. Each method's false-discovery rate
under no heterogeneity, by sample size; the dashed line marks a nominal 0.05
reference.

```{r gfx-fdr, fig.width = 9, fig.height = 5}
ggplot(fdr_dt,
       aes(x = FDR, y = reorder(group, FDR), fill = Analysis)) +
  geom_col(width = 0.7) +
  geom_vline(xintercept = 0.05, linetype = "dashed", colour = "grey40") +
  facet_wrap(~ paste0("n = ", n), nrow = 1) +
  scale_x_continuous(limits = c(0, 1)) +
  labs(x = "false-discovery rate (null)", y = NULL,
       title = "False-discovery rate under no heterogeneity",
       subtitle = "Dashed line = 0.05 reference; GRF arms inflate severely at dmin.grf = 0") +
  theme_bw() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

```{r plan-teardown}
future::plan("sequential")
gc(verbose = FALSE)
```

---

# Interpretation

```{r interpretation, echo=FALSE, results='asis'}
fmt_pct <- function(x) if (length(x) && is.finite(x)) sprintf("%.0f%%", 100 * x) else "NA"
getm <- function(dt, m) { r <- dt[method == m]; if (nrow(r)) r else dt[0] }
cell <- agg[n == max(n) & tau_in == CONFIG$tau_in[["strong"]]]
dn_ef <- getm(cell, "DINA-effMaxSG"); dn_e <- getm(cell, "DINA-eff")
fs_ef <- getm(cell, "FS-effMaxSG");   fs_e <- getm(cell, "FS-eff")
grf   <- getm(cell, "FS-grf-tree"); grf_fr <- getm(cell, "FS-grf-frontier")

cat(sprintf("
At the strongest signal (n = %d, tau_in = %.1f, maxk = max_depth = grf_depth = 1):

* **DINA (effMaxSG)** detected in %s, recovered the {x1} covariate in %s of
  detections, sensitivity %s, specificity %s, |S-hat|/n %s.
* **DINA (eff)** detected in %s, cov-set hit %s, sensitivity %s, specificity %s.
* **FS-consistency (effMaxSG)** detected in %s, cov-set hit %s, sensitivity %s,
  specificity %s.
* **FS-consistency (eff)** detected in %s, cov-set hit %s, sensitivity %s,
  specificity %s.
* **FS-grf-tree** detected in %s, cov-set hit %s, sensitivity %s, specificity %s.
* **FS-grf-frontier (effMaxSG)** detected in %s, cov-set hit %s, sensitivity %s,
  specificity %s.

This is the single-factor case, the most favourable setting for DINA's
univariate search. All engines target the same one-covariate truth; the
consistency and DINA arms vary the selection rule (`effMaxSG` vs `eff`), while
the two GRF arms differ in how they select from the same doubly-robust scores
(policy tree vs Pareto frontier under effMaxSG).
",
  max(CONFIG$n_levels), CONFIG$tau_in[["strong"]],
  fmt_pct(dn_ef$detection_rate), fmt_pct(dn_ef$covariate_set_hit),
  fmt_pct(dn_ef$sensitivity), fmt_pct(dn_ef$specificity), fmt_pct(dn_ef$mean_sg_frac),
  fmt_pct(dn_e$detection_rate), fmt_pct(dn_e$covariate_set_hit),
  fmt_pct(dn_e$sensitivity), fmt_pct(dn_e$specificity),
  fmt_pct(fs_ef$detection_rate), fmt_pct(fs_ef$covariate_set_hit),
  fmt_pct(fs_ef$sensitivity), fmt_pct(fs_ef$specificity),
  fmt_pct(fs_e$detection_rate), fmt_pct(fs_e$covariate_set_hit),
  fmt_pct(fs_e$sensitivity), fmt_pct(fs_e$specificity),
  fmt_pct(grf$detection_rate), fmt_pct(grf$covariate_set_hit),
  fmt_pct(grf$sensitivity), fmt_pct(grf$specificity),
  fmt_pct(grf_fr$detection_rate), fmt_pct(grf_fr$covariate_set_hit),
  fmt_pct(grf_fr$sensitivity), fmt_pct(grf_fr$specificity)))
```

Points the tables support:

* **Matched hypothesis space.** With `fs_maxk = 1`, `dina_max_depth = 1`, and
  `grf_depth = 1`, every engine targets the single-covariate truth, so
  differences reflect the search and selection rule rather than an
  under-specified model. This is the ideal case for DINA's univariate search;
  the two-factor companion shows the cost when the truth is a genuine
  conjunction.
* **Selection rule, not search, drives the size.** For the consistency and DINA
  arms, `effMaxSG` tightens the subgroup toward the truth (higher specificity,
  `|S-hat|/n` near the true 0.5), while the max-effect rule `eff` chases the
  largest effect. The `*-effMaxSG` vs `*-eff` pairs isolate this.
* **Two GRF arms on the same scores.** `FS-grf-tree` selects via the depth-1
  policy tree; `FS-grf-frontier` selects from the Pareto frontier of the same
  doubly-robust scores under effMaxSG. Comparing them isolates the effect of the
  selection step alone (tree vs frontier) holding the forest fixed. Neither uses
  `sg_focus` to vary a rule across rows, so each appears once.
* **Multi-focus cost.** `dina_foci` reuses one cross-fitted `dina()` fit across
  foci (cheap); `fs_foci` re-runs `forestsearch()` per focus. The per-method
  timing tables quantify this directly.
* **Threshold recovery.** For `dina_subgroup()`, among reps recovering the axis,
  bias against $c_1 = 0$ is reported in the threshold table; `effMaxSG` keeps
  the cut near the truth.

This study isolates **discovery** behaviour on a single-factor truth. It does
not address post-selection inference (the conditional-on-signature CIs and the
selection-adjusted bootstrap), which are separate questions.

# Graphical summary

The tables above are reshaped onto the operating-characteristic display
convention used by the cross-cut simulation articles
(`_utils/plot_oc_metrics.R`), so the same bubble-chart and Pareto-frontier
machinery applies. `sigrec_to_display()` renames the per-method metrics to the
`Detection / Sen / Spec / PPV` convention, maps `method` to color and the method
*family* (FS-consistency / FS-grf / DINA) to shape, and carries the
recovery-specific metrics (`CovHit`, `SigCorrect`, `Shat_frac`)
alongside.

```{r gfx-display-dt}
display_dt <- sigrec_to_display(agg)
# Pareto frontier on the recovery-relevant axes (all higher-is-better):
# sensitivity, specificity, and covariate-set hit.  A method is on the frontier
# if no other matches it on all three and beats it on at least one.
display_dt <- compute_oc_pareto_flag(
  display_dt, metrics = c("Sen", "Spec", "CovHit"))
```

## Sensitivity x specificity (bubble size = covariate-set hit)

The single most informative view: the x-axis is sensitivity, the y-axis is
specificity, and bubble size is the covariate-set hit rate. A method that
recovers the truth tightly sits in the upper-right with a large bubble; a method
that over-extends (captures the harmed plus much of the unharmed) drifts toward
high sensitivity but low specificity, with a smaller bubble. Faceted by `n`;
color = method, shape = method family.

```{r gfx-sen-spec, fig.width=10, fig.height=4.2}
plot_oc_metrics(
  display_dt,
  x = "Sen", y = "Spec", z = "CovHit",
  size_limits = c(0, 1),
  title    = "Signature recovery: sensitivity x specificity (bubble size = cov-set hit)",
  subtitle = "Color = method; shape = family (FS-consistency / FS-grf / DINA). Upper-right with a large bubble = tight recovery."
) + ggplot2::facet_wrap(~ n, labeller = ggplot2::labeller(n = function(v) paste0("n = ", v)))
```

## Pareto frontier on {sensitivity, specificity, cov-set hit}

The same axes, highlighting the Pareto frontier (solid, thicker stroke) and
dimming dominated methods. The frontier picks out the defensible operating
points across the three recovery axes simultaneously.

```{r gfx-pareto, fig.width=10, fig.height=4.2}
plot_oc_metrics(
  display_dt,
  x = "Sen", y = "Spec", z = "CovHit",
  pareto_col  = "pareto_frontier",
  size_limits = c(0, 1),
  title    = "Pareto frontier on {Sen, Spec, cov-set hit} -- frontier solid, dominated dimmed",
  subtitle = "Frontier computed on the three recovery axes, not on the two plotted axes alone."
) + ggplot2::facet_wrap(~ n, labeller = ggplot2::labeller(n = function(v) paste0("n = ", v)))
```

## Subgroup-size accuracy

Discovered subgroup size as a fraction of `n` (`Shat_frac`), against the true
proportion (dashed line). This exposes over-extension as a *size* failure: a
method whose point sits well above the true fraction is flagging more subjects
as harmed than the truth contains, even if its sensitivity looks high.

```{r gfx-sizeb, fig.width=9, fig.height=4.5}
# True |S|/n for S = {x1 >= 0} with x1 ~ U[-1, 1] is 0.5.  Use a CONFIG override
# if present, else default to 0.5 (this single-factor DGM).
true_frac <- if (!is.null(CONFIG$true_frac)) CONFIG$true_frac else 0.5
ggplot2::ggplot(
  display_dt,
  ggplot2::aes(x = Shat_frac, y = reorder(method, Shat_frac),
               color = group, shape = Analysis)) +
  ggplot2::geom_vline(xintercept = true_frac, linetype = "dashed",
                      color = "grey40") +
  ggplot2::geom_point(size = 3) +
  ggplot2::facet_wrap(~ n, labeller = ggplot2::labeller(n = function(v) paste0("n = ", v))) +
  ggplot2::labs(
    x = "|S-hat| / n  (discovered subgroup size as fraction of n)",
    y = NULL, color = "method", shape = "family",
    title = "Subgroup-size accuracy vs. the true fraction (dashed)",
    subtitle = sprintf("True |S|/n = %.2f. Points far to the right over-extend the subgroup.", true_frac)) +
  ggplot2::theme_bw(base_size = 11) +
  ggplot2::theme(legend.position = "right")
```

# Timing

## Time per method

How the study's wall-clock time divides across the methods. Each method's
per-call elapsed time is recorded inside `run_one_rep()`; because replications
run concurrently across workers under a parallel plan, the raw sum of per-call
times is total *compute*, not elapsed time. To report real time, that compute
is expressed as a **share** and applied to the measured wall-clock for the run
(`timings$study`), so the `wall-clock min` column sums to the actual elapsed
study time. (Under `parallel = "none"` the share and the raw compute coincide.)

```{r timing-per-method}
method_time <- results[, .(compute_min = sum(fit_seconds, na.rm = TRUE) / 60),
                        by = .(method)]
method_time[, share := compute_min / sum(compute_min)]
# Apportion the real elapsed wall-clock (timings$study, in seconds) by share.
method_time[, wall_min := share * (timings$study / 60)]
method_time <- method_time[order(-share)]

method_time[, .(method, share, wall_min)] |>
  gt::gt() |>
  gt::tab_header(
    title = gt::md("**Time per method**"),
    subtitle = sprintf("study wall-clock = %.2f min, apportioned by compute share",
                       timings$study / 60)) |>
  gt::fmt_percent(columns = share, decimals = 1) |>
  gt::fmt_number(columns = wall_min, decimals = 2) |>
  gt::cols_label(method = "method", share = "share of compute",
                 wall_min = "wall-clock min") |>
  gt::tab_source_note(gt::md(
    "*`share of compute` is the method's fraction of total fit/selection time;
     `wall-clock min` applies that share to the measured study wall-clock, so
     the column sums to the real elapsed time. DINA's shared `dina()` fit is
     split across its foci.*")) |>
  gt::tab_options(table.font.size = 12)
```

## Overall timing

```{r timing}
timings$total <- Reduce(`+`, timings)
data.frame(Step = names(timings),
           Minutes = round(unlist(timings) / 60, 2)) |>
  gt::gt() |>
  gt::tab_header(title = "Computation timing (wall-clock)") |>
  gt::tab_style(style = gt::cell_text(weight = "bold"),
                locations = gt::cells_body(rows = Step == "total"))
```

# Session information

```{r session-info}
sessionInfo()
```
