---
title: "CRAN Submission Preparation — `forestsearch` v0.1.0"
subtitle: "Branch: `cran-review-1` | Final Status: Ready for win-builder"
date: "2026-03-18"
format:
  html:
    toc: true
    toc-depth: 3
    theme: flatly
    code-fold: true
    embed-resources: true
---

## Final Check Result

```
── R CMD check results ──────────── forestsearch 0.1.0 ────
Duration: 4m 39.2s

0 errors ✔ | 0 warnings ✔ | 0 notes ✔
```

**Environment:** macOS Tahoe 26.3.1, R 4.5.2 (aarch64-apple-darwin20)

All checks pass cleanly — including vignette building, standard examples, and
`--run-donttest` examples (3m 20s).

---

## CRAN Reviewer Response

The CRAN reviewer raised eight issues. All have been resolved.

```{r}
#| label: tbl-cran-fixes
#| echo: false
#| tbl-cap: "CRAN reviewer issues — all resolved"

cran_fixes <- data.frame(
  Fix = 1:8,
  `Reviewer Issue` = c(
    "Single quotes around acronyms; extra spaces in Description field",
    "Commented-out lines in examples",
    "options(warn = -1) is not allowed",
    "par() not restored via on.exit() in function",
    "Functions write to user's home filespace by default",
    "par() not restored in vignette examples",
    "Missing \\\\value tags in 13 .Rd files",
    "Examples for unexported functions"
  ),
  File = c(
    "DESCRIPTION",
    "R/bootstrap_ystar.R",
    "R/bootstrap_dofuture_main.R",
    "R/subgroup_consistency_helpers.R",
    "R/cox_ahr_cde_wrapper.R",
    "vignettes/forestsearch.Rmd",
    "13 R source files",
    "4 R source files"
  ),
  Resolution = c(
    "Removed quotes from GRF, MRCT, AFT; collapsed Description whitespace",
    "Removed commented-out example lines",
    "Removed old_warn / options(warn = -1) entirely",
    "Added oldpar <- par(); on.exit(par(oldpar), add = TRUE) before par(mfrow)",
    "Changed default to output_dir = tempdir()",
    "Same on.exit pattern in grf-trees chunk; removed manual par(mfrow = c(1,1))",
    "Added @return for all functions (named list, side-effect, or invisible(x))",
    "Promoted 4 functions from @keywords internal to @export"
  ),
  check.names = FALSE
)

knitr::kable(cran_fixes, align = c("c", "l", "l", "l"))
```

### Fix 7 — `@return` Tag Details

| Return type | Functions | Tag pattern |
|:---|:---|:---|
| Named list | `generate_aft_dgm_flex`, `get_FSdata` | `\\item{name}{description}` for each component |
| Side-effect only | `plot_spline_treatment_effect`, `print_grf_details` | `No return value, called for side effects` |
| S3 invisible return | 9 print/plot methods | `Invisibly returns \\code{x}.` with `invisible(x)` in body |

**Revision:** The `@return` block for `get_FSdata()` originally documented
six components that did not match the actual return value. Corrected to
document the five actual components: `df`, `confs_names`, `confs`,
`lassokeep`, `lassoomit`.

### Fix 8 — Exported Functions

| Function | File |
|:---|:---|
| `create_sample_size_table` | R/truefind_asymptotic.R |
| `detect_variable_types` | R/synthetic_data_perturbation.R |
| `extract_selected_tree_cuts` | R/grf_helpers.R |
| `find_quantile_for_proportion` | R/find_k_inter_main.R |

All four verified: `@export` present, parameter documentation matches
signature, `@return` documented, `@importFrom` tags present where needed.

---

## Additional Fixes Discovered During `--run-donttest`

Running `devtools::check(args = "--as-cran")` with `--run-donttest`
revealed a series of example failures beyond the original eight issues.
These were resolved iteratively across multiple check cycles.

### Non-Self-Contained Examples

Ten `\donttest{}` blocks referenced objects never defined within the example
(`mydata`, `trial_data`, `df_large`, `fs_bootstrap`, bare `fs`, etc.).
Each was made self-contained using a standard GBSG setup pattern:

```r
df <- survival::gbsg
df$grade3 <- as.integer(df$grade == "3")
df$id <- seq_len(nrow(df))
fs <- forestsearch(df,
  confounders.name = c("age", "meno", "size", "grade3", "nodes", "pgr", "er"),
  outcome.name = "rfstime", event.name = "status", treat.name = "hormon",
  fs.splits = 100, max_subgroups_search = 5,
  use_twostage = TRUE, details = FALSE, plot.sg = FALSE)
```

Three examples could not be made self-contained due to complex prerequisites
and were replaced with descriptive stubs:

- `extract_selected_tree_cuts()` — requires fitted policy tree objects
- `plot_spline_treatment_effect()` — requires a DGM with spline-based
  treatment heterogeneity
- `print.cox_ahr_cde` / `summary.cox_ahr_cde` — `cox_ahr_cde_analysis()`
  requires specialised columns not present in raw GBSG

### Bare `gbsg` References

Since `survival` is in Imports (not Depends), bare `gbsg` is not available
in the check environment. Fifteen `\donttest{}` blocks across 6 files were
converted to use `survival::gbsg` or the `df <- survival::gbsg` setup
pattern.

### GBSG Dataset Integration

All examples using `survival::gbsg` required two accommodations:

**1. ID column.** `forestsearch()` defaults to `id.name = "id"`, but
`survival::gbsg` has `pid`. The internal `merge()` at line 720 of
`forestsearch_main.R` is hardcoded to `by = "id"`, so `id.name = "pid"`
does not work. The workaround is `df$id <- seq_len(nrow(df))`. This is
a latent bug documented below for a future patch.

**2. Column names for downstream functions.** `plot_sg_weighted_km()`
defaults to internal names (`Y`, `Event`, `Treat`), but `df.est` retains
original column names after the merge. Examples now pass column names
explicitly.

### Wrong Parameter Names

| File | Problem | Resolution |
|:---|:---|:---|
| `R/cox_ahr_cde_wrapper.R` (print/summary methods) | `outcome.name` etc. don't exist on `cox_ahr_cde_analysis()` (uses `tte_name`) | Replaced with descriptive stubs |
| `R/generate_aft_dgm_helpers_outcome.R` | Old param names for `generate_aft_dgm_flex()` | Replaced with descriptive stub |
| `R/plot_sg_results.R` (print/plot methods) | `plot_subgroup()` is internal, not exported | Replaced with `plot_sg_results()` |

### Speed Optimisation

Default parameters produce correct but slow examples. All `forestsearch()`
calls in examples use fast settings to keep total `--run-donttest` time
under 5 minutes:

| Parameter | Default | Example value | Effect |
|:---|:---|:---|:---|
| `fs.splits` | 1000 | 100 | 10x fewer consistency splits |
| `max_subgroups_search` | 10 | 5 | 2x fewer candidates evaluated |
| `use_twostage` | TRUE | TRUE | Early stopping in consistency evaluation |
| `nb_boots` | — | 10 | Minimal bootstrap (illustration only) |
| `Kfolds` | nrow(df) | 5 | 5-fold instead of 686 LOO |
| `sims` (tenfold) | — | 2 | 2 repetitions instead of 200 |
| `parallel_args` | callr/6 workers | sequential | Avoids worker startup overhead in examples |

### Parallel Args Bug

`parallel_args = list(plan = "sequential")` without `show_message` causes
`if (show_message)` to fail on `NULL`. All example calls now include
`show_message = FALSE` explicitly.

### Code Bug Fix

`generate_gbsg_bootstrap_general()` in `R/synthetic_data_perturbation.R`
used `data("gbsg", package = "survival", envir = environment())` which
fails when `survival` is in Imports (not Depends). Replaced with
`gbsg <- survival::gbsg`.

### Deprecation Warnings

`create_gbsg_dgm()` and `simulate_from_gbsg_dgm()` emit `.Deprecated()`
warnings. All 8 calls in `R/sim_aft_gbsg.R` `\donttest{}` examples
wrapped in `suppressWarnings()`.

---

## Complete File Inventory

```{r}
#| label: tbl-files
#| echo: false
#| tbl-cap: "All modified files and change types"

files <- data.frame(
  File = c(
    "DESCRIPTION",
    "vignettes/forestsearch.Rmd",
    "R/bootstrap_dofuture_main.R",
    "R/cox_ahr_cde_wrapper.R",
    "R/forestsearch_cross_validation.R",
    "R/generate_aft_dgm_helpers_outcome.R",
    "R/get_fsdata.R",
    "R/grf_helpers.R",
    "R/plot_sg_results.R",
    "R/plot_sg_weighted_km.R",
    "R/render_forestplot.R",
    "R/sim_aft_gbsg.R",
    "R/subgroup_consistency_helpers.R",
    "R/synthetic_data_perturbation.R"
  ),
  Changes = c(
    "Acronym quotes removed; Description whitespace fixed",
    "par() on.exit fix in grf-trees chunk",
    "warn=-1 removed; example made self-contained with speed params",
    "tempdir() default; main example rewritten; print/summary stubs",
    "6 example blocks: self-contained, speed params, Kfolds reduced",
    "@return tag; example replaced with descriptive stub",
    "@return corrected to match actual return value",
    "extract_selected_tree_cuts example replaced with stub",
    "3 example blocks: self-contained, plot_subgroup -> plot_sg_results",
    "2 example blocks: self-contained, column names passed explicitly",
    "5 example blocks: self-contained, trimmed, bare gbsg fixed",
    "8 deprecated calls wrapped in suppressWarnings",
    "par() on.exit fix; 3 example blocks made self-contained",
    "data() bug fixed; bare gbsg converted; broken data() in example removed"
  ),
  `Function Code Changed` = c(
    "—", "—", "Yes", "Yes", "No", "No", "No", "No",
    "No", "No", "No", "No", "Yes", "Yes"
  ),
  check.names = FALSE
)

knitr::kable(files, align = c("l", "l", "c"))
```

Four files had changes to function bodies (not just examples/documentation):

1. **`bootstrap_dofuture_main.R`** — `options(warn = -1)` and `old_warn`
   restore removed
2. **`cox_ahr_cde_wrapper.R`** — `output_dir` default changed from `NULL`
   to `tempdir()`
3. **`subgroup_consistency_helpers.R`** — `on.exit(par(oldpar), add = TRUE)`
   added to `plot_subgroup()`
4. **`synthetic_data_perturbation.R`** — `data("gbsg", ...)` replaced with
   `survival::gbsg` in `generate_gbsg_bootstrap_general()`

---

## Known Issues for Future Releases

These items do not block the current submission but should be addressed
in a patch release.

### Hardcoded `merge(by = "id")` in `forestsearch()`

Line 720 of `R/forestsearch_main.R`:

```r
df.est_out <- merge(df, temp, by = "id", all.x = TRUE)
```

This ignores the `id.name` parameter. Users passing `id.name = "pid"` will
hit an error. The fix is `by = id.name` on lines 720 and 724, plus an
audit of downstream code that assumes column name `"id"`.

### `oc_analyses_gbsg.R` Retains `\dontrun{}`

Four examples use `\dontrun{}`. If CRAN asks to convert to `\donttest{}`,
they will need `suppressWarnings()` wrappers around `create_gbsg_dgm()`
calls.

### Vignette Packaging on R 4.3.3

`R CMD build` on Linux with R 4.3.3 produces an empty `inst/doc/`,
causing installation failure. This does not occur on macOS with R 4.5.2.
Not a submission blocker but may affect users building from source on
older R versions.

---

## Submission Roadmap

```{r}
#| label: tbl-checklist
#| echo: false
#| tbl-cap: "Pre-submission checklist"

checklist <- data.frame(
  Step = c(
    "All 8 CRAN reviewer issues resolved",
    "R CMD check --as-cran: 0 errors, 0 warnings, 0 notes",
    "--run-donttest examples pass (3m 20s)",
    "Vignette builds and installs correctly",
    "Changes committed to cran-review-1 branch",
    "Submit to win-builder R-devel",
    "Submit to win-builder R-release",
    "Update cran-comments.md with win-builder results",
    "Submit via CRAN submission portal"
  ),
  Status = c(
    "Done", "Done", "Done", "Done", "Done",
    "Next", "Next", "Pending", "Pending"
  )
)

knitr::kable(checklist, align = c("l", "c"))
```

### win-builder Submission

```r
devtools::check_win_devel()
devtools::check_win_release()
```

Results will be emailed to `larry.leon.05@post.harvard.edu`. Once both
pass, update `cran-comments.md` and submit via
`https://cran.r-project.org/submit.html`.
