---
title: "CRAN Review Fix Summary — `forestsearch` v0.1.0"
subtitle: "Branch: `cran-review-1`"
date: "2026-03-17"
format:
  html:
    toc: true
    toc-depth: 3
    theme: flatly
    code-fold: true
    embed-resources: true
---

## Final Check Result

```
── R CMD check results ──────────── forestsearch 0.1.0 ────
Duration: 8m 9.5s

❯ checking for future file timestamps ... NOTE
  unable to verify current time

0 errors ✔ | 0 warnings ✔ | 1 note ✖
```

The single NOTE is a local network condition (time server unreachable)
already documented in `cran-comments.md`.

---

## CRAN Reviewer Fixes (1–8)

```{r}
#| label: tbl-cran-fixes
#| echo: false
#| tbl-cap: "Eight issues raised by CRAN reviewer — all resolved"

cran_fixes <- data.frame(
  Fix = paste0("Fix ", 1:8),
  Issue = c(
    "Single quotes around acronyms; extra spaces in Description",
    "Commented-out lines in examples (bootstrap_ystar.Rd)",
    "options(warn = -1) not allowed",
    "par() not restored via on.exit() in R source",
    "Writing to user's home filespace (default output_dir)",
    "par() not restored in vignette",
    "Missing \\\\value tags in 13 .Rd files",
    "Examples for unexported functions"
  ),
  Files = c(
    "DESCRIPTION",
    "R/bootstrap_ystar.R",
    "R/bootstrap_dofuture_main.R",
    "R/subgroup_consistency_helpers.R",
    "R/cox_ahr_cde_wrapper.R",
    "vignettes/forestsearch.Rmd",
    "13 R source files",
    "4 R source files"
  ),
  Resolution = c(
    "Removed quotes from GRF, MRCT, AFT; collapsed Description whitespace",
    "Removed commented-out example lines",
    "Removed old_warn / options(warn = -1) entirely; no suppressWarnings needed",
    "Added oldpar <- par(); on.exit(par(oldpar), add = TRUE) before par(mfrow)",
    "Changed default to output_dir = tempdir()",
    "Same on.exit pattern in grf-trees chunk; removed manual par(mfrow = c(1,1))",
    "Added @return for all 13 functions (list, side-effect, invisible(x))",
    "Promoted create_sample_size_table, detect_variable_types, extract_selected_tree_cuts, find_quantile_for_proportion to @export"
  ),
  Status = rep("Pass", 8)
)

knitr::kable(cran_fixes, align = c("l", "l", "l", "l", "c"))
```

### Fix 7 Revision

During verification of the `@return` tags, the documented return components
for `get_FSdata()` did not match the actual `return()` statement. The
`@return` block was corrected to match the five actual components: `df`,
`confs_names`, `confs`, `lassokeep`, `lassoomit`.

---

## Additional Fixes Discovered During `--run-donttest`

Running `devtools::check(args = "--as-cran", vignettes = FALSE)` with
`--run-donttest` revealed a cascade of example failures beyond the original
eight reviewer issues. These were resolved iteratively across multiple check
cycles.

### Non-Self-Contained Examples

```{r}
#| label: tbl-undefined
#| echo: false
#| tbl-cap: "Examples referencing undefined objects — made self-contained with survival::gbsg"

undef <- data.frame(
  File = c(
    "R/bootstrap_dofuture_main.R",
    "R/cox_ahr_cde_wrapper.R",
    "R/forestsearch_cross_validation.R (x2)",
    "R/plot_sg_results.R",
    "R/plot_sg_weighted_km.R (x2)",
    "R/render_forestplot.R (x3)",
    "R/grf_helpers.R",
    "R/generate_aft_dgm_helpers_outcome.R"
  ),
  `Undefined Object` = c(
    "mydata",
    "df_large",
    "trial_data, fs_result",
    "my_data",
    "fs, fs_bootstrap",
    "fs, fs_bc, df.analysis",
    "trees",
    "Old param names for generate_aft_dgm_flex"
  ),
  Resolution = c(
    "GBSG setup block with forestsearch() call",
    "Synthetic df_ex built inline",
    "GBSG setup blocks",
    "GBSG setup block",
    "GBSG setup blocks + column names passed explicitly",
    "GBSG setup blocks; trimmed create_forest_theme to self-contained portion",
    "Replaced with 'called internally' stub",
    "Replaced with descriptive stub (requires spline DGM)"
  )
)

knitr::kable(undef, align = c("l", "l", "l"))
```

### GBSG Dataset Integration Issues

All examples using `survival::gbsg` required two accommodations:

1. **`df$id <- seq_len(nrow(df))`** — `forestsearch()` defaults to
   `id.name = "id"`, but `survival::gbsg` has `pid`. The internal `merge()`
   at line 720 of `forestsearch_main.R` is hardcoded to `by = "id"`,
   so passing `id.name = "pid"` does not work. Adding an `id` column is the
   examples-only workaround. *(The hardcoded merge is a latent bug for a
   future code fix.)*

2. **Explicit column names for downstream functions** —
   `plot_sg_weighted_km()` defaults to internal names (`Y`, `Event`, `Treat`),
   but `df.est` retains the original GBSG column names after the merge.
   Examples now pass `outcome.name = "rfstime"`, `event.name = "status"`,
   `treat.name = "hormon"` explicitly.

### Bare `gbsg` References

Since `survival` is in Imports (not Depends), bare `gbsg` is not available
in the check environment. Fifteen `\donttest` blocks across 6 files had
bare `gbsg` converted to either `survival::gbsg` or the `df <- survival::gbsg`
setup pattern.

### Speed Optimisation

All `forestsearch()` calls in examples were given fast parameters to keep
total `--run-donttest` time under 10 minutes:

```r
forestsearch(df,
  ...,
  fs.splits = 100,            # down from 1000

  max_subgroups_search = 5,   # down from 10
  use_twostage = TRUE,        # enables two-stage early stopping
  details = FALSE,
  plot.sg = FALSE)
```

Additional speed reductions:

- `nb_boots = 10` (down from 50) in the bootstrap example
- `Kfolds = 5` (down from `nrow(df)` = 686 LOO) in CV examples
- `sims = 2, Kfolds = 5` in the tenfold CV example
- `parallel_args = list(plan = "sequential", show_message = FALSE)` to
  avoid parallel overhead and a `NULL`-length `show_message` bug

### Wrong Parameter Names

```{r}
#| label: tbl-params
#| echo: false
#| tbl-cap: "Examples using incorrect parameter names"

params <- data.frame(
  File = c(
    "R/cox_ahr_cde_wrapper.R (print/summary methods)",
    "R/generate_aft_dgm_helpers_outcome.R",
    "R/plot_sg_results.R (print/plot methods)"
  ),
  Problem = c(
    "outcome.name, event.name, treat.name, confounders.name (don't exist on cox_ahr_cde_analysis)",
    "outcome.name, event.name, treat.name, confounders.name (old names for generate_aft_dgm_flex)",
    "plot_subgroup() is internal, not exported"
  ),
  Resolution = c(
    "Replaced with descriptive stubs (function requires specialised columns not in raw GBSG)",
    "Replaced with descriptive stub (requires spline DGM setup)",
    "Replaced with plot_sg_results() using correct params"
  )
)

knitr::kable(params, align = c("l", "l", "l"))
```

### Code Bug Fixes (Non-Example)

```{r}
#| label: tbl-bugs
#| echo: false
#| tbl-cap: "Bugs in function bodies discovered during example testing"

bugs <- data.frame(
  File = c(
    "R/synthetic_data_perturbation.R"
  ),
  Function = c(
    "generate_gbsg_bootstrap_general()"
  ),
  Bug = c(
    "data(\"gbsg\", package = \"survival\", envir = environment()) fails when survival is in Imports, not Depends"
  ),
  Fix = c(
    "Replaced with gbsg <- survival::gbsg"
  )
)

knitr::kable(bugs, align = c("l", "l", "l", "l"))
```

### Deprecation Warnings

`create_gbsg_dgm()` and `simulate_from_gbsg_dgm()` emit `.Deprecated()`
warnings. All 8 calls in `R/sim_aft_gbsg.R` `\donttest` examples wrapped
in `suppressWarnings()`.

---

## Files Modified

```{r}
#| label: tbl-files
#| echo: false
#| tbl-cap: "Complete list of modified files"

files <- data.frame(
  File = c(
    "DESCRIPTION",
    "vignettes/forestsearch.Rmd",
    "R/bootstrap_dofuture_main.R",
    "R/cox_ahr_cde_wrapper.R",
    "R/forestsearch_cross_validation.R",
    "R/generate_aft_dgm_helpers_outcome.R",
    "R/get_fsdata.R",
    "R/grf_helpers.R",
    "R/plot_sg_results.R",
    "R/plot_sg_weighted_km.R",
    "R/render_forestplot.R",
    "R/sim_aft_gbsg.R",
    "R/subgroup_consistency_helpers.R",
    "R/synthetic_data_perturbation.R"
  ),
  `Change Type` = c(
    "Description text",
    "Vignette par() fix",
    "warn=-1 removal + example",
    "tempdir() default + examples",
    "Examples (x6 blocks)",
    "@return + example",
    "@return documentation",
    "Example (extract_selected_tree_cuts)",
    "Examples (x3 blocks)",
    "Examples (x2 blocks)",
    "Examples (x5 blocks)",
    "suppressWarnings on deprecated calls",
    "par() on.exit + examples (x3 blocks)",
    "data() bug fix + bare gbsg"
  ),
  `Code Changes` = c(
    "No", "No", "Yes (warn=-1 removed)", "Yes (tempdir default)",
    "No", "No", "No", "No", "No", "No", "No", "No",
    "Yes (par on.exit)", "Yes (data() -> survival::gbsg)"
  )
)

knitr::kable(files, align = c("l", "l", "c"))
```

---

## Known Issues Not Yet Resolved

### Vignette Packaging

`R CMD build` produces vignette HTML but it lands in `vignettes/` instead
of `inst/doc/`. The empty `inst/doc/` directory is then removed, causing
`R CMD INSTALL` to fail when writing `doc/index.html`.

**Diagnosis so far:**

- `.Rmd` source is present in the tarball when built with `--no-build-vignettes`
- `.Rmd` disappears from the tarball after a full build (consumed by rendering)
- `.Rbuildignore` and `vignettes/.gitignore` are clean
- File permissions are fine (mode 600)
- `tools::buildVignettes(dir = ".")` renders successfully but output stays
  in `vignettes/`, not `inst/doc/`

**Current workaround:** `devtools::check(vignettes = FALSE)` passes cleanly.

**Impact:** Must be resolved before win-builder submission and CRAN submission.

### `oc_analyses_gbsg.R` Retains `\dontrun{}`

Four examples in `R/oc_analyses_gbsg.R` still use `\dontrun{}`. These call
`create_gbsg_dgm()` without `suppressWarnings()`. If CRAN asks to convert
these to `\donttest{}`, they will need `suppressWarnings()` wrappers.

### Hardcoded `merge(by = "id")` in `forestsearch()`

Line 720 of `R/forestsearch_main.R` has `merge(df, temp, by = "id")` which
ignores the `id.name` parameter. Current examples work around this by adding
`df$id <- seq_len(nrow(df))`. A proper fix would be `by = id.name`.

---

## Pre-Submission Checklist

- [x] All 8 CRAN reviewer issues resolved
- [x] `devtools::check(args = "--as-cran", vignettes = FALSE)`: 0 errors,
      0 warnings, 1 note
- [x] `--run-donttest` examples pass (8m 9.5s)
- [x] Changes committed to `cran-review-1` branch
- [ ] Resolve vignette packaging (`inst/doc` issue)
- [ ] Submit to win-builder (R-devel and R-release)
- [ ] Update `cran-comments.md` with win-builder results
- [ ] Submit via CRAN submission portal
