---
title: "GBSG maxeff subgroup analysis: Naive, Full Bootstrap, Multiplier Resampling, and Guo & He"
format:
  html:
    embed-resources: true
    toc: true
    code-fold: true
bibliography: []
---

<!--
  analysis_gbsg_survival_maxeff_mrconfirm.qmd

  A simplified single-track version of analysis_gbsg_cox_multimethod_*.qmd,
  reduced to the sg_focus = "maxeff" path where Guo & He (2021) Algorithm 3
  applies, and adding Guo & He as a fourth de-biasing correction alongside the
  Naive estimate, the Full Bootstrap (FB), and Multiplier Resampling (MR).

  Relative to the multimethod document, this DROPS: the DINA and GRF selection
  tracks (Guo & He's argmax comparator is not the relevant correction for
  data-driven regenerating families), and the CV/LOO validation blocks.

  Guo & He enters through the adapter fs_to_guohe() (package API, exported by
  forestsearch), which exports the maxeff candidate family as membership
  columns and runs guohe_algorithm3() on the SAME selected subgroup
  forestsearch chose.

  VENUE. The forestsearch fit, the full bootstrap, and the Guo & He pair
  bootstrap are compute-heavy and require the forestsearch package installed.
  Run on the analysis machine (Linux, many cores) -- see the parallelism note in
  the setup chunk. Not a chat-sandbox document.
-->

```{r setup}
#| message: false
#| warning: false
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(doFuture)
library(kableExtra)

timings <- list()

# Guo & He comparator: guohe_algorithm3() / fs_to_guohe() are package API
# (exported by forestsearch). gh_dir is kept only for the LOO cache path below.
gh_dir <- Sys.getenv("GH_DIR", unset = "../../../quarto/GuoHe")

# ---------------------------------------------------------------------------
# Output location for this document's results:  <base>/<dirout>/
#   results_dir = NULL -> base is <qmd dir>/_payloads
#   dirout      = NULL -> the .qmd stem
# Setting results_dir redirects the base only; each run still gets its own
# subdirectory named by dirout.
#
# Examples:
#   results_dir <- NULL;  dirout <- NULL
#     -> <qmd dir>/_payloads/<stem>/<stem>_payload.rds
#   results_dir <- NULL;  dirout <- "gbsg_survival_maxeff_mrconfirm"
#     -> <qmd dir>/_payloads/gbsg_survival_maxeff_mrconfirm/gbsg_survival_maxeff_mrconfirm_payload.rds
#   results_dir <- "~/GitHub/fs-post-selection-manuscript_jasa/_payloads"; dirout <- NULL
#     -> ~/GitHub/fs-post-selection-manuscript_jasa/_payloads/<stem>/<stem>_payload.rds
# ---------------------------------------------------------------------------
results_dir <- NULL
dirout      <- NULL
```

```{r parallel-setup}
# PARALLELISM NOTES (see the doFuture discussion):
#   * FB  -- forestsearch_bootstrap_dofuture() uses %dofuture%; it inherits the
#            plan set here and restores the previous plan on exit. Any plan works.
#   * MR  -- the de-biased gate is vectorized matrix algebra (a single crossprod
#            over all draws), NOT a parallel loop; it needs no plan, only a
#            threaded BLAS to benefit. Do not wrap it in a future.
#   * GH  -- guohe_algorithm3()'s pair bootstrap is parallelised via its own
#            `parallel` argument. guohe_algorithm3() is package API, so any
#            plan works (workers see the installed forestsearch); forked
#            workers (plan(multicore), Linux/Mac) remain the cheapest, hence
#            the unix-only default below.
n_cores <- max(1L, floor(0.80 * parallel::detectCores(logical = FALSE)))
cat("Using", n_cores, "physical cores for parallel sections\n")

NB         <- 1000L     # full-bootstrap resamples (production: 500-2000)
mr_draws <- 5000L     # MR multiplier draws
GH_B       <- 2000L     # Guo & He pair-bootstrap resamples
GH_r       <- 0.03      # Guo & He tuning parameter (their adaptive default region)
GH_PARALLEL <- .Platform$OS.type == "unix"   # forked GH bootstrap where available
# --- MR toggles -----------------------------------------------------------
# run_mr: the in-fit MR de-biased gate, computed inside forestsearch(). TRUE
#   passes mr_inference = TRUE to the primary fit; FALSE skips it entirely
#   (the fit is then called with mr_inference = FALSE, so no MR compute runs).
run_mr     <- TRUE
timings$mr <- NA_real_
# run_mr_confirm: a SEPARATE confirmation fit -- a second forestsearch() with
#   mr_inference = TRUE forced on, run REGARDLESS of run_mr, replaying the
#   primary fit's captured arguments at the same seed. It reproduces the in-fit
#   MR exactly and thereby confirms the internal analysis. Costs a full second
#   fit; set FALSE to skip for a fast pass.
run_mr_confirm     <- TRUE
timings$mr_confirm <- NA_real_
# --------------------------------------------------------------------------
run_loo    <- TRUE      # leave-one-out (N-fold) honest out-of-sample subgroup estimate
timings$loo <- NA_real_
# LOO fold-loop workers: 80% of physical cores (same rule as n_cores above), so
# this adapts to whichever machine renders. No nested parallelism to guard
# against: forestsearch_Kfold() pins each fold's inner forestsearch() refit to
# plan = "sequential", workers = 1, so these are the ONLY workers in play.
loo_workers <- n_cores
# LOO cache: TRUE reuses a saved cv_out (LOO_CACHE env var, else gh_dir/
#   cv_out_maxeff.rds) on re-render, skipping the N-fold recompute. Set FALSE to
#   force a fresh run -- e.g. a reproducibility check. The N-fold search is
#   deterministic (forestsearch_Kfold() sets seed 8316951 with reproducible
#   parallel streams), so a forced recompute reproduces the cached result; a
#   fresh run then overwrites the cache.
loo_use_cache <- FALSE
run_cv_kfold <- TRUE    # repeated 10-fold cross-validation (stability diagnostics)
Ksims       <- 50L      # repeated 10-fold partitions (>= 50 for production)
timings$kfold <- NA_real_
```

```{r data}
df.analysis <- gbsg
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
})

confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name   <- "status"
id.name      <- "id"
treat.name   <- "hormon"

cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
```

## Maxeff identification {#sec-fit}

The forest search is run under `sg_focus = "maxeff"`: the selected subgroup is
the effect maximiser over the enumerated candidate family -- the argmax
primitive of Guo and He (2021), and the selection Guo & He's de-biasing is built
to correct. Under `maxeff` the candidate pool is not truncated, so the family
the correction ranges over is the full enumerated family. The receptor cuts
$\{er \le 0\}$ and $\{pgr \le 0\}$ are pre-specified as forced candidates, and
$J = 10$ cuts are used for the continuous `er` and `pgr` covariates.

```{r forestsearch-maxeff}
#| fig-width: 10
#| fig-height: 7
t0 <- proc.time()

fs <- forestsearch(
  df.analysis,
  outcome.name          = outcome.name,
  event.name            = event.name,
  treat.name            = treat.name,
  id.name               = id.name,
  confounders.name      = confounders.name,
  is.RCT                = TRUE,
  seedit                = 8316951,
  est.scale             = "hr",
  # candidate generation: consistency search only (no LASSO/GRF/DINA screens)
  use_lasso             = FALSE,
  use_grf               = FALSE,
  use_dina              = FALSE,
  # continuous-cut construction + pre-specified receptor candidates
  conf.cont_jcuts       = list(er = 10, pgr = 10),
  collapse_cuts         = TRUE,
  conf_force            = c("er <= 0", "pgr <= 0"),
  # selection: the effect-maximiser argmax (Guo & He primitive)
  subgroup_method       = "consistency",
  sg_focus              = "maxeff",
  # subgroup constraints
  n.min                 = 60, d0.min = 10, d1.min = 10, maxk = 2,
  # Tier-2 de-biased gate (MR) computed inside the fit; run_mr toggles it. The
  # independent confirmation fit (see the MR-confirmation section) forces
  # mr_inference = TRUE regardless, by replaying fs$args_call_all.
  mr_inference           = run_mr,
  mr_inference_args      = list(draws = mr_draws),
  # consistency search
  consistency_method    = "resample",
  fs.splits             = 1000, stop_threshold = 0.95,
  use_twostage          = TRUE, max.minutes = 3,
  show_candidate_summary = TRUE, details = TRUE, quiet = FALSE,
  plot.sg               = TRUE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE)
)
plan("sequential")
timings$fit <- (proc.time() - t0)[["elapsed"]]

stopifnot(!is.null(fs$sg.harm))
cat("\nSelected subgroup:", paste(fs$sg.harm, collapse = " & "),
    "  |  fit", round(timings$fit / 60, 1), "min\n")
```

## Full bootstrap (FB)

```{r bootstrap}
t0 <- proc.time()
fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est   = fs,
  nb_boots = NB,
  seed     = 8316951,   # explicit default: FB is reproducible render-to-render
  show_three = FALSE,
  details  = FALSE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE)
)
plan("sequential")
timings$fb <- (proc.time() - t0)[["elapsed"]]

fb_sum <- summarize_bootstrap_results(
  sgharm = fs$sg.harm, boot_results = fs_bc,
  create_plots = FALSE, est.scale = "hr"
)
cat("FB completed in", round(timings$fb / 60, 1), "min\n")
```

## Multiplier resampling (MR)

The Tier-2 de-biased gate is the MR correction, computed inside the fit above
(vectorized over `r format(mr_draws, big.mark=",")` draws) when `run_mr = TRUE`.
It is read from `fs$mr_inference`; no separate run is needed. Setting
`run_mr <- FALSE` skips it -- the fit is then called with `mr_inference = FALSE`,
so no MR compute happens inside the fit -- mirroring the `run_loo` /
`run_cv_kfold` toggles. The separate confirmation fit in the next section still
supplies an MR estimate even when `run_mr = FALSE`.

```{r mr-skip-note, eval = !run_mr, results = "asis"}
cat("> **In-fit MR gate skipped** (`run_mr = FALSE`). The MR estimate reported",
    "below and in the comparison table comes from the separate confirmation fit.\n")
```

```{r mr, eval = run_mr}
g <- fs$mr_inference
stopifnot(!is.null(g))
timings$mr <- g$timing_seconds
cat("MR gate:", round(timings$mr, 2), "s over",
    format(mr_draws, big.mark = ","), "draws\n")
```

## MR confirmation (independent second fit) {#sec-mr-confirm}

A **separate multiplier-resampling run that executes regardless of `run_mr`**, as
a check on the in-fit gate. There is no package entry point that re-runs MR on an
existing fit -- the gate is built from the fit's internal cut matrix and working
data, which the returned object does not expose -- so the confirmation is a
second `forestsearch()` fit with `mr_inference = TRUE` forced on. It replays the
primary fit's own captured arguments (`fs$args_call_all`, the list the package
itself uses to reconstruct the call for its bootstrap and cross-validation),
overriding only `mr_inference` (plus display-only flags), so the two fits are
identical in every computational respect. Because the search seed (`seedit`) and
the MR draw seed are unchanged, the confirmation reproduces the same selected
subgroup, the same candidate family, and the same multiplier draws -- so when the
in-fit gate also ran (`run_mr = TRUE`), the two MR results must match to full
precision. A mismatch would indicate the in-fit result was altered downstream, or
that the pipeline is not deterministic under the stated seed.

This runs a full second forest search, so it roughly doubles the fit cost; set
`run_mr_confirm <- FALSE` in the configuration to skip it for a fast pass.

```{r mr-confirm-skip-note, eval = !run_mr_confirm, results = "asis"}
cat("> **MR confirmation fit skipped** (`run_mr_confirm = FALSE`). Set it to",
    "`TRUE` to run the independent second fit that reproduces the in-fit MR.\n")
```

```{r mr-confirm, eval = run_mr_confirm}
t0 <- proc.time()
# Replay the primary fit via its own captured argument list, forcing the MR gate
# on. do.call(forestsearch, args_call_all) is the package's documented
# reconstruction path (see forestsearch_main.R Section 1B): a faithful re-run
# differing ONLY in the flags set below.
confirm_args <- fs$args_call_all
confirm_args$mr_inference           <- TRUE   # the only computational change
confirm_args$quiet                  <- TRUE   # display-only suppressions (these do
confirm_args$details                <- FALSE  #   not affect the search or MR RNG):
confirm_args$show_candidate_summary <- FALSE  #   keep the confirmation quiet and
confirm_args$plot.sg                <- FALSE  #   avoid a duplicate subgroup plot.
fs_confirm <- do.call(forestsearch, confirm_args)
plan("sequential")
timings$mr_confirm <- (proc.time() - t0)[["elapsed"]]

g_confirm <- fs_confirm$mr_inference
stopifnot(!is.null(g_confirm))
cat(sprintf("MR confirmation fit: debiased HR %.4f (%.4f, %.4f) | %.1f min\n",
            g_confirm$debiased$est, g_confirm$debiased$lower,
            g_confirm$debiased$upper, timings$mr_confirm / 60))

# When the in-fit gate also ran, confirm the two agree to full precision.
if (isTRUE(run_mr) && exists("g") && !is.null(g)) {
  agree <- all(vapply(c("est", "lower", "upper"), function(f)
    isTRUE(all.equal(g$debiased[[f]], g_confirm$debiased[[f]])), logical(1)))
  cat(sprintf("  in-fit MR: debiased HR %.4f (%.4f, %.4f)  ->  %s\n",
              g$debiased$est, g$debiased$lower, g$debiased$upper,
              if (agree) "MATCH (in-fit MR reproduced exactly)"
              else "MISMATCH -- investigate"))
  if (!agree)
    warning("MR confirmation MISMATCH: the independent fit did not reproduce the ",
            "in-fit MR de-biased HR; investigate determinism or downstream ",
            "mutation of fs$mr_inference.")
} else {
  cat("  in-fit MR gate was skipped (run_mr = FALSE); this confirmation fit is",
      "the sole MR estimate.\n")
}
```

## Guo & He (GH)

The adapter exports the maxeff candidate family as membership columns and runs
Guo & He Algorithm 3 on the same selected subgroup. The membership integrity
check asserts that the exported selected-subgroup column matches forestsearch's
own `sg.harm.id` exactly before the estimator runs.

```{r guohe}
t0 <- proc.time()
gh <- fs_to_guohe(
  fit       = fs,
  time      = "time_months", event = "status", treatment = "hormon",
  B         = GH_B, r = GH_r, level = 0.05, seed = 8316951,
  parallel  = GH_PARALLEL, verify = TRUE
)
timings$gh <- (proc.time() - t0)[["elapsed"]]

cat("GH family size:", length(gh$export$candidates),
    " | selected:", gh$selected,
    " | GH", round(timings$gh / 60, 1), "min\n")
cat(sprintf("GH  naive HR %.3f  debiased HR %.3f  lower bound %.3f\n",
            gh$naive_hr, gh$debiased_hr, gh$bound_hr))
```

## Repeated 10-fold cross-validation (stability diagnostics)

Repeated 10-fold cross-validation partitions the data into ten folds, refits the
selection on each nine-fold training set, and records which subgroup is
identified -- repeated over `Ksims` random partitions. It is a **stability
diagnostic on the subgroup _definition_**: across `Ksims` × 10 training splits,
does the maxeff argmax converge on a single dominant subgroup, or does the
selection wander with the particular split? This is `forestsearch_tenfold()`;
because the search is data-adjacent, the recovered-subgroup distribution is the
quantity of interest.

This is a pure maxeff, fixed-family analysis (no GRF, DINA, or LASSO screens), so
the GRF-specific fold diagnostics are omitted; the universal ones -- identified-
subgroup distribution, selected-candidate consistency, per-fold numerics, and
agreement with the full-data selection -- are reported.

```{r cv-kfold-skip-note, eval = !run_cv_kfold, results = "asis"}
cat("> **10-fold CV skipped.** Set `run_cv_kfold <- TRUE` in the",
    "configuration to enable.\n")
```

```{r kfold-cv, eval = run_cv_kfold}
t0 <- proc.time()
fs_kfold <- forestsearch_tenfold(
  fs.est  = fs,
  sims    = Ksims,
  Kfolds  = 10,
  details = FALSE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = FALSE)
)
plan("sequential")
timings$kfold <- (proc.time() - t0)[["elapsed"]]
cat("10-fold CV:", Ksims, "x 10 partitions in",
    round(timings$kfold / 60, 1), "min\n")
metrics_tables <- cv_metrics_tables(fs_kfold)
metrics_tables
```

### Diagnostic tabulations

`summarize_cv_results()` consolidates the per-fold record into diagnostic tables.
For this maxeff fixed-family analysis the meaningful (non-GRF) diagnostics are:
the distribution of identified subgroups across resamples, the consistency of the
selected candidates, the per-fold numeric summary, and agreement with the
full-data selection.

```{r cv-diagnostics-run, eval = run_cv_kfold}
cv_diag <- summarize_cv_results(
  cv_output    = fs_kfold,
  original_sg  = fs$sg.harm,
  top_n        = 10L,
  create_plots = TRUE
)
```

**Which subgroups emerged across resamples?** The identification summary shows
the distribution of selected subgroups across all `r Ksims` × 10 sim × fold
pairs. A single dominant subgroup with high share indicates the maxeff argmax
converges on a stable selection; a long tail indicates the selection is sensitive
to the training split.

```{r cv-identification, eval = run_cv_kfold}
cv_diag$identification_summary
```

**How confidently were the successful identifications made?** Among folds that
identified a subgroup, the `pconsistency_distribution` shows where the selected
candidate's consistency probability (`Pcons`) landed. (Under maxeff no
consistency threshold is applied at selection, so this is descriptive of the
selected argmax, not a gate.)

```{r cv-pcons, eval = run_cv_kfold}
cv_diag$pconsistency_distribution
```

**How do the numeric fold-level diagnostics summarize?** The
`fold_numeric_summary` tabulates median / IQR / range for each numeric fold-level
column: fold size (`n_test`), training-fold HR of the selected subgroup
(`training_fs_hr`, in-sample and optimistically biased -- for diagnostic
comparison only), and candidate count (`n_candidates_evaluated`).

```{r cv-fold-numeric, eval = run_cv_kfold}
cv_diag$fold_numeric_summary
```

**Agreement with the full-data analysis.** How often does a training fold
reproduce the original full-data subgroup exactly, or share a factor with it?

```{r cv-original-agreement, eval = run_cv_kfold}
cv_diag$original_agreement
```

**Identified-subgroup distribution (plot).** The bar chart of the top identified
subgroups across the `r Ksims` × 10 partitions visualizes the stability of the
maxeff selection: a dominant bar is a stable definition, a flat spread is an
unstable one.

```{r cv-identification-plot, eval = run_cv_kfold}
#| fig-width: 9
#| fig-height: 5
#| fig-cap: "Distribution of identified subgroups across repeated 10-fold training splits."
if (!is.null(cv_diag$plots) && !is.null(cv_diag$plots$identification)) {
  cv_diag$plots$identification
}
```


## Leave-one-out (honest out-of-sample estimate)

Leave-one-out re-runs the entire selection on each hold-out of one subject and
predicts the held-out subject's subgroup membership from its covariates. It is an
out-of-sample **stability check on the subgroup _definition_**, not a de-biasing
correction of the _effect_: it asks whether the same subgroup is recovered when
no subject informs its own label. This is `forestsearch_Kfold()` at N-fold
(`Kfolds = nrow(df)`), applied to the selected maxeff fit.

Unlike the coarse frozen-family companion analysis -- where the median-cut family
is small, so the argmax has few candidates to choose among and the recovered
definition is mechanically stable -- the maxeff family here is the full,
finer-grained candidate set (quantile cuts on the continuous covariates). The
argmax therefore has genuine room to land on different definitions across folds,
so this is the **discriminating** leave-one-out comparison: whether the
out-of-sample estimate tracks Naive (definition perfectly stable) or one of the
de-biasing corrections (definition varies fold to fold, so out-of-sample
membership differs from in-sample) is informative rather than predetermined. The
recovered-definition distribution and the out-of-sample hazard ratio are reported
below; read them against the agreement metrics.

Note this is the heaviest step: N-fold means one full forest search per subject.
It is the long pole; on the analysis box, not the sandbox.

The fold loop is distributed across `r loo_workers` workers. `parallel_args`
controls **fold-level** parallelism only -- each fold's inner `forestsearch()`
refit is pinned to `plan = "sequential"`, so there is no nested oversubscription
-- and the fold loop draws parallel-safe L'Ecuyer streams
(`.options.future = list(seed = TRUE)`). In the frozen-family companion analysis
the parallel fold loop was verified against a sequential run of the same
analysis: **identical** per-subject predictions and an identical out-of-sample
HR, at 5.2x the speed on 11 workers.

```{r loo-skipped-note, eval = !run_loo}
cat("Leave-one-out skipped (run_loo = FALSE).",
    "Set run_loo <- TRUE in parallel-setup to run it.\n")
```

```{r loo, eval = run_loo}
# Leave-one-out (N-fold) is the long pole -- one full forest search per subject.
# If loo_use_cache is TRUE and a precomputed cv_out exists (LOO_CACHE env var,
# else the default path in gh_dir), load it; otherwise run the N-fold search and
# cache it. The fold loop runs in parallel on loo_workers workers; each fold's
# inner search is pinned sequential by forestsearch_Kfold().
loo_cache <- Sys.getenv("LOO_CACHE",
                        unset = file.path(gh_dir, "cv_out_maxeff.rds"))
if (loo_use_cache && file.exists(loo_cache)) {
  cv_out <- readRDS(loo_cache)
  timings$loo <- NA_real_
  cat("LOO loaded from cache:", loo_cache, "\n")
} else {
  t0 <- proc.time()
  fs_OOB <- forestsearch_Kfold(
    fs.est  = fs,
    details = FALSE,
    Kfolds  = nrow(df.analysis),   # N-fold = leave-one-out
    parallel_args = list(plan = "multisession", workers = loo_workers,
                         show_message = TRUE)
  )
  plan("sequential")
  timings$loo <- (proc.time() - t0)[["elapsed"]]
  cv_out <- forestsearch_KfoldOut(res = fs_OOB, details = FALSE, outall = TRUE)
  saveRDS(cv_out, loo_cache)
}

# Out-of-sample harm-subgroup HR, read BY IDENTITY (the harm row of the K-fold
# table, matched on its subgroup label sg0.name = "Not recommend"), never
# positionally -- the table layout depends on n_sgfound. Two structural cases:
#   (i)  stable subgroup: SG_tab_Kfold is a matrix whose harm row carries the
#        sg0.name label -> its out-of-sample HR (identical to tab_all's KfA_0);
#   (ii) ITT fallback: no stable subgroup survives out-of-sample (n_sgfound == 1)
#        and SG_tab_Kfold collapses to the 1-D ITT vector -> no harm row ->
#        report the fallback explicitly. In the finer maxeff family this branch
#        may genuinely trigger if the argmax definition varies across folds.
.loo_harm_row <- function(cv_out, harm_label = "Not recommend") {
  sgk <- cv_out$SG_tab_Kfold
  if (is.matrix(sgk) && "Subgroup" %in% colnames(sgk)) {
    j <- which(sgk[, "Subgroup"] == harm_label)
    if (length(j) == 1L)
      return(list(case = "stable",
                  hr   = as.character(sgk[j, "HR (95% CI)"]),
                  n    = as.character(sgk[j, "n"])))
  }
  list(case = "itt_fallback",
       hr   = "— (no stable subgroup out-of-sample; ITT)",
       n    = NA_character_)
}
loo <- .loo_harm_row(cv_out)

# Definition stability: fold recovery and agreement with the full-data subgroup.
metr <- cv_out$sens_metrics_original
cat(sprintf(
  "LOO [%s]: out-of-sample harm HR %s (n = %s)\n  agreement: sens_H = %.2f, ppv_H = %.2f (exact-match = %.2f)\n",
  loo$case, loo$hr, loo$n, metr[["sens_H"]], metr[["ppv_H"]],
  cv_out$find_metrics[["Exact"]]))

# Full CV tables (ITT / full-analysis / K-fold, plus agreement metrics).
tables <- cv_summary_tables(cv_out)
tables$combined_table
tables$metrics_table
```

## Comparison: four corrections on the selected subgroup {#sec-table}

```{r table}
#| tbl-cap: "Subgroup hazard ratios for the GBSG selected (maxeff) subgroup. Parenthetical values are 95% intervals; Guo & He reports a one-sided lower bound. Leave-one-out is NOT a de-biasing correction: it is an out-of-sample stability check on the subgroup definition. Whether it tracks Naive or a correction is informative in this finer maxeff family (see the leave-one-out section)."
#| echo: false
fci <- function(e, l, u, d = 2) {
  ifelse(is.na(e), "—", sprintf("%.*f (%.*f, %.*f)", d, e, d, l, d, u))
}

# Selected-subgroup (H) HR + 95% CI on the HR scale, read straight from the
# actual result objects (verified against str(fs_bc$H_estimates) and
# str(fs$mr_inference); the estimates are unchanged, this is display plumbing):
#   FB bias-corrected: the bootstrap object's H_estimates carries H0 (naive HR)
#     and H2 (the "HR*" infinitesimal-jackknife bias-corrected HR) with
#     H2_lower/H2_upper -- summarize_bootstrap_results() formats these into a gt
#     table and exposes no top-level est_H/lo_H/hi_H fields.
#   MR de-biased gate: fs$mr_inference$debiased$est / lower / upper (there is no
#     g$Hc_estimates; the gate's H-subgroup de-biased HR lives under $debiased).
naive_hr <- gh$naive_hr    # naive is identical across methods (same fit)

# MR row: the in-fit gate (g) when run_mr = TRUE, else the always-run confirmation
# fit (g_confirm). Both live on the HR scale under $debiased; em-dash only if BOTH
# the in-fit gate and the confirmation fit were skipped.
g_report <- if (isTRUE(run_mr) && exists("g") && !is.null(g)) g else
            if (exists("g_confirm") && !is.null(g_confirm)) g_confirm else NULL
mr_row <- if (!is.null(g_report))
  fci(g_report$debiased$est, g_report$debiased$lower, g_report$debiased$upper) else
  "— (no MR: run_mr and run_mr_confirm both FALSE)"

# LOO row read by identity from the harm-subgroup extraction above (both the
# stable-subgroup and ITT-fallback cases resolve to loo$hr).
loo_row <- if (isTRUE(run_loo) && exists("loo")) loo$hr else "— (run_loo = FALSE)"

tab <- data.frame(
  Method = c("Naive", "Full bootstrap", "Multiplier resampling",
             "Guo & He (2021)", "Leave-one-out (out-of-sample)"),
  Estimate = c(
    sprintf("%.2f", naive_hr),
    fci(fs_bc$H_estimates$H2, fs_bc$H_estimates$H2_lower, fs_bc$H_estimates$H2_upper),
    mr_row,
    sprintf("%.2f (LB %.2f)", gh$debiased_hr, gh$bound_hr),
    loo_row
  ),
  check.names = FALSE
)
kbl(tab, booktabs = TRUE, escape = FALSE, align = "ll") |>
  kable_styling()
```

::: {.callout-note}
## Reading the table

Naive is the uncorrected subgroup hazard ratio. Full bootstrap and Multiplier
resampling both target the conditional estimand and report two-sided intervals;
they differ in that FB re-runs the entire selection on each resample (so its
cost scales with the resample count) while MR holds the fitted family fixed and
is vectorized (so its cost does not). Guo & He corrects the winner's-curse
optimism of the argmax selection and reports a de-biased point estimate with a
one-sided lower confidence bound -- the object their Theorems 4/5 and
simulations certify.

Guo & He assumes a fixed candidate family; the maxeff family here is large
(hundreds to ~1744 candidates depending on cut construction) and data-adjacent,
which is outside their stated small-fixed-$k$ regime. The Section 5 reproduction
established that their estimator is faithfully implemented on disjoint families
up to $k = 12$; its behaviour on a family this size is the open question, and is
exactly why the head-to-head with MR is of interest.

Leave-one-out is **not a de-biasing correction** and should not be read as one.
It is an out-of-sample **stability check on the subgroup _definition_**: it asks
whether the same subgroup is recovered when no subject informs its own label. Its
value here depends on the fold-recovery behaviour, which the agreement metrics
report. If every fold recovers the *identical* definition, out-of-sample
membership equals in-sample membership and the leave-one-out HR is mechanically
the naive one -- so `LOO = Naive` would indicate a perfectly stable definition,
not a failure to correct. If the argmax instead lands on *different* definitions
across folds (which the finer maxeff family makes possible, unlike the coarse
frozen-family companion), out-of-sample membership departs from in-sample and the
leave-one-out HR moves away from Naive -- and *where* it moves is the informative
signal. Read the leave-one-out row together with the agreement metrics
(`sens_H`, `ppv_H`, exact-match): they say whether the definition is stable, and
the HR says what the honest out-of-sample effect is for that (possibly varying)
membership. Leave-one-out validates the *definition*; the de-biasing corrections
address the *effect* optimism -- distinct questions, so agreement or disagreement
between the leave-one-out row and the corrections is not a contradiction.
:::

```{r timing}
#| echo: false
td <- data.frame(
  Step = c("Fit (+ in-fit MR gate)", "Full bootstrap", "MR gate (within fit)",
           "MR confirmation fit (independent)",
           "Guo & He bootstrap", "10-fold CV",
           sprintf("Leave-one-out (N-fold, %d workers)", loo_workers)),
  Seconds = c(timings$fit, timings$fb, timings$mr, timings$mr_confirm, timings$gh,
              timings$kfold, timings$loo)
)
td$Minutes <- round(td$Seconds / 60, 2)
kbl(td, booktabs = TRUE, align = "lrr") |> kable_styling()
```

# Reproducibility payload {#sec-export-payload}

The chunk below harvests the objects this render already computed into a single
small list and writes it as an `.rds`, so the results survive the HTML. No
analysis is repeated. Every element is guarded, so a section this render skipped
(`run_mr`, `run_mr_confirm`, `run_cv_kfold`, `run_loo`) stores `NULL` rather
than erroring, and the top-level names match the other GBSG payloads
(`table`, `labels`, `meta`, `extras`, `est_scale`, `built_at`,
`forestsearch_version`) so a reader can treat them uniformly.

```{r export-payload}
#| echo: false
#| message: false

.qmd_dir  <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) getwd())
.qmd_stem <- tryCatch(tools::file_path_sans_ext(basename(knitr::current_input())),
                      error = function(e) "analysis")
.dirout   <- if (is.null(dirout)) .qmd_stem else dirout
.base     <- if (is.null(results_dir)) file.path(.qmd_dir, "_payloads") else results_dir
.out_dir  <- file.path(.base, .dirout)
dir.create(.out_dir, recursive = TRUE, showWarnings = FALSE)
.payload_file <- file.path(.out_dir, paste0(.dirout, "_payload.rds"))

.get <- function(nm) if (exists(nm, inherits = TRUE)) get(nm, inherits = TRUE) else NULL
.try <- function(expr) tryCatch(expr, error = function(e) NULL)

.fs   <- .get("fs")
.fsbc <- .get("fs_bc")
.g    <- if (is.null(.fs)) NULL else .fs$mr_inference
.fsc  <- .get("fs_confirm")
.gc   <- .get("g_confirm")
.gh   <- .get("gh")
.cvd  <- .get("cv_diag")
.cvo  <- .get("cv_out")
.loo  <- .get("loo")

payload <- list(
  table  = .get("tab"),
  labels = list(
    sg_harm  = if (is.null(.fs)) NULL else .fs$sg.harm,
    sg_focus = if (is.null(.fs)) NULL else .fs$sg_focus,
    focus    = .get("focus")),   # no focus profile in this document: NULL
  meta = list(
    NB          = .get("NB"),
    mr_draws    = .get("mr_draws"),
    Ksims       = .get("Ksims"),
    n_cores     = .get("n_cores"),
    loo_workers = .get("loo_workers"),
    n_total     = .try(nrow(.get("df.analysis"))),
    n_events    = .try(sum(.get("df.analysis")[[.get("event.name")]])),
    outcome     = .get("outcome.name"),
    event       = .get("event.name"),
    treat       = .get("treat.name"),
    covariates  = .get("confounders.name"),
    timings     = .get("timings"),
    # This document carries no fs_param_provenance() helper; the fit's own
    # captured argument list is the provenance record it does produce.
    provenance  = .try(fs_param_provenance(.fs)),
    args_call_all = if (is.null(.fs)) NULL else .fs$args_call_all),
  extras = list(
    mr = if (is.null(.g)) NULL else list(
      naive          = .g$naive,
      debiased       = .g$debiased,
      settings       = .g$settings,
      n_family       = .g$n_family,
      selection_bias = .g$selection_bias,
      timing_seconds = .g$timing_seconds),
    # The independent confirmation fit -- this document's distinguishing result.
    mr_confirm = if (is.null(.gc)) NULL else list(
      naive          = .gc$naive,
      debiased       = .gc$debiased,
      settings       = .gc$settings,
      n_family       = .gc$n_family,
      selection_bias = .gc$selection_bias,
      timing_seconds = .gc$timing_seconds,
      sg_harm        = if (is.null(.fsc)) NULL else .fsc$sg.harm,
      agrees_with_in_fit = .try(
        all(vapply(c("est", "lower", "upper"), function(f)
          isTRUE(all.equal(.g$debiased[[f]], .gc$debiased[[f]])), logical(1))))),
    fb = if (is.null(.fsbc)) NULL else list(
      H_estimates  = .fsbc$H_estimates,
      Hc_estimates = .fsbc$Hc_estimates),
    pareto = NULL,   # no banded focus here: the frontier block does not run
    cv = if (isTRUE(.get("run_cv_kfold")) && !is.null(.cvd)) list(
      metrics                   = .get("metrics_tables"),
      identification_summary    = .cvd$identification_summary,
      pconsistency_distribution = .cvd$pconsistency_distribution,
      fold_numeric_summary      = .cvd$fold_numeric_summary,
      original_agreement        = .cvd$original_agreement) else NULL,
    loo = if (isTRUE(.get("run_loo")) && !is.null(.loo)) list(
      row          = .loo,
      sens_metrics = if (is.null(.cvo)) NULL else .cvo$sens_metrics_original,
      find_metrics = if (is.null(.cvo)) NULL else .cvo$find_metrics,
      cache        = .get("loo_cache")) else NULL,
    gh = if (is.null(.gh)) NULL else list(
      naive_hr    = .gh$naive_hr,
      debiased_hr = .gh$debiased_hr,
      bound_hr    = .gh$bound_hr,
      selected    = .gh$selected,
      n_family    = .try(length(.gh$export$candidates)),
      B           = .get("GH_B"),
      r           = .get("GH_r"))),
  est_scale = "hr",
  built_at  = Sys.time(),
  forestsearch_version = tryCatch(as.character(utils::packageVersion("forestsearch")),
                                  error = function(e) NA_character_))

saveRDS(payload, .payload_file)
cat("Wrote payload to:", .payload_file, "\n")
```
