---
title: "GBSG subgroup analysis across selection foci: Naive, Full Bootstrap, Multiplier Resampling, and Guo & He"
subtitle: "Survival (Cox / HR) outcome; `sg_focus` is a single knob"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
bibliography: []
---

<!--
  analysis_gbsg_survival_effMaxSG.qmd

  Supersedes analysis_gbsg_survival_maxeff.qmd.  At fs_sg_focus <- "maxeff"
  this document reproduces that analysis; the difference is that fs_sg_focus
  now accepts ANY legal focus and the document adapts around it.

  WHAT ADAPTS TO THE FOCUS
    * Pareto frontier block (figure + CIs + table + narrative) runs only for
      the banded foci effMaxSG / effMinSG (canonical hrMaxSG / hrMinSG),
      where the (effect, N) frontier IS the selection geometry rather than a
      post-hoc curiosity.
    * Guo & He runs only under sg_focus = "maxeff".  This is a hard
      constraint, not a stylistic choice: fs_export_maxeff_family()
      (fs_to_guohe.R:141-143) stops unless fit$sg_focus == "maxeff", and it
      needs the untruncated family that only maxeff guarantees.
    * selection_rule is passed through only for the banded foci.  Elsewhere
      .validate_selection_rule() RAISES AN ERROR for a non-default value
      rather than ignoring it, so the document coerces it to "neighborhood".
    * stop_threshold is supplied only for maxeffCons, the one focus whose
      sort key contains no Pcons term.  Everywhere else forestsearch()
      resets it to NULL; passing NULL up front keeps the render warning-free
      and the provenance capture honest.
    * Under maxeff the five engine-forced overrides are pre-resolved in the
      call itself, so args_call_all records what the engine ran with rather
      than what was typed.  The provenance table VERIFIES this rather than
      assuming it.

  VENUE.  The forestsearch fit, the full bootstrap, the Guo & He pair
  bootstrap and the leave-one-out loop are compute-heavy and require the
  forestsearch package installed (not load_all(): doFuture workers spawn
  separate R processes).  Run on the analysis machine.
-->

```{r setup}
#| message: false
#| warning: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(doFuture)
library(kableExtra)

timings <- list()

# Guo & He comparator: guohe_algorithm3() / fs_to_guohe() are package API
# (exported by forestsearch).  gh_dir is kept only for the LOO cache path.
gh_dir <- Sys.getenv("GH_DIR", unset = "../../../quarto/GuoHe")

# ---------------------------------------------------------------------------
# Output location for this document's results:  <base>/<dirout>/
#   results_dir = NULL -> base is <qmd dir>/_payloads
#   dirout      = NULL -> the .qmd stem
# Setting results_dir redirects the base only; each run still gets its own
# subdirectory named by dirout.
#
# Examples:
#   results_dir <- NULL;  dirout <- NULL
#     -> <qmd dir>/_payloads/<stem>/<stem>_payload.rds
#   results_dir <- NULL;  dirout <- "gbsg_survival_effMaxSG"
#     -> <qmd dir>/_payloads/gbsg_survival_effMaxSG/gbsg_survival_effMaxSG_payload.rds
#   results_dir <- "~/GitHub/fs-post-selection-manuscript_jasa/_payloads"; dirout <- NULL
#     -> ~/GitHub/fs-post-selection-manuscript_jasa/_payloads/<stem>/<stem>_payload.rds
# ---------------------------------------------------------------------------
results_dir <- NULL
dirout      <- NULL
```

## Selection focus {#sec-focus}

`sg_focus` is the single knob that decides which surviving candidate is
returned. Two runs that differ only in `sg_focus` estimate different things, so
the resolved focus is recorded in the table below and every downstream section
keys off it.

```{r focus-config}
#| code-fold: false

# ---------------------------------------------------------------------------
# THE ONLY KNOB THAT CHANGES THE ESTIMAND.  Any legal value is accepted:
#
#   "maxcons" / "eff" / "hr"   consistency argmax, effect breaks ties
#   "maxeff"                   effect argmax over ALL candidates, ungated
#   "maxeffCons"               effect argmax among consistency qualifiers
#   "maxSG" / "minSG"          size extremes among qualifiers, no band
#   "effMaxSG" / "hrMaxSG"     largest N within the effect band
#   "effMinSG" / "hrMinSG"     smallest N within the effect band
#
# Prefer the eff*/maxcons spellings: they name the rule and survive the move
# to GLM outcomes.  Note maxSG/minSG are NOT spellings of effMaxSG/effMinSG.
# ---------------------------------------------------------------------------
fs_sg_focus            <- "effMaxSG"

# selection_rule and effect_neighborhood shape the inclusion band.  They are
# consulted ONLY for effMaxSG / effMinSG; the profile below coerces the rule
# to "neighborhood" elsewhere, because forestsearch() treats a non-default
# rule under a non-band focus as an error rather than a no-op.
fs_selection_rule      <- "neighborhood"        # "neighborhood" | "pareto" | "both"
fs_effect_neighborhood <- 0.10

# Thresholds and search settings (inert under maxeff -- see the profile).
fs_hr_threshold         <- 1.0
fs_hr_consistency       <- 1.0
fs_pconsistency         <- 0.90
fs_minp                 <- 0.025
fs_use_twostage         <- TRUE
fs_stop_threshold       <- 0.95   # consumed by maxeffCons only
fs_consistency_method   <- "resample"
```

```{r focus-profile}
#| code-fold: true

# Mirror of the package's .normalize_sg_focus() (forestsearch_helpers.R:916).
.fs_focus_canonical <- function(x) {
  switch(x,
         effMaxSG = "hrMaxSG",
         effMinSG = "hrMinSG",
         eff      = "hr",
         maxcons  = "hr",
         x)
}

#' Resolve a supplied sg_focus into the settings this document must honour
#'
#' Returns the canonical focus plus the derived flags every downstream chunk
#' keys on, and pre-resolves the arguments forestsearch() would otherwise
#' override internally.  Pre-resolving matters for provenance: the maxeff
#' override block (forestsearch_main.R:1309) does NOT re-sync args_call_all,
#' so a value typed into the call and then overridden is captured in its
#' pre-override form.  Supplying the engine-forced value directly removes the
#' discrepancy at source.
.fs_focus_profile <- function(sg_focus,
                              selection_rule,
                              effect_neighborhood,
                              hr_threshold,
                              hr_consistency,
                              pconsistency,
                              minp,
                              use_twostage,
                              stop_threshold) {

  valid_user <- c("hr", "eff", "maxcons",
                  "hrMaxSG", "effMaxSG", "hrMinSG", "effMinSG",
                  "maxSG", "minSG", "maxeff", "maxeffCons")
  if (!is.character(sg_focus) || length(sg_focus) != 1L ||
      !sg_focus %in% valid_user) {
    stop("fs_sg_focus must be one of: ",
         paste(sQuote(valid_user), collapse = ", "),
         ".  Got: ", paste(deparse(sg_focus), collapse = " "),
         call. = FALSE)
  }

  canonical     <- .fs_focus_canonical(sg_focus)
  is_banded     <- canonical %in% c("hrMaxSG", "hrMinSG")
  is_maxeff     <- identical(canonical, "maxeff")
  is_maxeffcons <- identical(canonical, "maxeffCons")

  # selection_rule: ERROR (not a silent no-op) outside the banded foci.
  rule <- if (is_banded) selection_rule else "neighborhood"
  if (!is_banded && !identical(selection_rule, "neighborhood")) {
    message("selection_rule = '", selection_rule, "' is inert for sg_focus = '",
            sg_focus, "' and would be rejected by forestsearch(); ",
            "coerced to 'neighborhood'.")
  }

  # stop_threshold is meaningful for maxeffCons alone.  forestsearch() resets
  # it to NULL for hrMaxSG / hrMinSG / hr / maxSG / minSG (warning only when
  # supplied explicitly) and under maxeff.  Passing NULL avoids the warning
  # and records the operative value.
  stop_thr <- if (is_maxeffcons) stop_threshold else NULL

  # maxeff engine-forced settings (forestsearch_main.R:1309).  Supplying them
  # directly suppresses the override warning AND keeps args_call_all truthful.
  if (is_maxeff) {
    pcons_out  <- 0
    minp_out   <- 0
    twostg_out <- FALSE
  } else {
    pcons_out  <- pconsistency
    minp_out   <- minp
    twostg_out <- use_twostage
  }

  key <- switch(canonical,
    hr         = "(-Pcons, -hr, K)",
    maxeff     = "(-hr, K)",
    maxeffCons = "(-hr, K)",
    maxSG      = "(-N, -Pcons, K)",
    minSG      = "(N, -Pcons, K)",
    hrMaxSG    = "(-in_band, -N, -Pcons, -hr, K)",
    hrMinSG    = "(-in_band, N, -Pcons, -hr, K)")

  ranges_over <- switch(canonical,
    maxeff     = "ALL candidates, ungated",
    maxeffCons = "consistency-qualifying candidates only",
    "all qualifiers")

  list(
    supplied              = sg_focus,
    canonical             = canonical,
    key                   = key,
    ranges_over           = ranges_over,
    is_banded             = is_banded,
    is_maxeff             = is_maxeff,
    is_maxeffcons         = is_maxeffcons,
    run_pareto            = is_banded,
    run_guohe             = is_maxeff,
    thresholds_live       = !is_maxeff,
    selection_rule        = rule,
    effect_neighborhood   = effect_neighborhood,
    hr_threshold          = hr_threshold,
    hr_consistency        = hr_consistency,
    pconsistency          = pcons_out,
    minp                  = minp_out,
    use_twostage          = twostg_out,
    stop_threshold        = stop_thr
  )
}

focus <- .fs_focus_profile(
  sg_focus             = fs_sg_focus,
  selection_rule       = fs_selection_rule,
  effect_neighborhood  = fs_effect_neighborhood,
  hr_threshold         = fs_hr_threshold,
  hr_consistency       = fs_hr_consistency,
  pconsistency         = fs_pconsistency,
  minp                 = fs_minp,
  use_twostage         = fs_use_twostage,
  stop_threshold       = fs_stop_threshold
)

# Chunk-level gates (chunk options are evaluated in the global environment).
run_pareto <- focus$run_pareto
```

```{r focus-record}
#| echo: false
.fmt1 <- function(x) if (is.null(x)) "NULL" else
  paste(deparse(x, width.cutoff = 200L), collapse = " ")

kbl(data.frame(
  Item = c("sg_focus (supplied)", "sg_focus (canonical)", "Sort key",
           "Ranges over", "selection_rule", "effect_neighborhood",
           "Effect floor (hr.threshold)", "Consistency screen (pconsistency)",
           "stop_threshold", "use_twostage",
           "Pareto frontier section", "Guo & He section"),
  Value = c(focus$supplied, focus$canonical, focus$key, focus$ranges_over,
            focus$selection_rule, .fmt1(focus$effect_neighborhood),
            if (focus$thresholds_live) .fmt1(focus$hr_threshold)
              else "disabled under maxeff",
            if (focus$thresholds_live) .fmt1(focus$pconsistency)
              else "0 (screen admits everything)",
            .fmt1(focus$stop_threshold),
            .fmt1(focus$use_twostage),
            if (focus$run_pareto) "RUN (banded focus)"
              else "skipped (band not the selection geometry)",
            if (focus$run_guohe) "RUN (argmax primitive)"
              else "skipped (requires sg_focus = 'maxeff')"),
  check.names = FALSE),
  booktabs = TRUE, align = "ll",
  caption = "Resolved selection configuration for this render.") |>
  kable_styling(full_width = FALSE)
```

```{r focus-narrative}
#| echo: false
#| results: asis
cat(switch(focus$canonical,
  hr = paste(
    "The selected subgroup is the **consistency argmax**: the candidate with the",
    "highest split-consistency probability, with effect size breaking ties. This is",
    "the rule of León et al. (2024). Because enumeration proceeds by effect while",
    "the key is led by consistency, the scan order carries no information about the",
    "ranking -- every candidate must be scored, and a candidate far down the",
    "enumeration can carry the highest `Pcons` and win.\n"),
  maxeff = paste(
    "The selected subgroup is the **effect maximiser** over the enumerated candidate",
    "family -- the argmax primitive of Guo and He (2021), and the selection their",
    "de-biasing is built to correct. No threshold can reject the winner: the",
    "consistency screen is set to admit everything, the effect floor is disabled,",
    "early stopping and two-stage narrowing are off, and the candidate pool is not",
    "truncated. Consistency is computed for the winner and reported, but",
    "descriptively -- so `prop_SG_found = 100%` is structural rather than an",
    "empirical finding.\n"),
  maxeffCons = paste(
    "The selected subgroup is the **effect maximiser among consistency qualifiers**.",
    "The sort key is identical to `maxeff`; what differs is the population it ranges",
    "over, since only candidates clearing `pconsistency.threshold` reach the sort.",
    "This is the one focus whose key contains no `Pcons` term, so early stopping is",
    "sound and `stop_threshold` is live for this render.\n"),
  maxSG = paste(
    "The selected subgroup is the **largest qualifying candidate**, with consistency",
    "breaking ties. There is no inclusion band and no effect term until the final",
    "`K` tiebreak, so the rule can run out to a large but weak subgroup.\n"),
  minSG = paste(
    "The selected subgroup is the **smallest qualifying candidate**, with consistency",
    "breaking ties. There is no inclusion band and no effect term until the final",
    "`K` tiebreak.\n"),
  hrMaxSG = paste(
    "The selected subgroup is the **largest candidate inside the effect band**:",
    "in-band candidates first, then size, then consistency, then effect. The band is",
    "what stops a size-seeking rule from running out to the weak-but-large candidate",
    "`maxSG` would take. Because the band is the selection geometry rather than a",
    "post-hoc diagnostic, the (effect, N) Pareto frontier is reported below as a",
    "**selection** diagnostic.\n"),
  hrMinSG = paste(
    "The selected subgroup is the **smallest candidate inside the effect band**:",
    "in-band candidates first, then size ascending, then consistency, then effect.",
    "Because this focus deliberately prefers small subgroups, the selected subgroup",
    "may be N-dominated by a larger neighbour and therefore may not itself appear on",
    "the frontier -- that is expected, not a defect. The frontier is reported below",
    "as a **selection** diagnostic.\n")))
```

```{r parallel-setup}
# PARALLELISM NOTES:
#   * FB  -- forestsearch_bootstrap_dofuture() uses %dofuture%; it inherits the
#            plan set here and restores the previous plan on exit.
#   * MR  -- the de-biased gate is vectorized matrix algebra (a single crossprod
#            over all draws), NOT a parallel loop; it needs no plan, only a
#            threaded BLAS to benefit.  Do not wrap it in a future.
#   * GH  -- guohe_algorithm3()'s pair bootstrap is parallelised via its own
#            `parallel` argument; forked workers (Linux/Mac) are cheapest.
n_cores <- max(1L, floor(0.80 * parallel::detectCores(logical = FALSE)))
cat("Using", n_cores, "physical cores for parallel sections\n")

NB          <- 1000L    # full-bootstrap resamples (production: 500-2000)
mr_draws    <- 5000L    # MR multiplier draws
GH_B        <- 2000L    # Guo & He pair-bootstrap resamples
GH_r        <- 0.03     # Guo & He tuning parameter
GH_PARALLEL <- .Platform$OS.type == "unix"

# Guo & He is available only under the argmax focus; gate the request on both
# the user toggle and the structural precondition.
gh_requested <- TRUE
run_gh       <- isTRUE(gh_requested) && focus$run_guohe

run_loo      <- TRUE    # leave-one-out (N-fold) out-of-sample subgroup estimate
run_cv_kfold <- TRUE    # repeated 10-fold cross-validation
Ksims        <- 50L     # repeated 10-fold partitions (>= 50 for production)

# LOO fold-loop workers.  No nested parallelism to guard against:
# forestsearch_Kfold() pins each fold's inner refit to plan = "sequential".
loo_workers <- n_cores

# Pre-seed so the timing table is well-formed whichever sections run.
timings$fit   <- NA_real_
timings$fb    <- NA_real_
timings$mr    <- NA_real_
timings$gh    <- NA_real_
timings$loo   <- NA_real_
timings$kfold <- NA_real_
timings$frontier <- NA_real_

# ---------------------------------------------------------------------------
# Pareto-frontier diagnostic parameters.  Consulted only when the focus is
# banded.  compute_frontier_cis() is the expensive call (its cost scales with
# frontier size x n_splits); set compute_cis <- FALSE to render the figure,
# table and narrative without interval bars.
# ---------------------------------------------------------------------------
frontier_params <- list(
  compute_cis = TRUE,
  n_splits    = 1000L,
  seed        = 1L,
  xlim        = NULL,    # HR is a natural-scale measure; NULL lets it auto-fit
  xlim_trim   = TRUE,    # zoom to point estimates, ignoring wide split-CI tails
  digits      = 2L
)
```

```{r data}
df.analysis <- gbsg
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
})

confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name   <- "status"
id.name      <- "id"
treat.name   <- "hormon"

cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
```

## Parameter provenance {#sec-param}

Each `forestsearch()` fit carries `args_call_all`: the argument list the call
actually resolved to, kept in sync by the package's `.sync_args_call_all()`. The
table below reconciles it against `formals(forestsearch)` at render time, so the
values reported are the ones the fit used rather than restated comments that can
drift.

```{r param-provenance-helper}
#| code-fold: false

.fs_fmt_val <- function(v) {
  if (is.null(v)) return("NULL")
  if (is.data.frame(v)) return(sprintf("<data.frame %d x %d>", nrow(v), ncol(v)))
  if (is.function(v)) return("<function>")
  if (is.list(v)) {
    if (length(v) == 0L) return("list()")
    nms <- names(v); if (is.null(nms)) nms <- rep("", length(v))
    inner <- vapply(seq_along(v), function(i) {
      val <- tryCatch(paste(deparse(v[[i]], width.cutoff = 200L), collapse = " "),
                      error = function(e) "<?>")
      if (nzchar(nms[i])) paste0(nms[i], " = ", val) else val
    }, character(1))
    return(paste0("list(", paste(inner, collapse = ", "), ")"))
  }
  if (length(v) > 6L) return(sprintf("<%s[%d]>", class(v)[1], length(v)))
  paste(deparse(v, width.cutoff = 200L), collapse = " ")
}

.fs_fmt_default <- function(x) {
  if (identical(x, quote(expr = ))) return("<required; no default>")
  paste(deparse(x, width.cutoff = 200L), collapse = " ")
}

# match.arg-style values resolve to their first element.  Applied to BOTH sides
# of the comparison below: a value captured unresolved (grf_selection,
# grf_select_statistic) still sits at its default and must not read as differing.
.fs_collapse_matcharg <- function(x) {
  ifelse(!is.na(x) & grepl('^c[(]"', x),
         sub('^c[(]"([^"]*)".*$', '"\\1"', x), x)
}

#' Reconcile a fit's resolved arguments against forestsearch()'s formals
fs_param_provenance <- function(fit, fn = forestsearch) {
  stopifnot(!is.null(fit), !is.null(fit$args_call_all))
  args <- fit$args_call_all
  fm   <- formals(fn)
  nm   <- setdiff(names(fm), "...")

  defaults <- vapply(fm[nm], .fs_fmt_default, character(1), USE.NAMES = FALSE)
  used <- vapply(nm, function(p) {
    if (!p %in% names(args)) return(NA_character_)
    .fs_fmt_val(args[[p]])
  }, character(1), USE.NAMES = FALSE)

  eff      <- .fs_collapse_matcharg(defaults)
  eff_used <- .fs_collapse_matcharg(used)

  out <- data.frame(
    parameter  = nm,
    default    = defaults,
    value_used = ifelse(is.na(used), "(not captured)", used),
    status     = ifelse(is.na(used), "not captured",
                 ifelse(eff_used == eff, "at default", "differs from default")),
    stringsAsFactors = FALSE
  )

  # forestsearch_main.R:1309 overrides five settings under sg_focus = "maxeff"
  # and does NOT re-sync args_call_all afterwards (the package notes this
  # itself at ~2763).  This document pre-resolves those five to the
  # engine-forced values in the call, so the capture should already agree.
  # VERIFY that rather than assuming it: a row that still carries a
  # pre-override value is flagged loudly.
  if (identical(args$sg_focus, "maxeff")) {
    forced <- list(pconsistency.threshold = 0, stop_threshold = NULL,
                   use_twostage = FALSE, max_subgroups_search = Inf, minp = 0)
    for (p in names(forced)) {
      i <- which(out$parameter == p)
      if (!length(i)) next
      agrees <- isTRUE(all.equal(args[[p]], forced[[p]]))
      out$status[i] <- if (agrees)
        "maxeff engine-forced value (pre-resolved in this call)" else
        "STALE -- captured BEFORE the maxeff override; not what the engine used"
    }
  }

  # stop_threshold is reset to NULL for these foci, but that branch DOES call
  # .sync_args_call_all(), so the capture is already truthful; label it.
  if (identical(args$sg_focus, "hrMaxSG") ||
      identical(args$sg_focus, "hrMinSG") ||
      args$sg_focus %in% c("hr", "maxSG", "minSG")) {
    i <- which(out$parameter == "stop_threshold")
    if (length(i) && is.null(args$stop_threshold))
      out$status[i] <- "NULL: early stopping unsound for this key (re-synced)"
  }

  out
}

#' Render a provenance table for one fit
fs_param_table <- function(fit, title) {
  pv <- fs_param_provenance(fit)
  n_diff <- sum(pv$status == "differs from default")
  sub <- sprintf(
    "%d of %d formals differ from their package default (forestsearch %s, read at render time).",
    n_diff, nrow(pv),
    tryCatch(as.character(utils::packageVersion("forestsearch")),
             error = function(e) "version unknown"))
  if (any(grepl("^STALE", pv$status))) {
    sub <- paste(sub,
      "WARNING: one or more rows were captured before the sg_focus = \"maxeff\"",
      "override and do not report the value the engine used.")
  }
  knitr::kable(pv, caption = paste0(title, " -- ", sub))
}
```

## Identification {#sec-fit}

The receptor cuts $\{er \le 0\}$ and $\{pgr \le 0\}$ are pre-specified as forced
candidates, and $J = 10$ cuts are used for the continuous `er` and `pgr`
covariates. Candidate generation is the consistency search alone -- no LASSO,
GRF or DINA screens contribute cuts -- so the family the selection ranges over is
determined entirely by the cut construction above.

```{r forestsearch-fit}
#| fig-width: 10
#| fig-height: 7
t0 <- proc.time()

fs <- forestsearch(
  df.analysis,
  outcome.name          = outcome.name,
  event.name            = event.name,
  treat.name            = treat.name,
  id.name               = id.name,
  confounders.name      = confounders.name,
  is.RCT                = TRUE,
  seedit                = 8316951,
  est.scale             = "hr",
  # candidate generation: consistency search only (no LASSO/GRF/DINA screens)
  use_lasso             = FALSE,
  use_grf               = FALSE,
  use_dina              = FALSE,
  # continuous-cut construction + pre-specified receptor candidates
  conf.cont_jcuts       = list(er = 10, pgr = 10),
  collapse_cuts         = TRUE,
  conf_force            = c("er <= 0", "pgr <= 0"),
  # selection: driven entirely by the focus profile
  subgroup_method       = "consistency",
  sg_focus              = focus$supplied,
  selection_rule        = focus$selection_rule,
  effect_neighborhood   = focus$effect_neighborhood,
  # Tier-2 de-biased gate (MR) computed inside the fit.  MR re-selects under
  # the SAME rule the search used -- .fs_mr_reselection_from_focus() maps
  # hrMaxSG -> effMaxSG, hrMinSG -> effMinSG, hr -> maxcons, and so on -- so
  # no reselection argument is needed here.
  mr_inference          = TRUE,
  mr_inference_args     = list(draws = mr_draws),
  # subgroup constraints (retained under every focus: they define which
  # candidates are ESTIMABLE, not which one wins)
  n.min                 = 60, d0.min = 10, d1.min = 10, maxk = 2,
  # thresholds + search settings, pre-resolved by the focus profile
  hr.threshold          = focus$hr_threshold,
  hr.consistency        = focus$hr_consistency,
  pconsistency.threshold = focus$pconsistency,
  minp                  = focus$minp,
  use_twostage          = focus$use_twostage,
  stop_threshold        = focus$stop_threshold,
  consistency_method    = fs_consistency_method,
  fs.splits             = 1000,
  max.minutes           = 3,
  show_candidate_summary = TRUE, details = TRUE, quiet = FALSE,
  plot.sg               = TRUE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE)
)
plan("sequential")
timings$fit <- (proc.time() - t0)[["elapsed"]]

stopifnot(!is.null(fs$sg.harm))
cat("\nSelected subgroup:", paste(fs$sg.harm, collapse = " & "),
    "\n  sg_focus (canonical):", fs$sg_focus,
    "\n  fit:", round(timings$fit / 60, 1), "min\n")

# The engine's canonical focus must agree with the profile; if it does not,
# every gate below is keyed on the wrong value.
stopifnot(identical(fs$sg_focus, focus$canonical))
```

```{r param-provenance}
#| echo: false
fs_param_table(fs, sprintf("Resolved parameters for the %s fit", focus$supplied))
```

### Identified subgroups and estimates

```{r fs-results}
res_tabs <- sg_tables(fs, ndecimals = 3, which_df = "est")
res_tabs$sg10_out
```

```{r fs-estimates}
res_tabs$tab_estimates
```

```{r fs-subgroup}
cat("Identified subgroup (H):", paste(fs$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs$df.est$treat.recommend == 0)))
```

## Selection diagnostic: Pareto frontier {#sec-pareto}

```{r pareto-skip-note, eval = !run_pareto}
#| echo: false
#| results: asis
cat(sprintf(paste(
  "_The Pareto section is skipped for `sg_focus = \"%s\"`._\n\n",
  "The frontier block runs only for the banded foci `effMaxSG` / `effMinSG`",
  "(canonical `hrMaxSG` / `hrMinSG`), where `.compute_inclusion_band()` makes",
  "the (effect, N) geometry part of the **selection rule** -- `in_band` leads",
  "the sort key, and `selection_rule` switches between the effect",
  "neighbourhood, the frontier, and their intersection. Under `%s` the key",
  "contains no `in_band` term, so a frontier plot would be a post-hoc curiosity",
  "rather than a diagnostic of how the winner was chosen.\n\n",
  "Under `maxeff` there is a second reason: the winner-only fast path scores",
  "consistency for the argmax alone and carries every other candidate with",
  "`Pcons = NA`, so the frontier can run to hundreds of members with an",
  "all-missing consistency column, and `compute_frontier_cis()` costs scale",
  "with frontier size x `n_splits`.\n"),
  focus$supplied, focus$supplied))
```

```{r pareto-intro, eval = run_pareto}
#| echo: false
#| results: asis
cat(sprintf(paste(
  "The *Pareto frontier* on (effect, N) is the set of candidate subgroups not",
  "dominated on both axes simultaneously: a candidate is dominated when another",
  "has at least as large an effect **and** at least as large a sample size, with",
  "at least one strict inequality.\n\n",
  "For `sg_focus = \"%s\"` with `selection_rule = \"%s\"`, this is not a post-hoc",
  "diagnostic -- it is the selection geometry. `.compute_inclusion_band()`",
  "builds `in_nbhd` (effect within %.0f%% of the strongest surviving effect) and",
  "`in_pareto` (non-dominated), and %s. The `in_band` indicator then **leads**",
  "the sort key %s, so what the figure and table below display is the space the",
  "rule actually chose within.%s\n"),
  focus$supplied, focus$selection_rule, 100 * focus$effect_neighborhood,
  switch(focus$selection_rule,
    neighborhood = "the rule uses the effect band alone",
    pareto       = "the rule uses the frontier alone",
    both         = "the rule uses their intersection"),
  focus$key,
  if (identical(focus$canonical, "hrMinSG"))
    paste(" Because this focus minimises size within the band, the selected",
          "subgroup may itself be N-dominated by a larger in-band neighbour and",
          "so may not appear on the frontier; that is the rule working as",
          "specified.") else ""))
```

```{r pareto-cis, eval = run_pareto}
# Split-derived CIs for the frontier members.  This is the expensive step; the
# same intervals feed the figure, the table and the narrative, so it is
# computed once.  Set frontier_params$compute_cis <- FALSE to skip.
t0 <- proc.time()
ci_tab <- if (isTRUE(frontier_params$compute_cis)) {
  compute_frontier_cis(
    fs,
    n_splits = frontier_params$n_splits,
    seed     = frontier_params$seed
  )
} else {
  NULL
}
timings$frontier <- (proc.time() - t0)[["elapsed"]]
cat("Frontier CIs:", if (is.null(ci_tab)) "skipped" else
      sprintf("%d member(s) over %s splits in %.1f s",
              nrow(ci_tab), format(frontier_params$n_splits, big.mark = ","),
              timings$frontier), "\n")
```

```{r pareto-plot, eval = run_pareto}
#| fig-width: 9
#| fig-height: 6
#| fig-cap: "Pareto frontier on (effect, N). The shaded region is the effect-size inclusion band; the selected subgroup is marked. Horizontal bars are split-derived intervals when computed."
p_pareto <- plot_pareto_frontier(
  fs,
  ci_table            = ci_tab,
  show_band           = TRUE,
  effect_neighborhood = focus$effect_neighborhood,
  xlim                = frontier_params$xlim,
  xlim_trim           = frontier_params$xlim_trim
)
print(p_pareto)
```

```{r pareto-table, eval = run_pareto}
# Formatted gt table; the selected subgroup is highlighted and marked with a
# star.  Effect column label and scale derive from fs$effect_measure, so the
# same call serves survival (HR) and GLM (OR, IRR, MD, ...) outcomes.
pareto_frontier_table(
  fs,
  ci_table = ci_tab,
  digits   = frontier_params$digits
)
```

```{r pareto-explain, eval = run_pareto}
#| results: asis
# Narrative comparison of the winner against each non-dominated alternative,
# phrased in terms of the focus's own lexicographic key.
cat(explain_pareto_selection(fs, ci_table = ci_tab,
                             digits = frontier_params$digits))
```

```{r pareto-data, eval = FALSE}
# Programmatic access (effect on the natural scale, plus is_selected /
# on_frontier / in_band flags):
pareto_frontier_table(fs, format = "data.table", include_dominated = TRUE)
```

## Full bootstrap (FB)

```{r bootstrap}
t0 <- proc.time()
fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est   = fs,
  nb_boots = NB,
  show_three = FALSE,
  details  = FALSE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE)
)
plan("sequential")
timings$fb <- (proc.time() - t0)[["elapsed"]]

fb_sum <- summarize_bootstrap_results(
  sgharm = fs$sg.harm, boot_results = fs_bc,
  create_plots = FALSE, est.scale = "hr"
)
cat("FB completed in", round(timings$fb / 60, 1), "min\n")
```

FB re-runs the **entire** selection on each resample, so it inherits whatever
focus this render uses without any further configuration -- and its cost scales
with `NB`.

## Multiplier resampling (MR)

The Tier-2 de-biased gate is the MR correction, computed inside the fit above
(vectorized over `r format(mr_draws, big.mark = ",")` draws). It is read from
`fs$mr_inference`; no separate run is needed.

MR re-selects under the same rule the search used. `sg_focus` is the single
source of truth for that mapping, so the re-selection reported below should
match this render's focus rather than a default.

```{r mr}
g <- fs$mr_inference
stopifnot(!is.null(g))
timings$mr <- g$timing_seconds
cat("MR gate:", round(timings$mr, 2), "s over",
    format(mr_draws, big.mark = ","), "draws\n")
cat("MR re-selection rule:", g$settings$reselection,
    sprintf("(search focus: %s / canonical %s)\n",
            focus$supplied, focus$canonical))
cat("Candidate family size:", g$n_family, "\n")

# Scale cross-check: the gate's naive HR and the bootstrap's naive H0 are the
# same quantity computed two ways.
.h0 <- tryCatch(as.numeric(fs_bc$H_estimates$H0), error = function(e) NA_real_)
if (is.finite(.h0) && is.finite(g$naive$est)) {
  cat(sprintf("Scale cross-check (MR naive vs FB H0): %s (%.3f vs %.3f)\n",
              if (abs(log(g$naive$est) - log(.h0)) < 1e-2) "PASS" else "CHECK",
              g$naive$est, .h0))
}
```

## Guo & He (GH)

```{r gh-skip-note, eval = !run_gh}
#| echo: false
#| results: asis
cat(if (!focus$run_guohe) sprintf(paste(
  "_The Guo & He section is skipped for `sg_focus = \"%s\"`._\n\n",
  "This is a structural constraint rather than a configuration choice.",
  "`fs_to_guohe()` calls `fs_export_maxeff_family()`, which stops unless",
  "`fit$sg_focus == \"maxeff\"` (`fs_to_guohe.R:141-143`), and it materializes",
  "the **untruncated** deduplicated family that only `maxeff` guarantees",
  "(`max_subgroups_search -> Inf`).\n\n",
  "The deeper reason is that Guo & He's estimand is defined against the argmax",
  "primitive: their correction removes the winner's-curse optimism of selecting",
  "the maximiser of an estimated effect. A rule that optimises **size** subject",
  "to an effect band is a different selection event, and their Theorems 4/5 say",
  "nothing about it. Running a second `maxeff` fit purely to obtain a",
  "comparator would report a correction for an estimator this document did not",
  "use, and pooling across foci silently mixes estimators.\n\n",
  "Set `fs_sg_focus <- \"maxeff\"` for the head-to-head with MR.\n"),
  focus$supplied) else
  "_Guo & He skipped by request (`gh_requested <- FALSE`)._\n")
```

```{r guohe, eval = run_gh}
# The adapter exports the maxeff candidate family as membership columns and runs
# Guo & He Algorithm 3 on the SAME selected subgroup forestsearch chose.  The
# membership integrity check asserts that the exported selected-subgroup column
# matches forestsearch's own sg.harm.id exactly before the estimator runs.
t0 <- proc.time()
gh <- fs_to_guohe(
  fit       = fs,
  time      = "time_months", event = "status", treatment = "hormon",
  B         = GH_B, r = GH_r, level = 0.05, seed = 8316951,
  parallel  = GH_PARALLEL, verify = TRUE
)
timings$gh <- (proc.time() - t0)[["elapsed"]]

cat("GH family size:", length(gh$export$candidates),
    " | selected:", gh$selected,
    " | GH", round(timings$gh / 60, 1), "min\n")
cat(sprintf("GH  naive HR %.3f  debiased HR %.3f  lower bound %.3f\n",
            gh$naive_hr, gh$debiased_hr, gh$bound_hr))
```

## Repeated 10-fold cross-validation (stability diagnostics)

Repeated 10-fold cross-validation partitions the data into ten folds, refits the
selection on each nine-fold training set, and records which subgroup is
identified -- repeated over `Ksims` random partitions. It is a **stability
diagnostic on the subgroup _definition_**: across `Ksims` × 10 training splits,
does the selection converge on a single dominant subgroup, or does it wander with
the particular split?

This is a fixed-family analysis with no GRF, DINA or LASSO screens, so the
GRF-specific fold diagnostics are omitted; the universal ones -- identified-
subgroup distribution, selected-candidate consistency, per-fold numerics, and
agreement with the full-data selection -- are reported.

```{r cv-kfold-skip-note, eval = !run_cv_kfold, results = "asis"}
cat("> **10-fold CV skipped.** Set `run_cv_kfold <- TRUE` in the",
    "configuration to enable.\n")
```

```{r kfold-cv, eval = run_cv_kfold}
t0 <- proc.time()
fs_kfold <- forestsearch_tenfold(
  fs.est  = fs,
  sims    = Ksims,
  Kfolds  = 10,
  details = FALSE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = FALSE)
)
plan("sequential")
timings$kfold <- (proc.time() - t0)[["elapsed"]]
cat("10-fold CV:", Ksims, "x 10 partitions in",
    round(timings$kfold / 60, 1), "min\n")
metrics_tables <- cv_metrics_tables(fs_kfold)
metrics_tables
```

### Diagnostic tabulations

```{r cv-diagnostics-run, eval = run_cv_kfold}
cv_diag <- summarize_cv_results(
  cv_output    = fs_kfold,
  original_sg  = fs$sg.harm,
  top_n        = 10L,
  create_plots = TRUE
)
```

**Which subgroups emerged across resamples?** A single dominant subgroup with
high share indicates the selection converges; a long tail indicates it is
sensitive to the training split.

```{r cv-identification, eval = run_cv_kfold}
cv_diag$identification_summary
```

**How confidently were the successful identifications made?** Among folds that
identified a subgroup, `pconsistency_distribution` shows where the selected
candidate's consistency probability landed.

```{r cv-pcons-note}
#| echo: false
#| results: asis
cat(if (focus$is_maxeff) paste(
  "Under `maxeff` no consistency threshold is applied at selection, so this is",
  "descriptive of the selected argmax rather than a gate.\n") else if (focus$is_banded) paste(
  "Under a banded focus consistency is a **filter** at admission and then a",
  "third-tier key after `in_band` and size, so the distribution reflects both",
  "roles.\n") else paste(
  "Consistency is a filter at admission for this focus, so every value shown",
  "cleared `pconsistency.threshold`.\n"))
```

```{r cv-pcons, eval = run_cv_kfold}
cv_diag$pconsistency_distribution
```

**How do the numeric fold-level diagnostics summarize?** Median / IQR / range for
each numeric fold-level column: fold size (`n_test`), training-fold HR of the
selected subgroup (`training_fs_hr`, in-sample and optimistically biased -- for
diagnostic comparison only), and candidate count
(`n_candidates_evaluated`).

```{r cv-fold-numeric, eval = run_cv_kfold}
cv_diag$fold_numeric_summary
```

**Agreement with the full-data analysis.** How often does a training fold
reproduce the original full-data subgroup exactly, or share a factor with it?

```{r cv-original-agreement, eval = run_cv_kfold}
cv_diag$original_agreement
```

```{r cv-identification-plot, eval = run_cv_kfold}
#| fig-width: 9
#| fig-height: 5
#| fig-cap: "Distribution of identified subgroups across repeated 10-fold training splits."
if (!is.null(cv_diag$plots) && !is.null(cv_diag$plots$identification)) {
  cv_diag$plots$identification
}
```

## Leave-one-out (honest out-of-sample estimate)

Leave-one-out re-runs the entire selection on each hold-out of one subject and
predicts the held-out subject's subgroup membership from its covariates. It is an
out-of-sample **stability check on the subgroup _definition_**, not a de-biasing
correction of the _effect_: it asks whether the same subgroup is recovered when
no subject informs its own label. This is `forestsearch_Kfold()` at N-fold
(`Kfolds = nrow(df)`).

```{r loo-focus-note}
#| echo: false
#| results: asis
cat(sprintf(paste(
  "How discriminating leave-one-out is depends on the focus. Under `%s` the key",
  "is %s. %s\n"),
  focus$supplied, focus$key,
  switch(focus$canonical,
    maxeff = paste(
      "With the full, finer-grained candidate family and no auxiliary condition,",
      "the argmax has genuine room to land on different definitions across folds,",
      "so this is the discriminating comparison: whether the out-of-sample",
      "estimate tracks Naive (definition perfectly stable) or one of the",
      "corrections (definition varies fold to fold) is informative rather than",
      "predetermined."),
    maxeffCons = paste(
      "The consistency filter removes the least replicable candidates before the",
      "effect argmax runs, so the definition is typically steadier than under",
      "`maxeff` -- but the winner can still move when a marginal qualifier drops",
      "out of a fold."),
    hrMaxSG = ,
    hrMinSG = paste(
      "The band restricts the competition to near-optimal candidates and size then",
      "decides among them, which tends to make the definition steadier than a pure",
      "effect argmax: fold-to-fold movement requires the band membership or the",
      "size ordering to change, not merely the effect ranking."),
    maxSG = ,
    minSG = paste(
      "Size leads the key and is the most stable quantity in the table, so the",
      "definition is usually recovered exactly and leave-one-out will often",
      "coincide with Naive by construction."),
    paste(
      "Consistency leads the key, and consistency is estimated from repeated",
      "splits rather than a single fit, so the definition is usually steady;",
      "coincidence with Naive should be read as stability, not as a failure to",
      "correct."))))
```

Note this is the heaviest step: N-fold means one full forest search per subject.
The fold loop is distributed across `r loo_workers` workers; `parallel_args`
controls **fold-level** parallelism only -- each fold's inner `forestsearch()`
refit is pinned to `plan = "sequential"`, so there is no nested oversubscription
-- and the fold loop draws parallel-safe L'Ecuyer streams
(`.options.future = list(seed = TRUE)`). In the frozen-family companion analysis
the parallel fold loop was verified against a sequential run of the same
analysis: identical per-subject predictions and an identical out-of-sample HR, at
5.2x the speed on 11 workers.

```{r loo-skipped-note, eval = !run_loo}
cat("Leave-one-out skipped (run_loo = FALSE).",
    "Set run_loo <- TRUE in parallel-setup to run it.\n")
```

```{r loo, eval = run_loo}
# If a precomputed cv_out is available (LOO_CACHE env var, else the default
# path in gh_dir) load it; otherwise run the N-fold search and cache it for
# re-renders.  The cache key carries the focus, so switching fs_sg_focus does
# NOT silently reuse another focus's leave-one-out result.
loo_cache <- Sys.getenv(
  "LOO_CACHE",
  unset = file.path(gh_dir, sprintf("cv_out_%s.rds", focus$canonical)))
if (file.exists(loo_cache)) {
  cv_out <- readRDS(loo_cache)
  timings$loo <- NA_real_
  cat("LOO loaded from cache:", loo_cache, "\n")
} else {
  t0 <- proc.time()
  fs_OOB <- forestsearch_Kfold(
    fs.est  = fs,
    details = FALSE,
    Kfolds  = nrow(df.analysis),   # N-fold = leave-one-out
    parallel_args = list(plan = "multisession", workers = loo_workers,
                         show_message = TRUE)
  )
  plan("sequential")
  timings$loo <- (proc.time() - t0)[["elapsed"]]
  cv_out <- forestsearch_KfoldOut(res = fs_OOB, details = FALSE, outall = TRUE)
  saveRDS(cv_out, loo_cache)
}

# Out-of-sample harm-subgroup HR, read BY IDENTITY (the harm row of the K-fold
# table, matched on its subgroup label sg0.name = "Not recommend"), never
# positionally -- the table layout depends on n_sgfound.  Two structural cases:
#   (i)  stable subgroup: SG_tab_Kfold is a matrix whose harm row carries the
#        sg0.name label -> its out-of-sample HR;
#   (ii) ITT fallback: no stable subgroup survives out-of-sample
#        (n_sgfound == 1) and SG_tab_Kfold collapses to the 1-D ITT vector.
.loo_harm_row <- function(cv_out, harm_label = "Not recommend") {
  sgk <- cv_out$SG_tab_Kfold
  if (is.matrix(sgk) && "Subgroup" %in% colnames(sgk)) {
    j <- which(sgk[, "Subgroup"] == harm_label)
    if (length(j) == 1L)
      return(list(case = "stable",
                  hr   = as.character(sgk[j, "HR (95% CI)"]),
                  n    = as.character(sgk[j, "n"])))
  }
  list(case = "itt_fallback",
       hr   = "\u2014 (no stable subgroup out-of-sample; ITT)",
       n    = NA_character_)
}
loo <- .loo_harm_row(cv_out)

metr <- cv_out$sens_metrics_original
cat(sprintf(
  "LOO [%s]: out-of-sample harm HR %s (n = %s)\n  agreement: sens_H = %.2f, ppv_H = %.2f (exact-match = %.2f)\n",
  loo$case, loo$hr, loo$n, metr[["sens_H"]], metr[["ppv_H"]],
  cv_out$find_metrics[["Exact"]]))

tables <- cv_summary_tables(cv_out)
tables$combined_table
tables$metrics_table
```

## Comparison: corrections on the selected subgroup {#sec-table}

```{r table}
#| echo: false
fci <- function(e, l, u, d = 2) {
  ifelse(is.na(e), "\u2014", sprintf("%.*f (%.*f, %.*f)", d, e, d, l, d, u))
}

# Naive is read from the bootstrap object's H_estimates ($H0 with
# H0_lower/H0_upper), which exists under every focus.  Guo & He reports the
# same quantity when it runs; cross-check rather than depend on it.
naive_hr <- tryCatch(as.numeric(fs_bc$H_estimates$H0), error = function(e) NA_real_)
if (isTRUE(run_gh) && exists("gh")) {
  if (is.finite(naive_hr) && is.finite(gh$naive_hr) &&
      abs(log(gh$naive_hr) - log(naive_hr)) > 1e-2) {
    warning("Naive HR disagrees between FB H0 (", round(naive_hr, 4),
            ") and Guo & He (", round(gh$naive_hr, 4), ").", call. = FALSE)
  }
}

loo_row <- if (isTRUE(run_loo) && exists("loo")) loo$hr else "\u2014 (run_loo = FALSE)"

method <- c("Naive", "Full bootstrap", "Multiplier resampling")
estimate <- c(
  sprintf("%.2f", naive_hr),
  fci(fs_bc$H_estimates$H2, fs_bc$H_estimates$H2_lower, fs_bc$H_estimates$H2_upper),
  fci(g$debiased$est, g$debiased$lower, g$debiased$upper)
)
if (isTRUE(run_gh) && exists("gh")) {
  method   <- c(method, "Guo & He (2021)")
  estimate <- c(estimate,
                sprintf("%.2f (LB %.2f)", gh$debiased_hr, gh$bound_hr))
}
method   <- c(method, "Leave-one-out (out-of-sample)")
estimate <- c(estimate, loo_row)

kbl(data.frame(Method = method, Estimate = estimate, check.names = FALSE),
    booktabs = TRUE, escape = FALSE, align = "ll",
    caption = sprintf(paste(
      "Subgroup hazard ratios for the GBSG selected subgroup under",
      "sg_focus = '%s'. Parenthetical values are 95%% intervals%s.",
      "Leave-one-out is NOT a de-biasing correction: it is an out-of-sample",
      "stability check on the subgroup definition."),
      focus$supplied,
      if (isTRUE(run_gh)) "; Guo & He reports a one-sided lower bound" else "")) |>
  kable_styling()
```

::: {.callout-note}
## Reading the table

Naive is the uncorrected subgroup hazard ratio. Full bootstrap and Multiplier
resampling both target the conditional estimand and report two-sided intervals;
they differ in that FB re-runs the entire selection on each resample (so its cost
scales with the resample count) while MR holds the fitted family fixed and is
vectorized (so its cost does not). FB and MR are a trade, not a strict ordering:
FB marginalizes over family regeneration, MR conditions on the realized family.

Leave-one-out is **not a de-biasing correction** and should not be read as one.
It is an out-of-sample **stability check on the subgroup _definition_**. If every
fold recovers the *identical* definition, out-of-sample membership equals
in-sample membership and the leave-one-out HR is mechanically the naive one -- so
`LOO = Naive` indicates a perfectly stable definition, not a failure to correct.
If the selection instead lands on *different* definitions across folds,
out-of-sample membership departs from in-sample and the leave-one-out HR moves
away from Naive -- and *where* it moves is the informative signal. Read the
leave-one-out row together with the agreement metrics (`sens_H`, `ppv_H`,
exact-match). Leave-one-out validates the *definition*; the de-biasing
corrections address the *effect* optimism -- distinct questions, so agreement or
disagreement between them is not a contradiction.
:::

```{r gh-callout, eval = run_gh}
#| echo: false
#| results: asis
# Built as a character vector rather than paste(): a fenced div and an ATX
# heading must begin at column 1, and paste()'s default sep = " " would insert
# a leading space before ':::' and '##'.
cat(c(
  "::: {.callout-note}",
  "## Reading the Guo & He row",
  "",
  paste(
    "Guo & He corrects the winner's-curse optimism of the argmax selection and",
    "reports a de-biased point estimate with a one-sided lower confidence bound --",
    "the object their Theorems 4/5 and simulations certify. They assume a fixed",
    "candidate family; the maxeff family here is large (hundreds to ~1744",
    "candidates depending on cut construction) and data-adjacent, which is outside",
    "their stated small-fixed-$k$ regime. The Section 5 reproduction established",
    "that their estimator is faithfully implemented on disjoint families up to",
    "$k = 12$; its behaviour on a family this size is the open question, and is",
    "exactly why the head-to-head with MR is of interest."),
  ":::",
  ""), sep = "\n")
```

```{r timing}
#| echo: false
td <- data.frame(
  Step = c("Fit + MR gate", "Frontier CIs", "Full bootstrap",
           "MR gate (within fit)", "Guo & He bootstrap", "10-fold CV",
           sprintf("Leave-one-out (N-fold, %d workers)", loo_workers)),
  Seconds = c(timings$fit, timings$frontier, timings$fb, timings$mr,
              timings$gh, timings$kfold, timings$loo)
)
td$Minutes <- round(td$Seconds / 60, 2)
kbl(td, booktabs = TRUE, align = "lrr",
    caption = sprintf("Computational timing (sg_focus = '%s'). NA marks a step this focus or configuration did not run.",
                      focus$supplied)) |>
  kable_styling()
```

## Session information

```{r session-info}
#| code-fold: true
sessionInfo()
```

# Reproducibility payload {#sec-export-payload}

The chunk below harvests the objects this render already computed into a single
small list and writes it as an `.rds`, so the results survive the HTML. No
analysis is repeated. Every element is guarded, so a section this render skipped
(`run_pareto`, `run_gh`, `run_cv_kfold`, `run_loo`) stores `NULL` rather than
erroring, and the top-level names match the other GBSG payloads
(`table`, `labels`, `meta`, `extras`, `est_scale`, `built_at`,
`forestsearch_version`) so a reader can treat them uniformly.

```{r export-payload}
#| echo: false
#| message: false

.qmd_dir  <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) getwd())
.qmd_stem <- tryCatch(tools::file_path_sans_ext(basename(knitr::current_input())),
                      error = function(e) "analysis")
.dirout   <- if (is.null(dirout)) .qmd_stem else dirout
.base     <- if (is.null(results_dir)) file.path(.qmd_dir, "_payloads") else results_dir
.out_dir  <- file.path(.base, .dirout)
dir.create(.out_dir, recursive = TRUE, showWarnings = FALSE)
.payload_file <- file.path(.out_dir, paste0(.dirout, "_payload.rds"))

.get <- function(nm) if (exists(nm, inherits = TRUE)) get(nm, inherits = TRUE) else NULL
.try <- function(expr) tryCatch(expr, error = function(e) NULL)

.fs   <- .get("fs")
.fsbc <- .get("fs_bc")
.g    <- if (is.null(.fs)) NULL else .fs$mr_inference
.gh   <- .get("gh")
.cvd  <- .get("cv_diag")
.cvo  <- .get("cv_out")
.loo  <- .get("loo")

# The comparison chunk passes its Method/Estimate vectors straight into kbl()
# without binding a frame, so rebuild the frame here rather than re-deriving
# any estimate.
.method <- .get("method"); .estimate <- .get("estimate")
.table <- if (is.character(.method) && is.character(.estimate) &&
              length(.method) == length(.estimate))
  data.frame(Method = .method, Estimate = .estimate,
             check.names = FALSE, stringsAsFactors = FALSE) else NULL

payload <- list(
  table  = .table,
  labels = list(
    sg_harm  = if (is.null(.fs)) NULL else .fs$sg.harm,
    sg_focus = if (is.null(.fs)) NULL else .fs$sg_focus,
    focus    = .get("focus")),
  meta = list(
    NB          = .get("NB"),
    mr_draws    = .get("mr_draws"),
    Ksims       = .get("Ksims"),
    n_cores     = .get("n_cores"),
    loo_workers = .get("loo_workers"),
    n_total     = .try(nrow(.get("df.analysis"))),
    n_events    = .try(sum(.get("df.analysis")[[.get("event.name")]])),
    outcome     = .get("outcome.name"),
    event       = .get("event.name"),
    treat       = .get("treat.name"),
    covariates  = .get("confounders.name"),
    timings     = .get("timings"),
    provenance  = .try(fs_param_provenance(.fs))),
  extras = list(
    mr = if (is.null(.g)) NULL else list(
      naive          = .g$naive,
      debiased       = .g$debiased,
      settings       = .g$settings,
      n_family       = .g$n_family,
      selection_bias = .g$selection_bias,
      timing_seconds = .g$timing_seconds),
    fb = if (is.null(.fsbc)) NULL else list(
      H_estimates  = .fsbc$H_estimates,
      Hc_estimates = .fsbc$Hc_estimates),
    pareto = if (isTRUE(.get("run_pareto"))) list(
      frontier = .try(pareto_frontier_table(.fs, format = "data.table",
                                            include_dominated = TRUE)),
      ci_tab   = .get("ci_tab"),
      params   = .get("frontier_params")) else NULL,
    cv = if (isTRUE(.get("run_cv_kfold")) && !is.null(.cvd)) list(
      metrics                   = .get("metrics_tables"),
      identification_summary    = .cvd$identification_summary,
      pconsistency_distribution = .cvd$pconsistency_distribution,
      fold_numeric_summary      = .cvd$fold_numeric_summary,
      original_agreement        = .cvd$original_agreement) else NULL,
    loo = if (isTRUE(.get("run_loo")) && !is.null(.loo)) list(
      row          = .loo,
      sens_metrics = if (is.null(.cvo)) NULL else .cvo$sens_metrics_original,
      find_metrics = if (is.null(.cvo)) NULL else .cvo$find_metrics,
      cache        = .get("loo_cache")) else NULL,
    gh = if (is.null(.gh)) NULL else list(
      naive_hr    = .gh$naive_hr,
      debiased_hr = .gh$debiased_hr,
      bound_hr    = .gh$bound_hr,
      selected    = .gh$selected,
      n_family    = .try(length(.gh$export$candidates)),
      B           = .get("GH_B"),
      r           = .get("GH_r"))),
  est_scale = "hr",
  built_at  = Sys.time(),
  forestsearch_version = tryCatch(as.character(utils::packageVersion("forestsearch")),
                                  error = function(e) NA_character_))

saveRDS(payload, .payload_file)
cat("Wrote payload to:", .payload_file, "\n")
```
