---
title: "Depth-2 Subgroup Identification with DINA and ForestSearch"
subtitle: "German Breast Cancer Study Group (GBSG) Analysis -- AND-Conjunctions of Two Covariates"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# Initialize timing
timings <- list()
t_vignette_start <- proc.time()
```

## Parallel Processing Configuration

ForestSearch supports parallel processing for the bootstrap. The depth-2
search is only modestly more expensive than depth-1 (the candidate count
grows by roughly `choose(K, 2) * (2 * length(grid_probs))^2`), so the same
configuration applies.

```{r parallel-setup}
# Detect available cores (capped for routine rendering / CRAN checks)
n_cores_total <- parallel::detectCores()
n_cores <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))
cat("Using", n_cores, "of", n_cores_total, "total cores for parallel processing")

# Number of bootstrap iterations.
# Use 500-2000 for production; reduced here for the vignette.
NB <- 20

# ---------------------------------------------------------------------------
# Depth-2 controls.  max_depth = 2L lets a selected subgroup be an
# AND-conjunction of two distinct covariates; grid_probs is the per-covariate
# quantile grid used to enumerate the conjunction thresholds.  Setting
# max_depth = 1L reproduces the classic single-covariate behaviour exactly.
# ---------------------------------------------------------------------------
dina_max_depth  <- 2L
dina_grid_probs <- seq(0.1, 0.9, by = 0.1)   # interior deciles (default)

# ---------------------------------------------------------------------------
# Per-analysis identification parameters (tune each analysis independently).
#   sg_focus      : which candidate subgroup the selection step chooses
#   hr_threshold  : harm/benefit floor on the HR scale
#   hr_consistency: minimum consistency rate (consistency search only)
# ---------------------------------------------------------------------------

# (A) ForestSearch + DINA screening (consistency search selects; DINA only
#     contributes candidate cuts -- here, a depth-2 conjunction's components)
dina_screen_sg_focus             <- "effMaxSG"
dina_screen_hr_threshold         <- 0.90
dina_screen_hr_consistency       <- 0.80
dina_screen_max_subgroups_search <- 30
# Screening cut set: FALSE => full DINA frontier; TRUE => only the single
# dina_subgroup()-selected subgroup (a conjunction at max_depth = 2), whose
# component cuts are injected into the screening pool.
dina_screen_selected_only        <- TRUE

# (B) Standalone DINA and (C) ForestSearch subgroup_method = "dina".  (B) and
# (C) share these so the standalone selection and the DINA-selection pipeline
# match.
fs_dina_sg_focus     <- "effMaxSG"
fs_dina_hr_threshold <- 0.80
```



# Introduction

This document is a companion to the main GBSG ForestSearch analysis. It
exercises the **depth-2** extension of the DINA subgroup engine end-to-end,
in the three ways DINA plugs into the ForestSearch workflow, using the same
data, confounders and selection focus as the depth-1 analysis.

A *depth-1* subgroup is a single covariate threshold, e.g. `{nodes >= 12}`.
A *depth-2* subgroup is an **AND-conjunction of two distinct covariates**,
e.g. `{grade3 >= 1 & pgr <= 32.5}`. Depth-2 is opt-in via `max_depth = 2L`
and is constructed so that a conjunction is selected only when it strictly
outranks every single-covariate candidate under the configured `sg_focus`;
with `max_depth = 1L` the behaviour is byte-identical to the classic engine.

The three modes demonstrated:

* **(A)** ForestSearch *using DINA for screening* (`use_dina = TRUE`,
  `dina_args$max_depth = 2`) -- DINA's selected conjunction contributes its
  component cuts to the candidate pool; the consistency search still selects.
* **(B)** *Standalone DINA* at depth 2 -- `dina_subgroup(max_depth = 2)`
  returns a conjunction, and `dina_subgroup_bootstrap(max_depth = 2)`
  provides per-component threshold stability and a cut-set selection
  frequency.
* **(C)** ForestSearch with `subgroup_method = "dina"` and
  `dina_args$max_depth = 2` -- selection is delegated to the depth-2
  `dina_subgroup()`, then the standard ForestSearch bootstrap produces the
  selection bias-corrected Cox estimate and the per-resample stability
  diagnostics (now populated for DINA mode and depth-aware).



# Setup

## Load Required Packages

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(doFuture)

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```


# Data: German Breast Cancer Study Group Trial

```{r data-setup}
# Load GBSG data (included in the forestsearch package)
df.analysis <- gbsg

# Prepare analysis variables
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
  treat <- hormon
})

# Define variable roles
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name   <- "status"
id.name      <- "id"
treat.name   <- "hormon"

cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```


# DINA Depth-2 Configuration

```{r dina-config}
#| code-fold: false

# Shared DINA *fit* settings (Cox / survival outcome => family mapped to "cox").
dina_seed   <- 8316951
dina_nfolds <- 2

# Harm/benefit floor on the natural-parameter (log-HR) scale for ratio
# measures; (B) and (C) use this floor.
m_diff_dina <- log(fs_dina_hr_threshold)

cat("max_depth:  ", dina_max_depth, "\n")
cat("grid_probs: ", paste(dina_grid_probs, collapse = ", "), "\n")
cat("sg_focus:   ", fs_dina_sg_focus, "\n")
cat("Harm floor: m_diff =", round(m_diff_dina, 4),
    sprintf("(hr.threshold = %.2f)\n", fs_dina_hr_threshold))
```


# (A) ForestSearch Using DINA for Screening (Depth-2)

ForestSearch is run exactly as in the main analysis, but with
`use_dina = TRUE` and `dina_args$max_depth = 2`. With
`dina_args$selected_only = TRUE`, DINA contributes the single subgroup it
would select -- here a depth-2 conjunction -- by injecting that conjunction's
**component cuts** into the candidate pool. The consistency search remains the
gatekeeper and composes cuts on its own, so the final subgroup may be a single
factor or a conjunction depending on what survives consistency selection.

```{r fs-dina-screen}
t0 <- proc.time()

fs_dina_screen <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis,
  outcome.name              = outcome.name,
  event.name                = event.name,
  treat.name                = treat.name,
  id.name                   = id.name,
  confounders.name          = confounders.name,
  is.RCT                    = TRUE,
  seedit                    = 8316951,
  est.scale                 = "hr",
  # --- Screening engines: GRF + LASSO + DINA ------------------------------
  use_lasso                 = TRUE,
  use_grf                   = TRUE,
  grf_depth                 = 2,
  dmin.grf                  = 4,
  frac.tau                  = 0.80,
  return_selected_cuts_only = TRUE,
  use_dina                  = TRUE,
  dina_args = list(
    seed = dina_seed, n_folds = dina_nfolds,
    scope = "wide", max_per_covariate = 3, max_subgroups = 10,
    selected_only = dina_screen_selected_only,
    max_depth     = dina_max_depth,      # depth-2 screening
    grid_probs    = dina_grid_probs
  ),
  # --- Selection engine ---------------------------------------------------
  subgroup_method           = "consistency",
  sg_focus                  = dina_screen_sg_focus,
  selection_rule            = "neighborhood",
  effect_neighborhood       = 0.10,
  # --- Subgroup constraints -----------------------------------------------
  n.min                     = 60,
  maxk                      = 2,
  max_subgroups_search      = dina_screen_max_subgroups_search,
  # --- Thresholds (survival / HR scale) -----------------------------------
  hr.threshold              = dina_screen_hr_threshold,
  hr.consistency            = dina_screen_hr_consistency,
  pconsistency.threshold    = 0.90,
  # --- Output / diagnostics -----------------------------------------------
  show_candidate_summary    = TRUE,
  details                   = TRUE,
  quiet                     = FALSE,
  by.risk                   = 12,
  # --- Parallel -----------------------------------------------------------
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$fs_dina_screen <- (proc.time() - t0)["elapsed"]

cat("Final subgroup (consistency-selected):",
    paste(fs_dina_screen$sg.harm, collapse = " & "), "\n")
if (!is.null(fs_dina_screen$sg.harm)) {
  cat("Subgroup size:", sum(fs_dina_screen$df.est$treat.recommend == 0),
      sprintf("(%.1f%% of ITT)\n",
              100 * mean(fs_dina_screen$df.est$treat.recommend == 0)))
}
```

The cuts DINA contributed to the candidate pool are recorded on the result
object. At `max_depth = 2` the selected DINA subgroup is a conjunction, so its
two component cuts appear here -- these are the candidates the consistency
search may then re-compose.

```{r fs-dina-screen-cuts}
dina_screen_cuts <- fs_dina_screen$dina_cuts
if (length(dina_screen_cuts) > 0) {
  gt(data.frame(
    `#` = seq_along(dina_screen_cuts),
    `DINA candidate cut` = dina_screen_cuts,
    check.names = FALSE
  )) |>
    tab_header(
      title = "Depth-2 DINA Screening Candidates Injected into ForestSearch",
      subtitle = "Component cuts of DINA's selected conjunction"
    ) |>
    cols_align(align = "left")
} else {
  cat("DINA contributed no candidate cuts under the configured frontier.\n")
}
```

```{r fs-dina-screen-pareto}
# Selection diagnostic for the DINA-screened run (selected subgroup marked).
pareto_frontier_table(fs_dina_screen)
```


# (B) Standalone DINA (Depth-2)

The same DINA fit run outside ForestSearch. Its `dina_frontier()` output is
the per-covariate candidate set; `dina_subgroup(max_depth = 2)` returns the
single best subgroup, now allowed to be a two-covariate conjunction; and
`dina_subgroup_bootstrap(max_depth = 2)` provides depth-aware stability
diagnostics.

```{r dina-standalone-fit}
t0 <- proc.time()

dina_fit <- dina(
  df         = df.analysis,
  outcome    = outcome.name,
  treatment  = treat.name,
  covariates = confounders.name,
  family     = "cox",
  status     = event.name,
  seed       = dina_seed,
  n_folds    = dina_nfolds
)

timings$dina_standalone <- (proc.time() - t0)["elapsed"]
cat("Standalone DINA fit complete (family = cox).\n")
```

## DINA Frontier (Per-Covariate Candidates)

```{r dina-standalone-frontier}
dina_fr <- dina_frontier(
  fit               = dina_fit,
  df                = df.analysis,
  covariates        = confounders.name,
  scope             = "wide",
  n_min             = 60,
  max_per_covariate = 3,
  max_subgroups     = 10,
  digits            = 3
)

gt(dina_fr[, c("covariate", "direction", "threshold",
               "n_subgroup", "effect", "cut_expr")]) |>
  tab_header(
    title = "Standalone DINA Frontier",
    subtitle = "Per-covariate non-dominated (effect, N) candidates; effect on HR scale"
  ) |>
  fmt_number(columns = c(threshold, effect), decimals = 3) |>
  cols_label(
    covariate = "Covariate", direction = "Direction",
    threshold = "Threshold", n_subgroup = "N", effect = "HR",
    cut_expr = "Cut"
  ) |>
  cols_align(align = "left", columns = c(covariate, direction, cut_expr))
```

## DINA-Selected Subgroup (Depth-2 Conjunction)

`dina_subgroup(max_depth = 2)` ranks the full-resolution single-covariate
candidates together with the gridded distinct-covariate conjunctions and
returns whichever wins under `sg_focus`. A conjunction is reported only when
it strictly outranks every singleton.

```{r dina-standalone-select}
dina_sel2 <- dina_subgroup(
  fit                 = dina_fit,
  df                  = df.analysis,
  covariates          = confounders.name,
  m_diff              = m_diff_dina,
  n_min               = 60,
  direction           = "both",
  max_depth           = dina_max_depth,    # depth-2
  grid_probs          = dina_grid_probs,
  sg_focus            = fs_dina_sg_focus,
  selection_rule      = "neighborhood",
  effect_neighborhood = 0.10,
  alpha               = 0.05
)

if (isTRUE(dina_sel2$found)) {
  cat(sprintf("DINA-selected subgroup (depth %d): %s\n",
              dina_sel2$depth, dina_sel2$label))
  cat(sprintf("N = %d  |  mean tau-hat = %.3f\n",
              dina_sel2$n_subgroup, dina_sel2$mean_tau_hat))
} else {
  cat("Standalone DINA selected no subgroup meeting the criteria.\n")
}
```

When the selection is a conjunction, `covariate` / `direction` / `threshold`
are length-2 vectors -- one entry per component cut.

```{r dina-standalone-components, eval = isTRUE(dina_sel2$found)}
op_comp <- ifelse(dina_sel2$direction == "left", "<=", ">=")
gt(data.frame(
  Component = seq_along(dina_sel2$covariate),
  Covariate = dina_sel2$covariate,
  Direction = dina_sel2$direction,
  Operator  = op_comp,
  Threshold = dina_sel2$threshold,
  check.names = FALSE
)) |>
  tab_header(
    title = "Component Cuts of the DINA-Selected Subgroup",
    subtitle = sprintf("Depth %d: AND-conjunction membership = intersection of components",
                       dina_sel2$depth)
  ) |>
  fmt_number(columns = Threshold, decimals = 3) |>
  cols_align(align = "left", columns = c(Covariate, Direction, Operator))
```

## Standalone DINA Bootstrap (Depth-2 Stability)

`dina_subgroup_bootstrap(max_depth = 2)` resamples the DINA selection. For a
conjunction it reports the bootstrap **threshold CI per component covariate**
(the two thresholds are not on a common scale) and a **cut-set selection
frequency** keyed by a canonical, threshold-independent key
(e.g. `"grade3 >= & pgr <="`). The within-subgroup standard-model refit is
currently depth-1 only and is therefore disabled here.

```{r dina-standalone-bootstrap, eval = isTRUE(dina_sel2$found)}
t0 <- proc.time()

dina_boot2 <- dina_subgroup_bootstrap(
  df          = df.analysis,
  outcome     = outcome.name,
  treatment   = treat.name,
  covariates  = confounders.name,
  family      = "cox",
  status      = event.name,
  m_diff      = m_diff_dina,
  n_min       = 60L,
  max_depth   = dina_max_depth,    # follows dina_sel2 (depth-2)
  grid_probs  = dina_grid_probs,
  sg_focus    = fs_dina_sg_focus,
  n_boot      = NB,
  seed        = dina_seed,
  fit         = dina_fit,          # pin the point estimate to dina_sel2
  sg          = dina_sel2,
  refit       = FALSE              # depth-2 refit deferred
)

timings$dina_boot2 <- (proc.time() - t0)["elapsed"]

# Full text summary (effect CI, per-component threshold CIs, cut-set frequency)
print(dina_boot2)
```

Per-component threshold stability, rendered as a table:

```{r dina-boot2-threshold-ci, eval = isTRUE(dina_sel2$found)}
tci <- dina_boot2$threshold_ci
if (is.matrix(tci)) {
  tci_df <- data.frame(Covariate = rownames(tci),
                       Lower = tci[, "lower"], Upper = tci[, "upper"],
                       check.names = FALSE)
} else {
  tci_df <- data.frame(Covariate = dina_sel2$covariate[1],
                       Lower = tci[["lower"]], Upper = tci[["upper"]],
                       check.names = FALSE)
}
gt(tci_df) |>
  tab_header(
    title = sprintf("Bootstrap Threshold CIs (%.0f%% percentile)",
                    100 * (1 - dina_boot2$alpha)),
    subtitle = sprintf("Restricted to the %d resamples matching the modal cut-set",
                       dina_boot2$n_modal_match)
  ) |>
  fmt_number(columns = c(Lower, Upper), decimals = 3) |>
  cols_align(align = "left", columns = Covariate)
```

Cut-set selection frequency across resamples (the depth-aware generalization
of the single-covariate frequency table):

```{r dina-boot2-selfreq, eval = isTRUE(dina_sel2$found)}
sf <- dina_boot2$selection_frequency
if (!is.null(sf) && length(sf) > 0) {
  sf_df <- as.data.frame(sf, stringsAsFactors = FALSE)
  names(sf_df) <- c("Cut-set", "Count")
  sf_df <- sf_df[order(-sf_df$Count), , drop = FALSE]
  sf_df$Percent <- round(100 * sf_df$Count / dina_boot2$n_boot_found, 1)
  gt(utils::head(sf_df, 8)) |>
    tab_header(title = "Cut-set Selection Frequency Across Bootstraps") |>
    cols_align(align = "left", columns = `Cut-set`)
} else {
  cat("No found-resample selections to tabulate.\n")
}
```


# (C) ForestSearch with `subgroup_method = "dina"` (Depth-2)

ForestSearch delegates *selection* to the depth-2 `dina_subgroup()`,
bypassing GRF, LASSO and the consistency search. The returned object is a
standard `forestsearch` result, so the existing bootstrap machinery applies
unchanged -- the bias correction reflects the depth-2 DINA selection process.

```{r fs-dina-select}
t0 <- proc.time()

fs_dina2 <- forestsearch(
  df.analysis,
  outcome.name        = outcome.name,
  event.name          = event.name,
  treat.name          = treat.name,
  id.name             = id.name,
  confounders.name    = confounders.name,
  df.predict          = df.analysis,        # original data for bias correction
  is.RCT              = TRUE,
  seedit              = 8316951,
  # --- DINA-selection mode (depth-2) --------------------------------------
  subgroup_method     = "dina",
  dina_args           = list(
    seed = dina_seed, n_folds = dina_nfolds,
    max_depth  = dina_max_depth,            # depth-2 selection
    grid_probs = dina_grid_probs
  ),
  # --- Selection criteria (passed through to dina_subgroup) ---------------
  sg_focus            = fs_dina_sg_focus,
  selection_rule      = "neighborhood",
  effect_neighborhood = 0.10,
  n.min               = 60,
  hr.threshold        = fs_dina_hr_threshold,  # -> log-HR harm floor
  details             = TRUE
)

timings$fs_dina_select <- (proc.time() - t0)["elapsed"]

cat("DINA-selected subgroup (H):",
    paste(fs_dina2$sg.harm, collapse = " & "), "\n")
if (!is.null(fs_dina2$sg.harm)) {
  cat("Number of component factors:", length(fs_dina2$sg.harm), "\n")
  cat("Subgroup size:", sum(fs_dina2$df.est$treat.recommend == 0),
      sprintf("(%.1f%% of ITT)\n",
              100 * mean(fs_dina2$df.est$treat.recommend == 0)))
}
```

## Bootstrap Bias-Corrected Cox Estimates

Each replicate re-fits DINA and re-runs the depth-2 `dina_subgroup()` from
scratch, so the correction reflects the depth-2 selection. The bias-corrected
estimate is membership-based, hence identical in form whether the selected
subgroup is a single cut or a conjunction.

```{r fs-dina-bootstrap, eval = !is.null(fs_dina2$sg.harm)}
t0 <- proc.time()

fs_dina2_bc <- forestsearch_bootstrap_dofuture(
  fs.est     = fs_dina2,
  nb_boots   = NB,
  show_three = FALSE,
  details    = FALSE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$fs_dina_bootstrap <- (proc.time() - t0)["elapsed"]

cat("DINA-mode depth-2 bootstrap completed in",
    round(timings$fs_dina_bootstrap / 60, 1), "minutes\n")
```

```{r fs-dina-no-subgroup, eval = is.null(fs_dina2$sg.harm)}
timings$fs_dina_bootstrap <- 0
cat("DINA (subgroup_method = \"dina\", max_depth = 2) identified no subgroup",
    "at hr.threshold =", fs_dina_hr_threshold, "-- bootstrap skipped.\n",
    "Lower the harm floor (raise the benefit) or widen grid_probs if a",
    "bootstrap demonstration is desired.\n")
```

```{r fs-dina-bootstrap-summary, eval = !is.null(fs_dina2$sg.harm)}
dina2_summaries <- summarize_bootstrap_results(
  sgharm       = fs_dina2$sg.harm,
  boot_results = fs_dina2_bc,
  create_plots = FALSE,
  est.scale    = "hr"
)

dina2_summaries$table
```

## Depth-2 Stability Diagnostics

In DINA-selection mode the per-resample stability diagnostics
(`K_sg`, `N_sg`, `M.1`, `M.2`, ...) are populated directly from the selected
subgroup, so `summarize_bootstrap_subgroups()` is non-blank and depth-aware:
a depth-2 selection fills both `M.1` and `M.2`, and conjunctions appear in the
agreement table as `factor & factor`.

```{r fs-dina-Ksg, eval = !is.null(fs_dina2$sg.harm)}
boot_dt <- data.table::as.data.table(fs_dina2_bc$results)
k_tab <- table(K_sg = boot_dt$K_sg, useNA = "ifany")
gt(as.data.frame(k_tab, stringsAsFactors = FALSE)) |>
  tab_header(
    title = "Per-resample Subgroup Depth (K_sg)",
    subtitle = "Number of component factors re-selected per bootstrap resample"
  ) |>
  cols_label(K_sg = "K_sg (factors)", Freq = "Resamples") |>
  cols_align(align = "left")
```

```{r fs-dina-subgroup-stability, eval = !is.null(fs_dina2$sg.harm)}
dina2_sbs <- summarize_bootstrap_subgroups(
  results     = fs_dina2_bc$results,
  nb_boots    = NB,
  original_sg = fs_dina2$sg.harm,
  maxk        = 2L
)

# Factor-position frequencies (M.1, M.2) -- previously blank in DINA mode.
if (!is.null(dina2_sbs$factor_freq) && nrow(dina2_sbs$factor_freq) > 0) {
  gt(utils::head(as.data.frame(dina2_sbs$factor_freq), 12)) |>
    tab_header(
      title = "Factor-Position Frequencies Across Bootstraps",
      subtitle = "DINA-mode depth-2 diagnostics (M.1 / M.2 now populated)"
    ) |>
    cols_align(align = "left")
} else {
  cat("No factor-frequency rows to display.\n")
}
```

```{r fs-dina-subgroup-agreement, eval = !is.null(fs_dina2$sg.harm)}
# Subgroup-level agreement: depth-2 conjunctions appear as "factor & factor".
if (!is.null(dina2_sbs$agreement) && nrow(dina2_sbs$agreement) > 0) {
  gt(utils::head(as.data.frame(dina2_sbs$agreement), 10)) |>
    tab_header(
      title = "Bootstrap Subgroup Agreement",
      subtitle = "Canonical cut-sets re-selected across resamples (sorted, collapsed)"
    ) |>
    cols_align(align = "left")
} else {
  cat("No subgroup-agreement rows to display.\n")
}
```

::: {.callout-note}
## Depth-2 notes

* **Where conjunctions can arise.** In modes (B) and (C) `dina_subgroup()`
  *selects* the conjunction directly. In mode (A) DINA only *screens*: it
  injects the conjunction's component cuts, and the consistency search decides
  the final subgroup (which may differ from DINA's conjunction).
* **`max_depth = 1L` is exact legacy behaviour.** Depth-2 only changes the
  result when a conjunction strictly outranks every single-covariate
  candidate under `sg_focus`.
* **GLM outcomes.** The same depth-2 path applies to binary (`OR`, `RD`),
  count (`IRR`, `IRD`) and continuous (`MD`) outcomes via
  `outcome_type` / `effect_measure`; the per-resample subgroup effect is
  reported on the ratio scale for `OR`/`RR`/`IRR` and on the identity scale
  for `MD`/`RD`/`IRD`.
* **Refit.** The within-subgroup standard-model refit in
  `dina_subgroup_bootstrap()` is currently depth-1 only; it is auto-disabled
  (with a message) for conjunctions.
:::


# Computational Timing

```{r timing-table}
timing_df <- data.frame(
  Step = c("Standalone DINA fit",
           "Standalone DINA depth-2 bootstrap",
           "(A) ForestSearch + DINA screening",
           "(C) ForestSearch subgroup_method = dina",
           "(C) DINA-mode depth-2 bootstrap"),
  Minutes = c(
    timings$dina_standalone,
    if (!is.null(timings$dina_boot2)) timings$dina_boot2 else NA_real_,
    timings$fs_dina_screen,
    timings$fs_dina_select,
    if (!is.null(timings$fs_dina_bootstrap)) timings$fs_dina_bootstrap else NA_real_
  ) / 60,
  check.names = FALSE
)

gt(timing_df) |>
  tab_header(title = "Computational Timing",
             subtitle = sprintf("Total: %.1f minutes",
                                 (proc.time() - t_vignette_start)["elapsed"] / 60)) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  cols_align(align = "left", columns = Step)
```


# Session Information

```{r session-info}
sessionInfo()
```
