---
title: "DINA Cox Survival Analysis — GBSG"
subtitle: "DINA analog to the ForestSearch GBSG Cox analysis"
author: "forestsearch v0.2.0"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

timings <- list()
t_doc_start <- proc.time()
```

# Introduction

This vignette is the DINA analog of the *Cox* branch of
`gbsg_survival_cox_poisson.qmd`. It applies the **DINA**
(Doubly-robust Iterative Neyman-Orthogonal) estimator to the
German Breast Cancer Study Group trial via the partial-likelihood
Cox extension.

The full inference toolkit applied here:

* Cross-fit point estimation via `dina()` (sandwich variance).
* Threshold-search harm subgroup identification via
  `dina_subgroup()` with sandwich-variance Wald CI.
* Selection-adjusted bootstrap percentile CI on the fixed-subgroup
  linear functional via `dina_subgroup_bootstrap()`.
* Stability diagnostics: selection-frequency table,
  `n_subgroup_ci`, `threshold_ci`.

The intent is to mirror the FS doc's Cox-analysis flow on the same
data with the same parameter conventions, *not* to repeat the FS
doc's downstream artifacts (CV, KM by subgroup, forest plots). Those
exist in the FS doc; this document focuses on the DINA inference
pipeline.

## Deferred: Poisson + offset analog

The FS doc also runs a **Poisson + offset** branch as the natural
exponential-family approximation to Cox. The analogous DINA branch
requires offset support in `dina_fit()`, which is not yet
implemented — `dina_fit()`'s Poisson family treats `Y` as raw counts
without person-time adjustment. Extending `dina_fit()` to accept an
`offset = ` argument is a well-scoped enhancement (offset slicing
through cross-fitting, baseline GLM fits, and step-2; the math is
clean because the treatment-effect contrast is invariant under a
shared offset). It's deferred to a future session.

The ITT-level Cox-vs-Poisson sanity check below uses plain
`survival::coxph()` and `stats::glm()` (no DINA) and is still
informative as a *population-level* diagnostic of the
proportional-hazards approximation, even though the subgroup-level
DINA Poisson + offset analysis awaits the extension.

# Setup

## Load Required Packages

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(doFuture)

theme_set(theme_minimal(base_size = 12))
```

## Analysis Parameters

```{r analysis-params}
#| code-fold: false

# ─── Shared parameters ───────────────────────────────────────────────────
n_cores         <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))
confounders     <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
id_col          <- "id"
treat_col       <- "hormon"

# ─── DINA Cox outcome columns ────────────────────────────────────────────
cox_outcome     <- "time_months"
cox_status      <- "status"

# ─── DINA tuning ─────────────────────────────────────────────────────────
dina_n_folds    <- 2L
dina_seed       <- 8316951L

# ─── Harm threshold (natural-parameter scale: log-HR for Cox) ────────────
# DINA's Cox target is log-HR (the natural parameter of the partial-
# likelihood model), NOT risk difference.  So m_diff is on the log-HR
# scale.  First-pass exploration uses 0.0 -- "any harm signal", HR >= 1.0.
# For clinically-meaningful harm matching the FS hr.threshold = 1.25,
# use log(1.25) approx 0.2231 in a follow-up run.
m_diff          <- 0.0
n_min           <- 60L               # matches FS n.min

# ─── Bootstrap ───────────────────────────────────────────────────────────
NB              <- 500L              # bootstrap iterations (use 1000+ for production)
boot_seed       <- 8316951L

# ─── Confidence level ────────────────────────────────────────────────────
alpha           <- 0.05
```

# Data: German Breast Cancer Study Group Trial

```{r data-setup}
df.analysis <- gbsg

df.analysis <- within(df.analysis, {
  id          <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3      <- as.integer(ifelse(grade == "3", 1, 0))
})

cat("Sample size:    ", nrow(df.analysis), "\n")
cat("Events:         ", sum(df.analysis[[cox_status]]),
    sprintf(" (%.1f%%)\n",
            100 * mean(df.analysis[[cox_status]])))
cat("Covariates:     ", paste(confounders, collapse = ", "), "\n")
cat("Cox outcome:    ", cox_outcome, "(+ status: ", cox_status, ")\n", sep = "")
```

# ITT-Level Cox vs. Poisson + Offset Sanity Check

A *population-level* sanity check that the Poisson + offset
approximation lines up with Cox at the ITT level on this dataset.
This uses plain `survival::coxph()` and `stats::glm()` (no DINA),
so it does not depend on the deferred DINA offset extension. The
purpose is to flag any gross mismatch at the population level that
would already cast doubt on a downstream Cox-vs-Poisson alignment
analysis at the subgroup level.

```{r itt-sanity}
df.analysis$time_months_log <- log(pmax(df.analysis$time_months, 1e-6))

cox_itt <- survival::coxph(
  survival::Surv(time_months, status) ~ hormon,
  data = df.analysis
)
pois_itt <- stats::glm(
  status ~ hormon,
  family = stats::poisson(),
  offset = df.analysis$time_months_log,
  data = df.analysis
)

itt_table <- data.table(
  Quantity = c("Cox  HR",
               "Poisson IRR",
               "Absolute difference"),
  Value    = c(
    sprintf("%.4f", exp(stats::coef(cox_itt))),
    sprintf("%.4f", exp(stats::coef(pois_itt)[["hormon"]])),
    sprintf("%.4f", abs(exp(stats::coef(cox_itt)) -
                        exp(stats::coef(pois_itt)[["hormon"]])))
  )
)

itt_table |>
  gt::gt() |>
  gt::tab_header(
    title    = "ITT-level Cox vs. Poisson + offset",
    subtitle = "Population-level proportional-hazards sanity check"
  )
```

# DINA Cox Analysis

## Fit cross-fit DINA Cox model

```{r dina-cox-fit}
t0 <- proc.time()

fit_dina_cox <- dina(
  df         = df.analysis,
  outcome    = cox_outcome,
  treatment  = treat_col,
  covariates = confounders,
  family     = "cox",
  status     = cox_status,
  n_folds    = dina_n_folds,
  seed       = dina_seed
)

timings$dina_cox_fit <- (proc.time() - t0)[["elapsed"]]
cat("DINA Cox fit completed in",
    round(timings$dina_cox_fit, 1), "seconds\n")
```

## Subgroup search (sandwich Wald CI)

### Scale: log-HR for Cox

DINA targets the natural-parameter contrast
$\tau(x) = \beta_0 + x^\top \beta_{1:d}$, which lies on the
natural-parameter scale of the family. The interpretation
therefore differs across families:

| Family | Natural-parameter scale | Interpretation of $\tau(x)$ |
|---|---|---|
| Gaussian | mean | mean difference (i.e., "risk difference" in the colloquial sense) |
| Binomial | logit | log odds ratio |
| Poisson | log rate | log incidence rate ratio |
| Cox | log hazard | log hazard ratio |

For the Cox family, the natural parameter is the log hazard (the
linear predictor of the partial-likelihood model), so `m_diff` is
on the **log-HR scale** — not a risk difference. The "DIfference in
NAtural parameters" name is precise: it's a difference of *natural
parameters*, and what those parameters mean depends on the family's
GLM link.

### Threshold: 0.0 as a permissive first pass

`m_diff = 0.0` asks for "the largest subgroup with mean
log-HR $\geq 0$", equivalently HR $\geq 1.0$ — any indication of
harm at all, regardless of magnitude. For clinically-meaningful
harm matching the FS `hr.threshold = 1.25`, use
$\log(1.25) \approx 0.223$ in a follow-up run.

Two practical notes about what to expect at this permissive
threshold:

1. **The discovered subgroup may differ from the FS finding.** A
   more permissive threshold often picks up *larger* subgroups
   (since more of the population satisfies the weaker harm
   condition), and may also surface a *different covariate* if
   small effects accumulate across a big subgroup. The FS Cox
   finding (Estrogen $\leq 0$ with HR $\approx 1.5$) should still
   be visible if it's real, but the threshold-search optimization
   may prefer a different cut at the lower threshold.

2. **Selection stability may drop.** With $m_{\text{diff}}$ near
   zero, multiple covariates can satisfy the harm condition on
   bootstrap replicates, so `selection_frequency` may be more
   diffuse and `n_modal_match` smaller than at a stricter
   threshold. The `effect_ci` interpretation stays clean — it is a
   percentile CI on a fixed-subgroup linear functional regardless
   of threshold — but the stability diagnostics
   (`n_subgroup_ci`, `threshold_ci`) may look noisier.

Both behaviors are expected and informative; they are what makes
$m_{\text{diff}} = 0.0$ a useful first-pass exploration.

### Direction

`direction = "both"` allows either tail of any covariate. The FS
Cox analysis discovered Estrogen $\leq 0$ (a *left*-direction
subgroup on `er`), so we expect DINA to surface something in that
neighborhood.

```{r dina-cox-subgroup}
sg_cox <- dina_subgroup(
  fit        = fit_dina_cox,
  df         = df.analysis,
  covariates = confounders,
  m_diff     = m_diff,
  n_min      = n_min,
  direction  = "both",
  alpha      = alpha
)

print(sg_cox)

# Gate the downstream bootstrap and summary on whether a subgroup was
# found.  The bootstrap targets the fixed-subgroup linear functional
# a*' beta with a* derived from the original-data subgroup, so it is
# undefined when no subgroup is identified.
run_boot <- isTRUE(sg_cox$found)
timings$dina_cox_boot <- NA_real_   # placeholder; overwritten if bootstrap runs

if (run_boot) {
  cat("\nSubgroup identified -- bootstrap inference will run below.\n")
} else {
  cat("\nNo subgroup met the criteria at m_diff =", m_diff,
      ", n_min =", n_min, ".\n",
      "Bootstrap inference will be skipped.\n", sep = "")
}
```

## Selection-adjusted bootstrap inference

The bootstrap targets the fixed-subgroup linear functional
$a^{*\top} \beta$, where $a^* = (1, \bar{x}_{S^*})$ is fixed from
the original-data subgroup $S^*$. If `dina_subgroup()` did not
identify a subgroup, $a^*$ is undefined and the bootstrap is
skipped; the chunks below are evaluated conditionally on
`run_boot`.

```{r dina-cox-bootstrap}
#| eval: !expr run_boot
t0 <- proc.time()

future::plan(future::multisession, workers = n_cores)

bs_cox <- dina_subgroup_bootstrap(
  df         = df.analysis,
  outcome    = cox_outcome,
  treatment  = treat_col,
  covariates = confounders,
  family     = "cox",
  status     = cox_status,
  m_diff     = m_diff,
  n_min      = n_min,
  direction  = "both",
  alpha      = alpha,
  n_boot     = NB,
  parallel   = "boots",
  n_folds    = dina_n_folds,
  seed       = boot_seed
)

future::plan("sequential")
gc(verbose = FALSE)

timings$dina_cox_boot <- (proc.time() - t0)[["elapsed"]]
cat("DINA Cox bootstrap completed in",
    round(timings$dina_cox_boot / 60, 2), "minutes\n")
```

```{r dina-cox-bootstrap-print}
#| eval: !expr run_boot
print(bs_cox)
```

## Summary table

```{r dina-cox-summary-helpers}
`%||%` <- function(a, b) if (is.null(a)) b else a

fmt_signature <- function(sg) {
  if (!isTRUE(sg$found)) return("—")
  cmp <- if (sg$direction == "left") "<=" else ">="
  sprintf("%s %s %.3f", sg$covariate, cmp, sg$threshold)
}

fmt_ratio_ci <- function(ci, digits = 3L) {
  if (any(!is.finite(ci))) return("—")
  sprintf("[%.*f, %.*f]",
          digits, exp(ci[["lower"]]),
          digits, exp(ci[["upper"]]))
}

fmt_log_ci <- function(ci, digits = 3L) {
  if (any(!is.finite(ci))) return("—")
  sprintf("[%.*f, %.*f]",
          digits, ci[["lower"]],
          digits, ci[["upper"]])
}

fmt_int_ci <- function(ci) {
  if (any(!is.finite(ci))) return("—")
  sprintf("[%.0f, %.0f]", ci[["lower"]], ci[["upper"]])
}

fmt_top_selections <- function(bs, k = 3L) {
  if (is.null(bs$selection_frequency) || bs$n_boot_found == 0L) return("—")
  sel_long <- as.data.frame(bs$selection_frequency,
                            stringsAsFactors = FALSE)
  sel_long <- sel_long[sel_long$Freq > 0L, , drop = FALSE]
  sel_long <- sel_long[order(-sel_long$Freq), , drop = FALSE]
  top <- head(sel_long, k)
  paste(sprintf("%s (%s): %d/%d (%.0f%%)",
                top$covariate, top$direction, top$Freq,
                bs$n_boot_found,
                100 * top$Freq / bs$n_boot_found),
        collapse = "; ")
}

# Defensive Wald-CI extractor from a dina_subgroup result.
# Tries several plausible field-name conventions before falling back to NA.
extract_wald_ci <- function(sg) {
  na_ci <- c(lower = NA_real_, upper = NA_real_)
  if (!isTRUE(sg$found)) return(na_ci)
  vec_candidates <- c("mean_tau_ci", "wald_ci", "ci")
  for (cand in vec_candidates) {
    if (cand %in% names(sg) && length(sg[[cand]]) == 2L) {
      v <- as.numeric(sg[[cand]])
      return(c(lower = v[1L], upper = v[2L]))
    }
  }
  pair_candidates <- list(
    c("mean_tau_ci_lower", "mean_tau_ci_upper"),
    c("mean_tau_ci_lo",    "mean_tau_ci_hi"),
    c("ci_lower",          "ci_upper"),
    c("lower",             "upper")
  )
  for (cand in pair_candidates) {
    if (all(cand %in% names(sg))) {
      return(c(lower = as.numeric(sg[[cand[1L]]]),
               upper = as.numeric(sg[[cand[2L]]])))
    }
  }
  na_ci
}
```

```{r dina-cox-summary-table}
#| eval: !expr run_boot
pt_cox   <- bs_cox$point
wald_cox <- extract_wald_ci(pt_cox)

summary_tbl <- data.table(
  Field = c(
    "Discovered subgroup signature",
    "n_subgroup",
    "Bootstrap found / total",
    "Bootstrap failed",
    "Modal-match iters / found",
    "Point: mean tau-hat (log-HR)",
    "Point: exp(mean tau-hat) (HR)",
    "Sandwich Wald CI (log-HR)",
    "Sandwich Wald CI (HR)",
    "Bootstrap effect CI (log-HR)",
    "Bootstrap effect CI (HR)",
    "n_subgroup CI (modal-match)",
    "Threshold CI (modal-match)",
    "Top (covariate, direction) selections"
  ),
  Value = c(
    fmt_signature(pt_cox),
    sprintf("%d / %d", pt_cox$n_subgroup %||% NA, pt_cox$n_total %||% NA),
    sprintf("%d / %d", bs_cox$n_boot_found, bs_cox$n_boot),
    sprintf("%d", bs_cox$n_boot_failed),
    if (!is.na(bs_cox$n_modal_match))
      sprintf("%d / %d", bs_cox$n_modal_match, bs_cox$n_boot_found)
    else "—",
    sprintf("%.3f", pt_cox$mean_tau_hat %||% NA_real_),
    sprintf("%.3f", exp(pt_cox$mean_tau_hat %||% NA_real_)),
    fmt_log_ci(wald_cox),
    fmt_ratio_ci(wald_cox),
    fmt_log_ci(bs_cox$effect_ci),
    fmt_ratio_ci(bs_cox$effect_ci),
    fmt_int_ci(bs_cox$n_subgroup_ci),
    fmt_log_ci(bs_cox$threshold_ci),
    fmt_top_selections(bs_cox)
  )
)

summary_tbl |>
  gt::gt() |>
  gt::tab_header(
    title    = gt::md("**DINA Cox analysis — summary**"),
    subtitle = sprintf("m_diff = %.3f (permissive first pass; log-HR scale);  n_boot = %d;  %d cross-fitting folds",
                       m_diff, NB, dina_n_folds)
  ) |>
  gt::tab_source_note(
    source_note = gt::md(
      "*Sandwich Wald CI is from `dina_subgroup()` on the original data.
       Bootstrap effect CI is the percentile CI on `a*'beta_b` with `a*`
       fixed from the original-data subgroup, via
       `dina_subgroup_bootstrap()`.  If Sandwich CI rows display '—', the
       field-name conventions in `extract_wald_ci()` need updating to match
       this DINA build.*"
    )
  )
```

```{r dina-cox-no-subgroup}
#| eval: !expr !run_boot
no_sg_tbl <- data.table(
  Field = c(
    "Subgroup found",
    "m_diff (log-HR scale)",
    "exp(m_diff) (HR scale)",
    "n_min",
    "direction",
    "Covariates searched",
    "n",
    "Note"
  ),
  Value = c(
    "No",
    sprintf("%.3f", m_diff),
    sprintf("%.3f", exp(m_diff)),
    sprintf("%d", n_min),
    "both",
    paste(confounders, collapse = ", "),
    sprintf("%d (events: %d)",
            nrow(df.analysis),
            sum(df.analysis[[cox_status]])),
    paste0("Bootstrap inference skipped because the fixed-subgroup ",
           "linear functional is undefined when no subgroup is ",
           "identified.")
  )
)

no_sg_tbl |>
  gt::gt() |>
  gt::tab_header(
    title    = gt::md("**DINA Cox analysis -- no subgroup identified**"),
    subtitle = sprintf("m_diff = %.3f (log-HR);  n_min = %d;  cross-fitting folds = %d",
                       m_diff, n_min, dina_n_folds)
  ) |>
  gt::tab_source_note(
    source_note = gt::md(
      "*To get a stable result on data where no harm subgroup exists,
       consider (i) loosening the search constraints (smaller `n_min`,
       different `direction`), (ii) checking the ITT-level diagnostic
       above for whether the overall effect direction is consistent
       with harm at all, or (iii) re-running on a different covariate
       set.*"
    )
  )
```

# Future Work: DINA Poisson + Offset Extension

The FS doc's Cox-vs-Poisson alignment analysis is the natural
next deliverable here. It requires:

1. **Extend `dina_fit()` with an `offset` argument.** Pass the
   offset vector to `.dina_one_pass()`, slice it by train/test
   indices, and forward to:
   * `.dina_fit_baselines_glm()` — include the offset in the
     within-arm `stats::glm()` calls so the returned
     `eta0_fn` / `eta1_fn` predict on the offset-inclusive
     linear-predictor scale.
   * `.dina_step2_glm()` — already passes `offset = nu_a$nu`
     to `stats::glm.fit()`; if step 1 incorporates the offset
     into `nu`, step 2 needs no further change.
2. **Forward `offset` through the wrappers** `dina()`,
   `dina_bagged()`, and `dina_subgroup_bootstrap()` as a
   pass-through argument.
3. **Add a Poisson + offset family** to the
   `dina_subgroup_bootstrap_smoke_test.qmd` rotation. Synthetic
   data with person-time exposure, log offset.
4. **Re-render this document with the Poisson + offset branch
   restored** and a Part III Cox-Poisson alignment section
   mirroring the FS doc.

The treatment-effect contrast $\tau(x) = \eta_1(x) - \eta_0(x)$ is
invariant under a shared offset added to both arms, so DINA's
estimand is unchanged; only the nuisance fits and step-2 likelihood
need the offset. Bias-correction would not be needed beyond what
cross-fitting already handles.

# Timings

```{r timings-table}
fmt_secs <- function(x) {
  if (is.null(x) || !is.finite(x)) "—" else sprintf("%.1f", x)
}

tim_df <- data.table(
  Step       = c("DINA Cox fit",
                 "DINA Cox bootstrap",
                 "Total"),
  `Time (s)` = c(
    fmt_secs(timings$dina_cox_fit),
    fmt_secs(timings$dina_cox_boot),
    fmt_secs((proc.time() - t_doc_start)[["elapsed"]])
  )
)

tim_df |>
  gt::gt() |>
  gt::tab_header(title = "Timing summary")
```

# Session Information

```{r session-info}
sessionInfo()
```
