---
title: "Exploratory Subgroup Identification with ForestSearch"
subtitle: "German Breast Cancer Study Group (GBSG) Analysis"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
execute:
  warning: true
  message: true
---

> **Method-general analysis.** The selection engine for the main analysis is set
> once via `fs_subgroup_method` in the setup chunk — `"consistency"` (default),
> `"dina"`, or `"grf"` (with `fs_grf_selection` = `"tree"`/`"frontier"`). The
> headline fit and every downstream step — full bootstrap, Tier-2 de-biased gate,
> cross-validation, leave-one-out, and the Kaplan–Meier band — inherit that engine
> from the single fitted object, so the whole analysis re-renders under a
> different method by changing one knob. The comparative DINA-screening,
> standalone-DINA, and GRF sections that follow are fixed illustrations of the
> individual engines and are unaffected by `fs_subgroup_method`.

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# Initialize timing
timings <- list()
t_vignette_start <- proc.time()
```

## Parallel Processing Configuration

ForestSearch supports parallel processing for computationally intensive
operations (bootstrap, cross-validation).

```{r parallel-setup}
# Detect available cores (limited to 2 cores for CRAN checks)
n_cores_total <- parallel::detectCores()
n_cores <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))
cat("Using", n_cores, "of", n_cores_total, "total cores for parallel processing")

# Number of bootstrap iterations
# Use 500-2000 for production; reduced here for vignette
NB <- 500

# ---------------------------------------------------------------------------
# Validation toggles (the slowest parts of the document).  Independent:
# run_cv  controls K-Fold cross-validation; run_loo controls leave-one-out
# (N-fold / OOB).  Either can be turned off without affecting the other.
# ---------------------------------------------------------------------------
run_cv  <- FALSE    # K-Fold cross-validation (kfold-cv + diagnostics)
run_loo <- FALSE    # Leave-one-out / N-fold OOB (oob-cv)
# 10-fold CV with multiple iterations
# Use Ksims >= 50 for production
Ksims <- 50


# Pre-seed CV/LOO timings so the Computational Timing table is well-formed
# whether or not these run (skipped analyses report NA).
timings$kfold <- NA_real_
timings$oob   <- NA_real_

# ---------------------------------------------------------------------------
# Per-analysis identification parameters (tune each analysis independently).
#   sg_focus      : which candidate subgroup the selection step chooses
#   hr_threshold  : harm/benefit floor on the HR scale
#   hr_consistency: minimum consistency rate (consistency search only)
# ---------------------------------------------------------------------------
# Main ForestSearch (consistency search)
fs_sg_focus              <- "maxSG"
fs_hr_threshold          <- 1.25
fs_hr_consistency        <- 1.0
fs_max_subgroups_search  <- 20
fs_consistency_method    <- "resample"
# Tier-2 de-biased gate: a fast multiplier-bootstrap approximation of the
# bootstrap bias-correction, computed inside the fit and compared to the full
# bootstrap below.  It runs for whichever selection engine fs_subgroup_method
# names, de-biasing against that engine's own candidate family (it re-derives
# the per-candidate Cox dfbeta itself).
run_debias_gate <- TRUE     # add debias_gate = TRUE to the forestsearch call
gate_draws      <- 2000L    # multiplier-bootstrap draws for the gate

# ---------------------------------------------------------------------------
# Selection engine for the MAIN analysis.  The headline fit, its bootstrap,
# the de-biased gate, CV/LOO, and the KM plot all flow from this single fit,
# so flipping subgroup_method re-runs the entire main analysis under that
# engine (each downstream step inherits the method from the fitted object):
#   "consistency" - the splitting-consistency search (default);
#   "dina"        - DINA performs selection (consistency search bypassed);
#   "grf"         - GRF performs selection (policy tree or DR-frontier).
# The Tier-2 gate de-biases against whichever engine's OWN candidate family,
# with re-selection matching that engine's selection rule.
# ---------------------------------------------------------------------------
fs_subgroup_method <- "consistency"  # "consistency" | "dina" | "grf"
fs_grf_selection   <- "frontier"      # only if fs_subgroup_method == "grf": "tree" | "frontier"
fs_grf_depth       <- 2L              # GRF depth (screening + GRF selection)
fs_dmin_grf        <- 4               # GRF DR-score screening floor
fs_dina_args       <- list()          # extra args forwarded to DINA selection
# Tier-2 gate (main analysis): the gate's re-selection family for "dina"/"grf" is
# restricted to candidates competitive on the method's NATIVE statistic (DINA
# tau-hat / GRF DR-score).  NULL = use the recommended default, which matches the
# band to effect_neighborhood automatically (the band the engine selects within);
# give a number in (0, 1) to set the band explicitly, or >= 1 to de-bias against
# the full family.  Ignored for "consistency" (its maxcons re-selection already
# matches the search).
fs_family_native_nbhd <- NULL

# (A) ForestSearch + DINA screening (consistency search selects; DINA only
#     contributes candidate cuts)
dina_screen_sg_focus             <- "effMaxSG"
dina_screen_hr_threshold         <- 0.9
dina_screen_hr_consistency       <- 0.8
dina_screen_max_subgroups_search <- 30
# Screening cut set: FALSE => full DINA frontier; TRUE => only the single
# dina_subgroup()-selected cut (uses dina_screen_sg_focus / selection_rule).
dina_screen_selected_only        <- TRUE

# (C) ForestSearch subgroup_method = "dina" (DINA performs selection; the
#     consistency search is bypassed, so hr_consistency does not apply here).
#     The standalone DINA analysis (B) reuses these so (B) and (C) match.
fs_dina_sg_focus     <- "effMaxSG"
fs_dina_hr_threshold <- 0.90
# Tier-2 gate: the DINA re-selection family is restricted to candidates competitive
# on DINA's native tau-hat.  NULL = recommended default (matches effect_neighborhood
# automatically); set a number in (0, 1) for an explicit band, or >= 1 for the full
# family.  Note: this only tunes candidate competition -- the gate holds the forest
# fixed, so the full bootstrap stays the reference for DINA regardless.
fs_dina_family_native_nbhd <- NULL

# (D) ForestSearch subgroup_method = "grf" (GRF performs selection via a causal
#     forest policy tree or DR-score frontier; the consistency search, LASSO and
#     DINA are bypassed).  Dedicated demo: always runs GRF regardless of
#     fs_subgroup_method, mirroring the standalone DINA selection (C).
fs_grf_demo_selection     <- "frontier"  # "frontier" | "tree"
fs_grf_demo_depth         <- 2L          # GRF tree / DR-pair depth
fs_grf_demo_dmin          <- 0.0         # DR-score harm floor (0 = permissive default)
fs_grf_sg_focus           <- "effMaxSG"  # -> GRF frontier/tree selection rule
# Tier-2 gate: the GRF re-selection family is restricted to candidates competitive
# on GRF's native (RMST-based) DR-score.  NULL = recommended default (matches
# effect_neighborhood automatically); set a number in (0, 1) for an explicit band,
# or >= 1 for the full family.  Note: this only tunes candidate competition -- the
# gate holds the forest fixed, so the full bootstrap stays the reference for GRF
# regardless (a dominant DR-score winner can collapse the family to 1).
fs_grf_family_native_nbhd <- NULL
```



# Introduction

This vignette demonstrates the **ForestSearch** methodology for exploratory
subgroup identification in survival analysis, as described in
León et al. (2024) *Statistics in Medicine*.

## Motivation

In clinical trials, particularly oncology, subgroup analyses are essential for:

- Evaluating treatment effect consistency across patient populations
- Identifying subgroups where treatment may be detrimental (harm)
- Characterizing subgroups with enhanced benefit
- Informing regulatory decisions and clinical practice

While prespecified subgroups provide stronger evidence, important subgroups
based on patient characteristics may not be anticipated. ForestSearch provides
a principled approach to **exploratory** subgroup identification with proper
statistical inference.

## Methodology Overview

ForestSearch identifies subgroups through:

1. **Candidate factor selection**: Using LASSO and/or Generalized Random Forests (GRF)
2. **Exhaustive subgroup search**: Evaluating all combinations up to `maxk` factors
3. **Consistency-based selection**: Applying splitting consistency criteria
4. **Bootstrap bias correction**: Adjusting for selection-induced optimism
5. **Cross-validation**: Assessing algorithm stability

The key innovation is the **splitting consistency criterion**: a subgroup is
considered "consistent with harm" if, when randomly split 50/50 many times,
both halves consistently show hazard ratios ≥ 1.0 (for example if 1.0
represents a meaningful "harm threshold").


# Setup

## Load Required Packages

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(policytree)
library(doFuture)

# Optional packages for enhanced output
library(patchwork)
library(weightedsurv)

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```


# Data: German Breast Cancer Study Group Trial

## Study Background

The GBSG trial evaluated hormonal treatment (tamoxifen) versus chemotherapy
in node-positive breast cancer patients. Key characteristics:

- **Sample size**: N = 686
- **Outcome**: Recurrence-free survival time
- **Censoring rate**: ~56%
- **Treatment**: Hormonal therapy (tamoxifen) vs. chemotherapy

## Data Preparation

```{r data-setup}
# Load GBSG data (included in forestsearch package)
df.analysis <- gbsg

# Prepare analysis variables
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
  treat <- hormon
})

# Define variable roles
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name <- "status"
id.name <- "id"
treat.name <- "hormon"

# Display data structure
cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```

## Baseline Characteristics

```{r baseline-table}
create_summary_table(
  data = df.analysis,
  treat_var = treat.name,
  table_title = "GBSG Baseline Characteristics by Treatment Arm",
  vars_continuous = c("age", "nodes", "size", "er", "pgr"),
  vars_categorical = c("grade", "meno"),
  font_size = 12
)
```

## Kaplan-Meier Analysis (ITT Population)

```{r km-itt}
#| fig-width: 8
#| fig-height: 5

# Prepare counting process data for KM plot
dfcount <- df_counting(
  df = df.analysis,
  by.risk = 6,
  tte.name = outcome.name,
  event.name = event.name,
  treat.name = treat.name
)

# Plot with confidence intervals and log-rank test
plot_weighted_km(
  dfcount,
  conf.int = TRUE,
  show.logrank = TRUE,
  ymax = 1.05,
  xmed.fraction = 0.775,
  ymed.offset = 0.125
)
```

The ITT Cox hazard ratio estimate is approximately 0.69 (95% CI: 0.54, 0.89),
suggesting an overall benefit for hormonal therapy.


# Preliminary Analysis: Generalized Random Forests

Before running ForestSearch, we can use GRF to explore potential treatment
effect heterogeneity and identify candidate factors.

```{r grf-analysis}
t0 <- proc.time()

grf_est <- grf.subg.harm.survival(
  data = df.analysis,
  confounders.name = confounders.name,
  outcome.name = outcome.name,
  event.name = event.name,
  id.name = id.name,
  treat.name = treat.name,
  maxdepth = 2,
  n.min = 60,
  dmin.grf = 12,
  frac.tau = 0.60,
  details = TRUE,
  return_selected_cuts_only = FALSE
)

timings$grf <- (proc.time() - t0)["elapsed"]
```

```{r grf-trees}
#| fig-width: 10
#| fig-height: 4

# Display policy trees
# leaf1 = recommend control, leaf2 = recommend treatment
oldpar <- par(mfrow = c(1, 2))
plot(grf_est$tree1, leaf.labels = c("Control", "Treat"), main = "Depth 1")
plot(grf_est$tree2, leaf.labels = c("Control", "Treat"), main = "Depth 2")
par(oldpar)
```

GRF identifies estrogen receptor status (ER) as a key factor, with ER ≤ 0
suggesting potential harm from hormonal therapy.


# ForestSearch Analysis




## Running ForestSearch

The selection engine for this main analysis is set by `fs_subgroup_method` in
the setup chunk (`"consistency"`, `"dina"`, or `"grf"` with `fs_grf_selection`).
Everything downstream — the full bootstrap, the Tier-2 de-biased gate, the
cross-validation/leave-one-out diagnostics, and the Kaplan–Meier band — runs off
the single fitted object and therefore inherits that engine automatically, so
the entire analysis re-renders under a different method by changing one knob.
The consistency search (the default below) performs an exhaustive search over
candidate subgroup combinations with up to `maxk` factors; the parameters in the
table apply to that engine. Key parameters:

| Parameter | Value | Description |
|-----------|-------|-------------|
| `hr.threshold` | 1.25 | Minimum HR for consistency evaluation |
| `hr.consistency` | 1.0 | Minimum consistency rate for candidates |
| `pconsistency.threshold` | 0.90 | Required consistency for selection |
| `maxk` | 2 | Maximum factors in subgroup definition |
| `n.min` | 60 | Minimum subgroup sample size |
| `d0.min`, `d1.min` | 12 | Minimum events per treatment arm |

::: {.callout-note}
## Selection criteria (`sg_focus`)

`sg_focus` controls which candidate subgroup is chosen once consistency
has been evaluated. The example below uses `"maxSG"` (largest passing
subgroup). Other options:

* `"hr"` -- highest consistency, then largest effect.
* `"minSG"` -- smallest passing subgroup, ties broken by consistency.
* `"hrMaxSG"` -- among candidates whose effect is within
  `effect_neighborhood` (default 10%) of the maximum, pick the
  *largest* sample size.
* `"hrMinSG"` -- among candidates within `effect_neighborhood` of the
  maximum effect, pick the *smallest* sample size.

`"hrMaxSG"` and `"hrMinSG"` use a *lexicographic-with-tolerance* rule:
effect size is primary, but a small relative loss in effect is accepted
to optimize sample size. Setting `effect_neighborhood = 0` reduces these
to a strict max-effect filter. For GLM ratio measures (OR, IRR), the
neighborhood test is applied on the natural scale.
:::

```{r forestsearch-main}
#| fig-width: 10
#| fig-height: 7

t0 <- proc.time()

fs <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis,
  outcome.name              = outcome.name,
  event.name                = event.name,
  treat.name                = treat.name,
  id.name                   = id.name,
  confounders.name          = confounders.name,
  potentialOutcome.name     = NULL,           # default NULL
  flag_harm.name            = NULL,           # default NULL
  df.predict                = NULL,           # default NULL
  df.test                   = NULL,           # default NULL
  is.RCT                    = TRUE,           # default TRUE
  seedit                    = 8316951,        # default 8316951
  est.scale                 = "hr",           # default "hr"
  # --- Screening engines (candidate generation) ---------------------------
  use_lasso                 = TRUE,           # default TRUE
  use_grf                   = TRUE,           # default TRUE
  grf_res                   = NULL,           # default NULL
  grf_cuts                  = NULL,           # default NULL
  grf_depth                 = fs_grf_depth,   # set in setup chunk (default 2)
  dmin.grf                  = fs_dmin_grf,     # set in setup chunk (default 0.0)
  frac.tau                  = 0.80,           # default 0.8
  vi.grf.min                = -0.2,           # default -0.2
  return_selected_cuts_only = TRUE,           # default TRUE
  tune_grf                  = FALSE,          # default FALSE
  use_dina                  = FALSE,          # default FALSE
  dina_res                  = NULL,           # default NULL
  dina_cuts                 = NULL,           # default NULL
  dina_args                 = fs_dina_args,    # set in setup chunk (DINA selection)
  # --- Candidate-cut construction -----------------------------------------
  max_n_confounders         = 1000,           # default 1000
  conf_force                = NULL,           # default NULL
  defaultcut_names          = NULL,           # default NULL
  cut_type                  = "default",      # default "default"
  exclude_cuts              = NULL,           # default NULL
  replace_med_grf           = FALSE,          # default FALSE
  cont.cutoff               = 4,              # default 4
  conf.cont_medians         = NULL,           # default NULL
  conf.cont_medians_force   = NULL,           # default NULL
  conf.cont_jcuts           = NULL,           # default NULL
  # --- Selection engine ---------------------------------------------------
  subgroup_method           = fs_subgroup_method,  # set in setup chunk: consistency | dina | grf
  grf_selection             = fs_grf_selection,    # set in setup chunk (used when subgroup_method = "grf")
  sg_focus                  = fs_sg_focus,         # set in setup chunk
  selection_rule            = "neighborhood",  # default "neighborhood"
  effect_neighborhood       = 0.10,            # default 0.10
  # --- Tier-2 de-biased gate ----------------------------------------------
  debias_gate               = run_debias_gate,           # set in setup chunk
  debias_gate_args          = list(draws = gate_draws,
                                   family_native_neighborhood = fs_family_native_nbhd),
  # --- Subgroup constraints -----------------------------------------------
  n.min                     = 60,             # default 60
  d0.min                    = 10,             # default 10
  d1.min                    = 10,             # default 10
  maxk                      = 2,              # default 2
  max_subgroups_search      = fs_max_subgroups_search,  # set in setup chunk
  # --- Thresholds (survival / HR scale) -----------------------------------
  hr.threshold              = fs_hr_threshold,    # set in setup chunk
  hr.consistency            = fs_hr_consistency,  # set in setup chunk
  effect.threshold          = NULL,           # default NULL (alias of hr.threshold)
  consistency.threshold     = NULL,           # default NULL (alias of hr.consistency)
  pconsistency.threshold    = 0.90,           # default 0.90
  consistency_method        = fs_consistency_method,
  m1.threshold              = Inf,            # default Inf
  minp                      = 0.025,          # default 0.025
  # --- Consistency search -------------------------------------------------
  fs.splits                 = 1000,           # default 1000
  stop_threshold            = 0.95,           # default 0.95
  use_twostage              = TRUE,           # default TRUE
  twostage_args             = list(),         # default list()
  max.minutes               = 3,              # default 3
  # --- Output / diagnostics -----------------------------------------------
  show_candidate_summary    = TRUE,           # default FALSE
  details                   = TRUE,           # default FALSE
  quiet                     = FALSE,          # default FALSE
  by.risk                   = 12,             # default 12
  plot.sg                   = TRUE,           # default FALSE
  plot.grf                  = FALSE,          # default FALSE
  # --- Parallel -----------------------------------------------------------
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$forestsearch <- (proc.time() - t0)["elapsed"]

cat("\nForestSearch completed in",
    round(timings$forestsearch, 1), "seconds\n")
```




## ForestSearch Results

### Identified Subgroups

```{r fs-results}
# Generate results tables
res_tabs <- sg_tables(fs, ndecimals = 3, which_df = "est")

# Display top subgroups meeting criteria
res_tabs$sg10_out
```

### Treatment Effect Estimates

```{r fs-estimates}
# ITT and subgroup estimates
res_tabs$tab_estimates
```

### Identified Subgroup Definition

```{r fs-subgroup}
cat("Identified subgroup (H):", paste(fs$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs$df.est$treat.recommend == 0)))
```

ForestSearch identifies **Estrogen ≤ 0** (ER-negative) as the subgroup with
potential harm. This is biologically plausible: tamoxifen is a selective
estrogen receptor modulator with limited efficacy in ER-negative tumors.

### Selection Diagnostic: Pareto Frontier

Beyond the single subgroup chosen by `sg_focus`, it is often informative to
inspect the *Pareto frontier* on (effect, N) -- the set of candidate subgroups
that are not dominated on both axes simultaneously. A candidate is
dominated when another candidate has at least as large an effect *and*
at least as large a sample size, with at least one strict inequality.
The frontier is a post-hoc diagnostic that lists genuine trade-off
alternatives to the selected subgroup; it does **not** affect selection.

For `"hrMinSG"` the selected subgroup may not appear on the frontier --
that focus deliberately prefers small subgroups, which can be N-dominated
by their larger neighbors.

```{r fs-pareto}
# Formatted gt table; selected subgroup is highlighted and marked with ★.
# Works uniformly for survival (HR) and GLM (OR, RR, IRR, RD, IRD, MD)
# outcomes -- effect column label and scale derived from fs$effect_measure.
pareto_frontier_table(fs)
```

```{r fs-pareto-data, eval = FALSE}
# Programmatic access (effect on natural scale, plus is_selected flag):
pareto_frontier_table(fs, format = "data.table")
```

### Agreement with GRF

ForestSearch and GRF identify subgroups by different mechanisms --
FS via exhaustive search with consistency filtering, GRF via causal
survival forests plus policy trees -- so agreement on subject-level
membership provides an informal robustness check.  `compare_fs_subgroups()`
produces a publication-ready cross-tabulation plus Cohen's kappa; it
accepts forestsearch and GRF result objects directly, and handles
non-detection (method identifies no subgroup) by assigning all
subjects to the complement.

```{r fs-grf-concordance}
sg_compare <- compare_fs_subgroups(
  fs1               = fs,
  fs2               = grf_est,
  label1            = "ForestSearch",
  label2            = "GRF Standalone",
  id.name           = id.name,
  subgroup_notation = "harm",
  title             = "FS vs GRF: Subgroup Membership Concordance",
  subtitle          = "GBSG observed data (Cox, harm search)",
  font_size         = 13
)
sg_compare$table
```

```{r fs-grf-concordance-stats}
cat("Concordance statistics:\n")
cat(sprintf("  Agreement:       %.1f%%\n",
    100 * sg_compare$concordance$agreement))
cat(sprintf("  Cohen's kappa:   %.3f\n",
    sg_compare$concordance$kappa))
cat(sprintf("  Concordant:      %d / Discordant: %d / Total: %d\n",
    sg_compare$concordance$n_agree,
    sg_compare$concordance$n_disagree,
    sg_compare$concordance$n_total))
```


# Bootstrap Bias Correction

## Rationale

Cox model estimates from identified subgroups are **upwardly biased** due to
the selection process (subgroups are selected *because* they show extreme effects).
Bootstrap bias correction addresses this by:

1. Resampling with replacement
2. Re-running the entire ForestSearch algorithm
3. Computing bias terms from bootstrap vs. observed estimates
4. Applying infinitesimal jackknife variance estimation

## Running Bootstrap Analysis

```{r bootstrap}

t0 <- proc.time()

fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est = fs,
  nb_boots = NB,
  show_three = FALSE,
  details = FALSE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nBootstrap completed in",
    round(timings$bootstrap / 60, 1), "minutes\n")
```

## Bootstrap Summary and Diagnostics

```{r bootstrap-summary}
# Comprehensive summary with diagnostics
summaries <- summarize_bootstrap_results(
  sgharm = fs$sg.harm,
  boot_results = fs_bc,
  create_plots = TRUE,
  est.scale = "hr"
)

# Display bias-corrected estimates table
summaries$table
```

## Fast de-biased gate vs. full bootstrap

The Tier-2 de-biased gate is a fast multiplier-bootstrap approximation of the
bootstrap bias-correction. It reuses per-candidate Cox influence (`dfbeta`) over
the selection engine's **own** candidate family — the enumerated consistency
candidates, DINA's qualifying cuts, or GRF's DR-score frontier, depending on
`fs_subgroup_method` — and re-selects under multiplier draws, so its
selection-bias term mirrors the same search optimism the full bootstrap
corrects — without re-running the search per resample. The table below compares
the gate (computed inside the fit) against the full bootstrap from the previous
section. The naive-HR row is a scale sanity check; agreement on the de-biased
row is the Tier-2-vs-Tier-1 validation. By default the de-biased CI uses the same
infinitesimal-jackknife variance as the full bootstrap (the `ci_method = 'ij'` default),
so the de-biased row compares both point estimate and interval like-for-like. The gate de-biases both the selected
subgroup and its complement, so both appear in the comparison and the table below.

```{r debias-gate-comparison, eval = !is.null(fs$sg.harm)}
#| echo: false
#| results: asis
g <- fs$debias_gate
if (is.null(g)) {
  cat("_The de-biased gate did not run (set `run_debias_gate <- TRUE`)._\n")
} else {
  fmt_ci <- function(e, l, u) sprintf("%.3f (%.3f, %.3f)", e, l, u)
  fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0  <- fl(fs_bc$H_estimates, "H0"); H0l <- fl(fs_bc$H_estimates, "H0_lower"); H0u <- fl(fs_bc$H_estimates, "H0_upper")
  H2  <- fl(fs_bc$H_estimates, "H2"); H2l <- fl(fs_bc$H_estimates, "H2_lower"); H2u <- fl(fs_bc$H_estimates, "H2_upper")

  # Complement: Tier-2 from g$complement (include_complement = TRUE), Tier-1 from
  # the bootstrap's Hc_estimates.
  cc  <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  Hc0 <- fl(fs_bc$Hc_estimates, "H0"); Hc0l <- fl(fs_bc$Hc_estimates, "H0_lower"); Hc0u <- fl(fs_bc$Hc_estimates, "H0_upper")
  Hc2 <- fl(fs_bc$Hc_estimates, "H2"); Hc2l <- fl(fs_bc$Hc_estimates, "H2_lower"); Hc2u <- fl(fs_bc$Hc_estimates, "H2_upper")

  q  <- c("naive HR (95% CI)", "de-biased HR (95% CI)")
  t2 <- c(fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  t1 <- c(fmt_ci(H0, H0l, H0u), fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    q  <- c(q, "complement naive HR (95% CI)", "complement de-biased HR (95% CI)")
    t2 <- c(t2, fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    t1 <- c(t1, fmt_ci(Hc0, Hc0l, Hc0u), fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `Tier 2 (gate)` = t2,
                     `Tier 1 (bootstrap)` = t1, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE,
        caption = sprintf("Selected subgroup: %s%s", paste(fs$sg.harm, collapse = " & "),
                          if (has_cc) " (complement = remaining subjects)" else "")),
      sep = "\n"); cat("\n\n")

  # Scale cross-check: the gate's naive HR and the bootstrap's naive H0 are the
  # same quantity (raw subgroup HR) computed two ways; they should agree.
  if (is.finite(H0) && is.finite(g$naive$est)) {
    scale_ok <- abs(log(g$naive$est) - log(H0)) < 1e-2
    cat(sprintf("Scale cross-check (naive gate vs bootstrap H0): **%s** (%.3f vs %.3f).\n\n",
                if (scale_ok) "PASS" else "CHECK", g$naive$est, H0))
  }

  # Family is the full enumerated space, not the screened survivors.
  n_surv <- tryCatch(nrow(fs$find.grps$out.found$hr.subgroups), error = function(e) NA_integer_)
  cat(sprintf("Candidate family: %d (full <= maxk space)%s.\n\n", g$n_family,
              if (!is.na(n_surv)) sprintf("; screened survivors = %d", n_surv) else ""))

  cat(sprintf("Re-selection rule = `%s`%s.\n\n",
              if (!is.null(g$gate$reselection)) g$gate$reselection else "?",
              if (fs_subgroup_method %in% c("dina", "grf"))
                sprintf(" (a Cox-effect proxy for the %s; native-family band = %s)",
                        if (identical(fs_subgroup_method, "dina")) "DINA tau-hat" else "GRF DR-score",
                        if (is.null(fs_family_native_nbhd)) "default (matched to effect_neighborhood)"
                        else if (fs_family_native_nbhd >= 1) "off (full family)"
                        else sprintf("%.2f", fs_family_native_nbhd))
              else " (matches the consistency search statistic)"))

  boot_s <- as.numeric(timings$bootstrap)
  cat(sprintf("Timing: Tier 1 bootstrap (%d boots) %.1f s (%.1f min); Tier 2 gate %.2f s",
              NB, boot_s, boot_s / 60, g$timing_seconds))
  if (is.finite(boot_s) && g$timing_seconds > 0)
    cat(sprintf(" -> **%.0fx** faster (multiplier loop only; the full bootstrap scales with `NB`, the gate does not)", boot_s / g$timing_seconds))
  cat(".\n\n")

  cat(sprintf("Gate `harm_flag_debiased` = **%s** (de-biased HR %.3f %s gate %.2f); selection bias %.4f (log-HR).\n",
              fs$harm_flag_debiased, g$debiased$est,
              if (g$gate$type == "point") "vs" else "lower-bound vs",
              g$gate$t_gate, g$selection_bias))
  if (!is.null(g$ci_method) && !is.null(g$debiased$se_ij))
    cat(sprintf("De-biased CI: %s (IJ SE %.3f vs robust SE %.3f, log-HR).\n",
                if (identical(g$ci_method, "ij")) "infinitesimal jackknife (Tier-1 analogue)" else "robust SE",
                g$debiased$se_ij, g$debiased$se_wald))
  if (has_cc)
    cat(sprintf("Complement (n = %d): naive HR %.3f -> de-biased %.3f (selection bias %.4f, log-HR).\n",
                cc$n, cc$naive$est, cc$debiased$est, cc$selection_bias))
}
```

```{r debias-gate-no-subgroup, eval = is.null(fs$sg.harm)}
#| echo: false
#| results: asis
cat("_No harm subgroup was selected, so the de-biased gate comparison does not apply._\n")
```

The same bias-corrected estimates table, rebuilt from the Tier-2 gate object. The
descriptive columns and naive effect are tier-independent (inherited from the
bootstrap result used as a template); the bias-corrected column carries the gate's
de-biased estimate for **both** the selected subgroup and its complement.

```{r debias-gate-table, eval = !is.null(fs$sg.harm) && !is.null(fs$debias_gate)}
gate_estimates_table(fs, boot_results = fs_bc, est.scale = "hr", digits = 2)
```

## Kaplan-Meier by Identified Subgroups

```{r fs-results-figure}
#| fig-width: 10
#| fig-height: 7
#| fig-cap: "Kaplan-Meier survival curves by identified subgroup"

km_result <- plot_sg_weighted_km(
  fs.est = fs,
  outcome.name = "time_months",
  event.name = "status",
  treat.name = "hormon",
  show.logrank = FALSE,
  conf.int = TRUE,
  by.risk = 12,
  show.cox = FALSE, show.cox.bc = TRUE,
  fs_bc = fs_bc,
  hr_bc_position = "topright"
)
```

::: {.callout-note}
`r figure_note(km_result, prefix = "**Note:** ", custom_text = "Medians [95% CI] for arms are un-adjusted.")`
:::


### Event Count Summary

Low event counts can lead to unstable HR estimates. This summary helps
identify potential issues:

```{r event-summary}
event_summary <- summarize_bootstrap_events(fs_bc, threshold = 10)
```

### Bootstrap Diagnostics

```{r bootstrap-diagnostics}
# Quality metrics
summaries$diagnostics_table_gt
```

### Subgroup Agreement

How consistently does bootstrap identify the same subgroup?

```{r subgroup-agreement}
# Agreement with original analysis
if (!is.null(summaries$subgroup_summary$original_agreement)) {
  summaries$subgroup_summary$original_agreement
}

# Factor presence across bootstrap iterations
if (!is.null(summaries$subgroup_summary$factor_presence)) {
  summaries$subgroup_summary$factor_presence
}
```

### GRF-Cut Stability Across Bootstraps

Separately from whether bootstrap identifies the *same subgroup*, we can
ask what raw cut GRF itself surfaced on each resample, independent of
whether the downstream consistency stage accepted it.  A dominant row
here paired with a low `(no GRF cut)` rate indicates that GRF's
upstream discovery is stable across resamples; a frequent
`(no GRF cut)` row indicates GRF itself is the source of
variability.  This distinction matters because the two failure modes
call for different tuning responses — relaxing `dmin.grf` /
`frac.tau` (for "GRF produced no cut") vs. relaxing
`pconsistency.threshold` (for "GRF produced a cut that consistency
rejected").

```{r bootstrap-grf-cuts}
if (!is.null(summaries$subgroup_summary$grf_cut_freq)) {
  summaries$subgroup_summary$grf_cut_freq |>
    utils::head(10) |>
    gt::gt() |>
    gt::tab_header(
      title = gt::md("**GRF cuts across bootstrap iterations**"),
      subtitle = "Top 10 by frequency; spans all bootstraps (not just successful ones)"
    ) |>
    gt::cols_label(
      Rank    = "Rank",
      grf_cut = "GRF cuts",
      N       = "N",
      Percent = "% of bootstraps"
    ) |>
    gt::fmt_number(columns = "Percent", decimals = 2) |>
    gt::tab_options(table.font.size = gt::px(13))
}
```

### Bootstrap Distributions

```{r bootstrap-plots}
#| fig-width: 10
#| fig-height: 4

if (!is.null(summaries$plots)) {
  summaries$plots$H_distribution + summaries$plots$Hc_distribution
}
```


# DINA-Based Subgroup Identification

DINA (Doubly-robust Identification of Non-Additive effects) screens and
selects subgroups from an influence-function representation of the
covariate-specific treatment effect, fit with cross-fitting. It plugs into
the ForestSearch workflow in three complementary ways, all configured to
mirror the main Cox analysis above (same confounders, harm threshold
`hr.threshold = 1.25`, minimum size `n.min = 60`, and `sg_focus`):

* **(A)** ForestSearch *using DINA for screening* -- DINA's per-covariate
  frontier contributes candidate cuts to the screening pool; the consistency
  search still performs selection.
* **(B)** *Standalone DINA* with parameters analogous to (A) -- the same
  fit's frontier (the candidates (A) screens) and its single selected
  subgroup via `dina_subgroup()`.
* **(C)** ForestSearch with `subgroup_method = "dina"` -- subgroup
  *selection* is delegated to `dina_subgroup()`, then the standard
  ForestSearch bootstrap provides the selection bias-corrected Cox estimate.

```{r dina-config}
#| code-fold: false

# Shared DINA *fit* settings (Cox / survival outcome => family mapped to "cox").
# Selection parameters (sg_focus, hr_threshold) live in the setup chunk under
# the fs_dina_* knobs so they can be tuned independently; (B) and (C) read them.
dina_seed   <- 8316951
dina_nfolds <- 2
# Harm/benefit floor on the natural-parameter (log-HR) scale for ratio measures
m_diff_dina <- log(fs_dina_hr_threshold)
```

## (A) ForestSearch Using DINA for Screening

Here ForestSearch is run exactly as in the main analysis above, but with
`use_dina = TRUE`: DINA's frontier cuts are merged into the candidate pool
alongside the GRF and LASSO cuts. `dina_args` is a flat, fully-named list;
in the screening role it supplies the DINA *fit* tuning (`seed`, `n_folds`)
and the *frontier* controls (`scope`, `max_per_covariate`, `max_subgroups`).
`scope = "wide"` casts a deliberately wide net -- the consistency stage
remains the gatekeeper. By default `dina_args$selected_only = TRUE` (matching
GRF's `return_selected_cuts_only`), so screening contributes the single cut
`dina_subgroup()` would select -- using this call's `sg_focus` /
`selection_rule` / `effect_neighborhood` / `hr.threshold`, i.e. the same cut
`subgroup_method = "dina"` (C) chooses. Set `dina_args$selected_only = FALSE`
to contribute the full frontier instead (more candidates).

```{r fs-dina-screen}
t0 <- proc.time()

fs_dina_screen <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis,
  outcome.name              = outcome.name,
  event.name                = event.name,
  treat.name                = treat.name,
  id.name                   = id.name,
  confounders.name          = confounders.name,
  potentialOutcome.name     = NULL,           # default NULL
  flag_harm.name            = NULL,           # default NULL
  df.predict                = NULL,           # default NULL
  df.test                   = NULL,           # default NULL
  is.RCT                    = TRUE,           # default TRUE
  seedit                    = 8316951,        # default 8316951
  est.scale                 = "hr",           # default "hr"
  # --- Screening engines: GRF + LASSO + DINA ------------------------------
  use_lasso                 = TRUE,           # default TRUE
  use_grf                   = TRUE,           # default TRUE
  grf_res                   = NULL,           # default NULL
  grf_cuts                  = NULL,           # default NULL
  grf_depth                 = 2,              # default 2
  dmin.grf                  = 4,              # default 0.0
  frac.tau                  = 0.80,           # default 0.8
  vi.grf.min                = -0.2,           # default -0.2
  return_selected_cuts_only = TRUE,           # default TRUE
  tune_grf                  = FALSE,          # default FALSE
  use_dina                  = TRUE,           # default FALSE
  dina_res                  = NULL,           # default NULL
  dina_cuts                 = NULL,           # default NULL
  dina_args = list(                           # default list()
    seed = dina_seed, n_folds = dina_nfolds,
    scope = "wide", max_per_covariate = 3, max_subgroups = 10,
    selected_only = dina_screen_selected_only # set in setup chunk
  ),
  # --- Candidate-cut construction -----------------------------------------
  max_n_confounders         = 1000,           # default 1000
  conf_force                = NULL,           # default NULL
  defaultcut_names          = NULL,           # default NULL
  cut_type                  = "default",      # default "default"
  exclude_cuts              = NULL,           # default NULL
  replace_med_grf           = FALSE,          # default FALSE
  cont.cutoff               = 4,              # default 4
  conf.cont_medians         = NULL,           # default NULL
  conf.cont_medians_force   = NULL,           # default NULL
  conf.cont_jcuts           = NULL,           # default NULL
  # --- Selection engine ---------------------------------------------------
  subgroup_method           = "consistency",  # default "consistency"; or "dina"
  sg_focus                  = dina_screen_sg_focus,    # set in setup chunk
  selection_rule            = "neighborhood",  # default "neighborhood"
  effect_neighborhood       = 0.10,            # default 0.10
  # --- Subgroup constraints -----------------------------------------------
  n.min                     = 60,             # default 60
  d0.min                    = 10,             # default 10
  d1.min                    = 10,             # default 10
  maxk                      = 2,              # default 2
  max_subgroups_search      = dina_screen_max_subgroups_search,  # set in setup chunk
  # --- Thresholds (survival / HR scale) -----------------------------------
  hr.threshold              = dina_screen_hr_threshold,    # set in setup chunk
  hr.consistency            = dina_screen_hr_consistency,  # set in setup chunk
  effect.threshold          = NULL,           # default NULL (alias of hr.threshold)
  consistency.threshold     = NULL,           # default NULL (alias of hr.consistency)
  pconsistency.threshold    = 0.90,           # default 0.90
  m1.threshold              = Inf,            # default Inf
  minp                      = 0.025,          # default 0.025
  # --- Consistency search -------------------------------------------------
  fs.splits                 = 1000,           # default 1000
  stop_threshold            = 0.95,           # default 0.95
  use_twostage              = TRUE,           # default TRUE
  twostage_args             = list(),         # default list()
  max.minutes               = 3,              # default 3
  # --- Output / diagnostics -----------------------------------------------
  show_candidate_summary    = TRUE,           # default FALSE
  details                   = TRUE,           # default FALSE
  quiet                     = FALSE,          # default FALSE
  by.risk                   = 12,             # default 12
  plot.sg                   = FALSE,          # default FALSE
  plot.grf                  = FALSE,          # default FALSE
  # --- Parallel -----------------------------------------------------------
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$fs_dina_screen <- (proc.time() - t0)["elapsed"]

cat("Identified subgroup (H):",
    paste(fs_dina_screen$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs_dina_screen$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs_dina_screen$df.est$treat.recommend == 0)))
```

The cuts DINA contributed to the candidate pool are recorded on the result
object. These are exactly the frontier candidates reproduced standalone in
(B).

```{r fs-dina-screen-cuts}
dina_screen_cuts <- fs_dina_screen$dina_cuts
if (length(dina_screen_cuts) > 0) {
  gt(data.frame(
    `#` = seq_along(dina_screen_cuts),
    `DINA candidate cut` = dina_screen_cuts,
    check.names = FALSE
  )) |>
    tab_header(title = "DINA Screening Candidates Injected into ForestSearch") |>
    cols_align(align = "left")
} else {
  cat("DINA contributed no candidate cuts under the configured frontier.\n")
}
```

```{r fs-dina-screen-pareto}
# Selection diagnostic for the DINA-screened run (selected subgroup marked).
pareto_frontier_table(fs_dina_screen)
```

## (B) Standalone DINA

The same DINA fit, run outside ForestSearch. Its `dina_frontier()` output is
the candidate set fed into screening in (A); its `dina_subgroup()` output is
the single subgroup selected by DINA itself -- the mechanism (C) uses
internally. Parameters are analogous to (A): same confounders, harm floor
`m_diff = log(1.25)`, `n_min = 60`, and `sg_focus`.

```{r dina-standalone-fit}
t0 <- proc.time()

dina_fit <- dina(
  df         = df.analysis,
  outcome    = outcome.name,
  treatment  = treat.name,
  covariates = confounders.name,
  family     = "cox",          # default "gaussian"; gaussian/binomial/poisson/cox
  status     = event.name,     # default NULL (required for family = "cox")
  # ... forwarded to dina_fit():
  seed              = dina_seed,   # default NULL
  n_folds           = dina_nfolds, # default 2L
  propensity_method = "logistic",  # default "logistic"
  baseline_method   = NULL,        # default NULL
  cross_fitting     = TRUE,        # default TRUE
  cens_type         = NULL,        # default NULL (no censoring model)
  cens_params       = list(),      # default list()
  n_grid            = 500L,        # default 500L
  eps               = 1e-6         # default 1e-6
)

timings$dina_standalone <- (proc.time() - t0)["elapsed"]
cat("Standalone DINA fit complete (family = cox).\n")
```

### DINA Frontier (Screening Candidates)

```{r dina-standalone-frontier}
dina_fr <- dina_frontier(
  fit               = dina_fit,
  df                = df.analysis,
  covariates        = confounders.name,
  scope             = "wide",     # default "wide"; or "harm" (requires m_diff)
  m_diff            = NULL,        # default NULL (no effect floor under scope = "wide")
  n_min             = 60,         # default 60L
  direction         = "both",     # default "both"; or "left" / "right"
  max_per_covariate = 3,          # default 3L
  max_subgroups     = 10,         # default 10L
  digits            = 3           # default 3L
)

gt(dina_fr[, c("covariate", "direction", "threshold",
               "n_subgroup", "effect", "cut_expr")]) |>
  tab_header(
    title = "Standalone DINA Frontier",
    subtitle = "Per-covariate non-dominated (effect, N) candidates; effect on HR scale"
  ) |>
  fmt_number(columns = c(threshold, effect), decimals = 3) |>
  cols_label(
    covariate = "Covariate", direction = "Direction",
    threshold = "Threshold", n_subgroup = "N", effect = "HR",
    cut_expr = "Cut"
  ) |>
  cols_align(align = "left", columns = c(covariate, direction, cut_expr))
```

### DINA-Selected Subgroup

```{r dina-standalone-select}
dina_sel <- dina_subgroup(
  fit                 = dina_fit,
  df                  = df.analysis,
  covariates          = confounders.name,
  m_diff              = m_diff_dina,
  n_min               = 60,             # default 60L
  direction           = "both",         # default "both"; or "left" / "right"
  sg_focus            = fs_dina_sg_focus,  # set in setup chunk
  selection_rule      = "neighborhood",  # default "neighborhood"; pareto/both (hrMaxSG/hrMinSG)
  effect_neighborhood = 0.10,            # default 0.10
  alpha               = 0.05             # default 0.05
)

if (isTRUE(dina_sel$found)) {
  op <- if (identical(dina_sel$direction, "left")) "<=" else ">="
  cat(sprintf("DINA-selected subgroup: %s %s %s\n",
              dina_sel$covariate, op, format(dina_sel$threshold)))
  cat(sprintf("Direction: %s | N = %d\n",
              dina_sel$direction, dina_sel$n_subgroup))
} else {
  cat("Standalone DINA selected no subgroup meeting the criteria.\n")
}
```

## (C) ForestSearch with `subgroup_method = "dina"`

In this mode ForestSearch delegates subgroup *selection* to
`dina_subgroup()`, bypassing GRF, LASSO and the consistency search. The
selection criteria are inherited from this call's `sg_focus`, `n.min` and
`hr.threshold` (mapped to the log-HR harm floor); `dina_args` supplies only
the fit tuning. The returned object is a standard `forestsearch` result, so
the existing bootstrap machinery applies unchanged.

```{r fs-dina-select}
t0 <- proc.time()

fs_dina <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis,
  outcome.name        = outcome.name,
  event.name          = event.name,
  treat.name          = treat.name,
  id.name             = id.name,
  confounders.name    = confounders.name,
  df.predict          = NULL,            # default NULL
  df.test             = NULL,            # default NULL
  is.RCT              = TRUE,            # default TRUE
  seedit              = 8316951,         # default 8316951
  # --- DINA-selection mode ------------------------------------------------
  subgroup_method     = "dina",          # default "consistency"
  dina_args           = list(seed = dina_seed, n_folds = dina_nfolds),  # fit tuning only
  dina_res            = NULL,            # default NULL (reuse a fitted dina object)
  # --- Selection criteria (passed through to dina_subgroup) ---------------
  sg_focus            = fs_dina_sg_focus,  # set in setup chunk
  selection_rule      = "neighborhood",  # default "neighborhood"
  effect_neighborhood = 0.10,            # default 0.10
  n.min               = 60,              # default 60
  hr.threshold        = fs_dina_hr_threshold,  # set in setup chunk (-> log-HR floor)
  # --- Tier-2 de-biased gate ----------------------------------------------
  debias_gate         = run_debias_gate,           # set in setup chunk
  debias_gate_args    = list(draws = gate_draws,
                             family_native_neighborhood = fs_dina_family_native_nbhd),
  # --- Output -------------------------------------------------------------
  details             = TRUE             # default FALSE
  # NOTE: subgroup_method = "dina" bypasses GRF, LASSO and the consistency
  # search; use_grf / use_lasso / use_twostage / fs.splits / maxk /
  # stop_threshold and related parameters do not apply in this mode.
)

timings$fs_dina_select <- (proc.time() - t0)["elapsed"]

cat("DINA-selected subgroup (H):",
    paste(fs_dina$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs_dina$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs_dina$df.est$treat.recommend == 0)))
```

### Bootstrap Bias-Corrected Cox Estimates

The within-subgroup Cox hazard ratio is upwardly biased by selection. The
ForestSearch bootstrap corrects this by re-running the full pipeline on each
resample; because the result above was produced in DINA-selection mode, each
replicate re-fits DINA and re-runs `dina_subgroup()` from scratch -- so the
correction reflects the DINA selection process itself.

```{r fs-dina-bootstrap, eval = !is.null(fs_dina$sg.harm)}
# Runs only when DINA identified a subgroup.  The bootstrap bias-corrects an
# identified subgroup's effect, so there is nothing to do otherwise.
t0 <- proc.time()

# Diagnostic run: sequential plan + details = TRUE so per-iteration output is
# visible and errors surface in order (helpful while validating DINA-mode
# bootstrap). Switch plan back to "multisession" for production speed.

fs_dina_bc <- forestsearch_bootstrap_dofuture(
  fs.est = fs_dina,
  nb_boots = NB,
  show_three = FALSE,
  details = TRUE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$fs_dina_bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nDINA-mode bootstrap completed in",
    round(timings$fs_dina_bootstrap / 60, 1), "minutes\n")
```

```{r fs-dina-no-subgroup, eval = is.null(fs_dina$sg.harm)}
# Shown only when DINA selected no subgroup at the configured harm floor.
timings$fs_dina_bootstrap <- 0
cat("DINA (subgroup_method = \"dina\") identified no subgroup at",
    "hr.threshold =", fs_dina_hr_threshold, "-- bootstrap skipped.\n",
    "DINA's model-based effect is attenuated relative to the raw Cox HR,",
    "so this floor may not be reached; see the standalone DINA frontier (B)",
    "for the achievable effect range, and lower the DINA harm floor if a",
    "bootstrap demonstration is desired.\n")
```

```{r fs-dina-bootstrap-summary, eval = !is.null(fs_dina$sg.harm)}
dina_summaries <- summarize_bootstrap_results(
  sgharm = fs_dina$sg.harm,
  boot_results = fs_dina_bc,
  create_plots = FALSE,
  est.scale = "hr"
)

dina_summaries$table
```

#### Tier-2 de-biased gate (counterpart to the Tier-1 bootstrap)

The same Tier-2-vs-Tier-1 comparison as the main analysis, for the
DINA-selection fit. The gate de-biases against DINA's **own** qualifying
candidate family (re-running nothing), while the bootstrap above re-runs DINA on
every resample.

For DINA the two tiers are **not** expected to agree as tightly as for the
consistency method. DINA selects on its survival-forest **tau-hat**, but the gate
can only re-select on the perturbed **Cox** effect; over a large depth-2 candidate
family the gate's winner's-curse term can therefore exceed the bootstrap's and
over-correct the selected subgroup. To keep the gate's re-selection neighbourhood
close to DINA's own selection band, the gate by default restricts that family to
candidates competitive on DINA's native tau-hat, matching the band to
`effect_neighborhood` (the recommended default, used here via
`fs_dina_family_native_nbhd <- NULL`; set an explicit number for a different band,
or `>= 1` to de-bias against the full unrestricted family). The complement is the
induced residual rather than a max-over-family selection, so it agrees across
tiers regardless. This band tunes only the *candidate-competition* part of the
bias -- the gate holds the survival forest fixed, so the forest-refit optimism is
invisible to it; treat the full bootstrap as the reference for DINA and the gate
as a fast, conservative check.

```{r fs-dina-debias-gate-comparison, eval = !is.null(fs_dina$sg.harm)}
#| echo: false
#| results: asis
g <- fs_dina$debias_gate
if (is.null(g)) {
  cat("_The de-biased gate did not run (set `run_debias_gate <- TRUE`)._\n")
} else {
  fmt_ci <- function(e, l, u) sprintf("%.3f (%.3f, %.3f)", e, l, u)
  fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0 <- fl(fs_dina_bc$H_estimates, "H0"); H0l <- fl(fs_dina_bc$H_estimates, "H0_lower"); H0u <- fl(fs_dina_bc$H_estimates, "H0_upper")
  H2 <- fl(fs_dina_bc$H_estimates, "H2"); H2l <- fl(fs_dina_bc$H_estimates, "H2_lower"); H2u <- fl(fs_dina_bc$H_estimates, "H2_upper")
  cc <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  q  <- c("naive HR (95% CI)", "de-biased HR (95% CI)")
  t2 <- c(fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  t1 <- c(fmt_ci(H0, H0l, H0u), fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    Hc0 <- fl(fs_dina_bc$Hc_estimates, "H0"); Hc0l <- fl(fs_dina_bc$Hc_estimates, "H0_lower"); Hc0u <- fl(fs_dina_bc$Hc_estimates, "H0_upper")
    Hc2 <- fl(fs_dina_bc$Hc_estimates, "H2"); Hc2l <- fl(fs_dina_bc$Hc_estimates, "H2_lower"); Hc2u <- fl(fs_dina_bc$Hc_estimates, "H2_upper")
    q  <- c(q, "complement naive HR (95% CI)", "complement de-biased HR (95% CI)")
    t2 <- c(t2, fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    t1 <- c(t1, fmt_ci(Hc0, Hc0l, Hc0u), fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `Tier 2 (gate)` = t2,
                     `Tier 1 (bootstrap)` = t1, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE,
        caption = sprintf("DINA-selected subgroup: %s%s",
                          paste(fs_dina$sg.harm, collapse = " & "),
                          if (has_cc) " (complement = remaining subjects)" else "")),
      sep = "\n"); cat("\n\n")

  boot_s <- as.numeric(timings$fs_dina_bootstrap)
  cat(sprintf("Timing: Tier 1 bootstrap (%d boots) %.1f s (%.1f min); Tier 2 gate %.2f s",
              NB, boot_s, boot_s / 60, g$timing_seconds))
  if (is.finite(boot_s) && g$timing_seconds > 0)
    cat(sprintf(" -> **%.0fx** faster (the bootstrap scales with `NB`, the gate does not)",
                boot_s / g$timing_seconds))
  cat(".\n\n")
  cat(sprintf("Gate harm flag = **%s** (de-biased HR %.3f %s gate %.2f); selection bias %.4f (log-HR); candidate family = %d.\n",
              isTRUE(g$harm_flag), g$debiased$est,
              if (g$gate$type == "point") "vs" else "lower-bound vs",
              g$gate$t_gate, g$selection_bias, g$n_family))
  cat(sprintf("Re-selection rule = `%s` (a Cox-effect proxy for DINA's tau-hat); native-family band = %s.\n",
              if (!is.null(g$gate$reselection)) g$gate$reselection else "?",
              if (is.null(fs_dina_family_native_nbhd)) "default (matched to effect_neighborhood)"
              else if (fs_dina_family_native_nbhd >= 1) "off (full family)"
              else sprintf("%.2f (restricted to tau-hat-competitive candidates)",
                           fs_dina_family_native_nbhd)))
  if (!is.null(g$ci_method) && !is.null(g$debiased$se_ij))
    cat(sprintf("De-biased CI: %s (IJ SE %.3f vs robust SE %.3f, log-HR).\n",
                if (identical(g$ci_method, "ij")) "infinitesimal jackknife (Tier-1 analogue)" else "robust SE",
                g$debiased$se_ij, g$debiased$se_wald))
}
```

The Tier-1 bias-corrected table above, rebuilt from the Tier-2 gate object (the
descriptive columns and naive effect are tier-independent; the bias-corrected
column carries the gate's de-biased estimate for the subgroup and its complement).

```{r fs-dina-debias-gate-table, eval = !is.null(fs_dina$sg.harm) && !is.null(fs_dina$debias_gate)}
gate_estimates_table(fs_dina, boot_results = fs_dina_bc, est.scale = "hr", digits = 2)
```

::: {.callout-note}
## Screening vs. selection roles for DINA

`dina_args` behaves differently by role. In the **screening** path (A,
`use_dina = TRUE`) it defaults to `selected_only = TRUE` (matching GRF's
`return_selected_cuts_only`): the screening path calls `dina_subgroup()` and
contributes only its selected cut, so it *does* consult this call's
`sg_focus` / `selection_rule` / `effect_neighborhood` -- yielding the same
cut the selection path (C) would choose. Setting `selected_only = FALSE`
instead drives `dina_frontier()`, which is selection-agnostic (`sg_focus` /
`selection_rule` are *not* consulted) and contributes the full candidate set.
In the **selection** path (C, `subgroup_method = "dina"`) the selection is
always driven by this call's `sg_focus` / `selection_rule` /
`effect_neighborhood` via `dina_subgroup()`, and `dina_args` supplies only
the fit. This mode is currently depth-1 (a
single covariate threshold).
:::



## (D) ForestSearch with `subgroup_method = "grf"`

In this mode ForestSearch delegates subgroup *selection* to GRF -- a causal-forest
policy tree (`grf_selection = "tree"`) or the doubly-robust-score frontier
(`"frontier"`) -- bypassing LASSO, DINA and the consistency search. The selection
criteria are inherited from this call's `sg_focus` (mapped to the GRF frontier/tree
rule), `n.min`, `grf_depth`, `grf_selection` and the DR-score harm floor
`dmin.grf`. The returned object is a standard `forestsearch` result, so the
existing bootstrap machinery applies unchanged.

```{r fs-grf-select}
t0 <- proc.time()

fs_grf <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis,
  outcome.name        = outcome.name,
  event.name          = event.name,
  treat.name          = treat.name,
  id.name             = id.name,
  confounders.name    = confounders.name,
  df.predict          = NULL,            # default NULL
  df.test             = NULL,            # default NULL
  is.RCT              = TRUE,            # default TRUE
  seedit              = 8316951,         # default 8316951 (also seeds the forest)
  # --- GRF-selection mode -------------------------------------------------
  subgroup_method     = "grf",           # default "consistency"
  grf_selection       = fs_grf_demo_selection,  # set in setup chunk ("tree" | "frontier")
  grf_depth           = fs_grf_demo_depth,      # set in setup chunk (tree / DR-pair depth)
  dmin.grf            = fs_grf_demo_dmin,        # set in setup chunk (DR-score harm floor)
  frac.tau            = 0.8,             # default 0.8 (RMST truncation fraction)
  grf_res             = NULL,            # default NULL (reuse a fitted forest)
  # --- Selection criteria -------------------------------------------------
  sg_focus            = fs_grf_sg_focus,  # set in setup chunk (-> GRF frontier/tree rule)
  selection_rule      = "neighborhood",  # default "neighborhood"
  effect_neighborhood = 0.10,            # default 0.10
  n.min               = 60,              # default 60
  # --- Tier-2 de-biased gate ----------------------------------------------
  debias_gate         = run_debias_gate,           # set in setup chunk
  debias_gate_args    = list(draws = gate_draws,
                             family_native_neighborhood = fs_grf_family_native_nbhd),
  # --- Output -------------------------------------------------------------
  details             = TRUE             # default FALSE
  # NOTE: subgroup_method = "grf" bypasses LASSO, DINA and the consistency
  # search; use_lasso / use_dina / use_twostage / fs.splits / maxk /
  # hr.threshold / pconsistency.threshold do not apply in this mode.
)

timings$fs_grf_select <- (proc.time() - t0)["elapsed"]

cat("GRF-selected subgroup (H):",
    paste(fs_grf$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs_grf$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs_grf$df.est$treat.recommend == 0)))
```

### Bootstrap Bias-Corrected Cox Estimates

The within-subgroup Cox hazard ratio is upwardly biased by selection. The
ForestSearch bootstrap corrects this by re-running the full pipeline on each
resample; because the result above was produced in GRF-selection mode, each
replicate re-fits the causal forest and re-runs GRF selection from scratch -- so
the correction reflects the GRF selection process itself.

```{r fs-grf-bootstrap, eval = !is.null(fs_grf$sg.harm)}
# Runs only when GRF identified a subgroup.  The bootstrap bias-corrects an
# identified subgroup's effect, so there is nothing to do otherwise.
t0 <- proc.time()

fs_grf_bc <- forestsearch_bootstrap_dofuture(
  fs.est = fs_grf,
  nb_boots = NB,
  show_three = FALSE,
  details = TRUE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$fs_grf_bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nGRF-mode bootstrap completed in",
    round(timings$fs_grf_bootstrap / 60, 1), "minutes\n")
```

```{r fs-grf-no-subgroup, eval = is.null(fs_grf$sg.harm)}
# Shown only when GRF selected no subgroup at the configured DR-score floor.
timings$fs_grf_bootstrap <- 0
cat("GRF (subgroup_method = \"grf\") identified no subgroup at",
    "dmin.grf =", fs_grf_demo_dmin, "-- bootstrap skipped.\n",
    "Lower the DR-score floor (dmin.grf) or switch grf_selection /",
    "grf_depth if a bootstrap demonstration is desired.\n")
```

```{r fs-grf-bootstrap-summary, eval = !is.null(fs_grf$sg.harm)}
grf_summaries <- summarize_bootstrap_results(
  sgharm = fs_grf$sg.harm,
  boot_results = fs_grf_bc,
  create_plots = FALSE,
  est.scale = "hr"
)

grf_summaries$table
```

#### Tier-2 de-biased gate (counterpart to the Tier-1 bootstrap)

The same Tier-2-vs-Tier-1 comparison as the main analysis, for the
GRF-selection fit. The gate de-biases against GRF's **own** DR-score candidate
family (re-running nothing), while the bootstrap above re-runs GRF on every
resample.

As with DINA, the two tiers are **not** expected to agree as tightly as for the
consistency method. GRF selects on its **doubly-robust (RMST-based) score**, but
the gate can only re-select on the perturbed **Cox** effect; the additive RMST
scale and the multiplicative hazard scale need not rank candidates the same way,
so over a large candidate family the gate's winner's-curse term can exceed the
bootstrap's and over-correct the selected subgroup. To keep the gate's
re-selection neighbourhood close to GRF's own selection band, the gate by default
restricts that family to candidates competitive on GRF's native DR-score, matching
the band to `effect_neighborhood` (the recommended default, used here via
`fs_grf_family_native_nbhd <- NULL`; set an explicit number for a different band,
or `>= 1` to de-bias against the full family). The complement is the induced
residual rather than a max-over-family selection, so it agrees across tiers
regardless. Two caveats specific to GRF: a *dominant* DR-score winner can collapse
the restricted family to a single candidate (`candidate family = 1` in the note
below), in which case the gate removes only the gate-threshold optimism; and,
more fundamentally, the band tunes only the *candidate-competition* part of the
bias -- the gate holds the causal forest fixed, so the forest-refit optimism that
the bootstrap captures is invisible to it. Treat the full bootstrap as the
reference for GRF and the gate as a fast, conservative check.

```{r fs-grf-debias-gate-comparison, eval = !is.null(fs_grf$sg.harm)}
#| echo: false
#| results: asis
g <- fs_grf$debias_gate
if (is.null(g)) {
  cat("_The de-biased gate did not run (set `run_debias_gate <- TRUE`)._\n")
} else {
  fmt_ci <- function(e, l, u) sprintf("%.3f (%.3f, %.3f)", e, l, u)
  fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0 <- fl(fs_grf_bc$H_estimates, "H0"); H0l <- fl(fs_grf_bc$H_estimates, "H0_lower"); H0u <- fl(fs_grf_bc$H_estimates, "H0_upper")
  H2 <- fl(fs_grf_bc$H_estimates, "H2"); H2l <- fl(fs_grf_bc$H_estimates, "H2_lower"); H2u <- fl(fs_grf_bc$H_estimates, "H2_upper")
  cc <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  q  <- c("naive HR (95% CI)", "de-biased HR (95% CI)")
  t2 <- c(fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  t1 <- c(fmt_ci(H0, H0l, H0u), fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    Hc0 <- fl(fs_grf_bc$Hc_estimates, "H0"); Hc0l <- fl(fs_grf_bc$Hc_estimates, "H0_lower"); Hc0u <- fl(fs_grf_bc$Hc_estimates, "H0_upper")
    Hc2 <- fl(fs_grf_bc$Hc_estimates, "H2"); Hc2l <- fl(fs_grf_bc$Hc_estimates, "H2_lower"); Hc2u <- fl(fs_grf_bc$Hc_estimates, "H2_upper")
    q  <- c(q, "complement naive HR (95% CI)", "complement de-biased HR (95% CI)")
    t2 <- c(t2, fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    t1 <- c(t1, fmt_ci(Hc0, Hc0l, Hc0u), fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `Tier 2 (gate)` = t2,
                     `Tier 1 (bootstrap)` = t1, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE,
        caption = sprintf("GRF-selected subgroup: %s%s",
                          paste(fs_grf$sg.harm, collapse = " & "),
                          if (has_cc) " (complement = remaining subjects)" else "")),
      sep = "\n"); cat("\n\n")

  boot_s <- as.numeric(timings$fs_grf_bootstrap)
  cat(sprintf("Timing: Tier 1 bootstrap (%d boots) %.1f s (%.1f min); Tier 2 gate %.2f s",
              NB, boot_s, boot_s / 60, g$timing_seconds))
  if (is.finite(boot_s) && g$timing_seconds > 0)
    cat(sprintf(" -> **%.0fx** faster (the bootstrap scales with `NB`, the gate does not)",
                boot_s / g$timing_seconds))
  cat(".\n\n")
  cat(sprintf("Gate harm flag = **%s** (de-biased HR %.3f %s gate %.2f); selection bias %.4f (log-HR); candidate family = %d.\n",
              isTRUE(g$harm_flag), g$debiased$est,
              if (g$gate$type == "point") "vs" else "lower-bound vs",
              g$gate$t_gate, g$selection_bias, g$n_family))
  cat(sprintf("Re-selection rule = `%s` (a Cox-effect proxy for GRF's DR-score); native-family band = %s.\n",
              if (!is.null(g$gate$reselection)) g$gate$reselection else "?",
              if (is.null(fs_grf_family_native_nbhd)) "default (matched to effect_neighborhood)"
              else if (fs_grf_family_native_nbhd >= 1) "off (full family)"
              else sprintf("%.2f (restricted to DR-score-competitive candidates)",
                           fs_grf_family_native_nbhd)))
  if (!is.null(g$ci_method) && !is.null(g$debiased$se_ij))
    cat(sprintf("De-biased CI: %s (IJ SE %.3f vs robust SE %.3f, log-HR).\n",
                if (identical(g$ci_method, "ij")) "infinitesimal jackknife (Tier-1 analogue)" else "robust SE",
                g$debiased$se_ij, g$debiased$se_wald))
}
```

The Tier-1 bias-corrected table above, rebuilt from the Tier-2 gate object (the
descriptive columns and naive effect are tier-independent; the bias-corrected
column carries the gate's de-biased estimate for the subgroup and its complement).

```{r fs-grf-debias-gate-table, eval = !is.null(fs_grf$sg.harm) && !is.null(fs_grf$debias_gate)}
gate_estimates_table(fs_grf, boot_results = fs_grf_bc, est.scale = "hr", digits = 2)
```

::: {.callout-note}
## Selection statistic vs. the gate's re-selection (GRF)

GRF ranks candidates by the **doubly-robust score** (RMST-based for survival),
while the Tier-2 gate re-selects on the perturbed **Cox** effect. The two scales
(additive RMST vs. multiplicative hazard) need not agree, so the gate is a
*proxy* for GRF's selection -- the same approximation noted for DINA's tau-hat.
`family_native_neighborhood` restricts the gate family to DR-score-competitive
candidates to keep that proxy close to GRF's actual selection neighbourhood; the
full bootstrap remains the reference because it re-runs the DR-score selection
itself.
:::



# Cross-Validation

Cross-validation assesses the **stability** of the ForestSearch algorithm.
Two approaches are available:

## K-Fold Cross-Validation

```{r cv-skipped-note, eval = !run_cv}
cat("K-Fold cross-validation skipped (run_cv = FALSE). Set run_cv <- TRUE",
    "in the parallel-setup chunk to run it.\n")
```

```{r kfold-cv, eval = run_cv}

t0 <- proc.time()

fs_kfold <- forestsearch_tenfold(
  fs.est = fs,
  sims = Ksims,
  Kfolds = 10,
  details = FALSE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = FALSE
  )
)

plan("sequential")
timings$kfold <- (proc.time() - t0)["elapsed"]
metrics_tables <- cv_metrics_tables(fs_kfold)
metrics_tables
```

### Diagnostic Tabulations

`summarize_cv_results()` consolidates the per-fold record into a panel
of diagnostic tables that together answer the core questions an
analyst asks of a repeated K-fold run: which subgroups did the
pipeline identify across resamples, which GRF cuts drove those
identifications, when identification failed what was the proximate
cause, and how confidently were the successful identifications made.

```{r cv-diagnostics-run, eval = run_cv}
cv_diag <- summarize_cv_results(
  cv_output         = fs_kfold,
  original_sg       = fs$sg.harm,
  original_grf_cuts = fs$grf_cuts,
  top_n             = 10L
)
```

**Which subgroups emerged across resamples?**  The identification
summary shows the distribution of selected subgroups across all
`r paste0(Ksims, " × 10 = ", Ksims * 10)` sim × fold pairs.  A
single dominant subgroup with high share indicates the pipeline
converges on a stable selection; a long tail suggests selection is
sensitive to the particular training split.

```{r cv-identification, eval = run_cv}
cv_diag$identification_summary
```

**Which GRF cuts drove those selections?**  The GRF cut summary
tabulates the raw policy-tree output on each training fold,
independent of whether the downstream consistency stage accepted
it.  Comparison against the identification summary reveals whether
GRF's upstream discovery is stable (a dominant cut shared across
folds) or noisy (a long tail of one-off cuts).

```{r cv-grf-cut, eval = run_cv}
cv_diag$grf_cut_summary
```

**When a fold failed to identify a subgroup, why?**  The no-subgroup
decomposition splits the non-identifying folds into two actionable
buckets: *GRF returned no cut* (nothing for consistency to evaluate,
typically pointing at `dmin.grf` or `frac.tau` thresholds) versus
*GRF returned a cut but consistency rejected it* (candidates existed
but none met `pconsistency.threshold`).  The former argues for
relaxing the GRF upstream; the latter for relaxing the consistency
threshold or reviewing candidate quality.

```{r cv-no-subgroup, eval = run_cv}
cv_diag$no_subgroup_decomposition
```

**How confidently were the successful identifications made?**  Among
folds that *did* identify a subgroup, the `pconsistency_distribution`
shows where the selected candidate's consistency probability
(`Pcons`) landed relative to the 0.90 threshold.  Identifications in
the `[0.85, 0.95)` bin are marginal; those concentrated at or near
1.0 are robust.

```{r cv-pcons, eval = run_cv}
cv_diag$pconsistency_distribution
```

**How do all the numeric fold-level diagnostics summarize together?**
The `fold_numeric_summary` tabulates median / IQR / range for each
numeric column in the per-fold record: fold size (`n_test`),
consistency probability (`pconsistency`), training-fold HR of the
selected subgroup (`training_fs_hr`, optimistically biased — for
diagnostic comparison only), and candidate count
(`n_candidates_evaluated`).

```{r cv-fold-numeric, eval = run_cv}
cv_diag$fold_numeric_summary
```

**Agreement with the full-data analysis.**  How often does a training
fold reproduce the original full-data subgroup exactly, share a
factor with it, or match its GRF cut?

```{r cv-original-agreement, eval = run_cv}
cv_diag$original_agreement
```


## Out-of-Bag (N-Fold) Cross-Validation

N-fold CV (leave-one-out):

```{r loo-skipped-note, eval = !run_loo}
cat("Leave-one-out (N-fold / OOB) skipped (run_loo = FALSE). Set run_loo <- TRUE",
    "in the parallel-setup chunk to run it.\n")
```

```{r oob-cv, eval = run_loo}

t0 <- proc.time()

fs_OOB <- forestsearch_Kfold(
  fs.est = fs,
  details = FALSE,
  Kfolds = nrow(df.analysis),  # N-fold = leave-one-out
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$oob <- (proc.time() - t0)["elapsed"]

# Summarize OOB results
cv_out <- forestsearch_KfoldOut(
  res = fs_OOB,
  details = FALSE,
  outall = TRUE
)

tables <- cv_summary_tables(cv_out)

tables$combined_table

tables$metrics_table
```


# Results Visualization

## Forest Plot

The forest plot summarizes treatment effects across the ITT population,
reference subgroups, and identified subgroups with cross-validation metrics.

```{r forest-plot}
#| fig-width: 18
#| fig-height: 12
#| fig-cap: "Subgroup forest plot including identified subgroups"

# Define reference subgroups for comparison
subgroups <- list(
  age_gt65 = list(
    subset_expr = "age > 65",
    name = "Age > 65",
    type = "reference"
  ),
  age_le65 = list(
    subset_expr = "age <= 65",
    name = "Age <= 65",
    type = "reference"
  ),
  pgr_positive = list(
    subset_expr = "pgr > 0",
    name = "PgR > 0",
    type = "reference"
  ),
  pgr_negative = list(
    subset_expr = "pgr <= 0",
    name = "PgR <= 0",
    type = "reference"
  )
)


my_theme <- create_forest_theme(base_size = 24,
  footnote_fontsize = 17, cv_fontsize = 22)


# Create forest plot
# Include fs_kfold and fs_OOB if available for CV metrics
result <- plot_subgroup_results_forestplot(
  fs_results = list(
    fs.est = fs,
    fs_bc = fs_bc,
    fs_OOB   = if (exists("fs_OOB"))   fs_OOB   else NULL,
    fs_kfold = if (exists("fs_kfold")) fs_kfold else NULL
  ),
  df_analysis = df.analysis,
  subgroup_list = subgroups,
  outcome.name = outcome.name,
  event.name = event.name,
  treat.name = treat.name,
  E.name = "Hormonal",
  C.name = "Chemo",
  ci_column_spaces = 25,
  xlog = TRUE,
  theme = my_theme
)


# Render forest plot
render_forestplot(result)
```


### KM Difference Plots: ITT and Subgroups

The solid black line denotes the ITT Kaplan-Meier treatment difference
estimates along with 95% CIs (the grey shaded region). K-M differences
corresponding to subgroups are displayed.

```{r KMdiffs}
#| fig-width: 8
#| fig-height: 6

# Add additional subgroups along with ITT and identified subgroups
ref_sgs <- list(
  age_young = list(subset_expr = "age < 65", color = "brown"),
  age_old = list(subset_expr = "age >= 65", color = "orange")
)

plot_km_band_forestsearch(
  df = df.analysis,
  fs.est = fs,
  ref_subgroups = ref_sgs,
  outcome.name = outcome.name,
  event.name = event.name,
  treat.name = treat.name,
  draws_band = 20
)
```


# Summary and Interpretation

## Key Findings

```{r summary-findings}
cat(strrep("=", 60), "\n")
cat("FORESTSEARCH ANALYSIS SUMMARY\n")
cat(strrep("=", 60), "\n\n")

cat("Dataset: GBSG (N =", nrow(df.analysis), ")\n")
cat("Outcome: Recurrence-free survival\n\n")

cat("ITT Analysis:\n")
cat("  HR (95% CI): 0.69 (0.54, 0.89)\n\n")

cat("Identified Subgroup (H):\n")
cat("  Definition:", paste(fs$sg.harm, collapse = " & "), "\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 0)))
cat("  Unadjusted HR:",
    sprintf("%.2f", exp(fs$grp.consistency$out_sg$result$hr[1])), "\n")

cat("\nComplement Subgroup (Hc):\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 1),
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 1)))
```

## Clinical Interpretation

The ForestSearch analysis identifies **estrogen receptor-negative (ER ≤ 0)**
patients as a subgroup with potential lack of benefit from hormonal therapy.

**Biological plausibility**: Tamoxifen is a selective estrogen receptor
modulator. Its efficacy depends on ER expression. The finding that ER-negative
patients may not benefit is consistent with:

- Mechanistic understanding of tamoxifen action
- Meta-analyses showing no tamoxifen benefit in ER-negative breast cancer
- Clinical guidelines recommending tamoxifen primarily for ER-positive tumors

**Caveats**:

1. This is an **exploratory** analysis requiring independent validation
2. The bias-corrected estimates have wider confidence intervals
3. Cross-validation metrics should be evaluated for algorithm stability

## Computational Timing

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Analysis = c("GRF", "ForestSearch", "Bootstrap",
               "FS (DINA screening)", "DINA (standalone)",
               "FS (subgroup_method=dina)", "FS DINA bootstrap",
               "kfold", "oob", "Total"),
  Seconds = c(
    timings$grf,
    timings$forestsearch,
    timings$bootstrap,
    timings$fs_dina_screen,
    timings$dina_standalone,
    timings$fs_dina_select,
    timings$fs_dina_bootstrap,
    timings$kfold,
    timings$oob,
    timings$total
  )
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(title = "Computational Timing") |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1) |>
  cols_label(
    Analysis = "Component",
    Seconds = "Time (sec)",
    Minutes = "Time (min)"
  )
```


# References

León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024).
"Exploratory subgroup identification in the heterogeneous Cox model:
A relatively simple procedure." *Statistics in Medicine*.
DOI: [10.1002/sim.10163](https://doi.org/10.1002/sim.10163)


# Session Information

```{r session-info}
sessionInfo()
```
