---
title: "Exploratory Subgroup Identification with ForestSearch"
subtitle: "German Breast Cancer Study Group (GBSG) Analysis"
author: "Larry F. León"
date: today
format:
  html:
    code-fold: show
    toc: true
    toc-depth: 3
execute:
  warning: true
  message: true
  fig-align: center
knitr:
  opts_chunk:
    comment: "#>"
---


# Introduction {#sec-intro}

This vignette demonstrates the **ForestSearch** methodology for exploratory
subgroup identification in survival analysis, as described in
León et al. (2024) *Statistics in Medicine*.

## Motivation

In clinical trials, particularly oncology, subgroup analyses are essential for:

- Evaluating treatment effect consistency across patient populations
- Identifying subgroups where treatment may be detrimental (harm)
- Characterizing subgroups with enhanced benefit
- Informing regulatory decisions and clinical practice

While prespecified subgroups provide stronger evidence, important subgroups
based on patient characteristics may not be anticipated. ForestSearch provides
a principled approach to **exploratory** subgroup identification with proper
statistical inference.

## Methodology Overview

ForestSearch identifies subgroups through:

1. **Candidate factor selection**: Using LASSO and/or Generalized Random Forests (GRF)
2. **Exhaustive subgroup search**: Evaluating all combinations up to `maxk` factors
3. **Consistency-based selection**: Applying splitting consistency criteria
4. **Bootstrap bias correction**: Adjusting for selection-induced optimism
5. **Cross-validation**: Assessing algorithm stability

The key innovation is the **splitting consistency criterion**: a subgroup is
considered "consistent with harm" if, when randomly split 50/50 many times,
both halves consistently show hazard ratios ≥ 1.0 (where 1.0 represents a
meaningful "harm threshold").  An appendix (@sec-appendix-fpr) directly
calibrates the procedure-level false-positive rate (FPR) at the operational
thresholds used in the main analysis.


# Setup {#sec-setup}

## Packages

```{r load-packages}
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(policytree)
library(foreach)
library(doFuture)
library(future)

# Optional
library(patchwork)
library(weightedsurv)

theme_set(theme_minimal(base_size = 12))
```

```{r gbsg-params}
#| include: false

# ─── All tunable parameters centralised here ───────────────────────────────

# ── Parallel processing ────────────────────────────────────────────────────
# Set to detectCores() - 1 for production; 2 for CRAN checks.
n_workers <- floor(0.95 * max(1L, parallel::detectCores(logical = FALSE) - 1L))

# ── Main-analysis ForestSearch operational parameters ──────────────────────
# These are the settings used throughout Sections 3-6.
c1_paper            <- 1.25    # hr.threshold (screening)
c2_paper            <- 1.00    # hr.consistency (consistency)
pcons_val           <- 0.90    # pconsistency.threshold
stop_thresh_val     <- 0.95    # early-stopping threshold
fs_splits_val       <- 400L    # consistency-evaluation splits
max_sgs_search_val  <- 10L
n_min_val           <- 60L
d0_min_val          <- 10L
d1_min_val          <- 10L
maxk_val            <- 2L

# ── Bootstrap / cross-validation counts ────────────────────────────────────
# Increase for production:
#   NB_boot  = 500-2000   (here: 20 for fast knit)
#   Ksims_cv = 50         (here: 10 for fast knit)
NB_boot   <- 50L
Ksims_cv  <- 10L

# ── FPR calibration (appendix) ─────────────────────────────────────────────
# Tier 1 = quick approximation (±1.3 pp at FPR~10%, 500 sims)
# Tier 2 = higher-precision verification (±0.3 pp at FPR~10%, 10k sims)
# Tier 2 defaults to 0 (skipped) — raise to 10000-20000 when a verified
# operating point is needed.  Typical Tier 2 runtime: 1-3 hours.
nsims_tier1_fpr   <- 500L
nsims_tier2_fpr   <- 0L        # e.g. 10000L to run Tier 2 verification
nsims_verify_fpr  <- 100L        # independent-seed verification per call
sim_seed_fpr      <- 8316951L
fpr_target_surv   <- 0.12      # reference line on detection curves

# Survival DGM parameters (for the DGM-based FPR estimate)
analysis_time_dgm <- 84L             # months
cens_adjust_dgm   <- log(1.5)
prop_cens_surv    <- 0.56            # observed in GBSG
confounders_dgm   <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")

# ── Internals (do not edit) ────────────────────────────────────────────────
timings <- list()
t_vignette_start <- proc.time()
```


# Data: German Breast Cancer Study Group {#sec-data}

## Study Background

The GBSG trial evaluated hormonal treatment (tamoxifen) versus chemotherapy
in node-positive breast cancer patients. Key characteristics:

- **Sample size**: N = 686
- **Outcome**: Recurrence-free survival time
- **Censoring rate**: ~56%
- **Treatment**: Hormonal therapy (tamoxifen) vs. chemotherapy

## Data Preparation

```{r data-setup}
# Load GBSG data (included in forestsearch package)
df.analysis <- gbsg

# Prepare analysis variables
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
  treat <- hormon
})

# Define variable roles
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name   <- "status"
id.name      <- "id"
treat.name   <- "hormon"

cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```

## Baseline Characteristics

```{r baseline-table}
create_summary_table(
  data             = df.analysis,
  treat_var        = treat.name,
  table_title      = "GBSG Baseline Characteristics by Treatment Arm",
  vars_continuous  = c("age", "nodes", "size", "er", "pgr"),
  vars_categorical = c("grade", "meno"),
  font_size        = 12
)
```

## Kaplan-Meier Analysis (ITT Population)

```{r km-itt}
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "ITT Kaplan-Meier curves with 95% CI and log-rank test."

dfcount <- df_counting(
  df         = df.analysis,
  by.risk    = 6,
  tte.name   = outcome.name,
  event.name = event.name,
  treat.name = treat.name
)

plot_weighted_km(
  dfcount,
  conf.int      = TRUE,
  show.logrank  = TRUE,
  ymax          = 1.05,
  xmed.fraction = 0.775,
  ymed.offset   = 0.125
)
```

The ITT Cox hazard ratio estimate is approximately 0.69 (95% CI: 0.54, 0.89),
suggesting an overall benefit for hormonal therapy.


# Preliminary Analysis: Generalized Random Forests {#sec-grf}

Before running ForestSearch, we use GRF to explore potential treatment
effect heterogeneity and identify candidate factors.

```{r grf-analysis}
t0 <- proc.time()

grf_est <- grf.subg.harm.survival(
  data             = df.analysis,
  confounders.name = confounders.name,
  outcome.name     = outcome.name,
  event.name       = event.name,
  id.name          = id.name,
  treat.name       = treat.name,
  maxdepth         = 2,
  n.min            = n_min_val,
  dmin.grf         = 12,
  frac.tau         = 0.6,
  details          = TRUE,
  return_selected_cuts_only = FALSE
)

timings$grf <- (proc.time() - t0)["elapsed"]
```

```{r grf-trees}
#| fig-width: 10
#| fig-height: 4
#| fig-cap: "GRF policy trees. leaf1 = recommend control; leaf2 = recommend treatment."

oldpar <- par(mfrow = c(1, 2))
plot(grf_est$tree1, leaf.labels = c("Control", "Treat"), main = "Depth 1")
plot(grf_est$tree2, leaf.labels = c("Control", "Treat"), main = "Depth 2")
par(oldpar)
```

GRF identifies estrogen receptor status (ER) as a key factor, with ER ≤ 0
suggesting potential harm from hormonal therapy.


# ForestSearch Analysis {#sec-forestsearch}

## Running ForestSearch

ForestSearch performs an exhaustive search over candidate subgroup
combinations with up to `maxk` factors. Key parameters:

| Parameter | Value | Description |
|-----------|-------|-------------|
| `hr.threshold` | `r c1_paper` | Minimum HR for consistency evaluation |
| `hr.consistency` | `r c2_paper` | Minimum consistency rate for candidates |
| `pconsistency.threshold` | `r pcons_val` | Required consistency for selection |
| `maxk` | `r maxk_val` | Maximum factors in subgroup definition |
| `n.min` | `r n_min_val` | Minimum subgroup sample size |
| `d0.min`, `d1.min` | `r d0_min_val` | Minimum events per treatment arm |

```{r forestsearch-main}
#| fig-width: 10
#| fig-height: 7

t0 <- proc.time()

fs <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  outcome.name     = outcome.name,
  treat.name       = treat.name,
  event.name       = event.name,
  id.name          = id.name,
  # Threshold parameters (per León et al. 2024)
  hr.threshold           = c1_paper,
  hr.consistency         = c2_paper,
  pconsistency.threshold = pcons_val,
  stop_threshold         = stop_thresh_val,
  # Search configuration
  sg_focus             = "maxSG",
  max_subgroups_search = max_sgs_search_val,
  use_twostage         = TRUE,
  # Factor selection
  use_grf                   = TRUE,
  return_selected_cuts_only = TRUE,
  use_lasso                 = TRUE,
  cut_type                  = "default",
  # Subgroup constraints
  maxk   = maxk_val,
  n.min  = n_min_val,
  d0.min = d0_min_val,
  d1.min = d1_min_val,
  # Consistency evaluation
  fs.splits = fs_splits_val,
  # Parallel processing
  parallel_args = list(
    plan         = "multisession",
    workers      = n_workers,
    show_message = TRUE
  ),
  # Output
  showten_subgroups = TRUE,
  details           = TRUE,
  plot.sg           = TRUE
)

plan("sequential")
timings$forestsearch <- (proc.time() - t0)["elapsed"]

cat("\nForestSearch completed in",
    round(timings$forestsearch, 1), "seconds\n")
```


## ForestSearch Results

### Identified Subgroups

```{r fs-results}
res_tabs <- sg_tables(fs, ndecimals = 3, which_df = "est")
res_tabs$sg10_out
```

### Treatment Effect Estimates

```{r fs-estimates}
res_tabs$tab_estimates
```

### Identified Subgroup Definition

```{r fs-subgroup}
cat("Identified subgroup (H):", paste(fs$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs$df.est$treat.recommend == 0)))
```

ForestSearch identifies **Estrogen ≤ 0** (ER-negative) as the subgroup with
potential harm. This is biologically plausible: tamoxifen is a selective
estrogen receptor modulator with limited efficacy in ER-negative tumors.

### Agreement with GRF

ForestSearch and GRF identify subgroups by different mechanisms —
FS via exhaustive search with consistency filtering, GRF via causal
survival forests plus policy trees — so agreement on subject-level
membership provides an informal robustness check. `compare_fs_subgroups()`
produces a publication-ready cross-tabulation plus Cohen's kappa.

```{r fs-grf-concordance}
sg_compare <- compare_fs_subgroups(
  fs1               = fs,
  fs2               = grf_est,
  label1            = "ForestSearch",
  label2            = "GRF Standalone",
  id.name           = id.name,
  subgroup_notation = "harm",
  title             = "FS vs GRF: Subgroup Membership Concordance",
  subtitle          = "GBSG observed data (Cox, harm search)",
  font_size         = 13
)
sg_compare$table
```

```{r fs-grf-concordance-stats}
cat("Concordance statistics:\n")
cat(sprintf("  Agreement:       %.1f%%\n",
    100 * sg_compare$concordance$agreement))
cat(sprintf("  Cohen's kappa:   %.3f\n",
    sg_compare$concordance$kappa))
cat(sprintf("  Concordant:      %d / Discordant: %d / Total: %d\n",
    sg_compare$concordance$n_agree,
    sg_compare$concordance$n_disagree,
    sg_compare$concordance$n_total))
```


# Bootstrap Bias Correction {#sec-bootstrap}

## Rationale

Cox model estimates from identified subgroups are **upwardly biased** due to
the selection process (subgroups are selected *because* they show extreme
effects). Bootstrap bias correction addresses this by:

1. Resampling with replacement
2. Re-running the entire ForestSearch algorithm
3. Computing bias terms from bootstrap vs. observed estimates
4. Applying infinitesimal jackknife variance estimation

## Running Bootstrap Analysis

```{r bootstrap}
t0 <- proc.time()

fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est        = fs,
  nb_boots      = NB_boot,
  show_three    = FALSE,
  details       = FALSE,
  parallel_args = list(
    plan         = "multisession",
    workers      = n_workers,
    show_message = TRUE
  )
)

plan("sequential")
timings$bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nBootstrap completed in",
    round(timings$bootstrap / 60, 1), "minutes\n")
```

## Bootstrap Summary and Diagnostics

```{r bootstrap-summary}
summaries <- summarize_bootstrap_results(
  sgharm       = fs$sg.harm,
  boot_results = fs_bc,
  create_plots = TRUE,
  est.scale    = "hr"
)
summaries$table
```

## Kaplan-Meier by Identified Subgroups

```{r fs-results-figure}
#| fig-width: 10
#| fig-height: 7
#| fig-cap: "Kaplan-Meier survival curves by identified subgroup"

km_result <- plot_sg_weighted_km(
  fs.est         = fs,
  outcome.name   = "time_months",
  event.name     = "status",
  treat.name     = "hormon",
  show.logrank   = FALSE,
  conf.int       = TRUE,
  by.risk        = 12,
  show.cox       = FALSE,
  show.cox.bc    = TRUE,
  fs_bc          = fs_bc,
  hr_bc_position = "topright"
)
```

::: {.callout-note}
`r figure_note(km_result, prefix = "**Note:** ", custom_text = "Medians [95% CI] for arms are un-adjusted.")`
:::


### Event Count Summary

Low event counts can lead to unstable HR estimates.

```{r event-summary}
event_summary <- summarize_bootstrap_events(fs_bc, threshold = 15)
```

### Bootstrap Diagnostics

```{r bootstrap-diagnostics}
summaries$diagnostics_table_gt
```

### Subgroup Agreement

How consistently does bootstrap identify the same subgroup?

```{r subgroup-agreement}
if (!is.null(summaries$subgroup_summary$original_agreement)) {
  summaries$subgroup_summary$original_agreement
}

if (!is.null(summaries$subgroup_summary$factor_presence)) {
  summaries$subgroup_summary$factor_presence
}
```

### Bootstrap Distributions

```{r bootstrap-plots}
#| fig-width: 10
#| fig-height: 4

if (!is.null(summaries$plots)) {
  summaries$plots$H_distribution + summaries$plots$Hc_distribution
}
```


# Cross-Validation {#sec-cv}

Cross-validation assesses the **stability** of the ForestSearch algorithm.
Two approaches are available.

## K-Fold Cross-Validation

```{r kfold-cv}
t0 <- proc.time()

fs_kfold <- forestsearch_tenfold(
  fs.est        = fs,
  sims          = Ksims_cv,
  Kfolds        = 10,
  details       = FALSE,
  parallel_args = list(
    plan         = "multisession",
    workers      = n_workers,
    show_message = FALSE
  )
)

plan("sequential")
timings$kfold <- (proc.time() - t0)["elapsed"]
metrics_tables <- cv_metrics_tables(fs_kfold)
metrics_tables
```

## Out-of-Bag (N-Fold) Cross-Validation

```{r oob-cv}
#| eval: true

t0 <- proc.time()

fs_OOB <- forestsearch_Kfold(
  fs.est        = fs,
  details       = FALSE,
  Kfolds        = round(nrow(df.analysis) / 100, 0),
  parallel_args = list(
    plan         = "multisession",
    workers      = n_workers,
    show_message = TRUE
  )
)

plan("sequential")
timings$oob <- (proc.time() - t0)["elapsed"]

cv_out <- forestsearch_KfoldOut(
  res     = fs_OOB,
  details = FALSE,
  outall  = TRUE
)

tables <- cv_summary_tables(cv_out)
tables$combined_table
tables$metrics_table
```


# Results Visualization {#sec-viz}

## Forest Plot

```{r forest-plot}
#| fig-width: 18
#| fig-height: 12
#| fig-cap: "Subgroup forest plot including identified subgroups"

subgroups <- list(
  age_gt65     = list(subset_expr = "age > 65",  name = "Age > 65",  type = "reference"),
  age_le65     = list(subset_expr = "age <= 65", name = "Age <= 65", type = "reference"),
  pgr_positive = list(subset_expr = "pgr > 0",   name = "PgR > 0",   type = "reference"),
  pgr_negative = list(subset_expr = "pgr <= 0",  name = "PgR <= 0",  type = "reference")
)

my_theme <- create_forest_theme(base_size = 24,
                                 footnote_fontsize = 17,
                                 cv_fontsize = 22)

result <- plot_subgroup_results_forestplot(
  fs_results = list(
    fs.est   = fs,
    fs_bc    = fs_bc,
    fs_OOB   = NULL,
    fs_kfold = fs_kfold
  ),
  df_analysis      = df.analysis,
  subgroup_list    = subgroups,
  outcome.name     = outcome.name,
  event.name       = event.name,
  treat.name       = treat.name,
  E.name           = "Hormonal",
  C.name           = "Chemo",
  ci_column_spaces = 25,
  xlog             = TRUE,
  theme            = my_theme
)

render_forestplot(result)
```

## KM Difference Plots: ITT and Subgroups

```{r KMdiffs}
#| fig-width: 8
#| fig-height: 6
#| fig-cap: "ITT (black) and subgroup KM differences with 95% confidence band."

ref_sgs <- list(
  age_young = list(subset_expr = "age < 65",  color = "brown"),
  age_old   = list(subset_expr = "age >= 65", color = "orange")
)

plot_km_band_forestsearch(
  df            = df.analysis,
  fs.est        = fs,
  ref_subgroups = ref_sgs,
  outcome.name  = outcome.name,
  event.name    = event.name,
  treat.name    = treat.name,
  draws_band    = 20
)
```


# Summary and Interpretation {#sec-summary}

## Key Findings

```{r summary-findings}
cat(strrep("=", 60), "\n")
cat("FORESTSEARCH ANALYSIS SUMMARY\n")
cat(strrep("=", 60), "\n\n")

cat("Dataset: GBSG (N =", nrow(df.analysis), ")\n")
cat("Outcome: Recurrence-free survival\n\n")

cat("ITT Analysis:\n")
cat("  HR (95% CI): 0.69 (0.54, 0.89)\n\n")

cat("Identified Subgroup (H):\n")
cat("  Definition:", paste(fs$sg.harm, collapse = " & "), "\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 0)))
cat("  Unadjusted HR:",
    sprintf("%.2f", exp(fs$grp.consistency$out_sg$result$hr[1])), "\n")

cat("\nComplement Subgroup (Hc):\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 1),
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 1)))
```

## Clinical Interpretation

The ForestSearch analysis identifies **estrogen receptor-negative (ER ≤ 0)**
patients as a subgroup with potential lack of benefit from hormonal therapy.

**Biological plausibility**: Tamoxifen is a selective estrogen receptor
modulator. Its efficacy depends on ER expression. The finding that ER-negative
patients may not benefit is consistent with:

- Mechanistic understanding of tamoxifen action
- Meta-analyses showing no tamoxifen benefit in ER-negative breast cancer
- Clinical guidelines recommending tamoxifen primarily for ER-positive tumors

**Caveats**:

1. This is an **exploratory** analysis requiring independent validation.
2. The bias-corrected estimates have wider confidence intervals.
3. Cross-validation metrics should be evaluated for algorithm stability.
4. The procedure-level false-positive rate at the operational
   thresholds (c1 = `r c1_paper`, c2 = `r sprintf("%.2f", c2_paper)`)
   is directly quantified in @sec-appendix-fpr below.


# Appendix: Direct FPR Calibration at the Operational Thresholds {#sec-appendix-fpr}

## Motivation {#sec-appendix-rationale}

The main analysis uses threshold pair
(c1 = `r c1_paper`, c2 = `r sprintf("%.2f", c2_paper)`) with
`pconsistency.threshold = `r pcons_val`` and
`stop_threshold = `r stop_thresh_val``.  A natural question follows
from any exploratory subgroup-identification analysis:

> Under the null hypothesis of *no treatment-by-subgroup interaction*,
> how often would ForestSearch spuriously identify **any** subgroup
> at these operational settings?

Both approaches used here generate null replicates from the **GBSG
parametric DGM** via `setup_gbsg_dgm(model = "null")`, consistent
with prior simulation experience using this DGM.  The two approaches
differ only in the **analysis wrapper** applied to each null
replicate:

- **Approach A — `fpr_calibration()`.** A single-purpose wrapper that
  runs `forestsearch()` under H0 at the analyst's exact threshold
  pair, returning a direct Monte Carlo FPR estimate with Wilson 95\%
  CI, plus the pair-specific implied L\_eff.  Builds null replicates
  via a user-supplied `null_fn`.
- **Approach B — `run_simulation_analysis()`.** The full simulation
  pipeline used in the León (2024) Table 1 reproduction documents.
  For each replicate it runs one or more of {FSl, FSlg, GRF, GRF60}
  and returns `any.H`, classification metrics, and effect estimates,
  feeding the standard `format_oc_results()` and
  `build_classification_table()` reporting helpers.

Both approaches are run below at the same operational configuration.
**Consistency between the two is the primary diagnostic** — large
disagreement would indicate a discrepancy in how the two wrappers
apply the ForestSearch pipeline.  Once consistency is confirmed, a
clean final version using Approach B alone is recommended because
it leverages the package's supported simulation API and produces
publication-ready operating-characteristics output.


## Analysis setup {#sec-appendix-setup}

```{r fpr-setup}
#| code-fold: false

# ── Parallel plan (same workers as main analysis) ─────────────────
plan(multisession, workers = n_workers)
cat("Workers:", n_workers, "\n")

# DGM-side fs_params are constructed in the DGM setup chunk below
# (fs_params_gbsg_dgm).  No permutation-side fs_params are needed.
```


## Approach A: `fpr_calibration()` wrapper {#sec-appendix-dgm}

We use `setup_gbsg_dgm(model = "null")` as the null DGM — parameters
calibrated to the GBSG AFT structure with treatment effect set to the
null and no treatment-by-subgroup interaction.  `simulate_from_dgm()`
then generates the replicates.

The null generator is constructed via a **closure factory** (rather
than as a plain anonymous function capturing globals by name) to
force the DGM S3 object and auxiliary parameters into a dedicated
enclosing environment.  This guarantees that `{future}` parallel
workers receive a self-contained function rather than one that
relies on lexical resolution in the knit session's `globalenv()`.

```{r fpr-dgm-setup}
#| code-fold: false

dgm_gbsg_null <- setup_gbsg_dgm(
  model   = "null",
  verbose = FALSE
)

# Null generator.
#
# We build null_fn_gbsg via a FACTORY FUNCTION so that dgm_gbsg_null,
# the target N, analysis_time, and cens_adjust are captured in a
# DEDICATED enclosing environment — NOT globalenv().  The {future}
# global-detection scanner exports enclosing environments that are
# not globalenv() by default, but objects living directly in
# globalenv() that are referenced only via lexical scope inside a
# closure are NOT always exported.  Wrapping construction in a
# factory forces every dependency (especially the S3 aft_dgm_flex
# object dgm_gbsg_null, which simulate_from_dgm() dispatches on) to
# live in the closure's own env and be auto-exported to workers.
#
# See the fpr_calibration_demo.qmd's null_fn_cox: in that document
# the DGM call succeeds at scale because the knit environment
# happens to serialise dgm_surv_null correctly; here we make the
# export explicit to eliminate the ambiguity.
make_null_fn <- function(dgm, n_target, analysis_time_, cens_adjust_) {
  force(dgm)
  force(n_target)
  force(analysis_time_)
  force(cens_adjust_)
  function(seed) {
    simulate_from_dgm(
      dgm,
      n             = n_target,
      analysis_time = analysis_time_,
      cens_adjust   = cens_adjust_,
      seed          = seed
    )
  }
}

null_fn_gbsg <- make_null_fn(
  dgm            = dgm_gbsg_null,
  n_target       = nrow(df.analysis),
  analysis_time_ = analysis_time_dgm,
  cens_adjust_   = cens_adjust_dgm
)

# Sanity: verify the factory-built closure captures what we expect.
cat("null_fn_gbsg enclosing environment:\n")
print(environment(null_fn_gbsg))
cat("Names in that environment:\n")
print(ls(envir = environment(null_fn_gbsg)))

# DGM-side fs_params.  Built to mirror the parameter set the package
# developer already verified works in fpr_calibration_demo.qmd's
# fs_cox list — specifically, we drop return_selected_cuts_only,
# cut_type, and showten_subgroups (which the demo did not set) and
# add is.RCT, by.risk, and vi.grf.min (which the demo did set).
# We retain the operational pcons / stop_threshold from the main
# GBSG analysis so the DGM-based FPR estimate is calibrated at the
# same operational setting used for subgroup identification.
fs_params_gbsg_dgm <- list(
  outcome.name           = "y_sim",
  event.name             = "event_sim",
  treat.name             = "treat_sim",
  id.name                = "id",
  confounders.name       = confounders_dgm,
  pconsistency.threshold = pcons_val,
  stop_threshold         = stop_thresh_val,
  sg_focus               = "hr",
  max_subgroups_search   = max_sgs_search_val,
  use_twostage           = TRUE,
  fs.splits              = fs_splits_val,
  n.min                  = n_min_val,
  d0.min                 = d0_min_val,
  d1.min                 = d1_min_val,
  maxk                   = maxk_val,
  by.risk                = 12L,
  vi.grf.min             = -0.2,
  use_lasso              = TRUE,
  use_grf                = TRUE,
  is.RCT                 = TRUE,
  details                = FALSE,
  quiet                  = TRUE,
  plot.sg                = FALSE,
  parallel_args          = list(plan = "sequential", workers = 1,
                                 show_message = FALSE)
)

cat(sprintf("DGM null ready | N = %d | analysis_time = %d months | prop_cens ~ %.2f\n",
            nrow(df.analysis), analysis_time_dgm, prop_cens_surv))
```

### DGM sanity check: single-replicate probe {#sec-appendix-dgm-probe}

Before running 500 replicates in parallel (where `foreach(.errorhandling = "pass")`
silently captures any error), we run a single DGM replicate inline and
inspect the result.  This surfaces any column-name mismatch, parameter
incompatibility, or package-loading issue in one place, with the full
R error visible if anything goes wrong.

```{r fpr-dgm-probe}
#| code-fold: false

# (1) Generate a single null replicate from the DGM.
probe_df <- null_fn_gbsg(seed = sim_seed_fpr + 2000L)

cat("DGM replicate N:", nrow(probe_df), "\n")
cat("DGM columns (first 20):",
    paste(head(names(probe_df), 20), collapse = ", "), "\n")
cat("Confounders requested:", paste(confounders_dgm, collapse = ", "), "\n")
missing_cols <- setdiff(confounders_dgm, names(probe_df))
if (length(missing_cols) > 0) {
  cat("** MISSING columns in DGM output: ",
      paste(missing_cols, collapse = ", "), " **\n")
} else {
  cat("All confounder columns present in DGM output.\n")
}
cat("Confounder types:\n")
print(sapply(probe_df[confounders_dgm], function(x)
  paste(class(x), collapse = "/")))
cat("Event rate:", sprintf("%.1f%%", 100 * mean(probe_df$event_sim)), "\n")

# (2) Run forestsearch on the single replicate and show any error.
probe_call <- c(
  list(df.analysis    = probe_df,
       hr.threshold   = c1_paper,
       hr.consistency = c2_paper,
       seedit         = sim_seed_fpr + 2000L),
  fs_params_gbsg_dgm
)

probe_fs <- tryCatch(do.call(forestsearch, probe_call),
                     error = function(e) e)

if (inherits(probe_fs, "error")) {
  cat("\n*** DGM single-replicate forestsearch() FAILED ***\n")
  cat("Error message:\n  ", conditionMessage(probe_fs), "\n", sep = "")
} else {
  cat("\nDGM single-replicate forestsearch() completed.\n")
  cat("  sg.harm:",
      if (length(probe_fs$sg.harm) > 0)
        paste(probe_fs$sg.harm, collapse = " & ")
      else "(none)", "\n")
}
```

### DGM parallel-worker probe: surface hidden errors {#sec-appendix-dgm-worker-probe}

The previous probe ran sequentially in the knit process and succeeded.
When we run the same call through `fpr_calibration()` with
`n_workers > 1`, every replicate fails silently because the internal
`foreach(.errorhandling = "pass")` captures but does not re-throw
errors.  To diagnose, we reproduce the parallel path here with 8
replicates and **print the first returned error object** verbatim.

Common root causes surfaced by this probe:

1. **`forestsearch` not installed in the worker's library path**.
   Parallel workers start fresh R processes and can only `library()`
   installed packages — `devtools::load_all()` state does not
   transfer.  Fix: `devtools::install()` before knitting.
2. **A closure-captured object is not exported to workers**.  Rare;
   `{future}` auto-detects globals, but unusual S3 objects sometimes
   slip through.  Fix: use `future.globals = TRUE` explicitly or
   refactor `null_fn_gbsg()` to build the DGM internally.
3. **Worker-specific package/library path misconfiguration** (e.g.
   user library not on `.libPaths()` inside the worker).

```{r fpr-dgm-worker-probe}
#| code-fold: false

suppressPackageStartupMessages({
  library(foreach)
  library(doFuture)
  library(future)
})

# Use the SAME plan that fpr_calibration would set internally.
future::plan("multisession", workers = min(8L, n_workers))

probe_batch <- tryCatch(
  foreach::foreach(
    i = seq_len(8L),
    .errorhandling = "pass",
    .options.future = list(packages = c("forestsearch"), seed = TRUE)
  ) %dofuture% {
    future::plan("sequential")

    # Step A: does the worker see the forestsearch package?
    fs_available <- "forestsearch" %in% loadedNamespaces() ||
                    requireNamespace("forestsearch", quietly = TRUE)

    # Step B: can the worker build a null replicate?
    null_df <- tryCatch(
      null_fn_gbsg(seed = sim_seed_fpr + 2000L + i),
      error = function(e) e
    )
    if (inherits(null_df, "error")) {
      return(list(stage = "null_fn_gbsg",
                  fs_available = fs_available,
                  msg = conditionMessage(null_df)))
    }

    # Step C: can the worker run forestsearch?
    fs_args <- c(list(df.analysis    = null_df,
                      hr.threshold   = c1_paper,
                      hr.consistency = c2_paper,
                      seedit         = sim_seed_fpr + 2000L + i),
                 fs_params_gbsg_dgm)
    fs_out <- tryCatch(do.call(forestsearch::forestsearch, fs_args),
                       error = function(e) e)
    if (inherits(fs_out, "error")) {
      return(list(stage = "forestsearch",
                  fs_available = fs_available,
                  msg = conditionMessage(fs_out)))
    }

    list(stage = "ok",
         fs_available = fs_available,
         found = as.integer(!is.null(fs_out$sg.harm) &&
                              length(fs_out$sg.harm) > 0))
  },
  error = function(e) e
)

future::plan("sequential")

# Report results
if (inherits(probe_batch, "error")) {
  cat("*** foreach loop itself errored ***\n",
      conditionMessage(probe_batch), "\n", sep = "")
} else {
  # Classify each result
  stages <- sapply(probe_batch, function(x) {
    if (inherits(x, "error")) "captured-error"
    else if (is.list(x) && !is.null(x$stage)) x$stage
    else "unknown"
  })
  cat("Parallel-worker probe (8 replicates):\n")
  print(table(stages, useNA = "ifany"))

  # Show first captured error verbatim
  err_idx <- which(stages == "captured-error")
  if (length(err_idx) > 0) {
    err <- probe_batch[[err_idx[1]]]
    cat("\nFirst captured foreach-error (",
        length(err_idx), " of 8):\n  ",
        conditionMessage(err), "\n", sep = "")
  }
  # Show first non-ok structured result
  bad_idx <- which(stages %in% c("null_fn_gbsg", "forestsearch"))
  if (length(bad_idx) > 0) {
    bad <- probe_batch[[bad_idx[1]]]
    cat("\nFirst structured failure (",
        length(bad_idx), " of 8):\n",
        "  stage:         ", bad$stage, "\n",
        "  fs_available:  ", bad$fs_available, "\n",
        "  error message: ", bad$msg, "\n", sep = "")
  }
  ok_idx <- which(stages == "ok")
  if (length(ok_idx) > 0) {
    cat("\n", length(ok_idx), " of 8 replicates completed successfully.\n",
        sep = "")
  }
}
```

### Tier 1 at paper defaults (c1 = `r c1_paper`, c2 = `r sprintf("%.2f", c2_paper)`)

```{r fpr-dgm-paper}
#| code-fold: false

t0 <- proc.time()[3]
cal_dgm_paper <- fpr_calibration(
  c1           = c1_paper,
  c2           = c2_paper,
  null_fn      = null_fn_gbsg,
  n_sample     = nrow(df.analysis),
  fs_params    = fs_params_gbsg_dgm,
  nsims        = nsims_tier1_fpr,
  nsims_verify = nsims_verify_fpr,
  n_min        = n_min_val,
  outcome_type_approx = "survival",
  prop_cens    = prop_cens_surv,
  sim_seed     = sim_seed_fpr + 2000L,
  n_workers    = n_workers
)
print(cal_dgm_paper)
timings$fpr_dgm_paper <- proc.time()[3] - t0
```


## Approach B: `run_simulation_analysis()` pipeline {#sec-appendix-runsim}

Approach B uses the package's full simulation pipeline as configured
in the León (2024) Table 1 reproduction documents.  For each null
replicate, `run_simulation_analysis()` simulates from the DGM and
runs one or more analysis variants; the per-replicate results are
collected via `collect_results()` and summarised with
`format_oc_results()`.  This is the supported path and produces the
publication-ready operating-characteristics output used in the
package's internal simulation studies.

### Worker pool reset {#sec-appendix-runsim-reset}

Before a second large foreach batch, reset the worker pool to
prevent memory accumulation from the Approach A calls (a pattern
directly from `reproduce_table1_m1.qmd`).

```{r runsim-reset}
#| code-fold: false
plan(sequential)
gc(verbose = FALSE)
Sys.sleep(1)
plan(multisession, workers = n_workers)
cat(sprintf("Workers reset: %d cores\n", n_workers))
```

### ForestSearch parameter list (Approach B side) {#sec-appendix-runsim-params}

```{r runsim-params}
#| code-fold: false

# fs_params for run_simulation_analysis().  Structurally identical to
# fs_params_gbsg_dgm used in Approach A, but formatted in the
# reproduce_table1 style (seedit is embedded in the list rather than
# passed externally by fpr_calibration).  We run the FSlg variant
# (LASSO + GRF), which is closest to the main analysis configuration.
fs_params_runsim <- list(
  outcome.name           = "y_sim",
  event.name             = "event_sim",
  treat.name             = "treat_sim",
  id.name                = "id",
  hr.threshold           = c1_paper,
  hr.consistency         = c2_paper,
  pconsistency.threshold = pcons_val,
  stop_threshold         = stop_thresh_val,
  sg_focus               = "hr",
  max_subgroups_search   = max_sgs_search_val,
  use_twostage           = TRUE,
  fs.splits              = fs_splits_val,
  n.min                  = n_min_val,
  d0.min                 = d0_min_val,
  d1.min                 = d1_min_val,
  maxk                   = maxk_val,
  vi.grf.min             = -0.2,
  seedit                 = sim_seed_fpr,
  use_lasso              = TRUE,
  use_grf                = TRUE,       # FSlg: LASSO + GRF
  tune_grf               = FALSE,
  is.RCT                 = TRUE,
  details                = FALSE,
  quiet                  = TRUE,
  plot.sg                = FALSE,
  parallel_args          = list(
    plan = "sequential", workers = 1, show_message = FALSE
  )
)

# Cox formula (treatment-only, per paper Section 2)
cox_formula_adj_runsim <- survival::Surv(y_sim, event_sim) ~ treat_sim
```

### Null simulation loop {#sec-appendix-runsim-loop}

```{r runsim-null}
#| code-fold: false

t0 <- proc.time()[3]

results_null_raw <- foreach(
  i = seq_len(nsims_tier1_fpr),
  .errorhandling = "pass",
  .options.future = list(packages = c("forestsearch"), seed = TRUE)
) %dofuture% {
  future::plan("sequential")

  r_fslg <- tryCatch(
    run_simulation_analysis(
      sim_id           = i,
      dgm              = dgm_gbsg_null,
      n_sample         = nrow(df.analysis),
      confounders_base = confounders_dgm,
      n_add_noise      = 0L,
      run_fs           = TRUE,
      run_fs_grf       = FALSE,
      run_grf          = FALSE,
      fs_params        = fs_params_runsim,
      cox_formula_adj  = cox_formula_adj_runsim,
      analysis_time    = analysis_time_dgm,
      cens_adjust      = cens_adjust_dgm,
      seed_base        = sim_seed_fpr + 2000L,
      verbose          = FALSE
    ),
    error = function(e) NULL
  )
  if (!is.null(r_fslg)) r_fslg$analysis <- "FSlg"

  data.table::rbindlist(
    Filter(Negate(is.null), list(r_fslg)),
    fill = TRUE
  )
}

plan("sequential")
timings$fpr_runsim <- proc.time()[3] - t0

results_null <- forestsearch::collect_results(results_null_raw, "H0 null")
n_failed_null <- attr(results_null, "n_failed")
cat(sprintf("Approach B: %d sims completed in %.1f min\n",
            nsims_tier1_fpr, timings$fpr_runsim / 60))
if (n_failed_null > 0L)
  cat(sprintf("  *** %d of %d replicates failed ***\n",
              n_failed_null, nsims_tier1_fpr))
```

### Approach B operating characteristics {#sec-appendix-runsim-oc}

```{r runsim-oc}
# Standard operating-characteristics table from reproduce_table1_m1.qmd.
# Under H0, any.H is the FPR; sens / ppv are null by construction
# (the unconditional averaging convention).
format_oc_results(
  results_null,
  analyses = "FSlg",
  metrics  = c("detection", "classification"),
  digits   = 3,
  title    = "GBSG Null — Approach B (run_simulation_analysis)",
  subtitle = sprintf(
    "N = %d, %d sims, θ†(ITT) ≈ %.2f",
    nrow(df.analysis), nsims_tier1_fpr, dgm_gbsg_null$hr_causal
  ),
  subgroup_notation = "harm"
)
```

### Extract single FPR number for cross-comparison {#sec-appendix-runsim-fpr}

```{r runsim-fpr-extract}
#| code-fold: false

# Extract the FSlg any.H rate for direct comparison with Approach A.
rn_fslg <- results_null[results_null$analysis == "FSlg", ]
n_reps  <- nrow(rn_fslg)
n_det   <- sum(rn_fslg$any.H == 1L, na.rm = TRUE)
fpr_B   <- n_det / n_reps

# Wilson 95% CI (same formula fpr_calibration uses internally)
.wilson_ci <- function(x, n, z = 1.96) {
  if (n == 0) return(c(NA_real_, NA_real_))
  p <- x / n
  denom  <- 1 + z^2 / n
  centre <- (p + z^2 / (2 * n)) / denom
  margin <- z * sqrt(p * (1 - p) / n + z^2 / (4 * n^2)) / denom
  c(max(0, centre - margin), min(1, centre + margin))
}
fpr_B_ci <- .wilson_ci(n_det, n_reps)

cat(sprintf("Approach B (FSlg) FPR:  %.1f%%  [95%% CI: %.1f%% -- %.1f%%]\n",
            100 * fpr_B, 100 * fpr_B_ci[1], 100 * fpr_B_ci[2]))
cat(sprintf("  based on %d / %d successful replicates\n", n_det, n_reps))
```


## Results {#sec-appendix-consistency}

```{r fpr-consistency-table}
#| tbl-cap: "DGM-based FPR calibration at the GBSG operational thresholds. Approach A = fpr_calibration() wrapper. Approach B = run_simulation_analysis() pipeline (FSlg variant). Close agreement validates both approaches; once satisfied, the final version will use Approach B alone."

# Shared row builder.  Accepts either an fpr_calibration-like object
# (Approach A) or a list with fpr, fpr_ci, nsims, P1, L_eff_implied
# fields that we fabricate for Approach B from results_null.
make_row <- function(res, approach_label) {
  fpr_corr <- if (!is.na(res$L_eff_implied))
    1 - (1 - res$P1)^res$L_eff_implied else NA_real_
  data.table(
    Approach          = approach_label,
    `Cal sims`        = res$nsims,
    `Sim FPR`         = sprintf("%.1f%%", 100 * res$fpr),
    `95% CI`          = sprintf("[%.1f%%, %.1f%%]",
                                 100 * res$fpr_ci[1],
                                 100 * res$fpr_ci[2]),
    `P1 (uncorrected)` = if (!is.na(res$P1))
                          sprintf("%.1f%%", 100 * res$P1) else NA_character_,
    `Implied L_eff`    = if (!is.na(res$L_eff_implied))
                          round(res$L_eff_implied, 3) else NA_real_,
    `Corrected FPR`    = if (!is.na(fpr_corr))
                          sprintf("%.1f%%", 100 * fpr_corr) else NA_character_,
    fpr_num            = res$fpr,
    fpr_corr_num       = fpr_corr
  )
}

# Build a parallel result object for Approach B, mimicking the
# fpr_calibration return shape so the table builder can reuse make_row.
# P1 and implied L_eff come from the Approach A calibration (they are
# properties of the threshold pair + d_eff, not of the wrapper), so
# they are reported once and shared.
cal_runsim_paper <- list(
  fpr           = fpr_B,
  fpr_ci        = fpr_B_ci,
  nsims         = n_reps,
  P1            = cal_dgm_paper$P1,
  L_eff_implied = if (!is.na(fpr_B) && !is.na(cal_dgm_paper$P1) &&
                       fpr_B > 0 && fpr_B < 1 &&
                       cal_dgm_paper$P1 > 0 && cal_dgm_paper$P1 < 1)
    log(1 - fpr_B) / log(1 - cal_dgm_paper$P1) else NA_real_
)

consistency_tbl <- rbind(
  make_row(cal_dgm_paper,     "Approach A — fpr_calibration()"),
  make_row(cal_runsim_paper,  "Approach B — run_simulation_analysis()")
)

gt(consistency_tbl) |>
  tab_header(
    title = sprintf(
      "GBSG: DGM-based FPR Calibration | c1 = %.2f, c2 = %.2f | pcons = %.2f, stop = %.2f",
      c1_paper, c2_paper, pcons_val, stop_thresh_val),
    subtitle = sprintf("N = %d | n.min = %d | up to %d calibration sims per row",
                        nrow(df.analysis), n_min_val, nsims_tier1_fpr)
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `Corrected FPR`,
                           rows = !is.na(fpr_corr_num) &
                                  fpr_corr_num <= fpr_target_surv)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = `Corrected FPR`,
                           rows = !is.na(fpr_corr_num) &
                                  fpr_corr_num > fpr_target_surv)
  ) |>
  cols_hide(columns = c(fpr_num, fpr_corr_num)) |>
  tab_source_note(sprintf(
    paste(
      "Both rows simulate from setup_gbsg_dgm(model = 'null').",
      "Approach A uses the fpr_calibration() wrapper;",
      "Approach B uses run_simulation_analysis() (FSlg variant).",
      "Corrected FPR = 1-(1-P1)^L_eff_implied using the pair-specific L_eff.",
      "Green shading: corrected FPR <= target %.0f%%.",
      "Red shading: corrected FPR > target."),
    100 * fpr_target_surv))
```


## Detection curves at the operational threshold pair {#sec-appendix-detection}

The detection curve shows, for any true hazard ratio $\theta$, the
probability that ForestSearch detects a subgroup with that true effect.
At $\theta = 1$ (H0) the curve passes through the direct simulation FPR
estimate by construction.

```{r plot-fpr-calibration-fn}
#| code-fold: true

# Local wrapper (to be exported from forestsearch in a future release).
plot_fpr_calibration <- function(cal_result,
                                  theta_seq    = seq(0.5, 4.0, by = 0.05),
                                  fpr_target   = NULL,
                                  label        = NULL,
                                  effect_label = "Effect ratio (theta)") {
  stopifnot(inherits(cal_result, "fpr_calibration"))

  c1       <- cal_result$c1
  c2       <- cal_result$c2
  d_eff    <- cal_result$d_eff
  L_eff    <- cal_result$L_eff_implied
  fpr_hat  <- cal_result$fpr
  ci_lo    <- cal_result$fpr_ci[1]
  ci_hi    <- cal_result$fpr_ci[2]
  N        <- cal_result$N
  nsims    <- cal_result$nsims

  p1_vec <- sapply(theta_seq, function(th)
    tryCatch(
      compute_detection_probability_glm(
        theta = th, d_eff = d_eff, c1 = c1, c2 = c2,
        effect_scale = "ratio", method = "monte_carlo"),
      error = function(e) NA_real_))

  p_corr_vec <- if (!is.na(L_eff))
    1 - (1 - p1_vec)^L_eff
  else rep(NA_real_, length(theta_seq))

  curve_dt <- data.frame(
    theta = rep(theta_seq, 2),
    prob  = c(p1_vec, p_corr_vec),
    curve = rep(c("Per-subgroup P1 (correct subgroup)",
                  "Corrected (any detection)"),
                each = length(theta_seq))
  )

  sim_pt <- data.frame(
    theta = 1, prob = fpr_hat, ci_lo = ci_lo, ci_hi = ci_hi,
    curve = "Direct simulation (theta=1)")

  sub_txt <- sprintf(
    "c1=%.2f, c2=%.2f | d_eff=%.1f | L_eff(implied)=%.3f | N=%d | %d sims",
    c1, c2, d_eff,
    if (!is.na(L_eff)) L_eff else 0,
    N, nsims)
  if (!is.null(label)) sub_txt <- paste0(label, " | ", sub_txt)

  p <- ggplot(curve_dt, aes(x = theta, y = prob,
                             colour = curve, linetype = curve)) +
    geom_line(linewidth = 0.9, na.rm = TRUE) +
    geom_vline(xintercept = 1, linetype = "dotted",
               colour = "grey60", linewidth = 0.4) +
    geom_point(data = sim_pt, aes(x = theta, y = prob),
               colour = "steelblue", size = 4, shape = 18,
               inherit.aes = FALSE) +
    geom_errorbar(data = sim_pt,
                  aes(x = theta, ymin = ci_lo, ymax = ci_hi),
                  colour = "steelblue", width = 0.04, linewidth = 0.7,
                  inherit.aes = FALSE) +
    annotate("text", x = 1.05, y = fpr_hat,
             label = sprintf("Sim FPR = %.1f%%", 100 * fpr_hat),
             hjust = 0, size = 3.2, colour = "steelblue") +
    scale_colour_manual(values = c(
      "Per-subgroup P1 (correct subgroup)" = "grey50",
      "Corrected (any detection)"          = "firebrick")) +
    scale_linetype_manual(values = c(
      "Per-subgroup P1 (correct subgroup)" = "dashed",
      "Corrected (any detection)"          = "solid")) +
    scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
    labs(x        = effect_label,
         y        = "Detection probability",
         title    = sprintf("Detection Curve: c1 = %.2f, c2 = %.2f", c1, c2),
         subtitle = sub_txt,
         colour   = NULL, linetype = NULL) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top",
          plot.title    = element_text(face = "bold"),
          plot.subtitle = element_text(size = 9, colour = "grey40"))

  if (!is.null(fpr_target)) {
    p <- p +
      geom_hline(yintercept = fpr_target, linetype = "dashed",
                 colour = "darkgreen", linewidth = 0.5) +
      annotate("text",
               x = min(theta_seq) + 0.05, y = fpr_target + 0.01,
               label = sprintf("FPR target = %.0f%%", 100 * fpr_target),
               hjust = 0, size = 3, colour = "darkgreen")
  }
  p
}
```

```{r fpr-detection-curves}
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "DGM-based detection curve at the paper-default threshold pair. Diamond = direct-simulation FPR with 95% CI at theta=1 (H0). Dashed grey = per-subgroup P1; solid red = corrected detection probability using the pair-specific implied L_eff."

plot_fpr_calibration(
  cal_dgm_paper,
  fpr_target   = fpr_target_surv,
  effect_label = "Hazard ratio (theta)",
  label        = sprintf("DGM | N = %d | prop_cens ~ %.2f | %d sims",
                          nrow(df.analysis), prop_cens_surv, nsims_tier1_fpr)
)
```


## Optional: Tier 2 verification at paper defaults {#sec-appendix-tier2}

If Tier 1 suggests the corrected FPR is close to the target (and the
operational thresholds will be used for interpretation), a Tier 2
high-precision DGM run is worthwhile.  This typically takes 1–3 hours.
Set `nsims_tier2_fpr <- 0L` in the setup chunk to skip.

```{r fpr-tier2}
#| code-fold: false

if (nsims_tier2_fpr > 0L) {
  t0 <- proc.time()[3]
  cal_dgm_paper_t2 <- fpr_calibration(
    c1           = c1_paper,
    c2           = c2_paper,
    null_fn      = null_fn_gbsg,
    n_sample     = nrow(df.analysis),
    fs_params    = fs_params_gbsg_dgm,
    nsims        = nsims_tier2_fpr,
    nsims_verify = 0L,
    n_min        = n_min_val,
    outcome_type_approx = "survival",
    prop_cens    = prop_cens_surv,
    sim_seed     = sim_seed_fpr + 50000L,
    n_workers    = n_workers
  )
  print(cal_dgm_paper_t2)
  timings$fpr_dgm_tier2 <- proc.time()[3] - t0
} else {
  message("Tier 2 skipped (nsims_tier2_fpr = 0L). ",
          "Set nsims_tier2_fpr to a positive integer in the setup chunk to run.")
}
```


## Interpretation {#sec-appendix-interp}

**Reading the result.**  The two rows in @sec-appendix-consistency
report the directly simulated FPR at the main-analysis operating
point: threshold pair (`r c1_paper`, `r sprintf("%.2f", c2_paper)`)
with `pconsistency.threshold = `r pcons_val`` and
`stop_threshold = `r stop_thresh_val``.  Row 1 uses the
`fpr_calibration()` wrapper (Approach A); row 2 uses
`run_simulation_analysis()` → `collect_results()` →
`format_oc_results()` (Approach B), the pipeline used in the package's
Table 1 reproduction documents.  The 95\% CI reflects Monte Carlo
uncertainty from up to `r nsims_tier1_fpr` DGM replicates (Wilson
interval).

**Consistency between approaches.**  Because both rows simulate from
the same DGM (`setup_gbsg_dgm(model = "null")`) at the same N, with
the same `fs_params` content routed to the same
`forestsearch()` engine, the Sim FPR estimates should agree up to
Monte Carlo noise.  Close agreement (overlapping CIs) validates both
approaches.  Residual differences are attributable to (i) different
random seed streams (Approach A uses `sim_seed + i`; Approach B uses
`seed_base + sim_id`), (ii) different treatment-assignment handling
(`fpr_calibration` uses the DGM's own treatment; `run_simulation_analysis`
assigns treatment inside `simulate_from_dgm` with
`draw_treatment = TRUE`), and (iii) small differences in what counts
as a "detection" in edge cases.

**Implied L\_eff**.  P1 is the per-subgroup asymptotic probability
that a single pre-specified subgroup passes screening + consistency
under H0; `L_eff_implied = log(1 − FPR) / log(1 − P1)` is the
pair-specific effective multiplicity — it replaces the universal
power-law model and does not assume threshold-independence.  `Corrected
FPR = 1 − (1 − P1)^L_eff_implied` is tautological at the calibration
pair by construction; the **Sim FPR** is the operational figure to
quote.  For survival with events-based filters (`d0.min`, `d1.min`),
L\_eff is typically close to or below 1 when the filters prune the
candidate pool tightly; values well above 1 indicate the candidate
pool is larger than a single test.

**Interpreting the identified ER ≤ 0 subgroup.**  The FPR figure
above is the procedure-level Type-I error rate: the probability that
ForestSearch identifies *any* subgroup under H0 at these operational
settings.  It is not subgroup-specific and should be interpreted as a
property of the analysis procedure rather than of the specific
subgroup found.  A high calibrated FPR indicates that the identified
subgroup (ER ≤ 0) warrants particularly cautious interpretation and
independent validation — the biological plausibility of the tamoxifen
/ ER-status mechanism supports the finding, but the statistical
evidence alone is weaker than the nominal split-sample HR
consistency suggests.

**Migration plan.**  Once consistency between Approach A and
Approach B is confirmed across a few runs, the recommended clean
version of this appendix should retain only Approach B: it (i) uses
the package's supported simulation API, (ii) produces
publication-ready operating-characteristics output via
`format_oc_results()`, (iii) naturally extends to multiple analysis
variants (FSl, FSlg, GRF, GRF60), and (iv) avoids the need for a
separate `fpr_calibration()` wrapper.  Approach A remains useful as
a quick single-number FPR check when only one threshold pair is of
interest.


# Computational Timing {#sec-timing}

```{r timing-summary}
# Local null-coalescing operator (the one in forestsearch is unexported)
`%||%` <- function(a, b) if (!is.null(a)) a else b

timings$total <- (proc.time() - t_vignette_start)["elapsed"]

# Sum FPR sub-components into a single row for display
timings$fpr_total <- sum(c(
  timings$fpr_dgm_paper  %||% 0,
  timings$fpr_runsim     %||% 0,
  timings$fpr_dgm_tier2  %||% 0
))

timing_df <- data.frame(
  Analysis = c("GRF", "ForestSearch", "Bootstrap",
               "kfold CV", "OOB CV",
               "FPR calibration (appendix)", "Total"),
  Seconds  = c(
    timings$grf,
    timings$forestsearch,
    timings$bootstrap,
    timings$kfold,
    timings$oob,
    if (timings$fpr_total > 0) timings$fpr_total else NA_real_,
    timings$total
  )
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(title = "Computational Timing") |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1,
             rows = !is.na(Seconds)) |>
  cols_label(
    Analysis = "Component",
    Seconds  = "Time (sec)",
    Minutes  = "Time (min)"
  ) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_body(rows = Analysis == "Total"))
```


# References {#sec-refs}

León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024).
"Exploratory subgroup identification in the heterogeneous Cox model:
A relatively simple procedure." *Statistics in Medicine*.
DOI: [10.1002/sim.10163](https://doi.org/10.1002/sim.10163)


# Session Info {#sec-session}

```{r session-info}
plan(sequential)
sessionInfo()
```
