---
title: "DINA Cox Survival Analysis — GBSG"
subtitle: "DINA analog to the ForestSearch GBSG Cox analysis"
author: "forestsearch v0.2.0"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

timings <- list()
t_doc_start <- proc.time()
```

# Introduction

This vignette is the DINA analog of the *Cox* branch of
`gbsg_survival_cox_poisson.qmd`. It applies the **DINA**
(Doubly-robust Iterative Neyman-Orthogonal) estimator to the
German Breast Cancer Study Group trial via the partial-likelihood
Cox extension.

The full inference toolkit applied here:

* Cross-fit point estimation via `dina()` (sandwich variance).
* Per-covariate linear-projection coefficient table — the DINA
  parametric analog of the best linear projection (BLP) reported
  in Cui et al. (2023, *JRSS-B* 85:179–211) Table 5. See the
  Appendix for the methodological background.
* Threshold-search harm subgroup identification via
  `dina_subgroup()` with sandwich-variance Wald CI.
* Selection-adjusted bootstrap percentile CI on the fixed-subgroup
  linear functional via `dina_subgroup_bootstrap()`.
* Stability diagnostics: selection-frequency table,
  `n_subgroup_ci`, `threshold_ci`.
* Feeding DINA's per-covariate Pareto frontier into
  `forestsearch()` as screening candidates, via `dina_frontier()`
  and the `use_dina` gate (see *DINA Screening Candidates for
  ForestSearch*).

The intent is to mirror the FS doc's Cox-analysis flow on the same
data with the same parameter conventions, *not* to repeat the FS
doc's downstream artifacts (CV, KM by subgroup, forest plots). Those
exist in the FS doc; this document focuses on the DINA inference
pipeline.

## Deferred: Poisson + offset analog

The FS doc also runs a **Poisson + offset** branch as the natural
exponential-family approximation to Cox. The analogous DINA branch
requires offset support in `dina_fit()`, which is not yet
implemented — `dina_fit()`'s Poisson family treats `Y` as raw counts
without person-time adjustment. Extending `dina_fit()` to accept an
`offset = ` argument is a well-scoped enhancement (offset slicing
through cross-fitting, baseline GLM fits, and step-2; the math is
clean because the treatment-effect contrast is invariant under a
shared offset). It's deferred to a future session.

The ITT-level Cox-vs-Poisson sanity check below uses plain
`survival::coxph()` and `stats::glm()` (no DINA) and is still
informative as a *population-level* diagnostic of the
proportional-hazards approximation, even though the subgroup-level
DINA Poisson + offset analysis awaits the extension.

# Setup

## Load Required Packages

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(doFuture)

theme_set(theme_minimal(base_size = 12))
```

## Analysis Parameters

```{r analysis-params}
#| code-fold: false

# ─── Shared parameters ───────────────────────────────────────────────────
n_cores         <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))
confounders     <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
id_col          <- "id"
treat_col       <- "hormon"

# ─── DINA Cox outcome columns ────────────────────────────────────────────
cox_outcome     <- "time_months"
cox_status      <- "status"

# ─── DINA tuning ─────────────────────────────────────────────────────────
dina_n_folds    <- 2L
dina_seed       <- 8316951L

# ─── Harm threshold (natural-parameter scale: log-HR for Cox) ────────────
# DINA's Cox target is log-HR (the natural parameter of the partial-
# likelihood model), NOT risk difference.  So m_diff is on the log-HR
# scale.  First-pass exploration uses 0.0 -- "any harm signal", HR >= 1.0.
# For clinically-meaningful harm matching the FS hr.threshold = 1.25,
# use log(1.25) approx 0.2231 in a follow-up run.
m_diff          <- log(0.80)
n_min           <- 60L               # matches FS n.min

# ─── Bootstrap ───────────────────────────────────────────────────────────
NB              <- 500L              # bootstrap iterations (use 1000+ for production)
boot_seed       <- 8316951L

# ─── Confidence level ────────────────────────────────────────────────────
alpha           <- 0.05
```

# Data: German Breast Cancer Study Group Trial

```{r data-setup}
df.analysis <- gbsg

df.analysis <- within(df.analysis, {
  id          <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3      <- as.integer(ifelse(grade == "3", 1, 0))
})

cat("Sample size:    ", nrow(df.analysis), "\n")
cat("Events:         ", sum(df.analysis[[cox_status]]),
    sprintf(" (%.1f%%)\n",
            100 * mean(df.analysis[[cox_status]])))
cat("Covariates:     ", paste(confounders, collapse = ", "), "\n")
cat("Cox outcome:    ", cox_outcome, "(+ status: ", cox_status, ")\n", sep = "")
```

# ITT-Level Cox vs. Poisson + Offset Sanity Check

A *population-level* sanity check that the Poisson + offset
approximation lines up with Cox at the ITT level on this dataset.
This uses plain `survival::coxph()` and `stats::glm()` (no DINA),
so it does not depend on the deferred DINA offset extension. The
purpose is to flag any gross mismatch at the population level that
would already cast doubt on a downstream Cox-vs-Poisson alignment
analysis at the subgroup level.

```{r itt-sanity}
df.analysis$time_months_log <- log(pmax(df.analysis$time_months, 1e-6))

cox_itt <- survival::coxph(
  survival::Surv(time_months, status) ~ hormon,
  data = df.analysis
)
pois_itt <- stats::glm(
  status ~ hormon,
  family = stats::poisson(),
  offset = df.analysis$time_months_log,
  data = df.analysis
)

itt_table <- data.table(
  Quantity = c("Cox  HR",
               "Poisson IRR",
               "Absolute difference"),
  Value    = c(
    sprintf("%.4f", exp(stats::coef(cox_itt))),
    sprintf("%.4f", exp(stats::coef(pois_itt)[["hormon"]])),
    sprintf("%.4f", abs(exp(stats::coef(cox_itt)) -
                        exp(stats::coef(pois_itt)[["hormon"]])))
  )
)

itt_table |>
  gt::gt() |>
  gt::tab_header(
    title    = "ITT-level Cox vs. Poisson + offset",
    subtitle = "Population-level proportional-hazards sanity check"
  )
```

# DINA Cox Analysis

## Fit cross-fit DINA Cox model

```{r dina-cox-fit}
t0 <- proc.time()

fit_dina_cox <- dina(
  df         = df.analysis,
  outcome    = cox_outcome,
  treatment  = treat_col,
  covariates = confounders,
  family     = "cox",
  status     = cox_status,
  n_folds    = dina_n_folds,
  seed       = dina_seed
)

timings$dina_cox_fit <- (proc.time() - t0)[["elapsed"]]
cat("DINA Cox fit completed in",
    round(timings$dina_cox_fit, 1), "seconds\n")
```

## Linear projection of the CATE (BLP analog)

DINA's coefficient table on the log-HR scale, paired with
sandwich-variance Wald CIs, is the parametric counterpart of the
best linear projection (BLP) reported by Cui et al. (2023)
Table 5. Under linearity of the CATE in the chosen covariates,
the DINA β coincides asymptotically with the BLP β; under
non-linearity, the DINA β is the parametric MLE under a
(potentially misspecified) linear model while the BLP β is the
L²-best linear approximation. The two should agree closely in
practice; see Appendix A for the detailed methodological
comparison.

The table below reports per-covariate β̂ with 95% Wald CIs on
both the log-HR (natural-parameter) scale and the exponentiated
HR scale.

```{r dina-cox-blp-table}
coef_dina <- stats::coef(fit_dina_cox)
se_dina   <- sqrt(diag(stats::vcov(fit_dina_cox)))
z_975     <- stats::qnorm(0.975)

ci_lo_log <- coef_dina - z_975 * se_dina
ci_hi_log <- coef_dina + z_975 * se_dina

blp_tbl <- data.table(
  Covariate            = names(coef_dina),
  `beta-hat (log-HR)`  = sprintf("%.3f", coef_dina),
  `Std. Error`         = sprintf("%.3f", se_dina),
  `95% Wald CI (log-HR)` = sprintf("[%.3f, %.3f]",
                                    ci_lo_log, ci_hi_log),
  `exp(beta-hat) (HR)` = sprintf("%.3f", exp(coef_dina)),
  `95% Wald CI (HR)`   = sprintf("[%.3f, %.3f]",
                                  exp(ci_lo_log), exp(ci_hi_log))
)

blp_tbl |>
  gt::gt() |>
  gt::tab_header(
    title    = gt::md("**DINA coefficient table -- parametric BLP analog**"),
    subtitle = gt::md(sprintf(
      "Per-covariate linear-projection coefficients on the log-HR scale;
       n = %d, %d cross-fitting folds, sandwich variance",
      nrow(df.analysis), dina_n_folds))
  ) |>
  gt::tab_source_note(
    source_note = gt::md(
      "*Coefficients are the DINA estimate of the linear functional
       tau(x) = beta_0 + x' beta on the log-HR scale, where the natural
       parameter for Cox PH is the log hazard. Under linearity these
       correspond asymptotically to the Cui et al. (2023) Table 5 best
       linear projection coefficients (Semenova & Chernozhukov 2021).
       Wald CIs use the cross-fit sandwich variance from `vcov(fit_dina_cox)`,
       which reduces to the Cox partial-likelihood information variance for
       the Cox family.*"
    )
  )
```

## Subgroup search (sandwich Wald CI)

### Scale: log-HR for Cox

DINA targets the natural-parameter contrast
$\tau(x) = \beta_0 + x^\top \beta_{1:d}$, which lies on the
natural-parameter scale of the family. The interpretation
therefore differs across families:

| Family | Natural-parameter scale | Interpretation of $\tau(x)$ |
|---|---|---|
| Gaussian | mean | mean difference (i.e., "risk difference" in the colloquial sense) |
| Binomial | logit | log odds ratio |
| Poisson | log rate | log incidence rate ratio |
| Cox | log hazard | log hazard ratio |

For the Cox family, the natural parameter is the log hazard (the
linear predictor of the partial-likelihood model), so `m_diff` is
on the **log-HR scale** — not a risk difference. The "DIfference in
NAtural parameters" name is precise: it's a difference of *natural
parameters*, and what those parameters mean depends on the family's
GLM link.

### Threshold: 0.0 as a permissive first pass

`m_diff = 0.0` asks for "the largest subgroup with mean
log-HR $\geq 0$", equivalently HR $\geq 1.0$ — any indication of
harm at all, regardless of magnitude. For clinically-meaningful
harm matching the FS `hr.threshold = 1.25`, use
$\log(1.25) \approx 0.223$ in a follow-up run.

Two practical notes about what to expect at this permissive
threshold:

1. **The discovered subgroup may differ from the FS finding.** A
   more permissive threshold often picks up *larger* subgroups
   (since more of the population satisfies the weaker harm
   condition), and may also surface a *different covariate* if
   small effects accumulate across a big subgroup. The FS Cox
   finding (Estrogen $\leq 0$ with HR $\approx 1.5$) should still
   be visible if it's real, but the threshold-search optimization
   may prefer a different cut at the lower threshold.

2. **Selection stability may drop.** With $m_{\text{diff}}$ near
   zero, multiple covariates can satisfy the harm condition on
   bootstrap replicates, so `selection_frequency` may be more
   diffuse and `n_modal_match` smaller than at a stricter
   threshold. The `effect_ci` interpretation stays clean — it is a
   percentile CI on a fixed-subgroup linear functional regardless
   of threshold — but the stability diagnostics
   (`n_subgroup_ci`, `threshold_ci`) may look noisier.

Both behaviors are expected and informative; they are what makes
$m_{\text{diff}} = 0.0$ a useful first-pass exploration.

### Direction

`direction = "both"` allows either tail of any covariate. The FS
Cox analysis discovered Estrogen $\leq 0$ (a *left*-direction
subgroup on `er`), so we expect DINA to surface something in that
neighborhood.

```{r dina-cox-subgroup}
sg_cox <- dina_subgroup(
  fit        = fit_dina_cox,
  df         = df.analysis,
  covariates = confounders,
  m_diff     = m_diff,
  n_min      = n_min,
  direction  = "both",
  alpha      = alpha
)

print(sg_cox)

# Gate the downstream bootstrap and summary on whether a subgroup was
# found.  The bootstrap targets the fixed-subgroup linear functional
# a*' beta with a* derived from the original-data subgroup, so it is
# undefined when no subgroup is identified.
run_boot    <- isTRUE(sg_cox$found)
run_no_boot <- !run_boot            # complement, used by the no-subgroup chunk
timings$dina_cox_boot <- NA_real_   # placeholder; overwritten if bootstrap runs

if (run_boot) {
  cat("\nSubgroup identified -- bootstrap inference will run below.\n")
} else {
  cat("\nNo subgroup met the criteria at m_diff =", m_diff,
      ", n_min =", n_min, ".\n",
      "Bootstrap inference will be skipped.\n", sep = "")
}
```

## Selection-adjusted bootstrap inference

The bootstrap targets the fixed-subgroup linear functional
$a^{*\top} \beta$, where $a^* = (1, \bar{x}_{S^*})$ is fixed from
the original-data subgroup $S^*$. If `dina_subgroup()` did not
identify a subgroup, $a^*$ is undefined and the bootstrap is
skipped; the chunks below are evaluated conditionally on
`run_boot`. The upstream `fit_dina_cox` and `sg_cox` are passed
directly to `dina_subgroup_bootstrap()` via its `fit` and `sg`
arguments, so the original-data point estimate is reused rather
than recomputed --- guaranteeing it matches the subgroup search
above exactly.

With `refit = TRUE`, the bootstrap additionally reports the
**standard within-subgroup Cox model** --- a plain `coxph()`
comparing the treatment arms inside the discovered signature
$\{\texttt{nodes} \ge 12\}$. This is the conventional
clinical-trial estimate of the treatment hazard ratio in the
subgroup, distinct from the DINA BLP-analog $a^{*\top}\beta$
(a projection of the cross-fitted CATE surface); the two coincide
only under correct linearity. Its bootstrap CI refits the same plain
Cox model within the *fixed* signature on each resample, so it is
conditional on the discovered signature in exactly the way the DINA
`effect_ci` is. Neither interval adjusts for the data-driven
selection of the signature --- a selection-adjusted (de-biased)
interval would require the more involved bias-corrected bootstrap of
Le&oacute;n et al. (2024, Section 3.2), not attempted here; selection
*stability* is instead summarized by the selection-frequency rows.

```{r dina-cox-bootstrap}
#| eval: !expr run_boot
t0 <- proc.time()

future::plan(future::multisession, workers = n_cores)

bs_cox <- dina_subgroup_bootstrap(
  df         = df.analysis,
  outcome    = cox_outcome,
  treatment  = treat_col,
  covariates = confounders,
  family     = "cox",
  status     = cox_status,
  m_diff     = m_diff,
  n_min      = n_min,
  direction  = "both",
  alpha      = alpha,
  n_boot     = NB,
  parallel   = "boots",
  n_folds    = dina_n_folds,
  seed       = boot_seed,
  fit        = fit_dina_cox,   # reuse upstream fit (no internal refit)
  sg         = sg_cox,         # reuse upstream subgroup -> point estimate identical
  refit             = TRUE,    # also fit the standard within-subgroup Cox model
  refit_confounders = "none",  # unadjusted; NULL = DINA covariates minus subgroup factor;
                               #   or a character vector of column names
  refit_strata      = NULL     # e.g. c("var") for Cox strata() terms
)

future::plan("sequential")
gc(verbose = FALSE)

timings$dina_cox_boot <- (proc.time() - t0)[["elapsed"]]
cat("DINA Cox bootstrap completed in",
    round(timings$dina_cox_boot / 60, 2), "minutes\n")
```

```{r dina-cox-bootstrap-print}
#| eval: !expr run_boot
print(bs_cox)
```

## Summary table

```{r dina-cox-summary-helpers}
`%||%` <- function(a, b) if (is.null(a)) b else a

fmt_signature <- function(sg) {
  if (!isTRUE(sg$found)) return("—")
  cmp <- if (sg$direction == "left") "<=" else ">="
  sprintf("%s %s %.3f", sg$covariate, cmp, sg$threshold)
}

fmt_ratio_ci <- function(ci, digits = 3L) {
  if (any(!is.finite(ci))) return("—")
  sprintf("[%.*f, %.*f]",
          digits, exp(ci[["lower"]]),
          digits, exp(ci[["upper"]]))
}

fmt_log_ci <- function(ci, digits = 3L) {
  if (any(!is.finite(ci))) return("—")
  sprintf("[%.*f, %.*f]",
          digits, ci[["lower"]],
          digits, ci[["upper"]])
}

fmt_int_ci <- function(ci) {
  if (any(!is.finite(ci))) return("—")
  sprintf("[%.0f, %.0f]", ci[["lower"]], ci[["upper"]])
}

fmt_top_selections <- function(bs, k = 3L) {
  if (is.null(bs$selection_frequency) || bs$n_boot_found == 0L) return("—")
  sel_long <- as.data.frame(bs$selection_frequency,
                            stringsAsFactors = FALSE)
  sel_long <- sel_long[sel_long$Freq > 0L, , drop = FALSE]
  sel_long <- sel_long[order(-sel_long$Freq), , drop = FALSE]
  top <- head(sel_long, k)
  paste(sprintf("%s (%s): %d/%d (%.0f%%)",
                top$covariate, top$direction, top$Freq,
                bs$n_boot_found,
                100 * top$Freq / bs$n_boot_found),
        collapse = "; ")
}

# Defensive Wald-CI extractor from a dina_subgroup result.
# Tries several plausible field-name conventions before falling back to NA.
extract_wald_ci <- function(sg) {
  na_ci <- c(lower = NA_real_, upper = NA_real_)
  if (!isTRUE(sg$found)) return(na_ci)
  vec_candidates <- c("mean_tau_ci", "wald_ci", "ci")
  for (cand in vec_candidates) {
    if (cand %in% names(sg) && length(sg[[cand]]) == 2L) {
      v <- as.numeric(sg[[cand]])
      return(c(lower = v[1L], upper = v[2L]))
    }
  }
  pair_candidates <- list(
    c("mean_tau_ci_lower", "mean_tau_ci_upper"),
    c("mean_tau_ci_lo",    "mean_tau_ci_hi"),
    c("ci_lower",          "ci_upper"),
    c("lower",             "upper")
  )
  for (cand in pair_candidates) {
    if (all(cand %in% names(sg))) {
      return(c(lower = as.numeric(sg[[cand[1L]]]),
               upper = as.numeric(sg[[cand[2L]]])))
    }
  }
  na_ci
}
```

```{r dina-cox-summary-table}
#| eval: !expr run_boot
pt_cox   <- bs_cox$point
wald_cox <- extract_wald_ci(pt_cox)

summary_tbl <- data.table(
  Field = c(
    "Discovered subgroup signature",
    "n_subgroup",
    "Bootstrap found / total",
    "Bootstrap failed",
    "Modal-match iters / found",
    "DINA point: mean tau-hat (log-HR)",
    "DINA point: exp(mean tau-hat) (HR)",
    "DINA Sandwich Wald CI (log-HR)",
    "DINA Sandwich Wald CI (HR)",
    "DINA bootstrap effect CI (log-HR)",
    "DINA bootstrap effect CI (HR)",
    "n_subgroup CI (modal-match)",
    "Threshold CI (modal-match)",
    "Top (covariate, direction) selections"
  ),
  Value = c(
    fmt_signature(pt_cox),
    sprintf("%d / %d", pt_cox$n_subgroup %||% NA, pt_cox$n_total %||% NA),
    sprintf("%d / %d", bs_cox$n_boot_found, bs_cox$n_boot),
    sprintf("%d", bs_cox$n_boot_failed),
    if (!is.na(bs_cox$n_modal_match))
      sprintf("%d / %d", bs_cox$n_modal_match, bs_cox$n_boot_found)
    else "—",
    sprintf("%.3f", pt_cox$mean_tau_hat %||% NA_real_),
    sprintf("%.3f", exp(pt_cox$mean_tau_hat %||% NA_real_)),
    fmt_log_ci(wald_cox),
    fmt_ratio_ci(wald_cox),
    fmt_log_ci(bs_cox$effect_ci),
    fmt_ratio_ci(bs_cox$effect_ci),
    fmt_int_ci(bs_cox$n_subgroup_ci),
    fmt_log_ci(bs_cox$threshold_ci),
    fmt_top_selections(bs_cox)
  )
)

# ---- Within-subgroup standard Cox model -------------------------------
# bs_cox$refit is a dina_subgroup_refit object (the plain Cox model fit
# within the discovered signature); bs_cox$refit_effect_ci is its
# bootstrap percentile CI, refit within the SAME fixed signature on each
# resample.  This is the clinically interpretable treatment HR, distinct
# from the DINA BLP-analog above.  Both are conditional on the signature.
rf_cox <- bs_cox$refit
refit_rows <- if (!is.null(rf_cox)) {
  adj <- if (length(rf_cox$confounders_used) == 0L) "unadjusted"
         else paste(rf_cox$confounders_used, collapse = ", ")
  strata_note <- if (length(rf_cox$strata_used) > 0L)
    sprintf("; strata: %s", paste(rf_cox$strata_used, collapse = ", "))
  else ""
  data.table(
    Field = c(
      "Within-subgroup model",
      "Within-subgroup Cox: events (trt / ctrl)",
      "Within-subgroup Cox point (log-HR)",
      "Within-subgroup Cox point (HR)",
      "Within-subgroup Cox Wald CI (log-HR)",
      "Within-subgroup Cox Wald CI (HR)",
      "Within-subgroup Cox bootstrap CI (log-HR)",
      "Within-subgroup Cox bootstrap CI (HR)"
    ),
    Value = c(
      sprintf("standard Cox within %s (%s%s)",
              rf_cox$signature, adj, strata_note),
      sprintf("%d (%d / %d)", rf_cox$n_events %||% NA_integer_,
              rf_cox$n_events_treated %||% NA_integer_,
              rf_cox$n_events_control %||% NA_integer_),
      sprintf("%.3f", rf_cox$effect),
      sprintf("%.3f", exp(rf_cox$effect)),
      fmt_log_ci(rf_cox$ci),
      fmt_ratio_ci(rf_cox$ci),
      fmt_log_ci(bs_cox$refit_effect_ci),
      fmt_ratio_ci(bs_cox$refit_effect_ci)
    )
  )
} else {
  data.table(
    Field = "Within-subgroup standard model",
    Value = "— (refit disabled or unavailable)"
  )
}

summary_tbl <- rbind(summary_tbl, refit_rows)

summary_tbl |>
  gt::gt() |>
  gt::tab_header(
    title    = gt::md("**DINA Cox analysis — summary**"),
    subtitle = sprintf("m_diff = %.3f (permissive first pass; log-HR scale);  n_boot = %d;  %d cross-fitting folds",
                       m_diff, NB, dina_n_folds)
  ) |>
  gt::tab_source_note(
    source_note = gt::md(
      "*Two treatment-effect estimates for the discovered subgroup, on the
       log-HR scale. **DINA** rows are the BLP-analog `a*'beta` (a
       projection of the cross-fitted CATE surface): Sandwich Wald CI from
       `dina_subgroup()`, bootstrap percentile CI on `a*'beta_b` with `a*`
       fixed from the original subgroup. **Within-subgroup Cox** rows are a
       standard `coxph()` comparing arms inside the discovered signature ---
       the conventional clinical estimate --- with its own Wald CI and a
       bootstrap percentile CI refit within the same fixed signature on each
       resample. The two coincide only under correct linearity. Both CIs are
       conditional on the discovered signature treated as pre-specified;
       neither adjusts for signature selection. If Sandwich CI rows display
       '—', update the field-name conventions in `extract_wald_ci()` to match
       this DINA build.*"
    )
  )
```

```{r dina-cox-no-subgroup}
#| eval: !expr run_no_boot
no_sg_tbl <- data.table(
  Field = c(
    "Subgroup found",
    "m_diff (log-HR scale)",
    "exp(m_diff) (HR scale)",
    "n_min",
    "direction",
    "Covariates searched",
    "n",
    "Note"
  ),
  Value = c(
    "No",
    sprintf("%.3f", m_diff),
    sprintf("%.3f", exp(m_diff)),
    sprintf("%d", n_min),
    "both",
    paste(confounders, collapse = ", "),
    sprintf("%d (events: %d)",
            nrow(df.analysis),
            sum(df.analysis[[cox_status]])),
    paste0("Bootstrap inference skipped because the fixed-subgroup ",
           "linear functional is undefined when no subgroup is ",
           "identified.")
  )
)

no_sg_tbl |>
  gt::gt() |>
  gt::tab_header(
    title    = gt::md("**DINA Cox analysis -- no subgroup identified**"),
    subtitle = sprintf("m_diff = %.3f (log-HR);  n_min = %d;  cross-fitting folds = %d",
                       m_diff, n_min, dina_n_folds)
  ) |>
  gt::tab_source_note(
    source_note = gt::md(
      "*To get a stable result on data where no harm subgroup exists,
       consider (i) loosening the search constraints (smaller `n_min`,
       different `direction`), (ii) checking the ITT-level diagnostic
       above for whether the overall effect direction is consistent
       with harm at all, or (iii) re-running on a different covariate
       set.*"
    )
  )
```

# DINA Screening Candidates for ForestSearch

The Cox DINA fit above does double duty. Beyond the BLP and the
single-subgroup views, its **per-covariate Pareto frontier** can seed
`forestsearch()`'s screening stage as candidate cuts — the DINA analog
of GRF tree cuts, gated by `use_dina`. The cuts are *candidates*, not
forced selections: `forestsearch()` composes them (up to `maxk`) and the
consistency stage remains the gatekeeper.

::: {.callout-note}
## `dina_args` is selection-agnostic

The screening path uses `dina_frontier()`, which returns the *entire*
per-covariate Pareto frontier. It does **not** use `forestsearch()`'s
`sg_focus` or `selection_rule`: those control how a *single* subgroup is
*selected* (in the consistency stage, and in `dina_subgroup()`), whereas
screening emits the whole frontier as a candidate net. Accordingly
`dina_args` recognises only the DINA **fit** keys (`family`, `seed`,
`n_folds`, `cens_type`, `cens_params`) and the **frontier** keys
(`scope`, `m_diff`, `n_min`, `direction`, `max_per_covariate`,
`max_subgroups`, `digits`); `sg_focus`/`selection_rule` are not among
them and raise an "unrecognised key" error if passed.
:::

## Per-covariate Pareto frontier (`dina_frontier`)

`dina_frontier()` reuses the fitted `fit_dina_cox` and returns, for each
covariate, the non-dominated (effect, size) cuts rendered as forestsearch
cut expressions. `scope = "wide"` casts a broad net (every candidate with
size $\ge$ `n_min`); the `effect` column is on the natural HR scale for
the Cox family. (`scope = "harm"` with `m_diff` would instead gate to the
harm tail on the log-HR scale — the same `m_diff` used by
`dina_subgroup()` above.)

```{r dina-frontier}
t0 <- proc.time()

frontier <- dina_frontier(
  fit        = fit_dina_cox,
  df         = df.analysis,
  covariates = confounders,
  scope      = "wide",
  n_min      = n_min,          # mirrors forestsearch n.min
  direction  = "both"
)                              # caps default to max_per_covariate = 3, max_subgroups = 10

timings$dina_frontier <- (proc.time() - t0)[["elapsed"]]

frontier |>
  gt::gt() |>
  gt::fmt_number(columns = c(threshold, mean_tau_hat, effect), decimals = 3) |>
  gt::cols_label(
    covariate = "Covariate", direction = "Dir", threshold = "Cutpoint",
    n_subgroup = "n", mean_tau_hat = "log-HR", effect = "HR",
    cut_expr = "Cut expression"
  ) |>
  gt::tab_header(
    title    = "DINA per-covariate Pareto frontier",
    subtitle = sprintf(
      "%d cuts across %d covariate(s); n_frontier = %d, n_qualifying = %d (scope = wide)",
      nrow(frontier), length(unique(frontier$covariate)),
      attr(frontier, "n_frontier"), attr(frontier, "n_qualifying"))
  )
```

The `cut_expr` column is exactly what feeds `forestsearch()`.

## How the candidates populate the screening pool (`get_FSdata`)

`get_FSdata()` builds the candidate cut pool the consistency search ranges
over. Calling it with and without `dina_cuts` shows precisely what DINA
contributes: the frontier cuts are merged into the pool alongside the
default median/categorical cuts, deduplicated against cuts already present.

```{r dina-screening-pool}
t0 <- proc.time()

fsdata_args <- list(
  df.analysis = df.analysis, use_lasso = FALSE, use_grf = FALSE,
  confounders.name = confounders, outcome.name = cox_outcome,
  event.name = cox_status, outcome_type = "survival", details = FALSE
)

pool_base <- do.call(get_FSdata, c(fsdata_args, list(dina_cuts = NULL)))$confs
pool_dina <- do.call(get_FSdata,
                     c(fsdata_args, list(dina_cuts = frontier$cut_expr)))$confs

timings$dina_screening_pool <- (proc.time() - t0)[["elapsed"]]

added <- setdiff(pool_dina, pool_base)

data.table(
  Quantity = c("Pool size — base (median + categorical cuts)",
               "Pool size — with DINA screening",
               "Net DINA-added cuts",
               "DINA-added cut expressions"),
  Value    = c(as.character(length(pool_base)),
               as.character(length(pool_dina)),
               as.character(length(added)),
               if (length(added)) paste(added, collapse = "; ") else "(none — all already present)")
) |>
  gt::gt() |>
  gt::tab_header(
    title    = "DINA candidates populating the get_FSdata() pool",
    subtitle = "use_grf = FALSE, use_lasso = FALSE: pool = median/categorical cuts + DINA frontier cuts"
  )
```

## End-to-end via `forestsearch(use_dina = TRUE)`

Two equivalent ways to drive this from `forestsearch()` itself. Runs are
kept light (`use_grf = FALSE`, sequential) because the point here is the
screening candidates, not the consistency search.

**Turnkey — `forestsearch()` fits DINA internally.** With `use_dina = TRUE`
and no `dina_res`, `forestsearch()` fits its own DINA model at the
screening stage (family mapped from `outcome_type`). Here `dina_args`
passes the same `seed`/`n_folds` as `fit_dina_cox`, so the internal fit
reproduces the standalone frontier.

```{r fs-use-dina-internal}
t0 <- proc.time()

fs_dina_int <- forestsearch(
  df.analysis      = df.analysis,
  outcome.name     = cox_outcome,
  event.name       = cox_status,
  treat.name       = treat_col,
  id.name          = id_col,
  confounders.name = confounders,
  outcome_type     = "survival",
  is.RCT           = TRUE,
  n.min            = n_min,
  use_lasso        = FALSE,
  use_grf          = TRUE,
  use_dina         = TRUE,                      # DINA fit happens inside forestsearch()
  dina_args        = list(seed = dina_seed, n_folds = dina_n_folds,
                          scope = "wide", n_min = n_min),
  parallel_args    = list(plan = "sequential"),
  maxk             = 2,
  fs.splits        = 1000,
  details          = TRUE
)

timings$fs_use_dina_internal <- (proc.time() - t0)[["elapsed"]]

cat("dina_cuts injected:", length(fs_dina_int$dina_cuts), "\n")
cat("internal dina_res class:",
    paste(class(fs_dina_int$dina_res), collapse = "/"), "\n")
cat("matches standalone frontier:",
    identical(sort(fs_dina_int$dina_cuts), sort(frontier$cut_expr)), "\n")
fs_dina_int$dina_cuts
```

**Reuse — supply the fit via `dina_res`.** Passing `dina_res = fit_dina_cox`
skips the internal refit and reuses the model already fitted above. The
injected cuts are identical; the saving is the DINA fit time (compare the
two rows in the Timings table).

```{r fs-use-dina-reuse}
t0 <- proc.time()

fs_dina_reuse <- forestsearch(
  df.analysis      = df.analysis,
  outcome.name     = cox_outcome,
  event.name       = cox_status,
  treat.name       = treat_col,
  id.name          = id_col,
  confounders.name = confounders,
  outcome_type     = "survival",
  is.RCT           = TRUE,
  n.min            = n_min,
  use_lasso        = FALSE,
  use_grf          = FALSE,
  use_dina         = TRUE,
  dina_res         = fit_dina_cox,              # reuse the fit; no refit
  dina_args        = list(scope = "wide", n_min = n_min),
  parallel_args    = list(plan = "sequential"),
  maxk             = 2,
  fs.splits        = 200,
  details          = FALSE
)

timings$fs_use_dina_reuse <- (proc.time() - t0)[["elapsed"]]

cat("dina_cuts injected:", length(fs_dina_reuse$dina_cuts), "\n")
cat("dina_res is the supplied fit (not refit):",
    identical(fs_dina_reuse$dina_res, fit_dina_cox), "\n")
cat("identical cuts to the internal-fit path:",
    identical(sort(fs_dina_reuse$dina_cuts), sort(fs_dina_int$dina_cuts)), "\n")
```

In both paths the injected `dina_cuts` are merged into the screening
candidate pool shown above and then handed to the consistency search
exactly like GRF cuts — candidates for, not commitments to, the final
subgroup.

# Future Work: DINA Poisson + Offset Extension

The FS doc's Cox-vs-Poisson alignment analysis is the natural
next deliverable here. It requires:

1. **Extend `dina_fit()` with an `offset` argument.** Pass the
   offset vector to `.dina_one_pass()`, slice it by train/test
   indices, and forward to:
   * `.dina_fit_baselines_glm()` — include the offset in the
     within-arm `stats::glm()` calls so the returned
     `eta0_fn` / `eta1_fn` predict on the offset-inclusive
     linear-predictor scale.
   * `.dina_step2_glm()` — already passes `offset = nu_a$nu`
     to `stats::glm.fit()`; if step 1 incorporates the offset
     into `nu`, step 2 needs no further change.
2. **Forward `offset` through the wrappers** `dina()`,
   `dina_bagged()`, and `dina_subgroup_bootstrap()` as a
   pass-through argument.
3. **Add a Poisson + offset family** to the
   `dina_subgroup_bootstrap_smoke_test.qmd` rotation. Synthetic
   data with person-time exposure, log offset.
4. **Re-render this document with the Poisson + offset branch
   restored** and a Part III Cox-Poisson alignment section
   mirroring the FS doc.

The treatment-effect contrast $\tau(x) = \eta_1(x) - \eta_0(x)$ is
invariant under a shared offset added to both arms, so DINA's
estimand is unchanged; only the nuisance fits and step-2 likelihood
need the offset. Bias-correction would not be needed beyond what
cross-fitting already handles.

# Appendix A: Linear projections of the CATE, and extensions of the threshold search

This appendix collects the methodological background for the DINA
coefficient table reported above and discusses how the covariate-
thresholding subgroup search (`dina_subgroup()`) compares to and
complements the linear-projection view. It also discusses a
proposed depth-2 extension of the threshold search.

## A.1 Background: best linear projection (BLP) of Cui et al.

Cui, Kosorok, Sverdrup, Wager & Zhu (2023, *JRSS-B* 85:179–211)
develop *causal survival forests* (CSF) and report two
complementary outputs from a fitted forest:

1. **Pointwise CATE estimates** $\hat\tau(X_i)$ — a non-parametric
   estimate of the treatment-effect contrast for each patient.
   Their Table 4 shows these for ten patients with HC3-style SEs.
2. **Best linear projection (BLP)** — a low-dimensional linear
   summary of the CATE, motivated by Semenova & Chernozhukov
   (2021, *Econometrics J* 24:264–289). Their Table 5 reports
   these.

The BLP estimand (Cui et al. eq. 26) is

$$\{\beta_0^*,\, \beta^*\}\;=\;\arg\min_{\beta_0,\,\beta}\;
\mathbb{E}\big[\big(\tau(X_i) - \beta_0 - A_i\beta\big)^2\big],$$

where $A_i$ is a user-chosen low-dimensional subset of the
covariates used in fitting the forest. The BLP is the L²-best
linear approximation to the (possibly non-linear) CATE function
$\tau(\cdot)$, projected onto the linear span of $A_i$.

CSF's implementation of BLP follows three steps:

1. Construct doubly-robust scores
   $\hat\Gamma_i = \hat\tau(X_i) + \psi_{\hat\tau(X_i)}(\,\cdot\,) / [\hat e(X_i)(1 - \hat e(X_i))]$
   where $\hat\tau(X_i)$ are CSF estimates and $\psi$ is the
   doubly-robust score function (Cui et al. eq. 11).
2. Run linear regression of $\hat\Gamma_i$ on $A_i$.
3. Compute HC3 robust standard errors (MacKinnon & White 1985,
   *J Econometrics* 29:305–325).

The motivation: pointwise non-parametric inference on $\tau(x)$ is
fundamentally difficult — intervals are often very wide, and
undersmoothing is hard to verify in practice. Linear projections
give a low-dimensional summary that admits robust inference under
nuisance-rate conditions.

## A.2 DINA's β̂ as a parametric BLP

DINA targets the linear functional
$\tau(x) = \beta_0 + x^\top \beta_{1:d}$ directly as a parametric
model. When the true CATE is linear in the chosen covariates, the
DINA estimand coincides with the BLP estimand exactly:

* **CSF + BLP**: estimate $\tau(\cdot)$ non-parametrically, then
  project onto the linear span of $A_i$.
* **DINA**: assume the linear form upfront, estimate via
  cross-fitted GLM/Cox step 2.

Asymptotically, with cross-fitting and good nuisance rates, both
target the same population coefficient under linearity. When the
true CATE is non-linear, BLP gives the L²-best linear approximation
while DINA's coefficient is the parametric MLE under a (potentially
misspecified) linear model. In practice the two should agree
closely.

The DINA coefficient table presented above —
`coef(fit_dina_cox)` paired with `sqrt(diag(vcov(fit_dina_cox)))` —
is therefore the parametric counterpart of Cui et al.'s Table 5.
The Wald CIs are constructed from DINA's sandwich variance for the
GLM step 2, or the Cox partial-likelihood information variance for
the Cox step 2; these correspond closely to the HC3 SEs used by
Cui et al.

## A.3 BLP vs. covariate-threshold search

The BLP/coefficient view (Table 5 / DINA coefficient table) and
the threshold-search view (`dina_subgroup()`) answer fundamentally
different questions:

| | BLP / DINA coefficient table | `dina_subgroup` (threshold search) |
|---|---|---|
| Output form | One coefficient per covariate | A single (covariate, direction, threshold) signature |
| Estimand | $\arg\min E[(\tau(X) - \beta_0 - A\beta)^2]$ | $\arg\max\{n_S : \bar{\tau}_S \ge m_{\text{diff}}\}$ |
| Granularity | All covariates simultaneously | One covariate, picked by maximization |
| Inference | HC3 / sandwich CI on each coefficient | Wald CI on within-subgroup mean; bootstrap percentile CI |
| Selection adjustment | Not needed | Required (selected subgroup → fixed-subgroup bootstrap) |
| Behavior under null | Reports small/noisy coefficients | May fail to identify a subgroup |
| Clinical use | Predictive-biomarker hypothesis generation | Actionable decision rule |
| Robustness to non-linearity | Best linear approximation regardless | Aligned with sharp threshold structure |

**Where they agree.** If one covariate dominates the heterogeneity
(say, a sharp threshold on `er`), the BLP shows a large coefficient
on `er` and `dina_subgroup` picks a threshold on `er`.

**Where they diverge:**

* Genuinely *multivariate* heterogeneity (older + ER-negative +
  node-positive jointly drive harm): BLP captures the joint linear
  story; univariate `dina_subgroup` may miss the structure (and is
  the motivation for the depth-2 extension below).
* Sharply *thresholded* heterogeneity in a single biomarker: the
  BLP averages over the threshold and reports a damped coefficient;
  `dina_subgroup` aligns with the true structure.

The two views are **complementary**, not competing. A complete
analysis reports both, as this document does.

## A.4 Multi-projection support

Cui et al.'s Table 5 shows two columns: an "all covariates"
projection and an "age only" projection from the same fitted CSF.
The BLP framework supports this by separating two stages:

1. **CATE estimation** — fit forest with rich covariates for good
   nuisance estimation.
2. **CATE summary** — project the resulting $\hat\tau(\cdot)$ onto
   any chosen $A_i$ for interpretation.

The same forest fit supports many BLP summaries by varying $A_i$.
DINA conflates these — the covariate set in `dina_fit()` IS the
projection set, so getting an "age only" projection requires
refitting DINA. This is not necessarily wrong (the parametric
linear assumption is over the chosen covariate set), but it is
less flexible than CSF + BLP when multiple projections are of
interest.

A future enhancement could add `dina_project(fit, df, A)` that
takes a fitted full-covariate DINA and projects $\hat\beta$ onto a
subset $A$ via an appropriate linear transformation — supporting
multi-projection workflows like CSF + BLP does.

## A.5 Toward depth-2 covariate thresholding

The current `dina_subgroup()` searches over single-covariate
thresholds — a *depth-1* cut. A natural extension is *depth-2*:
search over pairs of covariate thresholds, yielding subgroups
defined by conjunctions like
"$\text{ER} \le 0$ AND $\text{age} \ge 50$". This is structurally
analogous to `policytree::policy_tree(depth = 2)` (Athey & Wager
2021) and to CAPITAL's constrained tree search at depth $d$
(Wang et al. 2021, arXiv:2110.05636).

### A.5.1 Computational scaling

Naive depth-2 search over $p$ covariates, two directions per
covariate, and $n$ candidate thresholds per covariate is
$O(p^2 \cdot 4 \cdot n^2)$. For the GBSG analysis
($p = 7$, $n = 686$) this is roughly 92M candidate evaluations
versus ~9.6K for depth-1 — a 10,000× increase. Substantial but
feasible with thoughtful caching: within-subgroup sums (numerator
and denominator of $\bar{\tau}_S$) can be updated incrementally
across nested thresholds, reducing the dominant cost. With
$n_{\min}$ as a hard floor and explicit early-termination when the
size constraint is violated, the effective search space is much
smaller.

### A.5.2 Inferential scaling

The current fixed-subgroup bootstrap target generalizes naturally
to depth-2:

* For a depth-2 subgroup $S^*$ defined by
  $(j_1, d_1, q_1, j_2, d_2, q_2)$, the linear functional is still
  $a^{*\top}\beta$ with $a^* = (1, \bar{x}_{S^*})$.
* The bootstrap iterates as before: refit DINA on resample, run
  depth-2 search, compute $a^{*\top}\beta_b$.
* `selection_frequency` records covariate-pair $(j_1, j_2)$ tuples
  rather than single covariates.
* `threshold_ci` becomes 2-dimensional (marginal CIs on $q_1$ and
  $q_2$ restricted to modal-pair iterations).
* The modal-match restriction is on the pair $(j_1, j_2)$.

So the inferential framework scales without conceptual change.

### A.5.3 Search-space and statistical considerations

1. **Search space definition.** Three reasonable options:
   * **AND-conjunction only** — recommended starting point.
     Patients satisfying both constraints. Most interpretable.
     Symmetric in the two covariates.
   * **OR-disjunction** — patients satisfying either constraint.
     Less interpretable; a different estimand (the union rather
     than the intersection).
   * **Full depth-2 tree** — outer split + per-leaf inner splits.
     Richest expressive class but least interpretable. Closest to
     `policy_tree(depth = 2)`.
2. **Multiplicity.** Depth-2 search has substantially more
   selection-induced optimism than depth-1. The fixed-subgroup
   bootstrap correctly accounts for this for the *effect CI*, but
   the bias of $\overline{\hat\tau}_{S^*}$ as an estimator of the
   *population mean within* $S^*$ grows with depth. This is the
   territory where the forestsearch-style two-bias-term correction
   (Section 6 of the theory notes) becomes more relevant.
3. **Sample-size requirements.** Small samples like GBSG
   (~300 events) probably do not support reliable depth-2
   identification — the $n_{\min}$ constraint becomes binding for
   most candidate cuts. Depth-2 is recommended for $n \ge 2000$
   with adequate event counts (say, $\ge 500$ events).

### A.5.4 Implementation effort

The bootstrap infrastructure scales without changes; the main code
change is the search loop within `dina_subgroup()` itself and a
matching update inside `dina_subgroup_bootstrap()`. A `max_depth`
argument with default `1L` keeps backward compatibility. Estimated
effort: one to two focused development sessions to add the
AND-conjunction search, including tests for the new search-space
mode.

### A.5.5 Comparison with existing depth-2 methods

* **CAPITAL** (Wang, Cai, Zeng, Song & Kosorok 2021): constrained
  policy tree search at depth $d$. Closest existing comparator —
  CAPITAL's optimization target is "largest subgroup with mean
  effect $\ge$ threshold", identical in spirit to `dina_subgroup`,
  with a tree-search solver rather than an exhaustive grid.
* **`policytree::policy_tree(depth = 2)`** (Athey & Wager 2021,
  *Econometrica* 89:133–161): doubly-robust empirical welfare
  maximization over depth-2 trees. Different optimization target
  (welfare maximization vs. constrained subgroup search) but the
  same structural form.

A depth-2 `dina_subgroup` would sit between these:
constrained-subgroup semantics (matching CAPITAL and the v0.1.0 FS
work), with the simpler AND-conjunction form for interpretability.

### A.5.6 Recommended priority

Deferred to v0.3.0 alongside the offset extension. The
interpretability gain for clinical applications is real — a depth-2
conjunction is closer to how clinicians think about biomarker
profiles than a depth-1 cut — but the inferential framework
benefits from being battle-tested at depth-1 first. Once the v0.2.0
release is stable, depth-2 is a natural next addition.

# Timings

```{r timings-table}
fmt_secs <- function(x) {
  if (is.null(x) || !is.finite(x)) "—" else sprintf("%.1f", x)
}

tim_df <- data.table(
  Step       = c("DINA Cox fit",
                 "DINA Cox bootstrap",
                 "DINA frontier",
                 "Screening pool (get_FSdata)",
                 "ForestSearch (use_dina, internal fit)",
                 "ForestSearch (use_dina, reuse fit)",
                 "Total"),
  `Time (s)` = c(
    fmt_secs(timings$dina_cox_fit),
    fmt_secs(timings$dina_cox_boot),
    fmt_secs(timings$dina_frontier),
    fmt_secs(timings$dina_screening_pool),
    fmt_secs(timings$fs_use_dina_internal),
    fmt_secs(timings$fs_use_dina_reuse),
    fmt_secs((proc.time() - t_doc_start)[["elapsed"]])
  )
)

tim_df |>
  gt::gt() |>
  gt::tab_header(title = "Timing summary")
```

# Session Information

```{r session-info}
sessionInfo()
```
