---
title: "DINA Subgroup Smoke Test — Threshold Search Across All Four Families"
subtitle: "End-to-end pipeline check for dina_subgroup() with variance propagation from both dina() (sandwich) and dina_bagged() (IJ)"
author: "forestsearch v0.2.0"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: true
  message: true
---

# Purpose

Smoke-test `forestsearch::dina_subgroup()` end-to-end across all four
response families.  The function bridges DINA's continuous per-patient
HTE estimate $\hat\tau(x_i)$ with forestsearch's interpretable
threshold-based subgroup framework: for each candidate covariate `x_j`
and threshold `q`, it computes the subgroup-mean $\hat\tau$ and returns
the *largest* subgroup whose mean meets the harm threshold `m_diff`.
Variance is propagated through the linear-functional identity
$\widehat{\mathrm{Var}}(\bar{\hat\tau}_S) = a_S^\top \mathrm{vcov}(\mathrm{fit}) a_S$,
which automatically tracks the variance source — sandwich for `"dina"`,
infinitesimal jackknife for `"dina_bagged"`.

For each family we:

1. Generate a single synthetic dataset with a **deliberately strong
   linear treatment contrast** $\tau(x) = \beta_0 + \beta_1 x_1$ so that
   a harm subgroup at one end of $x_1$ genuinely exists.
2. Fit plain `dina()` (sandwich variance) on the data.
3. Call `dina_subgroup(fit_sandwich, ...)` with a realistic `m_diff`,
   asserting structural and mathematical correctness of the returned
   `"dina_subgroup"` object.
4. Fit `dina_bagged()` (IJ variance) on the same data with
   `parallel = "bags"`.
5. Call `dina_subgroup(fit_ij, ...)` and assert the IJ-variance path
   gives a similar subgroup and a sensible Wald CI.
6. Negative control: call `dina_subgroup` with an impossibly large
   `m_diff` and assert `found == FALSE`.

Assertions accumulate via the standard `.tc()` helper; a gt
pass/fail summary is rendered at the bottom.

---

# Setup

```{r setup}
library(forestsearch)
library(data.table)
library(gt)

set.seed(20260528)

SMOKE <- list(
  n                = 800L,
  d                = 5L,
  n_bags           = 200L,
  n_folds_per_bag  = 2L,
  plain_n_folds    = 2L,
  parallel         = "bags",     # "bags" or "none"
  n_workers_frac   = 0.9,
  n_workers        = NULL,
  n_min            = 60L,
  alpha            = 0.05,
  cox_T0           = 1.0
)

# Treatment-effect contrast: large beta_1, others zero.  With X ~ U[-1,1]:
#   tau(x) = beta_0 + beta_1 * x_1
# so for x_1 sufficiently negative, tau(x) is well below beta_0; for x_1
# sufficiently positive, tau(x) is well above.  This gives the search a
# genuine signal to find on x_1.
BETA_TRUE <- c(0.40, 1.20, 0.00, 0.00, 0.00, 0.00)
stopifnot(length(BETA_TRUE) == SMOKE$d + 1L)

# Family-specific m_diff values, chosen as ~50% of the maximum tau(x)
# expected on the right tail (x_1 -> 1):  beta_0 + beta_1 * 1 = 1.60.
# Using 0.80 puts the threshold squarely within the harm-tail region so a
# right-direction subgroup on x_1 is recoverable.
M_DIFF_BY_FAMILY <- c(
  gaussian = 0.80,
  binomial = 0.80,   # natural-parameter (log-OR) scale
  poisson  = 0.80,   # natural-parameter (log-IRR) scale
  cox      = 0.80    # natural-parameter (log-HR) scale
)
```

## `.tc()` accumulating helper

```{r tc-helper}
.tc_env <- new.env(parent = emptyenv())
.tc_env$rows <- list()

.tc <- function(label, condition, family, details = "") {
  passed <- isTRUE(tryCatch(condition, error = function(e) FALSE))
  msg <- if (!passed && nzchar(details)) details
         else if (!passed) "condition FALSE or errored"
         else ""
  .tc_env$rows[[length(.tc_env$rows) + 1L]] <- data.table(
    family  = family,
    label   = label,
    status  = if (passed) "PASS" else "FAIL",
    details = msg
  )
  invisible(passed)
}

.tc_table <- function() {
  if (length(.tc_env$rows) == 0L) {
    return(data.table(family = character(), label = character(),
                      status = character(), details = character()))
  }
  rbindlist(.tc_env$rows)
}

`%||%` <- function(a, b) if (is.null(a)) b else a
```

## Parallel plan setup

```{r plan-setup}
if (SMOKE$parallel == "bags") {
  total_cores <- parallelly::availableCores()
  n_workers <- if (is.null(SMOKE$n_workers)) {
    max(1L, as.integer(floor(SMOKE$n_workers_frac * total_cores)))
  } else {
    as.integer(SMOKE$n_workers)
  }
  future::plan(future::multisession, workers = n_workers)
  message(sprintf(
    "Parallel plan: %d workers (%.0f%% of %d available cores)",
    n_workers, 100 * n_workers / total_cores, total_cores
  ))
} else {
  future::plan("sequential")
  message("Parallel plan: sequential")
}
```

# Data simulation

Same general DGM as the other DINA smoke tests, with $\beta_1 = 1.20$
to make the heterogeneity strong enough that a subgroup at the high-$x_1$
end is genuinely detectable.

```{r data-helpers}
simulate_dina_data <- function(family, n, d, beta_true, cox_T0 = 1.0) {
  X <- matrix(stats::runif(n * d, -1, 1), n, d)
  colnames(X) <- paste0("x", seq_len(d))

  e_x  <- stats::plogis(0.3 * X[, 1])
  W    <- stats::rbinom(n, 1L, e_x)

  mu_x  <- 0.2 * X[, 1] - 0.1 * X[, 2]
  tau_x <- as.numeric(beta_true[1L] + X %*% beta_true[-1L])
  eta_w <- mu_x + W * tau_x

  out <- list(X = X, W = W, beta_true = beta_true, eta_w = eta_w,
              tau_x_true = tau_x)

  out$Y <- switch(
    family,
    gaussian = as.numeric(eta_w + stats::rnorm(n)),
    binomial = stats::rbinom(n, 1L, stats::plogis(eta_w)),
    poisson  = stats::rpois(n, exp(pmin(eta_w, 5))),
    cox = {
      t_event  <- stats::rexp(n, rate = exp(eta_w))
      t_censor <- stats::runif(n, 0, cox_T0)
      time   <- pmin(t_event, t_censor)
      status <- as.integer(t_event <= t_censor)
      data.frame(time = time, status = status)
    }
  )
  out
}

# Pack a single family's matrix-form data into a data frame for the
# wrapper API (dina(), dina_bagged(), dina_subgroup()).
build_df <- function(dat, family) {
  if (family == "cox") {
    cbind(
      data.frame(time = dat$Y$time, status = dat$Y$status, w = dat$W),
      as.data.frame(dat$X)
    )
  } else {
    cbind(
      data.frame(y = dat$Y, w = dat$W),
      as.data.frame(dat$X)
    )
  }
}
```

## Per-family test runner

The runner exercises `dina_subgroup()` against both the sandwich-variance
and IJ-variance fits and stores everything needed for downstream
diagnostic tables.

```{r runner}
run_subgroup_family_tests <- function(family) {
  m_diff <- M_DIFF_BY_FAMILY[[family]]

  dat <- simulate_dina_data(
    family    = family,
    n         = SMOKE$n,
    d         = SMOKE$d,
    beta_true = BETA_TRUE,
    cox_T0    = SMOKE$cox_T0
  )
  df_fam       <- build_df(dat, family)
  outcome_col  <- if (family == "cox") "time" else "y"
  status_col   <- if (family == "cox") "status" else NULL
  covariates_v <- colnames(dat$X)

  cens_type   <- if (family == "cox") "uniform" else NULL
  cens_params <- if (family == "cox") list(T0 = SMOKE$cox_T0) else list()

  # ---- Plain DINA (sandwich variance) -------------------------------------
  fit_sandwich <- tryCatch(
    dina(df = df_fam, outcome = outcome_col, treatment = "w",
         covariates = covariates_v, family = family, status = status_col,
         n_folds = SMOKE$plain_n_folds,
         cens_type = cens_type, cens_params = cens_params,
         seed = 20260528L),
    error = function(e) NULL
  )
  .tc("dina() completes without error", !is.null(fit_sandwich), family)

  # ---- Bagged DINA (IJ variance) ------------------------------------------
  fit_ij <- tryCatch(
    dina_bagged(df = df_fam, outcome = outcome_col, treatment = "w",
                covariates = covariates_v, family = family, status = status_col,
                n_bags = SMOKE$n_bags,
                n_folds_per_bag = SMOKE$n_folds_per_bag,
                parallel = SMOKE$parallel,
                cens_type = cens_type, cens_params = cens_params,
                seed = 20260528L, verbose = FALSE),
    error = function(e) NULL
  )
  .tc("dina_bagged() completes without error", !is.null(fit_ij), family)

  # ---- Subgroup search on sandwich fit ------------------------------------
  sg_sand <- if (!is.null(fit_sandwich)) {
    tryCatch(
      dina_subgroup(fit = fit_sandwich, df = df_fam,
                    covariates = covariates_v,
                    m_diff = m_diff, n_min = SMOKE$n_min,
                    alpha = SMOKE$alpha),
      error = function(e) {
        attr(e, "details") <- conditionMessage(e)
        e
      }
    )
  } else NULL
  sand_ok <- !is.null(sg_sand) && !inherits(sg_sand, "error")
  .tc("dina_subgroup() on sandwich fit completes without error", sand_ok, family,
      details = if (!sand_ok && inherits(sg_sand, "error")) attr(sg_sand, "details") else "")

  # ---- Subgroup search on IJ-bagged fit -----------------------------------
  sg_ij <- if (!is.null(fit_ij)) {
    tryCatch(
      dina_subgroup(fit = fit_ij, df = df_fam,
                    covariates = covariates_v,
                    m_diff = m_diff, n_min = SMOKE$n_min,
                    alpha = SMOKE$alpha),
      error = function(e) {
        attr(e, "details") <- conditionMessage(e)
        e
      }
    )
  } else NULL
  ij_ok <- !is.null(sg_ij) && !inherits(sg_ij, "error")
  .tc("dina_subgroup() on IJ-bagged fit completes without error", ij_ok, family,
      details = if (!ij_ok && inherits(sg_ij, "error")) attr(sg_ij, "details") else "")

  # ---- Structural assertions on the sandwich result -----------------------
  if (sand_ok) {
    .tc("Sandwich subgroup: returns object of class 'dina_subgroup'",
        inherits(sg_sand, "dina_subgroup"), family)
    .tc("Sandwich subgroup: 'found' field present and logical",
        is.logical(sg_sand$found) && length(sg_sand$found) == 1L, family)
    .tc("Sandwich subgroup: n_candidates_searched > 0",
        is.numeric(sg_sand$n_candidates_searched) &&
          sg_sand$n_candidates_searched > 0L, family)
    .tc("Sandwich subgroup: print() runs without error",
        !inherits(tryCatch(utils::capture.output(print(sg_sand)),
                           error = function(e) e), "error"), family)

    if (isTRUE(sg_sand$found)) {
      # Math sanity: the returned subgroup satisfies its own constraints
      .tc("Sandwich subgroup: chosen mean_tau_hat >= m_diff",
          sg_sand$mean_tau_hat >= m_diff - 1e-10, family,
          details = sprintf("mean = %.3f, m_diff = %.3f",
                            sg_sand$mean_tau_hat, m_diff))
      .tc("Sandwich subgroup: chosen n_subgroup >= n_min",
          sg_sand$n_subgroup >= SMOKE$n_min, family,
          details = sprintf("n = %d, n_min = %d",
                            sg_sand$n_subgroup, SMOKE$n_min))
      .tc("Sandwich subgroup: SE is positive and finite",
          is.finite(sg_sand$se_mean_tau_hat) &&
            sg_sand$se_mean_tau_hat > 0, family)
      .tc("Sandwich subgroup: CI lower <= mean <= CI upper",
          sg_sand$ci[["lower"]] <= sg_sand$mean_tau_hat &&
            sg_sand$mean_tau_hat <= sg_sand$ci[["upper"]], family)
      .tc("Sandwich subgroup: mask is logical of length n",
          is.logical(sg_sand$mask) && length(sg_sand$mask) == SMOKE$n &&
            sum(sg_sand$mask) == sg_sand$n_subgroup, family)
      .tc("Sandwich subgroup: chosen covariate is among input covariates",
          sg_sand$covariate %in% covariates_v, family)
      .tc("Sandwich subgroup: threshold within range of chosen covariate",
          {
            v <- df_fam[[sg_sand$covariate]]
            sg_sand$threshold >= min(v) && sg_sand$threshold <= max(v)
          }, family)
    } else {
      .tc(sprintf("Sandwich subgroup: NOT found at m_diff = %.2f (informational)",
                  m_diff),
          TRUE, family,
          details = "No qualifying subgroup; subsequent IJ check independent.")
    }
  }

  # ---- Structural assertions on the IJ result -----------------------------
  if (ij_ok) {
    .tc("IJ subgroup: returns object of class 'dina_subgroup'",
        inherits(sg_ij, "dina_subgroup"), family)
    if (isTRUE(sg_ij$found)) {
      .tc("IJ subgroup: chosen mean_tau_hat >= m_diff",
          sg_ij$mean_tau_hat >= m_diff - 1e-10, family,
          details = sprintf("mean = %.3f, m_diff = %.3f",
                            sg_ij$mean_tau_hat, m_diff))
      .tc("IJ subgroup: SE is non-negative and finite",
          is.finite(sg_ij$se_mean_tau_hat) &&
            sg_ij$se_mean_tau_hat >= 0, family)
      .tc("IJ subgroup: CI lower <= mean <= CI upper",
          sg_ij$ci[["lower"]] <= sg_ij$mean_tau_hat &&
            sg_ij$mean_tau_hat <= sg_ij$ci[["upper"]], family)
    }
  }

  # ---- Cross-fit agreement: both search results pick the same covariate ---
  # (Same data, same m_diff, slightly different coefficient estimates between
  #  sandwich and IJ paths.  The chosen covariate should typically agree.)
  if (sand_ok && ij_ok && isTRUE(sg_sand$found) && isTRUE(sg_ij$found)) {
    .tc("Sandwich and IJ subgroup searches pick the same covariate",
        sg_sand$covariate == sg_ij$covariate, family,
        details = sprintf("sandwich = %s, ij = %s",
                          sg_sand$covariate, sg_ij$covariate))
  }

  # ---- Negative control: impossibly large m_diff => found == FALSE --------
  if (!is.null(fit_sandwich)) {
    sg_null <- tryCatch(
      dina_subgroup(fit = fit_sandwich, df = df_fam,
                    covariates = covariates_v,
                    m_diff = 100, n_min = SMOKE$n_min,
                    alpha = SMOKE$alpha),
      error = function(e) NULL
    )
    .tc("Negative control: large m_diff returns found == FALSE",
        !is.null(sg_null) && isFALSE(sg_null$found), family,
        details = if (!is.null(sg_null))
          sprintf("found = %s", as.character(sg_null$found))
        else "errored")
  }

  list(
    family       = family,
    m_diff       = m_diff,
    sg_sandwich  = sg_sand,
    sg_ij        = sg_ij,
    fit_sandwich = fit_sandwich,
    fit_ij       = fit_ij
  )
}

# Per-family gt table comparing the two subgroup-search results
family_subgroup_gt <- function(res, family) {

  pull <- function(sg, field) {
    if (is.null(sg) || inherits(sg, "error")) return(NA)
    val <- sg[[field]]
    if (is.null(val)) NA else val
  }

  fmt_ci <- function(sg) {
    if (is.null(sg) || inherits(sg, "error") || !isTRUE(sg$found)) return("—")
    sprintf("[%.3f, %.3f]", sg$ci[["lower"]], sg$ci[["upper"]])
  }

  rows <- data.table(
    Field = c("Found?", "Covariate", "Direction", "Threshold",
              "n_subgroup", "mean_tau_hat", "SE", "Wald CI",
              "Candidates searched", "Candidates qualifying"),
    `Sandwich (dina)` = c(
      as.character(pull(res$sg_sandwich, "found")),
      as.character(pull(res$sg_sandwich, "covariate") %||% "—"),
      as.character(pull(res$sg_sandwich, "direction") %||% "—"),
      sprintf("%.3f", pull(res$sg_sandwich, "threshold")),
      sprintf("%d",   pull(res$sg_sandwich, "n_subgroup")),
      sprintf("%.3f", pull(res$sg_sandwich, "mean_tau_hat")),
      sprintf("%.3f", pull(res$sg_sandwich, "se_mean_tau_hat")),
      fmt_ci(res$sg_sandwich),
      sprintf("%d",   pull(res$sg_sandwich, "n_candidates_searched")),
      sprintf("%d",   pull(res$sg_sandwich, "n_candidates_qualifying"))
    ),
    `IJ (dina_bagged)` = c(
      as.character(pull(res$sg_ij, "found")),
      as.character(pull(res$sg_ij, "covariate") %||% "—"),
      as.character(pull(res$sg_ij, "direction") %||% "—"),
      sprintf("%.3f", pull(res$sg_ij, "threshold")),
      sprintf("%d",   pull(res$sg_ij, "n_subgroup")),
      sprintf("%.3f", pull(res$sg_ij, "mean_tau_hat")),
      sprintf("%.3f", pull(res$sg_ij, "se_mean_tau_hat")),
      fmt_ci(res$sg_ij),
      sprintf("%d",   pull(res$sg_ij, "n_candidates_searched")),
      sprintf("%d",   pull(res$sg_ij, "n_candidates_qualifying"))
    )
  )

  rows |>
    gt::gt() |>
    gt::tab_header(
      title    = sprintf("%s family — dina_subgroup() results",
                         tools::toTitleCase(family)),
      subtitle = sprintf("m_diff = %.2f; harm direction = positive tau on the natural-parameter scale",
                         res$m_diff)
    ) |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_source_note(
      source_note = gt::md(
        "*True heterogeneity is on `x1` only (β₁ = 1.20); a right-direction
         subgroup at high x1 is the expected discovery.*"
      )
    )
}
```

---

# Gaussian family

```{r family-gaussian}
res_gauss <- run_subgroup_family_tests("gaussian")
family_subgroup_gt(res_gauss, "gaussian")
```

---

# Binomial family

```{r family-binomial}
res_binom <- run_subgroup_family_tests("binomial")
family_subgroup_gt(res_binom, "binomial")
```

---

# Poisson family

```{r family-poisson}
res_pois <- run_subgroup_family_tests("poisson")
family_subgroup_gt(res_pois, "poisson")
```

---

# Cox family

```{r family-cox}
res_cox <- run_subgroup_family_tests("cox")
family_subgroup_gt(res_cox, "cox")
```

```{r plan-teardown}
future::plan("sequential")
gc(verbose = FALSE)
```

---

# Summary

```{r summary}
results_tab <- .tc_table()
n_total <- nrow(results_tab)
n_pass  <- results_tab[status == "PASS", .N]
n_fail  <- n_total - n_pass

subtitle_text <- sprintf(
  "%d / %d assertions passed (%d failures)", n_pass, n_total, n_fail
)

results_tab |>
  gt::gt(groupname_col = "family") |>
  gt::tab_header(
    title    = gt::md("**DINA subgroup smoke-test summary**"),
    subtitle = subtitle_text
  ) |>
  gt::cols_label(
    label   = "Assertion",
    status  = "Status",
    details = "Details"
  ) |>
  gt::tab_style(
    style    = gt::cell_fill(color = "#d4f4dd"),
    locations = gt::cells_body(rows = status == "PASS")
  ) |>
  gt::tab_style(
    style    = gt::cell_fill(color = "#f9d4d4"),
    locations = gt::cells_body(rows = status == "FAIL")
  ) |>
  gt::tab_style(
    style    = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |>
  gt::tab_options(table.font.size = 12)
```

---

# Session information

```{r session-info}
sessionInfo()
```
