---
title: "Smoke test: DINA screening in forestsearch() (use_dina)"
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: true
  message: false
---

This living doc exercises the Phase-2 `use_dina` integration end to end on
synthetic data, across **all three outcome families** that map to a DINA fit:
survival (cox), binary (binomial), and continuous (gaussian). Each family runs
the same battery of checks via the canonical `.tc()` helper, and the results are
rendered as a `gt` summary at the end.

The DINA path is isolated with `use_grf = FALSE` and `use_lasso = FALSE`, so the
candidate pool is just median cuts plus the DINA per-covariate frontier cuts and
the test needs no `grf`/`glmnet`. Runs are sequential and deliberately light
(`maxk = 2`, reduced `fs.splits`, `max.minutes = 1`); the DINA cuts are produced
during screening, before the consistency search, so the assertions do not depend
on a subgroup actually being found.

```{r}
#| label: setup
library(forestsearch)

# --- .tc() accumulator -----------------------------------------------------
TC <- local({
  rows <- list()
  add <- function(check, result, note = "") {
    rows[[length(rows) + 1L]] <<- data.frame(
      check = check, result = result, note = note, stringsAsFactors = FALSE)
    invisible()
  }
  list(
    tc   = function(check, cond, note = "")
      add(check, if (isTRUE(cond)) "PASS" else "FAIL", note),
    skip = function(check, note = "") add(check, "SKIP", note),
    table = function() do.call(rbind, rows)
  )
})
.tc   <- TC$tc
.skip <- TC$skip

ok     <- function(x) !inherits(x, "error")
errmsg <- function(x) if (inherits(x, "error")) conditionMessage(x) else ""
varof  <- function(cuts) sub(" .*", "", cuts)     # "x1 <= 0.3" -> "x1"
```

```{r}
#| label: dgm
# All DGMs share a known harmful subgroup at x1 > 0.3 (treatment worsens the
# outcome there). x2/x3 are weak prognostic; x4/x5 are noise.
.mkX <- function(n) matrix(runif(n * 5, -1, 1), n, 5,
                           dimnames = list(NULL, paste0("x", 1:5)))

make_surv <- function(n = 400L, seed = 101L) {
  set.seed(seed); X <- .mkX(n); treat <- rbinom(n, 1, 0.5)
  harm <- as.integer(X[, "x1"] > 0.3)
  rate <- 0.08 * exp(0.4 * X[, "x2"] - 0.3 * X[, "x3"] +
                       treat * (-0.4 + 1.2 * harm))
  tt <- rexp(n, rate); cc <- rexp(n, 0.02)
  data.frame(id = seq_len(n), y = pmin(tt, cc),
             d = as.integer(tt <= cc), treat = treat, X)
}

make_bin <- function(n = 400L, seed = 303L) {
  set.seed(seed); X <- .mkX(n); treat <- rbinom(n, 1, 0.5)
  harm <- as.integer(X[, "x1"] > 0.3)
  p <- plogis(0.3 * X[, "x2"] - 0.2 * X[, "x3"] +
                treat * (-0.4 + 1.2 * harm))
  data.frame(id = seq_len(n), y = rbinom(n, 1, p), treat = treat, X)
}

make_cont <- function(n = 400L, seed = 202L) {
  set.seed(seed); X <- .mkX(n); treat <- rbinom(n, 1, 0.5)
  harm <- as.integer(X[, "x1"] > 0.3)
  y <- 0.5 * X[, "x2"] + treat * (-0.3 + 1.0 * harm) + rnorm(n)
  data.frame(id = seq_len(n), y = y, treat = treat, X)
}

dsurv <- make_surv(); dbin <- make_bin(); dcont <- make_cont()
confs <- paste0("x", 1:5)
```

```{r}
#| label: battery
# One battery of checks, run per outcome family.  harm_m_diff is on the family's
# natural-parameter (link) scale: log-HR / log-OR for ratio families, mean
# difference for gaussian.
battery <- function(label, df, outcome_type, expected_family, harm_m_diff,
                    event.name = "y") {
  base <- list(
    df.analysis = df, outcome.name = "y", event.name = event.name,
    treat.name = "treat", id.name = "id", confounders.name = confs,
    outcome_type = outcome_type, use_lasso = FALSE, use_grf = FALSE,
    parallel_args = list(plan = "sequential"),
    maxk = 2, fs.splits = 100, max.minutes = 1, details = FALSE)
  run <- function(extra = list())
    tryCatch(do.call(forestsearch, modifyList(base, extra)),
             error = function(e) e)
  p <- function(s) paste0(label, ": ", s)

  # Backward compatibility
  f0 <- run(list(use_dina = FALSE))
  .tc(p("use_dina=FALSE runs"), ok(f0), errmsg(f0))
  .tc(p("use_dina=FALSE leaves dina_cuts NULL"), ok(f0) && is.null(f0$dina_cuts))

  # Core DINA screening
  f1 <- run(list(use_dina = TRUE)); dc <- if (ok(f1)) f1$dina_cuts else NULL
  .tc(p("use_dina=TRUE runs"), ok(f1), errmsg(f1))
  .tc(p("dina_cuts non-empty character"), is.character(dc) && length(dc) >= 1L)
  .tc(p("dina_cuts canonical 'var <= value'"),
      length(dc) >= 1L && all(grepl("^x[1-5] <= -?[0-9.]+$", dc)))
  .tc(p("dina_cuts reference real confounders"),
      length(dc) >= 1L && all(varof(dc) %in% confs))
  .tc(p("dina_cuts span >1 covariate"),
      length(dc) >= 1L && length(unique(varof(dc))) > 1L)
  .tc(p("dina_res family mapped correctly"),
      ok(f1) && inherits(f1$dina_res, "dina") &&
        identical(f1$dina_res$family, expected_family))

  # dina_args caps
  fc <- run(list(use_dina = TRUE,
                 dina_args = list(max_subgroups = 3L, max_per_covariate = 1L)))
  dcc <- if (ok(fc)) fc$dina_cuts else NULL
  .tc(p("max_subgroups caps total (<= 3)"), ok(fc) && length(dcc) <= 3L)
  .tc(p("max_per_covariate respected (<= 1)"),
      ok(fc) && (length(dcc) == 0L || all(table(varof(dcc)) <= 1L)))

  # Harm-gated scope
  fh <- run(list(use_dina = TRUE,
                 dina_args = list(scope = "harm", m_diff = harm_m_diff)))
  .tc(p("scope='harm' runs"), ok(fh), errmsg(fh))
  .tc(p("scope='harm' returns character dina_cuts"),
      ok(fh) && is.character(fh$dina_cuts))

  # Pre-supplied cuts skip the fit
  fp <- run(list(use_dina = TRUE, dina_cuts = c("x1 <= 0.3")))
  .tc(p("pre-supplied dina_cuts used verbatim"),
      ok(fp) && identical(fp$dina_cuts, "x1 <= 0.3"))
  .tc(p("pre-supplied => dina_res NULL"), ok(fp) && is.null(fp$dina_res))

  invisible(NULL)
}
```

## Survival (cox)

```{r}
#| label: survival
battery("survival", dsurv, "survival", "cox", log(1.25), event.name = "d")
```

## Binary (binomial)

```{r}
#| label: binary
battery("binary", dbin, "binary", "binomial", log(1.25), event.name = "y")
```

## Continuous (gaussian)

```{r}
#| label: continuous
# Continuous has no event column; forestsearch() placeholders the (absent) one.
battery("continuous", dcont, "continuous", "gaussian", 0.2, event.name = "event")
```

## Screening diagnostics (`details = TRUE`)

```{r}
#| label: details
cap <- capture.output(
  tryCatch(forestsearch(
    df.analysis = dsurv, outcome.name = "y", event.name = "d",
    treat.name = "treat", id.name = "id", confounders.name = confs,
    outcome_type = "survival", use_lasso = FALSE, use_grf = FALSE,
    parallel_args = list(plan = "sequential"),
    maxk = 2, fs.splits = 100, max.minutes = 1,
    use_dina = TRUE, details = TRUE),
    error = function(e) e))
.tc("details: 'DINA cuts identified' line printed",
    any(grepl("DINA cuts identified", cap)))
```

## Coexistence with GRF (only if `grf` is installed)

```{r}
#| label: grf-coexist
if (requireNamespace("grf", quietly = TRUE)) {
  fg <- tryCatch(forestsearch(
    df.analysis = dsurv, outcome.name = "y", event.name = "d",
    treat.name = "treat", id.name = "id", confounders.name = confs,
    outcome_type = "survival", use_lasso = FALSE, use_grf = TRUE,
    parallel_args = list(plan = "sequential"),
    maxk = 2, fs.splits = 100, max.minutes = 1, use_dina = TRUE),
    error = function(e) e)
  .tc("GRF + DINA run together", ok(fg), errmsg(fg))
  .tc("both grf_cuts and dina_cuts populated",
      ok(fg) && length(fg$grf_cuts) >= 1L && length(fg$dina_cuts) >= 1L)
} else {
  .skip("GRF + DINA run together", "grf not installed")
  .skip("both grf_cuts and dina_cuts populated", "grf not installed")
}
```

## Summary

```{r}
#| label: summary
res <- TC$table()
n_pass <- sum(res$result == "PASS")
n_fail <- sum(res$result == "FAIL")
n_skip <- sum(res$result == "SKIP")
cat(sprintf("PASS: %d   FAIL: %d   SKIP: %d\n", n_pass, n_fail, n_skip))

if (requireNamespace("gt", quietly = TRUE)) {
  library(gt)
  gt(res) |>
    tab_header(title = gt::md("**DINA screening smoke test**"),
               subtitle = gt::md(sprintf("%d pass / %d fail / %d skip",
                                         n_pass, n_fail, n_skip))) |>
    cols_label(check = "Check", result = "Result", note = "Note") |>
    tab_style(style = cell_fill(color = "#d9f2d9"),
              locations = cells_body(rows = result == "PASS")) |>
    tab_style(style = cell_fill(color = "#ffd6d6"),
              locations = cells_body(rows = result == "FAIL")) |>
    tab_style(style = cell_fill(color = "#fff2cc"),
              locations = cells_body(rows = result == "SKIP"))
} else {
  knitr::kable(res)
}
```
