---
title: "The mean-difference closed-form fixture"
subtitle: "Binding fs_mr_inference() to values computable independently"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: true
  message: true
---

## What this document is

`dev/identifier-alignment/verification/` establishes that the closed forms are
**correct** --- the algebra gives the numbers it claims. Nothing there fails if
`forestsearch()` regresses, because none of it touches the pipeline. This
document is the missing half: does `fs_mr_inference()`, run end to end,
reproduce those values?

Three rules, from `SPEC_closedform_fixture.md`, each recording a mistake already
made once in this project:

1. **No re-implementing the pipeline.** Every comparison is against a value the
   package returned. Where a closed form is computed here it is computed from
   the *observed* per-candidate objects, never from a parallel construction of
   $D_g(b)$.
2. **No asserting at the leaf.** Each level contains at least one assertion that
   runs the top-level function and reads its returned object, so the fixture
   proves the value *arrives* rather than merely that some helper can compute it.
3. **No control that passes because the quantity is easy.** Every scenario is
   configured so the target is distinguishable from zero, from $\sigma_D$, and
   from the naive estimate.

The MR correction is settled as of `dad0415`. This document asserts **against**
it, not about it. A failing assertion is a finding, not a licence to adjust the
code.

### Why the mean difference

The MD score is linear in $(\alpha, \beta)$, so Lemma 1's mean-value curvature
term vanishes identically --- the linearization carries no curvature error and
every per-candidate object is closed form. It is also the only measure fitted
with `lm()`, the path `411a448` repaired: a $\sigma_D$ assertion fails **on
contact** if that path breaks again, rather than degrading silently to
`n_family = 0`.

```{r setup}
#| include: false
library(forestsearch)
library(gt)

mr  <- forestsearch:::fs_mr_inference
asm <- forestsearch:::.fs_mr_assemble
ijv <- forestsearch:::.fs_mr_ij_var

SPEC_MD <- list(outcome_type = "continuous", effect_measure = "MD",
                treat.name = "trt", outcome.name = "y", event.name = NULL,
                offset.name = NULL, adjust_covariates = NULL,
                adverse_outcome = TRUE)
NOFLOOR <- list(effect_floor = NULL, consistency = NULL)

DRAWS <- 20000L          # spec: B >= 20000 keeps the O(B^-1/2) envelope inside
Z975  <- stats::qnorm(0.975)

# Deterministic centered scores: constructed, not sampled, so arm means and arm
# variances are fixed before any fit runs.
fx_z <- function(m) { z <- stats::qnorm((seq_len(m) - 0.5) / m); z - mean(z) }

fx_blocks <- function(n_per_arm, delta, s) {
  out <- NULL
  for (b in seq_along(delta)) {
    zb <- fx_z(n_per_arm)
    out <- rbind(out, data.frame(trt = rep(c(1L, 0L), each = n_per_arm),
                                 y = c(delta[b] + s[b] * zb, s[b] * zb)))
  }
  out$id <- seq_len(nrow(out))
  k <- 2 * n_per_arm
  list(df = out,
       rows = lapply(seq_along(delta), function(b) ((b - 1) * k + 1):(b * k)))
}

# Closed forms. The (n_a - 1) denominators are the one-step influence
# (X'X)^-1 x_i e_i / (1 - h_i) with h_i = 1/n_a; see @sec-findings.
fx_cf_beta <- function(df, idx) {
  d <- df[idx, ]; mean(d$y[d$trt == 1L]) - mean(d$y[d$trt == 0L])
}
fx_cf_db <- function(df, idx) {
  d <- df[idx, ]
  y1 <- d$y[d$trt == 1L]; y0 <- d$y[d$trt == 0L]
  n1 <- length(y1); n0 <- length(y0)
  db <- numeric(nrow(d))
  db[d$trt == 1L] <-  (y1 - mean(y1)) / (n1 - 1)
  db[d$trt == 0L] <- -(y0 - mean(y0)) / (n0 - 1)
  db
}
fx_cf_sigma <- function(df, idx) sqrt(sum(fx_cf_db(df, idx)^2))

# Pass/fail bookkeeping. Levels are gated: Level 1 must pass before Level 2
# means anything, and Level 2 before Level 3.
LEVEL_OK <- c(L1 = TRUE, L2 = TRUE, L3 = TRUE)
mark <- function(level, ok) {
  if (!isTRUE(ok)) LEVEL_OK[[level]] <<- FALSE
  if (isTRUE(ok)) "PASS" else "**FAIL**"
}
gate <- function(level) {
  if (!LEVEL_OK[[level]])
    stop("Level ", level, " failed; later levels are not meaningful. ",
         "Report levels in order and stop on the first failure.", call. = FALSE)
  invisible(TRUE)
}
tbl <- function(d, title, subtitle = NULL) {
  g <- gt(d) |> tab_header(title = title, subtitle = subtitle) |>
    cols_align(align = "left", columns = 1)
  g
}
```

## The four scenarios

| | shape | isolates |
|---|---|---|
| **S1** | two candidates, floor inactive | the main selection-bias check |
| **S2** | $M = 1$, floor active | the conditional denominator convention --- the load-bearing scenario |
| **S3** | $M = 1$, floor inactive | the degenerate IJ case, $\widetilde V/\sigma^2_D \to 4$ |
| **S4** | $M$ disjoint candidates, no separation | the $\mathbb{E}[\max]$ constants |

`multiplier = "gaussian"` throughout. The package default is `"poisson"`; under
Gaussian multipliers $D_g(b)$ is exactly multivariate normal at every $n$, which
is what every closed form below assumes. A fixture left on the Poisson default
misses by a few percent and the miss looks like a defect.

```{r build-frames}
#| code-fold: true

# S1 -- two OVERLAPPING candidates, so Sigma_12 != 0 and tau is not trivial.
s1  <- fx_blocks(40, delta = c(1.2, 0.4), s = c(1.0, 1.3))
s1_cands <- list(
  g1 = c(s1$rows[[1]], s1$rows[[2]][1:20], s1$rows[[2]][41:60]),
  g2 = c(s1$rows[[2]], s1$rows[[1]][1:20], s1$rows[[1]][41:60]))

# S3 / S2 -- a single candidate.
s3  <- fx_blocks(60, delta = 1.0, s = 1.0)
s3_idx <- s3$rows[[1]]
```

# Level 1 --- the per-candidate objects {#sec-level1}

Algebraic identities, so equality at `1e-10` rather than an envelope. A failure
here is a broken influence path, which is `411a448`'s signature.

```{r level1}
#| code-fold: true
A1 <- asm(s1$df, s1_cands, SPEC_MD)
Sig <- crossprod(A1$B)

rows <- list()
for (g in seq_along(s1_cands)) {
  idx <- s1_cands[[g]]
  ok_b <- isTRUE(all.equal(A1$beta_hat[g], fx_cf_beta(s1$df, idx), tolerance = 1e-10))
  ok_s <- isTRUE(all.equal(A1$sigma_D[g],  fx_cf_sigma(s1$df, idx), tolerance = 1e-10))
  ok_B <- isTRUE(all.equal(unname(A1$B[idx, g]), unname(fx_cf_db(s1$df, idx)),
                           tolerance = 1e-10))
  tr <- s1$df$trt[idx]
  ok_z <- max(abs(c(sum(A1$B[idx, g][tr == 1L]), sum(A1$B[idx, g][tr == 0L])))) < 1e-10
  rows[[length(rows) + 1]] <- data.frame(
    quantity = sprintf("beta_hat(g%d)", g),
    closed_form = fx_cf_beta(s1$df, idx), observed = A1$beta_hat[g],
    abs_diff = abs(A1$beta_hat[g] - fx_cf_beta(s1$df, idx)),
    result = mark("L1", ok_b))
  rows[[length(rows) + 1]] <- data.frame(
    quantity = sprintf("sigma_D(g%d)", g),
    closed_form = fx_cf_sigma(s1$df, idx), observed = A1$sigma_D[g],
    abs_diff = abs(A1$sigma_D[g] - fx_cf_sigma(s1$df, idx)),
    result = mark("L1", ok_s))
  rows[[length(rows) + 1]] <- data.frame(
    quantity = sprintf("B[, g%d] = e_i/(n_a - 1)", g),
    closed_form = NA_real_, observed = NA_real_,
    abs_diff = max(abs(A1$B[idx, g] - fx_cf_db(s1$df, idx))),
    result = mark("L1", ok_B))
  rows[[length(rows) + 1]] <- data.frame(
    quantity = sprintf("within-arm sums of B[, g%d]", g),
    closed_form = 0, observed = NA_real_,
    abs_diff = max(abs(c(sum(A1$B[idx, g][tr == 1L]), sum(A1$B[idx, g][tr == 0L])))),
    result = mark("L1", ok_z))
}

Bcf <- matrix(0, nrow(s1$df), length(s1_cands))
for (g in seq_along(s1_cands)) Bcf[s1_cands[[g]], g] <- fx_cf_db(s1$df, s1_cands[[g]])
ok_S <- isTRUE(all.equal(unname(Sig), unname(crossprod(Bcf)), tolerance = 1e-10))
rows[[length(rows) + 1]] <- data.frame(
  quantity = "Sigma = crossprod(B)", closed_form = NA_real_, observed = NA_real_,
  abs_diff = max(abs(Sig - crossprod(Bcf))), result = mark("L1", ok_S))

L1 <- do.call(rbind, rows)
tbl(L1, "Level 1 --- per-candidate objects (S1 frame)",
    "Equality at 1e-10. Read from .fs_mr_assemble().") |>
  fmt_number(columns = c("closed_form", "observed"), decimals = 6) |>
  fmt_scientific(columns = "abs_diff", decimals = 2) |>
  sub_missing(missing_text = "--")
```

## Level 1 reaches the top level

Reading `.fs_mr_assemble()` proves the assembler works; it does not prove
`fs_mr_inference()` calls it, or that the value reaches `debiased$se_wald`.

```{r level1-toplevel}
#| code-fold: true
r1 <- mr(s1$df, s1_cands, SPEC_MD, selected_members = s1_cands[[1]],
         admission = NOFLOOR, reselection = "maxeff",
         draws = DRAWS, multiplier = "gaussian", seed = 42L)

ok_tl <- isTRUE(all.equal(r1$debiased$se_wald, fx_cf_sigma(s1$df, s1_cands[[1]]),
                          tolerance = 1e-10))
L1b <- data.frame(
  quantity = c("debiased$se_wald vs closed-form sigma_D(g1)", "n_family"),
  closed_form = c(fx_cf_sigma(s1$df, s1_cands[[1]]), 2),
  observed = c(r1$debiased$se_wald, r1$n_family),
  abs_diff = c(abs(r1$debiased$se_wald - fx_cf_sigma(s1$df, s1_cands[[1]])), 0),
  result = c(mark("L1", ok_tl), mark("L1", r1$n_family == 2L)))
tbl(L1b, "Level 1 --- top-level tie-in",
    "fs_mr_inference() run end to end; values read from its returned object.") |>
  fmt_number(columns = c("closed_form", "observed"), decimals = 6) |>
  fmt_scientific(columns = "abs_diff", decimals = 2)
```

```{r gate1}
#| echo: false
gate("L1")
```

# Level 2 --- the correction {#sec-level2}

## S1 --- two candidates, floor inactive

With $d = \widehat\beta(g_1) - \widehat\beta(g_2)$ and
$\tau^2 = \sigma_1^2 + \sigma_2^2 - 2\sigma_{12}$, both taken from the Level 1
objects,
$$\widehat{\mathrm{bias}}_{\mathrm{sel}} \longrightarrow \tau\,\varphi(d/\tau).$$

```{r s1-bias}
#| code-fold: true
d_s1   <- A1$beta_hat[1] - A1$beta_hat[2]
tau_s1 <- sqrt(Sig[1, 1] + Sig[2, 2] - 2 * Sig[1, 2])
cf_s1  <- tau_s1 * stats::dnorm(d_s1 / tau_s1)
env_s1 <- 4 * max(A1$sigma_D) / sqrt(DRAWS)
err_s1 <- r1$selection_bias - cf_s1

S1 <- data.frame(
  quantity = c("d = beta1 - beta2", "tau", "selection_bias"),
  closed_form = c(d_s1, tau_s1, cf_s1),
  observed = c(NA_real_, NA_real_, r1$selection_bias),
  error = c(NA_real_, NA_real_, err_s1),
  envelope = c(NA_real_, NA_real_, env_s1),
  selection_rate = c(NA_real_, NA_real_, r1$selection_rate),
  result = c("--", "--", mark("L2", abs(err_s1) <= env_s1)))
tbl(S1, "S1 --- selection bias against tau * phi(d / tau)",
    sprintf("B = %d, gaussian multipliers. Envelope 4 max sigma_j B^-1/2.", DRAWS)) |>
  fmt_number(columns = c("closed_form", "observed", "error", "envelope",
                         "selection_rate"), decimals = 6) |>
  sub_missing(missing_text = "--")
```

The target is `r sprintf("%.4f", cf_s1)`, against $\sigma_D =$
`r sprintf("%.4f", A1$sigma_D[1])` and a naive estimate of
`r sprintf("%.4f", A1$beta_hat[1])`. It is distinguishable from zero, from
$\sigma_D$, and from the naive value, so a pipeline that computed nothing would
not pass.

### Rate check --- the error must fall at $B^{-1/2}$

A systematic offset that does not shrink is the failure a single comparison
cannot see. Averaged over eight seeds per $B$, since a single seed's error is
itself noise.

```{r s1-rate}
#| code-fold: true
seeds <- 101:108
rate_rows <- lapply(c(DRAWS, 4L * DRAWS, 16L * DRAWS), function(B) {
  errs <- vapply(seeds, function(sd_) {
    rr <- mr(s1$df, s1_cands, SPEC_MD, selected_members = s1_cands[[1]],
             admission = NOFLOOR, reselection = "maxeff",
             draws = B, multiplier = "gaussian", seed = sd_)
    rr$selection_bias - cf_s1
  }, numeric(1))
  data.frame(B = B, rms_error = sqrt(mean(errs^2)))
})
RT <- do.call(rbind, rate_rows)
RT$ratio_vs_first <- RT$rms_error / RT$rms_error[1]
RT$predicted <- sqrt(RT$B[1] / RT$B)
ok_rate <- RT$rms_error[3] < RT$rms_error[1]
RT$result <- c("--", "--", mark("L2", ok_rate))
tbl(RT, "S1 rate check --- RMS error over 8 seeds",
    "predicted = sqrt(B_1 / B), the B^-1/2 rate.") |>
  fmt_number(columns = c("rms_error"), decimals = 7) |>
  fmt_number(columns = c("ratio_vs_first", "predicted"), decimals = 4)
```

### Tie sanity --- $d = 0$, orthogonal unit-scale influences

At $d = 0$ with $\Sigma_{12} = 0$ and $\sigma_1 = \sigma_2 = \sigma$,
$\tau = \sqrt2\,\sigma$ and $\tau\varphi(0) = \sigma/\sqrt\pi$ --- the expected
maximum of two normals.

```{r s1-tie}
#| code-fold: true
tie <- fx_blocks(40, delta = c(1.0, 1.0), s = c(1.0, 1.0))
tie_c <- list(g1 = tie$rows[[1]], g2 = tie$rows[[2]])
At <- asm(tie$df, tie_c, SPEC_MD)
rt <- mr(tie$df, tie_c, SPEC_MD, selected_members = tie_c[[1]],
         admission = NOFLOOR, reselection = "maxeff",
         draws = DRAWS, multiplier = "gaussian", seed = 5L)
cf_tie <- mean(At$sigma_D) / sqrt(pi)
env_tie <- 4 * max(At$sigma_D) / sqrt(DRAWS)
ok_tie <- abs(rt$selection_bias - cf_tie) <= env_tie
TIE <- data.frame(quantity = "selection_bias at d = 0",
                  closed_form = cf_tie, observed = rt$selection_bias,
                  error = rt$selection_bias - cf_tie, envelope = env_tie,
                  selection_rate = rt$selection_rate,
                  result = mark("L2", ok_tie))
tbl(TIE, "S1 tie sanity --- sigma / sqrt(pi)") |>
  fmt_number(columns = 2:6, decimals = 6)
```

## S2 --- $M = 1$, floor active: the denominator convention {#sec-s2}

The load-bearing scenario. With $M = 1$, $c = t_g - \widehat\beta(g)$ and
$\sigma = \sigma_{D,g}$, the **conditional** convention the package implements
as of `dad0415` gives
$$\widehat{\mathrm{bias}}_{\mathrm{sel}} \longrightarrow
  \frac{\sigma\,\varphi(c/\sigma)}{\overline\Phi(c/\sigma)},
  \qquad p = \overline\Phi(c/\sigma),$$
against the **unconditional** divide-by-$B$ value $\sigma\varphi(c/\sigma)$.
The two differ by exactly $1/p$. A fixture that only checks the value the code
produces cannot tell a convention change from a regression, so S2 asserts both
that the conditional form holds **and** that the unconditional form does not.

```{r s2}
#| code-fold: true
A3   <- asm(s3$df, list(g1 = s3_idx), SPEC_MD)
sg   <- A3$sigma_D[1]; bh <- A3$beta_hat[1]
zcs  <- c(0.8416, 0.4307, 0.0)          # p ~ 0.20, 0.33, 0.50

s2_rows <- lapply(zcs, function(zc) {
  fl <- bh + zc * sg
  rr <- mr(s3$df, list(g1 = s3_idx), SPEC_MD, selected_members = s3_idx,
           admission = list(effect_floor = fl, consistency = NULL),
           reselection = "maxeff", draws = DRAWS, multiplier = "gaussian",
           seed = 7L)
  p_obs  <- rr$selection_rate
  cond   <- sg * stats::dnorm(zc) / stats::pnorm(zc, lower.tail = FALSE)
  uncond <- sg * stats::dnorm(zc)
  env    <- 4 * sg / sqrt(DRAWS * p_obs)
  ok_c   <- abs(rr$selection_bias - cond) <= env
  ok_u   <- abs(rr$selection_bias - uncond) > env     # must NOT match
  data.frame(c_over_sigma = zc, p = p_obs,
             conditional = cond, observed = rr$selection_bias,
             error = rr$selection_bias - cond, envelope = env,
             unconditional = uncond, ratio_cond_uncond = cond / uncond,
             result = mark("L2", ok_c && ok_u))
})
S2 <- do.call(rbind, s2_rows)
tbl(S2, "S2 --- conditional closed form asserted, unconditional asserted to fail",
    sprintf("B = %d. Envelope 4 sigma / sqrt(B p) -- the average is over Bp draws, not B.", DRAWS)) |>
  fmt_number(columns = 2:8, decimals = 6)
```

The separation is `r sprintf("%.2f", S2$ratio_cond_uncond[2])`-fold at
$c = \sigma/2$, so the two conventions are not distinguishable only in
principle.

### `fixed_bias`: the inverted pair {#sec-inverted}

The mean-zero envelope
$|\widehat{\mathrm{bias}}_{\mathrm{fix}}| \le z_{0.975}\,\sigma_D B^{-1/2}$
holds at $p = 1$ **and only at $p = 1$**. Under conditioning, admission selects
draws with upward perturbation, so $\widehat{\mathrm{bias}}_{\mathrm{fix}}$ is a
genuine selection effect and must *breach* the envelope.

This inverted pair is the strongest assertion in the fixture: an implementation
that reverted to the unconditional convention would pass at rate 1 and fail S2,
and nothing else here detects that.

```{r fixed-bias-pair}
#| code-fold: true
env_mz <- function(sd_, B) Z975 * sd_ / sqrt(B)

fb_rows <- list()
for (nm in c("S1", "S3", "S4 (M=2)")) {
  rr <- switch(nm,
    "S1" = r1,
    "S3" = mr(s3$df, list(g1 = s3_idx), SPEC_MD, selected_members = s3_idx,
              admission = NOFLOOR, reselection = "maxeff",
              draws = DRAWS, multiplier = "gaussian", seed = 7L),
    "S4 (M=2)" = rt)
  sd_ <- rr$debiased$se_wald
  e <- env_mz(sd_, DRAWS)
  fb_rows[[nm]] <- data.frame(
    scenario = nm, selection_rate = rr$selection_rate,
    fixed_bias = rr$fixed_bias, envelope = e,
    expectation = "within", result = mark("L2", abs(rr$fixed_bias) <= e))
}
for (i in seq_along(zcs)) {
  zc <- zcs[i]; fl <- bh + zc * sg
  rr <- mr(s3$df, list(g1 = s3_idx), SPEC_MD, selected_members = s3_idx,
           admission = list(effect_floor = fl, consistency = NULL),
           reselection = "maxeff", draws = DRAWS, multiplier = "gaussian",
           seed = 7L)
  e <- env_mz(sg, DRAWS)
  fb_rows[[paste0("S2_", i)]] <- data.frame(
    scenario = sprintf("S2 (c/sigma = %.4f)", zc),
    selection_rate = rr$selection_rate,
    fixed_bias = rr$fixed_bias, envelope = e,
    expectation = "breaches", result = mark("L2", abs(rr$fixed_bias) > e))
}
FB <- do.call(rbind, fb_rows)
tbl(FB, "fixed_bias --- mean-zero envelope at rate 1, breach below it",
    "selection_rate reported on every row so a vacuous rate-1 pass is visible as such.") |>
  fmt_number(columns = c("selection_rate", "fixed_bias", "envelope"), decimals = 6)
```

### Monotonicity in $p$, on a fixed draw matrix

`fs_mr_inference()` builds its multipliers from `seed`, so holding the seed,
the data and `draws` fixed while varying only `admission$effect_floor` varies
$p$ **on the same draw matrix**. Redrawing per floor would confound the trend
with sampling noise.

```{r monotone}
#| code-fold: true
zc_grid <- c(1.2816, 0.8416, 0.4307, 0.0, -0.8416)
mono <- do.call(rbind, lapply(zc_grid, function(zc) {
  rr <- mr(s3$df, list(g1 = s3_idx), SPEC_MD, selected_members = s3_idx,
           admission = list(effect_floor = bh + zc * sg, consistency = NULL),
           reselection = "maxeff", draws = DRAWS, multiplier = "gaussian",
           seed = 7L)
  data.frame(c_over_sigma = zc, p = rr$selection_rate,
             fixed_bias = rr$fixed_bias)
}))
mono <- mono[order(mono$p), ]
ok_mono <- all(diff(mono$fixed_bias) < 0)
mono$result <- c(rep("--", nrow(mono) - 1), mark("L2", ok_mono))
tbl(mono, "fixed_bias is monotone decreasing in p, approaching 0 as p -> 1",
    "Same seed, same data, same draws: only the floor varies.") |>
  fmt_number(columns = 1:3, decimals = 6)
```

## S4 --- $M$ disjoint candidates, no separation

The $\Sigma \ge 0$ gate runs **before** the $\mathbb{E}[\max]$ comparison,
since the bound is Slepian's and holds under non-negative pairwise covariance
rather than universally.

```{r s4}
#| code-fold: true
Emax <- c(1 / sqrt(pi), 3 / (2 * sqrt(pi)), 1.02938)
s4 <- do.call(rbind, lapply(2:4, function(M) {
  blM <- fx_blocks(40, delta = rep(1.0, M), s = rep(1.0, M))
  cM  <- stats::setNames(blM$rows, paste0("g", seq_len(M)))
  AM  <- asm(blM$df, cM, SPEC_MD)
  SigM <- crossprod(AM$B)
  min_off <- min(SigM[upper.tri(SigM)])
  gate_ok <- min_off >= -1e-12
  rr <- mr(blM$df, cM, SPEC_MD, selected_members = cM[[1]],
           admission = NOFLOOR, reselection = "maxeff",
           draws = DRAWS, multiplier = "gaussian", seed = 3L)
  cf  <- mean(AM$sigma_D) * Emax[M - 1]
  env <- 4 * max(AM$sigma_D) / sqrt(DRAWS)
  data.frame(M = M, min_offdiag_Sigma = min_off, gate = if (gate_ok) "ok" else "**FAIL**",
             closed_form = cf, observed = rr$selection_bias,
             error = rr$selection_bias - cf, envelope = env,
             selection_rate = rr$selection_rate,
             result = mark("L2", gate_ok && abs(rr$selection_bias - cf) <= env))
}))
tbl(s4, "S4 --- E[max] of M iid normals, after the Sigma >= 0 gate",
    "Constants 1/sqrt(pi), 3/(2 sqrt(pi)), 1.02938 for M = 2, 3, 4.") |>
  fmt_number(columns = c("closed_form", "observed", "error", "envelope",
                         "selection_rate"), decimals = 6) |>
  fmt_scientific(columns = "min_offdiag_Sigma", decimals = 2)
```

```{r gate2}
#| echo: false
gate("L2")
```

# Level 3 --- the interval {#sec-level3}

## S3 --- $\widetilde V / \sigma^2_{D,\widehat H} \to 4$

The assertion that certifies **both** terms of the residual reach the variance.
With $M = 1$ the residual is $r_b = \text{const} - 2 D_1(b)$, so its variance is
$4\sigma^2$. An implementation dropping
$\widehat{\mathrm{bias}}_{\mathrm{fix}}$ returns $1$ here and passes every other
check in this document. Reported at three values of $B$, confirming approach
rather than asserting a point value at one $B$.

```{r s3-ratio}
#| code-fold: true
s2d <- A3$sigma_D[1]^2
S3 <- do.call(rbind, lapply(c(DRAWS, 4L * DRAWS, 16L * DRAWS), function(B) {
  rr <- mr(s3$df, list(g1 = s3_idx), SPEC_MD, selected_members = s3_idx,
           admission = NOFLOOR, reselection = "maxeff",
           draws = B, multiplier = "gaussian", seed = 7L)
  data.frame(B = B, var_ij = rr$debiased$var_ij, sigma2_D = s2d,
             ratio = rr$debiased$var_ij / s2d,
             selection_rate = rr$selection_rate,
             ij_source = rr$debiased$ij_source)
}))
ok_s3 <- S3$ratio[3] > 3.8 && S3$ratio[3] < 4.2 && S3$ratio[3] > S3$ratio[1]
S3$result <- c("--", "--", mark("L3", ok_s3))
tbl(S3, "S3 --- IJ variance ratio approaches 4",
    "Value 1 would indicate a dropped fixed_bias term.") |>
  fmt_number(columns = c("var_ij", "sigma2_D", "ratio", "selection_rate"), decimals = 6)
```

## Row-centering: $\widetilde V$ is invariant to a constant shift in $r$

`.fs_mr_ij_var()` centers the multiplier matrix by row, so
$\sum_b (\Xi - \bar\Xi)_{ib} = 0$ and adding a constant to every $r_b$ cannot
move $\widetilde V$. A wiring check on the centering, not a statistical one.
This binds the **shipped** function, as the reference scripts do --- not a copy.

```{r centering}
#| code-fold: true
set.seed(99)
N_ <- 40L; Bd <- 800L
Xi_ <- matrix(stats::rnorm(N_ * Bd), N_, Bd)
r_  <- stats::rnorm(Bd)
full <- seq_len(Bd)
sub  <- sort(sample(Bd, 300L))          # the selection_rate < 1 case

CEN <- do.call(rbind, lapply(list(list("all draws", full), list("subset (rate < 1)", sub)),
  function(cc) {
    base <- ijv(Xi_, r_, cc[[2]])
    shifted <- vapply(c(-3.5, 0.75, 10), function(s) ijv(Xi_, r_ + s, cc[[2]])$tilde_V,
                      numeric(1))
    data.frame(draws_used = cc[[1]], B_ok = base$B_ok, tilde_V = base$tilde_V,
               max_abs_change = max(abs(shifted - base$tilde_V)),
               result = mark("L3", max(abs(shifted - base$tilde_V)) < 1e-12))
  }))
tbl(CEN, "tilde_V under r -> r + c, for c in {-3.5, 0.75, 10}",
    "Exact invariance follows from row-centering of Xi.") |>
  fmt_number(columns = "tilde_V", decimals = 8) |>
  fmt_scientific(columns = "max_abs_change", decimals = 2)
```

## What this level cannot assert, and why {#sec-rbar-gap}

The spec requires $\bar r$ over the draws the IJ uses, asserted against zero at
$10^{-10}$, with `selection_rate` beside it. **That assertion cannot be made
from the returned object.** `fs_mr_inference()` returns `selection_bias`,
`fixed_bias`, `selection_rate`, and `debiased$var_ij`/`se_ij`/`ij_source`; it
returns neither $r_b$, nor $\Xi$, nor the per-draw winner vector. Recovering
$\bar r$ requires rebuilding $D_g(b)$ inside the fixture, which is the first
thing the spec forbids.

The invariance check above is the strongest available substitute --- it
exercises the shipped IJ arithmetic and would fail if row-centering were
removed --- but it is a check on `.fs_mr_ij_var()`, not on the pipeline's
choice of draw set. Under the conditional convention $\bar r = 0$ is an
algebraic identity of the code as written (`fs_mr_inference.R:465-475`: both
bias terms average over `ok_H`, and the IJ runs on `ok_H`), so what the
assertion would actually detect is a *reintroduced* denominator mismatch --- the
F13 defect returning. That is exactly the failure mode with no other detector
here.

Closing it needs one of: a `mean_r` (or `r`/`ok`) element on the returned
object, or an internal accessor the fixture may call. Both are package changes,
which this task excludes. **Recorded as an open gap, not worked around.**

```{r gate3}
#| echo: false
gate("L3")
```

# Findings {#sec-findings}

## The spec's Level 1 closed forms omit the leverage factor

`SPEC_closedform_fixture.md` states
$\mathrm{db}_{g,i} = e_i/n_1$ (treated), $-e_i/n_0$ (control), and
$\sigma^2_{D,g} = \widehat s_1^2/n_1 + \widehat s_0^2/n_0$, "the Neyman
variance". **The package computes neither.**

`.dfbeta_glm()` (`R/consistency_resample.R:168-176`) returns the one-step
influence $(X'X)^{-1}x_i e_i/(1 - h_i)$. For a two-group model $h_i = 1/n_a$,
so the leverage factor is $n_a/(n_a - 1)$ and

$$\mathrm{db}_i = \frac{e_i}{n_a - 1},
\qquad
\sigma^2_{D,g} = \frac{\mathrm{SS}_1}{(n_1-1)^2} + \frac{\mathrm{SS}_0}{(n_0-1)^2}.$$

Measured at $n_a = 100$: package `0.2696379`, spec's Neyman form `0.2669415` ---
a 1.0% gap, against a stated tolerance of `1e-10`. The spec's forms would have
failed **on contact**, and the failure would have looked like a broken influence
path rather than a mis-stated reference.

This is a spec correction, not a package defect. The one-step form is the
documented and verified construction: `411a448` validated `.dfbeta_glm()`
against `stats::dfbeta()` to ~1e-16 on `lm()` unweighted, `lm()` with prior
weights, and three glm families. The discrepancy is $O(n_a^{-1})$ relative ---
the same order as the perturbed-information remainder the spec itself notes ---
so the Neyman form is the correct *asymptotic* statement and the wrong *exact*
one. This document and the testthat file assert the exact form, and the
testthat file additionally asserts that the Neyman form does **not** match, so a
future change toward it is caught rather than absorbed.

## The verification README's V5/V6 paragraph was stale

It described the F13 denominator question as "undecided, and these scripts do
not decide it", three paragraphs above its own V8/V9 section stating that the
package implements the conditional convention. Corrected when this fixture was
built; the reference scripts themselves were not touched.

# Summary

```{r summary}
#| echo: false
SUM <- data.frame(
  level = c("Level 1 --- per-candidate objects",
            "Level 2 --- the correction",
            "Level 3 --- the interval"),
  status = ifelse(LEVEL_OK, "PASS", "**FAIL**"),
  note = c("beta_hat, sigma_D, B, Sigma; plus se_wald at the top level",
           "S1 + rate + tie, S2 conditional vs unconditional, the inverted fixed_bias pair, S4",
           "S3 ratio -> 4, row-centering invariance; rbar gap recorded"))
tbl(SUM, "Fixture result", format(Sys.time(), "%Y-%m-%d %H:%M"))
```

```{r sessioninfo}
#| echo: false
cat("forestsearch", as.character(utils::packageVersion("forestsearch")), "|",
    R.version.string, "|", "B =", DRAWS, "| multiplier = gaussian\n")
```
