---
title: "FS Parameter Bundles: Comparison and Differences"
subtitle: "Side-by-side view of fs1, fs2, ... regimes"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
---

This document loads every `fs*_params.R` file in this directory and
renders two views of the parameter values:

- **Section 2 — Differences only**: the compact view. One row per
  parameter that varies across the bundles. Best for "what's
  different between fs1 and fs2?"
- **Section 3 — All parameters**: the comprehensive view. Every
  parameter, with cells that differ across bundles highlighted.
  Best for an audit-trail / one-page reference.

The document auto-discovers `fs*_params.R` files. Add a new bundle
by dropping `fs2_params.R` into this directory and re-rendering.

```{r setup}
#| code-fold: true
#| code-summary: "Setup, discovery, and parameter loading (click to expand)"

library(data.table)
library(dplyr)
library(gt)

# Base R adds %||% in 4.4; define a fallback for older R.
if (!exists("%||%", mode = "function")) {
  `%||%` <- function(x, y) if (is.null(x)) y else x
}

# -- Auto-discover fs*_params.R files ---------------------------------------
# Files are sourced from the current working directory of this qmd
# (i.e., quarto/simulations/<study>/_fsparams/). Each must define a
# function get_fsN_params(sg_focus, ...).
fs_files <- list.files(pattern = "^fs[0-9]+_params\\.R$",
                       full.names = FALSE)

if (length(fs_files) == 0L) {
  stop("No fs*_params.R files found in this directory.", call. = FALSE)
}

# Extract bundle names (e.g., "fs1", "fs2") from filenames
fs_names <- sub("_params\\.R$", "", fs_files)
names(fs_files) <- fs_names

cat(sprintf("Discovered %d FS bundle(s): %s\n",
            length(fs_files), paste(fs_names, collapse = ", ")))
```

```{r load-bundles}
#| code-fold: true
#| code-summary: "Source each file, call its function, capture parameters (click to expand)"

# We call each bundle's function with TWO sg_focus values to surface any
# sg_focus-dependent parameters (e.g., stop_threshold returns NULL for
# hrMaxSG/hrMinSG and 0.90 otherwise). Choosing one of each:
#   - "hr"      -> the early-stopping branch
#   - "hrMaxSG" -> the neighborhood-criterion branch
sg_focus_probes <- c(early_stopping = "hr", neighborhood = "hrMaxSG")

# Source each file in a fresh environment so functions don't collide
load_bundle <- function(file, fs_name) {
  env <- new.env()
  source(file, local = env)
  fn_name <- paste0("get_", fs_name, "_params")
  if (!exists(fn_name, envir = env, inherits = FALSE)) {
    stop(sprintf("File %s does not define function %s().",
                 file, fn_name), call. = FALSE)
  }
  fn <- get(fn_name, envir = env)
  # Call with each sg_focus probe; return as named list of parameter lists
  lapply(sg_focus_probes, function(sf) fn(sg_focus = sf))
}

bundles <- mapply(load_bundle,
                  file    = fs_files,
                  fs_name = fs_names,
                  SIMPLIFY = FALSE)
# bundles[[fs_name]][[probe_name]] is a parameter list

# Extract regime descriptions from each file's header comment
extract_regime_description <- function(file) {
  lines <- readLines(file, warn = FALSE)
  # Find lines starting with "# Regime:" and capture until next blank
  # comment line (a line that is just "#" with optional spaces)
  start <- grep("^#\\s+Regime:", lines)
  if (length(start) == 0L) return(NA_character_)
  start <- start[1]
  end <- start
  for (i in (start + 1):length(lines)) {
    if (!grepl("^#", lines[i])) break
    if (grepl("^#\\s*$", lines[i])) break
    end <- i
  }
  desc <- paste(sub("^#\\s?", "", lines[start:end]), collapse = "\n")
  desc
}

regime_descriptions <- vapply(fs_files, extract_regime_description,
                              character(1))
```

# 1. Regime Descriptions {#sec-regimes}

What each FS bundle is designed for. Pulled from the header comment
block of each `fs*_params.R` file.

```{r regime-descriptions}
#| results: asis
for (i in seq_along(fs_names)) {
  cat(sprintf("\n## %s\n\n", fs_names[i]))
  if (is.na(regime_descriptions[i])) {
    cat("_No `Regime:` line found in header comment._\n")
  } else {
    cat("```\n", regime_descriptions[i], "\n```\n", sep = "")
  }
}
```

```{r build-comparison-data}
#| code-fold: true
#| code-summary: "Build the wide comparison table (click to expand)"

# Collect all parameter names across all bundles. We use the
# "early_stopping" probe as the canonical view (so stop_threshold
# appears as 0.90 rather than NULL).
canonical_probe <- "early_stopping"

format_value <- function(x) {
  if (is.null(x)) return("NULL")
  if (is.list(x)) {
    # parallel_args / conf.cont_jcuts are lists — render compactly as
    # "key=val, key=val".  The "list(...)" wrapper is dropped for
    # display readability; the values still parse unambiguously.
    parts <- vapply(seq_along(x),
                    function(i) sprintf("%s=%s",
                                        names(x)[i],
                                        format_value(x[[i]])),
                    character(1))
    return(paste(parts, collapse = ", "))
  }
  if (length(x) > 1L) return(paste0("c(",
                                    paste(vapply(x, format_value,
                                                 character(1)),
                                          collapse = ", "),
                                    ")"))
  if (is.character(x)) return(sprintf('"%s"', x))
  if (is.logical(x))   return(if (x) "TRUE" else "FALSE")
  format(x)
}

# Get the union of parameter names, in the order they appear in the
# first bundle (with any extra params from later bundles appended)
param_order <- character(0)
for (fs_name in fs_names) {
  pl <- bundles[[fs_name]][[canonical_probe]]
  param_order <- union(param_order, names(pl))
}

# Build a data.table: one row per parameter, one column per bundle.
# Use data.table::set() rather than `[[<-` to avoid the shallow-copy
# warning that base-R column assignment on a data.table triggers.
wide_dt <- data.table::data.table(parameter = param_order)
for (fs_name in fs_names) {
  pl <- bundles[[fs_name]][[canonical_probe]]
  data.table::set(
    wide_dt,
    j     = fs_name,
    value = vapply(param_order,
                   function(p) format_value(pl[[p]]),
                   character(1))
  )
}

# Add a "differs" flag: TRUE if any pair of bundle values differs
if (length(fs_names) >= 2L) {
  wide_dt[, differs := apply(.SD, 1, function(row) length(unique(row)) > 1L),
          .SDcols = fs_names]
} else {
  wide_dt[, differs := FALSE]
}

# Also add a note column for parameters with sg_focus-dependent values
# (i.e., those that differ between the two probes within a bundle)
note_for <- function(p) {
  notes <- character(0)
  for (fs_name in fs_names) {
    val_a <- bundles[[fs_name]][["early_stopping"]][[p]]
    val_b <- bundles[[fs_name]][["neighborhood"]][[p]]
    if (!identical(val_a, val_b)) {
      notes <- c(notes, sprintf(
        "%s: %s for hr/maxSG/minSG; %s for hrMaxSG/hrMinSG",
        fs_name, format_value(val_a), format_value(val_b)
      ))
    }
  }
  if (length(notes) == 0L) "" else paste(notes, collapse = "; ")
}
wide_dt[, sg_focus_dependent := vapply(parameter, note_for, character(1))]

# -- Helpers for cell-level "odd-one-out" highlighting in the diffs table --
#
# For each row, identify which bundle columns hold a value different from
# the row's *mode* (most common value).  These are the cells the reader's
# eye most needs to find; they get a soft fill in the rendered table.
#
# Tie-breaking: if multiple values tie for most-common, the first-encountered
# is treated as the mode (so highlighting biases toward the lower-indexed
# bundles being "normal").  In practice this matters only for 50/50 splits.
row_mode <- function(values) {
  tab <- table(values)
  names(tab)[which.max(tab)]
}

# Returns, for one row of the diffs table, a logical vector over fs_names:
# TRUE = cell value differs from the row's mode.
row_odd_mask <- function(row_values) {
  mode_val <- row_mode(row_values)
  row_values != mode_val
}
```

# 2. Differences Only {#sec-diffs}

The compact view: only parameters whose value differs across at
least one pair of bundles.

```{r diffs-only}
diffs_dt <- wide_dt[differs == TRUE]

if (nrow(diffs_dt) == 0L) {
  cat("**All bundles share identical parameter values.**\n\n",
      "(This is expected when only one bundle is defined, or when ",
      "two bundles happen to be identical.)")
} else {
  # -- Per-cell highlighting: identify odd-one-out cells row-by-row ---------
  # For each parameter row, mark cells whose value differs from the row's
  # mode.  These are the cells worth visually flagging; same-as-mode cells
  # stay un-styled so they recede.
  #
  # odd_cells_long is a list with one entry per row of diffs_dt, each
  # entry a character vector of the *column names* (bundle ids) flagged
  # odd for that row.
  odd_cells_long <- lapply(seq_len(nrow(diffs_dt)), function(i) {
    row_vals <- unlist(diffs_dt[i, fs_names, with = FALSE], use.names = FALSE)
    mask     <- row_odd_mask(row_vals)
    fs_names[mask]
  })

  # -- Detect cell value types for alignment --------------------------------
  # Right-align numeric-looking columns, center logicals, left-align the
  # rest.  Detection is per-column on the formatted character values, so
  # we don't have to track the original R types.
  is_numeric_str  <- function(v) all(grepl("^-?[0-9]+(\\.[0-9]+)?$", v))
  is_logical_str  <- function(v) all(v %in% c("TRUE", "FALSE"))

  num_cols  <- fs_names[vapply(fs_names,
                               function(c) is_numeric_str(diffs_dt[[c]]),
                               logical(1))]
  log_cols  <- fs_names[vapply(fs_names,
                               function(c) is_logical_str(diffs_dt[[c]]),
                               logical(1))]
  text_cols <- setdiff(fs_names, c(num_cols, log_cols))

  tbl <- diffs_dt[, .SD, .SDcols = c("parameter", fs_names)] |>
    gt::gt() |>
    gt::tab_header(
      title    = "Parameters that differ across FS bundles",
      subtitle = sprintf(
        "%d of %d parameters vary; highlighted cells differ from the row's most-common value",
        nrow(diffs_dt), nrow(wide_dt)
      )
    ) |>
    gt::cols_label(parameter = "Parameter") |>
    # Per-type alignment
    gt::cols_align(align = "left",   columns = parameter) |>
    gt::cols_align(align = "right",  columns = dplyr::all_of(num_cols))  |>
    gt::cols_align(align = "center", columns = dplyr::all_of(log_cols))  |>
    gt::cols_align(align = "left",   columns = dplyr::all_of(text_cols)) |>
    # Parameter column bold
    gt::tab_style(
      style    = gt::cell_text(weight = "bold"),
      locations = gt::cells_body(columns = parameter)
    ) |>
    # Tighter padding & monospace data cells so the wide rows breathe
    gt::tab_options(
      table.font.size            = gt::px(13),
      data_row.padding           = gt::px(4),
      heading.title.font.size    = gt::px(15),
      heading.subtitle.font.size = gt::px(12),
      column_labels.font.weight  = "bold"
    ) |>
    gt::tab_style(
      style     = gt::cell_text(font = gt::google_font("JetBrains Mono")),
      locations = gt::cells_body(columns = dplyr::all_of(fs_names))
    )

  # -- Apply the odd-one-out highlight, one row at a time -------------------
  # cells_body() can take rows + columns simultaneously; the loop is over
  # rows because each row's "odd" column set is different.
  for (i in seq_along(odd_cells_long)) {
    odd_cols <- odd_cells_long[[i]]
    if (length(odd_cols) == 0L) next
    tbl <- tbl |>
      gt::tab_style(
        style     = list(
          gt::cell_fill(color = "#fff3cd"),     # soft amber
          gt::cell_text(weight = "bold")
        ),
        locations = gt::cells_body(
          columns = dplyr::all_of(odd_cols),
          rows    = i
        )
      )
  }

  tbl
}
```

# 3. All Parameters {#sec-all}

The comprehensive view: every parameter, grouped into thematic
categories.  Rows where values differ across bundles are filled in
soft amber and the individual cells that deviate from the row's most
common value are highlighted with bolder amber so the eye can find
"what changed" within a long row.  Rows that are uniform across all
bundles are left un-styled and recede visually.

Parameters that vary by `sg_focus` *within* a bundle (currently
`stop_threshold` and `sg_focus` itself) are flagged in a footnote
beneath the table rather than in a per-row column — the rule is the
same for every bundle, so repeating it 8 times across the row added
no information.

```{r all-params}
# -- Category assignment for row grouping ----------------------------------
# Each of the 29 parameters falls into one of seven thematic groups.
# Unknown parameters (added in future bundles) fall into "Other" so the
# table still renders rather than failing.
param_category <- function(p) {
  if (p %in% c("outcome.name", "event.name", "treat.name", "id.name",
               "outcome_type", "effect_measure"))
    return("Outcome / data binding")
  if (p %in% c("use_lasso", "use_grf", "return_selected_cuts_only",
               "max_subgroups_search", "use_twostage"))
    return("Method switches")
  if (p %in% c("hr.threshold", "hr.consistency", "pconsistency.threshold",
               "stop_threshold", "sg_focus"))
    return("FS thresholds & criteria")
  if (p %in% c("conf_force", "conf.cont_jcuts"))
    return("Confounder configuration")
  if (p %in% c("fs.splits", "n.min", "d0.min", "d1.min", "maxk"))
    return("Splits & subgroup-size constraints")
  if (p %in% c("dmin.grf"))
    return("GRF")
  if (p %in% c("is.RCT", "seedit", "details", "quiet", "parallel_args"))
    return("Run controls")
  "Other"
}

# Preserve the natural category order (declared in param_category logic).
category_order <- c(
  "Outcome / data binding",
  "Method switches",
  "FS thresholds & criteria",
  "Confounder configuration",
  "Splits & subgroup-size constraints",
  "GRF",
  "Run controls",
  "Other"
)

# -- Build display data ----------------------------------------------------
# Drop sg_focus_dependent (now a footnote) and differs (signal carried by
# row fill).  Keep parameter + 8 bundle columns + a hidden category for
# row grouping.
display_dt <- wide_dt[, .SD, .SDcols = c("parameter", fs_names, "differs")]
display_dt[, category := vapply(parameter, param_category, character(1))]
display_dt[, category := factor(category, levels = category_order)]
data.table::setorder(display_dt, category)

# -- Per-cell odd-one-out for rows that differ -----------------------------
# Same logic as the diffs table.  Cells whose value matches the row mode
# get only the row-level amber fill; cells that deviate get the brighter
# amber + bold treatment.
differing_rows <- which(display_dt$differs)
odd_cells_per_row <- lapply(differing_rows, function(i) {
  row_vals <- unlist(display_dt[i, fs_names, with = FALSE], use.names = FALSE)
  mask     <- row_odd_mask(row_vals)
  fs_names[mask]
})
names(odd_cells_per_row) <- as.character(differing_rows)

# -- Per-column alignment by type ------------------------------------------
is_numeric_str  <- function(v) all(grepl("^-?[0-9]+(\\.[0-9]+)?$", v))
is_logical_str  <- function(v) all(v %in% c("TRUE", "FALSE"))
num_cols  <- fs_names[vapply(fs_names,
                             function(c) is_numeric_str(display_dt[[c]]),
                             logical(1))]
log_cols  <- fs_names[vapply(fs_names,
                             function(c) is_logical_str(display_dt[[c]]),
                             logical(1))]
text_cols <- setdiff(fs_names, c(num_cols, log_cols))

# -- Build the gt table ----------------------------------------------------
tbl <- display_dt |>
  gt::gt(
    rowname_col  = "parameter",
    groupname_col = "category"
  ) |>
  gt::cols_hide(columns = differs) |>
  gt::tab_header(
    title    = "All FS parameters",
    subtitle = sprintf(
      "%d parameters across %d bundle(s); row fill = differs across bundles, cell highlight = differs from row's most-common value",
      nrow(display_dt), length(fs_names)
    )
  ) |>
  gt::cols_align(align = "right",  columns = dplyr::all_of(num_cols))  |>
  gt::cols_align(align = "center", columns = dplyr::all_of(log_cols))  |>
  gt::cols_align(align = "left",   columns = dplyr::all_of(text_cols)) |>
  gt::tab_options(
    table.font.size            = gt::px(13),
    data_row.padding           = gt::px(3),
    row_group.font.weight      = "bold",
    row_group.background.color = "#f0f0f0",
    heading.title.font.size    = gt::px(15),
    heading.subtitle.font.size = gt::px(12),
    column_labels.font.weight  = "bold",
    stub.font.weight           = "bold"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(font = gt::google_font("JetBrains Mono")),
    locations = gt::cells_body(columns = dplyr::all_of(fs_names))
  ) |>
  # Row-level soft amber on every differing row (recedes; signals "this
  # row has variability without yet saying where").
  gt::tab_style(
    style     = gt::cell_fill(color = "#fff8e1"),
    locations = gt::cells_body(rows = differing_rows)
  ) |>
  # Footnote tying the "sg_focus dependent" rule to its two affected
  # parameters.  This replaces the column we removed.
  gt::tab_source_note(
    source_note = gt::md(
      "**`sg_focus`-dependent values** (within each bundle): `stop_threshold` is `0.95` when `sg_focus` is one of `\"hr\"`, `\"maxSG\"`, `\"minSG\"` and `NULL` when `sg_focus` is `\"hrMaxSG\"` or `\"hrMinSG\"` (early stopping is incompatible with neighborhood-based selection).  The `sg_focus` parameter itself takes the value passed in.  All other parameters are invariant in `sg_focus`."
    )
  )

# -- Apply the brighter per-cell highlight on the odd-one-out cells --------
# Loop over differing rows and style each cell individually.  This goes
# *after* the row-level fill so it visually wins (gt resolves later
# tab_style() calls on top).
for (i in seq_along(differing_rows)) {
  row_idx  <- differing_rows[i]
  odd_cols <- odd_cells_per_row[[i]]
  if (length(odd_cols) == 0L) next
  tbl <- tbl |>
    gt::tab_style(
      style     = list(
        gt::cell_fill(color = "#ffd54f"),     # brighter amber
        gt::cell_text(weight = "bold")
      ),
      locations = gt::cells_body(
        columns = dplyr::all_of(odd_cols),
        rows    = row_idx
      )
    )
}

tbl
```

# 4. Notes {#sec-notes}

- **`sg_focus`-dependent parameters** vary by `sg_focus` within a
  single bundle. The most important example is `stop_threshold`,
  which the package requires to be `NULL` when `sg_focus` is
  `"hrMaxSG"` or `"hrMinSG"` (neighborhood-based selection is
  incompatible with early stopping). The `fs*_params.R` function
  applies this rule internally, and Section 3 documents it as a
  footnote under the All Parameters table.
- **Probe values used**: this article calls each bundle's function
  with `sg_focus = "hr"` (early-stopping branch, used for the
  display values in Sections 2-3) and `sg_focus = "hrMaxSG"`
  (neighborhood branch, used to detect `sg_focus`-dependent
  parameters). Other `sg_focus` values produce identical results
  to one of these two branches.
- **To add a new bundle**: drop `fsN_params.R` into this directory
  with a function `get_fsN_params()` and re-render. The article
  picks it up automatically.
