---
title: "Exploratory Subgroup Identification with ForestSearch"
subtitle: "German Breast Cancer Study Group (GBSG) Analysis"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
execute:
  warning: true
  message: true
---

> **Method-general analysis.** The selection engine for the main analysis is set
> once via `fs_subgroup_method` in the setup chunk — `"consistency"` (default),
> `"dina"`, or `"grf"` (with `fs_grf_selection` = `"tree"`/`"frontier"`). The
> headline fit and every downstream step — full bootstrap, MR,
> cross-validation, leave-one-out, and the Kaplan–Meier band — inherit that engine
> from the single fitted object, so the whole analysis re-renders under a
> different method by changing one knob. The comparative DINA-screening,
> standalone-DINA, and GRF sections that follow are fixed illustrations of the
> individual engines and are unaffected by `fs_subgroup_method`.

::: {.callout-note title="Two resampling procedures: FB and MR"}
Both correct the selected subgroup's effect estimate for selection bias, and
both appear throughout this document.

* **FB --- full bootstrap.** `forestsearch_bootstrap_dofuture()` re-runs the
  entire identification on each of `NB` nonparametric resamples, regenerating
  the candidate family every time, and applies the Leon (2024) Eq. 7
  correction. Cost scales with `NB`.
* **MR --- multiplier resampling.** `fs_mr_inference()`, reached via
  `forestsearch(mr_inference = TRUE)`, is a refit-free multiplier bootstrap
  computed inside the fit. It holds the fitted candidate family fixed and
  returns an infinitesimal-jackknife (IJ) interval. Cost is one vectorised
  `crossprod`, effectively independent of `NB`.

The two therefore target different estimands whenever the family regenerates:
FB marginalises over family regeneration, MR conditions on the realised family.
For the consistency engine with a fixed family they coincide to leading order,
which is what the FB-vs-MR panels below check.

**Where MR runs.** MR is computed once, inside the analysis fit itself. It does
**not** run inside bootstrap replicates or cross-validation folds. The argument
`mr_in_replicates` defaults to `FALSE` in `forestsearch_bootstrap_dofuture()`,
`forestsearch_Kfold()` and `forestsearch_tenfold()`, and each strips
`mr_inference` from the replicate call unless it is set. This document pins it
to `FALSE` at every one of those call sites -- the package default, so the pin
is a no-op that only makes the setting visible -- and never sets it to `TRUE`.
Every MR figure below therefore comes from the single fit for that engine.

Setting `mr_in_replicates = TRUE` restores the earlier behaviour, in which
`mr_inference = TRUE` was inherited from the parent call into every replicate.
That is worth doing only if a per-replicate MR object is wanted for its own
sake --- nothing downstream consumes one --- and it costs roughly 2.3x on the
main bootstrap at `NB` = 1000. The change is otherwise inert: with the outer
seed fixed, `TRUE` and `FALSE` give bit-identical resample indices, per-replicate
selected subgroups, and bias-corrected estimates.
:::

<!--
  analysis_gbsg_survival_multimethod.qmd

  The multimethod GBSG analysis: one dataset, one outcome type (survival),
  three identifier engines (FS / DINA / GRF) compared side by side, each with
  Naive, Full Bootstrap (FB) and Multiplier Resampling (MR) corrections.

  Relative to analysis_gbsg_survival_maxeff_mrconfirm.qmd, this document ADDS:
  the DINA and GRF selection tracks, the DINA-screened FS variant, and the
  CV / LOO validation blocks.  It DROPS: the Guo & He (2021) Algorithm 3
  comparator (its argmax primitive is the relevant correction only on the
  maxeff path) and the independent MR confirmation fit.

  VENUE.  The four forestsearch fits and their bootstraps are compute-heavy and
  require the forestsearch package INSTALLED (devtools::install(), not
  load_all() -- doFuture workers spawn separate R processes).  Run on the
  analysis machine; see the parallelism notes in the setup chunk.

  PARAMETER PROVENANCE.  Every fit renders a table reconciling
  fit$args_call_all against formals(forestsearch), so the values actually used
  are visible in the rendered document without consulting ?forestsearch.
-->

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# Initialize timing
timings <- list()
t_vignette_start <- proc.time()
```

## Parallel Processing Configuration

ForestSearch supports parallel processing for computationally intensive
operations (bootstrap, cross-validation).

```{r parallel-setup}
# PARALLELISM NOTES
#   * FS search -- forestsearch() parallelises the consistency search via its
#     own parallel_args; the plan is set per call and restored on exit.
#   * FB  -- forestsearch_bootstrap_dofuture() uses %dofuture% with
#     .options.future = list(seed = TRUE), so each replicate holds its own
#     L'Ecuyer-CMRG substream.  It inherits the plan set here.
#   * MR  -- vectorised matrix algebra (a single crossprod over all draws), NOT
#     a parallel loop.  It needs no plan, only a threaded BLAS to benefit.  Do
#     not wrap it in a future.
#   * CV / LOO -- forestsearch_Kfold() and forestsearch_tenfold() pin each
#     fold's inner refit to plan = "sequential", workers = 1, so the fold loop
#     workers are the only ones in play.  No nested parallelism to guard.
# Every parallel block below is followed by plan("sequential").

# Worker count for every parallel block below: 80% of PHYSICAL cores (at least
# one).  This is a render-time setting, not the package default -- forestsearch's
# own .default_parallel_workers() is what applies the CRAN 2-core cap under
# R CMD check, and it is not in play here.
n_cores_total <- parallel::detectCores()
n_cores <- max(1L, floor(0.80 * parallel::detectCores(logical = FALSE)))
cat("Using", n_cores, "of", n_cores_total, "total cores for parallel processing")

# Number of bootstrap iterations
# Use 500-2000 for production; reduced here for vignette
NB <- 1000

# ---------------------------------------------------------------------------
# Validation toggles (the slowest parts of the document).  Independent:
# run_cv  controls K-Fold cross-validation; run_loo controls leave-one-out
# (N-fold / OOB).  Either can be turned off without affecting the other.
# ---------------------------------------------------------------------------
#   run_cv  : TRUE runs repeated K-fold CV (kfold-cv plus the diagnostic
#             tabulations and the identification plot).  FALSE skips all of
#             them; the cv-skipped-note chunk states so in the rendered output
#             and timings$kfold stays NA.
#   run_loo : TRUE runs leave-one-out / N-fold OOB.  FALSE skips it, the
#             loo-skipped-note chunk says so, and timings$oob stays NA.
run_cv  <- FALSE    # K-Fold cross-validation (kfold-cv + diagnostics)
run_loo <- FALSE    # Leave-one-out / N-fold OOB (oob-cv)
# K-fold CV with multiple iterations
# Use Ksims >= 50 for production
Ksims    <- 50
Kfolds_cv <- 10     # folds per iteration; feeds BOTH the call and the narrative,
                    # so the "Ksims x Kfolds" count in the prose cannot drift

# Pre-seed CV/LOO timings so the Computational Timing table is well-formed
# whether or not these run (skipped analyses report NA).
timings$kfold <- NA_real_
timings$oob   <- NA_real_

# ---------------------------------------------------------------------------
# Per-analysis identification parameters (tune each analysis independently).
#   sg_focus      : which candidate subgroup the selection step chooses
#   hr_threshold  : harm/benefit floor on the HR scale
#   hr_consistency: minimum consistency rate (consistency search only)
# ---------------------------------------------------------------------------
# Main ForestSearch (consistency search)
fs_sg_focus              <- "effMaxSG"
fs_hr_threshold          <- 1.0
fs_hr_consistency        <- 1.0
# Uncapped.  The cap has no counterpart in the method: the candidate family is
# every conjunction of at most two conditions, less empty intersections and
# size/event failures, and selection ranges over that family.  Truncation
# applies a PREVIEW ordering, so the candidate that is optimal under the actual
# selection criterion can be discarded before it is ever evaluated.  Measured
# cost of removing it here: 73 extra candidates at ~4.8 ms each, under a second.
# It previously bound (pool 123 truncated to 50), so the FS label may differ from earlier runs --
# that is a result change, not a defect.
fs_max_subgroups_search  <- Inf

fs_conf_force <- c("er <= 0", "pgr <= 0")

fs_consistency_method    <- "resample"
# MR: a fast multiplier-bootstrap approximation of the
# bootstrap bias-correction, computed inside the fit and compared to the full
# bootstrap below.  It runs for whichever selection engine fs_subgroup_method
# names, de-biasing against that engine's own candidate family (it re-derives
# the per-candidate Cox dfbeta itself).
# run_mr : TRUE passes mr_inference = TRUE to every forestsearch() call, so the
#          MR de-biased estimate is computed inside each fit.  FALSE passes
#          mr_inference = FALSE and no MR compute runs at all; the MR rows and
#          tables then report that it was skipped rather than showing blanks.
run_mr <- TRUE     # add mr_inference = TRUE to the forestsearch call
mr_draws      <- 5000L    # multiplier-bootstrap draws for MR

# ---------------------------------------------------------------------------
# Output directory for the Section 6.2 reproducibility payload written at the
# END of this document (manuscript Table + results overview).
#   NULL (default) -> alongside this .qmd, resolved at render time
#   or an explicit directory, e.g. "~/results" or "/abs/path/out"
#
# Pointed at ../rerun so this render writes NOWHERE NEAR the manuscript-era
# reference bundles that sit alongside this document.  This directory is
# read-only for the identifier-alignment re-analysis: gbsg_table_payload.rds
# is the comparison target, and rendering into the default NULL location is
# how a reference gets clobbered.  (The filename this document writes,
# gbsg_survival_multimethod_payload.rds, does not itself collide -- but its
# sibling ACTG175 document's does, so both are redirected for the same reason
# and by the same rule.)
results_dir <- "../rerun"

# ---------------------------------------------------------------------------
# Selection engine for the MAIN analysis.  The headline fit, its bootstrap,
# the MR, CV/LOO, and the KM plot all flow from this single fit,
# so flipping subgroup_method re-runs the entire main analysis under that
# engine (each downstream step inherits the method from the fitted object):
#   "consistency" - the splitting-consistency search (default);
#   "dina"        - DINA performs selection (consistency search bypassed);
#   "grf"         - GRF performs selection (policy tree or DR-frontier).
# The MR de-biases against whichever engine's OWN candidate family,
# with re-selection matching that engine's selection rule.
# ---------------------------------------------------------------------------
fs_subgroup_method <- "consistency"  # "consistency" | "dina" | "grf"
fs_grf_selection   <- "frontier"      # only if fs_subgroup_method == "grf": "tree" | "frontier"
fs_grf_select_statistic <- "effect"       # only if subgroup_method=="grf" & grf_selection=="frontier": "dr" | "effect".
                                      #   "dr"     = native DR-score frontier winner;
                                      #   "effect" = re-rank the DR frontier on the inferential Cox HR (the
                                      #              statistic the MR de-biases).  No-op for the policy tree.
fs_grf_depth       <- 2L              # GRF depth (screening + GRF selection)
fs_dmin_grf        <- 4               # GRF DR-score screening floor
fs_dina_args       <- list()          # extra args forwarded to DINA selection
fs_dina_select_statistic <- "effect"    # only if subgroup_method=="dina": "dina" | "effect".
                                      #   "dina"   = native DINA tau-hat winner;
                                      #   "effect" = re-rank the DINA family on the inferential Cox HR (the
                                      #              statistic the MR de-biases).
# MR (main analysis): MR's de-biasing family for "dina"/"grf" is
# restricted to candidates competitive on the method's NATIVE statistic (DINA
# tau-hat / GRF DR-score).  NULL = use the recommended default, which matches the
# band to effect_neighborhood automatically (the band the engine selects within);
# give a number in (0, 1) to set the band explicitly, or >= 1 to de-bias against
# the full family.  Ignored for "consistency" (its maxcons rule already
# matches the search).
fs_family_native_nbhd <- NULL

# (A) ForestSearch + DINA screening (consistency search selects; DINA only
#     contributes candidate cuts)
dina_screen_sg_focus             <- "effMaxSG"
dina_screen_hr_threshold         <- 0.9
dina_screen_hr_consistency       <- 0.8
# Uncapped, for the same reason as fs_max_subgroups_search above: no
# counterpart in the method, truncation under a preview ordering, and a
# measured cost of 12 extra candidates at ~4.8 ms each.  It previously bound
# (pool 42 truncated to 30).
dina_screen_max_subgroups_search <- Inf
# Screening cut set: FALSE => full DINA frontier; TRUE => only the single
# dina_subgroup()-selected cut (uses dina_screen_sg_focus / selection_rule).
dina_screen_selected_only        <- TRUE

# (C) ForestSearch subgroup_method = "dina" (DINA performs selection; the
#     consistency search is bypassed, so hr_consistency does not apply here).
#     The standalone DINA analysis (B) reuses these so (B) and (C) match.
fs_dina_sg_focus     <- "effMaxSG"
fs_dina_hr_threshold <- 0.90
# MR: the DINA de-biasing family is restricted to candidates competitive
# on DINA's native tau-hat.  NULL = recommended default (matches effect_neighborhood
# automatically); set a number in (0, 1) for an explicit band, or >= 1 for the full
# family.  Note: this only tunes candidate competition -- MR holds the forest
# fixed, so the full bootstrap stays the reference for DINA regardless.
fs_dina_family_native_nbhd <- NULL

# (D) ForestSearch subgroup_method = "grf" (GRF performs selection via a causal
#     forest policy tree or DR-score frontier; the consistency search, LASSO and
#     DINA are bypassed).  Dedicated demo: always runs GRF regardless of
#     fs_subgroup_method, mirroring the standalone DINA selection (C).
fs_grf_demo_selection     <- "frontier"  # "frontier" | "tree"
fs_grf_demo_depth         <- 2L          # GRF tree / DR-pair depth
fs_grf_demo_dmin          <- 0.0         # DR-score harm floor (0 = permissive default)
fs_grf_sg_focus           <- "effMaxSG"  # -> GRF frontier/tree selection rule
# MR: the GRF de-biasing family is restricted to candidates competitive
# on GRF's native (RMST-based) DR-score.  NULL = recommended default (matches
# effect_neighborhood automatically); set a number in (0, 1) for an explicit band,
# or >= 1 for the full family.  Note: this only tunes candidate competition -- the
# MR holds the forest fixed, so the full bootstrap stays the reference for GRF
# regardless (a dominant DR-score winner can collapse the family to 1).
fs_grf_family_native_nbhd <- NULL
```



# Introduction

This vignette demonstrates the **ForestSearch** methodology for exploratory
subgroup identification in survival analysis, as described in
León et al. (2024) *Statistics in Medicine*.

## Motivation

In clinical trials, particularly oncology, subgroup analyses are essential for:

- Evaluating treatment effect consistency across patient populations
- Identifying subgroups where treatment may be detrimental (harm)
- Characterizing subgroups with enhanced benefit
- Informing regulatory decisions and clinical practice

While prespecified subgroups provide stronger evidence, important subgroups
based on patient characteristics may not be anticipated. ForestSearch provides
a principled approach to **exploratory** subgroup identification with proper
statistical inference.

## Methodology Overview

ForestSearch identifies subgroups through:

1. **Candidate factor selection**: Using LASSO and/or Generalized Random Forests (GRF)
2. **Exhaustive subgroup search**: Evaluating all combinations up to `maxk` factors
3. **Consistency-based selection**: Applying splitting consistency criteria
4. **Bootstrap bias correction**: Adjusting for selection-induced optimism
5. **Cross-validation**: Assessing algorithm stability

The key innovation is the **splitting consistency criterion**: a subgroup is
considered "consistent with harm" if, when randomly split 50/50 many times,
both halves consistently show hazard ratios ≥ 1.0 (for example if 1.0
represents a meaningful "harm threshold").


# Setup

## Load Required Packages

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(policytree)
library(doFuture)

# Optional packages for enhanced output
library(patchwork)
library(weightedsurv)

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```

## Parameter provenance {#sec-param-helper}

Each `forestsearch()` fit carries `args_call_all`: the argument list the call
actually resolved to, kept in sync by the package's own `.sync_args_call_all()`
and used by `forestsearch()` itself for reconstruction. The helper below
reconciles that list against `formals(forestsearch)` at render time, so the
values reported are the ones the fit used, read from the fit rather than
restated in comments that can drift.

```{r param-provenance-helper}
#| code-fold: false

.fs_fmt_val <- function(v) {
  if (is.null(v)) return("NULL")
  if (is.data.frame(v)) return(sprintf("<data.frame %d x %d>", nrow(v), ncol(v)))
  if (is.function(v)) return("<function>")
  if (is.list(v)) {
    if (length(v) == 0L) return("list()")
    nms <- names(v); if (is.null(nms)) nms <- rep("", length(v))
    inner <- vapply(seq_along(v), function(i) {
      val <- tryCatch(paste(deparse(v[[i]], width.cutoff = 200L), collapse = " "),
                      error = function(e) "<?>")
      if (nzchar(nms[i])) paste0(nms[i], " = ", val) else val
    }, character(1))
    return(paste0("list(", paste(inner, collapse = ", "), ")"))
  }
  if (length(v) > 6L) return(sprintf("<%s[%d]>", class(v)[1], length(v)))
  paste(deparse(v, width.cutoff = 200L), collapse = " ")
}

.fs_fmt_default <- function(x) {
  if (identical(x, quote(expr = ))) return("<required; no default>")
  paste(deparse(x, width.cutoff = 200L), collapse = " ")
}

# match.arg-style values resolve to their first element.  Applied to BOTH sides
# of the comparison below: a value captured unresolved (grf_selection,
# grf_select_statistic) still sits at its default and must not read as differing.
.fs_collapse_matcharg <- function(x) {
  ifelse(!is.na(x) & grepl('^c[(]"', x),
         sub('^c[(]"([^"]*)".*$', '"\\1"', x), x)
}

#' Reconcile a fit's resolved arguments against forestsearch()'s formals
#'
#' @param fit A fitted `forestsearch()` object carrying `args_call_all`.
#' @param fn Function whose formals define the reference set.
#' @return One row per formal: the package default, the value the fit resolved
#'   to, and whether that value sits at the default.
fs_param_provenance <- function(fit, fn = forestsearch) {
  stopifnot(!is.null(fit), !is.null(fit$args_call_all))
  args <- fit$args_call_all
  fm   <- formals(fn)
  nm   <- setdiff(names(fm), "...")

  defaults <- vapply(fm[nm], .fs_fmt_default, character(1), USE.NAMES = FALSE)
  used <- vapply(nm, function(p) {
    if (!p %in% names(args)) return(NA_character_)
    .fs_fmt_val(args[[p]])
  }, character(1), USE.NAMES = FALSE)

  eff      <- .fs_collapse_matcharg(defaults)
  eff_used <- .fs_collapse_matcharg(used)

  out <- data.frame(
    parameter  = nm,
    default    = defaults,
    value_used = ifelse(is.na(used), "(not captured)", used),
    status     = ifelse(is.na(used), "not captured",
                 ifelse(eff_used == eff, "at default", "differs from default")),
    stringsAsFactors = FALSE
  )

  # forestsearch_main.R:1304 overrides these five under sg_focus = "maxeff" and
  # does NOT re-sync args_call_all afterwards (the package notes this itself at
  # ~2763), so what was captured is the value typed into the call, not the value
  # the engine ran with.
  if (identical(args$sg_focus, "maxeff")) {
    .ov <- c("pconsistency.threshold", "stop_threshold", "use_twostage",
             "max_subgroups_search", "minp")
    i <- out$parameter %in% .ov
    out$status[i] <- "captured BEFORE maxeff override -- not what the engine used"
  }

  out
}

#' Render a provenance table for one fit
# Selection-engine knobs: the formals that govern candidate generation,
# admission and selection.  The per-fit tables below show only these, and only
# where the resolved value is not the package default -- that is what a reader
# needs in order to interpret the fit.  The full 85-formal reconciliation for
# every fit is in @sec-param-appendix.
.fs_selection_knobs <- c(
  # engine and focus
  "subgroup_method", "sg_focus", "selection_rule", "effect_neighborhood",
  # admission thresholds
  "hr.threshold", "hr.consistency", "effect.threshold", "consistency.threshold",
  "pconsistency.threshold", "stop_threshold", "consistency_method", "fs.splits",
  # candidate generation
  "use_lasso", "use_grf", "use_dina", "use_twostage", "twostage_args",
  "conf_force", "conf.cont_jcuts", "collapse_cuts", "cont.cutoff",
  "maxk", "max_subgroups_search", "minp",
  # size constraints
  "n.min", "n.min.frac", "d0.min", "d1.min",
  # engine-specific
  "grf_selection", "grf_select_statistic", "grf_depth", "dmin.grf",
  "vi.grf.min", "dina_args",
  # post-selection inference
  "mr_inference", "mr_inference_args"
)

fs_param_table <- function(fit, title, knobs = .fs_selection_knobs) {
  pv   <- fs_param_provenance(fit)
  keep <- pv$parameter %in% knobs & pv$status != "at default"
  tab  <- pv[keep, , drop = FALSE]

  sub <- sprintf(
    "Selection-engine knobs differing from their package default: %d of %d such knobs, out of %d formals in all (forestsearch %s, read at render time). Full reconciliation in the appendix.",
    nrow(tab), sum(pv$parameter %in% knobs), nrow(pv),
    tryCatch(as.character(utils::packageVersion("forestsearch")),
             error = function(e) "version unknown"))
  # Only meaningful when the maxeff override actually fired; under any other
  # sg_focus nothing is overridden and the caveat would be false.
  if (any(grepl("^captured BEFORE", tab$status))) {
    sub <- paste(sub,
      "args_call_all is not re-synced after the sg_focus = \"maxeff\" override block",
      "(forestsearch_main.R:1304; the package notes this itself at ~2763), so the",
      "rows flagged as such report the value typed into the call rather than the",
      "value the engine used -- the render's own maxeff warning lists those.")
  }

  if (!nrow(tab)) {
    return(gt::gt(data.frame(note = "Every selection-engine knob is at its package default.")) |>
             gt::tab_header(title = title, subtitle = sub) |>
             gt::cols_label(note = "") |>
             gt::tab_options(table.font.size = gt::px(11)))
  }

  gt::gt(tab) |>
    gt::tab_header(title = title, subtitle = sub) |>
    gt::cols_label(parameter = "Parameter", default = "Package default",
                   value_used = "Value used", status = "Status") |>
    gt::tab_style(
      style = gt::cell_text(font = gt::google_font("JetBrains Mono"), size = "x-small"),
      locations = gt::cells_body(columns = c(default, value_used))) |>
    gt::tab_options(table.font.size = gt::px(11), data_row.padding = gt::px(2))
}
```


# Data: German Breast Cancer Study Group Trial

## Study Background

The GBSG trial evaluated hormonal treatment (tamoxifen) versus chemotherapy
in node-positive breast cancer patients. Key characteristics:

- **Sample size**: N = 686
- **Outcome**: Recurrence-free survival time
- **Censoring rate**: ~56%
- **Treatment**: Hormonal therapy (tamoxifen) vs. chemotherapy

## Data Preparation

```{r data-setup}
# Load GBSG data (included in forestsearch package)
df.analysis <- gbsg

# Prepare analysis variables
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
  treat <- hormon
})

# Define variable roles
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name <- "status"
id.name <- "id"
treat.name <- "hormon"

# Display data structure
cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```

## Baseline Characteristics

```{r baseline-table}
create_summary_table(
  data = df.analysis,
  treat_var = treat.name,
  table_title = "GBSG Baseline Characteristics by Treatment Arm",
  vars_continuous = c("age", "nodes", "size", "er", "pgr"),
  vars_categorical = c("grade", "meno"),
  font_size = 12
)
```

## Kaplan-Meier Analysis (ITT Population)

```{r km-itt}
#| fig-width: 8
#| fig-height: 5

# Prepare counting process data for KM plot
dfcount <- df_counting(
  df = df.analysis,
  by.risk = 6,
  tte.name = outcome.name,
  event.name = event.name,
  treat.name = treat.name
)

# Plot with confidence intervals and log-rank test
plot_weighted_km(
  dfcount,
  conf.int = TRUE,
  show.logrank = TRUE,
  ymax = 1.05,
  xmed.fraction = 0.775,
  ymed.offset = 0.125
)
```

The ITT Cox hazard ratio estimate is approximately 0.69 (95% CI: 0.54, 0.89),
suggesting an overall benefit for hormonal therapy.


# Preliminary Analysis: Generalized Random Forests

Before running ForestSearch, we can use GRF to explore potential treatment
effect heterogeneity and identify candidate factors.

```{r grf-analysis}
t0 <- proc.time()

grf_est <- grf.subg.harm.survival(
  data = df.analysis,
  confounders.name = confounders.name,
  outcome.name = outcome.name,
  event.name = event.name,
  id.name = id.name,
  treat.name = treat.name,
  maxdepth = 2,
  n.min = 60,
  dmin.grf = 12,
  frac.tau = 0.60,
  details = TRUE,
  return_selected_cuts_only = FALSE
)

timings$grf <- (proc.time() - t0)["elapsed"]
```

```{r grf-trees}
#| fig-width: 10
#| fig-height: 4

# Display policy trees
# leaf1 = recommend control, leaf2 = recommend treatment
oldpar <- par(mfrow = c(1, 2))
plot(grf_est$tree1, leaf.labels = c("Control", "Treat"), main = "Depth 1")
plot(grf_est$tree2, leaf.labels = c("Control", "Treat"), main = "Depth 2")
par(oldpar)
```

GRF identifies estrogen receptor status (ER) as a key factor, with ER ≤ 0
suggesting potential harm from hormonal therapy.


# ForestSearch Analysis




## Running ForestSearch

The selection engine for this main analysis is set by `fs_subgroup_method` in
the setup chunk (`"consistency"`, `"dina"`, or `"grf"` with `fs_grf_selection`).
Everything downstream — the full bootstrap, the MR, the
cross-validation/leave-one-out diagnostics, and the Kaplan–Meier band — runs off
the single fitted object and therefore inherits that engine automatically, so
the entire analysis re-renders under a different method by changing one knob.
The consistency search (the default below) performs an exhaustive search over
candidate subgroup combinations with up to `maxk` factors.

The parameters this fit actually used are not restated here. They are rendered
from `fit$args_call_all` by the provenance table immediately after the call
(@sec-param-helper), so they cannot drift from the code that produced the
results; the full reconciliation across all fits is in @sec-param-appendix.

::: {.callout-note}
## Selection criteria (`sg_focus`)

`sg_focus` chooses among the candidates that survive admission. This analysis
uses `"effMaxSG"`. The `eff*` spellings are preferred over the internal
Cox-flavoured `hr*` names because they carry over to GLM outcomes; both
normalise to the same rule.

**Foci with no band.** These rank all qualifiers directly:

* `"maxcons"` (alias `"eff"`) --- consistency argmax: highest `Pcons`, ties
  broken by effect. Canonical name `"hr"`.
* `"maxeff"` --- effect argmax over *all* candidates, ungated: key
  $(-\mathrm{hr}, K)$. No consistency condition is applied at selection, and the
  effect floor is disabled.
* `"maxeffCons"` --- effect argmax *among consistency-qualifying* candidates.
  The only focus for which first-passer early stopping is sound, because its key
  matches the enumeration order exactly.
* `"maxSG"` / `"minSG"` --- largest / smallest qualifier, ties by consistency.
  These are **not** spellings of `"effMaxSG"`/`"effMinSG"`: they rank by size
  among all qualifiers with no band at all.

**Foci with a band.** `"effMaxSG"` / `"effMinSG"` (canonical
`"hrMaxSG"`/`"hrMinSG"`) do *not* simply take the largest effect. Candidates are
first partitioned into a **candidate inclusion band**, then ranked within it by
sample size. The full sort for `"effMaxSG"` is five keys:

$$(-\texttt{in\_band},\; -N,\; -\texttt{Pcons},\; -\texttt{hr},\; K)$$

so: in-band first, then largest $N$, then higher consistency, then higher
effect, then fewer factors. `"effMinSG"` is identical except $N$ is minimised.
Only `in_band` depends on `selection_rule`; the sort itself is fixed.

**`selection_rule` draws the band.** `.compute_inclusion_band()` forms two
indicators and multiplies them:

| `selection_rule` | `in_nbhd` | `in_pareto` | `in_band` is |
|---|---|---|---|
| `"neighborhood"` (default) | active | all ones | the effect band only |
| `"pareto"` | all ones | active | the frontier only |
| `"both"` | active | active | their intersection |

`in_nbhd` is a *relative* tolerance: with `effect_neighborhood = 0.10` a
candidate is in-band iff its effect is within 10% of the strongest effect among
the survivors. Setting it to `0` collapses the band to a strict max-effect
filter. `in_pareto` is the two-dimensional Pareto-non-dominated set on
(effect, $N$) --- candidate $i$ is dominated when some $j$ is at least as good on
both axes and strictly better on one. For ratio measures stored on the log scale
(GLM OR, IRR) both tests are applied after exponentiating, so the comparison is
on the natural scale; Cox HR is already natural.

**Scope.** `selection_rule` and `effect_neighborhood` are consulted *only* by
`"effMaxSG"`/`"effMinSG"`. For every other focus they are inert --- and this is
enforced, not ignored: passing a `selection_rule` other than `"neighborhood"`
together with a non-band focus is an **error**, not a silent no-op.
:::

```{r forestsearch-main}
#| fig-width: 10
#| fig-height: 7

t0 <- proc.time()

fs <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis,
  outcome.name              = outcome.name,
  event.name                = event.name,
  treat.name                = treat.name,
  id.name                   = id.name,
  confounders.name          = confounders.name,
  potentialOutcome.name     = NULL,           # default NULL
  flag_harm.name            = NULL,           # default NULL
  df.predict                = NULL,           # default NULL
  df.test                   = NULL,           # default NULL
  is.RCT                    = TRUE,           # default TRUE
  seedit                    = 8316951,        # default 8316951
  est.scale                 = "hr",           # default "hr"
  # --- Screening engines (candidate generation) ---------------------------
  use_lasso                 = FALSE,           # default TRUE
  use_grf                   = FALSE,           # default TRUE
  grf_res                   = NULL,           # default NULL
  grf_cuts                  = NULL,           # default NULL
  grf_depth                 = fs_grf_depth,   # set in setup chunk (default 2)
  dmin.grf                  = fs_dmin_grf,     # set in setup chunk (default 0.0)
  frac.tau                  = 0.80,           # default 0.8
  vi.grf.min                = -0.2,           # default -0.2
  return_selected_cuts_only = TRUE,           # default TRUE
  tune_grf                  = FALSE,          # default FALSE
  use_dina                  = FALSE,          # default FALSE
  dina_res                  = NULL,           # default NULL
  dina_cuts                 = NULL,           # default NULL
  dina_args                 = fs_dina_args,   # extra DINA tuning (setup chunk)
  # --- Candidate-cut construction -----------------------------------------
  max_n_confounders         = 1000,           # default 1000
  conf_force                = fs_conf_force,           # default NULL
  defaultcut_names          = NULL,           # default NULL
  cut_type                  = "default",      # default "default"
  exclude_cuts              = NULL,           # default NULL
  replace_med_grf           = FALSE,          # default FALSE
  cont.cutoff               = 4,              # default 4
  conf.cont_medians         = NULL,           # default NULL
  conf.cont_medians_force   = NULL,           # default NULL
  conf.cont_jcuts           = list(er = 10, pgr = 10),           # default NULL
  collapse_cuts = TRUE,
  # --- Selection engine ---------------------------------------------------
  subgroup_method           = fs_subgroup_method,  # set in setup chunk: consistency | dina | grf
  grf_selection             = fs_grf_selection,    # set in setup chunk (used when subgroup_method = "grf")
  grf_select_statistic      = fs_grf_select_statistic,  # set in setup chunk (frontier only)
  dina_select_statistic     = fs_dina_select_statistic,  # set in setup chunk (used when subgroup_method = "dina")
  sg_focus                  = fs_sg_focus,         # set in setup chunk
  selection_rule            = "neighborhood",  # default "neighborhood"
  effect_neighborhood       = 0.10,            # default 0.10
  # --- MR ----------------------------------------------
  mr_inference               = run_mr,           # set in setup chunk
  mr_inference_args          = list(draws = mr_draws,
                                   family_native_neighborhood = fs_family_native_nbhd),
  # --- Subgroup constraints -----------------------------------------------
  n.min                     = 60,                       # default 60
  d0.min                    = 10,                       # default 10
  d1.min                    = 10,                       # default 10
  maxk                      = 2,                        # default 2
  max_subgroups_search      = fs_max_subgroups_search,  # set in setup chunk
  # --- Thresholds (survival / HR scale) -----------------------------------
  hr.threshold              = fs_hr_threshold,    # set in setup chunk
  hr.consistency            = fs_hr_consistency,  # set in setup chunk
  effect.threshold          = NULL,           # default NULL (alias of hr.threshold)
  consistency.threshold     = NULL,           # default NULL (alias of hr.consistency)
  pconsistency.threshold    = 0.90,           # default 0.90
  consistency_method        = fs_consistency_method,
  m1.threshold              = Inf,            # default Inf
  minp                      = 0.025,          # default 0.025
  # --- Consistency search -------------------------------------------------
  fs.splits                 = 1000,           # default 1000
  stop_threshold            = NULL,           # NULL is what this configuration RESOLVES to, not a
                                              #   preference.  This call's sg_focus ("effMaxSG")
                                              #   normalizes to "hrMaxSG", and forestsearch_main.R:1403
                                              #   resets stop_threshold to NULL for that focus: a
                                              #   neighborhood band needs every candidate evaluated,
                                              #   so a truncated scan is unsound.  Any non-NULL value
                                              #   reaches the same NULL but warns on every render;
                                              #   passing NULL skips the reset branch outright (its
                                              #   guard is !is.null(stop_threshold)).  Named here so
                                              #   the resolved setting stays inspectable.
  use_twostage              = TRUE,           # default TRUE
  twostage_args             = list(),         # default list()
  max.minutes               = 3,              # default 3
  # --- Output / diagnostics -----------------------------------------------
  show_candidate_summary    = TRUE,           # default FALSE
  details                   = TRUE,           # default FALSE
  quiet                     = FALSE,          # default FALSE
  by.risk                   = 12,             # default 12
  plot.sg                   = TRUE,           # default FALSE
  plot.grf                  = FALSE,          # default FALSE
  # --- Parallel -----------------------------------------------------------
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$forestsearch <- (proc.time() - t0)["elapsed"]

# Fit-integrity guard.  NOT stopifnot(!is.null(fs$sg.harm)): a NULL subgroup is
# a legitimate outcome this document handles explicitly (see the *-no-subgroup
# chunks).  Assert only that the fit returned and captured its arguments, which
# every downstream provenance table depends on.
stopifnot(!is.null(fs), !is.null(fs$args_call_all))

# Guard for the GRF-dependent diagnostics below, read from the FIT rather than
# from the call, so it tracks whatever the fit actually resolved to.  With
# use_grf = FALSE no GRF cut is ever produced, and the bootstrap GRF-cut
# stability table would render a single "(no GRF cut) -- 100%" row presented as
# a result.
fs_used_grf <- isTRUE(fs$args_call_all$use_grf)

cat("\nForestSearch completed in",
    round(timings$forestsearch, 1), "seconds\n")
```

```{r fs-param-provenance}
#| echo: false
fs_param_table(fs, "Main ForestSearch fit: resolved parameters")
```




## ForestSearch Results

### Identified Subgroups

```{r fs-results}
# Generate results tables
res_tabs <- sg_tables(fs, ndecimals = 3, which_df = "est")

# Display top subgroups meeting criteria
res_tabs$sg10_out
```

### Treatment Effect Estimates

```{r fs-estimates}
# ITT and subgroup estimates
res_tabs$tab_estimates
```

### Identified Subgroup Definition

```{r fs-subgroup}
cat("Identified subgroup (H):", paste(fs$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs$df.est$treat.recommend == 0)))
```

ForestSearch identifies **Estrogen ≤ 0** (ER-negative) as the subgroup with
potential harm. This is biologically plausible: tamoxifen is a selective
estrogen receptor modulator with limited efficacy in ER-negative tumors.

### Selection Diagnostic: Pareto Frontier

Beyond the single subgroup chosen by `sg_focus`, it is often informative to
inspect the *Pareto frontier* on (effect, N) -- the set of candidate subgroups
that are not dominated on both axes simultaneously. A candidate is
dominated when another candidate has at least as large an effect *and*
at least as large a sample size, with at least one strict inequality.
The frontier is a post-hoc diagnostic that lists genuine trade-off
alternatives to the selected subgroup; it does **not** affect selection.

For `"hrMinSG"` the selected subgroup may not appear on the frontier --
that focus deliberately prefers small subgroups, which can be N-dominated
by their larger neighbors.

```{r fs-pareto}
# Formatted gt table; selected subgroup is highlighted and marked with ★.
# Works uniformly for survival (HR) and GLM (OR, RR, IRR, RD, IRD, MD)
# outcomes -- effect column label and scale derived from fs$effect_measure.
pareto_frontier_table(fs)
```

```{r fs-pareto-data, eval = FALSE}
# Programmatic access (effect on natural scale, plus is_selected flag):
pareto_frontier_table(fs, format = "data.table")
```

### Agreement with GRF

ForestSearch and GRF identify subgroups by different mechanisms --
FS via exhaustive search with consistency filtering, GRF via causal
survival forests plus policy trees -- so agreement on subject-level
membership provides an informal robustness check.  `compare_fs_subgroups()`
produces a publication-ready cross-tabulation plus Cohen's kappa; it
accepts forestsearch and GRF result objects directly, and handles
non-detection (method identifies no subgroup) by assigning all
subjects to the complement.

```{r fs-grf-concordance}
sg_compare <- compare_fs_subgroups(
  fs1               = fs,
  fs2               = grf_est,
  label1            = "ForestSearch",
  label2            = "GRF Standalone",
  id.name           = id.name,
  subgroup_notation = "harm",
  title             = "FS vs GRF: Subgroup Membership Concordance",
  subtitle          = "GBSG observed data (Cox, harm search)",
  font_size         = 13
)
sg_compare$table
```

```{r fs-grf-concordance-stats}
cat("Concordance statistics:\n")
cat(sprintf("  Agreement:       %.1f%%\n",
    100 * sg_compare$concordance$agreement))
cat(sprintf("  Cohen's kappa:   %.3f\n",
    sg_compare$concordance$kappa))
cat(sprintf("  Concordant:      %d / Discordant: %d / Total: %d\n",
    sg_compare$concordance$n_agree,
    sg_compare$concordance$n_disagree,
    sg_compare$concordance$n_total))
```


# Bootstrap Bias Correction

## Rationale

Cox model estimates from identified subgroups are **upwardly biased** due to
the selection process (subgroups are selected *because* they show extreme effects).
Bootstrap bias correction addresses this by:

1. Resampling with replacement
2. Re-running the entire ForestSearch algorithm
3. Computing bias terms from bootstrap vs. observed estimates
4. Applying infinitesimal jackknife variance estimation

## Running Bootstrap Analysis

```{r bootstrap}
#| message: false
# message: false suppresses forestsearch_bootstrap_dofuture()'s per-iteration
# progress and its "Low events in NEW subgroup" notes (both message(), not
# cat()).  Roughly 200 lines of console noise in a NB = 1000 run; the
# low-event condition is summarised properly in the diagnostics chunk below.

t0 <- proc.time()

fs_bc <- forestsearch_bootstrap_dofuture(
  # Pinned to the package default.  The fit carries mr_inference = run_mr
  # (TRUE here), and each replicate/fold call copies the fit's argument
  # list, so without this the flag is inherited and MR runs inside every
  # replicate.  Nothing downstream consumes a per-replicate MR object.
  # No-op as written.
  mr_in_replicates = FALSE,
  fs.est = fs,
  nb_boots = NB,
  show_three = FALSE,
  details = FALSE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nBootstrap completed in",
    round(timings$bootstrap / 60, 1), "minutes\n")
```

## Bootstrap Summary and Diagnostics

```{r bootstrap-summary}
# Comprehensive summary with diagnostics
summaries <- summarize_bootstrap_results(
  sgharm = fs$sg.harm,
  boot_results = fs_bc,
  create_plots = TRUE,
  est.scale = "hr"
)

# Display bias-corrected estimates table
summaries$table
```

## MR vs. FB

The MR is a fast multiplier-bootstrap approximation of the
bootstrap bias-correction. It reuses per-candidate Cox influence (`dfbeta`) over
the selection engine's **own** candidate family — the enumerated consistency
candidates, DINA's qualifying cuts, or GRF's DR-score frontier, depending on
`fs_subgroup_method`. Its selection-bias term is computed over that family
under multiplier draws, mirroring the same search optimism the full bootstrap
corrects — without re-running the search per resample. The subgroup reported
throughout remains the one the fit selected; nothing below re-identifies it. The table below compares
MR (computed inside the fit) against the full bootstrap from the previous
section. The naive-HR row is a scale sanity check; agreement on the de-biased
row is the MR-vs-FB validation. By default the de-biased CI uses the same
infinitesimal-jackknife variance as the full bootstrap (the `ci_method = 'ij'` default),
so the de-biased row compares both point estimate and interval like-for-like. The MR de-biases both the selected
subgroup and its complement, so both appear in the comparison and the table below.

```{r debias-MR-comparison, eval = !is.null(fs$sg.harm)}
#| echo: false
#| results: asis
g <- fs$mr_inference
if (is.null(g)) {
  cat("_The MR did not run (set `run_mr <- TRUE`)._\n")
} else {
  fmt_ci <- function(e, l, u) sprintf("%.3f (%.3f, %.3f)", e, l, u)
  fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0  <- fl(fs_bc$H_estimates, "H0"); H0l <- fl(fs_bc$H_estimates, "H0_lower"); H0u <- fl(fs_bc$H_estimates, "H0_upper")
  H2  <- fl(fs_bc$H_estimates, "H2"); H2l <- fl(fs_bc$H_estimates, "H2_lower"); H2u <- fl(fs_bc$H_estimates, "H2_upper")

  # Complement: MR from g$complement (include_complement = TRUE), FB from
  # the bootstrap's Hc_estimates.
  cc  <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  Hc0 <- fl(fs_bc$Hc_estimates, "H0"); Hc0l <- fl(fs_bc$Hc_estimates, "H0_lower"); Hc0u <- fl(fs_bc$Hc_estimates, "H0_upper")
  Hc2 <- fl(fs_bc$Hc_estimates, "H2"); Hc2l <- fl(fs_bc$Hc_estimates, "H2_lower"); Hc2u <- fl(fs_bc$Hc_estimates, "H2_upper")

  q  <- c("naive HR (95% CI)", "de-biased HR (95% CI)")
  mr <- c(fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  fb <- c(fmt_ci(H0, H0l, H0u), fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    q  <- c(q, "complement naive HR (95% CI)", "complement de-biased HR (95% CI)")
    mr <- c(mr, fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    fb <- c(fb, fmt_ci(Hc0, Hc0l, Hc0u), fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `MR` = mr,
                     `FB` = fb, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE,
        caption = sprintf("Selected subgroup: %s%s", paste(fs$sg.harm, collapse = " & "),
                          if (has_cc) " (complement = remaining subjects)" else "")),
      sep = "\n"); cat("\n\n")

  # Scale cross-check: MR's naive HR and the bootstrap's naive H0 are the
  # same quantity (raw subgroup HR) computed two ways; they should agree.
  if (is.finite(H0) && is.finite(g$naive$est)) {
    scale_ok <- abs(log(g$naive$est) - log(H0)) < 1e-2
    cat(sprintf("Scale cross-check (naive MR vs bootstrap H0): **%s** (%.3f vs %.3f).\n\n",
                if (scale_ok) "PASS" else "CHECK", g$naive$est, H0))
  }

  # Family is the full enumerated space, not the screened survivors.
  n_surv <- tryCatch(nrow(fs$find.grps$out.found$hr.subgroups), error = function(e) NA_integer_)
  cat(sprintf("Candidate family: %d (full <= maxk space)%s.\n\n", g$n_family,
              if (!is.na(n_surv)) sprintf("; screened survivors = %d", n_surv) else ""))

  cat(sprintf("Selection rule replayed within multiplier draws = `%s`%s.\n\n",
              if (!is.null(g$settings$reselection)) g$settings$reselection else "?",
              if (fs_subgroup_method %in% c("dina", "grf"))
                sprintf(" (a Cox-effect proxy for the %s; native-family band = %s)",
                        if (identical(fs_subgroup_method, "dina")) "DINA tau-hat" else "GRF DR-score",
                        if (is.null(fs_family_native_nbhd)) "default (matched to effect_neighborhood)"
                        else if (fs_family_native_nbhd >= 1) "off (full family)"
                        else sprintf("%.2f", fs_family_native_nbhd))
              else " (matches the consistency search statistic)"))

  boot_s <- as.numeric(timings$bootstrap)
  cat(sprintf("Timing: FB bootstrap (%d boots) %.1f s (%.1f min); MR %.2f s",
              NB, boot_s, boot_s / 60, g$timing_seconds))
  if (is.finite(boot_s) && g$timing_seconds > 0)
    cat(sprintf(" -> **%.0fx** faster (multiplier loop only; the full bootstrap scales with `NB`, MR does not)", boot_s / g$timing_seconds))
  cat(".\n\n")

  cat(sprintf("MR `mr_harm_confirmed` = **%s** (de-biased HR %.3f %s MR %.2f); selection bias %.4f (log-HR).\n",
              fs$mr_harm_confirmed, g$debiased$est,
              if (g$settings$confirm_rule == "point") "vs" else "lower-bound vs",
              g$settings$t_confirm, g$selection_bias))
  if (!is.null(g$ci_method) && !is.null(g$debiased$se_ij))
    cat(sprintf("De-biased CI: %s (IJ SE %.3f vs robust SE %.3f, log-HR).\n",
                if (identical(g$ci_method, "ij")) "infinitesimal jackknife (FB analogue)" else "robust SE",
                g$debiased$se_ij, g$debiased$se_wald))
  if (has_cc)
    cat(sprintf("Complement (n = %d): naive HR %.3f -> de-biased %.3f (selection bias %.4f, log-HR).\n",
                cc$n, cc$naive$est, cc$debiased$est, cc$selection_bias))
}
```

```{r debias-MR-no-subgroup, eval = is.null(fs$sg.harm)}
#| echo: false
#| results: asis
cat("_No harm subgroup was selected, so the MR comparison does not apply._\n")
```

The same bias-corrected estimates table, rebuilt from the MR object. The
descriptive columns and naive effect are the same under either correction
(inherited from the bootstrap result used as a template); the bias-corrected
column carries MR's
de-biased estimate for **both** the selected subgroup and its complement.

```{r debias-MR-table, eval = !is.null(fs$sg.harm) && !is.null(fs$mr_inference)}
mr_estimates_table(fs, boot_results = fs_bc, est.scale = "hr", digits = 2)
```

## Kaplan-Meier by Identified Subgroups

```{r fs-results-figure}
#| fig-width: 10
#| fig-height: 7
#| fig-cap: "Kaplan-Meier survival curves by identified subgroup"

km_result <- plot_sg_weighted_km(
  fs.est = fs,
  outcome.name = "time_months",
  event.name = "status",
  treat.name = "hormon",
  show.logrank = FALSE,
  conf.int = TRUE,
  by.risk = 12,
  show.cox = FALSE, show.cox.bc = TRUE,
  fs_bc = fs_bc,
  hr_bc_position = "topright"
)
```

::: {.callout-note}
`r figure_note(km_result, prefix = "**Note:** ", custom_text = "Medians [95% CI] for arms are un-adjusted.")`
:::


### Event Count Summary

Low event counts can lead to unstable HR estimates. This summary helps
identify potential issues:

```{r event-summary}
event_summary <- summarize_bootstrap_events(fs_bc, threshold = 10)
```

### Bootstrap Diagnostics

```{r bootstrap-diagnostics}
# Quality metrics
summaries$diagnostics_table_gt
```

### Subgroup Agreement

How consistently does bootstrap identify the same subgroup?

```{r subgroup-agreement}
# Agreement with original analysis
if (!is.null(summaries$subgroup_summary$original_agreement)) {
  summaries$subgroup_summary$original_agreement
}

# Factor presence across bootstrap iterations
if (!is.null(summaries$subgroup_summary$factor_presence)) {
  summaries$subgroup_summary$factor_presence
}
```

### GRF-Cut Stability Across Bootstraps

```{r grf-cuts-skipped-note, eval = !fs_used_grf}
#| echo: false
#| results: asis
cat("_The main fit ran with `use_grf = FALSE`, so GRF contributed no candidate",
    "cuts and there is no GRF-cut stability to report. This block renders when",
    "the fit is built with `use_grf = TRUE`._\n")
```

```{r grf-cuts-note, eval = fs_used_grf}
#| echo: false
#| results: asis
cat(paste(
  "Separately from whether bootstrap identifies the *same subgroup*, we can",
  "ask what raw cut GRF itself surfaced on each resample, independent of",
  "whether the downstream consistency stage accepted it.  A dominant row",
  "here paired with a low `(no GRF cut)` rate indicates that GRF's",
  "upstream discovery is stable across resamples; a frequent",
  "`(no GRF cut)` row indicates GRF itself is the source of",
  "variability.  This distinction matters because the two failure modes",
  "call for different tuning responses -- relaxing `dmin.grf` /",
  "`frac.tau` (for \"GRF produced no cut\") vs. relaxing",
  "`pconsistency.threshold` (for \"GRF produced a cut that consistency",
  "rejected\")."), "\n")
```

```{r bootstrap-grf-cuts, eval = fs_used_grf}
if (!is.null(summaries$subgroup_summary$grf_cut_freq)) {
  summaries$subgroup_summary$grf_cut_freq |>
    utils::head(10) |>
    gt::gt() |>
    gt::tab_header(
      title = gt::md("**GRF cuts across bootstrap iterations**"),
      subtitle = "Top 10 by frequency; spans all bootstraps (not just successful ones)"
    ) |>
    gt::cols_label(
      Rank    = "Rank",
      grf_cut = "GRF cuts",
      N       = "N",
      Percent = "% of bootstraps"
    ) |>
    gt::fmt_number(columns = "Percent", decimals = 2) |>
    gt::tab_options(table.font.size = gt::px(13))
}
```

### Bootstrap Distributions

```{r bootstrap-plots}
#| fig-width: 10
#| fig-height: 4

if (!is.null(summaries$plots)) {
  summaries$plots$H_distribution + summaries$plots$Hc_distribution
}
```


# DINA-Based Subgroup Identification

DINA (Doubly-robust Identification of Non-Additive effects) screens and
selects subgroups from an influence-function representation of the
covariate-specific treatment effect, fit with cross-fitting. It plugs into
the ForestSearch workflow in three complementary ways, all configured to
mirror the main Cox analysis above: same confounders, same minimum size
(`n.min = 60`), same `sg_focus`. The harm threshold is the one thing they do
*not* inherit from the main fit -- each track sets its own in the setup chunk
(`dina_screen_hr_threshold` for (A), `fs_dina_hr_threshold` for (B) and (C)),
because DINA's model-based effect is attenuated relative to the raw Cox HR. The
resolved values are in the per-fit provenance tables below.

* **(A)** ForestSearch *using DINA for screening* -- DINA's per-covariate
  frontier contributes candidate cuts to the screening pool; the consistency
  search still performs selection.
* **(B)** *Standalone DINA* with parameters analogous to (A) -- the same
  fit's frontier (the candidates (A) screens) and its single selected
  subgroup via `dina_subgroup()`.
* **(C)** ForestSearch with `subgroup_method = "dina"` -- subgroup
  *selection* is delegated to `dina_subgroup()`, then the standard
  ForestSearch bootstrap provides the selection bias-corrected Cox estimate.

```{r dina-config}
#| code-fold: false

# Shared DINA *fit* settings (Cox / survival outcome => family mapped to "cox").
# Selection parameters (sg_focus, hr_threshold) live in the setup chunk under
# the fs_dina_* knobs so they can be tuned independently; (B) and (C) read them.
dina_seed   <- 8316951
dina_nfolds <- 2
# Harm/benefit floor on the natural-parameter (log-HR) scale for ratio measures
m_diff_dina <- log(fs_dina_hr_threshold)
```

## (A) ForestSearch Using DINA for Screening

Here ForestSearch is run exactly as in the main analysis above, but with
`use_dina = TRUE`: DINA's frontier cuts are merged into the candidate pool
alongside the GRF and LASSO cuts. `dina_args` is a flat, fully-named list;
in the screening role it supplies the DINA *fit* tuning (`seed`, `n_folds`)
and the *frontier* controls (`scope`, `max_per_covariate`, `max_subgroups`).
`scope = "wide"` casts a deliberately wide net -- the consistency stage
remains the gatekeeper. By default `dina_args$selected_only = TRUE` (matching
GRF's `return_selected_cuts_only`), so screening contributes the single cut
`dina_subgroup()` would select -- using this call's `sg_focus` /
`selection_rule` / `effect_neighborhood` / `hr.threshold`, i.e. the same cut
`subgroup_method = "dina"` (C) chooses. Set `dina_args$selected_only = FALSE`
to contribute the full frontier instead (more candidates).

```{r fs-dina-screen}
t0 <- proc.time()

fs_dina_screen <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis,
  outcome.name              = outcome.name,
  event.name                = event.name,
  treat.name                = treat.name,
  id.name                   = id.name,
  confounders.name          = confounders.name,
  potentialOutcome.name     = NULL,           # default NULL
  flag_harm.name            = NULL,           # default NULL
  df.predict                = NULL,           # default NULL
  df.test                   = NULL,           # default NULL
  is.RCT                    = TRUE,           # default TRUE
  seedit                    = 8316951,        # default 8316951
  est.scale                 = "hr",           # default "hr"
  # --- Screening engines: GRF + LASSO + DINA ------------------------------
  use_lasso                 = TRUE,           # default TRUE
  use_grf                   = TRUE,           # default TRUE
  grf_res                   = NULL,           # default NULL
  grf_cuts                  = NULL,           # default NULL
  grf_depth                 = 2,              # default 2
  dmin.grf                  = 4,              # default 0.0
  frac.tau                  = 0.80,           # default 0.8
  vi.grf.min                = -0.2,           # default -0.2
  return_selected_cuts_only = TRUE,           # default TRUE
  tune_grf                  = FALSE,          # default FALSE
  use_dina                  = TRUE,           # default FALSE
  dina_res                  = NULL,           # default NULL
  dina_cuts                 = NULL,           # default NULL
  dina_args = list(                           # default list()
    seed = dina_seed, n_folds = dina_nfolds,
    scope = "wide", max_per_covariate = 3, max_subgroups = 10,
    selected_only = dina_screen_selected_only # set in setup chunk
  ),
  # --- Candidate-cut construction -----------------------------------------
  max_n_confounders         = 1000,           # default 1000
  conf_force                = NULL,           # default NULL
  defaultcut_names          = NULL,           # default NULL
  cut_type                  = "default",      # default "default"
  exclude_cuts              = NULL,           # default NULL
  replace_med_grf           = FALSE,          # default FALSE
  cont.cutoff               = 4,              # default 4
  conf.cont_medians         = NULL,           # default NULL
  conf.cont_medians_force   = NULL,           # default NULL
  conf.cont_jcuts           = NULL,           # default NULL
  # --- Selection engine ---------------------------------------------------
  subgroup_method           = "consistency",  # default "consistency"; or "dina"
  sg_focus                  = dina_screen_sg_focus,    # set in setup chunk
  selection_rule            = "neighborhood",  # default "neighborhood"
  effect_neighborhood       = 0.10,            # default 0.10
  # --- Subgroup constraints -----------------------------------------------
  n.min                     = 60,             # default 60
  d0.min                    = 10,             # default 10
  d1.min                    = 10,             # default 10
  maxk                      = 2,              # default 2
  max_subgroups_search      = dina_screen_max_subgroups_search,  # set in setup chunk
  # --- Thresholds (survival / HR scale) -----------------------------------
  hr.threshold              = dina_screen_hr_threshold,    # set in setup chunk
  hr.consistency            = dina_screen_hr_consistency,  # set in setup chunk
  effect.threshold          = NULL,           # default NULL (alias of hr.threshold)
  consistency.threshold     = NULL,           # default NULL (alias of hr.consistency)
  pconsistency.threshold    = 0.90,           # default 0.90
  consistency_method        = fs_consistency_method,  # ALIGN: match main run + ACTG175 (MR/consistency parity)
  m1.threshold              = Inf,            # default Inf
  minp                      = 0.025,          # default 0.025
  # --- Consistency search -------------------------------------------------
  fs.splits                 = 1000,           # default 1000
  stop_threshold            = NULL,           # NULL is what this configuration RESOLVES to, not a
                                              #   preference.  This call's sg_focus ("effMaxSG")
                                              #   normalizes to "hrMaxSG", and forestsearch_main.R:1403
                                              #   resets stop_threshold to NULL for that focus: a
                                              #   neighborhood band needs every candidate evaluated,
                                              #   so a truncated scan is unsound.  Any non-NULL value
                                              #   reaches the same NULL but warns on every render;
                                              #   passing NULL skips the reset branch outright (its
                                              #   guard is !is.null(stop_threshold)).  Named here so
                                              #   the resolved setting stays inspectable.
  use_twostage              = TRUE,           # default TRUE
  twostage_args             = list(),         # default list()
  max.minutes               = 3,              # default 3
  # --- Output / diagnostics -----------------------------------------------
  show_candidate_summary    = TRUE,           # default FALSE
  details                   = TRUE,           # default FALSE
  quiet                     = FALSE,          # default FALSE
  by.risk                   = 12,             # default 12
  plot.sg                   = FALSE,          # default FALSE
  plot.grf                  = FALSE,          # default FALSE
  # --- Parallel -----------------------------------------------------------
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$fs_dina_screen <- (proc.time() - t0)["elapsed"]
stopifnot(!is.null(fs_dina_screen), !is.null(fs_dina_screen$args_call_all))

cat("Identified subgroup (H):",
    paste(fs_dina_screen$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs_dina_screen$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs_dina_screen$df.est$treat.recommend == 0)))
```

```{r fs-dina-screen-param-provenance}
#| echo: false
fs_param_table(fs_dina_screen, "(A) DINA-screened fit: resolved parameters")
```

The cuts DINA contributed to the candidate pool are recorded on the result
object. These are exactly the frontier candidates reproduced standalone in
(B).

```{r fs-dina-screen-cuts}
dina_screen_cuts <- fs_dina_screen$dina_cuts
if (length(dina_screen_cuts) > 0) {
  gt(data.frame(
    `#` = seq_along(dina_screen_cuts),
    `DINA candidate cut` = dina_screen_cuts,
    check.names = FALSE
  )) |>
    tab_header(title = "DINA Screening Candidates Injected into ForestSearch") |>
    cols_align(align = "left")
} else {
  cat("DINA contributed no candidate cuts under the configured frontier.\n")
}
```

```{r fs-dina-screen-pareto}
# Selection diagnostic for the DINA-screened run (selected subgroup marked).
pareto_frontier_table(fs_dina_screen)
```

## (B) Standalone DINA

The same DINA fit, run outside ForestSearch. Its `dina_frontier()` output is
the candidate set fed into screening in (A); its `dina_subgroup()` output is
the single subgroup selected by DINA itself -- the mechanism (C) uses
internally. Parameters are analogous to (A): same confounders, `n_min = 60`,
and `sg_focus`, with the harm floor `m_diff = log(fs_dina_hr_threshold)` set in
the dina-config chunk below (so (B) and (C) share one floor by construction).

```{r dina-standalone-fit}
t0 <- proc.time()

dina_fit <- dina(
  df         = df.analysis,
  outcome    = outcome.name,
  treatment  = treat.name,
  covariates = confounders.name,
  family     = "cox",          # default "gaussian"; gaussian/binomial/poisson/cox
  status     = event.name,     # default NULL (required for family = "cox")
  # ... forwarded to dina_fit():
  seed              = dina_seed,   # default NULL
  n_folds           = dina_nfolds, # default 2L
  propensity_method = "logistic",  # default "logistic"
  baseline_method   = NULL,        # default NULL
  cross_fitting     = TRUE,        # default TRUE
  cens_type         = NULL,        # default NULL (no censoring model)
  cens_params       = list(),      # default list()
  n_grid            = 500L,        # default 500L
  eps               = 1e-6         # default 1e-6
)

timings$dina_standalone <- (proc.time() - t0)["elapsed"]
cat("Standalone DINA fit complete (family = cox).\n")
```

### DINA Frontier (Screening Candidates)

```{r dina-standalone-frontier}
dina_fr <- dina_frontier(
  fit               = dina_fit,
  df                = df.analysis,
  covariates        = confounders.name,
  scope             = "wide",     # default "wide"; or "harm" (requires m_diff)
  m_diff            = NULL,        # default NULL (no effect floor under scope = "wide")
  n_min             = 60,         # default 60L
  direction         = "both",     # default "both"; or "left" / "right"
  max_per_covariate = 3,          # default 3L
  max_subgroups     = 10,         # default 10L
  digits            = 3           # default 3L
)

gt(dina_fr[, c("covariate", "direction", "threshold",
               "n_subgroup", "effect", "cut_expr")]) |>
  tab_header(
    title = "Standalone DINA Frontier",
    subtitle = "Per-covariate non-dominated (effect, N) candidates; effect on HR scale"
  ) |>
  fmt_number(columns = c(threshold, effect), decimals = 3) |>
  cols_label(
    covariate = "Covariate", direction = "Direction",
    threshold = "Threshold", n_subgroup = "N", effect = "HR",
    cut_expr = "Cut"
  ) |>
  cols_align(align = "left", columns = c(covariate, direction, cut_expr))
```

### DINA-Selected Subgroup

```{r dina-standalone-select}
dina_sel <- dina_subgroup(
  fit                 = dina_fit,
  df                  = df.analysis,
  covariates          = confounders.name,
  m_diff              = m_diff_dina,
  n_min               = 60,             # default 60L
  direction           = "both",         # default "both"; or "left" / "right"
  sg_focus            = fs_dina_sg_focus,  # set in setup chunk
  selection_rule      = "neighborhood",  # default "neighborhood"; pareto/both (hrMaxSG/hrMinSG)
  effect_neighborhood = 0.10,            # default 0.10
  alpha               = 0.05             # default 0.05
)

if (isTRUE(dina_sel$found)) {
  op <- if (identical(dina_sel$direction, "left")) "<=" else ">="
  cat(sprintf("DINA-selected subgroup: %s %s %s\n",
              dina_sel$covariate, op, format(dina_sel$threshold)))
  cat(sprintf("Direction: %s | N = %d\n",
              dina_sel$direction, dina_sel$n_subgroup))
} else {
  cat("Standalone DINA selected no subgroup meeting the criteria.\n")
}
```

## (C) ForestSearch with `subgroup_method = "dina"`

In this mode ForestSearch delegates subgroup *selection* to
`dina_subgroup()`, bypassing GRF, LASSO and the consistency search. The
selection criteria are inherited from this call's `sg_focus`, `n.min` and
`hr.threshold` (mapped to the log-HR harm floor); `dina_args` supplies only
the fit tuning. The returned object is a standard `forestsearch` result, so
the existing bootstrap machinery applies unchanged.

```{r fs-dina-select}
t0 <- proc.time()

fs_dina <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis,
  outcome.name        = outcome.name,
  event.name          = event.name,
  treat.name          = treat.name,
  id.name             = id.name,
  confounders.name    = confounders.name,
  df.predict          = NULL,            # default NULL
  df.test             = NULL,            # default NULL
  is.RCT              = TRUE,            # default TRUE
  seedit              = 8316951,         # default 8316951
  # --- DINA-selection mode ------------------------------------------------
  subgroup_method     = "dina",          # default "consistency"
  dina_args           = list(seed = dina_seed, n_folds = dina_nfolds),  # fit tuning only
  dina_select_statistic = fs_dina_select_statistic,  # setup-chunk selector (top-level)
  dina_res            = NULL,            # default NULL (reuse a fitted dina object)
  # --- Selection criteria (passed through to dina_subgroup) ---------------
  sg_focus            = fs_dina_sg_focus,  # set in setup chunk
  selection_rule      = "neighborhood",  # default "neighborhood"
  effect_neighborhood = 0.10,            # default 0.10
  n.min               = 60,              # default 60
  hr.threshold        = fs_dina_hr_threshold,  # set in setup chunk (-> log-HR floor)
  # --- MR ----------------------------------------------
  mr_inference         = run_mr,           # set in setup chunk
  mr_inference_args    = list(draws = mr_draws,
                             family_native_neighborhood = fs_dina_family_native_nbhd),
  # --- Output -------------------------------------------------------------
  details             = TRUE             # default FALSE
  # NOTE: subgroup_method = "dina" bypasses GRF, LASSO and the consistency
  # search; use_grf / use_lasso / use_twostage / fs.splits / maxk /
  # stop_threshold and related parameters do not apply in this mode.
)

timings$fs_dina_select <- (proc.time() - t0)["elapsed"]
stopifnot(!is.null(fs_dina), !is.null(fs_dina$args_call_all))

cat("DINA-selected subgroup (H):",
    paste(fs_dina$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs_dina$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs_dina$df.est$treat.recommend == 0)))
```

```{r fs-dina-param-provenance}
#| echo: false
fs_param_table(fs_dina, "(C) subgroup_method = dina: resolved parameters")
```

Under `subgroup_method = "dina"` the consistency search, LASSO and GRF screens
are bypassed. The table still lists every formal, so the inert ones appear as
resolved defaults rather than silently absent.

### Bootstrap Bias-Corrected Cox Estimates

The within-subgroup Cox hazard ratio is upwardly biased by selection. The
ForestSearch bootstrap corrects this by re-running the full pipeline on each
resample; because the result above was produced in DINA-selection mode, each
replicate re-fits DINA and re-runs `dina_subgroup()` from scratch -- so the
correction reflects the DINA selection process itself.

```{r fs-dina-bootstrap, eval = !is.null(fs_dina$sg.harm)}
#| message: false
# message: false suppresses forestsearch_bootstrap_dofuture()'s per-iteration
# progress and its "Low events in NEW subgroup" notes (both message(), not
# cat()).  Roughly 200 lines of console noise in a NB = 1000 run; the
# low-event condition is summarised properly in the diagnostics chunk below.
# Runs only when DINA identified a subgroup.  The bootstrap bias-corrects an
# identified subgroup's effect, so there is nothing to do otherwise.
t0 <- proc.time()

# details = TRUE keeps per-iteration output visible while validating DINA-mode
# bootstrap.  The plan is "multisession" (production speed); switch it to
# "sequential" if errors need to surface in loop order.

fs_dina_bc <- forestsearch_bootstrap_dofuture(
  # Pinned to the package default, as at the main bootstrap above -- see the
  # comment there for why this is a no-op.
  mr_in_replicates = FALSE,
  fs.est = fs_dina,
  nb_boots = NB,
  show_three = FALSE,
  details = TRUE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$fs_dina_bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nDINA-mode bootstrap completed in",
    round(timings$fs_dina_bootstrap / 60, 1), "minutes\n")
```

```{r fs-dina-no-subgroup, eval = is.null(fs_dina$sg.harm)}
# Shown only when DINA selected no subgroup at the configured harm floor.
timings$fs_dina_bootstrap <- 0
cat("DINA (subgroup_method = \"dina\") identified no subgroup at",
    "hr.threshold =", fs_dina_hr_threshold, "-- bootstrap skipped.\n",
    "DINA's model-based effect is attenuated relative to the raw Cox HR,",
    "so this floor may not be reached; see the standalone DINA frontier (B)",
    "for the achievable effect range, and lower the DINA harm floor if a",
    "bootstrap demonstration is desired.\n")
```

```{r fs-dina-bootstrap-summary, eval = !is.null(fs_dina$sg.harm)}
dina_summaries <- summarize_bootstrap_results(
  sgharm = fs_dina$sg.harm,
  boot_results = fs_dina_bc,
  create_plots = FALSE,
  est.scale = "hr"
)

dina_summaries$table
```

#### MR (counterpart to FB)

The same MR-vs-FB comparison as the main analysis, for the
DINA-selection fit. The MR de-biases against DINA's **own** qualifying
candidate family (re-running nothing), while the bootstrap above re-runs DINA on
every resample.

For DINA, FB and MR are **not** expected to agree as tightly as for the
consistency method. DINA selects on its survival-forest **tau-hat**, but MR's de-biasing is
computed on the perturbed **Cox** effect; over a large depth-2 candidate family
the winner's-curse term can therefore exceed the bootstrap's and over-correct
the selected subgroup. To keep that band close to DINA's own selection band,
MR by default restricts its family to
candidates competitive on DINA's native tau-hat, matching the band to
`effect_neighborhood` (the recommended default, used here via
`fs_dina_family_native_nbhd <- NULL`; set an explicit number for a different band,
or `>= 1` to de-bias against the full unrestricted family). The complement is the
induced residual rather than a max-over-family selection, so FB and MR agree on
it regardless. This band tunes only the *candidate-competition* part of the
bias -- MR holds the survival forest fixed, so the forest-refit optimism is
invisible to it; treat the full bootstrap as the reference for DINA and MR
as a fast, conservative check.

```{r fs-dina-debias-MR-comparison, eval = !is.null(fs_dina$sg.harm)}
#| echo: false
#| results: asis
g <- fs_dina$mr_inference
if (is.null(g)) {
  cat("_The MR did not run (set `run_mr <- TRUE`)._\n")
} else {
  fmt_ci <- function(e, l, u) sprintf("%.3f (%.3f, %.3f)", e, l, u)
  fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0 <- fl(fs_dina_bc$H_estimates, "H0"); H0l <- fl(fs_dina_bc$H_estimates, "H0_lower"); H0u <- fl(fs_dina_bc$H_estimates, "H0_upper")
  H2 <- fl(fs_dina_bc$H_estimates, "H2"); H2l <- fl(fs_dina_bc$H_estimates, "H2_lower"); H2u <- fl(fs_dina_bc$H_estimates, "H2_upper")
  cc <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  q  <- c("naive HR (95% CI)", "de-biased HR (95% CI)")
  mr <- c(fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  fb <- c(fmt_ci(H0, H0l, H0u), fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    Hc0 <- fl(fs_dina_bc$Hc_estimates, "H0"); Hc0l <- fl(fs_dina_bc$Hc_estimates, "H0_lower"); Hc0u <- fl(fs_dina_bc$Hc_estimates, "H0_upper")
    Hc2 <- fl(fs_dina_bc$Hc_estimates, "H2"); Hc2l <- fl(fs_dina_bc$Hc_estimates, "H2_lower"); Hc2u <- fl(fs_dina_bc$Hc_estimates, "H2_upper")
    q  <- c(q, "complement naive HR (95% CI)", "complement de-biased HR (95% CI)")
    mr <- c(mr, fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    fb <- c(fb, fmt_ci(Hc0, Hc0l, Hc0u), fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `MR` = mr,
                     `FB` = fb, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE,
        caption = sprintf("DINA-selected subgroup: %s%s",
                          paste(fs_dina$sg.harm, collapse = " & "),
                          if (has_cc) " (complement = remaining subjects)" else "")),
      sep = "\n"); cat("\n\n")

  boot_s <- as.numeric(timings$fs_dina_bootstrap)
  cat(sprintf("Timing: FB bootstrap (%d boots) %.1f s (%.1f min); MR %.2f s",
              NB, boot_s, boot_s / 60, g$timing_seconds))
  if (is.finite(boot_s) && g$timing_seconds > 0)
    cat(sprintf(" -> **%.0fx** faster (the bootstrap scales with `NB`, MR does not)",
                boot_s / g$timing_seconds))
  cat(".\n\n")
  cat(sprintf("MR harm flag = **%s** (de-biased HR %.3f %s MR %.2f); selection bias %.4f (log-HR); candidate family = %d.\n",
              isTRUE(g$harm_flag), g$debiased$est,
              if (g$settings$confirm_rule == "point") "vs" else "lower-bound vs",
              g$settings$t_confirm, g$selection_bias, g$n_family))
  cat(sprintf("Selection rule replayed within multiplier draws = `%s` (a Cox-effect proxy for DINA's tau-hat); native-family band = %s.\n",
              if (!is.null(g$settings$reselection)) g$settings$reselection else "?",
              if (is.null(fs_dina_family_native_nbhd)) "default (matched to effect_neighborhood)"
              else if (fs_dina_family_native_nbhd >= 1) "off (full family)"
              else sprintf("%.2f (restricted to tau-hat-competitive candidates)",
                           fs_dina_family_native_nbhd)))
  if (!is.null(g$ci_method) && !is.null(g$debiased$se_ij))
    cat(sprintf("De-biased CI: %s (IJ SE %.3f vs robust SE %.3f, log-HR).\n",
                if (identical(g$ci_method, "ij")) "infinitesimal jackknife (FB analogue)" else "robust SE",
                g$debiased$se_ij, g$debiased$se_wald))
}
```

The FB bias-corrected table above, rebuilt from the MR object (the
descriptive columns and naive effect are the same under either correction; the
bias-corrected column carries MR's de-biased estimate for the subgroup and its
complement).

```{r fs-dina-debias-MR-table, eval = !is.null(fs_dina$sg.harm) && !is.null(fs_dina$mr_inference)}
mr_estimates_table(fs_dina, boot_results = fs_dina_bc, est.scale = "hr", digits = 2)
```

::: {.callout-note}
## Screening vs. selection roles for DINA

`dina_args` behaves differently by role. In the **screening** path (A,
`use_dina = TRUE`) it defaults to `selected_only = TRUE` (matching GRF's
`return_selected_cuts_only`): the screening path calls `dina_subgroup()` and
contributes only its selected cut, so it *does* consult this call's
`sg_focus` / `selection_rule` / `effect_neighborhood` -- yielding the same
cut the selection path (C) would choose. Setting `selected_only = FALSE`
instead drives `dina_frontier()`, which is selection-agnostic (`sg_focus` /
`selection_rule` are *not* consulted) and contributes the full candidate set.
In the **selection** path (C, `subgroup_method = "dina"`) the selection is
always driven by this call's `sg_focus` / `selection_rule` /
`effect_neighborhood` via `dina_subgroup()`, and `dina_args` supplies only
the fit. This mode is currently depth-1 (a
single covariate threshold).
:::



## (D) ForestSearch with `subgroup_method = "grf"`

In this mode ForestSearch delegates subgroup *selection* to GRF -- a causal-forest
policy tree (`grf_selection = "tree"`) or the doubly-robust-score frontier
(`"frontier"`) -- bypassing LASSO, DINA and the consistency search. The selection
criteria are inherited from this call's `sg_focus` (mapped to the GRF frontier/tree
rule), `n.min`, `grf_depth`, `grf_selection` and the DR-score harm floor
`dmin.grf`. The returned object is a standard `forestsearch` result, so the
existing bootstrap machinery applies unchanged.

```{r fs-grf-select}
t0 <- proc.time()

fs_grf <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis,
  outcome.name        = outcome.name,
  event.name          = event.name,
  treat.name          = treat.name,
  id.name             = id.name,
  confounders.name    = confounders.name,
  df.predict          = NULL,            # default NULL
  df.test             = NULL,            # default NULL
  is.RCT              = TRUE,            # default TRUE
  seedit              = 8316951,         # default 8316951 (also seeds the forest)
  # --- GRF-selection mode -------------------------------------------------
  subgroup_method     = "grf",           # default "consistency"
  grf_selection       = fs_grf_demo_selection,  # set in setup chunk ("tree" | "frontier")
  grf_select_statistic = fs_grf_select_statistic,  # set in setup chunk (frontier only)
  grf_depth           = fs_grf_demo_depth,      # set in setup chunk (tree / DR-pair depth)
  dmin.grf            = fs_grf_demo_dmin,        # set in setup chunk (DR-score harm floor)
  frac.tau            = 0.8,             # default 0.8 (RMST truncation fraction)
  grf_res             = NULL,            # default NULL (reuse a fitted forest)
  # --- Selection criteria -------------------------------------------------
  sg_focus            = fs_grf_sg_focus,  # set in setup chunk (-> GRF frontier/tree rule)
  selection_rule      = "neighborhood",  # default "neighborhood"
  effect_neighborhood = 0.10,            # default 0.10
  n.min               = 60,              # default 60
  # --- MR ----------------------------------------------
  mr_inference         = run_mr,           # set in setup chunk
  mr_inference_args    = list(draws = mr_draws,
                             family_native_neighborhood = fs_grf_family_native_nbhd),
  # --- Output -------------------------------------------------------------
  details             = TRUE             # default FALSE
  # NOTE: subgroup_method = "grf" bypasses LASSO, DINA and the consistency
  # search; use_lasso / use_dina / use_twostage / fs.splits / maxk /
  # hr.threshold / pconsistency.threshold do not apply in this mode.
)

timings$fs_grf_select <- (proc.time() - t0)["elapsed"]
stopifnot(!is.null(fs_grf), !is.null(fs_grf$args_call_all))

cat("GRF-selected subgroup (H):",
    paste(fs_grf$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs_grf$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs_grf$df.est$treat.recommend == 0)))
```

```{r fs-grf-param-provenance}
#| echo: false
fs_param_table(fs_grf, "(D) subgroup_method = grf: resolved parameters")
```

### Bootstrap Bias-Corrected Cox Estimates

The within-subgroup Cox hazard ratio is upwardly biased by selection. The
ForestSearch bootstrap corrects this by re-running the full pipeline on each
resample; because the result above was produced in GRF-selection mode, each
replicate re-fits the causal forest and re-runs GRF selection from scratch -- so
the correction reflects the GRF selection process itself.

```{r fs-grf-bootstrap, eval = !is.null(fs_grf$sg.harm)}
#| message: false
# message: false suppresses forestsearch_bootstrap_dofuture()'s per-iteration
# progress and its "Low events in NEW subgroup" notes (both message(), not
# cat()).  Roughly 200 lines of console noise in a NB = 1000 run; the
# low-event condition is summarised properly in the diagnostics chunk below.
# Runs only when GRF identified a subgroup.  The bootstrap bias-corrects an
# identified subgroup's effect, so there is nothing to do otherwise.
t0 <- proc.time()

fs_grf_bc <- forestsearch_bootstrap_dofuture(
  # Pinned to the package default, as at the main bootstrap above -- see the
  # comment there for why this is a no-op.
  mr_in_replicates = FALSE,
  fs.est = fs_grf,
  nb_boots = NB,
  show_three = FALSE,
  details = TRUE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$fs_grf_bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nGRF-mode bootstrap completed in",
    round(timings$fs_grf_bootstrap / 60, 1), "minutes\n")
```

```{r fs-grf-no-subgroup, eval = is.null(fs_grf$sg.harm)}
# Shown only when GRF selected no subgroup at the configured DR-score floor.
timings$fs_grf_bootstrap <- 0
cat("GRF (subgroup_method = \"grf\") identified no subgroup at",
    "dmin.grf =", fs_grf_demo_dmin, "-- bootstrap skipped.\n",
    "Lower the DR-score floor (dmin.grf) or switch grf_selection /",
    "grf_depth if a bootstrap demonstration is desired.\n")
```

```{r fs-grf-bootstrap-summary, eval = !is.null(fs_grf$sg.harm)}
grf_summaries <- summarize_bootstrap_results(
  sgharm = fs_grf$sg.harm,
  boot_results = fs_grf_bc,
  create_plots = FALSE,
  est.scale = "hr"
)

grf_summaries$table
```

#### MR (counterpart to FB)

The same MR-vs-FB comparison as the main analysis, for the
GRF-selection fit. The MR de-biases against GRF's **own** DR-score candidate
family (re-running nothing), while the bootstrap above re-runs GRF on every
resample.

As with DINA, FB and MR are **not** expected to agree as tightly as for the
consistency method. GRF selects on its **doubly-robust (RMST-based) score**, but
MR's de-biasing is computed on the perturbed **Cox** effect; the additive RMST
scale and the multiplicative hazard scale need not rank candidates the same way,
so over a large candidate family the winner's-curse term can exceed the
bootstrap's and over-correct the selected subgroup. To keep that band close to
GRF's own selection band, MR by default restricts its family to candidates competitive on GRF's native DR-score, matching
the band to `effect_neighborhood` (the recommended default, used here via
`fs_grf_family_native_nbhd <- NULL`; set an explicit number for a different band,
or `>= 1` to de-bias against the full family). The complement is the induced
residual rather than a max-over-family selection, so FB and MR agree on it
regardless. Two caveats specific to GRF: a *dominant* DR-score winner can collapse
the restricted family to a single candidate (`candidate family = 1` in the note
below), in which case MR removes only MR-threshold optimism; and,
more fundamentally, the band tunes only the *candidate-competition* part of the
bias -- MR holds the causal forest fixed, so the forest-refit optimism that
the bootstrap captures is invisible to it. Treat the full bootstrap as the
reference for GRF and MR as a fast, conservative check.

```{r fs-grf-debias-MR-comparison, eval = !is.null(fs_grf$sg.harm)}
#| echo: false
#| results: asis
g <- fs_grf$mr_inference
if (is.null(g)) {
  cat("_The MR did not run (set `run_mr <- TRUE`)._\n")
} else {
  fmt_ci <- function(e, l, u) sprintf("%.3f (%.3f, %.3f)", e, l, u)
  fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0 <- fl(fs_grf_bc$H_estimates, "H0"); H0l <- fl(fs_grf_bc$H_estimates, "H0_lower"); H0u <- fl(fs_grf_bc$H_estimates, "H0_upper")
  H2 <- fl(fs_grf_bc$H_estimates, "H2"); H2l <- fl(fs_grf_bc$H_estimates, "H2_lower"); H2u <- fl(fs_grf_bc$H_estimates, "H2_upper")
  cc <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  q  <- c("naive HR (95% CI)", "de-biased HR (95% CI)")
  mr <- c(fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  fb <- c(fmt_ci(H0, H0l, H0u), fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    Hc0 <- fl(fs_grf_bc$Hc_estimates, "H0"); Hc0l <- fl(fs_grf_bc$Hc_estimates, "H0_lower"); Hc0u <- fl(fs_grf_bc$Hc_estimates, "H0_upper")
    Hc2 <- fl(fs_grf_bc$Hc_estimates, "H2"); Hc2l <- fl(fs_grf_bc$Hc_estimates, "H2_lower"); Hc2u <- fl(fs_grf_bc$Hc_estimates, "H2_upper")
    q  <- c(q, "complement naive HR (95% CI)", "complement de-biased HR (95% CI)")
    mr <- c(mr, fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    fb <- c(fb, fmt_ci(Hc0, Hc0l, Hc0u), fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `MR` = mr,
                     `FB` = fb, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE,
        caption = sprintf("GRF-selected subgroup: %s%s",
                          paste(fs_grf$sg.harm, collapse = " & "),
                          if (has_cc) " (complement = remaining subjects)" else "")),
      sep = "\n"); cat("\n\n")

  boot_s <- as.numeric(timings$fs_grf_bootstrap)
  cat(sprintf("Timing: FB bootstrap (%d boots) %.1f s (%.1f min); MR %.2f s",
              NB, boot_s, boot_s / 60, g$timing_seconds))
  if (is.finite(boot_s) && g$timing_seconds > 0)
    cat(sprintf(" -> **%.0fx** faster (the bootstrap scales with `NB`, MR does not)",
                boot_s / g$timing_seconds))
  cat(".\n\n")
  cat(sprintf("MR harm flag = **%s** (de-biased HR %.3f %s MR %.2f); selection bias %.4f (log-HR); candidate family = %d.\n",
              isTRUE(g$harm_flag), g$debiased$est,
              if (g$settings$confirm_rule == "point") "vs" else "lower-bound vs",
              g$settings$t_confirm, g$selection_bias, g$n_family))
  cat(sprintf("Selection rule replayed within multiplier draws = `%s` (a Cox-effect proxy for GRF's DR-score); native-family band = %s.\n",
              if (!is.null(g$settings$reselection)) g$settings$reselection else "?",
              if (is.null(fs_grf_family_native_nbhd)) "default (matched to effect_neighborhood)"
              else if (fs_grf_family_native_nbhd >= 1) "off (full family)"
              else sprintf("%.2f (restricted to DR-score-competitive candidates)",
                           fs_grf_family_native_nbhd)))
  if (!is.null(g$ci_method) && !is.null(g$debiased$se_ij))
    cat(sprintf("De-biased CI: %s (IJ SE %.3f vs robust SE %.3f, log-HR).\n",
                if (identical(g$ci_method, "ij")) "infinitesimal jackknife (FB analogue)" else "robust SE",
                g$debiased$se_ij, g$debiased$se_wald))
}
```

The FB bias-corrected table above, rebuilt from the MR object (the
descriptive columns and naive effect are the same under either correction; the
bias-corrected column carries MR's de-biased estimate for the subgroup and its
complement).

```{r fs-grf-debias-MR-table, eval = !is.null(fs_grf$sg.harm) && !is.null(fs_grf$mr_inference)}
mr_estimates_table(fs_grf, boot_results = fs_grf_bc, est.scale = "hr", digits = 2)
```

::: {.callout-note}
## Selection statistic vs. MR's de-biasing scale (GRF)

GRF ranks candidates by the **doubly-robust score** (RMST-based for survival),
while MR's de-biasing is computed on the perturbed **Cox** effect. The two scales
(additive RMST vs. multiplicative hazard) need not agree, so MR is a
*proxy* for GRF's selection -- the same approximation noted for DINA's tau-hat.
`family_native_neighborhood` restricts MR family to DR-score-competitive
candidates to keep that proxy close to GRF's actual selection neighbourhood; the
full bootstrap remains the reference because it re-runs the DR-score selection
itself.
:::



# Cross-Validation

Cross-validation assesses the **stability** of the ForestSearch algorithm.
Two approaches are available:

## K-Fold Cross-Validation

```{r cv-skipped-note, eval = !run_cv}
cat("K-Fold cross-validation skipped (run_cv = FALSE). Set run_cv <- TRUE",
    "in the parallel-setup chunk to run it.\n")
```

```{r kfold-cv, eval = run_cv}

t0 <- proc.time()

fs_kfold <- forestsearch_tenfold(
  # Pinned to the package default.  The fit carries mr_inference = run_mr
  # (TRUE here), and each replicate/fold call copies the fit's argument
  # list, so without this the flag is inherited and MR runs inside every
  # replicate.  Nothing downstream consumes a per-replicate MR object.
  # No-op as written.
  mr_in_replicates = FALSE,
  fs.est = fs,
  sims = Ksims,
  Kfolds = Kfolds_cv,
  details = FALSE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = FALSE
  )
)

plan("sequential")
timings$kfold <- (proc.time() - t0)["elapsed"]
metrics_tables <- cv_metrics_tables(fs_kfold)
metrics_tables
```

```{r cv-diag-heading, eval = run_cv}
#| echo: false
#| results: asis
cat("### Diagnostic Tabulations\n\n")
cat(paste(
  "`summarize_cv_results()` consolidates the per-fold record into a panel",
  "of diagnostic tables that together answer the core questions an",
  "analyst asks of a repeated K-fold run: which subgroups did the",
  "pipeline identify across resamples, which GRF cuts drove those",
  "identifications, when identification failed what was the proximate",
  "cause, and how confidently were the successful identifications made."), "\n\n")
```

```{r cv-diagnostics-run, eval = run_cv}
cv_diag <- summarize_cv_results(
  cv_output         = fs_kfold,
  original_sg       = fs$sg.harm,
  original_grf_cuts = fs$grf_cuts,
  top_n             = 10L
)
```

```{r cv-identification-note, eval = run_cv}
#| echo: false
#| results: asis
cat(sprintf(paste(
  "**Which subgroups emerged across resamples?**  The identification",
  "summary shows the distribution of selected subgroups across all",
  "%d x %d = %d sim x fold pairs.  A",
  "single dominant subgroup with high share indicates the pipeline",
  "converges on a stable selection; a long tail suggests selection is",
  "sensitive to the particular training split.\n\n"),
  Ksims, Kfolds_cv, Ksims * Kfolds_cv))
```

```{r cv-identification, eval = run_cv}
cv_diag$identification_summary
```

```{r cv-grf-cut-note, eval = run_cv}
#| echo: false
#| results: asis
cat(paste(
  "**Which GRF cuts drove those selections?**  The GRF cut summary",
  "tabulates the raw policy-tree output on each training fold,",
  "independent of whether the downstream consistency stage accepted",
  "it.  Comparison against the identification summary reveals whether",
  "GRF's upstream discovery is stable (a dominant cut shared across",
  "folds) or noisy (a long tail of one-off cuts)."), "\n\n")
```

```{r cv-grf-cut, eval = run_cv}
cv_diag$grf_cut_summary
```

```{r cv-no-subgroup-note, eval = run_cv}
#| echo: false
#| results: asis
cat(paste(
  "**When a fold failed to identify a subgroup, why?**  The no-subgroup",
  "decomposition splits the non-identifying folds into two actionable",
  "buckets: *GRF returned no cut* (nothing for consistency to evaluate,",
  "typically pointing at `dmin.grf` or `frac.tau` thresholds) versus",
  "*GRF returned a cut but consistency rejected it* (candidates existed",
  "but none met `pconsistency.threshold`).  The former argues for",
  "relaxing the GRF upstream; the latter for relaxing the consistency",
  "threshold or reviewing candidate quality."), "\n\n")
```

```{r cv-no-subgroup, eval = run_cv}
cv_diag$no_subgroup_decomposition
```

```{r cv-pcons-note, eval = run_cv}
#| echo: false
#| results: asis
cat(sprintf(paste(
  "**How confidently were the successful identifications made?**  Among",
  "folds that *did* identify a subgroup, the `pconsistency_distribution`",
  "shows where the selected candidate's consistency probability",
  "(`Pcons`) landed relative to the %s threshold.  Identifications in",
  "the `[0.85, 0.95)` bin are marginal; those concentrated at or near",
  "1.0 are robust.\n\n"),
  sprintf("%.2f", fs$args_call_all$pconsistency.threshold)))
```

```{r cv-pcons, eval = run_cv}
cv_diag$pconsistency_distribution
```

```{r cv-fold-numeric-note, eval = run_cv}
#| echo: false
#| results: asis
cat(paste(
  "**How do all the numeric fold-level diagnostics summarize together?**",
  "The `fold_numeric_summary` tabulates median / IQR / range for each",
  "numeric column in the per-fold record: fold size (`n_test`),",
  "consistency probability (`pconsistency`), training-fold HR of the",
  "selected subgroup (`training_fs_hr`, optimistically biased -- for",
  "diagnostic comparison only), and candidate count",
  "(`n_candidates_evaluated`)."), "\n\n")
```

```{r cv-fold-numeric, eval = run_cv}
cv_diag$fold_numeric_summary
```

```{r cv-original-agreement-note, eval = run_cv}
#| echo: false
#| results: asis
cat(paste(
  "**Agreement with the full-data analysis.**  How often does a training",
  "fold reproduce the original full-data subgroup exactly, share a",
  "factor with it, or match its GRF cut?"), "\n\n")
```

```{r cv-original-agreement, eval = run_cv}
cv_diag$original_agreement
```


## Out-of-Bag (N-Fold) Cross-Validation

```{r loo-note, eval = run_loo}
#| echo: false
#| results: asis
cat("N-fold CV (leave-one-out):\n\n")
```

```{r loo-skipped-note, eval = !run_loo}
cat("Leave-one-out (N-fold / OOB) skipped (run_loo = FALSE). Set run_loo <- TRUE",
    "in the parallel-setup chunk to run it.\n")
```

```{r oob-cv, eval = run_loo}

t0 <- proc.time()

fs_OOB <- forestsearch_Kfold(
  # Pinned to the package default.  The fit carries mr_inference = run_mr
  # (TRUE here), and each replicate/fold call copies the fit's argument
  # list, so without this the flag is inherited and MR runs inside every
  # replicate.  Nothing downstream consumes a per-replicate MR object.
  # No-op as written.
  mr_in_replicates = FALSE,
  fs.est = fs,
  details = FALSE,
  Kfolds = nrow(df.analysis),  # N-fold = leave-one-out
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$oob <- (proc.time() - t0)["elapsed"]

# Summarize OOB results
cv_out <- forestsearch_KfoldOut(
  res = fs_OOB,
  details = FALSE,
  outall = TRUE
)

tables <- cv_summary_tables(cv_out)

tables$combined_table

tables$metrics_table
```


# Results Visualization

## Forest Plot

The forest plot summarizes treatment effects across the ITT population,
reference subgroups, and identified subgroups with cross-validation metrics.

```{r forest-plot}
#| fig-width: 18
#| fig-height: 12
#| fig-cap: "Subgroup forest plot including identified subgroups"

# Define reference subgroups for comparison
subgroups <- list(
  age_gt65 = list(
    subset_expr = "age > 65",
    name = "Age > 65",
    type = "reference"
  ),
  age_le65 = list(
    subset_expr = "age <= 65",
    name = "Age <= 65",
    type = "reference"
  ),
  pgr_positive = list(
    subset_expr = "pgr > 0",
    name = "PgR > 0",
    type = "reference"
  ),
  pgr_negative = list(
    subset_expr = "pgr <= 0",
    name = "PgR <= 0",
    type = "reference"
  )
)


my_theme <- create_forest_theme(base_size = 24,
  footnote_fontsize = 17, cv_fontsize = 22)


# Create forest plot
# Include fs_kfold and fs_OOB if available for CV metrics
result <- plot_subgroup_results_forestplot(
  fs_results = list(
    fs.est = fs,
    fs_bc = fs_bc,
    fs_OOB   = if (exists("fs_OOB"))   fs_OOB   else NULL,
    fs_kfold = if (exists("fs_kfold")) fs_kfold else NULL
  ),
  df_analysis = df.analysis,
  subgroup_list = subgroups,
  outcome.name = outcome.name,
  event.name = event.name,
  treat.name = treat.name,
  E.name = "Hormonal",
  C.name = "Chemo",
  ci_column_spaces = 25,
  xlog = TRUE,
  theme = my_theme
)


# Render forest plot
render_forestplot(result)
```


### KM Difference Plots: ITT and Subgroups

The solid black line denotes the ITT Kaplan-Meier treatment difference
estimates along with 95% CIs (the grey shaded region). K-M differences
corresponding to subgroups are displayed.

```{r KMdiffs}
#| fig-width: 8
#| fig-height: 6

# Add additional subgroups along with ITT and identified subgroups
ref_sgs <- list(
  age_young = list(subset_expr = "age < 65", color = "brown"),
  age_old = list(subset_expr = "age >= 65", color = "orange")
)

plot_km_band_forestsearch(
  df = df.analysis,
  fs.est = fs,
  ref_subgroups = ref_sgs,
  outcome.name = outcome.name,
  event.name = event.name,
  treat.name = treat.name,
  draws_band = 20
)
```


# Summary and Interpretation

## Key Findings

```{r summary-findings}
cat(strrep("=", 60), "\n")
cat("FORESTSEARCH ANALYSIS SUMMARY\n")
cat(strrep("=", 60), "\n\n")

cat("Dataset: GBSG (N =", nrow(df.analysis), ")\n")
cat("Outcome: Recurrence-free survival\n\n")

cat("ITT Analysis:\n")
cat("  HR (95% CI): 0.69 (0.54, 0.89)\n\n")

cat("Identified Subgroup (H):\n")
cat("  Definition:", paste(fs$sg.harm, collapse = " & "), "\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 0)))
cat("  Unadjusted HR:",
    sprintf("%.2f", exp(fs$grp.consistency$out_sg$result$hr[1])), "\n")

cat("\nComplement Subgroup (Hc):\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 1),
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 1)))
```

## Clinical Interpretation

The ForestSearch analysis identifies **estrogen receptor-negative (ER ≤ 0)**
patients as a subgroup with potential lack of benefit from hormonal therapy.

**Biological plausibility**: Tamoxifen is a selective estrogen receptor
modulator. Its efficacy depends on ER expression. The finding that ER-negative
patients may not benefit is consistent with:

- Mechanistic understanding of tamoxifen action
- Meta-analyses showing no tamoxifen benefit in ER-negative breast cancer
- Clinical guidelines recommending tamoxifen primarily for ER-positive tumors

**Caveats**:

1. This is an **exploratory** analysis requiring independent validation
2. The bias-corrected estimates have wider confidence intervals
3. Cross-validation metrics should be evaluated for algorithm stability

## Computational Timing

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

# The rows are DRIVEN BY names(timings), not typed out.  A hand-written row list
# is how the (D) GRF fit and its bootstrap went missing from this table while
# still counting toward Total.  This map supplies readable labels and the
# display order; anything timed above but not labelled here STOPS the render
# rather than being silently dropped.
.timing_labels <- c(
  grf               = "GRF (preliminary)",
  forestsearch      = "ForestSearch (main fit)",
  bootstrap         = "FB bootstrap (main)",
  fs_dina_screen    = "FS (DINA screening)",
  dina_standalone   = "DINA (standalone)",
  fs_dina_select    = "FS (subgroup_method=dina)",
  fs_dina_bootstrap = "FB bootstrap (DINA)",
  fs_grf_select     = "FS (subgroup_method=grf)",
  fs_grf_bootstrap  = "FB bootstrap (GRF)",
  kfold             = "K-fold CV",
  oob               = "Leave-one-out CV",
  total             = "Total"
)

.unlabelled <- setdiff(names(timings), names(.timing_labels))
if (length(.unlabelled))
  stop("Computational Timing: no label for timings$",
       paste(.unlabelled, collapse = ", timings$"),
       " -- add it to .timing_labels in the timing-summary chunk.")

# Display order follows the map; Total last.  Skipped analyses (run_cv /
# run_loo FALSE) are pre-seeded NA and still get a row, so the reader sees
# that they were skipped rather than that they were forgotten.
.ord <- intersect(names(.timing_labels), names(timings))
.ord <- c(setdiff(.ord, "total"), intersect("total", .ord))

timing_df <- data.frame(
  Analysis  = unname(.timing_labels[.ord]),
  Seconds   = vapply(.ord, function(k) as.numeric(timings[[k]]), numeric(1)),
  row.names = NULL, stringsAsFactors = FALSE)
timing_df$Minutes <- timing_df$Seconds / 60

# Reconciliation: Total is the whole-document wall clock, so it legitimately
# exceeds the sum of the timed blocks (data prep, tables, figures, rendering are
# not timed).  Stating the residual makes an unexplained gap visible.
.parts_s <- sum(timing_df$Seconds[.ord != "total"], na.rm = TRUE)
.total_s <- as.numeric(timings$total)
.resid_s <- .total_s - .parts_s

gt(timing_df) |>
  tab_header(title = "Computational Timing") |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1) |>
  cols_label(
    Analysis = "Component",
    Seconds = "Time (sec)",
    Minutes = "Time (min)"
  ) |>
  tab_footnote(footnote = md(sprintf(
    paste("Rows are generated from `names(timings)`; a timed component with no",
          "label stops the render. Timed blocks account for %.1f s of the %.1f s",
          "total, leaving %.1f s (%.0f%%) in untimed work (data prep, tables,",
          "figures, document rendering). `NA` marks an analysis a toggle skipped."),
    .parts_s, .total_s, .resid_s, 100 * .resid_s / .total_s)))
```


# References

León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024).
"Exploratory subgroup identification in the heterogeneous Cox model:
A relatively simple procedure." *Statistics in Medicine*.
DOI: [10.1002/sim.10163](https://doi.org/10.1002/sim.10163)


# Parameter provenance (all fits) {#sec-param-appendix}

Formals whose resolved value differs from the package default, stacked across
the four fits. Values are read from each fit's `args_call_all`, so this table
cannot drift from the code that produced the results.

```{r param-provenance-appendix}
#| echo: false
.fits <- list("FS (main)" = fs, "(A) DINA-screened" = fs_dina_screen,
              "(C) DINA-selected" = fs_dina, "(D) GRF-selected" = fs_grf)
.pp <- do.call(rbind, lapply(names(.fits), function(nm) {
  d <- fs_param_provenance(.fits[[nm]])
  d <- d[d$status == "differs from default", c("parameter", "default", "value_used")]
  if (nrow(d)) cbind(fit = nm, d) else NULL
}))
gt::gt(.pp, groupname_col = "fit") |>
  gt::tab_header(
    title = "Parameters differing from package defaults, by fit",
    subtitle = sprintf("forestsearch %s, read from args_call_all at render time.",
                       tryCatch(as.character(utils::packageVersion("forestsearch")),
                                error = function(e) "version unknown"))) |>
  gt::cols_label(parameter = "Parameter", default = "Package default",
                 value_used = "Value used") |>
  gt::tab_style(
    style = gt::cell_text(font = gt::google_font("JetBrains Mono"), size = "x-small"),
    locations = gt::cells_body(columns = c(default, value_used))) |>
  gt::tab_options(table.font.size = gt::px(11), data_row.padding = gt::px(2))
```

# Session Information

```{r session-info}
sessionInfo()
```


# Reproducibility payload (Section 6.2) {#sec-export-payload-gbsg}

The chunk below harvests the already-computed objects from the three engine runs
into one small list and writes it as an `.rds` for the manuscript Section 6.2
reader document to reproduce the table and the results overview. No analysis is
repeated; the output path, core count, and package version resolve at run time.

```{r export-section6-2-payload}
#| echo: false
#| message: false

.out_dir <- if (is.null(results_dir)) {
  tryCatch(dirname(knitr::current_input(dir = TRUE)), error = function(e) getwd())
} else {
  results_dir
}
dir.create(.out_dir, recursive = TRUE, showWarnings = FALSE)

.get <- function(nm) if (exists(nm, inherits = TRUE)) get(nm, inherits = TRUE) else NULL

.triple <- function(x) {
  if (is.null(x) || is.null(x$est)) c(NA_real_, NA_real_, NA_real_) else c(x$est, x$lower, x$upper)
}
.fb_triple <- function(est_list) {
  if (is.null(est_list)) c(NA_real_, NA_real_, NA_real_) else c(est_list$H2, est_list$H2_lower, est_list$H2_upper)
}

.engine_rows <- function(method, fit, boot) {
  if (is.null(fit) || is.null(fit$mr_inference)) return(NULL)
  g  <- fit$mr_inference
  cc <- g$complement
  tr <- fit$df.est$treat.recommend
  n_tot <- length(tr); n_H <- sum(tr == 0, na.rm = TRUE)
  nv_H  <- .triple(g$naive)
  nv_Hc <- .triple(if (!is.null(cc)) cc$naive)
  mr_H  <- .triple(g$debiased)
  mr_Hc <- .triple(if (!is.null(cc)) cc$debiased)
  fb_H  <- .fb_triple(if (!is.null(boot)) boot$H_estimates)
  fb_Hc <- .fb_triple(if (!is.null(boot)) boot$Hc_estimates)
  data.frame(
    method = method, region = c("H", "Hc"),
    n = c(n_H, n_tot - n_H),
    pct = round(100 * c(n_H, n_tot - n_H) / n_tot),
    naive_est = c(nv_H[1], nv_Hc[1]), naive_lo = c(nv_H[2], nv_Hc[2]), naive_hi = c(nv_H[3], nv_Hc[3]),
    fb_est = c(fb_H[1], fb_Hc[1]), fb_lo = c(fb_H[2], fb_Hc[2]), fb_hi = c(fb_H[3], fb_Hc[3]),
    mr_est = c(mr_H[1], mr_Hc[1]), mr_lo = c(mr_H[2], mr_Hc[2]), mr_hi = c(mr_H[3], mr_Hc[3]),
    stringsAsFactors = FALSE)
}

.engine_meta <- function(fit, boot_seconds) {
  if (is.null(fit) || is.null(fit$mr_inference))
    return(list(label = NA_character_, n_family = NA_integer_, selection_bias = NA_real_,
                boot_sec = NA_real_, mr_sec = NA_real_, speedup = NA_real_))
  g <- fit$mr_inference
  mr_sec <- if (!is.null(g$timing_seconds)) as.numeric(g$timing_seconds) else NA_real_
  list(label = paste(fit$sg.harm, collapse = " & "),
       n_family = g$n_family, selection_bias = g$selection_bias,
       boot_sec = boot_seconds, mr_sec = mr_sec,
       speedup = if (is.finite(boot_seconds) && is.finite(mr_sec) && mr_sec > 0) boot_seconds / mr_sec else NA_real_)
}

.tm <- .get("timings")
.boot_sec <- function(key) {
  v <- if (!is.null(.tm) && !is.null(.tm[[key]])) suppressWarnings(as.numeric(.tm[[key]])) else NA_real_
  if (length(v) == 1L && is.finite(v)) v else NA_real_
}

# FS vs GRF agreement on the harm flag (treat.recommend == 0), aligned by id, with Cohen's kappa.
.concordance <- function(a, b, idn) {
  if (is.null(a) || is.null(b)) return(NULL)
  da <- a$df.est; db <- b$df.est
  if (is.null(da) || is.null(db) || !(idn %in% names(da)) || !(idn %in% names(db))) return(NULL)
  m <- merge(data.frame(.id = da[[idn]], ha = as.integer(da$treat.recommend == 0)),
             data.frame(.id = db[[idn]], hb = as.integer(db$treat.recommend == 0)), by = ".id")
  if (nrow(m) == 0L) return(NULL)
  po <- mean(m$ha == m$hb); pa <- mean(m$ha); pb <- mean(m$hb)
  pe <- pa * pb + (1 - pa) * (1 - pb)
  list(n_common = nrow(m), n_agree = sum(m$ha == m$hb), pct_agree = 100 * po,
       kappa = if (pe < 1) (po - pe) / (1 - pe) else NA_real_)
}

# Observed event rate by arm inside an engine's harm region (treat.recommend == 0).
# Cox analog of the binary version: the "rate" is the proportion experiencing the
# event (status == 1), reported alongside n and event counts so a reader can
# recompute incidence later if person-time is added.
.harm_rates <- function(fit, ename, tname) {
  if (is.null(fit) || is.na(ename) || is.na(tname)) return(NULL)
  d <- fit$df.est
  if (is.null(d) || !all(c(ename, tname, "treat.recommend") %in% names(d))) return(NULL)
  h <- d[d$treat.recommend == 0, , drop = FALSE]
  if (nrow(h) == 0L) return(NULL)
  et <- h[[ename]][h[[tname]] == 1]; ec <- h[[ename]][h[[tname]] == 0]
  list(treated_n      = length(et),
       control_n      = length(ec),
       treated_events = if (length(et)) sum(et == 1, na.rm = TRUE) else NA_integer_,
       control_events = if (length(ec)) sum(ec == 1, na.rm = TRUE) else NA_integer_,
       treated_pct    = if (length(et)) 100 * mean(et == 1, na.rm = TRUE) else NA_real_,
       control_pct    = if (length(ec)) 100 * mean(ec == 1, na.rm = TRUE) else NA_real_)
}

# Machine / hardware fingerprint of whatever computer ran this render, so that
# timings pooled across different machines stay attributable. Pure base R plus
# guarded platform calls (Linux / macOS / Windows); any field that cannot be
# determined returns NA rather than erroring.
.machine_info <- function(n_workers = NULL) {
  si   <- Sys.info()
  sn   <- unname(si[["sysname"]])               # "Linux" | "Darwin" | "Windows"
  sess <- tryCatch(sessionInfo(), error = function(e) NULL)

  .sys <- function(cmd) tryCatch(
    trimws(paste(system(cmd, intern = TRUE, ignore.stderr = TRUE), collapse = " ")),
    error = function(e) NA_character_)

  cpu_model <- tryCatch({
    if (sn == "Linux") {
      ln <- grep("^model name", readLines("/proc/cpuinfo", warn = FALSE), value = TRUE)[1]
      if (length(ln) && !is.na(ln)) trimws(sub("^model name\\s*:\\s*", "", ln)) else NA_character_
    } else if (sn == "Darwin") {
      .sys("sysctl -n machdep.cpu.brand_string")
    } else if (sn == "Windows") {
      out <- strsplit(.sys("wmic cpu get name"), "  +")[[1]]
      out <- out[nzchar(out) & out != "Name"]
      if (length(out)) out[length(out)] else NA_character_
    } else NA_character_
  }, error = function(e) NA_character_)

  ram_gb <- tryCatch({
    if (sn == "Linux") {
      kb <- as.numeric(sub("\\D*(\\d+).*", "\\1",
              grep("^MemTotal", readLines("/proc/meminfo", warn = FALSE), value = TRUE)[1]))
      round(kb / 1024^2, 1)
    } else if (sn == "Darwin") {
      round(as.numeric(.sys("sysctl -n hw.memsize")) / 1024^3, 1)
    } else if (sn == "Windows") {
      round(as.numeric(gsub("\\D", "", .sys("wmic ComputerSystem get TotalPhysicalMemory"))) / 1024^3, 1)
    } else NA_real_
  }, error = function(e) NA_real_)

  list(
    hostname           = unname(si[["nodename"]]),
    user               = unname(si[["user"]]),
    os                 = if (!is.null(sess) && !is.null(sess$running)) sess$running else NA_character_,
    sysname            = sn,
    os_release         = unname(si[["release"]]),
    arch               = unname(si[["machine"]]),
    cpu_model          = if (length(cpu_model) == 1L) cpu_model else NA_character_,
    cores_physical     = tryCatch(parallel::detectCores(logical = FALSE), error = function(e) NA_integer_),
    cores_logical      = tryCatch(parallel::detectCores(logical = TRUE),  error = function(e) NA_integer_),
    cores_used         = if (!is.null(n_workers)) as.integer(n_workers) else NA_integer_,
    workers_registered = tryCatch(future::nbrOfWorkers(), error = function(e) NA_integer_),
    future_plan        = tryCatch(class(future::plan())[1], error = function(e) NA_character_),
    ram_gb             = if (length(ram_gb) == 1L) ram_gb else NA_real_,
    r_version          = R.version.string,
    r_platform         = R.version$platform,
    blas               = if (!is.null(sess) && !is.null(sess$BLAS)) sess$BLAS else NA_character_,
    lapack             = if (!is.null(sess) && !is.null(sess$LAPACK)) sess$LAPACK else NA_character_,
    timestamp          = format(Sys.time(), tz = "UTC", usetz = TRUE)
  )
}

.fs      <- .get("fs");      .fs_bc      <- .get("fs_bc")
.fs_dina <- .get("fs_dina"); .fs_dina_bc <- .get("fs_dina_bc")
.fs_grf  <- .get("fs_grf");  .fs_grf_bc  <- .get("fs_grf_bc")

.meta_fs   <- .engine_meta(.fs,      .boot_sec("bootstrap"))
.meta_dina <- .engine_meta(.fs_dina, .boot_sec("fs_dina_bootstrap"))
.meta_grf  <- .engine_meta(.fs_grf,  .boot_sec("fs_grf_bootstrap"))

table1 <- do.call(rbind, list(
  .engine_rows("FS",   .fs,      .fs_bc),
  .engine_rows("DINA", .fs_dina, .fs_dina_bc),
  .engine_rows("GRF",  .fs_grf,  .fs_grf_bc)))

.dsel <- .get("dina_sel")
dina_standalone <- if (!is.null(.dsel) && isTRUE(.dsel$found)) {
  op <- if (identical(.dsel$direction, "left")) "<=" else ">="
  list(found = TRUE, label = sprintf("%s %s %s", .dsel$covariate, op, format(.dsel$threshold)), n = .dsel$n_subgroup)
} else {
  list(found = FALSE, label = NA_character_, n = NA_integer_)
}

.idname <- if (exists("id.name")) id.name else "id"
.oname  <- if (exists("outcome.name")) outcome.name else NA_character_
.ename  <- if (exists("event.name")) event.name else NA_character_
.tname  <- if (exists("treat.name")) treat.name else NA_character_

payload <- list(
  table  = table1,
  labels = list(fs = .meta_fs$label, dina = .meta_dina$label, grf = .meta_grf$label,
                dina_standalone = dina_standalone),
  meta   = list(
    B          = if (exists("NB")) NB else NA_integer_,
    mr_draws = if (exists("mr_draws")) mr_draws else NA_integer_,
    n_total    = if (!is.null(.fs)) nrow(.fs$df.est) else NA_integer_,
    cores      = tryCatch(unname(future::availableCores()), error = function(e) NA_integer_),
    machine    = .machine_info(n_workers = if (exists("n_cores")) n_cores else NULL),
    outcome    = .oname,
    event      = .ename,
    covariates = if (exists("confounders.name")) confounders.name else NA_character_,
    timings    = list(fs = .meta_fs, dina = .meta_dina, grf = .meta_grf)),
  extras = list(
    concordance = .concordance(.fs, .fs_grf, .idname),
    harm_rates  = .harm_rates(.fs, .ename, .tname)),
  est_scale = "hr",
  built_at  = Sys.time(),
  forestsearch_version = tryCatch(as.character(utils::packageVersion("forestsearch")), error = function(e) NA_character_))

# Payload filename -- analog of actg175's "actg175_table_payload.rds".
# (If your gbsg reader/manuscript doc expects "gbsg_table1_payload.rds", change it here.)
.payload_file <- file.path(.out_dir, "gbsg_survival_multimethod_payload.rds")
saveRDS(payload, .payload_file)
cat("Wrote Section 6.2 payload to:", .payload_file, "\n")
```
