---
title: "Threshold Calibration for Binary Outcomes via the GLM Approximation"
subtitle: "Extending the Section 2.1 Framework to Logistic Regression"
author: "Larry F. Leon"
date: today
format:
  html:
    fig-width: 10
    fig-height: 7
execute:
  warning: true
  message: true
  fig-align: center
knitr:
  opts_chunk:
    comment: "#>"
---

```{r setup}
#| include: false

# ── Simulation control ──────────────────────────────────────────
# ~ 600/hr mac
nsims       <- 2000L       # sims per (c1, c2, N, hypothesis)
nsims_prod  <- 5000L     # production target for projection (may differ from nsims)
sim_seed    <- 8316951L
n_workers   <- ceiling(0.95 * max(1L,
  parallel::detectCores(logical = FALSE) - 1L))
run_h1      <- FALSE     # FALSE = skip H1 sims, use approximation

# ── Output ──────────────────────────────────────────────────────
output_dir  <- "../_data"

# ── FPR approximation ──────────────────────────────────────────
n_min_val   <- 60L       # n.min = reference subgroup size for P1
fpr_target  <- 0.12      # target procedure-level FPR (must match S1)

# ── DGM / outcome parameters ──────────────────────────────────
p_event     <- 0.30      # event probability in the binary DGM
p_harm      <- 0.30      # harm subgroup prevalence (metadata only)
theta_h1    <- 2.0       # OR for harm subgroup under H1
confounders_name <- c("bm1", "bm2", "age", "ecog")

# ── ForestSearch parameters ───────────────────────────────────
pcons       <- 0.90      # pconsistency.threshold
fs_splits   <- 400L
maxk_val    <- 2L

# ── Sample sizes (shared with S1) ────────────────────────────
N_cal_values <- c(300, 500, 700)

# ── Threshold grid for simulation validation ──────────────────
# IMPORTANT: the grid must lie in the *operational* threshold range —
# not at low values where the procedure is saturated.
#
# For binary outcomes with d_eff = n_H * p * (1-p) = 12.6
# (n_H=60, p=0.30), the per-subgroup SD of log(OR) ≈ 0.56.  Below
# c1 ≈ 1.5, virtually every candidate subgroup exceeds c1 by chance,
# so the simulated FPR is flat (insensitive to threshold).  Calibrating
# L_eff in that saturated regime inflates the estimate and causes the
# corrected approximation to overestimate FPR at the higher thresholds
# used in selection.
#
# Rule of thumb: set min(c1_grid) so that P1 at n_H = n_min_val is
# no greater than ~0.15.  For this DGM (p_event=0.30, n_H=60):
#   c1 = 1.50 → P1 ≈ 0.16  (borderline)
#   c1 = 1.60 → P1 ≈ 0.12  (acceptable floor)
# The range should match what S1 will sweep for selection.
c1_grid     <- round(seq(1.50, 2.50, length.out = 10), 2)
c2_grid     <- round(seq(1.00, 2.00, length.out = 10), 2)

```


# The GLM Approximation {#sec-approx}

## Parallel to Section 2.1

For the Cox model, the Section 2.1 approximation uses the
asymptotic normality of the score statistic:
$\hat\beta_s \approx (4/d) \tilde L_d(0) \sim N(\beta, 4/d)$.
For split-halves: $\hat\beta^1_s, \hat\beta^2_s \sim N(\beta, 8/d)$
independently.

For a canonical GLM with treatment as the sole covariate, the
MLE satisfies exactly the same structure via Fisher information.
With balanced arms ($n_1 = n_0 = n_H/2$):

$$
\text{Var}(\hat\beta_{\text{full}}) = \frac{4}{d_{\text{eff}}},
\qquad
\text{Var}(\hat\beta_{\text{half}}) = \frac{8}{d_{\text{eff}}}
$$

where $d_{\text{eff}}$ is the effective information:

| Family | $d_{\text{eff}}$ | Analogous to |
|---|---|---|
| Cox | $d = n_H(1-p_c)$ | Number of events |
| Binomial (logit) | $n_H \bar p(1-\bar p)$ | Bernoulli information |
| Poisson (log) | $D$ (total events) | Number of events |
| Gaussian (identity) | $n_H / \sigma^2$ | Precision-weighted n |

The detection probability for a **specific subgroup** with
true effect $\beta = \log(\theta^\dagger(H))$ is:

$$
p(c_1, c_2; d_{\text{eff}}, \theta^\dagger(H)) =
P(W_1 + W_2 \ge 2\log c_1, \;\min(W_1, W_2) \ge \log c_2)
$$

with $W_1, W_2 \stackrel{\text{iid}}{\sim}
N(\log \theta^\dagger(H), \; 8/d_{\text{eff}})$.

This gives:

- **Type-I error** (per-subgroup): $p(c_1, c_2; d_{\text{eff}}, 1)$
  with $\beta = 0$
- **Power**: $p(c_1, c_2; d_{\text{eff}}, \theta^\dagger(H))$
  for the true harm subgroup

The thresholds are then chosen so that
$p(c_1, c_2; d_{\text{eff}}, 1) \approx 10\%$ and
$p(c_1, c_2; d_{\text{eff}}, \theta^\dagger(H))$ is
acceptably high, and **validated by simulation**.


## Implementation

```{r approx-function}
#| code-fold: false

# Use compute_detection_probability_glm() and d_eff_*() helpers
# from forestsearch (installed in R/truefind_asymptotic_glm.R).
# These implement equation (3) extended to GLM via d_eff.
#
# Shorthand wrapper for concise calls in this document:
p_detect <- function(theta, d_eff, c1, c2) {
  compute_detection_probability_glm(
    theta = theta, d_eff = d_eff, c1 = c1, c2 = c2,
    effect_scale = "ratio", method = "monte_carlo"
  )
}
```


## Pipeline Overview {#sec-pipeline}

The figure below shows the complete two-document calibration workflow.
This document (C1) fits the $L_{\text{eff}}$ model from simulation;
`selection_binary_frontier.qmd` (S1) consumes it to sweep the threshold
space and validate selected pairs. The amber nodes are the two checkpoints
that guard against the saturated-regime failure described in
@sec-threshold-indep.

```{=html}
<figure style="margin:1.5rem 0 2rem;">
<svg id="cal-pipeline" width="100%" viewBox="0 0 680 940" role="img" xmlns="http://www.w3.org/2000/svg">
<title>Binary outcome calibration pipeline: C1 to S1</title>
<desc>Flowchart of the two-document L_eff calibration pipeline. C1 runs H0 simulations, checks threshold-independence with a CV diagnostic, fits the power law, and saves to .rds. S1 loads the calibration, verifies the grid is above the saturation floor, computes the corrected FPR on a dense grid, selects and validates frontier pairs, and reports the optimal threshold pair.</desc>
<defs>
<style>
#cal-pipeline .th{font-family:ui-sans-serif,system-ui,sans-serif;font-size:14px;font-weight:500}
#cal-pipeline .ts{font-family:ui-sans-serif,system-ui,sans-serif;font-size:12px;font-weight:400}
#cal-pipeline .arr{stroke:#888780;stroke-width:1.5;fill:none}
#cal-pipeline .leader{stroke:#B4B2A9;stroke-width:0.5;stroke-dasharray:4 3;fill:none}
#cal-pipeline .node{cursor:default}
#cal-pipeline .c-teal>rect{fill:#E1F5EE;stroke:#0F6E56;stroke-width:0.5}
#cal-pipeline .c-teal .th{fill:#085041}
#cal-pipeline .c-teal .ts{fill:#0F6E56}
#cal-pipeline .c-purple>rect{fill:#EEEDFE;stroke:#534AB7;stroke-width:0.5}
#cal-pipeline .c-purple .th{fill:#3C3489}
#cal-pipeline .c-purple .ts{fill:#534AB7}
#cal-pipeline .c-amber>rect{fill:#FAEEDA;stroke:#BA7517;stroke-width:0.5}
#cal-pipeline .c-amber .th{fill:#633806}
#cal-pipeline .c-amber .ts{fill:#854F0B}
#cal-pipeline .c-gray>rect{fill:#F1EFE8;stroke:#888780;stroke-width:0.5}
#cal-pipeline .c-gray .th{fill:#444441}
#cal-pipeline .c-gray .ts{fill:#5F5E5A}
#cal-pipeline .c-green>rect{fill:#EAF3DE;stroke:#3B6D11;stroke-width:0.5}
#cal-pipeline .c-green .th{fill:#27500A}
#cal-pipeline .c-green .ts{fill:#3B6D11}
</style>
<marker id="cp-arrow" viewBox="0 0 10 10" refX="8" refY="5" markerWidth="6" markerHeight="6" orient="auto-start-reverse">
<path d="M2 1L8 5L2 9" fill="none" stroke="#888780" stroke-width="1.5" stroke-linecap="round" stroke-linejoin="round"/>
</marker>
</defs>
<!-- Arrows -->
<line x1="340" y1="106" x2="340" y2="122" class="arr" marker-end="url(#cp-arrow)"/>
<line x1="340" y1="172" x2="340" y2="188" class="arr" marker-end="url(#cp-arrow)"/>
<line x1="340" y1="238" x2="340" y2="254" class="arr" marker-end="url(#cp-arrow)"/>
<line x1="340" y1="322" x2="340" y2="338" class="arr" marker-end="url(#cp-arrow)"/>
<line x1="340" y1="388" x2="340" y2="404" class="arr" marker-end="url(#cp-arrow)"/>
<line x1="340" y1="444" x2="340" y2="508" class="arr" marker-end="url(#cp-arrow)"/>
<line x1="340" y1="558" x2="340" y2="574" class="arr" marker-end="url(#cp-arrow)"/>
<line x1="340" y1="638" x2="340" y2="654" class="arr" marker-end="url(#cp-arrow)"/>
<line x1="340" y1="704" x2="340" y2="720" class="arr" marker-end="url(#cp-arrow)"/>
<line x1="340" y1="770" x2="340" y2="786" class="arr" marker-end="url(#cp-arrow)"/>
<line x1="340" y1="836" x2="340" y2="852" class="arr" marker-end="url(#cp-arrow)"/>
<!-- Separator -->
<line x1="80" y1="453" x2="600" y2="453" stroke="#B4B2A9" stroke-width="0.8" stroke-dasharray="5 3"/>
<!-- Section headers -->
<g class="c-teal">
<rect x="60" y="14" width="560" height="28" rx="6"/>
<text class="th" x="340" y="28" text-anchor="middle" dominant-baseline="central">C1 — calibration_binary_leff.qmd</text>
</g>
<g class="c-purple">
<rect x="60" y="464" width="560" height="28" rx="6"/>
<text class="th" x="340" y="478" text-anchor="middle" dominant-baseline="central">S1 — selection_binary_frontier.qmd</text>
</g>
<!-- C1 nodes -->
<g class="node c-teal">
<rect x="140" y="56" width="400" height="50" rx="8"/>
<text class="th" x="340" y="74" text-anchor="middle" dominant-baseline="central">Define calibration grid</text>
<text class="ts" x="340" y="91" text-anchor="middle" dominant-baseline="central">c1_grid ≥ 1.5 for p = 0.30  ·  P1 ≲ 0.15 at n_min</text>
</g>
<g class="node c-teal">
<rect x="140" y="122" width="400" height="50" rx="8"/>
<text class="th" x="340" y="140" text-anchor="middle" dominant-baseline="central">Run H0 simulations</text>
<text class="ts" x="340" y="157" text-anchor="middle" dominant-baseline="central">nsims per (c1, c2, N) pair  ·  H0 only</text>
</g>
<g class="node c-teal">
<rect x="140" y="188" width="400" height="50" rx="8"/>
<text class="th" x="340" y="206" text-anchor="middle" dominant-baseline="central">Compute P1 and implied L_eff</text>
<text class="ts" x="340" y="223" text-anchor="middle" dominant-baseline="central">L_eff = log(1 − FPR) / log(1 − P1)  per (c1, c2, N)</text>
</g>
<g class="node c-amber">
<rect x="140" y="254" width="400" height="68" rx="8"/>
<text class="th" x="340" y="272" text-anchor="middle" dominant-baseline="central">CV diagnostic — threshold-independence check</text>
<text class="ts" x="340" y="290" text-anchor="middle" dominant-baseline="central">CV &lt; 20%: proceed  ·  CV 20–30%: borderline</text>
<text class="ts" x="340" y="308" text-anchor="middle" dominant-baseline="central">CV &gt; 30%: raise min(c1_grid) and re-run from step 1</text>
</g>
<g class="node c-teal">
<rect x="140" y="338" width="400" height="50" rx="8"/>
<text class="th" x="340" y="356" text-anchor="middle" dominant-baseline="central">Filter saturated rows · fit power law</text>
<text class="ts" x="340" y="373" text-anchor="middle" dominant-baseline="central">L_eff(N) = C · (N/n_min)^α  ·  exclude FPR_sim ≥ 0.50</text>
</g>
<g class="node c-gray">
<rect x="190" y="404" width="300" height="40" rx="8"/>
<text class="th" x="340" y="419" text-anchor="middle" dominant-baseline="central">Save calibration to .rds</text>
<text class="ts" x="340" y="435" text-anchor="middle" dominant-baseline="central">C, α, saturation_floor_c1, c1_range</text>
</g>
<!-- S1 nodes -->
<g class="node c-purple">
<rect x="140" y="508" width="400" height="50" rx="8"/>
<text class="th" x="340" y="526" text-anchor="middle" dominant-baseline="central">Load .rds · check saturation floor</text>
<text class="ts" x="340" y="543" text-anchor="middle" dominant-baseline="central">C, α, saturation_floor_c1  ·  warn if grid below floor</text>
</g>
<g class="node c-amber">
<rect x="140" y="574" width="400" height="64" rx="8"/>
<text class="th" x="340" y="591" text-anchor="middle" dominant-baseline="central">Saturation floor check</text>
<text class="ts" x="340" y="609" text-anchor="middle" dominant-baseline="central">Grid must stay above saturation_floor_c1</text>
<text class="ts" x="340" y="626" text-anchor="middle" dominant-baseline="central">FPR_corr overestimates if c1_grid dips below floor</text>
</g>
<g class="node c-purple">
<rect x="140" y="654" width="400" height="50" rx="8"/>
<text class="th" x="340" y="672" text-anchor="middle" dominant-baseline="central">Compute FPR_corr on dense grid</text>
<text class="ts" x="340" y="689" text-anchor="middle" dominant-baseline="central">FPR_corr = 1 − (1 − P1)^(C · (N/n_min)^α)</text>
</g>
<g class="node c-purple">
<rect x="140" y="720" width="400" height="50" rx="8"/>
<text class="th" x="340" y="738" text-anchor="middle" dominant-baseline="central">Select frontier pairs</text>
<text class="ts" x="340" y="755" text-anchor="middle" dominant-baseline="central">FPR_corr ≤ target at worst-case N  ·  binding at N_max</text>
</g>
<g class="node c-purple">
<rect x="140" y="786" width="400" height="50" rx="8"/>
<text class="th" x="340" y="804" text-anchor="middle" dominant-baseline="central">Validate with H0 simulations</text>
<text class="ts" x="340" y="821" text-anchor="middle" dominant-baseline="central">Sim_FPR vs FPR_corr  ·  check In CI (95%)</text>
</g>
<g class="node c-green">
<rect x="140" y="852" width="400" height="50" rx="8"/>
<text class="th" x="340" y="870" text-anchor="middle" dominant-baseline="central">Optimal pair + power curve</text>
<text class="ts" x="340" y="887" text-anchor="middle" dominant-baseline="central">c1_opt, c2_opt  ·  H1 power assessment at theta_h1</text>
</g>
<!-- Side notes -->
<g class="c-gray">
<rect x="548" y="262" width="112" height="48" rx="4"/>
<text class="ts" x="604" y="280" text-anchor="middle" dominant-baseline="central">No (CV &gt; 30%):</text>
<text class="ts" x="604" y="297" text-anchor="middle" dominant-baseline="central">↑ raise c1_grid</text>
</g>
<line x1="540" y1="283" x2="548" y2="283" class="leader"/>
<g class="c-gray">
<rect x="548" y="580" width="112" height="44" rx="4"/>
<text class="ts" x="604" y="598" text-anchor="middle" dominant-baseline="central">No (grid too low):</text>
<text class="ts" x="604" y="614" text-anchor="middle" dominant-baseline="central">warning emitted</text>
</g>
<line x1="540" y1="606" x2="548" y2="606" class="leader"/>
<!-- nsims annotations -->
<g class="c-gray">
<rect x="18" y="126" width="108" height="44" rx="4"/>
<text class="ts" x="72" y="144" text-anchor="middle" dominant-baseline="central">Production:</text>
<text class="ts" x="72" y="160" text-anchor="middle" dominant-baseline="central">2,000–5,000</text>
</g>
<line x1="126" y1="148" x2="140" y2="148" class="leader"/>
<g class="c-gray">
<rect x="18" y="790" width="108" height="44" rx="4"/>
<text class="ts" x="72" y="808" text-anchor="middle" dominant-baseline="central">Production:</text>
<text class="ts" x="72" y="824" text-anchor="middle" dominant-baseline="central">2,000–5,000</text>
</g>
<line x1="126" y1="812" x2="140" y2="812" class="leader"/>
<!-- Legend -->
<g class="c-teal"><rect x="98" y="913" width="12" height="12" rx="2"/></g>
<text class="ts" x="116" y="919" dominant-baseline="central" style="fill:#444441">C1 step</text>
<g class="c-purple"><rect x="186" y="913" width="12" height="12" rx="2"/></g>
<text class="ts" x="204" y="919" dominant-baseline="central" style="fill:#444441">S1 step</text>
<g class="c-amber"><rect x="274" y="913" width="12" height="12" rx="2"/></g>
<text class="ts" x="292" y="919" dominant-baseline="central" style="fill:#444441">Decision / check</text>
<g class="c-gray"><rect x="414" y="913" width="12" height="12" rx="2"/></g>
<text class="ts" x="432" y="919" dominant-baseline="central" style="fill:#444441">Data / note</text>
<g class="c-green"><rect x="510" y="913" width="12" height="12" rx="2"/></g>
<text class="ts" x="528" y="919" dominant-baseline="central" style="fill:#444441">Output</text>
</svg>
<figcaption style="font-size:0.82em;color:#5F5E5A;margin-top:0.4rem;text-align:center;">
Figure: Two-document calibration pipeline.
The long arrow crossing the dashed line represents the <code>.rds</code> handoff.
Left annotations show recommended production simulation budgets.
</figcaption>
</figure>
```


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)

# -- Safe foreach result collection -------------------------------------------
# Replaces .errorhandling = "remove" (which silently drops failures)
# with .errorhandling = "pass" + post-hoc separation of errors from results.
collect_results <- function(raw_list, label = "") {
  is_err <- vapply(raw_list, inherits, logical(1), "error")
  n_failed <- sum(is_err)
  if (n_failed > 0) {
    err_msgs <- vapply(raw_list[is_err], conditionMessage, character(1))
    warning(sprintf("%s%d of %d tasks failed.",
        if (nzchar(label)) paste0(label, ": ") else "",
        n_failed, length(raw_list)))
    message("  First errors:\n    ",
        paste(head(unique(err_msgs), 3), collapse = "\n    "))
  }
  good <- raw_list[!is_err]
  if (length(good) == 0) {
    result <- NULL
  } else if (requireNamespace("data.table", quietly = TRUE)) {
    result <- data.table::rbindlist(good, use.names = TRUE, fill = TRUE)
    result <- as.data.frame(result)
  } else {
    result <- do.call(rbind, good)
  }
  attr(result, "n_failed") <- n_failed
  attr(result, "n_total") <- length(raw_list)
  result
}


# L_eff save/load helpers (exported from forestsearch >= 0.2.0)
plan(multisession, workers = n_workers)
cat("Workers:", n_workers, "| Sims per config:", nsims, "\n")
```

```{r timing-init}
#| include: false
timings <- list()
t_vignette_start <- proc.time()
```


# Part I: Approximation-Guided Threshold Selection {#sec-guide}

## Survival Benchmark

Reproduce the survival result: with $d = 33$ (n = 60, 45
percent censoring), the defaults c1 = 1.25, c2 = 1.0 give
$p \approx 10\%$ type-I error and $\approx 80\%$ power at
$\theta^\dagger(H) = 1.94$.

```{r survival-benchmark}
#| tbl-cap: "Survival benchmark: Section 2.1 reproduced"

surv_bench <- data.frame(
  n_sg = c(60, 80, 100),
  stringsAsFactors = FALSE
)
surv_bench$d <- round(d_eff_survival(surv_bench$n_sg, prop_cens = 0.45))
surv_bench$`P(theta=1.0)` <- sapply(surv_bench$d, function(d)
  round(p_detect(1.0, d, 1.25, 1.0), 3))
surv_bench$`P(theta=0.75)` <- sapply(surv_bench$d, function(d)
  round(p_detect(0.75, d, 1.25, 1.0), 3))
surv_bench$`P(theta=2.0)` <- sapply(surv_bench$d, function(d)
  round(p_detect(theta_h1, d, 1.25, 1.0), 3))

gt(surv_bench) |>
  tab_header(
    title = "Survival: c1=1.25, c2=1.0 (Paper Defaults)",
    subtitle = "Reproduces Figure 2 and Table 1 footnotes"
  ) |>
  cols_label(n_sg = "n(H)", d = "d (events)",
    `P(theta=1.0)` = "P(OR=1.0)",
    `P(theta=0.75)` = "P(OR=0.75)",
    `P(theta=2.0)` = "Power(OR=2)")
```


## Binary: d_eff Is Lower, So the Thresholds Need Adjustment

For binary outcomes with $\bar p = 0.30$:
$d_{\text{eff}} = n_H \cdot 0.30 \cdot 0.70 = 0.21 n_H$.
For $n_H = 60$: $d_{\text{eff}} = 12.6$ (vs $d = 33$ for survival).

```{r binary-default-check}
#| tbl-cap: "Binary: survival defaults applied to binary d_eff"

bin_check <- data.frame(
  n_sg = c(60, 100, 150, 200, 240),
  p = p_event,
  stringsAsFactors = FALSE
)
bin_check$d_eff <- round(d_eff_binary(bin_check$n_sg, p_event = p_event), 1)

# Survival defaults: c1=1.25, c2=1.0
bin_check$P_10 <- sapply(bin_check$d_eff, function(d)
  round(p_detect(1.0, d, 1.25, 1.0), 3))
bin_check$P_075 <- sapply(bin_check$d_eff, function(d)
  round(p_detect(0.75, d, 1.25, 1.0), 3))
bin_check$Power_20 <- sapply(bin_check$d_eff, function(d)
  round(p_detect(theta_h1, d, 1.25, 1.0), 3))

gt(bin_check) |>
  tab_header(
    title = "Binary with Survival Defaults (c1=1.25, c2=1.0)",
    subtitle = "Per-subgroup P(theta=1.0) is the type-I error analog"
  ) |>
  cols_label(n_sg = "n(H)", p = "p", d_eff = "d_eff",
    P_10 = "P(OR=1.0)", P_075 = "P(OR=0.75)",
    Power_20 = "Power(OR=2)")
```


## Per-Subgroup P1 at Key Thresholds

A brief check of $P_1$ at `n.min` for the grid endpoints
and the paper default.  Full threshold selection (frontier,
power curves, validation) is performed in
`selection_binary_frontier.qmd` (S1).

```{r approx-summary}
#| tbl-cap: !expr 'sprintf("Per-subgroup P1 at n.min = %d (p = %.2f, d_eff = %.0f)", n_min_val, p_event, d_eff_binary(n_min_val, p_event))'

d_ref <- round(d_eff_binary(n_min_val, p_event))

# Representative pairs: paper default + grid corners
check_pairs <- data.frame(
  c1 = c(1.25, min(c1_grid), max(c1_grid)),
  c2 = c(1.00, min(c2_grid), max(c2_grid)),
  Label = c("Paper default", "Grid min", "Grid max"),
  stringsAsFactors = FALSE
)
check_pairs <- check_pairs[check_pairs$c2 <= check_pairs$c1, ]

check_pairs$`P(OR=1)` <- sapply(seq_len(nrow(check_pairs)), function(i)
  round(p_detect(1.0, d_ref, check_pairs$c1[i], check_pairs$c2[i]), 3))
check_pairs[[sprintf("Power(OR=%.1f)", theta_h1)]] <-
  sapply(seq_len(nrow(check_pairs)), function(i)
    round(p_detect(theta_h1, d_ref, check_pairs$c1[i], check_pairs$c2[i]), 3))

gt(check_pairs) |>
  tab_header(
    title = "Per-Subgroup Detection at n.min",
    subtitle = sprintf("d_eff = %d | See S1 for full threshold selection",
                        d_ref)
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `P(OR=1)`,
      rows = `P(OR=1)` >= 0.03 & `P(OR=1)` <= fpr_target))
```


# Part II: Simulation Validation {#sec-sim}

## Binary DGM

```{r sim-dgm}
#| code-fold: false

sim_binary <- function(n, seed, has_hte = TRUE) {
  set.seed(seed)
  df <- data.frame(
    id = seq_len(n), treat = rbinom(n, 1, 0.5),
    bm1 = as.factor(rbinom(n, 1, 0.70)),
    bm2 = as.factor(rbinom(n, 1, 0.50)),
    age = round(rnorm(n, 55, 10)),
    ecog = as.factor(sample(0:1, n, replace = TRUE, prob = c(0.6, 0.4)))
  )
  # Baseline event rate from setup: p_event
  lp <- with(df, qlogis(p_event) + 0.02 * (age - 55) + 0.3 * (ecog == 1))
  if (has_hte) {
    lp <- lp + with(df,
      log(theta_h1) * treat * (bm1 == 0) +
      log(1 / theta_h1) * treat * (bm1 == 1))
  }
  df$progressed <- rbinom(n, 1, plogis(lp))
  df$flag_harm  <- as.integer(df$bm1 == 0)
  df
}
```


## Simulation Engine

```{r sim-engine}
#| code-fold: false

run_sim <- function(sim_num, n_total, has_hte, c1, c2, pcons) {
  seed <- sim_seed + sim_num
  df <- sim_binary(n_total, seed, has_hte)
  fs_error <- NA_character_
  fs <- tryCatch(
    suppressWarnings(
      forestsearch(
        df.analysis = df,
        confounders.name = confounders_name,
        outcome.name = "progressed", event.name = "progressed",
        treat.name = "treat", id.name = "id",
        outcome_type = "binary", effect_measure = "OR",
        adverse_outcome = TRUE,
        hr.threshold = c1, hr.consistency = c2,
        pconsistency.threshold = pcons, fs.splits = fs_splits,
        n.min = n_min_val, maxk = maxk_val,
        use_lasso = TRUE, use_grf = TRUE, use_twostage = TRUE,
        is.RCT = TRUE, seedit = seed, details = FALSE,
        quiet = TRUE, plot.sg = FALSE,
        parallel_args = list(plan = "sequential", workers = 1,
                             show_message = FALSE)
      )
    ),
    error = function(e) { fs_error <<- conditionMessage(e); NULL }
  )
  found <- !is.null(fs) && !is.null(fs$sg.harm) && length(fs$sg.harm) > 0
  out <- data.table(sim = sim_num, n_total = n_total, has_hte = has_hte,
                    c1 = c1, c2 = c2, pcons = pcons,
                    any_found = as.integer(found), error_msg = fs_error)
  if (found && !is.null(fs$df.est) && "treat.recommend" %in% names(fs$df.est)) {
    sg_ind <- fs$df.est$treat.recommend; flag <- df$flag_harm
    tp <- sum(sg_ind == 0 & flag == 1); fp <- sum(sg_ind == 0 & flag == 0)
    tn <- sum(sg_ind == 1 & flag == 0); fn <- sum(sg_ind == 1 & flag == 1)
    out$sens <- if ((tp+fn)>0) tp/(tp+fn) else NA_real_
    out$spec <- if ((tn+fp)>0) tn/(tn+fp) else NA_real_
    out$ppv  <- if ((tp+fp)>0) tp/(tp+fp) else NA_real_
    out$size_H <- sum(sg_ind == 0)
  } else { out$sens <- out$spec <- out$ppv <- NA_real_; out$size_H <- NA_integer_ }
  out
}
```


## Run Simulations

Validate the approximation-selected threshold pairs.
Fix pconsistency = 0.90.  The threshold grid is defined
by `c1_grid` and `c2_grid` in the setup chunk.

::: {.callout-tip}
For initial runs, use a coarser grid (`by = 0.10`) to
identify the viable region quickly, then refine with
`by = 0.05` or `by = 0.02` around the frontier.
:::

```{r run-sims}
t0 <- proc.time()[3]

# Build threshold pairs from the grid defined in the setup chunk
pair_grid <- expand.grid(c1 = c1_grid, c2 = c2_grid,
                          stringsAsFactors = FALSE)
pair_grid <- pair_grid[pair_grid$c2 <= pair_grid$c1, ]
sim_pairs <- lapply(seq_len(nrow(pair_grid)), function(i)
  list(c1 = pair_grid$c1[i], c2 = pair_grid$c2[i]))

cat(sprintf("Threshold grid: %d c1 x %d c2 -> %d valid pairs (c2 <= c1)\n",
    length(c1_grid), length(c2_grid), length(sim_pairs)))

# Build task grid: H0 always; H1 only if run_h1 = TRUE
hte_levels <- if (run_h1) c(FALSE, TRUE) else FALSE
tasks <- do.call(rbind, lapply(sim_pairs, function(pr) {
  expand.grid(sim = seq_len(nsims), c1 = pr$c1, c2 = pr$c2,
              n_total = N_cal_values, has_hte = hte_levels,
              stringsAsFactors = FALSE)
}))

n_hyp <- if (run_h1) 2L else 1L
cat(sprintf("Simulation: %d pairs x %d N x %d hypothesis x %d sims = %d tasks\n",
    length(sim_pairs), length(N_cal_values), n_hyp, nsims, nrow(tasks)))
if (!run_h1) cat("(H1 skipped: power computed from approximation)\n")

res <- foreach(i = seq_len(nrow(tasks)), .errorhandling = "pass",
  .options.future = list(packages = c("forestsearch", "data.table"),
                         seed = TRUE)
) %dofuture% {
  future::plan("sequential")
  run_sim(tasks$sim[i], tasks$n_total[i], tasks$has_hte[i],
          tasks$c1[i], tasks$c2[i], pcons)
}
timings$simulation <- proc.time()[3] - t0

# Free parallel workers before sequential post-processing
future::plan("sequential"); gc(verbose = FALSE)

sim_raw <- forestsearch:::collect_results(res, "grid sims")
sim_dt  <- data.table::as.data.table(sim_raw)

# ── Post-hoc task-count check ──────────────────────────────────
n_expected <- nrow(tasks)
n_received <- nrow(sim_dt)
if (n_received < n_expected) {
  warning(sprintf(
    "%d of %d tasks silently failed (%.1f%% loss). FPR/power estimates may be noisy.",
    n_expected - n_received, n_expected,
    100 * (1 - n_received / n_expected)
  ))
}
cat(sprintf("Simulation: %.1f min | %d/%d results | %d internal errors\n",
    timings$simulation / 60, n_received, n_expected,
    sum(!is.na(sim_dt$error_msg))))

# ── FSdata failure diagnostic ────────────────────────────────────
n_fsdata_err <- sum(!is.na(sim_dt$error_msg))
if (n_fsdata_err > 0) {
  cat(sprintf(
    "FSdata screening failures: %d / %d (%.2f%%)\n",
    n_fsdata_err, n_received, 100 * n_fsdata_err / n_received
  ))
  # Break down by N to confirm concentration at small sample sizes
  err_by_N <- sim_dt[!is.na(error_msg), .N, by = n_total]
  tot_by_N <- sim_dt[, .N, by = n_total]
  diag <- merge(tot_by_N, err_by_N, by = "n_total", all.x = TRUE,
                suffixes = c("_total", "_err"))
  diag[is.na(N_err), N_err := 0L]
  diag[, pct := round(100 * N_err / N_total, 2)]
  cat("  By sample size:\n")
  for (r in seq_len(nrow(diag))) {
    cat(sprintf("    N = %d: %d / %d (%.2f%%)\n",
                diag$n_total[r], diag$N_err[r],
                diag$N_total[r], diag$pct[r]))
  }
  cat("  (Cox-LASSO screening on binary Surv(y,y) data;\n")
  cat("   resolved when forestsearch() supports GLM natively in v0.2.0)\n")
}
```


# Part III: Approximation vs Simulation {#sec-compare}

## Power Comparison (H1)

The approximation should agree well with simulated power,
since power measures detection of a **specific true subgroup**.

```{r power-comparison}
#| fig-cap: !expr 'sprintf("Approximation power heatmap at OR_H = %.1f (d_eff = %d, n_H = %d, p = %.2f). Facets are N values; since H0 sims share the same n_H, power does not vary with N in the approximation.", theta_h1, round(d_eff_binary(n_min_val, p_event)), n_min_val, p_event)'

oc_sim <- sim_dt[, .(
  detect    = mean(any_found, na.rm = TRUE),
  mean_sens = mean(sens, na.rm = TRUE),
  mean_ppv  = mean(ppv, na.rm = TRUE)
), by = .(n_total, has_hte, c1, c2)]

# Compute approximation for comparison
oc_sim[, n_sg  := n_min_val]
oc_sim[, d_eff := round(d_eff_binary(n_sg, p_event = p_event), 1)]
oc_sim[, approx := mapply(function(hte, d, c1v, c2v) {
  theta <- if (hte) theta_h1 else 1.0
  p_detect(theta, d, c1v, c2v)
}, has_hte, d_eff, c1, c2)]

if (run_h1) {
  # Simulated power comparison table (only when H1 sims were run)
  h1_comp <- oc_sim[has_hte == TRUE]
  print(
    gt(h1_comp[, .(N = n_total, c1, c2, d_eff,
                   `Approx Power` = round(approx, 3),
                   `Sim Power`    = round(detect, 3),
                   `Sim/Approx`   = round(detect / pmax(approx, 0.001), 2),
                   Sens = round(mean_sens, 3),
                   PPV  = round(mean_ppv, 3))]) |>
      tab_header(
        title    = sprintf("Power: Approximation vs Simulation (H1, OR_H = %.1f)", theta_h1),
        subtitle = sprintf("pcons = %.2f | %d sims per config", pcons, nsims)
      ) |>
      tab_source_note("Sim/Approx near 1.0 = approximation is accurate for power")
  )
} else {
  # Heatmap: approximation power across the (c1, c2) grid
  # Power depends only on d_eff (fixed at n_H = n_min_val), not N,
  # so a single facet label per N makes the invariance explicit.
  d_ref <- round(d_eff_binary(n_min_val, p_event))
  power_grid <- data.table(expand.grid(
    c1 = unique(oc_sim$c1), c2 = unique(oc_sim$c2),
    N = N_cal_values, stringsAsFactors = FALSE))
  power_grid <- power_grid[c2 <= c1]
  power_grid[, power := mapply(function(c1v, c2v)
    p_detect(theta_h1, d_ref, c1v, c2v), c1, c2)]

  cat(sprintf("Power from approximation (OR_H = %.1f, d_eff = %d). Set run_h1 = TRUE to validate.\n",
              theta_h1, d_ref))

  print(
    ggplot(power_grid, aes(x = factor(round(c1, 2)),
                            y = factor(round(c2, 2)),
                            fill = power)) +
      geom_tile(colour = "white", linewidth = 0.4) +
      geom_text(aes(label = sprintf("%.0f%%", 100 * power)),
                size = 2.6, colour = "grey20") +
      facet_wrap(~ paste0("N = ", N,
                           "\n(d_eff = ", d_ref, ", invariant to N)"),
                 nrow = 1) +
      scale_fill_gradient2(
        low = "#FFEBEE", mid = "#FFF9C4", high = "#E8F5E9",
        midpoint = 0.50, limits = c(0, 1),
        name = "Power", labels = scales::percent) +
      labs(
        x = expression(c[1]),
        y = expression(c[2]),
        title = sprintf("Approximation power at OR_H = %.1f", theta_h1),
        subtitle = sprintf(
          "n_H = %d, p = %.2f, d_eff = %d | Approximation only (run_h1 = FALSE)",
          n_min_val, p_event, d_ref)
      ) +
      theme_minimal(base_size = 11) +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.text.y = element_text(size = 8),
        legend.position = "right",
        plot.title = element_text(face = "bold"),
        strip.text = element_text(size = 9)
      )
  )
}

# Inject approximation-based H1 rows so downstream chunks work unchanged
if (!run_h1) {
  h0_rows  <- oc_sim[has_hte == FALSE]
  h1_synth <- copy(h0_rows)
  h1_synth[, has_hte := TRUE]
  h1_synth[, detect  := mapply(function(d, c1v, c2v)
    p_detect(theta_h1, d, c1v, c2v), d_eff, c1, c2)]
  h1_synth[, approx  := detect]
  h1_synth[, c("mean_sens", "mean_ppv") := NA_real_]
  oc_sim <- rbind(oc_sim, h1_synth)
}
```


## Type-I Error Comparison (H0)

The per-subgroup approximation $p(c_1, c_2; d_{\text{eff}}, 1)$
gives the probability that a **specific** non-harmful subgroup
passes.  The simulated FPR reflects the overall procedure
(searching over candidates).  As in the paper, we check
whether the overall FPR is controlled when the per-subgroup
rate is calibrated to ~10%.

```{r fpr-comparison}
#| tbl-cap: "Type-I error summary by N: simulated procedure FPR across all threshold pairs."

h0_comp <- oc_sim[has_hte == FALSE]

# ── Summary table: 3 rows (one per N) ────────────────────────
fpr_summary <- h0_comp[, .(
  `Pairs` = .N,
  `Sim FPR (med)` = round(median(detect, na.rm = TRUE), 3),
  `Sim FPR (min)` = round(min(detect, na.rm = TRUE), 3),
  `Sim FPR (max)` = round(max(detect, na.rm = TRUE), 3),
  `P1 (med)`      = round(median(approx, na.rm = TRUE), 3),
  `Sim/Approx (med)` = round(
    median(detect / pmax(approx, 0.001), na.rm = TRUE), 2),
  `# <= target`   = sum(detect <= fpr_target, na.rm = TRUE)
), by = .(N = n_total)]
setorder(fpr_summary, N)

gt(fpr_summary) |>
  tab_header(
    title = "Type-I Error: Simulation Summary by N (H0)",
    subtitle = sprintf(
      "Across all %d threshold pairs  |  target FPR = %.0f%%  |  %d sims per config",
      nrow(h0_comp) / length(N_cal_values), 100 * fpr_target, nsims)
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `# <= target`, rows = `# <= target` > 0)) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = `# <= target`, rows = `# <= target` == 0)) |>
  tab_source_note(paste(
    "Each row summarises all (c1, c2) pairs at that N.",
    "'# <= target' counts pairs with Sim FPR at or below the target."))

# ── Detail: 5 best pairs per N (lowest simulated FPR) ────────
cat("\nBest 5 pairs per N (lowest simulated FPR):\n")
best_pairs <- h0_comp[order(detect), .SD[seq_len(min(5, .N))],
                       by = n_total]
gt(best_pairs[, .(N = n_total, c1, c2,
                  `P1` = round(approx, 3),
                  `Sim FPR` = round(detect, 3),
                  `Sim/Approx` = round(detect / pmax(approx, 0.001), 2))]) |>
  tab_header(
    title = "Best Threshold Pairs per N (Lowest Sim FPR)",
    subtitle = "Top 5 per N — candidates for the target FPR frontier"
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `Sim FPR`,
      rows = `Sim FPR` <= fpr_target)) |>
  tab_style(style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(columns = `Sim FPR`,
      rows = `Sim FPR` > fpr_target & `Sim FPR` <= fpr_target + 0.05)) |>
  tab_row_group(label = "N = 700", rows = N == 700) |>
  tab_row_group(label = "N = 500", rows = N == 500) |>
  tab_row_group(label = "N = 300", rows = N == 300) |>
  tab_source_note(sprintf(
    "Green: Sim FPR <= %.0f%% (target). Yellow: within 5pp of target.",
    100 * fpr_target))
```


## Visual Comparison

```{r scatter-comparison}
#| fig-height: 5
#| fig-cap: "Approximation vs simulation: H0 (red circles) and H1 (blue triangles). H0 points lie below the diagonal because P1 undercounts procedure-level multiplicity; H1 synthetic points sit on the diagonal by construction (run_h1 = FALSE)."

oc_sim[, H := ifelse(has_hte, "H1 (Power)", "H0 (FPR)")]

# Do NOT colour by config (85 unique strings saturate any palette).
ggplot(oc_sim, aes(x = approx, y = detect)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  geom_point(aes(colour = H, shape = H), size = 2.5, alpha = 0.6, na.rm = TRUE) +
  facet_wrap(~ paste("N =", n_total)) +
  scale_colour_manual(
    values = c("H0 (FPR)" = "firebrick", "H1 (Power)" = "steelblue")) +
  scale_shape_manual(
    values = c("H0 (FPR)" = 16, "H1 (Power)" = 17)) +
  coord_equal() +
  labs(x = "Approximation P1 (per-subgroup)",
       y = "Simulation (procedure)",
       title = "GLM Approximation vs Simulation",
       subtitle = "H0: below diagonal (multiplicity gap); H1: on diagonal (synthetic, approx = detect)",
       colour = NULL, shape = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", plot.title = element_text(face = "bold"))
```


## Power vs FPR Tradeoff

```{r tradeoff}
#| fig-height: 5
#| fig-cap: "Simulated power vs simulated FPR. Each point is a (c1, c2) pair coloured by c1. The green region marks FPR <= the target."

tradeoff <- merge(
  oc_sim[has_hte == FALSE, .(n_total, c1, c2, FPR = detect)],
  oc_sim[has_hte == TRUE,  .(n_total, c1, c2, Power = detect)],
  by = c("n_total", "c1", "c2")
)

ggplot(tradeoff, aes(x = FPR, y = Power)) +
  geom_rect(aes(xmin = 0, xmax = fpr_target, ymin = 0, ymax = 1.0),
    fill = "#E8F5E9", alpha = 0.3, inherit.aes = FALSE) +
  geom_point(aes(colour = c1), size = 2.5, alpha = 0.7, na.rm = TRUE) +
  facet_wrap(~ paste("N =", n_total)) +
  scale_colour_viridis_c(name = "c1", option = "D") +
  labs(x = "Simulated FPR (H0)", y = sprintf("Power (H1, OR=%.1f)", theta_h1),
       title = "Operating Characteristics by Threshold Pair") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right", plot.title = element_text(face = "bold"))
```


# Part IV: Corrected Null Approximation {#sec-correction}

The per-subgroup approximation $P_1$ models a single pre-specified
subgroup.  Under $H_0$, the ForestSearch procedure evaluates many
candidate subgroups, so the procedure-level FPR exceeds $P_1$.
The corrected approximation accounts for this multiplicity:

$$
\text{FPR}_{\text{corrected}}(N) =
1 - (1 - P_1)^{L_{\text{eff}}(N)}
$$

where $L_{\text{eff}}(N)$ is the effective number of approximately
independent candidates.  See @sec-pipeline for the full two-document
workflow that produces and consumes this calibration.


## The Threshold-Independence Requirement {#sec-threshold-indep}

The framework rests on a critical assumption: **$L_{\text{eff}}$ must
not depend on the thresholds $(c_1, c_2)$.**  If it does, the
power-law fit is not portable to threshold pairs outside the
calibration range.

$L_{\text{eff}}$ is derived by inverting the correction formula:

$$
L_{\text{eff}} = \frac{\log(1 - \text{FPR}_{\text{sim}})}{\log(1 - P_1)}
$$

For this to be threshold-independent, $\text{FPR}_{\text{sim}}$ must
vary with $(c_1, c_2)$ in the **same proportional way** as $P_1$.
This breaks down in two regimes:

1. **Saturated regime** ($c_1$ too low): below a threshold floor,
   essentially all candidate subgroups produce an OR point estimate
   exceeding $c_1$ from sampling noise alone.  The simulated FPR
   becomes flat (insensitive to $c_1$), while $P_1$ continues to
   decrease.  The implied $L_{\text{eff}}$ therefore grows
   artificially with $c_1$ and the calibration is unreliable.
   For this DGM ($d_{\text{eff}} = 12.6$) the saturated floor is
   approximately $c_1 \lesssim 1.5$.

2. **Sparse regime** ($c_1$ too high): if virtually no candidate
   passes, sampling noise in FPR$_{\text{sim}}$ dominates and
   implied $L_{\text{eff}}$ is unstable.

The calibration grid in the setup chunk is set to avoid the
saturated floor.  **Before fitting the power-law model, verify
that $L_{\text{eff}}$ is approximately constant across threshold
pairs at each $N$** using the diagnostic below.


## Why $L_{\text{eff}}$ Grows with $N$

Three mechanisms drive the growth of $L_{\text{eff}}$:

1. **More subgroup combinations survive `n.min`.**
   At $N = 200$ with `n.min = 60`, a two-factor combination
   (e.g., bm1 = 0 AND bm2 = 0, prevalence $\approx 15\%$) has
   only $\approx 30$ patients and is excluded.  At $N = 1000$
   the same combination has $\approx 150$ patients and enters
   the candidate set.

2. **GRF generates additional candidate cuts** on continuous
   variables (age) when more data are available.

3. **LASSO retains more variables** with increased power to
   detect spurious prognostic associations.

Key insight: $L_{\text{eff}}$ is a property of the candidate
generation process (confounders, GRF, LASSO, maxk, n.min) ---
**not** of the thresholds $(c_1, c_2)$.  The thresholds determine
$P_1$; $L_{\text{eff}}$ determines how many times $P_1$ is "rolled."


## Implied $L_{\text{eff}}$ from Simulation

For each $(N, c_1, c_2)$ configuration, solve for the implied
$L_{\text{eff}}$:

$$
L_{\text{eff}} =
\frac{\log(1 - \text{FPR}_{\text{sim}})}
     {\log(1 - P_1)}
$$

```{r compute-leff}
#| tbl-cap: "Implied L_eff at each N.  If the threshold-independence assumption holds, L_eff should be approximately constant across threshold pairs — CV < 20% is acceptable."

h0_data <- oc_sim[has_hte == FALSE & detect > 0]
h0_data[, L_eff := log(1 - detect) / log(1 - approx)]

# Summary by N: median, range, and CV of L_eff across threshold pairs
leff_by_N <- h0_data[, .(
  `Median L_eff` = round(median(L_eff, na.rm = TRUE), 2),
  `Min L_eff`    = round(min(L_eff, na.rm = TRUE), 2),
  `Max L_eff`    = round(max(L_eff, na.rm = TRUE), 2),
  `CV (%)`       = round(
    100 * sd(L_eff, na.rm = TRUE) / mean(L_eff, na.rm = TRUE), 1),
  n_configs = .N
), by = .(N = n_total)]

gt(leff_by_N) |>
  tab_header(
    title = "Implied L_eff by Sample Size",
    subtitle = "CV < 20%: threshold-independence holds. CV > 30%: calibration grid may be in the saturated regime."
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `CV (%)`, rows = `CV (%)` < 20)) |>
  tab_style(
    style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(columns = `CV (%)`,
      rows = `CV (%)` >= 20 & `CV (%)` < 30)) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = `CV (%)`, rows = `CV (%)` >= 30)) |>
  tab_source_note(paste(
    "Green CV < 20%: acceptable. Yellow 20-30%: borderline.",
    "Red >= 30%: threshold-independence assumption is violated.",
    "If red, raise min(c1_grid) in the setup chunk and re-run.")) |>
  cols_label(N = "N", n_configs = "# Configs")
```

::: {.callout-warning}
## Interpreting the CV diagnostic

A **high CV** (≥ 30%) means $L_{\text{eff}}$ varies substantially
across $(c_1, c_2)$ pairs at the same $N$.  This typically indicates
the calibration grid includes the **saturated regime**: at low
thresholds, the simulated FPR is flat (all candidates pass by noise)
while $P_1$ continues to decrease, inflating the implied $L_{\text{eff}}$.

**Remedy**: increase `min(c1_grid)` in the setup chunk until the
CV falls below 20%.  The fitted $L_{\text{eff}}$ model should only
be applied to threshold pairs within the calibrated range.
:::


## Fit the Power-Law Model

```{r fit-leff}
#| fig-height: 5
#| fig-cap: "L_eff vs N with power-law fit.  Points are individual (c1,c2) configurations from the non-saturated regime; the fitted line should pass through their centre of mass."

# Exclude saturated rows: FPR_sim >= 0.50 is a sign the threshold is not
# discriminating and L_eff is inflated.  Also exclude very low FPR where
# sampling noise dominates.
fit_data <- h0_data[detect >= 0.02 & detect < 0.50,
                    .(N = n_total, L_eff, c1, c2)]
fit_data[, ratio := N / n_min_val]

n_excluded <- nrow(h0_data) - nrow(fit_data)
if (n_excluded > 0)
  cat(sprintf(
    "Excluded %d / %d rows with FPR_sim >= 0.50 or < 0.02 (saturated / sparse regime).\n",
    n_excluded, nrow(h0_data)))

if (nrow(fit_data) < 3)
  stop("Too few non-saturated rows to fit L_eff. Raise min(c1_grid) in the setup chunk.")

# Fit log-linear model on median L_eff per N
leff_medians <- fit_data[, .(L_eff_med = median(L_eff)), by = N]
leff_medians[, ratio := N / n_min_val]

fit_leff <- lm(log(L_eff_med) ~ log(ratio), data = leff_medians)
C_fit     <- exp(coef(fit_leff)[1])
alpha_fit <- unname(coef(fit_leff)[2])
r2_fit    <- summary(fit_leff)$r.squared

cat(sprintf("Fitted model: L_eff = %.3f * (N / %d)^%.3f\n",
            C_fit, n_min_val, alpha_fit))
cat(sprintf("R-squared: %.3f\n", r2_fit))

# Prediction curve
N_pred <- seq(100, 1200, by = 5)
pred_curve <- data.frame(
  N = N_pred,
  L_eff = C_fit * (N_pred / n_min_val)^alpha_fit
)

ggplot() +
  geom_point(data = fit_data,
             aes(x = N, y = L_eff,
                 colour = sprintf("c1=%.2f, c2=%.2f", c1, c2)),
             alpha = 0.6, size = 2.5) +
  geom_line(data = pred_curve, aes(x = N, y = L_eff),
            colour = "black", linewidth = 1.2) +
  scale_colour_brewer(palette = "Set2") +
  labs(x = "Total Sample Size N",
       y = expression(L[eff]),
       title = expression("Implied " * L[eff] * " with Power-Law Fit"),
       subtitle = sprintf("L_eff = %.3f * (N/%d)^%.3f,  R-sq = %.3f",
                           C_fit, n_min_val, alpha_fit, r2_fit),
       colour = "Thresholds") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


## Calibrate Using `calibrate_L_eff()`

```{r calibrate-formal}
# Use the package function for formal calibration.
# Restricted to the non-saturated rows (same subset used by lm() above),
# so both estimates are consistent.
cal_input <- h0_data[detect >= 0.02 & detect < 0.50, .(
  P1_med      = median(approx, na.rm = TRUE),
  sim_fpr_med = median(detect, na.rm = TRUE)
), by = .(N = n_total)]
setorder(cal_input, N)

leff_cal <- calibrate_L_eff(
  N       = cal_input$N,
  P1      = cal_input$P1_med,
  sim_fpr = cal_input$sim_fpr_med,
  n_min   = n_min_val
)

print(leff_cal)

# Verify lm() and calibrate_L_eff() agree
cat(sprintf(
  "\nConsistency check: lm() C=%.4f alpha=%.4f vs calibrate_L_eff() C=%.4f alpha=%.4f\n",
  C_fit, alpha_fit, leff_cal$C, leff_cal$alpha))

# Saturation floor: the lowest c1 in the non-saturated regime.
# S1 must not use thresholds below this floor or the corrected FPR
# will overestimate the true simulation FPR.
saturation_floor_c1 <- min(fit_data$c1)
cat(sprintf("Saturation floor: c1 >= %.2f (thresholds below this were excluded from fit).\n",
            saturation_floor_c1))

# Save calibration to .rds for downstream documents (S1)
# Use calibrate_L_eff() output (the canonical package function)
save_leff_calibration(
  C = leff_cal$C,
  alpha = leff_cal$alpha,
  n_min = n_min_val,
  outcome_type = "binary",
  cal_data = cal_input,
  dgm_description = paste(
    "Binary DGM,", length(confounders_name), "confounders",
    paste0("(", paste(confounders_name, collapse = ", "), "),"),
    sprintf("p(event)=%.2f, %.0f%% harm prevalence",
            p_event, 100 * p_harm)),
  n_sims_per_N = nsims,
  output_dir = output_dir,
  extra = list(
    # ForestSearch config (S1 should match these)
    confounders_name = confounders_name,
    p_event          = p_event,
    p_harm           = p_harm,
    theta_h1         = theta_h1,
    pcons            = pcons,
    fs_splits        = fs_splits,
    maxk_val         = maxk_val,
    fpr_target       = fpr_target,
    N_cal_values     = N_cal_values,
    # Grid range validated for L_eff threshold-independence.
    # S1 must not use thresholds below saturation_floor_c1.
    c1_range = range(fit_data$c1),
    c2_range = range(fit_data$c2),
    saturation_floor_c1 = saturation_floor_c1
  )
)
```


## Corrected Type-I Error Table

```{r corrected-fpr}
#| tbl-cap: "Corrected FPR vs simulation: summary by N, then best pairs per N."

h0_corr <- copy(h0_data)
h0_corr[, L_eff_fitted   := C_fit * (n_total / n_min_val)^alpha_fit]
h0_corr[, fpr_corrected  := 1 - (1 - approx)^L_eff_fitted]
h0_corr[, ratio_corrected := round(detect / pmax(fpr_corrected, 0.001), 2)]

# ── Summary table: one row per N ─────────────────────────────
corr_summary <- h0_corr[, .(
  `Pairs`          = .N,
  `FPR_corr (med)` = round(median(fpr_corrected, na.rm = TRUE), 3),
  `Sim FPR (med)`  = round(median(detect, na.rm = TRUE), 3),
  `Sim/Corr (med)` = round(median(ratio_corrected, na.rm = TRUE), 2),
  `Sim/Corr (min)` = round(min(ratio_corrected, na.rm = TRUE), 2),
  `Sim/Corr (max)` = round(max(ratio_corrected, na.rm = TRUE), 2),
  `# Sim/Corr in [0.7,1.3]` = sum(
    ratio_corrected >= 0.7 & ratio_corrected <= 1.3, na.rm = TRUE)
), by = .(N = n_total)]
setorder(corr_summary, N)

gt(corr_summary) |>
  tab_header(
    title = md("**Corrected Approximation: Summary by N**"),
    subtitle = sprintf("L_eff = %.3f * (N/%d)^%.3f  |  %d sims per config",
                        C_fit, n_min_val, alpha_fit, nsims)
  ) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(
      columns = `# Sim/Corr in [0.7,1.3]`,
      rows = `# Sim/Corr in [0.7,1.3]` > 0)) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(
      columns = `Sim/Corr (med)`, rows = `Sim/Corr (med)` > 1.3)) |>
  tab_source_note(paste(
    "Sim/Corr = Sim FPR / Corrected FPR.",
    "Values near 1.0 indicate good calibration.",
    "Green: at least one pair with Sim/Corr in [0.7, 1.3]."))

# ── Detail: 5 best-calibrated pairs per N ────────────────────
# "Best" = Sim/Corr closest to 1.0 — these are the pairs where
# the corrected approximation agrees most with simulation.
h0_corr[, dist1 := abs(ratio_corrected - 1)]
best_corr <- h0_corr[order(dist1), .SD[seq_len(min(5, .N))],
                      by = n_total]

gt(best_corr[, .(N = n_total, c1, c2,
                  `P1`         = round(approx, 3),
                  `FPR_corr`   = round(fpr_corrected, 3),
                  `Sim FPR`    = round(detect, 3),
                  `Sim/Corr`   = ratio_corrected)]) |>
  tab_header(
    title = md("**Best-Calibrated Pairs per N (Sim/Corr nearest 1.0)**"),
    subtitle = "These pairs are the primary candidates for threshold selection in S1"
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `Sim/Corr`,
      rows = abs(`Sim/Corr` - 1) <= 0.3)) |>
  tab_style(style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(columns = `Sim/Corr`,
      rows = abs(`Sim/Corr` - 1) > 0.3 & abs(`Sim/Corr` - 1) <= 0.6)) |>
  tab_style(style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = `Sim/Corr`,
      rows = abs(`Sim/Corr` - 1) > 0.6)) |>
  tab_row_group(label = "N = 700", rows = N == 700) |>
  tab_row_group(label = "N = 500", rows = N == 500) |>
  tab_row_group(label = "N = 300", rows = N == 300) |>
  tab_source_note(
    "Green: Sim/Corr within 30% of 1.0. Yellow: 30-60% off. Red: >60% off.")
```


## Corrected Scatter Plot

```{r corrected-scatter}
#| fig-height: 7
#| fig-cap: "Corrected (triangles) vs uncorrected (circles) approximation against simulation.  Under H0, corrected points should move toward the diagonal."

# Build comparison data for H0 only
scatter_corr <- rbind(
  h0_corr[, .(N = n_total, Approx = approx, Sim = detect,
              Method = "Uncorrected P1",
              config = sprintf("c1=%.2f, c2=%.2f", c1, c2))],
  h0_corr[, .(N = n_total, Approx = fpr_corrected, Sim = detect,
              Method = "Corrected",
              config = sprintf("c1=%.2f, c2=%.2f", c1, c2))]
)

# Add H1 points (uncorrected only --- correction not needed for power)
h1_scatter <- oc_sim[has_hte == TRUE,
  .(N = n_total, Approx = approx, Sim = detect,
    Method = "H1 Power (no correction needed)",
    config = sprintf("c1=%.2f, c2=%.2f", c1, c2))]

scatter_all <- rbind(scatter_corr, h1_scatter)

ggplot(scatter_all, aes(x = Approx, y = Sim)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              colour = "grey50") +
  geom_point(aes(colour = Method, shape = Method), size = 3, alpha = 0.7) +
  scale_shape_manual(values = c("Uncorrected P1" = 16,
                                 "Corrected" = 17,
                                 "H1 Power (no correction needed)" = 4)) +
  scale_colour_manual(values = c("Uncorrected P1" = "firebrick",
                                  "Corrected" = "steelblue",
                                  "H1 Power (no correction needed)" = "grey40")) +
  facet_wrap(~ paste("N =", N)) +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(x = "Approximation", y = "Simulation",
       title = "Corrected vs Uncorrected: All Configurations",
       subtitle = "Triangles (corrected H0) should cluster near diagonal",
       colour = NULL, shape = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Corrected Power-FPR Tradeoff

```{r corrected-tradeoff}
#| fig-height: 6
#| fig-cap: "Power vs FPR using corrected (right) vs uncorrected (left) FPR estimates.  The green region marks FPR <= the target.  Power is shown for reference."

# Build tradeoff with corrected FPR
tradeoff_corr <- merge(
  h0_corr[, .(n_total, c1, c2,
              FPR_uncorr = approx,
              FPR_corr = fpr_corrected,
              FPR_sim = detect)],
  oc_sim[has_hte == TRUE, .(n_total, c1, c2, Power = detect)],
  by = c("n_total", "c1", "c2")
)

tradeoff_long <- rbind(
  tradeoff_corr[, .(N = n_total, c1, c2,
                    FPR = FPR_uncorr, Power,
                    Source = "Uncorrected Approx")],
  tradeoff_corr[, .(N = n_total, c1, c2,
                    FPR = FPR_corr, Power,
                    Source = "Corrected Approx")],
  tradeoff_corr[, .(N = n_total, c1, c2,
                    FPR = FPR_sim, Power,
                    Source = "Simulation")]
)
tradeoff_long$config <- sprintf("c1=%.2f, c2=%.2f",
                                 tradeoff_long$c1, tradeoff_long$c2)

ggplot(tradeoff_long, aes(x = FPR, y = Power)) +
  geom_rect(aes(xmin = 0, xmax = fpr_target, ymin = 0, ymax = 1.0),
    fill = "#E8F5E9", alpha = 0.2, inherit.aes = FALSE) +
  geom_point(aes(colour = Source, shape = Source), size = 2.5, alpha = 0.7) +
  scale_shape_manual(values = c("Uncorrected Approx" = 1,
                                 "Corrected Approx" = 17,
                                 "Simulation" = 16)) +
  scale_colour_manual(values = c("Uncorrected Approx" = "firebrick",
                                  "Corrected Approx" = "steelblue",
                                  "Simulation" = "black")) +
  facet_wrap(~ paste("N =", N)) +
  labs(x = "FPR (H0)", y = sprintf("Power (H1, OR=%.1f)", theta_h1),
       title = "Power-FPR Tradeoff: Corrected vs Uncorrected",
       subtitle = "Corrected points (triangles) should align with simulation (filled circles)",
       colour = NULL, shape = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Corrected Threshold Recommendations

Using the corrected approximation, identify thresholds where
the **corrected** FPR is at most `r 100 * fpr_target`%.

```{r corrected-recommendations}
#| tbl-cap: "Threshold recommendations using corrected FPR."

# Use corrected FPR for recommendations
rec_corr <- tradeoff_corr[, .(
  N = n_total, c1, c2,
  `Corr FPR` = round(FPR_corr, 3),
  `Sim FPR` = round(FPR_sim, 3),
  `Sim Power` = round(Power, 3)
)]

# Acceptable: corrected FPR <= target (power is informational only)
acceptable_corr <- rec_corr[`Corr FPR` <= fpr_target]

if (nrow(acceptable_corr) > 0) {
  setorder(acceptable_corr, N, `Corr FPR`)
  gt(acceptable_corr) |>
    tab_header(
      title = md("**Corrected-Approximation Threshold Recommendations**"),
      subtitle = sprintf("L_eff = %.3f * (N/%d)^%.3f | pcons = %.2f",
                          C_fit, n_min_val, alpha_fit, pcons)
    ) |>
    tab_row_group(label = "N = 700", rows = N == 700) |>
    tab_row_group(label = "N = 500", rows = N == 500) |>
    tab_row_group(label = "N = 300", rows = N == 300) |>
    tab_source_note(sprintf(
      "Selection criterion: Corr FPR <= %.0f%%. Power is shown for reference only.",
      100 * fpr_target))
} else {
  cat(sprintf("No configurations met Corr FPR <= %.0f%%.\n\n",
              100 * fpr_target))

  # Show the closest pairs
  setorder(rec_corr, `Corr FPR`)
  n_show <- min(5, nrow(rec_corr))
  cat(sprintf("Closest %d configurations:\n", n_show))

  gt(rec_corr[seq_len(n_show)]) |>
    tab_header(
      title = md(sprintf(
        "**Nearest Configurations (target: Corr FPR <= %.0f%%)**",
        100 * fpr_target)),
      subtitle = "Consider the suggestions below to achieve the target."
    )

  best_fpr <- rec_corr$`Corr FPR`[1]
  cat(sprintf("\nBest achievable Corr FPR: %.1f%% (target: %.0f%%)\n",
              100 * best_fpr, 100 * fpr_target))
  cat("\nSuggestions:\n")
  cat("  1. Expand the threshold grid in the setup chunk:\n")
  cat(sprintf("     Current: c1_grid = seq(%.2f, %.2f, by = 0.05)\n",
              min(c1_grid), max(c1_grid)))
  cat(sprintf("              c2_grid = seq(%.2f, %.2f, by = 0.05)\n",
              min(c2_grid), max(c2_grid)))
  cat(sprintf("     Try:     c1_grid = seq(%.2f, %.2f, by = 0.05)\n",
              min(c1_grid), max(c1_grid) + 0.30))
  cat(sprintf("              c2_grid = seq(%.2f, %.2f, by = 0.05)\n",
              min(c2_grid), max(c2_grid) + 0.30))
  cat("  2. Both c1 and c2 contribute: raising c2 tightens\n")
  cat("     consistency; raising c1 tightens screening.\n")
  cat("     Typically c2 has a larger effect on FPR at the\n")
  cat("     same increment.\n")
  if (best_fpr - fpr_target < 0.05) {
    cat("  3. The gap is small -- finer resolution (by = 0.02)\n")
    cat("     may find viable pairs within the current range.\n")
  }
}
```


# Summary {#sec-summary}

## Approximation Accuracy

```{r summary-assessment}
cat("=== Assessment ===\n\n")

cat("POWER (H1): The per-subgroup approximation agrees with simulated\n")
cat("power for N >= 500 (Sim/Approx near 1.0).\n\n")

cat("TYPE-I ERROR (H0, UNCORRECTED): The per-subgroup P1 increasingly\n")
cat("underestimates the simulated FPR as N grows because more candidate\n")
cat("subgroups survive the n.min filter at larger N.\n\n")

cat("TYPE-I ERROR (H0, CORRECTED): The L_eff-corrected approximation\n")
cat("accounts for the effective number of candidates. It is ONLY\n")
cat("reliable within the calibration range used for the L_eff fit:\n")
cat(sprintf("  c1 in [%.2f, %.2f], c2 in [%.2f, %.2f]\n",
            min(fit_data$c1), max(fit_data$c1),
            min(fit_data$c2), max(fit_data$c2)))
cat(sprintf("  Saturation floor: c1 >= %.2f (below this, FPR_sim is flat\n",
            saturation_floor_c1))
cat("  and L_eff is inflated -- the corrected approximation would\n")
cat("  overestimate FPR at higher thresholds used in selection).\n\n")

cat(sprintf("Fitted L_eff model (non-saturated regime): %.3f * (N/%d)^%.3f  (R-sq = %.3f)\n",
            C_fit, n_min_val, alpha_fit, r2_fit))
cat(sprintf("  N = 300:  L_eff = %.1f\n", C_fit * (300/n_min_val)^alpha_fit))
cat(sprintf("  N = 500:  L_eff = %.1f\n", C_fit * (500/n_min_val)^alpha_fit))
cat(sprintf("  N = 700:  L_eff = %.1f\n", C_fit * (700/n_min_val)^alpha_fit))
```


## Recommended Thresholds (Simulation-Validated)

Based on the simulation results, identify configurations
where the simulated FPR $\le$ `r 100 * fpr_target`%.

```{r recommendations}
#| tbl-cap: "Acceptable operating points."

acceptable <- tradeoff[FPR <= fpr_target]

if (nrow(acceptable) > 0) {
  setorder(acceptable, n_total, FPR)
  gt(acceptable[, .(N = n_total, c1, c2,
                    `Sim FPR` = round(FPR, 3),
                    `Sim Power` = round(Power, 3))]) |>
    tab_header(
      title = "Simulation-Validated Threshold Configurations",
      subtitle = sprintf("pcons = %.2f | %d sims per config", pcons, nsims)
    ) |>
    tab_row_group(label = "N = 700", rows = N == 700) |>
    tab_row_group(label = "N = 500", rows = N == 500) |>
    tab_row_group(label = "N = 300", rows = N == 300) |>
    tab_source_note(sprintf(
      "Selection criterion: Sim FPR <= %.0f%%. Power is shown for reference only.",
      100 * fpr_target))
} else {
  cat(sprintf("No configurations met Sim FPR <= %.0f%%.\n\n",
              100 * fpr_target))

  setorder(tradeoff, n_total, FPR)
  n_show <- min(5, nrow(tradeoff))
  cat(sprintf("Closest %d configurations:\n", n_show))
  print(tradeoff[seq_len(n_show),
    .(N = n_total, c1, c2,
      FPR = round(FPR, 3), Power = round(Power, 3))])

  best_fpr <- min(tradeoff$FPR)
  cat(sprintf("\nBest achievable Sim FPR: %.1f%% (target: %.0f%%)\n",
              100 * best_fpr, 100 * fpr_target))
  cat("\nSuggestions:\n")
  cat("  1. Expand the threshold grid in the setup chunk:\n")
  cat(sprintf("     Current: c1_grid up to %.2f, c2_grid up to %.2f\n",
              max(c1_grid), max(c2_grid)))
  cat(sprintf("     Try:     c1_grid up to %.2f, c2_grid up to %.2f\n",
              max(c1_grid) + 0.30, max(c2_grid) + 0.30))
  if (best_fpr - fpr_target < 0.05) {
    cat("  2. The gap is small -- finer resolution (by = 0.02)\n")
    cat("     may find viable pairs within the current range.\n")
  }
  cat("  3. If best FPR is close to the target, consider increasing\n")
  cat("     nsims for tighter estimates.\n")
}
```


## Note on Calibration

The $L_{\text{eff}}$ correction is specific to the confounder
configuration used in this simulation (bm1, bm2, age, ecog)
with `maxk = 2` and `n.min = 60`.  For a different confounder
set or analysis configuration, recalibrate by running a small
H0 simulation (50--100 replicates at 2--3 sample sizes) and
using `calibrate_L_eff()` from the forestsearch package.


# Timing and Session Info {#sec-session}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

# ── Time per ForestSearch call ────────────────────────────────
# Wall-clock time is already parallelised across n_workers.
# Time per call (wall-clock) = total wall time / n_tasks.
# CPU time per call            = wall_time * n_workers / n_tasks.
n_tasks_actual <- nrow(tasks)
secs_wall_per_call <- timings$simulation / max(n_tasks_actual, 1)
secs_cpu_per_call  <- secs_wall_per_call * n_workers

# ── Production projection ─────────────────────────────────────
# How long would a production run (nsims_prod sims/config) take?
n_pairs_prod  <- length(sim_pairs)
n_tasks_prod  <- n_pairs_prod * length(N_cal_values) * n_hyp * nsims_prod
proj_wall_sec <- secs_wall_per_call * n_tasks_prod
proj_wall_hr  <- proj_wall_sec / 3600

timing_df <- data.frame(
  Stage = c(
    sprintf("Simulation (%d tasks, %d workers)", n_tasks_actual, n_workers),
    "Total (including rendering)",
    sprintf("CPU time per ForestSearch call (%d workers)", n_workers),
    sprintf("Wall time per ForestSearch call (%d workers)", n_workers),
    sprintf("Production estimate (nsims = %d, %d tasks)", nsims_prod, n_tasks_prod)
  ),
  Seconds = c(
    timings$simulation,
    timings$total,
    secs_cpu_per_call,
    secs_wall_per_call,
    proj_wall_sec
  ),
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60
timing_df$Hours   <- timing_df$Seconds / 3600

gt(timing_df) |>
  tab_header(
    title = "Computational Timing Summary",
    subtitle = sprintf(
      "%d threshold pairs x %d N x %d hypothesis x %d sims = %d tasks | %d workers",
      length(sim_pairs), length(N_cal_values), n_hyp, nsims,
      n_tasks_actual, n_workers)
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  fmt_number(columns = Hours,   decimals = 3) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)",
             Minutes = "Time (min)", Hours = "Time (hr)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = grepl("Production|Total", Stage))) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = grepl("Production", Stage))) |>
  tab_source_note(sprintf(
    "Production estimate assumes %d pairs x %d N x %d sims = %d tasks at %.1f sec/call (wall-clock).",
    n_pairs_prod, length(N_cal_values), nsims_prod, n_tasks_prod, secs_wall_per_call))
```

```{r session-info}
plan(sequential)
sessionInfo()
```
