---
title: "Cox vs GLM Approximation: Null Operating Characteristics"
subtitle: "Threshold Calibration for the GBSG Survival DGM"
author: "Larry F. Leon"
date: today
execute:
  warning: true
  message: true
  fig-align: center
knitr:
  opts_chunk:
    comment: "#>"
---

```{r setup}
#| include: false
# ── Simulation control ──────────────────────────────────────────
# ~750/hr linux
nsims_null  <- 2000L       # H0 sims for Part II (paper-defaults Cox & IRR)
nsims_cal   <- 1000L       # H0 sims per N for L_eff calibration (Part III)
nsims_grid  <- 2000L       # H0 sims per (c1,c2) pair for grid sweep (Part IV)
sim_seed    <- 8316951L
n_workers   <- ceiling(0.95 * max(1L,
  parallel::detectCores(logical = FALSE) - 1L))

# ── Production targets (for timing projection only) ─────────────
# These control what the "Projected production run" row reports.
# They do not affect what is actually simulated above.
nsims_prod_null <- 4000L   # target null sims (Parts II)
nsims_prod_cal  <- 2000L    # target cal sims per N (Part III)
nsims_prod_grid <- 2000L    # target grid sims per pair (Part IV)

# ── Output ──────────────────────────────────────────────────────
output_dir  <- "../_data"

# ── FPR target ─────────────────────────────────────────────────
fpr_target  <- 0.12

# ── DGM / outcome parameters ──────────────────────────────────
prop_cens   <- 0.46      # GBSG censoring proportion
p_harm      <- 0.13      # harm subgroup prevalence (metadata only)
n_min_val   <- 60L       # n.min = reference subgroup size for P1
hr_h1       <- 2.0       # HR for harm subgroup under H1 (reference)

# ── ForestSearch parameters ───────────────────────────────────
confounders_base <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")
pcons       <- 0.90      # pconsistency.threshold
fs_splits   <- 400L
maxk_val    <- 2L
n_sample    <- 700L      # reference sample size for main sims
analysis_time <- 84L     # GBSG analysis time (months)
cens_adjust   <- log(1.5)
fs_hr_threshold    <- 1.25   # c1: screening threshold (paper default)
fs_hr_consistency  <- 1.0    # c2: consistency threshold (paper default)
fs_d0_min          <- 12L    # min events, control arm
fs_d1_min          <- 12L    # min events, treatment arm
fs_by_risk         <- 12L    # risk-table interval (months)
fs_vi_grf_min      <- -0.2   # GRF variable importance threshold

# -- Adjusted Cox formula (used in run_simulation_analysis) ---
cox_formula_adj <- survival::Surv(y_sim, event_sim) ~
  treat_sim + z1 + z2 + z3

# ── L_eff calibration ────────────────────────────────────────
N_cal_values <- c(300, 500, 700)

# ── Threshold grid for null sweep (Part IV) ──────────────────
# C3 validates L_eff threshold-independence across this grid.
# S2's approximation grid should fall within this range.
c1_grid <- c(0.90,1.00, 1.10, 1.15, 1.20, 1.25, 1.30, 1.40, 1.50)
c2_grid <- c(0.80,0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20, 1.25)
```


# Introduction {#sec-intro}

Leon et al. (2024, Section 2.1) approximate the per-subgroup
detection probability as:

$$
P_1(c_1, c_2; d, \theta) =
P(W_1 + W_2 \ge 2\log c_1, \;\min(W_1, W_2) \ge \log c_2)
\tag{3}
$$

with $W_1, W_2 \stackrel{\text{iid}}{\sim} N(\log\theta, 8/d)$.

This document focuses on **null operating characteristics**
for the GBSG survival DGM: we evaluate how the false-positive
rate (FPR) varies across a grid of $(c_1, c_2)$ threshold
pairs to identify the region where the procedure-level FPR
is at most 10--12%.

We compare:

- **(B)** GLM approximation (per-subgroup $P_1$ via
  `compute_detection_probability_glm()`)
- **(C)** Cox ForestSearch simulation
- **(A)** Poisson+offset ForestSearch simulation
- **$L_{\text{eff}}$ correction**: $1 - (1 - P_1)^{L_{\text{eff}}}$

All simulations are under $H_0$ (uniform treatment effect,
no subgroup heterogeneity).


## Pipeline Overview {#sec-pipeline}

The figure below shows the complete workflow across the C3 and S2 documents.
C3 (this document) establishes the null operating characteristics and fits
the $L_{\text{eff}}$ power law. S2 (`selection_survival_frontier.qmd`)
consumes the calibration to sweep a dense threshold grid and select
operating points.

```{=html}
<figure style="margin:1.5rem 0 2rem;">
<svg id="surv-pipeline" width="100%" viewBox="0 0 680 860" role="img" xmlns="http://www.w3.org/2000/svg">
<title>Survival calibration pipeline: C3 to S2</title>
<desc>Flowchart of the two-document survival L_eff calibration pipeline. C3 runs Cox and Poisson H0 simulations at paper defaults, then multi-N calibration to fit L_eff, then a grid sweep at the reference N. S2 loads the calibration, applies the corrected FPR formula across a dense grid, selects and validates frontier pairs.</desc>
<defs>
<style>
#surv-pipeline .th{font-family:ui-sans-serif,system-ui,sans-serif;font-size:14px;font-weight:500}
#surv-pipeline .ts{font-family:ui-sans-serif,system-ui,sans-serif;font-size:12px;font-weight:400}
#surv-pipeline .arr{stroke:#888780;stroke-width:1.5;fill:none}
#surv-pipeline .leader{stroke:#B4B2A9;stroke-width:0.5;stroke-dasharray:4 3;fill:none}
#surv-pipeline .node{cursor:default}
#surv-pipeline .c-teal>rect{fill:#E1F5EE;stroke:#0F6E56;stroke-width:0.5}
#surv-pipeline .c-teal .th{fill:#085041}
#surv-pipeline .c-teal .ts{fill:#0F6E56}
#surv-pipeline .c-purple>rect{fill:#EEEDFE;stroke:#534AB7;stroke-width:0.5}
#surv-pipeline .c-purple .th{fill:#3C3489}
#surv-pipeline .c-purple .ts{fill:#534AB7}
#surv-pipeline .c-amber>rect{fill:#FAEEDA;stroke:#BA7517;stroke-width:0.5}
#surv-pipeline .c-amber .th{fill:#633806}
#surv-pipeline .c-amber .ts{fill:#854F0B}
#surv-pipeline .c-gray>rect{fill:#F1EFE8;stroke:#888780;stroke-width:0.5}
#surv-pipeline .c-gray .th{fill:#444441}
#surv-pipeline .c-gray .ts{fill:#5F5E5A}
#surv-pipeline .c-green>rect{fill:#EAF3DE;stroke:#3B6D11;stroke-width:0.5}
#surv-pipeline .c-green .th{fill:#27500A}
#surv-pipeline .c-green .ts{fill:#3B6D11}
#surv-pipeline .c-coral>rect{fill:#FAECE7;stroke:#993C1D;stroke-width:0.5}
#surv-pipeline .c-coral .th{fill:#4A1B0C}
#surv-pipeline .c-coral .ts{fill:#993C1D}
</style>
<marker id="sv-arrow" viewBox="0 0 10 10" refX="8" refY="5" markerWidth="6" markerHeight="6" orient="auto-start-reverse">
<path d="M2 1L8 5L2 9" fill="none" stroke="#888780" stroke-width="1.5" stroke-linecap="round" stroke-linejoin="round"/>
</marker>
</defs>
<!-- Section headers -->
<g class="c-teal">
<rect x="60" y="14" width="560" height="28" rx="6"/>
<text class="th" x="340" y="28" text-anchor="middle" dominant-baseline="central">C3 — calibration_survival_leff_grid.qmd</text>
</g>
<g class="c-purple">
<rect x="60" y="444" width="560" height="28" rx="6"/>
<text class="th" x="340" y="458" text-anchor="middle" dominant-baseline="central">S2 — selection_survival_frontier.qmd</text>
</g>
<!-- Separator -->
<line x1="80" y1="433" x2="600" y2="433" stroke="#B4B2A9" stroke-width="0.8" stroke-dasharray="5 3"/>
<!-- C3 arrows -->
<line x1="340" y1="106" x2="340" y2="122" class="arr" marker-end="url(#sv-arrow)"/>
<line x1="340" y1="172" x2="340" y2="188" class="arr" marker-end="url(#sv-arrow)"/>
<line x1="340" y1="238" x2="340" y2="254" class="arr" marker-end="url(#sv-arrow)"/>
<line x1="340" y1="304" x2="340" y2="320" class="arr" marker-end="url(#sv-arrow)"/>
<line x1="340" y1="370" x2="340" y2="386" class="arr" marker-end="url(#sv-arrow)"/>
<line x1="340" y1="420" x2="340" y2="488" class="arr" marker-end="url(#sv-arrow)"/>
<!-- S2 arrows -->
<line x1="340" y1="538" x2="340" y2="554" class="arr" marker-end="url(#sv-arrow)"/>
<line x1="340" y1="604" x2="340" y2="620" class="arr" marker-end="url(#sv-arrow)"/>
<line x1="340" y1="670" x2="340" y2="686" class="arr" marker-end="url(#sv-arrow)"/>
<line x1="340" y1="736" x2="340" y2="752" class="arr" marker-end="url(#sv-arrow)"/>
<!-- C3 nodes -->
<g class="node c-teal">
<rect x="140" y="56" width="400" height="50" rx="8"/>
<text class="th" x="340" y="74" text-anchor="middle" dominant-baseline="central">Part I: Per-subgroup approximation</text>
<text class="ts" x="340" y="91" text-anchor="middle" dominant-baseline="central">P1(c1, c2; d_eff)  ·  d_eff = n_H × (1 − prop_cens)</text>
</g>
<g class="node c-teal">
<rect x="140" y="122" width="400" height="50" rx="8"/>
<text class="th" x="340" y="140" text-anchor="middle" dominant-baseline="central">Part II: Baseline null simulations</text>
<text class="ts" x="340" y="157" text-anchor="middle" dominant-baseline="central">Cox PH + Poisson IRR at paper defaults (c1=1.25, c2=1.0)</text>
</g>
<g class="node c-teal">
<rect x="140" y="188" width="400" height="50" rx="8"/>
<text class="th" x="340" y="206" text-anchor="middle" dominant-baseline="central">Part III: Multi-N L_eff calibration</text>
<text class="ts" x="340" y="223" text-anchor="middle" dominant-baseline="central">L_eff = log(1−FPR) / log(1−P1)  ·  fit C·(N/n_min)^α</text>
</g>
<g class="node c-amber">
<rect x="140" y="254" width="400" height="50" rx="8"/>
<text class="th" x="340" y="272" text-anchor="middle" dominant-baseline="central">CV diagnostic — threshold-independence check</text>
<text class="ts" x="340" y="289" text-anchor="middle" dominant-baseline="central">CV of implied L_eff across grid  ·  flag if CV &gt; 30%</text>
</g>
<g class="node c-teal">
<rect x="140" y="320" width="400" height="50" rx="8"/>
<text class="th" x="340" y="338" text-anchor="middle" dominant-baseline="central">Part IV: Null threshold grid sweep</text>
<text class="ts" x="340" y="355" text-anchor="middle" dominant-baseline="central">Sim FPR at each (c1, c2) for N = n_sample</text>
</g>
<g class="node c-gray">
<rect x="190" y="386" width="300" height="34" rx="8"/>
<text class="th" x="340" y="398" text-anchor="middle" dominant-baseline="central">Save to .rds</text>
<text class="ts" x="340" y="414" text-anchor="middle" dominant-baseline="central">C, α, c1_range, grid_fpr</text>
</g>
<!-- S2 nodes -->
<g class="node c-purple">
<rect x="140" y="488" width="400" height="50" rx="8"/>
<text class="th" x="340" y="506" text-anchor="middle" dominant-baseline="central">Load .rds · verify grid range</text>
<text class="ts" x="340" y="523" text-anchor="middle" dominant-baseline="central">C, α  ·  S2 grid must stay within C3's validated range</text>
</g>
<g class="node c-purple">
<rect x="140" y="554" width="400" height="50" rx="8"/>
<text class="th" x="340" y="572" text-anchor="middle" dominant-baseline="central">Compute FPR_corr on dense grid</text>
<text class="ts" x="340" y="589" text-anchor="middle" dominant-baseline="central">FPR_corr = 1 − (1 − P1)^(C · (N/n_min)^α)</text>
</g>
<g class="node c-purple">
<rect x="140" y="620" width="400" height="50" rx="8"/>
<text class="th" x="340" y="638" text-anchor="middle" dominant-baseline="central">Select frontier pairs</text>
<text class="ts" x="340" y="655" text-anchor="middle" dominant-baseline="central">FPR_corr ≤ target at worst-case N  ·  binding at N_max</text>
</g>
<g class="node c-purple">
<rect x="140" y="686" width="400" height="50" rx="8"/>
<text class="th" x="340" y="704" text-anchor="middle" dominant-baseline="central">Validate with H0 simulations</text>
<text class="ts" x="340" y="721" text-anchor="middle" dominant-baseline="central">Sim FPR vs FPR_corr  ·  check In CI (95%)</text>
</g>
<g class="node c-green">
<rect x="140" y="752" width="400" height="50" rx="8"/>
<text class="th" x="340" y="770" text-anchor="middle" dominant-baseline="central">Optimal pair + power curve</text>
<text class="ts" x="340" y="787" text-anchor="middle" dominant-baseline="central">c1_opt, c2_opt  ·  H1 power at HR_h1 = 2.0</text>
</g>
<!-- Side notes: Cox vs IRR -->
<g class="c-coral">
<rect x="548" y="126" width="114" height="42" rx="4"/>
<text class="ts" x="605" y="143" text-anchor="middle" dominant-baseline="central">Cox PH primary</text>
<text class="ts" x="605" y="159" text-anchor="middle" dominant-baseline="central">IRR cross-check</text>
</g>
<line x1="540" y1="148" x2="548" y2="148" class="leader"/>
<!-- nsims annotations -->
<g class="c-gray">
<rect x="18" y="130" width="110" height="42" rx="4"/>
<text class="ts" x="73" y="147" text-anchor="middle" dominant-baseline="central">Prod: 4,000</text>
<text class="ts" x="73" y="163" text-anchor="middle" dominant-baseline="central">null sims</text>
</g>
<line x1="128" y1="148" x2="140" y2="148" class="leader"/>
<g class="c-gray">
<rect x="18" y="196" width="110" height="42" rx="4"/>
<text class="ts" x="73" y="213" text-anchor="middle" dominant-baseline="central">Prod: 500</text>
<text class="ts" x="73" y="229" text-anchor="middle" dominant-baseline="central">per N × 3 N</text>
</g>
<line x1="128" y1="214" x2="140" y2="214" class="leader"/>
<g class="c-gray">
<rect x="18" y="328" width="110" height="42" rx="4"/>
<text class="ts" x="73" y="345" text-anchor="middle" dominant-baseline="central">Prod: 500</text>
<text class="ts" x="73" y="361" text-anchor="middle" dominant-baseline="central">per pair × 63</text>
</g>
<line x1="128" y1="346" x2="140" y2="346" class="leader"/>
<g class="c-gray">
<rect x="18" y="694" width="110" height="42" rx="4"/>
<text class="ts" x="73" y="711" text-anchor="middle" dominant-baseline="central">Prod: 2,000</text>
<text class="ts" x="73" y="727" text-anchor="middle" dominant-baseline="central">per pair × N</text>
</g>
<line x1="128" y1="712" x2="140" y2="712" class="leader"/>
<!-- Legend -->
<g class="c-teal"><rect x="88" y="824" width="12" height="12" rx="2"/></g>
<text class="ts" x="106" y="830" dominant-baseline="central" style="fill:#444441">C3 step</text>
<g class="c-purple"><rect x="176" y="824" width="12" height="12" rx="2"/></g>
<text class="ts" x="194" y="830" dominant-baseline="central" style="fill:#444441">S2 step</text>
<g class="c-amber"><rect x="264" y="824" width="12" height="12" rx="2"/></g>
<text class="ts" x="282" y="830" dominant-baseline="central" style="fill:#444441">Check</text>
<g class="c-gray"><rect x="352" y="824" width="12" height="12" rx="2"/></g>
<text class="ts" x="370" y="830" dominant-baseline="central" style="fill:#444441">Data / note</text>
<g class="c-coral"><rect x="454" y="824" width="12" height="12" rx="2"/></g>
<text class="ts" x="472" y="830" dominant-baseline="central" style="fill:#444441">Model detail</text>
<g class="c-green"><rect x="556" y="824" width="12" height="12" rx="2"/></g>
<text class="ts" x="574" y="830" dominant-baseline="central" style="fill:#444441">Output</text>
</svg>
<figcaption style="font-size:0.82em;color:#5F5E5A;margin-top:0.4rem;text-align:center;">
Figure: Two-document survival calibration pipeline. Left annotations show recommended
production simulation budgets. The long arrow crossing the dashed line is the
<code>.rds</code> handoff.
</figcaption>
</figure>
```


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(weightedsurv)
library(data.table)
library(survival)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)

# -- Safe foreach result collection -------------------------------------------
# Replaces .errorhandling = "remove" (which silently drops failures)
# with .errorhandling = "pass" + post-hoc separation of errors from results.
collect_results <- function(raw_list, label = "") {
  is_err <- vapply(raw_list, inherits, logical(1), "error")
  n_failed <- sum(is_err)
  if (n_failed > 0) {
    err_msgs <- vapply(raw_list[is_err], conditionMessage, character(1))
    warning(sprintf("%s%d of %d tasks failed.",
        if (nzchar(label)) paste0(label, ": ") else "",
        n_failed, length(raw_list)))
    message("  First errors:\n    ",
        paste(head(unique(err_msgs), 3), collapse = "\n    "))
  }
  good <- raw_list[!is_err]
  if (length(good) == 0) {
    result <- NULL
  } else if (requireNamespace("data.table", quietly = TRUE)) {
    result <- data.table::rbindlist(good, use.names = TRUE, fill = TRUE)
    result <- as.data.frame(result)
  } else {
    result <- do.call(rbind, good)
  }
  attr(result, "n_failed") <- n_failed
  attr(result, "n_total") <- length(raw_list)
  result
}


# L_eff save/load helpers
#source(here::here("../R", "leff_calibration_io.R"))
plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")
cat("H0 sims per config:", nsims_null, "\n")
```

```{r timing-init}
timings <- list()
t_vignette_start <- proc.time()
```


# Null DGM {#sec-dgm}

```{r create-dgm-null}
t0 <- proc.time()[3]

dgm_null <- setup_gbsg_dgm(model = "null", k_treat = 1.0, verbose = TRUE)

cat("\nNull DGM:\n")
cat("  HR(ITT):", round(dgm_null$hr_causal, 3), "\n")
cat("  AHR:", round(dgm_null$AHR, 3), "\n")

timings$dgm <- proc.time()[3] - t0
```


# Part I: Per-Subgroup Approximation {#sec-approx}

Instant calculations --- no simulation required.

```{r approx-glm-fn}
#| code-fold: false

# Shorthand wrapper for compute_detection_probability_glm()
p_detect_glm <- function(theta, d_eff, c1, c2) {
  compute_detection_probability_glm(
    theta = theta, d_eff = d_eff, c1 = c1, c2 = c2,
    effect_scale = "ratio", method = "monte_carlo"
  )
}
```


## Per-Subgroup Approximation

```{r approx-grid-null}
# Reference subgroup: n_H = n.min (from setup)
n_H_dgm <- n_min_val
d_dgm   <- round(d_eff_survival(n_H_dgm, prop_cens))

approx_null <- expand.grid(c1 = c1_grid, c2 = c2_grid,
                            stringsAsFactors = FALSE)
approx_null <- approx_null[approx_null$c2 <= approx_null$c1, ]
approx_null$P1 <- mapply(function(c1v, c2v)
  round(p_detect_glm(1.0, d_dgm, c1v, c2v), 4),
  approx_null$c1, approx_null$c2)

cat(sprintf("d_eff = %d (n_H=%d, prop_cens=%.2f) | %d valid (c1,c2) pairs\n",
            d_dgm, n_H_dgm, prop_cens, nrow(approx_null)))
cat(sprintf("P1 range: %.3f -- %.3f (paper default c1=1.25, c2=1.0: %.3f)\n",
            min(approx_null$P1), max(approx_null$P1),
            approx_null$P1[approx_null$c1 == 1.25 & approx_null$c2 == 1.00]))
```


## Per-Subgroup Approximation Heatmap

```{r approx-heatmap, fig.width=8, fig.height=6}
#| fig-cap: "Per-subgroup P1(theta=1) heatmap.  This is the single-subgroup probability; the procedure-level FPR is evaluated by simulation below."

ggplot(approx_null, aes(x = factor(c1), y = factor(c2), fill = P1)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.1f%%", 100 * P1)),
            size = 3.5, colour = "black") +
  scale_fill_gradient2(
    low = "#E8F5E9", mid = "#FFF9C4", high = "#E53935",
    midpoint = 0.10, name = "P1"
  ) +
  labs(x = expression(c[1] ~ "(screening)"),
       y = expression(c[2] ~ "(consistency)"),
       title = "Per-Subgroup Approximation: P1(theta=1)",
       subtitle = sprintf("n(H)=%d, d=%d", n_H_dgm, d_dgm)) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```


# Part II: Null Simulations {#sec-sim}

## Simulation Parameters

```{r sim-params}
# Build compound parameter objects from setup-chunk scalars
fs_params_base <- list(
  hr.threshold = fs_hr_threshold,
  hr.consistency = fs_hr_consistency,
  pconsistency.threshold = pcons,
  fs.splits = fs_splits,
  n.min = n_min_val,
  d0.min = fs_d0_min, d1.min = fs_d1_min,
  maxk = maxk_val, by.risk = fs_by_risk,
  vi.grf.min = fs_vi_grf_min,
  use_lasso = TRUE, use_grf = TRUE,
  use_twostage = TRUE, twostage_args = list(),
    parallel_args = list(plan="sequential", workers = 1, show_message = FALSE)
)

fs_params_cox <- modifyList(fs_params_base, list(
  outcome.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id"
))


fs_params_irr <- modifyList(fs_params_base, list(
  outcome.name = "event_sim",   # binary event indicator (0/1)
  event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id",
  outcome_type = "binary", effect_measure = "IRR",
  offset.name = "y_sim",        # follow-up time for log(time) offset
  dmin.grf = 0.0
))

sim_config <- list(
  n_sample = n_sample,
  analysis_time = analysis_time,
  seed_base = sim_seed,
  cens_adjust = cens_adjust
)

cat("Simulation config:\n")
cat("  H0 sims:", nsims_null, "\n")
cat("  N:", sim_config$n_sample, "\n")
cat("  analysis_time:", sim_config$analysis_time, "months\n")
cat("  n.min:", fs_params_base$n.min, "(paper default)\n")
```


## Cox PH Null Simulations (Paper Defaults)

```{r run-cox-null}
cat("Running Cox H0 simulations (c1=1.25, c2=1.0)...\n")
t0 <- proc.time()[3]

results_null_cox_raw <- foreach(
  sim = seq_len(nsims_null),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE)
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id = sim, dgm = dgm_null,
    n_sample = sim_config$n_sample,
    analysis_time = sim_config$analysis_time,
    cens_adjust = sim_config$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = cox_formula_adj,
    n_add_noise = 0L,
    run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
    fs_params = fs_params_cox, verbose = FALSE
  )
}
timings$null_cox <- proc.time()[3] - t0
results_null_cox <- collect_results(results_null_cox_raw, "Cox H0")
if (nrow(results_null_cox) < nsims_null) {
  warning(sprintf(
    "Cox H0: %d of %d tasks silently failed.",
    nsims_null - nrow(results_null_cox), nsims_null
  ))
}
cat("Cox H0:", round(timings$null_cox / 60, 1), "min |",
    nrow(results_null_cox), "/", nsims_null, "rows\n")

# Reset workers to prevent memory accumulation
future::plan("sequential"); gc(verbose = FALSE)
future::plan("multisession", workers = n_workers)
```


## Poisson IRR Null Simulations (Paper Defaults)

```{r run-irr-null}
cat("Running Poisson IRR H0 simulations (c1=1.25, c2=1.0)...\n")
t0 <- proc.time()[3]

results_null_irr_raw <- foreach(
  sim = seq_len(nsims_null),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE)
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id = sim, dgm = dgm_null,
    n_sample = sim_config$n_sample,
    analysis_time = sim_config$analysis_time,
    cens_adjust = sim_config$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = cox_formula_adj,
    n_add_noise = 0L,
    run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
    fs_params = fs_params_irr, verbose = FALSE
  )
}
timings$null_irr <- proc.time()[3] - t0
results_null_irr <- collect_results(results_null_irr_raw, "IRR H0")
if (nrow(results_null_irr) < nsims_null) {
  warning(sprintf(
    "IRR H0: %d of %d tasks silently failed.",
    nsims_null - nrow(results_null_irr), nsims_null
  ))
}
cat("IRR H0:", round(timings$null_irr / 60, 1), "min |",
    nrow(results_null_irr), "/", nsims_null, "rows\n")

# Reset workers before calibration phases
future::plan("sequential"); gc(verbose = FALSE)
future::plan("multisession", workers = n_workers)
```


## Baseline H0 Comparison: Cox vs Poisson (Paper Defaults)

```{r baseline-h0}
#| tbl-cap: "H0 comparison at paper defaults (c1=1.25, c2=1.0)"

cox_h0 <- results_null_cox[results_null_cox$analysis == "FS", ]
irr_h0 <- results_null_irr[results_null_irr$analysis == "FS", ]

P1_default <- p_detect_glm(1.0, d_dgm, 1.25, 1.0)

baseline_h0 <- data.frame(
  Metric = c("Simulated FPR", "Per-subgroup P1", "Sim/P1 Ratio",
             "N sims"),
  `Cox PH` = c(
    round(mean(cox_h0$any.H), 4),
    round(P1_default, 4),
    round(mean(cox_h0$any.H) / P1_default, 2),
    nrow(cox_h0)
  ),
  `Poisson IRR` = c(
    round(mean(irr_h0$any.H), 4),
    round(P1_default, 4),
    round(mean(irr_h0$any.H) / P1_default, 2),
    nrow(irr_h0)
  ),
  check.names = FALSE
)

gt(baseline_h0) |>
  tab_header(
    title = md("**H0 Baseline: Cox vs Poisson (c1=1.25, c2=1.0)**"),
    subtitle = sprintf("N=%d, %d sims, HR(ITT)=%.2f",
      sim_config$n_sample, nsims_null,
      dgm_null$hr_causal)
  )
```


# Part III: $L_{\text{eff}}$ Calibration {#sec-leff}

## Multi-N Null Simulations

```{r null-multi-N}
n_sims_cal <- nsims_cal

cat("Running multi-N null calibration simulations...\n")
if (n_sims_cal < nsims_null) {
  cat(sprintf("  Note: using nsims_cal = %d (not nsims_null = %d).\n",
              n_sims_cal, nsims_null))
  cat("  For higher precision, increase nsims_cal in the setup chunk.\n")
  cat("  For production-grade calibration, use C2 instead.\n")
}
cat(sprintf("N values: %s | Sims per N: %d\n",
            paste(N_cal_values, collapse = ", "), n_sims_cal))
t0 <- proc.time()[3]

cal_results <- list()
for (N_val in N_cal_values) {
  t_N <- proc.time()[3]

  res_N_raw <- foreach(
    sim = seq_len(n_sims_cal),
    .errorhandling = "pass",
    .options.future = list(
      packages = c("forestsearch", "survival", "data.table", "future"),
      seed = TRUE)
  ) %dofuture% {
    future::plan("sequential")
    run_simulation_analysis(
      sim_id = sim, dgm = dgm_null,
      n_sample = N_val,
      analysis_time = sim_config$analysis_time,
      cens_adjust = sim_config$cens_adjust,
      confounders_base = confounders_base,
      cox_formula_adj = cox_formula_adj,
      n_add_noise = 0L,
      run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
      fs_params = fs_params_cox, verbose = FALSE
    )
  }

  res_N <- collect_results(res_N_raw, sprintf("cal N=%d", N_val))
  fs_rows <- res_N[res_N$analysis == "FS", ]
  if (nrow(fs_rows) < n_sims_cal) {
    warning(sprintf(
      "Calibration N=%d: %d of %d tasks silently failed.",
      N_val, n_sims_cal - nrow(fs_rows), n_sims_cal
    ))
  }
  fpr_N <- mean(fs_rows$any.H, na.rm = TRUE)

  cal_results[[as.character(N_val)]] <- data.frame(
    N = N_val, sim_fpr = fpr_N, n_sims = nrow(fs_rows)
  )

  cat(sprintf("  N = %4d: FPR = %.3f (%d sims) | %.1f sec\n",
      N_val, fpr_N, nrow(fs_rows), proc.time()[3] - t_N))
}

cal_df <- do.call(rbind, cal_results)
timings$null_cal <- proc.time()[3] - t0
cat(sprintf("Multi-N calibration: %.1f min\n", timings$null_cal / 60))

# Free workers before L_eff fitting (re-established before grid phase)
future::plan("sequential"); gc(verbose = FALSE)
```


## Implied $L_{\text{eff}}$ from Simulation

For each $N$, solve for the implied $L_{\text{eff}}$:

$$
L_{\text{eff}} =
\frac{\log(1 - \text{FPR}_{\text{sim}})}
     {\log(1 - P_1)}
$$

```{r compute-leff}
#| tbl-cap: "Implied L_eff at each N.  Consistency across threshold pairs supports L_eff as a candidate-count property."

cal_df$n_H <- n_min_val
cal_df$d_eff <- round(d_eff_survival(cal_df$n_H, prop_cens))
cal_df$P1 <- sapply(cal_df$d_eff, function(d) {
  p_detect_glm(1.0, d, 1.25, 1.0)
})
cal_df$L_eff <- log(1 - cal_df$sim_fpr) / log(1 - cal_df$P1)

cat("=== Implied L_eff ===\n")
print(cal_df[, c("N", "n_H", "d_eff", "P1", "sim_fpr", "L_eff")])
```


## Fit $L_{\text{eff}}$ Power Law

```{r calibrate-leff}
#| fig-height: 5
#| fig-cap: "Implied L_eff with power-law fit for the GBSG survival DGM."

# Fit power law
cal_valid <- cal_df[cal_df$sim_fpr > 0 & cal_df$P1 > 0 &
                     is.finite(cal_df$L_eff), ]

if (nrow(cal_valid) >= 2) {
  leff_cal <- calibrate_L_eff(
    N = cal_valid$N, P1 = cal_valid$P1,
    sim_fpr = cal_valid$sim_fpr, n_min = n_min_val
  )
  print(leff_cal)

  C_fit <- leff_cal$C
  alpha_fit <- leff_cal$alpha

  N_pred <- seq(200, 1200, by = 5)
  pred_curve <- data.frame(
    N = N_pred, L_eff = C_fit * (N_pred / n_min_val)^alpha_fit
  )

  print(
    ggplot() +
      geom_point(data = cal_valid, aes(x = N, y = L_eff),
                 size = 4, colour = "firebrick") +
      geom_line(data = pred_curve, aes(x = N, y = L_eff),
                colour = "steelblue", linewidth = 1) +
      labs(x = "Total Sample Size N",
           y = expression(L[eff]),
           title = expression(L[eff] * " Calibration: GBSG Survival DGM"),
           subtitle = sprintf("L_eff = %.3f * (N/%d)^%.3f  (R-sq = %.3f)",
                               C_fit, n_min_val, alpha_fit,
                               leff_cal$r_squared)) +
      theme_minimal(base_size = 12) +
      theme(plot.title = element_text(face = "bold"))
  )
} else {
  cat("Insufficient calibration points. Using L_eff = 1.\n")
  C_fit <- 1; alpha_fit <- 0
}

cat(sprintf("\nFitted model: L_eff = %.3f * (N/%d)^%.3f\n",
            C_fit, n_min_val, alpha_fit))
cat(sprintf("  N = 300:  L_eff = %.2f\n", C_fit * (300/n_min_val)^alpha_fit))
cat(sprintf("  N = 500:  L_eff = %.2f\n", C_fit * (500/n_min_val)^alpha_fit))
cat(sprintf("  N = 700:  L_eff = %.2f\n", C_fit * (700/n_min_val)^alpha_fit))
```


## Save Calibration

```{r save-leff}
# Save calibration to .rds for downstream documents (S2)
save_leff_calibration(
  C = C_fit,
  alpha = alpha_fit,
  n_min = n_min_val,
  outcome_type = "survival",
  cal_data = cal_df,
  dgm_description = paste(
    "GBSG survival DGM,", length(confounders_base), "confounders",
    paste0("(", paste(confounders_base, collapse = ", "), "),"),
    sprintf("prop_cens=%.2f, p_harm=%.2f", prop_cens, p_harm)),
  n_sims_per_N = n_sims_cal,
  output_dir = output_dir,
  extra = list(
    # ForestSearch config (S2 should match these)
    confounders_base = confounders_base,
    prop_cens        = prop_cens,
    p_harm           = p_harm,
    hr_h1            = hr_h1,
    pcons            = pcons,
    fs_splits        = fs_splits,
    maxk_val         = maxk_val,
    n_min_val        = n_min_val,
    analysis_time    = analysis_time,
    cens_adjust      = cens_adjust,
    fpr_target       = fpr_target,
    N_cal_values     = N_cal_values,
    # Grid range validated for L_eff threshold-independence
    c1_range = range(c1_grid),
    c2_range = range(c2_grid)
  )
)
```

::: {.callout-important}
## Shared `.rds` With C2

Both this document (C3) and `calibration_survival_fourway.qmd` (C2)
save to `calibration_survival_leff.rds`.
**C2 is the primary source** (up to 20K replicates per N).
C3 uses fewer replicates and is designed as a faster grid sweep.
If both are rendered, the last-rendered file wins.
To avoid overwriting C2's calibration, set `recalibrate = FALSE`
in this document or render C2 after C3.
:::


## Corrected Baseline

```{r corrected-baseline}
L_eff_main <- C_fit * (n_sample / n_min_val)^alpha_fit
P1_cox <- p_detect_glm(1.0, d_dgm, 1.25, 1.0)
fpr_corr <- 1 - (1 - P1_cox)^L_eff_main

cat(sprintf("=== Corrected Baseline (N=%d, c1=1.25, c2=1.0) ===\n",
            n_sample))
cat(sprintf("  L_eff:         %.2f\n", L_eff_main))
cat(sprintf("  P1 (uncorr):   %.3f\n", P1_cox))
cat(sprintf("  FPR (corr):    %.3f\n", fpr_corr))
cat(sprintf("  Sim FPR (Cox): %.3f\n", mean(cox_h0$any.H)))
cat(sprintf("  Sim FPR (IRR): %.3f\n", mean(irr_h0$any.H)))
cat(sprintf("  Sim/Corr:      %.2f\n",
            mean(cox_h0$any.H) / max(fpr_corr, 0.001)))
```


# Part IV: Null Threshold Grid {#sec-null-grid}

Systematic evaluation of $(c_1, c_2)$ pairs under $H_0$ at
$N = `r n_sample`$ to identify the region where
FPR $\le$ `r 100 * fpr_target`%.

Grid defined in the setup chunk: `c1_grid` and `c2_grid`.

## Grid Design

```{r null-grid-config}
#| code-fold: false

grid_pairs <- expand.grid(c1 = c1_grid, c2 = c2_grid,
                           stringsAsFactors = FALSE)
grid_pairs <- grid_pairs[grid_pairs$c2 <= grid_pairs$c1, ]

N_grid     <- n_sample

cat(sprintf("Null grid: %d (c1,c2) pairs x %d sims = %d tasks at N = %d\n",
    nrow(grid_pairs), nsims_grid,
    nrow(grid_pairs) * nsims_grid, N_grid))
if (nsims_grid < nsims_null) {
  cat(sprintf("  Note: using nsims_grid = %d (not nsims_null = %d).\n",
              nsims_grid, nsims_null))
  cat("  To increase grid precision, raise nsims_grid in the setup chunk.\n")
}
cat(sprintf("(c1 range: %.2f--%.2f, c2 range: %.2f--%.2f)\n",
    min(c1_grid), max(c1_grid), min(c2_grid), max(c2_grid)))
```


## Run Grid Simulations

```{r run-null-grid}
# Reset workers before grid phase
future::plan("sequential"); gc(verbose = FALSE)
future::plan("multisession", workers = n_workers)

t0 <- proc.time()[3]

grid_tasks <- do.call(rbind, lapply(seq_len(nrow(grid_pairs)), function(g) {
  data.frame(
    grid_id = g, sim = seq_len(nsims_grid),
    c1 = grid_pairs$c1[g], c2 = grid_pairs$c2[g],
    stringsAsFactors = FALSE
  )
}))

cat(sprintf("Running %d null grid tasks...\n", nrow(grid_tasks)))

grid_results_raw <- foreach(
  i = seq_len(nrow(grid_tasks)),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE)
) %dofuture% {
  future::plan("sequential")
  sim_id <- grid_tasks$sim[i]
  c1_val <- grid_tasks$c1[i]
  c2_val <- grid_tasks$c2[i]

  fs_params_grid <- modifyList(fs_params_cox, list(
    hr.threshold = c1_val,
    hr.consistency = c2_val
  ))

  res <- run_simulation_analysis(
    sim_id = sim_id, dgm = dgm_null,
    n_sample = N_grid,
    analysis_time = sim_config$analysis_time,
    cens_adjust = sim_config$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = cox_formula_adj,
    n_add_noise = 0L,
    run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
    fs_params = fs_params_grid, verbose = FALSE
  )

  fs_row <- res[res$analysis == "FS", ]
  data.frame(
    sim = sim_id, c1 = c1_val, c2 = c2_val,
    any_H = if (nrow(fs_row) > 0) fs_row$any.H[1] else 0L,
    stringsAsFactors = FALSE
  )
}

timings$null_grid <- proc.time()[3] - t0
grid_results <- collect_results(grid_results_raw, "null grid")
n_expected_grid <- nrow(grid_tasks)
n_failed_grid   <- attr(grid_results, "n_failed")
cat(sprintf("Null grid: %.1f min | %d/%d results (%d failed)\n",
    timings$null_grid / 60, nrow(grid_results), n_expected_grid, n_failed_grid))

# Free workers before sequential post-processing
future::plan("sequential"); gc(verbose = FALSE)
```


# Part V: Results {#sec-results}

## FPR Summary and Best Threshold Pairs

```{r grid-fpr-table}
grid_dt  <- data.table(grid_results)
grid_fpr <- grid_dt[, .(
  FPR    = mean(any_H, na.rm = TRUE),
  n_found = sum(any_H, na.rm = TRUE),
  n_sims  = .N
), by = .(c1, c2)]

# Approximations
grid_fpr[, P1 := mapply(function(c1v, c2v)
  p_detect_glm(1.0, d_dgm, c1v, c2v), c1, c2)]
L_eff_N <- C_fit * (N_grid / n_min_val)^alpha_fit
grid_fpr[, FPR_corr      := 1 - (1 - P1)^L_eff_N]
grid_fpr[, Sim_over_P1   := round(FPR / pmax(P1, 0.001), 2)]
grid_fpr[, Sim_over_Corr := round(FPR / pmax(FPR_corr, 0.001), 2)]

# ── Overview summary ─────────────────────────────────────────
fpr_range_str <- function(vals) {
  if (length(vals) == 0 || all(is.na(vals))) return("—")
  sprintf("%.1f–%.1f%%", 100 * min(vals, na.rm = TRUE),
                          100 * max(vals, na.rm = TRUE))
}
fpr_overview <- data.frame(
  Region = c(
    "FPR <= 10%",
    sprintf("10%% < FPR <= %.0f%%", 100 * fpr_target),
    sprintf("FPR > %.0f%%", 100 * fpr_target)
  ),
  `Pairs (n)` = c(
    sum(grid_fpr$FPR <= 0.10),
    sum(grid_fpr$FPR > 0.10 & grid_fpr$FPR <= fpr_target),
    sum(grid_fpr$FPR > fpr_target)
  ),
  `FPR range` = c(
    fpr_range_str(grid_fpr$FPR[grid_fpr$FPR <= 0.10]),
    fpr_range_str(grid_fpr$FPR[grid_fpr$FPR > 0.10 & grid_fpr$FPR <= fpr_target]),
    fpr_range_str(grid_fpr$FPR[grid_fpr$FPR > fpr_target])
  ),
  check.names = FALSE
)

gt(fpr_overview) |>
  tab_header(
    title = md("**Simulated FPR: Grid Overview**"),
    subtitle = sprintf("N=%d | %d sims/pair | n.min=%d | L_eff=%.2f",
                        N_grid, nsims_grid, n_min_val, L_eff_N)
  ) |>
  tab_style(style = list(cell_fill(color = "#E8F5E9"),
                          cell_text(weight = "bold")),
    locations = cells_body(rows = 1)) |>
  tab_style(style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(rows = 2)) |>
  tab_style(style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(rows = 3)) |>
  tab_source_note(sprintf(
    "%d total (c1,c2) pairs | target FPR <= %.0f%% | paper defaults c1=1.25, c2=1.0",
    nrow(grid_fpr), 100 * fpr_target))

# ── Best pairs: lowest simulated FPR ─────────────────────────
n_best <- min(10L, nrow(grid_fpr))
best_fpr <- grid_fpr[order(FPR)][seq_len(n_best)]

gt(best_fpr[, .(c1, c2,
    `Sim FPR`  = sprintf("%.1f%%", 100 * FPR),
    `P1`       = sprintf("%.1f%%", 100 * P1),
    `FPR_corr` = sprintf("%.1f%%", 100 * FPR_corr),
    `Sim/Corr` = Sim_over_Corr,
    fpr_num    = FPR)]) |>    # keep numeric for row predicate, hidden below
  tab_header(
    title = md("**Best Threshold Pairs (Lowest Sim FPR)**"),
    subtitle = sprintf(
      "Top %d of %d pairs | paper defaults (c1=1.25, c2=1.0) shown for reference",
      n_best, nrow(grid_fpr))
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `Sim FPR`,
      rows = fpr_num <= fpr_target)) |>
  tab_style(style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(columns = `Sim FPR`,
      rows = fpr_num > fpr_target & fpr_num <= fpr_target + 0.05)) |>
  cols_hide(columns = fpr_num) |>
  tab_source_note(sprintf(
    "Green: FPR <= %.0f%% (target). Yellow: within 5pp of target.",
    100 * fpr_target))
```


## $L_{\text{eff}}$ Consistency Across Threshold Pairs

If $L_{\text{eff}}$ is truly a property of the candidate-generation
process and not the threshold pair, the implied $L_{\text{eff}}$
should be approximately constant across $(c_1, c_2)$ at the same $N$.
A coefficient of variation (CV) below 20% is acceptable.

```{r leff-consistency}
#| fig-height: 4
#| fig-cap: "Implied vs calibrated L_eff across all (c1, c2) pairs. Points near the horizontal line support threshold-independence."

grid_fpr[, implied_L_eff := ifelse(
  P1 > 0 & FPR > 0 & FPR < 1,
  log(1 - FPR) / log(1 - P1), NA_real_)]

leff_valid <- grid_fpr[!is.na(implied_L_eff)]
cv_val <- 100 * sd(leff_valid$implied_L_eff) /
           median(leff_valid$implied_L_eff)

cat(sprintf("Calibrated L_eff at N=%d: %.2f\n", N_grid, L_eff_N))
cat(sprintf("Implied L_eff: median=%.2f, range=[%.2f, %.2f], CV=%.0f%%\n",
    median(leff_valid$implied_L_eff),
    min(leff_valid$implied_L_eff),
    max(leff_valid$implied_L_eff), cv_val))
if (cv_val > 30) {
  warning(sprintf(
    "CV = %.0f%% > 30%% — L_eff may not be threshold-independent at this grid.",
    cv_val))
} else {
  cat(sprintf("CV = %.0f%% (< 30%%) — threshold-independence assumption holds.\n",
              cv_val))
}

# Scatter: implied vs calibrated, coloured by Sim FPR
ggplot(leff_valid, aes(x = interaction(factor(c1), factor(c2), sep=","),
                        y = implied_L_eff, colour = FPR)) +
  geom_hline(yintercept = L_eff_N, linetype = "dashed",
             colour = "steelblue", linewidth = 0.9) +
  geom_point(size = 2.5, alpha = 0.8) +
  annotate("label", x = 1, y = L_eff_N,
           label = sprintf("Calibrated = %.2f", L_eff_N),
           colour = "steelblue", size = 3, hjust = 0, vjust = -0.3) +
  scale_colour_gradient(low = "#E8F5E9", high = "#E53935",
                         name = "Sim FPR", labels = scales::percent) +
  labs(x = "(c1, c2) pair",
       y = expression("Implied " * L[eff]),
       title = expression(L[eff] * " Consistency Across Threshold Grid"),
       subtitle = sprintf("CV = %.0f%% | Calibrated = %.2f | N = %d",
                           cv_val, L_eff_N, N_grid)) +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(face = "bold"),
        legend.position = "right")
```

## FPR Heatmap

```{r grid-heatmap, fig.width=8, fig.height=6}
#| fig-cap: !expr 'sprintf("Simulated FPR heatmap by (c1, c2).  Values <= %.0f%% are in the acceptable region.", 100 * fpr_target)'

ggplot(grid_fpr, aes(x = factor(c1), y = factor(c2), fill = FPR)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.1f%%", 100 * FPR)),
            size = 3.5, colour = "black") +
  scale_fill_gradient2(
    low = "#E8F5E9", mid = "#FFF9C4", high = "#E53935",
    midpoint = fpr_target,
    limits = c(0, max(grid_fpr$FPR, 0.20)),
    name = "FPR"
  ) +
  labs(x = expression(c[1] ~ "(screening threshold)"),
       y = expression(c[2] ~ "(consistency threshold)"),
       title = "Simulated FPR Under H0: GBSG Survival DGM",
       subtitle = sprintf("N=%d, %d sims/pair, n.min=%d",
                           N_grid, nsims_grid, n_min_val)) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        axis.text = element_text(size = 11))
```


## FPR Sensitivity to $c_2$ (Fixing $c_1$)

```{r grid-c2-profiles, fig.width=9, fig.height=5}
#| fig-cap: "FPR vs c2 for each fixed c1.  The target band is highlighted."

ggplot(grid_fpr, aes(x = c2, y = FPR,
                      colour = c1, group = factor(c1))) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0.10, linetype = "dashed", colour = "darkgreen") +
  geom_hline(yintercept = fpr_target, linetype = "dashed", colour = "orange") +
  annotate("rect", xmin = -Inf, xmax = Inf,
           ymin = 0.10, ymax = fpr_target,
           alpha = 0.15, fill = "gold") +
  scale_colour_viridis_c(name = expression(c[1]), option = "D") +
  labs(x = expression(c[2] ~ "(consistency threshold)"),
       y = "Simulated FPR",
       title = "FPR Sensitivity to Consistency Threshold",
       subtitle = sprintf("N=%d, each line is a fixed c1", N_grid)) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


## FPR Sensitivity to $c_1$ (Fixing $c_2$)

```{r grid-c1-profiles, fig.width=9, fig.height=5}
#| fig-cap: "FPR vs c1 for each fixed c2.  Shows the screening threshold's independent effect."

ggplot(grid_fpr, aes(x = c1, y = FPR,
                      colour = c2, group = factor(c2))) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0.10, linetype = "dashed", colour = "darkgreen") +
  geom_hline(yintercept = fpr_target, linetype = "dashed", colour = "orange") +
  annotate("rect", xmin = -Inf, xmax = Inf,
           ymin = 0.10, ymax = fpr_target,
           alpha = 0.15, fill = "gold") +
  scale_colour_viridis_c(name = expression(c[2]), option = "C") +
  labs(x = expression(c[1] ~ "(screening threshold)"),
       y = "Simulated FPR",
       title = "FPR Sensitivity to Screening Threshold",
       subtitle = sprintf("N=%d, each line is a fixed c2", N_grid)) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


## Approximation Accuracy Across the Grid

```{r grid-approx-scatter, fig.width=8, fig.height=6}
#| fig-cap: "Approximation vs simulation across the (c1,c2) grid.  Corrected points (triangles) should cluster near the diagonal."

ggplot(grid_fpr, aes(x = FPR)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              colour = "grey50") +
  geom_point(aes(y = P1, colour = "Uncorrected P1"),
             size = 2.5, alpha = 0.7) +
  geom_point(aes(y = FPR_corr, colour = "L_eff Corrected"),
             size = 2.5, alpha = 0.7, shape = 17) +
  scale_colour_manual(values = c("Uncorrected P1" = "firebrick",
                                  "L_eff Corrected" = "steelblue")) +
  coord_equal(xlim = c(0, max(grid_fpr$FPR, 0.25)),
              ylim = c(0, max(grid_fpr$P1, 0.25))) +
  labs(x = "Simulated FPR", y = "Approximated FPR",
       title = "Approximation Accuracy Across Threshold Grid",
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Acceptable Region (FPR $\le$ `r 100 * fpr_target`%)

```{r grid-acceptable}
#| tbl-cap: !expr 'sprintf("Threshold pairs with simulated FPR <= %.0f%%", 100 * fpr_target)'

acceptable_grid <- grid_fpr[FPR <= fpr_target]
setorder(acceptable_grid, FPR)

if (nrow(acceptable_grid) > 0) {
  gt_acc <- gt(acceptable_grid[, .(c1, c2,
    `Sim FPR` = sprintf("%.1f%%", 100 * FPR),
    `P1 (uncorr)` = sprintf("%.1f%%", 100 * P1),
    `FPR (corr)` = sprintf("%.1f%%", 100 * FPR_corr))]) |>
    tab_header(
      title = md(sprintf(
        "**Acceptable Threshold Configurations (FPR <= %.0f%%)**",
        100 * fpr_target)),
      subtitle = sprintf("N=%d, GBSG survival DGM, n.min=%d",
                          N_grid, n_min_val)
    ) |>
    tab_source_note("Paper defaults: c1=1.25, c2=1.0")
  knitr::asis_output(gt::as_raw_html(gt_acc))
} else {
  cat(sprintf("No configurations achieved FPR <= %.0f%% in this grid.\n",
              100 * fpr_target))
}
```


# Summary {#sec-summary}

```{r summary}
# ── Key configurations ────────────────────────────────────────
paper_row   <- grid_fpr[c1 == 1.25 & c2 == 1.00]
target_band <- grid_fpr[FPR >= 0.08 & FPR <= fpr_target]
setorder(target_band, -FPR)
conservative <- grid_fpr[FPR <= 0.10]
setorder(conservative, c1, c2)

# Build display table: paper default + target band + conservative
make_row <- function(dt, label) {
  if (nrow(dt) == 0) return(NULL)
  dt[, .(
    Region   = label,
    c1       = c1,
    c2       = c2,
    `Sim FPR`   = sprintf("%.1f%%", 100 * FPR),
    `P1`        = sprintf("%.1f%%", 100 * P1),
    `FPR_corr`  = sprintf("%.1f%%", 100 * FPR_corr),
    `Sim/Corr`  = round(Sim_over_Corr, 2),
    fpr_num     = FPR   # numeric column for gt row predicates
  )]
}

summary_dt <- rbind(
  make_row(paper_row,   "Paper default"),
  make_row(target_band, sprintf("Target band (8-%d%%)", round(100 * fpr_target))),
  make_row(conservative,"Conservative (FPR <= 10%)")
)

if (is.null(summary_dt) || nrow(summary_dt) == 0) {
  cat("No rows met any of the three FPR criteria at this grid resolution.\n")
} else {
  target_label <- sprintf("Target band (8-%d%%)", round(100 * fpr_target))
  gt_out <- gt(summary_dt) |>
    tab_header(
      title = md("**Null Operating Characteristics Summary**"),
      subtitle = sprintf(
        "GBSG survival DGM | N=%d | n.min=%d | pcons=%.2f | L_eff=%.2f",
        N_grid, n_min_val, pcons, L_eff_N)
    ) |>
    tab_row_group(label = "Conservative (FPR \u2264 10%)",
      rows = Region == "Conservative (FPR <= 10%)") |>
    tab_row_group(label = sprintf("Target band (8\u2013%d%%)", round(100*fpr_target)),
      rows = Region == target_label) |>
    tab_row_group(label = "Paper default (c1=1.25, c2=1.0)",
      rows = Region == "Paper default") |>
    cols_hide(columns = c(Region, fpr_num)) |>
    tab_style(style = cell_fill(color = "#E8F5E9"),
      locations = cells_body(columns = `Sim FPR`,
        rows = fpr_num <= fpr_target)) |>
    tab_source_note(paste(
      "L_eff note: for survival, L_eff < 1 is common because events-based filters",
      "(d0.min, d1.min) aggressively prune candidates,",
      "making paper defaults already conservative for survival endpoints."))
  knitr::asis_output(gt::as_raw_html(gt_out))
}
```


# Save All Simulation Results {#sec-save}

```{r save-all-results}
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

sim_bundle <- list(

  # Metadata
  params = list(
    nsims_null = nsims_null,
    nsims_grid = nsims_grid,
    n_sims_cal = nsims_cal,
    sim_seed = sim_seed,
    N_grid = N_grid,
    N_cal_values = N_cal_values,
    sim_config = sim_config,
    fs_params_base = fs_params_base,
    fs_params_cox = fs_params_cox,
    fs_params_irr = fs_params_irr,
    c1_grid = c1_grid,
    c2_grid = c2_grid,
    n_min_val = n_min_val,
    prop_cens = prop_cens,
    p_harm = p_harm,
    hr_h1 = hr_h1,
    fpr_target = fpr_target,
    timestamp = Sys.time(),
    R_version = R.version.string,
    forestsearch_version = tryCatch(
      as.character(utils::packageVersion("forestsearch")),
      error = function(e) "unknown")
  ),

  # DGM
  dgm_null = dgm_null,

  # Paper-defaults null sims (Part II)
  results_null_cox = results_null_cox,
  results_null_irr = results_null_irr,

  # L_eff calibration (Part III)
  leff = list(
    C = C_fit,
    alpha = alpha_fit,
    n_min = n_min_val,
    cal_df = cal_df
  ),

  # Grid sweep results (Part IV)
  grid_results = grid_results,
  grid_fpr = grid_fpr,

  # Timing
  timings = timings
)

rds_path <- file.path(output_dir,
  sprintf("calibration_survival_leff_grid_%s.rds",
          format(Sys.Date(), "%Y%m%d")))
saveRDS(sim_bundle, rds_path)

cat(sprintf("=== All results saved ===\n"))
cat(sprintf("  File: %s\n", rds_path))
cat(sprintf("  Size: %.1f MB\n", file.size(rds_path) / 1024^2))
cat(sprintf("  Contents:\n"))
cat(sprintf("    results_null_cox: %d rows\n", nrow(results_null_cox)))
cat(sprintf("    results_null_irr: %d rows\n", nrow(results_null_irr)))
cat(sprintf("    cal_df:           %d rows (N = %s)\n",
            nrow(cal_df), paste(cal_df$N, collapse = ", ")))
cat(sprintf("    grid_results:     %d rows (%d configs x %d sims)\n",
            nrow(grid_results), nrow(grid_pairs), nsims_grid))
cat(sprintf("    grid_fpr:         %d (c1,c2) pairs\n", nrow(grid_fpr)))
cat(sprintf("    L_eff:            C=%.4f, alpha=%.4f\n",
            C_fit, alpha_fit))
```


## How to Reload

```{r reload-example}
#| eval: false
#| code-fold: false

res <- readRDS(latest_rds("calibration_survival_leff_grid"))

# Paper-defaults FPR
cox_h0 <- res$results_null_cox[res$results_null_cox$analysis == "FS", ]
cat("Cox H0 FPR:", mean(cox_h0$any.H), "\n")

# L_eff calibration
res$leff$C            # fitted C
res$leff$alpha        # fitted alpha
res$leff$cal_df       # raw calibration points

# Grid sweep
res$grid_fpr          # FPR by (c1, c2)
res$grid_fpr[FPR <= res$params$fpr_target]   # acceptable configs

# Config
res$params$nsims_null       # how many sims
res$params$c1_grid          # which thresholds swept
res$params$fpr_target       # FPR target
```


# Timing and Session Info {#sec-session}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

# ── Per-call wall-clock rates ─────────────────────────────────
# All times are already parallelised wall-clock seconds.
n_tasks_cox  <- nsims_null
n_tasks_irr  <- nsims_null
n_tasks_cal  <- nsims_cal * length(N_cal_values)
n_tasks_grid <- nsims_grid * nrow(grid_pairs)

rate_cox  <- timings$null_cox  / n_tasks_cox
rate_irr  <- timings$null_irr  / n_tasks_irr
rate_cal  <- timings$null_cal  / n_tasks_cal
rate_grid <- timings$null_grid / n_tasks_grid

# ── Actual run ────────────────────────────────────────────────
timing_df <- data.frame(
  Stage = c(
    sprintf("DGM setup"),
    sprintf("Cox H0 — paper defaults (%d sims)", nsims_null),
    sprintf("IRR H0 — paper defaults (%d sims)", nsims_null),
    sprintf("L_eff calibration (%d sims × %d N)", nsims_cal, length(N_cal_values)),
    sprintf("Null grid sweep (%d sims × %d pairs)", nsims_grid, nrow(grid_pairs)),
    "Total (including rendering)"
  ),
  Seconds = c(
    timings$dgm, timings$null_cox, timings$null_irr,
    timings$null_cal, timings$null_grid, timings$total),
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60
timing_df$Hours   <- timing_df$Seconds / 3600

# ── Per-call rates row ────────────────────────────────────────
rates_row <- data.frame(
  Stage   = sprintf("Wall-clock per FS call (%d workers): Cox %.2fs | IRR %.2fs | Grid %.2fs",
                     n_workers, rate_cox, rate_irr, rate_grid),
  Seconds = NA_real_, Minutes = NA_real_, Hours = NA_real_,
  stringsAsFactors = FALSE)

# ── Production projection ─────────────────────────────────────
proj_cox   <- rate_cox  * nsims_prod_null
proj_irr   <- rate_irr  * nsims_prod_null
proj_cal   <- rate_cal  * nsims_prod_cal * length(N_cal_values)
proj_grid  <- rate_grid * nsims_prod_grid * nrow(grid_pairs)
proj_total <- proj_cox + proj_irr + proj_cal + proj_grid + timings$dgm

prod_row <- data.frame(
  Stage   = sprintf("Production estimate: %d null / %d cal / %d grid  (%d pairs)",
                     nsims_prod_null, nsims_prod_cal, nsims_prod_grid, nrow(grid_pairs)),
  Seconds = proj_total,
  Minutes = proj_total / 60,
  Hours   = proj_total / 3600,
  stringsAsFactors = FALSE)

timing_df <- rbind(timing_df, rates_row, prod_row)

gt(timing_df) |>
  tab_header(
    title = "Computational Timing Summary",
    subtitle = sprintf(
      "%d workers | grid: %d pairs × %d sims | cal: %d sims × %d N",
      n_workers, nrow(grid_pairs), nsims_grid,
      nsims_cal, length(N_cal_values))
  ) |>
  fmt_number(columns = Seconds, decimals = 1, rows = !is.na(Seconds)) |>
  fmt_number(columns = Minutes, decimals = 2, rows = !is.na(Minutes)) |>
  fmt_number(columns = Hours,   decimals = 3, rows = !is.na(Hours))   |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)",
             Minutes = "Time (min)", Hours = "Time (hr)") |>
  tab_style(style = cell_text(weight = "bold"),
    locations = cells_body(rows = grepl("Total|Production", Stage))) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = grepl("Production", Stage))) |>
  tab_style(style = cell_fill(color = "#F1EFE8"),
    locations = cells_body(rows = grepl("Wall-clock", Stage))) |>
  tab_source_note(sprintf(
    "Production estimate: %d null + %d irr + %d cal×%dN + %d grid×%d pairs | same machine, same workers.",
    nsims_prod_null, nsims_prod_null, nsims_prod_cal, length(N_cal_values),
    nsims_prod_grid, nrow(grid_pairs)))
```

```{r session-info}
plan(sequential)
sessionInfo()
```


# Appendix: Simulation Budget Guidance {#sec-simbudget .appendix}

The three simulation phases have independent precision requirements.
The table below summarises the 95% confidence interval half-width
for a simulated FPR estimate as a function of the number of
replicates, together with the implied total task counts and
recommended production targets.

```{r sim-budget}
#| tbl-cap: "Simulation budget guidance: 95% CI half-width for simulated FPR by phase and replicate count."
#| code-fold: true

# ── Helper ────────────────────────────────────────────────────
ci_hw <- function(fpr, n, z = 1.96)
  z * sqrt(fpr * (1 - fpr) / n)

se_leff <- function(fpr, P1, n) {
  se_f <- sqrt(fpr * (1 - fpr) / n)
  se_f / (P1 * abs(log(1 - P1)))
}

cv_leff <- function(fpr, P1, n) {
  L <- log(1 - fpr) / log(1 - P1)
  100 * se_leff(fpr, P1, n) / L
}

# ── Parameters from this document ────────────────────────────
P1_ref   <- p_detect_glm(1.0, d_dgm, 1.25, 1.0)   # P1 at paper defaults
n_pairs  <- nrow(grid_pairs)                          # 63 pairs

# ── Part II: baseline null sims ───────────────────────────────
fpr_base <- c(mean(cox_h0$any.H), fpr_target)         # actual + target
fpr_ii   <- mean(fpr_base)                            # representative value

budget_ii <- data.frame(
  `nsims_null` = c(500L, 1000L, 2000L, 4000L),
  `95% CI ±(pp)` = round(100 * sapply(c(500,1000,2000,4000),
                    function(n) ci_hw(fpr_ii, n)), 1),
  check.names = FALSE
)

# ── Part III: L_eff calibration ───────────────────────────────
# Representative FPR at N=500 (middle cal point)
fpr_cal <- if (exists("cal_df") && nrow(cal_df) > 0)
  cal_df$sim_fpr[which.min(abs(cal_df$N - 500))] else 0.07

budget_iii <- data.frame(
  `nsims_cal` = c(200L, 500L, 1000L, 2000L),
  `CV of L_eff (%)` = round(sapply(c(200,500,1000,2000),
                       function(n) cv_leff(fpr_cal, P1_ref, n)), 0),
  check.names = FALSE
)

# ── Part IV: grid sweep ────────────────────────────────────────
budget_iv <- data.frame(
  `nsims_grid` = c(500L, 1000L, 2000L, 4000L),
  `95% CI ±(pp) @ FPR=10%` = round(100 * sapply(c(500,1000,2000,4000),
                              function(n) ci_hw(0.10, n)), 1),
  `95% CI ±(pp) @ FPR=5%`  = round(100 * sapply(c(500,1000,2000,4000),
                              function(n) ci_hw(0.05, n)), 1),
  `Total tasks (×%d pairs)` = c(500,1000,2000,4000) * n_pairs,
  check.names = FALSE
)
names(budget_iv)[4] <- sprintf("Total tasks (×%d pairs)", n_pairs)

# ── Recommended settings ──────────────────────────────────────
recs <- data.frame(
  Phase = c(
    "Part II — Baseline null (Cox + IRR)",
    "Part III — L_eff calibration",
    "Part IV — Grid sweep"
  ),
  Parameter = c("nsims_null", "nsims_cal", "nsims_grid"),
  `Current (actual)` = c(nsims_null, nsims_cal, nsims_grid),
  `Minimum` = c(2000L, 1000L, 2000L),
  `Production` = c(4000L, 2000L, 2000L),
  `Precision at production` = c(
    sprintf("±%.1fpp at FPR=%.0f%%",
            100 * ci_hw(fpr_ii, 4000), 100 * fpr_ii),
    sprintf("CV~%.0f%% on L_eff",
            cv_leff(fpr_cal, P1_ref, 2000)),
    sprintf("±%.1fpp at FPR=10%%",
            100 * ci_hw(0.10, 2000))
  ),
  check.names = FALSE
)

# ── Render tables ─────────────────────────────────────────────
tbl_recs <- gt(recs) |>
  tab_header(
    title = "Recommended Simulation Budgets",
    subtitle = sprintf(
      "GBSG survival DGM | %d cal N values | %d grid pairs | %d workers",
      length(N_cal_values), n_pairs, n_workers)
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(
      columns = Production,
      rows = `Current (actual)` >= Minimum)) |>
  tab_style(style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(
      columns = `Current (actual)`,
      rows = `Current (actual)` < Minimum)) |>
  tab_source_note(paste(
    "Green: current setting meets or exceeds minimum.",
    "Red: current setting is below minimum — raise before production run.",
    sprintf("Part III note: C2 (calibration_survival_fourway.qmd) runs up to 20K sims/N",
            ""))) |>
  cols_label(`Current (actual)` = "Current")

knitr::asis_output(gt::as_raw_html(tbl_recs))

cat("\n")

# Detail tables
tbl_ii <- gt(budget_ii) |>
  tab_header(
    title = "Part II — Baseline Null Precision",
    subtitle = sprintf("Single threshold pair | representative FPR = %.1f%%",
                        100 * fpr_ii)
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = `nsims_null` >= 2000L)) |>
  tab_source_note("Green: meets minimum. 2,000 = minimum; 4,000 = production.")

tbl_iii <- gt(budget_iii) |>
  tab_header(
    title = "Part III — L_eff Calibration Precision",
    subtitle = sprintf(
      "3-point power-law fit | P1=%.3f at paper defaults | FPR~%.1f%% at N=500",
      P1_ref, 100 * fpr_cal)
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = `nsims_cal` >= 1000L)) |>
  tab_source_note(paste(
    "CV is high because L_eff < 1 for survival (events-based pruning).",
    "C2 is the primary calibration source; C3 is a cross-check.",
    "1,000 = minimum standalone; 2,000 = production standalone."))

tbl_iv <- gt(budget_iv) |>
  tab_header(
    title = "Part IV — Grid Sweep Precision",
    subtitle = sprintf(
      "%d threshold pairs | boundary detection requires CI < 2-3pp gap",
      n_pairs)
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = `nsims_grid` >= 2000L)) |>
  tab_source_note(
    "2,000 sims/pair: ±1.3pp at FPR=10% — adequate for boundary classification.")

knitr::asis_output(gt::as_raw_html(tbl_ii))
cat("\n")
knitr::asis_output(gt::as_raw_html(tbl_iii))
cat("\n")
knitr::asis_output(gt::as_raw_html(tbl_iv))
```
