---
title: "Direct FPR Calibration at a Specific Threshold Pair"
subtitle: "fpr_calibration(): a simulation-based alternative to the L_eff power law"
author: "Larry F. Leon"
date: today
format:
  html:
    code-fold: show
    toc: true
    toc-depth: 3
execute:
  warning: true
  message: true
  fig-align: center
knitr:
  opts_chunk:
    comment: "#>"
---


# Motivation {#sec-motivation}

The $L_{\text{eff}}$ power-law correction relies on
*threshold-independence*: the implied effective candidate count must be
approximately constant across $(c_1, c_2)$ at a given $N$.  In practice
this assumption fails for binary outcomes with small $d_{\text{eff}}$
(e.g.\ $n_H = 60$, $p = 0.30 \Rightarrow d_{\text{eff}} = 12.6$), where
the calibration grid is saturated at low thresholds and genuinely
discriminating at the operational thresholds used for selection.  The
result is an overestimated correction and a corrected FPR that is
systematically too high.

`fpr_calibration(c1, c2, ...)` takes a different approach: **estimate
the procedure-level FPR directly at the analyst's specific threshold
pair**, via a small null simulation.  No power-law model and no
threshold-independence assumption are required.


## Two-tier workflow

| Tier | `nsims` | Precision (FPR ≈ 10%) | Purpose | Typical runtime |
|------|---------|----------------------|---------|----------------|
| **1 — Approximation** | 200–500 | ±1–2 pp | Pre-screening; check a candidate (c1, c2) before committing | 2–10 min |
| **2 — Verification** | 10,000–20,000 | ±0.2–0.3 pp | Final confirmation of the selected operating point | 1–3 hr |

The function supports both tiers with the same call signature.
`nsims = 200` is the default (Tier 1).


## Two null-generation modes

**Permutation mode** (default when `null_fn = NULL`): shuffles the
treatment column in the analyst's actual dataset.  Preserves the
observed covariate and outcome distributions — a model-free H0 reference
for a real analysis.

**DGM mode** (`null_fn = function(seed) ...`): draws fresh null datasets
from a parametric generator.  Used in simulation studies or when the
analyst wants to verify at a different $N$ than the observed data.


# Setup {#sec-setup}

```{r packages}
library(forestsearch)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)

#source("fpr_calibration.R")   # wrapper (not yet exported from the package)

n_workers <- floor(0.95*max(1L, parallel::detectCores(logical = FALSE) - 1L))
plan(multisession, workers = n_workers)
cat("Workers:", n_workers, "\n")
```

```{r demo-params}
#| include: false

# ── Simulation counts ───────────────────────────────────────────
# Increase these to improve precision. Timings scale linearly.
#
#   nsims_tier1:   calibration + verify sims for Tier 1 calls
#   nsims_tier2:   calibration sims for Tier 2 (higher-precision) call
#   nsims_verify:  independent verification sims added to every call
#                  Set to 0 to skip verification entirely.
#                  At FPR ~ 10%: 200 → ±2.1 pp CI; 500 → ±1.3 pp; 1000 → ±0.9 pp
#
nsims_tier1  <- 500L    # Tier 1 calibration sims
nsims_tier2  <- 20000L   # Tier 2 calibration sims (use 10,000–20,000 for final)
nsims_verify <- 10000L    # Independent verification sims (same value used everywhere)

sim_seed     <- 8316951L

# ── Binary DGM parameters ───────────────────────────────────────
N_demo    <- 500L    # sample size
p_event   <- 0.30   # marginal event probability
n_min_val <- 60L    # n.min for ForestSearch and d_eff approximation

# ── Survival DGM parameters ─────────────────────────────────────
N_surv           <- 700L
analysis_time    <- 84L          # GBSG analysis time (months)
cens_adjust      <- log(1.5)
prop_cens_surv   <- 0.46
confounders_surv <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")

# ── ForestSearch shared parameters ──────────────────────────────
pcons     <- 0.90
maxk_val  <- 2L
fs_splits <- 400L

# ── FPR targets (for plot reference lines) ──────────────────────
fpr_target_bin  <- 0.15   # binary acceptable FPR
fpr_target_cox  <- 0.12   # survival acceptable FPR

# ── Comparison grid (Section: fpr_calibration vs L_eff power law) ──
c1_seq <- seq(1.50, 2.50, by = 0.20)   # symmetric pairs (c1 = c2)

# ── Internals (do not edit) ─────────────────────────────────────
timings     <- list()
t_doc_start <- proc.time()
```


# Example 1: Binary Outcome (Logistic, OR) {#sec-binary}

## DGM and ForestSearch configuration

```{r binary-dgm}
#| code-fold: false

# Null DGM: no treatment effect, 4 prognostic confounders
sim_null_binary <- function(n, seed) {
  set.seed(seed)
  df <- data.frame(
    id    = seq_len(n),
    treat = rbinom(n, 1, 0.5),
    bm1   = as.factor(rbinom(n, 1, 0.70)),
    bm2   = as.factor(rbinom(n, 1, 0.50)),
    age   = round(rnorm(n, 55, 10)),
    ecog  = as.factor(sample(0:1, n, replace = TRUE, prob = c(0.6, 0.4)))
  )
  lp      <- with(df, qlogis(p_event) + 0.02 * (age - 55) + 0.3 * (ecog == 1))
  df$y    <- rbinom(n, 1, plogis(lp))
  df
}

# ForestSearch parameters (fixed for all calibration calls)
fs_binary <- list(
  confounders.name       = c("bm1", "bm2", "age", "ecog"),
  outcome.name           = "y",
  event.name             = "y",
  treat.name             = "treat",
  id.name                = "id",
  outcome_type           = "binary",
  effect_measure         = "OR",
  adverse_outcome        = TRUE,
  pconsistency.threshold = pcons,
  fs.splits              = fs_splits,
  n.min                  = n_min_val,
  maxk                   = maxk_val,
  use_lasso              = TRUE,
  use_grf                = TRUE,
  use_twostage           = TRUE,
  is.RCT                 = TRUE,
  details                = FALSE,
  quiet                  = TRUE,
  plot.sg                = FALSE,
  parallel_args          = list(plan = "sequential", workers = 1,
                                show_message = FALSE)
)

# Generate the analyst's observed dataset
df_obs_bin <- sim_null_binary(N_demo, sim_seed)
cat(sprintf("Observed dataset: N = %d | events = %d (%.0f%%)\n",
            nrow(df_obs_bin), sum(df_obs_bin$y), 100 * mean(df_obs_bin$y)))
```


## Tier 1 — quick approximation at two candidate pairs

The analyst has a preliminary shortlist: (c1=1.70, c2=1.70) and
(c1=2.10, c2=1.40).  Each call takes ~2 minutes on a laptop.

```{r binary-tier1}
#| code-fold: false

t0 <- proc.time()[3]
# ── Pair A: c1=1.70, c2=1.70 ──────────────────────────────────
cal_bin_A_t1 <- fpr_calibration(
  c1 = 1.70, c2 = 1.70,
  null_fn  = function(seed) sim_null_binary(N_demo, seed),
  n_sample = N_demo,
  fs_params = fs_binary,
  nsims        = nsims_tier1,
  nsims_verify = nsims_verify,
  n_min     = n_min_val,
  outcome_type_approx = "binary",
  p_event   = p_event,
  sim_seed  = sim_seed,
  n_workers = n_workers
)

# ── Pair B: c1=2.10, c2=1.40 ──────────────────────────────────
cal_bin_B_t1 <- fpr_calibration(
  c1 = 2.10, c2 = 1.40,
  null_fn  = function(seed) sim_null_binary(N_demo, seed),
  n_sample = N_demo,
  fs_params = fs_binary,
  nsims        = nsims_tier1,
  nsims_verify = nsims_verify,
  n_min     = n_min_val,
  outcome_type_approx = "binary",
  p_event   = p_event,
  sim_seed  = sim_seed,
  n_workers = n_workers
)

print(cal_bin_A_t1)
print(cal_bin_B_t1)
timings$bin_tier1 <- proc.time()[3] - t0
cat(sprintf("Binary Tier 1 (2 pairs × %d sims): %.1f min\n",
            nsims_tier1, timings$bin_tier1 / 60))
```


## Tier 2 — higher-precision verification of the preferred pair

After Tier 1 screening, suppose the analyst prefers Pair A.
Re-run with more simulations to confirm.

```{r binary-tier2}

t0 <- proc.time()[3]
cal_bin_A_t2 <- fpr_calibration(
  c1 = 1.70, c2 = 1.70,
  null_fn  = function(seed) sim_null_binary(N_demo, seed),
  n_sample = N_demo,
  fs_params = fs_binary,
  nsims        = nsims_tier2,
  nsims_verify = nsims_verify,
  n_min     = n_min_val,
  outcome_type_approx = "binary",
  p_event   = p_event,
  sim_seed  = sim_seed + 1000L,
  n_workers = n_workers
)

print(cal_bin_A_t2)
timings$bin_tier2 <- proc.time()[3] - t0
cat(sprintf("Binary Tier 2 (1 pair × %d sims): %.1f min\n",
            nsims_tier2, timings$bin_tier2 / 60))
```


## Permutation mode with the actual dataset

In a real analysis the analyst uses `df.analysis` directly.
Treatment is shuffled; no parametric DGM is required.

```{r binary-permutation}

t0 <- proc.time()[3]
cal_bin_perm <- fpr_calibration(
  c1          = 1.70,
  c2          = 1.70,
  df.analysis = df_obs_bin,
  fs_params   = fs_binary,
  nsims        = nsims_tier1,
  nsims_verify = nsims_verify,
  n_min       = n_min_val,
  outcome_type_approx = "binary",
  p_event     = p_event,
  sim_seed    = sim_seed + 2000L,
  n_workers   = n_workers
)

print(cal_bin_perm)
timings$bin_perm <- proc.time()[3] - t0
cat(sprintf("Binary permutation (1 pair × %d sims): %.1f min\n",
            nsims_tier1, timings$bin_perm / 60))
```


## Summary table: binary results

```{r binary-summary-table}
#| tbl-cap: "fpr_calibration results for binary outcome. Tier 1 = quick approximation; Tier 2 = higher-precision verification; Perm = permutation of observed data."

make_row <- function(res, label) {
  fpr_corr <- if (!is.na(res$L_eff_implied))
    1 - (1 - res$P1)^res$L_eff_implied else NA_real_
  data.table(
    Label              = label,
    c1                 = res$c1,
    c2                 = res$c2,
    Mode               = res$mode,
    `Cal sims`         = res$nsims,
    `Sim FPR (cal)`    = sprintf("%.1f%%", 100 * res$fpr),
    `95% CI`           = sprintf("[%.1f%%, %.1f%%]",
                                 100 * res$fpr_ci[1], 100 * res$fpr_ci[2]),
    `P1 (uncorrected)` = sprintf("%.1f%%", 100 * res$P1),
    `Implied L_eff`    = round(res$L_eff_implied, 2),
    `Corrected FPR`    = sprintf("%.1f%%", 100 * fpr_corr),
    `Sim FPR (verify)` = if (!is.null(res$fpr_verify) && !is.na(res$fpr_verify))
      sprintf("%.1f%%  [%.1f%%, %.1f%%]",
              100 * res$fpr_verify,
              100 * res$fpr_verify_ci[1],
              100 * res$fpr_verify_ci[2]) else "\u2014",
    fpr_num      = res$fpr,
    fpr_corr_num = fpr_corr,
    fpr_ver_num  = if (!is.null(res$fpr_verify)) res$fpr_verify else NA_real_
  )
}

bin_tbl <- rbind(
  make_row(cal_bin_A_t1,  sprintf("Pair A \u2014 Tier 1 (%d sims)", nsims_tier1)),
  make_row(cal_bin_A_t2,  sprintf("Pair A \u2014 Tier 2 (%d sims)", nsims_tier2)),
  make_row(cal_bin_perm,  sprintf("Pair A \u2014 Permutation (%d sims)", nsims_tier1)),
  make_row(cal_bin_B_t1,  sprintf("Pair B \u2014 Tier 1 (%d sims)", nsims_tier1))
)

gt(bin_tbl) |>
  tab_header(
    title    = "Binary Outcome: fpr_calibration Summary",
    subtitle = sprintf(
      "N=%d | n.min=%d | p_event=%.2f | %d cal + %d verify sims",
      N_demo, n_min_val, p_event, nsims_tier1, nsims_verify)
  ) |>
  tab_spanner(label = "Calibration batch",
    columns = c(`Sim FPR (cal)`, `95% CI`,
                `P1 (uncorrected)`, `Implied L_eff`, `Corrected FPR`)) |>
  tab_spanner(label = "Verification batch (independent)",
    columns = `Sim FPR (verify)`) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `Corrected FPR`,
      rows = fpr_corr_num <= fpr_target_bin)) |>
  tab_style(style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = `Corrected FPR`,
      rows = fpr_corr_num > fpr_target_bin)) |>
  tab_style(style = cell_fill(color = "#E6F1FB"),
    locations = cells_body(columns = `Sim FPR (verify)`)) |>
  cols_hide(columns = c(fpr_num, fpr_corr_num, fpr_ver_num)) |>
  tab_source_note(paste(
    "Calibration batch: derives P1, L_eff_implied, Corrected FPR = 1-(1-P1)^L_eff.",
    "Verification batch: independent sims (different seed) to validate Corrected FPR.",
    "Pair A = (c1=1.70, c2=1.70). Pair B = (c1=2.10, c2=1.40)."))
```


# Example 2: Cox PH (Survival, HR) {#sec-cox}

## DGM and ForestSearch configuration

```{r cox-dgm}
#| code-fold: false

# Null survival DGM using the GBSG structure
dgm_surv_null <- setup_gbsg_dgm(model = "null", k_treat = 1.0,
                                 verbose = FALSE)

cox_formula_adj <- survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3

fs_cox <- list(
  outcome.name           = "y_sim",
  event.name             = "event_sim",
  treat.name             = "treat_sim",
  id.name                = "id",
  confounders.name       = confounders_surv,
  pconsistency.threshold = pcons,
  fs.splits              = fs_splits,
  n.min                  = n_min_val,
  d0.min                 = 12L,
  d1.min                 = 12L,
  maxk                   = maxk_val,
  by.risk                = 12L,
  vi.grf.min             = -0.2,
  use_lasso              = TRUE,
  use_grf                = TRUE,
  use_twostage           = TRUE,
  is.RCT                 = TRUE,
  details                = FALSE,
  quiet                  = TRUE,
  plot.sg                = FALSE,
  parallel_args          = list(plan = "sequential", workers = 1,
                                show_message = FALSE)
)

# Null generator: directly simulate from the GBSG DGM under H0.
# Do NOT call run_simulation_analysis() here — that runs ForestSearch,
# which is fpr_calibration()'s job.  We only need the raw dataset.
null_fn_cox <- function(seed) {
  simulate_from_dgm(dgm_surv_null, n = N_surv,
                    analysis_time = analysis_time,
                    cens_adjust   = cens_adjust,
                    seed          = seed)
}

cat(sprintf("Survival DGM ready | N=%d | analysis_time=%d months | prop_cens~%.2f\n",
            N_surv, analysis_time, prop_cens_surv))
```


## Tier 1 — calibration at paper defaults and at a tighter pair

```{r cox-tier1}

t0 <- proc.time()[3]
# ── Paper defaults: c1=1.25, c2=1.0 ───────────────────────────
cal_cox_paper <- fpr_calibration(
  c1       = 1.25,
  c2       = 1.0,
  null_fn  = null_fn_cox,
  n_sample = N_surv,
  fs_params = fs_cox,
  nsims        = nsims_tier1,
  nsims_verify = nsims_verify,
  n_min     = n_min_val,
  outcome_type_approx = "survival",
  prop_cens = prop_cens_surv,
  sim_seed  = sim_seed,
  n_workers = n_workers
)

# ── Tighter pair: c1=1.40, c2=1.10 ────────────────────────────
cal_cox_tight <- fpr_calibration(
  c1       = 1.40,
  c2       = 1.10,
  null_fn  = null_fn_cox,
  n_sample = N_surv,
  fs_params = fs_cox,
  nsims        = nsims_tier1,
  nsims_verify = nsims_verify,
  n_min     = n_min_val,
  outcome_type_approx = "survival",
  prop_cens = prop_cens_surv,
  sim_seed  = sim_seed,
  n_workers = n_workers
)

print(cal_cox_paper)
print(cal_cox_tight)
timings$cox_tier1 <- proc.time()[3] - t0
cat(sprintf("Cox Tier 1 (2 pairs × %d sims): %.1f min\n",
            nsims_tier1, timings$cox_tier1 / 60))
```


## Summary table: Cox results

```{r cox-summary-table}
#| tbl-cap: "fpr_calibration results for Cox PH outcome."

cox_tbl <- rbind(
  make_row(cal_cox_paper, "Paper defaults (c1=1.25, c2=1.0)"),
  make_row(cal_cox_tight, "Tighter pair  (c1=1.40, c2=1.10)")
)

gt(cox_tbl) |>
  tab_header(
    title    = "Survival Outcome (Cox PH): fpr_calibration Summary",
    subtitle = sprintf("N=%d | n.min=%d | prop_cens=%.2f | %d cal + %d verify sims",
                        N_surv, n_min_val, prop_cens_surv, nsims_tier1, nsims_verify)
  ) |>
  tab_spanner(label = "Calibration batch",
    columns = c(`Sim FPR (cal)`, `95% CI`,
                `P1 (uncorrected)`, `Implied L_eff`, `Corrected FPR`)) |>
  tab_spanner(label = "Verification batch (independent)",
    columns = `Sim FPR (verify)`) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `Corrected FPR`,
      rows = fpr_corr_num <= fpr_target_cox)) |>
  tab_style(style = cell_fill(color = "#E6F1FB"),
    locations = cells_body(columns = `Sim FPR (verify)`)) |>
  cols_hide(columns = c(fpr_num, fpr_corr_num, fpr_ver_num)) |>
  tab_source_note(paste(
    "Calibration batch: derives P1, L_eff_implied, Corrected FPR = 1-(1-P1)^L_eff.",
    "Verification batch: independent sims (different seed) to validate Corrected FPR.",
    "Implied L_eff identical (0.19) for both pairs: threshold-independence holds for survival.",
    "L_eff < 1: events-based filters prune candidates before threshold check."))
```


# Mathematical Framework {#sec-math}

## Per-subgroup detection probability $P_1$

Leon et al. (2024, eq. 3) derive the asymptotic probability that a
single pre-specified subgroup $H$ passes both the screening and
consistency criteria under a true effect $\theta$:

$$
P_1(c_1, c_2;\, d_{\text{eff}},\, \theta) =
P\!\left(W_1 + W_2 \ge 2\log c_1,\;\min(W_1, W_2) \ge \log c_2\right)
$$

where $W_1, W_2 \stackrel{\text{iid}}{\sim}
N\!\left(\log\theta,\; \tfrac{8}{d_{\text{eff}}}\right)$.

### Role of each component

| Component | Symbol | Role |
|-----------|--------|------|
| Screening threshold | $c_1$ | Minimum effect in both half-samples for a candidate to advance |
| Consistency threshold | $c_2$ | Minimum effect in each individual half-sample (enforces split-sample stability) |
| Effective information | $d_{\text{eff}}$ | Governs the variance $8/d_{\text{eff}}$ of each half-sample log-effect |
| True effect | $\theta$ | Log-effect shift of the $N(\log\theta, 8/d_{\text{eff}})$ distribution |

Under $H_0$, $\theta = 1$ so $\log\theta = 0$ and both half-sample
estimates are centred at zero.  $P_1(\theta=1)$ is the per-subgroup
false-positive probability — the chance a single null candidate
accidentally clears both thresholds.

The effective information by outcome type is:

$$
d_{\text{eff}} = \begin{cases}
n_H (1 - p_c) & \text{survival (events)} \\
n_H \bar{p}(1-\bar{p}) & \text{binary (Bernoulli info)} \\
n_H / \sigma_Y^2 & \text{continuous (precision-weighted)} \\
\end{cases}
$$

where $n_H$ is the subgroup size (typically `n.min`), $p_c$ is the
censoring proportion, and $\bar{p}$ is the marginal event probability.


## Procedure-level FPR and the $L_{\text{eff}}$ correction

ForestSearch evaluates multiple candidates, not just one.  If
$L_{\text{eff}}$ candidates are approximately independent under $H_0$:

$$
\text{FPR} = 1 - (1 - P_1)^{L_{\text{eff}}}
$$

The power-law approach fits $L_{\text{eff}}(N) = C(N/n_{\min})^\alpha$
from a reference simulation at a fixed $(c_1, c_2)$, then assumes this
holds across all threshold pairs — the **threshold-independence**
assumption.  For survival, events-based filters make $L_{\text{eff}} < 1$
(the procedure is more conservative than a single test).  For binary
outcomes with small $d_{\text{eff}}$, $L_{\text{eff}}$ varies
substantially across the threshold grid, violating the assumption.


## Direct simulation via `fpr_calibration()`

`fpr_calibration()` sidesteps both assumptions.  For $B$ null replicates
at the analyst's specific $(c_1, c_2)$:

$$
\widehat{\text{FPR}} = \frac{1}{B}\sum_{b=1}^B \mathbf{1}[\text{subgroup found in replicate } b]
$$

with Wilson 95% CI.  No power law and no threshold-independence are
required — this is a direct Monte Carlo estimate at the operating point
of interest.

The **implied $L_{\text{eff}}$** is then a by-product:

$$
\widehat{L}_{\text{eff}}(c_1, c_2) =
\frac{\log(1 - \widehat{\text{FPR}})}{\log(1 - P_1(c_1, c_2))}
$$

This is pair-specific and makes no portability claim.  It is reported as
a diagnostic — if it differs substantially from the power-law value at
the same $N$, the threshold-independence assumption was violated for this
pair.


## Detection curve: $P_1(\theta)$ and the corrected curve

Using the estimated $\widehat{L}_{\text{eff}}(c_1, c_2)$ from
`fpr_calibration()`, the detection probability at any true effect
$\theta$ is:

$$
P_{\text{corr}}(\theta;\, c_1, c_2) =
1 - \bigl(1 - P_1(\theta;\, c_1, c_2)\bigr)^{\widehat{L}_{\text{eff}}}
$$

Under $H_0$ ($\theta = 1$), this recovers $\widehat{\text{FPR}}$ by
construction.  Under $H_1$ ($\theta > 1$), it gives the corrected power
for detecting a subgroup with true effect $\theta$.  The wrapper
`plot_fpr_calibration()` plots both $P_1(\theta)$ and
$P_{\text{corr}}(\theta)$ together with the simulation estimate and CI
at $\theta = 1$.


# Detection Curve Wrapper {#sec-plot-wrapper}

## `plot_fpr_calibration()`

```{r plot-wrapper-fn}
#| code-fold: false

#' Plot Detection Curve from fpr_calibration Result
#'
#' Plots P1(theta) and the L_eff-corrected curve across a range of true
#' effect sizes, anchored by the direct simulation FPR estimate at theta=1.
#'
#' @param cal_result  An object of class "fpr_calibration".
#' @param theta_seq   Numeric vector of effect sizes to plot (ratio scale,
#'   e.g. OR or HR). Default seq(0.5, 4.0, by = 0.05).
#' @param fpr_target  Optional numeric. Draws a horizontal reference line.
#' @param label       Optional character. Added to the plot subtitle.
#' @param effect_label Character. Label for x-axis (e.g. "OR", "HR").
#'
#' @return A ggplot2 object.
plot_fpr_calibration <- function(cal_result,
                                  theta_seq    = seq(0.5, 4.0, by = 0.05),
                                  fpr_target   = NULL,
                                  label        = NULL,
                                  effect_label = "Effect ratio (theta)") {

  stopifnot(inherits(cal_result, "fpr_calibration"))

  c1      <- cal_result$c1
  c2      <- cal_result$c2
  d_eff   <- cal_result$d_eff
  L_eff   <- cal_result$L_eff_implied
  fpr_hat <- cal_result$fpr
  ci_lo   <- cal_result$fpr_ci[1]
  ci_hi   <- cal_result$fpr_ci[2]
  N       <- cal_result$N
  nsims   <- cal_result$nsims

  # Compute P1 and corrected curve across theta_seq
  p1_vec <- sapply(theta_seq, function(th)
    tryCatch(
      compute_detection_probability_glm(
        theta = th, d_eff = d_eff, c1 = c1, c2 = c2,
        effect_scale = "ratio", method = "monte_carlo"),
      error = function(e) NA_real_))

  p_corr_vec <- if (!is.na(L_eff))
    1 - (1 - p1_vec)^L_eff
  else
    rep(NA_real_, length(theta_seq))

  curve_dt <- data.frame(
    theta  = rep(theta_seq, 2),
    prob   = c(p1_vec, p_corr_vec),
    curve  = rep(c("Per-subgroup P1 (correct subgroup)",
                   "Corrected (any detection)"),
                 each = length(theta_seq))
  )

  # Simulation point at theta = 1
  sim_pt <- data.frame(
    theta = 1, prob = fpr_hat, ci_lo = ci_lo, ci_hi = ci_hi,
    curve = "Direct simulation (theta=1)")

  # Subtitle
  sub_txt <- sprintf(
    "c1=%.2f, c2=%.2f | d_eff=%.1f | L_eff(implied)=%.3f | N=%d | %d sims",
    c1, c2, d_eff,
    if (!is.na(L_eff)) L_eff else 0,
    N, nsims)
  if (!is.null(label)) sub_txt <- paste0(label, " | ", sub_txt)

  p <- ggplot(curve_dt, aes(x = theta, y = prob,
                              colour = curve, linetype = curve)) +
    geom_line(linewidth = 0.9, na.rm = TRUE) +
    geom_vline(xintercept = 1, linetype = "dotted",
               colour = "grey60", linewidth = 0.4) +
    # Simulation FPR point + CI at theta = 1
    geom_point(data = sim_pt, aes(x = theta, y = prob),
               colour = "steelblue", size = 4, shape = 18,
               inherit.aes = FALSE) +
    geom_errorbar(data = sim_pt,
                  aes(x = theta, ymin = ci_lo, ymax = ci_hi),
                  colour = "steelblue", width = 0.04, linewidth = 0.7,
                  inherit.aes = FALSE) +
    annotate("text", x = 1.05, y = fpr_hat,
             label = sprintf("Sim FPR = %.1f%%", 100 * fpr_hat),
             hjust = 0, size = 3.2, colour = "steelblue") +
    scale_colour_manual(values = c(
      "Per-subgroup P1 (correct subgroup)" = "grey50",
      "Corrected (any detection)"          = "firebrick")) +
    scale_linetype_manual(values = c(
      "Per-subgroup P1 (correct subgroup)" = "dashed",
      "Corrected (any detection)"          = "solid")) +
    scale_y_continuous(labels = scales::percent,
                       limits = c(0, NA)) +
    labs(x        = effect_label,
         y        = "Detection probability",
         title    = sprintf("Detection Curve: c1 = %.2f, c2 = %.2f", c1, c2),
         subtitle = sub_txt,
         colour   = NULL, linetype = NULL) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top",
          plot.title      = element_text(face = "bold"),
          plot.subtitle   = element_text(size = 9, colour = "grey40"))

  if (!is.null(fpr_target)) {
    p <- p +
      geom_hline(yintercept = fpr_target, linetype = "dashed",
                 colour = "darkgreen", linewidth = 0.5) +
      annotate("text",
               x = min(theta_seq) + 0.05,
               y = fpr_target + 0.01,
               label = sprintf("FPR target = %.0f%%", 100 * fpr_target),
               hjust = 0, size = 3, colour = "darkgreen")
  }

  p
}
```


## Binary example: detection curves at two pairs

```{r binary-detection-curves, fig.width=9, fig.height=5}
#| fig-cap: "Detection curves for binary outcome. Diamond = direct simulation FPR with 95% CI at theta=1 (H0). Solid red = corrected curve using implied L_eff; dashed grey = per-subgroup P1. The corrected curve is anchored to the simulation point by construction."

p_A <- plot_fpr_calibration(
  cal_bin_A_t1,
  fpr_target   = fpr_target_bin,
  effect_label = "Odds ratio (theta)",
  label        = "Pair A"
)

p_B <- plot_fpr_calibration(
  cal_bin_B_t1,
  fpr_target   = fpr_target_bin,
  effect_label = "Odds ratio (theta)",
  label        = "Pair B"
)

p_A + p_B +
  plot_annotation(
    title    = "Binary Outcome: Detection Curves by Threshold Pair",
    subtitle = sprintf("N=%d | n.min=%d | d_eff=%.1f | p_event=%.2f",
                        N_demo, n_min_val,
                        d_eff_binary(n_min_val, p_event), p_event),
    theme    = theme(plot.title    = element_text(face = "bold"),
                     plot.subtitle = element_text(size = 10, colour = "grey40"))
  )
```


## Cox example: detection curve at paper defaults

```{r cox-detection-curve, fig.width=7, fig.height=4.5}
#| fig-cap: "Detection curve for Cox PH outcome at paper defaults. L_eff < 1 for survival means the corrected curve lies BELOW the per-subgroup P1 — the procedure is more conservative than a single test."

plot_fpr_calibration(
  cal_cox_paper,
  fpr_target   = fpr_target_cox,
  effect_label = "Hazard ratio (theta)",
  label        = "GBSG survival, paper defaults"
)
```


# Comparison: fpr_calibration vs L_eff power law {#sec-compare}

```{r compare-plot}
#| fig-height: 5
#| fig-cap: "Direct simulation FPR (point + 95% CI) vs the L_eff-corrected approximation for binary outcome. The corrected approximation overestimates because the calibration grid was in the saturated regime. fpr_calibration() recovers the true FPR without any power-law assumption."

# Compute corrected approximation using the loaded C/alpha
# (these come from calibration_binary_leff.qmd if the .rds is available;
#  otherwise use representative values from the HTML output)
C_fit     <- tryCatch(
  load_leff_calibration("binary", output_dir = "../_data")$C,
  error = function(e) 2.283)
alpha_fit <- tryCatch(
  load_leff_calibration("binary", output_dir = "../_data")$alpha,
  error = function(e) 0.228)

# Build comparison for a range of c1 pairs (c2 = c1 for simplicity)
# c1_seq is defined in the setup chunk (demo-params)

# Run fpr_calibration at each pair (Tier 1: nsims_tier1 sims each)
t0 <- proc.time()[3]
compare_results <- lapply(c1_seq, function(c1v) {
  res <- fpr_calibration(
    c1       = c1v,
    c2       = c1v,
    null_fn  = function(seed) sim_null_binary(N_demo, seed),
    n_sample = N_demo,
    fs_params = fs_binary,
    nsims     = nsims_tier1,
    n_min     = n_min_val,
    outcome_type_approx = "binary",
    p_event   = p_event,
    sim_seed  = sim_seed,
    n_workers = n_workers,
    quiet     = TRUE
  )
  data.table(
    c1         = c1v,
    c2         = c1v,
    sim_fpr    = res$fpr,
    ci_lo      = res$fpr_ci[1],
    ci_hi      = res$fpr_ci[2],
    P1         = res$P1,
    d_eff      = d_eff_binary(n_min_val, p_event),
    L_eff      = C_fit * (N_demo / n_min_val)^alpha_fit
  )
})
compare_dt <- rbindlist(compare_results)
compare_dt[, fpr_corr := 1 - (1 - P1)^L_eff]

elapsed <- proc.time()[3] - t0
timings$compare <- elapsed
cat(sprintf("Grid comparison: %d pairs in %.1f min\n",
            length(c1_seq), elapsed / 60))

# Reshape for plotting
plot_dt <- rbind(
  compare_dt[, .(c1, fpr = sim_fpr, lo = ci_lo, hi = ci_hi,
                 method = "Direct simulation (fpr_calibration)")],
  compare_dt[, .(c1, fpr = fpr_corr, lo = NA_real_, hi = NA_real_,
                 method = "L_eff corrected approximation")],
  compare_dt[, .(c1, fpr = P1, lo = NA_real_, hi = NA_real_,
                 method = "Uncorrected P1")]
)

ggplot(plot_dt, aes(x = c1, y = fpr, colour = method, linetype = method)) +
  geom_line(linewidth = 0.9) +
  geom_point(data = plot_dt[method == "Direct simulation (fpr_calibration)"],
             size = 3) +
  geom_errorbar(
    data = plot_dt[method == "Direct simulation (fpr_calibration)"],
    aes(ymin = lo, ymax = hi), width = 0.03, linewidth = 0.6) +
  geom_hline(yintercept = fpr_target_bin, linetype = "dashed",
             colour = "grey40", linewidth = 0.5) +
  annotate("text", x = max(c1_seq) - 0.05, y = fpr_target_bin + 0.005,
           label = sprintf("FPR target = %.0f%%", 100 * fpr_target_bin), hjust = 1, size = 3,
           colour = "grey40") +
  scale_colour_manual(values = c(
    "Direct simulation (fpr_calibration)" = "steelblue",
    "L_eff corrected approximation"       = "firebrick",
    "Uncorrected P1"                       = "grey60"
  )) +
  scale_linetype_manual(values = c(
    "Direct simulation (fpr_calibration)" = "solid",
    "L_eff corrected approximation"       = "dashed",
    "Uncorrected P1"                       = "dotted"
  )) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x        = expression(c[1] ~ "(= " ~ c[2] ~ "for this sweep)"),
    y        = "Procedure-level FPR",
    title    = "Direct Simulation vs L_eff Correction: Binary Outcome",
    subtitle = sprintf(
      "N=%d | %d sims/pair (Tier 1) | d_eff=%.1f | L_eff=%.2f (power-law)",
      N_demo, nsims_tier1, d_eff_binary(n_min_val, p_event),
      C_fit * (N_demo / n_min_val)^alpha_fit),
    colour   = NULL, linetype = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


# Precision Guide {#sec-precision}

```{r precision-table}
#| tbl-cap: "95% CI half-width for fpr_calibration() at typical FPR levels. The Tier 1 (approximation) range is highlighted green; Tier 2 (verification) in blue."
#| code-fold: true

ci_hw <- function(fpr, n) 1.96 * sqrt(fpr * (1 - fpr) / n)

prec_dt <- data.table(expand.grid(
  `nsims`   = c(50L, 100L, 200L, 500L, 1000L, 5000L, 20000L),
  `FPR ref` = c(0.05, 0.10, 0.15)
))
prec_dt[, `CI half-width (pp)` := round(100 * ci_hw(`FPR ref`, nsims), 1)]
prec_dt[, Tier := fcase(
  nsims <= 500,  "Tier 1 — approximation",
  nsims <= 2000, "Tier 1/2 transition",
  default =      "Tier 2 — verification"
)]
prec_wide <- dcast(prec_dt, nsims + Tier ~ paste0("FPR=", `FPR ref`*100, "%"),
                   value.var = "CI half-width (pp)")
setorder(prec_wide, nsims)

gt(prec_wide) |>
  tab_header(
    title    = "Precision of fpr_calibration() by nsims",
    subtitle = "95% CI half-width in percentage points"
  ) |>
  tab_spanner(label = "95% CI half-width (pp)",
              columns = starts_with("FPR=")) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(rows = Tier == "Tier 1 — approximation")) |>
  tab_style(style = cell_fill(color = "#E6F1FB"),
    locations = cells_body(rows = Tier == "Tier 2 — verification")) |>
  tab_source_note(paste(
    "Green = Tier 1 (quick approximation, 200–500 sims).",
    "Blue  = Tier 2 (high-precision verification, 5,000–20,000 sims).",
    "Recommended workflow: Tier 1 to screen candidate (c1,c2);",
    "Tier 2 to confirm the final operating point."))
```


# Timing Summary {#sec-timing}

```{r timing-summary}
# Local fallback in case fpr_calibration.R is not sourced in this session
`%||%` <- function(a, b) if (!is.null(a)) a else b

timings$total <- (proc.time() - t_doc_start)["elapsed"]

# ── Per-sim rates ──────────────────────────────────────────────
# Each binary Tier 1 call: 2 pairs × nsims_tier1 sims
# Each Cox Tier 1 call:    2 pairs × nsims_tier1 sims
rate_bin <- timings$bin_tier1 / (2 * nsims_tier1)
rate_cox <- timings$cox_tier1 / (2 * nsims_tier1)
n_cmp_pairs <- length(seq(1.50, 2.50, by = 0.20))
rate_cmp <- ifelse(!is.null(timings$compare),
                   timings$compare / (n_cmp_pairs * nsims_tier1),
                   NA_real_)

# ── Tier 2 projection from Tier 1 rate ────────────────────────
proj_tier2_bin <- rate_bin * nsims_tier2   # per pair
proj_tier2_cox <- rate_cox * 10000L        # per pair at 10K sims

timing_tbl <- data.frame(
  Stage = c(
    sprintf("Binary Tier 1 — 2 pairs × %d sims", nsims_tier1),
    sprintf("Binary Tier 2 — 1 pair  × %d sims", nsims_tier2),
    sprintf("Binary permutation — 1 pair × %d sims", nsims_tier1),
    sprintf("Cox Tier 1    — 2 pairs × %d sims", nsims_tier1),
    sprintf("Comparison grid — %d pairs × %d sims", n_cmp_pairs, nsims_tier1),
    "Total (document)"
  ),
  Seconds = c(
    timings$bin_tier1 %||% NA_real_,
    timings$bin_tier2 %||% NA_real_,
    timings$bin_perm  %||% NA_real_,
    timings$cox_tier1 %||% NA_real_,
    timings$compare   %||% NA_real_,
    timings$total
  ),
  stringsAsFactors = FALSE
)
timing_tbl$Minutes <- timing_tbl$Seconds / 60

gt(timing_tbl) |>
  tab_header(
    title    = "Computational Timing Summary",
    subtitle = sprintf("%d workers | binary N=%d, d_eff=%.1f | Cox N=%d, d_eff=%.1f",
                        n_workers, N_demo,
                        d_eff_binary(n_min_val, p_event),
                        N_surv,
                        d_eff_survival(n_min_val, prop_cens_surv))
  ) |>
  fmt_number(columns = Seconds, decimals = 1, rows = !is.na(Seconds)) |>
  fmt_number(columns = Minutes, decimals = 2, rows = !is.na(Minutes)) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)", Minutes = "Time (min)") |>
  tab_style(style = cell_text(weight = "bold"),
    locations = cells_body(rows = grepl("Total", Stage))) |>
  tab_row_group(label = "Binary", rows = grepl("Binary|permutation|Comparison", Stage)) |>
  tab_row_group(label = "Cox (survival)", rows = grepl("Cox", Stage)) |>
  tab_row_group(label = "Document total", rows = grepl("Total", Stage)) |>
  tab_source_note(sprintf(
    paste("Per-sim wall-clock: binary %.2f sec, Cox %.2f sec  (%d workers).",
          "Tier 2 projection (10K sims, 1 pair): binary %.0f min, Cox %.0f min."),
    rate_bin, rate_cox, n_workers,
    proj_tier2_bin / 60, proj_tier2_cox / 60))
```


# Session Info {#sec-session}

```{r cleanup}
plan(sequential)
sessionInfo()
```
