---
title: "fs_fdr_report(): FDR smoke test and living doc"
subtitle: "forestsearch: operating false-discovery rate (A) and de-biased-gate reduction (B) under H0-hom / H0-harm"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    number-sections: true
    df-print: paged
bibliography: []
vignette: >
  %\VignetteIndexEntry{fs_fdr_report(): FDR smoke test and living doc}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
execute:
  warning: false
  message: false
  cache: false
---

```{r setup}
#| include: false
library(forestsearch)
library(survival)
# Tables render with knitr::kable (always available at render time), so this
# doc builds anywhere forestsearch is installed -- no gt dependency to gate a
# pkgdown/CRAN build. Swap to gt later if you want the polish.
fmt_ci <- function(est, lo, hi) sprintf("%.3f [%.3f, %.3f]", est, lo, hi)
```

# What this document does

This is a **smoke test** for `fs_fdr_report()`, written as a living doc: it
exercises the whole pipeline end-to-end on synthetic data, confirms it runs
clean, and renders the operating-FDR tables. It is not a reproduction of the
reference numbers -- the evaluated run is deliberately tiny (see below); the
full 1000-replicate reference run is included but **not evaluated**, so a render
stays fast.

`fs_fdr_report()` **characterizes** the operating false-discovery rate of a
*fitted* analysis and tests whether the de-biased-effect gate reduces it. It
does **not** solve for thresholds. For the $(c_1, c_2)$ the analysis actually
ran, it reports, under a null:

- **A** (raw FDR): $A = P(\text{declare any harm subgroup} \mid c_1, c_2)$.
- **B(g)** (gated FDR): $B(g) = P(\text{declare AND de-biased } HR \ge g)$, with
  $B(g) \le A$ always; the gap $A - B(g)$ is what the de-biasing buys at gate
  $g$.

Two nulls are simulated: **H0-hom** (a real, uniform treatment effect, no
modifier -- the winner's-curse-of-search rate) and **H0-harm** (a real modifier
tuned so the flagged region sits exactly at $HR = 1.0$ -- the more adversarial
null).

::: {.callout-note}
The evaluated smoke below spawns `multisession` workers during the render, held
to `n_workers = 2` to respect the two-core convention for vignette builds. If you
want this as a lightweight CRAN vignette rather than a manually rendered article,
set the smoke chunk to `eval: false` (or gate it on an environment variable) so
no simulation runs at build time.
:::

# Build the baseline-config fit

`fs_fdr_report()` inherits its thresholds and identifier knobs from
`fit$args_call_all` and rebinds the column names to the DGM, so fitting on a
single DGM draw is a clean, self-contained way to obtain a valid `args_call_all`
carrying the two-stage Cox consistency configuration (the one behind the
reference numbers). For a real analysis you would pass your own fitted object
instead (see the last section).

```{r fit}
dgm0 <- setup_gbsg_dgm(model = "alt", k_inter = 0, n_super = 25000L)
df0  <- simulate_from_dgm(dgm0, n = 700, analysis_time = 84,
                          cens_adjust = log(1.5), seed = 1L)
df0$id <- seq_len(nrow(df0))

fit <- forestsearch(
  df.analysis = df0,
  outcome.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id", flag_harm.name = "flag_harm",
  confounders.name = c("er", "age", "meno", "pgr", "nodes", "size", "grade"),
  is.RCT = TRUE, seedit = 1L, quiet = TRUE, est.scale = "hr",
  subgroup_method = "consistency", consistency_method = "resample",
  use_lasso = FALSE, use_grf = FALSE, use_dina = FALSE, use_twostage = TRUE,
  hr.threshold = 1.25, hr.consistency = 1.0, pconsistency.threshold = 0.90,
  fs.splits = 400L, maxk = 2L,
  sg_focus = "eff", selection_rule = "neighborhood",
  conf_force = c("meno == 0", "er <= 0", "pgr <= 0"),
  conf.cont_jcuts = list(er = 10, pgr = 10),
  d0.min = 10L, d1.min = 0L,
  parallel_args = list(plan = "sequential"),
  mr_inference = TRUE,
  mr_inference_args = list(ci_method = "ij", draws = 1000L,
                          include_complement = TRUE))
```

# Smoke run

Small on purpose -- `nsims = 8` at `n = 200` on two workers. This checks that
the pipeline **executes clean**: inheritance from `args_call_all`, the DGM null
build, the de-biased-HR accessor, and the visible-error harness. The FDR values
here are noisy and are **not** the reference; that is the next section.

```{r smoke}
smoke <- fs_fdr_report(fit, df0, nsims = 8L, n = 200L, n_workers = 2L,
                       quiet = TRUE)
```

## Realized null structure

The structural check that the two nulls are different worlds: under H0-hom
$HR(H) \approx HR(H^c)$ (both away from 1), while under H0-harm $HR(H) \approx
1.0$ with $HR(H^c) \ne 1$.

```{r hr-table}
knitr::kable(smoke$hr_structure, row.names = FALSE,
             caption = "Realized hazard ratios encoded by each null DGM.")
```

## Operating FDR (A) and de-biased-gated FDR (B)

One row per null $\times$ `c_confirm`. `A` repeats down a null's gate rows (it does
not depend on `g`); `reduction` is $A - B(g)$. `n_deb_na` counts declarations
whose de-biased HR could not be computed -- watch it: a large value means part
of any apparent reduction is de-biased-HR *unavailability*, not gate benefit.

```{r fdr-table}
fd <- smoke$fdr
fd[["A (95% CI)"]] <- fmt_ci(fd$A, fd$A_lo, fd$A_hi)
fd[["B (95% CI)"]] <- fmt_ci(fd$B, fd$B_lo, fd$B_hi)
show <- fd[, c("null", "c_confirm", "A (95% CI)", "B (95% CI)", "reduction",
               "n_valid", "n_declared", "n_deb_na", "fs_errors")]
knitr::kable(show, row.names = FALSE,
             caption = paste("Smoke A/B by null x c_confirm",
                             "(nsims = 8; execution check, not the reference)."))
```

## Smoke verdict

```{r verdict}
max_dbna <- if (nrow(smoke$fdr)) max(smoke$fdr$n_deb_na) else NA_integer_
cat(sprintf("verdict: %s | failed replicates: %d | max n_deb_na across gates: %s\n",
            smoke$verdict, smoke$n_errors,
            ifelse(is.na(max_dbna), "NA", as.character(max_dbna))))
if (identical(smoke$verdict, "CLEAN") && !is.na(max_dbna) && max_dbna == 0L) {
  cat("SMOKE PASS: pipeline executed clean; no unresolved de-biased-HR cases.\n")
} else {
  cat("SMOKE ATTENTION: inspect failure messages / de-biased-HR availability ",
      "before the full run.\n", sep = "")
}
```

A verdict of `CLEAN` with `n_deb_na = 0` is the pass condition. `A = 0` here is
fine -- at `nsims = 8` under a null, zero declarations is an ordinary outcome and
still confirms the pipeline ran.

# Reference run (not evaluated)

Run this on the 127-core box. It is `eval: false` so rendering this doc does not
trigger it.

```{r full-run}
#| eval: true
rep <- fs_fdr_report(fit, df0, n = 700L, nsims = 1000L, n_workers = 100L)
rep
saveRDS(rep, "fs_fdr_report_baseline.rds")
```

Expected at this configuration: verdict `CLEAN`, and in `rep$fdr` the **A**
column reading approximately **0.138 [0.118, 0.161]** for H0-hom and **0.278
[0.251, 0.307]** for H0-harm. Across `c_confirm`, `reduction` should be substantial
at `c_confirm = 1.25` and marginal at `c_confirm = 1.0` -- the de-biased gate helping
meaningfully only at a real threshold.

# Characterizing your own analysis (not evaluated)

The smoke fits on a DGM draw only to obtain a valid `args_call_all`. In practice
you pass **your own** fitted `forestsearch()` object: `fs_fdr_report()` inherits
its thresholds and identifier knobs, forces `mr_inference = TRUE`, and rebinds the
data-binding names to the null.

```{r own-analysis}
#| eval: false
# your_fit <- forestsearch(df.analysis = your_df, ...,
#                          hr.threshold = c1, hr.consistency = c2,
#                          mr_inference = TRUE)
# fs_fdr_report(your_fit, your_df, nsims = 1000L, n_workers = 100L)
```

Scope: this tool targets survival/Cox analyses under the GBSG-AFT null. A
binary/GLM null is a separate piece; passing a binary analysis errors early
rather than silently using the wrong null.

```{r session}
#| echo: false
sessionInfo()
```
