---
title: "The GBSG AFT data-generating mechanism"
subtitle: "An introduction / appendix for FDR operating characteristics with fs_fdr_report()"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    number-sections: true
    df-print: paged
bibliography: []
vignette: >
  %\VignetteIndexEntry{The GBSG AFT data-generating mechanism}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
execute:
  warning: false
  message: false
  cache: false
---

```{r setup}
#| include: false
library(forestsearch)
library(survival)
```

# Why this document

`fs_fdr_report()` reports the *operating* false-discovery rate of a fitted
`forestsearch()` analysis: for the analysis's own $(c_1, c_2)$, it simulates
under a null and returns **A** = $P(\text{declare any harm subgroup})$ and, at
each de-biased-HR gate $g$, **B(g)** = $P(\text{declare AND de-biased } HR \ge
g)$. Those probabilities are only as meaningful as the null they are computed
under. This appendix documents that null: the GBSG accelerated-failure-time
(AFT) data-generating mechanism built by `setup_gbsg_dgm()`, all of its options,
and — a point that is easy to lose track of — how its **censoring** model is
obtained.

Everything below is drawn from the package source (`setup_gbsg_dgm.R`,
`sim_aft_gbsg.R`, `simulate_from_dgm.R`, `find_k_inter_main.R`), not from
recollection.

# The outcome model is a Weibull AFT fitted to GBSG

`setup_gbsg_dgm()` is a thin wrapper: it calls the internal
`.create_gbsg_dgm_()`, then reshapes the result into an `aft_dgm_flex` object
that `simulate_from_dgm()` can draw from. The substance is in
`.create_gbsg_dgm_()`.

It fits a Weibull AFT to the GBSG data,

$$\log T = \mu + \sigma\,\varepsilon + \eta(X), \qquad \eta(X) = X'\gamma,$$

via `survreg(Surv(y, event) ~ z_true, dist = "weibull")`, and extracts the
fitted coefficients $\gamma$ and scale $\sigma$. It then **scales three of those
coefficients** by the effect multipliers you supply:

$$\gamma_{\text{treat}} \leftarrow k_{\text{treat}}\,\gamma_{\text{treat}},
\qquad
\gamma_{z_3} \leftarrow k_{z_3}\,\gamma_{z_3},
\qquad
\gamma_{z_h} \leftarrow k_{\text{inter}}\,\gamma_{z_h},$$

where $z_h = \text{treat} \times z_1 \times z_3$ is the treatment-by-subgroup
interaction. Converting the AFT to a proportional-hazards reading
($b_0 = -\gamma/\sigma$) gives the two hazard ratios that anchor everything:

$$\mathrm{HR}(H) = \exp\!\big(-\gamma_{\text{treat}}/\sigma - \gamma_{z_h}/\sigma\big),
\qquad
\mathrm{HR}(H^c) = \exp\!\big(-\gamma_{\text{treat}}/\sigma\big).$$

The single most useful consequence: **`k_inter = 0` makes $\gamma_{z_h} = 0$, so
$\mathrm{HR}(H) = \mathrm{HR}(H^c)$** — a real, uniform treatment effect with no
modifier. That is the homogeneous null. A nonzero `k_inter` bends the harm
region's HR away from the complement's.

`model = "alt"` activates the interaction and defines the harm region; `model =
"null"` sets `flag_harm = 0` throughout (a no-subgroup world).

# The harm subgroup

The flagged region is fixed and clinical: `flag_harm = (z1 == 1 & z3 == 1)`,
where `z1 = 1` marks **low estrogen receptor** (ER at or below the
`z1_quantile` quantile) and `z3 = 1` marks **premenopausal** status. The source
labels it *"low ER & premenopausal."* `z1_quantile` (default 0.25) therefore
controls how large the candidate harm region is.

# The censoring model — fitted from the data by default

This is the part worth stating precisely. For the default `cens_type =
"weibull"`, `.create_gbsg_dgm_()` **fits a censoring model to the data**:

```r
survreg(Surv(y, 1 - event) ~ z_cens, dist = "weibull")
```

Flipping the event indicator to `1 - event` turns the outcome fit into a
*time-to-censoring* fit, and `z_cens` are covariates — so the censoring is a
**covariate-dependent Weibull**, estimated from GBSG, not a fixed rate. Its
intercept $\mu_c$, scale $\tau_c$, and covariate effects are stored in the DGM's
`cens_params`.

So your recollection is right, with one refinement: it is not that the DGM
*auto-selects a censoring family from the data*. You choose the family via
`cens_type` (default `"weibull"`), and the default family's **parameters** are
what get estimated from the data. The alternative, `cens_type = "uniform"`, is
analytical rather than fitted.

At draw time, `simulate_from_dgm()` generates censoring times as

$$\log C = \mu_c + \texttt{cens\_adjust} + \tau_c\,\varepsilon_c + \eta_c(X),$$

so **`cens_adjust` is a log-scale shift of the censoring intercept** that tunes
the censoring rate *without refitting*: `cens_adjust = log(2)` doubles expected
censoring times (fewer events censored), `log(0.5)` halves them (more
censoring). On top of this, `analysis_time` imposes **administrative**
censoring — follow-up is capped at `analysis_time - entry_time` — which is why
the source warns that a time-unit mismatch (e.g. `analysis_time` in months but
the fitted $\mu$ implying years) can drive censoring toward 100%. A companion
diagnostic in `simulate_from_dgm.R` compares the reference (GBSG) and simulated
censoring rates and KM curves so you can check the draw is aligned with the real
data.

::: {.callout-note}
## The general engine has an explicit toggle
`setup_gbsg_dgm()` uses the legacy GBSG path, which always fits the Weibull
censoring model from data. The more general `generate_aft_dgm()` (which works on
any dataset) exposes a `select_censoring` argument: `TRUE` (default) fits the
censoring model from data; `FALSE` lets you supply `cens_params` analytically or
force an intercept-only fit. `setup_gbsg_dgm()`'s own header notes a migration
path toward that engine. For the GBSG null behind `fs_fdr_report()`, the
fit-from-data Weibull is what you get.
:::

# Calibrating `k_inter` to a target HR

You rarely set `k_inter` directly; you set the HR you want and solve for it.
`calibrate_k_inter()` uses root-finding (`uniroot`, `extendInt = "yes"`, over a
default range of $[-10, 10]$) to find the `k_inter` at which the realized
$\mathrm{HR}(H)$ equals a target. `use_ahr = FALSE` calibrates the Cox-scale
$\mathrm{HR}(H)$; `use_ahr = TRUE` calibrates the average HR.

This is exactly how the two `fs_fdr_report()` nulls arise:

- **H0-hom:** `k_inter = 0` directly — $\mathrm{HR}(H) = \mathrm{HR}(H^c)$, a
  uniform effect. False discovery = manufacturing heterogeneity from selection.
- **H0-harm:** `calibrate_k_inter(target_hr_harm = 1.0)` — a *real* modifier
  tuned so the flagged region sits exactly at $\mathrm{HR}(H) = 1.0$ while the
  complement still responds. False discovery = calling that region harmful when
  it is null there. This is the more adversarial null.

# All the options at a glance

`setup_gbsg_dgm()`:

| Argument | Default | Role |
|---|---|---|
| `model` | `"alt"` | `"alt"` = real effect/interaction and a defined harm region; `"null"` = no subgroup (`flag_harm = 0`). |
| `k_treat` | `1` | Multiplier on the fitted treatment main effect $\gamma_{\text{treat}}$. |
| `k_inter` | `1` | Multiplier on the interaction coefficient $\gamma_{z_h}$; `0` gives homogeneity. |
| `k_z3` | `1` | Multiplier on the $z_3$ (menopausal-status) effect. |
| `z1_quantile` | `0.25` | ER quantile defining `z1 = 1` (low ER), hence the harm-region size. |
| `n_super` | `5000` | Super-population size the AFT and censoring models are built on. |
| `cens_type` | `"weibull"` | Censoring family: `"weibull"` (fitted from data) or `"uniform"` (analytical). |
| `use_rand_params` | `FALSE` | Refit the nuisance covariate coefficients from a richer model. |
| `seed` | `8316951` | Seed for the super-population construction. |
| `verbose` | `FALSE` | Print the fitted coefficients and subgroup summary. |

Key `simulate_from_dgm()` draw controls:

| Argument | Default | Role |
|---|---|---|
| `n` | — | Trial size to draw. |
| `analysis_time` | `48` | Calendar time of analysis; sets administrative censoring via `analysis_time - entry_time`. |
| `cens_adjust` | `0` | Log-scale shift of the censoring intercept; tunes the censoring rate without refitting. |
| `seed` | — | Per-draw seed (draws are reproducible and plan-invariant). |

Emitted columns: `y_sim` = $\min(T, C)$, `event_sim` = $\mathbb{1}(T \le C)$,
`treat_sim`, `flag_harm`, and `c_time` — the exact names `fs_fdr_report()`
binds to.

# Seeing it concretely

The two nulls, with their realized structure. (`n_super` is modest here for a
fast render; `fs_fdr_report()` defaults to 25000.)

```{r build-dgms}
n_super_demo <- 5000L
dgm_hom  <- setup_gbsg_dgm(model = "alt", k_inter = 0, n_super = n_super_demo)
k_harm   <- calibrate_k_inter(target_hr_harm = 1.0, model = "alt",
                              use_ahr = FALSE)
dgm_harm <- setup_gbsg_dgm(model = "alt", k_inter = k_harm,
                           n_super = n_super_demo)
```

```{r hr-structure}
# k_inter lives in effect_modifiers (NOT model_params); HR/AHR are top-level.
hr_row <- function(tag, dgm) {
  em <- dgm$effect_modifiers
  data.frame(
    null    = tag,
    k_inter = round(em$k_inter, 3),
    HR_H    = round(dgm$hr_H_true,  3),
    HR_Hc   = round(dgm$hr_Hc_true, 3),
    AHR_H   = round(dgm$AHR_H_true,  3),
    AHR_Hc  = round(dgm$AHR_Hc_true, 3),
    prop_H  = round(dgm$subgroup_info$proportion, 3),
    stringsAsFactors = FALSE)
}
knitr::kable(
  rbind(hr_row("H0-hom  (k_inter = 0)",       dgm_hom),
        hr_row("H0-harm (target HR_H = 1.0)", dgm_harm)),
  row.names = FALSE,
  caption = "Realized structure of the two nulls (prop_H = harm-region share).")
```

Under H0-hom the two HRs should be $\approx$ equal and away from 1; under
H0-harm, $\mathrm{HR}(H) \approx 1.0$ with $\mathrm{HR}(H^c)$ unchanged.

The fitted censoring model, and how `cens_adjust` moves the censoring rate:

```{r censoring}
# Parameters of the Weibull censoring model fitted from the data:
str(dgm_hom$cens_params[c("type", "mu", "sigma")])

# cens_adjust shifts the censoring intercept; higher -> longer censoring times
# -> fewer censored observations.
cens_rate <- function(ca) {
  d <- simulate_from_dgm(dgm_hom, n = 1000L, analysis_time = 84,
                         cens_adjust = ca, seed = 1L)
  mean(d$event_sim == 0L)
}
adj <- c(log(0.5), 0, log(1.5), log(2))
knitr::kable(
  data.frame(cens_adjust = c("log(0.5)", "0", "log(1.5)", "log(2)"),
             value       = round(adj, 3),
             cens_rate   = round(vapply(adj, cens_rate, numeric(1)), 3)),
  row.names = FALSE,
  caption = "Censoring rate vs cens_adjust at analysis_time = 84 (n = 1000).")
```

# Connecting back to `fs_fdr_report()`

Operating characteristics build directly on this DGM: `fs_fdr_report()` draws
H0-hom and/or H0-harm from `setup_gbsg_dgm()`, runs the *analysis's own*
identifier configuration (inherited from `args_call_all`) on each draw, and
tabulates A and B(g). The DGM options above are the levers behind those nulls —
`k_inter` and `calibrate_k_inter()` set the effect structure; `cens_adjust` and
`analysis_time` set the censoring; `n` (defaulting to your analysis's sample
size) sets the trial size.

```{r fdr-pointer}
#| eval: false
# The operating-characteristics run (see the fs_fdr_report smoke doc for the
# full harness). This is where the DGM above becomes an FDR:
# rep <- fs_fdr_report(fit, df0, n = 700L, nsims = 1000L, n_workers = 100L)
# rep$fdr          # A (raw FDR) and B(g) (de-biased-gated FDR) by null x c_gate
```

```{r session}
#| echo: false
sessionInfo()
```
