---
title: "Exploratory Subgroup Identification with ForestSearch"
subtitle: "German Breast Cancer Study Group (GBSG) Analysis"
author: "Larry F. León"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
    theme: cosmo
    highlight: tango
    fig_width: 8
    fig_height: 6
vignette: >
  %\VignetteIndexEntry{Exploratory Subgroup Identification with ForestSearch}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(

  echo = TRUE,
  warning = FALSE,
  message = FALSE,

  fig.align = "center",
  fig.retina = 2,
  fig.width = 8,
  fig.height = 6
)

# Bootstrap 5 code-folding hook for pkgdown compatibility
# Usage: add `echo=FALSE, code_fold=TRUE` to any chunk header
knitr::knit_hooks$set(code_fold = function(before, options, envir) {
  if (!before && isTRUE(options$code_fold)) {
    id <- gsub("[^a-zA-Z0-9]", "", options$label)
    code_lines <- knitr::knit_code$get(options$label)
    code_text  <- htmltools::htmlEscape(paste(code_lines, collapse = "\n"))
    sprintf(
      '<p><a class="btn btn-outline-secondary btn-sm" data-bs-toggle="collapse" href="#collapse-%s" role="button" aria-expanded="false" aria-controls="collapse-%s"><i class="bi bi-code-slash"></i> Code</a></p>\n<div class="collapse" id="collapse-%s"><div class="card card-body p-0"><pre class="r"><code class="hljs">%s</code></pre></div></div>',
      id, id, id, code_text
    )
  }
})

# Initialize timing
timings <- list()
t_vignette_start <- proc.time()
```

# Introduction

This vignette demonstrates the **ForestSearch** methodology for exploratory 
subgroup identification in survival analysis, as described in 
León et al. (2024) *Statistics in Medicine*.

## Motivation

In clinical trials, particularly oncology, subgroup analyses are essential for:

- Evaluating treatment effect consistency across patient populations
- Identifying subgroups where treatment may be detrimental (harm)
- Characterizing subgroups with enhanced benefit
- Informing regulatory decisions and clinical practice

While prespecified subgroups provide stronger evidence, important subgroups 
based on patient characteristics may not be anticipated. ForestSearch provides 
a principled approach to **exploratory** subgroup identification with proper 
statistical inference.

## Methodology Overview

ForestSearch identifies subgroups through:
 
1. **Candidate factor selection**: Using LASSO and/or Generalized Random Forests (GRF)
2. **Exhaustive subgroup search**: Evaluating all combinations up to `maxk` factors
3. **Consistency-based selection**: Applying splitting consistency criteria
4. **Bootstrap bias correction**: Adjusting for selection-induced optimism
5. **Cross-validation**: Assessing algorithm stability

The key innovation is the **splitting consistency criterion**: a subgroup is 
considered "consistent with harm" if, when randomly split 50/50 many times, 
both halves consistently show hazard ratios ≥ 1.0 (for example if 1.0 represents a meaningful "harm threshold").

# Setup

## Load Required Packages

```{r load-packages}
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(policytree)
library(doFuture)

# Optional packages for enhanced output
library(patchwork)
library(weightedsurv)

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```


# Data: German Breast Cancer Study Group Trial

## Study Background

The GBSG trial evaluated hormonal treatment (tamoxifen) versus chemotherapy 
in node-positive breast cancer patients. Key characteristics:

- **Sample size**: N = 686
- **Outcome**: Recurrence-free survival time
- **Censoring rate**: ~56%
- **Treatment**: Hormonal therapy (tamoxifen) vs. chemotherapy

## Data Preparation

```{r data-setup}
# Load GBSG data (included in forestsearch package)
df.analysis <- gbsg

# Prepare analysis variables
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
  treat <- hormon
})

# Define variable roles
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name <- "status"
id.name <- "id"
treat.name <- "hormon"

# Display data structure
cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]), 
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```

## Baseline Characteristics

```{r baseline-table}
create_summary_table(
  data                  = df.analysis,                                       # required
  treat_var             = treat.name,                                        # default "treat"
  vars_continuous       = c("age", "nodes", "size", "er", "pgr"),            # default NULL
  vars_categorical      = c("grade", "meno"),                                # default NULL
  vars_binary           = NULL,                                              # default NULL
  var_labels            = NULL,                                              # default NULL
  digits                = 1,                                                 # default 1
  show_pvalue           = TRUE,                                              # default TRUE
  show_smd              = TRUE,                                              # default TRUE
  show_missing          = TRUE,                                              # default TRUE
  table_title           = "GBSG Baseline Characteristics by Treatment Arm",  # default "Baseline Characteristics by Treatment Arm"
  table_subtitle        = NULL,                                              # default NULL
  source_note           = NULL,                                              # default NULL
  font_size             = 12,                                                # default 12
  header_font_size      = 14,                                                # default 14
  footnote_font_size    = 10,                                                # default 10
  use_alternating_rows  = TRUE,                                              # default TRUE
  stripe_color          = "#f9f9f9",                                         # default "#f9f9f9"
  indent_size           = 20,                                                # default 20
  highlight_pval        = 0.05,                                              # default 0.05
  highlight_smd         = 0.2,                                               # default 0.2
  highlight_color       = "#fff3cd",                                         # default "#fff3cd"
  compact_mode          = FALSE,                                             # default FALSE
  column_width_var      = 200,                                               # default 200
  column_width_stats    = 120,                                               # default 120
  show_column_borders   = FALSE                                              # default FALSE
)
```

## Kaplan-Meier Analysis (ITT Population)

```{r km-itt, fig.width=8, fig.height=5}
# Prepare counting process data for KM plot
dfcount <- df_counting(
  df = df.analysis,
  by.risk = 6,
  tte.name = outcome.name,
  event.name = event.name,
  treat.name = treat.name
)

# Plot with confidence intervals and log-rank test
plot_weighted_km(
  dfcount,
  conf.int = TRUE,
  show.logrank = TRUE,
  ymax = 1.05,
  xmed.fraction = 0.775,
  ymed.offset = 0.125
)
```

The ITT Cox hazard ratio estimate is approximately 0.69 (95% CI: 0.54, 0.89), 
suggesting an overall benefit for hormonal therapy.

# Preliminary Analysis: Generalized Random Forests

Before running ForestSearch, we can use GRF to explore potential treatment 
effect heterogeneity and identify candidate factors.

```{r grf-analysis}
t0 <- proc.time()

# All grf.subg.harm.survival() parameters shown with their default values.
# Required: data, confounders.name, outcome.name, event.name, id.name, treat.name.
grf_est <- grf.subg.harm.survival(
  data                      = df.analysis,         # required
  confounders.name          = confounders.name,    # required
  outcome.name              = outcome.name,        # required
  event.name                = event.name,          # required
  id.name                   = id.name,             # required
  treat.name                = treat.name,          # required
  frac.tau                  = 0.6,                 # default 1.0; max follow-up fraction
  n.min                     = 60,                  # default 60; min subgroup size
  dmin.grf                  = 12,                  # default 0.0; min RMST diff (months)
  RCT                       = TRUE,                # default TRUE
  details                   = TRUE,                # default FALSE
  sg.criterion              = "mDiff",             # default "mDiff"; or "Nsg"
  maxdepth                  = 2,                   # default 2; max policy tree depth (<=3)
  seedit                    = 8316951,             # default 8316951
  return_selected_cuts_only = FALSE,               # default FALSE
  tune_grf                  = FALSE                # default FALSE; CV hyperparameter tuning
)

timings$grf <- (proc.time() - t0)["elapsed"]
```

```{r grf-trees, fig.width=10, fig.height=4}
# Display policy trees
# leaf1 = recommend control, leaf2 = recommend treatment
par(mfrow = c(1, 2))
plot(grf_est$tree1, leaf.labels = c("Control", "Treat"), main = "Depth 1")
plot(grf_est$tree2, leaf.labels = c("Control", "Treat"), main = "Depth 2")
par(mfrow = c(1, 1))
```

GRF identifies estrogen receptor status (ER) as a key factor, with ER ≤ 0 
suggesting potential harm from hormonal therapy.

# ForestSearch Analysis


## Parallel Processing Configuration
 
ForestSearch supports parallel processing for computationally intensive 
operations (bootstrap, cross-validation).

```{r parallel-setup}
# Detect available cores (limited to 2 cores for CRAN checks)
n_cores <- 2
n_cores_total <- parallel::detectCores()
cat("Using", n_cores, "of", n_cores_total, "total cores for parallel processing")
```



## Running ForestSearch

ForestSearch performs an exhaustive search over candidate subgroup combinations 
with up to `maxk` factors. Key parameters:

| Parameter | Value | Description |
|-----------|-------|-------------|
| `hr.threshold` | 1.25 | Minimum HR for consistency evaluation |
| `hr.consistency` | 1.0 | Minimum consistency rate for candidates |
| `pconsistency.threshold` | 0.90 | Required consistency for selection |
| `maxk` | 2 | Maximum factors in subgroup definition |
| `n.min` | 60 | Minimum subgroup sample size |
| `d0.min`, `d1.min` | 12 | Minimum events per treatment arm |

```{r forestsearch-main, fig.width=10, fig.height=7}
t0 <- proc.time()

# All forestsearch() parameters shown with their default values.
# Comments indicate the package default; the value used here may differ.
fs <- forestsearch(
  df.analysis                = df.analysis,                  # required

  # ─── Variable names ─────────────────────────────────────────────────────
  outcome.name               = outcome.name,                 # default "tte"
  event.name                 = event.name,                   # default "event"
  treat.name                 = treat.name,                   # default "treat"
  id.name                    = id.name,                      # default "id"
  potentialOutcome.name      = NULL,                         # default NULL
  flag_harm.name             = NULL,                         # default NULL
  confounders.name           = confounders.name,             # default NULL

  # ─── Parallel processing ────────────────────────────────────────────────
  parallel_args              = list(plan = "multisession",
                                    workers = n_cores,
                                    show_message = TRUE),    # default plan="multisession"

  # ─── Prediction / RCT flag ──────────────────────────────────────────────
  df.predict                 = NULL,                         # default NULL
  df.test                    = NULL,                         # default NULL
  is.RCT                     = TRUE,                         # default TRUE
  seedit                     = 8316951,                      # default 8316951
  est.scale                  = "hr",                         # default "hr"

  # ─── Factor selection: LASSO + GRF ──────────────────────────────────────
  use_lasso                  = TRUE,                         # default TRUE
  use_grf                    = TRUE,                         # default TRUE
  grf_res                    = NULL,                         # default NULL; pass cached GRF result
  grf_cuts                   = NULL,                         # default NULL
  max_n_confounders          = 1000,                         # default 1000
  grf_depth                  = 2,                            # default 2
  dmin.grf                   = 0.0,                          # default 0.0
  frac.tau                   = 0.8,                          # default 0.8
  return_selected_cuts_only  = TRUE,                         # default TRUE
  vi.grf.min                 = -0.2,                         # default -0.2; GRF VI threshold
  tune_grf                   = FALSE,                        # default FALSE

  # ─── Cuts / discretization ──────────────────────────────────────────────
  conf_force                 = NULL,                         # default NULL
  defaultcut_names           = NULL,                         # default NULL
  cut_type                   = "default",                    # default "default"
  exclude_cuts               = NULL,                         # default NULL
  replace_med_grf            = FALSE,                        # default FALSE
  cont.cutoff                = 4,                            # default 4
  conf.cont_medians          = NULL,                         # default NULL
  conf.cont_medians_force    = NULL,                         # default NULL
  conf.cont_jcuts            = NULL,                         # default NULL

  # ─── Subgroup constraints ───────────────────────────────────────────────
  n.min                      = 60,                           # default 60; min subgroup size
  d0.min                     = 12,                           # default 10; min events arm 0
  d1.min                     = 12,                           # default 10; min events arm 1
  maxk                       = 2,                            # default 2; max factors per SG
  max_subgroups_search       = 3,                            # default 10
  max.minutes                = 3,                            # default 3; per-stage timeout

  # ─── Thresholds (preferred names take precedence if both supplied) ─────
  effect.threshold           = NULL,                         # default NULL; alias for hr.threshold
  consistency.threshold      = NULL,                         # default NULL; alias for hr.consistency
  hr.threshold               = 1.25,                         # default 1.25; screening threshold
  hr.consistency             = 1.0,                          # default 1.0; consistency threshold

  # ─── Selection (sg_focus / Pareto neighborhood) ─────────────────────────
  sg_focus                   = "maxSG",                    # default "hr"; "hr"/"eff", "maxSG", "minSG", "hrMaxSG"/"effMaxSG", "hrMinSG"/"effMinSG"
  selection_rule             = "neighborhood",               # default "neighborhood"; or "pareto", "both" (only for hrMaxSG/hrMinSG)
  effect_neighborhood        = 0.10,                         # default 0.10; tol for hrMaxSG/hrMinSG

  # ─── Consistency evaluation ─────────────────────────────────────────────
  fs.splits                  = 100,                          # default 1000; split-half repeats
  m1.threshold               = Inf,                          # default Inf
  pconsistency.threshold     = 0.80,                         # default 0.90
  stop_threshold             = 0.80,                         # default 0.95; early-stop consistency

  # ─── Two-stage consistency ──────────────────────────────────────────────
  use_twostage               = TRUE,                         # default TRUE
  twostage_args              = list(),                       # default list()

  # ─── GLM outcome support (survival analysis here) ───────────────────────
  outcome_type               = "survival",                   # default "survival"; or "binary", "continuous", "count"
  effect_measure             = NULL,                         # default NULL; auto-set from outcome_type
  offset.name                = NULL,                         # default NULL; required for count
  adverse_outcome            = NULL,                         # default NULL; auto-set
  overdispersion             = "none",                       # default "none"; or "quasi", "negbin"
  grf_count_transform        = "log",                        # default "log"; or "identity"

  # ─── Propensity-score adjustment (observational analyses) ───────────────
  ps_method                  = NULL,                         # default NULL
  ps_adjust_method           = "none",                       # default "none"; or "iptw", "dr_gcomp"
  ps_hat                     = NULL,                         # default NULL

  # ─── Output / diagnostics ───────────────────────────────────────────────
  show_candidate_summary     = TRUE,                         # default FALSE; pre-/post-consistency previews
  minp                       = 0.025,                        # default 0.025
  details                    = TRUE,                         # default FALSE
  quiet                      = FALSE,                        # default FALSE
  by.risk                    = 12,                           # default 12
  plot.sg                    = TRUE,                         # default FALSE; KM plots for SG
  plot.grf                   = FALSE                         # default FALSE
)

plan("sequential")
timings$forestsearch <- (proc.time() - t0)["elapsed"]

cat("\nForestSearch completed in", 
    round(timings$forestsearch, 1), "seconds\n")
```




## ForestSearch Results

### Identified Subgroups

```{r fs-results}
# Generate results tables
# All sg_tables() parameters shown with their default values.
res_tabs <- sg_tables(
  fs                    = fs,                                # required
  which_df              = "est",                             # default "est"
  est_title             = "Treatment Effect Estimates",      # default "Treatment Effect Estimates"
  est_caption           = "Training data estimates",         # default "Training data estimates"
  sg_title              = "Identified Subgroups",            # default "Identified Subgroups"
  sg_subtitle           = NULL,                              # default NULL
  potentialOutcome.name = NULL,                              # default NULL
  hr_1a                 = NA,                                # default NA
  hr_0a                 = NA,                                # default NA
  ndecimals             = 3,                                 # default 3
  include_search_info   = TRUE,                              # default TRUE
  subgroup_notation     = NULL,                              # default NULL; or "harm", "benefit"
  font_size             = 12                                 # default 12
)

# Display top subgroups meeting criteria
res_tabs$sg10_out
```

### Treatment Effect Estimates

```{r fs-estimates}
# ITT and subgroup estimates
res_tabs$tab_estimates
```

### Identified Subgroup Definition

```{r fs-subgroup}
cat("Identified subgroup (H):", paste(fs$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs$df.est$treat.recommend == 0), 
    sprintf("(%.1f%% of ITT)\n", 
            100 * mean(fs$df.est$treat.recommend == 0)))
```

ForestSearch identifies **Estrogen ≤ 0** (ER-negative) as the subgroup with 
potential harm. This is biologically plausible: tamoxifen is a selective 
estrogen receptor modulator with limited efficacy in ER-negative tumors.

# Bootstrap Bias Correction

## Rationale

Cox model estimates from identified subgroups are **upwardly biased** due to 
the selection process (subgroups are selected *because* they show extreme effects). 
Bootstrap bias correction addresses this by:

1. Resampling with replacement
2. Re-running the entire ForestSearch algorithm
3. Computing bias terms from bootstrap vs. observed estimates
4. Applying infinitesimal jackknife variance estimation

## Running Bootstrap Analysis

```{r bootstrap, eval=TRUE}
# Number of bootstrap iterations
# Use 500-2000 for production; reduced here for vignette
NB <- 2

t0 <- proc.time()

# All forestsearch_bootstrap_dofuture() parameters shown with their defaults.
fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est        = fs,                                        # required
  nb_boots      = NB,                                        # required (use 500-2000 for production)
  seed          = 8316951L,                                  # default 8316951L
  details       = FALSE,                                     # default FALSE
  show_three    = FALSE,                                     # default FALSE
  parallel_args = list(plan = "multisession",
                       workers = n_cores,
                       show_message = TRUE),                 # default list()
  digits        = 4                                          # default 4
)

plan("sequential")
timings$bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nBootstrap completed in", 
    round(timings$bootstrap / 60, 1), "minutes\n")
```

## Bootstrap Summary and Diagnostics

```{r bootstrap-summary}
# All summarize_bootstrap_results() parameters shown with their defaults.
summaries <- summarize_bootstrap_results(
  sgharm       = fs$sg.harm,                                 # required
  boot_results = fs_bc,                                      # required
  create_plots = TRUE,                                       # default FALSE
  est.scale    = "hr",                                       # default "hr"
  digits       = 2                                           # default 2
)

# Display bias-corrected estimates table
summaries$table
```

## Kaplan-Meier by Identified Subgroups

```{r fs-results_figure, fig.width=10, fig.height=7, fig.cap="Kaplan-Meier survival curves by identified subgroup"}

# All plot_sg_weighted_km() parameters shown with their defaults.
km_result <- plot_sg_weighted_km(
  fs.est         = fs,                                       # required
  fs_bc          = fs_bc,                                    # default NULL
  outcome.name   = "time_months",                            # default "Y"
  event.name     = "status",                                 # default "Event"
  treat.name     = "hormon",                                 # default "Treat"
  by.risk        = 12,                                       # default NULL
  sg0_name       = NULL,                                     # default NULL
  sg1_name       = NULL,                                     # default NULL
  conf.int       = TRUE,                                     # default TRUE
  show.logrank   = FALSE,                                    # default TRUE
  show.cox       = FALSE,                                    # default TRUE
  show.cox.bc    = TRUE,                                     # default TRUE
  put.legend.lr  = "topleft",                                # default "topleft"
  ymax           = 1.05,                                     # default 1.05
  xmed.fraction  = 0.65,                                     # default 0.65
  hr_bc_position = "topright",                               # default "bottomright"
  hr_bc_cex      = 0.725,                                    # default 0.725
  title          = NULL,                                     # default NULL
  verbose        = FALSE                                     # default FALSE
)
```
::: {.figure-note}
`r figure_note(km_result, prefix = "**Note:** ", include_definition = TRUE, include_hr_explanation = TRUE, custom_text = "Medians [95% CI] for arms are un-adjusted.")`
:::


### Event Count Summary

Low event counts can lead to unstable HR estimates. This summary helps 
identify potential issues:

```{r event-summary}
# note that default required minimum events is 12 for subgroup candidate
# Here we evaluate frequency of subgroup candidates in bootstrap samples less than 15
# All summarize_bootstrap_events() parameters shown with their defaults.
event_summary <- summarize_bootstrap_events(
  boot_results = fs_bc,                                      # required
  threshold    = 15                                          # default 5
)
```

### Bootstrap Diagnostics

```{r bootstrap-diagnostics}
# Quality metrics
summaries$diagnostics_table_gt
```

### Subgroup Agreement

How consistently does bootstrap identify the same subgroup?

```{r subgroup-agreement}
# Agreement with original analysis
if (!is.null(summaries$subgroup_summary$original_agreement)) {
  summaries$subgroup_summary$original_agreement
}

# Factor presence across bootstrap iterations
if (!is.null(summaries$subgroup_summary$factor_presence)) {
  summaries$subgroup_summary$factor_presence
}
```

### Bootstrap Distributions

```{r bootstrap-plots, fig.width=10, fig.height=4, eval=TRUE}
if (!is.null(summaries$plots)) {
  summaries$plots$H_distribution + summaries$plots$Hc_distribution
}
```

# Cross-Validation

Cross-validation assesses the **stability** of the ForestSearch algorithm. 
Two approaches are available:

## K-Fold Cross-Validation

```{r kfold-cv, eval = TRUE}
# 10-fold CV with multiple iterations
# Use Ksims >= 50 for production
Ksims <- 1

t0 <- proc.time()

# All forestsearch_tenfold() parameters shown with their defaults.
fs_kfold <- forestsearch_tenfold(
  fs.est        = fs,                                        # required
  sims          = Ksims,                                     # required (use >=50 for production)
  Kfolds        = 2,                                         # default 10
  details       = FALSE,                                     # default TRUE
  seed          = 8316951L,                                  # default 8316951L
  parallel_args = list(plan = "multisession",
                       workers = n_cores,
                       show_message = FALSE),                # default plan="multisession", workers=6
  keep_resCV    = FALSE                                      # default FALSE
)

plan("sequential")
timings$kfold <- (proc.time() - t0)["elapsed"]

# All cv_metrics_tables() parameters shown with their defaults.
metrics_tables <- cv_metrics_tables(
  cv_result        = fs_kfold,                               # required
  sg_definition    = NULL,                                   # default NULL
  title            = "Cross-Validation Metrics",             # default "Cross-Validation Metrics"
  show_percentages = TRUE,                                   # default TRUE
  digits           = 1,                                      # default 1
  include_raw      = FALSE,                                  # default FALSE
  table_style      = "combined",                             # default "combined"; or "separate", "minimal"
  use_gt           = TRUE                                    # default TRUE
)
metrics_tables

```

## Out-of-Bag (N-Fold) Cross-Validation

N-fold CV (leave-one-out):

```{r oob-cv, eval = FALSE}
t0 <- proc.time()

# All forestsearch_Kfold() parameters shown with their defaults.
fs_OOB <- forestsearch_Kfold(
  fs.est        = fs,                                        # required
  Kfolds        = round(nrow(df.analysis)/100, 0),           # default nrow(fs.est$df.est) (LOO)
  seedit        = 8316951L,                                  # default 8316951L
  parallel_args = list(plan = "multisession",
                       workers = n_cores,
                       show_message = TRUE),                 # default plan="multisession", workers=6
  sg0.name      = "Not recommend",                           # default "Not recommend"
  sg1.name      = "Recommend",                               # default "Recommend"
  details       = FALSE                                      # default FALSE
)

plan("sequential")
timings$oob <- (proc.time() - t0)["elapsed"]

# All forestsearch_KfoldOut() parameters shown with their defaults.
cv_out <- forestsearch_KfoldOut(
  res     = fs_OOB,                                          # required
  details = FALSE,                                           # default FALSE
  outall  = TRUE,                                            # default FALSE
  digits  = 4                                                # default 4
)

# All cv_summary_tables() parameters shown with their defaults.
tables <- cv_summary_tables(
  kfold_out    = cv_out,                                     # required
  title        = "Cross-Validation Summary",                 # default "Cross-Validation Summary"
  subtitle     = NULL,                                       # default NULL
  show_metrics = TRUE,                                       # default TRUE
  digits       = 3,                                          # default 3
  font_size    = 12,                                         # default 12
  use_gt       = TRUE                                        # default TRUE
)

tables$combined_table

tables$metrics_table




```

# Results Visualization

## Forest Plot

The forest plot summarizes treatment effects across the ITT population, 
reference subgroups, and identified subgroups with cross-validation metrics.

```{r forest-plot, fig.width=18, fig.height=12, fig.cap="Subgroup forest plot including identified subgroups"}


# Define reference subgroups for comparison
subgroups <- list(
  age_gt65 = list(
    subset_expr = "age > 65",
    name = "Age > 65",
    type = "reference"
  ),
  age_le65 = list(
    subset_expr = "age <= 65",
    name = "Age ≤ 65",
    type = "reference"
  ),
  pgr_positive = list(
    subset_expr = "pgr > 0",
    name = "PgR > 0",
    type = "reference"
  ),
  pgr_negative = list(
    subset_expr = "pgr <= 0",
    name = "PgR ≤ 0",
    type = "reference"
  )
)


# All create_forest_theme() parameters shown with their defaults.
my_theme <- create_forest_theme(
  base_size         = 24,                                    # default 10
  scale             = 1.0,                                   # default 1.0
  row_padding       = NULL,                                  # default NULL
  ci_pch            = 15,                                    # default 15
  ci_lwd            = NULL,                                  # default NULL
  ci_Theight        = NULL,                                  # default NULL
  ci_col            = "black",                               # default "black"
  header_fontsize   = NULL,                                  # default NULL (auto)
  body_fontsize     = NULL,                                  # default NULL (auto)
  footnote_fontsize = 17,                                    # default NULL
  footnote_col      = "darkcyan",                            # default "darkcyan"
  title_fontsize    = NULL,                                  # default NULL
  cv_fontsize       = 22,                                    # default NULL
  cv_col            = "gray30",                              # default "gray30"
  refline_lwd       = NULL,                                  # default NULL
  refline_lty       = "dashed",                              # default "dashed"
  refline_col       = "gray30",                              # default "gray30"
  vertline_lwd      = NULL,                                  # default NULL
  vertline_lty      = "dashed",                              # default "dashed"
  vertline_col      = "gray20",                              # default "gray20"
  arrow_type        = "closed",                              # default "closed"
  arrow_col         = "black",                               # default "black"
  summary_fill      = "black",                               # default "black"
  summary_col       = "black"                                # default "black"
)


# Create forest plot
# Include fs_kfold and fs_OOB if available for CV metrics
# All plot_subgroup_results_forestplot() parameters shown with their defaults.
result <- plot_subgroup_results_forestplot(
  fs_results        = list(fs.est   = fs,
                           fs_bc    = fs_bc,
                           fs_OOB   = NULL,
                           fs_kfold = fs_kfold),             # required
  df_analysis       = df.analysis,                           # required
  subgroup_list     = subgroups,                             # default NULL
  outcome.name      = outcome.name,                          # required
  event.name        = event.name,                            # required
  treat.name        = treat.name,                            # required
  E.name            = "Hormonal",                            # default "Experimental"
  C.name            = "Chemo",                               # default "Control"
  est.scale         = "hr",                                  # default "hr"
  xlog              = TRUE,                                  # default TRUE
  title_text        = NULL,                                  # default NULL
  arrow_text        = c("Favors Experimental", "Favors Control"),  # default same
  footnote_text     = c("Eg 80% of training found SG: 70% of B (+) also B in CV testing"),  # default same
  xlim              = c(0.25, 1.5),                          # default c(0.25, 1.5)
  ticks_at          = c(0.25, 0.70, 1.0, 1.5),               # default c(0.25, 0.70, 1.0, 1.5)
  show_cv_metrics   = TRUE,                                  # default TRUE
  cv_source         = "auto",                                # default "auto"; or "kfold", "oob", "both"
  posthoc_colors    = c("powderblue", "beige"),              # default same
  reference_colors  = c("yellow", "powderblue"),             # default same
  ci_column_spaces  = 25,                                    # default 20
  conf.level        = 0.95,                                  # default 0.95
  theme             = my_theme,                              # default NULL
  outcome_type      = NULL,                                  # default NULL; auto-detect
  effect_measure    = NULL,                                  # default NULL; auto-detect
  offset.name       = NULL,                                  # default NULL
  extreme_ci_cap    = 1.5,                                   # default 1.5
  xlim_method       = "clinical"                             # default "clinical"; or "data"
)


# Option 2: Custom sizing
# All render_forestplot() parameters shown with their defaults.
render_forestplot(
  x       = result,                                          # required
  newpage = TRUE                                             # default TRUE
)

```


### KM Difference plots: ITT and subgroups 

The solid black line denotes the ITT Kaplan-Meier treatment difference estimates along with 95%95% CIs (the grey shaded region). K-M differences corresponding to subgroups are displayed.


```{r KMdiffs, fig.width = 8, fig.height = 6, fig.align="center"}

# Add additional subgroups along with ITT and identified subgroups
ref_sgs <- list(
age_young = list(subset_expr = "age < 65", color = "brown"),
age_old = list(subset_expr = "age >= 65", color = "orange")
)

# All plot_km_band_forestsearch() parameters shown with their defaults.
plot_km_band_forestsearch(
  df             = df.analysis,                              # required
  fs.est         = fs,                                       # default NULL
  sg_cols        = NULL,                                     # default NULL
  sg_labels      = NULL,                                     # default NULL
  sg_colors      = NULL,                                     # default NULL
  itt_color      = "azure3",                                 # default "azure3"
  outcome.name   = outcome.name,                             # default "tte"
  event.name     = event.name,                               # default "event"
  treat.name     = treat.name,                               # default "treat"
  xlabel         = "Time",                                   # default "Time"
  ylabel         = "Survival differences",                   # default "Survival differences"
  yseq_length    = 5,                                        # default 5
  draws_band     = 20,                                       # default 1000 (lower for speed)
  tau_add        = NULL,                                     # default NULL
  by_risk        = 6,                                        # default 6
  risk_cex       = 0.75,                                     # default 0.75
  risk_delta     = 0.035,                                    # default 0.035
  risk_pad       = 0.015,                                    # default 0.015
  ymax_pad       = 0.11,                                     # default 0.11
  show_legend    = TRUE,                                     # default TRUE
  legend_pos     = "topleft",                                # default "topleft"
  legend_cex     = 0.75,                                     # default 0.75
  ref_subgroups  = ref_sgs,                                  # default NULL
  verbose        = FALSE                                     # default FALSE
)
 
# # Example with more subgroups
# ref_sgs <- list(
# pgr_positive = list(subset_expr = "pgr > 0", color ="green"),
# pgr_negative = list(subset_expr = "pgr <= 0", color = "purple"),
# age_young = list(subset_expr = "age < 65", color = "brown"),
# age_old = list(subset_expr = "age >= 65", color = "orange")
# )

```




# Summary and Interpretation

## Key Findings

```{r summary-findings}
# Extract key results
cat("=" %>% rep(60) %>% paste(collapse = ""), "\n")
cat("FORESTSEARCH ANALYSIS SUMMARY\n")
cat("=" %>% rep(60) %>% paste(collapse = ""), "\n\n")

cat("Dataset: GBSG (N =", nrow(df.analysis), ")\n")
cat("Outcome: Recurrence-free survival\n\n")

cat("ITT Analysis:\n")
cat("  HR (95% CI): 0.69 (0.54, 0.89)\n\n")

cat("Identified Subgroup (H):\n")
cat("  Definition:", paste(fs$sg.harm, collapse = " & "), "\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 0), 
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 0)))
cat("  Unadjusted HR:", sprintf("%.2f", exp(fs$grp.consistency$out_sg$result$hr[1])), "\n")

cat("\nComplement Subgroup (Hc):\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 1),
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 1)))
```

## Clinical Interpretation

The ForestSearch analysis identifies **estrogen receptor-negative (ER ≤ 0)** 
patients as a subgroup with potential lack of benefit from hormonal therapy.

**Biological plausibility**: Tamoxifen is a selective estrogen receptor 
modulator. Its efficacy depends on ER expression. The finding that ER-negative 
patients may not benefit is consistent with:

- Mechanistic understanding of tamoxifen action
- Meta-analyses showing no tamoxifen benefit in ER-negative breast cancer
- Clinical guidelines recommending tamoxifen primarily for ER-positive tumors

**Caveats**:

1. This is an **exploratory** analysis requiring independent validation
2. The bias-corrected estimates have wider confidence intervals
3. Cross-validation metrics should be evaluated for algorithm stability

## Computational Timing

```{r timing-summary, echo=FALSE, code_fold=TRUE}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Analysis = c("GRF", "ForestSearch", "Bootstrap", "Total"),
  Seconds = c(
    timings$grf,
    timings$forestsearch,
    timings$bootstrap,
    timings$total
  )
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(title = "Computational Timing") |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1) |>
  cols_label(
    Analysis = "Component",
    Seconds = "Time (sec)",
    Minutes = "Time (min)"
  )
```

# References

León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024). 
"Exploratory subgroup identification in the heterogeneous Cox model: 
A relatively simple procedure." *Statistics in Medicine*. 
DOI: 10.1002/sim.10163

# Session Information

```{r session-info}
sessionInfo()
```
