---
title: "Survival post-selection inference: the certified products"
subtitle: "A worked GBSG analysis, end to end"
author: "Larry Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    code-fold: false
vignette: >
  %\VignetteIndexEntry{Survival post-selection inference: the certified products}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
#| message: false
#| warning: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")
library(forestsearch)
library(survival)
```

`forestsearch()` identifies a subgroup by searching; anything estimated on that
subgroup afterwards is estimated on a subgroup the data chose. This vignette
runs the pipeline end to end on a public dataset and shows what the package
reports about that selection, and -- as importantly -- what it declines to
certify.

Every claim about coverage below comes from
`dev/notes/NOTE_survival_products_2026-09-09.md`, which is the source of truth
for the certified products; the simulation evidence behind it is named at the
end. Nothing here is new methodology, and no number in this document is
computed by it: the reporting block reads elements the fit already carries.

## The data and the fit

The GBSG node-positive breast-cancer trial shipped with **survival**: 686
patients, hormonal therapy against none, recurrence-free survival. The
covariate set is the usual seven; `er` and `pgr` are forced in as candidate
receptor cutpoints, and continuous covariates get 10 quantile cuts each.

```{r data}
df.analysis <- gbsg
df.analysis <- within(df.analysis, {
  id          <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3      <- ifelse(grade == "3", 1, 0)
})
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
cat("Sample size:", nrow(df.analysis),
    "| events:", sum(df.analysis$status),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis$status)))
```

The fit below is the applications configuration cut to vignette size: the
consistency search with no LASSO/GRF/DINA screen, `sg_focus = "effMaxSG"`
(largest subgroup within a 20% band of the strongest observed effect), 200
split-sample replicates and 500 multiplier draws. `mr_inference = TRUE` turns
on the post-selection gate, whose result is attached to the fit as
`fs$mr_inference`; no separate run is needed. There is no bootstrap, no
cross-validation and no leave-one-out here -- those live in the applications
documents, not in a vignette.

```{r fit}
t0 <- proc.time()
fs <- forestsearch(
  df.analysis,
  outcome.name = "time_months", event.name = "status",
  treat.name = "hormon", id.name = "id",
  confounders.name = confounders.name,
  is.RCT = TRUE, seedit = 8316951, est.scale = "hr",
  use_lasso = FALSE, use_grf = FALSE, use_dina = FALSE,
  conf.cont_jcuts = list(er = 10, pgr = 10), collapse_cuts = TRUE,
  conf_force = c("er <= 0", "pgr <= 0"),
  subgroup_method = "consistency", sg_focus = "effMaxSG",
  selection_rule = "neighborhood", effect_neighborhood = 0.20,
  mr_inference = TRUE, mr_inference_args = list(draws = 500L),
  n.min = 60, d0.min = 10, d1.min = 10, maxk = 2,
  hr.threshold = 1.0, hr.consistency = 1.0, pconsistency.threshold = 0.80,
  fs.splits = 200, max.minutes = 3,
  details = FALSE, quiet = TRUE, plot.sg = FALSE,
  parallel_args = list(plan = "sequential"))
cat(sprintf("fit time: %.1f seconds\n", (proc.time() - t0)[["elapsed"]]))
```

## The identified subgroup

```{r subgroup}
cat("Selected subgroup:", paste(fs$sg.harm, collapse = " & "), "\n")
```

The subgroup is a conjunction of at most `maxk = 2` binary cutpoints drawn from
the candidate family. `family_status` records how that family was built --
`"no-front-end"` here, meaning no fitted model on the observed data shaped it,
which is the weakest form of selection the package will report on.

## The certified products

`print()` reports them; nothing extra needs to be called.

```{r print}
print(fs)
```

Read top to bottom, this is the package's order of standing.

**The harm subgroup's one-sided lower bound** on `HR(Hhat)` is the *field*
construction. Its coverage is 0.944--0.974 across ten harm cells at 12.4% and
31% prevalence, n = 500 / 1000 / 1500, and it is flat in n.

**The complement's one-sided upper bound** on `HR(Hhat^c)` is *field-s*, the
studentized complement field. Its small-sample shortfall closes with n:
0.912 -> 0.942 -> 0.947 at 31% prevalence with HR 1.50, and 0.919 -> 0.942 ->
0.946 at HR 1.75; it is at or above 0.94 at every cell of the 12.4% regime.
field-s is at or above the unscaled field in all thirteen certification cells.

**The joint two-subgroup claim** is the Bonferroni pair at gamma = 0.025 each
side, covering 0.939--0.963 at every cell. Use it when both statements are
made together.

Both field bounds exist only under `ci_method = "field"`. That is the default;
`ci_method = "ij"` omits the field block entirely and leaves only the
two-sided interval.

`summary()` adds the standard errors, the re-selection mass and the
certified/not-certified paragraph:

```{r summary}
summary(fs)
```

## How to read a bound: by location

These bounds are read **by location against a clinically meaningful effect
size**, never as significance at the null.

The harm-side question is *how much harm the data can rule out*. A one-sided
lower bound in the high 0.8s or above, against a naive subgroup estimate well
above 1, says the adjusted evidence does not establish clinically meaningful
harm -- the naive estimate was largely selection.

The complement-side question is a benefit claim: *is the one-sided upper bound
below a threshold that would matter?* Name the threshold you are reading
against -- 0.80 and 0.85 are the usual reading aids in these records -- and say
whether the bound clears it. A threshold quoted this way is a reading aid, not
a decision rule.

```{r locations}
g <- fs$mr_inference
cat(sprintf("naive HR on Hhat            %.3f\n", g$naive$est))
cat(sprintf("de-biased HR on Hhat        %.3f\n", g$debiased$est))
cat(sprintf("field one-sided lower       %.3f\n", g$field$lower_1s))
cat(sprintf("field-s one-sided upper Hc  %.3f  (vs 0.80 / 0.85)\n",
            g$field$complement$upper_1s_s))
```

The distance between the naive estimate and the de-biased one is what
selection cost; the spread across the constructions is what the adjustments
pay for it.

## p-hat as a diagnostic

`p_hat(Hhat)` is the frequency with which the gate's own re-selection map
picks the same subgroup across the multiplier draws. It is a **recorded
diagnostic**: it is computed after the bounds are formed, and **no construction
reads it**. It is also estimated from the same draws that build the bound, so
it is a flag, not an independent instrument -- which is a reason not to let any
interval depend on it.

```{r phat}
ph  <- g$reselection$p_hat
lab <- g$selected_label
cat(sprintf("p-hat(Hhat) = %.3f over a family of %d candidates\n",
            unname(ph[[lab]]), g$n_family))
print(round(sort(ph, decreasing = TRUE)[1:5], 3))
```

Harm-block bias is a monotone increasing function of p-hat, crossing zero near
p-hat ~ 0.5:

- **Low p-hat, below 0.20** -- the pick is unstable and the correction runs the
  other way, *over*-correcting (bias -0.11 to -0.28 log units at 12.4%
  prevalence, n = 1500). The one-sided lower bound on Hhat is the conservative
  product here; the two-sided interval is the exposed one. `print()` says so.
- **High p-hat, 0.5 and above** -- the stable-pick regime, where the correction
  *under*-corrects (+0.02 log units) and the bounds are somewhat optimistic.
  `print()` says so.
- **In between, 0.20 to 0.5** -- no line is printed, because the NOTE supports
  no claim there.

Those two edges come from the NOTE: 0.5 is where it puts the bias crossing
zero, and 0.20 is the top of the band it characterises as low. They are
**descriptive band edges, not calibrated thresholds** -- they tell you which
pole an analysis sits at, and the flag is directional, not a decision rule.
The complement's coverage, by contrast, is a stable function of p-hat.

This fit sits at the low pole (p-hat = 0.006), so the printed output carries
the unstable-pick line and not the stable-pick one, and the harm-side lower
bound should be read as the conservative product.

## What is not certified

**No two-sided interval is certified.** The IJ two-term two-sided interval is
retained and reported as the secondary, conservative option, and it is the only
two-sided construction offered -- but its harm-block coverage falls to
0.913--0.917 at 12.4% prevalence once n is 1000 or more (against 0.971--0.981
at 31%). The field's own two-sided interval is lower still, 0.878--0.930. Read
two-sided statements at low prevalence and large n with that caveat.

Also on record: `effect_neighborhood > 0.25` is not adoptable; under a pure-size
pick (`sg_focus = "maxSG"`) the naive SE mis-calibrates, at 0.876 of the error
SD, and field-s inherits that.

The one-sided products certify and the two-sided one does not for a reason
worth stating: each one-sided bound is exposed to one pole of the p-hat bias
only, and in the conservative direction. A two-sided interval is exposed to
both.

## Where the evidence lives

- `dev/notes/NOTE_survival_products_2026-09-09.md` -- the certified products
  and their caveats; the source of truth for every claim above.
- `REPORT_cert20_2026-09-08.md` -- the 31%-prevalence sample-size profile.
- `REPORT_tier2_2026-09-08.md` -- the 12.4% dominated regime.
- `REPORT_fixedphat_ij2s_2026-09-09.md` -- the fixed-p-hat diagnostics, the
  harm-block two-sided decay and detection conditioning.
- `REPORT_cimethod_flip_2026-09-09.md` -- why `ci_method = "field"` is the
  default.

```{r session}
sessionInfo()
```
