---
title: "MRCT Consistency Evaluation via Subgroup Identification"
subtitle: "A ForestSearch Framework with a View Towards Biomarker Causal Effects"
author: "Larry Leon, Masashi Shimura, Xiaojing Cui, Kenichi Takahashi, Shuping Jiang, and William Wang"
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: "Madrid"
    colortheme: "default"
    fonttheme: "structurebold"
    slide_level: 2
    keep_tex: true
    latex_engine: pdflatex
    includes:
      in_header: MRCT_beamer_preamble.tex
classoption: "aspectratio=169"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  comment = "#>",
  dev = "pdf",
  out.width = "95%",
  fig.width = 9,
  fig.height = 5
)
options(warn = -1, width = 72)
rm(list = ls())
```

```{r load-packages, include=FALSE}
suppressWarnings({
  library(survival)
  library(data.table)
  library(ggplot2)
  library(gt)
  library(dplyr)
  library(grf)
  library(grid)
  library(forestploter)
  library(DiagrammeR)
  library(DiagrammeRsvg)
  library(rsvg)
  library(knitr)
  library(kableExtra)
  library(doFuture)
  library(doRNG)
  library(forestsearch)
  library(weightedsurv)
})

registerDoFuture()
registerDoRNG()
theme_set(theme_minimal(base_size = 11))
```


```{r case-study-data}
# Load synthetic case-study dataset
dfsynth <- read.table("data/dfsynthetic.csv",
                       header = TRUE, sep = ",")
df.case <- dfsynth
df.case$AP <- df.case$region_asia
```


```{r create-dgm-spline}
dgm_spline <- generate_aft_dgm_flex(
  data = df.case,
  continuous_vars = c("age", "bm"),
  factor_vars = c("male", "histology",
                   "prior_treat", "AP"),
  set_beta_spec = list(set_var = c("z_AP"),
                       beta_var = -log(5)),
  continuous_vars_cens = c("age"),
  factor_vars_cens = c("prior_treat"),
  cens_type = "weibull",
  outcome_var = "tte", event_var = "event",
  treatment_var = "treat",
  subgroup_vars = NULL, subgroup_cuts = NULL,
  k_inter = 0.0, model = "alt",
  spline_spec = list(
    var = "z_bm", knot = 5, zeta = 10,
    log_hrs = log(c(3, 1.25, 0.5))
  ),
  verbose = FALSE, standardize = FALSE
)
```



# Outline

0. **Regional Consistency**: The MRCT regional consistency challenge
1. **Methodology**: Weibull AFT / Cox model with spline biomarker effects
2. **Estimands**: Marginal HR, Average HR, and Controlled Direct Effect
3. **ForestSearch Algorithm**: Exploratory subgroup identification 
4. **Case Study**: Synthetic MRCT with biomarker-driven treatment heterogeneity
5. **Simulation Design**: DGM specification and per-replicate workflow
6. **Operating Characteristics**: HTE and Null scenarios 
7. **Summary and Discussion**


# The MRCT Consistency Challenge


## Exploratory Subgroup Identification Analyses

\begin{columns}[T]
\begin{column}{0.48\textwidth}

\textbf{The Scenario}

\begin{itemize}
\item A \textbf{global oncology trial} meets statistical significance for the ITT population
\item \textbf{Asia-Pacific (AP)} comprises 15--20\% of the trial
\item \textbf{Consistency criterion}: positive point estimate (Cox model HR $< 1.0$) in AP region
\item \textbf{Problem}: AP regional HR estimate $\approx 1.0$ --- consistency appears to fail
\end{itemize}

\end{column}
\begin{column}{0.48\textwidth}

\begin{block}{Key Question}
Can a subgroup $\mathcal{G}$ with \textbf{enhanced benefit} be identified in non-AP such that AP$(\mathcal{G})$ meets regional consistency?
\end{block}

\smallskip

\textbf{Forest Search (León et al., 2024)}

\smallskip

Imagine a forest plot evaluating all-possible combinations of covariates (e.g. ecog = 0, ecog = 1, bm low, bm high, ecog = 0 \& bm low, etc) 

\smallskip

Signals of harm (Cox estimate $> 1.25$ threshold); \textit{Reproducibility}? 

\end{column}
\end{columns}


# Statistical Framework


## Weibull AFT / Cox Dual Representation

\begin{columns}[T]
\begin{column}{0.48\textwidth}

\textbf{AFT Model} (\texttt{survreg}):
$$\log(T) = \log(\theta) + L'\gamma + \tau\varepsilon$$
where $\tau = 1/\nu$ and $\varepsilon \sim$ extreme value.

\medskip

\textbf{Cox Hazard Model}:
$$\lambda(t;\, L) = \lambda_{0}(t)\,\exp(L'\beta)$$
with Weibull baseline $\lambda_{0}(t) = \nu\theta^{-\nu}t^{\nu-1}$.

\end{column}
\begin{column}{0.48\textwidth}

\begin{alertblock}{AFT-to-Hazard Transformation}
$$\beta = -\gamma\,/\,\tau$$
R's \texttt{survreg} estimates $(\log\hat{\theta}, \hat{\gamma}, \hat{\tau})$ on the AFT scale; hazard-ratio coefficients are obtained by dividing by the scale parameter.
\end{alertblock}

\medskip

Potential outcomes are \emph{simulated} on the AFT scale while treatment effects are \emph{interpreted} on the hazard-ratio scale.

\end{column}
\end{columns}


## Two-Phase Spline Model {.shrink}

Cox linear predictor with biomarker $Z$ and treatment $X$:

$$L'\beta := \beta_{1} X + \beta_{2} Z + \beta_{3} ZX + \beta_{4}(Z-k)\mathbf{1}(Z>k) + \beta_{5}(Z-k)\mathbf{1}(Z>k)X$$

\small

| Parameter | Term | Role |
|:---------:|------|------|
| $\beta_{1}$ | $X$ | Main treatment effect (intercept of log-HR) |
| $\beta_{2}$ | $Z$ | Prognostic biomarker effect (both arms) |
| $\beta_{3}$ | $ZX$ | Biomarker $\times$ treatment interaction ($Z \leq k$) |
| $\beta_{4}$ | $(Z{-}k)\mathbf{1}(Z{>}k)$ | Spline: prognostic slope change above $k$ |
| $\beta_{5}$ | $(Z{-}k)\mathbf{1}(Z{>}k)X$ | Spline: treatment-effect slope change above $k$ |

\normalsize



## Causal Log-HR and Spline Parameterization

\vspace{6pt}

\begin{center}
\small
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{p{0.94\textwidth}}
\hline
\textbf{Causal log-hazard-ratio} for biomarker level $z$: \\[4pt]
$\displaystyle \psi^{0}(z) = \begin{cases} \beta_{1}^{0} + \beta_{3}^{0}\, z, & z \leq k \\[4pt] \beta_{1}^{0} + \beta_{3}^{0}\, z + \beta_{5}^{0}(z-k), & z > k \end{cases}$ \\[6pt]
\hline
\begin{minipage}[t]{0.48\textwidth}
\vspace{4pt}
\textbf{Spline Parameterization:} \\[4pt]
$\displaystyle \beta_{1}^{0} = \psi^{0}(0)$ \\[6pt]
$\displaystyle \beta_{3}^{0} = \frac{\psi^{0}(k) - \psi^{0}(0)}{k}$ \\[6pt]
$\displaystyle \beta_{5}^{0} = \frac{\psi^{0}(\zeta) - \psi^{0}(0) - \beta_{3}^{0}\zeta}{\zeta - k}$
\vspace{4pt}
\end{minipage}
\hfill
\begin{minipage}[t]{0.44\textwidth}
\vspace{4pt}
\textbf{Working Example:} \\[4pt]
\centering
\begin{tabular}{ccc}
$z$ & $\psi^{0}(z)$ & HR$(z)$ \\
\hline
0            & $\log(3.00)$  & 3.00 \\
5 (knot)     & $\log(1.25)$  & 1.25 \\
10 ($\zeta$) & $\log(0.50)$  & 0.50 \\
\end{tabular}
\\[8pt]
\raggedright
Harm at low $z$, crossover near knot, \\
benefit at high $z$.
\vspace{4pt}
\end{minipage} \\
\hline
\end{tabular}
\end{center}


## Potential Outcomes and Estimands {.shrink}

\begin{columns}[T]
\begin{column}{0.52\textwidth}

\textbf{Potential Outcome (PO)} $T[x,z]$, under treatment condition $x$ (set to $x=1[0]$ for treatment[control]) and biomarker level $z$


\begin{alertblock}{PO contrast} $\psi^{0}(z) = \{m(0,z) - m(1,z)\}/\tau$: Difference in true means of POs $\log(T[0,z]) - \log(T[1,z])$.
\end{alertblock}

\textbf{Strong prognostic $W$:} Linear predictor $l_{x,z,w} = \beta_{1}^{0}x + \beta_{2}^{0}z + \beta_{3}^{0}zx + \beta_{4}^{0}(z-k)I(z>k) + \beta_{5}^{0}(z-k)I(z>k)x + \beta_{w}^{0}w$

\begin{alertblock}{PO hazard function (omitting baseline)}
$\theta^{x}(z,v) = \exp(l(x,z,v))$
\end{alertblock}

\end{column}
\begin{column}{0.43\textwidth}

\begin{alertblock}{Average Hazard Ratio (AHR)}
$$\text{AHR}(z^{+}) = \exp\!\bigl\{E_{Z \geq z}\;\psi^{0}(Z)\bigr\}$$
Exponentiated average of differences in true means of POs (causal log-hazard ratios across $z$)
\end{alertblock}

\begin{alertblock}{Controlled Direct Effect (CDE)}
 Let $\bar{\theta}^{x}(z+) = E_{Z \geq z,V}\theta^{x}(Z,V)$


$$\text{CDE}(z^{+}) = \bar{\theta}^{1}(z+)\;/\;\bar{\theta}^{0}(z+)$$

Ratio of average exponentiated hazards
\end{alertblock}

\end{column}
\end{columns}


## Three Treatment Effect Estimands {.fragile .shrink}

\begin{columns}[T]
\begin{column}{0.48\textwidth}
\begin{lstlisting}
dgm_spline <- generate_aft_dgm_flex(
  data = df.case,
  continuous_vars = c("age", "bm"),
  factor_vars = c("male", "histology",
                   "prior_treat", "AP"),
  set_beta_spec = list(set_var = c("z_AP"),
                       beta_var = -log(5)),
  continuous_vars_cens = c("age"),
  factor_vars_cens = c("prior_treat"),
  cens_type = "weibull",
  outcome_var = "tte", event_var = "event",
  treatment_var = "treat",
  subgroup_vars = NULL, subgroup_cuts = NULL,
  k_inter = 0.0, model = "alt",
  spline_spec = list(
    var = "z_bm", knot = 5, zeta = 10,
    log_hrs = log(c(3, 1.25, 0.5))
  ),
  verbose = FALSE, standardize = FALSE
)
\end{lstlisting}
\end{column}
\begin{column}{0.52\textwidth}
```{r, fig.width = 4, fig.height = 3.5}
df_large <- simulate_from_dgm(
  dgm = dgm_spline, n = 5000,
  rand_ratio = 1, draw_treatment = FALSE,
  entry_var = "entrytime",
  analysis_time = 60, cens_adjust = 0,
  seed = 12345
)
results_ahr <- cox_ahr_cde_analysis(
  df = df_large,
  verbose = FALSE,
  tte_name = "y_sim", event_name = "event_sim",
  treat_name = "treat_sim",
  z_name = "z_bm", hr_threshold = NULL,
  plot_style = "grid",
  plot_select = "ahr_only"
)
```
\end{column}
\end{columns}





## Subgroup Defined by (general) $H$ {.shrink}

\begin{columns}[T]
\begin{column}{0.48\textwidth}

\textbf{AFT Model Specification}
$\log T_i(x=1) = \mu + \mathbf{L}_i(1)^{\top} \boldsymbol{\gamma} + \tau\ \varepsilon_i$

The covariate vector includes treatment, prognostic effects ($V$, say), and an interaction:
$$\mathbf{L}_i(x=a)^{\top}\boldsymbol{\gamma} = a \cdot \gamma_{\text{treat}} + \mathbf{V}_i^{\top}\boldsymbol{\gamma}_V + a \cdot H_i \cdot \gamma_{\text{int}}$$
where $H_i \in \{0,1\}$ flags harm-subgroup membership ($\mathcal{S}$).

$$\text{AHR}(\mathcal{S}) = \exp\!\left\{\frac{1}{|\mathcal{S}|}\sum_{i \in \mathcal{S}} \psi_i^{0}\right\}$$

\end{column}

\begin{column}{0.48\textwidth}

The \textbf{marginal hazard ratio} for a subgroup $\mathcal{S}$ is obtained from a Cox model fit to stacked potential outcomes:

\begin{alertblock}

$\text{HR}_{\text{marg}}(\mathcal{S}) = \exp\!\bigl(\hat{\beta}_{\text{Cox}} \;\big|\; \text{data} = \{(T_i(1), 1, 1),\, (T_i(0), 1, 0) : i \in \mathcal{S}\}\bigr)$

\end{alertblock}

\textbf{Computation in ForestSearch}:
\begin{enumerate}
\item Generate shared error: $\varepsilon_i = \log(\text{Exp}(1))$
\item Compute potential survival times:
\begin{align*}
\log T_i(1) &= \mu + \sigma\varepsilon_i + \mathbf{L}_i(1)^{\top}\boldsymbol{\gamma} \\
\log T_i(0) &= \mu + \sigma\varepsilon_i + \mathbf{L}_i(0)^{\top}\boldsymbol{\gamma}
\end{align*}
\item Fit: $\text{HR}_{\text{marg}} = \exp(\hat{\beta}_{\text{treat}})$
\end{enumerate}

\end{column}
\end{columns}



## ForestSearch Algorithm (León et al., 2024, \emph{SIM})

\begin{enumerate}
\item \textbf{Candidate Selection} --- From baseline covariates to $L$ candidate subgroup indicators
  \begin{itemize}
  \item \textit{LASSO}: Cox regularization for prognostic selection; mitigates false discovery from noise
  \item \textit{GRF}: Causal survival forest targets RMST to identify treatment effect heterogeneity
  \item Continuous covariates $\rightarrow$ binary splits (medians, quartiles, GRF cuts)
  \end{itemize}

\item \textbf{Exhaustive Subgroup Search} --- All combos of $L$ candidates (default max of 2-way)
  \begin{itemize}
  \item $L(L-1)/2 + L$ possible combinations from $L$ subgroup indicators (max = 2)
  \item Filter: $n \geq 60$, events $d_0 \geq 12$, and $d_1 \geq 12$
  \item Screen: Cox HR $\geq 1.25$ (experimental vs.\ control) ``signal of harm''
  \end{itemize}

\item \textbf{Splitting Consistency} --- Validate each candidate ($R = 1,000$ splits)
  \begin{itemize}
  \item Randomly split subgroup 50/50; require \emph{both halves} HR $\geq 1.0$ ``consistent with harm''
  \item $\hat{p}_{\text{cons}} = R^{-1}\sum_{r} I(\min(\hat{\beta}_s^{1r}, \hat{\beta}_s^{2r}) \geq 0) \geq 90\%$
  \item Select $\widehat{H}$: subgroup ``maximally consistent with harm''
  \end{itemize}
\end{enumerate}

## Three Subgroup Identification Objectives

\small

| Objective | Goal | Criterion | Strategy |
|:---------:|------|-----------|----------|
| **A** | Largest detrimental | $\hat{\beta} \geq \log(0.90)$ | Harm SG; complement = benefit |
| **B** | **Smallest** harmful | $\hat{\beta} \geq \log(0.90)$ | Min $\hat{\mathcal{H}}$; $\hat{\mathcal{H}}^{c} = \hat{\mathcal{G}}$ |
| **C** | **Largest** beneficial | $\hat{\beta} \leq \log(0.60)$ | Directly target strong benefit |

\normalsize

\medskip

- In **(A)** and **(B)**: identify harm $\hat{\mathcal{H}}$, then the complement $\hat{\mathcal{G}} = \hat{\mathcal{H}}^{c}$ is generally beneficial
- In **(C)**: directly target a clinically compelling beneficial subgroup effect



## Part 1: DGM Specification & Simulation Engine

```{=latex}
\centering
\resizebox{!}{0.85\textheight}{%
\begin{tikzpicture}[
  node distance=0.4cm and 0.7cm,
  font=\small,
  setup/.style={rectangle, rounded corners=4pt, draw=blue!60, fill=blue!8,
    text width=5.0cm, align=center, minimum height=0.65cm, inner sep=3pt},
  dgm/.style={rectangle, rounded corners=4pt, draw=orange!80, fill=orange!8,
    text width=4.6cm, align=center, minimum height=0.65cm, inner sep=3pt},
  scenario/.style={rectangle, rounded corners=4pt, draw=orange!80, fill=orange!15,
    text width=3.6cm, align=center, minimum height=0.6cm, inner sep=3pt},
  simbox/.style={rectangle, rounded corners=4pt, draw=green!60!black, fill=green!8,
    text width=5.0cm, align=center, minimum height=0.65cm, inner sep=3pt},
  arrow/.style={-{Stealth[length=5pt]}, thick, draw=gray!70},
]
\node[setup] (data) {\textbf{Case-Study Dataset}\\[-1pt]
  {\scriptsize Synthetic trial ($N \approx 2000$)}\\[-1pt]
  {\scriptsize Covariates, entry-times, treatment}};
\node[dgm, below=of data] (dgm) {\textbf{generate\_aft\_dgm\_flex()}\\[-1pt]
  {\scriptsize Weibull AFT + spline biomarker effects}\\[-1pt]
  {\scriptsize Censoring model selection (AIC/BIC)}};
\node[scenario, below left=0.45cm and 0.3cm of dgm] (alt) {\textbf{HTE Scenario}\\[-1pt]
  {\scriptsize $\log(3,\; 1.25,\; 0.50)$}\\[-1pt]
  {\scriptsize \textit{Strong biomarker effect}}};
\node[scenario, below right=0.45cm and 0.3cm of dgm] (null) {\textbf{Null Scenario}\\[-1pt]
  {\scriptsize $\log(0.7,\; 0.7,\; 0.7)$}\\[-1pt]
  {\scriptsize \textit{Uniform benefit}}};
\node[dgm, below=1.35cm of dgm] (dgmobj) {\textbf{DGM Object}\\[-1pt]
  {\scriptsize Model params ($\beta^{0}, \tau, \theta$), super-population, censoring}};
\node[simbox, below=of dgmobj] (engine) {\textbf{mrct\_region\_sims()}\\[-1pt]
  {\scriptsize $N_{\text{sims}} = 1{,}000 - 10{,}000$ replicates}};
\node[simbox, below=of engine] (simdata) {\textbf{simulate\_from\_dgm()}\\[-1pt]
  {\scriptsize Draw $n = 500$; Survival times}\\[-1pt]
  {\scriptsize Censoring + admin (analysis\_time $= 60$)}};
\draw[arrow] (data) -- (dgm);
\draw[arrow] (dgm) -- (alt);
\draw[arrow] (dgm) -- (null);
\draw[arrow] (alt) -- (dgmobj);
\draw[arrow] (null) -- (dgmobj);
\draw[arrow] (dgmobj) -- (engine);
\draw[arrow] (engine) -- (simdata);
\end{tikzpicture}%
}
```



## Part 2: Per-Replicate Analysis

```{=latex}
\centering
\resizebox{!}{0.85\textheight}{%
\begin{tikzpicture}[
  node distance=0.25cm and 0.6cm,
  font=\scriptsize,
  simbox/.style={rectangle, rounded corners=3pt, draw=green!60!black, fill=green!8,
    text width=3.8cm, align=center, minimum height=0.5cm, inner sep=2pt},
  region/.style={rectangle, rounded corners=3pt, draw=green!60!black, fill=green!12,
    text width=2.4cm, align=center, minimum height=0.4cm, inner sep=2pt},
  fsbox/.style={rectangle, rounded corners=3pt, draw=red!60, fill=red!6,
    text width=2.8cm, align=center, minimum height=0.4cm, inner sep=2pt},
  fsstep/.style={rectangle, rounded corners=2pt, draw=red!40, fill=red!4,
    text width=2.6cm, align=center, minimum height=0.35cm, inner sep=2pt},
  decision/.style={diamond, draw=purple!70, fill=purple!8,
    text width=1.0cm, align=center, inner sep=1pt, aspect=2.0},
  record/.style={rectangle, rounded corners=3pt, draw=green!60!black, fill=green!8,
    text width=4.8cm, align=center, minimum height=0.5cm, inner sep=2pt},
  arrow/.style={-{Stealth[length=4pt]}, thick, draw=gray!70},
]
\node[simbox] (simdata) {\textbf{simulate\_from\_dgm()}\\[-1pt]
  Simulated trial ($n = 500$)};
\node[region, below left=0.3cm and 0.3cm of simdata] (nonap) {\textbf{Non-AP} (Training)\\[-1pt]
  ${\sim}80$--$85\%$};
\node[region, below right=0.3cm and 0.3cm of simdata] (ap) {\textbf{AP} (Testing)\\[-1pt]
  ${\sim}15$--$20\%$};
\node[fsbox, below=0.35cm of nonap] (fs) {\textbf{forestsearch()}\\[-1pt]
  on Non-AP data};
\node[fsstep, below=of fs] (grf) {GRF cut selection};
\node[fsstep, below=of grf] (lasso) {Cox-LASSO selection};
\node[fsstep, below=of lasso] (enum) {Enumerate all-possible\\[-1pt]
  subgroup candidates};
\node[fsstep, below=of enum] (screen) {Screen: HR $\geq 0.90$\\[-1pt]
  Consistency $\geq 80\%$};
\node[fsstep, below=of screen] (select) {Select smallest subgroup\\[-1pt]
  (\texttt{sg\_focus="minSG"})};
\node[decision, below=0.3cm of select] (found) {$\hat{\mathcal{G}}$\\found?};
\node[fsbox, right=0.8cm of found] (apply) {Apply $\hat{\mathcal{G}}$ to AP\\[-1pt]
  HR(AP, $\hat{\mathcal{G}}$)};
\node[fsstep, left=0.8cm of found] (nofound) {any\_found $= 0$};
\node[record, below=0.55cm of found] (record) {\textbf{Record per-replicate results}\\[-1pt]
  any\_found, sg definition,
  HR(non-AP), HR(AP), HR(ITT), HR(AP,\,$\hat{\mathcal{G}}$)};
\draw[arrow] (simdata) -- (nonap);
\draw[arrow] (simdata) -- (ap);
\draw[arrow] (nonap) -- (fs);
\draw[arrow] (fs) -- (grf);
\draw[arrow] (grf) -- (lasso);
\draw[arrow] (lasso) -- (enum);
\draw[arrow] (enum) -- (screen);
\draw[arrow] (screen) -- (select);
\draw[arrow] (select) -- (found);
\draw[arrow] (found) -- node[above] {Yes} (apply);
\draw[arrow] (found) -- node[above] {No} (nofound);
\draw[arrow] (apply) |- (record);
\draw[arrow] (nofound) |- (record);
\draw[arrow, dashed, gray] (ap) |- (apply);
\end{tikzpicture}%
}
```




# Case Study and Simulations

## Super population 

\small

```{r superpop, echo=FALSE, fig.align ='center'}

create_summary_table(
  data = dgm_spline$df_super,
  treat_var = "treat",
  table_title = "Super-population Characteristics by Treatment Arm",
  vars_continuous = c("z_age", "z_bm", "entrytime", "lin_pred_1", "lin_pred_0"),
  vars_categorical = c("z_male", "z_histology", "z_prior_treat", "z_AP", "ecog", "strat"),
  font_size = 12
) |>
  gt::tab_options(table.align = "center", table.width = gt::pct(80))


```


```{r simulate-from-dgm, echo=FALSE}
df_example <- simulate_from_dgm(
  dgm = dgm_spline, n = 500,
  rand_ratio = 1, draw_treatment = FALSE,
  entry_var = "entrytime",
  analysis_time = 60, cens_adjust = 0,
  seed = 123
)

df_nonAP <- subset(df_example, z_AP == 0)
df_AP    <- subset(df_example, z_AP == 1)

```

\normalsize


## Kaplan-Meier: ITT Ideal (No Censoring) and Censoring

```{r km-itt-ideal, fig.height=6, fig.width=11}
dfc <- df_counting(df = df_example, by.risk = 6,
  tte.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim")

df_example$t_true <- pmin(df_example$t_true, 100)
df_example$event_mod <- ifelse(df_example$t_true <= 100, 1, 0)
dfideal <- df_counting(df = df_example, by.risk = 6,
  tte.name = "t_true", event.name = "event_mod",
  treat.name = "treat_sim")

par(mfrow = c(1, 2))
plot_weighted_km(dfideal, conf.int = FALSE,
  show.logrank = TRUE, ymax = 1.1,
  xmed.fraction = 0.7, ymed.offset = 0.15)
title("ITT no censoring (through 100 months)")

plot_weighted_km(dfc, conf.int = FALSE,
  show.logrank = TRUE, ymax = 1.1,
  xmed.fraction = 0.7, ymed.offset = 0.15)
title("ITT censored")
```


## Kaplan-Meier: Non-AP and AP Populations

```{r km-regional, fig.height=6, fig.width=12}
dfc0 <- df_counting(df = df_nonAP, by.risk = 6,
  tte.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim")
dfc1 <- df_counting(df = df_AP, by.risk = 6,
  tte.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim")

par(mfrow = c(1, 2))
plot_weighted_km(dfc0, conf.int = FALSE,
  show.logrank = TRUE, ymax = 1.1,
  xmed.fraction = 0.65, ymed.offset = 0.15)
title("Non-AP (Training)")

plot_weighted_km(dfc1, conf.int = FALSE,
  show.logrank = TRUE, ymax = 1.1,
  xmed.fraction = 0.65, ymed.offset = 0.15)
title("AP (Testing)")
```


## Spline Fit to Observed Data {.fragile}

\begin{columns}[T]
\begin{column}{0.48\textwidth}
\begin{lstlisting}
cox_cs_fit(
df = df_example,
tte_name = "y_sim", event_name = "event_sim",
treat_name = "treat_sim",
truebeta_name = "loghr_po",
z_name = "z_bm", z_by = 1, z_window = 0.5,
alpha = 0.20,
verbose = FALSE,
plot_params = list(
xlab = "Biomarker Level",
main_title = "Biomarker Heterogeneity (n=500)")
)
\end{lstlisting}

\end{column}

\begin{column}{0.52\textwidth}

```{r spline-fit, fig.width = 6, fig.height = 6}
result <- cox_cs_fit(
  df = df_example,
  tte_name = "y_sim", event_name = "event_sim",
  treat_name = "treat_sim",
  truebeta_name = "loghr_po",
  z_name = "z_bm", z_by = 1, z_window = 0.5,
  alpha = 0.20,
  verbose = FALSE,
  plot_params = list(
    xlab = "Biomarker Level",
    main_title = "Biomarker Heterogeneity (n=500)")
)
```


\end{column}
\end{columns}




# ForestSearch: Subgroup Identification


```{r forestsearch-alt, echo=FALSE, eval=TRUE}
confounders.name <- c("z_age", "z_bm", "z_male",
  "ecog", "z_histology", "z_prior_treat", "strat")

dfa <- as.data.frame(df_nonAP)

fs_output <- capture.output(
  fs <- forestsearch(
    dfa,
    confounders.name = confounders.name,
    outcome.name = "y_sim",
    treat.name = "treat_sim",
    event.name = "event_sim", id.name = "id",
    potentialOutcome.name = "loghr_po",
    df.test = as.data.frame(df_AP),
    hr.threshold = 0.9, hr.consistency = 0.80,
    stop_threshold = NULL,
    details = TRUE,
    pconsistency.threshold = 0.90,
    sg_focus = "minSG", max_subgroups_search = 30,
    showten_subgroups = TRUE,
    conf_force = c("z_age <= 65", "z_bm < 1",
      "z_bm < 2", "z_bm < 3", "z_bm < 4", "z_bm < 5", "z_bm < 6", "z_bm < 7", "z_bm < 8", "z_bm < 9", "z_bm < 10"),
    use_grf = TRUE, use_lasso = TRUE,
    maxk = 1, n.min = 60, d0.min = 10,
    d1.min = 10, plot.sg = FALSE, by.risk = 6,
    parallel_args = list(plan = "multisession",
      workers = 100, show_message = TRUE)
  )
)

plan("sequential")

```


## ForestSearch on Non-AP Data 


```{r, results='asis', echo=FALSE, eval = TRUE}
# Split after filtering summary instead
format_fs_details(
  fs_output,
  split_after = "Found.*candidate",
  fontsize = "tiny",
  col_widths = c(0.55, 0.45)
)
```



\normalsize

## Subgroups Identified: Non-AP Training {.shrink}

```{r fs-sg10, results="asis"}
res_tabs <- sg_tables(fs, ndecimals = 2,
  which_df = "est",
  est_title = "Treatment Effect Estimates",
  est_caption = "Non-AP (training)")
res_tabs$sg10_out
```



## Training and Testing Estimates 

```{r fs-estimates,fig.align = 'center'}

res_tabs$tab_estimates %>%
  gt::tab_options(
    table.align = "center",
    table.font.size = gt::px(9),
    table.width     = gt::pct(100)
  )

res_tabs_ap <- sg_tables(fs, ndecimals = 2,
  which_df = "testing",
  est_title = NULL,
  est_caption = "AP (testing)")

res_tabs_ap$tab_estimates %>%
  gt::tab_options(
    table.align = "center",
    table.font.size = gt::px(9),
    table.width     = gt::pct(100)
  )


```



## AP (ITT) vs Biomarker High (per identified $\geq 2$)

```{r km-ap-bm, fig.height=4.5, fig.width=9}

# Instead of low, all AP
df_AP_lo <- df_AP

df_AP_hi <- subset(df_AP, z_bm >= 2)

if (nrow(df_AP_lo) > 10 & nrow(df_AP_hi) > 10) {
  dfc_lo <- df_counting(df = df_AP_lo, by.risk = 6,
    tte.name = "y_sim", event.name = "event_sim",
    treat.name = "treat_sim")
  dfc_hi <- df_counting(df = df_AP_hi, by.risk = 6,
    tte.name = "y_sim", event.name = "event_sim",
    treat.name = "treat_sim")

  par(mfrow = c(1, 2))
  plot_weighted_km(dfc_lo, conf.int = FALSE,
    show.logrank = TRUE, ymax = 1.1,
    xmed.fraction = 0.725, ymed.offset = 0.15)
  title("AP Population")
  plot_weighted_km(dfc_hi, conf.int = FALSE,
    show.logrank = TRUE, ymax = 1.1,
    xmed.fraction = 0.725, ymed.offset = 0.15)
  title("AP Population: BM >= 2")
}
```


# Null Hypothesis Evaluation (Uniform HR = 0.7)


## Null: Regional Kaplan-Meier Plots

```{r km-null, fig.height=4.5, fig.width=9}

dgm_null <- generate_aft_dgm_flex(
  data = df.case,
  continuous_vars = c("age", "bm"),
  factor_vars = c("male", "histology",
                   "prior_treat", "AP"),
  set_beta_spec = list(set_var = c("z_AP"),
                       beta_var = -log(5)),
  continuous_vars_cens = c("age"),
  factor_vars_cens = c("prior_treat"),
  outcome_var = "tte", event_var = "event",
  treatment_var = "treat",
  cens_type = "weibull",
  subgroup_vars = NULL, subgroup_cuts = NULL,
  model = "alt",
  spline_spec = list(var = "z_bm", knot = 5,
    zeta = 10, log_hrs = log(c(0.7, 0.7, 0.7))),
  k_inter = 0.0, verbose = FALSE
)


df_null <- simulate_from_dgm(
  dgm = dgm_null, n = 500, rand_ratio = 1,
  draw_treatment = FALSE, entry_var = "entrytime",
  analysis_time = 60, cens_adjust = 0, seed = 12345
)

df_null_nonAP <- subset(df_null, z_AP == 0)
df_null_AP    <- subset(df_null, z_AP == 1)

dfc_null0 <- df_counting(df = df_null_nonAP, by.risk = 6,
  tte.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim")
dfc_null1 <- df_counting(df = df_null_AP, by.risk = 6,
  tte.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim")

par(mfrow = c(1, 2))
plot_weighted_km(dfc_null0, conf.int = FALSE,
  show.logrank = TRUE, ymax = 1.1,
  xmed.fraction = 0.6, ymed.offset = 0.15)
title("Null: non-AP (Training)")
plot_weighted_km(dfc_null1, conf.int = FALSE,
  show.logrank = TRUE, ymax = 1.1,
  xmed.fraction = 0.6, ymed.offset = 0.15)
title("Null: AP (Testing)")
```


## ForestSearch Under Null 

```{r fs-null, eval=TRUE}
dfa_null <- as.data.frame(df_null_nonAP)

fs_null_output <- capture.output(
  fs <- forestsearch(
    dfa_null,
    confounders.name = confounders.name,
    outcome.name = "y_sim",
    treat.name = "treat_sim",
    event.name = "event_sim", id.name = "id",
    potentialOutcome.name = "loghr_po",
    hr.threshold = 0.9, hr.consistency = 0.80,
    stop_threshold = NULL,
    details = TRUE,
    pconsistency.threshold = 0.90,
    sg_focus = "minSG", max_subgroups_search = 30,
    showten_subgroups = TRUE,
    conf_force = c("z_age <= 65", "z_bm < 1",
      "z_bm < 2", "z_bm < 3", "z_bm < 4", "z_bm < 5", "z_bm < 6", "z_bm < 7", "z_bm < 8", "z_bm < 9", "z_bm < 10"),
    use_grf = TRUE, use_lasso = TRUE,
    maxk = 1, n.min = 60, d0.min = 10,
    d1.min = 10, plot.sg = FALSE, by.risk = 6,
    parallel_args = list(plan = "multisession",
      workers = 100, show_message = TRUE)
  )
)

plan("sequential")


```

```{r, results='asis', echo=FALSE, eval = TRUE}
# Split after filtering summary instead
format_fs_details(
  fs_null_output,
  split_after = "Found.*candidate",
  fontsize = "tiny",
  col_widths = c(0.55, 0.45)
)
```

\normalsize


# Simulation Studies: Operating Characteristics


## HTE Scenario: Simulation Setup {.fragile}

\begin{columns}[T]
\begin{column}{0.48\textwidth}
\begin{lstlisting}
options(future.globals.onReference = "ignore")
results_alt <- mrct_region_sims(
  dgm       = dgm_spline,
  n_sims    = 5000,
  verbose_n_sims = 1L,
  n_sample  = 500,
  region_var = "z_AP",

  sg_focus  = "minSG", maxk = 1,
  hr.threshold  = 0.90,
  hr.consistency = 0.80,
  pconsistency.threshold = 0.80,

  confounders.name = c("z_age", "z_bm",
    "z_male", "ecog", "z_histology",
    "z_prior_treat", "strat"),
  conf_force = c("z_age <= 65",
    "z_bm <= 0", "z_bm <= 1",
    "z_bm <= 2", "z_bm <= 5"),
\end{lstlisting}
\end{column}
\begin{column}{0.48\textwidth}
\begin{lstlisting}
  fs_args = list(
    use_grf = TRUE, use_lasso = TRUE,
    cut_type = "default", n.min = 60,
    stop_threshold = 0.95,
    d0.min = 12, d1.min = 12,
    max_subgroups_search = 30,
    use_twostage = TRUE,
    twostage_args = list(
      n.splits.screen = 50,
      batch.size = 25,
      conf.level = 0.99)),

  sim_args = list(rand_ratio = 1,
                  draw_treatment = FALSE),
  analysis_time = 60, cens_adjust = 0,
  parallel_args = list(
    plan = "multisession",
    workers = 10, show_message = TRUE),
  details = TRUE, seed = 20240101
)
\end{lstlisting}
\end{column}
\end{columns}

\normalsize


## HTE Scenario: Summary Results {.shrink}

```{r alt-summarize, echo = FALSE, fig.align = 'center'}
results_alt <- readRDS("results/results_alt_k=1.rds")
results_null <- readRDS("results/results_null_k=1.rds")

temp <- summaryout_mrct(
  pop_summary = NULL,
  mrct_sims   = results_alt,
  mrct_sims_null = results_null,
  scenario_labels = c("HTE","Uniform"),
  showtable   = FALSE,
  table_width = 600,
  font_size = 12
) 


temp$out_table %>%
  gt::tab_options(
    table.align = "center",
    table.font.size = gt::px(9),
    table.width     = gt::pct(100)
  )


```


## HTE Scenario: HR Distributions

```{r alt-hrplot, fig.height = 4, fig.width=8}
plot_out <- SGplot_estimates(
  df             = results_alt,
  label_training = "Non-AP, ITT",
  label_itt      = "Overall, ITT",
  label_testing  = "AP, ITT",
  label_sg       = "AP, identified subgroup"
)
print(plot_out$plot_estimates)
```


## HTE Scenario: Subgroup Distribution

```{r alt-sgdist, fig.height=4, fig.width=8}
plot_sg_distribution(results_alt, min_pct = 0.5,
  title = "HTE Scenario: Subgroups Found",
  wrap_width = 30)
```




## Null Scenario: HR Distributions

```{r null-hrplot, fig.height=3.5, fig.width=8}
plot_out_null <- SGplot_estimates(
  df             = results_null,
  label_training = "Non-AP, ITT",
  label_itt      = "Overall, ITT",
  label_testing  = "AP, ITT",
  label_sg       = "AP, identified subgroup",
  trim_fraction = 0.01
)
print(plot_out_null$plot_estimates)
```


## Null Scenario: Subgroup Distribution

```{r null-sgdist, fig.height=4.5, fig.width=8}
plot_sg_distribution(results_null, min_pct = 0.5,
  title = "Uniform Benefit Scenario: Subgroups Found",
  wrap_width = 30)
```




# Key Takeaways


## Summary

1. **ForestSearch** provides a principled approach to subgroup identification that directly addresses MRCT regional consistency challenges

2. The **Weibull AFT/Cox dual framework** enables causal interpretation --- simulated on AFT scale, interpreted via hazard ratios

3. Three complementary estimands (**AHR**, **CDE**, **Marginal HR**) robustly characterize treatment effect heterogeneity

4. The **train/test paradigm** (non-AP $\to$ AP) avoids overfitting

5. **Data-driven simulation** evaluates full operating characteristics --- false discovery, estimation bias, CI coverage --- grounded in real trial data

6. **Simulation results** demonstrate the approach can recover AP consistency under heterogeneous effects while controlling false-discovery under the null


## {.plain}

\begin{center}
\vspace{1.5cm}

{\Large \textbf{Thank You}}

\vspace{0.8cm}

\textbf{ForestSearch R Package}

\medskip

Leon, L.F., Jemielita, T., Guo, Z., Marceau West, R. \& Anderson, K.M. (2024)

\emph{Exploratory subgroup identification in the heterogeneous Cox model.}

Statistics in Medicine. doi:10.1002/sim.10163

\vspace{0.8cm}

\texttt{remotes::install\_github("larry-leon/forestsearch")}

\smallskip
\url{https://larry-leon.github.io/forestsearch/}

\vspace{0.5cm}

\small
R code throughout this deck is directly executable with the ForestSearch
package and the synthetic case-study dataset.

\end{center}

# Additional Notes

## Some More Details {.shrink}

\begin{columns}[T]
\begin{column}{0.48\textwidth}

\textbf{Dual Causal Status}

The Weibull AFT scale-change parameter $\exp(-\beta)$ equals the \textbf{geometric mean ratio} of potential survival times:
$$\exp\!\bigl(E[\log T_{i}(1) - \log T_{i}(0)]\bigr)$$
Simultaneously \textbf{individual-level} and \textbf{population-level} causal.

\medskip

\textbf{AHR Inherits This Status}:
$$\text{AHR}(\mathcal{S}) = \exp\!\left(\frac{1}{|\mathcal{S}|}\sum_{i \in \mathcal{S}} [\theta_{i}(1) - \theta_{i}(0)]\right)$$

\end{column}
\begin{column}{0.48\textwidth}

\textbf{Cox HR as Operational Tool}

\begin{itemize}
\item \emph{True} effects: defined through $\psi^{0}(z)$
\item Subgroup \emph{selection}: Cox HR as efficient screening
\item Subgroup \emph{evaluation}: AHR and CDE
\item \emph{Correction}: bootstrap bias correction (\texttt{hr.H.bc})
\end{itemize}

\medskip

\textbf{De-Biased Estimation}:
$$\hat{\beta}^{*}(\hat{\mathcal{H}}) = \hat{\beta}(\hat{\mathcal{H}}) - \frac{1}{B}\sum_{b=1}^{B}\bigl[\eta^{*}_{b}(\hat{\mathcal{H}}^{*}_{b}) + \eta^{*}_{b}(\hat{\mathcal{H}})\bigr]$$

Infinitesimal jackknife for valid CIs.

\end{column}
\end{columns}


## ForestSearch Implementation Details {.shrink}

\begin{columns}[T]
\begin{column}{0.48\textwidth}

\textbf{Feature Discovery Pipeline}

\begin{enumerate}
\item \textbf{GRF} (causal survival forests): treatment effect heterogeneity via RMST
\item \textbf{LASSO} regularization: variable screening via Cox penalized regression
\item \textbf{Exhaustive combinatorial search}: across factors and cutpoints
\item \textbf{Forced cutpoints}: \texttt{conf\_force} for domain-guided biomarker cuts
\end{enumerate}

\medskip

\textbf{Selection Criteria}

\begin{itemize}
\item HR threshold: $\hat{\beta} \geq \log(0.90)$ 
\item Consistency rate $\geq$ 80--90\% 
\item Minimum sizes: $n \geq 60$, $d \geq 12$ 
\item \texttt{sg\_focus}: \texttt{"minSG"} $|$ \texttt{"hr"} $|$ \texttt{"maxSG"}
\end{itemize}

\end{column}
\begin{column}{0.48\textwidth}

\textbf{Cross-Validation}
\begin{itemize}
\item \texttt{leave-one-out}
\item \texttt{10-fold}
\item \texttt{agreement metrics: cv-testing vs full} sens, ppv, subgroup match (cv-training vs full) metrics
\end{itemize}

\medskip
\textbf{Bootstrap Inference}
\begin{itemize}
\item parallelised (dofuture) bootstrap
\item Full algorithm mimicked in each replicate ($B = 300$--$2000$)
\item IJ variance estimation for valid confidence intervals
\end{itemize}

\end{column}
\end{columns}

