---
title: "Synthetic CV Benchmark for ForestSearch"
subtitle: "Known-truth validation of subgroup identification and CV reproducibility"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    code-tools: true
    code-fold: true
    mathjax: true
    theme: cosmo
    fontsize: 11pt
    number-sections: true
---

# Purpose {#sec-purpose}

This document validates the ForestSearch subgroup-identification and
cross-validation machinery against a synthetic dataset where the
*ground truth is known* and the detection task is made deliberately
easy.  It is complementary to `gbsg_cv_qc.qmd`: that document asks
"does the current implementation reproduce the León et al. (2024)
paper on GBSG?", whereas this one asks "does the current
implementation identify the *correct* harm subgroup when all
difficulty is stripped out of the problem?".

The data-generating mechanism (DGM) is a Weibull proportional-hazards
model with **three binary covariates only**: one true harm driver
(`z1`) with a strong treatment-by-subgroup interaction, plus two
prognostic-only factors (`z2`, `z3`) that shift the baseline hazard
but do *not* interact with treatment.  There are **no noise
covariates, no continuous variables, and no cut grid** --- the
algorithm has only six candidate subgroups to evaluate (three
singletons + three pairs under `maxk = 2`), and `{z1}` is among
them.  See the roxygen header of `make_benchmark_dataset.R` for
full specification.

Because this is a deliberately easy test, a correctly-functioning
implementation should score at the top of every metric in the
**Pass/Fail thresholds** tables below *without* any pre-screening
(LASSO, GRF) turned on.  Performance substantially below these
thresholds points to a real issue in the core algorithm or CV
wrapper, not a dataset-difficulty issue.

```{mermaid}
flowchart LR
  A[make_benchmark_dataset] --> B[Full-data forestsearch]
  B --> C{sg.harm == '{z1}'?}
  C -->|YES| D[10-fold CV × Ksims]
  C -->|NO| X[FAIL: benchmark unusable]
  D --> E{sens_H ≥ 0.90?}
  E -->|YES| F[LOO CV]
  E -->|NO| Y[FAIL: CV unstable]
  F --> G[Compare FA_0 HR vs KfA_0 HR]
```


# Setup {#sec-setup}

```{r load-packages}
#| message: false
#| warning: false
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(future)
library(foreach)
library(doFuture)
```

```{r source-dgm}
# Source the DGM generator.  Must be alongside this .qmd.
source("make_benchmark_dataset.R")
```

```{r benchmark-config}
# Dataset
N                <- 700L       # sample size (comparable to GBSG's 686)
prev_z1          <- 0.20       # true harm prevalence
hr_harm          <- 2.20       # HR in harm subgroup
hr_benefit       <- 0.65       # HR in benefit group
dgm_seed         <- 20250421L

# Cross-validation
Ksims_cv         <- 50L        # repetitions for 10-fold CV
Kfolds_tenfold   <- 10L
run_loo_cv       <- TRUE       # set FALSE to skip LOO (≈ 5–10 min)

# ForestSearch parameters -- NO pre-screening, matching the "easy"
# benchmark design.  With binary-only candidates and no noise, the
# algorithm should identify {z1} without any help.
fs_sg_focus            <- "maxSG"
fs_hr_threshold        <- 1.25
fs_hr_consistency      <- 1.00
fs_pconsistency_thresh <- 0.90
fs_stop_threshold      <- 0.95
fs_splits              <- 400L
fs_n_min               <- 30L
fs_d0_min              <- 5L
fs_d1_min              <- 5L
fs_maxk                <- 2L
fs_max_subgroups       <- 10L

# Parallel execution
n_workers  <- max(1L, parallel::detectCores() - 1L)
```


# Data Generation {#sec-data}

```{r generate-data}
df <- make_benchmark_dataset(
  n          = N,
  prev_z1    = prev_z1,
  hr_harm    = hr_harm,
  hr_benefit = hr_benefit,
  seed       = dgm_seed
)

stopifnot(
  nrow(df) == N,
  all(c("time", "status", "treat", "z1") %in% names(df))
)
```

```{r dgm-summary-table}
.summary <- summarize_benchmark_dataset(df)

.summary_df <- data.frame(
  Statistic = c(
    "Sample size (n)",
    "Harm prevalence (realized)",
    "Overall event rate",
    "Events in harm subgroup, treated",
    "Events in harm subgroup, control",
    "Events in benefit group, treated",
    "Events in benefit group, control",
    "Empirical HR, overall",
    "Empirical HR, harm subgroup (expect ≈ 2.00)",
    "Empirical HR, benefit group (expect ≈ 0.70)"
  ),
  Value = c(
    format(.summary$n),
    sprintf("%.3f (nominal %.2f)", .summary$harm_prevalence, prev_z1),
    sprintf("%.1f%%", 100 * .summary$event_rate),
    format(.summary$events_by_arm["1", "1"]),
    format(.summary$events_by_arm["0", "1"]),
    format(.summary$events_by_arm["1", "0"]),
    format(.summary$events_by_arm["0", "0"]),
    sprintf("%.3f", .summary$hr_overall),
    sprintf("%.3f", .summary$hr_harm),
    sprintf("%.3f", .summary$hr_benefit)
  ),
  stringsAsFactors = FALSE
)

gt::gt(.summary_df) |>
  gt::tab_header(
    title = "Synthetic dataset: realized summary",
    subtitle = "z1 = true harm driver; z2, z3 = prognostic-only (no treatment interaction)"
  ) |>
  gt::cols_label(Statistic = "Statistic", Value = "Value")
```


# Full-data ForestSearch {#sec-full-fs}

```{r fs-full}
#| message: false
candidate_names <- c("z1", "z2", "z3")

fs <- forestsearch(
  df.analysis            = df,
  confounders.name       = candidate_names,
  outcome.name           = "time",
  event.name             = "status",
  treat.name             = "treat",
  id.name                = "id",
  is.RCT                 = TRUE,
  use_lasso              = FALSE,
  use_grf                = FALSE,
  use_twostage           = FALSE,
  sg_focus               = fs_sg_focus,
  hr.threshold           = fs_hr_threshold,
  hr.consistency         = fs_hr_consistency,
  pconsistency.threshold = fs_pconsistency_thresh,
  stop_threshold         = fs_stop_threshold,
  fs.splits              = fs_splits,
  n.min                  = fs_n_min,
  d0.min                 = fs_d0_min,
  d1.min                 = fs_d1_min,
  maxk                   = fs_maxk,
  max_subgroups_search   = fs_max_subgroups,
  parallel_args          = list(plan         = "sequential",
                                workers      = 1L,
                                show_message = FALSE),
  quiet                  = TRUE,
  details                = FALSE
)
plan("sequential")

# The label we expect is "{z1}" (a single binary factor cut).
fs_sg <- paste(fs$sg.harm, collapse = " & ")
cat("Identified subgroup (sg.harm): ", fs_sg, "\n", sep = "")
```

```{r fs-benchmark-row-1}
.bench1 <- data.frame(
  Check       = "Full-data sg.harm identifies the true harm driver",
  Expected    = "{z1}",
  Observed    = if (is.null(fs$sg.harm)) "(none)" else fs_sg,
  Pass        = identical(fs_sg, "{z1}"),
  stringsAsFactors = FALSE
)
.bench1$Result <- ifelse(.bench1$Pass, "\U2713 PASS", "\U2717 FAIL")

gt::gt(.bench1[, c("Check", "Expected", "Observed", "Result")]) |>
  gt::tab_header(title = "Benchmark 1: Full-data identification") |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#ffcccc"),
    locations = gt::cells_body(rows = !.bench1$Pass)
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#ccffcc"),
    locations = gt::cells_body(rows = .bench1$Pass)
  )

if (!.bench1$Pass) {
  warning(
    "Full-data benchmark FAILED.  Downstream CV benchmarks will ",
    "interpret mismatched 'sg.harm' incorrectly.  Investigate ",
    "before continuing."
  )
}
```


# Ground-truth HR in the harm subgroup {#sec-truth-hr}

For comparison with the CV-based KfA_0 HR below.

```{r fa-truth-hr}
.fit_truth <- survival::coxph(
  survival::Surv(time, status) ~ treat,
  data = df[df$z1 == 1L, ], robust = FALSE
)
.sm <- summary(.fit_truth)$conf.int
.truth_hr  <- unname(.sm[1, 1])
.truth_lo  <- unname(.sm[1, 3])
.truth_hi  <- unname(.sm[1, 4])
.truth_n   <- sum(df$z1 == 1L)
.truth_d   <- sum(df$status[df$z1 == 1L])

sprintf(
  "True harm subgroup (z1 = 1): n = %d, events = %d, HR = %.2f (%.2f, %.2f)",
  .truth_n, .truth_d, .truth_hr, .truth_lo, .truth_hi
)
```


# 10-fold Cross-Validation Benchmark {#sec-cv-kfold}

```{r kfold-cv}
#| message: false
.t0 <- proc.time()
fs_kfold <- forestsearch_tenfold(
  fs.est        = fs,
  sims          = Ksims_cv,
  Kfolds        = Kfolds_tenfold,
  details       = FALSE,
  keep_resCV    = FALSE,         # fold_summary is enough for benchmarks
  parallel_args = list(plan = "multisession",
                       workers = n_workers,
                       show_message = FALSE)
)
plan("sequential")
.kfold_seconds <- as.numeric((proc.time() - .t0)["elapsed"])
sprintf("10-fold CV finished in %.1f s", .kfold_seconds)
```

## Aggregate agreement metrics

```{r kfold-aggregate}
.kfold_metrics <- data.frame(
  metric   = c("sens_H", "sens_Hc", "ppv_H", "ppv_Hc",
               "Any Found", "Exact Match"),
  value    = c(
    unname(fs_kfold$sens_summary["sens_H"]),
    unname(fs_kfold$sens_summary["sens_Hc"]),
    unname(fs_kfold$sens_summary["ppv_H"]),
    unname(fs_kfold$sens_summary["ppv_Hc"]),
    unname(fs_kfold$find_summary["Any"]),
    unname(fs_kfold$find_summary["Exact"])
  ),
  threshold = c(0.95, 0.95, 0.95, 0.95, 0.99, 0.95),
  stringsAsFactors = FALSE
)
.kfold_metrics$Pass   <- .kfold_metrics$value >= .kfold_metrics$threshold
.kfold_metrics$Result <- ifelse(.kfold_metrics$Pass,
                                "\U2713 PASS", "\U2717 FAIL")

gt::gt(.kfold_metrics) |>
  gt::tab_header(
    title    = sprintf("Benchmark 2: 10-fold CV (%d sims)", Ksims_cv),
    subtitle = "Pass thresholds per DGM design"
  ) |>
  gt::fmt_number(columns = c("value", "threshold"), decimals = 3) |>
  gt::cols_label(
    metric    = "Metric",
    value     = "Observed (median)",
    threshold = "Threshold",
    Pass      = "Meets?",
    Result    = "Result"
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#ffcccc"),
    locations = gt::cells_body(rows = !.kfold_metrics$Pass)
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#ccffcc"),
    locations = gt::cells_body(rows = .kfold_metrics$Pass)
  )
```

## Fold-level distribution (uses the new `fold_summary` slot)

Because the package update preserves `fold_summary` for every
`forestsearch_tenfold()` call, we can inspect the empirical
distribution of identified subgroups across all sim × fold pairs
directly --- no re-running required.

```{r kfold-fold-summary}
if (!is.null(fs_kfold$fold_summary)) {

  fsum <- fs_kfold$fold_summary

  # Compact display: subgroup -> number of folds that identified it
  .sg_table <- as.data.frame(
    table(
      subgroup = ifelse(
        is.na(fsum$sg1),
        "(no subgroup)",
        ifelse(
          is.na(fsum$sg2),
          fsum$sg1,
          paste(fsum$sg1, fsum$sg2, sep = " & ")
        )
      )
    )
  )
  names(.sg_table) <- c("Training-FS subgroup", "n folds")
  .sg_table <- .sg_table[
    order(-.sg_table[["n folds"]]), , drop = FALSE
  ]
  .sg_table[["% folds"]] <- round(
    100 * .sg_table[["n folds"]] / sum(.sg_table[["n folds"]]), 2
  )
  .sg_table[["Matches truth?"]] <- ifelse(
    .sg_table[["Training-FS subgroup"]] == "{z1}", "YES", ""
  )

  gt::gt(.sg_table) |>
    gt::tab_header(
      title = sprintf(
        "Identified subgroups across all %d sim × fold pairs",
        nrow(fsum)
      ),
      subtitle = "Ground truth: {z1}"
    )

} else {
  cat("fold_summary slot not available. ",
      "Install the v0.3.0 update to forestsearch_cross_validation.R.\n",
      sep = "")
}
```


# N-fold (LOO) Cross-Validation Benchmark {#sec-cv-oob}

```{r oob-cv}
#| message: false
if (isTRUE(run_loo_cv)) {
  .t0 <- proc.time()
  fs_oob <- forestsearch_Kfold(
    fs.est        = fs,
    Kfolds        = nrow(df),
    parallel_args = list(plan = "multisession",
                         workers = n_workers,
                         show_message = FALSE)
  )
  plan("sequential")
  .oob_seconds <- as.numeric((proc.time() - .t0)["elapsed"])

  fs_oob_out <- forestsearch_KfoldOut(fs_oob, outall = TRUE, details = FALSE)
  sprintf("LOO CV finished in %.1f s", .oob_seconds)
} else {
  cat("LOO CV skipped (run_loo_cv = FALSE).\n")
}
```

## Aggregate agreement metrics

```{r oob-aggregate}
#| eval: !expr run_loo_cv
.oob_metrics <- data.frame(
  metric    = c("sens_H", "sens_Hc", "ppv_H", "ppv_Hc",
                "Any Found", "Exact Match"),
  value     = c(
    unname(fs_oob_out$sens_metrics_original["sens_H"]),
    unname(fs_oob_out$sens_metrics_original["sens_Hc"]),
    unname(fs_oob_out$sens_metrics_original["ppv_H"]),
    unname(fs_oob_out$sens_metrics_original["ppv_Hc"]),
    unname(fs_oob_out$find_metrics["Any"]),
    unname(fs_oob_out$find_metrics["Exact"])
  ),
  threshold = c(0.95, 0.95, 0.95, 0.95, 0.99, 0.98),
  stringsAsFactors = FALSE
)
.oob_metrics$Pass   <- .oob_metrics$value >= .oob_metrics$threshold
.oob_metrics$Result <- ifelse(.oob_metrics$Pass,
                              "\U2713 PASS", "\U2717 FAIL")

gt::gt(.oob_metrics) |>
  gt::tab_header(
    title    = "Benchmark 3: LOO (N-fold) CV",
    subtitle = "Pass thresholds per DGM design"
  ) |>
  gt::fmt_number(columns = c("value", "threshold"), decimals = 3) |>
  gt::cols_label(
    metric    = "Metric",
    value     = "Observed",
    threshold = "Threshold",
    Pass      = "Meets?",
    Result    = "Result"
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#ffcccc"),
    locations = gt::cells_body(rows = !.oob_metrics$Pass)
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#ccffcc"),
    locations = gt::cells_body(rows = .oob_metrics$Pass)
  )
```

## FA_0 vs KfA_0 HR comparison {#sec-fa-vs-kfa}

Under the synthetic DGM with 100% (or nearly so) exact match, the
KfA_0 subgroup should contain the same subjects as FA_0, so the two
Cox HRs should agree within Monte-Carlo noise.  A large discrepancy
here reproduces the disturbing pattern observed in the GBSG LOO
analysis (Section 5.3.2 of `gbsg_cv_qc.qmd`) and would indicate the
CV procedure is destabilizing the harm estimate even on easy data.

```{r oob-fa-vs-kfa}
#| eval: !expr run_loo_cv
.tab <- fs_oob_out$tab_all
# tab_all rows: "Overall", "FA_0", "KfA_0", "FA_1", "KfA_1"

.fa_kfa <- data.frame(
  Analysis = c("FA_0 (full-data harm)",
               "KfA_0 (LOO-CV harm)"),
  n        = c(.tab["FA_0", "n"], .tab["KfA_0", "n"]),
  HR       = c(.tab["FA_0", "Hazard ratio"],
               .tab["KfA_0", "Hazard ratio"]),
  stringsAsFactors = FALSE
)

gt::gt(.fa_kfa) |>
  gt::tab_header(
    title    = "Full-analysis vs LOO-CV: harm-subgroup Cox HR",
    subtitle = "True HR in harm subgroup ≈ 2.00"
  ) |>
  gt::tab_source_note(paste(
    "Expected: both HRs in [1.5, 3.0] with similar CIs.",
    "Large directional disagreement (e.g., FA > 1.5 but KfA < 1.0)",
    "signals CV instability despite an easy DGM."
  ))
```


# Benchmark summary {#sec-summary}

```{r overall-summary}
.all_passes <- c(
  full_fs = .bench1$Pass,
  kfold   = all(.kfold_metrics$Pass)
)

if (isTRUE(run_loo_cv)) {
  .all_passes <- c(.all_passes, oob = all(.oob_metrics$Pass))
}

.total <- all(.all_passes)
.status <- ifelse(.total,
                  "\U2713 ALL BENCHMARKS PASSED",
                  "\U2717 ONE OR MORE BENCHMARKS FAILED")

.overall_df <- data.frame(
  Stage = c(
    "Full-data identification",
    "10-fold CV (all metrics)",
    if (isTRUE(run_loo_cv)) "LOO CV (all metrics)" else NULL
  ),
  Result = c(
    ifelse(.all_passes["full_fs"], "\U2713 PASS", "\U2717 FAIL"),
    ifelse(.all_passes["kfold"],   "\U2713 PASS", "\U2717 FAIL"),
    if (isTRUE(run_loo_cv))
      ifelse(.all_passes["oob"],   "\U2713 PASS", "\U2717 FAIL")
  ),
  stringsAsFactors = FALSE
)

gt::gt(.overall_df) |>
  gt::tab_header(title = .status) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#ccffcc"),
    locations = gt::cells_body(rows = grepl("PASS", .overall_df$Result))
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#ffcccc"),
    locations = gt::cells_body(rows = grepl("FAIL", .overall_df$Result))
  )
```


# Session info {#sec-session}

```{r session}
sessionInfo()
```
