---
title: "GBSG 10-fold CV: dmin.grf Sensitivity Analysis"
subtitle: "dmin.grf = {12, 6, 4} months, with/without conf_force, at frac.tau = 0.8"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    code-tools: true
    code-fold: true
    mathjax: true
    theme: cosmo
    fontsize: 11pt
    number-sections: true
---

# Purpose {#sec-purpose}

The prior 10-fold diagnostic (`gbsg_10fold_cv_diagnostic.qmd`, run at
`dmin.grf = 12` months, `frac.tau = 0.6`) revealed three facts:

1. **GRF returns no `er` cut on 41% of 10-fold training sets** (410 of
   1000 sim × fold pairs).  When GRF fails to surface `er <= 0`, the
   downstream pipeline typically finds no subgroup at all, or drifts
   to weak alternatives.
2. **On folds where GRF does surface `er <= 0` (58%)**, the pipeline
   recovers `{er <= 0}` exactly 80% of the time.
3. **Forcing `er <= 0` via `conf_force`** rescues ~82% of the
   no-GRF-`er`-cut folds, lifting `sens_H` from 0.524 to 0.841.

A separately rendered LOO diagnostic at `dmin.grf = 4` months and
`frac.tau = 0.8` (more relaxed GRF settings) showed near-perfect
recovery: 684/686 LOO folds identifying `{er <= 0}`, `sens_H = 1.000`,
no contribution from `conf_force`.  The strategic read: GRF's job is
candidate *discovery*; the FS consistency stage is the gatekeeper.
Let GRF cast a wide net.

Two questions addressed here:

**A.** When GRF at `dmin.grf = 12` "returned no `er` cut" on 41% of
10-fold training sets, did it produce cuts on *other* variables
(`pgr`, `grade3`, `age`), or did the policy tree fail to produce any
qualifying subgroup at all?  This is a straight tabulation of the
full `grf_cuts` column now preserved in `fold_summary`.

**B.** Does relaxing `dmin.grf` close the gap between full-data/LOO
and 10-fold CV?  We run three `dmin.grf` settings (12, 6, 4 months)
corresponding to (6, 3, 2)-month RMST difference thresholds, each
with and without `conf_force`, all at `frac.tau = 0.8`.

Six 10-fold CV variants side-by-side:

| Config        | `dmin.grf` | `conf_force`     | Approx. RMST threshold |
|:--------------|:-----------|:-----------------|:-----------------------|
| A: strict base   | 12         | NULL             | 6 months |
| B: strict forced | 12         | `c("er <= 0")`   | 6 months |
| C: mid base      | 6          | NULL             | 3 months |
| D: mid forced    | 6          | `c("er <= 0")`   | 3 months |
| E: relaxed base  | 4          | NULL             | 2 months |
| F: relaxed forced| 4          | `c("er <= 0")`   | 2 months |

Expected runtime: ~45–60 minutes for all six runs at 100 sims × 10
folds each.


# Setup {#sec-setup}

```{r load-packages}
#| message: false
#| warning: false
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(future)
library(foreach)
library(doFuture)
```

```{r config}
# Variable names (match gbsg_cv_qc.qmd and prior 10-fold diagnostic)
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name     <- "time_months"
event.name       <- "status"
treat.name       <- "hormon"
id.name          <- "id"

# Base args shared across all four runs.  dmin.grf and conf_force vary;
# everything else mirrors the prior diagnostic (which in turn matches
# gbsg_cv_qc.qmd paper-reproduction config).
fs_args_base <- list(
  # Threshold parameters (Leon et al. 2024)
  hr.threshold              = 1.25,
  hr.consistency            = 1.00,
  pconsistency.threshold    = 0.90,
  stop_threshold            = 0.95,
  # Search configuration
  sg_focus                  = "maxSG",
  max_subgroups_search      = 10L,
  use_twostage              = FALSE,
  # Factor selection / GRF
  use_grf                   = TRUE,
  return_selected_cuts_only = TRUE,
  use_lasso                 = TRUE,
  cut_type                  = "default",
  frac.tau                  = 0.8,        # relaxed RMST integration horizon
  # Subgroup constraints
  maxk                      = 2L,
  n.min                     = 60L,
  d0.min                    = 10L,
  d1.min                    = 10L,
  # Consistency evaluation
  fs.splits                 = 400L,
  # Output
  details                   = FALSE,
  quiet                     = TRUE,
  plot.sg                   = FALSE,
  plot.grf                  = FALSE
)

# 10-fold CV
Ksims_cv  <- 100L
Kfolds_cv <- 10L

# Parallel execution
n_workers <- max(1L, parallel::detectCores() - 1L)
```

```{r data-setup}
df <- survival::gbsg
df$grade3      <- as.integer(df$grade == "3")
df$id          <- seq_len(nrow(df))
df$time_months <- df$rfstime / 30.4375
```


# Full-data ForestSearch (four configs) {#sec-full}

Brief sanity checks; all six should identify `{er <= 0}` on the
full data.

```{r fs-full-all}
#| message: false
.run_fs <- function(dmin_val, conf_force_val) {
  args <- c(list(
    df.analysis      = df,
    confounders.name = confounders.name,
    outcome.name     = outcome.name,
    event.name       = event.name,
    treat.name       = treat.name,
    id.name          = id.name,
    dmin.grf         = dmin_val,
    conf_force       = conf_force_val,
    parallel_args    = list(plan         = "multisession",
                            workers      = n_workers,
                            show_message = FALSE)
  ), fs_args_base)
  do.call(forestsearch, args)
}

fs_A <- .run_fs(12, NULL);          plan("sequential")  # A: dmin=12 base
fs_B <- .run_fs(12, c("er <= 0"));  plan("sequential")  # B: dmin=12 forced
fs_C <- .run_fs(6,  NULL);          plan("sequential")  # C: dmin=6  base
fs_D <- .run_fs(6,  c("er <= 0"));  plan("sequential")  # D: dmin=6  forced
fs_E <- .run_fs(4,  NULL);          plan("sequential")  # E: dmin=4  base
fs_F <- .run_fs(4,  c("er <= 0"));  plan("sequential")  # F: dmin=4  forced

.config_labels <- c("A: dmin=12 base",   "B: dmin=12 forced",
                    "C: dmin=6  base",   "D: dmin=6  forced",
                    "E: dmin=4  base",   "F: dmin=4  forced")

.full_summary <- data.frame(
  Config   = .config_labels,
  sg.harm  = c(paste(fs_A$sg.harm, collapse = " & "),
               paste(fs_B$sg.harm, collapse = " & "),
               paste(fs_C$sg.harm, collapse = " & "),
               paste(fs_D$sg.harm, collapse = " & "),
               paste(fs_E$sg.harm, collapse = " & "),
               paste(fs_F$sg.harm, collapse = " & ")),
  grf_cuts = c(paste(fs_A$grf_cuts, collapse = " | "),
               paste(fs_B$grf_cuts, collapse = " | "),
               paste(fs_C$grf_cuts, collapse = " | "),
               paste(fs_D$grf_cuts, collapse = " | "),
               paste(fs_E$grf_cuts, collapse = " | "),
               paste(fs_F$grf_cuts, collapse = " | ")),
  stringsAsFactors = FALSE
)

gt::gt(.full_summary) |>
  gt::tab_header(title = "Full-data identification across six configs")
```


# 10-fold CV runs (six configs) {#sec-tenfold}

```{r run-all-tenfold}
#| message: false
.run_tenfold <- function(fs_est, label) {
  t0 <- proc.time()
  out <- forestsearch_tenfold(
    fs.est        = fs_est,
    sims          = Ksims_cv,
    Kfolds        = Kfolds_cv,
    details       = FALSE,
    keep_resCV    = FALSE,
    parallel_args = list(plan         = "multisession",
                         workers      = n_workers,
                         show_message = FALSE)
  )
  plan("sequential")
  secs <- as.numeric((proc.time() - t0)["elapsed"])
  message(sprintf("  %s: %.1f s", label, secs))
  attr(out, "runtime_seconds") <- secs
  out
}

message("Running six 10-fold CV configs:")
tf_A <- .run_tenfold(fs_A, "A: dmin=12 base   ")
tf_B <- .run_tenfold(fs_B, "B: dmin=12 forced ")
tf_C <- .run_tenfold(fs_C, "C: dmin=6  base   ")
tf_D <- .run_tenfold(fs_D, "D: dmin=6  forced ")
tf_E <- .run_tenfold(fs_E, "E: dmin=4  base   ")
tf_F <- .run_tenfold(fs_F, "F: dmin=4  forced ")

.runtime_summary <- data.frame(
  Config  = .config_labels,
  Seconds = c(attr(tf_A, "runtime_seconds"),
              attr(tf_B, "runtime_seconds"),
              attr(tf_C, "runtime_seconds"),
              attr(tf_D, "runtime_seconds"),
              attr(tf_E, "runtime_seconds"),
              attr(tf_F, "runtime_seconds")),
  stringsAsFactors = FALSE
)

gt::gt(.runtime_summary) |>
  gt::tab_header(
    title = sprintf(
      "10-fold CV runtimes (%d sims × %d folds each)",
      Ksims_cv, Kfolds_cv)
  ) |>
  gt::fmt_number(columns = "Seconds", decimals = 1)
```


# Aggregate metrics {#sec-metrics}

```{r metrics-table}
.metric_names <- c("sens_H", "sens_Hc", "ppv_H", "ppv_Hc")
.find_names   <- c("Any", "Exact", "At least 1",
                   "Cov1", "Cov2", "Cov 1 & 2",
                   "Cov1 exact", "Cov2 exact")

.extract_metrics <- function(tf, label) {
  vals <- c(tf$sens_summary[.metric_names],
            tf$find_summary[.find_names])
  data.frame(
    metric = c(.metric_names, .find_names),
    value  = unname(vals),
    Config = label,
    stringsAsFactors = FALSE
  )
}

.all_metrics <- rbind(
  .extract_metrics(tf_A, "A: dmin=12 base"),
  .extract_metrics(tf_B, "B: dmin=12 forced"),
  .extract_metrics(tf_C, "C: dmin=6  base"),
  .extract_metrics(tf_D, "D: dmin=6  forced"),
  .extract_metrics(tf_E, "E: dmin=4  base"),
  .extract_metrics(tf_F, "F: dmin=4  forced")
)

# Pivot to wide (metric x config)
.metrics_wide <- reshape(
  .all_metrics,
  timevar   = "Config",
  idvar     = "metric",
  direction = "wide"
)
names(.metrics_wide) <- gsub("^value\\.", "", names(.metrics_wide))

gt::gt(.metrics_wide) |>
  gt::tab_header(
    title    = "10-fold CV agreement and finding metrics across six configs",
    subtitle = sprintf(
      "GBSG, N = %d, %d sims × %d folds = %d sim × fold pairs, frac.tau = 0.8",
      nrow(df), Ksims_cv, Kfolds_cv, Ksims_cv * Kfolds_cv)
  ) |>
  gt::fmt_number(columns = -metric, decimals = 3) |>
  gt::cols_label(metric = "Metric") |>
  gt::tab_source_note(paste(
    "Δ(base -> forced) at fixed dmin.grf measures the conf_force rescue.",
    "Scan along the base configs (A, C, E) to see whether relaxing",
    "dmin.grf closes the base-vs-forced gap without needing conf_force."
  ))
```


# Full `grf_cuts` distribution across 10-fold training sets {#sec-grf-full}

This section answers Question A from §1: at each `dmin.grf`
setting, what did GRF actually return across the 1000 sim × fold
training sets?  Specifically, when the `er`-cut projection shows
no entry, did GRF still produce cuts on other variables, or did
the policy tree return nothing at all?

## Full `grf_cuts` strings, all six configs

```{r grf-full-dist}
.clean_grf <- function(x) {
  if (is.na(x) || x == "") return("(no GRF cut)")
  x
}

.grf_dist <- function(tf, label, top_n = 15L) {
  fsum <- tf$fold_summary
  cuts <- vapply(fsum$grf_cuts, .clean_grf, character(1))
  tab <- as.data.frame(table(cuts = cuts), stringsAsFactors = FALSE)
  tab <- tab[order(-tab$Freq), , drop = FALSE]
  tab$Config <- label
  tab$Pct    <- round(100 * tab$Freq / nrow(fsum), 2)
  head(tab, top_n)
}

.grf_all <- rbind(
  .grf_dist(tf_A, "A: dmin=12 base"),
  .grf_dist(tf_B, "B: dmin=12 forced"),
  .grf_dist(tf_C, "C: dmin=6  base"),
  .grf_dist(tf_D, "D: dmin=6  forced"),
  .grf_dist(tf_E, "E: dmin=4  base"),
  .grf_dist(tf_F, "F: dmin=4  forced")
)
names(.grf_all)[names(.grf_all) == "cuts"] <- "GRF cuts"
names(.grf_all)[names(.grf_all) == "Freq"] <- "n folds"
names(.grf_all)[names(.grf_all) == "Pct"]  <- "% folds"

gt::gt(.grf_all,
       groupname_col = "Config",
       row_group_as_column = TRUE) |>
  gt::tab_header(
    title    = "Top GRF cut strings across 10-fold training sets",
    subtitle = "Full grf_cuts column; not projected to any single variable"
  )
```

## `er`-variable projection (for comparison with prior doc)

```{r grf-er-projection}
.extract_er_cut <- function(cuts_str) {
  if (is.na(cuts_str)) return("(no GRF cut)")
  parts <- strsplit(cuts_str, " \\| ", fixed = FALSE)[[1]]
  er_parts <- parts[grepl("^er\\b", parts)]
  if (length(er_parts) == 0L) return("(no er cut)")
  paste(er_parts, collapse = "; ")
}

.er_proj <- function(tf, label) {
  fsum <- tf$fold_summary
  er_cuts <- vapply(fsum$grf_cuts, .extract_er_cut, character(1))
  tab <- as.data.frame(table(er_cut = er_cuts), stringsAsFactors = FALSE)
  tab <- tab[order(-tab$Freq), , drop = FALSE]
  tab$Config <- label
  tab$Pct    <- round(100 * tab$Freq / nrow(fsum), 2)
  tab
}

.er_all <- rbind(
  .er_proj(tf_A, "A: dmin=12 base"),
  .er_proj(tf_B, "B: dmin=12 forced"),
  .er_proj(tf_C, "C: dmin=6  base"),
  .er_proj(tf_D, "D: dmin=6  forced"),
  .er_proj(tf_E, "E: dmin=4  base"),
  .er_proj(tf_F, "F: dmin=4  forced")
)
names(.er_all)[names(.er_all) == "er_cut"] <- "er-cut projection"
names(.er_all)[names(.er_all) == "Freq"]   <- "n folds"
names(.er_all)[names(.er_all) == "Pct"]    <- "% folds"

gt::gt(.er_all,
       groupname_col = "Config",
       row_group_as_column = TRUE) |>
  gt::tab_header(
    title    = "GRF `er`-cut projection (for continuity with prior doc)",
    subtitle = "'(no er cut)' means GRF split on other variable(s); '(no GRF cut)' means GRF returned nothing"
  )
```


# Cross-tab: full GRF cut × identified subgroup {#sec-xtab}

For each config, counts of (GRF `grf_cuts` string) × (identified
training-FS subgroup).  The informative rows are those where GRF
returned a cut that does **not** involve `er` and yet the pipeline
still identified `{er <= 0}` (only possible under a forced config)
or a non-`er` subgroup based on the GRF cut.

```{r xtab-helper}
.build_sg_full <- function(fsum) {
  ifelse(
    is.na(fsum$sg1),
    "(no subgroup)",
    ifelse(
      is.na(fsum$sg2),
      fsum$sg1,
      paste(fsum$sg1, fsum$sg2, sep = " & ")
    )
  )
}

.xtab <- function(tf, label) {
  fsum <- tf$fold_summary
  fsum$grf_clean <- vapply(fsum$grf_cuts, .clean_grf, character(1))
  fsum$sg_full   <- .build_sg_full(fsum)

  tab <- as.data.frame(
    table(`GRF cuts`  = fsum$grf_clean,
          `Identified subgroup` = fsum$sg_full),
    stringsAsFactors = FALSE
  )
  tab <- tab[tab$Freq > 0, , drop = FALSE]
  names(tab)[names(tab) == "Freq"] <- "n folds"
  tab$Config <- label
  tab <- tab[order(-tab[["n folds"]]), , drop = FALSE]
  tab
}
```

## Config A: `dmin.grf = 12`, baseline

```{r xtab-A}
.xA <- .xtab(tf_A, "A: dmin=12 base")
gt::gt(.xA[, c("GRF cuts", "Identified subgroup", "n folds")]) |>
  gt::tab_header(
    title    = "A: dmin.grf = 12, baseline",
    subtitle = sprintf("%d total sim × fold pairs",
                       sum(.xA[["n folds"]]))
  )
```

## Config B: `dmin.grf = 12`, forced

```{r xtab-B}
.xB <- .xtab(tf_B, "B: dmin=12 forced")
gt::gt(.xB[, c("GRF cuts", "Identified subgroup", "n folds")]) |>
  gt::tab_header(
    title    = "B: dmin.grf = 12, forced",
    subtitle = sprintf("%d total sim × fold pairs",
                       sum(.xB[["n folds"]]))
  )
```

## Config C: `dmin.grf = 6`, baseline

```{r xtab-C}
.xC <- .xtab(tf_C, "C: dmin=6  base")
gt::gt(.xC[, c("GRF cuts", "Identified subgroup", "n folds")]) |>
  gt::tab_header(
    title    = "C: dmin.grf = 6, baseline",
    subtitle = sprintf("%d total sim × fold pairs",
                       sum(.xC[["n folds"]]))
  )
```

## Config D: `dmin.grf = 6`, forced

```{r xtab-D}
.xD <- .xtab(tf_D, "D: dmin=6  forced")
gt::gt(.xD[, c("GRF cuts", "Identified subgroup", "n folds")]) |>
  gt::tab_header(
    title    = "D: dmin.grf = 6, forced",
    subtitle = sprintf("%d total sim × fold pairs",
                       sum(.xD[["n folds"]]))
  )
```

## Config E: `dmin.grf = 4`, baseline

```{r xtab-E}
.xE <- .xtab(tf_E, "E: dmin=4  base")
gt::gt(.xE[, c("GRF cuts", "Identified subgroup", "n folds")]) |>
  gt::tab_header(
    title    = "E: dmin.grf = 4, baseline",
    subtitle = sprintf("%d total sim × fold pairs",
                       sum(.xE[["n folds"]]))
  )
```

## Config F: `dmin.grf = 4`, forced

```{r xtab-F}
.xF <- .xtab(tf_F, "F: dmin=4  forced")
gt::gt(.xF[, c("GRF cuts", "Identified subgroup", "n folds")]) |>
  gt::tab_header(
    title    = "F: dmin.grf = 4, forced",
    subtitle = sprintf("%d total sim × fold pairs",
                       sum(.xF[["n folds"]]))
  )
```


# Headline summary {#sec-summary}

```{r headline}
.headline <- function(tf, label) {
  fsum <- tf$fold_summary
  grf  <- vapply(fsum$grf_cuts, .clean_grf, character(1))
  er   <- vapply(fsum$grf_cuts, .extract_er_cut, character(1))
  sg   <- .build_sg_full(fsum)

  data.frame(
    Config                               = label,
    `Folds with GRF producing any cut`  = sprintf(
      "%d (%.1f%%)",
      sum(grf != "(no GRF cut)"),
      100 * mean(grf != "(no GRF cut)")),
    `Folds with GRF cut containing er`  = sprintf(
      "%d (%.1f%%)",
      sum(grepl("^er\\b", fsum$grf_cuts) |
          grepl("\\| *er\\b", fsum$grf_cuts)),
      100 * mean(grepl("^er\\b", fsum$grf_cuts) |
                 grepl("\\| *er\\b", fsum$grf_cuts))),
    `Folds identifying {er <= 0} exactly` = sprintf(
      "%d (%.1f%%)",
      sum(sg == "{er <= 0}"),
      100 * mean(sg == "{er <= 0}")),
    `Folds with no subgroup identified`   = sprintf(
      "%d (%.1f%%)",
      sum(sg == "(no subgroup)"),
      100 * mean(sg == "(no subgroup)")),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}

.all_headline <- rbind(
  .headline(tf_A, "A: dmin=12 base"),
  .headline(tf_B, "B: dmin=12 forced"),
  .headline(tf_C, "C: dmin=6  base"),
  .headline(tf_D, "D: dmin=6  forced"),
  .headline(tf_E, "E: dmin=4  base"),
  .headline(tf_F, "F: dmin=4  forced")
)

gt::gt(.all_headline) |>
  gt::tab_header(title = "Headline figures across six configs") |>
  gt::tab_source_note(paste(
    "Compare the base configs (A -> C -> E) to see the effect of",
    "relaxing dmin.grf on GRF's output and downstream identification.",
    "Compare (A -> B), (C -> D), (E -> F) to see conf_force's",
    "contribution at each dmin.grf setting; if base and forced agree",
    "at a given dmin.grf, the threshold is relaxed enough that GRF",
    "alone suffices."
  ))
```


# Session info {#sec-session}

```{r session}
sessionInfo()
```
