---
title: "GBSG Diagnostic: Forcing `er <= 0` as a Candidate in LOO-CV"
subtitle: "Does injecting the full-data cut into every fold stabilize LOO subgroup recovery?"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    code-tools: true
    code-fold: true
    mathjax: true
    theme: cosmo
    fontsize: 11pt
    number-sections: true
---

# Purpose {#sec-purpose}

The `gbsg_cv_qc.qmd` LOO analysis (§5.3.3) showed that 27 of 686
training folds identified a subgroup that *still involves* `er` but
with a different cut point --- typically `er <= 8` paired with a
second factor on `size` or `meno`.  The cut `er <= 0` that appears
in the full-data analysis is selected by the GRF causal-survival-forest
+ policy-tree stage, which chooses a split value specific to each
fold's training data.  Because GRF's split choice is data-dependent,
holding out a specific subject can flip the policy tree's preferred
split on `er` from 0 to a different value, after which the
consistency-passing candidate on that fold is `{er <= 8} & ...`
instead of `{er <= 0}`.

This document tests that hypothesis directly.  `forestsearch()`
accepts a `conf_force` parameter that forces a specific cut
expression into the candidate pool, regardless of what GRF produces.
Setting `conf_force = c("er <= 0")` on the full-data run makes
`er <= 0` available on every LOO fold.  If GRF split-point
instability is the dominant driver of the non-exact-match folds, we
expect:

1. **LOO exact-match rate jumps** from 96.1% to essentially 100%.
2. **`sens_H`, `PPV_H` improve** toward paper-level values.
3. **The `FA_0 = 1.95 (1.04, 3.67)` vs `KfA_0 = 0.50 (0.22, 1.16)` HR
   disagreement** in §5.3.2 disappears --- FA_0 and KfA_0 should
   have nearly identical subject membership.

If those three predictions hold, GRF split-point instability is
confirmed as the dominant driver, and concrete remedies become
available (GRF seed-averaging, a wider default cut grid on the
offending variable, or a user-facing recommendation to set
`conf_force` when stability is a priority).  If the predictions
don't hold, something else is contributing --- most likely LASSO
screening sometimes dropping `er`, or the consistency threshold
boundary fluctuating with the specific fold --- and the next
diagnostic should target those.

**This is a *diagnostic* configuration, not a recommended analysis
setting.**  Pre-specifying a cut point that was learned from the
same data contaminates the subsequent CV as an inference tool.
We use it here purely to isolate GRF's contribution to LOO
variance.


# Setup {#sec-setup}

```{r load-packages}
#| message: false
#| warning: false
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(future)
library(foreach)
library(doFuture)
```

```{r config}
# Data
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name     <- "rfstime"
event.name       <- "status"
treat.name       <- "hormon"
id.name          <- "id"

# ForestSearch paper-reproduction settings (match gbsg_cv_qc.qmd)
fs_args_common <- list(
  is.RCT                 = TRUE,
  use_lasso              = TRUE,
  use_grf                = TRUE,
  use_twostage           = FALSE,
  sg_focus               = "maxSG",
  hr.threshold           = 1.25,
  hr.consistency         = 1.00,
  pconsistency.threshold = 0.90,
  stop_threshold         = 0.95,
  fs.splits              = 400L,
  n.min                  = 60L,
  d0.min                 = 10L,
  d1.min                 = 10L,
  maxk                   = 2L,
  max_subgroups_search   = 10L,
  dmin.grf               = 4,
  frac.tau               = 0.8,
  quiet                  = TRUE,
  details                = FALSE,
  plot.sg                = FALSE,
  plot.grf               = FALSE
)

# Parallel execution
n_workers <- max(1L, parallel::detectCores() - 1L)
```

```{r data-setup}
df <- survival::gbsg
df$grade3 <- as.integer(df$grade == "3")
df$id     <- seq_len(nrow(df))
df$rfstime <- df$rfstime/30.4375
```


# Full-data ForestSearch (two runs) {#sec-full}

## Baseline (no forced cuts)

```{r fs-baseline}
#| message: false
.t0 <- proc.time()
fs_baseline <- do.call(
  forestsearch,
  c(list(
    df.analysis      = df,
    confounders.name = confounders.name,
    outcome.name     = outcome.name,
    event.name       = event.name,
    treat.name       = treat.name,
    id.name          = id.name,
    conf_force       = NULL,
    parallel_args    = list(plan         = "multisession",
                            workers      = n_workers,
                            show_message = FALSE)
  ), fs_args_common)
)
plan("sequential")
.baseline_full_seconds <- as.numeric((proc.time() - .t0)["elapsed"])

cat(sprintf(
  "Baseline full-data sg.harm: %s  (%.1f s)\n",
  if (is.null(fs_baseline$sg.harm)) "(none)" else
    paste(fs_baseline$sg.harm, collapse = " & "),
  .baseline_full_seconds
))
```

## With `conf_force = c("er <= 0")`

```{r fs-forced}
#| message: false
.t0 <- proc.time()
fs_forced <- do.call(
  forestsearch,
  c(list(
    df.analysis      = df,
    confounders.name = confounders.name,
    outcome.name     = outcome.name,
    event.name       = event.name,
    treat.name       = treat.name,
    id.name          = id.name,
    conf_force       = c("er <= 0"),         # <<-- the only change
    parallel_args    = list(plan         = "multisession",
                            workers      = n_workers,
                            show_message = FALSE)
  ), fs_args_common)
)
plan("sequential")
.forced_full_seconds <- as.numeric((proc.time() - .t0)["elapsed"])

cat(sprintf(
  "Forced-cut full-data sg.harm: %s  (%.1f s)\n",
  if (is.null(fs_forced$sg.harm)) "(none)" else
    paste(fs_forced$sg.harm, collapse = " & "),
  .forced_full_seconds
))
```

Both configurations should identify `{er <= 0}` on the full data.
The forced-cut run guarantees it is in the candidate pool; the
baseline run reaches it via GRF.


# N-fold (LOO) Cross-Validation: baseline {#sec-loo-baseline}

```{r loo-baseline}
#| message: false
.t0 <- proc.time()
loo_baseline <- forestsearch_Kfold(
  fs.est        = fs_baseline,
  Kfolds        = nrow(df),
  parallel_args = list(plan         = "multisession",
                       workers      = n_workers,
                       show_message = FALSE)
)
plan("sequential")
.baseline_loo_seconds <- as.numeric((proc.time() - .t0)["elapsed"])

loo_baseline_out <- forestsearch_KfoldOut(
  loo_baseline, outall = TRUE, details = FALSE
)

sprintf("Baseline LOO finished in %.1f s", .baseline_loo_seconds)
```


# N-fold (LOO) Cross-Validation: forced cut {#sec-loo-forced}

```{r loo-forced}
#| message: false
.t0 <- proc.time()
loo_forced <- forestsearch_Kfold(
  fs.est        = fs_forced,
  Kfolds        = nrow(df),
  parallel_args = list(plan         = "multisession",
                       workers      = n_workers,
                       show_message = FALSE)
)
plan("sequential")
.forced_loo_seconds <- as.numeric((proc.time() - .t0)["elapsed"])

loo_forced_out <- forestsearch_KfoldOut(
  loo_forced, outall = TRUE, details = FALSE
)

sprintf("Forced-cut LOO finished in %.1f s", .forced_loo_seconds)
```


# Side-by-side comparison {#sec-compare}

## Agreement and subgroup-finding metrics

```{r compare-metrics}
.metric_names <- c("sens_H", "sens_Hc", "ppv_H", "ppv_Hc")
.find_names   <- c("Any", "Exact", "At least 1",
                   "Cov1", "Cov2",
                   "Cov 1 & 2", "Cov1 exact", "Cov2 exact")

.compare <- rbind(
  data.frame(
    metric   = .metric_names,
    Baseline = unname(loo_baseline_out$sens_metrics_original[.metric_names]),
    Forced   = unname(loo_forced_out$sens_metrics_original[.metric_names]),
    stringsAsFactors = FALSE
  ),
  data.frame(
    metric   = .find_names,
    Baseline = unname(loo_baseline_out$find_metrics[.find_names]),
    Forced   = unname(loo_forced_out$find_metrics[.find_names]),
    stringsAsFactors = FALSE
  )
)
.compare$Delta <- .compare$Forced - .compare$Baseline

gt::gt(.compare) |>
  gt::tab_header(
    title    = "LOO-CV agreement metrics: baseline vs conf_force('er <= 0')",
    subtitle = sprintf(
      "GBSG, N = %d, paper-reproduction config",
      nrow(df)
    )
  ) |>
  gt::fmt_number(columns = c("Baseline", "Forced", "Delta"),
                 decimals = 3) |>
  gt::cols_label(
    metric   = "Metric",
    Baseline = "Baseline",
    Forced   = "conf_force",
    Delta    = "Δ (Forced − Baseline)"
  ) |>
  gt::tab_source_note(paste(
    "Positive Δ on sens_H, ppv_H, and Exact supports the hypothesis",
    "that GRF split-point instability is the dominant driver of the",
    "non-exact-match LOO folds in gbsg_cv_qc.qmd §5.3.3.  Near-zero Δ",
    "would indicate the instability is NOT primarily GRF-driven."
  ))
```

## FA_0 vs KfA_0 hazard-ratio comparison

This is the §5.3.2-equivalent table.  The concerning baseline result
in `gbsg_cv_qc.qmd` was FA_0 = 1.95 (1.04, 3.67) vs KfA_0 = 0.50
(0.22, 1.16) --- point estimates disagreeing directionally with
barely-overlapping CIs.

```{r fa-vs-kfa-compare}
.extract_row <- function(tab_all, row_name) {
  if (is.null(tab_all) || !row_name %in% rownames(tab_all)) {
    return(list(n = NA, HR = NA_character_))
  }
  list(
    n  = tab_all[row_name, "n"],
    HR = tab_all[row_name, "Hazard ratio"]
  )
}

.b_fa0 <- .extract_row(loo_baseline_out$tab_all, "FA_0")
.b_kf0 <- .extract_row(loo_baseline_out$tab_all, "KfA_0")
.f_fa0 <- .extract_row(loo_forced_out$tab_all,   "FA_0")
.f_kf0 <- .extract_row(loo_forced_out$tab_all,   "KfA_0")

.harm_tab <- data.frame(
  Config   = c("Baseline", "Baseline",
               "Forced",   "Forced"),
  Analysis = c("FA_0 (full-data)", "KfA_0 (LOO-CV)",
               "FA_0 (full-data)", "KfA_0 (LOO-CV)"),
  n        = c(.b_fa0$n, .b_kf0$n, .f_fa0$n, .f_kf0$n),
  HR       = c(.b_fa0$HR, .b_kf0$HR, .f_fa0$HR, .f_kf0$HR),
  stringsAsFactors = FALSE
)

gt::gt(.harm_tab) |>
  gt::tab_header(
    title    = "Harm-subgroup HR: full-analysis (FA_0) vs LOO-CV (KfA_0)",
    subtitle = "Both configurations"
  ) |>
  gt::tab_source_note(paste(
    "If forcing 'er <= 0' aligns FA_0 and KfA_0 to nearly identical",
    "HRs, the §5.3.2 disagreement is attributable to GRF split-point",
    "drift across folds.  A persistent disagreement under the forced",
    "cut would indicate a different mechanism."
  ))
```

## Per-fold identified-subgroup distribution

Compares the empirical distribution of fold-level subgroups between
the two configurations.  Under the forced cut, almost every fold
should either identify `{er <= 0}` (the forced candidate wins) or
report no subgroup at all.

```{r compare-per-fold}
.fold_dist <- function(kfold_out, label) {
  resCV <- kfold_out$resCV
  fold_sgs <- vapply(
    unique(resCV$cvindex),
    function(k) {
      rows <- resCV[resCV$cvindex == k, , drop = FALSE]
      sg1 <- as.character(rows$sg1[1])
      sg2 <- as.character(rows$sg2[1])
      if (is.na(sg1) && is.na(sg2)) return("(no subgroup)")
      if (is.na(sg2)) return(sg1)
      paste(sg1, sg2, sep = " & ")
    },
    character(1)
  )

  tab <- as.data.frame(table(sg = fold_sgs), stringsAsFactors = FALSE)
  tab$Config <- label
  tab <- tab[order(-tab$Freq), , drop = FALSE]
  names(tab)[names(tab) == "sg"]   <- "Training-FS subgroup"
  names(tab)[names(tab) == "Freq"] <- "n folds"
  tab
}

.dist_baseline <- .fold_dist(loo_baseline, "Baseline")
.dist_forced   <- .fold_dist(loo_forced,   "Forced")

gt::gt(.dist_baseline) |>
  gt::tab_header(
    title = "Baseline: LOO training-fold subgroups"
  ) |>
  gt::cols_move_to_start(columns = "Config")

gt::gt(.dist_forced) |>
  gt::tab_header(
    title = "Forced conf_force('er <= 0'): LOO training-fold subgroups"
  ) |>
  gt::cols_move_to_start(columns = "Config")
```


# Summary checkpoint {#sec-summary}

```{r summary-table}
.summary_df <- data.frame(
  Check = c(
    "Baseline full-data sg.harm contains '{er <= 0}'",
    "Forced    full-data sg.harm contains '{er <= 0}'",
    "Forced LOO Exact match - Baseline Exact match > 2 pp",
    "Forced LOO sens_H     - Baseline sens_H     > 0.05",
    "Forced LOO ppv_H      - Baseline ppv_H      > 0.05"
  ),
  Observed = c(
    any(grepl("er <= 0", fs_baseline$sg.harm, fixed = TRUE)),
    any(grepl("er <= 0", fs_forced$sg.harm,   fixed = TRUE)),
    (loo_forced_out$find_metrics["Exact"]
      - loo_baseline_out$find_metrics["Exact"]) > 0.02,
    (loo_forced_out$sens_metrics_original["sens_H"]
      - loo_baseline_out$sens_metrics_original["sens_H"]) > 0.05,
    (loo_forced_out$sens_metrics_original["ppv_H"]
      - loo_baseline_out$sens_metrics_original["ppv_H"]) > 0.05
  ),
  stringsAsFactors = FALSE
)
.summary_df$Result <- ifelse(.summary_df$Observed,
                             "\U2713 YES", "\U2717 NO")

gt::gt(.summary_df[, c("Check", "Result")]) |>
  gt::tab_header(title = "Checks relevant to the GRF-instability hypothesis")
```


# Session info {#sec-session}

```{r session}
sessionInfo()
```
