---
title: "ForestSearch Cross-Validation QC"
subtitle: "GBSG: matching León et al. (2024) Section 4.1 CV properties"
author: "Larry F. León"
date: today
format:
  html:
    code-fold: true
    toc: true
    toc-depth: 3
execute:
  warning: false
  message: false
  fig-align: center
knitr:
  opts_chunk:
    comment: "#>"
---

# Purpose {#sec-purpose}

This document exists solely to **QC the ForestSearch cross-validation
machinery** against the benchmarks reported in León et al. (2024)
Section 4.1.  All bootstrap, FPR-calibration, theta-grid, sensitivity-sweep,
visualisation, and summary sections have been stripped out — only the
components needed to:

1. Reproduce the full-data ForestSearch analysis (to provide the `fs`
   object that CV consumes), and
2. Run the two CV procedures (10-fold repeated and N-fold leave-one-out),

are retained here.

**Paper targets (León et al. 2024, Section 4.1, GBSG, `maxSG` search):**

| Quantity | Paper value |
|---------|-------------|
| Identified Ĥ (full data) | `er <= 0` (ER-negative) |
| θ̂(Ĥ)                     | 1.95 (1.05, 3.61) |
| θ̂(Ĥc)                    | 0.61 (0.47, 0.80) |
| N-fold (LOO) subgroup recovery | 686 / 686 folds identify `er <= 0` |
| 10-fold CV: median folds with identified subgroup | 9 / 10 |
| 10-fold CV: sensCV(Ĥ)    | ≈ 73% |
| 10-fold CV: ppvCV(Ĥ)     | ≈ 83% |
| 10-fold CV: sensCV(Ĥc)   | ≈ 98% |
| 10-fold CV: ppvCV(Ĥc)    | ≈ 96% |
| 10-fold CV: exact-match reproduction rate | 70% (median) |

**Configuration note.**  `use_twostage` post-dates the paper and
`sg_focus = "maxSG"` was the paper's search direction, so
`use_twostage = FALSE` is the reproducibility default here.  If a run
produces materially different CV metrics from the paper benchmarks,
re-check the `CV Configuration Sanity Check` section before
investigating the CV code itself.


# Setup {#sec-setup}

## Packages

```{r load-packages}
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(foreach)
library(doFuture)
library(future)
```

## Parameters

Only the parameters relevant to the FS + CV pipeline are retained.  All
FPR / theta-grid / sweep parameters have been removed.

```{r qc-params}
#| include: false

# ── Parallel processing ────────────────────────────────────────────────────
# Set to detectCores() - 1 for production; 2 for CRAN checks.
n_workers <- floor(0.95 * max(1L, parallel::detectCores(logical = FALSE) - 1L))

# ── ForestSearch operational parameters (per León et al. 2024) ─────────────
c1_paper            <- 1.25    # hr.threshold (screening)
c2_paper            <- 1.00    # hr.consistency (consistency)
pcons_val           <- 0.90    # pconsistency.threshold
stop_thresh_val     <- 0.95    # early-stopping threshold
fs_splits_val       <- 400L    # consistency-evaluation splits
max_sgs_search_val  <- 10L
n_min_val           <- 60L
d0_min_val          <- 10L
d1_min_val          <- 10L
maxk_val            <- 2L

# ── Cross-validation counts ────────────────────────────────────────────────
# León et al. (2024) Section 4.1 uses 200 random 10-fold CV analyses;
# for fast knits, lower Ksims_cv to 10-20 at the cost of stability
# in the median sensitivity / ppv estimates.
Ksims_cv <- 1L
```


# Data: German Breast Cancer Study Group {#sec-data}

```{r data-setup}
# Load GBSG data (included in forestsearch package)
df.analysis <- gbsg

# Prepare analysis variables
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
  treat <- hormon
})

# Define variable roles
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name   <- "status"
id.name      <- "id"
treat.name   <- "hormon"

cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```


# ForestSearch Analysis {#sec-forestsearch}

## Running ForestSearch

Paper configuration: `sg_focus = "maxSG"`, LASSO + GRF screening,
single-stage consistency evaluation (`use_twostage = FALSE`).

```{r forestsearch-main}
#| fig-width: 10
#| fig-height: 7

t0 <- proc.time()

fs <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  outcome.name     = outcome.name,
  treat.name       = treat.name,
  event.name       = event.name,
  id.name          = id.name,
  # Threshold parameters (per León et al. 2024)
  hr.threshold           = c1_paper,
  hr.consistency         = c2_paper,
  pconsistency.threshold = pcons_val,
  stop_threshold         = stop_thresh_val,
  # Search configuration (paper: sg_focus = "maxSG", single-stage)
  sg_focus             = "maxSG",
  max_subgroups_search = max_sgs_search_val,
  use_twostage         = FALSE,        # paper predates two-stage
  # Factor selection
  use_grf                   = TRUE,
  return_selected_cuts_only = TRUE,
  dmin.grf = 4,
  frac.tau = 0.8,
  use_lasso                 = TRUE,
  cut_type                  = "default",
  # Subgroup constraints
  maxk   = maxk_val,
  n.min  = n_min_val,
  d0.min = d0_min_val,
  d1.min = d1_min_val,
  # Consistency evaluation
  fs.splits = fs_splits_val,
  # Parallel processing
  parallel_args = list(
    plan         = "multisession",
    workers      = n_workers,
    show_message = TRUE
  ),
  # Output
  showten_subgroups = TRUE,
  details           = TRUE,
  quiet = TRUE,
  plot.sg           = FALSE
)

plan("sequential")
fs_seconds <- as.numeric((proc.time() - t0)["elapsed"])
cat(sprintf("\nForestSearch completed in %.1f seconds\n", fs_seconds))
```

## ForestSearch Results

### Identified Subgroup

```{r fs-subgroup}
cat("Identified subgroup (H):", paste(fs$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs$df.est$treat.recommend == 0)))
```

### Treatment Effect Estimates

```{r fs-estimates}
res_tabs <- sg_tables(fs, ndecimals = 3, which_df = "est")
res_tabs$tab_estimates
```


# Cross-Validation {#sec-cv}

## CV Configuration Sanity Check {#sec-cv-config}

The CV functions execute with `cv_args <- fs$args_call_all` (plus a
few forced overrides: sequential plan, `details = FALSE`,
`quiet = TRUE`, `plot.sg = FALSE`, `ps_hat = NULL`).  So *every*
parameter from the original `forestsearch()` call propagates into the
CV folds — not just the paper-critical ones.  This section reports
the effective settings two ways:

1. **Paper-reproducibility focus** — the handful of parameters that
   drive paper reproducibility, flagged against their paper-target
   values.
2. **Full propagated configuration** — every non-infrastructure entry
   in `fs$args_call_all`, auto-discovered so nothing can drift out of
   the sanity-check when new parameters are added to
   `forestsearch()`.

Paper references (León et al. 2024, Section 4.1):

- **N-fold CV** = leave-one-out (`Kfolds = N`).
- **10-fold CV** across 200 random simulations.
- Subgroup search uses LASSO first-stage screen + GRF candidate
  factors, single-stage consistency (`use_twostage = FALSE`).

### Paper-reproducibility focus

```{r cv-config-paper-focus}
# Map of paper-critical arguments to their paper-target values.
# NOTE: adding a new entry here just adds a row; the table is not
# the thing the CV run actually consumes — that is fs$args_call_all
# directly via cv_args <- fs_args inside forestsearch_Kfold/tenfold.
.paper_targets <- list(
  sg_focus               = "maxSG",
  use_twostage           = FALSE,
  use_lasso              = FALSE,
  use_grf                = TRUE,
  hr.threshold           = 1.25,
  hr.consistency         = 1.00,
  pconsistency.threshold = 0.90,
  maxk                   = 2L
)

.args <- fs$args_call_all

.paper_df <- data.frame(
  Setting = names(.paper_targets),
  Value   = vapply(names(.paper_targets), function(nm) {
    val <- .args[[nm]]
    if (is.null(val)) "NULL" else as.character(val)
  }, character(1)),
  Paper_target = vapply(.paper_targets, function(v) as.character(v),
                        character(1)),
  stringsAsFactors = FALSE
)
.paper_df$Matches <- ifelse(.paper_df$Value == .paper_df$Paper_target,
                            "YES", "NO")

# CV-runner parameters (separate from forestsearch() args, set in this QMD)
.runner_df <- data.frame(
  Setting      = c("Kfolds_kfold", "Kfolds_oob", "Ksims_cv"),
  Value        = c("10", as.character(nrow(df.analysis)),
                   as.character(Ksims_cv)),
  Paper_target = c("10", as.character(nrow(df.analysis)), "200"),
  stringsAsFactors = FALSE
)
.runner_df$Matches <- ifelse(.runner_df$Value == .runner_df$Paper_target,
                             "YES", "NO")

.focus_df <- rbind(.runner_df, .paper_df)
rownames(.focus_df) <- NULL

gt::gt(.focus_df) |>
  gt::tab_header(title = "Paper-critical CV settings vs León et al. (2024)") |>
  gt::cols_label(
    Setting      = "Setting",
    Value        = "In effect",
    Paper_target = "Paper target",
    Matches      = "Match?"
  ) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold", color = "darkred"),
    locations = gt::cells_body(columns = "Matches", rows = Matches == "NO")
  )
```

### Full propagated configuration (auto-discovered)

Every non-infrastructure entry in `fs$args_call_all`.  Infrastructure
entries (data, parallel plan, reusable-object caches, display flags)
are excluded because they don't affect reproducibility — the CV code
explicitly overrides them.

```{r cv-config-full-discover}
# Parameters the CV code deliberately overrides or ignores.  Exclude
# from the reproducibility audit so the table focuses on things that
# actually affect results.
.cv_exclude <- c(
  "df.analysis",         # data
  "parallel_args",       # infrastructure
  "grf_res", "grf_cuts", # reusable GRF objects, not config
  "ps_hat",              # reusable PS object, recomputed per fold
  "details", "quiet", "plot.sg", "plot.grf",
  "showten_subgroups"    # display flags
)

.all_args   <- fs$args_call_all
.keep_names <- setdiff(names(.all_args), .cv_exclude)
.keep_args  <- .all_args[.keep_names]

# Flatten each value to a single-cell character representation.
.flatten_value <- function(x) {
  if (is.null(x))                       return("NULL")
  if (is.list(x) && length(x) == 0)     return("list()")
  if (is.list(x))                       return(sprintf("<list of %d>", length(x)))
  if (is.function(x))                   return("<function>")
  if (length(x) == 0)                   return("(empty)")
  if (length(x) > 1)                    return(paste(as.character(x),
                                                     collapse = ", "))
  as.character(x)
}

.full_df <- data.frame(
  Setting = names(.keep_args),
  Value   = vapply(.keep_args, .flatten_value, character(1)),
  stringsAsFactors = FALSE
)
.full_df <- .full_df[order(.full_df$Setting), ]
rownames(.full_df) <- NULL

gt::gt(.full_df) |>
  gt::tab_header(
    title = sprintf(
      "All %d propagated forestsearch() arguments", nrow(.full_df)),
    subtitle = "(auto-discovered from fs$args_call_all)"
  ) |>
  gt::tab_source_note(paste(
    "Any parameter added to forestsearch() in a future package version",
    "appears here automatically without editing this chunk.  The CV code",
    "itself uses fs$args_call_all directly, so this table reflects",
    "exactly what the CV folds see (modulo the forced infrastructure",
    "overrides listed above)."))
```


## 10-Fold Cross-Validation {#sec-cv-kfold}

Paper benchmark: median 9/10 folds identify a subgroup, sensCV(Ĥ) ≈ 73%,
ppvCV(Ĥ) ≈ 83%, sensCV(Ĥc) ≈ 98%, ppvCV(Ĥc) ≈ 96%, exact-match rate
≈ 70% (median).

### Inspection: first simulation's 10 training folds {#sec-cv-kfold-inspect}

Before the aggregate `forestsearch_tenfold()` run, this block replicates
the **exact** data preparation and fold assignment that the function
would apply to `ksim = 1` (seed `8316951 + 1000 * 1 = 8317951`), then
calls `forestsearch()` directly on each of the 10 training folds with
`details = TRUE` and `quiet = FALSE`.  The printout below shows the
full training-FS diagnostic for every fold of the first simulation.

```{r kfold-cv-inspect-sim1}
#| results: "markup"

# Replicate the data-prep stanza inside forestsearch_tenfold
.get_names <- c(confounders.name, outcome.name, event.name, id.name, treat.name)
.dfa <- fs$df.est[, c(.get_names, "treat.recommend")]
names(.dfa)[names(.dfa) == "treat.recommend"] <- "treat.recommend.original"
.dfnew <- as.data.frame(.dfa)

# Replicate sim = 1 shuffling: set.seed(seed + 1000 * ksim)
set.seed(8316951L + 1000L * 1L)
.id_sample <- sample(seq_len(nrow(.dfnew)), replace = FALSE)
.df_scrambled <- .dfnew[.id_sample, ]

# Replicate fold assignment: cut(seq_len(N), breaks = Kfolds)
.folds_sim1 <- cut(seq_len(nrow(.df_scrambled)), breaks = 10, labels = FALSE)

# Build verbose cv_args (mirrors forestsearch_tenfold SECTION 3
# except details/quiet flags are flipped ON for inspection).
.inspect_cv_args <- fs$args_call_all
.inspect_cv_args$parallel_args <- list(
  plan         = "multisession",
  workers      = n_workers,
  show_message = FALSE
)
.inspect_cv_args$details <- TRUE     # <<-- print full training FS details
.inspect_cv_args$quiet   <- FALSE    # <<-- include startup config banner
.inspect_cv_args$plot.sg <- FALSE
.inspect_cv_args$ps_hat  <- NULL

# Loop over all 10 folds of sim = 1
.inspect_rows_kfold <- vector("list", 10L)

for (.cv_i in seq_len(10L)) {
  cat(sprintf(
    "\n\n=========================================================\n"))
  cat(sprintf("  sim = 1  |  fold = %d / 10  |  seed = %d\n",
              .cv_i, 8316951L + 1000L * 1L))
  cat(sprintf(
    "=========================================================\n"))

  .test_idx <- which(.folds_sim1 == .cv_i, arr.ind = TRUE)
  .x_train  <- .df_scrambled[-.test_idx, ]

  cat(sprintf("Training-fold size:  n = %d\n", nrow(.x_train)))
  cat(sprintf("Held-out test size:  n = %d\n\n", length(.test_idx)))

  .inspect_cv_args$df.analysis <- .x_train

  .fs_train_i <- try(
    do.call(forestsearch, .inspect_cv_args),
    silent = FALSE
  )
  plan("sequential")

  if (inherits(.fs_train_i, "try-error")) {
    .sg_found_i <- "(forestsearch ERROR)"
  } else if (is.null(.fs_train_i$sg.harm)) {
    .sg_found_i <- "(no subgroup)"
  } else {
    .sg_found_i <- paste(.fs_train_i$sg.harm, collapse = " & ")
  }

  .inspect_rows_kfold[[.cv_i]] <- data.frame(
    fold            = .cv_i,
    n_train         = nrow(.x_train),
    training_SG     = .sg_found_i,
    stringsAsFactors = FALSE
  )
}

.inspect_summary_kfold <- do.call(rbind, .inspect_rows_kfold)
```

#### Scan summary across the 10 training folds (sim = 1)

```{r kfold-cv-inspect-summary}
gt::gt(.inspect_summary_kfold) |>
  gt::tab_header(title = "sim = 1: training-fold subgroups") |>
  gt::cols_label(
    fold        = "Fold",
    n_train     = "n (train)",
    training_SG = "Training-FS subgroup"
  ) |>
  gt::tab_source_note(paste(
    "Full-data subgroup:",
    paste(fs$sg.harm, collapse = " & "),
    "| Any row equal to this is an 'exact match' for sim = 1."
  ))
```


### Aggregate 10-fold CV across simulations

```{r kfold-cv}
t0 <- proc.time()
fs_kfold <- forestsearch_tenfold(
  fs.est        = fs,
  sims          = Ksims_cv,
  Kfolds        = 10,
  details       = FALSE,
  parallel_args = list(
    plan         = "multisession",
    workers      = n_workers,
    show_message = FALSE
  )
)
plan("sequential")
kfold_seconds <- as.numeric((proc.time() - t0)["elapsed"])

metrics_tables <- cv_metrics_tables(fs_kfold)
metrics_tables
```

### 10-fold CV: paper-comparable summary

```{r kfold-cv-compare}
# Per-simulation "folds with identified subgroup" is not directly stored
# in the fs_tenfold object's summary outputs; derive it from the raw
# metrics matrix if available.  Otherwise, just print the median sens /
# ppv summary side-by-side with the paper values.

.paper_10fold <- data.frame(
  metric = c("sens_H", "ppv_H", "sens_Hc", "ppv_Hc"),
  paper  = c(0.73, 0.83, 0.98, 0.96)
)

.qc_10fold <- data.frame(
  metric = names(fs_kfold$sens_summary),
  qc     = as.numeric(fs_kfold$sens_summary)
)

compare_df <- merge(.paper_10fold, .qc_10fold, by = "metric", sort = FALSE)
compare_df$diff_pp <- (compare_df$qc - compare_df$paper) * 100

gt::gt(compare_df) |>
  gt::fmt_number(columns = c("paper", "qc"), decimals = 3) |>
  gt::fmt_number(columns = "diff_pp", decimals = 1) |>
  gt::tab_header(title = "10-fold CV: this QC vs León et al. (2024)") |>
  gt::cols_label(
    metric  = "Metric",
    paper   = "Paper (median)",
    qc      = "This QC (median)",
    diff_pp = "Diff (pp)"
  ) |>
  gt::tab_source_note(paste(
    "A gap of > ~5 pp on sens_H after the sanity-check passes suggests a",
    "change in the CV procedure itself; verify package version and",
    "forestsearch_tenfold() behavior before investigating further."))
```


### Diagnostic: legacy-style splitter sensitivity check {#sec-cv-splitter-debug}

The legacy (paper-vintage) ForestSearch used an **exact 50/50 split
without replacement** when evaluating subgroup consistency:
`sample(seq_len(N), size = round(N/2), replace = FALSE)`.  The current
package uses **independent Bernoulli trials**
(`sample(c(TRUE, FALSE), N, replace = TRUE, prob = c(0.5, 0.5))`),
which gives binomially-distributed split sizes
(SD ≈ √(N/4)).  For small subgroups this extra per-split variance
plausibly destabilises fold-to-fold subgroup recovery in CV, as
discussed in `legacy_vs_current_audit.md` §1.1.

This block monkey-patches `forestsearch::run_single_consistency_split()`
via `assignInNamespace()` to use the legacy exact-50/50 splitter,
re-runs a small number of 10-fold CV simulations, and compares
sens_H / sens_Hc / ppv_H / ppv_Hc against the current Bernoulli
version.  If the *Legacy* row is materially closer to the *Paper
target* row than *Current* is, the splitter difference is the
dominant cause of the paper-vs-current CV gap.  If Legacy and Current
are roughly equal, §1.2 (Pcons denominator convention) or §1.3
(per-candidate seeding) from the audit is the more likely
explanation.

Both versions run **sequentially** (no multisession workers) so the
namespace override is in effect — workers would reload the package
and bypass the patch.  Cost is roughly 1–2 minutes per sim × 2
versions.  Default `debug_splitter_sims = 10L` → ~15–20 minutes total.
Set it to `0L` to skip the diagnostic entirely.

```{r cv-splitter-diagnostic}
#| results: "markup"

debug_splitter_sims <- 10L   # set to 0L to skip the whole diagnostic

if (debug_splitter_sims > 0L) {

  # -------------------------------------------------------------------
  # 1.  Define legacy-style splitter
  # -------------------------------------------------------------------
  # Matches subgroup_consistency_v0.R lines 157-169: exact 50/50 via
  # sample(...,replace=FALSE).  Non-estimable splits return 0 (not NA),
  # matching the legacy convention that
  #   flag.consistency <- rep(0, n.splits)
  #   ... only overwritten when BOTH splits estimable AND HRs > threshold
  # so the denominator in Pcons = mean(flag.consistency) is always
  # n.splits.
  legacy_run_single_consistency_split <- function(
      df.x, N.x, hr.consistency, cox_init = 0,
      estimator_fn = NULL, consistency_threshold = NULL) {

    # Exact 50/50 split (no replacement)
    n_half <- round(N.x / 2)
    idx_split1 <- tryCatch(
      sample(seq_len(N.x), size = n_half, replace = FALSE),
      error = function(e) NULL
    )
    if (is.null(idx_split1)) return(0)           # legacy: error -> 0

    df.x <- as.data.frame(df.x)
    df.x.split1 <- df.x[ idx_split1, , drop = FALSE]
    df.x.split2 <- df.x[-idx_split1, , drop = FALSE]

    if (nrow(df.x.split1) < 5 || nrow(df.x.split2) < 5) return(0)

    # ---- Survival path (Larry's GBSG setup) ------------------------
    if (is.null(estimator_fn)) {
      if (sum(df.x.split1$Event) < 2 ||
          sum(df.x.split2$Event) < 2) return(0)

      fit1 <- tryCatch(
        suppressWarnings(survival::coxph(
          survival::Surv(Y, Event) ~ Treat, data = df.x.split1,
          init = cox_init, robust = FALSE,
          model = FALSE, x = FALSE, y = FALSE)),
        error = function(e) NULL
      )
      fit2 <- tryCatch(
        suppressWarnings(survival::coxph(
          survival::Surv(Y, Event) ~ Treat, data = df.x.split2,
          init = cox_init, robust = FALSE,
          model = FALSE, x = FALSE, y = FALSE)),
        error = function(e) NULL
      )
      if (is.null(fit1) || is.null(fit2)) return(0)
      hr.split1 <- exp(fit1$coefficients[1])
      hr.split2 <- exp(fit2$coefficients[1])
      if (is.na(hr.split1) || is.na(hr.split2)) return(0)
      return(as.numeric(
        hr.split1 > hr.consistency && hr.split2 > hr.consistency))
    }

    # ---- GLM path (not exercised for GBSG survival) -----------------
    for (half in list(df.x.split1, df.x.split2)) {
      treat_col <- intersect(c("Treat", "treat"), names(half))
      if (length(treat_col) > 0L) {
        tc <- treat_col[1L]
        if (sum(half[[tc]] == 1L) < 3L ||
            sum(half[[tc]] == 0L) < 3L) return(0)
      }
    }
    res1 <- estimator_fn(df.x.split1)
    res2 <- estimator_fn(df.x.split2)
    if (is.na(res1$estimate) || is.na(res2$estimate)) return(0)
    as.numeric(res1$estimate >= consistency_threshold &&
               res2$estimate >= consistency_threshold)
  }

  # -------------------------------------------------------------------
  # 2.  Save original + install legacy override in forestsearch ns
  # -------------------------------------------------------------------
  .original_run_single <- forestsearch::run_single_consistency_split
  assignInNamespace(
    "run_single_consistency_split",
    legacy_run_single_consistency_split,
    ns = "forestsearch"
  )

  # -------------------------------------------------------------------
  # 3.  Run CV with the legacy splitter (sequential)
  # -------------------------------------------------------------------
  #     Must stay in-process so the namespace override is live --
  #     multisession workers would reload forestsearch and bypass it.
  .t0 <- proc.time()
  fs_kfold_legacy <- tryCatch(
    forestsearch_tenfold(
      fs.est        = fs,
      sims          = debug_splitter_sims,
      Kfolds        = 10,
      details       = FALSE,
      parallel_args = list(plan = "sequential",
                           workers = 1, show_message = FALSE)
    ),
    error = function(e) {
      # Restore even on error
      assignInNamespace(
        "run_single_consistency_split",
        .original_run_single, ns = "forestsearch")
      stop("Legacy-splitter CV run failed: ", conditionMessage(e))
    }
  )
  plan("sequential")
  .legacy_seconds <- as.numeric((proc.time() - .t0)["elapsed"])

  # -------------------------------------------------------------------
  # 4.  Restore the original splitter
  # -------------------------------------------------------------------
  assignInNamespace(
    "run_single_consistency_split",
    .original_run_single,
    ns = "forestsearch"
  )

  # -------------------------------------------------------------------
  # 5.  Run matched-sim-count CV with the CURRENT (Bernoulli) splitter
  # -------------------------------------------------------------------
  #     Same sequential plan, same seed base -> fair side-by-side.
  .t0 <- proc.time()
  fs_kfold_current <- forestsearch_tenfold(
    fs.est        = fs,
    sims          = debug_splitter_sims,
    Kfolds        = 10,
    details       = FALSE,
    parallel_args = list(plan = "sequential",
                         workers = 1, show_message = FALSE)
  )
  plan("sequential")
  .current_seconds <- as.numeric((proc.time() - .t0)["elapsed"])

  # -------------------------------------------------------------------
  # 6.  Comparison table
  # -------------------------------------------------------------------
  cat(sprintf(
    "\nSplitter diagnostic: %d sims × 10 folds, sequential execution\n",
    debug_splitter_sims))
  cat(sprintf("  Legacy  splitter:  elapsed = %6.1f s\n", .legacy_seconds))
  cat(sprintf("  Current splitter:  elapsed = %6.1f s\n", .current_seconds))

  .paper_vals <- c(sens_H = 0.73, sens_Hc = 0.98,
                   ppv_H  = 0.83, ppv_Hc  = 0.96)

  .splitter_compare <- data.frame(
    Splitter = c("Legacy (exact 50/50)",
                 "Current (Bernoulli)",
                 "Paper target"),
    sens_H  = c(unname(fs_kfold_legacy$sens_summary["sens_H"]),
                unname(fs_kfold_current$sens_summary["sens_H"]),
                unname(.paper_vals["sens_H"])),
    sens_Hc = c(unname(fs_kfold_legacy$sens_summary["sens_Hc"]),
                unname(fs_kfold_current$sens_summary["sens_Hc"]),
                unname(.paper_vals["sens_Hc"])),
    ppv_H   = c(unname(fs_kfold_legacy$sens_summary["ppv_H"]),
                unname(fs_kfold_current$sens_summary["ppv_H"]),
                unname(.paper_vals["ppv_H"])),
    ppv_Hc  = c(unname(fs_kfold_legacy$sens_summary["ppv_Hc"]),
                unname(fs_kfold_current$sens_summary["ppv_Hc"]),
                unname(.paper_vals["ppv_Hc"])),
    stringsAsFactors = FALSE
  )
  rownames(.splitter_compare) <- NULL

  gt::gt(.splitter_compare) |>
    gt::tab_header(
      title = sprintf(
        "Splitter diagnostic: medians over %d sims × 10 folds",
        debug_splitter_sims),
      subtitle = "Sequential execution; namespace patch via assignInNamespace()"
    ) |>
    gt::fmt_number(
      columns  = c("sens_H", "sens_Hc", "ppv_H", "ppv_Hc"),
      decimals = 3
    ) |>
    gt::tab_source_note(paste(
      "If the 'Legacy' row is materially closer to 'Paper target' than",
      "the 'Current' row is, the exact-50/50 splitter is a key driver of",
      "the paper-vs-current CV gap.  If Legacy ≈ Current, the splitter is",
      "NOT the dominant cause, and §1.2 (Pcons denominator) or §1.3",
      "(per-candidate seeding) from legacy_vs_current_audit.md is the",
      "more likely explanation."
    ))

} else {
  cat("Splitter diagnostic skipped (debug_splitter_sims = 0L).\n")
  cat("Set debug_splitter_sims = 10L (or higher) to run the comparison.\n")
}
```


## N-Fold (Leave-One-Out) Cross-Validation {#sec-cv-oob}

León et al. (2024) Section 4.1 defines N-fold CV as leave-one-out:
*"across the N = 686 training sets (based on deleting a single
subject) all analyses identified Estrogen<=0"*.  We therefore set
`Kfolds = nrow(df.analysis)`.  A smaller `Kfolds` (e.g. 7) would be a
coarse K-fold CV, **not** the paper's N-fold / LOO — each training
fold would be missing ~98 subjects rather than 1.

### Inspection: first 10 "delete-1" training datasets {#sec-cv-oob-inspect}

Before the full N = `r nrow(df.analysis)` leave-one-out run, this
block replicates the exact LOO fold assignment used by
`forestsearch_Kfold()` (no scrambling when `Kfolds = N`; fold *k*
holds out observation *k* by position in `fs$df.est`), then calls
`forestsearch()` directly on each of the first 10 delete-1 training
datasets with `details = TRUE` and `quiet = FALSE`.  The printout
below shows the full training-FS diagnostic for each of those 10
folds.

```{r oob-cv-inspect-first10}
#| results: "markup"

# Replicate forestsearch_Kfold data prep (same as the tenfold version)
.dfa_oob <- fs$df.est[, c(.get_names, "treat.recommend")]
names(.dfa_oob)[names(.dfa_oob) == "treat.recommend"] <- "treat.recommend.original"
.dfnew_oob <- as.data.frame(.dfa_oob)

# LOO: NO scrambling (forestsearch_Kfold scrambles only when
# Kfolds < nrow).  Fold k = observation k by position.
.N_oob <- nrow(.dfnew_oob)
.folds_oob <- cut(seq_len(.N_oob), breaks = .N_oob, labels = FALSE)

# Reuse the verbose cv_args from the k-fold inspection (same details /
# quiet / plot.sg / ps_hat / parallel_args settings apply).
.oob_inspect_cv_args <- .inspect_cv_args

# Loop over the first 10 LOO folds
n_oob_inspect <- 10L
.inspect_rows_oob <- vector("list", n_oob_inspect)

for (.cv_i in seq_len(n_oob_inspect)) {
  cat(sprintf(
    "\n\n=========================================================\n"))
  cat(sprintf("  LOO fold = %d / %d  |  holding out row %d of fs$df.est\n",
              .cv_i, .N_oob, .cv_i))
  cat(sprintf(
    "=========================================================\n"))

  .test_idx <- which(.folds_oob == .cv_i, arr.ind = TRUE)
  .x_train  <- .dfnew_oob[-.test_idx, ]

  cat(sprintf("Training-fold size:  n = %d\n", nrow(.x_train)))
  cat(sprintf("Held-out subject id: %s\n\n",
              as.character(.dfnew_oob[[id.name]][.test_idx])))

  .oob_inspect_cv_args$df.analysis <- .x_train

  .fs_train_i <- try(
    do.call(forestsearch, .oob_inspect_cv_args),
    silent = FALSE
  )
  plan("sequential")

  if (inherits(.fs_train_i, "try-error")) {
    .sg_found_i <- "(forestsearch ERROR)"
  } else if (is.null(.fs_train_i$sg.harm)) {
    .sg_found_i <- "(no subgroup)"
  } else {
    .sg_found_i <- paste(.fs_train_i$sg.harm, collapse = " & ")
  }

  .inspect_rows_oob[[.cv_i]] <- data.frame(
    fold            = .cv_i,
    held_out_id     = .dfnew_oob[[id.name]][.test_idx],
    n_train         = nrow(.x_train),
    training_SG     = .sg_found_i,
    stringsAsFactors = FALSE
  )
}

.inspect_summary_oob <- do.call(rbind, .inspect_rows_oob)
```

#### Scan summary across the first 10 LOO training folds

```{r oob-cv-inspect-summary}
gt::gt(.inspect_summary_oob) |>
  gt::tab_header(title = sprintf(
    "First 10 of %d LOO training folds: identified subgroups",
    .N_oob)) |>
  gt::cols_label(
    fold        = "Fold",
    held_out_id = "Held-out id",
    n_train     = "n (train)",
    training_SG = "Training-FS subgroup"
  ) |>
  gt::tab_source_note(paste(
    "Full-data subgroup:",
    paste(fs$sg.harm, collapse = " & "),
    "| Paper target: all 686 rows identify 'er <= 0'."
  ))
```


### Aggregate N-fold (LOO) CV across all folds

```{r oob-cv}
.oob_kfolds <- nrow(df.analysis)

t0 <- proc.time()
fs_OOB <- forestsearch_Kfold(
  fs.est        = fs,
  details       = FALSE,
  Kfolds        = .oob_kfolds,
  parallel_args = list(
    plan         = "multisession",
    workers      = n_workers,
    show_message = TRUE
  )
)
plan("sequential")
oob_seconds <- as.numeric((proc.time() - t0)["elapsed"])

cv_out <- forestsearch_KfoldOut(
  res     = fs_OOB,
  details = FALSE,
  outall  = TRUE
)

tables <- cv_summary_tables(cv_out)
tables$combined_table
tables$metrics_table
```

### Per-fold subgroup distribution across all N LOO training sets

The previous chunk's `forestsearch_KfoldOut(outall = TRUE)` already
collected the subgroup definition identified by each of the
`r nrow(df.analysis)` LOO training folds into `cv_out$SGs_found`
(a `Kfolds × 2` matrix with columns `sg1`, `sg2`).  The chunks
below tabulate that matrix.  No re-running of `forestsearch()` —
we are summarising work the main OOB run already did.

```{r oob-sg-frequency}
# Build a single "identified subgroup" label per fold by pasting
# sg1 and sg2 (sg2 is NA for depth-1 subgroups; both NA = no subgroup).
.sg_found_df <- as.data.frame(cv_out$SGs_found, stringsAsFactors = FALSE)

.sg_label_per_fold <- vapply(
  seq_len(nrow(.sg_found_df)),
  function(i) {
    s1 <- .sg_found_df$sg1[i]
    s2 <- .sg_found_df$sg2[i]
    if (is.na(s1) && is.na(s2)) return("(no subgroup)")
    parts <- c(s1, s2)
    parts <- parts[!is.na(parts) & parts != ""]
    paste(parts, collapse = " & ")
  },
  character(1)
)

# Full-data subgroup string for comparison
sg_full_label <- paste(fs$sg.harm, collapse = " & ")

# Frequency table, sorted by count descending
.sg_freq <- as.data.frame(
  table(training_SG = .sg_label_per_fold),
  stringsAsFactors = FALSE
)
.sg_freq <- .sg_freq[order(-.sg_freq$Freq), ]
.sg_freq$pct        <- 100 * .sg_freq$Freq / sum(.sg_freq$Freq)
.sg_freq$match_full <- ifelse(
  .sg_freq$training_SG == sg_full_label, "YES", ""
)
rownames(.sg_freq) <- NULL

gt::gt(.sg_freq) |>
  gt::tab_header(
    title = sprintf(
      "LOO training subgroups across all N = %d folds",
      sum(.sg_freq$Freq)),
    subtitle = sprintf("Full-data subgroup: %s", sg_full_label)
  ) |>
  gt::fmt_number(columns = "pct", decimals = 1) |>
  gt::cols_label(
    training_SG = "Training-FS subgroup",
    Freq        = "n folds",
    pct         = "% folds",
    match_full  = "Matches full?"
  ) |>
  gt::tab_source_note(paste(
    "Paper target: all 686 folds identify",
    sprintf("'%s'", sg_full_label),
    "(León et al. 2024, Section 4.1)."
  ))
```

### LOO subgroup-recovery summary statistics

```{r oob-sg-summary-stats}
.n_folds_total   <- length(.sg_label_per_fold)
.n_no_subgroup   <- sum(.sg_label_per_fold == "(no subgroup)")
.n_with_subgroup <- .n_folds_total - .n_no_subgroup
.n_match_full    <- sum(.sg_label_per_fold == sg_full_label)
.n_distinct_sgs  <- length(unique(
  .sg_label_per_fold[.sg_label_per_fold != "(no subgroup)"]
))

.loo_summary <- data.frame(
  Metric = c(
    "Total LOO folds",
    "Folds with any subgroup found",
    "Folds with no subgroup found",
    "Folds exactly matching full-data subgroup",
    "Distinct subgroup definitions across folds"
  ),
  Value = c(
    sprintf("%d", .n_folds_total),
    sprintf("%d (%.1f%%)",
            .n_with_subgroup,
            100 * .n_with_subgroup / .n_folds_total),
    sprintf("%d (%.1f%%)",
            .n_no_subgroup,
            100 * .n_no_subgroup / .n_folds_total),
    sprintf("%d (%.1f%%)",
            .n_match_full,
            100 * .n_match_full / .n_folds_total),
    sprintf("%d", .n_distinct_sgs)
  ),
  stringsAsFactors = FALSE
)

gt::gt(.loo_summary) |>
  gt::tab_header(title = "LOO (delete-1) subgroup-recovery summary") |>
  gt::cols_label(Metric = "Metric", Value = "Value") |>
  gt::tab_source_note(paste(
    "Paper target: 'Folds with any subgroup found' = 100%;",
    "'Folds exactly matching full-data subgroup' = 100%;",
    "'Distinct subgroup definitions' = 1."
  ))
```

### N-fold (LOO) CV: paper-comparable summary

```{r oob-cv-compare}
sg_full   <- paste(fs$sg.harm, collapse = " & ")
pct_found <- fs_OOB$prop_SG_found

cat(sprintf("Full-data subgroup: %s\n", sg_full))
cat(sprintf("Percent of N = %d LOO training folds with an identified subgroup: %s%%\n",
            .oob_kfolds, format(pct_found)))

cat(sprintf("Paper target: 100%% (all 686 folds identify %s)\n",
            "er <= 0"))

# Exact-match rate to the full-data Ĥ across folds
exact_rate <- 100 * mean(fs_OOB$resCV$treat.recommend ==
                         fs_OOB$resCV$treat.recommend.original)
cat(sprintf("Subject-level agreement between LOO and full-data predictions: %.1f%%\n",
            exact_rate))
cat("Paper target: ~100% (agreement implied by identical subgroup recovery)\n")
```


# N-fold (LOO) Cross-Validation: baseline 

```{r loo-baseline}
#| message: false
.t0 <- proc.time()
loo_baseline <- forestsearch_Kfold(
  fs.est        = fs,
  Kfolds        = nrow(df.analysis),
  parallel_args = list(plan         = "multisession",
                       workers      = n_workers,
                       show_message = FALSE)
)
plan("sequential")
.baseline_loo_seconds <- as.numeric((proc.time() - .t0)["elapsed"])

sprintf("Baseline LOO finished in %.1f s", .baseline_loo_seconds)

loo_baseline_out <- forestsearch_KfoldOut(
  res     = loo_baseline,
  details = FALSE,
  outall  = TRUE
)

tables <- cv_summary_tables(loo_baseline_out)
tables$combined_table
tables$metrics_table

```




# Timing {#sec-timing}

```{r timing-summary}
timing_df <- data.frame(
  Component = c("ForestSearch (full data)",
                "10-fold CV",
                "N-fold (LOO) CV"),
  Seconds   = c(fs_seconds, kfold_seconds, oob_seconds),
  Sims      = c(NA_integer_, Ksims_cv, .oob_kfolds)
)
timing_df$Minutes <- timing_df$Seconds / 60

gt::gt(timing_df) |>
  gt::fmt_number(columns = c("Seconds", "Minutes"), decimals = 1) |>
  gt::fmt_integer(columns = "Sims") |>
  gt::sub_missing(missing_text = "—") |>
  gt::tab_header(title = "CV QC: Computational Timing") |>
  gt::tab_source_note(paste(
    "Every knit re-runs ForestSearch and both CV procedures from scratch.",
    "To cache expensive CV runs across knits, wrap the forestsearch_tenfold()",
    "and forestsearch_Kfold() calls in the .cache_compute() helper from",
    "the production gbsg_analysis.qmd."))
```


# References {#sec-refs}

León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024).
"Exploratory subgroup identification in the heterogeneous Cox model:
A relatively simple procedure." *Statistics in Medicine*.
DOI: [10.1002/sim.10163](https://doi.org/10.1002/sim.10163)


# Session Info {#sec-session}

```{r session-info}
plan(sequential)
sessionInfo()
```
