---
title: "GBSG Diagnostic: 10-fold CV with `conf_force('er <= 0')` and GRF Cut Distribution"
subtitle: "Does GRF drift on larger training-set perturbations, and does conf_force recover stability?"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    code-tools: true
    code-fold: true
    mathjax: true
    theme: cosmo
    fontsize: 11pt
    number-sections: true
---

# Purpose {#sec-purpose}

Context for this document: the prior LOO diagnostic
(`gbsg_er_forced_cut_diagnostic.qmd`) ran with `outcome.name = "rfstime"`
(days scale) and `dmin.grf = 60` (60 days ≈ 2 months), which is a
substantially looser GRF filter than `gbsg_cv_qc.qmd`, which runs
with `outcome.name = "time_months"` and the signature default
`dmin.grf = 12` (12 months).  That unit/filter difference accounts
for the apparent disagreement between the two documents' LOO
results (the LOO-baseline in the forced-cut diagnostic showed
686/686 exact match; `gbsg_cv_qc.qmd` §5.3 and §6 showed partial
mismatch).

This document reruns the baseline-vs-forced comparison at 10-fold CV,
**using the same months-scale outcome and `dmin.grf = 12` as
`gbsg_cv_qc.qmd`**, so its results are directly comparable to
§5.2 of that document.

Two questions:

**Q1.**  Under the paper-reproduction config (months scale,
`dmin.grf = 12`), how often does GRF return `er <= 0` as its
policy-tree cut across 10-fold training sets (which are ~90% of
the full sample)?  This is the direct measurement the earlier LOO
diagnostic could not provide, enabled by the `grf_cuts` column
newly preserved in `fold_summary`.

**Q2.**  On folds where GRF drifts to a different `er` cut (e.g.,
`er <= 8`), does forcing `er <= 0` into the candidate pool via
`conf_force = c("er <= 0")` recover a stable `{er <= 0}`
identification?  Or does the consistency stage still prefer the
drifted cut?

If GRF is stable at 10-fold under this config, the forced comparison
will show near-zero Δ.  If GRF drifts on a substantial fraction of
folds, the forced comparison tells us whether the drift is
recoverable by supplying `er <= 0` as an alternative candidate.


# Setup {#sec-setup}

```{r load-packages}
#| message: false
#| warning: false
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(future)
library(foreach)
library(doFuture)
```

```{r config}
# Variable names (match gbsg_cv_qc.qmd exactly)
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name     <- "time_months"      # MONTHS scale to match cv_qc
event.name       <- "status"
treat.name       <- "hormon"
id.name          <- "id"

# ForestSearch paper-reproduction settings (match gbsg_cv_qc.qmd exactly).
# Key alignment: outcome is on months scale, so dmin.grf = 12 means
# 12 months -- matching cv_qc's implicit default.
fs_args_common <- list(
  # Threshold parameters (Leon et al. 2024)
  hr.threshold              = 1.25,
  hr.consistency            = 1.00,
  pconsistency.threshold    = 0.90,
  stop_threshold            = 0.95,
  # Search configuration
  sg_focus                  = "maxSG",
  max_subgroups_search      = 10L,
  use_twostage              = FALSE,
  # Factor selection / GRF
  use_grf                   = TRUE,
  return_selected_cuts_only = TRUE,
  use_lasso                 = TRUE,
  cut_type                  = "default",
  dmin.grf                  = 12,        # <<-- months scale; matches cv_qc default
  # Subgroup constraints
  maxk                      = 2L,
  n.min                     = 60L,
  d0.min                    = 10L,
  d1.min                    = 10L,
  # Consistency evaluation
  fs.splits                 = 400L,
  # Output
  details                   = FALSE,
  quiet                     = TRUE,
  plot.sg                   = FALSE,
  plot.grf                  = FALSE
)

# 10-fold CV
Ksims_cv  <- 100L    # repetitions of 10-fold split (total 1000 sim x fold pairs)
Kfolds_cv <- 10L

# Parallel execution
n_workers <- max(1L, parallel::detectCores() - 1L)
```

```{r data-setup}
df <- survival::gbsg
df$grade3      <- as.integer(df$grade == "3")
df$id          <- seq_len(nrow(df))
df$time_months <- df$rfstime / 30.4375     # <<-- match cv_qc unit conversion
```


# Full-data ForestSearch (two runs) {#sec-full}

Provided for self-containment; these are the same objects the LOO
diagnostic produced.  Both runs should identify `{er <= 0}`.

## Baseline (no forced cuts)

```{r fs-baseline}
#| message: false
fs_baseline <- do.call(
  forestsearch,
  c(list(
    df.analysis      = df,
    confounders.name = confounders.name,
    outcome.name     = outcome.name,
    event.name       = event.name,
    treat.name       = treat.name,
    id.name          = id.name,
    conf_force       = NULL,
    parallel_args    = list(plan         = "multisession",
                            workers      = n_workers,
                            show_message = FALSE)
  ), fs_args_common)
)
plan("sequential")

cat(sprintf(
  "Baseline full-data sg.harm: %s\n",
  if (is.null(fs_baseline$sg.harm)) "(none)" else
    paste(fs_baseline$sg.harm, collapse = " & ")
))
cat(sprintf(
  "Baseline full-data GRF cuts: %s\n",
  if (is.null(fs_baseline$grf_cuts)) "(none)" else
    paste(fs_baseline$grf_cuts, collapse = " | ")
))
```

## With `conf_force = c("er <= 0")`

```{r fs-forced}
#| message: false
fs_forced <- do.call(
  forestsearch,
  c(list(
    df.analysis      = df,
    confounders.name = confounders.name,
    outcome.name     = outcome.name,
    event.name       = event.name,
    treat.name       = treat.name,
    id.name          = id.name,
    conf_force       = c("er <= 0"),
    parallel_args    = list(plan         = "multisession",
                            workers      = n_workers,
                            show_message = FALSE)
  ), fs_args_common)
)
plan("sequential")

cat(sprintf(
  "Forced full-data sg.harm: %s\n",
  if (is.null(fs_forced$sg.harm)) "(none)" else
    paste(fs_forced$sg.harm, collapse = " & ")
))
cat(sprintf(
  "Forced full-data GRF cuts: %s\n",
  if (is.null(fs_forced$grf_cuts)) "(none)" else
    paste(fs_forced$grf_cuts, collapse = " | ")
))
```


# 10-fold CV: baseline {#sec-tenfold-baseline}

```{r tenfold-baseline}
#| message: false
.t0 <- proc.time()
tenfold_baseline <- forestsearch_tenfold(
  fs.est        = fs_baseline,
  sims          = Ksims_cv,
  Kfolds        = Kfolds_cv,
  details       = FALSE,
  keep_resCV    = FALSE,
  parallel_args = list(plan         = "multisession",
                       workers      = n_workers,
                       show_message = FALSE)
)
plan("sequential")
.baseline_seconds <- as.numeric((proc.time() - .t0)["elapsed"])
sprintf("Baseline 10-fold CV (%d sims x %d folds) finished in %.1f s",
        Ksims_cv, Kfolds_cv, .baseline_seconds)
```


# 10-fold CV: forced cut {#sec-tenfold-forced}

```{r tenfold-forced}
#| message: false
.t0 <- proc.time()
tenfold_forced <- forestsearch_tenfold(
  fs.est        = fs_forced,
  sims          = Ksims_cv,
  Kfolds        = Kfolds_cv,
  details       = FALSE,
  keep_resCV    = FALSE,
  parallel_args = list(plan         = "multisession",
                       workers      = n_workers,
                       show_message = FALSE)
)
plan("sequential")
.forced_seconds <- as.numeric((proc.time() - .t0)["elapsed"])
sprintf("Forced 10-fold CV (%d sims x %d folds) finished in %.1f s",
        Ksims_cv, Kfolds_cv, .forced_seconds)
```


# Side-by-side comparison {#sec-compare}

## Aggregate agreement and subgroup-finding metrics

```{r compare-metrics}
.metric_names <- c("sens_H", "sens_Hc", "ppv_H", "ppv_Hc")
.find_names   <- c("Any", "Exact", "At least 1",
                   "Cov1", "Cov2",
                   "Cov 1 & 2", "Cov1 exact", "Cov2 exact")

.compare <- rbind(
  data.frame(
    metric   = .metric_names,
    Baseline = unname(tenfold_baseline$sens_summary[.metric_names]),
    Forced   = unname(tenfold_forced$sens_summary[.metric_names]),
    stringsAsFactors = FALSE
  ),
  data.frame(
    metric   = .find_names,
    Baseline = unname(tenfold_baseline$find_summary[.find_names]),
    Forced   = unname(tenfold_forced$find_summary[.find_names]),
    stringsAsFactors = FALSE
  )
)
.compare$Delta <- .compare$Forced - .compare$Baseline

gt::gt(.compare) |>
  gt::tab_header(
    title    = "10-fold CV agreement metrics: baseline vs conf_force('er <= 0')",
    subtitle = sprintf(
      "GBSG, N = %d, %d sims x %d folds = %d sim x fold pairs",
      nrow(df), Ksims_cv, Kfolds_cv, Ksims_cv * Kfolds_cv
    )
  ) |>
  gt::fmt_number(columns = c("Baseline", "Forced", "Delta"),
                 decimals = 3) |>
  gt::cols_label(
    metric   = "Metric",
    Baseline = "Baseline (median)",
    Forced   = "conf_force (median)",
    Delta    = "Δ (Forced − Baseline)"
  ) |>
  gt::tab_source_note(paste(
    "Positive Δ on sens_H, ppv_H, and Exact supports the hypothesis",
    "that GRF drifts on some 10-fold training sets (dropping ~10%",
    "of subjects), and that forcing 'er <= 0' as a candidate",
    "recovers stability.  Near-zero Δ would indicate GRF is stable",
    "even under 10-fold perturbations."
  ))
```

## Per-fold identified-subgroup distribution

Uses the `fold_summary` slot to show the empirical distribution of
training-FS subgroups across all sim × fold pairs.

```{r compare-fold-dist}
.fold_dist_10fold <- function(tenfold_out, label) {
  fsum <- tenfold_out$fold_summary
  fold_sgs <- ifelse(
    is.na(fsum$sg1),
    "(no subgroup)",
    ifelse(
      is.na(fsum$sg2),
      fsum$sg1,
      paste(fsum$sg1, fsum$sg2, sep = " & ")
    )
  )
  tab <- as.data.frame(table(sg = fold_sgs), stringsAsFactors = FALSE)
  tab <- tab[order(-tab$Freq), , drop = FALSE]
  names(tab) <- c("Training-FS subgroup", "n folds")
  tab[["% folds"]] <- round(
    100 * tab[["n folds"]] / nrow(fsum), 2
  )
  tab$Config <- label
  tab
}

.dist_base  <- .fold_dist_10fold(tenfold_baseline, "Baseline")
.dist_force <- .fold_dist_10fold(tenfold_forced,   "Forced")

gt::gt(.dist_base[, c("Training-FS subgroup", "n folds", "% folds")]) |>
  gt::tab_header(
    title = sprintf("Baseline 10-fold: subgroups across %d sim × fold pairs",
                    nrow(tenfold_baseline$fold_summary))
  )

gt::gt(.dist_force[, c("Training-FS subgroup", "n folds", "% folds")]) |>
  gt::tab_header(
    title = sprintf("Forced conf_force('er <= 0') 10-fold: subgroups across %d pairs",
                    nrow(tenfold_forced$fold_summary))
  )
```


# GRF cut distribution across 10-fold training sets {#sec-grf-cuts}

Uses the newly-preserved `grf_cuts` column in `fold_summary` to
tabulate what GRF returned on each of the `r Ksims_cv * Kfolds_cv`
10-fold training sets.  This is the direct measurement of GRF's
cut-choice distribution at the 90%-of-N training-set size.

## `er`-cut values chosen by GRF --- baseline

```{r grf-cuts-baseline}
.extract_er_cut <- function(cuts_str) {
  if (is.na(cuts_str)) return("(no GRF cut)")
  parts <- strsplit(cuts_str, " \\| ", fixed = FALSE)[[1]]
  er_parts <- parts[grepl("^er\\b", parts)]
  if (length(er_parts) == 0L) return("(no er cut)")
  paste(er_parts, collapse = "; ")
}

fsum_base <- tenfold_baseline$fold_summary
fsum_base$er_cut <- vapply(
  fsum_base$grf_cuts, .extract_er_cut, character(1)
)

.er_freq_base <- as.data.frame(
  table(er_cut = fsum_base$er_cut),
  stringsAsFactors = FALSE
)
names(.er_freq_base) <- c("GRF er-cut", "n folds")
.er_freq_base <- .er_freq_base[order(-.er_freq_base[["n folds"]]), ,
                                drop = FALSE]
.er_freq_base[["% folds"]] <- round(
  100 * .er_freq_base[["n folds"]] / nrow(fsum_base), 2
)

gt::gt(.er_freq_base) |>
  gt::tab_header(
    title    = "Baseline 10-fold: GRF `er`-cut distribution",
    subtitle = sprintf(
      "Across %d sim × fold training sets (~90%% of N each)",
      nrow(fsum_base)
    )
  ) |>
  gt::tab_source_note(paste(
    "If one row dominates at near-100%, GRF is stable even under",
    "10-fold perturbations.  Multiple rows with non-trivial",
    "frequency would indicate GRF drift --- and in that case, folds",
    "with non-zero cuts are the ones where conf_force would be",
    "expected to help."
  ))
```

## `er`-cut values chosen by GRF --- forced-cut run

Sanity check: because `conf_force = c("er <= 0")` does not affect
GRF itself (it injects the cut *after* GRF runs, into `get_FSdata`),
the GRF cut distribution here should match the baseline within
Monte-Carlo noise.  If it doesn't, something unexpected is going
on with the parameter propagation.

```{r grf-cuts-forced}
fsum_forc <- tenfold_forced$fold_summary
fsum_forc$er_cut <- vapply(
  fsum_forc$grf_cuts, .extract_er_cut, character(1)
)

.er_freq_forc <- as.data.frame(
  table(er_cut = fsum_forc$er_cut),
  stringsAsFactors = FALSE
)
names(.er_freq_forc) <- c("GRF er-cut", "n folds")
.er_freq_forc <- .er_freq_forc[order(-.er_freq_forc[["n folds"]]), ,
                                drop = FALSE]
.er_freq_forc[["% folds"]] <- round(
  100 * .er_freq_forc[["n folds"]] / nrow(fsum_forc), 2
)

gt::gt(.er_freq_forc) |>
  gt::tab_header(
    title    = "Forced 10-fold: GRF `er`-cut distribution (should match baseline)",
    subtitle = sprintf(
      "Across %d sim × fold training sets",
      nrow(fsum_forc)
    )
  )
```

## Cross-tab: GRF `er`-cut vs identified subgroup (baseline)

Shows, for each GRF cut value, which final subgroup the algorithm
identified.  Reveals whether non-`er <= 0` GRF cuts actually
translated into non-`{er <= 0}` identifications, or whether the
consistency stage was able to recover the canonical subgroup from
a different GRF starting point.

```{r xtab-baseline}
fsum_base$sg_full <- ifelse(
  is.na(fsum_base$sg1),
  "(no subgroup)",
  ifelse(
    is.na(fsum_base$sg2),
    fsum_base$sg1,
    paste(fsum_base$sg1, fsum_base$sg2, sep = " & ")
  )
)

.xtab_base <- as.data.frame(
  table(
    `GRF er-cut`  = fsum_base$er_cut,
    `Identified subgroup` = fsum_base$sg_full
  ),
  stringsAsFactors = FALSE
)
.xtab_base <- .xtab_base[.xtab_base$Freq > 0, , drop = FALSE]
names(.xtab_base)[names(.xtab_base) == "Freq"] <- "n folds"

gt::gt(.xtab_base) |>
  gt::tab_header(
    title = "Baseline: GRF `er`-cut vs identified subgroup"
  )
```

## Cross-tab: GRF `er`-cut vs identified subgroup (forced)

```{r xtab-forced}
fsum_forc$sg_full <- ifelse(
  is.na(fsum_forc$sg1),
  "(no subgroup)",
  ifelse(
    is.na(fsum_forc$sg2),
    fsum_forc$sg1,
    paste(fsum_forc$sg1, fsum_forc$sg2, sep = " & ")
  )
)

.xtab_forc <- as.data.frame(
  table(
    `GRF er-cut`  = fsum_forc$er_cut,
    `Identified subgroup` = fsum_forc$sg_full
  ),
  stringsAsFactors = FALSE
)
.xtab_forc <- .xtab_forc[.xtab_forc$Freq > 0, , drop = FALSE]
names(.xtab_forc)[names(.xtab_forc) == "Freq"] <- "n folds"

gt::gt(.xtab_forc) |>
  gt::tab_header(
    title    = "Forced: GRF `er`-cut vs identified subgroup",
    subtitle = "Rows where GRF drifted but subgroup is still {er <= 0} are where conf_force rescued the result"
  )
```


# Summary checkpoint {#sec-summary}

```{r summary}
# Fraction of folds where GRF chose er <= 0
.frac_grf_er0_base <- mean(fsum_base$er_cut == "er <= 0", na.rm = TRUE)
.frac_grf_er0_forc <- mean(fsum_forc$er_cut == "er <= 0", na.rm = TRUE)

# Fraction of folds where final identified subgroup is {er <= 0}
.frac_sg_er0_base <- mean(fsum_base$sg_full == "{er <= 0}", na.rm = TRUE)
.frac_sg_er0_forc <- mean(fsum_forc$sg_full == "{er <= 0}", na.rm = TRUE)

.summary_df <- data.frame(
  Check = c(
    "Baseline: fraction of folds where GRF returned er <= 0",
    "Baseline: fraction of folds where identified sg is {er <= 0}",
    "Forced:   fraction of folds where GRF returned er <= 0",
    "Forced:   fraction of folds where identified sg is {er <= 0}",
    "GRF drift visible at 10-fold? (baseline GRF er <= 0 < 0.99)",
    "Does conf_force recover sg consistency where GRF drifted?"
  ),
  Value = c(
    sprintf("%.3f", .frac_grf_er0_base),
    sprintf("%.3f", .frac_sg_er0_base),
    sprintf("%.3f", .frac_grf_er0_forc),
    sprintf("%.3f", .frac_sg_er0_forc),
    ifelse(.frac_grf_er0_base < 0.99, "YES", "NO"),
    ifelse(.frac_sg_er0_forc - .frac_sg_er0_base > 0.02, "YES", "NO / N/A")
  ),
  stringsAsFactors = FALSE
)

gt::gt(.summary_df) |>
  gt::tab_header(title = "Key 10-fold CV summary figures")
```


# Session info {#sec-session}

```{r session}
sessionInfo()
```
