---
title: "Smoke test — Cox/GLM consistency `resample` fallback alignment"
format:
  html:
    toc: true
    code-fold: false
bibliography: []
---

```{r setup, include=FALSE}
# warning = TRUE so any package-emitted warning surfaces in the render; the
# runner additionally captures per-call warnings and reports them explicitly.
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = TRUE)
```

## Purpose

`consistency_method = "resample"` computes the split-consistency rate from a
closed form. When that closed form is inestimable for a candidate (it returns
`rate_closed = NA`), the evaluator is supposed to **fall back to literal
splitting** rather than dropping the candidate.

Previously the two families behaved differently on that NA branch: the GLM path
fell through to splitting, while the Cox path returned `NULL` (dropping the
candidate). This test confirms they are now **aligned** — a candidate whose
closed-form rate is `NA` is retained via splitting for *both* Cox and GLM, and
the fallback rate matches what native `consistency_method = "split"` produces.

The trigger is made deterministic by temporarily replacing
`consistency_resample()` with a stub that returns `NA`, which forces every
candidate down the fallback path. GRF/LASSO/DINA screening are disabled (as in
the real applications' main runs), so only default candidate cuts are used.

```{r libs}
library(forestsearch)
library(survival)

SEED <- 8316951L
```

## Synthetic data with a clear harm subgroup

Both generators place a strong *harm* signal (treatment is detrimental) inside
`x1 > 0.5 & x2 > 0.5` and a benefit elsewhere, so the consistency screen should
identify that subgroup robustly under every method.

```{r data-gen}
make_cox <- function(n = 800, seed = 1) {
  set.seed(seed)
  x1 <- runif(n); x2 <- runif(n)
  treat <- rbinom(n, 1, 0.5)
  harm  <- as.integer(x1 > 0.5 & x2 > 0.5)          # harm subgroup
  # log-HR of treatment: +1.1 (harm) inside, -0.5 (benefit) outside
  lhr   <- ifelse(harm == 1, 1.1, -0.5) * treat
  rate  <- 0.15 * exp(lhr)
  tt    <- rexp(n, rate)
  cens  <- pmin(rexp(n, 0.05), 24)                  # admin + random censoring
  time  <- pmin(tt, cens)
  event <- as.integer(tt <= cens)
  data.frame(id = seq_len(n), time = time, event = event,
             treat = treat, x1 = x1, x2 = x2)
}

make_binary <- function(n = 800, seed = 2) {
  set.seed(seed)
  x1 <- runif(n); x2 <- runif(n)
  treat <- rbinom(n, 1, 0.5)
  harm  <- as.integer(x1 > 0.5 & x2 > 0.5)
  # log-OR of treatment on the adverse outcome: +1.3 (harm) inside, -0.6 outside
  lodds <- -0.2 + ifelse(harm == 1, 1.3, -0.6) * treat
  y     <- rbinom(n, 1, plogis(lodds))              # 1 = adverse (no benefit)
  data.frame(id = seq_len(n), y = y, treat = treat, x1 = x1, x2 = x2)
}

cox_df <- make_cox()
bin_df <- make_binary()

cat("Cox: n =", nrow(cox_df), "| events =", sum(cox_df$event),
    "| harm-subgroup n =", sum(cox_df$x1 > 0.5 & cox_df$x2 > 0.5), "\n")
cat("Binary: n =", nrow(bin_df), "| adverse =", sum(bin_df$y),
    "| harm-subgroup n =", sum(bin_df$x1 > 0.5 & bin_df$x2 > 0.5), "\n")
```

## Fallback-forcing stub

`with_resample_NA()` swaps `forestsearch:::consistency_resample()` for a stub
returning `rate_closed = NA` for the duration of a single expression, then
restores the original. Under `consistency_method = "resample"`, this forces
every candidate onto the NA→split fallback.

```{r stub}
with_resample_NA <- function(expr) {
  orig <- forestsearch::consistency_resample
  stub <- function(...) list(rate_closed = NA_real_, rate_mc = NA_real_,
                             beta_hat = NA_real_, sigma_D = NA_real_)
  assignInNamespace("consistency_resample", stub, ns = "forestsearch")
  on.exit(assignInNamespace("consistency_resample", orig, ns = "forestsearch"),
          add = TRUE)
  force(expr)
}
```

## Runner

Runs `forestsearch()` with GRF/LASSO/DINA off and the two-stage consistency
engine (the path the applications use), and returns the identified subgroup plus
its consistency proportion. `force_na = TRUE` wraps the call in the stub.

```{r runner}
run_fs <- function(df, family, method, force_na = FALSE) {
  common <- list(
    df.analysis      = df,
    treat.name       = "treat",
    id.name          = "id",
    confounders.name = c("x1", "x2"),
    is.RCT           = TRUE,
    use_lasso        = FALSE,
    use_grf          = FALSE,
    use_dina         = FALSE,
    subgroup_method  = "consistency",
    consistency_method = method,
    use_twostage     = TRUE,
    maxk             = 2L,
    n.min            = 50L,
    d0.min           = 5L,
    d1.min           = 5L,
    hr.threshold     = 1.0,
    hr.consistency   = 1.0,
    pconsistency.threshold = 0.80,
    seedit           = SEED,
    details          = FALSE,
    mr_inference      = FALSE
  )
  args <- if (family == "cox") {
    c(common, list(outcome.name = "time", event.name = "event",
                   outcome_type = "survival", effect_measure = "HR"))
  } else {
    c(common, list(outcome.name = "y", event.name = "y",
                   outcome_type = "binary", effect_measure = "OR"))
  }
  call_fs <- function() do.call(forestsearch, args)
  warns <- character(0)
  fs <- tryCatch(
    withCallingHandlers(
      if (force_na) with_resample_NA(call_fs()) else call_fs(),
      warning = function(w) {
        warns <<- c(warns, conditionMessage(w))
        invokeRestart("muffleWarning")
      }
    ),
    error = function(e) list(.error = conditionMessage(e))
  )
  if (!is.null(fs$.error)) {
    return(list(found = NA, subgroup = paste("ERROR:", fs$.error),
                pcons = NA_real_, result = NULL, harm_id = NULL, warnings = warns))
  }
  sg <- fs$sg.harm
  found <- !is.null(sg)
  # out_sg is a LIST: $result (candidate table, top row = selected) + pareto_frontier
  # + sg.harm/sg.harm_label/df_flag/sg.harm.id.  Pcons lives in out_sg$result.
  os <- tryCatch(fs$grp.consistency$out_sg, error = function(e) NULL)
  result_tab <- if (!is.null(os) && !is.null(os$result)) as.data.frame(os$result) else NULL
  pc <- if (!is.null(result_tab) && "Pcons" %in% names(result_tab)) {
    as.numeric(result_tab$Pcons[1])
  } else NA_real_
  # per-subject membership indicator (length nrow(df)) -- strongest discriminator
  harm_id <- tryCatch(fs$grp.consistency$sg.harm.id, error = function(e) NULL)
  list(found = found,
       subgroup = if (found) paste(deparse(sg), collapse = " ") else "(none)",
       pcons = pc, result = result_tab, harm_id = harm_id, warnings = warns)
}
```

## Run all six scenarios

```{r run-all}
scenarios <- expand.grid(
  family = c("cox", "binary"),
  mode   = c("resample (normal)", "resample (forced NA -> fallback)", "split (native)"),
  stringsAsFactors = FALSE
)

RUN <- list()   # full run objects, keyed by "family|mode", for the assertions
rows <- vector("list", nrow(scenarios))
for (i in seq_len(nrow(scenarios))) {
  fam  <- scenarios$family[i]
  mode <- scenarios$mode[i]
  method   <- if (grepl("split", mode)) "split" else "resample"
  force_na <- grepl("forced NA", mode)
  r <- run_fs(if (fam == "cox") cox_df else bin_df, fam, method, force_na)
  RUN[[paste(fam, mode, sep = "|")]] <- r
  rows[[i]] <- data.frame(
    family = fam, mode = mode, found = r$found,
    pcons = round(r$pcons, 3), n_pass = if (is.null(r$result)) 0L else nrow(r$result),
    n_warn = length(r$warnings), subgroup = r$subgroup, stringsAsFactors = FALSE)
}
res <- do.call(rbind, rows)
knitr::kable(res, caption = "Identified subgroup by family x consistency mode (n_pass = candidates passing consistency; n_warn = warnings captured)")
```

## Assertions

The fix is confirmed when, for **each** family:

1. `resample (normal)` finds a subgroup — the closed-form path is intact.
2. `resample (forced NA -> fallback)` **also** finds a subgroup — the NA branch
   falls back to splitting instead of dropping the candidate. *(Pre-fix, Cox
   would return no subgroup here; GLM already passed.)*
3. `split (native)` finds a subgroup.
4. The forced-NA fallback **reproduces** native splitting: the *full* candidate
   table (`out_sg$result`, all rows/columns) **and** the per-subject membership
   vector are identical between the two runs. This discriminates even when
   `Pcons` saturates at 1.0 — unlike a top-row `Pcons` comparison, which is
   vacuous under saturation.

```{r assertions}
check_family <- function(fam) {
  rn <- RUN[[paste(fam, "resample (normal)", sep = "|")]]
  rf <- RUN[[paste(fam, "resample (forced NA -> fallback)", sep = "|")]]
  sp <- RUN[[paste(fam, "split (native)", sep = "|")]]

  a1 <- isTRUE(rn$found)                                   # closed-form path intact
  a2 <- isTRUE(rf$found)                                   # KEY: fallback fires (not dropped)
  a3 <- isTRUE(sp$found)                                   # native split works

  # a4: fallback genuinely reproduces native splitting. Require BOTH the full
  # candidate table and the per-subject membership to be non-NULL and identical.
  # (Guards against a vacuous pass: e.g. two NULLs, or saturated Pcons matching
  #  by coincidence -- neither can satisfy this.)
  tab_match <- !is.null(rf$result) && !is.null(sp$result) &&
               isTRUE(all.equal(rf$result, sp$result))
  id_match  <- !is.null(rf$harm_id) && !is.null(sp$harm_id) &&
               length(rf$harm_id) == length(sp$harm_id) &&
               identical(rf$harm_id, sp$harm_id)
  a4 <- tab_match && id_match

  data.frame(
    family = fam,
    `1_resample_found`  = a1,
    `2_fallback_fires`  = a2,
    `3_split_found`     = a3,
    PASS = all(a1, a2, a3),                                # core alignment gate
    `4_fallback_reproduces_split` = a4,                    # table + membership identical
    tab_match = tab_match, id_match = id_match,
    check.names = FALSE
  )
}

report <- rbind(check_family("cox"), check_family("binary"))
knitr::kable(report, caption = "Fallback-alignment assertions (PASS gates on 1-3; col 4 = rigorous reproduction check)")

## Surface every warning captured across all runs -- this is the codebase check
all_warn <- unlist(lapply(RUN, function(r) r$warnings), use.names = FALSE)
cat("\n--- Warnings captured across all six runs ---\n")
if (length(all_warn) == 0) {
  cat("None. No package warning was emitted by any Cox or GLM run.\n")
} else {
  for (w in unique(all_warn)) cat(" * ", w, "\n", sep = "")
  cat("(", length(all_warn), " total,", length(unique(all_warn)), "unique) -- review these; they are not test failures per se but indicate package-level conditions.\n")
}

cat("\n--- Verdict ---\n")
if (all(report$PASS)) {
  cat("\u2705 CORE PASS - Cox and GLM both retain the candidate on an NA closed-form",
      "(fall back to splitting instead of dropping).\n")
  if (all(report$`4_fallback_reproduces_split`)) {
    cat("\u2705 REPRODUCTION PASS - forced-NA fallback reproduces native splitting",
        "exactly (full candidate table + per-subject membership identical) for both families.\n")
  } else {
    cat("\u26a0\ufe0f  REPRODUCTION CHECK FAILED for a family (col 4 = FALSE):",
        "the fallback did NOT reproduce native splitting. Inspect tab_match / id_match",
        "columns -- this would be a genuine discrepancy worth investigating, not a cosmetic one.\n")
  }
} else {
  cat("\u274c CORE FAILURE - see the assertion table above.\n")
  stop("Consistency fallback alignment smoke test failed (core assertions 1-3).")
}
```

## Notes

- This exercises the **two-stage** evaluator (`use_twostage = TRUE`), which is
  the path both applications use and where the Cox-vs-GLM asymmetry lived. To
  additionally cover the single-stage evaluator, set `use_twostage = FALSE` in
  `run_fs()`'s `common` list and re-render.
- The stub forces *every* candidate through the fallback, which is the strongest
  form of the test. In production the fallback fires only for candidates whose
  closed form is genuinely inestimable (e.g. a non-convergent fit).
- If assertion 4 shows a benign mismatch, check that `seedit` is identical across
  the two calls (it is, via `SEED`) — the split RNG must match for the fallback
  and native-split rates to be bit-identical.
