---
title: "forestsearch — no-subgroup & robustness smoke suite"
format: html
bibliography: []
execute:
  warning: false
  message: false
  error: true
---

## What this is

A smoke suite that exercises the CV pipeline **end-to-end on real data** to stress the
robustness of the recent `feature/glm-extension` revisions:

1. **Loud failures** — a genuine fold-refit *error* must be fatal and diagnostic
   (`.collect_cv_results` re-raises rather than silently falling back to ITT).
2. **No-subgroup completion** — when the primary fit identifies no subgroup
   (`df.est == NULL`, `sg.harm == NULL`), the CV entry points must complete as a
   valid 0%-identification result instead of crashing. This is checked **per
   outcome family**, because the family-specific placeholder-event handling is
   exactly what regressed for continuous/count.
3. **Happy-path regression** — subgroup-found runs still complete and return sane
   structure.

Every check is wrapped in `tryCatch` and prints `[PASS] / [FAIL] / [SKIP] / [INFO]`,
with a summary table and an overall verdict at the end. Nothing here mutates your
package or repo.

::: {.callout-important}
## Read before running

- **Written without execution.** I could not run R against the built package, so
  treat first-run hiccups as expected. The **survival / GBSG** row is the
  high-confidence anchor. The **binary / continuous / count** rows are best-effort:
  if one fails on an *argument* error (unknown `effect_measure`, an `event.name`
  the family doesn't want, a missing column), the fix is almost always a one-line
  edit in the **Config** block below — or point each family at your own
  `tests/testthat/helper-synthetic-dgm.R` (`.fs_args_for()` / `.make_*_data()`),
  which already encode the correct per-family arguments.
- **No-subgroup is forced via `pconsistency.threshold = 0.999`** — a near-impossible
  split-consistency bar that no candidate clears, *family-agnostically*. If a family
  still finds a subgroup under that, the no-subgroup checks `SKIP` (with a hint to
  push the threshold higher) rather than giving a false pass.
- **Sequential by design** (`plan = "sequential"`) so it runs fast and sidesteps the
  doFuture worker/install rule. Run **after** `devtools::install()` so the installed
  package matches the branch.
- **Count** is a commented stub (no standard count dataset here) — fill it from your
  own count DGM; the shape is shown at the bottom of the Config block.
:::

```{r setup}
library(forestsearch)
set.seed(8316951L)

SEQ <- list(plan = "sequential", workers = 1, show_message = FALSE)  # no parallel during smoke
```

## Config — EDIT THIS BLOCK if column names / measures differ

```{r config}
# Per family: how to load the dataset, the fixed forestsearch() args, and two
# threshold sets — one that (loosely) permits a subgroup, one that forbids all.
FAMILIES <- list(

  survival = list(
    label = "survival/Cox (GBSG)",
    load  = function() {
      d <- survival::gbsg           # N = 686; hormon = treatment, rfstime/status = outcome
      d$id <- seq_len(nrow(d))
      d
    },
    args = list(
      outcome.name     = "rfstime",
      event.name       = "status",
      treat.name       = "hormon",
      id.name          = "id",
      confounders.name = c("age", "meno", "size", "grade", "nodes", "pgr", "er"),
      outcome_type     = "survival",
      est.scale        = "hr",
      sg_focus         = "hr"
    ),
    thr_found = list(hr.threshold = 1.05, pconsistency.threshold = 0.50),
    thr_nosg  = list(hr.threshold = 100,  pconsistency.threshold = 0.999)
  ),

  # ---- GLM rows: BEST-EFFORT. Confirm against your ACTG175 analysis / helpers. ----

  binary = list(
    label = "binary/OR (ACTG175)",
    load  = function() {
      if (!requireNamespace("speff2trial", quietly = TRUE))
        stop("package 'speff2trial' not installed")
      d <- speff2trial::ACTG175      # treat = 0/1; cens = 0/1 outcome
      d$id <- seq_len(nrow(d))
      d
    },
    args = list(
      outcome.name     = "cens",     # 0/1 response
      # event.name deliberately omitted -> defaults to "event" (absent);
      #   the binary path does not use a survival event column.
      treat.name       = "treat",
      id.name          = "id",
      confounders.name = c("age", "wtkg", "karnof", "cd40", "cd80"),
      outcome_type     = "binary",
      effect_measure   = "OR"
    ),
    thr_found = list(effect.threshold = 1.05, pconsistency.threshold = 0.50),
    thr_nosg  = list(effect.threshold = 100,  pconsistency.threshold = 0.999)
  ),

  continuous = list(
    label = "continuous/MD (ACTG175 CD4)",
    load  = function() {
      if (!requireNamespace("speff2trial", quietly = TRUE))
        stop("package 'speff2trial' not installed")
      d <- speff2trial::ACTG175
      d$id <- seq_len(nrow(d))
      d
    },
    args = list(
      outcome.name     = "cd420",    # CD4 at 20 weeks (continuous)
      # event.name omitted -> forestsearch synthesizes ".event_placeholder" (= 1L);
      #   THIS is the path the Layer-B fix must handle on the no-subgroup fallback.
      treat.name       = "treat",
      id.name          = "id",
      confounders.name = c("age", "wtkg", "karnof", "cd40", "cd80"),
      outcome_type     = "continuous",
      effect_measure   = "MD"
    ),
    thr_found = list(effect.threshold = 1.0,  pconsistency.threshold = 0.50),
    thr_nosg  = list(effect.threshold = 1e6,  pconsistency.threshold = 0.999)
  )

  # ,count = list(
  #   label = "count/IRR (your count DGM)",
  #   load  = function() { <your count data frame with an offset column> },
  #   args  = list(outcome.name = "<count>", treat.name = "treat", id.name = "id",
  #                confounders.name = c(...), outcome_type = "count",
  #                effect_measure = "IRR", offset.name = "<log-exposure>"),
  #   thr_found = list(effect.threshold = 1.05, pconsistency.threshold = 0.50),
  #   thr_nosg  = list(effect.threshold = 100,  pconsistency.threshold = 0.999)
  # )
)
```

## Harness

```{r harness}
.smoke <- new.env()
.smoke$rows <- list()

record <- function(family, scenario, status, detail = "") {
  .smoke$rows[[length(.smoke$rows) + 1L]] <-
    data.frame(family = family, scenario = scenario, status = status,
               detail = detail, stringsAsFactors = FALSE)
  cat(sprintf("[%-4s] %-26s | %-34s %s\n",
              status, family, scenario,
              if (nzchar(detail)) paste0("- ", detail) else ""))
  invisible(NULL)
}

# Fit forestsearch() for a given family/mode; returns the fs object OR the error.
fit_fs <- function(cfg, mode) {
  df  <- cfg$load()
  thr <- if (identical(mode, "nosg")) cfg$thr_nosg else cfg$thr_found
  args <- c(list(df.analysis = df, parallel_args = SEQ, seedit = 8316951L),
            cfg$args, thr)
  do.call(forestsearch, args)
}

is_no_subgroup <- function(fs) is.null(fs$sg.harm) && is.null(fs$df.est)
```

## Per-family checks

```{r per-family}
for (fam in names(FAMILIES)) {
  cfg <- FAMILIES[[fam]]
  lab <- cfg$label

  # -- dataset availability --
  df0 <- tryCatch(cfg$load(), error = function(e) e)
  if (inherits(df0, "error")) {
    record(lab, "dataset loads", "SKIP", conditionMessage(df0))
    next
  }

  # ============================ NO-SUBGROUP BRANCH ============================
  fs_nosg <- tryCatch(fit_fs(cfg, "nosg"), error = function(e) e)

  if (inherits(fs_nosg, "error")) {
    record(lab, "fit, no-subgroup config", "FAIL", conditionMessage(fs_nosg))
  } else if (!is_no_subgroup(fs_nosg)) {
    record(lab, "fit, no-subgroup config", "SKIP",
           "still found a subgroup; push pconsistency.threshold toward 1")
  } else {
    record(lab, "fit, no-subgroup config", "PASS", "sg.harm/df.est NULL as intended")

    # ***** crown jewel: tenfold must COMPLETE on a no-subgroup fit *****
    tf <- tryCatch(
      forestsearch_tenfold(fs_nosg, sims = 3L, Kfolds = 3L,
                           details = FALSE, parallel_args = SEQ),
      error = function(e) e)
    if (inherits(tf, "error")) {
      record(lab, "tenfold completes (no-subgroup)", "FAIL", conditionMessage(tf))
    } else {
      all_itt <- tryCatch(all(tf$fold_summary$any_found == 0), error = function(e) NA)
      record(lab, "tenfold completes (no-subgroup)", "PASS",
             sprintf("class=%s; all any_found==0: %s",
                     paste(class(tf), collapse = "/"), all_itt))
    }

    # Kfold on a no-subgroup fit: Option 1 -> completes; else must be a CLEAR
    # message, never the cryptic NULL-attribute / undefined-columns crash.
    kf <- tryCatch(
      forestsearch_Kfold(fs_nosg, Kfolds = 5L, parallel_args = SEQ),
      error = function(e) e)
    if (inherits(kf, "error")) {
      msg     <- conditionMessage(kf)
      cryptic <- grepl("attempt to set an attribute on NULL|undefined columns selected", msg)
      record(lab, "Kfold on no-subgroup", if (cryptic) "FAIL" else "INFO",
             if (cryptic) paste("CRYPTIC crash:", msg)
             else paste("errored with a message (Option 2 / clear reject?):", msg))
    } else {
      record(lab, "Kfold on no-subgroup", "PASS", "completed (0%-identification result)")
    }
  }

  # ============================== HAPPY PATH ================================
  fs_found <- tryCatch(fit_fs(cfg, "found"), error = function(e) e)
  if (inherits(fs_found, "error")) {
    record(lab, "fit, subgroup-permitting config", "FAIL", conditionMessage(fs_found))
  } else if (is.null(fs_found$sg.harm)) {
    record(lab, "happy-path (subgroup found)", "SKIP",
           "permissive config found nothing; lower pconsistency.threshold to exercise")
  } else {
    tf2 <- tryCatch(
      forestsearch_tenfold(fs_found, sims = 3L, Kfolds = 3L,
                           details = FALSE, parallel_args = SEQ),
      error = function(e) e)
    record(lab, "happy-path tenfold (subgroup present)",
           if (inherits(tf2, "error")) "FAIL" else "PASS",
           if (inherits(tf2, "error")) conditionMessage(tf2)
           else "subgroup-found CV completed")
  }
}
```

## Cross-cutting: fatal-on-any (loud-failures guard)

Directly checks that `.collect_cv_results` re-raises a captured error instead of
swallowing it — the core of the loud-failures fix, verified deterministically so it
does not depend on provoking a real refit crash.

```{r collector-fatal}
collector_fatal_check <- function() {
  fn <- tryCatch(get(".collect_cv_results", envir = asNamespace("forestsearch")),
                 error = function(e) NULL)
  if (is.null(fn)) {
    record("core", "collector present", "SKIP",
           ".collect_cv_results not found (renamed / not exported internally?)")
    return(invisible())
  }
  simulated <- tryCatch(stop("SIMULATED refit failure on fold X of Y"),
                        error = function(e) e)
  mixed_list <- list(data.frame(a = 1), simulated, data.frame(a = 2))
  out <- tryCatch(fn(mixed_list), error = function(e) conditionMessage(e))
  if (is.character(out) && grepl("SIMULATED refit failure", out)) {
    record("core", "fatal-on-any: collector re-raises", "PASS",
           "captured error propagated (not swallowed to a fallback)")
  } else {
    record("core", "fatal-on-any: collector re-raises", "FAIL",
           "returned a value instead of re-raising -> silent-failure regression")
  }
}
collector_fatal_check()
```

## Summary

```{r summary}
summ <- do.call(rbind, .smoke$rows)
knitr::kable(summ, row.names = FALSE)

n <- function(s) sum(summ$status == s)
cat(sprintf("\nOVERALL: %d PASS | %d FAIL | %d SKIP | %d INFO\n",
            n("PASS"), n("FAIL"), n("SKIP"), n("INFO")))
if (n("FAIL") == 0L) {
  cat("No robustness failures detected.\n")
} else {
  cat("Robustness FAILURES present — inspect the FAIL rows above.\n")
}

# SKIP/INFO are not failures: SKIP usually means a threshold needs tuning or a
# dataset/helper is unavailable; INFO on Kfold means it errored with a *clear*
# (non-cryptic) message, which is acceptable under the reject-with-message policy.
```
