---
title: "Testing `conf.cont_jcuts`: J-quantile cuts on the GBSG dataset"
author: "forestsearch development"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    code-fold: false
    code-tools: true
execute:
  warning: false
  message: false
---

## Overview

This document is a smoke test for the new `conf.cont_jcuts` argument
to `get_FSdata()`.  It exercises the feature on the GBSG breast-cancer
trial dataset across six scenarios:

1. **Baseline** — default cut behaviour, no override.
2. **Single-variable override** — `pgr` with J = 5 (5 binary cuts).
3. **Multi-variable override** — `pgr` with J = 5, `age` with J = 10.
4. **Empirical verification** — resolved cut values match the
   corresponding empirical quantiles.
5. **Coverage check** — proportion of patients flagged at the k-th
   cut tracks k/J.
6. **Validation errors** — bad inputs fire informative errors.

The feature replaces the default `{mean, median, Q1, Q3}` cut set with
J − 1 binary cuts at the (k/J)-th quantiles of the variable, but
**only** for variables explicitly named in `conf.cont_jcuts`.  All
other continuous covariates retain default behaviour.

## 1. Setup

```{r setup}
#| label: setup

# Load forestsearch with the conf.cont_jcuts feature.  Choose ONE of
# the following depending on your setup:

library(forestsearch)

# OR — active development without install:
# devtools::load_all(".")

library(survival)
library(knitr)

# ---- Test parameters --------------------------------------------------
.SEED     <- 8316951
.CONT_CUT <- 4L     # cont.cutoff for is.continuous()
.CONFS    <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
```

```{r prep-data}
#| label: prep-data

set.seed(.SEED)
df <- survival::gbsg
df$grade3 <- as.integer(df$grade == "3")

# Inspect continuous status using the same cont.cutoff that get_FSdata()
# will apply downstream.
n_unique  <- vapply(df[, .CONFS], function(x) length(unique(x)), integer(1))
cont_flag <- n_unique >= .CONT_CUT
kable(data.frame(
  variable    = .CONFS,
  n_unique    = unname(n_unique),
  continuous  = ifelse(cont_flag, "yes", "no")
), caption = "Confounder inventory in GBSG (n = 686)")
```

## 2. Baseline — default behaviour

No `conf.cont_jcuts` supplied.  Each continuous variable produces up
to 4 distinct expressions via `cut_var()` (mean / median / Q1 / Q3).
A redundant median fallback from the `conf.cont_Medcuts` block is
deduplicated against `cut_var()` output by `unique()` when the
literal values match (which is typical).

```{r baseline}
#| label: baseline

fs_default <- get_FSdata(
  df.analysis      = df,
  confounders.name = .CONFS,
  outcome.name     = "rfstime",
  event.name       = "status",
  use_lasso        = FALSE,
  use_grf          = FALSE,
  details          = FALSE
)

confs_default <- fs_default$confs
length(confs_default)

# Per-variable cut count.  The leading token of each expression is the
# variable name.
default_counts <- table(sub(" .*$", "", confs_default))
kable(as.data.frame(default_counts),
      col.names = c("variable", "n_cuts"),
      caption = "Default cut counts per variable")
```

## 3. Single-variable override — `pgr` with J = 5

Expectation: `pgr` produces J = 5 cuts at the 1/6, 2/6, 3/6, 4/6, 5/6
quantiles; all other continuous variables are untouched.

```{r jcuts-single}
#| label: jcuts-single

fs_pgr5 <- get_FSdata(
  df.analysis      = df,
  confounders.name = .CONFS,
  outcome.name     = "rfstime",
  event.name       = "status",
  conf.cont_jcuts  = list(pgr = 5),
  use_lasso        = FALSE,
  use_grf          = FALSE,
  details          = FALSE
)

# Cuts referencing pgr only
pgr_cuts <- fs_pgr5$confs[grepl("^pgr ", fs_pgr5$confs)]
pgr_cuts

# Number of pgr-cuts (4 expected) + comparison to other variables
pgr5_counts <- table(sub(" .*$", "", fs_pgr5$confs))
kable(as.data.frame(pgr5_counts),
      col.names = c("variable", "n_cuts"),
      caption = "Cut counts with conf.cont_jcuts = list(pgr = 5)")

stopifnot(sum(grepl("^pgr ", fs_pgr5$confs)) == 5L)
# Untouched variables retain the same cut count as in the baseline.
stopifnot(sum(grepl("^age ", fs_pgr5$confs))  ==
          sum(grepl("^age ", confs_default)))
stopifnot(sum(grepl("^size ", fs_pgr5$confs)) ==
          sum(grepl("^size ", confs_default)))
```

## 4. Multi-variable override — `list(pgr = 5, age = 10)`

`pgr` should produce 5 cuts; `age` should produce 10; the rest stay at
default.  The `details = TRUE` setting also exercises the new
"J-quantile cuts applied to:" diagnostic line.

```{r jcuts-multi}
#| label: jcuts-multi

fs_multi <- get_FSdata(
  df.analysis      = df,
  confounders.name = .CONFS,
  outcome.name     = "rfstime",
  event.name       = "status",
  conf.cont_jcuts  = list(pgr = 5, age = 10),
  use_lasso        = FALSE,
  use_grf          = FALSE,
  details          = TRUE
)

multi_counts <- table(sub(" .*$", "", fs_multi$confs))
kable(as.data.frame(multi_counts),
      col.names = c("variable", "n_cuts"),
      caption = "Cut counts with conf.cont_jcuts = list(pgr = 5, age = 10)")

stopifnot(sum(grepl("^pgr ", fs_multi$confs))  == 5L)    # J=5  -> 5 cuts
stopifnot(sum(grepl("^age ", fs_multi$confs))  == 10L)   # J=10 -> 10 cuts

# Untouched variables retain default counts
for (v in c("size", "nodes", "er")) {
  stopifnot(
    sum(grepl(paste0("^", v, " "), fs_multi$confs)) ==
    sum(grepl(paste0("^", v, " "), confs_default))
  )
}
```

## 5. Empirical verification

The literal cut values resolved by `process_conf_force_expr()` should
equal the empirical (k/(J+1))-th quantiles (rounded to 1 decimal).

```{r empirical}
#| label: empirical

# pgr cuts: J = 5  (placed at probs 1/6, 2/6, 3/6, 4/6, 5/6)
pgr_cut_vals <- sort(as.numeric(sub("pgr <= ", "",
                       fs_multi$confs[grepl("^pgr ", fs_multi$confs)])))
expected_pgr <- as.numeric(round(quantile(df$pgr, probs = (1:5) / 6,
                                          na.rm = TRUE), 1))
kable(data.frame(
  k             = 1:5,
  prob          = (1:5) / 6,
  cut_observed  = pgr_cut_vals,
  cut_expected  = expected_pgr,
  match         = abs(pgr_cut_vals - expected_pgr) < 1e-9
), caption = "pgr cut values vs. empirical (k/6)-quantiles")

# age cuts: J = 10  (placed at probs 1/11, 2/11, ..., 10/11)
age_cut_vals <- sort(as.numeric(sub("age <= ", "",
                       fs_multi$confs[grepl("^age ", fs_multi$confs)])))
expected_age <- as.numeric(round(quantile(df$age, probs = (1:10) / 11,
                                          na.rm = TRUE), 1))
kable(data.frame(
  k             = 1:10,
  prob          = (1:10) / 11,
  cut_observed  = age_cut_vals,
  cut_expected  = expected_age,
  match         = abs(age_cut_vals - expected_age) < 1e-9
), caption = "age cut values vs. empirical (k/11)-quantiles")

stopifnot(all(abs(pgr_cut_vals - expected_pgr) < 1e-9))
stopifnot(all(abs(age_cut_vals - expected_age) < 1e-9))
```

### Coverage check

Proportion of patients flagged by each cut should track k/J.  Small
deviations are expected because the rounding-to-1-decimal step in
`process_conf_force_expr()` and the discreteness of the data (ties)
shift exact equality.

```{r coverage}
#| label: coverage

cov_pgr <- vapply(seq_along(pgr_cut_vals),
                  function(k) mean(df$pgr <= pgr_cut_vals[k], na.rm = TRUE),
                  numeric(1))
cov_age <- vapply(seq_along(age_cut_vals),
                  function(k) mean(df$age <= age_cut_vals[k], na.rm = TRUE),
                  numeric(1))

kable(data.frame(
  variable = c(rep("pgr", 5), rep("age", 10)),
  k        = c(1:5, 1:10),
  target   = c((1:5) / 6, (1:10) / 11),
  observed = round(c(cov_pgr, cov_age), 3)
), caption = "Empirical coverage at each cut (target vs. observed)")
```

## 6. Validation errors

Each block below should produce an informative error.

```{r validation-bad-name}
#| label: validation-bad-name
#| error: true

# Variable name not in confounders.name
get_FSdata(df.analysis = df, confounders.name = .CONFS,
           outcome.name = "rfstime", event.name = "status",
           conf.cont_jcuts = list(NOPE = 5),
           use_lasso = FALSE, use_grf = FALSE, details = FALSE)
```

```{r validation-bad-J}
#| label: validation-bad-J
#| error: true

# J < 1 (not positive); J = 1 is now valid (1 cut at the median)
get_FSdata(df.analysis = df, confounders.name = .CONFS,
           outcome.name = "rfstime", event.name = "status",
           conf.cont_jcuts = list(pgr = 0),
           use_lasso = FALSE, use_grf = FALSE, details = FALSE)
```

```{r validation-conflict-default}
#| label: validation-conflict-default
#| error: true

# Conflict: same variable in both conf.cont_jcuts and defaultcut_names
get_FSdata(df.analysis = df, confounders.name = .CONFS,
           outcome.name = "rfstime", event.name = "status",
           conf.cont_jcuts  = list(pgr = 5),
           defaultcut_names = "pgr",
           use_lasso = FALSE, use_grf = FALSE, details = FALSE)
```

```{r validation-unnamed}
#| label: validation-unnamed
#| error: true

# Unnamed list
get_FSdata(df.analysis = df, confounders.name = .CONFS,
           outcome.name = "rfstime", event.name = "status",
           conf.cont_jcuts = list(5),
           use_lasso = FALSE, use_grf = FALSE, details = FALSE)
```

## 7. Categorical / non-continuous variable

Variables that don't pass the continuous check are silently skipped
(with a warning).  `meno` (binary) is in `confounders.name` but should
be skipped if mistakenly listed in `conf.cont_jcuts`.

```{r non-continuous}
#| label: non-continuous

fs_skip <- suppressWarnings(get_FSdata(
  df.analysis      = df,
  confounders.name = .CONFS,
  outcome.name     = "rfstime",
  event.name       = "status",
  conf.cont_jcuts  = list(meno = 5),    # binary -> skipped
  use_lasso        = FALSE,
  use_grf          = FALSE,
  details          = FALSE
))

# meno was not continuous -> no j-cuts emitted; falls through to its
# default (categorical indicator) handling.
sum(grepl("^meno ", fs_skip$confs))
sum(grepl("\\bmeno\\b", fs_skip$confs))   # may match "meno." factor cols too

# Compare with default run
identical(
  sort(fs_skip$confs[grepl("^meno", fs_skip$confs)]),
  sort(confs_default[grepl("^meno", confs_default)])
)
```

## Summary

```{r summary}
#| label: summary
#| echo: false
#| results: asis

cat("All assertions passed.  The `conf.cont_jcuts` feature is functioning as designed:\n\n",
    "- J = 5 produces 5 cuts at empirical (k/6)-quantiles (`pgr`).\n",
    "- J = 10 produces 10 cuts at empirical (k/11)-quantiles (`age`).\n",
    "- Unlisted continuous variables retain default behaviour.\n",
    "- Validation errors fire on bad inputs.\n",
    "- Non-continuous variables are skipped with a warning.\n",
    sep = "")
```

## Session info

```{r session-info}
#| label: session-info
#| echo: false

sessionInfo()
```
