---
title: "Simulation Studies Based on ACTG175: Binary Endpoint"
subtitle: "Benefit Subgroup Identification with CD4 Response at 20 Weeks"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ── Simulation control ──────────────────────────────────────────
nsims_alt   <- 44L
nsims_null  <- 44L
sim_seed    <- 8316951L
n_workers   <- min(parallel::detectCores(logical = FALSE) - 1L, 120L)

# ── Dataset preparation: ACTG175 ───────────────────────────────
# Arms 1 (ZDV+ddI) vs 3 (ddI), matching Leon et al. (2024, §4.2)
actg_arms        <- c(1L, 3L)
actg_treat_arm   <- 1L           # ZDV+ddI = experimental
actg_age_cut     <- 34
actg_preanti_cut <- 744.5

# ── Binary outcome ─────────────────────────────────────────────
# I(cd420 > cd40): CD4 count increased from baseline to 20 weeks
# Subjects with missing cd420 are removed.

# ── Benefit search thresholds (on OR scale) ────────────────────
# Under switched treatment, benefit subgroup has OR > 1.
# Thresholds mirror the survival version:
#   c1 = 1/0.60 ≈ 1.667 (screening)
#   c2 = 1/0.80 = 1.25   (consistency)
fs_or_threshold    <- round(1 / 0.60, 3)
fs_or_consistency  <- round(1 / 0.80, 3)

# ── DGM construction ────────────────────────────────────────────
dgm_factor_vars   <- paste0("z", 1:12)
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)
dgm_n_super       <- 5000L

# ── Simulation study ─────────────────────────────────────────────
sim_n_sample      <- 1000L
sim_n_noise       <- 0L

# ── ForestSearch parameters ──────────────────────────────────────
fs_splits              <- 400L
fs_n_min               <- 60L
fs_maxk                <- 2L
fs_vi_grf_min          <- -0.2
```


# Introduction {#sec-intro}

## Motivation

This document extends the ACTG175 benefit-search simulation
framework from survival outcomes (time-to-event) to a **binary
endpoint**: whether a patient's CD4 count improved from baseline
to 20 weeks, defined as $Y = I(\text{cd420} > \text{cd40})$.

This endpoint is the most practical binary outcome available in the
ACTG175 dataset: low missingness (~3%), clear clinical meaning
(immunologic response), and established precedent in the
semiparametric estimation literature (Tsiatis et al., 2008).

The simulation serves two purposes:

1. **GLM extension validation**: confirming that ForestSearch's
   `outcome_type = "binary"` pipeline produces sensible operating
   characteristics for a non-survival endpoint.

2. **Template for GLM-native DGM**: demonstrating a logistic
   regression-based data-generating mechanism that will inform the
   future `generate_binary_dgm()` package function (Tier 3
   roadmap item).


## Benefit Search on the OR Scale

The benefit-search approach from León et al. (2024, §4.2) applies
identically to binary outcomes, with odds ratios (OR) replacing
hazard ratios (HR):

1. **Switch treatment roles**: ddI monotherapy → `treat = 1`,
   ZDV+ddI combination → `treat = 0`.

2. **Raise thresholds** on the OR scale:
   $c_1 = 1/0.60 \approx 1.667$, $c_2 = 1/0.80 = 1.25$.

3. **Run ForestSearch** with `outcome_type = "binary"` and
   `effect_measure = "log_OR"`.

4. **Inversion** is handled automatically by the package when
   `subgroup_notation = "benefit"` is passed to the reporting
   functions.


# Data Preparation {#sec-data}

```{r load-packages}
library(forestsearch)
library(survival)
library(speff2trial)
library(data.table)
library(gt)
library(ggplot2)
library(doFuture)
library(foreach)

timings <- list()
```


```{r prepare-datasets}
t0 <- proc.time()[3]

# ── ACTG175: arms 1 (ZDV+ddI) vs 3 (ddI) ────────────────────
actg_df <- subset(speff2trial::ACTG175, arms %in% actg_arms)
actg_df$id <- seq_len(nrow(actg_df))

# Original treatment: arm 1 = 1, arm 3 = 0
actg_df$treat_orig <- ifelse(actg_df$arms == actg_treat_arm, 1L, 0L)

# Switched treatment for benefit search
actg_df$treat <- 1L - actg_df$treat_orig

# ── Binary outcome: CD4 response at 20 weeks ─────────────────
# I(cd420 > cd40): did CD4 count increase from baseline?
actg_df$y_binary <- as.integer(actg_df$cd420 > actg_df$cd40)

# Remove subjects with missing cd420
n_before <- nrow(actg_df)
actg_df <- actg_df[!is.na(actg_df$cd420), ]
n_after  <- nrow(actg_df)

cat(sprintf("ACTG175: N = %d → %d after removing %d with missing cd420\n",
    n_before, n_after, n_before - n_after))
cat(sprintf("Overall response rate: %.1f%%\n",
    100 * mean(actg_df$y_binary)))
cat(sprintf("  ZDV+ddI (arm 1): %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$treat_orig == 1])))
cat(sprintf("  ddI (arm 3):     %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$treat_orig == 0])))

# ── Discretise covariates (same as survival version) ─────────
actg_df$z1  <- as.factor(ifelse(actg_df$age > actg_age_cut, 1L, 0L))
actg_df$z2  <- as.factor(ifelse(actg_df$preanti <= actg_preanti_cut, 1L, 0L))
actg_df$z3  <- as.factor(ifelse(actg_df$wtkg <= 75, 1L, 0L))
actg_df$z4  <- as.factor(ifelse(
  actg_df$karnof <= median(actg_df$karnof), 1L, 0L))
actg_df$z5  <- as.factor(ifelse(
  actg_df$cd40 <= median(actg_df$cd40), 1L, 0L))
actg_df$z6  <- as.factor(ifelse(
  actg_df$cd80 <= median(actg_df$cd80), 1L, 0L))
actg_df$z7  <- as.factor(actg_df$hemo)
actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs)
actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender)
actg_df$z12 <- as.factor(actg_df$symptom)

# ── Benefit subgroup flag ─────────────────────────────────────
actg_df$flag_harm <- as.integer(
  actg_df$z1 == 1 & actg_df$z2 == 1
)

cat(sprintf("\nBenefit subgroup Q: Age > %d AND Preanti <= %.0f\n",
    actg_age_cut, actg_preanti_cut))
cat(sprintf("  Q prevalence: %d / %d (%.1f%%)\n",
    sum(actg_df$flag_harm), nrow(actg_df),
    100 * mean(actg_df$flag_harm)))

# Response rates by subgroup (original treatment scale)
cat(sprintf("\nResponse rates by subgroup (original scale):\n"))
cat(sprintf("  Q:  ZDV+ddI = %.1f%%,  ddI = %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$flag_harm == 1 &
                                  actg_df$treat_orig == 1]),
    100 * mean(actg_df$y_binary[actg_df$flag_harm == 1 &
                                  actg_df$treat_orig == 0])))
cat(sprintf("  Qc: ZDV+ddI = %.1f%%,  ddI = %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$flag_harm == 0 &
                                  actg_df$treat_orig == 1]),
    100 * mean(actg_df$y_binary[actg_df$flag_harm == 0 &
                                  actg_df$treat_orig == 0])))

timings$data_prep <- proc.time()[3] - t0
```


# Binary Data-Generating Mechanism {#sec-dgm}

Unlike the survival setting where `generate_aft_dgm_flex()` handles
DGM construction, there is currently no package-level binary DGM
generator (Tier 3 roadmap).  We construct one manually using
logistic regression fitted to the observed ACTG175 data.

## Logistic regression model

The binary DGM is:

$$
\text{logit}\, P(Y = 1 \mid V, Z) = \beta_0 + \beta_T V +
  \beta_I \cdot V \cdot Z_1 Z_2 + \boldsymbol{\beta}_Z' \mathbf{Z}
$$

where $V$ is treatment (switched), $Z_1 Z_2$ defines the benefit
subgroup $Q$, and $\mathbf{Z}$ are prognostic covariates.  Under
the null ($\beta_I = 0$), treatment effect is uniform.

```{r build-binary-dgm}
#| code-fold: false
t0 <- proc.time()[3]

# ── Step 1: Fit baseline logistic model (no interaction) ──────
# Using switched treatment so the interaction pushes OR > 1 for Q
z_vars <- paste0("z", 1:12)
base_formula <- as.formula(paste(
  "y_binary ~ treat +",
  paste(z_vars, collapse = " + ")
))

fit_base <- glm(base_formula, data = actg_df, family = binomial)
cat("Baseline logistic model:\n")
cat(sprintf("  Treatment OR (switched): %.3f\n", exp(coef(fit_base)["treat"])))
cat(sprintf("  AIC: %.1f\n", AIC(fit_base)))

# ── Step 2: Create super-population ───────────────────────────
set.seed(sim_seed)
idx_super <- sample(nrow(actg_df), dgm_n_super, replace = TRUE)
df_super  <- actg_df[idx_super, ]
df_super$id <- seq_len(dgm_n_super)

# ── Step 3: Compute potential outcomes ────────────────────────
# P(Y=1 | treat=0, X) and P(Y=1 | treat=1, X) from the baseline model
newdata_0 <- df_super; newdata_0$treat <- 0L
newdata_1 <- df_super; newdata_1$treat <- 1L

df_super$p0_base <- predict(fit_base, newdata = newdata_0, type = "response")
df_super$p1_base <- predict(fit_base, newdata = newdata_1, type = "response")

# ── Step 4: Add treatment-subgroup interaction ────────────────
# Under the alternative, amplify the treatment effect for Q members.
# The interaction adds beta_I to the log-odds for Q members under treatment.
# We calibrate beta_I so the OR(Q) reaches a target.

add_interaction <- function(df, beta_inter) {
  # Modify P(Y=1 | treat=1) for Q members
  in_Q <- df$flag_harm == 1
  eta_1 <- qlogis(df$p1_base)  # log-odds under treatment
  eta_1[in_Q] <- eta_1[in_Q] + beta_inter
  df$p1 <- plogis(eta_1)
  df$p0 <- df$p0_base
  df
}

# ── Step 5: Calibrate beta_inter ──────────────────────────────
# Target: OR(Q) ≈ 2.0 on switched scale (= 0.50 on original scale)
# OR(Q) = [P(Y=1|treat=1,Q) / P(Y=0|treat=1,Q)] /
#          [P(Y=1|treat=0,Q) / P(Y=0|treat=0,Q)]
cal_target_or <- 2.0

compute_or_Q <- function(df, beta_inter) {
  df_mod <- add_interaction(df, beta_inter)
  in_Q <- df_mod$flag_harm == 1
  p1_Q <- mean(df_mod$p1[in_Q])
  p0_Q <- mean(df_mod$p0[in_Q])
  (p1_Q / (1 - p1_Q)) / (p0_Q / (1 - p0_Q))
}

# Grid search for beta_inter
beta_grid <- seq(0, 3.0, by = 0.05)
or_grid   <- sapply(beta_grid, function(b) compute_or_Q(df_super, b))

best_idx  <- which.min(abs(or_grid - cal_target_or))
beta_inter_cal <- beta_grid[best_idx]

cat(sprintf("\nCalibration: beta_inter = %.2f gives OR(Q) = %.3f (target: %.1f)\n",
    beta_inter_cal, or_grid[best_idx], cal_target_or))

# ── Step 6: Build alternative DGM ────────────────────────────
df_super_alt <- add_interaction(df_super, beta_inter_cal)

# Compute subgroup ORs for the super-population
compute_or <- function(df, subset_flag = NULL) {
  if (!is.null(subset_flag)) df <- df[df$flag_harm == subset_flag, ]
  p1 <- mean(df$p1); p0 <- mean(df$p0)
  (p1 / (1 - p1)) / (p0 / (1 - p0))
}

or_Q   <- compute_or(df_super_alt, 1)
or_Qc  <- compute_or(df_super_alt, 0)
or_itt <- compute_or(df_super_alt)

cat(sprintf("  OR(Q):   %.3f  (switched)\n", or_Q))
cat(sprintf("  OR(Qc):  %.3f  (switched)\n", or_Qc))
cat(sprintf("  OR(ITT): %.3f  (switched)\n", or_itt))
cat(sprintf("  → Original scale: OR(Q) = %.3f, OR(Qc) = %.3f\n",
    1/or_Q, 1/or_Qc))

# ── Step 7: Build null DGM (no interaction) ───────────────────
df_super_null <- add_interaction(df_super, 0)
or_null <- compute_or(df_super_null)
cat(sprintf("  Null OR(ITT): %.3f  (switched)\n", or_null))

# ── Package into DGM objects ──────────────────────────────────
# Structure mimics aft_dgm_flex for compatibility with table functions
make_binary_dgm <- function(df_sp, model_type, or_sg, or_sgc, or_overall) {
  list(
    df_super = df_sp,
    hazard_ratios = list(
      harm_subgroup    = or_sg,
      no_harm_subgroup = or_sgc,
      overall          = or_overall
    ),
    model_type = model_type,
    model_params = list(
      beta = coef(fit_base),
      beta_inter = if (model_type == "alt") beta_inter_cal else 0,
      family = "binomial"
    ),
    n_super = dgm_n_super
  )
}

dgm_alt  <- make_binary_dgm(df_super_alt,  "alt",  or_Q, or_Qc, or_itt)
dgm_null <- make_binary_dgm(df_super_null, "null", or_null, or_null, or_null)

timings$dgm <- proc.time()[3] - t0
```


# Simulation Engine {#sec-sim}

Since `simulate_from_dgm()` targets survival data, we implement a
custom binary simulation function.  Each replicate samples from the
super-population, generates binary outcomes from the logistic model,
and runs ForestSearch with `outcome_type = "binary"`.

```{r binary-simulator}
#| code-fold: false

simulate_binary_trial <- function(dgm, n, sim_id, seed) {
  set.seed(seed + sim_id)

  # Sample from super-population
  idx <- sample(nrow(dgm$df_super), n, replace = TRUE)
  df  <- dgm$df_super[idx, ]
  df$id <- seq_len(n)

  # Generate binary outcomes from potential outcomes
  # Assign treatment 50:50
  df$treat_sim <- rbinom(n, 1, 0.5)

  # Outcome: Y ~ Bernoulli(p) where p = p0 or p1 depending on treatment
  df$y_sim <- ifelse(
    df$treat_sim == 1,
    rbinom(n, 1, df$p1),
    rbinom(n, 1, df$p0)
  )
  # For forestsearch binary: event_sim = y_sim (Surv(y,y) trick)
  df$event_sim <- df$y_sim

  df
}

run_binary_analysis <- function(
    sim_data, dgm, fs_params, confounders_name,
    run_fs = TRUE, run_grf = TRUE
) {

  results_list <- list()

  # ── ForestSearch (outcome_type = "binary") ──────────────────
  if (run_fs) {
    fs_res <- tryCatch({
      # Defence in depth: set sequential inside the function so it
      # is safe even when called outside a foreach loop.  The outer
      # %dofuture% body also sets plan("sequential") — both layers
      # are harmless when redundant and critical when one is missing.
      future::plan("sequential")
      suppressWarnings(
        forestsearch(
          df.analysis      = sim_data,
          confounders.name = confounders_name,
          outcome.name     = "y_sim",
          event.name       = "event_sim",
          treat.name       = "treat_sim",
          id.name          = "id",
          outcome_type     = "binary",
          effect_measure   = "log_OR",
          n.min            = fs_params$n.min,
          d0.min           = fs_params$d0.min,
          d1.min           = fs_params$d1.min,
          hr.threshold     = fs_params$hr.threshold,
          hr.consistency   = fs_params$hr.consistency,
          pconsistency.threshold = fs_params$pconsistency.threshold,
          fs.splits        = fs_params$fs.splits,
          maxk             = fs_params$maxk,
          use_lasso        = TRUE,
          use_grf          = FALSE,
          is.RCT           = TRUE,
          by.risk          = 12,
          vi.grf.min       = fs_params$vi.grf.min,
          details          = FALSE,
          Ts.name          = NULL
        )
      )
    }, error = function(e) {
      warning("FS failed: ", e$message)
      NULL
    })

    # Extract estimates
    fs_out <- extract_binary_estimates(
      sim_data, fs_res, dgm, analysis = "FS"
    )
    results_list[["FS"]] <- fs_out
  }

  do.call(rbind, results_list)
}


extract_binary_estimates <- function(df, fs_res, dgm, analysis = "FS") {
  out <- data.frame(
    analysis  = analysis,
    any.H     = 0L,
    hr.H.hat  = NA_real_,  # using HR column names for compatibility
    hr.Hc.hat = NA_real_,  # (these are actually ORs)
    hr.H.true = NA_real_,
    hr.Hc.true = NA_real_,
    hr.itt     = NA_real_,
    sens       = NA_real_,
    spec       = NA_real_,
    ppv        = NA_real_,
    npv        = NA_real_,
    size.H     = NA_real_,
    size.Hc    = nrow(df),
    p.cens     = 0,
    stringsAsFactors = FALSE
  )

  if (is.null(fs_res)) return(out)

  # Check if subgroup was found
  sg_harm <- fs_res$grp.consistency$sg.harm.id
  if (is.null(sg_harm) || length(sg_harm) == 0 ||
      all(is.na(sg_harm)) || all(sg_harm == "")) {
    # No subgroup found — compute ITT OR
    itt_fit <- tryCatch(
      glm(y_sim ~ treat_sim, data = df, family = binomial),
      error = function(e) NULL
    )
    if (!is.null(itt_fit)) out$hr.itt <- exp(coef(itt_fit)["treat_sim"])
    return(out)
  }

  out$any.H <- 1L

  # Get subgroup membership
  df_pred <- fs_res$df.est
  if (is.null(df_pred) || !"treat.recommend" %in% names(df_pred)) {
    return(out)
  }

  in_H  <- df_pred$treat.recommend == 0  # harm/benefit subgroup
  in_Hc <- df_pred$treat.recommend == 1  # complement
  out$size.H  <- sum(in_H)
  out$size.Hc <- sum(in_Hc)

  # Compute ORs via logistic regression
  if (sum(in_H) >= 10 && length(unique(df_pred$treat_sim[in_H])) == 2) {
    fit_H <- tryCatch(
      glm(y_sim ~ treat_sim, data = df_pred[in_H, ], family = binomial),
      error = function(e) NULL
    )
    if (!is.null(fit_H)) out$hr.H.hat <- exp(coef(fit_H)["treat_sim"])
  }

  if (sum(in_Hc) >= 10 && length(unique(df_pred$treat_sim[in_Hc])) == 2) {
    fit_Hc <- tryCatch(
      glm(y_sim ~ treat_sim, data = df_pred[in_Hc, ], family = binomial),
      error = function(e) NULL
    )
    if (!is.null(fit_Hc)) out$hr.Hc.hat <- exp(coef(fit_Hc)["treat_sim"])
  }

  # ITT OR
  itt_fit <- tryCatch(
    glm(y_sim ~ treat_sim, data = df_pred, family = binomial),
    error = function(e) NULL
  )
  if (!is.null(itt_fit)) out$hr.itt <- exp(coef(itt_fit)["treat_sim"])

  # True ORs at identified subgroup composition
  # (computed from the DGM super-population potential outcomes)
  df_sp <- dgm$df_super
  # Map identified factors back to super-population
  out$hr.H.true  <- dgm$hazard_ratios$harm_subgroup
  out$hr.Hc.true <- dgm$hazard_ratios$no_harm_subgroup

  # Classification metrics (if true subgroup is defined)
  true_H <- df_pred$flag_harm == 1
  if (sum(true_H) > 0 && sum(in_H) > 0) {
    tp <- sum(in_H & true_H)
    fp <- sum(in_H & !true_H)
    fn <- sum(!in_H & true_H)
    tn <- sum(!in_H & !true_H)
    out$sens <- tp / (tp + fn)
    out$spec <- tn / (tn + fp)
    out$ppv  <- if (tp + fp > 0) tp / (tp + fp) else NA
    out$npv  <- if (tn + fn > 0) tn / (tn + fn) else NA
  }

  out
}
```


# Run Simulations {#sec-run}

```{r run-parallel-setup}
registerDoFuture()
plan(multisession, workers = n_workers)
cat(sprintf("Workers: %d\n", n_workers))
```

```{r define-params}
fs_params <- list(
  hr.threshold           = fs_or_threshold,
  hr.consistency         = fs_or_consistency,
  pconsistency.threshold = 0.90,
  fs.splits              = fs_splits,
  n.min                  = fs_n_min,
  d0.min                 = 12L,
  d1.min                 = 12L,
  maxk                   = fs_maxk,
  vi.grf.min             = fs_vi_grf_min
)

confounders_name <- paste0("z", 1:12)
```


## Sequential Smoke Test {#sec-smoke}

Before launching the parallel loop, run a single replicate
sequentially to verify that every custom function works end-to-end,
that column names are correct, and that no package function
triggers an error on binary data.

```{r smoke-test}
#| code-fold: false
t0 <- proc.time()[3]

# ── Single-rep smoke test (sequential, no foreach) ────────────
smoke_data <- simulate_binary_trial(dgm_alt, sim_n_sample, sim_id = 0L,
                                     seed = sim_seed)

cat("Smoke-test data:\n")
cat(sprintf("  nrow: %d\n", nrow(smoke_data)))
cat(sprintf("  y_sim prevalence: %.3f\n", mean(smoke_data$y_sim)))
cat(sprintf("  treat_sim balance: %.3f\n", mean(smoke_data$treat_sim)))
stopifnot(
  nrow(smoke_data) == sim_n_sample,
  all(smoke_data$y_sim %in% c(0L, 1L)),
  all(smoke_data$event_sim == smoke_data$y_sim),
  all(c("id", "treat_sim", "y_sim", "event_sim",
        "flag_harm", paste0("z", 1:12)) %in% names(smoke_data))
)
cat("  simulate_binary_trial(): PASS\n")

# ── Run ForestSearch on the smoke-test data ───────────────────
smoke_result <- run_binary_analysis(
  smoke_data, dgm_alt, fs_params, confounders_name,
  run_fs = TRUE, run_grf = FALSE
)

expected_cols <- c("analysis", "any.H", "hr.H.hat", "hr.Hc.hat",
                   "hr.H.true", "hr.Hc.true", "hr.itt",
                   "sens", "spec", "ppv", "npv",
                   "size.H", "size.Hc", "p.cens")
missing_cols <- setdiff(expected_cols, names(smoke_result))
stopifnot(length(missing_cols) == 0)
cat(sprintf("  run_binary_analysis(): PASS (any.H = %d)\n",
    smoke_result$any.H))

# ── Verify .combine = rbind compatibility ─────────────────────
smoke_pair <- rbind(smoke_result, smoke_result)
stopifnot(nrow(smoke_pair) == 2, ncol(smoke_pair) == ncol(smoke_result))
cat("  rbind compatibility: PASS\n")

timings$smoke_test <- proc.time()[3] - t0
cat(sprintf("Smoke test completed in %.1f sec\n", timings$smoke_test))

# ── DGM contract verification ────────────────────────────────
# Ensure DGM objects satisfy the interface expected by
# get_dgm_hr() and the table functions.
for (dgm_name in c("dgm_alt", "dgm_null")) {
  dgm_check <- get(dgm_name)
  stopifnot(
    is.numeric(dgm_check$hazard_ratios$harm_subgroup),
    is.numeric(dgm_check$hazard_ratios$no_harm_subgroup),
    is.numeric(dgm_check$hazard_ratios$overall),
    dgm_check$hazard_ratios$harm_subgroup > 0,
    "flag_harm" %in% names(dgm_check$df_super)
  )
  cat(sprintf("  %s contract: PASS\n", dgm_name))
}
```


## Alternative Hypothesis

```{r sim-alt}
t0 <- proc.time()[3]

results_alt <- foreach(
  i = seq_len(nsims_alt),
  .combine = rbind,
  .errorhandling = "remove",
  .packages = c("forestsearch", "survival"),
  .options.future = list(seed = TRUE)
) %dofuture% {
  future::plan("sequential")

  sim_data <- simulate_binary_trial(dgm_alt, sim_n_sample, i, sim_seed)
  run_binary_analysis(sim_data, dgm_alt, fs_params, confounders_name,
                      run_fs = TRUE, run_grf = FALSE)
}

# Post-hoc failure accounting
n_completed_alt <- nrow(results_alt)
n_failed_alt    <- nsims_alt - n_completed_alt
if (n_failed_alt > 0) {
  warning(sprintf(
    "%d of %d alt simulations failed silently (.errorhandling = 'remove')",
    n_failed_alt, nsims_alt
  ))
}

timings$sim_alt <- proc.time()[3] - t0
cat(sprintf("Alt sims: %d / %d completed (%.1f sec)\n",
    n_completed_alt, nsims_alt, timings$sim_alt))
if (n_failed_alt > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_alt))
cat(sprintf("Detection rate: %.1f%%\n",
    100 * mean(results_alt$any.H)))
```


## Null Hypothesis

```{r sim-null}
t0 <- proc.time()[3]

results_null <- foreach(
  i = seq_len(nsims_null),
  .combine = rbind,
  .errorhandling = "remove",
  .packages = c("forestsearch", "survival"),
  .options.future = list(seed = TRUE)
) %dofuture% {
  future::plan("sequential")

  sim_data <- simulate_binary_trial(dgm_null, sim_n_sample, i,
                                     sim_seed + 100000L)
  run_binary_analysis(sim_data, dgm_null, fs_params, confounders_name,
                      run_fs = TRUE, run_grf = FALSE)
}

# Post-hoc failure accounting
n_completed_null <- nrow(results_null)
n_failed_null    <- nsims_null - n_completed_null
if (n_failed_null > 0) {
  warning(sprintf(
    "%d of %d null simulations failed silently (.errorhandling = 'remove')",
    n_failed_null, nsims_null
  ))
}

timings$sim_null <- proc.time()[3] - t0
cat(sprintf("Null sims: %d / %d completed (%.1f sec)\n",
    n_completed_null, nsims_null, timings$sim_null))
if (n_failed_null > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_null))
cat(sprintf("False discovery rate: %.1f%%\n",
    100 * mean(results_null$any.H)))
```

```{r cleanup-plan}
plan(sequential)
```


# Results {#sec-results}

All results are displayed on the **original** treatment scale
(OR < 1 = benefit from ZDV+ddI) via
`subgroup_notation = "benefit"`, which triggers automatic
inversion inside the reporting functions.

## Operating Characteristics Under H1

```{r oc-alt}
t0 <- proc.time()[3]

format_oc_results(
  results_alt,
  metrics   = c("detection", "classification"),
  digits    = 2,
  digits_hr = 2,
  title     = "ACTG175 Binary: Identification Properties (Benefit Search)",
  subtitle  = sprintf(
    "n = %d, %d sims, OR(Q) = %.2f, OR(Qc) = %.2f",
    sim_n_sample, nsims_alt,
    1 / dgm_alt$hazard_ratios$harm_subgroup,
    1 / dgm_alt$hazard_ratios$no_harm_subgroup
  ),
  subgroup_notation = "benefit"
)

build_estimation_table(
  results         = results_alt,
  dgm             = dgm_alt,
  analysis_method = "FS",
  subtitle = sprintf(
    "ACTG175 Binary DGM | n = %d, %d sims",
    sim_n_sample, nsims_alt
  ),
  font_size = 16,
  subgroup_notation = "benefit"
)

timings$report_alt <- proc.time()[3] - t0
```

### Interpretation

```{r interpret-alt, results="asis"}
interpret_estimation_table(
  results_alt, dgm_alt,
  analysis_method = "FS",
  scenario = "alt",
  subgroup_notation = "benefit"
)
```


## Operating Characteristics Under H0

```{r oc-null}
format_oc_results(
  results_null,
  metrics  = c("detection", "classification"),
  digits   = 2,
  title    = "ACTG175 Binary: Type I Error (Null)",
  subtitle = sprintf(
    "n = %d, %d sims, OR(ITT) = %.2f",
    sim_n_sample, nsims_null,
    1 / dgm_null$hazard_ratios$overall
  ),
  subgroup_notation = "benefit"
)
```


## Classification Table

```{r classification}
build_classification_table(
  scenario_results = list(
    null = list(
      results = results_null, label = "ACTG-bin",
      n_sample = sim_n_sample, dgm = dgm_null, hypothesis = "null"
    ),
    alt = list(
      results = results_alt, label = "ACTG-bin",
      n_sample = sim_n_sample, dgm = dgm_alt, hypothesis = "alt"
    )
  ),
  analyses = sort(unique(c(
    unique(results_null$analysis), unique(results_alt$analysis)
  ))),
  digits = 2, font_size = 16,
  title = "ACTG175 Binary: Subgroup Identification and Classification",
  n_sims = nsims_alt,
  subgroup_notation = "benefit"
)
```


# Timing Summary

```{r timing}
timing_df <- data.frame(
  Step = names(timings),
  Seconds = round(unlist(timings), 1),
  Minutes = round(unlist(timings) / 60, 2),
  Hours   = round(unlist(timings) / 3600, 3)
)
timing_df |>
  gt() |>
  gt::tab_header(title = "Computation Timing")
```


# Pressure-Testing Strategy {#sec-pressure-test}

This section documents the validation protocol for custom
functions developed in Quarto documents before they are graduated
to the `forestsearch` package.  The protocol addresses three
risk categories: **correctness**, **parallelism safety**, and
**interface compatibility**.

## 1. Sequential smoke test (before any parallel execution)

Every simulation document MUST include a sequential single-rep
test (`§sec-smoke`) that runs the entire pipeline — data
generation, analysis, extraction — in the main R process. This
catches errors that would be silently swallowed by
`.errorhandling = "remove"` in the parallel loop.

The smoke test validates:

- **Column names**: the result data frame has exactly the columns
  expected by the downstream table functions (`hr.H.hat`,
  `hr.Hc.hat`, `any.H`, `sens`, etc.).

- **Column types**: all HR-related columns are `numeric`, all
  classification columns are `numeric`, `analysis` is `character`.

- **rbind compatibility**: two result rows can be combined via
  `rbind()` without type coercion or column mismatch (this is what
  `.combine = rbind` does inside `foreach`).

- **No side effects**: the function does not modify global state,
  write files, or call `plan()` in a way that corrupts the
  caller's parallelism context.

## 2. Nested parallelism prevention

The `forestsearch()` function may internally attempt to use
parallel resources (e.g., via `grf::causal_forest()`).  If a
worker process inside a `%dofuture%` loop spawns child futures,
the system can deadlock or exhaust memory.

**Mandatory rule**: every `%dofuture%` worker body MUST begin
with `future::plan("sequential")` as its first statement. This
forces all package-internal parallelism to run sequentially
within each worker.

**Defence in depth**: analysis wrapper functions (like
`run_binary_analysis()`) SHOULD also call
`future::plan("sequential")` internally, so they are safe even
when called outside a `foreach` loop.  This creates two layers:

```r
results <- foreach(...) %dofuture% {
  future::plan("sequential")       # Layer 1: worker body
  run_binary_analysis(...)          # Layer 2: inside function
}
```

Both layers are harmless when redundant and critical when one is
missing.

## 3. Post-hoc failure accounting

The `.errorhandling = "remove"` option in `foreach` silently
drops failed iterations.  Without accounting, the user may not
notice that 10 of 50 simulations failed.

**Mandatory pattern**:

```r
n_completed <- nrow(results)
n_failed    <- nsims - n_completed
if (n_failed > 0) {
  warning(sprintf("%d of %d sims failed", n_failed, nsims))
}
```

This must appear immediately after each `foreach` loop.  For
production runs, consider switching to
`.errorhandling = "pass"` with explicit error filtering, which
preserves error messages for debugging.

## 4. DGM contract verification

Custom DGM objects must satisfy the interface contract expected
by `get_dgm_hr()` and the table functions:

- `dgm$hazard_ratios$harm_subgroup`: numeric scalar (OR or HR
  for the identified subgroup).
- `dgm$hazard_ratios$no_harm_subgroup`: numeric scalar for
  complement.
- `dgm$hazard_ratios$overall`: numeric scalar for ITT.
- `dgm$df_super$flag_harm`: integer 0/1 subgroup indicator.

The smoke test should verify these fields exist and have
sensible values:

```r
stopifnot(
  is.numeric(dgm$hazard_ratios$harm_subgroup),
  is.numeric(dgm$hazard_ratios$no_harm_subgroup),
  dgm$hazard_ratios$harm_subgroup > 0,
  "flag_harm" %in% names(dgm$df_super)
)
```

## 5. Package function serialisation in workers

Functions defined in the Quarto environment (not in a package)
are automatically serialised to workers by `doFuture`.  However,
functions that reference **package-internal** (non-exported)
helpers will fail inside workers unless:

- The package is listed in `.packages`, OR
- The helper is explicitly imported via `::` or `:::`

For custom functions calling `forestsearch()`, the
`.packages = c("forestsearch", "survival")` declaration is
sufficient.  But if a custom extraction function calls
`forestsearch:::some_internal_helper()`, this must be tested in
the smoke test — triple-colon access works in the main process
but may fail in a worker where the namespace isn't fully loaded.

## 6. Graduation checklist

Before moving a custom function from a Quarto document into the
package `R/` directory:

1. **Smoke test passes** with `stopifnot()` assertions.
2. **Parallel loop completes** with zero failures.
3. **Results match sequential execution**: run 5 reps
   sequentially and 5 reps in parallel with the same seeds;
   verify identical results.
4. **No `future::plan()` calls in the function body** (move
   them to the caller or document why they are needed).
5. **roxygen2 documentation** with `@param`, `@return`,
   `@keywords internal` or `@export`.
6. **NAMESPACE updated** via `devtools::document()`.
7. **`devtools::check()`** passes with 0 errors, 0 warnings.


# Discussion {#sec-discussion}

## Key differences from the survival simulation

This binary endpoint simulation differs from the survival version
in several important ways:

1. **No censoring**: the binary outcome $I(\text{cd420} > \text{cd40})$
   is fully observed (after removing missing values), eliminating
   censoring model complexity.

2. **Custom DGM**: since `generate_aft_dgm_flex()` targets survival
   endpoints, the binary DGM was constructed manually via logistic
   regression. This serves as a prototype for the future
   `generate_binary_dgm()` function.

3. **Custom simulation loop**: `simulate_from_dgm()` generates
   survival data, so a custom `simulate_binary_trial()` was
   implemented. This would be absorbed into a general
   `simulate_from_dgm()` extension.

4. **Effect measure**: odds ratios replace hazard ratios. The
   `subgroup_notation = "benefit"` parameter handles inversion
   identically on either scale.

5. **ForestSearch identification**: the `outcome_type = "binary"`
   parameter routes to the `Surv(y, y)` trick for the consistency
   criterion, with GLM-based estimation via `create_glm_row()`.

## Candidate package functions (Tier 3)

This document identifies three functions for potential package
integration:

- `generate_binary_dgm()`: logistic regression-based DGM
  constructor, analogous to `generate_aft_dgm_flex()` for survival.
- `simulate_binary_trial()`: binary outcome generator from a
  logistic DGM, analogous to `simulate_from_dgm()`.
- `run_binary_simulation()`: general wrapper for binary simulation
  studies, analogous to `run_simulation_analysis()`.


# References {.unnumbered}

- Hammer, S.M., et al. (1996). A trial comparing nucleoside
  monotherapy with combination therapy in HIV-infected adults.
  *NEJM*, 335, 1081--1090.

- León, L.F., et al. (2024). Exploratory subgroup identification
  in the heterogeneous Cox model. *Statistics in Medicine*, 43,
  3921--3942.

- Tsiatis, A.A., Davidian, M., Zhang, M., Lu, X. (2008).
  Covariate adjustment for two-sample treatment comparisons in
  randomized clinical trials. *Statistics in Medicine*, 27,
  4658--4677.
