---
title: "Frozen-family check: FB vs MR correspondence under non-regenerating cuts"
format:
  html:
    embed-resources: true
    toc: true
    code-fold: true
bibliography: []
---

<!--
  analysis_gbsg_frozen_family.qmd

  A controlled variant of the GBSG maxeff analysis whose sole purpose is to
  test the estimand claim that the Full Bootstrap (FB) and Multiplier Resampling
  (MR) corrections differ BECAUSE FB regenerates the candidate family on each
  resample (recomputing quantile cutpoints), while MR holds the fitted family
  fixed.

  MECHANISM (established by reading the package). The bootstrap replays the
  original forestsearch call arguments (fs.est$args_call_all) on each resample.
  Under conf.cont_jcuts, the *instruction* "cut er at its J quantiles" is
  replayed, so the quantiles are recomputed on each bootstrap sample and the cut
  LOCATIONS move -- the family regenerates. Under a conf_force expression with a
  fixed numeric threshold ("er <= 36"), the cut does NOT move: it is a fixed
  string, exempt from cut collapse, replayed verbatim.

  DESIGN. Freeze every cut. Drop the continuous factors from confounders.name
  (so nothing is auto-cut) and supply, via conf_force, a fixed cut per
  continuous factor at its OBSERVED-DATA median, plus the clinically meaningful
  receptor-negative cuts {er <= 0} and {pgr <= 0}. Keep the categorical factors
  (meno, grade3). Remove conf.cont_jcuts entirely. In this family FB has nothing
  to regenerate, so the prediction is FB ~ MR (up to Monte Carlo error and MR's
  leading-order approximation).

  WHAT THIS DOES AND DOES NOT SHOW. It does NOT reproduce the headline maxeff
  numbers -- the family is a coarser median-cut family, so the selected subgroup
  and its effect estimates will differ. Do not compare point estimates across
  the two analyses. It DOES test whether, in a provably frozen family, FB and MR
  correspond. The headline analysis's FB-MR gap (regenerating family) is the
  contrast: regeneration present -> divergence; regeneration absent ->
  correspondence. Two runs, one claim.

  VENUE. Requires the forestsearch package installed. Compute-heavy (a
  forestsearch fit + an FB bootstrap). Run on the analysis machine, not the chat
  sandbox.
-->

```{r setup}
#| message: false
#| warning: false
library(forestsearch)
library(survival)
library(data.table)
library(doFuture)
library(kableExtra)

timings <- list()

# Guo & He comparator: sourced, not installed (as in the headline analysis).
gh_dir <- Sys.getenv("GH_DIR", unset = "quarto/GuoHe")
source(file.path(gh_dir, "guohe_algorithm3.R"))
source(file.path(gh_dir, "fs_to_guohe.R"))
```

```{r parallel-setup}
n_cores <- max(1L, floor(0.80 * parallel::detectCores(logical = FALSE)))
NB         <- 1000L     # full-bootstrap resamples, matching the headline run
gate_draws <- 5000L     # MR multiplier draws
GH_B       <- 2000L     # Guo & He pair-bootstrap resamples
GH_r       <- 0.03      # Guo & He tuning parameter
GH_PARALLEL <- .Platform$OS.type == "unix"   # forked GH bootstrap where available
run_loo    <- TRUE      # leave-one-out (N-fold) honest out-of-sample subgroup estimate
timings$loo <- NA_real_
cat("Using", n_cores, "physical cores\n")
```

```{r data}
df.analysis <- gbsg
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
})

# The forced cuts are computed ONCE on the observed data and hard-coded as
# fixed-threshold expressions, so the bootstrap replay cannot move them.
med <- sapply(c("age", "size", "nodes", "pgr", "er"),
              function(v) stats::median(df.analysis[[v]]))
frozen_cuts <- c(
  sprintf("age <= %g",   med[["age"]]),     # observed median
  sprintf("size <= %g",  med[["size"]]),
  sprintf("nodes <= %g", med[["nodes"]]),
  sprintf("pgr <= %g",   med[["pgr"]]),
  sprintf("er <= %g",    med[["er"]]),
  "er <= 0",                                 # receptor-negative (clinical)
  "pgr <= 0"
)
cat("Frozen cuts (fixed thresholds, replayed verbatim under bootstrap):\n")
cat(paste0("  ", frozen_cuts), sep = "\n")

# CATEGORICAL factors only in confounders.name: the continuous factors are
# DROPPED so forestsearch never auto-cuts (and never quantile-recomputes) them.
confounders.name <- c("meno", "grade3")

outcome.name <- "time_months"
event.name   <- "status"
id.name      <- "id"
treat.name   <- "hormon"
```

## Frozen-family fit {#sec-fit}

Every continuous factor is represented only by fixed median cuts (plus the
receptor-negative cuts); the continuous factors themselves are excluded from the
confounder set, and `conf.cont_jcuts` is not used. The candidate family is
therefore fully determined by the fixed cuts and cannot regenerate under
bootstrap resampling.

```{r fit}
#| fig-width: 10
#| fig-height: 7
t0 <- proc.time()

fs <- forestsearch(
  df.analysis,
  outcome.name       = outcome.name,
  event.name         = event.name,
  treat.name         = treat.name,
  id.name            = id.name,
  confounders.name   = confounders.name,   # categorical only
  is.RCT             = TRUE,
  seedit             = 8316951,
  est.scale          = "hr",
  use_lasso          = FALSE,
  use_grf            = FALSE,
  use_dina           = FALSE,
  conf_force         = frozen_cuts,        # all cuts fixed
  # conf.cont_jcuts intentionally OMITTED -- no quantile cuts
  collapse_cuts      = TRUE,
  subgroup_method    = "consistency",
  sg_focus           = "maxeff",
  n.min              = 60, d0.min = 10, d1.min = 10, maxk = 2,
  debias_gate        = TRUE,
  debias_gate_args   = list(draws = gate_draws),
  consistency_method = "resample",
  fs.splits          = 1000, stop_threshold = 0.95,
  use_twostage       = TRUE, max.minutes = 3,
  show_candidate_summary = TRUE, details = TRUE, quiet = FALSE,
  plot.sg            = TRUE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE)
)
plan("sequential")
timings$fit <- (proc.time() - t0)[["elapsed"]]
stopifnot(!is.null(fs$sg.harm))
cat("\nSelected subgroup:", paste(fs$sg.harm, collapse = " & "),
    "  |  fit", round(timings$fit / 60, 1), "min\n")
```

### Gate 0 -- confirm the family is actually frozen

The whole check is void if any candidate cut is quantile-derived (hence
regenerating). Verify that the realized candidate family contains only the fixed
cuts: every cut label must correspond to one of the forced expressions or a
categorical factor level, and no continuous factor may appear as an auto-derived
cut. This is a hard gate -- if it fails, stop; the comparison below is not
interpretable.

```{r gate0}
fam <- as.data.frame(fs$grp.consistency$out_sg$result)
mcols <- grep("^M\\.[0-9]+$", names(fam), value = TRUE)
labels <- unique(unlist(lapply(mcols, function(cn) fam[[cn]])))
labels <- labels[nzchar(labels)]

# Normalize a family cut label ("{er <= 0}", "!{age <= 53}") to its inner form.
norm_lab <- function(x) {
  x <- sub("^!", "", trimws(x))
  trimws(sub("^\\{", "", sub("\\}$", "", x)))
}
inner <- sort(unique(norm_lab(labels)))

# Allowed inner forms: the forced cuts, plus categorical factor cuts on
# meno / grade3 (which forestsearch may encode as "meno <= 0" / bare levels).
allowed_fixed <- frozen_cuts
cat("Distinct candidate cut forms in the realized family:\n")
cat(paste0("  ", inner), sep = "\n")

# Flag any cut on a continuous factor that is NOT one of the fixed thresholds.
cont_vars <- c("age", "size", "nodes", "pgr", "er")
offending <- inner[
  vapply(inner, function(s) {
    v <- sub("\\s*(<=|>=|<|>|==).*$", "", s)
    v <- trimws(v)
    (v %in% cont_vars) && !(s %in% trimws(sub("\\s+", " ", allowed_fixed)))
  }, logical(1))
]

if (length(offending)) {
  stop("GATE 0 FAILED -- continuous-factor cuts not in the fixed set are ",
       "present (family is NOT frozen): ",
       paste(offending, collapse = "; "))
}
cat("\nGATE 0 PASS -- every continuous-factor cut is a fixed threshold; ",
    "the family cannot regenerate under bootstrap.\n", sep = "")
```

## Full bootstrap (FB)

```{r bootstrap}
t0 <- proc.time()
fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est   = fs, nb_boots = NB, show_three = FALSE, details = FALSE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE)
)
plan("sequential")
timings$fb <- (proc.time() - t0)[["elapsed"]]
cat("FB completed in", round(timings$fb / 60, 1), "min\n")
```

## Multiplier resampling (MR)

```{r mr}
g <- fs$debias_gate
stopifnot(!is.null(g))
timings$mr <- g$timing_seconds
cat("MR gate:", round(timings$mr, 2), "s over",
    format(gate_draws, big.mark = ","), "draws\n")
```

## Guo & He (GH)

The adapter exports the frozen candidate family as membership columns and runs
Guo & He Algorithm 3 on the same selected subgroup, asserting membership
integrity against `fit$grp.consistency$sg.harm.id` before the estimator runs.
The frozen family is coarser than the headline maxeff family (about 116
candidates versus ~1340), which remains well within the range where the argmax
de-biasing is meaningful.

```{r guohe}
t0 <- proc.time()
gh <- fs_to_guohe(
  fit       = fs,
  time      = "time_months", event = "status", treatment = "hormon",
  B         = GH_B, r = GH_r, level = 0.05, seed = 8316951,
  parallel  = GH_PARALLEL, verify = TRUE
)
timings$gh <- (proc.time() - t0)[["elapsed"]]
cat("GH family size:", length(gh$export$candidates),
    " | selected:", gh$selected,
    " | GH", round(timings$gh / 60, 1), "min\n")
cat(sprintf("GH  naive HR %.3f  debiased HR %.3f  lower bound %.3f\n",
            gh$naive_hr, gh$debiased_hr, gh$bound_hr))
```

## Leave-one-out (honest out-of-sample estimate)

Leave-one-out re-runs the entire selection on each hold-out of one subject and
predicts the held-out subject's subgroup membership from its covariates. Because
no subject informs its own subgroup label, the resulting subgroup effect is free
of selection bias at the source, rather than corrected for it after the fact --
a quasi-external reference. This is `forestsearch_Kfold()` at N-fold
(`Kfolds = nrow(df)`), the same leave-one-out analysis carried in the multimethod
document, applied to the frozen-family fit. The out-of-sample subgroup hazard
ratio is added to the comparison below.

Note this is the heaviest step: N-fold means one full forest search per subject
(here 686 re-runs). It is tractable on the small frozen family but is the long
pole; on the analysis box, not the sandbox.

```{r loo-skipped-note, eval = !run_loo}
cat("Leave-one-out skipped (run_loo = FALSE).",
    "Set run_loo <- TRUE in parallel-setup to run it.\n")
```

```{r loo, eval = run_loo}
t0 <- proc.time()
fs_OOB <- forestsearch_Kfold(
  fs.est  = fs,
  details = FALSE,
  Kfolds  = nrow(df.analysis),   # N-fold = leave-one-out
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE)
)
plan("sequential")
timings$loo <- (proc.time() - t0)[["elapsed"]]

cv_out <- forestsearch_KfoldOut(res = fs_OOB, details = FALSE, outall = TRUE)

# The honest out-of-sample subgroup estimate is the K-fold (Kf) row of
# SG_tab_Kfold; the harm subgroup is row 1. The "HR (95% CI)" column is a
# preformatted string, so it is displayed as-is in the comparison table.
loo_hr_str <- as.character(cv_out$SG_tab_Kfold[1, "HR (95% CI)"])
cat("LOO out-of-sample subgroup HR:", loo_hr_str,
    " | LOO", round(timings$loo / 60, 1), "min\n")

# Full CV tables (ITT / full-analysis / K-fold, plus agreement metrics).
tables <- cv_summary_tables(cv_out)
tables$combined_table
tables$metrics_table
```

## Corrections in the frozen family {#sec-compare}

```{r table}
#| tbl-cap: "Subgroup hazard ratios in the FROZEN (non-regenerating) family. The FB-MR gap is the quantity of interest (see below). Guo & He reports a one-sided lower bound. Leave-one-out is an honest out-of-sample estimate (not a de-biasing correction), shown to see which correction it tracks."
#| echo: false
fci <- function(e, l, u, d = 2) {
  ifelse(is.na(e), "\u2014", sprintf("%.*f (%.*f, %.*f)", d, e, d, l, d, u))
}
naive_hr <- fs_bc$H_estimates$H0

loo_row <- if (isTRUE(run_loo) && exists("cv_out")) {
  as.character(cv_out$SG_tab_Kfold[1, "HR (95% CI)"])
} else "\u2014 (run_loo = FALSE)"

tab <- data.frame(
  Method = c("Naive", "Full bootstrap", "Multiplier resampling",
             "Guo & He (2021)", "Leave-one-out (out-of-sample)"),
  Estimate = c(
    sprintf("%.2f", naive_hr),
    fci(fs_bc$H_estimates$H2, fs_bc$H_estimates$H2_lower, fs_bc$H_estimates$H2_upper),
    fci(g$debiased$est, g$debiased$lower, g$debiased$upper),
    sprintf("%.2f (LB %.2f)", gh$debiased_hr, gh$bound_hr),
    loo_row
  ),
  check.names = FALSE
)
kbl(tab, booktabs = TRUE, escape = FALSE, align = "ll") |> kable_styling()
```

```{r gap}
#| echo: false
fb_hr <- fs_bc$H_estimates$H2
mr_hr <- g$debiased$est
cat(sprintf("FB debiased HR: %.3f\n", fb_hr))
cat(sprintf("MR debiased HR: %.3f\n", mr_hr))
cat(sprintf("FB - MR (log-HR): %.4f\n", log(fb_hr) - log(mr_hr)))
cat(sprintf("FB / MR (HR ratio): %.4f\n", fb_hr / mr_hr))
```

::: {.callout-note}
## Interpretation

If the frozen-family FB and MR debiased HRs agree up to Monte Carlo error, that
corroborates the account that their divergence in the headline maxeff analysis
is the family-regeneration component: FB there recomputes quantile cutpoints on
each resample and so marginalizes over family regeneration (targeting the
marginal $\theta^\dagger(H)$), while MR holds the fitted family fixed
(targeting the conditional $\beta(\widehat{H})$). Here, with every cut fixed,
there is nothing to regenerate, so the two estimands coincide and the estimators
should correspond.

A small residual FB-MR difference may persist even frozen, from MR's
leading-order (infinitesimal-jackknife) approximation to the bootstrap -- a
distinct term from regeneration. A large residual would instead implicate that
approximation and would be its own finding. This run does not, by itself,
decompose those two terms; it tests whether the dominant regeneration term is
removed by freezing the family.

The comparison is within-family only. The selected subgroup here differs from
the headline maxeff subgroup (a coarser median-cut family), so point estimates
are not comparable across the two analyses -- only the FB-MR gap within each is.

Leave-one-out is included as an honest, selection-bias-free out-of-sample
reference: it estimates the subgroup effect without any subject informing its
own subgroup label, so it corrects the winner's-curse optimism at the source
rather than after the fact. In the frozen family, where the candidate set cannot
regenerate, the conditional and marginal targets coincide, so LOO is expected to
sit near both FB and MR (which themselves nearly coincide here). The more
discriminating comparison is in the headline maxeff family, where regeneration
separates FB (marginal) from MR (conditional): there, which correction LOO
tracks is informative rather than predetermined. Read the LOO row as a reference
point, not a fifth correction.
:::

```{r timing}
#| echo: false
td <- data.frame(
  Step = c("Fit + MR gate", "Full bootstrap", "MR gate (within fit)",
           "Guo & He bootstrap", "Leave-one-out (N-fold)"),
  Seconds = c(timings$fit, timings$fb, timings$mr, timings$gh, timings$loo)
)
td$Minutes <- round(td$Seconds / 60, 2)
kbl(td, booktabs = TRUE, align = "lrr") |> kable_styling()
```
