---
title: "GBSG maxeff subgroup analysis: Naive, Full Bootstrap, Multiplier Resampling, and Guo & He"
format:
  html:
    embed-resources: true
    toc: true
    code-fold: true
bibliography: []
---

<!--
  analysis_gbsg_cox_maxeff.qmd

  A simplified single-track version of analysis_gbsg_cox_multimethod_*.qmd,
  reduced to the sg_focus = "maxeff" path where Guo & He (2021) Algorithm 3
  applies, and adding Guo & He as a fourth de-biasing correction alongside the
  Naive estimate, the Full Bootstrap (FB), and Multiplier Resampling (MR).

  Relative to the multimethod document, this DROPS: the DINA and GRF selection
  tracks (Guo & He's argmax comparator is not the relevant correction for
  data-driven regenerating families), and the CV/LOO validation blocks.

  Guo & He enters through the sourced adapter fs_to_guohe.R, which exports the
  maxeff candidate family as membership columns and runs guohe_algorithm3() on
  the SAME selected subgroup forestsearch chose. Neither guohe_algorithm3.R nor
  the forestsearch package is modified.

  VENUE. The forestsearch fit, the full bootstrap, and the Guo & He pair
  bootstrap are compute-heavy and require the forestsearch package installed.
  Run on the analysis machine (Linux, many cores) -- see the parallelism note in
  the setup chunk. Not a chat-sandbox document.
-->

```{r setup}
#| message: false
#| warning: false
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(doFuture)
library(kableExtra)

timings <- list()

# ---------------------------------------------------------------------------
# Guo & He comparator: sourced, not installed. Point gh_dir at the directory
# holding guohe_algorithm3.R and fs_to_guohe.R (e.g. the repo's quarto/GuoHe/).
# ---------------------------------------------------------------------------
gh_dir <- Sys.getenv("GH_DIR", unset = "quarto/GuoHe")
source(file.path(gh_dir, "guohe_algorithm3.R"))
source(file.path(gh_dir, "fs_to_guohe.R"))
```

```{r parallel-setup}
# PARALLELISM NOTES (see the doFuture discussion):
#   * FB  -- forestsearch_bootstrap_dofuture() uses %dofuture%; it inherits the
#            plan set here and restores the previous plan on exit. Any plan works.
#   * MR  -- the de-biased gate is vectorized matrix algebra (a single crossprod
#            over all draws), NOT a parallel loop; it needs no plan, only a
#            threaded BLAS to benefit. Do not wrap it in a future.
#   * GH  -- guohe_algorithm3()'s pair bootstrap is parallelised via its own
#            `parallel` argument. Because guohe_algorithm3() is SOURCED (not
#            installed), multisession workers would not see it; use forked
#            workers (plan(multicore), Linux/Mac) so the sourced function is
#            inherited, OR run GH serially. On the Linux analysis box, multicore
#            is available.
n_cores <- max(1L, floor(0.80 * parallel::detectCores(logical = FALSE)))
cat("Using", n_cores, "physical cores for parallel sections\n")

NB         <- 1000L     # full-bootstrap resamples (production: 500-2000)
gate_draws <- 5000L     # MR multiplier draws
GH_B       <- 2000L     # Guo & He pair-bootstrap resamples
GH_r       <- 0.03      # Guo & He tuning parameter (their adaptive default region)
GH_PARALLEL <- .Platform$OS.type == "unix"   # forked GH bootstrap where available
```

```{r data}
df.analysis <- gbsg
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
})

confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name   <- "status"
id.name      <- "id"
treat.name   <- "hormon"

cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
```

## Maxeff identification {#sec-fit}

The forest search is run under `sg_focus = "maxeff"`: the selected subgroup is
the effect maximiser over the enumerated candidate family -- the argmax
primitive of Guo and He (2021), and the selection Guo & He's de-biasing is built
to correct. Under `maxeff` the candidate pool is not truncated, so the family
the correction ranges over is the full enumerated family. The receptor cuts
$\{er \le 0\}$ and $\{pgr \le 0\}$ are pre-specified as forced candidates, and
$J = 10$ cuts are used for the continuous `er` and `pgr` covariates.

```{r forestsearch-maxeff}
#| fig-width: 10
#| fig-height: 7
t0 <- proc.time()

fs <- forestsearch(
  df.analysis,
  outcome.name          = outcome.name,
  event.name            = event.name,
  treat.name            = treat.name,
  id.name               = id.name,
  confounders.name      = confounders.name,
  is.RCT                = TRUE,
  seedit                = 8316951,
  est.scale             = "hr",
  # candidate generation: consistency search only (no LASSO/GRF/DINA screens)
  use_lasso             = FALSE,
  use_grf               = FALSE,
  use_dina              = FALSE,
  # continuous-cut construction + pre-specified receptor candidates
  conf.cont_jcuts       = list(er = 10, pgr = 10),
  collapse_cuts         = TRUE,
  conf_force            = c("er <= 0", "pgr <= 0"),
  # selection: the effect-maximiser argmax (Guo & He primitive)
  subgroup_method       = "consistency",
  sg_focus              = "maxeff",
  # subgroup constraints
  n.min                 = 60, d0.min = 10, d1.min = 10, maxk = 2,
  # Tier-2 de-biased gate (MR) computed inside the fit
  debias_gate           = TRUE,
  debias_gate_args      = list(draws = gate_draws),
  # consistency search
  consistency_method    = "resample",
  fs.splits             = 1000, stop_threshold = 0.95,
  use_twostage          = TRUE, max.minutes = 3,
  show_candidate_summary = TRUE, details = TRUE, quiet = FALSE,
  plot.sg               = TRUE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE)
)
plan("sequential")
timings$fit <- (proc.time() - t0)[["elapsed"]]

stopifnot(!is.null(fs$sg.harm))
cat("\nSelected subgroup:", paste(fs$sg.harm, collapse = " & "),
    "  |  fit", round(timings$fit / 60, 1), "min\n")
```

## Full bootstrap (FB)

```{r bootstrap}
t0 <- proc.time()
fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est   = fs,
  nb_boots = NB,
  show_three = FALSE,
  details  = FALSE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE)
)
plan("sequential")
timings$fb <- (proc.time() - t0)[["elapsed"]]

fb_sum <- summarize_bootstrap_results(
  sgharm = fs$sg.harm, boot_results = fs_bc,
  create_plots = FALSE, est.scale = "hr"
)
cat("FB completed in", round(timings$fb / 60, 1), "min\n")
```

## Multiplier resampling (MR)

The Tier-2 de-biased gate is the MR correction, computed inside the fit above
(vectorized over `r format(gate_draws, big.mark=",")` draws). It is read from
`fs$debias_gate`; no separate run is needed.

```{r mr}
g <- fs$debias_gate
stopifnot(!is.null(g))
timings$mr <- g$timing_seconds
cat("MR gate:", round(timings$mr, 2), "s over",
    format(gate_draws, big.mark = ","), "draws\n")
```

## Guo & He (GH)

The adapter exports the maxeff candidate family as membership columns and runs
Guo & He Algorithm 3 on the same selected subgroup. The membership integrity
check asserts that the exported selected-subgroup column matches forestsearch's
own `sg.harm.id` exactly before the estimator runs.

```{r guohe}
t0 <- proc.time()
gh <- fs_to_guohe(
  fit       = fs,
  time      = "time_months", event = "status", treatment = "hormon",
  B         = GH_B, r = GH_r, level = 0.05, seed = 8316951,
  parallel  = GH_PARALLEL, verify = TRUE
)
timings$gh <- (proc.time() - t0)[["elapsed"]]

cat("GH family size:", length(gh$export$candidates),
    " | selected:", gh$selected,
    " | GH", round(timings$gh / 60, 1), "min\n")
cat(sprintf("GH  naive HR %.3f  debiased HR %.3f  lower bound %.3f\n",
            gh$naive_hr, gh$debiased_hr, gh$bound_hr))
```

## Comparison: four corrections on the selected subgroup {#sec-table}

```{r table}
#| tbl-cap: "De-biased subgroup hazard ratios for the GBSG selected (maxeff) subgroup. Parenthetical values are 95% intervals; Guo & He reports a one-sided lower bound."
#| echo: false
fci <- function(e, l, u, d = 2) {
  ifelse(is.na(e), "\u2014", sprintf("%.*f (%.*f, %.*f)", d, e, d, l, d, u))
}

# Selected-subgroup (H) HR + 95% CI on the HR scale, read straight from the
# actual result objects (verified against str(fs_bc$H_estimates) and
# str(fs$debias_gate); the estimates are unchanged, this is display plumbing):
#   FB bias-corrected: the bootstrap object's H_estimates carries H0 (naive HR)
#     and H2 (the "HR*" infinitesimal-jackknife bias-corrected HR) with
#     H2_lower/H2_upper -- summarize_bootstrap_results() formats these into a gt
#     table and exposes no top-level est_H/lo_H/hi_H fields.
#   MR de-biased gate: fs$debias_gate$debiased$est / lower / upper (there is no
#     g$Hc_estimates; the gate's H-subgroup de-biased HR lives under $debiased).
naive_hr <- gh$naive_hr    # naive is identical across methods (same fit)

tab <- data.frame(
  Method = c("Naive", "Full bootstrap", "Multiplier resampling", "Guo & He (2021)"),
  Estimate = c(
    sprintf("%.2f", naive_hr),
    fci(fs_bc$H_estimates$H2, fs_bc$H_estimates$H2_lower, fs_bc$H_estimates$H2_upper),
    fci(g$debiased$est, g$debiased$lower, g$debiased$upper),
    sprintf("%.2f (LB %.2f)", gh$debiased_hr, gh$bound_hr)
  ),
  check.names = FALSE
)
kbl(tab, booktabs = TRUE, escape = FALSE, align = "ll") |>
  kable_styling()
```

::: {.callout-note}
## Reading the table

Naive is the uncorrected subgroup hazard ratio. Full bootstrap and Multiplier
resampling both target the conditional estimand and report two-sided intervals;
they differ in that FB re-runs the entire selection on each resample (so its
cost scales with the resample count) while MR holds the fitted family fixed and
is vectorized (so its cost does not). Guo & He corrects the winner's-curse
optimism of the argmax selection and reports a de-biased point estimate with a
one-sided lower confidence bound -- the object their Theorems 4/5 and
simulations certify.

Guo & He assumes a fixed candidate family; the maxeff family here is large
(hundreds to ~1744 candidates depending on cut construction) and data-adjacent,
which is outside their stated small-fixed-$k$ regime. The Section 5 reproduction
established that their estimator is faithfully implemented on disjoint families
up to $k = 12$; its behaviour on a family this size is the open question, and is
exactly why the head-to-head with MR is of interest.
:::

```{r timing}
#| echo: false
td <- data.frame(
  Step = c("Fit + MR gate", "Full bootstrap", "MR gate (within fit)",
           "Guo & He bootstrap"),
  Seconds = c(timings$fit, timings$fb, timings$mr, timings$gh)
)
td$Minutes <- round(td$Seconds / 60, 2)
kbl(td, booktabs = TRUE, align = "lrr") |> kable_styling()
```
