---
title: "Consistency screen: `split` vs `resample`"
subtitle: "What the MR approximation inside the search actually buys (GBSG, Cox)"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: true
  message: true
bibliography: []
---

`forestsearch()` offers `consistency_method = c("split", "resample")`.

- **`"split"`** is the literal calculation: repeated 50/50 splits, each requiring
  two refits per candidate.
- **`"resample"`** replaces that with a single subgroup fit plus a multiplier
  representation of the split pair. The per-subject treatment influence
  contributions are the treatment column of the model's `dfbeta`, the half
  effects are $\hat\beta \pm D$ with $D=\sum_i G_i\,\mathrm{dfbeta}_i$, and
  because $\sigma_D=\sqrt{\sum_i \mathrm{dfbeta}_i^2}$ is exactly the robust SE,
  the consistency rate collapses to a closed form,
  $2\Phi\!\left((\hat\beta-c)/\sigma_D\right)-1$.

This is MR machinery operating *inside* the search, so it is inherited by every
bootstrap resample --- the full bootstrap re-runs the whole search per draw.
This document measures what that substitution buys and whether it changes any
conclusion.

Two tiers, independently toggleable:

- **Tier A (cheap, minutes):** a single search under each method. Compares the
  selected subgroup, the candidate/survivor counts, the consistency rates, and
  the naive effects.
- **Tier B (expensive, ~10--25 min per arm):** the full bootstrap under each
  method. This is the number that matters, because the screen runs inside every
  one of the `NB` resamples --- Tier A alone understates the saving by roughly a
  factor of `NB`.

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")
library(forestsearch)
library(survival)
library(future)
timings <- list()
```

```{r config}
n_cores <- max(1L, floor(0.80 * parallel::detectCores(logical = FALSE)))
NB      <- 1000L        # bootstrap draws for Tier B
RUN_TIER_B <- TRUE      # set FALSE to skip the full-bootstrap arms

# Knobs mirrored from analysis_gbsg_cox_multimethod_psi_v2_2new.qmd so the
# comparison sits on the same configuration as the main analysis.
fs_sg_focus             <- "effMaxSG"
fs_hr_threshold         <- 1.0
fs_hr_consistency       <- 1.0
fs_max_subgroups_search <- 50
fs_conf_force           <- c("er <= 0", "pgr <= 0")
fs_subgroup_method      <- "consistency"   # the screen only applies here
run_mr         <- TRUE
mr_draws              <- 5000L
cat("cores:", n_cores, "| NB:", NB, "| Tier B:", RUN_TIER_B, "\n")
```

## Data

```{r data-setup}
df.analysis <- gbsg
df.analysis <- within(df.analysis, {
  id          <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3      <- ifelse(grade == "3", 1, 0)
  treat       <- hormon
})
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"; event.name <- "status"
id.name      <- "id";          treat.name <- "hormon"
cat("n =", nrow(df.analysis), "| events =", sum(df.analysis$status), "\n")
```

## Tier A --- one search under each method

```{r tier-a}
fit_one <- function(method, seed = 8316951) {
  t0 <- proc.time()
  fs <- forestsearch(
    df.analysis,
    outcome.name = outcome.name, event.name = event.name,
    treat.name = treat.name, id.name = id.name,
    confounders.name = confounders.name,
    is.RCT = TRUE, seedit = seed, est.scale = "hr",
    use_lasso = FALSE, use_grf = FALSE, use_dina = FALSE,
    conf_force = fs_conf_force,
    conf.cont_jcuts = list(er = 10, pgr = 10),
    collapse_cuts = TRUE, cont.cutoff = 4,
    subgroup_method = fs_subgroup_method,
    sg_focus = fs_sg_focus,
    selection_rule = "neighborhood", effect_neighborhood = 0.10,
    mr_inference = run_mr,
    mr_inference_args = list(draws = mr_draws),
    n.min = 60, d0.min = 10, d1.min = 10, maxk = 2,
    max_subgroups_search = fs_max_subgroups_search,
    hr.threshold = fs_hr_threshold, hr.consistency = fs_hr_consistency,
    pconsistency.threshold = 0.90,
    consistency_method = method,          # <-- the comparison
    fs.splits = 1000, stop_threshold = 0.95,
    use_twostage = TRUE, max.minutes = 3,
    details = FALSE, quiet = TRUE, plot.sg = FALSE,
    parallel_args = list(plan = "multisession", workers = n_cores,
                         show_message = FALSE)
  )
  plan("sequential")
  list(fs = fs, secs = unname((proc.time() - t0)["elapsed"]))
}

a_split  <- fit_one("split")
a_resamp <- fit_one("resample")
timings$search_split  <- a_split$secs
timings$search_resamp <- a_resamp$secs

descr <- function(o) {
  fs <- o$fs
  sg <- tryCatch(paste(fs$sg.harm, collapse = " & "), error = function(e) NA_character_)
  g  <- fs$mr_inference
  data.frame(
    search_secs = round(o$secs, 2),
    subgroup    = if (length(sg) && !is.na(sg) && nzchar(sg)) sg else "(none)",
    n_family    = tryCatch(as.integer(g$n_family), error = function(e) NA_integer_),
    sel_bias    = tryCatch(round(as.numeric(g$selection_bias), 4), error = function(e) NA_real_),
    mr_secs   = tryCatch(round(as.numeric(g$timing_seconds), 3), error = function(e) NA_real_),
    stringsAsFactors = FALSE
  )
}
tierA <- rbind(cbind(method = "split", descr(a_split)),
               cbind(method = "resample", descr(a_resamp)))
knitr::kable(tierA, caption = "Tier A: single search under each consistency method")
```

```{r tier-a-agreement}
same_sg <- identical(tierA$subgroup[1], tierA$subgroup[2])
cat(sprintf("Same subgroup selected: %s\n", same_sg))
cat(sprintf("Search speed-up (split / resample): %.1fx\n",
            timings$search_split / timings$search_resamp))
if (!same_sg) {
  cat("\nNOTE: the two methods selected DIFFERENT subgroups. That is the",
      "\nsubstantive result -- the approximation changed the search outcome,",
      "\nnot merely its speed. Report it rather than averaging over it.\n")
}
```

## Tier B --- full bootstrap under each method

The screen runs inside every resample, so this is where the substitution pays.
Both arms use the same `NB` and the same worker count, so the ratio is
attributable to the screen.

```{r tier-b}
#| eval: !expr RUN_TIER_B
boot_one <- function(fs_obj) {
  t0 <- proc.time()
  bc <- forestsearch_bootstrap_dofuture(
    fs.est = fs_obj, nb_boots = NB, show_three = FALSE, details = FALSE,
    parallel_args = list(plan = "multisession", workers = n_cores,
                         show_message = FALSE)
  )
  plan("sequential")
  list(bc = bc, secs = unname((proc.time() - t0)["elapsed"]))
}
b_split  <- boot_one(a_split$fs)
b_resamp <- boot_one(a_resamp$fs)
timings$boot_split  <- b_split$secs
timings$boot_resamp <- b_resamp$secs

pull <- function(o) {
  e <- o$bc$H_estimates
  f <- function(n) tryCatch(as.numeric(e[[n]]), error = function(x) NA_real_)
  data.frame(boot_min = round(o$secs / 60, 2),
             naive_H0 = round(f("H0"), 4),
             debiased_H2 = round(f("H2"), 4),
             H2_lo = round(f("H2_lower"), 4), H2_hi = round(f("H2_upper"), 4))
}
tierB <- rbind(cbind(method = "split", pull(b_split)),
               cbind(method = "resample", pull(b_resamp)))
knitr::kable(tierB, caption = "Tier B: full bootstrap under each consistency method")

cat(sprintf("\nBootstrap speed-up (split / resample): %.1fx\n",
            timings$boot_split / timings$boot_resamp))
cat(sprintf("De-biased HR: split %.4f vs resample %.4f (absolute diff %.4f)\n",
            tierB$debiased_H2[1], tierB$debiased_H2[2],
            abs(tierB$debiased_H2[1] - tierB$debiased_H2[2])))
```

```{r tier-b-skipped}
#| eval: !expr (!RUN_TIER_B)
cat("Tier B skipped (RUN_TIER_B = FALSE).\n")
```

## Reading this

The comparison answers two separate questions, and they should not be
collapsed:

1. **Speed.** The Tier B ratio is the honest figure for what the MR
   approximation buys in practice, because the screen is re-run inside every
   bootstrap draw. The Tier A ratio understates it by roughly `NB`.
2. **Fidelity.** Whether `"resample"` reaches the same subgroup and the same
   de-biased estimate as the exact `"split"` calculation. Agreement supports
   treating the approximation as sound; disagreement is a finding about the
   approximation, not a bug.

Both are single-application results. A claim that `"resample"` is a reliable
substitute in general needs replication across simulated datasets --- the same
distinction that applies to every other one-dataset comparison in this project.

```{r session}
#| echo: false
cat("\n--- timings (seconds) ---\n"); print(unlist(timings))
```
