---
title: "Guo & He (2021) comparison: analysis record"
subtitle: "Table 7 (a nested family cut at every observed value): Guo & He at fixed $r$ against MR, identification by the argmax, and the complement/joint capability"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: false
  message: false
bibliography: []
---

```{r setup}
#| include: false
# Analysis record for the Guo & He comparison. The manuscript section is
# drafted separately, from the payload written in "Manuscript payload" below
# (dev/tasks/claude_cc_task_guohe_record_payload_2026-09-11.md). Originally the
# T3 assembly document for the fs-post-selection supplement.
# Task: dev/tasks/claude_cc_task_guohe_supplement_2026-09-09.md section 6, as
# amended by ..._v2.md (A2 for the B6 citations, A5 for the Adaptive B).
#
# TRANSPLANTED from mr_field_vs_guohe.qmd (50d6b641): structure, bundle
# reading, and the Wilson machinery. Never re-authored, never format-converted.
#
# Scope decision D1: t7 only (six cells), with one cross-reference sentence to
# the committed 16-cell record.
#
# Bundle sources, all read-only:
#   B2  guohe_repro_t7_*.rds        (G&H fixed-r, arithmetic only, never re-run)
#       mr_field_vs_guohe_t7_*.rds  (Naive, MR (IJ), MR (field))
#       guohe_adaptive_t7_*.rds     (Adaptive column -- PHASE B, may be absent)
#   B3  mr_field_complement_vs_guohe_t7_*.rds  (T1 -- may be absent)
#   B5  mr_field_vs_guohe_t7_*.rds  (c-hat, p-hat, M_eff, SDs)
#
# RENDERING: render with RStudio's bundled Quarto binary on the executing
# machine and record the binary path used in the render record. Preserve
# embed-resources: true. The HTML is untracked per the GuoHe/*.html convention.
source("mr_vs_guohe_sim.R")

CELLS <- sprintf("t7_beta2_%02d", 0:5)
BETA2 <- (0:5) / 10

have <- function(f) file.exists(f)
read_if <- function(f) if (have(f)) readRDS(f) else NULL

# ---- B2 / B5 inputs: committed, present ------------------------------------
bun <- lapply(CELLS, function(id) readRDS(sprintf("mr_field_vs_guohe_%s.rds", id)))
names(bun) <- CELLS

for (b in bun) {
  g <- b$gate2
  if (g$n_rep_done != g$n_rep_expected || g$n_errored > 0L ||
      g$naive_mismatch > 0L || g$cur_mismatch > 0L) {
    stop("Gate 2 violation in ", b$id, ": done ", g$n_rep_done, "/",
         g$n_rep_expected, ", errored ", g$n_errored, ", naive_mismatch ",
         g$naive_mismatch, ", cur_mismatch ", g$cur_mismatch, call. = FALSE)
  }
}

NREP  <- bun[[1]]$gate2$n_rep_done
DRAWS <- bun[[1]]$draws
R_OUT <- bun[[1]]$field_R_out
R_IN  <- bun[[1]]$field_R_in

# ---- Phase-dependent inputs -------------------------------------------------
# Committed Guo & He reproduction bundles: the source of the naive column and
# every fixed-r column, uniformly across all six cells (ITEM 1). Verified to
# carry r{1..4}_cover/_dist/_bias and naive_point/_lower/_cover on all six.
REPRO <- lapply(CELLS, function(id) readRDS(sprintf("guohe_repro_%s.rds", id)))
names(REPRO) <- CELLS

# Fixed r = 1/12 production bundles (REPORT_guohe_fixedr_t7_2026-09-09.md).
# Present for t7_beta2_00 and t7_beta2_03 only; other cells fall back to the
# committed reproduction's stored r = 1/12 column.
FIXR <- lapply(CELLS, function(id)
  read_if(sprintf("guohe_adaptive_%s_fixedr00833.rds", id)))
names(FIXR) <- CELLS
FIXR_CELLS <- CELLS[!vapply(FIXR, is.null, logical(1))]

CMP <- lapply(CELLS, function(id)
  read_if(sprintf("mr_field_complement_vs_guohe_%s.rds", id)))
names(CMP) <- CELLS
HAVE_COMPLEMENT <- all(!vapply(CMP, is.null, logical(1)))

# The E1 range citations and their read from dev/notes/ were removed on
# 2026-09-11 (see B6, limit 2, and the task document named above).

z95 <- qnorm(0.95)

wilson <- function(x, n, z = 1.959964) {
  p <- x / n
  den <- 1 + z^2 / n
  ctr <- (p + z^2 / (2 * n)) / den
  hw  <- z * sqrt(p * (1 - p) / n + z^2 / (4 * n^2)) / den
  c(lo = ctr - hw, hi = ctr + hw)
}
fmt_cov <- function(x, n) {
  w <- wilson(x, n); sprintf("%.3f (%.3f, %.3f)", x / n, w["lo"], w["hi"])
}
MCSE <- sqrt(0.95 * 0.05 / NREP)
```

# What this section is

A paired comparison of selection-adjusted one-sided lower bounds for
$\gamma_{\hat c}$ — the true effect in the *selected* subgroup — on the Guo & He
(2021) Section 5.2 / Table 7 design: $W \sim \mathrm{Unif}(0, 80)$, a treatment
effect of $\beta_2$ on $\{W \le 30\}$ and $0$ elsewhere, $n = 400$, and a
**nested** candidate family $S(c) = \{W \le c\}$ with $c$ running
over $30$ and every observed $w \in (30, 60]$ — the distinct members of Guo and
He's continuum $c \in [30, 60]$, since the estimate changes only at observed
values; roughly 151 candidates per replicate. Six cells, $\beta_2 = 0.0, \ldots, 0.5$, `r NREP` paired replicates
each, all on the replication's own seeds.

The same three-way comparison across all sixteen replicated cells (Tables 3–5,
Table 6 and Table 7) is recorded in `mr_field_vs_guohe.qmd` and
`REPORT_mr_field_vs_guohe_2026-09-05.md`; this section restricts to Table 7,
the only design whose candidate family is a nested continuum of cutpoints and therefore
the only one that speaks to the post-selection setting of interest (D1).

All values are on the oriented log-hazard-ratio scale. Coverage is one-sided
unless stated. **Margin** means **estimate minus lower bound**, throughout this
document and its companion reports; for the Guo & He columns that is identically
bias $+$ dist, since $\hat\gamma = \gamma + \mathrm{bias}$ and
$L = \gamma - \mathrm{dist}$. Monte Carlo standard error at `r NREP` replicates and coverage
near $0.95$ is $\approx `r sprintf("%.3f", MCSE)`$.

# B2 — Coverage, margin, and point-estimate bias for $\gamma_{\hat c}$

```{r b2-build}
METHODS <- c("Naive", "G&H r=1/3", "G&H r=1/12", "G&H r=1/21", "G&H r=1/30",
             "MR (field)", "MR (IJ)")

# est / SE / one-sided lower / cover per method. G&H "SE" = implied margin
# (deb - low)/z_.95 (the method publishes a bound); MR (field) est = est2,
# lower = beta_deb - q95.
# MARGIN CONVENTION (single definition, used everywhere in this document and in
# the companion reports): margin = estimate - lower bound. For the G&H columns
# this equals bias + dist identically, since est = gamma + bias and
# low = gamma - dist; verified equal to all.equal on all six cells.
method_cols <- function(res, meth, rp, fx) {
  gs <- rp$results$gamma_s
  switch(meth,
    # ITEM 1: naive read from the committed reproduction bundle, all six cells.
    "Naive"      = list(est = rp$results$naive_point,
                        se  = (rp$results$naive_point - rp$results$naive_lower) / z95,
                        low = rp$results$naive_lower, cov = rp$results$naive_cover),
    "MR (IJ)"    = list(est = res$mr_est, se = res$mr_se_ij,
                        low = res$mr_lower_1s, cov = res$mr_cover),
    "MR (field)" = list(est = res$fld_est2, se = res$fld_lambda_sd,
                        low = res$fld_lower_1s, cov = res$fld_cover_1s),
    # r = 1/12 comes from the fixed-r production run where that bundle exists
    # (its OWN Algorithm-3 refit, not the stored lookup), otherwise from the
    # committed reproduction's stored column.
    # ITEM 1: every fixed-r column read from the committed reproduction bundle,
    # uniformly across all six cells, so the head-to-head is on one footing.
    { i <- match(meth, METHODS) - 1L
      rw <- rp$results
      est <- gs + rw[[sprintf("r%d_bias", i)]]
      low <- gs - rw[[sprintf("r%d_dist", i)]]
      list(est = est, se = (est - low) / z95, low = low,
           cov = rw[[sprintf("r%d_cover", i)]]) })
}

cell_summary <- function(id) {
  res <- bun[[id]]$results
  do.call(rbind, lapply(METHODS, function(meth) {
    v <- method_cols(res, meth, REPRO[[id]], FIXR[[id]])
    if (is.null(v)) return(data.frame(
      cell = id, method = meth, mean_est = NA_real_, bias = NA_real_,
      sd_emp = NA_real_, mean_margin = NA_real_, cover = NA_real_,
      cover_lo = NA_real_, cover_hi = NA_real_, n = NA_integer_,
      row.names = NULL))
    n <- length(v$est); w <- wilson(sum(v$cov), n)
    data.frame(
      cell = id, method = meth,
      mean_est = mean(v$est), bias = mean(v$est - REPRO[[id]]$results$gamma_s),
      sd_emp = sd(v$est), mean_margin = mean(v$est - v$low),
      cover = mean(v$cov), cover_lo = w["lo"], cover_hi = w["hi"], n = n,
      row.names = NULL)
  }))
}
SUM <- do.call(rbind, lapply(CELLS, cell_summary))
```

```{r b2-table}
#| output: asis
for (k in seq_along(CELLS)) {
  id <- CELLS[k]
  s <- SUM[SUM$cell == id, ]
  tab <- data.frame(
    Method = s$method,
    `Coverage of the 95% lower bound (Wilson)` = ifelse(
      is.na(s$cover), "—",
      sprintf("%.3f (%.3f, %.3f)", s$cover, s$cover_lo, s$cover_hi)),
    `Mean margin` = ifelse(is.na(s$mean_margin), "—",
                           sprintf("%.3f", s$mean_margin)),
    `Point-estimate bias` = ifelse(is.na(s$bias), "—",
                                   sprintf("%+.4f", s$bias)),
    check.names = FALSE)
  # MR (IJ) is context, not a comparator: gray it.
  tab$Method[tab$Method == "MR (IJ)"] <- "*MR (IJ), context*"
  print(knitr::kable(
    tab, row.names = FALSE,
    caption = sprintf(
      paste("**%s** ($\\beta_2 = %.1f$, %d paired replicates).",
            "Coverage of the one-sided 95%% lower bound for $\\gamma_{\\hat c}$;",
            "mean margin = mean(est $-$ lower); point-estimate bias against",
            "$\\gamma_{\\hat c}$ (naive $\\hat\\beta$; G&H bias-reduced;",
            "MR $\\hat\\beta_2$ = est2)."),
      id, BETA2[k], NREP)))
  cat("\n")
}
```

::: {.callout-note appearance="simple"}
**Source of the naive and Guo & He columns.** Both are read from the **committed
reproduction bundles** `guohe_repro_t7_beta2_0{0..5}.rds`, uniformly across all
six cells, so the head-to-head sits on one footing across the whole $r$ grid.
Those bundles were verified to carry `r{1..4}_cover/_dist/_bias` and
`naive_point/_lower/_cover` on every cell, all at $B = 2000$.

`t7_beta2_00` and `t7_beta2_03` were **additionally confirmed by an
independent-stream production run** at $r = 1/12$ — its own Algorithm-3 refit at
$B = 2000$, $\mathrm{orient} = +1$, 2000 replicates per cell
(`REPORT_guohe_fixedr_t7_2026-09-09.md`). The produced bound reproduces the
stored column **identically** (`identical()` on 2000/2000 replicates in both
cells), and an independent bootstrap stream at the same $B$ reaches the same
coverage decision on 99.0% and 99.7% of replicates. No cell was re-run for this
table.

**Two qualifications on the G&H columns.**

1. The replicated fixed-$r$ columns run $\approx 0.010$ *below* the published
   Table 7 values — a replication deficit recorded on 2026-09-04 and carried
   here unchanged. The published columns are themselves conservative
   ($0.947$–$0.973$ against a nominal $0.95$), and increasingly so as $r$
   falls ($0.947$–$0.964$ at $r = 1/3$; $0.961$–$0.973$ at $r = 1/12$ to
   $1/30$) — the finite-sample behaviour Guo and He describe for small $r$
   (their Section 2.5).
2. Monte Carlo standard error at `r NREP` replicates is
   $\approx `r sprintf("%.3f", MCSE)`$, so differences smaller than about one
   percentage point are not resolvable here.
:::

```{r b2-adaptive-note}
#| output: asis
# The Adaptive column is not produced and is not pending: it was measured on a
# pilot and deliberately not adopted.
cat(paste0(
  "**On the Adaptive column.** The adaptive choice of $r$ was measured on a ",
  "pilot rather than reported: it cost **10.3×** the fixed-$r$ bound ",
  "(356.5 against 34.7 core-seconds per replicate) and selected between ",
  "candidates whose cross-validation objective gaps were **within 2% of the ",
  "objective's own spread** across replicates (median gap 2.18e-03 against a ",
  "spread of $\\approx$ 0.28) — so the fixed-$r$ column at $r = 1/12$ is ",
  "reported instead. A seed-offset test sharpened this: changing only the ",
  "adaptive seed offset flipped $\\hat r$ on 5 of 10 replicates with the data ",
  "and the selection untouched, so $\\hat r$ on this design is substantially ",
  "seed-determined rather than data-driven ",
  "(`REPORT_guohe_adaptive_gate1b_2026-09-09.md`; ",
  "`dev/notes/NOTE_adaptive_seed_parity_2026-09-09.md`).\n\n"))
```

## Consolidated operating characteristics

```{r b2-consolidated}
#| output: asis
KEEP <- c("Naive", "G&H r=1/12", "MR (field)", "MR (IJ)")
con <- do.call(rbind, lapply(seq_along(CELLS), function(k) {
  s6 <- SUM[SUM$cell == CELLS[k] & SUM$method %in% KEEP, ]
  s6 <- s6[match(KEEP, s6$method), ]
  data.frame(
    `$\\beta_2$` = sprintf("%.1f", BETA2[k]),
    Method = ifelse(s6$method == "MR (IJ)", "MR (IJ), context", s6$method),
    Coverage = sprintf("%.3f (%.3f, %.3f)", s6$cover, s6$cover_lo, s6$cover_hi),
    Margin = sprintf("%.3f", s6$mean_margin),
    Bias = sprintf("%+.4f", s6$bias),
    check.names = FALSE, row.names = NULL)
}))
print(knitr::kable(con, row.names = FALSE, caption = paste(
  "Operating characteristics across all six cells: one-sided 95% coverage of",
  "$\\gamma_{\\hat c}$ with Wilson intervals, margin (estimate $-$ lower) and",
  "point-estimate bias, for the naive bound, Guo & He at $r = 1/12$, MR (field)",
  "and MR (IJ) as context. Naive and Guo & He are read from the committed",
  "reproduction bundles on all six cells; MR columns from the paired",
  "`mr_field_vs_guohe_t7_*` bundles on the same replicates.")))
cat("\n")
```

# B3 — Complement and joint bounds {#sec-b3}

::: {.callout-tip appearance="simple"}
## Capability, not comparison

**This section is a capability demonstration, not a comparison.** Nothing in it
is scored against Guo & He, because their framework does not define the
quantities involved. The two paragraphs below say why, and they are the
substance of the section — the numbers merely show that our construction
delivers what it claims on this design.

> Table 8 of Guo and He (2021) reports an upper confidence bound on the
> log-hazard-ratio scale; this is the identical one-sided lower-bound construction
> under their Section 4 convention βᵢ = −log HR, applied to the argmax-selected
> subgroup, and does not constitute an upper-bound capability for a second,
> complementary subgroup.

> The Guo–He correction is defined through the maximum functional over the
> supplied candidate family. The complement Ĥᶜ is not a member of that family
> and is not the argmax of any functional, so no analogous correction exists
> within their framework; the absence of a complement bound, and hence of a
> joint two-subgroup claim, is structural rather than an implementation gap.
:::

On this design the complement's true effect is **exactly 0**: the DGM sets
$b(W) = \beta_2$ on $\{W \le 30\}$ and $0$ elsewhere
(`guohe_sec52_truth.R:75`), and the family starts at $30$, so every
$\hat H^c = \{W > \hat c\}$ lies inside the null region and complement
coverage is simply upper $\ge 0$. The structural argument above concerns this
design's nested family, in which $\hat H^c$ is never a candidate. It is not
claimed for Guo and He's two-subgroup design (their Section 5.1), where the
complement of the selected subgroup is the other candidate — with two
candidates, the argmin — nor for their MONET1 families (their Section 4), which
contain each subgroup and its complement.

```{r b3-table}
#| output: asis
if (!HAVE_COMPLEMENT) {
  cat(paste0(
    "::: {.callout-important appearance='simple'}\n",
    "**Pending: T1 bundles absent.** `mr_field_complement_vs_guohe_t7_*.rds` ",
    "have not been produced, so the field-s complement upper coverage, its ",
    "mean bound location, and the joint coverage are *pending*. The two ",
    "paragraphs above stand on their own and are not contingent on them.\n\n",
    "On this design the complement's true effect is **exactly 0**: the DGM ",
    "sets $b(W) = \\beta_2$ on $\\{W \\le 30\\}$ and $0$ elsewhere ",
    "(`guohe_sec52_truth.R:75`), and the cutpoint grid starts at $30$, so ",
    "every $\\hat H^c = \\{W > \\hat c\\}$ lies inside the null region. ",
    "Complement coverage is therefore $\\text{upper} \\ge 0$, with no ",
    "dilution term.\n",
    ":::\n\n"))
} else {
  tab <- do.call(rbind, lapply(seq_along(CELLS), function(k) {
    r <- CMP[[CELLS[k]]]$results; n <- nrow(r)
    data.frame(
      Cell = CELLS[k], `$\\beta_2$` = sprintf("%.1f", BETA2[k]),
      `Complement upper coverage vs 0, field-s (Wilson)` =
        fmt_cov(sum(r$c_cover_95_s), n),
      `Mean complement bound location, field-s` =
        sprintf("%.4f", mean(r$c_upper_95_s, na.rm = TRUE)),
      `Joint Bonferroni coverage, joint_s (Wilson)` = fmt_cov(sum(r$joint_cover_s), n),
      `Joint, unscaled (Wilson)` = fmt_cov(sum(r$joint_cover), n),
      check.names = FALSE, row.names = NULL)
  }))
  print(knitr::kable(tab, row.names = FALSE, caption = paste(
    "Capability, not comparison. Per cell: the field-s one-sided 95% upper bound",
    "on $\\beta(\\hat H^c)$ against a truth of exactly 0, its mean location, and",
    "the joint (0.975 lower on $\\hat H$, 0.975 upper on $\\hat H^c$) Bonferroni",
    "pair in both scalings — `joint_s` is the documented product, `joint` the",
    "unscaled comparator. Guo & He columns are absent by construction, not by",
    "omission.")))
  cat("\n")
}
```

# B5 — What was selected, and how stably

```{r b5}
#| output: asis
# N7: report the closed forms plus c-hat's distribution and the per-cell means
# of the two NON-degenerate metrics. Specificity and PPV are functions of c-hat
# alone on this design, so they are computed from c-hat, not re-estimated.
# NPV is degenerate by construction and is stated, never tabulated.
spec_of <- function(ch) (80 - ch) / 50      # P(W > c-hat | W > 30)
ppv_of  <- function(ch) 30 / ch             # P(W <= 30 | W <= c-hat)

b5 <- do.call(rbind, lapply(seq_along(CELLS), function(k) {
  r <- bun[[CELLS[k]]]$results
  err_sd <- sd(r$fld_est2 - r$theta)
  data.frame(
    Cell = CELLS[k], `$\\beta_2$` = sprintf("%.1f", BETA2[k]),
    `$\\hat c$ median (IQR)` = sprintf("%.1f (%.1f, %.1f)",
      median(r$c_hat), quantile(r$c_hat, .25), quantile(r$c_hat, .75)),
    `$\\hat c$ min/max` = sprintf("%.1f / %.1f", min(r$c_hat), max(r$c_hat)),
    `Mean specificity` = if (BETA2[k] == 0) "—" else
      sprintf("%.3f", mean(spec_of(r$c_hat))),
    `Mean PPV` = if (BETA2[k] == 0) "—" else
      sprintf("%.3f", mean(ppv_of(r$c_hat))),
    `Mean $\\gamma_{\\hat c}$` = sprintf("%.3f", mean(r$theta)),
    `$\\hat p(\\hat H)$ median (IQR)` = sprintf("%.3f (%.3f, %.3f)",
      median(r$p_hat_H), quantile(r$p_hat_H, .25), quantile(r$p_hat_H, .75)),
    `$M_{\\mathrm{eff}}$ median` = sprintf("%.1f", median(r$M_eff)),
    `Marginal SD` = sprintf("%.3f", sd(r$fld_est2)),
    `Error SD` = sprintf("%.3f", err_sd),
    check.names = FALSE, row.names = NULL)
}))
print(knitr::kable(b5, row.names = FALSE, caption = paste(
  "Selection and stability per cell. Specificity $= (80 - \\hat c)/50$ and",
  "PPV $= 30/\\hat c$ are exact functions of $\\hat c$ on this design and are",
  "computed as such; sensitivity and NPV are degenerate at 1 by nesting and are",
  "not tabulated. At $\\beta_2 = 0$ there is no harm region, so specificity and",
  "PPV are shown as —. Mean $\\gamma_{\\hat c}$ is the true effect in the",
  "selected subgroup, averaged over replicates.",
  "Marginal SD is $\\mathrm{sd}(\\hat\\beta_2)$ across replicates;",
  "error SD is $\\mathrm{sd}(\\hat\\beta_2 - \\gamma_{\\hat c})$ — they differ",
  "because $\\gamma_{\\hat c}$ itself moves with the selection.")))
cat("\n")
```

::: {.callout-note appearance="simple"}
**On this design identification has one dimension, and its arithmetic says
why.** The true harm region is $H = \{W \le 30\}$ and every candidate is
$S(c) = \{W \le c\}$ with $\hat c \ge 30$, so $H = S(30) \subseteq S(\hat c)$
for every replicate. With $W \sim \mathrm{Unif}(0, 80)$ the classification
operating characteristics are therefore closed forms in $\hat c$ alone:

- **Sensitivity $\equiv 1$** and **NPV $\equiv 1$** (for $\beta_2 > 0$) —
  *identically*, by nesting, not by merit. Both are degenerate by construction
  on this design, so neither is tabulated: a column of ones would read as a
  result when it is an artefact of $H \subseteq S(\hat c)$.
- **Specificity** $= (80 - \hat c)/50$ — reported per cell above.
- **PPV** $= 30/\hat c$ — reported per cell above.

The whole of the identification story is how far $\hat c$ overshoots $30$,
which the $\hat c$ distribution above reports directly and which specificity
and PPV then restate as monotone functions of it. Two further readings follow
from the same arithmetic. At $\beta_2 = 0$ there is no harm region: every
selection is a false identification, specificity and PPV are not defined as
accuracy there, and that cell's $\hat c$ distribution is the reference for the
argmax's own pull toward the smallest candidate, $c = 30$ — where the true
boundary of the other cells sits. And overshoot reaches the estimand:
$\gamma_{\hat c} = \beta(\hat c)$ is a weighted average of $b(\cdot)$ over
$[0, \hat c]$ (Guo and He, Section 5.2), so it falls below $\beta_2$ as
$\hat c$ overshoots, which the mean $\gamma_{\hat c}$ column shows.
:::

```{r b5-chat-plot}
#| fig-width: 9
#| fig-height: 4.2
op <- par(mar = c(4, 4, 2, 1))
cl <- lapply(CELLS, function(id) bun[[id]]$results$c_hat)
boxplot(cl, names = sprintf("%.1f", BETA2), outline = FALSE,
        xlab = expression(beta[2]), ylab = expression(hat(c)),
        main = expression("Selected cutpoint " * hat(c) * " by " * beta[2]))
abline(h = 30, lty = 3)
mtext("dotted line: the true boundary c = 30", side = 3, line = -1.2, cex = 0.8)
par(op)
```

# B6 — Honest limits

```{r b6}
#| output: asis
# The field's coverage figures are COMPUTED from the rendered bundles, never
# transcribed, so this paragraph cannot go stale against the tables above.
# Rewritten 2026-09-11 (dev/tasks/claude_cc_task_guohe_record_payload_2026-09-11.md):
# limit 1 is grounded on this design's own records (its earlier regime
# attribution came from campaign e1stud, not from this design); the E1 range
# citations and the dev/notes read are removed; limit 3 acknowledges Guo & He
# Section 6.
fld_cov   <- SUM$cover[SUM$method == "MR (field)"]
fld_bias  <- SUM$bias[SUM$method == "MR (field)"]
fld_lo3   <- sort(fld_cov)[1:3]
fld_min   <- sprintf("%.3f", min(fld_lo3))
fld_range <- sprintf("%.3f–%.3f", min(fld_lo3), max(fld_lo3))
fld_bias_range <- sprintf("%+.3f to %+.3f", min(fld_bias), max(fld_bias))
stopifnot(all(fld_bias > 0))  # limit 1 states the bias is positive in every cell
phat_med   <- vapply(CELLS, function(id) median(bun[[id]]$results$p_hat_H), numeric(1))
phat_range <- sprintf("%.3f–%.3f", min(phat_med), max(phat_med))
.cu <- vapply(CMP, function(b) mean(b$results$c_cover_95_s), numeric(1))
stopifnot(BETA2[which.min(.cu)] == 0)  # limit 1 places the complement's low point at the null cell
cmp_low <- sprintf("%.3f", min(.cu))

cat(sprintf(paste0(
"Five limits belong on the record, stated plainly.\n\n",

"1. **The field's low points are under-correction on a nested family.** ",
"Across these cells the MR (field) one-sided coverage **bottoms out at %s**, ",
"with the three lowest cells spanning %s, and its point estimate est2 keeps a ",
"positive bias in every cell (%s). The committed 16-cell record reads the ",
"nested family as the hardest case for the second-order correction, ",
"consistent with the first proof-of-concept's reading that the excess is a ",
"property of unequal candidate variances under nesting ",
"(`REPORT_mr_field_vs_guohe_2026-09-05.md:34`). The complement bound's low ",
"point, %s, is the null cell, where the selection is loosest ",
"(`REPORT_mr_complement_vs_guohe_2026-09-09.md:174`), and re-selection is ",
"spread thin throughout (median $\\hat p(\\hat H)$ %s; B5): these shortfalls ",
"sit where the pick is least settled. The design is also Guo and He's own ",
"ground. It illustrates their continuum results (their Section 3, Theorems ",
"4–5), whereas its ~151 candidates, each differing from its neighbour by one ",
"subject, are the near-duplicate case that strains the no-tie condition (A5) ",
"behind our construction's theory (supplement Section S5, \"No-tie ",
"regularity\"). The field's calibration here is an empirical finding on this ",
"design, not a consequence of that theory.\n\n",

"2. **The complement and joint figures in @sec-b3 belong to this design ",
"alone.** They say nothing about forestsearch's own screened identifier, ",
"which is evaluated separately (campaign `e1stud`, ",
"`REVIEW_E1_fields_2026-09-08.md`, whose pre-registered complement criterion ",
"was not met, `:10`).\n\n",

"3. **Guo & He is not run on forestsearch's screened family.** Their method ",
"is applied exactly as published, on the family their Section 5.2 specifies. ",
"Guo and He suggest (their Section 6) that when identification involves other ",
"considerations their bound serves as a conservative approach; that ",
"suggestion is not tested here, and no claim is made either way.\n\n",

"4. **Everything here is one-sided.** Every construction compared in this ",
"section is one-sided, so none of it depends on the choice of two-sided ",
"interval method.\n\n",

"5. **Identification is flattered by construction.** The true boundary is ",
"the family's smallest member, toward which the argmax is drawn even at ",
"$\\beta_2 = 0$; ",
"sensitivity and NPV are identically 1 for $\\beta_2 > 0$, and overshoot is ",
"the only identification error (B5). The screen and forestsearch's family ",
"construction are not exercised, the complement's truth is 0 in every cell ",
"(so a benefit claim is never tested), and everything is at $n = 400$.\n"),
fld_min, fld_range, fld_bias_range, cmp_low, phat_range))
```

# Manuscript payload

The manuscript section is drafted from this payload, not from the tables above.
It is a superset, so trimming the section never forces a rebuild, and every
element is computed here from the committed bundles: nothing is re-run and
nothing is typed. Bounds are on the oriented log-hazard-ratio scale; each `_hr`
column is $\exp$ of the corresponding mean log-scale bound.

```{r export-payload}
#| output: asis
# TASK dev/tasks/claude_cc_task_guohe_record_payload_2026-09-11.md.
# Location pattern transplanted from the export-payload chunk of
# quarto/applications/actg175/analysis_actg175_continuous_oc.qmd:
# <qmd dir>/_payloads/<stem>/<stem>_payload.rds.
.qmd_dir  <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) getwd())
.qmd_stem <- tryCatch(tools::file_path_sans_ext(basename(knitr::current_input())),
                      error = function(e) "guohe_supp_section")
.out_dir  <- file.path(.qmd_dir, "_payloads", .qmd_stem)
dir.create(.out_dir, recursive = TRUE, showWarnings = FALSE)
.payload_file <- file.path(.out_dir, paste0(.qmd_stem, "_payload.rds"))

# (1) The bound for gamma_c-hat on t7: SUM (B2, unchanged) plus the mean bound.
.low <- do.call(rbind, lapply(CELLS, function(id) data.frame(
  cell = id, method = METHODS,
  mean_lower = vapply(METHODS, function(m)
    mean(method_cols(bun[[id]]$results, m, REPRO[[id]], FIXR[[id]])$low),
    numeric(1)),
  row.names = NULL)))
bound_t7 <- merge(SUM, .low, by = c("cell", "method"), sort = FALSE)
bound_t7 <- bound_t7[order(match(bound_t7$cell, CELLS),
                           match(bound_t7$method, METHODS)), ]
bound_t7$beta2 <- BETA2[match(bound_t7$cell, CELLS)]
bound_t7$mean_lower_hr <- exp(bound_t7$mean_lower)
rownames(bound_t7) <- NULL

# (2) Identification on t7 (H = {W <= 30}; no harm region at beta2 = 0).
ident_t7 <- do.call(rbind, lapply(seq_along(CELLS), function(k) {
  r <- bun[[CELLS[k]]]$results; ch <- r$c_hat
  data.frame(cell = CELLS[k], beta2 = BETA2[k],
    chat_min = min(ch), chat_q25 = unname(quantile(ch, 0.25)),
    chat_med = median(ch), chat_q75 = unname(quantile(ch, 0.75)),
    chat_max = max(ch), p_chat_30 = mean(abs(ch - 30) < 1e-9),
    spec_mean = if (BETA2[k] == 0) NA_real_ else mean((80 - ch) / 50),
    ppv_mean  = if (BETA2[k] == 0) NA_real_ else mean(30 / ch),
    n_sel_mean = mean(r$n_sel), n_sel_med = median(r$n_sel),
    gamma_mean = mean(r$theta), dilution = BETA2[k] - mean(r$theta),
    phat_med = median(r$p_hat_H), phat_q25 = unname(quantile(r$p_hat_H, 0.25)),
    phat_q75 = unname(quantile(r$p_hat_H, 0.75)), phat_mean = mean(r$p_hat_H),
    M_eff_med = median(r$M_eff),
    sd_marg = sd(r$fld_est2), sd_err = sd(r$fld_est2 - r$theta),
    row.names = NULL)
}))

# (3) Complement and joint on t7 (complement truth exactly 0: cover = upper >= 0).
#     naive = cgate_naive_est + z95 * cgate_se_wald, the engine's own naive
#     convention (se_wald = the candidate SE; R/fs_mr_inference.R naive block);
#     IJ two-term = cgate_est + z95 * cgate_se_ij, mirroring the harm side's
#     lower_1s. Definitions are quoted in the Stage 0 record.
.up_row <- function(u, n) { x <- sum(u >= 0); w <- wilson(x, n)
  c(cover = x / n, lo = unname(w["lo"]), hi = unname(w["hi"])) }
comp_t7 <- do.call(rbind, lapply(seq_along(CELLS), function(k) {
  r <- CMP[[CELLS[k]]]$results; n <- nrow(r)
  up <- list(naive   = r$cgate_naive_est + z95 * r$cgate_se_wald,
             ij      = r$cgate_est + z95 * r$cgate_se_ij,
             field   = r$c_upper_95,
             field_s = r$c_upper_95_s)
  do.call(rbind, lapply(names(up), function(m) {
    cv <- .up_row(up[[m]], n)
    data.frame(cell = CELLS[k], beta2 = BETA2[k], method = m,
      cover = cv[["cover"]], cover_lo = cv[["lo"]], cover_hi = cv[["hi"]],
      mean_upper = mean(up[[m]]), mean_upper_hr = exp(mean(up[[m]])),
      row.names = NULL)
  }))
}))
joint_t7 <- do.call(rbind, lapply(seq_along(CELLS), function(k) {
  r <- CMP[[CELLS[k]]]$results; n <- nrow(r)
  ws <- wilson(sum(r$joint_cover_s), n); wu <- wilson(sum(r$joint_cover), n)
  data.frame(cell = CELLS[k], beta2 = BETA2[k],
    joint_s = mean(r$joint_cover_s), joint_s_lo = unname(ws["lo"]),
    joint_s_hi = unname(ws["hi"]),
    joint = mean(r$joint_cover), joint_lo = unname(wu["lo"]),
    joint_hi = unname(wu["hi"]),
    corr_mean = mean(r$joint_corr), corr_s_mean = mean(r$joint_corr_s),
    pair_lower_H = mean(r$h_lower_975),
    pair_lower_H_hr = exp(mean(r$h_lower_975)),
    pair_upper_Hc_s = mean(r$c_upper_975_s),
    pair_upper_Hc_s_hr = exp(mean(r$c_upper_975_s)),
    row.names = NULL)
}))

# (3b) The naive convention on the harm side: the section's naive row is the
#      replication's (Cox SE); the complement's naive row uses the engine's SE.
#      Both conventions on the harm side, so the difference is on record.
naive_convention <- do.call(rbind, lapply(seq_along(CELLS), function(k) {
  r <- bun[[CELLS[k]]]$results
  data.frame(cell = CELLS[k], beta2 = BETA2[k],
    cover_replication_se = mean(r$naive_cover),
    cover_engine_se = mean(r$naive_est - z95 * r$mr_se_wald <= r$theta),
    mean_se_replication = mean(r$naive_se), mean_se_engine = mean(r$mr_se_wald),
    row.names = NULL)
}))

# (4) The committed 16-cell paired record, harm side only, at r = 1/12 and 1/30.
#     Column reads transplanted from mr_field_vs_guohe.qmd (method_cols).
P16 <- c(sprintf("t35_beta2_%02d", 0:5), sprintf("t6_k%02d", c(2L, 6L, 10L, 12L)),
         CELLS)
.b16 <- c(lapply(stats::setNames(P16[1:10], P16[1:10]), function(id)
  readRDS(sprintf("mr_field_vs_guohe_%s.rds", id))), bun)
.m16 <- c("Naive", "G&H r=1/12", "G&H r=1/30", "MR (field)", "MR (IJ)")
.c16 <- function(res, meth) switch(meth,
  "Naive"      = list(est = res$naive_est, low = res$naive_lower, cov = res$naive_cover),
  "G&H r=1/12" = list(est = res$gh_r2_deb, low = res$gh_r2_low, cov = res$gh_r2_cov),
  "G&H r=1/30" = list(est = res$gh_r4_deb, low = res$gh_r4_low, cov = res$gh_r4_cov),
  "MR (field)" = list(est = res$fld_est2,  low = res$fld_lower_1s, cov = res$fld_cover_1s),
  "MR (IJ)"    = list(est = res$mr_est,    low = res$mr_lower_1s,  cov = res$mr_cover))
pre16 <- do.call(rbind, lapply(P16, function(id) {
  res <- .b16[[id]]$results
  do.call(rbind, lapply(.m16, function(meth) {
    v <- .c16(res, meth); n <- length(v$est); w <- wilson(sum(v$cov), n)
    data.frame(cell = id, method = meth, cover = mean(v$cov),
      cover_lo = unname(w["lo"]), cover_hi = unname(w["hi"]),
      margin = mean(v$est - v$low), bias = mean(v$est - res$theta), n = n,
      row.names = NULL)
  }))
}))
pre16$design <- ifelse(grepl("^t35", pre16$cell), "Tables 3-5: two complementary subgroups",
                ifelse(grepl("^t6", pre16$cell), "Table 6: k disjoint null subgroups",
                       "Table 7: nested continuum"))

payload <- list(
  bound_t7 = bound_t7, ident_t7 = ident_t7, comp_t7 = comp_t7,
  joint_t7 = joint_t7, naive_convention = naive_convention, pre16 = pre16,
  meta = list(
    nrep = NREP, mcse = MCSE, draws = DRAWS, field_R_out = R_OUT,
    field_R_in = R_IN, r_fixed = "1/12 (G&H r=1/12: r2 / gh_r2 columns)",
    scale = "oriented log-HR; *_hr = exp(mean log-scale bound)",
    complement_truth = 0,
    head_at_render = tryCatch(system("git rev-parse HEAD", intern = TRUE),
                              error = function(e) NA_character_),
    inputs_md5 = tools::md5sum(c(sprintf("mr_field_vs_guohe_%s.rds", P16),
                                 sprintf("guohe_repro_%s.rds", CELLS),
                                 sprintf("mr_field_complement_vs_guohe_%s.rds", CELLS))),
    built_at = Sys.time(), R = R.version.string,
    forestsearch_version = tryCatch(as.character(utils::packageVersion("forestsearch")),
                                    error = function(e) NA_character_)))
saveRDS(payload, .payload_file)

for (nm in c("bound_t7", "ident_t7", "comp_t7", "joint_t7", "naive_convention",
             "pre16")) {
  print(knitr::kable(payload[[nm]], digits = 4, row.names = FALSE,
                     caption = sprintf("Payload element `%s`.", nm)))
  cat("\n")
}
cat("Payload written to `", file.path("_payloads", .qmd_stem,
    paste0(.qmd_stem, "_payload.rds")), "`.\n\n", sep = "")
```


# Provenance

```{r provenance}
prov <- data.frame(
  Item = c("Cells", "Replicates per cell", "MR draws", "Field $R_{out}$ / $R_{in}$",
           "MCSE at coverage 0.95", "Adaptive column", "Complement / joint (B3)",
           "Manuscript payload"),
  Value = c(
    paste(CELLS, collapse = ", "),
    format(NREP),
    sprintf("%d (%s)", DRAWS, bun[[1]]$multiplier),
    sprintf("%d / %d", R_OUT, R_IN),
    sprintf("%.4f", MCSE),
    if (length(FIXR_CELLS)) sprintf("not adopted; fixed r = 1/12 on %s",
      paste(FIXR_CELLS, collapse = ", ")) else "not adopted",
    if (HAVE_COMPLEMENT) "present" else "PENDING (T1 bundles absent)",
    file.path("_payloads", .qmd_stem, paste0(.qmd_stem, "_payload.rds"))),
  check.names = FALSE)
knitr::kable(prov, row.names = FALSE, caption = paste(
  "Inputs and phase state. Every bundle is read-only; the G&H fixed-$r$",
  "columns are arithmetic on the stored replication bundles and are never",
  "re-run."))
```

```{r provenance-platform}
#| output: asis
# N7 item 3: one provenance sentence on the two build platforms. The N3 tally is
# reported here and, being zero, closes the matter -- no further caveat appears
# anywhere else in this section.
if (HAVE_COMPLEMENT) {
  n3 <- sum(vapply(CMP, function(b) b$gate2$sel_mismatch, numeric(1)))
  n3n <- sum(vapply(CMP, function(b) b$gate2$n_rep_done, numeric(1)))
  wa <- max(vapply(CMP, function(b) b$gate2$worst_abs, numeric(1)))
  cat(sprintf(paste0(
    "The complement and joint bundles in @sec-b3 were produced on arm64 with ",
    "Accelerate, while the head-to-head bundles behind the B2 tables were ",
    "produced on x86_64 with reference BLAS; across every column with a ",
    "counterpart in both, the worst absolute deviation is %.2e (all within ",
    "`all.equal` at 1e-8), and the selected cutpoint agreed on %d of %d ",
    "replicates.\n\n"), wa, n3n - n3, n3n))
  if (n3 > 0) cat(sprintf(paste0(
    "::: {.callout-warning appearance='simple'}\n",
    "**%d selection mismatch(es)** between the two builds. Every comparison ",
    "involving the complement columns is affected and is caveated accordingly.\n",
    ":::\n\n"), n3))
}
```
