---
title: "Guo & He (2021) comparison for the post-selection supplement"
subtitle: "Table 7 (nested, data-built family): the fixed-$r$ and Adaptive columns against MR, and the complement/joint capability"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: false
  message: false
bibliography: []
---

```{r setup}
#| include: false
# T3 assembly document for the fs-post-selection supplement.
# Task: dev/tasks/claude_cc_task_guohe_supplement_2026-09-09.md section 6, as
# amended by ..._v2.md (A2 for the B6 citations, A5 for the Adaptive B).
#
# TRANSPLANTED from mr_field_vs_guohe.qmd (50d6b641): structure, bundle
# reading, and the Wilson machinery. Never re-authored, never format-converted.
#
# Scope decision D1: t7 only (six cells), with one cross-reference sentence to
# the committed 16-cell record.
#
# Bundle sources, all read-only:
#   B2  guohe_repro_t7_*.rds        (G&H fixed-r, arithmetic only, never re-run)
#       mr_field_vs_guohe_t7_*.rds  (Naive, MR (IJ), MR (field))
#       guohe_adaptive_t7_*.rds     (Adaptive column -- PHASE B, may be absent)
#   B3  mr_field_complement_vs_guohe_t7_*.rds  (T1 -- may be absent)
#   B5  mr_field_vs_guohe_t7_*.rds  (c-hat, p-hat, M_eff, SDs)
#
# RENDERING: render with RStudio's bundled Quarto binary on the executing
# machine and record the binary path used in the render record. Preserve
# embed-resources: true. The HTML is untracked per the GuoHe/*.html convention.
source("mr_vs_guohe_sim.R")

CELLS <- sprintf("t7_beta2_%02d", 0:5)
BETA2 <- (0:5) / 10

have <- function(f) file.exists(f)
read_if <- function(f) if (have(f)) readRDS(f) else NULL

# ---- B2 / B5 inputs: committed, present ------------------------------------
bun <- lapply(CELLS, function(id) readRDS(sprintf("mr_field_vs_guohe_%s.rds", id)))
names(bun) <- CELLS

for (b in bun) {
  g <- b$gate2
  if (g$n_rep_done != g$n_rep_expected || g$n_errored > 0L ||
      g$naive_mismatch > 0L || g$cur_mismatch > 0L) {
    stop("Gate 2 violation in ", b$id, ": done ", g$n_rep_done, "/",
         g$n_rep_expected, ", errored ", g$n_errored, ", naive_mismatch ",
         g$naive_mismatch, ", cur_mismatch ", g$cur_mismatch, call. = FALSE)
  }
}

NREP  <- bun[[1]]$gate2$n_rep_done
DRAWS <- bun[[1]]$draws
R_OUT <- bun[[1]]$field_R_out
R_IN  <- bun[[1]]$field_R_in

# ---- Phase-dependent inputs -------------------------------------------------
ADA <- lapply(CELLS, function(id) read_if(sprintf("guohe_adaptive_%s.rds", id)))
names(ADA) <- CELLS
HAVE_ADAPTIVE <- all(!vapply(ADA, is.null, logical(1)))

CMP <- lapply(CELLS, function(id)
  read_if(sprintf("mr_field_complement_vs_guohe_%s.rds", id)))
names(CMP) <- CELLS
HAVE_COMPLEMENT <- all(!vapply(CMP, is.null, logical(1)))

# ---- certified ranges: cite the record that EXISTS ---------------------------
# REVIEW_certification_2026-09-09.md is in no reachable commit (verified against
# find, git ls-files and git log --all at merge f221f75e), so a "pending sync"
# marker would be misleading -- it implies a document that is merely late.
# The certified ranges actually on record are in REVIEW_E1_fields_2026-09-08.md:
#   :78  | Ĥᶜ (upper) | **field-s** | **0.912 / 0.921 / 0.929** | ...
#        Standard table 2 is headed "across cells (six cells)" (:69) and this
#        column is "1s cov min / mean / max", so 0.912-0.929 is the min-to-max
#        across ALL SIX cells. No subset.
#   :86  The sentence is prefixed "Band cells:", so its four-value lists are the
#        four epsilon cells: joint_s 0.942 / 0.939 / 0.949 / 0.951-0.952 vs
#        joint 0.932 / 0.933 / 0.935 / 0.935. The trailing clause
#        "ends unchanged (0.939 vs 0.941-0.942; 0.930 vs 0.929-0.930)" adds the
#        two END cells (maxSG HR 1.75, minSG HR 1.75), following the sentence's
#        own "joint_s ... vs joint ..." order: joint_s ends are 0.939 and 0.930.
#        => joint_s across ALL SIX cells is 0.930-0.952; the four band cells
#           alone span the narrower 0.939-0.952. Citing 0.939-0.952 as the
#           certified range would silently drop the 0.930 end cell.
#        (Under the opposite reading of that clause the low end would be 0.929
#         rather than 0.930; either way it lies below the band-cell minimum.)
# The wider 0.912-0.960 / 0.939-0.963 quoted in the task documents appear only
# in an uncommitted handoff and are not reproducible from this repository.
CERT_FILE  <- "REVIEW_E1_fields_2026-09-08.md"
CERT_PATH  <- file.path("..", "..", "dev", "notes", CERT_FILE)
CERT_PRESENT <- file.exists(CERT_PATH)
cert_upper <- sprintf("0.912-0.929 (`%s:78`)", CERT_FILE)
cert_joint <- sprintf("0.930-0.952 (`%s:86`)", CERT_FILE)

z95 <- qnorm(0.95)

wilson <- function(x, n, z = 1.959964) {
  p <- x / n
  den <- 1 + z^2 / n
  ctr <- (p + z^2 / (2 * n)) / den
  hw  <- z * sqrt(p * (1 - p) / n + z^2 / (4 * n^2)) / den
  c(lo = ctr - hw, hi = ctr + hw)
}
fmt_cov <- function(x, n) {
  w <- wilson(x, n); sprintf("%.3f (%.3f, %.3f)", x / n, w["lo"], w["hi"])
}
MCSE <- sqrt(0.95 * 0.05 / NREP)
```

# What this section is

A paired comparison of selection-adjusted one-sided lower bounds for
$\gamma_{\hat c}$ — the true effect in the *selected* subgroup — on the Guo & He
(2021) Section 5.2 / Table 7 design: $W \sim \mathrm{Unif}(0, 80)$, a treatment
effect of $\beta_2$ on $\{W \le 30\}$ and $0$ elsewhere, $n = 400$, and a
**nested, data-built** candidate family $S(c) = \{W \le c\}$ with $c$ running
over $30$ and every observed $w \in (30, 60]$ — roughly 151 candidates per
replicate. Six cells, $\beta_2 = 0.0, \ldots, 0.5$, `r NREP` paired replicates
each, all on the replication's own seeds.

The same three-way comparison across all sixteen replicated cells (Tables 3–5,
Table 6 and Table 7) is recorded in `mr_field_vs_guohe.qmd` and
`REPORT_mr_field_vs_guohe_2026-09-05.md`; this section restricts to Table 7,
the only design whose candidate family is nested and data-built and therefore
the only one that speaks to the post-selection setting of interest (D1).

All values are on the oriented log-hazard-ratio scale. Coverage is one-sided
unless stated. Monte Carlo standard error at `r NREP` replicates and coverage
near $0.95$ is $\approx `r sprintf("%.3f", MCSE)`$.

# B2 — Coverage, margin, and point-estimate bias for $\gamma_{\hat c}$

```{r b2-build}
METHODS <- c("Naive", "G&H r=1/3", "G&H r=1/12", "G&H r=1/21", "G&H r=1/30",
             "Adaptive", "MR (field)", "MR (IJ)")

# est / SE / one-sided lower / cover per method. G&H "SE" = implied margin
# (deb - low)/z_.95 (the method publishes a bound); MR (field) est = est2,
# lower = beta_deb - q95.
method_cols <- function(res, meth, ada) {
  switch(meth,
    "Naive"      = list(est = res$naive_est, se = res$naive_se,
                        low = res$naive_lower, cov = res$naive_cover),
    "MR (IJ)"    = list(est = res$mr_est, se = res$mr_se_ij,
                        low = res$mr_lower_1s, cov = res$mr_cover),
    "MR (field)" = list(est = res$fld_est2, se = res$fld_lambda_sd,
                        low = res$fld_lower_1s, cov = res$fld_cover_1s),
    "Adaptive"   = if (is.null(ada)) NULL else
                     list(est = ada$results$gamma_s - ada$results$ad_dist_primary +
                            ada$results$ad_bias_primary,
                          se = ada$results$ad_dist_primary / z95,
                          low = ada$results$ad_lower_primary,
                          cov = ada$results$ad_cover_primary),
    { i <- match(meth, METHODS) - 1L
      list(est = res[[sprintf("gh_r%d_deb", i)]],
           se  = (res[[sprintf("gh_r%d_deb", i)]] - res[[sprintf("gh_r%d_low", i)]]) / z95,
           low = res[[sprintf("gh_r%d_low", i)]],
           cov = res[[sprintf("gh_r%d_cov", i)]]) })
}

cell_summary <- function(id) {
  res <- bun[[id]]$results
  do.call(rbind, lapply(METHODS, function(meth) {
    v <- method_cols(res, meth, ADA[[id]])
    if (is.null(v)) return(data.frame(
      cell = id, method = meth, mean_est = NA_real_, bias = NA_real_,
      sd_emp = NA_real_, mean_margin = NA_real_, cover = NA_real_,
      cover_lo = NA_real_, cover_hi = NA_real_, n = NA_integer_,
      row.names = NULL))
    n <- length(v$est); w <- wilson(sum(v$cov), n)
    data.frame(
      cell = id, method = meth,
      mean_est = mean(v$est), bias = mean(v$est - res$theta),
      sd_emp = sd(v$est), mean_margin = mean(v$est - v$low),
      cover = mean(v$cov), cover_lo = w["lo"], cover_hi = w["hi"], n = n,
      row.names = NULL)
  }))
}
SUM <- do.call(rbind, lapply(CELLS, cell_summary))
```

```{r b2-table}
#| output: asis
for (k in seq_along(CELLS)) {
  id <- CELLS[k]
  s <- SUM[SUM$cell == id, ]
  tab <- data.frame(
    Method = s$method,
    `Coverage of the 95% lower bound (Wilson)` = ifelse(
      is.na(s$cover), "*pending Phase B*",
      sprintf("%.3f (%.3f, %.3f)", s$cover, s$cover_lo, s$cover_hi)),
    `Mean margin` = ifelse(is.na(s$mean_margin), "—",
                           sprintf("%.3f", s$mean_margin)),
    `Point-estimate bias` = ifelse(is.na(s$bias), "—",
                                   sprintf("%+.4f", s$bias)),
    check.names = FALSE)
  # MR (IJ) is context, not a comparator: gray it.
  tab$Method[tab$Method == "MR (IJ)"] <- "*MR (IJ), context*"
  print(knitr::kable(
    tab, row.names = FALSE,
    caption = sprintf(
      paste("**%s** ($\\beta_2 = %.1f$, %d paired replicates).",
            "Coverage of the one-sided 95%% lower bound for $\\gamma_{\\hat c}$;",
            "mean margin = mean(est $-$ lower); point-estimate bias against",
            "$\\gamma_{\\hat c}$ (naive $\\hat\\beta$; G&H bias-reduced;",
            "MR $\\hat\\beta_2$ = est2)."),
      id, BETA2[k], NREP)))
  cat("\n")
}
```

::: {.callout-note appearance="simple"}
**Two qualifications on the G&H columns.**

1. The replicated fixed-$r$ columns run $\approx 0.010$ *below* the published
   Table 7 values — a replication deficit recorded on 2026-09-04 and carried
   here unchanged. The published columns are themselves conservative
   ($0.947$–$0.973$ against a nominal $0.95$), because adjacent nested
   candidates differ by a single subject and the effective number of
   independent comparisons is far below the family size.
2. Monte Carlo standard error at `r NREP` replicates is
   $\approx `r sprintf("%.3f", MCSE)`$, so differences smaller than about one
   percentage point are not resolvable here.
:::

```{r b2-adaptive-note}
#| output: asis
if (!HAVE_ADAPTIVE) cat(paste0(
  "::: {.callout-important appearance='simple'}\n",
  "**Adaptive column: pending Phase B.** `guohe_adaptive_t7_*.rds` are not ",
  "present. The Adaptive row renders as *pending Phase B* above and the ",
  "column is excluded from every comparison in this document. It is produced ",
  "by `guohe_sec52_adaptive_run.R` on the Linux box under Gate 1b, at ",
  "$B = 2000$ (A5), $r$-grid $\\{1/3, 1/12, 1/21, 1/30\\}$, $v = 5$, ",
  "$\\mathrm{orient} = +1$.\n\n",
  "Two caveats will accompany it when it lands: Guo & He's own Table 6 ",
  "caution on the adaptive procedure, and the fact that `guohe_adaptive_r()` ",
  "draws **independently across $r$** (no common random numbers on the grid), ",
  "which inflates $\\mathrm{Var}(\\hat r)$ — and the Adaptive column is ",
  "precisely a measurement of $\\hat r$.\n",
  ":::\n\n"))
```

# B3 — Complement and joint bounds {#sec-b3}

::: {.callout-tip appearance="simple"}
## Capability, not comparison

**This section is a capability demonstration, not a comparison.** Nothing in it
is scored against Guo & He, because their framework does not define the
quantities involved. The two paragraphs below say why, and they are the
substance of the section — the numbers merely show that our construction
delivers what it claims on this design.

> Table 8 of Guo and He (2021) reports an upper confidence bound on the
> hazard-ratio scale; this is the identical one-sided lower-bound construction
> under their Section 4 convention βᵢ = −log HR, applied to the argmax-selected
> subgroup, and does not constitute an upper-bound capability for a second,
> complementary subgroup.

> The Guo–He correction is defined through the maximum functional over the
> supplied candidate family. The complement Ĥᶜ is not a member of that family
> and is not the argmax of any functional, so no analogous correction exists
> within their framework; the absence of a complement bound, and hence of a
> joint two-subgroup claim, is structural rather than an implementation gap.
:::

```{r b3-table}
#| output: asis
if (!HAVE_COMPLEMENT) {
  cat(paste0(
    "::: {.callout-important appearance='simple'}\n",
    "**Pending: T1 bundles absent.** `mr_field_complement_vs_guohe_t7_*.rds` ",
    "have not been produced, so the field-s complement upper coverage, its ",
    "mean bound location, and the joint coverage are *pending*. The two ",
    "paragraphs above stand on their own and are not contingent on them.\n\n",
    "On this design the complement's true effect is **exactly 0**: the DGM ",
    "sets $b(W) = \\beta_2$ on $\\{W \\le 30\\}$ and $0$ elsewhere ",
    "(`guohe_sec52_truth.R:75`), and the cutpoint grid starts at $30$, so ",
    "every $\\hat H^c = \\{W > \\hat c\\}$ lies inside the null region. ",
    "Complement coverage is therefore $\\text{upper} \\ge 0$, with no ",
    "dilution term.\n",
    ":::\n\n"))
} else {
  tab <- do.call(rbind, lapply(seq_along(CELLS), function(k) {
    r <- CMP[[CELLS[k]]]$results; n <- nrow(r)
    data.frame(
      Cell = CELLS[k], `$\\beta_2$` = sprintf("%.1f", BETA2[k]),
      `Complement upper coverage vs 0 (field-s, Wilson)` =
        fmt_cov(sum(r$c_cover_95_s), n),
      `Mean upper-bound location` = sprintf("%.3f", mean(r$c_upper_95_s, na.rm = TRUE)),
      `Joint coverage (Bonferroni, Wilson)` = fmt_cov(sum(r$joint_cover_s), n),
      check.names = FALSE, row.names = NULL)
  }))
  print(knitr::kable(tab, row.names = FALSE, caption = paste(
    "Capability, not comparison. Field-s one-sided 95% upper bound on",
    "$\\beta(\\hat H^c)$ against a truth of exactly 0, and the joint",
    "(0.975 lower on $\\hat H$, 0.975 upper on $\\hat H^c$) Bonferroni pair.",
    "Guo & He columns are absent by construction, not by omission.")))
  cat("\n")
}
```

# B5 — What was selected, and how stably

```{r b5}
#| output: asis
# N7: report the closed forms plus c-hat's distribution and the per-cell means
# of the two NON-degenerate metrics. Specificity and PPV are functions of c-hat
# alone on this design, so they are computed from c-hat, not re-estimated.
# NPV is degenerate by construction and is stated, never tabulated.
spec_of <- function(ch) (80 - ch) / 50      # P(W > c-hat | W > 30)
ppv_of  <- function(ch) 30 / ch             # P(W <= 30 | W <= c-hat)

b5 <- do.call(rbind, lapply(seq_along(CELLS), function(k) {
  r <- bun[[CELLS[k]]]$results
  err_sd <- sd(r$fld_est2 - r$theta)
  data.frame(
    Cell = CELLS[k], `$\\beta_2$` = sprintf("%.1f", BETA2[k]),
    `$\\hat c$ median (IQR)` = sprintf("%.1f (%.1f, %.1f)",
      median(r$c_hat), quantile(r$c_hat, .25), quantile(r$c_hat, .75)),
    `$\\hat c$ min/max` = sprintf("%.1f / %.1f", min(r$c_hat), max(r$c_hat)),
    `Mean specificity` = sprintf("%.3f", mean(spec_of(r$c_hat))),
    `Mean PPV` = sprintf("%.3f", mean(ppv_of(r$c_hat))),
    `$\\hat p(\\hat H)$ median (IQR)` = sprintf("%.3f (%.3f, %.3f)",
      median(r$p_hat_H), quantile(r$p_hat_H, .25), quantile(r$p_hat_H, .75)),
    `$M_{\\mathrm{eff}}$ median` = sprintf("%.1f", median(r$M_eff)),
    `Marginal SD` = sprintf("%.3f", sd(r$fld_est2)),
    `Error SD` = sprintf("%.3f", err_sd),
    check.names = FALSE, row.names = NULL)
}))
print(knitr::kable(b5, row.names = FALSE, caption = paste(
  "Selection and stability per cell. Specificity $= (80 - \\hat c)/50$ and",
  "PPV $= 30/\\hat c$ are exact functions of $\\hat c$ on this design and are",
  "computed as such; sensitivity and NPV are degenerate at 1 by nesting and are",
  "not tabulated. Marginal SD is $\\mathrm{sd}(\\hat\\beta_2)$ across replicates;",
  "error SD is $\\mathrm{sd}(\\hat\\beta_2 - \\gamma_{\\hat c})$ — they differ",
  "because $\\gamma_{\\hat c}$ itself moves with the selection.")))
cat("\n")
```

::: {.callout-note appearance="simple"}
**Classification is not the right lens on this design, and its arithmetic says
why.** The true harm region is $H = \{W \le 30\}$ and every candidate is
$S(c) = \{W \le c\}$ with $\hat c \ge 30$, so $H = S(30) \subseteq S(\hat c)$
for every replicate. With $W \sim \mathrm{Unif}(0, 80)$ the classification
operating characteristics are therefore closed forms in $\hat c$ alone:

- **Sensitivity $\equiv 1$** and **NPV $\equiv 1$** — *identically*, by nesting,
  not by merit. Both are degenerate by construction on this design, so neither
  is tabulated: a column of ones would read as a result when it is an artefact
  of $H \subseteq S(\hat c)$.
- **Specificity** $= (80 - \hat c)/50$ — reported per cell above.
- **PPV** $= 30/\hat c$ — reported per cell above.

Sensitivity carries no information in this design. The whole of the
identification story is how far $\hat c$ overshoots $30$, which the $\hat c$
distribution above reports directly and which specificity and PPV then restate
as monotone functions of it.
:::

```{r b5-chat-plot}
#| fig-width: 9
#| fig-height: 4.2
op <- par(mar = c(4, 4, 2, 1))
cl <- lapply(CELLS, function(id) bun[[id]]$results$c_hat)
boxplot(cl, names = sprintf("%.1f", BETA2), outline = FALSE,
        xlab = expression(beta[2]), ylab = expression(hat(c)),
        main = expression("Selected cutpoint " * hat(c) * " by " * beta[2]))
abline(h = 30, lty = 3)
mtext("dotted line: the true boundary c = 30", side = 3, line = -1.2, cex = 0.8)
par(op)
```

# B6 — Honest limits

```{r b6}
#| output: asis
# The field's coverage figures are COMPUTED from the rendered bundles, never
# transcribed, so this paragraph cannot go stale against the tables above.
# N7: lead with the MINIMUM, then give the span of the three lowest cells --
# quoting the span alone reads as if its upper end were a low point when it is
# mid-pack.
# Measured T1 ranges, computed from the committed complement bundles so the
# contrast in B6 limit 2 cannot drift from @sec-b3.
if (HAVE_COMPLEMENT) {
  .cu <- vapply(CMP, function(b) mean(b$results$c_cover_95_s), numeric(1))
  .cj <- vapply(CMP, function(b) mean(b$results$joint_cover_s), numeric(1))
  cert_upper_meas <- sprintf("%.3f-%.3f", min(.cu), max(.cu))
  cert_joint_meas <- sprintf("%.3f-%.3f", min(.cj), max(.cj))
} else {
  cert_upper_meas <- "pending"; cert_joint_meas <- "pending"
}

fld_cov <- SUM$cover[SUM$method == "MR (field)"]
fld_lo3 <- sort(fld_cov)[1:3]
fld_min <- sprintf("%.3f", min(fld_lo3))
fld_range <- sprintf("%.3f–%.3f", min(fld_lo3), max(fld_lo3))

cat(sprintf(paste0(
"Four limits belong on the record, stated plainly.\n\n",

"1. **The field's low points are ours, and they are under-correction, not ",
"noise.** Across these cells the MR (field) one-sided coverage **bottoms out ",
"at %s**, with the three lowest cells spanning %s. That shortfall is the ",
"stable-pick regime — replicates where ",
"$\\hat p(\\hat H)$ is high and the selection barely moves — and it is a ",
"design that does not flatter us: a nested, data-built family of ~151 ",
"candidates differing by one subject at a time is close to the worst case for ",
"a re-selection-based correction. We report the number rather than the ",
"regime-averaged one.\n\n",

"2. **The complement and joint figures in @sec-b3 are new measurements, not ",
"the certified ones.** The certified ranges on record are field-s one-sided ",
"upper coverage **%s** and joint Bonferroni coverage **%s**, both across all ",
"six cells of campaign `e1stud` at n = 500 (four ε band cells plus the maxSG ",
"and minSG ends), the joint pair's four band cells alone spanning the narrower ",
"0.939–0.952. That is a single design at one sample size — the cells vary the ",
"band width ε, the hazard ratio and the selection focus, not the prevalence or ",
"n — so it certifies those constructions on that design, not across ",
"prevalences or sample sizes. The wider figures 0.912–0.960 and 0.939–0.963, ",
"quoted in the task documents, appear only in an uncommitted handoff and are ",
"not reproducible from anything in this repository. Measured here on Table 7, ",
"for contrast: field-s upper **%s**, joint **%s**. The Table 7 design is a ",
"different regime and the numbers here stand or fall on their own.\n\n",

"3. **Guo & He is not run on our screened, data-built family.** Their method ",
"is applied here exactly as published, on the family their Section 5.2 ",
"specifies. It is not applied to the forestsearch-screened family, which lies ",
"outside its stated scope; no claim is made about how it would behave there.\n\n",

"4. **Everything here is one-sided.** Every construction compared in this ",
"section is one-sided, so none of it depends on the choice of two-sided ",
"interval method.\n"),
fld_min, fld_range, cert_upper, cert_joint, cert_upper_meas, cert_joint_meas))
```

```{r b6-cert-note}
#| output: asis
if (!CERT_PRESENT) cat(paste0(
  "::: {.callout-warning appearance='simple'}\n",
  "**Certified-range source missing.** `", CERT_FILE, "` is not in the tree, so ",
  "the two certified figures in limit 2 could not be verified against their ",
  "source at render time.\n",
  ":::\n\n"))
```

# Provenance

```{r provenance}
prov <- data.frame(
  Item = c("Cells", "Replicates per cell", "MR draws", "Field $R_{out}$ / $R_{in}$",
           "MCSE at coverage 0.95", "Adaptive column", "Complement / joint (B3)",
           "Certified-range source (B6)"),
  Value = c(
    paste(CELLS, collapse = ", "),
    format(NREP),
    sprintf("%d (%s)", DRAWS, bun[[1]]$multiplier),
    sprintf("%d / %d", R_OUT, R_IN),
    sprintf("%.4f", MCSE),
    if (HAVE_ADAPTIVE) "present" else "PENDING PHASE B",
    if (HAVE_COMPLEMENT) "present" else "PENDING (T1 bundles absent)",
    if (CERT_PRESENT) sprintf("%s (present)", CERT_FILE) else
      sprintf("%s (MISSING)", CERT_FILE)),
  check.names = FALSE)
knitr::kable(prov, row.names = FALSE, caption = paste(
  "Inputs and phase state. Every bundle is read-only; the G&H fixed-$r$",
  "columns are arithmetic on the stored replication bundles and are never",
  "re-run."))
```

```{r provenance-platform}
#| output: asis
# N7 item 3: one provenance sentence on the two build platforms. The N3 tally is
# reported here and, being zero, closes the matter -- no further caveat appears
# anywhere else in this section.
if (HAVE_COMPLEMENT) {
  n3 <- sum(vapply(CMP, function(b) b$gate2$sel_mismatch, numeric(1)))
  n3n <- sum(vapply(CMP, function(b) b$gate2$n_rep_done, numeric(1)))
  wa <- max(vapply(CMP, function(b) b$gate2$worst_abs, numeric(1)))
  cat(sprintf(paste0(
    "The complement and joint bundles in @sec-b3 were produced on arm64 with ",
    "Accelerate, while the head-to-head bundles behind the B2 tables were ",
    "produced on x86_64 with reference BLAS; across every column with a ",
    "counterpart in both, the worst absolute deviation is %.2e (all within ",
    "`all.equal` at 1e-8), and the selected cutpoint agreed on %d of %d ",
    "replicates.\n\n"), wa, n3n - n3, n3n))
  if (n3 > 0) cat(sprintf(paste0(
    "::: {.callout-warning appearance='simple'}\n",
    "**%d selection mismatch(es)** between the two builds. Every comparison ",
    "involving the complement columns is affected and is caveated accordingly.\n",
    ":::\n\n"), n3))
}
```
