---
title: "Smoke test --- GLM-generalized Guo & He (2021) de-biasing comparator"
subtitle: "End-to-end pipeline check on synthetic survival, binary, and continuous data"
date: 2026-07-17
format: html
bibliography: []
---

This living document exercises `guohe_debias()` (in `guohe_comparator.R`) end to
end on self-generated synthetic data for all three outcome types. It is a
fail-loud integration check: each section asserts that the mechanics run and
that de-biasing moves the selected effect toward the null with a conservative
one-sided bound. No external data are required. If `synthetic code.R` and
`syntheticdata.csv` (the MONET1 survival reference) are present alongside this
file, the final section additionally confirms an **exact** reproduction of that
reference.

```{r}
#| label: setup
#| message: false
library(survival)
source("guohe_comparator.R")

# One synthetic trial with eight binary candidate covariates; v1 is the true
# effect modifier carrying an extra treatment effect. `link` shapes the outcome.
simulate_trial <- function(n, seed, link = c("survival", "binary", "continuous")) {
  link <- match.arg(link)
  set.seed(seed)
  trt <- rbinom(n, 1, 0.5)
  x   <- as.data.frame(matrix(rbinom(n * 8, 1, 0.4), n, 8))
  names(x) <- paste0("v", 1:8)
  modifier <- x$v1
  base <- 0.4 * x$v2 - 0.3 * x$v3

  if (link == "survival") {
    loghaz <- -3 + base + trt * (-0.2 - 1.0 * modifier)   # protective, +benefit in v1
    tlat   <- rexp(n, rate = exp(loghaz))
    cens   <- rexp(n, rate = exp(-3.2))
    out <- data.frame(trt = trt,
                      time  = pmin(tlat, cens),
                      event = as.integer(tlat <= cens))
  } else if (link == "binary") {
    lp <- -0.3 + base + trt * (-0.2 - 1.0 * modifier)      # protective OR in v1
    out <- data.frame(trt = trt, y = rbinom(n, 1, plogis(lp)))
  } else {
    mu <- 1.0 + base + trt * (0.1 + 1.2 * modifier)        # +MD in v1
    out <- data.frame(trt = trt, y = rnorm(n, mu, 1.0))
  }
  cbind(out, x)
}

candidates <- paste0("v", 1:8)   # v1 is the planted modifier
```

## Survival (hazard ratio)

```{r}
#| label: survival
d_surv <- simulate_trial(1000, seed = 11, link = "survival")
res_surv <- guohe_debias(
  d_surv, outcome = "survival", treatment = "trt", candidates = candidates,
  time = "time", event = "event", orient = -1,
  B = 500L, r = 0.03, level = 0.05, seed = 1L, symmetric = FALSE
)
print(res_surv)

stopifnot(
  is.finite(res_surv$naive),
  res_surv$debiased > res_surv$naive,   # HR shrinks toward 1
  res_surv$bound    > res_surv$debiased # upper bound is conservative
)
```

## Binary (odds ratio)

```{r}
#| label: binary
d_bin <- simulate_trial(1000, seed = 22, link = "binary")
res_bin <- guohe_debias(
  d_bin, outcome = "binary", treatment = "trt", candidates = candidates,
  y = "y", orient = -1, B = 500L, r = 0.03, level = 0.05,
  seed = 1L, symmetric = FALSE
)
print(res_bin)

stopifnot(
  is.finite(res_bin$naive),
  res_bin$debiased > res_bin$naive,   # OR shrinks toward 1
  res_bin$bound    > res_bin$debiased # upper bound is conservative
)
```

## Continuous (mean difference)

```{r}
#| label: continuous
d_cont <- simulate_trial(1000, seed = 33, link = "continuous")
res_cont <- guohe_debias(
  d_cont, outcome = "continuous", treatment = "trt", candidates = candidates,
  y = "y", orient = +1, B = 500L, r = 0.03, level = 0.05,
  seed = 1L, symmetric = FALSE
)
print(res_cont)

stopifnot(
  is.finite(res_cont$naive),
  res_cont$debiased < res_cont$naive,   # MD shrinks toward 0
  res_cont$bound    < res_cont$debiased # toward-zero bound is conservative
)
```

## Optional: exact reproduction of the MONET1 survival reference

Runs only if `synthetic code.R` and `syntheticdata.csv` sit alongside this file.
Confirms that `outcome = "survival"`, `orient = -1`, `symmetric = TRUE`
reproduces `BC()` and `CB()` to machine precision on identical resamples.

```{r}
#| label: monet1-exact
if (file.exists("synthetic code.R") && file.exists("syntheticdata.csv")) {
  # Evaluate only the reference's function definitions (up to its read.csv line),
  # so BC()/CB() are defined without triggering its slow B = 2000 demonstration.
  ref_txt <- readLines("synthetic code.R")
  cut <- grep("read.csv", ref_txt)[1] - 1L
  eval(parse(text = paste(ref_txt[seq_len(cut)], collapse = "\n")))

  D <- read.csv("syntheticdata.csv"); n <- nrow(D)
  B <- 500L; r <- 0.03; level <- 0.05; S <- 20260717L
  inds <- names(D)[4:11]

  set.seed(S); ref_bound <- BC(D, B, r, n, level)
  set.seed(S); ref_point <- CB(D, B, r, n)
  mine <- guohe_debias(D, outcome = "survival", treatment = "treatment",
                       candidates = inds, time = "aval1", event = "CNSR",
                       orient = -1, B = B, r = r, level = level,
                       seed = S, min_events = 1L, symmetric = TRUE)

  stopifnot(
    abs(ref_bound - mine$bound)    < 1e-9,
    abs(ref_point - mine$debiased) < 1e-9
  )
  cat(sprintf("Exact match confirmed: bound %.6f, debiased HR %.6f (selected %s)\n",
              mine$bound, mine$debiased, mine$selected))
} else {
  cat("MONET1 reference files not found alongside this document; section skipped.\n")
}
```
