---
title: "Smoke test --- Guo & He Algorithm 3 (enumerated family) comparator"
subtitle: "Structural invariants, three outcome paths, and scale/calibration tiers"
date: 2026-07-18
format: html
bibliography: []
---

Fail-loud integration test for `guohe_algorithm3()` in `guohe_algorithm3.R`. Every
section asserts with `stopifnot`, so a completed render *is* a pass.

**Two tiers, set by `SCALE` below.**

- `SCALE = "quick"` --- structural invariants, all three outcome paths, small
  families, modest `B`. Runs in roughly a minute; suitable anywhere.
- `SCALE = "full"` --- adds a realistic enumerated conjunction family
  (several hundred candidates) at production `B`, and a replicated null
  calibration check. **Run this tier in Claude Code / on the 127-core box**: the
  full-family cell alone is on the order of an hour single-core, and the
  calibration cell is a replication loop that wants parallel workers.

The invariants in the quick tier are the load-bearing ones: they lock the
validation chain back to the authors' published Table 2, and they pin the
relationship between the certified one-sided bound and the two-sided extension.

**Closed forms.** All three full-tier cells are GLM paths (scale and separated
are binary, calibration is continuous), so `guohe_algorithm3()` auto-enables its
exact closed-form evaluation: with a single binary covariate and no adjustment,
the logistic model is saturated -- its MLE is the 2x2 log-odds-ratio -- and the
Gaussian coefficient is the difference of arm means. These are the SAME
estimators `forestsearch()` fits in its unadjusted mode (`.estimate_or()` with
`adjust_covariates = NULL, ps_adjust_method = "none"`, and
`.make_lm_estimator()`), not approximations to them. Verified against `glm`/`lm`:
continuous agrees to 7e-15 and binary to 9e-9 (the latter being `glm`'s own IRLS
tolerance, so the closed form is if anything the more accurate of the two), with
identical selected candidates and identical drop-rate vectors -- at roughly
**700x** the speed. The survival cells are unaffected: Cox has no closed form and
the gate refuses one.

```{r}
#| label: setup
#| message: false
library(survival)
library(forestsearch)  # guohe_algorithm3() / guohe_adaptive_r() are package API

SCALE <- "full"           # "quick" (any machine) or "full" (Claude Code / 127-core)
B_Q   <- 200L             # bootstrap draws, quick tier

# ---- Full-tier settings -----------------------------------------------------
# PRODUCTION values, for a real operating-characteristic evaluation rather than
# a smoke test. Reduce for a quick code check (B_F 300L, REPS 100L, B_CAL 500L);
# the calibration cell stamps a PROVISIONAL banner whenever it is run below
# REPS 300 / B_CAL 1000.
#
# What is safe to reduce, and what is not:
#   * B_F drives only the enumerated-family scale cell, which makes no coverage
#     claim. Reducing it costs nothing statistically.
#   * REPS controls only the PRECISION of the calibration coverage estimate, not
#     the estimator. At REPS = 100 the Monte Carlo SE on a 0.95 coverage is
#     about 0.022 -- ample to clear the 0.80 assertion, too coarse to publish.
#   * B_CAL is the inner bootstrap for the calibration cell and DOES affect the
#     answer: the bound is a tail quantile (q_{1-alpha}), and noisy quantiles pull
#     coverage slightly downward. The de-biased point estimate is a mean and
#     converges much faster, so it is comparatively insensitive to B_CAL.
B_F   <- 2000L            # bootstrap draws, full-family scale cell
REPS  <- 1000L            # replications, calibration cell
B_CAL <- 1000L            # inner bootstrap draws, calibration cell

# ---- Parallelism (calibration cell only) -----------------------------------
# Outermost loop only. The invariant/regression cells stay SERIAL so their
# exact-match assertions remain valid regression tests.
PARALLEL <- TRUE
WORKERS  <- round(0.9*max(1L, parallel::detectCores() - 2L))

# One synthetic trial: `k` binary candidate covariates, v1 the true modifier.
simulate_trial <- function(n, k, seed,
                           link = c("survival", "binary", "continuous"),
                           delta = 1.0) {
  link <- match.arg(link)
  set.seed(seed)
  trt <- rbinom(n, 1, 0.5)
  x   <- as.data.frame(matrix(rbinom(n * k, 1, 0.4), n, k))
  names(x) <- paste0("v", seq_len(k))
  modifier <- x$v1
  base <- 0.4 * x$v2 - 0.3 * x$v3

  out <- switch(link,
    survival = {
      loghaz <- -3 + base + trt * (-0.2 - delta * modifier)   # protective in v1
      tlat <- rexp(n, rate = exp(loghaz))
      cens <- rexp(n, rate = exp(-3.2))
      data.frame(trt = trt, time = pmin(tlat, cens),
                 event = as.integer(tlat <= cens))
    },
    binary = {
      lp <- -0.3 + base + trt * (-0.2 + delta * modifier)     # HARMFUL in v1
      data.frame(trt = trt, y = rbinom(n, 1, plogis(lp)))
    },
    continuous = {
      mu <- 1.0 + base + trt * (0.1 + delta * modifier)       # +MD in v1
      data.frame(trt = trt, y = rnorm(n, mu, 1.0))
    }
  )
  cbind(out, x)
}

cand_names <- function(k) paste0("v", seq_len(k))
cat("SCALE =", SCALE, "\n")
```

## Invariant A --- continuity with the validated Section 2 engine

`guohe_algorithm3()` must reproduce `guohe_debias()` exactly on identical seeds:
the arithmetic is the same, so this transitively preserves the machine-precision
match to the authors' `synthetic code.R` and the ~1e-3 agreement with their
published Table 2. If this fails, the validation chain is broken.

```{r}
#| label: invariant-continuity
if (file.exists("guohe_comparator.R")) {
  source("guohe_comparator.R")
  d <- simulate_trial(900, 8, seed = 11, link = "survival")
  cn <- cand_names(8); S <- 4242L

  a3 <- guohe_algorithm3(d, "survival", treatment = "trt", candidates = cn,
                         time = "time", event = "event", orient = -1,
                         B = B_Q, r = 0.03, level = 0.05, seed = S,
                         min_events = 5L, diagnostics = FALSE)
  s2 <- guohe_debias(d, "survival", treatment = "trt", candidates = cn,
                     time = "time", event = "event", orient = -1,
                     B = B_Q, r = 0.03, level = 0.05, seed = S,
                     min_events = 5L, symmetric = FALSE)

  stopifnot(
    identical(a3$selected, s2$selected),
    abs(a3$naive           - s2$naive)    < 1e-12,
    abs(a3$debiased        - s2$debiased) < 1e-12,
    abs(a3$bound_one_sided - s2$bound)    < 1e-12
  )
  cat(sprintf("PASS: A3 reproduces the validated engine exactly (bound %.10f)\n",
              a3$bound_one_sided))
} else {
  cat("guohe_comparator.R not alongside; continuity check skipped.\n")
}
```

## Invariant B --- the one-sided / two-sided quantile relationship

The certified one-sided limit inverts $T^*$ at $q_{1-\alpha}$; the two-sided
interval inverts the *same* $T^*$ at $(q_{\alpha/2}, q_{1-\alpha/2})$. So the
two-sided conservative endpoint must equal the one-sided bound run at
`level/2` **exactly**, and must be strictly more conservative than the one-sided
bound at `level`. This pins the distinction the interval labelling depends on.

```{r}
#| label: invariant-quantile
d  <- simulate_trial(900, 8, seed = 12, link = "survival")
cn <- cand_names(8); S <- 777L

full <- guohe_algorithm3(d, "survival", treatment = "trt", candidates = cn,
                         time = "time", event = "event", orient = -1,
                         B = B_Q, level = 0.05, seed = S, diagnostics = FALSE)
half <- guohe_algorithm3(d, "survival", treatment = "trt", candidates = cn,
                         time = "time", event = "event", orient = -1,
                         B = B_Q, level = 0.025, seed = S, diagnostics = FALSE)

# orient = -1 on a ratio scale: the UPPER endpoint is the conservative one
cons <- full$ci_two_sided[2]
stopifnot(
  abs(cons - half$bound_one_sided) < 1e-12,   # identity
  cons > full$bound_one_sided,                # strictly wider
  full$ci_two_sided[1] < full$ci_two_sided[2]
)
cat(sprintf("PASS: two-sided(0.05) upper %.6f == one-sided(0.025) %.6f; one-sided(0.05) %.6f\n",
            cons, half$bound_one_sided, full$bound_one_sided))

# the Wald interval is a DIFFERENT object (T* is skewed) -- report, do not assert equality
cat(sprintf("      Wald upper %.6f (heuristic; differs from quantile inversion)\n",
            full$ci_wald[2]))
```

## Survival path (hazard ratio, benefit orientation)

```{r}
#| label: survival
d <- simulate_trial(1000, 10, seed = 21, link = "survival")
res_surv <- guohe_algorithm3(d, "survival", treatment = "trt",
                             candidates = cand_names(10),
                             time = "time", event = "event", orient = -1,
                             B = B_Q, seed = 1L)
print(res_surv)

stopifnot(
  is.finite(res_surv$naive),
  res_surv$debiased        > res_surv$naive,      # HR shrinks toward 1
  res_surv$bound_one_sided > res_surv$debiased,   # bound conservative
  res_surv$ci_two_sided[1] < res_surv$ci_two_sided[2]
)
```

## Binary path (odds ratio, HARM orientation)

The forest-search harm setting: `orient = +1` selects the most *harmful* odds
ratio, and de-biasing must shrink it **down** toward 1.

```{r}
#| label: binary-harm
d <- simulate_trial(1000, 10, seed = 22, link = "binary")
res_bin <- guohe_algorithm3(d, "binary", treatment = "trt",
                            candidates = cand_names(10), y = "y",
                            orient = +1, B = B_Q, seed = 1L)
print(res_bin)

stopifnot(
  is.finite(res_bin$naive),
  res_bin$debiased        < res_bin$naive,       # OR shrinks toward 1 from above
  res_bin$bound_one_sided < res_bin$debiased,    # lower bound conservative
  res_bin$ci_two_sided[1] < res_bin$ci_two_sided[2]
)
```

## Continuous path, and orientation symmetry

Negating the outcome *and* flipping `orient` leaves the internal oriented score
`orient * b` unchanged, so the selected candidate must be identical and every
reported mean difference must flip sign exactly. This is a genuine algebraic
invariant of the orientation handling, not a soft check.

```{r}
#| label: continuous-symmetry
d <- simulate_trial(1000, 10, seed = 23, link = "continuous")
cn <- cand_names(10); S <- 5L

pos <- guohe_algorithm3(d, "continuous", treatment = "trt", candidates = cn,
                        y = "y", orient = +1, B = B_Q, seed = S,
                        diagnostics = FALSE)
d_neg <- d; d_neg$y <- -d$y
neg <- guohe_algorithm3(d_neg, "continuous", treatment = "trt", candidates = cn,
                        y = "y", orient = -1, B = B_Q, seed = S,
                        diagnostics = FALSE)
print(pos)

stopifnot(
  is.finite(pos$naive),
  pos$debiased        < pos$naive,       # MD shrinks toward 0
  pos$bound_one_sided < pos$debiased,    # toward-zero bound conservative
  identical(pos$selected, neg$selected),
  abs(pos$naive    + neg$naive)    < 1e-10,
  abs(pos$debiased + neg$debiased) < 1e-10
)
cat(sprintf("PASS: orientation symmetry (selected %s; MD %.6f vs %.6f)\n",
            pos$selected, pos$debiased, neg$debiased))
```

## Reproducibility and estimability diagnostics

Same seed must give bit-identical output. The drop-rate diagnostic must be
reported for every candidate and lie in $[0,1]$; a deliberately tiny candidate
should show a strictly higher drop rate than a large one, since small subgroups
lose an arm in resamples and are silently removed from that resample's maximum.

```{r}
#| label: repro-diagnostics
d <- simulate_trial(800, 6, seed = 24, link = "continuous")
cn <- cand_names(6)

r1 <- guohe_algorithm3(d, "continuous", treatment = "trt", candidates = cn,
                       y = "y", orient = +1, B = B_Q, seed = 99L)
r2 <- guohe_algorithm3(d, "continuous", treatment = "trt", candidates = cn,
                       y = "y", orient = +1, B = B_Q, seed = 99L)
stopifnot(
  identical(r1$selected, r2$selected),
  abs(r1$debiased        - r2$debiased)        < 1e-12,
  abs(r1$bound_one_sided - r2$bound_one_sided) < 1e-12
)

# Add a small candidate sized to produce PARTIAL drop-out -- the insidious case:
# estimable on the observed data, yet absent from a sizeable share of resamples,
# which silently shrinks the competitor set and UNDER-states the correction.
# (A candidate small enough to be dropped outright is less dangerous: it is
# excluded up front and raises a warning.)
d2 <- d
set.seed(1); d2$small <- 0L; d2$small[sample.int(nrow(d2), 26)] <- 1L
d2$big <- as.integer(runif(nrow(d2)) < 0.45)
r3 <- guohe_algorithm3(d2, "continuous", treatment = "trt",
                       candidates = c(cn, "small", "big"), y = "y",
                       orient = +1, B = B_Q, seed = 3L, min_events = 5L)
dr <- r3$drop_rate
stopifnot(
  !is.null(dr),
  length(dr) == length(cn) + 2L,
  all(dr >= 0 & dr <= 1),
  r3$n_estimable == r3$n_candidates,   # fine on the observed data ...
  dr[["small"]] > 0.05,                # ... yet dropping in many resamples
  dr[["small"]] < 1,                   # partial, not outright exclusion
  dr[["small"]] > dr[["big"]]
)
cat(sprintf("PASS: reproducible; partial drop-out detected -- small(n=26) %.3f vs big %.3f\n",
            dr[["small"]], dr[["big"]]))
```

## Full tier --- enumerated conjunction family at production `B`

Realistic forest-search geometry: cuts on continuous covariates plus their
pairwise conjunctions, giving several hundred overlapping candidates. **Claude
Code / 127-core box.** Skipped unless `SCALE == "full"`.

```{r}
#| label: full-scale
if (identical(SCALE, "full")) {
  set.seed(9090); n <- 1200
  trt <- rbinom(n, 1, 0.5)
  z <- matrix(rnorm(n * 4), n, 4)
  harm <- as.integer(z[, 1] > 0.5 & z[, 2] > 0)
  lp <- -0.4 + 0.3 * z[, 3] + trt * (-0.25 + 1.1 * harm)
  ybin <- rbinom(n, 1, plogis(lp))

  base <- list(); nm <- character(0)
  for (j in 1:4) for (q in c(.3, .4, .5, .6, .7, .8)) {
    base[[length(base) + 1L]] <- as.integer(z[, j] > quantile(z[, j], q))
    nm <- c(nm, sprintf("z%d.gt.q%02d", j, round(100 * q)))
  }
  base <- as.data.frame(base); names(base) <- nm

  prs <- list(); pnm <- character(0); ix <- combn(ncol(base), 2)
  for (k in seq_len(ncol(ix))) {
    v <- as.integer(base[[ix[1, k]]] == 1 & base[[ix[2, k]]] == 1)
    if (sum(v) >= 60) {
      prs[[length(prs) + 1L]] <- v
      pnm <- c(pnm, paste0(names(base)[ix[1, k]], "_AND_", names(base)[ix[2, k]]))
    }
  }
  prs <- as.data.frame(prs); names(prs) <- pnm
  fam <- cbind(base, prs)
  dat <- cbind(data.frame(trt = trt, y = ybin), fam)

  # No outer loop here, so parallelism goes over the bootstrap draws instead.
  # Safe because nothing else is parallel at this point.
  if (isTRUE(PARALLEL)) future::plan(future::multisession, workers = WORKERS)
  res_full <- guohe_algorithm3(dat, "binary", treatment = "trt",
                               candidates = names(fam), y = "y",
                               orient = +1, B = B_F, seed = 11L,
                               min_events = 5L, parallel = PARALLEL)
  if (isTRUE(PARALLEL)) future::plan(future::sequential)
  print(res_full)
  stopifnot(
    res_full$n_candidates > 200L,
    res_full$debiased < res_full$naive,          # large family -> large correction
    res_full$ci_two_sided[1] < res_full$ci_two_sided[2]
  )
  cat(sprintf("PASS: %d candidates; naive OR %.3f -> de-biased %.3f\n",
              res_full$n_candidates, res_full$naive, res_full$debiased))
} else {
  cat("SCALE != 'full'; enumerated-conjunction cell skipped (Claude Code tier).\n")
}
```

## Full tier --- null calibration

Under a null with **no** effect modification, every candidate has the same true
effect, so the naive maximum is pure winner's curse. The de-biased estimate
should return near the truth and the one-sided bound should cover at roughly the
nominal rate. This is the statistically meaningful check and needs replication;
parallelize over `REPS` with `doFuture`/`future` on the 127-core box.

```{r}
#| label: full-calibration
if (identical(SCALE, "full")) {
  TRUE_MD <- 0.1
  one_rep <- function(rep_id) {
    # Force the RNG kind: a parallel backend may leave L'Ecuyer-CMRG current,
    # and set.seed() inherits the CURRENT kind. Pinning it here makes each
    # replication byte-identical to the serial run.
    RNGkind("Mersenne-Twister")
    set.seed(10000 + rep_id)
    n <- 800; k <- 25
    trt <- rbinom(n, 1, 0.5)
    x <- as.data.frame(matrix(rbinom(n * k, 1, 0.4), n, k))
    names(x) <- paste0("v", seq_len(k))
    y <- 1 + TRUE_MD * trt + rnorm(n)            # no modification anywhere
    dd <- cbind(data.frame(trt = trt, y = y), x)
    r <- guohe_algorithm3(dd, "continuous", treatment = "trt",
                          candidates = names(x), y = "y", orient = +1,
                          B = B_CAL, seed = rep_id, diagnostics = FALSE)
    c(naive = r$naive, debiased = r$debiased,
      covered_one = as.numeric(r$bound_one_sided <= TRUE_MD),
      covered_two = as.numeric(r$ci_two_sided[1] <= TRUE_MD &&
                                 TRUE_MD <= r$ci_two_sided[2]))
  }
  # Parallelism, outermost loop only. Each replication seeds itself, so the
  # parallel result is byte-identical to the serial one -- which is what makes
  # the speed-up verifiable. Never also set parallel = TRUE inside
  # guohe_algorithm3() here: that would nest multisession within multisession.
  if (isTRUE(PARALLEL)) {
    future::plan(future::multisession, workers = WORKERS)
    reps <- future.apply::future_lapply(
      seq_len(REPS),
      function(i) {
        library(forestsearch)   # guohe_algorithm3() is package API
        one_rep(i)
      },
      future.seed = TRUE,
      future.globals = c("one_rep", "TRUE_MD", "B_CAL")
    )
  } else {
    reps <- lapply(seq_len(REPS), one_rep)
  }
  if (isTRUE(PARALLEL)) future::plan(future::sequential)   # release workers
  out <- do.call(rbind, reps)
  cat(sprintf("truth %.3f | mean naive %.3f (bias %+.3f) | mean de-biased %.3f (bias %+.3f)\n",
              TRUE_MD, mean(out[, "naive"]), mean(out[, "naive"]) - TRUE_MD,
              mean(out[, "debiased"]), mean(out[, "debiased"]) - TRUE_MD))
  cat(sprintf("one-sided bound coverage %.3f | two-sided CI coverage %.3f (nominal 0.95)\n",
              mean(out[, "covered_one"]), mean(out[, "covered_two"])))
  cat(sprintf("settings: REPS = %d, B_CAL = %d | Monte Carlo SE on coverage ~ %.3f\n",
              REPS, B_CAL, sqrt(0.95 * 0.05 / REPS)))
  if (REPS < 300L || B_CAL < 1000L) {
    cat("*** PROVISIONAL: code-check settings. Restore REPS >= 300 and",
        "B_CAL >= 1000 before quoting these coverage figures. ***\n")
  }
  stopifnot(
    mean(out[, "naive"]) > TRUE_MD,                                  # winner's curse present
    abs(mean(out[, "debiased"]) - TRUE_MD) < abs(mean(out[, "naive"]) - TRUE_MD),
    mean(out[, "covered_one"]) > 0.80
  )
  cat("PASS: de-biasing reduces bias and the one-sided bound covers.\n")
} else {
  cat("SCALE != 'full'; calibration cell skipped (Claude Code tier).\n")
}
```

## Full tier --- separated-signal operating characteristics

The calibration cell above is the **all-tied null**: every candidate shares the
same true effect, which is the configuration the shrinkage offsets are designed
for. This cell is its complement --- a **real signal inside a large overlapping
conjunction family**, i.e. the regime a forest-search application actually
lands in. It replicates the enumerated-family geometry of the scale cell across
many datasets and evaluates bias and coverage at several fixed `r`.

Truth is the least-false marginal odds ratio of the **selected** rule, which
differs from replication to replication (the winner is a data-driven
approximation to the generating rule and generally does not equal it). Each
worker therefore returns the numeric thresholds of its selected rule, and the
master evaluates those against one large fixed population --- which keeps the
truth computation out of the workers entirely.

**Cost.** With the closed-form path this cell is no longer expensive: the
binary outcome is unadjusted, so each resample scores the whole family with one
matrix product rather than `candidates` model fits. What was roughly 47 minutes
on 125 workers under `glm` is now a few minutes. Adaptive-`r` (Algorithm 2) was
excluded here because it multiplies cost by a further `v * length(r_grid)` per
dataset, which was prohibitive under `glm`; that objection no longer holds, and
adding it is the natural next extension of this cell.

```{r}
#| label: full-separated
SEP_REPS <- 200L                    # replications
SEP_N    <- 1200L                   # trial size per replication
SEP_B    <- 300L                    # bootstrap draws
SEP_R    <- c(0.03, 0.15, 0.30)     # fixed tuning values to compare
SEP_V    <- 5L                      # CV folds for the adaptive-r arm
TRUTH_N  <- 300000L                 # population size for the truth evaluation

# Build the enumerated family AND record each candidate's numeric thresholds,
# so the selected rule can be re-applied to the truth population later.
build_family <- function(z) {
  qgrid <- c(.3, .4, .5, .6, .7, .8)
  cols <- list(); nm <- character(0); thr <- list()
  for (j in 1:4) for (q in qgrid) {
    tv <- unname(stats::quantile(z[, j], q))
    cols[[length(cols) + 1L]] <- as.integer(z[, j] > tv)
    nm <- c(nm, sprintf("z%d.gt.q%02d", j, round(100 * q)))
    t4 <- rep(-Inf, 4); t4[j] <- tv
    thr[[length(thr) + 1L]] <- t4
  }
  nb <- length(cols)
  ix <- utils::combn(nb, 2)
  for (k in seq_len(ncol(ix))) {
    a <- ix[1, k]; b <- ix[2, k]
    v <- as.integer(cols[[a]] == 1 & cols[[b]] == 1)
    if (sum(v) >= 60) {
      cols[[length(cols) + 1L]] <- v
      nm  <- c(nm, paste0(nm[a], "_AND_", nm[b]))
      thr[[length(thr) + 1L]] <- pmax(thr[[a]], thr[[b]])   # conjunction = tighter
    }
  }
  fam <- as.data.frame(cols); names(fam) <- nm
  list(fam = fam, thr = stats::setNames(thr, nm))
}

# One replication: generate, enumerate, then de-bias at each fixed r.
one_sep_rep <- function(rep_id) {
  RNGkind("Mersenne-Twister")
  set.seed(50000 + rep_id)
  trt  <- rbinom(SEP_N, 1, 0.5)
  z    <- matrix(rnorm(SEP_N * 4), SEP_N, 4)
  harm <- as.integer(z[, 1] > 0.5 & z[, 2] > 0)
  lp   <- -0.4 + 0.3 * z[, 3] + trt * (-0.25 + 1.1 * harm)
  y    <- rbinom(SEP_N, 1, plogis(lp))
  bf  <- build_family(z)
  dat <- cbind(data.frame(trt = trt, y = y), bf$fam)

  row_of <- function(arm, r_used, res) {
    tv <- bf$thr[[res$selected]]
    data.frame(rep_id = rep_id, arm = arm, r_used = r_used,
               selected = res$selected,
               t1 = tv[1], t2 = tv[2], t3 = tv[3], t4 = tv[4],
               naive = res$naive, debiased = res$debiased,
               bound = res$bound_one_sided,
               ci_lo = res$ci_two_sided[1], ci_hi = res$ci_two_sided[2],
               stringsAsFactors = FALSE)
  }

  # fixed-r arms
  fixed <- do.call(rbind, lapply(SEP_R, function(rv) {
    res <- guohe_algorithm3(dat, "binary", treatment = "trt",
                            candidates = names(bf$fam), y = "y",
                            orient = +1, B = SEP_B, r = rv, level = 0.05,
                            seed = rep_id, min_events = 5L, diagnostics = FALSE)
    row_of(sprintf("r=%.2f", rv), rv, res)
  }))

  # adaptive-r arm (Algorithm 2), selecting r from the same grid per dataset
  ar <- try(suppressWarnings(
    guohe_adaptive_r(dat, "binary", treatment = "trt",
                     candidates = names(bf$fam), y = "y", orient = +1,
                     r_grid = SEP_R, v = SEP_V, B = SEP_B, level = 0.05,
                     seed = rep_id, min_events = 5L, refit = TRUE)
  ), silent = TRUE)
  if (inherits(ar, "try-error")) return(fixed)
  rbind(fixed, row_of("adaptive", ar$r_hat, ar$fit))
}

if (identical(SCALE, "full")) {
  if (isTRUE(PARALLEL)) {
    future::plan(future::multisession, workers = WORKERS)
    sep <- future.apply::future_lapply(
      seq_len(SEP_REPS),
      function(i) {
        library(forestsearch)   # guohe_algorithm3() / guohe_adaptive_r() are package API
        one_sep_rep(i)
      },
      future.seed = TRUE,
      future.globals = c("one_sep_rep", "build_family",
                         "SEP_N", "SEP_B", "SEP_R", "SEP_V")
    )
    future::plan(future::sequential)
  } else {
    sep <- lapply(seq_len(SEP_REPS), one_sep_rep)
  }
  sep <- do.call(rbind, sep)

  # ---- truth for each selected rule, evaluated in the master ---------------
  set.seed(987654)
  Ztr   <- matrix(rnorm(TRUTH_N * 4), TRUTH_N, 4)
  trttr <- rbinom(TRUTH_N, 1, 0.5)
  hmtr  <- as.integer(Ztr[, 1] > 0.5 & Ztr[, 2] > 0)
  ytr   <- rbinom(TRUTH_N, 1,
                  plogis(-0.4 + 0.3 * Ztr[, 3] + trttr * (-0.25 + 1.1 * hmtr)))

  truth_of <- function(t1, t2, t3, t4) {
    keep <- (Ztr[, 1] > t1) & (Ztr[, 2] > t2) & (Ztr[, 3] > t3) & (Ztr[, 4] > t4)
    if (sum(keep) < 200L) return(NA_real_)
    ok <- tryCatch(suppressWarnings(
      stats::glm(ytr[keep] ~ trttr[keep], family = stats::binomial())),
      error = function(e) NULL)
    if (is.null(ok)) NA_real_ else unname(exp(stats::coef(ok)[2]))
  }
  key <- paste(sep$t1, sep$t2, sep$t3, sep$t4, sep = "|")
  uk  <- !duplicated(key)
  tru <- vapply(which(uk), function(i)
    truth_of(sep$t1[i], sep$t2[i], sep$t3[i], sep$t4[i]), numeric(1))
  sep$truth <- tru[match(key, key[uk])]

  # ---- summarise by r ------------------------------------------------------
  ok <- is.finite(sep$truth)
  arms <- c(sprintf("r=%.2f", SEP_R), "adaptive")
  summ <- do.call(rbind, lapply(arms, function(av) {
    s <- sep[ok & sep$arm == av, ]
    if (!nrow(s)) return(NULL)
    data.frame(
      arm            = av,
      reps           = nrow(s),
      mean_r         = mean(s$r_used),
      mean_truth     = mean(s$truth),
      bias_naive     = mean(s$naive - s$truth),
      bias_debiased  = mean(s$debiased - s$truth),
      rmse_naive     = sqrt(mean((s$naive - s$truth)^2)),
      rmse_debiased  = sqrt(mean((s$debiased - s$truth)^2)),
      cover_onesided = mean(s$bound <= s$truth),
      cover_twosided = mean(s$ci_lo <= s$truth & s$truth <= s$ci_hi)
    )
  }))
  cat(sprintf("Separated-signal OCs | %d reps, %d candidates, B = %d, n = %d\n",
              SEP_REPS, length(unique(sep$selected)), SEP_B, SEP_N))
  cat(sprintf("distinct selected rules across replications: %d\n",
              length(unique(sep$selected))))
  cat(sprintf("Monte Carlo SE on coverage ~ %.3f\n\n", sqrt(0.95 * 0.05 / SEP_REPS)))
  print(summ, row.names = FALSE, digits = 4)

  ad <- sep[ok & sep$arm == "adaptive", ]
  if (nrow(ad)) {
    cat("\nadaptive-r: distribution of the selected r\n")
    print(table(sprintf("%.2f", ad$r_used)))
  }

  stopifnot(
    nrow(summ) == length(SEP_R) + 1L,
    all(summ$reps > 0.8 * SEP_REPS),          # truth resolvable for most reps
    all(summ$bias_naive > 0)                  # winner's curse present at every r
  )
  cat("\nRead: `bias_naive` is the winner's curse; `bias_debiased` is what remains\n",
      "after correction. A sign flip means over-correction past truth. Coverage is\n",
      "against the SELECTED rule's own truth, which varies by replication.\n", sep = "")
} else {
  cat("SCALE != 'full'; separated-signal cell skipped (Claude Code tier).\n")
}
```
