---
title: "forestsearch efficiency evaluation"
subtitle: "Measurement only --- no package code modified"
format: html
bibliography: []
execute:
  echo: false
  warning: false
---

```{r}
#| label: load
res_dir <- "results"
rd <- function(f) {
  p <- file.path(res_dir, f)
  if (file.exists(p)) readRDS(p) else NULL
}
env <- rd("environment.rds"); G <- rd("G.rds"); M <- rd("M.rds")
P <- rd("P.rds"); E <- rd("E.rds"); ALL <- rd("all.rds")
fmt <- function(x, d = 2) formatC(x, format = "f", digits = d)
```

## Environment

```{r}
if (!is.null(env)) knitr::kable(t(env), col.names = "value") else
  cat("No environment record found --- run `run_eval()` first.")
```

```{r}
#| results: asis
if (!is.null(ALL$guard)) {
  cat("**Read-only contract:** ", ALL$guard$note, "\n\n", sep = "")
}
```

## Stage G --- equivalence gates

Speedups below are reportable only for candidates whose gate passed. A failed
gate means the candidate does not compute the same thing, and its timing is
not evidence of anything.

```{r}
if (!is.null(G)) {
  knitr::kable(G[, c("gate", "what", "n", "pass", "detail")])
} else cat("Stage G not run.")
```

```{r}
#| results: asis
if (!is.null(G) && !all(G$pass, na.rm = TRUE)) {
  cat("\n::: {.callout-important}\n## Gate failure\n",
      "One or more candidates did not reproduce the package result. ",
      "Do not incorporate those candidates on the strength of their timing.\n:::\n")
}
```

## Stage M --- unit costs

```{r}
if (!is.null(M)) {
  m <- M
  m$speedup <- paste0(fmt(m$speedup, 2), "x")
  m$current_ms <- fmt(m$current_ms, 3)
  m$candidate_ms <- fmt(m$candidate_ms, 3)
  knitr::kable(m[, c("id", "what", "n", "current_ms", "candidate_ms",
                     "speedup", "note")])
} else cat("Stage M not run.")
```

## Stage P --- parallel behaviour

### P2: batch_size sweep (the decision point)

`forestsearch()` defaults `stop_threshold = 0.95`; with `sg_focus = "hr"` the
batching logic forces `batch_size = 1`, so each `future_lapply()` call carries
one candidate. The question is empirical: does the finer early stopping save
more than the idle workers and per-candidate serialization cost?

Read the `extra` column: if `evaluated` stays low across all batch sizes,
`batch_size = 1` is doing its job and should be left alone. If `evaluated`
is large and wall-clock falls sharply as `batch_size` rises, the default is
costing real time on this hardware.

**The `selected` field must be identical across every row.** If it is not,
batch size is changing the answer, and that is a correctness question that
takes precedence over any timing.

```{r}
if (!is.null(P)) {
  p2 <- P[P$id == "P2", ]
  if (nrow(p2)) knitr::kable(p2[, c("x", "seconds", "extra")],
                             col.names = c("batch_size", "seconds", "detail"))
} else cat("Stage P not run.")
```

### P1 / P3 / P4

```{r}
if (!is.null(P)) {
  rest <- P[P$id != "P2", ]
  if (nrow(rest)) knitr::kable(rest[, c("id", "what", "x", "xlab",
                                        "seconds", "extra")])
}
```

## Stage E --- measured end-to-end attribution

Shares come from `Rprof` on a sequential run, so they are observed rather than
modelled. `E6` isolates pure formula dispatch --- the ceiling on what swapping
in `coxph.fit()` can remove.

```{r}
if (!is.null(E)) knitr::kable(E) else cat("Stage E not run.")
```

## Decision table

```{r}
gate_ok <- function(ids) {
  if (is.null(G)) return(NA)
  sel <- G$gate %in% ids
  if (!any(sel)) NA else all(G$pass[sel], na.rm = TRUE)
}
spd <- function(id) {
  if (is.null(M)) return(NA_real_)
  v <- M$speedup[M$id == id]
  if (!length(v)) NA_real_ else stats::median(v, na.rm = TRUE)
}

dec <- data.frame(
  stringsAsFactors = FALSE,
  change = c("Fast Cox engine (coxph.fit)",
             "Exact futility stopping",
             "Search scorer without summary.survfit",
             "Vectorised combination pre-screen",
             "GLM via glm.fit",
             "batch_size default at high worker counts",
             "plan() hygiene in subgroup_search.R"),
  gate = c(gate_ok(c("G1", "G2")), gate_ok(c("G3a", "G3b")),
           gate_ok(c("G4a", "G4b")), gate_ok(grep("^G5", G$gate, value = TRUE)),
           gate_ok("G6"), NA, NA),
  measured = c(paste0(fmt(spd("M1")), "x"),
               "see P2 (data-dependent)",
               paste0(fmt(spd("M2")), "x"),
               paste0(fmt(spd("M3.k2")), "x / ", fmt(spd("M3.k3")), "x"),
               paste0(fmt(spd("M4")), "x"),
               "see P2", "see P1"),
  behaviour = c("Pcons bit-identical",
                "decision identical; Pcons becomes a bound for rejected candidates",
                "HR/medians identical; CI differs at ~1e-6 (qnorm vs 1.96)",
                "passing set identical",
                "estimate and SE identical",
                "must verify selected subgroup is invariant",
                "restores caller's plan; no numerical effect")
)
knitr::kable(dec)
```

## What this evaluation did not measure

- The **adjusted** Cox path (`adjust_covariates`, especially `strata()`).
  The fast engine does not apply there and no claim is made about it.
- **DINA and GRF.** These findings concern the FS search and consistency
  stages only.
- Any **GLM outcome family beyond the one gated** (`run_glm = TRUE` covers
  binary/OR). Continuous/MD and count/IRR are untested here.
- Interaction with `use_twostage = TRUE`, which has its own screening logic.
