---
title: "ACTG175 Binary Harm Search: Simulation Study"
format: html
---

This article runs the simulation study for configuration
**`actg175_binary_m1_harm_hrMaxSG_fs1`** (study = ACTG175,
outcome = binary, DGM = m1 [calibrated GLM, target OR(Q) = 2.0],
objective = harm, sg_focus = `hrMaxSG`, FS bundle = fs1).

The article supports three rendering modes, controlled by `sim_mode`
in the setup chunk:

- **`import`** — load a pre-computed bundle (.rds) produced by the
  companion runner script `actg175_binary_m1_harm_hrMaxSG_fs1.R`. Fast.
- **`demo`** — run the simulations inline at small *N* (100 + 100).
  Suitable for self-contained execution by readers who want to
  reproduce the workflow without committing to a full run.
- **`full`** — run the simulations inline at production *N*
  (mirrors the runner script: 1000 + 100). Slow.
- **`auto`** (default) — `import` if a bundle is present at the
  expected path, otherwise `demo`.

Data preparation, DGM calibration, and a single-trial sanity check
always run, regardless of mode, so the tutorial narrative remains
intact. In `import` mode the freshly-built DGM objects are
overridden with the bundle's versions to keep the OC tables
consistent with the simulation output they describe.

```{r setup}
#| code-fold: true
#| code-summary: "Setup and parameters (click to expand)"

library(forestsearch)
library(speff2trial)
library(data.table)
library(gt)
library(ggplot2)
library(doFuture)
library(foreach)

set.seed(8316951L)

# -- Mode --------------------------------------------------------------------
# One of: "import", "demo", "full", "auto".
# "auto" picks "import" if BUNDLE_PATH exists, else "demo".
sim_mode    <- "auto"
BUNDLE_PATH <- "actg175_binary_m1_harm_hrMaxSG_fs1.rds"

if (sim_mode == "auto") {
  sim_mode <- if (file.exists(BUNDLE_PATH)) "import" else "demo"
}
if (sim_mode == "import" && !file.exists(BUNDLE_PATH)) {
  stop(sprintf("sim_mode = 'import' but no bundle found at '%s'.",
               BUNDLE_PATH), call. = FALSE)
}

# -- Simulation control ------------------------------------------------------
# nsims_* are derived from sim_mode.  Edit the per-mode values here if
# different reference scales are needed.
nsims_alt <- switch(sim_mode,
  import = 0L,            # not used (loaded from bundle)
  demo   = 100L,
  full   = 1000L,
  stop(sprintf("Unknown sim_mode: '%s'", sim_mode))
)
nsims_null <- switch(sim_mode,
  import = 0L,
  demo   = 100L,
  full   = 100L
)
n_workers   <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))
sim_n       <- 1000L

cat(sprintf("sim_mode = %s | nsims_alt = %d | nsims_null = %d | sim_n = %d\n",
            sim_mode, nsims_alt, nsims_null, sim_n))

# -- FS parameters -----------------------------------------------------------
c1 <- 1.25           # hr.threshold (OR scale)
c2 <- 1.0            # hr.consistency
fs_pconsistency    <- 0.80
max_look <- 10
fs_splits          <- 1000L
fs_n_min           <- 60L
fs_d0_min          <- 10L
fs_d1_min          <- 0L
fs_maxk            <- 2L
fs_dmin_grf        <- 0.0
fs_seedit          <- 8316951L
fs_sg_focus        <- "hrMaxSG"
fs_use_lasso       <- FALSE
fs_use_grf         <- TRUE
fs_return_selected_cuts_only <- FALSE
fs_is_rct          <- TRUE

# -- GRF standalone parameters ------------------------------------------------
grf_dmin           <- 0.10
grf_maxdepth       <- 2L
grf_sg_criterion   <- "mDiff"

# -- Reporting ---------------------------------------------------------------
sg_notation        <- "harm"

# -- Calibration -------------------------------------------------------------
cal_target_or      <- 2.0

# -- Diagnostic capture ------------------------------------------------------
diag_keep          <- c("fs_full", "candidate_table")
diag_keep_first_n  <- 5L

# -- Plot thresholds ---------------------------------------------------------
sgdist_min_pct_alt  <- 1
sgdist_min_pct_null <- 1
sgdist_top_k        <- 10L
ordist_ylim         <- c(0, 3)

# -- Consolidated parameter lists (used for both single-trial and sims) -------
# fs_params: passed to do.call(forestsearch, ...) for single-trial analyses
# and to run_simulation_analysis(fs_params = ...) for the simulation loop.
# run_simulation_analysis() overrides details/quiet internally.
# conf_force is set after data prep (depends on actg_df factor variables).
fs_params <- list(
  outcome.name           = "y_sim",
  event.name             = "y_sim",
  treat.name             = "treat_sim",
  id.name                = "id",
  outcome_type           = "binary",
  effect_measure         = "OR",
  use_lasso              = fs_use_lasso,
  use_grf                = fs_use_grf,
  return_selected_cuts_only = fs_return_selected_cuts_only,
  max_subgroups_search   = max_look,
  use_twostage           = TRUE,
  hr.threshold           = c1,
  hr.consistency         = c2,
  pconsistency.threshold = fs_pconsistency,
  stop_threshold         = 0.90,
  sg_focus               = fs_sg_focus,
  fs.splits              = fs_splits,
  n.min                  = fs_n_min,
  d0.min                 = fs_d0_min,
  d1.min                 = fs_d1_min,
  maxk                   = fs_maxk,
  dmin.grf               = fs_dmin_grf,
  is.RCT                 = fs_is_rct,
  seedit                 = fs_seedit,
  showten_subgroups      = TRUE,
  details                = TRUE,
  quiet                  = FALSE,
  parallel_args          = list(plan = "sequential", workers = 1,
                                show_message = FALSE)
)

grf_params <- list(
  dmin.grf     = grf_dmin,
  maxdepth     = grf_maxdepth,
  sg.criterion = grf_sg_criterion
)
```

# Data Preparation {#sec-data}

```{r data-prep}
#| code-fold: true
#| code-summary: "ACTG175 data prep (click to expand)"

# ACTG175: arm 1 = ZDV+ddI (experimental), arm 3 = ddI (control)
actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$id <- seq_len(nrow(actg_df))

# Treatment: ZDV+ddI = 1, ddI = 0 (NO switching)
actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)

# Binary outcome: adverse (failure to improve)
actg_df$y_binary <- as.integer(actg_df$cd420 <= actg_df$cd40)
actg_df <- actg_df[!is.na(actg_df$cd420), ]

# -- DGM factor variables -----------------------------------------------------
# z1, z2: TRUE subgroup Q = {cd40 > 421 & wtkg > 82.3}
# (from observed-data FS result — see companion analysis document)

 actg_df$z1  <- as.factor(ifelse(actg_df$cd40 > 375, 1L, 0L))
 actg_df$z2  <- as.factor(ifelse(actg_df$wtkg > 75, 1L, 0L))

#actg_df$z1  <- as.factor(ifelse(actg_df$age > 34, 1L, 0L))
#actg_df$z2  <- as.factor(ifelse(actg_df$preanti <= 750, 1L, 0L))


# z7-z12: truly binary indicators (no continuous counterpart)
actg_df$z7  <- as.factor(actg_df$hemo)
actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs)
actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender)
actg_df$z12 <- as.factor(actg_df$symptom)

actg_df$ar_naive   <- as.factor(ifelse(actg_df$str2 == 0, 1L, 0L))
actg_df$prior_6mo  <- as.factor(ifelse(actg_df$preanti <= 6 * 30.4375, 1L, 0L))

actg_df$flag_harm_ref <- as.integer(actg_df$z1 == 1 & actg_df$z2 == 1)

# -- Analysis covariates (what the analyst sees) -------------------------------
cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
bin_vars  <- c("hemo", "homo", "drugs", "race", "gender", "symptom","ar_naive","prior_6mo")
for (v in bin_vars) actg_df[[v]] <- as.factor(actg_df[[v]])
confounders <- c(cont_vars, bin_vars)

fs_conf_force <- c("ar_naive == 1","prior_6mo == 1")

# Inject conf_force into the consolidated fs_params (defined in setup,
# but conf_force depends on data-prep factor variables)
fs_params$conf_force <- fs_conf_force

# -- DGM variable sets ---------------------------------------------------------
dgm_factor_vars   <- c("z1", "z2", paste0("z", 7:12))
dgm_cont_vars     <- cont_vars
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)

# -- Summary -------------------------------------------------------------------
Q_n   <- sum(actg_df$flag_harm_ref)
Q_N   <- nrow(actg_df)
Q_pct <- 100 * Q_n / Q_N

cat(sprintf("N = %d, Q prevalence = %d / %d (%.1f%%)\n", Q_N, Q_n, Q_N, Q_pct))
cat(sprintf("Adverse rate: overall=%.1f%%, ZDV+ddI=%.1f%%, ddI=%.1f%%\n",
    100 * mean(actg_df$y_binary),
    100 * mean(actg_df$y_binary[actg_df$treat == 1]),
    100 * mean(actg_df$y_binary[actg_df$treat == 0])))
```

# DGM Calibration {#sec-calibration}

Target: OR(Q) = `r cal_target_or` on the adverse scale.  The DGM
includes factor variables (for subgroup definition) and raw continuous
covariates (for a richer prognostic baseline), consistent with the
survival `generate_aft_dgm_flex()` design.

```{r calibrate}
dgm_calibrated <- calibrate_glm_interaction(
  data            = actg_df,
  factor_vars     = dgm_factor_vars,
  continuous_vars = dgm_cont_vars,
  outcome_var     = "y_binary",
  treatment_var   = "treat",
  target_effect   = cal_target_or,
  outcome_type    = "binary",
  effect_measure  = "OR",
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  k_inter_range   = c(0, 3),
  grid_step       = 0.05,
  n_super         = 5000L,
  seed            = 8316951L,
  verbose         = TRUE
)

cat(sprintf("\n  OR(Q):   %.3f (target: %.1f)\n",
    dgm_calibrated$hazard_ratios$harm_subgroup, cal_target_or))
cat(sprintf("  OR(Qc):  %.3f\n", dgm_calibrated$hazard_ratios$no_harm_subgroup))
cat(sprintf("  OR(ITT): %.3f\n", dgm_calibrated$hazard_ratios$overall))
```

```{r null-dgm}
dgm_null <- generate_glm_dgm(
  data            = actg_df,
  factor_vars     = dgm_factor_vars,
  continuous_vars = dgm_cont_vars,
  outcome_var     = "y_binary",
  treatment_var   = "treat",
  outcome_type    = "binary",
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  model           = "null",
  n_super         = 5000L,
  seed            = 8316951L,
  verbose         = TRUE
)
```

# Single Trial Sanity Check {#sec-single-trial}

```{r single-trial-alt}
#| code-fold: true
#| code-summary: "Single simulated trial + FS analysis alt (click to expand)"

df_sim <- simulate_from_glm_dgm(dgm_calibrated, n = sim_n, seed = 42)

itt_fit <- glm(y_sim ~ treat_sim, family = "binomial", data = df_sim)
itt_OR <- exp(coef(itt_fit)["treat_sim"])
itt_event <- with(df_sim,mean(y_sim))

fpr_c1 <- fpr_approx(effect_true = itt_OR, p_event = itt_event, n_min = 60, c1 = c1, 
           n_confounders_cont = length(cont_vars), n_confounders_bin = length(bin_vars),
           effect_type = "ratio", n_eff = max_look)

fs_single <- do.call(forestsearch, c(
  list(df.analysis = df_sim, confounders.name = confounders),
  fs_params
))

if(!is.null(fs_single$sg.harm)){
fs_single_tabs <- sg_tables(fs_single, ndecimals = 3)
fs_single_tabs$sg10_out
fs_single_tabs$tab_estimates
}

```


```{r single-trial-null}
#| code-fold: true
#| code-summary: "Single simulated trial + FS analysis null (click to expand)"

df_sim <- simulate_from_glm_dgm(dgm_null, n = sim_n, seed = 42)

itt_fit <- glm(y_sim ~ treat_sim, family = "binomial", data = df_sim)
itt_OR <- exp(coef(itt_fit)["treat_sim"])
itt_event <- with(df_sim,mean(y_sim))


fpr_c1 <- fpr_approx(effect_true = itt_OR, p_event = itt_event, n_min = 60, c1 = 1.25, 
           n_confounders_cont = length(cont_vars), n_confounders_bin = length(bin_vars),
           effect_type = "ratio", n_eff = max_look)

fs_single <- do.call(forestsearch, c(
  list(df.analysis = df_sim, confounders.name = confounders),
  fs_params
))

if(!is.null(fs_single$sg.harm)){
fs_single_tabs <- sg_tables(fs_single, ndecimals = 3)
fs_single_tabs$sg10_out
fs_single_tabs$tab_estimates
}

```

# Full Simulation Study {#sec-sims}

## Parameter Lists

```{r sim-params}
#| eval: !expr sim_mode != "import"
#| code-fold: true
#| code-summary: "Parallel plan setup (click to expand)"

plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")
cat("fs_params and grf_params defined in setup chunk\n")
```

## H1 Simulations (Calibrated HTE)

The foreach loop below mirrors the H1 loop in
`sim_actg175_binary_m1_harm_hrMaxSG_fs1.R` (the simulation core called
by the companion runner script).  Updates should be applied in both
places.

```{r run-alt}
#| eval: !expr sim_mode != "import"
#| code-fold: true
#| code-summary: "H1 simulation loop (click to expand)"
#| output: false

cat("Running", nsims_alt, "simulations under H1...\n")
t0 <- proc.time()[3]

results_alt_raw <- foreach(
  sim = seq_len(nsims_alt),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_calibrated,
    n_sample         = sim_n,
    confounders_base = confounders,
    run_fs           = TRUE,
    run_fs_grf       = FALSE,
    run_grf          = TRUE,
    fs_params        = fs_params,
    grf_params       = grf_params,
    keep             = diag_keep,
    keep_first_n     = diag_keep_first_n,
    verbose          = FALSE
  )
}

t_alt <- proc.time()[3] - t0
results_alt <- collect_results(results_alt_raw, "H1 alt",
                               keep_diagnostics = diag_keep_first_n > 0L)
n_failed_alt <- attr(results_alt, "n_failed")

cat(sprintf("Completed in %.1f min (%d rows, %d failed)\n",
    t_alt / 60, nrow(results_alt), n_failed_alt))
```

## H0 Simulations (Null — Homogeneous Effect)

```{r run-null}
#| eval: !expr sim_mode != "import"
#| code-fold: true
#| code-summary: "H0 simulation loop (click to expand)"
#| output: false

cat("Running", nsims_null, "simulations under H0...\n")
t0 <- proc.time()[3]

results_null_raw <- foreach(
  sim = seq_len(nsims_null),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_null,
    n_sample         = sim_n,
    confounders_base = confounders,
    run_fs           = TRUE,
    run_fs_grf       = FALSE,
    run_grf          = TRUE,
    fs_params        = fs_params,
    grf_params       = grf_params,
    verbose          = FALSE
  )
}

t_null <- proc.time()[3] - t0
results_null <- collect_results(results_null_raw, "H0 null")
n_failed_null <- attr(results_null, "n_failed")

cat(sprintf("Completed in %.1f min (%d rows, %d failed)\n",
    t_null / 60, nrow(results_null), n_failed_null))
```

## Mode Handoff

```{r mode-handoff}
#| code-fold: true
#| code-summary: "Bundle import or sim-result handoff (click to expand)"

if (sim_mode == "import") {
  bundle <- readRDS(BUNDLE_PATH)

  # Override the freshly-built DGMs with the bundle's versions so that
  # downstream OC tables match the simulation results they describe.
  dgm_calibrated <- bundle$dgm$calibrated
  dgm_null       <- bundle$dgm$null

  results_alt    <- bundle$results$alt
  results_null   <- bundle$results$null
  t_alt          <- bundle$meta$t_alt_sec
  t_null         <- bundle$meta$t_null_sec
  nsims_alt      <- bundle$meta$n_alt
  nsims_null     <- bundle$meta$n_null
  n_failed_alt   <- bundle$meta$n_failed_alt
  n_failed_null  <- bundle$meta$n_failed_null

  cat(sprintf("Imported bundle: %s\n", BUNDLE_PATH))
  cat(sprintf("  config_id : %s\n", bundle$meta$config_id))
  cat(sprintf("  timestamp : %s\n", format(bundle$meta$timestamp)))
  cat(sprintf("  n_alt     : %d (failed: %d)\n",
              nsims_alt, n_failed_alt))
  cat(sprintf("  n_null    : %d (failed: %d)\n",
              nsims_null, n_failed_null))
  if (!is.null(bundle$meta$package_version)) {
    cat(sprintf("  forestsearch %s\n", bundle$meta$package_version))
  }
  if (!is.null(bundle$meta$config_id) &&
      bundle$meta$config_id != "actg175_binary_m1_harm_hrMaxSG_fs1") {
    warning(sprintf(
      "Bundle config_id ('%s') does not match this article ('%s').",
      bundle$meta$config_id, "actg175_binary_m1_harm_hrMaxSG_fs1"
    ), call. = FALSE)
  }
} else {
  cat(sprintf("Inline simulation complete (%s mode).\n", sim_mode))
}
```

# Results {#sec-results}

## Metrics Glossary {#sec-glossary}

Classification metrics follow the definitions in León et al.
(2024, Section 3).

### Detection

**Found** ($\text{any}(Q)$): proportion of simulations where
the method identifies *any* harm subgroup
($\texttt{any.H} = 1$).

- **Under H1** (alternative): $Q$ exists with a differential
  treatment effect.  $\text{any}(Q)$ is **detection power**.
- **Under H0** (null): $Q = \emptyset$ and $Q^c$ is the ITT
  population.  $\text{any}(Q)$ is the **false positive rate
  (FPR)** for falsely declaring that a harm subgroup exists.

### Classification Under H1

Under the alternative, the true harm subgroup $Q$ exists
($Q = \{\text{cd40} > 421 \cap \text{wtkg} > 82.3\}$,
prevalence $\approx$ `r sprintf("%.0f", Q_pct)`%)
with a genuinely different treatment effect (OR(Q) =
`r cal_target_or`).  Following León et al. (2024, Section 3),
define for estimated subgroup $\hat{Q}$:

$$
\text{sens}(\hat{Q}) = \frac{\#\{i \in \hat{Q} \cap Q\}}{\#\{i \in Q\}},
\quad
\text{ppv}(\hat{Q}) = \frac{\#\{i \in \hat{Q} \cap Q\}}{\#\{i \in \hat{Q}\}},
$$

with analogous measures for the complement $\hat{Q}^c$.

The operating-characteristics tables report these as:

| Table column | León et al. notation | Interpretation |
|---|---|---|
| **Sen** | $\text{sens}(\hat{Q})$ | Fraction of true $Q$ captured by $\hat{Q}$ |
| **Spec** | $\text{sens}(\hat{Q}^c)$ | Fraction of true $Q^c$ correctly in $\hat{Q}^c$ |
| **PPV** | $\text{ppv}(\hat{Q})$ | Fraction of $\hat{Q}$ truly in $Q$ |
| **NPV** | $\text{ppv}(\hat{Q}^c)$ | Fraction of $\hat{Q}^c$ truly in $Q^c$ |

Classification metrics are averaged across **all** simulations
(León et al. 2024, Table 1).  When no subgroup is detected
($\hat{Q} = \emptyset$), the simulation contributes
$\text{sens}(\hat{Q}) = 0$ and $\text{ppv}(\hat{Q}) = 0$
(since no member of $Q$ is captured and no subjects are
assigned to $\hat{Q}$).  This convention is implemented
at the source in `run_simulation_analysis()`, matching
the original `fs.estimates.out()` function used in
León et al. (2024).

::: {.callout-note}
## Classification metrics with discovered cuts

Because cuts are discovered from the data in this setup, a
$\hat{Q}$ that captures "roughly the right region" but with
slightly shifted boundaries will still receive high sens/PPV.
Expect these metrics to be lower than under the discretised
(oracle-cut) setup — the gap quantifies the cost of not knowing
the true cuts in advance.
:::

### Classification Under H0

Under the null, $Q = \emptyset$ --- no subgroup with a
differential treatment effect exists.  The complement $Q^c$
is the entire ITT population.  Per León et al. (2024,
Section 3), $\text{sens}(\hat{Q})$ is **undefined** (shown
as "—" in the table) and $\text{ppv}(\hat{Q}) \equiv 0$.
For the complement, $\text{sens}(\hat{Q}^c) = |\hat{Q}^c|/N$
and $\text{ppv}(\hat{Q}^c) = 1$ since $Q^c$ = ITT.

The only operationally meaningful metric under H0 is the
**detection rate (FPR)**.

### Effect Estimates

All effect estimates are on the adverse scale:
OR $> 1$ = treatment (ZDV+ddI) increases the adverse outcome
(non-improvement) = **harm**.  No inversion or treatment
switching is applied.

# Operating Characteristics {#sec-oc}

## H1: Detection + Classification

```{r oc-alt}
format_oc_results(
  results_alt,
  metrics   = c("detection", "classification", "subgroup_size"),
  digits    = 2,
  digits_hr = 2,
  title     = "ACTG175 Binary: Harm Subgroup Identification",
  dgm = dgm_calibrated,
  subgroup_notation = sg_notation
)
```

### True Subgroup Sizes Across Replicates

```{r true-size-summary}
#| code-fold: true

# One row per replicate-method combination, but sizeH_true is identical
# across method rows within the same sim.  Deduplicate on sim.
truth_by_sim <- unique(results_alt[, c("sim", "sizeH_true", "sizeHc_true",
                                        "propH_true", "propHc_true")])

callout_Q <- sprintf(
  "|Q|:  mean = %.1f (SD %.2f),  range = [%d, %d],  prop = %.1f%%",
  mean(truth_by_sim$sizeH_true),
  sd(truth_by_sim$sizeH_true),
  min(truth_by_sim$sizeH_true),
  max(truth_by_sim$sizeH_true),
  100 * mean(truth_by_sim$propH_true))

callout_Qc <- sprintf(
  "|Qc|: mean = %.1f (SD %.2f),  range = [%d, %d],  prop = %.1f%%",
  mean(truth_by_sim$sizeHc_true),
  sd(truth_by_sim$sizeHc_true),
  min(truth_by_sim$sizeHc_true),
  max(truth_by_sim$sizeHc_true),
  100 * mean(truth_by_sim$propHc_true))

cat(sprintf("True subgroup sizes across %d H1 replicates (N = %d each):\n",
            nrow(truth_by_sim), sim_n))
cat("  ", callout_Q,  "\n")
cat("  ", callout_Qc, "\n")
```

::: {.callout-note title="Reference sizes for the estimation tables below"}
- **|Q| (true harm subgroup)**: mean `r sprintf("%.1f", mean(truth_by_sim$sizeH_true))`
  per trial (SD `r sprintf("%.2f", sd(truth_by_sim$sizeH_true))`,
  range `r min(truth_by_sim$sizeH_true)`–`r max(truth_by_sim$sizeH_true)`,
  average proportion `r sprintf("%.1f%%", 100 * mean(truth_by_sim$propH_true))`).
- **|Qc| (complement)**: mean `r sprintf("%.1f", mean(truth_by_sim$sizeHc_true))`
  per trial (SD `r sprintf("%.2f", sd(truth_by_sim$sizeHc_true))`,
  range `r min(truth_by_sim$sizeHc_true)`–`r max(truth_by_sim$sizeHc_true)`,
  average proportion `r sprintf("%.1f%%", 100 * mean(truth_by_sim$propHc_true))`).

Compare to $\widehat{|H|}$ (the FS-identified subgroup size) reported in
the row-group labels of the estimation tables below.  A ratio
$\widehat{|H|} \big/ \overline{|Q|} \approx 1$ indicates the method is
recovering a subgroup close to the true size; values much smaller
suggest FS is identifying a strict subset of $Q$; values larger suggest
FS is absorbing some $Q^c$ subjects into $\hat H$.
:::

## H1: Estimation Tables

```{r oc-alt-estimation}
build_estimation_table(
  results         = results_alt,
  dgm             = dgm_calibrated,
  analysis_method = "FS",
  subgroup_notation = sg_notation,
  title           = "FS: Effect Estimates Under H1"
)

build_estimation_table(
  results         = results_alt,
  dgm             = dgm_calibrated,
  analysis_method = "GRF",
  subgroup_notation = sg_notation,
  title           = "GRF: Effect Estimates Under H1"
)
```

## H1: Classification Tables

```{r oc-alt-classification}
scenario_alt <- list(
  H1 = list(
    results    = results_alt,
    dgm        = dgm_calibrated,
    hypothesis = "alt",
    n_sample   = sim_n,
    label      = "ACTG175"
  )
)

build_classification_table(
  scenario_alt,
  analyses          = "FS",
  subgroup_notation = sg_notation,
  title             = "FS: Classification Under H1"
)

build_classification_table(
  scenario_alt,
  analyses          = "GRF",
  subgroup_notation = sg_notation,
  title             = "GRF: Classification Under H1"
)
```

## H1: Subgroup Distribution

```{r oc-alt-sgdist, fig.width=9, fig.height=6}
plot_sg_distribution(
  results_alt[results_alt$analysis == "FS", ],
  any_col   = "any.H",
  label_col = "sg.def",
  min_pct   = sgdist_min_pct_alt,
  top_k     = sgdist_top_k,
  title     = "FS: Identified Subgroup Distribution (H1)"
)
```

## H1: OR Distribution

```{r oc-alt-ordist, fig.width=9, fig.height=4.5}
or_Q_true <- dgm_calibrated$hazard_ratios$harm_subgroup

plot_effect_distribution(
  results           = results_alt,
  analysis_method   = "FS",
  effect_measure    = "OR",
  subgroup_notation = sg_notation,
  reference_effect  = or_Q_true,
  ylim              = ordist_ylim,
  title             = "FS: OR Estimates Across Simulations (H1)"
)
```

## H0: Type I Error

```{r oc-null}
format_oc_results(
  results_null,
  dgm = dgm_null,
  metrics   = c("detection", "classification"),
  digits    = 2,
  title     = "ACTG175 Binary: Type I Error (Null)",
   subgroup_notation = sg_notation
)
```

## H0: Estimation Tables

```{r oc-null-estimation}
build_estimation_table(
  results         = results_null,
  dgm             = dgm_null,
  analysis_method = "FS",
  subgroup_notation = sg_notation,
  title           = "FS: Effect Estimates Under H0"
)

build_estimation_table(
  results         = results_null,
  dgm             = dgm_null,
  analysis_method = "GRF",
  subgroup_notation = sg_notation,
  title           = "GRF: Effect Estimates Under H0"
)
```

## H0: Subgroup Distribution

```{r oc-null-sgdist, fig.width=9, fig.height=6}
plot_sg_distribution(
  results_null[results_null$analysis == "FS", ],
  any_col   = "any.H",
  label_col = "sg.def",
  min_pct   = sgdist_min_pct_null,
  top_k     = sgdist_top_k,
  title     = "FS: Identified Subgroup Distribution (H0)"
)
```

# Timing

```{r timing}
cat(sprintf("Mode: %s\n", sim_mode))
cat(sprintf("H1: %.1f min (%d sims)\n", t_alt / 60, nsims_alt))
cat(sprintf("H0: %.1f min (%d sims)\n", t_null / 60, nsims_null))
plan(sequential)
```
