---
title: "FS Parameter Bundles: Comparison and Differences"
subtitle: "Side-by-side view of fs1, fs2, ... regimes"
format:
  html:
    toc: true
    toc-depth: 3
---

This document loads every `fs*_params.R` file in this directory and
renders two views of the parameter values:

- **Section 2 — Differences only**: the compact view. One row per
  parameter that varies across the bundles. Best for "what's
  different between fs1 and fs2?"
- **Section 3 — All parameters**: the comprehensive view. Every
  parameter, with cells that differ across bundles highlighted.
  Best for an audit-trail / one-page reference.

The document auto-discovers `fs*_params.R` files. Add a new bundle
by dropping `fs2_params.R` into this directory and re-rendering.

```{r setup}
#| code-fold: true
#| code-summary: "Setup, discovery, and parameter loading (click to expand)"

library(data.table)
library(gt)

# Base R adds %||% in 4.4; define a fallback for older R.
if (!exists("%||%", mode = "function")) {
  `%||%` <- function(x, y) if (is.null(x)) y else x
}

# -- Auto-discover fs*_params.R files ---------------------------------------
# Files are sourced from the current working directory of this qmd
# (i.e., quarto/simulations/<study>/_fsparams/). Each must define a
# function get_fsN_params(sg_focus, ...).
fs_files <- list.files(pattern = "^fs[0-9]+_params\\.R$",
                       full.names = FALSE)

if (length(fs_files) == 0L) {
  stop("No fs*_params.R files found in this directory.", call. = FALSE)
}

# Extract bundle names (e.g., "fs1", "fs2") from filenames
fs_names <- sub("_params\\.R$", "", fs_files)
names(fs_files) <- fs_names

cat(sprintf("Discovered %d FS bundle(s): %s\n",
            length(fs_files), paste(fs_names, collapse = ", ")))
```

```{r load-bundles}
#| code-fold: true
#| code-summary: "Source each file, call its function, capture parameters (click to expand)"

# We call each bundle's function with TWO sg_focus values to surface any
# sg_focus-dependent parameters (e.g., stop_threshold returns NULL for
# hrMaxSG/hrMinSG and 0.90 otherwise). Choosing one of each:
#   - "hr"      -> the early-stopping branch
#   - "hrMaxSG" -> the neighborhood-criterion branch
sg_focus_probes <- c(early_stopping = "hr", neighborhood = "hrMaxSG")

# Source each file in a fresh environment so functions don't collide
load_bundle <- function(file, fs_name) {
  env <- new.env()
  source(file, local = env)
  fn_name <- paste0("get_", fs_name, "_params")
  if (!exists(fn_name, envir = env, inherits = FALSE)) {
    stop(sprintf("File %s does not define function %s().",
                 file, fn_name), call. = FALSE)
  }
  fn <- get(fn_name, envir = env)
  # Call with each sg_focus probe; return as named list of parameter lists
  lapply(sg_focus_probes, function(sf) fn(sg_focus = sf))
}

bundles <- mapply(load_bundle,
                  file    = fs_files,
                  fs_name = fs_names,
                  SIMPLIFY = FALSE)
# bundles[[fs_name]][[probe_name]] is a parameter list

# Extract regime descriptions from each file's header comment
extract_regime_description <- function(file) {
  lines <- readLines(file, warn = FALSE)
  # Find lines starting with "# Regime:" and capture until next blank
  # comment line (a line that is just "#" with optional spaces)
  start <- grep("^#\\s+Regime:", lines)
  if (length(start) == 0L) return(NA_character_)
  start <- start[1]
  end <- start
  for (i in (start + 1):length(lines)) {
    if (!grepl("^#", lines[i])) break
    if (grepl("^#\\s*$", lines[i])) break
    end <- i
  }
  desc <- paste(sub("^#\\s?", "", lines[start:end]), collapse = "\n")
  desc
}

regime_descriptions <- vapply(fs_files, extract_regime_description,
                              character(1))
```

# 1. Regime Descriptions {#sec-regimes}

What each FS bundle is designed for. Pulled from the header comment
block of each `fs*_params.R` file.

```{r regime-descriptions}
#| results: asis
for (i in seq_along(fs_names)) {
  cat(sprintf("\n## %s\n\n", fs_names[i]))
  if (is.na(regime_descriptions[i])) {
    cat("_No `Regime:` line found in header comment._\n")
  } else {
    cat("```\n", regime_descriptions[i], "\n```\n", sep = "")
  }
}
```

```{r build-comparison-data}
#| code-fold: true
#| code-summary: "Build the wide comparison table (click to expand)"

# Collect all parameter names across all bundles. We use the
# "early_stopping" probe as the canonical view (so stop_threshold
# appears as 0.90 rather than NULL).
canonical_probe <- "early_stopping"

format_value <- function(x) {
  if (is.null(x)) return("NULL")
  if (is.list(x)) {
    # parallel_args is a list — render compactly
    parts <- vapply(seq_along(x),
                    function(i) sprintf("%s=%s",
                                        names(x)[i],
                                        format_value(x[[i]])),
                    character(1))
    return(paste0("list(", paste(parts, collapse = ", "), ")"))
  }
  if (length(x) > 1L) return(paste0("c(",
                                    paste(vapply(x, format_value,
                                                 character(1)),
                                          collapse = ", "),
                                    ")"))
  if (is.character(x)) return(sprintf('"%s"', x))
  if (is.logical(x))   return(if (x) "TRUE" else "FALSE")
  format(x)
}

# Get the union of parameter names, in the order they appear in the
# first bundle (with any extra params from later bundles appended)
param_order <- character(0)
for (fs_name in fs_names) {
  pl <- bundles[[fs_name]][[canonical_probe]]
  param_order <- union(param_order, names(pl))
}

# Build a data.table: one row per parameter, one column per bundle
wide_dt <- data.table::data.table(parameter = param_order)
for (fs_name in fs_names) {
  pl <- bundles[[fs_name]][[canonical_probe]]
  wide_dt[[fs_name]] <- vapply(param_order,
                               function(p) format_value(pl[[p]]),
                               character(1))
}

# Add a "differs" flag: TRUE if any pair of bundle values differs
if (length(fs_names) >= 2L) {
  wide_dt[, differs := apply(.SD, 1, function(row) length(unique(row)) > 1L),
          .SDcols = fs_names]
} else {
  wide_dt[, differs := FALSE]
}

# Also add a note column for parameters with sg_focus-dependent values
# (i.e., those that differ between the two probes within a bundle)
note_for <- function(p) {
  notes <- character(0)
  for (fs_name in fs_names) {
    val_a <- bundles[[fs_name]][["early_stopping"]][[p]]
    val_b <- bundles[[fs_name]][["neighborhood"]][[p]]
    if (!identical(val_a, val_b)) {
      notes <- c(notes, sprintf(
        "%s: %s for hr/maxSG/minSG; %s for hrMaxSG/hrMinSG",
        fs_name, format_value(val_a), format_value(val_b)
      ))
    }
  }
  if (length(notes) == 0L) "" else paste(notes, collapse = "; ")
}
wide_dt[, sg_focus_dependent := vapply(parameter, note_for, character(1))]
```

# 2. Differences Only {#sec-diffs}

The compact view: only parameters whose value differs across at
least one pair of bundles.

```{r diffs-only}
diffs_dt <- wide_dt[differs == TRUE]

if (nrow(diffs_dt) == 0L) {
  cat("**All bundles share identical parameter values.**\n\n",
      "(This is expected when only one bundle is defined, or when ",
      "two bundles happen to be identical.)")
} else {
  diffs_dt[, .SD, .SDcols = c("parameter", fs_names)] |>
    gt::gt() |>
    gt::tab_header(
      title    = "Parameters that differ across FS bundles",
      subtitle = sprintf("%d of %d parameters vary",
                         nrow(diffs_dt), nrow(wide_dt))
    ) |>
    gt::cols_label(parameter = "Parameter") |>
    gt::tab_style(
      style    = gt::cell_text(weight = "bold"),
      locations = gt::cells_body(columns = parameter)
    )
}
```

# 3. All Parameters {#sec-all}

The comprehensive view. Every parameter; rows where values differ
across bundles are flagged in the "Differs" column. The
`sg_focus_dependent` column flags parameters that vary by sg_focus
*within* a bundle (e.g., `stop_threshold`).

```{r all-params}
display_dt <- wide_dt[, .SD,
                     .SDcols = c("parameter", fs_names,
                                 "differs", "sg_focus_dependent")]

display_dt |>
  gt::gt() |>
  gt::tab_header(
    title    = "All FS parameters",
    subtitle = sprintf("%d parameters across %d bundle(s)",
                       nrow(wide_dt), length(fs_names))
  ) |>
  gt::cols_label(
    parameter          = "Parameter",
    differs            = "Differs?",
    sg_focus_dependent = "sg_focus dependent (within bundle)"
  ) |>
  gt::fmt(columns = differs,
          fns = function(x) ifelse(x, "✓", "")) |>
  gt::tab_style(
    style     = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(columns = parameter)
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#ffe8cc"),
    locations = gt::cells_body(rows = which(display_dt$differs))
  )
```

# 4. Notes {#sec-notes}

- **`sg_focus_dependent` parameters** vary by `sg_focus` within a
  single bundle. The most important example is `stop_threshold`,
  which the package requires to be `NULL` when `sg_focus` is
  `"hrMaxSG"` or `"hrMinSG"` (neighborhood-based selection is
  incompatible with early stopping). The `fs*_params.R` function
  applies this rule internally.
- **Probe values used**: this article calls each bundle's function
  with `sg_focus = "hr"` (early-stopping branch, used for the
  display values in Sections 2-3) and `sg_focus = "hrMaxSG"`
  (neighborhood branch, used to detect `sg_focus`-dependent
  parameters). Other `sg_focus` values produce identical results
  to one of these two branches.
- **To add a new bundle**: drop `fsN_params.R` into this directory
  with a function `get_fsN_params()` and re-render. The article
  picks it up automatically.
