---
title: "Extreme Subgroup Effects Under a Uniform Treatment Benefit"
subtitle: "A Simulation Study Based on Observed Trial Data"
author: "Larry F. León"
date: today
format:
  beamer:
    theme: Madrid
    colortheme: default
    aspectratio: 169
    fontsize: 11pt
    navigation: empty
    section-titles: true
    toc: true
    toc-depth: 1
    keep-tex: true
    include-in-header:
      text: |
        \definecolor{UBCblue}{RGB}{12, 35, 68}
        \definecolor{forestgreen}{RGB}{34, 139, 34}
        \definecolor{caliborange}{RGB}{192, 57, 43}
        \definecolor{resulteal}{RGB}{20, 143, 119}
        \definecolor{randomorange}{RGB}{230, 126, 34}
        \definecolor{trueHRred}{RGB}{178, 34, 34}
        \setbeamercolor{title}{fg=white,bg=UBCblue}
        \setbeamercolor{frametitle}{fg=white,bg=UBCblue}
        \setbeamercolor{structure}{fg=UBCblue}
        \setbeamercolor{block title}{bg=UBCblue!85,fg=white}
        \setbeamercolor{block body}{bg=UBCblue!5}
        \setbeamercolor{block title alerted}{bg=caliborange!85,fg=white}
        \setbeamercolor{block body alerted}{bg=caliborange!5}
        \setbeamercolor{block title example}{bg=forestgreen!75,fg=white}
        \setbeamercolor{block body example}{bg=forestgreen!5}
        \usepackage{booktabs}
        \setbeamertemplate{navigation symbols}{}
execute:
  echo: false
  warning: false
  message: false
---


```{r setup}
#| include: false
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = "center", dev = "pdf")

# =========================================================================
# CONTROL PARAMETERS
# =========================================================================
n_sims_null <- 1000L          # 5000 for publication quality
results_rds <- NULL          # set to e.g. "results/extreme_sims_5000.rds"
hr_true     <- 0.70

# Forest-plot sizing (beamer-friendly)
fp_row_height  <- 0.32
fp_overhead_in <- 1.2
fp_base_size   <- 9
fp_point_size  <- 1.8
fp_line_size   <- 0.55
```


```{r load-utilities}
#| include: false
library(forestsearch)
library(survival)
library(ggplot2)
library(dplyr)
library(patchwork)
theme_set(theme_minimal(base_size = 10))

# Source the standalone utilities file (must be in same directory)
# Provides: summarize_extreme_sims(), plot_hr_forest(), plot_ub_forest(),
#           save_sim_results(), load_sim_results(), compare_extreme_sims(),
#           forest_fig_height()
source("summarize_extreme_sims.R")
```


```{r run-or-load-simulations}
#| include: false

# =========================================================================
# Either load pre-computed results or run fresh simulations.
# NEW-RESULTS WORKFLOW:
#   1. Run once with n_sims_null = 5000 and results_rds set
#   2. On subsequent renders the RDS loads in seconds
#   3. To update: delete the RDS and re-render
# =========================================================================

loaded_from_rds <- FALSE
if (!is.null(results_rds) && file.exists(results_rds)) {
  saved <- load_sim_results(results_rds)
  sim_hrs <- saved$sim_hrs; sim_ubs <- saved$sim_ubs
  sim_ns  <- saved$sim_ns;  subgroups <- saved$subgroups
  n_sims_null <- saved$n_sims
  if (!is.null(saved$hr_true)) hr_true <- saved$hr_true
  loaded_from_rds <- TRUE; rm(saved)
}

if (!loaded_from_rds) {
  data(gbsg, package = "survival")
  gbsg$time_months <- gbsg$rfstime / 30.4375
  gbsg$treat <- gbsg$hormon

  set.seed(99)
  dgm_null <- generate_aft_dgm_flex(
    data = gbsg,
    continuous_vars = c("age","size","nodes","pgr","er"),
    factor_vars = c("meno","grade"),
    outcome_var = "time_months", event_var = "status",
    treatment_var = "hormon", subgroup_vars = NULL,
    model = "null", n_super = 5000, seed = 99, verbose = FALSE)

  cal_null <- calibrate_cens_adjust(
    dgm = dgm_null, target = "rate", n = 800,
    analysis_time = 84, max_entry = 24,
    seed = 42, n_eval = 1500, verbose = FALSE)

  gbsg$flag_itt <- 1L
  age_med <- median(gbsg$age); size_med <- median(gbsg$size)
  nodes_q3 <- quantile(gbsg$nodes, 0.75)
  pgr_med <- median(gbsg$pgr); er_cut <- 20

  subgroups <- list(
    list(id="flag_itt == 1", name="All Patients", grp="ITT"),
    list(id="meno == 1", name="Post-menopausal", grp="Clinical"),
    list(id="meno == 0", name="Pre-menopausal", grp="Clinical"),
    list(id="grade == 3", name="Grade 3", grp="Clinical"),
    list(id="grade != 3", name="Grade 1/2", grp="Clinical"),
    list(id="age <= age_med", name="Age (young)", grp="Continuous"),
    list(id="age > age_med", name="Age (older)", grp="Continuous"),
    list(id="age <= 50", name="Age <= 50", grp="Continuous"),
    list(id="age > 50", name="Age > 50", grp="Continuous"),
    list(id="size <= size_med", name="Tumour size (small)", grp="Continuous"),
    list(id="size > size_med", name="Tumour size (large)", grp="Continuous"),
    list(id="nodes == 0", name="Node-negative", grp="Continuous"),
    list(id="nodes > 0 & nodes <= 3", name="Nodes 1-3", grp="Continuous"),
    list(id="nodes > nodes_q3", name="High nodes (>Q3)", grp="Continuous"),
    list(id="pgr <= pgr_med", name="PGR low", grp="Continuous"),
    list(id="pgr > pgr_med", name="PGR high", grp="Continuous"),
    list(id="er <= er_cut", name="ER-low (<20)", grp="Continuous"),
    list(id="er > er_cut", name="ER-high (>=20)", grp="Continuous"),
    list(id="meno==0 & grade==3", name="Pre-meno / Grade 3", grp="Interaction: meno x grade"),
    list(id="meno==1 & grade==3", name="Post-meno / Grade 3", grp="Interaction: meno x grade"),
    list(id="meno==0 & grade!=3", name="Pre-meno / Grade 1-2", grp="Interaction: meno x grade"),
    list(id="meno==1 & grade!=3", name="Post-meno / Grade 1-2", grp="Interaction: meno x grade"),
    list(id="meno==0 & age<=50", name="Pre-meno / Age<=50", grp="Interaction: meno x age"),
    list(id="meno==0 & age>50", name="Pre-meno / Age>50", grp="Interaction: meno x age"),
    list(id="meno==1 & age<=age_med", name="Post-meno / Age(young)", grp="Interaction: meno x age"),
    list(id="meno==1 & age>age_med", name="Post-meno / Age(older)", grp="Interaction: meno x age"),
    list(id="meno==0 & er<=er_cut", name="Pre-meno / ER-low", grp="Interaction: meno x ER"),
    list(id="meno==0 & er>er_cut", name="Pre-meno / ER-high", grp="Interaction: meno x ER"),
    list(id="meno==1 & er<=er_cut", name="Post-meno / ER-low", grp="Interaction: meno x ER"),
    list(id="meno==1 & er>er_cut", name="Post-meno / ER-high", grp="Interaction: meno x ER"),
    list(id="grade==3 & nodes==0", name="Grade 3 / Node-neg", grp="Interaction: grade x nodes"),
    list(id="grade==3 & nodes>0", name="Grade 3 / Node-pos", grp="Interaction: grade x nodes"),
    list(id="grade!=3 & nodes==0", name="Grade 1-2 / Node-neg", grp="Interaction: grade x nodes"),
    list(id="grade!=3 & nodes>0", name="Grade 1-2 / Node-pos", grp="Interaction: grade x nodes"),
    list(id="grade==3 & nodes>nodes_q3", name="Grade 3 / High nodes", grp="Interaction: grade x nodes"),
    list(id="grade==3 & er<=er_cut", name="Grade 3 / ER-low", grp="Interaction: grade x ER"),
    list(id="grade==3 & er>er_cut", name="Grade 3 / ER-high", grp="Interaction: grade x ER"),
    list(id="grade!=3 & er<=er_cut", name="Grade 1-2 / ER-low", grp="Interaction: grade x ER"),
    list(id="grade!=3 & er>er_cut", name="Grade 1-2 / ER-high", grp="Interaction: grade x ER"),
    list(id="grade==3 & pgr<=pgr_med", name="Grade 3 / PGR low", grp="Interaction: grade x PGR"),
    list(id="grade==3 & pgr>pgr_med", name="Grade 3 / PGR high", grp="Interaction: grade x PGR"),
    list(id="grade!=3 & pgr<=pgr_med", name="Grade 1-2 / PGR low", grp="Interaction: grade x PGR"),
    list(id="grade!=3 & pgr>pgr_med", name="Grade 1-2 / PGR high", grp="Interaction: grade x PGR"),
    list(id="meno==0 & age<=50 & er<=er_cut", name="Pre-meno/Yng/ER-low", grp="3-way"),
    list(id="meno==0 & age<=50 & er>er_cut", name="Pre-meno/Yng/ER-high", grp="3-way"),
    list(id="meno==1 & grade==3 & er<=er_cut", name="Post-meno/G3/ER-low", grp="3-way"),
    list(id="meno==1 & grade==3 & nodes>0", name="Post-meno/G3/Node+", grp="3-way"),
    list(id="grade==3 & nodes>0 & er<=er_cut", name="G3/Node+/ER-low", grp="3-way"),
    list(id="size>size_med & nodes>0", name="Large/Node-pos", grp="Interaction: size x nodes"),
    list(id="size>size_med & nodes==0", name="Large/Node-neg", grp="Interaction: size x nodes"),
    list(id="size<=size_med & nodes>0", name="Small/Node-pos", grp="Interaction: size x nodes"),
    list(id="size<=size_med & nodes==0", name="Small/Node-neg", grp="Interaction: size x nodes"),
    list(id="random60==1", name="random60", grp="Random (N~60)"),
    list(id="random40==1", name="random40", grp="Random (N~40)"),
    list(id="random20==1", name="random20", grp="Random (N~20)"),
    list(id="random15==1", name="random15", grp="Random (N~15)"))

  cox_sR <- function(data) {
    fit <- tryCatch(
      coxph(Surv(y_sim, event_sim) ~ treat_sim + strata(grade), data = data),
      error = function(e) NULL)
    if (is.null(fit) || nrow(summary(fit)$conf.int) == 0)
      return(c(NA_real_, NA_real_))
    ci <- summary(fit)$conf.int[1, ]
    c(ci["exp(coef)"], ci["upper .95"])
  }

  sg_names <- sapply(subgroups, `[[`, "name")
  sg_ids <- sapply(subgroups, `[[`, "id")
  n_sg <- length(subgroups)
  sim_hrs <- matrix(NA_real_, n_sims_null, n_sg, dimnames=list(NULL, sg_names))
  sim_ubs <- matrix(NA_real_, n_sims_null, n_sg, dimnames=list(NULL, sg_names))
  sim_ns  <- matrix(NA_real_, n_sims_null, n_sg, dimnames=list(NULL, sg_names))

  for (ss in seq_len(n_sims_null)) {
    df_s <- simulate_from_dgm(dgm=dgm_null, n=nrow(gbsg),
      analysis_time=84, max_entry=24,
      cens_adjust=cal_null$cens_adjust, seed=ss)
    df_s$flag_itt <- 1L
    df_s$age_med <- age_med; df_s$size_med <- size_med
    df_s$nodes_q3 <- nodes_q3; df_s$pgr_med <- pgr_med; df_s$er_cut <- er_cut
    set.seed(ss + 1e6L)
    n_s <- nrow(df_s)
    r_idx <- sample.int(n_s, min(60L, n_s), replace=FALSE)
    df_s$random60 <- as.integer(seq_len(n_s) %in% r_idx)
    df_s$random40 <- as.integer(seq_len(n_s) %in% r_idx[seq_len(min(40,length(r_idx)))])
    df_s$random20 <- as.integer(seq_len(n_s) %in% r_idx[seq_len(min(20,length(r_idx)))])
    df_s$random15 <- as.integer(seq_len(n_s) %in% r_idx[seq_len(min(15,length(r_idx)))])
    for (gg in seq_len(n_sg)) {
      df_sg <- tryCatch(subset(df_s, eval(parse(text=sg_ids[[gg]]))),
                        error=function(e) NULL)
      if (is.null(df_sg) || nrow(df_sg) < 5) next
      sim_ns[ss,gg] <- nrow(df_sg)
      r <- cox_sR(df_sg)
      sim_hrs[ss,gg] <- r[1]; sim_ubs[ss,gg] <- r[2]
    }
  }

  if (!is.null(results_rds))
    save_sim_results(sim_hrs, sim_ubs, sim_ns, subgroups,
                     file=results_rds, hr_true=hr_true,
                     metadata=list(dgm_seed=99, cens_adjust=cal_null$cens_adjust))
}
```


```{r compute-summaries}
#| include: false

# One call produces everything for tables and plots.
S <- summarize_extreme_sims(sim_hrs, sim_ubs, sim_ns, subgroups,
                             hr_true = hr_true)

fh_ub_single <- forest_fig_height(S, "single", "ub", fp_row_height, fp_overhead_in)
fh_ub_combo  <- forest_fig_height(S, "combo",  "ub", fp_row_height, fp_overhead_in)
fh_hr_single <- forest_fig_height(S, "single", "hr", fp_row_height, fp_overhead_in)
fh_hr_combo  <- forest_fig_height(S, "combo",  "hr", fp_row_height, fp_overhead_in)
```


# Motivation & Overview

## The Core Question

\begin{block}{Subgroup analysis in regulatory submissions}
When many subgroups are examined in a single trial, small subgroups can
produce hazard ratio estimates that look dramatic \emph{purely by chance}.
\end{block}

\begin{alertblock}{This study asks directly}
Under a DGM with a \textbf{perfectly uniform treatment benefit}
(true HR\,=\,0.70 for every patient), how often do standard Cox model
analyses produce extreme subgroup results?
\end{alertblock}

- Clinical subgroups of a given $N$ are statistically indistinguishable
  from meaningless *random* groupings of the same size
- The \textcolor{randomorange}{\textbf{random}} benchmarks random combinations 
and should fluctuate around overall results with "extremeness" due to variability of small samples
as opposed to true subgroup effects


## Setting: Uniform Treatment Effect

The DGM is built from a real trial (GBSG breast cancer, $N = 686$) using
`generate_aft_dgm_flex()`.

\begin{block}{Key feature --- constant log hazard ratio}
$$\psi^{0}(\mathbf{L}) = \log(\mathrm{HR}_{\mathrm{true}}) \quad \text{for all } \mathbf{L}.$$
\end{block}

- True overall HR = 0.70 (moderate benefit)
- Every subgroup shares *exactly* this HR
- Any apparent subgroup finding is a false positive by construction

\begin{exampleblock}{Spline parameterisation}
\texttt{log.hrs = log(c(0.70, 0.70, 0.70))} --- uniform HR at biomarker
values $0$, knot $k$, and upper anchor $\zeta$.
\end{exampleblock}


# Simulation Workflow

## Pipeline Overview

\small
\begin{center}
\begin{tabular}{clp{7cm}}
\toprule
\textbf{Step} & \textbf{Function} & \textbf{Purpose} \\
\midrule
1 & --- & Convert source data to consistent time scale \\
2 & \texttt{generate\_aft\_dgm\_flex()} & Fit Weibull AFT; build super population \\
3 & \texttt{check\_censoring\_dgm()} & Verify censoring matches source data \\
4 & \texttt{calibrate\_cens\_adjust()} & Find \texttt{cens\_adjust} via \texttt{uniroot} \\
5 & \texttt{simulate\_from\_dgm()} & Draw synthetic trials from calibrated DGM \\
6 & \texttt{coxph()} loop & Fit Cox in 56 subgroups; record HR and UB(HR) \\
7 & \texttt{summarize\_extreme\_sims()} & Compute all summary statistics \\
8 & \texttt{plot\_ub\_forest()} & Visualise ECI distributions \\
\bottomrule
\end{tabular}
\end{center}





## Workflow: DGM Setup & Calibration
```{r flow-diagram-part1-create, eval=FALSE}
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)

g1 <- grViz('
digraph part1 {
  graph [rankdir = TB, fontname = Helvetica, fontsize = 14,
         nodesep = 0.6, ranksep = 0.5, splines = true]
  node  [fontname = Helvetica, fontsize = 12, margin = "0.20,0.12"]
  edge  [fontname = Helvetica, fontsize = 11, color = "#444444"]

  Source [
    label     = "Source Data\\n(GBSG, N = 686)\\nTime → months",
    shape     = cylinder,
    style     = filled,
    fillcolor = "#dce8f5",
    color     = "#2471a3"
  ]

  PartI [
    label     = "Part I — Build Null DGM\\n\\ngenerate_aft_dgm_flex()\\nWeibull AFT fit  |  Censoring model (AIC)\\nSuper population (N = 5,000)",
    shape     = rectangle,
    style     = filled,
    fillcolor = "#eaf4ea",
    color     = "#1e8449"
  ]

  PartII [
    label     = "Part II — Censoring Diagnostics\\n\\ncheck_censoring_dgm()\\nOverall rate  |  Time quantiles  |  KM median",
    shape     = rectangle,
    style     = filled,
    fillcolor = "#fef9e7",
    color     = "#b7950b"
  ]

  CalibOK [
    label     = "Censoring\\nrate OK?",
    shape     = diamond,
    style     = filled,
    fillcolor = "#fdf2f8",
    color     = "#76448a",
    width     = 1.6,
    height    = 1.0
  ]

  Calibrate [
    label     = "calibrate_cens_adjust()\\nuniroot grid search\\nTarget: sim rate ≈ obs rate",
    shape     = rectangle,
    style     = "filled,dashed",
    fillcolor = "#fff3e0",
    color     = "#c0392b"
  ]

  PartIII [
    label     = "Part III — Calibrated DGM Ready\\n\\nsimulate_from_dgm()\\nOne RCT per call  |  Uniform HR = 0.70",
    shape     = rectangle,
    style     = "filled,bold",
    fillcolor = "#eaf4ea",
    color     = "#1e8449"
  ]

  # Layout: Source and PartI side by side, then down
  { rank = same; Source; PartI }
  { rank = same; PartII; Calibrate }

  Source  -> PartI    [label = " fit model"]
  PartI   -> PartII   [label = " pilot sim"]
  PartII  -> CalibOK
  CalibOK -> PartIII  [label = "Yes"]
  CalibOK -> Calibrate [label = "No ", color = "#c0392b", style = dashed]
  Calibrate -> PartII  [label = " retry", color = "#c0392b", style = dashed]
}
')

svg1 <- charToRaw(export_svg(g1))
rsvg_pdf(svg1, "sim_flow_part1.pdf", width = 14, height = 8)
rsvg_png(svg1, "sim_flow_part1.png", width = 2800)
```


```{r flow-diagram-part1, fig.cap="Workflow Part 1: DGM construction and censoring calibration.", out.width="95%", out.height="85%"}
knitr::include_graphics("sim_flow_part1.pdf")
```


## Workflow: Simulation Loop & Results
```{r flow-diagram-part2-create, eval=FALSE}
g2 <- grViz('
digraph part2 {
  graph [rankdir = TB, fontname = Helvetica, fontsize = 14,
         nodesep = 0.6, ranksep = 0.5, splines = true]
  node  [fontname = Helvetica, fontsize = 12, margin = "0.20,0.12"]
  edge  [fontname = Helvetica, fontsize = 11, color = "#444444"]

  PartIII [
    label     = "Calibrated DGM\\n\\nsimulate_from_dgm()\\nN = 686 per trial  |  HR = 0.70",
    shape     = rectangle,
    style     = "filled,bold",
    fillcolor = "#eaf4ea",
    color     = "#1e8449"
  ]

  PartIV [
    label     = "Part IV — Simulation Loop (n = 5,000)\\n\\nFor each trial:\\n• Fit stratified Cox in 56 subgroups\\n• Record HR point estimate and UB(HR)\\n• Re-sample random benchmarks (N ≈ 15, 20, 40, 60)",
    shape     = rectangle,
    style     = filled,
    fillcolor = "#eaf4ea",
    color     = "#1e8449"
  ]

  Results [
    label     = "Results — Empirical CI Distributions\\n\\nMedian HR + 1st–99th percentile ECI\\nPr(UB ≥ 1.0)  |  Med UB | UB ≥ 1\\n\\nECI width driven entirely by subgroup N",
    shape     = rectangle,
    style     = "filled,rounded",
    fillcolor = "#e8f8f5",
    color     = "#148f77"
  ]

  Plots [
    label     = "Visualization — gg_forest()\\n\\nPanel A: single-variable subgroups\\nPanel B: combinations + random benchmarks\\nHR distribution plot  |  UB(HR) distribution plot",
    shape     = rectangle,
    style     = filled,
    fillcolor = "#dce8f5",
    color     = "#2471a3"
  ]

  { rank = same; PartIII; PartIV }
  { rank = same; Results; Plots }

  PartIII -> PartIV  [label = " repeat × 5,000"]
  PartIV  -> Results [label = " summarize_extreme_sims()"]
  Results -> Plots   [label = " plot_ub_forest()"]

  # Loop arrow
  PartIV -> PartIII  [label = "next trial ",
                       color = "#1e8449", style = dashed,
                       constraint = false]
}
')

svg2 <- charToRaw(export_svg(g2))
rsvg_pdf(svg2, "sim_flow_part2.pdf", width = 14, height = 8)
rsvg_png(svg2, "sim_flow_part2.png", width = 2800)
```

```{r flow-diagram-part2, fig.cap="Workflow Part 2: Simulation loop and results.", out.width="95%", out.height="85%"}
knitr::include_graphics("sim_flow_part2.pdf")
```


## Subgroup Definitions (1/3): ITT, Clinical, Continuous

```{r subgroup-table-1}
library(gt)

data.frame(
  Num  = 1:18,
  Category = c("ITT", rep("Clinical",4), rep("Continuous",13)),
  Subgroup = c(
    "All Patients",
    "Post-menopausal", "Pre-menopausal", "Grade 3", "Grade 1/2",
    "Age (young)", "Age (older)", "Age ≤ 50", "Age > 50",
    "Tumour size (small)", "Tumour size (large)",
    "Node-negative", "Nodes 1–3", "High nodes (>Q3)",
    "PGR low", "PGR high", "ER-low (<20)", "ER-high (≥20)"
  ),
  Definition = c(
    "All subjects (flag_itt == 1)",
    "meno == 1", "meno == 0", "grade == 3", "grade != 3",
    "age ≤ median", "age > median", "age ≤ 50", "age > 50",
    "size ≤ median", "size > median",
    "nodes == 0", "0 < nodes ≤ 3", "nodes > Q3",
    "pgr ≤ median", "pgr > median", "er ≤ 20", "er > 20"
  ),
  stringsAsFactors = FALSE
) |>
  gt(groupname_col = "Category") |>
  cols_label(Num = "#") |>
  cols_width(Num ~ px(30), Subgroup ~ px(160), Definition ~ px(160)) |>
  tab_options(
    table.font.size   = px(11),
    row_group.font.weight = "bold",
    column_labels.font.weight = "bold",
    data_row.padding   = px(2),
    table.width        = pct(100)
  )
```


## Subgroup Definitions (2/3): Two-Way Interactions 
\small
```{r subgroup-table-2}
data.frame(
  Num  = 19:52,
  Category = c(
    rep("Meno × Grade", 4), rep("Meno × Age", 4),
    rep("Meno × ER", 4), rep("Grade × Nodes", 5),
    rep("Grade × ER", 4), rep("Grade × PGR", 4),
    rep("3-way", 5), rep("Size × Nodes", 4)
  ),
  Subgroup = c(
    "Pre-meno / Grade 3", "Post-meno / Grade 3",
    "Pre-meno / Grade 1–2", "Post-meno / Grade 1–2",
    "Pre-meno / Age ≤ 50", "Pre-meno / Age > 50",
    "Post-meno / Age (young)", "Post-meno / Age (older)",
    "Pre-meno / ER-low", "Pre-meno / ER-high",
    "Post-meno / ER-low", "Post-meno / ER-high",
    "Grade 3 / Node-neg", "Grade 3 / Node-pos",
    "Grade 1–2 / Node-neg", "Grade 1–2 / Node-pos",
    "Grade 3 / High nodes",
    "Grade 3 / ER-low", "Grade 3 / ER-high",
    "Grade 1–2 / ER-low", "Grade 1–2 / ER-high",
    "Grade 3 / PGR low", "Grade 3 / PGR high",
    "Grade 1–2 / PGR low", "Grade 1–2 / PGR high",
    "Pre-meno / Yng / ER-low", "Pre-meno / Yng / ER-high",
    "Post-meno / G3 / ER-low", "Post-meno / G3 / Node+",
    "G3 / Node+ / ER-low",
    "Large / Node-pos", "Large / Node-neg",
    "Small / Node-pos", "Small / Node-neg"
  ),
  Definition = c(
    "meno==0 & grade==3", "meno==1 & grade==3",
    "meno==0 & grade!=3", "meno==1 & grade!=3",
    "meno==0 & age≤50", "meno==0 & age>50",
    "meno==1 & age≤med", "meno==1 & age>med",
    "meno==0 & er≤20", "meno==0 & er>20",
    "meno==1 & er≤20", "meno==1 & er>20",
    "grade==3 & nodes==0", "grade==3 & nodes>0",
    "grade!=3 & nodes==0", "grade!=3 & nodes>0",
    "grade==3 & nodes>Q3",
    "grade==3 & er≤20", "grade==3 & er>20",
    "grade!=3 & er≤20", "grade!=3 & er>20",
    "grade==3 & pgr≤med", "grade==3 & pgr>med",
    "grade!=3 & pgr≤med", "grade!=3 & pgr>med",
    "meno==0 & age≤50 & er≤20", "meno==0 & age≤50 & er>20",
    "meno==1 & grade==3 & er≤20", "meno==1 & grade==3 & nodes>0",
    "grade==3 & nodes>0 & er≤20",
    "size>med & nodes>0", "size>med & nodes==0",
    "size≤med & nodes>0", "size≤med & nodes==0"
  ),
  stringsAsFactors = FALSE
) |>
  gt(groupname_col = "Category") |>
  cols_label(Num = "#") |>
  cols_width(Num ~ px(30), Subgroup ~ px(180), Definition ~ px(200)) |>
  tab_options(
    table.font.size   = px(9),
    row_group.font.size = px(10),
    row_group.font.weight = "bold",
    column_labels.font.weight = "bold",
    data_row.padding   = px(1),
    table.width        = pct(100)
  )
```


## Subgroup Definitions (3/3): Random Benchmarks

```{r subgroup-table-3}
data.frame(
  Num  = 53:56,
  Subgroup = c("random60", "random40", "random20", "random15"),
  N = c("~60", "~40", "~20", "~15"),
  Construction = c(
    "Draw 60 random indices from trial",
    "First 40 of the random60 indices",
    "First 20 of the random60 indices",
    "First 15 of the random60 indices"
  ),
  stringsAsFactors = FALSE
) |>
  gt() |>
  cols_label(Num = "#", N = "Approx N") |>
  tab_options(
    table.font.size   = px(12),
    column_labels.font.weight = "bold",
    data_row.padding   = px(4),
    table.width        = pct(90)
  ) |>
  tab_source_note(
    source_note = paste(
      "Re-sampled every simulation.",
      "Nested: random15 ⊂ random20 ⊂ random40 ⊂ random60.",
      "No clinical meaning — true HR = 0.70 by construction."
    )
  )
```





# Part I --- Building the Null DGM

## Fitting `generate_aft_dgm_flex()`

- Fits a **Weibull AFT model** to source data (GBSG)
- Generates a large *super population* ($N_{\text{super}} = 5000$) from which
  all future trials are drawn
- `model = "null"` --- treatment effect identical for every patient
- Censoring model estimated by AIC selection (Weibull / log-normal)

\begin{exampleblock}{R code}
\small\ttfamily
dgm\_null <- generate\_aft\_dgm\_flex(\\
\quad data = gbsg, model = "null",\\
\quad subgroup\_vars = NULL, ~~ n\_super = 5000)
\end{exampleblock}


<!-- ## Time-Scale Consistency -->
<!-- \begin{alertblock}{The single most common simulation failure} -->
<!-- Building the DGM on \emph{days} (\texttt{rfstime}) then passing -->
<!-- \texttt{analysis\_time} in \emph{months} $\Rightarrow$ universal -->
<!-- administrative censoring. -->
<!-- \end{alertblock} -->

<!-- \begin{block}{Quick diagnostic} -->
<!-- Compare $\exp(\mu)$ to the observed median event time: -->

<!-- \medskip -->
<!-- \centering -->
<!-- \begin{tabular}{ll} -->
<!-- \toprule -->
<!-- Quantity & Value (months) \\ -->
<!-- \midrule -->
<!-- Observed median event time & $\sim 33$ \\ -->
<!-- $\exp(\hat\mu)$ from DGM & should be similar \\ -->
<!-- \bottomrule -->
<!-- \end{tabular} -->
<!-- \end{block} -->

<!-- **Rule:** convert `rfstime` to months *before* fitting. -->


# Part II --- Censoring Diagnostics & Calibration

## Censoring Consistency

- Credible results require censoring in synthetic trials to match source data
- `check_censoring_dgm()` compares three quantities:
  1. **Overall censoring rate** vs `rate_tol` (default 10 pp)
  2. **Censoring-time quantiles** among censored subjects
  3. **KM median censoring time** vs `median_tol` (default 25%)


## Calibrating `cens_adjust`

`calibrate_cens_adjust()` uses `uniroot` to find the log-scale shift
that equates a target statistic between reference and simulated data.

\begin{block}{Objective function}
At each candidate \texttt{cens\_adjust}:
(1) generate synthetic subjects,
(2) extract target metric,
(3) return sim\_metric $-$ ref\_metric.
\end{block}

\begin{center}
\begin{tabular}{ll}
\toprule
Target & Equates \\
\midrule
\texttt{"rate"} & Overall censoring rate \\
\texttt{"km\_median"} & KM-based median censoring time \\
\bottomrule
\end{tabular}
\end{center}

**Monotonicity:** larger `cens_adjust` $\Rightarrow$ longer censoring times.


# Part III --- Generating Synthetic Trials

## `simulate_from_dgm()`

With the null DGM calibrated, each simulated trial is a single function call:

1. Samples from super population with replacement
2. Assigns treatment ($1{:}1$ randomisation)
3. Generates latent survival times from the AFT model
4. Applies calibrated censoring model
5. Truncates at `analysis_time`

Under the null DGM: `flag_harm` $\equiv$ 0 for every subject.


# Part IV --- Simulation Results

## Subgroups Under Study (56 total)

\footnotesize
\begin{columns}[T]
\begin{column}{0.48\textwidth}
\begin{tabular}{lp{3.5cm}}
\toprule
\textbf{Group} & \textbf{Examples} \\
\midrule
ITT & All Patients ($N\!\approx\!686$) \\
Clinical & Post-/Pre-menopausal, Grade \\
Continuous & Age, Size, Nodes, PGR, ER \\
Interaction & Meno $\times$ Grade, ER, Age; Grade $\times$ Nodes, ER, PGR \\
\bottomrule
\end{tabular}
\end{column}
\begin{column}{0.48\textwidth}
\begin{tabular}{lp{3.5cm}}
\toprule
\textbf{Group} & \textbf{Examples} \\
\midrule
3-way & Pre-meno/Young/ER-low, \ldots \\
Size $\times$ Nodes & Large/Node-pos, \ldots \\
\textcolor{randomorange}{\textbf{Random}} & \textcolor{randomorange}{random60, 40, 20, 15} \\
\bottomrule
\end{tabular}

\medskip

\begin{alertblock}{\footnotesize Key benchmark}
\textcolor{randomorange}{\texttt{random*}} subgroups have
\textbf{no clinical meaning}. Re-sampled each simulation.
\end{alertblock}
\end{column}
\end{columns}


## Summary Statistics

:::::: {.columns}
::: {.column width="48%"}
\begin{block}{HR point estimate distribution}
\begin{itemize}\small
\item Median + 1st--99th percentile (ECI)
\item Under truth, median $\approx 0.70$ for every subgroup
\item ECI width driven by subgroup $N$
\end{itemize}
\end{block}
:::

::: {.column width="48%"}
\begin{block}{Upper confidence bound UB(HR)}
\begin{itemize}\small
\item Upper end of the Cox 95\% CI
\item $\Pr(\text{UB} < 1.0)$: power to show benefit
\item Med UB $|$ UB$\ge$1: worst-case for the safety reviewer
\end{itemize}
\end{block}
:::
::::::


## Simulation: `r n_sims_null` trials completed

```{r sim-summary-text}
#| results: asis
cat(sprintf("- **%d** simulated trials from the null DGM (true HR = %.2f)\n",
            S$n_sims, hr_true))
cat(sprintf("- **%d** subgroups evaluated per trial\n", length(subgroups)))
cat(sprintf("- Subgroup sizes range from $N \\approx 15$ (random15) to $N \\approx %d$ (ITT)\n",
            max(S$mean_n, na.rm = TRUE)))
```

\begin{alertblock}{Re-use with new results}
\footnotesize
\texttt{res <- load\_sim\_results("new\_file.rds")}\\
\texttt{S <- summarize\_extreme\_sims(res\$sim\_hrs, res\$sim\_ubs, res\$sim\_ns, res\$subgroups)}\\
All downstream tables and plots update automatically.
\end{alertblock}


# Results --- Forest Plots

## HR Distributions: Single-Variable Subgroups

```{r hr-forest-single}
#| fig.width: 10
#| fig.height: !expr fh_hr_single
#| fig.cap: "HR distribution — single-variable subgroups. True HR = 0.70 (red dashed)."
#| out.height: "85%"
plot_hr_forest(S, panel = "single",
               base_size = fp_base_size, point_size = fp_point_size,
               line_size = fp_line_size)
```


## HR Distributions: Combinations + Random Benchmarks

```{r hr-forest-combo}
#| fig.width: 10
#| fig.height: !expr fh_hr_combo
#| fig.cap: "HR distribution — combination and random-benchmark subgroups."
#| out.height: "85%"
plot_hr_forest(S, panel = "combo",
               base_size = fp_base_size, point_size = fp_point_size,
               line_size = fp_line_size)
```


## UB(HR) Distributions: Single-Variable Subgroups {.shrink=5}

```{r ub-forest-single}
#| fig.width: 10
#| fig.height: !expr fh_ub_single
#| fig.cap: "UB(HR) — single-variable subgroups. Dotted: UB=1.0; Dashed: true HR=0.70."
#| out.height: "82%"
plot_ub_forest(S, panel = "single",
               base_size = fp_base_size, point_size = fp_point_size,
               line_size = fp_line_size)
```


## UB(HR) Distributions: Combinations + Random Benchmarks {.shrink=5}

```{r ub-forest-combo}
#| fig.width: 10
#| fig.height: !expr fh_ub_combo
#| fig.cap: "UB(HR) — combination and random-benchmark subgroups."
#| out.height: "82%"
plot_ub_forest(S, panel = "combo",
               base_size = fp_base_size, point_size = fp_point_size,
               line_size = fp_line_size)
```


# Key Findings

## HR Point Estimates

\begin{block}{Large subgroups ($N > 200$)}
Median HR tracks truth ($\approx 0.70$) closely; ECI is narrow.
\end{block}

\begin{alertblock}{Small subgroups ($N \lesssim 40$)}
\begin{itemize}
\item ECI for \texttt{random15}/\texttt{random20} spans $\sim\!0.25$ to $>1.0$
\item Clinical subgroups of similar size show \textbf{indistinguishable} ECI widths
\item Clinical rationale provides \textbf{no statistical protection}
\end{itemize}
\end{alertblock}

\begin{center}
\fcolorbox{UBCblue}{UBCblue!5}{\parbox{0.85\textwidth}{\centering
\textbf{Clinical subgroups of a given $N$ are statistically
indistinguishable from meaningless random groupings of the same size.}
}}
\end{center}


## Upper Bounds UB(HR)

:::::: {.columns}
::: {.column width="55%"}
\begin{block}{The upper-bound distribution}
\begin{itemize}\small
\item \texttt{random40} ($N\!\approx\!40$): 1\% of trials produce UB(HR)\,$>$\,3.0--4.0
\item \texttt{random15} ($N\!\approx\!15$): values above 5.0 not rare
\item A reviewer seeing UB(HR)\,=\,3.5 in 15 patients cannot distinguish
      chance from a genuine safety signal
\end{itemize}
\end{block}
:::

::: {.column width="42%"}
\begin{exampleblock}{\footnotesize Illustrative 99th pctl of UB(HR)}
\centering\footnotesize
\begin{tabular}{lr}
\toprule
Subgroup & P99 UB \\
\midrule
ITT ($N\!\approx\!686$) & $<1.1$ \\
$N\!\approx\!200$ & $\sim\!1.5$ \\
$N\!\approx\!60$ & $\sim\!2.5$ \\
random40 & $\sim\!3.5$ \\
random20 & $\sim\!5.0$ \\
random15 & $>5.0$ \\
\bottomrule
\end{tabular}
\end{exampleblock}
:::
::::::


## Cox Model Works Correctly

\begin{block}{What the simulation confirms}
\begin{enumerate}
\item Cox model is \textbf{unbiased} --- median HR $\approx$ 0.70 for every subgroup
\item Confidence intervals have correct \textbf{coverage}
\item Variability driven \textbf{entirely by sample size}
\end{enumerate}
\end{block}

\begin{alertblock}{The interpretive hazard}
\begin{itemize}
\item UB(HR)\,=\,3.5 in a 15-patient subgroup is \emph{expected} under uniform benefit
\item Without simulation-based calibration, no way to quantify how ``extreme'' a result is
\item The \texttt{random*} benchmarks provide exactly this calibration
\end{itemize}
\end{alertblock}


# Conclusions

## Conclusions

1. **Small subgroups produce extreme results by chance.**
   Under uniform HR = 0.70, subgroups with $N \lesssim 40$ regularly produce
   UB(HR) $> 3.0$.

2. **Clinical rationale does not protect against noise.**
   Clinical subgroups have the *same* HR/UB distribution as
   random benchmarks of that size.

3. **The Cox model is working correctly.**
   Wide CIs in small subgroups are accurate --- they reflect genuine uncertainty.

4. **Simulation-based calibration is essential.**
   Without a null-DGM benchmark, reviewers cannot quantify "extreme" for a
   given subgroup size.

5. **The `forestsearch` package** provides a complete, reproducible workflow.


## {.plain}

\vfill
\begin{center}
{\Large\bfseries\color{UBCblue} Thank You}

\bigskip

{\large \texttt{forestsearch} R package}

\medskip

\url{https://larry-leon.github.io/forestsearch/}

\medskip

\texttt{pak::pak("larry-leon/forestsearch")}
\end{center}
\vfill


# Appendix {.unnumbered}

## New-Results Workflow

\footnotesize

```{r show-workflow}
#| echo: true
#| eval: false

source("summarize_extreme_sims.R")

# ── Option A: Load existing results ─────────────────────────────
res <- load_sim_results("results/extreme_sims_5000.rds")

# ── Option B: After running a new simulation loop ────────────────
save_sim_results(sim_hrs, sim_ubs, sim_ns, subgroups,
                 file = "results/extreme_sims_new.rds",
                 hr_true = 0.70,
                 metadata = list(dgm_seed = 99, n_per_trial = 686))

# ── Summarize (works with either option) ─────────────────────────
S <- summarize_extreme_sims(res$sim_hrs, res$sim_ubs,
                             res$sim_ns, res$subgroups)

# ── All plots driven by S ────────────────────────────────────────
plot_hr_forest(S, panel = "single")
plot_hr_forest(S, panel = "combo")
plot_ub_forest(S, panel = "single")   # <-- ub-forest-single
plot_ub_forest(S, panel = "combo")    # <-- ub-forest-combo
```


## `summarize_extreme_sims()` API

\footnotesize

\begin{tabular}{lp{8.5cm}}
\toprule
\textbf{Field} & \textbf{Description} \\
\midrule
\texttt{hr\_q}, \texttt{ub\_q} & Quantile matrices [\texttt{n\_sg} $\times$ 3]: 1st, median, 99th \\
\texttt{mean\_n} & Mean subgroup $N$ across simulations \\
\texttt{pr\_ub\_lt1}, \texttt{pr\_ub\_ge1} & Power / failure rate per subgroup \\
\texttt{cond\_ub} & Conditional UB quantiles given UB $\ge$ 1 \\
\texttt{results\_tbl} & Formatted \texttt{data.frame} for printing \\
\texttt{ok}, \texttt{ok\_ub} & Logical: valid subgroups \\
\texttt{idx\_hr\_single}, \texttt{idx\_ub\_combo}, \ldots & Index vectors for forest panels \\
\texttt{cat\_ok}, \texttt{cat\_cols} & Category labels and colour map \\
\texttt{lab\_cmed\_ub} & Pre-formatted conditional median labels \\
\bottomrule
\end{tabular}


## Comparing Two Simulation Runs

\footnotesize

```{r show-compare}
#| echo: true
#| eval: false

S_old <- summarize_extreme_sims(...)   # e.g. n_sims = 1000
S_new <- summarize_extreme_sims(...)   # e.g. n_sims = 5000

comp <- compare_extreme_sims(S_old, S_new,
                              metric = "both",
                              label1 = "1000 sims",
                              label2 = "5000 sims")
print(comp, row.names = FALSE)
# Subgroup | N | MedHR | Delta_MedHR | PrUBge1 | Delta_PrUBge1 | MedUBcond
```

Useful for verifying convergence, comparing DGM variants, or assessing
sensitivity to censoring calibration.


## Helper Utilities

\footnotesize

```{r show-helpers}
#| echo: true
#| eval: false

# ── Optimal figure height ────────────────────────────────────────
fh <- forest_fig_height(S, panel = "combo", type = "ub",
                         row_height = 0.45, overhead = 1.5)

# ── Save / load round-trip ───────────────────────────────────────
save_sim_results(sim_hrs, sim_ubs, sim_ns, subgroups,
                 file = "results/my_run.rds", hr_true = 0.70,
                 metadata = list(cens_adjust = 0.42))

res <- load_sim_results("results/my_run.rds")
# prints: Loaded: ... (5000 sims x 56 subgroups, saved ...)

# ── Print method ──────────────────────────────────────────────────
print(S)   # shows n_sims, n_sg, hr_true, then results_tbl
```
