---
title: "GLM Extension: QC Simulation and CRAN Submission Workflow"
subtitle: "forestsearch v0.2.0 — feature/glm-extension branch"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
    theme: cosmo
execute:
  eval: false
---

# Overview

This document covers the complete path from QC validation through
CRAN submission for the forestsearch GLM extension.

**Current state:**

- Branch: `feature/glm-extension`
- CRAN published version: v0.1.0 (survival only)
- Target version: v0.2.0 (adds GLM binary/continuous outcomes)
- Local `devtools::check(args = "--as-cran")`: 0/0/0


# Step 1: Install the Branch Version {#sec-install}

`multisession` parallelization spawns fresh R processes that only see
**installed** packages.  `devtools::load_all()` loads into the current
session only — workers cannot find those functions.

## 1.1: Install from the branch

Run once from the package root (`~/GitHub/forestsearch/`):

```{r}
# Install the current branch into your R library
# This makes all exports (including GLM functions) visible to
# multisession workers
devtools::install(upgrade = "never", quick = TRUE)
```

After this, `library(forestsearch)` in any R session (including
parallel workers) will load the GLM extension code.

## 1.2: Verify installation

```{r}
# Confirm the installed version has GLM exports
library(forestsearch)
stopifnot(is.function(make_effect_estimator))
stopifnot(is.function(estimate_propensity_scores))
stopifnot(is.function(plot_sg_glm_outcomes))
cat("forestsearch installed with GLM exports\n")
cat("Version:", as.character(packageVersion("forestsearch")), "\n")
```

## 1.3: Key distinction

| Command | Scope | Parallel workers see it? |
|---|---|---|
| `devtools::load_all()` | Current session only | **No** |
| `devtools::install()` | System R library | **Yes** |
| `source("dev/glm/glm_effect_estimators.R")` | Current session only | **No** (unless `source()` inside worker) |

For QC simulations with `multisession`, you **must** use
`devtools::install()`.  The `source()` workaround used in dev
Quarto documents is only appropriate for non-parallelized or
sequential testing.


# Step 2: QC Simulation — Cox vs Poisson Comparison {#sec-qc}

The GLM extension approximates survival analysis using Poisson
regression with a log(time) offset.  The QC simulation validates this
by comparing:

- **Cox PH model** (survival endpoint, existing pipeline)
- **Poisson GLM with IRR** (binary + offset, GLM pipeline)

on the same GBSG data, using the full ForestSearch pipeline with
bootstrap bias correction.

## 2.1: Setup

```{r}
library(forestsearch)
library(survival)
library(doFuture)
library(foreach)

n_cores <- max(1L, floor(parallel::detectCores() * 0.80))
cat("Using", n_cores, "cores\n")
```

## 2.2: Data preparation

```{r}
df <- survival::gbsg
df$id <- seq_len(nrow(df))
df$grade3 <- as.integer(df$grade == "3")
confs <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")

cat("GBSG: N =", nrow(df), "\n")
```

## 2.3: Run A — Survival (Cox PH)

```{r}
cat("=== Run A: Survival (Cox PH) ===\n")
fs_cox <- forestsearch(
  df.analysis = df, confounders.name = confs, conf_force = c("er <= 0"),
  outcome.name = "rfstime", event.name = "status",
  treat.name = "hormon", id.name = "id",
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE),
  maxk = 2, sg_focus = "hr"
)
cat("Cox SG:", paste(fs_cox$sg.harm, collapse = " & "), "\n")
```

```{r}
# Bootstrap with multisession (requires devtools::install)
fs_bc_cox <- forestsearch_bootstrap_dofuture(
  fs.est = fs_cox, nb_boots = 500,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE),
  details = TRUE
)
```

```{r}
# Summarize
summ_cox <- summarize_bootstrap_results(
  sgharm = fs_cox$sg.harm,
  boot_results = fs_bc_cox,
  create_plots = TRUE, est.scale = "hr"
)
summ_cox$table
```

## 2.4: Run B — Poisson GLM (IRR with offset)

```{r}
cat("=== Run B: Poisson GLM (IRR + offset) ===\n")
fs_irr <- forestsearch(
  df.analysis = df, confounders.name = confs, conf_force = c("er <= 0"),
  outcome.name = "status", event.name = "status",
  treat.name = "hormon", id.name = "id",
  outcome_type = "binary", effect_measure = "IRR", offset.name = "rfstime",
  dmin.grf = 0.0,
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE),
  maxk = 2, sg_focus = "hr"
)
cat("IRR SG:", paste(fs_irr$sg.harm, collapse = " & "), "\n")
```

```{r}
# Bootstrap with multisession
fs_bc_irr <- forestsearch_bootstrap_dofuture(
  fs.est = fs_irr, nb_boots = 500,
  parallel_args = list(plan = "multisession", workers = n_cores,
                       show_message = TRUE),
  details = TRUE
)
```

```{r}
# Summarize
summ_irr <- summarize_bootstrap_results(
  sgharm = fs_irr$sg.harm,
  boot_results = fs_bc_irr,
  create_plots = TRUE, est.scale = "hr"
)
summ_irr$table
```

## 2.5: Compare Results

```{r}
cat("\n=== QC Comparison: Cox vs Poisson IRR ===\n\n")
cat("Cox PH subgroup:     ", paste(fs_cox$sg.harm, collapse = " & "), "\n")
cat("Poisson IRR subgroup:", paste(fs_irr$sg.harm, collapse = " & "), "\n\n")

cat("Cox bootstrap table:\n")
print(fs_bc_cox$FSsg_tab)
cat("\nIRR bootstrap table:\n")
print(fs_bc_irr$FSsg_tab)
```

## 2.6: Visualizations

```{r}
# Cox: KM curves
km_cox <- plot_sg_weighted_km(
  fs.est = fs_cox, fs_bc = fs_bc_cox,
  outcome.name = "rfstime", event.name = "status", treat.name = "hormon",
  show.logrank = FALSE, conf.int = TRUE
)

# IRR: bar chart
bar_irr <- plot_sg_glm_outcomes(
  fs.est = fs_irr, fs_bc = fs_bc_irr,
  outcome.name = "status", treat.name = "hormon",
  E.name = "Hormon", C.name = "Control"
)
print(bar_irr)

# IRR: forest plot
fp_irr <- plot_subgroup_results_forestplot(
  fs_results = list(fs.est = fs_irr, fs_bc = fs_bc_irr),
  df_analysis = df,
  outcome.name = "status", event.name = "status", treat.name = "hormon",
  E.name = "Hormon", C.name = "Control",
  title_text = "Poisson IRR Forest Plot (GBSG)"
)
plot(fp_irr)
```


# Step 3: Version Bump {#sec-version}

Edit `DESCRIPTION` to update the version:

```
Version: 0.2.0
```

Run `devtools::document()` after editing.


# Step 4: Update NEWS.md {#sec-news}

Add to the top of `NEWS.md`:

```
# forestsearch 0.2.0

## New Features

* GLM extension: subgroup identification for binary and continuous
  outcomes using odds ratios (OR), risk differences (RD), relative
  risks (RR), incidence rate ratios (IRR), incidence rate differences
  (IRD), and mean differences (MD).

* Propensity score adjustment for observational studies via stabilized
  IPTW (inverse probability of treatment weighting) with GRF, LASSO,
  or logistic regression PS estimation.

* New exported functions:
  - `make_effect_estimator()`: factory for GLM estimator closures
  - `estimate_propensity_scores()`: PS estimation with trimming
  - `plot_sg_glm_outcomes()`: bar charts (binary) and point-and-errorbar
    (continuous) outcome visualization

* Forest plot (`plot_subgroup_results_forestplot()`) extended to support
  GLM effect measures with auto-detected display defaults, clinically
  meaningful axis limits, and extreme CI capping.

* Cross-validation (`forestsearch_Kfold`, `forestsearch_tenfold`)
  extended to handle GLM parameters and propensity score re-estimation.

## Bug Fixes

* Fixed `merge(by = "id")` ignoring `id.name` parameter in
  `forestsearch()` output.

* Fixed `ps_adjust_method` missing `"none"` in `match.arg()` choices,
  which caused silent bootstrap failures for survival/RCT analyses.

* Removed `stringr` dependency (replaced with base R in cross-validation).

* Fixed non-ASCII characters in R source files.

* Added `sandwich` to Suggests for conditional robust SE.

## Internal

* New internal helpers for GLM effect estimation, Poisson rate models,
  and propensity score computation.

* `grf.subg.harm.glm()` for GRF screening with binary/continuous outcomes.

* Bootstrap bias correction extended for GLM outcomes with
  log-scale/identity-scale handling.
```


# Step 5: Update cran-comments.md {#sec-cran-comments}

Replace contents of `cran-comments.md`:

```
## Test environments

* macOS aarch64 (local), R 4.5.2
* Windows Server 2022, R-devel (win-builder)
* Windows Server 2022, R-release (win-builder)

## R CMD check results

0 errors | 0 warnings | 0 notes

## Changes since v0.1.0

This release adds a GLM extension for subgroup identification with
binary and continuous outcomes (OR, RD, RR, IRR, IRD, MD), propensity
score adjustment via stabilized IPTW, and updated visualization
functions.  All existing survival analysis functionality is unchanged.
See NEWS.md for details.

## Downstream dependencies

There are currently no downstream dependencies for this package.
```


# Step 6: Final Check on the Branch {#sec-check}

```{r}
devtools::document()
devtools::check(args = "--as-cran")
```

Confirm 0 errors, 0 warnings, 0 notes.


# Step 7: Win-Builder Testing {#sec-winbuilder}

Submit to win-builder from the branch (before merging to master):

```{r}
devtools::check_win_devel()    # R-devel
devtools::check_win_release()  # R-release
```

Results are emailed to `larry.leon.05@post.harvard.edu`.
Review for any Windows-specific issues (path separators, encoding,
missing system libraries).


# Step 8: Merge to Master {#sec-merge}

After win-builder passes, merge via GitHub Desktop:

1. Switch to `master` branch
2. Merge `feature/glm-extension` into `master`
3. Push to GitHub

Or from the terminal:

```bash
git checkout master
git merge feature/glm-extension
git push origin master
```


# Step 9: Build and Submit to CRAN {#sec-submit}

```{r}
# Build the tarball
devtools::build()
```

Submit the `.tar.gz` file at:
`https://cran.r-project.org/submit.html`

Confirm the email from CRAN to complete the submission.


# Session Information

```{r}
#| eval: true
sessionInfo()
```
