---
title: "Exploratory Subgroup Identification with ForestSearch: GLM Extension"
subtitle: "GBSG Analysis — Poisson-Offset (Incidence Rate Ratio)"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false

# Navigate to package root — find it first, then set as knitr root
# (setwd alone resets between chunks; root.dir persists)
pkg_root <- getwd()
while (!file.exists(file.path(pkg_root, "DESCRIPTION")) &&
       pkg_root != dirname(pkg_root)) {
  pkg_root <- dirname(pkg_root)
}
stopifnot("Could not find package root" = file.exists(file.path(pkg_root, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = pkg_root)

knitr::opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.retina = 2,
  fig.width = 8,
  fig.height = 6
)

timings <- list()
t_vignette_start <- proc.time()
```

# Introduction

This vignette demonstrates the **ForestSearch GLM extension** using the
Poisson-offset approach as a built-in sanity check against the Cox
proportional hazards model.  Under the Poisson-offset approximation:

$$
\log\bigl(E[\text{events}]\bigr) = \log(\text{follow-up time}) + \beta_0 + \beta_1 \cdot \text{Treat}
$$

the incidence rate ratio (IRR) $\exp(\hat\beta_1)$ approximates the
Cox hazard ratio when the baseline hazard is locally constant.  This
allows direct comparison of GLM-based subgroup identification against the
established survival methodology from León et al. (2024).

## GLM Effect Measure: IRR

| Property | Cox (reference) | Poisson-offset (GLM) |
|---|---|---|
| Estimand | Hazard ratio | Incidence rate ratio |
| Scale | $\exp(\hat\beta)$ | $\exp(\hat\beta)$ |
| Outcome | Time-to-event + censoring | Binary event indicator |
| Offset | None | $\log(\text{follow-up time})$ |
| `outcome_type` | `"survival"` (default) | `"binary"` |
| `effect_measure` | Not specified | `"IRR"` |


# Setup

## Load Required Packages

```{r load-packages}
devtools::document(quiet = TRUE)
devtools::load_all(".", quiet = TRUE)

library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(policytree)
library(doFuture)
library(patchwork)

# GLM estimator closures (not yet graduated to R/)
source("dev/glm/glm_effect_estimators.R")

theme_set(theme_minimal(base_size = 12))
```


# Data: German Breast Cancer Study Group Trial

## Study Background

The GBSG trial evaluated hormonal treatment (tamoxifen) versus chemotherapy
in node-positive breast cancer patients.  Key characteristics:

- **Sample size**: N = 686
- **Outcome**: Recurrence-free survival time
- **Censoring rate**: ~56%
- **Treatment**: Hormonal therapy (tamoxifen) vs. chemotherapy

## Data Preparation

```{r data-setup}
#| code-fold: false

df.analysis <- survival::gbsg

df.analysis <- within(df.analysis, {
  id          <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3      <- ifelse(grade == "3", 1, 0)
  treat       <- hormon
})

confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
id.name          <- "id"
treat.name       <- "hormon"

# Survival outcome (for Cox reference)
outcome.name.surv <- "time_months"
event.name        <- "status"

# GLM Poisson-offset outcome
outcome.name.glm  <- "status"        # binary event indicator
offset.name       <- "time_months"   # follow-up time as offset

cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```

## ITT Effect Comparison

```{r itt-comparison}
# Cox HR
cox_itt <- coxph(Surv(time_months, status) ~ hormon, data = df.analysis)

# Poisson IRR
pois_itt <- glm(status ~ hormon, family = poisson(link = "log"),
                offset = log(time_months), data = df.analysis)

cat("ITT Treatment Effect:\n")
cat("  Cox  log-HR: ", round(coef(cox_itt)["hormon"], 4),
    " (HR =", round(exp(coef(cox_itt)["hormon"]), 3), ")\n")
cat("  Pois log-IRR:", round(coef(pois_itt)["hormon"], 4),
    " (IRR =", round(exp(coef(pois_itt)["hormon"]), 3), ")\n")
cat("  Difference:  ", round(coef(pois_itt)["hormon"] - coef(cox_itt)["hormon"], 4), "\n")
```


# Parallel Processing Configuration

```{r parallel-setup}
n_cores <- max(1L, parallel::detectCores(logical = FALSE) - 1L)
n_cores_total <- parallel::detectCores()
cat("Using", n_cores, "of", n_cores_total, "total cores\n")

# Sequential for load_all() compatibility; switch to multisession
# after devtools::install()
par_args <- list(
  plan         = "sequential",
  workers      = n_cores,
  show_message = TRUE
)
```


# Preliminary Analysis: GRF Candidate Selection {#sec-grf}

## GRF Sign Convention for Adverse Outcomes

`causal_forest` estimates `E[Y(1) - Y(0) | X]` where higher Y is
implicitly better.  Since our binary outcome is adverse (event = bad),
the `adverse_outcome = TRUE` flag (default for binary outcomes) flips
the outcome to `1 - Y` internally so that the harm-detection sign
convention aligns with `causal_survival_forest`.

## GRF with Multiple `dmin.grf` Thresholds

The `dmin.grf` parameter gates on the minimum doubly-robust score
difference within a policy tree leaf node.  For binary outcomes, this
is on the probability scale.  We explore how different thresholds affect
GRF subgroup identification.

### Cox Reference (RMST-based)

```{r grf-cox}
t0 <- proc.time()

grf_cox <- grf.subg.harm.survival(
  data = df.analysis,
  confounders.name = confounders.name,
  outcome.name = outcome.name.surv,
  event.name = event.name,
  id.name = id.name,
  treat.name = treat.name,
  maxdepth = 2,
  n.min = 60,
  dmin.grf = 12,
  frac.tau = 0.6,
  details = TRUE,
  return_selected_cuts_only = FALSE
)

timings$grf_cox <- (proc.time() - t0)["elapsed"]

cat("\nCox GRF cuts:", paste(grf_cox$tree.cuts, collapse = ", "), "\n")
if (!is.null(grf_cox$sg.harm.id)) {
  cat("Cox GRF subgroup:", paste(grf_cox$sg.harm.id, collapse = " & "), "\n")
}
```

```{r grf-cox-trees, fig.width=10, fig.height=4}
par(mfrow = c(1, 2))
plot(grf_cox$tree1, leaf.labels = c("Control", "Treat"), main = "Cox: Depth 1")
plot(grf_cox$tree2, leaf.labels = c("Control", "Treat"), main = "Cox: Depth 2")
par(mfrow = c(1, 1))
```


### GLM (Binary) with `dmin.grf` Sweep

```{r grf-glm-sweep}
dmin_values <- c(0.00, 0.01, 0.025, 0.05, 0.075, 0.10, 0.15, 0.20)

cat("=== GRF GLM dmin.grf sweep (adverse_outcome = TRUE) ===\n\n")
cat(sprintf("  %-10s  %-15s  %-8s  %s\n",
            "dmin.grf", "Subgroup", "N cuts", "Cuts"))
cat("  ", paste(rep("-", 70), collapse = ""), "\n")

grf_glm_results <- list()

for (d in dmin_values) {
  res <- tryCatch(
    grf.subg.harm.glm(
      data = df.analysis,
      confounders.name = confounders.name,
      outcome.name = outcome.name.glm,
      treat.name = treat.name,
      id.name = id.name,
      n.min = 60,
      dmin.grf = d,
      RCT = TRUE,
      maxdepth = 2,
      seedit = 8316951,
      return_selected_cuts_only = FALSE,
      adverse_outcome = TRUE,
      details = FALSE
    ),
    error = function(e) NULL
  )

  grf_glm_results[[as.character(d)]] <- res

  found <- !is.null(res) && !is.null(res$sg.harm.id) &&
           length(res$sg.harm.id) > 0 && !all(is.na(res$sg.harm.id))

  if (found) {
    sg <- paste(res$sg.harm.id, collapse = " & ")
    nc <- length(res$tree.cuts)
    cuts <- paste(res$tree.cuts, collapse = ", ")
    cat(sprintf("  %-10.3f  %-15s  %-8d  %s\n", d, sg, nc, cuts))
  } else {
    cat(sprintf("  %-10.3f  %-15s  %-8s  %s\n", d, "(none)", "-", "-"))
  }
}
```

### GRF Comparison Table

```{r grf-comparison-table}
grf_summary <- data.frame(
  dmin.grf = dmin_values,
  Subgroup = vapply(dmin_values, function(d) {
    res <- grf_glm_results[[as.character(d)]]
    found <- !is.null(res) && !is.null(res$sg.harm.id) &&
             length(res$sg.harm.id) > 0 && !all(is.na(res$sg.harm.id))
    if (found) paste(res$sg.harm.id, collapse = " & ") else "(none)"
  }, character(1)),
  N_cuts = vapply(dmin_values, function(d) {
    res <- grf_glm_results[[as.character(d)]]
    if (!is.null(res$tree.cuts)) length(res$tree.cuts) else 0L
  }, integer(1))
)

gt(grf_summary) |>
  tab_header(
    title = md("**GRF Subgroup Selection by `dmin.grf`**"),
    subtitle = "causal_forest with adverse_outcome = TRUE"
  ) |>
  cols_label(
    dmin.grf = "dmin.grf",
    Subgroup = "Identified Subgroup",
    N_cuts   = "N Cuts"
  )
```

::: {.callout-note}
## Interpreting the `dmin.grf` Sweep

- **`dmin.grf = 0`**: Any positive treatment harm qualifies.  Most
  permissive — GRF may find subgroups driven by noise.
- **`dmin.grf = 0.05`**: Requires ≥ 5 percentage-point difference in
  event probability.  A clinically meaningful threshold for many
  binary outcomes.
- **Higher values**: Increasingly restrictive; may miss real subgroups
  with moderate effects.  In the GBSG data, the binary outcome (event
  at censoring) has weaker signal than the survival RMST, so the
  threshold at which GRF stops finding subgroups is informative.

The Cox reference uses `dmin.grf = 12` on the RMST scale (months).
:::


# ForestSearch Analysis {#sec-forestsearch}

## Parameter Mapping

| Parameter | Cox | Poisson-offset | Notes |
|---|---|---|---|
| `hr.threshold` | 1.25 | 1.25 | Converted to log(1.25) internally |
| `hr.consistency` | 1.0 | 1.0 | Converted to log(1.0) = 0 internally |
| `dmin.grf` | 12 (RMST, months) | 0.0 (probability) | Scale-specific; default auto-resolved |
| `d0.min` / `d1.min` | Min events per arm | Min subjects per arm | Repurposed for GLM |
| `use_grf` | `TRUE` | `TRUE` | Uses `causal_forest` for GLM |
| `use_lasso` | `TRUE` | `TRUE` | Uses `family = "poisson"` for GLM |

## Running ForestSearch (Poisson-Offset)

```{r forestsearch-glm, fig.width=10, fig.height=7}
#| code-fold: false

t0 <- proc.time()

fs_pois <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  outcome.name     = outcome.name.glm,
  event.name       = outcome.name.glm,
  treat.name       = treat.name,
  id.name          = id.name,
  # GLM specification
  outcome_type     = "binary",
  effect_measure   = "IRR",
  offset.name      = offset.name,
  # Threshold parameters
  hr.threshold           = 1.25,
  hr.consistency         = 1.0,
  pconsistency.threshold = 0.80,
  stop_threshold         = 0.80,
  # Search configuration
  sg_focus             = "hr",
  max_subgroups_search = 3,
  use_twostage         = TRUE,
  # Factor selection
  use_grf                   = TRUE,
  return_selected_cuts_only = TRUE,
  use_lasso                 = TRUE,
  cut_type                  = "default",
  dmin.grf                  = 0.0,    # probability scale for GLM
  # Subgroup constraints
  maxk   = 2,
  n.min  = 60,
  d0.min = 12,
  d1.min = 12,
  # Consistency evaluation
  fs.splits = 100,
  # Parallel processing
  parallel_args = par_args,
  # Output options
  showten_subgroups = TRUE,
  details           = TRUE,
  plot.sg           = FALSE
)

plan("sequential")
timings$forestsearch <- (proc.time() - t0)["elapsed"]

cat("\nForestSearch (Poisson-offset) completed in",
    round(timings$forestsearch, 1), "seconds\n")
```


## ForestSearch Results

### Identified Subgroups

```{r fs-results}
res_tabs <- sg_tables(fs_pois, ndecimals = 3, which_df = "est")
res_tabs$sg10_out
```

### Treatment Effect Estimates

```{r fs-estimates}
res_tabs$tab_estimates
```

### Identified Subgroup Definition

```{r fs-subgroup}
cat("Identified subgroup (H):", paste(fs_pois$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs_pois$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs_pois$df.est$treat.recommend == 0)))
```


# Bootstrap Bias Correction {#sec-bootstrap}

## Rationale

The GLM effect estimates from identified subgroups are **upwardly biased**
due to the selection process, just as Cox HRs are.  The same bootstrap
bias correction framework applies:

$$
\hat\theta_{\text{adj}} = 2\hat\theta_{\text{obs}} - (\hat\theta_{*} + \hat\theta^*_{*} - \hat\theta^*_{\text{obs}})
$$

where $\hat\theta$ is the log-IRR (or any GLM effect estimate).

## Running Bootstrap Analysis

```{r bootstrap}
#| code-fold: false

# Use 500–2000 for production; reduced here for the vignette
NB <- 100

t0 <- proc.time()

fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est        = fs_pois,
  nb_boots      = NB,
  show_three    = FALSE,
  details       = TRUE,
  parallel_args = par_args
)

plan("sequential")
timings$bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nBootstrap completed in",
    round(timings$bootstrap / 60, 1), "minutes\n")
```

## Bootstrap Summary

```{r bootstrap-results}
# Publication-quality gt table with bias-corrected estimates
format_bootstrap_table_gt(
  fs_bc,
  fs.est    = fs_pois,
  title     = "Bootstrap Bias-Corrected IRR Estimates",
  font_size = 12
)
```

## Bootstrap CIs

```{r bootstrap-cis}
cat("H (Questionable):\n")
cat("  Raw:            ", fs_bc$SG_CIs$H_raw, "\n")
cat("  Bias-corrected: ", fs_bc$SG_CIs$H_bc, "\n\n")
cat("Hc (Recommend):\n")
cat("  Raw:            ", fs_bc$SG_CIs$Hc_raw, "\n")
cat("  Bias-corrected: ", fs_bc$SG_CIs$Hc_bc, "\n")
```


## Bootstrap Diagnostics

```{r bootstrap-diagnostics}
boot_success <- sum(!is.na(fs_bc$results$H_biasadj_2))
cat("Success rate:", boot_success, "/", NB,
    sprintf("(%.0f%%)\n", 100 * boot_success / NB))

# Subgroup agreement across bootstrap iterations
m_cols <- paste0("M.", 1:7)
m_present <- intersect(m_cols, names(fs_bc$results))

if (length(m_present) > 0) {
  sg_strings <- apply(
    fs_bc$results[, m_present, with = FALSE], 1,
    function(row) paste(row[row != "" & !is.na(row)], collapse = " & ")
  )
  sg_tab <- sort(table(sg_strings), decreasing = TRUE)
  cat("\nSubgroups found across bootstrap iterations:\n")
  for (i in seq_along(sg_tab)) {
    cat(sprintf("  %3d/%d  %s\n", sg_tab[i], NB, names(sg_tab)[i]))
  }
}
```

## Bootstrap Distributions

```{r bootstrap-distributions, fig.width=10, fig.height=5}
h_bc  <- fs_bc$results$H_biasadj_2[!is.na(fs_bc$results$H_biasadj_2)]
hc_bc <- fs_bc$results$Hc_biasadj_2[!is.na(fs_bc$results$Hc_biasadj_2)]

if (length(h_bc) > 5 && length(hc_bc) > 5) {
  par(mfrow = c(1, 2))

  hist(h_bc, breaks = 25, col = "salmon", border = "white",
       main = "H (Questionable)\nBias-Corrected log-IRR",
       xlab = "log-IRR")
  abline(v = 0, lty = 2, lwd = 2)
  abline(v = median(h_bc), col = "red", lwd = 2)

  hist(hc_bc, breaks = 25, col = "lightblue", border = "white",
       main = "Hc (Recommend)\nBias-Corrected log-IRR",
       xlab = "log-IRR")
  abline(v = 0, lty = 2, lwd = 2)
  abline(v = median(hc_bc), col = "blue", lwd = 2)

  par(mfrow = c(1, 1))
}
```


# Cox vs. Poisson-Offset Comparison {#sec-comparison}

## Run Cox Reference

```{r cox-reference}
t0 <- proc.time()

fs_cox <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  outcome.name     = outcome.name.surv,
  event.name       = event.name,
  treat.name       = treat.name,
  id.name          = id.name,
  hr.threshold           = 1.25,
  hr.consistency         = 1.0,
  pconsistency.threshold = 0.80,
  stop_threshold         = 0.80,
  sg_focus             = "hr",
  max_subgroups_search = 3,
  use_twostage         = TRUE,
  use_grf              = TRUE,
  return_selected_cuts_only = TRUE,
  use_lasso            = TRUE,
  cut_type             = "default",
  dmin.grf             = 12,          # RMST scale (months) for survival
  maxk = 2, n.min = 60, d0.min = 12, d1.min = 12,
  fs.splits = 100,
  parallel_args = par_args,
  details = FALSE, plot.sg = FALSE
)
plan("sequential")
timings$cox <- (proc.time() - t0)["elapsed"]
```

```{r cox-bootstrap}
fs_bc_cox <- forestsearch_bootstrap_dofuture(
  fs.est     = fs_cox,
  nb_boots   = NB,
  details    = FALSE,
  parallel_args = par_args
)
plan("sequential")
```

## Side-by-Side Results

```{r side-by-side}
cat("=" |> rep(70) |> paste(collapse = ""), "\n")
cat("GBSG Subgroup Identification: Cox vs. Poisson-Offset\n")
cat("=" |> rep(70) |> paste(collapse = ""), "\n\n")

cat(sprintf("%-25s  %-25s  %-25s\n", "", "Cox (survival)", "Poisson (GLM)"))
cat("-" |> rep(70) |> paste(collapse = ""), "\n")

cat(sprintf("%-25s  %-25s  %-25s\n", "Subgroup",
    paste(fs_cox$sg.harm, collapse = " & "),
    paste(fs_pois$sg.harm, collapse = " & ")))

cat(sprintf("%-25s  %-25s  %-25s\n", "N(harm) / N(benefit)",
    paste(sum(fs_cox$df.est$treat.recommend == 0), "/",
          sum(fs_cox$df.est$treat.recommend == 1)),
    paste(sum(fs_pois$df.est$treat.recommend == 0), "/",
          sum(fs_pois$df.est$treat.recommend == 1))))

cat("-" |> rep(70) |> paste(collapse = ""), "\n")

cat(sprintf("%-25s  %-25s  %-25s\n", "H raw",
    fs_bc_cox$SG_CIs$H_raw, fs_bc$SG_CIs$H_raw))
cat(sprintf("%-25s  %-25s  %-25s\n", "H bias-corrected",
    fs_bc_cox$SG_CIs$H_bc, fs_bc$SG_CIs$H_bc))
cat(sprintf("%-25s  %-25s  %-25s\n", "Hc raw",
    fs_bc_cox$SG_CIs$Hc_raw, fs_bc$SG_CIs$Hc_raw))
cat(sprintf("%-25s  %-25s  %-25s\n", "Hc bias-corrected",
    fs_bc_cox$SG_CIs$Hc_bc, fs_bc$SG_CIs$Hc_bc))

cat("-" |> rep(70) |> paste(collapse = ""), "\n")
```

## Comparative Bootstrap Tables

### Cox (Survival)

```{r comparison-cox-boot}
format_bootstrap_table_gt(
  fs_bc_cox,
  fs.est    = fs_cox,
  title     = "Cox: Bootstrap Bias-Corrected HR Estimates",
  font_size = 12
)
```

### Poisson-Offset (GLM)

```{r comparison-pois-boot}
format_bootstrap_table_gt(
  fs_bc,
  fs.est    = fs_pois,
  title     = "Poisson: Bootstrap Bias-Corrected IRR Estimates",
  font_size = 12
)
```

## Comparative ForestSearch Tables

```{r comparison-tables}
cat("=== Cox ===\n")
res_cox <- sg_tables(fs_cox, ndecimals = 3, which_df = "est")
res_cox$tab_estimates

cat("\n=== Poisson-Offset (GLM) ===\n")
res_tabs$tab_estimates
```


# Summary and Interpretation

## Key Findings

```{r summary-findings}
cat("=" |> rep(60) |> paste(collapse = ""), "\n")
cat("FORESTSEARCH GLM EXTENSION — ANALYSIS SUMMARY\n")
cat("=" |> rep(60) |> paste(collapse = ""), "\n\n")

cat("Dataset: GBSG (N =", nrow(df.analysis), ")\n")
cat("Outcome type: Binary (event indicator)\n")
cat("Effect measure: Incidence Rate Ratio (IRR)\n")
cat("Offset: Follow-up time (months)\n\n")

cat("ITT Analysis:\n")
cat("  Cox HR (95% CI):   ", round(exp(coef(cox_itt)), 3),
    sprintf("(%.3f, %.3f)\n",
            exp(confint(cox_itt))[1], exp(confint(cox_itt))[2]))
cat("  Poisson IRR (95% CI):", round(exp(coef(pois_itt)["hormon"]), 3),
    sprintf("(%.3f, %.3f)\n",
            exp(confint(pois_itt)["hormon", 1]),
            exp(confint(pois_itt)["hormon", 2])))

cat("\nIdentified Subgroup (H):\n")
cat("  Definition:", paste(fs_pois$sg.harm, collapse = " & "), "\n")
cat("  Size:", sum(fs_pois$df.est$treat.recommend == 0),
    sprintf("(%.1f%%)\n", 100 * mean(fs_pois$df.est$treat.recommend == 0)))
cat("  IRR (unadjusted):", fs_bc$SG_CIs$H_raw, "\n")
cat("  IRR (bias-corrected):", fs_bc$SG_CIs$H_bc, "\n")

cat("\nComplement Subgroup (Hc):\n")
cat("  Size:", sum(fs_pois$df.est$treat.recommend == 1),
    sprintf("(%.1f%%)\n", 100 * mean(fs_pois$df.est$treat.recommend == 1)))
cat("  IRR (unadjusted):", fs_bc$SG_CIs$Hc_raw, "\n")
cat("  IRR (bias-corrected):", fs_bc$SG_CIs$Hc_bc, "\n")

cat("\nAgreement:\n")
cat("  Both paths identify same subgroup:", 
    identical(sort(fs_cox$sg.harm), sort(fs_pois$sg.harm)), "\n")
```

## Clinical Interpretation

The Poisson-offset GLM analysis identifies the same subgroup as the
Cox model: **estrogen receptor-negative patients with tumor size ≤ 35mm**
show potential harm from hormonal therapy.

The IRR estimates track the Cox HRs closely, with the expected slight
attenuation due to the non-constant baseline hazard in the GBSG data.
The bias-corrected IRR provides a conservative estimate that accounts
for selection-induced optimism.

## Computational Timing

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Analysis = c("GRF (Cox)", "ForestSearch (Poisson)", "Bootstrap (Poisson)",
               "Cox Reference", "Total"),
  Seconds = c(
    timings$grf_cox,
    timings$forestsearch,
    timings$bootstrap,
    timings$cox,
    timings$total
  )
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(title = "Computational Timing") |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1) |>
  cols_label(
    Analysis = "Component",
    Seconds = "Time (sec)",
    Minutes = "Time (min)"
  )
```


# References

León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024).
"Exploratory subgroup identification in the heterogeneous Cox model:
A relatively simple procedure." *Statistics in Medicine*.
DOI: 10.1002/sim.10163


# Session Information

```{r session-info}
sessionInfo()
```
