---
title: "Phase 4 Validation: GLM Summary Tables"
subtitle: "ForestSearch GLM Extension — Visualization"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: flatly
execute:
  warning: false
  message: false
---

```{r setup}
#| code-fold: false

knitr::opts_chunk$set(comment = "#>")

while (!file.exists("DESCRIPTION") && getwd() != dirname(getwd())) {
  setwd("..")
}
stopifnot("Could not find package root" = file.exists("DESCRIPTION"))

devtools::document(quiet = TRUE)
devtools::load_all(".", quiet = TRUE)

source("dev/glm/glm_effect_estimators.R")

suppressPackageStartupMessages({
  library(survival)
  library(future)
})

cat("Package root:", getwd(), "\n")
cat("R version:", R.version.string, "\n")
```


# Overview

Phase 4 adds GLM-aware summary tables via `SG_tab_estimates_glm()`,
`analyze_subgroup_glm()`, and branching in `sg_tables()`.  The tables
show per-arm outcome rates and the GLM effect estimate (RD, OR, IRR,
etc.) instead of the survival-specific columns (medians, RMST, HR).

The bootstrap `FSsg_tab` now returns a proper summary matrix with
a bias-corrected effect column (`IRR*`, `RD*`, etc.) instead of the
Phase 2 placeholder list.


# Data Preparation

```{r data-prep}
#| code-fold: false

df.analysis <- survival::gbsg
df.analysis$id <- seq_len(nrow(df.analysis))
df.analysis$grade3 <- as.integer(df.analysis$grade == "3")

confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
par_args <- list(plan = "sequential", workers = 1L, show_message = FALSE)
```


# Cox Survival Tables (Regression Guard) {#sec-cox}

```{r cox-run}
fs_cox <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  conf_force       = c("er <= 0"),
  outcome.name     = "rfstime",
  event.name       = "status",
  treat.name       = "hormon",
  id.name          = "id",
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE,
  use_grf = TRUE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = par_args,
  details = FALSE, maxk = 2, sg_focus = "hr"
)
plan("sequential")
cat("Cox subgroup:", paste(fs_cox$sg.harm, collapse = " & "), "\n")
```

## Treatment Effect Estimates

```{r cox-estimates}
res_cox <- sg_tables(fs_cox, ndecimals = 3, which_df = "est")
res_cox$tab_estimates
```

## Identified Subgroups

```{r cox-subgroups}
res_cox$sg10_out
```


# Poisson-Offset GLM Tables {#sec-poisson}

```{r pois-run}
fs_pois <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  conf_force       = c("er <= 0"),
  outcome.name     = "status",
  event.name       = "status",
  treat.name       = "hormon",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "IRR",
  offset.name      = "rfstime",
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE,
  use_grf = FALSE, use_lasso = TRUE,
  parallel_args = par_args,
  details = FALSE, maxk = 2, sg_focus = "hr", vi.grf.min = NULL
)
plan("sequential")
cat("Poisson subgroup:", paste(fs_pois$sg.harm, collapse = " & "), "\n")
```

## Treatment Effect Estimates

```{r pois-estimates}
res_pois <- sg_tables(fs_pois, ndecimals = 3, which_df = "est")
res_pois$tab_estimates
```

## Identified Subgroups

```{r pois-subgroups}
res_pois$sg10_out
```


# Bootstrap with GLM Summary Table {#sec-bootstrap}

```{r boot-pois}
boot_pois <- forestsearch_bootstrap_dofuture(
  fs.est        = fs_pois,
  nb_boots      = 100L,
  parallel_args = par_args,
  details       = TRUE
)
```

## Bootstrap Summary Table (FSsg_tab)

```{r boot-table}
if (is.data.frame(boot_pois$FSsg_tab) || is.matrix(boot_pois$FSsg_tab)) {
  cat("FSsg_tab:\n")
  print(boot_pois$FSsg_tab)
} else {
  str(boot_pois$FSsg_tab)
}
```

## Bootstrap CIs

```{r boot-cis}
cat("H (Questionable):\n")
cat("  Raw:            ", boot_pois$SG_CIs$H_raw, "\n")
cat("  Bias-corrected: ", boot_pois$SG_CIs$H_bc, "\n\n")
cat("Hc (Recommend):\n")
cat("  Raw:            ", boot_pois$SG_CIs$Hc_raw, "\n")
cat("  Bias-corrected: ", boot_pois$SG_CIs$Hc_bc, "\n")
```


# Side-by-Side Comparison {#sec-comparison}

```{r comparison}
cat("=" |> rep(70) |> paste(collapse = ""), "\n")
cat("GBSG Summary Tables: Cox vs. Poisson-Offset\n")
cat("=" |> rep(70) |> paste(collapse = ""), "\n\n")

cat("Cox (survival) table columns:\n")
cat("  Subgroup | n | n1 | events | m1 | m0 | RMST | HR (95% CI)\n\n")

cat("GLM (Poisson-offset) table columns:\n")
cat("  Subgroup | n | n1 | Rate(C) | Rate(T) | IRR (95% CI)\n\n")

cat("Both tables show:\n")
cat("  - ITT row (full population)\n")
cat("  - Questionable (H) row — identified subgroup where treatment is harmful\n")
cat("  - Recommend (Hc) row — complement where treatment is beneficial\n")
cat("  - When bootstrap results available: bias-corrected effect column\n")
```


# Session Info

```{r session-info}
sessionInfo()
```
