---
title: "Phase 2 Bootstrap Validation: Cox vs. Poisson-Offset"
subtitle: "ForestSearch GLM Extension — Bias Correction"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: flatly
execute:
  warning: false
  message: false
---

```{r setup}
#| code-fold: false

knitr::opts_chunk$set(comment = "#>")

# Navigate to package root
while (!file.exists("DESCRIPTION") && getwd() != dirname(getwd())) {
  setwd("..")
}
stopifnot(
  "Could not find package root" = file.exists("DESCRIPTION")
)

devtools::document(quiet = TRUE)
devtools::load_all(".", quiet = TRUE)

source("dev/glm/glm_effect_estimators.R")
source("dev/glm/glm_simulate.R")

suppressPackageStartupMessages({
  library(survival)
  library(future)
})

n_cores <- max(1L, parallel::detectCores(logical = FALSE) - 1L)
cat("Package root:", getwd(), "\n")
cat("R version:", R.version.string, "\n")
```


# Overview

This document validates Phase 2 of the GLM extension: bootstrap bias
correction via `forestsearch_bootstrap_dofuture()`.  Both the Cox
(survival) and Poisson-offset (GLM) paths are run on the GBSG dataset,
and the bias-corrected estimates are compared side by side.

**Expected result:** Both paths identify the same subgroup, produce
closely tracking effect estimates, and achieve high bootstrap success
rates.


# Data Preparation

```{r data-prep}
#| code-fold: false

df.analysis <- survival::gbsg
df.analysis$id <- seq_len(nrow(df.analysis))
df.analysis$grade3 <- as.integer(df.analysis$grade == "3")

confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
id.name          <- "id"

df.analysis$rfstime <- df.analysis$rfstime/30.4375

cat("GBSG: N =", nrow(df.analysis),
    ", events =", sum(df.analysis$status),
    sprintf("(%.0f%%)\n", 100 * mean(df.analysis$status)))
```

```{r bootstrap-config}
#| code-fold: false

# Bootstrap configuration
nb_boots <- 100L

# Sequential for load_all() compatibility.
# For multisession, use devtools::install() first.
boot_parallel <- list(
  plan         = "sequential",
  workers      = 1L,
  show_message = FALSE
)
```


# Cox Survival Path {#sec-cox}

## Subgroup Identification

```{r cox-forestsearch}
#| code-fold: false

fs_cox <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  conf_force       = c("er <= 0"),
  outcome.name     = "rfstime",
  event.name       = "status",
  treat.name       = "hormon",
  id.name          = id.name,
  hr.threshold           = 1.25,
  hr.consistency         = 1.0,
  pconsistency.threshold = 0.80,
  stop_threshold         = 0.80,
  sg_focus             = "hr",
  max_subgroups_search = 3,
  use_twostage         = TRUE,
  use_grf              = TRUE,
  return_selected_cuts_only = TRUE,
  use_lasso            = TRUE,
  cut_type             = "default",
  maxk   = 2,
  n.min  = 60,
  d0.min = 12,
  d1.min = 12,
  fs.splits = 100,
  parallel_args = boot_parallel,
  showten_subgroups = TRUE,
  details           = TRUE,
  plot.sg           = TRUE
)
plan("sequential")
```

```{r cox-subgroup}
cat("Cox subgroup:", paste(fs_cox$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs_cox$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs_cox$df.est$treat.recommend == 0)))
```

## Bootstrap Bias Correction

```{r cox-bootstrap}
#| code-fold: false

t0_cox <- proc.time()
boot_cox <- forestsearch_bootstrap_dofuture(
  fs.est        = fs_cox,
  nb_boots      = nb_boots,
  parallel_args = boot_parallel,
  details       = TRUE
)
elapsed_cox <- (proc.time() - t0_cox)["elapsed"]
```

```{r cox-boot-results}
cat("Elapsed:", round(elapsed_cox / 60, 1), "min\n")
cat("Success rate:",
    sum(!is.na(boot_cox$results$H_biasadj_2)), "/", nb_boots, "\n\n")
cat("H (Questionable):\n")
cat("  Raw:            ", boot_cox$SG_CIs$H_raw, "\n")
cat("  Bias-corrected: ", boot_cox$SG_CIs$H_bc, "\n\n")
cat("Hc (Recommend):\n")
cat("  Raw:            ", boot_cox$SG_CIs$Hc_raw, "\n")
cat("  Bias-corrected: ", boot_cox$SG_CIs$Hc_bc, "\n")
```

## Cox Results Tables

```{r cox-tables}
res_cox <- sg_tables(fs_cox, ndecimals = 3, which_df = "est")
res_cox$sg10_out
```

```{r cox-estimates}
res_cox$tab_estimates
```


# Poisson-Offset GLM Path {#sec-poisson}

::: {.callout-note}
## GRF and LASSO not yet wired for GLM (Phase 1)
The Poisson-offset path disables GRF and LASSO.  All confounders are
cut at quartiles/median via `cut_type = "default"`.
:::

## Subgroup Identification

```{r pois-forestsearch}
#| code-fold: false

fs_pois <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  conf_force       = c("er <= 0"),
  outcome.name     = "status",
  event.name       = "status",
  treat.name       = "hormon",
  id.name          = id.name,
  outcome_type     = "binary",
  effect_measure   = "IRR",
  offset.name      = "rfstime",
  hr.threshold           = 1.25,
  hr.consistency         = 1.0,
  pconsistency.threshold = 0.80,
  stop_threshold         = 0.80,
  sg_focus             = "hr",
  max_subgroups_search = 3,
  use_twostage         = TRUE,
  use_grf   = FALSE,
  use_lasso = FALSE,
  cut_type  = "default",
  maxk   = 2,
  n.min  = 60,
  d0.min = 12,
  d1.min = 12,
  fs.splits = 100,
  parallel_args = boot_parallel,
  showten_subgroups = TRUE,
  details           = TRUE,
  plot.sg           = FALSE
)
plan("sequential")
```

```{r pois-subgroup}
cat("Poisson subgroup:", paste(fs_pois$sg.harm, collapse = " & "), "\n")
if (!is.null(fs_pois$df.est)) {
  cat("Subgroup size:", sum(fs_pois$df.est$treat.recommend == 0),
      sprintf("(%.1f%% of ITT)\n",
              100 * mean(fs_pois$df.est$treat.recommend == 0)))
}
```

## Bootstrap Bias Correction

```{r pois-bootstrap}
#| code-fold: false

t0_pois <- proc.time()
boot_pois <- forestsearch_bootstrap_dofuture(
  fs.est        = fs_pois,
  nb_boots      = nb_boots,
  parallel_args = boot_parallel,
  details       = TRUE
)
elapsed_pois <- (proc.time() - t0_pois)["elapsed"]
```

```{r pois-boot-results}
cat("Elapsed:", round(elapsed_pois / 60, 1), "min\n")
cat("Success rate:",
    sum(!is.na(boot_pois$results$H_biasadj_2)), "/", nb_boots, "\n\n")
cat("H (Questionable):\n")
cat("  Raw:            ", boot_pois$SG_CIs$H_raw, "\n")
cat("  Bias-corrected: ", boot_pois$SG_CIs$H_bc, "\n\n")
cat("Hc (Recommend):\n")
cat("  Raw:            ", boot_pois$SG_CIs$Hc_raw, "\n")
cat("  Bias-corrected: ", boot_pois$SG_CIs$Hc_bc, "\n")
```


# Side-by-Side Comparison {#sec-comparison}

```{r comparison}
cat("=" |> rep(70) |> paste(collapse = ""), "\n")
cat("GBSG Bootstrap Bias Correction: Cox vs. Poisson-Offset\n")
cat("=" |> rep(70) |> paste(collapse = ""), "\n\n")

cat(sprintf("%-25s  %-25s  %-25s\n", "", "Cox (survival)", "Poisson (GLM)"))
cat("-" |> rep(70) |> paste(collapse = ""), "\n")

cat(sprintf("%-25s  %-25s  %-25s\n",
    "Subgroup",
    paste(fs_cox$sg.harm, collapse = " & "),
    paste(fs_pois$sg.harm, collapse = " & ")))

cat(sprintf("%-25s  %-25s  %-25s\n",
    "N(harm) / N(benefit)",
    paste(sum(fs_cox$df.est$treat.recommend == 0), "/",
          sum(fs_cox$df.est$treat.recommend == 1)),
    paste(sum(fs_pois$df.est$treat.recommend == 0), "/",
          sum(fs_pois$df.est$treat.recommend == 1))))

cat(sprintf("%-25s  %-25d  %-25d\n",
    "Bootstrap N",
    nb_boots, nb_boots))

cat(sprintf("%-25s  %-25d  %-25d\n",
    "Success count",
    sum(!is.na(boot_cox$results$H_biasadj_2)),
    sum(!is.na(boot_pois$results$H_biasadj_2))))

cat("-" |> rep(70) |> paste(collapse = ""), "\n")

cat(sprintf("%-25s  %-25s  %-25s\n",
    "H raw", boot_cox$SG_CIs$H_raw, boot_pois$SG_CIs$H_raw))
cat(sprintf("%-25s  %-25s  %-25s\n",
    "H bias-corrected", boot_cox$SG_CIs$H_bc, boot_pois$SG_CIs$H_bc))
cat(sprintf("%-25s  %-25s  %-25s\n",
    "Hc raw", boot_cox$SG_CIs$Hc_raw, boot_pois$SG_CIs$Hc_raw))
cat(sprintf("%-25s  %-25s  %-25s\n",
    "Hc bias-corrected", boot_cox$SG_CIs$Hc_bc, boot_pois$SG_CIs$Hc_bc))

cat("-" |> rep(70) |> paste(collapse = ""), "\n")
```

## ITT Agreement

```{r itt-agreement}
cox_itt <- survival::coxph(
  survival::Surv(rfstime, status) ~ hormon, data = df.analysis
)
pois_itt <- glm(
  status ~ hormon, family = poisson(link = "log"),
  offset = log(rfstime), data = df.analysis
)

cat("Cox  log-HR: ", round(coef(cox_itt)["hormon"], 4), "\n")
cat("Pois log-IRR:", round(coef(pois_itt)["hormon"], 4), "\n")
cat("Difference:  ", round(coef(pois_itt)["hormon"] - coef(cox_itt)["hormon"], 4), "\n")
```

## Bootstrap Distribution Comparison

```{r boot-distributions, fig.width=10, fig.height=5}
par(mfrow = c(1, 2))

# H subgroup
h_cox  <- boot_cox$results$H_biasadj_2
h_pois <- boot_pois$results$H_biasadj_2
h_cox  <- h_cox[!is.na(h_cox)]
h_pois <- h_pois[!is.na(h_pois)]

if (length(h_cox) > 0 && length(h_pois) > 0) {
  xlim_h <- range(c(h_cox, h_pois))
  hist(h_cox, breaks = 20, col = rgb(0.2, 0.4, 0.8, 0.5),
       main = "H Subgroup (Bias-Corrected)",
       xlab = "log-HR / log-IRR", xlim = xlim_h)
  hist(h_pois, breaks = 20, col = rgb(0.8, 0.3, 0.2, 0.5), add = TRUE)
  legend("topright", c("Cox", "Poisson"), pch = 15,
         col = c(rgb(0.2, 0.4, 0.8, 0.5), rgb(0.8, 0.3, 0.2, 0.5)))
  abline(v = 0, lty = 2)
}

# Hc complement
hc_cox  <- boot_cox$results$Hc_biasadj_2
hc_pois <- boot_pois$results$Hc_biasadj_2
hc_cox  <- hc_cox[!is.na(hc_cox)]
hc_pois <- hc_pois[!is.na(hc_pois)]

if (length(hc_cox) > 0 && length(hc_pois) > 0) {
  xlim_hc <- range(c(hc_cox, hc_pois))
  hist(hc_cox, breaks = 20, col = rgb(0.2, 0.4, 0.8, 0.5),
       main = "Hc Complement (Bias-Corrected)",
       xlab = "log-HR / log-IRR", xlim = xlim_hc)
  hist(hc_pois, breaks = 20, col = rgb(0.8, 0.3, 0.2, 0.5), add = TRUE)
  legend("topright", c("Cox", "Poisson"), pch = 15,
         col = c(rgb(0.2, 0.4, 0.8, 0.5), rgb(0.8, 0.3, 0.2, 0.5)))
  abline(v = 0, lty = 2)
}

par(mfrow = c(1, 1))
```

## Bootstrap Success Rate by Subgroup

```{r boot-subgroups}
# Which subgroups were found across bootstrap iterations?
m_cols <- paste0("M.", 1:7)
m_cols_present <- intersect(m_cols, names(boot_cox$results))

if (length(m_cols_present) > 0) {
  cat("Cox bootstrap — subgroups found:\n")
  sg_strings_cox <- apply(
    boot_cox$results[, m_cols_present, with = FALSE], 1,
    function(row) paste(row[row != "" & !is.na(row)], collapse = " & ")
  )
  sg_tab_cox <- sort(table(sg_strings_cox), decreasing = TRUE)
  for (i in seq_along(sg_tab_cox)) {
    cat(sprintf("  %3d/%d  %s\n", sg_tab_cox[i], nb_boots, names(sg_tab_cox)[i]))
  }

  cat("\nPoisson bootstrap — subgroups found:\n")
  sg_strings_pois <- apply(
    boot_pois$results[, m_cols_present, with = FALSE], 1,
    function(row) paste(row[row != "" & !is.na(row)], collapse = " & ")
  )
  sg_tab_pois <- sort(table(sg_strings_pois), decreasing = TRUE)
  for (i in seq_along(sg_tab_pois)) {
    cat(sprintf("  %3d/%d  %s\n", sg_tab_pois[i], nb_boots, names(sg_tab_pois)[i]))
  }
}
```


# Session Info

```{r session-info}
sessionInfo()
```
