---
title: "GBSG Analysis: Cox vs. Poisson-Offset Comparison"
subtitle: "ForestSearch GLM Extension — Sanity Check"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: flatly
execute:
  warning: false
  message: false
---

```{r setup}
#| code-fold: false

knitr::opts_chunk$set(comment = "#>")

# Navigate to package root
while (!file.exists("DESCRIPTION") && getwd() != dirname(getwd())) {
  setwd("..")
}
stopifnot(
  "Could not find package root" = file.exists("DESCRIPTION")
)

devtools::document(quiet = TRUE)
devtools::load_all(".", quiet = TRUE)

# Source dev/glm closures (not yet graduated to R/)
source("dev/glm/glm_effect_estimators.R")
source("dev/glm/glm_simulate.R")

suppressPackageStartupMessages({
  library(survival)
  library(future)
})

n_cores <- max(1L, parallel::detectCores(logical = FALSE) - 1L)
cat("Package root:", getwd(), "\n")
cat("R version:", R.version.string, "\n")
cat("Workers:", n_cores, "\n")
```

# Data Preparation

The GBSG breast cancer dataset (`survival::gbsg`, N = 686) is prepared
identically for both the Cox and Poisson-offset analyses.

```{r data-prep}
#| code-fold: false

df.analysis <- survival::gbsg
df.analysis$id <- seq_len(nrow(df.analysis))
df.analysis$grade3 <- as.integer(df.analysis$grade == "3")

confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name     <- "rfstime"
event.name       <- "status"
treat.name       <- "hormon"
id.name          <- "id"

cat("GBSG: N =", nrow(df.analysis),
    ", events =", sum(df.analysis[[event.name]]),
    sprintf("(%.0f%%)\n", 100 * mean(df.analysis[[event.name]])))
```


# Cox Survival Analysis (Reference) {#sec-cox}

This is the standard ForestSearch analysis from the primary vignette.

```{r forestsearch-cox, fig.width=10, fig.height=7}
#| code-fold: false

fs_cox <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  outcome.name     = outcome.name,
  treat.name       = treat.name,
  event.name       = event.name,
  id.name          = id.name,
  # Threshold parameters (per León et al. 2024)
  hr.threshold           = 1.25,
  hr.consistency         = 1.0,
  pconsistency.threshold = 0.80,
  stop_threshold         = 0.80,
  # Search configuration
  sg_focus             = "hr",
  max_subgroups_search = 3,
  use_twostage         = TRUE,
  # Factor selection
  use_grf                    = TRUE,
  return_selected_cuts_only  = TRUE,
  use_lasso                  = TRUE,
  cut_type                   = "default",
  # Subgroup constraints
  maxk   = 2,
  n.min  = 60,
  d0.min = 12,
  d1.min = 12,
  # Consistency evaluation
  fs.splits = 100,
  # Sequential for load_all() compatibility
  parallel_args = list(
    plan         = "sequential",
    workers      = n_cores,
    show_message = TRUE
  ),
  # Output options
  showten_subgroups = TRUE,
  details           = TRUE,
  plot.sg           = TRUE
)
plan("sequential")
```

## Cox Results

### Identified Subgroups

```{r cox-results}
res_cox <- sg_tables(fs_cox, ndecimals = 3, which_df = "est")
res_cox$sg10_out
```

### Treatment Effect Estimates

```{r cox-estimates}
res_cox$tab_estimates
```

### Identified Subgroup Definition

```{r cox-subgroup}
cat("Identified subgroup (H):", paste(fs_cox$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs_cox$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs_cox$df.est$treat.recommend == 0)))
```


# Poisson-Offset GLM Analysis (Sanity Check) {#sec-poisson}

The Poisson regression with `log(rfstime)` as offset approximates the Cox
model when the baseline hazard is locally constant.  Under this
approximation, `exp(log-IRR)` should track the Cox hazard ratio, and the
identified subgroup should be similar or identical.

The key API difference: `outcome.name` is the binary event indicator,
`offset.name` points to the follow-up time column, and `outcome_type = "binary"`
with `effect_measure = "IRR"`.

::: {.callout-note}
## GRF and LASSO not yet wired for GLM outcomes (Phase 1)

The Poisson-offset analysis disables GRF and LASSO candidate selection
because these components use `causal_survival_forest()` and
`glmnet(family = "cox")` respectively, which require survival-specific
inputs.  Wiring `causal_forest()` and `glmnet(family = "poisson")` is
Phase 3 work.  For now, all confounders are cut at quartiles/median
(the `cut_type = "default"` path).
:::

```{r forestsearch-poisson}
#| code-fold: false

fs_pois <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  conf_force = c("er <= 0"),
  outcome.name     = event.name,        # binary event indicator (0/1)
  event.name       = event.name,        # same column (binary IS the event)
  treat.name       = treat.name,
  id.name          = id.name,
  # GLM outcome specification
  outcome_type     = "binary",
  effect_measure   = "IRR",
  offset.name      = outcome.name,      # "rfstime" — log(rfstime) used as offset
  # Threshold parameters — same as Cox for comparability
  # IRR thresholds map directly to HR thresholds
  hr.threshold           = 1.25,        # log(1.25) screening on log-IRR scale
  hr.consistency         = 1.0,         # log(1.0) = 0 consistency threshold
  pconsistency.threshold = 0.80,
  stop_threshold         = 0.80,
  # Search configuration
  sg_focus             = "hr",
  max_subgroups_search = 3,
  use_twostage         = TRUE,
  # Factor selection — GRF/LASSO not yet wired for GLM (Phase 1)
  use_grf   = FALSE,
  use_lasso = FALSE,
  cut_type  = "default",
  # Subgroup constraints
  maxk   = 2,
  n.min  = 60,
  d0.min = 12,
  d1.min = 12,
  # Consistency evaluation
  fs.splits = 100,
  # Sequential for load_all() compatibility
  parallel_args = list(
    plan         = "sequential",
    workers      = n_cores,
    show_message = TRUE
  ),
  # Output options
  showten_subgroups = TRUE,
  details           = TRUE,
  plot.sg           = FALSE              # KM plots not applicable for GLM
)
plan("sequential")
```

## Poisson-Offset Results

### Identified Subgroup Definition

```{r pois-subgroup}
if (!is.null(fs_pois$sg.harm)) {
  cat("Identified subgroup (H):", paste(fs_pois$sg.harm, collapse = " & "), "\n")
  cat("Subgroup size:", sum(fs_pois$df.est$treat.recommend == 0),
      sprintf("(%.1f%% of ITT)\n",
              100 * mean(fs_pois$df.est$treat.recommend == 0)))
} else {
  cat("No subgroup identified by Poisson-offset path\n")
}
```

### Candidate Subgroups from Search

```{r pois-candidates}
pois_cands <- fs_pois$find.grps$out.found$hr.subgroups
if (!is.null(pois_cands) && nrow(pois_cands) > 0) {
  # HR column contains log-IRR in the GLM path; exponentiate for comparison
  pois_cands$IRR <- exp(pois_cands$HR)

  factor_cols <- setdiff(
    names(pois_cands),
    c("grp", "K", "n", "E", "d1", "m1", "m0", "HR", "L(HR)", "U(HR)", "IRR")
  )

  cat("Top candidates (sorted by IRR):\n\n")
  top <- head(pois_cands[order(-pois_cands$IRR), ], 10)
  for (i in seq_len(nrow(top))) {
    active <- factor_cols[
      as.numeric(unlist(top[i, factor_cols, with = FALSE])) == 1
    ]
    cat(sprintf("  #%d: log-IRR = %6.3f  (IRR = %5.2f)  n = %3d  %s\n",
                i, top$HR[i], top$IRR[i], top$n[i],
                paste(active, collapse = " & ")))
  }
} else {
  cat("No candidates found\n")
}
```


# Side-by-Side Comparison {#sec-comparison}

```{r comparison}
cat("=" |> rep(70) |> paste(collapse = ""), "\n")
cat("GBSG Subgroup Identification: Cox vs. Poisson-Offset\n")
cat("=" |> rep(70) |> paste(collapse = ""), "\n\n")

# Cox results
cat("Cox (survival) path:\n")
if (!is.null(fs_cox$sg.harm)) {
  cat("  Subgroup: ", paste(fs_cox$sg.harm, collapse = " & "), "\n")
  cat("  N(harm):  ", sum(fs_cox$df.est$treat.recommend == 0), "\n")
  cat("  N(benefit):", sum(fs_cox$df.est$treat.recommend == 1), "\n")
} else {
  cat("  No subgroup identified\n")
}

cat("\nPoisson-offset (GLM) path:\n")
if (!is.null(fs_pois$sg.harm)) {
  cat("  Subgroup: ", paste(fs_pois$sg.harm, collapse = " & "), "\n")
  cat("  N(harm):  ", sum(fs_pois$df.est$treat.recommend == 0), "\n")
  cat("  N(benefit):", sum(fs_pois$df.est$treat.recommend == 1), "\n")
} else {
  cat("  No subgroup identified\n")
}

# ITT-level comparison of Cox log-HR vs Poisson log-IRR
cat("\nITT-level effect estimates:\n")
cox_itt <- survival::coxph(
  survival::Surv(rfstime, status) ~ hormon, data = df.analysis
)
pois_itt <- glm(
  status ~ hormon, family = poisson(link = "log"),
  offset = log(rfstime), data = df.analysis
)
cat("  Cox  log-HR: ", round(coef(cox_itt)["hormon"], 4), "\n")
cat("  Pois log-IRR:", round(coef(pois_itt)["hormon"], 4), "\n")
cat("  Difference:  ", round(coef(pois_itt)["hormon"] - coef(cox_itt)["hormon"], 4), "\n")
```

::: {.callout-note}
## Interpreting the Comparison

The Cox and Poisson-offset paths may identify **different subgroups** for
two reasons:

1. **Candidate factor sets differ.** The Cox path uses GRF + LASSO for
   variable selection; the Poisson path uses quartile cuts only (Phase 1).
   This means the Cox path evaluates a smaller, more targeted set of
   candidates.

2. **The Poisson approximation is imperfect.** The GBSG data has a Weibull-like
   hazard (not constant), so the Poisson log-IRR diverges slightly from the
   Cox log-HR, especially in subgroups with heterogeneous follow-up.

Despite these differences, the ITT-level estimates should agree closely,
and the identified subgroups should share common factors (particularly
estrogen receptor status, which is the dominant prognostic/predictive
variable in this dataset).

When GRF and LASSO are wired for GLM outcomes (Phase 3), the candidate
sets will align and the comparison will be tighter.
:::


# Session Info

```{r session-info}
sessionInfo()
```
