---
title: "Phase 3 Validation: GRF + LASSO for GLM Outcomes"
subtitle: "ForestSearch GLM Extension — Variable Selection"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: flatly
execute:
  warning: false
  message: false
---

```{r setup}
#| code-fold: false

knitr::opts_chunk$set(comment = "#>")

# Navigate to package root
while (!file.exists("DESCRIPTION") && getwd() != dirname(getwd())) {
  setwd("..")
}
stopifnot("Could not find package root" = file.exists("DESCRIPTION"))

devtools::document(quiet = TRUE)
devtools::load_all(".", quiet = TRUE)

source("dev/glm/glm_effect_estimators.R")
source("dev/glm/glm_simulate.R")

suppressPackageStartupMessages({
  library(survival)
  library(future)
})

cat("Package root:", getwd(), "\n")
cat("R version:", R.version.string, "\n")
```


# Overview

Phase 3 wires GRF and LASSO candidate selection for GLM outcomes:

| Component | Survival (existing) | GLM (new) |
|---|---|---|
| **GRF** | `causal_survival_forest(X, Y, W, D, horizon)` | `causal_forest(X, Y, W)` |
| **LASSO** | `glmnet(family = "cox")` | `family = "binomial"` / `"poisson"` / `"gaussian"` |
| **VI screening** | `causal_survival_forest` | `causal_forest` |

This document tests all combinations on the GBSG dataset.


# Data Preparation

```{r data-prep}
#| code-fold: false

df.analysis <- survival::gbsg
df.analysis$id <- seq_len(nrow(df.analysis))
df.analysis$grade3 <- as.integer(df.analysis$grade == "3")

confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
id.name          <- "id"

# Common parallel config (sequential for load_all compatibility)
par_args <- list(plan = "sequential", workers = 1L, show_message = FALSE)

cat("GBSG: N =", nrow(df.analysis), "\n")
```


# LASSO Family Branching {#sec-lasso}

Direct tests of `lasso_selection()` with each family.

```{r lasso-cox}
cat("=== Cox LASSO (family = 'cox') ===\n")
lasso_cox <- lasso_selection(
  df = df.analysis,
  confounders.name = confounders.name,
  outcome.name = "rfstime", event.name = "status",
  outcome_type = "survival"
)
cat("  Selected:", paste(lasso_cox$selected, collapse = ", "), "\n")
cat("  Omitted: ", paste(lasso_cox$omitted, collapse = ", "), "\n")
```

```{r lasso-binomial}
cat("=== Binomial LASSO (family = 'binomial') ===\n")
lasso_bin <- lasso_selection(
  df = df.analysis,
  confounders.name = confounders.name,
  outcome.name = "status", event.name = "status",
  outcome_type = "binary"
)
cat("  Selected:", paste(lasso_bin$selected, collapse = ", "), "\n")
cat("  Omitted: ", paste(lasso_bin$omitted, collapse = ", "), "\n")
```

```{r lasso-poisson}
cat("=== Poisson LASSO (family = 'poisson', offset = log(rfstime)) ===\n")
lasso_pois <- lasso_selection(
  df = df.analysis,
  confounders.name = confounders.name,
  outcome.name = "status", event.name = "status",
  outcome_type = "binary",
  offset.name = "rfstime"
)
cat("  Selected:", paste(lasso_pois$selected, collapse = ", "), "\n")
cat("  Omitted: ", paste(lasso_pois$omitted, collapse = ", "), "\n")
```

```{r lasso-comparison}
cat("\n=== LASSO Comparison ===\n")
all_vars <- confounders.name
lasso_tbl <- data.frame(
  Variable  = all_vars,
  Cox       = ifelse(all_vars %in% lasso_cox$selected, "Y", "-"),
  Binomial  = ifelse(all_vars %in% lasso_bin$selected, "Y", "-"),
  Poisson   = ifelse(all_vars %in% lasso_pois$selected, "Y", "-")
)
print(lasso_tbl, row.names = FALSE)
```


# GRF: Causal Forest for GLM Outcomes {#sec-grf}

Direct test of `grf.subg.harm.glm()` vs `grf.subg.harm.survival()`.

```{r grf-survival}
cat("=== GRF: causal_survival_forest ===\n")
grf_surv <- grf.subg.harm.survival(
  data = df.analysis,
  confounders.name = confounders.name,
  outcome.name = "rfstime", event.name = "status",
  treat.name = "hormon", id.name = "id",
  n.min = 60, dmin.grf = 0.0,
  RCT = TRUE, maxdepth = 2, seedit = 8316951,
  return_selected_cuts_only = TRUE, details = TRUE
)
cat("  Cuts:", paste(grf_surv$tree.cuts, collapse = ", "), "\n")
if (!is.null(grf_surv$sg.harm.id)) {
  cat("  Subgroup:", paste(grf_surv$sg.harm.id, collapse = " & "), "\n")
}
```

```{r grf-glm}
cat("=== GRF: causal_forest (binary outcome) ===\n")
grf_glm <- grf.subg.harm.glm(
  data = df.analysis,
  confounders.name = confounders.name,
  outcome.name = "status",
  treat.name = "hormon", id.name = "id",
  n.min = 60, dmin.grf = 0.0,
  RCT = TRUE, maxdepth = 2, seedit = 8316951,
  return_selected_cuts_only = TRUE, details = TRUE
)
cat("  Cuts:", paste(grf_glm$tree.cuts, collapse = ", "), "\n")
if (!is.null(grf_glm$sg.harm.id)) {
  cat("  Subgroup:", paste(grf_glm$sg.harm.id, collapse = " & "), "\n")
} else {
  cat("  No subgroup identified (expected — see note below)\n")
}
```

::: {.callout-note}
## Why `causal_forest` may not identify the same subgroup

`causal_survival_forest` estimates RMST-based treatment effects, which
capture time-to-event heterogeneity.  `causal_forest` estimates CATE on
the binary outcome (event yes/no), which is less sensitive to hazard
timing.  In the GBSG data where the treatment effect manifests through
delayed recurrence rather than event prevention, the survival GRF
typically finds stronger heterogeneity.  This is expected behavior, not
a bug — the two forests answer different questions.

The key point: when `causal_forest` doesn't find a subgroup, LASSO and
forced confounders carry the candidate set, and the downstream
search + consistency pipeline operates identically.
:::


# Full Pipeline Comparison {#sec-pipeline}

Four configurations on the same GBSG data.

## Config A: Cox + GRF + LASSO (reference)

```{r config-a}
#| code-fold: false

fs_a <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  conf_force       = c("er <= 0"),
  outcome.name     = "rfstime",
  event.name       = "status",
  treat.name       = "hormon",
  id.name          = id.name,
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE,
  use_grf = TRUE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = par_args,
  details = TRUE, maxk = 2, sg_focus = "hr"
)
plan("sequential")
cat("Config A subgroup:", paste(fs_a$sg.harm, collapse = " & "), "\n")
```

## Config B: Poisson-offset + GRF + LASSO

```{r config-b}
#| code-fold: false

fs_b <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  conf_force       = c("er <= 0"),
  outcome.name     = "status",
  event.name       = "status",
  treat.name       = "hormon",
  id.name          = id.name,
  outcome_type = "binary", effect_measure = "IRR",
  offset.name  = "rfstime",
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE,
  use_grf = TRUE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = par_args,
  details = TRUE, maxk = 2, sg_focus = "hr", vi.grf.min = NULL
)
plan("sequential")
cat("Config B subgroup:", paste(fs_b$sg.harm, collapse = " & "), "\n")
```

## Config C: Poisson-offset + LASSO only (no GRF)

```{r config-c}
#| code-fold: false

fs_c <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  conf_force       = c("er <= 0"),
  outcome.name     = "status",
  event.name       = "status",
  treat.name       = "hormon",
  id.name          = id.name,
  outcome_type = "binary", effect_measure = "IRR",
  offset.name  = "rfstime",
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE,
  use_grf = FALSE, use_lasso = TRUE,
  parallel_args = par_args,
  details = TRUE, maxk = 2, sg_focus = "hr", vi.grf.min = NULL
)
plan("sequential")
cat("Config C subgroup:", paste(fs_c$sg.harm, collapse = " & "), "\n")
```

## Config D: Poisson-offset + default cuts only (no GRF, no LASSO)

```{r config-d}
#| code-fold: false

fs_d <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  conf_force       = c("er <= 0"),
  outcome.name     = "status",
  event.name       = "status",
  treat.name       = "hormon",
  id.name          = id.name,
  outcome_type = "binary", effect_measure = "IRR",
  offset.name  = "rfstime",
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 20,
  use_twostage = TRUE,
  use_grf = FALSE, use_lasso = FALSE,
  parallel_args = par_args,
  details = TRUE, maxk = 2, sg_focus = "hr", vi.grf.min = NULL
)
plan("sequential")
cat("Config D subgroup:", paste(fs_d$sg.harm, collapse = " & "), "\n")
```

## Side-by-Side Summary

```{r summary-table}
configs <- list(
  list(label = "A: Cox + GRF + LASSO",         fs = fs_a),
  list(label = "B: Poisson + GRF + LASSO",     fs = fs_b),
  list(label = "C: Poisson + LASSO only",       fs = fs_c),
  list(label = "D: Poisson + default cuts only", fs = fs_d)
)

cat("=" |> rep(80) |> paste(collapse = ""), "\n")
cat("GBSG Phase 3 Comparison\n")
cat("=" |> rep(80) |> paste(collapse = ""), "\n\n")

cat(sprintf("%-35s  %-30s  %5s  %5s\n",
            "Config", "Subgroup", "N(H)", "Cand"))
cat("-" |> rep(80) |> paste(collapse = ""), "\n")

for (cfg in configs) {
  sg <- if (!is.null(cfg$fs$sg.harm)) {
    paste(cfg$fs$sg.harm, collapse = " & ")
  } else {
    "(none)"
  }
  n_harm <- if (!is.null(cfg$fs$df.est)) {
    sum(cfg$fs$df.est$treat.recommend == 0)
  } else {
    NA
  }
  n_cand <- if (!is.null(cfg$fs$find.grps$out.found$hr.subgroups)) {
    nrow(cfg$fs$find.grps$out.found$hr.subgroups)
  } else {
    0
  }
  cat(sprintf("%-35s  %-30s  %5s  %5d\n",
              cfg$label, sg,
              if (!is.na(n_harm)) as.character(n_harm) else "-",
              n_cand))
}
cat("-" |> rep(80) |> paste(collapse = ""), "\n")
```

::: {.callout-note}
## Interpreting the results

All four configurations should identify the same (or very similar)
subgroup, demonstrating that the variable selection method affects the
candidate set size but not the final answer when the signal is strong.

Key observations:

- **Config A** (Cox, reference): Uses `causal_survival_forest` + Cox LASSO.
  The gold standard.
- **Config B** (Poisson + GRF + LASSO): Uses `causal_forest` (binary CATE)
  + Poisson LASSO.  GRF may find different or no cuts, but LASSO +
  `conf_force` ensure the candidate set is adequate.
- **Config C** (Poisson + LASSO only): Skips GRF entirely.  LASSO alone
  selects the right confounders.
- **Config D** (default cuts only): No variable selection.  The largest
  candidate set — all quartile/median cuts of all confounders.  Slower
  search but same answer.
:::


# Session Info

```{r session-info}
sessionInfo()
```
