---
title: "GLM Forest Plot Testing"
subtitle: "Comprehensive Evaluation of plot_subgroup_results_forestplot for GLM Outcomes"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 11
    fig-height: 8
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false

pkg_root <- getwd()
while (!file.exists(file.path(pkg_root, "DESCRIPTION")) &&
       pkg_root != dirname(pkg_root)) {
  pkg_root <- dirname(pkg_root)
}
stopifnot("Could not find package root" = file.exists(
  file.path(pkg_root, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = pkg_root)
knitr::opts_chunk$set(comment = "#>", fig.align = "center")
```

```{r load}
devtools::document(quiet = TRUE)
devtools::load_all(".", quiet = TRUE)
source("dev/glm/glm_effect_estimators.R")

suppressPackageStartupMessages({
  library(survival)
  library(data.table)
  library(grf)
  library(glmnet)
  library(gt)
  library(forestploter)
})

# Test tracker
test_log <- data.frame(
  Test = character(), Description = character(),
  Status = character(), Details = character(),
  stringsAsFactors = FALSE
)

log_test <- function(name, desc, status, details = "") {
  test_log <<- rbind(test_log, data.frame(
    Test = name, Description = desc,
    Status = status, Details = details,
    stringsAsFactors = FALSE
  ))
  icon <- if (status == "PASS") "\u2705" else "\u274C"
  cat(icon, name, ":", desc, "--", status, "\n")
  if (nzchar(details)) cat("    ", details, "\n")
}
```


# Synthetic Data {#sec-data}

Two datasets are prepared:

- **GBSG (real):** Used for survival and binary/rate-based GLM
  outcomes (event = `status`, offset = `rfstime`).
- **Simulated continuous:** Synthetic data with a known continuous
  outcome and treatment-subgroup interaction.

```{r data-prep}
#| code-fold: false

# --- GBSG setup ---
df <- survival::gbsg
df$id <- seq_len(nrow(df))
df$grade3 <- as.integer(df$grade == "3")
confs <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")

cat("GBSG: N =", nrow(df), ", events =", sum(df$status),
    ", P(hormon) =", round(mean(df$hormon), 3), "\n")

# --- Simulated continuous outcome ---
set.seed(42)
n_sim <- 800
df_cont <- data.frame(
  id     = seq_len(n_sim),
  age    = rnorm(n_sim, 55, 10),
  bm     = rnorm(n_sim, 0, 1),
  treat  = rbinom(n_sim, 1, 0.5)
)
# True subgroup: bm > 0 has harm (positive MD)
df_cont$bm_pos <- as.integer(df_cont$bm > 0)
df_cont$outcome <- with(df_cont,
  10 + 0.5 * age + ifelse(bm > 0, 3.0, -2.0) * treat + rnorm(n_sim, 0, 3)
)

cat("Continuous sim: N =", n_sim,
    ", subgroup {bm>0}: n =", sum(df_cont$bm_pos),
    ", true MD(sg) = 3.0, true MD(co) = -2.0\n")
```


# Test Group 1: Survival Forest Plot (Regression Baseline) {#sec-survival}

Verify the existing survival (HR) pathway works unchanged.

## 1.1: ForestSearch + Bootstrap + Kfold (Survival)

```{r fs-survival}
n_cores <- max(1L, floor(parallel::detectCores() * 0.80))

fs_surv <- forestsearch(
  df.analysis = df, confounders.name = confs, conf_force = c("er <= 0"),
  outcome.name = "rfstime", event.name = "status",
  treat.name = "hormon", id.name = "id",
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
  maxk = 2, sg_focus = "hr"
)
cat("Survival SG:", paste(fs_surv$sg.harm, collapse = " & "), "\n")
```

```{r boot-survival}
#| eval: !expr '!is.null(fs_surv$sg.harm)'

fs_bc_surv <- forestsearch_bootstrap_dofuture(
  fs.est = fs_surv, nb_boots = 20,
  parallel_args = list(plan = "sequential", workers = n_cores,
                       show_message = FALSE),
  details = FALSE
)
cat("Bootstrap success:",
    round(100 * mean(!is.na(fs_bc_surv$results$H_biasadj_2))), "%\n")
```

```{r cv-survival}
#| eval: !expr '!is.null(fs_surv$sg.harm)'

fs_kfold_surv <- forestsearch_Kfold(
  fs.est = fs_surv, Kfolds = 5,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
)
cat("CV found:", fs_kfold_surv$prop_SG_found, "%\n")
```

## 1.2: Survival Forest Plot

```{r fp-survival, fig.width=11, fig.height=7}
#| eval: !expr '!is.null(fs_surv$sg.harm)'

test_id <- "1.2"
result <- tryCatch({
  fp_surv <- plot_subgroup_results_forestplot(
    fs_results = list(fs.est = fs_surv, fs_bc = fs_bc_surv,
                      fs_kfold = fs_kfold_surv),
    df_analysis = df,
    outcome.name = "rfstime", event.name = "status", treat.name = "hormon",
    E.name = "Hormon", C.name = "Control",
    title_text = "Test 1.2: Survival (HR) Forest Plot"
  )
  plot(fp_surv)
  paste0("rows=", nrow(fp_surv$data), ", class=", class(fp_surv)[1])
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "Survival HR forest plot renders", "PASS", result)
} else {
  log_test(test_id, "Survival HR forest plot renders", "FAIL", result)
}
```


# Test Group 2: Binary GLM -- Odds Ratio (OR) {#sec-or}

## 2.1: ForestSearch + Bootstrap (OR)

```{r fs-or}
fs_or <- forestsearch(
  df.analysis = df, confounders.name = confs, conf_force = c("er <= 0"),
  outcome.name = "status", event.name = "status",
  treat.name = "hormon", id.name = "id",
  outcome_type = "binary", effect_measure = "OR", dmin.grf = 0.0,
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
  maxk = 2, sg_focus = "hr"
)
cat("OR SG:", paste(fs_or$sg.harm, collapse = " & "), "\n")
```

```{r boot-or}
#| eval: !expr '!is.null(fs_or$sg.harm)'

fs_bc_or <- forestsearch_bootstrap_dofuture(
  fs.est = fs_or, nb_boots = 20,
  parallel_args = list(plan = "sequential", workers = n_cores,
                       show_message = FALSE),
  details = FALSE
)
```

```{r cv-or}
#| eval: !expr '!is.null(fs_or$sg.harm)'

fs_kfold_or <- forestsearch_Kfold(
  fs.est = fs_or, Kfolds = 5,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
)
```

## 2.2: OR Forest Plot

```{r fp-or, fig.width=11, fig.height=7}
#| eval: !expr '!is.null(fs_or$sg.harm)'

test_id <- "2.2"
result <- tryCatch({
  fp_or <- plot_subgroup_results_forestplot(
    fs_results = list(fs.est = fs_or, fs_bc = fs_bc_or,
                      fs_kfold = fs_kfold_or),
    df_analysis = df,
    outcome.name = "status", event.name = "status", treat.name = "hormon",
    E.name = "Hormon", C.name = "Control",
    title_text = "Test 2.2: Binary Outcome -- OR Forest Plot"
  )
  plot(fp_or)

  # Verify auto-detected display defaults
  ci_col <- grep("^OR", names(fp_or$data), value = TRUE)
  checks <- c(
    has_or_label = length(ci_col) > 0,
    outcome_type = fs_or$outcome_type == "binary",
    measure      = fs_or$effect_measure == "OR"
  )
  paste0("rows=", nrow(fp_or$data), ", ci_col=", ci_col[1],
         ", checks=", paste(names(checks[checks]), collapse = "+"))
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "OR forest plot renders with correct labels", "PASS", result)
} else {
  log_test(test_id, "OR forest plot renders with correct labels", "FAIL", result)
}
```


# Test Group 3: Binary GLM -- Incidence Rate Ratio (IRR) {#sec-irr}

## 3.1: ForestSearch + Bootstrap (IRR)

```{r fs-irr}
fs_irr <- forestsearch(
  df.analysis = df, confounders.name = confs, conf_force = c("er <= 0"),
  outcome.name = "status", event.name = "status",
  treat.name = "hormon", id.name = "id",
  outcome_type = "binary", effect_measure = "IRR", offset.name = "rfstime",
  dmin.grf = 0.0,
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
  maxk = 2, sg_focus = "hr"
)
cat("IRR SG:", paste(fs_irr$sg.harm, collapse = " & "), "\n")
```

```{r boot-irr}
#| eval: !expr '!is.null(fs_irr$sg.harm)'

fs_bc_irr <- forestsearch_bootstrap_dofuture(
  fs.est = fs_irr, nb_boots = 50,
  parallel_args = list(plan = "sequential", workers = n_cores,
                       show_message = FALSE),
  details = FALSE
)
```

## 3.2: IRR Forest Plot

```{r fp-irr, fig.width=11, fig.height=7}
#| eval: !expr '!is.null(fs_irr$sg.harm)'

test_id <- "3.2"
result <- tryCatch({
  fp_irr <- plot_subgroup_results_forestplot(
    fs_results = list(fs.est = fs_irr, fs_bc = fs_bc_irr),
    df_analysis = df,
    outcome.name = "status", event.name = "status", treat.name = "hormon",
    E.name = "Hormon", C.name = "Control",
    title_text = "Test 3.2: Binary Outcome -- IRR Forest Plot"
  )
  plot(fp_irr)

  ci_col <- grep("^IRR", names(fp_irr$data), value = TRUE)
  paste0("rows=", nrow(fp_irr$data), ", ci_col=", ci_col[1])
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "IRR forest plot renders (log scale, ref=1)", "PASS", result)
} else {
  log_test(test_id, "IRR forest plot renders (log scale, ref=1)", "FAIL", result)
}
```


# Test Group 4: Binary GLM -- Risk Difference (RD) {#sec-rd}

## 4.1: ForestSearch (RD)

```{r fs-rd}
fs_rd <- forestsearch(
  df.analysis = df, confounders.name = confs, conf_force = c("er <= 0"),
  outcome.name = "status", event.name = "status",
  treat.name = "hormon", id.name = "id",
  outcome_type = "binary", effect_measure = "RD", dmin.grf = 0.0,
  hr.threshold = 0.05, hr.consistency = 0.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
  maxk = 2, sg_focus = "hr"
)
cat("RD SG:", paste(fs_rd$sg.harm, collapse = " & "), "\n")
```

## 4.2: RD Forest Plot (Identity Scale)

```{r fp-rd, fig.width=11, fig.height=7}
#| eval: !expr '!is.null(fs_rd$sg.harm)'

test_id <- "4.2"
result <- tryCatch({
  fp_rd <- plot_subgroup_results_forestplot(
    fs_results = list(fs.est = fs_rd),
    df_analysis = df,
    outcome.name = "status", event.name = "status", treat.name = "hormon",
    E.name = "Hormon", C.name = "Control",
    title_text = "Test 4.2: Binary Outcome -- RD Forest Plot (identity scale)"
  )
  plot(fp_rd)

  ci_col <- grep("^RD", names(fp_rd$data), value = TRUE)
  paste0("rows=", nrow(fp_rd$data), ", ci_col=", ci_col[1],
         ", identity_scale=TRUE")
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "RD forest plot renders (linear scale, ref=0)", "PASS", result)
} else {
  log_test(test_id, "RD forest plot renders (linear scale, ref=0)", "FAIL", result)
}
```


# Test Group 5: Binary GLM -- IRR + IPTW (Observational) {#sec-iptw}

## 5.1: ForestSearch + Bootstrap (IRR + IPTW)

```{r fs-iptw}
fs_iptw <- forestsearch(
  df.analysis = df, confounders.name = confs, conf_force = c("er <= 0"),
  outcome.name = "status", event.name = "status",
  treat.name = "hormon", id.name = "id",
  outcome_type = "binary", effect_measure = "IRR", offset.name = "rfstime",
  dmin.grf = 0.0,
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
  maxk = 2, sg_focus = "hr",
  is.RCT = FALSE, ps_method = "grf", ps_adjust_method = "iptw"
)
cat("IPTW IRR SG:", paste(fs_iptw$sg.harm, collapse = " & "), "\n")
```

```{r boot-iptw}
#| eval: !expr '!is.null(fs_iptw$sg.harm)'

fs_bc_iptw <- forestsearch_bootstrap_dofuture(
  fs.est = fs_iptw, nb_boots = 20,
  parallel_args = list(plan = "sequential", workers = n_cores,
                       show_message = FALSE),
  details = FALSE
)
```

## 5.2: IPTW Forest Plot

```{r fp-iptw, fig.width=11, fig.height=7}
#| eval: !expr '!is.null(fs_iptw$sg.harm)'

test_id <- "5.2"
result <- tryCatch({
  fp_iptw <- plot_subgroup_results_forestplot(
    fs_results = list(fs.est = fs_iptw, fs_bc = fs_bc_iptw),
    df_analysis = df,
    outcome.name = "status", event.name = "status", treat.name = "hormon",
    E.name = "Hormon", C.name = "Control",
    title_text = "Test 5.2: IRR + IPTW -- Observational Forest Plot"
  )
  plot(fp_iptw)
  paste0("rows=", nrow(fp_iptw$data))
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "IPTW forest plot renders", "PASS", result)
} else {
  log_test(test_id, "IPTW forest plot renders", "FAIL", result)
}
```


# Test Group 6: Continuous Outcome -- Mean Difference (MD) {#sec-md}

## 6.1: ForestSearch (MD)

```{r fs-md}
fs_md <- forestsearch(
  df.analysis = df_cont, confounders.name = c("age", "bm_pos"),
  outcome.name = "outcome", event.name = "outcome",
  treat.name = "treat", id.name = "id",
  outcome_type = "continuous", effect_measure = "MD", dmin.grf = 0.0,
  hr.threshold = 1.0, hr.consistency = 0.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
  maxk = 2, sg_focus = "hr"
)
cat("MD SG:", paste(fs_md$sg.harm, collapse = " & "), "\n")
```

## 6.2: MD Forest Plot (Identity Scale)

```{r fp-md, fig.width=11, fig.height=7}
#| eval: !expr '!is.null(fs_md$sg.harm)'

test_id <- "6.2"
result <- tryCatch({
  fp_md <- plot_subgroup_results_forestplot(
    fs_results = list(fs.est = fs_md),
    df_analysis = df_cont,
    outcome.name = "outcome", event.name = "outcome", treat.name = "treat",
    E.name = "Treatment", C.name = "Control",
    title_text = "Test 6.2: Continuous Outcome -- MD Forest Plot"
  )
  plot(fp_md)

  ci_col <- grep("^MD", names(fp_md$data), value = TRUE)
  paste0("rows=", nrow(fp_md$data), ", ci_col=", ci_col[1])
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "MD forest plot renders (linear scale, ref=0)", "PASS", result)
} else {
  log_test(test_id, "MD forest plot renders (linear scale, ref=0)", "FAIL", result)
}
```


# Test Group 7: Display Defaults Verification {#sec-defaults}

## 7.1: Log-Scale Measures (OR, RR, IRR) Get xlog=TRUE, ref=1

```{r test-defaults-log}
test_id <- "7.1"
result <- tryCatch({
  # Mock fs result with OR -- df.est needs treat.recommend column
  df_mock <- df
  df_mock$treat.recommend <- as.integer(df$er > 0)
  mock_fs <- list(outcome_type = "binary", effect_measure = "OR",
                  sg.harm = c("{er <= 0}"), df.est = df_mock,
                  args_call_all = list(offset.name = NULL,
                                       treat.name = "hormon"))
  mock_results <- list(fs.est = mock_fs)

  fp_test <- plot_subgroup_results_forestplot(
    fs_results = mock_results,
    df_analysis = df,
    outcome.name = "status", event.name = "status", treat.name = "hormon",
    E.name = "E", C.name = "C",
    show_cv_metrics = FALSE
  )
  ci_cols <- names(fp_test$data)
  has_or <- any(grepl("^OR", ci_cols))
  paste0("ci_label_has_OR=", has_or)
}, error = function(e) e$message)

if (grepl("TRUE", result)) {
  log_test(test_id, "OR auto-detects xlog=TRUE, ci_label='OR (...)'", "PASS", result)
} else {
  log_test(test_id, "OR auto-detects xlog=TRUE, ci_label='OR (...)'", "FAIL", result)
}
```

## 7.2: Identity-Scale Measures (RD, IRD, MD) Get xlog=FALSE, ref=0

```{r test-defaults-identity}
test_id <- "7.2"
result <- tryCatch({
  df_mock <- df
  df_mock$treat.recommend <- as.integer(df$er > 0)
  mock_fs <- list(outcome_type = "binary", effect_measure = "RD",
                  sg.harm = c("{er <= 0}"), df.est = df_mock,
                  args_call_all = list(offset.name = NULL,
                                       treat.name = "hormon"))
  mock_results <- list(fs.est = mock_fs)

  fp_test <- plot_subgroup_results_forestplot(
    fs_results = mock_results,
    df_analysis = df,
    outcome.name = "status", event.name = "status", treat.name = "hormon",
    E.name = "E", C.name = "C",
    show_cv_metrics = FALSE
  )
  ci_cols <- names(fp_test$data)
  has_rd <- any(grepl("^RD", ci_cols))

  # Check formatting: RD should use 3 decimal places
  rd_col <- grep("^RD", ci_cols, value = TRUE)[1]
  sample_val <- fp_test$data[[rd_col]][1]
  has_3dp <- grepl("\\.[0-9]{3}", sample_val)

  paste0("ci_label_has_RD=", has_rd, ", 3dp_format=", has_3dp)
}, error = function(e) e$message)

if (grepl("RD=TRUE.*3dp.*TRUE", result)) {
  log_test(test_id, "RD auto-detects xlog=FALSE, 3dp formatting", "PASS", result)
} else {
  log_test(test_id, "RD auto-detects xlog=FALSE, 3dp formatting", "FAIL", result)
}
```

## 7.3: Survival Falls Through to HR Defaults (Regression)

```{r test-defaults-survival}
test_id <- "7.3"
result <- tryCatch({
  df_mock <- df
  df_mock$treat.recommend <- as.integer(df$er > 0)
  mock_fs <- list(outcome_type = "survival", effect_measure = NULL,
                  sg.harm = c("{er <= 0}"), df.est = df_mock,
                  args_call_all = list(treat.name = "hormon"))
  mock_results <- list(fs.est = mock_fs)

  fp_test <- plot_subgroup_results_forestplot(
    fs_results = mock_results,
    df_analysis = df,
    outcome.name = "rfstime", event.name = "status", treat.name = "hormon",
    E.name = "E", C.name = "C",
    show_cv_metrics = FALSE
  )
  ci_cols <- names(fp_test$data)
  has_hr <- any(grepl("^HR", ci_cols))
  paste0("ci_label_has_HR=", has_hr)
}, error = function(e) e$message)

if (grepl("TRUE", result)) {
  log_test(test_id, "Survival auto-detects HR label (unchanged)", "PASS", result)
} else {
  log_test(test_id, "Survival auto-detects HR label (unchanged)", "FAIL", result)
}
```

## 7.4: User-Specified xlim/ticks_at Override Auto-Defaults

```{r test-user-override}
#| eval: !expr '!is.null(fs_or$sg.harm)'

test_id <- "7.4"
result <- tryCatch({
  fp_custom <- plot_subgroup_results_forestplot(
    fs_results = list(fs.est = fs_or),
    df_analysis = df,
    outcome.name = "status", event.name = "status", treat.name = "hormon",
    E.name = "E", C.name = "C",
    xlim = c(0.10, 10.0),
    ticks_at = c(0.10, 0.50, 1.0, 2.0, 10.0),
    show_cv_metrics = FALSE,
    title_text = "Test 7.4: OR with custom xlim"
  )
  plot(fp_custom)
  "User xlim/ticks_at accepted without error"
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "User xlim/ticks_at overrides auto-defaults", "PASS", result)
} else {
  log_test(test_id, "User xlim/ticks_at overrides auto-defaults", "FAIL", result)
}
```


# Test Group 8: Reference Subgroups with GLM {#sec-ref}

## 8.1: OR Forest Plot with Reference Subgroups

```{r fp-or-refs, fig.width=11, fig.height=9}
#| eval: !expr '!is.null(fs_or$sg.harm)'

test_id <- "8.1"
result <- tryCatch({
  ref_list <- list(
    er_neg = list(
      subset_expr = "er <= 0",
      name = "ER-negative",
      type = "reference"
    ),
    er_pos = list(
      subset_expr = "er > 0",
      name = "ER-positive",
      type = "reference"
    ),
    young = list(
      subset_expr = "age < 50",
      name = "Age < 50",
      type = "reference"
    )
  )

  fp_ref <- plot_subgroup_results_forestplot(
    fs_results = list(fs.est = fs_or, fs_bc = fs_bc_or),
    df_analysis = df,
    subgroup_list = ref_list,
    outcome.name = "status", event.name = "status", treat.name = "hormon",
    E.name = "Hormon", C.name = "Control",
    title_text = "Test 8.1: OR Forest Plot with Reference Subgroups"
  )
  plot(fp_ref)
  paste0("rows=", nrow(fp_ref$data),
         ", ref_rows=", sum(fp_ref$row_types == "reference"))
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "OR forest plot with reference subgroups", "PASS", result)
} else {
  log_test(test_id, "OR forest plot with reference subgroups", "FAIL", result)
}
```


# Test Group 9: Extreme Estimate Trimming {#sec-trimming}

Verify that extreme GLM estimates (e.g., OR > 50 from sparse cells) are
clipped to `xlim` for the CI bars while preserving the actual numeric
value in the text column.

## 9.1: Synthetic Extreme OR -- Trimming Applied

```{r test-trimming, fig.width=11, fig.height=7}
test_id <- "9.1"
result <- tryCatch({
  # Create a dataset where the subgroup has sparse events that will
  # trigger the per-arm reliability check or produce capped CI bounds
  set.seed(99)
  n_ext <- 400
  df_extreme <- data.frame(
    id     = seq_len(n_ext),
    treat  = rep(0:1, n_ext / 2),
    x1     = c(rep(1, 40), rep(0, n_ext - 40))
  )
  # Subgroup {x1=1}: moderate events in both arms (estimable but extreme)
  df_extreme$status <- 0L
  df_extreme$status[df_extreme$x1 == 1 & df_extreme$treat == 1] <-
    rbinom(sum(df_extreme$x1 == 1 & df_extreme$treat == 1), 1, 0.90)
  df_extreme$status[df_extreme$x1 == 1 & df_extreme$treat == 0] <-
    rbinom(sum(df_extreme$x1 == 1 & df_extreme$treat == 0), 1, 0.10)
  # Complement: balanced events
  df_extreme$status[df_extreme$x1 == 0] <-
    rbinom(sum(df_extreme$x1 == 0), 1, 0.30)
  df_extreme$rfstime <- rexp(n_ext, 0.01) + 1
  df_extreme$treat.recommend <- 1 - df_extreme$x1

  mock_fs <- list(
    outcome_type = "binary", effect_measure = "OR",
    sg.harm = c("{x1 == 1}"), df.est = df_extreme,
    args_call_all = list(offset.name = NULL, treat.name = "treat")
  )

  fp_trim <- plot_subgroup_results_forestplot(
    fs_results = list(fs.est = mock_fs),
    df_analysis = df_extreme,
    outcome.name = "status", event.name = "status", treat.name = "treat",
    E.name = "Treat", C.name = "Control",
    xlim = c(0.25, 4.0),
    title_text = "Test 9.1: OR with Extreme Estimates (trimmed/capped)"
  )
  plot(fp_trim)

  # Check for capped bounds (">4.0") or NE in text column
  ci_col <- grep("^OR", names(fp_trim$data), value = TRUE)[1]
  ci_vals <- fp_trim$data[[ci_col]]
  has_cap <- any(grepl(">|NE", ci_vals), na.rm = TRUE)

  # Verify bars are within xlim
  est_vals <- fp_trim$data$est[!is.na(fp_trim$data$est)]
  all_within <- length(est_vals) == 0 ||
    all(est_vals >= 0.25 & est_vals <= 4.0)

  paste0("has_cap_or_NE=", has_cap, ", bars_within_xlim=", all_within)
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "Extreme ORs capped in text, bars within xlim", "PASS", result)
} else {
  log_test(test_id, "Extreme ORs capped in text, bars within xlim", "FAIL", result)
}
```

## 9.2: Normal Range OR -- No Trimming

```{r test-no-trimming}
#| eval: !expr '!is.null(fs_or$sg.harm)'

test_id <- "9.2"
result <- tryCatch({
  # Use the real OR forest plot -- estimates should be within normal
  # range and not trigger capping
  fp_normal <- plot_subgroup_results_forestplot(
    fs_results = list(fs.est = fs_or),
    df_analysis = df,
    outcome.name = "status", event.name = "status", treat.name = "hormon",
    E.name = "E", C.name = "C",
    xlim = c(0.25, 4.0),
    show_cv_metrics = FALSE
  )

  ci_col <- grep("^OR", names(fp_normal$data), value = TRUE)[1]
  has_cap <- any(grepl(">|<|NE", fp_normal$data[[ci_col]]), na.rm = TRUE)

  paste0("no_capping_needed=", !has_cap)
}, error = function(e) e$message)

if (grepl("TRUE", result)) {
  log_test(test_id, "Normal-range ORs not capped", "PASS", result)
} else {
  log_test(test_id, "Normal-range ORs not capped", "FAIL", result)
}
```


# Results Summary {#sec-summary}

```{r summary-table}
n_pass <- sum(test_log$Status == "PASS")
n_fail <- sum(test_log$Status == "FAIL")
n_total <- nrow(test_log)

cat(sprintf("\n=== GLM Forest Plot Test Results: %d/%d PASSED ===\n\n",
            n_pass, n_total))

gt(test_log) |>
  tab_header(
    title = md("**GLM Forest Plot Test Suite Results**"),
    subtitle = sprintf("%d tests: %d passed, %d failed", n_total, n_pass, n_fail)
  ) |>
  tab_style(
    style = cell_fill(color = "#e8f5e9"),
    locations = cells_body(rows = Status == "PASS")
  ) |>
  tab_style(
    style = cell_fill(color = "#ffcdd2"),
    locations = cells_body(rows = Status == "FAIL")
  )
```

```{r final-check}
if (n_fail > 0) {
  cat("\n\u274C FAILURES DETECTED:\n")
  for (i in which(test_log$Status == "FAIL")) {
    cat("  ", test_log$Test[i], ":", test_log$Description[i], "\n")
    cat("    ", test_log$Details[i], "\n")
  }
} else {
  cat("\n\u2705 ALL TESTS PASSED\n")
}
```


# Session Information

```{r session-info}
sessionInfo()
```
