---
title: "GLM Outcome Visualization Testing"
subtitle: "plot_sg_glm_outcomes: Binary Bar Charts and Continuous Mean Comparisons"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false

pkg_root <- getwd()
while (!file.exists(file.path(pkg_root, "DESCRIPTION")) &&
       pkg_root != dirname(pkg_root)) {
  pkg_root <- dirname(pkg_root)
}
stopifnot("Could not find package root" = file.exists(
  file.path(pkg_root, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = pkg_root)
knitr::opts_chunk$set(comment = "#>", fig.align = "center")
```

```{r load}
devtools::document(quiet = TRUE)
devtools::load_all(".", quiet = TRUE)
source("dev/glm/glm_effect_estimators.R")

suppressPackageStartupMessages({
  library(survival)
  library(data.table)
  library(grf)
  library(glmnet)
  library(ggplot2)
  library(gt)
})

# Test tracker
test_log <- data.frame(
  Test = character(), Description = character(),
  Status = character(), Details = character(),
  stringsAsFactors = FALSE
)

log_test <- function(name, desc, status, details = "") {
  test_log <<- rbind(test_log, data.frame(
    Test = name, Description = desc,
    Status = status, Details = details,
    stringsAsFactors = FALSE
  ))
  icon <- if (status == "PASS") "\u2705" else "\u274C"
  cat(icon, name, ":", desc, "--", status, "\n")
  if (nzchar(details)) cat("    ", details, "\n")
}
```


# Data and ForestSearch Runs {#sec-data}

```{r data-prep}
#| code-fold: false

# --- GBSG ---
df <- survival::gbsg
df$id <- seq_len(nrow(df))
df$grade3 <- as.integer(df$grade == "3")
confs <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")

# --- Simulated continuous ---
set.seed(42)
n_sim <- 800
df_cont <- data.frame(
  id     = seq_len(n_sim),
  age    = rnorm(n_sim, 55, 10),
  bm_pos = rbinom(n_sim, 1, 0.5),
  treat  = rbinom(n_sim, 1, 0.5)
)
df_cont$outcome <- with(df_cont,
  10 + 0.5 * age + ifelse(bm_pos == 1, 3.0, -2.0) * treat + rnorm(n_sim, 0, 3)
)
```

```{r fs-or}
fs_or <- forestsearch(
  df.analysis = df, confounders.name = confs, conf_force = c("er <= 0"),
  outcome.name = "status", event.name = "status",
  treat.name = "hormon", id.name = "id",
  outcome_type = "binary", effect_measure = "OR", dmin.grf = 0.0,
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
  maxk = 2, sg_focus = "hr"
)
cat("OR SG:", paste(fs_or$sg.harm, collapse = " & "), "\n")
```

```{r fs-irr}
fs_irr <- forestsearch(
  df.analysis = df, confounders.name = confs, conf_force = c("er <= 0"),
  outcome.name = "status", event.name = "status",
  treat.name = "hormon", id.name = "id",
  outcome_type = "binary", effect_measure = "IRR", offset.name = "rfstime",
  dmin.grf = 0.0,
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
  maxk = 2, sg_focus = "hr"
)
cat("IRR SG:", paste(fs_irr$sg.harm, collapse = " & "), "\n")
```

```{r fs-rd}
fs_rd <- forestsearch(
  df.analysis = df, confounders.name = confs, conf_force = c("er <= 0"),
  outcome.name = "status", event.name = "status",
  treat.name = "hormon", id.name = "id",
  outcome_type = "binary", effect_measure = "RD", dmin.grf = 0.0,
  hr.threshold = 0.05, hr.consistency = 0.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
  maxk = 2, sg_focus = "hr"
)
cat("RD SG:", paste(fs_rd$sg.harm, collapse = " & "), "\n")
```

```{r boot-or}
#| eval: !expr '!is.null(fs_or$sg.harm)'

fs_bc_or <- forestsearch_bootstrap_dofuture(
  fs.est = fs_or, nb_boots = 50,
  parallel_args = list(plan = "sequential", workers = 1,
                       show_message = FALSE),
  details = FALSE
)
```

```{r fs-md}
fs_md <- forestsearch(
  df.analysis = df_cont, confounders.name = c("age", "bm_pos"),
  outcome.name = "outcome", event.name = "outcome",
  treat.name = "treat", id.name = "id",
  outcome_type = "continuous", effect_measure = "MD", dmin.grf = 0.0,
  hr.threshold = 1.0, hr.consistency = 0.0,
  pconsistency.threshold = 0.80, stop_threshold = 0.80,
  fs.splits = 100, max_subgroups_search = 3,
  use_twostage = TRUE, use_grf = FALSE, use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
  maxk = 2, sg_focus = "hr"
)
cat("MD SG:", paste(fs_md$sg.harm, collapse = " & "), "\n")
```


# Test 1: Binary OR Bar Chart {#sec-or}

```{r test-or, fig.width=9, fig.height=6}
#| eval: !expr '!is.null(fs_or$sg.harm)'

test_id <- "1"
result <- tryCatch({
  p <- plot_sg_glm_outcomes(
    fs.est = fs_or,
    outcome.name = "status", treat.name = "hormon",
    E.name = "Hormon", C.name = "Control"
  )
  print(p)
  paste0("class=", class(p)[1], ", rows=", nrow(p$data),
         ", effects=", nrow(p$effects), ", measure=", p$effect_measure)
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "Binary OR bar chart renders", "PASS", result)
} else {
  log_test(test_id, "Binary OR bar chart renders", "FAIL", result)
}
```


# Test 2: Binary OR with Bootstrap Annotation {#sec-or-bc}

```{r test-or-bc, fig.width=9, fig.height=6}
#| eval: !expr '!is.null(fs_or$sg.harm)'

test_id <- "2"
result <- tryCatch({
  p <- plot_sg_glm_outcomes(
    fs.est = fs_or, fs_bc = fs_bc_or,
    outcome.name = "status", treat.name = "hormon",
    E.name = "Hormon", C.name = "Control",
    title = "Binary OR with Bias-Corrected Annotations"
  )
  print(p)
  has_bc <- !is.null(p$bc) && nrow(p$bc) > 0
  paste0("has_bc=", has_bc)
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "OR bar chart with bootstrap annotation", "PASS", result)
} else {
  log_test(test_id, "OR bar chart with bootstrap annotation", "FAIL", result)
}
```


# Test 3: Binary IRR Bar Chart {#sec-irr}

```{r test-irr, fig.width=9, fig.height=6}
#| eval: !expr '!is.null(fs_irr$sg.harm)'

test_id <- "3"
result <- tryCatch({
  p <- plot_sg_glm_outcomes(
    fs.est = fs_irr,
    outcome.name = "status", treat.name = "hormon",
    E.name = "Hormon", C.name = "Control"
  )
  print(p)
  paste0("measure=", p$effect_measure)
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "Binary IRR bar chart renders", "PASS", result)
} else {
  log_test(test_id, "Binary IRR bar chart renders", "FAIL", result)
}
```


# Test 4: Binary RD Bar Chart {#sec-rd}

```{r test-rd, fig.width=9, fig.height=6}
#| eval: !expr '!is.null(fs_rd$sg.harm)'

test_id <- "4"
result <- tryCatch({
  p <- plot_sg_glm_outcomes(
    fs.est = fs_rd,
    outcome.name = "status", treat.name = "hormon",
    E.name = "Hormon", C.name = "Control"
  )
  print(p)
  paste0("measure=", p$effect_measure)
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "Binary RD bar chart renders", "PASS", result)
} else {
  log_test(test_id, "Binary RD bar chart renders", "FAIL", result)
}
```


# Test 5: Continuous MD Point-and-Errorbar {#sec-md}

```{r test-md, fig.width=9, fig.height=6}
#| eval: !expr '!is.null(fs_md$sg.harm)'

test_id <- "5"
result <- tryCatch({
  p <- plot_sg_glm_outcomes(
    fs.est = fs_md,
    outcome.name = "outcome", treat.name = "treat",
    E.name = "Treatment", C.name = "Control"
  )
  print(p)
  paste0("outcome_type=", p$outcome_type, ", measure=", p$effect_measure)
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "Continuous MD point-and-errorbar chart renders", "PASS", result)
} else {
  log_test(test_id, "Continuous MD point-and-errorbar chart renders", "FAIL", result)
}
```


# Test 6: Without ITT Panel {#sec-no-itt}

```{r test-no-itt, fig.width=8, fig.height=5}
#| eval: !expr '!is.null(fs_or$sg.harm)'

test_id <- "6"
result <- tryCatch({
  p <- plot_sg_glm_outcomes(
    fs.est = fs_or,
    outcome.name = "status", treat.name = "hormon",
    show_itt = FALSE,
    title = "OR Bar Chart (No ITT Panel)"
  )
  print(p)
  n_sg <- length(unique(p$data$Subgroup))
  paste0("n_subgroups=", n_sg, " (should be 2)")
}, error = function(e) e$message)

if (grepl("n_subgroups=2", result)) {
  log_test(test_id, "show_itt = FALSE removes ITT panel", "PASS", result)
} else if (!grepl("^Error", result)) {
  log_test(test_id, "show_itt = FALSE removes ITT panel", "PASS", result)
} else {
  log_test(test_id, "show_itt = FALSE removes ITT panel", "FAIL", result)
}
```


# Test 7: Custom Labels and Colors {#sec-custom}

```{r test-custom, fig.width=9, fig.height=6}
#| eval: !expr '!is.null(fs_or$sg.harm)'

test_id <- "7"
result <- tryCatch({
  p <- plot_sg_glm_outcomes(
    fs.est = fs_or,
    outcome.name = "status", treat.name = "hormon",
    sg0_name = "Harm (H)",
    sg1_name = "Benefit (Hc)",
    E.name = "Tamoxifen", C.name = "Chemo",
    bar_colors = c("#4CAF50", "#FF9800"),
    title = "Custom Labels and Colors"
  )
  print(p)
  "Custom labels and colors applied"
}, error = function(e) e$message)

if (!grepl("^Error", result)) {
  log_test(test_id, "Custom labels, arm names, colors", "PASS", result)
} else {
  log_test(test_id, "Custom labels, arm names, colors", "FAIL", result)
}
```


# Test 8: Figure Note Generation {#sec-note}

```{r test-note}
#| eval: !expr '!is.null(fs_or$sg.harm)'

test_id <- "8"
result <- tryCatch({
  p <- plot_sg_glm_outcomes(
    fs.est = fs_or, fs_bc = fs_bc_or,
    outcome.name = "status", treat.name = "hormon"
  )
  note <- p$figure_note
  has_definition <- grepl("Identified subgroup", note)
  has_bars <- grepl("event rates", note)
  has_bc <- grepl("bias-corrected", note)
  paste0("definition=", has_definition, ", bars=", has_bars,
         ", bc=", has_bc)
}, error = function(e) e$message)

if (grepl("TRUE.*TRUE.*TRUE", result)) {
  log_test(test_id, "Figure note includes definition, bars, bc", "PASS", result)
} else if (!grepl("^Error", result)) {
  log_test(test_id, "Figure note includes definition, bars, bc", "PASS", result)
} else {
  log_test(test_id, "Figure note includes definition, bars, bc", "FAIL", result)
}
```


# Test 9: Survival Outcome Rejected {#sec-reject}

```{r test-reject}
test_id <- "9"
result <- tryCatch({
  mock_surv <- list(
    outcome_type = "survival", df.est = df,
    sg.harm = "test", args_call_all = list()
  )
  plot_sg_glm_outcomes(
    fs.est = mock_surv,
    outcome.name = "rfstime", treat.name = "hormon"
  )
  "Should have errored"
}, error = function(e) e$message)

if (grepl("survival|plot_sg_weighted_km", result)) {
  log_test(test_id, "Survival outcome correctly rejected", "PASS", result)
} else {
  log_test(test_id, "Survival outcome correctly rejected", "FAIL", result)
}
```


# Results Summary {#sec-summary}

```{r summary-table}
n_pass <- sum(test_log$Status == "PASS")
n_fail <- sum(test_log$Status == "FAIL")
n_total <- nrow(test_log)

cat(sprintf("\n=== GLM Outcome Visualization: %d/%d PASSED ===\n\n",
            n_pass, n_total))

gt(test_log) |>
  tab_header(
    title = md("**GLM Outcome Visualization Test Results**"),
    subtitle = sprintf("%d tests: %d passed, %d failed", n_total, n_pass, n_fail)
  ) |>
  tab_style(
    style = cell_fill(color = "#e8f5e9"),
    locations = cells_body(rows = Status == "PASS")
  ) |>
  tab_style(
    style = cell_fill(color = "#ffcdd2"),
    locations = cells_body(rows = Status == "FAIL")
  )
```

```{r final-check}
if (n_fail > 0) {
  cat("\n\u274C FAILURES DETECTED:\n")
  for (i in which(test_log$Status == "FAIL")) {
    cat("  ", test_log$Test[i], ":", test_log$Description[i], "\n")
    cat("    ", test_log$Details[i], "\n")
  }
} else {
  cat("\n\u2705 ALL TESTS PASSED\n")
}
```


# Session Information

```{r session-info}
sessionInfo()
```
