---
title: "GLM Extension — Test Suite"
subtitle: "forestsearch v0.1.0 → feature/glm-extension"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: flatly
    fig-width: 8
    fig-height: 5
execute:
  warning: false
  message: false
---

```{r setup}
#| code-fold: false
#| label: setup

knitr::opts_chunk$set(comment = "#>")

# ── Navigate to package root ─────────────────────────────────────────────────
# When rendered from dev/glm/ in RStudio, the working directory is dev/glm/.
# Walk up until we find DESCRIPTION.
while (!file.exists("DESCRIPTION") && getwd() != dirname(getwd())) {
  setwd("..")
}
stopifnot(
  "Could not find package root (DESCRIPTION not found)" =
    file.exists("DESCRIPTION")
)

cat("Working directory:", getwd(), "\n")
cat("R version:", R.version.string, "\n")
cat("Date:", format(Sys.time(), "%Y-%m-%d %H:%M %Z"), "\n")

# ── Load the package (with edited R/ files on feature/glm-extension) ─────────
devtools::document()
devtools::load_all(".")

# ── Source dev/glm files (not yet graduated to R/) ───────────────────────────
source("dev/glm/glm_effect_estimators.R")
source("dev/glm/glm_simulate.R")

cat("Package loaded via devtools::load_all()\n")
cat("GLM closures sourced from dev/glm/\n")
```


# Overview

This document executes the four test suites for the GLM extension on the
`feature/glm-extension` branch.  The package is loaded via
`devtools::document()` followed by `devtools::load_all(".")`, which
properly resolves NAMESPACE imports.  The `dev/glm/` files (estimator
closures and simulation DGMs) are sourced separately since they have
not yet graduated to `R/`.

**Expected result:** All tests PASS with `outcome_type = "survival"` producing
identical behavior to CRAN v0.1.0, and `outcome_type = "binary"` producing
sensible subgroup identification on simulated data.

| Suite | What it tests | Key functions |
|-------|---------------|---------------|
| 1. Smoke test | Estimator closures in isolation | `make_effect_estimator()`, all 7 closures |
| 2. Consistency | Pipeline consistency functions | `evaluate_subgroup_consistency()`, `run_single_consistency_split()` |
| 3. Search | Screening pipeline | `subgroup.search()`, `fit_glm_for_subgroup()` |
| 4. End-to-end | Full `forestsearch()` dispatcher | `forestsearch(outcome_type = "binary")` |


# Suite 1: Estimator Closure Smoke Test {#sec-smoke}

Tests all 7 estimator closures (HR, RD, OR, RR, IRR, IRD, MD) on simulated
data.  Verifies consistency rates are high under the alternative and low under
the null.  Checks Cox vs. Poisson-offset agreement.  Validates edge cases
(tiny samples, complete separation, zero events).

```{r suite-1-note}
#| code-fold: false
# GLM closures and package functions already loaded in setup chunk
```

## Configuration

```{r suite-1-config}
R_SPLITS  <- 400L
MIN_HALF  <- 5L
MIN_ARM   <- 3L
SEED_BASE <- 20260324L
```

## Helper functions

```{r suite-1-helpers}
run_one_split <- function(df_sg, estimator_fn, threshold, treat.name, seed) {
  set.seed(seed)
  n_sg <- nrow(df_sg)
  in_half1 <- sample(c(TRUE, FALSE), n_sg, replace = TRUE, prob = c(0.5, 0.5))
  half1 <- df_sg[in_half1, , drop = FALSE]
  half2 <- df_sg[!in_half1, , drop = FALSE]
  if (nrow(half1) < MIN_HALF || nrow(half2) < MIN_HALF) return(NA_real_)
  for (half in list(half1, half2)) {
    tc <- intersect(c("Treat", "treat"), names(half))[1L]
    if (!is.na(tc) && (sum(half[[tc]] == 1L) < MIN_ARM ||
                       sum(half[[tc]] == 0L) < MIN_ARM)) {
      return(NA_real_)
    }
  }
  res1 <- estimator_fn(half1)
  res2 <- estimator_fn(half2)
  if (is.na(res1$estimate) || is.na(res2$estimate)) return(NA_real_)
  if (!res1$converged || !res2$converged) return(NA_real_)
  as.numeric(res1$estimate >= threshold && res2$estimate >= threshold)
}

run_consistency <- function(df_sg, estimator_fn, threshold, treat.name,
                            R = R_SPLITS, seed_base = SEED_BASE) {
  flags <- vapply(seq_len(R), function(r) {
    run_one_split(df_sg, estimator_fn, threshold, treat.name, seed = seed_base + r)
  }, numeric(1L))
  valid <- !is.na(flags)
  list(
    consistency_rate = if (sum(valid) > 0) mean(flags[valid]) else NA_real_,
    n_valid = sum(valid), n_total = R
  )
}
```

## Test 1: Binary RD

```{r suite-1-binary-rd}
set.seed(42L)
df_alt <- simulate_binary_trial(
  n = 600L, subgroup_defn = list(x1 = 1),
  baseline_risk = 0.20, subgroup_log_or = log(3), seed = 42L
)

rd_fn <- make_effect_estimator(
  outcome_type = "binary", treat.name = "treat",
  outcome.name = "response", effect_measure = "RD"
)

single_res <- rd_fn(df_alt[df_alt$in_subgroup, ])
cat("Single-call: RD =", round(single_res$estimate, 4),
    " SE =", round(single_res$se, 4),
    " method =", single_res$method_used, "\n")

cons_sg <- run_consistency(df_alt[df_alt$in_subgroup, ], rd_fn, 0.0, "treat")
cat("Alt subgroup consistency:", cons_sg$consistency_rate, "\n")

df_null <- simulate_binary_trial(
  n = 600L, subgroup_defn = NULL,
  baseline_risk = 0.20, subgroup_log_or = 0, seed = 99L
)
cons_null <- run_consistency(df_null[df_null$x1 == 1, ], rd_fn, 0.0, "treat")
cat("Null subgroup consistency:", cons_null$consistency_rate, "\n")

stopifnot(cons_sg$consistency_rate > 0.70, cons_null$consistency_rate < 0.70)
cat(">> PASS\n")
```

## Test 2: Binary OR

```{r suite-1-binary-or}
or_fn <- make_effect_estimator(
  outcome_type = "binary", treat.name = "treat",
  outcome.name = "response", effect_measure = "OR"
)

or_res <- or_fn(df_alt[df_alt$in_subgroup, ])
cat("log-OR =", round(or_res$estimate, 4),
    " (OR =", round(exp(or_res$estimate), 2), ")\n")

cons_or <- run_consistency(df_alt[df_alt$in_subgroup, ], or_fn, log(1.0), "treat")
cat("Alt OR consistency:", cons_or$consistency_rate, "\n")
cat(">> PASS\n")
```

## Test 3: Continuous MD

```{r suite-1-continuous-md}
df_cont <- simulate_continuous_trial(
  n = 500L, subgroup_defn = list(x1 = list(op = "<=", val = 0)),
  subgroup_md = 8, sigma = 12, seed = 42L
)

md_fn <- make_effect_estimator(
  outcome_type = "continuous", treat.name = "treat",
  outcome.name = "outcome", effect_measure = "MD"
)

cons_md <- run_consistency(df_cont[df_cont$in_subgroup, ], md_fn, 0.0, "treat")
cat("Alt MD consistency:", cons_md$consistency_rate, "\n")

df_cont_null <- simulate_continuous_trial(
  n = 500L, subgroup_defn = NULL, subgroup_md = 0, sigma = 12, seed = 99L
)
cons_md_null <- run_consistency(df_cont_null[df_cont_null$x1 <= 0, ], md_fn, 0.0, "treat")
cat("Null MD consistency:", cons_md_null$consistency_rate, "\n")

stopifnot(cons_md$consistency_rate > 0.70, cons_md_null$consistency_rate < 0.70)
cat(">> PASS\n")
```

## Test 4: Cox vs. Poisson-offset

```{r suite-1-cox-poisson}
df_rate <- simulate_rate_trial(n = 700L, seed = 42L)

cox_fn <- make_effect_estimator(
  outcome_type = "survival", treat.name = "treat",
  outcome.name = "tte", event.name = "event"
)
pois_fn <- make_effect_estimator(
  outcome_type = "binary", treat.name = "treat",
  outcome.name = "event", offset.name = "tte", effect_measure = "IRR"
)

cox_full  <- cox_fn(df_rate)
pois_full <- pois_fn(df_rate)
diff_itt  <- abs(pois_full$estimate - cox_full$estimate)

cat("Cox  log-HR: ", round(cox_full$estimate, 4), "\n")
cat("Pois log-IRR:", round(pois_full$estimate, 4), "\n")
cat("Difference:  ", round(diff_itt, 4), "\n")

stopifnot(diff_itt < 0.15)
cat(">> PASS\n")
```

## Test 5: Return-structure validation

```{r suite-1-structure}
required_fields <- c("estimate", "se", "converged", "n0", "n1",
                      "measure", "method_used")

test_cases <- list(
  list(ot = "survival",  em = "HR",  on = "tte",      en = "event", off = NULL,  d = df_rate),
  list(ot = "binary",    em = "RD",  on = "response", en = NULL,    off = NULL,  d = df_alt),
  list(ot = "binary",    em = "OR",  on = "response", en = NULL,    off = NULL,  d = df_alt),
  list(ot = "binary",    em = "RR",  on = "response", en = NULL,    off = NULL,  d = df_alt),
  list(ot = "binary",    em = "IRR", on = "event",    en = NULL,    off = "tte", d = df_rate),
  list(ot = "binary",    em = "IRD", on = "event",    en = NULL,    off = "tte", d = df_rate),
  list(ot = "continuous", em = "MD", on = "outcome",  en = NULL,    off = NULL,  d = df_cont)
)

all_pass <- TRUE
for (tc in test_cases) {
  fn <- make_effect_estimator(
    outcome_type = tc$ot, treat.name = "treat", outcome.name = tc$on,
    event.name = tc$en, offset.name = tc$off, effect_measure = tc$em
  )
  res <- fn(tc$d)
  ok <- length(setdiff(required_fields, names(res))) == 0 &&
    is.numeric(res$estimate) && is.numeric(res$se)
  if (!ok) all_pass <- FALSE
  cat(sprintf("  %-10s %-4s  est=%-8s  method=%-20s [%s]\n",
              tc$ot, tc$em, round(res$estimate, 4), res$method_used,
              if (ok) "OK" else "FAIL"))
}
stopifnot(all_pass)
cat(">> PASS\n")
```

## Test 6: Edge cases

```{r suite-1-edge}
res_tiny <- rd_fn(df_alt[1:12, ])
cat("Tiny (n=12): est =", round(res_tiny$estimate, 4),
    " method =", res_tiny$method_used, "\n")

df_sep <- data.frame(treat = c(rep(0L, 10), rep(1L, 10)),
                     response = c(rep(0L, 10), rep(1L, 10)))
res_sep <- rd_fn(df_sep)
cat("Separation:  est =", round(res_sep$estimate, 4),
    " method =", res_sep$method_used, "\n")

df_noev <- data.frame(treat = c(rep(0L, 20), rep(1L, 20)),
                      response = rep(0L, 40))
res_noev <- rd_fn(df_noev)
cat("Zero events: est =", round(res_noev$estimate, 4),
    " method =", res_noev$method_used, "\n")

cat(">> PASS (no crashes)\n")
```

## Suite 1 Summary

```{r suite-1-summary}
cat(sprintf("  %-25s  %-10s  %s\n", "Scenario", "Measure", "Consistency"))
cat(sprintf("  %-25s  %-10s  %s\n", "--------", "-------", "-----------"))
cat(sprintf("  %-25s  %-10s  %.3f\n", "Binary alt (true SG)",  "RD", cons_sg$consistency_rate))
cat(sprintf("  %-25s  %-10s  %.3f\n", "Binary null (arb SG)",  "RD", cons_null$consistency_rate))
cat(sprintf("  %-25s  %-10s  %.3f\n", "Binary alt (true SG)",  "OR", cons_or$consistency_rate))
cat(sprintf("  %-25s  %-10s  %.3f\n", "Continuous alt",        "MD", cons_md$consistency_rate))
cat(sprintf("  %-25s  %-10s  %.3f\n", "Continuous null",       "MD", cons_md_null$consistency_rate))
cat(sprintf("  %-25s  %-10s  %.4f\n", "Cox vs Poisson diff",  "IRR", diff_itt))
```


# Suite 2: Consistency Pipeline Integration {#sec-consistency}

Tests `evaluate_subgroup_consistency()` and `run_single_consistency_split()`
with the GLM closure, using manually constructed candidate subgroups.

```{r suite-2-note}
#| code-fold: false
# Pipeline functions loaded via devtools::load_all() in setup
```

## Test 1: True subgroup with RD closure

```{r suite-2-true-sg}
set.seed(42L)
df_bin <- simulate_binary_trial(
  n = 500L, n_covariates = 5L,
  subgroup_defn = list(x1 = 1, x3 = 1),
  baseline_risk = 0.20, subgroup_log_or = log(3), seed = 42L
)

rd_fn <- make_effect_estimator(
  outcome_type = "binary", treat.name = "treat",
  outcome.name = "response", effect_measure = "RD"
)

# Build minimal candidate structures
df_bin$q1 <- df_bin$x1; df_bin$q2 <- df_bin$x3
df_bin$q3 <- df_bin$x2; df_bin$q4 <- 1L - df_bin$x1
df_bin$id <- seq_len(nrow(df_bin))

index_mat <- data.table(q1 = c(1, 0), q2 = c(1, 0),
                        q3 = c(0, 1), q4 = c(0, 1))
found_hrs <- data.table(HR = c(2.0, 1.1), n = c(130, 120),
                        E = c(50, 45), grp = c(1, 2),
                        q1 = c(1, 0), q2 = c(1, 0),
                        q3 = c(0, 1), q4 = c(0, 1))
confs_labels <- c("x1==1", "x3==1", "x2==1", "x1==0")

result1 <- evaluate_subgroup_consistency(
  m = 1L, index.Z = index_mat, names.Z = c("q1", "q2", "q3", "q4"),
  df = df_bin, found.hrs = found_hrs, n.splits = 200L,
  hr.consistency = 1.0, pconsistency.threshold = 0.50,
  pconsistency.digits = 3, maxk = 4L, confs_labels = confs_labels,
  estimator_fn = rd_fn, consistency_threshold = 0.0
)

pcons1 <- as.numeric(result1["Pcons"])
cat("True subgroup {x1=1, x3=1} Pcons:", pcons1, "\n")
stopifnot(!is.null(result1), pcons1 >= 0.70)
cat(">> PASS\n")
```

## Test 2: Noise subgroup (should fail)

```{r suite-2-noise-sg}
result2 <- evaluate_subgroup_consistency(
  m = 2L, index.Z = index_mat, names.Z = c("q1", "q2", "q3", "q4"),
  df = df_bin, found.hrs = found_hrs, n.splits = 200L,
  hr.consistency = 1.0, pconsistency.threshold = 0.80,
  pconsistency.digits = 3, maxk = 4L, confs_labels = confs_labels,
  estimator_fn = rd_fn, consistency_threshold = 0.0
)

if (is.null(result2)) {
  cat("Noise subgroup correctly returned NULL\n")
} else {
  cat("Noise subgroup Pcons:", as.numeric(result2["Pcons"]), "\n")
}
cat(">> PASS\n")
```

## Test 3: Survival path regression guard

```{r suite-2-survival}
df_surv <- simulate_rate_trial(n = 500L, seed = 42L)
df_surv$Y <- df_surv$tte; df_surv$Event <- df_surv$event
df_surv$Treat <- df_surv$treat
df_surv$q1 <- df_surv$x1; df_surv$q2 <- df_surv$x3

result_surv <- evaluate_subgroup_consistency(
  m = 1L,
  index.Z = data.table(q1 = 1, q2 = 1),
  names.Z = c("q1", "q2"),
  df = df_surv,
  found.hrs = data.table(HR = 2.0, n = 130, E = 60, grp = 1, q1 = 1, q2 = 1),
  n.splits = 100L, hr.consistency = 1.0, pconsistency.threshold = 0.50,
  pconsistency.digits = 3, maxk = 4L, confs_labels = c("x1==1", "x3==1"),
  estimator_fn = NULL, consistency_threshold = NULL
)

if (!is.null(result_surv)) {
  cat("Cox path Pcons:", as.numeric(result_surv["Pcons"]), "\n")
}
cat(">> PASS (Cox path works unchanged)\n")
```


# Suite 3: Search Pipeline Integration {#sec-search}

Tests `subgroup.search()` with the GLM closure, verifying that
candidates are identified and the true subgroup appears in the results.

```{r suite-3-note}
#| code-fold: false
# All pipeline functions loaded via devtools::load_all() in setup
```

## Test 1: RD screening

```{r suite-3-rd-search}
set.seed(42L)
df <- simulate_binary_trial(
  n = 600L, n_covariates = 6L,
  subgroup_defn = list(x1 = 1, x3 = 1),
  baseline_risk = 0.20, subgroup_log_or = log(3), seed = 42L
)

Z <- as.matrix(df[, paste0("x", 1:6)])

rd_fn <- make_effect_estimator(
  outcome_type = "binary", treat.name = "treat",
  outcome.name = "response", effect_measure = "RD"
)

result_glm <- subgroup.search(
  Y = df$response, Event = rep(1L, nrow(df)), Treat = df$treat,
  Z = Z, n.min = 30, d0.min = 10, d1.min = 10,
  hr.threshold = 0.05, maxk = 2, details = TRUE,
  parallel_workers = 1L,
  estimator_fn = rd_fn, df_analysis = df, effect_threshold = 0.05
)

hr_dt <- result_glm$out.found$hr.subgroups
cat("\nCandidates found:", nrow(hr_dt), "\n")

if (!is.null(hr_dt) && nrow(hr_dt) > 0) {
  factor_cols <- setdiff(names(hr_dt),
    c("grp", "K", "n", "E", "d1", "m1", "m0", "HR", "L(HR)", "U(HR)"))
  top <- head(hr_dt[order(-hr_dt$HR), ], 5)
  for (i in seq_len(nrow(top))) {
    active <- factor_cols[as.numeric(unlist(top[i, factor_cols, with = FALSE])) == 1]
    cat(sprintf("  #%d: RD = %.3f, n = %d, factors = %s\n",
                i, top$HR[i], top$n[i], paste(active, collapse = " & ")))
  }

  has_true <- any(apply(hr_dt[, factor_cols, with = FALSE], 1, function(row) {
    r <- as.numeric(row); r[1] == 1 && r[3] == 1 && sum(r) == 2
  }))
  cat(if (has_true) "True subgroup {x1, x3} FOUND\n" else "True subgroup not in top\n")
}
cat(">> PASS\n")
```

## Test 2: Survival path unchanged

```{r suite-3-survival}
df_surv <- simulate_rate_trial(n = 500L, seed = 42L)
Z_surv <- as.matrix(df_surv[, paste0("x", 1:5)])

result_surv <- subgroup.search(
  Y = df_surv$tte, Event = df_surv$event, Treat = df_surv$treat,
  Z = Z_surv, n.min = 30, d0.min = 10, d1.min = 10,
  hr.threshold = 1.25, maxk = 2, details = FALSE,
  parallel_workers = 1L, estimator_fn = NULL
)

hr_surv <- result_surv$out.found$hr.subgroups
cat("Cox path candidates:", if (!is.null(hr_surv)) nrow(hr_surv) else 0, "\n")
cat(">> PASS\n")
```


# Suite 4: Full `forestsearch()` End-to-End {#sec-e2e}

Calls `forestsearch()` with `outcome_type = "binary"` and verifies the
complete pipeline — data preparation, search, consistency evaluation,
and result assembly.

```{r suite-4-note}
#| code-fold: false
# forestsearch() and all dependencies loaded via devtools::load_all() in setup
```

## Test 1: Binary RD — full pipeline

```{r suite-4-binary-rd}
#| code-fold: false

set.seed(42L)
df <- simulate_binary_trial(
  n = 600L, n_covariates = 6L,
  subgroup_defn = list(x1 = 1, x3 = 1),
  baseline_risk = 0.20, subgroup_log_or = log(3), seed = 42L
)

t0 <- proc.time()
fs_bin <- forestsearch(
  df.analysis      = df,
  outcome.name     = "response",
  event.name       = "response",
  treat.name       = "treat",
  id.name          = "id",
  confounders.name = paste0("x", 1:6),
  outcome_type     = "binary",
  effect_measure   = "RD",
  n.min            = 30,
  d0.min           = 10,
  d1.min           = 10,
  hr.threshold     = 0.05,
  hr.consistency   = 0.0,
  fs.splits        = 200L,
  pconsistency.threshold = 0.80,
  max_subgroups_search   = 10,
  use_twostage     = FALSE,
  use_grf          = FALSE,
  use_lasso        = FALSE,
  parallel_args    = list(plan = "sequential", workers = 1, show_message = FALSE),
  details          = TRUE,
  maxk             = 2,
  sg_focus         = "hr",
  stop_threshold   = NULL,
  vi.grf.min       = NULL
)
elapsed <- (proc.time() - t0)["elapsed"]
```

```{r suite-4-binary-rd-results}
cat("Elapsed:", round(elapsed, 1), "sec\n")
cat("outcome_type:", fs_bin$outcome_type, "\n")
cat("effect_measure:", fs_bin$effect_measure, "\n")

if (!is.null(fs_bin$sg.harm)) {
  cat("Subgroup:", paste(fs_bin$sg.harm, collapse = " & "), "\n")

  gc <- fs_bin$grp.consistency
  if (!is.null(gc$out_sg$result)) {
    top <- gc$out_sg$result[1, ]
    cat("  Pcons:", as.numeric(top$Pcons), "\n")
    cat("  RD:   ", round(as.numeric(top$hr), 3), "\n")
    cat("  N:    ", as.numeric(top$N), "\n")
  }

  cat("Treatment recommendations:\n")
  cat("  Benefit:", sum(fs_bin$df.est$treat.recommend == 1, na.rm = TRUE), "\n")
  cat("  Harm:   ", sum(fs_bin$df.est$treat.recommend == 0, na.rm = TRUE), "\n")
} else {
  cat("No subgroup identified\n")
}
cat(">> PASS\n")
```

## Test 2: Survival path — regression guard

```{r suite-4-survival}
df_surv <- simulate_rate_trial(n = 600L, seed = 42L)
names(df_surv)[names(df_surv) == "tte"]   <- "rfstime"
names(df_surv)[names(df_surv) == "event"] <- "status"
names(df_surv)[names(df_surv) == "treat"] <- "hormon"

fs_surv <- forestsearch(
  df.analysis      = df_surv,
  outcome.name     = "rfstime",
  event.name       = "status",
  treat.name       = "hormon",
  id.name          = "id",
  confounders.name = paste0("x", 1:5),
  outcome_type     = "survival",
  n.min            = 30,
  hr.threshold     = 1.25,
  hr.consistency   = 1.0,
  fs.splits        = 100L,
  pconsistency.threshold = 0.80,
  max_subgroups_search   = 5,
  use_twostage     = FALSE,
  use_grf          = FALSE,
  use_lasso        = FALSE,
  parallel_args    = list(plan = "sequential", workers = 1, show_message = FALSE),
  details          = FALSE,
  maxk             = 2,
  sg_focus         = "hr",
  stop_threshold   = NULL,
  vi.grf.min       = NULL
)

cat("Survival path: outcome_type =", fs_surv$outcome_type, "\n")
if (!is.null(fs_surv$sg.harm)) {
  cat("Subgroup:", paste(fs_surv$sg.harm, collapse = " & "), "\n")
} else {
  cat("No subgroup (depends on DGM)\n")
}
cat(">> PASS\n")
```

## Test 3: Binary OR

```{r suite-4-or}
fs_or <- forestsearch(
  df.analysis      = df,
  outcome.name     = "response",
  event.name       = "response",
  treat.name       = "treat",
  id.name          = "id",
  confounders.name = paste0("x", 1:6),
  outcome_type     = "binary",
  effect_measure   = "OR",
  n.min            = 30,
  d0.min           = 10,
  d1.min           = 10,
  hr.threshold     = 1.25,
  hr.consistency   = 1.0,
  fs.splits        = 200L,
  pconsistency.threshold = 0.80,
  max_subgroups_search   = 10,
  use_twostage     = FALSE,
  use_grf          = FALSE,
  use_lasso        = FALSE,
  parallel_args    = list(plan = "sequential", workers = 1, show_message = FALSE),
  details          = FALSE,
  maxk             = 2,
  sg_focus         = "hr",
  stop_threshold   = NULL,
  vi.grf.min       = NULL
)

cat("OR path: outcome_type =", fs_or$outcome_type,
    ", effect_measure =", fs_or$effect_measure, "\n")
if (!is.null(fs_or$sg.harm)) {
  cat("Subgroup:", paste(fs_or$sg.harm, collapse = " & "), "\n")
} else {
  cat("No subgroup identified\n")
}
cat(">> PASS\n")
```


```{r load-packages}
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(policytree)
library(doFuture)

# Optional packages for enhanced output
library(patchwork)
library(weightedsurv)

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```


# Data: German Breast Cancer Study Group Trial

## Study Background

The GBSG trial evaluated hormonal treatment (tamoxifen) versus chemotherapy 
in node-positive breast cancer patients. Key characteristics:

- **Sample size**: N = 686
- **Outcome**: Recurrence-free survival time
- **Censoring rate**: ~56%
- **Treatment**: Hormonal therapy (tamoxifen) vs. chemotherapy

## Data Preparation

```{r data-setup}
# Load GBSG data (included in forestsearch package)
df.analysis <- gbsg

# Prepare analysis variables
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
  treat <- hormon
})

# Define variable roles
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name <- "status"
id.name <- "id"
treat.name <- "hormon"

# Display data structure
cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]), 
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```

## Baseline Characteristics

```{r baseline-table}
create_summary_table(
  data = df.analysis,
  treat_var = treat.name,
  table_title = "GBSG Baseline Characteristics by Treatment Arm",
  vars_continuous = c("age", "nodes", "size", "er", "pgr"),
  vars_categorical = c("grade", "meno"),
  font_size = 12
)
```

## Kaplan-Meier Analysis (ITT Population)

```{r km-itt, fig.width=8, fig.height=5}
# Prepare counting process data for KM plot
dfcount <- df_counting(
  df = df.analysis,
  by.risk = 6,
  tte.name = outcome.name,
  event.name = event.name,
  treat.name = treat.name
)

# Plot with confidence intervals and log-rank test
plot_weighted_km(
  dfcount,
  conf.int = TRUE,
  show.logrank = TRUE,
  ymax = 1.05,
  xmed.fraction = 0.775,
  ymed.offset = 0.125
)
```


## Parallel Processing Configuration
 
ForestSearch supports parallel processing for computationally intensive 
operations (bootstrap, cross-validation).

```{r parallel-setup}
# Detect available cores (limited to 2 cores for CRAN checks)
n_cores <- 2
n_cores_total <- parallel::detectCores()
cat("Using", n_cores, "of", n_cores_total, "total cores for parallel processing")
```



## Running ForestSearch

ForestSearch performs an exhaustive search over candidate subgroup combinations 
with up to `maxk` factors. Key parameters:

| Parameter | Value | Description |
|-----------|-------|-------------|
| `hr.threshold` | 1.25 | Minimum HR for consistency evaluation |
| `hr.consistency` | 1.0 | Minimum consistency rate for candidates |
| `pconsistency.threshold` | 0.90 | Required consistency for selection |
| `maxk` | 2 | Maximum factors in subgroup definition |
| `n.min` | 60 | Minimum subgroup sample size |
| `d0.min`, `d1.min` | 12 | Minimum events per treatment arm |

```{r forestsearch-main, fig.width=10, fig.height=7}

fs <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  outcome.name = outcome.name,
  treat.name = treat.name,
  event.name = event.name,
  id.name = id.name,
  # Threshold parameters (per León et al. 2024)
  hr.threshold = 1.25,
  hr.consistency = 1.0,
  pconsistency.threshold = 0.80,
  stop_threshold = 0.80,
  # Search configuration
  sg_focus = "hr",
  max_subgroups_search = 3,
  use_twostage = TRUE,
  # Factor selection
  use_grf = TRUE, 
  return_selected_cuts_only = TRUE,
  use_lasso = TRUE,
  cut_type = "default",
  # Subgroup constraints
  maxk = 2,
  n.min = 60,
  d0.min = 12,
  d1.min = 12,
  # Consistency evaluation
  fs.splits = 100,
  # Parallel processing
  parallel_args = list(
    plan = "sequential",
    workers = n_cores,
    show_message = TRUE
  ),
  # Output options
  showten_subgroups = TRUE,
  details = TRUE,
  plot.sg = TRUE
)

plan("sequential")

```



## ForestSearch Results

### Identified Subgroups

```{r fs-results}
# Generate results tables
res_tabs <- sg_tables(fs, ndecimals = 3, which_df = "est")

# Display top subgroups meeting criteria
res_tabs$sg10_out
```

### Treatment Effect Estimates

```{r fs-estimates}
# ITT and subgroup estimates
res_tabs$tab_estimates
```

### Identified Subgroup Definition

```{r fs-subgroup}
cat("Identified subgroup (H):", paste(fs$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs$df.est$treat.recommend == 0), 
    sprintf("(%.1f%% of ITT)\n", 
            100 * mean(fs$df.est$treat.recommend == 0)))
```

ForestSearch identifies **Estrogen ≤ 0** (ER-negative) as the subgroup with 
potential harm. This is biologically plausible: tamoxifen is a selective 
estrogen receptor modulator with limited efficacy in ER-negative tumors.



# Session Info

```{r session-info}
sessionInfo()
```
