---
title: "GRF `dmin.grf` Calibration Study"
subtitle: "Operating Characteristics: RMST vs. Risk Difference"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: flatly
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| code-fold: false

knitr::opts_chunk$set(comment = "#>")

while (!file.exists("DESCRIPTION") && getwd() != dirname(getwd())) {
  setwd("..")
}
stopifnot("Could not find package root" = file.exists("DESCRIPTION"))

devtools::document(quiet = TRUE)
devtools::load_all(".", quiet = TRUE)

source("dev/glm/glm_effect_estimators.R")
source("dev/glm/glm_simulate.R")

suppressPackageStartupMessages({
  library(survival)
  library(grf)
  library(policytree)
})

cat("R version:", R.version.string, "\n")
cat("grf version:", as.character(packageVersion("grf")), "\n")
```


# Objective

Verify empirically what `dmin.grf` is compared against in the policy tree
node metric `diff`.  Specifically:

1. Does `diff >= dmin.grf` gate on the full RMST difference, or half?
2. How does the DR-score-based `diff` relate to the true RMST difference?
3. What is the analogous relationship for `causal_forest` with binary outcomes?

We generate trial data with a known subgroup, compute the true RMST and
RD differences, and compare them to the `diff` values reported by the
policy tree machinery.


# Data-Generating Mechanism

Exponential survival with a strong treatment-harm subgroup.

```{r dgm}
#| code-fold: false

generate_trial <- function(n = 1500, seed = 42L,
                            baseline_hazard = 0.02,
                            sg_log_hr = 0.8,
                            comp_log_hr = -0.3) {
  set.seed(seed)

  x1 <- rbinom(n, 1, 0.5)
  x2 <- rbinom(n, 1, 0.5)
  x3 <- rbinom(n, 1, 0.5)
  x4 <- rbinom(n, 1, 0.5)
  x5 <- rbinom(n, 1, 0.5)
  treat <- rbinom(n, 1, 0.5)

  in_sg <- (x1 == 1) & (x3 == 1)
  log_hr <- ifelse(in_sg, sg_log_hr, comp_log_hr)
  hazard <- baseline_hazard * exp(log_hr * treat)

  surv_time <- rexp(n, rate = hazard)
  cens_time <- runif(n, 30, 50)
  obs_time  <- pmin(surv_time, cens_time)
  event     <- as.integer(surv_time <= cens_time)

  # Binary landmark at 24 months
  event_24m <- as.integer(surv_time <= 24)

  data.frame(
    id = seq_len(n),
    x1 = x1, x2 = x2, x3 = x3, x4 = x4, x5 = x5,
    treat = treat,
    tte = obs_time,
    event = event,
    event_24m = event_24m,
    in_sg = in_sg
  )
}

df <- generate_trial(n = 1500, seed = 42L)
confounders.name <- c("x1", "x2", "x3", "x4", "x5")

cat("N =", nrow(df), ", events =", sum(df$event),
    sprintf("(%.0f%%)\n", 100 * mean(df$event)))
cat("True subgroup {x1=1, x3=1}: n =", sum(df$in_sg), "\n\n")
```


# True Treatment Effects {#sec-true}

Compute the true RMST differences and RD at the landmark time.

```{r true-effects}
#| code-fold: false

# ---- True RMST differences at various tau ----
taus <- c(12, 18, 24, 30, 36)

cat("=== True RMST differences (months lost under treatment) ===\n\n")
cat(sprintf("  %-8s  %12s  %12s\n", "tau", "Subgroup", "Complement"))
cat("  ", paste(rep("-", 35), collapse = ""), "\n")

for (tau in taus) {
  # Theoretical RMST under exponential:
  # RMST(tau) = (1 - exp(-lambda*tau)) / lambda
  lambda0 <- 0.02
  # Subgroup
  lam_sg_c <- lambda0
  lam_sg_t <- lambda0 * exp(0.8)
  rmst_sg_c <- (1 - exp(-lam_sg_c * tau)) / lam_sg_c
  rmst_sg_t <- (1 - exp(-lam_sg_t * tau)) / lam_sg_t
  rmst_diff_sg <- rmst_sg_c - rmst_sg_t  # positive = treatment harmful

  # Complement
  lam_co_c <- lambda0
  lam_co_t <- lambda0 * exp(-0.3)
  rmst_co_c <- (1 - exp(-lam_co_c * tau)) / lam_co_c
  rmst_co_t <- (1 - exp(-lam_co_t * tau)) / lam_co_t
  rmst_diff_co <- rmst_co_c - rmst_co_t

  cat(sprintf("  %-8d  %12.2f  %12.2f\n", tau, rmst_diff_sg, rmst_diff_co))
}

# ---- True RD at landmark ----
cat("\n=== True Risk Difference at t* = 24 months ===\n")
p_ctrl_sg <- 1 - exp(-0.02 * 24)
p_trt_sg  <- 1 - exp(-0.02 * exp(0.8) * 24)
rd_sg <- p_trt_sg - p_ctrl_sg

p_ctrl_co <- 1 - exp(-0.02 * 24)
p_trt_co  <- 1 - exp(-0.02 * exp(-0.3) * 24)
rd_co <- p_trt_co - p_ctrl_co

cat("  Subgroup RD:   ", round(rd_sg, 4), " (treatment increases risk)\n")
cat("  Complement RD: ", round(rd_co, 4), " (treatment decreases risk)\n")
```


# GRF: `grf.subg.harm.survival` {#sec-grf-survival}

## Run with Full Diagnostics

```{r grf-survival-run}
#| code-fold: false

# Match the GBSG analysis configuration
grf_surv <- grf.subg.harm.survival(
  data = df,
  confounders.name = confounders.name,
  outcome.name = "tte",
  event.name = "event",
  id.name = "id",
  treat.name = "treat",
  maxdepth = 2,
  n.min = 60,
  dmin.grf = 0,          # start with 0 to see all candidates
  frac.tau = 0.6,
  RCT = TRUE,
  details = TRUE,
  return_selected_cuts_only = FALSE
)
```

## Inspect Node Metrics Directly

The key question: what does the `diff` column in `compute_node_metrics`
actually represent, and how does it compare to the true RMST difference?

```{r grf-survival-internals}
#| code-fold: false

# Re-fit the causal survival forest to inspect DR scores directly
X <- as.matrix(df[, confounders.name])
Y <- df$tte
W <- df$treat
D <- df$event

tau_csf <- 0.6 * min(max(Y[W == 1 & D == 1]), max(Y[W == 0 & D == 1]))
cat("Horizon (frac.tau = 0.6):", round(tau_csf, 2), "months\n\n")

cs.forest <- causal_survival_forest(
  X, Y, W, D, W.hat = 0.5,
  horizon = tau_csf, seed = 8316951
)

# DR scores: 2-column matrix [control_score, treated_score]
dr_scores <- double_robust_scores(cs.forest)
cat("DR score dimensions:", dim(dr_scores), "\n")
cat("DR score column names:", colnames(dr_scores), "\n\n")

# Individual-level CATE-like quantity
cate_dr <- dr_scores[, 1] - dr_scores[, 2]

cat("=== DR-score-based CATE (control - treated) ===\n")
cat("  Overall mean:", round(mean(cate_dr), 3), "\n")
cat("  Subgroup mean:", round(mean(cate_dr[df$in_sg]), 3), "\n")
cat("  Complement mean:", round(mean(cate_dr[!df$in_sg]), 3), "\n\n")

# GRF's own CATE estimates (for comparison)
tau_hat <- predict(cs.forest)$predictions
cat("=== GRF predict() CATE estimates ===\n")
cat("  Overall mean:", round(mean(tau_hat), 3), "\n")
cat("  Subgroup mean:", round(mean(tau_hat[df$in_sg]), 3), "\n")
cat("  Complement mean:", round(mean(tau_hat[!df$in_sg]), 3), "\n\n")

cat("=== Theoretical RMST diff at tau =", round(tau_csf, 1), "===\n")
lam0 <- 0.02
rmst_c <- (1 - exp(-lam0 * tau_csf)) / lam0
rmst_t <- (1 - exp(-lam0 * exp(0.8) * tau_csf)) / (lam0 * exp(0.8))
cat("  Subgroup RMST diff (control - treated):", round(rmst_c - rmst_t, 3), "\n")
rmst_co_c <- (1 - exp(-lam0 * tau_csf)) / lam0
rmst_co_t <- (1 - exp(-lam0 * exp(-0.3) * tau_csf)) / (lam0 * exp(-0.3))
cat("  Complement RMST diff (control - treated):", round(rmst_co_c - rmst_co_t, 3), "\n")
```

## Policy Tree Node Metrics

```{r grf-survival-nodes}
#| code-fold: false

# Fit policy tree (same as grf.subg.harm.survival does internally)
tree2 <- policy_tree(X, dr_scores, depth = 2, min.node.size = 60)
metrics <- compute_node_metrics(df, dr_scores, tree2, X, n.min = 60)

cat("=== Policy Tree Node Metrics ===\n")
cat("  (diff = mean control DR score - mean treated DR score)\n\n")
cat(sprintf("  %-8s  %10s  %6s  %s\n",
            "Node", "diff", "N", "Interpretation"))
cat("  ", paste(rep("-", 50), collapse = ""), "\n")
for (i in seq_len(nrow(metrics))) {
  node_ids <- which(predict(tree2, X, type = "node.id") == metrics$leaf.node[i])
  pct_in_sg <- round(100 * mean(df$in_sg[node_ids]), 0)
  interp <- if (pct_in_sg > 80) {
    sprintf("~subgroup (%d%% true SG)", pct_in_sg)
  } else if (pct_in_sg < 20) {
    sprintf("~complement (%d%% true SG)", pct_in_sg)
  } else {
    sprintf("mixed (%d%% true SG)", pct_in_sg)
  }

  cat(sprintf("  %-8d  %10.3f  %6d  %s\n",
              metrics$leaf.node[i],
              metrics$diff[i],
              metrics$Nsg[i],
              interp))
}

# The critical comparison
cat("\n=== dmin.grf Comparison ===\n")
max_diff <- max(metrics$diff)
cat("  Max diff across nodes:", round(max_diff, 3), "\n")
cat("  True RMST diff (subgroup, tau =", round(tau_csf, 1), "):",
    round(rmst_c - rmst_t, 3), "\n")
cat("  Ratio (observed diff / true RMST diff):",
    round(max_diff / (rmst_c - rmst_t), 3), "\n\n")
cat("  => If ratio ≈ 1.0, dmin.grf compares directly to RMST diff\n")
cat("  => If ratio ≈ 0.5, dmin.grf compares to RMST diff / 2\n")
```


## `dmin.grf` Sweep: Survival

```{r dmin-sweep-survival}
#| code-fold: false

dmin_vals <- c(0, 1, 2, 3, 5, 8, 10, 12, 15, 20)

cat("=== dmin.grf sweep: grf.subg.harm.survival ===\n")
cat(sprintf("  %-10s  %-12s  %s\n", "dmin.grf", "Subgroup?", "Cuts"))
cat("  ", paste(rep("-", 60), collapse = ""), "\n")

for (d in dmin_vals) {
  res <- tryCatch(
    grf.subg.harm.survival(
      data = df,
      confounders.name = confounders.name,
      outcome.name = "tte", event.name = "event",
      id.name = "id", treat.name = "treat",
      maxdepth = 2, n.min = 60, dmin.grf = d,
      frac.tau = 0.6, RCT = TRUE, details = FALSE,
      return_selected_cuts_only = FALSE
    ),
    error = function(e) NULL
  )

  if (!is.null(res) && !is.null(res$sg.harm.id) &&
      length(res$sg.harm.id) > 0 && !all(is.na(res$sg.harm.id))) {
    sg <- paste(res$sg.harm.id, collapse = " & ")
    cuts <- paste(res$tree.cuts, collapse = ", ")
    cat(sprintf("  %-10.0f  %-12s  %s\n", d, sg, cuts))
  } else {
    cat(sprintf("  %-10.0f  %-12s  %s\n", d, "(none)", "-"))
  }
}
```


# GRF: `grf.subg.harm.glm` (Binary) {#sec-grf-binary}

## Inspect DR Scores

```{r grf-binary-internals}
#| code-fold: false

Y_bin <- df$event_24m

c.forest <- causal_forest(
  X, Y_bin, W, W.hat = 0.5, seed = 8316951
)
dr_bin <- double_robust_scores(c.forest)

cate_bin <- dr_bin[, 1] - dr_bin[, 2]

cat("=== causal_forest (binary 24m event) DR scores ===\n")
cat("  Overall mean CATE:", round(mean(cate_bin), 4), "\n")
cat("  Subgroup mean CATE:", round(mean(cate_bin[df$in_sg]), 4), "\n")
cat("  Complement mean CATE:", round(mean(cate_bin[!df$in_sg]), 4), "\n\n")

# GRF's own CATE
tau_hat_bin <- predict(c.forest)$predictions
cat("=== GRF predict() CATE ===\n")
cat("  Overall mean:", round(mean(tau_hat_bin), 4), "\n")
cat("  Subgroup mean:", round(mean(tau_hat_bin[df$in_sg]), 4), "\n")
cat("  Complement mean:", round(mean(tau_hat_bin[!df$in_sg]), 4), "\n\n")

cat("=== True RD (P(event|trt) - P(event|ctrl)) ===\n")
cat("  Subgroup:  ", round(rd_sg, 4), "\n")
cat("  Complement:", round(rd_co, 4), "\n")
```

## Policy Tree Node Metrics

```{r grf-binary-nodes}
#| code-fold: false

tree2_bin <- policy_tree(X, dr_bin, depth = 2, min.node.size = 60)
metrics_bin <- compute_node_metrics(df, dr_bin, tree2_bin, X, n.min = 60)

cat("=== Policy Tree Node Metrics (binary) ===\n")
cat("  (diff = mean control DR score - mean treated DR score)\n\n")
for (i in seq_len(nrow(metrics_bin))) {
  node_ids <- which(predict(tree2_bin, X, type = "node.id") == metrics_bin$leaf.node[i])
  pct_in_sg <- round(100 * mean(df$in_sg[node_ids]), 0)

  cat(sprintf("  Node %d: diff = %7.4f  N = %4d  (%d%% true SG)\n",
              metrics_bin$leaf.node[i],
              metrics_bin$diff[i],
              metrics_bin$Nsg[i],
              pct_in_sg))
}

max_diff_bin <- max(metrics_bin$diff)
cat("\n  Max diff:", round(max_diff_bin, 4), "\n")
cat("  True RD (subgroup):", round(rd_sg, 4),
    "  (sign-flipped:", round(-rd_sg, 4), ")\n")
cat("  Ratio (observed diff / |true RD|):",
    round(max_diff_bin / abs(rd_sg), 3), "\n")
```

## `dmin.grf` Sweep: Binary

```{r dmin-sweep-binary}
#| code-fold: false

dmin_bin_vals <- c(0, 0.01, 0.025, 0.05, 0.08, 0.10, 0.15, 0.20, 0.25, 0.30)

cat("=== dmin.grf sweep: grf.subg.harm.glm (binary 24m) ===\n")
cat(sprintf("  %-10s  %-12s  %s\n", "dmin.grf", "Subgroup?", "Cuts"))
cat("  ", paste(rep("-", 60), collapse = ""), "\n")

for (d in dmin_bin_vals) {
  res <- tryCatch(
    grf.subg.harm.glm(
      data = df,
      confounders.name = confounders.name,
      outcome.name = "event_24m",
      treat.name = "treat", id.name = "id",
      n.min = 60, dmin.grf = d,
      RCT = TRUE, maxdepth = 2, seedit = 8316951,
      return_selected_cuts_only = FALSE, details = FALSE
    ),
    error = function(e) NULL
  )

  if (!is.null(res) && !is.null(res$sg.harm.id) &&
      length(res$sg.harm.id) > 0 && !all(is.na(res$sg.harm.id))) {
    sg <- paste(res$sg.harm.id, collapse = " & ")
    cuts <- paste(res$tree.cuts, collapse = ", ")
    cat(sprintf("  %-10.3f  %-12s  %s\n", d, sg, cuts))
  } else {
    cat(sprintf("  %-10.3f  %-12s  %s\n", d, "(none)", "-"))
  }
}
```


# DR Score Sign Convention {#sec-sign}

Understanding the sign convention is critical for interpreting `diff`
and calibrating `dmin.grf`.

```{r sign-convention}
cat("=== Sign Convention Analysis ===\n\n")
cat("double_robust_scores() returns a 2-column matrix:\n")
cat("  Column 1: 'control' — DR score for action 'do not treat'\n")
cat("  Column 2: 'treated' — DR score for action 'treat'\n\n")

cat("compute_node_metrics() computes:\n")
cat("  diff = mean(control_score) - mean(treated_score)\n\n")

cat("Interpretation:\n")
cat("  diff > 0 => control action has higher value => treatment is HARMFUL\n")
cat("  diff < 0 => treatment action has higher value => treatment is BENEFICIAL\n\n")

cat("For causal_survival_forest (RMST):\n")
cat("  Positive diff = treatment reduces RMST by ~diff months\n")
cat("  Units: months of restricted mean survival\n\n")

cat("For causal_forest (binary outcome):\n")
cat("  Positive diff = treatment increases event probability by ~diff\n")
cat("  BUT: sign depends on whether higher Y is good or bad\n")
cat("  For event = bad (status/death): diff > 0 means treatment is harmful\n\n")

# Verify with subgroup-level DR scores
cat("=== Empirical verification ===\n\n")

cat("Survival (RMST):\n")
cat("  Subgroup (treatment harmful):\n")
cat("    mean control_score:", round(mean(dr_scores[df$in_sg, 1]), 3), "\n")
cat("    mean treated_score:", round(mean(dr_scores[df$in_sg, 2]), 3), "\n")
cat("    diff:", round(mean(dr_scores[df$in_sg, 1]) - mean(dr_scores[df$in_sg, 2]), 3), "\n\n")

cat("  Complement (treatment beneficial):\n")
cat("    mean control_score:", round(mean(dr_scores[!df$in_sg, 1]), 3), "\n")
cat("    mean treated_score:", round(mean(dr_scores[!df$in_sg, 2]), 3), "\n")
cat("    diff:", round(mean(dr_scores[!df$in_sg, 1]) - mean(dr_scores[!df$in_sg, 2]), 3), "\n\n")

cat("Binary (event at 24m):\n")
cat("  Subgroup (treatment harmful — increases events):\n")
cat("    mean control_score:", round(mean(dr_bin[df$in_sg, 1]), 4), "\n")
cat("    mean treated_score:", round(mean(dr_bin[df$in_sg, 2]), 4), "\n")
cat("    diff:", round(mean(dr_bin[df$in_sg, 1]) - mean(dr_bin[df$in_sg, 2]), 4), "\n\n")

cat("  Complement (treatment beneficial — decreases events):\n")
cat("    mean control_score:", round(mean(dr_bin[!df$in_sg, 1]), 4), "\n")
cat("    mean treated_score:", round(mean(dr_bin[!df$in_sg, 2]), 4), "\n")
cat("    diff:", round(mean(dr_bin[!df$in_sg, 1]) - mean(dr_bin[!df$in_sg, 2]), 4), "\n")
```


# Summary Table {#sec-summary}

```{r summary}
cat("=" |> rep(75) |> paste(collapse = ""), "\n")
cat("dmin.grf Calibration Summary\n")
cat("=" |> rep(75) |> paste(collapse = ""), "\n\n")

cat(sprintf("%-30s  %12s  %12s\n", "", "Survival", "Binary (24m)"))
cat("-" |> rep(75) |> paste(collapse = ""), "\n")

cat(sprintf("%-30s  %12.1f  %12s\n",
    "Horizon / landmark",
    tau_csf, "24 months"))

cat(sprintf("%-30s  %12.3f  %12.4f\n",
    "True effect (subgroup)",
    rmst_c - rmst_t, rd_sg))

cat(sprintf("%-30s  %12.3f  %12.4f\n",
    "Max node diff",
    max(metrics$diff), max(metrics_bin$diff)))

cat(sprintf("%-30s  %12.3f  %12.3f\n",
    "Ratio (node diff / true)",
    max(metrics$diff) / (rmst_c - rmst_t),
    max(metrics_bin$diff) / abs(rd_sg)))

cat("-" |> rep(75) |> paste(collapse = ""), "\n\n")

cat("If ratio ≈ 1.0: dmin.grf gates directly on the true effect size\n")
cat("If ratio ≈ 0.5: dmin.grf gates on half the true effect size\n")
cat("If ratio differs: relationship depends on DR score estimation\n")
```


# Critical Finding: Sign Convention {#sec-critical}

::: {.callout-warning}
## Binary outcome sign convention is inverted

For `causal_survival_forest`, longer survival is better, so the DR scores
naturally align with the harm convention: `diff > 0` means treatment is
harmful (control is better).

For `causal_forest` with a binary **adverse** outcome (Y = event = bad),
the forest tries to **maximize Y**, so the treated action scores higher
in the harmful subgroup (because treatment produces more events).  This
means `diff = control_score - treated_score < 0` in the harmful subgroup.

The `select_best_subgroup` check (`diff >= dmin.grf`) would never select
the harmful subgroup — it selects the complement instead.
:::

## Demonstration

```{r sign-demo}
cat("=== Sign Convention Demonstration ===\n\n")
cat("Survival (higher Y = longer life = GOOD):\n")
cat("  Subgroup node (treatment harmful): diff =",
    round(metrics$diff[which.max(metrics$diff)], 3), "> 0 => CORRECT\n")
cat("  dmin.grf gate: diff >= dmin works as intended\n\n")

cat("Binary (Y = event = BAD):\n")
sg_node <- which.min(metrics_bin$diff)
co_node <- which.max(metrics_bin$diff)
cat("  Subgroup node (treatment harmful): diff =",
    round(metrics_bin$diff[sg_node], 4), "< 0 => WRONG SIGN\n")
cat("  Complement node (treatment beneficial): diff =",
    round(metrics_bin$diff[co_node], 4), "> 0\n")
cat("  dmin.grf gate: selects COMPLEMENT, not harm subgroup!\n")
```

## Fix: Flip Outcome for Adverse Events

For binary outcomes where Y = 1 is adverse (the standard clinical trial
convention), we should pass `1 - Y` to `causal_forest` so that higher
values = better (no event).

```{r sign-fix}
cat("=== Fix: Pass (1 - Y) to causal_forest ===\n\n")

Y_flipped <- 1L - Y_bin

cf_fixed <- causal_forest(
  X, Y_flipped, W, W.hat = 0.5, seed = 8316951
)
dr_fixed <- double_robust_scores(cf_fixed)

tree_fixed <- policy_tree(X, dr_fixed, depth = 2, min.node.size = 60)
metrics_fixed <- compute_node_metrics(df, dr_fixed, tree_fixed, X, n.min = 60)

cat("Node metrics with flipped outcome (1 - event):\n")
for (i in seq_len(nrow(metrics_fixed))) {
  node_ids <- which(predict(tree_fixed, X, type = "node.id") == metrics_fixed$leaf.node[i])
  pct_in_sg <- round(100 * mean(df$in_sg[node_ids]), 0)
  cat(sprintf("  Node %d: diff = %7.4f  N = %4d  (%d%% true SG)\n",
              metrics_fixed$leaf.node[i],
              metrics_fixed$diff[i],
              metrics_fixed$Nsg[i],
              pct_in_sg))
}

max_diff_fixed <- max(metrics_fixed$diff)
cat("\n  Max diff:", round(max_diff_fixed, 4), "\n")
cat("  True |RD|:", round(abs(rd_sg), 4), "\n")
cat("  Ratio:", round(max_diff_fixed / abs(rd_sg), 3), "\n\n")

# Verify the fix works through grf.subg.harm.glm
cat("=== Verify fix: pass (1 - event) to grf.subg.harm.glm ===\n")
df$no_event_24m <- 1L - df$event_24m

grf_fixed <- grf.subg.harm.glm(
  data = df,
  confounders.name = confounders.name,
  outcome.name = "no_event_24m",
  treat.name = "treat", id.name = "id",
  n.min = 60, dmin.grf = 0.0,
  RCT = TRUE, maxdepth = 2, seedit = 8316951,
  return_selected_cuts_only = TRUE, details = TRUE
)

if (!is.null(grf_fixed$sg.harm.id)) {
  cat("  Subgroup found:", paste(grf_fixed$sg.harm.id, collapse = " & "), "\n")
  cat("  Cuts:", paste(grf_fixed$tree.cuts, collapse = ", "), "\n")
} else {
  cat("  No subgroup found\n")
}
```

::: {.callout-important}
## Implementation Implication

`grf.subg.harm.glm()` should automatically flip the outcome when the
effect measure implies Y = 1 is adverse (RD, OR, RR).  This flip should
happen inside the GRF function only — it does not affect the downstream
search/consistency/bootstrap pipeline, which operates on the original
outcome scale.

Proposed change to `grf.subg.harm.glm()`:

```r
# If outcome is an adverse event, flip for causal_forest
# so that higher = better (matches causal_survival_forest convention)
if (adverse_outcome) {
  Y <- 1L - data[, outcome.name]
} else {
  Y <- data[, outcome.name]
}
```
:::


# Session Info

```{r session-info}
sessionInfo()
```
