---
title: "Phase 5 Test Suite: Cross-Validation Threading for GLM Extension"
subtitle: "forestsearch_Kfold and forestsearch_tenfold with outcome_type, effect_measure, PS adjustment"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false

pkg_root <- getwd()
while (!file.exists(file.path(pkg_root, "DESCRIPTION")) &&
       pkg_root != dirname(pkg_root)) {
  pkg_root <- dirname(pkg_root)
}
stopifnot("Could not find package root" = file.exists(
  file.path(pkg_root, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = pkg_root)
knitr::opts_chunk$set(comment = "#>")
```

```{r load}
devtools::document(quiet = TRUE)
devtools::load_all(".", quiet = TRUE)
source("dev/glm/glm_effect_estimators.R")

suppressPackageStartupMessages({
  library(survival)
  library(data.table)
  library(grf)
  library(glmnet)
  library(doFuture)
  library(foreach)
  library(gt)
})

# Track test results
test_log <- data.frame(
  Test = character(), Description = character(),
  Status = character(), Details = character(),
  stringsAsFactors = FALSE
)

log_test <- function(name, desc, status, details = "") {
  test_log <<- rbind(test_log, data.frame(
    Test = name, Description = desc,
    Status = status, Details = details,
    stringsAsFactors = FALSE
  ))
  icon <- if (status == "PASS") "\u2705" else "\u274C"
  cat(icon, name, ":", desc, "—", status, "\n")
  if (nzchar(details)) cat("    ", details, "\n")
}
```


# Data Preparation {#sec-data}

```{r data-prep}
df <- survival::gbsg
df$id <- seq_len(nrow(df))
df$grade3 <- as.integer(df$grade == "3")
confs <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")

cat("GBSG data: N =", nrow(df), "\n")
cat("Confounders:", paste(confs, collapse = ", "), "\n")
cat("Treatment (hormon): n1 =", sum(df$hormon), ", n0 =", sum(1 - df$hormon), "\n")
```


# Baseline ForestSearch Runs {#sec-baseline}

Build a set of `forestsearch()` result objects covering all parameter
combinations that Phase 5 must handle.  These are the inputs to the
CV functions.

```{r fs-common-args}
# Shared arguments for all forestsearch calls
fs_common <- list(
  df.analysis = df,
  confounders.name = confs,
  conf_force = c("er <= 0"),
  treat.name = "hormon",
  id.name = "id",
  hr.threshold = 1.25,
  hr.consistency = 1.0,
  pconsistency.threshold = 0.80,
  stop_threshold = 0.80,
  fs.splits = 100,
  max_subgroups_search = 3,
  use_twostage = TRUE,
  use_grf = FALSE,
  use_lasso = TRUE,
  return_selected_cuts_only = TRUE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
  maxk = 2,
  sg_focus = "hr"
)
```


## Run 1: Survival (regression baseline)

```{r fs-survival}
args_surv <- c(fs_common, list(
  outcome.name = "rfstime",
  event.name = "status"
))
fs_surv <- do.call(forestsearch, args_surv)
cat("Survival subgroup:", paste(fs_surv$sg.harm, collapse = " & "), "\n")
cat("outcome_type:", fs_surv$args_call_all$outcome_type, "\n")
```


## Run 2: GLM Binary IRR (RCT, no PS)

```{r fs-glm-rct}
args_glm_rct <- c(fs_common, list(
  outcome.name = "status",
  event.name = "status",
  outcome_type = "binary",
  effect_measure = "IRR",
  offset.name = "rfstime",
  dmin.grf = 0.0
))
fs_glm_rct <- do.call(forestsearch, args_glm_rct)
cat("GLM IRR (RCT) subgroup:", paste(fs_glm_rct$sg.harm, collapse = " & "), "\n")
cat("ps_method:", fs_glm_rct$args_call_all$ps_method, "\n")
```


## Run 3: GLM Binary IRR + IPTW (observational)

```{r fs-glm-iptw}
args_glm_iptw <- c(fs_common, list(
  outcome.name = "status",
  event.name = "status",
  outcome_type = "binary",
  effect_measure = "IRR",
  offset.name = "rfstime",
  dmin.grf = 0.0,
  is.RCT = FALSE,
  ps_method = "grf",
  ps_adjust_method = "iptw"
))
fs_glm_iptw <- do.call(forestsearch, args_glm_iptw)
cat("GLM IRR + IPTW subgroup:", paste(fs_glm_iptw$sg.harm, collapse = " & "), "\n")
cat("ps_method:", fs_glm_iptw$args_call_all$ps_method, "\n")
cat("ps_adjust_method:", fs_glm_iptw$args_call_all$ps_adjust_method, "\n")
cat("sw in df.est:", "sw" %in% names(fs_glm_iptw$df.est), "\n")
```


## Run 4: GLM Binary OR (RCT, no PS)

```{r fs-glm-or}
args_glm_or <- c(fs_common, list(
  outcome.name = "status",
  event.name = "status",
  outcome_type = "binary",
  effect_measure = "OR",
  dmin.grf = 0.0
))
fs_glm_or <- do.call(forestsearch, args_glm_or)
cat("GLM OR (RCT) subgroup:", paste(fs_glm_or$sg.harm, collapse = " & "), "\n")
```


## Run 5: GLM Binary OR + IPTW (observational)

```{r fs-glm-or-iptw}
args_glm_or_iptw <- c(fs_common, list(
  outcome.name = "status",
  event.name = "status",
  outcome_type = "binary",
  effect_measure = "OR",
  dmin.grf = 0.0,
  is.RCT = FALSE,
  ps_method = "grf",
  ps_adjust_method = "iptw"
))
fs_glm_or_iptw <- do.call(forestsearch, args_glm_or_iptw)
cat("GLM OR + IPTW subgroup:", paste(fs_glm_or_iptw$sg.harm, collapse = " & "), "\n")
```


# Test Group 1: forestsearch_Kfold {#sec-kfold}

## Test 1.1: Survival → Kfold (regression baseline)

```{r test-1-1}
test_id <- "1.1"
test_desc <- "Survival → Kfold (5-fold)"

result <- tryCatch({
  if (is.null(fs_surv$sg.harm)) stop("No subgroup found in baseline")
  cv <- forestsearch_Kfold(
    fs.est = fs_surv, Kfolds = 5,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
    details = TRUE
  )
  stopifnot(is.list(cv))
  stopifnot(!is.null(cv$resCV))
  stopifnot(nrow(cv$resCV) == nrow(df))
  stopifnot(cv$Kfolds == 5)
  paste0("prop_SG=", cv$prop_SG_found, "%, sens_H=",
         round(cv$sens_summary["sens_H"], 3))
}, error = function(e) e$message)

if (!grepl("^Error|^No sub", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 1.2: GLM IRR (RCT) → Kfold

```{r test-1-2}
test_id <- "1.2"
test_desc <- "GLM IRR (RCT, no PS) → Kfold (5-fold)"

result <- tryCatch({
  if (is.null(fs_glm_rct$sg.harm)) stop("No subgroup found in baseline")
  cv <- forestsearch_Kfold(
    fs.est = fs_glm_rct, Kfolds = 5,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
    details = TRUE
  )
  stopifnot(nrow(cv$resCV) == nrow(df))
  # Verify offset column was available (no crash)
  paste0("prop_SG=", cv$prop_SG_found, "%, sens_H=",
         round(cv$sens_summary["sens_H"], 3))
}, error = function(e) e$message)

if (!grepl("^Error|^No sub", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 1.3: GLM IRR + IPTW → Kfold

```{r test-1-3}
test_id <- "1.3"
test_desc <- "GLM IRR + IPTW (observational) → Kfold (5-fold)"

result <- tryCatch({
  if (is.null(fs_glm_iptw$sg.harm)) stop("No subgroup found in baseline")
  cv <- forestsearch_Kfold(
    fs.est = fs_glm_iptw, Kfolds = 5,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
    details = TRUE
  )
  stopifnot(nrow(cv$resCV) == nrow(df))
  # Verify PS was re-estimated (not carried from original)
  cv_ps_method <- cv$cv_args$ps_method
  cv_ps_hat <- cv$cv_args$ps_hat
  stopifnot(cv_ps_method == "grf")
  stopifnot(is.null(cv_ps_hat))  # Must be NULL'd for re-estimation
  paste0("prop_SG=", cv$prop_SG_found, "%, ps_method=", cv_ps_method,
         ", ps_hat=NULL (re-estimated)")
}, error = function(e) e$message)

if (!grepl("^Error|^No sub", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 1.4: GLM OR (RCT) → Kfold

```{r test-1-4}
test_id <- "1.4"
test_desc <- "GLM OR (RCT, no PS, no offset) → Kfold (5-fold)"

result <- tryCatch({
  if (is.null(fs_glm_or$sg.harm)) stop("No subgroup found in baseline")
  cv <- forestsearch_Kfold(
    fs.est = fs_glm_or, Kfolds = 5,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
    details = TRUE
  )
  stopifnot(nrow(cv$resCV) == nrow(df))
  paste0("prop_SG=", cv$prop_SG_found, "%")
}, error = function(e) e$message)

if (!grepl("^Error|^No sub", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 1.5: GLM OR + IPTW → Kfold

```{r test-1-5}
test_id <- "1.5"
test_desc <- "GLM OR + IPTW (observational) → Kfold (5-fold)"

result <- tryCatch({
  if (is.null(fs_glm_or_iptw$sg.harm)) stop("No subgroup found in baseline")
  cv <- forestsearch_Kfold(
    fs.est = fs_glm_or_iptw, Kfolds = 5,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
    details = TRUE
  )
  stopifnot(nrow(cv$resCV) == nrow(df))
  paste0("prop_SG=", cv$prop_SG_found, "%")
}, error = function(e) e$message)

if (!grepl("^Error|^No sub", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 1.6: print_cv_params displays GLM and PS params

```{r test-1-6}
test_id <- "1.6"
test_desc <- "print_cv_params shows outcome_type, effect_measure, ps_method"

result <- tryCatch({
  if (is.null(fs_glm_iptw$sg.harm)) stop("No subgroup found in baseline")
  output <- capture.output(
    forestsearch_Kfold(
      fs.est = fs_glm_iptw, Kfolds = 3,
      parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
      details = TRUE
    )
  )
  has_outcome <- any(grepl("outcome_type: binary", output))
  has_measure <- any(grepl("effect_measure: IRR", output))
  has_offset  <- any(grepl("offset.name: rfstime", output))
  has_ps      <- any(grepl("ps_method: grf", output))
  has_adjust  <- any(grepl("ps_adjust_method: iptw", output))
  has_reest   <- any(grepl("re-estimated on each fold", output))

  checks <- c(outcome = has_outcome, measure = has_measure, offset = has_offset,
               ps = has_ps, adjust = has_adjust, reest = has_reest)
  if (all(checks)) {
    "All 6 params printed"
  } else {
    paste0("Missing: ", paste(names(checks)[!checks], collapse = ", "))
  }
}, error = function(e) e$message)

if (result == "All 6 params printed") {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 1.7: print_cv_params omits PS params for RCT

```{r test-1-7}
test_id <- "1.7"
test_desc <- "print_cv_params omits ps_method for RCT (ps_method=none)"

result <- tryCatch({
  if (is.null(fs_glm_rct$sg.harm)) stop("No subgroup found in baseline")
  output <- capture.output(
    forestsearch_Kfold(
      fs.est = fs_glm_rct, Kfolds = 3,
      parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
      details = TRUE
    )
  )
  has_ps <- any(grepl("ps_method:", output))
  has_outcome <- any(grepl("outcome_type: binary", output))

  if (!has_ps && has_outcome) {
    "ps_method absent (correct), outcome_type present"
  } else {
    paste0("ps_method shown=", has_ps, ", outcome_type=", has_outcome)
  }
}, error = function(e) e$message)

if (grepl("correct", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 1.8: print_cv_params omits GLM params for survival

```{r test-1-8}
test_id <- "1.8"
test_desc <- "print_cv_params omits GLM params for survival outcome"

result <- tryCatch({
  if (is.null(fs_surv$sg.harm)) stop("No subgroup found in baseline")
  output <- capture.output(
    forestsearch_Kfold(
      fs.est = fs_surv, Kfolds = 3,
      parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE),
      details = TRUE
    )
  )
  has_outcome_type <- any(grepl("outcome_type:", output))
  has_effect_meas  <- any(grepl("effect_measure:", output))
  has_ps           <- any(grepl("ps_method:", output))

  if (!has_outcome_type && !has_effect_meas && !has_ps) {
    "GLM and PS params correctly omitted"
  } else {
    paste0("outcome_type=", has_outcome_type, ", effect_measure=",
           has_effect_meas, ", ps=", has_ps)
  }
}, error = function(e) e$message)

if (grepl("correctly omitted", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


# Test Group 2: forestsearch_tenfold {#sec-tenfold}

## Test 2.1: Survival → tenfold (regression baseline)

```{r test-2-1}
test_id <- "2.1"
test_desc <- "Survival → tenfold (2 sims, 3-fold)"

result <- tryCatch({
  if (is.null(fs_surv$sg.harm)) stop("No subgroup found in baseline")
  tf <- forestsearch_tenfold(
    fs.est = fs_surv, sims = 2, Kfolds = 3,
    details = TRUE,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  stopifnot(is.list(tf))
  stopifnot(!is.null(tf$sens_summary))
  stopifnot(!is.null(tf$find_summary))
  stopifnot(length(tf$sens_summary) == 4)
  paste0("2/2 sims, sens_H=", round(tf$sens_summary["sens_H"], 3))
}, error = function(e) e$message)

if (!grepl("^Error|^No sub|^All sim", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 2.2: GLM IRR + IPTW → tenfold

```{r test-2-2}
test_id <- "2.2"
test_desc <- "GLM IRR + IPTW → tenfold (2 sims, 3-fold)"

result <- tryCatch({
  if (is.null(fs_glm_iptw$sg.harm)) stop("No subgroup found in baseline")
  tf <- forestsearch_tenfold(
    fs.est = fs_glm_iptw, sims = 2, Kfolds = 3,
    details = TRUE,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  stopifnot(is.list(tf))
  stopifnot(length(tf$sens_summary) == 4)
  paste0("2/2 sims, sens_H=", round(tf$sens_summary["sens_H"], 3))
}, error = function(e) e$message)

if (!grepl("^Error|^No sub|^All sim", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 2.3: GLM OR (RCT) → tenfold

```{r test-2-3}
test_id <- "2.3"
test_desc <- "GLM OR (RCT, no PS) → tenfold (2 sims, 3-fold)"

result <- tryCatch({
  if (is.null(fs_glm_or$sg.harm)) stop("No subgroup found in baseline")
  tf <- forestsearch_tenfold(
    fs.est = fs_glm_or, sims = 2, Kfolds = 3,
    details = TRUE,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  stopifnot(is.list(tf))
  paste0("2/2 sims, sens_H=", round(tf$sens_summary["sens_H"], 3))
}, error = function(e) e$message)

if (!grepl("^Error|^No sub|^All sim", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 2.4: GLM OR + IPTW → tenfold

```{r test-2-4}
test_id <- "2.4"
test_desc <- "GLM OR + IPTW → tenfold (2 sims, 3-fold)"

result <- tryCatch({
  if (is.null(fs_glm_or_iptw$sg.harm)) stop("No subgroup found in baseline")
  tf <- forestsearch_tenfold(
    fs.est = fs_glm_or_iptw, sims = 2, Kfolds = 3,
    details = TRUE,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  stopifnot(is.list(tf))
  paste0("2/2 sims, sens_H=", round(tf$sens_summary["sens_H"], 3))
}, error = function(e) e$message)

if (!grepl("^Error|^No sub|^All sim", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


# Test Group 3: Parameter Flow Verification {#sec-param-flow}

## Test 3.1: ps_hat is NULL'd for CV folds

```{r test-3-1}
test_id <- "3.1"
test_desc <- "cv_args$ps_hat is NULL (prevents length mismatch)"

result <- tryCatch({
  if (is.null(fs_glm_iptw$sg.harm)) stop("No subgroup found in baseline")
  # Kfold internally builds cv_args — check via the returned object
  cv <- forestsearch_Kfold(
    fs.est = fs_glm_iptw, Kfolds = 3,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  ps_hat_in_cv <- cv$cv_args$ps_hat
  if (is.null(ps_hat_in_cv)) {
    "ps_hat correctly NULL'd in cv_args"
  } else {
    paste0("ps_hat has length ", length(ps_hat_in_cv), " — should be NULL")
  }
}, error = function(e) e$message)

if (grepl("correctly NULL", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 3.2: offset.name included in fold data

```{r test-3-2}
test_id <- "3.2"
test_desc <- "offset.name column present in CV fold data"

result <- tryCatch({
  if (is.null(fs_glm_iptw$sg.harm)) stop("No subgroup found in baseline")
  # The fact that Kfold completed without error proves offset was present.
  # Verify by checking cv_args contains the offset.name
  cv <- forestsearch_Kfold(
    fs.est = fs_glm_iptw, Kfolds = 3,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  off <- cv$cv_args$offset.name
  if (!is.null(off) && off == "rfstime") {
    "offset.name='rfstime' preserved in cv_args"
  } else {
    paste0("offset.name=", deparse(off))
  }
}, error = function(e) e$message)

if (grepl("preserved", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 3.3: ps_method flows through to CV folds

```{r test-3-3}
test_id <- "3.3"
test_desc <- "ps_method=grf and ps_adjust_method=iptw flow to cv_args"

result <- tryCatch({
  if (is.null(fs_glm_iptw$sg.harm)) stop("No subgroup found in baseline")
  cv <- forestsearch_Kfold(
    fs.est = fs_glm_iptw, Kfolds = 3,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  ps_m <- cv$cv_args$ps_method
  ps_a <- cv$cv_args$ps_adjust_method
  if (ps_m == "grf" && ps_a == "iptw") {
    "ps_method=grf, ps_adjust_method=iptw — correct"
  } else {
    paste0("ps_method=", ps_m, ", ps_adjust_method=", ps_a)
  }
}, error = function(e) e$message)

if (grepl("correct", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 3.4: outcome_type and effect_measure flow to CV folds

```{r test-3-4}
test_id <- "3.4"
test_desc <- "outcome_type=binary and effect_measure=IRR flow to cv_args"

result <- tryCatch({
  if (is.null(fs_glm_rct$sg.harm)) stop("No subgroup found in baseline")
  cv <- forestsearch_Kfold(
    fs.est = fs_glm_rct, Kfolds = 3,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  ot <- cv$cv_args$outcome_type
  em <- cv$cv_args$effect_measure
  if (ot == "binary" && em == "IRR") {
    "outcome_type=binary, effect_measure=IRR — correct"
  } else {
    paste0("outcome_type=", ot, ", effect_measure=", em)
  }
}, error = function(e) e$message)

if (grepl("correct", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 3.5: adverse_outcome flows to CV folds

```{r test-3-5}
test_id <- "3.5"
test_desc <- "adverse_outcome resolved and flows to cv_args"

result <- tryCatch({
  if (is.null(fs_glm_rct$sg.harm)) stop("No subgroup found in baseline")
  cv <- forestsearch_Kfold(
    fs.est = fs_glm_rct, Kfolds = 3,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  ao <- cv$cv_args$adverse_outcome
  if (isTRUE(ao)) {
    "adverse_outcome=TRUE — correct (binary outcome)"
  } else {
    paste0("adverse_outcome=", deparse(ao))
  }
}, error = function(e) e$message)

if (grepl("correct", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 3.6: dmin.grf auto-resolved flows to CV folds

```{r test-3-6}
test_id <- "3.6"
test_desc <- "dmin.grf=0.0 (GLM auto-resolved) flows to cv_args"

result <- tryCatch({
  if (is.null(fs_glm_rct$sg.harm)) stop("No subgroup found in baseline")
  cv <- forestsearch_Kfold(
    fs.est = fs_glm_rct, Kfolds = 3,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  dmin <- cv$cv_args$dmin.grf
  if (!is.null(dmin) && dmin == 0.0) {
    "dmin.grf=0.0 — correct (GLM auto-resolved)"
  } else {
    paste0("dmin.grf=", deparse(dmin))
  }
}, error = function(e) e$message)

if (grepl("correct", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


# Test Group 4: KfoldOut GLM Guard {#sec-kfoldout}

## Test 4.1: KfoldOut with outall=TRUE on GLM result

```{r test-4-1}
test_id <- "4.1"
test_desc <- "KfoldOut(outall=TRUE) on GLM result — no crash"

result <- tryCatch({
  if (is.null(fs_glm_iptw$sg.harm)) stop("No subgroup found in baseline")
  cv <- forestsearch_Kfold(
    fs.est = fs_glm_iptw, Kfolds = 5,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  # Build the res object that KfoldOut expects
  res_for_out <- list(
    resCV = cv$resCV,
    Kfolds = cv$Kfolds,
    cv_args = cv$cv_args,
    sg_analysis = fs_glm_iptw$sg.harm,
    sg0.name = "Not recommend",
    sg1.name = "Recommend"
  )
  out <- forestsearch_KfoldOut(res = res_for_out, outall = TRUE, details = TRUE)
  stopifnot(!is.null(out$sens_metrics_original))
  stopifnot(!is.null(out$find_metrics))
  # Should have outcome_type in result (GLM guard path)
  has_ot <- !is.null(out$outcome_type) && out$outcome_type == "binary"
  if (has_ot) {
    "GLM guard path executed, no crash"
  } else {
    "Completed but outcome_type not returned"
  }
}, error = function(e) e$message)

if (grepl("no crash|Completed", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


## Test 4.2: KfoldOut with outall=TRUE on survival result

```{r test-4-2}
test_id <- "4.2"
test_desc <- "KfoldOut(outall=TRUE) on survival result — survival path"

result <- tryCatch({
  if (is.null(fs_surv$sg.harm)) stop("No subgroup found in baseline")
  cv <- forestsearch_Kfold(
    fs.est = fs_surv, Kfolds = 5,
    parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
  )
  res_for_out <- list(
    resCV = cv$resCV,
    Kfolds = cv$Kfolds,
    cv_args = cv$cv_args,
    sg_analysis = fs_surv$sg.harm,
    sg0.name = "Not recommend",
    sg1.name = "Recommend"
  )
  out <- forestsearch_KfoldOut(res = res_for_out, outall = TRUE, details = TRUE)
  # Survival path should produce tab_all
  has_tab <- !is.null(out$tab_all)
  if (has_tab) {
    "Survival outall path: tab_all present"
  } else {
    "tab_all missing"
  }
}, error = function(e) e$message)

if (grepl("tab_all present", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


# Test Group 5: stringr Removal {#sec-stringr}

## Test 5.1: CV_sgs works without stringr

```{r test-5-1}
test_id <- "5.1"
test_desc <- "CV_sgs uses base R (nchar/substr), not stringr"

result <- tryCatch({
  # Call CV_sgs directly with test data
  sg1_test <- c("er <= 0", "size <= 35", NA, "nodes <= 3")
  sg2_test <- c("size <= 35", NA, NA, "er <= 0")
  sg_analysis_test <- c("er <= 0", "size <= 35")
  out <- CV_sgs(sg1 = sg1_test, sg2 = sg2_test,
                confs = confs, sg_analysis = sg_analysis_test)
  stopifnot(is.list(out))
  stopifnot(!is.null(out$any_found))
  stopifnot(!is.null(out$exact_match))
  paste0("any_found=", paste(out$any_found, collapse = ","),
         ", exact=", paste(out$exact_match, collapse = ","))
}, error = function(e) e$message)

if (!grepl("^Error|stringr", result)) {
  log_test(test_id, test_desc, "PASS", result)
} else {
  log_test(test_id, test_desc, "FAIL", result)
}
```


# Results Summary {#sec-summary}

```{r summary-table}
n_pass <- sum(test_log$Status == "PASS")
n_fail <- sum(test_log$Status == "FAIL")
n_total <- nrow(test_log)

cat(sprintf("\n=== Phase 5 Test Results: %d/%d PASSED ===\n\n",
            n_pass, n_total))

gt(test_log) |>
  tab_header(
    title = md("**Phase 5 Test Suite Results**"),
    subtitle = sprintf("%d tests: %d passed, %d failed", n_total, n_pass, n_fail)
  ) |>
  tab_style(
    style = cell_fill(color = "#e8f5e9"),
    locations = cells_body(rows = Status == "PASS")
  ) |>
  tab_style(
    style = cell_fill(color = "#ffcdd2"),
    locations = cells_body(rows = Status == "FAIL")
  )
```

```{r final-check}
if (n_fail > 0) {
  cat("\n\u274C FAILURES DETECTED — review failed tests above.\n")
  cat("Failed tests:\n")
  for (i in which(test_log$Status == "FAIL")) {
    cat("  ", test_log$Test[i], ":", test_log$Description[i], "\n")
    cat("    ", test_log$Details[i], "\n")
  }
} else {
  cat("\n\u2705 ALL TESTS PASSED — Phase 5 CV threading is complete.\n")
}
```


# Session Information

```{r session-info}
sessionInfo()
```
