---
title: "Propensity Score Adjustment for ForestSearch GLM Extension"
subtitle: "Implementation Design and Simulation Evaluation"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false

pkg_root <- getwd()
while (!file.exists(file.path(pkg_root, "DESCRIPTION")) &&
       pkg_root != dirname(pkg_root)) {
  pkg_root <- dirname(pkg_root)
}
stopifnot("Could not find package root" = file.exists(file.path(pkg_root, "DESCRIPTION")))
knitr::opts_knit$set(root.dir = pkg_root)

knitr::opts_chunk$set(comment = "#>", fig.align = "center")
```

```{r load}
devtools::document(quiet = TRUE)
devtools::load_all(".", quiet = TRUE)
source("dev/glm/glm_effect_estimators.R")

suppressPackageStartupMessages({
  library(survival)
  library(data.table)
  library(grf)
  library(glmnet)
  library(gt)
})
```


# Design Overview

ForestSearch's GLM extension now supports optional propensity score (PS)
adjustment for subgroup effect estimation.  The PS is included as an
adjustment covariate in the GLM estimator closure, adding a doubly-robust
flavour to the treatment effect estimates.

## Architecture

The doubly-robust approach follows Bang & Robins (2005): the propensity
score $\hat{e}(X)$ is estimated on the full sample, then the **inverse
propensity score (IPS) covariate** is constructed as $1/\hat{e}(X_i)$
for treated subjects and $1/(1 - \hat{e}(X_i))$ for controls.  This IPS
covariate is included in the outcome GLM, yielding a treatment effect
estimate that is consistent if either the propensity model or the
outcome model is correctly specified.

```
forestsearch()
  └── Section 2C: estimate_propensity_scores()
        ├── method = "grf"      →  grf::regression_forest(X, W)
        ├── method = "lasso"    →  glmnet::cv.glmnet(X, W, family="binomial")
        ├── method = "logistic" →  glm(W ~ X, family=binomial)
        └── method = "none"     →  skip (default for RCT)
      ↓
  ps_hat = e(X) estimated on full sample
  ips_covar = 1/e(X) for treated, 1/(1-e(X)) for control  [Bang & Robins 2005]
      ↓
  make_effect_estimator(..., adjust_covariates = "ips_covar")
      ↓
  closure: glm(Y ~ treat + ips_covar, ...)  →  doubly-robust estimate
```

## Parameter Defaults

| Setting | `ps_method` default | Rationale |
|---|---|---|
| `is.RCT = TRUE` | `"none"` | Randomization ensures balance |
| `is.RCT = FALSE` | `"grf"` | Non-parametric, cross-fitted |
| User-supplied PS | Pass `ps_hat` vector | Maximum flexibility |

## PS Estimation Methods

| Method | Implementation | Pros | Cons |
|---|---|---|---|
| `"grf"` | `regression_forest(X, W)` | Non-parametric, OOB cross-fitted | Requires `grf` |
| `"lasso"` | `cv.glmnet(X, W, family="binomial")` | Penalized, handles many covariates | Parametric (log-odds) |
| `"logistic"` | `glm(W ~ X, family=binomial)` | Simple, interpretable | May misspecify |
| `"none"` | Skip | No overhead | No confounding adjustment |

All methods include PS trimming at [2.5%, 97.5%] quantiles by default
to avoid positivity violations.


# GBSG Demonstration {#sec-gbsg}

## PS Estimation Comparison

```{r ps-estimation}
#| code-fold: false

df <- survival::gbsg
df$id <- seq_len(nrow(df))
df$grade3 <- as.integer(df$grade == "3")
confs <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")

ps_grf    <- estimate_propensity_scores(df, "hormon", confs, method = "grf")
ps_lasso  <- estimate_propensity_scores(df, "hormon", confs, method = "lasso")
ps_logit  <- estimate_propensity_scores(df, "hormon", confs, method = "logistic")

ps_compare <- data.frame(
  Method     = c("GRF", "LASSO", "Logistic"),
  Mean       = round(c(mean(ps_grf$ps_hat), mean(ps_lasso$ps_hat),
                        mean(ps_logit$ps_hat)), 3),
  SD         = round(c(sd(ps_grf$ps_hat), sd(ps_lasso$ps_hat),
                        sd(ps_logit$ps_hat)), 3),
  Min        = round(c(min(ps_grf$ps_hat), min(ps_lasso$ps_hat),
                        min(ps_logit$ps_hat)), 3),
  Max        = round(c(max(ps_grf$ps_hat), max(ps_lasso$ps_hat),
                        max(ps_logit$ps_hat)), 3),
  Trimmed    = c(ps_grf$trimmed, ps_lasso$trimmed, ps_logit$trimmed),
  Cor_GRF    = round(c(1, cor(ps_grf$ps_hat, ps_lasso$ps_hat),
                        cor(ps_grf$ps_hat, ps_logit$ps_hat)), 3)
)

gt(ps_compare) |>
  tab_header(title = md("**PS Estimation Methods on GBSG Data**"),
             subtitle = "P(hormon = 1 | X)") |>
  cols_label(Cor_GRF = "Cor with GRF")
```

```{r ps-histograms, fig.width=10, fig.height=4}
par(mfrow = c(1, 3))
hist(ps_grf$ps_hat, breaks = 30, col = "steelblue", border = "white",
     main = "GRF", xlab = "P(treat = 1 | X)", xlim = c(0, 1))
abline(v = mean(df$hormon), col = "red", lwd = 2, lty = 2)
hist(ps_lasso$ps_hat, breaks = 30, col = "forestgreen", border = "white",
     main = "LASSO", xlab = "P(treat = 1 | X)", xlim = c(0, 1))
abline(v = mean(df$hormon), col = "red", lwd = 2, lty = 2)
hist(ps_logit$ps_hat, breaks = 30, col = "darkorange", border = "white",
     main = "Logistic", xlab = "P(treat = 1 | X)", xlim = c(0, 1))
abline(v = mean(df$hormon), col = "red", lwd = 2, lty = 2)
par(mfrow = c(1, 1))
```

## Adjusted vs Unadjusted Estimates

The doubly-robust adjustment follows Bang & Robins (2005): the inverse
propensity score (IPS) covariate is defined as $1/\hat{e}(X_i)$ for
treated subjects and $1/(1 - \hat{e}(X_i))$ for controls.  Including
this as a covariate in the outcome GLM yields consistency if **either**
the propensity or outcome model is correctly specified.

```{r adj-vs-unadj}
df_ps <- ps_grf$data

fn_unadj <- make_effect_estimator("binary", "hormon", "status",
  offset.name = "rfstime", effect_measure = "IRR")
fn_adj <- make_effect_estimator("binary", "hormon", "status",
  offset.name = "rfstime", effect_measure = "IRR",
  adjust_covariates = "ips_covar")

# ITT
res_u <- fn_unadj(df_ps)
res_a <- fn_adj(df_ps)
cat("ITT Estimates:\n")
cat("  Unadjusted IRR:", round(exp(res_u$estimate), 3), "\n")
cat("  DR-adjusted IRR:", round(exp(res_a$estimate), 3),
    " (IPS covariate, Bang & Robins 2005)\n")

# Subgroup {er <= 0}
sg <- df_ps[df_ps$er <= 0, ]
res_u_sg <- fn_unadj(sg)
res_a_sg <- fn_adj(sg)
cat("\nSubgroup {er <= 0}:\n")
cat("  Unadjusted IRR:", round(exp(res_u_sg$estimate), 3), "\n")
cat("  DR-adjusted IRR:", round(exp(res_a_sg$estimate), 3), "\n")

# Complement {er > 0}
co <- df_ps[df_ps$er > 0, ]
res_u_co <- fn_unadj(co)
res_a_co <- fn_adj(co)
cat("\nComplement {er > 0}:\n")
cat("  Unadjusted IRR:", round(exp(res_u_co$estimate), 3), "\n")
cat("  DR-adjusted IRR:", round(exp(res_a_co$estimate), 3), "\n")
```

::: {.callout-note}
In the GBSG RCT, PS adjustment has minimal impact because treatment
was randomized.  The benefit of PS adjustment is most apparent in
observational data where confounding is present.
:::


# Simulation: Confounded Treatment Assignment {#sec-simulation}

To demonstrate the value of PS adjustment, we simulate an observational
study where treatment assignment depends on covariates.

## Data-Generating Mechanism

```{r sim-dgm}
#| code-fold: false

simulate_confounded_trial <- function(n = 2000, seed = 42L) {
  set.seed(seed)

  # Covariates
  x1 <- rbinom(n, 1, 0.5)
  x2 <- rbinom(n, 1, 0.5)
  x3 <- rnorm(n)
  x4 <- rnorm(n)

  # Confounded treatment: P(W=1) depends on x1 and x3
  ps_true <- plogis(-0.5 + 0.8 * x1 + 0.6 * x3)
  treat   <- rbinom(n, 1, ps_true)

  # True subgroup: x1 == 1
  in_sg <- x1 == 1

  # Outcome: binary event with treatment-subgroup interaction
  # x3 is both a confounder AND a prognostic factor
  lp <- -1.5 + 0.5 * x3 + 0.3 * x4 +
        ifelse(in_sg, 0.8, -0.4) * treat
  event <- rbinom(n, 1, plogis(lp))

  # Follow-up time (for Poisson-offset)
  tte <- rexp(n, rate = 0.05) + 1

  data.frame(
    id = seq_len(n),
    x1 = x1, x2 = x2, x3 = x3, x4 = x4,
    treat = treat, event = event, tte = tte,
    in_sg = in_sg, ps_true = ps_true
  )
}

df_sim <- simulate_confounded_trial(n = 2000, seed = 42L)

cat("Simulated data:\n")
cat("  N =", nrow(df_sim), "\n")
cat("  P(treat=1):", round(mean(df_sim$treat), 3), "\n")
cat("  Event rate:", round(mean(df_sim$event), 3), "\n")
cat("  True subgroup {x1=1}: n =", sum(df_sim$in_sg),
    sprintf("(%.0f%%)\n", 100 * mean(df_sim$in_sg)))

cat("\nTrue PS range:", round(range(df_sim$ps_true), 3), "\n")
cat("  Mean PS (in_sg=TRUE):", round(mean(df_sim$ps_true[df_sim$in_sg]), 3), "\n")
cat("  Mean PS (in_sg=FALSE):", round(mean(df_sim$ps_true[!df_sim$in_sg]), 3), "\n")
```

## PS Estimation Quality

```{r sim-ps-quality, fig.width=10, fig.height=4}
confs_sim <- c("x1", "x2", "x3", "x4")

ps_sim_grf   <- estimate_propensity_scores(df_sim, "treat", confs_sim, method = "grf")
ps_sim_lasso <- estimate_propensity_scores(df_sim, "treat", confs_sim, method = "lasso")
ps_sim_logit <- estimate_propensity_scores(df_sim, "treat", confs_sim, method = "logistic")

cat("PS estimation quality (correlation with true PS):\n")
cat("  GRF:     ", round(cor(ps_sim_grf$ps_hat, df_sim$ps_true), 3), "\n")
cat("  LASSO:   ", round(cor(ps_sim_lasso$ps_hat, df_sim$ps_true), 3), "\n")
cat("  Logistic:", round(cor(ps_sim_logit$ps_hat, df_sim$ps_true), 3), "\n")

par(mfrow = c(1, 3))
plot(df_sim$ps_true, ps_sim_grf$ps_hat, pch = 16, cex = 0.3, col = "steelblue",
     xlab = "True PS", ylab = "Estimated PS", main = "GRF")
abline(0, 1, col = "red", lwd = 2)
plot(df_sim$ps_true, ps_sim_lasso$ps_hat, pch = 16, cex = 0.3, col = "forestgreen",
     xlab = "True PS", ylab = "Estimated PS", main = "LASSO")
abline(0, 1, col = "red", lwd = 2)
plot(df_sim$ps_true, ps_sim_logit$ps_hat, pch = 16, cex = 0.3, col = "darkorange",
     xlab = "True PS", ylab = "Estimated PS", main = "Logistic")
abline(0, 1, col = "red", lwd = 2)
par(mfrow = c(1, 1))
```

## Bias Comparison: Unadjusted vs PS-Adjusted

```{r sim-bias}
# True treatment effects (from the DGM)
cat("True treatment effects (log-OR scale):\n")
cat("  Subgroup {x1=1}:   log-OR = 0.8  (OR = 2.23)\n")
cat("  Complement {x1=0}: log-OR = -0.4  (OR = 0.67)\n\n")

df_ps_sim <- ps_sim_grf$data

# Build estimators
fn_u <- make_effect_estimator("binary", "treat", "event",
  effect_measure = "OR")
fn_a <- make_effect_estimator("binary", "treat", "event",
  effect_measure = "OR", adjust_covariates = "ips_covar")

# ITT
cat("ITT estimates:\n")
cat("  Unadjusted OR: ", round(exp(fn_u(df_ps_sim)$estimate), 3), "\n")
cat("  PS-adjusted OR:", round(exp(fn_a(df_ps_sim)$estimate), 3), "\n\n")

# Subgroup
sg_sim <- df_ps_sim[df_ps_sim$x1 == 1, ]
co_sim <- df_ps_sim[df_ps_sim$x1 == 0, ]

cat("Subgroup {x1=1} (true OR = 2.23):\n")
cat("  Unadjusted OR: ", round(exp(fn_u(sg_sim)$estimate), 3), "\n")
cat("  PS-adjusted OR:", round(exp(fn_a(sg_sim)$estimate), 3), "\n\n")

cat("Complement {x1=0} (true OR = 0.67):\n")
cat("  Unadjusted OR: ", round(exp(fn_u(co_sim)$estimate), 3), "\n")
cat("  PS-adjusted OR:", round(exp(fn_a(co_sim)$estimate), 3), "\n")
```


## Monte Carlo Evaluation

```{r sim-mc, fig.width=10, fig.height=5}
n_sims <- 100

results <- data.frame(
  sim = integer(), method = character(),
  sg_est = numeric(), co_est = numeric(),
  stringsAsFactors = FALSE
)

for (s in seq_len(n_sims)) {
  df_s <- simulate_confounded_trial(n = 2000, seed = s)
  ps_s <- estimate_propensity_scores(df_s, "treat", confs_sim, method = "grf")
  df_s <- ps_s$data

  fn_u_s <- make_effect_estimator("binary", "treat", "event",
    effect_measure = "OR")
  fn_a_s <- make_effect_estimator("binary", "treat", "event",
    effect_measure = "OR", adjust_covariates = "ips_covar")

  sg_s <- df_s[df_s$x1 == 1, ]
  co_s <- df_s[df_s$x1 == 0, ]

  res_u_sg <- fn_u_s(sg_s)
  res_a_sg <- fn_a_s(sg_s)
  res_u_co <- fn_u_s(co_s)
  res_a_co <- fn_a_s(co_s)

  results <- rbind(results, data.frame(
    sim = s, method = "Unadjusted",
    sg_est = res_u_sg$estimate, co_est = res_u_co$estimate
  ))
  results <- rbind(results, data.frame(
    sim = s, method = "PS-adjusted",
    sg_est = res_a_sg$estimate, co_est = res_a_co$estimate
  ))
}

# Summary statistics
cat("=== Monte Carlo Results (", n_sims, "simulations) ===\n\n")
cat("Subgroup {x1=1} (true log-OR = 0.80):\n")
unadj_sg <- results$sg_est[results$method == "Unadjusted"]
adj_sg   <- results$sg_est[results$method == "PS-adjusted"]
cat(sprintf("  Unadjusted: mean = %.3f, bias = %.3f, RMSE = %.3f\n",
    mean(unadj_sg, na.rm = TRUE),
    mean(unadj_sg, na.rm = TRUE) - 0.8,
    sqrt(mean((unadj_sg - 0.8)^2, na.rm = TRUE))))
cat(sprintf("  PS-adjusted: mean = %.3f, bias = %.3f, RMSE = %.3f\n",
    mean(adj_sg, na.rm = TRUE),
    mean(adj_sg, na.rm = TRUE) - 0.8,
    sqrt(mean((adj_sg - 0.8)^2, na.rm = TRUE))))

cat("\nComplement {x1=0} (true log-OR = -0.40):\n")
unadj_co <- results$co_est[results$method == "Unadjusted"]
adj_co   <- results$co_est[results$method == "PS-adjusted"]
cat(sprintf("  Unadjusted: mean = %.3f, bias = %.3f, RMSE = %.3f\n",
    mean(unadj_co, na.rm = TRUE),
    mean(unadj_co, na.rm = TRUE) - (-0.4),
    sqrt(mean((unadj_co - (-0.4))^2, na.rm = TRUE))))
cat(sprintf("  PS-adjusted: mean = %.3f, bias = %.3f, RMSE = %.3f\n",
    mean(adj_co, na.rm = TRUE),
    mean(adj_co, na.rm = TRUE) - (-0.4),
    sqrt(mean((adj_co - (-0.4))^2, na.rm = TRUE))))

# Boxplots
par(mfrow = c(1, 2))
boxplot(sg_est ~ method, data = results,
        main = "Subgroup {x1=1}\n(true log-OR = 0.80)",
        ylab = "log-OR", col = c("salmon", "lightblue"))
abline(h = 0.8, col = "red", lwd = 2, lty = 2)

boxplot(co_est ~ method, data = results,
        main = "Complement {x1=0}\n(true log-OR = -0.40)",
        ylab = "log-OR", col = c("salmon", "lightblue"))
abline(h = -0.4, col = "red", lwd = 2, lty = 2)
par(mfrow = c(1, 1))
```


# Integration with ForestSearch {#sec-integration}

## API Usage

```r
# RCT (default: no PS adjustment)
forestsearch(df, outcome_type = "binary", effect_measure = "IRR",
             dmin.grf = 0.0, ...)

# Observational with GRF-based PS (default when is.RCT = FALSE)
forestsearch(df, outcome_type = "binary", effect_measure = "IRR",
             dmin.grf = 0.0, is.RCT = FALSE, ps_method = "grf", ...)

# Observational with LASSO PS
forestsearch(df, outcome_type = "binary", is.RCT = FALSE,
             ps_method = "lasso", ...)

# User-supplied PS
forestsearch(df, outcome_type = "binary", is.RCT = FALSE,
             ps_hat = my_ps_vector, ...)
```

## Key Design Decisions

1. **IPS covariate, not weight.**  Following Bang & Robins (2005), the
   inverse propensity score is included as an adjustment covariate
   (`Y ~ treat + ips_covar`) rather than as IPTW.  The IPS covariate
   takes the value $1/\hat{e}(X)$ for treated and $1/(1-\hat{e}(X))$
   for controls.  This gives doubly-robust estimates within each
   subgroup slice and avoids extreme weight instability in small
   subgroups.

2. **Full-sample PS used within subgroups.**  Propensity scores are
   estimated once on the full sample and carried into subgroup slices
   as-is.  The PS is a design quantity reflecting the treatment
   assignment mechanism — re-estimating within slices would be
   methodologically incorrect and statistically unstable.

3. **GRF PS from the screening step.**  When `use_grf = TRUE`, the
   causal forest already estimates `W.hat`.  The `grf.subg.harm.glm()`
   function now returns the fitted forest object, enabling PS extraction
   without re-estimation.

4. **Trimming.**  PS values are trimmed to [2.5%, 97.5%] before computing
   the IPS covariate, preventing extreme values from near-zero
   denominators in subgroup slices where one treatment arm dominates.

5. **Backward compatibility.**  For `is.RCT = TRUE` (default), `ps_method`
   resolves to `"none"` and the estimator closure is identical to the
   unadjusted version.  Existing analyses are unaffected.


# Session Information

```{r session-info}
sessionInfo()
```
