---
title: "QC Simulation: Cox PH vs Poisson IRR Approximation"
subtitle: "Validating the GLM Extension via Operating Characteristics Comparison"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# Number of simulations (increase for production runs)
nsims_alt  <- 100L
nsims_null <- 100L
```


# Introduction {#sec-intro}

This document validates the GLM extension by comparing two analysis
pipelines on *identical* simulated datasets drawn from the same
survival-based DGM:
\
- **Cox PH** (the established survival pipeline via
  `run_simulation_analysis()`)
- **Poisson IRR** (the GLM extension using `forestsearch()` with
  `outcome_type = "binary"`, `effect_measure = "IRR"`,
  `offset.name = "y_sim"`)

The Poisson regression with a log(time) offset approximates the Cox
proportional hazards model.  Under this approximation, the incidence
rate ratio (IRR) targets the same quantity as the hazard ratio (HR).
If the GLM pipeline is correctly implemented, the two approaches should
produce comparable subgroup identification rates, classification
metrics, and effect estimates.

**Validation targets:**

| Metric | Expectation |
|---|---|
| Subgroup identification (power) | Similar |
| Sensitivity, Specificity, PPV, NPV | Similar |
| IRR estimates vs HR estimates | Numerically close |
| Type I error (null) | Comparable |


# Setup {#sec-setup}

```{r load-packages}
library(forestsearch)
library(weightedsurv)
library(data.table)
library(survival)
library(ggplot2)
library(gt)
library(foreach)
library(doFuture)
library(future)
```

```{r timing-init}
timings <- list()
t_vignette_start <- proc.time()
```

```{r setup-parallel}
n_workers <- min(parallel::detectCores() - 1, 120)
plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")
```


# Data Generating Mechanism {#sec-dgm}

We use the same GBSG-based DGM from León et al. (2024).  The DGM
generates survival data with treatment effect heterogeneity: a harm
subgroup H = {ER-low AND premenopausal} has HR > 1, while the
complement Hc benefits from treatment (HR < 1).

## Alternative Hypothesis DGM

```{r create-dgm-alt}
t0 <- proc.time()[3]

dgm_alt <- setup_gbsg_dgm(
  model = "alt",
  k_treat = 1.0,
  k_inter = 2.0,
  k_z3 = 1.0,
  z1_quantile = 0.25,
  n_super = 5000,
  cens_type = "weibull",
  seed = 8316951,
  verbose = TRUE
)

dgm_alt <- compute_dgm_cde(dgm_alt)
print(dgm_alt)

timings$dgm_creation <- proc.time()[3] - t0
```

## Calibrate for Target HR = 2.0 in H

```{r calibrate}
t0 <- proc.time()[3]

k_inter_cox <- calibrate_k_inter(
  target_hr_harm = 2.0,
  model = "alt",
  k_treat = 1.0,
  cens_type = "weibull",
  use_ahr = FALSE,
  verbose = TRUE
)

dgm_calibrated <- setup_gbsg_dgm(
  model = "alt",
  k_treat = 1.0,
  k_inter = k_inter_cox,
  verbose = TRUE
)
dgm_calibrated <- compute_dgm_cde(dgm_calibrated)

cat("\nCalibrated DGM:\n")
cat("  HR(H):", round(dgm_calibrated$hr_H_true, 3), "\n")
cat("  HR(Hc):", round(dgm_calibrated$hr_Hc_true, 3), "\n")
cat("  HR(overall):", round(dgm_calibrated$hr_causal, 3), "\n")
cat("  AHR(H):", round(dgm_calibrated$AHR_H_true, 3), "\n")
cat("  AHR(Hc):", round(dgm_calibrated$AHR_Hc_true, 3), "\n")

timings$calibration <- proc.time()[3] - t0
```

## Null Hypothesis DGM

```{r create-dgm-null}
t0 <- proc.time()[3]

dgm_null <- setup_gbsg_dgm(
  model = "null",
  k_treat = 1.0,
  verbose = TRUE
)

cat("\nNull DGM:\n")
cat("  HR(overall):", round(dgm_null$hr_causal, 3), "\n")
cat("  AHR:", round(dgm_null$AHR, 3), "\n")

timings$dgm_null <- proc.time()[3] - t0
```


# Single-Trial Comparison {#sec-single}

Before running the full simulation, verify the two pipelines on a
single trial to confirm alignment.

```{r single-trial}
sim_data <- simulate_from_dgm(
  dgm = dgm_calibrated,
  n = 700,
  rand_ratio = 1,
  seed = 1,
  analysis_time = 84,
  cens_adjust = log(1.5)
)

cat("Simulated trial: N =", nrow(sim_data),
    ", Events =", sum(sim_data$event_sim),
    "(", round(100 * mean(sim_data$event_sim), 1), "%)\n")

# Add required id column
sim_data$id <- seq_len(nrow(sim_data))
```

## Cox PH on Single Trial

```{r single-cox}
fit_cox <- coxph(Surv(y_sim, event_sim) ~ treat_sim, data = sim_data)
cat("Cox ITT HR:", round(exp(coef(fit_cox)), 3), "\n")
```

## Poisson IRR on Single Trial

```{r single-poisson}
fit_pois <- glm(event_sim ~ treat_sim + offset(log(y_sim)),
                family = poisson, data = sim_data)
cat("Poisson ITT IRR:", round(exp(coef(fit_pois)["treat_sim"]), 3), "\n")

cat("\nComparison: Cox HR =", round(exp(coef(fit_cox)), 3),
    "vs Poisson IRR =", round(exp(coef(fit_pois)["treat_sim"]), 3), "\n")
```

## ForestSearch: Cox vs Poisson on Single Trial

```{r single-fs-cox}
confounders <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")

fs_cox <- forestsearch(
  df.analysis = sim_data, confounders.name = confounders,
  outcome.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id",
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.90, fs.splits = 400,
  n.min = 60, maxk = 2,
  use_twostage = TRUE, use_lasso = TRUE, use_grf = FALSE,
  parallel_args = list(plan = "sequential", workers = 1,
                       show_message = FALSE)
)
cat("Cox FS SG:", paste(fs_cox$sg.harm, collapse = " & "), "\n")
```

```{r single-fs-irr}
fs_irr <- forestsearch(
  df.analysis = sim_data, confounders.name = confounders,
  outcome.name = "event_sim", event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id",
  outcome_type = "binary", effect_measure = "IRR",
  offset.name = "y_sim", dmin.grf = 0.0,
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.90, fs.splits = 400,
  n.min = 60, maxk = 2,
  use_twostage = TRUE, use_lasso = TRUE, use_grf = FALSE,
  parallel_args = list(plan = "sequential", workers = 1,
                       show_message = FALSE)
)
cat("IRR FS SG:", paste(fs_irr$sg.harm, collapse = " & "), "\n")
```


# Simulation Parameters {#sec-params}

```{r define-params}
sim_config_alt <- list(
  n_sims = nsims_alt,
  n_sample = 700,
  analysis_time = 84,
  seed_base = 8316951,
  cens_adjust = log(1.5)
)

sim_config_null <- list(
  n_sims = nsims_null,
  n_sample = 700,
  analysis_time = 84,
  seed_base = 8316951,
  cens_adjust = log(1.5)
)

confounders_base <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")

# Shared ForestSearch parameters (common to both pipelines)
fs_params_base <- list(
  hr.threshold = 1.25,
  hr.consistency = 1.0,
  pconsistency.threshold = 0.90,
  fs.splits = 400,
  n.min = 60,
  d0.min = 12,
  d1.min = 12,
  maxk = 2,
  by.risk = 12,
  vi.grf.min = -0.2,
  use_lasso = TRUE,
  use_grf = TRUE,
  use_twostage = TRUE,
  twostage_args = list()
)

# Cox-specific params
fs_params_cox <- modifyList(fs_params_base, list(
  outcome.name = "y_sim",
  event.name = "event_sim",
  treat.name = "treat_sim",
  id.name = "id"
))

# IRR-specific params (binary + offset)
fs_params_irr <- modifyList(fs_params_base, list(
  outcome.name = "event_sim",
  event.name = "event_sim",
  treat.name = "treat_sim",
  id.name = "id",
  outcome_type = "binary",
  effect_measure = "IRR",
  offset.name = "y_sim",
  dmin.grf = 0.0
))
```


# GLM Simulation Analysis Helper {#sec-helper}

Since `run_simulation_analysis()` does not yet support GLM parameters,
we define a helper that runs the Poisson IRR pipeline and extracts
results in a compatible `data.table` format.

```{r glm-helper}
run_glm_sim_analysis <- function(
    sim_id, dgm, n_sample, analysis_time = Inf,
    cens_adjust = 0, confounders_base, fs_params,
    seed_base = 8316951L, verbose = FALSE
) {
  # Simulate data from the survival DGM

  sim_data <- simulate_from_dgm(
    dgm = dgm, n = n_sample,
    seed = seed_base + sim_id,
    analysis_time = analysis_time,
    cens_adjust = cens_adjust
  )
  sim_data$id <- seq_len(nrow(sim_data))

  outcome_name <- fs_params$outcome.name
  event_name   <- fs_params$event.name
  treat_name   <- fs_params$treat.name
  offset_name  <- fs_params$offset.name

  # True subgroup properties
  has_harm <- "flag_harm" %in% names(sim_data)
  size_H_true  <- if (has_harm) sum(sim_data$flag_harm) else NA_integer_
  prop_H_true  <- if (has_harm) mean(sim_data$flag_harm) else NA_real_

  # Run ForestSearch with GLM parameters
  fs_args <- list(
    df.analysis      = sim_data,
    confounders.name = confounders_base,
    outcome.name     = outcome_name,
    event.name       = event_name,
    treat.name       = treat_name,
    id.name          = "id",
    outcome_type     = fs_params$outcome_type,
    effect_measure   = fs_params$effect_measure,
    offset.name      = offset_name,
    dmin.grf         = fs_params$dmin.grf,
    use_lasso        = if (!is.null(fs_params$use_lasso)) fs_params$use_lasso else TRUE,
    use_grf          = if (!is.null(fs_params$use_grf)) fs_params$use_grf else FALSE,
    hr.threshold     = fs_params$hr.threshold,
    hr.consistency   = fs_params$hr.consistency,
    pconsistency.threshold = fs_params$pconsistency.threshold,
    fs.splits        = fs_params$fs.splits,
    n.min            = fs_params$n.min,
    d0.min           = fs_params$d0.min,
    d1.min           = fs_params$d1.min,
    maxk             = fs_params$maxk,
    use_twostage     = fs_params$use_twostage,
    twostage_args    = fs_params$twostage_args,
    parallel_args    = list(plan = "sequential", workers = 1,
                            show_message = FALSE),
    details          = FALSE,
    plot.sg          = FALSE
  )

  fs_result <- tryCatch(
    do.call(forestsearch, fs_args),
    error = function(e) { warning("FS-IRR failed sim ", sim_id, ": ", e$message); NULL }
  )

  # Initialize output
  out <- data.table::data.table(
    sim         = sim_id,
    analysis    = "FS-IRR",
    any.H       = 0L,
    sizeH_true  = size_H_true,
    propH_true  = prop_H_true,
    size.H      = NA_integer_,
    size.Hc     = nrow(sim_data),
    hr.H.hat    = NA_real_,
    hr.Hc.hat   = NA_real_,
    hr.itt      = NA_real_,
    ppv = NA_real_, npv = NA_real_,
    sens = NA_real_, spec = NA_real_,
    p.cens = 1 - mean(sim_data[[event_name]])
  )

  # ITT IRR
  out$hr.itt <- tryCatch({
    fit <- glm(
      as.formula(paste(event_name, "~", treat_name,
                       "+ offset(log(", offset_name, "))")),
      family = poisson, data = sim_data
    )
    exp(coef(fit)[treat_name])
  }, error = function(e) NA_real_)

  # Check if subgroup was found
  has_result <- !is.null(fs_result) && !is.null(fs_result$sg.harm)

  if (!has_result) {
    out$hr.Hc.hat <- out$hr.itt
    return(out)
  }

  out$any.H <- 1L

  # Extract subgroup indicator
  df_est <- fs_result$df.est
  if (!is.null(df_est) && "treat.recommend" %in% names(df_est)) {
    sim_data$sg_hat <- as.integer(df_est$treat.recommend == 0)
  } else {
    out$hr.Hc.hat <- out$hr.itt
    return(out)
  }

  out$size.H  <- sum(sim_data$sg_hat, na.rm = TRUE)
  out$size.Hc <- sum(sim_data$sg_hat == 0L, na.rm = TRUE)

  # IRR in each subgroup
  .poisson_irr <- function(sub_df) {
    tryCatch({
      fit <- glm(
        as.formula(paste(event_name, "~", treat_name,
                         "+ offset(log(", offset_name, "))")),
        family = poisson, data = sub_df
      )
      exp(coef(fit)[treat_name])
    }, error = function(e) NA_real_)
  }

  if (out$size.H > 10)
    out$hr.H.hat <- .poisson_irr(subset(sim_data, sg_hat == 1))
  if (out$size.Hc > 10)
    out$hr.Hc.hat <- .poisson_irr(subset(sim_data, sg_hat == 0))

  # Classification metrics vs true subgroup
  if (has_harm) {
    true_H <- sim_data$flag_harm == 1L
    pred_H <- sim_data$sg_hat == 1L
    tp <- sum(true_H & pred_H)
    fp <- sum(!true_H & pred_H)
    fn <- sum(true_H & !pred_H)
    tn <- sum(!true_H & !pred_H)
    out$sens <- tp / max(tp + fn, 1)
    out$spec <- tn / max(tn + fp, 1)
    out$ppv  <- tp / max(tp + fp, 1)
    out$npv  <- tn / max(tn + fn, 1)
  }

  return(out)
}
```


# Alternative Hypothesis Simulations {#sec-alt}

## Cox PH Pipeline (Established)

```{r run-alt-cox}
cat("Running", sim_config_alt$n_sims,
    "Cox simulations under H1...\n")

t0 <- proc.time()[3]

results_alt_cox <- foreach(
  sim = seq_len(sim_config_alt$n_sims),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table"),
    seed = TRUE
  )
) %dofuture% {
  run_simulation_analysis(
    sim_id = sim,
    dgm = dgm_calibrated,
    n_sample = sim_config_alt$n_sample,
    analysis_time = sim_config_alt$analysis_time,
    cens_adjust = sim_config_alt$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE,
    run_fs_grf = FALSE,
    run_grf = FALSE,
    fs_params = fs_params_cox,
    verbose = FALSE
  )
}

timings$alt_cox <- proc.time()[3] - t0
cat("Cox H1 completed in", round(timings$alt_cox / 60, 1), "minutes\n")
cat("Results:", nrow(results_alt_cox), "rows\n")
```

## Poisson IRR Pipeline (GLM Extension)

```{r run-alt-irr}
cat("Running", sim_config_alt$n_sims,
    "Poisson IRR simulations under H1...\n")

t0 <- proc.time()[3]

results_alt_irr <- foreach(
  sim = seq_len(sim_config_alt$n_sims),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "data.table"),
    seed = TRUE
  )
) %dofuture% {
  run_glm_sim_analysis(
    sim_id = sim,
    dgm = dgm_calibrated,
    n_sample = sim_config_alt$n_sample,
    analysis_time = sim_config_alt$analysis_time,
    cens_adjust = sim_config_alt$cens_adjust,
    confounders_base = confounders_base,
    fs_params = fs_params_irr
  )
}

timings$alt_irr <- proc.time()[3] - t0
cat("IRR H1 completed in", round(timings$alt_irr / 60, 1), "minutes\n")
cat("Results:", nrow(results_alt_irr), "rows\n")
```


# Null Hypothesis Simulations {#sec-null}

## Cox PH Pipeline

```{r run-null-cox}
cat("Running", sim_config_null$n_sims,
    "Cox simulations under H0...\n")

t0 <- proc.time()[3]

results_null_cox <- foreach(
  sim = seq_len(sim_config_null$n_sims),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table"),
    seed = TRUE
  )
) %dofuture% {
  run_simulation_analysis(
    sim_id = sim,
    dgm = dgm_null,
    n_sample = sim_config_null$n_sample,
    analysis_time = sim_config_null$analysis_time,
    cens_adjust = sim_config_null$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE,
    run_fs_grf = FALSE,
    run_grf = FALSE,
    fs_params = fs_params_cox,
    verbose = FALSE
  )
}

timings$null_cox <- proc.time()[3] - t0
cat("Cox H0 completed in", round(timings$null_cox / 60, 1), "minutes\n")
```

## Poisson IRR Pipeline

```{r run-null-irr}
cat("Running", sim_config_null$n_sims,
    "Poisson IRR simulations under H0...\n")

t0 <- proc.time()[3]

results_null_irr <- foreach(
  sim = seq_len(sim_config_null$n_sims),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "data.table"),
    seed = TRUE
  )
) %dofuture% {
  run_glm_sim_analysis(
    sim_id = sim,
    dgm = dgm_null,
    n_sample = sim_config_null$n_sample,
    analysis_time = sim_config_null$analysis_time,
    cens_adjust = sim_config_null$cens_adjust,
    confounders_base = confounders_base,
    fs_params = fs_params_irr
  )
}

timings$null_irr <- proc.time()[3] - t0
cat("IRR H0 completed in", round(timings$null_irr / 60, 1), "minutes\n")
```


# Results: Alternative Hypothesis {#sec-results-alt}

## Cox PH Operating Characteristics

```{r oc-alt-cox}
format_oc_results(
  results_alt_cox,
  metrics = c("detection", "classification", "hr_estimates"),
  digits = 2, digits_hr = 2,
  title = "Cox PH: Identification Properties (H1)",
  subtitle = sprintf("n = %d, %d simulations, HR(overall) = %.2f",
                      sim_config_alt$n_sample, sim_config_alt$n_sims,
                      dgm_calibrated$hr_causal)
)
```

```{r est-alt-cox}
build_estimation_table(
  results = results_alt_cox,
  dgm = dgm_calibrated,
  analysis_method = "FS",
  subtitle = sprintf("Cox PH: n = %d, %d sims",
                      sim_config_alt$n_sample, sim_config_alt$n_sims),
  font_size = 14
)
```

## Poisson IRR Operating Characteristics

```{r oc-alt-irr}
# Use format_oc_results with the IRR results
# Note: hr.H.hat / hr.Hc.hat columns contain IRR estimates
format_oc_results(
  results_alt_irr,
  metrics = c("detection", "classification", "hr_estimates"),
  digits = 2, digits_hr = 2,
  title = "Poisson IRR: Identification Properties (H1)",
  subtitle = sprintf("n = %d, %d simulations, HR(overall) = %.2f",
                      sim_config_alt$n_sample, sim_config_alt$n_sims,
                      dgm_calibrated$hr_causal)
)
```

## Side-by-Side Comparison (H1)

```{r compare-alt}
# Extract key metrics for comparison
compare_metrics <- function(res_cox, res_irr, label_cox = "Cox PH",
                            label_irr = "Poisson IRR") {
  # Cox results (FS method only)
  cox <- res_cox[res_cox$analysis == "FS", ]
  irr <- res_irr

  cox_found <- cox[cox$any.H == 1, ]
  irr_found <- irr[irr$any.H == 1, ]

  data.frame(
    Metric = c("Detection Rate", "Sensitivity", "Specificity", "PPV", "NPV",
               "Mean HR/IRR (H)", "Mean HR/IRR (Hc)", "Mean HR/IRR (ITT)",
               "Mean SG Size (H)"),
    Cox_PH = c(
      round(mean(cox$any.H), 3),
      round(mean(cox$sens, na.rm = TRUE), 3),
      round(mean(cox$spec, na.rm = TRUE), 3),
      round(mean(cox$ppv, na.rm = TRUE), 3),
      round(mean(cox$npv, na.rm = TRUE), 3),
      round(mean(cox_found$hr.H.hat, na.rm = TRUE), 3),
      round(mean(cox_found$hr.Hc.hat, na.rm = TRUE), 3),
      round(mean(cox$hr.itt, na.rm = TRUE), 3),
      round(mean(cox_found$size.H, na.rm = TRUE), 0)
    ),
    Poisson_IRR = c(
      round(mean(irr$any.H), 3),
      round(mean(irr$sens, na.rm = TRUE), 3),
      round(mean(irr$spec, na.rm = TRUE), 3),
      round(mean(irr$ppv, na.rm = TRUE), 3),
      round(mean(irr$npv, na.rm = TRUE), 3),
      round(mean(irr_found$hr.H.hat, na.rm = TRUE), 3),
      round(mean(irr_found$hr.Hc.hat, na.rm = TRUE), 3),
      round(mean(irr$hr.itt, na.rm = TRUE), 3),
      round(mean(irr_found$size.H, na.rm = TRUE), 0)
    ),
    stringsAsFactors = FALSE
  )
}

comp_alt <- compare_metrics(results_alt_cox, results_alt_irr)

gt(comp_alt) |>
  tab_header(
    title = md("**H1 Comparison: Cox PH vs Poisson IRR**"),
    subtitle = sprintf("n = %d, %d simulations, true HR(H) = %.2f",
                        sim_config_alt$n_sample, sim_config_alt$n_sims,
                        dgm_calibrated$hr_H_true)
  ) |>
  cols_label(
    Metric = "Metric",
    Cox_PH = "Cox PH (HR)",
    Poisson_IRR = "Poisson IRR"
  ) |>
  tab_source_note(
    md(paste0("**True values:** HR(H) = ",
              round(dgm_calibrated$hr_H_true, 3),
              ", HR(Hc) = ",
              round(dgm_calibrated$hr_Hc_true, 3)))
  )
```


# Results: Null Hypothesis {#sec-results-null}

## Side-by-Side Comparison (H0)

```{r compare-null}
comp_null <- compare_metrics(results_null_cox, results_null_irr)

gt(comp_null) |>
  tab_header(
    title = md("**H0 Comparison: Cox PH vs Poisson IRR (Type I Error)**"),
    subtitle = sprintf("n = %d, %d simulations, true HR = %.2f (uniform)",
                        sim_config_null$n_sample, sim_config_null$n_sims,
                        dgm_null$hr_causal)
  ) |>
  cols_label(
    Metric = "Metric",
    Cox_PH = "Cox PH (HR)",
    Poisson_IRR = "Poisson IRR"
  )
```


# Per-Simulation Concordance {#sec-concordance}

Since both pipelines analyze the same DGM (and with the same seed
structure), we can assess how often they agree on whether a subgroup
is identified.

```{r concordance}
# Match simulations by sim_id
cox_detect <- results_alt_cox[results_alt_cox$analysis == "FS",
                              c("sim", "any.H")]
names(cox_detect) <- c("sim", "cox_found")

irr_detect <- results_alt_irr[, c("sim", "any.H")]
names(irr_detect) <- c("sim", "irr_found")

concordance <- merge(cox_detect, irr_detect, by = "sim")

# Agreement matrix
agree_both   <- sum(concordance$cox_found == 1 & concordance$irr_found == 1)
agree_neither <- sum(concordance$cox_found == 0 & concordance$irr_found == 0)
cox_only     <- sum(concordance$cox_found == 1 & concordance$irr_found == 0)
irr_only     <- sum(concordance$cox_found == 0 & concordance$irr_found == 1)
n_matched    <- nrow(concordance)

cat("=== Per-Simulation Concordance ===\n")
cat("Matched simulations:", n_matched, "\n")
cat("Both found SG:      ", agree_both, sprintf("(%.0f%%)\n", 100 * agree_both / n_matched))
cat("Neither found SG:   ", agree_neither, sprintf("(%.0f%%)\n", 100 * agree_neither / n_matched))
cat("Cox only:           ", cox_only, sprintf("(%.0f%%)\n", 100 * cox_only / n_matched))
cat("IRR only:           ", irr_only, sprintf("(%.0f%%)\n", 100 * irr_only / n_matched))
cat("Agreement rate:     ", round(100 * (agree_both + agree_neither) / n_matched, 1), "%\n")
```


# Visualization {#sec-viz}

## Detection Rate Comparison

```{r viz-detection, fig.width=7, fig.height=4}
det_df <- data.frame(
  Pipeline = c("Cox PH", "Poisson IRR"),
  H1_Power = c(
    mean(results_alt_cox[results_alt_cox$analysis == "FS", ]$any.H),
    mean(results_alt_irr$any.H)
  ),
  H0_TypeI = c(
    mean(results_null_cox[results_null_cox$analysis == "FS", ]$any.H),
    mean(results_null_irr$any.H)
  )
)

det_long <- data.frame(
  Pipeline = rep(det_df$Pipeline, 2),
  Hypothesis = rep(c("H1 (Power)", "H0 (Type I)"), each = 2),
  Rate = c(det_df$H1_Power, det_df$H0_TypeI)
)

ggplot(det_long, aes(x = Pipeline, y = Rate, fill = Pipeline)) +
  geom_col(width = 0.5) +
  facet_wrap(~ Hypothesis, scales = "free_y") +
  geom_text(aes(label = sprintf("%.3f", Rate)), vjust = -0.5, size = 3.5) +
  scale_fill_manual(values = c("Cox PH" = "steelblue", "Poisson IRR" = "coral")) +
  labs(title = "Detection Rates: Cox PH vs Poisson IRR",
       y = "Rate", x = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

## ITT Estimate Comparison

```{r viz-itt, fig.width=7, fig.height=5}
cox_itt <- results_alt_cox[results_alt_cox$analysis == "FS", ]$hr.itt
irr_itt <- results_alt_irr$hr.itt

itt_df <- data.frame(
  Cox_HR = cox_itt[seq_len(min(length(cox_itt), length(irr_itt)))],
  IRR = irr_itt[seq_len(min(length(cox_itt), length(irr_itt)))]
)

ggplot(itt_df, aes(x = Cox_HR, y = IRR)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Per-Simulation ITT Estimates: Cox HR vs Poisson IRR",
       x = "Cox HR", y = "Poisson IRR") +
  coord_equal() +
  theme_minimal(base_size = 12)
```


# Computational Timing Summary {#sec-timing}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Stage = c("DGM creation (H1)", "Calibration", "DGM creation (H0)",
            "H1 sims (Cox)", "H1 sims (IRR)",
            "H0 sims (Cox)", "H0 sims (IRR)",
            "Total"),
  Seconds = c(
    timings$dgm_creation, timings$calibration, timings$dgm_null,
    timings$alt_cox, timings$alt_irr,
    timings$null_cox, timings$null_irr,
    timings$total
  ),
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(
    title = "Computational Timing Summary",
    subtitle = sprintf("%d H1 + %d H0 sims per pipeline, %d workers",
                        sim_config_alt$n_sims, sim_config_null$n_sims,
                        n_workers)
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)", Minutes = "Time (min)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total")
  )
```


# Summary {#sec-summary}

This QC simulation validates the Poisson IRR approximation to the Cox
PH model within the ForestSearch framework.

**Key validation points:**

| Aspect | Expected | Observed |
|---|---|---|
| ITT estimates | IRR approximates HR | Check scatter plot above |
| Detection rate (H1) | Similar power | Check comparison table |
| Classification metrics | Comparable sens/spec/PPV/NPV | Check comparison table |
| Type I error (H0) | Comparable | Check null comparison |
| Per-simulation concordance | High agreement | Check concordance matrix |

**Interpretation:**

The Poisson regression with log(time) offset approximates the Cox PH model
when the event rate is not too high (the rare-events approximation).
In the GBSG setting with ~50% event rate, some divergence is expected.
The key question is whether ForestSearch's subgroup identification
machinery produces comparable operating characteristics under both
approaches, which the tables and concordance analysis above address.


# Session Information {#sec-session}

```{r session-info}
plan(sequential)
sessionInfo()
```
