---
title: "QC Simulation: Cox PH vs Poisson IRR Approximation"
subtitle: "Validating the GLM Extension via Operating Characteristics Comparison"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# Number of simulations (increase for production runs)
nsims_alt  <- 150L
nsims_null <- 150L
```


# Introduction {#sec-intro}

This document validates the GLM extension by comparing two analysis
pipelines on *identical* simulated datasets drawn from the same
survival-based DGM:
\
- **Cox PH** (the established survival pipeline via
  `run_simulation_analysis()`)
- **Poisson IRR** (the GLM extension using `forestsearch()` with
  `outcome_type = "binary"`, `effect_measure = "IRR"`,
  `offset.name = "y_sim"`)

The Poisson regression with a log(time) offset approximates the Cox
proportional hazards model.  Under this approximation, the incidence
rate ratio (IRR) targets the same quantity as the hazard ratio (HR).
If the GLM pipeline is correctly implemented, the two approaches should
produce comparable subgroup identification rates, classification
metrics, and effect estimates.

**Validation targets:**

| Metric | Expectation |
|---|---|
| Subgroup identification (power) | Identical (same Cox-based forestsearch) |
| Sensitivity, Specificity, PPV, NPV | Identical (same subgroups) |
| IRR estimates vs HR estimates | Numerically close |
| Type I error (null) | Identical |


# Setup {#sec-setup}

```{r load-packages}
library(forestsearch)
library(weightedsurv)
library(data.table)
library(survival)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)
```

```{r timing-init}
timings <- list()
t_vignette_start <- proc.time()
```

```{r setup-parallel}
n_workers <- min(parallel::detectCores() - 1, 120)
plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")
```


# Data Generating Mechanism {#sec-dgm}

We use the same GBSG-based DGM from León et al. (2024).  The DGM
generates survival data with treatment effect heterogeneity: a harm
subgroup H = {ER-low AND premenopausal} has HR > 1, while the
complement Hc benefits from treatment (HR < 1).

## Alternative Hypothesis DGM

```{r create-dgm-alt}
t0 <- proc.time()[3]

dgm_alt <- setup_gbsg_dgm(
  model = "alt",
  k_treat = 1.0,
  k_inter = 2.0,
  k_z3 = 1.0,
  z1_quantile = 0.25,
  n_super = 5000,
  cens_type = "weibull",
  seed = 8316951,
  verbose = TRUE
)

dgm_alt <- compute_dgm_cde(dgm_alt)
print(dgm_alt)

timings$dgm_creation <- proc.time()[3] - t0
```

## Calibrate for Target HR = 2.0 in H

```{r calibrate}
t0 <- proc.time()[3]

k_inter_cox <- calibrate_k_inter(
  target_hr_harm = 2.0,
  model = "alt",
  k_treat = 1.0,
  cens_type = "weibull",
  use_ahr = FALSE,
  verbose = TRUE
)

dgm_calibrated <- setup_gbsg_dgm(
  model = "alt",
  k_treat = 1.0,
  k_inter = k_inter_cox,
  verbose = TRUE
)
dgm_calibrated <- compute_dgm_cde(dgm_calibrated)

cat("\nCalibrated DGM:\n")
cat("  HR(H):", round(dgm_calibrated$hr_H_true, 3), "\n")
cat("  HR(Hc):", round(dgm_calibrated$hr_Hc_true, 3), "\n")
cat("  HR(overall):", round(dgm_calibrated$hr_causal, 3), "\n")
cat("  AHR(H):", round(dgm_calibrated$AHR_H_true, 3), "\n")
cat("  AHR(Hc):", round(dgm_calibrated$AHR_Hc_true, 3), "\n")

timings$calibration <- proc.time()[3] - t0
```

## Null Hypothesis DGM

```{r create-dgm-null}
t0 <- proc.time()[3]

dgm_null <- setup_gbsg_dgm(
  model = "null",
  k_treat = 1.0,
  verbose = TRUE
)

cat("\nNull DGM:\n")
cat("  HR(overall):", round(dgm_null$hr_causal, 3), "\n")
cat("  AHR:", round(dgm_null$AHR, 3), "\n")

timings$dgm_null <- proc.time()[3] - t0
```


# Single-Trial Comparison {#sec-single}

Before running the full simulation, verify the two pipelines on a
single trial to confirm alignment.

```{r single-trial}
sim_data <- simulate_from_dgm(
  dgm = dgm_calibrated,
  n = 700,
  rand_ratio = 1,
  seed = 1,
  analysis_time = 84,
  cens_adjust = log(1.5)
)

cat("Simulated trial: N =", nrow(sim_data),
    ", Events =", sum(sim_data$event_sim),
    "(", round(100 * mean(sim_data$event_sim), 1), "%)\n")

# Add required id column
sim_data$id <- seq_len(nrow(sim_data))
```

## Cox PH on Single Trial

```{r single-cox}
fit_cox <- coxph(Surv(y_sim, event_sim) ~ treat_sim, data = sim_data)
cat("Cox ITT HR:", round(exp(coef(fit_cox)), 3), "\n")
```

## Poisson IRR on Single Trial

```{r single-poisson}
fit_pois <- glm(event_sim ~ treat_sim + offset(log(y_sim)),
                family = poisson, data = sim_data)
cat("Poisson ITT IRR:", round(exp(coef(fit_pois)["treat_sim"]), 3), "\n")

cat("\nComparison: Cox HR =", round(exp(coef(fit_cox)), 3),
    "vs Poisson IRR =", round(exp(coef(fit_pois)["treat_sim"]), 3), "\n")
```

## ForestSearch: Cox vs Poisson on Single Trial

```{r single-fs-cox}
confounders <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")

fs_cox <- forestsearch(
  df.analysis = sim_data, confounders.name = confounders,
  outcome.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id",
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.90, fs.splits = 400,
  n.min = 60, maxk = 2,
  use_twostage = TRUE, use_lasso = TRUE, use_grf = FALSE,
  parallel_args = list(plan = "multisession", workers = 1,
                       show_message = FALSE)
)
cat("Cox FS SG:", paste(fs_cox$sg.harm, collapse = " & "), "\n")
```

```{r single-fs-irr}
# Identification uses the SAME Cox PH forestsearch as the Cox pipeline.
# The IRR estimation happens downstream in .extract_fs_estimates_gen().
# Here we verify that both pipelines identify the same subgroup.
fs_irr <- forestsearch(
  df.analysis = sim_data, confounders.name = confounders,
  outcome.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id",
  hr.threshold = 1.25, hr.consistency = 1.0,
  pconsistency.threshold = 0.90, fs.splits = 400,
  n.min = 60, maxk = 2,
  use_twostage = TRUE, use_lasso = TRUE, use_grf = FALSE,
  parallel_args = list(plan = "multisession", workers = 1,
                       show_message = FALSE)
)
cat("IRR FS SG:", paste(fs_irr$sg.harm, collapse = " & "), "\n")
cat("(Same as Cox — identification is identical by design)\n")
```


# Simulation Parameters {#sec-params}

```{r define-params}
sim_config_alt <- list(
  n_sims = nsims_alt,
  n_sample = 700,
  analysis_time = 84,
  seed_base = 8316951,
  cens_adjust = log(1.5)
)

sim_config_null <- list(
  n_sims = nsims_null,
  n_sample = 700,
  analysis_time = 84,
  seed_base = 8316951,
  cens_adjust = log(1.5)
)

confounders_base <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")

# Shared ForestSearch parameters (common to both pipelines)
fs_params_base <- list(
  hr.threshold = 1.25,
  hr.consistency = 1.0,
  pconsistency.threshold = 0.90,
  fs.splits = 400,
  n.min = 60,
  d0.min = 12,
  d1.min = 12,
  maxk = 2,
  by.risk = 12,
  vi.grf.min = -0.2,
  use_lasso = TRUE,
  use_grf = TRUE,
  use_twostage = TRUE,
  twostage_args = list()
)

# Cox-specific params
fs_params_cox <- modifyList(fs_params_base, list(
  outcome.name = "y_sim",
  event.name = "event_sim",
  treat.name = "treat_sim",
  id.name = "id"
))

# IRR-specific params: forestsearch *identification* still uses Cox PH
# (outcome.name = "y_sim"), but the *estimation* step uses Poisson GLM
# (controlled by outcome_type / effect_measure / offset.name)
fs_params_irr <- modifyList(fs_params_base, list(
  outcome.name = "y_sim",
  event.name = "event_sim",
  treat.name = "treat_sim",
  id.name = "id",
  outcome_type = "binary",
  effect_measure = "IRR",
  offset.name = "y_sim",
  dmin.grf = 0.0
))
```


# Simulation Approach {#sec-approach}

Both pipelines share the same two-phase architecture inside
`run_simulation_analysis()`:

1. **Identification** — `forestsearch()` always uses Cox PH to discover
   subgroups (`outcome.name = "y_sim"`).  Both pipelines use identical
   seeds, so they identify the *same* subgroups in every replicate.

2. **Estimation** — `.extract_fs_estimates_gen()` computes effect sizes
   *within* identified subgroups.  When `outcome_type`, `effect_measure`,
   and `offset.name` are present in `fs_params`, the estimator switches
   from `coxph()` to `glm()` (Poisson with log-link for IRR).
   Survival-specific estimands (AHR, CDE) are skipped for GLM.

This design isolates the comparison: detection and classification are
*identical* by construction, so the QC focuses on whether IRR estimates
approximate HR estimates within the same subgroups.


# Alternative Hypothesis Simulations {#sec-alt}

## Cox PH Pipeline (Established)

```{r run-alt-cox}
cat("Running", sim_config_alt$n_sims,
    "Cox simulations under H1...\n")

t0 <- proc.time()[3]

results_alt_cox <- foreach(
  sim = seq_len(sim_config_alt$n_sims),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table"),
    seed = TRUE
  )
) %dofuture% {
  run_simulation_analysis(
    sim_id = sim,
    dgm = dgm_calibrated,
    n_sample = sim_config_alt$n_sample,
    analysis_time = sim_config_alt$analysis_time,
    cens_adjust = sim_config_alt$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE,
    run_fs_grf = FALSE,
    run_grf = FALSE,
    fs_params = fs_params_cox,
    verbose = FALSE
  )
}

timings$alt_cox <- proc.time()[3] - t0
cat("Cox H1 completed in", round(timings$alt_cox / 60, 1), "minutes\n")
cat("Results:", nrow(results_alt_cox), "rows\n")
```

## Poisson IRR Pipeline (GLM Extension)

```{r run-alt-irr}
cat("Running", sim_config_alt$n_sims,
    "Poisson IRR simulations under H1...\n")

t0 <- proc.time()[3]

results_alt_irr <- foreach(
  sim = seq_len(sim_config_alt$n_sims),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table"),
    seed = TRUE
  )
) %dofuture% {
  run_simulation_analysis(
    sim_id = sim,
    dgm = dgm_calibrated,
    n_sample = sim_config_alt$n_sample,
    analysis_time = sim_config_alt$analysis_time,
    cens_adjust = sim_config_alt$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE,
    run_fs_grf = FALSE,
    run_grf = FALSE,
    fs_params = fs_params_irr,
    verbose = FALSE
  )
}

timings$alt_irr <- proc.time()[3] - t0
cat("IRR H1 completed in", round(timings$alt_irr / 60, 1), "minutes\n")
cat("Results:", nrow(results_alt_irr), "rows\n")
```


# Null Hypothesis Simulations {#sec-null}

## Cox PH Pipeline

```{r run-null-cox}
cat("Running", sim_config_null$n_sims,
    "Cox simulations under H0...\n")

t0 <- proc.time()[3]

results_null_cox <- foreach(
  sim = seq_len(sim_config_null$n_sims),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table"),
    seed = TRUE
  )
) %dofuture% {
  run_simulation_analysis(
    sim_id = sim,
    dgm = dgm_null,
    n_sample = sim_config_null$n_sample,
    analysis_time = sim_config_null$analysis_time,
    cens_adjust = sim_config_null$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE,
    run_fs_grf = FALSE,
    run_grf = FALSE,
    fs_params = fs_params_cox,
    verbose = FALSE
  )
}

timings$null_cox <- proc.time()[3] - t0
cat("Cox H0 completed in", round(timings$null_cox / 60, 1), "minutes\n")
```

## Poisson IRR Pipeline

```{r run-null-irr}
cat("Running", sim_config_null$n_sims,
    "Poisson IRR simulations under H0...\n")

t0 <- proc.time()[3]

results_null_irr <- foreach(
  sim = seq_len(sim_config_null$n_sims),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table"),
    seed = TRUE
  )
) %dofuture% {
  run_simulation_analysis(
    sim_id = sim,
    dgm = dgm_null,
    n_sample = sim_config_null$n_sample,
    analysis_time = sim_config_null$analysis_time,
    cens_adjust = sim_config_null$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE,
    run_fs_grf = FALSE,
    run_grf = FALSE,
    fs_params = fs_params_irr,
    verbose = FALSE
  )
}

timings$null_irr <- proc.time()[3] - t0
cat("IRR H0 completed in", round(timings$null_irr / 60, 1), "minutes\n")
```


# Results: Alternative Hypothesis {#sec-results-alt}

## Cox PH Operating Characteristics

```{r oc-alt-cox}
format_oc_results(
  results_alt_cox,
  metrics = c("detection", "classification", "hr_estimates"),
  digits = 2, digits_hr = 2,
  title = "Cox PH: Identification Properties (H1)",
  subtitle = sprintf("n = %d, %d simulations, HR(overall) = %.2f",
                      sim_config_alt$n_sample, sim_config_alt$n_sims,
                      dgm_calibrated$hr_causal)
)
```

```{r est-alt-cox}
build_estimation_table(
  results = results_alt_cox,
  dgm = dgm_calibrated,
  analysis_method = "FS",
  subtitle = sprintf("Cox PH: n = %d, %d sims",
                      sim_config_alt$n_sample, sim_config_alt$n_sims),
  font_size = 14
)
```

## Poisson IRR Operating Characteristics

```{r oc-alt-irr}
# Use format_oc_results with the IRR results
# Note: hr.H.hat / hr.Hc.hat columns now contain IRR estimates
# produced by run_simulation_analysis() with GLM dispatch
format_oc_results(
  results_alt_irr,
  metrics = c("detection", "classification", "hr_estimates"),
  digits = 2, digits_hr = 2,
  title = "Poisson IRR: Identification Properties (H1)",
  subtitle = sprintf("n = %d, %d simulations, HR(overall) = %.2f",
                      sim_config_alt$n_sample, sim_config_alt$n_sims,
                      dgm_calibrated$hr_causal)
)
```

## Side-by-Side Comparison (H1)

```{r compare-alt}
# Extract key metrics for comparison
compare_metrics <- function(res_cox, res_irr, label_cox = "Cox PH",
                            label_irr = "Poisson IRR") {
  # Both pipelines label their results as "FS"
  cox <- res_cox[res_cox$analysis == "FS", ]
  irr <- res_irr[res_irr$analysis == "FS", ]

  cox_found <- cox[cox$any.H == 1, ]
  irr_found <- irr[irr$any.H == 1, ]

  data.frame(
    Metric = c("Detection Rate", "Sensitivity", "Specificity", "PPV", "NPV",
               "Mean HR/IRR (H)", "Mean HR/IRR (Hc)", "Mean HR/IRR (ITT)",
               "Mean SG Size (H)"),
    Cox_PH = c(
      round(mean(cox$any.H), 3),
      round(mean(cox$sens, na.rm = TRUE), 3),
      round(mean(cox$spec, na.rm = TRUE), 3),
      round(mean(cox$ppv, na.rm = TRUE), 3),
      round(mean(cox$npv, na.rm = TRUE), 3),
      round(mean(cox_found$hr.H.hat, na.rm = TRUE), 3),
      round(mean(cox_found$hr.Hc.hat, na.rm = TRUE), 3),
      round(mean(cox$hr.itt, na.rm = TRUE), 3),
      round(mean(cox_found$size.H, na.rm = TRUE), 0)
    ),
    Poisson_IRR = c(
      round(mean(irr$any.H), 3),
      round(mean(irr$sens, na.rm = TRUE), 3),
      round(mean(irr$spec, na.rm = TRUE), 3),
      round(mean(irr$ppv, na.rm = TRUE), 3),
      round(mean(irr$npv, na.rm = TRUE), 3),
      round(mean(irr_found$hr.H.hat, na.rm = TRUE), 3),
      round(mean(irr_found$hr.Hc.hat, na.rm = TRUE), 3),
      round(mean(irr$hr.itt, na.rm = TRUE), 3),
      round(mean(irr_found$size.H, na.rm = TRUE), 0)
    ),
    stringsAsFactors = FALSE
  )
}

comp_alt <- compare_metrics(results_alt_cox, results_alt_irr)

gt(comp_alt) |>
  tab_header(
    title = md("**H1 Comparison: Cox PH vs Poisson IRR**"),
    subtitle = sprintf("n = %d, %d simulations, true HR(H) = %.2f",
                        sim_config_alt$n_sample, sim_config_alt$n_sims,
                        dgm_calibrated$hr_H_true)
  ) |>
  cols_label(
    Metric = "Metric",
    Cox_PH = "Cox PH (HR)",
    Poisson_IRR = "Poisson IRR"
  ) |>
  tab_source_note(
    md(paste0("**True values:** HR(H) = ",
              round(dgm_calibrated$hr_H_true, 3),
              ", HR(Hc) = ",
              round(dgm_calibrated$hr_Hc_true, 3)))
  )
```


# Results: Null Hypothesis {#sec-results-null}

## Side-by-Side Comparison (H0)

```{r compare-null}
comp_null <- compare_metrics(results_null_cox, results_null_irr)

gt(comp_null) |>
  tab_header(
    title = md("**H0 Comparison: Cox PH vs Poisson IRR (Type I Error)**"),
    subtitle = sprintf("n = %d, %d simulations, true HR = %.2f (uniform)",
                        sim_config_null$n_sample, sim_config_null$n_sims,
                        dgm_null$hr_causal)
  ) |>
  cols_label(
    Metric = "Metric",
    Cox_PH = "Cox PH (HR)",
    Poisson_IRR = "Poisson IRR"
  )
```


# Per-Simulation Concordance {#sec-concordance}

Since both pipelines use the *same* Cox PH–based `forestsearch()` call
with identical seeds, subgroup identification must agree in every
replicate.  This concordance check serves as a sanity test confirming
the design; the substantive comparison is in the HR vs IRR estimates.

```{r concordance}
# Match simulations by sim_id
cox_detect <- results_alt_cox[results_alt_cox$analysis == "FS",
                              c("sim", "any.H")]
names(cox_detect) <- c("sim", "cox_found")

irr_detect <- results_alt_irr[results_alt_irr$analysis == "FS",
                               c("sim", "any.H")]
names(irr_detect) <- c("sim", "irr_found")

concordance <- merge(cox_detect, irr_detect, by = "sim")

# Agreement matrix
agree_both   <- sum(concordance$cox_found == 1 & concordance$irr_found == 1)
agree_neither <- sum(concordance$cox_found == 0 & concordance$irr_found == 0)
cox_only     <- sum(concordance$cox_found == 1 & concordance$irr_found == 0)
irr_only     <- sum(concordance$cox_found == 0 & concordance$irr_found == 1)
n_matched    <- nrow(concordance)

cat("=== Per-Simulation Concordance ===\n")
cat("Matched simulations:", n_matched, "\n")
cat("Both found SG:      ", agree_both, sprintf("(%.0f%%)\n", 100 * agree_both / n_matched))
cat("Neither found SG:   ", agree_neither, sprintf("(%.0f%%)\n", 100 * agree_neither / n_matched))
cat("Cox only:           ", cox_only, sprintf("(%.0f%%)\n", 100 * cox_only / n_matched))
cat("IRR only:           ", irr_only, sprintf("(%.0f%%)\n", 100 * irr_only / n_matched))
cat("Agreement rate:     ", round(100 * (agree_both + agree_neither) / n_matched, 1), "%\n")
```


# Visualization {#sec-viz}

## Detection Rate Comparison

```{r viz-detection, fig.width=7, fig.height=4}
det_df <- data.frame(
  Pipeline = c("Cox PH", "Poisson IRR"),
  H1_Power = c(
    mean(results_alt_cox[results_alt_cox$analysis == "FS", ]$any.H),
    mean(results_alt_irr[results_alt_irr$analysis == "FS", ]$any.H)
  ),
  H0_TypeI = c(
    mean(results_null_cox[results_null_cox$analysis == "FS", ]$any.H),
    mean(results_null_irr[results_null_irr$analysis == "FS", ]$any.H)
  )
)

det_long <- data.frame(
  Pipeline = rep(det_df$Pipeline, 2),
  Hypothesis = rep(c("H1 (Power)", "H0 (Type I)"), each = 2),
  Rate = c(det_df$H1_Power, det_df$H0_TypeI)
)

ggplot(det_long, aes(x = Pipeline, y = Rate, fill = Pipeline)) +
  geom_col(width = 0.5) +
  facet_wrap(~ Hypothesis, scales = "free_y") +
  geom_text(aes(label = sprintf("%.3f", Rate)), vjust = -0.5, size = 3.5) +
  scale_fill_manual(values = c("Cox PH" = "steelblue", "Poisson IRR" = "coral")) +
  labs(title = "Detection Rates: Cox PH vs Poisson IRR",
       y = "Rate", x = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

## ITT Estimate Comparison

```{r viz-itt, fig.width=7, fig.height=5}
cox_itt <- results_alt_cox[results_alt_cox$analysis == "FS", ]$hr.itt
irr_itt <- results_alt_irr[results_alt_irr$analysis == "FS", ]$hr.itt

itt_df <- data.frame(
  Cox_HR = cox_itt[seq_len(min(length(cox_itt), length(irr_itt)))],
  IRR = irr_itt[seq_len(min(length(cox_itt), length(irr_itt)))]
)

ggplot(itt_df, aes(x = Cox_HR, y = IRR)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Per-Simulation ITT Estimates: Cox HR vs Poisson IRR",
       x = "Cox HR", y = "Poisson IRR") +
  coord_equal() +
  theme_minimal(base_size = 12)
```

## Subgroup Estimate Comparison (Key Validation)

Since both pipelines identify the *same* subgroups, we can directly compare
Cox HR vs Poisson IRR within those subgroups.  This is the core QC target.

```{r viz-subgroup, fig.width=10, fig.height=5}
cox_sg <- results_alt_cox[results_alt_cox$analysis == "FS" &
                           results_alt_cox$any.H == 1, ]
irr_sg <- results_alt_irr[results_alt_irr$analysis == "FS" &
                           results_alt_irr$any.H == 1, ]

sg_df <- merge(
  cox_sg[, c("sim", "hr.H.hat", "hr.Hc.hat")],
  irr_sg[, c("sim", "hr.H.hat", "hr.Hc.hat")],
  by = "sim", suffixes = c("_cox", "_irr")
)

p_H <- ggplot(sg_df, aes(x = hr.H.hat_cox, y = hr.H.hat_irr)) +
  geom_point(alpha = 0.6, size = 2, color = "firebrick") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40") +
  labs(title = expression(hat(H) ~ ": Harm Subgroup"),
       x = "Cox HR", y = "Poisson IRR") +
  theme_minimal(base_size = 12)

p_Hc <- ggplot(sg_df, aes(x = hr.Hc.hat_cox, y = hr.Hc.hat_irr)) +
  geom_point(alpha = 0.6, size = 2, color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40") +
  labs(title = expression(hat(H)^c ~ ": Complement"),
       x = "Cox HR", y = "Poisson IRR") +
  theme_minimal(base_size = 12)

p_H + p_Hc +
  patchwork::plot_annotation(
    title = "Per-Simulation Subgroup Estimates: Cox HR vs Poisson IRR",
    subtitle = sprintf("%d simulations where subgroup was identified", nrow(sg_df))
  )
```


# Computational Timing Summary {#sec-timing}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Stage = c("DGM creation (H1)", "Calibration", "DGM creation (H0)",
            "H1 sims (Cox)", "H1 sims (IRR)",
            "H0 sims (Cox)", "H0 sims (IRR)",
            "Total"),
  Seconds = c(
    timings$dgm_creation, timings$calibration, timings$dgm_null,
    timings$alt_cox, timings$alt_irr,
    timings$null_cox, timings$null_irr,
    timings$total
  ),
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(
    title = "Computational Timing Summary",
    subtitle = sprintf("%d H1 + %d H0 sims per pipeline, %d workers",
                        sim_config_alt$n_sims, sim_config_null$n_sims,
                        n_workers)
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)", Minutes = "Time (min)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total")
  )
```


# Summary {#sec-summary}

This QC simulation validates the Poisson IRR approximation to the Cox
PH model within the ForestSearch framework.

**Key validation points:**

| Aspect | Expected | Observed |
|---|---|---|
| ITT estimates | IRR approximates HR | Check scatter plot above |
| Detection rate (H1) | Identical (same identification) | Check comparison table |
| Classification metrics | Identical sens/spec/PPV/NPV | Check comparison table |
| Type I error (H0) | Identical | Check null comparison |
| Per-simulation concordance | 100% agreement | Check concordance matrix |

**Interpretation:**

Both pipelines use the same Cox PH–based `forestsearch()` for subgroup
*identification*, so detection, classification, and concordance must be
identical by construction.  The comparison that matters is the *estimation*
step: the Poisson IRR (with log-time offset) should approximate the Cox HR
within the identified subgroups.  In the GBSG setting with ~50% event rate,
some divergence is expected due to the rare-events approximation.  The
comparison tables and scatter plots above quantify this approximation error.


# Session Information {#sec-session}

```{r session-info}
plan(sequential)
sessionInfo()
```
