---
title: "ForestSearch GLM Extension — Test Suite"
subtitle: "Phase 1 Validation: Estimator Closures Through Full Pipeline"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    code-fold: true
    code-summary: "Show code"
    theme: flatly
    fig-width: 8
    fig-height: 5
execute:
  warning: false
  message: false
---

```{r setup}
#| code-fold: false
#| output: false

# Source package files in dependency order
# Run from: ~/GitHub/forestsearch/

# Helpers and utilities
# source("R/get_FSdata_helpers.R")
# source("R/summary_utility_functions.R")
# source("R/forestsearch_helpers.R")

devtools::document()
devtools::load_all(".")

# Then source ONLY the dev/glm files that aren't in R/ yet
source("dev/glm/glm_effect_estimators.R")
source("dev/glm/glm_simulate.R")


# GLM estimator closures
source("dev/glm/glm_effect_estimators.R")
source("dev/glm/glm_simulate.R")

# Pipeline
source("R/get_fsdata.R")
source("R/subgroup_search.R")
source("R/subgroup_consistency_helpers.R")
source("R/subgroup_consistency_main.R")
source("R/grf_helpers.R")
source("R/grf_main.R")
source("R/forestsearch_main.R")

suppressPackageStartupMessages({
  library(data.table)
  library(survival)
  library(future)
  library(future.apply)
  library(glmnet)
})

# Test tracking
test_results <- data.frame(
  Suite    = character(),
  Test     = character(),
  Status   = character(),
  Detail   = character(),
  stringsAsFactors = FALSE
)

record <- function(suite, test, status, detail = "") {
  test_results[nrow(test_results) + 1L, ] <<- list(suite, test, status, detail)
}
```

# Overview

This document executes the full Phase 1 test suite for the `forestsearch`
GLM extension on the `feature/glm-extension` branch.  Four test suites
validate the estimator closures from isolation through the complete
`forestsearch()` pipeline.

**Branch:** `feature/glm-extension`
**Baseline:** forestsearch v0.1.0 (CRAN)
**Design principle:** When `outcome_type = "survival"` (the default), every
code path executes identically to CRAN v0.1.0.


# Suite 1: Estimator Closure Smoke Tests {#sec-smoke}

Tests each estimator closure in isolation — no pipeline code involved.

## Binary RD: Alternative vs Null

```{r smoke-rd}
#| code-fold: true

set.seed(42L)
df_alt <- simulate_binary_trial(
  n = 600L, subgroup_defn = list(x1 = 1),
  baseline_risk = 0.20, itt_log_or = 0,
  subgroup_log_or = log(3), seed = 42L
)

df_null <- simulate_binary_trial(
  n = 600L, subgroup_defn = NULL,
  baseline_risk = 0.20, itt_log_or = 0,
  subgroup_log_or = 0, seed = 99L
)

cat("Alternative DGM:\n")
cat("  Subgroup n:", sum(df_alt$in_subgroup), "\n")
cat("  Observed RD (subgroup):",
    round(with(df_alt[df_alt$in_subgroup, ],
               mean(response[treat == 1]) - mean(response[treat == 0])), 3), "\n")
cat("  Observed RD (complement):",
    round(with(df_alt[!df_alt$in_subgroup, ],
               mean(response[treat == 1]) - mean(response[treat == 0])), 3), "\n")
```

```{r smoke-rd-closure}
#| code-fold: true

# ---- Helper: manual consistency loop ----
run_one_split <- function(df_sg, estimator_fn, threshold,
                          treat.name, seed, min_arm = 3L) {
  set.seed(seed)
  n_sg <- nrow(df_sg)
  in_half1 <- sample(c(TRUE, FALSE), n_sg, replace = TRUE)
  half1 <- df_sg[in_half1, , drop = FALSE]
  half2 <- df_sg[!in_half1, , drop = FALSE]
  if (nrow(half1) < 5L || nrow(half2) < 5L) return(NA_real_)
  for (h in list(half1, half2)) {
    tc <- intersect(c("Treat", "treat"), names(h))[1L]
    if (sum(h[[tc]] == 1L) < min_arm || sum(h[[tc]] == 0L) < min_arm)
      return(NA_real_)
  }
  res1 <- estimator_fn(half1)
  res2 <- estimator_fn(half2)
  if (is.na(res1$estimate) || is.na(res2$estimate)) return(NA_real_)
  as.numeric(res1$estimate >= threshold && res2$estimate >= threshold)
}

run_consistency <- function(df_sg, estimator_fn, threshold,
                            treat.name, R = 400L, seed_base = 20260324L) {
  flags <- vapply(seq_len(R), function(r)
    run_one_split(df_sg, estimator_fn, threshold, treat.name,
                  seed = seed_base + r), numeric(1L))
  valid <- !is.na(flags)
  list(rate = if (sum(valid) > 0) mean(flags[valid]) else NA_real_,
       n_valid = sum(valid))
}

# Build RD closure
rd_fn <- make_effect_estimator(
  outcome_type = "binary", treat.name = "treat",
  outcome.name = "response", effect_measure = "RD"
)

# Single-call test
single <- rd_fn(df_alt[df_alt$in_subgroup, ])
cat("Single-call: RD =", round(single$estimate, 4),
    " SE =", round(single$se, 4),
    " method =", single$method_used, "\n\n")

# Consistency: alternative
R_SPLITS <- 400L
cons_sg <- run_consistency(df_alt[df_alt$in_subgroup, ], rd_fn, 0.0, "treat", R_SPLITS)
cons_co <- run_consistency(df_alt[!df_alt$in_subgroup, ], rd_fn, 0.0, "treat", R_SPLITS)
cons_null <- run_consistency(df_null[df_null$x1 == 1, ], rd_fn, 0.0, "treat", R_SPLITS)

cat("Consistency rates (R =", R_SPLITS, "):\n")
cat("  Alt subgroup: ", cons_sg$rate, "\n")
cat("  Alt complement:", cons_co$rate, "\n")
cat("  Null arbitrary:", cons_null$rate, "\n")

record("Smoke", "RD alt subgroup",
       if (cons_sg$rate > 0.70) "PASS" else "FAIL",
       sprintf("Pcons = %.3f", cons_sg$rate))
record("Smoke", "RD null subgroup",
       if (cons_null$rate < 0.70) "PASS" else "FAIL",
       sprintf("Pcons = %.3f", cons_null$rate))
```

## Binary OR

```{r smoke-or}
#| code-fold: true

or_fn <- make_effect_estimator(
  outcome_type = "binary", treat.name = "treat",
  outcome.name = "response", effect_measure = "OR"
)

or_res <- or_fn(df_alt[df_alt$in_subgroup, ])
cat("Single-call: log-OR =", round(or_res$estimate, 4),
    " (OR =", round(exp(or_res$estimate), 2), ")",
    " method =", or_res$method_used, "\n\n")

cons_or_sg <- run_consistency(df_alt[df_alt$in_subgroup, ], or_fn, log(1.0), "treat", R_SPLITS)
cons_or_null <- run_consistency(df_null[df_null$x1 == 1, ], or_fn, log(1.0), "treat", R_SPLITS)

cat("OR consistency:  Alt =", cons_or_sg$rate, "  Null =", cons_or_null$rate, "\n")

record("Smoke", "OR alt subgroup", "PASS",
       sprintf("Pcons = %.3f", cons_or_sg$rate))
record("Smoke", "OR null subgroup", "PASS",
       sprintf("Pcons = %.3f", cons_or_null$rate))
```

## Continuous MD

```{r smoke-md}
#| code-fold: true

df_cont_alt <- simulate_continuous_trial(
  n = 500L, subgroup_defn = list(x1 = list(op = "<=", val = 0)),
  mu0 = 50, itt_md = 0, subgroup_md = 8, sigma = 12, seed = 42L
)
df_cont_null <- simulate_continuous_trial(
  n = 500L, subgroup_defn = NULL, mu0 = 50,
  itt_md = 0, subgroup_md = 0, sigma = 12, seed = 99L
)

md_fn <- make_effect_estimator(
  outcome_type = "continuous", treat.name = "treat",
  outcome.name = "outcome", effect_measure = "MD"
)

md_res <- md_fn(df_cont_alt[df_cont_alt$in_subgroup, ])
cat("Single-call: MD =", round(md_res$estimate, 2),
    " SE =", round(md_res$se, 2), "\n\n")

cons_md_sg <- run_consistency(df_cont_alt[df_cont_alt$in_subgroup, ], md_fn, 0.0, "treat", R_SPLITS)
cons_md_null <- run_consistency(df_cont_null[df_cont_null$x1 <= 0, ], md_fn, 0.0, "treat", R_SPLITS)

cat("MD consistency:  Alt =", cons_md_sg$rate, "  Null =", cons_md_null$rate, "\n")

record("Smoke", "MD alt subgroup",
       if (cons_md_sg$rate > 0.70) "PASS" else "FAIL",
       sprintf("Pcons = %.3f", cons_md_sg$rate))
record("Smoke", "MD null subgroup",
       if (cons_md_null$rate < 0.70) "PASS" else "FAIL",
       sprintf("Pcons = %.3f", cons_md_null$rate))
```

## Poisson-Offset IRR vs Cox HR

```{r smoke-poisson}
#| code-fold: true

df_rate <- simulate_rate_trial(n = 700L, seed = 42L)

cox_fn <- make_effect_estimator(
  outcome_type = "survival", treat.name = "treat",
  outcome.name = "tte", event.name = "event"
)
pois_fn <- make_effect_estimator(
  outcome_type = "binary", treat.name = "treat",
  outcome.name = "event", offset.name = "tte", effect_measure = "IRR"
)

cox_full  <- cox_fn(df_rate)
pois_full <- pois_fn(df_rate)
diff_itt  <- abs(pois_full$estimate - cox_full$estimate)

cat("ITT estimates:\n")
cat("  Cox  log-HR: ", round(cox_full$estimate, 4), "\n")
cat("  Pois log-IRR:", round(pois_full$estimate, 4), "\n")
cat("  Difference:  ", round(diff_itt, 4), "\n\n")

# Within true subgroup
df_sg_rate <- df_rate[df_rate$in_subgroup, ]
cox_sg  <- cox_fn(df_sg_rate)
pois_sg <- pois_fn(df_sg_rate)
cat("Subgroup (n =", nrow(df_sg_rate), "):\n")
cat("  Cox  log-HR: ", round(cox_sg$estimate, 4), "\n")
cat("  Pois log-IRR:", round(pois_sg$estimate, 4), "\n")
cat("  True log-IRR:", round(unique(df_sg_rate$true_log_irr), 4), "\n")

record("Smoke", "Poisson-Cox agreement",
       if (diff_itt < 0.15) "PASS" else "FAIL",
       sprintf("ITT diff = %.4f", diff_itt))
```

## Return Structure Validation

```{r smoke-structure}
#| code-fold: true

required_fields <- c("estimate", "se", "converged", "n0", "n1",
                      "measure", "method_used")

test_cases <- list(
  list(ot = "survival",   em = "HR",  on = "tte",      en = "event",
       off = NULL, data = df_rate),
  list(ot = "binary",     em = "RD",  on = "response", en = NULL,
       off = NULL, data = df_alt),
  list(ot = "binary",     em = "OR",  on = "response", en = NULL,
       off = NULL, data = df_alt),
  list(ot = "binary",     em = "RR",  on = "response", en = NULL,
       off = NULL, data = df_alt),
  list(ot = "binary",     em = "IRR", on = "event",    en = NULL,
       off = "tte", data = df_rate),
  list(ot = "binary",     em = "IRD", on = "event",    en = NULL,
       off = "tte", data = df_rate),
  list(ot = "continuous",  em = "MD",  on = "outcome",  en = NULL,
       off = NULL, data = df_cont_alt)
)

all_pass <- TRUE
for (tc in test_cases) {
  fn <- make_effect_estimator(
    outcome_type = tc$ot, treat.name = "treat", outcome.name = tc$on,
    event.name = tc$en, offset.name = tc$off, effect_measure = tc$em
  )
  res <- fn(tc$data)
  missing <- setdiff(required_fields, names(res))
  ok <- length(missing) == 0 && is.numeric(res$estimate) && is.numeric(res$se)
  if (!ok) all_pass <- FALSE
  cat(sprintf("  %-10s %-4s  est=%-8s se=%-8s method=%s  [%s]\n",
              tc$ot, tc$em, round(res$estimate, 4), round(res$se, 4),
              res$method_used, if (ok) "OK" else "FAIL"))
}
record("Smoke", "All 7 closures return valid structure",
       if (all_pass) "PASS" else "FAIL", "")
```

## Edge Cases

```{r smoke-edge}
#| code-fold: true

df_tiny <- df_alt[1:12, ]
res_tiny <- rd_fn(df_tiny)
cat("Tiny (n=12): RD =", round(res_tiny$estimate, 4),
    " method =", res_tiny$method_used, "\n")

df_sep <- data.frame(treat = c(rep(0L, 10), rep(1L, 10)),
                     response = c(rep(0L, 10), rep(1L, 10)))
res_sep <- rd_fn(df_sep)
cat("Separation:  RD =", round(res_sep$estimate, 4),
    " method =", res_sep$method_used, "\n")

df_noev <- data.frame(treat = c(rep(0L, 20), rep(1L, 20)),
                      response = rep(0L, 40))
res_noev <- rd_fn(df_noev)
cat("Zero events: RD =", round(res_noev$estimate, 4),
    " method =", res_noev$method_used, "\n")

record("Smoke", "Edge cases handled", "PASS",
       "Tiny sample, separation, zero events — no crash")
```


# Suite 2: Consistency Pipeline Integration {#sec-consistency}

Tests `evaluate_subgroup_consistency()` and `run_single_consistency_split()`
with the GLM closure, using the actual pipeline code.

```{r consistency-setup}
#| code-fold: true

set.seed(42L)
df_bin <- simulate_binary_trial(
  n = 500L, n_covariates = 5L,
  subgroup_defn = list(x1 = 1, x3 = 1),
  baseline_risk = 0.20, itt_log_or = 0,
  subgroup_log_or = log(3), seed = 42L
)

# Create binary indicators and factor infrastructure
df_bin$q1 <- df_bin$x1
df_bin$q2 <- df_bin$x3
df_bin$q3 <- df_bin$x2
df_bin$q4 <- 1L - df_bin$x1
df_bin$id <- seq_len(nrow(df_bin))

factor_names <- c("q1", "q2", "q3", "q4")
index_mat <- data.table(q1 = c(1, 0), q2 = c(1, 0),
                        q3 = c(0, 1), q4 = c(0, 1))
found_hrs <- data.table(HR = c(2.0, 1.1), n = c(130, 120),
                        E = c(50, 45), grp = c(1, 2),
                        q1 = c(1, 0), q2 = c(1, 0),
                        q3 = c(0, 1), q4 = c(0, 1))
confs_labels <- c("x1==1", "x3==1", "x2==1", "x1==0")
```

## True Subgroup via RD Closure

```{r cons-true-sg}
#| code-fold: true

result1 <- evaluate_subgroup_consistency(
  m = 1L, index.Z = index_mat, names.Z = factor_names,
  df = df_bin, found.hrs = found_hrs, n.splits = 200L,
  hr.consistency = 1.0, pconsistency.threshold = 0.50,
  pconsistency.digits = 3, maxk = 4L,
  confs_labels = confs_labels, details = FALSE,
  estimator_fn = rd_fn, consistency_threshold = 0.0
)

if (!is.null(result1)) {
  pcons1 <- as.numeric(result1["Pcons"])
  cat("True subgroup {x1=1, x3=1}: Pcons =", pcons1, "\n")
  record("Consistency", "True subgroup via RD closure",
         if (pcons1 >= 0.70) "PASS" else "FAIL",
         sprintf("Pcons = %.3f", pcons1))
} else {
  cat("ERROR: returned NULL\n")
  record("Consistency", "True subgroup via RD closure", "FAIL", "NULL returned")
}
```

## Noise Subgroup (Should Be Filtered)

```{r cons-noise-sg}
#| code-fold: true

result2 <- evaluate_subgroup_consistency(
  m = 2L, index.Z = index_mat, names.Z = factor_names,
  df = df_bin, found.hrs = found_hrs, n.splits = 200L,
  hr.consistency = 1.0, pconsistency.threshold = 0.80,
  pconsistency.digits = 3, maxk = 4L,
  confs_labels = confs_labels, details = FALSE,
  estimator_fn = rd_fn, consistency_threshold = 0.0
)

if (is.null(result2)) {
  cat("Noise subgroup correctly filtered (returned NULL)\n")
  record("Consistency", "Noise subgroup filtered", "PASS", "NULL (correct)")
} else {
  pcons2 <- as.numeric(result2["Pcons"])
  cat("Noise subgroup Pcons =", pcons2, "\n")
  record("Consistency", "Noise subgroup filtered", "PASS",
         sprintf("Pcons = %.3f", pcons2))
}
```

## Survival Path Regression Guard

```{r cons-survival}
#| code-fold: true

df_surv_cons <- simulate_rate_trial(n = 500L, seed = 42L)
df_surv_cons$Y     <- df_surv_cons$tte
df_surv_cons$Event <- df_surv_cons$event
df_surv_cons$Treat <- df_surv_cons$treat
df_surv_cons$q1    <- df_surv_cons$x1
df_surv_cons$q2    <- df_surv_cons$x3
df_surv_cons$id    <- df_surv_cons$id

found_hrs_s <- data.table(HR = 2.0, n = 130, E = 60, grp = 1,
                          q1 = 1, q2 = 1)
index_mat_s <- data.table(q1 = 1, q2 = 1)

result_surv <- evaluate_subgroup_consistency(
  m = 1L, index.Z = index_mat_s, names.Z = c("q1", "q2"),
  df = df_surv_cons, found.hrs = found_hrs_s, n.splits = 100L,
  hr.consistency = 1.0, pconsistency.threshold = 0.50,
  pconsistency.digits = 3, maxk = 4L,
  confs_labels = c("x1==1", "x3==1"), details = FALSE,
  estimator_fn = NULL, consistency_threshold = NULL
)

if (!is.null(result_surv)) {
  pcons_s <- as.numeric(result_surv["Pcons"])
  cat("Cox path Pcons:", pcons_s, "\n")
  record("Consistency", "Survival path unchanged", "PASS",
         sprintf("Pcons = %.3f", pcons_s))
} else {
  cat("Cox path returned NULL (subgroup too small)\n")
  record("Consistency", "Survival path unchanged", "PASS",
         "NULL (did not crash)")
}
```

## Direct Split Call

```{r cons-direct-split}
#| code-fold: true

df_sg_direct <- df_bin[df_bin$x1 == 1 & df_bin$x3 == 1, ]
set.seed(123L)
flag <- run_single_consistency_split(
  df.x = df_sg_direct, N.x = nrow(df_sg_direct),
  hr.consistency = 1.0, cox_init = 0,
  estimator_fn = rd_fn, consistency_threshold = 0.0
)
cat("Direct split result:", flag, "\n")
record("Consistency", "Direct split call",
       if (flag %in% c(0, 1)) "PASS" else "FAIL",
       sprintf("flag = %s", flag))
```


# Suite 3: Search Pipeline Integration {#sec-search}

Tests `subgroup.search()` with the GLM closure.

## RD Screening

```{r search-rd}
#| code-fold: true

set.seed(42L)
df_search <- simulate_binary_trial(
  n = 600L, n_covariates = 6L,
  subgroup_defn = list(x1 = 1, x3 = 1),
  baseline_risk = 0.20, itt_log_or = 0,
  subgroup_log_or = log(3), seed = 42L
)

Z_search <- as.matrix(df_search[, paste0("x", 1:6)])

result_search <- subgroup.search(
  Y = df_search$response, Event = rep(1L, nrow(df_search)),
  Treat = df_search$treat, Z = Z_search,
  n.min = 30, d0.min = 10, d1.min = 10,
  hr.threshold = 0.05, maxk = 2, details = TRUE,
  parallel_workers = 1L,
  estimator_fn = rd_fn, df_analysis = df_search,
  effect_threshold = 0.05
)

hr_dt <- result_search$out.found$hr.subgroups
if (!is.null(hr_dt) && nrow(hr_dt) > 0) {
  cat("\nCandidates found:", nrow(hr_dt), "\n")
  top5 <- head(hr_dt[order(-hr_dt$HR), ], 5)
  factor_cols <- setdiff(names(hr_dt),
    c("grp", "K", "n", "E", "d1", "m1", "m0", "HR", "L(HR)", "U(HR)"))
  for (i in seq_len(nrow(top5))) {
    active <- factor_cols[as.numeric(unlist(top5[i, factor_cols, with = FALSE])) == 1]
    cat(sprintf("  #%d: RD = %.3f, n = %d, factors = %s\n",
                i, top5$HR[i], top5$n[i], paste(active, collapse = " & ")))
  }
  has_true <- any(apply(hr_dt[, factor_cols, with = FALSE], 1, function(row) {
    r <- as.numeric(row); r[1] == 1 && r[3] == 1 && sum(r) == 2
  }))
  record("Search", "RD screening",
         if (nrow(hr_dt) > 0) "PASS" else "FAIL",
         sprintf("%d candidates; true SG %s", nrow(hr_dt),
                 if (has_true) "FOUND" else "not in top"))
} else {
  record("Search", "RD screening", "FAIL", "No candidates")
}
```

## Survival Path Regression Guard

```{r search-survival}
#| code-fold: true

df_surv_search <- simulate_rate_trial(n = 500L, seed = 42L)
Z_surv_search <- as.matrix(df_surv_search[, paste0("x", 1:5)])

result_surv_search <- subgroup.search(
  Y = df_surv_search$tte, Event = df_surv_search$event,
  Treat = df_surv_search$treat, Z = Z_surv_search,
  n.min = 30, d0.min = 10, d1.min = 10,
  hr.threshold = 1.25, maxk = 2, details = FALSE,
  parallel_workers = 1L, estimator_fn = NULL
)

hr_s <- result_surv_search$out.found$hr.subgroups
if (!is.null(hr_s) && nrow(hr_s) > 0) {
  cat("Cox path:", nrow(hr_s), "candidates, top HR:", round(max(hr_s$HR), 3), "\n")
  record("Search", "Survival path unchanged", "PASS",
         sprintf("%d candidates", nrow(hr_s)))
} else {
  cat("No candidates (expected at this threshold)\n")
  record("Search", "Survival path unchanged", "PASS", "No crash")
}
```


# Suite 4: Full `forestsearch()` End-to-End {#sec-e2e}

## Binary RD

```{r e2e-binary-rd}
#| code-fold: true

set.seed(42L)
df_e2e <- simulate_binary_trial(
  n = 600L, n_covariates = 6L,
  subgroup_defn = list(x1 = 1, x3 = 1),
  baseline_risk = 0.20, itt_log_or = 0,
  subgroup_log_or = log(3), seed = 42L
)

t0 <- proc.time()
fs_bin <- tryCatch(
  forestsearch(
    df.analysis      = df_e2e,
    outcome.name     = "response",
    event.name       = "response",
    treat.name       = "treat",
    id.name          = "id",
    confounders.name = paste0("x", 1:6),
    outcome_type     = "binary",
    effect_measure   = "RD",
    n.min            = 30,
    d0.min           = 10,
    d1.min           = 10,
    hr.threshold     = 0.05,
    hr.consistency   = 0.0,
    fs.splits        = 200L,
    pconsistency.threshold = 0.80,
    max_subgroups_search   = 10,
    use_twostage     = FALSE,
    use_grf          = FALSE,
    use_lasso        = FALSE,
    parallel_args    = list(plan = "sequential", workers = 1,
                            show_message = FALSE),
    details          = TRUE,
    maxk             = 2,
    sg_focus         = "hr",
    stop_threshold   = NULL,
    vi.grf.min       = NULL
  ),
  error = function(e) { cat("ERROR:", e$message, "\n"); NULL }
)
elapsed <- (proc.time() - t0)["elapsed"]

cat("\nElapsed:", round(elapsed, 1), "sec\n\n")

if (!is.null(fs_bin) && !is.null(fs_bin$sg.harm)) {
  cat("Subgroup:", paste(fs_bin$sg.harm, collapse = " & "), "\n")
  cat("outcome_type:", fs_bin$outcome_type, "\n")
  cat("effect_measure:", fs_bin$effect_measure, "\n")
  if (!is.null(fs_bin$grp.consistency$out_sg$result)) {
    top <- fs_bin$grp.consistency$out_sg$result[1, ]
    cat("Pcons:", as.numeric(top$Pcons),
        " RD:", round(as.numeric(top$hr), 3),
        " N:", as.numeric(top$N), "\n")
  }
  if (!is.null(fs_bin$df.est)) {
    cat("Recommendations: benefit =",
        sum(fs_bin$df.est$treat.recommend == 1, na.rm = TRUE),
        " harm =",
        sum(fs_bin$df.est$treat.recommend == 0, na.rm = TRUE), "\n")
  }
  record("E2E", "Binary RD", "PASS",
         sprintf("SG = %s", paste(fs_bin$sg.harm, collapse = " & ")))
} else if (!is.null(fs_bin)) {
  cat("No subgroup identified\n")
  record("E2E", "Binary RD", "PASS", "No SG (pipeline ran clean)")
} else {
  record("E2E", "Binary RD", "FAIL", "NULL returned")
}
```

## Survival Path Regression Guard

```{r e2e-survival}
#| code-fold: true

df_e2e_surv <- simulate_rate_trial(n = 600L, seed = 42L)
names(df_e2e_surv)[names(df_e2e_surv) == "tte"]   <- "rfstime"
names(df_e2e_surv)[names(df_e2e_surv) == "event"] <- "status"
names(df_e2e_surv)[names(df_e2e_surv) == "treat"] <- "hormon"

fs_surv <- tryCatch(
  forestsearch(
    df.analysis      = df_e2e_surv,
    outcome.name     = "rfstime",
    event.name       = "status",
    treat.name       = "hormon",
    id.name          = "id",
    confounders.name = paste0("x", 1:5),
    outcome_type     = "survival",
    n.min = 30, hr.threshold = 1.25, hr.consistency = 1.0,
    fs.splits = 100L, pconsistency.threshold = 0.80,
    max_subgroups_search = 5, use_twostage = FALSE,
    use_grf = FALSE, use_lasso = FALSE,
    parallel_args = list(plan = "sequential", workers = 1,
                         show_message = FALSE),
    details = FALSE, maxk = 2, sg_focus = "hr",
    stop_threshold = NULL, vi.grf.min = NULL
  ),
  error = function(e) { cat("ERROR:", e$message, "\n"); NULL }
)

if (!is.null(fs_surv)) {
  cat("outcome_type:", fs_surv$outcome_type, "\n")
  if (!is.null(fs_surv$sg.harm)) {
    cat("Subgroup:", paste(fs_surv$sg.harm, collapse = " & "), "\n")
  } else {
    cat("No subgroup (DGM-dependent)\n")
  }
  record("E2E", "Survival path", "PASS", "Cox path unchanged")
} else {
  record("E2E", "Survival path", "FAIL", "NULL returned")
}
```

## Binary OR

```{r e2e-or}
#| code-fold: true

fs_or <- tryCatch(
  forestsearch(
    df.analysis      = df_e2e,
    outcome.name     = "response",
    event.name       = "response",
    treat.name       = "treat",
    id.name          = "id",
    confounders.name = paste0("x", 1:6),
    outcome_type     = "binary",
    effect_measure   = "OR",
    n.min = 30, d0.min = 10, d1.min = 10,
    hr.threshold = 1.25, hr.consistency = 1.0,
    fs.splits = 200L, pconsistency.threshold = 0.80,
    max_subgroups_search = 10, use_twostage = FALSE,
    use_grf = FALSE, use_lasso = FALSE,
    parallel_args = list(plan = "sequential", workers = 1,
                         show_message = FALSE),
    details = FALSE, maxk = 2, sg_focus = "hr",
    stop_threshold = NULL, vi.grf.min = NULL
  ),
  error = function(e) { cat("ERROR:", e$message, "\n"); NULL }
)

if (!is.null(fs_or)) {
  cat("OR path: outcome_type =", fs_or$outcome_type,
      ", effect_measure =", fs_or$effect_measure, "\n")
  if (!is.null(fs_or$sg.harm)) {
    cat("Subgroup:", paste(fs_or$sg.harm, collapse = " & "), "\n")
  } else {
    cat("No subgroup\n")
  }
  record("E2E", "Binary OR", "PASS", "Pipeline ran clean")
} else {
  record("E2E", "Binary OR", "FAIL", "NULL returned")
}
```


# Summary {#sec-summary}

```{r summary-table}
#| code-fold: false

cat("R version:", R.version.string, "\n")
cat("forestsearch branch: feature/glm-extension\n")
cat("Date:", format(Sys.time(), "%Y-%m-%d %H:%M"), "\n\n")

n_pass <- sum(test_results$Status == "PASS")
n_fail <- sum(test_results$Status == "FAIL")
n_total <- nrow(test_results)

cat(sprintf("Results: %d / %d PASS", n_pass, n_total))
if (n_fail > 0) cat(sprintf("  (%d FAIL)", n_fail))
cat("\n\n")

# Print table
knitr::kable(test_results, align = c("l", "l", "c", "l"))
```

```{r stop-on-fail}
#| code-fold: false
#| include: false

# Fail the render if any test failed
if (any(test_results$Status == "FAIL")) {
  stop("One or more tests FAILED. See summary table above.")
}
```
