---
title: "Subgroup Identification Analysis"
subtitle: "Applications to consistency analyses in MRCTs"
author: "Larry Leon"
toc: true
number-sections: true
toc-depth: 3
format: 
  html:
    self-contained: true
    code-fold: true
    bibliography: subgroups_ref3.bib
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    citation_package: natbib
    fig_width: 10
    fig_height: 8
    fontsize: 10pt
    extra_dependencies: "subfig"
---


```{r Setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

options(warn = -1)

rm(list=ls())

library(tinytex)
library(ggplot2)
library(table1)

library(gt)

# test these packages
# just for testing any conflicts
# library(tidyverse)
# library(plyr)
# library(dplyr)
# library(glmnet)

library(survival)
library(data.table)
library(randomForest)
library(grf)
library(policytree)
library(DiagrammeR)

library(grid)
library(forestploter)
library(randomizr)

codepath <- c("Rcode/forestsearch/R/")
source(paste0(codepath,"source_forestsearch_v0.R"))
source_fs_functions(file_loc=codepath)

source("Rcode/kmplotting/R/KMplotting_functions_v0.R")

# Save output flag
output <- FALSE
# Loading results
loadit <- TRUE

# Record time of all analyses
tALL.start<-proc.time()[3]
 
#library(doParallel)
#cat("Number of cores=",c(detectCores()),"\n")

# Note: leave workers unspecified
# on my linux machine workers(=100) needs to be set
plan(multisession)
#plan(multisession, workers=120)

nsims <- 1000
nboots <- 1000


```


# Summary

We evaluate the potential for utilizing subgroup analyses for enhancing regional consistency analyses in oncology multi-regional clinical trials (MRCTs).   We consider a setting 
where asia pacific (AP) countries consist of approximately $15-20\%$ of a global trial and the criteria for consistency is a positive point estimate (Cox model hazard-ratio $< 1.0$) while 
the global trial meets statistical significance.   

Suppose the following -- 

**ITT meets statistical significance but the AP regional analysis appears problematic (hazard-ratio estimate $\approx 1.0$)**

Can a subgroup with enhanced benefit be identified in the non-AP population which then translates to the AP population?   

That is, suppose there exists a subgroup ${\cal G}$ such that the treatment affect 
is more pronounced and (standard) Cox model analyses based on the ${\cal G}$ subgroup of AP (AP$({\cal G})$, say) enables meeting consistency.

We assume Cox model analyses are the primary analysis where only the treatment arm is included as a covariate (with or without stratification by randomization factors). 

While there are a plethora of subgroup identification approaches available we consider the *forestSearch* approach, @Leon_2024, which was developed for identifying strong subgroup treatment effects (i.e., strong HTEs).  

We consider three objectives:  (A) Identify a subgroup with the largest detrimental effect (*harm*, say) defined in terms of the largest hazard-ratio estimate exceeding a clinical threshold in favor of control (e.g., $\hat\beta \geq \log(0.90)$);  (B) Identify the *smallest* subgroup with a hazard-ratio estimate indicative of *harm* (e.g., $\hat\beta \geq \log(0.90)$); and (C) Identify the *largest* subgroup with strong benefit (e.g., $\hat\beta \leq \log(0.60)$).
These objectives are related with the overarching goal of identifying a subgroup with enhanced benefit;   In (A) and (B) we identify a subgroup indicative of *harm*, $\hat{\cal H}$ say, in which case the complementary subgroup (${\hat{\cal H}}^{c}:= {\hat{\cal G}}$) may be generally beneficial.  Whereas in (C) we are more directly *targeting* a strong (clinically compelling) beneficial subgroup effect.  

In this note we provide a brief review of the Weibull model in Section 2 where we describe a (2-phase) spline model that we utilize for inducing differential treatment effects by way of a ''biomarker''.  Section 3 provides examples of how to generate ''biomarker profiles'' of interest where one can define subgroups based on biomarker cutpoints which correspond to underlying causal hazard ratio effects (e.g., causal hazard-ratio effect for biomarker $\geq 2$ is approximately $0.5$, say).  Section 4 describes the simulation setup (data generating model/process) and generates an example dataset used for analysis illustration where in Section 5 we discuss details of implementing (A)-(C) above.  In Section 6 we evaluate the operating characteristics for objective (B) under conditions where there exists a strong biomarker effect, and under the null of a uniform treatment benefit (i.e., ${\cal G} = \emptyset$).   Lastly, Section 7 briefly discusses estimation (de-biased point and CI estimates) -- taking into account the identification algorithm -- where there is interest in inference regarding the population from which the subgroup was identified (e.g., non-AP).  

Throughout this note R code is provided and illustrated based on real trial data (subject-level baseline characteristics are fixed at their observed quantities).  Our hope is that these notes can serve as a framework for future trial planning as well as for simulating from study data to understand the operating characteristics for teams and external discussions.

A diagram illustrating the general flow follows.


```{r}
#| fig-cap: "Subgroup identification analysis based on non-AP data to identify subgroups of marked benefit in AP. Operating characteristics evaluates whole process starting from analysis option"
#| label: analysis-diagram
library(png)
library(grid)
img <- readPNG('MRCT-Flow.png')
grid.raster(img)
```



# Weibull AFT/Cox model with biomarker effects

## Brief Review
For the Weibull distribution with shape parameter $\nu$ and scale parameter $\theta$ the
density, cdf, survival, hazard and cumulative hazard functions are:
\begin{eqnarray*}
f(t) &=& \nu \theta^{-\nu}t^{\nu-1}\exp(-(t/\theta)^{\nu}), \cr F(t)
&=& \int_{0}^{t}\nu \theta^{-\nu}s^{\nu-1}\exp(-(s/\theta)^{\nu})ds
\cr &=& \int_{0}^{(t/\theta)^{\nu}}e^{-w}dw \cr & &
(w=(s/\theta)^{\nu},dw=\nu s^{\nu-1}\theta^{-\nu}ds) \cr
&=&1-\exp(-(t/\theta)^{\nu}), \cr S(t) &=& \exp(-(t/\theta)^{\nu}),
\cr \lambda(t)&=&\nu \theta^{-\nu}t^{\nu-1}, \cr \Lambda(t) &=&
-\log(S(t))=(t/\theta)^{\nu}.
\end{eqnarray*}
Note that here we define the density to correspond with R's definition.
For shape parameter $\nu \in (0,1)$ the hazard is strictly decreasing in $t \geq 0$,
whereas for $\nu >1$ the hazard is strictly increasing in $t \geq 0$.  

*Note*: $\Lambda(T) \sim E(1)$

The cumulative hazard function $\Lambda(\cdot)$ evaluated at $T$, $\Lambda(T)$, as a random variable has cdf
$$\eqalign{ \Pr(\Lambda(T) \leq t) &=\Pr(-\log(1-F(T)) \leq t) =\Pr(1-F(T) \geq e^{-t}) \cr
&=\Pr(T \leq F^{-1}(1-e^{-t})) =F(F^{-1}(1-e^{-t}))  = 1-e^{-t}, \cr}$$
\noindent
which is the CDF of the exponential distribution, $E(1)$ (say). 


In the following we use 

\begin{equation}
\tag{1}
\Lambda(T) \sim E(1)
\end{equation}
\noindent
to represent the Weibull regression model as an AFT model which is also a Cox model. 

Now, $\Lambda(T)=(T/\theta)^{\nu}$ and write
$Q=-\log(S(T))=\Lambda(T)=(T/\theta)^{\nu}$, where from (1), $Q \sim E(1)$.  That is
$\log(Q)=\nu(\log(T)-\log(\theta))$ can be expressed as

\begin{equation}
\tag{2}
\log(T)=\log(\theta)+ \tau \log(Q) := \log(\theta) + \tau \epsilon,
\end{equation}
\noindent
where $\tau=1/\nu$ and it is easy to show that $\epsilon=\log(Q)$ has the ``extreme
value'' distribution with density $f_{\epsilon}(x)=\exp(x-e^{x})$ for $x \in {\cal R}$.
Here the range of $\log(T) \in {\cal R}$ is un-restricted.  The *survReg* routine uses the parameterization $(2)$ and therefore estimates
$\log(\theta)$ and $\tau=1/\nu$.

To incorporate covariates $L$ (say) we specify
$$\lambda(t;L)=\big(\nu \theta^{-\nu}t^{\nu-1} \big) \exp(L'\beta)
:= \lambda_{0}(t)\exp(L'\beta),$$
\noindent
where $\lambda_{0}(t)$ is the hazard, say, for $T_{0} \sim \hbox{Weibull}(\nu,\theta)$.
This is a special case of the proportional hazards model. The chf (conditional chf with covariate vector
$L$) is $\Lambda(t;Z)=(t/\theta)^{\nu}\exp(L'\beta)$ so that
analogous to above this leads to the representation

\begin{equation}
\tag{3}
\log(T) =\log(\theta)+\tau[-L'\beta+\epsilon] =\log(\theta)+L'\gamma + \tau \epsilon,
\end{equation}
\noindent
where $\gamma=-\tau\beta$, with $\tau$ and $\epsilon$ defined in (2).  R *survReg* uses this AFT parameterization so that the estimated components of
$\gamma$, $\gamma_{p}$ say, are that of $-\tau\beta_{p}$ for $p=1,\ldots,m$ ($m$ is dimension of $\beta$).

When fitting the AFT model (3) via suvreg we therefore transform parameters $\hat\gamma$ to the Weibull hazard-ratio parameterization (2) via

\begin{equation}
\tag{4}
\hat\beta = -\hat\gamma / \hat{\tau}.
\end{equation}


As an illustration we compare the *survReg* model fits for the case-study dataset.  The following table below compares the Weibull *survReg* model fits with covariates Treat and Ecog1 (Ecog = 1 vs 0) where components of $\hat\gamma$ from model (3) are calculated according to *survReg* and $\hat\beta$ are formed via (4).  In the table below Weibull estimates of $\hat\beta$ are compared to Cox model versions.
  
```{r Case-study}
# case-study example
df.case <- read.table("simdatasource/dfexplore.csv", header=TRUE, sep=",")
#names(df.case)
```


```{r Weibull-vs-Cox}
# Comparing Weibull vs Cox with case-study where outcomes are artifical
# This is just for illustration to show conversion of Weibull parameters from 
# AFT regression to Weibull hazard 
fit.weib_ex <- survreg(Surv(tte,pmax(event,1)) ~ treat + ecog1, dist='weibull', data=df.case)
tauhat <- fit.weib_ex$scale
# convert (treat,ecog1) regression parms to Weibull hazard-ratio
bhat.weib <- -(1)*coef(fit.weib_ex)[c(2,3)]/tauhat
# Compare to Cox 
fit.cox_ex <- coxph(Surv(tte,pmax(event,1)) ~ treat + ecog1, data=df.case)
res <- cbind(bhat.weib,coef(fit.cox_ex))
res <- as.data.frame(res)
colnames(res)<-c("Weibull","Cox")
res |> gt() |>
fmt_number(columns=1:2,decimals=6) |>
tab_header(title="Comparing Weibull to Cox hazard ratio estimates",
subtitle="Case-study dataset with artificial outcomes")

```



## Biomarker effects with spline model  

We now outline how potential outcomes are simulated according to parameters fit to the case-study dataset but with parameters specified to induce biomarker effects.  That is, causal treatment effects (on log(hazard-ratio) scale) that follow 
a spline model according to patterns where biomarker effects increase with biomarker levels;  Including various degrees of limited treatment effects for low biomarker levels.

We first consider a Weibull model with treatment and a single biomarker covariate $Z$ where we write the linear predictor of the Cox model $L'\beta$ (say) as 

\begin{equation}
\tag{5}
L'\beta  := \beta_{1}\hbox{Treat} + \beta_{2}\hbox{Z} + \beta_{3}\hbox{Z}\hbox{Treat} + \beta_{4}(\hbox{Z}-k)I(\hbox{Z}>k) + \beta_{5}(\hbox{Z}-k)I(\hbox{Z}>k)\hbox{Treat}.
\end{equation}

Following the potential-outcome approach let $l_{x,z}$ denote subject's hazard-function "had they followed treatment regimen $Treat=x$ while having biomarker level $Z=z$".   That is, for subject with biomarker level $Z=z$ we can simulate 
their survival outcomes under both treatment ($x=1$) and control ($x=0$) conditions.   Let $\beta^{0} = (\beta_{1}^{0},\ldots,\beta_{5}^{0})'$ denote the true coefficients and denote the hazard function as 

$$\lambda_{x,z}(t) = \lambda_{0}(t)\exp(l_{x,z}), \quad \hbox{say}.$$  

Writing 

\begin{equation}
l_{x,z} = \beta^{0}_{1}x + \beta^{0}_{2}z + \beta^{0}_{3}zx + \beta^{0}_{4}(z-k)I(z>k) + 
 \beta^{0}_{5}(z-k)I(z>k)x,
\end{equation}
\noindent
the log of the hazard ratio for biomarker level $z$ under treatment ($x=1$) relative to control ($x=0$) is given by 

\begin{equation}
\tag{6}
\psi^{0}(z) := \log(\lambda_{1,z}(t)/\lambda_{0,z}(t)) = \beta^{0}_{1} + \beta^{0}_{3}z + \beta^{0}_{5}(z-k)I(z>k).
\end{equation}


## Causal log-hazard-ratio

The log(hazard-ratio) for biomarker level $z$ is a linear function of $z$ with a change-point (in slope) at $z=k$ given by 

\begin{eqnarray*}
\psi^{0}(z) &=& \beta^{0}_{1} + \beta^{0}_{3}z, \quad \hbox{for} \ z \leq k, \cr
            &=& \beta^{0}_{1} + \beta^{0}_{3}z + \beta^{0}_{5}(z-k), \quad \hbox{for} \ z > k.
\end{eqnarray*}

Log hazard-ratio parameters $(\beta^{0}_{1},\beta^{0}_{3},\beta^{0}_{5})$ can be chosen to generate "treatment effect patterns" by specifying $\psi^{0}(z)$ values at 
$z=0$, $z=k$, and $z=\zeta$ for $\zeta > k$.  For specified $\psi^{0}(0)$, $\psi^{0}(k)$, and $\psi^{0}(\zeta)$ we have 

\begin{eqnarray}
\beta^{0}_{1} &=& \psi^{0}(0), \cr
\beta^{0}_{3} &=& {(\psi^{0}(k) - \beta^{0}_{1}) \over k}, \cr
\beta^{0}_{5} &=& {(\psi^{0}(\zeta) - \beta^{0}_{1} - \beta^{0}_{3}\zeta) \over (\zeta -k)}.
\end{eqnarray}


The function *get_dgm_stratified* generates $\psi^{0}(z)$ according to desired "biomarker treatment effect patterns" as follows.

- Let $X$ and $Z$ denote the treatment and biomarker variables in the case-study dataset and for specified $k$, form the covariates $L:=(X,Z,ZX,(Z-k)I(Z>k),(Z-k)I(Z>k)X))$;
- Fit the Weibull model (recall on AFT scale) to get $\log(\hat\theta)$, $\hat\tau=1/\hat{\nu}$, and $\hat\gamma$ corresponding to $L$;
- $\hat\gamma$ is in terms of the AFT parameterization given by model (3)
- Next transform to the Weibull (Cox) log hazard-ratio parameterization (4): $\hat\beta = -\hat\gamma/\hat\tau$
- Set "true" parameters $\theta^{0}=\hat\theta$, and $\tau^{0}=\hat\tau$
- Initialize the "true" parameter $\beta^{0} = \hat\beta$ and re-define parameters 1, 3, and 5 in order to satisfy specified $\psi^{0}(0)$, $\psi^{0}(k)$, and $\psi^{0}(\zeta)$:
$\beta^{0}[1] = \psi^{0}(0)$, $\beta^{0}[3] = (\psi^{0}(k) - \beta^{0}[1])/k$, and $\beta^{0}[5] = (\psi^{0}(\zeta) - \beta^{0}[1] - \beta^{0}[3]\zeta)/(\zeta -k)$;
- Form corresponding $\gamma^{0}= -\beta^{0}\tau^{0}$
- For simulations we use the AFT parameterization (3) to generate $\log(T)$ outcomes according to $\log(T) = \log(\theta^{0}) + L'\gamma^{0} + \tau^{0}\epsilon$ where recall $\epsilon$ has the ``extreme value'' distribution.


The inputs of the *get_dgm_stratified* function are: 

- The case-study dataset ("df")
- "knot"=$k$, "zeta"=$\zeta$, and "log.hrs"= ($\psi^{0}(0),\psi^{0}(k),\psi^{0}(\zeta))$
- Note that *get_dgm_stratified* allows for outcomes to follow a stratified Weibull (Cox) models in which case the log(hazard-ratio) effects will depend on the strata, "strata_tte", where the default is non-stratified ("strata_tte=NULL")
- If stratified the $\tau$ parameters and hence $\gamma^{0}$ depend on the strata
- If stratified, then to simplify, the $\gamma^{0}$ effects are calculated based on the median of the $\tau^{0}$'s ($\tau^{0}=\tau_{med}$, say).  



To-do --> explain outputs and how used in *draw_sim_stratified* ...


## Example where treatment effects increase with increasing biomarker 

We assume that $z=0$ indicates biomarker negative values with $z>1$ indicating positive levels.  As an example, suppose the log(hazard ratio) at
$z=0$ is $\psi^{0}(0)=\log(3)$ and decreases linearly for $z \leq 5$ (with slope $\beta_{3}^{0}$) such that at $z=5$, $\psi^{0}(5)=\log(1.25)$ and 
for $z>5$ decreases linearly (with a change in slope and intercept) such that at $z=10$, $\psi^{0}(10) = \log(0.5)$.  In the following we will describe summary measures of the treatment effects as a function of increasing [or decreasing] values of the biomarker.  In this example treatment is detrimental 
for lower values of the biomarker with treatment effects increasing fairly quickly as the biomarker increases with an overall ``average hazard ratio'' (AHR) of $\approx 0.74$ (see Working Example below).  


## Including prognostic factors $W$

We extend the model to include a baseline prognostic factor $W$ within $L$ (effect parameter $\beta^{0}_{w}$) where

\begin{equation}
l_{x,z,w} = \beta^{0}_{1}x + \beta^{0}_{2}z + \beta^{0}_{3}zx + \beta^{0}_{4}(z-k)I(z>k) + 
 \beta^{0}_{5}(z-k)I(z>k)x + \beta^{0}_{w}w.
\end{equation}


Note that defining $m(1,z,w)$ [$m(0,z,w)$)] as the conditional expected value of $\log(T)$ with treatment set to experimental [control] given biomarker level $Z=z$ and $W=w$ 

$$\psi^{0}(z,w) = \{m(0,z,w)-m(1,z,w)\}/\tau$$
\noindent
which corresponds to the difference in the (true) means of the potential outcomes under experimental and control conditions ($\log(T[0,z,w]$ versus $\log(T[1,z,w]$, say) 


# Specifying biomarker treatment effects

## Average hazard ratios (AHRs)

We define the *biomarker average hazard ratio (AHR)* as the expected value of $\psi^{0}(\cdot)$ across "increasing biomarker" sub-populations.  For example, $\hbox{AHR}(2^{+})$ represents the AHR for subjects with biomarker values $\geq 2$ via  


\begin{equation}
\tag{6}
\hbox{AHR}(z^{+}):= \exp\left\{E_{Z \geq z} \psi^{0}(Z) \right\};
\end{equation}

With the opposite direction -- across "decreasing biomarker" sub-populations

\begin{equation}
\tag{7}
\hbox{AHR}(z^{-}):= \exp\left\{E_{Z \leq z} \psi^{0}(Z) \right\}.
\end{equation}

Here the expectations are over the distribution of $Z$, however if there is a prognostic factor $W$ then the expectations are over $(Z,W)$.  In our calculations we calculate empirical averages.


## Controlled direct-effect (CDE) versions (@ACR_2015)

Averaging across hazard ratios:  Define the hazard (omitting baseline hazard) $\theta^{x}(z,w) = \exp(l(x,z,w))$ setting treatment to $X=x$ given $Z=z$ and $W=w$.  Aalen et al. (@ACR_2015) define the controlled direct-effect (CDE) as the ratio of the expected hazards.  Here we consider the above cumulative versions (omitting possible dependence on $W$).   Let $\bar\theta^{x}(z+) = E_{Z \geq z}\theta^{x}(Z)$ for $x=0,1$, and define

$$\hbox{CDE}(z^{+}):= \bar\theta^{1}(z+)/\bar\theta^{0}(z+), \quad \hbox{and} \quad \hbox{CDE}(z^{-}):= \bar\theta^{1}(z-)/\bar\theta^{0}(z-)$$



```{r}
df.case <- within(df.case,{
  z <- pmin(biomarker,20)  # eg, PDL-1 expression truncated at 20
  CTregimen <- ifelse(mAD_CTregimen1==0,"socA","socB") # Outcome stratification (Weibull/Cox model)
  })

# These are confounders (fixed) that were characteristics from observed data
confounders.name <- c("age","biomarker","male","EU_region","AP_region","histology","surgery","ecog1","mAD_CTregimen1")
```


## Compare 'AHR' versions under various biomarker specifications 

### Under PH (uniform effects, hr=0.7), with strong prognostic effect $\beta_{w}= -\log(5)$


```{r}
# Outcome model is NON-stratified (strata_tte=NULL)
# PH 

log.hrs <- log(c(0.7,0.7,0.7))

dgm <- get_dgm_stratified(df=df.case,zeta=10,knot=5,log.hrs=log.hrs,strata_tte=NULL,details=FALSE)

# Simulate with randomization factors "stratum" 
# Note that strata_tte="CTregimen" is a simplification (collapsing over region and metastatic) of "stratum"
# xname="AP_region" specifies prognostic effect factor
# bx=log(5) denotes log(hazard ratio) relative effect with respect to xname factor

# Draw very large-sample to get approximation to bias
# Note: return_df will NOT return the large simulated dataset but the "large-sample" estimates as a list
# checking=TRUE will compare the estimated tau (stratified) parameters based on the case-study dataset to 
# the versions based on the simulated dataset
# details=TRUE will output the approximations to Cox model estimators from several models ("population summaries")
# return_df =FALSE will return the "population summaries" as a list as well as the biomarker AHR functional (see below)

pop_summary1 <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=100000,wname="AP_region",bw=-log(5),strata_rand="stratum",checking=TRUE,details=TRUE,return_df=FALSE)

```

```{r,  fig.width=12, fig.height=10}
plot_AHRs(popsummary=pop_summary1,dfcase=df.case)
```




### Under ''moderate'' biomarker effects 


```{r}
# Moderate biomarker
log.hrs <- log(c(0.9,0.8,0.65))
dgm <- get_dgm_stratified(df=df.case,zeta=10,knot=5,log.hrs=log.hrs,strata_tte=NULL,details=FALSE)
# Strong prognostic factor 
pop_summary5 <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=100000,wname="AP_region",bw=-log(5),strata_rand="stratum",checking=TRUE,details=TRUE,return_df=FALSE)

```


```{r,  fig.width=12, fig.height=10}

plot_AHRs(popsummary=pop_summary5,dfcase=df.case)

```


## Calculating causal 'AHR' effects for subgroups
To evaluate the estimation properties of our procedures we calculate the empirical average within subgroup $\cal G$ (say) which can represent 
the estimated subgroup or can be the ITT population (or any underlying subgroup).  That is, for each subject with characteristics 
${\bf L}=(Z,W)$ say, we have their individual log hazard-ratio contrast $\psi^{0}({\bf L})$; We estimate 
$\hbox{AHR}({\cal G}):= \exp\left\{E_{\cal G} \psi^{0}\right\}$ by calculating the empirical average 

\begin{equation}
\tag{8}
\hbox{AHR}({\cal G}):= \exp(\bar\beta({\cal G})), \quad \hbox{where} \quad \bar\beta({\cal G}) =(1/n({\cal G}))\sum_{g \in {\cal G}} \psi^{0}({\bf L}_{g}), 
\end{equation}
with $n({\cal G})$ denoting the size of ${\cal G}$.


# Working example for subgroup identification strategies

In the remainder we will be evaluating the performance of subgroup identification procedures when the underlying biomarker profile (``differential treatment effect mechanism'') is quite strong 
and challenges with establishing consistency within the AP region are induced by a strong prognostic effect ($\beta_{w}=-\log(5)$).

Note that in the following we draw a very large sample ($n = 100,000$) to approximate the asymptotic properties of Cox model analyses based on ITT and AP region subgroups (non-AP is denoted $W=0$, and AP is $W=1$).  The code outputs the overall AHR which represent (8) applied to the ITT population (i.e., 
$\cal G$=ITT) as well as (8) applied to non-AP and AP subgroups ("AHR W=1", and "AHR W=0", resp.).


```{r}
# Outcome model is NON-stratified (strata_tte=NULL)
# Strong biomarker 

log.hrs <- log(c(3,1.25,0.50))

dgm <- get_dgm_stratified(df=df.case,zeta=10,knot=5,log.hrs=log.hrs,strata_tte=NULL,details=FALSE)

# Simulate with randomization factors "stratum" 
# Note that strata_tte="CTregimen" is a simplification (collapsing over region and metastatic) of "stratum"
# xname="AP_region" specifies prognostic effect factor
# bx=log(5) denotes log(hazard ratio) relative effect with respect to xname factor

# Draw very large-sample to get approximation to bias
# Note: return_df will NOT return the large simulated dataset but the "large-sample" estimates as a list
# checking=TRUE will compare the estimated tau (stratified) parameters based on the case-study dataset to 
# the versions based on the simulated dataset
# details=TRUE will output the approximations to Cox model estimators from several models ("population summaries")
# return_df =FALSE will return the "population summaries" as a list as well as the biomarker AHR functional (see below)

# Strong prognostic factor 
pop_summary <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=100000,wname="AP_region",bw=-log(5),strata_rand="stratum",checking=TRUE,details=TRUE,return_df=FALSE)

# An illustrative simulated example dataset
df_example <- draw_sim_stratified(dgm=dgm,ss=123,wname="AP_region",bw=-log(5),strata_rand="stratum")

# Subgroups identified with nonAP population
df_nonAP <- subset(df_example,AP_region==0)
# Applied to AP
df_AP <- subset(df_example,AP_region==1)
df_ITT <- df_example


# True causal AHR
ahr_true = c(round(pop_summary$AHR,4))
# AHR Cox un-adjusted (only treatment)
ahr_unadj = c(round(pop_summary$ITT_unadj,4))
# Treatment plus randomization stratificaton sR
ahr_sR = c(round(pop_summary$ITT_sR,4))
# sR including adjustment by x
ahr_sRw = c(round(pop_summary$ITT_sRw,4))

# Relative biases 
bias_unadj <-round(100*c(pop_summary$ITT_unadj-pop_summary$AHR)/pop_summary$AHR,1)
bias_sR <-round(100*c(pop_summary$ITT_sR-pop_summary$AHR)/pop_summary$AHR,1)
bias_sRw <-round(100*c(pop_summary$ITT_sRw-pop_summary$AHR)/pop_summary$AHR,1)

# The bias within X=1 population
ahr_w1 = c(round(pop_summary$W_1,4))
bias_w1 <-round(100*c(pop_summary$W_1-pop_summary$AHR)/pop_summary$AHR,1)


```


```{r,  fig.width=12, fig.height=10}

plot_AHRs(popsummary=pop_summary,dfcase=df.case)

```



- Overall Population (large-sample) --
  - The overall true (causal) average hazard-ratio (AHR) = `r ahr_true`
  - The $\approx$ asymptotic limit of un-adjusted Cox (only treatment) = `r ahr_unadj` 
  - The $\approx$ asymptotic limit of stratified Cox (treatment and stratified by *stratum*) = `r ahr_sR`
  - The $\approx$ asymptotic limit of stratified Cox with adjustment by $x$ = `r ahr_sRw`
  
The relative over-estimation biases for the Cox model estimates when un-adjusted, stratified (by randomization factors), and stratified + adjusted by the strong prognostic factor $X$ are: `r bias_unadj`\%, `r bias_sR`\%, and 
`r bias_sRw`\%, respectively.  

## Dilution of marginal Cox model within AP region

Moreover, within the *AP_region* ($X=1$) the $\approx$ limit of the Cox model estimator based on the (*AP_region*) regional local data is `r ahr_w1` with corresponding bias of `r bias_w1`\%.   


**We would therefore expect challenges with establishing consistency based on evaluating the local regional data.**


Lets check if there is any imbalance (asymptotically) by biomarker by drawing a large sample (N=100,000) and viewing summary tables: 
```{r}
# This time return large dataset (corresponding to that used in the above asymptotic approximations)
dflarge <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=100000,wname="AP_region",bw=-log(5),strata_rand="stratum",checking=FALSE,details=FALSE,return_df=TRUE)
# create factor versions of treatment and AP region
dflarge$trtsim <- ifelse(dflarge$treat.sim==1,"Experimental","Control")
dflarge$apregion <- ifelse(dflarge$AP_region==1,"AP","non-AP")
dflarge$bm_lt2 <- ifelse(dflarge$biomarker<2,"biomarker<2","biomarker>=2")
table1( ~ biomarker + bm_lt2 | trtsim, data=dflarge, caption=c("ITT by treatment arm"))
table1( ~ biomarker + bm_lt2 | apregion, data=dflarge, caption=c("ITT by AP region"))
table1( ~ biomarker + bm_lt2 | trtsim, data=subset(dflarge,AP_region==1), caption=c("AP region by treatment arm"))
rm("dflarge")
```

Next, a simulated example with same sample size as the case-study   

-  Here we simulated a study according to the biomarker effects described above
-  Summarize the non-AP and AP populations  
-  Apply subgroup identification procedures to the non-AP data and apply to AP  



```{r KMitt,fig.width=8, fig.height=6}
# non-stratified
kmH.fit<-KM.plot.2sample.weighted(df=df_example, 
tte.name="y.sim", event.name="event.sim", treat.name="treat.sim",
risk.set=TRUE, by.risk=6,
details=TRUE,
Xlab="Months",Ylab="Overall survival",
arms=c("Experimental","Control"))
title(main="ITT Population")

```



## Now evaluate non-AP and AP ---> AP is "diluted"


```{r KMregions,fig.width=12, fig.height=6}
# non-stratified
par(mfrow=c(1,2))
kmH.fit<-KM.plot.2sample.weighted(df=df_nonAP, 
tte.name="y.sim", event.name="event.sim", treat.name="treat.sim",
risk.set=TRUE, by.risk=6, details=FALSE,
show.logrank=FALSE, put.legend.arms="top",
Xlab="Months",Ylab="Overall survival",
arms=c("Experimental","Control"))
title(main="Non-AP Population")

kmH.fit<-KM.plot.2sample.weighted(df=df_AP, 
tte.name="y.sim", event.name="event.sim", treat.name="treat.sim",
risk.set=TRUE, by.risk=6, details=FALSE,
show.logrank=FALSE, put.legend.arms="top",
Xlab="Months",Ylab="Overall survival",
arms=c("Experimental","Control"))
title(main="AP Population")
```

Note that results stratified by randomization (or true outcome strata) are similar

```{r KMregions2,fig.width=12, fig.height=6, eval=FALSE}
# stratified
par(mfrow=c(1,2))
kmH.fit<-KM.plot.2sample.weighted(df=df_nonAP, 
tte.name="y.sim", event.name="event.sim", treat.name="treat.sim",strata.name="strata.simR",
risk.set=TRUE, by.risk=6, details=FALSE,
show.logrank=FALSE, put.legend.arms="top",
Xlab="Months",Ylab="Overall survival",
arms=c("Experimental","Control"))
title(main="Non-AP Population")

kmH.fit<-KM.plot.2sample.weighted(df=df_AP, 
tte.name="y.sim", event.name="event.sim", treat.name="treat.sim",strata.name="strata.simR",
risk.set=TRUE, by.risk=6, details=FALSE,
show.logrank=FALSE, put.legend.arms="top",
Xlab="Months",Ylab="Overall survival",
arms=c("Experimental","Control"))
title(main="AP Population")
```

Check balance within simulated trial


```{r}
# This time return large dataset (corresponding to that used in the above asymptotic approximations)
df <- df_example
# create factor versions of treatment and AP region
df$trtsim <- ifelse(df$treat.sim==1,"Experimental","Control")
df$apregion <- ifelse(df$AP_region==1,"AP","non-AP")
df$bm_lt2 <- ifelse(df$biomarker<2,"biomarker<2","biomarker>=2")
table1( ~ biomarker + bm_lt2 | trtsim, data=df, caption=c("ITT by treatment arm"))
table1( ~ biomarker + bm_lt2 | apregion, data=df, caption=c("ITT by AP region"))
table1( ~ biomarker + bm_lt2 | trtsim, data=subset(df,AP_region==1), caption=c("AP region by treatment arm"))
```

## Looking at 9 simulated AP trials when AP region is a strong factor ($\beta_{w} = -\log(5)$) and null ($\beta_{w}=0$)

### Nine simulated AP realizations under $\beta_{w}= -\log(5)$

```{r, fig.width=12, fig.height=12}

par(mfrow=c(3,3))

for(ss in 1:9){
df_ex <- draw_sim_stratified(dgm=dgm,ss=ss,wname="AP_region",bw=-log(5),strata_rand="stratum")
kmH.fit<-KM.plot.2sample.weighted(df=subset(df_ex,AP_region==1), 
tte.name="y.sim", event.name="event.sim", treat.name="treat.sim",strata.name="strata.simR",
risk.set=TRUE, by.risk=6, details=FALSE, show.med=FALSE, cox.cex= 1.25,
show.logrank=FALSE, put.legend.arms="top",
Xlab="Months",Ylab="Overall survival",
arms=c("Experimental","Control"))
title(main="AP Population")
}


```


### Nine simulated AP realizations under $\beta_{w} = 0$


```{r, fig.width=12, fig.height=12}

par(mfrow=c(3,3))

for(ss in 1:9){
df_ex <- draw_sim_stratified(dgm=dgm,ss=ss,wname="AP_region",bw=-log(1),strata_rand="stratum")
kmH.fit<-KM.plot.2sample.weighted(df=subset(df_ex,AP_region==1), 
tte.name="y.sim", event.name="event.sim", treat.name="treat.sim",strata.name="strata.simR",
risk.set=TRUE, by.risk=6, details=FALSE,show.med=FALSE, cox.cex= 1.25,
show.logrank=FALSE, put.legend.arms="top",
Xlab="Months",Ylab="Overall survival",
arms=c("Experimental","Control"))
title(main="AP Population")
}


```

## Compare non-parametric (cubic-spline) fit with true $\psi^{0}$ 

For ITT dataset

```{r}
# ydel (default=0.25) is how much room to provide in space for counts
temp <- cox_cs_fit2(df=df_example,tte.name="y.sim",event.name="event.sim",treat.name="treat.sim",
strata.name="strata.simO",z.name=c("z"), details=FALSE ,boots=0, xlab=c("Biomarker"),maxz = 25, ydel=0.5, show_plot=TRUE,truebeta.name="loghr.po")
title("ITT")
```


Within non-AP and AP regions


```{r,fig.width=12, fig.height=6}

par(mfrow=c(1,2))

temp <- cox_cs_fit(df=df_nonAP,tte.name="y.sim",event.name="event.sim",treat.name="treat.sim",
strata.name="strata.simO",z.name=c("z"), details=FALSE ,boots=0, xlab=c("Biomarker"),maxz = 25, ydel=0.5, show_plot=TRUE,truebeta.name="loghr.po")
title("Non-AP")


temp <- cox_cs_fit(df=df_AP,tte.name="y.sim",event.name="event.sim",treat.name="treat.sim",
strata.name="strata.simO",z.name=c("z"), details=FALSE ,boots=0, xlab=c("Biomarker"),maxz = 25, ydel=0.5, show_plot=TRUE,truebeta.name="loghr.po")
title("AP")

```

## Examine basic forest plot of non-AP data used for identification

```{r Forestplot, echo=TRUE, message=FALSE, warning=TRUE,fig.width=8, fig.height=12}

fp <- forestplot_baseline(df=df_nonAP,confounders.name=confounders.name,
arrow_text=c("favors Treatment","Control"),E.name=c("Treatment"),C.name=c("Control"),outcome.name="y.sim",event.name="event.sim",treat.name="treat.sim",
xlim = c(0.25, 2.0), ticks_at = c(0.65, 1.0, 1.3))

plot(fp$p)

aa <- paste(confounders.name, collapse=" + ")
aa <- paste0("~ " ,aa)
aa <- paste0(aa," | treat_name")
aa <- as.formula(aa)

table1( aa , data=df_nonAP, caption=c("Simulated dataset: non-AP (training data)"))

```

# Subgroup identification criteria:  Options for targeting subgroup effects

## Finding Non-AP subgroup with highest consistency rate (in favor of control)  

In the following we evaluate subgroups based on single-factors (e.g., "biomarker < J" say, or "Age <= 65", etc) and 
select the subgroup with the highest consistency rate with Cox hazard-ratio estimate meeting selection criteria;  The selected subgroup with the highest 
consistency rate meeting hazard ratio estimate criterion (log hazard-ratio estimate $\log(\hat\beta) \geq \log(0.90)$, say) where the "consistency rate" is at least $90$\%.

Here we are seeking to identify subgroups where the hazard ratio estimate is at least $\log(0.90)$ which may indicate limited benefit.  Further, subgroup candidates are ranked according 
to highest hazard ratio estimate and highest consistency rate.

*This is an option, and we discuss alternatives below*  


To do --->  Refer criterion (hazard ratio thresholds) to equations in paper @Leon_2024


```{r ForestSearch_k1_10_hr, echo=TRUE, message=FALSE, warning=TRUE, fig.width=8, fig.height=6, eval=FALSE}

# Show the first (up to) 10 subgroups ("showten_subgroups=TRUE") meeting identification criteria 
# Subgroups based on only single-factors ("maxk=1")

# Note that sg_focus= "hr" sorts the subgroups by consistency rate and largest hazard-ratio estimates  
# Selects the subgroup with the largest consistency rate;  If there are ties among consistency rates, then 
# the subgroup with largest HR estimate is selected estimate satisfying the (minimum) consisency criteria 
# (pconsistency.threshold)

# Note: we exclude cuts for biomarker which are "<=" in order to only consider "<" 
fs <- forestsearch(df.analysis=df_nonAP, df.test=df_AP, confounders.name=confounders.name,
outcome.name="y.sim",treat.name="treat.sim", event.name="event.sim",
potentialOutcome.name="loghr.po",
hr.threshold = 0.90, hr.consistency = 0.80, pconsistency.threshold = 0.90,
sg_focus = "hr",
showten_subgroups=TRUE, details=TRUE,
conf_force=c("age <= 65","biomarker < 1","biomarker < 2","biomarker < 3",
"biomarker < 4","biomarker < 5","biomarker < 6","biomarker < 7","biomarker < 8",
"biomarker < 9","biomarker < 10"), exclude_cuts = c("biomarker <="),
maxk=1, plot.sg=FALSE)

if(output) save(fs,file="output/sim3_k1tenhr.Rdata")

```

Display the first 10 subgroups and Kaplan-Meier plots within the estimated subgroups --

Notice that the first 10 subgroups are sorted by decreasing HR estimates and the **selected subgroup ** corresponds to the highest hazard ratio estimate.

```{r, echo=TRUE, message=FALSE, warning=TRUE}
if(loadit) load(file="output/sim3_k1tenhr.Rdata")
# Consistency for subgroups contained in "grp.consistency"
sg_tables(fs=fs,which_df="est",est_caption="Non-AP (training) estimates")$sg10_out

#sg_tables(fs=fs,which_df="est",est_caption="Non-AP (training) estimates",potentialOutcome.name="loghr.po")$tab_estimates

# Show detailed estimates for the next analysis
id_harm <- paste(fs$sg.harm,collapse=" & ")


```


The **selected subgroup** is `r id_harm`

Next plot the K-M curves within the estimated non-AP (training) subgroups.

```{r, fig.width=12, fig.height=10}
plot_found.subgroups(fs=fs, df=fs$df.est, title_itt=c("Non-AP ITT"))

```


## Finding Non-AP subgroup with smallest sample size meeting selection criteria  

We now select the **smallest subgroup** meeting:  log hazard-ratio estimate $\log(\hat\beta) \geq log(0.90)$ with "consistency rate" at least $90$\%.

```{r ForestSearch_k1_10_minSG, echo=TRUE, message=FALSE, warning=TRUE, fig.width=8, fig.height=6, eval=FALSE}

# sg_focus="MinSG" selects smallest subgroup meeting selection criteria

fs <- forestsearch(df.analysis=df_nonAP, df.test=df_AP, confounders.name=confounders.name,
outcome.name="y.sim",treat.name="treat.sim", event.name="event.sim",
potentialOutcome.name="loghr.po",
hr.threshold = 0.90, hr.consistency = 0.80, pconsistency.threshold = 0.90,
sg_focus = "minSG",
showten_subgroups=TRUE, details=TRUE,
conf_force=c("age <= 65","biomarker < 1","biomarker < 2","biomarker < 3",
"biomarker < 4","biomarker < 5","biomarker < 6","biomarker < 7","biomarker < 8",
"biomarker < 9","biomarker < 10"), exclude_cuts=c("biomarker <="),
maxk=1, plot.sg=TRUE)

if(output) save(fs,file="output/sim3_k1tenmin.Rdata")

```

```{r, echo=TRUE, message=FALSE, warning=TRUE}
if(loadit) load(file="output/sim3_k1tenmin.Rdata")

# Training summaries
SG_est <- sg_tables(fs=fs,which_df="est",est_caption="Non-AP (training) estimates",potentialOutcome.name="loghr.po")

id_harm <- paste(fs$sg.harm,collapse=" & ")
```

Top 10 (up to 10) subgroups meeting criteria:

```{r, echo=TRUE, message=FALSE, warning=TRUE}
# Top 10 (up to 10) SG's
SG_est$sg10_out
```

Estimates within training (un-adjusted).  Note that "AHR(po)" denotes (we refer to the log hazard ratio contrast as "potential-outcome") denotes 
equation (8) applied to the respective subgroups in the table (e.g., ${\cal G}$=ITT).


```{r, echo=TRUE, message=FALSE, warning=TRUE}
SG_est$tab_estimates
```

The **selected subgroup** is `r id_harm`

Next plot the K-M curves within the estimated non-AP subgroups.

```{r, fig.width=12, fig.height=10}
plot_found.subgroups(fs=fs, df=fs$df.est, title_itt=c("Non-AP ITT"))
```

Next, display the estimates within the AP population (testing data)

```{r, echo=TRUE, message=FALSE, warning=TRUE}
# Training summaries
SG_test <- sg_tables(fs=fs,which_df="testing",est_caption="AP (testing) estimates",potentialOutcome.name="loghr.po")
```

```{r, echo=TRUE, message=FALSE, warning=TRUE}
SG_test$tab_estimates
```

K-M curves applied to AP subgroups.

```{r, fig.width=12, fig.height=10}
plot_found.subgroups(fs=fs, df=fs$df.test, title_itt=c("AP ITT"))
```



Now, if we are only concerned with finding the selected subgroup then the above calculations can be much faster by removing the option to output the first 10 subgroups meeting the selection criteria ("showten_subgroups=FALSE" is the default).  This is the default and strongly 
recommended if running simulations or if there is interest in calculating adjusted hazard-ratio estimates for the selected subgroup.   However it is crucial to note that in our applications we are using the non-AP data for subgroup identification and interested in the estimates 
based on the AP local data.  Therefore the non-AP data represents the *training dataset* and the AP data represents the *testing dataset* where such adjustment is not necessary (that is, the AP data is not utilized for the subgroup selection).

Here we turn-off showing the first ten subgroups (default in forestsearch) and apply the estimated subgroups to the AP data.  

```{r, k1minSG}
# df.test = df_AP indicates that the selected subgroup identification flags 
# (based on df_nonAP analysis) will be appended to df_AP

# Speed is increased by first sorting by subgroup size and once a subgroup 
# meets the consistency criterion the overall criteria is met and the algorithm stops

fs <- forestsearch(df.analysis=df_nonAP, df.test=df_AP, 
confounders.name=confounders.name,
outcome.name="y.sim", treat.name="treat.sim", event.name="event.sim",
potentialOutcome.name="loghr.po",
hr.threshold = 0.90, hr.consistency = 0.80, pconsistency.threshold = 0.90,
sg_focus="minSG", details=TRUE,
conf_force=c("age <= 65","biomarker < 1","biomarker < 2","biomarker < 3",
"biomarker < 4","biomarker < 5","biomarker < 6","biomarker < 7","biomarker < 8",
"biomarker < 9","biomarker < 10"), exclude_cuts=c("biomarker <="), maxk=1)

if(output) save(fs,file="output/sim3_k1min.Rdata")

```


Apply estimated subgroup to the AP data ---


```{r, echo=TRUE, message=FALSE, warning=TRUE}
if(loadit) load(file="output/sim3_k1min.Rdata")

#sg_tables(fs=fs,which_df="est",est_caption="Non-AP (training) estimates",potentialOutcome.name="loghr.po")$tab_estimates

# Testing summaries
SG_test <- sg_tables(fs=fs,which_df="testing",est_caption="AP (testing) estimates",potentialOutcome.name="loghr.po")
```


```{r, echo=TRUE, message=FALSE, warning=TRUE}
SG_test$tab_estimates
```

K-M curves applied to AP subgroups (repeat of above).

```{r, fig.width=12, fig.height=10}
plot_found.subgroups(fs=fs, df=fs$df.test, title_itt=c("AP ITT"))
```


## Finding the largest non-AP subgroup with *strong benefit* 

Here we are *directly* targeting the largest subgroup with a strong benefit.  This is done by simply reversing the role of treatment so that we seek the largest subgroup for which the hazard ratio is below a clinical threshold (e.g., $\leq 0.6$).

```{r ForestSearch_k1_10_maxSG, echo=TRUE, message=FALSE, warning=TRUE, fig.width=8, fig.height=6, eval=FALSE}
# sg_focus="MaxSG" selects smallest subgroup meeting selection criteria
# Reverse role of treatment so that we find "worst control" subgroup
dfnew <- copy(df_nonAP)
dfnew$control.sim <- 1-dfnew$treat.sim

dftest <- copy(df_AP)
dftest$control.sim <- 1-dftest$treat.sim

fs_k1_10_maxSG <- forestsearch(df.analysis=dfnew, df.test=dftest, confounders.name=confounders.name,
outcome.name="y.sim",treat.name="control.sim", event.name="event.sim",
potentialOutcome.name="loghr.po",
est.scale="1/hr", hr.threshold = 1/0.6, hr.consistency = 1/0.7, pconsistency.threshold = 0.90,
sg_focus = "maxSG",
showten_subgroups=TRUE, details=TRUE,
conf_force=c("age <= 65","biomarker < 1","biomarker < 2","biomarker < 3",
"biomarker < 4","biomarker < 5","biomarker < 6","biomarker < 7","biomarker < 8",
"biomarker < 9","biomarker < 10"), exclude_cuts=c("biomarker <="),
maxk=1, plot.sg=TRUE)

if(output) save(fs_k1_10_maxSG,file="output/sim3_k1tenmax.Rdata")

```

```{r, echo=TRUE, message=FALSE, warning=TRUE}
if(loadit) load(file="output/sim3_k1tenmax.Rdata")
fs <- fs_k1_10_maxSG
# Testing summaries
SG_est <- sg_tables(fs=fs,which_df="est",est_caption="Non-AP (training) estimates",potentialOutcome.name="loghr.po")
```

Top 10 (up to 10) subgroups meeting criteria:

```{r, echo=TRUE, message=FALSE, warning=TRUE}
# Top 10 (up to 10) SG's
SG_est$sg10_out
```

Estimates within training (un-adjusted)


```{r, echo=TRUE, message=FALSE, warning=TRUE}
SG_est$tab_estimates
```

The **selected subgroup** is `r id_harm`

Next plot the K-M curves within the estimated non-AP subgroups.

```{r, fig.width=12, fig.height=10}
plot_found.subgroups(fs=fs, df=fs$df.est, title_itt=c("Non-AP ITT"))

```

Next, display the estimates within the AP population (testing data)

```{r, echo=TRUE, message=FALSE, warning=TRUE}
# Training summaries
SG_test <- sg_tables(fs=fs,which_df="testing",est_caption="AP (testing) estimates",potentialOutcome.name="loghr.po")
```

```{r, echo=TRUE, message=FALSE, warning=TRUE}
SG_test$tab_estimates
```

K-M curves applied to AP subgroups.

```{r, fig.width=12, fig.height=10}
plot_found.subgroups(fs=fs, df=fs$df.test, title_itt=c("AP ITT"))
```


## What happens under the ("null") where treatment is uniform?


```{r Simnull1}
# Null uniform benefit hr=0.7
log.hrs <- log(c(0.7,0.7,0.7))

dgm <- get_dgm_stratified(df=df.case,zeta=10,knot=5,log.hrs=log.hrs,strata_tte=NULL,details=TRUE)

df_null <- draw_sim_stratified(dgm=dgm,ss=1,wname="AP_region",bw=-log(5),strata_rand="stratum",checking=FALSE,details=FALSE,return_df=TRUE)
```

Note that under the null of uniform treatment the challenges of meeting consistency for the AP region are not prevalent in the first place ---


Nevertheless, let's see what happens when we apply the same subgroup identification criteria (to non-AP) as above.

```{r,fig.width=8, fig.height=6}
kmH.fit<-KM.plot.2sample.weighted(df=subset(df_null,AP_region==1), 
tte.name="y.sim", event.name="event.sim", treat.name="treat.sim",
risk.set=TRUE, by.risk=6, details=FALSE,
show.logrank=FALSE, put.legend.arms="top",
Xlab="Months",Ylab="Overall survival",
arms=c("Experimental","Control"))
title(main="AP Population")
```



```{r}
fs <- forestsearch(df.analysis=subset(df_null,AP_region==0),  
confounders.name=confounders.name,
outcome.name="y.sim", treat.name="treat.sim", event.name="event.sim",
potentialOutcome.name="loghr.po",
hr.threshold = 0.90, hr.consistency = 0.80, pconsistency.threshold = 0.90,
sg_focus="minSG", details=TRUE,
conf_force=c("age <= 65","biomarker < 1","biomarker < 2","biomarker < 3",
"biomarker < 4","biomarker < 5","biomarker < 6","biomarker < 7","biomarker < 8",
"biomarker < 9","biomarker < 10"), exclude_cuts=c("biomarker <="), maxk=1)

if(output) save(fs, file="output/sim3_k1min_null.Rdata")

```

```{r}
if(loadit) load("output/sim3_k1min_null.Rdata")
if(is.null(fs$sg.harm)) cat("Subgroup NOT found","\n")

```


We have been focusing on identifying subgroups based only on single factors, such as biomarker cuts, or other factors.   However, in theory, subgroups effects could involve more than a single factor (In our forestsearch algorithm this is set by *maxk=1*).

For illustration we now consider selecting the smallest subgroup among 2-factor combinations (*maxk=2*).


```{r ForestSearch_k2_10_minSG, echo=TRUE, message=FALSE, warning=TRUE, fig.width=8, fig.height=6, eval=FALSE}

# sg_focus="MinSG" selects smallest subgroup meeting selection criteria

fs <- forestsearch(df.analysis=df_nonAP, confounders.name=confounders.name,
outcome.name="y.sim",treat.name="treat.sim", event.name="event.sim",
potentialOutcome.name="loghr.po",
hr.threshold = 0.90, hr.consistency = 0.80, pconsistency.threshold = 0.90,
sg_focus = "minSG",
showten_subgroups=TRUE, details=TRUE,
conf_force=c("age <= 65","biomarker < 1","biomarker < 2","biomarker < 3",
"biomarker < 4","biomarker < 5","biomarker < 6","biomarker < 7","biomarker < 8",
"biomarker < 9","biomarker < 10"), exclude_cuts=c("biomarker <="),
maxk=2, plot.sg=TRUE)

if(output) save(fs,file="output/sim3_k2tenmin.Rdata")

```


```{r, echo=TRUE, message=FALSE, warning=TRUE}
if(loadit) load(file="output/sim3_k2tenmin.Rdata")
# Testing summaries
SG_est <- sg_tables(fs=fs,which_df="est",est_caption="Non-AP (training) estimates",potentialOutcome.name="loghr.po")
```

Top 10 (up to 10) subgroups meeting criteria:

```{r, echo=TRUE, message=FALSE, warning=TRUE}
# Top 10 (up to 10) SG's
SG_est$sg10_out
```


# Simulations

Next we setup simulations to evaluate ability to identify subgroups based on non-AP data and 
the accuracy to estimate treatment effects in the AP data.

The simulation function *mrct_APregion_sims* simulates (**n_sims** times) from specified data-generating-model (**dgm**):

-  ITT sample is drawn with subgroups identified (as above)
-  Cox estimates based on the AP local data;   
-  Subgroups identified based on non-AP and applied to AP
-  For identified subgroups the Cox model estimates are created (**hr_sg**) 
   -  Note that if a subgroup is NOT found then "subgroups" are taken as the entire AP population  
   -  That is, if a subgroup is NOT identified then we interpret the **hr_sg** as the reported AP estimates which would be the overall AP (since no subgroup is found)
   -  In tables below we also summarize **hr_sg_null** where we define **hr_sg** to be missing if no subgroup is found 
       -  This represents the AP subgroup estimates conditional on a subgroup being identified

```{r Simsetup}
# Simulation function to automate
mrct_APregion_sims <- function(dgm,n_sims,wname="AP_region",bw=-log(5),sg_focus="minSG",maxk=1, 
hr.threshold=0.90, hr.consistency=0.80, pconsistency.threshold=0.9,details=FALSE){
t.start<-proc.time()[3]
  # message for backends
  # borrowed from simtrials (sim_fixed_n)
  if (!is(plan(), "sequential")) {
    # future backend
    message("Using ", nbrOfWorkers(), " cores with backend ", attr(plan("list")[[1]], "class")[2])
  } else if (foreach::getDoParWorkers() > 1) {
    message("Using ", foreach::getDoParWorkers(), " cores with backend ", foreach::getDoParName())
    message("Warning: ")
    message("doFuture may exhibit suboptimal performance when using a doParallel backend.")
  } else {
    message("Backend uses sequential processing.")
  }
results <- foreach(
    sim = seq_len(n_sims),
    .options.future=list(seed=TRUE),
    .combine="rbind", 
    .errorhandling="pass"
  ) %dofuture% {
# simulate data
dfs <- draw_sim_stratified(dgm=dgm,ss=sim,wname=wname,bw=bw,strata_rand="stratum")
# Subgroups identified with nonAP population
df_nonAP <- subset(dfs,AP_region==0)
# Applied to AP
df_AP <- subset(dfs,AP_region==1)
# For subgroup identified estimates     
# Return NA if not estimable
# First, AP results (does not require subgroup identification)
# Initialize to missing
n_test <- c(nrow(df_AP))
n_train <- c(nrow(df_nonAP))
n_itt <- c(nrow(dfs))

fit <- summary(coxph(Surv(y.sim,event.sim) ~ treat.sim, data=df_nonAP))$conf.int
hr_train <- c(fit[1])
rm("fit")

# Overall (ITT) standard stratified (by randomization strata)
fit <- summary(coxph(Surv(y.sim,event.sim) ~ treat.sim + strata(strata.simR), data=dfs))$conf.int
hr_itt <- c(fit[1])
rm("fit")

# Overall (ITT) standard stratified (by X)
fit <- summary(coxph(Surv(y.sim,event.sim) ~ treat.sim + strata(AP_region), data=dfs))$conf.int
hr_ittX <- c(fit[1])
rm("fit")


hr_test<- NA
any_found <- 0.0
sg_found <- "none"
n_sg <- n_test
hr_sg <- NA
prev_sg <- 1.0

# Restrict to AP estimable
fitAP <- try(summary(coxph(Surv(y.sim,event.sim) ~ treat.sim, data=df_AP))$conf.int,TRUE)
if(!inherits(fitAP,"try-error")){
hr_test <- c(fitAP[1])
rm("fitAP")
# For identified subgroups?
# initialize to testing
any_found <- 0.0 
sg_found <- "none" 
n_sg <- n_test 
prev_sg <- 1.0 # Prevalence of AP: n_sg/n_all
hr_sg <- NA
# Potential outcome
POhr_sg <- exp(mean(df_AP$loghr.po))
# Note: if not found then n_sg=n_all, hr_sg=hr_all 
fs_temp <- try(forestsearch(df.analysis=df_nonAP, df.test=df_AP, 
confounders.name=confounders.name,
outcome.name="y.sim", treat.name="treat.sim", event.name="event.sim",
potentialOutcome.name="loghr.po",
hr.threshold = hr.threshold, hr.consistency = hr.consistency, 
pconsistency.threshold = pconsistency.threshold,
sg_focus=sg_focus, 
conf_force=c("age <= 65","biomarker < 1","biomarker < 2","biomarker < 3",
"biomarker < 4","biomarker < 5","biomarker < 6","biomarker < 7","biomarker < 8",
"biomarker < 9","biomarker < 10"), exclude_cuts =c("biomarker <="), maxk=maxk),TRUE)
# FS with error output NA's (set above)
# FS without error
if(!inherits(fs_temp,"try-error")){
# FS done (w/o error) but NO sg found (replace NAs above with what is calculated)
if(is.null(fs_temp$sg.harm)){
# consider "sg = AP"  
any_found <- 0.0
sg_found <- "none"
n_sg <- n_test
hr_sg <- hr_test
POhr_sg <- exp(mean(df_AP$loghr.po))
prev_sg <- 1.0
}
# If sg found
if(!is.null(fs_temp$sg.harm)){
any_found <- 1.0  
df_test <- fs_temp$df.test 
sg_found <- c(paste(fs_temp$sg.harm,collapse=" & "))
if(details & sim <= 10) cat("Simulation subgroup found:",c(sim,sg_found),"\n")  
df_sg <- subset(df_test,treat.recommend==1)
n_sg <- c(nrow(df_sg))
prev_sg <- c(n_sg/n_test)
# Subgroup analysis
fitSG <- try(summary(coxph(Surv(y.sim,event.sim)~treat.sim,data=df_sg))$conf.int,TRUE)
if(!inherits(fitSG,"try_error") & is.numeric(fitSG[1])) hr_sg <- c(fitSG[1])
if(inherits(fitSG,"try_error") | !is.numeric(fitSG[1])) hr_sg <- NA
loghr.po <- df_sg$loghr.po
POhr_sg <- exp(mean(loghr.po))
rm("fitSG")
}
 }  # FS without error done
rm("fs_temp")
} # AP estimable in first place
# Results
dfres <- data.table::data.table(n_itt,hr_itt,hr_ittX,n_train,hr_train,n_test,hr_test,any_found,sg_found,n_sg,hr_sg,POhr_sg,prev_sg)
return(dfres)
  }
  t.now<-proc.time()[3]
  t.min<-(t.now-t.start)/60
  if(details){
    cat("Minutes for simulations",c(n_sims,t.min),"\n")
    cat("Projection per 1000",c(t.min*(1000/n_sims)),"\n")
    cat("Propn subgroups found =",c(sum(!is.na(results$any_found))/n_sims),"\n")
  }
# Define hr_sg_null as missing if any_found = 0
results$hr_sg_null <- results$hr_sg
results$hr_sg_null[results$any_found==0] <- NA
return(results)  
}

summaryout_mrct <- function(pop_summary,mrct_sims, 
out_sgs = c("sg_found","sg_biomarker","sg_age"),
out_sgs2 = c("sg_biomarker","sg_age","sg_male","sg_ecog1","sg_histology","sg_CTregimen","sg_EU","sg_surgery"),
out_est = c("+ reldelta + APflag + sg_le85 + APflag2 + APflag3 + hr_sg_null"),
sg_type=1,
tab_caption = c("Identified subgroups and estimation summaries under biomarker effects: 1-factor combinations"), showtable=TRUE){

outwhat1 <- c("~ hr_itt + hr_ittX + hr_test + found + prev_sg + hr_sg + POhr_sg +")
if(sg_type==1) outwhat2 <- paste(out_sgs, collapse="+")
if(sg_type==2) outwhat2 <- paste(out_sgs2, collapse="+")
out_what <- as.formula(paste(outwhat1,outwhat2,out_est))

res <- list()
# True causal AHR
res$ahr_true = c(round(pop_summary$AHR,4))
# AHR Cox un-adjusted (only treatment)
res$ahr_unadj = c(round(pop_summary$ITT_unadj,4))
# Treatment plus randomization stratificaton sR
res$ahr_sR = c(round(pop_summary$ITT_sR,4))
# adjustment by w
res$ahr_sRw = c(round(pop_summary$ITT_sRw,4))
# Relative biases 
res$bias_unadj <-round(100*c(pop_summary$ITT_unadj-pop_summary$AHR)/pop_summary$AHR,1)
res$bias_sR <-round(100*c(pop_summary$ITT_sR-pop_summary$AHR)/pop_summary$AHR,1)
res$bias_sRw <-round(100*c(pop_summary$ITT_sRw-pop_summary$AHR)/pop_summary$AHR,1)
# The bias within W=1 population
res$ahr_w1 = c(round(pop_summary$W_1,4))
res$bias_w1 <-round(100*c(pop_summary$W_1-pop_summary$AHR)/pop_summary$AHR,1)

mrct_sims$APflag <- ifelse(mrct_sims$hr_test > 0.9,"AP > 0.9","AP <=0.9")
mrct_sims$sg_le85 <- ifelse(mrct_sims$hr_sg <=0.85,"AP(sg)<=0.85","AP(sg)>0.85")

mrct_sims$APflag2 <- ifelse(mrct_sims$hr_test > 0.9 & mrct_sims$hr_sg <= 0.85,"AP > 0.9 & AP(sg) <= 0.85","Not")
mrct_sims$APflag3 <- ifelse(mrct_sims$hr_test > 0.9 & mrct_sims$hr_sg <= 0.80,"AP > 0.9 & AP(sg) <= 0.80","Not")

mrct_sims$reldelta <- 100*c(mrct_sims$hr_test - mrct_sims$hr_sg)/mrct_sims$hr_test

mrct_sims$found <- as.factor(mrct_sims$any_found)

# Classify by specific factors identified
mrct_sims$sg_biomarker <- ifelse(grepl("biomarker",mrct_sims$sg_found),
"biomarker",ifelse(mrct_sims$sg_found!="none","other","none"))

mrct_sims$sg_age <- ifelse(grepl("age",mrct_sims$sg_found),
"age",ifelse(mrct_sims$sg_found!="none","other","none"))

mrct_sims$sg_ecog1 <- ifelse(grepl("ecog1",mrct_sims$sg_found),
"ecog1",ifelse(mrct_sims$sg_found!="none","other","none"))

mrct_sims$sg_histology <- ifelse(grepl("histology",mrct_sims$sg_found),
"histology",ifelse(mrct_sims$sg_found!="none","other","none"))

mrct_sims$sg_CTregimen <- ifelse(grepl("mAD_CTregimen1",mrct_sims$sg_found),
"mAD_CT",ifelse(mrct_sims$sg_found!="none","other","none"))

mrct_sims$sg_male <- ifelse(grepl("male",mrct_sims$sg_found),
"male",ifelse(mrct_sims$sg_found!="none","other","none"))

mrct_sims$sg_surgery <- ifelse(grepl("surgery",mrct_sims$sg_found),
"surgery",ifelse(mrct_sims$sg_found!="none","other","none"))

mrct_sims$sg_EU <- ifelse(grepl("EU_region",mrct_sims$sg_found),
"EU region",ifelse(mrct_sims$sg_found!="none","other","none"))

out_table <- table1(out_what, data=mrct_sims, caption=tab_caption)

if(showtable) print(out_table)

return(list(res=res,out_table=out_table))
}

SGplot_estimates <- function(df,label_training="Training",label_testing="Testing", label_itt="ITT (sR)", label_sg="Testing(subgroup)"){
df_itt <- data.table(est=df$hr_itt, analysis=label_itt)
df_training <- data.table(est=df$hr_train, analysis=label_training)
df_testing <- data.table(est=df$hr_test, analysis=label_testing)
df_sg <- data.table(est=df$hr_sg, analysis=label_sg)

hr_estimates <- rbind(df_itt, df_training, df_testing, df_sg)  
est_order <- c(label_itt,label_training,label_testing,label_sg)
hr_estimates <- within(hr_estimates, {analysis <- factor(analysis,levels=est_order)})
p <- ggplot(hr_estimates, aes(x=analysis, y=est, fill=analysis)) + 
  geom_violin(trim=FALSE)
plot_estimates <- p + geom_boxplot(width=0.1)
return(list(dfPlot_estimates=hr_estimates, plot_estimates=plot_estimates))
} 

```  


## Strong biomarker effect (prognostic effect $\beta_{w}=-\log(5)$)

Simulations under strong biomarker effect 

```{r sims_alt, echo=TRUE, message=FALSE, warning=TRUE, eval=FALSE}
# Run sims and store results
t.start<-proc.time()[3]
# DGM 
#log.hrs <- log(c(2.5,1.25,0.50))
# example 2
log.hrs <- log(c(3,1.25,0.50))

dgm <- get_dgm_stratified(df=df.case,zeta=10,knot=5,log.hrs=log.hrs,strata_tte=NULL,details=TRUE)

pop_summary <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=100000,wname="AP_region",
                                   bw=-log(5),strata_rand="stratum",checking=FALSE,details=FALSE,return_df=FALSE)

# True causal AHR
ahr_true = c(round(pop_summary$AHR,4))
# AHR Cox un-adjusted (only treatment)
ahr_unadj = c(round(pop_summary$ITT_unadj,4))
# Treatment plus randomization stratificaton sR
ahr_sR = c(round(pop_summary$ITT_sR,4))
# sR including adjustment by x
ahr_sRw = c(round(pop_summary$ITT_sRw,4))
# Relative biases 
bias_unadj <-round(100*c(pop_summary$ITT_unadj-pop_summary$AHR)/pop_summary$AHR,1)
bias_sR <-round(100*c(pop_summary$ITT_sR-pop_summary$AHR)/pop_summary$AHR,1)
bias_sRw <-round(100*c(pop_summary$ITT_sRw-pop_summary$AHR)/pop_summary$AHR,1)
# The bias within W=1 population
ahr_w1 = c(round(pop_summary$W_1,4))
bias_w1 <-round(100*c(pop_summary$W_1-pop_summary$AHR)/pop_summary$AHR,1)

cat("AHR true (causal) and within AP:",c(ahr_true,ahr_w1),"\n")

n_sims <- nsims

library(doFuture)
library(doRNG)
registerDoFuture()
registerDoRNG()

mrct_sims <- mrct_APregion_sims(dgm=dgm,n_sims=n_sims,wname="AP_region",
                                bw=-log(5),sg_focus="minSG",details=TRUE)

t.now<-proc.time()[3]
t.min<-(t.now-t.start)/60
cat("Minutes (doFuture) simulations",c(t.min),"\n")

cat("How often was a subgroup identified=",c(mean(mrct_sims$any_found)),"\n")

if(output) save(n_sims,pop_summary,dgm,mrct_sims,file="output/mrct_sims3_v0.Rdata")

```

```{r}
if(loadit) load(file="output/mrct_sims3_v0.Rdata")

# Look at large-sample spline in non-AP
dflarge <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=10000,wname="AP_region",bw=-log(5),
                               strata_rand="stratum",checking=FALSE,details=FALSE,return_df=TRUE)

temp <- cox_cs_fit(df=subset(dflarge,AP_region==0),tte.name="y.sim",event.name="event.sim",treat.name="treat.sim",
strata.name="strata.simO",z.name=c("z"), details=FALSE ,boots=0, xlab=c("Biomarker"),maxz = 25, ydel=0.5, cex_legend=0.5, show_plot=TRUE,truebeta.name="loghr.po",
main_title=c("Non-AP large sample ~ training estimation properties"))

true_pos <- round(100*c(mean(mrct_sims$any_found)),4)

temp <- summaryout_mrct(pop_summary=pop_summary, mrct_sims=mrct_sims,showtable=FALSE)

ahr_true <- c(temp$res$ahr_true)
ahr_unadj <- c(temp$res$ahr_unadj)
ahr_sR <- c(temp$res$ahr_sR)
ahr_sRw <- c(temp$res$ahr_sRw)
ahr_w1 <- c(temp$res$ahr_w1)

```

- The underlying true ITT causal hazard ratio is `r ahr_true` 
  - The large-sample (single draw) approximation to ITT un-adjusted is `r ahr_unadj`
  - The large-sample ~ to ITT stratified is `r ahr_sR`
  - The large-sample ~ to ITT stratified + x-adjustment is `r ahr_sRw`
  - The large-sample ~ to AP region analysis is `r ahr_w1`


The true-positive rate (identifying a subgroup under HTE (heterogeneous treatment effects)) 

  - Across `r format(n_sims,nsmall=0,big.mark=",")` simulations (a subgroup was identified) = `r true_pos`\% 
  
Summary of subgroups identified and estimation properties are:


```{r}
# Show the table
temp$out_table
```


```{r, fig.width=10, fig.height=8}
temp <- SGplot_estimates(df=mrct_sims, label_training="Non-AP, ITT", label_itt="Overall, ITT", label_testing="AP, ITT", label_sg="AP, identified subgroup")
plot(temp$plot_estimates)
```




## Null model where benefit is uniform (prognostic effect bw=log5)

Simulations under the **null** where treatment benefit is uniform;  Treatment does not depend on any baseline factors (hazard-ratio is 0.70)

```{r sims_null, echo=TRUE, message=FALSE, warning=TRUE, eval=FALSE}
# Increase simulations to 5,000 under "null"
t.start<-proc.time()[3]

# Null uniform benefit hr=0.7
log.hrs <- log(c(0.7,0.7,0.7))
dgm <- get_dgm_stratified(df=df.case,zeta=10,knot=5,log.hrs=log.hrs,strata_tte=NULL,details=TRUE)

pop_summary <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=100000,wname="AP_region",
                                   bw=-log(5),strata_rand="stratum",checking=FALSE,details=FALSE,return_df=FALSE)

# True causal AHR
ahr_true = c(round(pop_summary$AHR,4))
# AHR Cox un-adjusted (only treatment)
ahr_unadj = c(round(pop_summary$ITT_unadj,4))
# Treatment plus randomization stratificaton sR
ahr_sR = c(round(pop_summary$ITT_sR,4))
# sR including adjustment by x
ahr_sRw = c(round(pop_summary$ITT_sRw,4))
# Relative biases 
bias_unadj <-round(100*c(pop_summary$ITT_unadj-pop_summary$AHR)/pop_summary$AHR,1)
bias_sR <-round(100*c(pop_summary$ITT_sR-pop_summary$AHR)/pop_summary$AHR,1)
bias_sRw <-round(100*c(pop_summary$ITT_sRw-pop_summary$AHR)/pop_summary$AHR,1)
# The bias within X=1 population
ahr_w1 = c(round(pop_summary$W_1,4))
bias_w1 <-round(100*c(pop_summary$W_1-pop_summary$AHR)/pop_summary$AHR,1)

cat("AHR true (causal) and within AP:",c(ahr_true,ahr_w1),"\n")

n_sims <- nsims

library(doFuture)
library(doRNG)
registerDoFuture()
registerDoRNG()


mrct_sims <- mrct_APregion_sims(dgm=dgm,n_sims=n_sims,wname="AP_region",
                                bw=-log(5),sg_focus="minSG",details=TRUE)

t.now<-proc.time()[3]
t.min<-(t.now-t.start)/60
cat("Minutes (doFuture) simulations",c(t.min),"\n")


cat("How often was a subgroup identified=",c(mean(mrct_sims$any_found)),"\n")

if(output) save(n_sims,pop_summary,dgm,mrct_sims,file="output/mrct_sims03_v0.Rdata")

```


```{r}
if(loadit) load(file="output/mrct_sims03_v0.Rdata")

false_pos <- round(100*c(mean(mrct_sims$any_found)),4)

# Look at large-sample spline in non-AP
dflarge <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=10000,wname="AP_region",bw=-log(5),
                               strata_rand="stratum",checking=FALSE,details=FALSE,return_df=TRUE)

temp <- cox_cs_fit(df=subset(dflarge,AP_region==0),tte.name="y.sim",event.name="event.sim",treat.name="treat.sim",
strata.name="strata.simO",z.name=c("z"), details=FALSE ,boots=0, xlab=c("Biomarker"),maxz = 25, ydel=0.5, cex_legend=0.5, show_plot=TRUE,truebeta.name="loghr.po",
main_title=c("Non-AP large sample ~ training estimation properties"))

# Note, here too many individual subgroup cuts, set sg_type=2
temp <- summaryout_mrct(pop_summary=pop_summary, mrct_sims=mrct_sims,showtable=FALSE, 
                        sg_type=2, tab_caption=c("Identified subgroups and estimation summaries under uniform treatment effects: 1-factor combinations"))

ahr_true <- c(temp$res$ahr_true)
ahr_unadj <- c(temp$res$ahr_unadj)
ahr_sR <- c(temp$res$ahr_sR)
ahr_sRw <- c(temp$res$ahr_sRw)
ahr_w1 <- c(temp$res$ahr_w1)
```


- The underlying true ITT causal hazard ratio is `r ahr_true` 
  - The large-sample (single draw) approximation to ITT un-adjusted is `r ahr_unadj`
  - The large-sample ~ to ITT stratified is `r ahr_sR`
  - The large-sample ~ to ITT stratified + x-adjustment is `r ahr_sRw`
  - The large-sample ~ to AP region analysis is `r ahr_w1`


The false-positive rate (identifying a subgroup under uniform benefit) 

  - Across `r format(n_sims,nsmall=0,big.mark=",")` simulations (a subgroup was falsely identified) = `r false_pos`\% 
  
Summary of subgroups identified and estimation properties are:


```{r}
# Show the table
temp$out_table
```


```{r, fig.width=10, fig.height=8}
temp <- SGplot_estimates(df=mrct_sims, label_training="Non-AP, ITT", label_itt="Overall, ITT", label_testing="AP, ITT", label_sg="AP, identified subgroup")
plot(temp$plot_estimates)
```



## Null model where benefit is uniform (prognostic effect bw=-log5): maxk=2 

Simulations under the **null** where treatment benefit is uniform (maxk=2);  Treatment does not depend on any baseline factors (hazard-ratio is 0.7)

Note: Allowing for 2-factor combinations may require more stringent hazard-ratio thresholds;  Here we increase to 1.25 (1.0) as in paper


```{r sims_k2_null, echo=TRUE, message=FALSE, warning=TRUE, eval=FALSE}
t.start<-proc.time()[3]

# Null uniform benefit hr=0.7
log.hrs <- log(c(0.7,0.7,0.7))
dgm <- get_dgm_stratified(df=df.case,zeta=10,knot=5,log.hrs=log.hrs,strata_tte=NULL,details=TRUE)

pop_summary <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=100000,wname="AP_region",
                                   bw=-log(5),strata_rand="stratum",checking=FALSE,details=FALSE,return_df=FALSE)

# True causal AHR
ahr_true = c(round(pop_summary$AHR,4))
# AHR Cox un-adjusted (only treatment)
ahr_unadj = c(round(pop_summary$ITT_unadj,4))
# Treatment plus randomization stratificaton sR
ahr_sR = c(round(pop_summary$ITT_sR,4))
# sR including adjustment by x
ahr_sRw = c(round(pop_summary$ITT_sRw,4))
# Relative biases 
bias_unadj <-round(100*c(pop_summary$ITT_unadj-pop_summary$AHR)/pop_summary$AHR,1)
bias_sR <-round(100*c(pop_summary$ITT_sR-pop_summary$AHR)/pop_summary$AHR,1)
bias_sRw <-round(100*c(pop_summary$ITT_sRw-pop_summary$AHR)/pop_summary$AHR,1)
# The bias within X=1 population
ahr_w1 = c(round(pop_summary$W_1,4))
bias_w1 <-round(100*c(pop_summary$W_1-pop_summary$AHR)/pop_summary$AHR,1)

cat("AHR true (causal) and within AP:",c(ahr_true,ahr_w1),"\n")

n_sims <- nsims

library(doFuture)
library(doRNG)
registerDoFuture()
registerDoRNG()

mrct_sims <- mrct_APregion_sims(dgm=dgm,n_sims=n_sims,wname="AP_region",bw=-log(5),
sg_focus="hr", maxk=2, hr.threshold=1.25, hr.consistency=1.0, details=TRUE)

t.now<-proc.time()[3]
t.min<-(t.now-t.start)/60
cat("Minutes (doFuture) simulations",c(t.min),"\n")


cat("How often was a subgroup identified=",c(mean(mrct_sims$any_found)),"\n")

if(output) save(n_sims,pop_summary,dgm,mrct_sims,file="output/mrct_sims03_k2_v0.Rdata")

```


```{r}
if(loadit) load(file="output/mrct_sims03_k2_v0.Rdata")
false_pos <- round(100*c(mean(mrct_sims$any_found)),4)

cat("What do a few non-null look like","\n")
head(mrct_sims$sg_found[which(mrct_sims$any_found==1)])

# Look at large-sample spline in non-AP
# This is same as above so do not repeat
# dflarge <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=10000,xname="AP_region",bx=log(5),strata_rand="stratum",checking=FALSE,details=FALSE,return_df=TRUE)
# temp <- cox_cs_fit(df=subset(dflarge,AP_region==0),tte.name="y.sim",event.name="event.sim",treat.name="treat.sim",
# strata.name="strata.simO",z.name=c("z"), details=FALSE ,boots=0, xlab=c("Biomarker"),maxz = 25, ydel=0.5, cex_legend=0.5, show_plot=TRUE,truebeta.name="loghr.po",
# main_title=c("Non-AP large sample ~ training estimation properties"))

# Note, here too many individual subgroup cuts, set sg_type=2
temp <- summaryout_mrct(pop_summary=pop_summary, mrct_sims=mrct_sims,showtable=FALSE, 
                        sg_type=2, tab_caption=c("Identified subgroups and estimation summaries under uniform treatment effects: 2-factor combinations"))

ahr_true <- c(temp$res$ahr_true)
ahr_unadj <- c(temp$res$ahr_unadj)
ahr_sR <- c(temp$res$ahr_sR)
ahr_sRw <- c(temp$res$ahr_sRw)
ahr_w1 <- c(temp$res$ahr_w1)
```


- The underlying true ITT causal hazard ratio is `r ahr_true` 
  - The large-sample (single draw) approximation to ITT un-adjusted is `r ahr_unadj`
  - The large-sample ~ to ITT stratified is `r ahr_sR`
  - The large-sample ~ to ITT stratified + x-adjustment is `r ahr_sRw`
  - The large-sample ~ to AP region analysis is `r ahr_w1`


The false-positive rate (identifying a subgroup under uniform benefit) 

  - Across `r format(n_sims,nsmall=0,big.mark=",")` simulations (a subgroup was falsely identified) = `r false_pos`\% 
  
Summary of subgroups identified and estimation properties are:

```{r}
# Show the table
temp$out_table
```


```{r, fig.width=10, fig.height=8}
temp <- SGplot_estimates(df=mrct_sims, label_training="Non-AP, ITT", label_itt="Overall, ITT", label_testing="AP, ITT", label_sg="AP, identified subgroup")
plot(temp$plot_estimates)
```


## Strong biomarker effect (prognostic effect bw=-log5) maxk=2

Simulations under strong biomarker effect and stronger prognostic effect allowing for 2-factor combination (maxk=2)

```{r sims_k2_alt1, echo=TRUE, message=FALSE, warning=TRUE, eval=FALSE}
t.start<-proc.time()[3]

# DGM 
#log.hrs <- log(c(2,1.0,0.60))
log.hrs <- log(c(3,1.25,0.50))

dgm <- get_dgm_stratified(df=df.case,zeta=10,knot=5,log.hrs=log.hrs,strata_tte=NULL,details=TRUE)

pop_summary <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=100000,wname="AP_region",bw=-log(5),
                                   strata_rand="stratum",checking=FALSE,details=FALSE,return_df=FALSE)

# True causal AHR
ahr_true = c(round(pop_summary$AHR,4))
# AHR Cox un-adjusted (only treatment)
ahr_unadj = c(round(pop_summary$ITT_unadj,4))
# Treatment plus randomization stratificaton sR
ahr_sR = c(round(pop_summary$ITT_sR,4))
# sR including adjustment by x
ahr_sRw = c(round(pop_summary$ITT_sRw,4))
# Relative biases 
bias_unadj <-round(100*c(pop_summary$ITT_unadj-pop_summary$AHR)/pop_summary$AHR,1)
bias_sR <-round(100*c(pop_summary$ITT_sR-pop_summary$AHR)/pop_summary$AHR,1)
bias_sRw <-round(100*c(pop_summary$ITT_sRw-pop_summary$AHR)/pop_summary$AHR,1)
# The bias within X=1 population
ahr_w1 = c(round(pop_summary$W_1,4))
bias_w1 <-round(100*c(pop_summary$W_1-pop_summary$AHR)/pop_summary$AHR,1)

cat("AHR true (causal) and within AP:",c(ahr_true,ahr_w1),"\n")

n_sims <- nsims

library(doFuture)
library(doRNG)
registerDoFuture()
registerDoRNG()

mrct_sims <- mrct_APregion_sims(dgm=dgm,n_sims=n_sims,wname="AP_region",bw=-log(5),
sg_focus="hr",maxk=2, hr.threshold=1.25, hr.consistency=1.0, details=TRUE)

t.now<-proc.time()[3]
t.min<-(t.now-t.start)/60
cat("Minutes (doFuture) simulations",c(t.min),"\n")

cat("How often was a subgroup identified=",c(mean(mrct_sims$any_found)),"\n")

if(output) save(n_sims,pop_summary,dgm,mrct_sims,file="output/mrct_sims3_k2_v0.Rdata")

```


```{r}
if(loadit) load(file="output/mrct_sims3_k2_v0.Rdata")
true_pos <- round(100*c(mean(mrct_sims$any_found)),4)

cat("What do a few non-null look like","\n")
head(mrct_sims$sg_found[which(mrct_sims$any_found==1)])

# Note, here too many individual subgroup cuts, set sg_type=2
temp <- summaryout_mrct(pop_summary=pop_summary, mrct_sims=mrct_sims,showtable=FALSE, sg_type=2, 
                        tab_caption=c("Identified subgroups and estimation summaries under biomarker effects: 2-factor combinations"))

ahr_true <- c(temp$res$ahr_true)
ahr_unadj <- c(temp$res$ahr_unadj)
ahr_sR <- c(temp$res$ahr_sR)
ahr_sRw <- c(temp$res$ahr_sRw)
ahr_w1 <- c(temp$res$ahr_w1)
```


- The underlying true ITT causal hazard ratio is `r ahr_true` 
  - The large-sample (single draw) approximation to ITT un-adjusted is `r ahr_unadj`
  - The large-sample ~ to ITT stratified is `r ahr_sR`
  - The large-sample ~ to ITT stratified + x-adjustment is `r ahr_sRw`
  - The large-sample ~ to AP region analysis is `r ahr_w1`


The true-positive rate (identifying a subgroup under strong biomarker effects) 

  - Across `r format(n_sims,nsmall=0,big.mark=",")` simulations (a subgroup was identified) = `r true_pos`\% 
  
Summary of subgroups identified and estimation properties are:

```{r}
# Show the table
temp$out_table
```


```{r, fig.width=10, fig.height=8}
temp <- SGplot_estimates(df=mrct_sims, label_training="Non-AP, ITT", label_itt="Overall, ITT", label_testing="AP, ITT", label_sg="AP, identified subgroup")
plot(temp$plot_estimates)
```

## Strong biomarker effect (prognostic effect $\beta_{w}=0$)

Simulations under strong biomarker effect but null prognostic effect 

```{r sims_alt_Wnull, echo=TRUE, message=FALSE, warning=TRUE, eval=FALSE}
# Run sims and store results
t.start<-proc.time()[3]
# DGM 
#log.hrs <- log(c(2.5,1.25,0.50))
# example 2
log.hrs <- log(c(3,1.25,0.50))

dgm <- get_dgm_stratified(df=df.case,zeta=10,knot=5,log.hrs=log.hrs,strata_tte=NULL,details=TRUE)

pop_summary <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=100000,wname="AP_region",
                                   bw=0.0,strata_rand="stratum",checking=FALSE,details=FALSE,return_df=FALSE)

# True causal AHR
ahr_true = c(round(pop_summary$AHR,4))
# AHR Cox un-adjusted (only treatment)
ahr_unadj = c(round(pop_summary$ITT_unadj,4))
# Treatment plus randomization stratificaton sR
ahr_sR = c(round(pop_summary$ITT_sR,4))
# sR including adjustment by x
ahr_sRw = c(round(pop_summary$ITT_sRw,4))
# Relative biases 
bias_unadj <-round(100*c(pop_summary$ITT_unadj-pop_summary$AHR)/pop_summary$AHR,1)
bias_sR <-round(100*c(pop_summary$ITT_sR-pop_summary$AHR)/pop_summary$AHR,1)
bias_sRw <-round(100*c(pop_summary$ITT_sRw-pop_summary$AHR)/pop_summary$AHR,1)
# The bias within W=1 population
ahr_w1 = c(round(pop_summary$W_1,4))
bias_w1 <-round(100*c(pop_summary$W_1-pop_summary$AHR)/pop_summary$AHR,1)

cat("AHR true (causal) and within AP:",c(ahr_true,ahr_w1),"\n")

n_sims <- nsims

library(doFuture)
library(doRNG)
registerDoFuture()
registerDoRNG()

mrct_sims <- mrct_APregion_sims(dgm=dgm,n_sims=n_sims,wname="AP_region",
                                bw=0.0,sg_focus="minSG",details=TRUE)

t.now<-proc.time()[3]
t.min<-(t.now-t.start)/60
cat("Minutes (doFuture) simulations",c(t.min),"\n")

cat("How often was a subgroup identified=",c(mean(mrct_sims$any_found)),"\n")

if(output) save(n_sims,pop_summary,dgm,mrct_sims,file="output/mrct_sims3Wnull_v0.Rdata")

```

```{r}
if(loadit) load(file="output/mrct_sims3Wnull_v0.Rdata")

# Look at large-sample spline in non-AP
dflarge <- draw_sim_stratified(dgm=dgm,ss=1,Ndraw=10000,wname="AP_region",bw=0.0,
                               strata_rand="stratum",checking=FALSE,details=FALSE,return_df=TRUE)

temp <- cox_cs_fit(df=subset(dflarge,AP_region==0),tte.name="y.sim",event.name="event.sim",treat.name="treat.sim",
strata.name="strata.simO",z.name=c("z"), details=FALSE ,boots=0, xlab=c("Biomarker"),maxz = 25, ydel=0.5, cex_legend=0.5, show_plot=TRUE,truebeta.name="loghr.po",
main_title=c("Non-AP large sample ~ training estimation properties"))

true_pos <- round(100*c(mean(mrct_sims$any_found)),4)

temp <- summaryout_mrct(pop_summary=pop_summary, mrct_sims=mrct_sims,showtable=FALSE)

ahr_true <- c(temp$res$ahr_true)
ahr_unadj <- c(temp$res$ahr_unadj)
ahr_sR <- c(temp$res$ahr_sR)
ahr_sRw <- c(temp$res$ahr_sRw)
ahr_w1 <- c(temp$res$ahr_w1)

```

- The underlying true ITT causal hazard ratio is `r ahr_true` 
  - The large-sample (single draw) approximation to ITT un-adjusted is `r ahr_unadj`
  - The large-sample ~ to ITT stratified is `r ahr_sR`
  - The large-sample ~ to ITT stratified + x-adjustment is `r ahr_sRw`
  - The large-sample ~ to AP region analysis is `r ahr_w1`


The true-positive rate (identifying a subgroup under HTE (heterogeneous treatment effects)) 

  - Across `r format(n_sims,nsmall=0,big.mark=",")` simulations (a subgroup was identified) = `r true_pos`\% 
  
Summary of subgroups identified and estimation properties are:


```{r}
# Show the table
temp$out_table
```


```{r, fig.width=10, fig.height=8}
temp <- SGplot_estimates(df=mrct_sims, label_training="Non-AP, ITT", label_itt="Overall, ITT", label_testing="AP, ITT", label_sg="AP, identified subgroup")
plot(temp$plot_estimates)
```

# Appendix

## Brief Overview of Methods

- *forestSearch* (@Leon_2024): 
  - Subgroups (SGs) with HR estimates indicative of "sub-optimal benefit" (e.g., $hr \geq 1.0$ threshold), are considered candidates with a "splitting consistency criteria" for selection^["splitting consistency criteria" judges the robustness/realness].  
     - In essence, if a subgroup that appears to derive the least benefit is identified, then the complement may 
  potentially be considered to derive benefit with a higher degree of confidence relative to the ITT population
   - SGs are formed by single factors (eg, SG1 = males, SG2 = age<65) and two-way combinations (e.g., SG3 = males & age<65)
   - SGs with a minimum size of $60$ and with a minimum of number of 10 events in each treatment arm will be considered
   - Reversing the role of treatment allows for identifying subgroups with "enhanced benefit" (e.g., $hr \leq 0.65$ threshold)
   - Continuous factors are cut at either medians (q2), or quartiles (q1, q2, q3, say)
   - Cuts are also identified by generalized random forests (GRF) which is an identification approach itself --   
  
- *Generalized random forests* (@ATW_2019, @CZ_2023) which is based on restricted mean survival time (RMST) comparisons as implemented in the R package @GRFpackage_2022:  
  - An RMST benefit of (at least) $3$ months for control is required for selection of a subgroup, where among tree depths of 1 and 2 the subgroup with the largest RMST benefit ($\geq 3$ months in favor of control) is selected
  - Similar to forestSearch, we are targeting relatively large effects
  - SGs with sample sizes of at least $60$ participants and a maximum tree depth of 2 is required


## Example of bootstrap adjusted estimates and CIs 

Example of bootstrap de-biased hazard-ratio (point) and CI estimation.   Note that this is relevant for inference with regard to the training population -- That is, we are interested in valid estimation when based on the same population from which subgroups were identified.  For example, if the AP data will also be used for identifying the subgroups then we would want to account for the algorithm used in identification.    Here we illustrate with the non-AP data.


```{r FSbootMinSG, echo=TRUE, message=FALSE, warning=TRUE, eval=FALSE}

fs_minSG <- forestsearch(df.analysis=df_nonAP, confounders.name=confounders.name,
outcome.name="y.sim", treat.name="treat.sim", event.name="event.sim",
potentialOutcome.name="loghr.po",
hr.threshold = 0.90, hr.consistency = 0.90, pconsistency.threshold = 0.90,
showten_subgroups=FALSE, details=FALSE, sg_focus="minSG",
conf_force=c("age <= 65","biomarker < 1","biomarker < 2","biomarker < 3",
"biomarker < 4","biomarker < 5","biomarker < 6","biomarker < 7","biomarker < 8",
"biomarker < 9","biomarker < 10"), exclude_cuts=c("biomarker <="), maxk=1, plot.sg=FALSE)

library(doFuture)
library(doRNG)
registerDoFuture()
registerDoRNG()

t.start<-proc.time()[3]

NB <- nboots

# Bootstrap bias-correction 
fs_bc <- forestsearch_bootstrap_dofuture(fs.est=fs_minSG,nb_boots=NB,show_three=TRUE,details=TRUE)
t.now<-proc.time()[3]
t.min<-(t.now-t.start)/60
cat("Minutes (doFuture) bootstrap",c(t.min),"\n")


if(output) save(fs_minSG, fs_bc, NB, file="output/boot3_minSG_k1_v0.Rdata")

```

```{r}
if(loadit) load(file="output/boot3_minSG_k1_v0.Rdata")

gt(as.data.frame(fs_bc$FSsg_tab), auto_align=FALSE) 


```


```{r FSbootMaxSG, echo=TRUE, message=FALSE, warning=TRUE, eval=FALSE}
dfnew <- copy(df_nonAP)
dfnew$control.sim <- 1-dfnew$treat.sim

fs_maxSG <- forestsearch(df.analysis=dfnew, confounders.name=confounders.name,
outcome.name="y.sim", treat.name="control.sim", event.name="event.sim",
potentialOutcome.name="loghr.po",
est.scale="1/hr", hr.threshold = 1/0.6, hr.consistency = 1/0.7, pconsistency.threshold = 0.90,
showten_subgroups=FALSE, details=TRUE, sg_focus="maxSG",
conf_force=c("age <= 65","biomarker < 1","biomarker < 2","biomarker < 3",
"biomarker < 4","biomarker < 5","biomarker < 6","biomarker < 7","biomarker < 8",
"biomarker < 9","biomarker < 10"), exclude_cuts=c("biomarker <="), maxk=1, plot.sg=FALSE)

library(doFuture)
library(doRNG)
registerDoFuture()
registerDoRNG()

t.start<-proc.time()[3]

NB <- nboots

# Bootstrap bias-correction 
fs_bc <- forestsearch_bootstrap_dofuture(fs.est=fs_maxSG,nb_boots=NB,show_three=TRUE,details=TRUE)
t.now<-proc.time()[3]
t.min<-(t.now-t.start)/60
cat("Minutes (doFuture) bootstrap",c(t.min),"\n")

if(output) save(fs_maxSG, fs_bc, NB, file="output/boot3_maxSG_k1_v0.Rdata")
```



```{r}
if(loadit) load(file="output/boot3_maxSG_k1_v0.Rdata")

gt(as.data.frame(fs_bc$FSsg_tab), auto_align=FALSE) 

```

```{r}
tALL.now<-proc.time()[3]
t.min<-(tALL.now-tALL.start)/60
cat("Minutes for ALL Analyses",c(t.min),"\n")
```


