---
title: "ACTG175 Re-analysis Under the Realigned Identifier Defaults"
subtitle: "Comparison against the manuscript-era payload, and consequences for reported claims"
format:
  pdf:
    documentclass: scrartcl
    papersize: letter
    geometry: margin=1in
    number-sections: true
    toc: true
    toc-depth: 2
bibliography: []
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(knitr)

# Used below for payloads that record the MR draw count under either the
# current name (mr_draws) or the pre-rename one (gate_draws).
`%||%` <- function(a, b) if (is.null(a)) b else a
```

# Purpose and scope

This report compares the ACTG175 odds-ratio analysis as re-run under the
realigned identifier defaults against the manuscript-era payload, and states
which reported claims are affected.

Two changes separate the runs. The first is deliberate: `use_lasso` and
`use_grf` were set to `FALSE` in the forest search (FS) fit, so that the
candidate family is fixed rather than estimated from the outcome data --- the
condition multiplier resampling (MR) requires, and the configuration the GBSG
analysis already used. The second was discovered during the comparison
and is not a change to this project's code at all: R 4.6.0 corrected
`stats::dfbeta()` for generalized linear model objects, and the manuscript-era
payload was built under R 4.5.2.

Everything below is computed from the two payload objects. No numbers are
carried in from the surrounding conversation.

```{r paths}
# Adjust if rendering from elsewhere.  Defaults assume this document sits in
# dev/identifier-alignment/.
path_ref <- "sim_analyses/actg175_table_payload.rds"   # manuscript-era
path_new <- "rerun/actg175_table_payload.rds"          # realigned re-run

ref_obj <- readRDS(path_ref)
new_obj <- readRDS(path_new)

ref <- ref_obj$table
new <- new_obj$table

stopifnot(identical(dim(ref), dim(new)),
          identical(ref$method, new$method),
          identical(ref$region, new$region))

row_id <- paste(ref$method, ref$region)
```

# Provenance

```{r provenance}
prov <- data.frame(
  quantity = c("bootstrap replicates (B)", "MR draws", "n total",
               "FS label", "FS MR family size"),
  reference = c(ref_obj$meta$B,
                ref_obj$meta$mr_draws %||% ref_obj$meta$gate_draws,
                ref_obj$meta$n_total,
                ref_obj$labels$fs,
                ref_obj$meta$timings$fs$n_family),
  new = c(new_obj$meta$B,
          new_obj$meta$mr_draws %||% new_obj$meta$gate_draws,
          new_obj$meta$n_total,
          new_obj$labels$fs,
          new_obj$meta$timings$fs$n_family)
)
kable(prov, caption = "Run provenance.")
```

The MR family size is identical. This is the decisive fact about the
front-end change: the prognostic lasso and the generalized random forest (GRF)
importance screen dropped nothing on the full ACTG175 sample, so the enumerated
family was already complete. Removing them therefore could not alter the FS
selection, and did not.

# What is unchanged

```{r unchanged}
same <- data.frame(
  row               = row_id,
  n_ref             = ref$n,
  n_new             = new$n,
  naive_est         = ref$naive_est,
  naive_est_reldiff = abs(new$naive_est - ref$naive_est) / abs(ref$naive_est)
)
kable(same, digits = c(0, 0, 0, 4, 16),
      caption = "Subgroup sizes and naive point estimates.")

labels_identical <- identical(unlist(ref_obj$labels[c("fs", "dina", "grf")]),
                              unlist(new_obj$labels[c("fs", "dina", "grf")]))
cat("All three identifier labels identical:", labels_identical, "\n")
```

Identification is untouched: labels, subgroup sizes, and the naive point
estimates agree to floating-point precision. The FS subgroup did **not** change
under the front-end removal, contrary to the expectation recorded before the
run.

# The naive interval discrepancy

Both intervals are symmetric on the log scale, so the implied standard error is
recoverable from the reported bounds.

```{r implied-se}
implied_se <- function(lo, hi) (log(hi) - log(lo)) / (2 * qnorm(0.975))

se_tab <- data.frame(
  row       = row_id,
  n         = ref$n,
  se_ref    = implied_se(ref$naive_lo, ref$naive_hi),
  se_new    = implied_se(new$naive_lo, new$naive_hi)
)
se_tab$ratio <- se_tab$se_ref / se_tab$se_new
kable(se_tab, digits = c(0, 0, 5, 5, 4),
      caption = "Implied naive standard errors, and the ratio between runs.")
```

Two features matter.

**The new values are the quantity the manuscript defines.** Section 4 of the
main paper specifies the naive fixed-subgroup variance as
$\sigma^2_{D,\hat{H}} = \sum_{i \in \hat{H}} \mathrm{db}^2_{\hat{H},i}$, the
diagonal of the influence-function covariance. Computed independently from the
raw `speff2trial::ACTG175` data, $\sqrt{\sum_i \mathrm{db}^2}$ equals the new
run's implied standard error exactly in every region --- 0.51181 (FS $\hat{H}$),
0.42753 (DINA $\hat{H}$), 0.42789 (GRF $\hat{H}$), 0.13098 (FS $\hat{H}^c$).

**The inflation is not a constant factor.** The ratios span roughly 2.3% and
are uncorrelated with subgroup size, which excludes any explanation in which
$\sigma_D$ is scaled by a constant.

## Root cause: an upstream change in base R {#sec-rootcause}

The discrepancy does not originate in `forestsearch`. It originates in
`stats::dfbeta()`.

R 4.6.0, released 2026-04-24, records the following under significant
user-visible changes:

> `weighted.residuals()` now returns weighted working residuals for `"glm"`
> objects. It used to return deviance residuals, but these do not give the
> one-step approximated dfbeta values. [...] Several influence measures for
> `"glm"` objects have been updated to use Pearson residuals rather than
> deviance residuals (the latter are not unbiased).

The manuscript-era payload was built under R 4.5.2; the re-run under R 4.6.1.
Substituting the two residual definitions into the influence formula reproduces
both runs, to five decimal places in all four regions:

| region | $n$ | deviance (pre-4.6.0) | working (4.6.0+) | payload, June | payload, re-run |
|---|---|---|---|---|---|
| FS $\hat{H}$ | 72 | 0.58019 | 0.51181 | 0.58019 | 0.51181 |
| DINA $\hat{H}$ | 98 | 0.49336 | 0.42753 | 0.49336 | 0.42753 |
| GRF $\hat{H}$ | 95 | 0.49643 | 0.42789 | 0.49643 | 0.42789 |
| FS $\hat{H}^c$ | 1011 | 0.15003 | 0.13098 | 0.15003 | 0.13098 |

Three consequences follow.

The `forestsearch` implementation is correct and was never changed:
`R/consistency_resample.R`, where $\sigma_D = \sqrt{\sum_i \mathrm{db}_i^2}$ is
formed, and `.consistency_glm_pieces()` are both byte-identical between the
June commit and the current one.

The manuscript's *definition* requires no revision. What changed is the base-R
function that evaluates it, and the change was a correction: deviance residuals
do not yield the one-step dfbeta approximation the definition intends. The
re-run values are therefore the correct ones.

The ratio's dependence on subgroup composition is explained. The ratio of
deviance to weighted working residuals depends on the fitted probabilities
within each subgroup, so it varies across regions rather than scaling
uniformly. This is precisely why no constant-factor explanation fitted.

## Demonstration

The mechanism is reproducible without ACTG175. The chunk below fits a logistic
model to simulated data and forms $\sigma_D$ under both residual definitions.

```{r rootcause-demo}
set.seed(11)
n <- 200
x <- rbinom(n, 1, 0.5)
z <- rnorm(n)
y <- rbinom(n, 1, plogis(-1 + 0.8 * x + 0.5 * z))
fit <- glm(y ~ x + z, family = binomial())

w <- fit$weights
h <- lm.influence(fit, do.coef = FALSE)$hat
X <- model.matrix(fit)
A <- solve(t(X) %*% (w * X))

# One-step dfbeta from an arbitrary residual vector.
dfb <- function(res) {
  t(sapply(seq_len(n), function(i) A %*% X[i, ] * sqrt(w[i]) * res[i] / (1 - h[i])))
}

res_working  <- sqrt(w) * (fit$y - fit$fitted.values) /
                (fit$fitted.values * (1 - fit$fitted.values))
res_deviance <- residuals(fit, type = "deviance")

sigma_new <- sqrt(colSums(dfb(res_working)^2))[2]
sigma_old <- sqrt(colSums(dfb(res_deviance)^2))[2]
sigma_pkg <- sqrt(colSums(dfbeta(fit)^2))[2]

data.frame(
  quantity = c("working residuals (R >= 4.6.0)",
               "deviance residuals (R < 4.6.0)",
               "stats::dfbeta() under this R",
               "ratio old / new"),
  value = c(sigma_new, sigma_old, sigma_pkg, sigma_old / sigma_new),
  row.names = NULL
)
cat("R version running this document:", as.character(getRversion()), "\n")
```

Under R 4.3.3 the third row equals the second (0.35054); under R 4.6.1 it
equals the first. The ratio, 1.13 here, sits in the same range as the four
ACTG175 regions.

```{r ratio-spread}
cat(sprintf("ratio range: %.4f to %.4f (spread %.1f%%)\n",
            min(se_tab$ratio), max(se_tab$ratio),
            100 * (max(se_tab$ratio) / min(se_tab$ratio) - 1)))
cat(sprintf("correlation of ratio with log(n): %.3f\n",
            cor(se_tab$ratio, log(se_tab$n))))
```

# Propagation to the multiplier-resampling estimates

The per-candidate admission threshold is
$t_g = \max(c_{\mathrm{screen}},\, c_{\mathrm{cons}} + z_{0.975}\,\sigma_{D,g})$.
An inflated $\sigma_D$ therefore raises the screening threshold, changes which
candidates survive on each multiplier draw, and shifts the estimated selection
bias. The consequence is visible in every MR column.

```{r mr-shift}
pctchg <- function(a, b) 100 * (b - a) / a

est_tab <- data.frame(
  row        = row_id,
  naive      = ref$naive_est,
  fb_ref     = ref$fb_est,  fb_new  = new$fb_est,
  fb_pct     = pctchg(ref$fb_est, new$fb_est),
  mr_ref     = ref$mr_est,  mr_new  = new$mr_est,
  mr_pct     = pctchg(ref$mr_est, new$mr_est)
)
kable(est_tab, digits = c(0, 4, 4, 4, 2, 4, 4, 2),
      caption = "Point estimates on the odds-ratio scale.")
```

```{r selection-bias}
sb <- data.frame(
  identifier = c("FS", "DINA", "GRF"),
  ref = sapply(c("fs", "dina", "grf"),
               function(k) ref_obj$meta$timings[[k]]$selection_bias),
  new = sapply(c("fs", "dina", "grf"),
               function(k) new_obj$meta$timings[[k]]$selection_bias)
)
sb$change_pct <- pctchg(sb$ref, sb$new)
kable(sb, digits = c(0, 4, 4, 1), row.names = FALSE,
      caption = "Estimated selection bias on the log-odds scale.")
```

Less bias is subtracted, so the de-biased estimates move upward, toward the
naive value and toward greater residual harm.

## Interval widths

```{r widths}
logwidth <- function(lo, hi) log(hi) - log(lo)
w_tab <- data.frame(
  row         = row_id,
  naive_ref   = logwidth(ref$naive_lo, ref$naive_hi),
  naive_new   = logwidth(new$naive_lo, new$naive_hi),
  naive_pct   = pctchg(logwidth(ref$naive_lo, ref$naive_hi),
                       logwidth(new$naive_lo, new$naive_hi)),
  mr_ref      = logwidth(ref$mr_lo, ref$mr_hi),
  mr_new      = logwidth(new$mr_lo, new$mr_hi),
  mr_pct      = pctchg(logwidth(ref$mr_lo, ref$mr_hi),
                       logwidth(new$mr_lo, new$mr_hi))
)
kable(w_tab, digits = c(0, 3, 3, 1, 3, 3, 1),
      caption = "Interval widths on the log scale.")
```

## Relation to the null

```{r null-relation}
classify <- function(lo, hi) ifelse(lo > 1, "excludes, harm",
                            ifelse(hi < 1, "excludes, protective", "includes 1"))
null_tab <- data.frame(
  row        = row_id,
  mr_ref     = sprintf("%.3f [%.3f, %.3f]", ref$mr_est, ref$mr_lo, ref$mr_hi),
  rel_ref    = classify(ref$mr_lo, ref$mr_hi),
  mr_new     = sprintf("%.3f [%.3f, %.3f]", new$mr_est, new$mr_lo, new$mr_hi),
  rel_new    = classify(new$mr_lo, new$mr_hi),
  sign_flip  = ifelse((ref$mr_est - 1) * (new$mr_est - 1) < 0, "YES", "")
)
kable(null_tab, caption = "MR estimates and intervals relative to OR = 1.")
```

No interval changes its relationship to the null: all six include $\mathrm{OR}
= 1$ in both runs. One point estimate crosses it --- the DINA harm region moves
from 0.956, on the protective side, to 1.072, on the harm side. This is the only
sign change in the table.

# Agreement between the full bootstrap and multiplier resampling

The full bootstrap (FB) replays the entire pipeline on each resample and places
no alignment requirement on the identifier. MR linearizes the selection event
and requires both that selection rank on the inferential coefficient and that
the candidate family be fixed. Section 4.2 states that MR reproduces the
bootstrap to first order when those conditions hold; Section 5 describes the
conditional-versus-marginal gap that opens when the family regenerates.

After the front-end change, the ACTG175 FS rows satisfy both conditions. The
DINA and GRF rows cannot: their families are generated by fitting a model to the
same data, so they target the marginal quantity while MR conditions on the
realized family.

```{r fb-mr-agreement}
disagree <- function(fb, mr) abs(log(fb / mr))

agree_tab <- data.frame(
  row    = row_id,
  family = c("fixed", "fixed", rep("conditional-inherent", 4)),
  d_ref  = disagree(ref$fb_est, ref$mr_est),
  d_new  = disagree(new$fb_est, new$mr_est)
)
# Movements below 0.001 on the log scale are not meaningful at this precision.
agree_tab$direction <- with(agree_tab,
  ifelse(abs(d_new - d_ref) < 1e-3, "unchanged",
  ifelse(d_new < d_ref, "closer to FB", "further from FB")))
kable(agree_tab, digits = c(0, 0, 3, 3, 0),
      caption = "Absolute log-ratio between the FB and MR point estimates. Zero is exact agreement.")
```

Both fixed-family rows move toward the bootstrap; the two regenerating-family
harm rows move away from it. This is the direction the framework predicts, and
it was produced by correcting $\sigma_D$ rather than by any adjustment made to
obtain it.

Three qualifications belong with this observation. Six rows is a small base.
The complement rows move very little, so the signal rests on the FS harm row's
substantial move in the predicted direction against the two model-based harm
rows moving oppositely. And the ACTG175 FS rows became fixed-family cases only
*because* of the front-end change made in this same re-run, so the comparison
is not a clean before-and-after on a single configuration.

# Consequences for reported claims

Section 6 of the main paper makes four assertions about ACTG175. Three are
unaffected or strengthened; one requires revision.

```{r claims}
claims <- data.frame(
  claim = c(
    "FS and GRF identify closely overlapping regions",
    "Naive odds ratios correct downward under both routes",
    "FS and GRF remain above the harm threshold",
    "DINA's apparent harm largely diminishes under correction"),
  status = c(
    "Unaffected -- labels identical",
    "Holds -- both routes below naive for FS and GRF",
    "Strengthened -- MR moves further from the null",
    "Requires revision -- see below")
)
kable(claims, caption = "Section 6 claims under the re-run.")
```

The fourth claim remains literally true: the DINA harm region's odds ratio falls
from 1.936 to 1.072 under MR, a large reduction. Its character changes, however.
In the manuscript-era payload, FB and MR agreed in placing the region below the
null (0.862 and 0.956). They now disagree in sign: FB at 0.862, MR at 1.072.
Since the paper's subject is the correspondence between the two routes, a
divergence in the one row where it occurs is worth stating rather than leaving
implicit --- particularly as Section 6 already frames the DINA and GRF results
as empirical evaluations on regenerating families rather than as instances of
Theorem 2.

# Required revisions

```{r revisions}
rev_tab <- data.frame(
  location = c("Supplement, ACTG175 table",
               "Supplement, ACTG175 table",
               "Supplement, ACTG175 table",
               "Supplement, reproducibility statement",
               "Main text, Section 6",
               "Companion GLM paper"),
  item     = c("naive_lo, naive_hi (all six rows)",
               "mr_est, mr_lo, mr_hi (all six rows)",
               "fb_est, fb_lo, fb_hi (FS rows only)",
               "State R >= 4.6.0 for GLM outcomes",
               "The DINA sentence",
               "Binary/OR, RR, RD, IRR results only"),
  cause    = c("stats::dfbeta() corrected in R 4.6.0",
               "Propagated through the admission threshold t_g",
               "Prefilter active within bootstrap replicates",
               "sigma_D on GLM outcomes is R-version dependent",
               "FB and MR now disagree in sign for DINA",
               "Same upstream change; MD and HR unaffected")
)
kable(rev_tab, caption = "Locations requiring revision. No definition changes; no package change.")
```

## Scope of the upstream change

The affected quantity is $\sigma_D$ on outcomes fitted with `stats::glm()`.
Continuous outcomes dispatch to `stats::lm()`, where working, Pearson and
deviance residuals coincide and the weights are unity, so the change is a
no-op; survival outcomes use `coxph()` and never reach the affected code.

| effect measure | fit | affected | ratio, old / new |
|---|---|---|---|
| Odds ratio, binomial logit | `glm` | yes | 1.165 |
| Risk ratio, binomial log or Poisson | `glm` | yes | 1.164 |
| Risk difference, binomial identity | `glm` | yes | 1.137 |
| Incidence rate ratio, Poisson log | `glm` | yes | 1.054 |
| Mean difference, continuous | `lm` | no | 1.000 |
| Hazard ratio, survival | `coxph` | no | not applicable |

For the companion generalized linear model paper this means the binary
odds-ratio application requires regeneration and the continuous mean-difference
application does not.

The GBSG
survival analysis reproduced to a maximum absolute difference of
$1.8 \times 10^{-15}$ across all 54 estimate and bound cells, and the three
simulation studies reproduced exactly --- FS and DINA at zero difference on
every column, GRF to $1.3 \times 10^{-9}$, with identification bit-exact in all
three. The evidence supporting Theorem 2, the alignment condition, and the
FB-versus-MR comparison is therefore untouched.

Two operational notes. The ACTG175 payload records no R version in its
metadata, whereas the GBSG payload and the simulation bundles do; adding it
would have made this difference visible at comparison time rather than after a
diagnostic search. And the FS run reports a candidate-pool truncation ---
`max_subgroups_search` limited evaluation to 20 of 37 enumerated candidates ---
which did not alter the selection here, both runs returning the same label, but
which is worth recording alongside the result.

# Session information

```{r session-info}
sessionInfo()
```
