---
title: "Validating `dfbeta_glm()` against exact leave-one-out"
subtitle: "Candidate replacement for `stats::dfbeta()` on the GLM influence path"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    code-fold: show
execute:
  echo: true
  warning: false
  message: false
bibliography: []
---

## Why

`sigma_D = sqrt(sum(dfbeta^2))` is the robust subgroup standard error the
consistency engine and multiplier resampling both rest on. On the GLM path it
is built from `stats::dfbeta(fit)` at `R/consistency_resample.R:206`.

**R 4.6.0 changed what that returns.** `weighted.residuals()` on a `glm` used
to return *deviance* residuals, which -- per R's own NEWS -- "do not give the
one-step approximated dfbeta values"; influence measures now use Pearson /
working residuals. Results computed under R < 4.6.0 therefore carry a
different `sigma_D`, inflated by 12--16% in the case measured in
`NOTE_actg175_naive_interval_rootcause.md`.

Computing the influence directly makes `sigma_D` correct on **every** R
version and removes the version dependency rather than documenting it.

**This document validates the candidate. It does not install it.**
`consistency_resample.R` is unchanged.

### The Cox path is not affected — confirmed, not assumed

`R/consistency_resample.R:79` uses `stats::residuals(fit, type = "dfbeta")` on
a `coxph` object. That dispatches to `survival:::residuals.coxph`, an
independent implementation, not to `stats::dfbeta`. It needs no change. (This
matches the empirical result that GBSG, a survival analysis, reproduced its
2026-06 reference to 1.8e-15 while the binary ACTG175 analysis did not.)

```{r cox-dispatch-check}
suppressMessages(library(survival))
set.seed(1)
n <- 60
.d <- data.frame(t = rexp(n), e = rbinom(n, 1, 0.7), treat = rep(0:1, n / 2))
.cx <- survival::coxph(survival::Surv(t, e) ~ treat, data = .d)
cat("class(fit):            ", paste(class(.cx), collapse = ", "), "\n")
cat("residuals() dispatches to: ",
    paste(format(utils::getS3method("residuals", "coxph"))[1], collapse = ""), "\n")
cat("stats::dfbeta has a coxph method? ",
    !is.null(tryCatch(utils::getS3method("dfbeta", "coxph"), error = function(e) NULL)),
    "\n")
```

## The three computations

```{r computations}
# ---- CANDIDATE -------------------------------------------------------------
# One-step influence built directly.  fit$residuals is the WORKING residual for
# any GLM family, so this is general rather than binomial-specific.
#
# The `is.null(w)` line is the 2026-08-05 correction.  An UNWEIGHTED lm() has
# fit$weights == NULL, not a vector of ones, so the original body errored on
# every lm fit -- and effect_measure = "MD" fits with lm().  See "What the
# original grid missed" below.  This body must stay byte-identical to
# forestsearch:::.dfbeta_glm(); the end-to-end section asserts that.
dfbeta_glm <- function(fit) {
  X <- stats::model.matrix(fit)
  w <- fit$weights
  if (is.null(w)) w <- rep(1, nrow(X))     # lm() with no prior weights
  h <- stats::lm.influence(fit, do.coef = FALSE)$hat
  A <- solve(crossprod(X, w * X))
  r <- sqrt(w) * fit$residuals          # weighted working residual
  t(A %*% t(X * (sqrt(w) * r / (1 - h))))
}

# ---- OLD (pre-R-4.6.0) -----------------------------------------------------
# lm.influence() computes dfbeta_i = (X'WX)^-1 x_i sqrt(w_i) * wt.res_i /
# (1 - h_i).  Before R 4.6.0, weighted.residuals() on a glm returned DEVIANCE
# residuals, so that is what went in as wt.res.
#
# Given the same NULL-weights treatment, so the lm rows added below produce a
# comparable C column rather than an uninformative NA.  On an lm the deviance
# residual IS the response residual, so this coincides with the candidate
# exactly -- which is the point: the lm path is one where old and new agree,
# and agreement there is therefore no evidence the path was exercised.
dfbeta_old_deviance <- function(fit) {
  X <- stats::model.matrix(fit)
  w <- fit$weights
  if (is.null(w)) w <- rep(1, nrow(X))
  h <- stats::lm.influence(fit, do.coef = FALSE)$hat
  A <- solve(crossprod(X, w * X))
  rdev <- stats::residuals(fit, type = "deviance")
  t(A %*% t(X * (sqrt(w) * rdev / (1 - h))))
}

# ---- GROUND TRUTH: exact leave-one-out -------------------------------------
# beta_hat - beta_hat_(i), refitting without observation i.  dfbeta is a
# one-step approximation to exactly this, so LOO -- not old-versus-new --
# decides which computation is right.
dfbeta_loo <- function(fit, coefname, idx = NULL) {
  # Refit via `subset = -i` rather than by handing update() a model frame.
  # model.frame() renames prior weights and offsets to "(weights)"/"(offset)",
  # so update(data = model.frame(fit)) fails to find them and every refit
  # errors -- silently, since the loop catches errors and returns NA.  Using
  # `subset` keeps the original call and data intact.
  b0 <- stats::coef(fit)[coefname]
  n  <- stats::nobs(fit)
  ii <- if (is.null(idx)) seq_len(n) else idx
  out <- rep(NA_real_, length(ii))
  cl  <- stats::getCall(fit)
  env <- environment(stats::formula(fit))
  for (k in seq_along(ii)) {
    # Splice the VALUE into the call.  update(fit, subset = -ii[k]) would
    # store the unevaluated expression `-ii[k]` and evaluate it in the model's
    # environment, where `ii` and `k` do not exist -- every refit then errors
    # and the loop returns all-NA.  Assigning the number directly avoids the
    # lookup entirely.
    cl$subset <- -ii[k]
    fi <- tryCatch(suppressWarnings(eval(cl, envir = env)),
                   error = function(e) NULL)
    if (is.null(fi)) next
    if (!is.null(fi$converged) && isFALSE(fi$converged)) next
    ci <- stats::coef(fi)[coefname]
    if (length(ci) != 1L || is.na(ci)) next
    out[k] <- unname(b0 - ci)
  }
  out
}

# Guard: a LOO vector that is entirely NA means the refits failed, which would
# make every downstream comparison vacuously NA.  Fail loudly instead.
check_loo <- function(loo, label) {
  if (all(is.na(loo)))
    stop("leave-one-out refits all failed for: ", label, call. = FALSE)
  invisible(TRUE)
}
```

## Error metrics

Elementwise relative error is unstable where the true influence is near zero,
so two metrics are reported. Both are scaled by the vector's own magnitude,
which is the scale `sigma_D` actually depends on.

```{r metrics}
err_stats <- function(approx, truth) {
  ok <- is.finite(approx) & is.finite(truth)
  if (!any(ok)) return(c(max_rel = NA_real_, med_rel = NA_real_, n_used = 0))
  a <- approx[ok]; b <- truth[ok]
  scale_b <- max(abs(b))
  if (!is.finite(scale_b) || scale_b == 0) return(c(max_rel = NA_real_,
                                                    med_rel = NA_real_,
                                                    n_used = sum(ok)))
  max_rel <- max(abs(a - b)) / scale_b               # worst error, vector-scaled
  big <- abs(b) > 1e-8 * scale_b                     # skip the numerically-zero tail
  med_rel <- if (any(big)) stats::median(abs(a[big] - b[big]) / abs(b[big])) else NA_real_
  c(max_rel = max_rel, med_rel = med_rel, n_used = sum(ok))
}
```

## Scenario generator

```{r scenarios}
make_fit <- function(spec) {
  set.seed(spec$seed)
  n <- spec$n
  z <- if (spec$balance == "1:1") rbinom(n, 1, 0.5) else rbinom(n, 1, 0.1)
  x1 <- rnorm(n)
  x2 <- if (isTRUE(spec$collinear)) x1 + rnorm(n, sd = 0.01) else rnorm(n)
  d <- data.frame(treat = z, x1 = x1, x2 = x2)
  if (isTRUE(spec$high_leverage)) d$x1[1] <- 40      # single extreme covariate value
  pw <- if (isTRUE(spec$prior_weights)) sample(1:3, n, TRUE) else rep(1, n)
  d$pw <- pw
  fam <- spec$family; lnk <- spec$link
  rate <- spec$rate
  fit <- NULL
  if (fam == "binomial") {
    eta0 <- stats::qlogis(rate)
    lp <- eta0 + 0.8 * d$treat + 0.4 * d$x1
    if (isTRUE(spec$separation)) lp <- -8 + 16 * d$treat   # fitted probs near 0/1
    d$y <- rbinom(n, 1, stats::plogis(lp))
    if (isTRUE(spec$few_events)) d$y[d$treat == 1][-(1:2)] <- 0
    f <- y ~ treat + x1 + x2
    fit <- switch(lnk,
      logit = stats::glm(f, data = d, family = stats::binomial("logit"), weights = pw),
      log = tryCatch(stats::glm(f, data = d, family = stats::binomial("log"),
                                weights = pw,
                                start = c(log(mean(d$y) + .01), 0, 0, 0)),
                     error = function(e) NULL, warning = function(w) NULL),
      identity = tryCatch(stats::glm(f, data = d, family = stats::binomial("identity"),
                                     weights = pw,
                                     start = c(mean(d$y), 0, 0, 0)),
                          error = function(e) NULL, warning = function(w) NULL),
      probit = stats::glm(f, data = d, family = stats::binomial("probit"), weights = pw))
  } else if (fam == "poisson" || fam == "quasipoisson") {
    d$logT <- if (isTRUE(spec$offset)) log(runif(n, 0.5, 2)) else 0
    d$y <- rpois(n, exp(log(rate * 10) + 0.5 * d$treat + 0.3 * d$x1 + d$logT))
    f <- if (isTRUE(spec$offset)) y ~ treat + x1 + x2 + offset(logT)
         else y ~ treat + x1 + x2
    fit <- stats::glm(f, data = d, weights = pw,
                      family = if (fam == "poisson") stats::poisson("log")
                               else stats::quasipoisson("log"))
  } else if (fam == "gaussian") {
    d$y <- 2 + 0.5 * d$treat + 0.3 * d$x1 + rnorm(n)
    fit <- stats::glm(y ~ treat + x1 + x2, data = d,
                      family = stats::gaussian("identity"), weights = pw)
  } else if (fam == "lm") {
    # THE PATH THE ORIGINAL GRID HAD NO ROW FOR.
    #
    # This is not a re-spelling of the gaussian branch above.  glm(gaussian)
    # and lm() give identical coefficients and identical influences, but they
    # differ in the one respect that mattered: glm() always populates
    # fit$weights (ones, when unweighted) and an unweighted lm() leaves it
    # NULL.  Note the branch below passes NO `weights` argument in the
    # unweighted case -- writing `weights = rep(1, n)` would populate
    # fit$weights and quietly restore the gaussian-glm situation, defeating
    # the whole purpose of the row.
    d$y <- 2 + 0.5 * d$treat + 0.3 * d$x1 + rnorm(n)
    f <- y ~ treat + x1 + x2
    fit <- if (isTRUE(spec$prior_weights)) stats::lm(f, data = d, weights = pw)
           else                            stats::lm(f, data = d)
  }
  fit
}

run_scenario <- function(spec) {
  fit <- tryCatch(make_fit(spec), error = function(e) NULL)
  base <- data.frame(scenario = spec$label, stringsAsFactors = FALSE)
  if (is.null(fit) || (!is.null(fit$converged) && isFALSE(fit$converged))) {
    return(cbind(base, status = "fit failed", A_exact = NA, B_max = NA, B_med = NA,
                 C_max = NA, C_med = NA, max_h = NA))
  }
  cn <- "treat"
  new <- tryCatch(dfbeta_glm(fit)[, cn], error = function(e) NULL)
  old <- tryCatch(dfbeta_old_deviance(fit)[, cn], error = function(e) NULL)
  bi  <- tryCatch(stats::dfbeta(fit)[, cn], error = function(e) NULL)
  if (is.null(new)) {
    return(cbind(base, status = "dfbeta_glm failed", A_exact = NA, B_max = NA,
                 B_med = NA, C_max = NA, C_med = NA, max_h = NA))
  }
  if (!all(is.finite(new))) {
    return(cbind(base, status = "dfbeta_glm non-finite", A_exact = NA, B_max = NA,
                 B_med = NA, C_max = NA, C_med = NA, max_h = NA))
  }
  loo <- dfbeta_loo(fit, cn)
  if (all(is.na(loo))) {
    return(cbind(base, status = "LOO refits failed", A_exact = NA, B_max = NA,
                 B_med = NA, C_max = NA, C_med = NA, max_h = NA))
  }
  # Arm A: candidate vs this R's stats::dfbeta()
  A_exact <- if (is.null(bi)) NA else isTRUE(all.equal(unname(new), unname(bi),
                                                       tolerance = 1e-10))
  B <- err_stats(new, loo)
  C <- if (is.null(old)) c(max_rel = NA, med_rel = NA) else err_stats(old, loo)
  h <- stats::lm.influence(fit, do.coef = FALSE)$hat
  cbind(base, status = "ok", A_exact = A_exact,
        B_max = unname(B["max_rel"]), B_med = unname(B["med_rel"]),
        C_max = unname(C["max_rel"]), C_med = unname(C["med_rel"]),
        max_h = max(h))
}
```

## The grid

```{r grid}
specs <- list()
add <- function(...) specs[[length(specs) + 1L]] <<- list(...)

# Common cases: family/link x n x rate x balance
for (fl in list(c("binomial", "logit"), c("binomial", "log"),
                c("binomial", "identity"), c("binomial", "probit"),
                c("poisson", "log"), c("quasipoisson", "log"),
                c("gaussian", "identity"))) {
  for (n in c(40, 100, 500)) {
    for (rate in c(0.5, 0.1, 0.02)) {
      for (bal in c("1:1", "1:9")) {
        add(label = sprintf("%s/%s n=%d rate=%.2f %s", fl[1], fl[2], n, rate, bal),
            family = fl[1], link = fl[2], n = n, rate = rate, balance = bal,
            seed = 1000 + length(specs))
      }
    }
  }
}
# Poisson with offset
for (n in c(40, 100, 500)) {
  add(label = sprintf("poisson/log OFFSET n=%d", n), family = "poisson",
      link = "log", n = n, rate = 0.5, balance = "1:1", offset = TRUE,
      seed = 2000 + n)
}
# Edge cases
add(label = "EDGE quasi-separation (binomial logit)", family = "binomial",
    link = "logit", n = 100, rate = 0.5, balance = "1:1", separation = TRUE,
    seed = 3001)
add(label = "EDGE high leverage (h near 1)", family = "binomial", link = "logit",
    n = 60, rate = 0.5, balance = "1:1", high_leverage = TRUE, seed = 3002)
add(label = "EDGE few events in one arm", family = "binomial", link = "logit",
    n = 120, rate = 0.2, balance = "1:1", few_events = TRUE, seed = 3003)
add(label = "EDGE prior weights supplied", family = "binomial", link = "logit",
    n = 100, rate = 0.4, balance = "1:1", prior_weights = TRUE, seed = 3004)
add(label = "EDGE near-collinear covariates", family = "binomial", link = "logit",
    n = 100, rate = 0.4, balance = "1:1", collinear = TRUE, seed = 3005)
add(label = "EDGE high leverage (poisson)", family = "poisson", link = "log",
    n = 60, rate = 0.5, balance = "1:1", high_leverage = TRUE, seed = 3006)
add(label = "EDGE prior weights (gaussian)", family = "gaussian",
    link = "identity", n = 80, rate = 0.5, balance = "1:1",
    prior_weights = TRUE, seed = 3007)

# lm() rows -- ADDED 2026-08-05, DISTINCT FROM THE glm(gaussian) ROWS ABOVE.
# effect_measure = "MD" is fitted with lm(), and no row in the original grid
# reached lm().  Unweighted and prior-weighted are separate rows because only
# the unweighted one has fit$weights == NULL, which is the condition that
# broke.
for (n in c(40, 100, 500)) {
  add(label = sprintf("lm/identity UNWEIGHTED n=%d  [MD path]", n),
      family = "lm", link = "identity", n = n, rate = 0.5, balance = "1:1",
      seed = 4000 + n)
}
add(label = "lm/identity PRIOR WEIGHTS n=100", family = "lm",
    link = "identity", n = 100, rate = 0.5, balance = "1:1",
    prior_weights = TRUE, seed = 4500)
add(label = "lm/identity UNWEIGHTED n=40 1:9 imbalance", family = "lm",
    link = "identity", n = 40, rate = 0.5, balance = "1:9", seed = 4600)
add(label = "lm/identity UNWEIGHTED high leverage", family = "lm",
    link = "identity", n = 60, rate = 0.5, balance = "1:1",
    high_leverage = TRUE, seed = 4601)

cat("scenarios:", length(specs), "\n")
cat("  of which lm():", sum(vapply(specs, function(s) s$family == "lm", logical(1))), "\n")
```

```{r run-grid}
res <- do.call(rbind, lapply(specs, run_scenario))
res$B_max <- as.numeric(res$B_max); res$C_max <- as.numeric(res$C_max)
res$B_med <- as.numeric(res$B_med); res$C_med <- as.numeric(res$C_med)
res$B_better <- res$B_max < res$C_max
knitr::kable(res, digits = 6, row.names = FALSE)
```

## Hard control: gaussian identity must be EXACT

For a gaussian identity-link GLM the one-step influence is not an
approximation -- it is the closed form of the leave-one-out change. If arm B is
not exact to machine precision here, the algebra is wrong.

```{r gaussian-control}
g <- res[grepl("^gaussian", res$scenario) | grepl("gaussian", res$scenario), ]
cat("gaussian scenarios:", nrow(g), "\n")
cat("max over gaussian scenarios of arm-B max relative error:",
    format(max(g$B_max, na.rm = TRUE), scientific = TRUE), "\n")
stopifnot(max(g$B_max, na.rm = TRUE) < 1e-8)
cat("PASS: one-step is exact for gaussian identity.\n")
```

## Hard control: the `lm()` path must run at all {#sec-lm-control}

The gaussian control above is a control on the *algebra*. This one is a control
on the *path*: `effect_measure = "MD"` fits with `stats::lm()`, and an
unweighted `lm()` is the only fit in this package for which `fit$weights` is
`NULL`. The assertions below are, in order: the function returns without
error; it returns finite values; it agrees with `stats::dfbeta()` to 1e-8; and
it is exact against leave-one-out.

The `stats::dfbeta()` comparison is a legitimate assertion **here** and not for
`binomial`, because on an identity-link gaussian fit the working residual and
the deviance residual coincide -- so this comparison is version-independent,
whereas for `binomial` the two differ on R < 4.6.0 by design.

```{r lm-control}
lm_rows <- res[startsWith(res$scenario, "lm/"), ]
cat("lm() scenarios:", nrow(lm_rows), "\n")
cat("all completed  :", all(lm_rows$status == "ok"), "\n")
cat("max arm-B rel error over lm() rows:",
    format(max(lm_rows$B_max, na.rm = TRUE), scientific = TRUE), "\n")
cat("all match this R's stats::dfbeta():", all(lm_rows$A_exact), "\n")

stopifnot(nrow(lm_rows) > 0L)                       # the row must EXIST
stopifnot(all(lm_rows$status == "ok"))              # ... and must not error
stopifnot(all(lm_rows$A_exact))                     # ... and must match dfbeta()
stopifnot(max(lm_rows$B_max, na.rm = TRUE) < 1e-8)  # ... and be exact vs LOO

# Direct check of the precondition, independent of the grid machinery: an
# unweighted lm() really does have NULL weights, and the function copes.
.lm_probe <- local({
  set.seed(99); n <- 120
  d <- data.frame(treat = rbinom(n, 1, 0.5), x1 = rnorm(n))
  d$y <- 2 + 0.5 * d$treat + 0.3 * d$x1 + rnorm(n)
  stats::lm(y ~ treat + x1, data = d)
})
cat("\nunweighted lm(): is.null(fit$weights) =", is.null(.lm_probe$weights), "\n")
stopifnot(is.null(.lm_probe$weights))               # the precondition holds
stopifnot(all(is.finite(dfbeta_glm(.lm_probe))))
stopifnot(isTRUE(all.equal(unname(dfbeta_glm(.lm_probe)),
                           unname(stats::dfbeta(.lm_probe)),
                           tolerance = 1e-8)))
cat("PASS: the lm() path runs, and is exact.\n")
```

## What the original grid missed {#sec-miss}

This section exists because the grid *passed* while the function was broken for
an entire outcome type, and that is worth recording precisely rather than
quietly fixing.

**The defect.** The first version of `dfbeta_glm()` opened with
`w <- fit$weights` and used `w` unconditionally. For an unweighted `lm()`,
`fit$weights` is `NULL`, so `crossprod(X, w * X)` was non-conformable and the
function errored -- on every `lm` fit, every time.

**The blast radius, all of it silent.** `effect_measure = "MD"` is the only
measure fitted with `lm()`, and nothing downstream reports a failure:

| Level | Behaviour with the defect |
|---|---|
| `.dfbeta_glm()` | errors, "non-conformable arguments" |
| `.consistency_glm_pieces()` | catches it, returns `NULL` |
| `consistency_resample()` | returns all-`NA` rates |
| the consistency engine | falls back to literal splitting, ignoring `consistency_method = "resample"` without a message |
| `fs_mr_inference()` | fits no candidate; returns the short variant with `n_family = 0` and no `debiased` element |

So a continuous-outcome analysis run with `mr_inference = TRUE` produced no MR
correction at all, and said nothing about it.

**Why this grid did not catch it.** The grid had a `gaussian identity` row, and
that row was the *strictest* check in the whole document -- exact to 7.2e-13,
enforced by `stopifnot`. But it was built with
`glm(family = gaussian("identity"))`, where `fit$weights` is a vector of ones.
It exercised the gaussian **family** and never the `lm` **path**. The two are
numerically identical and structurally different, and only the structural
difference mattered.

That is the general lesson, and it is not about weights:

> A control that passes because the *quantity* is easy is not evidence that the
> *code path* is reachable. The gaussian row was chosen for being the case
> where the algebra is exact -- which is exactly why it could not detect a
> failure that had nothing to do with the algebra.

**What the end-to-end section did and did not buy.** The end-to-end check added
in the original commit was the right instinct and it did catch two wiring bugs.
But it ran one `forestsearch()` fit on ACTG175 -- a **binary** outcome, hence
`glm`. One end-to-end path is not end-to-end coverage; the continuous path had
no run of any kind, at any level.

**What now covers it.**

1. The `lm/` rows added to the grid above -- unweighted and prior-weighted as
   separate scenarios, since only the unweighted one has `NULL` weights.
2. @sec-lm-control, which asserts the rows exist, complete, and are exact, and
   separately probes the `is.null(fit$weights)` precondition directly, so the
   control fails loudly if a future R changes it.
3. `tests/testthat/test-dfbeta-lm-influence-path.R` in the package suite, which
   runs `forestsearch()` on a continuous outcome with `mr_inference = TRUE` and
   asserts the MR object is the full variant (`n_family > 0`, no `note`, a
   finite `debiased$est`). Verified against the pre-fix body: it fails all
   three.

## The ACTG175 regions from the actual analysis

```{r actg175}
suppressMessages(library(speff2trial))
ad <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
ad$treat <- ifelse(ad$arms == 1L, 1L, 0L)
ad <- ad[!is.na(ad$cd420), ]
ad$y <- 1L - as.integer(ad$cd420 > ad$cd40)
ad$id <- seq_len(nrow(ad))          # forestsearch() requires an id column

regions <- list(
  "FS H  (wtkg>86 & cd40>380)"        = (ad$wtkg > 86) & (ad$cd40 > 380),
  "FS Hc (complement)"                = !((ad$wtkg > 86) & (ad$cd40 > 380)),
  "DINA H (preanti>=849.4 & cd40>=338)" =
    (ad$preanti >= 849.40000000000009) & (ad$cd40 >= 338),
  "GRF H (wtkg>84.37 & cd40>368.2)"   =
    (ad$wtkg > 84.369600000000005) & (ad$cd40 > 368.20000000000005))

actg <- do.call(rbind, lapply(names(regions), function(nm) {
  s <- regions[[nm]]
  fit <- stats::glm(y ~ treat, data = ad[s, ], family = stats::binomial())
  new <- dfbeta_glm(fit)[, "treat"]
  old <- dfbeta_old_deviance(fit)[, "treat"]
  bi  <- stats::dfbeta(fit)[, "treat"]
  loo <- dfbeta_loo(fit, "treat")
  B <- err_stats(new, loo); C <- err_stats(old, loo)
  data.frame(region = nm, n = sum(s),
             A_exact = isTRUE(all.equal(unname(new), unname(bi), tolerance = 1e-10)),
             B_max = unname(B["max_rel"]), B_med = unname(B["med_rel"]),
             C_max = unname(C["max_rel"]), C_med = unname(C["med_rel"]),
             sigma_new = sqrt(sum(new^2)), sigma_old = sqrt(sum(old^2)),
             sigma_loo = sqrt(sum(loo^2)), stringsAsFactors = FALSE)
}))
knitr::kable(actg, digits = 6, row.names = FALSE)
```

## Summary

```{r summary}
ok <- res[res$status == "ok", ]

# Classify B against C.  A strict `<` counts an exact TIE as "not better",
# which misreports gaussian: there the deviance and working residuals coincide,
# so the two computations are identical by construction, not merely close.
rel <- function(a, b) abs(a - b) / pmax(abs(b), .Machine$double.eps)
ok$cmp <- ifelse(rel(ok$B_max, ok$C_max) < 1e-8, "tie",
                 ifelse(ok$B_max < ok$C_max, "better", "worse"))

cat("scenarios run   :", nrow(res), "\n")
cat("scenarios ok    :", nrow(ok), "\n")
cat("arm B vs arm C  : better", sum(ok$cmp == "better"),
    "| tie", sum(ok$cmp == "tie"), "| worse", sum(ok$cmp == "worse"), "\n\n")

cat("arm B (candidate vs LOO) max rel -- median",
    format(stats::median(ok$B_max, na.rm = TRUE), digits = 3, scientific = TRUE), "\n")
cat("arm C (old vs LOO)       max rel -- median",
    format(stats::median(ok$C_max, na.rm = TRUE), digits = 3, scientific = TRUE), "\n\n")

# Arm A: candidate must equal THIS R's stats::dfbeta() to machine precision.
badA <- ok[!(ok$A_exact %in% TRUE), ]
cat("ARM A (candidate vs stats::dfbeta on this R) not exact in",
    nrow(badA), "scenario(s):\n")
if (nrow(badA)) print(badA[, c("scenario", "max_h")], row.names = FALSE)

# Where the ONE-STEP ITSELF is poor: both computations far from LOO.  This is a
# limitation of sigma_D as a construct, independent of which residual is used.
cat("\nONE-STEP BREAKDOWN (arm B max rel > 0.10) -- a sigma_D limitation,\n")
cat("not a property of this fix.  Both computations are unusable here:\n")
poor <- ok[ok$B_max > 0.10, c("scenario", "B_max", "C_max", "max_h", "cmp")]
if (nrow(poor)) print(poor, row.names = FALSE) else cat("  none\n")

cat("\nAmong scenarios where the one-step is usable (arm B max rel <= 0.10):\n")
good <- ok[ok$B_max <= 0.10, ]
cat("  n =", nrow(good), "| B better", sum(good$cmp == "better"),
    "| tie", sum(good$cmp == "tie"), "| worse", sum(good$cmp == "worse"), "\n")
cat("  arm B max rel: median",
    format(stats::median(good$B_max), digits = 3, scientific = TRUE),
    " worst", format(max(good$B_max), digits = 3, scientific = TRUE), "\n")
cat("  arm C max rel: median",
    format(stats::median(good$C_max), digits = 3, scientific = TRUE),
    " worst", format(max(good$C_max), digits = 3, scientific = TRUE), "\n")

failed <- res[res$status != "ok", c("scenario", "status")]
cat("\nNON-OK SCENARIOS (", nrow(failed), "):\n", sep = "")
if (nrow(failed)) print(unique(failed[, c("status")]), row.names = FALSE)
if (nrow(failed)) print(table(failed$status))

saveRDS(list(grid = res, actg = actg), "dfbeta_glm_validation_results.rds")
```



## End-to-end: the installed path, not the leaf

The GRF threading work taught this the hard way -- 30,000 comparisons against
a leaf function passed while the argument never reached it. Verifying
`.dfbeta_glm()` in isolation says nothing about whether `forestsearch()`
actually uses it. This section runs a full fit and checks the number that
comes out the other end.

```{r end-to-end}
suppressMessages(library(forestsearch))

# The installed internal must BE the candidate validated above.
cat("installed .dfbeta_glm identical to the candidate here: ",
    isTRUE(all.equal(deparse(body(forestsearch:::.dfbeta_glm)),
                     deparse(body(dfbeta_glm)))), "\n")

fs <- forestsearch(
  df.analysis = ad, confounders.name = c("age","preanti","wtkg","karnof",
                                         "cd40","cd80","hemo","homo","drugs",
                                         "race","gender","symptom"),
  outcome.name = "y", event.name = "y", treat.name = "treat", id.name = "id",
  is.RCT = TRUE, seedit = 8316951L,
  outcome_type = "binary", effect_measure = "OR", adverse_outcome = TRUE,
  use_lasso = FALSE, use_grf = FALSE, use_dina = FALSE,
  cut_type = "default", cont.cutoff = 4L,
  conf.cont_jcuts = list(cd40 = 10, wtkg = 10),
  subgroup_method = "consistency", sg_focus = "effMaxSG",
  selection_rule = "neighborhood", effect_neighborhood = 0.10,
  mr_inference = TRUE, mr_inference_args = list(draws = 2000L),
  n.min = 60L, d0.min = 10L, d1.min = 10L, maxk = 2L,
  effect.threshold = 1.25, consistency.threshold = 1.0,
  pconsistency.threshold = 0.90, consistency_method = "resample",
  fs.splits = 500L, use_twostage = TRUE,
  details = FALSE, quiet = TRUE, plot.sg = FALSE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE))

lab <- paste(fs$sg.harm, collapse = " & ")
cat("selected subgroup :", lab, "\n")
cat("family status     :", fs$family_status, "\n")

# sigma_D that came OUT of the pipeline, via MR's naive Wald interval
g  <- fs$mr_inference
se_pipeline <- (log(g$naive$upper) - log(g$naive$lower)) / (2 * stats::qnorm(0.975))

# sigma_D computed independently on the selected members
mem <- fs$df.est$treat.recommend == 0
f_ind <- stats::glm(y ~ treat, data = ad[mem, ], family = stats::binomial())
se_candidate <- sqrt(sum(dfbeta_glm(f_ind)[, "treat"]^2))
se_old       <- sqrt(sum(dfbeta_old_deviance(f_ind)[, "treat"]^2))
se_loo       <- sqrt(sum(dfbeta_loo(f_ind, "treat")^2))

cat(sprintf("\nn in H                          : %d\n", sum(mem)))
cat(sprintf("sigma_D out of forestsearch()   : %.6f\n", se_pipeline))
cat(sprintf("sigma_D candidate, independent  : %.6f\n", se_candidate))
cat(sprintf("sigma_D exact leave-one-out     : %.6f\n", se_loo))
cat(sprintf("sigma_D old (deviance-based)    : %.6f\n", se_old))
cat(sprintf("\npipeline vs candidate rel diff  : %.3e\n",
            abs(se_pipeline - se_candidate) / se_candidate))
cat(sprintf("pipeline vs exact LOO rel diff  : %.3e\n",
            abs(se_pipeline - se_loo) / se_loo))
cat(sprintf("old      vs exact LOO rel diff  : %.3e\n",
            abs(se_old - se_loo) / se_loo))

stopifnot(abs(se_pipeline - se_candidate) / se_candidate < 1e-8)
cat("\nPASS: the value forestsearch() reports is the validated computation.\n")
```

## Findings

**The candidate is correct and the old computation is wrong**, on the regions
that matter and across the grid. Installed as `.dfbeta_glm()` in `c68fe70`;
corrected for the `lm()` path on 2026-08-05 (@sec-miss). Two limitations
below are known and accepted rather than fixed.

::: {.callout-note}
The counts in this section are computed from the grid at render time rather
than written in. They were previously hard-coded, and adding the `lm/` rows
made them stale on the spot -- a small instance of the same "one quantity,
two places" failure this whole document is about.
:::

### On the four ACTG175 regions -- the decisive result

`sigma_loo` is `sqrt(sum(dfbeta_LOO^2))` built from the exact refits.

| region | n | sigma old | sigma candidate | sigma exact-LOO | old vs LOO | candidate vs LOO |
|---|---|---|---|---|---|---|
| FS H | 72 | 0.580187 | 0.511807 | **0.515491** | +12.5% | -0.7% |
| FS Hc | 1011 | 0.150031 | 0.130977 | **0.131022** | +14.5% | -0.03% |
| DINA H | 98 | 0.493361 | 0.427534 | **0.428863** | +15.0% | -0.3% |
| GRF H | 95 | 0.496428 | 0.427890 | **0.428751** | +15.8% | -0.2% |

The old deviance-based `sigma_D` is 12--16% too large on every region; the
candidate is within 0.7%. This is the same 12--16% inflation measured in
`NOTE_actg175_naive_interval_rootcause.md`, now confirmed against ground truth
rather than against the other computation.

### Across the grid

Of `r nrow(res)` scenarios, `r nrow(ok)` produced a usable fit. Splitting on
whether the one-step approximation is usable at all (arm B max relative error
<= 0.10):

```{r grid-regimes, echo=FALSE}
.bad <- ok[ok$B_max > 0.10, ]
knitr::kable(data.frame(
  regime = c("one-step usable", "one-step broken down"),
  n      = c(nrow(good), nrow(.bad)),
  candidate_better = c(sum(good$cmp == "better"), sum(.bad$cmp == "better")),
  tie              = c(sum(good$cmp == "tie"),    sum(.bad$cmp == "tie")),
  candidate_worse  = c(sum(good$cmp == "worse"),  sum(.bad$cmp == "worse"))),
  row.names = FALSE)
```

In the usable regime the candidate is **never worse**
(`r sum(good$cmp == "worse")` scenarios), and the `r sum(good$cmp == "tie")`
ties are the identity-link gaussian and `lm` rows, where the deviance and
working residuals coincide so the two computations are identical by
construction. Median max relative error: candidate
**`r format(stats::median(good$B_max), digits = 3, scientific = TRUE)`**, old
**`r format(stats::median(good$C_max), digits = 3, scientific = TRUE)`**.

Note what the tie count means and does not mean. A tie is a scenario where the
two computations agree exactly -- which is informative about the algebra and
says nothing about whether the code path was exercised. That distinction is
what @sec-miss is about.

### Hard control passed

Gaussian identity is exact to machine precision (worst arm-B max relative
error 7.2e-13 across all gaussian scenarios, enforced by `stopifnot` above).
The algebra is right.

### The one-step itself breaks down at high leverage, independent of this fix

In `r sum(ok$B_max > 0.10)` of `r nrow(ok)` scenarios **both** computations are
far from the exact LOO. These
are sparse-event, small-n, imbalanced-arm fits with high leverage: worst cases
reach relative errors of 1e+12 (binomial/logit n=100 rate=0.02 1:9) and 9e+05
(binomial/probit n=40 rate=0.02 1:1, max hat = 0.99999992). Where a single
observation approaches leverage 1, `1/(1 - h_i)` explodes and no choice of
residual rescues it.

**This is a property of `sigma_D` as a construct, not of the fix.** It bounds
how much any influence-based standard error can be trusted in small, sparse
subgroups -- which is exactly the regime forestsearch searches in. Worth
knowing independently. The scenarios where the candidate is nominally worse are
all inside this regime, where both numbers are unusable and the comparison is
meaningless.

### Behaviour at exact leverage 1 -- and why no guard was added

`binomial/probit n=40 rate=0.50 1:9` has one subject in the treated arm and a
hat value of **exactly 1**. There:

- the candidate returns **`NaN`** (it is `0/0`, not a division by a small
  number -- so `NaN`, not `Inf`);
- `stats::dfbeta()` returns **exactly `0`** for that observation.

**Zero is wrong, and wrong in the direction that matters.** The observation
which alone determines the coefficient is recorded as having no influence on
it, so `sigma_D` comes out *smaller* than with two such observations:

```{r leverage-one, echo=TRUE}
set.seed(99); n <- 40
d0 <- data.frame(x1 = rnorm(n), x2 = rnorm(n)); d0$y <- rbinom(n, 1, 0.5)
d0$treat <- 0L
for (k in 1:3) {
  dd <- d0; dd$treat[seq_len(k)] <- 1L; dd$y[seq_len(k)] <- 1L
  f <- stats::glm(y ~ treat + x1 + x2, data = dd, family = stats::binomial())
  h <- stats::lm.influence(f, do.coef = FALSE)$hat
  b <- stats::dfbeta(f)[, "treat"]
  cat(sprintf("treated=%d  max hat=%.4f  sigma_D=%.5f  influence of treated pt=%.5g\n",
              k, max(h), sqrt(sum(b^2)), b[1]))
}
```

One observation yields a *smaller* `sigma_D` than two -- backwards. A subgroup
whose effect rests on a single observation is therefore scored with a falsely
narrow interval and a falsely low `t_g`: anticonservative exactly where
caution is warranted. **Dropping such a subgroup is a fix, not a regression.**

**No epsilon guard is applied**, deliberately. At `h = 1` the leave-one-out
estimate does not exist -- removing the observation leaves the coefficient
unidentifiable -- so `NaN` is the honest answer and a clamped value pretends
otherwise. A guard would also make `sigma_D` scale as `1/epsilon`, letting an
arbitrary constant set the magnitude of a reported quantity. And
`.consistency_glm_pieces()` already rejects non-finite `sigma_D`, so the
candidate is dropped -- the same treatment every other non-estimable candidate
receives.

### The difference is NOT systematically in one direction

The ACTG175 regions all had the old computation *larger* (+12.5% to +15.8%),
but that is a property of those fits, not a general bias. Across families and
event rates the sign reverses:

```{r direction, echo=TRUE}
set.seed(4); rows <- list()
for (fam in c("logit", "probit", "poisson", "gaussian")) {
  for (rate in c(0.5, 0.2, 0.05)) for (nn in c(60, 200)) {
    d <- data.frame(treat = rbinom(nn, 1, .5), x1 = rnorm(nn))
    f <- switch(fam,
      logit = { d$y <- rbinom(nn, 1, plogis(qlogis(rate) + .7 * d$treat))
                glm(y ~ treat + x1, d, family = binomial()) },
      probit = { d$y <- rbinom(nn, 1, plogis(qlogis(rate) + .7 * d$treat))
                 glm(y ~ treat + x1, d, family = binomial("probit")) },
      poisson = { d$y <- rpois(nn, exp(log(rate * 10) + .4 * d$treat))
                  glm(y ~ treat + x1, d, family = poisson()) },
      gaussian = { d$y <- 2 + .5 * d$treat + rnorm(nn)
                   glm(y ~ treat + x1, d, family = gaussian()) })
    sc <- sqrt(sum(dfbeta_glm(f)[, "treat"]^2))
    so <- sqrt(sum(dfbeta_old_deviance(f)[, "treat"]^2))
    if (is.finite(sc) && is.finite(so))
      rows[[length(rows) + 1L]] <- data.frame(fam = fam, rate = rate, n = nn,
                                              ratio_old_over_cand = so / sc)
  }
}
rr <- do.call(rbind, rows)
cat("old larger :", sum(rr$ratio > 1.0001),
    "| ~equal :", sum(abs(rr$ratio - 1) <= 1e-4),
    "| CANDIDATE larger :", sum(rr$ratio < 0.9999), "of", nrow(rr), "\n")
cat("ratio range:", format(range(rr$ratio), digits = 4), "\n")
```

At low event rates the candidate is substantially *larger*; near rate 0.5 the
old computation is larger. **The claim this validation supports is accuracy
against leave-one-out, not a direction of bias.** Regenerating a pre-fix result
can move `sigma_D` either way.

### Not defects

- **`r sum(res$status == "fit failed")` "fit failed" scenarios** are
  `binomial(log)` / `binomial(identity)` convergence failures in the generator,
  before any dfbeta is computed.
- **1 "dfbeta_glm failed"** (`gaussian/identity n=40 rate=0.50 1:9`) has
  **zero subjects in the treated arm**: the design is rank-deficient, the
  `treat` coefficient is `NA`, and `stats::dfbeta()` errors too.
  `.consistency_glm_pieces()` already rejects this upstream via its
  `any(is.na(cf))` guard, so it cannot reach either computation in production.
- **The single arm-A "inexactness"** (`binomial/logit n=40 rate=0.02 1:1`, one
  event in 40) is a relative difference of **2.06e-09** -- numerical noise from
  a near-degenerate fit. It passes `all.equal` at tolerance 1e-6 and fails only
  the strict 1e-10 used here. Arm A is otherwise exact in
  `r sum(ok$A_exact %in% TRUE)` of `r nrow(ok)`.

### Corrected 2026-08-05: the `lm()` path

The version first installed errored on every unweighted `lm()` fit, taking the
whole `effect_measure = "MD"` influence path down with it -- silently, at every
level. The correction is one added line; the `X` assignment already existed and
only moves up, because the new line needs `nrow(X)`:

```r
  X <- stats::model.matrix(fit)          # moved up from below `h`
  w <- fit$weights
+ if (is.null(w)) w <- rep(1, nrow(X))   # lm() with no prior weights
```

The grid did not catch it because its `gaussian identity` control was a
`glm(family = gaussian())`, where `fit$weights` is a vector of ones -- so the
control exercised the gaussian *family* and never the `lm` *path*. **The
gaussian control passing was therefore not the control it appeared to be.**
Full account in @sec-miss; coverage now in the `lm/` grid rows,
@sec-lm-control, and `tests/testthat/test-dfbeta-lm-influence-path.R`.

Nothing above this section changes: the accuracy findings, the ACTG175 result,
the leverage-1 behaviour and the no-epsilon-guard decision all concerned `glm`
fits and are unaffected. What changes is the scope over which they hold --
previously `glm` only, in fact rather than in intent.

### Outcome

**Installed**, without an epsilon guard, as `.dfbeta_glm()` in
`R/consistency_resample.R`, replacing `stats::dfbeta(fit)` at the GLM
influence site. It is exact where exactness is provable, materially closer to
ground truth everywhere the one-step is usable, never worse there, and removes
the R-version dependency rather than documenting it. The leverage-1 case
returns `NaN` and the caller drops the candidate, which is the correct
treatment for a subgroup whose effect is not identified without one
observation. As of the correction above it also handles `lm()`, weighted and
unweighted, which is the `MD` path.

The remaining limitation -- that the one-step itself breaks down at high
leverage regardless of residual definition -- is unchanged by this and bounds
how far `sigma_D` can be trusted in small sparse subgroups generally.

```{r session}
sessionInfo()
```
