---
title: "Validating `dfbeta_glm()` against exact leave-one-out"
subtitle: "Candidate replacement for `stats::dfbeta()` on the GLM influence path"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    code-fold: show
execute:
  echo: true
  warning: false
  message: false
bibliography: []
---

## Why

`sigma_D = sqrt(sum(dfbeta^2))` is the robust subgroup standard error the
consistency engine and multiplier resampling both rest on. On the GLM path it
is built from `stats::dfbeta(fit)` at `R/consistency_resample.R:206`.

**R 4.6.0 changed what that returns.** `weighted.residuals()` on a `glm` used
to return *deviance* residuals, which -- per R's own NEWS -- "do not give the
one-step approximated dfbeta values"; influence measures now use Pearson /
working residuals. Results computed under R < 4.6.0 therefore carry a
different `sigma_D`, inflated by 12--16% in the case measured in
`NOTE_actg175_naive_interval_rootcause.md`.

Computing the influence directly makes `sigma_D` correct on **every** R
version and removes the version dependency rather than documenting it.

**This document validates the candidate. It does not install it.**
`consistency_resample.R` is unchanged.

### The Cox path is not affected — confirmed, not assumed

`R/consistency_resample.R:79` uses `stats::residuals(fit, type = "dfbeta")` on
a `coxph` object. That dispatches to `survival:::residuals.coxph`, an
independent implementation, not to `stats::dfbeta`. It needs no change. (This
matches the empirical result that GBSG, a survival analysis, reproduced its
2026-06 reference to 1.8e-15 while the binary ACTG175 analysis did not.)

```{r cox-dispatch-check}
suppressMessages(library(survival))
set.seed(1)
n <- 60
.d <- data.frame(t = rexp(n), e = rbinom(n, 1, 0.7), treat = rep(0:1, n / 2))
.cx <- survival::coxph(survival::Surv(t, e) ~ treat, data = .d)
cat("class(fit):            ", paste(class(.cx), collapse = ", "), "\n")
cat("residuals() dispatches to: ",
    paste(format(utils::getS3method("residuals", "coxph"))[1], collapse = ""), "\n")
cat("stats::dfbeta has a coxph method? ",
    !is.null(tryCatch(utils::getS3method("dfbeta", "coxph"), error = function(e) NULL)),
    "\n")
```

## The three computations

```{r computations}
# ---- CANDIDATE -------------------------------------------------------------
# One-step influence built directly.  fit$residuals is the WORKING residual for
# any GLM family, so this is general rather than binomial-specific.
dfbeta_glm <- function(fit) {
  w <- fit$weights
  h <- stats::lm.influence(fit, do.coef = FALSE)$hat
  X <- stats::model.matrix(fit)
  A <- solve(crossprod(X, w * X))
  r <- sqrt(w) * fit$residuals          # weighted working residual
  t(A %*% t(X * (sqrt(w) * r / (1 - h))))
}

# ---- OLD (pre-R-4.6.0) -----------------------------------------------------
# lm.influence() computes dfbeta_i = (X'WX)^-1 x_i sqrt(w_i) * wt.res_i /
# (1 - h_i).  Before R 4.6.0, weighted.residuals() on a glm returned DEVIANCE
# residuals, so that is what went in as wt.res.
dfbeta_old_deviance <- function(fit) {
  w <- fit$weights
  h <- stats::lm.influence(fit, do.coef = FALSE)$hat
  X <- stats::model.matrix(fit)
  A <- solve(crossprod(X, w * X))
  rdev <- stats::residuals(fit, type = "deviance")
  t(A %*% t(X * (sqrt(w) * rdev / (1 - h))))
}

# ---- GROUND TRUTH: exact leave-one-out -------------------------------------
# beta_hat - beta_hat_(i), refitting without observation i.  dfbeta is a
# one-step approximation to exactly this, so LOO -- not old-versus-new --
# decides which computation is right.
dfbeta_loo <- function(fit, coefname, idx = NULL) {
  # Refit via `subset = -i` rather than by handing update() a model frame.
  # model.frame() renames prior weights and offsets to "(weights)"/"(offset)",
  # so update(data = model.frame(fit)) fails to find them and every refit
  # errors -- silently, since the loop catches errors and returns NA.  Using
  # `subset` keeps the original call and data intact.
  b0 <- stats::coef(fit)[coefname]
  n  <- stats::nobs(fit)
  ii <- if (is.null(idx)) seq_len(n) else idx
  out <- rep(NA_real_, length(ii))
  cl  <- stats::getCall(fit)
  env <- environment(stats::formula(fit))
  for (k in seq_along(ii)) {
    # Splice the VALUE into the call.  update(fit, subset = -ii[k]) would
    # store the unevaluated expression `-ii[k]` and evaluate it in the model's
    # environment, where `ii` and `k` do not exist -- every refit then errors
    # and the loop returns all-NA.  Assigning the number directly avoids the
    # lookup entirely.
    cl$subset <- -ii[k]
    fi <- tryCatch(suppressWarnings(eval(cl, envir = env)),
                   error = function(e) NULL)
    if (is.null(fi)) next
    if (!is.null(fi$converged) && isFALSE(fi$converged)) next
    ci <- stats::coef(fi)[coefname]
    if (length(ci) != 1L || is.na(ci)) next
    out[k] <- unname(b0 - ci)
  }
  out
}

# Guard: a LOO vector that is entirely NA means the refits failed, which would
# make every downstream comparison vacuously NA.  Fail loudly instead.
check_loo <- function(loo, label) {
  if (all(is.na(loo)))
    stop("leave-one-out refits all failed for: ", label, call. = FALSE)
  invisible(TRUE)
}
```

## Error metrics

Elementwise relative error is unstable where the true influence is near zero,
so two metrics are reported. Both are scaled by the vector's own magnitude,
which is the scale `sigma_D` actually depends on.

```{r metrics}
err_stats <- function(approx, truth) {
  ok <- is.finite(approx) & is.finite(truth)
  if (!any(ok)) return(c(max_rel = NA_real_, med_rel = NA_real_, n_used = 0))
  a <- approx[ok]; b <- truth[ok]
  scale_b <- max(abs(b))
  if (!is.finite(scale_b) || scale_b == 0) return(c(max_rel = NA_real_,
                                                    med_rel = NA_real_,
                                                    n_used = sum(ok)))
  max_rel <- max(abs(a - b)) / scale_b               # worst error, vector-scaled
  big <- abs(b) > 1e-8 * scale_b                     # skip the numerically-zero tail
  med_rel <- if (any(big)) stats::median(abs(a[big] - b[big]) / abs(b[big])) else NA_real_
  c(max_rel = max_rel, med_rel = med_rel, n_used = sum(ok))
}
```

## Scenario generator

```{r scenarios}
make_fit <- function(spec) {
  set.seed(spec$seed)
  n <- spec$n
  z <- if (spec$balance == "1:1") rbinom(n, 1, 0.5) else rbinom(n, 1, 0.1)
  x1 <- rnorm(n)
  x2 <- if (isTRUE(spec$collinear)) x1 + rnorm(n, sd = 0.01) else rnorm(n)
  d <- data.frame(treat = z, x1 = x1, x2 = x2)
  if (isTRUE(spec$high_leverage)) d$x1[1] <- 40      # single extreme covariate value
  pw <- if (isTRUE(spec$prior_weights)) sample(1:3, n, TRUE) else rep(1, n)
  d$pw <- pw
  fam <- spec$family; lnk <- spec$link
  rate <- spec$rate
  fit <- NULL
  if (fam == "binomial") {
    eta0 <- stats::qlogis(rate)
    lp <- eta0 + 0.8 * d$treat + 0.4 * d$x1
    if (isTRUE(spec$separation)) lp <- -8 + 16 * d$treat   # fitted probs near 0/1
    d$y <- rbinom(n, 1, stats::plogis(lp))
    if (isTRUE(spec$few_events)) d$y[d$treat == 1][-(1:2)] <- 0
    f <- y ~ treat + x1 + x2
    fit <- switch(lnk,
      logit = stats::glm(f, data = d, family = stats::binomial("logit"), weights = pw),
      log = tryCatch(stats::glm(f, data = d, family = stats::binomial("log"),
                                weights = pw,
                                start = c(log(mean(d$y) + .01), 0, 0, 0)),
                     error = function(e) NULL, warning = function(w) NULL),
      identity = tryCatch(stats::glm(f, data = d, family = stats::binomial("identity"),
                                     weights = pw,
                                     start = c(mean(d$y), 0, 0, 0)),
                          error = function(e) NULL, warning = function(w) NULL),
      probit = stats::glm(f, data = d, family = stats::binomial("probit"), weights = pw))
  } else if (fam == "poisson" || fam == "quasipoisson") {
    d$logT <- if (isTRUE(spec$offset)) log(runif(n, 0.5, 2)) else 0
    d$y <- rpois(n, exp(log(rate * 10) + 0.5 * d$treat + 0.3 * d$x1 + d$logT))
    f <- if (isTRUE(spec$offset)) y ~ treat + x1 + x2 + offset(logT)
         else y ~ treat + x1 + x2
    fit <- stats::glm(f, data = d, weights = pw,
                      family = if (fam == "poisson") stats::poisson("log")
                               else stats::quasipoisson("log"))
  } else if (fam == "gaussian") {
    d$y <- 2 + 0.5 * d$treat + 0.3 * d$x1 + rnorm(n)
    fit <- stats::glm(y ~ treat + x1 + x2, data = d,
                      family = stats::gaussian("identity"), weights = pw)
  }
  fit
}

run_scenario <- function(spec) {
  fit <- tryCatch(make_fit(spec), error = function(e) NULL)
  base <- data.frame(scenario = spec$label, stringsAsFactors = FALSE)
  if (is.null(fit) || (!is.null(fit$converged) && isFALSE(fit$converged))) {
    return(cbind(base, status = "fit failed", A_exact = NA, B_max = NA, B_med = NA,
                 C_max = NA, C_med = NA, max_h = NA))
  }
  cn <- "treat"
  new <- tryCatch(dfbeta_glm(fit)[, cn], error = function(e) NULL)
  old <- tryCatch(dfbeta_old_deviance(fit)[, cn], error = function(e) NULL)
  bi  <- tryCatch(stats::dfbeta(fit)[, cn], error = function(e) NULL)
  if (is.null(new)) {
    return(cbind(base, status = "dfbeta_glm failed", A_exact = NA, B_max = NA,
                 B_med = NA, C_max = NA, C_med = NA, max_h = NA))
  }
  if (!all(is.finite(new))) {
    return(cbind(base, status = "dfbeta_glm non-finite", A_exact = NA, B_max = NA,
                 B_med = NA, C_max = NA, C_med = NA, max_h = NA))
  }
  loo <- dfbeta_loo(fit, cn)
  if (all(is.na(loo))) {
    return(cbind(base, status = "LOO refits failed", A_exact = NA, B_max = NA,
                 B_med = NA, C_max = NA, C_med = NA, max_h = NA))
  }
  # Arm A: candidate vs this R's stats::dfbeta()
  A_exact <- if (is.null(bi)) NA else isTRUE(all.equal(unname(new), unname(bi),
                                                       tolerance = 1e-10))
  B <- err_stats(new, loo)
  C <- if (is.null(old)) c(max_rel = NA, med_rel = NA) else err_stats(old, loo)
  h <- stats::lm.influence(fit, do.coef = FALSE)$hat
  cbind(base, status = "ok", A_exact = A_exact,
        B_max = unname(B["max_rel"]), B_med = unname(B["med_rel"]),
        C_max = unname(C["max_rel"]), C_med = unname(C["med_rel"]),
        max_h = max(h))
}
```

## The grid

```{r grid}
specs <- list()
add <- function(...) specs[[length(specs) + 1L]] <<- list(...)

# Common cases: family/link x n x rate x balance
for (fl in list(c("binomial", "logit"), c("binomial", "log"),
                c("binomial", "identity"), c("binomial", "probit"),
                c("poisson", "log"), c("quasipoisson", "log"),
                c("gaussian", "identity"))) {
  for (n in c(40, 100, 500)) {
    for (rate in c(0.5, 0.1, 0.02)) {
      for (bal in c("1:1", "1:9")) {
        add(label = sprintf("%s/%s n=%d rate=%.2f %s", fl[1], fl[2], n, rate, bal),
            family = fl[1], link = fl[2], n = n, rate = rate, balance = bal,
            seed = 1000 + length(specs))
      }
    }
  }
}
# Poisson with offset
for (n in c(40, 100, 500)) {
  add(label = sprintf("poisson/log OFFSET n=%d", n), family = "poisson",
      link = "log", n = n, rate = 0.5, balance = "1:1", offset = TRUE,
      seed = 2000 + n)
}
# Edge cases
add(label = "EDGE quasi-separation (binomial logit)", family = "binomial",
    link = "logit", n = 100, rate = 0.5, balance = "1:1", separation = TRUE,
    seed = 3001)
add(label = "EDGE high leverage (h near 1)", family = "binomial", link = "logit",
    n = 60, rate = 0.5, balance = "1:1", high_leverage = TRUE, seed = 3002)
add(label = "EDGE few events in one arm", family = "binomial", link = "logit",
    n = 120, rate = 0.2, balance = "1:1", few_events = TRUE, seed = 3003)
add(label = "EDGE prior weights supplied", family = "binomial", link = "logit",
    n = 100, rate = 0.4, balance = "1:1", prior_weights = TRUE, seed = 3004)
add(label = "EDGE near-collinear covariates", family = "binomial", link = "logit",
    n = 100, rate = 0.4, balance = "1:1", collinear = TRUE, seed = 3005)
add(label = "EDGE high leverage (poisson)", family = "poisson", link = "log",
    n = 60, rate = 0.5, balance = "1:1", high_leverage = TRUE, seed = 3006)
add(label = "EDGE prior weights (gaussian)", family = "gaussian",
    link = "identity", n = 80, rate = 0.5, balance = "1:1",
    prior_weights = TRUE, seed = 3007)

cat("scenarios:", length(specs), "\n")
```

```{r run-grid}
res <- do.call(rbind, lapply(specs, run_scenario))
res$B_max <- as.numeric(res$B_max); res$C_max <- as.numeric(res$C_max)
res$B_med <- as.numeric(res$B_med); res$C_med <- as.numeric(res$C_med)
res$B_better <- res$B_max < res$C_max
knitr::kable(res, digits = 6, row.names = FALSE)
```

## Hard control: gaussian identity must be EXACT

For a gaussian identity-link GLM the one-step influence is not an
approximation -- it is the closed form of the leave-one-out change. If arm B is
not exact to machine precision here, the algebra is wrong.

```{r gaussian-control}
g <- res[grepl("^gaussian", res$scenario) | grepl("gaussian", res$scenario), ]
cat("gaussian scenarios:", nrow(g), "\n")
cat("max over gaussian scenarios of arm-B max relative error:",
    format(max(g$B_max, na.rm = TRUE), scientific = TRUE), "\n")
stopifnot(max(g$B_max, na.rm = TRUE) < 1e-8)
cat("PASS: one-step is exact for gaussian identity.\n")
```

## The ACTG175 regions from the actual analysis

```{r actg175}
suppressMessages(library(speff2trial))
ad <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
ad$treat <- ifelse(ad$arms == 1L, 1L, 0L)
ad <- ad[!is.na(ad$cd420), ]
ad$y <- 1L - as.integer(ad$cd420 > ad$cd40)
ad$id <- seq_len(nrow(ad))          # forestsearch() requires an id column

regions <- list(
  "FS H  (wtkg>86 & cd40>380)"        = (ad$wtkg > 86) & (ad$cd40 > 380),
  "FS Hc (complement)"                = !((ad$wtkg > 86) & (ad$cd40 > 380)),
  "DINA H (preanti>=849.4 & cd40>=338)" =
    (ad$preanti >= 849.40000000000009) & (ad$cd40 >= 338),
  "GRF H (wtkg>84.37 & cd40>368.2)"   =
    (ad$wtkg > 84.369600000000005) & (ad$cd40 > 368.20000000000005))

actg <- do.call(rbind, lapply(names(regions), function(nm) {
  s <- regions[[nm]]
  fit <- stats::glm(y ~ treat, data = ad[s, ], family = stats::binomial())
  new <- dfbeta_glm(fit)[, "treat"]
  old <- dfbeta_old_deviance(fit)[, "treat"]
  bi  <- stats::dfbeta(fit)[, "treat"]
  loo <- dfbeta_loo(fit, "treat")
  B <- err_stats(new, loo); C <- err_stats(old, loo)
  data.frame(region = nm, n = sum(s),
             A_exact = isTRUE(all.equal(unname(new), unname(bi), tolerance = 1e-10)),
             B_max = unname(B["max_rel"]), B_med = unname(B["med_rel"]),
             C_max = unname(C["max_rel"]), C_med = unname(C["med_rel"]),
             sigma_new = sqrt(sum(new^2)), sigma_old = sqrt(sum(old^2)),
             sigma_loo = sqrt(sum(loo^2)), stringsAsFactors = FALSE)
}))
knitr::kable(actg, digits = 6, row.names = FALSE)
```

## Summary

```{r summary}
ok <- res[res$status == "ok", ]

# Classify B against C.  A strict `<` counts an exact TIE as "not better",
# which misreports gaussian: there the deviance and working residuals coincide,
# so the two computations are identical by construction, not merely close.
rel <- function(a, b) abs(a - b) / pmax(abs(b), .Machine$double.eps)
ok$cmp <- ifelse(rel(ok$B_max, ok$C_max) < 1e-8, "tie",
                 ifelse(ok$B_max < ok$C_max, "better", "worse"))

cat("scenarios run   :", nrow(res), "\n")
cat("scenarios ok    :", nrow(ok), "\n")
cat("arm B vs arm C  : better", sum(ok$cmp == "better"),
    "| tie", sum(ok$cmp == "tie"), "| worse", sum(ok$cmp == "worse"), "\n\n")

cat("arm B (candidate vs LOO) max rel -- median",
    format(stats::median(ok$B_max, na.rm = TRUE), digits = 3, scientific = TRUE), "\n")
cat("arm C (old vs LOO)       max rel -- median",
    format(stats::median(ok$C_max, na.rm = TRUE), digits = 3, scientific = TRUE), "\n\n")

# Arm A: candidate must equal THIS R's stats::dfbeta() to machine precision.
badA <- ok[!(ok$A_exact %in% TRUE), ]
cat("ARM A (candidate vs stats::dfbeta on this R) not exact in",
    nrow(badA), "scenario(s):\n")
if (nrow(badA)) print(badA[, c("scenario", "max_h")], row.names = FALSE)

# Where the ONE-STEP ITSELF is poor: both computations far from LOO.  This is a
# limitation of sigma_D as a construct, independent of which residual is used.
cat("\nONE-STEP BREAKDOWN (arm B max rel > 0.10) -- a sigma_D limitation,\n")
cat("not a property of this fix.  Both computations are unusable here:\n")
poor <- ok[ok$B_max > 0.10, c("scenario", "B_max", "C_max", "max_h", "cmp")]
if (nrow(poor)) print(poor, row.names = FALSE) else cat("  none\n")

cat("\nAmong scenarios where the one-step is usable (arm B max rel <= 0.10):\n")
good <- ok[ok$B_max <= 0.10, ]
cat("  n =", nrow(good), "| B better", sum(good$cmp == "better"),
    "| tie", sum(good$cmp == "tie"), "| worse", sum(good$cmp == "worse"), "\n")
cat("  arm B max rel: median",
    format(stats::median(good$B_max), digits = 3, scientific = TRUE),
    " worst", format(max(good$B_max), digits = 3, scientific = TRUE), "\n")
cat("  arm C max rel: median",
    format(stats::median(good$C_max), digits = 3, scientific = TRUE),
    " worst", format(max(good$C_max), digits = 3, scientific = TRUE), "\n")

failed <- res[res$status != "ok", c("scenario", "status")]
cat("\nNON-OK SCENARIOS (", nrow(failed), "):\n", sep = "")
if (nrow(failed)) print(unique(failed[, c("status")]), row.names = FALSE)
if (nrow(failed)) print(table(failed$status))

saveRDS(list(grid = res, actg = actg), "dfbeta_glm_validation_results.rds")
```



## End-to-end: the installed path, not the leaf

The GRF threading work taught this the hard way -- 30,000 comparisons against
a leaf function passed while the argument never reached it. Verifying
`.dfbeta_glm()` in isolation says nothing about whether `forestsearch()`
actually uses it. This section runs a full fit and checks the number that
comes out the other end.

```{r end-to-end}
suppressMessages(library(forestsearch))

# The installed internal must BE the candidate validated above.
cat("installed .dfbeta_glm identical to the candidate here: ",
    isTRUE(all.equal(deparse(body(forestsearch:::.dfbeta_glm)),
                     deparse(body(dfbeta_glm)))), "\n")

fs <- forestsearch(
  df.analysis = ad, confounders.name = c("age","preanti","wtkg","karnof",
                                         "cd40","cd80","hemo","homo","drugs",
                                         "race","gender","symptom"),
  outcome.name = "y", event.name = "y", treat.name = "treat", id.name = "id",
  is.RCT = TRUE, seedit = 8316951L,
  outcome_type = "binary", effect_measure = "OR", adverse_outcome = TRUE,
  use_lasso = FALSE, use_grf = FALSE, use_dina = FALSE,
  cut_type = "default", cont.cutoff = 4L,
  conf.cont_jcuts = list(cd40 = 10, wtkg = 10),
  subgroup_method = "consistency", sg_focus = "effMaxSG",
  selection_rule = "neighborhood", effect_neighborhood = 0.10,
  mr_inference = TRUE, mr_inference_args = list(draws = 2000L),
  n.min = 60L, d0.min = 10L, d1.min = 10L, maxk = 2L,
  effect.threshold = 1.25, consistency.threshold = 1.0,
  pconsistency.threshold = 0.90, consistency_method = "resample",
  fs.splits = 500L, use_twostage = TRUE,
  details = FALSE, quiet = TRUE, plot.sg = FALSE,
  parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE))

lab <- paste(fs$sg.harm, collapse = " & ")
cat("selected subgroup :", lab, "\n")
cat("family status     :", fs$family_status, "\n")

# sigma_D that came OUT of the pipeline, via MR's naive Wald interval
g  <- fs$mr_inference
se_pipeline <- (log(g$naive$upper) - log(g$naive$lower)) / (2 * stats::qnorm(0.975))

# sigma_D computed independently on the selected members
mem <- fs$df.est$treat.recommend == 0
f_ind <- stats::glm(y ~ treat, data = ad[mem, ], family = stats::binomial())
se_candidate <- sqrt(sum(dfbeta_glm(f_ind)[, "treat"]^2))
se_old       <- sqrt(sum(dfbeta_old_deviance(f_ind)[, "treat"]^2))
se_loo       <- sqrt(sum(dfbeta_loo(f_ind, "treat")^2))

cat(sprintf("\nn in H                          : %d\n", sum(mem)))
cat(sprintf("sigma_D out of forestsearch()   : %.6f\n", se_pipeline))
cat(sprintf("sigma_D candidate, independent  : %.6f\n", se_candidate))
cat(sprintf("sigma_D exact leave-one-out     : %.6f\n", se_loo))
cat(sprintf("sigma_D old (deviance-based)    : %.6f\n", se_old))
cat(sprintf("\npipeline vs candidate rel diff  : %.3e\n",
            abs(se_pipeline - se_candidate) / se_candidate))
cat(sprintf("pipeline vs exact LOO rel diff  : %.3e\n",
            abs(se_pipeline - se_loo) / se_loo))
cat(sprintf("old      vs exact LOO rel diff  : %.3e\n",
            abs(se_old - se_loo) / se_loo))

stopifnot(abs(se_pipeline - se_candidate) / se_candidate < 1e-8)
cat("\nPASS: the value forestsearch() reports is the validated computation.\n")
```

## Findings

**The candidate is correct and the old computation is wrong**, on the regions
that matter and across the grid. It is not yet installed; two limitations
below need a decision first.

### On the four ACTG175 regions -- the decisive result

`sigma_loo` is `sqrt(sum(dfbeta_LOO^2))` built from the exact refits.

| region | n | sigma old | sigma candidate | sigma exact-LOO | old vs LOO | candidate vs LOO |
|---|---|---|---|---|---|---|
| FS H | 72 | 0.580187 | 0.511807 | **0.515491** | +12.5% | -0.7% |
| FS Hc | 1011 | 0.150031 | 0.130977 | **0.131022** | +14.5% | -0.03% |
| DINA H | 98 | 0.493361 | 0.427534 | **0.428863** | +15.0% | -0.3% |
| GRF H | 95 | 0.496428 | 0.427890 | **0.428751** | +15.8% | -0.2% |

The old deviance-based `sigma_D` is 12--16% too large on every region; the
candidate is within 0.7%. This is the same 12--16% inflation measured in
`NOTE_actg175_naive_interval_rootcause.md`, now confirmed against ground truth
rather than against the other computation.

### Across the grid

Of 136 scenarios, 112 produced a usable fit. Splitting on whether the one-step
approximation is usable at all (arm B max relative error <= 0.10):

| regime | n | candidate better | tie | candidate worse |
|---|---|---|---|---|
| one-step usable | 67 | 49 | 18 | **0** |
| one-step broken down | 45 | 38 | 0 | 7 |

In the usable regime the candidate is **never worse**, and the 18 ties are
gaussian, where the deviance and working residuals coincide so the two
computations are identical by construction. Median max relative error:
candidate **2.15e-02**, old **1.94e-01** -- about a 9x improvement.

### Hard control passed

Gaussian identity is exact to machine precision (worst arm-B max relative
error 7.2e-13 across all gaussian scenarios, enforced by `stopifnot` above).
The algebra is right.

### The one-step itself breaks down at high leverage, independent of this fix

In 45 of 112 scenarios **both** computations are far from the exact LOO. These
are sparse-event, small-n, imbalanced-arm fits with high leverage: worst cases
reach relative errors of 1e+12 (binomial/logit n=100 rate=0.02 1:9) and 9e+05
(binomial/probit n=40 rate=0.02 1:1, max hat = 0.99999992). Where a single
observation approaches leverage 1, `1/(1 - h_i)` explodes and no choice of
residual rescues it.

**This is a property of `sigma_D` as a construct, not of the fix.** It bounds
how much any influence-based standard error can be trusted in small, sparse
subgroups -- which is exactly the regime forestsearch searches in. Worth
knowing independently. The 7 scenarios where the candidate is nominally worse
are all inside this regime, where both numbers are unusable and the comparison
is meaningless.

### Behaviour at exact leverage 1 -- and why no guard was added

`binomial/probit n=40 rate=0.50 1:9` has one subject in the treated arm and a
hat value of **exactly 1**. There:

- the candidate returns **`NaN`** (it is `0/0`, not a division by a small
  number -- so `NaN`, not `Inf`);
- `stats::dfbeta()` returns **exactly `0`** for that observation.

**Zero is wrong, and wrong in the direction that matters.** The observation
which alone determines the coefficient is recorded as having no influence on
it, so `sigma_D` comes out *smaller* than with two such observations:

```{r leverage-one, echo=TRUE}
set.seed(99); n <- 40
d0 <- data.frame(x1 = rnorm(n), x2 = rnorm(n)); d0$y <- rbinom(n, 1, 0.5)
d0$treat <- 0L
for (k in 1:3) {
  dd <- d0; dd$treat[seq_len(k)] <- 1L; dd$y[seq_len(k)] <- 1L
  f <- stats::glm(y ~ treat + x1 + x2, data = dd, family = stats::binomial())
  h <- stats::lm.influence(f, do.coef = FALSE)$hat
  b <- stats::dfbeta(f)[, "treat"]
  cat(sprintf("treated=%d  max hat=%.4f  sigma_D=%.5f  influence of treated pt=%.5g\n",
              k, max(h), sqrt(sum(b^2)), b[1]))
}
```

One observation yields a *smaller* `sigma_D` than two -- backwards. A subgroup
whose effect rests on a single observation is therefore scored with a falsely
narrow interval and a falsely low `t_g`: anticonservative exactly where
caution is warranted. **Dropping such a subgroup is a fix, not a regression.**

**No epsilon guard is applied**, deliberately. At `h = 1` the leave-one-out
estimate does not exist -- removing the observation leaves the coefficient
unidentifiable -- so `NaN` is the honest answer and a clamped value pretends
otherwise. A guard would also make `sigma_D` scale as `1/epsilon`, letting an
arbitrary constant set the magnitude of a reported quantity. And
`.consistency_glm_pieces()` already rejects non-finite `sigma_D`, so the
candidate is dropped -- the same treatment every other non-estimable candidate
receives.

### The difference is NOT systematically in one direction

The ACTG175 regions all had the old computation *larger* (+12.5% to +15.8%),
but that is a property of those fits, not a general bias. Across families and
event rates the sign reverses:

```{r direction, echo=TRUE}
set.seed(4); rows <- list()
for (fam in c("logit", "probit", "poisson", "gaussian")) {
  for (rate in c(0.5, 0.2, 0.05)) for (nn in c(60, 200)) {
    d <- data.frame(treat = rbinom(nn, 1, .5), x1 = rnorm(nn))
    f <- switch(fam,
      logit = { d$y <- rbinom(nn, 1, plogis(qlogis(rate) + .7 * d$treat))
                glm(y ~ treat + x1, d, family = binomial()) },
      probit = { d$y <- rbinom(nn, 1, plogis(qlogis(rate) + .7 * d$treat))
                 glm(y ~ treat + x1, d, family = binomial("probit")) },
      poisson = { d$y <- rpois(nn, exp(log(rate * 10) + .4 * d$treat))
                  glm(y ~ treat + x1, d, family = poisson()) },
      gaussian = { d$y <- 2 + .5 * d$treat + rnorm(nn)
                   glm(y ~ treat + x1, d, family = gaussian()) })
    sc <- sqrt(sum(dfbeta_glm(f)[, "treat"]^2))
    so <- sqrt(sum(dfbeta_old_deviance(f)[, "treat"]^2))
    if (is.finite(sc) && is.finite(so))
      rows[[length(rows) + 1L]] <- data.frame(fam = fam, rate = rate, n = nn,
                                              ratio_old_over_cand = so / sc)
  }
}
rr <- do.call(rbind, rows)
cat("old larger :", sum(rr$ratio > 1.0001),
    "| ~equal :", sum(abs(rr$ratio - 1) <= 1e-4),
    "| CANDIDATE larger :", sum(rr$ratio < 0.9999), "of", nrow(rr), "\n")
cat("ratio range:", format(range(rr$ratio), digits = 4), "\n")
```

At low event rates the candidate is substantially *larger*; near rate 0.5 the
old computation is larger. **The claim this validation supports is accuracy
against leave-one-out, not a direction of bias.** Regenerating a pre-fix result
can move `sigma_D` either way.

### Not defects

- **22 "fit failed" scenarios** are `binomial(log)` / `binomial(identity)`
  convergence failures in the generator, before any dfbeta is computed.
- **1 "dfbeta_glm failed"** (`gaussian/identity n=40 rate=0.50 1:9`) has
  **zero subjects in the treated arm**: the design is rank-deficient, the
  `treat` coefficient is `NA`, and `stats::dfbeta()` errors too.
  `.consistency_glm_pieces()` already rejects this upstream via its
  `any(is.na(cf))` guard, so it cannot reach either computation in production.
- **The single arm-A "inexactness"** (`binomial/logit n=40 rate=0.02 1:1`, one
  event in 40) is a relative difference of **2.06e-09** -- numerical noise from
  a near-degenerate fit. It passes `all.equal` at tolerance 1e-6 and fails only
  the strict 1e-10 used here. Arm A is otherwise exact in 111 of 112.

### Outcome

**Installed**, without an epsilon guard, as `.dfbeta_glm()` in
`R/consistency_resample.R`, replacing `stats::dfbeta(fit)` at the GLM
influence site. It is exact where exactness is provable, materially closer to
ground truth everywhere the one-step is usable, never worse there, and removes
the R-version dependency rather than documenting it. The leverage-1 case
returns `NaN` and the caller drops the candidate, which is the correct
treatment for a subgroup whose effect is not identified without one
observation.

The remaining limitation -- that the one-step itself breaks down at high
leverage regardless of residual definition -- is unchanged by this and bounds
how far `sigma_D` can be trusted in small sparse subgroups generally.

```{r session}
sessionInfo()
```
