---
title: "ACTG175 subgroup analysis across selection foci: Naive, Full Bootstrap, and Multiplier Resampling"
subtitle: "Binary harm search (GLM odds-ratio outcome); `sg_focus` is a single knob"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
execute:
  warning: true
  message: true
bibliography: []
---

<!--
  analysis_actg175_binary_sgfocus.qmd

  The binary-outcome (odds-ratio) counterpart to
  analysis_gbsg_survival_sgfocus.qmd.  Data preparation, GLM plumbing and the
  Pareto block follow analysis_actg175_binary_multimethod.qmd; the document
  scope is the REDUCED single-track shape of the survival focus document --
  Naive / FB / MR, plus the Pareto selection diagnostic and CV/LOO.  The
  DINA-screening, standalone-DINA and GRF-demo comparators of the multimethod
  file are deliberately not carried across.

  SINGLE ADVERSE SCALE.  Everything operates on y_binary_neg (the indicator of
  NO CD4 improvement), so an OR above the screening threshold indicates harm --
  the direct analogue of an HR above the threshold on the survival path.

  WHAT ADAPTS TO THE FOCUS
    * Pareto frontier block runs only for the banded foci effMaxSG / effMinSG,
      where the (effect, N) frontier IS the selection geometry.
    * selection_rule is passed through only for the banded foci.  Elsewhere
      .validate_selection_rule() RAISES AN ERROR for a non-default value.
    * stop_threshold is supplied only for maxeffCons.
    * Under maxeff the five engine-forced overrides are pre-resolved in the
      call so args_call_all stays truthful.

  NO GUO & HE HERE.  fs_to_guohe() is a survival-path adapter (it takes
  time / event / treatment and runs a Cox argmax correction) AND requires
  sg_focus = "maxeff".  Neither condition is met by a binary OR analysis, so
  the GH track has no counterpart in this document.

  IMPORTANT: on the GLM path the Tier-2 de-biased gate runs ONLY under
  consistency_method = "resample".  Under the default "split" engine it is
  silently skipped for GLM outcomes.

  VENUE.  Compute-heavy; requires the forestsearch package installed (not
  load_all(): doFuture workers spawn separate R processes).
-->

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

timings <- list()
t_vignette_start <- proc.time()
```

# Setup

## Load required packages

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(speff2trial)
library(data.table)
library(ggplot2)
library(gt)
library(doFuture)
library(future)
library(kableExtra)

theme_set(theme_minimal(base_size = 12))
```

## Selection focus {#sec-focus}

`sg_focus` is the single knob that decides which surviving candidate is
returned. Two runs that differ only in `sg_focus` estimate different things, so
the resolved focus is recorded in the table below and every downstream section
keys off it.

```{r focus-config}
#| code-fold: false

# ---------------------------------------------------------------------------
# THE ONLY KNOB THAT CHANGES THE ESTIMAND.  Any legal value is accepted:
#
#   "maxcons" / "eff" / "hr"   consistency argmax, effect breaks ties
#   "maxeff"                   effect argmax over ALL candidates, ungated
#   "maxeffCons"               effect argmax among consistency qualifiers
#   "maxSG" / "minSG"          size extremes among qualifiers, no band
#   "effMaxSG" / "hrMaxSG"     largest N within the effect band
#   "effMinSG" / "hrMinSG"     smallest N within the effect band
#
# Prefer the eff*/maxcons spellings on the GLM path: "hazard ratio" is
# meaningless for a binary OR outcome, and .normalize_sg_focus() collapses the
# two vocabularies to the same canonical rule before anything keys on it.
# Note maxSG/minSG are NOT spellings of effMaxSG/effMinSG.
# ---------------------------------------------------------------------------
fs_sg_focus            <- "effMaxSG"

# selection_rule and effect_neighborhood shape the inclusion band.  They are
# consulted ONLY for effMaxSG / effMinSG; the profile below coerces the rule to
# "neighborhood" elsewhere, because forestsearch() treats a non-default rule
# under a non-band focus as an error rather than a no-op.  For OR (a ratio
# measure stored on the log scale) both band tests are applied after
# exponentiating, so the comparison is on the natural OR scale.
fs_selection_rule      <- "both"        # "neighborhood" | "pareto" | "both"
fs_effect_neighborhood <- 0.10

# Thresholds and search settings on the OR scale (inert under maxeff).
# Adverse scale: OR > threshold = harm.
fs_or_threshold         <- 1.25
fs_or_consistency       <- 1.0
fs_pconsistency         <- 0.90
fs_max_subgroups_search <- 20L
fs_minp                 <- 0.025
fs_use_twostage         <- TRUE
fs_stop_threshold       <- 0.95   # consumed by maxeffCons only
# The GLM Tier-2 gate requires "resample"; under "split" it is silently
# skipped for GLM outcomes.
fs_consistency_method   <- "resample"
```

```{r focus-profile}
#| code-fold: false

# Mirror of the package's .normalize_sg_focus() (forestsearch_helpers.R:916).
.fs_focus_canonical <- function(x) {
  switch(x,
         effMaxSG = "hrMaxSG",
         effMinSG = "hrMinSG",
         eff      = "hr",
         maxcons  = "hr",
         x)
}

#' Resolve a supplied sg_focus into the settings this document must honour
#'
#' Returns the canonical focus plus the derived flags every downstream chunk
#' keys on, and pre-resolves the arguments forestsearch() would otherwise
#' override internally.  Pre-resolving matters for provenance: the maxeff
#' override block (forestsearch_main.R:1309) does NOT re-sync args_call_all,
#' so a value typed into the call and then overridden is captured in its
#' pre-override form.  Supplying the engine-forced value directly removes the
#' discrepancy at source.
.fs_focus_profile <- function(sg_focus,
                              selection_rule,
                              effect_neighborhood,
                              effect_threshold,
                              consistency_threshold,
                              pconsistency,
                              max_subgroups_search,
                              minp,
                              use_twostage,
                              stop_threshold) {

  valid_user <- c("hr", "eff", "maxcons",
                  "hrMaxSG", "effMaxSG", "hrMinSG", "effMinSG",
                  "maxSG", "minSG", "maxeff", "maxeffCons")
  if (!is.character(sg_focus) || length(sg_focus) != 1L ||
      !sg_focus %in% valid_user) {
    stop("fs_sg_focus must be one of: ",
         paste(sQuote(valid_user), collapse = ", "),
         ".  Got: ", paste(deparse(sg_focus), collapse = " "),
         call. = FALSE)
  }

  canonical     <- .fs_focus_canonical(sg_focus)
  is_banded     <- canonical %in% c("hrMaxSG", "hrMinSG")
  is_maxeff     <- identical(canonical, "maxeff")
  is_maxeffcons <- identical(canonical, "maxeffCons")

  rule <- if (is_banded) selection_rule else "neighborhood"
  if (!is_banded && !identical(selection_rule, "neighborhood")) {
    message("selection_rule = '", selection_rule, "' is inert for sg_focus = '",
            sg_focus, "' and would be rejected by forestsearch(); ",
            "coerced to 'neighborhood'.")
  }

  stop_thr <- if (is_maxeffcons) stop_threshold else NULL

  if (is_maxeff) {
    pcons_out  <- 0
    msg_out    <- Inf
    minp_out   <- 0
    twostg_out <- FALSE
  } else {
    pcons_out  <- pconsistency
    msg_out    <- max_subgroups_search
    minp_out   <- minp
    twostg_out <- use_twostage
  }

  key <- switch(canonical,
    hr         = "(-Pcons, -eff, K)",
    maxeff     = "(-eff, K)",
    maxeffCons = "(-eff, K)",
    maxSG      = "(-N, -Pcons, K)",
    minSG      = "(N, -Pcons, K)",
    hrMaxSG    = "(-in_band, -N, -Pcons, -eff, K)",
    hrMinSG    = "(-in_band, N, -Pcons, -eff, K)")

  ranges_over <- switch(canonical,
    maxeff     = "ALL candidates, ungated",
    maxeffCons = "consistency-qualifying candidates only",
    "all qualifiers")

  list(
    supplied              = sg_focus,
    canonical             = canonical,
    key                   = key,
    ranges_over           = ranges_over,
    is_banded             = is_banded,
    is_maxeff             = is_maxeff,
    is_maxeffcons         = is_maxeffcons,
    run_pareto            = is_banded,
    thresholds_live       = !is_maxeff,
    selection_rule        = rule,
    effect_neighborhood   = effect_neighborhood,
    effect_threshold      = effect_threshold,
    consistency_threshold = consistency_threshold,
    pconsistency          = pcons_out,
    max_subgroups_search  = msg_out,
    minp                  = minp_out,
    use_twostage          = twostg_out,
    stop_threshold        = stop_thr
  )
}

focus <- .fs_focus_profile(
  sg_focus              = fs_sg_focus,
  selection_rule        = fs_selection_rule,
  effect_neighborhood   = fs_effect_neighborhood,
  effect_threshold      = fs_or_threshold,
  consistency_threshold = fs_or_consistency,
  pconsistency          = fs_pconsistency,
  max_subgroups_search  = fs_max_subgroups_search,
  minp                  = fs_minp,
  use_twostage          = fs_use_twostage,
  stop_threshold        = fs_stop_threshold
)

# Chunk-level gates (chunk options are evaluated in the global environment).
run_pareto <- focus$run_pareto
```

```{r focus-record}
#| echo: false
.fmt1 <- function(x) if (is.null(x)) "NULL" else
  paste(deparse(x, width.cutoff = 200L), collapse = " ")

kbl(data.frame(
  Item = c("sg_focus (supplied)", "sg_focus (canonical)", "Sort key",
           "Ranges over", "selection_rule", "effect_neighborhood",
           "Effect floor (effect.threshold, OR)",
           "Consistency screen (pconsistency)",
           "stop_threshold", "max_subgroups_search", "use_twostage",
           "Pareto frontier section"),
  Value = c(focus$supplied, focus$canonical, focus$key, focus$ranges_over,
            focus$selection_rule, .fmt1(focus$effect_neighborhood),
            if (focus$thresholds_live) .fmt1(focus$effect_threshold)
              else "disabled under maxeff",
            if (focus$thresholds_live) .fmt1(focus$pconsistency)
              else "0 (screen admits everything)",
            .fmt1(focus$stop_threshold),
            .fmt1(focus$max_subgroups_search),
            .fmt1(focus$use_twostage),
            if (focus$run_pareto) "RUN (banded focus)"
              else "skipped (band not the selection geometry)"),
  check.names = FALSE),
  booktabs = TRUE, align = "ll",
  caption = "Resolved selection configuration for this render.") |>
  kable_styling(full_width = FALSE)
```

```{r focus-narrative}
#| echo: false
#| results: asis
cat(switch(focus$canonical,
  hr = paste(
    "The selected subgroup is the **consistency argmax**: the candidate with the",
    "highest split-consistency probability, with the odds ratio breaking ties.",
    "Because enumeration proceeds by effect while the key is led by consistency,",
    "the scan order carries no information about the ranking -- every candidate",
    "must be scored.\n"),
  maxeff = paste(
    "The selected subgroup is the **effect maximiser** over the enumerated",
    "candidate family, with no auxiliary condition: the consistency screen is set",
    "to admit everything, the OR floor is disabled, early stopping and two-stage",
    "narrowing are off, and the candidate pool is not truncated. Consistency is",
    "computed for the winner and reported, but descriptively -- so",
    "`prop_SG_found = 100%` is structural rather than an empirical finding.\n"),
  maxeffCons = paste(
    "The selected subgroup is the **effect maximiser among consistency",
    "qualifiers**. The sort key is identical to `maxeff`; what differs is the",
    "population it ranges over, since only candidates clearing",
    "`pconsistency.threshold` reach the sort. This is the one focus whose key",
    "contains no `Pcons` term, so early stopping is sound and `stop_threshold` is",
    "live for this render.\n"),
  maxSG = paste(
    "The selected subgroup is the **largest qualifying candidate**, with",
    "consistency breaking ties. There is no inclusion band and no effect term",
    "until the final `K` tiebreak, so the rule can run out to a large but weak",
    "subgroup.\n"),
  minSG = paste(
    "The selected subgroup is the **smallest qualifying candidate**, with",
    "consistency breaking ties. There is no inclusion band and no effect term",
    "until the final `K` tiebreak.\n"),
  hrMaxSG = paste(
    "The selected subgroup is the **largest candidate inside the effect band**:",
    "in-band candidates first, then size, then consistency, then effect. The band",
    "is what stops a size-seeking rule from running out to the weak-but-large",
    "candidate `maxSG` would take. Because the band is the selection geometry",
    "rather than a post-hoc diagnostic, the (OR, N) Pareto frontier is reported",
    "below as a **selection** diagnostic.\n"),
  hrMinSG = paste(
    "The selected subgroup is the **smallest candidate inside the effect band**:",
    "in-band candidates first, then size ascending, then consistency, then effect.",
    "Because this focus deliberately prefers small subgroups, the selected",
    "subgroup may be N-dominated by a larger neighbour and therefore may not",
    "itself appear on the frontier -- that is expected, not a defect.\n")))
```

## Parallel processing configuration

```{r parallel-setup}
n_cores_total <- parallel::detectCores()
n_cores <- max(1L, floor(0.80 * parallel::detectCores(logical = FALSE)))
cat("Using", n_cores, "of", n_cores_total, "total cores for parallel processing\n")

NB       <- 1000L   # full-bootstrap resamples (production: 500-2000)
run_mr   <- TRUE    # Tier-2 de-biased gate (requires consistency_method = "resample")
mr_draws <- 2000L   # multiplier draws

run_cv  <- FALSE    # repeated 10-fold cross-validation
run_loo <- FALSE    # leave-one-out / N-fold OOB
Ksims   <- 50L      # repeated 10-fold partitions (>= 50 for production)

# Pre-seed so the timing table is well-formed whichever sections run.
timings$forestsearch <- NA_real_
timings$frontier     <- NA_real_
timings$bootstrap    <- NA_real_
timings$mr           <- NA_real_
timings$kfold        <- NA_real_
timings$oob          <- NA_real_

# ---------------------------------------------------------------------------
# Numeric precision (two-stage control).
#   construction_digits : upstream sprintf precision in the bootstrap / CV
#     table builders.  Acts as a CEILING for downstream display.
#   display_digits      : precision rendered in user-facing tables.
# ---------------------------------------------------------------------------
construction_digits <- 2L
display_digits      <- 2L

# Display-only tidy of a subgroup label for printing: rounds post-operator
# thresholds via the package's tidy_cut_display().  The membership-exact
# sg.harm is never altered -- this only affects what is printed.
fmt_sg <- function(x, d = display_digits) {
  if (is.null(x) || length(x) == 0L) return(x)
  paste(forestsearch:::tidy_cut_display(x, digits = as.integer(d)),
        collapse = " & ")
}

# ---------------------------------------------------------------------------
# Pareto-frontier diagnostic parameters.  Consulted only when the focus is
# banded.  compute_frontier_cis() is the expensive call (cost scales with
# frontier size x n_splits); set compute_cis <- FALSE to render the figure,
# table and narrative without interval bars.  For OR the effect axis is
# auto-detected as a ratio scale and rendered on log scale.
# ---------------------------------------------------------------------------
frontier_params <- list(
  compute_cis = TRUE,
  n_splits    = 1000L,
  seed        = 1L,
  xlim        = c(0.35, 10),
  xlim_trim   = NULL,   # NULL -> auto (TRUE for ratio measures such as OR)
  digits      = 2L
)
```

## Confounders and variable roles

```{r analysis-roles}
#| code-fold: false

# --- Reproducibility -------------------------------------------------------
analysis_seed <- 8316951L
set.seed(analysis_seed)

# --- Confounders -----------------------------------------------------------
cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
bin_vars  <- c("hemo", "homo", "drugs", "race", "gender", "symptom")
confounders.name <- c(cont_vars, bin_vars)
id.name   <- "id"

# --- Variables (single adverse scale) --------------------------------------
treat.name   <- "treat"
outcome.name <- "y_binary_neg"   # 1 = no improvement (adverse)

# --- ForestSearch GLM parameters (binary, OR scale) ------------------------
fs_outcome_type   <- "binary"
fs_effect_measure <- "OR"
```

# Data: ACTG175 trial

## Treatment assignment

```{r data-treat}
# ACTG175: arm 1 = ZDV+ddI (experimental), arm 3 = ddI (control)
actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$id <- seq_len(nrow(actg_df))

# Treatment: ZDV+ddI = 1, ddI = 0 (no switching)
actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)
```

## Outcome construction

```{r data-outcomes}
# Drop subjects missing the week-20 CD4 measurement
actg_df <- actg_df[!is.na(actg_df$cd420), ]

# Beneficial scale: CD4 improvement indicator (1 = improved, higher = better)
actg_df$y_binary_pos <- as.integer(actg_df$cd420 > actg_df$cd40)

# Adverse scale: NO improvement indicator (1 = did not improve, higher = worse).
# This is the single analysis scale for the entire document.
actg_df$y_binary_neg <- 1L - actg_df$y_binary_pos

cat("Sample size:", nrow(actg_df), "\n")
cat("Adverse events (no improvement):", sum(actg_df$y_binary_neg),
    sprintf("(%.1f%%)\n", 100 * mean(actg_df$y_binary_neg)))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```

## Covariates

```{r data-covariates}
# Keep ALL candidate covariates numeric.  forestsearch() builds and evaluates
# "var <= value" candidate cuts, so binary indicators left as factors trigger
# "'<=' not meaningful for factors" warnings during the consistency search.
# With 0/1 numeric coding, cont.cutoff (= 4 unique values) treats the binary
# candidates as categorical automatically -- no factor conversion is needed.
for (v in bin_vars) actg_df[[v]] <- as.numeric(actg_df[[v]])

cat("Continuous candidates:", paste(cont_vars, collapse = ", "), "\n")
cat("Binary candidates:    ", paste(bin_vars, collapse = ", "), "\n")
```

## Baseline characteristics

```{r baseline-table}
create_summary_table(
  data = actg_df,
  treat_var = treat.name,
  table_title = "ACTG175 Baseline Characteristics and Outcome by Treatment Arm",
  vars_continuous = c("y_binary_neg", cont_vars),
  vars_categorical = bin_vars,
  font_size = 12
)
```

## ITT analysis (adverse scale)

```{r itt-analysis}
# Adverse-scale ITT OR (y_binary_neg = no improvement; OR > 1 = treatment
# increases the odds of no improvement = harm).
itt_neg <- glm(y_binary_neg ~ treat, data = actg_df, family = binomial)

beta      <- coef(itt_neg)["treat"]
se        <- summary(itt_neg)$coefficients["treat", "Std. Error"]
z         <- beta / se
p         <- 2 * pnorm(-abs(z))
itt_or    <- exp(beta)
itt_ci_lo <- exp(beta - qnorm(0.975) * se)
itt_ci_hi <- exp(beta + qnorm(0.975) * se)

cat("=== ITT Odds Ratio, adverse scale (y_binary_neg) ===\n")
cat(sprintf("  log-OR (treat): %7.4f  (SE %.4f)\n", beta, se))
cat(sprintf("  OR:             %6.3f  (95%% CI: %.3f, %.3f)\n",
            itt_or, itt_ci_lo, itt_ci_hi))
cat(sprintf("  p-value:        %.4f\n", p))
cat("\nOn the adverse scale, OR > 1 indicates the treatment raises the odds of\n")
cat("no improvement (harm); OR < 1 indicates fewer non-improvers (benefit).\n")
```

# ForestSearch analysis

## Parameter provenance {#sec-param}

Each `forestsearch()` fit carries `args_call_all`: the argument list the call
actually resolved to, kept in sync by the package's `.sync_args_call_all()`. The
table below reconciles it against `formals(forestsearch)` at render time, so the
values reported are the ones the fit used rather than restated comments that can
drift.

```{r param-provenance-helper}
#| code-fold: false

.fs_fmt_val <- function(v) {
  if (is.null(v)) return("NULL")
  if (is.data.frame(v)) return(sprintf("<data.frame %d x %d>", nrow(v), ncol(v)))
  if (is.function(v)) return("<function>")
  if (is.list(v)) {
    if (length(v) == 0L) return("list()")
    nms <- names(v); if (is.null(nms)) nms <- rep("", length(v))
    inner <- vapply(seq_along(v), function(i) {
      val <- tryCatch(paste(deparse(v[[i]], width.cutoff = 200L), collapse = " "),
                      error = function(e) "<?>")
      if (nzchar(nms[i])) paste0(nms[i], " = ", val) else val
    }, character(1))
    return(paste0("list(", paste(inner, collapse = ", "), ")"))
  }
  if (length(v) > 6L) return(sprintf("<%s[%d]>", class(v)[1], length(v)))
  paste(deparse(v, width.cutoff = 200L), collapse = " ")
}

.fs_fmt_default <- function(x) {
  if (identical(x, quote(expr = ))) return("<required; no default>")
  paste(deparse(x, width.cutoff = 200L), collapse = " ")
}

# match.arg-style values resolve to their first element.  Applied to BOTH sides
# of the comparison so a value captured unresolved does not read as differing.
.fs_collapse_matcharg <- function(x) {
  ifelse(!is.na(x) & grepl('^c[(]"', x),
         sub('^c[(]"([^"]*)".*$', '"\\1"', x), x)
}

#' Reconcile a fit's resolved arguments against forestsearch()'s formals
fs_param_provenance <- function(fit, fn = forestsearch) {
  stopifnot(!is.null(fit), !is.null(fit$args_call_all))
  args <- fit$args_call_all
  fm   <- formals(fn)
  nm   <- setdiff(names(fm), "...")

  defaults <- vapply(fm[nm], .fs_fmt_default, character(1), USE.NAMES = FALSE)
  used <- vapply(nm, function(p) {
    if (!p %in% names(args)) return(NA_character_)
    .fs_fmt_val(args[[p]])
  }, character(1), USE.NAMES = FALSE)

  eff      <- .fs_collapse_matcharg(defaults)
  eff_used <- .fs_collapse_matcharg(used)

  out <- data.frame(
    parameter  = nm,
    default    = defaults,
    value_used = ifelse(is.na(used), "(not captured)", used),
    status     = ifelse(is.na(used), "not captured",
                 ifelse(eff_used == eff, "at default", "differs from default")),
    stringsAsFactors = FALSE
  )

  # forestsearch_main.R:1309 overrides five settings under sg_focus = "maxeff"
  # and does NOT re-sync args_call_all afterwards.  This document pre-resolves
  # those five to the engine-forced values, so the capture should already
  # agree.  VERIFY rather than assume.
  if (identical(args$sg_focus, "maxeff")) {
    forced <- list(pconsistency.threshold = 0, stop_threshold = NULL,
                   use_twostage = FALSE, max_subgroups_search = Inf, minp = 0)
    for (p in names(forced)) {
      i <- which(out$parameter == p)
      if (!length(i)) next
      agrees <- isTRUE(all.equal(args[[p]], forced[[p]]))
      out$status[i] <- if (agrees)
        "maxeff engine-forced value (pre-resolved in this call)" else
        "STALE -- captured BEFORE the maxeff override; not what the engine used"
    }
  }

  # stop_threshold is reset to NULL for these foci, but that branch DOES call
  # .sync_args_call_all(), so the capture is already truthful; label it.
  if (args$sg_focus %in% c("hrMaxSG", "hrMinSG", "hr", "maxSG", "minSG")) {
    i <- which(out$parameter == "stop_threshold")
    if (length(i) && is.null(args$stop_threshold))
      out$status[i] <- "NULL: early stopping unsound for this key (re-synced)"
  }

  out
}

#' Render a provenance table for one fit
fs_param_table <- function(fit, title) {
  pv <- fs_param_provenance(fit)
  n_diff <- sum(pv$status == "differs from default")
  sub <- sprintf(
    "%d of %d formals differ from their package default (forestsearch %s, read at render time).",
    n_diff, nrow(pv),
    tryCatch(as.character(utils::packageVersion("forestsearch")),
             error = function(e) "version unknown"))
  if (any(grepl("^STALE", pv$status))) {
    sub <- paste(sub,
      "WARNING: one or more rows were captured before the sg_focus = \"maxeff\"",
      "override and do not report the value the engine used.")
  }
  knitr::kable(pv, caption = paste0(title, " -- ", sub))
}
```

## Running ForestSearch

Candidate generation uses the consistency search with LASSO and GRF screening
contributing cuts; $J = 10$ cuts are used for the continuous `cd40` and `wtkg`
covariates. Selection is driven entirely by the focus profile resolved above.

::: {.callout-note}
## API note: GLM-natural vocabulary

This document uses the GLM-natural parameter names: `sg_focus = "effMaxSG"`
(alias for the legacy `"hrMaxSG"`), `effect.threshold` (alias for
`hr.threshold`), and `consistency.threshold` (alias for `hr.consistency`). The
`"eff*"` and `effect.*` forms read more naturally where there is no hazard
ratio. The legacy `"hr*"` and `hr.*` forms continue to work and produce
identical results.
:::

```{r forestsearch-fit}
#| fig-width: 10
#| fig-height: 7

t0 <- proc.time()

fs <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis               = actg_df,
  confounders.name          = confounders.name,
  outcome.name              = outcome.name,
  event.name                = outcome.name,    # binary: outcome IS the event
  treat.name                = treat.name,
  id.name                   = id.name,
  is.RCT                    = TRUE,
  seedit                    = analysis_seed,
  # --- GLM parameters -----------------------------------------------------
  outcome_type              = fs_outcome_type,
  effect_measure            = fs_effect_measure,
  adverse_outcome           = TRUE,
  # --- Screening engines (candidate generation) ---------------------------
  use_lasso                 = TRUE,
  use_grf                   = TRUE,
  grf_depth                 = 2L,
  dmin.grf                  = 0,
  frac.tau                  = 0.80,
  vi.grf.min                = -0.2,
  return_selected_cuts_only = TRUE,
  tune_grf                  = FALSE,
  use_dina                  = FALSE,
  # --- Candidate-cut construction -----------------------------------------
  cut_type                  = "default",
  cont.cutoff               = 4L,
  conf.cont_jcuts           = list(cd40 = 10, wtkg = 10),
  # --- Selection engine: driven entirely by the focus profile -------------
  subgroup_method           = "consistency",
  sg_focus                  = focus$supplied,
  selection_rule            = focus$selection_rule,
  effect_neighborhood       = focus$effect_neighborhood,
  # --- Tier-2 de-biased gate ----------------------------------------------
  # MR re-selects under the SAME rule the search used --
  # .fs_mr_reselection_from_focus() maps hrMaxSG -> effMaxSG,
  # hrMinSG -> effMinSG, hr -> maxcons, and so on -- so no reselection
  # argument is needed here.
  mr_inference              = run_mr,
  mr_inference_args         = list(draws = mr_draws),
  # --- Subgroup constraints (retained under every focus) ------------------
  n.min                     = 60L,
  d0.min                    = 10L,
  d1.min                    = 10L,
  maxk                      = 2L,
  # --- Thresholds + search settings, pre-resolved by the focus profile ----
  effect.threshold          = focus$effect_threshold,
  consistency.threshold     = focus$consistency_threshold,
  pconsistency.threshold    = focus$pconsistency,
  max_subgroups_search      = focus$max_subgroups_search,
  minp                      = focus$minp,
  use_twostage              = focus$use_twostage,
  stop_threshold            = focus$stop_threshold,
  consistency_method        = fs_consistency_method,
  fs.splits                 = 500L,
  # --- Output / diagnostics -----------------------------------------------
  show_candidate_summary    = TRUE,
  details                   = TRUE,
  quiet                     = FALSE,
  plot.sg                   = FALSE,   # KM not applicable; see plot_sg_glm_outcomes()
  # --- Parallel -----------------------------------------------------------
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential"); gc(verbose = FALSE)
timings$forestsearch <- (proc.time() - t0)[["elapsed"]]

cat("\nForestSearch completed in",
    round(timings$forestsearch, 1), "seconds\n")
cat("sg_focus (canonical):", fs$sg_focus, "\n")

# The engine's canonical focus must agree with the profile; if it does not,
# every gate below is keyed on the wrong value.
stopifnot(identical(fs$sg_focus, focus$canonical))
```

```{r param-provenance}
#| echo: false
fs_param_table(fs, sprintf("Resolved parameters for the %s fit", focus$supplied))
```

## ForestSearch results

### Identified subgroups

```{r fs-results}
res_tabs <- sg_tables(fs, ndecimals = display_digits, which_df = "est")
res_tabs$sg10_out
```

### Treatment effect estimates

```{r fs-estimates}
res_tabs$tab_estimates
```

### Identified subgroup definition

```{r fs-subgroup}
if (!is.null(fs$sg.harm) && length(fs$sg.harm) > 0) {
  cat("Identified subgroup (H):", fmt_sg(fs$sg.harm), "\n")
  cat("Subgroup size:", sum(fs$df.est$treat.recommend == 0),
      sprintf("(%.1f%% of ITT)\n",
              100 * mean(fs$df.est$treat.recommend == 0)))
} else {
  cat("ForestSearch: no subgroup identified at the configured thresholds.\n")
}
```

## Selection diagnostic: Pareto frontier {#sec-pareto}

```{r pareto-skip-note, eval = !run_pareto}
#| echo: false
#| results: asis
cat(sprintf(paste(
  "_The Pareto section is skipped for `sg_focus = \"%s\"`._\n\n",
  "The frontier block runs only for the banded foci `effMaxSG` / `effMinSG`",
  "(canonical `hrMaxSG` / `hrMinSG`), where `.compute_inclusion_band()` makes",
  "the (OR, N) geometry part of the **selection rule** -- `in_band` leads the",
  "sort key, and `selection_rule` switches between the effect neighbourhood, the",
  "frontier, and their intersection. Under `%s` the key contains no `in_band`",
  "term, so a frontier plot would be a post-hoc curiosity rather than a",
  "diagnostic of how the winner was chosen.\n\n",
  "Under `maxeff` there is a second reason: the winner-only fast path scores",
  "consistency for the argmax alone and carries every other candidate with",
  "`Pcons = NA`, so the frontier can run to hundreds of members with an",
  "all-missing consistency column, and `compute_frontier_cis()` costs scale with",
  "frontier size x `n_splits`.\n"),
  focus$supplied, focus$supplied))
```

```{r pareto-intro, eval = run_pareto}
#| echo: false
#| results: asis
cat(sprintf(paste(
  "The *Pareto frontier* on (OR, N) is the set of candidate subgroups not",
  "dominated on both axes simultaneously: a candidate is dominated when another",
  "has at least as large an effect **and** at least as large a sample size, with",
  "at least one strict inequality.\n\n",
  "For `sg_focus = \"%s\"` with `selection_rule = \"%s\"`, this is not a post-hoc",
  "diagnostic -- it is the selection geometry. `.compute_inclusion_band()` builds",
  "`in_nbhd` (effect within %.0f%% of the strongest surviving effect) and",
  "`in_pareto` (non-dominated), and %s. The `in_band` indicator then **leads** the",
  "sort key %s, so what the figure and table below display is the space the rule",
  "actually chose within.%s\n\n",
  "OR is a ratio measure stored on the log scale, so both band tests are applied",
  "after exponentiating and the rendered values are on the natural OR scale; the",
  "effect axis is auto-rendered on log scale.\n"),
  focus$supplied, focus$selection_rule, 100 * focus$effect_neighborhood,
  switch(focus$selection_rule,
    neighborhood = "the rule uses the effect band alone",
    pareto       = "the rule uses the frontier alone",
    both         = "the rule uses their intersection"),
  focus$key,
  if (identical(focus$canonical, "hrMinSG"))
    paste(" Because this focus minimises size within the band, the selected",
          "subgroup may itself be N-dominated by a larger in-band neighbour and",
          "so may not appear on the frontier; that is the rule working as",
          "specified.") else ""))
```

```{r pareto-cis, eval = run_pareto}
# Split-derived CIs for the frontier members.  This is the expensive step; the
# same intervals feed the figure, the table and the narrative, so it is
# computed once.  Set frontier_params$compute_cis <- FALSE to skip.
t0 <- proc.time()
ci_tab <- if (isTRUE(frontier_params$compute_cis)) {
  compute_frontier_cis(
    fs,
    n_splits = frontier_params$n_splits,
    seed     = frontier_params$seed
  )
} else {
  NULL
}
timings$frontier <- (proc.time() - t0)[["elapsed"]]
cat("Frontier CIs:", if (is.null(ci_tab)) "skipped" else
      sprintf("%d member(s) over %s splits in %.1f s",
              nrow(ci_tab), format(frontier_params$n_splits, big.mark = ","),
              timings$frontier), "\n")
```

```{r pareto-plot, eval = run_pareto}
#| fig-width: 9
#| fig-height: 6
#| fig-cap: "Pareto frontier on (OR, N). The shaded region is the effect-size inclusion band; the selected subgroup is marked. Horizontal bars are split-derived intervals when computed."
p_pareto <- plot_pareto_frontier(
  fs,
  ci_table            = ci_tab,
  show_band           = TRUE,
  effect_neighborhood = focus$effect_neighborhood,
  xlim                = frontier_params$xlim,
  xlim_trim           = frontier_params$xlim_trim
)
print(p_pareto)
```

```{r pareto-table, eval = run_pareto}
# Formatted gt table; the selected subgroup is highlighted and marked with a
# star.  Effect column label and scale derive from fs$effect_measure.
pareto_frontier_table(
  fs,
  ci_table = ci_tab,
  digits   = frontier_params$digits
)
```

```{r pareto-explain, eval = run_pareto}
#| results: asis
# Narrative comparison of the winner against each non-dominated alternative,
# phrased in terms of the focus's own lexicographic key.
cat(explain_pareto_selection(fs, ci_table = ci_tab,
                             digits = frontier_params$digits))
```

```{r pareto-data, eval = FALSE}
# Programmatic access (effect on the natural OR scale, plus is_selected /
# on_frontier / in_band flags):
pareto_frontier_table(fs, format = "data.table", include_dominated = TRUE)
```

# Bootstrap bias correction

## Rationale

GLM estimates from identified subgroups are **upwardly biased** because
subgroups are selected *because* they show extreme effects. The full bootstrap
re-runs the entire selection on each resample, computes bias terms from
bootstrap versus observed estimates, and applies infinitesimal-jackknife
variance estimation. Because it re-runs the selection, it inherits whatever
focus this render uses without further configuration -- and its cost scales with
`NB`.

## Running the bootstrap

```{r bootstrap}
t0 <- proc.time()

fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est = fs,
  nb_boots = NB,
  show_three = FALSE,
  details = FALSE,
  digits = construction_digits,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential"); gc(verbose = FALSE)
timings$bootstrap <- (proc.time() - t0)[["elapsed"]]

cat("\nBootstrap completed in",
    round(timings$bootstrap / 60, 1), "minutes\n")
```

## Bootstrap summary and diagnostics

```{r bootstrap-summary}
# The OR column (with bias-corrected OR*) is auto-detected from FSsg_tab; the
# est.scale argument only affects the legacy HR-flavored caption helper.
summaries <- summarize_bootstrap_results(
  sgharm = fs$sg.harm,
  boot_results = fs_bc,
  create_plots = TRUE,
  est.scale = "hr",
  digits = display_digits
)

summaries$table
```

# Multiplier resampling (MR)

The Tier-2 de-biased gate is a fast multiplier-bootstrap approximation of the
bootstrap bias correction. It reuses per-candidate GLM influence (`dfbeta`) over
the consistency search's own candidate family and re-selects under multiplier
draws, so its selection-bias term mirrors the same search optimism the full
bootstrap corrects -- without re-running the search per resample. It holds the
fitted family fixed and is vectorized, so its cost does not scale with the
resample count.

MR re-selects under the same rule the search used. `sg_focus` is the single
source of truth for that mapping, so the re-selection reported below should match
this render's focus rather than a default.

```{r mr-gate, eval = !is.null(fs$sg.harm)}
#| echo: false
#| results: asis
g <- fs$mr_inference
if (is.null(g)) {
  cat("_The de-biased gate did not run (set `run_mr <- TRUE`, and note the GLM",
      "path requires `consistency_method = \"resample\"`)._\n")
} else {
  timings$mr <- g$timing_seconds
  fmt_ci <- function(e, l, u) sprintf("%.3f (%.3f, %.3f)", e, l, u)
  fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0  <- fl(fs_bc$H_estimates, "H0")
  H0l <- fl(fs_bc$H_estimates, "H0_lower"); H0u <- fl(fs_bc$H_estimates, "H0_upper")
  H2  <- fl(fs_bc$H_estimates, "H2")
  H2l <- fl(fs_bc$H_estimates, "H2_lower"); H2u <- fl(fs_bc$H_estimates, "H2_upper")

  # Complement: Tier-2 from g$complement (include_complement = TRUE); Tier-1
  # from the bootstrap's Hc_estimates.
  cc     <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  Hc0  <- fl(fs_bc$Hc_estimates, "H0")
  Hc0l <- fl(fs_bc$Hc_estimates, "H0_lower"); Hc0u <- fl(fs_bc$Hc_estimates, "H0_upper")
  Hc2  <- fl(fs_bc$Hc_estimates, "H2")
  Hc2l <- fl(fs_bc$Hc_estimates, "H2_lower"); Hc2u <- fl(fs_bc$Hc_estimates, "H2_upper")

  q  <- c("naive OR (95% CI)", "de-biased OR (95% CI)")
  t2 <- c(fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  t1 <- c(fmt_ci(H0, H0l, H0u), fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    q  <- c(q, "complement naive OR (95% CI)", "complement de-biased OR (95% CI)")
    t2 <- c(t2, fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    t1 <- c(t1, fmt_ci(Hc0, Hc0l, Hc0u), fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `MR (Tier 2)` = t2,
                     `Full bootstrap (Tier 1)` = t1, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE,
        caption = sprintf("Selected subgroup: %s%s (sg_focus = '%s')",
                          fmt_sg(fs$sg.harm),
                          if (has_cc) " (complement = remaining subjects)" else "",
                          focus$supplied)),
      sep = "\n"); cat("\n\n")

  # Scale cross-check: the gate's naive OR and the bootstrap's naive H0 are the
  # same quantity (raw subgroup OR) computed two ways.
  if (is.finite(H0) && is.finite(g$naive$est)) {
    scale_ok <- abs(log(g$naive$est) - log(H0)) < 1e-2
    cat(sprintf("Scale cross-check (MR naive vs bootstrap H0): **%s** (%.3f vs %.3f).\n\n",
                if (scale_ok) "PASS" else "CHECK", g$naive$est, H0))
  }

  cat(sprintf("Candidate family: %d (full <= maxk space).\n\n", g$n_family))
  cat(sprintf("Re-selection rule = `%s`, matching sg_focus = `%s` (canonical `%s`).\n\n",
              if (!is.null(g$settings$reselection)) g$settings$reselection else "?",
              focus$supplied, focus$canonical))

  boot_s <- as.numeric(timings$bootstrap)
  cat(sprintf("Timing: full bootstrap (%d boots) %.1f s (%.1f min); MR gate %.2f s",
              NB, boot_s, boot_s / 60, g$timing_seconds))
  if (is.finite(boot_s) && g$timing_seconds > 0)
    cat(sprintf(" -> **%.0fx** faster (the full bootstrap scales with `NB`, MR does not)",
                boot_s / g$timing_seconds))
  cat(".\n\n")

  cat(sprintf("Gate `mr_harm_confirmed` = **%s** (de-biased OR %.3f %s gate %.2f); selection bias %.4f (log-OR).\n",
              fs$mr_harm_confirmed, g$debiased$est,
              if (g$settings$confirm_rule == "point") "vs" else "lower-bound vs",
              g$settings$t_confirm, g$selection_bias))
  if (!is.null(g$ci_method) && !is.null(g$debiased$se_ij))
    cat(sprintf("De-biased CI: %s (IJ SE %.3f vs robust SE %.3f, log-OR).\n",
                if (identical(g$ci_method, "ij")) "infinitesimal jackknife (Tier-1 analogue)" else "robust SE",
                g$debiased$se_ij, g$debiased$se_wald))
  if (has_cc)
    cat(sprintf("Complement (n = %d): naive OR %.3f -> de-biased %.3f (selection bias %.4f, log-OR).\n",
                cc$n, cc$naive$est, cc$debiased$est, cc$selection_bias))
}
```

```{r mr-no-subgroup, eval = is.null(fs$sg.harm)}
#| echo: false
#| results: asis
cat("_No harm subgroup was selected, so the de-biased gate comparison does not apply._\n")
```

The same bias-corrected estimates table, rebuilt from the MR object. The
descriptive columns and naive effect are tier-independent (inherited from the
bootstrap result used as a template); the bias-corrected column carries MR's
de-biased estimate for **both** the selected subgroup and its complement.

```{r mr-table, eval = !is.null(fs$sg.harm) && !is.null(fs$mr_inference)}
mr_estimates_table(fs, boot_results = fs_bc, est.scale = "hr",
                   digits = display_digits)
```

## Subgroup outcome visualization

```{r fs-glm-plot}
#| fig-width: 10
#| fig-height: 7
#| fig-cap: "Rate of no improvement by treatment arm and identified subgroup (adverse scale)"

fs_plot <- plot_sg_glm_outcomes(
  fs.est         = fs,
  outcome.name   = outcome.name,
  treat.name     = treat.name,
  effect_measure = fs_effect_measure,
  outcome_type   = fs_outcome_type,
  E.name         = "ZDV+ddI",
  C.name         = "ddI",
  show_itt       = TRUE,
  show_effect    = TRUE,
  verbose        = TRUE
)
```

## Bootstrap diagnostics

```{r bootstrap-diagnostics}
summaries$diagnostics_table_gt
```

### Subgroup agreement

How consistently does the bootstrap identify the same subgroup?

```{r subgroup-agreement}
if (!is.null(summaries$subgroup_summary$original_agreement)) {
  summaries$subgroup_summary$original_agreement
}

if (!is.null(summaries$subgroup_summary$factor_presence)) {
  summaries$subgroup_summary$factor_presence
}
```

### Bootstrap distributions

```{r bootstrap-plots}
#| fig-width: 10
#| fig-height: 8
if (!is.null(summaries$plots)) summaries$plots
```

# Cross-validation

Cross-validation assesses the **stability of the subgroup definition** under
this focus. It is not a de-biasing correction of the effect.

## Repeated 10-fold cross-validation

```{r cv-skipped-note, eval = !run_cv}
cat("K-Fold cross-validation skipped (run_cv = FALSE). Set run_cv <- TRUE",
    "in the parallel-setup chunk to run it.\n")
```

```{r kfold-cv, eval = run_cv}
t0 <- proc.time()

fs_kfold <- forestsearch_tenfold(
  fs.est = fs,
  sims = Ksims,
  Kfolds = 10L,
  details = FALSE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = FALSE
  )
)

plan("sequential"); gc(verbose = FALSE)
timings$kfold <- (proc.time() - t0)[["elapsed"]]
cat("10-fold CV:", Ksims, "x 10 partitions in",
    round(timings$kfold / 60, 1), "min\n")
metrics_tables <- cv_metrics_tables(fs_kfold)
metrics_tables
```

### Diagnostic tabulations

`summarize_cv_results()` consolidates the per-fold record into a panel of
diagnostic tables: which subgroups the pipeline identified across resamples,
which GRF cuts drove those identifications, when identification failed what the
proximate cause was, and how confidently the successful identifications were
made.

```{r cv-diagnostics-run, eval = run_cv}
cv_diag <- summarize_cv_results(
  cv_output         = fs_kfold,
  original_sg       = fs$sg.harm,
  original_grf_cuts = fs$grf_cuts,
  top_n             = 10L,
  create_plots      = TRUE
)
```

```{r cv-identification, eval = run_cv}
cv_diag$identification_summary
```

```{r cv-grf-cut, eval = run_cv}
cv_diag$grf_cut_summary
```

```{r cv-no-subgroup, eval = run_cv}
cv_diag$no_subgroup_decomposition
```

```{r cv-pcons-note}
#| echo: false
#| results: asis
cat(if (focus$is_maxeff) paste(
  "Under `maxeff` no consistency threshold is applied at selection, so the",
  "distribution below is descriptive of the selected argmax rather than a",
  "gate.\n") else if (focus$is_banded) paste(
  "Under a banded focus consistency is a **filter** at admission and then a",
  "third-tier key after `in_band` and size, so the distribution reflects both",
  "roles.\n") else paste(
  "Consistency is a filter at admission for this focus, so every value shown",
  "cleared `pconsistency.threshold`.\n"))
```

```{r cv-pcons, eval = run_cv}
cv_diag$pconsistency_distribution
```

```{r cv-fold-numeric, eval = run_cv}
cv_diag$fold_numeric_summary
```

```{r cv-original-agreement, eval = run_cv}
cv_diag$original_agreement
```

```{r cv-identification-plot, eval = run_cv}
#| fig-width: 9
#| fig-height: 5
#| fig-cap: "Distribution of identified subgroups across repeated 10-fold training splits."
if (!is.null(cv_diag$plots) && !is.null(cv_diag$plots$identification)) {
  cv_diag$plots$identification
}
```

## Leave-one-out (out-of-bag, N-fold)

Leave-one-out re-runs the entire selection on each hold-out of one subject and
predicts the held-out subject's subgroup membership from its covariates. It asks
whether the same subgroup is recovered when no subject informs its own label.

```{r loo-focus-note}
#| echo: false
#| results: asis
cat(sprintf(paste(
  "How discriminating leave-one-out is depends on the focus. Under `%s` the key",
  "is %s. %s\n"),
  focus$supplied, focus$key,
  switch(focus$canonical,
    maxeff = paste(
      "With the full candidate family and no auxiliary condition, the argmax has",
      "genuine room to land on different definitions across folds, so this is the",
      "discriminating comparison: whether the out-of-sample estimate tracks Naive",
      "(definition perfectly stable) or one of the corrections (definition varies",
      "fold to fold) is informative rather than predetermined."),
    maxeffCons = paste(
      "The consistency filter removes the least replicable candidates before the",
      "effect argmax runs, so the definition is typically steadier than under",
      "`maxeff` -- but the winner can still move when a marginal qualifier drops",
      "out of a fold."),
    hrMaxSG = ,
    hrMinSG = paste(
      "The band restricts the competition to near-optimal candidates and size then",
      "decides among them, which tends to make the definition steadier than a pure",
      "effect argmax: fold-to-fold movement requires the band membership or the",
      "size ordering to change, not merely the effect ranking."),
    maxSG = ,
    minSG = paste(
      "Size leads the key and is the most stable quantity in the table, so the",
      "definition is usually recovered exactly and leave-one-out will often",
      "coincide with Naive by construction."),
    paste(
      "Consistency leads the key, and consistency is estimated from repeated",
      "splits rather than a single fit, so the definition is usually steady;",
      "coincidence with Naive should be read as stability, not as a failure to",
      "correct."))))
```

```{r loo-skipped-note, eval = !run_loo}
cat("Leave-one-out (N-fold / OOB) skipped (run_loo = FALSE). Set run_loo <- TRUE",
    "in the parallel-setup chunk to run it.\n")
```

```{r oob-cv, eval = run_loo}
t0 <- proc.time()

fs_OOB <- forestsearch_Kfold(
  fs.est = fs,
  details = FALSE,
  Kfolds = nrow(fs$df.est),  # N-fold = leave-one-out
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential"); gc(verbose = FALSE)
timings$oob <- (proc.time() - t0)[["elapsed"]]

cv_out <- forestsearch_KfoldOut(
  res = fs_OOB,
  details = FALSE,
  outall = TRUE
)

tables <- cv_summary_tables(cv_out)
tables$combined_table
tables$metrics_table
```

# Comparison: corrections on the selected subgroup {#sec-table}

```{r comparison-table}
#| echo: false
fci <- function(e, l, u, d = display_digits) {
  ifelse(is.na(e), "\u2014", sprintf("%.*f (%.*f, %.*f)", d, e, d, l, d, u))
}
fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)

naive_or <- fl(fs_bc$H_estimates, "H0")
g <- fs$mr_inference

method   <- c("Naive", "Full bootstrap")
estimate <- c(sprintf("%.2f", naive_or),
              fci(fl(fs_bc$H_estimates, "H2"),
                  fl(fs_bc$H_estimates, "H2_lower"),
                  fl(fs_bc$H_estimates, "H2_upper")))
if (!is.null(g)) {
  method   <- c(method, "Multiplier resampling")
  estimate <- c(estimate,
                fci(g$debiased$est, g$debiased$lower, g$debiased$upper))
}

kbl(data.frame(Method = method, Estimate = estimate, check.names = FALSE),
    booktabs = TRUE, escape = FALSE, align = "ll",
    caption = sprintf(paste(
      "Subgroup odds ratios (adverse scale) for the ACTG175 selected subgroup",
      "under sg_focus = '%s'. Parenthetical values are 95%% intervals."),
      focus$supplied)) |>
  kable_styling()
```

::: {.callout-note}
## Reading the table

Naive is the uncorrected subgroup odds ratio. Full bootstrap and Multiplier
resampling both target the conditional estimand and report two-sided intervals;
they differ in that FB re-runs the entire selection on each resample (so its cost
scales with the resample count) while MR holds the fitted family fixed and is
vectorized (so its cost does not). FB and MR are a trade, not a strict ordering:
FB marginalizes over family regeneration, MR conditions on the realized family.

There is no Guo & He row in this document. `fs_to_guohe()` is a survival-path
adapter and additionally requires `sg_focus = "maxeff"`; neither condition is met
by a binary OR analysis under an arbitrary focus. The survival companion
(`analysis_gbsg_survival_sgfocus.qmd`) carries that comparison.
:::

```{r timing}
#| echo: false
td <- data.frame(
  Step = c("Fit + MR gate", "Frontier CIs", "Full bootstrap",
           "MR gate (within fit)", "10-fold CV", "Leave-one-out (N-fold)"),
  Seconds = c(timings$forestsearch, timings$frontier, timings$bootstrap,
              timings$mr, timings$kfold, timings$oob)
)
td$Minutes <- round(td$Seconds / 60, 2)
kbl(td, booktabs = TRUE, align = "lrr",
    caption = sprintf("Computational timing (sg_focus = '%s'). NA marks a step this focus or configuration did not run.",
                      focus$supplied)) |>
  kable_styling()

cat(sprintf("\nTotal document time: %.1f min\n",
            (proc.time() - t_vignette_start)[["elapsed"]] / 60))
```

# Session information

```{r session-info}
#| code-fold: true
sessionInfo()
```
