---
title: "Exploratory Subgroup Identification with ForestSearch"
subtitle: "ACTG175 Binary Harm Search (GLM Odds-Ratio Outcome)"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
execute:
  warning: true
  message: true
---

> **Method-general analysis.** The selection engine for the main analysis is set
> once via `fs_subgroup_method` in the setup chunk — `"consistency"` (default),
> `"dina"`, or `"grf"` (with `fs_grf_selection` = `"tree"`/`"frontier"`). The
> headline fit and every downstream step — full bootstrap, Tier-2 de-biased gate,
> cross-validation, leave-one-out, and the subgroup outcome plot — inherit that
> engine from the single fitted object, so the whole analysis re-renders under a
> different method by changing one knob. The comparative DINA-screening,
> standalone-DINA, and GRF sections that follow are fixed illustrations of the
> individual engines and are unaffected by `fs_subgroup_method`.
>
> This is the **binary-outcome (odds-ratio) counterpart** to the GBSG Cox
> multimethod vignette. Everything operates on a single **adverse scale**
> (`y_binary_neg`, the indicator of *no* CD4 improvement), so an OR above the
> screening threshold indicates harm — the direct analogue of an HR above the
> threshold on the survival path. The Tier-1 (full bootstrap) vs Tier-2 (fast
> multiplier-bootstrap de-biased gate) comparison is reported for each engine.

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# Initialize timing
timings <- list()
t_vignette_start <- proc.time()
```

## Parallel Processing Configuration

ForestSearch supports parallel processing for computationally intensive
operations (bootstrap, cross-validation).

```{r parallel-setup}
# Detect available cores
n_cores_total <- parallel::detectCores()
n_cores <- max(1L, floor(0.80 * parallel::detectCores(logical = FALSE)))
cat("Using", n_cores, "of", n_cores_total, "total cores for parallel processing")

# Number of bootstrap iterations
# Use 500-2000 for production; reduced here for demo / pressure test
NB <- 1000L

# ---------------------------------------------------------------------------
# Validation toggles (the slowest parts of the document).  Independent:
# run_cv  controls K-Fold cross-validation; run_loo controls leave-one-out
# (N-fold / OOB).  Either can be turned off without affecting the other.
# ---------------------------------------------------------------------------
run_cv  <- FALSE    # K-Fold cross-validation (kfold-cv + diagnostics)
run_loo <- FALSE    # Leave-one-out / N-fold OOB (oob-cv)
# 10-fold CV with multiple iterations.  Use Ksims >= 50 for production.
Ksims <- 50L

# Pre-seed CV/LOO timings so the Computational Timing table is well-formed
# whether or not these run (skipped analyses report NA).
timings$kfold <- NA_real_
timings$oob   <- NA_real_

# ---------------------------------------------------------------------------
# Per-analysis identification parameters (tune each analysis independently).
#   sg_focus       : which candidate subgroup the selection step chooses
#   or_threshold   : harm floor on the OR scale (adverse scale: OR > c1 = harm)
#   or_consistency : minimum per-split consistency OR (consistency search only)
# ---------------------------------------------------------------------------
# Main ForestSearch (consistency search)
fs_sg_focus              <- "effMaxSG"   # GLM-natural alias of "hrMaxSG"
fs_selection_rule        <- "both"       # neighborhood ∩ Pareto (effMaxSG/effMinSG)
fs_or_threshold          <- 1.25
fs_or_consistency        <- 1.0
fs_pconsistency          <- 0.90
fs_max_subgroups_search  <- 20L
# IMPORTANT: the Tier-2 de-biased gate on the GLM path runs ONLY under the
# resample-consistency engine (consistency_method = "resample").  Under the
# default "split" engine the gate is silently skipped for GLM outcomes.
fs_consistency_method    <- "resample"
# Tier-2 de-biased gate: a fast multiplier-bootstrap approximation of the
# bootstrap bias-correction, computed inside the fit and compared to the full
# bootstrap below.  It runs for whichever selection engine fs_subgroup_method
# names, de-biasing against that engine's own candidate family (it re-derives
# the per-candidate GLM dfbeta itself).
run_mr <- TRUE     # add mr_inference = TRUE to the forestsearch call
mr_draws      <- 2000L    # multiplier-bootstrap draws for the gate

# ---------------------------------------------------------------------------
# Output location for this document's results:  <base>/<dirout>/
#   results_dir = NULL -> base is <qmd dir>/_payloads
#   dirout      = NULL -> the .qmd stem
# Setting results_dir redirects the base only; each run still gets its own
# subdirectory named by dirout.
#
# Examples:
#   results_dir <- NULL;  dirout <- NULL
#     -> <qmd dir>/_payloads/<stem>/<stem>_payload.rds
#   results_dir <- NULL;  dirout <- "gbsg_survival_multimethod"
#     -> <qmd dir>/_payloads/gbsg_survival_multimethod/gbsg_survival_multimethod_payload.rds
#   results_dir <- "~/GitHub/fs-post-selection-manuscript_jasa/_payloads"; dirout <- NULL
#     -> ~/GitHub/fs-post-selection-manuscript_jasa/_payloads/<stem>/<stem>_payload.rds
# ---------------------------------------------------------------------------
results_dir <- NULL
dirout      <- NULL

# ---------------------------------------------------------------------------
# MODEL-SCREENED FAMILY CONFIGURATION (use_lasso = TRUE, use_grf = TRUE)
# ---------------------------------------------------------------------------
# Conditional-on-family regime: the correction captures within-family
# selection optimism but not family regeneration (model screening happens
# once, before the search).
#
# Selects: {wtkg>83.46} & {cd40>413}, n=71
#   Naive OR 6.34, multiplier 4.00
#
# Uses 2000 MR draws, not the 5000 used for the published run.
#
# NOT the manuscript-reproducing document - see fixed_family.qmd for that.
# ---------------------------------------------------------------------------

# ---------------------------------------------------------------------------
# Selection engine for the MAIN analysis.  The headline fit, its bootstrap,
# the de-biased gate, CV/LOO, and the subgroup outcome plot all flow from this
# single fit, so flipping subgroup_method re-runs the entire main analysis
# under that engine (each downstream step inherits the method from the fitted
# object):
#   "consistency" - the splitting-consistency search (default);
#   "dina"        - DINA performs selection (consistency search bypassed);
#   "grf"         - GRF performs selection (policy tree or DR-frontier).
# The Tier-2 gate de-biases against whichever engine's OWN candidate family,
# with re-selection matching that engine's selection rule.
# ---------------------------------------------------------------------------
fs_subgroup_method <- "consistency"  # "consistency" | "dina" | "grf"
fs_grf_selection   <- "frontier"     # only if subgroup_method == "grf": "tree" | "frontier"
fs_grf_select_statistic <- "effect"  # only if subgroup_method=="grf" & grf_selection=="frontier": "dr" | "effect".
                                     #   "dr"     = native DR-score frontier winner;
                                     #   "effect" = re-rank the DR frontier on the inferential GLM OR (the
                                     #              statistic the Tier-2 gate de-biases).  No-op for the policy tree.
fs_grf_depth       <- 2L             # GRF depth (screening + GRF selection)
fs_dmin_grf        <- 0              # GRF DR-score screening floor (RD scale for binary; 0 = permissive)
fs_dina_args       <- list()         # extra args forwarded to DINA selection
fs_dina_select_statistic <- "effect" # only if subgroup_method=="dina": "dina" | "effect".
                                     #   "dina"   = native DINA tau-hat winner;
                                     #   "effect" = re-rank the DINA family on the inferential GLM OR (the
                                     #              statistic the Tier-2 gate de-biases).
# Tier-2 gate (main analysis): the gate's re-selection family for "dina"/"grf" is
# restricted to candidates competitive on the method's NATIVE statistic (DINA
# tau-hat / GRF DR-score).  NULL = use the recommended default, which matches the
# band to effect_neighborhood automatically (the band the engine selects within);
# give a number in (0, 1) to set the band explicitly, or >= 1 to de-bias against
# the full family.  Ignored for "consistency" (its maxcons re-selection already
# matches the search).
fs_family_native_nbhd <- NULL

# (A) ForestSearch + DINA screening (consistency search selects; DINA only
#     contributes candidate cuts)
dina_screen_sg_focus             <- "effMaxSG"
dina_screen_or_threshold         <- 1.0
dina_screen_or_consistency       <- 0.9
dina_screen_max_subgroups_search <- 30L
# Screening cut set: FALSE => full DINA frontier; TRUE => only the single
# dina_subgroup()-selected cut (uses dina_screen_sg_focus / selection_rule).
dina_screen_selected_only        <- TRUE

# (C) ForestSearch subgroup_method = "dina" (DINA performs selection; the
#     consistency search is bypassed, so or_consistency does not apply here).
#     The standalone DINA analysis (B) reuses these so (B) and (C) match.
fs_dina_sg_focus     <- "effMaxSG"
fs_dina_or_threshold <- 1.0
# Tier-2 gate: the DINA re-selection family is restricted to candidates competitive
# on DINA's native tau-hat.  NULL = recommended default (matches effect_neighborhood
# automatically); set a number in (0, 1) for an explicit band, or >= 1 for the full
# family.  Note: this only tunes candidate competition -- the gate holds the forest
# fixed, so the full bootstrap stays the reference for DINA regardless.
fs_dina_family_native_nbhd <- NULL

# (D) ForestSearch subgroup_method = "grf" (GRF performs selection via a causal
#     forest policy tree or DR-score frontier; the consistency search, LASSO and
#     DINA are bypassed).  Dedicated demo: always runs GRF regardless of
#     fs_subgroup_method, mirroring the standalone DINA selection (C).
fs_grf_demo_selection     <- "frontier"  # "frontier" | "tree"
fs_grf_demo_depth         <- 2L          # GRF tree / DR-pair depth
fs_grf_demo_dmin          <- 0.0         # DR-score harm floor (RD scale; 0 = permissive default)
fs_grf_sg_focus           <- "effMaxSG"  # -> GRF frontier/tree selection rule
# Tier-2 gate: the GRF re-selection family is restricted to candidates competitive
# on GRF's native (RD-based) DR-score.  NULL = recommended default (matches
# effect_neighborhood automatically); set a number in (0, 1) for an explicit band,
# or >= 1 for the full family.  Note: this only tunes candidate competition -- the
# gate holds the forest fixed, so the full bootstrap stays the reference for GRF
# regardless (a dominant DR-score winner can collapse the family to 1).
fs_grf_family_native_nbhd <- NULL

# ---------------------------------------------------------------------------
# Numeric precision (two-stage control).
#   construction_digits : upstream sprintf precision in the bootstrap / CV
#     table builders.  Acts as a CEILING for downstream display.
#   display_digits      : precision rendered in user-facing tables.
# ---------------------------------------------------------------------------
construction_digits <- 2L
display_digits      <- 2L

# Display-only tidy of a subgroup label for printing: rounds the post-operator
# thresholds to `d` decimals via the package's tidy_cut_display().  The
# membership-exact sg.harm (re-parsed by get_dfpred and carried through the
# bootstrap) is never altered -- this only affects what is printed.  Used at
# every site in this document that renders sg.harm.
fmt_sg <- function(x, d = display_digits) {
  if (is.null(x) || length(x) == 0L) return(x)
  paste(forestsearch:::tidy_cut_display(x, digits = as.integer(d)), collapse = " & ")
}

# ---------------------------------------------------------------------------
# Pareto-frontier diagnostic parameters (drives compute_frontier_cis(),
# plot_pareto_frontier(), pareto_frontier_table(), explain_pareto_selection()).
# For OR (ratio-scale), the effect axis is auto-rendered on log scale.
# ---------------------------------------------------------------------------
frontier_params <- list(
  n_splits      = 1000L,
  seed          = 1L,
  xlim          = c(0.35, 10),   # OR auto-detected as ratio scale
  digits        = 2L
)
```



# Introduction

This vignette demonstrates the **ForestSearch** methodology for exploratory
subgroup identification with a **binary outcome on the odds-ratio (OR) scale**,
applied to the ACTG175 trial. It uses the GLM extension of ForestSearch
(León et al. 2024 *Statistics in Medicine*) and is the binary-outcome
counterpart to the GBSG Cox multimethod analysis.

## Motivation

In clinical trials, particularly oncology and infectious-disease studies,
subgroup analyses are essential for:

- Evaluating treatment effect consistency across patient populations
- Identifying subgroups where treatment may be detrimental (harm)
- Characterizing subgroups with enhanced benefit
- Informing regulatory decisions and clinical practice

While prespecified subgroups provide stronger evidence, important subgroups
based on patient characteristics may not be anticipated. ForestSearch provides
a principled approach to **exploratory** subgroup identification with proper
statistical inference.

## Methodology Overview

ForestSearch identifies subgroups through:

1. **Candidate factor selection**: Using LASSO and/or Generalized Random Forests (GRF)
2. **Exhaustive subgroup search**: Evaluating all combinations up to `maxk` factors
3. **Consistency-based selection**: Applying splitting consistency criteria
4. **Bootstrap bias correction**: Adjusting for selection-induced optimism
5. **Cross-validation**: Assessing algorithm stability

The key innovation is the **splitting consistency criterion**: a subgroup is
considered "consistent with harm" if, when randomly split 50/50 many times,
both halves consistently show odds ratios at or above the harm threshold (for
example, 1.0 representing a nominal "harm" boundary on the adverse scale).

## Single-Scale (Adverse) Framing

::: {.callout-note}
**ForestSearch and threshold direction.** ForestSearch was developed for
survival data where the outcome is typically the timing of an adverse event and
thresholds (e.g., indicative of harm) are defined on the hazard-ratio scale.
For a binary endpoint, the same harm logic applies on the **adverse scale**: the
outcome `y_binary_neg = 1 - y_binary_pos` is the indicator of *no* CD4
improvement (1 = worse), and with `adverse_outcome = TRUE` an OR above the
screening threshold means the treatment **increases the odds of no improvement**
(harm). This is the direct binary analogue of an HR above the threshold on the
survival path, so the entire multimethod workflow — consistency, DINA, and GRF —
runs on this one scale.
:::

## Trial Background

ACTG175 (Hammer et al. 1996) randomized HIV-positive patients with CD4
counts of 200–500 per mm³ to four treatment arms. This analysis uses the
two pure-monotherapy/combination arms:

- **Arm 1**: ZDV + ddI combination therapy (treatment, T = 1)
- **Arm 3**: ddI monotherapy (control, T = 0)

The outcome is the change in CD4 count from baseline to week 20 (`cd40` →
`cd420`), dichotomized into the improvement indicator.


# Setup

## Load Required Packages

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(speff2trial)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(policytree)
library(doFuture)
library(future)

# Optional packages for enhanced output
library(patchwork)
library(forestploter)

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```

## Confounders and Variable Roles

```{r analysis-roles}
#| code-fold: false

# ─── Reproducibility ────────────────────────────────────────────────────
analysis_seed <- 8316951L
set.seed(analysis_seed)

# ─── Confounders ──────────────────────────────────────────────────────
cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
bin_vars  <- c("hemo", "homo", "drugs", "race", "gender", "symptom")
confounders.name <- c(cont_vars, bin_vars)
id.name   <- "id"

# ─── Variables (single adverse scale) ──────────────────────────────────
treat.name   <- "treat"
outcome.name <- "y_binary_neg"   # FS adverse-scale outcome (1 = no improvement)

# ─── ForestSearch GLM parameters (binary, OR scale) ────────────────────
fs_outcome_type   <- "binary"
fs_effect_measure <- "OR"
```


# Data: ACTG175 Trial

## Treatment Assignment

```{r data-treat}
# ACTG175: arm 1 = ZDV+ddI (experimental), arm 3 = ddI (control)
actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$id <- seq_len(nrow(actg_df))

# Treatment: ZDV+ddI = 1, ddI = 0 (no switching)
actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)
```

## Outcome Construction

```{r data-outcomes}
# Drop subjects missing the week-20 CD4 measurement
actg_df <- actg_df[!is.na(actg_df$cd420), ]

# Beneficial scale: CD4 improvement indicator (1 = improved, higher = better)
actg_df$y_binary_pos <- as.integer(actg_df$cd420 > actg_df$cd40)

# Adverse scale: NO improvement indicator (1 = did not improve, higher = worse).
# This is the single analysis scale for the entire multimethod workflow.
actg_df$y_binary_neg <- 1L - actg_df$y_binary_pos

cat("Sample size:", nrow(actg_df), "\n")
cat("Adverse events (no improvement):", sum(actg_df$y_binary_neg),
    sprintf("(%.1f%%)\n", 100 * mean(actg_df$y_binary_neg)))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```

## Covariates

```{r data-covariates}
# Keep ALL candidate covariates numeric.  forestsearch() builds and evaluates
# "var <= value" candidate cuts, and its DINA paths require numeric columns, so
# binary indicators left as factors trigger hundreds of "'<=' not meaningful for
# factors" warnings during the consistency search and "covariate must be numeric"
# failures in DINA (both subgroup_method = "dina" selection and use_dina
# screening).  With 0/1 numeric coding, forestsearch's cont.cutoff (= 4 unique
# values) treats the binary candidates as categorical automatically -- no factor
# conversion is needed, and every engine (consistency, DINA, GRF, standalone
# DINA) receives numeric covariates.
for (v in bin_vars) actg_df[[v]] <- as.numeric(actg_df[[v]])

cat("Continuous candidates:", paste(cont_vars, collapse = ", "), "\n")
cat("Binary candidates:    ", paste(bin_vars, collapse = ", "), "\n")
```

## Baseline Characteristics

```{r baseline-table}
create_summary_table(
  data = actg_df,
  treat_var = treat.name,
  table_title = "ACTG175 Baseline Characteristics and Outcome by Treatment Arm",
  vars_continuous = c("y_binary_neg", cont_vars),
  vars_categorical = bin_vars,
  font_size = 12
)
```

## ITT Analysis (Adverse Scale)

```{r itt-analysis}
# Adverse-scale ITT OR (y_binary_neg = no improvement; OR > 1 = treatment
# increases the odds of no improvement = harm).
itt_neg <- glm(y_binary_neg ~ treat, data = actg_df, family = binomial)

beta  <- coef(itt_neg)["treat"]
se    <- summary(itt_neg)$coefficients["treat", "Std. Error"]
z     <- beta / se
p     <- 2 * pnorm(-abs(z))
itt_or    <- exp(beta)
itt_ci_lo <- exp(beta - qnorm(0.975) * se)
itt_ci_hi <- exp(beta + qnorm(0.975) * se)

cat("=== ITT Odds Ratio, adverse scale (y_binary_neg) ===\n")
cat(sprintf("  log-OR (treat): %7.4f  (SE %.4f)\n", beta, se))
cat(sprintf("  OR:             %6.3f  (95%% CI: %.3f, %.3f)\n",
            itt_or, itt_ci_lo, itt_ci_hi))
cat(sprintf("  p-value:        %.4f\n", p))
cat("\nOn the adverse scale, OR > 1 indicates the treatment raises the odds of\n")
cat("no improvement (harm); OR < 1 indicates fewer non-improvers (benefit).\n")
```


# Preliminary Analysis: Generalized Random Forests

Before running ForestSearch, we can use GRF to explore potential treatment
effect heterogeneity and identify candidate factors. GRF runs here on the same
adverse scale (`y_binary_neg`) with `adverse_outcome = TRUE`, so its
risk-difference (RD) CATE is aligned with the harm direction used throughout.

```{r grf-analysis}
t0 <- proc.time()

grf_est <- grf.subg.harm.glm(
  data             = actg_df,
  confounders.name = confounders.name,
  outcome.name     = outcome.name,
  treat.name       = treat.name,
  id.name          = id.name,
  outcome_type     = fs_outcome_type,
  effect_measure   = fs_effect_measure,
  adverse_outcome  = TRUE,
  dmin.grf         = 0.05,     # RD-scale exploration floor (5 percentage points)
  maxdepth         = 2L,
  RCT              = TRUE,
  sg.criterion     = "mDiff",
  details          = TRUE
)

timings$grf <- (proc.time() - t0)["elapsed"]
```

```{r grf-trees}
#| fig-width: 10
#| fig-height: 4
#| fig-cap: "Policy trees from GRF (depth 1 and 2), adverse scale"

# leaf1 = recommend control, leaf2 = recommend treatment
if (!is.null(grf_est$trees)) {
  oldpar <- par(mfrow = c(1, 2))
  if (length(grf_est$trees) >= 1 && !is.null(grf_est$trees[[1]])) {
    plot(grf_est$trees[[1]], leaf.labels = c("Control", "Treat"), main = "Depth 1")
  }
  if (length(grf_est$trees) >= 2 && !is.null(grf_est$trees[[2]])) {
    plot(grf_est$trees[[2]], leaf.labels = c("Control", "Treat"), main = "Depth 2")
  }
  par(oldpar)
} else {
  cat("GRF: no policy trees available.\n")
}
```

GRF surfaces candidate factors driving treatment-effect heterogeneity on the
adverse scale; ForestSearch then evaluates subgroup candidates with consistency
filtering.


# ForestSearch Analysis

## Running ForestSearch

The selection engine for this main analysis is set by `fs_subgroup_method` in
the setup chunk (`"consistency"`, `"dina"`, or `"grf"` with `fs_grf_selection`).
Everything downstream — the full bootstrap, the Tier-2 de-biased gate, the
cross-validation/leave-one-out diagnostics, and the subgroup outcome plot — runs
off the single fitted object and therefore inherits that engine automatically,
so the entire analysis re-renders under a different method by changing one knob.
The consistency search (the default below) performs an exhaustive search over
candidate subgroup combinations with up to `maxk` factors; the parameters in the
table apply to that engine.

| Parameter | Value | Description |
|-----------|-------|-------------|
| `effect_measure` | `OR` | Odds ratio (binary) |
| `effect.threshold` | 1.25 | Screening OR threshold (adverse scale) |
| `consistency.threshold` | 1.0 | Per-split consistency OR threshold |
| `pconsistency.threshold` | 0.90 | Required consistency rate across splits |
| `maxk` | 2 | Maximum factors per subgroup |
| `n.min` | 60 | Minimum subgroup size |
| `d0.min`, `d1.min` | 10 | Minimum sample size per arm |
| `consistency_method` | `resample` | Required for the Tier-2 gate (GLM path) |

::: {.callout-note collapse="true"}
**API note: GLM-natural vocabulary.** This template uses the GLM-natural
parameter names: `sg_focus = "effMaxSG"` (alias for the legacy `"hrMaxSG"`),
`effect.threshold` (alias for `hr.threshold`), and `consistency.threshold`
(alias for `hr.consistency`). The `"eff*"` and `effect.*` forms read more
naturally for GLM outcomes (continuous MD, binary OR/RR/RD, count IRR) where
there is no hazard ratio. The legacy `"hr*"` and `hr.*` forms continue to work
and produce identical results. See `NEWS.md` for the full list of aliases.
:::

::: {.callout-note}
## Selection criteria (`sg_focus`)

`sg_focus` controls which candidate subgroup is chosen once consistency has been
evaluated. The example below uses `"effMaxSG"` with `selection_rule = "both"`:
the rule selects from the **intersection** of the effect-size neighborhood
(candidates within `effect_neighborhood` of the maximum effect) and the Pareto
non-dominated set on (effect, N), then picks the largest-N candidate. For OR
(ratio measure), the neighborhood test is applied on the natural OR scale.
:::

```{r forestsearch-main}
#| fig-width: 10
#| fig-height: 7

t0 <- proc.time()

fs <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis               = actg_df,
  confounders.name          = confounders.name,
  outcome.name              = outcome.name,
  event.name                = outcome.name,    # binary: outcome IS the event
  treat.name                = treat.name,
  id.name                   = id.name,
  is.RCT                    = TRUE,
  seedit                    = analysis_seed,
  # --- GLM parameters -----------------------------------------------------
  outcome_type              = fs_outcome_type,
  effect_measure            = fs_effect_measure,
  adverse_outcome           = TRUE,
  # --- Screening engines (candidate generation) ---------------------------
  use_lasso                 = TRUE,
  use_grf                   = TRUE,
  grf_depth                 = fs_grf_depth,
  dmin.grf                  = fs_dmin_grf,
  frac.tau                  = 0.80,
  vi.grf.min                = -0.2,
  return_selected_cuts_only = TRUE,
  tune_grf                  = FALSE,
  use_dina                  = FALSE,
  dina_args                 = fs_dina_args,
  # --- Candidate-cut construction -----------------------------------------
  cut_type                  = "default",
  cont.cutoff               = 4L,
  conf.cont_jcuts           = list(cd40 = 10, wtkg = 10),
  # --- Selection engine ---------------------------------------------------
  subgroup_method           = fs_subgroup_method,        # consistency | dina | grf
  grf_selection             = fs_grf_selection,          # used when subgroup_method = "grf"
  grf_select_statistic      = fs_grf_select_statistic,   # frontier only
  dina_select_statistic     = fs_dina_select_statistic,  # used when subgroup_method = "dina"
  sg_focus                  = fs_sg_focus,
  selection_rule            = fs_selection_rule,
  effect_neighborhood       = 0.10,
  # --- Tier-2 de-biased gate ----------------------------------------------
  mr_inference               = run_mr,
  mr_inference_args          = list(draws = mr_draws,
                                   family_native_neighborhood = fs_family_native_nbhd),
  # --- Subgroup constraints -----------------------------------------------
  n.min                     = 60L,
  d0.min                    = 10L,
  d1.min                    = 10L,
  maxk                      = 2L,
  max_subgroups_search      = fs_max_subgroups_search,
  # --- Thresholds (OR scale) ----------------------------------------------
  effect.threshold          = fs_or_threshold,
  consistency.threshold     = fs_or_consistency,
  pconsistency.threshold    = fs_pconsistency,
  consistency_method        = fs_consistency_method,   # "resample" (gate requires it)
  # --- Consistency search -------------------------------------------------
  fs.splits                 = 500L,
  use_twostage              = TRUE,
  # --- Output / diagnostics -----------------------------------------------
  show_candidate_summary    = TRUE,
  details                   = TRUE,
  quiet                     = FALSE,
  plot.sg                   = FALSE,   # KM not applicable; use plot_sg_glm_outcomes()
  # --- Parallel -----------------------------------------------------------
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential"); gc(verbose = FALSE)
timings$forestsearch <- (proc.time() - t0)["elapsed"]

cat("\nForestSearch completed in",
    round(timings$forestsearch, 1), "seconds\n")
```


## ForestSearch Results

### Identified Subgroups

```{r fs-results}
# Generate results tables
res_tabs <- sg_tables(fs, ndecimals = display_digits, which_df = "est")

# Display top subgroups meeting criteria
res_tabs$sg10_out
```

### Treatment Effect Estimates

```{r fs-estimates}
# ITT and subgroup estimates
res_tabs$tab_estimates
```

### Identified Subgroup Definition

```{r fs-subgroup}
if (!is.null(fs$sg.harm) && length(fs$sg.harm) > 0) {
  cat("Identified subgroup (H):", fmt_sg(fs$sg.harm), "\n")
  cat("Subgroup size:", sum(fs$df.est$treat.recommend == 0),
      sprintf("(%.1f%% of ITT)\n",
              100 * mean(fs$df.est$treat.recommend == 0)))
} else {
  cat("ForestSearch: no subgroup identified at the configured thresholds.\n")
}
```

### Selection Diagnostic: Pareto Frontier

Beyond the single subgroup chosen by `sg_focus`, it is often informative to
inspect the *Pareto frontier* on (effect, N) — the set of candidate subgroups
that are not dominated on both axes simultaneously. A candidate is dominated
when another candidate has at least as large an effect *and* at least as large a
sample size, with at least one strict inequality. Under `effMaxSG` with
`selection_rule = "both"` the selected subgroup is typically a frontier member
that also sits in the effect-size band; the frontier serves both as the
selection geometry and as a diagnostic of the trade-off space.

For OR outcomes the effect axis is auto-rendered on log scale by
`plot_pareto_frontier()` and `pareto_frontier_table()` (auto-detected from
`fs$effect_measure`); the rendered values are on the natural OR scale.

```{r fs-pareto}
#| fig-width: 9
#| fig-height: 6

# 1. Compute the split-derived CIs once (the expensive step).
ci_tab <- compute_frontier_cis(
  fs,
  n_splits = frontier_params$n_splits,
  seed     = frontier_params$seed
)

# 2. Plot, with split-derived CI bars overlaid; same intervals as the table.
p_pareto <- plot_pareto_frontier(
  fs,
  ci_table  = ci_tab,
  show_band = TRUE,
  xlim      = frontier_params$xlim
)
print(p_pareto)

# 3. Render the table.  Selected subgroup is highlighted and marked with ★.
pareto_frontier_table(
  fs,
  ci_table = ci_tab,
  digits   = frontier_params$digits
)
```

### Selection Narrative

`explain_pareto_selection()` describes the chosen subgroup relative to the
non-dominated alternatives on the frontier.

```{r fs-pareto-explain, results = "asis"}
cat(explain_pareto_selection(fs, ci_table = ci_tab))
```

### Agreement with GRF

ForestSearch and GRF identify subgroups by different mechanisms — FS via
exhaustive search with consistency filtering, GRF via causal forests plus policy
trees — so agreement on subject-level membership provides an informal robustness
check. `compare_fs_subgroups()` produces a publication-ready cross-tabulation
plus Cohen's kappa; it accepts forestsearch and GRF result objects directly, and
handles non-detection (method identifies no subgroup) by assigning all subjects
to the complement.

```{r fs-grf-concordance}
sg_compare <- compare_fs_subgroups(
  fs1               = fs,
  fs2               = grf_est,
  label1            = "ForestSearch",
  label2            = "GRF Standalone",
  id.name           = id.name,
  subgroup_notation = "harm",
  title             = "FS vs GRF: Subgroup Membership Concordance",
  subtitle          = "ACTG175 observed data (binary OR, adverse scale)",
  font_size         = 13
)
sg_compare$table
```

```{r fs-grf-concordance-stats}
cat("Concordance statistics:\n")
cat(sprintf("  Agreement:       %.1f%%\n",
    100 * sg_compare$concordance$agreement))
if (!is.null(sg_compare$concordance$kappa)) {
  cat(sprintf("  Cohen's kappa:   %.3f\n", sg_compare$concordance$kappa))
}
cat(sprintf("  Concordant:      %d / Discordant: %d / Total: %d\n",
    sg_compare$concordance$n_agree,
    sg_compare$concordance$n_disagree,
    sg_compare$concordance$n_total))
```


# Bootstrap Bias Correction

## Rationale

GLM estimates from identified subgroups are **upwardly biased** due to the
selection process (subgroups are selected *because* they show extreme effects).
Bootstrap bias correction addresses this by:

1. Resampling with replacement
2. Re-running the entire ForestSearch algorithm
3. Computing bias terms from bootstrap vs. observed estimates
4. Applying infinitesimal jackknife variance estimation

## Running Bootstrap Analysis

```{r bootstrap}
t0 <- proc.time()

fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est = fs,
  nb_boots = NB,
  show_three = FALSE,
  details = FALSE,
  digits = construction_digits,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential"); gc(verbose = FALSE)
timings$bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nBootstrap completed in",
    round(timings$bootstrap / 60, 1), "minutes\n")
```

## Bootstrap Summary and Diagnostics

```{r bootstrap-summary}
# The OR column (with bias-corrected OR*) is auto-detected from FSsg_tab; the
# est.scale argument only affects the legacy HR-flavored caption helper.
summaries <- summarize_bootstrap_results(
  sgharm = fs$sg.harm,
  boot_results = fs_bc,
  create_plots = TRUE,
  est.scale = "hr",
  digits = display_digits
)

summaries$table
```

## Fast de-biased gate vs. full bootstrap

The Tier-2 de-biased gate is a fast multiplier-bootstrap approximation of the
bootstrap bias-correction. It reuses per-candidate GLM influence (`dfbeta`) over
the selection engine's **own** candidate family — the enumerated consistency
candidates, DINA's qualifying cuts, or GRF's DR-score frontier, depending on
`fs_subgroup_method` — and re-selects under multiplier draws, so its
selection-bias term mirrors the same search optimism the full bootstrap
corrects — without re-running the search per resample. The table below compares
the gate (computed inside the fit) against the full bootstrap from the previous
section. The naive-OR row is a scale sanity check; agreement on the de-biased
row is the Tier-2-vs-Tier-1 validation. By default the de-biased CI uses the same
infinitesimal-jackknife variance as the full bootstrap (the `ci_method = 'ij'`
default), so the de-biased row compares both point estimate and interval
like-for-like. The gate de-biases both the selected subgroup and its complement,
so both appear in the comparison and the table below.

```{r debias-gate-comparison, eval = !is.null(fs$sg.harm)}
#| echo: false
#| results: asis
g <- fs$mr_inference
if (is.null(g)) {
  cat("_The de-biased gate did not run (set `run_mr <- TRUE`, and note",
      "the GLM path requires `consistency_method = \"resample\"`)._\n")
} else {
  fmt_ci <- function(e, l, u) sprintf("%.3f (%.3f, %.3f)", e, l, u)
  fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0  <- fl(fs_bc$H_estimates, "H0"); H0l <- fl(fs_bc$H_estimates, "H0_lower"); H0u <- fl(fs_bc$H_estimates, "H0_upper")
  H2  <- fl(fs_bc$H_estimates, "H2"); H2l <- fl(fs_bc$H_estimates, "H2_lower"); H2u <- fl(fs_bc$H_estimates, "H2_upper")

  # Complement: Tier-2 from g$complement (include_complement = TRUE), Tier-1 from
  # the bootstrap's Hc_estimates.
  cc  <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  Hc0 <- fl(fs_bc$Hc_estimates, "H0"); Hc0l <- fl(fs_bc$Hc_estimates, "H0_lower"); Hc0u <- fl(fs_bc$Hc_estimates, "H0_upper")
  Hc2 <- fl(fs_bc$Hc_estimates, "H2"); Hc2l <- fl(fs_bc$Hc_estimates, "H2_lower"); Hc2u <- fl(fs_bc$Hc_estimates, "H2_upper")

  q  <- c("naive OR (95% CI)", "de-biased OR (95% CI)")
  t2 <- c(fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  t1 <- c(fmt_ci(H0, H0l, H0u), fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    q  <- c(q, "complement naive OR (95% CI)", "complement de-biased OR (95% CI)")
    t2 <- c(t2, fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    t1 <- c(t1, fmt_ci(Hc0, Hc0l, Hc0u), fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `Tier 2 (gate)` = t2,
                     `Tier 1 (bootstrap)` = t1, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE,
        caption = sprintf("Selected subgroup: %s%s", fmt_sg(fs$sg.harm),
                          if (has_cc) " (complement = remaining subjects)" else "")),
      sep = "\n"); cat("\n\n")

  # Scale cross-check: the gate's naive OR and the bootstrap's naive H0 are the
  # same quantity (raw subgroup OR) computed two ways; they should agree.
  if (is.finite(H0) && is.finite(g$naive$est)) {
    scale_ok <- abs(log(g$naive$est) - log(H0)) < 1e-2
    cat(sprintf("Scale cross-check (naive gate vs bootstrap H0): **%s** (%.3f vs %.3f).\n\n",
                if (scale_ok) "PASS" else "CHECK", g$naive$est, H0))
  }

  # Family is the full enumerated space, not the screened survivors.
  n_surv <- tryCatch(nrow(fs$find.grps$out.found$hr.subgroups), error = function(e) NA_integer_)
  cat(sprintf("Candidate family: %d (full <= maxk space)%s.\n\n", g$n_family,
              if (!is.na(n_surv)) sprintf("; screened survivors = %d", n_surv) else ""))

  cat(sprintf("Re-selection rule = `%s`%s.\n\n",
              if (!is.null(g$settings$reselection)) g$settings$reselection else "?",
              " (matches the consistency search statistic)"))

  boot_s <- as.numeric(timings$bootstrap)
  cat(sprintf("Timing: Tier 1 bootstrap (%d boots) %.1f s (%.1f min); Tier 2 gate %.2f s",
              NB, boot_s, boot_s / 60, g$timing_seconds))
  if (is.finite(boot_s) && g$timing_seconds > 0)
    cat(sprintf(" -> **%.0fx** faster (multiplier loop only; the full bootstrap scales with `NB`, the gate does not)", boot_s / g$timing_seconds))
  cat(".\n\n")

  cat(sprintf("Gate `mr_harm_confirmed` = **%s** (de-biased OR %.3f %s gate %.2f); selection bias %.4f (log-OR).\n",
              fs$mr_harm_confirmed, g$debiased$est,
              if (g$settings$confirm_rule == "point") "vs" else "lower-bound vs",
              g$settings$t_confirm, g$selection_bias))
  if (!is.null(g$ci_method) && !is.null(g$debiased$se_ij))
    cat(sprintf("De-biased CI: %s (IJ SE %.3f vs robust SE %.3f, log-OR).\n",
                if (identical(g$ci_method, "ij")) "infinitesimal jackknife (Tier-1 analogue)" else "robust SE",
                g$debiased$se_ij, g$debiased$se_wald))
  if (has_cc)
    cat(sprintf("Complement (n = %d): naive OR %.3f -> de-biased %.3f (selection bias %.4f, log-OR).\n",
                cc$n, cc$naive$est, cc$debiased$est, cc$selection_bias))
}
```

```{r debias-gate-no-subgroup, eval = is.null(fs$sg.harm)}
#| echo: false
#| results: asis
cat("_No harm subgroup was selected, so the de-biased gate comparison does not apply._\n")
```

The same bias-corrected estimates table, rebuilt from the Tier-2 gate object. The
descriptive columns and naive effect are tier-independent (inherited from the
bootstrap result used as a template); the bias-corrected column carries the gate's
de-biased estimate for **both** the selected subgroup and its complement.

```{r debias-gate-table, eval = !is.null(fs$sg.harm) && !is.null(fs$mr_inference)}
mr_estimates_table(fs, boot_results = fs_bc, est.scale = "hr", digits = 2)
```

## Subgroup Outcome Visualization

```{r fs-glm-plot}
#| fig-width: 10
#| fig-height: 7
#| fig-cap: "Rate of no improvement by treatment arm and identified subgroup (FS, adverse scale)"

fs_plot <- plot_sg_glm_outcomes(
  fs.est         = fs,
  outcome.name   = outcome.name,
  treat.name     = treat.name,
  effect_measure = fs_effect_measure,
  outcome_type   = fs_outcome_type,
  E.name         = "ZDV+ddI",
  C.name         = "ddI",
  show_itt       = TRUE,
  show_effect    = TRUE,
  verbose        = TRUE
)
```

### Bootstrap Diagnostics

```{r bootstrap-diagnostics}
summaries$diagnostics_table_gt
```

### Subgroup Agreement

How consistently does bootstrap identify the same subgroup?

```{r subgroup-agreement}
if (!is.null(summaries$subgroup_summary$original_agreement)) {
  summaries$subgroup_summary$original_agreement
}

if (!is.null(summaries$subgroup_summary$factor_presence)) {
  summaries$subgroup_summary$factor_presence
}
```

### Bootstrap Distributions

```{r bootstrap-plots}
#| fig-width: 10
#| fig-height: 4
#| fig-cap: "Bootstrap distributions of OR for H (questionable) and Hc (recommend)"

if (!is.null(summaries$plots)) {
  summaries$plots$H_distribution + summaries$plots$Hc_distribution
}
```


# DINA-Based Subgroup Identification

DINA (Doubly-robust Identification of Non-Additive effects) screens and selects
subgroups from an influence-function representation of the covariate-specific
treatment effect, fit with cross-fitting. For a binary endpoint the DINA family
is `"binomial"` and the effect is reported on the OR scale. It plugs into the
ForestSearch workflow in three complementary ways, all configured to mirror the
main analysis above (same confounders, harm threshold, minimum size `n.min = 60`,
and `sg_focus`):

* **(A)** ForestSearch *using DINA for screening* — DINA's per-covariate frontier
  contributes candidate cuts to the screening pool; the consistency search still
  performs selection.
* **(B)** *Standalone DINA* with parameters analogous to (A) — the same fit's
  frontier (the candidates (A) screens) and its single selected subgroup via
  `dina_subgroup()`.
* **(C)** ForestSearch with `subgroup_method = "dina"` — subgroup *selection* is
  delegated to `dina_subgroup()`, then the standard ForestSearch bootstrap
  provides the selection bias-corrected OR estimate.

```{r dina-config}
#| code-fold: false

# Shared DINA *fit* settings (binary outcome => family mapped to "binomial").
# Selection parameters (sg_focus, or_threshold) live in the setup chunk under
# the fs_dina_* knobs so they can be tuned independently; (B) and (C) read them.
dina_seed   <- 8316951L
dina_nfolds <- 2L
# Harm/benefit floor on the natural-parameter (log-OR) link scale for the
# binomial family.
m_diff_dina <- log(fs_dina_or_threshold)
```

## (A) ForestSearch Using DINA for Screening

Here ForestSearch is run as in the main analysis, but with `use_dina = TRUE`:
DINA's frontier cuts are merged into the candidate pool alongside the GRF and
LASSO cuts. `dina_args` is a flat, fully-named list; in the screening role it
supplies the DINA *fit* tuning (`seed`, `n_folds`) and the *frontier* controls
(`scope`, `max_per_covariate`, `max_subgroups`). `scope = "wide"` casts a
deliberately wide net — the consistency stage remains the gatekeeper. By default
`dina_args$selected_only = TRUE`, so screening contributes the single cut
`dina_subgroup()` would select — using this call's `sg_focus` / `selection_rule`
/ `effect_neighborhood` / `effect.threshold`, i.e. the same cut
`subgroup_method = "dina"` (C) chooses. Set `dina_args$selected_only = FALSE` to
contribute the full frontier instead.

```{r fs-dina-screen}
t0 <- proc.time()

fs_dina_screen <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis               = actg_df,
  confounders.name          = confounders.name,
  outcome.name              = outcome.name,
  event.name                = outcome.name,
  treat.name                = treat.name,
  id.name                   = id.name,
  is.RCT                    = TRUE,
  seedit                    = analysis_seed,
  # --- GLM parameters -----------------------------------------------------
  outcome_type              = fs_outcome_type,
  effect_measure            = fs_effect_measure,
  adverse_outcome           = TRUE,
  # --- Screening engines: GRF + LASSO + DINA ------------------------------
  use_lasso                 = TRUE,
  use_grf                   = TRUE,
  grf_depth                 = 2L,
  dmin.grf                  = 0,
  frac.tau                  = 0.80,
  vi.grf.min                = -0.2,
  return_selected_cuts_only = TRUE,
  tune_grf                  = FALSE,
  use_dina                  = TRUE,
  dina_args = list(
    seed = dina_seed, n_folds = dina_nfolds,
    scope = "wide", max_per_covariate = 3L, max_subgroups = 10L,
    selected_only = dina_screen_selected_only
  ),
  # --- Candidate-cut construction -----------------------------------------
  cut_type                  = "default",
  cont.cutoff               = 4L,
  conf.cont_jcuts           = list(cd40 = 10, wtkg = 10),
  # --- Selection engine ---------------------------------------------------
  subgroup_method           = "consistency",
  sg_focus                  = dina_screen_sg_focus,
  selection_rule            = "neighborhood",
  effect_neighborhood       = 0.10,
  # --- Subgroup constraints -----------------------------------------------
  n.min                     = 60L,
  d0.min                    = 10L,
  d1.min                    = 10L,
  maxk                      = 2L,
  max_subgroups_search      = dina_screen_max_subgroups_search,
  # --- Thresholds (OR scale) ----------------------------------------------
  effect.threshold          = dina_screen_or_threshold,
  consistency.threshold     = dina_screen_or_consistency,
  pconsistency.threshold    = fs_pconsistency,
  consistency_method        = fs_consistency_method,
  # --- Consistency search -------------------------------------------------
  fs.splits                 = 500L,
  use_twostage              = TRUE,
  # --- Output / diagnostics -----------------------------------------------
  show_candidate_summary    = TRUE,
  details                   = TRUE,
  quiet                     = FALSE,
  plot.sg                   = FALSE,
  # --- Parallel -----------------------------------------------------------
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential"); gc(verbose = FALSE)
timings$fs_dina_screen <- (proc.time() - t0)["elapsed"]

if (!is.null(fs_dina_screen$sg.harm) && length(fs_dina_screen$sg.harm) > 0) {
  cat("Identified subgroup (H):",
      fmt_sg(fs_dina_screen$sg.harm), "\n")
  cat("Subgroup size:", sum(fs_dina_screen$df.est$treat.recommend == 0),
      sprintf("(%.1f%% of ITT)\n",
              100 * mean(fs_dina_screen$df.est$treat.recommend == 0)))
} else {
  cat("DINA-screened ForestSearch: no subgroup identified.\n")
}
```

The cuts DINA contributed to the candidate pool are recorded on the result
object. These are exactly the frontier candidates reproduced standalone in (B).

```{r fs-dina-screen-cuts}
dina_screen_cuts <- fs_dina_screen$dina_cuts
if (length(dina_screen_cuts) > 0) {
  gt(data.frame(
    `#` = seq_along(dina_screen_cuts),
    `DINA candidate cut` = dina_screen_cuts,
    check.names = FALSE
  )) |>
    tab_header(title = "DINA Screening Candidates Injected into ForestSearch") |>
    cols_align(align = "left")
} else {
  cat("DINA contributed no candidate cuts under the configured frontier.\n")
}
```

```{r fs-dina-screen-pareto, eval = !is.null(fs_dina_screen$sg.harm)}
# Selection diagnostic for the DINA-screened run (selected subgroup marked).
pareto_frontier_table(fs_dina_screen)
```

## (B) Standalone DINA

The same DINA fit, run outside ForestSearch. Its `dina_frontier()` output is the
candidate set fed into screening in (A); its `dina_subgroup()` output is the
single subgroup selected by DINA itself — the mechanism (C) uses internally.
Parameters are analogous to (A): same confounders, harm floor
`m_diff = log(fs_dina_or_threshold)`, `n_min = 60`, and `sg_focus`.

```{r dina-standalone-fit}
t0 <- proc.time()

# Standalone DINA -- dina(), dina_frontier(), dina_subgroup() -- on the numeric
# covariate frame (all candidates are numeric per `data-covariates`).
dina_fit <- dina(
  df         = actg_df,
  outcome    = outcome.name,
  treatment  = treat.name,
  covariates = confounders.name,
  family     = "binomial",     # binary outcome (no status column required)
  # ... forwarded to dina_fit():
  seed              = dina_seed,   # default NULL
  n_folds           = dina_nfolds, # default 2L
  propensity_method = "logistic",  # default "logistic"
  cross_fitting     = TRUE         # default TRUE
)

timings$dina_standalone <- (proc.time() - t0)["elapsed"]
cat("Standalone DINA fit complete (family = binomial).\n")
```

### DINA Frontier (Screening Candidates)

```{r dina-standalone-frontier}
dina_fr <- dina_frontier(
  fit               = dina_fit,
  df                = actg_df,
  covariates        = confounders.name,
  scope             = "wide",     # default "wide"; or "harm" (requires m_diff)
  m_diff            = NULL,        # default NULL (no effect floor under scope = "wide")
  n_min             = 60L,        # default 60L
  direction         = "both",     # default "both"; or "left" / "right"
  max_per_covariate = 3L,         # default 3L
  max_subgroups     = 10L,        # default 10L
  digits            = 3L          # default 3L
)

gt(dina_fr[, c("covariate", "direction", "threshold",
               "n_subgroup", "effect", "cut_expr")]) |>
  tab_header(
    title = "Standalone DINA Frontier",
    subtitle = "Per-covariate non-dominated (effect, N) candidates; effect on OR scale"
  ) |>
  fmt_number(columns = c(threshold, effect), decimals = 3) |>
  cols_label(
    covariate = "Covariate", direction = "Direction",
    threshold = "Threshold", n_subgroup = "N", effect = "OR",
    cut_expr = "Cut"
  ) |>
  cols_align(align = "left", columns = c(covariate, direction, cut_expr))
```

### DINA-Selected Subgroup

```{r dina-standalone-select}
dina_sel <- dina_subgroup(
  fit                 = dina_fit,
  df                  = actg_df,
  covariates          = confounders.name,
  m_diff              = m_diff_dina,
  n_min               = 60L,            # default 60L
  direction           = "both",         # default "both"; or "left" / "right"
  sg_focus            = fs_dina_sg_focus,
  selection_rule      = "neighborhood", # default "neighborhood"; pareto/both
  effect_neighborhood = 0.10,           # default 0.10
  alpha               = 0.05            # default 0.05
)

if (isTRUE(dina_sel$found)) {
  op <- if (identical(dina_sel$direction, "left")) "<=" else ">="
  cat(sprintf("DINA-selected subgroup: %s %s %s\n",
              dina_sel$covariate, op, format(dina_sel$threshold)))
  cat(sprintf("Direction: %s | N = %d\n",
              dina_sel$direction, dina_sel$n_subgroup))
} else {
  cat("Standalone DINA selected no subgroup meeting the criteria.\n")
}
```

## (C) ForestSearch with `subgroup_method = "dina"`

In this mode ForestSearch delegates subgroup *selection* to `dina_subgroup()`,
bypassing GRF, LASSO and the consistency search. The selection criteria are
inherited from this call's `sg_focus`, `n.min` and `effect.threshold` (mapped to
the log-OR harm floor); `dina_args` supplies only the fit tuning. The returned
object is a standard `forestsearch` result, so the existing bootstrap machinery
applies unchanged.

```{r fs-dina-select}
t0 <- proc.time()

fs_dina <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis         = actg_df,
  confounders.name    = confounders.name,
  outcome.name        = outcome.name,
  event.name          = outcome.name,
  treat.name          = treat.name,
  id.name             = id.name,
  is.RCT              = TRUE,
  seedit              = analysis_seed,
  # --- GLM parameters -----------------------------------------------------
  outcome_type        = fs_outcome_type,
  effect_measure      = fs_effect_measure,
  adverse_outcome     = TRUE,
  # --- DINA-selection mode ------------------------------------------------
  subgroup_method     = "dina",
  dina_args           = list(seed = dina_seed, n_folds = dina_nfolds),  # fit tuning only
  dina_select_statistic = fs_dina_select_statistic,
  # --- Selection criteria (passed through to dina_subgroup) ---------------
  sg_focus            = fs_dina_sg_focus,
  selection_rule      = "neighborhood",
  effect_neighborhood = 0.10,
  n.min               = 60L,
  effect.threshold    = fs_dina_or_threshold,  # -> log-OR harm floor
  # --- Tier-2 de-biased gate ----------------------------------------------
  mr_inference         = run_mr,
  mr_inference_args    = list(draws = mr_draws,
                             family_native_neighborhood = fs_dina_family_native_nbhd),
  # --- Output -------------------------------------------------------------
  details             = TRUE
  # NOTE: subgroup_method = "dina" bypasses GRF, LASSO and the consistency
  # search; use_grf / use_lasso / use_twostage / fs.splits / maxk and related
  # parameters do not apply in this mode.
)

timings$fs_dina_select <- (proc.time() - t0)["elapsed"]

if (!is.null(fs_dina$sg.harm) && length(fs_dina$sg.harm) > 0) {
  cat("DINA-selected subgroup (H):",
      fmt_sg(fs_dina$sg.harm), "\n")
  cat("Subgroup size:", sum(fs_dina$df.est$treat.recommend == 0),
      sprintf("(%.1f%% of ITT)\n",
              100 * mean(fs_dina$df.est$treat.recommend == 0)))
} else {
  cat("DINA selection: no subgroup met the criteria.\n")
}
```

### Bootstrap Bias-Corrected OR Estimates

The within-subgroup OR is upwardly biased by selection. The ForestSearch
bootstrap corrects this by re-running the full pipeline on each resample;
because the result above was produced in DINA-selection mode, each replicate
re-fits DINA and re-runs `dina_subgroup()` from scratch — so the correction
reflects the DINA selection process itself.

```{r fs-dina-bootstrap, eval = !is.null(fs_dina$sg.harm)}
# Runs only when DINA identified a subgroup.  The bootstrap bias-corrects an
# identified subgroup's effect, so there is nothing to do otherwise.
t0 <- proc.time()

fs_dina_bc <- forestsearch_bootstrap_dofuture(
  fs.est = fs_dina,
  nb_boots = NB,
  show_three = FALSE,
  details = FALSE,
  digits = construction_digits,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential"); gc(verbose = FALSE)
timings$fs_dina_bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nDINA-mode bootstrap completed in",
    round(timings$fs_dina_bootstrap / 60, 1), "minutes\n")
```

```{r fs-dina-no-subgroup, eval = is.null(fs_dina$sg.harm)}
# Shown only when DINA selected no subgroup at the configured harm floor.
timings$fs_dina_bootstrap <- 0
cat("DINA (subgroup_method = \"dina\") identified no subgroup at",
    "effect.threshold =", fs_dina_or_threshold, "-- bootstrap skipped.\n",
    "DINA's model-based effect is attenuated relative to the raw OR,",
    "so this floor may not be reached; see the standalone DINA frontier (B)",
    "for the achievable effect range, and lower the DINA harm floor if a",
    "bootstrap demonstration is desired.\n")
```

```{r fs-dina-bootstrap-summary, eval = !is.null(fs_dina$sg.harm)}
dina_summaries <- summarize_bootstrap_results(
  sgharm = fs_dina$sg.harm,
  boot_results = fs_dina_bc,
  create_plots = FALSE,
  est.scale = "hr",
  digits = display_digits
)

dina_summaries$table
```

#### Tier-2 de-biased gate (counterpart to the Tier-1 bootstrap)

The same Tier-2-vs-Tier-1 comparison as the main analysis, for the
DINA-selection fit. The gate de-biases against DINA's **own** qualifying
candidate family (re-running nothing), while the bootstrap above re-runs DINA on
every resample.

For DINA the two tiers are **not** expected to agree as tightly as for the
consistency method. DINA selects on its model-based **tau-hat**, but the gate can
only re-select on the perturbed **GLM** OR; over a large depth-2 candidate family
the gate's winner's-curse term can therefore exceed the bootstrap's and
over-correct the selected subgroup. To keep the gate's re-selection neighbourhood
close to DINA's own selection band, the gate by default restricts that family to
candidates competitive on DINA's native tau-hat, matching the band to
`effect_neighborhood` (the recommended default, used here via
`fs_dina_family_native_nbhd <- NULL`; set an explicit number for a different band,
or `>= 1` to de-bias against the full unrestricted family). The complement is the
induced residual rather than a max-over-family selection, so it agrees across
tiers regardless. This band tunes only the *candidate-competition* part of the
bias — the gate holds the DINA fit fixed, so the refit optimism is invisible to
it; treat the full bootstrap as the reference for DINA and the gate as a fast,
conservative check.

```{r fs-dina-debias-gate-comparison, eval = !is.null(fs_dina$sg.harm)}
#| echo: false
#| results: asis
g <- fs_dina$mr_inference
if (is.null(g)) {
  cat("_The de-biased gate did not run (set `run_mr <- TRUE`)._\n")
} else {
  fmt_ci <- function(e, l, u) sprintf("%.3f (%.3f, %.3f)", e, l, u)
  fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0 <- fl(fs_dina_bc$H_estimates, "H0"); H0l <- fl(fs_dina_bc$H_estimates, "H0_lower"); H0u <- fl(fs_dina_bc$H_estimates, "H0_upper")
  H2 <- fl(fs_dina_bc$H_estimates, "H2"); H2l <- fl(fs_dina_bc$H_estimates, "H2_lower"); H2u <- fl(fs_dina_bc$H_estimates, "H2_upper")
  cc <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  q  <- c("naive OR (95% CI)", "de-biased OR (95% CI)")
  t2 <- c(fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  t1 <- c(fmt_ci(H0, H0l, H0u), fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    Hc0 <- fl(fs_dina_bc$Hc_estimates, "H0"); Hc0l <- fl(fs_dina_bc$Hc_estimates, "H0_lower"); Hc0u <- fl(fs_dina_bc$Hc_estimates, "H0_upper")
    Hc2 <- fl(fs_dina_bc$Hc_estimates, "H2"); Hc2l <- fl(fs_dina_bc$Hc_estimates, "H2_lower"); Hc2u <- fl(fs_dina_bc$Hc_estimates, "H2_upper")
    q  <- c(q, "complement naive OR (95% CI)", "complement de-biased OR (95% CI)")
    t2 <- c(t2, fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    t1 <- c(t1, fmt_ci(Hc0, Hc0l, Hc0u), fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `Tier 2 (gate)` = t2,
                     `Tier 1 (bootstrap)` = t1, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE,
        caption = sprintf("DINA-selected subgroup: %s%s",
                          fmt_sg(fs_dina$sg.harm),
                          if (has_cc) " (complement = remaining subjects)" else "")),
      sep = "\n"); cat("\n\n")

  boot_s <- as.numeric(timings$fs_dina_bootstrap)
  cat(sprintf("Timing: Tier 1 bootstrap (%d boots) %.1f s (%.1f min); Tier 2 gate %.2f s",
              NB, boot_s, boot_s / 60, g$timing_seconds))
  if (is.finite(boot_s) && g$timing_seconds > 0)
    cat(sprintf(" -> **%.0fx** faster (the bootstrap scales with `NB`, the gate does not)",
                boot_s / g$timing_seconds))
  cat(".\n\n")
  cat(sprintf("Gate harm flag = **%s** (de-biased OR %.3f %s gate %.2f); selection bias %.4f (log-OR); candidate family = %d.\n",
              isTRUE(g$harm_flag), g$debiased$est,
              if (g$settings$confirm_rule == "point") "vs" else "lower-bound vs",
              g$settings$t_confirm, g$selection_bias, g$n_family))
  cat(sprintf("Re-selection rule = `%s` (a GLM-OR proxy for DINA's tau-hat); native-family band = %s.\n",
              if (!is.null(g$settings$reselection)) g$settings$reselection else "?",
              if (is.null(fs_dina_family_native_nbhd)) "default (matched to effect_neighborhood)"
              else if (fs_dina_family_native_nbhd >= 1) "off (full family)"
              else sprintf("%.2f (restricted to tau-hat-competitive candidates)",
                           fs_dina_family_native_nbhd)))
  if (!is.null(g$ci_method) && !is.null(g$debiased$se_ij))
    cat(sprintf("De-biased CI: %s (IJ SE %.3f vs robust SE %.3f, log-OR).\n",
                if (identical(g$ci_method, "ij")) "infinitesimal jackknife (Tier-1 analogue)" else "robust SE",
                g$debiased$se_ij, g$debiased$se_wald))
}
```

The Tier-1 bias-corrected table above, rebuilt from the Tier-2 gate object (the
descriptive columns and naive effect are tier-independent; the bias-corrected
column carries the gate's de-biased estimate for the subgroup and its complement).

```{r fs-dina-debias-gate-table, eval = !is.null(fs_dina$sg.harm) && !is.null(fs_dina$mr_inference)}
mr_estimates_table(fs_dina, boot_results = fs_dina_bc, est.scale = "hr", digits = 2)
```

::: {.callout-note}
## Screening vs. selection roles for DINA

`dina_args` behaves differently by role. In the **screening** path (A,
`use_dina = TRUE`) it defaults to `selected_only = TRUE`: the screening path
calls `dina_subgroup()` and contributes only its selected cut, so it *does*
consult this call's `sg_focus` / `selection_rule` / `effect_neighborhood` —
yielding the same cut the selection path (C) would choose. Setting
`selected_only = FALSE` instead drives `dina_frontier()`, which is
selection-agnostic and contributes the full candidate set. In the **selection**
path (C, `subgroup_method = "dina"`) the selection is always driven by this
call's `sg_focus` / `selection_rule` / `effect_neighborhood` via
`dina_subgroup()`, and `dina_args` supplies only the fit. This mode is currently
depth-1 (a single covariate threshold).
:::


## (D) ForestSearch with `subgroup_method = "grf"`

In this mode ForestSearch delegates subgroup *selection* to GRF — a causal-forest
policy tree (`grf_selection = "tree"`) or the doubly-robust-score frontier
(`"frontier"`) — bypassing LASSO, DINA and the consistency search. The selection
criteria are inherited from this call's `sg_focus` (mapped to the GRF
frontier/tree rule), `n.min`, `grf_depth`, `grf_selection` and the DR-score harm
floor `dmin.grf` (on the RD scale for binary). The returned object is a standard
`forestsearch` result, so the existing bootstrap machinery applies unchanged.

```{r fs-grf-select}
t0 <- proc.time()

fs_grf <- forestsearch(
  # --- Required data + roles ----------------------------------------------
  df.analysis         = actg_df,
  confounders.name    = confounders.name,
  outcome.name        = outcome.name,
  event.name          = outcome.name,
  treat.name          = treat.name,
  id.name             = id.name,
  is.RCT              = TRUE,
  seedit              = analysis_seed,   # also seeds the forest
  # --- GLM parameters -----------------------------------------------------
  outcome_type        = fs_outcome_type,
  effect_measure      = fs_effect_measure,
  adverse_outcome     = TRUE,
  # --- GRF-selection mode -------------------------------------------------
  subgroup_method     = "grf",
  grf_selection       = fs_grf_demo_selection,   # "tree" | "frontier"
  grf_select_statistic = fs_grf_select_statistic, # frontier only
  grf_depth           = fs_grf_demo_depth,
  dmin.grf            = fs_grf_demo_dmin,          # DR-score harm floor (RD scale)
  frac.tau            = 0.8,
  # --- Selection criteria -------------------------------------------------
  sg_focus            = fs_grf_sg_focus,          # -> GRF frontier/tree rule
  selection_rule      = "neighborhood",
  effect_neighborhood = 0.10,
  n.min               = 60L,
  # --- Tier-2 de-biased gate ----------------------------------------------
  mr_inference         = run_mr,
  mr_inference_args    = list(draws = mr_draws,
                             family_native_neighborhood = fs_grf_family_native_nbhd),
  # --- Output -------------------------------------------------------------
  details             = TRUE
  # NOTE: subgroup_method = "grf" bypasses LASSO, DINA and the consistency
  # search; use_lasso / use_dina / use_twostage / fs.splits / maxk and related
  # parameters do not apply in this mode.
)

timings$fs_grf_select <- (proc.time() - t0)["elapsed"]

if (!is.null(fs_grf$sg.harm) && length(fs_grf$sg.harm) > 0) {
  cat("GRF-selected subgroup (H):",
      fmt_sg(fs_grf$sg.harm), "\n")
  cat("Subgroup size:", sum(fs_grf$df.est$treat.recommend == 0),
      sprintf("(%.1f%% of ITT)\n",
              100 * mean(fs_grf$df.est$treat.recommend == 0)))
} else {
  cat("GRF selection: no subgroup met the criteria.\n")
}
```

### Bootstrap Bias-Corrected OR Estimates

The within-subgroup OR is upwardly biased by selection. The ForestSearch
bootstrap corrects this by re-running the full pipeline on each resample;
because the result above was produced in GRF-selection mode, each replicate
re-fits the causal forest and re-runs GRF selection from scratch — so the
correction reflects the GRF selection process itself.

```{r fs-grf-bootstrap, eval = !is.null(fs_grf$sg.harm)}
# Runs only when GRF identified a subgroup.
t0 <- proc.time()

fs_grf_bc <- forestsearch_bootstrap_dofuture(
  fs.est = fs_grf,
  nb_boots = NB,
  show_three = FALSE,
  details = FALSE,
  digits = construction_digits,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential"); gc(verbose = FALSE)
timings$fs_grf_bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nGRF-mode bootstrap completed in",
    round(timings$fs_grf_bootstrap / 60, 1), "minutes\n")
```

```{r fs-grf-no-subgroup, eval = is.null(fs_grf$sg.harm)}
# Shown only when GRF selected no subgroup at the configured DR-score floor.
timings$fs_grf_bootstrap <- 0
cat("GRF (subgroup_method = \"grf\") identified no subgroup at",
    "dmin.grf =", fs_grf_demo_dmin, "-- bootstrap skipped.\n",
    "Lower the DR-score floor (dmin.grf) or switch grf_selection /",
    "grf_depth if a bootstrap demonstration is desired.\n")
```

```{r fs-grf-bootstrap-summary, eval = !is.null(fs_grf$sg.harm)}
grf_summaries <- summarize_bootstrap_results(
  sgharm = fs_grf$sg.harm,
  boot_results = fs_grf_bc,
  create_plots = FALSE,
  est.scale = "hr",
  digits = display_digits
)

grf_summaries$table
```

#### Tier-2 de-biased gate (counterpart to the Tier-1 bootstrap)

The same Tier-2-vs-Tier-1 comparison as the main analysis, for the GRF-selection
fit. The gate de-biases against GRF's **own** DR-score candidate family
(re-running nothing), while the bootstrap above re-runs GRF on every resample.

As with DINA, the two tiers are **not** expected to agree as tightly as for the
consistency method. GRF selects on its **doubly-robust (RD-based) score**, but
the gate can only re-select on the perturbed **GLM** OR; the additive RD scale
and the multiplicative odds scale need not rank candidates the same way, so over
a large candidate family the gate's winner's-curse term can exceed the
bootstrap's and over-correct the selected subgroup. To keep the gate's
re-selection neighbourhood close to GRF's own selection band, the gate by default
restricts that family to candidates competitive on GRF's native DR-score,
matching the band to `effect_neighborhood` (the recommended default, used here
via `fs_grf_family_native_nbhd <- NULL`; set an explicit number for a different
band, or `>= 1` to de-bias against the full family). The complement is the
induced residual rather than a max-over-family selection, so it agrees across
tiers regardless. Two caveats specific to GRF: a *dominant* DR-score winner can
collapse the restricted family to a single candidate (`candidate family = 1` in
the note below), in which case the gate removes only the gate-threshold optimism;
and, more fundamentally, the band tunes only the *candidate-competition* part of
the bias — the gate holds the causal forest fixed, so the forest-refit optimism
that the bootstrap captures is invisible to it. Treat the full bootstrap as the
reference for GRF and the gate as a fast, conservative check.

```{r fs-grf-debias-gate-comparison, eval = !is.null(fs_grf$sg.harm)}
#| echo: false
#| results: asis
g <- fs_grf$mr_inference
if (is.null(g)) {
  cat("_The de-biased gate did not run (set `run_mr <- TRUE`)._\n")
} else {
  fmt_ci <- function(e, l, u) sprintf("%.3f (%.3f, %.3f)", e, l, u)
  fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0 <- fl(fs_grf_bc$H_estimates, "H0"); H0l <- fl(fs_grf_bc$H_estimates, "H0_lower"); H0u <- fl(fs_grf_bc$H_estimates, "H0_upper")
  H2 <- fl(fs_grf_bc$H_estimates, "H2"); H2l <- fl(fs_grf_bc$H_estimates, "H2_lower"); H2u <- fl(fs_grf_bc$H_estimates, "H2_upper")
  cc <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  q  <- c("naive OR (95% CI)", "de-biased OR (95% CI)")
  t2 <- c(fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  t1 <- c(fmt_ci(H0, H0l, H0u), fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    Hc0 <- fl(fs_grf_bc$Hc_estimates, "H0"); Hc0l <- fl(fs_grf_bc$Hc_estimates, "H0_lower"); Hc0u <- fl(fs_grf_bc$Hc_estimates, "H0_upper")
    Hc2 <- fl(fs_grf_bc$Hc_estimates, "H2"); Hc2l <- fl(fs_grf_bc$Hc_estimates, "H2_lower"); Hc2u <- fl(fs_grf_bc$Hc_estimates, "H2_upper")
    q  <- c(q, "complement naive OR (95% CI)", "complement de-biased OR (95% CI)")
    t2 <- c(t2, fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    t1 <- c(t1, fmt_ci(Hc0, Hc0l, Hc0u), fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `Tier 2 (gate)` = t2,
                     `Tier 1 (bootstrap)` = t1, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE,
        caption = sprintf("GRF-selected subgroup: %s%s",
                          fmt_sg(fs_grf$sg.harm),
                          if (has_cc) " (complement = remaining subjects)" else "")),
      sep = "\n"); cat("\n\n")

  boot_s <- as.numeric(timings$fs_grf_bootstrap)
  cat(sprintf("Timing: Tier 1 bootstrap (%d boots) %.1f s (%.1f min); Tier 2 gate %.2f s",
              NB, boot_s, boot_s / 60, g$timing_seconds))
  if (is.finite(boot_s) && g$timing_seconds > 0)
    cat(sprintf(" -> **%.0fx** faster (the bootstrap scales with `NB`, the gate does not)",
                boot_s / g$timing_seconds))
  cat(".\n\n")
  cat(sprintf("Gate harm flag = **%s** (de-biased OR %.3f %s gate %.2f); selection bias %.4f (log-OR); candidate family = %d.\n",
              isTRUE(g$harm_flag), g$debiased$est,
              if (g$settings$confirm_rule == "point") "vs" else "lower-bound vs",
              g$settings$t_confirm, g$selection_bias, g$n_family))
  cat(sprintf("Re-selection rule = `%s` (a GLM-OR proxy for GRF's DR-score); native-family band = %s.\n",
              if (!is.null(g$settings$reselection)) g$settings$reselection else "?",
              if (is.null(fs_grf_family_native_nbhd)) "default (matched to effect_neighborhood)"
              else if (fs_grf_family_native_nbhd >= 1) "off (full family)"
              else sprintf("%.2f (restricted to DR-score-competitive candidates)",
                           fs_grf_family_native_nbhd)))
  if (!is.null(g$ci_method) && !is.null(g$debiased$se_ij))
    cat(sprintf("De-biased CI: %s (IJ SE %.3f vs robust SE %.3f, log-OR).\n",
                if (identical(g$ci_method, "ij")) "infinitesimal jackknife (Tier-1 analogue)" else "robust SE",
                g$debiased$se_ij, g$debiased$se_wald))
}
```

The Tier-1 bias-corrected table above, rebuilt from the Tier-2 gate object.

```{r fs-grf-debias-gate-table, eval = !is.null(fs_grf$sg.harm) && !is.null(fs_grf$mr_inference)}
mr_estimates_table(fs_grf, boot_results = fs_grf_bc, est.scale = "hr", digits = 2)
```

::: {.callout-note}
## Selection statistic vs. the gate's re-selection (GRF)

GRF ranks candidates by the **doubly-robust score** (RD-based for binary), while
the Tier-2 gate re-selects on the perturbed **GLM** OR. The two scales (additive
RD vs. multiplicative odds) need not agree, so the gate is a *proxy* for GRF's
selection — the same approximation noted for DINA's tau-hat.
`family_native_neighborhood` restricts the gate family to DR-score-competitive
candidates to keep that proxy close to GRF's actual selection neighbourhood; the
full bootstrap remains the reference because it re-runs the DR-score selection
itself.
:::



# Cross-Validation

Cross-validation assesses the **stability** of the ForestSearch algorithm.
Two approaches are available.

## K-Fold Cross-Validation

```{r cv-skipped-note, eval = !run_cv}
cat("K-Fold cross-validation skipped (run_cv = FALSE). Set run_cv <- TRUE",
    "in the parallel-setup chunk to run it.\n")
```

```{r kfold-cv, eval = run_cv}
t0 <- proc.time()

fs_kfold <- forestsearch_tenfold(
  fs.est = fs,
  sims = Ksims,
  Kfolds = 10L,
  details = FALSE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = FALSE
  )
)

plan("sequential"); gc(verbose = FALSE)
timings$kfold <- (proc.time() - t0)["elapsed"]
metrics_tables <- cv_metrics_tables(fs_kfold)
metrics_tables
```

### Diagnostic Tabulations

`summarize_cv_results()` consolidates the per-fold record into a panel of
diagnostic tables: which subgroups the pipeline identified across resamples,
which GRF cuts drove those identifications, when identification failed what the
proximate cause was, and how confidently the successful identifications were
made.

```{r cv-diagnostics-run, eval = run_cv}
cv_diag <- summarize_cv_results(
  cv_output         = fs_kfold,
  original_sg       = fs$sg.harm,
  original_grf_cuts = fs$grf_cuts,
  top_n             = 10L
)
```

```{r cv-identification, eval = run_cv}
cv_diag$identification_summary
```

```{r cv-grf-cut, eval = run_cv}
cv_diag$grf_cut_summary
```

```{r cv-no-subgroup, eval = run_cv}
cv_diag$no_subgroup_decomposition
```

```{r cv-pcons, eval = run_cv}
cv_diag$pconsistency_distribution
```

```{r cv-fold-numeric, eval = run_cv}
cv_diag$fold_numeric_summary
```

```{r cv-original-agreement, eval = run_cv}
cv_diag$original_agreement
```


## Out-of-Bag (N-Fold) Cross-Validation

N-fold CV (leave-one-out):

```{r loo-skipped-note, eval = !run_loo}
cat("Leave-one-out (N-fold / OOB) skipped (run_loo = FALSE). Set run_loo <- TRUE",
    "in the parallel-setup chunk to run it.\n")
```

```{r oob-cv, eval = run_loo}
t0 <- proc.time()

fs_OOB <- forestsearch_Kfold(
  fs.est = fs,
  details = FALSE,
  Kfolds = nrow(fs$df.est),  # N-fold = leave-one-out
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential"); gc(verbose = FALSE)
timings$oob <- (proc.time() - t0)["elapsed"]

# Summarize OOB results
cv_out <- forestsearch_KfoldOut(
  res = fs_OOB,
  details = FALSE,
  outall = TRUE
)

tables <- cv_summary_tables(cv_out)

tables$combined_table

tables$metrics_table
```


# Results Visualization

## Forest Plot

The forest plot summarizes treatment effects across the ITT population and the
identified subgroup with bootstrap-bias-corrected OR (and cross-validation
metrics when CV was run). OR is a ratio measure, so the axis is on the log scale.

```{r forest-plot}
#| fig-width: 14
#| fig-height: 6
#| fig-cap: "Subgroup forest plot; OR on the adverse scale (no improvement)"

# Build the fs_results list dynamically based on what's been computed.
fs_results_list <- list(fs.est = fs)
if (exists("fs_bc")) fs_results_list$fs_bc <- fs_bc

cv_label_for_plot <- NULL
if (exists("fs_kfold")) {
  fs_results_list$fs_kfold <- fs_kfold
  cv_label_for_plot <- "10-fold CV"
} else if (exists("fs_OOB")) {
  fs_results_list$fs_kfold <- fs_OOB
  cv_label_for_plot <- "LOO CV"
}

forest_obj <- plot_subgroup_results_forestplot(
  fs_results     = fs_results_list,
  df_analysis    = actg_df,
  outcome.name   = outcome.name,
  event.name     = outcome.name,
  treat.name     = treat.name,
  E.name         = "ZDV+ddI",
  C.name         = "ddI",
  outcome_type   = fs_outcome_type,
  effect_measure = fs_effect_measure,
  xlog           = TRUE,                 # OR is ratio scale -> log axis
  arrow_text     = c("Favors ZDV+ddI", "Favors ddI"),
  title_text     = "ACTG175 Subgroup Results — Binary (OR, adverse scale)",
  show_cv_metrics = !is.null(cv_label_for_plot),
  cv_source      = "kfold",
  xlim_method    = "data"
)

render_forestplot(forest_obj)

if (!is.null(cv_label_for_plot)) {
  cat(sprintf("\nCV annotations source: %s.\n", cv_label_for_plot))
}
```


# Summary and Interpretation

## Key Findings

```{r summary-findings}
cat(strrep("=", 60), "\n")
cat("FORESTSEARCH ANALYSIS SUMMARY\n")
cat(strrep("=", 60), "\n\n")

cat("Dataset: ACTG175 (N =", nrow(actg_df), ")\n")
cat("Outcome: binary, no CD4 improvement (adverse scale), OR measure\n\n")

cat("ITT Analysis (adverse scale):\n")
cat(sprintf("  OR (95%% CI): %.2f (%.2f, %.2f)\n\n", itt_or, itt_ci_lo, itt_ci_hi))

if (!is.null(fs$sg.harm) && length(fs$sg.harm) > 0) {
  cat("Identified Subgroup (H):\n")
  cat("  Definition:", fmt_sg(fs$sg.harm), "\n")
  cat("  Size:", sum(fs$df.est$treat.recommend == 0),
      sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 0)))

  cat("\nComplement Subgroup (Hc):\n")
  cat("  Size:", sum(fs$df.est$treat.recommend == 1),
      sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 1)))
} else {
  cat("No harm subgroup was identified at the configured thresholds.\n")
}
```

## Interpretation Notes

The ForestSearch analysis identifies a subgroup where the treatment is
associated with elevated odds of no CD4 improvement (harm on the adverse scale).
As an exploratory result it requires the usual caveats:

1. This is an **exploratory** analysis requiring independent validation.
2. The bias-corrected (Tier-1 bootstrap and Tier-2 gate) estimates have wider
   confidence intervals than the naive subgroup OR.
3. Cross-validation metrics should be evaluated for algorithm stability.

The Tier-1/Tier-2 comparison is the methodological focus: for the consistency
engine the gate is expected to track the full bootstrap closely on both point
estimate and IJ interval; for DINA and GRF the gate is a fast, conservative proxy
(it holds the engine's fit fixed), so the full bootstrap remains the reference.

## Computational Timing

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Analysis = c("GRF (preliminary)", "ForestSearch", "Bootstrap",
               "FS (DINA screening)", "DINA (standalone)",
               "FS (subgroup_method=dina)", "FS DINA bootstrap",
               "FS (subgroup_method=grf)", "FS GRF bootstrap",
               "kfold", "oob", "Total"),
  Seconds = c(
    timings$grf,
    timings$forestsearch,
    timings$bootstrap,
    timings$fs_dina_screen,
    timings$dina_standalone,
    timings$fs_dina_select,
    if (!is.null(timings$fs_dina_bootstrap)) timings$fs_dina_bootstrap else NA_real_,
    timings$fs_grf_select,
    if (!is.null(timings$fs_grf_bootstrap)) timings$fs_grf_bootstrap else NA_real_,
    timings$kfold,
    timings$oob,
    timings$total
  )
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(title = "Computational Timing") |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1) |>
  cols_label(
    Analysis = "Component",
    Seconds = "Time (sec)",
    Minutes = "Time (min)"
  )
```


# References

León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024).
"Exploratory subgroup identification in the heterogeneous Cox model:
A relatively simple procedure." *Statistics in Medicine*.
DOI: [10.1002/sim.10163](https://doi.org/10.1002/sim.10163)

Hammer SM, Katzenstein DA, Hughes MD, et al. (1996). A trial comparing
nucleoside monotherapy with combination therapy in HIV-infected adults with CD4
cell counts from 200 to 500 per cubic millimeter. AIDS Clinical Trials Group
Study 175 Study Team. *N Engl J Med*. 335(15):1081-90.


# Session Information

```{r session-info}
sessionInfo()
```


# Reproducibility payload (Section 6.2) {#sec-export-payload-actg}

The chunk below harvests the already-computed objects from the three engine runs
into one small list and writes it as an `.rds` for the manuscript Section 6.2
reader document to reproduce the table and the results overview. No analysis is
repeated; the output path, core count, and package version resolve at run time.

```{r export-section6-2-payload}
#| echo: false
#| message: false

.qmd_dir  <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) getwd())
.qmd_stem <- tryCatch(tools::file_path_sans_ext(basename(knitr::current_input())),
                      error = function(e) "analysis")
.dirout   <- if (is.null(dirout)) .qmd_stem else dirout
.base     <- if (is.null(results_dir)) file.path(.qmd_dir, "_payloads") else results_dir
.out_dir  <- file.path(.base, .dirout)
dir.create(.out_dir, recursive = TRUE, showWarnings = FALSE)

.get <- function(nm) if (exists(nm, inherits = TRUE)) get(nm, inherits = TRUE) else NULL

.triple <- function(x) {
  if (is.null(x) || is.null(x$est)) c(NA_real_, NA_real_, NA_real_) else c(x$est, x$lower, x$upper)
}
.fb_triple <- function(est_list) {
  if (is.null(est_list)) c(NA_real_, NA_real_, NA_real_) else c(est_list$H2, est_list$H2_lower, est_list$H2_upper)
}

.engine_rows <- function(method, fit, boot) {
  if (is.null(fit) || is.null(fit$mr_inference)) return(NULL)
  g  <- fit$mr_inference
  cc <- g$complement
  tr <- fit$df.est$treat.recommend
  n_tot <- length(tr); n_H <- sum(tr == 0, na.rm = TRUE)
  nv_H  <- .triple(g$naive)
  nv_Hc <- .triple(if (!is.null(cc)) cc$naive)
  mr_H  <- .triple(g$debiased)
  mr_Hc <- .triple(if (!is.null(cc)) cc$debiased)
  fb_H  <- .fb_triple(if (!is.null(boot)) boot$H_estimates)
  fb_Hc <- .fb_triple(if (!is.null(boot)) boot$Hc_estimates)
  data.frame(
    method = method, region = c("H", "Hc"),
    n = c(n_H, n_tot - n_H),
    pct = round(100 * c(n_H, n_tot - n_H) / n_tot),
    naive_est = c(nv_H[1], nv_Hc[1]), naive_lo = c(nv_H[2], nv_Hc[2]), naive_hi = c(nv_H[3], nv_Hc[3]),
    fb_est = c(fb_H[1], fb_Hc[1]), fb_lo = c(fb_H[2], fb_Hc[2]), fb_hi = c(fb_H[3], fb_Hc[3]),
    mr_est = c(mr_H[1], mr_Hc[1]), mr_lo = c(mr_H[2], mr_Hc[2]), mr_hi = c(mr_H[3], mr_Hc[3]),
    stringsAsFactors = FALSE)
}

.engine_meta <- function(fit, boot_seconds) {
  if (is.null(fit) || is.null(fit$mr_inference))
    return(list(label = NA_character_, n_family = NA_integer_, selection_bias = NA_real_,
                boot_sec = NA_real_, mr_sec = NA_real_, speedup = NA_real_))
  g <- fit$mr_inference
  mr_sec <- if (!is.null(g$timing_seconds)) as.numeric(g$timing_seconds) else NA_real_
  list(label = paste(fit$sg.harm, collapse = " & "),
       n_family = g$n_family, selection_bias = g$selection_bias,
       boot_sec = boot_seconds, mr_sec = mr_sec,
       speedup = if (is.finite(boot_seconds) && is.finite(mr_sec) && mr_sec > 0) boot_seconds / mr_sec else NA_real_)
}

.tm <- .get("timings")
.boot_sec <- function(key) {
  v <- if (!is.null(.tm) && !is.null(.tm[[key]])) suppressWarnings(as.numeric(.tm[[key]])) else NA_real_
  if (length(v) == 1L && is.finite(v)) v else NA_real_
}

# FS vs GRF agreement on the harm flag (treat.recommend == 0), aligned by id, with Cohen's kappa.
.concordance <- function(a, b, idn) {
  if (is.null(a) || is.null(b)) return(NULL)
  da <- a$df.est; db <- b$df.est
  if (is.null(da) || is.null(db) || !(idn %in% names(da)) || !(idn %in% names(db))) return(NULL)
  m <- merge(data.frame(.id = da[[idn]], ha = as.integer(da$treat.recommend == 0)),
             data.frame(.id = db[[idn]], hb = as.integer(db$treat.recommend == 0)), by = ".id")
  if (nrow(m) == 0L) return(NULL)
  po <- mean(m$ha == m$hb); pa <- mean(m$ha); pb <- mean(m$hb)
  pe <- pa * pb + (1 - pa) * (1 - pb)
  list(n_common = nrow(m), n_agree = sum(m$ha == m$hb), pct_agree = 100 * po,
       kappa = if (pe < 1) (po - pe) / (1 - pe) else NA_real_)
}

# Adverse-outcome rate by arm inside an engine's harm region (treat.recommend == 0).
.harm_rates <- function(fit, oname, tname) {
  if (is.null(fit) || is.na(oname) || is.na(tname)) return(NULL)
  d <- fit$df.est
  if (is.null(d) || !all(c(oname, tname, "treat.recommend") %in% names(d))) return(NULL)
  h <- d[d$treat.recommend == 0, , drop = FALSE]
  if (nrow(h) == 0L) return(NULL)
  yt <- h[[oname]][h[[tname]] == 1]; yc <- h[[oname]][h[[tname]] == 0]
  list(treated_pct = if (length(yt)) 100 * mean(yt) else NA_real_,
       control_pct = if (length(yc)) 100 * mean(yc) else NA_real_)
}

.fs      <- .get("fs");      .fs_bc      <- .get("fs_bc")
.fs_dina <- .get("fs_dina"); .fs_dina_bc <- .get("fs_dina_bc")
.fs_grf  <- .get("fs_grf");  .fs_grf_bc  <- .get("fs_grf_bc")

.meta_fs   <- .engine_meta(.fs,      .boot_sec("bootstrap"))
.meta_dina <- .engine_meta(.fs_dina, .boot_sec("fs_dina_bootstrap"))
.meta_grf  <- .engine_meta(.fs_grf,  .boot_sec("fs_grf_bootstrap"))

table1 <- do.call(rbind, list(
  .engine_rows("FS",   .fs,      .fs_bc),
  .engine_rows("DINA", .fs_dina, .fs_dina_bc),
  .engine_rows("GRF",  .fs_grf,  .fs_grf_bc)))

.dsel <- .get("dina_sel")
dina_standalone <- if (!is.null(.dsel) && isTRUE(.dsel$found)) {
  op <- if (identical(.dsel$direction, "left")) "<=" else ">="
  list(found = TRUE, label = sprintf("%s %s %s", .dsel$covariate, op, format(.dsel$threshold)), n = .dsel$n_subgroup)
} else {
  list(found = FALSE, label = NA_character_, n = NA_integer_)
}

.idname <- if (exists("id.name")) id.name else "id"
.oname  <- if (exists("outcome.name")) outcome.name else NA_character_
.tname  <- if (exists("treat.name")) treat.name else NA_character_

payload <- list(
  table  = table1,
  labels = list(fs = .meta_fs$label, dina = .meta_dina$label, grf = .meta_grf$label,
                dina_standalone = dina_standalone),
  meta   = list(
    B          = if (exists("NB")) NB else NA_integer_,
    mr_draws = if (exists("mr_draws")) mr_draws else NA_integer_,
    n_total    = if (!is.null(.fs)) nrow(.fs$df.est) else NA_integer_,
    cores      = tryCatch(unname(future::availableCores()), error = function(e) NA_integer_),
    covariates = if (exists("confounders.name")) confounders.name else NA_character_,
    timings    = list(fs = .meta_fs, dina = .meta_dina, grf = .meta_grf)),
  extras = list(
    concordance = .concordance(.fs, .fs_grf, .idname),
    harm_rates  = .harm_rates(.fs, .oname, .tname)),
  est_scale = "or",
  built_at  = Sys.time(),
  forestsearch_version = tryCatch(as.character(utils::packageVersion("forestsearch")), error = function(e) NA_character_))

.payload_file <- file.path(.out_dir, paste0(.qmd_stem, "_payload.rds"))
saveRDS(payload, .payload_file)
cat("Wrote Section 6.2 payload to:", .payload_file, "\n")
```
