---
title: "ACTG175 Continuous Analysis: Applied Operating-Characteristic Evaluation"
format: html
---

```{r setup}
#| message: false
library(gt)

# ---------------------------------------------------------------------------
# Output location for this document's results:  <base>/<dirout>/
#   results_dir = NULL -> base is <qmd dir>/_payloads
#   dirout      = NULL -> the .qmd stem
# ---------------------------------------------------------------------------
results_dir <- NULL
dirout      <- NULL

# This document reads the stored evaluation objects only; it computes nothing
# heavy and types no number.  The draw sets behind it were produced by
# dev/glm-continuous-sims/oc_applied_eval_rung_2026-08-31.R (eleven rungs,
# 2e5 draws each) and assembled by oc_applied_eval_summary_2026-08-31.R.
.qmd_dir <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                     error = function(e) getwd())
repo_root <- normalizePath(file.path(.qmd_dir, "..", "..", ".."))
S <- readRDS(file.path(repo_root, "dev", "glm-continuous-sims",
                       "oc_applied_eval_summary_2026-08-31.rds"))

fmt  <- function(x, d = 3) formatC(x, format = "f", digits = d)
fmt0 <- function(x) formatC(x, format = "d")
pct  <- function(x, d = 1) paste0(fmt(100 * x, d), "%")
```

## 1. The analysis and its anchor

The applied analysis this evaluation attaches to is the ACTG175 continuous
harm search of `analysis_actg175_continuous_compare_all.qmd`: outcome
`y_decline` (CD4 decline, adverse scale) on the ZDV+ddI vs ddI arms
(N = `r fmt0(S$meta$n)`), twelve confounders (six continuous with
`jcuts = 10`, six binary), thresholds c1 = `r fmt0(S$meta$analyst_c1)` and
c2 = `r fmt0(S$meta$analyst_c2)` on the mean-difference scale,
`consistency_method = "`r S$meta$consistency_method`"` with
`pconsistency = `r fmt(S$meta$pconsistency, 2)``, `n.min = 60`, `maxk = 2`.

The evaluation attaches to the **fixed-family `maxeffCons` configuration** —
the MR-anchor configuration of that document, with
`use_lasso = use_grf = use_dina = FALSE`.  That is the document's own
rationale for its anchor: multiplier-resampling inference and the lasso/GRF
front ends are mutually exclusive under `subgroup_method = "consistency"`
(the family would be estimated from the same data the search runs on), so
the inferential anchor runs with the front ends off, and the operating
characteristics evaluated here are the operating characteristics of exactly
that procedure.  With the front ends off the candidate family is enumerated
from the covariates and their quantile cuts alone; the cut locations are
still quantile-derived ("no-front-end" rather than literally fixed), and the
evaluation holds this family fixed — see the limitations.

The anchor that analysis found:
**Ĥ = `` `r S$meta$anchor$def` ``**, n(Ĥ) = `r fmt0(S$meta$anchor$n_H)`
of `r fmt0(S$meta$n)`, fitted mean difference
**T̂~obs~ = `r fmt(S$meta$anchor$T_obs, 6)`** CD4 units, consistency
`r fmt(S$meta$anchor$p_cons, 2)`.  The ITT mean difference on `y_decline`
is `r fmt(S$meta$itt[1, 1], 2)` (SE `r fmt(S$meta$itt[1, 2], 2)`) — on
average the combination arm *reduces* CD4 decline, so a harm subgroup, if
real, sits against a beneficial background.

## 2. The anchored DGM

Truths are anchored to the analysis data by a two-point mixture
stylization, stated plainly: the super-population
(n = `r fmt0(S$meta$scale$regions$n_g[3])` draws of the covariate rows) has
exactly two treatment-effect values — the fitted background effect
`r fmt(S$meta$beta_treat, 6)` on Q^c^ and a planted effect q on Q — with
the baseline outcome surface and residual scale
(σ = `r fmt(S$meta$scale$sigma, 2)`) taken from the baseline GLM fit to the
analysis data.  Nothing else about the joint covariate law is modeled; the
covariate rows are the data's own.

Q is the found rule carried into the generator by its clause mapping
(`age = 37`; `cd40 = list(type = "greater", value = 507)`), reproducing
Ĥ's membership subject-for-subject on the analysis frame
(`r fmt0(S$meta$anchor$n_H)`/`r fmt0(S$meta$anchor$n_H)`).  On the
super-population Q has prevalence `r fmt(S$meta$scale$regions$P_g[1], 4)`.

**Q's representability in the discovery family.**  The evaluation family is
the analyst's own discovery grid at n = `r fmt0(S$meta$n)`
(M = `r fmt0(S$meta$M)` candidate rules after floors and deduplication),
kept pure: Q is *not* forced in.  Ĥ's `cd40` threshold (507) lies just off
the population quantile grid (nearest grid cut 506), so no member is
identical to Ĥ; the nearest member
(`` `r S$meta$purity$nearest_lab` ``) matches it with purity and Jaccard
`r fmt(S$meta$purity$nearest_purity, 4)` — the grid expresses the found
rule to within about `r pct(1 - S$meta$purity$nearest_purity, 2)` of its
membership.  This is reported as a feature of the evaluation: the search is
scored on the family it actually searches.

**The rungs.**  One effect ladder,
q ∈ {`r paste(sapply(S$meta$q_rungs, function(x) fmt(x, ifelse(x %% 1 == 0, 0, 2))), collapse = ", ")`}
CD4 units: from essentially zero (q = `r fmt(S$meta$q_rungs[1], 2)`, the
zero-plus structural null) through the observed statistic
(q = T̂~obs~ = `r fmt(S$meta$T_obs, 2)`).  Each rung is one
`r format(S$meta$draws, big.mark = ",")`-draw set (seed
`r fmt0(S$meta$seed)`) with every threshold c1 ∈ {0, 1, …, 200, T̂~obs~}
swept against it at the analyst's c2 = `r fmt0(S$meta$analyst_c2)`.  Across
rungs the family's memberships, prevalences, standard errors and overlap
matrix are bit-identical (gated); only the planted effect moves, and each
candidate's oriented mean moves linearly in q through its purity.

## 3. Type-I under the structural null

**Why the zero-plus rung, not the null branch.**  Version 0.3.2 orients
every candidate's mean by the harm direction s = sign(m~τ~[Q]), so the
oriented means are the signed mixture s·(m~τ~[Q^c^] + (m~τ~[Q] −
m~τ~[Q^c^])·PQg): benefit-direction rules carry oriented-*negative* means,
and a planted effect of exactly zero is rejected (no harm direction to
orient by) — deliberately.  The wrapper's null *branch*, by contrast, keeps
its pre-0.3.2 oriented-absolute reading of the single common effect: on
this population it would turn the fitted −27 benefit into a +27 "harm",
and stage 0 measured exactly that — false declaration
`r fmt(S$meta$stage0_null_det, 3)` under the null branch at the analyst's
gate.  That reading is nonsense here (it is correct for the MD40-style
fixtures whose generators hand the search a positive-adverse outcome), so
the null branch is not edited and not used.  The operational structural
null is the alternative branch at q = `r fmt(S$meta$q_rungs[1], 2)` —
correctly signed under 0.3.2, and the same truth as "no subgroup" to a
hundredth of a CD4 unit.

At the analyst's gate (c1 = `r fmt0(S$meta$analyst_c1)`,
c2 = `r fmt0(S$meta$analyst_c2)`), the search's family-wise false-declaration
rate under this structural null is
**`r fmt(S$type1$at_analyst$det_rate, 5)`**
(MC SE `r fmt(S$type1$at_analyst$det_rate_se, 5)`).

Read at 1-unit resolution — the resolution of this document's integer c1
grid — the smallest integer c1 holding type-I at or below 0.05 is
**c1~05~ = `r fmt0(S$meta$c1_05)`**
(rate `r fmt(S$type1$rate_at_c1_05$det_rate, 5)` ±
`r fmt(S$type1$rate_at_c1_05$det_rate_se, 5)`), and at or below 0.10 is
**c1~10~ = `r fmt0(S$meta$c1_10)`**
(rate `r fmt(S$type1$rate_at_c1_10$det_rate, 5)` ±
`r fmt(S$type1$rate_at_c1_10$det_rate_se, 5)`).

```{r type1-curve}
#| fig-width: 7
#| fig-height: 4.5
#| fig-cap: "Type-I error of the search under the zero-plus structural null, against the screening threshold c1 (c2 fixed at the analyst's value). Guides: the analyst's c1, and c1_05 / c1_10 at the 0.05 / 0.10 lines."
tc <- S$type1$curve
op <- par(mar = c(4.2, 4.2, 1, 1))
plot(tc$c1, tc$det_rate, type = "l", lwd = 2,
     xlab = "c1 (CD4 units, mean difference)",
     ylab = "P(any declaration | structural null)")
abline(h = c(0.05, 0.10), lty = 3, col = "grey40")
abline(v = S$meta$analyst_c1, lty = 2, col = "firebrick")
abline(v = c(S$meta$c1_05, S$meta$c1_10), lty = 2, col = "steelblue")
text(S$meta$analyst_c1, max(tc$det_rate), "analyst c1", pos = 4, cex = 0.8,
     col = "firebrick")
text(S$meta$c1_05, 0.05, expression(c1[0.05]), pos = 4, cex = 0.8,
     col = "steelblue")
par(op)
```

## 4. The effect ladder: power at fixed type-I

Per rung, the declaration rate at the analyst's c1 and the power at the
fixed-type-I thresholds c1~05~ and c1~10~; the last three columns give the
largest integer c1 at which the rung still holds 80 / 90 / 95% power (read
from the grid at 1-unit resolution; no separate inversion calls).

```{r ladder-table}
ld <- S$ladder
gt(data.frame(
  q = sapply(ld$q, function(x) fmt(x, ifelse(x %% 1 == 0, 1, 2))),
  det10 = paste0(fmt(ld$det_at_analyst, 4), " ± ",
                 fmt(ld$det_at_analyst_se, 4)),
  p05 = paste0(fmt(ld$power_at_c1_05, 4), " ± ",
               fmt(ld$power_at_c1_05_se, 4)),
  p10 = paste0(fmt(ld$power_at_c1_10, 4), " ± ",
               fmt(ld$power_at_c1_10_se, 4)),
  c80 = ifelse(is.na(ld$c1_pow80), "—", fmt0(ld$c1_pow80)),
  c90 = ifelse(is.na(ld$c1_pow90), "—", fmt0(ld$c1_pow90)),
  c95 = ifelse(is.na(ld$c1_pow95), "—", fmt0(ld$c1_pow95)))) |>
  cols_label(q = "q (planted MD)",
             det10 = md(paste0("rate at c1 = ",
                               fmt0(S$meta$analyst_c1))),
             p05 = md(paste0("power at c1&#8320;&#8325; = ",
                             fmt0(S$meta$c1_05))),
             p10 = md(paste0("power at c1&#8321;&#8320; = ",
                             fmt0(S$meta$c1_10))),
             c80 = "c1 at 80%", c90 = "c1 at 90%", c95 = "c1 at 95%") |>
  tab_source_note(md(paste0(
    "One ", format(S$meta$draws, big.mark = ","),
    "-draw set per rung; ± is the Monte-Carlo SE. ",
    "— = the target rate is not attained at any integer c1 ≥ 0.")))
```

```{r declaration-curves}
#| fig-width: 7
#| fig-height: 5
#| fig-cap: "Declaration curves: the probability of any declaration against c1, one curve per rung (the zero-plus structural null lowest). Guides: c1_05, the analyst's c1, and the 0.80 power line."
cv <- S$curves
qs <- unique(cv$q)
cols <- hcl.colors(length(qs), "Zissou 1")
op <- par(mar = c(4.2, 4.2, 1, 1))
plot(NA, xlim = range(cv$c1), ylim = c(0, 1),
     xlab = "c1 (CD4 units, mean difference)", ylab = "P(any declaration)")
abline(h = 0.80, lty = 3, col = "grey40")
abline(v = c(S$meta$c1_05, S$meta$analyst_c1), lty = 2,
       col = c("steelblue", "firebrick"))
for (k in seq_along(qs)) {
  d <- cv[cv$q == qs[k], ]
  lines(d$c1, d$det_rate, lwd = 1.6, col = cols[k])
}
legend("topright", legend = sapply(qs, function(x) paste0("q = ", fmt(x, 2))),
       col = cols, lwd = 1.6, cex = 0.7, bg = "white")
par(op)
```

At the strict fixed-type-I threshold c1~05~ = `r fmt0(S$meta$c1_05)`,
`r if (any(S$ladder$power_at_c1_05 >= 0.80)) paste0("the smallest rung reaching 80% power is q = ", fmt(min(S$ladder$q[S$ladder$power_at_c1_05 >= 0.80]), 2), " CD4 units of planted harm") else paste0("no rung on this ladder reaches 80% power — even the observed-statistic rung attains only ", pct(S$ladder$power_at_c1_05[nrow(S$ladder)], 1))`;
at the analyst's own c1 = `r fmt0(S$meta$analyst_c1)`
(type-I `r fmt(S$type1$at_analyst$det_rate, 3)`), the rate reaches
`r pct(S$ladder$det_at_analyst[nrow(S$ladder)], 1)` at the observed-statistic
rung.  The search as configured is liberal: its declaration threshold sits
far below the level that would control the family-wise false-declaration
rate, and holding that rate to 5% costs most of the power at every planted
effect the ladder covers.

## 5. What is declared, per rung

Conditional on declaring at the analyst's gate: the expected size of the
declared subgroup, its expected overlap quality against Q (PPV, sensitivity,
specificity), the expected true oriented effect on the declared rule, the
naive selection bias, and the selection mass on rules whose true mean is
below c1.

```{r declared-table}
dc <- S$declared
gt(data.frame(
  q = sapply(dc$q, function(x) fmt(x, ifelse(x %% 1 == 0, 1, 2))),
  EnH = paste0(fmt(dc$EnH, 1), " ± ", fmt(dc$EnH_se, 2)),
  Eppv = paste0(fmt(dc$Eppv, 3), " ± ", fmt(dc$Eppv_se, 4)),
  Esens = paste0(fmt(dc$Esens, 3), " ± ", fmt(dc$Esens_se, 4)),
  Espec = paste0(fmt(dc$Espec, 3), " ± ", fmt(dc$Espec_se, 4)),
  EbetaH = paste0(fmt(dc$EbetaH, 2), " ± ", fmt(dc$EbetaH_se, 3)),
  naive = paste0(fmt(dc$Enaive_bias, 2), " ± ~",
                 fmt(dc$Enaive_bias_se_approx, 3)),
  mb = paste0(fmt(dc$mass_below, 3), " ± ", fmt(dc$mass_below_se, 4)))) |>
  cols_label(q = "q", EnH = md("E\\|Ĥ\\|"), Eppv = "E[PPV]",
             Esens = "E[sens]", Espec = "E[spec]",
             EbetaH = md("E[β(Ĥ)]"), naive = "E[naive − β(Ĥ)]",
             mb = "mass below c1") |>
  tab_source_note(md(paste0(
    "All conditional on declaration at (c1 = ", fmt0(S$meta$analyst_c1),
    ", c2 = ", fmt0(S$meta$analyst_c2), "). ± is the delta-method MC SE ",
    "from the multinomial selection counts; the naive-bias SE (marked ~) ",
    "treats the winner's noise scale as se_g and is approximate.")))
```

The zero-plus rung's row is the honest false-positive reading: when the
structural null declares (rate `r fmt(S$coherence$zero_plus_det, 4)`), the
declared rule's expected true oriented effect is
`r fmt(S$coherence$zero_plus_EbetaH, 2)` — a *benefit*, not a harm, pulled
up from the pure background `r fmt(S$meta$beta_treat, 2)` only by the
selection mass that lands on partially-Q-overlapping rules
(E[PPV] = `r fmt(S$declared$Eppv[1], 3)`) — and the entire selection mass
sits on rules whose true mean is below c1.

## 6. The calibration curve for the observed statistic

Per rung, P(T ≥ T̂~obs~ | q): the probability that the search's declared
statistic reaches the observed `r fmt(S$meta$T_obs, 2)` when the true
planted effect is q — read directly off each rung's grid at
c1 = T̂~obs~.

```{r calibration-table}
cb <- S$calibration
gt(data.frame(
  q = sapply(cb$q, function(x) fmt(x, ifelse(x %% 1 == 0, 1, 2))),
  p = paste0(fmt(cb$p_ge_Tobs, 4), " ± ", fmt(cb$p_ge_Tobs_se, 4)))) |>
  cols_label(q = "q (planted MD)", p = md("P(T ≥ T̂~obs~ | q)")) |>
  tab_source_note("Grid read-off at c1 = T̂obs; no extra machinery.")
```

Interpolating linearly between adjacent rungs (stated as interpolation, at
the ladder's own resolution), the curve crosses
0.05 near **q ≈ `r fmt(S$crossings$q_cross[1], 1)`**,
0.5 near **q ≈ `r fmt(S$crossings$q_cross[2], 1)`**, and
0.95 near **q ≈ `r fmt(S$crossings$q_cross[3], 1)`** CD4 units.

```{r calibration-curve}
#| fig-width: 7
#| fig-height: 4.5
#| fig-cap: "The calibration curve: P(T ≥ T̂_obs | q) against the planted effect q, with the observed statistic marked. Dotted guides at 0.05, 0.5, 0.95."
op <- par(mar = c(4.2, 4.2, 1, 1))
plot(cb$q, cb$p_ge_Tobs, type = "b", pch = 16, lwd = 1.6,
     xlab = "q: planted effect on Q (CD4 units, mean difference)",
     ylab = expression(P(T >= hat(T)[obs] ~ "|" ~ q)))
abline(h = c(0.05, 0.5, 0.95), lty = 3, col = "grey40")
abline(v = S$meta$T_obs, lty = 2, col = "firebrick")
text(S$meta$T_obs, 0.1, expression(hat(T)[obs]), pos = 2, col = "firebrick",
     cex = 0.9)
par(op)
```

This is *a consonance curve for the planted effect under the anchored
two-point mixture model; not a posterior, not adjusted for anything outside
the model*.

## 7. The `se_g` band

The family's standard errors are the prevalence-scaled anchor
se~g~ = √(V~eff~[Q]/(n·P~g~)); the direct per-member value replaces
V~eff~[Q] with the member's own V~eff~[g].  The ratio
se~direct~/se~g~ at three rungs, overall and by purity band:

```{r seg-band-table}
sb <- S$seg_band
bands <- do.call(rbind, lapply(as.character(c(1L, 7L, 11L)), function(k) {
  b <- sb[[k]]
  data.frame(q = fmt(b$q, 2),
             range = paste0("[", fmt(b$range[1], 4), ", ",
                            fmt(b$range[2], 4), "]"),
             median = fmt(b$median, 4),
             hi  = fmt(b$by_band$median[b$by_band$band == "PQg >= 0.95"], 4),
             mid = fmt(b$by_band$median[b$by_band$band == "mid"], 4),
             lo  = fmt(b$by_band$median[b$by_band$band == "PQg < 0.25"], 4))
}))
gt(bands) |>
  cols_label(q = "q", range = "ratio range", median = "median",
             hi = md("median, PQg ≥ 0.95"), mid = "median, mid",
             lo = md("median, PQg < 0.25")) |>
  tab_source_note(md(paste0(
    "se~direct~/se~g~ over the ", fmt0(S$meta$M), " members; purity bands ",
    "PQg ≥ 0.95 / [0.25, 0.95) / < 0.25.")))
```

At the top rung the largest deviation of the ratio from 1 is
`r pct(S$seg_band$max_dev_top, 2)` —
`r if (isTRUE(S$seg_band$sensitivity_triggered)) "above" else "within"` the
~5% reporting band, so the prevalence-scaled approximation is
`r if (isTRUE(S$seg_band$sensitivity_triggered)) "accompanied by the sensitivity run below" else "reported, not acted on: no sensitivity run was required"`.

```{r seg-sensitivity}
#| eval: !expr (!is.null(S$sensitivity))
#| include: !expr (!is.null(S$sensitivity))
sv <- S$sensitivity$table
gt(data.frame(
  run = sv$run,
  det10 = paste0(fmt(sv$det_at_analyst, 4), " ± ",
                 fmt(sv$det_at_analyst_se, 4)),
  p05 = paste0(fmt(sv$power_at_c1_05, 4), " ± ",
               fmt(sv$power_at_c1_05_se, 4)),
  p10 = paste0(fmt(sv$power_at_c1_10, 4), " ± ",
               fmt(sv$power_at_c1_10_se, 4)),
  pT = paste0(fmt(sv$p_ge_Tobs, 4), " ± ", fmt(sv$p_ge_Tobs_se, 4)),
  EnH = fmt(sv$EnH, 1), EbetaH = fmt(sv$EbetaH, 2),
  Eppv = fmt(sv$Eppv, 3))) |>
  cols_label(run = "", det10 = md(paste0("rate at c1 = ",
                                         fmt0(S$meta$analyst_c1))),
             p05 = md(paste0("power at ", fmt0(S$meta$c1_05))),
             p10 = md(paste0("power at ", fmt0(S$meta$c1_10))),
             pT = md("P(T ≥ T̂~obs~)"),
             EnH = md("E\\|Ĥ\\|"), EbetaH = md("E[β(Ĥ)]"),
             Eppv = "E[PPV]") |>
  tab_source_note(md(paste0(
    "Top rung (q = T̂~obs~) re-evaluated with the family's se~g~ replaced ",
    "by the member-direct se~direct~ — one draw set, same seed. Labeled as ",
    "the breadth ladder labeled it: **a sensitivity, not a constructor, ",
    "not adopted**.")))
```

## 8. Limitations

- **The between-rule E|Ĥ| gap.**  In the MD40/MD120 verification fixtures
  the wrapper's expected declared-subgroup size differed from the measured
  search's by about two subjects, an offset that did not scale with the
  planted harm.  The same wrapper machinery produces the E|Ĥ| column here;
  a constant-order offset of that size should be assumed rather than
  precision at the single-subject level.
- **Q is not in the family.**  With this analysis's numbers: Ĥ's `cd40`
  threshold (507) is off the population grid (nearest cut 506), the nearest
  member matches with purity `r fmt(S$meta$purity$nearest_purity, 4)`, and
  no member is identical.  Every "power" number is therefore the power to
  declare *some* rule, scored against a Q the family can express only to
  ~`r pct(1 - S$meta$purity$nearest_purity, 2)` of its membership — the
  honest representation of the search as the analyst configured it.
- **The mixture stylization.**  Truths are a two-point effect mixture on
  the data's own covariate rows with a GLM baseline surface; real
  heterogeneity is not two-valued.  The calibration curve inherits this:
  it is a consonance statement inside the model, per the caveat above.
- **Fixed-family conditioning.**  The evaluation conditions on the
  no-front-end family (`use_lasso = use_grf = FALSE`, the MR-anchor
  configuration).  The compare-all document's front-ends-on arms search a
  data-dependent family; their operating characteristics are a different
  (conditional-on-family) estimand and are not evaluated here.  The cut
  locations themselves are quantile-derived and would move under
  resampling; the evaluation holds them at their population values.

# Reproducibility payload {#sec-export-payload}

```{r export-payload}
#| echo: false
#| message: false

.qmd_stem <- tryCatch(tools::file_path_sans_ext(basename(knitr::current_input())),
                      error = function(e) "analysis")
.dirout   <- if (is.null(dirout)) .qmd_stem else dirout
.base     <- if (is.null(results_dir)) file.path(.qmd_dir, "_payloads") else results_dir
.out_dir  <- file.path(.base, .dirout)
dir.create(.out_dir, recursive = TRUE, showWarnings = FALSE)
.payload_file <- file.path(.out_dir, paste0(.dirout, "_payload.rds"))

payload <- list(
  table  = S$ladder,
  labels = list(
    sg_harm  = S$meta$anchor$def,
    sg_focus = "maxeffCons",
    focus    = NULL),
  meta = list(
    n_total    = S$meta$n,
    n_events   = NULL,
    outcome    = "y_decline",
    event      = NULL,
    treat      = "treat",
    covariates = c("age", "preanti", "wtkg", "karnof", "cd40", "cd80",
                   "hemo", "homo", "drugs", "race", "gender", "symptom"),
    c1         = S$meta$analyst_c1,
    c2         = S$meta$analyst_c2,
    args_call  = list(consistency_method = S$meta$consistency_method,
                      pconsistency = S$meta$pconsistency,
                      draws = S$meta$draws, block = S$meta$block,
                      seed = S$meta$seed, M = S$meta$M,
                      q_rungs = S$meta$q_rungs, T_obs = S$meta$T_obs)),
  extras = list(
    type1 = list(at_analyst = S$type1$at_analyst,
                 c1_05 = S$meta$c1_05, rate_at_c1_05 = S$type1$rate_at_c1_05,
                 c1_10 = S$meta$c1_10, rate_at_c1_10 = S$type1$rate_at_c1_10),
    declared    = S$declared,
    calibration = S$calibration,
    crossings   = S$crossings,
    coherence   = S$coherence,
    seg_band    = list(max_dev_top = S$seg_band$max_dev_top,
                       triggered   = S$seg_band$sensitivity_triggered,
                       by_band     = lapply(S$seg_band[
                         as.character(c(1L, 7L, 11L))], `[[`, "by_band")),
    anchor      = S$meta$anchor,
    purity      = S$meta$purity),
  est_scale = "md",
  built_at  = Sys.time(),
  forestsearch_version = tryCatch(as.character(utils::packageVersion("forestsearch")),
                                  error = function(e) NA_character_))

saveRDS(payload, .payload_file)
cat("Wrote payload to:", .payload_file, "\n")
```
