---
title: "ACTG175: Cross-Analysis Subgroup Summary"
subtitle: "Selected subgroups across continuous (MD) and binary (OR) outcomes"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")
```


# Introduction

This document loads the saved `selected_subgroups_<analysis>.rds`
bundles produced by the per-outcome selection-rule comparison vignettes
and renders a single side-by-side summary across all
`(sg_focus, selection_rule)` configurations and both outcome scales.

The expected source files (relative to this document's `_data/`
directory) are:

| File                                  | Produced by                                       |
|---------------------------------------|---------------------------------------------------|
| `selected_subgroups_continuous.rds`   | `analysis_actg175_continuous_compare_all.qmd`     |
| `selected_subgroups_binary.rds`       | `analysis_actg175_binary_compare_all.qmd`         |
| `comparison_continuous.rds`           | (full object, optional drill-down)                |
| `comparison_binary.rds`               | (full object, optional drill-down)                |

If a tidy file is missing, the section that depends on it reports the
gap and skips its content; the rest of the document still renders.


# Setup

## Packages

```{r load-packages}
suppressPackageStartupMessages({
  library(forestsearch)   # for compare_fs_subgroups() in the drill-down
  library(data.table)
  library(gt)
})
```

## Input Directory

```{r input-dir}
#| code-fold: false

# Resolved relative to *this qmd's directory*, not the caller's getwd().
# Matches the convention used by the comparison vignettes: artifacts
# are saved into <qmd-dir>/_data/ and this post-hoc document reads from
# the same place.  Override `input_dir` here if your two comparison
# qmds rendered into a different directory.
qmd_input  <- knitr::current_input(dir = TRUE)
qmd_dir    <- if (is.null(qmd_input)) getwd() else dirname(qmd_input)
input_dir  <- file.path(qmd_dir, "_data")

cat("Input directory: ", normalizePath(input_dir, mustWork = FALSE), "\n",
    sep = "")
```

## Load Tidy Summaries

```{r load-tidy}
# Helper: load an .rds if it exists, NULL otherwise.  Reports which
# files were found / missing so the reader sees the state explicitly.
load_if_exists <- function(path, label) {
  if (file.exists(path)) {
    obj <- readRDS(path)
    cat(sprintf("  [OK]      %-40s  (%d rows)\n", basename(path), nrow(obj)))
    obj
  } else {
    cat(sprintf("  [MISSING] %-40s  -- skipped\n", basename(path)))
    NULL
  }
}

cat("Loading tidy selected-subgroup summaries:\n")
tidy_cont <- load_if_exists(
  file.path(input_dir, "selected_subgroups_continuous.rds"),
  "continuous")
tidy_bin  <- load_if_exists(
  file.path(input_dir, "selected_subgroups_binary.rds"),
  "binary")

# Stack what's available.  rbind() works because both data.frames have
# identical column schemas (per extract_selected_subgroups() contract).
tidy_all <- do.call(rbind,
                    Filter(Negate(is.null), list(tidy_cont, tidy_bin)))

if (is.null(tidy_all) || nrow(tidy_all) == 0L) {
  stop("No tidy summaries found in '", input_dir, "'.  Render the ",
       "comparison vignettes first.")
}

cat(sprintf("\nCombined tidy table: %d rows across %d analyses (%s).\n",
            nrow(tidy_all),
            length(unique(tidy_all$analysis)),
            paste(unique(tidy_all$analysis), collapse = ", ")))
```


# Selected Subgroups Across Analyses

The headline table.  One row per `(analysis, sg_focus, selection_rule)`
configuration, with the selected subgroup definition, sample sizes, and
the effect estimate on the natural scale.

::: {.callout-note}
The `effect` column is on each analysis's natural scale (MD for
continuous, OR for binary -- check the `effect_measure` column).
Magnitudes are not directly comparable across analyses; this table is
for **side-by-side inspection of which subgroup each configuration
selected**, not for effect-size comparison.
:::

```{r headline-table}
# Order rows so configurations group together across analyses:
# all "effMaxSG / pareto" rows adjacent, etc.
tidy_all <- tidy_all[order(tidy_all$combo_label,
                           tidy_all$analysis), , drop = FALSE]

# Add a per-row effect-with-measure label so visual comparison across
# rows doesn't accidentally mix scales.
tidy_disp <- tidy_all
tidy_disp$effect_fmt <- ifelse(
  is.na(tidy_disp$effect),
  NA_character_,
  sprintf("%.3f (%s)", tidy_disp$effect, tidy_disp$effect_measure)
)

show_cols <- c("analysis", "combo_label", "sg_focus", "selection_rule",
               "subgroup", "N_H", "N_Hc", "effect_fmt",
               "Pcons", "K", "on_frontier", "n_passed")

gt::gt(tidy_disp[, show_cols, drop = FALSE], groupname_col = "combo_label") |>
  gt::tab_header(
    title    = "Selected Subgroups Across Analyses and Configurations",
    subtitle = sprintf("ACTG175: %s",
                       paste(unique(tidy_all$analysis), collapse = " vs "))
  ) |>
  gt::cols_label(
    analysis       = "Analysis",
    sg_focus       = "sg_focus",
    selection_rule = "Rule",
    subgroup       = "Selected subgroup",
    N_H            = "N(H)",
    N_Hc           = "N(H\u1d9c)",
    effect_fmt     = "Effect (measure)",
    Pcons          = "Pcons",
    K              = "K",
    on_frontier    = "Frontier?",
    n_passed       = "Passed"
  ) |>
  gt::fmt_number(columns = c("Pcons"), decimals = 3) |>
  gt::sub_missing(missing_text = "\u2014")
```


# Cross-Analysis Agreement

Does a given `(sg_focus, selection_rule)` configuration select the
*same* subgroup definition under both outcome scales?  This is a
sanity check on the **selection rule itself**: a rule that produces
wildly different winners under closely-related outcome encodings (here
CD4 decline vs the indicator of no improvement) is less stable than
one that produces similar winners.

```{r agreement-by-config}
# Cast wide: one row per combo_label, columns = analyses, value = subgroup
configs <- unique(tidy_all$combo_label)
analyses <- unique(tidy_all$analysis)

wide <- data.frame(
  combo_label = configs,
  stringsAsFactors = FALSE
)
for (a in analyses) {
  col_sub <- sprintf("subgroup_%s", a)
  col_NH  <- sprintf("N_H_%s",     a)
  col_eff <- sprintf("effect_%s",  a)
  for (i in seq_along(configs)) {
    row <- tidy_all[tidy_all$combo_label == configs[i] &
                    tidy_all$analysis    == a, , drop = FALSE]
    wide[i, col_sub] <- if (nrow(row) > 0L) row$subgroup[1L] else NA_character_
    wide[i, col_NH]  <- if (nrow(row) > 0L) row$N_H[1L]      else NA_integer_
    wide[i, col_eff] <- if (nrow(row) > 0L) {
      sprintf("%.3f (%s)", row$effect[1L], row$effect_measure[1L])
    } else {
      NA_character_
    }
  }
}

# Same-subgroup flag (only meaningful when both analyses have a row)
if (length(analyses) == 2L) {
  s1 <- wide[[sprintf("subgroup_%s", analyses[1L])]]
  s2 <- wide[[sprintf("subgroup_%s", analyses[2L])]]
  wide$same_subgroup <- ifelse(
    is.na(s1) | is.na(s2), NA, s1 == s2
  )
}

gt::gt(wide) |>
  gt::tab_header(
    title    = "Same Configuration, Different Outcomes",
    subtitle = "Does each (sg_focus, selection_rule) cell pick the same subgroup definition under both outcome scales?"
  ) |>
  gt::sub_missing(missing_text = "\u2014")

if ("same_subgroup" %in% names(wide)) {
  agreed <- sum(wide$same_subgroup, na.rm = TRUE)
  total  <- sum(!is.na(wide$same_subgroup))
  cat(sprintf("\nExact-match agreement across analyses: %d / %d configurations (%.0f%%).\n",
              agreed, total, 100 * agreed / max(1L, total)))
}
```


# Per-Subject Agreement (Optional Drill-Down)

The tidy summaries above compare **subgroup definitions** as strings.
A finer check: for a given configuration, do the two analyses classify
the *same subjects* into H?

This requires the full `comparison_<analysis>.rds` bundles (not just
the tidy summaries) because per-subject membership lives in
`comparison$fs[[i]]$df.est$treat.recommend`.  The section below runs
only if both bundles are available.

```{r load-full-bundles}
cmp_cont <- if (file.exists(file.path(input_dir, "comparison_continuous.rds"))) {
  readRDS(file.path(input_dir, "comparison_continuous.rds"))
} else NULL
cmp_bin  <- if (file.exists(file.path(input_dir, "comparison_binary.rds"))) {
  readRDS(file.path(input_dir, "comparison_binary.rds"))
} else NULL

have_full <- !is.null(cmp_cont) && !is.null(cmp_bin)
if (!have_full) {
  cat("Skipping per-subject agreement: ",
      if (is.null(cmp_cont)) "comparison_continuous.rds " else "",
      if (is.null(cmp_bin))  "comparison_binary.rds "     else "",
      "not available.\n", sep = "")
}
```

```{r per-subject-agreement, eval = !is.null(cmp_cont) && !is.null(cmp_bin)}
# Local null-coalesce so this chunk doesn't depend on R >= 4.4 (where
# %||% became base) or on a package internal.
`%||%` <- function(a, b) if (is.null(a) || length(a) == 0L) b else a

# Compute Cohen's kappa between H/Hc assignments for each shared
# combo_label.  Requires both bundles to have at least one common
# configuration AND for both fs objects to have produced a subgroup.

common_combos <- intersect(cmp_cont$combos$label, cmp_bin$combos$label)

if (length(common_combos) == 0L) {
  cat("No shared (sg_focus, selection_rule) configurations between the two analyses.\n")
} else {
  agreement_rows <- lapply(common_combos, function(lbl) {
    i_cont <- which(cmp_cont$combos$label == lbl)[1L]
    i_bin  <- which(cmp_bin$combos$label  == lbl)[1L]
    fs_c   <- cmp_cont$fs[[i_cont]]
    fs_b   <- cmp_bin$fs[[i_bin]]

    # Skip combos where either side has no subgroup
    if (is.null(fs_c) || is.null(fs_b) ||
        is.null(fs_c$sg.harm) || length(fs_c$sg.harm) == 0L ||
        is.null(fs_b$sg.harm) || length(fs_b$sg.harm) == 0L) {
      return(data.frame(
        combo_label = lbl,
        n_subjects  = NA_integer_,
        agreement   = NA_real_,
        kappa       = NA_real_,
        H_both      = NA_integer_,
        H_cont_only = NA_integer_,
        H_bin_only  = NA_integer_,
        Hc_both     = NA_integer_,
        stringsAsFactors = FALSE
      ))
    }

    cmp_pair <- tryCatch(
      compare_fs_subgroups(
        fs1     = fs_c,
        fs2     = fs_b,
        label1  = "continuous",
        label2  = "binary",
        id.name = "id",
        subgroup_notation = "harm",
        title   = sprintf("Continuous vs Binary: %s", lbl)
      ),
      error = function(e) NULL
    )

    if (is.null(cmp_pair)) {
      return(data.frame(
        combo_label = lbl,
        n_subjects  = NA_integer_,
        agreement   = NA_real_,
        kappa       = NA_real_,
        H_both      = NA_integer_,
        H_cont_only = NA_integer_,
        H_bin_only  = NA_integer_,
        Hc_both     = NA_integer_,
        stringsAsFactors = FALSE
      ))
    }

    # Extract cell counts from the 2x2 cross-tab (H/Hc x H/Hc)
    cnc <- cmp_pair$concordance
    data.frame(
      combo_label = lbl,
      n_subjects  = as.integer(cnc$n_total %||% NA_integer_),
      agreement   = as.numeric(cnc$agreement %||% NA_real_),
      kappa       = as.numeric(cnc$kappa     %||% NA_real_),
      H_both      = as.integer(cnc$H_both    %||% NA_integer_),
      H_cont_only = as.integer(cnc$H_only1   %||% NA_integer_),
      H_bin_only  = as.integer(cnc$H_only2   %||% NA_integer_),
      Hc_both     = as.integer(cnc$Hc_both   %||% NA_integer_),
      stringsAsFactors = FALSE
    )
  })

  agreement_tab <- do.call(rbind, agreement_rows)

  gt::gt(agreement_tab) |>
    gt::tab_header(
      title    = "Per-Subject Agreement: Continuous vs Binary",
      subtitle = "Cohen's kappa for H/H\u1d9c classification across the two outcome scales"
    ) |>
    gt::cols_label(
      combo_label = "Configuration",
      n_subjects  = "N",
      agreement   = "Agreement",
      kappa       = "\u03ba",
      H_both      = "H \u2229 H",
      H_cont_only = "H (cont only)",
      H_bin_only  = "H (bin only)",
      Hc_both     = "H\u1d9c \u2229 H\u1d9c"
    ) |>
    gt::fmt_number(columns = c("agreement", "kappa"), decimals = 3) |>
    gt::sub_missing(missing_text = "\u2014")
}
```

::: {.callout-note}
**Interpreting kappa**: \u03ba = 1 is perfect agreement; \u03ba = 0 is
chance-level; values above 0.6 are typically called "substantial" and
above 0.8 "almost perfect".  Low \u03ba with high agreement happens
when both analyses classify most subjects into H\u1d9c (recommend) --
the marginal imbalance inflates raw agreement but kappa adjusts for
it.
:::


# Stability Diagnostics

Two stability views derived directly from the tidy table.

## Configurations Where No Subgroup Was Selected

Rows where the analysis ran but yielded no winner (`subgroup` is NA).
Often a sign that the consistency threshold was too strict for the
candidate pool, or that the selection rule's queue was ordered
unfavorably (see the SUMMARY banner of the source rendering for the
specific reason).

```{r no-subgroup}
no_sub <- tidy_all[is.na(tidy_all$subgroup), , drop = FALSE]
if (nrow(no_sub) == 0L) {
  cat("All configurations selected a subgroup.\n")
} else {
  gt::gt(no_sub[, c("analysis", "combo_label", "sg_focus", "selection_rule",
                    "n_passed", "error"), drop = FALSE]) |>
    gt::tab_header(
      title = sprintf("Configurations with no selected subgroup (%d of %d rows)",
                      nrow(no_sub), nrow(tidy_all))
    ) |>
    gt::sub_missing(missing_text = "\u2014")
}
```

## Frontier Membership

Did the selected subgroup lie on the Pareto frontier?  Under
`selection_rule = "both"` and `"pareto"` this is always TRUE by
construction; under `"neighborhood"` it can be FALSE (the rule picks
in-band largest-N regardless of dominance).

```{r frontier-membership}
fr_tab <- tidy_all[, c("analysis", "combo_label", "sg_focus",
                       "selection_rule", "on_frontier"), drop = FALSE]
gt::gt(fr_tab) |>
  gt::tab_header(title = "Selected subgroup on Pareto frontier?") |>
  gt::sub_missing(missing_text = "\u2014")
```


# Session Information

```{r session-info}
sessionInfo()
```
