---
title: "ACTG175 Continuous Analysis: Applied Operating Characteristics, Self-Contained"
format: html
params:
  draws: 20000
  n_workers: 14
  c2_ratio: 0.8
  seed: 8316951
---

```{r setup}
#| message: false
suppressPackageStartupMessages({
  library(forestsearch)
  library(speff2trial)
  library(gt)
})

# ---------------------------------------------------------------------------
# Output location for this document's results:  <base>/<dirout>/
#   results_dir = NULL -> base is <qmd dir>/_payloads
#   dirout      = NULL -> the .qmd stem
# ---------------------------------------------------------------------------
results_dir <- NULL
dirout      <- NULL

# Everything below is computed in this document: the analysis, the anchored
# truths (the primary Q and its two supersets — the Q_variants knob, defined
# visibly in section 3 once the population frame exists), the candidate
# family, and the operating characteristics.  Nothing is read from disk.
# Re-render at other precisions via the YAML params (draws, n_workers,
# c2_ratio, seed); the archived companion
# analysis_actg175_continuous_oc_evaluation.qmd holds the deep fixed-c2 run.
q_rungs   <- c(0.01, 5, 10, 15, 20, 40, 60, 87 + 11 / 12)   # primary rungs; last = T_obs, gated below
q_shared  <- c(0.01, 5, 10, 20, 40, 87 + 11 / 12)           # rungs run for each Q superset, incl. its own sub-threshold null
c1_ladder <- sort(c(5, 10, 15, 20, 30, 40, 50, 60, 80, 100, 120, 87 + 11 / 12))
c2_vec    <- sort(unique(c(5, params$c2_ratio * c1_ladder)))

fmt  <- function(x, d = 3) formatC(x, format = "f", digits = d)
fmt0 <- function(x) formatC(x, format = "d")
pct  <- function(x, d = 1) paste0(fmt(100 * x, d), "%")
pm   <- function(x, s, d = 4) paste0(fmt(x, d), " ± ", fmt(s, d))
at   <- function(d, c1v, c2v) d[abs(d$c1 - c1v) < 1e-9 & abs(d$c2 - c2v) < 1e-9, ]
qlab <- function(x) fmt(x, ifelse(x %% 1 == 0, 1, 2))
t_doc <- proc.time()[["elapsed"]]
```

## 1. The data

The analysis frame is the compare-all document's: ACTG175 arms 1 (ZDV+ddI)
and 3 (ddI), outcome `y_decline` = CD4 decline from baseline to week 20
(adverse scale), six continuous and six binary candidate confounders.

```{r data-prep}
cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
bin_vars  <- c("hemo", "homo", "drugs", "race", "gender", "symptom")
confounders.name <- c(cont_vars, bin_vars)

actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$id <- seq_len(nrow(actg_df))
actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)
actg_df <- actg_df[!is.na(actg_df$cd420), ]
actg_df$y_decline <- actg_df$cd40 - actg_df$cd420
for (v in bin_vars) actg_df[[v]] <- as.factor(actg_df[[v]])

N <- nrow(actg_df)
stopifnot(N == 1083L)
cat("Analysis N =", N, "\n")

# the unadjusted ITT mean difference — distinct from the covariate-adjusted
# GLM coefficient beta_treat the generator anchors to (section 3)
itt <- coef(summary(lm(y_decline ~ treat, data = actg_df)))["treat", ]
cat(sprintf("unadjusted ITT mean difference = %.3f (SE %.3f)\n",
            itt["Estimate"], itt["Std. Error"]))
```

## 2. The analysis

The search whose operating characteristics this document evaluates is the
fixed-family `maxeffCons` configuration — the compare-all document's
MR-anchor arm.  Multiplier-resampling inference and the lasso/GRF front ends
are mutually exclusive under `subgroup_method = "consistency"` (the family
would be estimated from the same data the search runs on), so the
inferential anchor runs with the front ends off, and the operating
characteristics evaluated here are the operating characteristics of exactly
that procedure.

```{r anchor}
fs_anchor <- forestsearch(
  df.analysis      = actg_df,
  confounders.name = confounders.name,
  outcome.name     = "y_decline",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "continuous",
  effect_measure   = "MD",
  adverse_outcome  = TRUE,
  seedit           = params$seed,
  sg_focus         = "maxeffCons",
  selection_rule   = "neighborhood",
  consistency_method = "resample",
  effect.threshold       = 10,
  consistency.threshold  = 5,
  pconsistency.threshold = 0.90,
  use_twostage     = TRUE,
  conf.cont_jcuts  = list(age = 10, preanti = 10, wtkg = 10,
                          karnof = 10, cd40 = 10, cd80 = 10),
  cut_type         = "default",
  maxk             = 2L,
  n.min            = 60L,
  d0.min           = 10L,
  d1.min           = 10L,
  fs.splits        = 500L,
  use_lasso        = FALSE,
  use_grf          = FALSE,
  use_dina         = FALSE,
  is.RCT           = TRUE,
  parallel_args    = list(plan = "sequential"),
  details          = FALSE,
  quiet            = TRUE
)

H_def  <- paste(fs_anchor$sg.harm, collapse = " & ")
n_H    <- sum(fs_anchor$df.est$treat.recommend == 0L)
win    <- fs_anchor$grp.consistency$out_sg$result[1L, ]
T_obs  <- as.numeric(win$hr)      # MD scale: the hr column holds the MD
p_cons <- as.numeric(win$Pcons)
cat(sprintf("H-hat = %s;  n(H-hat) = %d;  T_obs = %.6f;  p.consistency = %.2f\n",
            H_def, n_H, T_obs, p_cons))

# gates: the anchor must reproduce the compare-all / stage-0 result exactly
stopifnot(setequal(fs_anchor$sg.harm, c("{age <= 37}", "!{cd40 <= 507}")),
          n_H == 66L, abs(T_obs - (87 + 11 / 12)) < 5e-7)
q_rungs[length(q_rungs)] <- T_obs
q_shared[length(q_shared)] <- T_obs
c1_ladder[abs(c1_ladder - (87 + 11 / 12)) < 1e-9] <- T_obs
c2_vec <- sort(unique(c(5, params$c2_ratio * c1_ladder)))
```

The anchor: **Ĥ = `` `r H_def` ``**, n(Ĥ) = `r fmt0(n_H)` of `r fmt0(N)`,
fitted mean difference **T̂~obs~ = `r fmt(T_obs, 6)`** CD4 units,
consistency `r fmt(p_cons, 2)`.

## 3. The anchored truth, and the Q knob

Truths are anchored to the analysis data by a two-point mixture
stylization: the super-population (draws of the covariate rows themselves)
has exactly two treatment-effect values — the fitted background effect on
Q^c^ and a planted effect q on Q — with the baseline outcome surface and
residual scale taken from the baseline GLM fit to the analysis data.
Nothing else about the joint covariate law is modeled.  The primary Q is Ĥ
carried into the generator by its clause mapping (`age = 37`;
`cd40 = list(type = "greater", value = 507)`), and on the identity link
`k_inter(q) = q − beta_treat` — which is Q-independent, since the within-Q
effect is `beta_treat + k_inter` for any Q.

```{r anchored-truth}
fml_base <- as.formula(paste("y_decline ~ treat +",
                             paste(c(bin_vars, cont_vars), collapse = " + ")))
beta_treat <- unname(coef(glm(fml_base, data = actg_df,
                              family = gaussian()))["treat"])
cat(sprintf("baseline GLM beta_treat = %.6f;  k_inter(q) = q + %.6f\n",
            beta_treat, -beta_treat))

Q_primary <- list(age = 37, cd40 = list(type = "greater", value = 507))

dgm_at <- function(q, cuts) generate_glm_dgm(
  data            = actg_df,
  factor_vars     = bin_vars,
  continuous_vars = cont_vars,
  outcome_var     = "y_decline",
  treatment_var   = "treat",
  outcome_type    = "continuous",
  effect_measure  = "MD",
  subgroup_vars   = c("age", "cd40"),
  subgroup_cuts   = cuts,
  model           = "alt",
  k_treat         = 1,
  k_inter         = q - beta_treat,
  adverse_outcome = TRUE,   # inert for a continuous outcome: the negation branch is binary-only
  n_super         = 5000L,
  seed            = params$seed,
  verbose         = FALSE
)

dgm_top <- dgm_at(T_obs, Q_primary)
print(fs_dgm_scale(dgm_top))
```

**The knob.**  The true subgroup is a named list of cutpoint sets — this
list is the document's knob.  The primary is Ĥ's exact clauses; the two
supersets keep the same factors with more permissive thresholds, chosen
from the family's own population cut grid (the J = 10 quantile cuts the
enumeration uses) to land nearest 2× and 3× Ĥ's prevalence.  The truth is
a modeling choice: the primary plants the found rule, the variants ask
what the same analysis means if the truly harmed population is broader
than the sharp rule the search returned.

```{r q-knob}
ds   <- dgm_top$df_super
P_of <- function(cuts) mean(ds$age <= cuts$age & ds$cd40 > cuts$cd40$value)
P0   <- P_of(Q_primary)

# the family's population cut grid: quantile cuts at k/(J+1), k = 1..J = 10
age_grid  <- unname(quantile(ds$age,  (1:10) / 11))
cd40_grid <- unname(quantile(ds$cd40, (1:10) / 11))
cand <- expand.grid(a = age_grid[age_grid > 37], c = cd40_grid[cd40_grid < 507])
cand$P <- mapply(function(a, c) mean(ds$age <= a & ds$cd40 > c), cand$a, cand$c)
pick <- function(target, ok) { d <- cand[ok, ]; d[which.min(abs(d$P - target)), ] }
w <- pick(2 * P0, rep(TRUE, nrow(cand)))
b <- pick(3 * P0, cand$a >= w$a & cand$c <= w$c & cand$P > w$P)

Q_variants <- list(
  primary = Q_primary,
  wider   = list(age = w$a, cd40 = list(type = "greater", value = w$c)),
  broad   = list(age = b$a, cd40 = list(type = "greater", value = b$c)))
prev <- sapply(Q_variants, P_of)
for (v in names(Q_variants)) cat(sprintf(
  "%-8s age <= %-6s & cd40 > %-6s  P = %.4f  (%.2fx primary)\n", v,
  format(Q_variants[[v]]$age), format(Q_variants[[v]]$cd40$value),
  prev[v], prev[v] / P0))

# gate: strict nesting Q_primary < Q_wider < Q_broad on df_super
memb3 <- sapply(Q_variants, function(cu) ds$age <= cu$age & ds$cd40 > cu$cd40$value)
stopifnot(all(memb3[, 1] <= memb3[, 2]), all(memb3[, 2] <= memb3[, 3]),
          sum(memb3[, 1]) < sum(memb3[, 2]), sum(memb3[, 2]) < sum(memb3[, 3]))
```

## 4. The family

The evaluation family is the analyst's own discovery grid at
n = `r fmt0(N)`, enumerated on the super-population from the same
covariates, quantile cuts, and floors the search uses.  It is kept pure —
Q is not forced in.

```{r family}
fs_args <- list(
  confounders.name = confounders.name,
  conf.cont_jcuts  = list(age = 10, preanti = 10, wtkg = 10,
                          karnof = 10, cd40 = 10, cd80 = 10),
  cut_type         = "default",
  cont.cutoff      = 4L,
  maxk             = 2L,
  n.min            = 60L
)
fam <- fs_oc_family_enumerate(dgm_top, fs_args, n = N, max_M = 10000L,
                              verbose = FALSE)
print(fam)
stopifnot(fam$M == 4508L, fam$orientation$s == 1)

# Q's representability: nearest member by Jaccard against H-hat on the
# super-population rows
H_super <- as.logical(dgm_top$df_super$flag_harm)
inter   <- colMeans(fam$memb & H_super)
jac     <- inter / (fam$Pg + mean(H_super) - inter)
i_near  <- which.max(jac)
purity_near <- inter[i_near] / fam$Pg[i_near]
cat(sprintf("nearest member: %s  (purity %.4f, Jaccard %.4f)\n",
            fam$lab[i_near], purity_near, jac[i_near]))
```

Ĥ's `cd40` threshold (507) lies just off the population quantile grid, so
no family member is identical to Ĥ; the nearest member
(`` `r fam$lab[i_near]` ``) matches it with purity
`r fmt(purity_near, 4)`.  This is a feature of the evaluation, not a
defect: the search is scored on the family it actually searches.

## 5. The fundamental call

This one call is the wrapper; everything below is this call swept over
truths and thresholds.  It is evaluated at the analyst's screening floor
c1 = 10 with the scaled consistency floor
c2 = `r fmt(params$c2_ratio, 1)`·c1 = `r fmt(params$c2_ratio * 10, 0)`, on
the top-rung family (planted q = T̂~obs~).

```{r fundamental-call}
pred <- fs_oc_predict(family = fam, n = N,
                      c1 = 10, c2 = params$c2_ratio * 10,
                      consistency_method = "resample", pconsistency = 0.90,
                      draws = params$draws, seed = params$seed)
print(pred)
```

## 6. The evaluation loop

The primary rungs: q ∈ {`r paste(sapply(q_rungs, qlab), collapse = ", ")`}
CD4 units — the sub-threshold-subgroup null (a true Q with zero effect
against the beneficial background), the analyst's c1 floor, two
intermediates, and the observed statistic.  Each superset variant runs
q ∈ {`r paste(sapply(q_shared, qlab), collapse = ", ")`} — each superset
carrying its own sub-threshold rung — twenty (variant, rung) jobs in
all, the ladder refined so every 0.05 consonance crossing is bracketed
within five units.  Each job rebuilds the anchored DGM for its
variant, re-enumerates the family (gated identical in size and
orientation; the Q-referencing fields are what change), and evaluates one
`fs_oc_grid()` — the full
`r fmt0(length(c1_ladder))` × `r fmt0(length(c2_vec))` =
`r fmt0(length(c1_ladder) * length(c2_vec))`-point (c1, c2) crossing on one
draw set of `r format(params$draws, big.mark = ",")`.

```{r evaluation-loop}
jobs <- rbind(data.frame(variant = "primary", q = q_rungs),
              data.frame(variant = "wider",   q = q_shared),
              data.frame(variant = "broad",   q = q_shared))
eval_job <- function(i) {
  v <- jobs$variant[i]
  fam_q <- fs_oc_family_enumerate(dgm_at(jobs$q[i], Q_variants[[v]]), fs_args,
                                  n = N, max_M = 10000L, verbose = FALSE)
  # per-rung truth gate: the planted within-Q effect must equal this job's q,
  # so a mis-set k_inter fails loudly rather than only through its sign
  stopifnot(fam_q$M == 4508L, fam_q$orientation$s == 1,
            abs(abs(fam_q$orientation$m_tau_Q) - jobs$q[i]) < 1e-9)
  g <- fs_oc_grid(family = fam_q, n = N, c1 = c1_ladder, c2 = c2_vec,
                  consistency_method = "resample", pconsistency = 0.90,
                  draws = params$draws, block = 5e4, seed = params$seed)
  cbind(variant = v, q = jobs$q[i], g$table)
}
t0 <- proc.time()[["elapsed"]]
oc <- do.call(rbind, parallel::mclapply(seq_len(nrow(jobs)), eval_job,
                                        mc.cores = params$n_workers))
t_loop <- proc.time()[["elapsed"]] - t0
stopifnot(nrow(oc) == nrow(jobs) * length(c1_ladder) * length(c2_vec))
cat(sprintf("evaluation loop: %.1f s (%.1f min) over %d jobs, %d workers\n",
            t_loop, t_loop / 60, nrow(jobs), params$n_workers))
ocp <- oc[oc$variant == "primary", ]
```

**The no-subgroup null, by construction.**  Under a homogeneous truth
every candidate's true mean *is* the common effect, so the no-subgroup
null is built by direct substitution on the primary family object — the
same object-level substitution the archived document's `se_direct`
sensitivity used — and evaluated over the same (c1, c2) crossing with the
same draws, block, and seed.  It cannot be reached through the generator
at `k_inter = 0`: both region effects are then equal and negative, the
orientation reads s = sign(m~τ~[Q]) = −1, and every candidate's oriented
mean becomes +`r fmt(-beta_treat, 2)`, clearing c1 = 10 trivially
(declaration ≈ 1 — the pathology stage 0 measured on the null branch).

```{r homogeneous-null}
fam_hom <- fam                             # the primary family, already built
fam_hom$beta_g <- rep(beta_treat, fam$M)   # homogeneous truth: no subgroup
g_hom <- fs_oc_grid(family = fam_hom, n = N, c1 = c1_ladder, c2 = c2_vec,
                    consistency_method = "resample", pconsistency = 0.90,
                    draws = params$draws, block = 5e4, seed = params$seed)
# fam_hom inherits PQg/sens_g/spec_g from the primary family, so the
# Q-classification functionals are computed against a Q that does not exist
# under this truth: NA them out (EnH, EbetaH, mass_below use only Pg/beta_g)
g_hom$table[, c("Eppv", "Esens", "Espec", "Enpv")] <- NA_real_
oc <- rbind(oc, cbind(variant = "homogeneous", q = NA_real_, g_hom$table))
```

**The sub-threshold null is not shared across variants.**  At q → 0 a
candidate's true mean is `beta_treat`·(1 − P(Q|g)), which depends on the
planted Q through P(Q|g) — a candidate largely inside Q~broad~ but
outside Q~primary~ has a mean near zero under broad against near
`r fmt(beta_treat, 2)` under primary — so the sub-threshold null is a
property of the planted Q and each variant runs its own
q = `r fmt(q_rungs[1], 2)` rung.  An earlier draft read the primary's
grid as every variant's left calibration endpoint under a shared-null
shortcut whose check compared each family's means to its own
Q-dependent background — an algebraic identity, unable to detect the
between-variant difference — so that shortcut was invalid and is
removed; the measured between-variant differences and each variant's own
declaration rate at (T̂~obs~, `r fmt(params$c2_ratio * 10, 0)`) are:

```{r null-not-shared}
fam0 <- parallel::mclapply(Q_variants, function(cu)
  fs_oc_family_enumerate(dgm_at(q_rungs[1], cu), fs_args, n = N,
                         max_M = 10000L, verbose = FALSE), mc.cores = 3L)
stopifnot(identical(fam0$wider$lab, fam0$primary$lab),
          identical(fam0$broad$lab, fam0$primary$lab))  # member alignment
dev_vs_primary <- sapply(fam0[c("wider", "broad")], function(f)
  c(max = max(abs(f$beta_g - fam0$primary$beta_g)),
    median = median(abs(f$beta_g - fam0$primary$beta_g))))
print(round(dev_vs_primary, 3))
det0 <- sapply(names(Q_variants), function(v) {
  d <- oc[oc$variant == v, ]
  at(d[d$q == q_rungs[1], ], T_obs, params$c2_ratio * 10)$det_rate
})
print(round(det0, 4))
# gate: the sub-threshold null must actually differ between variants;
# if no superset moves any member by > 1 CD4 unit the premise is wrong
stopifnot(max(dev_vs_primary["max", ]) > 1)
```

**Two nulls, bracketing the question.**  The q = `r fmt(q_rungs[1], 2)`
rung is *not* the no-subgroup null: with `k_inter(q) = q − beta_treat` the
within-Q effect is q ≈ 0 while Q^c^ keeps the fitted benefit
`r fmt(beta_treat, 2)` — a true subgroup exists, carrying
`r fmt(-beta_treat, 0)` CD4 units of genuine effect heterogeneity, and
sits far below any threshold the analysis uses.  That is **the
sub-threshold-subgroup null** (a true Q with zero effect against the
beneficial background).  The homogeneous family above is **the
no-subgroup null**.  Both of the analyst's nulls are therefore present,
and they bracket the question — no subgroup at all, and a subgroup real
but below the thresholds; the second is the more adversarial of the two,
its Q-enriched candidates carrying elevated means, and is where the larger
false-declaration rate comes from.

## 7. The diagonal: c2 = `r fmt(params$c2_ratio, 1)`·c1

The headline policy scales the consistency floor with the screening floor.
All views in this section sit on the diagonal
c2 = `r fmt(params$c2_ratio, 1)`·c1.

### 7.1 Type-I under the two nulls

```{r type1-diagonal}
diag_of <- function(qv) do.call(rbind, lapply(c1_ladder, function(c1v)
  at(ocp[ocp$q == qv, ], c1v, params$c2_ratio * c1v)))
tI  <- diag_of(q_rungs[1])
och <- oc[oc$variant == "homogeneous", ]
tIh <- do.call(rbind, lapply(c1_ladder, function(c1v)
  at(och, c1v, params$c2_ratio * c1v)))

type1_at_10 <- at(tI, 10, params$c2_ratio * 10)
hom_at_10   <- at(tIh, 10, params$c2_ratio * 10)
sub_at_5    <- at(ocp[ocp$q == q_rungs[1], ], 10, 5)
hom_at_5    <- at(och, 10, 5)
stopifnot(is.finite(type1_at_10$det_rate), is.finite(hom_at_10$det_rate))

c1_05_diag <- suppressWarnings(min(tI$c1[tI$det_rate <= 0.05]))
c1_10_diag <- suppressWarnings(min(tI$c1[tI$det_rate <= 0.10]))
c1_05_hom  <- suppressWarnings(min(tIh$c1[tIh$det_rate <= 0.05]))
c1_10_hom  <- suppressWarnings(min(tIh$c1[tIh$det_rate <= 0.10]))
stopifnot(is.finite(c1_05_diag), is.finite(c1_10_diag),
          is.finite(c1_05_hom), is.finite(c1_10_hom))

# gate: the no-subgroup null can only declare less often than the
# sub-threshold null, whose Q-enriched candidates carry elevated means —
# c1_05^hom <= c1_05^sub-threshold and pointwise over the whole crossing
mrg <- merge(och[, c("c1", "c2", "det_rate")],
             ocp[ocp$q == q_rungs[1], c("c1", "c2", "det_rate")],
             by = c("c1", "c2"), suffixes = c("_hom", "_sub"))
stopifnot(nrow(mrg) == length(c1_ladder) * length(c2_vec),
          c1_05_hom <= c1_05_diag,
          all(mrg$det_rate_hom <= mrg$det_rate_sub + 1e-9))

# each superset's own sub-threshold null, on the diagonal
tIv <- lapply(c(wider = "wider", broad = "broad"), function(v) {
  d0 <- oc[oc$variant == v & oc$q == q_rungs[1], ]
  do.call(rbind, lapply(c1_ladder, function(c1v)
    at(d0, c1v, params$c2_ratio * c1v)))
})
sub_w_10 <- at(tIv$wider, 10, params$c2_ratio * 10)
sub_b_10 <- at(tIv$broad, 10, params$c2_ratio * 10)

# the no-subgroup null's tail probability at the observed statistic, read
# at the analyst's operating c2 as in section 7.2 (previously unread)
hom_at_Tobs <- at(och, T_obs, params$c2_ratio * 10)
```

At the analyst's screening floor c1 = 10 (diagonal c2 =
`r fmt(params$c2_ratio * 10, 0)`), the family-wise false-declaration rate
is **`r fmt(type1_at_10$det_rate, 5)`**
(MC SE `r fmt(type1_at_10$det_rate_se, 5)`) under the
sub-threshold-subgroup null and **`r fmt(hom_at_10$det_rate, 5)`**
(MC SE `r fmt(hom_at_10$det_rate_se, 5)`) under the no-subgroup null; at
the fixed comparator (10, 5) the two read
`r pm(sub_at_5$det_rate, sub_at_5$det_rate_se)` and
`r pm(hom_at_5$det_rate, hom_at_5$det_rate_se)`.  Read at the ladder's own
resolution, the smallest ladder c1 holding the rate at or below 0.05 on
the diagonal is **c1~05~^diag^ = `r fmt0(c1_05_diag)`**
(rate `r pm(at(tI, c1_05_diag, params$c2_ratio * c1_05_diag)$det_rate,
            at(tI, c1_05_diag, params$c2_ratio * c1_05_diag)$det_rate_se)`)
under the sub-threshold null against
**c1~05~^hom^ = `r fmt0(c1_05_hom)`**
(rate `r pm(at(tIh, c1_05_hom, params$c2_ratio * c1_05_hom)$det_rate,
            at(tIh, c1_05_hom, params$c2_ratio * c1_05_hom)$det_rate_se)`)
under the no-subgroup null; at or below 0.10,
**c1~10~^diag^ = `r fmt0(c1_10_diag)`** against
**c1~10~^hom^ = `r fmt0(c1_10_hom)`**.  For the no-subgroup null only the
declaration event is reported: with no true Q, `PQg`, sensitivity,
specificity, PPV, and every other classification functional are
undefined, so the homogeneous rows carry `det_rate` alone.  Each
superset's own sub-threshold null carries its own type-I — at the
analyst's (10, `r fmt(params$c2_ratio * 10, 0)`): wider
`r pm(sub_w_10$det_rate, sub_w_10$det_rate_se)`, broad
`r pm(sub_b_10$det_rate, sub_b_10$det_rate_se)`, with the full diagonal
curves in the payload — and the headline above is the primary's.

```{r type1-diagonal-curve}
#| fig-width: 7
#| fig-height: 4.5
#| fig-cap: "Type-I error of the search under the two nulls along the diagonal c2 = 0.8·c1, at ladder resolution: the sub-threshold-subgroup null (filled) and the no-subgroup null (open, dashed). Guides: the analyst's c1 = 10, and the sub-threshold c1_05/c1_10 at the 0.05/0.10 lines."
op <- par(mar = c(4.2, 4.2, 1, 1))
plot(tI$c1, tI$det_rate, type = "b", pch = 16, lwd = 2,
     ylim = range(0, tI$det_rate, tIh$det_rate),
     xlab = "c1 (CD4 units, mean difference); c2 = 0.8·c1",
     ylab = "P(any declaration | null)")
lines(tIh$c1, tIh$det_rate, type = "b", pch = 1, lty = 2, lwd = 2)
abline(h = c(0.05, 0.10), lty = 3, col = "grey40")
abline(v = 10, lty = 2, col = "firebrick")
abline(v = c(c1_05_diag, c1_10_diag), lty = 2, col = "steelblue")
text(10, max(tI$det_rate), "analyst c1", pos = 4, cex = 0.8, col = "firebrick")
text(c1_05_diag, 0.05, expression(c1[0.05]^diag), pos = 4, cex = 0.8,
     col = "steelblue")
legend("topright", legend = c("sub-threshold-subgroup null",
                              "no-subgroup null"),
       pch = c(16, 1), lty = c(1, 2), lwd = 2, cex = 0.8, bg = "white")
par(op)
```

### 7.2 The effect ladder

Per rung: the declaration rate at the analyst's diagonal gate
(c1 = 10, c2 = `r fmt(params$c2_ratio * 10, 0)`), the power at the
diagonal fixed-type-I thresholds, and the calibration columns
P(T ≥ T̂~obs~ | q) — read at c1 = T̂~obs~ with the analyst's operating
consistency floor c2 = `r fmt(params$c2_ratio * 10, 0)` and, alongside,
the archived comparator c2 = 5.  c2 sets candidate *eligibility*, and the
calibration interrogates the analysis as actually conducted, so it is
read at the analyst's own consistency floor while type-I and power are
read on the policy diagonal.

```{r ladder-table}
lad <- do.call(rbind, lapply(q_rungs, function(qv) {
  d  <- diag_of(qv)
  dq <- ocp[ocp$q == qv, ]
  data.frame(
    q      = qv,
    det10  = at(d, 10, params$c2_ratio * 10)$det_rate,
    det10s = at(d, 10, params$c2_ratio * 10)$det_rate_se,
    p05    = at(d, c1_05_diag, params$c2_ratio * c1_05_diag)$det_rate,
    p05s   = at(d, c1_05_diag, params$c2_ratio * c1_05_diag)$det_rate_se,
    p10    = at(d, c1_10_diag, params$c2_ratio * c1_10_diag)$det_rate,
    p10s   = at(d, c1_10_diag, params$c2_ratio * c1_10_diag)$det_rate_se,
    pT     = at(dq, T_obs, params$c2_ratio * 10)$det_rate,
    pTs    = at(dq, T_obs, params$c2_ratio * 10)$det_rate_se,
    pT5    = at(dq, T_obs, 5)$det_rate,
    pT5s   = at(dq, T_obs, 5)$det_rate_se)
}))

# gate: power columns monotone in q within 2 MC SEs
mono_ok <- function(p, s) all(diff(p) >= -2 * sqrt(s[-1]^2 + s[-length(s)]^2))
stopifnot(mono_ok(lad$det10, lad$det10s), mono_ok(lad$p05, lad$p05s),
          mono_ok(lad$p10, lad$p10s))

gt(data.frame(
  q     = sapply(lad$q, qlab),
  det10 = pm(lad$det10, lad$det10s),
  p05   = pm(lad$p05, lad$p05s),
  p10   = pm(lad$p10, lad$p10s),
  pT    = pm(lad$pT, lad$pTs),
  pT5   = pm(lad$pT5, lad$pT5s))) |>
  cols_label(q = "q (planted MD)",
             det10 = paste0("rate at (10, ", fmt(params$c2_ratio * 10, 0), ")"),
             p05 = html(paste0("power at c1<sub>05</sub><sup>diag</sup> = ",
                               fmt0(c1_05_diag))),
             p10 = html(paste0("power at c1<sub>10</sub><sup>diag</sup> = ",
                               fmt0(c1_10_diag))),
             pT = html(paste0("P(T ≥ T̂<sub>obs</sub> | q), c2 = ",
                              fmt(params$c2_ratio * 10, 0))),
             pT5 = html("P(T ≥ T̂<sub>obs</sub> | q), c2 = 5")) |>
  tab_source_note(md(paste0(
    "Rate and power columns on the diagonal c2 = ", fmt(params$c2_ratio, 1),
    "·c1; the calibration columns at c1 = T̂~obs~ with c2 at the analyst's ",
    "operating floor (", fmt(params$c2_ratio * 10, 0), ") and the archived ",
    "comparator (5); one ", format(params$draws, big.mark = ","),
    "-draw set per rung; ± is the Monte-Carlo SE.")))
```

### 7.3 What is declared at (c1 = 10, c2 = `r fmt(params$c2_ratio * 10, 0)`)

Conditional on declaring at the analyst's diagonal gate: the expected size
of the declared subgroup, its expected overlap quality against Q, the
expected true oriented effect on the declared rule, the naive selection
bias, and the selection mass on rules whose true mean is below c1.

```{r declared-table}
dec <- do.call(rbind, lapply(q_rungs, function(qv)
  at(ocp[ocp$q == qv, ], 10, params$c2_ratio * 10)))

# gate: the sub-threshold row's declared rules carry an oriented-NEGATIVE
# expected true effect (the honest false-positive reading)
stopifnot(dec$EbetaH[1] < 0)

gt(data.frame(
  q     = sapply(dec$q, qlab),
  EnH   = fmt(dec$EnH, 1),
  Eppv  = fmt(dec$Eppv, 3),
  Esens = fmt(dec$Esens, 3),
  Espec = fmt(dec$Espec, 3),
  Ebeta = fmt(dec$EbetaH, 2),
  naive = fmt(dec$Enaive_bias, 2),
  mb    = fmt(dec$mass_below, 3))) |>
  cols_label(q = "q", EnH = md("E\\|Ĥ\\|"), Eppv = "E[PPV]",
             Esens = "E[sens]", Espec = "E[spec]", Ebeta = md("E[β(Ĥ)]"),
             naive = "naive bias", mb = "mass below c1") |>
  tab_source_note(md(paste0(
    "All conditional on declaration at (c1 = 10, c2 = ",
    fmt(params$c2_ratio * 10, 0), "). The archived companion carries ",
    "delta-method MC SEs for these functionals at 2e5 draws.")))
```

The sub-threshold row is the honest false-positive reading: when the
sub-threshold-subgroup null declares (rate `r fmt(dec$det_rate[1], 4)`), the declared
rule's expected true oriented effect is `r fmt(dec$EbetaH[1], 2)` — a
*benefit*, not a harm, pulled up from the pure background
`r fmt(beta_treat, 2)` only by the selection mass on
partially-Q-overlapping rules (E[PPV] = `r fmt(dec$Eppv[1], 3)`).

## 8. The c2 = 5 comparator

The analyst's original gate held c2 fixed at 5 against c1 = 10 — and
against every other c1, which is the configuration the archived deep run
evaluates.  A fixed c2 = 5 against c1 = 50 makes no sense as practice — the
consistency floor should grow with the claim being screened — so the fixed
row is kept here only as the liberal comparator reconciling with the
archived run.

```{r comparator-table}
cmp <- do.call(rbind, lapply(q_rungs, function(qv)
  at(ocp[ocp$q == qv, ], 10, 5)))
gt(data.frame(
  q     = sapply(cmp$q, qlab),
  det   = pm(cmp$det_rate, cmp$det_rate_se),
  EnH   = fmt(cmp$EnH, 1),
  Eppv  = fmt(cmp$Eppv, 3),
  Ebeta = fmt(cmp$EbetaH, 2))) |>
  cols_label(q = "q", det = "rate at (10, 5)", EnH = md("E\\|Ĥ\\|"),
             Eppv = "E[PPV]", Ebeta = md("E[β(Ĥ)]")) |>
  tab_source_note("The analyst's original fixed-c2 gate, at this document's draw count.")
```

The sub-threshold-null rate at (10, 5) here is
`r pm(cmp$det_rate[1], cmp$det_rate_se[1])`, against the archived deep
run's 0.5850 ± 0.0011 at 2e5 draws — the two agree to Monte-Carlo
precision, and remaining differences across the tables are MC precision
and ladder-resolution effects, not configuration.

## 9. The type-I surface

The false-declaration rate under the sub-threshold-subgroup null over the
full (c1, c2) crossing, with the fixed c2 = 5 row and the
c2 = `r fmt(params$c2_ratio, 1)`·c1 diagonal marked.

```{r type1-surface}
#| fig-width: 7
#| fig-height: 5.5
#| fig-cap: "Type-I error of the search under the sub-threshold-subgroup null over the (c1, c2) crossing. White contour lines at 0.05 and 0.10; dashed lines mark the fixed c2 = 5 comparator row and the c2 = 0.8·c1 diagonal."
z <- ocp[ocp$q == q_rungs[1], ]
zm <- matrix(NA_real_, length(c1_ladder), length(c2_vec))
for (i in seq_along(c1_ladder)) for (j in seq_along(c2_vec))
  zm[i, j] <- at(z, c1_ladder[i], c2_vec[j])$det_rate
op <- par(mar = c(4.2, 4.2, 1.5, 1))
image(c1_ladder, c2_vec, zm, col = hcl.colors(24, "YlOrRd", rev = TRUE),
      xlab = "c1 (CD4 units, mean difference)", ylab = "c2",
      main = "P(any declaration | sub-threshold null)")
contour(c1_ladder, c2_vec, zm, levels = c(0.05, 0.10), add = TRUE,
        col = "white", lwd = 2, labcex = 0.8)
abline(h = 5, lty = 2, col = "grey25")
abline(a = 0, b = params$c2_ratio, lty = 2, col = "grey25")
text(max(c1_ladder) * 0.75, 5, "c2 = 5", pos = 3, cex = 0.8, col = "grey25")
text(max(c1_ladder) * 0.8, params$c2_ratio * max(c1_ladder) * 0.8,
     "c2 = 0.8·c1", pos = 3, srt = 33, cex = 0.8, col = "grey25")
par(op)
```

```{r surface-reading}
tI5 <- do.call(rbind, lapply(c1_ladder, function(c1v) at(z, c1v, 5)))
c1_05_fix <- suppressWarnings(min(tI5$c1[tI5$det_rate <= 0.05]))
```

The picture is the point of the scaled policy: along the diagonal the
consistency screen binds (for noisy candidates
`r fmt(params$c2_ratio, 1)`·c1 + z~p~·se~g~ overtakes c1), so 5% type-I
control arrives at c1~05~^diag^ = `r fmt0(c1_05_diag)` on the diagonal
versus `r if (is.finite(c1_05_fix)) fmt0(c1_05_fix) else "beyond the ladder"`
on the fixed c2 = 5 row at the same ladder resolution.

## 10. Turning the true-subgroup knob

The primary evaluation plants Ĥ itself.  This section turns the knob of
section 3: the same analysis, the same family, the same thresholds, scored
against each variant's own Q — power at the diagonal fixed-type-I
threshold for the common rungs (q = 20, 40, T̂~obs~), the declared rule's overlap quality at the
analyst's (10, `r fmt(params$c2_ratio * 10, 0)`) on the observed-statistic
rung, and the calibration crossings, read as in section 7.2 at
c1 = T̂~obs~ with the analyst's operating c2 =
`r fmt(params$c2_ratio * 10, 0)` and the archived comparator c2 = 5
alongside.  Each variant's calibration curve is anchored by its own
sub-threshold null rung (section 6).

```{r knob-table}
cal <- lapply(names(Q_variants), function(v) {
  qs <- if (v == "primary") q_rungs else q_shared
  read_pT <- function(c2v) sapply(qs, function(qv) {
    d <- oc[oc$variant == v, ]   # every rung is the variant's own, incl. 0.01
    at(d[d$q == qv, ], T_obs, c2v)$det_rate
  })
  list(qs = qs, pT = read_pT(params$c2_ratio * 10), pT5 = read_pT(5))
})
names(cal) <- names(Q_variants)

cross <- function(qs, pT, lev) {
  i <- which(pT >= lev)[1]
  if (is.na(i)) NA_real_
  else if (i == 1) qs[1]
  else qs[i - 1] + (lev - pT[i - 1]) * (qs[i] - qs[i - 1]) / (pT[i] - pT[i - 1])
}
knob <- do.call(rbind, lapply(names(Q_variants), function(v) {
  d   <- oc[oc$variant == v, ]
  pow <- sapply(c(20, 40, T_obs), function(qv)
    at(d[d$q == qv, ], c1_05_diag, params$c2_ratio * c1_05_diag)$det_rate)
  top <- at(d[abs(d$q - T_obs) < 1e-9, ], 10, params$c2_ratio * 10)
  data.frame(variant = v,
             def   = sprintf("age ≤ %s & cd40 > %s",
                             format(Q_variants[[v]]$age),
                             format(Q_variants[[v]]$cd40$value)),
             P     = prev[v], ratio = prev[v] / prev["primary"],
             p20   = pow[1], p40 = pow[2], ptop = pow[3],
             ppv   = top$Eppv, sens = top$Esens,
             q05   = cross(cal[[v]]$qs, cal[[v]]$pT, 0.05),
             q50   = cross(cal[[v]]$qs, cal[[v]]$pT, 0.50),
             q05c  = cross(cal[[v]]$qs, cal[[v]]$pT5, 0.05),
             q50c  = cross(cal[[v]]$qs, cal[[v]]$pT5, 0.50))
}))

# report, not a gate: with rungs near the 0.05 level and MC SE ~ 0.004 a
# calibration column can be non-monotone in q by chance, which would make
# cross()'s first-crossing read noisy — check within 2 MC SEs and print
cal_se <- lapply(names(Q_variants), function(v) {
  qs <- if (v == "primary") q_rungs else q_shared
  sapply(qs, function(qv) {
    d <- oc[oc$variant == v, ]
    at(d[d$q == qv, ], T_obs, params$c2_ratio * 10)$det_rate_se
  })
})
names(cal_se) <- names(Q_variants)
mono_cal <- sapply(names(Q_variants), function(v)
  mono_ok(cal[[v]]$pT, cal_se[[v]]))
cat("calibration column non-decreasing in q within 2 MC SEs:\n")
print(mono_cal)

qfmt <- function(x) ifelse(is.na(x), "beyond the ladder",
                    ifelse(x <= q_rungs[1], paste0("≤ ", qlab(q_rungs[1])),
                           fmt(x, 1)))
gt(data.frame(
  v    = knob$variant,
  def  = knob$def,
  P    = paste0(fmt(knob$P, 4), " (", fmt(knob$ratio, 2), "×)"),
  p20  = fmt(knob$p20, 4), p40 = fmt(knob$p40, 4), ptop = fmt(knob$ptop, 4),
  ppv  = fmt(knob$ppv, 3), sens = fmt(knob$sens, 3),
  q05  = qfmt(knob$q05),
  q05c = qfmt(knob$q05c),
  q50  = qfmt(knob$q50),
  q50c = qfmt(knob$q50c))) |>
  cols_label(v = "Q variant", def = "definition", P = "P(Q)",
             p20 = "power, q = 20", p40 = "power, q = 40",
             ptop = html("power, q = T̂<sub>obs</sub>"),
             ppv = "E[PPV]", sens = "E[sens]",
             q05 = paste0("q at 0.05 (c2 = ", fmt(params$c2_ratio * 10, 0), ")"),
             q05c = "q at 0.05 (c2 = 5)",
             q50 = paste0("q at 0.5 (c2 = ", fmt(params$c2_ratio * 10, 0), ")"),
             q50c = "q at 0.5 (c2 = 5)") |>
  tab_source_note(md(paste0(
    "Power columns at the diagonal (c1 = ", fmt0(c1_05_diag), ", c2 = ",
    fmt(params$c2_ratio * c1_05_diag, 1), "); E[PPV]/E[sens] scored against ",
    "the variant's own Q, conditional on declaration at (10, ",
    fmt(params$c2_ratio * 10, 0), ") on the q = T̂~obs~ rung; the last four ",
    "columns are the q where P(T ≥ T̂~obs~ | q, Q), read at c1 = T̂~obs~ ",
    "and the stated c2, crosses 0.05 / 0.5, interpolated linearly at ladder ",
    "resolution with each variant's own sub-threshold null point.")))
```

`r if (all(mono_cal)) "Each variant's calibration column is non-decreasing in q within 2 MC SEs, so the crossing reads are clean." else paste0("Note: the calibration column is non-monotone in q beyond 2 MC SEs for the ", paste(names(mono_cal)[!mono_cal], collapse = " and "), " variant(s); those crossings are first crossings of a noisy column and should be read with that in mind.")`

```{r knob-figure}
#| fig-width: 7
#| fig-height: 4.5
#| fig-cap: "Calibration curves P(T ≥ T̂_obs | q, Q) for the three Q-variants at the analyst's operating c2 = 8 (filled, solid) and the archived comparator c2 = 5 (open, dashed), each anchored by its own sub-threshold null point. Dotted guides at 0.05 and 0.5; the vertical dashed line marks q = T̂_obs."
cols3 <- c(primary = "grey20", wider = "steelblue", broad = "firebrick")
op <- par(mar = c(4.2, 4.2, 1, 1))
plot(NA, xlim = c(0, max(q_rungs)), ylim = c(0, 1),
     xlab = "q: planted effect on Q (CD4 units, mean difference)",
     ylab = expression(P(T >= hat(T)[obs] ~ "|" ~ q, Q)))
abline(h = c(0.05, 0.5), lty = 3, col = "grey40")
abline(v = T_obs, lty = 2, col = "grey40")
for (v in names(cal)) {
  lines(cal[[v]]$qs, cal[[v]]$pT, type = "b", pch = 16, lwd = 1.6,
        col = cols3[v])
  lines(cal[[v]]$qs, cal[[v]]$pT5, type = "b", pch = 1, lty = 2, lwd = 1.2,
        col = cols3[v])
}
legend("bottomright",
       legend = c(sprintf("%s (P = %.3f)", names(cal), prev),
                  paste0("filled/solid: c2 = ", fmt(params$c2_ratio * 10, 0)),
                  "open/dashed: c2 = 5"),
       col = c(cols3, "grey40", "grey40"), lwd = c(rep(1.6, 3), NA, NA),
       pch = c(rep(16, 3), NA, NA), cex = 0.8, bg = "white")
par(op)
```

The reading this section exists for: **the observed statistic constrains a
trade between breadth and severity.**  At the analyst's operating gate
(c2 = `r fmt(params$c2_ratio * 10, 0)`), under the primary Q the
calibration curve crosses 0.5 at q ≈ `r qfmt(knob$q50[1])` CD4 units of
planted harm; under the `r fmt(knob$ratio[2], 1)`×-prevalence wider Q it
crosses at q ≈ `r qfmt(knob$q50[2])`, and under the
`r fmt(knob$ratio[3], 1)`×-prevalence broad Q at
q ≈ `r qfmt(knob$q50[3])` — a narrow Q needs a large planted harm to make
T̂~obs~ = `r fmt(T_obs, 1)` unsurprising, a broad Q a smaller one.  The
c2 = 5 comparator columns should sit close to the c2 =
`r fmt(params$c2_ratio * 10, 0)` columns, and the archived deep run
(fixed c2 = 5, 2×10^5^ draws) put the primary 0.05 crossing at q ≈ 7.5.  "What harm is consistent with
what we saw" has no single answer without fixing who is harmed; the table
says what each choice implies.

## 11. Reading and limitations

- **Three different questions.**  The declaration event (sections 7.1–7.2),
  the declared rule's quality and magnitude (7.3), and the calibration
  columns for the observed statistic (7.2, last two columns) answer
  different questions; none is a posterior, and the calibration columns
  are consonance statements inside the anchored model only.  The c2
  policy applies to the first two — type-I and power are threshold-policy
  quantities read on the diagonal — and not to the calibration read-off,
  which is taken at the analyst's operating consistency floor.
- **Two nulls.**  The no-subgroup null (section 6, by substitution) and
  the sub-threshold-subgroup null (the q = `r fmt(q_rungs[1], 2)` rung)
  bracket the null question; the generator's `k_inter = 0` route to the
  first is unusable under the 0.3.2 orientation (every oriented mean
  flips to +`r fmt(-beta_treat, 2)`), a recorded limitation of the
  orientation at a homogeneous truth, documented here in prose and not
  patched.
- **The mixture stylization.**  Truths are a two-point effect mixture on
  the data's own covariate rows with a GLM baseline surface; real
  heterogeneity is not two-valued.
- **Q is off the grid.**  No family member is identical to Ĥ (nearest
  purity `r fmt(purity_near, 4)`); every "power" number is the power to
  declare *some* rule.
- **Fixed-family conditioning.**  The evaluation conditions on the
  no-front-end family; the front-ends-on arms of the compare-all document
  search a data-dependent family, a different estimand.  The cut locations
  are quantile-derived and held at their population values.
- **The E|Ĥ| machinery offset.**  In the verification fixtures the
  wrapper's expected declared-subgroup size ran about two subjects from
  the measured search's; assume that order of offset here.
- **Precision.**  All rates carry MC SE ≈ √(p(1−p)/draws) ≈ 0.004 at the
  default `r format(params$draws, big.mark = ",")` draws; the archived
  companion holds the fixed-c2 analysis at 2e5 draws.

## 12. Interpretation: what the analysis may claim

The analysis found **Ĥ = `` `r H_def` ``** — n(Ĥ) =
`r fmt0(n_H)` of `r fmt0(N)`, fitted mean difference
T̂~obs~ = `r fmt(T_obs, 1)` CD4 units, consistency `r fmt(p_cons, 2)` —
against an unadjusted ITT mean difference of
`r fmt(itt["Estimate"], 1)` CD4 units (SE `r fmt(itt["Std. Error"], 1)`),
a benefit on average.  (The covariate-adjusted GLM coefficient
`beta_treat` = `r fmt(beta_treat, 1)`, which the generator anchors to, is
a related but distinct quantity; the two are not conflated here.)  The
sections above ask what the procedure as configured does under truths
anchored to this data; this section says what may therefore be claimed.
The caveats governing all of it are section 11's and are not repeated
here.

### 12.1 The declaration event is close to uninformative

At (c1 = 10, c2 = `r fmt(params$c2_ratio * 10, 0)`) the search declares
*some* harm subgroup with probability `r fmt(hom_at_10$det_rate, 3)`
under the no-subgroup null and `r fmt(type1_at_10$det_rate, 3)` under
the sub-threshold-subgroup null; at the fixed (10, 5) gate the latter is
`r fmt(sub_at_5$det_rate, 3)`.  Two consequences.  First, "ForestSearch
identified a subgroup" carries almost no evidential weight at this
configuration.  Second, because the two nulls differ by only
`r fmt(type1_at_10$det_rate - hom_at_10$det_rate, 3)`, the rate is
driven by multiplicity over M = `r fmt0(fam$M)` correlated candidates
and a liberal threshold, not by a lurking weak subgroup.  The thresholds
are clinical-relevance floors, not error controls.

### 12.2 The observed magnitude — a range, not a p-value

```{r interp-tail-table}
tail_tab <- data.frame(
  truth = c("no subgroup (homogeneous)",
            "Q = H-hat (primary), zero effect",
            "Q = wider, zero effect",
            "Q = broad, zero effect"),
  rate  = c(hom_at_Tobs$det_rate, unname(det0)),
  se    = c(hom_at_Tobs$det_rate_se,
            sapply(names(Q_variants), function(v) {
              d <- oc[oc$variant == v, ]
              at(d[d$q == q_rungs[1], ], T_obs,
                 params$c2_ratio * 10)$det_rate_se
            })))
gt(data.frame(truth = tail_tab$truth, p = pm(tail_tab$rate, tail_tab$se))) |>
  cols_label(truth = "truth", p = html("P(T ≥ T̂<sub>obs</sub>)")) |>
  tab_source_note(md(paste0(
    "Read at (c1 = T̂~obs~, c2 = ", fmt(params$c2_ratio * 10, 0),
    ") as in section 7.2; ± is the Monte-Carlo SE.")))
```

This is the selection-honest tail probability of the observed statistic,
computed against the distribution of the search's own maximum, so
multiplicity is already inside it.  The answer is a **range** — from
`r fmt(tail_tab$rate[1], 4)` with no subgroup at all to
`r fmt(tail_tab$rate[4], 4)` under the broad reading — that depends on
how broad the truly affected population may be, and under the broad
reading the observation is not surprising even at zero harm.  As an
external cross-check (computed in the archived
`analysis_actg175_continuous_oc_evaluation` deep run, not here), the
primary row's value at 2×10^5^ draws was 0.041.  The observed magnitude
is suggestive, not decisive, under every null in the family.

### 12.3 The magnitude of the harm is weakly identified

The 0.05 consonance crossings of section 10 put the smallest planted
harm consistent with T̂~obs~ at q ≈ `r qfmt(knob$q05[1])` CD4 units if
the harmed population is exactly Ĥ, but q ≈ `r qfmt(knob$q05[2])` under
the wider reading and `r qfmt(knob$q05[3])` under the broad one —
collapsing toward the first rung — against the naive
T̂~obs~ = `r fmt(T_obs, 1)` and the analysis's own floor c1 = 10.  If
the harmed population is not exactly the rule the search returned, the
data are consistent with arbitrarily small harm.  **Do not quote
T̂~obs~ as an effect size**: the de-biased (multiplier-resampling /
bootstrap) estimates are the estimation-side response, and this curve is
the design-side statement of why they are needed.

### 12.4 What error control costs

Holding the family-wise false-declaration rate at 0.05 requires
c1 = `r fmt0(c1_05_diag)` on the scaled diagonal
(c1 = `r fmt0(c1_05_hom)` against the no-subgroup null alone, and
`r if (is.finite(c1_05_fix)) fmt0(c1_05_fix) else "beyond the ladder"`
on the fixed c2 = 5 row) — thresholds far above the clinical floor —
and at that gate power is `r fmt(lad$p05[nrow(lad)], 2)` even at the
observed-statistic rung and lower at every smaller planted effect
(section 7.2).  No configuration on this ladder is a test: strict
control costs most of the power at every planted effect examined, so
the finding is hypothesis-generating and needs prespecified
confirmation — which this same machinery can size.

### 12.5 Breadth versus severity

For T̂~obs~ to be *typical* (the 0.5 crossing), a subgroup exactly Ĥ
must be harmed by q ≈ `r qfmt(knob$q50[1])` CD4 units, the wider
population by q ≈ `r qfmt(knob$q50[2])`, the broad one by
q ≈ `r qfmt(knob$q50[3])`; meanwhile top-rung power rises
`r fmt(knob$ptop[1], 2)` → `r fmt(knob$ptop[3], 2)` across the variants
as the declared rule's sensitivity to the true Q falls
`r fmt(knob$sens[1], 2)` → `r fmt(knob$sens[3], 2)`.  A narrow true
subgroup must be severely harmed to make T̂~obs~ typical; a broad one
need only be mildly harmed — section 10's table carries the full
statement.

### 12.6 A draft results paragraph

*A draft for the analyst to adapt — not a conclusion of this document:*

> ForestSearch, configured with clinical-relevance floors rather than
> error controls, identified Ĥ = `r H_def`
> (n = `r fmt0(n_H)` of `r fmt0(N)`) with fitted mean difference
> `r fmt(T_obs, 1)` CD4 units; since the procedure declares some
> subgroup with probability
> `r fmt(hom_at_10$det_rate, 2)`–`r fmt(type1_at_10$det_rate, 2)` under
> nulls anchored to these data, the identification is
> hypothesis-generating.  The observed statistic's
> selection-honest tail probability ranges from
> `r fmt(hom_at_Tobs$det_rate, 4)` (no subgroup) to
> `r fmt(max(det0), 3)` (a broad mildly-affected population), and the
> harm magnitude consistent with it at the 0.05 consonance level ranges
> from essentially zero if the affected population is the broad variant
> to `r qfmt(knob$q05[1])` CD4 units if it is Ĥ itself, so the fitted
> magnitude should not be quoted as an effect estimate.  A
> prespecified confirmation in independent data — sized with the same
> operating-characteristic machinery — is required before any clinical
> claim.

Three external facts cited above and in section 11, named as external: the
archived deep run's calibration values
(`analysis_actg175_continuous_oc_evaluation`), and the ~2-subject
E\|Ĥ\| machinery offset (the verification fixtures, section 11); the
archived `se_direct` sensitivity's ≤ 1.4-point movement of the
calibration column (same archived document) bounds the se-model choice.

```{r doc-clock}
cat(sprintf("document compute wall-clock so far: %.1f min\n",
            (proc.time()[["elapsed"]] - t_doc) / 60))
```

# Reproducibility payload {#sec-export-payload}

```{r export-payload}
#| echo: false
#| message: false
.qmd_dir  <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) getwd())
.qmd_stem <- tryCatch(tools::file_path_sans_ext(basename(knitr::current_input())),
                      error = function(e) "analysis")
.dirout   <- if (is.null(dirout)) .qmd_stem else dirout
.base     <- if (is.null(results_dir)) file.path(.qmd_dir, "_payloads") else results_dir
.out_dir  <- file.path(.base, .dirout)
dir.create(.out_dir, recursive = TRUE, showWarnings = FALSE)
.payload_file <- file.path(.out_dir, paste0(.dirout, "_payload.rds"))

payload <- list(
  table  = lad,
  labels = list(
    sg_harm  = H_def,
    sg_focus = "maxeffCons",
    focus    = NULL),
  meta = list(
    n_total    = N,
    n_events   = NULL,
    outcome    = "y_decline",
    event      = NULL,
    treat      = "treat",
    covariates = confounders.name,
    c1         = 10,
    c2         = params$c2_ratio * 10,
    args_call  = list(consistency_method = "resample", pconsistency = 0.90,
                      draws = params$draws, block = 5e4, seed = params$seed,
                      M = fam$M, q_rungs = q_rungs, q_shared = q_shared,
                      T_obs = T_obs, c1_ladder = c1_ladder, c2_vec = c2_vec,
                      n_workers = params$n_workers, t_loop_secs = t_loop)),
  extras = list(
    type1 = list(diagonal = tI, c1_05_diag = c1_05_diag,
                 c1_10_diag = c1_10_diag, at_analyst_diag = type1_at_10,
                 surface = z, c1_05_fixed_c2_5 = c1_05_fix,
                 homogeneous = list(diagonal = tIh, c1_05_hom = c1_05_hom,
                                    c1_10_hom = c1_10_hom,
                                    at_analyst_diag = hom_at_10,
                                    at_comparator_c2_5 = hom_at_5,
                                    surface = och),
                 subthreshold_variants = tIv),
    declared    = dec,
    comparator_c2_5 = cmp,
    calibration = lad[, c("q", "pT", "pTs", "pT5", "pT5s")],
    anchor      = list(def = H_def, n_H = n_H, T_obs = T_obs,
                       p_cons = p_cons, beta_treat = beta_treat),
    purity      = list(nearest_lab = fam$lab[i_near],
                       nearest_purity = purity_near,
                       nearest_jaccard = jac[i_near]),
    c2_policy   = list(headline = "c2 = c2_ratio * c1",
                       c2_ratio = params$c2_ratio, comparator_c2 = 5,
                       applies_to = "type-I and power (threshold policy)",
                       calibration_read_at = c(analyst = params$c2_ratio * 10,
                                               comparator = 5),
                       reason = paste(
                         "c2 sets candidate eligibility; the calibration",
                         "interrogates the analysis as actually conducted,",
                         "so it is read at the analyst's operating",
                         "consistency floor, not on the policy diagonal")),
    q_variants  = list(defs = Q_variants, prevalence = prev,
                       table = knob, curves = cal,
                       null_not_shared = list(dev_vs_primary = dev_vs_primary,
                                              det_at_Tobs = det0)),
    interpretation = list(
      tail_table = tail_tab,
      itt        = itt,
      crossings  = knob[, c("variant", "P", "q05", "q05c", "q50", "q50c")])),
  est_scale = "md",
  built_at  = Sys.time(),
  forestsearch_version = tryCatch(as.character(utils::packageVersion("forestsearch")),
                                  error = function(e) NA_character_))

saveRDS(payload, .payload_file)
cat("Wrote payload to:", .payload_file, "\n")
```
