---
title: "ACTG175 Continuous Harm Search: Single-Data Analysis"
format: html
---

```{r setup}
library(forestsearch)
library(speff2trial)
library(gt)

set.seed(8316951L)

# ---------------------------------------------------------------------------
# Output location for this document's results:  <base>/<dirout>/
#   results_dir = NULL -> base is <qmd dir>/_payloads
#   dirout      = NULL -> the .qmd stem
# Setting results_dir redirects the base only; each run still gets its own
# subdirectory named by dirout.
#
# Examples:
#   results_dir <- NULL;  dirout <- NULL
#     -> <qmd dir>/_payloads/<stem>/<stem>_payload.rds
#   results_dir <- NULL;  dirout <- "actg175_continuous_fixed_family"
#     -> <qmd dir>/_payloads/actg175_continuous_fixed_family/actg175_continuous_fixed_family_payload.rds
#   results_dir <- "~/GitHub/fs-post-selection-manuscript_jasa/_payloads"; dirout <- NULL
#     -> ~/GitHub/fs-post-selection-manuscript_jasa/_payloads/<stem>/<stem>_payload.rds
# ---------------------------------------------------------------------------
results_dir <- NULL
dirout      <- NULL
```

## Data Preparation

```{r data-prep}
# ACTG175: arm 1 = ZDV+ddI (experimental), arm 3 = ddI (control)
actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$id <- seq_len(nrow(actg_df))

# Treatment: ZDV+ddI = 1, ddI = 0 (NO switching)
actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)

actg_df <- actg_df[!is.na(actg_df$cd420), ]

# -- Continuous outcomes -------------------------------------------------------
# Adverse scale: CD4 decline (higher = worse)
actg_df$y_decline <- actg_df$cd40 - actg_df$cd420

# Beneficial scale: CD4 improvement (higher = better)
actg_df$y_improve <- actg_df$cd420 - actg_df$cd40

cat(sprintf("N = %d\n", nrow(actg_df)))
cat(sprintf("Mean CD4 decline (y_decline):    %.1f\n", mean(actg_df$y_decline)))
cat(sprintf("  ZDV+ddI (T=1): %.1f\n", mean(actg_df$y_decline[actg_df$treat == 1])))
cat(sprintf("  ddI     (T=0): %.1f\n", mean(actg_df$y_decline[actg_df$treat == 0])))
cat(sprintf("Mean CD4 improvement (y_improve): %.1f\n", mean(actg_df$y_improve)))
cat(sprintf("  ZDV+ddI (T=1): %.1f\n", mean(actg_df$y_improve[actg_df$treat == 1])))
cat(sprintf("  ddI     (T=0): %.1f\n", mean(actg_df$y_improve[actg_df$treat == 0])))
```

## Covariates

```{r covariates}
cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")

bin_vars <- c("hemo", "homo", "drugs", "race", "gender", "symptom")
for (v in bin_vars) actg_df[[v]] <- as.factor(actg_df[[v]])

confounders <- c(cont_vars, bin_vars)
```

## ForestSearch (continuous, adverse scale)

FS uses `y_decline` (higher = worse) with `adverse_outcome = TRUE`.
No outcome flip — MD > c1 means treatment increases CD4 decline = **harm**.

### Fixed-family configuration

This template demonstrates the **fixed-family** configuration. With
`use_lasso = FALSE` and `use_grf = FALSE`, the candidate family is enumerated
directly from the supplied covariates and their `cont.cutoff` quantile splits,
and is then held fixed under resampling — no screening model is refit on each
split or bootstrap replicate, so the family cannot regenerate.

Enabling either screener puts the analysis in the **conditional-on-family**
regime instead: the family becomes a function of the resampled data, and the
subgroup that is selected is conditional on a family that moved with it. That
is a different estimand and a different inferential target, not a tuning choice.

Because no screener runs, the GRF screening parameters (`vi.grf.min`,
`dmin.grf`) are inert and are omitted from the call rather than passed and
ignored. `cont.cutoff` is retained: it governs family *construction*, which is
still active.

```{r fs-run}
# Thresholds on the mean-difference scale (CD4 count units)
c1 <- 10    # screening: MD >= 10
c2 <- 5     # consistency: MD >= 5

cat(sprintf("c1 (threshold) = %.0f, c2 (consistency) = %.0f\n", c1, c2))

fs_result <- forestsearch(
  df.analysis      = actg_df,
  confounders.name = confounders,
  outcome.name     = "y_decline",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "continuous",
  effect_measure   = "MD",
  adverse_outcome  = TRUE,
  seedit           = 8316951L,
  n.min            = 60L,
  d0.min           = 10L,
  d1.min           = 10L,
  hr.threshold     = c1,
  hr.consistency   = c2,
  pconsistency.threshold = 0.80,
  fs.splits        = 500L,
  maxk             = 2L,
  sg_focus         = "minSG",
  selection_rule   = "neighborhood",
  use_lasso        = FALSE,
  use_grf          = FALSE,
  is.RCT           = TRUE,
  cont.cutoff      = 4L,
  details          = TRUE,
  quiet            = FALSE,
  parallel_args    = list(plan = "sequential", workers = 1,
                          show_message = FALSE)
)
```

### Identified Subgroups

```{r fs-results}
fs_tabs <- sg_tables(fs_result, ndecimals = 3, which_df = "est")
fs_tabs$sg10_out
```

### Treatment Effect Estimates

```{r fs-estimates}
fs_tabs$tab_estimates
```

## GRF Standalone (continuous, beneficial scale)

GRF uses `y_improve` (higher = better) with `adverse_outcome = FALSE` (default).
CATE < 0 means treatment decreases improvement = **harm**.

```{r grf-run}
grf_result <- grf.subg.harm.glm(
  data             = actg_df,
  confounders.name = confounders,
  outcome.name     = "y_improve",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "continuous",
  effect_measure   = "MD",
  dmin.grf         = 0.1,
  maxdepth         = 2L,
  RCT              = TRUE,
  sg.criterion     = "mDiff",
  details          = TRUE
)

cat("\n=== GRF Result ===\n")
if (!is.null(grf_result$sg.harm.id) && length(grf_result$sg.harm.id) > 0) {
  cat("Subgroup:", paste(grf_result$sg.harm.id, collapse = " & "), "\n")
  grf_H <- grf_result$data$treat.recommend == 0
  cat(sprintf("Size(H): %d,  Size(Hc): %d\n", sum(grf_H), sum(!grf_H)))
  cat(sprintf("CATE(H):  %.2f (treatment effect on improvement in H)\n",
      grf_result$cate_sg))
  cat(sprintf("CATE(Hc): %.2f\n", grf_result$cate_sgc))
} else {
  cat("GRF: no subgroup identified.\n")
}
```

### GRF Treatment Effect Estimates

```{r grf-results}
grf_tabs <- sg_tables(grf_result, ndecimals = 3, which_df = "est")
grf_tabs$tab_estimates
```

## Side-by-Side Comparison

```{r comparison}
cat("=== FS (adverse Y = CD4 decline, adverse_outcome = TRUE) ===\n")
fs_tabs$tab_estimates

cat("\n=== GRF (beneficial Y = CD4 improvement, adverse_outcome = FALSE) ===\n")
grf_tabs$tab_estimates
```

# Reproducibility payload {#sec-export-payload}

The chunk below harvests the objects this render already computed into a single
small list and writes it as an `.rds`, so the results survive the HTML. No
analysis is repeated. Every element is guarded, so a section this render skipped
stores `NULL` rather than erroring, and the top-level names match the other
application payloads (`table`, `labels`, `meta`, `extras`, `est_scale`,
`built_at`, `forestsearch_version`) so a reader can treat them uniformly.
`est_scale` is `"md"` here: the mean-difference scale, the lowercase form of the
package's `effect_measure = "MD"`.

```{r export-payload}
#| echo: false
#| message: false

.qmd_dir  <- tryCatch(dirname(knitr::current_input(dir = TRUE)),
                      error = function(e) getwd())
.qmd_stem <- tryCatch(tools::file_path_sans_ext(basename(knitr::current_input())),
                      error = function(e) "analysis")
.dirout   <- if (is.null(dirout)) .qmd_stem else dirout
.base     <- if (is.null(results_dir)) file.path(.qmd_dir, "_payloads") else results_dir
.out_dir  <- file.path(.base, .dirout)
dir.create(.out_dir, recursive = TRUE, showWarnings = FALSE)
.payload_file <- file.path(.out_dir, paste0(.dirout, "_payload.rds"))

.get <- function(nm) if (exists(nm, inherits = TRUE)) get(nm, inherits = TRUE) else NULL
.try <- function(expr) tryCatch(expr, error = function(e) NULL)

.fs   <- .get("fs_result")
.grf  <- .get("grf_result")

payload <- list(
  table  = .try(.get("fs_tabs")$tab_estimates),
  labels = list(
    sg_harm  = if (is.null(.fs)) NULL else .fs$sg.harm,
    sg_focus = if (is.null(.fs)) NULL else .fs$sg_focus,
    focus    = .get("focus")),   # no focus profile in this document: NULL
  meta = list(
    n_total     = .try(nrow(.get("actg_df"))),
    # A continuous outcome has no event indicator, so this element is
    # structurally NULL here. It is kept so the payload shape matches the
    # survival and binary documents rather than dropping to six elements.
    n_events    = NULL,
    outcome     = "y_decline",
    event       = NULL,
    treat       = "treat",
    covariates  = .get("confounders"),
    c1          = .get("c1"),
    c2          = .get("c2"),
    # No `provenance` element: fs_param_provenance() is a per-document helper
    # defined inline in the gbsg and binary documents, not a package export.
    # Calling it here would resolve to NULL through .try() on every render and
    # record a silent absence as if the check had run. The resolved settings
    # that matter for this template are captured in extras$fixed_family below,
    # read from the fitted object rather than from the config above it.
    args_call   = .try(.fs$args_call_all)),
  extras = list(
    # The fixed-family condition this template demonstrates, recorded as the
    # settings that produce it rather than as prose.
    fixed_family = list(
      use_lasso      = .try(.fs$args_call_all$use_lasso),
      use_grf        = .try(.fs$args_call_all$use_grf),
      cont.cutoff    = .try(.fs$args_call_all$cont.cutoff),
      selection_rule = .try(.fs$args_call_all$selection_rule)),
    fs_tables = .try(list(
      sg10_out      = .get("fs_tabs")$sg10_out,
      tab_estimates = .get("fs_tabs")$tab_estimates)),
    # The standalone GRF run is a comparator, not a screener: it does not
    # enter the ForestSearch family.
    grf = if (is.null(.grf)) NULL else list(
      sg_harm_id = .grf$sg.harm.id,
      cate_sg    = .grf$cate_sg,
      cate_sgc   = .grf$cate_sgc,
      table      = .try(.get("grf_tabs")$tab_estimates))),
  est_scale = "md",
  built_at  = Sys.time(),
  forestsearch_version = tryCatch(as.character(utils::packageVersion("forestsearch")),
                                  error = function(e) NA_character_))

saveRDS(payload, .payload_file)
cat("Wrote payload to:", .payload_file, "\n")
```
