---
title: "Simulation Studies Based on ACTG175: Binary Endpoint"
subtitle: "ForestSearch Subgroup Identification with CD4 Response at 20 Weeks"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ── Simulation control ──────────────────────────────────────────
nsims_alt   <- 50L
nsims_null  <- 50L
sim_seed    <- 8316951L
n_workers   <- min(parallel::detectCores(logical = FALSE) - 1L, 120L)

# ── Dataset preparation: ACTG175 ───────────────────────────────
actg_arms        <- c(1L, 3L)     # ZDV+ddI vs ddI
actg_treat_arm   <- 1L            # ZDV+ddI = experimental
actg_age_cut     <- 34
actg_preanti_cut <- 744.5

# ── Binary outcome ─────────────────────────────────────────────
# I(cd420 > cd40): CD4 count increased from baseline to 20 weeks
# Subjects with missing cd420 are removed.

# ── Benefit search thresholds (on OR scale) ────────────────────
#   c1 = 1/0.60 ≈ 1.667 (screening)
#   c2 = 1/0.80 = 1.25   (consistency)
fs_or_threshold    <- round(1 / 0.60, 3)
fs_or_consistency  <- round(1 / 0.80, 3)

# ── DGM construction ────────────────────────────────────────────
dgm_n_super       <- 5000L
dgm_or_target     <- 2.0

# ── Simulation study ─────────────────────────────────────────────
sim_n_sample      <- 1000L

# ── ForestSearch parameters ──────────────────────────────────────
fs_splits              <- 400L
fs_n_min               <- 60L
fs_d0_min              <- 12L
fs_d1_min              <- 12L
fs_maxk                <- 2L
fs_vi_grf_min          <- -0.2
fs_pconsistency        <- 0.90
fs_seedit              <- 8316951L

# ── ForestSearch pipeline flags ──────────────────────────────────
fs_use_lasso           <- FALSE   # LASSO pre-screening
fs_use_grf             <- TRUE    # GRF pre-screening
fs_is_rct              <- TRUE    # RCT design (W.hat = 0.5)
fs_adverse_outcome     <- NULL    # NULL = package default (TRUE for binary)

# ── Reporting ────────────────────────────────────────────────────
sg_notation            <- "benefit"  # invert ORs to original scale
```


# Introduction {#sec-intro}

This document applies ForestSearch with `outcome_type = "binary"`
to the ACTG175 dataset using the CD4 response endpoint
$Y = I(\text{cd420} > \text{cd40})$.  Treatment is switched
(ddI = treat, ZDV+ddI = control) for the benefit-search framework.

Before running simulations, ForestSearch is applied to the
**actual ACTG175 data** with full diagnostic output (@sec-single)
so that every step of the procedure — LASSO screening, GRF
variable importance, subgroup enumeration, and bootstrap
consistency — can be reviewed directly.


# Data Preparation {#sec-data}

```{r load-packages}
library(forestsearch)
library(survival)
library(speff2trial)
library(data.table)
library(gt)
library(ggplot2)
library(doFuture)
library(foreach)

timings <- list()
```

```{r prepare-datasets}
t0 <- proc.time()[3]

# ── ACTG175: arms 1 (ZDV+ddI) vs 3 (ddI) ────────────────────
actg_df <- subset(speff2trial::ACTG175, arms %in% actg_arms)
actg_df$id <- seq_len(nrow(actg_df))

# Original treatment: arm 1 = 1, arm 3 = 0
actg_df$treat_orig <- ifelse(actg_df$arms == actg_treat_arm, 1L, 0L)

# Switched treatment for benefit search
actg_df$treat <- 1L - actg_df$treat_orig

# ── Binary outcome: CD4 response at 20 weeks ─────────────────
actg_df$y_binary <- as.integer(actg_df$cd420 > actg_df$cd40)

# Remove subjects with missing cd420
n_before <- nrow(actg_df)
actg_df <- actg_df[!is.na(actg_df$cd420), ]
n_after  <- nrow(actg_df)

cat(sprintf("ACTG175: N = %d → %d after removing %d with missing cd420\n",
    n_before, n_after, n_before - n_after))
cat(sprintf("Overall response rate: %.1f%%\n",
    100 * mean(actg_df$y_binary)))
cat(sprintf("  ZDV+ddI (arm 1): %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$treat_orig == 1])))
cat(sprintf("  ddI (arm 3):     %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$treat_orig == 0])))

# ── Discretise covariates ─────────────────────────────────────
actg_df$z1  <- as.factor(ifelse(actg_df$age > actg_age_cut, 1L, 0L))
actg_df$z2  <- as.factor(ifelse(actg_df$preanti <= actg_preanti_cut, 1L, 0L))
actg_df$z3  <- as.factor(ifelse(actg_df$wtkg <= 75, 1L, 0L))
actg_df$z4  <- as.factor(ifelse(
  actg_df$karnof <= median(actg_df$karnof), 1L, 0L))
actg_df$z5  <- as.factor(ifelse(
  actg_df$cd40 <= median(actg_df$cd40), 1L, 0L))
actg_df$z6  <- as.factor(ifelse(
  actg_df$cd80 <= median(actg_df$cd80), 1L, 0L))
actg_df$z7  <- as.factor(actg_df$hemo)
actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs)
actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender)
actg_df$z12 <- as.factor(actg_df$symptom)

# ── Subgroup flag ─────────────────────────────────────────────
actg_df$flag_harm <- as.integer(
  actg_df$z1 == 1 & actg_df$z2 == 1
)

cat(sprintf("\nSubgroup Q: Age > %d AND Preanti <= %.0f\n",
    actg_age_cut, actg_preanti_cut))
cat(sprintf("  Q prevalence: %d / %d (%.1f%%)\n",
    sum(actg_df$flag_harm), nrow(actg_df),
    100 * mean(actg_df$flag_harm)))

cat(sprintf("\nResponse rates by subgroup (original scale):\n"))
cat(sprintf("  Q:  ZDV+ddI = %.1f%%,  ddI = %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$flag_harm == 1 &
                                  actg_df$treat_orig == 1]),
    100 * mean(actg_df$y_binary[actg_df$flag_harm == 1 &
                                  actg_df$treat_orig == 0])))
cat(sprintf("  Qc: ZDV+ddI = %.1f%%,  ddI = %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$flag_harm == 0 &
                                  actg_df$treat_orig == 1]),
    100 * mean(actg_df$y_binary[actg_df$flag_harm == 0 &
                                  actg_df$treat_orig == 0])))

# ── Drop single-level factors ────────────────────────────────
z_vars_all <- paste0("z", 1:12)
z_vars <- z_vars_all[vapply(z_vars_all, function(v) {
  length(unique(actg_df[[v]])) >= 2
}, logical(1))]

dropped <- setdiff(z_vars_all, z_vars)
if (length(dropped) > 0) {
  cat(sprintf("Dropped single-level factors: %s\n",
      paste(dropped, collapse = ", ")))
}

confounders_name <- z_vars
cat(sprintf("Confounders (%d): %s\n",
    length(confounders_name),
    paste(confounders_name, collapse = ", ")))

timings$data_prep <- proc.time()[3] - t0
```


# Single-Data Analysis {#sec-single}

Before any simulation, we run ForestSearch on the **actual
ACTG175 data** with `details = TRUE` to examine every step of
the procedure: LASSO screening, GRF variable importance,
candidate enumeration, and bootstrap consistency.

```{r single-data-analysis}
#| code-fold: false
t0 <- proc.time()[3]

cat("=", rep("=", 60), "\n", sep = "")
cat("  ForestSearch on ACTG175 observed data\n")
cat("  outcome_type = 'binary', effect_measure = 'OR'\n")
cat("  Treatment: switched (ddI = 1, ZDV+ddI = 0)\n")
cat("  Y = I(cd420 > cd40) = improvement\n")
cat("  Thresholds: c1 =", fs_or_threshold,
    " c2 =", fs_or_consistency, "\n")
cat("=", rep("=", 60), "\n\n", sep = "")

fs_actg <- forestsearch(
  df.analysis      = actg_df,
  confounders.name = confounders_name,
  outcome.name     = "y_binary",
  event.name       = "y_binary",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "OR",
  adverse_outcome  = fs_adverse_outcome,
  seedit           = fs_seedit,
  n.min            = fs_n_min,
  d0.min           = fs_d0_min,
  d1.min           = fs_d1_min,
  effect.threshold     = fs_or_threshold,
  consistency.threshold   = fs_or_consistency,
  pconsistency.threshold = fs_pconsistency,
  fs.splits        = fs_splits,
  maxk             = fs_maxk,
  use_lasso        = fs_use_lasso,
  use_grf          = fs_use_grf,
  is.RCT           = fs_is_rct,
  vi.grf.min       = fs_vi_grf_min,
  details          = TRUE,
  quiet            = FALSE,
  parallel_args    = list(
    plan = "multisession", workers = 9, show_message = TRUE
  )
)

timings$single_data <- proc.time()[3] - t0
```


## Examine Result

```{r examine-single-result}
#| code-fold: false

cat("\n=== ForestSearch Result ===\n")

if (!is.null(fs_actg$grp.consistency)) {
  sg <- fs_actg$grp.consistency$sg.harm.id
  if (!is.null(sg) && length(sg) > 0 && !all(is.na(sg))) {
    cat("Subgroup identified:", paste(sg, collapse = " & "), "\n")

    df_est <- fs_actg$df.est
    in_H  <- df_est$treat.recommend == 0
    in_Hc <- df_est$treat.recommend == 1

    cat(sprintf("  Size(H): %d,  Size(Hc): %d\n",
        sum(in_H), sum(in_Hc)))

    # OR in identified subgroup (switched scale)
    if (sum(in_H) >= 10) {
      fit_H <- glm(y_binary ~ treat, data = df_est[in_H, ],
                    family = binomial)
      or_H <- exp(coef(fit_H)["treat"])
      cat(sprintf("  OR(H, switched):  %.3f  (inverted: %.3f)\n",
          or_H, 1/or_H))
    }
    if (sum(in_Hc) >= 10) {
      fit_Hc <- glm(y_binary ~ treat, data = df_est[in_Hc, ],
                     family = binomial)
      or_Hc <- exp(coef(fit_Hc)["treat"])
      cat(sprintf("  OR(Hc, switched): %.3f  (inverted: %.3f)\n",
          or_Hc, 1/or_Hc))
    }

    # ITT
    fit_itt <- glm(y_binary ~ treat, data = df_est, family = binomial)
    cat(sprintf("  OR(ITT, switched): %.3f\n",
        exp(coef(fit_itt)["treat"])))

    # Classification vs true Q
    true_H <- df_est$flag_harm == 1
    tp <- sum(in_H & true_H)
    fp <- sum(in_H & !true_H)
    fn <- sum(!in_H & true_H)
    tn <- sum(!in_H & !true_H)
    cat(sprintf("\n  Classification vs true Q:\n"))
    cat(sprintf("    TP: %d  FP: %d  FN: %d  TN: %d\n",
        tp, fp, fn, tn))
    cat(sprintf("    Sens: %.3f  Spec: %.3f  PPV: %.3f\n",
        tp / (tp + fn), tn / (tn + fp), tp / (tp + fp)))
  } else {
    cat("No subgroup identified by ForestSearch.\n")
  }
} else {
  cat("ForestSearch did not return grp.consistency.\n")
}

cat(sprintf("\nSingle-data analysis: %.1f sec\n", timings$single_data))
```


# Binary DGM {#sec-dgm}

```{r build-binary-dgm}
#| code-fold: false
t0 <- proc.time()[3]

# ── Baseline logistic model (improvement) ─────────────────────
base_formula <- as.formula(paste(
  "y_binary ~ treat +",
  paste(z_vars, collapse = " + ")
))

fit_base <- glm(base_formula, data = actg_df, family = binomial)
cat("Baseline logistic model (improvement):\n")
cat(sprintf("  Treatment OR (switched): %.3f\n",
    exp(coef(fit_base)["treat"])))

# ── Super-population ─────────────────────────────────────────
set.seed(sim_seed)
idx_super <- sample(nrow(actg_df), dgm_n_super, replace = TRUE)
df_super  <- actg_df[idx_super, ]
df_super$id <- seq_len(dgm_n_super)

# ── Potential outcomes ────────────────────────────────────────
newdata_0 <- df_super; newdata_0$treat <- 0L
newdata_1 <- df_super; newdata_1$treat <- 1L

df_super$p0_base <- predict(fit_base, newdata = newdata_0, type = "response")
df_super$p1_base <- predict(fit_base, newdata = newdata_1, type = "response")

# ── Interaction ───────────────────────────────────────────────
# Positive beta_inter increases P(improvement | ddI, Q)
add_interaction <- function(df, beta_inter) {
  in_Q <- df$flag_harm == 1
  eta_1 <- qlogis(df$p1_base)
  eta_1[in_Q] <- eta_1[in_Q] + beta_inter
  df$p1 <- plogis(eta_1)
  df$p0 <- df$p0_base
  df
}

# ── Calibrate ─────────────────────────────────────────────────
compute_or_Q <- function(df, beta_inter) {
  df_mod <- add_interaction(df, beta_inter)
  in_Q <- df_mod$flag_harm == 1
  p1_Q <- mean(df_mod$p1[in_Q])
  p0_Q <- mean(df_mod$p0[in_Q])
  (p1_Q / (1 - p1_Q)) / (p0_Q / (1 - p0_Q))
}

beta_grid <- seq(0, 3.0, by = 0.05)
or_grid   <- sapply(beta_grid, function(b) compute_or_Q(df_super, b))
best_idx  <- which.min(abs(or_grid - dgm_or_target))
beta_inter_cal <- beta_grid[best_idx]

cat(sprintf("\nCalibration: beta = %.2f → OR(Q) = %.3f (target: %.1f)\n",
    beta_inter_cal, or_grid[best_idx], dgm_or_target))

# ── Build DGMs ───────────────────────────────────────────────
df_super_alt  <- add_interaction(df_super, beta_inter_cal)
df_super_null <- add_interaction(df_super, 0)

compute_or <- function(df, subset_flag = NULL) {
  if (!is.null(subset_flag)) df <- df[df$flag_harm == subset_flag, ]
  p1 <- mean(df$p1); p0 <- mean(df$p0)
  (p1 / (1 - p1)) / (p0 / (1 - p0))
}

or_Q    <- compute_or(df_super_alt, 1)
or_Qc   <- compute_or(df_super_alt, 0)
or_itt  <- compute_or(df_super_alt)
or_null <- compute_or(df_super_null)

cat(sprintf("  OR(Q):   %.3f  (switched)\n", or_Q))
cat(sprintf("  OR(Qc):  %.3f  (switched)\n", or_Qc))
cat(sprintf("  OR(ITT): %.3f  (switched)\n", or_itt))
cat(sprintf("  Null OR: %.3f  (switched)\n", or_null))

make_binary_dgm <- function(df_sp, model_type, or_sg, or_sgc, or_all) {
  list(
    df_super = df_sp,
    hazard_ratios = list(
      harm_subgroup = or_sg, no_harm_subgroup = or_sgc, overall = or_all
    ),
    model_type = model_type,
    model_params = list(
      beta = coef(fit_base), beta_inter = beta_inter_cal,
      family = "binomial"
    ),
    n_super = dgm_n_super
  )
}

dgm_alt  <- make_binary_dgm(df_super_alt,  "alt",  or_Q, or_Qc, or_itt)
dgm_null <- make_binary_dgm(df_super_null, "null", or_null, or_null, or_null)

timings$dgm <- proc.time()[3] - t0
```


# Simulation Engine {#sec-sim}

```{r binary-simulator}
#| code-fold: false

simulate_binary_trial <- function(dgm, n, sim_id, seed) {
  set.seed(seed + sim_id)
  idx <- sample(nrow(dgm$df_super), n, replace = TRUE)
  df  <- dgm$df_super[idx, ]
  df$id <- seq_len(n)
  df$treat_sim <- rbinom(n, 1, 0.5)
  df$y_sim <- ifelse(
    df$treat_sim == 1,
    rbinom(n, 1, df$p1),
    rbinom(n, 1, df$p0)
  )
  df
}

run_binary_fs <- function(sim_data, dgm, fs_params, confounders_name) {
  fs_res <- tryCatch({
    future::plan("sequential")
    forestsearch(
      df.analysis      = sim_data,
      confounders.name = confounders_name,
      outcome.name     = "y_sim",
      event.name       = "y_sim",
      treat.name       = "treat_sim",
      id.name          = "id",
      outcome_type     = "binary",
      effect_measure   = "OR",
      adverse_outcome  = fs_params$adverse_outcome,
      seedit           = fs_params$seedit,
      n.min            = fs_params$n.min,
      d0.min           = fs_params$d0.min,
      d1.min           = fs_params$d1.min,
      hr.threshold     = fs_params$hr.threshold,
      hr.consistency   = fs_params$hr.consistency,
      pconsistency.threshold = fs_params$pconsistency.threshold,
      fs.splits        = fs_params$fs.splits,
      maxk             = fs_params$maxk,
      use_lasso        = fs_params$use_lasso,
      use_grf          = fs_params$use_grf,
      is.RCT           = fs_params$is_rct,
      vi.grf.min       = fs_params$vi.grf.min,
      details          = FALSE,
      quiet            = TRUE,
      parallel_args    = list(
        plan = "sequential", workers = 1, show_message = FALSE
      )
    )
  }, error = function(e) {
    warning("FS failed: ", e$message)
    NULL
  })

  extract_binary_estimates(sim_data, fs_res, dgm, analysis = "FS")
}


extract_binary_estimates <- function(df, fs_res, dgm, analysis = "FS") {
  out <- data.frame(
    analysis   = analysis,
    any.H      = 0L,
    hr.H.hat   = NA_real_,
    hr.Hc.hat  = NA_real_,
    hr.H.true  = NA_real_,
    hr.Hc.true = NA_real_,
    hr.itt     = NA_real_,
    sens       = NA_real_,
    spec       = NA_real_,
    ppv        = NA_real_,
    npv        = NA_real_,
    size.H     = NA_real_,
    size.Hc    = nrow(df),
    p.cens     = 0,
    stringsAsFactors = FALSE
  )

  if (is.null(fs_res)) return(out)

  sg_harm <- fs_res$grp.consistency$sg.harm.id
  if (is.null(sg_harm) || length(sg_harm) == 0 ||
      all(is.na(sg_harm)) || all(sg_harm == "")) {
    itt_fit <- tryCatch(
      glm(y_sim ~ treat_sim, data = df, family = binomial),
      error = function(e) NULL
    )
    if (!is.null(itt_fit)) out$hr.itt <- exp(coef(itt_fit)["treat_sim"])
    return(out)
  }

  out$any.H <- 1L
  df_pred <- fs_res$df.est
  if (is.null(df_pred) || !"treat.recommend" %in% names(df_pred))
    return(out)

  in_H  <- df_pred$treat.recommend == 0
  in_Hc <- df_pred$treat.recommend == 1
  out$size.H  <- sum(in_H)
  out$size.Hc <- sum(in_Hc)

  if (sum(in_H) >= 10 && length(unique(df_pred$treat_sim[in_H])) == 2) {
    fit_H <- tryCatch(
      glm(y_sim ~ treat_sim, data = df_pred[in_H, ], family = binomial),
      error = function(e) NULL
    )
    if (!is.null(fit_H)) out$hr.H.hat <- exp(coef(fit_H)["treat_sim"])
  }

  if (sum(in_Hc) >= 10 && length(unique(df_pred$treat_sim[in_Hc])) == 2) {
    fit_Hc <- tryCatch(
      glm(y_sim ~ treat_sim, data = df_pred[in_Hc, ], family = binomial),
      error = function(e) NULL
    )
    if (!is.null(fit_Hc)) out$hr.Hc.hat <- exp(coef(fit_Hc)["treat_sim"])
  }

  itt_fit <- tryCatch(
    glm(y_sim ~ treat_sim, data = df_pred, family = binomial),
    error = function(e) NULL
  )
  if (!is.null(itt_fit)) out$hr.itt <- exp(coef(itt_fit)["treat_sim"])

  out$hr.H.true  <- dgm$hazard_ratios$harm_subgroup
  out$hr.Hc.true <- dgm$hazard_ratios$no_harm_subgroup

  true_H <- df_pred$flag_harm == 1
  if (sum(true_H) > 0 && sum(in_H) > 0) {
    tp <- sum(in_H & true_H)
    fp <- sum(in_H & !true_H)
    fn <- sum(!in_H & true_H)
    tn <- sum(!in_H & !true_H)
    out$sens <- tp / (tp + fn)
    out$spec <- tn / (tn + fp)
    out$ppv  <- if (tp + fp > 0) tp / (tp + fp) else NA
    out$npv  <- if (tn + fn > 0) tn / (tn + fn) else NA
  }

  out
}
```


# Run Simulations {#sec-run}

```{r define-params}
fs_params <- list(
  # Thresholds
  hr.threshold           = fs_or_threshold,
  hr.consistency         = fs_or_consistency,
  pconsistency.threshold = fs_pconsistency,
  # Computational
  fs.splits              = fs_splits,
  n.min                  = fs_n_min,
  d0.min                 = fs_d0_min,
  d1.min                 = fs_d1_min,
  maxk                   = fs_maxk,
  vi.grf.min             = fs_vi_grf_min,
  seedit                 = fs_seedit,
  # Pipeline flags
  use_lasso              = fs_use_lasso,
  use_grf                = fs_use_grf,
  is_rct                 = fs_is_rct,
  adverse_outcome        = fs_adverse_outcome
)
```

```{r run-parallel-setup}
registerDoFuture()
plan(multisession, workers = n_workers)
cat(sprintf("Workers: %d\n", n_workers))
```

## Alternative Hypothesis

```{r sim-alt}
t0 <- proc.time()[3]

results_alt <- foreach(
  i = seq_len(nsims_alt),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  sim_data <- simulate_binary_trial(dgm_alt, sim_n_sample, i, sim_seed)
  run_binary_fs(sim_data, dgm_alt, fs_params, confounders_name)
}

n_completed_alt <- nrow(results_alt)
n_failed_alt    <- nsims_alt - n_completed_alt
if (n_failed_alt > 0) {
  warning(sprintf("%d of %d alt sims failed", n_failed_alt, nsims_alt))
}

timings$sim_alt <- proc.time()[3] - t0
cat(sprintf("Alt: %d / %d completed (%.1f sec)\n",
    n_completed_alt, nsims_alt, timings$sim_alt))
if (n_failed_alt > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_alt))
cat(sprintf("Detection rate: %.1f%%\n",
    100 * mean(results_alt$any.H)))
```

## Null Hypothesis

```{r sim-null}
t0 <- proc.time()[3]

results_null <- foreach(
  i = seq_len(nsims_null),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  sim_data <- simulate_binary_trial(dgm_null, sim_n_sample, i,
                                     sim_seed + 100000L)
  run_binary_fs(sim_data, dgm_null, fs_params, confounders_name)
}

n_completed_null <- nrow(results_null)
n_failed_null    <- nsims_null - n_completed_null
if (n_failed_null > 0) {
  warning(sprintf("%d of %d null sims failed", n_failed_null, nsims_null))
}

timings$sim_null <- proc.time()[3] - t0
cat(sprintf("Null: %d / %d completed (%.1f sec)\n",
    n_completed_null, nsims_null, timings$sim_null))
if (n_failed_null > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_null))
cat(sprintf("False discovery rate: %.1f%%\n",
    100 * mean(results_null$any.H)))
```

```{r cleanup-plan}
plan(sequential)
```


# Results {#sec-results}

Results are displayed with `subgroup_notation = "benefit"`, which
automatically inverts all OR values from the switched scale to the
original scale and uses Q/Qc labels.

## Operating Characteristics Under H1

```{r oc-alt}
t0 <- proc.time()[3]

format_oc_results(
  results_alt,
  metrics   = c("detection", "classification"),
  digits    = 2,
  digits_hr = 2,
  title     = "ACTG175 Binary: Identification (FS)",
  subtitle  = sprintf(
    "n = %d, %d sims, OR(Q) = %.2f, OR(Qc) = %.2f (switched)",
    sim_n_sample, nsims_alt, or_Q, or_Qc
  ),
  subgroup_notation = sg_notation
)

build_estimation_table(
  results         = results_alt,
  dgm             = dgm_alt,
  analysis_method = "FS",
  subtitle = sprintf(
    "ACTG175 Binary | n = %d, %d sims",
    sim_n_sample, nsims_alt
  ),
  font_size = 16,
  subgroup_notation = sg_notation
)

timings$report_alt <- proc.time()[3] - t0
```

### Interpretation

```{r interpret-alt, results="asis"}
interpret_estimation_table(
  results_alt, dgm_alt,
  analysis_method = "FS",
  scenario = "alt",
  subgroup_notation = sg_notation
)
```


## Operating Characteristics Under H0

```{r oc-null}
format_oc_results(
  results_null,
  metrics  = c("detection", "classification"),
  digits   = 2,
  title    = "ACTG175 Binary: Type I Error (FS)",
  subtitle = sprintf(
    "n = %d, %d sims, OR(ITT) = %.2f (switched)",
    sim_n_sample, nsims_null, or_null
  ),
  subgroup_notation = sg_notation
)
```


## Classification Table

```{r classification}
build_classification_table(
  scenario_results = list(
    null = list(
      results = results_null, label = "ACTG-bin",
      n_sample = sim_n_sample, dgm = dgm_null, hypothesis = "null"
    ),
    alt = list(
      results = results_alt, label = "ACTG-bin",
      n_sample = sim_n_sample, dgm = dgm_alt, hypothesis = "alt"
    )
  ),
  analyses = "FS",
  digits = 2, font_size = 16,
  title = "ACTG175 Binary: Classification (FS)",
  n_sims = nsims_alt,
  subgroup_notation = sg_notation
)
```


# Timing Summary

```{r timing}
timing_df <- data.frame(
  Step = names(timings),
  Seconds = round(unlist(timings), 1),
  Minutes = round(unlist(timings) / 60, 2)
)
timing_df |>
  gt() |>
  gt::tab_header(title = "Computation Timing")
```


# References {.unnumbered}

- Hammer, S.M., et al. (1996). *NEJM*, 335, 1081--1090.
- León, L.F., et al. (2024). *Statistics in Medicine*, 43,
  3921--3942.
- Tsiatis, A.A., et al. (2008). *Statistics in Medicine*, 27,
  4658--4677.
