---
title: "Simulation Studies Based on ACTG175: Survival Endpoint (Benefit Search)"
subtitle: "Adapting the ForestSearch Framework to Identify Treatment Benefit in HIV Clinical Trial Data"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ── Simulation control ──────────────────────────────────────────
# Demonstration run; increase to >= 500 for stable operating
# characteristics.  Monte Carlo SE for a 50% detection rate is
# approximately 2.2% at 500 sims and 1.0% at 2500 sims.
nsims_alt   <- 50L
nsims_null  <- 50L

# These are "production" run projection nsims
prod_alt  <- 10000L
prod_null <- 20000L
# 10k/20k ~ 1hr linux

sim_seed    <- 8316951L
n_workers   <- min(
  parallel::detectCores(logical = FALSE) - 1L, 120L
)

# ── Dataset preparation: GBSG ───────────────────────────────────
gbsg_er_quantile <- 0.25          # ER percentile cutpoint for ER-low definition

# ── Dataset preparation: ACTG175 ───────────────────────────────
#
# ARMS (matching Leon et al. 2024, Section 4.2):
#   speff2trial::ACTG175 arm coding:
#     0 = zidovudine (ZDV) alone
#     1 = ZDV + didanosine (ZDV+ddI)   <- experimental
#     2 = ZDV + zalcitabine (ZDV+ddC)
#     3 = didanosine (ddI) alone        <- control
#
# The paper compared arm 1 (ZDV+ddI, combination) vs arm 3 (ddI,
# monotherapy) and searched for a BENEFIT subgroup Q by switching
# treatment roles.
#
# Cutpoint rationale (from Leon et al. 2024, Section 4.2):
#   z1 (age):     age > 34    — median split; paper found Age > 34
#   z2 (preanti): preanti <= 744.5 — q3 split; paper found Preanti <= q3
#   z3 (wtkg):    wtkg <= 75  — GRF-identified prognostic split
#   z4 (Karnofsky): data-driven median
#   z5 (CD4):     cd40 <= median
#   z6 (CD8):     cd80 <= median
#   z7–z12:       binary indicators passed through as-is
#
actg_arms        <- c(1L, 3L)    # ZDV+ddI (arm 1) vs ddI alone (arm 3)
actg_treat_arm   <- 1L           # Arm coded as treatment = 1 (ZDV+ddI)
actg_age_cut     <- 34           # Age cutpoint (years): age > cut -> z1 = 1
actg_preanti_cut <- 744.5        # Preanti cutpoint (days, q3): preanti <= cut -> z2 = 1
actg_wtkg_cut    <- 75           # Weight cutpoint (kg): wtkg <= cut -> z3 = 1
# z4–z6 use data-driven medians; z7–z12 are binary indicators

# ── Benefit search via treatment switching ─────────────────────
# Leon et al. (2024) searched for benefit by SWITCHING treatment
# roles: the combination (ZDV+ddI) was recoded as control and the
# monotherapy (ddI) as treatment.  Under switched roles, the
# BENEFIT subgroup Q appears as a "harm" subgroup (ddI monotherapy
# is worse than combination for these patients).
#
# ForestSearch then runs with elevated thresholds:
#   c1 = 1/0.60 ≈ 1.667  (screening: HR ≥ 1.667 under switched roles
#                          = HR ≤ 0.60 under original roles)
#   c2 = 1/0.80 = 1.25   (consistency: each split HR ≥ 1.25
#                          = HR ≤ 0.80 under original roles)
#
# After identification, hazard ratios are inverted back to the
# original treatment scale.

# ── DGM construction ────────────────────────────────────────────
# The DGM is built on the SWITCHED treatment data so that the
# benefit subgroup Q appears as a harm subgroup in the DGM
# framework.  All downstream simulation and analysis uses the
# switched treatment indicator.
dgm_factor_vars   <- paste0("z", 1:12) # All binary confounder names
dgm_continuous_vars <- character(0)    # No continuous covariates
dgm_subgroup_vars <- c("z1", "z2")    # Variables defining benefit subgroup Q
# Within-level values that define membership in Q (under switched treatment,
# Q appears as "harm"):
#   z1 = 1 means age > actg_age_cut   (older, per paper's finding)
#   z2 = 1 means preanti <= actg_preanti_cut (less prior treatment)
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)
dgm_k_treat       <- 1.0             # Main treatment effect multiplier
dgm_n_super       <- 5000L           # Super-population size for DGM
dgm_k_inter_init  <- 2.0             # Starting k_inter before calibration

# ── Calibration ─────────────────────────────────────────────────
# Target HR(Q) = 2.0 under SWITCHED treatment (= 0.50 under original).
# This means the benefit subgroup has HR = 0.50 under the original
# treatment coding — a strong benefit.
cal_target_hr  <- 2.0
cal_k_grid     <- seq(1.0, 5.0, by = 0.25) # k_inter grid for grid search

# ── Single trial simulation ──────────────────────────────────────
single_n          <- 1000L  # Sample size for single-trial demonstration
single_seed       <- 1L     # Seed for single-trial demonstration
single_rand_ratio <- 1      # Randomisation ratio (1:1)

# ── Simulation study ─────────────────────────────────────────────
sim_n_sample      <- 1000L  # Per-simulation sample size
# Administrative censoring time on the DGM's time scale.
# CRITICAL: must use the SAME units as the outcome variable in the
# source data.  ACTG175 times are converted from days to months
# (days / 30.4375).
sim_analysis_time <- 48     # months 
sim_cens_adjust   <- 0      # Additional censoring adjustment (log-scale)
sim_n_noise       <- 0L     # Number of noise variables added per replicate

# ── Analysis flags ───────────────────────────────────────────────
run_fs      <- TRUE   # Run ForestSearch
run_fs_grf  <- FALSE  # Run ForestSearch with GRF pre-screening only
run_grf     <- TRUE   # Run GRF-only subgroup analysis

# ── ForestSearch parameters (benefit search thresholds) ─────────
# Under switched treatment roles, the thresholds are set to detect
# benefit on the ORIGINAL scale:
#   c1 = 1/0.60 ≈ 1.667: subgroup HR ≥ 1.667 (switched)
#                        = HR ≤ 0.60 (original)
#   c2 = 1/0.80 = 1.25:  split HRs ≥ 1.25 (switched)
#                        = HR ≤ 0.80 (original)
fs_hr_threshold        <- round(1 / 0.60, 3)  # c1 for benefit search
fs_hr_consistency      <- round(1 / 0.80, 3)  # c2 for benefit search
fs_pconsistency_thresh <- 0.90   # Bootstrap consistency proportion threshold
fs_splits              <- 400L   # Number of ForestSearch bootstrap splits
fs_n_min               <- 60L    # Minimum subgroup size (total)
fs_d0_min              <- 12L    # Minimum events in control arm
fs_d1_min              <- 12L    # Minimum events in treatment arm
fs_maxk                <- 2L     # Maximum number of subgroup-defining variables
fs_by_risk             <- 6L     # Risk-table interval (months)
fs_vi_grf_min          <- -0.2   # Minimum GRF variable importance

# ── GRF standalone parameters ───────────────────────────────────
# These control the GRF-only subgroup analysis (run_grf = TRUE).
# dmin.grf is on the RMST scale (months after time-unit conversion).
# The default in default_grf_params_gen() is dmin.grf = 12, which
# was calibrated for days (~0.4 months).  After converting ACTG175
# times to months, this must be rescaled.
grf_dmin               <- 2    # min RMST difference (months)
grf_frac_tau           <- 0.80   # fraction of max event time for horizon
grf_maxdepth           <- 2L     # max policy tree depth
grf_sg_criterion       <- "mDiff"

# ── Adjusted Cox formula (used in run_simulation_analysis) ───────
# Note: treat_sim is the SWITCHED treatment indicator in the
# simulated data.  Confounders z1 and z2 are the subgroup-
# defining variables; z3 is a strong prognostic covariate.
cox_formula_adj <- survival::Surv(y_sim, event_sim) ~
  treat_sim + z1 + z2 + z3

# ── Detection curve ──────────────────────────────────────────────
det_theta_range <- c(0.5, 3.0)  # HR range for detection curve
det_n_points    <- 50L          # Number of points on detection curve

# -- Output persistence -------------------------------------------------------
# Save simulation results to _data/ for downstream analysis.
# Set save_results = FALSE to skip saving (e.g., during development).
save_results           <- TRUE
output_dir             <- "_data"
```


# Introduction {#sec-intro}

## Motivation

Leon et al. (2024) demonstrated the ForestSearch methodology using
the German Breast Cancer Study Group (GBSG) dataset
(`survival::gbsg`) as both the applied analysis target and the
covariate template for simulation studies.  The `setup_gbsg_dgm()`
convenience function encodes all GBSG-specific preparation choices
--- variable discretisation, subgroup definition, Weibull
calibration --- into a single reproducible call.

This document adapts the same simulation framework to the **AIDS
Clinical Trials Group Study 175** (ACTG175) dataset from the
`speff2trial` package.  ACTG175 is a randomised trial of four
antiretroviral regimens in HIV-infected subjects (Hammer et al.,
*NEJM* 1996).  Following León et al. (2024, Section 4.2), we
compare **arm 1** (zidovudine + didanosine, ZDV+ddI) to **arm 3**
(didanosine alone, ddI) and search for a subgroup with
**pronounced treatment benefit** ($N = 1{,}083$).

The adaptation serves two purposes:

1. **Methodological validation**: confirming that the ForestSearch
   operating characteristics generalise beyond the GBSG breast
   cancer setting to a different disease, population, endpoint
   structure, and covariate set --- and specifically to the
   **benefit-search** variant of the algorithm.

2. **Template for new-dataset simulations**: demonstrating the
   recommended workflow when `setup_gbsg_dgm()` is not
   appropriate and the analyst must use
   `generate_aft_dgm_flex()` directly.


## Benefit Search via Treatment Switching {#sec-benefit-search}

ForestSearch's default mode identifies subgroups where treatment
is **detrimental** (hazard ratio $> 1$).  To identify subgroups
with **pronounced benefit** instead, León et al. (2024, §4.2)
introduced a treatment-switching approach:

1. **Switch treatment roles**: recode the treatment indicator so
   that control becomes "treatment" and vice versa.  In the
   ACTG175 context: ddI monotherapy (originally control) is
   recoded as `treat = 1`, and ZDV+ddI combination (originally
   experimental) becomes `treat = 0`.

2. **Raise the thresholds**: under switched roles, a subgroup
   that strongly benefits from the original treatment will appear
   to have a high HR (ddI alone is worse for these patients).
   The paper used:
   - $c_1 = 1/0.60 \approx 1.667$ (screening: original HR
     $\leq 0.60$)
   - $c_2 = 1/0.80 = 1.25$ (consistency: original HR
     $\leq 0.80$)

3. **Run ForestSearch** with these thresholds on the switched
   data.  The identified "harm" subgroup $\hat{H}$ under
   switched roles is the **benefit** subgroup $\hat{Q}$ under
   the original treatment coding.

4. **Invert the results**: hazard ratios reported by
   ForestSearch are inverted ($1/\text{HR}$) to return to the
   original treatment scale.

In the paper's analysis, this approach identified
$\hat{Q} = \{\text{Age} > 34 \cap \text{Preanti} \leq 744.5\}$
with a bias-corrected HR of $0.59$ ($0.37$, $0.94$) ---
a substantially stronger benefit than the ITT HR of $0.84$.

The simulation DGM in this document is built on the
**switched** treatment data.  All ForestSearch analyses run on
the switched indicator.  The calibration target HR $= 2.0$
(under switched treatment) corresponds to an original-scale
HR $= 0.50$, representing a strong treatment benefit in the
subgroup.


## Adaptation Checklist {#sec-checklist}

To substitute your own trial data for ACTG175, you must address
each of the following steps.  The sections that follow implement
each one for ACTG175; cross-references are provided so you can
see the concrete implementation alongside each checklist item.

| Step | What to do | Section |
|------|-----------|---------|
| **1. Load your data** | Create `id`, `y` (observed time), `event` (0/1 indicator), and `treat` (0 = control, 1 = treatment) columns. Time units for `y` set the time scale for the entire simulation. | @sec-comparison |
| **2. Choose confounders** | Identify the candidate variables for the subgroup search. These become `factor_vars` in the DGM and `confounders_base` in the simulation. | @sec-comparison |
| **3. Discretise continuous variables** | Create binary `z1`, `z2`, ... columns from continuous covariates. Document your cutpoint rationale (see @sec-cutpoints). | @sec-comparison |
| **4. Define the harm subgroup** | Set `dgm_subgroup_vars` and `dgm_subgroup_cuts` consistently with your factor coding. Verify with an explicit `stopifnot()` check. | @sec-dgm |
| **5. Set time-scale parameters** | Ensure `sim_analysis_time` is in the same units as your outcome variable. Run the time-scale diagnostic after DGM creation. | @sec-dgm |
| **6. Update `cox_formula_adj`** | Include confounders appropriate for your study. The formula must use simulated column names (`y_sim`, `event_sim`, `treat_sim`). | Setup chunk |
| **7. Update `confounders_base`** | This must list all `z*` variable names passed to ForestSearch and GRF. It must match `factor_vars` in the DGM call. | @sec-sims |
| **8. Calibrate `k_inter`** | Run the grid search targeting a clinically meaningful HR(H). | @sec-calibrate |
| **9. Run time-scale diagnostic** | Verify `exp(dgm$model_params$mu)` is close to your observed median event time. | @sec-timescale-check |
| **10. Run censoring diagnostic** | Use `check_censoring_dgm()` to compare simulated vs observed censoring. | @sec-cens-check |
| **11. Set simulation count** | Use $\geq 500$ replicates for stable operating characteristics. | @sec-sims |


## Choosing Discretisation Cutpoints {#sec-cutpoints}

ForestSearch operates on binary (factor) covariates.  Continuous
variables must be discretised before building the DGM.  There are
three common strategies:

1. **Clinical cutpoints** --- use thresholds defined by domain
   knowledge (e.g., CD4 $< 200$ cells/mm³ for AIDS-defining
   immunosuppression).  Preferred when available.

2. **Data-driven medians** --- split at the sample median;
   appropriate when no clinical threshold exists.  Used here
   for `z3` (Karnofsky performance score) and `z4` (CD4
   baseline count).

3. **GRF-guided splits** --- run a preliminary GRF analysis on the
   observed data (`grf::causal_forest()`) to identify the
   variable-importance-weighted split points that maximise
   treatment effect heterogeneity.  This was the approach for
   `z1` (age $> 34$) and `z2` (preanti $\leq 744.5$), derived from
   the analysis in `actg175_analysis_v1.Rnw`.

For your own dataset, we recommend starting with strategy (1)
where available, falling back to (2), and optionally using (3) as
a sensitivity analysis.  Whichever approach you choose, document
the rationale alongside the cutpoint value in your setup chunk.


## Steps for Swapping GBSG with ACTG175 {#sec-swap-steps}

Replacing the GBSG template with ACTG175 requires changes at every
stage of the simulation pipeline.  The table below summarises each
change; subsequent sections implement them in detail.

| Step | GBSG (original) | ACTG175 (this document) |
|------|-----------------|-------------------------|
| **Data source** | `survival::gbsg` (686 patients) | `speff2trial::ACTG175`, arms 1 vs 3 (~1,083 patients) |
| **Outcome** | Recurrence-free survival in months (`rfstime / 30.4375`) | Time to first event in months (`days / 30.4375`) |
| **Event indicator** | `status` (1 = recurrence/death) | `cens` (1 = event) |
| **Treatment** | `hormon` (0/1) | `arms`: 1 = ZDV+ddI (combination), 3 = ddI (monotherapy); **switched** for benefit search |
| **Confounders** | 7 binary variables: ER, age, meno, PgR, nodes, size, grade | 12 variables: age, preanti, weight, Karnofsky, CD4, CD8, hemophilia, homosexual, IV drugs, race, gender, symptoms |
| **Discretisation** | ER $\leq$ Q25, age $\leq$ median, meno = premenopausal, etc. | Age $>$ 34, preanti $\leq$ 744.5, weight $\leq$ 75 kg, CD4 $\leq$ median, etc. (from León et al. 2024 analysis) |
| **Subgroup definition** | *H* = {ER-low and premenopausal} (~13% prevalence) | *Q* = {Age > 34 and Preanti ≤ 744.5} (~35% prevalence, benefit subgroup) || **DGM builder** | `setup_gbsg_dgm()` | `generate_aft_dgm_flex()` |
| **Censoring** | Weibull; `analysis_time = 84` months | Weibull; `analysis_time = 48` months ($\sim$4 years) |
| **ForestSearch parameters** | `n.min = 60`, `maxk = 2`, `hr.threshold = 1.25` | `hr.threshold = 1.667`, `hr.consistency = 1.25` (benefit search thresholds) |

The benefit subgroup definition for ACTG175 is derived from the
ForestSearch analysis in León et al. (2024, Section 4.2), which
identified Age $> `r actg_age_cut`$ and prior antiretroviral
therapy $\leq `r actg_preanti_cut`$ days as the combination
defining a subgroup with pronounced treatment benefit
($\hat{Q}$: HR $= 0.59$, $0.37$--$0.94$).


# Side-by-Side Dataset Comparison {#sec-comparison}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(survival)
library(speff2trial)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)

# -- Safe foreach result collection -------------------------------------------
# Replaces .errorhandling = "remove" (which silently drops failures)
# with .errorhandling = "pass" + post-hoc separation of errors from results.
collect_results <- function(raw_list, label = "") {
  is_err <- vapply(raw_list, inherits, logical(1), "error")
  n_failed <- sum(is_err)
  if (n_failed > 0) {
    err_msgs <- vapply(raw_list[is_err], conditionMessage, character(1))
    warning(sprintf("%s%d of %d tasks failed.",
        if (nzchar(label)) paste0(label, ": ") else "",
        n_failed, length(raw_list)))
    message("  First errors:\n    ",
        paste(head(unique(err_msgs), 3), collapse = "\n    "))
  }
  good <- raw_list[!is_err]
  if (length(good) == 0) {
    result <- NULL
  } else if (requireNamespace("data.table", quietly = TRUE)) {
    result <- data.table::rbindlist(good, use.names = TRUE, fill = TRUE)
    result <- as.data.frame(result)
  } else {
    result <- do.call(rbind, good)
  }
  attr(result, "n_failed") <- n_failed
  attr(result, "n_total") <- length(raw_list)
  result
}

```

```{r timing-init}
#| include: false
timings <- list()
t_vignette_start <- proc.time()
```

```{r compat-helper}
#| include: false
# Backward-compatibility aliases for generate_aft_dgm_flex() objects.
# Several downstream functions (build_classification_table, etc.)
# access top-level fields (hr_H_true, df_super_rand) that exist on
# gbsg_dgm objects but not on native aft_dgm_flex objects.  This
# helper adds the aliases so the pipeline works with either DGM type.
add_compat_aliases <- function(dgm) {
  if (is.null(dgm$hr_H_true) && !is.null(dgm$hazard_ratios)) {
    dgm$hr_H_true  <- dgm$hazard_ratios$harm_subgroup
    dgm$hr_Hc_true <- dgm$hazard_ratios$no_harm_subgroup
    dgm$hr_causal  <- dgm$hazard_ratios$overall
    dgm$AHR        <- dgm$hazard_ratios$AHR
    dgm$AHR_H_true <- dgm$hazard_ratios$AHR_harm
    dgm$AHR_Hc_true <- dgm$hazard_ratios$AHR_no_harm
  }
  if (is.null(dgm$df_super_rand) && !is.null(dgm$df_super)) {
    dgm$df_super_rand <- dgm$df_super
  }
  dgm
}
```


## Loading and Preparing Both Datasets

The code below prepares both datasets for the simulation framework.
For each dataset, the key steps are:

1. Create a standardised outcome variable (`y`) in the dataset's
   natural time units.
2. Create a binary treatment indicator (`treat`: 0 = control,
   1 = treatment).
3. Discretise continuous covariates into binary factors
   (`z1`, `z2`, ...) using the cutpoints defined in the setup
   chunk.
4. Construct the true harm-subgroup indicator (`flag_harm`)
   using the same variables and levels specified in
   `dgm_subgroup_vars` and `dgm_subgroup_cuts`.

```{r prepare-datasets}
#| code-fold: false

# ── GBSG ──────────────────────────────────────────────────────
gbsg_df        <- survival::gbsg
gbsg_df$y      <- gbsg_df$rfstime / 30.4375
gbsg_df$treat  <- gbsg_df$hormon

er_cut         <- quantile(gbsg_df$er, probs = gbsg_er_quantile)
gbsg_df$z1     <- as.factor(ifelse(gbsg_df$er <= er_cut, 1L, 0L))
gbsg_df$z3     <- as.factor(ifelse(gbsg_df$meno == 0, 1L, 0L))
gbsg_df$flag_harm <- as.integer(gbsg_df$z1 == 1 & gbsg_df$z3 == 1)

# ── ACTG175 ───────────────────────────────────────────────────
# Select arms 1 (ZDV+ddI) and 3 (ddI alone), matching the paper
actg_df        <- subset(speff2trial::ACTG175, arms %in% actg_arms)
actg_df$id     <- seq_len(nrow(actg_df))
actg_df$time_days   <- actg_df$days
actg_df$time_months <- actg_df$time_days / 30.4375

# Original treatment coding: arm 1 (ZDV+ddI) = 1, arm 3 (ddI) = 0
actg_df$treat_orig <- ifelse(actg_df$arms == actg_treat_arm, 1L, 0L)

# SWITCHED treatment coding for benefit search:
# ddI (arm 3) = 1 ("treatment"), ZDV+ddI (arm 1) = 0 ("control")
# Under switched roles, ForestSearch's harm-detection machinery
# identifies the BENEFIT subgroup Q.
actg_df$treat  <- 1L - actg_df$treat_orig

cat(sprintf("ACTG175: N = %d (arm 1: n=%d, arm 3: n=%d)\n",
    nrow(actg_df),
    sum(actg_df$treat_orig == 1), sum(actg_df$treat_orig == 0)))
cat(sprintf("Treatment switched: treat=1 is ddI (original control)\n"))

# ── Discretise continuous variables into binary factors ───────
# Subgroup-defining variables (from Leon et al. 2024, Section 4.2):
# z1: age > actg_age_cut  (subgroup factor 1 — older patients)
actg_df$z1  <- as.factor(ifelse(actg_df$age  > actg_age_cut,  1L, 0L))
# z2: preanti <= actg_preanti_cut  (subgroup factor 2 — less prior treatment)
actg_df$z2  <- as.factor(ifelse(actg_df$preanti <= actg_preanti_cut, 1L, 0L))
# z3: weight <= actg_wtkg_cut (prognostic)
actg_df$z3  <- as.factor(ifelse(actg_df$wtkg <= actg_wtkg_cut, 1L, 0L))
# z4: Karnofsky <= median (data-driven cutpoint)
actg_df$z4  <- as.factor(ifelse(
  actg_df$karnof <= median(actg_df$karnof), 1L, 0L
))
# z5: CD4 baseline <= median (data-driven cutpoint)
actg_df$z5  <- as.factor(ifelse(
  actg_df$cd40 <= median(actg_df$cd40), 1L, 0L
))
# z6: CD8 <= median (data-driven cutpoint)
actg_df$z6  <- as.factor(ifelse(
  actg_df$cd80 <= median(actg_df$cd80), 1L, 0L
))
# z7-z12: binary indicators (passed through as-is from the dataset)
actg_df$z7  <- as.factor(actg_df$hemo)
actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs)
actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender)
actg_df$z12 <- as.factor(actg_df$symptom)

# ── Factor coding summary ──────────────────────────────────────
# z1  = 1 => age > 34          (older — benefit subgroup factor)
# z2  = 1 => preanti <= 744.5  (less prior ART — benefit subgroup factor)
# z3  = 1 => weight <= 75 kg   (low weight)
# z4  = 1 => Karnofsky <= median (lower performance)
# z5  = 1 => CD4 <= median     (lower baseline CD4)
# z6  = 1 => CD8 <= median     (lower CD8)
# z7  = 1 => hemophilia yes
# z8  = 1 => homosexual activity yes
# z9  = 1 => IV drug use yes
# z10 = 1 => non-white
# z11 = 1 => male
# z12 = 1 => symptomatic yes

# ── Benefit subgroup: z1 = 1 AND z2 = 1 ───────────────────────
# Q = {Age > 34 AND Preanti <= 744.5}
# Under switched treatment, Q appears as the "harm" subgroup.
actg_df$flag_harm <- as.integer(
  actg_df[[dgm_subgroup_vars[1]]] == dgm_subgroup_cuts[[dgm_subgroup_vars[1]]] &
  actg_df[[dgm_subgroup_vars[2]]] == dgm_subgroup_cuts[[dgm_subgroup_vars[2]]]
)

# Verify subgroup definition matches expectations
stopifnot(
  identical(dgm_subgroup_vars, c("z1", "z2")),
  identical(dgm_subgroup_cuts, list(z1 = 1L, z2 = 1L))
)
cat(sprintf(
  "Benefit subgroup Q: z1 == 1 (age > %d) AND z2 == 1 (preanti <= %.1f)\n",
  actg_age_cut, actg_preanti_cut
))
cat(sprintf(
  "  Q prevalence: %d / %d (%.1f%%)\n",
  sum(actg_df$flag_harm), nrow(actg_df),
  100 * mean(actg_df$flag_harm)
))
```


## Structural Comparison

```{r dataset-comparison-table}
#| tbl-cap: "Side-by-side comparison of GBSG and ACTG175 datasets"

comp <- data.frame(
  Feature = c(
    "Disease area",
    "Trial name",
    "R package",
    "Total N (used)",
    "Treatment arms",
    "Outcome",
    "Time unit",
    "Median follow-up",
    "Event rate",
    "Treatment effect (ITT HR)",
    "Subgroup definition (H)",
    "H prevalence",
    "N confounders (binary)",
    "Confounder types"
  ),
  GBSG = c(
    "Breast cancer",
    "German Breast Cancer Study Group",
    "survival::gbsg",
    sprintf("%d", nrow(gbsg_df)),
    "Hormonal therapy vs. control",
    "Recurrence-free survival",
    "Months",
    sprintf("%.1f months", median(gbsg_df$y)),
    sprintf("%.0f%%", 100 * mean(gbsg_df$status)),
    sprintf("%.3f", exp(coef(coxph(
      Surv(y, status) ~ treat, data = gbsg_df
    )))),
    sprintf("ER-low (Q%d) AND premenopausal",
            as.integer(gbsg_er_quantile * 100)),
    sprintf("%.0f%%", 100 * mean(gbsg_df$flag_harm)),
    "7",
    "All pre-discretised from continuous/categorical"
  ),
  ACTG175 = c(
    "HIV / AIDS",
    "AIDS Clinical Trials Group Study 175",
    "speff2trial::ACTG175",
    sprintf("%d", nrow(actg_df)),
    "ZDV+ddI (combination) vs. ddI (monotherapy)",
    "Time to >= 50% CD4 decline, AIDS, or death",
    "Months",
    sprintf("%.1f months", median(actg_df$time_months)),
    sprintf("%.0f%%", 100 * mean(actg_df$cens)),
    sprintf("%.3f", exp(coef(coxph(
      Surv(time_months, cens) ~ treat_orig, data = actg_df
    )))),
    sprintf("Age > %d AND Preanti <= %.0f", actg_age_cut, actg_preanti_cut),
    sprintf("%.0f%%", 100 * mean(actg_df$flag_harm)),
    "12",
    "Mix of binary indicators and discretised continuous"
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt(comp) |>
  tab_header(
    title = "Dataset Comparison: GBSG vs ACTG175"
  ) |>
  cols_label(Feature = "") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Feature)
  ) |>
  tab_style(
    style = cell_fill(color = "#F5F5F5"),
    locations = cells_body(rows = seq(1, nrow(comp), by = 2))
  )
```


## Confounder Mapping

```{r confounder-mapping}
#| tbl-cap: "Mapping of confounder variables between datasets"

conf_map <- data.frame(
  Role = c(
    "Subgroup factor 1",
    "Subgroup factor 2",
    "Prognostic 1",
    "Prognostic 2",
    "Prognostic 3",
    "Prognostic 4",
    "Prognostic 5",
    "Demographic 1",
    "Demographic 2",
    "Demographic 3",
    "Demographic 4",
    "Demographic 5"
  ),
  `GBSG variable` = c(
    sprintf("z1: ER <= Q%d", as.integer(gbsg_er_quantile * 100)),
    "z3: premenopausal",
    "z2: age <= median",
    "z4: PgR <= median",
    "z5: nodes <= median",
    "v6: tumour size <= median",
    "v7: grade = 3",
    rep("---", 5)
  ),
  `ACTG175 variable` = c(
    sprintf("z1: age > %d", actg_age_cut),
    sprintf("z2: preanti <= %.0f days", actg_preanti_cut),
    sprintf("z3: weight <= %d kg", actg_wtkg_cut),
    "z4: Karnofsky <= median",
    "z5: CD4 baseline <= median",
    "z6: CD8 <= median",
    "z7: hemophilia",
    "z8: homosexual activity",
    "z9: IV drug use",
    "z10: race (non-white)",
    "z11: gender (male)",
    "z12: symptomatic"
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt(conf_map) |>
  tab_header(
    title = "Confounder Variable Mapping"
  ) |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(rows = 1:2)
  ) |>
  tab_source_note(
    "Highlighted rows define the harm subgroup H."
  )
```


## Survival Curves (Observed Data)

```{r km-comparison, fig.height=5, fig.width=10}
#| fig-cap: "Kaplan-Meier curves in the observed datasets. Left: GBSG (hormonal therapy). Right: ACTG175 (ZDV+ddI vs ddI, original treatment coding)."

# ── GBSG KM ──────────────────────────────────────────────────
km_gbsg    <- survfit(Surv(y, status) ~ treat, data = gbsg_df)
km_gbsg_df <- data.frame(
  time = c(km_gbsg$time[1:km_gbsg$strata[1]],
           km_gbsg$time[(km_gbsg$strata[1] + 1):sum(km_gbsg$strata)]),
  surv = c(km_gbsg$surv[1:km_gbsg$strata[1]],
           km_gbsg$surv[(km_gbsg$strata[1] + 1):sum(km_gbsg$strata)]),
  arm = c(
    rep("Control", km_gbsg$strata[1]),
    rep("Hormonal", km_gbsg$strata[2])
  )
)

p_gbsg <- ggplot(km_gbsg_df, aes(x = time, y = surv, colour = arm)) +
  geom_step(linewidth = 0.8) +
  labs(
    title = "GBSG",
    x = "Months", y = "Survival Probability",
    colour = "Treatment"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# ── ACTG175 KM ───────────────────────────────────────────────
km_actg <- survfit(Surv(time_months, cens) ~ treat_orig, data = actg_df)
km_actg_df <- data.frame(
  time = c(km_actg$time[1:km_actg$strata[1]],
           km_actg$time[(km_actg$strata[1] + 1):sum(km_actg$strata)]),
  surv = c(km_actg$surv[1:km_actg$strata[1]],
           km_actg$surv[(km_actg$strata[1] + 1):sum(km_actg$strata)]),
  arm = c(
    rep("ddI alone (control)", km_actg$strata[1]),
    rep("ZDV+ddI (experimental)", km_actg$strata[2])
  )
)

p_actg <- ggplot(km_actg_df, aes(x = time, y = surv, colour = arm)) +
  geom_step(linewidth = 0.8) +
  labs(
    title = "ACTG175",
    x = "Months", y = "Survival Probability",
    colour = "Treatment"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

p_gbsg + p_actg +
  plot_annotation(
    title = "Observed Survival in Source Datasets",
    theme = theme(plot.title = element_text(face = "bold"))
  )
```


## Subgroup Prevalence and Outcome

```{r subgroup-preview}
#| code-fold: false

cat("=== GBSG: Harm subgroup H = {ER-low AND premenopausal} ===\n")
cat("  N(H):", sum(gbsg_df$flag_harm), "/", nrow(gbsg_df),
    sprintf("(%.1f%%)\n", 100 * mean(gbsg_df$flag_harm)))
fit_gbsg_h <- coxph(
  Surv(y, status) ~ treat,
  data = gbsg_df[gbsg_df$flag_harm == 1, ]
)
cat("  Observed HR(H):",
    round(exp(coef(fit_gbsg_h)), 3), "\n\n")

cat(sprintf(
  "=== ACTG175: Benefit subgroup Q = {age>%d AND preanti<=%.0f} ===\n",
  actg_age_cut, actg_preanti_cut
))
cat("  N(Q):", sum(actg_df$flag_harm), "/", nrow(actg_df),
    sprintf("(%.1f%%)\n", 100 * mean(actg_df$flag_harm)))
fit_actg_h <- coxph(
  Surv(time_months, cens) ~ treat_orig,
  data = actg_df[actg_df$flag_harm == 1, ]
)
cat("  Observed HR(Q), original scale:",
    round(exp(coef(fit_actg_h)), 3), "\n")
fit_actg_hc <- coxph(
  Surv(time_months, cens) ~ treat_orig,
  data = actg_df[actg_df$flag_harm == 0, ]
)
cat("  Observed HR(Qc), original scale:",
    round(exp(coef(fit_actg_hc)), 3), "\n")
```


# Creating the ACTG175 DGM {#sec-dgm}

Because `setup_gbsg_dgm()` is specific to the GBSG dataset, we use
`generate_aft_dgm_flex()` directly.  This requires the analyst to
supply the prepared dataset, specify all modelling choices, and
define the subgroup explicitly.

The key function arguments are:

- `data`: the prepared analysis dataset with `y`, `event`,
  `treat`, and all `z*` factor columns.
- `factor_vars`: character vector of **all** binary confounder
  names (here, `z1`--`z12`).  These are used as covariates in
  the AFT model fitted to the observed data.
- `subgroup_vars` / `subgroup_cuts`: define the harm subgroup
  $H$.  `subgroup_cuts` specifies the factor level (here, `1L`)
  that indicates membership.
- `model`: `"alt"` embeds a treatment-by-subgroup interaction;
  `"null"` fits a uniform treatment effect (no HTE).
- `k_inter`: multiplier on the interaction coefficient (AFT
  log-time scale).  Higher values produce larger HR(H).
- `k_treat`: multiplier on the main treatment coefficient.
  Usually `1.0` (retain the observed ITT effect).
- `n_super`: size of the super-population drawn from the fitted
  AFT model.  $5000$ is sufficient for stable HR estimates.


## Step 1: Prepare the Analysis Dataset

The DGM is built on the **switched** treatment data so that the
benefit subgroup $Q$ appears as a "harm" subgroup in the standard
ForestSearch framework.  The `treat` column in `dfa` is already
the switched indicator (ddI = 1, ZDV+ddI = 0).

```{r prepare-actg-dgm}
#| code-fold: false

# Prepare the ACTG175 data for generate_aft_dgm_flex()
# Column names here must match the *_var arguments below.
# Note: 'treat' is already the SWITCHED indicator from the prep step.
dfa        <- actg_df
dfa$y      <- dfa$time_months # outcome in months (days / 30.4375)
dfa$event  <- dfa$cens        # 1 = event, 0 = censored
# 'treat' is the switched indicator (ddI=1, ZDV+ddI=0)

cat("DGM input data:\n")
cat("  N =", nrow(dfa), "\n")
cat("  treat = 1 (ddI monotherapy):", sum(dfa$treat == 1), "\n")
cat("  treat = 0 (ZDV+ddI combination):", sum(dfa$treat == 0), "\n")
cat("  (Treatment is SWITCHED for benefit search)\n")
```


## Step 2: Build the DGM (Alternative Hypothesis)

```{r create-dgm-alt}
t0 <- proc.time()[3]

dgm_alt <- generate_aft_dgm_flex(
  data            = dfa,
  outcome_var     = "y",
  event_var       = "event",
  treatment_var   = "treat",
  continuous_vars = dgm_continuous_vars,
  factor_vars     = dgm_factor_vars,
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  model           = "alt",
  k_inter         = dgm_k_inter_init,
  k_treat         = dgm_k_treat,
  n_super         = dgm_n_super,
  seed            = sim_seed,
  verbose         = TRUE
)

# Enrich the DGM with controlled direct effect (CDE) quantities.
# compute_dgm_cde() adds individual-level potential outcomes
# (theta_0, theta_1, loghr_po) to the super-population.  These are
# required by build_estimation_table() and interpret_estimation_table()
# to compute CDE bias and RMSE metrics.  Always call this after
# creating any alternative-hypothesis DGM.
dgm_alt <- compute_dgm_cde(dgm_alt)
dgm_alt <- add_compat_aliases(dgm_alt)

cat("\n=== ACTG175 Alternative DGM ===\n")
cat("  HR(H):",   round(dgm_alt$hazard_ratios$harm_subgroup,      3), "\n")
cat("  HR(Hc):",  round(dgm_alt$hazard_ratios$no_harm_subgroup,     3), "\n")
cat("  HR(ITT):", round(dgm_alt$hazard_ratios$overall, 3), "\n")
cat("  AHR(H):",  round(dgm_alt$hazard_ratios$AHR_harm,     3), "\n")
cat("  AHR(Hc):", round(dgm_alt$hazard_ratios$AHR_no_harm,    3), "\n")
cat("  H prevalence:",
    round(100 * mean(dgm_alt$df_super$flag_harm), 1), "%\n")

timings$dgm_alt <- proc.time()[3] - t0
```

### Time-Scale Sanity Check {#sec-timescale-check}

The single most common simulation failure is building the DGM on one
time scale (e.g., months) and then passing `analysis_time` in a different
scale.  This results in a follow-up window far shorter
or longer than the DGM event-time distribution, producing either
universal censoring or no censoring at all.

The diagnostic below verifies that the DGM's implied median event time
(`exp(mu)`) is in the same ballpark as the observed data and that
`analysis_time` is on the correct scale.

```{r timescale-check}
#| code-fold: false

observed_median <- median(dfa$y[dfa$event == 1])
dgm_median      <- exp(dgm_alt$model_params$mu)

cat(sprintf("Observed median event time:  %.1f months\n", observed_median))
cat(sprintf("DGM implied median (exp(mu)): %.1f months\n", dgm_median))
cat(sprintf("analysis_time:                %.0f months\n", sim_analysis_time))

if (sim_analysis_time < dgm_median * 0.5) {
  warning("analysis_time is much shorter than DGM median -- ",
          "expect heavy administrative censoring!")
}
if (sim_analysis_time > dgm_median * 10) {
  warning("analysis_time is much longer than DGM median -- ",
          "nearly all events will be observed (minimal censoring).")
}
```


## Step 3: Calibrate for Target HR {#sec-calibrate}

### Why calibrate?

The DGM's `k_inter` parameter controls the strength of the
treatment-by-subgroup interaction on the AFT (log-time) scale.
Because the mapping from `k_inter` to the Cox-model hazard ratio
HR(H) is non-linear and depends on the specific covariate
distribution, censoring structure, and baseline hazard, we cannot
set `k_inter` analytically to achieve a desired HR(H).  Instead,
we search over a grid of candidate values and select the one that
yields an HR(H) closest to the target.

A target HR(H) of `r cal_target_hr` is used here because it
represents a clinically meaningful level of harm --- treatment
doubles the hazard in the subgroup --- and matches the simulation
scenarios in León et al. (2024).  Users should choose their target
based on the effect size they want their simulation to be powered
to detect.

**Note:** `calibrate_k_inter()` automates this process for GBSG-based
DGMs.  For non-GBSG datasets, the manual grid search shown here is
required.  This is a known limitation and will be addressed in a
future release.

```{r calibrate-k-inter}
t0 <- proc.time()[3]

cal_results <- data.frame(
  k_inter = cal_k_grid,
  hr_H    = NA_real_,
  hr_Hc   = NA_real_,
  stringsAsFactors = FALSE
)

for (i in seq_along(cal_k_grid)) {
  dgm_k <- tryCatch(
    generate_aft_dgm_flex(
      data            = dfa,
      outcome_var     = "y",
      event_var       = "event",
      treatment_var   = "treat",
      continuous_vars = dgm_continuous_vars,
      factor_vars     = dgm_factor_vars,
      subgroup_vars   = dgm_subgroup_vars,
      subgroup_cuts   = dgm_subgroup_cuts,
      model           = "alt",
      k_inter         = cal_k_grid[i],
      k_treat         = dgm_k_treat,
      n_super         = dgm_n_super,
      seed            = sim_seed,
      verbose         = FALSE
    ),
    error = function(e) NULL
  )

  if (!is.null(dgm_k)) {
    cal_results$hr_H[i]  <- dgm_k$hazard_ratios$harm_subgroup
    cal_results$hr_Hc[i] <- dgm_k$hazard_ratios$no_harm_subgroup
  }
}

# Find k_inter closest to target
cal_valid  <- cal_results[!is.na(cal_results$hr_H), ]
best_idx   <- which.min(abs(cal_valid$hr_H - cal_target_hr))
k_inter_best <- cal_valid$k_inter[best_idx]

cat(sprintf(
  "Calibration: k_inter = %.2f gives HR(H) = %.3f (target: %.1f)\n",
  k_inter_best, cal_valid$hr_H[best_idx], cal_target_hr
))

timings$calibrate <- proc.time()[3] - t0
```


```{r calibration-plot, fig.height=5}
#| fig-cap: "Calibration of k_inter: HR(H) as a function of the interaction parameter."

ggplot(cal_valid, aes(x = k_inter)) +
  geom_line(aes(y = hr_H,  colour = "HR(H)"),  linewidth = 1) +
  geom_line(aes(y = hr_Hc, colour = "HR(Hc)"), linewidth = 1) +
  geom_hline(yintercept = cal_target_hr, linetype = "dashed",
             colour = "grey50") +
  geom_hline(yintercept = 1.0, linetype = "dotted",
             colour = "grey70") +
  geom_point(
    data = cal_valid[best_idx, ],
    aes(x = k_inter, y = hr_H),
    colour = "red", size = 4, shape = 16
  ) +
  scale_colour_manual(
    values = c("HR(H)" = "firebrick", "HR(Hc)" = "steelblue")
  ) +
  labs(
    x = "k_inter",
    y = "Hazard Ratio",
    title = "ACTG175 DGM Calibration",
    subtitle = sprintf(
      "Best: k_inter = %.2f -> HR(H) = %.3f (target: %.1f)",
      k_inter_best, cal_valid$hr_H[best_idx], cal_target_hr
    ),
    colour = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Step 4: Create Calibrated DGM

```{r dgm-calibrated}
t0 <- proc.time()[3]

dgm_calibrated <- generate_aft_dgm_flex(
  data            = dfa,
  outcome_var     = "y",
  event_var       = "event",
  treatment_var   = "treat",
  continuous_vars = dgm_continuous_vars,
  factor_vars     = dgm_factor_vars,
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  model           = "alt",
  k_inter         = k_inter_best,
  k_treat         = dgm_k_treat,
  n_super         = dgm_n_super,
  seed            = sim_seed,
  verbose         = TRUE
)
# Enrich with CDE potential outcomes (required for estimation tables)
dgm_calibrated <- compute_dgm_cde(dgm_calibrated)
dgm_calibrated <- add_compat_aliases(dgm_calibrated)

cat("\n=== Calibrated ACTG175 DGM ===\n")
cat("  HR(H):",   round(dgm_calibrated$hazard_ratios$harm_subgroup,      3), "\n")
cat("  HR(Hc):",  round(dgm_calibrated$hazard_ratios$no_harm_subgroup,     3), "\n")
cat("  HR(ITT):", round(dgm_calibrated$hazard_ratios$overall, 3), "\n")

timings$dgm_calibrated <- proc.time()[3] - t0
```


## Step 5: Create Null DGM

Under `model = "null"`, the treatment effect is uniform across
all patients (no treatment effect heterogeneity).  The `k_inter`
argument is irrelevant under the null and defaults to `1`.  Any
subgroup detected under the null is a false positive.

```{r dgm-null}
t0 <- proc.time()[3]

dgm_null <- generate_aft_dgm_flex(
  data            = dfa,
  outcome_var     = "y",
  event_var       = "event",
  treatment_var   = "treat",
  continuous_vars = dgm_continuous_vars,
  factor_vars     = dgm_factor_vars,
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  model           = "null",
  k_treat         = dgm_k_treat,
  n_super         = dgm_n_super,
  seed            = sim_seed,
  verbose         = TRUE
)
dgm_null <- add_compat_aliases(dgm_null)

cat("\n=== ACTG175 Null DGM ===\n")
cat("  HR(ITT):", round(dgm_null$hazard_ratios$overall, 3), "\n")

timings$dgm_null <- proc.time()[3] - t0
```


# Single Trial Simulation {#sec-single-trial}

Before running the full simulation study, we generate a single
simulated trial to verify that the pipeline produces sensible data.
This is a useful sanity check: the simulated event rate, subgroup
prevalence, and ITT HR should be broadly consistent with the DGM
parameters.

```{r simulate-single}
sim_data <- simulate_from_dgm(
  dgm           = dgm_calibrated,
  n             = single_n,
  rand_ratio    = single_rand_ratio,
  seed          = single_seed,
  analysis_time = sim_analysis_time,
  cens_adjust   = sim_cens_adjust
)

cat("Simulated ACTG175 trial:\n")
cat("  N =", nrow(sim_data), "\n")
cat("  Events =", sum(sim_data$event_sim),
    sprintf("(%.1f%%)\n", 100 * mean(sim_data$event_sim)))
cat("  Harm subgroup size =", sum(sim_data$flag_harm),
    sprintf("(%.1f%%)\n", 100 * mean(sim_data$flag_harm)))

fit_sim <- coxph(Surv(y_sim, event_sim) ~ treat_sim, data = sim_data)
cat("  Estimated ITT HR =", round(exp(coef(fit_sim)), 3), "\n")
```

## Censoring Diagnostic {#sec-cens-check}

The `check_censoring_dgm()` function compares the simulated censoring
distribution against the observed data used to build the DGM.  Large
discrepancies suggest a time-scale mismatch or a poorly calibrated
censoring model.

```{r censoring-diagnostic}
check_censoring_dgm(
  sim_data = sim_data,
  dgm      = dgm_calibrated
)
```


## Baseline Characteristics (Simulated Trial) {#sec-single-demog}

```{r single-demographics}
# Label treatment arms for the table
sim_data$treat_label <- factor(
  sim_data$treat_sim,
  levels = c(0, 1),
  labels = c("Combo (ZDV+ddI)", "ddI monotherapy")
)

# Simulated data has z1-z12 as factor columns from the DGM super-population.
# Identify which z-variables are present
z_vars <- grep("^z[0-9]+$", names(sim_data), value = TRUE)
z_vars_present <- z_vars[z_vars %in% names(sim_data)]

create_summary_table(
  data       = sim_data,
  treat_var  = "treat_label",
  vars_continuous  = "y_sim",
  vars_categorical = z_vars_present,
  var_labels = c(y_sim = "Follow-up time (months)"),
  table_title = "Simulated Trial: Baseline Characteristics by Treatment Arm",
  table_subtitle = sprintf("N = %d, Events = %d (%.1f%%)",
    nrow(sim_data), sum(sim_data$event_sim),
    100 * mean(sim_data$event_sim)),
  font_size = 11
)
```


## Kaplan-Meier Analysis (Simulated Trial) {#sec-single-km}

```{r single-km}
#| fig-width: 8
#| fig-height: 5.5
#| fig-cap: "Kaplan-Meier curves for a single simulated ACTG175 trial (switched treatment scale)."

dfcount_sim <- weightedsurv::df_counting(
  df        = sim_data,
  by.risk   = fs_by_risk,
  tte.name  = "y_sim",
  event.name = "event_sim",
  treat.name = "treat_sim"
)

weightedsurv::plot_weighted_km(
  dfcount_sim,
  conf.int     = TRUE,
  show.logrank = TRUE,
  ymax         = 1.05,
  xmed.fraction = 0.775,
  ymed.offset   = 0.125
)
title(main = "Simulated ACTG175 Trial: ITT Kaplan-Meier",
      sub  = sprintf("N = %d | Events = %d | Switched treatment (ddI = treat)",
                      nrow(sim_data), sum(sim_data$event_sim)),
      cex.sub = 0.85)
```

```{r single-cox-summary}
fit_sim_adj <- coxph(
  Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
  data = sim_data
)

cat("=== Single Simulated Trial: Cox PH Summary ===\n")
cat(sprintf("  Unadjusted HR: %.3f\n", exp(coef(coxph(
  Surv(y_sim, event_sim) ~ treat_sim, data = sim_data)))))
cat(sprintf("  Adjusted HR:   %.3f (z1 + z2 + z3)\n",
    exp(coef(fit_sim_adj)["treat_sim"])))
```


## ForestSearch on Simulated Trial {#sec-single-fs}

```{r single-fs}
t0 <- proc.time()[3]

fs_single <- tryCatch(
  suppressWarnings(forestsearch(
    df.analysis            = sim_data,
    confounders.name       = dgm_calibrated$factor_vars,
    outcome.name           = "y_sim",
    event.name             = "event_sim",
    treat.name             = "treat_sim",
    id.name                = "id",
    hr.threshold           = fs_hr_threshold,
    hr.consistency         = fs_hr_consistency,
    pconsistency.threshold = fs_pconsistency_thresh,
    fs.splits              = fs_splits,
    n.min                  = fs_n_min,
    d0.min                 = fs_d0_min,
    d1.min                 = fs_d1_min,
    maxk                   = fs_maxk,
    by.risk                = fs_by_risk,
    vi.grf.min             = fs_vi_grf_min,
    use_lasso              = TRUE,
    use_grf                = TRUE,
    is.RCT                 = TRUE,
    details                = TRUE,
    quiet                  = FALSE,
    seedit                 = sim_seed,
    parallel_args          = list(
      plan = "sequential", workers = 1, show_message = FALSE
    )
  )),
  error = function(e) {
    warning("ForestSearch failed: ", e$message)
    NULL
  }
)

timings$single_fs <- proc.time()[3] - t0
cat(sprintf("\nForestSearch completed in %.1f sec\n", timings$single_fs))
```

```{r single-fs-results}
fs_single_detected <- !is.null(fs_single) &&
  !is.null(fs_single$grp.consistency) &&
  !is.null(fs_single$grp.consistency$sg.harm.id) &&
  length(fs_single$grp.consistency$sg.harm.id) > 0

if (fs_single_detected) {
  sg_id <- fs_single$grp.consistency$sg.harm.id
  cat("=== ForestSearch Result (Single Simulated Trial) ===\n")
  cat("  Subgroup:", paste(sg_id, collapse = " & "), "\n")

  df_est <- fs_single$df.est
  in_H   <- df_est$treat.recommend == 0
  true_Q <- sim_data$flag_harm == 1

  tp <- sum(in_H & true_Q); fp <- sum(in_H & !true_Q)
  fn <- sum(!in_H & true_Q); tn <- sum(!in_H & !true_Q)
  cat(sprintf("  n(H) = %d (%.1f%%), n(Hc) = %d\n",
      sum(in_H), 100 * mean(in_H), sum(!in_H)))
  cat(sprintf("  Sensitivity: %.3f\n", tp / max(tp + fn, 1)))
  cat(sprintf("  Specificity: %.3f\n", tn / max(tn + fp, 1)))
  cat(sprintf("  PPV:         %.3f\n", tp / max(tp + fp, 1)))
} else {
  cat("ForestSearch did not identify a subgroup in this single trial.\n")
  cat("This can happen with moderate effect sizes at n =", single_n, "\n")
}
```


### KM by Identified Subgroups

```{r single-fs-km}
#| fig-width: 10
#| fig-height: 7
#| fig-cap: "Kaplan-Meier curves by ForestSearch-identified subgroup (single simulated trial)."

if (fs_single_detected) {
  km_result <- plot_sg_weighted_km(
    fs.est       = fs_single,
    outcome.name = "y_sim",
    event.name   = "event_sim",
    treat.name   = "treat_sim",
    by.risk      = fs_by_risk,
    conf.int     = TRUE,
    show.logrank = TRUE,
    show.cox     = TRUE,
    show.cox.bc  = FALSE,
    ymax         = 1.05,
    xmed.fraction = 0.65,
    title        = "Simulated ACTG175: KM by Identified Subgroup"
  )
} else {
  cat("No subgroup identified -- KM plot skipped.\n")
}
```

::: {.callout-note}
The subgroup identified in this single trial may differ from the true
$Q$ defined by the DGM. The simulation study below evaluates how often
ForestSearch correctly recovers the true subgroup across many replicates.
:::


# Running Simulation Studies {#sec-sims}

## Setup

```{r sim-setup}
#| code-fold: false

plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")

# ── Candidate variable set ────────────────────────────────────
# confounders_base tells run_simulation_analysis() which columns in
# the simulated data to pass to ForestSearch and GRF as candidate
# subgroup-defining variables.  These must:
#   1. Be present in the super-population (dgm$df_super)
#   2. Be binary factors (0/1) for ForestSearch
#   3. Match the names used in factor_vars when building the DGM
#
# If your dataset uses different variable names (e.g. "age_bin",
# "cd4_low"), update BOTH factor_vars (in the DGM call) and
# confounders_base (here) to match.
confounders_actg <- dgm_factor_vars

# ── ForestSearch parameters ───────────────────────────────────
# These are passed to run_simulation_analysis() and forwarded to
# forestsearch().  The annotations below explain each parameter
# and when you might change it for your own study.
fs_params <- list(
  # Column name mappings — must match simulate_from_dgm() output
  outcome.name           = "y_sim",
  event.name             = "event_sim",
  treat.name             = "treat_sim",
  id.name                = "id",

  # Variable screening flags
  use_lasso              = TRUE,    # LASSO pre-screening (recommended)
  use_grf                = TRUE,    # GRF pre-screening (recommended)

  # Subgroup detection thresholds (see Leon et al. 2024, Section 3)
  # c1: a candidate subgroup must have HR > hr.threshold to be flagged.
  # Increase for more conservative detection; decrease for more power.
  hr.threshold           = fs_hr_threshold,
  # c2: in each bootstrap replicate, the subgroup HR must exceed this
  # value.  Typically set to 1.0 (any harm signal).
  hr.consistency         = fs_hr_consistency,
  # The proportion of bootstrap replicates in which the subgroup HR
  # exceeds c2.  The paper default is 0.90.
  pconsistency.threshold = fs_pconsistency_thresh,

  # Computational and sample-size constraints
  fs.splits              = fs_splits,    # Bootstrap replicates (400 is default)
  n.min                  = fs_n_min,     # Minimum subgroup size (both arms)
  d0.min                 = fs_d0_min,    # Minimum events, control arm
  d1.min                 = fs_d1_min,    # Minimum events, treatment arm
  maxk                   = fs_maxk,      # Max subgroup-defining variables
  by.risk                = fs_by_risk,   # Risk-table granularity (time units)
  vi.grf.min             = fs_vi_grf_min, # GRF variable importance threshold

  # Two-stage estimation: after identifying a subgroup, re-estimates
  # the treatment effect using an independent hold-out sample.
  # Recommended for bias correction.
  use_twostage           = TRUE,
  twostage_args          = list(),

  # Inner parallelism: ForestSearch runs sequentially within each
  # worker.  The outer loop (foreach) parallelises across replicates.
  parallel_args          = list(
    plan = "sequential", workers = 1, show_message = FALSE
  )
)

# -- GRF standalone parameter overrides ----------------------------------------
grf_params <- list(
  dmin.grf     = grf_dmin,
  frac.tau     = grf_frac_tau,
  maxdepth     = grf_maxdepth,
  sg.criterion = grf_sg_criterion
)

# ── Simulation configuration ─────────────────────────────────
sim_config <- list(
  n_sample      = sim_n_sample,
  analysis_time = sim_analysis_time,
  cens_adjust   = sim_cens_adjust
)
```


## Alternative Hypothesis Simulations

```{r run-alt}
cat("Running", nsims_alt, "simulations under H1...\n")
t0 <- proc.time()[3]

results_alt_raw <- foreach(
  sim = seq_len(nsims_alt),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  # Reset plan inside workers to prevent nested parallelism.
  # Each worker runs ForestSearch sequentially; the outer foreach
  # loop parallelises across simulation replicates.
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_calibrated,
    n_sample         = sim_config$n_sample,
    analysis_time    = sim_config$analysis_time,
    cens_adjust      = sim_config$cens_adjust,
    confounders_base = confounders_actg,
    cox_formula_adj  = cox_formula_adj,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    grf_params       = grf_params,
    verbose          = FALSE
  )
}

timings$sims_alt <- proc.time()[3] - t0
results_alt <- collect_results(results_alt_raw, "H1 alt")
n_failed_alt <- attr(results_alt, "n_failed")
cat("Completed in", round(timings$sims_alt / 60, 1), "minutes\n")
if (n_failed_alt > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_alt))
cat("Results:", nrow(results_alt), "rows\n")
```


## Null Hypothesis Simulations

```{r run-null}
cat("Running", nsims_null, "simulations under H0...\n")
t0 <- proc.time()[3]

results_null_raw <- foreach(
  sim = seq_len(nsims_null),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  # Reset plan inside workers to prevent nested parallelism.
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_null,
    n_sample         = sim_config$n_sample,
    analysis_time    = sim_config$analysis_time,
    cens_adjust      = sim_config$cens_adjust,
    confounders_base = confounders_actg,
    cox_formula_adj  = cox_formula_adj,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    grf_params       = grf_params,
    verbose          = FALSE
  )
}

timings$sims_null <- proc.time()[3] - t0
results_null <- collect_results(results_null_raw, "H0 null")
n_failed_null <- attr(results_null, "n_failed")
cat("Completed in", round(timings$sims_null / 60, 1), "minutes\n")
if (n_failed_null > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_null))
```

```{r cleanup-plan}
plan(sequential)
```

```{r save-results}
#| code-fold: false

if (save_results) {
  if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

  # -- Failure accounting via collect_results() above ---------------------------

  bundle <- list(
    # -- Core results ----------------------------------------------------------
    results_alt  = results_alt,
    results_null = results_null,

    # -- DGM objects -----------------------------------------------------------
    dgm_calibrated = dgm_calibrated,
    dgm_null       = dgm_null,

    # -- Simulation configuration ----------------------------------------------
    config = list(
      nsims_alt     = nsims_alt,
      nsims_null    = nsims_null,
      n_sample      = sim_config$n_sample,
      analysis_time = sim_config$analysis_time,
      cens_adjust   = sim_config$cens_adjust,
      n_noise       = sim_n_noise,
      sim_seed      = sim_seed,
      n_workers     = n_workers,
      outcome_type  = "survival"
    ),

    # -- ForestSearch parameters -----------------------------------------------
    fs_params = fs_params,

    # -- DGM true effects (convenience) ----------------------------------------
    true_effects = list(
      hr_Q    = dgm_calibrated$hazard_ratios$harm_subgroup,
      hr_Qc   = dgm_calibrated$hazard_ratios$no_harm_subgroup,
      hr_ITT  = dgm_calibrated$hazard_ratios$overall,
      ahr_Q   = dgm_calibrated$hazard_ratios$AHR_harm,
      ahr_Qc  = dgm_calibrated$hazard_ratios$AHR_no_harm,
      k_inter = k_inter_best
    ),

    # -- Analysis flags --------------------------------------------------------
    analyses_run = list(
      run_fs     = run_fs,
      run_fs_grf = run_fs_grf,
      run_grf    = run_grf,
      grf_params = grf_params
    ),

    # -- Provenance ------------------------------------------------------------
    timestamp   = Sys.time(),
    R_version   = R.version.string,
    fs_version  = tryCatch(
      as.character(utils::packageVersion("forestsearch")),
      error = function(e) "unknown"
    ),
    timings     = timings,

    # -- Failure accounting ----------------------------------------------------
    n_failed_alt  = n_failed_alt,
    n_failed_null = n_failed_null
  )

  filename <- sprintf("actg175_survival_benefit_simulations_%dalts_%dnulls.rds",
                       nsims_alt, nsims_null)
  path <- file.path(output_dir, filename)
  saveRDS(bundle, path)

  cat(sprintf("Results saved: %s\n", path))
  cat(sprintf("  Size: %.1f KB\n", file.size(path) / 1024))
  cat(sprintf("  Alt sims: %d completed (%d failed)\n",
      nsims_alt - n_failed_alt, n_failed_alt))
  cat(sprintf("  Null sims: %d completed (%d failed)\n",
      nsims_null - n_failed_null, n_failed_null))
  cat(sprintf("  HR(Q) = %.3f, HR(Qc) = %.3f\n",
      dgm_calibrated$hazard_ratios$harm_subgroup,
      dgm_calibrated$hazard_ratios$no_harm_subgroup))
} else {
  cat("save_results = FALSE; skipping save.\n")
}
```


# Results {#sec-results}

## Metrics Glossary {#sec-glossary}

The operating-characteristics tables below report three groups of
metrics:

- **detection**: the proportion of simulations where ForestSearch
  (or GRF) identifies *any* subgroup ($\texttt{any.H} = 1$).
  Under the alternative, this is detection power; under the null,
  it is the false-positive rate (type I error).

- **classification**: among simulations where a subgroup was
  detected, how well the identified subgroup $\hat{H}$ overlaps
  with the true subgroup $H$.  Includes sensitivity (proportion
  of true $H$ members captured), specificity (proportion of
  $H^c$ members correctly excluded), and positive predictive value
  (PPV, proportion of $\hat{H}$ that truly belongs to $H$).

- **cde_estimates**: bias and root-mean-square error (RMSE) of
  the controlled direct effect (CDE) estimator for the treatment
  effect within $\hat{H}$ and $\hat{H}^c$, relative to the true
  DGM values.  The CDE is defined at the individual level via
  potential outcomes and represents the causal estimand
  (see León et al. 2024, Section 2.2).


## Operating Characteristics Under H1

The simulation engine runs on the **switched** treatment scale.
When `subgroup_notation = "benefit"` is passed to the table
functions, they automatically invert all hazard ratios to the
**original** scale (HR < 1 = benefit from ZDV+ddI) and apply
Q/Qc notation.

```{r oc-alt}
t0 <- proc.time()[3]

format_oc_results(
  results_alt,
  metrics   = c("detection", "classification", "cde_estimates"),
  digits    = 2,
  digits_hr = 2,
  title     = "ACTG175: Identification Properties (Benefit Search)",
  subtitle  = sprintf(
    "n = %d, %d sims, HR(Q) = %.2f, HR(Qc) = %.2f",
    sim_config$n_sample, nsims_alt,
    1 / dgm_calibrated$hazard_ratios$harm_subgroup,
    1 / dgm_calibrated$hazard_ratios$no_harm_subgroup
  ),
  subgroup_notation = "benefit"
)

build_estimation_table(
  results         = results_alt,
  dgm             = dgm_calibrated,
  analysis_method = "FS",
  subtitle = sprintf(
    "ACTG175 DGM | n = %d, %d sims",
    sim_config$n_sample, nsims_alt
  ),
  font_size = 16,
  subgroup_notation = "benefit"
)

build_estimation_table(
  results         = results_alt,
  dgm             = dgm_calibrated,
  analysis_method = "GRF",
  subtitle = sprintf(
    "ACTG175 DGM | n = %d, %d sims",
    sim_config$n_sample, nsims_alt
  ),
  font_size = 16,
  subgroup_notation = "benefit"
)


timings$summarize_alt <- proc.time()[3] - t0
```

### Interpretation

```{r interpret-alt, results="asis"}
interpret_estimation_table(
  results_alt, dgm_calibrated,
  analysis_method = "FS",
  scenario = "alt",
  subgroup_notation = "benefit"
)
```


## Operating Characteristics Under H0

```{r oc-null}
t0 <- proc.time()[3]

format_oc_results(
  results_null,
  metrics  = c("detection", "classification"),
  digits   = 2,
  title    = "ACTG175: Type I Error (Null)",
  subtitle = sprintf(
    "n = %d, %d sims, HR(ITT) = %.2f",
    sim_config$n_sample, nsims_null,
    1 / dgm_null$hazard_ratios$overall
  ),
  subgroup_notation = "benefit"
)

build_estimation_table(
  results         = results_null,
  dgm             = dgm_null,
  analysis_method = "FS",
  subtitle = sprintf(
    "ACTG175 DGM | n = %d, %d sims",
    sim_config$n_sample, nsims_null
  ),
  font_size = 16,
  subgroup_notation = "benefit"
)

build_estimation_table(
  results         = results_null,
  dgm             = dgm_null,
  analysis_method = "GRF",
  subtitle = sprintf(
    "ACTG175 DGM | n = %d, %d sims",
    sim_config$n_sample, nsims_null
  ),
  font_size = 16,
  subgroup_notation = "benefit"
)


timings$summarize_null <- proc.time()[3] - t0
```


## Classification Table

```{r classification}
scenario_list <- list(
  null = list(
    results    = results_null,
    label      = "ACTG",
    n_sample   = sim_config$n_sample,
    dgm        = dgm_null,
    hypothesis = "null"
  ),
  alt = list(
    results    = results_alt,
    label      = "ACTG",
    n_sample   = sim_config$n_sample,
    dgm        = dgm_calibrated,
    hypothesis = "alt"
  )
)

build_classification_table(
  scenario_results = scenario_list,
  analyses = sort(unique(c(
    unique(results_null$analysis),
    unique(results_alt$analysis)
  ))),
  digits    = 2,
  font_size = 16,
  title     = "ACTG175: Subgroup Identification and Classification",
  n_sims    = nsims_alt,
  subgroup_notation = "benefit"
)
```


# Theoretical Detection Probability {#sec-detection}

The detection probability is the theoretical power of ForestSearch's
bootstrap consistency rule to flag a subgroup with a given HR, given
the expected subgroup size and censoring proportion.  This curve
provides a reference for interpreting the empirical detection rate
observed in the simulation.

The `method = "cubature"` argument uses numerical integration
(via the `cubature` package) to compute the joint probability
that the subgroup HR exceeds `hr.threshold` (c1) and that the
proportion of bootstrap replicates exceeding `hr.consistency`
(c2) is at least `pconsistency.threshold`.  An alternative
`method = "normal"` uses a normal approximation.

```{r detection-curve, fig.height=6}
n_sg_expected <- sim_config$n_sample *
  mean(dgm_calibrated$df_super$flag_harm)
prop_cens <- mean(results_alt$p.cens)

cat("Expected subgroup size:", round(n_sg_expected), "\n")
cat("Censoring proportion:",   round(prop_cens, 3), "\n")
cat("True HR(H):",             round(dgm_calibrated$hazard_ratios$harm_subgroup, 3), "\n")

prob_detect <- compute_detection_probability(
  theta          = dgm_calibrated$hazard_ratios$harm_subgroup,
  n_sg           = round(n_sg_expected),
  prop_cens      = prop_cens,
  hr_threshold   = fs_params$hr.threshold,
  hr_consistency = fs_params$hr.consistency,
  method         = "cubature"
)

cat("\nTheoretical P(detect):", round(prob_detect, 3), "\n")
cat("Empirical FS:",
    round(mean(results_alt[
      results_alt$analysis == "FS", "any.H"
    ]), 3), "\n")

# Generate curve
detection_curve <- generate_detection_curve(
  theta_range       = det_theta_range,
  n_points          = det_n_points,
  n_sg              = round(n_sg_expected),
  prop_cens         = prop_cens,
  hr_threshold      = fs_params$hr.threshold,
  hr_consistency    = fs_params$hr.consistency,
  include_reference = TRUE,
  verbose           = FALSE
)

plot_detection_curve(
  detection_curve,
  add_reference_lines = TRUE,
  add_threshold_line  = TRUE,
  title = sprintf(
    "Detection Curve (n_sg=%d, cens=%.0f%%, c1=%.2f)",
    round(n_sg_expected), 100 * prop_cens,
    fs_params$hr.threshold
  )
)

abline(v = dgm_calibrated$hazard_ratios$harm_subgroup, lty = 2, col = "blue")
```


# Timing Summary {#sec-timing}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Stage = c(
    "DGM creation (H1)",
    "k_inter calibration",
    "DGM calibrated",
    "DGM creation (H0)",
    "Single-trial ForestSearch",
    "Simulations H1",
    "Simulations H0",
    "Summarize H1",
    "Summarize H0",
    "Total"
  ),
  Seconds = c(
    timings$dgm_alt,
    timings$calibrate,
    timings$dgm_calibrated,
    timings$dgm_null,
    timings$single_fs,
    timings$sims_alt,
    timings$sims_null,
    timings$summarize_alt,
    timings$summarize_null,
    timings$total
  ),
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60
timing_df$Hours   <- timing_df$Seconds / 3600

gt(timing_df) |>
  tab_header(
    title    = "Computational Timing Summary",
    subtitle = sprintf(
      "%d H1 + %d H0 sims | %d workers",
      nsims_alt, nsims_null, n_workers
    )
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  fmt_number(columns = Hours,   decimals = 3) |>
  cols_label(
    Stage   = "Stage",
    Seconds = "Time (sec)",
    Minutes = "Time (min)",
    Hours   = "Time (hr)"
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total")
  )
```

### Projection for Production Runs

```{r projection}
# Per-replicate cost by stage
time_per_alt  <- timings$sims_alt / nsims_alt
time_per_null <- timings$sims_null / nsims_null

proj_alt   <- time_per_alt * prod_alt
proj_null  <- time_per_null * prod_null
proj_other <- timings$total - timings$sims_alt - timings$sims_null
proj_total <- proj_alt + proj_null + proj_other

proj_df <- data.frame(
  Component = c("H1 simulations", "H0 simulations", "Other (DGM, calibration, summaries)",
                "Total"),
  Count     = c(
    sprintf("%d replicates", prod_alt),
    sprintf("%d replicates", prod_null),
    "---",
    sprintf("%d + %d", prod_alt, prod_null)
  ),
  `Per replicate` = c(
    sprintf("%.1f sec", time_per_alt),
    sprintf("%.1f sec", time_per_null),
    "---",
    "---"
  ),
  Projected = c(
    sprintf("%.1f min", proj_alt / 60),
    sprintf("%.1f min", proj_null / 60),
    sprintf("%.1f min", proj_other / 60),
    sprintf("%.1f min (%.1f hours)", proj_total / 60, proj_total / 3600)
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt(proj_df) |>
  tab_header(
    title = sprintf("Projected Time: %d alt + %d null", prod_alt, prod_null),
    subtitle = sprintf(
      "Based on current run (%d + %d on %d workers)",
      nsims_alt, nsims_null, n_workers
    )
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Component == "Total")
  )
```


# Discussion {#sec-discussion}

## Key Differences from the GBSG Simulation

The ACTG175 dataset introduces several features that test
ForestSearch under different conditions than the GBSG template:

1. **Larger sample size** ($N \approx 1{,}000$ vs $686$): more
   candidate subgroups survive the `n.min` filter, potentially
   increasing $L_{\text{eff}}$ and the procedure-level FPR.

2. **More confounders** ($12$ vs $7$): the combinatorial search
   space is larger, testing GRF and LASSO screening effectiveness.

3. **Different event rate**: ACTG175 has a lower event rate than
   GBSG, which reduces $d_{\text{eff}}$ per subgroup and may
   affect detection power.

4. **Different subgroup structure**: the age $\times$ CD8 subgroup
   is defined by continuous biomarkers discretised at clinically
   motivated cutpoints, rather than the GBSG's receptor-status
   $\times$ menopausal-status interaction.

5. **Time scale**: ACTG175 times have been converted from days to months for
   AFT model parametrisation but not the HR-scale operating
   characteristics.

## Reloading Saved Results

When `save_results = TRUE`, the simulation bundle is saved as an
`.rds` file in the `_data/` directory.  To reload for downstream
analysis:

```r
bundle <- readRDS("_data/actg175_survival_benefit_simulations_50alts_50nulls.rds")

# Core results (data.tables)
results_alt  <- bundle$results_alt
results_null <- bundle$results_null

# DGM objects (class "aft_dgm_flex")
dgm_cal  <- bundle$dgm_calibrated
dgm_null <- bundle$dgm_null

# True effects
bundle$true_effects$hr_Q     # HR(Q) on switched scale
bundle$true_effects$ahr_Q    # AHR(Q)
bundle$true_effects$k_inter  # calibrated interaction multiplier

# Reproduce tables
format_oc_results(results_alt, subgroup_notation = "benefit", ...)
build_estimation_table(results_alt, dgm = dgm_cal,
                       subgroup_notation = "benefit", ...)
```


## Limitations

- The ACTG175 benefit subgroup definition (age $> `r actg_age_cut`$
  AND preanti $\leq `r actg_preanti_cut`$) is based on the
  exploratory analysis in León et al. (2024, §4.2). The true
  treatment effect heterogeneity pattern in ACTG175 is unknown.

- The benefit-search approach (treatment switching) adds a layer
  of indirection: all DGM, simulation, and ForestSearch operations
  use the switched treatment indicator, and results must be
  inverted for interpretation on the original treatment scale.

- `generate_aft_dgm_flex()` requires more analyst judgment than
  `setup_gbsg_dgm()`. The variable discretisation, subgroup
  definition, and censoring calibration must all be specified
  manually.

- The simulation count (`r nsims_alt` sims) is for demonstration.
  Production studies should use $\geq 500$ simulations for stable
  operating characteristics (Monte Carlo SE for a 50% rate $\approx
  2.2\%$ at 500 sims).

- The adjusted Cox formula (`cox_formula_adj`) includes only 3 of
  12 confounders.  This is a deliberate choice for parsimony, but
  users should select confounders appropriate for their study
  design.

- `calibrate_k_inter()` does not yet support non-GBSG datasets;
  the manual grid search in @sec-calibrate is the current
  workaround.


# Session Info {#sec-session}

```{r session-info}
sessionInfo()
```


# References {#sec-references}

- Leon, L.F., Jemielita, T., Guo, Z., Marceau West, R.,
  Anderson, K.M. (2024). Exploratory subgroup identification
  in the heterogeneous Cox model. *Statistics in Medicine*,
  43, 4253--4275.

- Hammer, S.M., Katzenstein, D.A., Hughes, M.D., et al. (1996).
  A trial comparing nucleoside monotherapy with combination
  therapy in HIV-infected adults with CD4 cell counts from 200
  to 500 per cubic millimeter. *NEJM*, 335, 1081--1090.
