---
title: "ACTG175 Binary Harm Search: Multi-Engine Simulation Study"
subtitle: |
  Seven selection arms (fs / dina / grf x eff / effMaxSG, plus grf-tree)
  through the `forestsearch()` post-selection-inference backbone
author: "forestsearch v0.2.0"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
---

```{r config-id}
#| include: false

# ─────────────────────────────────────────────────────────────────────────
# RUN CONTROL — edit these to drive the document.
# ─────────────────────────────────────────────────────────────────────────
# sim_mode:  "demo"   small nsims, runs inline (self-test of the methods path)
#            "full"   production nsims, writes a bundle
#            "import" reload a previously-saved bundle (no simulation)
#            "auto"   import if BUNDLE_PATH exists, else demo
sim_mode  <- "demo"

sims_demo <- 10L
sims_full <- 1000L
sim_n     <- 700L

# ─────────────────────────────────────────────────────────────────────────
# METHODS — the seven comparison arms.
# ─────────────────────────────────────────────────────────────────────────
# Each "<engine>-<focus>" label names the *selector*; every arm is executed
# through forestsearch(), which supplies the shared bias-corrected
# post-selection inference, so no "FS" appears in any label.
#
#   fs-*    consistency search; GRF + DINA used as *screening* into the pool
#           (use_grf = TRUE, use_dina = TRUE, use_lasso = FALSE by default)
#   dina-*  DINA is the *selector*
#   grf-*   GRF frontier (grf-eff / grf-effMaxSG) or policy tree (grf-tree)
#
# This is the character-vector form of run_simulation_analysis(methods = ).
# For per-arm control (e.g. flipping the fs screening knobs) switch to the
# named-list form documented in ?run_simulation_analysis.

# METHODS <- c(
#   "fs-eff",   "fs-effMaxSG",
#   "dina-eff", "dina-effMaxSG",
#   "grf-eff",  "grf-effMaxSG",
#   "grf-tree"
# )


METHODS <- c(
  "fs-eff",   
  "grf-tree"
)


# Arm used for the single-trial / subgroup / effect-distribution displays.
# Must be a member of METHODS; falls back to the first arm if the named one
# isn't being run, so editing METHODS never breaks the display chunks.
FLAGSHIP_ARM <- "fs-effMaxSG"
if (!FLAGSHIP_ARM %in% METHODS) {
  warning(sprintf("FLAGSHIP_ARM '%s' not in METHODS; using '%s' instead.",
                  FLAGSHIP_ARM, METHODS[1L]))
  FLAGSHIP_ARM <- METHODS[1L]
}

# ─────────────────────────────────────────────────────────────────────────
# BUNDLE PATH — machine-local output for "full"/"import" modes.
# Edit to a directory that exists on the rendering machine.
# ─────────────────────────────────────────────────────────────────────────
scenario    <- "actg175_binary_m1_harm"
BUNDLE_DIR  <- "_bundles"
BUNDLE_PATH <- file.path(
  BUNDLE_DIR,
  sprintf("%s_methods_n%d_sims%d.rds", scenario, sim_n,
          if (sim_mode == "full") sims_full else sims_demo)
)
CONFIG_ID   <- tools::file_path_sans_ext(basename(BUNDLE_PATH))

if (sim_mode %in% c("full", "import") && !dir.exists(BUNDLE_DIR)) {
  dir.create(BUNDLE_DIR, recursive = TRUE, showWarnings = FALSE)
}
```

# Overview

This document is the **master per-config template** for ACTG175 harm-subgroup
operating-characteristic studies, built on the generalized `methods` interface
of `run_simulation_analysis()`. A single Monte Carlo loop runs all seven arms
in `METHODS` on the *same* simulated data each replication, so every contrast
is within-replication.

The design isolates two questions the León et al. (2024) precursor could not
address jointly:

1. **Selection focus.** Within the consistency ensemble, how do `eff`
   (stability-first), `effMaxSG` (broad, band-aware), and the engine-direct
   selectors trade off detection, the sensitivity/specificity/PPV/NPV quartet,
   bias, and realized subgroup size?
2. **Engine choice under a common inference backbone.** Because `dina-*` and
   `grf-*` are run *through* `forestsearch()`, they inherit the bias-corrected
   bootstrap + jackknife post-selection inference their native implementations
   lack. The arms therefore differ only in *who selects the subgroup*, with the
   inference layer held fixed — a within-engine comparison rather than a
   cross-method one.

All operating characteristics use the León et al. (2024) Table 1 conventions:
the true marginal effect is denoted $\theta^\dagger(\cdot)$ on the
natural-parameter scale (here the **odds ratio** for the binary adverse
outcome), detection is $\mathrm{any}(\cdot)$, and the classification quartet
$\{\text{sens}(\hat H), \text{ppv}(\hat H), \text{sens}(\hat H^c),
\text{ppv}(\hat H^c)\}$ is averaged **unconditionally** (a replication with no
detected subgroup contributes sens $= 0$, ppv $= 0$ for $\hat H$ and
spec $= 1$, npv $= 1$ for $\hat H^c$).

```{r setup}
#| code-fold: true
#| code-summary: "Libraries, base FS parameters, notation, helpers (click to expand)"

library(forestsearch)
library(speff2trial)
library(data.table)
library(gt)
library(ggplot2)
library(doFuture)
library(foreach)

set.seed(8316951L)

`%||%` <- function(a, b) if (is.null(a)) b else a

# Tidy a cut-string for DISPLAY only (rounds full-precision thresholds to
# `digits` significant figures and collapses ragged whitespace; integer cuts
# untouched).  Does NOT alter the stored sg.def used for evaluation.
tidy_sg_def <- function(s, digits = 4L) {
  if (length(s) != 1L || is.na(s) || !nzchar(s)) return(s)
  s <- gsub("[[:space:]]+", " ", s)
  m <- gregexpr("-?[0-9]+\\.[0-9]+", s)[[1L]]
  if (m[1L] == -1L) return(s)
  L <- attr(m, "match.length")
  for (i in rev(seq_along(m))) {
    num <- as.numeric(substr(s, m[i], m[i] + L[i] - 1L))
    rep <- sub("\\.$", "", formatC(signif(num, digits), format = "fg", flag = "#"))
    s <- paste0(substr(s, 1L, m[i] - 1L), rep, substr(s, m[i] + L[i], nchar(s)))
  }
  s
}

# -- Shared _utils (cross-method comparison plots) ---------------------------
# Locate quarto/_utils by walking up from the render directory, so this file
# works at any nesting depth (e.g. simulations/actg175/binary_methods/).
.find_dir_upwards <- function(target, start = ".", max_up = 7L) {
  d <- normalizePath(start, winslash = "/", mustWork = FALSE)
  for (i in seq_len(max_up)) {
    cand <- file.path(d, target)
    if (dir.exists(cand)) return(cand)
    parent <- dirname(d)
    if (identical(parent, d)) break          # reached filesystem root
    d <- parent
  }
  stop(sprintf("Could not locate '%s' within %d levels above '%s'.",
               target, max_up, normalizePath(start, mustWork = FALSE)),
       call. = FALSE)
}
.util_dir <- .find_dir_upwards("_utils")
source(file.path(.util_dir, "plot_oc_metrics.R"))
source(file.path(.util_dir, "compute_oc_pareto_flag.R"))
source(file.path(.util_dir, "compute_oc_score.R"))

# ===========================================================================
# BASE FORESTSEARCH PARAMETERS  (specified here, not sourced from _fsparams)
# ---------------------------------------------------------------------------
# Rationale: the `methods` interface collapses the former per-sg_focus qmd
# fleet into this single document, so the external bundle's reason to exist
# (keeping sibling qmds in sync) no longer applies.  These are the SHARED
# base; per-arm engine / focus / screening overrides are injected by the
# `methods` expander and take precedence (default_sim_params() -> fs_params
# -> per-arm overrides).  Revise values here.
#
# stop_threshold = NULL is the cross-focus-safe choice: the band-aware focuses
# (effMaxSG) require NULL, and for `eff` it simply disables early stopping.
# ===========================================================================
fs_base <- list(
  # -- Identifiers (match simulate_from_glm_dgm() output) --------------------
  outcome.name           = "y_sim",
  event.name             = "y_sim",
  treat.name             = "treat_sim",
  id.name                = "id",

  # -- Outcome typing (binary adverse outcome -> odds ratio) -----------------
  outcome_type           = "binary",
  effect_measure         = "OR",

  # -- Effect-size / stability thresholds ------------------------------------
  hr.threshold           = 1.25,     # OR scale; modest harm signal
  hr.consistency         = 1.0,
  pconsistency.threshold = 0.90,
  stop_threshold         = NULL,     # cross-focus-safe (see header)

  # -- Candidate construction ------------------------------------------------
  conf.cont_jcuts        = list(cd40 = 10, wtkg = 10),  # quantile cuts
  conf_force             = c("ar_naive == 1", "prior_6mo == 1"),
  maxk                   = 2L,       # up to two-factor conjunctions
  fs.splits              = 1000L,

  # -- Sample-size constraints -----------------------------------------------
  # n.min = NULL -> sample-size-adaptive rule (reaches forestsearch() as NULL
  # via the methods path's NULL-preserving merge).  Set an integer to override.
  n.min                  = NULL,
  d0.min                 = 10L,
  d1.min                 = 10L,

  # -- GRF screening defaults (fs-* arms; vi.grf.min is the SCREENING knob) ---
  # dmin.grf is deliberately NOT set here: it is a SELECTION floor and is set
  # per-arm in make_methods() (screen / tree / frontier), see setup below.
  use_grf                = TRUE,
  use_dina               = TRUE,
  use_lasso              = FALSE,
  vi.grf.min             = -0.2,

  # -- RCT / seed ------------------------------------------------------------
  is.RCT                 = TRUE,
  seedit                 = 8316951L,

  # -- Inner-loop parallel plan (forced sequential; avoids nested parallelism)
  parallel_args          = list(plan = "sequential", workers = 1,
                                show_message = FALSE)
)

# -- Reporting / calibration -------------------------------------------------
sg_notation    <- "harm"     # H / Ĥ (harm) notation in OC tables
cal_target_or  <- 2.0        # target OR(Q) on the adverse scale

# -- Diagnostic capture ------------------------------------------------------
diag_keep         <- c("fs_full", "candidate_table")
diag_keep_first_n <- 5L

# -- Plot thresholds ---------------------------------------------------------
sgdist_min_pct_alt  <- 0
sgdist_min_pct_null <- 0
sgdist_top_k        <- 10L
ordist_ylim         <- c(0, 4)

# -- nsims from sim_mode -----------------------------------------------------
if (sim_mode == "auto")
  sim_mode <- if (file.exists(BUNDLE_PATH)) "import" else "demo"
if (sim_mode == "import" && !file.exists(BUNDLE_PATH))
  stop(sprintf("sim_mode = 'import' but no bundle at '%s'.", BUNDLE_PATH),
       call. = FALSE)

nsims_alt <- switch(sim_mode, import = 0L, demo = sims_demo, full = sims_full,
                    stop(sprintf("Unknown sim_mode: '%s'", sim_mode)))
nsims_null <- switch(sim_mode, import = 0L, demo = sims_demo, full = sims_full)

cat(sprintf("sim_mode = %s | nsims_alt = %d | nsims_null = %d | sim_n = %d\n",
            sim_mode, nsims_alt, nsims_null, sim_n))
cat("Arms:", paste(METHODS, collapse = ", "), "\n")

# ===========================================================================
# GRF SELECTION KNOBS  (dmin.grf is a SELECTION floor, not screening)
# ---------------------------------------------------------------------------
# Following subgroup_method_signature_recovery_2factor.qmd, dmin.grf is split
# by GRF *use* into three independent knobs.  dmin.grf is consumed in
# select_best_subgroup() as the harm-eligibility floor (diff >= dmin.grf), so
# it bites only where GRF is the *selector* (grf-* arms); for fs-* arms GRF is
# a screening input governed by vi.grf.min and dmin.grf is inert.
#
# SCALE NOTE: the signature-recovery vignette is a SURVIVAL DGM, where its
# 0.30 floor is on the RMST / time scale.  This template is BINARY (OR), so
# the GRF DR scores are on a risk-difference scale and 0.30 does NOT transfer.
# The selection floors below are placeholders on the binary DR-score scale and
# MUST be calibrated for this outcome; 0.0 reproduces the permissive
# "any positive harm" baseline.  Raise tree / frontier to control GRF-arm FDR.
# ===========================================================================
grf_depth          <- 2L
grf_screen_dmin    <- 0.0   # GRF screening into consistency engine (fs-* arms)
grf_tree_dmin      <- 0.0   # GRF policy-tree SELECTION   (grf-tree)   [TEST: permissive baseline]
grf_frontier_dmin  <- 0.0   # GRF frontier SELECTION (grf-eff/grf-effMaxSG) [TEST: permissive baseline]

# Build the named-list `methods` spec from the canonical labels.  The GRF
# selection arms disable separate screening (use_grf = use_lasso = FALSE:
# "selection IS GRF") and carry grf_depth + their own dmin.grf.  fs-* arms
# keep GRF + DINA screening and use grf_screen_dmin.  n.min is inherited from
# fs_base (NULL -> adaptive); not set per arm.
make_methods <- function(labels) {
  spec <- list()
  for (lbl in labels) {
    parts  <- strsplit(lbl, "-", fixed = TRUE)[[1L]]
    engine <- parts[1L]
    focus  <- paste(parts[-1L], collapse = "-")
    if (engine == "fs") {
      spec[[lbl]] <- list(
        subgroup_method = "consistency", sg_focus = focus,
        selection_rule  = "neighborhood",
        use_grf = TRUE, use_dina = TRUE, use_lasso = FALSE,
        dmin.grf = grf_screen_dmin
      )
    } else if (engine == "dina") {
      spec[[lbl]] <- list(
        subgroup_method = "dina", sg_focus = focus,
        selection_rule  = "neighborhood"
      )
    } else if (engine == "grf") {
      if (identical(focus, "tree")) {
        spec[[lbl]] <- list(
          subgroup_method = "grf", grf_selection = "tree",
          grf_depth = grf_depth, dmin.grf = grf_tree_dmin,
          use_grf = FALSE, use_lasso = FALSE
        )
      } else {
        spec[[lbl]] <- list(
          subgroup_method = "grf", grf_selection = "frontier",
          sg_focus = focus, grf_depth = grf_depth, dmin.grf = grf_frontier_dmin,
          use_grf = FALSE, use_lasso = FALSE
        )
      }
    } else {
      stop(sprintf("Unknown engine in label '%s'.", lbl), call. = FALSE)
    }
  }
  spec
}
METHODS_SPEC <- make_methods(METHODS)

# ===========================================================================
# Label -> method-family bridge for the _utils comparison panel.
# The shared _utils key off the OLD labels (^FS-grf, ^FS-eff, ^DINA); these
# new "<engine>-<focus>" labels need their own mapping.  Kept HERE rather than
# editing the shared _utils file.
# ===========================================================================
method_family <- function(label) {
  fam <- rep("other", length(label))
  fam[grepl("^fs-",   label)] <- "FS-consistency"
  fam[grepl("^grf-",  label)] <- "FS-grf"
  fam[grepl("^dina-", label)] <- "DINA"
  factor(fam, levels = c("FS-consistency", "FS-grf", "DINA", "other"))
}

# Build a plot_oc_metrics()-ready display table directly from a
# run_simulation_analysis() result, applying the León unconditional-averaging
# convention.  One row per arm (`analysis` column).
oc_agg_from_results <- function(res) {
  dt <- data.table::as.data.table(res)
  dt[, `:=`(
    sens_u = ifelse(is.na(sens) & any.H == 0, 0, sens),
    ppv_u  = ifelse(is.na(ppv)  & any.H == 0, 0, ppv),
    spec_u = ifelse(is.na(spec) & any.H == 0, 1, spec),
    npv_u  = ifelse(is.na(npv)  & any.H == 0, 1, npv)
  )]
  agg <- dt[, .(
    Detection = mean(any.H,  na.rm = TRUE),
    Sen       = mean(sens_u, na.rm = TRUE),
    Spec      = mean(spec_u, na.rm = TRUE),
    PPV       = mean(ppv_u,  na.rm = TRUE),
    NPV       = mean(npv_u,  na.rm = TRUE)
  ), by = .(group = analysis)]
  agg[, Analysis := method_family(group)]
  agg[]
}

timings <- list()
```

# Data Preparation {#sec-data}

ACTG175 (`speff2trial::ACTG175`): the experimental arm is ZDV+ddI (`arms == 1`)
versus the ddI control (`arms == 3`); no treatment switching. The adverse binary
outcome is failure to improve CD4 (`cd420 <= cd40`). The embedded true harm
subgroup is $Q = \{\text{cd40} > 390 \,\cap\, \text{wtkg} > 80\}$.

```{r data-prep}
#| code-fold: true
#| code-summary: "ACTG175 data prep (click to expand)"

actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$id    <- seq_len(nrow(actg_df))
actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)

# Binary adverse outcome: failure to improve CD4 from baseline
actg_df$y_binary <- as.integer(actg_df$cd420 <= actg_df$cd40)
actg_df <- actg_df[!is.na(actg_df$cd420), ]

# DGM factor variables.  True subgroup Q = {cd40 > 390 & wtkg > 80}.
actg_df$z1 <- as.factor(ifelse(actg_df$cd40 > 390, 1L, 0L))
actg_df$z2 <- as.factor(ifelse(actg_df$wtkg > 80,  1L, 0L))

actg_df$z7  <- as.factor(actg_df$hemo)
actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs)
actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender)
actg_df$z12 <- as.factor(actg_df$symptom)

actg_df$ar_naive  <- as.integer(ifelse(actg_df$str2 == 0, 1L, 0L))
actg_df$prior_6mo <- as.integer(ifelse(actg_df$preanti <= 6 * 30.4375, 1L, 0L))

actg_df$flag_harm_ref <- as.integer(actg_df$z1 == 1 & actg_df$z2 == 1)

# Analysis covariates (what the analyst sees)
cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
bin_vars  <- c("hemo", "homo", "drugs", "race", "gender", "symptom",
               "ar_naive", "prior_6mo")
# Binary confounders kept NUMERIC 0/1: DINA requires numeric covariates and
# the GRF policy-tree emits numeric cut-strings (e.g. "symptom <= 0") that are
# invalid for factors.  z7-z12 (created above) stay factors for the DGM.
for (v in bin_vars) actg_df[[v]] <- as.integer(actg_df[[v]])
confounders <- c(cont_vars, bin_vars)

# DGM variable sets
dgm_factor_vars   <- c("z1", "z2", paste0("z", 7:12))
dgm_cont_vars     <- cont_vars
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)

Q_n <- sum(actg_df$flag_harm_ref); Q_N <- nrow(actg_df)
Q_pct <- 100 * Q_n / Q_N
cat(sprintf("N = %d, Q prevalence = %d / %d (%.1f%%)\n", Q_N, Q_n, Q_N, Q_pct))
cat(sprintf("Adverse rate: overall=%.1f%%, ZDV+ddI=%.1f%%, ddI=%.1f%%\n",
    100 * mean(actg_df$y_binary),
    100 * mean(actg_df$y_binary[actg_df$treat == 1]),
    100 * mean(actg_df$y_binary[actg_df$treat == 0])))
```

# DGM Calibration {#sec-calibration}

The alternative DGM is calibrated to a target $\theta^\dagger(Q) =$ OR(Q) $=$
`r cal_target_or` on the adverse scale; the null DGM embeds a homogeneous effect.

```{r calibrate}
#| eval: !expr sim_mode != "import"

dgm_calibrated <- calibrate_glm_interaction(
  data            = actg_df,
  factor_vars     = dgm_factor_vars,
  continuous_vars = dgm_cont_vars,
  outcome_var     = "y_binary",
  treatment_var   = "treat",
  target_effect   = cal_target_or,
  outcome_type    = "binary",
  effect_measure  = "OR",
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  k_inter_range   = c(0, 3),
  grid_step       = 0.05,
  n_super         = 5000L,
  seed            = 8316951L,
  verbose         = TRUE
)

dgm_null <- generate_glm_dgm(
  data            = actg_df,
  factor_vars     = dgm_factor_vars,
  continuous_vars = dgm_cont_vars,
  outcome_var     = "y_binary",
  treatment_var   = "treat",
  outcome_type    = "binary",
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  model           = "null",
  n_super         = 5000L,
  seed            = 8316951L,
  verbose         = TRUE
)

cat(sprintf("\n  theta-dagger(Q):   %.3f (target OR = %.1f)\n",
    dgm_calibrated$hazard_ratios$harm_subgroup, cal_target_or))
cat(sprintf("  theta-dagger(Qc):  %.3f\n",
    dgm_calibrated$hazard_ratios$no_harm_subgroup))
cat(sprintf("  theta-dagger(ITT): %.3f\n",
    dgm_calibrated$hazard_ratios$overall))
```

# Single-Trial Sanity Check {#sec-single-trial}

One simulated trial under H1, analyzed with the flagship arm
(`fs-effMaxSG`: GRF + DINA screening into the consistency search, broad
band-aware selection) to confirm the pipeline before the Monte Carlo loop.

```{r single-trial}
#| eval: !expr sim_mode != "import"
#| code-fold: true
#| code-summary: "Single simulated trial + flagship FS analysis (click to expand)"

df_sim <- simulate_from_glm_dgm(dgm_calibrated, n = sim_n, seed = 42)

# Flagship arm overrides, applied to the shared base.
fs_flag <- utils::modifyList(
  fs_base,
  list(subgroup_method = "consistency", sg_focus = "effMaxSG",
       selection_rule = "neighborhood")
)

# Filter to valid forestsearch() formals (the MC path does this internally
# via filter_valid_args(); the manual single-trial call must do it too).
fs_formals <- names(formals(forestsearch))
call_args <- c(
  list(df.analysis = df_sim, confounders.name = confounders,
       details = FALSE, plot.sg = FALSE, quiet = TRUE),
  fs_flag[intersect(names(fs_flag), fs_formals)]
)
call_args$outcome.name <- "y_sim"
call_args$event.name   <- "y_sim"

fs_single <- tryCatch(
  do.call(forestsearch, call_args),
  error = function(e) { message("single-trial FS failed: ", e$message); NULL }
)

if (!is.null(fs_single) && !is.null(fs_single$sg.harm)) {
  cat("Flagship single-trial selected subgroup:\n  ",
      paste(fs_single$sg.harm, collapse = " & "), "\n")
} else {
  cat("Flagship single-trial: no harm subgroup identified on this draw.\n")
}
```

# Monte Carlo Study {#sec-mc}

Each replication simulates a trial and runs all seven `METHODS` arms on it
through `run_simulation_analysis(methods = METHODS_SPEC)`. Arms are tagged in the
`analysis` column by their label; `fs_base` supplies the shared parameters and
the per-arm overrides take precedence.

The loops run in parallel across simulations via `multisession` workers
(`n_workers`), while each `forestsearch()` call stays single-threaded
(`fs_base$parallel_args`) to avoid nested-parallelism failures.

```{r parallel-setup}
#| eval: !expr sim_mode != "import"
#| code-fold: true
#| code-summary: "Worker pool (click to expand)"

# Use 90% of physical cores so the cap adapts per machine (Linux box vs Mac).
# Outer = multisession across simulations; inner forestsearch() forced
# sequential via parallel_args, with the in-worker plan("sequential") as the
# nested guard.
n_workers <- max(1L, ceiling(0.9 * parallel::detectCores(logical = FALSE)))
future::plan(future::multisession, workers = n_workers)
cat(sprintf("Parallel backend: multisession, %d workers\n", n_workers))
```

```{r run-alt}
#| eval: !expr sim_mode != "import"
#| code-fold: true
#| code-summary: "H1 simulation loop (click to expand)"
#| output: false

t0 <- proc.time()[3]
results_alt_raw <- foreach(
  sim = seq_len(nsims_alt),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_calibrated,
    n_sample         = sim_n,
    confounders_base = confounders,
    methods          = METHODS_SPEC,
    fs_params        = fs_base,
    keep             = diag_keep,
    keep_first_n     = diag_keep_first_n,
    verbose          = FALSE
  )
}
t_alt <- proc.time()[3] - t0
results_alt <- forestsearch:::collect_results(results_alt_raw, "H1 alt")
n_failed_alt <- attr(results_alt, "n_failed")
timings$sims_alt <- t_alt
cat(sprintf("H1 done in %.1f min (%d rows, %d failed)\n",
            t_alt / 60, nrow(results_alt), n_failed_alt))
```

```{r run-null}
#| eval: !expr sim_mode != "import"
#| code-fold: true
#| code-summary: "H0 simulation loop (click to expand)"
#| output: false

t0 <- proc.time()[3]
results_null_raw <- foreach(
  sim = seq_len(nsims_null),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_null,
    n_sample         = sim_n,
    confounders_base = confounders,
    methods          = METHODS_SPEC,
    fs_params        = fs_base,
    verbose          = FALSE
  )
}
t_null <- proc.time()[3] - t0
results_null <- forestsearch:::collect_results(results_null_raw, "H0 null")
n_failed_null <- attr(results_null, "n_failed")
timings$sims_null <- t_null
cat(sprintf("H0 done in %.1f min (%d rows, %d failed)\n",
            t_null / 60, nrow(results_null), n_failed_null))

# Tear down the worker pool at the parallel -> sequential transition.
future::plan("sequential"); invisible(gc())
```

```{r save-bundle}
#| eval: !expr sim_mode == "full"
#| code-fold: true
#| code-summary: "Bundle assembly + saveRDS (click to expand)"

bundle <- list(
  meta = list(
    config_id       = CONFIG_ID, timestamp = Sys.time(),
    methods         = METHODS,   fs_base   = fs_base,
    n_alt = nsims_alt, n_null = nsims_null, sim_n = sim_n,
    confounders = confounders,
    t_alt_sec = unname(t_alt), t_null_sec = unname(t_null),
    n_failed_alt = n_failed_alt, n_failed_null = n_failed_null,
    package_version = as.character(utils::packageVersion("forestsearch")),
    sessionInfo = utils::sessionInfo()
  ),
  dgm     = list(calibrated = dgm_calibrated, null = dgm_null),
  results = list(alt = results_alt, null = results_null)
)
saveRDS(bundle, BUNDLE_PATH)
cat(sprintf("Bundle saved: %s\n", BUNDLE_PATH))
```

```{r mode-handoff}
#| code-fold: true
#| code-summary: "Bundle import / handoff (click to expand)"

if (sim_mode == "import") {
  bundle         <- readRDS(BUNDLE_PATH)
  dgm_calibrated <- bundle$dgm$calibrated
  dgm_null       <- bundle$dgm$null
  results_alt    <- bundle$results$alt
  results_null   <- bundle$results$null
  t_alt          <- bundle$meta$t_alt_sec
  t_null         <- bundle$meta$t_null_sec
  nsims_alt      <- bundle$meta$n_alt
  nsims_null     <- bundle$meta$n_null
  timings$sims_alt  <- t_alt
  timings$sims_null <- t_null
  cat(sprintf("Imported bundle: %s (%d H1, %d H0)\n",
              CONFIG_ID, nsims_alt, nsims_null))
  if (!is.null(bundle$meta$config_id) && bundle$meta$config_id != CONFIG_ID)
    warning(sprintf("Bundle config_id '%s' != article '%s'.",
                    bundle$meta$config_id, CONFIG_ID), call. = FALSE)
} else {
  cat(sprintf("Inline simulation complete (%s mode).\n", sim_mode))
}
```

# Results: H1 (Alternative) {#sec-results-alt}

## Detection, Classification, Subgroup Size — All Arms

```{r oc-alt}
format_oc_results(
  results_alt,
  metrics           = c("detection", "classification", "subgroup_size"),
  digits            = 2, digits_hr = 2,
  dgm               = dgm_calibrated,
  subgroup_notation = sg_notation,
  title             = "ACTG175 Binary (H1): All Arms"
)
```

## True Subgroup Sizes Across Replicates

```{r true-size-summary}
#| code-fold: true

truth_by_sim <- unique(results_alt[, c("sim", "sizeH_true", "sizeHc_true",
                                        "propH_true", "propHc_true")])
cat(sprintf("True subgroup sizes across %d H1 replicates (N = %d each):\n",
            nrow(truth_by_sim), sim_n))
cat(sprintf("  |Q|:  mean = %.1f (SD %.2f), range = [%d, %d], prop = %.1f%%\n",
    mean(truth_by_sim$sizeH_true), sd(truth_by_sim$sizeH_true),
    min(truth_by_sim$sizeH_true), max(truth_by_sim$sizeH_true),
    100 * mean(truth_by_sim$propH_true)))
```

## Per-Arm Estimation Tables (bias on the OR scale)

```{r oc-alt-estimation, results='asis'}
# build_estimation_table() returns a gt object (or NULL); inside a loop it
# must be rendered via gt::as_raw_html() + cat() (auto-print does not fire).
for (m in METHODS) {
  tbl <- build_estimation_table(
    results           = results_alt,
    dgm               = dgm_calibrated,
    analysis_method   = m,
    subgroup_notation = sg_notation,
    title             = sprintf("%s: Effect Estimates Under H1", m)
  )
  if (!is.null(tbl)) cat(gt::as_raw_html(tbl), "\n\n")
  else cat(sprintf("*%s: no estimable subgroup effects under H1.*\n\n", m))
}
```

## Per-Arm Classification Tables

```{r oc-alt-classification, results='asis'}
scenario_alt <- list(
  H1 = list(results = results_alt, dgm = dgm_calibrated,
            hypothesis = "alt", n_sample = sim_n, label = "ACTG175")
)
for (m in METHODS) {
  tbl <- build_classification_table(
    scenario_alt,
    analyses          = m,
    subgroup_notation = sg_notation,
    title             = sprintf("%s: Classification Under H1", m)
  )
  if (!is.null(tbl)) cat(gt::as_raw_html(tbl), "\n\n")
  else cat(sprintf("*%s: no estimable classification under H1.*\n\n", m))
}
```

## Cross-Arm Comparison Panel

The bubble chart places each arm in (Spec, NPV) space with detection encoded as
point size, colored by arm and shaped by engine family. Spec/NPV characterize
the *complement* $\hat H^c$ — the region declared to benefit — so the
upper-right corner is the protect-the-beneficiaries optimum.

```{r oc-alt-bubble, fig.width=8, fig.height=5.5}
oc_disp_alt <- oc_agg_from_results(results_alt)
oc_disp_alt <- compute_oc_pareto_flag(oc_disp_alt, metrics = c("Spec", "NPV"))
print(
  plot_oc_metrics(
    oc_disp_alt, x = "Spec", y = "NPV", z = "Detection",
    color_by = "group", shape_by = "Analysis",
    pareto_col = "pareto_frontier",
    title = "ACTG175 Binary (H1): arms in (Spec, NPV) space"
  )
)
```

## Subgroup & Effect Distributions (flagship arm)

```{r oc-alt-sgdist, fig.width=9, fig.height=6}
flag_rows_alt <- results_alt[results_alt$analysis == FLAGSHIP_ARM, ]
if (nrow(flag_rows_alt) == 0L || sum(flag_rows_alt$any.H, na.rm = TRUE) == 0L) {
  cat(sprintf("*No detected subgroups for flagship arm `%s` under H1; ",
              FLAGSHIP_ARM),
      "subgroup/effect distributions skipped.*\n")
} else {
  flag_rows_alt[["sg.def"]] <- vapply(flag_rows_alt[["sg.def"]],
                                       tidy_sg_def, character(1), USE.NAMES = FALSE)
  print(plot_sg_distribution(
    flag_rows_alt,
    any_col = "any.H", label_col = "sg.def",
    min_pct = sgdist_min_pct_alt, top_k = sgdist_top_k,
    title = sprintf("%s: Identified Subgroup Distribution (H1)", FLAGSHIP_ARM)
  ))
  print(plot_effect_distribution(
    results           = results_alt,
    analysis_method   = FLAGSHIP_ARM,
    effect_measure    = "OR",
    subgroup_notation = sg_notation,
    reference_effect  = dgm_calibrated$hazard_ratios$harm_subgroup,
    ylim              = ordist_ylim,
    title             = sprintf("%s: OR Estimates Across Simulations (H1)",
                                FLAGSHIP_ARM)
  ))
}
```

# Results: H0 (Null) — Type I Error {#sec-results-null}

```{r oc-null}
format_oc_results(
  results_null,
  metrics           = c("detection", "classification"),
  digits            = 2,
  dgm               = dgm_null,
  subgroup_notation = sg_notation,
  title             = "ACTG175 Binary (H0): Type I Error, All Arms"
)
```

```{r oc-null-bubble, fig.width=8, fig.height=5.5}
oc_disp_null <- oc_agg_from_results(results_null)
print(
  plot_oc_metrics(
    oc_disp_null, x = "Spec", y = "Detection", z = "Detection",
    color_by = "group", shape_by = "Analysis",
    title = "ACTG175 Binary (H0): false-positive (Detection) by arm"
  )
)
```

# Timing Summary {#sec-timing}

```{r timing}
#| echo: false
timing_dt <- data.frame(
  Stage   = c("Simulations H1", "Simulations H0"),
  Seconds = c(timings$sims_alt %||% NA_real_, timings$sims_null %||% NA_real_)
)
timing_dt$Minutes <- timing_dt$Seconds / 60
gt::gt(timing_dt) |>
  gt::tab_header(title = "Computational Timing") |>
  gt::fmt_number(columns = c(Seconds, Minutes), decimals = 2) |>
  gt::sub_missing(missing_text = "—")
```

# Session Information

```{r session-info}
sessionInfo()
```
