---
title: "ForestSearch de-biased gate — minimal test case (ACTG175 binary)"
format: html
---

```{r setup}
#| include: true
# ---------------------------------------------------------------------------
# Minimal, self-contained example of the INTEGRATED Tier-2 de-biased gate:
# forestsearch(..., debias_gate = TRUE) flags whether the selected harm
# subgroup's fast de-biased estimate is still consistent with harm, and we
# validate that flag against the full (Tier-1) bootstrap on one simulated trial.
#
# All knobs live in THIS chunk.  Results run live (no caching).  Requires only
# the installed forestsearch package (with the debias_gate feature) plus
# speff2trial (for ACTG175) and future.
# ---------------------------------------------------------------------------
suppressPackageStartupMessages({
  library(forestsearch)
  library(future)
})

## -- knobs -------------------------------------------------------------------
sim_n        <- 700L          # simulated trial size
cal_target_or<- 2.0           # true OR in the harm subgroup Q = {cd40>390 & wtkg>80}
alt_seed     <- 42L           # simulated alt trial seed
null_seed    <- 43L      # simulated null trial seed
dgm_seed     <- 8316952L      # DGM calibration seed

fs_effect_threshold       <- 1.25   # screening OR
fs_consistency_threshold  <- 1.0    # per-split OR
fs_pconsistency_threshold <- 0.90   # consistency rate
fs_sg_focus               <- "effMaxSG"
fs_selection_rule         <- "both"
effect_neighborhood       <- 0.10
fs_adjust_covariates      <- c("cd40", "wtkg")

gate_type  <- "point"   # "point" or "ci"
gate_t     <- 1.0       # near-null gate on the OR scale
gate_draws <- 3000L

nb_boots   <- 200L      # Tier-1 bootstrap iterations (reduce for speed)
run_tier1  <- TRUE     # FALSE: skip the full-bootstrap comparison (gate-only,
                        #        fast CI-style check); the gate still runs in-fit
run_null   <- TRUE      # also check the gate rejects under the null
display_digits <- 3

set.seed(dgm_seed)
n_workers <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))

.fmt_ci <- function(e, l, u, d = display_digits)
  sprintf("%.*f (%.*f, %.*f)", d, e, d, l, d, u)
```

## Data and data-generating model

```{r data}
#| code-fold: true
#| code-summary: "ACTG175 prep + DGM calibration"
# ACTG175: arm 1 = ZDV+ddI (experimental), arm 3 = ddI (control)
actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$id    <- seq_len(nrow(actg_df))
actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)
actg_df$y_binary <- as.integer(actg_df$cd420 <= actg_df$cd40)   # adverse = no improvement
actg_df <- actg_df[!is.na(actg_df$cd420), ]

# DGM markers for the true harm subgroup Q = {cd40 > 390 & wtkg > 80}
actg_df$z1 <- as.factor(ifelse(actg_df$cd40 > 390, 1L, 0L))
actg_df$z2 <- as.factor(ifelse(actg_df$wtkg > 80,  1L, 0L))
actg_df$ar_naive  <- as.integer(actg_df$str2 == 0)
actg_df$prior_6mo <- as.integer(actg_df$preanti <= 6 * 30.4375)

cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
bin_vars  <- c("hemo", "homo", "drugs", "race", "gender", "symptom", "ar_naive", "prior_6mo")
for (v in bin_vars) actg_df[[v]] <- as.integer(actg_df[[v]])
confounders <- c(cont_vars, bin_vars)

dgm_factor_vars   <- c("z1", "z2")
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)

# Alternative DGM: OR(Q) calibrated to cal_target_or; ITT / Qc near-null.
dgm_alt <- calibrate_glm_interaction(
  data = actg_df, factor_vars = dgm_factor_vars, continuous_vars = cont_vars,
  outcome_var = "y_binary", treatment_var = "treat",
  target_effect = cal_target_or, outcome_type = "binary", effect_measure = "OR",
  subgroup_vars = dgm_subgroup_vars, subgroup_cuts = dgm_subgroup_cuts,
  k_inter_range = c(0, 3), grid_step = 0.05, n_super = 5000L,
  seed = dgm_seed, verbose = FALSE)

# Null DGM: homogeneous (no interaction).
dgm_null <- generate_glm_dgm(
  data = actg_df, factor_vars = dgm_factor_vars, continuous_vars = cont_vars,
  outcome_var = "y_binary", treatment_var = "treat", outcome_type = "binary",
  subgroup_vars = dgm_subgroup_vars, subgroup_cuts = dgm_subgroup_cuts,
  model = "null", n_super = 5000L, seed = dgm_seed, verbose = FALSE)

cat(sprintf("DGM OR(Q) = %.3f  OR(Qc) = %.3f  OR(ITT) = %.3f\n",
    dgm_alt$hazard_ratios$harm_subgroup,
    dgm_alt$hazard_ratios$no_harm_subgroup,
    dgm_alt$hazard_ratios$overall))
```

The data-generating harm subgroup is $Q = \{\texttt{cd40} > 390 \text{ and } \texttt{wtkg} > 80\}$,
calibrated to OR $\approx$ `r sprintf("%.1f", cal_target_or)`; $Q^c$ and ITT are near-null.

## One alternative trial: fit with the integrated gate

```{r fit-gate}
df_sim <- simulate_from_glm_dgm(dgm_alt, n = sim_n, seed = alt_seed)

fs <- forestsearch(
  df.analysis = df_sim, confounders.name = confounders,
  outcome.name = "y_sim", treat.name = "treat_sim",
  outcome_type = "binary", effect_measure = "OR",
  hr.threshold = fs_effect_threshold, hr.consistency = fs_consistency_threshold,
  pconsistency.threshold = fs_pconsistency_threshold,
  sg_focus = fs_sg_focus, selection_rule = fs_selection_rule,
  effect_neighborhood = effect_neighborhood, maxk = 2L, n.min = 60,
  adjust_covariates = fs_adjust_covariates,
  consistency_method = "resample",
  debias_gate = TRUE,
  debias_gate_args = list(gate = gate_type, t_gate = gate_t, draws = gate_draws),
  parallel_args = list(plan = "multisession", workers = n_workers, show_message = FALSE),
  details = FALSE, quiet = TRUE)
plan("sequential")
```

```{r gate-report}
#| echo: false
#| results: asis
g <- fs$debias_gate
if (is.null(fs$sg.harm)) {
  cat("_No harm subgroup was selected in this trial._\n")
} else if (is.null(g)) {
  cat("_The gate did not run (check consistency_method = \"resample\" and debias_gate = TRUE)._\n")
} else {
  cat(sprintf("**Selected harm subgroup:** `%s` (n = %d)\n\n",
              paste(fs$sg.harm, collapse = " & "), g$n_selected))
  cat(sprintf("**`harm_flag_debiased` = %s** ", fs$harm_flag_debiased))
  cat(sprintf("(de-biased %s %.3f %s gate %.2f)\n\n",
              g$measure, g$debiased$est,
              if (gate_type == "point") "vs" else "lower-bound vs", g$gate$t_gate))
  cat(sprintf("Selection bias %.4f (log scale); bootstrap selection rate %.2f; ",
              g$selection_bias, g$selection_rate))
  cat(sprintf("gate compute time %.2f s.\n", g$timing_seconds))
  cc <- g$complement
  if (!is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est))
    cat(sprintf("\n**Complement** (n = %d): naive %s %.3f -> de-biased %.3f (selection bias %.4f).\n",
                cc$n, g$measure, cc$naive$est, cc$debiased$est, cc$selection_bias))
}
```

## Gate checks (internal consistency)

These validate the **wiring** and the **sound (full-space) family** by construction.
They assert nothing about the flag outcome on a single simulated trial — only that the
gate ran on the resample path, that its naive estimate reproduces an independent adjusted
refit on the selected subgroup (a scale check needing no bootstrap), that the candidate
family is the full enumerated space rather than the screened survivors, and that the
correction is finite and moves toward the null. A hard failure stops the render.

```{r checks}
#| echo: false
#| results: asis
g <- fs$debias_gate
if (is.null(fs$sg.harm)) {
  cat("_No subgroup selected in this trial; gate checks not applicable._\n")
} else if (is.null(g)) {
  cat("_Gate did not run; cannot perform checks (see report above)._\n")
} else {
  chk <- list()
  add <- function(name, pass, detail) {
    res <- if (isTRUE(pass)) "PASS" else if (isFALSE(pass)) "FAIL" else "SKIP"
    chk[[length(chk) + 1L]] <<- data.frame(check = name, result = res,
                                           detail = detail, stringsAsFactors = FALSE)
  }

  # (1) Gate engaged on the resample path (no silent fallback).
  add("gate engaged", !is.null(g) && !is.null(fs$harm_flag_debiased),
      "fs$debias_gate and harm_flag_debiased populated")

  # (2) Naive estimate == independent adjusted refit on the selected subgroup,
  #     on the effect (OR) scale.  Direct scale/wiring check; no bootstrap needed.
  sgid    <- fs$grp.consistency$sg.harm.id
  aligned <- !is.null(sgid) && length(sgid) == nrow(df_sim)
  form    <- stats::as.formula(paste("y_sim ~ treat_sim",
               if (length(fs_adjust_covariates))
                 paste("+", paste(fs_adjust_covariates, collapse = " + ")) else ""))
  refit_or <- function(rows) tryCatch(
    exp(stats::coef(stats::glm(form, family = stats::binomial(),
                               data = df_sim[rows, , drop = FALSE]))[["treat_sim"]]),
    error = function(e) NA_real_)
  if (aligned) {
    or_ref <- refit_or(sgid == 1)
    add("naive == independent refit",
        is.finite(or_ref) && abs(log(or_ref) - log(g$naive$est)) < 1e-3,
        sprintf("gate naive OR %.4f vs independent refit %.4f", g$naive$est, or_ref))
  } else {
    add("naive == independent refit", NA, "subgroup membership not row-aligned; skipped")
  }

  # (3) Family is the FULL enumerated space, not the screened survivors.
  n_surv <- tryCatch(nrow(fs$find.grps$out.found$hr.subgroups),
                     error = function(e) NA_integer_)
  add("family >= survivors", if (is.na(n_surv)) NA else g$n_family >= n_surv,
      sprintf("n_family = %d; screened survivors = %s%s", g$n_family,
              if (is.na(n_surv)) "NA" else n_surv,
              if (!is.na(n_surv) && g$n_family > n_surv) " (full space confirmed)"
              else if (!is.na(n_surv) && g$n_family == n_surv) " (equal -- verify enumeration)"
              else ""))

  # (4) Finiteness and bounds.
  add("estimates finite",
      all(is.finite(c(g$naive$est, g$debiased$est, g$selection_bias))),
      sprintf("naive %.3f, de-biased %.3f, sel.bias %.4f",
              g$naive$est, g$debiased$est, g$selection_bias))
  add("selection_rate in [0,1]",
      is.finite(g$selection_rate) && g$selection_rate >= 0 && g$selection_rate <= 1,
      sprintf("selection_rate = %.3f", g$selection_rate))

  # (5) Correction direction: de-bias shrinks a harm estimate toward the null.
  null_val <- if (g$measure %in% c("OR", "RR", "IRR")) 1 else 0
  dir_ok <- if (g$naive$est > null_val) g$debiased$est <= g$naive$est * 1.001 else TRUE
  add("correction toward null", dir_ok,
      sprintf("de-biased %.3f %s naive %.3f", g$debiased$est,
              if (dir_ok) "<=" else ">", g$naive$est))

  # (6) Complement: naive == independent refit on the complement, and the
  #     de-biased estimate is present and finite (include_complement = TRUE).
  cc <- g$complement
  if (!is.null(cc) && !is.null(cc$naive) && is.finite(cc$naive$est)) {
    if (aligned) {
      or_c <- refit_or(sgid == 0)
      add("complement naive == refit",
          is.finite(or_c) && abs(log(or_c) - log(cc$naive$est)) < 1e-3,
          sprintf("complement naive OR %.4f vs refit %.4f", cc$naive$est, or_c))
    } else {
      add("complement naive == refit", NA, "membership not row-aligned; skipped")
    }
    add("complement de-biased finite",
        !is.null(cc$debiased) && is.finite(cc$debiased$est),
        if (!is.null(cc$debiased))
          sprintf("complement naive %.3f -> de-biased %.3f", cc$naive$est, cc$debiased$est)
        else "complement de-biased missing")
  } else {
    add("complement de-biased", NA, "complement not computed (include_complement = FALSE?)")
  }

  tab <- do.call(rbind, chk)
  cat(knitr::kable(tab, row.names = FALSE), sep = "\n"); cat("\n\n")

  if (any(tab$result == "FAIL"))
    stop("Gate internal-consistency check(s) failed: ",
         paste(tab$check[tab$result == "FAIL"], collapse = "; "), call. = FALSE)
}
```

## Validation against the full (Tier-1) bootstrap

```{r tier1}
#| echo: false
#| results: asis
if (!is.null(fs$sg.harm) && isTRUE(run_tier1)) {
  t0 <- proc.time()
  fs_bc <- forestsearch_bootstrap_dofuture(
    fs.est = fs, nb_boots = nb_boots, show_three = FALSE, details = FALSE,
    parallel_args = list(plan = "multisession", workers = n_workers, show_message = FALSE))
  plan("sequential")
  t_tier1 <- as.numeric((proc.time() - t0)["elapsed"])

  .fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0  <- .fl(fs_bc$H_estimates, "H0");  H0l <- .fl(fs_bc$H_estimates, "H0_lower"); H0u <- .fl(fs_bc$H_estimates, "H0_upper")
  H2  <- .fl(fs_bc$H_estimates, "H2");  H2l <- .fl(fs_bc$H_estimates, "H2_lower"); H2u <- .fl(fs_bc$H_estimates, "H2_upper")
  g   <- fs$debias_gate

  cc  <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  Hc0 <- .fl(fs_bc$Hc_estimates, "H0"); Hc0l <- .fl(fs_bc$Hc_estimates, "H0_lower"); Hc0u <- .fl(fs_bc$Hc_estimates, "H0_upper")
  Hc2 <- .fl(fs_bc$Hc_estimates, "H2"); Hc2l <- .fl(fs_bc$Hc_estimates, "H2_lower"); Hc2u <- .fl(fs_bc$Hc_estimates, "H2_upper")

  q  <- c("naive OR (95% CI)", "de-biased OR (95% CI)")
  t2 <- c(.fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          .fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  t1 <- c(.fmt_ci(H0, H0l, H0u), .fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    q  <- c(q, "complement naive OR (95% CI)", "complement de-biased OR (95% CI)")
    t2 <- c(t2, .fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                .fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    t1 <- c(t1, .fmt_ci(Hc0, Hc0l, Hc0u), .fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `Tier 2 (gate)` = t2,
                     `Tier 1 (bootstrap)` = t1, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE), sep = "\n"); cat("\n\n")

  # Scale cross-check: the gate's naive OR and the bootstrap's naive H0 are the
  # same quantity computed two ways; they should agree up to rounding.
  if (is.finite(H0) && is.finite(g$naive$est)) {
    scale_ok <- abs(log(g$naive$est) - log(H0)) < 1e-2
    cat(sprintf("Scale cross-check (naive gate vs bootstrap H0): **%s** (%.4f vs %.4f).\n\n",
                if (scale_ok) "PASS" else "FAIL", g$naive$est, H0))
  }

  cat(sprintf("Timing: Tier 1 (%d boots) %.1f s (%.1f min); Tier 2 gate %.2f s -> **%.0fx** faster.\n",
              nb_boots, t_tier1, t_tier1 / 60, g$timing_seconds,
              if (g$timing_seconds > 0) t_tier1 / g$timing_seconds else NA))
  cat("\nThe naive-OR row is a scale sanity check; agreement on the de-biased row validates ",
      "Tier 2 against Tier 1. The gate's de-biased CI is approximate (subgroup robust SE) ",
      "and narrower than the Tier-1 infinitesimal-jackknife CI.\n", sep = "")
} else if (!is.null(fs$sg.harm)) {
  cat(sprintf("_Tier-1 bootstrap comparison skipped (`run_tier1 = FALSE`). The gate ran in-fit in %.2f s; ",
              fs$debias_gate$timing_seconds))
  cat(sprintf("de-biased %s = %.3f, `harm_flag_debiased` = %s._\n",
              fs$debias_gate$measure, fs$debias_gate$debiased$est, fs$harm_flag_debiased))
}
```

## Null trial: the gate should reject

```{r null-check}
#| echo: false
#| results: asis
#| eval: !expr run_null
df_null <- simulate_from_glm_dgm(dgm_null, n = sim_n, seed = null_seed)
fs0 <- forestsearch(
  df.analysis = df_null, confounders.name = confounders,
  outcome.name = "y_sim", treat.name = "treat_sim",
  outcome_type = "binary", effect_measure = "OR",
  hr.threshold = fs_effect_threshold, hr.consistency = fs_consistency_threshold,
  pconsistency.threshold = fs_pconsistency_threshold,
  sg_focus = fs_sg_focus, selection_rule = fs_selection_rule,
  effect_neighborhood = effect_neighborhood, maxk = 2L, n.min = 60,
  adjust_covariates = fs_adjust_covariates, consistency_method = "resample",
  debias_gate = TRUE,
  debias_gate_args = list(gate = gate_type, t_gate = gate_t, draws = gate_draws),
  parallel_args = list(plan = "multisession", workers = n_workers, show_message = FALSE),
  details = FALSE, quiet = TRUE)
plan("sequential")

if (is.null(fs0$sg.harm)) {
  cat("No subgroup selected under the null (the search itself found nothing).\n")
} else {
  g0 <- fs0$debias_gate
  cat(sprintf("Null trial selected `%s`; naive OR %.3f -> de-biased %.3f -> **harm_flag_debiased = %s**.\n",
              paste(fs0$sg.harm, collapse = " & "), g0$naive$est, g0$debiased$est,
              fs0$harm_flag_debiased))
  cat("\nUnder the null the de-biased estimate should collapse toward the ITT effect and fail the gate.\n")
}
```


## Validation against the full (Tier-1) bootstrap

```{r tier1-null}
#| echo: false
#| results: asis

fs <- fs0
if (!is.null(fs$sg.harm) && isTRUE(run_tier1)) {
  t0 <- proc.time()
  fs_bc <- forestsearch_bootstrap_dofuture(
    fs.est = fs, nb_boots = nb_boots, show_three = FALSE, details = FALSE,
    parallel_args = list(plan = "multisession", workers = n_workers, show_message = FALSE))
  plan("sequential")
  t_tier1 <- as.numeric((proc.time() - t0)["elapsed"])

  .fl <- function(o, n) tryCatch(as.numeric(o[[n]]), error = function(e) NA_real_)
  H0  <- .fl(fs_bc$H_estimates, "H0");  H0l <- .fl(fs_bc$H_estimates, "H0_lower"); H0u <- .fl(fs_bc$H_estimates, "H0_upper")
  H2  <- .fl(fs_bc$H_estimates, "H2");  H2l <- .fl(fs_bc$H_estimates, "H2_lower"); H2u <- .fl(fs_bc$H_estimates, "H2_upper")
  g   <- fs$debias_gate

  cc  <- g$complement
  has_cc <- !is.null(cc) && !is.null(cc$debiased) && is.finite(cc$debiased$est)
  Hc0 <- .fl(fs_bc$Hc_estimates, "H0"); Hc0l <- .fl(fs_bc$Hc_estimates, "H0_lower"); Hc0u <- .fl(fs_bc$Hc_estimates, "H0_upper")
  Hc2 <- .fl(fs_bc$Hc_estimates, "H2"); Hc2l <- .fl(fs_bc$Hc_estimates, "H2_lower"); Hc2u <- .fl(fs_bc$Hc_estimates, "H2_upper")

  q  <- c("naive OR (95% CI)", "de-biased OR (95% CI)")
  t2 <- c(.fmt_ci(g$naive$est, g$naive$lower, g$naive$upper),
          .fmt_ci(g$debiased$est, g$debiased$lower, g$debiased$upper))
  t1 <- c(.fmt_ci(H0, H0l, H0u), .fmt_ci(H2, H2l, H2u))
  if (has_cc) {
    q  <- c(q, "complement naive OR (95% CI)", "complement de-biased OR (95% CI)")
    t2 <- c(t2, .fmt_ci(cc$naive$est, cc$naive$lower, cc$naive$upper),
                .fmt_ci(cc$debiased$est, cc$debiased$lower, cc$debiased$upper))
    t1 <- c(t1, .fmt_ci(Hc0, Hc0l, Hc0u), .fmt_ci(Hc2, Hc2l, Hc2u))
  }
  comp <- data.frame(quantity = q, `Tier 2 (gate)` = t2,
                     `Tier 1 (bootstrap)` = t1, check.names = FALSE)
  cat(knitr::kable(comp, row.names = FALSE), sep = "\n"); cat("\n\n")

  # Scale cross-check: the gate's naive OR and the bootstrap's naive H0 are the
  # same quantity computed two ways; they should agree up to rounding.
  if (is.finite(H0) && is.finite(g$naive$est)) {
    scale_ok <- abs(log(g$naive$est) - log(H0)) < 1e-2
    cat(sprintf("Scale cross-check (naive gate vs bootstrap H0): **%s** (%.4f vs %.4f).\n\n",
                if (scale_ok) "PASS" else "FAIL", g$naive$est, H0))
  }

  cat(sprintf("Timing: Tier 1 (%d boots) %.1f s (%.1f min); Tier 2 gate %.2f s -> **%.0fx** faster.\n",
              nb_boots, t_tier1, t_tier1 / 60, g$timing_seconds,
              if (g$timing_seconds > 0) t_tier1 / g$timing_seconds else NA))
  cat("\nThe naive-OR row is a scale sanity check; agreement on the de-biased row validates ",
      "Tier 2 against Tier 1. The gate's de-biased CI is approximate (subgroup robust SE) ",
      "and narrower than the Tier-1 infinitesimal-jackknife CI.\n", sep = "")
} else if (!is.null(fs$sg.harm)) {
  cat(sprintf("_Tier-1 bootstrap comparison skipped (`run_tier1 = FALSE`). The gate ran in-fit in %.2f s; ",
              fs$debias_gate$timing_seconds))
  cat(sprintf("de-biased %s = %.3f, `harm_flag_debiased` = %s._\n",
              fs$debias_gate$measure, fs$debias_gate$debiased$est, fs$harm_flag_debiased))
}
```

## What this example checks

This is the ongoing smoke test for the integrated gate. The **Gate checks** section
asserts the wiring and the full-space family by construction (gate engaged, naive estimate
reproduces an independent adjusted refit, family larger than the screened survivors,
correction finite and toward the null) and stops the render on a hard failure. A passing
run shows, on the
alternative trial, that `forestsearch(debias_gate = TRUE)` selects the true harm subgroup,
that `harm_flag_debiased` is `TRUE`, that the gate's naive OR matches the bootstrap's naive
OR (scale check) and its de-biased OR tracks the bootstrap's bias-corrected OR (the
Tier-2-vs-Tier-1 validation), and that the gate is markedly faster than the full bootstrap.
Under the null trial, the de-biased estimate collapses toward the ITT effect and
`harm_flag_debiased` is `FALSE` (or no subgroup is selected) — the false-discovery
mitigation the gate is designed to provide.

Set `run_tier1 <- FALSE` for a fast gate-only run (a CI-style check): the gate still runs
inside the fit and reports `harm_flag_debiased`, but the multi-minute bootstrap comparison
is skipped. Keep `run_tier1 <- TRUE` for the full Tier-2-vs-Tier-1 validation, and
`run_null` toggles the null-trial check independently.
