---
title: "Improving `uniroot()` Performance in R"
subtitle: "Techniques, benchmarks, and alternatives"
author: "Larry Leon"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: false
    code-tools: true
    theme: cosmo
    df-print: paged
    embed-resources: true
execute:
  warning: false
  message: false
  eval: true
editor: source
---

# Overview

`stats::uniroot()` is R's standard univariate root-finder. It calls compiled
C code (`.External2("do_zeroin2")`) implementing Brent's method — a hybrid
of bisection, secant, and inverse quadratic interpolation. Because the
core algorithm is already in C, the usual performance bottleneck is
**not** the root-finder itself but one of:

1. An unnecessarily wide search bracket,
2. Avoidable function evaluations,
3. An overly tight tolerance,
4. An expensive user-supplied `f(x)` written in interpreted R,
5. Calling `uniroot()` repeatedly with a cold start.

This document lays out the main levers in rough effort-to-payoff order,
with runnable benchmarks. The techniques compound when `uniroot()` is
called inside a bootstrap or simulation loop, where per-call savings
multiply by `B × n_subgroups`.

```{r setup}
suppressPackageStartupMessages({
  library(microbenchmark)
})

# A "moderately expensive" objective we can reuse across sections.
# Think of it as a stand-in for a likelihood or estimating equation.
make_objective <- function(target, work = 0L) {
  force(target); force(work)
  function(x) {
    if (work > 0L) for (i in seq_len(work)) sqrt(i)  # fake CPU work
    pnorm(x) - target
  }
}

f <- make_objective(0.975)
```

# Key techniques

## Tighten the search bracket

Brent's method converges superlinearly, but the starting bracket still
matters — every doubling of the interval costs roughly one extra
iteration. If you can bound the root analytically (even loosely), pass
the narrowest honest `interval` you have.

```{r bracket-width}
microbenchmark(
  very_wide = uniroot(f, interval = c(-50, 50)),
  wide      = uniroot(f, interval = c(-10, 10)),
  tight     = uniroot(f, interval = c(1, 3)),
  times = 200L
)
```

## Pre-compute the endpoints via `f.lower` / `f.upper`

If you've already evaluated `f` at the bracket ends — very common, since
you often check signs to confirm a root is bracketed — pass them in.
This saves two function evaluations per call, which is typically a
15–25% speedup when `f` is the bottleneck.

```{r fcache}
fl <- f(1); fu <- f(3)

microbenchmark(
  without_endpoints = uniroot(f, interval = c(1, 3)),
  with_endpoints    = uniroot(f, interval = c(1, 3),
                              f.lower = fl, f.upper = fu),
  times = 500L
)
```

## Relax `tol` when appropriate

The default is `.Machine$double.eps^0.25` ≈ `1.22e-4` for the *x*
tolerance. For things like confidence-interval endpoints or HR
thresholds where 4–5 significant figures suffice, leave the default or
loosen it. Tightening costs ~3–4 extra evaluations per order of
magnitude.

```{r tolerance}
microbenchmark(
  tol_1e_10 = uniroot(f, interval = c(1, 3), tol = 1e-10),
  tol_1e_8  = uniroot(f, interval = c(1, 3), tol = 1e-8),
  default   = uniroot(f, interval = c(1, 3)),
  tol_1e_3  = uniroot(f, interval = c(1, 3), tol = 1e-3),
  times = 500L
)
```

## Avoid `extendInt` in hot loops

`extendInt = "yes"` (or `"upX"/"downX"`) is great for robustness but
each extension costs extra evaluations and risks an unbracketed
diverge. Inside a bootstrap where you know monotonicity, bracket
explicitly and only fall back to extension if a bracket check fails.

## Move `f` to C++ via Rcpp

When `f` itself is the bottleneck — for example, if it calls
`survfit()`, does numerical integration, or loops over observations in
R — rewriting it in C++ is usually the biggest single win. The
R-to-C callback overhead per evaluation is small relative to
interpreted R loops inside `f`.

The chunk below simulates an expensive `f`; the tighter bracket still
helps, but notice how the absolute timings rise — this is the regime
where Rcpp pays off.

```{r expensive-f}
f_slow <- make_objective(0.975, work = 1000L)
microbenchmark(
  wide_slow_f  = uniroot(f_slow, interval = c(-10, 10)),
  tight_slow_f = uniroot(f_slow, interval = c(1, 3)),
  times = 100L
)
```

## Warm-start across batched calls

Classic win when calling `uniroot()` many times with slowly varying
parameters — the common case inside a bootstrap, a grid search, or a
profile likelihood. Instead of starting each call with the same wide
bracket, initialize with a small window around the previous root.

```{r warm-start}
targets <- seq(0.50, 0.999, length.out = 1000)

naive_batch <- function(targets) {
  vapply(targets, function(tg) {
    uniroot(make_objective(tg), interval = c(-10, 10))$root
  }, numeric(1))
}

warm_batch <- function(targets, half_width = 0.5) {
  out  <- numeric(length(targets))
  prev <- 0
  for (i in seq_along(targets)) {
    out[i] <- uniroot(
      make_objective(targets[i]),
      interval  = c(prev - half_width, prev + half_width),
      extendInt = "yes"  # safety net for rare misses
    )$root
    prev <- out[i]
  }
  out
}

# Sanity check — the two approaches agree to within tol
stopifnot(max(abs(naive_batch(targets) - warm_batch(targets))) < 1e-3)

microbenchmark(
  naive = naive_batch(targets),
  warm  = warm_batch(targets),
  times = 5L
)
```

## Cache endpoint evaluations across batched calls

If many calls share bracket endpoints and the "expensive" part of `f`
is independent of the varying parameter (common after profiling out a
nuisance parameter in a survival estimating equation), compute the
endpoint pieces once and reuse them.

```{r endpoint-cache}
fixed_bracket_naive <- function(targets) {
  vapply(targets, function(tg) {
    g <- function(x) pnorm(x) - tg
    uniroot(g, interval = c(-5, 5))$root
  }, numeric(1))
}

fixed_bracket_cached <- function(targets) {
  lo <- -5; hi <- 5
  plo <- pnorm(lo); phi <- pnorm(hi)  # target-independent work
  vapply(targets, function(tg) {
    g <- function(x) pnorm(x) - tg
    uniroot(g, interval = c(lo, hi),
            f.lower = plo - tg, f.upper = phi - tg)$root
  }, numeric(1))
}

microbenchmark(
  no_cache = fixed_bracket_naive(targets),
  cached   = fixed_bracket_cached(targets),
  times = 10L
)
```

# Alternatives worth knowing

## `rootSolve::uniroot.all()`

Grids an interval, then polishes each sign change. Use when
monotonicity is **not** guaranteed — for example, when searching for
multiple crossings of a test statistic.

```{r rootsolve}
#| eval: !expr 'requireNamespace("rootSolve", quietly = TRUE)'
ff <- function(x) sin(x) - 0.3
rootSolve::uniroot.all(ff, interval = c(0, 4 * pi))
```

## `pracma::brentDekker()` / `pracma::fzero()`

Same Brent–Dekker family, sometimes with more diagnostic output or
slightly different behavior on pathological functions. Usually within
noise of base `uniroot()`.

```{r pracma}
#| eval: !expr 'requireNamespace("pracma", quietly = TRUE)'
microbenchmark(
  base   = uniroot(f, interval = c(1, 3)),
  pracma = pracma::brentDekker(f, a = 1, b = 3),
  times = 200L
)
```

## Hand-rolled Newton for monotone, smooth objectives

When you have a cheap derivative and know the objective is monotone and
$C^1$, Newton's method typically beats Brent because it throws away the
bracket-safety work that Brent does. The normal-quantile example below
illustrates the scale of the difference; `qnorm()` is included as the
reference "fully compiled" option.

```{r newton}
newton_qnorm <- function(p, x0 = 0, tol = 1e-8, maxit = 50L) {
  x <- x0
  for (i in seq_len(maxit)) {
    fx <- pnorm(x) - p
    if (abs(fx) < tol) return(x)
    dfx <- dnorm(x)
    if (dfx < 1e-300) break
    x <- x - fx / dfx
  }
  x
}

stopifnot(abs(newton_qnorm(0.975) - qnorm(0.975)) < 1e-6)

f975 <- make_objective(0.975)
microbenchmark(
  uniroot_call = uniroot(f975, interval = c(-5, 5))$root,
  newton_call  = newton_qnorm(0.975),
  qnorm_call   = qnorm(0.975),       # reference
  times = 500L
)
```

## Other directions

- **`nleqslv::nleqslv()`** — when you find yourself wrapping `uniroot()`
  to solve several coupled equations, switch to a proper nonlinear
  systems solver.
- **`RcppNumerical`** — Brent-style root-finders callable directly from
  C++, useful when the enclosing loop is also in C++.
- **Table-and-interpolate** — if the objective is smooth and called
  millions of times with the same structure, a precomputed spline over
  the parameter space can replace root-finding entirely.

# Decision checklist

When a `uniroot()`-heavy routine is slow, work through these in order:

1. **Profile first.** `profvis::profvis({ ... })` to confirm
   `uniroot()` (or its objective) is actually the hotspot.
2. **Narrow the bracket** using domain knowledge.
3. **Pass `f.lower` / `f.upper`** whenever available.
4. **Relax `tol`** if 4–5 significant figures is enough.
5. **Avoid `extendInt`** in hot loops.
6. **Warm-start** across batched calls.
7. **Move `f` into Rcpp** if it dominates the profile.
8. **Consider Newton/Halley** for monotone $C^1$ objectives with a
   cheap derivative.

# Session information

```{r sessioninfo}
sessionInfo()
```
