---
title: "ACTG175 Continuous Harm Search: Single-Data Analysis"
format: html
---

```{r setup}
library(forestsearch)
library(speff2trial)
library(gt)

set.seed(8316951L)
```

## Data Preparation

```{r data-prep}
# ACTG175: arm 1 = ZDV+ddI (experimental), arm 3 = ddI (control)
actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$id <- seq_len(nrow(actg_df))

# Treatment: ZDV+ddI = 1, ddI = 0 (NO switching)
actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)

actg_df <- actg_df[!is.na(actg_df$cd420), ]

# -- Continuous outcomes -------------------------------------------------------
# Adverse scale: CD4 decline (higher = worse)
actg_df$y_decline <- actg_df$cd40 - actg_df$cd420

# Beneficial scale: CD4 improvement (higher = better)
actg_df$y_improve <- actg_df$cd420 - actg_df$cd40

cat(sprintf("N = %d\n", nrow(actg_df)))
cat(sprintf("Mean CD4 decline (y_decline):    %.1f\n", mean(actg_df$y_decline)))
cat(sprintf("  ZDV+ddI (T=1): %.1f\n", mean(actg_df$y_decline[actg_df$treat == 1])))
cat(sprintf("  ddI     (T=0): %.1f\n", mean(actg_df$y_decline[actg_df$treat == 0])))
cat(sprintf("Mean CD4 improvement (y_improve): %.1f\n", mean(actg_df$y_improve)))
cat(sprintf("  ZDV+ddI (T=1): %.1f\n", mean(actg_df$y_improve[actg_df$treat == 1])))
cat(sprintf("  ddI     (T=0): %.1f\n", mean(actg_df$y_improve[actg_df$treat == 0])))
```

## Covariates

```{r covariates}
cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")

bin_vars <- c("hemo", "homo", "drugs", "race", "gender", "symptom")
for (v in bin_vars) actg_df[[v]] <- as.factor(actg_df[[v]])

confounders <- c(cont_vars, bin_vars)
```

## ForestSearch (continuous, adverse scale)

FS uses `y_decline` (higher = worse) with `adverse_outcome = TRUE`.
No outcome flip — MD > c1 means treatment increases CD4 decline = **harm**.

```{r fs-run}
# Thresholds on the mean-difference scale (CD4 count units)
c1 <- 10    # screening: MD >= 10
c2 <- 5     # consistency: MD >= 5

cat(sprintf("c1 (threshold) = %.0f, c2 (consistency) = %.0f\n", c1, c2))

fs_result <- forestsearch(
  df.analysis      = actg_df,
  confounders.name = confounders,
  outcome.name     = "y_decline",
  event.name       = "y_decline",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "continuous",
  effect_measure   = "MD",
  adverse_outcome  = TRUE,
  seedit           = 8316951L,
  n.min            = 60L,
  d0.min           = 10L,
  d1.min           = 10L,
  hr.threshold     = c1,
  hr.consistency   = c2,
  pconsistency.threshold = 0.80,
  fs.splits        = 500L,
  max_subgroups_search = 15,
  maxk             = 2L,
  sg_focus         = "minSG",
  use_lasso        = FALSE,
  use_grf          = TRUE,
  is.RCT           = TRUE,
  vi.grf.min       = -0.2,
  dmin.grf         = 0,
  cont.cutoff      = 4L,
  details          = TRUE,
  quiet            = FALSE,
  parallel_args    = list(plan = "sequential", workers = 1,
                          show_message = FALSE)
)
```

### Identified Subgroups

```{r fs-results}
fs_tabs <- sg_tables(fs_result, ndecimals = 3, which_df = "est")
fs_tabs$sg10_out
```

### Treatment Effect Estimates

```{r fs-estimates}
fs_tabs$tab_estimates
```

## GRF Standalone (continuous, beneficial scale)

GRF uses `y_improve` (higher = better) with `adverse_outcome = FALSE` (default).
CATE < 0 means treatment decreases improvement = **harm**.

```{r grf-run}
grf_result <- grf.subg.harm.glm(
  data             = actg_df,
  confounders.name = confounders,
  outcome.name     = "y_improve",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "continuous",
  effect_measure   = "MD",
  dmin.grf         = 0.1,
  maxdepth         = 2L,
  RCT              = TRUE,
  sg.criterion     = "mDiff",
  details          = TRUE
)

cat("\n=== GRF Result ===\n")
if (!is.null(grf_result$sg.harm.id) && length(grf_result$sg.harm.id) > 0) {
  cat("Subgroup:", paste(grf_result$sg.harm.id, collapse = " & "), "\n")
  grf_H <- grf_result$data$treat.recommend == 0
  cat(sprintf("Size(H): %d,  Size(Hc): %d\n", sum(grf_H), sum(!grf_H)))
  cat(sprintf("CATE(H):  %.2f (treatment effect on improvement in H)\n",
      grf_result$cate_sg))
  cat(sprintf("CATE(Hc): %.2f\n", grf_result$cate_sgc))
} else {
  cat("GRF: no subgroup identified.\n")
}
```

### GRF Treatment Effect Estimates

```{r grf-results}
grf_tabs <- sg_tables(grf_result, ndecimals = 3, which_df = "est")
grf_tabs$tab_estimates
```

## Side-by-Side Comparison

```{r comparison}
cat("=== FS (adverse Y = CD4 decline, adverse_outcome = TRUE) ===\n")
fs_tabs$tab_estimates

cat("\n=== GRF (beneficial Y = CD4 improvement, adverse_outcome = FALSE) ===\n")
grf_tabs$tab_estimates
```
