---
title: "Exploratory Subgroup Identification with ForestSearch"
subtitle: "German Breast Cancer Study Group (GBSG) Analysis"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# Initialize timing
timings <- list()
t_vignette_start <- proc.time()
```

# Analysis Configuration

All tunable parameters for this vignette are defined here in one place,
so that GRF, ForestSearch, frontier-CI, bootstrap, and cross-validation
settings are not scattered through the document.

## Run-time flags

Cross-validation defaults to **off** because it is the most expensive
part of the document.  Set the flags below to `TRUE` to enable.

```{r cv-flags}
run_cv_kfold <- FALSE   # K-fold CV (forestsearch_tenfold + summarize_cv_results)
run_cv_oob   <- FALSE   # Leave-one-out CV (forestsearch_Kfold)
# K-fold CV: number of repeated 10-fold partitions
# Use Ksims >= 50 for production
Ksims <- 50
```

## Parallel processing & iteration counts

```{r parallel-setup}
# Detect available cores
n_cores_total <- parallel::detectCores()
n_cores       <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))
cat("Using", n_cores, "of", n_cores_total, "total cores for parallel processing")

# Bootstrap iterations: 500-2000 for production; reduced here for vignette
NB <- 50
```

## GRF parameters

Used by the preliminary GRF analysis.  Note that `dmin.grf` and
`frac.tau` differ between the GRF standalone analysis and the
GRF candidate-selection inside ForestSearch (see `fs_params`).

```{r grf-params}
grf_params <- list(
  maxdepth                  = 2L,
  n.min                     = 60L,
  dmin.grf                  = 12L,
  frac.tau                  = 0.60,
  details                   = TRUE,
  return_selected_cuts_only = FALSE
)
```

## ForestSearch parameters

```{r fs-params}
fs_params <- list(
  # Threshold parameters (per León et al. 2024)
  hr.threshold           = 1.25,
  hr.consistency         = 1.0,
  pconsistency.threshold = 0.90,
  stop_threshold         = 0.95,
  # Search configuration
  sg_focus               = "hrMaxSG",
  selection_rule         = "pareto",
  max_subgroups_search   = 20L,
  use_twostage           = TRUE,
  # Factor selection
  use_grf                   = TRUE,
  dmin.grf                  = 4L,
  frac.tau                  = 0.80,
  return_selected_cuts_only = TRUE,
  use_lasso                 = TRUE,
  cut_type                  = "default",
  conf.cont_jcuts           = list(er = 10, pgr = 10),
  # Subgroup constraints
  maxk   = 2L,
  n.min  = 60L,
  d0.min = 10L,
  d1.min = 10L,
    # Consistency evaluation
  fs.splits = 1000L,
  # Output options
  show_candidate_summary = TRUE,
  details           = TRUE,
  plot.sg           = TRUE
)
```

## Frontier-CI and table-display parameters

Used by `compute_frontier_cis()`, `plot_pareto_frontier()`, and
`pareto_frontier_table()` in the Pareto frontier diagnostic.

```{r frontier-params}
frontier_params <- list(
  n_splits      = 1000L,
  seed          = 1L,
  xlim          = c(1.25, 3.0),
  digits_effect = 2L,
  digits_pcons  = 2L,
  digits_ci     = 2L
)
```

## CV-result placeholders

Initialised to `NULL` so downstream references (forest plot, timing
summary) work whether or not the CV steps actually run.

```{r cv-placeholders}
fs_kfold       <- NULL
fs_OOB         <- NULL
cv_diag        <- NULL
metrics_tables <- NULL
tables         <- NULL
timings$kfold  <- NA_real_
timings$oob    <- NA_real_
```



# Introduction

This vignette demonstrates the **ForestSearch** methodology for exploratory
subgroup identification in survival analysis, as described in
León et al. (2024) *Statistics in Medicine*.

## Motivation

In clinical trials, particularly oncology, subgroup analyses are essential for:

- Evaluating treatment effect consistency across patient populations
- Identifying subgroups where treatment may be detrimental (harm)
- Characterizing subgroups with enhanced benefit
- Informing regulatory decisions and clinical practice

While prespecified subgroups provide stronger evidence, important subgroups
based on patient characteristics may not be anticipated. ForestSearch provides
a principled approach to **exploratory** subgroup identification with proper
statistical inference.

## Methodology Overview

ForestSearch identifies subgroups through:

1. **Candidate factor selection**: Using LASSO and/or Generalized Random Forests (GRF)
2. **Exhaustive subgroup search**: Evaluating all combinations up to `maxk` factors
3. **Consistency-based selection**: Applying splitting consistency criteria
4. **Bootstrap bias correction**: Adjusting for selection-induced optimism
5. **Cross-validation**: Assessing algorithm stability

The key innovation is the **splitting consistency criterion**: a subgroup is
considered "consistent with harm" if, when randomly split 50/50 many times,
both halves consistently show hazard ratios ≥ 1.0 (for example if 1.0
represents a meaningful "harm threshold").


# Setup

## Load Required Packages

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(policytree)
library(doFuture)

# Optional packages for enhanced output
library(patchwork)
library(weightedsurv)

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```


# Data: German Breast Cancer Study Group Trial

## Study Background

The GBSG trial evaluated hormonal treatment (tamoxifen) versus chemotherapy
in node-positive breast cancer patients. Key characteristics:

- **Sample size**: N = 686
- **Outcome**: Recurrence-free survival time
- **Censoring rate**: ~56%
- **Treatment**: Hormonal therapy (tamoxifen) vs. chemotherapy

## Data Preparation

```{r data-setup}
# Load GBSG data (included in forestsearch package)
df.analysis <- gbsg

# Prepare analysis variables
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
  treat <- hormon
})

# Define variable roles
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name <- "status"
id.name <- "id"
treat.name <- "hormon"

# Display data structure
cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```

## Baseline Characteristics

```{r baseline-table}
create_summary_table(
  data = df.analysis,
  treat_var = treat.name,
  table_title = "GBSG Baseline Characteristics by Treatment Arm",
  vars_continuous = c("age", "nodes", "size", "er", "pgr"),
  vars_categorical = c("grade", "meno"),
  font_size = 12
)
```

## Kaplan-Meier Analysis (ITT Population)

```{r km-itt}
#| fig-width: 8
#| fig-height: 5

# Prepare counting process data for KM plot
dfcount <- df_counting(
  df = df.analysis,
  by.risk = 6,
  tte.name = outcome.name,
  event.name = event.name,
  treat.name = treat.name
)

# Plot with confidence intervals and log-rank test
plot_weighted_km(
  dfcount,
  conf.int = TRUE,
  show.logrank = TRUE,
  ymax = 1.05,
  xmed.fraction = 0.775,
  ymed.offset = 0.125
)
```

The ITT Cox hazard ratio estimate is approximately 0.69 (95% CI: 0.54, 0.89),
suggesting an overall benefit for hormonal therapy.


# Preliminary Analysis: Generalized Random Forests

Before running ForestSearch, we can use GRF to explore potential treatment
effect heterogeneity and identify candidate factors.

```{r grf-analysis}
t0 <- proc.time()

# Fixed args (data + names) come from the analysis context; tunable
# parameters come from grf_params in the Analysis Configuration.
grf_est <- do.call(grf.subg.harm.survival, c(
  list(
    data             = df.analysis,
    confounders.name = confounders.name,
    outcome.name     = outcome.name,
    event.name       = event.name,
    id.name          = id.name,
    treat.name       = treat.name
  ),
  grf_params
))

timings$grf <- (proc.time() - t0)["elapsed"]
```

```{r grf-trees}
#| fig-width: 10
#| fig-height: 4

# Display policy trees
# leaf1 = recommend control, leaf2 = recommend treatment
oldpar <- par(mfrow = c(1, 2))
plot(grf_est$tree1, leaf.labels = c("Control", "Treat"), main = "Depth 1")
plot(grf_est$tree2, leaf.labels = c("Control", "Treat"), main = "Depth 2")
par(oldpar)
```

GRF identifies estrogen receptor status (ER) as a key factor, with ER ≤ 0
suggesting potential harm from hormonal therapy.


# ForestSearch Analysis




## Running ForestSearch

ForestSearch performs an exhaustive search over candidate subgroup combinations
with up to `maxk` factors. Key parameters:

| Parameter | Value | Description |
|-----------|-------|-------------|
| `hr.threshold` | 1.25 | Minimum HR for consistency evaluation |
| `hr.consistency` | 1.0 | Minimum consistency rate for candidates |
| `pconsistency.threshold` | 0.90 | Required consistency for selection |
| `maxk` | 2 | Maximum factors in subgroup definition |
| `n.min` | 60 | Minimum subgroup sample size |
| `d0.min`, `d1.min` | 12 | Minimum events per treatment arm |

::: {.callout-note}
## Selection criteria (`sg_focus`)

`sg_focus` controls which candidate subgroup is chosen once consistency
has been evaluated. The example below uses `"maxSG"` (largest passing
subgroup). Other options:

* `"hr"` -- highest consistency, then largest effect.
* `"minSG"` -- smallest passing subgroup, ties broken by consistency.
* `"hrMaxSG"` -- among candidates whose effect is within
  `effect_neighborhood` (default 10%) of the maximum, pick the
  *largest* sample size.
* `"hrMinSG"` -- among candidates within `effect_neighborhood` of the
  maximum effect, pick the *smallest* sample size.

`"hrMaxSG"` and `"hrMinSG"` use a *lexicographic-with-tolerance* rule:
effect size is primary, but a small relative loss in effect is accepted
to optimize sample size. Setting `effect_neighborhood = 0` reduces these
to a strict max-effect filter. For GLM ratio measures (OR, IRR), the
neighborhood test is applied on the natural scale.
:::

```{r forestsearch-main}
#| fig-width: 10
#| fig-height: 7

t0 <- proc.time()

# Fixed args (data + names) come from the analysis context.  All tunable
# thresholds, search settings, and selection options come from fs_params
# in the Analysis Configuration.  parallel_args depends on n_cores so it
# is built here rather than stored in fs_params.
fs <- do.call(forestsearch, c(
  list(
    df.analysis      = df.analysis,
    confounders.name = confounders.name,
    outcome.name     = outcome.name,
    treat.name       = treat.name,
    event.name       = event.name,
    id.name          = id.name,
    parallel_args    = list(
      plan         = "multisession",
      workers      = n_cores,
      show_message = TRUE
    )
  ),
  fs_params
))

plan("sequential")
timings$forestsearch <- (proc.time() - t0)["elapsed"]

cat("\nForestSearch completed in",
    round(timings$forestsearch, 1), "seconds\n")
```




## ForestSearch Results

### Identified Subgroups

```{r fs-results}
# Generate results tables
res_tabs <- sg_tables(fs, ndecimals = 3, which_df = "est")

# Display top subgroups meeting criteria
res_tabs$sg10_out
```

### Treatment Effect Estimates

```{r fs-estimates}
# ITT and subgroup estimates
res_tabs$tab_estimates
```

### Identified Subgroup Definition

```{r fs-subgroup}
cat("Identified subgroup (H):", paste(fs$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs$df.est$treat.recommend == 0)))
```

ForestSearch identifies **Estrogen ≤ 0** (ER-negative) as the subgroup with
potential harm. This is biologically plausible: tamoxifen is a selective
estrogen receptor modulator with limited efficacy in ER-negative tumors.

### Selection Diagnostic: Pareto Frontier

Beyond the single subgroup chosen by `sg_focus`, it is often informative to
inspect the *Pareto frontier* on (effect, N) -- the set of candidate subgroups
that are not dominated on both axes simultaneously. A candidate is
dominated when another candidate has at least as large an effect *and*
at least as large a sample size, with at least one strict inequality.
The frontier is a post-hoc diagnostic that lists genuine trade-off
alternatives to the selected subgroup; it does **not** affect selection.

For `"hrMinSG"` the selected subgroup may not appear on the frontier --
that focus deliberately prefers small subgroups, which can be N-dominated
by their larger neighbors.

```{r fs-pareto}
# Formatted gt table; selected subgroup is highlighted and marked with ★.
# Works uniformly for survival (HR) and GLM (OR, RR, IRR, RD, IRD, MD)
# outcomes -- effect column label and scale derived from fs$effect_measure.

# 1. Compute the CIs once (this is the expensive step, ~1-3 min on GBSG).
#    Parameters come from frontier_params in the Analysis Configuration.
ci_tab <- compute_frontier_cis(
  fs,
  n_splits = frontier_params$n_splits,
  seed     = frontier_params$seed
)

# 2. Plot, with split-derived CI bars overlaid; same intervals as the table.
p <- plot_pareto_frontier(
  fs,
  ci_table  = ci_tab,
  show_band = TRUE,
  xlim      = frontier_params$xlim
)
print(p)

# 3. Render the table. Pass the same ci_tab so the table and plot
#    show identical intervals.
pareto_frontier_table(
  fs,
  ci_table      = ci_tab,
  digits_effect = frontier_params$digits_effect,
  digits_pcons  = frontier_params$digits_pcons,
  digits_ci     = frontier_params$digits_ci,
  include_dominated= TRUE
)

```


### Selection Narrative

`explain_pareto_selection()` produces a verbal account of why the selected
subgroup wins on the configured lexicographic criterion relative to the
other non-dominated candidates.  It is descriptive (reads the frontier and
config off `fs`), not algorithmic, and weaves the selected subgroup's Naive
95% CI into the selected-candidate paragraph when `ci_table` is supplied.

```{r fs-pareto-explain, results = "asis"}
cat(explain_pareto_selection(fs, ci_table = ci_tab))
```


```{r fs-pareto-data, eval = FALSE}
# Programmatic access — two paths.

# (a) Frontier table with pre-formatted CI strings, one row per frontier
#     member. The CI columns ("Naive 95% CI", "Split ~ 95% CI",
#     "FSBC ~ 95% CI") are character strings like "(1.45, 2.87)".
pareto_frontier_table(
  fs,
  ci_table      = ci_tab,
  format        = "data.table",
  digits_effect = frontier_params$digits_effect,
  digits_pcons  = frontier_params$digits_pcons,
  digits_ci     = frontier_params$digits_ci
)

# (b) Raw numeric CIs in a separate keyed data.table. Same content as (a)
#     but each bound is a numeric column (naive_lcl, naive_ucl, split_lcl,
#     split_ucl, fsbc_estimate, fsbc_lcl, fsbc_ucl, plus split_sd and
#     n_valid_splits). This is what step 1 of fs-pareto produced.
ci_tab
```


### Agreement with GRF

ForestSearch and GRF identify subgroups by different mechanisms --
FS via exhaustive search with consistency filtering, GRF via causal
survival forests plus policy trees -- so agreement on subject-level
membership provides an informal robustness check.  `compare_fs_subgroups()`
produces a publication-ready cross-tabulation plus Cohen's kappa; it
accepts forestsearch and GRF result objects directly, and handles
non-detection (method identifies no subgroup) by assigning all
subjects to the complement.

```{r fs-grf-concordance}
sg_compare <- compare_fs_subgroups(
  fs1               = fs,
  fs2               = grf_est,
  label1            = "ForestSearch",
  label2            = "GRF Standalone",
  id.name           = id.name,
  subgroup_notation = "harm",
  title             = "FS vs GRF: Subgroup Membership Concordance",
  subtitle          = "GBSG observed data (Cox, harm search)",
  font_size         = 13
)
sg_compare$table
```

```{r fs-grf-concordance-stats}
cat("Concordance statistics:\n")
cat(sprintf("  Agreement:       %.1f%%\n",
    100 * sg_compare$concordance$agreement))
cat(sprintf("  Cohen's kappa:   %.3f\n",
    sg_compare$concordance$kappa))
cat(sprintf("  Concordant:      %d / Discordant: %d / Total: %d\n",
    sg_compare$concordance$n_agree,
    sg_compare$concordance$n_disagree,
    sg_compare$concordance$n_total))
```


# Bootstrap Bias Correction

## Rationale

Cox model estimates from identified subgroups are **upwardly biased** due to
the selection process (subgroups are selected *because* they show extreme effects).
Bootstrap bias correction addresses this by:

1. Resampling with replacement
2. Re-running the entire ForestSearch algorithm
3. Computing bias terms from bootstrap vs. observed estimates
4. Applying infinitesimal jackknife variance estimation

## Running Bootstrap Analysis

```{r bootstrap}

t0 <- proc.time()

fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est = fs,
  nb_boots = NB,
  show_three = FALSE,
  details = FALSE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nBootstrap completed in",
    round(timings$bootstrap / 60, 1), "minutes\n")
```

## Bootstrap Summary and Diagnostics

```{r bootstrap-summary}
# Comprehensive summary with diagnostics
summaries <- summarize_bootstrap_results(
  sgharm = fs$sg.harm,
  boot_results = fs_bc,
  create_plots = TRUE,
  est.scale = "hr"
)

# Display bias-corrected estimates table
summaries$table
```

## Kaplan-Meier by Identified Subgroups

```{r fs-results-figure}
#| fig-width: 10
#| fig-height: 7
#| fig-cap: "Kaplan-Meier survival curves by identified subgroup"

km_result <- plot_sg_weighted_km(
  fs.est = fs,
  outcome.name = "time_months",
  event.name = "status",
  treat.name = "hormon",
  show.logrank = FALSE,
  conf.int = TRUE,
  by.risk = 12,
  show.cox = FALSE, show.cox.bc = TRUE,
  fs_bc = fs_bc,
  hr_bc_position = "topright"
)
```

::: {.callout-note}
`r figure_note(km_result, prefix = "**Note:** ", custom_text = "Medians [95% CI] for arms are un-adjusted.")`
:::


### Event Count Summary

Low event counts can lead to unstable HR estimates. This summary helps
identify potential issues:

```{r event-summary}
event_summary <- summarize_bootstrap_events(fs_bc, threshold = 10)
```

### Bootstrap Diagnostics

```{r bootstrap-diagnostics}
# Quality metrics
summaries$diagnostics_table_gt
```

### Subgroup Agreement

How consistently does bootstrap identify the same subgroup?

```{r subgroup-agreement}
# Agreement with original analysis
if (!is.null(summaries$subgroup_summary$original_agreement)) {
  summaries$subgroup_summary$original_agreement
}

# Factor presence across bootstrap iterations
if (!is.null(summaries$subgroup_summary$factor_presence)) {
  summaries$subgroup_summary$factor_presence
}
```

### GRF-Cut Stability Across Bootstraps

Separately from whether bootstrap identifies the *same subgroup*, we can
ask what raw cut GRF itself surfaced on each resample, independent of
whether the downstream consistency stage accepted it.  A dominant row
here paired with a low `(no GRF cut)` rate indicates that GRF's
upstream discovery is stable across resamples; a frequent
`(no GRF cut)` row indicates GRF itself is the source of
variability.  This distinction matters because the two failure modes
call for different tuning responses — relaxing `dmin.grf` /
`frac.tau` (for "GRF produced no cut") vs. relaxing
`pconsistency.threshold` (for "GRF produced a cut that consistency
rejected").

```{r bootstrap-grf-cuts}
if (!is.null(summaries$subgroup_summary$grf_cut_freq)) {
  summaries$subgroup_summary$grf_cut_freq |>
    utils::head(10) |>
    gt::gt() |>
    gt::tab_header(
      title = gt::md("**GRF cuts across bootstrap iterations**"),
      subtitle = "Top 10 by frequency; spans all bootstraps (not just successful ones)"
    ) |>
    gt::cols_label(
      Rank    = "Rank",
      grf_cut = "GRF cuts",
      N       = "N",
      Percent = "% of bootstraps"
    ) |>
    gt::fmt_number(columns = "Percent", decimals = 2) |>
    gt::tab_options(table.font.size = gt::px(13))
}
```

### Bootstrap Distributions

```{r bootstrap-plots}
#| fig-width: 10
#| fig-height: 4

if (!is.null(summaries$plots)) {
  summaries$plots$H_distribution + summaries$plots$Hc_distribution
}
```


# Cross-Validation

Cross-validation assesses the **stability** of the ForestSearch algorithm.
Two approaches are available:

## K-Fold Cross-Validation

```{r cv-kfold-skip-note, eval = !run_cv_kfold, results = "asis"}
cat("> **K-fold CV skipped.** Set `run_cv_kfold <- TRUE` in the Analysis Configuration to enable. The forest plot below will omit kfold CV metrics.\n")
```

```{r kfold-cv, eval = run_cv_kfold}

t0 <- proc.time()

fs_kfold <- forestsearch_tenfold(
  fs.est = fs,
  sims = Ksims,
  Kfolds = 10,
  details = FALSE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = FALSE
  )
)

plan("sequential")
timings$kfold <- (proc.time() - t0)["elapsed"]
metrics_tables <- cv_metrics_tables(fs_kfold)
metrics_tables
```

### Diagnostic Tabulations

`summarize_cv_results()` consolidates the per-fold record into a panel
of diagnostic tables that together answer the core questions an
analyst asks of a repeated K-fold run: which subgroups did the
pipeline identify across resamples, which GRF cuts drove those
identifications, when identification failed what was the proximate
cause, and how confidently were the successful identifications made.

```{r cv-diagnostics-run, eval = run_cv_kfold}
cv_diag <- summarize_cv_results(
  cv_output         = fs_kfold,
  original_sg       = fs$sg.harm,
  original_grf_cuts = fs$grf_cuts,
  top_n             = 10L
)
```

**Which subgroups emerged across resamples?**  The identification
summary shows the distribution of selected subgroups across all
`r paste0(Ksims, " × 10 = ", Ksims * 10)` sim × fold pairs.  A
single dominant subgroup with high share indicates the pipeline
converges on a stable selection; a long tail suggests selection is
sensitive to the particular training split.

```{r cv-identification, eval = run_cv_kfold}
cv_diag$identification_summary
```

**Which GRF cuts drove those selections?**  The GRF cut summary
tabulates the raw policy-tree output on each training fold,
independent of whether the downstream consistency stage accepted
it.  Comparison against the identification summary reveals whether
GRF's upstream discovery is stable (a dominant cut shared across
folds) or noisy (a long tail of one-off cuts).

```{r cv-grf-cut, eval = run_cv_kfold}
cv_diag$grf_cut_summary
```

**When a fold failed to identify a subgroup, why?**  The no-subgroup
decomposition splits the non-identifying folds into two actionable
buckets: *GRF returned no cut* (nothing for consistency to evaluate,
typically pointing at `dmin.grf` or `frac.tau` thresholds) versus
*GRF returned a cut but consistency rejected it* (candidates existed
but none met `pconsistency.threshold`).  The former argues for
relaxing the GRF upstream; the latter for relaxing the consistency
threshold or reviewing candidate quality.

```{r cv-no-subgroup, eval = run_cv_kfold}
cv_diag$no_subgroup_decomposition
```

**How confidently were the successful identifications made?**  Among
folds that *did* identify a subgroup, the `pconsistency_distribution`
shows where the selected candidate's consistency probability
(`Pcons`) landed relative to the 0.90 threshold.  Identifications in
the `[0.85, 0.95)` bin are marginal; those concentrated at or near
1.0 are robust.

```{r cv-pcons, eval = run_cv_kfold}
cv_diag$pconsistency_distribution
```

**How do all the numeric fold-level diagnostics summarize together?**
The `fold_numeric_summary` tabulates median / IQR / range for each
numeric column in the per-fold record: fold size (`n_test`),
consistency probability (`pconsistency`), training-fold HR of the
selected subgroup (`training_fs_hr`, optimistically biased — for
diagnostic comparison only), and candidate count
(`n_candidates_evaluated`).

```{r cv-fold-numeric, eval = run_cv_kfold}
cv_diag$fold_numeric_summary
```

**Agreement with the full-data analysis.**  How often does a training
fold reproduce the original full-data subgroup exactly, share a
factor with it, or match its GRF cut?

```{r cv-original-agreement, eval = run_cv_kfold}
cv_diag$original_agreement
```


## Out-of-Bag (N-Fold) Cross-Validation

N-fold CV (leave-one-out):

```{r cv-oob-skip-note, eval = !run_cv_oob, results = "asis"}
cat("> **OOB (leave-one-out) CV skipped.** Set `run_cv_oob <- TRUE` in the Analysis Configuration to enable. This is the most expensive component of the document (~20-30 min on GBSG).\n")
```

```{r oob-cv, eval = run_cv_oob}

t0 <- proc.time()

fs_OOB <- forestsearch_Kfold(
  fs.est = fs,
  details = FALSE,
  Kfolds = nrow(df.analysis),  # N-fold = leave-one-out
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$oob <- (proc.time() - t0)["elapsed"]

# Summarize OOB results
cv_out <- forestsearch_KfoldOut(
  res = fs_OOB,
  details = FALSE,
  outall = TRUE
)

tables <- cv_summary_tables(cv_out)

tables$combined_table

tables$metrics_table
```


# Results Visualization

## Forest Plot

The forest plot summarizes treatment effects across the ITT population,
reference subgroups, and identified subgroups with cross-validation metrics.

```{r forest-plot}
#| fig-width: 18
#| fig-height: 12
#| fig-cap: "Subgroup forest plot including identified subgroups"

# Define reference subgroups for comparison
subgroups <- list(
  age_gt65 = list(
    subset_expr = "age > 65",
    name = "Age > 65",
    type = "reference"
  ),
  age_le65 = list(
    subset_expr = "age <= 65",
    name = "Age <= 65",
    type = "reference"
  ),
  pgr_positive = list(
    subset_expr = "pgr > 0",
    name = "PgR > 0",
    type = "reference"
  ),
  pgr_negative = list(
    subset_expr = "pgr <= 0",
    name = "PgR <= 0",
    type = "reference"
  )
)


my_theme <- create_forest_theme(base_size = 24,
  footnote_fontsize = 17, cv_fontsize = 22)


# Create forest plot.  fs_kfold / fs_OOB enter only when the
# corresponding CV step ran; otherwise NULL omits them from the
# plot's CV metrics columns.
result <- plot_subgroup_results_forestplot(
  fs_results = list(
    fs.est   = fs,
    fs_bc    = fs_bc,
    fs_OOB   = if (run_cv_oob)   fs_OOB   else NULL,
    fs_kfold = if (run_cv_kfold) fs_kfold else NULL
  ),
  df_analysis = df.analysis,
  subgroup_list = subgroups,
  outcome.name = outcome.name,
  event.name = event.name,
  treat.name = treat.name,
  E.name = "Hormonal",
  C.name = "Chemo",
  ci_column_spaces = 25,
  xlog = TRUE,
  theme = my_theme
)


# Render forest plot
render_forestplot(result)
```


### KM Difference Plots: ITT and Subgroups

The solid black line denotes the ITT Kaplan-Meier treatment difference
estimates along with 95% CIs (the grey shaded region). K-M differences
corresponding to subgroups are displayed.

```{r KMdiffs}
#| fig-width: 8
#| fig-height: 6

# Add additional subgroups along with ITT and identified subgroups
ref_sgs <- list(
  age_young = list(subset_expr = "age < 65", color = "brown"),
  age_old = list(subset_expr = "age >= 65", color = "orange")
)

plot_km_band_forestsearch(
  df = df.analysis,
  fs.est = fs,
  ref_subgroups = ref_sgs,
  outcome.name = outcome.name,
  event.name = event.name,
  treat.name = treat.name,
  draws_band = 20
)
```


# Summary and Interpretation

## Key Findings

```{r summary-findings}
cat(strrep("=", 60), "\n")
cat("FORESTSEARCH ANALYSIS SUMMARY\n")
cat(strrep("=", 60), "\n\n")

cat("Dataset: GBSG (N =", nrow(df.analysis), ")\n")
cat("Outcome: Recurrence-free survival\n\n")

cat("ITT Analysis:\n")
cat("  HR (95% CI): 0.69 (0.54, 0.89)\n\n")

cat("Identified Subgroup (H):\n")
cat("  Definition:", paste(fs$sg.harm, collapse = " & "), "\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 0)))
cat("  Unadjusted HR:",
    sprintf("%.2f", exp(fs$grp.consistency$out_sg$result$hr[1])), "\n")

cat("\nComplement Subgroup (Hc):\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 1),
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 1)))
```

## Clinical Interpretation

The ForestSearch analysis identifies **estrogen receptor-negative (ER ≤ 0)**
patients as a subgroup with potential lack of benefit from hormonal therapy.

**Biological plausibility**: Tamoxifen is a selective estrogen receptor
modulator. Its efficacy depends on ER expression. The finding that ER-negative
patients may not benefit is consistent with:

- Mechanistic understanding of tamoxifen action
- Meta-analyses showing no tamoxifen benefit in ER-negative breast cancer
- Clinical guidelines recommending tamoxifen primarily for ER-positive tumors

**Caveats**:

1. This is an **exploratory** analysis requiring independent validation
2. The bias-corrected estimates have wider confidence intervals
3. Cross-validation metrics should be evaluated for algorithm stability

## Computational Timing

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Analysis = c("GRF", "ForestSearch", "Bootstrap", "kfold", "oob", "Total"),
  Seconds = c(
    timings$grf,
    timings$forestsearch,
    timings$bootstrap,
    timings$kfold,
    timings$oob,
    timings$total
  )
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(title = "Computational Timing") |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1) |>
  cols_label(
    Analysis = "Component",
    Seconds = "Time (sec)",
    Minutes = "Time (min)"
  )
```


# References

León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024).
"Exploratory subgroup identification in the heterogeneous Cox model:
A relatively simple procedure." *Statistics in Medicine*.
DOI: [10.1002/sim.10163](https://doi.org/10.1002/sim.10163)


# Session Information

```{r session-info}
sessionInfo()
```
