---
title: "ACTG175 (Binary): Comparing Selection Rules"
subtitle: "OR-scale comparison across sg_focus × selection_rule cells"
format: html
---

```{r setup}
#| code-fold: true
#| code-summary: "Setup (click to expand)"

# Initialize timing
t_doc_start <- proc.time()
timings <- list()
```


# Introduction

This vignette compares the selection diagnostic side-by-side across two
choices of `selection_rule`, holding all other ForestSearch parameters
fixed.  It uses the package's
[`compare_selection_rules()`](../R/compare_selection_rules.R) wrapper to
run `forestsearch()` once per combination and bundle the per-combo
diagnostics for inspection.

The structure is:

1. **Data setup + ITT** — identical to the single-rule analyses.
2. **Comparison run** — one call to `compare_selection_rules()` that
   handles the loop, captures each run's
   `CANDIDATE EVALUATION PREVIEW` + `CANDIDATE EVALUATION SUMMARY`
   banners, builds the Pareto-frontier plots, and computes split-derived
   CIs.
3. **Per-combo diagnostics** — replay of captured console output,
   individual frontier plot/table/narrative for each combination.
4. **Side-by-side composition** — patchwork-composed plot panel
   for direct visual comparison.

Bootstrap, cross-validation, and the combined forest plot are
**deliberately omitted** so the document stays focused on the
*selection mechanism* and the comparison reads as a tight diagnostic
of the rule axis alone.


# Setup

## Load Required Packages

```{r load-packages}
suppressPackageStartupMessages({
  library(forestsearch)
  library(speff2trial)       # ACTG175 dataset
  library(data.table)
  library(ggplot2)
  library(gt)
  library(future)
  library(patchwork)         # for side-by-side plot composition
})

# Source the comparison wrapper.  When this graduates into the package
# (copy compare_selection_rules.R into R/ and run devtools::document()),
# this source() line can be removed.
#source("compare_selection_rules.R")

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```


## Analysis Parameters

All ForestSearch parameters are defined here once and reused for both
combinations.  The two axes of variation are `sg_focus` and
`selection_rule`, which the wrapper handles directly.

```{r analysis-params}
#| code-fold: false

# ─── Parallel processing ────────────────────────────────────────────────
n_cores <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))

# ─── Reproducibility ────────────────────────────────────────────────────
analysis_seed <- 8316951L
set.seed(analysis_seed)

# ─── Confounders ───────────────────────────────────────────────────────
cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
bin_vars  <- c("hemo", "homo", "drugs", "race", "gender", "symptom")
confounders.name <- c(cont_vars, bin_vars)
id.name <- "id"

# ─── Variables (dual-scale framing) ────────────────────────────────────
treat.name         <- "treat"
adverse_outcome    <- "y_binary_neg"   # FS adverse-scale outcome
beneficial_outcome <- "y_binary_pos"   # GRF beneficial-scale outcome

# ─── ForestSearch parameters (binary, OR scale) ────────────────────────
fs_outcome_type    <- "binary"
fs_effect_measure  <- "OR"

# NOTE: fs_sg_focus and fs_selection_rule are NOT defined here.
# They live in the `comparison-axes` chunk below as the two axes the
# wrapper varies.  All OTHER ForestSearch parameters are fixed across
# the comparison.

# Thresholds on the OR scale (adverse scale: OR > c1 = harm)
fs_or_threshold    <- 1.25  # screening: OR >= 1.25 indicates harm
fs_or_consistency  <- 1.0   # consistency: OR >= 1.0 (nominal harm)

# Binary OR estimates near the threshold are noisier than continuous MD;
# slightly relaxed pconsistency (matches the single-data binary template).
fs_pconsistency    <- 0.80
fs_n_min           <- 60L
fs_d0_min          <- 10L
fs_d1_min          <- 10L
fs_splits          <- 500L

fs_max_subgroups   <- 30

fs_maxk            <- 2L
fs_use_lasso       <- TRUE
fs_use_grf         <- TRUE
fs_is_RCT          <- TRUE
fs_vi_grf_min      <- -0.2
fs_dmin_grf        <- 0
fs_cont_cutoff     <- 4L

# ─── Pareto-frontier diagnostic parameters ──────────────────────────────
# For OR (ratio-scale), the effect axis is auto-rendered on log scale
# by plot_pareto_frontier(); leave xlim = NULL for auto bounds.
frontier_params <- list(
  n_splits      = 1000L,
  seed          = 1L,
  xlim          = NULL,
  digits        = 2L
)
```


## Comparison Axes

The two axes varied by the wrapper.  Tuple semantics: paired
element-by-element.

::: {.callout-note}
**`sg_focus = "effMaxSG"` notation.**  The `"eff*"` names
(`"eff"`, `"effMaxSG"`, `"effMinSG"`) are GLM-natural aliases for
the legacy Cox-flavored names (`"hr"`, `"hrMaxSG"`, `"hrMinSG"`).
Both vocabularies produce identical results.  The `"eff*"` forms
read more naturally for continuous (MD), binary (OR/RR/RD), and
count (IRR) outcomes where there is no hazard ratio.  Old code
using `"hrMaxSG"` continues to work without changes.
:::

```{r comparison-axes}
sg_focus_vec       <- c("effMaxSG", "effMaxSG", "effMinSG", "effMinSG", "eff", "maxSG","minSG")
selection_rule_vec <- c("pareto",   "both", "pareto", "both", "neighborhood", "neighborhood", "neighborhood")

# Pretty labels used in plot titles, list names, and prose
combo_labels       <- paste(sg_focus_vec, selection_rule_vec, sep = " / ")
n_combos           <- length(sg_focus_vec)

cat("Combinations to be run:\n")
print(data.frame(
  combo          = seq_len(n_combos),
  sg_focus       = sg_focus_vec,
  selection_rule = selection_rule_vec,
  label          = combo_labels
))
```


# Data Preparation

## Treatment Assignment

```{r data-treat}
# ACTG175: arm 1 = ZDV+ddI (experimental), arm 3 = ddI (control)
actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$id <- seq_len(nrow(actg_df))

# Treatment: ZDV+ddI = 1, ddI = 0 (NO switching)
actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)
```


## Outcome Construction

```{r data-outcomes}
# Drop subjects missing the week-20 CD4 measurement
actg_df <- actg_df[!is.na(actg_df$cd420), ]

# Beneficial scale: CD4 improvement indicator (1 = improved, higher = better)
actg_df$y_binary_pos <- as.integer(actg_df$cd420 > actg_df$cd40)

# Adverse scale: NO improvement indicator (1 = did not improve, higher = worse)
actg_df$y_binary_neg <- 1L - actg_df$y_binary_pos

outcome_vars <- c("y_binary_neg", "y_binary_pos")
```


## Covariates

```{r data-covariates}
# Coerce binary indicators to factors (analyst does not know true cuts
# for continuous covariates; ForestSearch handles those internally)
for (v in bin_vars) actg_df[[v]] <- as.factor(actg_df[[v]])

cat("Continuous candidates:", paste(cont_vars, collapse = ", "), "\n")
cat("Binary candidates:    ", paste(bin_vars, collapse = ", "), "\n")
cat("Analysis N:", nrow(actg_df), "\n")
```


## Baseline Characteristics

```{r baseline-table}
create_summary_table(
  data             = actg_df,
  treat_var        = treat.name,
  table_title      = "ACTG175 Baseline Characteristics and Outcomes by Treatment Arm",
  vars_continuous  = c(outcome_vars, cont_vars),
  vars_categorical = bin_vars,
  font_size        = 12
)
```


## ITT Analysis

```{r itt-analysis}
# Adverse-scale ITT OR (y_binary_neg = no improvement; OR > 1 = treatment
# increases the odds of no improvement = harm).
itt_neg <- glm(y_binary_neg ~ treat, data = actg_df, family = binomial)

# Helper to print log-OR, OR, 95% CI, p-value
print_or <- function(fit, label) {
  beta  <- coef(fit)["treat"]
  se    <- summary(fit)$coefficients["treat", "Std. Error"]
  z     <- beta / se
  p     <- 2 * pnorm(-abs(z))
  or    <- exp(beta)
  ci_lo <- exp(beta - qnorm(0.975) * se)
  ci_hi <- exp(beta + qnorm(0.975) * se)
  cat(sprintf("=== ITT Odds Ratio, %s ===\n", label))
  cat(sprintf("  log-OR (treat): %7.4f  (SE %.4f)\n", beta, se))
  cat(sprintf("  OR:             %6.3f  (95%% CI: %.3f, %.3f)\n",
              or, ci_lo, ci_hi))
  cat(sprintf("  p-value:        %.4f\n", p))
}

print_or(itt_neg, "adverse scale (y_binary_neg)")
```


# Selection-Rule Comparison

## Running `compare_selection_rules()`

One call to the wrapper runs `forestsearch()` for each
`(sg_focus, selection_rule)` combination with all other parameters
held fixed, and bundles the per-combo diagnostics into a single
object.

```{r run-comparison}
#| message: false
#| warning: false

t0 <- proc.time()

comparison <- compare_selection_rules(
  df.analysis      = actg_df,
  sg_focus         = sg_focus_vec,
  selection_rule   = selection_rule_vec,
  combo_labels     = combo_labels,

  # Wrapper-specific controls
  compute_cis      = TRUE,
  n_splits         = frontier_params$n_splits,
  ci_seed          = frontier_params$seed,
  plot_xlim        = frontier_params$xlim,
  show_band        = TRUE,
  verbose          = TRUE,

  # All other ForestSearch arguments — pass-through via ... .
  # These match the single-data binary analysis 1-for-1 except for
  # sg_focus and selection_rule (which the wrapper supplies).
  confounders.name = confounders.name,
  outcome.name     = adverse_outcome,
  event.name       = adverse_outcome,   # binary: outcome IS the event
  treat.name       = treat.name,
  id.name          = id.name,
  outcome_type     = fs_outcome_type,
  effect_measure   = fs_effect_measure,
  adverse_outcome  = TRUE,
  seedit           = analysis_seed,
  effect.threshold       = fs_or_threshold,
  consistency.threshold  = fs_or_consistency,
  pconsistency.threshold = fs_pconsistency,
  max_subgroups_search = fs_max_subgroups,
  use_twostage     = TRUE,
  conf.cont_jcuts  = list(cd40 = 10, wtkg = 10),
  use_grf          = fs_use_grf,
  return_selected_cuts_only = TRUE,
  use_lasso        = fs_use_lasso,
  cut_type         = "default",
  maxk             = fs_maxk,
  n.min            = fs_n_min,
  d0.min           = fs_d0_min,
  d1.min           = fs_d1_min,
  fs.splits        = fs_splits,
  parallel_args    = list(
    plan         = "multisession",
    workers      = n_cores,
    show_message = TRUE
  ),
  show_candidate_summary = TRUE,
  details          = TRUE,
  plot.sg          = FALSE,
  is.RCT           = fs_is_RCT,
  vi.grf.min       = fs_vi_grf_min,
  dmin.grf         = fs_dmin_grf,
  cont.cutoff      = fs_cont_cutoff,
  tune_grf         = FALSE,
  quiet            = FALSE
)

plan("sequential"); invisible(gc(verbose = FALSE))
timings$compare <- (proc.time() - t0)["elapsed"]

cat("\nComparison completed in",
    round(timings$compare / 60, 1), "minutes\n")
```


## Comparison Summary

```{r comparison-summary}
print(comparison)
```


# Per-Combo Diagnostics

## Captured Console Output

The wrapper captured the in-flight diagnostic output from each
`forestsearch()` call.  Each combination's two diagnostic banners --
the `CANDIDATE EVALUATION PREVIEW` (pre-consistency; all candidates
entering evaluation) and the `CANDIDATE EVALUATION SUMMARY`
(post-consistency; passing candidates with
Pcons / Frontier / InBand / Selected flags) -- are shown below as the
primary view.  The full captured output (GRF preview, Cox-LASSO
selection, filtering summary, parallel-config messages, post-run
timings, etc.) is folded into a collapsible block for readers who want
the complete context.

```{r console-loop, results = "asis"}
# Loop over all combos.  This replaces the previous two hardcoded
# sub-sections so the document handles n_combos > 2 correctly
# (e.g. effMaxSG/pareto, effMaxSG/both, effMinSG/pareto, effMinSG/both).
for (i in seq_len(n_combos)) {
  cat(sprintf("\n### %s\n\n", combo_labels[i]))

  diag <- comparison$diagnostics[[i]]

  cat("**Pre-consistency preview:**\n\n")
  if (!is.na(diag$preview)) {
    cat("```\n", diag$preview, "\n```\n\n", sep = "")
  } else {
    cat("_(not available)_\n\n")
  }

  cat("**Post-consistency summary:**\n\n")
  if (!is.na(diag$summary)) {
    cat("```\n", diag$summary, "\n```\n\n", sep = "")
  } else {
    cat("_(not available)_\n\n")
  }

  cat("<details><summary>Full forestsearch output (expand)</summary>\n\n```\n")
  cat(comparison$console[[i]], sep = "\n")
  cat("\n```\n\n</details>\n\n")
}
```


## Pareto-Frontier Plots

Each combination produces its own Pareto frontier on (OR, N).  The
plot for each combo is shown individually below, then composed
side-by-side via `patchwork` for direct visual comparison.  For OR
(ratio-scale), the effect axis is auto-rendered on log scale by
`plot_pareto_frontier()`.

### Individual Panels

```{r plot-combo-loop, results = "asis"}
#| fig-width: 8
#| fig-height: 5

# One panel per combo (replaces the previous two hardcoded chunks so the
# document handles n_combos > 2 correctly).
for (i in seq_len(n_combos)) {
  cat(sprintf("\n#### %s\n\n", combo_labels[i]))
  print(comparison$plots[[i]])
  cat("\n\n")
}
```


### Side-by-Side Composition

```{r plot-side-by-side}
#| fig-width: 14
#| fig-height: 6
#| fig-cap: "Pareto frontier across selection rules (side-by-side)"

print(comparison$plot_grid)
```


## Combined Plot(s) by Matching Passing-Set Group

`compare_selection_rules()` groups configurations by their
post-consistency passing set.  When all configurations share one
passing set, a single combined plot is built (`comparison$plot_combined`).
When configurations fall into two or more matching groups -- e.g. all
`effMaxSG` runs agree, all `effMinSG` runs agree, but the two
`sg_focus` values disagree on which candidates pass -- a combined plot
is built per group and stored in `comparison$plot_combined_subsets`.
Each group's plot labels its selected subgroups with `S<i>:` markers
identifying the configuration(s) that picked them.

```{r plot-combined}
#| fig-width: 10
#| fig-height: 6
#| fig-cap: "Combined Pareto frontier with S1/S2 winner annotations"

if (!is.null(comparison$plot_combined)) {
  cat("All configurations share one passing set; single combined plot:\n\n")
  print(comparison$plot_combined)
} else if (length(comparison$plot_combined_subsets) > 0L) {
  cat(sprintf("Passing sets fall into %d matching group(s); one combined plot per group:\n\n",
              length(comparison$plot_combined_subsets)))
  for (nm in names(comparison$plot_combined_subsets)) {
    cat(sprintf("Group: %s\n", nm))
    print(comparison$plot_combined_subsets[[nm]])
  }
  if (!is.null(comparison$combined_skip_reason)) {
    cat("\nNote: ", comparison$combined_skip_reason, "\n", sep = "")
  }
} else {
  cat("**No combined plot.**\n\n")
  if (!is.null(comparison$combined_skip_reason)) {
    cat("Reason:  ", comparison$combined_skip_reason, "\n\n", sep = "")
  }
  cat("See the side-by-side composition above as the alternative view.\n")
}
```


## Frontier Tables

Per-combination Pareto-frontier table with split-derived and FSBC
CIs.  Selected subgroup is highlighted with ★.

```{r table-combo-loop, results = "asis"}
for (i in seq_len(n_combos)) {
  cat(sprintf("\n### `%s`\n\n", combo_labels[i]))
  tbl <- pareto_frontier_table(
    comparison$fs[[i]],
    ci_table = comparison$ci_tab[[i]],
    digits   = frontier_params$digits
  )
  print(tbl)
  cat("\n\n")
}
```


## Selection Narratives

`explain_pareto_selection()` produces a verbal account of why the
selected subgroup wins on the configured lexicographic criterion
relative to the other non-dominated candidates.  Reading the two
narratives together highlights which axis of the lex key the two
rules tie-broke on differently.

```{r explain-loop, results = "asis"}
for (i in seq_len(n_combos)) {
  cat(sprintf("\n### `%s`\n\n", combo_labels[i]))
  cat(explain_pareto_selection(
    comparison$fs[[i]],
    ci_table = comparison$ci_tab[[i]],
    digits   = frontier_params$digits
  ))
  cat("\n\n")
}
```


# Rich Post-Consistency View

The `include_dominated = TRUE` argument to `pareto_frontier_table()`
shows **all** passing candidates (not just the frontier), with
`on_frontier` and `in_band` flag columns added.  This is the
post-hoc analog of the `CANDIDATE EVALUATION SUMMARY` banner printed
during the run.

```{r dominated-loop, results = "asis"}
for (i in seq_len(n_combos)) {
  cat(sprintf("\n### `%s`\n\n", combo_labels[i]))
  tbl <- pareto_frontier_table(
    comparison$fs[[i]],
    ci_table          = comparison$ci_tab[[i]],
    include_dominated = TRUE,
    digits            = frontier_params$digits
  )
  print(tbl)
  cat("\n\n")
}
```


# Programmatic Access (Reference)

```{r programmatic, eval = FALSE}
# Each combination's full fs object is available for downstream use.
# Index with [[i]] where i runs from 1 to n_combos; see comparison$combos
# for the (sg_focus, selection_rule, label) tuple at each index.
comparison$combos                    # data.frame: per-combo configuration
comparison$fs[[1]]                   # fs object for combo 1

# CI tables (data.tables, one per combo):
comparison$ci_tab[[1]]

# Plot objects (ggplot, one per combo):
comparison$plots[[1]]

# Side-by-side panel (single patchwork object):
comparison$plot_grid

# Combined plot (when passing sets match), and the reason if skipped:
comparison$plot_combined             # ggplot or NULL (all combos match)
comparison$plot_combined_subsets     # named list of ggplots, one per
                                     # matching subset of size >= 2
comparison$combined_skip_reason      # character or NULL

# Console output (full captured stdout, one character per line):
comparison$console[[1]]

# Diagnostics (filtered slices: preview, summary, full):
comparison$diagnostics[[1]]$preview  # PREVIEW block as single string
comparison$diagnostics[[1]]$summary  # SUMMARY block as single string
comparison$diagnostics[[1]]$full     # entire captured stdout (single string)

# Errors (NULL if combo succeeded):
comparison$errors[[1]]
```


# Computational Timing

```{r timing-summary}
timings$total <- (proc.time() - t_doc_start)["elapsed"]

timing_df <- data.frame(
  Component = c(sprintf("Comparison run (%d combos)", n_combos),
                "Total document"),
  Seconds   = c(timings$compare, timings$total)
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(title = "Computational Timing") |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1) |>
  cols_label(
    Component = "Component",
    Seconds   = "Time (sec)",
    Minutes   = "Time (min)"
  )
```


# References

León LA, Ye J, Zhao H. *A Bootstrap Framework for Subgroup
Identification in Randomized Clinical Trials with Bias Correction.*
Statistics in Medicine. 2024.  [doi:10.1002/sim.10163](https://doi.org/10.1002/sim.10163)
