---
title: "Exploratory Subgroup Identification with ForestSearch"
subtitle: "German Breast Cancer Study Group (GBSG) Analysis: Cox and Poisson Rate Model Comparison"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 8
    fig-height: 6
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# Initialize timing
timings <- list()
t_vignette_start <- proc.time()
```


# Introduction

This vignette demonstrates the **ForestSearch** methodology for exploratory
subgroup identification in survival analysis, as described in
León et al. (2024) *Statistics in Medicine*.

## Motivation

In clinical trials, particularly oncology, subgroup analyses are essential for:

- Evaluating treatment effect consistency across patient populations
- Identifying subgroups where treatment may be detrimental (harm)
- Characterizing subgroups with enhanced benefit
- Informing regulatory decisions and clinical practice

While prespecified subgroups provide stronger evidence, important subgroups
based on patient characteristics may not be anticipated. ForestSearch provides
a principled approach to **exploratory** subgroup identification with proper
statistical inference.

## Methodology Overview

ForestSearch identifies subgroups through:

1. **Candidate factor selection**: Using LASSO and/or Generalized Random Forests (GRF)
2. **Exhaustive subgroup search**: Evaluating all combinations up to `maxk` factors
3. **Consistency-based selection**: Applying splitting consistency criteria
4. **Bootstrap bias correction**: Adjusting for selection-induced optimism
5. **Cross-validation**: Assessing algorithm stability

The key innovation is the **splitting consistency criterion**: a subgroup is
considered "consistent with harm" if, when randomly split 50/50 many times,
both halves consistently show hazard ratios ≥ 1.0 (for example if 1.0
represents a meaningful "harm threshold").


# Setup

## Load Required Packages

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(policytree)
library(doFuture)

# Optional packages for enhanced output
library(patchwork)
library(weightedsurv)

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```

## Analysis Parameters

```{r analysis-params}
#| code-fold: false

# ─── Shared parameters ───────────────────────────────────────────────────

n_cores   <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))

confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
id.name    <- "id"

# ─── Cox parameters ──────────────────────────────────────────────────────
cox_outcome.name <- "time_months"
cox_event.name   <- "status"
cox_treat.name   <- "hormon"

fs_sg_focus <- "maxSG"

# ─── ForestSearch tuning (shared across Cox and Poisson) ─────────────────
fs_hr.threshold            <- 1.25
fs_hr.consistency          <- 1.0
fs_pconsistency.threshold  <- 0.80
fs_stop_threshold          <- 0.90
fs_maxk                    <- 2
fs_n.min                   <- 60
fs_d0.min                  <- 10
fs_d1.min                  <- 10
fs_splits                  <- 1000
fs_max_subgroups_search    <- 15

# ─── Bootstrap / CV ──────────────────────────────────────────────────────
NB    <- 500    # Bootstrap iterations (use 500–2000 for production)
# ─── GRF parameters (shared across Cox and Poisson) ─────────────────────
grf_tune           <- FALSE  # cross-validated GRF hyperparameter tuning (off by default)
grf_vi_min         <- -0.2   # GRF variable importance threshold
grf_dmin_cox       <- 4     # min RMST difference for Cox GRF (months)
grf_dmin_pois      <- 0.0    # min CATE difference for Poisson GRF
grf_maxdepth       <- 2L     # max policy tree depth

# ─── Poisson-specific parameters ─────────────────────────────────────────
pois_outcome.name   <- "status"         # event indicator (0/1 count)
pois_offset.name    <- "time_months"    # person-time offset
pois_treat.name     <- "hormon"
pois_effect_measure <- "IRR"
```


# Data: German Breast Cancer Study Group Trial

## Study Background

The GBSG trial evaluated hormonal treatment (tamoxifen) versus chemotherapy
in node-positive breast cancer patients. Key characteristics:

- **Sample size**: N = 686
- **Outcome**: Recurrence-free survival time
- **Censoring rate**: ~56%
- **Treatment**: Hormonal therapy (tamoxifen) vs. chemotherapy

## Data Preparation

```{r data-setup}
# Load GBSG data (included in forestsearch package)
df.analysis <- gbsg

# Prepare analysis variables
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
  treat <- hormon
})

# Display data structure
cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[cox_event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[cox_event.name]])))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```

## Baseline Characteristics

```{r baseline-table}
create_summary_table(
  data = df.analysis,
  treat_var = cox_treat.name,
  table_title = "GBSG Baseline Characteristics by Treatment Arm",
  vars_continuous = c("age", "nodes", "size", "er", "pgr"),
  vars_categorical = c("grade", "meno"),
  font_size = 12
)
```

## Kaplan-Meier Analysis (ITT Population)

```{r km-itt}
#| fig-width: 8
#| fig-height: 5

# Prepare counting process data for KM plot
dfcount <- df_counting(
  df = df.analysis,
  by.risk = 6,
  tte.name = cox_outcome.name,
  event.name = cox_event.name,
  treat.name = cox_treat.name
)

# Plot with confidence intervals and log-rank test
plot_weighted_km(
  dfcount,
  conf.int = TRUE,
  show.logrank = TRUE,
  ymax = 1.05,
  xmed.fraction = 0.775,
  ymed.offset = 0.125
)
```

The ITT Cox hazard ratio estimate is approximately 0.69 (95% CI: 0.54, 0.89),
suggesting an overall benefit for hormonal therapy.


# Preliminary Analysis: Generalized Random Forests

Before running ForestSearch, we can use GRF to explore potential treatment
effect heterogeneity and identify candidate factors.

```{r grf-analysis}
t0 <- proc.time()

grf_est <- grf.subg.harm.survival(
  data = df.analysis,
  confounders.name = confounders.name,
  outcome.name = cox_outcome.name,
  event.name = cox_event.name,
  id.name = id.name,
  treat.name = cox_treat.name,
  maxdepth = grf_maxdepth,
  n.min = fs_n.min,
  dmin.grf = grf_dmin_cox,
  frac.tau = 0.6,
  details = TRUE,
  return_selected_cuts_only = FALSE,
  tune_grf = grf_tune
)

timings$grf <- (proc.time() - t0)["elapsed"]
```

```{r grf-trees}
#| fig-width: 10
#| fig-height: 4

# Display policy trees
# leaf1 = recommend control, leaf2 = recommend treatment
oldpar <- par(mfrow = c(1, 2))
plot(grf_est$tree1, leaf.labels = c("Control", "Treat"), main = "Depth 1")
plot(grf_est$tree2, leaf.labels = c("Control", "Treat"), main = "Depth 2")
par(oldpar)
```

GRF identifies estrogen receptor status (ER) as a key factor, with ER ≤ 0
suggesting potential harm from hormonal therapy.


# Part I: Cox Proportional Hazards Analysis {#sec-cox}


## Parallel Processing Configuration

ForestSearch supports parallel processing for computationally intensive
operations (bootstrap, cross-validation).


## Running ForestSearch (Cox)

ForestSearch performs an exhaustive search over candidate subgroup combinations
with up to `maxk` factors. Key parameters:

| Parameter | Value | Description |
|-----------|-------|-------------|
| `hr.threshold` | 1.25 | Minimum HR for consistency evaluation |
| `hr.consistency` | 1.0 | Minimum consistency rate for candidates |
| `pconsistency.threshold` | 0.90 | Required consistency for selection |
| `maxk` | 2 | Maximum factors in subgroup definition |
| `n.min` | 60 | Minimum subgroup sample size |
| `d0.min`, `d1.min` | 12 | Minimum events per treatment arm |

```{r forestsearch-cox}
#| fig-width: 10
#| fig-height: 7

t0 <- proc.time()

fs_cox <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  outcome.name = cox_outcome.name,
  treat.name = cox_treat.name,
  event.name = cox_event.name,
  id.name = id.name,
  # Threshold parameters (per León et al. 2024)
  hr.threshold = fs_hr.threshold,
  hr.consistency = fs_hr.consistency,
  pconsistency.threshold = fs_pconsistency.threshold,
  stop_threshold = fs_stop_threshold,
  # Search configuration
  sg_focus = fs_sg_focus,
  max_subgroups_search = fs_max_subgroups_search,
  use_twostage = TRUE,
  # Factor selection
  use_grf = TRUE,
  return_selected_cuts_only = TRUE,
  use_lasso = TRUE,
  cut_type = "default",
  # Subgroup constraints
  maxk = fs_maxk,
  n.min = fs_n.min,
  d0.min = fs_d0.min,
  d1.min = fs_d1.min,
  # Consistency evaluation
  fs.splits = fs_splits,
  # Parallel processing
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  ),
  # Output options
  showten_subgroups = TRUE,
  details = TRUE,
  plot.sg = TRUE,
  vi.grf.min = grf_vi_min,
  tune_grf = grf_tune
)

plan("sequential")
timings$cox_fs <- (proc.time() - t0)["elapsed"]

cat("\nForestSearch (Cox) completed in",
    round(timings$cox_fs, 1), "seconds\n")
```



## Cox Results

### Identified Subgroups

```{r cox-results}
# Generate results tables
cox_tabs <- sg_tables(fs_cox, ndecimals = 3, which_df = "est")

# Display top subgroups meeting criteria
cox_tabs$sg10_out
```

### Treatment Effect Estimates

```{r cox-estimates}
cox_tabs$tab_estimates
```

### Identified Subgroup Definition

```{r cox-subgroup}
cat("Identified subgroup (H):", paste(fs_cox$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs_cox$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs_cox$df.est$treat.recommend == 0)))
```

ForestSearch identifies **Estrogen ≤ 0** (ER-negative) as the subgroup with
potential harm. This is biologically plausible: tamoxifen is a selective
estrogen receptor modulator with limited efficacy in ER-negative tumors.


## Cox Bootstrap Bias Correction

```{r cox-bootstrap}
t0 <- proc.time()

fs_cox_bc <- forestsearch_bootstrap_dofuture(
  fs.est = fs_cox,
  nb_boots = NB,
  show_three = FALSE,
  details = FALSE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$cox_bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nBootstrap completed in",
    round(timings$cox_bootstrap / 60, 1), "minutes\n")
```

### Bootstrap Summary

```{r cox-bootstrap-summary}
cox_summaries <- summarize_bootstrap_results(
  sgharm = fs_cox$sg.harm,
  boot_results = fs_cox_bc,
  create_plots = TRUE,
  est.scale = "hr"
)

cox_summaries$table
```

### Kaplan-Meier by Identified Subgroups

```{r cox-km-subgroups}
#| fig-width: 10
#| fig-height: 7
#| fig-cap: "Kaplan-Meier survival curves by identified subgroup (Cox analysis)"

km_result_cox <- plot_sg_weighted_km(
  fs.est = fs_cox,
  outcome.name = cox_outcome.name,
  event.name = cox_event.name,
  treat.name = cox_treat.name,
  show.logrank = FALSE,
  conf.int = TRUE,
  by.risk = 12,
  show.cox = FALSE, show.cox.bc = TRUE,
  fs_bc = fs_cox_bc,
  hr_bc_position = "topright"
)
```

::: {.callout-note}
`r figure_note(km_result_cox, prefix = "**Note:** ", custom_text = "Medians [95% CI] for arms are un-adjusted.")`
:::


### Event Count Summary

```{r cox-event-summary}
event_summary_cox <- summarize_bootstrap_events(fs_cox_bc, threshold = 15)
```

### Bootstrap Diagnostics

```{r cox-bootstrap-diagnostics}
cox_summaries$diagnostics_table_gt
```

### Subgroup Agreement

```{r cox-subgroup-agreement}
if (!is.null(cox_summaries$subgroup_summary$original_agreement)) {
  cox_summaries$subgroup_summary$original_agreement
}

if (!is.null(cox_summaries$subgroup_summary$factor_presence)) {
  cox_summaries$subgroup_summary$factor_presence
}
```

### Bootstrap Distributions

```{r cox-bootstrap-plots}
#| fig-width: 10
#| fig-height: 4

if (!is.null(cox_summaries$plots)) {
  cox_summaries$plots$H_distribution + cox_summaries$plots$Hc_distribution
}
```



# Part II: Poisson Rate Model Analysis {#sec-poisson}

## Rationale

The Poisson regression model with an offset for person-time provides an
alternative to the Cox PH model for estimating treatment effects on the
event rate scale (incidence rate ratio, IRR). Under an exponential
survival assumption, the Poisson IRR is approximately equal to the Cox HR.

This section applies ForestSearch with `outcome_type = "count"` and
`effect_measure = "IRR"` to the GBSG data, treating each patient's
binary event indicator (`status`) as a count (0 or 1) and
`log(time_months)` as the offset. This is the standard piecewise
exponential model approximation with a single interval per patient.

Comparing the Poisson and Cox analyses serves two purposes:

1. **Validation**: If both models identify the same subgroup with similar
   effect sizes, this strengthens confidence in the findings.
2. **GLM pipeline verification**: Demonstrates that the ForestSearch GLM
   extension produces coherent results on real clinical trial data.


## ITT Poisson Model

```{r pois-itt}
# Verify the Poisson approximation at the ITT level
cox_itt <- coxph(
  Surv(time_months, status) ~ hormon,
  data = df.analysis
)
pois_itt <- glm(
  status ~ hormon,
  family = poisson(),
  offset = log(pmax(time_months, 1e-6)),
  data = df.analysis
)

cat("ITT Cox  HR:      ", sprintf("%.4f", exp(coef(cox_itt))), "\n")
cat("ITT Poisson IRR:  ", sprintf("%.4f", exp(coef(pois_itt)[["hormon"]])), "\n")
cat("Absolute diff:    ",
    sprintf("%.4f", abs(exp(coef(cox_itt)) - exp(coef(pois_itt)[["hormon"]]))),
    "\n")
```


## Running ForestSearch (Poisson)

The Poisson analysis uses `outcome_type = "count"` with `effect_measure = "IRR"`
and `offset.name = "time_months"`. Screening and consistency thresholds are
set identically to the Cox analysis so that any differences in results
reflect the model, not tuning.

```{r forestsearch-poisson}
#| fig-width: 10
#| fig-height: 7

t0 <- proc.time()

fs_pois <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  outcome.name = pois_outcome.name,
  treat.name = pois_treat.name,
  id.name = id.name,
  # GLM parameters (event.name is not needed for count outcomes;
  # forestsearch() handles this internally)
  outcome_type = "count",
  effect_measure = pois_effect_measure,
  offset.name = pois_offset.name,
  # Threshold parameters (same as Cox for comparability)
  hr.threshold = fs_hr.threshold,
  hr.consistency = fs_hr.consistency,
  pconsistency.threshold = fs_pconsistency.threshold,
  stop_threshold = fs_stop_threshold,
  # Search configuration
  sg_focus = fs_sg_focus,
  max_subgroups_search = fs_max_subgroups_search,
  use_twostage = TRUE,
  # Factor selection
  use_grf = TRUE,
  return_selected_cuts_only = TRUE,
  use_lasso = TRUE,
  cut_type = "default",
  # Subgroup constraints
  maxk = fs_maxk,
  n.min = fs_n.min,
  d0.min = fs_d0.min,
  d1.min = fs_d1.min,
  # Consistency evaluation
  fs.splits = fs_splits,
  # Parallel processing
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  ),
  # Output options
  showten_subgroups = TRUE,
  details = TRUE,
  plot.sg = FALSE,   # KM plots not applicable; use plot_sg_glm_outcomes()
  vi.grf.min = grf_vi_min,
  tune_grf = grf_tune
)

plan("sequential")
timings$pois_fs <- (proc.time() - t0)["elapsed"]

cat("\nForestSearch (Poisson) completed in",
    round(timings$pois_fs, 1), "seconds\n")
```


## Poisson Results

### Identified Subgroups

```{r pois-results}
pois_tabs <- sg_tables(fs_pois, ndecimals = 3, which_df = "est")
pois_tabs$sg10_out
```

### Treatment Effect Estimates

```{r pois-estimates}
pois_tabs$tab_estimates
```

### Identified Subgroup Definition

```{r pois-subgroup}
cat("Identified subgroup (H):",
    paste(fs_pois$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs_pois$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs_pois$df.est$treat.recommend == 0)))
```

### Subgroup Event Rates

```{r pois-glm-plot}
#| fig-width: 10
#| fig-height: 7
#| fig-cap: "Event rates by treatment arm and identified subgroup (Poisson analysis)"

pois_plot <- plot_sg_glm_outcomes(
  fs.est = fs_pois,
  outcome.name = pois_outcome.name,
  treat.name = pois_treat.name,
  offset.name = pois_offset.name,
  effect_measure = pois_effect_measure,
  outcome_type = "count",
  E.name = "Hormonal",
  C.name = "Chemo",
  show_itt = TRUE,
  show_effect = TRUE,
  verbose = TRUE
)
```


## Poisson Bootstrap Bias Correction

```{r pois-bootstrap}
t0 <- proc.time()

fs_pois_bc <- forestsearch_bootstrap_dofuture(
  fs.est = fs_pois,
  nb_boots = NB,
  show_three = FALSE,
  details = FALSE,
  parallel_args = list(
    plan = "multisession",
    workers = n_cores,
    show_message = TRUE
  )
)

plan("sequential")
timings$pois_bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nPoisson bootstrap completed in",
    round(timings$pois_bootstrap / 60, 1), "minutes\n")
```

### Bootstrap Summary

```{r pois-bootstrap-summary}
pois_summaries <- summarize_bootstrap_results(
  sgharm = fs_pois$sg.harm,
  boot_results = fs_pois_bc,
  create_plots = TRUE,
  est.scale = "hr"   # IRR is on the same ratio scale as HR
)

pois_summaries$table
```

### Bootstrap Diagnostics

```{r pois-bootstrap-diagnostics}
pois_summaries$diagnostics_table_gt
```


## Computational Timing

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Analysis = c(
    "GRF Screening",
    "ForestSearch (Cox)", "Bootstrap (Cox)",
    "ForestSearch (Poisson)", "Bootstrap (Poisson)",
    "Total"
  ),
  Seconds = c(
    timings$grf,
    timings$cox_fs, timings$cox_bootstrap,
    timings$pois_fs, timings$pois_bootstrap,
    timings$total
  )
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(title = "Computational Timing") |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1) |>
  cols_label(
    Analysis = "Component",
    Seconds = "Time (sec)",
    Minutes = "Time (min)"
  )
```


# References

León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024).
"Exploratory subgroup identification in the heterogeneous Cox model:
A relatively simple procedure." *Statistics in Medicine*.
DOI: [10.1002/sim.10163](https://doi.org/10.1002/sim.10163)


# Session Information

```{r session-info}
sessionInfo()
```
