---
title: "Exploratory Subgroup Identification with ForestSearch"
subtitle: "German Breast Cancer Study Group (GBSG) Analysis"
author: "Larry F. León"
date: today
format:
  html:
    code-fold: show
    toc: true
    toc-depth: 3
execute:
  warning: true
  message: true
  fig-align: center
knitr:
  opts_chunk:
    comment: "#>"
---


# Introduction {#sec-intro}

This vignette demonstrates the **ForestSearch** methodology for exploratory
subgroup identification in survival analysis, as described in
León et al. (2024) *Statistics in Medicine*.

## Motivation

In clinical trials, particularly oncology, subgroup analyses are essential for:

- Evaluating treatment effect consistency across patient populations
- Identifying subgroups where treatment may be detrimental (harm)
- Characterizing subgroups with enhanced benefit
- Informing regulatory decisions and clinical practice

While prespecified subgroups provide stronger evidence, important subgroups
based on patient characteristics may not be anticipated. ForestSearch provides
a principled approach to **exploratory** subgroup identification with proper
statistical inference.

## Methodology Overview

ForestSearch identifies subgroups through:

1. **Candidate factor selection**: Using LASSO and/or Generalized Random Forests (GRF)
2. **Exhaustive subgroup search**: Evaluating all combinations up to `maxk` factors
3. **Consistency-based selection**: Applying splitting consistency criteria
4. **Bootstrap bias correction**: Adjusting for selection-induced optimism
5. **Cross-validation**: Assessing algorithm stability

The key innovation is the **splitting consistency criterion**: a subgroup is
considered "consistent with harm" if, when randomly split 50/50 many times,
both halves consistently show hazard ratios ≥ 1.0 (where 1.0 represents a
meaningful "harm threshold").  An appendix (@sec-appendix-fpr) directly
calibrates the procedure-level false-positive rate (FPR) at the operational
thresholds used in the main analysis.

## Executive Overview

**Goal.**  This analysis applies ForestSearch to the GBSG breast cancer
trial (N = 686) to ask whether there exists a sizeable subgroup of
patients for whom hormonal therapy may *not* be beneficial — or may
even be detrimental.  The analysis is run with `sg_focus = "maxSG"`,
which targets the **largest** subgroup satisfying the harm criterion
rather than the smallest or the one with the most extreme hazard
ratio.  This framing is intentional: from a risk-management standpoint,
the more actionable question is "how many patients in the trial
population could reasonably be recommended *against* this therapy?"
rather than "what is the worst-affected subgroup we can find?".  A
larger identified subgroup, if genuine, translates into a larger
population for whom alternative therapies should be considered.

**Key findings.**  At the León (2024) operational thresholds
(c₁ = 1.25 screening, c₂ = 1.00 consistency) with
`pconsistency.threshold = 0.90` and `stop_threshold = 0.95`,
ForestSearch identifies **ER-negative patients (estrogen receptor
status ≤ 0)** as the largest subgroup passing the harm criterion —
82 patients, 12% of the ITT population.  The identified subgroup shows
an estimated hazard ratio of 1.95 (95% CI: 1.04, 3.67) against the
therapy, while the complement (604 patients, 88%) shows a clear
benefit of 0.61 (0.47, 0.80).  Overall ITT is 0.69 (0.54, 0.89).  The
finding is biologically plausible: tamoxifen is a selective estrogen
receptor modulator with limited efficacy in ER-negative disease, and
meta-analyses and clinical guidelines already recommend tamoxifen
primarily for ER-positive tumors.

**Validation.**  Three independent checks support the identification:
(i) bootstrap bias correction (1,000 iterations) adjusts for
selection-induced optimism in the subgroup effect estimates;
(ii) K-fold and out-of-bag cross-validation across 50 repetitions
confirm the ER factor is recovered with high frequency rather than
being an artifact of a particular split; (iii) the FPR calibration
appendix (@sec-appendix-fpr) directly simulates the null hypothesis
via `setup_gbsg_dgm(model = "null")` and finds a procedure-level FPR
of 6.0% (95% CI: 5.4%, 6.7%) from 5,000 calibration replicates, with
an independent verification batch at 7.0% (6.3%, 7.7%) on disjoint
seeds — below the 12% tolerance and well-aligned with the asymptotic
León (2024) equation-3 approximation (P1 = 19.7%, implied L\_eff =
0.283).  Theta-grid verification at θ†(H) ∈ {0.75, 0.9, 1.0, 1.25,
1.5, 2.0} further tests the transportability of the correction across
the null and alternative regions.

**Caveats.**  These findings are specific to the analysis specification
used here — in particular the `maxSG` search direction, the
(c₁, c₂) = (1.25, 1.00) operational thresholds, the `pconsistency`
and `stop_threshold` values, the LASSO + GRF screening combination,
and the `n.min = 60` / `d0.min = d1.min = 12` subgroup-size filters.
Different specifications would produce different subgroups, different
FPR calibration, and potentially different conclusions.  The ER ≤ 0
finding is stable across the validation checks under *these*
specifications; readers adapting this workflow to their own trials
should re-calibrate at their chosen operational configuration and
verify robustness to specification changes.


# Setup {#sec-setup}

## Packages

```{r load-packages}
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(grf)
library(policytree)
library(foreach)
library(doFuture)
library(future)

# Optional
library(patchwork)
library(weightedsurv)

theme_set(theme_minimal(base_size = 12))
```

```{r gbsg-params}
#| include: false

# ─── All tunable parameters centralised here ───────────────────────────────

# ── Parallel processing ────────────────────────────────────────────────────
# Set to detectCores() - 1 for production; 2 for CRAN checks.
n_workers <- floor(0.95 * max(1L, parallel::detectCores(logical = FALSE) - 1L))

# ── Main-analysis ForestSearch operational parameters ──────────────────────
# These are the settings used throughout Sections 3-6.
c1_paper            <- 1.25    # hr.threshold (screening)
c2_paper            <- 1.00    # hr.consistency (consistency)
pcons_val           <- 0.90    # pconsistency.threshold
stop_thresh_val     <- 0.95    # early-stopping threshold
fs_splits_val       <- 400L    # consistency-evaluation splits
max_sgs_search_val  <- 10L
n_min_val           <- 60L
d0_min_val          <- 10L
d1_min_val          <- 10L
maxk_val            <- 2L

# ── Bootstrap / cross-validation counts ────────────────────────────────────
# Increase for production:
#   NB_boot  = 500-2000   (here: 20 for fast knit)
#   Ksims_cv = 50         (here: 10 for fast knit)
NB_boot   <- 1000L
Ksims_cv  <- 50L

# ── FPR calibration (appendix) ─────────────────────────────────────────────
# The appendix runs two independent null-simulation batches:
#   * Calibration batch: produces the primary FPR estimate and its CI,
#     plus the asymptotic P1 / implied L_eff diagnostics.
#   * Verification batch: an independent run on a disjoint seed range
#     that confirms the calibration FPR.  Skippable by setting
#     nsims_verify_fpr = 0L.  Scale up (e.g. 10000L–20000L) to achieve
#     ±0.2–0.3 pp precision for a final production result (~1–3 hr).
#
# CI half-width at FPR ~ 10%:
#   200 sims  ≈ ±2.1 pp  |  500 sims  ≈ ±1.3 pp  |  1000 sims ≈ ±0.9 pp
#   5000 sims ≈ ±0.4 pp  |  10000 sims ≈ ±0.3 pp | 20000 sims ≈ ±0.2 pp
nsims_tier1_fpr   <- 500L     # calibration batch
nsims_verify_fpr  <- 1000L    # verification batch (set to 0L to skip)
sim_seed_fpr      <- 8316951L
fpr_target_surv   <- 0.12     # reference line on detection curves

# ── Theta-grid verification (appendix detection curve) ─────────────────────
# Independent DGM simulations at multiple θ†(H) values (true harm-subgroup
# HR) to test whether the corrected-FPR prediction from the calibration
# batch (estimated at θ = 1) can be "transported" to neighbouring θ values.
# Each θ requires its own alternative DGM built via
# calibrate_k_inter(target_hr_harm = θ) -> setup_gbsg_dgm(model = "alt", ...).
#
# theta_merge_cut is the (working-hypothesis) upper edge of the region
# where the corrected prediction is expected to approximate the truth.
# Below theta_merge_cut the merged curve shows the corrected prediction;
# above it the merged curve shows the uncorrected per-subgroup P1.
# The diamonds from theta_grid_verify empirically test this assumption.
theta_grid_verify  <- c(0.75, 0.90, 1.25, 1.50, 2.00)   # θ = 1.00 comes from verification batch
theta_merge_cut    <- 1.25                   # boundary (working hypothesis)
nsims_theta_verify <- 500L                   # sims per θ (set 0L to skip grid)

# Survival DGM parameters (for the DGM-based FPR estimate)
# Note: the censoring structure is baked into setup_gbsg_dgm() and shifted
# by cens_adjust_dgm.  We do NOT set prop_cens as a parameter here — the
# observed censoring rate is measured empirically from the DGM replicates
# after the calibration batch completes (results_null$p.cens).
analysis_time_dgm <- 84L             # months
cens_adjust_dgm   <- log(1.5)
confounders_dgm   <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")

# ── Caching of expensive appendix simulations ──────────────────────────────
# The appendix simulation chunks (calibration batch, verification batch,
# theta-grid verification) are the dominant runtime contributors (often
# > 1 hour at production nsims settings).  Each chunk's results are cached
# to an RDS file keyed by a parameter fingerprint; a rebuild is triggered
# whenever any input to the cache key changes.
#
# To FORCE re-computation (e.g. after changing the DGM or fs_params):
#   - Set cache_refresh <- TRUE below, OR
#   - Delete the corresponding .rds file from cache_dir manually.
#
# Cache files live in the chat's working directory; they are small
# (< 5 MB each) and safe to commit or .gitignore as preferred.
cache_dir     <- "_cache_gbsg_appendix"
cache_refresh <- FALSE   # set TRUE to force re-running all cached chunks

if (!dir.exists(cache_dir)) dir.create(cache_dir, showWarnings = FALSE)

# Helper: read an RDS cache if present and compatible, else run `compute()`
# and write the result. `key` is a named list of all values that invalidate
# the cache when they change.  `label` is a short human-readable tag used
# in messages.
.cache_compute <- function(cache_file, key, compute, label = "chunk") {
  path <- file.path(cache_dir, cache_file)
  if (!cache_refresh && file.exists(path)) {
    cached <- readRDS(path)
    if (identical(cached$key, key)) {
      message(sprintf("[cache HIT ] %s: loaded from %s", label, path))
      return(cached$value)
    }
    message(sprintf("[cache MISS] %s: key changed, rebuilding", label))
  } else if (cache_refresh) {
    message(sprintf("[cache SKIP] %s: cache_refresh = TRUE", label))
  } else {
    message(sprintf("[cache MISS] %s: no cache file, building", label))
  }
  value <- compute()
  saveRDS(list(key = key, value = value), path)
  message(sprintf("[cache SAVE] %s: wrote %s", label, path))
  value
}

# ── Internals (do not edit) ────────────────────────────────────────────────
timings <- list()
t_vignette_start <- proc.time()
```


# Data: German Breast Cancer Study Group {#sec-data}

## Study Background

The GBSG trial evaluated hormonal treatment (tamoxifen) versus chemotherapy
in node-positive breast cancer patients. Key characteristics:

- **Sample size**: N = 686
- **Outcome**: Recurrence-free survival time
- **Censoring rate**: ~56%
- **Treatment**: Hormonal therapy (tamoxifen) vs. chemotherapy

## Data Preparation

```{r data-setup}
# Load GBSG data (included in forestsearch package)
df.analysis <- gbsg

# Prepare analysis variables
df.analysis <- within(df.analysis, {
  id <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3 <- ifelse(grade == "3", 1, 0)
  treat <- hormon
})

# Define variable roles
confounders.name <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
outcome.name <- "time_months"
event.name   <- "status"
id.name      <- "id"
treat.name   <- "hormon"

cat("Sample size:", nrow(df.analysis), "\n")
cat("Events:", sum(df.analysis[[event.name]]),
    sprintf("(%.1f%%)\n", 100 * mean(df.analysis[[event.name]])))
cat("Baseline factors:", paste(confounders.name, collapse = ", "), "\n")
```

## Baseline Characteristics

```{r baseline-table}
create_summary_table(
  data             = df.analysis,
  treat_var        = treat.name,
  table_title      = "GBSG Baseline Characteristics by Treatment Arm",
  vars_continuous  = c("age", "nodes", "size", "er", "pgr"),
  vars_categorical = c("grade", "meno"),
  font_size        = 12
)
```

## Kaplan-Meier Analysis (ITT Population)

```{r km-itt}
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "ITT Kaplan-Meier curves with 95% CI and log-rank test."

dfcount <- df_counting(
  df         = df.analysis,
  by.risk    = 6,
  tte.name   = outcome.name,
  event.name = event.name,
  treat.name = treat.name
)

plot_weighted_km(
  dfcount,
  conf.int      = TRUE,
  show.logrank  = TRUE,
  ymax          = 1.05,
  xmed.fraction = 0.775,
  ymed.offset   = 0.125
)
```

The ITT Cox hazard ratio estimate is approximately 0.69 (95% CI: 0.54, 0.89),
suggesting an overall benefit for hormonal therapy.


# Preliminary Analysis: Generalized Random Forests {#sec-grf}

Before running ForestSearch, we use GRF to explore potential treatment
effect heterogeneity and identify candidate factors.

```{r grf-analysis}
t0 <- proc.time()

grf_est <- grf.subg.harm.survival(
  data             = df.analysis,
  confounders.name = confounders.name,
  outcome.name     = outcome.name,
  event.name       = event.name,
  id.name          = id.name,
  treat.name       = treat.name,
  maxdepth         = 2,
  n.min            = n_min_val,
  dmin.grf         = 12,
  frac.tau         = 0.6,
  details          = TRUE,
  return_selected_cuts_only = FALSE
)

timings$grf <- (proc.time() - t0)["elapsed"]
```

```{r grf-trees}
#| fig-width: 10
#| fig-height: 4
#| fig-cap: "GRF policy trees. leaf1 = recommend control; leaf2 = recommend treatment."

oldpar <- par(mfrow = c(1, 2))
plot(grf_est$tree1, leaf.labels = c("Control", "Treat"), main = "Depth 1")
plot(grf_est$tree2, leaf.labels = c("Control", "Treat"), main = "Depth 2")
par(oldpar)
```

GRF identifies estrogen receptor status (ER) as a key factor, with ER ≤ 0
suggesting potential harm from hormonal therapy.


# ForestSearch Analysis {#sec-forestsearch}

## Running ForestSearch

ForestSearch performs an exhaustive search over candidate subgroup
combinations with up to `maxk` factors. Key parameters:

| Parameter | Value | Description |
|-----------|-------|-------------|
| `hr.threshold` | `r c1_paper` | Minimum HR for consistency evaluation |
| `hr.consistency` | `r c2_paper` | Minimum consistency rate for candidates |
| `pconsistency.threshold` | `r pcons_val` | Required consistency for selection |
| `maxk` | `r maxk_val` | Maximum factors in subgroup definition |
| `n.min` | `r n_min_val` | Minimum subgroup sample size |
| `d0.min`, `d1.min` | `r d0_min_val` | Minimum events per treatment arm |

```{r forestsearch-main}
#| fig-width: 10
#| fig-height: 7

t0 <- proc.time()

fs <- forestsearch(
  df.analysis,
  confounders.name = confounders.name,
  outcome.name     = outcome.name,
  treat.name       = treat.name,
  event.name       = event.name,
  id.name          = id.name,
  # Threshold parameters (per León et al. 2024)
  hr.threshold           = c1_paper,
  hr.consistency         = c2_paper,
  pconsistency.threshold = pcons_val,
  stop_threshold         = stop_thresh_val,
  # Search configuration
  sg_focus             = "maxSG",
  max_subgroups_search = max_sgs_search_val,
  use_twostage         = TRUE,
  # Factor selection
  use_grf                   = TRUE,
  return_selected_cuts_only = TRUE,
  use_lasso                 = TRUE,
  cut_type                  = "default",
  # Subgroup constraints
  maxk   = maxk_val,
  n.min  = n_min_val,
  d0.min = d0_min_val,
  d1.min = d1_min_val,
  # Consistency evaluation
  fs.splits = fs_splits_val,
  # Parallel processing
  parallel_args = list(
    plan         = "multisession",
    workers      = n_workers,
    show_message = TRUE
  ),
  # Output
  showten_subgroups = TRUE,
  details           = TRUE,
  plot.sg           = TRUE
)

plan("sequential")
timings$forestsearch <- (proc.time() - t0)["elapsed"]

cat("\nForestSearch completed in",
    round(timings$forestsearch, 1), "seconds\n")
```


## ForestSearch Results

### Identified Subgroups

```{r fs-results}
res_tabs <- sg_tables(fs, ndecimals = 3, which_df = "est")
res_tabs$sg10_out
```

### Treatment Effect Estimates

```{r fs-estimates}
res_tabs$tab_estimates
```

### Identified Subgroup Definition

```{r fs-subgroup}
cat("Identified subgroup (H):", paste(fs$sg.harm, collapse = " & "), "\n")
cat("Subgroup size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%% of ITT)\n",
            100 * mean(fs$df.est$treat.recommend == 0)))
```

ForestSearch identifies **Estrogen ≤ 0** (ER-negative) as the subgroup with
potential harm. This is biologically plausible: tamoxifen is a selective
estrogen receptor modulator with limited efficacy in ER-negative tumors.

### Agreement with GRF

ForestSearch and GRF identify subgroups by different mechanisms —
FS via exhaustive search with consistency filtering, GRF via causal
survival forests plus policy trees — so agreement on subject-level
membership provides an informal robustness check. `compare_fs_subgroups()`
produces a publication-ready cross-tabulation plus Cohen's kappa.

```{r fs-grf-concordance}
sg_compare <- compare_fs_subgroups(
  fs1               = fs,
  fs2               = grf_est,
  label1            = "ForestSearch",
  label2            = "GRF Standalone",
  id.name           = id.name,
  subgroup_notation = "harm",
  title             = "FS vs GRF: Subgroup Membership Concordance",
  subtitle          = "GBSG observed data (Cox, harm search)",
  font_size         = 13
)
sg_compare$table
```

```{r fs-grf-concordance-stats}
cat("Concordance statistics:\n")
cat(sprintf("  Agreement:       %.1f%%\n",
    100 * sg_compare$concordance$agreement))
cat(sprintf("  Cohen's kappa:   %.3f\n",
    sg_compare$concordance$kappa))
cat(sprintf("  Concordant:      %d / Discordant: %d / Total: %d\n",
    sg_compare$concordance$n_agree,
    sg_compare$concordance$n_disagree,
    sg_compare$concordance$n_total))
```


# Bootstrap Bias Correction {#sec-bootstrap}

## Rationale

Cox model estimates from identified subgroups are **upwardly biased** due to
the selection process (subgroups are selected *because* they show extreme
effects). Bootstrap bias correction addresses this by:

1. Resampling with replacement
2. Re-running the entire ForestSearch algorithm
3. Computing bias terms from bootstrap vs. observed estimates
4. Applying infinitesimal jackknife variance estimation

## Running Bootstrap Analysis

```{r bootstrap}
t0 <- proc.time()

fs_bc <- forestsearch_bootstrap_dofuture(
  fs.est        = fs,
  nb_boots      = NB_boot,
  show_three    = FALSE,
  details       = FALSE,
  parallel_args = list(
    plan         = "multisession",
    workers      = n_workers,
    show_message = TRUE
  )
)

plan("sequential")
timings$bootstrap <- (proc.time() - t0)["elapsed"]

cat("\nBootstrap completed in",
    round(timings$bootstrap / 60, 1), "minutes\n")
```

## Bootstrap Summary and Diagnostics

```{r bootstrap-summary}
summaries <- summarize_bootstrap_results(
  sgharm       = fs$sg.harm,
  boot_results = fs_bc,
  create_plots = TRUE,
  est.scale    = "hr"
)
summaries$table
```

## Kaplan-Meier by Identified Subgroups

```{r fs-results-figure}
#| fig-width: 10
#| fig-height: 7
#| fig-cap: "Kaplan-Meier survival curves by identified subgroup"

km_result <- plot_sg_weighted_km(
  fs.est         = fs,
  outcome.name   = "time_months",
  event.name     = "status",
  treat.name     = "hormon",
  show.logrank   = FALSE,
  conf.int       = TRUE,
  by.risk        = 12,
  show.cox       = FALSE,
  show.cox.bc    = TRUE,
  fs_bc          = fs_bc,
  hr_bc_position = "topright"
)
```

::: {.callout-note}
`r figure_note(km_result, prefix = "**Note:** ", custom_text = "Medians [95% CI] for arms are un-adjusted.")`
:::


### Event Count Summary

Low event counts can lead to unstable HR estimates.

```{r event-summary}
event_summary <- summarize_bootstrap_events(fs_bc, threshold = 15)
```

### Bootstrap Diagnostics

```{r bootstrap-diagnostics}
summaries$diagnostics_table_gt
```

### Subgroup Agreement

How consistently does bootstrap identify the same subgroup?

```{r subgroup-agreement}
if (!is.null(summaries$subgroup_summary$original_agreement)) {
  summaries$subgroup_summary$original_agreement
}

if (!is.null(summaries$subgroup_summary$factor_presence)) {
  summaries$subgroup_summary$factor_presence
}
```

### Bootstrap Distributions

```{r bootstrap-plots}
#| fig-width: 10
#| fig-height: 4

if (!is.null(summaries$plots)) {
  summaries$plots$H_distribution + summaries$plots$Hc_distribution
}
```


# Cross-Validation {#sec-cv}

Cross-validation assesses the **stability** of the ForestSearch algorithm.
Two approaches are available.

## K-Fold Cross-Validation

```{r kfold-cv}
t0 <- proc.time()

fs_kfold <- forestsearch_tenfold(
  fs.est        = fs,
  sims          = Ksims_cv,
  Kfolds        = 10,
  details       = FALSE,
  parallel_args = list(
    plan         = "multisession",
    workers      = n_workers,
    show_message = FALSE
  )
)

plan("sequential")
timings$kfold <- (proc.time() - t0)["elapsed"]
metrics_tables <- cv_metrics_tables(fs_kfold)
metrics_tables
```

## Out-of-Bag (N-Fold) Cross-Validation

```{r oob-cv}
#| eval: true

t0 <- proc.time()

fs_OOB <- forestsearch_Kfold(
  fs.est        = fs,
  details       = FALSE,
  Kfolds        = round(nrow(df.analysis) / 100, 0),
  parallel_args = list(
    plan         = "multisession",
    workers      = n_workers,
    show_message = TRUE
  )
)

plan("sequential")
timings$oob <- (proc.time() - t0)["elapsed"]

cv_out <- forestsearch_KfoldOut(
  res     = fs_OOB,
  details = FALSE,
  outall  = TRUE
)

tables <- cv_summary_tables(cv_out)
tables$combined_table
tables$metrics_table
```


# Results Visualization {#sec-viz}

## Forest Plot

```{r forest-plot}
#| fig-width: 18
#| fig-height: 12
#| fig-cap: "Subgroup forest plot including identified subgroups"

subgroups <- list(
  age_gt65     = list(subset_expr = "age > 65",  name = "Age > 65",  type = "reference"),
  age_le65     = list(subset_expr = "age <= 65", name = "Age <= 65", type = "reference"),
  pgr_positive = list(subset_expr = "pgr > 0",   name = "PgR > 0",   type = "reference"),
  pgr_negative = list(subset_expr = "pgr <= 0",  name = "PgR <= 0",  type = "reference")
)

my_theme <- create_forest_theme(base_size = 24,
                                 footnote_fontsize = 17,
                                 cv_fontsize = 22)

result <- plot_subgroup_results_forestplot(
  fs_results = list(
    fs.est   = fs,
    fs_bc    = fs_bc,
    fs_OOB   = NULL,
    fs_kfold = fs_kfold
  ),
  df_analysis      = df.analysis,
  subgroup_list    = subgroups,
  outcome.name     = outcome.name,
  event.name       = event.name,
  treat.name       = treat.name,
  E.name           = "Hormonal",
  C.name           = "Chemo",
  ci_column_spaces = 25,
  xlog             = TRUE,
  theme            = my_theme
)

render_forestplot(result)
```

## KM Difference Plots: ITT and Subgroups

```{r KMdiffs}
#| fig-width: 8
#| fig-height: 6
#| fig-cap: "ITT (black) and subgroup KM differences with 95% confidence band."

ref_sgs <- list(
  age_young = list(subset_expr = "age < 65",  color = "brown"),
  age_old   = list(subset_expr = "age >= 65", color = "orange")
)

plot_km_band_forestsearch(
  df            = df.analysis,
  fs.est        = fs,
  ref_subgroups = ref_sgs,
  outcome.name  = outcome.name,
  event.name    = event.name,
  treat.name    = treat.name,
  draws_band    = 20
)
```


# Summary and Interpretation {#sec-summary}

## Key Findings

```{r summary-findings}
cat(strrep("=", 60), "\n")
cat("FORESTSEARCH ANALYSIS SUMMARY\n")
cat(strrep("=", 60), "\n\n")

cat("Dataset: GBSG (N =", nrow(df.analysis), ")\n")
cat("Outcome: Recurrence-free survival\n\n")

cat("ITT Analysis:\n")
cat("  HR (95% CI): 0.69 (0.54, 0.89)\n\n")

cat("Identified Subgroup (H):\n")
cat("  Definition:", paste(fs$sg.harm, collapse = " & "), "\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 0),
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 0)))
cat("  Unadjusted HR:",
    sprintf("%.2f", exp(fs$grp.consistency$out_sg$result$hr[1])), "\n")

cat("\nComplement Subgroup (Hc):\n")
cat("  Size:", sum(fs$df.est$treat.recommend == 1),
    sprintf("(%.1f%%)\n", 100 * mean(fs$df.est$treat.recommend == 1)))
```

## Clinical Interpretation

The ForestSearch analysis identifies **estrogen receptor-negative (ER ≤ 0)**
patients as a subgroup with potential lack of benefit from hormonal therapy.

**Biological plausibility**: Tamoxifen is a selective estrogen receptor
modulator. Its efficacy depends on ER expression. The finding that ER-negative
patients may not benefit is consistent with:

- Mechanistic understanding of tamoxifen action
- Meta-analyses showing no tamoxifen benefit in ER-negative breast cancer
- Clinical guidelines recommending tamoxifen primarily for ER-positive tumors

**Caveats**:

1. This is an **exploratory** analysis requiring independent validation.
2. The bias-corrected estimates have wider confidence intervals.
3. Cross-validation metrics should be evaluated for algorithm stability.
4. The procedure-level false-positive rate at the operational
   thresholds (c1 = `r c1_paper`, c2 = `r sprintf("%.2f", c2_paper)`)
   is directly quantified in @sec-appendix-fpr below.


# Appendix: Direct FPR Calibration at the Operational Thresholds {#sec-appendix-fpr}

## Motivation {#sec-appendix-rationale}

The main analysis uses threshold pair
(c1 = `r c1_paper`, c2 = `r sprintf("%.2f", c2_paper)`) with
`pconsistency.threshold = `r pcons_val`` and
`stop_threshold = `r stop_thresh_val``.  A natural question follows
from any exploratory subgroup-identification analysis:

> Under the null hypothesis of *no treatment-by-subgroup interaction*,
> how often would ForestSearch spuriously identify **any** subgroup
> at these operational settings?

We answer this using the parametric GBSG DGM via
`setup_gbsg_dgm(model = "null")` and the package's supported
simulation pipeline — `run_simulation_analysis()` →
`collect_results()` → `format_oc_results()` — as configured in the
León (2024) Table 1 reproduction documents.  This yields a direct
Monte Carlo FPR estimate plus the full operating-characteristics
table (detection + classification) under H0 at the operational
configuration used for subgroup identification.


## Analysis setup {#sec-appendix-setup}

```{r fpr-setup}
#| code-fold: true

# ── Parallel plan (same workers as main analysis) ─────────────────
plan(multisession, workers = n_workers)
cat("Workers:", n_workers, "\n")

# DGM and ForestSearch params are constructed in @sec-appendix-dgm-setup below.

# Wilson CI helper (used by theta-grid verification and summary table).
.wilson_ci <- function(x, n, z = 1.96) {
  if (n == 0) return(c(NA_real_, NA_real_))
  p <- x / n
  denom  <- 1 + z^2 / n
  centre <- (p + z^2 / (2 * n)) / denom
  margin <- z * sqrt(p * (1 - p) / n + z^2 / (4 * n^2)) / denom
  c(max(0, centre - margin), min(1, centre + margin))
}
```


## DGM-based FPR calibration {#sec-appendix-dgm}

We use `setup_gbsg_dgm(model = "null")` as the null DGM — parameters
calibrated to the GBSG AFT structure with treatment effect set to the
null and no treatment-by-subgroup interaction.  `simulate_from_dgm()`
then generates the replicates inside `run_simulation_analysis()`.

### DGM and ForestSearch parameter setup {#sec-appendix-dgm-setup}

```{r fpr-dgm-setup}
#| code-fold: true

dgm_gbsg_null <- setup_gbsg_dgm(
  model   = "null",
  verbose = FALSE
)

# ForestSearch parameter list for run_simulation_analysis().  Retains
# the operational pcons / stop_threshold from the main GBSG analysis
# so the DGM-based FPR estimate is calibrated at the same operational
# setting used for subgroup identification.  Formatted in the style
# of reproduce_table1_m1.qmd's fs_cox list.
fs_params_runsim <- list(
  outcome.name           = "y_sim",
  event.name             = "event_sim",
  treat.name             = "treat_sim",
  id.name                = "id",
  hr.threshold           = c1_paper,
  hr.consistency         = c2_paper,
  pconsistency.threshold = pcons_val,
  stop_threshold         = stop_thresh_val,
  sg_focus               = "hr",
  max_subgroups_search   = max_sgs_search_val,
  use_twostage           = TRUE,
  fs.splits              = fs_splits_val,
  n.min                  = n_min_val,
  d0.min                 = d0_min_val,
  d1.min                 = d1_min_val,
  maxk                   = maxk_val,
  vi.grf.min             = -0.2,
  seedit                 = sim_seed_fpr,
  use_lasso              = TRUE,
  use_grf                = TRUE,        # FSlg: LASSO + GRF
  tune_grf               = FALSE,
  is.RCT                 = TRUE,
  details                = FALSE,
  quiet                  = TRUE,
  plot.sg                = FALSE,
  parallel_args          = list(
    plan = "sequential", workers = 1, show_message = FALSE
  )
)

# Cox formula (treatment-only, per León et al. 2024 Section 2)
cox_formula_adj_runsim <- survival::Surv(y_sim, event_sim) ~ treat_sim

# DGM-reference censoring rate: a single large-N draw from the DGM
# gives a low-variance Monte Carlo estimate of the theoretical
# censoring rate the DGM produces at (analysis_time_dgm, cens_adjust_dgm).
# At n = 10000, SE ~ 0.005 — precise enough to cross-check against the
# empirical rate averaged over the calibration replicates later on.
dgm_ref_draw <- simulate_from_dgm(
  dgm_gbsg_null,
  n             = 10000L,
  analysis_time = analysis_time_dgm,
  cens_adjust   = cens_adjust_dgm,
  seed          = sim_seed_fpr + 999L
)
prop_cens_ref <- 1 - mean(dgm_ref_draw$event_sim)

cat(sprintf("DGM null ready | N = %d | analysis_time = %d months | cens_adjust = %.3f\n",
            nrow(df.analysis), analysis_time_dgm, cens_adjust_dgm))
cat(sprintf("DGM reference prop_cens: %.3f  (from n = 10000 single draw)\n",
            prop_cens_ref))
```

### Calibration batch {#sec-appendix-dgm-loop}

Primary FPR estimate from `nsims_tier1_fpr` = `r nsims_tier1_fpr`
null replicates.

```{r fpr-dgm-loop}
#| code-fold: true

# Cache key: everything that changes the simulation output.
.cal_key <- list(
  nsims         = nsims_tier1_fpr,
  seed_base     = sim_seed_fpr + 2000L,
  fs_params     = fs_params_runsim,
  dgm_id        = list(model = "null", seed = 8316951L),
  n_sample      = nrow(df.analysis),
  analysis_time = analysis_time_dgm,
  cens_adjust   = cens_adjust_dgm,
  confounders   = confounders_dgm
)

.cal_cached <- .cache_compute(
  cache_file = "results_null.rds",
  key        = .cal_key,
  label      = sprintf("calibration batch (%d sims)", nsims_tier1_fpr),
  compute    = function() {
    t0 <- proc.time()[3]

    raw <- foreach(
      i = seq_len(nsims_tier1_fpr),
      .errorhandling = "pass",
      .options.future = list(packages = c("forestsearch"), seed = TRUE)
    ) %dofuture% {
      future::plan("sequential")

      r_fslg <- tryCatch(
        run_simulation_analysis(
          sim_id           = i,
          dgm              = dgm_gbsg_null,
          n_sample         = nrow(df.analysis),
          confounders_base = confounders_dgm,
          n_add_noise      = 0L,
          run_fs           = TRUE,
          run_fs_grf       = FALSE,
          run_grf          = FALSE,
          fs_params        = fs_params_runsim,
          cox_formula_adj  = cox_formula_adj_runsim,
          analysis_time    = analysis_time_dgm,
          cens_adjust      = cens_adjust_dgm,
          seed_base        = sim_seed_fpr + 2000L,
          verbose          = FALSE
        ),
        error = function(e) NULL
      )
      if (!is.null(r_fslg)) r_fslg$analysis <- "FSlg"

      data.table::rbindlist(
        Filter(Negate(is.null), list(r_fslg)),
        fill = TRUE
      )
    }

    plan("sequential")
    list(
      results = forestsearch::collect_results(raw, "H0 null"),
      elapsed = proc.time()[3] - t0
    )
  }
)

results_null       <- .cal_cached$results
timings$fpr_runsim <- .cal_cached$elapsed
n_failed_null      <- attr(results_null, "n_failed")
cat(sprintf("Calibration batch: %d sims (elapsed %.1f min, cache-aware)\n",
            nsims_tier1_fpr, timings$fpr_runsim / 60))
if (!is.null(n_failed_null) && n_failed_null > 0L)
  cat(sprintf("  *** %d of %d replicates failed ***\n",
              n_failed_null, nsims_tier1_fpr))
```


### Verification batch {#sec-appendix-dgm-verify}

Independent null-simulation run on a **disjoint seed range**
(`seed_base = sim_seed_fpr + 50000L`) that confirms the calibration
FPR is reproducible on a fresh sample.  Scale `nsims_verify_fpr`
upward (e.g. 10,000–20,000) to achieve ±0.2–0.3 pp precision for a
final production result.  Set `nsims_verify_fpr = 0L` in the setup
chunk to skip.

```{r fpr-dgm-verify}
#| code-fold: true

if (nsims_verify_fpr > 0L) {
  .ver_key <- list(
    nsims         = nsims_verify_fpr,
    seed_base     = sim_seed_fpr + 50000L,
    fs_params     = fs_params_runsim,
    dgm_id        = list(model = "null", seed = 8316951L),
    n_sample      = nrow(df.analysis),
    analysis_time = analysis_time_dgm,
    cens_adjust   = cens_adjust_dgm,
    confounders   = confounders_dgm
  )

  .ver_cached <- .cache_compute(
    cache_file = "results_verify.rds",
    key        = .ver_key,
    label      = sprintf("verification batch (%d sims)", nsims_verify_fpr),
    compute    = function() {
      t0 <- proc.time()[3]

      # Reset worker pool to prevent memory accumulation from a prior
      # batch before running this one.
      plan(sequential); gc(verbose = FALSE); Sys.sleep(1)
      plan(multisession, workers = n_workers)

      raw <- foreach(
        i = seq_len(nsims_verify_fpr),
        .errorhandling = "pass",
        .options.future = list(packages = c("forestsearch"), seed = TRUE)
      ) %dofuture% {
        future::plan("sequential")

        r_fslg <- tryCatch(
          run_simulation_analysis(
            sim_id           = i,
            dgm              = dgm_gbsg_null,
            n_sample         = nrow(df.analysis),
            confounders_base = confounders_dgm,
            n_add_noise      = 0L,
            run_fs           = TRUE,
            run_fs_grf       = FALSE,
            run_grf          = FALSE,
            fs_params        = fs_params_runsim,
            cox_formula_adj  = cox_formula_adj_runsim,
            analysis_time    = analysis_time_dgm,
            cens_adjust      = cens_adjust_dgm,
            seed_base        = sim_seed_fpr + 50000L,
            verbose          = FALSE
          ),
          error = function(e) NULL
        )
        if (!is.null(r_fslg)) r_fslg$analysis <- "FSlg"

        data.table::rbindlist(
          Filter(Negate(is.null), list(r_fslg)),
          fill = TRUE
        )
      }

      plan("sequential")
      list(
        results = forestsearch::collect_results(raw, "H0 null (verify)"),
        elapsed = proc.time()[3] - t0
      )
    }
  )

  results_verify      <- .ver_cached$results
  timings$fpr_verify  <- .ver_cached$elapsed
  n_failed_verify     <- attr(results_verify, "n_failed")
  cat(sprintf("Verification batch: %d sims (elapsed %.1f min, cache-aware)\n",
              nsims_verify_fpr, timings$fpr_verify / 60))
  if (!is.null(n_failed_verify) && n_failed_verify > 0L)
    cat(sprintf("  *** %d of %d replicates failed ***\n",
                n_failed_verify, nsims_verify_fpr))
} else {
  results_verify <- NULL
  timings$fpr_verify <- 0
  message("Verification batch skipped (nsims_verify_fpr = 0L). ",
          "Set nsims_verify_fpr to a positive integer in the setup chunk to run.")
}
```

### Theta-grid verification {#sec-appendix-dgm-theta}

Independent DGM simulations at θ†(H) ∈ `r paste(theta_grid_verify, collapse = ", ")`
test whether the corrected-FPR prediction from the calibration batch
(estimated at θ = 1) can be transported to neighbouring values of the
true harm-subgroup HR.  For each θ we (1) calibrate `k_inter` to
achieve `hr_H_true = θ`, (2) build an alternative DGM with
`setup_gbsg_dgm(model = "alt", k_inter = ...)`, and (3) run
`nsims_theta_verify` null replicates through `run_simulation_analysis()`.
The resulting detection rates are overlaid as diamonds on the merged
detection curve in @sec-appendix-detection.  The θ = 1 diamond there
comes from the verification batch above (no alternative DGM needed —
θ = 1 is the null).  Set `nsims_theta_verify = 0L` in the setup chunk
to skip this block.

```{r fpr-theta-grid}
#| code-fold: true

if (nsims_theta_verify > 0L && length(theta_grid_verify) > 0L) {
  .theta_key <- list(
    theta_grid    = theta_grid_verify,
    nsims         = nsims_theta_verify,
    fs_params     = fs_params_runsim,
    n_sample      = nrow(df.analysis),
    analysis_time = analysis_time_dgm,
    cens_adjust   = cens_adjust_dgm,
    confounders   = confounders_dgm,
    seed_offsets  = sim_seed_fpr + 80000L + seq_along(theta_grid_verify),
    sim_seeds     = sim_seed_fpr + 100000L + 1000L * seq_along(theta_grid_verify)
  )

  .theta_cached <- .cache_compute(
    cache_file = "theta_verify.rds",
    key        = .theta_key,
    label      = sprintf("theta-grid verification (%d θ × %d sims)",
                         length(theta_grid_verify), nsims_theta_verify),
    compute    = function() {
      t0 <- proc.time()[3]

      # ── Step 1: Calibrate k_inter and build one alt DGM per θ ──
      # Sequential computation — done BEFORE activating multisession
      # so workers don't sit idle during ~30–60 s of k_inter calibration.
      plan(sequential)

      alt_dgms <- vector("list", length(theta_grid_verify))
      names(alt_dgms) <- sprintf("theta_%.2f", theta_grid_verify)
      for (j in seq_along(theta_grid_verify)) {
        th   <- theta_grid_verify[j]
        kj   <- calibrate_k_inter(
          target_hr_harm = th,
          k_treat        = 1.0,
          seed           = sim_seed_fpr + 80000L + j,
          verbose        = FALSE
        )
        alt_dgms[[j]] <- setup_gbsg_dgm(
          model   = "alt",
          k_treat = 1.0,
          k_inter = kj,
          seed    = sim_seed_fpr + 80000L + j,
          verbose = FALSE
        )
        cat(sprintf("  θ=%.2f calibrated: k_inter=%.4f | hr_H_true=%.3f\n",
                    th, kj, alt_dgms[[j]]$hr_H_true))
      }

      # ── Step 2: Run sims per θ, resetting workers before each ──
      theta_rows <- vector("list", length(theta_grid_verify))
      for (j in seq_along(theta_grid_verify)) {
        th  <- theta_grid_verify[j]
        dgj <- alt_dgms[[j]]
        seed_offset_j <- 100000L + 1000L * j

        plan(sequential); gc(verbose = FALSE); Sys.sleep(1)
        plan(multisession, workers = n_workers)

        raw_j <- foreach(
          i = seq_len(nsims_theta_verify),
          .errorhandling = "pass",
          .options.future = list(packages = c("forestsearch"), seed = TRUE)
        ) %dofuture% {
          future::plan("sequential")

          r <- tryCatch(
            run_simulation_analysis(
              sim_id           = i,
              dgm              = dgj,
              n_sample         = nrow(df.analysis),
              confounders_base = confounders_dgm,
              n_add_noise      = 0L,
              run_fs           = TRUE,
              run_fs_grf       = FALSE,
              run_grf          = FALSE,
              fs_params        = fs_params_runsim,
              cox_formula_adj  = cox_formula_adj_runsim,
              analysis_time    = analysis_time_dgm,
              cens_adjust      = cens_adjust_dgm,
              seed_base        = sim_seed_fpr + seed_offset_j,
              verbose          = FALSE
            ),
            error = function(e) NULL
          )
          if (!is.null(r)) r$analysis <- "FSlg"
          data.table::rbindlist(
            Filter(Negate(is.null), list(r)), fill = TRUE)
        }

        res_j <- forestsearch::collect_results(
          raw_j, sprintf("theta=%.2f", th))
        rj    <- res_j[res_j$analysis == "FSlg", ]
        ndet  <- sum(rj$any.H == 1L, na.rm = TRUE)
        nrep  <- nrow(rj)
        ci    <- .wilson_ci(ndet, nrep)

        theta_rows[[j]] <- data.table(
          theta      = th,
          nsims      = nrep,
          detections = ndet,
          fpr        = ndet / nrep,
          ci_lo      = ci[1],
          ci_hi      = ci[2]
        )
        cat(sprintf("  θ=%.2f: detection %d/%d = %.1f%% [%.1f%%, %.1f%%]\n",
                    th, ndet, nrep, 100 * ndet / nrep,
                    100 * ci[1], 100 * ci[2]))
      }

      plan("sequential")
      list(
        df      = data.table::rbindlist(theta_rows),
        elapsed = proc.time()[3] - t0
      )
    }
  )

  theta_verify_df          <- .theta_cached$df
  timings$fpr_theta_verify <- .theta_cached$elapsed

  cat(sprintf("\nTheta-grid verification: %d θ values × %d sims (elapsed %.1f min, cache-aware)\n",
              length(theta_grid_verify), nsims_theta_verify,
              timings$fpr_theta_verify / 60))
} else {
  theta_verify_df <- NULL
  timings$fpr_theta_verify <- 0
  message("Theta-grid verification skipped (nsims_theta_verify = 0L).")
}
```


## Results {#sec-appendix-consistency}

### Operating characteristics (detection + classification) {#sec-appendix-oc}

```{r fpr-oc-table}
# Publication-ready OC table via format_oc_results().  Under H0,
# any.H is the FPR; sens / ppv are null by construction (the
# unconditional averaging convention used in the reproduction docs).
format_oc_results(
  results_null,
  analyses = "FSlg",
  metrics  = c("detection", "classification"),
  digits   = 3,
  title    = "GBSG Null — DGM-based FPR Calibration",
  subtitle = sprintf(
    "N = %d, %d sims, θ†(ITT) ≈ %.2f",
    nrow(df.analysis), nsims_tier1_fpr, dgm_gbsg_null$hr_causal
  ),
  subgroup_notation = "harm"
)
```

### Asymptotic approximation {#sec-appendix-asymp}

The León et al. (2024) asymptotic approximation (equation 3) gives
the per-subgroup detection probability P1 at the operational
threshold pair, using effective information d\_eff derived from the
minimum subgroup size `n.min` and the observed censoring proportion.
This is a **model-based** reference value computed independently of
the simulation — the implied L\_eff below quantifies how many
effectively independent candidate subgroups the procedure is
evaluating, obtained by matching this P1 against the simulated FPR.

We first cross-check the censoring rate from two independent sources:

- **DGM reference**: one large n = 10,000 draw from `dgm_gbsg_null`
  via `simulate_from_dgm()` (computed in the setup chunk).  This is
  a low-variance Monte Carlo estimate of the DGM's theoretical
  censoring rate at the current (`analysis_time_dgm`,
  `cens_adjust_dgm`) settings — independent of any ForestSearch
  call.
- **Empirical**: the mean of `p.cens` across the calibration batch
  replicates (each at n = `r nrow(df.analysis)`).

Agreement between the two (difference < 0.02 = 2 pp) confirms that
the DGM's censoring model is internally consistent at both sample
sizes.  We use the empirical rate for downstream d\_eff / P1
calculations since it reflects the exact `n` and seed handling used
in the calibration batch.

```{r fpr-asymptotic-block}
#| code-fold: true

# Empirical censoring proportion from the calibration batch.
# run_simulation_analysis() records p.cens = 1 - mean(event_sim) for
# each replicate, so we average across replicates to get the rate
# the DGM actually produces at n = nrow(df.analysis) with the
# current (analysis_time_dgm, cens_adjust_dgm) settings.
prop_cens_emp <- mean(results_null$p.cens, na.rm = TRUE)

# Cross-check: compare the empirical rate (averaged over the
# calibration batch at n = 686) against the DGM-reference rate
# (single n = 10000 draw computed in the setup chunk).  They should
# agree up to Monte Carlo noise — disagreement would indicate a
# sample-size-dependent censoring effect worth investigating.
prop_cens_diff <- abs(prop_cens_emp - prop_cens_ref)

cat("Censoring rate check:\n")
cat(sprintf("  DGM reference:  %.3f  (n = 10000 single draw)\n",
            prop_cens_ref))
cat(sprintf("  Empirical:      %.3f  (averaged over %d replicates at n = %d)\n",
            prop_cens_emp,
            sum(!is.na(results_null$p.cens)),
            nrow(df.analysis)))
cat(sprintf("  |difference|:   %.3f  (expected < 0.02 for agreement)\n\n",
            prop_cens_diff))

# Per-subgroup asymptotic P1 at (c1, c2, d_eff) under H0.
# Use the empirical censoring rate since it reflects the n, analysis_time,
# and cens_adjust used in the calibration batch directly.
d_eff_est <- forestsearch::d_eff_survival(n_min_val, prop_cens_emp)
P1_est    <- tryCatch(
  forestsearch::compute_detection_probability_glm(
    theta = 1.0, d_eff = d_eff_est,
    c1 = c1_paper, c2 = c2_paper,
    effect_scale = "ratio", method = "monte_carlo"),
  error = function(e) NA_real_)

cat("Asymptotic approximation (León 2024, eq. 3):\n")
cat(sprintf("  d_eff       = %.1f  (n.min = %d, prop_cens = %.3f)\n",
            d_eff_est, n_min_val, prop_cens_emp))
cat(sprintf("  P1(theta=1) = %.1f%%  (per-subgroup asymptotic FPR)\n",
            100 * P1_est))
```


### Summary {#sec-appendix-summary}

```{r fpr-summary-table}
#| tbl-cap: "DGM-based FPR calibration at the GBSG operational thresholds. The Calibration batch estimates L_eff from the simulated FPR and the asymptotic P1; that L_eff is then carried forward to predict the Corrected FPR.  The Verification batch (independent seeds) provides an out-of-sample test: its Sim FPR should fall close to the Corrected FPR predicted from the Calibration row."

# Note: .wilson_ci() is defined in the fpr-setup chunk (hoisted for
# use across theta-grid verification and this summary chunk).

# ── Step 1: Calibration batch — estimate L_eff from the data ─────
rn_cal   <- results_null[results_null$analysis == "FSlg", ]
n_cal    <- nrow(rn_cal)
ndet_cal <- sum(rn_cal$any.H == 1L, na.rm = TRUE)
fpr_cal  <- ndet_cal / n_cal
ci_cal   <- .wilson_ci(ndet_cal, n_cal)

# This is the ESTIMATED parameter carried forward to the Corrected
# FPR prediction.  L_eff_cal is the pair-specific effective
# multiplicity derived by matching the calibration-batch simulated
# FPR against the asymptotic P1.
L_eff_cal <- if (!is.na(fpr_cal) && !is.na(P1_est) &&
                 fpr_cal > 0 && fpr_cal < 1 &&
                 P1_est > 0 && P1_est < 1)
  log(1 - fpr_cal) / log(1 - P1_est) else NA_real_

# Corrected FPR prediction:  1 − (1 − P1)^L_eff_cal
# This equals fpr_cal for the Calibration row (tautology by
# construction) and serves as the PREDICTION for the Verification
# row.  Both rows show the same value; the verification row's Sim
# FPR is the out-of-sample target.
fpr_corr_pred <- if (!is.na(P1_est) && !is.na(L_eff_cal))
  1 - (1 - P1_est)^L_eff_cal else NA_real_

# ── Step 2: Verification batch (optional) ─────────────────────────
if (!is.null(results_verify)) {
  rn_ver   <- results_verify[results_verify$analysis == "FSlg", ]
  n_ver    <- nrow(rn_ver)
  ndet_ver <- sum(rn_ver$any.H == 1L, na.rm = TRUE)
  fpr_ver  <- ndet_ver / n_ver
  ci_ver   <- .wilson_ci(ndet_ver, n_ver)
} else {
  n_ver <- NA_integer_; fpr_ver <- NA_real_
  ci_ver <- c(NA_real_, NA_real_)
}

# ── Step 3: Build the two-row (or one-row) summary table ──────────
fmt_pct <- function(x) {
  if (is.na(x)) "—" else sprintf("%.1f%%", 100 * x)
}
fmt_ci <- function(lo, hi) {
  if (any(is.na(c(lo, hi)))) "—"
  else sprintf("[%.1f%%, %.1f%%]", 100 * lo, 100 * hi)
}

row_cal <- data.table(
  Batch            = sprintf("Calibration (%d sims)", n_cal),
  Sims             = n_cal,
  `Sim FPR`        = fmt_pct(fpr_cal),
  `95% CI`         = fmt_ci(ci_cal[1], ci_cal[2]),
  `Implied L_eff`  = if (!is.na(L_eff_cal)) sprintf("%.3f", L_eff_cal) else "—",
  `Corrected FPR`  = fmt_pct(fpr_corr_pred),
  fpr_num          = fpr_cal
)

if (!is.null(results_verify)) {
  row_ver <- data.table(
    Batch            = sprintf("Verification (%d sims)", n_ver),
    Sims             = n_ver,
    `Sim FPR`        = fmt_pct(fpr_ver),
    `95% CI`         = fmt_ci(ci_ver[1], ci_ver[2]),
    `Implied L_eff`  = "— (carried forward)",
    `Corrected FPR`  = fmt_pct(fpr_corr_pred),
    fpr_num          = fpr_ver
  )
  summary_tbl <- rbind(row_cal, row_ver)
} else {
  summary_tbl <- row_cal
}

# Out-of-sample diagnostic: is fpr_ver within the corrected FPR's
# CI?  Report the percentage-point gap in the source note.
ver_gap_pp <- if (!is.null(results_verify) &&
                  !is.na(fpr_ver) && !is.na(fpr_corr_pred)) {
  100 * abs(fpr_ver - fpr_corr_pred)
} else {
  NA_real_
}

# ── Step 4: Extract best estimates for downstream plotting ────────
# Use verification when available (higher precision); otherwise
# calibration.  L_eff_implied used by the detection curve plot is
# always the calibration estimate (L_eff_cal).
if (!is.null(results_verify) && !is.na(fpr_ver)) {
  fpr_hat <- fpr_ver; fpr_ci <- ci_ver; n_reps <- n_ver
} else {
  fpr_hat <- fpr_cal; fpr_ci <- ci_cal; n_reps <- n_cal
}
L_eff_implied <- L_eff_cal  # always the calibration estimate

# ── Step 5: Render the table ──────────────────────────────────────
src_note <- paste(
  "Null replicates simulated from setup_gbsg_dgm(model = 'null').",
  "FPR = mean(any.H) across replicates (FSlg: LASSO + GRF screening).",
  "Calibration batch: seed_base = sim_seed_fpr + 2000L.",
  "Verification batch: seed_base = sim_seed_fpr + 50000L (disjoint seed range).",
  "Implied L_eff = log(1 − Sim FPR) / log(1 − P1) from the Calibration batch",
  "— this is the pair-specific effective multiplicity, used as the estimated",
  "parameter for the Corrected FPR prediction.",
  "Corrected FPR = 1 − (1 − P1)^L_eff_implied.  For the Calibration row this",
  "equals Sim FPR by construction (tautology).  For the Verification row it",
  "is the PREDICTION carried forward from the Calibration batch;",
  "agreement with the Verification Sim FPR confirms the correction works out",
  "of sample."
)
if (!is.na(ver_gap_pp)) {
  src_note <- paste(
    src_note,
    sprintf("Out-of-sample gap: |Sim FPR (verify) − Corrected FPR (predicted)| = %.1f pp.",
            ver_gap_pp)
  )
}

gt(summary_tbl) |>
  tab_header(
    title    = sprintf(
      "GBSG: DGM-based FPR Calibration | c1 = %.2f, c2 = %.2f | pcons = %.2f, stop = %.2f",
      c1_paper, c2_paper, pcons_val, stop_thresh_val),
    subtitle = sprintf(
      "N = %d | n.min = %d | P1(asymp) = %s",
      nrow(df.analysis), n_min_val,
      if (!is.na(P1_est)) sprintf("%.1f%%", 100 * P1_est) else "—")
  ) |>
  cols_hide(columns = fpr_num) |>
  tab_source_note(src_note)
```


## Detection curves at the operational threshold pair {#sec-appendix-detection}

The detection curve below combines two views of the detection probability
as a function of the true harm-subgroup HR θ†(H):

- For **θ ≤ `r theta_merge_cut` (null region)**: the corrected prediction
  `1 − (1 − P1(θ))^L_eff_cal`, using the pair-specific effective
  multiplicity estimated from the calibration batch at θ = 1.
- For **θ ≥ `r theta_merge_cut` (alt region)**: the uncorrected
  per-subgroup asymptotic P1(θ) from the León (2024) equation-3
  approximation, which empirically tracks power well when the
  subgroup effect is strong enough to be reliably detected.

**Both predictions are shown at θ = `r theta_merge_cut`**, so the
reader can compare the corrected and uncorrected values directly
against the verification diamond at that θ — it is plausible that the
uncorrected P1 is already the more accurate predictor at the boundary.

The choice of `r theta_merge_cut` as the cut-off between the two
regions is a **working hypothesis** — we do not claim any procedural
or theoretical significance for that particular value.  Verification
simulations at θ ∈ {`r paste(sort(c(theta_grid_verify, 1.0)), collapse = ", ")`}
(diamonds with Wilson 95% CIs) empirically test the merged curve.
Agreement within CIs across the null region supports the
"transportability" of `L_eff_cal`; systematic departure would indicate
the correction should be re-estimated θ-by-θ rather than carried
forward from θ = 1.

```{r plot-fpr-calibration-fn}
#| code-fold: true

# Local wrapper (to be exported from forestsearch in a future release).
# Builds a MERGED detection curve: corrected prediction for theta <= theta_cut,
# uncorrected per-subgroup P1 for theta >= theta_cut.  Both segments INCLUDE
# the boundary so the reader can compare two predictions (corrected and
# uncorrected) directly against the verification diamond at theta = theta_cut.
# Verification diamonds are overlaid from an optional data.table with columns
# theta / fpr / ci_lo / ci_hi.
plot_fpr_calibration <- function(cal_result,
                                  theta_seq      = seq(0.5, 3.0, by = 0.05),
                                  theta_cut      = 1.25,
                                  theta_verify   = NULL,
                                  fpr_target     = NULL,
                                  label          = NULL,
                                  effect_label   = "Effect ratio (theta)") {
  stopifnot(inherits(cal_result, "fpr_calibration"))

  c1       <- cal_result$c1
  c2       <- cal_result$c2
  d_eff    <- cal_result$d_eff
  L_eff    <- cal_result$L_eff_implied
  fpr_hat  <- cal_result$fpr
  ci_lo    <- cal_result$fpr_ci[1]
  ci_hi    <- cal_result$fpr_ci[2]
  N        <- cal_result$N
  nsims    <- cal_result$nsims

  # ── Build the two raw curves ─────────────────────────────────────
  p1_vec <- sapply(theta_seq, function(th)
    tryCatch(
      compute_detection_probability_glm(
        theta = th, d_eff = d_eff, c1 = c1, c2 = c2,
        effect_scale = "ratio", method = "monte_carlo"),
      error = function(e) NA_real_))

  p_corr_vec <- if (!is.na(L_eff)) {
    1 - (1 - p1_vec)^L_eff
  } else {
    rep(NA_real_, length(theta_seq))
  }

  # ── Merge: corrected for theta <= cut, uncorrected for theta >= cut ─────
  # Both segments INCLUDE theta_cut so the reader sees two endpoints at
  # the merge boundary: the corrected prediction (red, ending at cut)
  # and the uncorrected P1 (grey, starting at cut).  This lets the
  # reader compare both model predictions directly against the
  # verification diamond at theta = theta_cut.
  null_idx <- theta_seq <= theta_cut
  alt_idx  <- theta_seq >= theta_cut

  curve_dt <- rbind(
    data.frame(theta = theta_seq[null_idx],
               prob  = p_corr_vec[null_idx],
               curve = sprintf("Corrected prediction (θ ≤ %.2f)", theta_cut)),
    data.frame(theta = theta_seq[alt_idx],
               prob  = p1_vec[alt_idx],
               curve = sprintf("Per-subgroup P1 (θ ≥ %.2f)", theta_cut))
  )

  # ── Verification diamonds ────────────────────────────────────────
  # Start with the calibration-batch θ = 1 result (always present),
  # then append any theta_verify data.table supplied by the caller.
  sim_pts <- data.frame(
    theta = 1, prob = fpr_hat, ci_lo = ci_lo, ci_hi = ci_hi
  )
  if (!is.null(theta_verify) && nrow(theta_verify) > 0L) {
    sim_pts <- rbind(
      sim_pts,
      data.frame(
        theta = theta_verify$theta,
        prob  = theta_verify$fpr,
        ci_lo = theta_verify$ci_lo,
        ci_hi = theta_verify$ci_hi
      )
    )
  }

  sub_txt <- sprintf(
    "c1=%.2f, c2=%.2f | d_eff=%.1f | L_eff(calibration)=%.3f | N=%d | calibration sims=%d",
    c1, c2, d_eff,
    if (!is.na(L_eff)) L_eff else 0,
    N, nsims)
  if (!is.null(label)) sub_txt <- paste0(label, " | ", sub_txt)

  p <- ggplot(curve_dt, aes(x = theta, y = prob,
                             colour = curve, linetype = curve)) +
    geom_line(linewidth = 0.9, na.rm = TRUE) +
    geom_vline(xintercept = 1, linetype = "dotted",
               colour = "grey60", linewidth = 0.4) +
    geom_vline(xintercept = theta_cut, linetype = "dashed",
               colour = "grey40", linewidth = 0.5) +
    annotate("text",
             x = theta_cut, y = 0.02,
             label = sprintf("θ = %.2f (merge boundary)", theta_cut),
             hjust = -0.05, vjust = 0, size = 3, colour = "grey40") +
    geom_point(data = sim_pts, aes(x = theta, y = prob),
               colour = "steelblue", size = 3.5, shape = 18,
               inherit.aes = FALSE) +
    geom_errorbar(data = sim_pts,
                  aes(x = theta, ymin = ci_lo, ymax = ci_hi),
                  colour = "steelblue", width = 0.04, linewidth = 0.7,
                  inherit.aes = FALSE) +
    scale_colour_manual(values = setNames(
      c("firebrick", "grey50"),
      c(sprintf("Corrected prediction (θ ≤ %.2f)", theta_cut),
        sprintf("Per-subgroup P1 (θ ≥ %.2f)", theta_cut))
    )) +
    scale_linetype_manual(values = setNames(
      c("solid", "dashed"),
      c(sprintf("Corrected prediction (θ ≤ %.2f)", theta_cut),
        sprintf("Per-subgroup P1 (θ ≥ %.2f)", theta_cut))
    )) +
    scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
    labs(x        = effect_label,
         y        = "Detection probability",
         title    = sprintf("Merged Detection Curve: c1 = %.2f, c2 = %.2f",
                            c1, c2),
         subtitle = sub_txt,
         colour   = NULL, linetype = NULL) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top",
          plot.title    = element_text(face = "bold"),
          plot.subtitle = element_text(size = 9, colour = "grey40"))

  if (!is.null(fpr_target)) {
    p <- p +
      geom_hline(yintercept = fpr_target, linetype = "dashed",
                 colour = "darkgreen", linewidth = 0.5) +
      annotate("text",
               x = min(theta_seq) + 0.05, y = fpr_target + 0.01,
               label = sprintf("FPR target = %.0f%%", 100 * fpr_target),
               hjust = 0, size = 3, colour = "darkgreen")
  }
  p
}
```

```{r fpr-detection-curves}
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "Merged DGM-based detection curve. Red solid = corrected FPR prediction (θ ≤ theta_merge_cut); grey dashed = uncorrected per-subgroup P1 (θ ≥ theta_merge_cut). Both predictions are shown at θ = theta_merge_cut so the reader can compare them directly against the verification diamond at that θ. Blue diamonds with Wilson 95% CIs = independent DGM simulations (calibration batch at θ = 1; theta-grid verification at θ ∈ theta_grid_verify). Agreement of diamonds with the corrected prediction across the null region empirically tests the transportability of L_eff_cal; at θ = theta_merge_cut, visual proximity of the diamond to either the red or grey endpoint indicates which model is the more accurate predictor at that boundary."

# Synthesize an fpr_calibration-shaped object from the run_simulation_analysis
# results so the plot wrapper can render the detection curve without change.
cal_for_plot <- structure(
  list(
    fpr           = fpr_hat,
    fpr_ci        = fpr_ci,
    nsims         = n_reps,
    P1            = P1_est,
    d_eff         = d_eff_est,
    L_eff_implied = L_eff_implied,
    c1            = c1_paper,
    c2            = c2_paper,
    N             = nrow(df.analysis)
  ),
  class = "fpr_calibration"
)

plot_fpr_calibration(
  cal_for_plot,
  theta_cut    = theta_merge_cut,
  theta_verify = theta_verify_df,
  fpr_target   = fpr_target_surv,
  effect_label = "True harm-subgroup HR (θ†(H))",
  label        = sprintf("DGM | N = %d | prop_cens ~ %.2f",
                          nrow(df.analysis), prop_cens_emp)
)
```


## Interpretation {#sec-appendix-interp}

**Reading the result.**  The summary table in @sec-appendix-summary
follows an **estimate-then-predict** structure at the main-analysis
operating point — threshold pair
(`r c1_paper`, `r sprintf("%.2f", c2_paper)`) with
`pconsistency.threshold = `r pcons_val`` and
`stop_threshold = `r stop_thresh_val``:

- **Calibration row** (`r nsims_tier1_fpr` sims, seed offset +2000) —
  supplies the **estimated parameter**.  The Sim FPR from this batch
  is matched against the asymptotic P1 from @sec-appendix-asymp to
  derive the implied L\_eff:
  `L_eff = log(1 − Sim FPR) / log(1 − P1)`.  The Corrected FPR in
  this row (`1 − (1 − P1)^L_eff`) equals the Sim FPR by construction
  — it's a tautological check that the round-trip is consistent.
- **Verification row** (`r nsims_verify_fpr` sims, seed offset
  +50000, disjoint seed range) — provides an **out-of-sample test**.
  Its Sim FPR is an independent simulation of the same quantity,
  while its Corrected FPR is the **prediction carried forward** from
  the Calibration row's L\_eff (same numeric value in both rows).
  Agreement between the Verification Sim FPR and the Corrected FPR
  prediction confirms the correction generalises beyond the
  calibration sample.

Set `nsims_verify_fpr = 0L` to skip the verification; scale up
(e.g.\ 10,000–20,000) for ±0.2–0.3 pp precision on a final
production result.  Wilson 95\% CI widths scale as
$1/\sqrt{N_{\text{sims}}}$.

**Operating characteristics table.**  @sec-appendix-oc shows the full
`format_oc_results()` output from the calibration batch.  The `any.H`
entry equals the calibration-batch Sim FPR; sens / ppv are null by
construction (the unconditional averaging convention from the paper —
the true harm subgroup is empty under H0).

**Asymptotic P1 and implied L\_eff.**  @sec-appendix-asymp reports P1
and d\_eff from the León (2024) equation-3 approximation — a
model-based per-subgroup reference computed independently of the
simulation.  The implied L\_eff is the pair-specific effective
multiplicity that reconciles the simulated FPR with the asymptotic
P1, quantifying how many effectively independent candidate subgroups
the procedure is evaluating after LASSO + GRF + two-stage screening
with `maxk = `r maxk_val``.  For survival with events-based filters
(`d0.min`, `d1.min`), L\_eff is typically close to or below 1 when
the filters prune the candidate pool tightly; values well above 1
indicate the candidate pool is larger than a single test.

**Merged detection curve and transportability (@sec-appendix-detection).**
The detection curve in @sec-appendix-detection combines two regimes:
the **corrected prediction** `1 − (1 − P1(θ))^L_eff_cal` for
θ ≤ `r theta_merge_cut` (null region), and the **uncorrected
per-subgroup P1(θ)** for θ ≥ `r theta_merge_cut` (alt region).  Both
predictions are shown at the boundary θ = `r theta_merge_cut`, so the
reader can compare them directly against the verification diamond at
that θ — the corrected and uncorrected curves typically differ by a
non-trivial gap at the boundary, and which one sits closer to the
empirical diamond indicates the more accurate predictor in that
neighbourhood.  The value `r theta_merge_cut` itself is a
working-hypothesis guess for where the corrected prediction stops
being the right thing to show; no procedural or theoretical
significance is claimed for it.  Independent verification simulations
at θ†(H) ∈
{`r paste(sort(c(theta_grid_verify, 1.0)), collapse = ", ")`}
(diamonds with Wilson 95% CIs) test the transportability of
`L_eff_cal` across the null region.  If those diamonds fall on the
corrected curve within CI width, the assumption that a single `L_eff`
estimated at θ = 1 transports across
θ ∈ [min(theta_grid_verify), `r theta_merge_cut`] is empirically
supported.  Systematic departure would indicate the correction should
be re-estimated θ-by-θ rather than carried forward — or that the
merge cut-off should be moved earlier, if the uncorrected P1 is
already the more accurate predictor below `r theta_merge_cut`.

**Interpreting the identified ER ≤ 0 subgroup.**  The FPR figure
above is the procedure-level Type-I error rate: the probability that
ForestSearch identifies *any* subgroup under H0 at these operational
settings.  It is not subgroup-specific and should be interpreted as
a property of the analysis procedure rather than of the specific
subgroup found.  The biological plausibility of the tamoxifen /
ER-status mechanism supports the finding, and the calibrated FPR
quantifies the Type-I error of the identification procedure to
complement the clinical evidence.


# Computational Timing {#sec-timing}

```{r timing-summary}
# Local null-coalescing operator (the one in forestsearch is unexported)
`%||%` <- function(a, b) if (!is.null(a)) a else b

timings$total <- (proc.time() - t_vignette_start)["elapsed"]

# Sum FPR sub-components into a single row for display
timings$fpr_total <- sum(c(
  timings$fpr_runsim       %||% 0,  # calibration batch
  timings$fpr_verify       %||% 0,  # verification batch at theta = 1
  timings$fpr_theta_verify %||% 0   # theta-grid verification
))

timing_df <- data.frame(
  Analysis = c("GRF", "ForestSearch", "Bootstrap",
               "kfold CV", "OOB CV",
               "FPR calibration (appendix)", "Total"),
  Seconds  = c(
    timings$grf,
    timings$forestsearch,
    timings$bootstrap,
    timings$kfold,
    timings$oob,
    if (timings$fpr_total > 0) timings$fpr_total else NA_real_,
    timings$total
  )
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(title = "Computational Timing") |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1,
             rows = !is.na(Seconds)) |>
  cols_label(
    Analysis = "Component",
    Seconds  = "Time (sec)",
    Minutes  = "Time (min)"
  ) |>
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_body(rows = Analysis == "Total"))
```


# References {#sec-refs}

León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024).
"Exploratory subgroup identification in the heterogeneous Cox model:
A relatively simple procedure." *Statistics in Medicine*.
DOI: [10.1002/sim.10163](https://doi.org/10.1002/sim.10163)


# Session Info {#sec-session}

```{r session-info}
plan(sequential)
sessionInfo()
```
