<!--
  Simulation estimation/coverage table (manuscript fragment).  Recreates the
  Section 8 table of the fs_t1_t2 operating-characteristics document for the
  ENUMERATED consistency search, in the manuscript reader style of
  _actg175_section6_2.qmd / _gbsg_section6_1a.qmd.

  Reads the combined results bundle produced by the simulation
  (fs_t1_t2_*_combined_*.rds: a list with `results`, `truth`, `meta`).
  FB = full nonparametric bootstrap (Leon Eq. 7); MR = refit-free
  de-biased gate (infinitesimal-jackknife interval).  Defines @tbl-sim-estimation-fs-h15-n1500.

  The identified-subgroup size n is constant within each block (it is the count
  of detected replicates), so it is reported in the caption rather than as a
  repeated column.  Keep the .rds beside this fragment or set `payload_dir`;
  when absent a TBD skeleton renders so the cross-reference still resolves.
-->

```{r sim-estimation-setup-fs-h15-n1500}
#| echo: false
#| message: false
#| warning: false
library(kableExtra)

payload_dir <- "_payloads/"
# Payload candidates, CURRENT stem first, legacy t1_t2 stem second: the first
# file that exists is used, so dropping a new-era bundle beside this fragment
# repoints it with no edit.
rds_candidates <- c("fs_maxeffCons_fb_mr_m1_h15_knoise3_n1500_combined_1_500.rds",
                   "fs_t1_t2_m1_h15_knoise3_n1500_combined_1_500.rds")
rds_file    <- rds_candidates[1]   # display name for the placeholder caption
show_len_med <- FALSE   # TRUE adds the median-CI-length column Len(med); default omits it

.pdir <- if (is.null(payload_dir)) {
  tryCatch(dirname(knitr::current_input(dir = TRUE)), error = function(e) getwd())
} else payload_dir
.paths    <- file.path(.pdir, rds_candidates)
.rds_path <- if (any(file.exists(.paths))) .paths[file.exists(.paths)][1] else .paths[1]
have_sim  <- file.exists(.rds_path)
if (have_sim) rds_file <- basename(.rds_path)

z975 <- qnorm(0.975)
`%||%` <- function(a, b) if (is.null(a) || length(a) == 0) b else a

# ---- payload compatibility -------------------------------------------------
# The analysis side renamed the resampling columns across eras:
#   t1_*  ->  fb_*   (full bootstrap, Leon Eq. 7)
#   t2_*  ->  mr_*   (multiplier resampling, IJ interval)
# and meta$gate_draws -> meta$mr_draws.  Resolve per bundle: try each candidate
# column name in order; a fully absent estimator yields an all-NA numeric vector
# (rendered as em-dashes) instead of NULL, which log() rejects.
.getcol <- function(r, nms) {
  for (nm in nms) { v <- r[[nm]]; if (!is.null(v)) return(as.numeric(v)) }
  rep(NA_real_, nrow(r))
}
.mfin  <- function(x) { x <- x[is.finite(x)]; if (length(x)) mean(x) else NA_real_ }
.sdfin <- function(x) { x <- x[is.finite(x)]; if (length(x) > 1L) sd(x) else NA_real_ }
.cover <- function(target, lo, hi) {
  target <- rep_len(target, length(lo))
  ok <- is.finite(target) & is.finite(lo) & is.finite(hi)
  if (!any(ok)) return(NA_real_)
  mean(target[ok] >= lo[ok] & target[ok] <= hi[ok])
}
.pbias <- function(est, target) {
  d <- (est - target) / target; d <- d[is.finite(d)]
  if (!length(d)) return(NA_real_); 100 * mean(d)
}
# Empirical (true) SD on the log-HR (beta) scale: SD of log(HR) across reps.
# Guarded so a schema change degrades to an em-dash, not a halt.
.sd_log <- function(e) {
  if (is.null(e) || !length(e)) return(NA_real_)
  l <- log(e[is.finite(e) & e > 0]); if (length(l) > 1L) sd(l) else NA_real_
}
# Stored per-replicate log-HR (beta-scale) SE for an estimator, read at the source
# (or_/nv_/t1_*_se; MR uses its IJ SE, t2_*_se_ij) rather than back-calculated from
# the CI bounds. A missing column yields all-NA so the table still renders.
.get_se <- function(r, pks, suffix, ij) {
  .getcol(r, paste0(pks, "_", suffix, if (ij) "_se_ij" else "_se"))
}

# One block (H or Hc): four estimator rows against three targets.
# Estimator labels: FB (full bootstrap), MR (multiplier resampling).
build_block <- function(r, suffix, tgt_cde, tgt_marg) {
  keys   <- list(oracle = "or", naive = "nv",
                 `FB` = c("fb", "t1"), `MR` = c("mr", "t2"))
  or_est <- .getcol(r, paste0("or_", suffix, "_est"))
  bH     <- .getcol(r, paste0("betaHhat_", suffix))              # per-rep beta(Hhat)
  do.call(rbind, lapply(names(keys), function(lab) {
    pks <- keys[[lab]]
    e   <- .getcol(r, paste0(pks, "_", suffix, "_est"))
    lo  <- .getcol(r, paste0(pks, "_", suffix, "_lo"))
    hi  <- .getcol(r, paste0(pks, "_", suffix, "_hi"))
    len <- hi - lo; len <- len[is.finite(len)]
    seb <- .get_se(r, pks, suffix, ij = identical(lab, "MR"))  # stored log-HR (beta) SE
    data.frame(
      Estimator  = lab,
      Avg        = .mfin(e),
      SD         = .sd_log(e),                     # empirical SD, log-HR (beta) scale  [reported]
      SD_hr      = .sdfin(e),                      # empirical SD, HR scale             [retained]
      SD_hat     = .mfin(seb),                     # mean stored model SE, log-HR scale [reported]
      SD_hat_hr  = .mfin(e * seb),                 # mean model SE, HR scale            [retained]
      Len_med    = if (length(len)) stats::median(len) else NA_real_,
      Length     = if (length(len)) mean(len) else NA_real_,
      b_oracle   = .pbias(e, or_est),
      b_betaHhat = .pbias(e, bH),
      b_cde      = .pbias(e, tgt_cde),
      b_marg     = .pbias(e, tgt_marg),
      Cov_oracle = .cover(or_est,  lo, hi),
      Cov_betaHhat = .cover(bH, lo, hi),
      Cov_cde    = .cover(tgt_cde, lo, hi),
      Cov_marg   = .cover(tgt_marg, lo, hi),
      stringsAsFactors = FALSE)
  }))
}

if (have_sim) {
  B <- readRDS(.rds_path); res <- B$results; truth <- B$truth; meta <- B$meta
  rH      <- res[res$detected %in% 1L, , drop = FALSE]  # %in%: NA-safe (CONFIG-ERROR rows)
  blockH  <- build_block(rH, "H",  truth$cde_H,  truth$marg_H)
  blockHc <- build_block(rH, "Hc", truth$cde_Hc, truth$marg_Hc)

  n_det      <- nrow(rH)
  n_sims     <- meta$n_sims
  n_sample   <- meta$n_sample
  nb_boots   <- meta$nb_boots
  mr_draws   <- meta$mr_draws %||% meta$gate_draws   # renamed gate_* -> mr_*; accept either vintage
  det_rate   <- mean(res$detected, na.rm = TRUE)
  meanH      <- mean(rH$n_harm, na.rm = TRUE)
  meanHc     <- n_sample - meanH
  lab_H  <- sprintf("Harm \u0124 (mean |\u0124| = %.0f, %.1f%% ITT)",
                    meanH, 100 * meanH / n_sample)
  lab_Hc <- sprintf("Complement \u0124\u1d9c (mean |\u0124\u1d9c| = %.0f, %.1f%% ITT)",
                    meanHc, 100 * meanHc / n_sample)
  cls <- sprintf("sens %.2f, spec %.2f, PPV %.2f, NPV %.2f",
                 .mfin(rH$sens), .mfin(rH$spec), .mfin(rH$ppv), .mfin(rH$npv))
  bH_H  <- rH$betaHhat_H[is.finite(rH$betaHhat_H)]
  betaH_H_mean <- mean(bH_H); betaH_H_med <- stats::median(bH_H); betaH_H_sd <- stats::sd(log(bH_H))
  # Pre-format the numbers the PROSE cites so a hand-typed figure can never drift
  # from the payload (matches the main-text _sim_estimation_fs_h10.qmd convention).
  det_str       <- sprintf("%.0f%%", 100 * det_rate)
  thH_str <- sprintf("%.2f", truth$marg_H); thcdeH_str <- sprintf("%.2f", truth$cde_H)
  nv_bO <- blockH$b_oracle[blockH$Estimator == "naive"]; nv_bM <- blockH$b_marg[blockH$Estimator == "naive"]
  nv_bB <- blockH$b_betaHhat[blockH$Estimator == "naive"]
  mr_bO <- blockH$b_oracle[blockH$Estimator == "MR"]; mr_bM <- blockH$b_marg[blockH$Estimator == "MR"]
  mr_bC <- blockH$b_cde[blockH$Estimator == "MR"]; mr_bB <- blockH$b_betaHhat[blockH$Estimator == "MR"]
  fb_bO <- blockH$b_oracle[blockH$Estimator == "FB"]; fb_bM <- blockH$b_marg[blockH$Estimator == "FB"]
  mr_cB <- blockH$Cov_betaHhat[blockH$Estimator == "MR"]
  nv_oracle_str <- sprintf("%.0f%%", abs(nv_bO)); nv_beta_str <- sprintf("%.0f%%", abs(nv_bB))
  nv_marg_str   <- sprintf("%.0f%%", abs(nv_bM))
  mr_oracle_str <- sprintf("%+.1f%%", mr_bO);      mr_marg_str <- sprintf("%+.1f%%", mr_bM)
  mr_cde_str    <- sprintf("%+.1f%%", mr_bC);      mr_beta_str <- sprintf("%+.1f%%", mr_bB)
  fb_marg_str   <- sprintf("%+.1f%%", fb_bM);      mr_covbeta_str <- sprintf("%.2f", mr_cB)
} else {
  .skel <- function() data.frame(
    Estimator = c("oracle", "naive", "FB", "MR"),
    Avg = NA_real_, SD = NA_real_, SD_hr = NA_real_, SD_hat = NA_real_, SD_hat_hr = NA_real_, Len_med = NA_real_, Length = NA_real_,
    b_oracle = NA_real_, b_betaHhat = NA_real_, b_cde = NA_real_, b_marg = NA_real_,
    Cov_oracle = NA_real_, Cov_betaHhat = NA_real_, Cov_cde = NA_real_, Cov_marg = NA_real_, stringsAsFactors = FALSE)
  blockH <- .skel(); blockHc <- .skel()
  # Skeleton mode: no payload, so no numbers — placeholders only, never
  # plausible-looking counts that could pass for results in a draft render.
  n_det <- NA_integer_; n_sims <- NA_integer_; n_sample <- NA_integer_
  nb_boots <- NA_integer_; mr_draws <- NA_integer_
  det_rate <- NA_real_; lab_H <- "Harm \u0124"; lab_Hc <- "Complement \u0124\u1d9c"; cls <- "\u2014"
  betaH_H_mean <- NA_real_; betaH_H_med <- NA_real_; betaH_H_sd <- NA_real_
  det_str <- "TBD"; nv_oracle_str <- "TBD"; nv_beta_str <- "TBD"; nv_marg_str <- "TBD"
  thH_str <- "TBD"; thcdeH_str <- "TBD"
  mr_oracle_str <- "TBD"; mr_marg_str <- "TBD"; mr_cde_str <- "TBD"; mr_beta_str <- "TBD"
  fb_marg_str <- "TBD"; mr_covbeta_str <- "TBD"
}

# n (finite reps contributing) is constant within each block, so it is reported
# here in the caption rather than as a repeated column.
if (have_sim) {
 cap_sim <- paste0(
  "Estimation and coverage for forest search over the ", format(n_det, big.mark = ","),
  " detected replicates (", sprintf("%.0f%% of %s simulations", 100 * det_rate, format(n_sims, big.mark = ",")),
  "; n = ", format(n_sample, big.mark = ","), "). ",
  sprintf("Scenario hazard-ratio effects $\\theta^{\\dagger}(H) = %.2f$, $\\theta^{\\dagger}(H^{c}) = %.2f$, $\\theta^{\\ddagger}(H) = %.2f$, $\\theta^{\\ddagger}(H^{c}) = %.2f$; ", truth$marg_H, truth$marg_Hc, truth$cde_H, truth$cde_Hc),
  sprintf("the conditional estimand $\\theta(\\widehat H)$ has mean %.2f, median %.2f (hazard-ratio scale) and SD %.2f ($\\beta$ scale). ", betaH_H_mean, betaH_H_med, betaH_H_sd),
  "Classification of $\\widehat H$ vs the true harm subgroup: ", cls, ". ",
  "Estimators, targets, and SD scales are defined with the main-text estimation tables (Section 7).")} else {
  cap_sim <- paste0(
  "Estimation and coverage for the enumerated consistency search. ",
  "**Placeholder \u2014 to be populated from the combined results bundle (`", rds_file, "`).**")
}
```

## Forest search under genuine harm: estimation and coverage {#sec-sim-estimation-fs-h15-n1500}

This design point moves from the borderline null to a genuine harm signal ($\theta^{\dagger}(H) = `r thH_str`$, $\theta^{\ddagger}(H) = `r thcdeH_str`$) at $n = 1{,}500$, with forest search on its fixed enumerated family, so the FB and MR de-biased procedures target the *same* estimand. @tbl-sim-estimation-fs-h15-n1500 reports estimation and coverage where forest search identified a subgroup in `r det_str` of the `r format(n_sims, big.mark = ",")` simulations; with a real effect present the identifier now recovers the true membership far more reliably than at the null (see the classification summary in the caption). The winner's curse weakens as the signal strengthens: the naive plug-in overstates the marginal hazard ratio by only `r nv_marg_str` here, against the much larger overstatement at the null (@tbl-sim-estimation-fs-h10). Scored against the marginal and controlled-direct-effect targets, multiplier resampling tips mildly negative (MR bias `r mr_marg_str` against the marginal and `r mr_cde_str` against the CDE) -- an artifact of scoring a selected region against a fixed *true*-subgroup target. Against the conditional estimand $\theta(\widehat H)$ that Theorem 2 of the main paper characterizes, however, the same intervals are near-nominal: MR point bias is `r mr_beta_str` and its coverage of $\theta(\widehat H)$ is `r mr_covbeta_str`. Switching the scoring target to $\theta(\widehat H)$ thus resolves the apparent over-correction seen against true-subgroup targets, and is the cleanest illustration that the conditional estimand is the parameter the correction is built to cover; its stability across sample size is traced in @fig-mr-coverage-h15c.

```{r}
#| label: tbl-sim-estimation-fs-h15-n1500
#| tbl-cap: !expr cap_sim
#| echo: false
#| message: false
#| warning: false
# Column headers as inline math via \( \): rendered by MathJax (HTML) and natively
# (LaTeX). Coverage uses a plain C (no hat), matching the bias b; SD carries a widehat.
h_sde <- "\\(\\text{SD}_{\\beta}\\)"
h_sd  <- "\\(\\widehat{\\text{SD}}_{\\beta}\\)"
h_bO <- "\\(b^{\\text{oracle}}\\)"; h_bB <- "\\(b^{\\theta}\\)"; h_bC <- "\\(b^{\\ddagger}\\)"; h_bM <- "\\(b^{\\dagger}\\)"
h_cO <- "\\(C^{\\text{oracle}}\\)"; h_cB <- "\\(C^{\\theta}\\)"; h_cC <- "\\(C^{\\ddagger}\\)"; h_cM <- "\\(C^{\\dagger}\\)"
# add_header_above() escapes these per format (% -> \% in LaTeX), so pass them plain.
sp_b <- "Relative bias (%)"; sp_c <- "95% coverage"
fmt2 <- function(x) ifelse(is.na(x), "\u2014", sprintf("%.2f", x))
fmtb <- function(x) ifelse(is.na(x), "\u2014", sprintf("%+.1f", x))
mk <- function(b) {
  d <- data.frame(
    Estimator = b$Estimator,
    Avg = fmt2(b$Avg), SD = fmt2(b$SD), SDh = fmt2(b$SD_hat),
    LenMed = fmt2(b$Len_med), Length = fmt2(b$Length),
    bO = fmtb(b$b_oracle), bB = fmtb(b$b_betaHhat), bC = fmtb(b$b_cde), bM = fmtb(b$b_marg),
    cO = fmt2(b$Cov_oracle), cB = fmt2(b$Cov_betaHhat), cC = fmt2(b$Cov_cde), cM = fmt2(b$Cov_marg),
    check.names = FALSE)
  if (!show_len_med) d$LenMed <- NULL
  d
}
tab <- rbind(mk(blockH), mk(blockHc))
loc_hdrs <- c("Estimator", "Avg", h_sde, h_sd, if (show_len_med) "Len(med)", "Length")
colnames(tab) <- c(loc_hdrs, h_bO, h_bB, h_bC, h_bM, h_cO, h_cB, h_cC, h_cM)
n_loc   <- length(loc_hdrs)                       # 6 with Len(med), 5 without
spanner <- setNames(c(n_loc, 4L, 4L), c(" ", sp_b, sp_c))
tbl <- kbl(tab, booktabs = TRUE, escape = FALSE,
           align = paste0("l", strrep("c", n_loc + 7L)), linesep = "") |>
  kable_styling(font_size = if (knitr::is_latex_output()) 10 else NULL) |>
  add_header_above(spanner) |>
  pack_rows(lab_H, 1, 4) |>
  pack_rows(lab_Hc, 5, 8)
# Rotate the (wide) table to landscape for PDF only; HTML stays in normal flow.
if (knitr::is_latex_output()) tbl <- landscape(tbl)
tbl
```

