---
title: "Bias vs coverage across the MR simulation cells"
subtitle: "One-sided and two-sided coverage of β(Ĥ) against residual bias, with the Gaussian reference (TASK_bias_coverage_display_2026-09-06)"
author: "Larry León"
date: today
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    code-fold: true
---

The display reads each simulation cell's (estimator, coverage) pair against a
Gaussian bookkeeping reference: with `b` = mean log error / empirical SD and
`r` = mean reported SE / empirical SD (both from the cell's own replicates), a
Gaussian error model predicts one-sided coverage $\Phi(z_{0.95} r - b)$ and
two-sided coverage $\Phi(z_{0.975} r - b) - \Phi(-z_{0.975} r - b)$. The
reference is bookkeeping, not theory: it says how much of each observed
coverage is explained by the (bias, SE-calibration) pair alone.

# Knobs

```{r setup, message=FALSE}
library(forestsearch)
library(ggplot2)
library(kableExtra)

# Campaign glob: which combined bundles to read (committed s7 + map1 cells).
# Self-contained: reads results/ beside this document only.
globs <- c("results/fs_*_s7_combined_1_2000.rds",
           "results/fs_*_map1_combined_1_2000.rds")
block      <- "H"                  # J1 default: harm block
estimators <- c("mr", "fld")       # IJ and field (naive available by knob)
curves     <- c(1, 1.25, 1.5, 2)   # J2 default

# Human-readable cell label from a bundle's meta.
cell_label <- function(meta) {
  hr <- meta$target_hr_harm
  base <- if (hr < 1) sprintf("protective %.2f", hr)
  else if (hr == 1) "null 1.0" else sprintf("harm %.4g", hr)
  kn <- if (!is.null(meta$k_random_noise) && meta$k_random_noise > 0)
    sprintf(", %d noise", meta$k_random_noise) else ""
  nn <- if (is.null(meta$n_sample) || meta$n_sample == 500)
    if (nzchar(kn) || hr < 1) "" else ", n=500" else sprintf(", n=%d", meta$n_sample)
  paste0(base, nn, kn)
}
```

# Cells

```{r read}
files <- sort(unlist(lapply(globs, Sys.glob)))
stopifnot(length(files) > 0)
tab <- do.call(rbind, lapply(files, function(f) {
  b <- readRDS(f)
  # meta may not carry target_hr_harm/k_random_noise (s7-era bundles); recover
  # them from the stem, which always does.
  m <- b$meta
  if (is.null(m$target_hr_harm))
    m$target_hr_harm <- as.numeric(sub("^.*_m1_h(\\d{3})_.*$", "\\1", basename(f))) / 100
  if (is.null(m$k_random_noise))
    m$k_random_noise <- as.integer(sub("^.*_knoise(\\d+)_.*$", "\\1", basename(f)))
  s <- fs_sim_bias_coverage(b$results, block = block, estimators = estimators)
  s$cell <- cell_label(m)
  s$bundle <- basename(f)
  s
}))
tab$estimator <- factor(tab$estimator, levels = c("naive", "mr", "fld"),
                        labels = c("naive", "IJ", "field"))
kbl(tab[, c("cell", "estimator", "n", "b", "r", "cov1", "cov1_wilson_lo",
            "cov1_wilson_hi", "cov2", "cov2_wilson_lo", "cov2_wilson_hi",
            "cov1_ref", "cov2_ref")],
    digits = 3, booktabs = TRUE, row.names = FALSE) |>
  kable_styling(font_size = 12)
```

# The display

```{r plot}
#| fig-width: 13
#| fig-height: 4.6
fs_plot_bias_coverage(tab, curves = curves)
```

# Reading

```{r reading, results='asis'}
#| echo: false
d1 <- tab$cov1 - tab$cov1_ref
d2 <- tab$cov2 - tab$cov2_ref
cat(sprintf(paste0(
  "Across the %d (cell, estimator) points, the Gaussian reference at each ",
  "point's own (b, r) explains the observed coverage to within %.3f ",
  "(one-sided, mean |gap| %.3f) and %.3f (two-sided, mean |gap| %.3f): the ",
  "coverage patterns in these cells are the arithmetic of retained bias and ",
  "SE calibration, not distributional exotica. Points below the nominal line ",
  "in the two-sided panel sit there because b is large relative to r (the ",
  "field rows at moderate separation), or because r has fallen to ~1 while b ",
  "remains (h1.5 at n=1500 -- the IJ row too); the one-sided panel's points ",
  "hug nominal because every estimator's (b, r) pair lands where ",
  "Φ(1.645r − b) ≈ 0.95.\n"),
  nrow(tab), max(abs(d1)), mean(abs(d1)), max(abs(d2)), mean(abs(d2))))
```
