---
title: "Decomposition of the complement's error variance across the regime sequence"
subtitle: "Naive error a, two-term correction c, de-biased error e = a − c, and the field's λ-SDᶜ, on the committed s7c / p30 / p30sg / nb20 bundles (TASK_complement_variance_banddial_2026-09-07, Part A)"
author: "Larry León"
date: today
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    code-fold: true
---

Per detected replicate, on the log scale: $a = \log(\hat\beta^c_{\text{nv}}) - \log\beta(\widehat H^c)$
(the naive error of the complement's Cox estimate against its own realized target),
$c = \log(\hat\beta^c_{\text{nv}}) - \log(\hat\beta^c_{\text{MR}})$ (the two-term correction the gate
subtracted), and $e = a - c = \log(\hat\beta^c_{\text{MR}}) - \log\beta(\widehat H^c)$ (the de-biased error).
Then $\mathrm{Var}(e) = \mathrm{Var}(a) + \mathrm{Var}(c) - 2\,\mathrm{Cov}(a, c)$, checked per cell, against the
three variances an analysis reports: the mean naive $\mathrm{SE}^2$, the mean field $\lambda\text{-SD}^{c\,2}$
(`fld_Hc_se`, the SD of $\Lambda^{*c}$ over the outer draws) and the mean IJ two-term $\mathrm{se}_{ij}^2$.
The complement's `selection_bias` / `fixed_bias` split of $c$ is not recorded in any bundle (the gate
returns the two components for the harm block only), so $c$ is read whole. Winner-only and winner-floor
rows are excluded (standing convention). Everything here reads the committed pooled bundles; nothing is
re-computed.

# Bundles and definitions

```{r setup, message=FALSE}
library(kableExtra)
`%||%` <- function(a, b) if (is.null(a) || length(a) == 0 || all(is.na(a))) b else a
globs <- strsplit(Sys.getenv("FS_SUMCV_GLOBS", paste(
  "results/fs_maxeffCons_fb_mr_field_m1_h*_knoise0_n500_s7c_combined_1_2000.rds",
  "results/fs_maxeffCons_fb_mr_field_m1_h*_z1q60_p30_combined_1_2000.rds",
  "results/fs_effMaxSG_fb_mr_field_m1_h*_z1q60_p30sg_combined_1_2000.rds",
  "results/fs_effMaxSG_fb_mr_field_m1_h*_z1q60_nb20_p30sgnb20_combined_1_2000.rds",
  "results/fs_effMaxSG_fb_mr_field_m1_h*_z1q60_nb20_j20_p30sgnb20j20_combined_1_2000.rds",
  sep = ";")), ";", fixed = TRUE)[[1]]
files <- unlist(lapply(globs, Sys.glob))
stopifnot(length(files) > 0)
bundles <- lapply(files, readRDS)
# Regime label from the campaign tag; eps / J from the batch metas where the
# pooled meta predates the knob fields (nb20; REPORT_nb20 side issue 1).
knob <- function(b, k, d) {
  v <- b$meta[[k]]
  if (is.null(v) && length(b$meta$source_files)) {
    bm <- lapply(b$meta$source_files, function(f) if (file.exists(file.path("results", f)))
      readRDS(file.path("results", f))$meta[[k]] else NULL)
    bm <- unique(unlist(bm)); if (length(bm) == 1L) v <- bm
  }
  v %||% d
}
regime <- function(b) {
  tag <- b$meta$campaign_tag; f <- b$meta$sg_focus %||% "maxeffCons"
  eps <- knob(b, "effect_neighborhood", 0.10); J <- knob(b, "er_jcuts", 10L)
  z <- b$meta$harm_z1_quantile %||% 0.25
  sprintf("%s (%s%s, J %d, prev %.0f%%)", tag, f,
          if (f %in% c("effMaxSG", "effMinSG")) sprintf(" eps %.2f", eps) else "", J,
          100 * (b$meta$harm_prevalence_super %||% if (z == 0.25) 0.124 else NA)) }
cell <- function(b) sprintf("HR %.2f n%d", b$meta$target_hr_harm, b$meta$n_sample)
order_key <- c(s7c = 1, p30 = 2, p30sg = 3, p30sgnb20 = 4, p30sgnb20j20 = 5)
meta_tab <- data.frame(regime = vapply(bundles, regime, ""), cell = vapply(bundles, cell, ""),
                       tag = vapply(bundles, function(b) b$meta$campaign_tag, ""),
                       hr = vapply(bundles, function(b) b$meta$target_hr_harm, 0),
                       n = vapply(bundles, function(b) b$meta$n_sample, 0L),
                       file = basename(files), stringsAsFactors = FALSE)
ord <- order(meta_tab$hr, meta_tab$n, order_key[meta_tab$tag])
bundles <- bundles[ord]; meta_tab <- meta_tab[ord, ]; files <- files[ord]
kbl(meta_tab[, c("cell", "regime", "file")], booktabs = TRUE, row.names = FALSE,
    caption = "Bundles read (committed pooled bundles; regime = campaign, focus, band, grid, prevalence).") |>
  kable_styling(font_size = 11)
```

```{r fns}
# The per-replicate objects on the detected replicates with every input finite.
abc <- function(r) {
  d <- r[r$detected %in% 1L & is.finite(r$nv_Hc_est) & is.finite(r$mr_Hc_est) &
           is.finite(r$betaHhat_Hc) & is.finite(r$nv_Hc_se) & is.finite(r$fld_Hc_se) &
           is.finite(r$mr_Hc_se_ij), , drop = FALSE]
  d$a <- log(d$nv_Hc_est) - log(d$betaHhat_Hc)
  d$c <- log(d$nv_Hc_est) - log(d$mr_Hc_est)
  d$e <- d$a - d$c
  d$ratio_H <- d$n_harm / d$n_true
  d
}
decomp <- function(d) {
  Va <- var(d$a); Vc <- var(d$c); Cac <- cov(d$a, d$c); Ve <- var(d$e)
  nse2 <- mean(d$nv_Hc_se^2); lam2 <- mean(d$fld_Hc_se^2); ij2 <- mean(d$mr_Hc_se_ij^2)
  data.frame(n = nrow(d),
             mean_a = mean(d$a), mean_c = mean(d$c), mean_e = mean(d$e),
             Var_a = Va, Var_c = Vc, Cov_ac = Cac, corr_ac = Cac / sqrt(Va * Vc), Var_e = Ve,
             check = (Va + Vc - 2 * Cac) - Ve,
             naive_SE2 = nse2, lam_SD2 = lam2, ij_se2 = ij2,
             Va_over_nSE2 = Va / nse2,
             excess_c = Vc - 2 * Cac,                # Var(e) - Var(a)
             Ve_over_nSE2 = Ve / nse2, r = sqrt(Ve / nse2),
             lam2_over_Ve = lam2 / Ve, lam2_over_Va = lam2 / Va, lam2_over_nSE2 = lam2 / nse2,
             ij2_over_Ve = ij2 / Ve,
             SD_a = sqrt(Va), SD_c = sqrt(Vc), SD_e = sqrt(Ve),
             naive_SE = sqrt(nse2), lam_SD = sqrt(lam2), ij_se = sqrt(ij2),
             Hhat_mean = mean(d$n_harm), Hhat_over_H = mean(d$ratio_H),
             p_hat_mean = if ("p_hat_H" %in% names(d)) mean(d$p_hat_H, na.rm = TRUE) else NA_real_)
}
D <- lapply(bundles, function(b) abc(b$results))
A1 <- do.call(rbind, lapply(seq_along(D), function(i) cbind(cell = meta_tab$cell[i], regime = meta_tab$regime[i], decomp(D[[i]]))))
```

# A1 — Per cell: the three components, the reported variances, the ratios

Variances on the log-HR scale. `check` is $[\mathrm{Var}(a) + \mathrm{Var}(c) - 2\mathrm{Cov}(a,c)] - \mathrm{Var}(e)$
(zero up to floating point). `excess_c` $= \mathrm{Var}(c) - 2\mathrm{Cov}(a,c) = \mathrm{Var}(e) - \mathrm{Var}(a)$: the
part of the de-biased error's variance the correction adds (positive) or removes (negative).

```{r a1-var}
kbl(A1[, c("cell", "regime", "n", "mean_a", "mean_c", "mean_e", "Var_a", "Var_c", "Cov_ac", "corr_ac", "Var_e", "check", "excess_c")],
    digits = c(0, 0, 0, 4, 4, 4, 5, 5, 5, 3, 5, 8, 5), booktabs = TRUE, row.names = FALSE,
    caption = "Means and variances of a (naive error), c (two-term correction) and e = a - c (de-biased error) on the complement, per cell.") |>
  kable_styling(font_size = 11)
```

```{r a1-ratio}
kbl(A1[, c("cell", "regime", "naive_SE2", "lam_SD2", "ij_se2", "Var_a", "Var_e", "Va_over_nSE2", "Ve_over_nSE2", "r", "lam2_over_Ve", "lam2_over_Va", "lam2_over_nSE2", "ij2_over_Ve")],
    digits = c(0, 0, 5, 5, 5, 5, 5, 3, 3, 3, 3, 3, 3, 3), booktabs = TRUE, row.names = FALSE,
    caption = "Reported variances (means of the squared per-replicate SEs) against the realized ones. r = sqrt(Var(e) / naive SE^2) is the regime diagnostic SD(beta-tilde^c) / naive SE; lam2_over_Ve = lambda-SD^c2 / Var(e); lam2_over_Va = lambda-SD^c2 / Var(a).") |>
  kable_styling(font_size = 11)
```

```{r a1-sd}
kbl(A1[, c("cell", "regime", "Hhat_mean", "Hhat_over_H", "p_hat_mean", "SD_a", "SD_c", "SD_e", "naive_SE", "lam_SD", "ij_se")],
    digits = c(0, 0, 1, 3, 3, 4, 4, 4, 4, 4, 4), booktabs = TRUE, row.names = FALSE,
    caption = "The same on the SD scale, with the region size and p-hat(Hhat) where recorded.") |>
  kable_styling(font_size = 11)
```

# The reported SD is marginal; the decomposition is on the error

The constructions tables of every report (`fs_sim_bias_coverage()`, `sd_emp = sd(log est)`; the template's
`.sd_log()`) take the SD of the **log estimate across replicates**, so their "SD_β", "SE/SD", "bias in SD units"
and the regime diagnostic "SD(β̃ᶜ)/naive SE" all carry the spread of the per-replicate target β(Ĥᶜ) as well as
the error: $\mathrm{Var}(\log\hat\beta^c_{\text{MR}}) = \mathrm{Var}(e) + \mathrm{Var}(\log\beta(\widehat H^c)) + 2\mathrm{Cov}(e, \log\beta(\widehat H^c))$.
The decomposition above is on $e$. Both are shown here so the two readings can be placed side by side.

```{r marginal}
M <- do.call(rbind, lapply(seq_along(D), function(i) { d <- D[[i]]
  lt <- log(d$betaHhat_Hc); lm <- log(d$mr_Hc_est); ln <- log(d$nv_Hc_est)
  data.frame(cell = meta_tab$cell[i], regime = meta_tab$regime[i], n = nrow(d),
             SD_log_mr = sd(lm), SD_log_nv = sd(ln), SD_log_target = sd(lt), SD_e = sd(d$e), SD_a = sd(d$a),
             corr_e_target = cor(d$e, lt), corr_a_target = cor(d$a, lt),
             naive_SE = sqrt(mean(d$nv_Hc_se^2)), lam_SD = sqrt(mean(d$fld_Hc_se^2)),
             reported_ratio = sd(lm) / sqrt(mean(d$nv_Hc_se^2)),   # SD(beta-tilde^c)/naive SE as reported
             error_ratio = sd(d$e) / sqrt(mean(d$nv_Hc_se^2)),      # sqrt(Var(e)/naive SE^2)
             lam_over_SDe = sqrt(mean(d$fld_Hc_se^2)) / sd(d$e),
             lam_over_SDmr = sqrt(mean(d$fld_Hc_se^2)) / sd(lm)) }))
kbl(M, digits = c(0, 0, 0, 4, 4, 4, 4, 4, 3, 3, 4, 4, 3, 3, 3, 3), booktabs = TRUE, row.names = FALSE,
    caption = "Marginal SD of the log estimates vs the error SDs. reported_ratio = SD(log MR est)/naive SE (the regime diagnostic as printed in the p30sg / nb20 reports); error_ratio = SD(e)/naive SE (this document's r); lam_over_SDe = lambda-SD^c / SD(e); lam_over_SDmr = lambda-SD^c / SD(log MR est) (the reports' 'lambda-SD^c / SD(beta-tilde^c)').") |>
  kable_styling(font_size = 10)
```

# A2 — Stratified by p̂(Ĥ) tertile (nb20 cells) and by |Ĥ|/|H| tertile (all cells)

```{r a2}
strat <- function(i, var, what) {
  d <- D[[i]]; v <- d[[var]]; if (!any(is.finite(v))) return(NULL)
  br <- stats::quantile(v, c(0, 1/3, 2/3, 1), na.rm = TRUE, names = FALSE)
  g <- cut(v, breaks = unique(br), include.lowest = TRUE, labels = FALSE)
  do.call(rbind, lapply(sort(unique(g)), function(k) {
    dk <- d[g == k, , drop = FALSE]
    cbind(cell = meta_tab$cell[i], regime = meta_tab$regime[i], stratum = sprintf("%s T%d [%.3f, %.3f]", what, k, min(v[g == k]), max(v[g == k])),
          decomp(dk)) }))
}
cols2 <- c("cell", "regime", "stratum", "n", "Hhat_over_H", "Var_a", "Var_c", "Cov_ac", "Var_e", "naive_SE2", "lam_SD2", "Va_over_nSE2", "r", "lam2_over_Ve", "lam2_over_Va")
dg2 <- c(0, 0, 0, 0, 3, 5, 5, 5, 5, 5, 5, 3, 3, 3, 3)
A2p <- do.call(rbind, lapply(which(vapply(D, function(d) "p_hat_H" %in% names(d) && any(is.finite(d$p_hat_H)), TRUE)), strat, var = "p_hat_H", what = "p-hat"))
kbl(A2p[, cols2], digits = dg2, booktabs = TRUE, row.names = FALSE,
    caption = "By p-hat(Hhat) tertile (nb20 cells, where p-hat is recorded): the decomposition within each tertile.") |>
  kable_styling(font_size = 10)
```

```{r a2b}
A2h <- do.call(rbind, lapply(seq_along(D), strat, var = "ratio_H", what = "|Hhat|/|H|"))
kbl(A2h[, cols2], digits = dg2, booktabs = TRUE, row.names = FALSE,
    caption = "By |Hhat|/|H| tertile (all cells).") |>
  kable_styling(font_size = 10)
```

# A3 — Across the regime sequence

Rows in the order the complement's variability grew: `s7c` (M1 prevalence 12.4%, `maxeffCons`) →
`p30` (31%, `maxeffCons`) → `p30sg` (`effMaxSG` ε 0.10) → nb20 arm A (ε 0.20, J 10) → nb20 arm B (ε 0.20, J 20).
`excess_total` $= \mathrm{Var}(e) - \text{naive SE}^2$ splits into `excess_a` $= \mathrm{Var}(a) - \text{naive SE}^2$
(the naive estimate's own variance beyond its reported SE: the complement's identity varying with the winner)
and `excess_c` $= \mathrm{Var}(c) - 2\mathrm{Cov}(a,c)$ (what the correction adds). `shortfall` $= \mathrm{Var}(e) - \lambda\text{-SD}^{c\,2}$,
and its share of `excess_total`.

```{r a3}
seq_tab <- function(hr, n) {
  S <- A1[A1$cell == sprintf("HR %.2f n%d", hr, n), ]
  S$excess_total <- S$Var_e - S$naive_SE2
  S$excess_a <- S$Var_a - S$naive_SE2
  S$shortfall <- S$Var_e - S$lam_SD2
  S$shortfall_share <- S$shortfall / S$excess_total
  S$lam_minus_nSE2 <- S$lam_SD2 - S$naive_SE2
  S[, c("regime", "n", "Hhat_mean", "p_hat_mean", "Var_a", "Var_c", "Cov_ac", "Var_e", "naive_SE2", "lam_SD2",
        "excess_a", "excess_c", "excess_total", "lam_minus_nSE2", "shortfall", "shortfall_share", "r", "lam2_over_Ve", "lam2_over_Va")]
}
dg3 <- c(0, 0, 1, 3, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 3, 3, 3, 3)
kbl(seq_tab(1.75, 500), digits = dg3, booktabs = TRUE, row.names = FALSE,
    caption = "HR 1.75 n500: the regime sequence.") |> kable_styling(font_size = 10)
kbl(seq_tab(1.50, 500), digits = dg3, booktabs = TRUE, row.names = FALSE,
    caption = "HR 1.50 n500: the regime sequence (no s7c cell at this HR).") |> kable_styling(font_size = 10)
kbl(seq_tab(1.00, 500), digits = dg3, booktabs = TRUE, row.names = FALSE,
    caption = "HR 1.00 n500: the null sequence.") |> kable_styling(font_size = 10)
kbl(rbind(seq_tab(1.50, 1000), seq_tab(1.00, 1000)), digits = dg3, booktabs = TRUE, row.names = FALSE,
    caption = "n = 1000 cells.") |> kable_styling(font_size = 10)
```

```{r a3-plot, fig.width = 11, fig.height = 4.5}
library(ggplot2)
P <- A1[A1$n > 0, ]; P$regime_short <- sub(" \\(.*$", "", P$regime); P$idx <- order_key[meta_tab$tag]
L <- rbind(data.frame(cell = P$cell, regime = P$regime_short, idx = P$idx, component = "Var(a)", value = P$Var_a),
           data.frame(cell = P$cell, regime = P$regime_short, idx = P$idx, component = "Var(c) - 2Cov(a,c)", value = P$excess_c),
           data.frame(cell = P$cell, regime = P$regime_short, idx = P$idx, component = "Var(e)", value = P$Var_e),
           data.frame(cell = P$cell, regime = P$regime_short, idx = P$idx, component = "naive SE^2", value = P$naive_SE2),
           data.frame(cell = P$cell, regime = P$regime_short, idx = P$idx, component = "lambda-SD^c2", value = P$lam_SD2))
L$regime <- factor(L$regime, levels = names(order_key))
ggplot(L, aes(x = regime, y = value, colour = component, group = component)) +
  geom_point(size = 2) + geom_line() + facet_wrap(~ cell, scales = "free_y", nrow = 1) +
  labs(x = NULL, y = "variance (log-HR)^2", colour = NULL) + theme_bw(base_size = 10) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "bottom")
```

# Numbers for the record

```{r reading, results='asis'}
#| echo: false
for (i in seq_len(nrow(A1))) with(A1[i, ], cat(sprintf(
  "- **%s, %s** (n %d, |Hhat| %.0f): Var(a) %.5f vs naive SE^2 %.5f (ratio %.3f); Var(c) %.5f, Cov(a,c) %.5f (corr %.3f), Var(c) - 2Cov %.5f; Var(e) %.5f (r %.3f, check %.1e); lambda-SD^c2 %.5f = %.3f Var(e) = %.3f Var(a) = %.3f naive SE^2; IJ se^2 %.5f = %.2f Var(e).\n",
  cell, regime, n, Hhat_mean, Var_a, naive_SE2, Va_over_nSE2, Var_c, Cov_ac, corr_ac, excess_c, Var_e, r, check, lam_SD2, lam2_over_Ve, lam2_over_Va, lam2_over_nSE2, ij_se2, ij2_over_Ve)))
```

# A5 — Location by stratum

Part L of `TASK_complement_location_2026-09-08`: the A2 stratification with **means** (each with its
standard error) in place of variances, and the field's two corrections beside $a$, $c$, $e$. Per detected
replicate with every input finite, on the log scale: $\lambda = $ `fld_Hc_lam_mean` (the field's additional
correction, the mean of $\Lambda^{*c}$ over the outer draws), $\lambda_s = $ `fld_Hc_lam_mean_s` (field-s),
and the field estimates' errors $e_f = \log(\texttt{fld\_Hc\_est2}) - \log\beta(\widehat H^c)$ and
$e_{f,s} = \log(\texttt{fld\_Hc\_est2\_s}) - \log\beta(\widehat H^c)$. In `fs_mr_inference.R` the complement
field sets `est2_w <- bdc - mean(lf)` and `est2 = to_eff(est2_w)` (and `est2s_w <- bdc - mean(lfs)`,
`est2_s = to_eff(est2s_w)`), with `mr_Hc_est = to_eff(bdc)`, so $e_f = e - \lambda$ and
$e_{f,s} = e - \lambda_s$ on the working scale; both identities are asserted below to $10^{-12}$.
The correction shortfall $a - c - \lambda_s$ of L3 is therefore $e_{f,s}$ itself.
Observed coverage is that of the field-s one-sided upper bound,
$\mathbb 1\{\beta(\widehat H^c) \le \texttt{fld\_Hc\_up1s\_s}\}$ (Wilson 95% limits), with the unscaled field's
beside it; the Gaussian-implied coverage is $\Phi(z_{0.95}\, r + b)$ with $b = \bar e_{f,s} / \mathrm{SD}(e_{f,s})$
and $r = \overline{\texttt{fld\_Hc\_se\_s}} / \mathrm{SD}(e_{f,s})$ from the stratum's own numbers (the **error**
SD, not the marginal SD of the constructions tables). $\rho^c$ = `fld_Hc_scale_ratio`. Winner-only and
winner-floor excluded as above. The section is skipped for bundles lacking the `_s` columns.

```{r a5-setup}
s_cols <- c("fld_Hc_est2", "fld_Hc_lam_mean", "fld_Hc_est2_s", "fld_Hc_lam_mean_s", "fld_Hc_se_s",
            "fld_Hc_up1s", "fld_Hc_up1s_s", "fld_Hc_scale_ratio", "p_hat_H")
has_s <- vapply(bundles, function(b) all(s_cols %in% names(b$results)) && any(is.finite(b$results$fld_Hc_est2_s)), TRUE)
# The A2 rows (abc()) with the field inputs also finite, plus the two corrections and the two field errors.
loc <- function(r) {
  d <- abc(r)
  d <- d[Reduce(`&`, lapply(s_cols, function(cn) is.finite(d[[cn]]))), , drop = FALSE]
  d$lam <- d$fld_Hc_lam_mean; d$lam_s <- d$fld_Hc_lam_mean_s
  d$ef   <- log(d$fld_Hc_est2)   - log(d$betaHhat_Hc)
  d$ef_s <- log(d$fld_Hc_est2_s) - log(d$betaHhat_Hc)
  d$cov_f <- d$betaHhat_Hc <= d$fld_Hc_up1s
  d$cov_s <- d$betaHhat_Hc <= d$fld_Hc_up1s_s
  d
}
D5 <- lapply(which(has_s), function(i) loc(bundles[[i]]$results)); names(D5) <- which(has_s)
# Regime sequence for this section: effMaxSG (eps ascending) -> maxSG -> minSG, HR ascending within.
ord5 <- if (length(D5)) { i5 <- as.integer(names(D5))
  order(match(vapply(bundles[i5], function(b) b$meta$sg_focus %||% "maxeffCons", ""), c("maxeffCons", "effMaxSG", "effMinSG", "maxSG", "minSG")),
        vapply(bundles[i5], function(b) knob(b, "effect_neighborhood", 0.10), 0), meta_tab$hr[i5]) } else integer(0)
D5 <- D5[ord5]
# Identity check of this record: ef = e - lam and ef_s = e - lam_s on the working scale, to 1e-12.
ident <- do.call(rbind, lapply(seq_along(D5), function(j) { d <- D5[[j]]; i <- as.integer(names(D5)[j])
  data.frame(cell = meta_tab$cell[i], regime = meta_tab$regime[i],
             n_detected = sum(bundles[[i]]$results$detected %in% 1L), n = nrow(d),
             max_abs_ef_minus_e_lam = max(abs(d$ef - (d$e - d$lam))),
             max_abs_efs_minus_e_lams = max(abs(d$ef_s - (d$e - d$lam_s)))) }))
stopifnot(length(D5) == 0 || (all(ident$max_abs_ef_minus_e_lam <= 1e-12) && all(ident$max_abs_efs_minus_e_lams <= 1e-12)))
if (length(D5)) kbl(ident, digits = c(0, 0, 0, 0, 16, 16), booktabs = TRUE, row.names = FALSE,
    caption = "Identity check: max |ef - (e - lam)| and max |ef_s - (e - lam_s)| per cell (asserted <= 1e-12); n = detected replicates with every input finite.") |>
  kable_styling(font_size = 10)
```

```{r a5-fns}
wil <- function(p, n, z = stats::qnorm(0.975)) {
  ctr <- (p + z^2 / (2 * n)) / (1 + z^2 / n)
  hw  <- z * sqrt(p * (1 - p) / n + z^2 / (4 * n^2)) / (1 + z^2 / n)
  c(ctr - hw, ctr + hw)
}
# decomp() of A1/A2 with means (and their SEs) in place of variances.
locate <- function(d) {
  m <- function(v) mean(v); s <- function(v) stats::sd(v) / sqrt(length(v))
  z1 <- stats::qnorm(0.95)
  b <- mean(d$ef_s) / stats::sd(d$ef_s); rr <- mean(d$fld_Hc_se_s) / stats::sd(d$ef_s)
  w <- wil(mean(d$cov_s), nrow(d))
  data.frame(n = nrow(d), Hhat_over_H = mean(d$ratio_H), p_hat_mean = mean(d$p_hat_H), rho_c = mean(d$fld_Hc_scale_ratio),
             mean_a = m(d$a), se_a = s(d$a), mean_c = m(d$c), se_c = s(d$c), mean_e = m(d$e), se_e = s(d$e),
             mean_lam = m(d$lam), se_lam = s(d$lam), mean_ef = m(d$ef), se_ef = s(d$ef),
             mean_lam_s = m(d$lam_s), se_lam_s = s(d$lam_s), mean_ef_s = m(d$ef_s), se_ef_s = s(d$ef_s),
             SD_ef_s = stats::sd(d$ef_s), naive_SE = mean(d$nv_Hc_se), se_field_s = mean(d$fld_Hc_se_s),
             cov_field = mean(d$cov_f), cov_field_s = mean(d$cov_s), wilson_lo = w[1], wilson_hi = w[2],
             b_s = b, r_s = rr, gauss_s = stats::pnorm(z1 * rr + b))
}
# strat() of A2 with locate() in place of decomp(); the whole-cell row first.
strat5 <- function(j, var, what) {
  d <- D5[[j]]; i <- as.integer(names(D5)[j]); v <- d[[var]]; if (!any(is.finite(v))) return(NULL)
  br <- stats::quantile(v, c(0, 1/3, 2/3, 1), na.rm = TRUE, names = FALSE)
  g <- cut(v, breaks = unique(br), include.lowest = TRUE, labels = FALSE)
  rbind(cbind(cell = meta_tab$cell[i], regime = meta_tab$regime[i], stratum = sprintf("%s all", what), locate(d)),
        do.call(rbind, lapply(sort(unique(g)), function(k) {
          dk <- d[g == k, , drop = FALSE]
          cbind(cell = meta_tab$cell[i], regime = meta_tab$regime[i],
                stratum = sprintf("%s T%d [%.3f, %.3f]", what, k, min(v[g == k]), max(v[g == k])),
                locate(dk)) })))
}
cols5a <- c("cell", "regime", "stratum", "n", "p_hat_mean", "Hhat_over_H", "rho_c",
            "mean_a", "se_a", "mean_c", "se_c", "mean_e", "se_e", "mean_lam", "se_lam", "mean_ef", "se_ef",
            "mean_lam_s", "se_lam_s", "mean_ef_s", "se_ef_s")
dg5a <- c(0, 0, 0, 0, 3, 3, 3, rep(4, 14))
cols5b <- c("cell", "regime", "stratum", "n", "naive_SE", "se_field_s", "SD_ef_s", "mean_ef_s", "b_s", "r_s",
            "cov_field", "cov_field_s", "wilson_lo", "wilson_hi", "gauss_s")
dg5b <- c(0, 0, 0, 0, 4, 4, 4, 4, 3, 3, 3, 3, 3, 3, 3)
A5p <- if (length(D5)) do.call(rbind, lapply(seq_along(D5), strat5, var = "p_hat_H", what = "p-hat")) else NULL
A5h <- if (length(D5)) do.call(rbind, lapply(seq_along(D5), strat5, var = "ratio_H", what = "|Hhat|/|H|")) else NULL
```

## L1 — Means by p̂(Ĥ) tertile, per cell

```{r a5-l1}
if (!is.null(A5p)) {
  kbl(A5p[, cols5a], digits = dg5a, booktabs = TRUE, row.names = FALSE,
      caption = "L1 (location): by p-hat(Hhat) tertile (tertiles within the cell as in A2). Means with their SEs of a (naive error), c (two-term correction), e = a - c, lam (field correction), ef = e - lam, lam_s (field-s correction), ef_s = e - lam_s; mean rho^c and |Hhat|/|H|.") |>
    kable_styling(font_size = 9)
}
```

```{r a5-l1-cov}
if (!is.null(A5p)) {
  kbl(A5p[, cols5b], digits = dg5b, booktabs = TRUE, row.names = FALSE,
      caption = "L1 (coverage): mean naive SE, mean fld_Hc_se_s and the error SD of ef_s in the stratum; observed one-sided upper coverage of the field (cov_field) and field-s (cov_field_s, Wilson) bounds; the Gaussian-implied field-s coverage Phi(z r + b) from the stratum's own (mean ef_s, SD ef_s, mean fld_Hc_se_s).") |>
    kable_styling(font_size = 9)
}
```

## L2 — The same by |Ĥ|/|H| tertile, per cell

```{r a5-l2}
if (!is.null(A5h)) {
  kbl(A5h[, cols5a], digits = dg5a, booktabs = TRUE, row.names = FALSE,
      caption = "L2 (location): by |Hhat|/|H| tertile.") |> kable_styling(font_size = 9)
  kbl(A5h[, cols5b], digits = dg5b, booktabs = TRUE, row.names = FALSE,
      caption = "L2 (coverage): by |Hhat|/|H| tertile.") |> kable_styling(font_size = 9)
}
```

## L3 — Regime sequence across the cells (p̂ tertiles)

Cells in the order effMaxSG ε ascending → maxSG → minSG (HR ascending within). For each cell the whole-cell
value and the three p̂-tertile values of the naive optimism $\bar a$, the two-term correction $\bar c$, the
field-s correction $\bar\lambda_s$, the residual $\bar e_{f,s} = \bar a - \bar c - \bar\lambda_s$
(the correction shortfall), and the residual as a fraction of the optimism, $\bar e_{f,s} / \bar a$.

```{r a5-l3}
if (!is.null(A5p)) {
  L3 <- do.call(rbind, lapply(seq_along(D5), function(j) { i <- as.integer(names(D5)[j])
    S <- A5p[A5p$cell == meta_tab$cell[i] & A5p$regime == meta_tab$regime[i], ]
    pick <- function(col) { v <- S[[col]]; names(v) <- sub("^p-hat ", "", sub(" \\[.*$", "", S$stratum)); v }
    q <- list(`mean a` = pick("mean_a"), `mean c` = pick("mean_c"), `mean lam_s` = pick("mean_lam_s"),
              `mean ef_s = a - c - lam_s` = pick("mean_ef_s"), `ef_s / a` = pick("mean_ef_s") / pick("mean_a"))
    do.call(rbind, lapply(names(q), function(nm) { v <- q[[nm]]
      data.frame(cell = meta_tab$cell[i], regime = meta_tab$regime[i], quantity = nm,
                 all = v[["all"]], T1 = v[["T1"]], T2 = v[["T2"]], T3 = v[["T3"]]) })) }))
  kbl(L3, digits = c(0, 0, 0, 4, 4, 4, 4), booktabs = TRUE, row.names = FALSE,
      caption = "L3: the regime sequence, p-hat tertiles (T1 low ... T3 high p-hat).") |> kable_styling(font_size = 10)
}
```

## Numbers for the record (A5)

```{r a5-record, results='asis'}
#| echo: false
if (!is.null(A5p)) {
  f3 <- function(x) formatC(x, digits = 3, format = "f"); f4 <- function(x) formatC(x, digits = 4, format = "f")
  ms <- function(m, s) sprintf("%s (%s)", f4(m), f4(s))
  st <- function(s) gsub("|Hhat|/|H|", "Hhat/H", s, fixed = TRUE)   # pipe-free label for the markdown tables
  cat("**Identity check** (max |ef - (e - lam)|, max |ef_s - (e - lam_s)|; n detected / n analysed):\n\n| cell | regime | detected | n | max abs ef | max abs ef_s |\n|---|---|---|---|---|---|\n")
  for (i in seq_len(nrow(ident))) with(ident[i, ], cat(sprintf("| %s | %s | %d | %d | %.1e | %.1e |\n", cell, regime, n_detected, n, max_abs_ef_minus_e_lam, max_abs_efs_minus_e_lams)))
  for (what in c("p-hat", "|Hhat|/|H|")) { A <- if (what == "p-hat") A5p else A5h
    cat(sprintf("\n**%s -- location by %s tertile** (mean (SE) on the log scale):\n\n| cell | regime | stratum | n | mean p-hat | mean Hhat/H | mean rho^c | a | c | e | lam | ef | lam_s | ef_s |\n|---|---|---|---|---|---|---|---|---|---|---|---|---|---|\n",
                if (what == "p-hat") "L1" else "L2", what))
    for (i in seq_len(nrow(A))) with(A[i, ], cat(sprintf("| %s | %s | %s | %d | %s | %s | %s | %s | %s | %s | %s | %s | %s | %s |\n", cell, regime, st(stratum), n, f3(p_hat_mean), f3(Hhat_over_H), f3(rho_c),
      ms(mean_a, se_a), ms(mean_c, se_c), ms(mean_e, se_e), ms(mean_lam, se_lam), ms(mean_ef, se_ef), ms(mean_lam_s, se_lam_s), ms(mean_ef_s, se_ef_s))))
    cat(sprintf("\n**%s -- coverage by %s tertile** (field-s upper bound; Gaussian-implied from the stratum's own mean ef_s, SD ef_s, mean fld_Hc_se_s):\n\n| cell | regime | stratum | n | mean naive SE | mean se_field_s | SD ef_s | mean ef_s | b | r | field cov | field-s cov [Wilson] | Gaussian-implied |\n|---|---|---|---|---|---|---|---|---|---|---|---|---|\n",
                if (what == "p-hat") "L1" else "L2", what))
    for (i in seq_len(nrow(A))) with(A[i, ], cat(sprintf("| %s | %s | %s | %d | %s | %s | %s | %s | %s | %s | %s | %s [%s, %s] | %s |\n", cell, regime, st(stratum), n, f4(naive_SE), f4(se_field_s), f4(SD_ef_s), f4(mean_ef_s), f3(b_s), f3(r_s),
      f3(cov_field), f3(cov_field_s), f3(wilson_lo), f3(wilson_hi), f3(gauss_s)))) }
  cat("\n**L3 -- regime sequence, p-hat tertiles** (T1 low ... T3 high p-hat):\n\n| cell | regime | quantity | all | T1 | T2 | T3 |\n|---|---|---|---|---|---|---|\n")
  for (i in seq_len(nrow(L3))) with(L3[i, ], cat(sprintf("| %s | %s | %s | %s | %s | %s | %s |\n", cell, regime, quantity, f4(all), f4(T1), f4(T2), f4(T3))))
}
```

# A6 — Harm-side location by stratum

`TASK_harm_location_2026-09-08`: the A5 chunk with the **harm-block** columns substituted, on the same
detected replicates (the A5 rows with the harm-block field inputs also finite). Per replicate, on the log
scale: $a_H = \log(\texttt{nv\_H\_est}) - \log\beta(\widehat H)$ (the naive error on $\widehat H$; positive =
harm over-estimated), $c_H = \log(\texttt{nv\_H\_est}) - \log(\texttt{mr\_H\_est})$ (the two-term
correction), $e_H = a_H - c_H$, $\lambda_H = $ `fld_H_lam_mean` (the harm field's own correction, the mean of
$\Lambda^*$ over the outer draws) and $e_{f,H} = \log(\texttt{fld\_H\_est2}) - \log\beta(\widehat H)$. In
`fs_mr_inference.R` the harm field sets `est2_w <- beta_deb - mean(lf)` and `est2 = to_eff(est2_w)` with
`mr_H_est = to_eff(beta_deb)`, so $e_{f,H} = e_H - \lambda_H$ on the working scale; asserted below to
$10^{-12}$. Observed coverage is that of the harm-side one-sided **lower** bounds: the field's stored
`fld_H_lo1s` ($= $ `to_eff(beta_deb - q95(Lambda*))`, the quantile form), $\mathbb 1\{\beta(\widehat H) \ge
\texttt{fld\_H\_lo1s}\}$, and the IJ two-term's $\exp(\log \texttt{mr\_H\_est} - z_{0.95}\,\texttt{mr\_H\_se\_ij})$
(the `e1stud_findings.R` construction); Wilson 95% limits. The Gaussian-implied field coverage is
$\Phi(z_{0.95}\, r - b)$ with $b = \bar e_{f,H} / \mathrm{SD}(e_{f,H})$ and
$r = \overline{\texttt{fld\_H\_se}} / \mathrm{SD}(e_{f,H})$ from the stratum's own numbers (the error SD; the
marginal SD of $\log \texttt{fld\_H\_est2}$ is shown beside it). No `_s` companion exists on the harm block
(`field_scale_complement` acts on the complement only). Tertiles within the cell as in A5.

```{r a6-setup}
h_cols <- c("nv_H_est", "nv_H_se", "mr_H_est", "mr_H_se_ij", "betaHhat_H", "fld_H_est2", "fld_H_lam_mean",
            "fld_H_se", "fld_H_lo1s")
has_h <- vapply(seq_along(bundles), function(i) has_s[i] && all(h_cols %in% names(bundles[[i]]$results)), TRUE)
# The A5 rows (loc()) with the harm-block inputs also finite, plus the harm-side quantities.
loch <- function(r) {
  d <- loc(r)
  d <- d[Reduce(`&`, lapply(h_cols, function(cn) is.finite(d[[cn]]))), , drop = FALSE]
  d$a_H  <- log(d$nv_H_est) - log(d$betaHhat_H)
  d$c_H  <- log(d$nv_H_est) - log(d$mr_H_est)
  d$e_H  <- d$a_H - d$c_H
  d$lam_H <- d$fld_H_lam_mean
  d$ef_H <- log(d$fld_H_est2) - log(d$betaHhat_H)
  d$cov_fH  <- d$betaHhat_H >= d$fld_H_lo1s
  d$cov_ijH <- d$betaHhat_H >= exp(log(d$mr_H_est) - stats::qnorm(0.95) * d$mr_H_se_ij)
  d
}
D6 <- lapply(which(has_h), function(i) loch(bundles[[i]]$results)); names(D6) <- which(has_h)
D6 <- D6[match(names(D5), names(D6))]; D6 <- D6[!is.na(names(D6))]   # A5's regime order
ident6 <- do.call(rbind, lapply(seq_along(D6), function(j) { d <- D6[[j]]; i <- as.integer(names(D6)[j])
  data.frame(cell = meta_tab$cell[i], regime = meta_tab$regime[i],
             n_detected = sum(bundles[[i]]$results$detected %in% 1L), n_A5 = nrow(D5[[as.character(i)]]), n = nrow(d),
             max_abs_efH_minus_eH_lamH = max(abs(d$ef_H - (d$e_H - d$lam_H)))) }))
stopifnot(length(D6) == 0 || all(ident6$max_abs_efH_minus_eH_lamH <= 1e-12))
if (length(D6)) kbl(ident6, digits = c(0, 0, 0, 0, 0, 16), booktabs = TRUE, row.names = FALSE,
    caption = "Identity check (harm block): max |ef_H - (e_H - lam_H)| per cell (asserted <= 1e-12); n = A5's replicates with the harm-block inputs also finite.") |>
  kable_styling(font_size = 10)
```

```{r a6-fns}
# locate() of A5 with the harm-block columns.
locate6 <- function(d) {
  m <- function(v) mean(v); s <- function(v) stats::sd(v) / sqrt(length(v))
  z1 <- stats::qnorm(0.95)
  b <- mean(d$ef_H) / stats::sd(d$ef_H); rr <- mean(d$fld_H_se) / stats::sd(d$ef_H)
  wf <- wil(mean(d$cov_fH), nrow(d)); wi <- wil(mean(d$cov_ijH), nrow(d))
  data.frame(n = nrow(d), Hhat_over_H = mean(d$ratio_H), p_hat_mean = mean(d$p_hat_H),
             mean_aH = m(d$a_H), se_aH = s(d$a_H), mean_cH = m(d$c_H), se_cH = s(d$c_H), mean_eH = m(d$e_H), se_eH = s(d$e_H),
             mean_lamH = m(d$lam_H), se_lamH = s(d$lam_H), mean_efH = m(d$ef_H), se_efH = s(d$ef_H),
             SD_efH = stats::sd(d$ef_H), SD_marg_fldH = stats::sd(log(d$fld_H_est2)),
             naive_SE_H = mean(d$nv_H_se), se_field_H = mean(d$fld_H_se), se_ij_H = mean(d$mr_H_se_ij),
             cov_field_H = mean(d$cov_fH), wf_lo = wf[1], wf_hi = wf[2],
             cov_ij_H = mean(d$cov_ijH), wi_lo = wi[1], wi_hi = wi[2],
             b_H = b, r_H = rr, gauss_H = stats::pnorm(z1 * rr - b),
             # the complement's A5 quantities on the same rows, for H3
             mean_a = m(d$a), mean_ef_s = m(d$ef_s))
}
strat6 <- function(j, var, what) {
  d <- D6[[j]]; i <- as.integer(names(D6)[j]); v <- d[[var]]; if (!any(is.finite(v))) return(NULL)
  br <- stats::quantile(v, c(0, 1/3, 2/3, 1), na.rm = TRUE, names = FALSE)
  g <- cut(v, breaks = unique(br), include.lowest = TRUE, labels = FALSE)
  rbind(cbind(cell = meta_tab$cell[i], regime = meta_tab$regime[i], stratum = sprintf("%s all", what), locate6(d)),
        do.call(rbind, lapply(sort(unique(g)), function(k) {
          dk <- d[g == k, , drop = FALSE]
          cbind(cell = meta_tab$cell[i], regime = meta_tab$regime[i],
                stratum = sprintf("%s T%d [%.3f, %.3f]", what, k, min(v[g == k]), max(v[g == k])),
                locate6(dk)) })))
}
cols6a <- c("cell", "regime", "stratum", "n", "p_hat_mean", "Hhat_over_H",
            "mean_aH", "se_aH", "mean_cH", "se_cH", "mean_eH", "se_eH", "mean_lamH", "se_lamH", "mean_efH", "se_efH")
dg6a <- c(0, 0, 0, 0, 3, 3, rep(4, 10))
cols6b <- c("cell", "regime", "stratum", "n", "naive_SE_H", "se_field_H", "se_ij_H", "SD_efH", "SD_marg_fldH", "mean_efH", "b_H", "r_H",
            "cov_field_H", "wf_lo", "wf_hi", "cov_ij_H", "wi_lo", "wi_hi", "gauss_H")
dg6b <- c(0, 0, 0, 0, 4, 4, 4, 4, 4, 4, 3, 3, 3, 3, 3, 3, 3, 3, 3)
A6p <- if (length(D6)) do.call(rbind, lapply(seq_along(D6), strat6, var = "p_hat_H", what = "p-hat")) else NULL
A6h <- if (length(D6)) do.call(rbind, lapply(seq_along(D6), strat6, var = "ratio_H", what = "|Hhat|/|H|")) else NULL
```

## H1 — Harm-block means by p̂(Ĥ) tertile, per cell

```{r a6-h1}
if (!is.null(A6p)) {
  kbl(A6p[, cols6a], digits = dg6a, booktabs = TRUE, row.names = FALSE,
      caption = "H1 (location, harm block): by p-hat(Hhat) tertile. Means with their SEs of a_H (naive error on Hhat), c_H (two-term correction), e_H = a_H - c_H, lam_H (harm field correction), ef_H = e_H - lam_H; mean p-hat and |Hhat|/|H|.") |>
    kable_styling(font_size = 9)
}
```

```{r a6-h1-cov}
if (!is.null(A6p)) {
  kbl(A6p[, cols6b], digits = dg6b, booktabs = TRUE, row.names = FALSE,
      caption = "H1 (coverage, harm block): mean naive SE, mean fld_H_se, mean mr_H_se_ij; the error SD of ef_H and the marginal SD of log fld_H_est2; observed one-sided lower coverage of the field bound fld_H_lo1s (Wilson) and of the IJ two-term bound exp(log mr_H_est - z mr_H_se_ij) (Wilson); the Gaussian-implied field coverage Phi(z r - b) from the stratum's own (mean ef_H, SD ef_H, mean fld_H_se).") |>
    kable_styling(font_size = 9)
}
```

## H2 — The same by |Ĥ|/|H| tertile, per cell

```{r a6-h2}
if (!is.null(A6h)) {
  kbl(A6h[, cols6a], digits = dg6a, booktabs = TRUE, row.names = FALSE,
      caption = "H2 (location, harm block): by |Hhat|/|H| tertile.") |> kable_styling(font_size = 9)
  kbl(A6h[, cols6b], digits = dg6b, booktabs = TRUE, row.names = FALSE,
      caption = "H2 (coverage, harm block): by |Hhat|/|H| tertile.") |> kable_styling(font_size = 9)
}
```

## H3 — The paradox test on both blocks

Per cell, on the same replicates: the correlation of $\hat p$ with the harm-side naive error $a_H$ and with
the complement's naive error $a$ (Pearson and Spearman), and the uncorrected fractions
$\bar e_{f,H} / \bar a_H$ (harm block) and $\bar e_{f,s} / \bar a$ (complement, A5's L3) by $\hat p$ tertile.

```{r a6-h3}
if (!is.null(A6p)) {
  H3c <- do.call(rbind, lapply(seq_along(D6), function(j) { d <- D6[[j]]; i <- as.integer(names(D6)[j])
    data.frame(cell = meta_tab$cell[i], regime = meta_tab$regime[i], n = nrow(d),
               pearson_p_aH = stats::cor(d$p_hat_H, d$a_H), spearman_p_aH = stats::cor(d$p_hat_H, d$a_H, method = "spearman"),
               pearson_p_a = stats::cor(d$p_hat_H, d$a), spearman_p_a = stats::cor(d$p_hat_H, d$a, method = "spearman")) }))
  kbl(H3c, digits = c(0, 0, 0, 3, 3, 3, 3), booktabs = TRUE, row.names = FALSE,
      caption = "H3 (correlations): corr(p-hat, a_H) [harm block] and corr(p-hat, a) [complement], Pearson and Spearman, per cell.") |> kable_styling(font_size = 10)
  H3f <- do.call(rbind, lapply(seq_along(D6), function(j) { i <- as.integer(names(D6)[j])
    S <- A6p[A6p$cell == meta_tab$cell[i] & A6p$regime == meta_tab$regime[i], ]
    pick <- function(col) { v <- S[[col]]; names(v) <- sub("^p-hat ", "", sub(" \\[.*$", "", S$stratum)); v }
    q <- list(`a_H (harm)` = pick("mean_aH"), `ef_H (harm)` = pick("mean_efH"), `ef_H / a_H (harm)` = pick("mean_efH") / pick("mean_aH"),
              `a (complement)` = pick("mean_a"), `ef_s (complement)` = pick("mean_ef_s"), `ef_s / a (complement)` = pick("mean_ef_s") / pick("mean_a"))
    do.call(rbind, lapply(names(q), function(nm) { v <- q[[nm]]
      data.frame(cell = meta_tab$cell[i], regime = meta_tab$regime[i], quantity = nm,
                 all = v[["all"]], T1 = v[["T1"]], T2 = v[["T2"]], T3 = v[["T3"]]) })) }))
  kbl(H3f, digits = c(0, 0, 0, 4, 4, 4, 4), booktabs = TRUE, row.names = FALSE,
      caption = "H3 (uncorrected fractions): harm block ef_H / a_H beside the complement's ef_s / a, by p-hat tertile (T1 low ... T3 high p-hat); the means they are ratios of shown above them.") |> kable_styling(font_size = 10)
}
```

## Numbers for the record (A6)

```{r a6-record, results='asis'}
#| echo: false
if (!is.null(A6p)) {
  f3 <- function(x) formatC(x, digits = 3, format = "f"); f4 <- function(x) formatC(x, digits = 4, format = "f")
  ms <- function(m, s) sprintf("%s (%s)", f4(m), f4(s))
  st <- function(s) gsub("|Hhat|/|H|", "Hhat/H", s, fixed = TRUE)
  cat("**Identity check (harm block)** (max |ef_H - (e_H - lam_H)|; n detected / n in A5 / n analysed):\n\n| cell | regime | detected | n A5 | n | max abs ef_H |\n|---|---|---|---|---|---|\n")
  for (i in seq_len(nrow(ident6))) with(ident6[i, ], cat(sprintf("| %s | %s | %d | %d | %d | %.1e |\n", cell, regime, n_detected, n_A5, n, max_abs_efH_minus_eH_lamH)))
  for (what in c("p-hat", "|Hhat|/|H|")) { A <- if (what == "p-hat") A6p else A6h
    cat(sprintf("\n**%s -- harm-block location by %s tertile** (mean (SE) on the log scale):\n\n| cell | regime | stratum | n | mean p-hat | mean Hhat/H | a_H | c_H | e_H | lam_H | ef_H |\n|---|---|---|---|---|---|---|---|---|---|---|\n",
                if (what == "p-hat") "H1" else "H2", what))
    for (i in seq_len(nrow(A))) with(A[i, ], cat(sprintf("| %s | %s | %s | %d | %s | %s | %s | %s | %s | %s | %s |\n", cell, regime, st(stratum), n, f3(p_hat_mean), f3(Hhat_over_H),
      ms(mean_aH, se_aH), ms(mean_cH, se_cH), ms(mean_eH, se_eH), ms(mean_lamH, se_lamH), ms(mean_efH, se_efH))))
    cat(sprintf("\n**%s -- harm-block coverage by %s tertile** (one-sided lower bounds; Gaussian-implied from the stratum's own mean ef_H, SD ef_H, mean fld_H_se):\n\n| cell | regime | stratum | n | mean naive SE | mean fld_H_se | mean se_ij | SD ef_H (error) | SD log fld_H_est2 (marginal) | mean ef_H | b | r | field cov [Wilson] | IJ cov [Wilson] | Gaussian-implied (field) |\n|---|---|---|---|---|---|---|---|---|---|---|---|---|---|---|\n",
                if (what == "p-hat") "H1" else "H2", what))
    for (i in seq_len(nrow(A))) with(A[i, ], cat(sprintf("| %s | %s | %s | %d | %s | %s | %s | %s | %s | %s | %s | %s | %s [%s, %s] | %s [%s, %s] | %s |\n", cell, regime, st(stratum), n, f4(naive_SE_H), f4(se_field_H), f4(se_ij_H), f4(SD_efH), f4(SD_marg_fldH), f4(mean_efH), f3(b_H), f3(r_H),
      f3(cov_field_H), f3(wf_lo), f3(wf_hi), f3(cov_ij_H), f3(wi_lo), f3(wi_hi), f3(gauss_H)))) }
  cat("\n**H3 -- correlations of p-hat with the naive errors** (Pearson / Spearman):\n\n| cell | regime | n | corr(p-hat, a_H) Pearson | Spearman | corr(p-hat, a) Pearson | Spearman |\n|---|---|---|---|---|---|---|\n")
  for (i in seq_len(nrow(H3c))) with(H3c[i, ], cat(sprintf("| %s | %s | %d | %s | %s | %s | %s |\n", cell, regime, n, f3(pearson_p_aH), f3(spearman_p_aH), f3(pearson_p_a), f3(spearman_p_a))))
  cat("\n**H3 -- uncorrected fractions, p-hat tertiles** (T1 low ... T3 high p-hat):\n\n| cell | regime | quantity | all | T1 | T2 | T3 |\n|---|---|---|---|---|---|---|\n")
  for (i in seq_len(nrow(H3f))) with(H3f[i, ], cat(sprintf("| %s | %s | %s | %s | %s | %s | %s |\n", cell, regime, quantity, f4(all), f4(T1), f4(T2), f4(T3))))
}
```
