---
title: "Prevalence ~30%: the current constructions across the p30 cells"
subtitle: "Naive, MR (IJ) and MR (field) for both blocks, and the joint pair, under a ~30% harm subgroup (TASK_prevalence30_2026-09-06)"
author: "Larry León"
date: today
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    code-fold: true
---

The four `p30` cells plant the harm subgroup in `{er ≤ q₀.₆₀(er)} ∩ {meno = 0}`
(prevalence 0.3065 on the super-population; M1's `{er ≤ 8} ∩ {meno = 0}` is
12.4%). Every complement result before this came from designs where near-tied
candidates have near-identical complements, so the complement's selection
variability was negligible — a regime, not a property of any method. The
constructions here are the current set only: naive and oracle as references,
MR (IJ, two-term), MR (field), and the joint pair; the displays use the Gaussian
bookkeeping reference (harm block: lower bound, $\Phi(z_{0.95} r - b)$;
complement block: upper bound, $\Phi(z_{0.95} r + b)$).

# Knobs

```{r setup, message=FALSE}
library(forestsearch)
library(ggplot2)
library(kableExtra)
`%||%` <- function(a, b) if (is.null(a) || length(a) == 0 || all(is.na(a))) b else a
globs <- strsplit(Sys.getenv("FS_SUMP_GLOBS",
                  "results/fs_*_z1q60_p30_combined_1_2000.rds"),
                  ";", fixed = TRUE)[[1]]
estimators <- c("naive", "mr", "fld")
curves     <- c(1, 1.25, 1.5, 2)
cell_label <- function(meta, f) {
  hr <- meta$target_hr_harm
  base <- if (hr < 1) sprintf("protective %.2f", hr)
  else if (hr == 1) "null 1.0" else sprintf("harm %.4g", hr)
  base <- paste0(base, sprintf(", prev %.0f%%", 100 * (meta$harm_prevalence_super %||% NA)))
  kn <- as.integer(sub("^.*_knoise(\\d+)_.*$", "\\1", basename(f)))
  nn <- if (meta$n_sample == 500) if (kn > 0 || hr < 1) "" else ", n=500" else sprintf(", n=%d", meta$n_sample)
  paste0(base, nn, if (kn > 0) sprintf(", %d noise", kn) else "")
}
files <- sort(unlist(lapply(globs, Sys.glob)))
stopifnot(length(files) > 0)
bundles <- lapply(files, readRDS)
labels  <- mapply(cell_label, lapply(bundles, `[[`, "meta"), files)
```

# Complement block (one-sided = upper bound)

```{r hc}
tabC <- do.call(rbind, lapply(seq_along(bundles), function(i) {
  s <- fs_sim_bias_coverage(bundles[[i]]$results, block = "Hc", estimators = estimators, side = "upper")
  s$cell <- labels[i]; s
}))
tabC$estimator <- factor(tabC$estimator, levels = c("naive", "mr", "fld"),
                         labels = c("naive", "IJ two-term", "field"))
kbl(tabC[, c("cell", "estimator", "n", "b", "r", "cov1", "cov1_wilson_lo", "cov1_wilson_hi",
             "cov2", "cov2_wilson_lo", "cov2_wilson_hi", "cov1_ref", "cov2_ref")],
    digits = 3, booktabs = TRUE, row.names = FALSE,
    caption = "Complement block; cov1 = one-sided 95% UPPER-bound coverage of beta(Hhat^c).") |>
  kable_styling(font_size = 12)
```

```{r hc-plot}
#| fig-width: 13
#| fig-height: 4.6
fs_plot_bias_coverage(tabC, curves = curves, side = "upper")
```

# Harm block (one-sided = lower bound)

```{r h}
tabH <- do.call(rbind, lapply(seq_along(bundles), function(i) {
  s <- fs_sim_bias_coverage(bundles[[i]]$results, block = "H", estimators = estimators, side = "lower")
  s$cell <- labels[i]; s
}))
tabH$estimator <- factor(tabH$estimator, levels = c("naive", "mr", "fld"),
                         labels = c("naive", "IJ two-term", "field"))
kbl(tabH[, c("cell", "estimator", "n", "b", "r", "cov1", "cov1_wilson_lo", "cov1_wilson_hi",
             "cov2", "cov2_wilson_lo", "cov2_wilson_hi", "cov1_ref", "cov2_ref")],
    digits = 3, booktabs = TRUE, row.names = FALSE,
    caption = "Harm block; cov1 = one-sided 95% LOWER-bound coverage of beta(Hhat).") |>
  kable_styling(font_size = 12)
```

```{r h-plot}
#| fig-width: 13
#| fig-height: 4.6
fs_plot_bias_coverage(tabH, curves = curves, side = "lower")
```

# The simultaneous pair

```{r joint}
wil <- function(x, n, z = qnorm(.975)) { c <- (x + z^2/(2*n))/(1+z^2/n); h <- z*sqrt(x*(1-x)/n + z^2/(4*n^2))/(1+z^2/n); c(c - h, c + h) }
tabJ <- do.call(rbind, lapply(seq_along(bundles), function(i) {
  r <- bundles[[i]]$results; d <- r[r$detected %in% 1L, ]
  d <- d[is.finite(d$betaHhat_H) & is.finite(d$betaHhat_Hc) & is.finite(d$fld_joint_gamma), ]
  pairs <- list(`separate 95%` = list(lo = d$fld_H_lo1s, up = d$fld_Hc_up1s),
                `Bonferroni`   = list(lo = d$fld_joint_bonf_loH, up = d$fld_joint_bonf_upHc),
                `calibrated`   = list(lo = d$fld_joint_loH, up = d$fld_joint_upHc))
  do.call(rbind, lapply(names(pairs), function(nm) { p <- pairs[[nm]]
    cH <- d$betaHhat_H >= p$lo; cC <- d$betaHhat_Hc <= p$up; jc <- mean(cH & cC); w <- wil(jc, nrow(d))
    data.frame(cell = labels[i], pair = nm, n = nrow(d), joint_cov = jc, wilson_lo = w[1], wilson_hi = w[2],
               cov_H = mean(cH), cov_Hc = mean(cC),
               margin_H = mean(log(d$mr_H_est) - log(p$lo)), margin_Hc = mean(log(p$up) - log(d$mr_Hc_est)),
               mean_gamma = mean(d$fld_joint_gamma), mean_corr = mean(d$fld_joint_corr), stringsAsFactors = FALSE) }))
}))
kbl(tabJ, digits = 3, booktabs = TRUE, row.names = FALSE,
    caption = "Joint coverage of (beta(Hhat) >= lower_H, beta(Hhat^c) <= upper_Hc) with Wilson intervals; margins on the log scale from the two-term de-biased estimates.") |>
  kable_styling(font_size = 12)
```

# Reading

```{r reading, results='asis'}
#| echo: false
fC <- function(k) tabC[tabC$estimator == k, ]; fH <- function(k) tabH[tabH$estimator == k, ]
cal <- tabJ[tabJ$pair == "calibrated", ]; bon <- tabJ[tabJ$pair == "Bonferroni", ]; sep <- tabJ[tabJ$pair == "separate 95%", ]
cat(sprintf(paste0(
  "Complement, one-sided upper coverage: naive %s (r = %s); IJ two-term %s (r = %s); field %s (r = %s). ",
  "Harm, one-sided lower coverage: naive %s; IJ two-term %s (r = %s); field %s (r = %s). ",
  "Joint pair: calibrated coverage %s (mean gamma %s, mean corr %s) vs Bonferroni %s vs separate 95%% bounds %s.\n"),
  paste(sprintf("%.3f", fC("naive")$cov1), collapse = "/"), paste(sprintf("%.2f", fC("naive")$r), collapse = "/"),
  paste(sprintf("%.3f", fC("IJ two-term")$cov1), collapse = "/"), paste(sprintf("%.2f", fC("IJ two-term")$r), collapse = "/"),
  paste(sprintf("%.3f", fC("field")$cov1), collapse = "/"), paste(sprintf("%.2f", fC("field")$r), collapse = "/"),
  paste(sprintf("%.3f", fH("naive")$cov1), collapse = "/"),
  paste(sprintf("%.3f", fH("IJ two-term")$cov1), collapse = "/"), paste(sprintf("%.2f", fH("IJ two-term")$r), collapse = "/"),
  paste(sprintf("%.3f", fH("field")$cov1), collapse = "/"), paste(sprintf("%.2f", fH("field")$r), collapse = "/"),
  paste(sprintf("%.3f", cal$joint_cov), collapse = "/"), paste(sprintf("%.4f", cal$mean_gamma), collapse = "/"), paste(sprintf("%+.3f", cal$mean_corr), collapse = "/"),
  paste(sprintf("%.3f", bon$joint_cov), collapse = "/"), paste(sprintf("%.3f", sep$joint_cov), collapse = "/")))
```
