---
title: "Simulation Studies Based on ACTG175: Binary Endpoint (GLM)"
subtitle: "Benefit Subgroup Identification with CD4 Response at 20 Weeks -- Raw Continuous Covariates"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# -- Simulation control -------------------------------------------------------
nsims_alt   <- 200L
nsims_null  <- 200L
sim_seed    <- 8316951L
n_workers   <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))

# -- FS diagnostic capture (priority D) ---------------------------------------
# run_simulation_analysis() can attach the full forestsearch() and GRF
# result objects to the first N replicates via keep = / keep_first_n =.
# Set diag_keep_first_n > 0 to enable diagnostic capture; set to 0L to
# disable (and save memory).  Default 20L gives Section 7's per-sim detail
# print and the cross-sim subgroup frequency summary in @sec-results.
diag_keep_first_n <- 20L
diag_keep         <- c("fs_full", "candidate_table")

# -- Subgroup-distribution plot thresholds (Section 8.2) ----------------------
# plot_sg_distribution() displays only subgroup labels that appear in at
# least min_pct of the detecting sims.  Under H1 (many finds) a larger
# threshold keeps the bar chart readable; under H0 (rare finds, heavy
# fragmentation across labels) a smaller threshold keeps any bars at all
# above the line.  Adjust to taste.  Values are in percent (0-100).
sgdist_min_pct_alt  <- 0   # H1: identified & complement bars
sgdist_min_pct_null <- 0   # H0: identified & complement bars

# Cap on the number of bars shown in the subgroup-distribution charts.
# If more labels clear the min_pct threshold, the top sgdist_top_k are
# shown and the rest are pooled into a single "Other (k labels)" bar.
# Prevents unreadable charts under H0 when finds are fragmented across
# many singleton labels.
sgdist_top_k         <- 15L

# Whether to render the pooled "Other (k labels)" bar alongside the top
# sgdist_top_k.  Set FALSE to drop the pooled bar entirely, which keeps
# the axis tightly focused on the top labels.  Useful under H0 when the
# "Other" bar is so large that it compresses the top labels visually.
# The subtitle continues to report the suppressed count for honesty.
# Per-scenario so you can show the pooled bar under H1 (where it is
# typically small and informative) and hide it under H0 (where it may
# dominate).
sgdist_show_other_alt  <- FALSE
sgdist_show_other_null <- FALSE

# -- Effect-distribution violin y-axis limits (Section 8.2.3) -----------------
# plot_effect_distribution() can produce extreme OR values that dominate the
# display.  ylim clips the viewport via coord_cartesian (data are NOT dropped;
# violins, boxplots, and mean/SD annotations use the full dataset).
# Set NULL to let ggplot2 determine limits automatically.
ordist_ylim <- c(0, 6)

# -- Production projection targets --------------------------------------------
prod_alt               <- 5000L
prod_null              <- 25000L


# -- Dataset preparation: ACTG175 ---------------------------------------------
#
# ARMS (matching Leon et al. 2024, Section 4.2):
#   speff2trial::ACTG175 arm coding:
#     0 = zidovudine (ZDV) alone
#     1 = ZDV + didanosine (ZDV+ddI)   <- experimental
#     2 = ZDV + zalcitabine (ZDV+ddC)
#     3 = didanosine (ddI) alone        <- control
#
# The paper compared arm 1 (ZDV+ddI, combination) vs arm 3 (ddI,
# monotherapy) and searched for a BENEFIT subgroup Q by switching
# treatment roles.
#
# True subgroup anchor (used for DGM construction and ground-truth
# classification ONLY; NOT exposed to the analysis methods):
#   z1 (age):     age > 34        -- median split
#   z2 (preanti): preanti <= 745  -- q3 split
#
# Analysis candidate covariates are provided as described in the
# "Realistic analysis setting" block below.
actg_arms        <- c(1L, 3L)
actg_treat_arm   <- 1L
actg_age_cut     <- 34
actg_preanti_cut <- 744.5

# -- Benefit search via treatment switching ------------------------------------
# This vignette implements a BENEFIT search for the experimental treatment
# (ZDV+ddI) using two combined conventions, so the package's harm-search
# machinery points in the right direction (full table in the Appendix):
#
#   1. Treatment is switched (ddI = 1, ZDV+ddI = 0) — so a benefit subgroup
#      of the original experimental appears as a "harm" subgroup of the
#      switched-treatment.
#
#   2. The binary outcome is encoded on the IMPROVEMENT scale: y_binary = 1
#      when the patient improved (cd420 > cd40).
#
# Under these conventions, OR(Q) > c1 in the analysis means: the switched
# treatment (ddI) increases the odds of improvement in Q — equivalently
# ZDV+ddI decreases the odds of improvement in Q — i.e. ZDV+ddI harms
# in Q.  This is the benefit subgroup of ZDV+ddI.  Thresholds:
#   c1 = fs_or_threshold    (screening)
#   c2 = fs_or_consistency  (consistency)
# After identification, ORs are inverted to the original scale via
# subgroup_notation = "benefit".

# -- DGM construction ---------------------------------------------------------
# The DGM uses the discretised z1-z12 factor formulation because
# generate_glm_dgm() requires factor covariates.  The TRUE benefit
# subgroup is defined by {z1 = 1 AND z2 = 1}, i.e. age > 34 AND
# preanti <= 744.5.  This provides the ground truth for the
# sensitivity / PPV computations in run_simulation_analysis().
dgm_factor_vars   <- paste0("z", 1:12)
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)
dgm_n_super       <- 5000L

# -- Treatment-effect scaling (k_treat) ---------------------------------------
# k_treat scales the fitted treatment main effect on the linear-predictor
# (log-odds) scale inside generate_glm_dgm():
#
#   eta_1 = eta_0 + k_treat * (eta_1_fitted - eta_0_fitted)
#
#   k_treat = 1 (default) preserves the fitted ACTG175 treatment effect.
#   k_treat > 1 amplifies the marginal ITT effect.
#   k_treat < 1 attenuates it.
#
# Two scenario-specific knobs so the H0 and H1 simulations can carry
# different baseline marginal effects independently:
#
#   dgm_k_treat_alt  -- applied to the calibrated alt DGM (used by both
#                       the initial alt build and the calibration grid).
#                       Calibration searches k_inter at this k_treat to
#                       hit cal_target_or, so changing this changes the
#                       k_inter that maps to OR(Q) = 2.0.
#
#   dgm_k_treat_null -- applied to the null (homogeneous-effect) DGM.
#                       Note: the null model represents a HOMOGENEOUS
#                       treatment effect (not a zero effect); k_treat
#                       scales the magnitude of that homogeneous effect.
#                       Useful for testing FPR under a stronger common
#                       benefit baseline.
#
# Both default to 1 so the existing operating-characteristic results are
# unchanged.  Set to a non-unit value to explore sensitivity.
dgm_k_treat_alt   <- 1.0
dgm_k_treat_null  <- 1.0

# -- Realistic analysis setting -----------------------------------------------
# The simulation mimics a real data analysis in which the analyst does
# NOT know which variables define Q nor the cuts that identify it.
# Continuous covariates are therefore supplied to forestsearch() and
# grf.subg.harm.glm() in their RAW form; the search methods discover
# cuts themselves (forestsearch: median / GRF-derived / forced cuts;
# grf.subg.harm.glm: policy-tree splits).
#
# NOTE: the true subgroup Q still involves age (continuous, cut at 34)
# and preanti (continuous, cut at 744.5).  Forestsearch will attempt
# cuts at medians and GRF-selected thresholds, which will generally
# DIFFER from the true cuts.  This is the realistic setting and is
# precisely what we want to evaluate: the method's ability to recover
# an effect subgroup when the cutpoints are unknown.
analysis_continuous_vars <- c(
  "age",      # continuous; TRUE cut (34) unknown to analyst
  "preanti",  # continuous; TRUE cut (744.5) unknown to analyst
  "wtkg",     # continuous
  "karnof",   # Karnofsky score (integer 70-100, treated as continuous)
  "cd40",     # baseline CD4 (continuous)
  "cd80"      # baseline CD8 (continuous)
)

# Truly binary indicators from ACTG175 (passed as factors to the
# analysis; no cut to discover).
analysis_binary_vars <- c(
  "hemo",     # hemophilia (0/1)
  "homo",     # homosexual activity (0/1)
  "drugs",    # IV drug use history (0/1)
  "race",     # race (0/1, white vs non-white as coded)
  "gender",   # gender (0/1)
  "symptom"   # symptomatic (0/1)
)

# -- Calibration ---------------------------------------------------------------
# Target OR(Q) = 2.0 on the switched scale (ddI = 1): in Q, ddI doubles
# the odds of improvement compared to ZDV+ddI.  Equivalently, ZDV+ddI
# has OR(Q) = 1/2.0 = 0.5 (halves improvement) — the benefit signal.
cal_target_or     <- 2.0
cal_k_grid_range  <- c(0, 3)
cal_grid_step     <- 0.05

# -- Simulation study ----------------------------------------------------------
sim_n_sample      <- 1000L
sim_n_noise       <- 0L

# -- Analysis flags ------------------------------------------------------------
run_fs            <- TRUE    # Run ForestSearch
run_fs_grf        <- FALSE   # Run FS with GRF pre-screening only
run_grf           <- TRUE    # Standalone GRF (grf.subg.harm.glm)

# -- ForestSearch parameters (benefit search thresholds) -----------------------
fs_or_threshold        <- round(1 / 0.60, 3)   # c1
fs_or_consistency      <- round(1 / 0.75, 3)   # c2
fs_pconsistency        <- 0.80
fs_splits              <- 1000L
fs_n_min               <- 60L
fs_d0_min              <- 10L
fs_d1_min              <- 10L
fs_maxk                <- 2L
fs_vi_grf_min          <- -0.2
fs_seedit              <- 8316951L
fs_cont_cutoff         <- 4L   # is.continuous() threshold (# unique values)

# -- FS-internal dmin.grf -----------------------------------------------------
# fs_dmin_grf is the dmin.grf value passed to forestsearch().  Inside
# forestsearch(), this parameter forwards to grf.subg.harm.glm() during
# the internal GRF candidate-cut extraction step (use_grf = TRUE).  The
# parameter MEANING is identical to the standalone GRF dmin.grf
# (minimum absolute CATE for a policy-tree node), but its ROLE differs:
#
#   - Standalone GRF  : dmin.grf is the FINAL subgroup-identification
#                       threshold.  Higher = stricter final subgroup.
#                       Controlled by grf_dmin (below).
#
#   - FS-internal GRF : dmin.grf is a CANDIDATE-CUT FILTER.  Cuts that
#                       fail this threshold are not passed back to FS's
#                       downstream enumeration.  Higher = fewer cuts
#                       reach FS, which may improve speed but can also
#                       remove cuts that FS would have used productively.
#                       Controlled by fs_dmin_grf (this variable).
#
# These two values are tuned independently.  The default 0.0 reproduces
# forestsearch()'s GLM default (any positive harm qualifies), so the
# existing simulation results are bit-identical.  Calibrate this
# independently from grf_dmin if you want a stricter pre-filter on FS
# candidate cuts.
fs_dmin_grf            <- 0.0

# -- Candidate-selection criterion (sg_focus) ---------------------------------
# `sg_focus` controls how forestsearch() sorts the candidate subgroups that
# meet the screening threshold, and therefore which one becomes the
# identified subgroup.  Five options (see ?subgroup_consistency):
#   "hr"       : primary sort = Pcons desc, secondary = effect size desc,
#                tertiary = K asc.  Package default.
#   "maxSG"    : primary sort = N desc (LARGEST candidate first),
#                secondary = Pcons desc, tertiary = K asc.  Use this when
#                the goal is to identify the LARGEST benefitting subgroup
#                among those meeting the consistency threshold.
#   "minSG"    : primary sort = N asc (smallest candidate first).
#   "hrMaxSG"  : compound (N desc, effect desc) — forces stop_threshold = NULL.
#   "hrMinSG"  : compound (N asc, effect desc) — forces stop_threshold = NULL.
#
# For this ACTG175 binary-benefit study we want to recover the LARGEST
# subgroup in which the ZDV+ddI combination is superior to ddI monotherapy,
# so we use "maxSG".  Early stopping is valid under "maxSG" (single sort
# criterion), though stop_threshold is set to NULL throughout this document.
fs_sg_focus            <- "maxSG"

# -- Continuous-variable cut strategy -----------------------------------------
# cut_type controls how candidate cuts are generated for each continuous
# covariate that forestsearch detects (i.e. any variable with more than
# fs_cont_cutoff unique values that is passed through confounders.name).
#
#   "default" : per continuous variable, get_conf_force() -> cut_var()
#               generates FOUR candidate cuts:
#                   var <= mean(var)
#                   var <= median(var)    (= Q2)
#                   var <= qlow(var)      (= Q1, 25th percentile)
#                   var <= qhigh(var)     (= Q3, 75th percentile)
#               This is the richer strategy and is the package default.
#
#   "median"  : per continuous variable, a SINGLE candidate cut:
#                   var <= median(var)
#               Use this to stress-test the "truly unknown cuts" regime
#               when the true cut is not at the median.
#
# Honest disclosure for this setup:
#   The true Q-defining cuts are age > 34 (~median of arms 1+3) and
#   preanti <= 744.5 (~Q3 of arms 1+3).  With cut_type = "default",
#   BOTH true cuts are present in the candidate set (age at median,
#   preanti at Q3).  With cut_type = "median", only the age cut is
#   recoverable; preanti's true Q3 cut is NOT in the candidate set,
#   forcing forestsearch to approximate Q via the wrong threshold.
#
# The default value ("default") is kept here so the simulation
# evaluates the realistic setting where the analyst uses the richer
# cut strategy a priori (without knowing the truth).  GRF candidate
# cuts (use_grf = TRUE) are additionally appended to this set.
fs_cut_type            <- "default"

# -- GRF standalone parameters ------------------------------------------------
# grf.subg.harm.glm() uses causal_forest() for candidate factor screening,
# then policy trees for subgroup identification.  With continuous inputs,
# policy trees automatically search over all candidate thresholds.
#
# dmin.grf: minimum absolute CATE (risk difference) for a policy-tree
#   node to qualify as a subgroup.  The calibration grids referenced
#   below were built on discretised inputs; with raw continuous inputs
#   the effective candidate-cut space is larger, which typically
#   INCREASES FPR slightly at a given dmin.grf.  Re-calibration is
#   recommended for production runs; dmin.grf = 0.10 is the current
#   conservative default carried over from the discretised setting.
grf_dmin               <- 0.1
grf_maxdepth           <- 2L
grf_sg_criterion       <- "mDiff"
grf_tune               <- FALSE

# -- ForestSearch pipeline flags -----------------------------------------------
fs_use_lasso           <- TRUE
fs_use_grf             <- TRUE
fs_is_rct              <- TRUE

# -- Single-inspection wrapper -------------------------------------------------
# fs_run() wraps forestsearch() with the parameter values defined above, so
# the two single-data / single-trial inspection calls (Sections 5 and 6) can
# share one definition instead of duplicating ~25 lines each.  The wrapper is
# intentionally narrow in scope: it is for interactive inspection runs where
# `details = TRUE` / `quiet = FALSE` are appropriate.  The simulation-loop
# path (Section 8) goes through run_simulation_analysis() instead, which
# constructs its own silent forestsearch() call internally.
#
# Arguments are the four column-name fields that differ between call sites;
# everything else (thresholds, pipeline flags, parallelism) is taken from
# the setup-chunk parameters.  Extra arguments are passed through to
# forestsearch() via `...` so a caller can override e.g. `maxk` without
# editing the wrapper.
fs_run <- function(data,
                   outcome_col,
                   event_col,
                   treat_col,
                   id_col = "id",
                   ...) {
  forestsearch(
    df.analysis      = data,
    confounders.name = confounders_analysis,
    outcome.name     = outcome_col,
    event.name       = event_col,
    treat.name       = treat_col,
    id.name          = id_col,
    outcome_type     = "binary",
    effect_measure   = "OR",
    seedit           = fs_seedit,
    n.min            = fs_n_min,
    d0.min           = fs_d0_min,
    d1.min           = fs_d1_min,
    hr.threshold     = fs_or_threshold,
    hr.consistency   = fs_or_consistency,
    pconsistency.threshold = fs_pconsistency,
    fs.splits        = fs_splits,
    max_subgroups_search = 15,
    maxk             = fs_maxk,
    sg_focus         = fs_sg_focus,
    use_lasso        = fs_use_lasso,
    use_grf          = fs_use_grf,
    is.RCT           = fs_is_rct,
    vi.grf.min       = fs_vi_grf_min,
    dmin.grf         = fs_dmin_grf,
    cont.cutoff      = fs_cont_cutoff,
    cut_type         = fs_cut_type,
    details          = TRUE,
    quiet            = FALSE,
    parallel_args    = list(plan = "sequential", workers = 1,
                             show_message = FALSE),
    ...
  )
}

# -- Reporting -----------------------------------------------------------------
sg_notation            <- "benefit"

# -- Output persistence --------------------------------------------------------
save_results           <- TRUE
output_dir             <- "../_data"

```

```{r timing-init}
#| include: false
t_vignette_start <- proc.time()
timings <- list()
```


# Introduction {#sec-intro}

## Motivation

This document extends the ACTG175 benefit-search simulation
framework from survival outcomes (time-to-event) to a **binary
endpoint**: whether a patient's CD4 count improved from baseline
to 20 weeks.  The outcome is encoded on the **improvement scale**:
$y\_binary = I(\text{cd420} > \text{cd40})$, where a 1 indicates
**improvement**.  Combined with treatment switching
(ddI $\to$ `treat = 1`, ZDV+ddI $\to$ `treat = 0`),
$\text{OR} > c_1$ in the analysis means ddI increases the odds
of improvement in the identified subgroup — equivalently, ZDV+ddI
*harms* improvement there.  ForestSearch identifies this as a harm
subgroup; `subgroup_notation = "benefit"` relabels it for display.

A full reference table of effect-measure / outcome-direction
combinations appears in the [Appendix](#sec-appendix-conventions).
@sec-stat-goals summarises the row that applies to this vignette.

The simulation uses the `forestsearch` GLM simulation suite
(`generate_glm_dgm`, `calibrate_glm_interaction`,
`simulate_from_glm_dgm`) and routes through
`run_simulation_analysis()` via class-based dispatch on the
`"glm_dgm"` object.

## Realistic Analysis Setting

A prior version of this simulation discretised every continuous
covariate into a binary factor **before** handing the data to
`forestsearch()` / `grf.subg.harm.glm()`.  The binarisation used
the same cuts that define the true subgroup $Q$ (e.g. age > 34,
preanti $\leq$ 744.5).  That setup implicitly gives the analysis
method knowledge of the true cutpoints, which is unrealistic.

The current setup **retains continuous covariates in their raw
form for the analysis step**.  The data-generating mechanism
(DGM) is built from the discretised factors (`z1`--`z12`) because
`generate_glm_dgm()` requires factor covariates and the true
subgroup must be defined algebraically.  The analysis methods,
however, see the **raw** continuous columns (`age`, `preanti`,
`wtkg`, `karnof`, `cd40`, `cd80`) along with the six truly
binary indicators (`hemo`, `homo`, `drugs`, `race`, `gender`,
`symptom`).  Cutpoints for the continuous covariates are
discovered by the search methods themselves (median splits,
GRF-derived cuts, policy-tree splits), exactly as in a real
analysis.

::: {.callout-important}
## What changes vs the prior (discretised) setup

- **Ground truth is unchanged**: the true benefit subgroup $Q$
  is still $\{\text{age} > 34\} \cap \{\text{preanti} \leq 744.5\}$
  with prevalence $\approx 34\%$, and the DGM's OR(Q) calibration
  is untouched.
- **Classification metrics are unchanged in definition**: `flag_harm`
  is computed identically from `z1` and `z2`.
- **What changes is the search problem**: ForestSearch must
  discover the cuts from candidate thresholds; GRF builds policy
  trees over raw continuous inputs.  Expect lower sensitivity and
  PPV than the discretised setting, because the analyst does not
  receive oracle cuts.

**Cut-set coverage for this run.**  ForestSearch is configured with
`cut_type = "default"` (the package default), which generates four
candidate cuts per continuous covariate: mean, median, Q1, and Q3.
For the ACTG175 covariate distribution the median of `age` is
approximately 34 and Q3 of `preanti` is approximately 744.5, so
**both true Q-defining cuts happen to fall inside the default
candidate set** — they are simply not privileged over the other
candidates.  The analyst has no oracle knowledge of which cuts
matter; ForestSearch must select them based on the data.  Switching
to `cut_type = "median"` would exclude preanti's Q3 cut from the
candidate set entirely, providing a harder stress test.  GRF policy
trees search over all continuous thresholds internally and are
unaffected by this setting.
:::

## Statistical Goals {#sec-stat-goals}

This simulation study addresses three questions about
ForestSearch and GRF standalone subgroup identification
applied to **binary outcomes** in the realistic (raw covariate)
setting:

1. **Detection power (H1)**: When a benefit subgroup $Q$
   exists with $\text{OR}(Q) = `r cal_target_or`$ -- i.e., in
   $Q$, the experimental treatment (ZDV+ddI) doubles the odds
   of improvement compared to ddI -- how often do ForestSearch
   and GRF identify *any* subgroup?
2. **False positive control (H0)**: When the treatment effect
   is homogeneous (no true $Q$), how often does each method
   falsely declare a benefit subgroup?
3. **Classification accuracy (H1)**: Conditional on detection,
   how well does the estimated $\hat{Q}$ recover the true $Q$?
   Measured by sensitivity ($\text{sens}(\hat{Q})$: fraction of
   true $Q$ captured) and positive predictive value
   ($\text{ppv}(\hat{Q})$: fraction of $\hat{Q}$ truly in $Q$).

::: {.callout-note}
## Directional configuration for this vignette

| Item | Value |
|------|-------|
| Outcome type | Binary |
| Native effect measure | OR |
| Y direction in the data | Improvement (`y_binary = 1` $\Leftrightarrow$ cd420 > cd40) |
| Search goal | Benefit (of ZDV+ddI) |
| Treatment coding | Switched (ddI = 1, ZDV+ddI = 0) |
| What "$\text{OR} > c_1$" means here | ddI increases P(improvement) in Q $\Leftrightarrow$ ZDV+ddI harms in Q |
| `subgroup_notation` | `"benefit"` |

This is one row of the Appendix master table
([@sec-appendix-conventions]) -- consult that table for any
other (outcome type $\times$ Y direction $\times$ search goal)
combination.
:::

::: {.callout-tip}
## Summary of results
See [Summary] at the end for a data-driven
summary of detection, classification, and timing from this run.
:::

## Simulation Pipeline

The simulation proceeds in six steps:

1. **Data preparation** (@sec-data): Load the ACTG175 trial,
   restrict to arms 1 (ZDV+ddI) and 3 (ddI), switch treatment
   labels so ddI = `treat = 1` and ZDV+ddI = `treat = 0`
   (benefit search convention), and construct BOTH:
   (a) the binary `z1`--`z12` factor set needed by the DGM; and
   (b) the realistic analysis candidate set
   (`age`, `preanti`, `wtkg`, `karnof`, `cd40`, `cd80`, `hemo`,
   `homo`, `drugs`, `race`, `gender`, `symptom`).

2. **DGM construction** (@sec-dgm): Fit a logistic regression
   to the observed data using the `z1`--`z12` factor
   formulation, create a 5,000-subject super-population, and
   define the true benefit subgroup
   $Q = \{z_1 = 1 \cap z_2 = 1\}$ (age > `r actg_age_cut`
   AND preanti $\leq$ `r actg_preanti_cut`).  The super-
   population retains the **raw** covariate columns as well as
   the `z1`--`z12` indicators.

3. **Calibration** (@sec-calibrate): Search over `k_inter`
   values to find the interaction strength that produces
   $\text{OR}(Q) = `r cal_target_or`$ -- a 2$\times$ odds-ratio
   of improvement for the experimental treatment in $Q$ (see
   @sec-stat-goals for the directional convention).

4. **Single-data analysis** (@sec-single): Run ForestSearch
   and GRF on the observed ACTG175 data using the **raw**
   analysis candidate set to verify the pipeline produces
   valid results before scaling to simulation.

5. **Monte Carlo simulation** (@sec-sims): For each of
   `r nsims_alt` alternative and `r nsims_null` null
   replicates, draw a random sample of $N = `r sim_n_sample`$
   from the DGM super-population, run ForestSearch and GRF
   on the raw analysis covariates, and record detection,
   classification, and effect estimates.

6. **Evaluation** (@sec-results): Compute operating
   characteristics (detection, sensitivity, PPV, effect
   estimation bias) under both hypotheses.

## Benefit Search on the OR Scale

The benefit-search approach from Leon et al. (2024, Section 4.2)
applies identically to binary outcomes, with odds ratios (OR)
replacing hazard ratios (HR).  The full configuration is shown in
the directional table in @sec-stat-goals and in the master
reference table in @sec-appendix-conventions; in brief, the
combination of switched treatment and adverse-coded `y_binary`
makes the package's harm-search machinery point at the benefit
direction, with screening threshold $c_1 =$ `r fs_or_threshold`
and consistency threshold $c_2 =$ `r fs_or_consistency` (both
sourced from the Quarto setup chunk).  Inversion to the original
scale is handled automatically by the package when
`subgroup_notation = "benefit"` is passed to the reporting
functions.


## Candidate Selection: `sg_focus = "maxSG"`

When several candidate subgroups simultaneously meet both the screening
threshold ($c_1 =$ `r fs_or_threshold`) and the consistency threshold
($c_2 =$ `r fs_or_consistency`), `forestsearch()` must pick one.  That
choice is governed by `sg_focus`.  Five options are available:

| `sg_focus`  | Primary sort   | Secondary        | Interpretation |
|-------------|----------------|------------------|----------------|
| `"hr"`      | Pcons      | effect       | Package default. Favours *most-consistent* candidate. |
| `"maxSG"`   | **N **     | Pcons        | **Largest** qualifying candidate. |
| `"minSG"`   | N           | Pcons        | Smallest qualifying candidate. |
| `"hrMaxSG"` | N          | effect       | Compound. Forces `stop_threshold = NULL`. |
| `"hrMinSG"` | N         | effect       | Compound. Forces `stop_threshold = NULL`. |

For the ACTG175 benefit search, the analytic goal is to recover the
**largest subgroup in which the ZDV+ddI combination is superior to ddI
monotherapy** — the population we would, in a real trial, most want to
describe and quantify.  Once the screening-plus-consistency filter has
passed, there is no further reason to pick a smaller, tighter subgroup;
the largest qualifying candidate is the most policy-relevant one.

We therefore use `fs_sg_focus <- "maxSG"` throughout this document (set
in the Quarto setup chunk). The concrete effect is:

- **Sort is by candidate N (descending)**, with Pcons as tiebreaker.
  Among candidates that pass screening and consistency, FS keeps the
  one with the largest $|H|$.
- **Single sort criterion**, so early stopping remains valid;
  `stop_threshold` can still be set if desired.  (In this document we
  set `stop_threshold = NULL` so all qualifying candidates are
  enumerated.)

This choice has two practical consequences that we track throughout
the operating-characteristics analysis:

1. $\widehat{|H|}$ will typically be closer to $|Q|$ than under the
   default `sg_focus = "hr"`, which trades size for consistency and
   tends to return strict subsets of $Q$.
2. The distribution of identified subgroups across simulations
   (reported in Section 8.2) should concentrate on a small number of
   definitions close to the true $Q$ anchor, rather than fragmenting
   into many small tight splits.


# GRF Standalone: Calibration Notes {#sec-grf-calibration}

The standalone GRF analysis uses `grf.subg.harm.glm()` with
`causal_forest()` for candidate variable screening and policy
trees for subgroup identification.  The key tuning parameter
is `dmin.grf`, the minimum absolute CATE (conditional average
treatment effect) on the **risk difference** scale required for
a policy-tree node to be valid.

## Configuration

For the **benefit + positive** scenario with switched treatment
(the ACTG175 binary configuration), the correct GRF setup is:

- `adverse_outcome = FALSE` (Y = improvement is positive)
- Treatment indicator = **switched** (ddI = 1, ZDV+ddI = 0)
- GRF identifies the subgroup where the switched "treatment"
  produces the most negative CATE --- i.e., where ddI monotherapy
  is worst relative to combination, which is the **benefit**
  subgroup on the original scale.

## Impact of Raw Continuous Inputs on Calibration

Prior `dmin.grf` calibration (in `grf_dmin_calibration_binary.qmd`)
was conducted on **discretised** inputs (`z1`--`z12` binary
factors) and recommended `dmin.grf = 0.02` for the benefit +
positive binary configuration.

With raw continuous inputs the policy-tree search evaluates
many more candidate thresholds per covariate, which tends to:

- **Widen the null distribution of the best CATE**, inflating
  the false positive rate at a given `dmin.grf`.
- **Improve sensitivity to the true cuts** when they align with
  natural splits in the covariate distribution.

The current document uses `dmin.grf = `r grf_dmin`` as a
conservative default carried over from the discretised setting.
A dedicated calibration run with raw continuous inputs is
recommended before production scaling.

::: {.callout-note}
The GRF FPR is higher than ForestSearch's because GRF lacks the
bootstrap consistency filter.  ForestSearch requires the subgroup
effect to exceed a threshold in 90% of bootstrap replicates,
providing stronger FPR control at the cost of lower detection.
The two methods are complementary: GRF provides a
threshold-free comparison.
:::


# Data Preparation {#sec-data}

```{r load-packages}
#| code-fold: true
#| code-summary: "Library loads (click to expand)"
library(forestsearch)
library(survival)
library(speff2trial)
library(data.table)
library(gt)
library(ggplot2)
library(doFuture)
library(foreach)

# Safe foreach result collection and worker resets are provided by
# forestsearch v0.2.0 via the exported helpers collect_results() and
# reset_workers() (R/forestsearch_helpers.R).  Call as
# forestsearch::collect_results(...) or just collect_results(...) after
# library(forestsearch).
```

```{r prepare-datasets}
#| code-fold: true
#| code-summary: "Data preparation (click to expand)"

t0 <- proc.time()[3]

# -- ACTG175: arms 1 (ZDV+ddI) vs 3 (ddI) ------------------------------------
actg_df <- subset(speff2trial::ACTG175, arms %in% actg_arms)
actg_df$id <- seq_len(nrow(actg_df))

# Original treatment: arm 1 = 1, arm 3 = 0
actg_df$treat_orig <- ifelse(actg_df$arms == actg_treat_arm, 1L, 0L)

# Switched treatment for benefit search
actg_df$treat <- 1L - actg_df$treat_orig

# -- Binary outcome: CD4 response at 20 weeks ---------------------------------
# Improvement scale: y_binary = 1 when the patient's CD4 count improved
# (cd420 > cd40).  With switched treatment (ddI = 1), OR > c1 in Q means
# ddI increases the odds of improvement — ForestSearch identifies this as
# the "harm" subgroup; subgroup_notation = "benefit" relabels for display.
actg_df$y_binary <- as.integer(actg_df$cd420 > actg_df$cd40)

# Remove subjects with missing cd420
n_before <- nrow(actg_df)
actg_df <- actg_df[!is.na(actg_df$cd420), ]
n_after  <- nrow(actg_df)

# =============================================================================
# Two parallel covariate views are constructed on the SAME data frame:
#
#   (1) DGM factor set: z1--z12 binary factors. Used by generate_glm_dgm()
#       and calibrate_glm_interaction() to fit the baseline logistic
#       regression and define the TRUE subgroup Q = {z1=1 & z2=1}.
#
#   (2) Analysis candidate set: raw continuous covariates + truly binary
#       indicators. Used by forestsearch() and grf.subg.harm.glm() in
#       both the observed-data analysis and every simulated replicate.
#       The analyst does NOT know the cuts that define Q.
# =============================================================================

# -- (1) DGM factor set: z1-z12 binary factors --------------------------------
#
# The z-series preserves the discretisation from León et al. (2024,
# Section 4.2). These are needed by the DGM because generate_glm_dgm()
# requires factor covariates, and because the true Q is defined
# algebraically in terms of z1 and z2.
actg_df$z1  <- as.factor(ifelse(actg_df$age > actg_age_cut, 1L, 0L))
actg_df$z2  <- as.factor(ifelse(actg_df$preanti <= actg_preanti_cut, 1L, 0L))
actg_df$z3  <- as.factor(ifelse(actg_df$wtkg <= 75, 1L, 0L))
actg_df$z4  <- as.factor(ifelse(
  actg_df$karnof <= median(actg_df$karnof), 1L, 0L))
actg_df$z5  <- as.factor(ifelse(
  actg_df$cd40 <= median(actg_df$cd40), 1L, 0L))
actg_df$z6  <- as.factor(ifelse(
  actg_df$cd80 <= median(actg_df$cd80), 1L, 0L))
actg_df$z7  <- as.factor(actg_df$hemo)
actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs)
actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender)
actg_df$z12 <- as.factor(actg_df$symptom)

# -- (2) Analysis candidate set: raw + binary ---------------------------------
#
# The continuous columns (age, preanti, wtkg, karnof, cd40, cd80) are
# ALREADY in the ACTG175 dataframe in their raw form. We make the six
# truly-binary indicators into factors so that forestsearch's
# is.continuous() detection treats them as categorical without
# attempting median/GRF cuts.
for (v in analysis_binary_vars) {
  actg_df[[v]] <- as.factor(actg_df[[v]])
}

confounders_analysis <- c(analysis_continuous_vars, analysis_binary_vars)

# Sanity: covariate-type inventory for the analyst-visible candidate set
cov_types <- vapply(confounders_analysis,
  function(v) if (is.factor(actg_df[[v]])) "factor" else class(actg_df[[v]])[1],
  character(1))
n_unique_per_cov <- vapply(confounders_analysis,
  function(v) length(unique(actg_df[[v]])), integer(1))

# -- True subgroup flag (for reference; generate_glm_dgm creates its own) -----
actg_df$flag_harm_ref <- as.integer(
  actg_df$z1 == 1 & actg_df$z2 == 1
)

# -- Summary scalars captured for downstream display ---------------------------
# y_binary = improvement: mean(y_binary) is the improvement rate directly.
n_ex_missing <- n_before - n_after
resp_overall <- 100 * mean(actg_df$y_binary)
resp_ZDV_ddI <- 100 * mean(actg_df$y_binary[actg_df$treat_orig == 1])
resp_ddI     <- 100 * mean(actg_df$y_binary[actg_df$treat_orig == 0])
Q_n          <- sum(actg_df$flag_harm_ref)
Q_N          <- nrow(actg_df)
Q_prev_pct   <- 100 * mean(actg_df$flag_harm_ref)

# Response-rate 2x2 grid (row = subgroup, col = treatment arm)
resp_grid <- data.frame(
  Subgroup  = c("Q (age > 34 & preanti \u2264 745)",
                "Qc (complement)"),
  `ZDV+ddI` = c(
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 1 & actg_df$treat_orig == 1]),
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 0 & actg_df$treat_orig == 1])),
  `ddI`     = c(
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 1 & actg_df$treat_orig == 0]),
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 0 & actg_df$treat_orig == 0])),
  check.names = FALSE
)

timings$data_prep <- proc.time()[3] - t0
```

::: {.callout-note title="Dataset summary"}
- **Source cohort**: ACTG175 arms 1 (ZDV+ddI, experimental) and 3
  (ddI, control). N = `r n_before` before exclusion,
  **N = `r Q_N`** after removing `r n_ex_missing` subject(s) with
  missing `cd420`.
- **Outcome**: binary response,
  `y_binary = I(cd420 > cd40)` (improvement scale).
  Overall **improvement** rate `r sprintf("%.1f%%", resp_overall)`;
  ZDV+ddI arm `r sprintf("%.1f%%", resp_ZDV_ddI)`; ddI arm
  `r sprintf("%.1f%%", resp_ddI)`.
- **True subgroup Q** (used for ground-truth classification only, not
  exposed to the analysis methods): age $>$ `r actg_age_cut` AND
  preanti $\le$ `r actg_preanti_cut`. Prevalence **`r Q_n` / `r Q_N`
  (`r sprintf("%.1f%%", Q_prev_pct)`)**.
- **Analysis candidate set** (`r length(confounders_analysis)` covariates):
  `r length(analysis_continuous_vars)` continuous
  (`r paste(analysis_continuous_vars, collapse = ", ")`) +
  `r length(analysis_binary_vars)` binary factors
  (`r paste(analysis_binary_vars, collapse = ", ")`).
- **DGM factor set**: z1–z12 binary factors, preserving the
  discretisation from León et al. (2024, Section 4.2). Kept on the same
  data frame so `generate_glm_dgm()` can build the super-population from
  the factor representation while the analysis methods see the raw
  continuous covariates.
:::

## Baseline Characteristics by Treatment Arm

```{r tbl1-demographics}
#| code-fold: true

# Label the original-scale arms for the header row (benefit-search
# "switched treatment" is a package-internal convention; the reader
# should see the original clinical labels in the demographics table).
actg_df$treat_label <- factor(
  ifelse(actg_df$treat_orig == 1L, "ZDV+ddI (experimental)",
                                    "ddI (control)"),
  levels = c("ddI (control)", "ZDV+ddI (experimental)")
)

create_summary_table(
  data             = actg_df,
  treat_var        = "treat_label",
  vars_continuous  = analysis_continuous_vars,
  vars_categorical = analysis_binary_vars,
  var_labels       = list(
    age     = "Age, years",
    preanti = "Days of prior antiretroviral therapy",
    wtkg    = "Weight, kg",
    karnof  = "Karnofsky score",
    cd40    = "Baseline CD4 count",
    cd80    = "Baseline CD8 count",
    hemo    = "Hemophilia",
    homo    = "Homosexual activity",
    drugs   = "IV drug use history",
    race    = "Non-white race",
    gender  = "Male",
    symptom = "Symptomatic"
  ),
  table_title      = "ACTG175: Baseline Characteristics by Treatment Arm",
  table_subtitle   = sprintf("N = %d (ZDV+ddI: %d; ddI: %d)",
                             nrow(actg_df),
                             sum(actg_df$treat_orig == 1L),
                             sum(actg_df$treat_orig == 0L)),
  source_note      = paste(
    "Analysis candidate set (what forestsearch / GRF see).",
    "Continuous variables are NOT discretised in the analysis step \u2014",
    "cut points are discovered by the methods."
  ),
  show_pvalue      = TRUE,
  show_smd         = TRUE,
  show_missing     = FALSE,
  font_size        = 11
)
```

## Response Rates by Subgroup and Arm

```{r tbl1-response-grid}
#| code-fold: true

gt::gt(resp_grid) |>
  gt::tab_header(
    title    = "Response Rate (%) by True Subgroup and Treatment Arm",
    subtitle = "Original treatment scale; subgroup Q = {age > 34} AND {preanti \u2264 745}"
  ) |>
  gt::fmt_number(columns = c("ZDV+ddI", "ddI"), decimals = 1) |>
  gt::cols_label(
    Subgroup    = gt::md("**Subgroup**"),
    `ZDV+ddI`   = gt::md("**ZDV+ddI**"),
    `ddI`       = gt::md("**ddI**")
  ) |>
  gt::tab_source_note(
    source_note = gt::md(sprintf(
      "For reference only \u2014 this is the observed *marginal* response-rate pattern in the true subgroup and its complement. The DGM calibrates to OR(Q) = %.2f on the switched (benefit-search) scale; see Section 4.",
      cal_target_or))
  ) |>
  gt::tab_options(table.font.size = 11)
```


# Creating the Binary DGM {#sec-dgm}

The GLM DGM uses `generate_glm_dgm()` to fit a logistic regression
on the observed ACTG175 data, create a super-population with
individual-level potential outcomes ($p_0$, $p_1$), and compute
true subgroup effects.  `calibrate_glm_interaction()` then searches
for the `k_inter` that produces the target OR(Q).

**Note on DGM inputs.**  The DGM continues to use the `z1`--`z12`
factor formulation.  This is a design requirement of
`generate_glm_dgm()` (which coerces any input to `factor` via
`as.factor()`) and is also the natural way to define the true
subgroup $Q$ algebraically.  The super-population returned by
the DGM *retains all original columns of `actg_df`*, including
the raw `age`, `preanti`, `wtkg`, `karnof`, `cd40`, `cd80`
columns, which are then used by the analysis step.

## Step 1: Build Initial DGM

```{r create-dgm-alt}
t0 <- proc.time()[3]

dgm_alt_init <- generate_glm_dgm(
  data          = actg_df,
  factor_vars   = dgm_factor_vars,
  outcome_var   = "y_binary",
  treatment_var = "treat",
  outcome_type  = "binary",
  effect_measure = "OR",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  model         = "alt",
  k_treat       = dgm_k_treat_alt,
  k_inter       = 0,    # no interaction before calibration
  n_super       = dgm_n_super,
  seed          = sim_seed,
  verbose       = TRUE
)

cat("\n=== Initial ACTG175 Binary DGM ===\n")
cat("  OR(Q):  ", round(dgm_alt_init$hazard_ratios$harm_subgroup, 3), "\n")
cat("  OR(Qc): ", round(dgm_alt_init$hazard_ratios$no_harm_subgroup, 3), "\n")
cat("  OR(ITT):", round(dgm_alt_init$hazard_ratios$overall, 3), "\n")

# Sanity check: super-population retains raw continuous columns
missing_raw <- setdiff(analysis_continuous_vars, names(dgm_alt_init$df_super))
if (length(missing_raw) > 0) {
  stop("DGM super-population is missing raw continuous columns: ",
       paste(missing_raw, collapse = ", "))
} else {
  cat("\nRaw continuous columns present in super-population: ",
      paste(analysis_continuous_vars, collapse = ", "), "\n")
}

timings$dgm_init <- proc.time()[3] - t0
```

## Step 2: Calibrate for Target OR {#sec-calibrate}

### Why calibrate?

The DGM's `k_inter` parameter controls the strength of the
treatment-by-subgroup interaction on the log-odds scale.  Because
the mapping from `k_inter` to the marginal OR(Q) is non-linear and
depends on the covariate distribution and baseline response rates,
we search over a grid of candidate values.

`calibrate_glm_interaction()` automates this process: it evaluates
`generate_glm_dgm()` at each grid point and selects the `k_inter`
closest to the target.

```{r calibrate}
t0 <- proc.time()[3]

dgm_calibrated <- calibrate_glm_interaction(
  data            = actg_df,
  factor_vars     = dgm_factor_vars,
  outcome_var     = "y_binary",
  treatment_var   = "treat",
  target_effect   = cal_target_or,
  outcome_type    = "binary",
  effect_measure  = "OR",
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  k_treat         = dgm_k_treat_alt,
  k_inter_range   = cal_k_grid_range,
  grid_step       = cal_grid_step,
  n_super         = dgm_n_super,
  seed            = sim_seed,
  verbose         = TRUE
)

cat("\n=== Calibrated ACTG175 Binary DGM ===\n")
cat("  OR(Q):  ", round(dgm_calibrated$hazard_ratios$harm_subgroup, 3),
    " (target:", cal_target_or, ")\n")
cat("  OR(Qc): ", round(dgm_calibrated$hazard_ratios$no_harm_subgroup, 3), "\n")
cat("  OR(ITT):", round(dgm_calibrated$hazard_ratios$overall, 3), "\n")
cat("  k_inter:", dgm_calibrated$model_params$k_inter, "\n")
cat("  -> Original scale: OR(Q) =",
    round(1 / dgm_calibrated$hazard_ratios$harm_subgroup, 3), "\n")

timings$calibrate <- proc.time()[3] - t0
```

### Calibration Curve

```{r calibration-plot, fig.height=5}
#| code-fold: true
#| code-summary: "Calibration-curve ggplot (click to expand)"
#| fig-cap: "Calibration of k_inter: OR(Q) on the improvement scale as a function of the interaction parameter."

cal_grid <- seq(cal_k_grid_range[1], cal_k_grid_range[2], by = cal_grid_step)
cal_or_Q <- vapply(cal_grid, function(k) {
  d <- generate_glm_dgm(
    data = actg_df, factor_vars = dgm_factor_vars,
    outcome_var = "y_binary", treatment_var = "treat",
    outcome_type = "binary", subgroup_vars = dgm_subgroup_vars,
    subgroup_cuts = dgm_subgroup_cuts, model = "alt",
    k_treat = dgm_k_treat_alt,
    k_inter = k, n_super = dgm_n_super, seed = sim_seed
  )
  d$hazard_ratios$harm_subgroup
}, numeric(1))

cal_df <- data.frame(k_inter = cal_grid, or_Q = cal_or_Q)
best_k <- dgm_calibrated$model_params$k_inter
or_Q_cal <- dgm_calibrated$hazard_ratios$harm_subgroup

ggplot(cal_df, aes(x = k_inter, y = or_Q)) +
  geom_line(linewidth = 1, colour = "firebrick") +
  geom_hline(yintercept = cal_target_or, linetype = "dashed",
             colour = "grey50") +
  geom_hline(yintercept = 1.0, linetype = "dotted",
             colour = "grey70") +
  geom_point(
    data = data.frame(k_inter = best_k, or_Q = or_Q_cal),
    colour = "red", size = 4, shape = 16
  ) +
  labs(
    x = "k_inter",
    y = "OR(Q), switched scale",
    title = "ACTG175 Binary DGM Calibration",
    subtitle = sprintf(
      "Best: k_inter = %.2f -> OR(Q) = %.3f (target: %.1f)",
      best_k, or_Q_cal, cal_target_or
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```


### Illustrative `k_treat` values: OR(ITT), OR(Q), and gap to c1

To give the reader a concrete sense of how `k_treat` shifts the
marginal and subgroup effects under H1, the table below holds the
calibrated `k_inter` fixed and varies `k_treat` over the same grid
$\{0.5, 1.0, 1.5\}$ as the null analog.  `k_treat` scales the
fitted main treatment effect directly, so $\text{OR}(\text{ITT})$
tracks `k_treat` smoothly; $\text{OR}(Q)$ inherits both the
scaled main effect and the held-fixed interaction, so it hits
the calibration target $\text{OR}(Q) = `r cal_target_or`$ only at
the calibration `k_treat` and drifts away as `k_treat` is
perturbed.  **Larger** `k_treat` amplifies the main effect,
pushing both $\text{OR}(\text{ITT})$ and $\text{OR}(Q)$ further
above c1 (improving FS detection); **smaller** `k_treat`
attenuates and may pull $\text{OR}(Q)$ back toward c1
(degrading detection).

```{r dgm-alt-illustrate}
#| code-fold: true
#| code-summary: "Build three illustrative alt DGMs and tabulate (click to expand)"

# Three illustrative alt DGMs.  k_inter is HELD FIXED at the value
# returned by calibration (dgm_calibrated$model_params$k_inter); only
# k_treat varies.  These DGMs are NOT used in the simulation loop
# below — the simulation uses dgm_calibrated, which is the calibrated
# alt DGM at k_treat = dgm_k_treat_alt.  The grid here is purely a
# didactic table.
k_inter_cal  <- dgm_calibrated$model_params$k_inter
k_treat_grid <- c(0.5, 1.0, 1.5)

illust_rows_alt <- lapply(k_treat_grid, function(k) {
  d <- generate_glm_dgm(
    data           = actg_df,
    factor_vars    = dgm_factor_vars,
    outcome_var    = "y_binary",
    treatment_var  = "treat",
    outcome_type   = "binary",
    effect_measure = "OR",
    subgroup_vars  = dgm_subgroup_vars,
    subgroup_cuts  = dgm_subgroup_cuts,
    model          = "alt",
    k_treat        = k,
    k_inter        = k_inter_cal,
    n_super        = dgm_n_super,
    seed           = sim_seed,
    verbose        = FALSE
  )
  or_ITT_switched <- as.numeric(d$hazard_ratios$overall)
  or_Q_switched   <- as.numeric(d$hazard_ratios$harm_subgroup)
  data.frame(
    k_treat       = k,
    or_itt_switch = or_ITT_switched,
    or_itt_orig   = 1 / or_ITT_switched,
    or_Q_switch   = or_Q_switched,
    or_Q_orig     = 1 / or_Q_switched,
    gap_to_c1     = fs_or_threshold - or_Q_switched
  )
})
illust_tbl_alt <- do.call(rbind, illust_rows_alt)

gt::gt(illust_tbl_alt) |>
  gt::tab_header(
    title    = "Illustrative alt DGMs: OR(ITT), OR(Q), and gap to c1 by k_treat",
    subtitle = sprintf(
      paste0("ACTG175 benefit search, switched scale; ",
             "c1 = %.3f, k_inter = %.3f (held fixed), n_super = %d"),
      fs_or_threshold, k_inter_cal, dgm_n_super
    )
  ) |>
  gt::tab_spanner(
    label   = "OR(ITT)",
    columns = c(or_itt_switch, or_itt_orig)
  ) |>
  gt::tab_spanner(
    label   = "OR(Q)",
    columns = c(or_Q_switch, or_Q_orig)
  ) |>
  gt::cols_label(
    k_treat       = "k_treat",
    or_itt_switch = "switched",
    or_itt_orig   = "original",
    or_Q_switch   = "switched",
    or_Q_orig     = "original",
    gap_to_c1     = gt::md("Gap (c1 \u2212 OR_Q_switched)")
  ) |>
  gt::fmt_number(
    columns  = c(or_itt_switch, or_itt_orig,
                 or_Q_switch,   or_Q_orig,   gap_to_c1),
    decimals = 3
  ) |>
  gt::fmt_number(columns = k_treat, decimals = 1) |>
  gt::tab_style(
    style    = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(rows = k_treat == 1.0)
  ) |>
  gt::tab_footnote(
    footnote = paste0(
      "k_treat = 1.0 (bold row) preserves the fitted main treatment ",
      "effect; this is the function default for generate_glm_dgm(). ",
      "This document's calibration was performed at k_treat = ",
      format(dgm_k_treat_alt, nsmall = 1),
      " (set in the setup chunk via dgm_k_treat_alt), and the resulting ",
      "k_inter = ", sprintf("%.3f", k_inter_cal),
      " is held fixed across all three rows.  k_treat scales the main ",
      "treatment effect directly, so OR(ITT) tracks k_treat smoothly; ",
      "OR(Q) hits the calibration target only at the calibration k_treat ",
      "because the held-fixed k_inter still adds to the linear predictor ",
      "for Q members regardless of k_treat. ",
      "Note: under H1, a NEGATIVE gap (last column) means OR(Q) exceeds ",
      "c1, which is detection-favorable -- the opposite sign convention ",
      "from the null table (where a positive gap is favorable, i.e., no ",
      "false alarm). Effects are computed exactly from the super-",
      "population's individual potential outcomes (p0, p1); marginal ORs ",
      "depend on the covariate distribution and are not, in general, ",
      "equal to any single conditional OR."
    )
  )
```


## Step 3: Create Null DGM

Under `model = "null"`, the treatment effect is **homogeneous**
across all patients (no treatment effect heterogeneity).
Per León et al. (2024, Section 3), $Q = \emptyset$ and
$Q^c$ is the ITT population.  The `flag_harm` column is
retained as a structural label in the simulated data (marking
$z_1 = 1 \cap z_2 = 1$) but has **no treatment-effect
meaning**: there is no true benefit subgroup under this
hypothesis.

```{r dgm-null}
t0 <- proc.time()[3]

dgm_null <- generate_glm_dgm(
  data          = actg_df,
  factor_vars   = dgm_factor_vars,
  outcome_var   = "y_binary",
  treatment_var = "treat",
  outcome_type  = "binary",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  model         = "null",
  k_treat       = dgm_k_treat_null,
  n_super       = dgm_n_super,
  seed          = sim_seed,
  verbose       = TRUE
)

cat("\n=== ACTG175 Null DGM ===\n")
cat("  OR(ITT):", round(dgm_null$hazard_ratios$overall, 3), "\n")
cat("  OR(Q):  ", round(dgm_null$hazard_ratios$harm_subgroup, 3), "\n")
cat("  OR(Qc): ", round(dgm_null$hazard_ratios$no_harm_subgroup, 3), "\n")

timings$dgm_null <- proc.time()[3] - t0
```


### Illustrative `k_treat` values: OR(ITT) and gap to c1

To give the reader a concrete sense of how `k_treat` controls the
marginal ITT effect and therefore the gap from c1, the table below
constructs three illustrative null DGMs at `k_treat` $\in \{0.5,
1.0, 1.5\}$ using the same data, factor list, super-population
size, and seed as the main `dgm_null` above.  The default
(`k_treat = 1.0`) preserves the fitted ITT effect; **larger**
`k_treat` amplifies the marginal benefit, pushing the
switched-scale OR further above 1 and *narrowing* the gap to c1;
**smaller** `k_treat` attenuates the marginal effect toward
OR $\approx 1$ and *widens* the gap.

```{r dgm-null-illustrate}
#| code-fold: true
#| code-summary: "Build three illustrative null DGMs and tabulate (click to expand)"

# Three illustrative null DGMs.  These are NOT used in the simulation
# loop below; the simulation uses `dgm_null` from the previous chunk
# (constructed with k_treat = dgm_k_treat_null).  The grid here is
# purely a didactic table so the reader can see how k_treat moves
# OR(ITT) and the gap to c1.
k_treat_grid <- c(0.5, 1.0, 1.5)

illust_rows <- lapply(k_treat_grid, function(k) {
  d <- generate_glm_dgm(
    data           = actg_df,
    factor_vars    = dgm_factor_vars,
    outcome_var    = "y_binary",
    treatment_var  = "treat",
    outcome_type   = "binary",
    effect_measure = "OR",
    subgroup_vars  = dgm_subgroup_vars,
    subgroup_cuts  = dgm_subgroup_cuts,
    model          = "null",
    k_treat        = k,
    n_super        = dgm_n_super,
    seed           = sim_seed,
    verbose        = FALSE
  )
  or_switched <- as.numeric(d$hazard_ratios$overall)
  data.frame(
    k_treat     = k,
    or_switched = or_switched,
    or_original = 1 / or_switched,
    gap_to_c1   = fs_or_threshold - or_switched
  )
})
illust_tbl <- do.call(rbind, illust_rows)

gt::gt(illust_tbl) |>
  gt::tab_header(
    title    = "Illustrative null DGMs: OR(ITT) and gap to c1 by k_treat",
    subtitle = sprintf(
      "ACTG175 benefit search, switched scale; c1 = %.3f, n_super = %d",
      fs_or_threshold, dgm_n_super
    )
  ) |>
  gt::cols_label(
    k_treat     = "k_treat",
    or_switched = "OR(ITT, switched)",
    or_original = "OR(ITT, original)",
    gap_to_c1   = gt::md("Gap (c1 \u2212 OR_switched)")
  ) |>
  gt::fmt_number(
    columns  = c(or_switched, or_original, gap_to_c1),
    decimals = 3
  ) |>
  gt::fmt_number(columns = k_treat, decimals = 1) |>
  gt::tab_style(
    style    = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(rows = k_treat == 1.0)
  ) |>
  gt::tab_footnote(
    footnote = paste0(
      "k_treat = 1.0 (bold row) preserves the fitted ITT effect; this is ",
      "the function default for generate_glm_dgm(). This document's null ",
      "simulation uses k_treat = ", format(dgm_k_treat_null, nsmall = 1),
      " (set in the setup chunk via dgm_k_treat_null). Smaller k_treat ",
      "attenuates the marginal effect toward OR \u2248 1 (widening the gap ",
      "to c1); larger k_treat amplifies the effect (narrowing the gap). ",
      "The realised OR is computed exactly from the super-population's ",
      "individual potential outcomes (p0, p1); the marginal OR depends ",
      "on the covariate distribution and is not, in general, equal to ",
      "any single conditional OR."
    )
  )
```


# Single-Data Analysis {#sec-single}

:::{.callout-important title="Real-data setting: no ground-truth classification"}
This section analyses the **actual ACTG175 data** -- not a simulated
trial.  The true benefit subgroup in a real trial is **unknown**; it is
the object the analysis is trying to estimate.  We therefore report the
subgroup ForestSearch identifies and the resulting arm-specific ORs,
but we **do not** compute classification metrics (sens / spec / PPV)
against any reference subgroup -- there is no clinical truth to
classify against.  Classification is only meaningful in the simulation
setting of @sec-sims, where the DGM fixes a known $Q$.
:::

Before any simulation, run ForestSearch on the **actual ACTG175
data** with `details = TRUE` to examine every step of the
procedure: LASSO screening, GRF variable importance, candidate
enumeration, and bootstrap consistency.

The analysis uses the **realistic candidate set**:

```{r show-analysis-set}
#| code-fold: true
#| code-summary: "Echo candidate covariate set (click to expand)"
#| output: false
cat("Analysis candidate covariates (no oracle cuts):\n")
cat("  Continuous:", paste(analysis_continuous_vars, collapse = ", "), "\n")
cat("  Binary:    ", paste(analysis_binary_vars, collapse = ", "), "\n")
cat("\nForestsearch / GRF will discover cuts on the continuous\n")
cat("covariates themselves; the TRUE cuts (age > 34, preanti <= 745)\n")
cat("are NOT provided.\n")
```

::: {.callout-note title="What the analysis methods see"}
The analysis candidate set (passed to `forestsearch()` / `grf.subg.harm.glm()`
as `confounders.name`) is:

- **Continuous** (`r length(analysis_continuous_vars)`):
  `r paste(analysis_continuous_vars, collapse = ", ")`
- **Binary** (`r length(analysis_binary_vars)`):
  `r paste(analysis_binary_vars, collapse = ", ")`

The methods discover cuts on the continuous covariates themselves;
the TRUE cuts (age $>$ `r actg_age_cut`, preanti $\le$
`r actg_preanti_cut`) are **not** provided.
:::

```{r single-data}
#| code-fold: true
#| code-summary: "fs_run() on real data (click to expand)"
t0 <- proc.time()[3]

# ForestSearch on ACTG175 observed data
#   outcome_type = 'binary', effect_measure = 'OR'
#   Treatment: switched (ddI = 1, ZDV+ddI = 0)
#   Y = I(cd420 > cd40) = improvement
#   Covariates: RAW continuous + binary indicators
#   Thresholds: c1 = fs_or_threshold, c2 = fs_or_consistency

fs_actg <- fs_run(
  data        = actg_df,
  outcome_col = "y_binary",
  event_col   = "y_binary",
  treat_col   = "treat",
  id_col      = "id"
)

timings$single_data <- proc.time()[3] - t0
```

## Examine Result

```{r examine-single}
#| code-fold: true
#| code-summary: "Examine FS result + subgroup OR fits (click to expand)"

cat("\n=== ForestSearch Result ===\n")

if (!is.null(fs_actg$grp.consistency)) {
  # NOTE: for a forestsearch object, the cut expressions defining the
  # identified subgroup live in `grp.consistency$sg.harm` (character
  # vector of factor-level cut names).  `grp.consistency$sg.harm.id`
  # is a different object: a per-subject 0/1 indicator of subgroup
  # membership with length nrow(df).  (For GRF objects, `sg.harm.id`
  # does carry the cut expressions — same name, different semantics.)
  sg <- fs_actg$grp.consistency$sg.harm
  if (!is.null(sg) && length(sg) > 0 && !all(is.na(sg))) {
    cat("Subgroup identified:", paste(sg, collapse = " & "), "\n")

    df_est <- fs_actg$df.est
    in_H  <- df_est$treat.recommend == 0
    in_Hc <- df_est$treat.recommend == 1

    cat(sprintf("  Size(H): %d,  Size(Hc): %d\n",
        sum(in_H), sum(in_Hc)))

    if (sum(in_H) >= 10) {
      fit_H <- glm(y_binary ~ treat, data = df_est[in_H, ],
                    family = binomial)
      or_H <- exp(coef(fit_H)["treat"])
      cat(sprintf("  OR(H):  %.3f\n", or_H))
    }
    if (sum(in_Hc) >= 10) {
      fit_Hc <- glm(y_binary ~ treat, data = df_est[in_Hc, ],
                     family = binomial)
      or_Hc <- exp(coef(fit_Hc)["treat"])
      cat(sprintf("  OR(Hc): %.3f\n", or_Hc))
    }

    fit_itt <- glm(y_binary ~ treat, data = df_est, family = binomial)
    cat(sprintf("  OR(ITT): %.3f\n",
        exp(coef(fit_itt)["treat"])))
  } else {
    cat("No subgroup identified by ForestSearch.\n")
  }
} else {
  cat("ForestSearch did not return grp.consistency.\n")
}

cat(sprintf("\nSingle-data analysis: %.1f sec\n", timings$single_data))
```

## GRF Standalone on Observed Data

```{r grf-single-data}
#| code-fold: true
#| code-summary: "GRF standalone call + output (click to expand)"

grf_actg <- grf.subg.harm.glm(
  data             = actg_df,
  confounders.name = confounders_analysis,
  outcome.name     = "y_binary",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "OR",
  dmin.grf         = grf_dmin,
  maxdepth         = grf_maxdepth,
  RCT              = TRUE,
  sg.criterion     = grf_sg_criterion,
  details          = TRUE
)

if (!is.null(grf_actg$sg.harm.id) && length(grf_actg$sg.harm.id) > 0) {
  cat("GRF subgroup:", paste(grf_actg$sg.harm.id, collapse = " & "), "\n")
  grf_H <- grf_actg$data$treat.recommend == 0
  cat(sprintf("  Size(H): %d,  Size(Hc): %d\n",
      sum(grf_H), sum(!grf_H)))
} else {
  cat("GRF: no subgroup identified.\n")
}
```


## FS vs GRF Concordance on Observed Data

When a method does not detect a subgroup, its classification is
"all subjects to $Q^c$" (`treat.recommend = 1` for everyone).
`compare_fs_subgroups()` handles this automatically — no manual
fallback is needed.

```{r fs-grf-concordance}
sg_compare <- compare_fs_subgroups(
  fs1                = fs_actg,
  fs2                = grf_actg,
  label1             = "ForestSearch",
  label2             = "GRF Standalone",
  id.name            = "id",
  subgroup_notation  = sg_notation,
  title              = "FS vs GRF: Subgroup Membership Concordance",
  subtitle           = "ACTG175 observed data (binary benefit search, raw continuous)",
  font_size          = 13
)

sg_compare$table
```

```{r fs-grf-concordance-detail}
cat("Concordance statistics:\n")
cat(sprintf("  Agreement: %.1f%%\n",
    100 * sg_compare$concordance$agreement))
cat(sprintf("  Cohen's kappa: %.3f\n",
    sg_compare$concordance$kappa))
cat(sprintf("  Concordant: %d / Discordant: %d / Total: %d\n",
    sg_compare$concordance$n_agree,
    sg_compare$concordance$n_disagree,
    sg_compare$concordance$n_total))
```


# Single Trial Simulation {#sec-single-trial}

Before the full simulation study, simulate a **single replicate** from
the calibrated DGM and run the two analysis methods end-to-end with
`details = TRUE`.  This mirrors what every replicate inside the Monte
Carlo loop will see and lets us sanity-check the full pipeline --
LASSO screening, GRF pre-screening, candidate enumeration, consistency
evaluation, subgroup-effect estimation -- before burning the budget of
thousands of replicates.

Unlike @sec-single, the **true subgroup here is known by construction**
(it is the DGM-embedded Q).  Per-replicate classification metrics are
reported in aggregate for the full simulation loop in @sec-results;
here we show only the identified subgroup definitions and estimated
ORs so the pipeline can be inspected end-to-end on a single replicate.

```{r single-trial-simulate}
#| code-fold: true
#| code-summary: "Simulate one trial from the calibrated DGM (click to expand)"
t0 <- proc.time()[3]

# ── Simulate one trial from the calibrated DGM ─────────────────────────
df_single <- simulate_from_glm_dgm(
  dgm  = dgm_calibrated,
  n    = sim_n_sample,
  seed = sim_seed + 1L
)

cat(sprintf("Single trial: N = %d, response rate = %.1f%%\n",
    nrow(df_single), 100 * mean(df_single$y_sim)))
cat(sprintf("  treat allocation: %.1f%%\n",
    100 * mean(df_single$treat_sim)))
cat(sprintf("  True Q prevalence: %.1f%%\n",
    100 * mean(df_single$flag_harm)))

# ── Verify OR in the simulated data ───────────────────────────────────
fit_single <- glm(y_sim ~ treat_sim, data = df_single, family = binomial)
cat(sprintf("  Simulated ITT OR: %.3f (DGM target: %.3f)\n",
    exp(coef(fit_single)["treat_sim"]),
    dgm_calibrated$hazard_ratios$overall))

# ── Confirm raw continuous columns are present in simulated data ──────
missing_in_sim <- setdiff(confounders_analysis, names(df_single))
if (length(missing_in_sim) > 0) {
  stop("Simulated data is missing analysis columns: ",
       paste(missing_in_sim, collapse = ", "))
}
```

## ForestSearch on the Simulated Trial

```{r single-trial-fs}
#| code-fold: true
#| code-summary: "fs_run() on the simulated trial (click to expand)"

# Ensure a stable id column for compare_fs_subgroups() alignment
df_single$id <- seq_len(nrow(df_single))

fs_single <- fs_run(
  data        = df_single,
  outcome_col = "y_sim",
  event_col   = "y_sim",
  treat_col   = "treat_sim",
  id_col      = "id"
)
```

```{r single-trial-fs-summary}
#| code-fold: true
#| code-summary: "Examine FS result on simulated trial (click to expand)"

cat("\n=== ForestSearch on simulated trial ===\n")

if (!is.null(fs_single$sg.harm) && length(fs_single$sg.harm) > 0L &&
    !all(is.na(fs_single$sg.harm))) {

  cat("Subgroup identified (cut expressions):\n")
  for (s in fs_single$sg.harm) cat("  ", s, "\n")

  df_est <- fs_single$df.est
  in_H   <- df_est$treat.recommend == 0

  # Subgroup effect on switched and original scales
  if (sum(in_H) >= 10L) {
    fit_H <- glm(y_sim ~ treat_sim,
                 data = df_est[in_H, ], family = binomial)
    or_H  <- exp(coef(fit_H)["treat_sim"])
    cat(sprintf("\n  OR(H_hat, switched) = %.3f  (inverted: %.3f)\n",
        or_H, 1/or_H))
  }
  if (sum(!in_H) >= 10L) {
    fit_Hc <- glm(y_sim ~ treat_sim,
                  data = df_est[!in_H, ], family = binomial)
    cat(sprintf("  OR(Hc_hat, switched) = %.3f  (inverted: %.3f)\n",
        exp(coef(fit_Hc)["treat_sim"]),
        1/exp(coef(fit_Hc)["treat_sim"])))
  }

} else {
  cat("No subgroup identified by ForestSearch on this replicate.\n")
}
```

## GRF Standalone on the Simulated Trial

```{r single-trial-grf}
#| code-fold: true
#| code-summary: "GRF call + output on simulated trial (click to expand)"

grf_single <- grf.subg.harm.glm(
  data             = df_single,
  confounders.name = confounders_analysis,
  outcome.name     = "y_sim",
  treat.name       = "treat_sim",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "OR",
  dmin.grf         = grf_dmin,
  maxdepth         = grf_maxdepth,
  RCT              = TRUE,
  sg.criterion     = grf_sg_criterion,
  details          = TRUE
)

cat("\n=== GRF standalone on simulated trial ===\n")
if (!is.null(grf_single$sg.harm.id) &&
    length(grf_single$sg.harm.id) > 0L &&
    !all(is.na(grf_single$sg.harm.id))) {
  cat("GRF subgroup (cut expressions):\n")
  for (s in grf_single$sg.harm.id) cat("  ", s, "\n")

  grf_H  <- grf_single$data$treat.recommend == 0
  cat(sprintf("  Size(H): %d,  Size(Hc): %d\n",
      sum(grf_H), sum(!grf_H)))
} else {
  cat("GRF: no subgroup identified on this replicate.\n")
}

timings$single_trial <- proc.time()[3] - t0
cat(sprintf("\nSingle trial analysis: %.1f sec\n", timings$single_trial))
```

## FS vs GRF Concordance on the Simulated Trial

```{r single-trial-concordance}
#| code-fold: true
#| code-summary: "FS vs GRF agreement on simulated trial (click to expand)"
sg_compare_sim <- compare_fs_subgroups(
  fs1                = fs_single,
  fs2                = grf_single,
  label1             = "ForestSearch",
  label2             = "GRF Standalone",
  id.name            = "id",
  subgroup_notation  = sg_notation,
  title              = "FS vs GRF on simulated trial",
  subtitle           = sprintf("N = %d, simulated from calibrated DGM",
                               nrow(df_single)),
  font_size          = 13
)
sg_compare_sim$table
```


# Running Simulation Studies {#sec-sims}

## Setup

```{r sim-setup}
#| code-fold: true
#| code-summary: "FS / GRF parameter lists for run_simulation_analysis() (click to expand)"

plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")

# -- Candidate variable set (RAW continuous + binary) --------------------------
# This is the key behavioural change vs the prior discretised setup:
# forestsearch() and grf.subg.harm.glm() receive raw continuous inputs
# and must discover cuts.
cat("Analysis confounders passed to FS/GRF:\n")
cat("  ", paste(confounders_analysis, collapse = ", "), "\n")

# -- ForestSearch parameters ---------------------------------------------------
fs_params <- list(
  # Column name mappings
  outcome.name           = "y_sim",
  event.name             = "y_sim",
  treat.name             = "treat_sim",
  id.name                = "id",

  # GLM pipeline
  outcome_type           = "binary",
  effect_measure         = "OR",

  # Variable screening
  use_lasso              = fs_use_lasso,
  use_grf                = fs_use_grf,

  # Consistency algorithm (v0.2.0 default: 3-10x speedup)
  use_twostage           = TRUE,

  # Continuous-variable detection threshold
  cont.cutoff            = fs_cont_cutoff,

  # Candidate cut strategy for continuous covariates
  # ("default" = mean + median + Q1 + Q3 per continuous variable;
  #  "median"  = median only). See setup chunk for full explanation.
  cut_type               = fs_cut_type,

  # Subgroup detection thresholds
  # Note: hr.threshold/hr.consistency are the generic ForestSearch
  # parameter names regardless of endpoint type. For binary outcomes,
  # these represent OR thresholds on the switched scale.
  hr.threshold           = fs_or_threshold,
  hr.consistency         = fs_or_consistency,
  pconsistency.threshold = fs_pconsistency,

  # Candidate-selection criterion: "maxSG" prefers the LARGEST
  # candidate subgroup meeting the consistency threshold (see setup
  # chunk for full description of sg_focus options).
  sg_focus               = fs_sg_focus,

  # Computational constraints
  fs.splits              = fs_splits,
  n.min                  = fs_n_min,
  d0.min                 = fs_d0_min,
  d1.min                 = fs_d1_min,
  maxk                   = fs_maxk,
  vi.grf.min             = fs_vi_grf_min,
  dmin.grf               = fs_dmin_grf,   # FS-internal GRF cut filter
  seedit                 = fs_seedit,

  # GRF tuning (applies to ForestSearch's internal GRF pre-screening)
  tune_grf               = grf_tune,

  # Inner parallelism: sequential within each worker
  parallel_args          = list(
    plan = "sequential", workers = 1, show_message = FALSE
  )
)

# -- GRF standalone parameter overrides ----------------------------------------
grf_params <- list(
  dmin.grf     = grf_dmin,
  maxdepth     = grf_maxdepth,
  sg.criterion = grf_sg_criterion,
  tune_grf     = grf_tune
)
```


## Alternative Hypothesis Simulations

```{r run-alt}
#| code-fold: true
#| code-summary: "H1 simulation loop — foreach wiring + run_simulation_analysis() (click to expand)"
#| output: false
cat("Running", nsims_alt, "simulations under H1...\n")
if (diag_keep_first_n > 0L) {
  cat(sprintf("  Diagnostic capture: keep = {%s}, first %d sim(s)\n",
      paste(diag_keep, collapse = ", "), diag_keep_first_n))
}
t0 <- proc.time()[3]

results_alt_raw <- foreach(
  sim = seq_len(nsims_alt),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_calibrated,
    n_sample         = sim_n_sample,
    confounders_base = confounders_analysis,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    grf_params       = grf_params,
    keep             = diag_keep,
    keep_first_n     = diag_keep_first_n,
    verbose          = FALSE
  )
}

timings$sims_alt <- proc.time()[3] - t0

# Post-hoc failure accounting + diagnostic harvest
# keep_diagnostics = TRUE preserves per-replicate fs_full / candidate_table
# attached by run_simulation_analysis() (stripped by rbindlist otherwise).
results_alt <- collect_results(results_alt_raw, "H1 alt",
                               keep_diagnostics = diag_keep_first_n > 0L)
n_failed_alt <- attr(results_alt, "n_failed")
diagnostics_alt <- attr(results_alt, "diagnostics")

cat("Completed in", round(timings$sims_alt / 60, 1), "minutes\n")
cat("Results:", nrow(results_alt), "rows\n")
if (n_failed_alt > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_alt))
if (!is.null(diagnostics_alt))
  cat(sprintf("  Diagnostic objects captured: %d replicate(s)\n",
      length(diagnostics_alt)))
if (nrow(results_alt) > 0) {
  cat(sprintf("Detection rate: %.1f%%\n",
      100 * mean(results_alt$any.H[results_alt$analysis == "FS"])))
} else {
  cat("All simulations failed — see error messages above.\n")
  cat("Common causes:\n")
  cat("  1. devtools::document() -> devtools::install() -> restart R needed\n")
  cat("  2. Package version mismatch between interactive and worker sessions\n")
}
```


## ForestSearch Sensitivity Diagnostic {#sec-fs-sensitivity}

:::{.callout-note title="Why this section exists"}
Empirically, FS sensitivity in this configuration hovers around 20%.
That raises the question: **which subgroups is FS actually
identifying?**  If it consistently picks up one component of the true
subgroup $Q = \{\text{age} > 34 \cap \text{preanti} \le 745\}$ but
misses the other, the low sensitivity is mechanical (one-of-two cuts
→ ≈ 50% true-Q overlap, further diluted by cut-point mismatch).  If
it picks noise variables, the low sensitivity reflects a different
problem.  The diagnostics below separate these hypotheses.
:::

### Per-replicate subgroup definitions (first 20)

The `sg.def` column on `results_alt` carries the cut-expression string
of the FS-identified subgroup for **every** replicate (a v0.2.0
addition to `run_simulation_analysis()`).  We print the first 20
alongside the corresponding sensitivity / PPV / size for context.

```{r fs-diag-per-sim}
#| code-fold: true
#| code-summary: "First N FS subgroup definitions (click to expand)"

fs_rows <- results_alt[results_alt$analysis == "FS", ]
show_n  <- min(20L, nrow(fs_rows))

diag_tbl <- data.frame(
  sim      = fs_rows$sim[seq_len(show_n)],
  any.H    = fs_rows$any.H[seq_len(show_n)],
  size.H   = fs_rows$size.H[seq_len(show_n)],
  sens     = round(fs_rows$sens[seq_len(show_n)], 3),
  ppv      = round(fs_rows$ppv[seq_len(show_n)],  3),
  sg.def   = fs_rows$sg.def[seq_len(show_n)],
  stringsAsFactors = FALSE
)

diag_tbl$sg.def[!nzchar(diag_tbl$sg.def)] <- "(none identified)"

knitr::kable(
  diag_tbl,
  col.names = c("Sim", "any.H", "|H_hat|", "Sens", "PPV",
                "Identified subgroup"),
  caption = sprintf(
    "First %d H1 replicates: FS-identified subgroup definitions.",
    show_n),
  align = c("r", "c", "r", "r", "r", "l")
)
```

### Cross-replicate subgroup-definition frequency (all sims)

```{r fs-diag-def-freq}
#| code-fold: true
#| code-summary: "Cross-replicate subgroup-definition frequency (click to expand)"

all_defs <- fs_rows$sg.def
all_defs[!nzchar(all_defs)] <- "(none identified)"

def_freq <- sort(table(all_defs), decreasing = TRUE)
def_df <- data.frame(
  subgroup = names(def_freq),
  n        = as.integer(def_freq),
  pct      = round(100 * as.integer(def_freq) / length(all_defs), 1),
  stringsAsFactors = FALSE
)

# Show all unique subgroups (usually <= 20) or top 20
show_rows <- min(20L, nrow(def_df))
knitr::kable(
  def_df[seq_len(show_rows), ],
  col.names = c("Identified subgroup",
                "Replicates",
                "% of H1 sims"),
  caption = sprintf(
    "Frequency of FS-identified subgroup definitions across %d H1 replicates.",
    length(all_defs)),
  row.names = FALSE,
  align = c("l", "r", "r")
)
```

### Variable-level frequency

Each subgroup definition is a conjunction of one or more cut
expressions.  Here we parse out the **variable names** appearing in
those cuts and count how often each variable is selected.  If FS is
recovering one component of true-$Q$ but missing the other, this
table will show a strong imbalance between `age` and `preanti`.

```{r fs-diag-var-freq}
#| code-fold: true
#| code-summary: "Variable-level frequency across FS finds (click to expand)"

extract_vars <- function(sg_str) {
  if (!nzchar(sg_str)) return(character(0))
  parts <- strsplit(sg_str, "\\s*&\\s*", perl = TRUE)[[1]]
  # A cut expression may be wrapped in braces and optionally negated:
  #   "{age <= 34}"       -> variable "age"
  #   "!{age <= 34}"      -> variable "age"
  #   "preanti"           -> variable "preanti"   (plain indicator)
  # Strip the outer !/{ / } wrappers, then take the first whitespace-
  # delimited token — the variable name.
  vars <- vapply(parts, function(p) {
    p <- trimws(p)
    p <- sub("^!", "",  p)           # optional negation
    p <- sub("^\\{", "", p)           # optional opening brace
    p <- sub("\\}$", "", p)           # optional closing brace
    tok <- strsplit(trimws(p), "\\s+", perl = TRUE)[[1]][1]
    if (is.na(tok)) "" else tok
  }, character(1))
  unique(vars[nzchar(vars)])
}

all_var_uses <- unlist(lapply(fs_rows$sg.def, extract_vars))
if (length(all_var_uses) == 0L) {
  cat("No identified subgroups across any replicate.\n")
} else {
  var_freq <- sort(table(all_var_uses), decreasing = TRUE)
  n_detect <- sum(nzchar(fs_rows$sg.def))
  var_df <- data.frame(
    variable       = names(var_freq),
    n_times_used   = as.integer(var_freq),
    pct_of_detects = round(100 * as.integer(var_freq) / n_detect, 1),
    stringsAsFactors = FALSE
  )
  knitr::kable(
    var_df,
    col.names = c("Variable",
                  "Times used",
                  "% of detecting sims"),
    caption = sprintf(
      paste0("Variables appearing in FS-identified subgroups across ",
             "%d detecting replicates (of %d total)."),
      n_detect, nrow(fs_rows)),
    row.names = FALSE,
    align = c("l", "r", "r")
  )
}
```

### FS-pipeline diagnostic columns

These columns (populated by `run_simulation_analysis()` from the full
`forestsearch()` result per replicate) let us check whether FS is
completing the full search or exiting via early stopping, and how
many candidates it's evaluating.

```{r fs-diag-pipeline}
#| code-fold: true
#| code-summary: "FS diagnostic columns (click to expand)"

if ("n_candidates_evaluated" %in% names(fs_rows)) {
  pipe_df <- data.frame(
    metric = c("Pconsistency threshold met (n_passed > 0)",
               "Early stopping triggered",
               "Candidates evaluated (mean)",
               "Candidates total available (mean)",
               "FS runtime, minutes (mean)"),
    value  = c(
      sprintf("%.1f%% of sims",
              100 * mean(fs_rows$n_passed > 0, na.rm = TRUE)),
      sprintf("%.1f%% of sims",
              100 * mean(fs_rows$early_stop_triggered, na.rm = TRUE)),
      sprintf("%.1f",
              mean(fs_rows$n_candidates_evaluated, na.rm = TRUE)),
      sprintf("%.1f",
              mean(fs_rows$n_candidates_total, na.rm = TRUE)),
      sprintf("%.2f",
              mean(fs_rows$fs_minutes, na.rm = TRUE))
    ),
    stringsAsFactors = FALSE
  )
  knitr::kable(
    pipe_df,
    col.names = c("Diagnostic", "Value"),
    caption = "FS pipeline diagnostics aggregated across H1 replicates.",
    align = c("l", "r")
  )
}
```

### Full `fs_full` objects for the first few sims (if captured)

When `diag_keep_first_n > 0` and `"fs_full" %in% diag_keep`, the full
`forestsearch()` result objects (minus `df.est`/`df.predict` to keep
memory bounded) are attached as `attr(results_alt, "diagnostics")`,
keyed by `sim_id`.  Below we pull the first one and inspect the
candidate-evaluation table — which lists every subgroup that passed
the screening threshold and their consistency p-values.

```{r fs-diag-candidate-peek}
#| code-fold: true
#| code-summary: "Peek at full fs_full objects for first few sims (click to expand)"

if (!is.null(diagnostics_alt) && length(diagnostics_alt) > 0L) {
  # diagnostics_alt is named by sim_id: $sim1 / $sim2 / ...
  first_key <- names(diagnostics_alt)[1L]
  d <- diagnostics_alt[[first_key]]

  cat(sprintf("Captured diagnostics for %d replicate(s): %s ... %s\n",
      length(diagnostics_alt),
      utils::head(names(diagnostics_alt), 1L),
      utils::tail(names(diagnostics_alt), 1L)))
  cat(sprintf("\nInspecting keeper set for '%s':\n", first_key))
  cat("  methods:", paste(names(d), collapse = ", "), "\n")
  if (!is.null(d$FS)) {
    cat("  FS keepers: ", paste(names(d$FS), collapse = ", "), "\n")
  }

  # Show the FS candidate table for the first captured replicate
  if (!is.null(d$FS$candidate_table) &&
      inherits(d$FS$candidate_table, "data.frame") &&
      nrow(d$FS$candidate_table) > 0L) {
    cat(sprintf(
      "\nCandidate subgroups evaluated for %s (top 10 by row):\n",
      first_key))
    print(utils::head(d$FS$candidate_table, 10L))
  }
} else {
  cat("No diagnostics captured (diag_keep_first_n == 0 or all sims failed).\n")
}
```

### Interpretation

Read the four tables above top-to-bottom:

- If the **subgroup-definition frequency** shows a handful of
  recurring cut expressions dominating, FS is converging on a small
  number of candidates.  If one of them matches true-$Q$ closely
  (e.g., `age <= 34 & preanti <= 745`), sensitivity on those sims
  should be high in the per-sim table.
- If the **variable frequency** shows `age` dominating and `preanti`
  rarely present (or vice versa), FS is recovering **one component
  only**.  Sensitivity of ≈ 0.50 on that sub-population of sims is
  expected; weighted by the fraction of times both are found, the
  average lands near 20%.  This is a candidate-enumeration issue,
  not a pathology.
- If noise variables (`noise1`, …) show up prominently, the threshold
  or consistency settings are too permissive for this sample size.
- If `n_passed > 0` is low (say < 30%), the identification threshold
  `fs_or_threshold` is binding and the subgroup is being **rejected
  outright** in most sims.  This would show up as `size.H = 0` and
  `any.H = 0L` for those replicates in the per-sim table.

The combination of the four tables is enough to pick between these
hypotheses without re-running anything.




```{r reset-workers-null}
# Reset worker pool between alt and null phases to prevent memory
# accumulation and worker bloat (null sims on bloated workers can be
# ~20x slower).  reset_workers() is exported from forestsearch v0.2.0.
forestsearch::reset_workers(workers = n_workers)
```

```{r run-null}
#| code-fold: true
#| code-summary: "H0 simulation loop (click to expand)"
#| output: false
cat("Running", nsims_null, "simulations under H0...\n")
t0 <- proc.time()[3]

results_null_raw <- foreach(
  sim = seq_len(nsims_null),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_null,
    n_sample         = sim_n_sample,
    confounders_base = confounders_analysis,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    grf_params       = grf_params,
    verbose          = FALSE
  )
}

timings$sims_null <- proc.time()[3] - t0

results_null <- collect_results(results_null_raw, "H0 null")
n_failed_null <- attr(results_null, "n_failed")

cat("Completed in", round(timings$sims_null / 60, 1), "minutes\n")
if (n_failed_null > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_null))
cat(sprintf("False positive rate: %.1f%%\n",
    100 * mean(results_null$any.H[results_null$analysis == "FS"])))
```

```{r cleanup-plan}
plan(sequential)
```

```{r save-results}
#| code-fold: true
#| code-summary: "Bundle assembly & save (click to expand)"

save_info <- NULL  # populated below when save_results = TRUE

if (save_results) {
  if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

  bundle <- list(
    # -- Core results ----------------------------------------------------------
    results_alt  = results_alt,
    results_null = results_null,

    # -- DGM objects -----------------------------------------------------------
    dgm_calibrated = dgm_calibrated,
    dgm_null       = dgm_null,

    # -- Simulation configuration ----------------------------------------------
    config = list(
      # Scenario counts
      nsims_alt    = nsims_alt,
      nsims_null   = nsims_null,
      n_sample     = sim_n_sample,
      n_noise      = sim_n_noise,
      sim_seed     = sim_seed,
      n_workers    = n_workers,
      outcome_type = "binary",
      effect_measure = "OR",

      # Document the analysis candidate set used (important provenance
      # for comparing to prior discretised runs)
      analysis_continuous_vars = analysis_continuous_vars,
      analysis_binary_vars     = analysis_binary_vars,
      confounders_analysis     = confounders_analysis,

      # ForestSearch thresholds and criteria (kept as top-level scalars
      # for easy inspection; fs_params below holds the full list)
      fs_or_threshold    = fs_or_threshold,
      fs_or_consistency  = fs_or_consistency,
      fs_sg_focus        = fs_sg_focus,
      fs_pconsistency    = fs_pconsistency,
      fs_splits          = fs_splits,
      fs_maxk            = fs_maxk,
      fs_use_lasso       = fs_use_lasso,
      fs_use_grf         = fs_use_grf,
      fs_is_rct          = fs_is_rct,
      fs_vi_grf_min      = fs_vi_grf_min,
      fs_cut_type        = fs_cut_type,
      fs_cont_cutoff     = fs_cont_cutoff,
      fs_n_min           = fs_n_min,
      fs_d0_min          = fs_d0_min,
      fs_d1_min          = fs_d1_min,
      fs_seedit          = fs_seedit,

      # GRF standalone parameters
      grf_dmin           = grf_dmin,
      grf_maxdepth       = grf_maxdepth,
      grf_sg_criterion   = grf_sg_criterion,
      grf_tune           = grf_tune,

      # Reporting parameters
      sg_notation        = sg_notation,
      sgdist_min_pct_alt  = sgdist_min_pct_alt,
      sgdist_min_pct_null = sgdist_min_pct_null,
      sgdist_top_k        = sgdist_top_k,
      sgdist_show_other_alt  = sgdist_show_other_alt,
      sgdist_show_other_null = sgdist_show_other_null,

      # Production projection targets (carried for reference only;
      # the run itself uses nsims_alt / nsims_null above)
      prod_alt  = prod_alt,
      prod_null = prod_null,

      # DGM construction parameters (for replayability)
      dgm_n_super        = dgm_n_super,
      dgm_subgroup_vars  = dgm_subgroup_vars,
      dgm_subgroup_cuts  = dgm_subgroup_cuts,
      dgm_factor_vars    = dgm_factor_vars,
      cal_target_or      = cal_target_or,
      cal_k_grid_range   = cal_k_grid_range,
      cal_grid_step      = cal_grid_step,

      # Diagnostic capture (Section 7)
      diag_keep          = diag_keep,
      diag_keep_first_n  = diag_keep_first_n
    ),

    # -- ForestSearch parameters -----------------------------------------------
    fs_params = fs_params,

    # -- GRF parameters --------------------------------------------------------
    grf_params = grf_params,

    # -- DGM true effects (convenience) ----------------------------------------
    true_effects = list(
      or_Q    = dgm_calibrated$hazard_ratios$harm_subgroup,
      or_Qc   = dgm_calibrated$hazard_ratios$no_harm_subgroup,
      or_ITT  = dgm_calibrated$hazard_ratios$overall,
      k_inter = dgm_calibrated$model_params$k_inter
    ),

    # -- Provenance ------------------------------------------------------------
    timestamp   = Sys.time(),
    R_version   = R.version.string,
    fs_version  = tryCatch(
      as.character(utils::packageVersion("forestsearch")),
      error = function(e) "unknown"
    ),
    timings     = timings,

    # -- Failure accounting ----------------------------------------------------
    n_failed_alt  = n_failed_alt,
    n_failed_null = n_failed_null
  )

  filename <- sprintf(
    "actg175_binary_benefit_simulations_rawcont_%dalts_%dnulls.rds",
    nsims_alt, nsims_null
  )
  path <- file.path(output_dir, filename)
  saveRDS(bundle, path)

  save_info <- data.frame(
    Field = c("Path", "Size (KB)", "H1 sims (completed / failed)",
              "H0 sims (completed / failed)",
              "OR(Q) on switched scale", "OR(Qc) on switched scale"),
    Value = c(
      path,
      sprintf("%.1f", file.size(path) / 1024),
      sprintf("%d / %d",
              nsims_alt - n_failed_alt,  n_failed_alt),
      sprintf("%d / %d",
              nsims_null - n_failed_null, n_failed_null),
      sprintf("%.3f",
              dgm_calibrated$hazard_ratios$harm_subgroup),
      sprintf("%.3f",
              dgm_calibrated$hazard_ratios$no_harm_subgroup)
    ),
    stringsAsFactors = FALSE
  )
}
```

```{r save-results-confirmation}
#| code-fold: true
#| results: asis

if (!is.null(save_info)) {
  cat(gt::as_raw_html(
    gt::gt(save_info) |>
      gt::tab_header(title = "Simulation bundle saved") |>
      gt::cols_label(Field = gt::md("**Field**"),
                     Value = gt::md("**Value**")) |>
      gt::tab_options(table.font.size = 11)
  ), "\n\n")
} else {
  cat("*`save_results = FALSE`; bundle not written.*\n\n")
}
```


# Results {#sec-results}

## Metrics Glossary {#sec-glossary}

Classification metrics follow the definitions in León et al.
(2024, Section 3) adapted for the benefit search.

### Detection

**Found** ($\text{any}(Q)$): proportion of simulations where
the method identifies *any* benefit subgroup
($\texttt{any.H} = 1$).

- **Under H1** (alternative): $Q$ exists with a differential
  treatment effect.  $\text{any}(Q)$ is **detection power**.
- **Under H0** (null): $Q = \emptyset$ and $Q^c$ is the ITT
  population.  $\text{any}(Q)$ is the **false positive rate
  (FPR)** for falsely declaring that a benefit subgroup exists.

### Classification Under H1

Under the alternative, the true benefit subgroup $Q$ exists
($Q = \{z_1 = 1 \cap z_2 = 1\}$, prevalence $\approx 34\%$)
with a genuinely different treatment effect.  Following
León et al. (2024, Section 3), define for estimated
subgroup $\hat{Q}$:

$$
\text{sens}(\hat{Q}) = \frac{\#\{i \in \hat{Q} \cap Q\}}{\#\{i \in Q\}},
\quad
\text{ppv}(\hat{Q}) = \frac{\#\{i \in \hat{Q} \cap Q\}}{\#\{i \in \hat{Q}\}},
$$

with analogous measures for the complement $\hat{Q}^c$.

The operating-characteristics tables report these as:

| Table column | León et al. notation | Interpretation |
|---|---|---|
| **Sen** | $\text{sens}(\hat{Q})$ | Fraction of true $Q$ captured by $\hat{Q}$ |
| **Spec** | $\text{sens}(\hat{Q}^c)$ | Fraction of true $Q^c$ correctly in $\hat{Q}^c$ |
| **PPV** | $\text{ppv}(\hat{Q})$ | Fraction of $\hat{Q}$ truly in $Q$ |
| **NPV** | $\text{ppv}(\hat{Q}^c)$ | Fraction of $\hat{Q}^c$ truly in $Q^c$ |

Classification metrics are averaged across **all** simulations
(León et al. 2024, Table 1).  When no subgroup is detected
($\hat{Q} = \emptyset$), the simulation contributes
$\text{sens}(\hat{Q}) = 0$ and $\text{ppv}(\hat{Q}) = 0$
(since no member of $Q$ is captured and no subjects are
assigned to $\hat{Q}$).  This convention is implemented
at the source in `run_simulation_analysis()`, matching
the original `fs.estimates.out()` function used in
León et al. (2024).

::: {.callout-note}
## Classification metrics with discovered cuts

Because cuts are discovered from the data in this setup, a
$\hat{Q}$ that captures "roughly the right region" but with
slightly shifted boundaries will still receive high sens/PPV.
Expect these metrics to be lower than under the discretised
(oracle-cut) setup — the gap quantifies the cost of not knowing
the true cuts in advance.
:::

### Classification Under H0

Under the null, $Q = \emptyset$ --- no subgroup with a
differential treatment effect exists.  The complement $Q^c$
is the entire ITT population.  Per León et al. (2024,
Section 3), $\text{sens}(\hat{Q})$ is **undefined** (shown
as "—" in the table) and $\text{ppv}(\hat{Q}) \equiv 0$.
For the complement, $\text{sens}(\hat{Q}^c) = |\hat{Q}^c|/N$
and $\text{ppv}(\hat{Q}^c) = 1$ since $Q^c$ = ITT.

The only operationally meaningful metric under H0 is the
**detection rate (FPR)**.

### Effect Estimates

All effect estimates are displayed on the **original** treatment
scale (OR $< 1$ = benefit from ZDV+ddI) via
`subgroup_notation = "benefit"`.

## Operating Characteristics Under H1

```{r oc-alt}
t0 <- proc.time()[3]

format_oc_results(
  results_alt,
  metrics   = c("detection", "classification", "subgroup_size"),
  digits    = 2,
  digits_hr = 2,
  title     = "ACTG175 Binary: Identification (Benefit Search, Raw Continuous)",
  subtitle  = sprintf(
    "n = %d, %d sims, OR(Q) = %.2f, OR(Qc) = %.2f (switched scale)",
    sim_n_sample, nsims_alt,
    dgm_calibrated$hazard_ratios$harm_subgroup,
    dgm_calibrated$hazard_ratios$no_harm_subgroup
  ),
  subgroup_notation = sg_notation
)
```

::: {.callout-tip title="Interpreting high Sens / low PPV"}
The combination of high sensitivity (~85%) with low specificity (~20%)
and low PPV (~35%) is internally consistent and follows directly from
Bayes' theorem.  The Bayes relationship is:

$$
\text{PPV}
= \frac{\text{Sens} \times \pi}
       {\text{Sens} \times \pi + (1 - \text{Spec}) \times (1 - \pi)}
$$

where $\pi \approx 34\%$ is the true-$Q$ prevalence.  Substituting the
observed metrics confirms PPV $\approx$ 35%.  The pattern arises because
$\hat{Q}$ is **much larger than $Q$** --- roughly 2--2.5$\times$ the
true subgroup size (check avg$|\hat Q|$ above).  The method recovers
most of $Q$ (high sensitivity) but also absorbs most of $Q^c$ (low
specificity), yielding a "diluted" $\hat{Q}$ with low purity (PPV).
This is expected when cuts are discovered from continuous covariates
rather than known *a priori*; the raw-continuous boundary
$\{z_1 = 1 \cap z_2 = 1\}$ is harder to recover precisely than a
discretised oracle-cut design.
:::

### True Subgroup Sizes Across Replicates

`run_simulation_analysis()` records the per-replicate size of the
**true** benefit subgroup $Q$ (`flag_harm == 1`) and its complement
$Q^c$ (`flag_harm == 0`) in the columns `sizeH_true`, `sizeHc_true`,
`propH_true`, `propHc_true`.  The estimation tables below report
$\widehat{|H|}$ (the identified subgroup size averaged across
replicates); the callout here gives the corresponding true-subgroup
sizes for comparison.

```{r true-size-summary}
#| code-fold: true

# One row per replicate-method combination, but sizeH_true is identical
# across method rows within the same sim. Deduplicate on sim.
truth_by_sim <- unique(results_alt[, c("sim", "sizeH_true", "sizeHc_true",
                                        "propH_true", "propHc_true")])

callout_Q <- sprintf(
  "|Q|:  mean = %.1f (SD %.2f),  range = [%d, %d],  prop = %.1f%%",
  mean(truth_by_sim$sizeH_true),
  sd(truth_by_sim$sizeH_true),
  min(truth_by_sim$sizeH_true),
  max(truth_by_sim$sizeH_true),
  100 * mean(truth_by_sim$propH_true))

callout_Qc <- sprintf(
  "|Qc|: mean = %.1f (SD %.2f),  range = [%d, %d],  prop = %.1f%%",
  mean(truth_by_sim$sizeHc_true),
  sd(truth_by_sim$sizeHc_true),
  min(truth_by_sim$sizeHc_true),
  max(truth_by_sim$sizeHc_true),
  100 * mean(truth_by_sim$propHc_true))

cat(sprintf("True subgroup sizes across %d H1 replicates (N = %d each):\n",
            nrow(truth_by_sim), sim_n_sample))
cat("  ", callout_Q,  "\n")
cat("  ", callout_Qc, "\n")
```

::: {.callout-note title="Reference sizes for the estimation tables below"}
- **|Q| (true benefit subgroup)**: mean `r sprintf("%.1f", mean(truth_by_sim$sizeH_true))`
  per trial (SD `r sprintf("%.2f", sd(truth_by_sim$sizeH_true))`,
  range `r min(truth_by_sim$sizeH_true)`–`r max(truth_by_sim$sizeH_true)`,
  average proportion `r sprintf("%.1f%%", 100 * mean(truth_by_sim$propH_true))`).
- **|Qc| (complement)**: mean `r sprintf("%.1f", mean(truth_by_sim$sizeHc_true))`
  per trial (SD `r sprintf("%.2f", sd(truth_by_sim$sizeHc_true))`,
  range `r min(truth_by_sim$sizeHc_true)`–`r max(truth_by_sim$sizeHc_true)`,
  average proportion `r sprintf("%.1f%%", 100 * mean(truth_by_sim$propHc_true))`).

Compare to $\widehat{|H|}$ (the FS-identified subgroup size) reported in
the row-group labels of the estimation tables below.  A ratio
$\widehat{|H|} \big/ \overline{|Q|} \approx 1$ indicates the method is
recovering a subgroup close to the true size; values much smaller
suggest FS is identifying a strict subset of $Q$; values larger suggest
FS is absorbing some $Q^c$ subjects into $\hat H$.
:::

### Distribution of Identified Subgroups

::: {.callout-tip title="Reminder: the true subgroup"}
The true subgroup $Q$ that FS is trying to recover is defined by
$\{\text{age} > `r actg_age_cut`\} \cap
 \{\text{preanti} \le `r actg_preanti_cut`\}$.

- **Observed ACTG175 prevalence**: `r Q_n` / `r Q_N`
  (`r sprintf("%.1f%%", Q_prev_pct)`).
- **Simulated-trial prevalence** ($n = `r sim_n_sample`$): mean
  `r sprintf("%.1f", mean(truth_by_sim$sizeH_true))` (SD
  `r sprintf("%.2f", sd(truth_by_sim$sizeH_true))`,
  range `r min(truth_by_sim$sizeH_true)`–`r max(truth_by_sim$sizeH_true)`,
  proportion `r sprintf("%.1f%%", 100 * mean(truth_by_sim$propH_true))`).

Under H1 the bar chart should concentrate on definitions close to this
anchor.  Under H0 the true subgroup does not exist, so any finds
reflect false discovery across the candidate space.
:::

The distribution of the identified subgroup $\hat H$ under H1
(calibrated HTE) and H0 (null) is rendered by
`forestsearch::plot_sg_distribution()`.  Three vignette-level settings
control how the charts behave:

1. **Top-K cap** (`sgdist_top_k` = `r sgdist_top_k`): at most that
   many bars are shown; additional passing labels pool into a single
   "Other (*k* labels)" row (or are suppressed -- see
   `sgdist_show_other_alt` / `sgdist_show_other_null`).  Without
   this, a diffuse H0 with one find per distinct label (the textbook
   false-positive signature) renders as overlapping, unreadable
   rows.
2. **Auto-halving** (`max_halvings = 2L`): the minimum-percentage
   thresholds (`sgdist_min_pct_alt` = `r sgdist_min_pct_alt`%,
   `sgdist_min_pct_null` = `r sgdist_min_pct_null`%) are halved up
   to twice if no label clears the initial value, so the chart does
   not render empty.
3. **Adaptive figure height** via `sgdist_fig_height()`, driven by
   `attr(p, "n_bars")`, keeps the vertical resolution consistent
   across scenarios regardless of how many distinct labels pass the
   filter.

```{r oc-sgdist-build}
#| code-fold: true
#| code-summary: "Build subgroup-distribution plots (click to expand)"

plot_sgdist_alt <- plot_sg_distribution(
  results              = results_alt[results_alt$analysis == "FS", ],
  any_col              = "any.H",
  label_col            = "sg.def",
  min_pct              = sgdist_min_pct_alt,
  top_k                = sgdist_top_k,
  show_other           = sgdist_show_other_alt,
  max_halvings         = 2L,
  bar_label_inside     = TRUE,
  placeholder_on_empty = TRUE,
  title                = "HTE: Subgroups Identified (H-hat)",
  wrap_width           = 30
)

plot_sgdist_null <- plot_sg_distribution(
  results              = results_null[results_null$analysis == "FS", ],
  any_col              = "any.H",
  label_col            = "sg.def",
  min_pct              = sgdist_min_pct_null,
  top_k                = sgdist_top_k,
  show_other           = sgdist_show_other_null,
  max_halvings         = 2L,
  bar_label_inside     = TRUE,
  placeholder_on_empty = TRUE,
  title                = "Null: Subgroups Identified (H-hat)",
  wrap_width           = 30
)

h_alt  <- sgdist_fig_height(attr(plot_sgdist_alt,  "n_bars"))
h_null <- sgdist_fig_height(attr(plot_sgdist_null, "n_bars"))
```

```{r oc-alt-sgdist, fig.width=10, fig.height=h_alt}
#| code-fold: true

print(plot_sgdist_alt)
```

```{r oc-null-sgdist, fig.width=10, fig.height=h_null}
#| code-fold: true

print(plot_sgdist_null)
```

Under `sg_focus = "maxSG"` and H1, the chart should concentrate on
definitions close to the true $Q$ anchor (age $> 34$ and preanti
$\le$ 745 on the discretised side of the DGM).  Under H0 the find
rate is low and any bars that clear the (possibly auto-halved)
threshold are scattered across the candidate space — the expected
false-discovery signature.  When the H0 chart is dominated by a large
"Other (*k* labels)" bar — i.e., many distinct labels each passing
the threshold at count 1 or 2 — that is itself the diagnostic: FS is
firing on unrelated noise with no preferred label.

### Distribution of Effect Estimates

The MRCT analogue is `forestsearch::SGplot_estimates()`, which
displays violins for four populations (ITT, non-AP training, AP
testing, AP-subgroup) defined by the train/test split used in the
MRCT consistency framework.  The ACTG175 binary-benefit study is a
single-population design (no train/test split), so we construct a
three-panel binary analogue using the columns `hr.itt` (ITT OR),
`hr.H.hat` (identified subgroup $\hat H$), and `hr.Hc.hat`
(identified complement $\hat H^c$) that `run_simulation_analysis()`
already populates.  All three are on the switched-treatment scale
and inverted to the original benefit scale for display.  On the
benefit scale, OR $> 1$ indicates the ZDV+ddI combination is superior
to ddI monotherapy.  As before, both H1 and H0 scenarios are shown.

```{r oc-alt-ordist, fig.width=9, fig.height=4.5}
#| code-fold: true
#| code-summary: "H1 OR-distribution violin (click to expand)"

or_Q_true <- 1 / dgm_calibrated$hazard_ratios$harm_subgroup

plot_effect_distribution(
  results           = results_alt,
  analysis_method   = "FS",
  effect_measure    = "OR",
  subgroup_notation = sg_notation,
  reference_effect  = or_Q_true,
  ylim              = ordist_ylim,
  title             = "ACTG175 Binary (HTE): OR Estimates Across Simulations",
  subtitle          = sprintf(
    "n = %d, %d sims | dotted green = true OR(Q) on benefit scale = %.2f",
    sim_n_sample, nsims_alt, or_Q_true)
)
```

```{r oc-null-ordist, fig.width=9, fig.height=4.5}
#| code-fold: true
#| code-summary: "H0 OR-distribution violin (click to expand)"

# Under H0 the DGM has no benefit subgroup -- no true-effect reference.
plot_effect_distribution(
  results           = results_null,
  analysis_method   = "FS",
  effect_measure    = "OR",
  subgroup_notation = sg_notation,
  ylim              = ordist_ylim,
  title             = "ACTG175 Binary (Null): OR Estimates Across Simulations",
  subtitle          = sprintf(
    "n = %d, %d sims | no true harm/benefit subgroup; dashed red = OR = 1",
    sim_n_sample, nsims_null)
)
```

Under H1 (top panel), the left violin (ITT) is centred near the DGM's
implied ITT benefit; the middle violin ($\widehat{\mathrm{OR}}(\hat
H)$) should concentrate near the dotted green reference line at the
true $\mathrm{OR}(Q)$; the right violin ($\widehat{\mathrm{OR}}(\hat
H^c)$) shows the residual effect in the complement and should sit
near $\mathrm{OR} = 1$ (i.e., no benefit in $\hat H^c$ once $\hat H$
is removed).  Under H0 (bottom panel), all three violins should
centre near $\mathrm{OR} = 1$; any concentration away from 1 in the
subgroup panels, conditional on a subgroup being found, reflects the
selection-driven bias that the bootstrap bias-correction in
`forestsearch()` is designed to correct.

```{r oc-alt-tables}
build_estimation_table(
  results         = results_alt,
  dgm             = dgm_calibrated,
  analysis_method = "FS",
  subtitle = sprintf(
    "ACTG175 Binary DGM (raw cont.) | n = %d, %d sims",
    sim_n_sample, nsims_alt
  ),
  font_size = 16,
  subgroup_notation = sg_notation
)

if ("GRF" %in% unique(results_alt$analysis)) {
  build_estimation_table(
    results         = results_alt,
    dgm             = dgm_calibrated,
    analysis_method = "GRF",
    subtitle = sprintf(
      "ACTG175 Binary DGM, GRF raw cont. (dmin=%.2f) | n = %d, %d sims",
      grf_dmin, sim_n_sample, nsims_alt
    ),
    font_size = 16,
    subgroup_notation = sg_notation
  )
}

timings$report_alt <- proc.time()[3] - t0
```

### Interpretation

```{r interpret-alt, results="asis"}
interpret_estimation_table(
  results_alt, dgm_calibrated,
  analysis_method = "FS",
  scenario = "alt",
  subgroup_notation = sg_notation
)
```


## Operating Characteristics Under H0

```{r oc-null}
format_oc_results(
  results_null,
  metrics  = c("detection", "classification"),
  digits   = 2,
  title    = "ACTG175 Binary: Type I Error (Null, Raw Continuous)",
  subtitle = sprintf(
    "n = %d, %d sims, OR(ITT) = %.2f (benefit scale)",
    sim_n_sample, nsims_null,
    1 / dgm_null$hazard_ratios$overall
  ),
  subgroup_notation = sg_notation
)
```

```{r oc-null-tables}
build_estimation_table(
  results         = results_null,
  dgm             = dgm_null,
  analysis_method = "FS",
  subtitle = sprintf(
    "ACTG175 Binary DGM (raw cont.) | n = %d, %d sims",
    sim_n_sample, nsims_null
  ),
  font_size = 16,
  subgroup_notation = sg_notation
)

if ("GRF" %in% unique(results_null$analysis)) {
  build_estimation_table(
    results         = results_null,
    dgm             = dgm_null,
    analysis_method = "GRF",
    subtitle = sprintf(
      "ACTG175 Binary DGM, GRF raw cont. (dmin=%.2f) | n = %d, %d sims",
      grf_dmin, sim_n_sample, nsims_null
    ),
    font_size = 16,
    subgroup_notation = sg_notation
  )
}
```




::: {.callout-note}
Under H0, $Q = \emptyset$.  Sen is undefined (shown as "—")
and PPV $\equiv 0$ (León et al. 2024, Section 3).
The detection rate (Found) is the false positive rate.
:::


## Classification Table

```{r classification}
build_classification_table(
  scenario_results = list(
    null = list(
      results  = results_null,
      label    = "ACTG-bin-raw",
      n_sample = sim_n_sample,
      dgm      = dgm_null,
      hypothesis = "null"
    ),
    alt = list(
      results  = results_alt,
      label    = "ACTG-bin-raw",
      n_sample = sim_n_sample,
      dgm      = dgm_calibrated,
      hypothesis = "alt"
    )
  ),
  analyses = sort(unique(c(
    unique(results_null$analysis),
    unique(results_alt$analysis)
  ))),
  digits    = 2,
  font_size = 16,
  title     = "ACTG175 Binary (Raw Continuous): Subgroup Identification and Classification",
  n_sims    = nsims_alt,
  subgroup_notation = sg_notation
)
```


# Timing Summary {#sec-timing}

```{r timing}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Step = c(names(timings)),
  Seconds = round(unlist(timings), 1),
  Minutes = round(unlist(timings) / 60, 2)
)

timing_df |>
  gt::gt() |>
  gt::tab_header(title = "Computation Timing") |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(rows = Step == "total")
  )
```

### Projection for Production Runs

```{r projection}
#| code-fold: true
#| code-summary: "Projection arithmetic (click to expand)"
#| results: asis

# Per-replicate cost (simulation loop only)
time_per_alt  <- timings$sims_alt / nsims_alt
time_per_null <- timings$sims_null / nsims_null

# prod_alt and prod_null defined in setup chunk
proj_alt  <- time_per_alt * prod_alt
proj_null <- time_per_null * prod_null
proj_other <- timings$total - timings$sims_alt - timings$sims_null
proj_total <- proj_alt + proj_null + proj_other

proj_df <- data.frame(
  Metric = c(
    sprintf("Current run (%d alt + %d null on %d workers)",
            nsims_alt, nsims_null, n_workers),
    "Per-replicate, H1",
    "Per-replicate, H0",
    sprintf("Projected H1 (%d sims)", prod_alt),
    sprintf("Projected H0 (%d sims)", prod_null),
    "Projected other (calibration, setup, reporting)",
    "Projected total"),
  Seconds = c(NA_real_,
              time_per_alt, time_per_null,
              proj_alt, proj_null, proj_other, proj_total),
  Minutes = c(NA_real_,
              time_per_alt / 60, time_per_null / 60,
              proj_alt / 60, proj_null / 60, proj_other / 60,
              proj_total / 60),
  Hours   = c(NA_real_, NA_real_, NA_real_, NA_real_, NA_real_,
              NA_real_, proj_total / 3600),
  stringsAsFactors = FALSE)

cat(gt::as_raw_html(
  gt::gt(proj_df) |>
    gt::tab_header(title = sprintf(
      "Projected runtime for production simulation (%d H1 + %d H0)",
      prod_alt, prod_null)) |>
    gt::fmt_number(columns = c("Seconds", "Minutes", "Hours"),
                   decimals = 1) |>
    gt::sub_missing(missing_text = "\u2014") |>
    gt::cols_label(Metric  = gt::md("**Metric**"),
                   Seconds = gt::md("**Seconds**"),
                   Minutes = gt::md("**Minutes**"),
                   Hours   = gt::md("**Hours**")) |>
    gt::tab_options(table.font.size = 11)
), "\n\n")
```


# Discussion {#sec-discussion}

## Key Differences from the Survival Simulation

This document mirrors the structure of `actg175_simulations_benefit.qmd`
(survival endpoint) with the following differences:

1. **No censoring**: the binary outcome $I(\text{cd420} > \text{cd40})$
   is fully observed, eliminating censoring model complexity.

2. **Package GLM simulation suite**: uses `generate_glm_dgm()`,
   `calibrate_glm_interaction()`, and `simulate_from_glm_dgm()`
   instead of `generate_aft_dgm_flex()`, manual grid search, and
   `simulate_from_dgm()`.  `run_simulation_analysis()` dispatches
   automatically based on the `"glm_dgm"` class.

3. **OR replaces HR**: thresholds and reporting use odds ratios on the
   switched scale.  The `subgroup_notation = "benefit"` inversion
   applies identically.

4. **No CDE/AHR estimands**: these are survival-specific.  The binary
   DGM stores OR-scale effects in the `hazard_ratios` list for
   reporting compatibility.

5. **GRF standalone included**: the standalone GRF analysis
   (`grf.subg.harm.glm`) uses `causal_forest()` for candidate
   factor screening and policy trees for subgroup identification.

## Key Difference from the Prior Discretised Setup

The prior version of this document discretised every continuous
covariate into a binary factor (using the same cuts that define
the true subgroup $Q$) before the analysis step.  That setup
implicitly gives the analysis methods knowledge of the true
cutpoints.  **This document feeds raw continuous covariates
(`age`, `preanti`, `wtkg`, `karnof`, `cd40`, `cd80`) to the
analysis methods**, so the search methods must discover cuts
themselves.  The DGM construction is unchanged: the true $Q$ is
still $\{\text{age} > 34\} \cap \{\text{preanti} \leq 744.5\}$
and the calibration targets the same OR(Q).

**Expected consequences** relative to the discretised run:

- Lower sensitivity and PPV for both FS and GRF, because
  discovered cuts will generally differ from the true cuts.
- Similar or slightly higher FPR for GRF, because the policy-tree
  search space is larger over continuous inputs.
- Detection power depends on how closely the median / GRF-derived
  cuts align with the true cuts.

## Limitations

- The binary DGM is based on a logistic regression fitted to
  the observed ACTG175 data.  The super-population preserves
  the empirical covariate distribution but assumes the logistic
  model is correctly specified.

- `k_inter` is a direct additive shift on the log-odds scale
  for Q members under treatment.  Positive values increase the
  outcome for Q; negative values decrease it.

- `dmin.grf` was calibrated on the discretised setup; a fresh
  calibration run with raw continuous inputs is recommended
  before production scaling.

## Reloading Saved Results

When `save_results = TRUE`, the simulation bundle is saved as an
`.rds` file in the `_data/` directory.  The bundle carries every
scalar, list, and object the vignette reads downstream of the
simulation run, which makes it straightforward to (a) reload for
interactive analysis or (b) split this document into a separate
"run" script and a "report" document in the future.

```r
bundle <- readRDS(
  "../_data/actg175_binary_benefit_simulations_rawcont_100alts_100nulls.rds"
)

# -- Core results (data.tables) --------------------------------------------
results_alt  <- bundle$results_alt
results_null <- bundle$results_null

# -- DGM objects (class "glm_dgm") -----------------------------------------
dgm_calibrated <- bundle$dgm_calibrated
dgm_null       <- bundle$dgm_null

# -- Restore ALL vignette-scope scalars consumed downstream of the run -----
# (Assigning locally lets Section 8 chunks run unchanged against the
#  reloaded bundle.)
with(bundle$config, {
  sim_n_sample        <<- n_sample
  nsims_alt           <<- nsims_alt
  nsims_null          <<- nsims_null
  n_workers           <<- n_workers
  confounders_analysis<<- confounders_analysis
  sg_notation         <<- sg_notation
  sgdist_min_pct_alt  <<- sgdist_min_pct_alt
  sgdist_min_pct_null <<- sgdist_min_pct_null
  sgdist_top_k        <<- sgdist_top_k
  sgdist_show_other_alt  <<- sgdist_show_other_alt
  sgdist_show_other_null <<- sgdist_show_other_null
  grf_dmin            <<- grf_dmin
  prod_alt            <<- prod_alt
  prod_null           <<- prod_null
})

timings      <- bundle$timings
n_failed_alt <- bundle$n_failed_alt
n_failed_null<- bundle$n_failed_null

# -- True effects (convenience) --------------------------------------------
bundle$true_effects$or_Q     # OR(Q) on switched scale
bundle$true_effects$or_Qc    # OR(Qc)
bundle$true_effects$or_ITT   # OR(ITT)
bundle$true_effects$k_inter  # calibrated interaction shift

# -- FS / GRF parameter lists (passed to run_simulation_analysis) ----------
bundle$fs_params
bundle$grf_params

# -- Provenance ------------------------------------------------------------
bundle$timestamp         # when the run completed
bundle$R_version         # R version used
bundle$fs_version        # forestsearch version used

# -- Reproduce Section 8 tables with the reloaded objects ------------------
format_oc_results(results_alt,    subgroup_notation = sg_notation, ...)
build_estimation_table(results_alt, dgm = dgm_calibrated,
                        subgroup_notation = sg_notation, ...)
```


# Summary {#sec-summary}

```{r summary-compute}
#| code-fold: true
#| code-summary: "Compute summary metrics"

mc_se <- function(p, n) sqrt(p * (1 - p) / n)
methods <- sort(unique(results_alt$analysis))

summ <- data.frame(Method = methods, stringsAsFactors = FALSE)
for (m in methods) {
  ra <- results_alt[results_alt$analysis == m, ]
  rn <- results_null[results_null$analysis == m, ]

  summ$alt_det[summ$Method == m]   <- mean(ra$any.H, na.rm = TRUE)
  summ$null_fpr[summ$Method == m]  <- mean(rn$any.H, na.rm = TRUE)
  summ$alt_sens[summ$Method == m]  <- mean(ra$sens, na.rm = TRUE)
  summ$alt_ppv[summ$Method == m]   <- mean(ra$ppv, na.rm = TRUE)
  summ$se_det[summ$Method == m]    <- mc_se(
    summ$alt_det[summ$Method == m], nrow(ra))
  summ$se_fpr[summ$Method == m]    <- mc_se(
    summ$null_fpr[summ$Method == m], nrow(rn))
}
```

## FS vs GRF Classification Concordance Across Simulations

Per-subject classification concordance between ForestSearch and GRF
is computed within each simulation replicate by
`run_simulation_analysis()` and stored as `agree.FS.GRF` and
`kappa.FS.GRF` columns.  When a method does not detect a subgroup,
it classifies all subjects to $Q^c$.

```{r sim-concordance}
#| code-fold: true
#| code-summary: "Classification concordance (click to expand)"
#| results: asis

conc_col <- "agree.FS.GRF"
kappa_col <- "kappa.FS.GRF"

if (all(c(conc_col, kappa_col) %in% names(results_alt)) &&
    all(c("FS", "GRF") %in% methods)) {

  # One row per sim (take FS rows — concordance is per-sim, same on both)
  fs_alt  <- results_alt[results_alt$analysis == "FS", ]
  grf_alt <- results_alt[results_alt$analysis == "GRF", ]
  n       <- nrow(fs_alt)

  both     <- sum(fs_alt$any.H == 1 & grf_alt$any.H == 1)
  fs_only  <- sum(fs_alt$any.H == 1 & grf_alt$any.H == 0)
  grf_only <- sum(fs_alt$any.H == 0 & grf_alt$any.H == 1)
  neither  <- sum(fs_alt$any.H == 0 & grf_alt$any.H == 0)
  both_idx <- fs_alt$any.H == 1 & grf_alt$any.H == 1

  conc_df <- data.frame(
    Metric = c(
      "Mean agreement (%)",
      "SD agreement (%)",
      "Mean kappa",
      "SD kappa",
      "Median kappa",
      "Both detect",
      "FS only",
      "GRF only",
      "Neither",
      sprintf("Mean agreement | both detect (n = %d)", both),
      sprintf("Mean kappa | both detect (n = %d)",     both)
    ),
    Value = c(
      sprintf("%.1f", 100 * mean(fs_alt[[conc_col]],  na.rm = TRUE)),
      sprintf("%.1f", 100 * sd(fs_alt[[conc_col]],    na.rm = TRUE)),
      sprintf("%.3f",       mean(fs_alt[[kappa_col]], na.rm = TRUE)),
      sprintf("%.3f",       sd(fs_alt[[kappa_col]],   na.rm = TRUE)),
      sprintf("%.3f",       median(fs_alt[[kappa_col]], na.rm = TRUE)),
      sprintf("%d (%.0f%%)", both,     100 * both     / n),
      sprintf("%d (%.0f%%)", fs_only,  100 * fs_only  / n),
      sprintf("%d (%.0f%%)", grf_only, 100 * grf_only / n),
      sprintf("%d (%.0f%%)", neither,  100 * neither  / n),
      if (both > 0)
        sprintf("%.1f%%", 100 * mean(fs_alt[[conc_col]][both_idx],  na.rm = TRUE))
      else "\u2014",
      if (both > 0)
        sprintf("%.3f",         mean(fs_alt[[kappa_col]][both_idx], na.rm = TRUE))
      else "\u2014"
    ),
    stringsAsFactors = FALSE)

  cat(gt::as_raw_html(
    gt::gt(conc_df,
           groupname_col = NULL) |>
      gt::tab_header(title = sprintf(
        "FS vs GRF classification concordance (%d sims)", n)) |>
      gt::tab_row_group(label = "Overall agreement",
                        rows = 1:5) |>
      gt::tab_row_group(label = "Detection patterns",
                        rows = 6:9) |>
      gt::tab_row_group(label = "Conditional on both detecting",
                        rows = 10:11) |>
      gt::row_group_order(groups = c(
        "Overall agreement",
        "Detection patterns",
        "Conditional on both detecting")) |>
      gt::cols_label(Metric = gt::md("**Metric**"),
                     Value  = gt::md("**Value**")) |>
      gt::tab_options(table.font.size = 11)
  ), "\n\n")

} else {
  cat("*Concordance columns not found in results.  Ensure",
      "`run_simulation_analysis()` was called with",
      "`run_fs = TRUE` and `run_grf = TRUE`.*\n\n")
}
```

## Key Findings

```{r summary-findings}
#| code-fold: true
#| code-summary: "Summary computation (click to expand)"
#| results: asis

# Top-line narrative (kept as markdown, not a table)
cat(sprintf(
  "**Simulation budget**: %d H1 + %d H0 sims.\n\n",
  nsims_alt, nsims_null))
cat("**Analysis setting**: raw continuous covariates;",
    "cuts discovered by the methods.\n\n")

# Per-method operating-characteristic summary as a gt table
truth_label <- ifelse(sg_notation == "benefit", "Q", "H")

kf_df <- data.frame(
  Method         = summ$Method,
  `Detection (%)` = sprintf("%.1f", 100 * summ$alt_det),
  `FPR (%)`       = sprintf("%.1f", 100 * summ$null_fpr),
  `Sens`          = sprintf("%.2f", summ$alt_sens),
  `PPV`           = sprintf("%.2f", summ$alt_ppv),
  check.names = FALSE,
  stringsAsFactors = FALSE)

cat(gt::as_raw_html(
  gt::gt(kf_df) |>
    gt::tab_header(title = "Operating characteristics by method") |>
    gt::tab_spanner(label = sprintf("Classification vs %s", truth_label),
                    columns = c("Sens", "PPV")) |>
    gt::cols_label(Method = gt::md("**Method**")) |>
    gt::tab_options(table.font.size = 11)
), "\n\n")

# Timing summary as a second compact gt table
time_per_alt  <- timings$sims_alt  / nsims_alt
time_per_null <- timings$sims_null / nsims_null
time_df <- data.frame(
  Metric = c(sprintf("Per-replicate (H1), %d workers",  n_workers),
             sprintf("Per-replicate (H0), %d workers",  n_workers),
             "This run (total)",
             sprintf("Projected production (%d H1 + %d H0)",
                     prod_alt, prod_null)),
  Value  = c(sprintf("%.1f sec",  time_per_alt),
             sprintf("%.1f sec",  time_per_null),
             sprintf("%.1f min",  as.numeric(timings$total) / 60),
             sprintf("%.1f hours",
                     (prod_alt  * time_per_alt +
                      prod_null * time_per_null) / 3600)),
  stringsAsFactors = FALSE)

cat(gt::as_raw_html(
  gt::gt(time_df) |>
    gt::tab_header(title = "Timing") |>
    gt::cols_label(Metric = gt::md("**Metric**"),
                   Value  = gt::md("**Value**")) |>
    gt::tab_options(table.font.size = 11)
), "\n\n")
```


# Appendix: Effect Measures and Outcome Direction Conventions {#sec-appendix-conventions}

*This appendix is a quick-lookup reference for this vignette's
directional configuration.  For comprehensive coverage across all
endpoint types -- including the eight-scenario decision table,
worked code examples for binary, continuous, count, and survival
outcomes, common pitfalls, and verification tests -- see the
standalone* [*ForestSearch Outcome Orientation Reference*](https://larry-leon.github.io/forestsearch/articles/forestsearch_outcome_orientation_guide.html) *vignette.*

## A.1 The unified harm-search convention

Every subgroup search in `forestsearch` uses a **single comparison
direction** throughout the pipeline.  In each candidate subgroup
$S$, the package checks whether the estimated treatment effect
satisfies

$$
\hat{\theta}(S) \;\geq\; c_1
$$

(`effect.threshold`), with the same direction reused for the
bootstrap consistency check, $\hat{\theta}(S_b) \geq c_2$.  This
comparison never flips.  The entire question of how to set up a
search reduces to: **how do I orient my data so the subgroup I
want to find produces $\hat{\theta}(S) > 0$ on the chosen effect
scale?**

Two transformations handle the orientations that don't satisfy
this naturally.  First, **benefit search** -- finding a subgroup
in which the original experimental treatment is *beneficial* --
is implemented by switching the treatment labels so the worse
arm becomes `treat = 1`.  A benefit subgroup of the original
experimental then appears as a harm subgroup of the
switched-treatment; magnitudes are inverted at reporting time
via `subgroup_notation = "benefit"`.  Second, when the user's
binary outcome is encoded on the positive scale (higher $Y$ =
better), the binary effect estimator can substitute $1 - Y$
before computing the OR -- this is what the `adverse_outcome`
flag controls in the FS estimator path.

The flag's *name* suggests an interpretive label ("the outcome
is adverse"), but mechanically what it does in the binary FS
estimator is **flip $Y$ to $1 - Y$ when `adverse_outcome = FALSE`,
and leave $Y$ as-is when `TRUE`** (the default for binary).  The
two framings -- pre-flip $Y$ at data prep and pass `TRUE`, or
keep $Y$ positive and pass `FALSE` -- are mechanically
equivalent for binary outcomes.  For **continuous** and **count**
outcomes the FS effect estimator does **not** auto-flip on the
flag (only the GRF candidate-cut step does); for those outcome
types you must encode $Y$ on the harm direction at data prep
and pass `adverse_outcome = TRUE`.  The simplest safe rule
across all GLM outcome types is therefore: **encode $Y$ on the
adverse scale at data prep and pass `adverse_outcome = TRUE`
everywhere**.  This is the convention used by the present
vignette; see @sec-stat-goals for the specific row that applies
and Section 1.1 for the data-prep step.

::: {.callout-tip title="DGM golden rule"}
Once the data is correctly oriented, build the DGM with **positive
`k_inter`**, a **positive calibration target**, and a **positive
`k_inter_range` grid**.  The DGM operates on the final, oriented
data; outcome and treatment manipulations (negation, switching)
happen at the data-prep step, *before* the DGM is constructed.
:::

## A.2 Effect scales and threshold conventions

`forestsearch` works on the **log scale** internally for ratio
measures and on the **identity scale** for additive measures.
This affects how `effect.threshold` is interpreted:

| Measure type | Examples | What `effect.threshold = c₁` becomes |
|--------------|----------|---------------------------------------|
| Ratio | HR, OR, IRR, RR | $\log(c_1)$ -- e.g., 1.25 → 0.223 |
| Additive (bounded) | RD, IRD | $c_1$ directly; bounded in $[-1, 1]$ |
| Additive (unbounded) | MD | $c_1$ directly; unbounded |

For RD and IRD, the package warns and remaps if the threshold
is set outside the bounded range.  For MD (mean difference) the
threshold is genuinely unbounded -- values like
30 cells/$\mu$L are legitimate, and the package no longer
remaps them as it once did when MD was incorrectly treated as a
bounded measure.  In all cases, the comparison the package
performs internally is $\hat{\theta} \geq$ (transformed
threshold), with $\hat{\theta}$ on the same scale (log for
ratios, identity for additives).


## A.3 Reference table

```{r appendix-conventions-table}
#| code-fold: true
#| code-summary: "Build the master conventions table (click to expand)"

conv <- data.frame(
  Outcome   = c(rep("Binary",     4),
                rep("Continuous", 4),
                rep("Count",      4),
                rep("Survival",   2)),
  Measure   = c(rep("OR",  4),
                rep("MD",  4),
                rep("IRR", 4),
                rep("HR",  2)),
  Direction = c(
    "Positive (e.g., improvement = 1)",
    "Positive (e.g., improvement = 1)",
    "Adverse (e.g., death = 1)",
    "Adverse (e.g., death = 1)",
    "Positive (e.g., QoL score, higher = better)",
    "Positive (e.g., QoL score, higher = better)",
    "Adverse (e.g., pain score, higher = worse)",
    "Adverse (e.g., pain score, higher = worse)",
    "Positive (rare: e.g., recovery events)\u00b9",
    "Positive (rare: e.g., recovery events)\u00b9",
    "Adverse (e.g., adverse-event count)",
    "Adverse (e.g., adverse-event count)",
    "Event = adverse (fixed by causal_survival_forest)",
    "Event = adverse (fixed)"
  ),
  Goal      = rep(c("Harm", "Benefit"), 7),
  EncodeY   = c(
    "Y as-is", "Y as-is",         # binary positive
    "Y as-is", "Y as-is",         # binary adverse
    "Pre-flip: Y_use = -Y\u00b2",  # cont positive
    "Pre-flip: Y_use = -Y\u00b2",
    "Y as-is", "Y as-is",         # cont adverse
    "(see footnote 1)",            # count positive
    "(see footnote 1)",
    "Y as-is", "Y as-is",         # count adverse
    "Y as-is", "Y as-is"          # survival
  ),
  Treat     = rep(c("Original", "Switched"), 7),
  AdvFlag   = c(
    "FALSE", "FALSE",   # binary positive
    "TRUE",  "TRUE",    # binary adverse
    "TRUE",  "TRUE",    # cont positive (after pre-flip)
    "TRUE",  "TRUE",    # cont adverse
    "TRUE",  "TRUE",    # count positive
    "TRUE",  "TRUE",    # count adverse
    "(N/A)", "(N/A)"     # survival
  ),
  Notation  = rep(c("\"harm\"", "\"benefit\""), 7),
  Meaning   = c(
    # Binary
    "Trt reduces P(positive Y) in subgroup",
    "Orig-experimental increases P(positive Y) in subgroup",
    "Trt increases P(adverse Y) in subgroup",
    "Orig-experimental reduces P(adverse Y) in subgroup",
    # Continuous
    "Trt reduces Y (positive scale) in subgroup",
    "Orig-experimental increases Y (positive scale) in subgroup",
    "Trt increases Y (adverse scale) in subgroup",
    "Orig-experimental reduces Y (adverse scale) in subgroup",
    # Count
    "Trt reduces rate of (positive) events in subgroup",
    "Orig-experimental increases rate of (positive) events",
    "Trt increases rate of adverse events in subgroup",
    "Orig-experimental reduces rate of adverse events",
    # Survival
    "Trt increases hazard of event in subgroup",
    "Orig-experimental reduces hazard of event in subgroup"
  ),
  stringsAsFactors = FALSE
)

# Build gt
gt_conv <- gt::gt(conv, groupname_col = "Outcome", rowname_col = NULL) |>
  gt::cols_label(
    Measure   = gt::md("**Measure**"),
    Direction = gt::md("**Native Y direction**"),
    Goal      = gt::md("**Search goal**"),
    EncodeY   = gt::md("**Encode Y at data prep**"),
    Treat     = gt::md("**Treatment coding**"),
    AdvFlag   = gt::md("**`adverse_outcome`**"),
    Notation  = gt::md("**`subgroup_notation`**"),
    Meaning   = gt::md("**\"Effect > c\u2081\" means in subgroup**")
  ) |>
  gt::tab_header(
    title    = "Effect measure / outcome-direction / search-goal combinations",
    subtitle = "User-facing settings for forestsearch() and grf.subg.harm.glm()"
  ) |>
  gt::tab_style(
    style    = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |>
  gt::tab_options(
    table.font.size = 10,
    row_group.padding = gt::px(4),
    data_row.padding  = gt::px(3)
  ) |>
  gt::tab_footnote(
    footnote = paste0(
      "Positive count outcomes are uncommon (count outcomes typically ",
      "model adverse events or rates).  If you have one, model the ",
      "complement (e.g., max-rate minus observed) on the adverse scale, ",
      "or consult the package authors."
    ),
    locations = gt::cells_body(
      columns = "Direction",
      rows    = which(conv$Outcome == "Count" & grepl("rare", conv$Direction))
    )
  ) |>
  gt::tab_footnote(
    footnote = paste0(
      "For continuous outcomes the FS effect estimator (`make_lm_estimator`) ",
      "does not respect `adverse_outcome`.  Pre-flip Y to the adverse direction ",
      "at data prep so OR / MD > threshold consistently indicates harm.  ",
      "GRF (`grf.subg.harm.glm`) does respect the flag for continuous; the ",
      "pre-flip + `adverse_outcome = TRUE` rule keeps both pipelines aligned."
    ),
    locations = gt::cells_body(
      columns = "EncodeY",
      rows    = which(conv$Outcome == "Continuous" &
                      grepl("Pre-flip", conv$EncodeY))
    )
  )

gt_conv
```

## A.4 Worked example: this vignette's configuration

The present vignette uses the row **Binary | OR | Adverse | Benefit**.
Concretely:

- The clinical concept is improvement (CD4 response by 20 weeks),
  but `y_binary` is encoded at data prep on the **adverse** scale
  as $I(\text{cd420} \leq \text{cd40})$ -- a 1 indicates failure
  to improve.
- Treatment is **switched** so ddI $\to$ `treat = 1` and
  ZDV+ddI $\to$ `treat = 0`.
- `adverse_outcome` defaults to `TRUE` for binary inside
  `forestsearch()` and is consistent with the adverse encoding
  of `y_binary`.
- The package's harm-search machinery flags subgroups where the
  switched treatment (ddI) increases the odds of non-improvement
  -- equivalently, where the original experimental (ZDV+ddI) is
  beneficial in giving improvement.
- Calibration target $\text{OR}(Q) = `r cal_target_or`$ corresponds
  to a $`r cal_target_or`\times$ odds-ratio of benefit for
  ZDV+ddI in $Q$.  This is the same magnitude as OR computed on
  the natural (original treatment, improvement-Y) scale, by the
  $\text{OR}(\text{switched, adverse}) = \text{OR}(\text{original, positive})$
  symmetry.
- Reporting uses `subgroup_notation = "benefit"` so summary tables
  display effect estimates on the original-experimental side.


## A.5 Common pitfalls

The standalone [*ForestSearch Outcome Orientation Reference*](https://larry-leon.github.io/forestsearch/articles/forestsearch_outcome_orientation_guide.html)
is the authoritative source for setup mechanics across all endpoint
types.  Two pitfalls show up often enough in practice to flag
here:

::: {.callout-warning title="Pitfall 1: calibration target on the wrong scale"}
Setting (e.g.) `cal_target_or = 0.5` together with a negative
`k_inter_range` produces a $Q$ that is only marginally more
extreme than the rest of the sample in the same direction --
no real contrast.  *Symptom:* near-100% detection paired with
very high false-positive rate (~97%).  *Fix:* use a target
above 1 with a positive `k_inter` grid for ratio measures, or
a positive target with a positive grid for MD.
:::

::: {.callout-warning title="Pitfall 2: `adverse_outcome` confused with outcome negation"}
Setting `adverse_outcome = FALSE` for a benefit search with
switched treatment and a *positive* binary outcome causes the
binary FS estimator to flip $Y \to 1 - Y$, reversing the OR
direction and breaking the $\hat{\theta} \geq c_1$ comparison.
*Symptom:* zero detection.  *Fix:* leave `adverse_outcome` at
its default `TRUE` for binary, *or* pre-flip $Y$ at data prep
(the convention used by this vignette -- see @sec-stat-goals).
:::

For the full list of pitfalls -- including MD-threshold remapping
behaviour, survival with positive events, and DGM construction
gotchas -- the eight-scenario decision table, the four
verification tests, and worked code examples for binary,
continuous, count, and survival outcomes, see the standalone
[*ForestSearch Outcome Orientation Reference*](https://larry-leon.github.io/forestsearch/articles/forestsearch_outcome_orientation_guide.html)
vignette.


# Session Info {#sec-session}

```{r session-info}
sessionInfo()
```


# References {#sec-references}

- Hammer, S.M., et al. (1996). *NEJM*, 335, 1081--1090.
- Leon, L.F., et al. (2024). *Statistics in Medicine*, 43,
  3921--3942.
- Tsiatis, A.A., et al. (2008). *Statistics in Medicine*, 27,
  4658--4677.
