---
title: "Simulation Studies Based on ACTG175: Binary Endpoint (GLM)"
subtitle: "Benefit Subgroup Identification with CD4 Response at 20 Weeks -- Raw Continuous Covariates"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# -- Simulation control -------------------------------------------------------
nsims_alt   <- 100L
nsims_null  <- 100L
sim_seed    <- 8316951L
n_workers   <- max(1L, floor(0.90 * parallel::detectCores(logical = FALSE)))

# -- Production projection targets --------------------------------------------
prod_alt               <- 5000L
prod_null              <- 25000L


# -- Dataset preparation: ACTG175 ---------------------------------------------
#
# ARMS (matching Leon et al. 2024, Section 4.2):
#   speff2trial::ACTG175 arm coding:
#     0 = zidovudine (ZDV) alone
#     1 = ZDV + didanosine (ZDV+ddI)   <- experimental
#     2 = ZDV + zalcitabine (ZDV+ddC)
#     3 = didanosine (ddI) alone        <- control
#
# The paper compared arm 1 (ZDV+ddI, combination) vs arm 3 (ddI,
# monotherapy) and searched for a BENEFIT subgroup Q by switching
# treatment roles.
#
# True subgroup anchor (used for DGM construction and ground-truth
# classification ONLY; NOT exposed to the analysis methods):
#   z1 (age):     age > 34        -- median split
#   z2 (preanti): preanti <= 745  -- q3 split
#
# Analysis candidate covariates are provided as described in the
# "Realistic analysis setting" block below.
actg_arms        <- c(1L, 3L)
actg_treat_arm   <- 1L
actg_age_cut     <- 34
actg_preanti_cut <- 744.5

# -- Benefit search via treatment switching ------------------------------------
# Under switched roles, the BENEFIT subgroup Q appears as a "harm"
# subgroup (ddI monotherapy is worse for Q).  Thresholds on OR scale:
#   c1 = 1/0.60 = 1.667  (screening: OR >= 1.667 under switched roles)
#   c2 = 1/0.80 = 1.25   (consistency)
# After identification, ORs are inverted to the original scale via
# subgroup_notation = "benefit".

# -- DGM construction ---------------------------------------------------------
# The DGM uses the discretised z1-z12 factor formulation because
# generate_glm_dgm() requires factor covariates.  The TRUE benefit
# subgroup is defined by {z1 = 1 AND z2 = 1}, i.e. age > 34 AND
# preanti <= 744.5.  This provides the ground truth for the
# sensitivity / PPV computations in run_simulation_analysis().
dgm_factor_vars   <- paste0("z", 1:12)
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)
dgm_n_super       <- 5000L

# -- Realistic analysis setting -----------------------------------------------
# The simulation mimics a real data analysis in which the analyst does
# NOT know which variables define Q nor the cuts that identify it.
# Continuous covariates are therefore supplied to forestsearch() and
# grf.subg.harm.glm() in their RAW form; the search methods discover
# cuts themselves (forestsearch: median / GRF-derived / forced cuts;
# grf.subg.harm.glm: policy-tree splits).
#
# NOTE: the true subgroup Q still involves age (continuous, cut at 34)
# and preanti (continuous, cut at 744.5).  Forestsearch will attempt
# cuts at medians and GRF-selected thresholds, which will generally
# DIFFER from the true cuts.  This is the realistic setting and is
# precisely what we want to evaluate: the method's ability to recover
# an effect subgroup when the cutpoints are unknown.
analysis_continuous_vars <- c(
  "age",      # continuous; TRUE cut (34) unknown to analyst
  "preanti",  # continuous; TRUE cut (744.5) unknown to analyst
  "wtkg",     # continuous
  "karnof",   # Karnofsky score (integer 70-100, treated as continuous)
  "cd40",     # baseline CD4 (continuous)
  "cd80"      # baseline CD8 (continuous)
)

# Truly binary indicators from ACTG175 (passed as factors to the
# analysis; no cut to discover).
analysis_binary_vars <- c(
  "hemo",     # hemophilia (0/1)
  "homo",     # homosexual activity (0/1)
  "drugs",    # IV drug use history (0/1)
  "race",     # race (0/1, white vs non-white as coded)
  "gender",   # gender (0/1)
  "symptom"   # symptomatic (0/1)
)

# -- Calibration ---------------------------------------------------------------
# Target OR(Q) = 2.0 under SWITCHED treatment (= 0.50 on original scale).
cal_target_or     <- 2.0
cal_k_grid_range  <- c(0, 3)
cal_grid_step     <- 0.05

# -- Simulation study ----------------------------------------------------------
sim_n_sample      <- 1000L
sim_n_noise       <- 0L

# -- Analysis flags ------------------------------------------------------------
run_fs            <- TRUE    # Run ForestSearch
run_fs_grf        <- FALSE   # Run FS with GRF pre-screening only
run_grf           <- TRUE    # Standalone GRF (grf.subg.harm.glm)

# -- ForestSearch parameters (benefit search thresholds) -----------------------
fs_or_threshold        <- round(1 / 0.60, 3)   # c1
fs_or_consistency      <- round(1 / 0.80, 3)   # c2
fs_pconsistency        <- 0.90
fs_splits              <- 400L
fs_n_min               <- 60L
fs_d0_min              <- 10L
fs_d1_min              <- 10L
fs_maxk                <- 2L
fs_vi_grf_min          <- -0.2
fs_seedit              <- 8316951L
fs_cont_cutoff         <- 4L   # is.continuous() threshold (# unique values)

# -- Continuous-variable cut strategy -----------------------------------------
# cut_type controls how candidate cuts are generated for each continuous
# covariate that forestsearch detects (i.e. any variable with more than
# fs_cont_cutoff unique values that is passed through confounders.name).
#
#   "default" : per continuous variable, get_conf_force() -> cut_var()
#               generates FOUR candidate cuts:
#                   var <= mean(var)
#                   var <= median(var)    (= Q2)
#                   var <= qlow(var)      (= Q1, 25th percentile)
#                   var <= qhigh(var)     (= Q3, 75th percentile)
#               This is the richer strategy and is the package default.
#
#   "median"  : per continuous variable, a SINGLE candidate cut:
#                   var <= median(var)
#               Use this to stress-test the "truly unknown cuts" regime
#               when the true cut is not at the median.
#
# Honest disclosure for this setup:
#   The true Q-defining cuts are age > 34 (~median of arms 1+3) and
#   preanti <= 744.5 (~Q3 of arms 1+3).  With cut_type = "default",
#   BOTH true cuts are present in the candidate set (age at median,
#   preanti at Q3).  With cut_type = "median", only the age cut is
#   recoverable; preanti's true Q3 cut is NOT in the candidate set,
#   forcing forestsearch to approximate Q via the wrong threshold.
#
# The default value ("default") is kept here so the simulation
# evaluates the realistic setting where the analyst uses the richer
# cut strategy a priori (without knowing the truth).  GRF candidate
# cuts (use_grf = TRUE) are additionally appended to this set.
fs_cut_type            <- "default"

# -- GRF standalone parameters ------------------------------------------------
# grf.subg.harm.glm() uses causal_forest() for candidate factor screening,
# then policy trees for subgroup identification.  With continuous inputs,
# policy trees automatically search over all candidate thresholds.
#
# dmin.grf: minimum absolute CATE (risk difference) for a policy-tree
#   node to qualify as a subgroup.  The calibration grids referenced
#   below were built on discretised inputs; with raw continuous inputs
#   the effective candidate-cut space is larger, which typically
#   INCREASES FPR slightly at a given dmin.grf.  Re-calibration is
#   recommended for production runs; dmin.grf = 0.10 is the current
#   conservative default carried over from the discretised setting.
grf_dmin               <- 0.10
grf_maxdepth           <- 2L
grf_sg_criterion       <- "mDiff"
grf_tune               <- FALSE

# -- ForestSearch pipeline flags -----------------------------------------------
fs_use_lasso           <- FALSE
fs_use_grf             <- TRUE
fs_is_rct              <- TRUE

# -- Reporting -----------------------------------------------------------------
sg_notation            <- "benefit"

# -- Output persistence --------------------------------------------------------
save_results           <- TRUE
output_dir             <- "../_data"

```

```{r timing-init}
#| include: false
t_vignette_start <- proc.time()
timings <- list()
```


# Introduction {#sec-intro}

## Motivation

This document extends the ACTG175 benefit-search simulation
framework from survival outcomes (time-to-event) to a **binary
endpoint**: whether a patient's CD4 count improved from baseline
to 20 weeks, defined as $Y = I(\text{cd420} > \text{cd40})$.

The simulation uses the `forestsearch` GLM simulation suite
(`generate_glm_dgm`, `calibrate_glm_interaction`,
`simulate_from_glm_dgm`) and routes through
`run_simulation_analysis()` via class-based dispatch on the
`"glm_dgm"` object.

## Realistic Analysis Setting

A prior version of this simulation discretised every continuous
covariate into a binary factor **before** handing the data to
`forestsearch()` / `grf.subg.harm.glm()`.  The binarisation used
the same cuts that define the true subgroup $Q$ (e.g. age > 34,
preanti $\leq$ 744.5).  That setup implicitly gives the analysis
method knowledge of the true cutpoints, which is unrealistic.

The current setup **retains continuous covariates in their raw
form for the analysis step**.  The data-generating mechanism
(DGM) is built from the discretised factors (`z1`--`z12`) because
`generate_glm_dgm()` requires factor covariates and the true
subgroup must be defined algebraically.  The analysis methods,
however, see the **raw** continuous columns (`age`, `preanti`,
`wtkg`, `karnof`, `cd40`, `cd80`) along with the six truly
binary indicators (`hemo`, `homo`, `drugs`, `race`, `gender`,
`symptom`).  Cutpoints for the continuous covariates are
discovered by the search methods themselves (median splits,
GRF-derived cuts, policy-tree splits), exactly as in a real
analysis.

::: {.callout-important}
## What changes vs the prior (discretised) setup

- **Ground truth is unchanged**: the true benefit subgroup $Q$
  is still $\{\text{age} > 34\} \cap \{\text{preanti} \leq 744.5\}$
  with prevalence $\approx 34\%$, and the DGM's OR(Q) calibration
  is untouched.
- **Classification metrics are unchanged in definition**: `flag_harm`
  is computed identically from `z1` and `z2`.
- **What changes is the search problem**: ForestSearch must
  discover the cuts from candidate thresholds; GRF builds policy
  trees over raw continuous inputs.  Expect lower sensitivity and
  PPV than the discretised setting, because the analyst does not
  receive oracle cuts.

**Cut-set coverage for this run.**  ForestSearch is configured with
`cut_type = "default"` (the package default), which generates four
candidate cuts per continuous covariate: mean, median, Q1, and Q3.
For the ACTG175 covariate distribution the median of `age` is
approximately 34 and Q3 of `preanti` is approximately 744.5, so
**both true Q-defining cuts happen to fall inside the default
candidate set** — they are simply not privileged over the other
candidates.  The analyst has no oracle knowledge of which cuts
matter; ForestSearch must select them based on the data.  Switching
to `cut_type = "median"` would exclude preanti's Q3 cut from the
candidate set entirely, providing a harder stress test.  GRF policy
trees search over all continuous thresholds internally and are
unaffected by this setting.
:::

## Statistical Goals

This simulation study addresses three questions about
ForestSearch and GRF standalone subgroup identification
applied to **binary outcomes** in the realistic (raw covariate)
setting:

1. **Detection power (H1)**: When a benefit subgroup $Q$
   exists with OR(Q) = `r cal_target_or` (switched scale),
   how often do ForestSearch and GRF identify *any* subgroup?
2. **False positive control (H0)**: When the treatment effect
   is homogeneous (no true $Q$), how often does each method
   falsely declare a benefit subgroup?
3. **Classification accuracy (H1)**: Conditional on detection,
   how well does the estimated $\hat{Q}$ recover the true $Q$?
   Measured by sensitivity ($\text{sens}(\hat{Q})$: fraction of
   true $Q$ captured) and positive predictive value
   ($\text{ppv}(\hat{Q})$: fraction of $\hat{Q}$ truly in $Q$).

::: {.callout-tip}
## Summary of results
See [Summary] at the end for a data-driven
summary of detection, classification, and timing from this run.
:::

## Simulation Pipeline

The simulation proceeds in six steps:

1. **Data preparation** (@sec-data): Load the ACTG175 trial,
   restrict to arms 1 (ZDV+ddI) and 3 (ddI), switch treatment
   labels so ddI = `treat = 1` and ZDV+ddI = `treat = 0`
   (benefit search convention), and construct BOTH:
   (a) the binary `z1`--`z12` factor set needed by the DGM; and
   (b) the realistic analysis candidate set
   (`age`, `preanti`, `wtkg`, `karnof`, `cd40`, `cd80`, `hemo`,
   `homo`, `drugs`, `race`, `gender`, `symptom`).

2. **DGM construction** (@sec-dgm): Fit a logistic regression
   to the observed data using the `z1`--`z12` factor
   formulation, create a 5,000-subject super-population, and
   define the true benefit subgroup
   $Q = \{z_1 = 1 \cap z_2 = 1\}$ (age > `r actg_age_cut`
   AND preanti $\leq$ `r actg_preanti_cut`).  The super-
   population retains the **raw** covariate columns as well as
   the `z1`--`z12` indicators.

3. **Calibration** (@sec-calibrate): Search over `k_inter`
   values to find the interaction strength that produces
   $\text{OR}(Q) = `r cal_target_or`$ on the switched scale
   (= $`r round(1/cal_target_or, 2)`$ on the original scale).

4. **Single-data analysis** (@sec-single): Run ForestSearch
   and GRF on the observed ACTG175 data using the **raw**
   analysis candidate set to verify the pipeline produces
   valid results before scaling to simulation.

5. **Monte Carlo simulation** (@sec-sims): For each of
   `r nsims_alt` alternative and `r nsims_null` null
   replicates, draw a random sample of $N = `r sim_n_sample`$
   from the DGM super-population, run ForestSearch and GRF
   on the raw analysis covariates, and record detection,
   classification, and effect estimates.

6. **Evaluation** (@sec-results): Compute operating
   characteristics (detection, sensitivity, PPV, effect
   estimation bias) under both hypotheses.

## Benefit Search on the OR Scale

The benefit-search approach from Leon et al. (2024, Section 4.2)
applies identically to binary outcomes, with odds ratios (OR)
replacing hazard ratios (HR):

1. **Switch treatment roles**: ddI monotherapy $\to$ `treat = 1`,
   ZDV+ddI combination $\to$ `treat = 0`.

2. **Raise thresholds** on the OR scale:
   $c_1 = 1/0.60 \approx 1.667$, $c_2 = 1/0.80 = 1.25$.

3. **Run ForestSearch** with `outcome_type = "binary"`,
   `effect_measure = "OR"`.

4. **Inversion** is handled automatically by the package when
   `subgroup_notation = "benefit"` is passed to the reporting
   functions.


# GRF Standalone: Calibration Notes {#sec-grf-calibration}

The standalone GRF analysis uses `grf.subg.harm.glm()` with
`causal_forest()` for candidate variable screening and policy
trees for subgroup identification.  The key tuning parameter
is `dmin.grf`, the minimum absolute CATE (conditional average
treatment effect) on the **risk difference** scale required for
a policy-tree node to be valid.

## Configuration

For the **benefit + positive** scenario with switched treatment
(the ACTG175 binary configuration), the correct GRF setup is:

- `adverse_outcome = FALSE` (Y = improvement is positive)
- Treatment indicator = **switched** (ddI = 1, ZDV+ddI = 0)
- GRF identifies the subgroup where the switched "treatment"
  produces the most negative CATE --- i.e., where ddI monotherapy
  is worst relative to combination, which is the **benefit**
  subgroup on the original scale.

## Impact of Raw Continuous Inputs on Calibration

Prior `dmin.grf` calibration (in `grf_dmin_calibration_binary.qmd`)
was conducted on **discretised** inputs (`z1`--`z12` binary
factors) and recommended `dmin.grf = 0.02` for the benefit +
positive binary configuration.

With raw continuous inputs the policy-tree search evaluates
many more candidate thresholds per covariate, which tends to:

- **Widen the null distribution of the best CATE**, inflating
  the false positive rate at a given `dmin.grf`.
- **Improve sensitivity to the true cuts** when they align with
  natural splits in the covariate distribution.

The current document uses `dmin.grf = `r grf_dmin`` as a
conservative default carried over from the discretised setting.
A dedicated calibration run with raw continuous inputs is
recommended before production scaling.

::: {.callout-note}
The GRF FPR is higher than ForestSearch's because GRF lacks the
bootstrap consistency filter.  ForestSearch requires the subgroup
effect to exceed a threshold in 90% of bootstrap replicates,
providing stronger FPR control at the cost of lower detection.
The two methods are complementary: GRF provides a
threshold-free comparison.
:::


# Data Preparation {#sec-data}

```{r load-packages}
library(forestsearch)
library(survival)
library(speff2trial)
library(data.table)
library(gt)
library(ggplot2)
library(doFuture)
library(foreach)

# -- Safe foreach result collection -------------------------------------------
# Replaces .errorhandling = "remove" (which silently drops failures)
# with .errorhandling = "pass" + post-hoc separation of errors from results.
collect_results <- function(raw_list, label = "") {
  is_err <- vapply(raw_list, inherits, logical(1), "error")
  n_failed <- sum(is_err)
  if (n_failed > 0) {
    err_msgs <- vapply(raw_list[is_err], conditionMessage, character(1))
    cat(sprintf("*** %s%d of %d tasks failed ***\n",
        if (nzchar(label)) paste0(label, ": ") else "",
        n_failed, length(raw_list)))
    cat("  First errors:\n")
    for (msg in head(unique(err_msgs), 3)) cat("    ", msg, "\n")
  }
  good <- raw_list[!is_err]
  if (length(good) == 0) {
    result <- data.frame()
    warning(sprintf("%sAll %d tasks failed. See error messages above.",
        if (nzchar(label)) paste0(label, ": ") else "",
        length(raw_list)))
  } else if (requireNamespace("data.table", quietly = TRUE)) {
    result <- data.table::rbindlist(good, use.names = TRUE, fill = TRUE)
    result <- as.data.frame(result)
  } else {
    result <- do.call(rbind, good)
  }
  attr(result, "n_failed") <- n_failed
  attr(result, "n_total") <- length(raw_list)
  result
}
```

```{r prepare-datasets}
t0 <- proc.time()[3]

# -- ACTG175: arms 1 (ZDV+ddI) vs 3 (ddI) ------------------------------------
actg_df <- subset(speff2trial::ACTG175, arms %in% actg_arms)
actg_df$id <- seq_len(nrow(actg_df))

# Original treatment: arm 1 = 1, arm 3 = 0
actg_df$treat_orig <- ifelse(actg_df$arms == actg_treat_arm, 1L, 0L)

# Switched treatment for benefit search
actg_df$treat <- 1L - actg_df$treat_orig

# -- Binary outcome: CD4 response at 20 weeks ---------------------------------
actg_df$y_binary <- as.integer(actg_df$cd420 > actg_df$cd40)

# Remove subjects with missing cd420
n_before <- nrow(actg_df)
actg_df <- actg_df[!is.na(actg_df$cd420), ]
n_after  <- nrow(actg_df)

cat(sprintf("ACTG175: N = %d -> %d after removing %d with missing cd420\n",
    n_before, n_after, n_before - n_after))
cat(sprintf("Overall response rate: %.1f%%\n",
    100 * mean(actg_df$y_binary)))
cat(sprintf("  ZDV+ddI (arm 1): %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$treat_orig == 1])))
cat(sprintf("  ddI (arm 3):     %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$treat_orig == 0])))

# =============================================================================
# Two parallel covariate views are constructed on the SAME data frame:
#
#   (1) DGM factor set: z1--z12 binary factors. Used by generate_glm_dgm()
#       and calibrate_glm_interaction() to fit the baseline logistic
#       regression and define the TRUE subgroup Q = {z1=1 & z2=1}.
#
#   (2) Analysis candidate set: raw continuous covariates + truly binary
#       indicators. Used by forestsearch() and grf.subg.harm.glm() in
#       both the observed-data analysis and every simulated replicate.
#       The analyst does NOT know the cuts that define Q.
# =============================================================================

# -- (1) DGM factor set: z1-z12 binary factors --------------------------------
#
# The z-series preserves the discretisation from León et al. (2024,
# Section 4.2). These are needed by the DGM because generate_glm_dgm()
# requires factor covariates, and because the true Q is defined
# algebraically in terms of z1 and z2.
actg_df$z1  <- as.factor(ifelse(actg_df$age > actg_age_cut, 1L, 0L))
actg_df$z2  <- as.factor(ifelse(actg_df$preanti <= actg_preanti_cut, 1L, 0L))
actg_df$z3  <- as.factor(ifelse(actg_df$wtkg <= 75, 1L, 0L))
actg_df$z4  <- as.factor(ifelse(
  actg_df$karnof <= median(actg_df$karnof), 1L, 0L))
actg_df$z5  <- as.factor(ifelse(
  actg_df$cd40 <= median(actg_df$cd40), 1L, 0L))
actg_df$z6  <- as.factor(ifelse(
  actg_df$cd80 <= median(actg_df$cd80), 1L, 0L))
actg_df$z7  <- as.factor(actg_df$hemo)
actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs)
actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender)
actg_df$z12 <- as.factor(actg_df$symptom)

# -- (2) Analysis candidate set: raw + binary ---------------------------------
#
# The continuous columns (age, preanti, wtkg, karnof, cd40, cd80) are
# ALREADY in the ACTG175 dataframe in their raw form. We make the six
# truly-binary indicators into factors so that forestsearch's
# is.continuous() detection treats them as categorical without
# attempting median/GRF cuts.
for (v in analysis_binary_vars) {
  actg_df[[v]] <- as.factor(actg_df[[v]])
}

confounders_analysis <- c(analysis_continuous_vars, analysis_binary_vars)

# Sanity: confirm types for the analyst-visible candidate set
cov_types <- vapply(confounders_analysis,
  function(v) if (is.factor(actg_df[[v]])) "factor" else class(actg_df[[v]])[1],
  character(1))

cat("\nAnalysis candidate covariates (seen by forestsearch / GRF):\n")
for (v in confounders_analysis) {
  cat(sprintf("  %-10s  %-9s  n_unique = %d\n",
      v, cov_types[v], length(unique(actg_df[[v]]))))
}

# -- True subgroup flag (for reference; generate_glm_dgm creates its own) -----
actg_df$flag_harm_ref <- as.integer(
  actg_df$z1 == 1 & actg_df$z2 == 1
)
cat(sprintf("\nSubgroup Q: age > %d AND preanti <= %.1f\n",
    actg_age_cut, actg_preanti_cut))
cat(sprintf("  Q prevalence: %d / %d (%.1f%%)\n",
    sum(actg_df$flag_harm_ref), nrow(actg_df),
    100 * mean(actg_df$flag_harm_ref)))

cat(sprintf("\nResponse rates by subgroup (original scale):\n"))
cat(sprintf("  Q:  ZDV+ddI = %.1f%%,  ddI = %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 1 &
                                  actg_df$treat_orig == 1]),
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 1 &
                                  actg_df$treat_orig == 0])))
cat(sprintf("  Qc: ZDV+ddI = %.1f%%,  ddI = %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 0 &
                                  actg_df$treat_orig == 1]),
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 0 &
                                  actg_df$treat_orig == 0])))

timings$data_prep <- proc.time()[3] - t0
```


# Creating the Binary DGM {#sec-dgm}

The GLM DGM uses `generate_glm_dgm()` to fit a logistic regression
on the observed ACTG175 data, create a super-population with
individual-level potential outcomes ($p_0$, $p_1$), and compute
true subgroup effects.  `calibrate_glm_interaction()` then searches
for the `k_inter` that produces the target OR(Q).

**Note on DGM inputs.**  The DGM continues to use the `z1`--`z12`
factor formulation.  This is a design requirement of
`generate_glm_dgm()` (which coerces any input to `factor` via
`as.factor()`) and is also the natural way to define the true
subgroup $Q$ algebraically.  The super-population returned by
the DGM *retains all original columns of `actg_df`*, including
the raw `age`, `preanti`, `wtkg`, `karnof`, `cd40`, `cd80`
columns, which are then used by the analysis step.

## Step 1: Build Initial DGM

```{r create-dgm-alt}
t0 <- proc.time()[3]

dgm_alt_init <- generate_glm_dgm(
  data          = actg_df,
  factor_vars   = dgm_factor_vars,
  outcome_var   = "y_binary",
  treatment_var = "treat",
  outcome_type  = "binary",
  effect_measure = "OR",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  model         = "alt",
  k_inter       = 0,    # no interaction before calibration
  n_super       = dgm_n_super,
  seed          = sim_seed,
  verbose       = TRUE
)

cat("\n=== Initial ACTG175 Binary DGM ===\n")
cat("  OR(Q):  ", round(dgm_alt_init$hazard_ratios$harm_subgroup, 3), "\n")
cat("  OR(Qc): ", round(dgm_alt_init$hazard_ratios$no_harm_subgroup, 3), "\n")
cat("  OR(ITT):", round(dgm_alt_init$hazard_ratios$overall, 3), "\n")

# Sanity check: super-population retains raw continuous columns
missing_raw <- setdiff(analysis_continuous_vars, names(dgm_alt_init$df_super))
if (length(missing_raw) > 0) {
  stop("DGM super-population is missing raw continuous columns: ",
       paste(missing_raw, collapse = ", "))
} else {
  cat("\nRaw continuous columns present in super-population: ",
      paste(analysis_continuous_vars, collapse = ", "), "\n")
}

timings$dgm_init <- proc.time()[3] - t0
```

## Step 2: Calibrate for Target OR {#sec-calibrate}

### Why calibrate?

The DGM's `k_inter` parameter controls the strength of the
treatment-by-subgroup interaction on the log-odds scale.  Because
the mapping from `k_inter` to the marginal OR(Q) is non-linear and
depends on the covariate distribution and baseline response rates,
we search over a grid of candidate values.

`calibrate_glm_interaction()` automates this process: it evaluates
`generate_glm_dgm()` at each grid point and selects the `k_inter`
closest to the target.

```{r calibrate}
t0 <- proc.time()[3]

dgm_calibrated <- calibrate_glm_interaction(
  data          = actg_df,
  factor_vars   = dgm_factor_vars,
  outcome_var   = "y_binary",
  treatment_var = "treat",
  target_effect = cal_target_or,
  outcome_type  = "binary",
  effect_measure = "OR",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  k_inter_range = cal_k_grid_range,
  grid_step     = cal_grid_step,
  n_super       = dgm_n_super,
  seed          = sim_seed,
  verbose       = TRUE
)

cat("\n=== Calibrated ACTG175 Binary DGM ===\n")
cat("  OR(Q):  ", round(dgm_calibrated$hazard_ratios$harm_subgroup, 3),
    " (target:", cal_target_or, ")\n")
cat("  OR(Qc): ", round(dgm_calibrated$hazard_ratios$no_harm_subgroup, 3), "\n")
cat("  OR(ITT):", round(dgm_calibrated$hazard_ratios$overall, 3), "\n")
cat("  k_inter:", dgm_calibrated$model_params$k_inter, "\n")
cat("  -> Original scale: OR(Q) =",
    round(1 / dgm_calibrated$hazard_ratios$harm_subgroup, 3), "\n")

timings$calibrate <- proc.time()[3] - t0
```

### Calibration Curve

```{r calibration-plot, fig.height=5}
#| fig-cap: "Calibration of k_inter: OR(Q) as a function of the interaction parameter."

cal_grid <- seq(cal_k_grid_range[1], cal_k_grid_range[2], by = cal_grid_step)
cal_or_Q <- vapply(cal_grid, function(k) {
  d <- generate_glm_dgm(
    data = actg_df, factor_vars = dgm_factor_vars,
    outcome_var = "y_binary", treatment_var = "treat",
    outcome_type = "binary", subgroup_vars = dgm_subgroup_vars,
    subgroup_cuts = dgm_subgroup_cuts, model = "alt",
    k_inter = k, n_super = dgm_n_super, seed = sim_seed
  )
  d$hazard_ratios$harm_subgroup
}, numeric(1))

cal_df <- data.frame(k_inter = cal_grid, or_Q = cal_or_Q)
best_k <- dgm_calibrated$model_params$k_inter

ggplot(cal_df, aes(x = k_inter, y = or_Q)) +
  geom_line(linewidth = 1, colour = "firebrick") +
  geom_hline(yintercept = cal_target_or, linetype = "dashed",
             colour = "grey50") +
  geom_hline(yintercept = 1.0, linetype = "dotted",
             colour = "grey70") +
  geom_point(
    data = data.frame(k_inter = best_k,
                      or_Q = dgm_calibrated$hazard_ratios$harm_subgroup),
    colour = "red", size = 4, shape = 16
  ) +
  labs(
    x = "k_inter",
    y = "OR(Q) on switched scale",
    title = "ACTG175 Binary DGM Calibration",
    subtitle = sprintf(
      "Best: k_inter = %.2f -> OR(Q) = %.3f (target: %.1f)",
      best_k, dgm_calibrated$hazard_ratios$harm_subgroup, cal_target_or
    )
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```


## Step 3: Create Null DGM

Under `model = "null"`, the treatment effect is **homogeneous**
across all patients (no treatment effect heterogeneity).
Per León et al. (2024, Section 3), $Q = \emptyset$ and
$Q^c$ is the ITT population.  The `flag_harm` column is
retained as a structural label in the simulated data (marking
$z_1 = 1 \cap z_2 = 1$) but has **no treatment-effect
meaning**: there is no true benefit subgroup under this
hypothesis.

```{r dgm-null}
t0 <- proc.time()[3]

dgm_null <- generate_glm_dgm(
  data          = actg_df,
  factor_vars   = dgm_factor_vars,
  outcome_var   = "y_binary",
  treatment_var = "treat",
  outcome_type  = "binary",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  model         = "null",
  n_super       = dgm_n_super,
  seed          = sim_seed,
  verbose       = TRUE
)

cat("\n=== ACTG175 Null DGM ===\n")
cat("  OR(ITT):", round(dgm_null$hazard_ratios$overall, 3), "\n")
cat("  OR(Q):  ", round(dgm_null$hazard_ratios$harm_subgroup, 3), "\n")
cat("  OR(Qc): ", round(dgm_null$hazard_ratios$no_harm_subgroup, 3), "\n")

timings$dgm_null <- proc.time()[3] - t0
```


# Single-Data Analysis {#sec-single}

Before any simulation, run ForestSearch on the **actual ACTG175
data** with `details = TRUE` to examine every step of the
procedure: LASSO screening, GRF variable importance, candidate
enumeration, and bootstrap consistency.

The analysis uses the **realistic candidate set**:

```{r show-analysis-set}
#| code-fold: false
cat("Analysis candidate covariates (no oracle cuts):\n")
cat("  Continuous:", paste(analysis_continuous_vars, collapse = ", "), "\n")
cat("  Binary:    ", paste(analysis_binary_vars, collapse = ", "), "\n")
cat("\nForestsearch / GRF will discover cuts on the continuous\n")
cat("covariates themselves; the TRUE cuts (age > 34, preanti <= 745)\n")
cat("are NOT provided.\n")
```

```{r single-data}
#| code-fold: false
t0 <- proc.time()[3]

cat("=", rep("=", 60), "\n", sep = "")
cat("  ForestSearch on ACTG175 observed data\n")
cat("  outcome_type = 'binary', effect_measure = 'OR'\n")
cat("  Treatment: switched (ddI = 1, ZDV+ddI = 0)\n")
cat("  Y = I(cd420 > cd40) = improvement\n")
cat("  Covariates: RAW continuous + binary indicators\n")
cat("  Thresholds: c1 =", fs_or_threshold,
    " c2 =", fs_or_consistency, "\n")
cat("=", rep("=", 60), "\n\n", sep = "")

fs_actg <- forestsearch(
  df.analysis      = actg_df,
  confounders.name = confounders_analysis,
  outcome.name     = "y_binary",
  event.name       = "y_binary",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "OR",
  seedit           = fs_seedit,
  n.min            = fs_n_min,
  d0.min           = fs_d0_min,
  d1.min           = fs_d1_min,
  hr.threshold     = fs_or_threshold,
  hr.consistency   = fs_or_consistency,
  pconsistency.threshold = fs_pconsistency,
  fs.splits        = fs_splits,
  maxk             = fs_maxk,
  use_lasso        = fs_use_lasso,
  use_grf          = fs_use_grf,
  is.RCT           = fs_is_rct,
  vi.grf.min       = fs_vi_grf_min,
  cont.cutoff      = fs_cont_cutoff,
  cut_type         = fs_cut_type,
  details          = TRUE,
  quiet            = FALSE,
  parallel_args    = list(
    plan = "sequential", workers = 1, show_message = FALSE
  )
)

timings$single_data <- proc.time()[3] - t0
```

## Examine Result

```{r examine-single}
#| code-fold: false

cat("\n=== ForestSearch Result ===\n")

if (!is.null(fs_actg$grp.consistency)) {
  # NOTE: for a forestsearch object, the cut expressions defining the
  # identified subgroup live in `grp.consistency$sg.harm` (character
  # vector of factor-level cut names).  `grp.consistency$sg.harm.id`
  # is a different object: a per-subject 0/1 indicator of subgroup
  # membership with length nrow(df).  (For GRF objects, `sg.harm.id`
  # does carry the cut expressions — same name, different semantics.)
  sg <- fs_actg$grp.consistency$sg.harm
  if (!is.null(sg) && length(sg) > 0 && !all(is.na(sg))) {
    cat("Subgroup identified:", paste(sg, collapse = " & "), "\n")

    df_est <- fs_actg$df.est
    in_H  <- df_est$treat.recommend == 0
    in_Hc <- df_est$treat.recommend == 1

    cat(sprintf("  Size(H): %d,  Size(Hc): %d\n",
        sum(in_H), sum(in_Hc)))

    if (sum(in_H) >= 10) {
      fit_H <- glm(y_binary ~ treat, data = df_est[in_H, ],
                    family = binomial)
      or_H <- exp(coef(fit_H)["treat"])
      cat(sprintf("  OR(H, switched):  %.3f  (inverted: %.3f)\n",
          or_H, 1/or_H))
    }
    if (sum(in_Hc) >= 10) {
      fit_Hc <- glm(y_binary ~ treat, data = df_est[in_Hc, ],
                     family = binomial)
      or_Hc <- exp(coef(fit_Hc)["treat"])
      cat(sprintf("  OR(Hc, switched): %.3f  (inverted: %.3f)\n",
          or_Hc, 1/or_Hc))
    }

    fit_itt <- glm(y_binary ~ treat, data = df_est, family = binomial)
    cat(sprintf("  OR(ITT, switched): %.3f\n",
        exp(coef(fit_itt)["treat"])))

    # Classification against true Q (defined by raw age & preanti cuts)
    true_H <- df_est$flag_harm_ref == 1
    if (sum(true_H) > 0 && sum(in_H) > 0) {
      tp <- sum(in_H & true_H);  fp <- sum(in_H & !true_H)
      fn <- sum(!in_H & true_H); tn <- sum(!in_H & !true_H)
      cat(sprintf("\n  Classification vs true Q (age > %d & preanti <= %.1f):\n",
          actg_age_cut, actg_preanti_cut))
      cat(sprintf("    TP: %d  FP: %d  FN: %d  TN: %d\n",
          tp, fp, fn, tn))
      cat(sprintf("    Sens: %.3f  Spec: %.3f  PPV: %.3f\n",
          tp/(tp+fn), tn/(tn+fp), tp/(tp+fp)))
    }
  } else {
    cat("No subgroup identified by ForestSearch.\n")
  }
} else {
  cat("ForestSearch did not return grp.consistency.\n")
}

cat(sprintf("\nSingle-data analysis: %.1f sec\n", timings$single_data))
```


## GRF Standalone on Observed Data

```{r grf-single-data}
#| code-fold: false

grf_actg <- grf.subg.harm.glm(
  data             = actg_df,
  confounders.name = confounders_analysis,
  outcome.name     = "y_binary",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "OR",
  dmin.grf         = grf_dmin,
  maxdepth         = grf_maxdepth,
  RCT              = TRUE,
  sg.criterion     = grf_sg_criterion,
  details          = TRUE
)

if (!is.null(grf_actg$sg.harm.id) && length(grf_actg$sg.harm.id) > 0) {
  cat("GRF subgroup:", paste(grf_actg$sg.harm.id, collapse = " & "), "\n")
  grf_H <- grf_actg$data$treat.recommend == 0
  cat(sprintf("  Size(H): %d,  Size(Hc): %d\n",
      sum(grf_H), sum(!grf_H)))
} else {
  cat("GRF: no subgroup identified.\n")
}
```


## FS vs GRF Concordance on Observed Data

When a method does not detect a subgroup, its classification is
"all subjects to $Q^c$" (`treat.recommend = 1` for everyone).
`compare_fs_subgroups()` handles this automatically — no manual
fallback is needed.

```{r fs-grf-concordance}
sg_compare <- compare_fs_subgroups(
  fs1                = fs_actg,
  fs2                = grf_actg,
  label1             = "ForestSearch",
  label2             = "GRF Standalone",
  id.name            = "id",
  subgroup_notation  = sg_notation,
  title              = "FS vs GRF: Subgroup Membership Concordance",
  subtitle           = "ACTG175 observed data (binary benefit search, raw continuous)",
  font_size          = 13
)

sg_compare$table
```

```{r fs-grf-concordance-detail}
cat("Concordance statistics:\n")
cat(sprintf("  Agreement: %.1f%%\n",
    100 * sg_compare$concordance$agreement))
cat(sprintf("  Cohen's kappa: %.3f\n",
    sg_compare$concordance$kappa))
cat(sprintf("  Concordant: %d / Discordant: %d / Total: %d\n",
    sg_compare$concordance$n_agree,
    sg_compare$concordance$n_disagree,
    sg_compare$concordance$n_total))
```


# Single Trial Simulation {#sec-single-trial}

Before the full simulation study, verify that `simulate_from_glm_dgm()`
and `run_simulation_analysis()` work end-to-end on a single replicate.

```{r single-trial}
#| code-fold: false
t0 <- proc.time()[3]

# Simulate one trial from the calibrated DGM
df_single <- simulate_from_glm_dgm(
  dgm  = dgm_calibrated,
  n    = sim_n_sample,
  seed = sim_seed + 1L
)

cat(sprintf("Single trial: N = %d, response rate = %.1f%%\n",
    nrow(df_single), 100 * mean(df_single$y_sim)))
cat(sprintf("  treat allocation: %.1f%%\n",
    100 * mean(df_single$treat_sim)))
cat(sprintf("  Q prevalence: %.1f%%\n",
    100 * mean(df_single$flag_harm)))

# Verify OR in the simulated data
fit_single <- glm(y_sim ~ treat_sim, data = df_single, family = binomial)
cat(sprintf("  Simulated ITT OR: %.3f (DGM: %.3f)\n",
    exp(coef(fit_single)["treat_sim"]),
    dgm_calibrated$hazard_ratios$overall))

# Confirm raw continuous columns are present in simulated data
missing_in_sim <- setdiff(confounders_analysis, names(df_single))
if (length(missing_in_sim) > 0) {
  stop("Simulated data is missing analysis columns: ",
       paste(missing_in_sim, collapse = ", "))
}

timings$single_trial <- proc.time()[3] - t0
```


# Running Simulation Studies {#sec-sims}

## Setup

```{r sim-setup}
#| code-fold: false

plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")

# -- Candidate variable set (RAW continuous + binary) --------------------------
# This is the key behavioural change vs the prior discretised setup:
# forestsearch() and grf.subg.harm.glm() receive raw continuous inputs
# and must discover cuts.
cat("Analysis confounders passed to FS/GRF:\n")
cat("  ", paste(confounders_analysis, collapse = ", "), "\n")

# -- ForestSearch parameters ---------------------------------------------------
fs_params <- list(
  # Column name mappings
  outcome.name           = "y_sim",
  event.name             = "y_sim",
  treat.name             = "treat_sim",
  id.name                = "id",

  # GLM pipeline
  outcome_type           = "binary",
  effect_measure         = "OR",

  # Variable screening
  use_lasso              = fs_use_lasso,
  use_grf                = fs_use_grf,

  # Continuous-variable detection threshold
  cont.cutoff            = fs_cont_cutoff,

  # Candidate cut strategy for continuous covariates
  # ("default" = mean + median + Q1 + Q3 per continuous variable;
  #  "median"  = median only). See setup chunk for full explanation.
  cut_type               = fs_cut_type,

  # Subgroup detection thresholds
  # Note: hr.threshold/hr.consistency are the generic ForestSearch
  # parameter names regardless of endpoint type. For binary outcomes,
  # these represent OR thresholds on the switched scale.
  hr.threshold           = fs_or_threshold,
  hr.consistency         = fs_or_consistency,
  pconsistency.threshold = fs_pconsistency,

  # Computational constraints
  fs.splits              = fs_splits,
  n.min                  = fs_n_min,
  d0.min                 = fs_d0_min,
  d1.min                 = fs_d1_min,
  maxk                   = fs_maxk,
  vi.grf.min             = fs_vi_grf_min,
  seedit                 = fs_seedit,

  # GRF tuning (applies to ForestSearch's internal GRF pre-screening)
  tune_grf               = grf_tune,

  # Inner parallelism: sequential within each worker
  parallel_args          = list(
    plan = "sequential", workers = 1, show_message = FALSE
  )
)

# -- GRF standalone parameter overrides ----------------------------------------
grf_params <- list(
  dmin.grf     = grf_dmin,
  maxdepth     = grf_maxdepth,
  sg.criterion = grf_sg_criterion,
  tune_grf     = grf_tune
)
```


## Alternative Hypothesis Simulations

```{r run-alt}
cat("Running", nsims_alt, "simulations under H1...\n")
t0 <- proc.time()[3]

results_alt_raw <- foreach(
  sim = seq_len(nsims_alt),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_calibrated,
    n_sample         = sim_n_sample,
    confounders_base = confounders_analysis,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    grf_params       = grf_params,
    verbose          = FALSE
  )
}

timings$sims_alt <- proc.time()[3] - t0

# Post-hoc failure accounting
results_alt <- collect_results(results_alt_raw, "H1 alt")
n_failed_alt <- attr(results_alt, "n_failed")

cat("Completed in", round(timings$sims_alt / 60, 1), "minutes\n")
cat("Results:", nrow(results_alt), "rows\n")
if (n_failed_alt > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_alt))
if (nrow(results_alt) > 0) {
  cat(sprintf("Detection rate: %.1f%%\n",
      100 * mean(results_alt$any.H[results_alt$analysis == "FS"])))
} else {
  cat("All simulations failed — see error messages above.\n")
  cat("Common causes:\n")
  cat("  1. devtools::document() -> devtools::install() -> restart R needed\n")
  cat("  2. Package version mismatch between interactive and worker sessions\n")
}
```


## Null Hypothesis Simulations

```{r reset-workers-null}
# Reset worker pool to prevent memory accumulation from alt sims
plan(sequential); gc(verbose = FALSE); Sys.sleep(1)
plan(multisession, workers = n_workers)
```

```{r run-null}
cat("Running", nsims_null, "simulations under H0...\n")
t0 <- proc.time()[3]

results_null_raw <- foreach(
  sim = seq_len(nsims_null),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_null,
    n_sample         = sim_n_sample,
    confounders_base = confounders_analysis,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    grf_params       = grf_params,
    verbose          = FALSE
  )
}

timings$sims_null <- proc.time()[3] - t0

results_null <- collect_results(results_null_raw, "H0 null")
n_failed_null <- attr(results_null, "n_failed")

cat("Completed in", round(timings$sims_null / 60, 1), "minutes\n")
if (n_failed_null > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_null))
cat(sprintf("False positive rate: %.1f%%\n",
    100 * mean(results_null$any.H[results_null$analysis == "FS"])))
```

```{r cleanup-plan}
plan(sequential)
```

```{r save-results}
#| code-fold: false

if (save_results) {
  if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

  bundle <- list(
    # -- Core results ----------------------------------------------------------
    results_alt  = results_alt,
    results_null = results_null,

    # -- DGM objects -----------------------------------------------------------
    dgm_calibrated = dgm_calibrated,
    dgm_null       = dgm_null,

    # -- Simulation configuration ----------------------------------------------
    config = list(
      nsims_alt    = nsims_alt,
      nsims_null   = nsims_null,
      n_sample     = sim_n_sample,
      n_noise      = sim_n_noise,
      sim_seed     = sim_seed,
      n_workers    = n_workers,
      outcome_type = "binary",
      effect_measure = "OR",
      # Document the analysis candidate set used (important provenance
      # for comparing to prior discretised runs)
      analysis_continuous_vars = analysis_continuous_vars,
      analysis_binary_vars     = analysis_binary_vars,
      confounders_analysis     = confounders_analysis
    ),

    # -- ForestSearch parameters -----------------------------------------------
    fs_params = fs_params,

    # -- GRF parameters --------------------------------------------------------
    grf_params = grf_params,

    # -- DGM true effects (convenience) ----------------------------------------
    true_effects = list(
      or_Q    = dgm_calibrated$hazard_ratios$harm_subgroup,
      or_Qc   = dgm_calibrated$hazard_ratios$no_harm_subgroup,
      or_ITT  = dgm_calibrated$hazard_ratios$overall,
      k_inter = dgm_calibrated$model_params$k_inter
    ),

    # -- Provenance ------------------------------------------------------------
    timestamp   = Sys.time(),
    R_version   = R.version.string,
    fs_version  = tryCatch(
      as.character(utils::packageVersion("forestsearch")),
      error = function(e) "unknown"
    ),
    timings     = timings,

    # -- Failure accounting ----------------------------------------------------
    n_failed_alt  = n_failed_alt,
    n_failed_null = n_failed_null
  )

  filename <- sprintf(
    "actg175_binary_benefit_simulations_rawcont_%dalts_%dnulls.rds",
    nsims_alt, nsims_null
  )
  path <- file.path(output_dir, filename)
  saveRDS(bundle, path)

  cat(sprintf("Results saved: %s\n", path))
  cat(sprintf("  Size: %.1f KB\n", file.size(path) / 1024))
  cat(sprintf("  Alt sims: %d completed (%d failed)\n",
      nsims_alt - n_failed_alt, n_failed_alt))
  cat(sprintf("  Null sims: %d completed (%d failed)\n",
      nsims_null - n_failed_null, n_failed_null))
  cat(sprintf("  OR(Q) = %.3f, OR(Qc) = %.3f\n",
      dgm_calibrated$hazard_ratios$harm_subgroup,
      dgm_calibrated$hazard_ratios$no_harm_subgroup))
} else {
  cat("save_results = FALSE; skipping save.\n")
}
```


# Results {#sec-results}

## Metrics Glossary {#sec-glossary}

Classification metrics follow the definitions in León et al.
(2024, Section 3) adapted for the benefit search.

### Detection

**Found** ($\text{any}(Q)$): proportion of simulations where
the method identifies *any* benefit subgroup
($\texttt{any.H} = 1$).

- **Under H1** (alternative): $Q$ exists with a differential
  treatment effect.  $\text{any}(Q)$ is **detection power**.
- **Under H0** (null): $Q = \emptyset$ and $Q^c$ is the ITT
  population.  $\text{any}(Q)$ is the **false positive rate
  (FPR)** for falsely declaring that a benefit subgroup exists.

### Classification Under H1

Under the alternative, the true benefit subgroup $Q$ exists
($Q = \{z_1 = 1 \cap z_2 = 1\}$, prevalence $\approx 34\%$)
with a genuinely different treatment effect.  Following
León et al. (2024, Section 3), define for estimated
subgroup $\hat{Q}$:

$$
\text{sens}(\hat{Q}) = \frac{\#\{i \in \hat{Q} \cap Q\}}{\#\{i \in Q\}},
\quad
\text{ppv}(\hat{Q}) = \frac{\#\{i \in \hat{Q} \cap Q\}}{\#\{i \in \hat{Q}\}},
$$

with analogous measures for the complement $\hat{Q}^c$.

The operating-characteristics tables report these as:

| Table column | León et al. notation | Interpretation |
|---|---|---|
| **Sen** | $\text{sens}(\hat{Q})$ | Fraction of true $Q$ captured by $\hat{Q}$ |
| **Spec** | $\text{sens}(\hat{Q}^c)$ | Fraction of true $Q^c$ correctly in $\hat{Q}^c$ |
| **PPV** | $\text{ppv}(\hat{Q})$ | Fraction of $\hat{Q}$ truly in $Q$ |
| **NPV** | $\text{ppv}(\hat{Q}^c)$ | Fraction of $\hat{Q}^c$ truly in $Q^c$ |

Classification metrics are averaged across **all** simulations
(León et al. 2024, Table 1).  When no subgroup is detected
($\hat{Q} = \emptyset$), the simulation contributes
$\text{sens}(\hat{Q}) = 0$ and $\text{ppv}(\hat{Q}) = 0$
(since no member of $Q$ is captured and no subjects are
assigned to $\hat{Q}$).  This convention is implemented
at the source in `run_simulation_analysis()`, matching
the original `fs.estimates.out()` function used in
León et al. (2024).

::: {.callout-note}
## Classification metrics with discovered cuts

Because cuts are discovered from the data in this setup, a
$\hat{Q}$ that captures "roughly the right region" but with
slightly shifted boundaries will still receive high sens/PPV.
Expect these metrics to be lower than under the discretised
(oracle-cut) setup — the gap quantifies the cost of not knowing
the true cuts in advance.
:::

### Classification Under H0

Under the null, $Q = \emptyset$ --- no subgroup with a
differential treatment effect exists.  The complement $Q^c$
is the entire ITT population.  Per León et al. (2024,
Section 3), $\text{sens}(\hat{Q})$ is **undefined** (shown
as "—" in the table) and $\text{ppv}(\hat{Q}) \equiv 0$.
For the complement, $\text{sens}(\hat{Q}^c) = |\hat{Q}^c|/N$
and $\text{ppv}(\hat{Q}^c) = 1$ since $Q^c$ = ITT.

The only operationally meaningful metric under H0 is the
**detection rate (FPR)**.

### Effect Estimates

All effect estimates are displayed on the **original** treatment
scale (OR $< 1$ = benefit from ZDV+ddI) via
`subgroup_notation = "benefit"`.

## Operating Characteristics Under H1

```{r oc-alt}
t0 <- proc.time()[3]

format_oc_results(
  results_alt,
  metrics   = c("detection", "classification"),
  digits    = 2,
  digits_hr = 2,
  title     = "ACTG175 Binary: Identification (Benefit Search, Raw Continuous)",
  subtitle  = sprintf(
    "n = %d, %d sims, OR(Q) = %.2f, OR(Qc) = %.2f (switched)",
    sim_n_sample, nsims_alt,
    dgm_calibrated$hazard_ratios$harm_subgroup,
    dgm_calibrated$hazard_ratios$no_harm_subgroup
  ),
  subgroup_notation = sg_notation
)

build_estimation_table(
  results         = results_alt,
  dgm             = dgm_calibrated,
  analysis_method = "FS",
  subtitle = sprintf(
    "ACTG175 Binary DGM (raw cont.) | n = %d, %d sims",
    sim_n_sample, nsims_alt
  ),
  font_size = 16,
  subgroup_notation = sg_notation
)

if ("GRF" %in% unique(results_alt$analysis)) {
  build_estimation_table(
    results         = results_alt,
    dgm             = dgm_calibrated,
    analysis_method = "GRF",
    subtitle = sprintf(
      "ACTG175 Binary DGM, GRF raw cont. (dmin=%.2f) | n = %d, %d sims",
      grf_dmin, sim_n_sample, nsims_alt
    ),
    font_size = 16,
    subgroup_notation = sg_notation
  )
}

timings$report_alt <- proc.time()[3] - t0
```

### Interpretation

```{r interpret-alt, results="asis"}
interpret_estimation_table(
  results_alt, dgm_calibrated,
  analysis_method = "FS",
  scenario = "alt",
  subgroup_notation = sg_notation
)
```


## Operating Characteristics Under H0

```{r oc-null}
format_oc_results(
  results_null,
  metrics  = c("detection", "classification"),
  digits   = 2,
  title    = "ACTG175 Binary: Type I Error (Null, Raw Continuous)",
  subtitle = sprintf(
    "n = %d, %d sims, OR(ITT) = %.2f (switched)",
    sim_n_sample, nsims_null,
    dgm_null$hazard_ratios$overall
  ),
  subgroup_notation = sg_notation
)
```

::: {.callout-note}
Under H0, $Q = \emptyset$.  Sen is undefined (shown as "—")
and PPV $\equiv 0$ (León et al. 2024, Section 3).
The detection rate (Found) is the false positive rate.
:::


## Classification Table

```{r classification}
build_classification_table(
  scenario_results = list(
    null = list(
      results  = results_null,
      label    = "ACTG-bin-raw",
      n_sample = sim_n_sample,
      dgm      = dgm_null,
      hypothesis = "null"
    ),
    alt = list(
      results  = results_alt,
      label    = "ACTG-bin-raw",
      n_sample = sim_n_sample,
      dgm      = dgm_calibrated,
      hypothesis = "alt"
    )
  ),
  analyses = sort(unique(c(
    unique(results_null$analysis),
    unique(results_alt$analysis)
  ))),
  digits    = 2,
  font_size = 16,
  title     = "ACTG175 Binary (Raw Continuous): Subgroup Identification and Classification",
  n_sims    = nsims_alt,
  subgroup_notation = sg_notation
)
```


# Timing Summary {#sec-timing}

```{r timing}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Step = c(names(timings)),
  Seconds = round(unlist(timings), 1),
  Minutes = round(unlist(timings) / 60, 2)
)

timing_df |>
  gt::gt() |>
  gt::tab_header(title = "Computation Timing") |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(rows = Step == "total")
  )
```

### Projection for Production Runs

```{r projection}
# Per-replicate cost (simulation loop only)
time_per_alt  <- timings$sims_alt / nsims_alt
time_per_null <- timings$sims_null / nsims_null

# prod_alt and prod_null defined in setup chunk
proj_alt  <- time_per_alt * prod_alt
proj_null <- time_per_null * prod_null
proj_other <- timings$total - timings$sims_alt - timings$sims_null
proj_total <- proj_alt + proj_null + proj_other

cat(sprintf("Current run: %d alt + %d null on %d workers\n",
    nsims_alt, nsims_null, n_workers))
cat(sprintf("  Per-replicate: %.1f sec (alt), %.1f sec (null)\n",
    time_per_alt, time_per_null))
cat(sprintf("\nProjected for %d alt + %d null:\n", prod_alt, prod_null))
cat(sprintf("  Alt sims:   %5.0f sec (%4.1f min)\n", proj_alt, proj_alt / 60))
cat(sprintf("  Null sims:  %5.0f sec (%4.1f min)\n", proj_null, proj_null / 60))
cat(sprintf("  Other:      %5.0f sec (%4.1f min)\n", proj_other, proj_other / 60))
cat(sprintf("  Total:      %5.0f sec (%4.1f min = %.1f hours)\n",
    proj_total, proj_total / 60, proj_total / 3600))
```


# Discussion {#sec-discussion}

## Key Differences from the Survival Simulation

This document mirrors the structure of `actg175_simulations_benefit.qmd`
(survival endpoint) with the following differences:

1. **No censoring**: the binary outcome $I(\text{cd420} > \text{cd40})$
   is fully observed, eliminating censoring model complexity.

2. **Package GLM simulation suite**: uses `generate_glm_dgm()`,
   `calibrate_glm_interaction()`, and `simulate_from_glm_dgm()`
   instead of `generate_aft_dgm_flex()`, manual grid search, and
   `simulate_from_dgm()`.  `run_simulation_analysis()` dispatches
   automatically based on the `"glm_dgm"` class.

3. **OR replaces HR**: thresholds and reporting use odds ratios on the
   switched scale.  The `subgroup_notation = "benefit"` inversion
   applies identically.

4. **No CDE/AHR estimands**: these are survival-specific.  The binary
   DGM stores OR-scale effects in the `hazard_ratios` list for
   reporting compatibility.

5. **GRF standalone included**: the standalone GRF analysis
   (`grf.subg.harm.glm`) uses `causal_forest()` for candidate
   factor screening and policy trees for subgroup identification.

## Key Difference from the Prior Discretised Setup

The prior version of this document discretised every continuous
covariate into a binary factor (using the same cuts that define
the true subgroup $Q$) before the analysis step.  That setup
implicitly gives the analysis methods knowledge of the true
cutpoints.  **This document feeds raw continuous covariates
(`age`, `preanti`, `wtkg`, `karnof`, `cd40`, `cd80`) to the
analysis methods**, so the search methods must discover cuts
themselves.  The DGM construction is unchanged: the true $Q$ is
still $\{\text{age} > 34\} \cap \{\text{preanti} \leq 744.5\}$
and the calibration targets the same OR(Q).

**Expected consequences** relative to the discretised run:

- Lower sensitivity and PPV for both FS and GRF, because
  discovered cuts will generally differ from the true cuts.
- Similar or slightly higher FPR for GRF, because the policy-tree
  search space is larger over continuous inputs.
- Detection power depends on how closely the median / GRF-derived
  cuts align with the true cuts.

## Limitations

- The binary DGM is based on a logistic regression fitted to
  the observed ACTG175 data.  The super-population preserves
  the empirical covariate distribution but assumes the logistic
  model is correctly specified.

- `k_inter` is a direct additive shift on the log-odds scale
  for Q members under treatment.  Positive values increase the
  outcome for Q; negative values decrease it.

- `dmin.grf` was calibrated on the discretised setup; a fresh
  calibration run with raw continuous inputs is recommended
  before production scaling.

## Reloading Saved Results

When `save_results = TRUE`, the simulation bundle is saved as an
`.rds` file in the `_data/` directory.  To reload for downstream
analysis:

```r
bundle <- readRDS(
  "../_data/actg175_binary_benefit_simulations_rawcont_100alts_100nulls.rds"
)

# Core results (data.tables)
results_alt  <- bundle$results_alt
results_null <- bundle$results_null

# DGM objects (class "glm_dgm")
dgm_cal  <- bundle$dgm_calibrated
dgm_null <- bundle$dgm_null

# True effects
bundle$true_effects$or_Q    # OR(Q) on switched scale
bundle$true_effects$k_inter # calibrated interaction shift

# Analysis candidate set used (for provenance)
bundle$config$confounders_analysis

# Reproduce tables
format_oc_results(results_alt, ...)
build_estimation_table(results_alt, dgm = dgm_cal, ...)
```


# Summary {#sec-summary}

```{r summary-compute}
#| code-fold: true
#| code-summary: "Compute summary metrics"

mc_se <- function(p, n) sqrt(p * (1 - p) / n)
methods <- sort(unique(results_alt$analysis))

summ <- data.frame(Method = methods, stringsAsFactors = FALSE)
for (m in methods) {
  ra <- results_alt[results_alt$analysis == m, ]
  rn <- results_null[results_null$analysis == m, ]

  summ$alt_det[summ$Method == m]   <- mean(ra$any.H, na.rm = TRUE)
  summ$null_fpr[summ$Method == m]  <- mean(rn$any.H, na.rm = TRUE)
  summ$alt_sens[summ$Method == m]  <- mean(ra$sens, na.rm = TRUE)
  summ$alt_ppv[summ$Method == m]   <- mean(ra$ppv, na.rm = TRUE)
  summ$se_det[summ$Method == m]    <- mc_se(
    summ$alt_det[summ$Method == m], nrow(ra))
  summ$se_fpr[summ$Method == m]    <- mc_se(
    summ$null_fpr[summ$Method == m], nrow(rn))
}
```

## FS vs GRF Classification Concordance Across Simulations

Per-subject classification concordance between ForestSearch and GRF
is computed within each simulation replicate by
`run_simulation_analysis()` and stored as `agree.FS.GRF` and
`kappa.FS.GRF` columns.  When a method does not detect a subgroup,
it classifies all subjects to $Q^c$.

```{r sim-concordance}
#| code-fold: true
#| code-summary: "Classification concordance"

conc_col <- "agree.FS.GRF"
kappa_col <- "kappa.FS.GRF"

if (all(c(conc_col, kappa_col) %in% names(results_alt)) &&
    all(c("FS", "GRF") %in% methods)) {

  # One row per sim (take FS rows — concordance is per-sim, same on both)
  fs_alt <- results_alt[results_alt$analysis == "FS", ]
  grf_alt <- results_alt[results_alt$analysis == "GRF", ]

  cat(sprintf("=== FS vs GRF Classification Concordance (%d sims) ===\n\n",
      nrow(fs_alt)))
  cat(sprintf("  Mean agreement:  %.1f%% (SD: %.1f%%)\n",
      100 * mean(fs_alt[[conc_col]], na.rm = TRUE),
      100 * sd(fs_alt[[conc_col]], na.rm = TRUE)))
  cat(sprintf("  Mean kappa:      %.3f (SD: %.3f)\n",
      mean(fs_alt[[kappa_col]], na.rm = TRUE),
      sd(fs_alt[[kappa_col]], na.rm = TRUE)))
  cat(sprintf("  Median kappa:    %.3f\n\n",
      median(fs_alt[[kappa_col]], na.rm = TRUE)))

  # Detection patterns
  both   <- sum(fs_alt$any.H == 1 & grf_alt$any.H == 1)
  fs_only <- sum(fs_alt$any.H == 1 & grf_alt$any.H == 0)
  grf_only <- sum(fs_alt$any.H == 0 & grf_alt$any.H == 1)
  neither <- sum(fs_alt$any.H == 0 & grf_alt$any.H == 0)
  n <- nrow(fs_alt)

  cat("Detection patterns:\n")
  cat(sprintf("  Both detect:    %3d (%.0f%%)\n", both, 100 * both / n))
  cat(sprintf("  FS only:        %3d (%.0f%%)\n", fs_only, 100 * fs_only / n))
  cat(sprintf("  GRF only:       %3d (%.0f%%)\n", grf_only, 100 * grf_only / n))
  cat(sprintf("  Neither:        %3d (%.0f%%)\n", neither, 100 * neither / n))

  if (both > 0) {
    both_idx <- fs_alt$any.H == 1 & grf_alt$any.H == 1
    cat(sprintf("\nWhen both detect (n=%d):\n", both))
    cat(sprintf("  Mean agreement: %.1f%%\n",
        100 * mean(fs_alt[[conc_col]][both_idx], na.rm = TRUE)))
    cat(sprintf("  Mean kappa:     %.3f\n",
        mean(fs_alt[[kappa_col]][both_idx], na.rm = TRUE)))
  }
} else {
  cat("Concordance columns not found in results.\n")
  cat("Ensure run_simulation_analysis() was called with run_fs=TRUE and run_grf=TRUE.\n")
}
```

## Key Findings

```{r summary-findings}
#| results: asis

cat(sprintf("**Simulation budget**: %d alt + %d null sims.\n\n", nsims_alt, nsims_null))
cat("**Analysis setting**: raw continuous covariates; cuts discovered by the methods.\n\n")

for (m in methods) {
  s <- summ[summ$Method == m, ]
  cat(sprintf("- **%s**: Detection = %.1f%%, FPR = %.1f%%, sens(%s) = %.2f, ppv(%s) = %.2f.\n",
      m, 100 * s$alt_det, 100 * s$null_fpr,
      ifelse(sg_notation == "benefit", "Q", "H"), s$alt_sens,
      ifelse(sg_notation == "benefit", "Q", "H"), s$alt_ppv))
}

time_per_alt  <- timings$sims_alt / nsims_alt
time_per_null <- timings$sims_null / nsims_null
cat(sprintf("\n- **Timing**: %.1f sec/sim (alt), %.1f sec/sim (null) on %d workers.\n",
    time_per_alt, time_per_null, n_workers))
cat(sprintf("- **Total runtime**: %.1f minutes.\n",
    as.numeric(timings$total) / 60))
cat(sprintf("- **Projected for %d alt + %d null**: %.1f hours.\n",
    prod_alt, prod_null,
    (prod_alt * time_per_alt + prod_null * time_per_null) / 3600))
```


# Session Info {#sec-session}

```{r session-info}
sessionInfo()
```


# References {#sec-references}

- Hammer, S.M., et al. (1996). *NEJM*, 335, 1081--1090.
- Leon, L.F., et al. (2024). *Statistics in Medicine*, 43,
  3921--3942.
- Tsiatis, A.A., et al. (2008). *Statistics in Medicine*, 27,
  4658--4677.
