---
title: "Classification Accuracy Metrics for Subgroup Identification"
subtitle: "Capture, reliability, set overlap, and single-number summaries"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    code-fold: false
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Purpose

This guide summarizes the metrics used to assess how well an *identified*
subgroup $\hat{H}$ recovers a *true* target subgroup $H$ (the harm subgroup),
and, symmetrically, how well its complement $\hat{H}^c$ recovers $H^c$. All
quantities are functions of a single $2 \times 2$ confusion table, so the
discussion below is organized around that table: first the four primary rates,
then the relationships between them, the set-overlap indices, and finally the
single-number omnibus summaries. A base-R reference implementation and a worked
example close the document.

# The confusion table

Take the harm subgroup $H$ as the positive class. Comparing the identified set
$\hat{H}$ against the truth $H$ partitions the sample into four cells:

| | Truly in $H$ | Truly in $H^c$ |
|:--|:--:|:--:|
| **Identified in $\hat{H}$** | $TP$ | $FP$ |
| **Identified in $\hat{H}^c$** | $FN$ | $TN$ |

with $N = TP + FP + FN + TN$. In set terms, $TP = |\hat{H} \cap H|$,
$FP = |\hat{H} \cap H^c|$, $FN = |\hat{H}^c \cap H|$, and
$TN = |\hat{H}^c \cap H^c|$.

# The four primary rates

The four standard rates each condition on a different margin of the table.

$$
\text{Sensitivity} = \frac{TP}{TP + FN}, \qquad
\text{Specificity} = \frac{TN}{TN + FP},
$$

$$
\text{PPV} = \frac{TP}{TP + FP}, \qquad
\text{NPV} = \frac{TN}{TN + FN}.
$$

They answer two distinct questions, and the distinction is what determines which
metric belongs to which target.

| Question | Denominator | Target $H$ | Target $H^c$ |
|:--|:--|:--:|:--:|
| **Capture** (recall-type): of those *truly* in the set, what fraction did we identify? | truth | Sensitivity | Specificity |
| **Reliability** (precision-type): of those we *placed* in the set, what fraction belong? | prediction | PPV | NPV |

Reading down the columns:

- **Capture of the true harm subgroup $H$** is *sensitivity* — truth in the
  denominator. This is the quantity meant by "how accurately are the true $H$
  members captured."
- **Capture of the complement $H^c$** is *specificity*: of the true non-members,
  the fraction correctly kept out of $\hat{H}$.
- **Reliability of the identified set $\hat{H}$** is *PPV*: of those we flag as
  harm, the fraction truly harmed.
- **Reliability of the identified complement $\hat{H}^c$** is *NPV*.

The common pitfall is to pair one capture-type rate with one reliability-type
rate across the two targets (for example, specificity for $H$ and NPV for
$H^c$). That combination describes only the complement and mixes the two
framings; it does not report the capture of true $H$ at all.

# Symmetry between $H$ and $H^c$

Relabeling $H^c$ as the positive class swaps the cells
$TP \leftrightarrow TN$ and $FP \leftrightarrow FN$, which gives the identities

$$
\text{sens}_{H^c} = \text{spec}_{H}, \qquad
\text{spec}_{H^c} = \text{sens}_{H},
$$

$$
\text{PPV}_{H^c} = \text{NPV}_{H}, \qquad
\text{NPV}_{H^c} = \text{PPV}_{H}.
$$

Consequently, reporting sensitivity *and* specificity already reports the
capture rate for *each* target, and reporting PPV *and* NPV already reports the
reliability of *each* identified label. A four-metric panel
(sensitivity, specificity, PPV, NPV) is therefore complete for both targets
under both framings; only the labeling needs to make the question explicit.

# Set-overlap indices: Jaccard and $F_1$

The Jaccard index measures overlap of the *positive* sets,

$$
J = \frac{|\hat{H} \cap H|}{|\hat{H} \cup H|}
  = \frac{TP}{TP + FP + FN},
$$

equivalent to the "intersection over union." It is a function of sensitivity and
PPV only and is monotonically related to the $F_1$ score,

$$
F_1 = \frac{2\,TP}{2\,TP + FP + FN}, \qquad
J = \frac{F_1}{2 - F_1}, \qquad
F_1 = \frac{2J}{1 + J}.
$$

The defining feature is that **$TN$ never appears**: Jaccard (and $F_1$) ignore
agreement on the negatives by construction. This is precisely why an overlap
index is attractive when the complement is large and $TN$ dominates the table —
but it is also why neither index can summarize specificity or NPV. Jaccard is a
recognized and useful *positive-set overlap* statistic; it is **not** an
omnibus across sensitivity, specificity, PPV, and NPV, and should be labeled as
the former.

# Single-number summaries that use all four cells

When a single figure is wanted, the following draw on the full table.

**Balanced accuracy** averages the two capture rates:

$$
\text{Balanced accuracy} = \tfrac{1}{2}\left(\text{sens} + \text{spec}\right).
$$

**Youden's $J$ (informedness)** and **markedness** collapse the capture pair and
the reliability pair, respectively:

$$
\text{Informedness} = \text{sens} + \text{spec} - 1, \qquad
\text{Markedness} = \text{PPV} + \text{NPV} - 1.
$$

**Matthews correlation coefficient (MCC)**, the $\phi$ coefficient for the
$2 \times 2$ table, is the omnibus summary; it ranges in $[-1, 1]$ and is well
behaved under class imbalance:

$$
\text{MCC} =
\frac{TP \cdot TN - FP \cdot FN}
     {\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}}.
$$

MCC is the (signed) geometric mean of informedness and markedness,

$$
\text{MCC}^2 = \text{Informedness} \times \text{Markedness},
$$

so it is the single figure that draws on all four of the primary rates at once.

**Cohen's $\kappa$** is chance-corrected agreement,

$$
\kappa = \frac{p_o - p_e}{1 - p_e}, \quad
p_o = \frac{TP + TN}{N}, \quad
p_e = \frac{(TP+FP)(TP+FN) + (FN+TN)(FP+TN)}{N^2}.
$$

# Reference implementation

The function below computes every metric above from the four cell counts, with
$H$ taken as the positive class. Zero denominators return `NA_real_` rather than
`NaN`.

```{r}
#| label: classification-metrics
classification_metrics <- function(tp, fp, fn, tn) {
  safe_ratio <- function(num, den) if (den > 0) num / den else NA_real_

  sens <- safe_ratio(tp, tp + fn)        # capture of true H
  spec <- safe_ratio(tn, tn + fp)        # capture of true H^c
  ppv  <- safe_ratio(tp, tp + fp)        # reliability of identified H
  npv  <- safe_ratio(tn, tn + fn)        # reliability of identified H^c

  jaccard <- safe_ratio(tp, tp + fp + fn)
  f1      <- safe_ratio(2 * tp, 2 * tp + fp + fn)

  informedness <- sens + spec - 1
  markedness   <- ppv + npv - 1
  bal_accuracy <- (sens + spec) / 2

  mcc_den <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
  mcc     <- if (mcc_den > 0) (tp * tn - fp * fn) / mcc_den else NA_real_

  n     <- tp + fp + fn + tn
  p_o   <- (tp + tn) / n
  p_e   <- ((tp + fp) * (tp + fn) + (fn + tn) * (fp + tn)) / n^2
  kappa <- safe_ratio(p_o - p_e, 1 - p_e)

  list(
    sensitivity  = sens,
    specificity  = spec,
    ppv          = ppv,
    npv          = npv,
    jaccard      = jaccard,
    f1           = f1,
    informedness = informedness,
    markedness   = markedness,
    bal_accuracy = bal_accuracy,
    mcc          = mcc,
    kappa        = kappa
  )
}
```

# Worked example

Consider a sample of $N = 1000$ with $200$ true harm-subgroup members. Suppose
$\hat{H}$ captures $150$ of them ($TP = 150$, $FN = 50$) and additionally flags
$80$ non-members ($FP = 80$, $TN = 720$).

```{r}
#| label: worked-example
m <- classification_metrics(tp = 150, fp = 80, fn = 50, tn = 720)

results <- data.frame(
  metric = names(m),
  value  = round(unlist(m), 4),
  row.names = NULL
)
knitr::kable(results, caption = "Metrics for the worked example.")
```

The two algebraic identities hold numerically:

```{r}
#| label: identity-checks
c(
  jaccard_from_f1 = m$f1 / (2 - m$f1),       # should equal m$jaccard
  jaccard_direct  = m$jaccard,
  mcc_squared     = m$mcc^2,                  # should equal informedness * markedness
  inf_times_mark  = m$informedness * m$markedness
)
```

# Recommended reporting

For comparing an identified subgroup against a known truth:

- Report the **four primary rates**, each labeled by the question it answers:
  sensitivity (capture of true $H$), specificity (capture of true $H^c$),
  PPV (reliability of $\hat{H}$), and NPV (reliability of $\hat{H}^c$). By the
  symmetry above, this panel covers both targets under both framings.
- If a **single summary** is wanted, prefer **MCC**, since it uses all four
  cells and behaves well under the class imbalance typical of subgroup
  identification (a large complement). Balanced accuracy, informedness, and
  markedness are interpretable partial summaries.
- Treat **Jaccard** ($F_1$) as a positive-set overlap statistic and label it as
  such; it deliberately ignores $TN$ and is not a substitute for specificity or
  NPV.

# References (canonical sources)

- Matthews, B. W. (1975). Comparison of the predicted and observed secondary
  structure of T4 phage lysozyme. *Biochimica et Biophysica Acta*.
- Youden, W. J. (1950). Index for rating diagnostic tests. *Cancer*.
- Cohen, J. (1960). A coefficient of agreement for nominal scales.
  *Educational and Psychological Measurement*.
- Powers, D. M. W. (2011). Evaluation: from precision, recall and F-measure to
  ROC, informedness, markedness and correlation. *Journal of Machine Learning
  Technologies*.
- Jaccard, P. (1912). The distribution of the flora in the alpine zone.
  *New Phytologist*.
