---
title: "Collapsing Near-Redundant Continuous Cuts: Synthetic ACTG175 Demo"
format: html
---

```{r setup}
#| include: false

# ===========================================================================
#  PARAMETERS  --  single source of truth for this demo.
# ===========================================================================

## -- Synthetic data ----------------------------------------------------------
demo_n    <- 1000L          # synthetic sample size (band scales as sd/sqrt(n))
demo_seed <- 8316951L

## -- Coarsening knobs to demonstrate -----------------------------------------
# band = c * sd(x) / sqrt(n);  representative = round-half-up(mean(cluster), digits);
# a cluster collapses only if <= tol subjects change membership (tol < 1 = fraction
# of n, tol >= 1 = absolute count).  These are the collapse_cuts_args defaults plus
# a small sweep on c so the band's effect is visible.
collapse_default <- list(c = 1.0, tol = 0.05, digits = 0L)
collapse_sweep_c <- c(1.0, 2.0, 3.0)

library(forestsearch)

# Internal helper (not exported); triple-colon per package convention.
collapse_fn <- forestsearch:::collapse_redundant_cuts
```

This demo shows how the opt-in cut-coarsening behaves on a candidate pool that
mimics the ACTG175 binary-harm run. Coarsening is governed by
`collapse_cuts = TRUE` and `collapse_cuts_args` in `forestsearch()` /
`get_FSdata()`; the engine is `forestsearch:::collapse_redundant_cuts()`.

# Synthetic ACTG175-like covariates

The synthetic generator reproduces the *marginal scale* of the real covariates
(means / sds), which is all the coarsening band depends on: `band = c · sd(x) / √n`.
Binary indicators get two levels (so they are never coarsened).

```{r synth-data}
set.seed(demo_seed)
n <- demo_n

synth_actg <- data.frame(
  # -- continuous (coarsenable; >= 4 unique values) --------------------------
  age     = round(rnorm(n, 35.0,   8.7)),
  wtkg    = round(rnorm(n, 75.0,  13.3), 1),
  cd40    = round(pmax(rnorm(n, 350,  120), 0)),
  cd80    = round(pmax(rnorm(n, 986,  480), 0)),
  karnof  = sample(c(70, 80, 90, 100), n, replace = TRUE,
                   prob = c(.03, .07, .45, .45)),
  preanti = round(pmax(rnorm(n, 380,  470), 0)),
  # -- binary indicators (pass through untouched) ---------------------------
  hemo = rbinom(n, 1, .08), homo = rbinom(n, 1, .66),
  drugs = rbinom(n, 1, .13), race = rbinom(n, 1, .29),
  gender = rbinom(n, 1, .82), symptom = rbinom(n, 1, .17),
  ar_naive = rbinom(n, 1, .54), prior_6mo = rbinom(n, 1, .50)
)

knitr::kable(
  data.frame(
    variable = c("age", "wtkg", "cd40", "cd80", "karnof", "preanti"),
    sd       = round(sapply(synth_actg[c("age","wtkg","cd40","cd80","karnof","preanti")],
                            sd), 2),
    `band (c=1)` = round(sapply(synth_actg[c("age","wtkg","cd40","cd80","karnof","preanti")],
                                function(x) sd(x) / sqrt(n)), 3),
    check.names = FALSE
  ),
  caption = "Per-variable standard-error band at c = 1 (the merge resolution)."
)
```

# The candidate pool from the attached run

These are the exact 44 candidate factors printed by the ACTG175 run, including
the J-quantile `cd40`/`preanti` cuts, the full-precision GRF/DGM cuts
(`wtkg <= 79.8336`), the DINA cut (`wtkg >= 87.2`), and the bare categorical
factors.

```{r candidate-pool}
candidate_cuts <- c(
  "cd40 <= 320",    "wtkg <= 79.8336", "wtkg >= 87.2",   "cd40 <= 202.8",
  "cd40 <= 239",    "cd40 <= 269.5",   "cd40 <= 295",    "cd40 <= 321",
  "cd40 <= 353.9",  "cd40 <= 384",     "cd40 <= 415",    "cd40 <= 450",
  "cd40 <= 504",    "preanti <= 0",    "preanti <= 39",  "preanti <= 203.9",
  "preanti <= 406", "preanti <= 676",  "preanti <= 872", "preanti <= 1049",
  "age <= 35.7",    "age <= 35",       "age <= 30",      "age <= 40",
  "wtkg <= 75.2",   "wtkg <= 74.5",    "wtkg <= 66.7",   "wtkg <= 82.6",
  "karnof <= 95.1", "karnof <= 90",    "cd80 <= 956.1",  "cd80 <= 895.5",
  "cd80 <= 639",    "cd80 <= 1170",    "hemo",           "homo",
  "drugs",          "race",            "gender",         "symptom",
  "ar_naive",       "prior_6mo",       "ar_naive == 1"
)
length(candidate_cuts)
```

# Coarsening at the default (`c = 1`, `tol = 0.05`, `digits = 0`)

```{r collapse-default}
collapsed <- collapse_fn(
  cuts             = candidate_cuts,
  df               = synth_actg,
  c_band           = collapse_default$c,
  safety_tol       = collapse_default$tol,
  digits           = collapse_default$digits,
  details          = TRUE   # prints the per-cluster collapse report
)
```

```{r collapse-default-side-by-side}
maxlen <- max(length(candidate_cuts), length(collapsed))
pad    <- function(x, k) c(x, rep("", k - length(x)))
knitr::kable(
  data.frame(
    `before (44)`            = pad(candidate_cuts, maxlen),
    `after  (collapsed)`     = pad(collapsed,       maxlen),
    check.names = FALSE
  ),
  caption = sprintf("Before vs after: %d -> %d candidate factors.",
                    length(candidate_cuts), length(collapsed))
)
```

Two things are happening, and the attached examples exercise both:

- **Rounding then de-duplication.** `wtkg <= 75.2` and `wtkg <= 74.5` are more
  than one standard-error apart, so the band does *not* cluster them — but at
  `digits = 0` both round to `wtkg <= 75`, and the duplicate is dropped. Full
  precision cuts are also cleaned up (`wtkg <= 79.8336 -> wtkg <= 80`,
  `cd40 <= 202.8 -> cd40 <= 203`, `karnof <= 95.1 -> karnof <= 95`).
- **Band clustering.** `cd40 <= 320` and `cd40 <= 321` sit inside the `cd40`
  band (~3-4 cells at `c = 1`), so they cluster and collapse to a single
  rounded-centroid threshold.

Binary indicators (`gender`, `hemo`, …) and equality tests (`ar_naive == 1`)
are never coarsened.

# Effect of the band multiplier `c`

`age <= 35.7` and `age <= 35` round to *different* integers (36 and 35), so
rounding alone leaves them distinct. They only merge once `c` widens the band
past their 0.7-year gap.

```{r collapse-sweep}
sweep <- lapply(collapse_sweep_c, function(cc) {
  collapse_fn(candidate_cuts, synth_actg,
              c_band = cc, safety_tol = collapse_default$tol,
              digits = collapse_default$digits, details = FALSE)
})
knitr::kable(
  data.frame(
    c                 = collapse_sweep_c,
    `n_candidates`    = vapply(sweep, length, integer(1)),
    `age cuts kept`   = vapply(sweep, function(s)
      paste(grep("^age", s, value = TRUE), collapse = "; "), character(1)),
    check.names = FALSE
  ),
  caption = "Raising c merges the age pair (35.7, 35) once the band exceeds their gap."
)
```

# Integration check through `get_FSdata()`

The same coarsening runs inside the package when `collapse_cuts = TRUE`. Here we
drive `get_FSdata()` directly with J-quantile cuts plus supplied GRF/DINA cuts
(no LASSO/GRF fitting, so the candidate pool is deterministic), and compare the
returned candidate factors with coarsening off vs on.

```{r end-to-end}
#| warning: false
synth_actg$y_sim <- rbinom(n, 1, 0.3)

covs <- c("age", "wtkg", "cd40", "cd80", "karnof", "preanti",
          "hemo", "homo", "drugs", "race", "gender", "symptom",
          "ar_naive", "prior_6mo")

# get_FSdata() builds the candidate-cut pool; it takes outcome/event names but
# NOT a treatment argument (treatment enters later, in forestsearch()).
fs_args <- list(
  df.analysis      = synth_actg,
  confounders.name = covs,
  outcome.name     = "y_sim", event.name = "y_sim",
  outcome_type     = "binary",
  use_lasso        = FALSE,
  use_grf          = TRUE,
  grf_cuts         = c("cd40 <= 320", "gender <= 0", "wtkg <= 79.8336"),
  dina_cuts        = "wtkg >= 87.2",
  conf.cont_jcuts  = list(cd40 = 10, preanti = 10),
  details          = FALSE
)

fs_off <- do.call(get_FSdata, c(fs_args, list(collapse_cuts = FALSE)))
fs_on  <- do.call(get_FSdata,
                  c(fs_args, list(collapse_cuts = TRUE,
                                  collapse_cuts_args = collapse_default)))

cat(sprintf("Candidate factors  off: %d   on: %d\n",
            length(fs_off$confs), length(fs_on$confs)))
```

```{r end-to-end-tables}
maxlen <- max(length(fs_off$confs), length(fs_on$confs))
knitr::kable(
  data.frame(
    `collapse_cuts = FALSE` = pad(sort(fs_off$confs), maxlen),
    `collapse_cuts = TRUE`  = pad(sort(fs_on$confs),  maxlen),
    check.names = FALSE
  ),
  caption = "get_FSdata() candidate factors: coarsening off vs on."
)
```
