---
title: "ForestSearch Bootstrap De-biasing: A Code-to-Theory Map"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
    theme: cosmo
vignette: >
  %\VignetteIndexEntry{ForestSearch Bootstrap De-biasing: A Code-to-Theory Map}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false
# This is an exposition article: every code block is an illustrative excerpt
# of package internals shown verbatim, not executed.  Internal (unexported)
# helpers are referenced with the package::: convention.
knitr::opts_chunk$set(eval = FALSE, comment = "#>")
```

This article maps the bootstrap selection-bias correction implemented in
`forestsearch_bootstrap_dofuture()` onto the two papers that define it, so that
each line of the de-biasing code can be read next to the equation it computes.

Two papers are involved, and they describe **the same estimator** by two routes:

- **León et al. (2024), *Statistics in Medicine*** (the "SIM" paper) defines the
  bootstrap bias-corrected Cox estimator with an infinitesimal-jackknife (IJ)
  interval — the *full bootstrap* that re-runs the entire enumerate--screen--select
  search on every resample. `forestsearch_bootstrap_dofuture()` is the literal
  implementation of its Eqs. (7)--(9). In the package vignettes this is **Tier-1**.

- **León & Anderson (2026)** (the "JASA" paper) shows that on a *fixed* candidate
  family the identical correction is obtained refit-free, by a multiplier
  resampling of the subgroup estimating function, reproducing the full bootstrap's
  point estimate **and** interval to first order. That is `fs_debias_gate()` — the
  **Tier-2** de-biased gate.

The point of this note is the correspondence: which code symbol is which paper
symbol, and which line computes which equation.

# The target of inference and the winner's curse

The reportable quantity is the **standard within-subgroup analysis**: the
treatment-only Cox log-hazard ratio (or, on the GLM path, the log odds ratio /
log incidence-rate ratio / mean difference) fit *inside* the identified subgroup
$\hat H$ — the effect the trial would report had $\hat H$ been prespecified
(JASA §1.2). Writing $\hat\beta(\hat H)$ for that estimate, selection by
maximizing over the candidate family makes it optimistic: $\hat H$ was chosen for
the extremity of the very quantity reported (the winner's curse, SIM §3.2,
JASA §1.1). The correction removes that optimism on the reporting scale and
attaches an interval that carries the variability of the selection itself.

All arithmetic is performed on the model's natural-parameter scale — the log
scale for ratio measures (HR, OR, RR, IRR), the identity scale for differences
(RD, IRD, MD) — and exponentiated at the end for ratio measures. In the code this
is the `est.loghr` flag threaded into `ci_est()`.

# The two-term correction

## The two discrepancies

For resample $b$, let $\hat\beta^{*}_{b}(g)$ be the within-subgroup estimate of a
*fixed* definition $g$ refit on the resample, and define the bootstrap-minus-observed
discrepancy

$$
\eta^{*}_{b}(g) \;=\; \hat\beta^{*}_{b}(g) - \hat\beta(g).
$$

The SIM correction (Eq. 7) subtracts the average of **two** such discrepancies,
evaluated at two different subgroups:

$$
\hat\beta^{*}(\hat H)
\;=\;
\hat\beta(\hat H)
\;-\;
\frac{1}{B}\sum_{b=1}^{B}\Bigl(\underbrace{\eta^{*}_{b}(\hat H^{*}_{b})}_{\text{selection optimism}}
\;+\;
\underbrace{\eta^{*}_{b}(\hat H)}_{\text{fixed-subgroup discrepancy}}\Bigr).
$$

- $\eta^{*}_{b}(\hat H^{*}_{b})$ is taken at the **re-selected** subgroup of each
  resample. Because the rule rewards extreme perturbed effects, this is positive
  on average — the selection optimism, the term in the lineage of Harrell et al.
  (1996). The JASA paper calls its average $\widehat{\text{bias}}_{\mathrm{sel}}$ (Eq. 9).
- $\eta^{*}_{b}(\hat H)$ is taken at the **observed** selected subgroup. It is
  mean-zero under centered resampling weights, but it is retained: it is the
  estimator's own finite-sample discrepancy at $\hat H$, the term Harrell drops,
  and — crucially — it supplies the per-draw residual the IJ variance is built
  from (JASA §3.1). Its average is $\widehat{\text{bias}}_{\mathrm{fix}}$.

In JASA notation the de-biased estimate is

$$
\tilde\beta(\hat H) \;=\; \hat\beta(\hat H) - \widehat{\text{bias}}_{\mathrm{sel}} - \widehat{\text{bias}}_{\mathrm{fix}}
\qquad\text{(JASA Eq. 10)},
$$

and substituting the dfbeta linearization (next section) turns SIM Eq. (7) into
this identically.

## The four per-resample fits in code

Each bootstrap iteration computes four within-subgroup effects. The mapping to
the paper symbols is exact (`bootstrap_analysis_dofuture.R`):

| Code symbol | Subgroup | Data | Paper symbol |
|---|---|---|---|
| `H_obs`     | observed $\hat H$        | original  | $\hat\beta(\hat H)$ |
| `H_star`    | observed $\hat H$        | bootstrap | $\hat\beta^{*}_{b}(\hat H)$ |
| `Hstar_obs` | re-selected $\hat H^{*}_{b}$ | original  | $\hat\beta(\hat H^{*}_{b})$ |
| `Hstar_star`| re-selected $\hat H^{*}_{b}$ | bootstrap | $\hat\beta^{*}_{b}(\hat H^{*}_{b})$ |

`H_obs` is computed once on the original data (it is the naive estimate); the
other three are computed inside the worker for each resample. The membership-based
evaluation — subsetting on `treat.recommend` carried from the *original* fit
(`H_star`) versus the *bootstrap* fit (`Hstar_*`) — is what cleanly separates the
fixed-subgroup term from the selection term:

```{r}
#| label: four-fits
# bootstrap_analysis_dofuture.R (worker body)

# observed subgroup H on the bootstrap data  ->  beta*_b(H)
df_H   <- subset(df_boot, treat.recommend == 0)
H_star <- fit_subgroup_effect(df_H, estimator_fn)$est_obs        # (GLM)  / get_Cox_sg(...) (survival)

# re-selected subgroup H*_b on the ORIGINAL data  ->  beta(H*_b)
df_Hstar  <- subset(run_bootstrap$df.predict, treat.recommend == 0)
Hstar_obs <- fit_subgroup_effect(df_Hstar, estimator_fn)$est_obs

# re-selected subgroup H*_b on the BOOTSTRAP data  ->  beta*_b(H*_b)
Hstar_star <- fit_subgroup_effect(
  subset(run_bootstrap$df.est, treat.recommend == 0), estimator_fn)$est_obs
```

Hence, term by term,
$\eta^{*}_{b}(\hat H^{*}_{b}) = \texttt{Hstar\_star} - \texttt{Hstar\_obs}$ and
$\eta^{*}_{b}(\hat H) = \texttt{H\_star} - \texttt{H\_obs}$.

## Harrell (H1) versus the two-term León estimate (H2)

The code stores two per-draw corrections:

```{r}
#| label: bias-formulas
# bootstrap_analysis_dofuture.R

# Method 1 (Harrell): selection optimism only
H_biasadj_1 <- H_obs - (Hstar_star - Hstar_obs)

# Method 2 (Leon, two-term): selection optimism + fixed-subgroup discrepancy
H_biasadj_2 <- 2 * H_obs - (H_star + Hstar_star - Hstar_obs)
```

These are per-draw *summands*; the estimate is their mean, taken later in
`get_targetEst()`. Expanding `H_biasadj_2` shows it is exactly SIM Eq. (7):

$$
\overline{\texttt{H\_biasadj\_2}}
= \hat\beta(\hat H)
- \tfrac1B\!\sum_b \eta^{*}_{b}(\hat H)
- \tfrac1B\!\sum_b \eta^{*}_{b}(\hat H^{*}_{b})
= \hat\beta(\hat H) - \widehat{\text{bias}}_{\mathrm{fix}} - \widehat{\text{bias}}_{\mathrm{sel}}
= \tilde\beta(\hat H).
$$

`H_biasadj_1` drops the $\eta^{*}_{b}(\hat H)$ term, giving the Harrell point
estimate. **H2 is the headline bias-corrected effect** — the `H2` / `OR*` / `HR*`
column in every table and the quantity the Tier-2 gate reproduces; H1 is a Harrell
comparator only. The complement $\hat H^{c}$ is corrected identically from
`Hc_obs`, `Hc_star`, `Hcstar_obs`, `Hcstar_star`.

# The infinitesimal-jackknife interval

The de-biased estimate is a smooth average over draws, so its variance is the IJ
of a bagged estimator (Efron 2014; Wager, Hastie & Efron 2014). Write the
per-draw residual as the centered total bias removed,

$$
r_{b} \;=\; \bigl(\widehat{\text{bias}}_{\mathrm{sel}} + \widehat{\text{bias}}_{\mathrm{fix}}\bigr)
- D_{\hat H^{*}_{b}}(b) - D_{\hat H}(b)
\qquad\text{(JASA Eq. 12)},
$$

which equals the centered H2 summand $\texttt{H\_biasadj\_2}_b - \tilde\beta(\hat H)$.
The IJ variance and its finite-$B$ correction are

$$
\tilde V = \sum_{i=1}^{N} \widetilde{\mathrm{cov}}_{i}^{\,2},
\qquad
\widetilde{\mathrm{cov}}_{i} = \frac{1}{B}\sum_{b=1}^{B}\bigl(K^{*}_{bi} - \bar K^{*}_{i}\bigr) r_{b},
\qquad
\hat V = \tilde V - \frac{N}{B}\,\tilde\sigma_{B}^{2}
$$

(SIM Eq. 9; JASA Eqs. 13--14), where $K^{*}_{bi}$ is the multiplicity of subject
$i$ in resample $b$ and $\tilde\sigma_{B}^{2} = B^{-1}\sum_b r_b^{2}$. The
multiplicity matrix $\{K^{*}_{bi}\}$ is exactly `Ystar_mat`, generated once on the
main process from the same indices that drive the resamples — so the IJ is read
from the very draws that produced the point estimate.

`get_targetEst()` (`bootstrap_calculations_helpers.R`) is this verbatim:

```{r}
#| label: ij-variance
# bootstrap_calculations_helpers.R

calc_cov <- function(x, Est) mean((x - mean(x, na.rm = TRUE)) * Est, na.rm = TRUE)

get_targetEst <- function(x, ystar, cov_method = "standard", cov_trim = 0.0) {
  mx <- mean(x, na.rm = TRUE)      #  = tilde beta(H)   (the bias-corrected estimate)
  xc <- c(x - mx)                  #  = r_b             (per-draw residual, JASA Eq. 12)
  N  <- ncol(ystar)

  cov_i  <- apply(ystar, 2, calc_cov, Est = xc)   #  = cov_tilde_i   (per-subject IJ influence)
  varhat <- N * mean(cov_i^2, na.rm = TRUE, trim = cov_trim)   #  = V_tilde = sum_i cov_i^2

  B.eval <- sum(!is.na(xc))
  termc  <- (N / B.eval^2) * B.eval * mean(xc^2, na.rm = TRUE, trim = cov_trim)  # = (N/B) sigma_B^2

  # V_hat = V_tilde - finite-B correction, with a degenerate-variance fallback
  if (is.na(varhat) || is.na(termc) || varhat <= termc) {
    varhat_new <- varhat; seH_new <- sqrt(varhat)
  } else {
    varhat_new <- varhat - termc; seH_new <- sqrt(varhat_new)
  }
  list(target_est = mx, sehat_new = seH_new, varhat = varhat_new)
}
```

Line by line: `mx` is $\tilde\beta(\hat H)$; `xc` is $r_b$; `calc_cov(ystar[, i], xc)`
is $\widetilde{\mathrm{cov}}_i$; `varhat` is $\tilde V$ (since
`N * mean(cov_i^2) = sum_i cov_i^2`); `termc` is the $(N/B)\tilde\sigma_B^2$
correction, using the *effective* successful-draw count `B.eval` in place of
nominal $B$; and `varhat_new` is $\hat V$, with the
`varhat <= termc` branch the JASA fallback to $\sqrt{\tilde V}$ (and, failing that,
the fixed-subgroup robust SE $\sigma_{D,\hat H}$) at small $B$.

`get_dfRes()` then forms the interval via `ci_est()`:

$$
\tilde\beta(\hat H) \pm z_{1-\alpha/2}\sqrt{\hat V},
$$

exponentiated for ratio measures (JASA Eq. 15). Because $r_b$ moves with the
*re-selected* winner $\hat H^{*}_{b}$, $\hat V$ carries the variability of which
subgroup was selected and is wider than the bare robust-SE interval that treats
$\hat H$ as prespecified — in the GBSG analysis, IJ SE $\approx 0.48$ against a
robust SE $\approx 0.36$ on the log-hazard scale (JASA §3.2).

# Tier-1 and Tier-2 are the same estimator

The bridge that makes `fs_debias_gate()` reproduce all of the above without
re-running the search is JASA §2.5. On a *fixed* family, refitting candidate $g$
on resample $b$ is, to first order, a dot product of the centered resampling
weights with the candidate's per-subject influence values (dfbeta):

$$
\hat\beta^{*}_{b}(g) - \hat\beta(g)
= \sum_{i}\bigl(K^{*}_{bi} - 1\bigr)\,\mathrm{db}_{g,i} + o_{p}(n_g^{-1/2})
\;=:\; D_{g}(b)
\qquad\text{(JASA Eq. 6).}
$$

So every $\eta^{*}_{b}(g)$ the full bootstrap computes by *literal refitting*
equals $D_g(b)$, re-selection becomes the deterministic rule applied to the
perturbed effects $\{\hat\beta(g) + D_g(b)\}$, and the whole competition is one
matrix product $P = B_{\mathrm{eff}}^{\top}\Xi$ over the influence columns and the
multiplier draws. `fs_debias_gate()` assembles the same
$\widehat{\text{bias}}_{\mathrm{sel}}$, $\widehat{\text{bias}}_{\mathrm{fix}}$, and
IJ residual $r$ from $P$ — which is why its de-biased estimate and interval match
`forestsearch_bootstrap_dofuture()`'s `H2` row to first order on a fixed
(consistency) family.

The two routes diverge only by what the family is allowed to do under resampling:

- **Consistency search (fixed enumerated family).** A genuine bootstrap re-selects
  over the same family on every resample, so the gate is *first-order exact*
  against it (JASA §5). The two columns should track on point, length, and
  coverage.
- **DINA / GRF selection (data-driven family).** The full bootstrap regenerates
  the candidate family on each resample (it re-fits the forest/DINA surface); the
  gate holds that fit fixed and so omits the family-regeneration component of the
  bias and variance. The gate is then *conditional on the proposed family* — a
  fast, conservative check, with the full bootstrap as the unconditional reference.

This is exactly the Tier-1 / Tier-2 distinction the multimethod vignettes report
per engine.

# Where each piece lives in the code

| Quantity / equation | Function | File (lines) |
|---|---|---|
| Orchestration, naive fit, scale, CIs | `forestsearch_bootstrap_dofuture()` | `bootstrap_dofuture_main.R` (≈144–658) |
| Per-resample four fits + H1/H2 (SIM Eq. 7) | `bootstrap_results()` worker body | `bootstrap_analysis_dofuture.R` (≈445–704) |
| Result collection (drop failures, rbind) | `.collect_bootstrap_results()` | `bootstrap_analysis_dofuture.R` (≈28–55) |
| Mean + IJ variance + finite-B (SIM Eq. 9) | `get_targetEst()` | `bootstrap_calculations_helpers.R` (≈331–373) |
| Per-subject IJ influence $\widetilde{\mathrm{cov}}_i$ | `calc_cov()` | `bootstrap_calculations_helpers.R` (≈134–136) |
| Assemble H0/H1/H2 + CIs | `get_dfRes()`, `ci_est()` | `bootstrap_calculations_helpers.R` (≈151–318) |
| Tier-2 refit-free reconstruction (JASA Eqs. 6, 10, 12–14) | `fs_debias_gate()` | `fs_debias_gate.R` |

# References

León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024). Exploratory
subgroup identification in the heterogeneous Cox model: a relatively simple
procedure. *Statistics in Medicine* 43(20):3921–3942. DOI:
[10.1002/sim.10163](https://doi.org/10.1002/sim.10163).

León LF, Anderson KM (2026). Post-selection inference for the standard trial
analysis of data-driven subgroups. Manuscript.

Harrell FE, Lee KL, Mark DB (1996). Multivariable prognostic models. *Statistics
in Medicine* 15(4):361–387.

Efron B (2014). Estimation and accuracy after model selection. *JASA*
109(507):991–1007.

Wager S, Hastie T, Efron B (2014). Confidence intervals for random forests: the
jackknife and the infinitesimal jackknife. *JMLR* 15:1625–1651.
