```{r lit-actg175-fallbacks}
#| include: false

# When this file is rendered as a stand-alone document for review,
# the parent vignette's setup chunk has not run, so `actg_age_cut`
# and `actg_preanti_cut` are not in scope.  We supply fallback
# values matching the parent vignette's setup chunk so inline
# references like `r actg_age_cut` resolve.  When this file is
# included as content in the parent vignette, the parent's setup
# chunk runs first and the fallbacks here become no-ops because
# the variables already exist.
if (!exists("actg_age_cut"))     actg_age_cut     <- 34
if (!exists("actg_preanti_cut")) actg_preanti_cut <- 744.5
```

## Literature on Subgroup Discovery in ACTG175 {#sec-literature-actg175}

ACTG175 [@hammer1996] has become the de-facto benchmark dataset for
HTE/subgroup-discovery methods in the precision-medicine literature.
Its appeal is structural: a moderate-sized randomised trial
(n = 2 139) with four well-defined treatment arms, a rich set of
baseline covariates, and both a continuous early CD4 endpoint
(`cd420`, the change in CD4 count at 20 ± 5 weeks) and a censored
composite survival endpoint (the time to a 50% CD4 decline,
AIDS-defining event, or death). Methods papers spanning the past
two decades have used ACTG175 to illustrate variable selection for
individualised treatment rules (ITRs), causal-forest CATE
estimation, doubly-robust regime estimation, classification-tree
subgroup detection, stable-subgroup distillation, mediation
analysis with heterogeneous effects, transportability of trial
findings to external populations, longitudinal recursive
partitioning, and Bayesian dynamic-treatment-regime evaluation.
Any survey of ACTG175 HTE work is therefore necessarily a
sampling rather than an enumeration; the table below covers a
spread of methodological traditions, but related work in each
tradition extends well beyond the specific entries cited.

Across these methodological traditions, the same two effect
modifiers recur: **age** and **prior antiretroviral therapy**
(stored in the data as `preanti`, the number of days of prior
therapy, or `str2`, the binary naive-vs-experienced indicator).
The roles they play differ in important ways. Prior antiretroviral
therapy was the *protocol-specified* stratification factor in the
original trial design [@hammer1996], whereas age has emerged as the
*data-discovered* effect modifier most consistently selected by
modern ITR and CATE-based methods.

```{r lit-actg175-table}
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Selected studies that have analysed ACTG175 for treatment-effect heterogeneity. Treatment contrasts use the original-trial labels: ZDV (zidovudine), ddI (didanosine), ZAL (zalcitabine). Outcome abbreviations: CD4-20w = continuous change in CD4 at 20 ± 5 weeks; surv. composite = time to 50% CD4 decline, AIDS-defining event, or death; t-yr surv. = t-year survival probability."

library(gt)

lit_actg175 <- data.frame(
  citation = c(
    "Hammer et al. (1996), NEJM",
    "Lu, Zhang & Zeng (2013), Stat Methods Med Res",
    "Fan, Lu, Song & Zhou (2017), JRSS-B",
    "Bai, Tsiatis, Lu & Song (2017), Lifetime Data Anal",
    "Jiang, Lu, Song & Davidian (2017), JRSS-B",
    "Cui, Kosorok, Sverdrup, Wager & Zhu (2023), JRSS-B",
    "Johnson, Lu & Davidian (2023), Biometrics",
    "Choi-Schreuer et al. (2021), J Stat Comput Simul",
    "Huang, Tang & Kenney (2025), arXiv (causalDT)",
    "Xi, Huang & Wang (2026), arXiv tutorial",
    "Kundu & Ghosh (2021), Stat Anal Data Min (SurvCART)",
    "Lee, Ghosh & Yang (2024), J Stat Comput Simul",
    "Sun et al. (2024), Stat Med",
    "Moodie & Stephens (2022), BayesDTR R package"
  ),
  method = c(
    "Stratified Cox PH; protocol-specified subgroup interaction test",
    "Penalised loss-based regression for ITR; adaptive LASSO",
    "Concordance-assisted learning (CAL); maximum-rank-correlation ITR",
    "Locally-efficient doubly-robust ITR (additive hazards)",
    "Doubly-robust ITR for t-year survival",
    "Causal survival forests (CSF); CATE + best-linear-projection",
    "One-step value-difference subgroup detection; classification tree",
    "Penalised Cox PH; HARE; comparison of PH-violating settings",
    "Causal distillation trees; stable-subgroup estimation",
    "Hybrid two-stage HTE inference + DR-policy workflow",
    "Survival CART (recursive partitioning via parameter-instability tests)",
    "Transportability via outcome regression / IPW / augmented calibration weighting (ACW)",
    "Heterogeneous mediation analysis for Cox PH with multiple mediators (Bayesian, sparsity-inducing prior)",
    "Bayesian dynamic treatment regimes; ACTG175-based plasmode simulation"
  ),
  contrast = c(
    "All four arms vs ZDV mono",
    "ZDV+ddI vs ddI; ZDV+ddI vs ZDV+ZAL",
    "ZDV+ddI vs ddI",
    "ZDV+ddI vs ZDV+ZAL",
    "Multiple",
    "ZDV+ddI vs ddI",
    "ZDV+ddI vs ZDV mono",
    "All four arms",
    "ZDV mono vs ZDV+ddI",
    "Various",
    "Subgroup detection on combined arms (no specific contrast)",
    "Combination vs ZDV mono, transported to external populations",
    "Combination vs ZDV (ZDV-only subset, n = 532)",
    "Sequential / dynamic treatment decisions"
  ),
  outcome = c(
    "Surv. composite",
    "CD4-20w",
    "CD4-20w",
    "Surv. composite",
    "t-yr surv.",
    "Surv. composite",
    "CD4-20w",
    "Surv. composite",
    "CD4-20w",
    "CD4-20w / surv.",
    "Surv. composite (also longitudinal CD4 in LongCART)",
    "Surv. composite (transported to external survival)",
    "Surv. composite (with mediator pathways)",
    "Longitudinal CD4 / surv."
  ),
  candidate_set = c(
    "Pre-specified: prior ART (str2), other strata",
    "12 covariates: age, wtkg, karnof, cd40, cd80, hemo, homo, drugs, race, gender, str2, symptom",
    "Single covariate (age) following Lu et al. (2013)",
    "log(age), homo (following Lu et al. 2013)",
    "Multiple covariates incl. age",
    "14 covariates incl. age, cd40, karnof, wtkg",
    "12 covariates as in Lu et al. (2013)",
    "12 covariates incl. preanti, age, drugs, symptom",
    "Multiple incl. age",
    "Multiple",
    "14 baseline covariates available in LongCART distribution incl. age, wtkg, karnof, cd40",
    "7 prognostic covariates incl. CD4 count and IV-drug-use history",
    "Baseline + post-randomisation mediators incl. age, sexual-activity (homo), CD4-20w",
    "Multiple incl. age, baseline CD4, prior ART"
  ),
  identified_role = c(
    paste0("Prior-ART subgroups: ZDV+ZAL vs ZDV HR = 0.39 (naive) vs 0.60 ",
           "(experienced); strong qualitative interaction with prior ART"),
    paste0("AGE selected as primary modifier; CD40 and HOMO secondary. ",
           "ZDV+ZAL favoured for young patients; ZDV+ddI for older. ",
           "Preanti not selected."),
    "AGE confirmed as primary modifier; consistent with Lu et al. (2013).",
    paste0("AGE (log-transformed) significant; HOMO marginal. ZDV+ddI ",
           "more beneficial for older patients; ZDV+ZAL for younger ",
           "with homo=1."),
    paste0("AGE the dominant modifier; reported cut at ~34 yrs ",
           "for ZDV+ZAL vs ZDV+ddI."),
    paste0("AGE shows positive monotone CATE: older patients benefit ",
           "more from ZDV+ddI vs ddI. BLP supports caution on ",
           "non-age heterogeneity."),
    paste0("Subgroup defined by classification tree on AGE; density ",
           "plot shows clear separation of benefit / no-benefit groups ",
           "by age."),
    paste0("PREANTI significant in Cox PH model for both off-trt ",
           "strata. AGE x off-trt interaction selected by HARE. ",
           "Karnof, symptom, CD40, CD80 also significant."),
    paste0("Stable subgroup includes AGE among top splits; method ",
           "emphasises stability across teacher models."),
    paste0("Used as illustrative ACTG175 reanalysis; AGE-centred ",
           "heterogeneity recovered."),
    paste0("Used as worked example for SurvCART; CD4 and other baseline ",
           "covariates emerge as splits via parameter-instability tests. ",
           "Methodologically distinct splitting principle from CART/causal forests."),
    paste0("Treatment effect found to differ across external target ",
           "populations; PH assumption *violated* in ACTG175 itself, ",
           "motivating non-PH transport methods. CD4 count and IV drug ",
           "use highlighted as key prognostic covariates."),
    paste0("Confirms heterogeneity by AGE and sexual-activity (homo) ",
           "level via mediation-pathway decomposition. ZDV-only subset ",
           "(n = 532); arrives at AGE finding through a route distinct ",
           "from direct CATE estimation."),
    paste0("ACTG175 used as plasmode backbone for Bayesian DTR ",
           "evaluation; specific subgroups not the primary deliverable.")
  ),
  stringsAsFactors = FALSE
)

gt(lit_actg175) |>
  cols_label(
    citation        = "Citation",
    method          = "Method",
    contrast        = "Treatment contrast",
    outcome         = "Outcome",
    candidate_set   = "Candidate covariates",
    identified_role = md("Identified role of **age** / **preanti**")
  ) |>
  tab_options(
    table.font.size       = px(11),
    column_labels.font.weight = "bold",
    data_row.padding      = px(4),
    table.width           = pct(100)
  ) |>
  cols_width(
    citation        ~ pct(15),
    method          ~ pct(17),
    contrast        ~ pct(10),
    outcome         ~ pct(8),
    candidate_set   ~ pct(20),
    identified_role ~ pct(30)
  )
```

### Methodological breadth

The studies in the table above are not interchangeable: they
address ACTG175 from genuinely different methodological starting
points, and they ask different questions. ITR-style methods
[@lu2013; @fan2017; @bai2017; @jiang2017; @cui2023; @johnson2023]
ask "for which patients is treatment $A$ better than treatment $B$,"
and they yield decision rules. Causal-forest [@cui2023] and
distillation-tree [@huang2025distilling; @xi2026] methods relax parametric
assumptions while still targeting CATE estimation. Recursive
partitioning under parameter-instability tests [@kundu_ghosh2021]
treats subgroup discovery as a tree-induction problem with
type-I-error control at each node, and uses a different splitting
principle than CART or causal forests. Transportability methods
[@lee_ghosh_yang2024] step outside the trial entirely and ask
whether the within-trial effect generalises to external populations,
finding the proportional-hazards assumption *violated* for ACTG175
itself. Heterogeneous mediation analysis [@sun2024_mediation]
decomposes the total effect into pathways through post-randomisation
mediators (e.g. CD4-20w) and identifies effect modifiers that drive
each pathway separately. Bayesian dynamic-treatment-regime
evaluation [@moodie_stephens_bayesdtr] uses ACTG175 as a plasmode
backbone for sequential decision-making rather than single-stage
subgroup identification.

These traditions disagree on what counts as the "right" answer: a
decision rule, a CATE function, a tree, a transported effect
estimate, a mediation pathway, or a sequential policy. They
nevertheless converge on a pair of effect modifiers — age and prior
antiretroviral therapy — even though no two of them are doing the
same statistical thing. That convergence is part of what motivates
treating the joint Age × Preanti subgroup as a substantively
interesting object to test, rather than as an artefact of any one
method's regularisation choices.

### Convergent and divergent findings

**The strongest, most consistent signal is age.** Among all
ITR-style methods that admit continuous covariates and use the
CD4-20w endpoint, age is selected as the dominant treatment-effect
modifier, with the qualitative pattern that ZDV+zalcitabine is
preferred for younger patients and ZDV+ddI for older patients
[@lu2013; @fan2017; @bai2017; @jiang2017]. The cut-points reported
across these papers cluster broadly: Lu et al. (2013) report
roughly 17.8 yrs (homo = 0) or 26.3 yrs (homo = 1) for the
ZDV+ddI vs ZDV+ZAL contrast; Jiang et al. (2017) report
approximately 34 yrs as a discriminating value for survival;
and the causal-survival
forest of @cui2023 recovers a smooth monotone CATE
function in age over the ZDV+ddI vs ddI contrast, with older
patients benefiting more. Subsequent qualitative-interaction tests
[@shi2021; @johnson2023] continue to confirm age as a true effect
modifier rather than a coincidental selection artefact.

**Preanti's role is more nuanced.** In the original NEJM analysis,
prior antiretroviral therapy (`str2`) was the protocol-stratified
subgroup variable; the subgroup-by-treatment interaction was
pre-specified. Hammer et al. (1996) reported notably different
relative hazards for ZDV+ZAL vs ZDV alone in the antiretroviral-
naive ($HR \approx 0.39$) vs antiretroviral-experienced
($HR \approx 0.60$) subgroups, supporting the importance of prior
treatment exposure. Modern ITR methods that operate on the
continuous CD4-20w endpoint, however, *typically do not select
preanti* as an effect modifier: Lu et al. (2013), for instance,
found age, CD40 and homo to be informative but did not retain
preanti or str2. By contrast, **survival-endpoint analyses
consistently surface preanti as a significant covariate**:
[@choi2021] find preanti to be significant in the
Cox PH model for both off-treatment strata, and several
censoring-distribution analyses surface preanti among the
significant predictors of censoring [@choi2023]. The most
plausible interpretation is that preanti acts more as a *prognostic*
covariate over the survival horizon than as a strong *predictive*
modifier of CD4 response at 20 weeks.

**A note on the candidate set.** Almost all of the methodology
papers above operate on the same 12-covariate "Lu et al. (2013)"
candidate set: 5 continuous (age, weight, Karnofsky, CD4 baseline,
CD8 baseline) and 7 binary (hemo, homo, drugs, race, gender, str2,
symptom). Notably, this set includes the binary `str2` indicator
but typically *not* the continuous `preanti` (days of prior
therapy). Methods constrained to this candidate set therefore have
limited ability to detect cuts on preanti at intermediate values
of prior treatment exposure. forestsearch's exploration of an
expanded candidate set that includes `preanti` directly may
therefore identify subgroup definitions that prior ITR methods
were structurally unable to recover.

### Connection to the present analysis

The benefit subgroup identified by `forestsearch` in this vignette
--- defined by AGE > $`r actg_age_cut`$ AND PREANTI $\leq
`r actg_preanti_cut`$ --- aligns with the literature on two
counts. First, the *direction* of the age effect (older patients
deriving greater benefit from the combination therapy) matches
@cui2023, @bai2017, and @jiang2017.
Second, the *involvement of preanti* connects the discovered
subgroup to the original Hammer et al. (1996) protocol-specified
finding that prior antiretroviral experience modifies treatment
efficacy --- a connection that ITR methods restricted to the
standard 12-covariate set cannot directly recover.

The discovered cut-point on age and the directionality of preanti
involvement should be regarded as *exploratory* findings consistent
with --- but not independently confirming --- two decades of HTE
analyses on ACTG175. In particular: the bias-corrected effect
estimates from `forestsearch` (e.g. $\hat\theta^*$ in the
estimation tables) are designed to debias the magnitude of
treatment effect *within the identified subgroup*; they do not
provide independent confirmation that the subgroup definition is
itself correct. External replication on an independent ACTG cohort
would be required for that purpose.

::: {.callout-note title="What this literature does NOT establish"}
The convergent age signal across many ITR methods does not, on its
own, validate any single threshold. Different papers have used
different treatment contrasts, different outcomes (continuous vs
survival), and different candidate sets, and the reported "best"
age cuts span a range from roughly 18 to 34 years. The
forestsearch threshold of $`r actg_age_cut`$ is best read as
*consistent with* this body of work rather than as the literature's
consensus value.
:::

<!-- ─────────────────────────────────────────────────────────────
     Bibliography entries (BibTeX format) for the @key citations
     used above. Move these into the vignette's references.bib or
     paste into a YAML `bibliography:` file. The keys match the
     in-text @cite calls. If the parent vignette already has a
     references.bib, deduplicate before merging.
     ───────────────────────────────────────────────────────── -->

<!--
@article{hammer1996,
  author  = {Hammer, S. M. and Katzenstein, D. A. and Hughes, M. D. and
             Gundacker, H. and Schooley, R. T. and Haubrich, R. H. and
             Henry, W. K. and Lederman, M. M. and Phair, J. P. and
             Niu, M. and Hirsch, M. S. and Merigan, T. C.},
  title   = {A trial comparing nucleoside monotherapy with combination
             therapy in {HIV}-infected adults with {CD4} cell counts
             from 200 to 500 per cubic millimeter},
  journal = {New England Journal of Medicine},
  volume  = {335},
  number  = {15},
  pages   = {1081--1090},
  year    = {1996},
  doi     = {10.1056/NEJM199610103351501}
}

@article{lu2013,
  author  = {Lu, W. and Zhang, H. H. and Zeng, D.},
  title   = {Variable selection for optimal treatment decision},
  journal = {Statistical Methods in Medical Research},
  volume  = {22}, number = {5}, pages = {493--504}, year = {2013},
  doi     = {10.1177/0962280211428383}
}

@article{fan2017,
  author  = {Fan, C. and Lu, W. and Song, R. and Zhou, Y.},
  title   = {Concordance-assisted learning for estimating optimal
             individualized treatment regimes},
  journal = {Journal of the Royal Statistical Society, Series B},
  volume  = {79}, number = {5}, pages = {1565--1582}, year = {2017},
  doi     = {10.1111/rssb.12216}
}

@article{bai2017,
  author  = {Bai, X. and Tsiatis, A. A. and Lu, W. and Song, R.},
  title   = {Optimal treatment regimes for survival endpoints using
             a locally-efficient doubly-robust estimator from a
             classification perspective},
  journal = {Lifetime Data Analysis},
  volume  = {23}, number = {4}, pages = {585--604}, year = {2017},
  doi     = {10.1007/s10985-016-9376-x}
}

@article{jiang2017,
  author  = {Jiang, R. and Lu, W. and Song, R. and Davidian, M.},
  title   = {On estimation of optimal treatment regimes for maximizing
             {$t$}-year survival probability},
  journal = {Journal of the Royal Statistical Society, Series B},
  volume  = {79}, pages = {1165--1185}, year = {2017},
  doi     = {10.1111/rssb.12201}
}

@article{cui2023,
  author  = {Cui, Y. and Kosorok, M. R. and Sverdrup, E. and
             Wager, S. and Zhu, R.},
  title   = {Estimating heterogeneous treatment effects with
             right-censored data via causal survival forests},
  journal = {Journal of the Royal Statistical Society, Series B},
  volume  = {85}, number = {2}, pages = {179--211}, year = {2023},
  doi     = {10.1093/jrsssb/qkac001}
}

@article{johnson2023,
  author  = {Johnson, D. and Lu, W. and Davidian, M.},
  title   = {A general framework for subgroup detection via
             one-step value difference estimation},
  journal = {Biometrics},
  volume  = {79}, number = {3}, pages = {2116--2126}, year = {2023},
  doi     = {10.1111/biom.13711}
}

@article{shi2021,
  author  = {Shi, C. and Lu, W. and Song, R.},
  title   = {On testing conditional qualitative treatment effects},
  journal = {Annals of Statistics},
  volume  = {49}, number = {4}, pages = {2298--2319}, year = {2021}
}

@article{choi2021,
  author  = {Choi-Schreuer, E. and Schreuer, A. and Schmid, M.},
  title   = {Effects of proportional hazard assumption on variable
             selection methods for censored data},
  journal = {Journal of Statistical Computation and Simulation},
  year    = {2021}
}

@article{choi2023,
  author  = {Choi, J. and others},
  title   = {Accountable survival contrast-learning for optimal
             dynamic treatment regimes},
  journal = {Scientific Reports},
  year    = {2023},
  doi     = {10.1038/s41598-023-29106-w}
}

@misc{huang2025distilling,
  author  = {Huang, M. and Tang, T. M. and Kenney, A. M.},
  title   = {Distilling heterogeneous treatment effects: Stable
             subgroup estimation in causal inference},
  year    = {2025},
  eprint  = {2502.07275},
  archivePrefix = {arXiv},
  primaryClass  = {stat.ME}
}

@misc{xi2026,
  author  = {Xi, N. M. and Huang, X. and Wang, L.},
  title   = {Subgroup identification and individualized treatment
             policies: A tutorial on the hybrid two-stage workflow},
  year    = {2026},
  eprint  = {2601.00136},
  archivePrefix = {arXiv}
}

@article{kundu_ghosh2021,
  author  = {Kundu, M. G. and Ghosh, S.},
  title   = {Survival trees based on heterogeneity in time-to-event
             and censoring distributions using parameter instability
             test},
  journal = {Statistical Analysis and Data Mining: The ASA Data
             Science Journal},
  volume  = {14},
  number  = {5},
  pages   = {466--483},
  year    = {2021},
  doi     = {10.1002/sam.11534}
}

@article{lee_ghosh_yang2024,
  author  = {Lee, D. and Ghosh, S. and Yang, S.},
  title   = {Transporting survival of an {HIV} clinical trial to the
             external target populations},
  journal = {Journal of Statistical Computation and Simulation},
  year    = {2024},
  note    = {Published online 23 March 2024; arXiv:2210.02571},
  doi     = {10.1080/00949655.2024.2330519}
}

@article{sun2024_mediation,
  author  = {Sun, J. and others},
  title   = {Heterogeneous mediation analysis for {Cox} proportional
             hazards model with multiple mediators},
  journal = {Statistics in Medicine},
  year    = {2024},
  doi     = {10.1002/sim.10239},
  note    = {PMC11588993}
}

@manual{moodie_stephens_bayesdtr,
  author  = {Moodie, E. E. M. and Stephens, D. A.},
  title   = {{BayesDTR}: {Bayesian} Methods for Dynamic Treatment
             Regimes (R package)},
  note    = {Uses ACTG175 (via the LongCART package) for plasmode
             simulation of dynamic treatment regimes}
}
-->
