---
title: "Reproduction: León et al. (2024) Table 1, M1 Block"
subtitle: "ForestSearch v0.2.0 — Pressure Test"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: true
  message: true
---


```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ═══════════════════════════════════════════════════════════════════
# ALL TUNABLE PARAMETERS
# ═══════════════════════════════════════════════════════════════════

# -- Simulation budget ---------------------------------------------------------
nsims              <- 20000L        # Paper uses 20000; scale up for production
n_workers          <- max(1L, floor(0.90 * parallel::detectCores(logical = FALSE)))

# -- DGM: M1 scenario (León et al. 2024, Section 3) ---------------------------
# N = 700, pH ≈ 13%, θ†(H) = 2.0, θ†(Hc) = 0.65, θ†(ITT) ≈ 0.71
# Censoring ≈ 46%, analysis_time = 84 months
sim_n_sample       <- 700L
sim_analysis_time  <- 84L         # GBSG follow-up (months)
sim_cens_adjust    <- log(1.5)    # Censoring adjustment

dgm_p_harm        <- 0.13        # target prevalence of H
dgm_hr_h1         <- 2.0         # target θ†(H)
dgm_z1_quantile   <- 0.25        # ER cutoff quantile → pH ≈ 13%
dgm_n_super        <- 5000L
dgm_seed           <- 8316951L

# -- ForestSearch parameters (Section 2, Step 3) -------------------------------
# c1 = 1.25 (screening), c2 = 1.0 (consistency)
# "at least 60 subjects and 10 events in each arm"
fs_hr_threshold    <- 1.25
fs_hr_consistency  <- 1.0
fs_pconsistency    <- 0.90
fs_splits          <- 400L
fs_n_min           <- 60L
fs_d0_min          <- 10L         # Paper: "10 events in each arm"
fs_d1_min          <- 10L
fs_maxk            <- 2L
fs_vi_grf_min      <- -0.2
fs_seedit          <- 8316951L

# -- GRF standalone parameters (Section 3.1) -----------------------------------
# Paper criterion: "RMST benefit of (at least) 6 months for control."
# DR scores from causal_survival_forest are anti-symmetric:
#   control[i] = -treated[i]  (verified in §1.2 DR Score Verification)
# Policy tree diff = control_mean - treated_mean = X - (-X) = 2X.
# A 6-month RMST benefit (X = 6) → diff = 12.
# Therefore dmin.grf = 12, matching default_grf_params_gen().
# GRF60: τ = 0.6 * min(τ0, τ1)
grf_dmin           <- 12          # 6-month RMST criterion (= 2 × 6 on DR scale)
grf_dmin_60        <- 12          # Same criterion for GRF60
grf_frac_tau       <- 0.6         # GRF60 truncation fraction
grf_maxdepth       <- 2L
grf_sg_criterion   <- "mDiff"
grf_tune           <- FALSE       # Paper predates tune_grf feature

# -- Confounders (no noise for first block) ------------------------------------
# Z1-Z5: prognostic; size, grade3: nonprognostic but correlated
confounders_base   <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")
n_add_noise        <- 0L          # First block: no noise factors

# -- Cox formula (treatment-only, per paper Section 2) -------------------------
cox_formula_adj    <- survival::Surv(y_sim, event_sim) ~ treat_sim
```


# Paper Reference

This document reproduces the first block of Table 1 from:

> León, L.F., Jemielita, T., Guo, Z., Marceau West, R., & Anderson, K.M.
> (2024). Exploratory subgroup identification in the heterogeneous Cox model:
> A relatively simple procedure. *Statistics in Medicine*, 43, 3921–3942.

**M1 scenario**: N = 700, pH ≈ 13%, θ†(H) = 2.0, θ†(Hc) = 0.65,
θ†(ITT) ≈ 0.71, censoring ≈ 46%.  Average size of true H = 89.

::: {.callout-tip}
## Summary of concordance

See [Summary and Concordance] at the end for a data-driven
comparison of all methods against the paper, including the
`dmin.grf` calibration rationale and the unconditional averaging
convention.
:::

## DR Score Verification: `dmin.grf = 12`

The paper's GRF criterion is "an estimated RMST benefit of
(at least) 6 months for control" (Section 3.1).  The DR scores
from `causal_survival_forest` are **anti-symmetric**
($\Gamma_{i,\text{control}} = -\Gamma_{i,\text{treated}}$),
so the policy tree leaf difference is twice the raw RMST effect:
`diff = X − (−X) = 2X`.  A 6-month benefit → `dmin.grf = 12`.

```{r dr-score-check}
#| code-fold: false

# Verify the anti-symmetric structure of causal_survival_forest DR scores
set.seed(42)
n_check <- 500
X_check <- matrix(rnorm(n_check * 5), n_check, 5)
W_check <- rbinom(n_check, 1, 0.5)
Y_check <- rexp(n_check) * (1 + 0.3 * W_check * (X_check[, 1] > 0))
D_check <- rbinom(n_check, 1, 0.8)
h_check <- quantile(Y_check[D_check == 1], 0.9)

csf_check <- grf::causal_survival_forest(
  X_check, Y_check, W_check, D_check,
  W.hat = 0.5, horizon = h_check, seed = 42, num.trees = 500
)
dr_check <- policytree::double_robust_scores(csf_check)

cat("=== DR Score Anti-Symmetry Verification ===\n")
cat(sprintf("  colnames: %s\n", paste(colnames(dr_check), collapse = ", ")))
cat(sprintf("  control + treated = 0?  max|sum| = %.2e\n",
    max(abs(dr_check[, 1] + dr_check[, 2]))))
cat(sprintf("  => diff = control - treated = 2 * control\n"))
cat(sprintf("  => dmin.grf = 12 implements a '6-month RMST benefit'\n"))

# Show the doubling relationship
cate_check <- predict(csf_check)$predictions
dr_diff <- dr_check[, 1] - dr_check[, 2]
cat(sprintf("\n  mean(CATE): %.4f\n", mean(cate_check)))
cat(sprintf("  mean(control - treated): %.4f\n", mean(dr_diff)))
cat(sprintf("  ratio (diff / CATE): %.2f  (expected ≈ 2.0)\n",
    mean(dr_diff) / mean(cate_check)))
```


## Expected Results (Paper Table 1, M1, No Noise, 20,000 Sims)

### M1 Null: N = 700, θ†(ITT) = 0.7

Under H0, H = ∅.  Per Section 3: sens(Ĥ) is undefined,
ppv(Ĥ) ≡ 0.  Only any(H), sens(Ĥ$^c$), ppv(Ĥ$^c$), and
avg|Ĥ| are reported.

|  | FSl | FSlg | GRF | GRF60 | VT(24) | VT(36) |
|---|---|---|---|---|---|---|
| **any(H)** | 0.02 | 0.03 | **0.25** | 0.05 | 0.03 | 0.04 |
| **sens(Ĥ$^c$)** | 1.00 | 1.00 | 0.97 | 0.99 | 1.00 | 1.00 |
| **ppv(Ĥ$^c$)** | 1.00 | 1.00 | 1.00 | 1.00 | 1.00 | 1.00 |
| **avg\|Ĥ\|** | 114 | 99 | 88 | 78 | 78 | 79 |

Boldface = any(H) ≥ 0.10 (inflated type I error per paper convention).
Asymptotic approximation (equation 3): 0.036.

### M1 Alt: N = 700, pH = 13%, θ†(H) = 2, θ†(Hc) = 0.65, θ†(ITT) = 0.71

|  | FSl | FSlg | GRF | GRF60 | VT(24) | VT(36) |
|---|---|---|---|---|---|---|
| **any(H)** | 0.77 | 0.86 | 0.94 | 0.72 | 0.49 | 0.47 |
| **sens(Ĥ)** | 0.72 | 0.82 | 0.84 | 0.66 | 0.46 | 0.42 |
| **sens(Ĥ$^c$)** | 0.99 | 0.99 | 0.97 | 0.98 | 0.99 | 0.99 |
| **ppv(Ĥ)** | 0.69 | 0.80 | 0.78 | 0.61 | 0.44 | 0.41 |
| **ppv(Ĥ$^c$)** | 0.96 | 0.98 | 0.98 | 0.96 | 0.93 | 0.93 |
| **avg\|Ĥ\|** | 94 | 92 | 102 | 99 | 92 | 93 |

Asymptotic approximation for detection (equation 3): 0.90.
Average size of true H = 89.

::: {.callout-note}
VT(24) and VT(36) (virtual twins) are included for reference
but are not reproduced here.  This document runs FSl, FSlg,
GRF, and GRF60 — the four methods implemented in `forestsearch`.
:::


# DGM Calibration

```{r dgm-calibration}
library(forestsearch)
library(survival)
library(data.table)
library(foreach)
library(doFuture)
library(ggplot2)

# Calibrate k_inter for θ†(H) ≈ 2.0
k_inter_cal <- calibrate_k_inter(
  target_hr_harm = dgm_hr_h1,
  z1_quantile    = dgm_z1_quantile,
  k_treat        = 1.0,
  n_super        = dgm_n_super,
  seed           = dgm_seed,
  verbose        = TRUE
)

cat(sprintf("Calibrated k_inter = %.4f\n", k_inter_cal))
```

```{r dgm-create}
# Alternative DGM
dgm_alt <- setup_gbsg_dgm(
  model       = "alt",
  k_treat     = 1.0,
  k_inter     = k_inter_cal,
  z1_quantile = dgm_z1_quantile,
  n_super     = dgm_n_super,
  seed        = dgm_seed,
  verbose     = TRUE
)
dgm_alt <- compute_dgm_cde(dgm_alt)

# Null DGM
dgm_null <- setup_gbsg_dgm(
  model   = "null",
  k_treat = 1.0,
  n_super = dgm_n_super,
  seed    = dgm_seed,
  verbose = TRUE
)

cat("\n=== M1 DGM Properties ===\n")
cat(sprintf("  pH:       %.1f%%\n", 100 * mean(dgm_alt$df_super$flag_harm)))
cat(sprintf("  θ†(H):    %.3f  (target: %.1f)\n",
    dgm_alt$hr_H_true, dgm_hr_h1))
cat(sprintf("  θ†(Hc):   %.3f\n", dgm_alt$hr_Hc_true))
cat(sprintf("  θ†(ITT):  %.3f  (alt)\n", dgm_alt$hr_causal))
cat(sprintf("  θ†(ITT):  %.3f  (null)\n", dgm_null$hr_causal))
```


# Simulation Functions

```{r sim-setup}
#| code-fold: false

plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")

# ── FS parameter lists ────────────────────────────────────────────
# FSl: LASSO only (no GRF screening)
fs_params_lasso <- list(
  outcome.name           = "y_sim",
  event.name             = "event_sim",
  treat.name             = "treat_sim",
  id.name                = "id",
  hr.threshold           = fs_hr_threshold,
  hr.consistency         = fs_hr_consistency,
  pconsistency.threshold = fs_pconsistency,
  fs.splits              = fs_splits,
  n.min                  = fs_n_min,
  d0.min                 = fs_d0_min,
  d1.min                 = fs_d1_min,
  maxk                   = fs_maxk,
  vi.grf.min             = fs_vi_grf_min,
  seedit                 = fs_seedit,
  use_lasso              = TRUE,
  use_grf                = FALSE,    # FSl: LASSO only
  tune_grf               = grf_tune,
  parallel_args          = list(
    plan = "sequential", workers = 1, show_message = FALSE
  )
)

# FSlg: LASSO + GRF (paper's combined algorithm)
fs_params_lasso_grf <- modifyList(fs_params_lasso, list(
  use_grf = TRUE     # FSlg: LASSO + GRF
))

# ── GRF standalone parameter lists ───────────────────────────────
# GRF: full τ = min(τ0, τ1)
grf_params_full <- list(
  dmin.grf     = grf_dmin,
  frac.tau     = 1.0,              # GRF: full horizon
  maxdepth     = grf_maxdepth,
  sg.criterion = grf_sg_criterion,
  tune_grf     = grf_tune
)

# GRF60: τ = 0.6 * min(τ0, τ1)
grf_params_60 <- list(
  dmin.grf     = grf_dmin_60,
  frac.tau     = grf_frac_tau,     # GRF60: 60% horizon
  maxdepth     = grf_maxdepth,
  sg.criterion = grf_sg_criterion,
  tune_grf     = grf_tune
)
```


# Run Simulations

```{r collect-helper}
#| include: false
# collect_results() now lives in forestsearch (R/forestsearch_helpers.R).
# Called via ::: since it is an internal (unexported) utility.
```

## Alternative Hypothesis (M1 Alt)

Four analyses per simulation: FSl, FSlg, GRF, GRF60.

```{r run-alt}
t0 <- proc.time()[3]

results_alt_raw <- foreach(
  i = seq_len(nsims),
  .errorhandling = "pass",
  .options.future = list(seed = TRUE)
) %dofuture% {
  future::plan("sequential")

  # FSl (LASSO only)
  r_fsl <- tryCatch(
    run_simulation_analysis(
      sim_id = i, dgm = dgm_alt,
      n_sample = sim_n_sample,
      confounders_base = confounders_base,
      n_add_noise = n_add_noise,
      run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
      fs_params = fs_params_lasso,
      cox_formula_adj = cox_formula_adj,
      analysis_time = sim_analysis_time,
      cens_adjust = sim_cens_adjust,
      verbose = FALSE
    ),
    error = function(e) NULL
  )
  if (!is.null(r_fsl)) r_fsl$analysis <- "FSl"

  # FSlg (LASSO + GRF)
  r_fslg <- tryCatch(
    run_simulation_analysis(
      sim_id = i, dgm = dgm_alt,
      n_sample = sim_n_sample,
      confounders_base = confounders_base,
      n_add_noise = n_add_noise,
      run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
      fs_params = fs_params_lasso_grf,
      cox_formula_adj = cox_formula_adj,
      analysis_time = sim_analysis_time,
      cens_adjust = sim_cens_adjust,
      verbose = FALSE
    ),
    error = function(e) NULL
  )
  if (!is.null(r_fslg)) r_fslg$analysis <- "FSlg"

  # GRF (full τ)
  r_grf <- tryCatch(
    run_simulation_analysis(
      sim_id = i, dgm = dgm_alt,
      n_sample = sim_n_sample,
      confounders_base = confounders_base,
      n_add_noise = n_add_noise,
      run_fs = FALSE, run_fs_grf = FALSE, run_grf = TRUE,
      grf_params = grf_params_full,
      cox_formula_adj = cox_formula_adj,
      analysis_time = sim_analysis_time,
      cens_adjust = sim_cens_adjust,
      verbose = FALSE
    ),
    error = function(e) NULL
  )
  if (!is.null(r_grf)) r_grf$analysis <- "GRF"

  # GRF60 (0.6τ)
  r_grf60 <- tryCatch(
    run_simulation_analysis(
      sim_id = i, dgm = dgm_alt,
      n_sample = sim_n_sample,
      confounders_base = confounders_base,
      n_add_noise = n_add_noise,
      run_fs = FALSE, run_fs_grf = FALSE, run_grf = TRUE,
      grf_params = grf_params_60,
      cox_formula_adj = cox_formula_adj,
      analysis_time = sim_analysis_time,
      cens_adjust = sim_cens_adjust,
      verbose = FALSE
    ),
    error = function(e) NULL
  )
  if (!is.null(r_grf60)) r_grf60$analysis <- "GRF60"

  data.table::rbindlist(
    Filter(Negate(is.null), list(r_fsl, r_fslg, r_grf, r_grf60)),
    fill = TRUE
  )
}

plan("sequential")
t_alt <- proc.time()[3] - t0

results_alt <- forestsearch:::collect_results(results_alt_raw, "H1 alt")
n_failed_alt <- attr(results_alt, "n_failed")
cat(sprintf("Alt: %d sims completed in %.1f min\n", nsims, t_alt / 60))
if (n_failed_alt > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_alt))
```


```{r reset-workers-null}
# Reset worker pool to prevent memory accumulation from alt sims.
# Without this, null sims run on bloated workers and are ~20x slower.
plan(sequential)
gc(verbose = FALSE)
Sys.sleep(1)
plan(multisession, workers = n_workers)
cat(sprintf("Workers reset: %d cores\n", n_workers))
```

## Null Hypothesis (M1 Null)

```{r run-null}
t0 <- proc.time()[3]

results_null_raw <- foreach(
  i = seq_len(nsims),
  .errorhandling = "pass",
  .options.future = list(seed = TRUE)
) %dofuture% {
  future::plan("sequential")

  r_fsl <- tryCatch(
    run_simulation_analysis(
      sim_id = i, dgm = dgm_null,
      n_sample = sim_n_sample,
      confounders_base = confounders_base,
      n_add_noise = n_add_noise,
      run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
      fs_params = fs_params_lasso,
      cox_formula_adj = cox_formula_adj,
      analysis_time = sim_analysis_time,
      cens_adjust = sim_cens_adjust,
      verbose = FALSE
    ),
    error = function(e) NULL
  )
  if (!is.null(r_fsl)) r_fsl$analysis <- "FSl"

  r_fslg <- tryCatch(
    run_simulation_analysis(
      sim_id = i, dgm = dgm_null,
      n_sample = sim_n_sample,
      confounders_base = confounders_base,
      n_add_noise = n_add_noise,
      run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
      fs_params = fs_params_lasso_grf,
      cox_formula_adj = cox_formula_adj,
      analysis_time = sim_analysis_time,
      cens_adjust = sim_cens_adjust,
      verbose = FALSE
    ),
    error = function(e) NULL
  )
  if (!is.null(r_fslg)) r_fslg$analysis <- "FSlg"

  r_grf <- tryCatch(
    run_simulation_analysis(
      sim_id = i, dgm = dgm_null,
      n_sample = sim_n_sample,
      confounders_base = confounders_base,
      n_add_noise = n_add_noise,
      run_fs = FALSE, run_fs_grf = FALSE, run_grf = TRUE,
      grf_params = grf_params_full,
      cox_formula_adj = cox_formula_adj,
      analysis_time = sim_analysis_time,
      cens_adjust = sim_cens_adjust,
      verbose = FALSE
    ),
    error = function(e) NULL
  )
  if (!is.null(r_grf)) r_grf$analysis <- "GRF"

  r_grf60 <- tryCatch(
    run_simulation_analysis(
      sim_id = i, dgm = dgm_null,
      n_sample = sim_n_sample,
      confounders_base = confounders_base,
      n_add_noise = n_add_noise,
      run_fs = FALSE, run_fs_grf = FALSE, run_grf = TRUE,
      grf_params = grf_params_60,
      cox_formula_adj = cox_formula_adj,
      analysis_time = sim_analysis_time,
      cens_adjust = sim_cens_adjust,
      verbose = FALSE
    ),
    error = function(e) NULL
  )
  if (!is.null(r_grf60)) r_grf60$analysis <- "GRF60"

  data.table::rbindlist(
    Filter(Negate(is.null), list(r_fsl, r_fslg, r_grf, r_grf60)),
    fill = TRUE
  )
}

plan("sequential")
t_null <- proc.time()[3] - t0

results_null <- forestsearch:::collect_results(results_null_raw, "H0 null")
n_failed_null <- attr(results_null, "n_failed")
cat(sprintf("Null: %d sims completed in %.1f min\n", nsims, t_null / 60))
if (n_failed_null > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_null))
```


# Results

## Operating Characteristics

```{r oc-tables}
# ── H1: Detection and Classification ─────────────────────────────
format_oc_results(
  results_alt,
  analyses = c("FSl", "FSlg", "GRF", "GRF60"),
  metrics  = c("detection", "classification"),
  digits   = 2,
  title    = "M1 Alt: Identification (Harm Search)",
  subtitle = sprintf(
    "N = %d, %d sims, θ†(H) = %.1f, θ†(Hc) = %.2f, θ†(ITT) ≈ %.2f",
    sim_n_sample, nsims, dgm_hr_h1,
    dgm_alt$hr_Hc_true, dgm_alt$hr_causal
  ),
  subgroup_notation = "harm"
)

# ── H0: Type I Error ─────────────────────────────────────────────
format_oc_results(
  results_null,
  analyses = c("FSl", "FSlg", "GRF", "GRF60"),
  metrics  = c("detection", "classification"),
  digits   = 2,
  title    = "M1 Null: Type I Error",
  subtitle = sprintf(
    "N = %d, %d sims, θ†(ITT) ≈ %.2f",
    sim_n_sample, nsims, dgm_null$hr_causal
  ),
  subgroup_notation = "harm"
)
```


## Classification Table (Paper Format)

```{r classification-table}
build_classification_table(
  scenario_results = list(
    null = list(
      results    = results_null,
      label      = "M1",
      n_sample   = sim_n_sample,
      dgm        = dgm_null,
      hypothesis = "null"
    ),
    alt = list(
      results    = results_alt,
      label      = "M1",
      n_sample   = sim_n_sample,
      dgm        = dgm_alt,
      hypothesis = "alt"
    )
  ),
  analyses = c("FSl", "FSlg", "GRF", "GRF60"),
  digits   = 2,
  title    = sprintf(
    "Table 1 Reproduction (M1, N=%d, %d sims) | θ†(H)=%.1f, θ†(Hc)=%.2f, pH≈%d%%",
    sim_n_sample, nsims, dgm_hr_h1, dgm_alt$hr_Hc_true,
    round(100 * mean(dgm_alt$df_super$flag_harm))
  ),
  subgroup_notation = "harm",
  font_size = 14
)
```


## Side-by-Side Comparison with Paper

```{r comparison}
#| code-fold: true
#| code-summary: "Build comparison tables"

# Paper's Table 1 M1 values (20,000 sims, no noise)
paper <- data.frame(
  Method  = c("FSl", "FSlg", "GRF", "GRF60"),
  # Null
  null_any     = c(0.02, 0.03, 0.25, 0.05),
  null_sens_Hc = c(1.00, 1.00, 0.97, 0.99),
  null_ppv_Hc  = c(1.00, 1.00, 1.00, 1.00),
  null_avg_H   = c(114,  99,   88,   78),
  # Alt
  alt_any      = c(0.77, 0.86, 0.94, 0.72),
  alt_sens_H   = c(0.72, 0.82, 0.84, 0.66),
  alt_sens_Hc  = c(0.99, 0.99, 0.97, 0.98),
  alt_ppv_H    = c(0.69, 0.80, 0.78, 0.61),
  alt_ppv_Hc   = c(0.96, 0.98, 0.98, 0.96),
  alt_avg_H    = c(94,   92,  102,   99),
  stringsAsFactors = FALSE
)

# Extract our results
our <- data.frame(Method = c("FSl", "FSlg", "GRF", "GRF60"),
                  stringsAsFactors = FALSE)
for (m in our$Method) {
  rn <- results_null[results_null$analysis == m, ]
  ra <- results_alt[results_alt$analysis == m, ]
  rn_det <- rn[rn$any.H == 1, ]
  ra_det <- ra[ra$any.H == 1, ]

  # Null
  our$null_any[our$Method == m]     <- mean(rn$any.H, na.rm = TRUE)
  our$null_sens_Hc[our$Method == m] <- if (nrow(rn_det) > 0)
    mean(rn_det$spec, na.rm = TRUE) else NA
  our$null_ppv_Hc[our$Method == m]  <- if (nrow(rn_det) > 0)
    mean(rn_det$npv, na.rm = TRUE) else NA
  our$null_avg_H[our$Method == m]   <- if (nrow(rn_det) > 0)
    mean(rn_det$size.H, na.rm = TRUE) else NA

  # Alt — unconditional averaging (León et al. 2024, Table 1 convention).
  # When Ĥ = ∅ under alt: sens(Ĥ) = 0/|H| = 0,  ppv(Ĥ) = 0 (convention).
  # sens(Ĥc) = 1 (everyone is in Ĥc),  ppv(Ĥc) = |Hc|/N.
  ra_sens <- ifelse(is.na(ra$sens) & ra$any.H == 0, 0, ra$sens)
  ra_ppv  <- ifelse(is.na(ra$ppv)  & ra$any.H == 0, 0, ra$ppv)
  ra_spec <- ifelse(is.na(ra$spec) & ra$any.H == 0, 1, ra$spec)
  ra_npv  <- ifelse(is.na(ra$npv)  & ra$any.H == 0, 1, ra$npv)

  our$alt_any[our$Method == m]      <- mean(ra$any.H, na.rm = TRUE)
  our$alt_sens_H[our$Method == m]   <- mean(ra_sens, na.rm = TRUE)
  our$alt_sens_Hc[our$Method == m]  <- mean(ra_spec, na.rm = TRUE)
  our$alt_ppv_H[our$Method == m]    <- mean(ra_ppv, na.rm = TRUE)
  our$alt_ppv_Hc[our$Method == m]   <- mean(ra_npv, na.rm = TRUE)
  our$alt_avg_H[our$Method == m]    <- if (nrow(ra_det) > 0)
    mean(ra_det$size.H, na.rm = TRUE) else NA
}
```

### M1 Alternative: Paper vs Reproduction

```{r comparison-alt-table}
fmt2 <- function(x) ifelse(is.na(x) | is.nan(x), "—", sprintf("%.2f", x))
fmt0 <- function(x) ifelse(is.na(x) | is.nan(x), "—", sprintf("%.0f", x))

alt_df <- data.frame(
  Method      = rep(paper$Method, each = 2),
  Source      = rep(c("Paper", sprintf("Ours (%d)", nsims)), 4),
  `any(H)`    = c(rbind(fmt2(paper$alt_any),    fmt2(our$alt_any))),
  `sens(H)`   = c(rbind(fmt2(paper$alt_sens_H), fmt2(our$alt_sens_H))),
  `sens(Hc)`  = c(rbind(fmt2(paper$alt_sens_Hc),fmt2(our$alt_sens_Hc))),
  `ppv(H)`    = c(rbind(fmt2(paper$alt_ppv_H),  fmt2(our$alt_ppv_H))),
  `ppv(Hc)`   = c(rbind(fmt2(paper$alt_ppv_Hc), fmt2(our$alt_ppv_Hc))),
  `avg|H|`    = c(rbind(fmt0(paper$alt_avg_H),  fmt0(our$alt_avg_H))),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt::gt(alt_df, groupname_col = "Method") |>
  gt::tab_header(
    title = "M1 Alternative: Paper (20,000 sims) vs Reproduction",
    subtitle = sprintf(
      "N = %d, %d sims | θ†(H) = %.1f, θ†(Hc) = %.2f, pH ≈ %d%%",
      sim_n_sample, nsims, dgm_hr_h1,
      dgm_alt$hr_Hc_true,
      round(100 * mean(dgm_alt$df_super$flag_harm))
    )
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#f0f7ff"),
    locations = gt::cells_body(
      rows = Source != "Paper"
    )
  ) |>
  gt::tab_options(table.font.size = gt::px(13))
```

### M1 Null: Paper vs Reproduction

```{r comparison-null-table}
null_df <- data.frame(
  Method      = rep(paper$Method, each = 2),
  Source      = rep(c("Paper", sprintf("Ours (%d)", nsims)), 4),
  `any(H)`    = c(rbind(fmt2(paper$null_any),     fmt2(our$null_any))),
  `sens(Hc)`  = c(rbind(fmt2(paper$null_sens_Hc), fmt2(our$null_sens_Hc))),
  `ppv(Hc)`   = c(rbind(fmt2(paper$null_ppv_Hc),  fmt2(our$null_ppv_Hc))),
  `avg|H|`    = c(rbind(fmt0(paper$null_avg_H),   fmt0(our$null_avg_H))),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt::gt(null_df, groupname_col = "Method") |>
  gt::tab_header(
    title = "M1 Null: Paper (20,000 sims) vs Reproduction",
    subtitle = sprintf(
      "N = %d, %d sims | θ†(ITT) ≈ %.2f",
      sim_n_sample, nsims, dgm_null$hr_causal
    )
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups()
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#fff5f0"),
    locations = gt::cells_body(
      rows = Source != "Paper"
    )
  ) |>
  gt::tab_options(table.font.size = gt::px(13))
```

### Summary of Discrepancies

```{r discrepancy-summary}
#| code-fold: true
#| code-summary: "Discrepancy analysis"

# Compute absolute differences for key metrics
disc <- data.frame(
  Method = paper$Method,
  stringsAsFactors = FALSE
)

disc$d_null_fpr   <- abs(our$null_any - paper$null_any)
disc$d_alt_any    <- abs(our$alt_any  - paper$alt_any)
disc$d_alt_sens   <- abs(our$alt_sens_H - paper$alt_sens_H)
disc$d_alt_ppv    <- abs(our$alt_ppv_H  - paper$alt_ppv_H)
disc$d_alt_sens_c <- abs(our$alt_sens_Hc - paper$alt_sens_Hc)
disc$d_alt_ppv_c  <- abs(our$alt_ppv_Hc  - paper$alt_ppv_Hc)

# Monte Carlo SE for a proportion at the paper's nsims
mc_se <- function(p, n) sqrt(p * (1 - p) / n)

cat("=== Discrepancy Analysis ===\n\n")
cat(sprintf("Paper: 20,000 sims  |  Ours: %d sims\n\n", nsims))

cat("--- Null (Type I Error) ---\n")
for (m in disc$Method) {
  d <- disc[disc$Method == m, ]
  p_val <- paper$null_any[paper$Method == m]
  se_ours <- mc_se(our$null_any[our$Method == m], nsims)
  cat(sprintf("  %-6s  |diff| = %.3f  (MC SE at %d sims ≈ %.3f)  %s\n",
      m, d$d_null_fpr, nsims, se_ours,
      if (d$d_null_fpr <= 2 * se_ours) "within 2 SE" else "** > 2 SE **"))
}

cat("\n--- Alternative (Detection & Classification) ---\n")
cat(sprintf("  %-6s  %8s  %8s  %8s  %8s\n",
    "", "|d|any(H)", "|d|sens", "|d|ppv", "Status"))
for (m in disc$Method) {
  d <- disc[disc$Method == m, ]
  se_any  <- mc_se(our$alt_any[our$Method == m], nsims)
  # For classification, conditional on detection
  n_det <- sum(results_alt$any.H[results_alt$analysis == m] == 1)
  se_sens <- if (n_det > 0) mc_se(our$alt_sens_H[our$Method == m], n_det) else NA
  max_d <- max(d$d_alt_any, d$d_alt_sens, d$d_alt_ppv, na.rm = TRUE)
  status <- if (max_d <= 0.10) "close" else if (max_d <= 0.20) "moderate" else "large"
  cat(sprintf("  %-6s  %8.3f  %8.3f  %8.3f  %8s\n",
      m, d$d_alt_any, d$d_alt_sens, d$d_alt_ppv, status))
}

cat(sprintf("\nNote: With %d sims, MC standard errors for proportions ≈ 0.03-0.05.\n",
    nsims))
cat("Discrepancies within ~0.10 are consistent with simulation variability.\n")
cat("Scale to nsims = 5000+ for tighter estimates.\n")
```


## Timing and Projections

```{r timing}
#| code-fold: false

# Per-replicate timing
sec_per_alt  <- t_alt / nsims
sec_per_null <- t_null / nsims
total_sec    <- t_alt + t_null

cat("=== Timing Summary ===\n\n")
cat(sprintf("  Alt simulations:  %d sims in %.1f min (%.1f sec/sim)\n",
    nsims, t_alt / 60, sec_per_alt))
cat(sprintf("  Null simulations: %d sims in %.1f min (%.1f sec/sim)\n",
    nsims, t_null / 60, sec_per_null))
cat(sprintf("  Total:            %.1f min\n", total_sec / 60))
cat(sprintf("  Workers:          %d\n\n", n_workers))

# Projections
for (n_target in c(1000, 2000, 5000, 20000)) {
  proj_alt  <- n_target * sec_per_alt
  proj_null <- n_target * sec_per_null
  proj_total <- proj_alt + proj_null
  cat(sprintf("  Projection for %5d sims: %.0f min (%.1f hours)\n",
      n_target, proj_total / 60, proj_total / 3600))
}

cat(sprintf("\n  Recommendation: nsims = 5000 provides MC SE ≈ 0.007 for\n"))
cat(sprintf("  proportions near 0.80, sufficient to distinguish all\n"))
cat(sprintf("  methods at the resolution of the paper's Table 1.\n"))
```


## Summary and Concordance {#summary}

```{r summary-concordance}
#| code-fold: true
#| code-summary: "Compute summary metrics"

# Helper: extract value for method m from our/paper
o <- function(m, col) our[[col]][our$Method == m]
p <- function(m, col) paper[[col]][paper$Method == m]
mc_se <- function(phat, n) sqrt(phat * (1 - phat) / n)

# Build concordance table
methods <- c("FSl", "FSlg", "GRF", "GRF60")
concordance <- data.frame(
  Method       = methods,
  Paper_any    = sapply(methods, p, "alt_any"),
  Ours_any     = sapply(methods, o, "alt_any"),
  Paper_sens   = sapply(methods, p, "alt_sens_H"),
  Ours_sens    = sapply(methods, o, "alt_sens_H"),
  Paper_ppv    = sapply(methods, p, "alt_ppv_H"),
  Ours_ppv     = sapply(methods, o, "alt_ppv_H"),
  Paper_FPR    = sapply(methods, p, "null_any"),
  Ours_FPR     = sapply(methods, o, "null_any"),
  stringsAsFactors = FALSE
)
concordance$d_any  <- abs(concordance$Ours_any - concordance$Paper_any)
concordance$d_sens <- abs(concordance$Ours_sens - concordance$Paper_sens)
concordance$d_ppv  <- abs(concordance$Ours_ppv - concordance$Paper_ppv)
concordance$max_d  <- pmax(concordance$d_any, concordance$d_sens, concordance$d_ppv)
concordance$grade  <- ifelse(concordance$max_d <= 0.05, "excellent",
                      ifelse(concordance$max_d <= 0.10, "close",
                      ifelse(concordance$max_d <= 0.15, "moderate", "investigate")))
```

### Alt: Detection and Classification (Unconditional)

```{r summary-alt-table}
fmt2 <- function(x) sprintf("%.2f", x)
fmt3 <- function(x) sprintf("%.3f", x)

summ_df <- data.frame(
  Method    = concordance$Method,
  `Paper any(H)` = fmt2(concordance$Paper_any),
  `Ours any(H)`  = fmt2(concordance$Ours_any),
  `|d|`          = fmt3(concordance$d_any),
  `Paper sens`   = fmt2(concordance$Paper_sens),
  `Ours sens`    = fmt2(concordance$Ours_sens),
  `|d| `         = fmt3(concordance$d_sens),
  `Paper ppv`    = fmt2(concordance$Paper_ppv),
  `Ours ppv`     = fmt2(concordance$Ours_ppv),
  `|d|  `        = fmt3(concordance$d_ppv),
  Grade          = concordance$grade,
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt::gt(summ_df) |>
  gt::tab_header(
    title = sprintf("Alt Concordance: Paper (20,000) vs Reproduction (%d sims)", nsims),
    subtitle = "Unconditional averaging (sens = 0 when undetected)"
  ) |>
  gt::tab_spanner(label = "Detection", columns = 2:4) |>
  gt::tab_spanner(label = "Sensitivity", columns = 5:7) |>
  gt::tab_spanner(label = "PPV", columns = 8:10) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#e8f5e9"),
    locations = gt::cells_body(rows = Grade %in% c("excellent", "close"))
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#fff3e0"),
    locations = gt::cells_body(rows = Grade == "moderate")
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#ffebee"),
    locations = gt::cells_body(rows = Grade == "investigate")
  ) |>
  gt::tab_options(table.font.size = gt::px(12))
```

### Null: Type I Error

```{r summary-null-table}
null_df <- data.frame(
  Method      = methods,
  `Paper FPR` = fmt2(concordance$Paper_FPR),
  `Ours FPR`  = fmt2(concordance$Ours_FPR),
  `|d|`       = fmt3(abs(concordance$Ours_FPR - concordance$Paper_FPR)),
  `MC SE`     = fmt3(sapply(methods, function(m)
    mc_se(o(m, "null_any"), sum(results_null$analysis == m)))),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt::gt(null_df) |>
  gt::tab_header(
    title = sprintf("Null Concordance: Paper (20,000) vs Reproduction (%d sims)", nsims)
  ) |>
  gt::tab_options(table.font.size = gt::px(12))
```


### Key Findings

```{r summary-findings}
#| results: asis

# Detection
best_det  <- concordance$Method[which.min(concordance$d_any)]
worst_det <- concordance$Method[which.max(concordance$d_any)]

# Classification
best_cls  <- concordance$Method[which.min(concordance$max_d)]
worst_cls <- concordance$Method[which.max(concordance$max_d)]

cat(sprintf("**Simulation budget**: %d sims (paper: 20,000). ", nsims))
cat(sprintf("MC SE ≈ %.3f for proportions near 0.80.\n\n",
    mc_se(0.80, nsims)))

cat(sprintf("**Best overall concordance**: %s (max |d| = %.3f, grade: %s).\n\n",
    best_cls,
    concordance$max_d[concordance$Method == best_cls],
    concordance$grade[concordance$Method == best_cls]))

cat(sprintf("**Largest discrepancy**: %s (max |d| = %.3f, grade: %s).\n\n",
    worst_cls,
    concordance$max_d[concordance$Method == worst_cls],
    concordance$grade[concordance$Method == worst_cls]))

cat("**`dmin.grf` calibration**: DR scores from `causal_survival_forest` ")
cat("are anti-symmetric ($\\Gamma_{\\text{control}} = -\\Gamma_{\\text{treated}}$), ")
cat("so the policy tree leaf difference equals 2× the raw RMST effect. ")
cat("A \"6-month RMST benefit\" → `dmin.grf = 12`.\n\n")

cat("**Averaging convention**: The paper reports **unconditional** ")
cat("classification metrics (sens = 0 when $\\hat{H} = \\emptyset$ under alt). ")
cat(sprintf("This matters most for GRF60 (detection = %.0f%%) where the ",
    100 * o("GRF60", "alt_any")))
cat("conditional/unconditional gap is largest.\n\n")

cat(sprintf("**Timing**: %.1f sec/sim (alt), %.1f sec/sim (null) on %d workers. ",
    t_alt / nsims, t_null / nsims, n_workers))
cat(sprintf("Projected for 20,000 sims: %.1f hours.\n",
    (20000 * (t_alt + t_null) / nsims) / 3600))
```


## Pressure Test: Null Classification

Under H0, `dgm$model == "null"` means Q = ∅ (León et al.,
Section 3).  With the DGM fix, `flag_harm = 0` for all
subjects, so `sens = NA` and `ppv = 0` naturally.

```{r null-classification-check}
#| code-fold: false

cat("=== Null Classification Pressure Test ===\n\n")
cat("dgm_null$model:", dgm_null$model, "\n\n")

for (m in c("FSl", "FSlg", "GRF", "GRF60")) {
  rn <- results_null[results_null$analysis == m, ]
  rn_det <- rn[rn$any.H == 1, ]
  n_det <- nrow(rn_det)

  cat(sprintf("%-6s  FPR=%.2f (%d/%d)\n", m,
      mean(rn$any.H, na.rm = TRUE), n_det, nrow(rn)))

  if (n_det > 0) {
    cat(sprintf("  sens: all NA? %s  (expected: TRUE)\n",
        all(is.na(rn_det$sens))))
    cat(sprintf("  ppv:  all 0?  %s  (expected: TRUE)\n",
        all(rn_det$ppv == 0, na.rm = TRUE)))
    cat(sprintf("  spec: mean=%.3f  (= avg |Hc_hat|/N)\n",
        mean(rn_det$spec, na.rm = TRUE)))
    cat(sprintf("  npv:  all 1?  %s  (expected: TRUE)\n",
        all(rn_det$npv == 1, na.rm = TRUE)))
  } else {
    cat("  No false positives — classification not computed\n")
  }
  cat("\n")
}
```


## Session Info

```{r session}
cat(sprintf("Total runtime: %.1f minutes\n", (t_alt + t_null) / 60))
sessionInfo()
```
