---
title: "The Pareto frontier in forestsearch subgroup selection"
subtitle: "How `selection_rule` shapes the candidate inclusion band for `sg_focus = \"effMaxSG\"`/`\"effMinSG\"`"
author: "forestsearch v0.2.0"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
---

# Purpose

`forestsearch()` (and, since v0.2.0, `dina_subgroup()`) selects among the
qualifying candidate subgroups using `sg_focus`. For the effect-anchored
foci `"effMaxSG"`/`"effMinSG"` (canonical `"hrMaxSG"`/`"hrMinSG"`), the
selection is *not* "take the candidate with the largest effect." Instead
the candidates are first partitioned into a **candidate inclusion band**,
and then ranked within that band by sample size. The `selection_rule`
argument controls how the band is drawn:

* `"neighborhood"` --- a one-dimensional effect band (default).
* `"pareto"` --- the two-dimensional Pareto-non-dominated set in
  (effect, $N$) space.
* `"both"` --- the intersection of the two (strictest).

This note explains how the **Pareto frontier** operates: what
"dominated" means, how the frontier becomes the `in_pareto` indicator,
how it combines with the neighborhood band, and how the resulting
`in_band` flag feeds the final selection sort. The single computation
shared by every path is `.compute_inclusion_band()`, called by both
`sort_subgroups()` (final, post-consistency selection) and
`sort_subgroups_preview()` (pre-consistency triage).

# Setup

```{r setup}
library(data.table)
library(ggplot2)
library(gt)

# Loaded for the validation section; the figure itself is self-contained.
have_fs <- requireNamespace("forestsearch", quietly = TRUE)

EFFECT_NEIGHBORHOOD <- 0.10   # the package default
```

# The dominance rule

Both objectives --- effect size and sample size --- are *maximized*. A
candidate $i$ is **dominated** if some other candidate $j$ is at least as
good on both axes and strictly better on at least one:

$$
\text{dominated}(i) \;\Longleftrightarrow\;
\exists\, j \neq i:\;
\mathrm{hr}_j \ge \mathrm{hr}_i \;\wedge\; N_j \ge N_i
\;\wedge\; (\mathrm{hr}_j > \mathrm{hr}_i \;\vee\; N_j > N_i).
$$

The **Pareto frontier** is the set of non-dominated candidates --- the
upper-right boundary of the cloud. The package implements this as an
$O(n^2)$ scan in `.pareto_dominated_xy()`; candidates with a missing
effect or size are treated as dominated. The function below is a
faithful mirror of that internal helper, kept local so this document
renders without the package; the validation section confirms the two
agree.

```{r dominance}
# Mirror of forestsearch internal .pareto_dominated_xy(): both axes maximized.
dominated_xy <- function(x, y) {
  n <- length(x)
  dom <- logical(n)
  for (i in seq_len(n)) {
    if (is.na(x[i]) || is.na(y[i])) { dom[i] <- TRUE; next }
    for (j in seq_len(n)) {
      if (i == j) next
      if (is.na(x[j]) || is.na(y[j])) next
      if (x[j] >= x[i] && y[j] >= y[i] && (x[j] > x[i] || y[j] > y[i])) {
        dom[i] <- TRUE
        break
      }
    }
  }
  dom
}
```

# How the band is formed

`.compute_inclusion_band()` builds two indicators and multiplies them:

```r
in_nbhd   <- (hr >= (1 - effect_neighborhood) * max(hr))   # 1-D effect band
in_pareto <- !dominated_xy(hr, N)                          # 2-D frontier
in_band   <- in_nbhd * in_pareto
```

Each `selection_rule` switches the two components on or off:

| `selection_rule` | `in_nbhd` | `in_pareto` | `in_band` is |
|---|---|---|---|
| `"neighborhood"` | active | all ones | the effect band only |
| `"pareto"` | all ones | active | the frontier only |
| `"both"` | active | active | their intersection |

For ratio measures stored on the log scale (GLM OR/IRR), `hr` is
exponentiated before either test so the comparison is on the natural
scale; Cox HR is already natural. The neighborhood floor is a *relative*
tolerance: with `effect_neighborhood = 0.10` a candidate is in-band iff
its effect is within 10% of the strongest effect among the survivors.
Setting it to `0` collapses the band to a strict max-effect filter.

# A worked candidate set

Five candidates that have already passed the consistency stage and the
effect threshold. Effect on the natural (HR) scale, size $N$:

```{r candidates}
cands <- data.table(
  id = c("A", "B", "C", "E", "F"),
  hr = c(2.50, 2.40, 2.30, 2.35, 2.00),
  N  = c(70L,  100L, 150L, 90L,  200L)
)

hr_max  <- max(cands$hr)
hr_floor <- (1 - EFFECT_NEIGHBORHOOD) * hr_max

cands[, in_nbhd     := hr >= hr_floor]
cands[, dominated   := dominated_xy(hr, N)]
cands[, on_frontier := !dominated]
cands[, in_both     := in_nbhd & on_frontier]
cands[, status      := fifelse(!in_nbhd, "neither (below floor)",
                        fifelse(in_both,  "both rules",
                                          "neighborhood only"))]

# effMaxSG within the "both" band: largest N among in-band-both candidates.
sel_id <- cands[in_both == TRUE][order(-N)][1L, id]
cands[, selected := id == sel_id]

cands[, .(id, hr, N, in_nbhd, on_frontier, in_both, selected)]
```

With `hr_max = 2.5`, the floor is `r round(hr_floor, 3)`. So:

* `A`, `B`, `C` are in the effect band **and** on the frontier --- in-band
  under every rule.
* `E` (HR 2.35, $N$ 90) is in the effect band but **dominated by `B`**
  (HR 2.40 $\ge$ 2.35 and $N$ 100 $\ge$ 90), so `"pareto"`/`"both"`
  exclude it while `"neighborhood"` keeps it.
* `F` (HR 2.00, $N$ 200) is on the frontier (largest $N$) but **below the
  floor**, so `"neighborhood"`/`"both"` exclude it while `"pareto"`
  keeps it --- this is the over-extension guard that the effect band
  provides and the bare frontier does not.

# The figure

```{r}
#| label: fig-frontier
#| fig-width: 7.2
#| fig-height: 4.6
#| fig-cap: "Candidate subgroups in the (effect, N) plane. The shaded region is the effect neighborhood band (HR at least the dashed floor); the dashed line traces the Pareto frontier. Colour encodes which selection_rule keeps each candidate. C is selected by effMaxSG (largest N in the band)."

pal <- c("both rules"           = "#1D9E75",
         "neighborhood only"    = "#EF9F27",
         "neither (below floor)" = "#888780")

frontier <- cands[on_frontier == TRUE][order(hr)]

# per-point label offsets so the letters sit clear of points/lines
labs <- data.table(
  id = c("A", "B", "C", "E", "F"),
  lx = c(2.50, 2.435, 2.30, 2.35, 2.00),
  ly = c(63,   108,   166,  79,   210))

ggplot(cands, aes(hr, N)) +
  annotate("rect", xmin = hr_floor, xmax = Inf, ymin = -Inf, ymax = Inf,
           fill = "#378ADD", alpha = 0.10) +
  geom_vline(xintercept = hr_floor, linetype = "dashed",
             colour = "#185FA5", linewidth = 0.4) +
  geom_line(data = frontier, aes(hr, N), linetype = "dashed",
            colour = "#888780", linewidth = 0.5) +
  geom_point(data = cands[selected == TRUE], aes(hr, N),
             shape = 21, size = 9, stroke = 1.3,
             colour = "#0F6E56", fill = NA) +
  geom_point(aes(colour = status), size = 3.6) +
  geom_text(data = labs, aes(lx, ly, label = id),
            size = 4, fontface = "bold") +
  annotate("text", x = hr_floor + 0.015, y = 219,
           label = sprintf("HR >= %.2f  (within %.0f%% of max)",
                           hr_floor, 100 * EFFECT_NEIGHBORHOOD),
           hjust = 0, vjust = 1, size = 3.1, colour = "#185FA5") +
  scale_colour_manual(values = pal, name = "in_band under") +
  scale_x_continuous(limits = c(1.9, 2.62)) +
  scale_y_continuous(limits = c(45, 222)) +
  labs(x = "effect (HR)", y = "subgroup size N") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        panel.grid.minor = element_blank())
```

Reading the figure: `"neighborhood"` keeps the whole shaded strip
$\{A, B, C, E\}$; `"both"` additionally drops `E` because it sits
*inside* the frontier (another point beats it on both axes); `"pareto"`
alone would keep `F` (top-left, on the frontier) despite its weak
effect. The selection then maximizes $N$ within the kept set.

# From inclusion to selection

`sort_subgroups()` orders the candidates by a five-key lexicographic
sort and takes the top row. For `"hrMaxSG"` the key is

$$
(-\,\texttt{in\_band},\; -N,\; -\texttt{Pcons},\; -\texttt{hr},\; K),
$$

so: in-band candidates first; then the largest $N$; then higher
consistency; then higher effect; then fewer factors. `"hrMinSG"` is
identical except $N$ is minimized. Only `in_band` depends on
`selection_rule` --- the sort itself is fixed.

For the worked set, every rule that keeps `C` selects it (largest $N$ in
the band), so the **top pick is `C`** under `"neighborhood"` and
`"both"` alike; the rules differ only in whether `E` carries an in-band
flag in the diagnostic table.

## Where the rules genuinely diverge

The top selection differs only in a narrow corner: when two candidates
**tie on the maximum band $N$**, because the sort breaks $N$-ties by
consistency *before* effect, whereas Pareto dominance looks only at
(effect, $N$).

```{r corner}
corner <- data.table(
  id    = c("T", "D"),
  Pcons = c(0.95, 0.80),
  hr    = c(2.30, 2.45),
  N     = c(150L, 150L),
  K     = c(1L,   1L)
)
corner[]
```

`D` dominates `T` (equal $N$, higher HR). Under `"neighborhood"` both are
in-band, they tie on $N$, and the consistency tier keeps the
higher-`Pcons` candidate `T`. Under `"both"`, `T` is off the frontier and
excluded, so the higher-effect `D` is selected.

```{r corner-eval}
if (have_fs) {
  top_pick <- function(rule) {
    s <- forestsearch::sort_subgroups(
      data.table::copy(corner), "hrMaxSG",
      selection_rule = rule, effect_neighborhood = EFFECT_NEIGHBORHOOD)
    s$id[1L]
  }
  data.table(selection_rule = c("neighborhood", "both"),
             top_pick = vapply(c("neighborhood", "both"), top_pick, character(1)))
} else {
  cat("forestsearch not installed; expected: neighborhood -> T, both -> D.\n")
}
```

So `"both"` refuses to settle on a point another candidate beats on both
axes, even when it won the consistency tie; `"neighborhood"` will keep
the higher-consistency point.

# Validation against the package

Confirm the local frontier and band computations agree with the
shipped internals, and that `compute_pareto_frontier()` returns the
expected non-dominated rows.

```{r validate}
if (have_fs) {
  .ok <- function(lbl, cond)
    cat(if (isTRUE(cond)) "PASS" else "FAIL", "-", lbl, "\n")

  dt_hn <- cands[, .(hr, N)]

  pkg_dom <- forestsearch:::pareto_dominated_flags(dt_hn)
  .ok("dominated flags match .pareto_dominated_xy mirror",
      identical(pkg_dom, cands$dominated))

  pkg_nbhd <- forestsearch:::.compute_inclusion_band(
    cands$hr, cands$N, "neighborhood", EFFECT_NEIGHBORHOOD)
  .ok("neighborhood band matches",
      identical(as.logical(pkg_nbhd), cands$in_nbhd))

  pkg_both <- forestsearch:::.compute_inclusion_band(
    cands$hr, cands$N, "both", EFFECT_NEIGHBORHOOD)
  .ok("both band matches in_nbhd & on_frontier",
      identical(as.logical(pkg_both), cands$in_both))

  fr <- forestsearch:::compute_pareto_frontier(dt_hn)
  .ok("compute_pareto_frontier returns the non-dominated set",
      setequal(fr$hr, cands[on_frontier == TRUE, hr]))
} else {
  cat("forestsearch not installed; validation skipped.\n")
}
```

# Practical guidance

* `"neighborhood"` (default) asks only "is the effect close to the
  best?" then maximizes $N$. It can keep a high-consistency candidate
  that another point beats on both effect and size.
* `"pareto"` removes the `effect_neighborhood` tuning parameter
  entirely and selects on the (effect, $N$) trade-off alone --- but
  without the effect floor it will admit large-$N$, weak-effect points
  (like `F`) that the band is designed to exclude.
* `"both"` is the conservative choice: it keeps the effect floor *and*
  refuses dominated points. The in-band set under `"both"` is always a
  subset of the one under `"neighborhood"`.

The frontier here is an inclusion filter, not a stand-alone selection
criterion: the chosen subgroup still comes from the five-key sort, and
under `"hrMinSG"` the selected point may deliberately be $N$-dominated
(small subgroups are preferred). For a post-hoc *reporting* view of the
frontier independent of `sg_focus`, use `compute_pareto_frontier()` /
`pareto_frontier_table()`.

## Relationship to `dina_subgroup()`

The DINA `effMaxSG`/`effMinSG` foci reuse the same
`.compute_inclusion_band()` core, but only with
`selection_rule = "neighborhood"` --- DINA's univariate threshold
candidates form a simple effect-versus-size trade-off where the 1-D band
suffices. The `"pareto"`/`"both"` frontier variants are forestsearch-only
today; extending them to DINA would be a deliberate addition.

# Session information

```{r session-info}
sessionInfo()
```
