---
title: "Selection criteria in forestsearch"
subtitle: "`sg_focus`, `selection_rule`, and the candidate inclusion band"
author: "forestsearch v0.2.0"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
bibliography: []
---

```{r setup}
#| code-fold: false
suppressPackageStartupMessages({
  library(data.table)
  library(ggplot2)
})
have_fs <- requireNamespace("forestsearch", quietly = TRUE)

EFFECT_NEIGHBORHOOD <- 0.10   # package default band half-width
```

# Purpose

`forestsearch()` enumerates candidate subgroups, screens them, and returns one.
This document is about the **last** of those three steps: given a set of
surviving candidates, which one is returned, and why.

Two arguments control it.

* **`sg_focus`** selects the ranking rule. Eight values are legal; they differ
  in what they optimise and in whether consistency enters the ranking at all.
* **`selection_rule`** shapes a *candidate inclusion band* --- but only for two
  of the eight foci. Everywhere else it is inert, and supplying a non-default
  value is an error rather than a silent no-op.

Everything below is stated against the shipped implementation:
`sort_subgroups()` in `R/subgroup_consistency_helpers.R`,
`.compute_inclusion_band()` and `.normalize_sg_focus()` in
`R/forestsearch_helpers.R`, and the override blocks in `R/forestsearch_main.R`.
Where a claim can be checked against the package, @sec-validate checks it.

## The identifier is the estimator

A recurring source of confusion is worth settling first. In this framework the
**identifier** --- the rule that picks a subgroup --- *is* the estimator. FB
(full bootstrap) and MR (multiplier resampling) are inference procedures applied
to that estimator; they correct its selection bias. They do not identify a
subgroup, and they do not change the one that was identified.

The identifier has two parts:

| part | argument | values |
|---|---|---|
| **engine** --- how candidates are generated and screened | `subgroup_method` | `"consistency"`, `"dina"`, `"grf"` |
| **focus** --- which survivor is returned | `sg_focus` | the eight below |

Both parts must be recorded to describe a run. Two analyses that differ in
`sg_focus` are estimating different things, however similar the output looks.

# Vocabulary: `eff*` and `hr*` are the same rules {#sec-vocab}

The internal names are Cox-flavoured (`hr` = hazard ratio). The `eff*` spellings
generalise to GLM outcomes --- binary/OR, continuous/MD, count/IRR --- where
"hazard ratio" is meaningless. `.normalize_sg_focus()` collapses them before
anything keys on the value:

```{r aliases}
#| code-fold: false
normalize <- function(x)
  switch(x, effMaxSG = "hrMaxSG", effMinSG = "hrMinSG",
            eff = "hr", maxcons = "hr", x)

vocab <- data.table(
  supplied  = c("eff", "maxcons", "hr", "effMaxSG", "hrMaxSG",
                "effMinSG", "hrMinSG", "maxeff", "maxeffCons", "maxSG", "minSG"))
vocab[, canonical := vapply(supplied, normalize, character(1))]
vocab[, note := c("alias", "alias (preferred: names the rule)", "canonical",
                  "alias (preferred: GLM-neutral)", "canonical",
                  "alias (preferred: GLM-neutral)", "canonical",
                  "canonical", "canonical", "canonical", "canonical")]
vocab[]
```

Three pairs are the same rule under two spellings. **Prefer the `eff*` and
`maxcons` spellings**; they say what the rule does and they survive the move to
GLM outcomes.

::: {.callout-warning}
## `maxSG` is not `effMaxSG`

`maxSG`/`minSG` are **not** spellings of `effMaxSG`/`effMinSG`. They rank by
size across *all* qualifiers with no band at all. The banded foci first restrict
to an inclusion band and only then rank by size. Different rules, similar names.
:::

# Where selection sits in the pipeline

```{r pipeline}
#| code-fold: true
data.table(
  stage = c("1. Generation", "2. Admission", "3. Selection"),
  what  = c("Enumerate covariate-cut conjunctions, in (-effect, K) order",
            "Apply size/event minima, the effect floor, and the consistency screen",
            "Rank the survivors by sg_focus; return the top row"),
  governed_by = c("maxk, cont.cutoff, conf_force, collapse_cuts, use_lasso/use_grf",
                  "n.min, d0.min, d1.min, hr.threshold, pconsistency.threshold",
                  "sg_focus, selection_rule, effect_neighborhood"))[]
```

Two facts about stage 1 matter later:

1. **Enumeration is in descending-effect order**, ties broken by `K` (the number
   of conjuncts). Written as a key: $(-\mathrm{HR},\, K)$.
2. **Admission is per candidate.** A candidate that fails the consistency screen
   never reaches stage 3 --- `evaluate_subgroup_consistency()` returns `NULL`
   below `pconsistency.threshold`. The screen is a *filter*, not a term in any
   ranking.

# The eight foci {#sec-foci}

`sort_subgroups()` applies a lexicographic sort and takes the first row. The
keys, verbatim from the implementation:

```{r keys}
#| code-fold: true
data.table(
  focus = c("maxcons (hr)", "maxeff", "maxeffCons", "maxSG", "minSG",
            "effMaxSG (hrMaxSG)", "effMinSG (hrMinSG)"),
  sort_key = c("(-Pcons, -hr, K)", "(-hr, K)", "(-hr, K)",
          "(-N, -Pcons, K)", "(N, -Pcons, K)",
          "(-in_band, -N, -Pcons, -hr, K)", "(-in_band, N, -Pcons, -hr, K)"),
  ranges_over = c("all qualifiers", "ALL candidates, ungated",
                  "consistency-qualifying candidates only",
                  "all qualifiers", "all qualifiers",
                  "all qualifiers", "all qualifiers"),
  band = c("no", "no", "no", "no", "no", "yes", "yes"))[]
```

`K` is always the final tiebreak and resolves exact ties toward the simpler
rule. It can never override the keys before it.

## `maxcons` --- consistency argmax {#sec-maxcons}

Key $(-\texttt{Pcons},\, -\texttt{hr},\, K)$. Consistency is primary; effect
breaks ties. This is the rule León et al. (2024) use.

Because enumeration is by *effect* and the key is led by *consistency*, the scan
order carries no information about the ranking. A candidate far down the
enumeration --- weaker effect --- can carry the highest `Pcons` and win. **Every
candidate must be evaluated**; nothing can be concluded early.

## `maxeff` --- unconditional effect argmax {#sec-maxeff}

Key $(-\texttt{hr},\, K)$ over *all* candidates. This focus is unusual enough to
deserve its own treatment (@sec-maxeff-detail): it applies **no thresholds at
all**, and it always returns a subgroup.

## `maxeffCons` --- effect argmax among qualifiers {#sec-maxeffcons}

Key $(-\texttt{hr},\, K)$ --- *identical* to `maxeff`. The difference is not in
the key but in the population it ranges over: only candidates that cleared
`pconsistency.threshold` reach the sort. The consistency condition is applied by
construction, upstream, and therefore needs no term in the key.

This is the only focus for which **early stopping is sound** (@sec-stopping).

## `maxSG` / `minSG` --- size extremes

Keys $(-N,\, -\texttt{Pcons},\, K)$ and $(N,\, -\texttt{Pcons},\, K)$. Largest
or smallest qualifier, consistency as tiebreak. No band, no effect term until
`K`.

## `effMaxSG` / `effMinSG` --- banded size extremes

Keys $(-\texttt{in\_band},\, \mp N,\, -\texttt{Pcons},\, -\texttt{hr},\, K)$.
In-band candidates first, then size, then consistency, then effect. Only
`in_band` depends on `selection_rule`; the rest of the key is fixed. These are
the subject of @sec-band onward.

# `maxeff` in detail {#sec-maxeff-detail}

`forestsearch()` treats `maxeff` as a special case and overrides five settings,
warning as it does so:

```{r maxeff-overrides}
#| code-fold: true
data.table(
  setting = c("pconsistency.threshold", "stop_threshold", "use_twostage",
              "max_subgroups_search", "minp", "rmin", "hr.threshold",
              "n.min / d0.min / d1.min"),
  under_maxeff = c("-> 0", "-> NULL", "-> FALSE", "-> Inf", "-> 0", "-> 0",
                   "disabled (disable_effect_floor = TRUE)", "RETAINED"),
  why = c("screen admits everything", "stopping unsound for this key",
          "no two-stage narrowing", "no truncation of the pool",
          "prevalence floor relaxed", "redundancy pruning relaxed",
          "no effect floor: pure argmax",
          "define which candidates are ESTIMABLE, not which one wins"))[]
```

Two consequences follow.

**Nothing can reject the winner.** There is no effect threshold to fail and no
consistency condition to fail. Consistency *is* computed for the winner
afterwards and reported, but descriptively --- it is not a filter. So
`prop_SG_found = 100%` is **structural**, not an empirical finding.

**The guarantee is conditional on a non-empty candidate family.** The size and
event minima are deliberately retained, and the source is explicit about the
reasoning: they define which candidates are estimable, not which one wins. Only
degenerate candidates --- empty cuts, exact duplicates --- are pruned, and those
share a retained twin's effect, so the argmax is unchanged. Dropping distinct
near-duplicates would have forced $(-\mathrm{hr}, K)$ to break genuine ties on
cut count, an arbitrary tiebreak in a rule designed to have none.

So `maxeff` returns empty only if no conjunction is estimable at all --- a data
condition, not a selection outcome.

::: {.callout-note}
## When `maxeff` is the right choice

`maxeff` is the argmax primitive that post-selection inference methods such as
Guo & He (2021) are defined against. Use it when the estimand is "the effect in
the subgroup that maximises the estimated effect," accepting that the subgroup
may be unstable. Do not use it when replicability is part of the claim --- it
can return a subgroup whose consistency is arbitrarily low.
:::

# Early stopping and why only one focus qualifies {#sec-stopping}

`stop_threshold` halts the consistency scan at the first candidate to reach it.
That is sound only if the scan order agrees with the ranking.

```{r stopping}
#| code-fold: true
data.table(
  focus = c("maxeffCons", "maxeff", "maxcons (hr)", "maxSG", "minSG",
            "effMaxSG", "effMinSG"),
  sort_key = c("(-hr, K)", "(-hr, K)", "(-Pcons, -hr, K)", "(-N, -Pcons, K)",
          "(N, -Pcons, K)", "(-in_band, ...)", "(-in_band, ...)"),
  matches_enumeration = c("YES", "n/a (winner-only path)", "no", "no", "no",
                          "no", "no"),
  stopping = c("SOUND, batch size 1", "not reached",
               rep("reset to NULL by forestsearch()", 5)))[]
```

The argument, from `subgroup_consistency_main.R`: `maxeffCons`'s key
$(-\mathrm{hr}, K)$ *is* the enumeration order $(-\mathrm{HR}, K)$, ties
included. So the first candidate to clear the threshold is provably the winner
--- everything unexamined has weaker or equal effect. Batch size drops to `1L`
to halt at the earliest sound point.

Any key with a `Pcons` term --- primary as in `maxcons`, or merely a size
tiebreak as in `maxSG`/`minSG` --- can be beaten by an unevaluated candidate.
`forestsearch()` therefore resets `stop_threshold` to `NULL` for
`c("hrMaxSG", "hrMinSG", "hr", "maxSG", "minSG")`, and **warns only when the
value was supplied explicitly**. Inherited defaults are reset silently, which is
worth knowing: a configuration can have early stopping disabled without saying
so anywhere.

::: {.callout-important}
## "Must evaluate every candidate" and "can stop early" are not in tension

`maxeffCons` gets no shortcut past consistency evaluation --- it must score
candidates one at a time, because it needs to know which qualify. What it gets
is permission to *halt* at the first qualifier.

`maxeff` is the opposite: it skips straight to the effect argmax and scores
consistency for that candidate alone (the winner-only fast path, worth roughly
20x), but it can return a subgroup that fails consistency entirely.
:::

# The inclusion band {#sec-band}

For `effMaxSG`/`effMinSG` only, `.compute_inclusion_band()` builds two
indicators and multiplies them:

```r
in_nbhd   <- (hr >= (1 - effect_neighborhood) * max(hr))   # 1-D effect band
in_pareto <- !dominated_xy(hr, N)                          # 2-D frontier
in_band   <- in_nbhd * in_pareto
```

`selection_rule` switches the components on or off:

| `selection_rule` | `in_nbhd` | `in_pareto` | `in_band` is |
|---|---|---|---|
| `"neighborhood"` (default) | active | all ones | the effect band only |
| `"pareto"` | all ones | active | the frontier only |
| `"both"` | active | active | their intersection |

`in_nbhd` is a **relative** tolerance: at `effect_neighborhood = 0.10` a
candidate is in-band iff its effect is within 10% of the strongest effect among
the survivors. Setting it to `0` collapses the band to a strict max-effect
filter. For ratio measures stored on the log scale (OR, IRR), both tests are
applied after exponentiating, so comparisons are on the natural scale; Cox HR is
already natural.

## Pareto dominance

Candidate $i$ is **dominated** when some $j$ is at least as good on both axes
and strictly better on one:

$$
\exists\, j \neq i:\quad
\mathrm{hr}_j \ge \mathrm{hr}_i \;\wedge\; N_j \ge N_i
\;\wedge\; (\mathrm{hr}_j > \mathrm{hr}_i \;\vee\; N_j > N_i).
$$

The non-dominated set is the **frontier**: candidates for which no alternative
is better on effect without being worse on size, or vice versa.

```{r dominance}
#| code-fold: false
# Mirror of the package internal .pareto_dominated_xy(); both axes maximised.
dominated_xy <- function(x, y) {
  n <- length(x)
  dom <- logical(n)
  for (i in seq_len(n)) {
    if (is.na(x[i]) || is.na(y[i])) { dom[i] <- TRUE; next }
    for (j in seq_len(n)) {
      if (i == j) next
      if (is.na(x[j]) || is.na(y[j])) next
      if (x[j] >= x[i] && y[j] >= y[i] && (x[j] > x[i] || y[j] > y[i])) {
        dom[i] <- TRUE
        break
      }
    }
  }
  dom
}
```

`NA` on either axis is treated as dominated --- an unscorable candidate cannot
be on the frontier.

# Worked example: the band {#sec-worked}

Five candidates that have already cleared consistency and the effect floor.

```{r candidates}
cands <- data.table(
  id = c("A", "B", "C", "E", "F"),
  hr = c(2.50, 2.40, 2.30, 2.35, 2.00),
  N  = c(70L,  100L, 150L, 90L,  200L))

hr_max   <- max(cands$hr)
hr_floor <- (1 - EFFECT_NEIGHBORHOOD) * hr_max

cands[, in_nbhd     := hr >= hr_floor]
cands[, dominated   := dominated_xy(hr, N)]
cands[, on_frontier := !dominated]
cands[, in_both     := in_nbhd & on_frontier]
cands[, .(id, hr, N, in_nbhd, on_frontier, in_both)]
```

With `hr_max = 2.5` the floor is `r round(hr_floor, 3)`:

* **`A`, `B`, `C`** are in the effect band *and* on the frontier --- in-band
  under every rule.
* **`E`** (HR 2.35, $N$ 90) is in the effect band but **dominated by `B`**
  (HR 2.40 $\ge$ 2.35 and $N$ 100 $\ge$ 90). `"pareto"` and `"both"` exclude it;
  `"neighborhood"` keeps it.
* **`F`** (HR 2.00, $N$ 200) is on the frontier --- nothing has larger $N$ ---
  but **below the floor**. `"neighborhood"` and `"both"` exclude it; `"pareto"`
  keeps it. This is the over-extension guard the effect band provides and the
  bare frontier does not.

```{r band-figure}
#| code-fold: true
#| fig-height: 4.2
pl <- copy(cands)
pl[, status := fifelse(!in_nbhd, "below floor (pareto only)",
                fifelse(in_both, "in band under all rules",
                                 "neighborhood only (dominated)"))]
ggplot(pl, aes(N, hr)) +
  geom_hline(yintercept = hr_floor, linetype = 2, colour = "grey45") +
  geom_step(data = pl[on_frontier == TRUE][order(N)],
            direction = "vh", colour = "grey70", linewidth = 0.4) +
  geom_point(aes(colour = status), size = 3.6) +
  geom_text(aes(label = id), nudge_y = 0.045, size = 3.6) +
  annotate("text", x = max(pl$N), y = hr_floor, vjust = -0.6, hjust = 1,
           label = sprintf("floor = %.3f", hr_floor), size = 3, colour = "grey40") +
  scale_colour_manual(values = c("in band under all rules" = "#2c7fb8",
                                 "neighborhood only (dominated)" = "#d95f02",
                                 "below floor (pareto only)" = "#7570b3")) +
  labs(x = "Subgroup size N", y = "Effect (HR)", colour = NULL,
       title = "Effect band, Pareto frontier, and their intersection") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

Under every rule that retains `C`, `C` wins --- it has the largest $N$ in the
band. So the **top pick is `C`** under `"neighborhood"` and `"both"` alike. The
rules differ here only in which candidates carry an in-band flag in the
diagnostic table.

## Where the rules genuinely diverge {#sec-corner}

Top selection differs only in a narrow corner: when two candidates **tie on the
maximum band $N$**. The sort breaks $N$-ties by consistency *before* effect,
whereas Pareto dominance looks only at (effect, $N$).

```{r corner}
corner <- data.table(
  id    = c("T", "D"),
  Pcons = c(0.95, 0.80),
  hr    = c(2.30, 2.45),
  N     = c(150L, 150L),
  K     = c(1L,   1L))
corner[]
```

`D` dominates `T` --- equal $N$, higher effect. Under `"neighborhood"` both are
in-band, they tie on $N$, and the consistency tier keeps `T`. Under `"both"`,
`T` is off the frontier and excluded, so `D` wins.

```{r corner-eval}
if (have_fs) {
  top_pick <- function(rule) {
    s <- forestsearch::sort_subgroups(
      data.table::copy(corner), "hrMaxSG",
      selection_rule = rule, effect_neighborhood = EFFECT_NEIGHBORHOOD)
    s$id[1L]
  }
  data.table(selection_rule = c("neighborhood", "both"),
             top_pick = vapply(c("neighborhood", "both"), top_pick, character(1)))
} else {
  cat("forestsearch not installed; expected: neighborhood -> T, both -> D.\n")
}
```

So `"both"` refuses to settle on a point another candidate beats on both axes,
even when it won the consistency tie. `"neighborhood"` keeps the
higher-consistency point.

# Worked example: the foci disagree {#sec-foci-example}

The band example holds `sg_focus` fixed. This one holds the candidate set fixed
and varies the focus, on a set constructed so that every rule picks differently.

```{r foci-set}
set <- data.table(
  id    = c("P", "Q", "R", "S"),
  hr    = c(3.10, 2.60, 2.55, 1.40),
  N     = c( 40L, 120L, 130L, 300L),
  Pcons = c(0.55, 0.97, 0.93, 0.99),
  K     = c(  2L,   2L,   1L,   1L))
set[]
```

Read the columns as competing claims. `P` has the strongest effect on the
smallest subgroup with poor replicability. `Q` and `R` are the middle ground.
`S` is large, highly replicable, and weak.

```{r foci-apply}
#| code-fold: false
pick <- function(dt, focus, p_star = 0.90,
                 rule = "neighborhood", nbhd = EFFECT_NEIGHBORHOOD) {
  d <- copy(dt)
  # Stage 2: the consistency screen is a FILTER, applied before ranking.
  # maxeff is the exception -- it overrides p_star to 0.
  qual <- if (focus == "maxeff") d else d[Pcons >= p_star]
  if (!nrow(qual)) return(NA_character_)

  if (focus %in% c("maxeff", "maxeffCons")) {
    setorder(qual, -hr, K)
  } else if (focus %in% c("maxcons", "hr")) {
    setorder(qual, -Pcons, -hr, K)
  } else if (focus == "maxSG") {
    setorder(qual, -N, -Pcons, K)
  } else if (focus == "minSG") {
    setorder(qual, N, -Pcons, K)
  } else if (focus %in% c("effMaxSG", "effMinSG")) {
    nb  <- qual$hr >= (1 - nbhd) * max(qual$hr)
    pa  <- if (rule == "neighborhood") rep(TRUE, nrow(qual)) else !dominated_xy(qual$hr, qual$N)
    nb  <- if (rule == "pareto") rep(TRUE, nrow(qual)) else nb
    qual[, in_band := as.integer(nb & pa)]
    ord <- if (focus == "effMaxSG")
      order(-qual$in_band, -qual$N, -qual$Pcons, -qual$hr, qual$K) else
      order(-qual$in_band,  qual$N, -qual$Pcons, -qual$hr, qual$K)
    qual <- qual[ord]
  }
  qual$id[1L]
}

foci <- c("maxeff", "maxeffCons", "maxcons", "maxSG", "minSG",
          "effMaxSG", "effMinSG")
res <- data.table(sg_focus = foci,
                  selected = vapply(foci, function(f) pick(set, f), character(1)))
res[set, on = c(selected = "id"),
    `:=`(hr = i.hr, N = i.N, Pcons = i.Pcons)]
res[]
```

Reading the result:

* **`maxeff` returns `P`** --- the strongest effect, on 40 subjects, with
  `Pcons` 0.55. No threshold could stop it; the consistency screen is overridden
  to 0. This is the winner's-curse case in its purest form.
* **`maxeffCons` returns `Q`** --- `P` is filtered out at admission (0.55 below
  $p^\star$ = 0.90), so the effect argmax over the survivors is `Q`. Same key as
  `maxeff`, different population, materially different answer.
* **`maxcons` returns `S`** --- highest `Pcons` at 0.99, despite the weakest
  effect among qualifiers. Consistency leads the key.
* **`maxSG` returns `S`** and **`minSG` returns `Q`** --- size extremes among
  qualifiers, with no band and no effect term until the final tiebreak.
* **`effMaxSG` returns `R`** --- the band is `hr >= 0.9 * 2.60 = 2.34`, which
  admits `Q` and `R` but excludes `S`; `R` then wins on size within the band
  ($N$ 130 against `Q`'s 120). Note this is *not* the effect argmax: `Q` has the
  stronger effect, but both are in-band, so the size key decides. The band is
  what stops the size-seeking rule from running out to the weak-but-large `S`
  that `maxSG` selects.
* **`effMinSG` returns `Q`** --- the same band, size minimised.

That last contrast is the practical point of the banded foci: **`maxSG` optimises
size subject to consistency; `effMaxSG` optimises size subject to consistency
*and* to staying near the best available effect.**

## Why `maxcons` cannot stop early, illustrated

The same set, in enumeration order, with the running argmax under each key:

```{r stopping-illustration}
#| code-fold: true
scan <- set[order(-hr, K)]                      # enumeration order
scan[, seen := seq_len(.N)]
scan[, best_maxeffCons := cummax(ifelse(Pcons >= 0.90, hr, -Inf))]
scan[, best_maxcons    := cummax(ifelse(Pcons >= 0.90, Pcons, -Inf))]
scan[, .(seen, id, hr, Pcons,
         qualifies = Pcons >= 0.90,
         running_effect_max = best_maxeffCons,
         running_Pcons_max  = best_maxcons)]
```

Under `maxeffCons` the first qualifier encountered --- `Q` at position 2 --- is
already the answer, because everything after it has weaker effect by
construction of the scan order. Under `maxcons` the answer is `S`, at position
4, the *last* candidate examined: its `Pcons` of 0.99 is invisible until it is
scored. Halting early on a `Pcons` key discards the winner.

# What is inert where {#sec-inert}

```{r inert}
#| code-fold: true
data.table(
  argument = c("selection_rule", "effect_neighborhood", "stop_threshold",
               "pconsistency.threshold", "hr.threshold"),
  active_for = c("effMaxSG, effMinSG only", "effMaxSG, effMinSG only",
                 "maxeffCons only", "all but maxeff", "all but maxeff"),
  elsewhere = c("ERROR if non-default", "ignored", "reset to NULL",
                "overridden to 0", "disabled"))[]
```

One of these is enforced rather than ignored: passing a `selection_rule` other
than `"neighborhood"` together with a non-band focus **raises an error**. The
rationale is that `selection_rule = "pareto"` under `sg_focus = "maxcons"` is
not a harmless redundancy --- it states an intent the run will not honour.

The rest degrade silently, which is why a run's resolved settings are worth
recording rather than inferring from the call.

# The preview sort carries an extra key {#sec-preview}

`sort_subgroups_preview()` triages candidates *before* consistency exists. It
inserts `-on_frontier` between `-in_band` and $\mp N$:

$$
(-\,\texttt{in\_band},\; -\,\texttt{on\_frontier},\; \mp N,\; -\mathrm{HR},\; K)
$$

with `on_frontier` computed irrespective of `selection_rule`. `Pcons` cannot
appear --- it does not exist yet.

The extra key protects the frontier from truncation. Only the top
`max_subgroups_search` candidates in preview order are sent for consistency
evaluation, and a restrictive rule such as `"both"` can otherwise let higher-$N$
*dominated* candidates crowd low-$N$ frontier members out of those slots, so a
candidate that would have won is never scored. The key affects **which
candidates are evaluated**, not which one wins.

# Validation against the package {#sec-validate}

```{r validate}
#| code-fold: false
if (have_fs) {
  .ok <- function(lbl, cond)
    cat(if (isTRUE(cond)) "PASS" else "FAIL", "-", lbl, "\n")

  dt_hn <- cands[, .(hr, N)]

  .ok("dominated flags match the .pareto_dominated_xy mirror",
      identical(forestsearch:::pareto_dominated_flags(dt_hn), cands$dominated))

  .ok("neighborhood band matches",
      identical(as.logical(forestsearch:::.compute_inclusion_band(
        cands$hr, cands$N, "neighborhood", EFFECT_NEIGHBORHOOD)), cands$in_nbhd))

  .ok("both band matches in_nbhd & on_frontier",
      identical(as.logical(forestsearch:::.compute_inclusion_band(
        cands$hr, cands$N, "both", EFFECT_NEIGHBORHOOD)), cands$in_both))

  .ok("compute_pareto_frontier returns the non-dominated set",
      setequal(forestsearch:::compute_pareto_frontier(dt_hn)$hr,
               cands[on_frontier == TRUE, hr]))

  # The alias table in @sec-vocab against the shipped normaliser.
  .ok("alias normalisation matches .normalize_sg_focus",
      all(vapply(vocab$supplied,
                 function(x) identical(forestsearch:::.normalize_sg_focus(x),
                                       normalize(x)), logical(1))))

  # The @sec-foci-example picks against the shipped sorter.
  chk <- function(focus, expected) {
    d <- copy(set)
    qual <- if (focus == "maxeff") d else d[Pcons >= 0.90]
    s <- forestsearch::sort_subgroups(data.table::copy(qual), focus,
                                      effect_neighborhood = EFFECT_NEIGHBORHOOD)
    identical(s$id[1L], expected)
  }
  for (f in foci)
    .ok(sprintf("sort_subgroups agrees on %s", f),
        chk(f, res[sg_focus == f, selected]))
} else {
  cat("forestsearch not installed; validation skipped.\n")
}
```

# Practical guidance

```{r guidance}
#| code-fold: true
data.table(
  if_you_want = c(
    "The maximal estimated effect, accepting instability",
    "The maximal effect among replicable subgroups",
    "The most replicable subgroup",
    "The largest replicable subgroup",
    "The largest subgroup that is also near-optimal on effect",
    "The smallest such subgroup"),
  use = c("maxeff", "maxeffCons", "maxcons", "maxSG", "effMaxSG", "effMinSG"),
  caveat = c(
    "No thresholds apply; always returns a subgroup; consistency is descriptive",
    "The only focus that can stop early; cost scales with where the winner sits",
    "Must evaluate the full pool; effect is only a tiebreak",
    "Can run out to a large, weak subgroup",
    "Band width (effect_neighborhood) is a real modelling choice",
    "Same, with size minimised"))[]
```

Two closing cautions.

**Record the focus.** Two runs that differ only in `sg_focus` estimate different
quantities. A results file that does not record which focus produced it cannot
be interpreted after the fact, and pooling across foci silently mixes estimators.

**`effect_neighborhood` is not a nuisance parameter.** At `0` the banded foci
collapse to a strict max-effect filter; as it grows they approach `maxSG`/
`minSG`. The default `0.10` sits between two qualitatively different rules, and
the choice deserves the same justification as any other modelling decision.

# Session information

```{r session-info}
#| code-fold: true
sessionInfo()
```
