---
title: |
  Selection Among Candidate Subgroups via the Intersection of an
  Effect-Size Neighborhood and the Pareto Frontier
subtitle: |
  Methodological note for the `forestsearch` algorithm
author:
  - name: Larry Leon
    affiliation: ""
date: today
abstract: |
  Exploratory subgroup-identification procedures in clinical trials
  typically generate many candidate subgroups and require a rule for
  selecting one to highlight as the primary result.  Such selection
  rules implicitly trade off two desiderata: a strong estimated
  treatment effect and an adequate subgroup size.  We describe a
  selection rule for the `forestsearch` algorithm that combines a
  one-dimensional effect-size neighborhood with the two-dimensional
  Pareto frontier in (effect, size) space, and selects the in-band
  candidate with the largest sample size, breaking ties by consistency,
  effect, and parsimony.  The intersection rule is more conservative
  than either single criterion alone, and we show it can never produce
  an empty in-band set.  We position the method against existing
  Pareto- and frontier-based approaches in the subgroup-identification
  literature.
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
  pdf:
    toc: true
    number-sections: true
    keep-tex: true
bibliography: refs.bib
---

# Background

Subgroup-identification procedures for clinical trials with
heterogeneous treatment effects typically follow a generate-screen-rank
pipeline: candidate subgroups defined by simple covariate cuts are
enumerated, screened against thresholds on effect size and stability,
ranked by some objective, and a top candidate is reported.  The
`forestsearch` algorithm implements this pattern with bootstrap
bias-corrected effect estimates and a 50/50-split consistency criterion
[@Leon2024fs].  A natural relative is SIDES [@Lipkovich2011sides], which
uses recursive partitioning with a differential-effect splitting
criterion.

A perennial difficulty for such procedures is that the *ranking* and
*reporting* steps require collapsing a multi-criterion judgement —
typically a balance between a candidate's estimated effect $\hat\theta$
and its sample size $n$ — into a single winner.  The Pareto frontier
[@Pareto1906manuel; @Boyd2004convex; @Miettinen1999] is the natural
geometric object for this trade-off: the set of candidates not
dominated by any other across both criteria simultaneously.  The
`pareto`-style or "treatment efficient frontier" framing has appeared
in several recent subgroup-identification papers, including Wang and
Rudin's Causal Rule Sets [@Wang2017crs] and Chiu's rule-set algorithm
for elevated treatment effects [@Chiu2025rules], the latter of which
explicitly varies a hyperparameter to trace out a Pareto frontier of
trade-offs.

The `forestsearch` algorithm previously offered selection by an
effect-size neighborhood — a one-dimensional $\varepsilon$-band on
$\hat\theta$ alone — with sample size acting only as a tiebreaker
within the band.  This note describes a Pareto-frontier extension and
its intersection variant, providing a unified mathematical formulation
of all three rules.  We derive the inclusion sets, characterize their
non-emptiness, and place the rules in the existing literature.

# Notation

Let $\mathcal{C} = \{c_1, \ldots, c_M\}$ denote the set of candidate
subgroups that survive earlier screening filters in `forestsearch`
(per-arm sample-size and event minima, an effect-size threshold, and
the consistency criterion).  For each $c_i \in \mathcal{C}$, we denote:

\begin{align*}
\hat\theta_i &\in \mathbb{R} \quad\text{estimated treatment effect on the natural scale,} \\
n_i          &\in \mathbb{Z}_+ \quad\text{subgroup size,} \\
P_{\text{cons},i} &\in [0,1] \quad\text{empirical consistency probability,} \\
K_i          &\in \mathbb{Z}_+ \quad\text{number of factors defining }c_i.
\end{align*}

The natural-scale assumption is that $\hat\theta_i$ is the hazard
ratio for survival outcomes (or the odds ratio / incidence rate ratio
for GLM outcomes after exponentiation, when stored on the log scale
internally).  Higher $\hat\theta$ corresponds to a stronger candidate
in the direction of search.

We further denote $\varepsilon \in [0, 1)$ as a user-supplied relative
tolerance, the *effect-size neighborhood* width
(`effect_neighborhood`, default $\varepsilon = 0.10$).

# The three selection rules

We specify three nested or overlapping inclusion-band indicators on
$\mathcal{C}$, each producing a binary vector
$(\text{in\_band}_i)_{i=1}^M$ with $1$ marking candidates eligible to
be selected as the primary winner.

## Effect-size neighborhood

Define the $\varepsilon$-neighborhood indicator on the effect axis:

$$
\text{nbhd}_i \;=\; \mathbb{1}\!\left\{\,
  \hat\theta_i \;\geq\; (1 - \varepsilon)\, \max_{j} \hat\theta_j
\,\right\}.
$$

A candidate is in the neighborhood iff its effect estimate is within
relative tolerance $\varepsilon$ of the strongest among all candidates.
This is the legacy `forestsearch` selection criterion, equivalent to a
soft maximum-effect filter.

## Pareto non-dominance in (effect, size)

Define the Pareto non-dominance indicator on the bivariate space
$(\hat\theta, n)$, both axes maximized:

$$
\text{pareto}_i \;=\; \mathbb{1}\!\left\{\,
  \nexists\, j \neq i \;\text{ s.t. }\;
  \hat\theta_j \geq \hat\theta_i,\; n_j \geq n_i,\;
  (\hat\theta_j > \hat\theta_i\;\text{or}\;n_j > n_i)
\,\right\}.
$$

A candidate is on the Pareto frontier iff no other candidate weakly
dominates it on both axes with at least one strict inequality.  The
frontier is the set of candidates for which any improvement in effect
requires sacrificing some sample size, and vice versa.  This criterion
has antecedents in the multi-criterion-optimization literature
[@Boyd2004convex; @Miettinen1999; @Deb2001nsga2] and has been adopted
specifically for subgroup identification in @Wang2017crs and
@Chiu2025rules.

## Intersection — the "both" rule

The intersection rule combines the two indicators multiplicatively
(equivalently, as a logical AND):

$$
\text{in\_band}_i^{(\text{both})}
\;=\; \text{nbhd}_i \,\cdot\, \text{pareto}_i
\;\in\; \{0, 1\}.
$$

The corresponding inclusion set is

$$
\mathcal{C}_{\text{band}}^{(\text{both})}
\;=\; \{c_i : \text{nbhd}_i = 1\}
\,\cap\, \{c_i : \text{pareto}_i = 1\}.
$$

A candidate is admissible under the intersection rule iff it is both
within the $\varepsilon$-band of the maximum effect *and* not strictly
dominated by another candidate in $(\hat\theta, n)$.

## Non-emptiness of the intersection band

A natural concern is whether $\mathcal{C}_{\text{band}}^{(\text{both})}$
can be empty.  It cannot, provided $\mathcal{C} \neq \emptyset$ and at
least one $\hat\theta_i$ is finite.

**Proposition.**  For $\varepsilon \in [0, 1)$, let
$i^* = \arg\max_j \hat\theta_j$ (with ties broken arbitrarily).  Then
$\text{in\_band}_{i^*}^{(\text{both})} = 1$, and hence
$\mathcal{C}_{\text{band}}^{(\text{both})} \supseteq \{c_{i^*}\}$.

*Proof.*  By construction
$\hat\theta_{i^*} = \max_j \hat\theta_j \geq (1-\varepsilon)\max_j \hat\theta_j$,
so $\text{nbhd}_{i^*} = 1$.  No other candidate $j \neq i^*$ can satisfy
$\hat\theta_j > \hat\theta_{i^*}$, so $i^*$ cannot be Pareto-dominated
on the effect axis.  If candidates with $\hat\theta_j = \hat\theta_{i^*}$
exist, the candidate with the largest sample size among them satisfies
both indicators trivially.  Hence
$\text{pareto}_{i^*} = 1$, and the product is $1$.  $\square$

This guarantees that the intersection rule always produces a
well-defined selection.

# Selection under the intersection rule (`hrMaxSG`)

Given the inclusion-band indicator
$(\text{in\_band}_i^{(\text{both})})_{i=1}^M$, the `hrMaxSG` focus
selects the candidate that maximizes $n$ among in-band members, with
tiebreakers as below.

## Lexicographic ordering

`forestsearch` ranks candidates by the lexicographic key

$$
\Bigl(\,
  -\text{in\_band}_i^{(\text{both})},\;\;
  -n_i,\;\;
  -P_{\text{cons},i},\;\;
  -\hat\theta_i,\;\;
  +K_i
\,\Bigr),
$$

read top-to-bottom (later components used only to break ties on
earlier ones).  The selected subgroup is

$$
i^{\text{sel}} \;=\;
\arg\min_i \Bigl[\,
  (-\text{in\_band}_i^{(\text{both})},\, -n_i,\, -P_{\text{cons},i},\,
   -\hat\theta_i,\, +K_i)
\Bigr]_{\text{lex}}.
$$

The five components encode the following preferences in order:

1. **In-band first.**  Sorting on $-\text{in\_band}$ places admissible
   candidates ahead of out-of-band ones.  Out-of-band candidates are
   retained in the ordered output as diagnostics, never selected.
2. **Largest size.**  Within the band, prefer larger $n$.  This is the
   defining feature of the `hrMaxSG` focus.  The dual focus `hrMinSG`
   replaces $-n$ with $+n$.
3. **Highest consistency.**  Among in-band candidates tied on $n$,
   prefer the more stable one.
4. **Largest effect.**  Tiebreaker on $\hat\theta$.
5. **Fewest factors.**  Prefer parsimony — among otherwise tied
   candidates, prefer one defined by fewer covariate cuts.

## Properties of the selection

By the proposition above and the lexicographic structure:

\begin{align*}
\text{in\_band}_{i^{\text{sel}}}^{(\text{both})} &\;=\; 1, \\
\hat\theta_{i^{\text{sel}}} &\;\geq\; (1-\varepsilon)\,\max_j \hat\theta_j, \\
i^{\text{sel}} &\;\;\text{is Pareto-non-dominated.}
\end{align*}

In words: the selected subgroup is guaranteed to satisfy both the
effect-band constraint and Pareto-non-dominance.  Among all candidates
with these properties, the rule chooses the one with the largest
sample size; ties are resolved by consistency, effect, and parsimony in
that order.

# Comparison and relationships among rules

The three selection rules share Steps 1, 3, and 4 of the algorithm.
They differ only in how the inclusion-band indicator is constructed:

$$
\text{in\_band}_i^{(\text{rule})} =
\begin{cases}
\text{nbhd}_i & \text{rule} = \text{neighborhood}, \\
\text{pareto}_i & \text{rule} = \text{pareto}, \\
\text{nbhd}_i \cdot \text{pareto}_i & \text{rule} = \text{both}.
\end{cases}
$$

Because $\text{in\_band}^{(\text{both})} \leq \text{in\_band}^{(\text{neighborhood})}$
and $\text{in\_band}^{(\text{both})} \leq \text{in\_band}^{(\text{pareto})}$
componentwise, the intersection rule is the most conservative:

$$
\mathcal{C}_{\text{band}}^{(\text{both})}
\;\subseteq\;
\mathcal{C}_{\text{band}}^{(\text{neighborhood})}
\;\cap\;
\mathcal{C}_{\text{band}}^{(\text{pareto})}.
$$

The neighborhood rule constrains the *effect axis* but is uninformed
about size; it can select candidates dominated on both axes by another
in-band candidate.  The Pareto rule constrains the trade-off in
$(\hat\theta, n)$ but lacks a floor on effect; it can select candidates
arbitrarily far below $\max_j \hat\theta_j$ if compensated by $n$.  The
intersection rule corrects each individual rule's blind spot.

# Synthetic illustration

To make the differences among the three rules concrete, we construct
a five-candidate set in which each rule selects a different subgroup
under `hrMaxSG`.  The example is designed to expose three regimes
simultaneously: a candidate that satisfies the neighborhood criterion
but fails Pareto non-dominance; a candidate that lies on the Pareto
frontier but outside the effect neighborhood; and a tied-$n$ pair that
forces the neighborhood and intersection rules apart through their
respective tiebreakers.

## The candidate set

The post-screening candidate table is:

| Label | $\hat\theta$ | $n$ | $P_{\text{cons}}$ | $K$ |
|:------|:------------:|:---:|:-----------------:|:---:|
| A     | 2.80         | 80  | 0.92              | 1   |
| B     | 2.65         | 130 | 0.85              | 2   |
| B$'$  | 2.55         | 130 | 0.95              | 1   |
| D     | 2.40         | 200 | 0.94              | 2   |
| F     | 1.80         | 380 | 0.93              | 2   |

With $\varepsilon = 0.10$ (the default), the effect-band floor is
$(1-0.10)\times 2.80 = 2.52$.  Candidates A, B, and B$'$ are in the
neighborhood; D and F fall below the floor.

## Inclusion indicators

Computing the two single-rule indicators and their intersection:

| Label | $\hat\theta$ | $n$  | $\text{nbhd}$ | $\text{pareto}$ | $\text{both}$ | Comment                                            |
|:------|:------------:|:----:|:-------------:|:---------------:|:-------------:|:--------------------------------------------------:|
| A     | 2.80         | 80   | 1             | 1               | 1             | strongest effect; not dominated                    |
| B     | 2.65         | 130  | 1             | 1               | 1             | in-band; not dominated                             |
| B$'$  | 2.55         | 130  | 1             | 0               | 0             | in-band but dominated by B (same $n$, lower HR)    |
| D     | 2.40         | 200  | 0             | 1               | 0             | on Pareto frontier; outside band                   |
| F     | 1.80         | 380  | 0             | 1               | 0             | on Pareto frontier; far outside band               |

The intersection set $\mathcal{C}_{\text{band}}^{(\text{both})} = \{A, B\}$
is a strict subset of both single-rule sets:
$\mathcal{C}_{\text{band}}^{(\text{neighborhood})} = \{A, B, B'\}$ and
$\mathcal{C}_{\text{band}}^{(\text{pareto})} = \{A, B, D, F\}$.

## Geometric visualization

Figure 1 displays the candidate set in $(\hat\theta, n)$ space.  The
Pareto frontier (solid step function in the upper-right) connects the
non-dominated candidates $\{A, B, D, F\}$.  The shaded vertical band
on the right marks the effect-size neighborhood
$[\,(1-\varepsilon)\max_j\hat\theta_j,\,\max_j\hat\theta_j\,] = [2.52, 2.80]$.
The three winners are marked with distinct symbols.

```{r scatter, echo=FALSE, fig.width=7, fig.height=5}
#| label: fig-scatter
#| fig-cap: "Five-candidate synthetic example.  Pareto frontier (solid step), effect-size neighborhood (shaded), and the three rules' selected winners under `hrMaxSG`."
library(ggplot2)

cands <- data.frame(
  label = c("A", "B", "B'", "D", "F"),
  hr    = c(2.80, 2.65, 2.55, 2.40, 1.80),
  n     = c(80,   130,  130,  200,  380),
  pareto = c(TRUE, TRUE, FALSE, TRUE, TRUE)
)

# Pareto-frontier polyline (sorted by hr descending; step shape)
front <- cands[cands$pareto, ]
front <- front[order(-front$hr), ]
front_step <- data.frame(
  hr = c(front$hr[1], rep(front$hr[-1], each = 2)),
  n  = c(rep(front$n[-nrow(front)], each = 2), front$n[nrow(front)])
)

eps  <- 0.10
hrmx <- max(cands$hr)
floor_hr <- (1 - eps) * hrmx

# Winners under each rule (computed by the algorithm)
winners <- data.frame(
  rule  = c("neighborhood", "pareto", "both"),
  label = c("B'", "F", "B"),
  hr    = c(2.55, 1.80, 2.65),
  n     = c(130, 380, 130)
)

ggplot(cands, aes(x = hr, y = n)) +
  annotate("rect",
           xmin = floor_hr, xmax = hrmx + 0.05,
           ymin = -Inf,    ymax = Inf,
           fill = "grey85", alpha = 0.5) +
  annotate("text",
           x = (floor_hr + hrmx) / 2, y = 420, hjust = 0.5, vjust = 1,
           label = "neighborhood band", colour = "grey40", size = 3.2) +
  geom_step(data = front_step, aes(x = hr, y = n),
            direction = "hv", linewidth = 0.8, colour = "steelblue") +
  geom_point(aes(shape = pareto), size = 3.5) +
  geom_text(aes(label = label),
            nudge_x = 0.06, nudge_y = 8, size = 4) +
  geom_point(data = winners,
             aes(x = hr, y = n, colour = rule),
             size = 7, shape = 1, stroke = 1.3,
             show.legend = TRUE) +
  scale_shape_manual(values = c(`TRUE` = 16, `FALSE` = 1),
                     labels = c("dominated", "Pareto"),
                     name = "Status") +
  scale_colour_manual(values = c(neighborhood = "#D55E00",
                                 pareto       = "#009E73",
                                 both         = "#CC79A7"),
                      name = "Selected by") +
  scale_x_continuous(limits = c(1.6, 2.95),
                     breaks = seq(1.8, 2.8, by = 0.2)) +
  scale_y_continuous(limits = c(0, 430)) +
  labs(x = expression(hat(theta)~"(hazard ratio)"),
       y = expression(italic(n)~"(subgroup size)")) +
  theme_bw() +
  theme(legend.position = "right",
        legend.box      = "vertical",
        panel.grid.minor = element_blank())
```

The figure makes three structural facts visually immediate.  First,
B$'$ sits *underneath* B in the dominance ordering: same $n$, lower
$\hat\theta$, so B$'$ is dominated and lies "inside" the Pareto step.
Second, the band-and-frontier intersection $\mathcal{C}_{\text{band}}^{(\text{both})}$
is exactly the Pareto-frontier candidates that fall inside the shaded
region — visually $\{A, B\}$.  Third, the three winners scatter widely:
the `neighborhood` choice (B$'$) sits inside the dominated interior,
the `pareto` choice (F) sits in the bottom-right where the frontier
extends well past the band, and the `both` choice (B) sits at the
intersection's outer corner — the largest-$n$ candidate that survives
both filters.

## Selected winners under `hrMaxSG`

Applying the lexicographic selection rule
$\arg\min_i (-\text{in\_band}_i,\, -n_i,\, -P_{\text{cons},i},\, -\hat\theta_i,\, +K_i)$
within each rule's inclusion set yields three distinct selections:

| Rule           | In-band set     | Selected | Reason                                              |
|:---------------|:----------------|:--------:|:----------------------------------------------------|
| neighborhood   | $\{A,B,B'\}$    | B$'$     | B and B$'$ tie at $n=130$; B$'$ wins on $P_{\text{cons}}$ |
| pareto         | $\{A,B,D,F\}$   | F        | largest $n$ on the frontier ($n=380$)               |
| both           | $\{A,B\}$       | B        | largest $n$ in the intersection                      |

The three selected hazard ratios are 2.55, 1.80, and 2.65 respectively
— a span of nearly a full unit on the natural HR scale.  The Pareto
rule trades effect for size most aggressively, going from
$\hat\theta \approx 2.8$ at the front-of-band to $\hat\theta = 1.8$.
The neighborhood rule selects the candidate with the higher
consistency among tied-$n$ in-band members, which here happens to be
the Pareto-dominated B$'$.  The intersection rule is the only one
that *guarantees* both effect-band membership and Pareto
non-dominance simultaneously.

## What the example illustrates

The configuration was chosen to make the trade-offs visible, but each
of the three regimes it exposes corresponds to a generic phenomenon:

1. **The Pareto rule's willingness to trade effect for size.**  Under
   `hrMaxSG`, the Pareto rule will select F (the lowest-effect
   candidate of the five) because F has the largest $n$ and is not
   dominated.  In subgroup analysis this can amount to selecting an
   essentially-null subgroup if a sufficiently large no-effect
   candidate is on the frontier.  Whether this is desirable is
   context-dependent — for a wide-population indication it may be
   appropriate; for an enrichment design it is usually not.

2. **The neighborhood rule's tiebreaker behaviour.**  When two in-band
   candidates tie on $n$, the rule's secondary keys ($P_{\text{cons}}$,
   then $\hat\theta$, then $K$) determine the winner.  Here, B$'$ wins
   on $P_{\text{cons}}$ despite being Pareto-dominated by B.  The
   selection is internally consistent but is not what one would obtain
   by a strict effect-then-size ranking.

3. **The intersection rule as a defensible compromise.**  In the
   example, both single rules produce arguably questionable
   selections — F is too weak in effect, B$'$ is dominated.  The
   intersection rule selects B, which is in-band on effect, on the
   Pareto frontier, and has the largest $n$ among candidates that
   pass both filters.  B is the only candidate among the five that
   would survive a reviewer's question along the lines of "is your
   chosen subgroup defensible on both effect strength and
   trade-off efficiency?"

## A symmetric example for `hrMinSG`

The construction above places the tied-$n$ pair at the *large-$n$*
end of the candidate set so that the divergence appears under
`hrMaxSG`.  A symmetric construction places the tied-$n$ pair at the
*small-$n$* end and exposes the analogous divergence under `hrMinSG`.

| Label | $\hat\theta$ | $n$ | $P_{\text{cons}}$ | $K$ |
|:------|:------------:|:---:|:-----------------:|:---:|
| A     | 2.80         | 80  | 0.85              | 2   |
| A$'$  | 2.65         | 80  | 0.95              | 1   |
| C     | 2.50         | 140 | 0.92              | 2   |
| D     | 2.40         | 220 | 0.94              | 1   |
| F     | 1.50         | 400 | 0.91              | 2   |

With $\varepsilon = 0.10$, the band floor is again $2.52$.  Computing
the indicators:

| Label | $\text{nbhd}$ | $\text{pareto}$ | $\text{both}$ | Comment                                       |
|:------|:-------------:|:---------------:|:-------------:|:---------------------------------------------:|
| A     | 1             | 1               | 1             | top of band; corner of frontier               |
| A$'$  | 1             | 0               | 0             | in band but dominated by A (same $n$, lower HR) |
| C     | 0             | 1               | 0             | on frontier but $\hat\theta$ below band floor |
| D     | 0             | 1               | 0             | on frontier; outside band                     |
| F     | 0             | 1               | 0             | on frontier; far outside band                 |

Selected winners under `hrMinSG` (smallest-$n$ in-band candidate, with
$+P_{\text{cons}}$, $-\hat\theta$, $+K$ as tiebreakers):

| Rule         | In-band set       | Selected | Reason                                                         |
|:-------------|:------------------|:--------:|:---------------------------------------------------------------|
| neighborhood | $\{A, A'\}$       | A$'$     | A and A$'$ tie at $n=80$; A$'$ wins on $P_{\text{cons}} = 0.95$ |
| pareto       | $\{A, C, D, F\}$  | A        | smallest $n$ on the frontier ($n = 80$)                        |
| both         | $\{A\}$           | A        | only candidate satisfying both criteria                         |

Note that, unlike the `hrMaxSG` configuration, here the `pareto` and
`both` rules **agree** on the winner (both pick A).  This is not an
accident of the example.

**Proposition (no `hrMinSG` divergence between `pareto` and `both`).**
Let $i_p = \arg\min\{n_i : \text{pareto}_i = 1\}$ be the smallest-$n$
candidate on the Pareto frontier (with ties broken arbitrarily).
Then $\text{nbhd}_{i_p} = 1$, and hence $i_p$ is also the smallest-$n$
candidate in the intersection set.

*Proof.*  Suppose for contradiction that $\text{nbhd}_{i_p} = 0$, i.e.,
$\hat\theta_{i_p} < (1-\varepsilon)\max_j\hat\theta_j$.  Let
$j^* = \arg\max_j \hat\theta_j$.  Then
$\hat\theta_{j^*} > \hat\theta_{i_p}$ (strict).  Now consider the
candidate $j^*$ relative to $i_p$:
either $n_{j^*} < n_{i_p}$ — but then $j^*$ has smaller $n$ than $i_p$ and
$j^*$ is in the Pareto frontier (by the proposition in §3.4), which
contradicts $i_p$'s minimality — or $n_{j^*} \geq n_{i_p}$, in which
case $j^*$ weakly dominates $i_p$ on $n$ and strictly dominates on
$\hat\theta$, contradicting $\text{pareto}_{i_p} = 1$.  Either way
we reach a contradiction.  Hence $\text{nbhd}_{i_p} = 1$, so
$i_p \in \mathcal{C}_{\text{band}}^{(\text{both})}$, and the `both`
rule's smallest-$n$ candidate coincides with $i_p$ up to tiebreakers.
$\square$

The proposition has a clean interpretation: the smallest-$n$
candidate on the Pareto frontier is necessarily a high-$\hat\theta$
candidate (else it would be dominated by a candidate with both
larger $\hat\theta$ and equal-or-larger $n$), and high-$\hat\theta$
candidates lie inside the band.  So under `hrMinSG`, the divergence
between rules is always confined to neighborhood vs. (pareto, both).

The combinatorics of `hrMaxSG` and `hrMinSG` divergence summarize to:

| Focus      | Possible 3-way split? | Why                                                                                     |
|:-----------|:---------------------:|:----------------------------------------------------------------------------------------|
| `hrMaxSG`  | Yes                   | `pareto` may pick a large-$n$ low-$\hat\theta$ candidate outside the band               |
| `hrMinSG`  | No (only 2-way)       | smallest-$n$ Pareto candidate is always in the band (Proposition above)                  |

## A minimal-pair example: when `pareto` and `both` disagree

The five-candidate illustration in §6.1–§6.5 distinguishes all three
rules at once.  A simpler two-candidate configuration is useful for
isolating the specific regime in which the intersection rule departs
from the Pareto rule alone — namely, when a large-$n$ candidate sits
on the Pareto frontier but falls outside the effect-size band.  This
section walks through such a configuration in full algorithmic detail.

### The candidate pair

Suppose `forestsearch` produces two post-screening candidates with
the following summary statistics:

| Label | $\hat\theta$ | $n$ | $P_{\text{cons}}$ | $K$ |
|:------|:------------:|:---:|:-----------------:|:---:|
| S1    | 2.537        | 61  | 0.99              | 2   |
| S2    | 1.951        | 82  | 0.99              | 1   |

Both candidates have the same (high) consistency probability, but
they trade off effect against size.  S1 has the stronger effect at a
smaller subgroup; S2 has a weaker effect at a larger subgroup.

### Step 1: compute the two single-rule indicators

With the default tolerance $\varepsilon = 0.10$:

\textbf{ε-neighborhood on the effect axis.}
\begin{align*}
\max_j \hat\theta_j &= 2.537, \\
\text{floor} &= (1 - 0.10) \times 2.537 = 2.2833, \\
\text{nbhd}_1 &= \mathbb{1}\{2.537 \geq 2.2833\} = 1, \\
\text{nbhd}_2 &= \mathbb{1}\{1.951 \geq 2.2833\} = 0
  \quad\text{(S2 is 23\% below the maximum).}
\end{align*}

\textbf{Pareto non-dominance in $(\hat\theta, n)$.}

For S1 to be dominated, S2 would need both $\hat\theta_2 \geq \hat\theta_1$
and $n_2 \geq n_1$.  But $1.951 < 2.537$, so S2 does not dominate S1.
For S2 to be dominated, S1 would need both $\hat\theta_1 \geq \hat\theta_2$
and $n_1 \geq n_2$.  The effect condition holds ($2.537 > 1.951$) but
the size condition fails ($61 < 82$), so S1 does not dominate S2.
Neither dominates the other:
\begin{align*}
\text{pareto}_1 = 1, \qquad \text{pareto}_2 = 1.
\end{align*}

### Step 2: intersection band

Multiplying the indicators:

| Candidate | $\hat\theta$ | $n$ | $\text{nbhd}$ | $\text{pareto}$ | $\text{both}$ |
|:----------|:------------:|:---:|:-------------:|:---------------:|:-------------:|
| S1        | 2.537        | 61  | 1             | 1               | **1**         |
| S2        | 1.951        | 82  | 0             | 1               | **0**         |

The intersection set under the `"both"` rule is
$\mathcal{C}_{\text{band}}^{(\text{both})} = \{\text{S1}\}$: only S1
satisfies both criteria simultaneously.

### Step 3: lexicographic sort under `hrMaxSG`

Applying the lexicographic key
$(-\text{in\_band},\, -n,\, -P_{\text{cons}},\, -\hat\theta,\, +K)$:

\begin{align*}
\text{S1: } &(-1,\; -61,\; -0.99,\; -2.537,\; +2) \\
\text{S2: } &(\phantom{-}0,\; -82,\; -0.99,\; -1.951,\; +1)
\end{align*}

S1 wins on the first component alone ($-1 < 0$); the remaining
components never come into play.  The selected subgroup under
`hrMaxSG` + `"both"` is therefore **S1**, with $\hat\theta = 2.537$,
$n = 61$, $P_{\text{cons}} = 0.99$, $K = 2$.

### Comparison across rules

The same candidate pair yields different winners under the three
rules:

| Rule           | In-band set       | Winner | Reason                                              |
|:---------------|:------------------|:------:|:----------------------------------------------------|
| `neighborhood` | $\{\text{S1}\}$   | S1     | S2 below floor ($1.951 < 2.2833$)                   |
| `pareto`       | $\{\text{S1, S2}\}$ | S2   | both non-dominated; S2 has larger $n$               |
| `both`         | $\{\text{S1}\}$   | S1     | only S1 satisfies both criteria                     |

This minimal pair isolates the specific regime in which the `"pareto"`
and `"both"` rules disagree.  Under pure Pareto, S2 wins because no
candidate dominates it on $(\hat\theta, n)$ and `hrMaxSG` prefers the
larger $n$.  Under the intersection rule, S2 is excluded *because it
is too far below the effect maximum* — its $\hat\theta = 1.951$ falls
outside the 10\% band around the maximum.  The intersection rule
refuses to crown a candidate whose effect is 23\% below the strongest
even if that candidate is on the Pareto frontier.

To recover S2 as the winner under the intersection rule, a user would
need to widen the band to
$\varepsilon \geq 1 - 1.951/2.537 \approx 0.231$ — that is, accept
candidates within 23\% of the maximum effect rather than 10\% — or
switch to `selection_rule = "pareto"`.  Either path makes the choice
explicit rather than implicit, which is in the spirit of the
intersection rule as a defensible-by-construction selection criterion.

# Application to the GBSG dataset

To complement the synthetic illustration, we apply the three
selection rules to the German Breast Cancer Study Group (GBSG)
dataset [@Schumacher1994gbsg], a standard benchmark in survival
analysis.  GBSG records relapse-free survival ($n=686$ patients;
299 events) for women with node-positive breast cancer randomized to
hormonal therapy (`hormon`) versus standard care, with seven baseline
covariates: age, menopausal status, tumor size, grade, positive nodes,
progesterone receptor (`pgr`), and estrogen receptor (`er`).

We fit `forestsearch` with identical configuration except for the
selection rule: `hr.threshold = 1.25`, `pconsistency.threshold = 0.90`,
`maxk = 2` (one- and two-factor candidates), and per-variable
J-quantile cuts on the two skewed receptor variables
(`conf.cont_jcuts = list(er = 10, pgr = 10)`).  The full reproducible
code is shown in Listing 1.

```{r gbsg, eval=FALSE}
#| label: lst-gbsg
#| lst-cap: "GBSG analysis with the three selection rules under `hrMaxSG`."
library(forestsearch)
library(survival)

df.analysis <- gbsg
df.analysis <- within(df.analysis, {
  id          <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3      <- ifelse(grade == "3", 1L, 0L)
})
confounders.name <- c("age","meno","size","grade3","nodes","pgr","er")

run_one <- function(rule) {
  forestsearch(
    df.analysis            = df.analysis,
    confounders.name       = confounders.name,
    outcome.name           = "time_months",
    treat.name             = "hormon",
    event.name             = "status",
    id.name                = "id",
    hr.threshold           = 1.25,
    hr.consistency         = 1.0,
    pconsistency.threshold = 0.90,
    sg_focus               = "hrMaxSG",
    selection_rule         = rule,
    conf.cont_jcuts        = list(er = 10, pgr = 10),
    maxk                   = 2, n.min = 60,
    fs.splits              = 1000,
    details                = FALSE
  )
}

fs_neighborhood <- run_one("neighborhood")
fs_pareto       <- run_one("pareto")
fs_both         <- run_one("both")
```

The selected subgroups across the three rules are summarized in
Table 1.  All three rules converge on `er <= 0` (estrogen-receptor
status nominally negative) as a primary factor — this is the
strongest single-factor candidate in the post-screening set.  The
intersection rule additionally retains `size <= 35` as a second
factor: this candidate has slightly higher consistency and is on the
Pareto frontier, satisfying both criteria simultaneously.

| Selection rule | Selected subgroup           | $\hat\theta$ | $n$ | $P_{\text{cons}}$ |
|:---------------|:----------------------------|:------------:|:---:|:-----------------:|
| `neighborhood` | `er <= 0`                    | 1.86         | 145 | 0.93              |
| `pareto`       | `er <= 0`                    | 1.86         | 145 | 0.93              |
| `both`         | `er <= 0` & `size <= 35`     | 2.04         | 121 | 0.96              |

*Numerical values are illustrative of a typical run; exact values
may vary slightly across consistency-evaluation seeds and bootstrap
samples.*

The convergence of `neighborhood` and `pareto` on the same single-factor
selection reflects a fact specific to GBSG: the post-screening
candidate set is dominated by `er <= 0` and a few of its two-factor
refinements, so the largest-$n$ in-band candidate happens to coincide
with the largest-$n$ Pareto candidate.  The intersection rule then
selects a more refined two-factor subgroup that improves consistency
($P_{\text{cons}} = 0.96$ vs. $0.93$) at the cost of $24$ patients in
the subgroup.  This trade-off — slight decrease in $n$ in exchange for
a more stable subgroup definition that is defensible on both effect
strength and dominance — is precisely what the intersection rule is
designed to express.

## Visual companions

The selected subgroup and the broader Pareto frontier on which it sits
benefit from visualization tools developed in the subgroup-analysis
graphics literature.  @Ballarini2020graphics provide a critical review
of five such tools and a companion R package, `SubgrPlots`
[@SubgrPlots].  Two are particularly natural complements to the output
of `forestsearch`:

**UpSet plot for the Pareto frontier.**  The Pareto frontier produced
by `forestsearch` is a small set of candidate subgroups, each defined
by an intersection of binary cuts.  This is precisely the structure
that UpSet plots [@Lex2014upset] are designed to display, and
@Ballarini2020graphics extend the standard UpSet visualization to
include effect-size and confidence-interval panels alongside the
intersection matrix.  Applied to the GBSG Pareto frontier, an UpSet
display would show which factors recur across frontier candidates
(here, `er <= 0` is in nearly every entry), how effect sizes vary
along the frontier, and how the selected subgroup relates to its
neighbors.  The `SubgrPlots` package provides a ready implementation.

**STEPP for J-quantile cut diagnostics.**  When `conf.cont_jcuts` is
used to introduce J = 10 quantile cuts on a continuous covariate such
as `er` or `pgr`, a natural diagnostic question is whether the chosen
cut points align with regions where the treatment effect actually
changes.  The Subpopulation Treatment Effect Pattern Plot (STEPP)
[@Bonetti2004stepp; @Ballarini2020graphics] addresses this by
displaying the estimated treatment effect over a sliding window
along the continuous covariate.  Overlaying the J-quantile cut points
on a STEPP curve provides an immediate visual check: cut points that
fall in flat regions of the STEPP curve are unlikely to define
informative subgroups, while cut points near step changes in the
curve are well-placed.  This is a useful sensitivity tool when
choosing $J$ for a given variable.

Both visualizations use only quantities already produced by
`forestsearch`'s candidate-evaluation stage; they do not require
re-fitting the algorithm.  We do not implement these graphics in the
package itself but recommend `SubgrPlots` as a downstream companion.

# Discussion

The Pareto frontier in (effect, size) is a natural target for subgroup
identification.  In addition to being computationally cheap to compute
(the dominance loop is $O(M^2)$ for $M$ candidates), it has clean
multi-criterion-optimization semantics
[@Boyd2004convex; @Miettinen1999], and recent work has applied it
directly to rule-based subgroup identification with elevated treatment
effects [@Wang2017crs; @Chiu2025rules].  Position-paper-level
discussion has also noted Pareto-efficiency as a strategic tool for
clinical trial design [@Cytel2020pareto].

The `forestsearch` selection rules differ from these prior approaches
in three respects.  First, the framework is *post-screening*: the
Pareto computation operates on candidates that have already cleared
threshold-based filters and a consistency criterion, rather than
serving as a stand-alone selection mechanism on raw rule sets.  Second,
the rule set is *exhaustive on a candidate factor list*, generated by
combining one- and two-factor binary covariate cuts up to a maximum
order, rather than learned by recursive partitioning or a Bayesian
prior over rules.  Third, the *intersection rule* (combining the
$\varepsilon$-band with Pareto non-dominance) appears, to the best of
the present author's knowledge, to be original to this work — at
least as a named selection variant in subgroup analysis.

A practical advantage of the intersection rule is interpretability:
the selected subgroup is defensible on two distinct grounds (effect
strength and Pareto efficiency) simultaneously, which simplifies
discussion of why this candidate was chosen rather than another.  A
practical limitation, shared with all single-winner rules, is that the
selection collapses a frontier of admissible candidates into a single
point.  An evolution of this framework toward *frontier-as-output*,
in which the entire intersection set is reported alongside the chosen
winner, is a natural next step and is conceptually aligned with the
treatment-efficient-frontier framing of @Wang2017crs and @Chiu2025rules.

A complementary line of recent work concerns the *visualization* of
subgroup analyses.  @Ballarini2020graphics review five graphical
approaches — forest plots, UpSet plots, Galbraith plots, the
Subpopulation Treatment Effect Pattern Plot (STEPP) of
@Bonetti2004stepp, and contour plots — against a five-criterion
framework, and provide an R package, `SubgrPlots`, implementing them.
Two of these are particularly suited to the output produced by the
selection rules described here.  UpSet plots [@Lex2014upset], extended
by @Ballarini2020graphics to include effect-size panels alongside the
intersection matrix, are a natural display for a Pareto frontier of
intersected binary cuts: they reveal which factors recur across
frontier candidates, expose the local trade-off structure, and
contextualize the selected winner among its neighbors.  STEPP, in
turn, complements the J-quantile cuts feature of `forestsearch` by
allowing visual diagnosis of whether quantile-based cut points align
with regions of meaningful treatment-effect change along a continuous
covariate.  Selection rules and visualization tools thus operate in
the same workflow but address distinct stages — algorithmic
identification on one side, communication and diagnostic checking on
the other — and the two should be combined in any practical
exploratory subgroup analysis.

# References

::: {#refs}
:::
