---
title: "Causal Treatment Effects in the Cox/AFT Framework"
subtitle: "Estimand Taxonomy, Biomarker-Dependent Effects, Causal Foundations, and ForestSearch Implementation"
author: "ForestSearch Package Documentation"
date: today
format:
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    embed-resources: true
    theme:
      light: cosmo
    html-math-method: mathjax
    smooth-scroll: true
    title-block-banner: "#2C3E50"
    title-block-banner-color: white
    mainfont: "Palatino, Georgia, serif"
    monofont: "Fira Code, JetBrains Mono, Consolas, monospace"
    fontsize: "1.05em"
    linestretch: 1.45
    link-external-newwindow: true
    code-fold: true
    code-tools: true
    anchor-sections: true
    fig-responsive: true
---


```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(gt)
library(knitr)
```

---

> **Scope.** This document is the unified reference for causal treatment-effect
> estimation in the ForestSearch package.  It synthesises three component
> vignettes—`treatment_effect_definitions`, `biomarker_effects`, and
> `causal_effects_brief_review`—together with recent methodological literature,
> into a single self-contained review.  The document covers:
>
> - The **Weibull AFT/Cox dual representation** and its causal advantages
> - The **Fay & Li (2024) estimand taxonomy** (individual-level vs. population-level)
> - **Three primary estimands**: Average Hazard Ratio (AHR), Controlled Direct
>   Effect (CDE), and Marginal Causal HR
> - **Biomarker-modulated treatment effects** via a two-phase spline model
> - **Causal critique** of the instantaneous Cox HR and its resolution
> - **Practical guidance** for estimand selection in ForestSearch simulations
>   and applied analyses
> - **DGM-level vs. simulation-level** truth and operating characteristics

---

# Introduction

The Cox proportional hazards model has been the de-facto standard for treatment
effect estimation in randomized clinical trials with censored time-to-event
outcomes since Cox (1972).  Its appeal is understandable: the partial
likelihood is semiparametrically efficient, hypothesis testing via the log-rank
score is exact, and the hazard ratio provides an intuitive multiplicative
summary.  Yet a growing body of methodological literature—from Hernán (2010)
through Aalen, Cook & Røysland (2015), Martinussen (2022), Fay & Li (2024), and
Knudsen et al. (2025)—has established that the instantaneous Cox hazard ratio
is **not** a causal estimand in the strict sense, even in perfectly randomized
trials with no unmeasured confounding.

This document addresses the question head-on: *which* hazard-ratio-scale
estimands are causally valid, under what conditions, and how are they computed
within the ForestSearch framework?  The answer is consequential for subgroup
identification: when treatment effects are heterogeneous across baseline
covariates, the choice of estimand determines what "true" effect each subgroup
is expected to recover, how bias is defined and measured, and how results
should be communicated to clinical audiences.

The core conclusions are:

1. The **Accelerated Failure Time (AFT) model** provides the natural causal
   data-generating mechanism for survival studies, with its scale-change
   parameter achieving both individual-level and population-level causal
   validity (Fay & Li 2024; Aalen et al. 2015).
2. The **Average Hazard Ratio (AHR)**—the geometric mean of individual causal
   log-hazard differences—is the primary simulation estimand in ForestSearch
   because it inherits the AFT's dual causal status and is deterministic given
   the super-population.
3. The **Controlled Direct Effect (CDE)** provides a complementary
   natural-scale perspective useful for biomarker-threshold analyses and as a
   secondary bias target.
4. The **Marginal Causal HR** from stacked potential outcomes serves as a
   validation target, providing the closest DGM-level analogue to the observed
   Cox HR.
5. The **Cumulative Hazard Ratio (CHR)** is recommended for applied reporting
   under non-proportional hazards.


# Notation and Foundations

## Potential Outcomes Framework

For subject $i$ in a two-arm randomized trial with treatment $A_i \in \{0, 1\}$,
the potential outcomes framework (Rubin 1974) posits a pair of latent survival
times:

$T_i(1): \text{ survival time had subject } i \text{ received treatment } (A = 1)$,

$T_i(0): \text{ survival time had subject } i \text{ received control } (A = 0).$

Only one potential outcome is observable per subject—the fundamental problem of
causal inference.  Let $\mathbf{X}_i$ denote baseline covariates, $C_i$ the
censoring time, and $Y_i = \min(T_i(A_i), C_i)$ the observed follow-up time
with event indicator $\Delta_i = \mathbf{1}(T_i(A_i) \leq C_i)$.

Throughout, $\psi^0(z) < 0$ denotes a **beneficial** treatment effect (HR $< 1$)
and $\psi^0(z) > 0$ a **detrimental** effect (HR $> 1$), following the standard
Cox convention where negative log-HR favors the experimental arm.

## Weibull Distribution Essentials

The Weibull distribution with shape $\nu > 0$ and scale $\theta > 0$ has:
$$
S(t) = \exp\!\bigl(-(t/\theta)^\nu\bigr),
\qquad
\lambda(t) = \nu\,\theta^{-\nu}\,t^{\nu-1},
\qquad
\Lambda(t) = (t/\theta)^\nu.
$$

A key identity underpins the simulation framework:

::: {.callout-tip icon=false}
## Probability-Integral Property
$$
\Lambda(T) = (T/\theta)^\nu \;\sim\; \mathrm{Exp}(1).
\tag{1}
$$
This holds because $\Pr(\Lambda(T) \leq t) = 1 - e^{-t}$.
It allows latent error terms to be drawn as $\varepsilon_i = \log Q_i$ with
$Q_i \sim \mathrm{Exp}(1)$ and survival times computed exactly on the AFT
scale.
:::


# The Weibull AFT / Cox Dual Representation

## The AFT Parameterization

Writing $Q = \Lambda(T) \sim \mathrm{Exp}(1)$ and $\varepsilon = \log Q$, the
**Extreme Value** error $\varepsilon$ has density
$f_\varepsilon(x) = \exp(x - e^x)$.  Incorporating covariates $L$, the
Weibull **AFT model** is:

$$
\log(T_i) = \mu + \mathbf{X}_i^\top \boldsymbol{\gamma} + \sigma\,\varepsilon_i,
\qquad \varepsilon_i \sim \text{Extreme Value},
\tag{2}
$$

where $\sigma = 1/\nu > 0$ is the scale parameter and $\boldsymbol{\gamma}$ are
regression coefficients on the log-time scale.  R's `survreg()` estimates
$(\hat{\mu}, \widehat{\boldsymbol{\gamma}}, \hat{\sigma})$ directly on the AFT
scale.

## The Hazard-Scale Parameterization

The same model can be written as a Cox proportional hazards model:

$$
\lambda(t \mid \mathbf{X}_i) = \lambda_0(t)\,\exp\!\bigl(\mathbf{X}_i^\top
\boldsymbol{\beta}_0\bigr),
$$

where the **AFT-to-hazard transformation** links the two parameterizations:

::: {.callout-important icon=false}
## AFT / Hazard-Scale Transformation
$$
\boldsymbol{\beta}_0 = -\boldsymbol{\gamma}\,/\,\sigma.
\tag{3}
$$
Hazard-ratio coefficients $\widehat{\boldsymbol{\beta}}_0$ are obtained from
`survreg()` output via (3).  This is the linchpin of the ForestSearch DGM:
potential outcomes are *simulated* on the AFT scale, while treatment effects
are *interpreted* on the hazard-ratio scale.
:::

## Why the Dual Representation Matters

The AFT-to-Cox correspondence means that a single fitted Weibull model provides:

- **Simulation engine**: survival times generated via the AFT error structure,
  guaranteeing exact exchangeability of latent frailties across treatment arms.
- **Interpretive frame**: treatment effects expressed as hazard ratios
  $\exp(\beta_0)$, connecting to clinical convention.
- **Causal ground truth**: as developed in [Section 9](#causal-aft), the AFT
  scale-change parameter achieves dual individual-/population-level causal
  status—a property the instantaneous Cox HR cannot claim.


# The ForestSearch AFT Data-Generating Mechanism

## Model Specification for Subgroup Simulations

The ForestSearch DGM embeds heterogeneous treatment effects via the covariate
vector $\mathbf{X}_i(a)$:

$$
\mathbf{X}_i(a)^\top \boldsymbol{\gamma}
= a\cdot\gamma_{\text{treat}}
+ \mathbf{Z}_i^\top \boldsymbol{\gamma}_Z
+ a\cdot H_i\cdot\gamma_{\text{int}},
$$

where $H_i \in \{0,1\}$ flags membership in the harm subgroup and
$\boldsymbol{\gamma}_Z$ captures prognostic covariate effects.  The
`k_inter` scalar modulates $\gamma_{\text{int}}$, providing direct control
over the magnitude of treatment effect heterogeneity.

## Individual Log-Hazard Contributions

From the AFT-to-hazard transformation (3), the individual conditional
log-hazard contributions under each treatment arm are:

$$
\theta_i(a) = \mathbf{X}_i(a)^\top \boldsymbol{\beta}_0, \quad a \in \{0,1\}.
$$

These are stored in the ForestSearch super-population data frame as `theta_1`
(treatment) and `theta_0` (control).  The individual causal log-hazard ratio is:

$$
\log \text{HR}_i^{(\text{po})} \;:=\; \theta_i(1) - \theta_i(0),
$$

stored as `loghr_po` in ForestSearch.

## Biomarker-Dependent Treatment Effects

To model treatment effects that vary continuously with a biomarker $Z$,
ForestSearch also supports a **two-phase spline model** with a single interior
knot at $Z = k$.  The Cox linear predictor is:

$$
L'\beta :=
\beta_1 X + \beta_2 Z + \beta_3 ZX
+ \beta_4(Z - k)\mathbf{1}(Z > k)
+ \beta_5(Z - k)\mathbf{1}(Z > k)X,
\tag{4}
$$

where $X \in \{0,1\}$ is the treatment indicator.

```{r}
#| label: param-table
#| echo: false
knitr::kable(
  data.frame(
    Parameter = c("$\\beta_1$", "$\\beta_2$", "$\\beta_3$",
                   "$\\beta_4$", "$\\beta_5$"),
    Term = c("$X$", "$Z$", "$ZX$",
             "$(Z-k)\\mathbf{1}(Z>k)$",
             "$(Z-k)\\mathbf{1}(Z>k)\\,X$"),
    Role = c("Main treatment effect (intercept of log-HR)",
             "Prognostic biomarker effect (both arms)",
             "Biomarker $\\times$ treatment interaction (slope of log-HR for $Z \\leq k$)",
             "Spline term: change in prognostic slope above $k$",
             "Spline term: change in treatment-effect slope above $k$")
  ),
  col.names = c("Parameter", "Term", "Role"),
  align = "cll",
  escape = FALSE,
  caption = "Parameters of the biomarker-treatment spline model."
)
```

### The Causal Log-Hazard-Ratio Function

Under the potential outcomes framework, the **causal log-hazard-ratio** at
biomarker level $z$ is:

$$
\psi^0(z) := \log\!\bigl(\lambda_{1,z}(t)\,/\,\lambda_{0,z}(t)\bigr)
= \beta_1^0 + \beta_3^0\,z + \beta_5^0\,(z - k)\,\mathbf{1}(z > k).
\tag{5}
$$

Because the baseline hazard cancels in the ratio, $\psi^0(z)$ is free of $t$
by the proportional hazards structure.  Expanding piecewise:

$$
\psi^0(z) = \begin{cases}
\beta_1^0 + \beta_3^0\,z, & z \leq k, \\[6pt]
\beta_1^0 + \beta_3^0\,z + \beta_5^0\,(z - k), & z > k.
\end{cases}
$$

### Anchor-Point Parameterization

Rather than specifying $(\beta_1^0, \beta_3^0, \beta_5^0)$ directly, the
framework specifies the log-HR at three interpretable **anchor points**
$\psi^0(0)$, $\psi^0(k)$, and $\psi^0(\zeta)$ for $\zeta > k$, and recovers:

$$
\beta_1^0 = \psi^0(0),\qquad
\beta_3^0 = \frac{\psi^0(k) - \psi^0(0)}{k},\qquad
\beta_5^0 = \frac{\psi^0(\zeta) - \psi^0(0) - \beta_3^0\,\zeta}{\zeta - k}.
\tag{6}
$$

### Connection to Potential-Outcome Means

Let $m(x,z) := E[\log T \mid \text{Treat} = x,\, Z = z]$ under the AFT model.
Then:

$$
\psi^0(z) = \frac{m(0,z) - m(1,z)}{\tau},
$$

confirming that $\psi^0(z)$ is the causal difference in expected log-survival
between arms, scaled by $\tau$.  When a prognostic covariate $W$ that does not
interact with treatment is included, $\psi^0$ depends on $z$ but not $W$.

### A Representative Biomarker Profile

```{r}
#| label: working-example
#| echo: false
knitr::kable(
  data.frame(
    Anchor = c("Low", "Knot", "High"),
    z = c("$0$", "$5$", "$10$"),
    psi = c("$\\log(3) \\approx 1.10$",
            "$\\log(1.25) \\approx 0.22$",
            "$\\log(0.5) \\approx -0.69$"),
    HR = c("$3.00$", "$1.25$", "$0.50$"),
    Interpretation = c("Strongly detrimental",
                       "Modestly detrimental",
                       "Strongly beneficial")
  ),
  col.names = c("Anchor", "$z$", "$\\psi^0(z)$", "HR", "Interpretation"),
  align = "cccrl",
  escape = FALSE,
  caption = "Representative oncology biomarker configuration. Treatment transitions from harm at low biomarker levels to substantial benefit at high levels; the overall AHR across the biomarker distribution is approximately 0.74."
)
```


# A Foundational Taxonomy: Individual-Level vs. Population-Level Estimands {#estimand-taxonomy}

Before examining specific estimands, Fay & Li (2024) establish a taxonomy that
clarifies *why* the hazard ratio causes difficulty — a framework absent from
earlier literature.

## The Core Distinction

A causal estimand requires comparing potential outcomes
$\{Y_i(0)\}$ and $\{Y_i(1)\}$ over a common comparison set $\Omega$.  The
pivotal question is: in what *order* does one apply comparison and
summarization?

- **Individual-level estimand**: Compare first within individuals, then
  summarize.  The ATE $E[Y_i(1) - Y_i(0)]$ is canonical — causal differences
  are formed per person, then averaged.
- **Population-level estimand**: Summarize each arm first, then compare.  The
  ratio $E[Y_i(1)] / E[Y_i(0)]$ is population-level.

**When these coincide.**  Linear estimands (ATE, survival difference, RMST
difference) are simultaneously individual- and population-level because
expectation is linear.  This dual status is precisely what makes them the most
*straightforwardly* causal quantities.

**When they diverge.**  For ratio estimands the order matters: under general
potential-outcome distributions, $E[Y_i(1)/Y_i(0)] \neq E[Y_i(1)]/E[Y_i(0)]$.
The left-hand individual ratio requires the *joint* distribution of
$(Y_i(0), Y_i(1))$, which is never observed.  Only $\rho = E[Y_i(1)]/E[Y_i(0)]$
is identifiable — but $\rho = 2$ does *not* imply the treatment doubles each
individual's outcome.

**The geometric mean exception.**  The geometric mean ratio for positive
outcomes is *both* individual-level and population-level:

$$
\exp\!\left(E_F\!\left[\log\frac{Y_i(1)}{Y_i(0)}\right]\right)
= \exp\!\left(E_{F_1}[\log Y_i(1)] - E_{F_0}[\log Y_i(0)]\right)
= \gamma_1/\gamma_0,
$$

where $\gamma_z$ is the geometric mean of $Y_i(z)$.  The $\log$ transformation
linearizes the ratio, making the geometric mean ratio identifiable *and*
individually interpretable.  This is why the AFT scale-change parameter and the
ForestSearch AHR achieve dual causal status.

## Classification of Key Survival Estimands

```{r}
#| label: estimand-taxonomy-table
#| echo: false
tax <- data.frame(
  Estimand = c(
    "Survival difference: $S_1(t^*) - S_0(t^*)$",
    "RMST difference: $\\int S_1 - \\int S_0$",
    "AFT scale-change: $\\exp(-\\beta)$",
    "Cumulative hazard ratio (CHR): $\\log S_1(t)/\\log S_0(t)$",
    "Cox AHR $\\theta_{\\text{CoxAHR}}$: partial likelihood limit",
    "Instantaneous HR $\\theta(t)$: $\\lambda_1(t)/\\lambda_0(t)$",
    "Constant Cox HR under PH: $\\lambda_1(t)/\\lambda_0(t) = \\theta$",
    "Marginal causal HR $\\theta_{\\text{mcHR}}$: conditioning on joint survival",
    "ForestSearch AHR: $\\exp\\!\\left(\\overline{\\log \\text{HR}_i^{(\\text{po})}}_{\\mathcal{S}}\\right)$"
  ),
  Individual = c("Yes", "Yes", "Yes", "No", "No", "No", "No", "Yes", "Yes"),
  Population = c(
    "Yes", "Yes", "Yes", "Yes", "Yes (under PH)",
    "Problematic", "Yes", "Yes", "Yes"
  ),
  Identifiable = c(
    "Yes", "Yes", "Yes", "Yes", "Yes",
    "Yes (but conditioning sets differ)", "Yes (under PH)",
    "No — requires joint $(T^0,T^1)$",
    "Yes (DGM direct; observed via AHR estimator)"
  ),
  Source = c(
    "Fay \\& Li (2024)",
    "Fay \\& Li (2024); Prentice \\& Aragaki (2022)",
    "Fay \\& Li (2024); Aalen et al. (2015)",
    "Fay \\& Li (2024); Wei \\& Schaubel (2008)",
    "Fay \\& Li (2024)",
    "Aalen (2015); Hernán (2010); Fay \\& Li (2024)",
    "Fay \\& Li (2024); Prentice \\& Aragaki (2022)",
    "Martinussen et al. (2020); Fay \\& Li (2024)",
    "ForestSearch; Fay \\& Li (2024)"
  ),
  stringsAsFactors = FALSE
)

gt(tax) |>
  fmt_markdown(columns = everything()) |>
  cols_label(
    Estimand     = md("**Estimand**"),
    Individual   = md("**Individual-level?**"),
    Population   = md("**Population-level?**"),
    Identifiable = md("**Identifiable from RCT?**"),
    Source       = md("**Key Source**")
  ) |>
  cols_width(
    Estimand     ~ px(220),
    Individual   ~ px(90),
    Population   ~ px(120),
    Identifiable ~ px(170),
    Source       ~ px(200)
  ) |>
  tab_header(
    title    = md("**Table 1. Causal Estimand Taxonomy**"),
    subtitle = "Classification by Fay & Li (2024) individual-level / population-level framework"
  ) |>
  tab_style(
    style     = cell_text(size = "small"),
    locations = cells_body()
  ) |>
  tab_style(
    style = list(cell_fill(color = "#d4edda")),
    locations = cells_body(rows = c(1, 2, 3, 9))
  ) |>
  tab_style(
    style = list(cell_fill(color = "#fff3cd")),
    locations = cells_body(rows = c(4, 5, 7))
  ) |>
  tab_style(
    style = list(cell_fill(color = "#f8d7da")),
    locations = cells_body(rows = c(6, 8))
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "#2C3E50"),
      cell_text(color = "white", weight = "bold")
    ),
    locations = cells_column_labels()
  ) |>
  tab_source_note(md(
    "Green = both individual- and population-level. Yellow = population-level only. Red = identifiability issues."
  )) |>
  tab_options(table.font.size = px(12))
```


# Three Primary Treatment Effect Estimands {#three-estimands}

## Average Hazard Ratio (AHR) {#ahr}

### Definition

The **AHR** for a subgroup $\mathcal{S}$ is the geometric mean of the individual
causal hazard ratios:

$$
\text{AHR}(\mathcal{S})
= \exp\left(\frac{1}{|\mathcal{S}|} \sum_{i \in \mathcal{S}}
  \left[\theta_i(1) - \theta_i(0)\right]\right)
= \exp\left(\frac{1}{|\mathcal{S}|} \sum_{i \in \mathcal{S}}
  \log \text{HR}_{i}^{(\text{po})}\right)
\tag{7}
$$

where $\log \text{HR}_{i}^{(\text{po})} = \theta_i(1) - \theta_i(0)$ is the individual causal log hazard ratio (stored as `loghr_po` in ForestSearch).

For biomarker-threshold subpopulations under the spline model:

$$
\text{AHR}(z^+) := \exp\!\bigl\{E_{Z \geq z}\;\psi^0(Z)\bigr\},\qquad
\text{AHR}(z^-) := \exp\!\bigl\{E_{Z \leq z}\;\psi^0(Z)\bigr\}.
$$

### Key Properties

- **Deterministic**: depends only on the DGM coefficients and covariate
  distribution, not on the random error $\varepsilon_i$.  Reproducible across
  simulation replications for a fixed super-population.
- **Dual causal status**: as a geometric mean of individual causal
  log-hazard-ratio differences, the AHR achieves both individual-level and
  population-level causal validity under the Fay & Li (2024) taxonomy.  The
  $\log$ transformation makes the comparison-then-summarize and
  summarize-then-compare operations equivalent.
- **No conditioning-set problem**: computed directly from potential outcomes,
  without conditioning on any observed survival time.

### Computation in ForestSearch

```r
# DGM-level AHR (super-population)
AHR_overall <- exp(mean(df_super$loghr_po))
AHR_H       <- exp(mean(df_super$loghr_po[df_super$flag_harm == 1]))
AHR_Hc      <- exp(mean(df_super$loghr_po[df_super$flag_harm == 0]))

# Per-replicate AHR in identified subgroup
ahr.H.hat   <- exp(mean(df_sim$loghr_po[df_sim$sg_hat == 1]))
```

**Package variables**: `loghr_po`, `ahr.H.hat`, `ahr.Hc.hat`, `AHR_H_true`,
`AHR_Hc_true`.


## Controlled Direct Effect (CDE) {#cde}

### Definition

The **CDE** for subgroup $\mathcal{S}$ is the ratio of average
exponentiated log-hazards (Aalen, Cook & Røysland 2015):

$$
\text{CDE}(\mathcal{S})
= \frac{\frac{1}{|\mathcal{S}|}
  \sum_{i \in \mathcal{S}} \exp\left(\theta_i(1)\right)}
  {\frac{1}{|\mathcal{S}|}
  \sum_{i \in \mathcal{S}} \exp\left(\theta_i(0)\right)}
= \frac{\overline{\exp(\theta_i(1))}_{\mathcal{S}}}
       {\overline{\exp(\theta_i(0))}_{\mathcal{S}}}
\tag{8}
$$

For biomarker-threshold subpopulations:
$\text{CDE}(z^+) := \overline{\exp(\theta_i(1))}_{Z \geq z} / \overline{\exp(\theta_i(0))}_{Z \geq z}$.

### Key Properties

- **Deterministic** (like AHR): no dependence on $\varepsilon_i$.
- **Natural-scale averaging**: by averaging on the hazard (exponential) scale
  rather than the log-hazard scale, the CDE weights subjects with larger
  absolute hazard contributions more heavily.
- **Differs from AHR under heterogeneity**: Jensen's inequality implies
  $\text{CDE}(\mathcal{S}) \neq \text{AHR}(\mathcal{S})$ unless all
  $\log \text{HR}_i^{(\text{po})}$ are identical.  The discrepancy grows with within-subgroup
  heterogeneity.
- **Parallels Prentice & Aragaki AHR**: the exponential-scale averaging mimics
  the control-arm weighting structure of the nonparametric AHR, making the CDE
  an informative natural-scale companion.

### Computation in ForestSearch

```r
# DGM-level CDE (super-population)
CDE_H  <- mean(exp(df_super$theta_1[idx_H])) / mean(exp(df_super$theta_0[idx_H]))
CDE_Hc <- mean(exp(df_super$theta_1[idx_Hc])) / mean(exp(df_super$theta_0[idx_Hc]))

# Equivalently via compute_dgm_cde():
dgm <- compute_dgm_cde(dgm)
```

**Package variables**: `theta_0`, `theta_1`, `cde.H.hat`, `cde.Hc.hat`,
`cde_H`, `cde_Hc`, `cde_results`.


## Marginal Causal Hazard Ratio {#marginal-hr}

### Definition

The **marginal (causal) HR** for subgroup $\mathcal{S}$ is obtained by fitting
a Cox model to *stacked potential outcomes*—a dataset in which each subject
contributes two rows (one under treatment, one under control), both with event
indicator 1:

$$
\text{HR}_{\text{marg}}(\mathcal{S})
= \exp\left(\hat{\beta}_{\text{Cox}} \mid
  \text{data} = \left\{\left(T_i(1), 1, 1\right), \left(T_i(0), 1, 0\right)
  : i \in \mathcal{S}\right\}\right)
$$

where $\hat{\beta}_{\text{Cox}}$ is the treatment coefficient from a Cox model fitted to
the stacked dataset in which each $i \in \mathcal{S}$ contributes two rows—one under
treatment, one under control—both with event indicator 1.

### Computation in ForestSearch

1. Generate a common set of errors: $\varepsilon_i = \log(\text{Exp}(1))$.
2. Compute potential survival times using the same $\varepsilon_i$ for both arms.
3. Stack into $2n$ rows; fit Cox model; extract $\exp(\hat{\beta}_{\text{treat}})$.

```r
df_stacked <- rbind(
  data.frame(time = T_1, event = 1, treat = 1, flag_harm = flag_harm),
  data.frame(time = T_0, event = 1, treat = 0, flag_harm = flag_harm)
)
hr_causal <- exp(coxph(Surv(time, event) ~ treat, data = df_stacked)$coef)
```

### Key Properties

- **Stochastic**: depends on the realized $\varepsilon_i$, introducing Monte
  Carlo variability across replications absent from AHR and CDE.
- **Validation role**: the DGM-level marginal HR is the ground truth
  $\theta^\dagger$ against which per-replicate estimator bias is assessed in
  `build_estimation_table()`.
- **Bypasses partial-likelihood risk sets**: by using stacked potential outcomes
  (all events observed, no censoring), the Cox fit here is not subject to the
  conditioning-set problem of the standard Cox HR.

**Package variables**: `hr_causal`, `hr_H_true`, `hr_Hc_true`,
`hr.H.hat`, `hr.Hc.hat`.


## When the Three Estimands Agree and When They Diverge

Under a **constant treatment effect** (no heterogeneity), the individual
log hazard ratio is identical for all subjects:

$$
\log \text{HR}_{i}^{(\text{po})} = \beta_{\text{treat}} \quad \forall\, i,
$$

so all three estimands coincide:
$\text{HR}_{\text{marg}} = \text{AHR} = \text{CDE} = \exp(\beta_{\text{treat}})$.

Under **heterogeneous effects**, three sources of divergence arise:

1. **Jensen's inequality** separates AHR from CDE:

$$
\exp\left(\overline{\log \text{HR}_i}\right) \neq
\overline{\exp(\theta_i(1))} / \overline{\exp(\theta_i(0))}
$$

   unless all $\log \text{HR}_{i}^{(\text{po})}$ are identical.

2. **Cox partial-likelihood weighting** separates the marginal HR from the AHR,
   because risk-set membership implicitly weights subjects non-uniformly.

3. **Monte Carlo variability** affects the marginal HR (which depends on
   $\varepsilon_i$) but not the AHR or CDE.


## Summary Comparison

```{r}
#| label: comparison-table
#| echo: false
comparison_df <- data.frame(
  Property = c(
    "Formal name",
    "Primary notation",
    "Formula (subgroup $\\mathcal{S}$)",
    "Averaging scale",
    "Depends on $\\varepsilon_i$",
    "Reproducible across seeds",
    "Individual-level causal status",
    "Population-level causal status",
    "Sensitive to extreme hazard values",
    "Standard estimation method",
    "Primary ForestSearch role",
    "Key ForestSearch variables"
  ),
  AHR = c(
    "Average hazard ratio",
    "$\\text{AHR}(\\mathcal{S})$",
    "$\\exp\\!\\left(\\overline{\\log \\text{HR}_i^{(\\text{po})}}_{\\mathcal{S}}\\right)$",
    "Log-hazard (then exponentiated)",
    "No",
    "Yes (deterministic)",
    "Yes",
    "Yes",
    "Moderate (log scale dampens extremes)",
    "Mean individual log-HR",
    "Primary simulation estimand; bias target $\\theta^\\dagger$",
    "<code>loghr_po</code>, <code>ahr.H.hat</code>, <code>AHR_H_true</code>"
  ),
  CDE = c(
    "Controlled direct effect",
    "$\\text{CDE}(\\mathcal{S})$, $\\theta^\\ddagger$",
    "$\\overline{\\exp(\\theta_i(1))}_{\\mathcal{S}} / \\overline{\\exp(\\theta_i(0))}_{\\mathcal{S}}$",
    "Hazard (natural scale ratio)",
    "No",
    "Yes (deterministic)",
    "Partial (ratio of means, not mean of ratios)",
    "Yes",
    "Yes (exponential scale amplifies large values)",
    "Ratio of average exponentiated log-hazards",
    "Secondary bias target $\\theta^\\ddagger$; biomarker-threshold analysis",
    "<code>theta_0</code>, <code>theta_1</code>, <code>cde.H.hat</code>, <code>cde_H</code>"
  ),
  MarginalHR = c(
    "Marginal (causal) hazard ratio",
    "$\\text{HR}_{\\text{marg}}(\\mathcal{S})$, $\\theta^\\dagger$",
    "$\\exp(\\hat{\\beta}_{\\text{Cox}})$ on stacked potential outcomes",
    "Hazard ratio (Cox partial likelihood)",
    "Yes",
    "No (Monte Carlo variability)",
    "No (population-averaged)",
    "Yes (under PH; stacked PO bypasses conditioning issue)",
    "No (rank-based partial likelihood)",
    "Cox PH on stacked potential outcomes",
    "DGM-level ground truth; validation target",
    "<code>hr_causal</code>, <code>hr_H_true</code>, <code>hr.H.hat</code>"
  ),
  stringsAsFactors = FALSE
)

gt(comparison_df) |>
  fmt_markdown(columns = everything()) |>
  cols_label(
    Property   = "",
    AHR        = md("**AHR**"),
    CDE        = md("**CDE** ($\\theta^\\ddagger$)"),
    MarginalHR = md("**Marginal HR** ($\\theta^\\dagger$)")
  ) |>
  cols_width(
    Property   ~ px(220),
    AHR        ~ px(240),
    CDE        ~ px(240),
    MarginalHR ~ px(240)
  ) |>
  tab_header(
    title    = md("**Table 2. Comparison of Three Treatment Effect Estimands**"),
    subtitle = "AHR, CDE, and Marginal Causal HR in ForestSearch"
  ) |>
  tab_row_group(label = md("**Definition**"), rows = 1:3) |>
  tab_row_group(label = md("**Statistical Properties**"), rows = 4:9) |>
  tab_row_group(label = md("**Use in ForestSearch**"), rows = 10:12) |>
  row_group_order(groups = c("**Definition**","**Statistical Properties**","**Use in ForestSearch**")) |>
  tab_style(
    style = list(cell_fill(color = "#F0F4F8"), cell_text(weight = "bold", size = "small")),
    locations = cells_row_groups()
  ) |>
  tab_style(
    style = cell_text(size = "small"),
    locations = cells_body()
  ) |>
  tab_style(
    style = cell_text(weight = "bold", size = "small"),
    locations = cells_body(columns = Property)
  ) |>
  tab_style(
    style = list(cell_fill(color = "#E8F4E8")),
    locations = cells_body(columns = AHR, rows = c(5,6,7,8))
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "#2C3E50"),
      cell_text(color = "white", weight = "bold")
    ),
    locations = cells_column_labels()
  ) |>
  tab_options(table.font.size = px(12))
```


# Causal Critique of the Cox Hazard Ratio {#causal-critique}

## Setup and the Fundamental Problem

Consider a two-arm randomized trial with treatment $A \in \{0,1\}$ and
unmeasured frailty $Z$ independent of $A$ at baseline ($Z \perp\!\!\!\perp A$
at $t=0$).  The Cox model for the marginal hazard is:

$$
\lambda(t \mid A = a) = \lambda_0(t)\,e^{\beta a}.
$$

Under randomization, Cox regression yields a **consistent estimator** of
$\beta$.  But consistency of estimation does not imply causal interpretability
of the estimand.

## Three Sources of Interpretive Difficulty (Fay & Li 2024)

Fay & Li (2024) identify three properties of $\lambda_z(t)$ that complicate
causal interpretation:

1. **It is a derivative** — a mathematical limit with no simple
   expectation-based definition.
2. **It is a local effect** — conveying information only at time $t$, unlike
   $S_z(t)$, which summarises cumulative experience from 0 to $t$.
3. **Its conditioning sets are not comparable.**  The hazard at $t$ in arm $z$
   conditions on $\Omega_z(t) = \{i : Y_i(z) \geq t\}$.  When treatment has an
   effect, $\Omega_0(t) \neq \Omega_1(t)$ for any $t > 0$.  This violates the
   identical comparison-set requirement for a causal estimand.

## The Collider Mechanism

Even though $Z \perp\!\!\!\perp A$ at baseline, among survivors at $t > 0$:

$$
Z \mathrel{\not\!\perp\!\!\!\perp} A \mid T \geq t,
$$

unless the joint hazard is additive in $A$ and $Z$ — a condition the Cox
multiplicative model violates.  Survival $S_t$ is a **collider** on the path
$A \to S_t \leftarrow Z \to S_{t+\Delta}$.  Conditioning opens a non-causal
association between $A$ and $Z$ that accumulates throughout follow-up.
Hernán (2010) termed this "built-in selection bias."

The joint distribution among survivors is:
$$
P(Z = z, A = a \mid T \geq t) \;\propto\;
\exp\!\!\left(-\int_0^t h(s, a, z)\,ds\right) P(Z = z)\,P(A = a),
$$
which factors only when $h$ is additive.  Under Cox's multiplicative structure
it does not, so $e^{\hat{\beta}}$ converges to a parameter that conflates the
treatment effect with a growing frailty imbalance between arms.

## Mathematical Elaboration (Martinussen 2022)

Martinussen (2022) shows the root cause lies in conditioning on survival at any
$t > 0$, regardless of model structure.  For a piecewise Cox model with change
point $\nu$:

$$
\frac{\lambda(t;\,A=1)}{\lambda(t;\,A=0)} =
\begin{cases} e^{\beta}, & t \leq \nu \\ 1, & t > \nu, \end{cases}
$$

the HR declining to 1 after $\nu$ is *fully consistent* with a beneficial
treatment at *all* times, when examined through the frailty-conditional HR.
The contradiction arises purely from selection.

The genuinely causal HR conditions on **joint survival** under both potential
outcomes:
$$
\text{HR}(t) =
\frac{P(t \leq T^1 < t+h \mid T^0 \geq t,\, T^1 \geq t) / h}
     {P(t \leq T^0 < t+h \mid T^0 \geq t,\, T^1 \geq t) / h}.
$$

When $e^\beta < 1$, this quantity satisfies $\text{HR}(t) < e^\beta < 1$ — the
true causal effect is *stronger* than the Cox HR implies.  This marginal causal
HR ($\theta_{\text{mcHR}}$) is not identifiable without untestable assumptions
about the joint distribution $(T^0, T^1)$.

**Hazard differences share the same limitation.**  Martinussen demonstrates that
the Aalen additive hazard model produces $\psi(t) = 0$ for $t > \nu$ for the
same selection reason — the problem lies in the hazard function itself, not in
its multiplicative structure.

## The Vaccine Waning Efficacy Example (Fay & Li 2024)

Fay & Li construct two data-generating models that produce **identical**
population cumulative incidence curves for both trial arms:

- **Model 1** (heterogeneous, constant individual efficacy): 95% healthy
  individuals (HR 0.04) and 5% frail (HR 0.04).  Constant individual efficacy,
  but frail subjects are differentially depleted from the placebo arm, causing
  the *observed* population HR to decline over time.
- **Model 2** (homogeneous, waning efficacy): genuinely declining individual
  protection over time.

Both models produce the same Kaplan-Meier curves.  Without frailty information,
the two are indistinguishable from time-to-event data alone — but the policy
implications differ completely.

::: {.callout-note icon=false}
## ForestSearch Relevance
This aliasing between population-level time-varying
effects and individual-level heterogeneity is exactly the problem ForestSearch's
subgroup identification machinery addresses.  By conditioning on baseline
covariates $\mathbf{X}_i$ via GRF and LASSO, the algorithm recovers
individual-level treatment effect variation that would otherwise be absorbed into
a time-varying population HR.  The Fay & Li example formalises why such
conditioning is both theoretically necessary and practically important.
:::


# The Cox Model Under and Without Proportional Hazards

## Under PH: A Valid Population-Level Causal Estimand

Fay & Li (2024) identify one setting where the constant Cox HR does achieve
clean population-level causal status.  When $\lambda_1(t)/\lambda_0(t) = \theta$
for all $t$, the PH assumption implies $S_1(t) = S_0(t)^\theta$, so:

$$
\theta = \frac{\log S_1(t)}{\log S_0(t)} \quad \text{for all } t > 0.
$$

This is the **cumulative hazard ratio (CHR)** at any time $t$ — a comparison of
two survival functions, both population-level causal estimands from the
randomized trial.  Under PH all three interpretive difficulties vanish.

**Important caveat.**  Even under PH, $\theta$ is *not* individual-level.
A HR of 0.5 implies the population hazard rate in the treatment arm is half the
control arm rate at every $t$, not that treatment halves each individual's
hazard.

## Without PH: The Cumulative Hazard Ratio Is Preferred

When $\theta(t)$ is not constant, Fay & Li (2024) recommend:

$$
\theta_{\text{CHR}}(t) = \frac{\log S_1(t)}{\log S_0(t)}
$$

as the most defensible population-level causal estimand, because it:
- Directly contrasts identifiable survival functions
- Removes all three interpretive difficulties (no limits, no local effects, same
  cumulative comparison window)
- Does not depend on the censoring distribution (unlike the Cox AHR under
  non-PH)

```{r}
#| label: causal-ordering
#| echo: false
ord <- data.frame(
  Estimand     = c(
    "RMST difference",
    "Survival difference at $t^*$",
    "CHR: $\\log S_1(t)/\\log S_0(t)$",
    "Nonparametric AHR (Prentice \\& Aragaki)",
    "Cox HR under PH",
    "Instantaneous HR $\\theta(t)$"
  ),
  causal_status = c(
    "Individual \\& population",
    "Individual \\& population",
    "Population (cleanest under non-PH)",
    "Population (approximately)",
    "Population (requires PH verification)",
    "Problematic"
  ),
  ph_required   = c("No", "No", "No", "No", "Yes", "N/A"),
  early_effects = c(
    "Lower (rare outcomes)",
    "Depends on $t^*$",
    "Yes",
    "Yes",
    "Yes",
    "Yes (but misleading)"
  ),
  stringsAsFactors = FALSE
)

gt(ord) |>
  fmt_markdown(columns = everything()) |>
  cols_label(
    Estimand      = md("**Estimand**"),
    causal_status = md("**Causal Status**"),
    ph_required   = md("**PH Required?**"),
    early_effects = md("**Sensitive to Early Effects?**")
  ) |>
  tab_header(
    title = md("**Table 3. Causal Cleanness Ordering Under Non-Proportional Hazards**"),
    subtitle = "Synthesised from Fay & Li (2024) and Prentice & Aragaki (2022)"
  ) |>
  tab_style(
    style = cell_text(size = "small"),
    locations = cells_body()
  ) |>
  tab_style(
    style = list(cell_fill(color = "#d4edda")),
    locations = cells_body(rows = 1:2)
  ) |>
  tab_style(
    style = list(cell_fill(color = "#fff3cd")),
    locations = cells_body(rows = 3:5)
  ) |>
  tab_style(
    style = list(cell_fill(color = "#f8d7da")),
    locations = cells_body(rows = 6)
  ) |>
  tab_style(
    style = list(cell_fill(color = "#2C3E50"), cell_text(color = "white", weight = "bold")),
    locations = cells_column_labels()
  ) |>
  tab_options(table.font.size = px(12))
```


# The AFT Model as a Causal Ground Truth {#causal-aft}

## Dual Causal Status of the Scale-Change Parameter

Both Aalen et al. (2015) and Fay & Li (2024) identify the AFT model as the
natural causal alternative to the Cox model.  The scale-change parameter
$\exp(-\beta_{\text{treat}})$ satisfies:

$$
\exp(-\beta) = \gamma_1 / \gamma_0,
$$

where $\gamma_z$ is the geometric mean of $Y_i(z)$.  The geometric mean ratio is
simultaneously:

- **Individual-level**: equals $\exp\bigl(E[\log Y_i(1) - \log Y_i(0)]\bigr)$,
  with comparison formed *within* individuals before averaging.
- **Population-level**: identifiable from marginal distributions in a randomized
  trial.

This **collapsibility**—invariance to marginalizing over the frailty
distribution—ensures the marginal AFT treatment effect is consistent regardless
of population heterogeneity.  Empirically, Aalen et al. (2015) show AFT
estimators are unbiased across a wide range of censoring rates and frailty
variances, while Cox estimators carry substantial variable bias.

## AFT and the ForestSearch AHR

The ForestSearch AHR computed from individual potential-outcome log-HR
differences:
$$
\text{AHR}(\mathcal{S}) = \exp\!\left(\frac{1}{|\mathcal{S}|}\sum_{i \in \mathcal{S}}
[\theta_i(1) - \theta_i(0)]\right),
$$
inherits the AFT's dual causal status by the Fay & Li (2024) framework.  It
is a geometric mean of individual causal log-hazard *differences*, making it
collapsible over frailty.  Crucially, the ForestSearch AHR is computed directly
from potential outcomes **without conditioning on any observed survival time**,
so it is entirely free of the conditioning-set problem documented by Fay & Li.

::: {.callout-note icon=false}
## Distinction from the Prentice & Aragaki (2022) AHR
Their nonparametric
AHR weights each instantaneous HR by the control-group failure probability at
that moment — a time-averaged summary estimated from observed data.  The
ForestSearch AHR is a DGM-level quantity averaging potential-outcome
log-hazards, providing the simulation ground truth.  Both avoid the collider
bias problem: Prentice & Aragaki through integration over follow-up time;
ForestSearch through direct use of potential outcomes.
:::


# A Practical Defense: What the Cox Model Can Legitimately Deliver

## The Prentice & Aragaki (2022) ITT Perspective

Prentice & Aragaki argue that hazard rate contrasts retain valid causal
interpretations in randomized trials — not as point-in-time HR quantities, but
as **functionals of hazard functions** that inherit causal standing from
randomization.

Under independent censoring, the hazard rate is consistently estimable via the
Nelson-Aalen estimator, and so is any smooth functional thereof.  The survival
function, RMST, and the nonparametric AHR all derive causal interpretation
*directly from randomization*.  The independence of $A$ from all
pre-randomization risk factors ensures outcome pattern differences are
attributable to treatment.

## AHR vs. RMST: Empirical Evidence from the WHI

Prentice & Aragaki's analyses of the CEE+MPA trial provide real-data evidence
for the relative sensitivity of AHR vs. RMST:

- For **breast cancer**, the **AHR** detected a statistically significant
  elevation by **6 years** from randomization (AHR $\approx$ 1.20).
- The **RMST** contrast did not show a clear reduction until approximately
  **12 years**, because the absolute risk difference accumulates slowly for a
  relatively rare outcome.

The breast cancer risk elevation was the *principal trigger* for early
stoppage of the CEE+MPA trial after 5.6 years.  Had RMST been the primary
reporting tool, the trial-stopping signal would not have been detectable.
This real-data finding validates the theoretical argument that AHR functionals
are more sensitive to early or moderate effects in prevention settings.

## What the Cox Model Can Legitimately Do

Combining all five foundational papers, the following Cox uses are defensible:

1. **Hypothesis testing** via the log-rank score test: valid under any
   specification.
2. **Under PH**: consistent estimation of $\theta$, a valid population-level
   causal estimand equalling $\log S_1(t)/\log S_0(t)$ at every $t$.
3. **CHR estimation**: plugging Kaplan-Meier estimates into
   $\log\hat{S}_1(t)/\log\hat{S}_0(t)$ provides the cleanest population-level
   causal summary under non-PH.
4. **AHR and RMST estimation**: Cox survival function estimates feed into
   functionals causally valid in randomized trials.
5. **Covariate adjustment**: adding baseline covariates moves the estimated
   conditional HR progressively toward an individual-level causal quantity.
   GRF/LASSO conditioning in ForestSearch is the mechanism by which individual-
   level heterogeneous treatment effects become estimable.
6. **Descriptive clinical communication**: "persons assigned to CEE+MPA had
   an 80% increase in CHD incidence over year 1" is a valid ITT statement,
   provided it is not interpreted as a causal HR at that specific time in
   isolation.


# Extension to Time-Varying Treatment (Knudsen et al. 2025)

## The Markov Property and Its Impossibility

When treatment $A(t)$ evolves over time, Cox marginal structural models require
the **Markov property** — that the marginal hazard depends only on current
treatment $A(t)$.  Knudsen et al. (2025) prove this requires a degenerate
(point-mass) frailty distribution.  With complete-data hazard
$\lambda(t \mid A(t), Z) = Z\,h(t, A(t))$:

$$
\lambda(t \mid \overline{A}(t)) =
h(t, A(t))\,E\!\left(Z \mid \overline{A}(t), T \geq t\right),
$$

and the conditional expectation of $Z$ depends on the entire cumulative hazard
history $H(t, \overline{A}(t))$, not just $A(t)$.  For Markovianity, the
Laplace transform of $Z$ must be exponential — corresponding to a point mass.
Since unmeasured heterogeneity is universal, the Markov property **cannot
realistically hold in any real dataset**.

## Practical Consequences

Knudsen et al. (2025) demonstrate that even with perfect confounding control,
causal interpretation of $e^{\hat{\beta}}$ in a Cox MSM requires the Markov
property.  In a concrete illness-death model, the rate-based survival estimate
substantially underestimates the true causal survival benefit: at $t=3$ the
rate-based estimate is 0.14 vs. the true causal contrast of 0.22 — a 57%
underestimate.  This is the time-varying treatment analogue of the Fay & Li
conditioning-set problem.

## Relevance to ForestSearch

ForestSearch uses **fixed baseline treatment assignment**, placing it in the
Fay & Li baseline-treatment setting where these time-varying complications do
not arise.  No Markov property is invoked, and the potential outcome approach
used to define the AHR and CDE directly avoids the conditioning-set distortion.

Any extension of ForestSearch to **time-varying treatment subgroup
identification** would need to contend with Knudsen et al.'s results:
subgroup-specific Cox MSM HRs would not be causally interpretable even with
perfect IPTW weighting.


# Reconciling the Literature: A Synthesis Table

```{r}
#| label: reconciliation-table
#| echo: false
df <- data.frame(
  Claim = c(
    "Cox HR $e^\\beta$ is NOT a causal HR",
    "Under PH, constant $\\theta$ IS a valid pop-level causal estimand",
    "CHR is causally valid under non-PH",
    "RMST and survival differences are both individual- and population-level",
    "AFT $\\exp(-\\beta)$ is both individual- and population-level",
    "Declining observed HR $\\neq$ genuine effect waning",
    "Hazard function functionals are causally valid in RCTs",
    "Markov property fails under frailty"
  ),
  Source = c(
    "Aalen (2015); Martinussen (2022); Hernán (2010); Fay \\& Li (2024)",
    "Fay \\& Li (2024); Prentice \\& Aragaki (2022)",
    "Fay \\& Li (2024); Wei \\& Schaubel (2008)",
    "Fay \\& Li (2024)",
    "Fay \\& Li (2024); Aalen et al. (2015)",
    "Martinussen (2022); Fay \\& Li (2024)",
    "Prentice \\& Aragaki (2022); Martinussen et al. (2020)",
    "Knudsen et al. (2025)"
  ),
  ForestSearch_Resolution = c(
    "AHR and CDE are computed from AFT potential outcomes without risk-set conditioning.  Stacked-PO marginal HR bypasses partial-likelihood risk sets.  Naive Cox HR <code>hr.H.hat</code> treated as biased comparator.",
    "When Cox models are used for AHR/RMST estimation in applied output, PH should be verified; CHR should be reported as primary summary.",
    "Applied ForestSearch output should report CHR alongside AHR; the nonparametric AHR tracks CHR closely in practice.",
    "RMST difference is a natural complement to AHR; less sensitive to early effects but cleanly dual-causal.",
    "Weibull AFT DGM provides ground truth with both causal levels.  The transformation $\\boldsymbol{\\beta}_0 = -\\boldsymbol{\\gamma}/\\sigma$ links all hazard-scale estimands.",
    "GRF/LASSO conditioning on baseline $\\mathbf{X}_i$ resolves the aliasing between frailty heterogeneity and time-varying observed HR.  This is the Fay \\& Li vaccine-example remedy.",
    "Randomization confers causal validity on hazard function estimates; functionals (AHR, RMST, survival contrasts) inherit this.  Cox is a valid estimation engine.",
    "Baseline-only treatment keeps ForestSearch in the safe zone.  No Markov assumption required."
  ),
  stringsAsFactors = FALSE
)

gt(df) |>
  fmt_markdown(columns = everything()) |>
  cols_label(
    Claim                   = md("**Claim**"),
    Source                  = md("**Source**"),
    ForestSearch_Resolution = md("**ForestSearch Design Response**")
  ) |>
  cols_width(
    Claim                   ~ px(170),
    Source                  ~ px(190),
    ForestSearch_Resolution ~ px(390)
  ) |>
  tab_header(
    title    = md("**Table 4. Reconciling the Causal HR Literature**"),
    subtitle = "From Aalen (2015) through Knudsen (2025), mapped to ForestSearch design choices"
  ) |>
  tab_style(style = cell_text(size = "small"), locations = cells_body()) |>
  tab_style(
    style = list(cell_fill(color = "#EEF2F7")),
    locations = cells_body(rows = c(1, 3, 5, 7))
  ) |>
  tab_style(
    style = list(cell_fill(color = "#2C3E50"), cell_text(color = "white", weight = "bold")),
    locations = cells_column_labels()
  ) |>
  tab_options(table.font.size = px(12))
```


# DGM-Level vs. Simulation-Level Truth {#dgm-vs-sim}

The variable names `hr.H.true` and `hr.Hc.true` appear in two distinct contexts
within the ForestSearch codebase.  Both use the same computational
core — a Cox model fit within the true subgroup defined by `flag_harm` — but
they operate on fundamentally different data.

## DGM-Level Truth (Super-Population)

When `setup_gbsg_dgm()` or `calculate_hazard_ratios()` constructs the DGM, the
marginal HR is computed from **stacked potential outcomes** in a large
super-population (typically $n_{\text{super}} = 5{,}000$).  Every subject
contributes uncensored survival times under *both* treatment and control,
using the **same** random error $\varepsilon_i$ for both arms:

```r
# In setup_gbsg_dgm() — stacked potential outcomes
df_po <- data.frame(
  time      = c(T_1, T_0),
  event     = 1L,
  treat     = c(rep(1, n_super), rep(0, n_super)),
  flag_harm = rep(flag_harm, 2)
)
hr_H_true <- exp(coxph(Surv(time, event) ~ treat_sim,
                       data = subset(df_po, flag_harm == 1))$coef)
```

These DGM-level values are essentially **fixed constants** for a given DGM
specification.  They converge to the theoretical values as
$n_{\text{super}} \to \infty$:

$$
\text{HR}_{\text{DGM}}(\mathcal{H})
= \exp\!\bigl(\beta_0[\text{treat}] + \beta_0[\text{zh}]\bigr),
\qquad
\text{HR}_{\text{DGM}}(\mathcal{H}^c)
= \exp\!\bigl(\beta_0[\text{treat}]\bigr).
$$

**Stored as:** `dgm$hr_H_true`, `dgm$hr_Hc_true`, `dgm$hr_causal`.

## Simulation-Level Truth (Per-Replicate)

During each simulation replicate, the true subgroup HRs are recomputed on the
**simulated sample** — with censoring, finite sample size, and randomized
treatment:

```r
# In run_fs_analysis() — observed data, single treatment arm per subject
hr.H.true  <- exp(coxph(Surv(y_sim, event_sim) ~ treat_sim,
                        data = subset(df, flag_harm == 1))$coef)
hr.Hc.true <- exp(coxph(Surv(y_sim, event_sim) ~ treat_sim,
                        data = subset(df, flag_harm == 0))$coef)
```

These per-replicate values are **random variables** that fluctuate across
iterations due to sampling variability, censoring patterns, and treatment
randomization.  **Stored as:** columns `hr.H.true`, `hr.Hc.true` in the
results `data.table`.

## Side-by-Side Comparison

```{r}
#| label: dgm-vs-sim-table
#| echo: false
dgm_sim_df <- data.frame(
  Aspect = c(
    "Data source",
    "Potential outcomes",
    "Censoring",
    "Sample size",
    "Treatment assignment",
    "Subgroup membership",
    "Nature of quantity",
    "Primary purpose",
    "Location in code",
    "Storage"
  ),
  DGM_Level = c(
    "Stacked potential outcomes from super-population",
    "Both $T(1)$ and $T(0)$ observed per subject",
    "None (event = 1 for all rows)",
    "$2 \\times n_\\text{super}$ (e.g. 10,000 rows)",
    "Both arms per subject (causal framework)",
    "Known <code>flag_harm</code> (truth)",
    "Essentially fixed (large $N$ constant)",
    "Define ground truth for bias calculation",
    "<code>setup_gbsg_dgm()</code>, <code>calculate_hazard_ratios()</code>",
    "<code>dgm&#36;hr_H_true</code>, <code>dgm&#36;hr_Hc_true</code>"
  ),
  Sim_Level = c(
    "Observed simulated trial data",
    "Only $T(A_i)$ observed per subject",
    "Yes (administrative + random censoring)",
    "$n_\\text{sample}$ (e.g. 300--700)",
    "Randomized (one arm per subject)",
    "Known <code>flag_harm</code> (truth)",
    "Random (varies across replicates)",
    "Per-replicate reference for estimator comparison",
    "<code>run_fs_analysis()</code>, <code>run_grf_analysis()</code>",
    "<code>results&#36;hr.H.true</code>, <code>results&#36;hr.Hc.true</code>"
  ),
  stringsAsFactors = FALSE
)

gt(dgm_sim_df) |>
  fmt_markdown(columns = everything()) |>
  cols_label(
    Aspect    = "",
    DGM_Level = md("**DGM-Level (Super-Population)**"),
    Sim_Level = md("**Simulation-Level (Per-Replicate)**")
  ) |>
  cols_width(Aspect ~ px(180), DGM_Level ~ px(320), Sim_Level ~ px(320)) |>
  tab_header(
    title    = md("**Table 5. DGM-Level vs. Simulation-Level True Hazard Ratios**"),
    subtitle = "Both use `coxph(Surv ~ treat)` within `flag_harm` subgroups, but on fundamentally different data"
  ) |>
  tab_style(style = cell_text(size = "small"), locations = cells_body()) |>
  tab_style(style = cell_text(weight = "bold", size = "small"), locations = cells_body(columns = Aspect)) |>
  tab_style(
    style = list(cell_fill(color = "#E8F4E8")),
    locations = cells_body(rows = c(7, 8))
  ) |>
  tab_style(
    style = list(cell_fill(color = "#2C3E50"), cell_text(color = "white", weight = "bold")),
    locations = cells_column_labels()
  ) |>
  tab_options(table.font.size = px(13))
```

## Role in Operating Characteristics

Multiple HR columns work together to quantify estimation performance:

| Column | Meaning |
|:-------|:--------|
| `hr.H.true` | Per-replicate Cox HR in the **true** $H$, observed data |
| `hr.H.hat` | Cox HR in the **identified** $\hat{H}$, observed data |
| `hr.H.bc` | Bootstrap bias-corrected HR in $\hat{H}$ |
| `ahr.H.hat` | AHR $= \exp\!\bigl(\overline{\log \text{HR}_i^{(\text{po})}}_{\hat{H}}\bigr)$ in the identified $\hat{H}$ |
| `cde.H.hat` | CDE in the identified $\hat{H}$ |
| `hr.Hc.true` | Per-replicate Cox HR in the **true** $H^c$ |
| `hr.Hc.hat` | Cox HR in the identified $\hat{H}^c$ |
| `hr.Hc.bc` | Bootstrap bias-corrected HR in $\hat{H}^c$ |
| `ahr.Hc.hat` | AHR in the identified $\hat{H}^c$ |
| `cde.Hc.hat` | CDE in the identified $\hat{H}^c$ |

**Relative bias** is computed against the DGM-level marginal HR truth
$\theta^\dagger_H = \mathtt{dgm\$hr\_H\_true}$:

$$
b^\dagger(\%) = 100 \times
\frac{\overline{\mathtt{hr.H.hat}} - \theta^\dagger_H}{\theta^\dagger_H}
$$

When CDE truth values $\theta^\ddagger$ are available,
`build_estimation_table()` reports a second bias column:

$$
b^\ddagger(\%) = 100 \times
\frac{\overline{\mathtt{hr.H.hat}} - \theta^\ddagger_H}{\theta^\ddagger_H}
$$

The `*` superscript in notation ($\hat{\theta}^*$) is reserved exclusively for
bootstrap bias-corrected estimates.  The decomposition separates: (1) **subgroup
misidentification** (identified $\hat{H} \neq H$) and (2) **estimation
variability** (finite-sample Cox coefficient noise).


# Complete Computation Pipeline

```{r}
#| label: pipeline-table
#| echo: false
formula_df <- data.frame(
  Step = c(
    "1. AFT model fit",
    "2. Hazard-scale transform",
    "3. Individual log-hazards",
    "4. Individual causal log-HR",
    "5a. AHR (subgroup $\\mathcal{S}$)",
    "5b. CDE (subgroup $\\mathcal{S}$)",
    "5c. Marginal HR (subgroup $\\mathcal{S}$)",
    "6. Theoretical HR (harm subgroup)",
    "7. Per-replicate truth (sim-level)",
    "8a. Per-replicate estimate (identified, Cox)",
    "8b. Per-replicate AHR (identified)",
    "8c. Per-replicate CDE (identified)"
  ),
  Formula = c(
    "$\\log T_i = \\mu + \\mathbf{X}_i'\\boldsymbol{\\gamma} + \\sigma\\varepsilon_i$",
    "$\\boldsymbol{\\beta}_0 = -\\boldsymbol{\\gamma}/\\sigma$",
    "$\\theta_i(x) = \\mathbf{X}_i(x)'\\boldsymbol{\\beta}_0$",
    "$\\log \\text{HR}_i^{(\\text{po})} = \\theta_i(1) - \\theta_i(0)$",
    "$\\text{AHR}(\\mathcal{S}) = \\exp\\!\\left(\\overline{\\log \\text{HR}_i^{(\\text{po})}}_{\\mathcal{S}}\\right)$",
    "$\\text{CDE}(\\mathcal{S}) = \\overline{\\exp(\\theta_1)}_{\\mathcal{S}} / \\overline{\\exp(\\theta_0)}_{\\mathcal{S}}$",
    "$\\exp(\\hat{\\beta}_{\\text{Cox}})$ from stacked potential outcomes",
    "$\\exp(\\beta_0[\\text{treat}] + \\beta_0[\\text{zh}])$",
    "<code>hr.H.true</code>: <code>coxph()</code> on observed data, <code>flag_harm == 1</code>",
    "<code>hr.H.hat</code>: <code>coxph()</code> on observed data, <code>sg_hat == 1</code>",
    "<code>ahr.H.hat</code>: $\\exp\\!\\bigl(\\overline{\\log \\text{HR}_i^{(\\text{po})}}_{\\hat{H}}\\bigr)$ for <code>sg_hat == 1</code>",
    "<code>cde.H.hat</code>: $\\overline{\\exp(\\theta_i(1))}_{\\hat{H}} / \\overline{\\exp(\\theta_i(0))}_{\\hat{H}}$ for <code>sg_hat == 1</code>"
  ),
  Code = c(
    "`survreg(Surv(y, event) ~ ..., dist = 'weibull')`",
    "`b0 <- -gamma / tau`",
    "`theta_0 <- X_ctrl %*% b0; theta_1 <- X_trt %*% b0`",
    "`log_hr_po <- theta_1 - theta_0  # stored as loghr_po`",
    "`exp(mean(df$loghr_po[idx]))`",
    "`mean(exp(df$theta_1[idx])) / mean(exp(df$theta_0[idx]))`",
    "`exp(coxph(Surv(t,e)~treat, data=df_stacked)$coef)`",
    "`exp(b0['treat'] + b0['zh'])`",
    "`exp(coxph(Surv(y,e)~treat, subset(df, flag_harm==1))$coef)`",
    "`exp(coxph(Surv(y,e)~treat, subset(df, sg_hat==1))$coef)`",
    "`exp(mean(df$loghr_po[df$sg_hat == 1]))`",
    "`mean(exp(df$theta_1[s])) / mean(exp(df$theta_0[s]))`"
  ),
  stringsAsFactors = FALSE
)

gt(formula_df) |>
  fmt_markdown(columns = c(Step, Formula)) |>
  cols_label(
    Step    = md("**Step**"),
    Formula = md("**Mathematical Formula**"),
    Code    = md("**R Implementation**")
  ) |>
  cols_width(Step ~ px(200), Formula ~ px(360), Code ~ px(350)) |>
  tab_header(
    title    = md("**Table 6. Treatment Effect Computation Pipeline**"),
    subtitle = "From AFT model fit to subgroup-level estimands"
  ) |>
  tab_style(style = cell_text(font = "monospace", size = "small"), locations = cells_body(columns = Code)) |>
  tab_style(style = cell_text(size = "small"), locations = cells_body(columns = c(Step, Formula))) |>
  tab_style(style = cell_text(weight = "bold", size = "small"), locations = cells_body(columns = Step)) |>
  tab_style(style = list(cell_fill(color = "#FFF8E1")), locations = cells_body(rows = 5:7)) |>
  tab_style(style = list(cell_fill(color = "#E8F4E8")), locations = cells_body(rows = 9:12)) |>
  tab_style(
    style = list(cell_fill(color = "#2C3E50"), cell_text(color = "white", weight = "bold")),
    locations = cells_column_labels()
  ) |>
  tab_options(table.font.size = px(12))
```


# Estimand Selection Guide {#selection-guide}

## Simulation Studies

**Use AHR as the primary estimand** for ForestSearch simulation studies.
It is deterministic, directly interpretable as the geometric-mean HR across
subjects, achieves dual causal status under the Fay & Li (2024) framework, and
is unaffected by conditioning-set distortion.  The WHI empirical evidence
supports the AHR/CHR scale's superior sensitivity to early or moderate effects
over RMST.

**Use the marginal HR** ($\theta^\dagger$, `dgm$hr_H_true`) for validation and
as the primary bias target in `build_estimation_table()`.  The stacked
potential-outcome Cox approach provides a familiar model-based estimand
confirming the DGM's causal structure produces the intended population-level
effects.

**Use CDE** ($\theta^\ddagger$) as a secondary bias target.  Its natural-scale
averaging parallels the Prentice & Aragaki AHR weighting structure, and it
appears as an estimation row in `build_estimation_table()` alongside AHR and
Cox HR rows.  The CDE is particularly informative when exploring treatment
effects across continuous biomarker thresholds via `cox_ahr_cde_analysis()`.

**Avoid interpreting per-replicate Cox HR causally.**  The `hr.H.hat` column
estimates the Cox AHR, which under non-PH depends on the censoring distribution,
is not individual-level, and conflates treatment effect with selection bias.
Its divergence from `dgm$hr_H_true` quantifies combined selection, finite-sample
noise, and censoring distortion.

## Applied Reporting

**Report CHR alongside AHR** as the primary population-level summary.  The
cumulative hazard ratio $\log\hat{S}_1(t)/\log\hat{S}_0(t)$ is the most
defensible under non-PH (Fay & Li; Wei & Schaubel).  The nonparametric AHR
tracks CHR closely in practice.

**Include RMST difference** as a clinically intuitive absolute time-to-event
summary.  It is simultaneously individual- and population-level causal, unambiguous
causally, but less sensitive than AHR/CHR to early effects in rare-outcome
prevention settings.

**GRF baseline covariate conditioning is theoretically justified** (Fay & Li
2024): adding covariates moves the estimated HR continuously toward an
individual-level causal quantity.  ForestSearch's GRF/LASSO conditioning is not
merely a statistical procedure — it is the mechanism by which individual-level
heterogeneous treatment effects become estimable and by which the frailty
aliasing in the Fay & Li vaccine example is resolved.

::: {.callout-warning icon=false}
## Notation Convention
Across all ForestSearch summary functions:
$H$ = true (oracle) subgroup; $\hat{H}$ = identified (data-driven) subgroup;
$*$ exclusively denotes bootstrap bias-corrected estimates ($\hat{\theta}^*$, never
"true subgroup").  The $\dagger$ superscript denotes the marginal HR truth;
$\ddagger$ denotes the CDE truth.
:::


# Key Takeaways

1. **The Fay & Li (2024) individual/population taxonomy** clarifies *why* the
   Cox HR causes difficulty: the instantaneous hazard at $t$ conditions on
   $\Omega_z(t) \neq \Omega_{z'}(t)$, violating the common comparison-set
   requirement for a causal estimand.

2. **Under PH**, the constant Cox HR achieves valid population-level causal
   status as $\log S_1(t)/\log S_0(t)$.  Even so, it is not individual-level.

3. **Under non-PH**, the cumulative hazard ratio is the preferred population-level
   causal estimand (Fay & Li; Wei & Schaubel).  The nonparametric AHR of
   Prentice & Aragaki tracks it closely in practice.

4. **The AFT $\exp(-\beta)$ is the only common survival ratio estimand that is
   simultaneously individual-level and population-level** (Fay & Li).  This
   geometric mean collapsibility underpins ForestSearch's choice of Weibull AFT
   as the causal DGM.

5. **The ForestSearch AHR** inherits the AFT's dual causal status, is
   deterministic, and is not subject to the conditioning-set problem.  It is the
   recommended primary simulation estimand.

6. **A declining observed HR over time cannot distinguish genuine effect waning
   from constant individual-level effects** in a heterogeneous population (Fay &
   Li vaccine example; Martinussen piecewise Cox example).  GRF/LASSO baseline
   conditioning resolves this aliasing.

7. **Hazard differences share the same causal limitation as hazard ratios**
   (Martinussen 2022).  The problem lies in interpreting the hazard function
   itself, not in its multiplicative structure.

8. **AHR is more sensitive than RMST** to early effects for rare outcomes.  The
   WHI breast cancer case — AHR detected risk elevation at 6 years vs. 12 for
   RMST — provides real-data validation that AHR/CHR-scale estimands are the
   right primary tools in prevention-trial subgroup detection.

9. **The Markov property fails under any non-degenerate frailty** (Knudsen et
   al. 2025), disqualifying Cox MSMs with time-varying treatment from clean
   causal interpretation.  ForestSearch's baseline-only design avoids this
   problem.

10. **Cox regression is a valid estimation engine but a problematic
    interpretive frame.**  All five foundational papers endorse the same division
    of labour: use Cox models for flexible hazard estimation; communicate causal
    effects through causally validated functionals (CHR, AHR, RMST, survival
    contrasts).


---

# References

- Aalen, O.O., Cook, R.J., & Røysland, K. (2015). Does Cox analysis of a
  randomized survival study yield a causal treatment effect? *Lifetime Data
  Analysis*, 21(4), 579--593.

- Abrahamowicz, M., Beauchamp, M.-E., Roberts, E.K., & Taylor, J.M.G. (2025).
  Revisiting the hazards of hazard ratios through simulations and case studies.
  *European Journal of Epidemiology*, 40, 611--629.

- Beyersmann, J., Schmoor, C., & Schumacher, M. (2025). Hazards constitute key
  quantities for analyzing, interpreting and understanding time-to-event data.
  *Biometrical Journal*, 67, e70057.

- Choi, J.-Y., Myung, R., & Lee, M.-J. (2025). Causal duration analysis based
  on survival probability ratio. *Sociological Methods & Research*.

- Edelmann, D. (2025). Revisiting hazard ratios: Can we define causal estimands
  for time-dependent treatment effects? *Biometrical Journal*, 67, e70100.

- Fay, M.P., & Li, F. (2024). Causal interpretation of the hazard ratio in
  randomized clinical trials. *Clinical Trials*, 21(5), 623--635.
  <https://doi.org/10.1177/17407745241243308>

- Hernán, M.A. (2010). The hazards of hazard ratios. *Epidemiology*, 21(1),
  13--15.

- Hernán, M.A., & Robins, J.M. (2020). *Causal Inference: What If*. Boca Raton:
  Chapman & Hall/CRC.

- Kalbfleisch, J.D., & Prentice, R.L. (1981). Estimation of the average hazard
  ratio. *Biometrika*, 68(1), 105--112.

- Kalbfleisch, J.D., & Prentice, R.L. (2002). *The Statistical Analysis of
  Failure Time Data* (2nd ed.). Wiley.

- Knudsen, M.B., Gabriel, E.E., Martinussen, T., Rytgaard, H.C.W., & Sjolander,
  A. (2025). On the limitations for causal inference in Cox models with
  time-varying treatment. *arXiv:2504.01524*.

- León, L.F., Jemielita, T., Guo, Z., Marceau West, R., & Anderson, K.M.
  (2024). Exploratory subgroup identification in the heterogeneous Cox model: A
  relatively simple procedure. *Statistics in Medicine*, 43(20), 3921--3942.
  <https://doi.org/10.1002/sim.10163>

- Martinussen, T. (2022). Causality and the Cox regression model. *Annual Review
  of Statistics and Its Application*, 9, 249--259.

- Martinussen, T., Vansteelandt, S., & Andersen, P.K. (2020). Subtleties in
  the interpretation of hazard contrasts. *Lifetime Data Analysis*, 26(4),
  833--855.

- Pearl, J. (2009). *Causality: Models, Reasoning, and Inference* (2nd ed.).
  Cambridge University Press.

- Prentice, R.L., & Aragaki, A.K. (2022). Intention-to-treat comparisons in
  randomized trials. *Statistical Science*, 37(3), 380--393.
  <https://doi.org/10.1214/21-STS830>

- Rubin, D.B. (1974). Estimating causal effects of treatments in randomized and
  nonrandomized studies. *Journal of Educational Psychology*, 66(5), 688--701.

- Simon, R., & Makuch, R.W. (1984). A non-parametric graphical representation
  of the relationship between survival and the occurrence of an event.
  *Statistics in Medicine*, 3(1), 35--44.

- Uno, H., Wittes, J., Fu, H., et al. (2015). Alternatives to hazard ratios for
  comparing the efficacy or safety of therapies in noninferiority studies.
  *Annals of Internal Medicine*, 163(2), 127--134.

- Wei, G., & Schaubel, D.E. (2008). Estimating cumulative treatment effects in
  the presence of nonproportional hazards. *Biometrics*, 64(3), 724--732.
