---
title: "Development Session Summary: GLM Extension Fixes"
subtitle: "ForestSearch v0.2.0 — Continuous Outcome Support and Reporting Infrastructure"
author: "Larry F. León"
date: 2025-04-14
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: true
    code-summary: "Show code"
    mainfont: "Palatino, Georgia, serif"
    monofont: "Fira Code, JetBrains Mono, Consolas, monospace"
    title-block-banner: "#2C3E50"
    title-block-banner-color: white
---

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = FALSE, comment = "#>")
```


# Overview

This document records the changes made during the April 14, 2025
development session on the `feature/glm-extension` branch.  The
session addressed three categories of issues: a critical estimator
bug in `run_simulation_analysis.R`, a family of reporting-layer
bugs in the benefit-search inversion and labeling code, and
structural cleanup of the Quarto document directory.

**Files modified:**

| File | Category | Severity |
|------|----------|----------|
| `R/run_simulation_analysis.R` | Estimator bug | Critical |
| `R/simulation_tables.R` | Reporting / labels | High |
| `R/oc_analyses.R` | Reporting / labels | High |
| `quarto/simulations/actg175_continuous_simulations.qmd` | Worker reset + `collect_results` | Moderate |
| `quarto/` (directory structure) | Organization | Maintenance |

All changes maintain backward compatibility with survival and
binary outcome types.  Existing simulation documents that use
HR or OR measures produce identical output.


# Bug 1: GLM Estimator Uses Wrong Response Variable {#sec-estimator}

## Symptom

The estimation table for continuous simulations showed all zeros:

| Estimator | Avg | SD | Min | Max | b† (%) |
|-----------|-----|----|-----|-----|--------|
| θ̂(Q̂)     | 0   | 0  | 0   | 0   | -100   |
| θ̂(Q̂ᶜ)    | 0   | 0  | 0   | 0   | -100   |

## Root Cause

Two estimator closures in `run_simulation_analysis.R` used
`event_name` as the GLM response variable instead of
`outcome_name`.

In `.extract_fs_estimates_gen()` (ForestSearch path, line 556):

```r
# BEFORE (buggy):
fit <- stats::glm(sub_df[[event_name]] ~ sub_df[[treat_name]],
                  family = glm_family)

# AFTER (fixed):
fit <- stats::glm(sub_df[[outcome_name]] ~ sub_df[[treat_name]],
                  family = glm_family)
```

The same bug existed in `.extract_grf_estimates_gen()` (GRF path,
line 861), where the closure parameter `ec` (event column) was
used instead of `oc` (outcome column).

## Why It Caused Zeros

For continuous outcomes, `run_simulation_analysis()` creates
`event_sim = 1L` for all observations (a compatibility shim for
downstream Cox code).  Fitting `glm(1 ~ treat_sim, family = gaussian())`
gives a treatment coefficient of exactly zero, because the
response is constant.

## Why Binary Was Unaffected

The binary simulation document explicitly sets
`event.name = "y_sim"` in `fs_params`, making
`event_name == outcome_name` by coincidence.  The continuous
document correctly omits `event.name` (no concept of events
for continuous data), so the default `"event_sim"` exposed
the bug.

## Verification

QC test on synthetic data with true MD = 40:

```r
# Fixed closure: MD = 41.13 (correct)
# Buggy closure: MD = 0.00  (constant response)
```


# Bug 2: Benefit-Search Inversion for Identity-Scale Measures {#sec-inversion}

## Symptom

The estimation table header showed `θ†(Q) = 0.03` instead of
the expected `θ†(Q) = -39.7` (negated from the switched-scale
value of +39.7).  Estimation values were `Inf` and `NaN`.

## Root Cause

`invert_hr_columns()` and `invert_dgm_hrs()` in
`simulation_tables.R` unconditionally used `1/x` to convert
from switched → original scale.  This is correct for
ratio-scale measures (HR, OR, IRR) but wrong for
identity-scale measures (MD, RD, IRD), where the correct
inversion is negation (`-x`).

| Measure | Scale | Correct Inversion | Old (buggy) |
|---------|-------|-------------------|-------------|
| HR      | Ratio | 1/1.85 = 0.54     | 1/1.85 = 0.54 ✓ |
| OR      | Ratio | 1/2.0 = 0.50      | 1/2.0 = 0.50 ✓ |
| **MD**  | Identity | **-(39.7) = -39.7** | **1/39.7 = 0.025** ✗ |
| **RD**  | Identity | **-(0.12) = -0.12** | **1/0.12 = 8.33** ✗ |
| IRR     | Ratio | 1/1.5 = 0.67      | 1/1.5 = 0.67 ✓ |
| **IRD** | Identity | **-(0.05) = -0.05** | **1/0.05 = 20** ✗ |

For MD values near zero, `1/x` produces `Inf`, explaining
the NaN values in the original output.

## Fix

Added `effect_measure` parameter to both functions with
auto-detection from `dgm$effect_measure`:

```r
# New helper
is_identity_scale <- function(effect_measure) {
  !is.null(effect_measure) && effect_measure %in% c("MD", "RD", "IRD")
}

# Updated inversion
invert_hr_columns <- function(res, effect_measure = NULL) {
  use_negate <- is_identity_scale(effect_measure)
  for (col in intersect(hr_cols, names(res))) {
    res[[col]] <- if (use_negate) -res[[col]] else 1 / res[[col]]
  }
  res
}
```

All six call sites across `simulation_tables.R` and
`oc_analyses.R` were updated to thread `effect_measure`.


# Bug 3: Hardcoded Survival Labels in Reporting Functions {#sec-labels}

## Symptom

Binary outcome estimation tables displayed footnotes referencing
"plugin Cox HR" and "average hazard ratio" — survival-specific
terminology applied to binary and continuous outcomes.

## Fix

Added `effect_measure_label()` helper and GLM-aware branching
throughout the reporting layer:

| Function | What changed |
|----------|-------------|
| `build_estimation_table()` | Footnote branches: GLM → "plugin OR estimate"; survival → "plugin Cox HR" |
| `interpret_estimation_table()` | All 5 narrative paragraphs use `em_label` instead of hardcoded "Cox HR" |
| `format_oc_results()` | HR column spanner adapts (e.g., "GLM OR" for binary) |

The `effect_measure_label()` function returns context-appropriate
labels:

| outcome_type | effect_measure | Label |
|-------------|----------------|-------|
| survival | HR | Cox HR |
| binary | OR | OR |
| binary | RD | risk difference |
| continuous | MD | mean difference |
| count | IRR | IRR |
| count | IRD | incidence rate difference |

AHR and CDE paragraphs in `interpret_estimation_table()` are
naturally gated by `has_ahr_H` and `has_cde_interp`, which are
`FALSE` for GLM outcomes (those columns are all `NA`).  No
explicit suppression was needed.


# Enhancement: `dgm` Parameter for `format_oc_results()` {#sec-dgm}

`format_oc_results()` previously had no way to auto-detect the
effect measure.  Added an optional `dgm` parameter:

```r
format_oc_results(
  results,
  dgm = dgm_calibrated,       # NEW: auto-detects effect_measure
  subgroup_notation = "benefit"
)
```

When `dgm` is provided, `effect_measure` is auto-detected from
`dgm$effect_measure`, matching the API of `build_estimation_table()`
and `build_classification_table()` which already accept `dgm`.
When `dgm = NULL` (default), backward-compatible ratio-scale
inversion is used.


# Continuous Simulations Document Updates {#sec-continuous}

## Worker Pool Reset

Added `reset-workers-null` chunk between the alt and null
simulation loops, matching the pattern in all three other
simulation documents:

```r
# Reset worker pool to prevent memory accumulation from alt sims.
# Without this, null sims run on bloated workers and can be ~20x slower.
plan(sequential); gc(verbose = FALSE); Sys.sleep(1)
plan(multisession, workers = n_workers)
```

## `collect_results()` Migration

Removed the 24-line inline `collect_results()` definition and
switched to the package-internal version:

```r
# Before:
collect_results <- function(raw_list, label = "") { ... }
results_alt <- collect_results(results_alt_raw, "H1 alt")

# After:
# collect_results() now lives in forestsearch (R/forestsearch_helpers.R).
results_alt <- forestsearch:::collect_results(results_alt_raw, "H1 alt")
```


# Quarto Directory Reorganization {#sec-reorg}

## Before

73 files in a flat `quarto/` directory: production documents,
working copies, QC scripts, guides, and archive material all
intermixed.

## After

| Directory | Contents | Files |
|-----------|----------|-------|
| `simulations/` | Production simulation studies | 8 |
| `calibration/` | L_eff calibration and threshold selection | 12 |
| `grf/` | GRF standalone calibration and review | 9 |
| `applications/` | Applied analyses and validation | 15 |
| `guides/` | Theory, literature, how-to guides | 22 |
| `qc/` | QC and test scripts (R) | 14 |
| `archive/` | Superseded working copies | 15 |

Root now contains only `_quarto.yml`, `_data/`, `README.md`,
`.gitignore`, and the 7 topic directories.

## Key Design Decisions

**Co-located output**: `output-dir` was removed from
`_quarto.yml`, so Quarto renders HTML alongside each `.qmd`
file.  Each topic directory contains both source and output.

**Shared data cache**: `_data/` remains at the `quarto/` root.
All moved documents had `output_dir <- "_data"` updated to
`output_dir <- "../_data"` via `sed`.  Functions like
`save_leff_calibration()` and `load_leff_calibration()` that
default to `output_dir = "_data"` will need explicit
`output_dir = "../_data"` when called from subdirectories.

**Git history preserved**: All moves used `git mv`, so GitHub
shows renames rather than delete-and-create.


# Summary of Outcome/Measure Coverage {#sec-matrix}

After all fixes, the reporting layer handles every
outcome/measure combination correctly:

| Outcome | Measure | Inversion | Footnote | Spanner | AHR/CDE |
|---------|---------|-----------|----------|---------|---------|
| Survival | HR | 1/x | plugin Cox HR | Cox Hazard Ratios | Shown |
| Binary | OR | 1/x | plugin OR estimate | GLM OR | Suppressed |
| Binary | RD | −x | plugin RD estimate | RD | Suppressed |
| Continuous | MD | −x | plugin MD estimate | MD | Suppressed |
| Count | IRR | 1/x | plugin IRR estimate | GLM IRR | Suppressed |
| Count | IRD | −x | plugin IRD estimate | IRD | Suppressed |


# Session Info

```{r}
#| echo: true
sessionInfo()
```
