---
title: "ForestSearch Outcome Orientation Reference"
subtitle: "Setting Up Harm and Benefit Searches Across All Endpoint Types"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")
library(forestsearch)
library(ggplot2)
library(gt)
```


# The Core Rule {#sec-rule}

ForestSearch uses a single comparison direction throughout the
entire search pipeline:

$$
\hat{\theta}(S) \;\geq\; c_1
$$

where $\hat{\theta}(S)$ is the estimated treatment effect in
candidate subgroup $S$ and $c_1$ is the screening threshold
(`effect.threshold`).  The consistency check uses the same
direction: each bootstrap split must satisfy
$\hat{\theta}(S_b) \geq c_2$.

**This comparison never changes.**  There is no internal sign flip
for the core search.  All scenario setup reduces to one question:

> *How do I orient my data so that the subgroup I want to find
> produces $\hat{\theta}(S) > 0$?*


## What the Parameters Control

| Parameter | Where it acts | What it does |
|-----------|---------------|--------------|
| `effect.threshold` | `subgroup_search()` | Screening: candidate must have $\hat{\theta} \geq c_1$ |
| `consistency.threshold` | `sg_consistency_out()` | Bootstrap: each split must have $\hat{\theta} \geq c_2$ |
| `adverse_outcome` | GRF variable importance only | Flips $Y$ for `causal_forest()` pre-screening |
| `outcome_type` | Estimator selection | Selects GLM family (logistic, Gaussian, Poisson) |
| `effect_measure` | Scale of $\hat{\theta}$ | OR/RR/RD for binary; MD for continuous; IRR/IRD for count |

**Critical**: `adverse_outcome` does **not** affect the core search
comparison.  It only controls whether GRF sees $Y$ or $1-Y$ when
computing variable importance scores.


# Effect Scale and Threshold Conventions {#sec-scales}

## Ratio-Scale Measures (OR, RR, IRR, HR)

ForestSearch works on the **log scale** internally.  When you set
`effect.threshold = 1.25`, it is converted to $\log(1.25) = 0.223$
before comparison.  The estimator returns $\log(\hat{\theta})$, so
the comparison is:

$$\log(\hat{\theta}) \geq \log(c_1)$$

- $\hat{\theta} > 1$: treatment increases the outcome (events, odds, rate)
- $\hat{\theta} < 1$: treatment decreases the outcome
- $\hat{\theta} = 1$: no treatment effect

## Identity-Scale Measures (MD, RD, IRD)

No log transformation.  The comparison is directly:

$$\hat{\theta} \geq c_1$$

For **RD** and **IRD**: bounded in $[-1, 1]$.  ForestSearch warns if
$c_1 > 1$ and remaps to a default.

For **MD** (mean difference): **unbounded**.  Thresholds like 30
cells/$\mu$L are legitimate.  ForestSearch does **not** warn or remap
for MD.  (This was a bug fixed in the current version — MD was
previously treated like RD.)


# The Eight Scenarios {#sec-scenarios}

Every subgroup search involves three binary choices:

1. **Search type**: Harm (find $S$ where treatment hurts) vs Benefit
   (find $S$ where treatment helps)
2. **Outcome orientation**: Adverse (higher $Y$ = worse, e.g., death,
   events) vs Positive (higher $Y$ = better, e.g., response, CD4)
3. **Endpoint type**: Survival (HR), Binary (OR), Continuous (MD),
   Count (IRR)

For **survival endpoints**, the "outcome orientation" maps to the
clinical meaning of the event:

- **Adverse event** (death, recurrence): HR > 1 = treatment
  increases hazard = harm.  This is the standard clinical trial
  convention.
- **Positive event** (recovery, remission): HR > 1 = treatment
  accelerates the good event = benefit.  Less common but valid.

This gives $2 \times 2 \times 4 = 16$ combinations, but the setup
rules reduce to a simple decision tree.

```{r decision-tree}
#| fig-height: 4
#| fig-cap: "Decision tree for ForestSearch setup."

scenarios <- data.frame(
  search      = rep(c("Harm", "Benefit"), each = 4),
  outcome     = rep(c("Adverse\n(higher=worse)", "Positive\n(higher=better)"), 4),
  action      = c(
    "Direct: effect > 0 ✓",            # Harm + Adverse
    "Negate Y or\nswitch treatment",    # Harm + Positive
    "Switch treatment\n+ negate Y",     # Benefit + Adverse
    "Direct: switch\ntreatment only"    # Benefit + Positive
  ),
  example     = c(
    "GBSG: status (event)\nhormon (treat)\nHR > 1 = harm",
    "Rare: negate Y so\nhigher = worse",
    "Rare: switch treat\nand negate Y",
    "ACTG: y_binary (improve)\nswitch treat\nOR > 1 on switched"
  ),
  color       = c("green3", "orange", "orange", "green3"),
  stringsAsFactors = FALSE
)

cat("Decision tree displayed as table below.\n")
```


## The Decision Table

```{r scenario-table1}
#| label: tbl-scenarios1
#| tbl-cap: "Complete setup guide for all ForestSearch scenarios."

tbl_data <- data.frame(
  Search = c(
    "Harm", "Harm", "Harm", "Harm",
    "Harm", "Harm",
    "Benefit", "Benefit", "Benefit", "Benefit",
    "Benefit", "Benefit"
  ),
  Outcome = c(
    "Adverse (↑Y = worse)", "Adverse (↑Y = worse)",
    "Adverse (↑Y = worse)",
    "Positive (↑Y = better)", "Positive (↑Y = better)",
    "Positive (↑Y = better)",
    "Adverse (↑Y = worse)", "Adverse (↑Y = worse)",
    "Adverse (↑Y = worse)",
    "Positive (↑Y = better)", "Positive (↑Y = better)",
    "Positive (↑Y = better)"
  ),
  Endpoint = c(
    "Survival (adverse event)", "Binary/Count", "Continuous",
    "Survival (positive event)", "Binary/Count", "Continuous",
    "Survival (adverse event)", "Binary/Count", "Continuous",
    "Survival (positive event)", "Binary/Count", "Continuous"
  ),
  Setup = c(
    "Direct: HR > 1 = harm",
    "Direct: no changes needed",
    "Direct: no changes needed",
    "Switch treatment (HR > 1 = delays good event = harm)",
    "Negate Y: use 1−Y (binary)",
    "Negate Y: use −Y",
    "Switch treatment (HR > 1 on switched = original helps)",
    "Switch treatment labels",
    "Switch treatment labels",
    "Direct: HR > 1 = accelerates good event = benefit",
    "Switch treatment labels",
    "Switch treatment labels"
  ),
  Effect_Sign = c(
    "HR(Q) > 1 naturally",
    "θ(Q) > 0 naturally",
    "θ(Q) > 0 naturally",
    "HR(Q) > 1 after switch",
    "θ(Q) > 0 after negation",
    "θ(Q) > 0 after negation",
    "HR(Q) > 1 after switch",
    "θ(Q) > 0 after switch",
    "θ(Q) > 0 after switch",
    "HR(Q) > 1 naturally",
    "θ(Q) > 0 after switch",
    "θ(Q) > 0 after switch"
  ),
  Example = c(
    "GBSG: status=recurrence, hormon=treat, HR>1=harm",
    "Poisson events, IRR > 1 = more events",
    "BP increase: MD > 0 = treatment raises BP",
    "Time to recovery: switch treat, HR>1=delays recovery",
    "Binary response: use 1−Y (failure)",
    "CD4 improvement: use −Y (CD4 loss)",
    "GBSG: switch treat, HR>1=no-hormone increases hazard",
    "ACTG: switch treat, OR>1 on switched scale",
    "ACTG: switch treat, MD>0 on switched scale",
    "Time to recovery: HR>1=accelerates recovery (benefit)",
    "Rare: switch treat for positive binary",
    "Rare: switch treat for positive continuous"
  ),
  stringsAsFactors = FALSE
)

names(tbl_data) <- c("Search", "Outcome", "Endpoint", "Setup",
                      "Effect Sign", "Example")

tbl_data |>
  gt() |>
  tab_header(
    title = "ForestSearch Setup Guide",
    subtitle = "How to orient data for each search scenario"
  ) |>
  tab_style(
    style = cell_fill(color = "#e8f5e9"),
    locations = cells_body(rows = c(1, 2, 7, 8))
  ) |>
  tab_style(
    style = cell_fill(color = "#fff3e0"),
    locations = cells_body(rows = c(3, 4, 5, 6))
  ) |>
  tab_footnote(
    footnote = "Green rows: direct setup (no outcome manipulation needed).",
    locations = cells_column_labels(columns = "Setup")
  ) |>
  tab_footnote(
    footnote = "Orange rows: outcome negation and/or treatment switching required.",
    locations = cells_column_labels(columns = "Setup")
  ) |>
  cols_width(Setup ~ px(250), Example ~ px(280)) |>
  tab_options(table.font.size = 13)
```


# Worked Examples {#sec-examples}

## Harm Search + Adverse Outcome (Direct) {#sec-harm-adverse}

**The simplest case.** Higher $Y$ is already worse, and we want to
find where treatment makes it even worse.

### Survival: GBSG Recurrence

```
# Y = time to recurrence (shorter = worse, but Cox HR > 1 = worse)
# treat = hormonal therapy
# Event = recurrence (adverse)
# Subgroup where HR > 1: treatment increases hazard
forestsearch(
  outcome.name = "rfstime",
  event.name   = "status",       # 1 = event (adverse)
  treat.name   = "hormon",
  effect.threshold = 1.25        # HR >= 1.25
)
```

**DGM setup**: `generate_aft_dgm_flex()` with positive `k_inter`.
Positive `k_inter` increases hazard for $Q$ under treatment → HR > 1.

**GRF pre-screening**: Uses `causal_survival_forest()`, which
takes `(time, event, horizon)` and internally estimates treatment
effects on survival probability.  Events are assumed adverse:
positive CATE = treatment increases survival = treatment helps.
Variable importance identifies covariates with the most treatment
effect heterogeneity.  `adverse_outcome` is not used in this path.

### Count: GBSG Poisson Rate Model

```
# Y = status (0/1 event count)
# offset = log(time_months)
# IRR > 1 = treatment increases event rate (adverse)
forestsearch(
  outcome.name     = "status",
  event.name       = "status",
  treat.name       = "hormon",
  outcome_type     = "count",
  effect_measure   = "IRR",
  offset.name      = "time_months",
  effect.threshold = 1.25        # IRR >= 1.25
)
```

**DGM setup**: `generate_glm_dgm(outcome_type = "count", k_inter = 0.5)`.
Positive `k_inter` shifts log-rate up for $Q$ → IRR > 1.

**Result**: 80% detection, 3.5% FPR (confirmed working).


## Harm Search + Positive Outcome (Negate Y) {#sec-harm-positive}

Higher $Y$ is better, but we want to find where treatment hurts.
This is **unusual** — typically if you're searching for harm, the
outcome is already adverse.  But it arises when, e.g., you want to
find where treatment reduces a beneficial outcome.

### Binary: Treatment Reduces Response Rate

```
# Y = 1 if patient responds (positive)
# We want: find S where treatment REDUCES P(response)
# Problem: OR < 1 when treatment hurts → fails θ ≥ c₁
# Fix: negate Y
data$y_failure <- 1L - data$y_response
forestsearch(
  outcome.name     = "y_failure",
  event.name       = "y_failure",
  outcome_type     = "binary",
  effect_measure   = "OR",
  effect.threshold = 1.5         # OR(failure) >= 1.5
)
```

### Continuous: Treatment Reduces CD4

```
# Y = cd420 - cd40 (positive = improvement)
# We want: find S where treatment REDUCES CD4 change
# Fix: negate Y
data$cd4_loss <- -(data$cd420 - data$cd40)
forestsearch(
  outcome.name     = "cd4_loss",
  outcome_type     = "continuous",
  effect_measure   = "MD",
  effect.threshold = 30          # MD(loss) >= 30
)
```


## Benefit Search + Positive Outcome (Switch Treatment) {#sec-benefit-positive}

**The ACTG175 pattern.** Higher $Y$ is better, and we want to find
where the *experimental* treatment helps more than control.

### The Treatment Switching Trick

ForestSearch always searches for $\hat{\theta} \geq c_1$.
If we keep the original treatment labels, the benefit subgroup has
$\hat{\theta} < 0$ (control is worse) and fails the comparison.

**Fix**: Swap treatment and control labels.  Under the switched
labels, the subgroup where the original treatment helps now appears
as a subgroup where the "new treatment" (actually the control) hurts.
The effect flips sign and exceeds the threshold.

### Survival: Benefit Search (Adverse Event)

This is the standard clinical trial benefit search.  The event is
adverse (death, recurrence), and we want to find $Q$ where the
experimental treatment reduces hazard.

```
# Event = recurrence (adverse)
# treat_orig = hormonal therapy
# Goal: find Q where hormonal therapy REDUCES recurrence

# Switch treatment labels
data$treat_sw <- 1L - data$hormon

# Now: under switched labels, HR > 1 in Q means
# the original control (no hormone) increases hazard
# → equivalently, hormone REDUCES hazard for Q

forestsearch(
  outcome.name     = "rfstime",
  event.name       = "status",
  treat.name       = "treat_sw",        # switched!
  effect.threshold = 1.25               # HR ≥ 1.25
)
```

**DGM setup**: `generate_aft_dgm_flex()` with positive `k_inter`
and the switched treatment column.  Positive `k_inter` increases
hazard for $Q$ under the switched treatment (= no hormone) →
HR > 1 on the switched scale.

### Survival: Positive Event (Time to Recovery)

In rare cases the survival event is positive — time to remission,
time to discharge.  Here HR > 1 = treatment **accelerates** the
good event, which is beneficial.

```
# Event = remission (positive)
# HR > 1 = treatment accelerates remission = treatment helps
# This is ALREADY a benefit search — no switching needed!

forestsearch(
  outcome.name     = "time_to_remission",
  event.name       = "remission_status",
  treat.name       = "treat",           # original labels!
  effect.threshold = 1.25               # HR ≥ 1.25 (beneficial)
)
```

For a **harm** search with a positive event (find $Q$ where
treatment delays recovery), you would switch treatment labels so
that HR > 1 under the switched labels means the original treatment
delays the positive event.

### Binary: ACTG175 CD4 Improvement

```
# Arms: ZDV+ddI (combo, arm 1) vs ddI (mono, arm 3)
# Y = I(cd420 > cd40)  (improvement, positive)
# Goal: find S where combo helps more than mono

# Step 1: Switch treatment labels
data$treat <- 1L - data$treat_orig   # ddI = 1, combo = 0

# Step 2: No outcome negation needed (Y is positive,
#          and switching makes the "harm" subgroup be
#          where ddI hurts, i.e., combo helps)

forestsearch(
  outcome.name     = "y_binary",     # improvement (0/1)
  event.name       = "y_binary",
  treat.name       = "treat",        # switched!
  outcome_type     = "binary",
  effect_measure   = "OR",
  effect.threshold = 1.667           # OR >= 1/0.60
)
```

**DGM setup**:

```
# Target OR(Q) = 2.0 on the switched scale
# Positive k_inter: ddI (treat=1) increases P(improve) for Q
# This means Q is where ddI helps — equivalently, combo doesn't help Q
# Wait: that's the OPPOSITE of what we want!
#
# Actually: with switched labels (ddI = treat),
# positive k_inter increases P(Y=1|Q, ddI)
# → OR(switched) > 1 in Q
# → ForestSearch finds Q as "harmed by ddI"
# → Reporting inverts: Q benefits from combo
calibrate_glm_interaction(
  target_effect = 2.0,           # OR(Q, switched) = 2.0
  k_inter_range = c(0, 3),      # positive grid
  outcome_type  = "binary",
  effect_measure = "OR"
)
```

**Result**: 75% detection, 6% FPR (confirmed working).

### Continuous: ACTG175 CD4 Change

```
# Y = cd420 - cd40 (CD4 change, positive = improvement)
# Goal: find S where combo produces more improvement

# Step 1: Switch treatment labels
data$treat <- 1L - data$treat_orig

# Step 2: No outcome negation — with switching,
#         positive MD means ddI increases CD4 change
#         for Q more than combo does.
#         ForestSearch finds Q where MD > threshold.

forestsearch(
  outcome.name     = "cd4_change",
  outcome_type     = "continuous",
  effect_measure   = "MD",
  effect.threshold = 30           # MD >= 30 cells/µL
)
```

**DGM setup**:

```
calibrate_glm_interaction(
  outcome_var   = "cd4_change",
  target_effect = 40,            # MD(Q, switched) = +40
  k_inter_range = c(0, 120),    # positive grid
  outcome_type  = "continuous",
  effect_measure = "MD"
)
```


## Benefit Search + Adverse Outcome (Switch + Negate) {#sec-benefit-adverse}

Higher $Y$ is worse (e.g., events), but we want to find where
the experimental treatment helps.  This requires both switching
treatment labels **and** negating the outcome.

This is the rarest scenario — typically if the outcome is adverse,
you'd search for harm rather than benefit.


# The `adverse_outcome` Parameter {#sec-adverse}

## Where It Acts

`adverse_outcome` has **two** effects, in different parts of the
pipeline:

### Effect 1: Binary Estimator Outcome Flip

In `make_effect_estimator()` → `.make_binary_estimator()` (line 303):

```r
if (!adverse_outcome) {
  data_slice[[outcome.name]] <- 1L - data_slice[[outcome.name]]
}
```

When `adverse_outcome = FALSE`: the estimator flips $Y \to 1 - Y$
before computing OR/RD/RR.  This converts a positive outcome
(response) to its failure complement, ensuring that OR > 1 means
"treatment increases failure" (harm).

When `adverse_outcome = TRUE`: no flip.  The estimator sees $Y$
as-is.  If $Y$ is truly adverse (events), OR > 1 = treatment
increases events = harm.  If $Y$ is positive (improvement) but
you're using switched treatment, OR > 1 = switched-treatment
increases improvement = the subgroup you want.

**This flip does NOT apply to continuous or count estimators.**
Those always use $Y$ as-is.

### Effect 2: GRF Variable Importance Flip

In `forestsearch()` (line ~1172):

```r
Y_vi <- if (adverse_outcome) 1L - Y else Y
```

This is the **opposite** direction from Effect 1.
`causal_forest()` estimates CATE where positive = treatment helps.
When `adverse_outcome = TRUE`, $Y$ represents bad outcomes, so we
flip to $1 - Y$ (good scale) for GRF.  When `adverse_outcome = FALSE`,
$Y$ is already on the good scale.

**This flip applies only to binary and continuous GLM outcomes.**
Survival uses `causal_survival_forest()` which handles the
survival outcome natively.  Count outcomes use a log transform.

### Effect 3: Survival — `causal_survival_forest()`

For Cox/survival analyses, `adverse_outcome` does not directly
appear in the code path.  The GRF pre-screening uses
`causal_survival_forest()`, which takes `(Y = time, Event, horizon)`
and estimates treatment effects on survival probability.

**Implicit assumption**: `causal_survival_forest()` treats events
as adverse — positive CATE = treatment increases survival
probability = treatment helps.  This is correct when the event
is death, recurrence, or any adverse outcome.

**When the event is positive** (e.g., recovery, remission): the
CATE sign is reversed.  Positive CATE = treatment delays recovery
= treatment hurts.  Variable importance is based on magnitude, so
pre-screening still identifies heterogeneous variables, but the
sign interpretation is flipped.  The Cox HR comparison
($\text{HR} \geq c_1$) remains correct because HR > 1 means
treatment accelerates the (positive) event.


## Summary Table

```{r adverse-outcome-table}
ao_tbl <- data.frame(
  Endpoint = c(
    "Survival (adverse event)",
    "Survival (positive event)",
    "Binary (adverse: events)",
    "Binary (positive: response)",
    "Continuous (adverse: pain)",
    "Continuous (positive: CD4)",
    "Count (adverse: events)",
    "Count (positive: recoveries)"
  ),
  Default = c(
    "N/A", "N/A",
    "TRUE", "TRUE",
    "FALSE", "FALSE",
    "TRUE", "TRUE"
  ),
  Correct_Setting = c(
    "N/A (handled by causal_survival_forest)",
    "N/A (HR direction is naturally correct)",
    "TRUE (no flip needed)",
    "TRUE if using switched treatment\nFALSE if NOT switching treatment",
    "TRUE (flip for GRF: higher pain = worse)",
    "FALSE (no flip: higher CD4 = better)",
    "TRUE (no flip needed)",
    "FALSE (flip for GRF: higher recoveries = better)"
  ),
  Estimator_Flip = c(
    "N/A", "N/A",
    "No", "No (TRUE) / Yes (FALSE)",
    "N/A", "N/A",
    "N/A", "N/A"
  ),
  GRF_Flip = c(
    "causal_survival_forest",
    "causal_survival_forest",
    "Yes (1−Y)", "No",
    "Yes (1−Y)", "No",
    "log transform", "log transform"
  ),
  stringsAsFactors = FALSE
)

ao_tbl |>
  gt::gt() |>
  gt::tab_header(
    title = "adverse_outcome: What It Does for Each Endpoint",
    subtitle = "Two independent effects: estimator flip (binary only) and GRF flip"
  ) |>
  gt::cols_label(
    Correct_Setting = "Recommended",
    Estimator_Flip = "Flips Y in Estimator?",
    GRF_Flip = "GRF Pre-screening"
  ) |>
  gt::tab_options(table.font.size = 12) |>
  gt::cols_width(Correct_Setting ~ gt::px(220))
```


## The ACTG175 Binary Benefit Case

The working ACTG175 binary benefit simulation uses:

- `Y = I(cd420 > cd40)` — improvement (positive outcome)
- `treat = 1 - treat_orig` — switched (ddI = treat)
- `adverse_outcome` — not set, defaults to `TRUE` for binary

With `adverse_outcome = TRUE`:

1. **Binary estimator**: no flip.  OR computed on improvement.
   OR > 1 = ddI increases P(improve) for Q.
2. **GRF**: flips to $1 - Y$ (failure).  `causal_forest` sees
   failure, identifies variables where treatment increases failure
   (= variables where ddI hurts = where combo helps).
3. **DGM**: positive `k_inter` increases P(improve|Q, ddI) →
   OR(improve) > 1 in Q.
4. **ForestSearch**: finds Q where OR ≥ threshold.
5. **Reporting**: `subgroup_notation = "benefit"` inverts labels.

This works because the estimator sees improvement directly, and
the DGM creates a positive effect on improvement for Q under the
switched treatment.

**What would happen with `adverse_outcome = FALSE`?**

The estimator would flip Y → 1−Y (failure).  Now:
- ddI increasing P(improve|Q) → ddI decreasing P(fail|Q)
- OR(fail) < 1 in Q → fails $\hat{\theta} \geq c_1$
- Zero detection.

This is why the "broken" binary simulation (which set
`adverse_outcome = FALSE`) produced garbage results.


# Common Pitfalls {#sec-pitfalls}

## Pitfall 1: Wrong Calibration Target Direction

::: {.callout-warning}
**Symptom**: 100% detection AND 97% false positive rate.

**Cause**: Calibration target on the wrong scale (e.g.,
`cal_target_or = 0.5` with `k_inter_range = c(-3, 0)`).
This makes Q only slightly more extreme in the same direction
as everyone — no real contrast.

**Fix**: Use target > 1 with positive `k_inter` for ratio measures,
or positive target with positive grid for MD.
:::

## Pitfall 2: MD Threshold Remapped to 0.05

::: {.callout-warning}
**Symptom**: Zero detections or near-100% detections for continuous
outcomes with `effect_measure = "MD"`.

**Cause**: The threshold validator treated MD like RD (bounded in
$[-1, 1]$) and remapped `effect.threshold = 30` to `0.05`.

**Fix**: Fixed in current version — MD is now correctly treated as
unbounded.
:::

## Pitfall 3: Confusing `adverse_outcome` with Outcome Negation

::: {.callout-warning}
**Symptom**: Zero detection with benefit search and positive outcome.

**Cause**: Setting `adverse_outcome = FALSE` when using a positive
binary outcome with switched treatment.  This causes the **binary
estimator** to flip $Y \to 1-Y$ internally, which reverses the
effect sign and breaks the $\hat{\theta} \geq c_1$ comparison.

**The working ACTG175 binary simulation** leaves `adverse_outcome`
at its default (`TRUE` for binary).  Despite the name suggesting
"the outcome is adverse," the actual meaning is: **keep $Y$ as-is
for the estimator** (no flip).

**Rule of thumb**: For benefit search with switched treatment,
use `adverse_outcome = TRUE` (or leave as default for binary).
Only set `FALSE` if you explicitly want the estimator to compute
effects on the $1-Y$ scale.
:::

## Pitfall 4: Survival with Positive Events

::: {.callout-warning}
**Symptom**: GRF pre-screening selects wrong variables for a
survival outcome where the event is positive (e.g., recovery).

**Cause**: `causal_survival_forest()` implicitly assumes events
are adverse.  When the event is positive, the CATE sign is
reversed, but variable importance is based on magnitude, so
pre-screening still identifies heterogeneous variables.

**Impact**: Low — variable importance is magnitude-based, not
sign-based.  The Cox HR comparison ($\text{HR} \geq c_1$) is
unaffected.  But if GRF pre-screening is the bottleneck, consider
switching to LASSO-only screening (`use_grf = FALSE`).
:::


# Quick Reference Card {#sec-quickref}

```{r quickref}
#| label: tbl-quickref
#| tbl-cap: "Quick reference: parameter settings for each scenario."

qr <- data.frame(
  Scenario = c(
    "GBSG Cox harm (adverse event)",
    "GBSG Cox benefit (adverse event, switched)",
    "Survival benefit (positive event, direct)",
    "Survival harm (positive event, switched)",
    "GBSG Poisson (harm, adverse)",
    "ACTG binary benefit (positive)",
    "ACTG continuous benefit (positive)",
    "Binary harm + positive outcome",
    "Continuous harm + positive outcome"
  ),
  Y = c(
    "rfstime/status",
    "rfstime/status",
    "time_to_recovery/recovered",
    "time_to_recovery/recovered",
    "status (0/1)",
    "I(cd420 > cd40)",
    "cd420 - cd40",
    "1 - response",
    "-(improvement)"
  ),
  Treatment = c(
    "hormon (original)",
    "1 - hormon (switched)",
    "original",
    "1 - treat (switched)",
    "hormon (original)",
    "1 - treat_orig (switched)",
    "1 - treat_orig (switched)",
    "original",
    "original"
  ),
  effect.threshold = c(
    "1.25 (HR)", "1.25 (HR)",
    "1.25 (HR)", "1.25 (HR)",
    "1.25 (IRR)",
    "1.667 (OR)", "30 (MD)",
    "1.5 (OR)", "30 (MD)"
  ),
  adverse_outcome = c(
    "N/A (Cox)", "N/A (Cox)",
    "N/A (Cox)", "N/A (Cox)",
    "TRUE (default)",
    "TRUE (default)", "FALSE (default)",
    "TRUE", "TRUE"
  ),
  DGM_target = c(
    "HR(Q) = 1.5", "HR(Q) = 1.5 (switched)",
    "HR(Q) = 1.5", "HR(Q) = 1.5 (switched)",
    "IRR(Q) = 2.0",
    "OR(Q) = 2.0", "MD(Q) = +40",
    "OR(Q) = 2.0", "MD(Q) = +30"
  ),
  k_inter_grid = c(
    "positive", "positive (switched treat)",
    "positive", "positive (switched treat)",
    "c(0, 2)",
    "c(0, 3)", "c(0, 120)",
    "c(0, 5)", "c(0, 100)"
  ),
  stringsAsFactors = FALSE
)

qr |>
  gt() |>
  tab_header(
    title = "Parameter Quick Reference",
    subtitle = "All confirmed-working scenarios"
  ) |>
  cols_label(
    effect.threshold = "effect.threshold",
    adverse_outcome = "adverse_outcome",
    DGM_target = "DGM Target",
    k_inter_grid = "k_inter Grid"
  ) |>
  tab_style(
    style = cell_text(size = "small"),
    locations = cells_body()
  ) |>
  tab_options(table.font.size = 12)
```


# Verification: Synthetic Tests {#sec-verify}

```{r synthetic-verify}
set.seed(42)
N <- 800
syn <- data.frame(
  id = seq_len(N), treat = rep(0:1, each = N / 2),
  z1 = as.factor(sample(0:1, N, TRUE)),
  z2 = as.factor(sample(0:1, N, TRUE)),
  z3 = as.factor(sample(0:1, N, TRUE))
)
in_Q <- syn$z1 == 1 & syn$z2 == 1

cat(sprintf("Q prevalence: %.1f%%\n\n", 100 * mean(in_Q)))
```

## Test 1: Harm Search — Adverse Binary (Direct)

```{r test-harm-adverse-v1}
# Treatment increases P(event) for Q
p_event <- ifelse(syn$treat == 1 & in_Q, 0.70,
            ifelse(syn$treat == 1, 0.35, 0.35))
syn$y_event <- rbinom(N, 1, p_event)

cat("=== Harm + Adverse Binary ===\n")
cat(sprintf("  OR(Q):  %.2f  (should be > 1)\n",
    {t <- table(syn$y_event[in_Q], syn$treat[in_Q])
     (t[2,2]*t[1,1]) / (t[1,2]*t[2,1])}))

fs1 <- suppressWarnings(forestsearch(
  df.analysis = syn, confounders.name = c("z1", "z2", "z3"),
  outcome.name = "y_event", event.name = "y_event",
  treat.name = "treat", id.name = "id",
  outcome_type = "binary", effect_measure = "OR",
  effect.threshold = 2.0, consistency.threshold = 1.5,
  pconsistency.threshold = 0.80,
  fs.splits = 200L, n.min = 40L, d0.min = 8L, d1.min = 8L,
  maxk = 2L, use_lasso = FALSE, use_grf = TRUE, is.RCT = TRUE,
  seedit = 42L, quiet = TRUE,
  parallel_args = list(plan = "sequential", workers = 1,
                       show_message = FALSE)
))

found <- !is.null(fs1) && !is.null(fs1$grp.consistency)
cat(sprintf("  Subgroup found: %s\n", found))
if (found) {
  sg <- names(fs1$grp.consistency$sg.harm$subgroup)
  cat(sprintf("  Variables: %s\n", paste(sg, collapse = ", ")))
  cat(sprintf("  Correct: %s\n", all(c("z1", "z2") %in% sg)))
}
```

## Test 2: Benefit Search — Positive Binary (Switch Treatment)

```{r test-benefit-positive}
# Combo (treat_orig = 1) helps Q improve more
p_improve <- ifelse(syn$treat == 0 & in_Q, 0.75,   # combo helps Q
              ifelse(syn$treat == 0, 0.45,           # combo baseline
               ifelse(in_Q, 0.40, 0.45)))            # ddI doesn't help Q

syn$y_improve <- rbinom(N, 1, p_improve)

# Switch treatment: ddI = 1, combo = 0
syn$treat_sw <- 1L - syn$treat

cat("\n=== Benefit + Positive Binary (switched treatment) ===\n")
cat(sprintf("  OR(Q, switched): %.2f  (should be > 1)\n",
    {t <- table(syn$y_improve[in_Q], syn$treat_sw[in_Q])
     (t[2,2]*t[1,1]) / (t[1,2]*t[2,1])}))

fs2 <- suppressWarnings(forestsearch(
  df.analysis = syn, confounders.name = c("z1", "z2", "z3"),
  outcome.name = "y_improve", event.name = "y_improve",
  treat.name = "treat_sw", id.name = "id",
  outcome_type = "binary", effect_measure = "OR",
  effect.threshold = 1.5, consistency.threshold = 1.2,
  pconsistency.threshold = 0.80,
  fs.splits = 200L, n.min = 40L, d0.min = 8L, d1.min = 8L,
  maxk = 2L, use_lasso = FALSE, use_grf = TRUE, is.RCT = TRUE,
  seedit = 42L, quiet = TRUE,
  parallel_args = list(plan = "sequential", workers = 1,
                       show_message = FALSE)
))

found2 <- !is.null(fs2) && !is.null(fs2$grp.consistency)
cat(sprintf("  Subgroup found: %s\n", found2))
if (found2) {
  sg2 <- names(fs2$grp.consistency$sg.harm$subgroup)
  cat(sprintf("  Variables: %s\n", paste(sg2, collapse = ", ")))
}
```

## Test 3: Harm Search — Positive Continuous (Negate Y)

```{r test-harm-positive-cont}
# Treatment reduces a beneficial outcome for Q
syn$y_good <- 50 + 5 * syn$treat + rnorm(N, sd = 12)
syn$y_good[in_Q & syn$treat == 1] <-
  syn$y_good[in_Q & syn$treat == 1] - 25

# Negate: higher = worse
syn$y_bad <- -syn$y_good

cat("\n=== Harm + Positive Continuous (negated) ===\n")
cat(sprintf("  Original MD(Q): %.1f (treat reduces Y for Q)\n",
    mean(syn$y_good[in_Q & syn$treat == 1]) -
    mean(syn$y_good[in_Q & syn$treat == 0])))
cat(sprintf("  Negated MD(Q):  %.1f (should be > 0)\n",
    mean(syn$y_bad[in_Q & syn$treat == 1]) -
    mean(syn$y_bad[in_Q & syn$treat == 0])))

fs3 <- suppressWarnings(forestsearch(
  df.analysis = syn, confounders.name = c("z1", "z2", "z3"),
  outcome.name = "y_bad", event.name = "y_bad",
  treat.name = "treat", id.name = "id",
  outcome_type = "continuous", effect_measure = "MD",
  effect.threshold = 10, consistency.threshold = 5,
  pconsistency.threshold = 0.80,
  fs.splits = 200L, n.min = 40L, d0.min = 8L, d1.min = 8L,
  maxk = 2L, use_lasso = TRUE, use_grf = TRUE, is.RCT = TRUE,
  seedit = 42L, quiet = TRUE,
  parallel_args = list(plan = "sequential", workers = 1,
                       show_message = FALSE)
))

found3 <- !is.null(fs3) && !is.null(fs3$grp.consistency)
cat(sprintf("  Subgroup found: %s\n", found3))
if (found3) {
  sg3 <- names(fs3$grp.consistency$sg.harm$subgroup)
  cat(sprintf("  Variables: %s\n", paste(sg3, collapse = ", ")))
}
```

## Test 4: Benefit Search — Positive Continuous (Switch Treatment)

```{r test-benefit-positive-cont}
# Combo helps Q more on a positive outcome
syn$y_cd4 <- 30 + 3 * (1 - syn$treat) + rnorm(N, sd = 15)
syn$y_cd4[in_Q & syn$treat == 0] <-
  syn$y_cd4[in_Q & syn$treat == 0] + 20

cat("\n=== Benefit + Positive Continuous (switched treatment) ===\n")

# Under switched treatment: MD(Q, switched) > 0
cat(sprintf("  MD(Q, switched): %.1f (should be > 0)\n",
    mean(syn$y_cd4[in_Q & syn$treat_sw == 1]) -
    mean(syn$y_cd4[in_Q & syn$treat_sw == 0])))

fs4 <- suppressWarnings(forestsearch(
  df.analysis = syn, confounders.name = c("z1", "z2", "z3"),
  outcome.name = "y_cd4", event.name = "y_cd4",
  treat.name = "treat_sw", id.name = "id",
  outcome_type = "continuous", effect_measure = "MD",
  effect.threshold = 10, consistency.threshold = 5,
  pconsistency.threshold = 0.80,
  fs.splits = 200L, n.min = 40L, d0.min = 8L, d1.min = 8L,
  maxk = 2L, use_lasso = TRUE, use_grf = TRUE, is.RCT = TRUE,
  seedit = 42L, quiet = TRUE,
  parallel_args = list(plan = "sequential", workers = 1,
                       show_message = FALSE)
))

found4 <- !is.null(fs4) && !is.null(fs4$grp.consistency)
cat(sprintf("  Subgroup found: %s\n", found4))
if (found4) {
  sg4 <- names(fs4$grp.consistency$sg.harm$subgroup)
  cat(sprintf("  Variables: %s\n", paste(sg4, collapse = ", ")))
}
```


## Summary of Tests

```{r test-summary}
results <- data.frame(
  Test = c(
    "1: Harm + Adverse binary",
    "2: Benefit + Positive binary (switch)",
    "3: Harm + Positive continuous (negate)",
    "4: Benefit + Positive continuous (switch)"
  ),
  Found = c(found, found2, found3, found4),
  stringsAsFactors = FALSE
)

results |>
  gt() |>
  tab_header(title = "Verification Summary") |>
  tab_style(
    style = cell_fill(color = "#c8e6c9"),
    locations = cells_body(rows = Found == TRUE)
  ) |>
  tab_style(
    style = cell_fill(color = "#ffcdd2"),
    locations = cells_body(rows = Found == FALSE)
  )
```


# DGM Setup Summary {#sec-dgm}

## The Golden Rule for DGMs

> **Positive `k_inter` + positive calibration target + positive
> `k_inter_range` grid.**

This creates a subgroup $Q$ where the effect estimate (on whatever
scale is being used) is large and positive.  ForestSearch detects it
because $\hat{\theta}(Q) \geq c_1$.

The outcome and treatment manipulations (negation, switching) happen
**before** the DGM is created — the DGM operates on the final,
oriented data.

```{r dgm-summary}
dgm_guide <- data.frame(
  Endpoint = c("Binary (OR)", "Continuous (MD)", "Count (IRR)",
               "Count + offset (IRR)"),
  outcome_var = c("y_failure or y_improve", "cd4_change or cd4_loss",
                  "event_count", "event_count"),
  k_inter = c("positive → ↑ P(Y|Q,T)", "positive → ↑ E[Y|Q,T]",
              "positive → ↑ rate(Q,T)", "positive → ↑ rate(Q,T)"),
  Target_scale = c("OR(Q) > 1", "MD(Q) > 0", "IRR(Q) > 1", "IRR(Q) > 1"),
  Grid = c("c(0, 3–5)", "c(0, 50–200)", "c(0, 2–3)", "c(0, 2–3)"),
  stringsAsFactors = FALSE
)

dgm_guide |>
  gt() |>
  tab_header(
    title = "DGM Configuration Guide",
    subtitle = "All targets and grids should be positive"
  ) |>
  cols_label(
    outcome_var = "outcome_var",
    k_inter = "k_inter Effect",
    Target_scale = "Target",
    Grid = "k_inter_range"
  )
```


# Session Info {#sec-session}

```{r session-info}
sessionInfo()
```
