---
title: "Review: Classification Metric Calculations in ForestSearch"
subtitle: "Exhaustive audit of sens, spec, PPV, NPV under harm and benefit searches"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    number-sections: true
    code-fold: true
    code-summary: "Show code"
execute:
  warning: true
  message: false
  fig-align: center
knitr:
  opts_chunk:
    comment: "#>"
---


# Purpose {#sec-purpose}

This document provides an exhaustive review of how classification
metrics (sensitivity, specificity, PPV, NPV) are calculated in the
`forestsearch` simulation pipeline, with specific attention to whether
the metrics are correctly labeled and computed under both
**harm searches** (`subgroup_notation = "harm"`) and
**benefit searches** (`subgroup_notation = "benefit"`).

The central question: in the binary benefit simulation, FS reports
`sens(Q) = 42%`.  Is this truly the sensitivity **for Q** (the
fraction of true benefit-subgroup members captured by $\hat{Q}$),
or is it accidentally the sensitivity for the complement $Q^c$?


# Code Audit: Tracing the Classification Pipeline {#sec-audit}

## Step 1: How `flag_harm` is created

The true subgroup indicator `flag_harm` is created by
`define_subgroups()` (in `generate_aft_dgm_helpers.R`, line 486):

```r
# All subgroup conditions met → flag_harm = 1
flag_harm <- as.numeric(Reduce("&", subgroup_indicators))
```

This is called by both `generate_glm_dgm()` (binary/continuous/count
DGMs) and `generate_aft_dgm_flex()` (survival DGMs).

**Critical point:** `flag_harm = 1` marks the **target subgroup**
regardless of search direction.  For a harm search, this is $H$; for
a benefit search with treatment switching, this is $Q$.  The name
`flag_harm` is inherited from the survival-only v0.1.x code but
functions as a generic "target subgroup" indicator.

Under the null model (`model = "null"`), `generate_glm_dgm()` explicitly
zeros out `flag_harm` (line 269):

```r
if (model == "null") {
  data$flag_harm <- 0L
}
```

So under $H_0$, no subject is in the target subgroup ($Q = \emptyset$).


## Step 2: How `sg_hat` is built from ForestSearch output

In `run_simulation_analysis.R` (line 624):

```r
df$sg_hat <- as.integer(fs_full$df.est$treat.recommend == 0)
```

The convention is:

- `treat.recommend == 0` → "do NOT recommend treatment" → subject is
  in the **estimated** subgroup ($\hat{H}$ or $\hat{Q}$)
- `treat.recommend == 1` → "recommend treatment" → subject is in the
  **complement** ($\hat{H}^c$ or $\hat{Q}^c$)

So `sg_hat == 1` means "in the estimated target subgroup."

For **GRF** (line 937), the identical convention is used:

```r
harm_ind <- as.integer(grf_data$treat.recommend == 0)
```


## Step 3: How TP/FP/TN/FN are computed

In `run_simulation_analysis.R` (lines 687--697), for **FS**:

```r
true_H <- df[[harm_col]] == 1L     # in TRUE target subgroup
hat_H  <- df$sg_hat == 1L          # in ESTIMATED target subgroup

tp <- sum(true_H & hat_H)          # correctly identified
fp <- sum(!true_H & hat_H)         # falsely included
tn <- sum(!true_H & !hat_H)        # correctly excluded
fn <- sum(true_H & !hat_H)         # missed
```

For **GRF** (lines 1010--1019), the identical logic is used after
merging `flag_harm` and `sg_hat` via subject IDs.


## Step 4: Metrics definitions

```r
sens = tp / (tp + fn)    # fraction of TRUE subgroup captured
spec = tn / (tn + fp)    # fraction of complement correctly excluded
ppv  = tp / (tp + fp)    # fraction of estimated subgroup that is correct
npv  = tn / (tn + fn)    # fraction of estimated complement that is correct
```


## Step 5: What happens in a benefit search?

In a benefit search, treatment roles are switched before calling
`forestsearch()`.  ForestSearch always identifies the "harm" subgroup
internally (subjects where the switched treatment appears harmful).
Under the original treatment scale, these are the subjects who
**benefit** from the original treatment.

The mapping:

| Code variable | Harm search meaning | Benefit search meaning |
|---|---|---|
| `flag_harm == 1` | Subject is in $H$ | Subject is in $Q$ |
| `treat.recommend == 0` | Subject is in $\hat{H}$ | Subject is in $\hat{Q}$ |
| `sg_hat == 1` | In $\hat{H}$ | In $\hat{Q}$ |
| `true_H` | In $H$ | In $Q$ |
| `hat_H` | In $\hat{H}$ | In $\hat{Q}$ |
| `TP` | In $H \cap \hat{H}$ | In $Q \cap \hat{Q}$ |
| `sens` | sens($H$) | sens($Q$) |
| `ppv` | ppv($H$) | ppv($Q$) |

**The classification computation is invariant to the search direction.**
The `subgroup_notation` parameter only affects display labels (H/Hc vs
Q/Qc) in `format_oc_results()` and `compare_fs_subgroups()`.


## Step 6: `compare_fs_subgroups()` alignment

`compare_fs_subgroups()` extracts subgroup membership via the same
convention:

```r
# forestsearch object (line 117-118):
sg = obj$df.est$treat.recommend     # 0 = in subgroup, 1 = complement
detected = any(obj$df.est$treat.recommend == 0L)

# GRF object (line 131-133):
sg = obj$data$treat.recommend       # same convention
```

Labels are applied via `subgroup_notation`:

- `"harm"`:  sg0 = "H (Questionable)", sg1 = "Hc (Recommend)"
- `"benefit"`: sg0 = "Q (Benefit)", sg1 = "Qc (No benefit)"

The cross-tabulation then compares two analyses' `treat.recommend`
vectors directly.  This is **aligned** with the classification metrics
in `run_simulation_analysis()` because both use the same
`treat.recommend == 0` → target subgroup convention.


# Pressure Test: Synthetic Data Verification {#sec-pressure}

```{r setup}
#| include: false
library(forestsearch)
library(data.table)
library(gt)
```


## Test 1: Perfect detection (TP = true subgroup)

```{r test-perfect}
# Synthetic data: 100 subjects, 30 in true subgroup
N <- 100L
n_Q <- 30L

df_test <- data.frame(
  id         = seq_len(N),
  flag_harm  = c(rep(1L, n_Q), rep(0L, N - n_Q)),
  sg_hat     = c(rep(1L, n_Q), rep(0L, N - n_Q))  # perfect detection
)

true_H <- df_test$flag_harm == 1L
hat_H  <- df_test$sg_hat == 1L

tp <- sum(true_H & hat_H);   fp <- sum(!true_H & hat_H)
tn <- sum(!true_H & !hat_H); fn <- sum(true_H & !hat_H)

cat("=== Test 1: Perfect detection ===\n")
cat(sprintf("  N = %d, |Q| = %d, |Q_hat| = %d\n", N, sum(true_H), sum(hat_H)))
cat(sprintf("  TP = %d, FP = %d, TN = %d, FN = %d\n", tp, fp, tn, fn))
cat(sprintf("  sens = %.2f (expect 1.00)\n", tp / (tp + fn)))
cat(sprintf("  spec = %.2f (expect 1.00)\n", tn / (tn + fp)))
cat(sprintf("  ppv  = %.2f (expect 1.00)\n", tp / (tp + fp)))
cat(sprintf("  npv  = %.2f (expect 1.00)\n", tn / (tn + fn)))

stopifnot(tp == 30, fp == 0, tn == 70, fn == 0)
stopifnot(tp / (tp + fn) == 1, tn / (tn + fp) == 1)
cat("  PASSED\n")
```


## Test 2: No detection (all predicted complement)

```{r test-no-detection}
df_test2 <- data.frame(
  id         = seq_len(N),
  flag_harm  = c(rep(1L, n_Q), rep(0L, N - n_Q)),
  sg_hat     = rep(0L, N)   # nothing detected
)

true_H <- df_test2$flag_harm == 1L
hat_H  <- df_test2$sg_hat == 1L

tp <- sum(true_H & hat_H);   fp <- sum(!true_H & hat_H)
tn <- sum(!true_H & !hat_H); fn <- sum(true_H & !hat_H)

cat("=== Test 2: No detection ===\n")
cat(sprintf("  TP = %d, FP = %d, TN = %d, FN = %d\n", tp, fp, tn, fn))
cat(sprintf("  sens = %.2f (expect 0.00)\n", tp / (tp + fn)))
cat(sprintf("  spec = %.2f (expect 1.00)\n", tn / (tn + fp)))
cat(sprintf("  ppv  = NA   (TP+FP=0, undefined)\n"))
cat(sprintf("  npv  = %.2f (expect %.2f = |Qc|/N)\n",
    tn / (tn + fn), (N - n_Q) / N))

stopifnot(tp == 0, fp == 0, tn == 70, fn == 30)
stopifnot(tp / (tp + fn) == 0)
cat("  PASSED\n")
```


## Test 3: Partial overlap (the realistic case)

```{r test-partial}
# True Q: subjects 1-30.  Estimated Q_hat: subjects 11-50 (overlap = 20)
df_test3 <- data.frame(
  id         = seq_len(N),
  flag_harm  = c(rep(1L, 30), rep(0L, 70)),
  sg_hat     = c(rep(0L, 10), rep(1L, 40), rep(0L, 50))
)

true_H <- df_test3$flag_harm == 1L
hat_H  <- df_test3$sg_hat == 1L

tp <- sum(true_H & hat_H);   fp <- sum(!true_H & hat_H)
tn <- sum(!true_H & !hat_H); fn <- sum(true_H & !hat_H)

sens <- tp / (tp + fn)
spec <- tn / (tn + fp)
ppv  <- tp / (tp + fp)
npv  <- tn / (tn + fn)

cat("=== Test 3: Partial overlap ===\n")
cat(sprintf("  True Q = {1,...,30}, Q_hat = {11,...,50}\n"))
cat(sprintf("  Overlap = {11,...,30} = 20 subjects\n"))
cat(sprintf("  TP = %d, FP = %d, TN = %d, FN = %d\n", tp, fp, tn, fn))
cat(sprintf("  sens = %.2f (= 20/30 = 0.667, fraction of Q in Q_hat)\n", sens))
cat(sprintf("  spec = %.2f (= 50/70 = 0.714, fraction of Qc excluded from Q_hat)\n", spec))
cat(sprintf("  ppv  = %.2f (= 20/40 = 0.500, fraction of Q_hat truly in Q)\n", ppv))
cat(sprintf("  npv  = %.2f (= 50/60 = 0.833, fraction of Q_hat_c truly in Qc)\n", npv))

stopifnot(tp == 20, fp == 20, tn == 50, fn == 10)
stopifnot(abs(sens - 20/30) < 1e-10)
stopifnot(abs(spec - 50/70) < 1e-10)
stopifnot(abs(ppv  - 20/40) < 1e-10)
stopifnot(abs(npv  - 50/60) < 1e-10)
cat("  PASSED\n")
```


## Test 4: Null hypothesis ($Q = \emptyset$)

```{r test-null}
df_test4 <- data.frame(
  id         = seq_len(N),
  flag_harm  = rep(0L, N),      # no true subgroup
  sg_hat     = c(rep(1L, 15), rep(0L, 85))  # false detection
)

true_H <- df_test4$flag_harm == 1L
hat_H  <- df_test4$sg_hat == 1L

tp <- sum(true_H & hat_H);   fp <- sum(!true_H & hat_H)
tn <- sum(!true_H & !hat_H); fn <- sum(true_H & !hat_H)

sens_val <- if ((tp + fn) > 0) tp / (tp + fn) else NA_real_
spec_val <- if ((tn + fp) > 0) tn / (tn + fp) else NA_real_
ppv_val  <- if ((tp + fp) > 0) tp / (tp + fp) else NA_real_
npv_val  <- if ((tn + fn) > 0) tn / (tn + fn) else NA_real_

cat("=== Test 4: Null hypothesis (Q = emptyset) ===\n")
cat(sprintf("  TP = %d, FP = %d, TN = %d, FN = %d\n", tp, fp, tn, fn))
cat(sprintf("  sens = %s  (TP+FN=0 → NA, correct)\n",
    ifelse(is.na(sens_val), "NA", sprintf("%.2f", sens_val))))
cat(sprintf("  spec = %.2f (= 85/100 = 0.85)\n", spec_val))
cat(sprintf("  ppv  = %.2f (= 0/15 = 0.00, all detections are false)\n", ppv_val))
cat(sprintf("  npv  = %.2f (= 85/85 = 1.00, everyone in Q_hat_c is truly Qc)\n", npv_val))

stopifnot(is.na(sens_val))
stopifnot(ppv_val == 0)
stopifnot(npv_val == 1)
cat("  PASSED\n")
```


## Test 5: Null with no detection (correct behavior)

```{r test-null-no-detect}
df_test5 <- data.frame(
  id         = seq_len(N),
  flag_harm  = rep(0L, N),   # no true subgroup
  sg_hat     = rep(0L, N)    # no detection either
)

true_H <- df_test5$flag_harm == 1L
hat_H  <- df_test5$sg_hat == 1L

tp <- sum(true_H & hat_H);   fp <- sum(!true_H & hat_H)
tn <- sum(!true_H & !hat_H); fn <- sum(true_H & !hat_H)

spec_val <- if ((tn + fp) > 0) tn / (tn + fp) else NA_real_
npv_val  <- if ((tn + fn) > 0) tn / (tn + fn) else NA_real_

cat("=== Test 5: Null + no detection ===\n")
cat(sprintf("  TP = %d, FP = %d, TN = %d, FN = %d\n", tp, fp, tn, fn))
cat(sprintf("  spec = %.2f (expect 1.00)\n", spec_val))
cat(sprintf("  npv  = %.2f (expect 1.00)\n", npv_val))
cat("  Matches run_simulation_analysis() lines 608-610\n")

stopifnot(spec_val == 1, npv_val == 1)
cat("  PASSED\n")
```


## Test 6: Inverted labels — what if Q and Qc were swapped?

This test verifies that swapping the definitions would produce
detectably wrong results.

```{r test-swap}
# Same partial-overlap data as Test 3
true_Q <- df_test3$flag_harm == 1L    # correct
true_Qc <- df_test3$flag_harm == 0L   # WRONG: this is the complement
hat_Q  <- df_test3$sg_hat == 1L

# Correct calculation (for Q)
tp_correct <- sum(true_Q & hat_Q)
fn_correct <- sum(true_Q & !hat_Q)
sens_correct <- tp_correct / (tp_correct + fn_correct)

# WRONG calculation (if we accidentally used Qc as the target)
tp_wrong <- sum(true_Qc & hat_Q)
fn_wrong <- sum(true_Qc & !hat_Q)
sens_wrong <- tp_wrong / (tp_wrong + fn_wrong)

cat("=== Test 6: What if we swapped Q and Qc? ===\n")
cat(sprintf("  Correct: sens(Q)  = %d/%d = %.3f\n",
    tp_correct, tp_correct + fn_correct, sens_correct))
cat(sprintf("  WRONG:   sens(Qc) = %d/%d = %.3f\n",
    tp_wrong, tp_wrong + fn_wrong, sens_wrong))
cat(sprintf("  These differ by %.3f — a swap would be detectable.\n",
    abs(sens_correct - sens_wrong)))

stopifnot(abs(sens_correct - 20/30) < 1e-10)
stopifnot(abs(sens_wrong - 20/70) < 1e-10)
cat("  PASSED: sens(Q) = 0.667 ≠ sens(Qc) = 0.286\n")
```


## Test 7: Verify `compare_fs_subgroups()` alignment

```{r test-compare}
# Create two mock indicator vectors (0 = in subgroup, 1 = complement)
# Matching the treat.recommend convention
mock_fs1 <- c(rep(0L, 40), rep(1L, 60))   # Q_hat = subjects 1-40
mock_fs2 <- c(rep(0L, 25), rep(1L, 75))   # Q_hat = subjects 1-25

result <- compare_fs_subgroups(
  mock_fs1, mock_fs2,
  label1 = "Method A", label2 = "Method B",
  subgroup_notation = "benefit"
)

cat("=== Test 7: compare_fs_subgroups() alignment ===\n")
cat("Cross-tabulation:\n")
print(result$crosstab)

cat(sprintf("\nAgreement: %.1f%%\n", 100 * result$concordance$agreement))

# Verify: both in Q → 25, both in Qc → 60, A=Q B=Qc → 15, A=Qc B=Q → 0
ct <- as.matrix(result$crosstab)
stopifnot(ct[1,1] == 25)   # both in Q
stopifnot(ct[2,2] == 60)   # both in Qc
stopifnot(ct[1,2] == 15)   # A=Q, B=Qc
stopifnot(ct[2,1] == 0)    # A=Qc, B=Q
cat("  Cross-tab values verified.\n")

# Key check: the "sg0" position (row 1, col 1) corresponds to
# treat.recommend == 0 for both methods.  With subgroup_notation = "benefit",
# this is labeled "Q (Benefit)".  This matches the convention in
# run_simulation_analysis() where sg_hat = (treat.recommend == 0).
cat("  Labels aligned with classification metric convention.\n")
cat("  PASSED\n")
```


## Test 8: Harm search — verify same logic with H/Hc labels

```{r test-harm}
result_harm <- compare_fs_subgroups(
  mock_fs1, mock_fs2,
  label1 = "Method A", label2 = "Method B",
  subgroup_notation = "harm"
)

cat("=== Test 8: Harm search labels ===\n")
cat("Cross-tabulation:\n")
print(result_harm$crosstab)

# Same numerical values, different labels
ct_harm <- as.matrix(result_harm$crosstab)
ct_benefit <- as.matrix(result$crosstab)
stopifnot(all(ct_harm == ct_benefit))
cat("  Numerical values identical to benefit search.\n")
cat("  Only labels differ: H/Hc vs Q/Qc.\n")
cat("  PASSED\n")
```


# Summary of Findings {#sec-findings}

## Answer to the central question

**`sens(Q) = 42%` is correctly computed and correctly labeled.**

The 42% is the fraction of true benefit-subgroup members $Q$ that
are captured by the estimated $\hat{Q}$.  This has been verified by
tracing through every step of the pipeline:

1. **`flag_harm`** is created by `define_subgroups()` and marks the
   **target subgroup** — $Q$ in a benefit search, $H$ in a harm
   search.

2. **`treat.recommend == 0`** in ForestSearch output marks subjects
   in the **estimated** target subgroup ($\hat{Q}$ or $\hat{H}$).

3. **`sg_hat = as.integer(treat.recommend == 0)`** converts this to a
   binary indicator where 1 = in estimated subgroup.

4. **`true_H = (flag_harm == 1)`** and **`hat_H = (sg_hat == 1)`**
   are then compared to produce TP/FP/TN/FN.

5. **`sens = TP/(TP+FN)`** is the fraction of the true target subgroup
   captured — this is sens($Q$) for a benefit search.

The computation is **invariant** to the search direction.  The
`subgroup_notation` parameter only affects display labels in
`format_oc_results()`, `build_estimation_table()`, and
`compare_fs_subgroups()`.


## `compare_fs_subgroups()` alignment

The cross-tabulation function uses the same `treat.recommend == 0` →
target subgroup convention.  With `subgroup_notation = "benefit"`,
the sg0 position is labeled "Q (Benefit)" — matching the classification
metric convention.  Verified with synthetic data in Test 7.


## Edge cases verified

| Scenario | sens | spec | ppv | npv | Source |
|---|---|---|---|---|---|
| Perfect detection | 1.00 | 1.00 | 1.00 | 1.00 | Test 1 |
| No detection (H1) | 0.00 | 1.00 | NA | $|Q^c|/N$ | Test 2, lines 600--604 |
| Partial overlap | 0.667 | 0.714 | 0.500 | 0.833 | Test 3 |
| Null + false detect | NA | 0.85 | 0.00 | 1.00 | Test 4 |
| Null + no detect | NA | 1.00 | NA | 1.00 | Test 5, lines 608--610 |
| Swapped Q/Qc | 0.286 ≠ 0.667 | — | — | — | Test 6 |


## Naming convention note

The name `flag_harm` is a legacy from the survival-only v0.1.x
codebase where the target subgroup was always a harm subgroup.
In the GLM extension (v0.2.0), `flag_harm` serves as a generic
"target subgroup indicator" — it marks $Q$ in benefit searches
and $H$ in harm searches.  Renaming to `flag_subgroup` or
`flag_target` would reduce confusion but is a Tier 3 cosmetic
change that should be deferred to avoid breaking downstream
code that references `flag_harm` by name.


# Session Info {#sec-session}

```{r session}
sessionInfo()
```
