---
title: "Threshold Selection Guide for ForestSearch"
subtitle: "Using `suggest_thresholds()` to Choose (c1, c2) for Non-Survival Outcomes"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---


# Background {#sec-background}

## The Problem

The paper defaults for ForestSearch (c1 = 1.25, c2 = 1.0)
were calibrated for **survival endpoints** using the GBSG DGM
(Leon et al., 2024).  For non-survival outcomes (binary,
count, continuous), these defaults may not provide adequate
FPR control because:

1. **Lower $d_{\text{eff}}$**: Binary outcomes with
   $\bar{p} = 0.30$ give $d_{\text{eff}} = 0.21 \cdot n_H$,
   roughly one-third the survival value
   ($d = 0.54 \cdot n_H$ at 46% censoring).  Lower
   $d_{\text{eff}}$ means higher per-subgroup $P_1$.

2. **Higher $L_{\text{eff}}$**: Without an events-based
   minimum (`d0.min`, `d1.min`), more candidate subgroups
   survive to the threshold check, so the procedure-level
   FPR is amplified relative to the per-subgroup $P_1$.

## The Solution

`suggest_thresholds()` uses the **corrected approximation**
to find the smallest $(c_1, c_2)$ that controls the
procedure-level FPR:

$$
\text{FPR}_{\text{corr}} = 1 - (1 - P_1)^{L_{\text{eff}}(N)}
\le \text{target}
$$

where:

- $P_1 = P(W_1 + W_2 \ge 2 \log c_1,\;
  \min(W_1, W_2) \ge \log c_2)$
  with $W_1, W_2 \sim N(0, 8/d_{\text{eff}})$ under $H_0$
- $L_{\text{eff}}(N) = C \cdot (N / n_{\min})^\alpha$ is
  the effective number of candidates

The function returns a **frontier table**: for each $c_1$
(ascending), the minimum $c_2$ that achieves the target.
The first row is the answer to "what is the smallest $c_1$?"

::: {.callout-note}
The frontier values are based on the corrected approximation
and do not involve simulation.  For simulation confirmation
with 95% confidence intervals (treating the corrected FPR as
the true rate), see:

- `selection_binary_frontier.qmd` (S1) for binary outcomes
- `selection_survival_frontier.qmd` (S2) for survival outcomes

Set `nsims_val > 0` in those documents to run ForestSearch
under $H_0$ and verify that the simulated FPR falls within
the expected range.
:::


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(ggplot2)
library(gt)

# Load forestsearch from source (picks up suggest_thresholds.R
# and other new files without needing devtools::install())
devtools::load_all(here::here(), quiet = TRUE)
```


# Installation {#sec-install}

The `suggest_thresholds()` function lives in
`R/suggest_thresholds.R`.  After placing it in the package
`R/` directory:

```{r install-note}
#| eval: false
#| code-fold: false

# From your package root:
devtools::document()   # generates man/suggest_thresholds.Rd
devtools::load_all()   # makes the function available
?suggest_thresholds    # view the help page
```


# Example 1: Survival (Paper Defaults) {#sec-survival}

For survival, the paper defaults should emerge naturally when
$L_{\text{eff}} = 1$ (the default).

```{r survival-example}
#| code-fold: false

# Subgroup: n_H = 91 (pH=13%, N=700), 46% censoring
d_surv <- d_eff_survival(n_sg = 91, prop_cens = 0.46)
cat(sprintf("d_eff (survival): %.0f events\n", d_surv))

# Without L_eff correction (default: C=1, alpha=0 -> L_eff=1)
frontier_surv <- suggest_thresholds(
  d_eff = d_surv, N = 700,
  fpr_target = 0.12
)
```

```{r survival-table}
#| tbl-cap: "Survival frontier: smallest c1 values achieving P1 <= 12% (L_eff = 1, so FPR_corr = P1)"

gt(head(frontier_surv[, c("c1", "min_c2", "P1")], 10)) |>
  tab_header(
    title = "Survival: Threshold Frontier (L_eff = 1)",
    subtitle = sprintf("d_eff = %d, N = 700", round(d_surv))
  ) |>
  fmt_number(columns = P1, decimals = 3) |>
  tab_source_note(
    "With L_eff = 1: corrected FPR = P1 (no multiplicity adjustment)."
  )
```

With $L_{\text{eff}} = 1$, the per-subgroup approximation
gives $P_1(1.25, 1.0) \approx 17\%$.  The paper found
simulated FPR of ~5.6% at N = 700 because the true
$L_{\text{eff}} < 1$ (the procedure is *more* conservative
than a single test due to the events-based filters).
The paper defaults are safe for survival.


# Example 2: Binary Outcome {#sec-binary}

## Without $L_{\text{eff}}$ Correction

```{r binary-no-leff}
#| code-fold: false

# Subgroup: 30% prevalence, N=700 -> n_H = 210, p = 0.30
d_bin <- d_eff_binary(n_sg = 210, p_event = 0.30)
cat(sprintf("d_eff (binary): %.1f\n", d_bin))

# Without correction: gives the per-subgroup P1 frontier
frontier_bin_raw <- suggest_thresholds(
  d_eff = d_bin, N = 700,
  fpr_target = 0.12
)
```

```{r binary-no-leff-table}
#| tbl-cap: "Binary frontier WITHOUT L_eff correction (P1 only)"

gt(head(frontier_bin_raw[, c("c1", "min_c2", "P1")], 8)) |>
  tab_header(
    title = "Binary: Per-Subgroup Frontier (L_eff = 1)",
    subtitle = sprintf("d_eff = %.1f, N = 700", d_bin)
  ) |>
  fmt_number(columns = P1, decimals = 3) |>
  tab_source_note(
    "With L_eff = 1: corrected FPR = P1. This UNDERESTIMATES the true procedure-level FPR."
  )
```

Without the $L_{\text{eff}}$ correction, the frontier suggests
lower thresholds than are actually needed.

## With $L_{\text{eff}}$ Correction

The $L_{\text{eff}}$ calibration from
`calibration_binary_leff.qmd` (5000 sims, 4 confounders):
$C = 0.220$, $\alpha = 1.298$.

```{r binary-with-leff}
#| code-fold: false

# With L_eff correction
frontier_bin <- suggest_thresholds(
  d_eff = d_bin, N = 700,
  fpr_target = 0.12,
  L_eff_C = 0.220,
  L_eff_alpha = 1.298
)
```

```{r binary-leff-table}
#| tbl-cap: "Binary frontier WITH L_eff correction.  Note: P1 is 2-3% but the corrected FPR is 10-12% due to the L_eff multiplier."

gt(head(frontier_bin, 10)) |>
  tab_header(
    title = "Binary: Corrected Frontier (L_eff calibrated)",
    subtitle = sprintf("d_eff = %.1f, N = 700, L_eff = %.1f",
      d_bin,
      0.220 * (700 / 60)^1.298)
  ) |>
  fmt_number(columns = c(FPR_corr, P1), decimals = 3) |>
  fmt_number(columns = L_eff, decimals = 1) |>
  tab_source_note(
    "P1 = per-subgroup probability. FPR_corr = 1-(1-P1)^L_eff = procedure-level FPR."
  )
```

The corrected frontier shows that c1 = 1.55--1.60 with
c2 = 1.50--1.55 is needed to achieve 12% FPR control.


## How It Changes With Sample Size

```{r binary-by-N, fig.height=6}
#| fig-cap: "Binary frontier across sample sizes.  Larger N requires higher thresholds because L_eff grows."
#| code-fold: false

N_values <- c(300, 500, 700, 1000)
p_harm <- 0.30
p_event <- 0.30

frontier_list <- lapply(N_values, function(N_val) {
  n_H <- round(p_harm * N_val)
  d <- d_eff_binary(n_H, p_event)
  ft <- suggest_thresholds(
    d_eff = d, N = N_val,
    fpr_target = 0.12,
    L_eff_C = 0.220, L_eff_alpha = 1.298
  )
  if (nrow(ft) > 0) {
    ft$N <- N_val
    ft$n_H <- n_H
    ft$d_eff_val <- d
  }
  ft
})

frontier_all <- do.call(rbind, frontier_list)

ggplot(frontier_all, aes(x = c1, y = min_c2,
                          colour = factor(N))) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression(c[1] ~ "(screening threshold)"),
       y = expression("Min " * c[2] ~ "(consistency)"),
       title = "Binary Outcome: Threshold Frontier by N",
       subtitle = "Smallest (c1, c2) achieving corrected FPR <= 12%",
       colour = "N") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


# Example 3: Continuous Outcome {#sec-continuous}

For continuous endpoints (e.g., change from baseline in blood
pressure, biomarker level), the effect is a **mean difference**
and thresholds are on the difference scale.

$$
d_{\text{eff}} = n_H / \sigma^2
$$

where $\sigma$ is the residual SD of the outcome.  Working in
**standardized units** ($\sigma = 1$) is recommended so that
$d_{\text{eff}} = n_H$ and the thresholds are on the Cohen's
$d$ scale.

```{r continuous-example}
#| code-fold: false

# Standardized continuous outcome (sigma = 1)
# Subgroup: 25% prevalence, N = 600 -> n_H = 150
n_H_cont <- 150
sigma_y <- 1.0
d_cont <- d_eff_continuous(n_sg = n_H_cont, sigma_y = sigma_y)
cat(sprintf("d_eff (continuous, sigma=%.1f): %.0f\n", sigma_y, d_cont))

frontier_cont <- suggest_thresholds(
  d_eff = d_cont, N = 600,
  fpr_target = 0.12,
  effect_scale = "difference",
  c1_range = c(0.01, 1.0),
  c1_step = 0.01,
  c2_candidates = seq(0.0, 0.50, by = 0.01)
)
```

```{r continuous-table}
#| tbl-cap: "Continuous frontier (standardized, sigma=1): thresholds are on the Cohen's d scale"

if (nrow(frontier_cont) > 0) {
  gt(head(frontier_cont[, c("c1", "min_c2", "P1")], 10)) |>
    tab_header(
      title = "Continuous: Threshold Frontier (L_eff = 1)",
      subtitle = sprintf("d_eff = %.0f (n_H=%d, sigma=%.1f), N = 600",
                          d_cont, n_H_cont, sigma_y)
    ) |>
    fmt_number(columns = P1, decimals = 3) |>
    tab_source_note(
      "Thresholds on the mean-difference scale.  With sigma=1, these are standardized effect sizes (Cohen's d)."
    )
} else {
  cat("No pairs found.  d_eff =", d_cont,
      "is large; try smaller c1_range.\n")
}
```

::: {.callout-note}
## Continuous: Threshold Interpretation

For continuous outcomes, $c_1$ and $c_2$ are on the
**mean difference** scale:

- **Standardized** ($\sigma = 1$): $c_1 = 0.05$ means
  "the half-sample mean difference must exceed 0.05 SD
  units."  These are small Cohen's $d$ values.

- **Unstandardized** ($\sigma = 15$, e.g., mmHg): scale
  thresholds accordingly.
  $d_{\text{eff}} = n_H / \sigma^2 = 150/225 = 0.67$.
  Thresholds would be in mmHg.

Because $d_{\text{eff}}$ for continuous outcomes can be very
large (when $\sigma$ is small relative to $\sqrt{n_H}$), the
per-subgroup $P_1$ may be low even at small thresholds.
:::


## Effect of Residual SD

```{r continuous-sigma, fig.height=5}
#| fig-cap: "Continuous outcome: how residual SD affects the threshold frontier.  Lower sigma = higher d_eff = lower thresholds needed."

sigma_values <- c(0.5, 1.0, 2.0, 5.0)
N_cont <- 600
n_H_cont <- 150

cont_list <- lapply(sigma_values, function(sig) {
  d <- d_eff_continuous(n_H_cont, sig)
  ft <- suggest_thresholds(
    d_eff = d, N = N_cont,
    fpr_target = 0.12,
    effect_scale = "difference",
    c1_range = c(0.01, 2.0), c1_step = 0.02,
    c2_candidates = seq(0.0, 1.5, by = 0.02)
  )
  if (nrow(ft) > 0) {
    ft$sigma <- sig
    ft$d_eff_val <- d
    ft$outcome <- sprintf("sigma=%.1f (d_eff=%.0f)", sig, d)
  }
  ft
})
cont_all <- do.call(rbind, cont_list[sapply(cont_list, nrow) > 0])

if (nrow(cont_all) > 0) {
  ggplot(cont_all, aes(x = c1, y = min_c2,
                        colour = outcome)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    scale_colour_brewer(palette = "Set1") +
    labs(x = expression(c[1] ~ "(mean difference threshold)"),
         y = expression("Min " * c[2] ~ "(consistency)"),
         title = "Continuous: Frontier by Residual SD",
         subtitle = sprintf("n(H)=%d, N=%d, FPR target=12%%",
                             n_H_cont, N_cont),
         colour = NULL) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "right",
          plot.title = element_text(face = "bold"))
}
```


# Example 4: Count Data {#sec-count}

For general count outcomes (e.g., number of adverse events,
exacerbations, hospital visits), the effect is an
**incidence rate ratio** (IRR) on the log scale.

$$
d_{\text{eff}} = D \quad \text{(total count across both arms
in the subgroup)}
$$

This uses `d_eff_count(total_events = D)` (not `d_eff_survival`).  When the
mean count per person is $\bar\mu$ and the subgroup has
$n_H$ subjects, $D \approx n_H \cdot \bar\mu$.

```{r count-example}
#| code-fold: false

# Count outcome: mean 3 events per person, n_H = 120
mu_bar <- 3.0
n_H_count <- 120
D_total <- round(n_H_count * mu_bar)
d_count <- d_eff_count(total_events = D_total)
cat(sprintf("d_eff (count): %.0f (D=%d, mu=%.1f, n_H=%d)\n",
            d_count, D_total, mu_bar, n_H_count))

frontier_count <- suggest_thresholds(
  d_eff = d_count, N = 600,
  fpr_target = 0.12,
  effect_scale = "ratio"
)
```

```{r count-table}
#| tbl-cap: "Count data frontier: thresholds on the IRR scale"

if (nrow(frontier_count) > 0) {
  gt(head(frontier_count[, c("c1", "min_c2", "P1")], 10)) |>
    tab_header(
      title = "Count Data: Threshold Frontier (L_eff = 1)",
      subtitle = sprintf(
        "d_eff = %d (D=%d events, mu=%.1f/person), N = 600",
        round(d_count), D_total, mu_bar)
    ) |>
    fmt_number(columns = P1, decimals = 3) |>
    tab_source_note(
      "Thresholds on the IRR scale (same interpretation as HR/OR)."
    )
} else {
  cat("No pairs found.\n")
}
```

::: {.callout-note}
## Count Data: d_eff Depends on Event Rate

The $d_{\text{eff}}$ for counts scales with the **total
event count** $D = n_H \cdot \bar\mu$:

- **High-rate** ($\bar\mu = 5$, $n_H = 100$):
  $d_{\text{eff}} = 500$.  Very high information; paper
  defaults likely adequate.

- **Low-rate** ($\bar\mu = 0.3$, $n_H = 100$):
  $d_{\text{eff}} = 30$.  Similar to survival; paper
  defaults may be adequate.

- **Very low-rate** ($\bar\mu = 0.1$, $n_H = 60$):
  $d_{\text{eff}} = 6$.  Lower than binary; higher
  thresholds needed.

The count setting with moderate-to-high event rates is the
most favorable for FPR control.
:::


## Effect of Event Rate

```{r count-rates, fig.height=5}
#| fig-cap: "Count data: how the mean event rate affects the threshold frontier.  Higher rates = more information = lower thresholds."

mu_values <- c(0.3, 1.0, 3.0, 10.0)
n_H_count <- 120
N_count <- 600

count_list <- lapply(mu_values, function(mu) {
  D <- round(n_H_count * mu)
  d <- d_eff_count(total_events = D)
  ft <- suggest_thresholds(
    d_eff = d, N = N_count,
    fpr_target = 0.12,
    effect_scale = "ratio"
  )
  if (nrow(ft) > 0) {
    ft$mu <- mu
    ft$d_eff_val <- d
    ft$outcome <- sprintf("mu=%.1f (d_eff=%d)", mu, round(d))
  }
  ft
})
count_all <- do.call(rbind, count_list[sapply(count_list, nrow) > 0])

if (nrow(count_all) > 0) {
  ggplot(count_all, aes(x = c1, y = min_c2,
                         colour = outcome)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    geom_point(data = data.frame(c1 = 1.25, min_c2 = 1.0),
               colour = "black", size = 4, shape = 8) +
    annotate("text", x = 1.28, y = 1.03,
             label = "Paper\ndefaults", size = 3, hjust = 0) +
    scale_colour_brewer(palette = "Dark2") +
    labs(x = expression(c[1] ~ "(IRR screening threshold)"),
         y = expression("Min " * c[2] ~ "(consistency)"),
         title = "Count Data: Frontier by Mean Event Rate",
         subtitle = sprintf("n(H)=%d, N=%d, FPR target=12%%",
                             n_H_count, N_count),
         colour = NULL) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "right",
          plot.title = element_text(face = "bold"))
}
```


# Example 5: Using the Result in ForestSearch {#sec-usage}

## Binary

```{r usage-binary}
#| eval: false
#| code-fold: false

d <- d_eff_binary(n_sg = 210, p_event = 0.30)
frontier <- suggest_thresholds(d, N = 700, fpr_target = 0.12,
  L_eff_C = 0.220, L_eff_alpha = 1.298)

fs <- forestsearch(
  df.analysis = my_data,
  confounders.name = c("bm1", "bm2", "age", "ecog"),
  outcome.name = "response", event.name = "response",
  treat.name = "treat", id.name = "id",
  outcome_type = "binary", effect_measure = "OR",
  adverse_outcome = TRUE,
  hr.threshold = frontier$c1[1],       # smallest c1
  hr.consistency = frontier$min_c2[1],
  pconsistency.threshold = 0.90,
  n.min = 60, maxk = 2, fs.splits = 400,
  use_lasso = TRUE, use_grf = TRUE, use_twostage = TRUE,
  is.RCT = TRUE, plot.sg = TRUE
)
```


## Continuous

```{r usage-continuous}
#| eval: false
#| code-fold: false

# Standardized outcome (or divide by SD first)
d <- d_eff_continuous(n_sg = 150, sigma_y = 1.0)
frontier <- suggest_thresholds(d, N = 600, fpr_target = 0.12,
  effect_scale = "difference",
  c1_range = c(0.01, 1.0), c1_step = 0.01,
  c2_candidates = seq(0.0, 0.50, by = 0.01))

fs <- forestsearch(
  df.analysis = my_data,
  confounders.name = c("x1", "x2", "x3"),
  outcome.name = "y_change", event.name = "y_change",
  treat.name = "treat", id.name = "id",
  outcome_type = "continuous", effect_measure = "MD",
  adverse_outcome = TRUE,
  hr.threshold = frontier$c1[1],
  hr.consistency = frontier$min_c2[1],
  pconsistency.threshold = 0.90,
  n.min = 60, maxk = 2, fs.splits = 400,
  use_lasso = TRUE, use_grf = TRUE, use_twostage = TRUE,
  is.RCT = TRUE, plot.sg = TRUE
)
```

::: {.callout-important}
## Continuous: Threshold Units

For `effect_measure = "MD"`, `hr.threshold` and
`hr.consistency` are on the **mean difference** scale
(not the HR/OR ratio scale).  ForestSearch automatically
sets the default to 0.0 when it detects an identity-link
measure.  The values from `suggest_thresholds()` with
`effect_scale = "difference"` are ready to use directly.
:::


## Count Data

```{r usage-count}
#| eval: false
#| code-fold: false

D <- round(120 * 3.0)  # n_H * mean_count
d <- d_eff_count(total_events = D)
frontier <- suggest_thresholds(d, N = 600, fpr_target = 0.12,
  effect_scale = "ratio")

fs <- forestsearch(
  df.analysis = my_data,
  confounders.name = c("x1", "x2", "x3"),
  outcome.name = "n_events", event.name = "n_events",
  treat.name = "treat", id.name = "id",
  outcome_type = "count", effect_measure = "IRR",
  adverse_outcome = TRUE,
  hr.threshold = frontier$c1[1],
  hr.consistency = frontier$min_c2[1],
  pconsistency.threshold = 0.90,
  n.min = 60, maxk = 2, fs.splits = 400,
  use_lasso = TRUE, use_grf = TRUE, use_twostage = TRUE,
  is.RCT = TRUE, plot.sg = TRUE
)
```


# Example 6: Cross-Outcome Comparison {#sec-compare}

```{r compare-outcomes, fig.height=7}
#| fig-cap: "Threshold frontiers by outcome type.  All on the ratio scale except continuous (difference scale, right panel)."

N <- 700
fpr_target <- 0.12

# Survival: n_H = 91 (pH=13%), 46% censoring
d_surv <- d_eff_survival(91, 0.46)
ft_surv <- suggest_thresholds(d_surv, N, fpr_target)
ft_surv$outcome <- sprintf("Survival (d_eff=%d)", round(d_surv))
ft_surv$scale <- "Ratio scale"

# Binary (with L_eff)
d_bin <- d_eff_binary(210, 0.30)
ft_bin <- suggest_thresholds(d_bin, N, fpr_target,
  L_eff_C = 0.220, L_eff_alpha = 1.298)
ft_bin$outcome <- sprintf("Binary (d_eff=%.0f, L_eff=%.1f)",
  d_bin, 0.220 * (N / 60)^1.298)
ft_bin$scale <- "Ratio scale"

# Count: low rate (mu=0.5, similar to survival)
d_cnt_lo <- d_eff_count(total_events = round(100 * 0.5))
ft_cnt_lo <- suggest_thresholds(d_cnt_lo, N, fpr_target)
ft_cnt_lo$outcome <- sprintf("Count mu=0.5 (d_eff=%d)",
  round(d_cnt_lo))
ft_cnt_lo$scale <- "Ratio scale"

# Count: high rate (mu=5)
d_cnt_hi <- d_eff_count(total_events = round(100 * 5))
ft_cnt_hi <- suggest_thresholds(d_cnt_hi, N, fpr_target)
ft_cnt_hi$outcome <- sprintf("Count mu=5 (d_eff=%d)",
  round(d_cnt_hi))
ft_cnt_hi$scale <- "Ratio scale"

# Continuous (standardized)
d_cont <- d_eff_continuous(150, 1.0)
ft_cont <- suggest_thresholds(d_cont, N, fpr_target,
  effect_scale = "difference",
  c1_range = c(0.01, 1.0), c1_step = 0.01,
  c2_candidates = seq(0.0, 0.50, by = 0.01))
ft_cont$outcome <- sprintf("Continuous (d_eff=%d, sigma=1)",
  round(d_cont))
ft_cont$scale <- "Difference scale"

# Ratio-scale outcomes
ratio_df <- rbind(ft_surv, ft_bin, ft_cnt_lo, ft_cnt_hi)

p_ratio <- ggplot(ratio_df, aes(x = c1, y = min_c2,
                                 colour = outcome)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_point(data = data.frame(c1 = 1.25, min_c2 = 1.0),
             colour = "black", size = 4, shape = 8) +
  annotate("text", x = 1.28, y = 1.03,
           label = "Paper\ndefaults", size = 3, hjust = 0) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression(c[1]),
       y = expression("Min " * c[2]),
       title = "Ratio-Scale Outcomes (HR / OR / IRR)",
       colour = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        plot.title = element_text(face = "bold"))

# Difference-scale outcome
p_diff <- ggplot(ft_cont, aes(x = c1, y = min_c2)) +
  geom_line(linewidth = 0.8, colour = "purple") +
  geom_point(size = 2, colour = "purple") +
  labs(x = expression(c[1] ~ "(mean difference)"),
       y = expression("Min " * c[2]),
       title = "Difference-Scale: Continuous (sigma=1)",
       subtitle = sprintf("d_eff=%d, N=%d", round(d_cont), N)) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

p_ratio + p_diff +
  plot_layout(widths = c(2, 1)) +
  plot_annotation(
    title = "Cross-Outcome Threshold Frontier Comparison",
    subtitle = sprintf("N=%d, FPR target=%.0f%%, L_eff=1 except where noted",
                        N, 100 * fpr_target))
```


# The $L_{\text{eff}}$ Calibration Step {#sec-leff}

## Why It Matters

Without the $L_{\text{eff}}$ correction, `suggest_thresholds()`
returns the **per-subgroup** frontier, which underestimates
the procedure-level FPR.  The underestimation grows with $N$
because more candidates are evaluated.

| DGM | Confounders | $L_{\text{eff}}$ at N=700 | Source |
|---|---|---|---|
| GBSG survival | 7 | ~0.3 | calibration_survival_fourway.qmd |
| Binary (4 conf) | 4 | ~5.3 | calibration_binary_leff.qmd |
| Count | -- | Not yet calibrated | Use binary as starting point |
| Continuous | -- | Not yet calibrated | Use binary as starting point |

## How to Calibrate

See `selection_binary_frontier.qmd` for the
full blueprint.  The short version:

```{r calibrate-example}
#| eval: false
#| code-fold: false

# --- Quick L_eff calibration (100 sims x 5 N values) ---
# 1. Define your H0 data generator
sim_null <- function(n, seed) { ... }

# 2. Run ForestSearch under H0 at multiple N values with
#    a reference threshold (e.g., c1=1.25, c2=1.0).
#    Record the simulated FPR at each N.

# 3. Compute P1 at each N
P1_vec <- sapply(N_values, function(N) {
  n_H <- round(p_harm * N)
  d <- d_eff_binary(n_H, p_event)
  compute_detection_probability_glm(1.0, d, 1.25, 1.0,
    effect_scale = "ratio")
})

# 4. Fit L_eff
cal <- calibrate_L_eff(
  N = N_values, P1 = P1_vec,
  sim_fpr = sim_fpr_vec, n_min = 60
)
# cal$C, cal$alpha

# 5. Use in suggest_thresholds()
suggest_thresholds(d_eff = d, N = 700,
  L_eff_C = cal$C, L_eff_alpha = cal$alpha,
  fpr_target = 0.12)
```


## When You Can Skip It

If you have no calibration data, calling
`suggest_thresholds()` with the defaults ($C = 1$,
$\alpha = 0$) gives the per-subgroup frontier.  This is
**conservative for survival** (where $L_{\text{eff}} < 1$)
and **optimistic for binary** (where $L_{\text{eff}} > 1$).

For a quick rule of thumb without calibration:

- **Survival**: paper defaults (c1 = 1.25, c2 = 1.0) are safe
- **Binary**: start with c1 = 1.60, c2 = 1.50 and validate


# Quick Reference {#sec-reference}

## Function Signature

```{r signature}
#| eval: false
#| code-fold: false

suggest_thresholds(
  d_eff,                              # from d_eff_*() helpers
  N,                                  # total sample size
  fpr_target = 0.10,                  # max procedure-level FPR
  n_min = 60,                         # min subgroup size
  L_eff_C = 1.0,                      # calibration constant
  L_eff_alpha = 0.0,                  # calibration exponent
  c1_range = c(1.05, 2.50),           # search range
  c1_step = 0.05,                     # grid resolution
  c2_candidates = seq(0.90, 2.0, 0.05),
  effect_scale = "ratio"              # or "difference"
)
```

## $d_{\text{eff}}$ by Outcome Type

```{r deff-table}
#| tbl-cap: "d_eff helpers and typical values"

deff_ref <- data.frame(
  `Outcome Type` = c("Survival (Cox PH)", "Binary (logit)",
                      "Count (Poisson + offset)", "Continuous"),
  `d_eff` = c("d_eff_survival(n_H, prop_cens)",
               "d_eff_binary(n_H, p_event)",
               "d_eff_count(total_events = D)",
               "d_eff_continuous(n_H, sigma_y)"),
  `Typical range` = c("33-65 (n_H=60-120, 46% cens)",
                       "13-44 (n_H=60-210, p=0.30)",
                       "Same as survival",
                       "Depends on sigma"),
  check.names = FALSE
)

gt(deff_ref) |>
  tab_header(title = "d_eff Helpers by Outcome Type")
```


## Reference $L_{\text{eff}}$ Calibrations

```{r leff-table}
#| tbl-cap: "Published L_eff calibrations"

leff_ref <- data.frame(
  DGM = c("GBSG survival", "Binary (4 confounders)"),
  C = c(0.029, 0.220),
  alpha = c(0.882, 1.298),
  `L_eff at N=700` = c(
    round(0.029 * (700/60)^0.882, 1),
    round(0.220 * (700/60)^1.298, 1)),
  Source = c("calibration_survival_fourway.qmd",
             "calibration_binary_leff.qmd"),
  check.names = FALSE
)

gt(leff_ref) |>
  tab_header(title = "Reference L_eff Calibrations")
```


## Recommended Thresholds Summary

```{r recommendations}
#| tbl-cap: "Recommended starting thresholds by outcome type"

rec <- data.frame(
  `Outcome Type` = c("Survival", "Binary (p~0.30)",
                      "Count (high rate, mu>=3)",
                      "Count (low rate, mu<1)",
                      "Continuous (sigma=1)"),
  `c1` = c("1.25", "1.55-1.60", "1.10-1.25",
           "1.25-1.40", "0.03-0.10"),
  `c2` = c("1.00", "1.50-1.55", "0.90-1.00",
           "1.00-1.15", "0.01-0.05"),
  `Scale` = c("HR", "OR", "IRR", "IRR", "Mean diff"),
  `FPR at N=700` = c("~6% (GBSG)", "~10-12% (4 conf)",
                      "Low (high d_eff)",
                      "~5-12%",
                      "Low (high d_eff)"),
  Source = c("Paper (Table 1)", "Calibration study",
             "Approximation", "Approximation",
             "Approximation"),
  check.names = FALSE
)

gt(rec) |>
  tab_header(
    title = "Recommended Starting Thresholds",
    subtitle = "Validate with suggest_thresholds() for your specific configuration"
  ) |>
  tab_source_note(
    "Paper defaults (c1=1.25, c2=1.0) remain the forestsearch() signature defaults."
  )
```

::: {.callout-important}
## Simulation Confirmation

The thresholds above are derived from the corrected
approximation $\text{FPR}_{\text{corr}} = 1 - (1 - P_1)^{L_{\text{eff}}}$.
Before using them in a production analysis, confirm via
simulation in the outcome-specific frontier documents:

- **Binary**: `selection_binary_frontier.qmd` — set `nsims_val = 500`
- **Survival**: `selection_survival_frontier.qmd` — set `nsims_val = 500`

The validation tables in those documents report 95% confidence
intervals treating the corrected FPR as the true rate:
$[\hat{p} - 1.96\sqrt{\hat{p}(1-\hat{p})/B}, \;
  \hat{p} + 1.96\sqrt{\hat{p}(1-\hat{p})/B}]$,
where $\hat{p} = \text{FPR}_{\text{corr}}$ and $B$ is the
number of simulation replicates.  If the simulated FPR falls
within this interval, the approximation is well-calibrated.
:::
