---
title: "Detection Probability for GLM Outcomes"
subtitle: "Extending the Section 2.1 Approximation to General Response Types"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
  fig-align: center
knitr:
  opts_chunk:
    comment: "#>"
---

```{r setup}
#| include: false
```


# Introduction {#sec-intro}

Leon et al. (2024, Section 2.1, eq. 3) approximate the
probability that ForestSearch identifies a specific subgroup $H$
with true effect $\beta = \log(\theta^\dagger(H))$:

$$
p(c_1, c_2; d, \theta^\dagger(H)) =
P(W_1 + W_2 \ge 2k_1, \;\min(W_1, W_2) \ge k_2)
\tag{3}
$$

with $W_1, W_2 \stackrel{\text{iid}}{\sim} N(\beta, 8/d)$,
$k_1 = \log(c_1)$, $k_2 = \log(c_2)$.

This document extends (3) to all GLM outcome types supported
by the `forestsearch` GLM extension and provides a unified
implementation suitable for inclusion in the package.


# Mathematical Framework {#sec-math}

## The Unifying Quantity: $d_{\text{eff}}$

The approximation (3) depends on the data only through two
quantities: the true effect $\beta$ and the variance
$\sigma^2_s = 8/d$.  For the Cox model, $d$ is the number of
events.  For a general GLM, we define $d_{\text{eff}}$ such that
the split-half variance is $\sigma^2_s = 8/d_{\text{eff}}$, and
the approximation becomes:

$$
p(c_1, c_2; d_{\text{eff}}, \theta) =
P(W_1 + W_2 \ge 2k_1, \;\min(W_1, W_2) \ge k_2)
$$

with $W_1, W_2 \stackrel{\text{iid}}{\sim}
N(\beta, \; 8/d_{\text{eff}})$.

Below we derive $d_{\text{eff}}$ for each GLM family from the
Fisher information of the treatment coefficient.


## Derivation by Family

### Cox Proportional Hazards (Survival)

For the Cox model $\lambda(t; V) = \lambda_0(t)\exp(\beta V)$,
the partial likelihood score statistic satisfies (Jennison and
Turnbull, 1984):

$$
\hat\beta \approx \frac{4}{d} \tilde{L}_d(0)
\stackrel{d}{\to} N(\beta, \; 4/d)
$$

where $d = n_H(1 - p_c)$ is the expected number of events
($p_c$ = censoring proportion).

For a random 50/50 split, each half has $d/2$ events, giving:

$$
\hat\beta^j_s \sim N(\beta, \; 8/d), \quad j = 1, 2
$$

Therefore $d_{\text{eff}}^{\text{Cox}} = d = n_H(1 - p_c)$.

**Thresholds**: $k_1 = \log(c_1)$, $k_2 = \log(c_2)$ (log scale).


### Poisson Regression with Offset (Count Data)

For $Y_i \sim \text{Poisson}(\mu_i)$ with
$\log \mu_i = \beta_0 + \beta V_i + \log(t_i)$,
the Fisher information for $\beta$ is (McCullagh and Nelder,
1989, Section 2.5):

$$
I(\beta) = \sum_{i=1}^{n_H} \mu_i \cdot V_i^2
\cdot \left(1 - \frac{\sum_j \mu_j V_j}{\sum_j \mu_j}\right)^{-1}
$$

Under balanced randomization and approximately equal expected
counts per arm, this simplifies to:

$$
\text{Var}(\hat\beta) \approx \frac{1}{D_0} + \frac{1}{D_1}
$$

where $D_j = \sum_{i: V_i = j} \mu_i$ is the expected event
count in arm $j$.  For balanced arms with total events
$D = D_0 + D_1$:

$$
\text{Var}(\hat\beta_{\text{full}}) \approx \frac{4}{D},
\qquad
\text{Var}(\hat\beta_{\text{half}}) = \frac{8}{D}
$$

Therefore $d_{\text{eff}}^{\text{Poisson}} = D$ (total events).

**Connection to Cox**: Under proportional hazards with
piecewise-constant baseline, $D = d$ (Laird and Olivier, 1981;
Whitehead, 1980), so the Poisson+offset approximation is
identical to the Cox approximation.  This equivalence is
validated empirically in the companion document
`calibration_survival_fourway.qmd`.

**Thresholds**: $k_1 = \log(c_1)$, $k_2 = \log(c_2)$ (log scale).


### Logistic Regression (Binary Outcomes)

For $Y_i \sim \text{Bernoulli}(p_i)$ with
$\text{logit}(p_i) = \beta_0 + \beta V_i$,
the Fisher information for $\beta$ is:

$$
I(\beta) = \sum_{i=1}^{n_H} p_i(1 - p_i) \cdot w_i
$$

where $w_i$ is the leverage of observation $i$ for the
treatment coefficient.  For the simple two-sample model with
balanced arms ($n_1 = n_0 = n_H/2$) and approximately equal
event probabilities ($p_0 \approx p_1 \approx \bar p$):

$$
\text{Var}(\hat\beta) \approx
\frac{1}{n_1 \bar p(1 - \bar p)} +
\frac{1}{n_0 \bar p(1 - \bar p)} =
\frac{4}{n_H \bar p(1 - \bar p)}
$$

Therefore $d_{\text{eff}}^{\text{Binary}} =
n_H \cdot \bar p (1 - \bar p)$.

**Comparison to survival**: For binary outcomes with
$\bar p = 0.30$ and $n_H = 100$:
$d_{\text{eff}} = 100 \times 0.21 = 21$.  For survival with
45\% censoring: $d = 100 \times 0.55 = 55$.  The binary
$d_{\text{eff}}$ is less than half the survival $d$ for the
same $n_H$, explaining why the survival defaults (c1 = 1.25,
c2 = 1.0) provide weaker type-I error control for binary
outcomes.

**Thresholds**: $k_1 = \log(c_1)$, $k_2 = \log(c_2)$ (log scale).


### Gaussian Regression (Continuous Outcomes)

For $Y_i \sim N(\mu_i, \sigma^2)$ with
$\mu_i = \beta_0 + \beta V_i$ (identity link),
the Fisher information for $\beta$ is:

$$
I(\beta) = \frac{n_H}{4\sigma^2}
$$

(balanced arms).  Therefore:

$$
\text{Var}(\hat\beta_{\text{full}}) = \frac{4\sigma^2}{n_H},
\qquad
\text{Var}(\hat\beta_{\text{half}}) = \frac{8\sigma^2}{n_H}
$$

and $d_{\text{eff}}^{\text{Continuous}} = n_H / \sigma^2$.

**Note on thresholds**: The treatment effect $\beta$ is a mean
difference (on the identity scale), not a log ratio.  Therefore
the thresholds $k_1$ and $k_2$ are on the **identity scale**:
$k_1 = c_1$ (not $\log(c_1)$) and $k_2 = c_2$ (not $\log(c_2)$).
The user specifies these as meaningful clinical differences
(e.g., $c_1 = 0.5$ SD units, $c_2 = 0$).


### Quasi-Poisson (Overdispersed Count Data)

For overdispersed counts with dispersion $\hat\phi > 1$:

$$
\text{Var}(\hat\beta) = \hat\phi \cdot \frac{4}{D}
$$

Therefore $d_{\text{eff}}^{\text{Quasi}} = D / \hat\phi$.


## Summary Table

| Outcome | Link | $\text{Var}(\hat\beta_{\text{full}})$ | $d_{\text{eff}}$ | Effect scale |
|---|---|---|---|---|
| Survival (Cox) | log-HR | $4/d$ | $n_H(1 - p_c)$ | log |
| Poisson + offset | log-IRR | $4/D$ | $D$ (events) | log |
| Binomial | logit | $4/(n_H \bar p(1-\bar p))$ | $n_H \bar p(1-\bar p)$ | log |
| Gaussian | identity | $4\sigma^2/n_H$ | $n_H / \sigma^2$ | identity |
| Quasi-Poisson | log-IRR | $4\hat\phi/D$ | $D / \hat\phi$ | log |


# Implementation {#sec-impl}

## Unified Function: `compute_detection_probability_glm()`

The function takes $d_{\text{eff}}$ directly, with the same
cubature and Monte Carlo methods as the existing
`compute_detection_probability()`.

```{r unified-function}
#| code-fold: false

#' Compute Detection Probability for GLM Outcomes
#'
#' Extends the Section 2.1 approximation (Leon et al., 2024,
#' eq. 3) to arbitrary GLM outcomes via the effective information
#' \eqn{d_{\text{eff}}}.
#'
#' @param theta Numeric (scalar or vector). True treatment effect
#'   in the subgroup on the **natural scale**:
#'   \itemize{
#'     \item Ratio measures (HR, OR, IRR): the ratio itself
#'       (e.g., 2.0 for doubling of hazard/odds/rate).
#'     \item Difference measures (MD): the mean difference
#'       (e.g., 0.5 for half a unit improvement).
#'   }
#' @param d_eff Numeric. Effective information in the subgroup.
#'   Computed via the \code{d_eff_*} helper functions:
#'   \itemize{
#'     \item \code{d_eff_survival(n_sg, prop_cens)}
#'     \item \code{d_eff_binary(n_sg, p_event)}
#'     \item \code{d_eff_count(n_sg, event_rate)} or
#'       \code{d_eff_count(total_events = D)}
#'     \item \code{d_eff_continuous(n_sg, sigma_y)}
#'   }
#' @param c1 Numeric. Screening threshold on the natural scale:
#'   \itemize{
#'     \item Ratio measures: the ratio (e.g., 1.25)
#'     \item Difference measures: the difference (e.g., 0.2)
#'   }
#' @param c2 Numeric. Consistency threshold on the natural scale.
#'   Same convention as \code{c1}. Default: 1.0 for ratio
#'   measures, 0.0 for difference measures.
#' @param effect_scale Character. \code{"ratio"} (default) for
#'   log-scale effects (HR, OR, IRR), \code{"difference"} for
#'   identity-scale effects (MD).
#' @param method Character. Integration method.
#' @param n_mc Integer. Monte Carlo samples.
#' @param seed Integer. RNG seed for Monte Carlo.
#' @param verbose Logical. Progress messages for vector theta.
#'
#' @return If \code{theta} is scalar, a single probability.
#'   If vector, a data.frame with columns \code{theta},
#'   \code{probability}.
#'
#' @details
#' The detection criterion (Leon et al., 2024, Section 2.1):
#' \enumerate{
#'   \item Screening: \eqn{W_1 + W_2 \ge 2k_1}
#'   \item Consistency: \eqn{\min(W_1, W_2) \ge k_2}
#' }
#' with \eqn{W_1, W_2 \sim N(\beta, 8/d_{\text{eff}})}
#' independently, where:
#' \itemize{
#'   \item For ratio measures:
#'     \eqn{\beta = \log(\theta)}, \eqn{k_j = \log(c_j)}
#'   \item For difference measures:
#'     \eqn{\beta = \theta}, \eqn{k_j = c_j}
#' }
#'
#' @examples
#' \donttest{
#' # Survival: reproduce Figure 2 of Leon et al. (2024)
#' d <- d_eff_survival(n_sg = 60, prop_cens = 0.45)
#' compute_detection_probability_glm(
#'   theta = 2.0, d_eff = d, c1 = 1.25, c2 = 1.0
#' )
#'
#' # Binary: OR = 2.0, event rate = 30%
#' d <- d_eff_binary(n_sg = 100, p_event = 0.30)
#' compute_detection_probability_glm(
#'   theta = 2.0, d_eff = d, c1 = 1.25, c2 = 1.0
#' )
#'
#' # Continuous: mean difference = 0.5, sigma = 1.5
#' d <- d_eff_continuous(n_sg = 100, sigma_y = 1.5)
#' compute_detection_probability_glm(
#'   theta = 0.5, d_eff = d, c1 = 0.2, c2 = 0.0,
#'   effect_scale = "difference"
#' )
#' }
#'
#' @export
compute_detection_probability_glm <- function(
    theta,
    d_eff,
    c1 = 1.25,
    c2 = 1.0,
    effect_scale = c("ratio", "difference"),
    method = c("cubature", "monte_carlo"),
    n_mc = 500000L,
    seed = 42L,
    verbose = FALSE
) {

  # --- Argument matching ---
  effect_scale <- match.arg(effect_scale)
  method <- match.arg(method)

  # --- Input validation ---
  if (d_eff <= 0) {
    stop("d_eff must be positive")
  }
  if (effect_scale == "ratio" && any(theta <= 0)) {
    stop("theta must be positive for ratio-scale effects")
  }
  if (effect_scale == "ratio" && (c1 <= 0 || c2 <= 0)) {
    stop("c1 and c2 must be positive for ratio-scale effects")
  }
  if (d_eff < 5) {
    warning("d_eff < 5: normal approximation may be unreliable")
  }

  # --- Transform to natural parameter scale ---
  if (effect_scale == "ratio") {
    beta <- log(theta)
    k1 <- log(c1)
    k2 <- log(c2)
  } else {
    beta <- theta
    k1 <- c1
    k2 <- c2
  }

  # --- Split-half variance ---
  sigma2_s <- 8 / d_eff

  # --- Handle vector theta ---
  if (length(beta) > 1) {
    probs <- vapply(beta, function(b) {
      .detect_prob_single(b, sigma2_s, k1, k2, method, n_mc, seed)
    }, numeric(1))
    return(data.frame(theta = theta, probability = probs))
  }

  # --- Single theta ---
  .detect_prob_single(beta, sigma2_s, k1, k2, method, n_mc, seed)
}


# Internal workhorse
.detect_prob_single <- function(beta, sigma2_s, k1, k2,
                                 method, n_mc, seed) {
  if (method == "cubature") {
    if (!requireNamespace("cubature", quietly = TRUE)) {
      stop("Package 'cubature' required for method = 'cubature'.",
           call. = FALSE)
    }
    sd_s <- sqrt(sigma2_s)
    integrand <- function(x) {
      meets <- (x[1] + x[2] >= 2 * k1) &&
               (x[1] >= k2) && (x[2] >= k2)
      if (!meets) return(0)
      dnorm(x[1], beta, sd_s) * dnorm(x[2], beta, sd_s)
    }
    result <- cubature::adaptIntegrate(
      f = integrand,
      lowerLimit = c(k2, k2),
      upperLimit = c(beta + 6 * sd_s, beta + 6 * sd_s),
      tol = 1e-4
    )
    result$integral
  } else {
    set.seed(seed)
    sd_s <- sqrt(sigma2_s)
    w1 <- rnorm(n_mc, mean = beta, sd = sd_s)
    w2 <- rnorm(n_mc, mean = beta, sd = sd_s)
    mean((w1 + w2 >= 2 * k1) & (w1 >= k2) & (w2 >= k2))
  }
}
```


## d_eff Helper Functions

```{r deff-helpers}
#| code-fold: false

#' Effective Information for Survival (Cox PH)
#'
#' @param n_sg Integer. Subgroup sample size.
#' @param prop_cens Numeric. Proportion censored (0--1).
#' @return Numeric. d_eff = n_sg * (1 - prop_cens).
#' @export
d_eff_survival <- function(n_sg, prop_cens) {
  stopifnot(n_sg > 0, prop_cens >= 0, prop_cens < 1)
  n_sg * (1 - prop_cens)
}


#' Effective Information for Binary Outcomes (Logistic)
#'
#' @param n_sg Integer. Subgroup sample size.
#' @param p_event Numeric. Event probability (0--1).
#' @return Numeric. d_eff = n_sg * p * (1 - p).
#' @export
d_eff_binary <- function(n_sg, p_event) {
  stopifnot(n_sg > 0, p_event > 0, p_event < 1)
  n_sg * p_event * (1 - p_event)
}


#' Effective Information for Count Outcomes (Poisson)
#'
#' @param n_sg Integer. Subgroup sample size (optional if
#'   total_events provided).
#' @param event_rate Numeric. Mean event rate per patient
#'   (optional if total_events provided).
#' @param total_events Numeric. Total observed events D
#'   (optional; overrides n_sg * event_rate).
#' @param dispersion Numeric. Overdispersion parameter
#'   (1.0 for Poisson, > 1 for quasi-Poisson). Default: 1.0.
#' @return Numeric. d_eff = D / dispersion.
#' @export
d_eff_count <- function(n_sg = NULL, event_rate = NULL,
                        total_events = NULL, dispersion = 1.0) {
  if (!is.null(total_events)) {
    D <- total_events
  } else if (!is.null(n_sg) && !is.null(event_rate)) {
    D <- n_sg * event_rate
  } else {
    stop("Provide either total_events or both n_sg and event_rate")
  }
  stopifnot(D > 0, dispersion > 0)
  D / dispersion
}


#' Effective Information for Continuous Outcomes (Gaussian)
#'
#' @param n_sg Integer. Subgroup sample size.
#' @param sigma_y Numeric. Residual standard deviation of the
#'   outcome.
#' @return Numeric. d_eff = n_sg / sigma_y^2.
#' @export
d_eff_continuous <- function(n_sg, sigma_y) {
  stopifnot(n_sg > 0, sigma_y > 0)
  n_sg / sigma_y^2
}
```


## Backward Compatibility

The existing `compute_detection_probability()` is a special
case with `effect_scale = "ratio"` and
$d_{\text{eff}} = n_H (1 - p_c)$:

```{r backward-compat}
#| code-fold: false

# Existing call:
# compute_detection_probability(theta=2.0, n_sg=60, prop_cens=0.45,
#                                hr_threshold=1.25, hr_consistency=1.0)

# Equivalent GLM call:
# d <- d_eff_survival(n_sg = 60, prop_cens = 0.45)
# compute_detection_probability_glm(theta=2.0, d_eff=d,
#                                    c1=1.25, c2=1.0)
```


# Validation {#sec-validation}

```{r load-packages}
library(forestsearch)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
```


## Survival: Reproduce Figure 2

```{r validate-survival}
#| fig-height: 5
#| fig-cap: "Reproduction of Figure 2 (Leon et al., 2024). Solid lines: existing compute_detection_probability(). Dashed: new GLM function."

hr_seq <- seq(0.5, 3.0, by = 0.05)
n_values <- c(60, 80, 100)
prop_cens <- 0.45

curves <- do.call(rbind, lapply(n_values, function(n) {
  d <- d_eff_survival(n, prop_cens)

  # (D) Existing Cox function
  p_cox <- sapply(hr_seq, function(hr) {
    compute_detection_probability(hr, n, prop_cens, 1.25, 1.0,
                                  method = "monte_carlo",
                                  n_mc = 500000)
  })

  # (B) New GLM function
  p_glm <- sapply(hr_seq, function(hr) {
    compute_detection_probability_glm(hr, d, 1.25, 1.0,
                                      method = "monte_carlo")
  })

  data.frame(
    HR = rep(hr_seq, 2),
    P = c(p_cox, p_glm),
    Source = rep(c("(D) Existing Cox", "(B) GLM (d_eff)"),
                each = length(hr_seq)),
    n_label = sprintf("n = %d (d = %d)", n, round(d))
  )
}))

ggplot(curves, aes(x = HR, y = P, colour = n_label,
                   linetype = Source)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = c(0.10, 0.80), linetype = "dashed",
             colour = "grey50") +
  geom_vline(xintercept = 1.0, linetype = "dotted",
             colour = "grey70") +
  scale_linetype_manual(values = c("(D) Existing Cox" = "solid",
                                    "(B) GLM (d_eff)" = "dashed")) +
  labs(x = expression(theta^dagger*(H)),
       y = "P(detect H)",
       title = "Validation: GLM Function Reproduces Figure 2",
       subtitle = "Censoring = 45%. Solid = existing, Dashed = new (should overlap)",
       colour = "Subgroup", linetype = "Function") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Numerical Agreement Table

```{r validate-table}
#| tbl-cap: "Point-by-point comparison at key theta values. Ratio should be 1.000."

checkpoints <- expand.grid(
  n_sg = c(60, 80, 100),
  theta = c(0.75, 1.0, 1.5, 2.0, 2.5)
)
checkpoints$d <- round(checkpoints$n_sg * (1 - prop_cens))

checkpoints$P_cox <- mapply(function(hr, n) {
  compute_detection_probability(hr, n, prop_cens, 1.25, 1.0,
                                method = "monte_carlo", n_mc = 500000)
}, checkpoints$theta, checkpoints$n_sg)

checkpoints$P_glm <- mapply(function(hr, d) {
  compute_detection_probability_glm(hr, d, 1.25, 1.0,
                                    method = "monte_carlo")
}, checkpoints$theta, checkpoints$d)

checkpoints$Ratio <- round(checkpoints$P_glm /
                             pmax(checkpoints$P_cox, 1e-6), 4)

gt(checkpoints[, c("n_sg", "d", "theta",
                    "P_cox", "P_glm", "Ratio")]) |>
  tab_header(
    title = "Existing vs GLM Function: Numerical Agreement",
    subtitle = "Ratio = P_glm / P_cox (should be ~1.000)"
  ) |>
  fmt_number(columns = c(P_cox, P_glm), decimals = 4) |>
  cols_label(n_sg = "n(H)", d = "d", theta = "theta",
             P_cox = "P (existing)", P_glm = "P (GLM)",
             Ratio = "Ratio")
```


## Cross-Family Comparison

```{r cross-family}
#| fig-height: 7
#| fig-cap: "Power curves across outcome types for the same subgroup size (n=100). The curves differ because d_eff differs."

hr_seq <- seq(0.5, 3.0, by = 0.05)
n_sg <- 100

families <- list(
  list(name = "Survival (45% cens)",
       d = d_eff_survival(n_sg, 0.45),
       scale = "ratio"),
  list(name = "Poisson (rate=0.8/yr)",
       d = d_eff_count(n_sg, event_rate = 0.8),
       scale = "ratio"),
  list(name = "Binary (p=0.30)",
       d = d_eff_binary(n_sg, 0.30),
       scale = "ratio"),
  list(name = "Binary (p=0.50)",
       d = d_eff_binary(n_sg, 0.50),
       scale = "ratio")
)

family_curves <- do.call(rbind, lapply(families, function(fam) {
  probs <- sapply(hr_seq, function(hr) {
    compute_detection_probability_glm(hr, fam$d, 1.25, 1.0,
                                      effect_scale = fam$scale,
                                      method = "monte_carlo")
  })
  data.frame(
    theta = hr_seq, P = probs,
    Family = sprintf("%s (d_eff=%.0f)", fam$name, fam$d)
  )
}))

ggplot(family_curves, aes(x = theta, y = P, colour = Family)) +
  geom_line(linewidth = 1.0) +
  geom_hline(yintercept = c(0.10, 0.80), linetype = "dashed",
             colour = "grey50") +
  geom_vline(xintercept = 1.0, linetype = "dotted",
             colour = "grey70") +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression(theta^dagger*(H) ~ "(true effect)"),
       y = "P(detect H)",
       title = "Detection Probability Across Outcome Types",
       subtitle = "n(H) = 100, c1 = 1.25, c2 = 1.0",
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


```{r cross-family-table}
#| tbl-cap: "d_eff and detection probabilities by outcome type (n=100, c1=1.25, c2=1.0)"

fam_table <- data.frame(
  Outcome = c("Survival (45% cens)", "Poisson (rate=0.8/yr)",
              "Binary (p=0.30)", "Binary (p=0.50)",
              "Continuous (sigma=1.5)"),
  d_eff = c(
    d_eff_survival(100, 0.45),
    d_eff_count(100, event_rate = 0.8),
    d_eff_binary(100, 0.30),
    d_eff_binary(100, 0.50),
    d_eff_continuous(100, 1.5)
  ),
  stringsAsFactors = FALSE
)

fam_table$`P(theta=1.0)` <- sapply(fam_table$d_eff, function(d) {
  round(compute_detection_probability_glm(1.0, d, 1.25, 1.0,
    method = "monte_carlo"), 3)
})

fam_table$`P(theta=2.0)` <- sapply(fam_table$d_eff, function(d) {
  round(compute_detection_probability_glm(2.0, d, 1.25, 1.0,
    method = "monte_carlo"), 3)
})

fam_table$`80% power at` <- sapply(fam_table$d_eff, function(d) {
  # Find theta where power ~ 80%
  hr_fine <- seq(1.0, 4.0, by = 0.01)
  probs <- sapply(hr_fine, function(hr) {
    compute_detection_probability_glm(hr, d, 1.25, 1.0,
      method = "monte_carlo")
  })
  idx <- which.min(abs(probs - 0.80))
  hr_fine[idx]
})

gt(fam_table) |>
  tab_header(
    title = "Cross-Family Comparison (n = 100)",
    subtitle = "c1 = 1.25, c2 = 1.0"
  ) |>
  fmt_number(columns = d_eff, decimals = 1) |>
  cols_label(
    Outcome = "Outcome Type",
    d_eff = "d_eff",
    `P(theta=1.0)` = "Type-I Error",
    `P(theta=2.0)` = "Power (OR=2)",
    `80% power at` = "80% Power at"
  )
```


## Continuous Outcome: Identity Scale

```{r continuous-demo}
#| fig-height: 5
#| fig-cap: "Continuous outcome: thresholds on the identity scale. Horizontal axis is true mean difference."

md_seq <- seq(-0.5, 2.0, by = 0.05)
sigma_values <- c(1.0, 1.5, 2.0)

cont_curves <- do.call(rbind, lapply(sigma_values, function(s) {
  d <- d_eff_continuous(100, s)
  probs <- sapply(md_seq, function(md) {
    compute_detection_probability_glm(
      theta = md, d_eff = d, c1 = 0.3, c2 = 0.0,
      effect_scale = "difference", method = "monte_carlo")
  })
  data.frame(MD = md_seq, P = probs,
             sigma = sprintf("sigma = %.1f (d_eff = %.0f)", s, d))
}))

ggplot(cont_curves, aes(x = MD, y = P, colour = sigma)) +
  geom_line(linewidth = 1.0) +
  geom_hline(yintercept = c(0.10, 0.80), linetype = "dashed",
             colour = "grey50") +
  geom_vline(xintercept = 0, linetype = "dotted",
             colour = "grey70") +
  scale_colour_brewer(palette = "Dark2") +
  labs(x = "True Mean Difference",
       y = "P(detect H)",
       title = "Continuous Outcome: Detection Probability",
       subtitle = "n(H) = 100, c1 = 0.3, c2 = 0.0 (identity scale)",
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


# File for Package Integration {#sec-package}

The functions `compute_detection_probability_glm()`,
`d_eff_survival()`, `d_eff_binary()`, `d_eff_count()`,
and `d_eff_continuous()` are defined above with full roxygen2
documentation.  To add to the package:

1. Copy the function definitions into
   `R/truefind_asymptotic_glm.R`
2. Add `@export` tags (included above)
3. Run `devtools::document()`

The existing `compute_detection_probability()` remains
unchanged for backward compatibility.


# References {#sec-references}

- Leon, L.F., Jemielita, T., Guo, Z., Marceau West, R.,
  Anderson, K.M. (2024). Exploratory subgroup identification
  in the heterogeneous Cox model. *Statistics in Medicine*,
  43, 4253--4275.
- Jennison, C. and Turnbull, B.W. (1984). Repeated confidence
  intervals for group sequential clinical trials. *Controlled
  Clinical Trials*, 5(1), 33--45.
- McCullagh, P. and Nelder, J.A. (1989). *Generalized Linear
  Models*, 2nd ed. Chapman and Hall.
- Laird, N. and Olivier, D. (1981). Covariance analysis of
  censored survival data using log-linear analysis techniques.
  *JASA*, 76(374), 231--240.
- Whitehead, J. (1980). Fitting Cox's regression model to
  survival data using GLIM. *JRSS-C*, 29(3), 268--275.
- Agresti, A. (2002). *Categorical Data Analysis*, 2nd ed.
  Wiley. (Variance of log-OR: Section 3.1.3.)


```{r session-info}
sessionInfo()
```
