---
title: "Cross-cut Summary: sg_focus × FS Bundle (ACTG175 Binary, m1)"
format: html
---

This article compares operating characteristics across the
cross-product of **`sg_focus`** and **FS parameter bundle** for a
fixed study × outcome × DGM × objective combination:

- study : ACTG175
- outcome : binary
- DGM : m1 (calibrated GLM, target OR(Q) = 2.0)
- objective : harm
- varied : `sg_focus` (e.g., `hr`, `hrMaxSG`, `maxSG`) × FS bundle (`fs*`)

Each cell of the grid corresponds to one bundle file produced by
its companion runner script. The full intended grid is declared
below; missing bundles are skipped with a message (set
`strict_bundles <- TRUE` in setup to error instead).

```{r setup}
#| code-fold: true
#| code-summary: "Setup, configs, and library loads (click to expand)"

library(forestsearch)
library(data.table)
library(gt)



# -- Fixed axes (must match every bundle's config) --------------------------
fixed_study     <- "actg175"
fixed_outcome   <- "binary"
fixed_dgm       <- "m1"
fixed_objective <- "harm"
fixed_n         <- 700L     # per-replicate sample size  -> "n{N}"     slot
fixed_sims      <- 1000L    # number of simulation reps  -> "sims{NS}" slot


# bundle_dir <- "../"
# list.files(bundle_dir, pattern = "\\.rds$")
# prefix <- sprintf("^%s_%s_%s_%s_.*\\.rds$",
#                   fixed_study, fixed_outcome, fixed_dgm, fixed_objective)
# list.files(bundle_dir, pattern = prefix, full.names = TRUE)
# # 
# [10] "..//actg175_binary_m1_harm_fs2_eff_neighborhood_n700_sims1000.rds"     
# [11] "..//actg175_binary_m1_harm_fs2_effMaxSG_both_n700_sims1000.rds"        
# [12] "..//actg175_binary_m1_harm_fs2_effMaxSG_neighborhood_n700_sims1000.rds"
# [13] "..//actg175_binary_m1_harm_fs2_effMaxSG_pareto_n700_sims1000.rds"      
# [14] "..//actg175_binary_m1_harm_fs2_effMinSG_both_n700_sims1000.rds"        
# [15] "..//actg175_binary_m1_harm_fs2_effMinSG_neighborhood_n700_sims1000.rds"
# [16] "..//actg175_binary_m1_harm_fs2_effMinSG_pareto_n700_sims1000.rds"      
# [17] "..//actg175_binary_m1_harm_fs2_maxSG_neighborhood_n700_sims1000.rds"   
# [18] "..//actg175_binary_m1_harm_fs2_minSG_neighborhood_n700_sims1000.rds"   


# → only actg175_binary_m1_harm_* bundles, full paths
# Base R adds %||% in 4.4; define a fallback for older R.
if (!exists("%||%", mode = "function")) {
  `%||%` <- function(x, y) if (is.null(x)) y else x
}

# -- Levels to compare ------------------------------------------------------
# Edit these vectors as bundles become available.  expand.grid() builds
# the full cross-product; missing bundles are skipped (or error if
# strict_bundles = TRUE).
focus_levels   <- c("eff_neighborhood",
                    "maxSG_neighborhood","effMaxSG_neighborhood","effMaxSG_pareto","effMaxSG_both",
                    "minSG_neighborhood","effMinSG_neighborhood","effMinSG_pareto","effMinSG_both"
                    )
fs_levels      <- c("fs3")
strict_bundles <- FALSE





# -- Build configs ----------------------------------------------------------
# Path is relative to this qmd's directory
# (i.e., quarto/simulations/actg175/_summaries/).
configs <- expand.grid(
  focus = focus_levels,
  fs    = fs_levels,
  KEEP.OUT.ATTRS   = FALSE,
  stringsAsFactors = FALSE
)
# Sort by focus first, then fs (matches the naming convention).
configs <- configs[order(
  match(configs$focus, focus_levels),
  match(configs$fs,    fs_levels)
), , drop = FALSE]
rownames(configs) <- NULL
configs$config_id <- sprintf(
  "%s_%s_%s_%s_%s_%s_n%d_sims%d",
  fixed_study, fixed_outcome, fixed_dgm, fixed_objective, configs$fs,
  configs$focus, 
  fixed_n, fixed_sims
)
configs$label       <- sprintf("%s × %s", configs$focus, configs$fs)
configs$bundle_path <- sprintf("../%s.rds", configs$config_id)

# -- Filter to existing bundles --------------------------------------------
exists_flag <- vapply(configs$bundle_path, file.exists, logical(1))
if (!all(exists_flag)) {
  miss_msg <- sprintf("Missing %d of %d bundles:\n  %s",
                      sum(!exists_flag), nrow(configs),
                      paste(configs$config_id[!exists_flag],
                            collapse = "\n  "))
  if (strict_bundles) {
    stop(miss_msg, call. = FALSE)
  } else {
    message(miss_msg)
    configs <- configs[exists_flag, , drop = FALSE]
  }
}

if (nrow(configs) == 0L) {
  stop("No bundles available to compare. Build at least one bundle first.",
       call. = FALSE)
}

# -- Display options --------------------------------------------------------
sg_notation <- "harm"

# -- Conditional trimming for OR estimates ---------------------------------
# When the |mean - median| / |median| ratio exceeds or_trim_threshold for
# a given OR estimate (computed across replicates with detection), the
# raw mean is being pulled by extreme values. In that case we report a
# trimmed mean instead. Cells where the rule does not trigger keep the
# raw mean (preserving the standard "on average" interpretation).
#
# Set or_trim_threshold = Inf to disable conditional trimming entirely.
or_trim_threshold <- 0.20    # 20% gap between mean and median triggers trim
or_trim           <- 0.05    # 5% trimmed mean from each tail when triggered

cat(sprintf("Comparing %d configuration(s):\n", nrow(configs)))
print(configs[, c("label", "config_id")])
```

# Load bundles {#sec-load}

```{r load-bundles}
#| code-fold: true
#| code-summary: "Bundle loading + validation (click to expand)"

bundles <- setNames(lapply(configs$bundle_path, readRDS), configs$label)

# Soft-validate config_id matches what's declared in `configs`
for (i in seq_len(nrow(configs))) {
  meta_id <- bundles[[i]]$meta$config_id
  if (!is.null(meta_id) && meta_id != configs$config_id[i]) {
    warning(sprintf(
      "Bundle '%s' has config_id '%s' but configs declares '%s'.",
      configs$label[i], meta_id, configs$config_id[i]
    ), call. = FALSE)
  }
}
```

## Bundle metadata

```{r bundle-metadata}
#| code-fold: true
#| code-summary: "Per-bundle metadata table (click to expand)"

meta_dt <- data.table::rbindlist(lapply(seq_along(bundles), function(i) {
  m <- bundles[[i]]$meta
  data.frame(
    label           = configs$label[i],
    focus           = configs$focus[i],
    fs              = configs$fs[i],
    timestamp       = format(m$timestamp     %||% NA),
    n_alt           = m$n_alt                %||% NA_integer_,
    n_null          = m$n_null               %||% NA_integer_,
    sim_n           = m$sim_n                %||% NA_integer_,
    n_failed_alt    = m$n_failed_alt         %||% NA_integer_,
    n_failed_null   = m$n_failed_null        %||% NA_integer_,
    t_alt_min       = round((m$t_alt_sec  %||% NA) / 60, 1),
    t_null_min      = round((m$t_null_sec %||% NA) / 60, 1),
    package_version = m$package_version      %||% NA_character_,
    stringsAsFactors = FALSE
  )
}))

gt::gt(meta_dt) |>
  gt::tab_header(title = "Bundle metadata") |>
  gt::sub_missing(missing_text = "—")
```

# Unified H1 Operating Characteristics

A single combined table with one row per (config × analysis method).
FS rows are grouped by `sg_focus × fs` so detection, classification,
and effect estimates can be compared across `sg_focus` criteria.

GRF does not depend on `sg_focus` (its results are invariant across
the focus criteria), so a single GRF row is shown at the end of the
table under its own group rather than redundantly within each focus
group.

The DGM truth (target OR, true subgroup, prevalence) is fixed across
all configurations and shown once in the table subtitle, rather than
repeated per cell.

```{r unified-oc-data}
#| code-fold: true
#| code-summary: "Compute per-(config × analysis) H1 OC metrics (click to expand)"

# DGM truth values are fixed across configs (all share m1).
# Use the first available bundle's calibrated DGM as the reference.
ref_dgm <- bundles[[1]]$dgm$calibrated
or_target_H  <- ref_dgm$hazard_ratios$harm_subgroup
or_target_Hc <- ref_dgm$hazard_ratios$no_harm_subgroup
or_target_ITT <- ref_dgm$hazard_ratios$overall

# -- Helper: conditional trimmed mean ----------------------------------------
# Returns a list with the chosen mean (raw or trimmed) and a flag
# indicating whether trimming was applied. The rule is the mean-vs-median
# ratio: if |mean - median| / |median| exceeds the threshold, the mean
# is being pulled by extreme values and we report a trimmed mean.
# Otherwise the raw mean is returned (preserving "on average" reading).
cond_trim_mean <- function(x,
                           threshold = or_trim_threshold,
                           trim      = or_trim,
                           na.rm     = TRUE) {
  x_clean <- if (na.rm) x[!is.na(x)] else x
  if (length(x_clean) == 0L) {
    return(list(value = NA_real_, trimmed = FALSE))
  }
  mu  <- mean(x_clean)
  med <- stats::median(x_clean)
  # Avoid division by zero / sign issues; require a meaningful median
  if (is.na(med) || med == 0 || !is.finite(med)) {
    return(list(value = mu, trimmed = FALSE))
  }
  if (abs(mu - med) / abs(med) > threshold) {
    list(value = mean(x_clean, trim = trim), trimmed = TRUE)
  } else {
    list(value = mu, trimmed = FALSE)
  }
}

# Compute the same summary statistics format_oc_results() does, but
# combined across all configurations into one long data.frame.
oc_rows <- data.table::rbindlist(lapply(seq_len(nrow(configs)), function(i) {
  b <- bundles[[i]]
  res <- b$results$alt
  if (is.null(res) || nrow(res) == 0L) return(NULL)
  dt <- data.table::as.data.table(res)

  per_analysis <- lapply(unique(dt$analysis), function(a) {
    r       <- dt[analysis == a]
    r_found <- r[any.H == 1L]
    n_sims  <- length(unique(r$sim))

    # Conditional trimming applied to the three OR estimates
    or_h_hat   <- if (nrow(r_found)) cond_trim_mean(r_found$hr.H.hat)  else list(value = NA_real_, trimmed = FALSE)
    or_hc_hat  <- if (nrow(r_found)) cond_trim_mean(r_found$hr.Hc.hat) else list(value = NA_real_, trimmed = FALSE)
    or_itt     <- cond_trim_mean(r$hr.itt)

    list(
      group              = sprintf("%s × %s", configs$focus[i], configs$fs[i]),
      Analysis           = a,
      N_sims             = n_sims,
      Detection          = mean(r$any.H == 1L, na.rm = TRUE),
      Sen                = mean(r$sens, na.rm = TRUE),
      Spec               = mean(r$spec, na.rm = TRUE),
      PPV                = mean(r$ppv, na.rm = TRUE),
      NPV                = mean(r$npv, na.rm = TRUE),
      OR_H_hat           = or_h_hat$value,
      OR_H_trimmed       = or_h_hat$trimmed,
      OR_Hc_hat          = or_hc_hat$value,
      OR_Hc_trimmed      = or_hc_hat$trimmed,
      OR_ITT             = or_itt$value,
      OR_ITT_trimmed     = or_itt$trimmed,
      Size_H_mean        = if (nrow(r_found)) mean(r_found$size.H, na.rm = TRUE) else NA_real_,
      Size_H_min         = if (nrow(r_found)) min(r_found$size.H,  na.rm = TRUE) else NA_integer_,
      Size_H_max         = if (nrow(r_found)) max(r_found$size.H,  na.rm = TRUE) else NA_integer_
    )
  })
  data.table::rbindlist(per_analysis)
}))

# Replace NaN (from mean of all-NA) with NA for clean display.
# Skip logical columns (the *_trimmed flags).
for (col in names(oc_rows)) {
  if (is.numeric(oc_rows[[col]])) {
    oc_rows[is.nan(get(col)), (col) := NA]
  }
}

# Combined trim-flag column for compact display
oc_rows[, Trim := {
  flags <- character(.N)
  for (i in seq_len(.N)) {
    parts <- character(0)
    if (isTRUE(OR_H_trimmed[i]))   parts <- c(parts, "H")
    if (isTRUE(OR_Hc_trimmed[i]))  parts <- c(parts, "Hᶜ")
    if (isTRUE(OR_ITT_trimmed[i])) parts <- c(parts, "ITT")
    flags[i] <- if (length(parts)) paste(parts, collapse = ",") else ""
  }
  flags
}]

# Order: by configs row order (focus × fs), then analysis (FS, GRF)
oc_rows[, group := factor(group,
                          levels = sprintf("%s × %s",
                                           configs$focus, configs$fs))]
data.table::setorder(oc_rows, group, Analysis)

# -- Collapse redundant GRF rows -----------------------------------------
# GRF does not depend on sg_focus, so its results are identical across
# the focus groups. We collapse the redundant rows into a single GRF
# row at the end of the table, displayed under its own group.
#
# Sanity check: warn if GRF metrics actually differ across focus groups,
# which would suggest a bundle inconsistency rather than a true GRF
# invariance.
grf_rows <- oc_rows[Analysis == "GRF"]
if (nrow(grf_rows) > 1L) {
  metric_cols <- c("Detection", "Sen", "Spec", "PPV", "NPV",
                   "OR_H_hat", "OR_Hc_hat", "OR_ITT",
                   "Size_H_mean", "Size_H_min", "Size_H_max")
  # Check (within tolerance for floating point)
  inconsistent <- vapply(metric_cols, function(col) {
    vals <- grf_rows[[col]]
    vals <- vals[!is.na(vals)]
    if (length(vals) <= 1L) return(FALSE)
    diff_range <- max(vals) - min(vals)
    # Tolerance: 1e-8 absolute or 1e-6 relative to median
    abs_med <- abs(stats::median(vals))
    diff_range > max(1e-8, 1e-6 * abs_med)
  }, logical(1))
  if (any(inconsistent)) {
    warning(sprintf(
      "GRF rows differ across focus groups in column(s): %s. Collapsing anyway using the first focus group's values; investigate if unexpected.",
      paste(names(which(inconsistent)), collapse = ", ")
    ), call. = FALSE)
  }
}

# Keep one GRF row (relabel its group), drop the rest, append at end
if (nrow(grf_rows) > 0L) {
  one_grf <- grf_rows[1L]
  one_grf[, group := "GRF (sg_focus-invariant)"]
  oc_rows <- rbind(
    oc_rows[Analysis != "GRF"],
    one_grf
  )
  # Re-establish factor levels so the GRF group sorts to the end
  fs_grp_levels <- sprintf("%s × %s", configs$focus, configs$fs)
  oc_rows[, group := factor(group,
                            levels = c(fs_grp_levels,
                                       "GRF (sg_focus-invariant)"))]
  data.table::setorder(oc_rows, group, Analysis)
}

# How many cells got trimmed (used to decide whether to expose Trim col)
n_trimmed_cells <- sum(oc_rows$Trim != "")
```

```{r unified-oc-table}
# Build subtitle: include trim rule wording only if any cells got trimmed.
oc_subtitle <- sprintf(
  "Truth (m1): OR(Q) = %.2f, OR(Qc) = %.2f, OR(ITT) = %.2f — same across all configs",
  or_target_H, or_target_Hc, or_target_ITT
)
if (n_trimmed_cells > 0L) {
  oc_subtitle <- paste0(
    oc_subtitle,
    sprintf(
      ". OR estimates are raw means; cells where |mean − median|/|median| > %.0f%% report a %.0f%% trimmed mean instead (flagged in Trim column).",
      100 * or_trim_threshold, 100 * or_trim
    )
  )
}

# Drop the per-OR boolean trim flags from display; the combined Trim
# column is the user-facing flag.
display_dt <- oc_rows[, !c("OR_H_trimmed", "OR_Hc_trimmed", "OR_ITT_trimmed")]

# If no cell needed trimming, drop the Trim column too (no signal to surface)
if (n_trimmed_cells == 0L) {
  display_dt[, Trim := NULL]
}

g <- display_dt |>
  gt::gt(groupname_col = "group") |>
  gt::tab_header(
    title    = "H1 Operating Characteristics — focus × fs comparison",
    subtitle = oc_subtitle
  ) |>
  gt::cols_label(
    Analysis    = "Analysis",
    N_sims      = "N",
    Detection   = "Detect",
    Sen         = "Sen",
    Spec        = "Spec",
    PPV         = "PPV",
    NPV         = "NPV",
    OR_H_hat    = "ÔR(Ĥ)",
    OR_Hc_hat   = "ÔR(Ĥᶜ)",
    OR_ITT      = "ÔR(ITT)",
    Size_H_mean = "|Ĥ| mean",
    Size_H_min  = "min",
    Size_H_max  = "max"
  ) |>
  gt::tab_spanner(label = "Detection / Classification",
                  columns = c(Detection, Sen, Spec, PPV, NPV)) |>
  gt::tab_spanner(label = "Effect estimates",
                  columns = c(OR_H_hat, OR_Hc_hat, OR_ITT)) |>
  gt::tab_spanner(label = "Subgroup size |Ĥ|",
                  columns = c(Size_H_mean, Size_H_min, Size_H_max)) |>
  gt::fmt_number(columns = c(Detection, Sen, Spec, PPV, NPV), decimals = 3) |>
  gt::fmt_number(columns = c(OR_H_hat, OR_Hc_hat, OR_ITT), decimals = 2) |>
  gt::fmt_number(columns = Size_H_mean, decimals = 1) |>
  gt::fmt_number(columns = c(Size_H_min, Size_H_max), decimals = 0) |>
  gt::sub_missing(missing_text = "—")

if (n_trimmed_cells > 0L) {
  g <- g |> gt::cols_label(Trim = "Trim")
}

g
```

# H1 Classification (grouped across configurations) {#sec-classification}

`build_classification_table()` accepts a list of scenarios and
renders one row group per scenario, producing a single combined
table per analysis method. This puts every configuration's
classification metrics side-by-side — the natural view for
comparing sensitivity, specificity, PPV, and NPV across `sg_focus`
criteria.

```{r build-scenario-list}
#| code-fold: true
#| code-summary: "Build the scenario list passed to build_classification_table() (click to expand)"

# One scenario per (focus, fs) cell. The scenario name becomes the row
# group label in the gt output.
scenarios_alt <- setNames(
  lapply(seq_len(nrow(configs)), function(i) {
    b <- bundles[[i]]
    list(
      results    = b$results$alt,
      dgm        = b$dgm$calibrated,
      hypothesis = "alt",
      n_sample   = b$meta$sim_n,
      label      = sprintf("%s × %s", configs$focus[i], configs$fs[i])
    )
  }),
  paste(configs$focus, configs$fs, sep = "_")
)
```

## FS classification

```{r class-fs}
build_classification_table(
  scenarios_alt,
  analyses          = "FS",
  subgroup_notation = sg_notation,
  title             = "FS — H1 Classification (grouped by sg_focus × fs)"
)
```

## GRF classification

```{r class-grf}
build_classification_table(
  scenarios_alt,
  analyses          = "GRF",
  subgroup_notation = sg_notation,
  title             = "GRF — H1 Classification (grouped by sg_focus × fs)"
)
```

# Per-config Detail

For deep-dives into any single configuration, see the corresponding
per-config article.

```{r per-config-pointers}
#| echo: false
pointers <- data.frame(
  label   = configs$label,
  article = sprintf("../%s.qmd", configs$config_id),
  bundle  = configs$bundle_path,
  stringsAsFactors = FALSE
)
gt::gt(pointers) |>
  gt::tab_header(title = "Per-config articles and bundles") |>
  gt::cols_label(
    label   = "Config",
    article = "Per-config article",
    bundle  = "Bundle (.rds)"
  )
```

# Notes

- This article assumes all bundles share the **same DGM truth**
  (study × outcome × DGM × objective). Cross-cuts that vary the
  DGM need a different summary template (with per-bundle reference
  values for target OR, true |Q|, etc.).
- The unified H1 OC table reports raw means (`mean(...)`) for each
  OR estimate by default — preserving the standard "on average"
  interpretation. When the |mean − median|/|median| ratio for a
  given OR exceeds `or_trim_threshold` (default 0.20), the raw
  mean is being pulled by extreme values; in those cells the
  table reports a trimmed mean (default 5% from each tail) and
  flags the row in the **Trim** column. The threshold and trim
  level are configurable in the setup chunk; the **Trim** column
  is hidden entirely when no cell needed trimming.
- **GRF rows are collapsed**: GRF does not depend on `sg_focus`,
  so its results are identical across the five focus groups. The
  table shows GRF once, at the end, under its own group. A
  consistency check warns if the GRF rows actually differ across
  focus groups (which would indicate a bundle issue rather than
  true invariance).
- The grouped classification tables (one per analysis method) use
  `build_classification_table()` with all configurations passed as
  scenarios in a single call — producing one table per analysis
  method with row groups per `sg_focus × fs` cell.
- Toggle `strict_bundles <- TRUE` in setup to fail when any cell
  in the declared grid is missing — useful as a completeness check
  once all configurations are expected to be built.
