---
title: "Simulation Studies Based on ACTG175"
subtitle: "Adapting the ForestSearch GBSG Simulation Framework to HIV Clinical Trial Data"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ── Simulation control ──────────────────────────────────────────
nsims_alt   <- 44L
nsims_null  <- 44L
sim_seed    <- 8316951L
n_workers   <- min(
  parallel::detectCores(logical = FALSE) - 1L, 120L
)
```


# Introduction {#sec-intro}

## Motivation

Leon et al. (2024) demonstrated the ForestSearch methodology using
the German Breast Cancer Study Group (GBSG) dataset
(`survival::gbsg`) as both the applied analysis target and the
covariate template for simulation studies.  The `setup_gbsg_dgm()`
convenience function encodes all GBSG-specific preparation choices
--- variable discretisation, subgroup definition, Weibull
calibration --- into a single reproducible call.

This document adapts the same simulation framework to the **AIDS
Clinical Trials Group Study 175** (ACTG175) dataset from the
`speff2trial` package.  ACTG175 is a randomised trial of four
antiretroviral regimens in HIV-infected subjects (Hammer et al.,
*NEJM* 1996).  The adaptation serves two purposes:

1. **Methodological validation**: confirming that the ForestSearch
   operating characteristics (detection rate, sensitivity,
   specificity, PPV) generalise beyond the GBSG breast cancer
   setting to a different disease, population, endpoint
   structure, and covariate set.

2. **Template for new-dataset simulations**: demonstrating the
   recommended workflow when `setup_gbsg_dgm()` is not
   appropriate and the analyst must use
   `generate_aft_dgm_flex()` directly.


## Steps for Swapping GBSG with ACTG175 {#sec-swap-steps}

Replacing the GBSG template with ACTG175 requires changes at every
stage of the simulation pipeline.  The table below summarises each
change; subsequent sections implement them in detail.

| Step | GBSG (original) | ACTG175 (this document) |
|------|-----------------|-------------------------|
| **Data source** | `survival::gbsg` (686 patients) | `speff2trial::ACTG175`, arms 1 vs 2 (~1,046 patients) |
| **Outcome** | Recurrence-free survival in months (`rfstime / 30.4375`) | Time to first event in days (`days`) |
| **Event indicator** | `status` (1 = recurrence/death) | `cens` (1 = event) |
| **Treatment** | `hormon` (0/1) | `arms`: 1 = didanosine, 2 = zidovudine; coded as `treat = ifelse(arms == 1, 1, 0)` |
| **Confounders** | 7 binary variables: ER, age, meno, PgR, nodes, size, grade | 12 variables: age, weight, Karnofsky, CD4 baseline, CD8 baseline, hemophilia, homosexual, IV drugs, race, gender, prior antiretroviral, symptoms |
| **Discretisation** | ER $\leq$ Q25, age $\leq$ median, meno = premenopausal, etc. | Age $\leq$ 34, weight $\leq$ 75 kg, CD4 $\leq$ median, CD8 $>$ 1292, etc. (guided by GRF screening in the `.Rnw` analysis) |
| **Subgroup definition** | $H = \{\text{ER-low} \cap \text{premenopausal}\}$ ($\sim$13% prevalence) | $H = \{\text{age} \leq 34 \cap \text{CD8} > 1292\}$ ($\sim$15--20% prevalence) |
| **DGM builder** | `setup_gbsg_dgm()` | `generate_aft_dgm_flex()` |
| **Censoring** | Weibull; `analysis_time = 84` months | Weibull; `analysis_time = 1200` days ($\sim$3.3 years) |
| **ForestSearch parameters** | `n.min = 60`, `maxk = 2`, `hr.threshold = 1.25` | Same defaults; adjusted if needed after initial runs |

The subgroup definition for ACTG175 is derived from the GRF
screening analysis in `actg175_analysis_v1.Rnw`, which identified
age $\leq 34$ and CD8 count $> 1292$ as the primary splits
delineating a subgroup recommended for control (i.e., potential harm
from the experimental treatment).


# Side-by-Side Dataset Comparison {#sec-comparison}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(survival)
library(speff2trial)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)
```

```{r timing-init}
#| include: false
timings <- list()
t_vignette_start <- proc.time()
```


## Loading and Preparing Both Datasets

```{r prepare-datasets}
#| code-fold: false

# ── GBSG ──────────────────────────────────────────────────────
gbsg_df <- survival::gbsg
gbsg_df$y <- gbsg_df$rfstime / 30.4375
gbsg_df$treat <- gbsg_df$hormon

er_cut <- quantile(gbsg_df$er, probs = 0.25)
gbsg_df$z1 <- as.factor(ifelse(gbsg_df$er <= er_cut, 1L, 0L))
gbsg_df$z3 <- as.factor(ifelse(gbsg_df$meno == 0, 1L, 0L))
gbsg_df$flag_harm <- as.integer(gbsg_df$z1 == 1 & gbsg_df$z3 == 1)

# ── ACTG175 ───────────────────────────────────────────────────
actg_df <- subset(speff2trial::ACTG175, arms %in% c(1, 2))
actg_df$id <- seq_len(nrow(actg_df))
actg_df$time_days <- actg_df$days
actg_df$treat <- ifelse(actg_df$arms == 1, 1L, 0L)

# Subgroup-defining binary variables (from GRF screening)
actg_df$z1 <- as.factor(ifelse(actg_df$age <= 34, 1L, 0L))
actg_df$z2 <- as.factor(ifelse(actg_df$wtkg <= 75, 1L, 0L))
actg_df$z3 <- as.factor(ifelse(
  actg_df$karnof <= median(actg_df$karnof), 1L, 0L
))
actg_df$z4 <- as.factor(ifelse(
  actg_df$cd40 <= median(actg_df$cd40), 1L, 0L
))
actg_df$z5 <- as.factor(ifelse(actg_df$cd80 > 1292, 1L, 0L))
actg_df$z6 <- as.factor(actg_df$hemo)
actg_df$z7 <- as.factor(actg_df$homo)
actg_df$z8 <- as.factor(actg_df$drugs)
actg_df$z9 <- as.factor(actg_df$race)
actg_df$z10 <- as.factor(actg_df$gender)
actg_df$z11 <- as.factor(actg_df$oprior)
actg_df$z12 <- as.factor(actg_df$symptom)

# Harm subgroup: age <= 34 AND CD8 > 1292
actg_df$flag_harm <- as.integer(actg_df$z1 == 1 & actg_df$z5 == 1)
```


## Structural Comparison

```{r dataset-comparison-table}
#| tbl-cap: "Side-by-side comparison of GBSG and ACTG175 datasets"

comp <- data.frame(
  Feature = c(
    "Disease area",
    "Trial name",
    "R package",
    "Total N (used)",
    "Treatment arms",
    "Outcome",
    "Time unit",
    "Median follow-up",
    "Event rate",
    "Treatment effect (ITT HR)",
    "Subgroup definition (H)",
    "H prevalence",
    "N confounders (binary)",
    "Confounder types"
  ),
  GBSG = c(
    "Breast cancer",
    "German Breast Cancer Study Group",
    "survival::gbsg",
    sprintf("%d", nrow(gbsg_df)),
    "Hormonal therapy vs. control",
    "Recurrence-free survival",
    "Months",
    sprintf("%.1f months", median(gbsg_df$y)),
    sprintf("%.0f%%", 100 * mean(gbsg_df$status)),
    sprintf("%.3f", exp(coef(coxph(
      Surv(y, status) ~ treat, data = gbsg_df
    )))),
    "ER-low AND premenopausal",
    sprintf("%.0f%%", 100 * mean(gbsg_df$flag_harm)),
    "7",
    "All pre-discretised from continuous/categorical"
  ),
  ACTG175 = c(
    "HIV / AIDS",
    "AIDS Clinical Trials Group Study 175",
    "speff2trial::ACTG175",
    sprintf("%d", nrow(actg_df)),
    "Didanosine (ddI) vs. zidovudine (ZDV)",
    "Time to >= 50% CD4 decline, AIDS, or death",
    "Days",
    sprintf("%.0f days", median(actg_df$time_days)),
    sprintf("%.0f%%", 100 * mean(actg_df$cens)),
    sprintf("%.3f", exp(coef(coxph(
      Surv(time_days, cens) ~ treat, data = actg_df
    )))),
    "Age <= 34 AND CD8 > 1292",
    sprintf("%.0f%%", 100 * mean(actg_df$flag_harm)),
    "12",
    "Mix of binary indicators and discretised continuous"
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt(comp) |>
  tab_header(
    title = "Dataset Comparison: GBSG vs ACTG175"
  ) |>
  cols_label(Feature = "") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Feature)
  ) |>
  tab_style(
    style = cell_fill(color = "#F5F5F5"),
    locations = cells_body(rows = seq(1, nrow(comp), by = 2))
  )
```


## Confounder Mapping

```{r confounder-mapping}
#| tbl-cap: "Mapping of confounder variables between datasets"

conf_map <- data.frame(
  Role = c(
    "Subgroup factor 1",
    "Subgroup factor 2",
    "Prognostic 1",
    "Prognostic 2",
    "Prognostic 3",
    "Prognostic 4",
    "Prognostic 5",
    "Demographic 1",
    "Demographic 2",
    "Demographic 3",
    "Demographic 4",
    "Demographic 5"
  ),
  `GBSG variable` = c(
    "z1: ER <= Q25",
    "z3: premenopausal",
    "z2: age <= median",
    "z4: PgR <= median",
    "z5: nodes <= median",
    "v6: tumour size <= median",
    "v7: grade = 3",
    rep("---", 5)
  ),
  `ACTG175 variable` = c(
    "z1: age <= 34",
    "z5: CD8 > 1292",
    "z2: weight <= 75 kg",
    "z3: Karnofsky <= median",
    "z4: CD4 baseline <= median",
    "z6: hemophilia",
    "z7: homosexual activity",
    "z8: IV drug use",
    "z9: race (non-white)",
    "z10: gender (male)",
    "z11: prior antiretroviral",
    "z12: symptomatic"
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt(conf_map) |>
  tab_header(
    title = "Confounder Variable Mapping"
  ) |>
  tab_style(
    style = cell_fill(color = "#E3F2FD"),
    locations = cells_body(rows = 1:2)
  ) |>
  tab_source_note(
    "Highlighted rows define the harm subgroup H."
  )
```


## Survival Curves (Observed Data)

```{r km-comparison, fig.height=5, fig.width=10}
#| fig-cap: "Kaplan-Meier curves in the observed datasets. Left: GBSG (hormonal therapy). Right: ACTG175 (didanosine vs zidovudine)."

p_gbsg <- ggplot() +
  geom_step(
    data = data.frame(survfit(
      Surv(y, status) ~ treat, data = gbsg_df
    )),
    aes(x = time, y = surv, colour = factor(strata)),
    linewidth = 0.8
  ) +
  labs(
    title = "GBSG",
    x = "Months", y = "Survival Probability",
    colour = "Treatment"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# Use survfit directly for ACTG175
km_actg <- survfit(Surv(time_days, cens) ~ treat, data = actg_df)
km_actg_df <- data.frame(
  time = c(km_actg$time[1:km_actg$strata[1]],
           km_actg$time[(km_actg$strata[1] + 1):sum(km_actg$strata)]),
  surv = c(km_actg$surv[1:km_actg$strata[1]],
           km_actg$surv[(km_actg$strata[1] + 1):sum(km_actg$strata)]),
  arm = c(
    rep("ZDV (control)", km_actg$strata[1]),
    rep("ddI (treatment)", km_actg$strata[2])
  )
)

p_actg <- ggplot(km_actg_df, aes(x = time, y = surv, colour = arm)) +
  geom_step(linewidth = 0.8) +
  labs(
    title = "ACTG175",
    x = "Days", y = "Survival Probability",
    colour = "Treatment"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

p_gbsg + p_actg +
  plot_annotation(
    title = "Observed Survival in Source Datasets",
    theme = theme(plot.title = element_text(face = "bold"))
  )
```


## Subgroup Prevalence and Outcome

```{r subgroup-preview}
#| code-fold: false

cat("=== GBSG: Harm subgroup H = {ER-low AND premenopausal} ===\n")
cat("  N(H):", sum(gbsg_df$flag_harm), "/", nrow(gbsg_df),
    sprintf("(%.1f%%)\n", 100 * mean(gbsg_df$flag_harm)))
fit_gbsg_h <- coxph(
  Surv(y, status) ~ treat,
  data = gbsg_df[gbsg_df$flag_harm == 1, ]
)
cat("  Observed HR(H):",
    round(exp(coef(fit_gbsg_h)), 3), "\n\n")

cat("=== ACTG175: Harm subgroup H = {age<=34 AND CD8>1292} ===\n")
cat("  N(H):", sum(actg_df$flag_harm), "/", nrow(actg_df),
    sprintf("(%.1f%%)\n", 100 * mean(actg_df$flag_harm)))
fit_actg_h <- coxph(
  Surv(time_days, cens) ~ treat,
  data = actg_df[actg_df$flag_harm == 1, ]
)
cat("  Observed HR(H):",
    round(exp(coef(fit_actg_h)), 3), "\n")
fit_actg_hc <- coxph(
  Surv(time_days, cens) ~ treat,
  data = actg_df[actg_df$flag_harm == 0, ]
)
cat("  Observed HR(Hc):",
    round(exp(coef(fit_actg_hc)), 3), "\n")
```


# Creating the ACTG175 DGM {#sec-dgm}

Because `setup_gbsg_dgm()` is specific to the GBSG dataset, we use
`generate_aft_dgm_flex()` directly.  This requires the analyst to
supply the prepared dataset, specify all modelling choices, and
define the subgroup explicitly.

## Step 1: Prepare the Analysis Dataset

```{r prepare-actg-dgm}
#| code-fold: false

# Prepare the ACTG175 data for generate_aft_dgm_flex()
dfa <- actg_df

# Outcome in days (keep on original scale for AFT)
dfa$y <- dfa$time_days
dfa$event <- dfa$cens
```


## Step 2: Build the DGM (Alternative Hypothesis)

```{r create-dgm-alt}
t0 <- proc.time()[3]

dgm_alt <- generate_aft_dgm_flex(
  data            = dfa,
  outcome_var     = "y",
  event_var       = "event",
  treatment_var   = "treat",
  continuous_vars = character(0),
  factor_vars     = c(
    "z1", "z2", "z3", "z4", "z5",
    "z6", "z7", "z8", "z9", "z10", "z11", "z12"
  ),
  subgroup_vars   = c("z1", "z5"),
  subgroup_cuts   = list(z1 = 1, z5 = 1),
  model           = "alt",
  k_inter         = 2.0,
  k_treat         = 1.0,
  n_super         = 5000,
  seed            = sim_seed,
  verbose         = TRUE
)

# Compute CDEs from super-population potential outcomes
dgm_alt <- compute_dgm_cde(dgm_alt)

cat("\n=== ACTG175 Alternative DGM ===\n")
cat("  HR(H):", round(dgm_alt$hr_H_true, 3), "\n")
cat("  HR(Hc):", round(dgm_alt$hr_Hc_true, 3), "\n")
cat("  HR(ITT):", round(dgm_alt$hr_causal, 3), "\n")
cat("  AHR(H):", round(dgm_alt$AHR_H_true, 3), "\n")
cat("  AHR(Hc):", round(dgm_alt$AHR_Hc_true, 3), "\n")
cat("  H prevalence:",
    round(100 * mean(dgm_alt$df_super$flag_harm), 1), "%\n")

timings$dgm_alt <- proc.time()[3] - t0
```


## Step 3: Calibrate for Target HR {#sec-calibrate}

```{r calibrate-k-inter}
t0 <- proc.time()[3]

# Calibrate k_inter for a target HR(H) = 2.0
# Note: calibrate_k_inter() currently wraps setup_gbsg_dgm(),
# so for a non-GBSG dataset, we do a manual grid search.

target_hr <- 2.0

k_values <- seq(1.0, 5.0, by = 0.25)
cal_results <- data.frame(
  k_inter = k_values,
  hr_H = NA_real_,
  hr_Hc = NA_real_,
  stringsAsFactors = FALSE
)

for (i in seq_along(k_values)) {
  dgm_k <- tryCatch(
    generate_aft_dgm_flex(
      data            = dfa,
      outcome_var     = "y",
      event_var       = "event",
      treatment_var   = "treat",
      continuous_vars = character(0),
      factor_vars     = c(
        "z1", "z2", "z3", "z4", "z5",
        "z6", "z7", "z8", "z9", "z10", "z11", "z12"
      ),
      subgroup_vars   = c("z1", "z5"),
      subgroup_cuts   = list(z1 = 1, z5 = 1),
      model           = "alt",
      k_inter         = k_values[i],
      k_treat         = 1.0,
      n_super         = 5000,
      seed            = sim_seed,
      verbose         = FALSE
    ),
    error = function(e) NULL
  )

  if (!is.null(dgm_k)) {
    cal_results$hr_H[i]  <- dgm_k$hr_H_true
    cal_results$hr_Hc[i] <- dgm_k$hr_Hc_true
  }
}

# Find k_inter closest to target
cal_valid <- cal_results[!is.na(cal_results$hr_H), ]
best_idx <- which.min(abs(cal_valid$hr_H - target_hr))
k_inter_best <- cal_valid$k_inter[best_idx]

cat(sprintf("Calibration: k_inter = %.2f gives HR(H) = %.3f (target: %.1f)\n",
            k_inter_best, cal_valid$hr_H[best_idx], target_hr))

timings$calibrate <- proc.time()[3] - t0
```


```{r calibration-plot, fig.height=5}
#| fig-cap: "Calibration of k_inter: HR(H) as a function of the interaction parameter."

ggplot(cal_valid, aes(x = k_inter)) +
  geom_line(aes(y = hr_H, colour = "HR(H)"), linewidth = 1) +
  geom_line(aes(y = hr_Hc, colour = "HR(Hc)"), linewidth = 1) +
  geom_hline(yintercept = target_hr, linetype = "dashed",
             colour = "grey50") +
  geom_hline(yintercept = 1.0, linetype = "dotted",
             colour = "grey70") +
  geom_point(
    data = cal_valid[best_idx, ],
    aes(x = k_inter, y = hr_H),
    colour = "red", size = 4, shape = 16
  ) +
  scale_colour_manual(
    values = c("HR(H)" = "firebrick", "HR(Hc)" = "steelblue")
  ) +
  labs(
    x = "k_inter",
    y = "Hazard Ratio",
    title = "ACTG175 DGM Calibration",
    subtitle = sprintf("Best: k_inter = %.2f -> HR(H) = %.3f",
                        k_inter_best, cal_valid$hr_H[best_idx]),
    colour = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Step 4: Create Calibrated DGM

```{r dgm-calibrated}
t0 <- proc.time()[3]

dgm_calibrated <- generate_aft_dgm_flex(
  data            = dfa,
  outcome_var     = "y",
  event_var       = "event",
  treatment_var   = "treat",
  continuous_vars = character(0),
  factor_vars     = c(
    "z1", "z2", "z3", "z4", "z5",
    "z6", "z7", "z8", "z9", "z10", "z11", "z12"
  ),
  subgroup_vars   = c("z1", "z5"),
  subgroup_cuts   = list(z1 = 1, z5 = 1),
  model           = "alt",
  k_inter         = k_inter_best,
  k_treat         = 1.0,
  n_super         = 5000,
  seed            = sim_seed,
  verbose         = TRUE
)
dgm_calibrated <- compute_dgm_cde(dgm_calibrated)

cat("\n=== Calibrated ACTG175 DGM ===\n")
cat("  HR(H):", round(dgm_calibrated$hr_H_true, 3), "\n")
cat("  HR(Hc):", round(dgm_calibrated$hr_Hc_true, 3), "\n")
cat("  HR(ITT):", round(dgm_calibrated$hr_causal, 3), "\n")

timings$dgm_calibrated <- proc.time()[3] - t0
```


## Step 5: Create Null DGM

```{r dgm-null}
t0 <- proc.time()[3]

dgm_null <- generate_aft_dgm_flex(
  data            = dfa,
  outcome_var     = "y",
  event_var       = "event",
  treatment_var   = "treat",
  continuous_vars = character(0),
  factor_vars     = c(
    "z1", "z2", "z3", "z4", "z5",
    "z6", "z7", "z8", "z9", "z10", "z11", "z12"
  ),
  subgroup_vars   = c("z1", "z5"),
  subgroup_cuts   = list(z1 = 1, z5 = 1),
  model           = "null",
  k_treat         = 1.0,
  n_super         = 5000,
  seed            = sim_seed,
  verbose         = TRUE
)

cat("\n=== ACTG175 Null DGM ===\n")
cat("  HR(ITT):", round(dgm_null$hr_causal, 3), "\n")

timings$dgm_null <- proc.time()[3] - t0
```


# Single Trial Simulation {#sec-single-trial}

```{r simulate-single}
sim_data <- simulate_from_dgm(
  dgm           = dgm_calibrated,
  n             = 1000,
  rand_ratio    = 1,
  seed          = 1,
  analysis_time = 1200,
  cens_adjust   = 0
)

cat("Simulated ACTG175 trial:\n")
cat("  N =", nrow(sim_data), "\n")
cat("  Events =", sum(sim_data$event_sim),
    sprintf("(%.1f%%)\n", 100 * mean(sim_data$event_sim)))
cat("  Harm subgroup size =", sum(sim_data$flag_harm),
    sprintf("(%.1f%%)\n", 100 * mean(sim_data$flag_harm)))

fit_sim <- coxph(Surv(y_sim, event_sim) ~ treat_sim, data = sim_data)
cat("  Estimated ITT HR =", round(exp(coef(fit_sim)), 3), "\n")
```


# Running Simulation Studies {#sec-sims}

## Setup

```{r sim-setup}
#| code-fold: false

plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")

# Confounders for ForestSearch
confounders_actg <- c(
  "z1", "z2", "z3", "z4", "z5",
  "z6", "z7", "z8", "z9", "z10", "z11", "z12"
)

# ForestSearch parameters
fs_params <- list(
  outcome.name = "y_sim",
  event.name   = "event_sim",
  treat.name   = "treat_sim",
  id.name      = "id",
  use_lasso    = TRUE,
  use_grf      = TRUE,
  hr.threshold = 1.25,
  hr.consistency = 1.0,
  pconsistency.threshold = 0.90,
  fs.splits    = 400,
  n.min        = 60,
  d0.min       = 12,
  d1.min       = 12,
  maxk         = 2,
  by.risk      = 12,
  vi.grf.min   = -0.2,
  use_twostage = TRUE,
  twostage_args = list(),
  parallel_args = list(
    plan = "sequential", workers = 1, show_message = FALSE
  )
)

sim_config <- list(
  n_sample      = 1000,
  analysis_time = 1200,
  cens_adjust   = 0
)
```


## Alternative Hypothesis Simulations

```{r run-alt}
cat("Running", nsims_alt, "simulations under H1...\n")
t0 <- proc.time()[3]

results_alt <- foreach(
  sim = seq_len(nsims_alt),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_calibrated,
    n_sample         = sim_config$n_sample,
    analysis_time    = sim_config$analysis_time,
    cens_adjust      = sim_config$cens_adjust,
    confounders_base = confounders_actg,
    cox_formula_adj  = survival::Surv(y_sim, event_sim) ~
      treat_sim + z1 + z2 + z3,
    n_add_noise      = 0L,
    run_fs           = TRUE,
    run_fs_grf       = FALSE,
    run_grf          = TRUE,
    fs_params        = fs_params,
    verbose          = FALSE
  )
}

timings$sims_alt <- proc.time()[3] - t0
cat("Completed in", round(timings$sims_alt / 60, 1), "minutes\n")
cat("Results:", nrow(results_alt), "rows\n")
```


## Null Hypothesis Simulations

```{r run-null}
cat("Running", nsims_null, "simulations under H0...\n")
t0 <- proc.time()[3]

results_null <- foreach(
  sim = seq_len(nsims_null),
  .combine = rbind,
  .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_null,
    n_sample         = sim_config$n_sample,
    analysis_time    = sim_config$analysis_time,
    cens_adjust      = sim_config$cens_adjust,
    confounders_base = confounders_actg,
    cox_formula_adj  = survival::Surv(y_sim, event_sim) ~
      treat_sim + z1 + z2 + z3,
    n_add_noise      = 0L,
    run_fs           = TRUE,
    run_fs_grf       = FALSE,
    run_grf          = TRUE,
    fs_params        = fs_params,
    verbose          = FALSE
  )
}

timings$sims_null <- proc.time()[3] - t0
cat("Completed in", round(timings$sims_null / 60, 1), "minutes\n")
```


# Results {#sec-results}

## Operating Characteristics Under H1

```{r oc-alt}
t0 <- proc.time()[3]

format_oc_results(
  results_alt,
  metrics  = c("detection", "classification", "cde_estimates"),
  digits   = 2,
  digits_hr = 2,
  title    = "ACTG175: Identification Properties (HTEs)",
  subtitle = sprintf(
    "n = %d, %d sims, HR(H) = %.2f, HR(Hc) = %.2f",
    sim_config$n_sample, nsims_alt,
    dgm_calibrated$hr_H_true, dgm_calibrated$hr_Hc_true
  )
)

build_estimation_table(
  results         = results_alt,
  dgm             = dgm_calibrated,
  analysis_method = "FS",
  subtitle = sprintf(
    "ACTG175 DGM | n = %d, %d sims",
    sim_config$n_sample, nsims_alt
  ),
  font_size = 16
)

timings$summarize_alt <- proc.time()[3] - t0
```

### Interpretation

```{r interpret-alt, results="asis"}
interpret_estimation_table(
  results_alt, dgm_calibrated,
  analysis_method = "FS",
  scenario = "alt"
)
```


## Operating Characteristics Under H0

```{r oc-null}
t0 <- proc.time()[3]

format_oc_results(
  results_null,
  metrics  = c("detection", "classification"),
  digits   = 2,
  title    = "ACTG175: Type I Error (Null)",
  subtitle = sprintf(
    "n = %d, %d sims, HR(ITT) = %.2f",
    sim_config$n_sample, nsims_null,
    dgm_null$hr_causal
  )
)

timings$summarize_null <- proc.time()[3] - t0
```


## Classification Table

```{r classification}
scenario_list <- list(
  null = list(
    results    = results_null,
    label      = "ACTG",
    n_sample   = sim_config$n_sample,
    dgm        = dgm_null,
    hypothesis = "null"
  ),
  alt = list(
    results    = results_alt,
    label      = "ACTG",
    n_sample   = sim_config$n_sample,
    dgm        = dgm_calibrated,
    hypothesis = "alt"
  )
)

build_classification_table(
  scenario_results = scenario_list,
  analyses = sort(unique(c(
    unique(results_null$analysis),
    unique(results_alt$analysis)
  ))),
  digits   = 2,
  font_size = 16,
  title    = "ACTG175: Subgroup Identification and Classification",
  n_sims   = nsims_alt
)
```


# Theoretical Detection Probability {#sec-detection}

```{r detection-curve, fig.height=6}
n_sg_expected <- sim_config$n_sample *
  mean(dgm_calibrated$df_super$flag_harm)
prop_cens <- mean(results_alt$p.cens)

cat("Expected subgroup size:", round(n_sg_expected), "\n")
cat("Censoring proportion:", round(prop_cens, 3), "\n")
cat("True HR(H):", round(dgm_calibrated$hr_H_true, 3), "\n")

prob_detect <- compute_detection_probability(
  theta          = dgm_calibrated$hr_H_true,
  n_sg           = round(n_sg_expected),
  prop_cens      = prop_cens,
  hr_threshold   = fs_params$hr.threshold,
  hr_consistency = fs_params$hr.consistency,
  method         = "cubature"
)

cat("\nTheoretical P(detect):", round(prob_detect, 3), "\n")
cat("Empirical FS:",
    round(mean(results_alt[
      results_alt$analysis == "FS", "any.H"
    ]), 3), "\n")

# Generate curve
detection_curve <- generate_detection_curve(
  theta_range      = c(0.5, 3.0),
  n_points         = 50,
  n_sg             = round(n_sg_expected),
  prop_cens        = prop_cens,
  hr_threshold     = fs_params$hr.threshold,
  hr_consistency   = fs_params$hr.consistency,
  include_reference = TRUE,
  verbose          = FALSE
)

plot_detection_curve(
  detection_curve,
  add_reference_lines = TRUE,
  add_threshold_line  = TRUE,
  title = sprintf(
    "Detection Curve (n_sg=%d, cens=%.0f%%, c1=%.2f)",
    round(n_sg_expected), 100 * prop_cens,
    fs_params$hr.threshold
  )
)

abline(v = dgm_calibrated$hr_H_true, lty = 2, col = "blue")
```


# Timing Summary {#sec-timing}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Stage = c(
    "DGM creation (H1)",
    "k_inter calibration",
    "DGM calibrated",
    "DGM creation (H0)",
    "Simulations H1",
    "Simulations H0",
    "Summarize H1",
    "Summarize H0",
    "Total"
  ),
  Seconds = c(
    timings$dgm_alt,
    timings$calibrate,
    timings$dgm_calibrated,
    timings$dgm_null,
    timings$sims_alt,
    timings$sims_null,
    timings$summarize_alt,
    timings$summarize_null,
    timings$total
  ),
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(
    title = "Computational Timing Summary",
    subtitle = sprintf(
      "%d H1 + %d H0 sims | %d workers",
      nsims_alt, nsims_null, n_workers
    )
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  cols_label(
    Stage = "Stage",
    Seconds = "Time (sec)",
    Minutes = "Time (min)"
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total")
  )
```


# Discussion {#sec-discussion}

## Key Differences from the GBSG Simulation

The ACTG175 dataset introduces several features that test
ForestSearch under different conditions than the GBSG template:

1. **Larger sample size** ($N \approx 1{,}000$ vs $686$): more
   candidate subgroups survive the `n.min` filter, potentially
   increasing $L_{\text{eff}}$ and the procedure-level FPR.

2. **More confounders** ($12$ vs $7$): the combinatorial search
   space is larger, testing GRF and LASSO screening effectiveness.

3. **Different event rate**: ACTG175 has a lower event rate than
   GBSG, which reduces $d_{\text{eff}}$ per subgroup and may
   affect detection power.

4. **Different subgroup structure**: the age $\times$ CD8 subgroup
   is defined by continuous biomarkers discretised at clinically
   motivated cutpoints, rather than the GBSG's receptor-status
   $\times$ menopausal-status interaction.

5. **Days vs months**: the time scale is different, affecting the
   AFT model parametrisation but not the HR-scale operating
   characteristics.

## Limitations

- The ACTG175 subgroup definition (age $\leq 34$ AND CD8 $> 1292$)
  is exploratory, derived from a single GRF analysis. The true
  treatment effect heterogeneity pattern in ACTG175 is unknown.

- `generate_aft_dgm_flex()` requires more analyst judgment than
  `setup_gbsg_dgm()`. The variable discretisation, subgroup
  definition, and censoring calibration must all be specified
  manually.

- The simulation count (`r nsims_alt` sims) is for demonstration.
  Production studies should use $\geq 500$ simulations for stable
  operating characteristics.


# Session Info {#sec-session}

```{r session-info}
plan(sequential)
sessionInfo()
```


# References {#sec-references}

- Leon, L.F., Jemielita, T., Guo, Z., Marceau West, R.,
  Anderson, K.M. (2024). Exploratory subgroup identification
  in the heterogeneous Cox model. *Statistics in Medicine*,
  43, 4253--4275.

- Hammer, S.M., Katzenstein, D.A., Hughes, M.D., et al. (1996).
  A trial comparing nucleoside monotherapy with combination
  therapy in HIV-infected adults with CD4 cell counts from 200
  to 500 per cubic millimeter. *NEJM*, 335, 1081--1090.
