---
title: "Simulation Studies: Continuous Outcome (ACTG175)"
subtitle: "Subgroup Identification with Mean Difference in CD4 Change"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ─── Simulation control ─────────────────────────────────────────────────
nsims_alt   <- 1000L
nsims_null  <- 1000L
sim_seed    <- 8316951L
n_workers   <- min(parallel::detectCores(logical = FALSE) - 1L, 12L)

# ─── Dataset: ACTG175 ──────────────────────────────────────────────────
# Arms 1 (ZDV+ddI) vs 3 (ddI) — same as the binary simulation.
# Outcome: CD4 change from baseline to 20 weeks
#          cd4_change = cd420 - cd40
# Positive values indicate improvement (CD4 increase).
# With positive k_inter, the DGM creates a subgroup Q where ddI
# has an extreme positive effect (MORE CD4 improvement than combo),
# mirroring the working binary simulation's approach.
actg_arms        <- c(1L, 3L)
actg_treat_arm   <- 1L
actg_age_cut     <- 34
actg_preanti_cut <- 744.5

# ─── Benefit search via treatment switching ─────────────────────────────
# Under switched roles (ddI = treat, combo = control), the DGM
# creates a subgroup Q where ddI has an extreme POSITIVE effect
# (ddI increases CD4 change for Q more than combo does).
# This mirrors the working binary simulation: cal_target > 1 on the
# switched scale with positive k_inter grid.
# ForestSearch searches for MD >= threshold:
#   c1 = 30   (screening: MD >= 30 in Q)
#   c2 = 10   (consistency: each split MD >= 10)

# ─── Subgroup definition ────────────────────────────────────────────────
dgm_factor_vars   <- paste0("z", 1:12)
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)
dgm_n_super       <- 5000L

# ─── Calibration ────────────────────────────────────────────────────────
# Target: MD(Q) = +40 on the switched scale.
# Positive MD means ddI (treat=1) INCREASES CD4 change for Q by 40
# more than combo.  This mirrors the working binary simulation
# (cal_target_or = 2.0 with positive k_inter grid).
# ForestSearch searches for MD >= threshold, so positive MD works.
cal_target_md     <- 40
cal_k_grid_range  <- c(0, 120)
cal_grid_step     <- 2

# ─── Simulation study ──────────────────────────────────────────────────
sim_n_sample      <- 1000L
sim_n_noise       <- 0L

# ─── Analysis flags ────────────────────────────────────────────────────
run_fs            <- TRUE
run_fs_grf        <- FALSE
run_grf           <- FALSE

# ─── ForestSearch parameters ───────────────────────────────────────────
# For continuous outcomes with identity-scale measure (MD), thresholds
# are on the natural difference scale.
# ForestSearch searches for subgroups where MD >= threshold.
# With the DGM creating MD(Q) = +40, the threshold = 30 detects Q.
fs_md_threshold        <- 30     # MD >= 30
fs_md_consistency      <- 10     # MD >= 10 in each split
fs_pconsistency        <- 0.90
fs_splits              <- 400L
fs_n_min               <- 60L
fs_d0_min              <- 12L
fs_d1_min              <- 12L
fs_maxk                <- 2L
fs_vi_grf_min          <- -0.2
fs_seedit              <- 8316951L

# ─── ForestSearch pipeline flags ───────────────────────────────────────
fs_use_lasso           <- TRUE    # Gaussian-LASSO
fs_use_grf             <- TRUE
fs_is_rct              <- TRUE

# ─── Reporting ─────────────────────────────────────────────────────────
sg_notation            <- "benefit"

# ─── Output persistence ───────────────────────────────────────────────
save_results           <- TRUE
output_dir             <- "_data"
```


# Introduction {#sec-intro}

## Motivation

This document validates the ForestSearch GLM extension for
**continuous outcomes** using the ACTG175 trial.  Instead of
binarising CD4 response (`cd420 > cd40`), we use the raw CD4
change (`cd420 - cd40`) as a continuous endpoint analysed via
Gaussian GLM with mean difference (MD) as the effect measure.

This tests the full pipeline:

1. `generate_glm_dgm()` with `outcome_type = "continuous"`:
   potential means $\mu_0$, $\mu_1$
2. `calibrate_glm_interaction()` to find `k_inter` producing a
   target MD in the subgroup
3. `simulate_from_glm_dgm()` generating $Y = \mu + \varepsilon$,
   $\varepsilon \sim N(0, \sigma^2)$
4. `run_simulation_analysis()` dispatching via `"glm_dgm"` class
5. Operating characteristics via `format_oc_results()` and
   `build_estimation_table()`

## Clinical Context

In the ACTG175 trial, CD4 count at 20 weeks is a well-established
surrogate for clinical outcomes in HIV.  The combination therapy
(ZDV+ddI) generally produces larger CD4 increases than monotherapy
(ddI).  The question is whether there exists a subgroup where the
benefit of combination therapy is particularly pronounced.

Using the switched-treatment convention (ddI = treat = 1), a
positive `k_inter` increases E[Y|Q, treat=1], creating a subgroup
Q where ddI has an extreme positive effect on CD4 change.
ForestSearch detects this subgroup because MD(Q) exceeds the
threshold, while the population baseline effect is negative
(ddI generally hurts).  This mirrors the working binary simulation.


# Data Preparation {#sec-data}

```{r load-packages}
library(forestsearch)
library(speff2trial)
library(data.table)
library(gt)
library(ggplot2)
library(doFuture)
library(foreach)

# -- Safe foreach result collection -------------------------------------------
# collect_results() now lives in forestsearch (R/forestsearch_helpers.R).
# Called via ::: since it is an internal (unexported) utility.

timings <- list()
```

```{r prepare-data}
t0 <- proc.time()[3]

actg_df <- subset(speff2trial::ACTG175, arms %in% actg_arms)
actg_df$id <- seq_len(nrow(actg_df))

# Original treatment: arm 1 = 1, arm 3 = 0
actg_df$treat_orig <- ifelse(actg_df$arms == actg_treat_arm, 1L, 0L)

# Switched treatment for benefit search
actg_df$treat <- 1L - actg_df$treat_orig

# Continuous outcome: CD4 change from baseline to 20 weeks
actg_df$cd4_change <- actg_df$cd420 - actg_df$cd40

# Remove missing cd420
n_before <- nrow(actg_df)
actg_df <- actg_df[!is.na(actg_df$cd420), ]
n_after  <- nrow(actg_df)

cat(sprintf("ACTG175 arms 1 vs 3: N = %d -> %d after removing %d missing\n",
    n_before, n_after, n_before - n_after))
cat(sprintf("CD4 change: mean = %.1f, SD = %.1f, range = [%d, %d]\n",
    mean(actg_df$cd4_change), sd(actg_df$cd4_change),
    min(actg_df$cd4_change), max(actg_df$cd4_change)))
cat(sprintf("  ZDV+ddI (arm 1): mean = %.1f\n",
    mean(actg_df$cd4_change[actg_df$treat_orig == 1])))
cat(sprintf("  ddI (arm 3):     mean = %.1f\n",
    mean(actg_df$cd4_change[actg_df$treat_orig == 0])))
cat(sprintf("  MD (combo - mono, original scale): %.1f\n",
    mean(actg_df$cd4_change[actg_df$treat_orig == 1]) -
    mean(actg_df$cd4_change[actg_df$treat_orig == 0])))

# Discretise covariates (same as binary simulation)
actg_df$z1  <- as.factor(ifelse(actg_df$age > actg_age_cut, 1L, 0L))
actg_df$z2  <- as.factor(ifelse(actg_df$preanti <= actg_preanti_cut, 1L, 0L))
actg_df$z3  <- as.factor(ifelse(actg_df$wtkg <= 75, 1L, 0L))
actg_df$z4  <- as.factor(ifelse(
  actg_df$karnof <= median(actg_df$karnof), 1L, 0L))
actg_df$z5  <- as.factor(ifelse(
  actg_df$cd40 <= median(actg_df$cd40), 1L, 0L))
actg_df$z6  <- as.factor(ifelse(
  actg_df$cd80 <= median(actg_df$cd80), 1L, 0L))
actg_df$z7  <- as.factor(actg_df$hemo)
actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs)
actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender)
actg_df$z12 <- as.factor(actg_df$symptom)

# Reference subgroup Q
actg_df$flag_Q_ref <- as.integer(actg_df$z1 == 1 & actg_df$z2 == 1)
cat(sprintf("\nSubgroup Q: age > %d AND preanti <= %.0f\n",
    actg_age_cut, actg_preanti_cut))
cat(sprintf("  Q prevalence: %d / %d (%.1f%%)\n",
    sum(actg_df$flag_Q_ref), nrow(actg_df),
    100 * mean(actg_df$flag_Q_ref)))

# CD4 change by subgroup (original scale)
cat("\nCD4 change by subgroup (original treatment scale):\n")
for (grp in c("Q", "Qc")) {
  idx <- if (grp == "Q") actg_df$flag_Q_ref == 1 else actg_df$flag_Q_ref == 0
  md <- mean(actg_df$cd4_change[idx & actg_df$treat_orig == 1]) -
    mean(actg_df$cd4_change[idx & actg_df$treat_orig == 0])
  cat(sprintf("  %s (n=%d): MD(combo - mono) = %.1f\n",
      grp, sum(idx), md))
}

timings$data_prep <- proc.time()[3] - t0
```


# Creating the Continuous DGM {#sec-dgm}

## Step 1: Initial DGM

```{r create-dgm-alt}
t0 <- proc.time()[3]

dgm_alt_init <- generate_glm_dgm(
  data          = actg_df,
  factor_vars   = dgm_factor_vars,
  outcome_var   = "cd4_change",
  treatment_var = "treat",
  outcome_type  = "continuous",
  effect_measure = "MD",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  model         = "alt",
  k_inter       = 0,
  n_super       = dgm_n_super,
  seed          = sim_seed,
  verbose       = TRUE
)

cat("\n=== Initial ACTG175 Continuous DGM ===\n")
cat("  MD(Q):  ", round(dgm_alt_init$hazard_ratios$harm_subgroup, 1), "\n")
cat("  MD(Qc): ", round(dgm_alt_init$hazard_ratios$no_harm_subgroup, 1), "\n")
cat("  MD(ITT):", round(dgm_alt_init$hazard_ratios$overall, 1), "\n")
cat("  sigma:  ", round(dgm_alt_init$model_params$sigma, 1), "\n")

timings$dgm_init <- proc.time()[3] - t0
```

## Step 2: Calibrate for Target MD {#sec-calibrate}

The `k_inter` parameter is a direct additive shift on the mean for
Q members under treatment.  Negative `k_inter` decreases E[Y|Q, T=1],
meaning ddI (switched treatment) produces worse CD4 gains for Q.

```{r calibrate}
t0 <- proc.time()[3]

dgm_calibrated <- calibrate_glm_interaction(
  data          = actg_df,
  factor_vars   = dgm_factor_vars,
  outcome_var   = "cd4_change",
  treatment_var = "treat",
  target_effect = cal_target_md,
  outcome_type  = "continuous",
  effect_measure = "MD",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  k_inter_range = cal_k_grid_range,
  grid_step     = cal_grid_step,
  n_super       = dgm_n_super,
  seed          = sim_seed,
  verbose       = TRUE
)

cat("\n=== Calibrated ACTG175 Continuous DGM ===\n")
cat("  Switched scale (ddI = treat):\n")
cat("    MD(Q):  ", round(dgm_calibrated$hazard_ratios$harm_subgroup, 1),
    " (target:", cal_target_md, ")\n")
cat("    MD(Qc): ", round(dgm_calibrated$hazard_ratios$no_harm_subgroup, 1), "\n")
cat("    MD(ITT):", round(dgm_calibrated$hazard_ratios$overall, 1), "\n")
cat("  -> Original scale: MD(Q) =",
    round(-dgm_calibrated$hazard_ratios$harm_subgroup, 1), "\n")
cat("  k_inter:", round(dgm_calibrated$model_params$k_inter, 2), "\n")
cat("  sigma:  ", round(dgm_calibrated$model_params$sigma, 1), "\n")

timings$calibrate <- proc.time()[3] - t0
```

### Calibration Curve

```{r calibration-plot, fig.height=5}
#| fig-cap: "Calibration of k_inter: MD(Q) as a function of the interaction parameter."

cal_grid <- seq(cal_k_grid_range[1], cal_k_grid_range[2], by = cal_grid_step)
cal_md_Q <- vapply(cal_grid, function(k) {
  d <- generate_glm_dgm(
    data = actg_df, factor_vars = dgm_factor_vars,
    outcome_var   = "cd4_change", treatment_var = "treat",
    outcome_type = "continuous", effect_measure = "MD",
    subgroup_vars = dgm_subgroup_vars,
    subgroup_cuts = dgm_subgroup_cuts, model = "alt",
    k_inter = k, n_super = dgm_n_super, seed = sim_seed
  )
  d$hazard_ratios$harm_subgroup
}, numeric(1))

cal_df <- data.frame(k_inter = cal_grid, md_Q = cal_md_Q)
best_k <- dgm_calibrated$model_params$k_inter

ggplot(cal_df, aes(x = k_inter, y = md_Q)) +
  geom_line(linewidth = 1, colour = "darkorange") +
  geom_hline(yintercept = cal_target_md, linetype = "dashed",
             colour = "grey50") +
  geom_hline(yintercept = 0, linetype = "dotted",
             colour = "grey70") +
  geom_point(
    data = data.frame(k_inter = best_k,
                      md_Q = dgm_calibrated$hazard_ratios$harm_subgroup),
    colour = "red", size = 4, shape = 16
  ) +
  labs(
    x = "k_inter (mean shift for Q under treatment)",
    y = "MD(Q) on switched scale",
    title = "ACTG175 Continuous DGM Calibration",
    subtitle = sprintf("Best: k_inter = %.0f -> MD(Q) = %.1f",
      best_k, dgm_calibrated$hazard_ratios$harm_subgroup)
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```

## Step 3: Null DGM

```{r dgm-null}
t0 <- proc.time()[3]

dgm_null <- generate_glm_dgm(
  data          = actg_df,
  factor_vars   = dgm_factor_vars,
  outcome_var   = "cd4_change",
  treatment_var = "treat",
  outcome_type  = "continuous",
  effect_measure = "MD",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  model         = "null",
  n_super       = dgm_n_super,
  seed          = sim_seed,
  verbose       = TRUE
)

cat("\n=== ACTG175 Null DGM ===\n")
cat("  MD(ITT):", round(dgm_null$hazard_ratios$overall, 1), "\n")
cat("  MD(Q):  ", round(dgm_null$hazard_ratios$harm_subgroup, 1), "\n")
cat("  MD(Qc): ", round(dgm_null$hazard_ratios$no_harm_subgroup, 1), "\n")

timings$dgm_null <- proc.time()[3] - t0
```


# Single Trial Verification {#sec-single-trial}

```{r single-trial}
#| code-fold: false
t0 <- proc.time()[3]

df_single <- simulate_from_glm_dgm(
  dgm  = dgm_calibrated,
  n    = sim_n_sample,
  seed = sim_seed + 1L
)

cat(sprintf("Single trial: N = %d\n", nrow(df_single)))
cat(sprintf("  CD4 change: mean = %.1f, SD = %.1f\n",
    mean(df_single$y_sim), sd(df_single$y_sim)))
cat(sprintf("  Q prevalence: %.1f%%\n",
    100 * mean(df_single$flag_harm)))

# Verify MD in the simulated data
md_sim <- mean(df_single$y_sim[df_single$treat_sim == 1]) -
  mean(df_single$y_sim[df_single$treat_sim == 0])
cat(sprintf("  Simulated ITT MD: %.1f (DGM: %.1f)\n",
    md_sim, dgm_calibrated$hazard_ratios$overall))

timings$single_trial <- proc.time()[3] - t0
```


# Running Simulation Studies {#sec-sims}

## Setup

```{r sim-setup}
#| code-fold: false

plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")

confounders_actg <- dgm_calibrated$factor_vars

fs_params <- list(
  outcome.name           = "y_sim",
  treat.name             = "treat_sim",
  id.name                = "id",

  # GLM pipeline
  outcome_type           = "continuous",
  effect_measure         = "MD",

  # Variable screening
  use_lasso              = fs_use_lasso,
  use_grf                = fs_use_grf,

  # Subgroup detection thresholds (identity scale)
  effect.threshold       = fs_md_threshold,
  consistency.threshold  = fs_md_consistency,
  pconsistency.threshold = fs_pconsistency,

  # Constraints
  fs.splits              = fs_splits,
  n.min                  = fs_n_min,
  d0.min                 = fs_d0_min,
  d1.min                 = fs_d1_min,
  maxk                   = fs_maxk,
  vi.grf.min             = fs_vi_grf_min,
  seedit                 = fs_seedit,

  parallel_args          = list(
    plan = "sequential", workers = 1, show_message = FALSE
  )
)
```


## Alternative Hypothesis Simulations

```{r run-alt}
cat("Running", nsims_alt, "simulations under H1...\n")
t0 <- proc.time()[3]

results_alt_raw <- foreach(
  sim = seq_len(nsims_alt),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_calibrated,
    n_sample         = sim_n_sample,
    confounders_base = confounders_actg,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    verbose          = FALSE
  )
}

timings$sims_alt <- proc.time()[3] - t0

results_alt <- forestsearch:::collect_results(results_alt_raw, "H1 alt")
n_failed_alt <- attr(results_alt, "n_failed")

cat("Completed in", round(timings$sims_alt / 60, 1), "minutes\n")
cat("Results:", nrow(results_alt), "rows\n")
if (n_failed_alt > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_alt))
cat(sprintf("Detection rate: %.1f%%\n",
    100 * mean(results_alt$any.H[results_alt$analysis == "FS"])))
```


## Null Hypothesis Simulations

```{r reset-workers-null}
# Reset worker pool to prevent memory accumulation from alt sims.
# Without this, null sims run on bloated workers and can be ~20x slower.
plan(sequential); gc(verbose = FALSE); Sys.sleep(1)
plan(multisession, workers = n_workers)
cat(sprintf("Workers reset: %d cores\n", n_workers))
```

```{r run-null}
cat("Running", nsims_null, "simulations under H0...\n")
t0 <- proc.time()[3]

results_null_raw <- foreach(
  sim = seq_len(nsims_null),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_null,
    n_sample         = sim_n_sample,
    confounders_base = confounders_actg,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    verbose          = FALSE
  )
}

timings$sims_null <- proc.time()[3] - t0

results_null <- forestsearch:::collect_results(results_null_raw, "H0 null")
n_failed_null <- attr(results_null, "n_failed")

cat("Completed in", round(timings$sims_null / 60, 1), "minutes\n")
if (n_failed_null > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_null))
cat(sprintf("False discovery rate: %.1f%%\n",
    100 * mean(results_null$any.H[results_null$analysis == "FS"])))
```

```{r cleanup-plan}
plan(sequential)
```

```{r save-results}
#| code-fold: false

if (save_results) {
  if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

  bundle <- list(
    results_alt    = results_alt,
    results_null   = results_null,
    dgm_calibrated = dgm_calibrated,
    dgm_null       = dgm_null,
    config = list(
      nsims_alt    = nsims_alt,
      nsims_null   = nsims_null,
      n_sample     = sim_n_sample,
      n_noise      = sim_n_noise,
      n_workers    = n_workers,
      fs_params    = fs_params,
      target_md    = cal_target_md,
      subgroup     = paste(dgm_subgroup_vars, collapse = " & ")
    ),
    true_effects = list(
      md_Q     = dgm_calibrated$hazard_ratios$harm_subgroup,
      md_Qc    = dgm_calibrated$hazard_ratios$no_harm_subgroup,
      md_ITT   = dgm_calibrated$hazard_ratios$overall,
      k_inter  = dgm_calibrated$model_params$k_inter,
      sigma    = dgm_calibrated$model_params$sigma
    ),
    timings = timings
  )

  fname <- sprintf("actg175_continuous_simulations_%dalts_%dnulls.rds",
      nsims_alt, nsims_null)
  saveRDS(bundle, file.path(output_dir, fname))
  cat("Saved:", file.path(output_dir, fname), "\n")
} else {
  cat("save_results = FALSE; skipping save.\n")
}
```


# Results {#sec-results}

## Operating Characteristics Under H1

```{r oc-alt}
t0 <- proc.time()[3]

format_oc_results(
  results_alt,
  metrics   = c("detection", "classification"),
  digits    = 2,
  digits_hr = 1,
  title     = "ACTG175 Continuous: Subgroup Identification",
  subtitle  = sprintf(
    "n = %d, %d sims, MD(Q) = %.1f, MD(Qc) = %.1f (switched)",
    sim_n_sample, nsims_alt,
    dgm_calibrated$hazard_ratios$harm_subgroup,
    dgm_calibrated$hazard_ratios$no_harm_subgroup
  ),
  subgroup_notation = sg_notation
)

build_estimation_table(
  results         = results_alt,
  dgm             = dgm_calibrated,
  analysis_method = "FS",
  subtitle = sprintf(
    "ACTG175 Continuous DGM | n = %d, %d sims",
    sim_n_sample, nsims_alt
  ),
  font_size = 16,
  subgroup_notation = sg_notation
)

timings$report_alt <- proc.time()[3] - t0
```


## Operating Characteristics Under H0

```{r oc-null}
format_oc_results(
  results_null,
  metrics  = c("detection", "classification"),
  digits   = 2,
  title    = "ACTG175 Continuous: Type I Error (Null)",
  subtitle = sprintf(
    "n = %d, %d sims, MD(ITT) = %.1f (switched)",
    sim_n_sample, nsims_null,
    dgm_null$hazard_ratios$overall
  ),
  subgroup_notation = sg_notation
)
```


## Classification Table

```{r classification}
build_classification_table(
  scenario_results = list(
    null = list(
      results    = results_null,
      label      = "ACTG-cont",
      n_sample   = sim_n_sample,
      dgm        = dgm_null,
      hypothesis = "null"
    ),
    alt = list(
      results    = results_alt,
      label      = "ACTG-cont",
      n_sample   = sim_n_sample,
      dgm        = dgm_calibrated,
      hypothesis = "alt"
    )
  ),
  analyses = sort(unique(c(
    unique(results_null$analysis),
    unique(results_alt$analysis)
  ))),
  digits    = 2,
  font_size = 16,
  title     = "ACTG175 Continuous: Subgroup Identification and Classification",
  n_sims    = nsims_alt,
  subgroup_notation = sg_notation
)
```


# Timing Summary {#sec-timing}

```{r timing}
timings$total <- Reduce(`+`, timings)

timing_df <- data.frame(
  Step = names(timings),
  Seconds = round(unlist(timings), 1),
  Minutes = round(unlist(timings) / 60, 2)
)

timing_df |>
  gt() |>
  gt::tab_header(title = "Computation Timing") |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(rows = Step == "total")
  )
```


# Discussion {#sec-discussion}

## Key Design Decisions

1. **Positive k_inter, positive target**: Following the working binary
   simulation's pattern, the DGM uses a positive `k_inter` to create
   a subgroup Q where ddI has an extreme positive effect (MD = +40).
   The population baseline effect is negative (ddI generally hurts),
   so there is a real Q vs Qc contrast that ForestSearch can detect.

2. **Identity-scale thresholds**: For continuous outcomes with
   `effect_measure = "MD"`, the screening and consistency thresholds
   are on the natural difference scale (e.g., MD >= 30 cells/$\mu$L).
   This differs from ratio-scale measures (HR, OR, IRR) where
   thresholds are multiplicative.

3. **Residual variance**: The DGM preserves the residual SD
   ($\sigma$) from the baseline Gaussian GLM fit, ensuring
   realistic noise levels in simulated data.

4. **Linear calibration**: For Gaussian GLM with identity link,
   `k_inter` maps linearly to MD(Q): $\text{MD}(Q) = \beta_T + k$
   where $\beta_T$ is the fitted treatment coefficient.  This makes
   calibration trivial compared to the non-linear binary (logistic)
   case.

## Comparison with the Binary ACTG175 Simulation

The same subgroup definition (z1 = age > 34, z2 = preanti <= 745)
is used.  Differences:

- **No floor/ceiling effects**: continuous CD4 change has no natural
  bounds, so the DGM does not need the careful probability-clipping
  logic of the binary version.
- **Gaussian LASSO**: variable screening uses `glmnet` with
  `family = "gaussian"` instead of `"binomial"`.
- **Effect size interpretation**: MD = +40 cells/$\mu$L means ddI
  increases CD4 change by 40 more than combo for Q.  On the original
  scale, this means combo provides 40 less CD4 benefit for Q.


# Session Info {#sec-session}

```{r session-info}
sessionInfo()
```


# References {#sec-references}

- Hammer SM, et al. (1996). *NEJM*, 335, 1081--1090.
- León LF, et al. (2024). *Statistics in Medicine*, 43, 3921--3942.
