---
title: "Null Threshold Search for Binary Outcomes"
subtitle: "Mapping the (c1, c2) Operating Region Where FPR <= 10-12%"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

nsims_null  <- 20L
sim_seed    <- 8316951L
n_workers   <- max(1L, parallel::detectCores(logical = FALSE) - 1L)
```


# Introduction {#sec-intro}

The companion document `binary_threshold_calibration.qmd`
establishes that:

- The per-subgroup approximation $P_1$ is accurate for
  **power** (H1) at $N \ge 500$
- Under **H0**, the procedure-level FPR exceeds $P_1$ by a
  factor $L_{\text{eff}}$ that grows with $N$
- The $L_{\text{eff}}$-corrected approximation
  $1 - (1-P_1)^{L_{\text{eff}}}$ aligns well with simulation

This document focuses exclusively on **null operating
characteristics** for the binary DGM.  We sweep a dense
$(c_1, c_2)$ grid at multiple sample sizes to identify
threshold configurations where the simulated FPR is at most
10--12%.

All simulations are under $H_0$ (uniform treatment effect,
no subgroup heterogeneity).


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)
```

```{r timing-init}
timings <- list()
t_vignette_start <- proc.time()
```

```{r setup-parallel}
plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")
cat("Sims per config:", nsims_null, "\n")
```


# Binary Null DGM {#sec-dgm}

Under $H_0$: no treatment-by-biomarker interaction.  The
outcome probability is $\approx 30\%$ with prognostic effects
from age and ECOG but no treatment effect heterogeneity.

```{r sim-null-fn}
#| code-fold: false

sim_binary_null <- function(n, seed) {
  set.seed(seed)
  df <- data.frame(
    id = seq_len(n),
    treat = rbinom(n, 1, 0.5),
    bm1 = as.factor(rbinom(n, 1, 0.70)),
    bm2 = as.factor(rbinom(n, 1, 0.50)),
    age = round(rnorm(n, 55, 10)),
    ecog = as.factor(sample(0:1, n, replace = TRUE, prob = c(0.6, 0.4)))
  )
  # No treatment effect (H0): prognostic covariates only
  lp <- with(df, qlogis(0.30) + 0.02 * (age - 55) + 0.3 * (ecog == 1))
  df$progressed <- rbinom(n, 1, plogis(lp))
  df
}
```


# Part I: Per-Subgroup Approximation {#sec-approx}

Instant calculations --- no simulation required.

```{r approx-fn}
#| code-fold: false

p_detect <- function(theta, d_eff, c1, c2) {
  compute_detection_probability_glm(
    theta = theta, d_eff = d_eff, c1 = c1, c2 = c2,
    effect_scale = "ratio", method = "monte_carlo"
  )
}
```


## Approximation Grid

```{r approx-grid}
#| tbl-cap: "Per-subgroup P1(theta=1) across (c1,c2) for n(H)=150, p=0.30"

# Representative subgroup: 30% prevalence at N=500 -> n_H=150
n_H_ref <- 150
p_event <- 0.30
d_eff_ref <- d_eff_binary(n_H_ref, p_event)

c1_seq <- c(1.00, 1.10, 1.15, 1.20, 1.25, 1.30, 1.40, 1.50,
            1.60, 1.70, 1.80)
c2_seq <- c(0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20, 1.25,
            1.30, 1.40, 1.50)

approx_null <- expand.grid(c1 = c1_seq, c2 = c2_seq,
                            stringsAsFactors = FALSE)
approx_null <- approx_null[approx_null$c2 <= approx_null$c1, ]

approx_null$P1 <- mapply(function(c1v, c2v) {
  round(p_detect(1.0, d_eff_ref, c1v, c2v), 4)
}, approx_null$c1, approx_null$c2)

cat(sprintf("Reference: n(H)=%d, p=%.2f, d_eff=%.1f\n",
            n_H_ref, p_event, d_eff_ref))
cat(sprintf("Grid: %d (c1,c2) pairs\n\n", nrow(approx_null)))

gt(approx_null[approx_null$P1 <= 0.20, ]) |>
  tab_header(
    title = "Per-Subgroup P1(theta=1.0) by Threshold Pair",
    subtitle = sprintf("n(H)=%d, p=%.2f, d_eff=%.1f (showing P1 <= 20%%)",
                        n_H_ref, p_event, d_eff_ref)
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = P1, rows = P1 <= 0.10)) |>
  tab_style(style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(columns = P1,
      rows = P1 > 0.10 & P1 <= 0.15))
```


## Approximation Heatmap

```{r approx-heatmap, fig.width=9, fig.height=7}
#| fig-cap: "Per-subgroup P1(theta=1) heatmap.  The procedure-level FPR will be higher by a factor of L_eff."

ggplot(approx_null, aes(x = factor(c1), y = factor(c2), fill = P1)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.1f%%", 100 * P1)),
            size = 3, colour = "black") +
  scale_fill_gradient2(
    low = "#E8F5E9", mid = "#FFF9C4", high = "#E53935",
    midpoint = 0.10, name = "P1"
  ) +
  labs(x = expression(c[1] ~ "(screening)"),
       y = expression(c[2] ~ "(consistency)"),
       title = "Per-Subgroup Approximation: P1(theta=1)",
       subtitle = sprintf("Binary, n(H)=%d, p=%.2f, d_eff=%.1f",
                           n_H_ref, p_event, d_eff_ref)) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        axis.text = element_text(size = 10))
```


# Part II: $L_{\text{eff}}$ Calibration {#sec-leff}

## Multi-N Null Simulations

Run ForestSearch under $H_0$ at the paper defaults (c1=1.25,
c2=1.0) across multiple sample sizes to calibrate
$L_{\text{eff}}$.

```{r leff-sims}
N_cal_values <- c(200, 300, 500, 700, 1000)
n_sims_cal <- min(nsims_null, 200L)
n_min_val <- 60
pcons <- 0.90

cat("Running L_eff calibration simulations...\n")
t0 <- proc.time()[3]

cal_results <- list()
for (N_val in N_cal_values) {
  t_N <- proc.time()[3]

  res_N <- foreach(
    s = seq_len(n_sims_cal),
    .combine = c, .errorhandling = "pass",
    .options.future = list(packages = c("forestsearch", "data.table"))
  ) %dofuture% {
    seed <- sim_seed + (N_val * 1000L) + s
    df_null <- sim_binary_null(N_val, seed)
    fs <- tryCatch(
      forestsearch(
        df.analysis = df_null,
        confounders.name = c("bm1", "bm2", "age", "ecog"),
        outcome.name = "progressed", event.name = "progressed",
        treat.name = "treat", id.name = "id",
        outcome_type = "binary", effect_measure = "OR",
        adverse_outcome = TRUE,
        hr.threshold = 1.25, hr.consistency = 1.0,
        pconsistency.threshold = pcons, fs.splits = 400,
        n.min = n_min_val, maxk = 2,
        use_lasso = TRUE, use_grf = TRUE, use_twostage = TRUE,
        is.RCT = TRUE, seedit = seed, details = FALSE, plot.sg = FALSE,
        parallel_args = list(plan = "sequential", workers = 1,
                             show_message = FALSE)
      ),
      error = function(e) NULL
    )
    found <- !is.null(fs) && !is.null(fs$sg.harm) && length(fs$sg.harm) > 0
    as.integer(found)
  }

  res_clean <- unlist(res_N[sapply(res_N, is.numeric)])
  fpr_N <- mean(res_clean, na.rm = TRUE)
  cal_results[[as.character(N_val)]] <- data.frame(
    N = N_val, sim_fpr = fpr_N, n_sims = length(res_clean)
  )

  cat(sprintf("  N = %4d: FPR = %.3f (%d/%d) | %.1f sec\n",
      N_val, fpr_N, sum(res_clean), length(res_clean),
      proc.time()[3] - t_N))
}

cal_df <- do.call(rbind, cal_results)
timings$leff_cal <- proc.time()[3] - t0
cat(sprintf("L_eff calibration: %.1f min\n", timings$leff_cal / 60))
```


## Fit $L_{\text{eff}}$

```{r fit-leff}
#| fig-height: 5
#| fig-cap: "Implied L_eff with power-law fit for the binary DGM (4 confounders)."

p_harm <- 0.30  # 30% prevalence of harm subgroup

cal_df$n_H <- round(p_harm * cal_df$N)
cal_df$d_eff <- d_eff_binary(cal_df$n_H, p_event)
cal_df$P1 <- sapply(cal_df$d_eff, function(d) {
  p_detect(1.0, d, 1.25, 1.0)
})

# Clamp sim_fpr away from 0 and 1
cal_df$sim_fpr <- pmax(cal_df$sim_fpr, 0.5 / n_sims_cal)
cal_df$sim_fpr <- pmin(cal_df$sim_fpr, 1 - 0.5 / n_sims_cal)

cal_df$L_eff <- log(1 - cal_df$sim_fpr) / log(1 - cal_df$P1)

cat("=== Implied L_eff ===\n")
print(cal_df[, c("N", "n_H", "d_eff", "P1", "sim_fpr", "L_eff")])

# Fit power law
cal_valid <- cal_df[is.finite(cal_df$L_eff) & cal_df$L_eff > 0, ]

if (nrow(cal_valid) >= 2) {
  leff_cal <- calibrate_L_eff(
    N = cal_valid$N, P1 = cal_valid$P1,
    sim_fpr = cal_valid$sim_fpr, n_min = n_min_val
  )
  print(leff_cal)

  C_fit <- leff_cal$C
  alpha_fit <- leff_cal$alpha

  N_pred <- seq(100, 1200, by = 5)
  pred_curve <- data.frame(
    N = N_pred, L_eff = C_fit * (N_pred / n_min_val)^alpha_fit
  )

  ggplot() +
    geom_point(data = cal_valid, aes(x = N, y = L_eff),
               size = 4, colour = "firebrick") +
    geom_line(data = pred_curve, aes(x = N, y = L_eff),
              colour = "steelblue", linewidth = 1) +
    labs(x = "Total Sample Size N",
         y = expression(L[eff]),
         title = expression(L[eff] * " Calibration: Binary DGM"),
         subtitle = sprintf("L_eff = %.3f * (N/%d)^%.3f   (4 confounders)",
                             C_fit, n_min_val, alpha_fit)) +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(face = "bold"))
} else {
  cat("Insufficient calibration points. Using L_eff = 1.\n")
  C_fit <- 1; alpha_fit <- 0
}
```


# Part III: Null Threshold Grid {#sec-grid}

Sweep a dense $(c_1, c_2)$ grid under $H_0$ at multiple
sample sizes.

## Grid Design

```{r grid-config}
#| code-fold: false

c1_grid <- c(1.00, 1.10, 1.15, 1.20, 1.25, 1.30, 1.40,
             1.50, 1.60, 1.70, 1.80)
c2_grid <- c(0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20,
             1.25, 1.30, 1.40, 1.50)

grid_pairs <- expand.grid(c1 = c1_grid, c2 = c2_grid,
                           stringsAsFactors = FALSE)
grid_pairs <- grid_pairs[grid_pairs$c2 <= grid_pairs$c1, ]

N_grid_values <- c(500, 700, 1000)
nsims_grid <- nsims_null

cat(sprintf("Null grid: %d (c1,c2) pairs x %d N values x %d sims = %d tasks\n",
    nrow(grid_pairs), length(N_grid_values), nsims_grid,
    nrow(grid_pairs) * length(N_grid_values) * nsims_grid))
```


## Run Grid Simulations

```{r run-grid}
t0 <- proc.time()[3]

# Build task list across all N values
grid_tasks <- do.call(rbind, lapply(N_grid_values, function(N_val) {
  do.call(rbind, lapply(seq_len(nrow(grid_pairs)), function(g) {
    data.frame(
      N = N_val, sim = seq_len(nsims_grid),
      c1 = grid_pairs$c1[g], c2 = grid_pairs$c2[g],
      stringsAsFactors = FALSE
    )
  }))
}))

cat(sprintf("Running %d null grid tasks...\n", nrow(grid_tasks)))

grid_results <- foreach(
  i = seq_len(nrow(grid_tasks)),
  .combine = rbind, .errorhandling = "remove",
  .options.future = list(packages = c("forestsearch", "data.table"))
) %dofuture% {
  N_val <- grid_tasks$N[i]
  sim_id <- grid_tasks$sim[i]
  c1_val <- grid_tasks$c1[i]
  c2_val <- grid_tasks$c2[i]

  seed <- sim_seed + (N_val * 1000L) + sim_id
  df_null <- sim_binary_null(N_val, seed)

  fs <- tryCatch(
    forestsearch(
      df.analysis = df_null,
      confounders.name = c("bm1", "bm2", "age", "ecog"),
      outcome.name = "progressed", event.name = "progressed",
      treat.name = "treat", id.name = "id",
      outcome_type = "binary", effect_measure = "OR",
      adverse_outcome = TRUE,
      hr.threshold = c1_val, hr.consistency = c2_val,
      pconsistency.threshold = pcons, fs.splits = 400,
      n.min = n_min_val, maxk = 2,
      use_lasso = TRUE, use_grf = TRUE, use_twostage = TRUE,
      is.RCT = TRUE, seedit = seed, details = FALSE, plot.sg = FALSE,
      parallel_args = list(plan = "sequential", workers = 1,
                           show_message = FALSE)
    ),
    error = function(e) NULL
  )

  found <- !is.null(fs) && !is.null(fs$sg.harm) && length(fs$sg.harm) > 0
  data.table(N = N_val, sim = sim_id, c1 = c1_val, c2 = c2_val,
             any_found = as.integer(found))
}

timings$null_grid <- proc.time()[3] - t0
cat(sprintf("Null grid: %.1f min | %d results\n",
    timings$null_grid / 60, nrow(grid_results)))
```


# Part IV: Results {#sec-results}

## FPR by Threshold Pair and N

```{r grid-fpr}
grid_dt <- data.table(grid_results)
grid_fpr <- grid_dt[, .(
  FPR = mean(any_found, na.rm = TRUE),
  n_found = sum(any_found, na.rm = TRUE),
  n_sims = .N
), by = .(N, c1, c2)]

# Add approximations
grid_fpr[, n_H := round(p_harm * N)]
grid_fpr[, d_eff := d_eff_binary(n_H, p_event)]
grid_fpr[, P1 := mapply(function(d, c1v, c2v) {
  p_detect(1.0, d, c1v, c2v)
}, d_eff, c1, c2)]

# L_eff corrected
grid_fpr[, L_eff_N := C_fit * (N / n_min_val)^alpha_fit]
grid_fpr[, FPR_corr := 1 - (1 - P1)^L_eff_N]
```


## FPR Heatmaps by N

```{r grid-heatmaps, fig.width=10, fig.height=12}
#| fig-cap: "Simulated FPR heatmaps by (c1, c2) at each sample size.  Green region: FPR <= 10%."

heatmap_list <- lapply(N_grid_values, function(N_val) {
  sub <- grid_fpr[N == N_val]
  ggplot(sub, aes(x = factor(c1), y = factor(c2), fill = FPR)) +
    geom_tile(colour = "white", linewidth = 0.5) +
    geom_text(aes(label = sprintf("%.0f%%", 100 * FPR)),
              size = 2.8, colour = "black") +
    scale_fill_gradient2(
      low = "#E8F5E9", mid = "#FFF9C4", high = "#E53935",
      midpoint = 0.10,
      limits = c(0, max(grid_fpr$FPR, 0.30)),
      name = "FPR"
    ) +
    labs(x = expression(c[1]),
         y = expression(c[2]),
         title = sprintf("N = %d", N_val)) +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold"),
          axis.text = element_text(size = 9))
})

wrap_plots(heatmap_list, ncol = 1) +
  plot_annotation(
    title = "Simulated FPR Under H0: Binary DGM",
    subtitle = sprintf("%d sims per pair, n.min=%d, pcons=%.2f",
                        nsims_grid, n_min_val, pcons))
```


## FPR Table (Acceptable Region)

```{r grid-fpr-table}
#| tbl-cap: "Threshold pairs with simulated FPR <= 15%.  Green: FPR <= 10%.  Yellow: 10-12%.  Orange: 12-15%."

acceptable <- grid_fpr[FPR <= 0.15]
setorder(acceptable, N, FPR)

if (nrow(acceptable) > 0) {
  gt(acceptable[, .(N, c1, c2,
    `Sim FPR` = round(FPR, 3),
    `P1` = round(P1, 3),
    `FPR (corr)` = round(FPR_corr, 3),
    `Sim/Corr` = round(FPR / pmax(FPR_corr, 0.001), 2))]) |>
    tab_header(
      title = md("**Threshold Configurations with FPR <= 15%**"),
      subtitle = sprintf("Binary DGM, n.min=%d, pcons=%.2f",
                          n_min_val, pcons)
    ) |>
    tab_style(style = cell_fill(color = "#E8F5E9"),
      locations = cells_body(columns = `Sim FPR`,
        rows = `Sim FPR` <= 0.10)) |>
    tab_style(style = cell_fill(color = "#FFF9C4"),
      locations = cells_body(columns = `Sim FPR`,
        rows = `Sim FPR` > 0.10 & `Sim FPR` <= 0.12)) |>
    tab_style(style = cell_fill(color = "#FFE0B2"),
      locations = cells_body(columns = `Sim FPR`,
        rows = `Sim FPR` > 0.12 & `Sim FPR` <= 0.15)) |>
    tab_row_group(label = "N = 1000", rows = N == 1000) |>
    tab_row_group(label = "N = 700", rows = N == 700) |>
    tab_row_group(label = "N = 500", rows = N == 500)
} else {
  cat("No configurations achieved FPR <= 15% in this grid.\n")
}
```


## FPR Sensitivity to $c_2$ (Fixing $c_1$)

```{r c2-profiles, fig.width=10, fig.height=8}
#| fig-cap: "FPR vs c2 for each fixed c1, by sample size.  The 10-12% target band is highlighted."

ggplot(grid_fpr, aes(x = c2, y = FPR,
                      colour = factor(c1), group = factor(c1))) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.10, linetype = "dashed", colour = "darkgreen") +
  geom_hline(yintercept = 0.12, linetype = "dashed", colour = "orange") +
  annotate("rect", xmin = -Inf, xmax = Inf,
           ymin = 0.10, ymax = 0.12,
           alpha = 0.12, fill = "gold") +
  facet_wrap(~ paste("N =", N), ncol = 1) +
  scale_colour_brewer(palette = "Paired") +
  labs(x = expression(c[2] ~ "(consistency threshold)"),
       y = "Simulated FPR",
       title = "FPR Sensitivity to Consistency Threshold",
       colour = expression(c[1])) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


## FPR Sensitivity to $c_1$ (Fixing $c_2$)

```{r c1-profiles, fig.width=10, fig.height=8}
#| fig-cap: "FPR vs c1 for each fixed c2, by sample size."

ggplot(grid_fpr, aes(x = c1, y = FPR,
                      colour = factor(c2), group = factor(c2))) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.10, linetype = "dashed", colour = "darkgreen") +
  geom_hline(yintercept = 0.12, linetype = "dashed", colour = "orange") +
  annotate("rect", xmin = -Inf, xmax = Inf,
           ymin = 0.10, ymax = 0.12,
           alpha = 0.12, fill = "gold") +
  facet_wrap(~ paste("N =", N), ncol = 1) +
  scale_colour_brewer(palette = "Set3") +
  labs(x = expression(c[1] ~ "(screening threshold)"),
       y = "Simulated FPR",
       title = "FPR Sensitivity to Screening Threshold",
       colour = expression(c[2])) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


## $L_{\text{eff}}$ Correction Accuracy Across the Grid

```{r corr-scatter, fig.width=9, fig.height=6}
#| fig-cap: "Corrected (triangles) vs uncorrected (circles) against simulation across the full grid."

ggplot(grid_fpr, aes(x = FPR)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              colour = "grey50") +
  geom_point(aes(y = P1, colour = "Uncorrected P1"),
             size = 2, alpha = 0.5) +
  geom_point(aes(y = FPR_corr, colour = "L_eff Corrected"),
             size = 2, alpha = 0.5, shape = 17) +
  scale_colour_manual(values = c("Uncorrected P1" = "firebrick",
                                  "L_eff Corrected" = "steelblue")) +
  facet_wrap(~ paste("N =", N)) +
  coord_equal(xlim = c(0, max(grid_fpr$FPR, 0.3)),
              ylim = c(0, max(grid_fpr$P1, 0.3))) +
  labs(x = "Simulated FPR", y = "Approximated FPR",
       title = "Approximation Accuracy Across Full Grid",
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Comparison to Survival

```{r compare-survival}
cat("=== Cross-DGM L_eff Comparison ===\n\n")
cat(sprintf("Binary DGM (4 confounders, maxk=2):\n"))
cat(sprintf("  L_eff = %.3f * (N/%d)^%.3f\n", C_fit, n_min_val, alpha_fit))
for (N_val in c(500, 700, 1000)) {
  L <- C_fit * (N_val / n_min_val)^alpha_fit
  cat(sprintf("  N = %4d: L_eff = %.1f\n", N_val, L))
}
cat("\nGBSG Survival DGM (7 confounders, maxk=2):\n")
cat("  See cox_vs_glm_approximation.qmd for comparison.\n")
cat("  Expected: larger L_eff (more confounders -> more candidates)\n")
```


# Summary {#sec-summary}

```{r summary}
cat("=== Null Threshold Search: Binary DGM ===\n\n")
cat(sprintf("DGM: Binary, p=%.2f, 4 confounders, n.min=%d, pcons=%.2f\n",
            p_event, n_min_val, pcons))
cat(sprintf("L_eff model: %.3f * (N/%d)^%.3f\n\n",
            C_fit, n_min_val, alpha_fit))

for (N_val in N_grid_values) {
  cat(sprintf("--- N = %d (L_eff = %.1f) ---\n", N_val,
              C_fit * (N_val / n_min_val)^alpha_fit))

  sub <- grid_fpr[N == N_val]

  cat("  FPR <= 10%:\n")
  ok10 <- sub[FPR <= 0.10]
  if (nrow(ok10) > 0) {
    setorder(ok10, -FPR)
    for (r in seq_len(min(8, nrow(ok10)))) {
      cat(sprintf("    c1=%.2f, c2=%.2f: %.1f%%\n",
                  ok10$c1[r], ok10$c2[r], 100 * ok10$FPR[r]))
    }
    if (nrow(ok10) > 8) cat(sprintf("    ... and %d more\n", nrow(ok10) - 8))
  } else {
    cat("    None\n")
  }

  cat("  FPR in 10-12%:\n")
  ok12 <- sub[FPR > 0.10 & FPR <= 0.12]
  if (nrow(ok12) > 0) {
    setorder(ok12, -FPR)
    for (r in seq_len(min(8, nrow(ok12)))) {
      cat(sprintf("    c1=%.2f, c2=%.2f: %.1f%%\n",
                  ok12$c1[r], ok12$c2[r], 100 * ok12$FPR[r]))
    }
  } else {
    cat("    None\n")
  }
  cat("\n")
}
```


# Timing and Session Info {#sec-session}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Stage = c("L_eff Calibration",
            "Null Grid Sweep",
            "Total"),
  Seconds = c(timings$leff_cal, timings$null_grid, timings$total),
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(
    title = "Computational Timing Summary",
    subtitle = sprintf("%d (c1,c2) pairs x %d N values x %d sims | %d workers",
      nrow(grid_pairs), length(N_grid_values), nsims_grid, n_workers)
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)",
             Minutes = "Time (min)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total")
  )
```

```{r session-info}
plan(sequential)
sessionInfo()
```
