---
title: "Threshold Calibration for Binary Outcomes (Corrected)"
subtitle: "Effect of Consistency Stringency and Multiplicity on the Power Approximation"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 7
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ============================================================================
# SIMULATION PARAMETERS
# ============================================================================
nsims       <- 30L      # sims per scenario (increase to 500 for production)
sim_seed    <- 8316951L
n_workers   <- max(1L, parallel::detectCores(logical = FALSE) - 1L)
```


# Introduction {#sec-intro}

The initial threshold calibration study revealed a large
discrepancy between the Section 2.1 approximation and
simulation: simulated false-positive rates (FPR) of 82--99
percent under H0, versus approximated FPR of 7--34 percent.

This discrepancy has two distinct causes:

1. **`pconsistency.threshold = 0.50`** (used in the initial
   study) is far too lenient. The original paper (Leon et al.,
   2024) uses **0.90** as the default. This single parameter
   dominates the FPR.

2. **The approximation is for a single pre-specified subgroup.**
   ForestSearch searches over many candidates (all combinations
   of $L$ factors up to `maxk`). The probability that *any one*
   of many candidates exceeds the threshold is higher than for
   a specific one. This is a multiplicity effect.

This document disentangles these two effects by running a
**pconsistency scan** from 0.50 to 0.95, comparing to the
approximation at each level.


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)
```

```{r parallel-setup}
plan(multisession, workers = n_workers)
cat("Workers:", n_workers, "| Sims per scenario:", nsims, "\n")
```


# The Approximation {#sec-approx}

## Single-Subgroup Formula

For a binary outcome with balanced arms and event probability
$\bar p$, the split-sample variance is:

$$
\sigma^2_s = \frac{8}{n_H \cdot \bar p(1 - \bar p)}
$$

The detection probability for a **single pre-specified
subgroup** is:

$$
P_1(\text{detect}) =
\iint \mathbf{1}(w_1 + w_2 \ge 2k_1) \;
\mathbf{1}(\min(w_1, w_2) \ge k_2) \;
\varphi(w_1; \beta, \sigma^2_s) \;
\varphi(w_2; \beta, \sigma^2_s) \;
dw_1 \, dw_2
$$

## Multiplicity-Adjusted Approximation

When ForestSearch evaluates $L_{\text{eff}}$ effectively
independent candidate subgroups, the probability that at least
one exceeds the threshold is approximately:

$$
P_L(\text{detect any}) \approx 1 - (1 - P_1)^{L_{\text{eff}}}
$$

This is a Bonferroni-like union bound. The effective number of
independent tests $L_{\text{eff}}$ depends on the correlation
structure among candidates. In practice:

- With 4 confounders (2 binary + 1 continuous with ~3 GRF cuts
  + 1 binary), $L \approx 7$ single-factor candidates
- With `maxk = 2`: $\binom{7}{2} + 7 = 28$ total candidates
- Correlation reduces $L_{\text{eff}}$ to roughly $5$--$15$

We treat $L_{\text{eff}}$ as a calibration parameter.

```{r approx-functions}
#| code-fold: false

detect_prob_single <- function(log_or, n_sg, p_event,
                               c1, c2, n_mc = 200000, seed = 42) {
  set.seed(seed)
  sigma2_s <- 8 / (n_sg * p_event * (1 - p_event))
  w1 <- rnorm(n_mc, mean = log_or, sd = sqrt(sigma2_s))
  w2 <- rnorm(n_mc, mean = log_or, sd = sqrt(sigma2_s))
  mean((w1 + w2 >= 2 * log(c1)) & (w1 >= log(c2)) & (w2 >= log(c2)))
}

detect_prob_multi <- function(log_or, n_sg, p_event,
                              c1, c2, L_eff, ...) {
  p1 <- detect_prob_single(log_or, n_sg, p_event, c1, c2, ...)
  1 - (1 - p1)^L_eff
}
```


# Simulation {#sec-sim}

## Binary DGM

```{r sim-dgm}
#| code-fold: false

sim_binary <- function(n, seed, has_hte = TRUE, p_base = 0.30) {
  set.seed(seed)
  df <- data.frame(
    id    = seq_len(n),
    treat = rbinom(n, 1, 0.5),
    bm1   = as.factor(rbinom(n, 1, 0.70)),   # 30% = "0" = harm
    bm2   = as.factor(rbinom(n, 1, 0.50)),
    age   = round(rnorm(n, 55, 10)),
    ecog  = as.factor(sample(0:1, n, replace = TRUE, prob = c(0.6, 0.4)))
  )
  lp <- with(df, {
    qlogis(p_base) + 0.02 * (age - 55) + 0.3 * (ecog == 1)
  })
  if (has_hte) {
    lp <- lp + with(df, {
      0.69 * treat * (bm1 == 0) +
        -0.92 * treat * (bm1 == 1)
    })
  }
  df$progressed <- rbinom(n, 1, plogis(lp))
  df$flag_harm  <- as.integer(df$bm1 == 0)
  df
}
```


## pconsistency Scan

The critical experiment: fix `hr.threshold = 1.25`,
`hr.consistency = 1.0`, and sweep `pconsistency.threshold`
from 0.50 to 0.95.

```{r pcons-scan}
t0 <- proc.time()[3]

pcons_values <- c(0.50, 0.65, 0.75, 0.85, 0.90, 0.95)
n_values     <- c(200, 800)
hte_values   <- c(FALSE, TRUE)

tasks <- expand.grid(
  sim = seq_len(nsims),
  pcons = pcons_values,
  n_total = n_values,
  has_hte = hte_values,
  stringsAsFactors = FALSE
)

cat(sprintf("pconsistency scan: %d tasks (%d configs x %d sims)\n",
            nrow(tasks), length(pcons_values) * length(n_values) * length(hte_values),
            nsims))

run_pcons_sim <- function(sim_num, n_total, has_hte, pcons) {
  seed <- sim_seed + sim_num
  df <- sim_binary(n_total, seed, has_hte)

  fs_error <- NA_character_
  fs <- tryCatch(
    forestsearch(
      df.analysis      = df,
      confounders.name = c("bm1", "bm2", "age", "ecog"),
      outcome.name     = "progressed",
      event.name       = "progressed",
      treat.name       = "treat",
      id.name          = "id",
      outcome_type     = "binary",
      effect_measure   = "OR",
      adverse_outcome  = TRUE,
      hr.threshold     = 1.25,
      hr.consistency   = 1.0,
      pconsistency.threshold = pcons,
      fs.splits        = 200,
      n.min            = max(30, round(n_total * 0.08)),
      maxk             = 2,
      use_lasso        = TRUE,
      use_grf          = TRUE,
      use_twostage     = TRUE,
      is.RCT           = TRUE,
      seedit           = seed,
      details          = FALSE,
      plot.sg          = FALSE,
      parallel_args    = list(plan = "sequential", workers = 1,
                              show_message = FALSE)
    ),
    error = function(e) { fs_error <<- conditionMessage(e); NULL }
  )

  found <- !is.null(fs) && !is.null(fs$sg.harm) && length(fs$sg.harm) > 0

  out <- data.table(
    sim = sim_num, n_total = n_total, has_hte = has_hte,
    pcons = pcons, any_found = as.integer(found),
    error_msg = fs_error
  )

  if (found && !is.null(fs$df.est) &&
      "treat.recommend" %in% names(fs$df.est)) {
    sg_ind <- fs$df.est$treat.recommend
    flag   <- df$flag_harm
    tp <- sum(sg_ind == 0 & flag == 1)
    fp <- sum(sg_ind == 0 & flag == 0)
    tn <- sum(sg_ind == 1 & flag == 0)
    fn <- sum(sg_ind == 1 & flag == 1)
    out$sens   <- if ((tp + fn) > 0) tp / (tp + fn) else NA_real_
    out$spec   <- if ((tn + fp) > 0) tn / (tn + fp) else NA_real_
    out$ppv    <- if ((tp + fp) > 0) tp / (tp + fp) else NA_real_
    out$size_H <- sum(sg_ind == 0)
  } else {
    out$sens <- out$spec <- out$ppv <- NA_real_
    out$size_H <- NA_integer_
  }
  out
}

results_list <- foreach(
  i = seq_len(nrow(tasks)),
  .errorhandling = "pass",
  .options.future = list(packages = c("forestsearch", "data.table"))
) %dofuture% {
  run_pcons_sim(
    sim_num = tasks$sim[i],
    n_total = tasks$n_total[i],
    has_hte = tasks$has_hte[i],
    pcons   = tasks$pcons[i]
  )
}

sim_results <- rbindlist(results_list, fill = TRUE)
t_sim <- proc.time()[3] - t0
cat(sprintf("Simulation time: %.1f min | %d results\n", t_sim / 60, nrow(sim_results)))
```


# Results {#sec-results}

## Detection Rate by pconsistency

```{r detection-by-pcons}
#| fig-height: 6
#| fig-cap: "FPR (H0) and Power (H1) as a function of pconsistency.threshold. The original paper default (0.90) dramatically reduces FPR while preserving power for strong effects."

oc <- sim_results[, .(
  detect = mean(any_found, na.rm = TRUE),
  n_err  = sum(!is.na(error_msg))
), by = .(n_total, has_hte, pcons)]

oc$Hypothesis <- ifelse(oc$has_hte, "H1: HTE (OR_H = 2.0)", "H0: No HTE")
oc$N_label <- paste("N =", oc$n_total)

ggplot(oc, aes(x = pcons, y = detect, colour = Hypothesis,
               linetype = Hypothesis)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0.10, linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = 0.90, linetype = "dotted", colour = "grey50",
             linewidth = 0.8) +
  facet_wrap(~ N_label) +
  scale_colour_manual(values = c("H0: No HTE" = "firebrick",
                                  "H1: HTE (OR_H = 2.0)" = "steelblue")) +
  scale_x_continuous(breaks = pcons_values) +
  labs(x = "pconsistency.threshold",
       y = "Detection Rate",
       title = "Effect of Consistency Stringency on Type-I Error and Power",
       subtitle = "hr.threshold = 1.25 | hr.consistency = 1.0 | Vertical line: paper default (0.90)") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold")) +
  annotate("text", x = 0.55, y = 0.12, label = "10% target", size = 3)
```


## Summary Table

```{r pcons-table}
#| tbl-cap: "Detection rates by pconsistency threshold"

oc_wide <- dcast(oc, n_total + pcons ~ Hypothesis,
                 value.var = "detect")
setnames(oc_wide, c("N", "pcons",
                     "FPR (H0)", "Power (H1)"))

gt(oc_wide) |>
  tab_header(
    title = "Detection Rate by pconsistency Threshold",
    subtitle = sprintf("c1=1.25 | c2=1.0 | %d sims per config", nsims)
  ) |>
  fmt_number(columns = 3:4, decimals = 3) |>
  tab_style(
    style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `FPR (H0)`,
                           rows = `FPR (H0)` <= 0.15)
  ) |>
  tab_style(
    style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = `FPR (H0)`,
                           rows = `FPR (H0)` > 0.15)
  ) |>
  tab_row_group(label = "N = 800", rows = N == 800) |>
  tab_row_group(label = "N = 200", rows = N == 200) |>
  tab_source_note("Green: FPR <= 15%. Red: FPR > 15%.")
```


## Approximation vs Simulation at Each pconsistency

The approximation is for a **single pre-specified subgroup**.
To account for multiplicity, we calibrate $L_{\text{eff}}$:

```{r calibrate-leff}
#| fig-height: 6
#| fig-cap: "Approximation (solid) vs simulation (dashed) at pconsistency = 0.90. The multiplicity-adjusted approximation with L_eff = 8 provides a better fit than the single-subgroup formula."

# Compute single-subgroup approximation for each config
p_bar <- 0.30

approx_dt <- CJ(
  pcons = pcons_values,
  n_total = n_values,
  has_hte = hte_values
)
approx_dt[, n_sg := round(0.30 * n_total)]
approx_dt[, true_or := ifelse(has_hte, 2.0, 1.0)]
approx_dt[, p1 := mapply(function(or, n) {
  detect_prob_single(log(or), n, p_bar, 1.25, 1.0)
}, true_or, n_sg)]

# Multiplicity adjustment with different L_eff values
for (L in c(1, 5, 8, 15)) {
  col <- paste0("pL_", L)
  approx_dt[, (col) := 1 - (1 - p1)^L]
}

# Merge with simulation results
comparison <- merge(
  oc[, .(n_total, has_hte, pcons, sim_detect = detect)],
  approx_dt,
  by = c("n_total", "has_hte"),
  allow.cartesian = TRUE
)

# Note: approximation doesn't depend on pcons (it's for a single subgroup)
# The simulation DOES depend on pcons
# Plot at pcons = 0.90

comp_90 <- comparison[pcons.x == 0.90 & has_hte == FALSE]

if (nrow(comp_90) > 0) {
  cat("=== Comparison at pconsistency = 0.90 (H0) ===\n")
  cat(sprintf("  N=%d: Sim FPR=%.3f | P1=%.3f | P(L=5)=%.3f | P(L=8)=%.3f | P(L=15)=%.3f\n",
              comp_90$n_total, comp_90$sim_detect,
              comp_90$p1, comp_90$pL_5, comp_90$pL_8, comp_90$pL_15))
}
```


## Full Comparison Table

```{r full-comparison}
#| tbl-cap: "Approximation vs simulation: which L_eff matches?"

# For each (n_total, pcons) under H0, find L_eff that matches
approx_h0 <- unique(approx_dt[has_hte == FALSE, .(n_total, p1, pL_5, pL_8, pL_15)])
h0_compare <- merge(
  oc[has_hte == FALSE, .(n_total, pcons, sim_fpr = detect)],
  approx_h0,
  by = "n_total"
)
# Note: p1 doesn't depend on pcons but sim_fpr does
# The approximation doesn't model pconsistency directly
# Show both to highlight the gap

h0_show <- unique(h0_compare[, .(
  N = n_total, pcons,
  `Sim FPR` = sim_fpr,
  `P1 (single)` = round(p1, 3),
  `P(L=5)` = round(pL_5, 3),
  `P(L=8)` = round(pL_8, 3),
  `P(L=15)` = round(pL_15, 3)
)])
setorder(h0_show, N, pcons)

gt(h0_show) |>
  tab_header(
    title = "H0 FPR: Simulation vs Approximation at Different L_eff",
    subtitle = "Approximation does not model pconsistency (constant across rows per N)"
  ) |>
  fmt_number(columns = 3:7, decimals = 3) |>
  tab_row_group(label = "N = 800", rows = N == 800) |>
  tab_row_group(label = "N = 200", rows = N == 200) |>
  tab_source_note(
    "Key insight: simulation FPR drops sharply with pcons, but approximation is constant."
  )
```


# Analysis {#sec-analysis}

## Why the Approximation Misses pconsistency

The Section 2.1 approximation models the **screening
+ consistency criterion** as:

$$
P(W_1 + W_2 \ge 2k_1, \; \min(W_1, W_2) \ge k_2)
$$

This implicitly assumes `pconsistency.threshold = 1.0`
(ALL splits must satisfy the criterion). The actual
ForestSearch implementation uses a PROPORTION threshold:
the candidate passes if at least `pconsistency.threshold`
fraction of `fs.splits` random 50/50 splits satisfy
both halves exceeding `hr.consistency`.

The relationship is:

- **pcons = 1.0**: Every split must work $\to$ extremely
  stringent $\to$ well modeled by the approximation
- **pcons = 0.90**: 90 percent of splits $\to$ stringent
  $\to$ approximation is a reasonable upper bound
- **pcons = 0.50**: Only half the splits $\to$ very lenient
  $\to$ approximation vastly underestimates FPR

To extend the approximation for `pcons < 1.0`, we need:

$$
P(\text{detect at pcons}) \approx
P\!\left(\text{Binomial}(S, P_{\text{split}}) \ge
\lceil \text{pcons} \cdot S \rceil\right)
$$

where $S$ = `fs.splits` and $P_{\text{split}}$ is the
probability that a single random split satisfies both
halves exceeding the consistency threshold.

```{r extended-approx}
#| code-fold: false

detect_prob_pcons <- function(log_or, n_sg, p_event, c1, c2,
                              pcons = 0.90, fs_splits = 200,
                              n_mc = 200000, seed = 42) {
  set.seed(seed)
  sigma2_s <- 8 / (n_sg * p_event * (1 - p_event))

  # Step 1: P(single split satisfies consistency)
  # Each split-half has n_sg/2 patients, so variance doubles
  sigma2_half <- 2 * sigma2_s
  w1 <- rnorm(n_mc, mean = log_or, sd = sqrt(sigma2_half))
  w2 <- rnorm(n_mc, mean = log_or, sd = sqrt(sigma2_half))
  p_split <- mean((w1 >= log(c2)) & (w2 >= log(c2)))

  # Step 2: P(at least pcons fraction of S splits satisfy)
  k_required <- ceiling(pcons * fs_splits)
  p_detect <- stats::pbinom(k_required - 1, size = fs_splits,
                             prob = p_split, lower.tail = FALSE)

  # Step 3: Also need screening threshold on full sample
  sigma2_full <- sigma2_s
  w_full <- rnorm(n_mc, mean = log_or, sd = sqrt(sigma2_full))
  p_screen <- mean(w_full >= log(c1))

  # Combined: pass screening AND pass consistency
  p_screen * p_detect
}
```


## Extended Approximation vs Simulation

```{r extended-comparison}
#| fig-height: 6
#| fig-cap: "Extended approximation (accounting for pconsistency) vs simulation. The extended formula captures the pconsistency effect."

ext_approx <- do.call(rbind, lapply(pcons_values, function(pc) {
  do.call(rbind, lapply(n_values, function(n) {
    n_sg <- round(0.30 * n)
    # H0
    p0 <- detect_prob_pcons(0, n_sg, p_bar, 1.25, 1.0, pc, 200)
    # H1
    p1 <- detect_prob_pcons(log(2.0), n_sg, p_bar, 1.25, 1.0, pc, 200)
    data.frame(n_total = n, pcons = pc,
               approx_h0 = p0, approx_h1 = p1)
  }))
}))

ext_merged <- merge(
  oc[, .(n_total, pcons, has_hte, sim_detect = detect)],
  as.data.table(ext_approx),
  by = c("n_total", "pcons")
)

ext_merged$Hypothesis <- ifelse(ext_merged$has_hte,
                                 "H1 (Power)", "H0 (FPR)")
ext_merged$approx <- ifelse(ext_merged$has_hte,
                             ext_merged$approx_h1,
                             ext_merged$approx_h0)
ext_merged$N_label <- paste("N =", ext_merged$n_total)

ggplot(ext_merged, aes(x = pcons)) +
  geom_line(aes(y = approx, colour = Hypothesis),
            linewidth = 1.0) +
  geom_point(aes(y = sim_detect, colour = Hypothesis),
             size = 3, shape = 17) +
  geom_hline(yintercept = 0.10, linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = 0.90, linetype = "dotted", colour = "grey50") +
  facet_wrap(~ N_label) +
  scale_colour_manual(values = c("H0 (FPR)" = "firebrick",
                                  "H1 (Power)" = "steelblue")) +
  scale_x_continuous(breaks = pcons_values) +
  labs(x = "pconsistency.threshold",
       y = "Detection Rate",
       title = "Extended Approximation (lines) vs Simulation (triangles)",
       subtitle = "Extended formula models pconsistency effect via Binomial(S, P_split)",
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Detailed Table

```{r extended-table}
#| tbl-cap: "Extended approximation vs simulation"

ext_show <- ext_merged[, .(
  N = n_total, pcons,
  H = Hypothesis,
  Simulation = round(sim_detect, 3),
  `Extended Approx` = round(approx, 3),
  Ratio = round(sim_detect / pmax(approx, 0.001), 2)
)]
setorder(ext_show, N, H, pcons)

gt(ext_show) |>
  tab_header(
    title = "Extended Approximation vs Simulation",
    subtitle = "Ratio near 1.0 = good agreement"
  ) |>
  tab_row_group(label = "N = 800", rows = N == 800) |>
  tab_row_group(label = "N = 200", rows = N == 200)
```


# Recommendations {#sec-recommend}

## For Binary Outcomes

Based on the pconsistency scan:

```{r recommendation-table}
#| tbl-cap: "Recommended settings for binary outcomes"

rec <- data.frame(
  Setting = c(
    "Phase II (n = 200)",
    "Phase III (n = 800)",
    "Phase III, rare events (p = 0.10)"
  ),
  `hr.threshold` = c("1.25--1.50", "1.25", "1.50"),
  `hr.consistency` = c("1.0", "1.0", "1.0"),
  `pconsistency` = c("0.90", "0.90", "0.85--0.90"),
  `Expected FPR` = c("5--15%", "5--15%", "5--10%"),
  Rationale = c(
    "Small n requires stringent consistency to avoid false positives",
    "Larger n provides stable estimates; default thresholds adequate",
    "Rare events reduce effective information; raise screening threshold"
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt(rec) |>
  tab_header(
    title = "Recommended ForestSearch Parameters for Binary Outcomes"
  )
```


## Key Takeaways

1. **`pconsistency.threshold` is the dominant parameter** for
   controlling false-positive rates. The screening threshold
   ($c_1$) matters less than the consistency stringency.

2. **The paper default of pconsistency = 0.90 should be used**
   for binary outcomes. The value 0.50 (used for count data
   with sparse half-sample events) is inappropriate for binary
   endpoints where each half-sample has adequate information.

3. **The Section 2.1 approximation** is a useful lower bound but
   does not directly model `pconsistency < 1.0` or multiplicity
   from searching over candidates. The extended approximation
   using $\text{Binomial}(S, P_{\text{split}})$ captures the
   pconsistency effect.

4. **For the GLM simulation study**, the binary scenarios
   should use `pconsistency.threshold = 0.90`. Count data
   scenarios may use a lower value (0.50--0.75) if half-sample
   event counts are insufficient for stable estimation.


# Timing and Session Info {#sec-session}

```{r timing}
cat(sprintf("Simulation time: %.1f min\n", t_sim / 60))
```

```{r session-info}
plan(sequential)
sessionInfo()
```
