---
title: "Cox vs GLM Approximation: A Four-Way Comparison"
subtitle: "Validating the GLM Power Approximation via the Poisson-Cox Equivalence"
author: "Larry F. Leon"
date: today
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ── Simulation control ──────────────────────────────────────────
# 100/200 ~ 3m linux
# 1k/2k ~ 16m
# 4k/8k ~ 1hr
nsims_alt  <- 10000L        # H1 sims (use 500+ for production)
nsims_null <- 25000L        # H0 sims (use 500+ for production)
sim_seed   <- 8316951L
n_workers  <- ceiling(0.95 * max(1L,
  parallel::detectCores(logical = FALSE) - 1L))

# ── Output ──────────────────────────────────────────────────────
output_dir <- "_data"

# ── FPR target ─────────────────────────────────────────────────
fpr_target <- 0.12

# ── DGM / outcome parameters ──────────────────────────────────
prop_cens  <- 0.46       # GBSG censoring proportion
p_harm     <- 0.13       # harm subgroup prevalence
n_min_val  <- 60L        # n.min = reference subgroup size for P1
hr_h1      <- 2.0        # target HR for harm subgroup under H1
analysis_time <- 84L     # GBSG analysis time (months)
cens_adjust   <- log(1.5)

# ── ForestSearch parameters ───────────────────────────────────
confounders_base <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")
pcons      <- 0.90       # pconsistency.threshold
fs_splits  <- 400L
maxk_val   <- 2L
n_sample   <- 700L       # sample size for main sims

# ── Approximation display ────────────────────────────────────
# n_H values for the per-subgroup approximation table
n_H_values <- c(60, 80, 91, 100, 120)
# Reference subgroup size for power curves (expected n_H at n_sample)
n_sg_ref   <- round(p_harm * n_sample)   # 0.13 * 700 = 91

# ── L_eff calibration ────────────────────────────────────────
N_cal_values <- c(300, 500, 700)
```


# Introduction {#sec-intro}

Leon et al. (2024, Section 2.1) developed an asymptotic
approximation for the probability that ForestSearch identifies a
subgroup $H$ with true log-hazard ratio $\beta$:

$$
p(c_1, c_2; d, \theta^\dagger(H)) =
P(W_1 + W_2 \ge 2\log c_1, \;\min(W_1, W_2) \ge \log c_2)
\tag{3}
$$

with $W_1, W_2 \stackrel{\text{iid}}{\sim} N(\beta, 8/d)$, where
$d$ is the expected number of events in $H$.

This document validates the extension of (3) to GLM outcomes by
exploiting the **Poisson-Cox equivalence**: on the same survival
data, Poisson regression with log(time) offset estimates the same
quantity as the Cox model, so both approximations should agree.

We compare four calculations:

| Label | Method | Source |
|---|---|---|
| **(A)** | ForestSearch Poisson+offset simulation | GLM pipeline |
| **(B)** | GLM approximation: $d_{\text{eff}} = D$ | GLM theory |
| **(C)** | ForestSearch Cox PH simulation | Established pipeline |
| **(D)** | Cox approximation: `compute_detection_probability()` | Package (Section 2.1) |

Since $d_{\text{eff}}^{\text{Poisson}} = D = d^{\text{Cox}}$
(both equal the number of events), **(B)** and **(D)** should be
identical, and **(A)** and **(C)** should agree closely.


# Theoretical Framework {#sec-theory}

## The Cox Score Statistic Argument (Section 2.1)

For the Cox model $\lambda(t; V) = \lambda_0(t)\exp(\beta V)$, the
score statistic $L_d(\beta)$ based on $d$ events satisfies
$\hat\beta_s \approx (4/d) L_d(0) \sim N(\beta, 4/d)$
(Jennison and Turnbull, 1984).

For a random 50/50 split of the subgroup, the split-half score
statistics $L_{d_1}$ and $L_{d_2}$ with $d_1 \approx d_2 \approx d/2$
give:

$$
\hat\beta^1_s, \hat\beta^2_s \stackrel{\text{iid}}{\sim}
N(\beta, \; 8/d)
$$

since $\text{Var}(\hat\beta_{\text{half}}) = 4/(d/2) = 8/d$.  By
construction $\hat\beta_s \approx (\hat\beta^1_s + \hat\beta^2_s)/2$,
so the screening condition $\hat\beta_s \ge \log(c_1)$ is equivalent
to $\hat\beta^1_s + \hat\beta^2_s \ge 2\log(c_1)$, yielding (3).


## The GLM Analog

For a canonical GLM with treatment as the sole covariate, the MLE
$\hat\beta$ is asymptotically normal by standard likelihood theory
(McCullagh and Nelder, 1989, Ch. 2):

$$
\hat\beta \stackrel{d}{\to}
N(\beta, \; I^{-1}(\beta))
$$

where $I(\beta)$ is the Fisher information.  For balanced arms
($n_1 = n_0 = n_H/2$), the Fisher information takes the form
$I(\beta) = (n_H/4) \cdot \kappa$, giving:

$$
\text{Var}(\hat\beta_{\text{full}}) = \frac{4}{n_H \cdot \kappa}
= \frac{4}{d_{\text{eff}}}
$$

The family-specific $\kappa$ and corresponding $d_{\text{eff}}$:


### Poisson (log link, with offset)

For the Poisson model $\log E[Y_i] = \beta_0 + \beta V_i + \log(t_i)$,
the Fisher information for $\beta$ is:

$$
I(\beta) = \sum_{j=0}^{1} d_j, \quad
d_j = \sum_{i: V_i = j} E[Y_i] = \sum_{i: V_i = j} \mu_i
$$

where $d_j$ is the expected number of events in arm $j$.  For balanced
arms with total events $D = d_0 + d_1$:

$$
\text{Var}(\hat\beta_{\text{full}}) = \frac{1}{d_0} + \frac{1}{d_1}
\approx \frac{4}{D}
$$

Therefore $d_{\text{eff}}^{\text{Poisson}} = D$ (total events).


### Cox PH

For the Cox model, the partial likelihood score gives
$\text{Var}(\hat\beta) \approx 4/d$ where $d$ is the number of events.
Therefore $d^{\text{Cox}} = d$ (events).


### The Equivalence

Under proportional hazards with a piecewise-constant baseline hazard,
the Poisson regression with log(time) offset and the Cox model
estimate the same log-hazard ratio (Laird and Olivier, 1981;
Whitehead, 1980).  The key identity is:

$$
d_{\text{eff}}^{\text{Poisson}} = D = d^{\text{Cox}}
$$

Both equal the number of observed events in the subgroup.  Therefore
the split-half variance $8/d_{\text{eff}}$ is identical, and the
approximation (3) gives the same result for both models.


### Other GLM Families

| Family | Link | $\kappa$ | $d_{\text{eff}}$ |
|---|---|---|---|
| Cox PH | log-HR | $(1-p_c)$ | $d$ (events) |
| Poisson + offset | log-IRR | $\bar\mu$ | $D$ (events) |
| Binomial | logit | $\bar p(1-\bar p)$ | $n_H \bar p(1-\bar p)$ |
| Gaussian | identity | $1/\sigma^2$ | $n_H / \sigma^2$ |

For binary outcomes (no offset), $d_{\text{eff}} = n_H \bar p(1-\bar p)$,
which is generally smaller than $D$ for survival data, explaining the
different operating characteristics observed in binary threshold
calibration.


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(weightedsurv)
library(data.table)
library(survival)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)

# L_eff save/load helpers (exported from forestsearch >= 0.2.0)
plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")
cat("H1 sims:", nsims_alt, "| H0 sims:", nsims_null, "\n")
```

```{r timing-init}
timings <- list()
t_vignette_start <- proc.time()
```


# Data Generating Mechanism {#sec-dgm}

We use the GBSG-based DGM from Leon et al. (2024).  The harm
subgroup H = {ER-low AND premenopausal} has $\theta^\dagger(H) > 1$,
while the complement benefits ($\theta^\dagger(H^c) < 1$).

## Alternative and Null DGMs

```{r create-dgms}
t0 <- proc.time()[3]

# Alternative: calibrate for HR(H) = hr_h1 (from setup)
k_inter_cox <- calibrate_k_inter(
  target_hr_harm = hr_h1,
  model = "alt",
  k_treat = 1.0,
  cens_type = "weibull",
  use_ahr = FALSE,
  verbose = TRUE
)

dgm_alt <- setup_gbsg_dgm(
  model = "alt", k_treat = 1.0, k_inter = k_inter_cox, verbose = TRUE
)
dgm_alt <- compute_dgm_cde(dgm_alt)

cat(sprintf("\nCalibrated Alternative DGM (target HR(H) = %.1f):\n", hr_h1))
cat("  HR(H):", round(dgm_alt$hr_H_true, 3), "\n")
cat("  HR(Hc):", round(dgm_alt$hr_Hc_true, 3), "\n")
cat("  HR(ITT):", round(dgm_alt$hr_causal, 3), "\n")

# Null: uniform treatment effect
dgm_null <- setup_gbsg_dgm(model = "null", k_treat = 1.0, verbose = TRUE)

cat("\nNull DGM:\n")
cat("  HR(ITT):", round(dgm_null$hr_causal, 3), "\n")

timings$dgm <- proc.time()[3] - t0
```


# Part I: Approximation Calculations {#sec-approx}

## (D) Cox Approximation (Section 2.1, from package)

```{r approx-cox}
#| code-fold: false

# n_H_values defined in setup chunk
cat("=== (D) Cox Approximation ===\n")
cat("Using compute_detection_probability() from forestsearch\n\n")

approx_cox <- data.frame(
  n_H = n_H_values,
  d = round(d_eff_survival(n_H_values, prop_cens)),
  stringsAsFactors = FALSE
)

# Type-I error: theta = 1.0
approx_cox$P_theta_1.0 <- sapply(approx_cox$n_H, function(n) {
  compute_detection_probability(
    theta = 1.0, n_sg = n, prop_cens = prop_cens,
    hr_threshold = 1.25, hr_consistency = 1.0
  )
})

# Power: theta = hr_h1
approx_cox[[sprintf("P_theta_%.1f", hr_h1)]] <-
  sapply(approx_cox$n_H, function(n) {
    compute_detection_probability(
      theta = hr_h1, n_sg = n, prop_cens = prop_cens,
      hr_threshold = 1.25, hr_consistency = 1.0
    )
  })

# Beneficial null: theta = 0.75
approx_cox$P_theta_0.75 <- sapply(approx_cox$n_H, function(n) {
  compute_detection_probability(
    theta = 0.75, n_sg = n, prop_cens = prop_cens,
    hr_threshold = 1.25, hr_consistency = 1.0
  )
})
```

## (B) GLM Approximation (d_eff-based)

```{r approx-glm}
#| code-fold: false

# Use compute_detection_probability_glm() and d_eff_*() helpers
# from forestsearch (R/truefind_asymptotic_glm.R).
# Shorthand wrapper for concise calls:
p_detect_glm <- function(theta, d_eff, c1, c2) {
  compute_detection_probability_glm(
    theta = theta, d_eff = d_eff, c1 = c1, c2 = c2,
    effect_scale = "ratio", method = "monte_carlo"
  )
}

cat("=== (B) GLM Approximation ===\n")
cat("Using d_eff = D (events) for Poisson+offset\n")
cat("Via compute_detection_probability_glm() from forestsearch\n\n")

approx_glm <- data.frame(
  n_H = n_H_values,
  d_eff = round(d_eff_survival(n_H_values, prop_cens)),
  stringsAsFactors = FALSE
)

approx_glm$P_theta_1.0 <- sapply(approx_glm$d_eff, function(d) {
  p_detect_glm(1.0, d, 1.25, 1.0)
})

approx_glm[[sprintf("P_theta_%.1f", hr_h1)]] <-
  sapply(approx_glm$d_eff, function(d) {
    p_detect_glm(hr_h1, d, 1.25, 1.0)
  })

approx_glm$P_theta_0.75 <- sapply(approx_glm$d_eff, function(d) {
  p_detect_glm(0.75, d, 1.25, 1.0)
})
```


## (B) vs (D): Approximation Agreement

```{r approx-comparison}
#| tbl-cap: "Cox approximation (D) vs GLM approximation (B). Since d_eff = d, these should be identical."

approx_comp <- data.frame(
  n_H = approx_cox$n_H,
  d = approx_cox$d,
  `Cox P(1.0)` = round(approx_cox$P_theta_1.0, 4),
  `GLM P(1.0)` = round(approx_glm$P_theta_1.0, 4),
  `Cox P(hr_h1)` = round(approx_cox[[sprintf("P_theta_%.1f", hr_h1)]], 4),
  `GLM P(hr_h1)` = round(approx_glm[[sprintf("P_theta_%.1f", hr_h1)]], 4),
  `Cox P(0.75)` = round(approx_cox$P_theta_0.75, 4),
  `GLM P(0.75)` = round(approx_glm$P_theta_0.75, 4),
  check.names = FALSE
)
names(approx_comp)[5] <- sprintf("Cox P(%.1f)", hr_h1)
names(approx_comp)[6] <- sprintf("GLM P(%.1f)", hr_h1)

gt(approx_comp) |>
  tab_header(
    title = "(B) vs (D): GLM and Cox Approximations",
    subtitle = "d_eff(Poisson) = d(Cox) = events. Values should match."
  ) |>
  tab_spanner(label = "Type-I Error (HR=1)", columns = 3:4) |>
  tab_spanner(label = sprintf("Power (HR=%.1f)", hr_h1), columns = 5:6) |>
  tab_spanner(label = "Beneficial (HR=0.75)", columns = 7:8) |>
  tab_source_note("c1=1.25, c2=1.0, censoring=46%")
```


## Power Curves: (B) and (D)

```{r power-curves}
#| fig-height: 6
#| fig-cap: "Power curves from both approximations. The two curves should be indistinguishable (validating d_eff = d)."

# n_sg_ref and d defined from setup-chunk scalars
d_ref <- round(d_eff_survival(n_sg_ref, prop_cens))

hr_seq <- seq(0.5, 3.0, by = 0.05)

cox_curve <- sapply(hr_seq, function(hr) {
  compute_detection_probability(hr, n_sg_ref, prop_cens, 1.25, 1.0)
})
glm_curve <- sapply(hr_seq, function(hr) {
  p_detect_glm(hr, d_ref, 1.25, 1.0)
})

curve_df <- data.frame(
  OR = rep(hr_seq, 2),
  P_detect = c(cox_curve, glm_curve),
  Method = rep(c("(D) Cox Approx", "(B) GLM Approx"), each = length(hr_seq))
)

ggplot(curve_df, aes(x = OR, y = P_detect, colour = Method, linetype = Method)) +
  geom_line(linewidth = 1.0) +
  geom_hline(yintercept = c(0.10, 0.80), linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = 1.0, linetype = "dotted", colour = "grey70") +
  scale_colour_manual(values = c("(D) Cox Approx" = "steelblue",
                                  "(B) GLM Approx" = "coral")) +
  labs(x = expression(theta^dagger*(H)),
       y = "P(detect H)",
       title = "Approximation Comparison: Cox vs GLM",
       subtitle = sprintf("n(H)=%d, d=%d events, c1=1.25, c2=1.0",
                          n_sg_ref, d_ref)) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", plot.title = element_text(face = "bold"))
```


# Part II: Simulation {#sec-sim}

## QC: Single-Trial Equivalence

```{r single-trial}
sim_data <- simulate_from_dgm(
  dgm = dgm_alt, n = n_sample, rand_ratio = 1,
  seed = 1, analysis_time = analysis_time,
  cens_adjust = cens_adjust
)
sim_data$id <- seq_len(nrow(sim_data))

cat("Simulated trial: N =", nrow(sim_data),
    "| Events =", sum(sim_data$event_sim),
    "(", round(100 * mean(sim_data$event_sim), 1), "%)\n")

# Cox HR
fit_cox <- coxph(Surv(y_sim, event_sim) ~ treat_sim, data = sim_data)
cox_hr <- exp(coef(fit_cox))
cox_se <- sqrt(vcov(fit_cox)[1, 1])

# Poisson IRR
fit_pois <- glm(event_sim ~ treat_sim + offset(log(y_sim)),
                family = poisson, data = sim_data)
pois_irr <- exp(coef(fit_pois)["treat_sim"])
pois_se <- sqrt(vcov(fit_pois)["treat_sim", "treat_sim"])

cat(sprintf("\nITT Estimates:\n  Cox HR:      %.4f (SE: %.4f)\n  Poisson IRR: %.4f (SE: %.4f)\n  Ratio:       %.4f\n",
            cox_hr, cox_se, pois_irr, pois_se, cox_hr / pois_irr))

d_obs <- sum(sim_data$event_sim)
cat(sprintf("\nVariance of log-effect:\n  Cox:     %.4f\n  Poisson: %.4f\n  4/d:     %.4f  (d = %d)\n",
            cox_se^2, pois_se^2, 4/d_obs, d_obs))
```


## Simulation Parameters

```{r sim-params}
# Build compound parameter objects from setup-chunk scalars
fs_params_base <- list(
  hr.threshold = 1.25,
  hr.consistency = 1.0,
  pconsistency.threshold = pcons,
  fs.splits = fs_splits,
  n.min = n_min_val,
  d0.min = 12, d1.min = 12,
  maxk = maxk_val, by.risk = 12,
  vi.grf.min = -0.2,
  use_lasso = TRUE, use_grf = TRUE,
  use_twostage = TRUE, twostage_args = list(),
  parallel_args = list(plan="sequential", workers = 1, show_message = FALSE)
)

fs_params_cox <- modifyList(fs_params_base, list(
  outcome.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id"
))

fs_params_irr <- modifyList(fs_params_base, list(
  outcome.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id",
  outcome_type = "binary", effect_measure = "IRR",
  offset.name = "y_sim", dmin.grf = 0.0
))

sim_config <- list(
  n_sims_alt = nsims_alt,
  n_sims_null = nsims_null,
  n_sample = n_sample,
  analysis_time = analysis_time,
  seed_base = sim_seed,
  cens_adjust = cens_adjust
)

cat("Simulation config:\n")
cat("  H1 sims:", sim_config$n_sims_alt, "\n")
cat("  H0 sims:", sim_config$n_sims_null, "\n")
cat("  N:", sim_config$n_sample, "\n")
cat("  analysis_time:", sim_config$analysis_time, "months\n")
cat("  n.min:", fs_params_base$n.min, "(paper default)\n")
```


## (C) Cox PH Simulations

```{r run-cox-alt}
cat("Running Cox H1 simulations...\n")
t0 <- proc.time()[3]

results_alt_cox <- foreach(
  sim = seq_len(sim_config$n_sims_alt),
  .combine = rbind, .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE)
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id = sim, dgm = dgm_alt,
    n_sample = sim_config$n_sample,
    analysis_time = sim_config$analysis_time,
    cens_adjust = sim_config$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
    fs_params = fs_params_cox, verbose = FALSE
  )
}
timings$alt_cox <- proc.time()[3] - t0
if (nrow(results_alt_cox) < sim_config$n_sims_alt) {
  warning(sprintf(
    "Cox H1: %d of %d tasks silently failed.",
    sim_config$n_sims_alt - nrow(results_alt_cox), sim_config$n_sims_alt
  ))
}
cat("Cox H1:", round(timings$alt_cox/60, 1), "min |",
    nrow(results_alt_cox), "/", sim_config$n_sims_alt, "rows\n")
```

```{r run-cox-null}
cat("Running Cox H0 simulations...\n")
t0 <- proc.time()[3]

results_null_cox <- foreach(
  sim = seq_len(sim_config$n_sims_null),
  .combine = rbind, .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE)
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id = sim, dgm = dgm_null,
    n_sample = sim_config$n_sample,
    analysis_time = sim_config$analysis_time,
    cens_adjust = sim_config$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
    fs_params = fs_params_cox, verbose = FALSE
  )
}
timings$null_cox <- proc.time()[3] - t0
if (nrow(results_null_cox) < sim_config$n_sims_null) {
  warning(sprintf(
    "Cox H0: %d of %d tasks silently failed.",
    sim_config$n_sims_null - nrow(results_null_cox), sim_config$n_sims_null
  ))
}
cat("Cox H0:", round(timings$null_cox/60, 1), "min |",
    nrow(results_null_cox), "/", sim_config$n_sims_null, "rows\n")
```


## (A) Poisson IRR Simulations

```{r run-irr-alt}
cat("Running Poisson IRR H1 simulations...\n")
t0 <- proc.time()[3]

results_alt_irr <- foreach(
  sim = seq_len(sim_config$n_sims_alt),
  .combine = rbind, .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE)
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id = sim, dgm = dgm_alt,
    n_sample = sim_config$n_sample,
    analysis_time = sim_config$analysis_time,
    cens_adjust = sim_config$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
    fs_params = fs_params_irr, verbose = FALSE
  )
}
timings$alt_irr <- proc.time()[3] - t0
if (nrow(results_alt_irr) < sim_config$n_sims_alt) {
  warning(sprintf(
    "IRR H1: %d of %d tasks silently failed.",
    sim_config$n_sims_alt - nrow(results_alt_irr), sim_config$n_sims_alt
  ))
}
cat("IRR H1:", round(timings$alt_irr/60, 1), "min |",
    nrow(results_alt_irr), "/", sim_config$n_sims_alt, "rows\n")
```

```{r run-irr-null}
cat("Running Poisson IRR H0 simulations...\n")
t0 <- proc.time()[3]

results_null_irr <- foreach(
  sim = seq_len(sim_config$n_sims_null),
  .combine = rbind, .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE)
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id = sim, dgm = dgm_null,
    n_sample = sim_config$n_sample,
    analysis_time = sim_config$analysis_time,
    cens_adjust = sim_config$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
    fs_params = fs_params_irr, verbose = FALSE
  )
}
timings$null_irr <- proc.time()[3] - t0
if (nrow(results_null_irr) < sim_config$n_sims_null) {
  warning(sprintf(
    "IRR H0: %d of %d tasks silently failed.",
    sim_config$n_sims_null - nrow(results_null_irr), sim_config$n_sims_null
  ))
}
cat("IRR H0:", round(timings$null_irr/60, 1), "min |",
    nrow(results_null_irr), "/", sim_config$n_sims_null, "rows\n")
```


# Part III: Four-Way Comparison {#sec-compare}

## Simulation Operating Characteristics

```{r sim-oc}
compare_metrics <- function(res_cox, res_irr) {
  cox <- res_cox[res_cox$analysis == "FS", ]
  irr <- res_irr[res_irr$analysis == "FS", ]
  cox_found <- cox[cox$any.H == 1, ]
  irr_found <- irr[irr$any.H == 1, ]

  data.frame(
    Metric = c("Detection Rate", "Sensitivity", "Specificity",
               "PPV", "NPV",
               "Mean Effect (H)", "Mean Effect (Hc)",
               "Mean Effect (ITT)", "Mean SG Size"),
    `(C) Cox PH` = c(
      round(mean(cox$any.H), 3),
      round(mean(cox$sens, na.rm = TRUE), 3),
      round(mean(cox$spec, na.rm = TRUE), 3),
      round(mean(cox$ppv, na.rm = TRUE), 3),
      round(mean(cox$npv, na.rm = TRUE), 3),
      round(mean(cox_found$hr.H.hat, na.rm = TRUE), 3),
      round(mean(cox_found$hr.Hc.hat, na.rm = TRUE), 3),
      round(mean(cox$hr.itt, na.rm = TRUE), 3),
      round(mean(cox_found$size.H, na.rm = TRUE), 0)),
    `(A) Poisson IRR` = c(
      round(mean(irr$any.H), 3),
      round(mean(irr$sens, na.rm = TRUE), 3),
      round(mean(irr$spec, na.rm = TRUE), 3),
      round(mean(irr$ppv, na.rm = TRUE), 3),
      round(mean(irr$npv, na.rm = TRUE), 3),
      round(mean(irr_found$hr.H.hat, na.rm = TRUE), 3),
      round(mean(irr_found$hr.Hc.hat, na.rm = TRUE), 3),
      round(mean(irr$hr.itt, na.rm = TRUE), 3),
      round(mean(irr_found$size.H, na.rm = TRUE), 0)),
    check.names = FALSE)
}
```


### H1 (Alternative): Cox vs Poisson

```{r compare-h1}
#| tbl-cap: "(A) vs (C): Simulation comparison under H1"

comp_h1 <- compare_metrics(results_alt_cox, results_alt_irr)

gt(comp_h1) |>
  tab_header(
    title = md("**H1: Cox PH (C) vs Poisson IRR (A)**"),
    subtitle = sprintf("N=%d, %d sims, true HR(H)=%.2f, HR(Hc)=%.2f",
      sim_config$n_sample, sim_config$n_sims_alt,
      dgm_alt$hr_H_true, dgm_alt$hr_Hc_true)
  ) |>
  tab_source_note("Detection/classification should be identical (same identification step)")
```


### H0 (Null): Cox vs Poisson

```{r compare-h0}
#| tbl-cap: "(A) vs (C): Simulation comparison under H0"

comp_h0 <- compare_metrics(results_null_cox, results_null_irr)

gt(comp_h0) |>
  tab_header(
    title = md("**H0: Cox PH (C) vs Poisson IRR (A)**"),
    subtitle = sprintf("N=%d, %d sims, uniform HR=%.2f",
      sim_config$n_sample, sim_config$n_sims_null, dgm_null$hr_causal)
  )
```


## Four-Way Summary Table

```{r four-way-table}
#| tbl-cap: "The four-way comparison: approximation and simulation for both methods"

# Extract simulation detection rates
cox_h1 <- results_alt_cox[results_alt_cox$analysis == "FS", ]
irr_h1 <- results_alt_irr[results_alt_irr$analysis == "FS", ]
cox_h0 <- results_null_cox[results_null_cox$analysis == "FS", ]
irr_h0 <- results_null_irr[results_null_irr$analysis == "FS", ]

# Approximation at n_H = n_min_val (from setup)
n_H_dgm <- n_min_val
d_dgm <- round(d_eff_survival(n_H_dgm, prop_cens))

four_way <- data.frame(
  Calculation = c(
    "(D) Cox Approximation",
    "(C) Cox Simulation",
    "(B) GLM Approximation",
    "(A) GLM (Poisson) Simulation"
  ),
  Method = c("Approx", "Simulation", "Approx", "Simulation"),
  `P(theta=1.0)` = c(
    round(compute_detection_probability(1.0, n_H_dgm, prop_cens, 1.25, 1.0), 3),
    round(mean(cox_h0$any.H), 3),
    round(p_detect_glm(1.0, d_dgm, 1.25, 1.0), 3),
    round(mean(irr_h0$any.H), 3)
  ),
  `P(theta=0.75)` = c(
    round(compute_detection_probability(0.75, n_H_dgm, prop_cens, 1.25, 1.0), 3),
    NA,
    round(p_detect_glm(0.75, d_dgm, 1.25, 1.0), 3),
    NA
  ),
  `P(theta=hr_h1)` = c(
    round(compute_detection_probability(hr_h1, n_H_dgm, prop_cens, 1.25, 1.0), 3),
    round(mean(cox_h1$any.H), 3),
    round(p_detect_glm(hr_h1, d_dgm, 1.25, 1.0), 3),
    round(mean(irr_h1$any.H), 3)
  ),
  check.names = FALSE
)
names(four_way)[5] <- sprintf("P(theta=%.1f)", hr_h1)

gt(four_way) |>
  tab_header(
    title = md("**Four-Way Comparison: Approximation vs Simulation**"),
    subtitle = sprintf("n(H)=%d, d=%d events, c1=1.25, c2=1.0",
                       n_H_dgm, d_dgm)
  ) |>
  sub_missing(missing_text = "\u2014") |>
  tab_source_note(md(paste0(
    "**Key:** (B)=(D) validates the GLM variance formula. ",
    "(A)=(C) validates the GLM pipeline. ",
    "Approx~Sim validates the Section 2.1 framework.")))
```


## Per-Simulation Concordance

```{r concordance}
cox_det <- cox_h1[, c("sim", "any.H")]
names(cox_det) <- c("sim", "cox")
irr_det <- irr_h1[, c("sim", "any.H")]
names(irr_det) <- c("sim", "irr")
conc <- merge(cox_det, irr_det, by = "sim")

cat("=== Per-Simulation Concordance (H1) ===\n")
cat("Both found:", sum(conc$cox == 1 & conc$irr == 1), "\n")
cat("Neither:   ", sum(conc$cox == 0 & conc$irr == 0), "\n")
cat("Cox only:  ", sum(conc$cox == 1 & conc$irr == 0), "\n")
cat("IRR only:  ", sum(conc$cox == 0 & conc$irr == 1), "\n")
cat("Agreement: ", round(100 * mean(conc$cox == conc$irr), 1), "%\n")
```


## Subgroup Estimate Scatter

```{r estimate-scatter, fig.width=10, fig.height=5}
#| fig-cap: "Cox HR vs Poisson IRR within identified subgroups. Points near the diagonal confirm the Poisson-Cox equivalence for estimation."

cox_sg <- cox_h1[cox_h1$any.H == 1, c("sim", "hr.H.hat", "hr.Hc.hat")]
irr_sg <- irr_h1[irr_h1$any.H == 1, c("sim", "hr.H.hat", "hr.Hc.hat")]
sg_df <- merge(cox_sg, irr_sg, by = "sim", suffixes = c("_cox", "_irr"))

p1 <- ggplot(sg_df, aes(x = hr.H.hat_cox, y = hr.H.hat_irr)) +
  geom_point(alpha = 0.5, size = 2, colour = "firebrick") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(title = "Harm Subgroup (H)", x = "Cox HR", y = "Poisson IRR") +
  coord_equal() + theme_minimal(base_size = 12)

p2 <- ggplot(sg_df, aes(x = hr.Hc.hat_cox, y = hr.Hc.hat_irr)) +
  geom_point(alpha = 0.5, size = 2, colour = "steelblue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(title = "Complement (Hc)", x = "Cox HR", y = "Poisson IRR") +
  coord_equal() + theme_minimal(base_size = 12)

p1 + p2 + plot_annotation(
  title = "Per-Simulation Estimates: Cox HR vs Poisson IRR",
  subtitle = sprintf("%d simulations with identified subgroup", nrow(sg_df)))
```


# Part IV: Corrected Null Approximation {#sec-correction}

The per-subgroup approximation $P_1$ models a single
pre-specified subgroup.  Under $H_0$, the procedure evaluates
many candidates, and the effective number of viable candidates
grows with $N$.

## Multi-N Null Simulations for Calibration

To calibrate $L_{\text{eff}}$ we need H0 simulations at
multiple sample sizes ($N \in \{`r paste(N_cal_values, collapse = ", ")`\}$,
defined in the setup chunk).

```{r null-multi-N}
n_sims_cal <- nsims_null  # use same count as main sims

cat("Running multi-N null calibration simulations...\n")
cat(sprintf("N values: %s | Sims per N: %d\n",
            paste(N_cal_values, collapse = ", "), n_sims_cal))
t0 <- proc.time()[3]

cal_results <- list()
for (N_val in N_cal_values) {
  t_N <- proc.time()[3]

  res_N <- foreach(
    sim = seq_len(n_sims_cal),
    .combine = rbind, .errorhandling = "remove",
    .options.future = list(
      packages = c("forestsearch", "survival", "data.table", "future"),
      seed = TRUE)
  ) %dofuture% {
    future::plan("sequential")
    run_simulation_analysis(
      sim_id = sim, dgm = dgm_null,
      n_sample = N_val,
      analysis_time = sim_config$analysis_time,
      cens_adjust = sim_config$cens_adjust,
      confounders_base = confounders_base,
      cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
      n_add_noise = 0L,
      run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
      fs_params = fs_params_cox, verbose = FALSE
    )
  }

  fs_rows <- res_N[res_N$analysis == "FS", ]
  if (nrow(fs_rows) < n_sims_cal) {
    warning(sprintf(
      "Calibration N=%d: %d of %d tasks silently failed.",
      N_val, n_sims_cal - nrow(fs_rows), n_sims_cal
    ))
  }
  fpr_N <- mean(fs_rows$any.H, na.rm = TRUE)

  cal_results[[as.character(N_val)]] <- data.frame(
    N = N_val, sim_fpr = fpr_N, n_sims = nrow(fs_rows)
  )

  cat(sprintf("  N = %4d: FPR = %.3f (%d sims) | %.1f sec\n",
      N_val, fpr_N, nrow(fs_rows), proc.time()[3] - t_N))
}

cal_df <- do.call(rbind, cal_results)
timings$null_cal <- proc.time()[3] - t0
cat(sprintf("Multi-N calibration: %.1f min\n", timings$null_cal / 60))
```


## Compute Per-Subgroup $P_1$ at Each N

```{r cal-P1}
cal_df$n_H <- n_min_val
cal_df$d_eff <- round(d_eff_survival(cal_df$n_H, prop_cens))

cal_df$P1 <- sapply(cal_df$d_eff, function(d) {
  p_detect_glm(1.0, d, 1.25, 1.0)
})

cat("=== Per-Subgroup P1 vs Simulated FPR ===\n")
cal_df$ratio <- round(cal_df$sim_fpr / pmax(cal_df$P1, 0.001), 2)
print(cal_df[, c("N", "n_H", "d_eff", "P1", "sim_fpr", "ratio")])
```


## Calibrate $L_{\text{eff}}$

```{r calibrate-leff}
#| fig-height: 5
#| fig-cap: "Implied L_eff at each N with power-law fit for the GBSG survival DGM."

# Implied L_eff
cal_df$L_eff <- log(1 - cal_df$sim_fpr) / log(1 - cal_df$P1)

cat("=== Implied L_eff ===\n")
print(cal_df[, c("N", "n_H", "d_eff", "P1", "sim_fpr", "L_eff")])

# Filter to valid rows (sim_fpr > 0 and P1 > 0)
cal_valid <- cal_df[cal_df$sim_fpr > 0 & cal_df$P1 > 0 &
                     is.finite(cal_df$L_eff), ]

if (nrow(cal_valid) >= 2) {
  leff_cal <- calibrate_L_eff(
    N       = cal_valid$N,
    P1      = cal_valid$P1,
    sim_fpr = cal_valid$sim_fpr,
    n_min   = n_min_val
  )
  print(leff_cal)

  C_fit <- leff_cal$C
  alpha_fit <- leff_cal$alpha

  N_pred <- seq(200, 1200, by = 5)
  pred_curve <- data.frame(
    N = N_pred,
    L_eff = C_fit * (N_pred / n_min_val)^alpha_fit
  )

  print(
    ggplot() +
      geom_point(data = cal_valid,
                 aes(x = N, y = L_eff), size = 4, colour = "firebrick") +
      geom_line(data = pred_curve,
                aes(x = N, y = L_eff), colour = "steelblue",
                linewidth = 1) +
      labs(x = "Total Sample Size N",
           y = expression(L[eff]),
           title = expression(L[eff] * " Calibration: GBSG Survival DGM"),
           subtitle = sprintf("L_eff = %.3f * (N/%d)^%.3f  (R-sq = %.3f)",
                               C_fit, n_min_val, alpha_fit,
                               leff_cal$r_squared)) +
      theme_minimal(base_size = 12) +
      theme(plot.title = element_text(face = "bold"))
  )

  cat(sprintf("\nFitted model: L_eff = %.3f * (N/%d)^%.3f\n",
              C_fit, n_min_val, alpha_fit))
  for (N_val in N_cal_values) {
    cat(sprintf("  N = %4d: L_eff = %.2f\n", N_val,
                C_fit * (N_val / n_min_val)^alpha_fit))
  }
} else {
  cat("Insufficient valid calibration points (need >= 2).\n")
  cat("Using fallback L_eff = 1 (no correction).\n")
  C_fit <- 1 / (n_sample / n_min_val)
  alpha_fit <- 1
}
```


## Save $L_{\text{eff}}$ Calibration

```{r save-leff}
# Save calibration to .rds for downstream documents (S2)
save_leff_calibration(
  C = C_fit,
  alpha = alpha_fit,
  n_min = n_min_val,
  outcome_type = "survival",
  cal_data = cal_df,
  dgm_description = paste(
    "GBSG survival DGM,", length(confounders_base), "confounders",
    paste0("(", paste(confounders_base, collapse = ", "), "),"),
    sprintf("prop_cens=%.2f, p_harm=%.2f", prop_cens, p_harm)),
  n_sims_per_N = n_sims_cal,
  output_dir = output_dir,
  extra = list(
    # ForestSearch config (S2 should match these)
    confounders_base = confounders_base,
    prop_cens        = prop_cens,
    p_harm           = p_harm,
    hr_h1            = hr_h1,
    pcons            = pcons,
    fs_splits        = fs_splits,
    maxk_val         = maxk_val,
    n_min_val        = n_min_val,
    analysis_time    = analysis_time,
    cens_adjust      = cens_adjust,
    fpr_target       = fpr_target,
    N_cal_values     = N_cal_values,
    # Grid range validated for L_eff threshold-independence
    # (not swept in C2; C3 provides this via its grid sweep)
    c1_range = c(1.00, 1.50),
    c2_range = c(0.90, 1.25)
  )
)
```

::: {.callout-important}
## Shared `.rds` With C3

Both this document (C2) and `calibration_survival_leff_grid.qmd` (C3)
save to `calibration_survival_leff.rds` via
`save_leff_calibration(outcome_type = "survival")`.
**C2 is the primary source** because it uses the full null simulation
(up to 20K replicates).  C3 uses fewer replicates and is
designed as a faster grid sweep.  If both are rendered, the
last-rendered file wins.  Render C2 last, or set
`recalibrate = FALSE` in C3 to skip its calibration.
:::


## Corrected Four-Way Table

```{r corrected-four-way}
#| tbl-cap: "Four-way comparison with L_eff-corrected null approximation.  The corrected row should align with simulation."

L_eff_main <- C_fit * (n_sample / n_min_val)^alpha_fit

P1_cox <- compute_detection_probability(
  1.0, n_H_dgm, prop_cens, 1.25, 1.0)
P1_glm <- p_detect_glm(1.0, d_dgm, 1.25, 1.0)

fpr_corr_cox <- 1 - (1 - P1_cox)^L_eff_main
fpr_corr_glm <- 1 - (1 - P1_glm)^L_eff_main

four_way_corr <- data.frame(
  Calculation = c(
    "(D) Cox Approx (uncorrected)",
    "(D') Cox Approx (L_eff corrected)",
    "(C) Cox Simulation",
    "(B) GLM Approx (uncorrected)",
    "(B') GLM Approx (L_eff corrected)",
    "(A) GLM (Poisson) Simulation"
  ),
  `P(theta=1.0)` = c(
    round(P1_cox, 3),
    round(fpr_corr_cox, 3),
    round(mean(cox_h0$any.H), 3),
    round(P1_glm, 3),
    round(fpr_corr_glm, 3),
    round(mean(irr_h0$any.H), 3)
  ),
  `P(theta=hr_h1)` = c(
    round(compute_detection_probability(hr_h1, n_H_dgm, prop_cens, 1.25, 1.0), 3),
    NA,
    round(mean(cox_h1$any.H), 3),
    round(p_detect_glm(hr_h1, d_dgm, 1.25, 1.0), 3),
    NA,
    round(mean(irr_h1$any.H), 3)
  ),
  check.names = FALSE
)
names(four_way_corr)[3] <- sprintf("P(theta=%.1f)", hr_h1)

gt(four_way_corr) |>
  tab_header(
    title = md("**Four-Way Comparison with L_eff Correction**"),
    subtitle = sprintf("N=%d, n(H)=%d, d=%d, L_eff=%.2f",
                        n_sample, n_H_dgm, d_dgm, L_eff_main)
  ) |>
  sub_missing(missing_text = "\u2014") |>
  tab_row_group(label = "GLM (Poisson+offset)",
    rows = grepl("GLM|\\(A\\)|\\(B\\)", Calculation)) |>
  tab_row_group(label = "Cox PH",
    rows = grepl("Cox|\\(C\\)|\\(D\\)", Calculation)) |>
  tab_source_note(md(paste0(
    "**Corrected rows (D', B')** should match simulation rows (C, A). ",
    sprintf("Power (theta=%.1f) does not need correction.", hr_h1))))
```


## Correction Assessment

```{r correction-assessment}
cat(sprintf("=== L_eff Correction Assessment (N = %d) ===\n\n", n_sample))
cat(sprintf("L_eff at N=%d: %.2f\n", n_sample, L_eff_main))
cat(sprintf("  (i.e., %.1f effectively independent candidates)\n\n",
            L_eff_main))

cat("Cox PH (H0):\n")
cat(sprintf("  Uncorrected P1:    %.3f\n", P1_cox))
cat(sprintf("  Corrected:         %.3f\n", fpr_corr_cox))
cat(sprintf("  Simulated:         %.3f\n", mean(cox_h0$any.H)))
cat(sprintf("  Sim/Uncorrected:   %.2f\n",
            mean(cox_h0$any.H) / max(P1_cox, 0.001)))
cat(sprintf("  Sim/Corrected:     %.2f\n\n",
            mean(cox_h0$any.H) / max(fpr_corr_cox, 0.001)))

cat("GLM Poisson (H0):\n")
cat(sprintf("  Uncorrected P1:    %.3f\n", P1_glm))
cat(sprintf("  Corrected:         %.3f\n", fpr_corr_glm))
cat(sprintf("  Simulated:         %.3f\n", mean(irr_h0$any.H)))
cat(sprintf("  Sim/Uncorrected:   %.2f\n",
            mean(irr_h0$any.H) / max(P1_glm, 0.001)))
cat(sprintf("  Sim/Corrected:     %.2f\n",
            mean(irr_h0$any.H) / max(fpr_corr_glm, 0.001)))
```


## Comparison to Binary Calibration

```{r compare-binary}
cat("=== Cross-DGM L_eff Comparison ===\n\n")
cat("The L_eff calibration depends on the candidate generation\n")
cat("process, not the outcome type.\n\n")
cat(sprintf("GBSG Survival DGM (%d confounders): C=%.3f, alpha=%.3f\n",
            length(confounders_base), C_fit, alpha_fit))
cat(sprintf("  L_eff(N=%d) = %.2f\n\n", n_sample, L_eff_main))
cat("Binary DGM (4 confounders): see calibration_binary_leff.qmd\n")
cat("  Expected: smaller L_eff (fewer confounders -> fewer candidates)\n")
```


# Summary {#sec-summary}

## Validation Conclusions

This four-way comparison validates the GLM approximation for
ForestSearch in two complementary ways:

1. **(B) = (D)**: The GLM approximation with
   $d_{\text{eff}} = D$ (events) produces identical results
   to the established Cox approximation with $d$ = events.
   This confirms the theoretical equivalence
   $d_{\text{eff}}^{\text{Poisson}} = d^{\text{Cox}}$.

2. **(A) = (C)**: The Poisson+offset simulation produces
   identical detection rates, classification metrics, and
   closely matching effect estimates compared to the Cox
   simulation.  This confirms the GLM pipeline is correctly
   implemented.

3. **Power approximation ~ Simulation**: The per-subgroup
   approximation agrees well with simulated power (H1),
   confirming the variance formula.

4. **$L_{\text{eff}}$ correction**: Under H0, the per-subgroup
   $P_1$ underestimates the procedure-level FPR because multiple
   candidates are evaluated.  The $L_{\text{eff}}$-corrected
   approximation $1 - (1-P_1)^{L_{\text{eff}}}$ brings
   the null prediction into alignment with simulation.
   $L_{\text{eff}}$ is calibrated via
   `calibrate_L_eff()` from a small H0 reference simulation.

## Implications for Other GLM Outcomes

The Poisson+offset case is a validation anchor: since it
targets the same quantity as the Cox model, agreement
confirms the GLM infrastructure.  For other outcome types:

| Outcome | $d_{\text{eff}}$ | Validation status |
|---|---|---|
| Poisson + offset | $D$ (events) = $d$ (Cox) | Validated here |
| Binary (logit) | $n_H \bar p(1-\bar p)$ | Requires separate calibration |
| Continuous | $n_H / \sigma^2$ | Requires separate calibration |

The binary case has $d_{\text{eff}} < d$ (Cox) for typical
event rates, explaining why the survival defaults (c1=1.25,
c2=1.0) do not provide adequate FPR control for binary
outcomes.

## References

- Leon, L.F., Jemielita, T., Guo, Z., Marceau West, R.,
  Anderson, K.M. (2024). Exploratory subgroup identification
  in the heterogeneous Cox model. *Statistics in Medicine*,
  43, 4253--4275.
- Jennison, C. and Turnbull, B.W. (1984). Repeated confidence
  intervals for group sequential clinical trials. *Controlled
  Clinical Trials*, 5(1), 33--45.
- McCullagh, P. and Nelder, J.A. (1989). *Generalized Linear
  Models*, 2nd ed. Chapman and Hall.
- Laird, N. and Olivier, D. (1981). Covariance analysis of
  censored survival data using log-linear analysis techniques.
  *Journal of the American Statistical Association*, 76(374),
  231--240.
- Whitehead, J. (1980). Fitting Cox's regression model to
  survival data using GLIM. *Journal of the Royal Statistical
  Society C*, 29(3), 268--275.


# Save All Simulation Results {#sec-save}

The simulation results are computationally expensive (hours).
Save everything to `.rds` so downstream documents and future
analyses can reload without re-running.

```{r save-all-results}
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

sim_bundle <- list(

  # Metadata
  params = list(
    nsims_alt = nsims_alt,
    nsims_null = nsims_null,
    sim_seed = sim_seed,
    n_workers = n_workers,
    n_sample = n_sample,
    hr_h1 = hr_h1,
    analysis_time = analysis_time,
    cens_adjust = cens_adjust,
    fpr_target = fpr_target,
    prop_cens = prop_cens,
    p_harm = p_harm,
    n_min_val = n_min_val,
    N_cal_values = N_cal_values,
    sim_config = sim_config,
    fs_params_base = fs_params_base,
    fs_params_cox = fs_params_cox,
    fs_params_irr = fs_params_irr,
    confounders_base = confounders_base,
    timestamp = Sys.time(),
    R_version = R.version.string,
    forestsearch_version = tryCatch(
      as.character(utils::packageVersion("forestsearch")),
      error = function(e) "unknown")
  ),

  # DGM objects
  dgm_alt = dgm_alt,
  dgm_null = dgm_null,

  # Raw simulation results (the expensive part)
  results_alt_cox = results_alt_cox,
  results_null_cox = results_null_cox,
  results_alt_irr = results_alt_irr,
  results_null_irr = results_null_irr,

  # L_eff calibration
  leff = list(
    C = C_fit,
    alpha = alpha_fit,
    n_min = n_min_val,
    cal_df = cal_df
  ),

  # Timing
  timings = timings
)

rds_path <- file.path(output_dir,
  sprintf("calibration_survival_fourway_%s.rds",
          format(Sys.Date(), "%Y%m%d")))
saveRDS(sim_bundle, rds_path)

cat(sprintf("=== All results saved ===\n"))
cat(sprintf("  File: %s\n", rds_path))
cat(sprintf("  Size: %.1f MB\n", file.size(rds_path) / 1024^2))
cat(sprintf("  Contents:\n"))
cat(sprintf("    results_alt_cox:  %d rows\n", nrow(results_alt_cox)))
cat(sprintf("    results_null_cox: %d rows\n", nrow(results_null_cox)))
cat(sprintf("    results_alt_irr:  %d rows\n", nrow(results_alt_irr)))
cat(sprintf("    results_null_irr: %d rows\n", nrow(results_null_irr)))
cat(sprintf("    cal_df:           %d rows (N = %s)\n",
            nrow(cal_df), paste(cal_df$N, collapse = ", ")))
cat(sprintf("    L_eff:            C=%.4f, alpha=%.4f\n",
            C_fit, alpha_fit))
```


## How to Reload

```{r reload-example}
#| eval: false
#| code-fold: false

# Load the most recent simulation bundle
res <- readRDS(latest_rds("calibration_survival_fourway"))

# Access simulation results directly
cox_h1 <- res$results_alt_cox[res$results_alt_cox$analysis == "FS", ]
cox_h0 <- res$results_null_cox[res$results_null_cox$analysis == "FS", ]
irr_h1 <- res$results_alt_irr[res$results_alt_irr$analysis == "FS", ]
irr_h0 <- res$results_null_irr[res$results_null_irr$analysis == "FS", ]

# Check what was run
res$params$nsims_alt        # number of H1 sims
res$params$nsims_null       # number of H0 sims
res$params$hr_h1            # HR for harm subgroup
res$params$analysis_time    # GBSG analysis time
res$params$fpr_target       # FPR target

# L_eff calibration
res$leff$C               # fitted C
res$leff$alpha           # fitted alpha
res$leff$cal_df          # raw calibration data

# Re-derive metrics without re-running
cat("Cox H0 FPR:", mean(cox_h0$any.H), "\n")
cat("Cox H1 Power:", mean(cox_h1$any.H), "\n")
cat("IRR H0 FPR:", mean(irr_h0$any.H), "\n")
cat("IRR H1 Power:", mean(irr_h1$any.H), "\n")
```


# Timing and Session Info {#sec-session}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Stage = c("DGM", "Cox H1", "Cox H0",
            "IRR H1", "IRR H0",
            "L_eff Calibration (multi-N)",
            "Total"),
  Seconds = c(timings$dgm, timings$alt_cox, timings$null_cox,
              timings$alt_irr, timings$null_irr,
              timings$null_cal, timings$total),
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(
    title = "Computational Timing Summary",
    subtitle = sprintf(
      "%d H1 sims + %d H0 sims + %d cal sims x %d N | %d workers",
      nsims_alt, nsims_null, n_sims_cal,
      length(N_cal_values), n_workers)
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)",
             Minutes = "Time (min)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total")
  )
```

```{r session-info}
plan(sequential)
sessionInfo()
```
