---
title: "Cox vs GLM Approximation: Null Operating Characteristics"
subtitle: "Threshold Calibration for the GBSG Survival DGM"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ── Simulation control ──────────────────────────────────────────
nsims_null  <- 30L       # H0 sims per config (use 500+ for production)
sim_seed    <- 8316951L
n_workers   <- ceiling(0.95 * max(1L,
  parallel::detectCores(logical = FALSE) - 1L))

# ── Output ──────────────────────────────────────────────────────
output_dir  <- "_data"

# ── FPR target ─────────────────────────────────────────────────
fpr_target  <- 0.12

# ── DGM / outcome parameters ──────────────────────────────────
prop_cens   <- 0.46      # GBSG censoring proportion
p_harm      <- 0.13      # harm subgroup prevalence (metadata only)
n_min_val   <- 60L       # n.min = reference subgroup size for P1
hr_h1       <- 2.0       # HR for harm subgroup under H1 (reference)

# ── ForestSearch parameters ───────────────────────────────────
confounders_base <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")
pcons       <- 0.90      # pconsistency.threshold
fs_splits   <- 400L
maxk_val    <- 2L
n_sample    <- 700L      # reference sample size for main sims
analysis_time <- 84L     # GBSG analysis time (months)
cens_adjust   <- log(1.5)

# ── L_eff calibration ────────────────────────────────────────
N_cal_values <- c(300, 500, 700)

# ── Threshold grid for null sweep (Part IV) ──────────────────
# C3 validates L_eff threshold-independence across this grid.
# S2's approximation grid should fall within this range.
c1_grid <- c(1.00, 1.10, 1.15, 1.20, 1.25, 1.30, 1.40, 1.50)
c2_grid <- c(0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20, 1.25)
```


# Introduction {#sec-intro}

Leon et al. (2024, Section 2.1) approximate the per-subgroup
detection probability as:

$$
P_1(c_1, c_2; d, \theta) =
P(W_1 + W_2 \ge 2\log c_1, \;\min(W_1, W_2) \ge \log c_2)
\tag{3}
$$

with $W_1, W_2 \stackrel{\text{iid}}{\sim} N(\log\theta, 8/d)$.

This document focuses on **null operating characteristics**
for the GBSG survival DGM: we evaluate how the false-positive
rate (FPR) varies across a grid of $(c_1, c_2)$ threshold
pairs to identify the region where the procedure-level FPR
is at most 10--12%.

We compare:

- **(B)** GLM approximation (per-subgroup $P_1$ via
  `compute_detection_probability_glm()`)
- **(C)** Cox ForestSearch simulation
- **(A)** Poisson+offset ForestSearch simulation
- **$L_{\text{eff}}$ correction**: $1 - (1 - P_1)^{L_{\text{eff}}}$

All simulations are under $H_0$ (uniform treatment effect,
no subgroup heterogeneity).


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(weightedsurv)
library(data.table)
library(survival)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)

# L_eff save/load helpers
#source(here::here("../R", "leff_calibration_io.R"))
plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")
cat("H0 sims per config:", nsims_null, "\n")
```

```{r timing-init}
timings <- list()
t_vignette_start <- proc.time()
```


# Null DGM {#sec-dgm}

```{r create-dgm-null}
t0 <- proc.time()[3]

dgm_null <- setup_gbsg_dgm(model = "null", k_treat = 1.0, verbose = TRUE)

cat("\nNull DGM:\n")
cat("  HR(ITT):", round(dgm_null$hr_causal, 3), "\n")
cat("  AHR:", round(dgm_null$AHR, 3), "\n")

timings$dgm <- proc.time()[3] - t0
```


# Part I: Per-Subgroup Approximation {#sec-approx}

Instant calculations --- no simulation required.

```{r approx-glm-fn}
#| code-fold: false

# Shorthand wrapper for compute_detection_probability_glm()
p_detect_glm <- function(theta, d_eff, c1, c2) {
  compute_detection_probability_glm(
    theta = theta, d_eff = d_eff, c1 = c1, c2 = c2,
    effect_scale = "ratio", method = "monte_carlo"
  )
}
```


## Approximation Across Threshold Pairs

```{r approx-grid-null}
#| tbl-cap: !expr 'sprintf("Per-subgroup P1(theta=1) across (c1,c2) pairs at n.min = %d", n_min_val)'

# Reference subgroup: n_H = n.min (from setup)
n_H_dgm <- n_min_val
d_dgm <- round(d_eff_survival(n_H_dgm, prop_cens))

approx_null <- expand.grid(c1 = c1_grid, c2 = c2_grid,
                            stringsAsFactors = FALSE)
approx_null <- approx_null[approx_null$c2 <= approx_null$c1, ]

approx_null$P1 <- mapply(function(c1v, c2v) {
  round(p_detect_glm(1.0, d_dgm, c1v, c2v), 4)
}, approx_null$c1, approx_null$c2)

gt(approx_null) |>
  tab_header(
    title = "Per-Subgroup P1(theta=1.0) by Threshold Pair",
    subtitle = sprintf("n(H)=%d, d=%d events, prop_cens=%.2f",
                        n_H_dgm, d_dgm, prop_cens)
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = P1, rows = P1 <= 0.10)) |>
  tab_style(style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(columns = P1,
      rows = P1 > 0.10 & P1 <= 0.15)) |>
  tab_source_note(
    "P1 is the per-subgroup probability.  Procedure-level FPR may differ."
  )
```


## Per-Subgroup Approximation Heatmap

```{r approx-heatmap, fig.width=8, fig.height=6}
#| fig-cap: "Per-subgroup P1(theta=1) heatmap.  This is the single-subgroup probability; the procedure-level FPR is evaluated by simulation below."

ggplot(approx_null, aes(x = factor(c1), y = factor(c2), fill = P1)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.1f%%", 100 * P1)),
            size = 3.5, colour = "black") +
  scale_fill_gradient2(
    low = "#E8F5E9", mid = "#FFF9C4", high = "#E53935",
    midpoint = 0.10, name = "P1"
  ) +
  labs(x = expression(c[1] ~ "(screening)"),
       y = expression(c[2] ~ "(consistency)"),
       title = "Per-Subgroup Approximation: P1(theta=1)",
       subtitle = sprintf("n(H)=%d, d=%d", n_H_dgm, d_dgm)) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```


# Part II: Null Simulations {#sec-sim}

## Simulation Parameters

```{r sim-params}
# Build compound parameter objects from setup-chunk scalars
fs_params_base <- list(
  hr.threshold = 1.25,
  hr.consistency = 1.0,
  pconsistency.threshold = pcons,
  fs.splits = fs_splits,
  n.min = n_min_val,
  d0.min = 12, d1.min = 12,
  maxk = maxk_val, by.risk = 12,
  vi.grf.min = -0.2,
  use_lasso = TRUE, use_grf = TRUE,
  use_twostage = TRUE, twostage_args = list()
)

fs_params_cox <- modifyList(fs_params_base, list(
  outcome.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id"
))

fs_params_irr <- modifyList(fs_params_base, list(
  outcome.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id",
  outcome_type = "binary", effect_measure = "IRR",
  offset.name = "y_sim", dmin.grf = 0.0
))

sim_config <- list(
  n_sample = n_sample,
  analysis_time = analysis_time,
  seed_base = sim_seed,
  cens_adjust = cens_adjust
)

cat("Simulation config:\n")
cat("  H0 sims:", nsims_null, "\n")
cat("  N:", sim_config$n_sample, "\n")
cat("  analysis_time:", sim_config$analysis_time, "months\n")
cat("  n.min:", fs_params_base$n.min, "(paper default)\n")
```


## Cox PH Null Simulations (Paper Defaults)

```{r run-cox-null}
cat("Running Cox H0 simulations (c1=1.25, c2=1.0)...\n")
t0 <- proc.time()[3]

results_null_cox <- foreach(
  sim = seq_len(nsims_null),
  .combine = rbind, .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table"),
    seed = TRUE)
) %dofuture% {
  run_simulation_analysis(
    sim_id = sim, dgm = dgm_null,
    n_sample = sim_config$n_sample,
    analysis_time = sim_config$analysis_time,
    cens_adjust = sim_config$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
    fs_params = fs_params_cox, verbose = FALSE
  )
}
timings$null_cox <- proc.time()[3] - t0
cat("Cox H0:", round(timings$null_cox / 60, 1), "min\n")
```


## Poisson IRR Null Simulations (Paper Defaults)

```{r run-irr-null}
cat("Running Poisson IRR H0 simulations (c1=1.25, c2=1.0)...\n")
t0 <- proc.time()[3]

results_null_irr <- foreach(
  sim = seq_len(nsims_null),
  .combine = rbind, .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table"),
    seed = TRUE)
) %dofuture% {
  run_simulation_analysis(
    sim_id = sim, dgm = dgm_null,
    n_sample = sim_config$n_sample,
    analysis_time = sim_config$analysis_time,
    cens_adjust = sim_config$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
    fs_params = fs_params_irr, verbose = FALSE
  )
}
timings$null_irr <- proc.time()[3] - t0
cat("IRR H0:", round(timings$null_irr / 60, 1), "min\n")
```


## Baseline H0 Comparison: Cox vs Poisson (Paper Defaults)

```{r baseline-h0}
#| tbl-cap: "H0 comparison at paper defaults (c1=1.25, c2=1.0)"

cox_h0 <- results_null_cox[results_null_cox$analysis == "FS", ]
irr_h0 <- results_null_irr[results_null_irr$analysis == "FS", ]

P1_default <- p_detect_glm(1.0, d_dgm, 1.25, 1.0)

baseline_h0 <- data.frame(
  Metric = c("Simulated FPR", "Per-subgroup P1", "Sim/P1 Ratio",
             "N sims"),
  `Cox PH` = c(
    round(mean(cox_h0$any.H), 4),
    round(P1_default, 4),
    round(mean(cox_h0$any.H) / P1_default, 2),
    nrow(cox_h0)
  ),
  `Poisson IRR` = c(
    round(mean(irr_h0$any.H), 4),
    round(P1_default, 4),
    round(mean(irr_h0$any.H) / P1_default, 2),
    nrow(irr_h0)
  ),
  check.names = FALSE
)

gt(baseline_h0) |>
  tab_header(
    title = md("**H0 Baseline: Cox vs Poisson (c1=1.25, c2=1.0)**"),
    subtitle = sprintf("N=%d, %d sims, HR(ITT)=%.2f",
      sim_config$n_sample, nsims_null,
      dgm_null$hr_causal)
  )
```


# Part III: $L_{\text{eff}}$ Calibration {#sec-leff}

## Multi-N Null Simulations

```{r null-multi-N}
n_sims_cal <- min(nsims_null, 500L)

cat("Running multi-N null calibration simulations...\n")
cat(sprintf("N values: %s | Sims per N: %d\n",
            paste(N_cal_values, collapse = ", "), n_sims_cal))
t0 <- proc.time()[3]

cal_results <- list()
for (N_val in N_cal_values) {
  t_N <- proc.time()[3]

  res_N <- foreach(
    sim = seq_len(n_sims_cal),
    .combine = rbind, .errorhandling = "remove",
    .options.future = list(
      packages = c("forestsearch", "survival", "data.table"),
      seed = TRUE)
  ) %dofuture% {
    run_simulation_analysis(
      sim_id = sim, dgm = dgm_null,
      n_sample = N_val,
      analysis_time = sim_config$analysis_time,
      cens_adjust = sim_config$cens_adjust,
      confounders_base = confounders_base,
      cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
      n_add_noise = 0L,
      run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
      fs_params = fs_params_cox, verbose = FALSE
    )
  }

  fs_rows <- res_N[res_N$analysis == "FS", ]
  fpr_N <- mean(fs_rows$any.H, na.rm = TRUE)

  cal_results[[as.character(N_val)]] <- data.frame(
    N = N_val, sim_fpr = fpr_N, n_sims = nrow(fs_rows)
  )

  cat(sprintf("  N = %4d: FPR = %.3f (%d sims) | %.1f sec\n",
      N_val, fpr_N, nrow(fs_rows), proc.time()[3] - t_N))
}

cal_df <- do.call(rbind, cal_results)
timings$null_cal <- proc.time()[3] - t0
cat(sprintf("Multi-N calibration: %.1f min\n", timings$null_cal / 60))
```


## Implied $L_{\text{eff}}$ from Simulation

For each $N$, solve for the implied $L_{\text{eff}}$:

$$
L_{\text{eff}} =
\frac{\log(1 - \text{FPR}_{\text{sim}})}
     {\log(1 - P_1)}
$$

```{r compute-leff}
#| tbl-cap: "Implied L_eff at each N.  Consistency across threshold pairs supports L_eff as a candidate-count property."

cal_df$n_H <- n_min_val
cal_df$d_eff <- round(d_eff_survival(cal_df$n_H, prop_cens))
cal_df$P1 <- sapply(cal_df$d_eff, function(d) {
  p_detect_glm(1.0, d, 1.25, 1.0)
})
cal_df$L_eff <- log(1 - cal_df$sim_fpr) / log(1 - cal_df$P1)

cat("=== Implied L_eff ===\n")
print(cal_df[, c("N", "n_H", "d_eff", "P1", "sim_fpr", "L_eff")])
```


## Fit $L_{\text{eff}}$ Power Law

```{r calibrate-leff}
#| fig-height: 5
#| fig-cap: "Implied L_eff with power-law fit for the GBSG survival DGM."

# Fit power law
cal_valid <- cal_df[cal_df$sim_fpr > 0 & cal_df$P1 > 0 &
                     is.finite(cal_df$L_eff), ]

if (nrow(cal_valid) >= 2) {
  leff_cal <- calibrate_L_eff(
    N = cal_valid$N, P1 = cal_valid$P1,
    sim_fpr = cal_valid$sim_fpr, n_min = n_min_val
  )
  print(leff_cal)

  C_fit <- leff_cal$C
  alpha_fit <- leff_cal$alpha

  N_pred <- seq(200, 1200, by = 5)
  pred_curve <- data.frame(
    N = N_pred, L_eff = C_fit * (N_pred / n_min_val)^alpha_fit
  )

  ggplot() +
    geom_point(data = cal_valid, aes(x = N, y = L_eff),
               size = 4, colour = "firebrick") +
    geom_line(data = pred_curve, aes(x = N, y = L_eff),
              colour = "steelblue", linewidth = 1) +
    labs(x = "Total Sample Size N",
         y = expression(L[eff]),
         title = expression(L[eff] * " Calibration: GBSG Survival DGM"),
         subtitle = sprintf("L_eff = %.3f * (N/%d)^%.3f  (R-sq = %.3f)",
                             C_fit, n_min_val, alpha_fit,
                             leff_cal$r_squared)) +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(face = "bold"))
} else {
  cat("Insufficient calibration points. Using L_eff = 1.\n")
  C_fit <- 1; alpha_fit <- 0
}

cat(sprintf("\nFitted model: L_eff = %.3f * (N/%d)^%.3f\n",
            C_fit, n_min_val, alpha_fit))
cat(sprintf("  N = 300:  L_eff = %.2f\n", C_fit * (300/n_min_val)^alpha_fit))
cat(sprintf("  N = 500:  L_eff = %.2f\n", C_fit * (500/n_min_val)^alpha_fit))
cat(sprintf("  N = 700:  L_eff = %.2f\n", C_fit * (700/n_min_val)^alpha_fit))
```


## Save Calibration

```{r save-leff}
# Save calibration to .rds for downstream documents (S2)
save_leff_calibration(
  C = C_fit,
  alpha = alpha_fit,
  n_min = n_min_val,
  outcome_type = "survival",
  cal_data = cal_df,
  dgm_description = paste(
    "GBSG survival DGM,", length(confounders_base), "confounders",
    paste0("(", paste(confounders_base, collapse = ", "), "),"),
    sprintf("prop_cens=%.2f, p_harm=%.2f", prop_cens, p_harm)),
  n_sims_per_N = n_sims_cal,
  output_dir = output_dir,
  extra = list(
    # ForestSearch config (S2 should match these)
    confounders_base = confounders_base,
    prop_cens        = prop_cens,
    p_harm           = p_harm,
    hr_h1            = hr_h1,
    pcons            = pcons,
    fs_splits        = fs_splits,
    maxk_val         = maxk_val,
    n_min_val        = n_min_val,
    analysis_time    = analysis_time,
    cens_adjust      = cens_adjust,
    fpr_target       = fpr_target,
    N_cal_values     = N_cal_values,
    # Grid range validated for L_eff threshold-independence
    c1_range = range(c1_grid),
    c2_range = range(c2_grid)
  )
)
```

::: {.callout-important}
## Shared `.rds` With C2

Both this document (C3) and `calibration_survival_fourway.qmd` (C2)
save to `calibration_survival_leff.rds`.
**C2 is the primary source** (up to 20K replicates per N).
C3 uses fewer replicates and is designed as a faster grid sweep.
If both are rendered, the last-rendered file wins.
To avoid overwriting C2's calibration, set `recalibrate = FALSE`
in this document or render C2 after C3.
:::


## Corrected Baseline

```{r corrected-baseline}
L_eff_main <- C_fit * (n_sample / n_min_val)^alpha_fit
P1_cox <- p_detect_glm(1.0, d_dgm, 1.25, 1.0)
fpr_corr <- 1 - (1 - P1_cox)^L_eff_main

cat(sprintf("=== Corrected Baseline (N=%d, c1=1.25, c2=1.0) ===\n",
            n_sample))
cat(sprintf("  L_eff:         %.2f\n", L_eff_main))
cat(sprintf("  P1 (uncorr):   %.3f\n", P1_cox))
cat(sprintf("  FPR (corr):    %.3f\n", fpr_corr))
cat(sprintf("  Sim FPR (Cox): %.3f\n", mean(cox_h0$any.H)))
cat(sprintf("  Sim FPR (IRR): %.3f\n", mean(irr_h0$any.H)))
cat(sprintf("  Sim/Corr:      %.2f\n",
            mean(cox_h0$any.H) / max(fpr_corr, 0.001)))
```


# Part IV: Null Threshold Grid {#sec-null-grid}

Systematic evaluation of $(c_1, c_2)$ pairs under $H_0$ at
$N = `r n_sample`$ to identify the region where
FPR $\le$ `r 100 * fpr_target`%.

Grid defined in the setup chunk: `c1_grid` and `c2_grid`.

## Grid Design

```{r null-grid-config}
#| code-fold: false

grid_pairs <- expand.grid(c1 = c1_grid, c2 = c2_grid,
                           stringsAsFactors = FALSE)
grid_pairs <- grid_pairs[grid_pairs$c2 <= grid_pairs$c1, ]

nsims_grid <- min(nsims_null, 500L)
N_grid     <- n_sample

cat(sprintf("Null grid: %d (c1,c2) pairs x %d sims = %d tasks at N = %d\n",
    nrow(grid_pairs), nsims_grid,
    nrow(grid_pairs) * nsims_grid, N_grid))
cat(sprintf("(c1 range: %.2f--%.2f, c2 range: %.2f--%.2f)\n",
    min(c1_grid), max(c1_grid), min(c2_grid), max(c2_grid)))
```


## Run Grid Simulations

```{r run-null-grid}
t0 <- proc.time()[3]

grid_tasks <- do.call(rbind, lapply(seq_len(nrow(grid_pairs)), function(g) {
  data.frame(
    grid_id = g, sim = seq_len(nsims_grid),
    c1 = grid_pairs$c1[g], c2 = grid_pairs$c2[g],
    stringsAsFactors = FALSE
  )
}))

cat(sprintf("Running %d null grid tasks...\n", nrow(grid_tasks)))

grid_results <- foreach(
  i = seq_len(nrow(grid_tasks)),
  .combine = rbind, .errorhandling = "remove",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table"),
    seed = TRUE)
) %dofuture% {
  sim_id <- grid_tasks$sim[i]
  c1_val <- grid_tasks$c1[i]
  c2_val <- grid_tasks$c2[i]

  fs_params_grid <- modifyList(fs_params_cox, list(
    hr.threshold = c1_val,
    hr.consistency = c2_val
  ))

  res <- run_simulation_analysis(
    sim_id = sim_id, dgm = dgm_null,
    n_sample = N_grid,
    analysis_time = sim_config$analysis_time,
    cens_adjust = sim_config$cens_adjust,
    confounders_base = confounders_base,
    cox_formula_adj = survival::Surv(y_sim, event_sim) ~ treat_sim + z1 + z2 + z3,
    n_add_noise = 0L,
    run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
    fs_params = fs_params_grid, verbose = FALSE
  )

  fs_row <- res[res$analysis == "FS", ]
  data.frame(
    sim = sim_id, c1 = c1_val, c2 = c2_val,
    any_H = if (nrow(fs_row) > 0) fs_row$any.H[1] else 0L,
    stringsAsFactors = FALSE
  )
}

timings$null_grid <- proc.time()[3] - t0
cat(sprintf("Null grid: %.1f min | %d results\n",
    timings$null_grid / 60, nrow(grid_results)))
```


# Part V: Results {#sec-results}

## FPR Table

```{r grid-fpr-table}
#| tbl-cap: !expr 'sprintf("Simulated FPR under H0 by (c1, c2).  Green: FPR <= 10%%.  Yellow: 10-%.0f%%.  Red: > %.0f%%.", 100 * fpr_target, 100 * fpr_target)'

grid_dt <- data.table(grid_results)
grid_fpr <- grid_dt[, .(
  FPR = mean(any_H, na.rm = TRUE),
  n_found = sum(any_H, na.rm = TRUE),
  n_sims = .N
), by = .(c1, c2)]

# Add approximations
grid_fpr[, P1 := mapply(function(c1v, c2v) {
  p_detect_glm(1.0, d_dgm, c1v, c2v)
}, c1, c2)]

# L_eff corrected
L_eff_N <- C_fit * (N_grid / n_min_val)^alpha_fit
grid_fpr[, FPR_corr := 1 - (1 - P1)^L_eff_N]
grid_fpr[, Sim_over_P1   := round(FPR / pmax(P1, 0.001), 2)]
grid_fpr[, Sim_over_Corr := round(FPR / pmax(FPR_corr, 0.001), 2)]

gt(grid_fpr[, .(c1, c2,
                `Sim FPR` = round(FPR, 3),
                `Found/Total` = sprintf("%d/%d", n_found, n_sims),
                `P1 (uncorr)` = round(P1, 3),
                `FPR (corr)` = round(FPR_corr, 3),
                `Sim/P1` = Sim_over_P1,
                `Sim/Corr` = Sim_over_Corr)]) |>
  tab_header(
    title = md("**Simulated FPR Under H0 by Threshold Pair**"),
    subtitle = sprintf("N=%d, %d sims/pair, n.min=%d, pcons=%.2f, L_eff=%.2f",
                        N_grid, nsims_grid, n_min_val, pcons, L_eff_N)
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `Sim FPR`,
      rows = `Sim FPR` <= 0.10)) |>
  tab_style(style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(columns = `Sim FPR`,
      rows = `Sim FPR` > 0.10 & `Sim FPR` <= fpr_target)) |>
  tab_style(style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = `Sim FPR`,
      rows = `Sim FPR` > fpr_target)) |>
  tab_source_note(sprintf(
    "Green: FPR <= 10%%. Yellow: 10-%.0f%%. Red: > %.0f%%.",
    100 * fpr_target, 100 * fpr_target))
```


## $L_{\text{eff}}$ Consistency Across Threshold Pairs

If $L_{\text{eff}}$ is truly a property of the candidate-generation
process (confounders, GRF, LASSO, maxk, n.min) and not the
threshold pair, the implied $L_{\text{eff}}$ should be
approximately constant across $(c_1, c_2)$ at the same $N$.

```{r leff-consistency}
#| tbl-cap: "Implied L_eff at each (c1, c2) pair.  If L_eff is threshold-independent, values should cluster near the calibrated value."

# Compute implied L_eff for each grid pair
grid_fpr[, implied_L_eff := ifelse(
  P1 > 0 & FPR > 0 & FPR < 1,
  log(1 - FPR) / log(1 - P1),
  NA_real_
)]

leff_summary <- grid_fpr[!is.na(implied_L_eff), .(
  `Median` = round(median(implied_L_eff), 2),
  `Min`    = round(min(implied_L_eff), 2),
  `Max`    = round(max(implied_L_eff), 2),
  `SD`     = round(sd(implied_L_eff), 3),
  `CV`     = sprintf("%.0f%%", 100 * sd(implied_L_eff) /
                      median(implied_L_eff)),
  `n_pairs` = .N
)]

cat(sprintf(
  "Calibrated L_eff at N=%d: %.2f\n", N_grid, L_eff_N))
cat(sprintf(
  "Implied L_eff across %d threshold pairs: median=%.2f, range=[%.2f, %.2f]\n",
  leff_summary$n_pairs, leff_summary$Median,
  leff_summary$Min, leff_summary$Max))

gt(grid_fpr[!is.na(implied_L_eff),
  .(c1, c2,
    `Sim FPR` = sprintf("%.1f%%", 100 * FPR),
    `P1` = sprintf("%.4f", P1),
    `Implied L_eff` = round(implied_L_eff, 2),
    `Calibrated L_eff` = round(L_eff_N, 2),
    `Ratio` = round(implied_L_eff / L_eff_N, 2)
  )]) |>
  tab_header(
    title = md("**L_eff Consistency Diagnostic**"),
    subtitle = sprintf("N=%d | Calibrated: L_eff = %.3f * (N/%d)^%.3f = %.2f",
                        N_grid, C_fit, n_min_val, alpha_fit, L_eff_N)
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = Ratio,
      rows = Ratio >= 0.7 & Ratio <= 1.3)) |>
  tab_style(style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = Ratio,
      rows = Ratio < 0.7 | Ratio > 1.3)) |>
  tab_source_note(paste(
    "Ratio = Implied / Calibrated.",
    "Green: within +/-30%.  Red: outside +/-30%.",
    "Constant ratio supports threshold-independence."))
```

## FPR Heatmap

```{r grid-heatmap, fig.width=8, fig.height=6}
#| fig-cap: !expr 'sprintf("Simulated FPR heatmap by (c1, c2).  Values <= %.0f%% are in the acceptable region.", 100 * fpr_target)'

ggplot(grid_fpr, aes(x = factor(c1), y = factor(c2), fill = FPR)) +
  geom_tile(colour = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.1f%%", 100 * FPR)),
            size = 3.5, colour = "black") +
  scale_fill_gradient2(
    low = "#E8F5E9", mid = "#FFF9C4", high = "#E53935",
    midpoint = fpr_target,
    limits = c(0, max(grid_fpr$FPR, 0.20)),
    name = "FPR"
  ) +
  labs(x = expression(c[1] ~ "(screening threshold)"),
       y = expression(c[2] ~ "(consistency threshold)"),
       title = "Simulated FPR Under H0: GBSG Survival DGM",
       subtitle = sprintf("N=%d, %d sims/pair, n.min=%d",
                           N_grid, nsims_grid, n_min_val)) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        axis.text = element_text(size = 11))
```


## FPR Sensitivity to $c_2$ (Fixing $c_1$)

```{r grid-c2-profiles, fig.width=9, fig.height=5}
#| fig-cap: "FPR vs c2 for each fixed c1.  The target band is highlighted."

ggplot(grid_fpr, aes(x = c2, y = FPR,
                      colour = factor(c1), group = factor(c1))) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0.10, linetype = "dashed", colour = "darkgreen") +
  geom_hline(yintercept = fpr_target, linetype = "dashed", colour = "orange") +
  annotate("rect", xmin = -Inf, xmax = Inf,
           ymin = 0.10, ymax = fpr_target,
           alpha = 0.15, fill = "gold") +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression(c[2] ~ "(consistency threshold)"),
       y = "Simulated FPR",
       title = "FPR Sensitivity to Consistency Threshold",
       subtitle = sprintf("N=%d, each line is a fixed c1", N_grid),
       colour = expression(c[1])) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


## FPR Sensitivity to $c_1$ (Fixing $c_2$)

```{r grid-c1-profiles, fig.width=9, fig.height=5}
#| fig-cap: "FPR vs c1 for each fixed c2.  Shows the screening threshold's independent effect."

ggplot(grid_fpr, aes(x = c1, y = FPR,
                      colour = factor(c2), group = factor(c2))) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0.10, linetype = "dashed", colour = "darkgreen") +
  geom_hline(yintercept = fpr_target, linetype = "dashed", colour = "orange") +
  annotate("rect", xmin = -Inf, xmax = Inf,
           ymin = 0.10, ymax = fpr_target,
           alpha = 0.15, fill = "gold") +
  scale_colour_brewer(palette = "Set2") +
  labs(x = expression(c[1] ~ "(screening threshold)"),
       y = "Simulated FPR",
       title = "FPR Sensitivity to Screening Threshold",
       subtitle = sprintf("N=%d, each line is a fixed c2", N_grid),
       colour = expression(c[2])) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


## Approximation Accuracy Across the Grid

```{r grid-approx-scatter, fig.width=8, fig.height=6}
#| fig-cap: "Approximation vs simulation across the (c1,c2) grid.  Corrected points (triangles) should cluster near the diagonal."

ggplot(grid_fpr, aes(x = FPR)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              colour = "grey50") +
  geom_point(aes(y = P1, colour = "Uncorrected P1"),
             size = 2.5, alpha = 0.7) +
  geom_point(aes(y = FPR_corr, colour = "L_eff Corrected"),
             size = 2.5, alpha = 0.7, shape = 17) +
  scale_colour_manual(values = c("Uncorrected P1" = "firebrick",
                                  "L_eff Corrected" = "steelblue")) +
  coord_equal(xlim = c(0, max(grid_fpr$FPR, 0.25)),
              ylim = c(0, max(grid_fpr$P1, 0.25))) +
  labs(x = "Simulated FPR", y = "Approximated FPR",
       title = "Approximation Accuracy Across Threshold Grid",
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Acceptable Region (FPR $\le$ `r 100 * fpr_target`%)

```{r grid-acceptable}
#| tbl-cap: !expr 'sprintf("Threshold pairs with simulated FPR <= %.0f%%", 100 * fpr_target)'

acceptable_grid <- grid_fpr[FPR <= fpr_target]
setorder(acceptable_grid, FPR)

if (nrow(acceptable_grid) > 0) {
  gt(acceptable_grid[, .(c1, c2,
    `Sim FPR` = sprintf("%.1f%%", 100 * FPR),
    `P1 (uncorr)` = sprintf("%.1f%%", 100 * P1),
    `FPR (corr)` = sprintf("%.1f%%", 100 * FPR_corr))]) |>
    tab_header(
      title = md(sprintf(
        "**Acceptable Threshold Configurations (FPR <= %.0f%%)**",
        100 * fpr_target)),
      subtitle = sprintf("N=%d, GBSG survival DGM, n.min=%d",
                          N_grid, n_min_val)
    ) |>
    tab_source_note("Paper defaults: c1=1.25, c2=1.0")
} else {
  cat(sprintf("No configurations achieved FPR <= %.0f%% in this grid.\n",
              100 * fpr_target))
}
```


# Summary {#sec-summary}

```{r summary}
cat("=== Null Operating Characteristics Summary ===\n\n")
cat(sprintf("DGM: GBSG survival, N=%d, n.min=%d, pcons=%.2f\n",
            N_grid, n_min_val, pcons))
cat(sprintf("L_eff model: %.3f * (N/%d)^%.3f\n",
            C_fit, n_min_val, alpha_fit))
cat(sprintf("L_eff at N=%d: %.2f\n\n", N_grid, L_eff_N))

cat("--- Paper Defaults (c1=1.25, c2=1.0) ---\n")
paper_row <- grid_fpr[c1 == 1.25 & c2 == 1.00]
if (nrow(paper_row) > 0) {
  cat(sprintf("  Sim FPR: %.1f%%\n", 100 * paper_row$FPR))
  cat(sprintf("  P1:      %.1f%%\n", 100 * paper_row$P1))
  cat(sprintf("  Corr:    %.1f%%\n\n", 100 * paper_row$FPR_corr))
}

cat(sprintf("--- Configurations with FPR in 8-%.0f%% ---\n",
            100 * fpr_target))
target_range <- grid_fpr[FPR >= 0.08 & FPR <= fpr_target]
if (nrow(target_range) > 0) {
  setorder(target_range, -FPR)
  for (r in seq_len(nrow(target_range))) {
    cat(sprintf("  c1=%.2f, c2=%.2f: FPR=%.1f%%\n",
                target_range$c1[r], target_range$c2[r],
                100 * target_range$FPR[r]))
  }
} else {
  cat("  None in this range.\n")
}

cat("\n--- All (c1,c2) with FPR <= 10% ---\n")
conservative <- grid_fpr[FPR <= 0.10]
if (nrow(conservative) > 0) {
  setorder(conservative, c1, c2)
  for (r in seq_len(nrow(conservative))) {
    cat(sprintf("  c1=%.2f, c2=%.2f: FPR=%.1f%%\n",
                conservative$c1[r], conservative$c2[r],
                100 * conservative$FPR[r]))
  }
}

cat(sprintf("\n--- Note on L_eff ---\n"))
cat("For survival, L_eff < 1 is common because the events-based\n")
cat("filters (d0.min, d1.min) aggressively prune candidates.\n")
cat("This means the paper defaults (c1=1.25, c2=1.0) are already\n")
cat("conservative for survival endpoints.\n")
```


# Save All Simulation Results {#sec-save}

```{r save-all-results}
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

sim_bundle <- list(

  # Metadata
  params = list(
    nsims_null = nsims_null,
    nsims_grid = nsims_grid,
    n_sims_cal = n_sims_cal,
    sim_seed = sim_seed,
    N_grid = N_grid,
    N_cal_values = N_cal_values,
    sim_config = sim_config,
    fs_params_base = fs_params_base,
    fs_params_cox = fs_params_cox,
    fs_params_irr = fs_params_irr,
    c1_grid = c1_grid,
    c2_grid = c2_grid,
    n_min_val = n_min_val,
    prop_cens = prop_cens,
    p_harm = p_harm,
    hr_h1 = hr_h1,
    fpr_target = fpr_target,
    timestamp = Sys.time(),
    R_version = R.version.string,
    forestsearch_version = tryCatch(
      as.character(utils::packageVersion("forestsearch")),
      error = function(e) "unknown")
  ),

  # DGM
  dgm_null = dgm_null,

  # Paper-defaults null sims (Part II)
  results_null_cox = results_null_cox,
  results_null_irr = results_null_irr,

  # L_eff calibration (Part III)
  leff = list(
    C = C_fit,
    alpha = alpha_fit,
    n_min = n_min_val,
    cal_df = cal_df
  ),

  # Grid sweep results (Part IV)
  grid_results = grid_results,
  grid_fpr = grid_fpr,

  # Timing
  timings = timings
)

rds_path <- file.path(output_dir,
  sprintf("calibration_survival_leff_grid_%s.rds",
          format(Sys.Date(), "%Y%m%d")))
saveRDS(sim_bundle, rds_path)

cat(sprintf("=== All results saved ===\n"))
cat(sprintf("  File: %s\n", rds_path))
cat(sprintf("  Size: %.1f MB\n", file.size(rds_path) / 1024^2))
cat(sprintf("  Contents:\n"))
cat(sprintf("    results_null_cox: %d rows\n", nrow(results_null_cox)))
cat(sprintf("    results_null_irr: %d rows\n", nrow(results_null_irr)))
cat(sprintf("    cal_df:           %d rows (N = %s)\n",
            nrow(cal_df), paste(cal_df$N, collapse = ", ")))
cat(sprintf("    grid_results:     %d rows (%d configs x %d sims)\n",
            nrow(grid_results), nrow(grid_pairs), nsims_grid))
cat(sprintf("    grid_fpr:         %d (c1,c2) pairs\n", nrow(grid_fpr)))
cat(sprintf("    L_eff:            C=%.4f, alpha=%.4f\n",
            C_fit, alpha_fit))
```


## How to Reload

```{r reload-example}
#| eval: false
#| code-fold: false

res <- readRDS(latest_rds("calibration_survival_leff_grid"))

# Paper-defaults FPR
cox_h0 <- res$results_null_cox[res$results_null_cox$analysis == "FS", ]
cat("Cox H0 FPR:", mean(cox_h0$any.H), "\n")

# L_eff calibration
res$leff$C            # fitted C
res$leff$alpha        # fitted alpha
res$leff$cal_df       # raw calibration points

# Grid sweep
res$grid_fpr          # FPR by (c1, c2)
res$grid_fpr[FPR <= res$params$fpr_target]   # acceptable configs

# Config
res$params$nsims_null       # how many sims
res$params$c1_grid          # which thresholds swept
res$params$fpr_target       # FPR target
```


# Timing and Session Info {#sec-session}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Stage = c("DGM", "Cox H0 (paper defaults)",
            "IRR H0 (paper defaults)",
            "L_eff Calibration (multi-N)",
            "Null Grid Sweep",
            "Total"),
  Seconds = c(timings$dgm, timings$null_cox, timings$null_irr,
              timings$null_cal, timings$null_grid, timings$total),
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(
    title = "Computational Timing Summary",
    subtitle = sprintf(
      "%d (c1,c2) pairs x %d sims + %d cal sims x %d N | %d workers",
      nrow(grid_pairs), nsims_grid, n_sims_cal,
      length(N_cal_values), n_workers)
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)",
             Minutes = "Time (min)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total")
  )
```

```{r session-info}
plan(sequential)
sessionInfo()
```
