---
title: "Threshold Calibration for Binary Outcomes via the GLM Approximation"
subtitle: "Extending the Section 2.1 Framework to Logistic Regression"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 7
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")
# ~ 500/hr on mac
nsims       <- 1000L
sim_seed    <- 8316951L
n_workers   <- ceiling(0.95*max(1L, parallel::detectCores(logical = FALSE) - 1L))

# Output directory for .rds results
output_dir  <- "_data"

# Reference subgroup size: n.min (most conservative FPR)
n_min_val   <- 60L
```


# The GLM Approximation {#sec-approx}

## Parallel to Section 2.1

For the Cox model, the Section 2.1 approximation uses the
asymptotic normality of the score statistic:
$\hat\beta_s \approx (4/d) \tilde L_d(0) \sim N(\beta, 4/d)$.
For split-halves: $\hat\beta^1_s, \hat\beta^2_s \sim N(\beta, 8/d)$
independently.

For a canonical GLM with treatment as the sole covariate, the
MLE satisfies exactly the same structure via Fisher information.
With balanced arms ($n_1 = n_0 = n_H/2$):

$$
\text{Var}(\hat\beta_{\text{full}}) = \frac{4}{d_{\text{eff}}},
\qquad
\text{Var}(\hat\beta_{\text{half}}) = \frac{8}{d_{\text{eff}}}
$$

where $d_{\text{eff}}$ is the effective information:

| Family | $d_{\text{eff}}$ | Analogous to |
|---|---|---|
| Cox | $d = n_H(1-p_c)$ | Number of events |
| Binomial (logit) | $n_H \bar p(1-\bar p)$ | Bernoulli information |
| Poisson (log) | $D$ (total events) | Number of events |
| Gaussian (identity) | $n_H / \sigma^2$ | Precision-weighted n |

The detection probability for a **specific subgroup** with
true effect $\beta = \log(\theta^\dagger(H))$ is:

$$
p(c_1, c_2; d_{\text{eff}}, \theta^\dagger(H)) =
P(W_1 + W_2 \ge 2\log c_1, \;\min(W_1, W_2) \ge \log c_2)
$$

with $W_1, W_2 \stackrel{\text{iid}}{\sim}
N(\log \theta^\dagger(H), \; 8/d_{\text{eff}})$.

This gives:

- **Type-I error** (per-subgroup): $p(c_1, c_2; d_{\text{eff}}, 1)$
  with $\beta = 0$
- **Power**: $p(c_1, c_2; d_{\text{eff}}, \theta^\dagger(H))$
  for the true harm subgroup

The thresholds are then chosen so that
$p(c_1, c_2; d_{\text{eff}}, 1) \approx 10\%$ and
$p(c_1, c_2; d_{\text{eff}}, \theta^\dagger(H))$ is
acceptably high, and **validated by simulation**.


## Implementation

```{r approx-function}
#| code-fold: false

# Use compute_detection_probability_glm() and d_eff_*() helpers
# from forestsearch (installed in R/truefind_asymptotic_glm.R).
# These implement equation (3) extended to GLM via d_eff.
#
# Shorthand wrapper for concise calls in this document:
p_detect <- function(theta, d_eff, c1, c2) {
  compute_detection_probability_glm(
    theta = theta, d_eff = d_eff, c1 = c1, c2 = c2,
    effect_scale = "ratio", method = "monte_carlo"
  )
}
```


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)

# L_eff save/load helpers
#source(here::here("../R", "leff_calibration_io.R"))
plan(multisession, workers = n_workers)
cat("Workers:", n_workers, "| Sims per config:", nsims, "\n")
```

```{r timing-init}
#| include: false
timings <- list()
t_vignette_start <- proc.time()
```


# Part I: Approximation-Guided Threshold Selection {#sec-guide}

## Survival Benchmark

Reproduce the survival result: with $d = 33$ (n = 60, 45
percent censoring), the defaults c1 = 1.25, c2 = 1.0 give
$p \approx 10\%$ type-I error and $\approx 80\%$ power at
$\theta^\dagger(H) = 1.94$.

```{r survival-benchmark}
#| tbl-cap: "Survival benchmark: Section 2.1 reproduced"

surv_bench <- data.frame(
  n_sg = c(60, 80, 100),
  stringsAsFactors = FALSE
)
surv_bench$d <- round(d_eff_survival(surv_bench$n_sg, prop_cens = 0.45))
surv_bench$`P(theta=1.0)` <- sapply(surv_bench$d, function(d)
  round(p_detect(1.0, d, 1.25, 1.0), 3))
surv_bench$`P(theta=0.75)` <- sapply(surv_bench$d, function(d)
  round(p_detect(0.75, d, 1.25, 1.0), 3))
surv_bench$`P(theta=2.0)` <- sapply(surv_bench$d, function(d)
  round(p_detect(2.0, d, 1.25, 1.0), 3))

gt(surv_bench) |>
  tab_header(
    title = "Survival: c1=1.25, c2=1.0 (Paper Defaults)",
    subtitle = "Reproduces Figure 2 and Table 1 footnotes"
  ) |>
  cols_label(n_sg = "n(H)", d = "d (events)",
    `P(theta=1.0)` = "P(OR=1.0)",
    `P(theta=0.75)` = "P(OR=0.75)",
    `P(theta=2.0)` = "Power(OR=2)")
```


## Binary: d_eff Is Lower, So the Thresholds Need Adjustment

For binary outcomes with $\bar p = 0.30$:
$d_{\text{eff}} = n_H \cdot 0.30 \cdot 0.70 = 0.21 n_H$.
For $n_H = 60$: $d_{\text{eff}} = 12.6$ (vs $d = 33$ for survival).

```{r binary-default-check}
#| tbl-cap: "Binary: survival defaults applied to binary d_eff"

bin_check <- data.frame(
  n_sg = c(60, 100, 150, 200, 240),
  p = 0.30,
  stringsAsFactors = FALSE
)
bin_check$d_eff <- round(d_eff_binary(bin_check$n_sg, p_event = 0.30), 1)

# Survival defaults: c1=1.25, c2=1.0
bin_check$P_10 <- sapply(bin_check$d_eff, function(d)
  round(p_detect(1.0, d, 1.25, 1.0), 3))
bin_check$P_075 <- sapply(bin_check$d_eff, function(d)
  round(p_detect(0.75, d, 1.25, 1.0), 3))
bin_check$Power_20 <- sapply(bin_check$d_eff, function(d)
  round(p_detect(2.0, d, 1.25, 1.0), 3))

gt(bin_check) |>
  tab_header(
    title = "Binary with Survival Defaults (c1=1.25, c2=1.0)",
    subtitle = "Per-subgroup P(theta=1.0) is the type-I error analog"
  ) |>
  cols_label(n_sg = "n(H)", p = "p", d_eff = "d_eff",
    P_10 = "P(OR=1.0)", P_075 = "P(OR=0.75)",
    Power_20 = "Power(OR=2)")
```


## Threshold Search: Find c1, c2 Where P(1.0) ~ 10%

```{r threshold-search}
#| fig-height: 8
#| fig-cap: "Power curves for binary outcomes (p=0.30) under candidate threshold pairs. Horizontal lines: 10% (type-I error target) and 80% (power target)."

or_seq <- seq(0.5, 3.0, by = 0.05)

# Candidate (c1, c2) pairs to evaluate
pairs <- list(
  list(c1 = 1.25, c2 = 1.00, label = "c1=1.25, c2=1.00 (survival)"),
  list(c1 = 1.50, c2 = 1.00, label = "c1=1.50, c2=1.00"),
  list(c1 = 1.25, c2 = 1.10, label = "c1=1.25, c2=1.10"),
  list(c1 = 1.25, c2 = 1.25, label = "c1=1.25, c2=1.25 (c2=c1)"),
  list(c1 = 1.50, c2 = 1.25, label = "c1=1.50, c2=1.25"),
  list(c1 = 1.50, c2 = 1.50, label = "c1=1.50, c2=1.50 (c2=c1)")
)

n_sg_vals <- c(60, 150, 240)

curve_data <- do.call(rbind, lapply(pairs, function(pr) {
  do.call(rbind, lapply(n_sg_vals, function(n) {
    d_eff <- d_eff_binary(n, p_event = 0.30)
    probs <- sapply(or_seq, function(or) p_detect(or, d_eff, pr$c1, pr$c2))
    data.frame(OR = or_seq, power = probs,
               Config = pr$label,
               n_sg = sprintf("n(H) = %d (d_eff = %.0f)", n, d_eff))
  }))
}))

ggplot(curve_data, aes(x = OR, y = power, colour = Config)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = c(0.10, 0.80), linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = 1.0, linetype = "dotted", colour = "grey70") +
  facet_wrap(~ n_sg, ncol = 1) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression(theta^dagger*(H) ~ "(true OR in subgroup)"),
       y = "P(detect H)",
       title = "GLM Approximation: Binary Outcomes (p = 0.30)",
       subtitle = "Per-subgroup detection probability",
       colour = "Thresholds") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", plot.title = element_text(face = "bold"),
        legend.text = element_text(size = 9))
```


## Approximation Summary Table

```{r approx-summary}
#| tbl-cap: "Per-subgroup type-I error and power for candidate threshold pairs (p=0.30)"

n_configs <- c(60, 100, 150, 200, 240)

approx_grid <- do.call(rbind, lapply(pairs, function(pr) {
  do.call(rbind, lapply(n_configs, function(n) {
    d_eff <- d_eff_binary(n, p_event = 0.30)
    data.frame(
      Config = pr$label, n_sg = n, d_eff = round(d_eff, 1),
      c1 = pr$c1, c2 = pr$c2,
      `P(1.0)` = round(p_detect(1.0, d_eff, pr$c1, pr$c2), 3),
      `P(0.75)` = round(p_detect(0.75, d_eff, pr$c1, pr$c2), 3),
      `Power(2.0)` = round(p_detect(2.0, d_eff, pr$c1, pr$c2), 3),
      check.names = FALSE)
  }))
}))

# Show configurations where P(1.0) ~ 10%
target <- approx_grid[approx_grid$`P(1.0)` <= 0.15 &
                       approx_grid$`P(1.0)` >= 0.03, ]

gt(target) |>
  tab_header(
    title = "Threshold Pairs with P(OR=1.0) in 3-15% Range",
    subtitle = "Candidates for simulation validation"
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `P(1.0)`,
      rows = `P(1.0)` >= 0.05 & `P(1.0)` <= 0.12)) |>
  cols_label(Config = "Thresholds", n_sg = "n(H)",
    d_eff = "d_eff", `P(1.0)` = "P(OR=1)",
    `P(0.75)` = "P(OR=0.75)", `Power(2.0)` = "Power(OR=2)")
```


# Part II: Simulation Validation {#sec-sim}

## Binary DGM

```{r sim-dgm}
#| code-fold: false

sim_binary <- function(n, seed, has_hte = TRUE) {
  set.seed(seed)
  df <- data.frame(
    id = seq_len(n), treat = rbinom(n, 1, 0.5),
    bm1 = as.factor(rbinom(n, 1, 0.70)),
    bm2 = as.factor(rbinom(n, 1, 0.50)),
    age = round(rnorm(n, 55, 10)),
    ecog = as.factor(sample(0:1, n, replace = TRUE, prob = c(0.6, 0.4)))
  )
  lp <- with(df, qlogis(0.30) + 0.02 * (age - 55) + 0.3 * (ecog == 1))
  if (has_hte) {
    lp <- lp + with(df,
      log(2.0) * treat * (bm1 == 0) + log(0.4) * treat * (bm1 == 1))
  }
  df$progressed <- rbinom(n, 1, plogis(lp))
  df$flag_harm  <- as.integer(df$bm1 == 0)
  df
}
```


## Simulation Engine

```{r sim-engine}
#| code-fold: false

run_sim <- function(sim_num, n_total, has_hte, c1, c2, pcons) {
  seed <- sim_seed + sim_num
  df <- sim_binary(n_total, seed, has_hte)
  fs_error <- NA_character_
  fs <- tryCatch(
    forestsearch(
      df.analysis = df,
      confounders.name = c("bm1", "bm2", "age", "ecog"),
      outcome.name = "progressed", event.name = "progressed",
      treat.name = "treat", id.name = "id",
      outcome_type = "binary", effect_measure = "OR",
      adverse_outcome = TRUE,
      hr.threshold = c1, hr.consistency = c2,
      pconsistency.threshold = pcons, fs.splits = 400,
      n.min = 60, maxk = 2,
      use_lasso = TRUE, use_grf = TRUE, use_twostage = TRUE,
      is.RCT = TRUE, seedit = seed, details = FALSE, plot.sg = FALSE,
      parallel_args = list(plan = "sequential", workers = 1,
                           show_message = FALSE)
    ),
    error = function(e) { fs_error <<- conditionMessage(e); NULL }
  )
  found <- !is.null(fs) && !is.null(fs$sg.harm) && length(fs$sg.harm) > 0
  out <- data.table(sim = sim_num, n_total = n_total, has_hte = has_hte,
                    c1 = c1, c2 = c2, pcons = pcons,
                    any_found = as.integer(found), error_msg = fs_error)
  if (found && !is.null(fs$df.est) && "treat.recommend" %in% names(fs$df.est)) {
    sg_ind <- fs$df.est$treat.recommend; flag <- df$flag_harm
    tp <- sum(sg_ind == 0 & flag == 1); fp <- sum(sg_ind == 0 & flag == 0)
    tn <- sum(sg_ind == 1 & flag == 0); fn <- sum(sg_ind == 1 & flag == 1)
    out$sens <- if ((tp+fn)>0) tp/(tp+fn) else NA_real_
    out$spec <- if ((tn+fp)>0) tn/(tn+fp) else NA_real_
    out$ppv  <- if ((tp+fp)>0) tp/(tp+fp) else NA_real_
    out$size_H <- sum(sg_ind == 0)
  } else { out$sens <- out$spec <- out$ppv <- NA_real_; out$size_H <- NA_integer_ }
  out
}
```


## Run Simulations

Validate the approximation-selected threshold pairs.
Fix pconsistency = 0.90.

```{r run-sims}
t0 <- proc.time()[3]

# Threshold pairs to validate (from approximation analysis)
sim_pairs <- list(
  list(c1 = 1.25, c2 = 1.00),
  list(c1 = 1.25, c2 = 1.10),
  list(c1 = 1.35, c2 = 1.00),
  list(c1 = 1.50, c2 = 1.00),
  list(c1 = 1.50, c2 = 1.25),
  list(c1 = 1.50, c2 = 1.35),
  list(c1 = 1.60, c2 = 1.35),
  list(c1 = 1.60, c2 = 1.40),
  list(c1= 1.70, c2 =1.50)
)

n_values <- c(300, 500, 700)
pcons <- 0.90

tasks <- do.call(rbind, lapply(sim_pairs, function(pr) {
  expand.grid(sim = seq_len(nsims), c1 = pr$c1, c2 = pr$c2,
              n_total = n_values, has_hte = c(FALSE, TRUE),
              stringsAsFactors = FALSE)
}))

cat(sprintf("Simulation: %d threshold pairs x %d N x 2 hypotheses x %d sims = %d tasks\n",
    length(sim_pairs), length(n_values), nsims, nrow(tasks)))

res <- foreach(i = seq_len(nrow(tasks)), .errorhandling = "pass",
  .options.future = list(packages = c("forestsearch", "data.table"))
) %dofuture% {
  run_sim(tasks$sim[i], tasks$n_total[i], tasks$has_hte[i],
          tasks$c1[i], tasks$c2[i], pcons)
}
sim_dt <- rbindlist(res, fill = TRUE)
timings$simulation <- proc.time()[3] - t0
cat(sprintf("Simulation: %.1f min | %d results | %d errors\n",
    timings$simulation/60, nrow(sim_dt), sum(!is.na(sim_dt$error_msg))))
```


# Part III: Approximation vs Simulation {#sec-compare}

## Power Comparison (H1)

The approximation should agree well with simulated power,
since power measures detection of a **specific true subgroup**.

```{r power-comparison}
#| tbl-cap: "Approximation vs simulated power (H1: true OR_H = 2.0, true n(H) ~ 30% of N)"

oc_sim <- sim_dt[, .(
  detect = mean(any_found, na.rm = TRUE),
  mean_sens = mean(sens, na.rm = TRUE),
  mean_ppv = mean(ppv, na.rm = TRUE)
), by = .(n_total, has_hte, c1, c2)]

# Compute approximation for comparison
oc_sim[, n_sg := n_min_val]
oc_sim[, d_eff := round(d_eff_binary(n_sg, p_event = 0.30), 1)]
oc_sim[, approx := mapply(function(hte, d, c1v, c2v) {
  theta <- if (hte) 2.0 else 1.0
  p_detect(theta, d, c1v, c2v)
}, has_hte, d_eff, c1, c2)]

# Power comparison
h1_comp <- oc_sim[has_hte == TRUE]

gt(h1_comp[, .(N = n_total, c1, c2, d_eff,
               `Approx Power` = round(approx, 3),
               `Sim Power` = round(detect, 3),
               `Sim/Approx` = round(detect / pmax(approx, 0.001), 2),
               Sens = round(mean_sens, 3),
               PPV = round(mean_ppv, 3))]) |>
  tab_header(
    title = "Power: Approximation vs Simulation (H1, OR_H = 2.0)",
    subtitle = sprintf("pcons = %.2f | %d sims per config", pcons, nsims)
  ) |>
  tab_row_group(label = "N = 800", rows = N == 800) |>
  tab_row_group(label = "N = 500", rows = N == 500) |>
  tab_row_group(label = "N = 200", rows = N == 200) |>
  tab_source_note("Sim/Approx near 1.0 = approximation is accurate for power")
```


## Type-I Error Comparison (H0)

The per-subgroup approximation $p(c_1, c_2; d_{\text{eff}}, 1)$
gives the probability that a **specific** non-harmful subgroup
passes.  The simulated FPR reflects the overall procedure
(searching over candidates).  As in the paper, we check
whether the overall FPR is controlled when the per-subgroup
rate is calibrated to ~10%.

```{r fpr-comparison}
#| tbl-cap: "Type-I error: per-subgroup approximation vs simulated procedure FPR"

h0_comp <- oc_sim[has_hte == FALSE]

gt(h0_comp[, .(N = n_total, c1, c2, d_eff,
               `Approx P(1.0)` = round(approx, 3),
               `Sim FPR` = round(detect, 3),
               `Sim/Approx` = round(detect / pmax(approx, 0.001), 2))]) |>
  tab_header(
    title = "Type-I Error: Approximation vs Simulation (H0)",
    subtitle = "Per-subgroup P(OR=1) vs procedure-level FPR"
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `Sim FPR`,
      rows = `Sim FPR` <= 0.12)) |>
  tab_style(style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(columns = `Sim FPR`,
      rows = `Sim FPR` > 0.12 & `Sim FPR` <= 0.20)) |>
  tab_style(style = cell_fill(color = "#FFEBEE"),
    locations = cells_body(columns = `Sim FPR`,
      rows = `Sim FPR` > 0.20)) |>
  tab_row_group(label = "N = 800", rows = N == 800) |>
  tab_row_group(label = "N = 500", rows = N == 500) |>
  tab_row_group(label = "N = 200", rows = N == 200) |>
  tab_source_note(
    "Green: Sim FPR <= 12%. Yellow: 12-20%. Red: > 20%."
  )
```


## Visual Comparison

```{r scatter-comparison}
#| fig-height: 7
#| fig-cap: "Approximation vs simulation for both H0 (circles) and H1 (triangles). Points near the diagonal indicate good agreement."

oc_sim$H <- ifelse(oc_sim$has_hte, "H1 (Power)", "H0 (FPR)")
oc_sim$config <- sprintf("c1=%.2f, c2=%.2f", oc_sim$c1, oc_sim$c2)

ggplot(oc_sim, aes(x = approx, y = detect)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  geom_point(aes(colour = config, shape = H), size = 3.5) +
  facet_wrap(~ paste("N =", n_total)) +
  scale_colour_brewer(palette = "Set1") +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(x = "Approximation (per-subgroup)", y = "Simulation (procedure)",
       title = "GLM Approximation vs Simulation",
       subtitle = "H1 (power) should agree; H0 (FPR) is validated empirically",
       colour = "Thresholds", shape = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right", plot.title = element_text(face = "bold"))
```


## Power vs FPR Tradeoff

```{r tradeoff}
#| fig-height: 5
#| fig-cap: "Simulated power vs simulated FPR. Each point is a (c1,c2) pair. The green region is the acceptable operating region."

tradeoff <- merge(
  oc_sim[has_hte == FALSE, .(n_total, c1, c2, FPR = detect)],
  oc_sim[has_hte == TRUE,  .(n_total, c1, c2, Power = detect)],
  by = c("n_total", "c1", "c2")
)
tradeoff$config <- sprintf("c1=%.2f\nc2=%.2f", tradeoff$c1, tradeoff$c2)

ggplot(tradeoff, aes(x = FPR, y = Power, label = config)) +
  geom_rect(aes(xmin = 0, xmax = 0.15, ymin = 0.50, ymax = 1.0),
    fill = "#E8F5E9", alpha = 0.3, inherit.aes = FALSE) +
  geom_point(aes(colour = sprintf("c1=%.2f, c2=%.2f", c1, c2)), size = 3) +
  geom_text(size = 2, nudge_y = -0.04, check_overlap = TRUE) +
  facet_wrap(~ paste("N =", n_total)) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Simulated FPR (H0)", y = "Simulated Power (H1, OR=2)",
       title = "Operating Characteristics by Threshold Pair",
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))
```


# Part IV: Corrected Null Approximation {#sec-correction}

The per-subgroup approximation $P_1$ models a single
pre-specified subgroup.  Under $H_0$, the ForestSearch
procedure evaluates many candidate subgroups, and the number
of viable candidates **grows with $N$** as more subgroup
combinations survive the `n.min` filter.  This explains the
widening gap between $P_1$ and the simulated FPR at larger $N$.

The corrected approximation accounts for this multiplicity:

$$
\text{FPR}_{\text{corrected}}(N) =
1 - (1 - P_1)^{L_{\text{eff}}(N)}
$$

where $L_{\text{eff}}(N)$ is the effective number of
approximately independent candidates evaluated by the procedure.


## Why $L_{\text{eff}}$ Grows with $N$

Three mechanisms drive the growth of $L_{\text{eff}}$:

1. **More subgroup combinations survive `n.min`**.
   At $N = 200$ with `n.min = 60`, a subgroup must comprise
   at least 30% of the sample.  Two-factor combinations (e.g.,
   bm1 = 0 AND bm2 = 0, prevalence $\approx 15\%$) have only
   $\approx 30$ patients and are excluded.  At $N = 1000$, the
   same combination has $\approx 150$ patients and enters the
   candidate set.

2. **GRF generates additional candidate cuts** on continuous
   variables (age) when more data is available.

3. **LASSO retains more variables** with increased power to
   detect spurious prognostic associations.

Key insight: $L_{\text{eff}}$ is a property of the candidate
generation process (confounders, GRF, LASSO, maxk, n.min)
--- **not** of the thresholds $(c_1, c_2)$.  The thresholds
determine $P_1$; $L_{\text{eff}}$ determines how many times
$P_1$ is "rolled."


## Implied $L_{\text{eff}}$ from Simulation

For each $(N, c_1, c_2)$ configuration, solve for the implied
$L_{\text{eff}}$:

$$
L_{\text{eff}} =
\frac{\log(1 - \text{FPR}_{\text{sim}})}
     {\log(1 - P_1)}
$$

```{r compute-leff}
#| tbl-cap: "Implied L_eff at each N.  If L_eff is a property of the candidate count, it should be approximately constant across threshold pairs at the same N."

h0_data <- oc_sim[has_hte == FALSE & detect > 0]
h0_data[, L_eff := log(1 - detect) / log(1 - approx)]

# Summary by N: median and range of L_eff across threshold pairs
leff_by_N <- h0_data[, .(
  `Median L_eff` = round(median(L_eff, na.rm = TRUE), 2),
  `Min L_eff` = round(min(L_eff, na.rm = TRUE), 2),
  `Max L_eff` = round(max(L_eff, na.rm = TRUE), 2),
  n_configs = .N
), by = .(N = n_total)]

gt(leff_by_N) |>
  tab_header(
    title = "Implied L_eff by Sample Size",
    subtitle = "Across all (c1, c2) pairs.  Consistency supports L_eff as a candidate-count property."
  ) |>
  cols_label(N = "N", n_configs = "# Configs")
```


## Fit the Power-Law Model

```{r fit-leff}
#| fig-height: 5
#| fig-cap: "L_eff vs N with power-law fit.  Points are individual (c1,c2) configurations; line is the fitted model."

n_min_val <- 60

# Use all H0 configurations for the fit
fit_data <- h0_data[, .(N = n_total, L_eff, c1, c2)]
fit_data[, ratio := N / n_min_val]

# Fit log-linear model on median L_eff per N
leff_medians <- fit_data[, .(L_eff_med = median(L_eff)), by = N]
leff_medians[, ratio := N / n_min_val]

fit_leff <- lm(log(L_eff_med) ~ log(ratio), data = leff_medians)
C_fit <- exp(coef(fit_leff)[1])
alpha_fit <- unname(coef(fit_leff)[2])
r2_fit <- summary(fit_leff)$r.squared

cat(sprintf("Fitted model: L_eff = %.3f * (N / %d)^%.3f\n",
            C_fit, n_min_val, alpha_fit))
cat(sprintf("R-squared: %.3f\n", r2_fit))

# Prediction curve
N_pred <- seq(100, 1200, by = 5)
pred_curve <- data.frame(
  N = N_pred,
  L_eff = C_fit * (N_pred / n_min_val)^alpha_fit
)

ggplot() +
  geom_point(data = fit_data, aes(x = N, y = L_eff, colour = sprintf("c1=%.2f, c2=%.2f", c1, c2)),
             alpha = 0.6, size = 2.5) +
  geom_line(data = pred_curve, aes(x = N, y = L_eff),
            colour = "black", linewidth = 1.2) +
  scale_colour_brewer(palette = "Set2") +
  labs(x = "Total Sample Size N",
       y = expression(L[eff]),
       title = expression("Implied " * L[eff] * " with Power-Law Fit"),
       subtitle = sprintf("L_eff = %.3f * (N/%d)^%.3f,  R-sq = %.3f",
                           C_fit, n_min_val, alpha_fit, r2_fit),
       colour = "Thresholds") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


## Calibrate Using `calibrate_L_eff()`

```{r calibrate-formal}
# Use the package function for formal calibration.
# Compute median P1 and median sim FPR at each N (across all threshold pairs)
# to ensure vectors are aligned.
cal_input <- h0_data[, .(
  P1_med      = median(approx, na.rm = TRUE),
  sim_fpr_med = median(detect, na.rm = TRUE)
), by = .(N = n_total)]
setorder(cal_input, N)

leff_cal <- calibrate_L_eff(
  N       = cal_input$N,
  P1      = cal_input$P1_med,
  sim_fpr = cal_input$sim_fpr_med,
  n_min   = n_min_val
)

print(leff_cal)

# Save calibration to .rds for downstream documents
save_leff_calibration(
  C = C_fit,
  alpha = alpha_fit,
  n_min = n_min_val,
  outcome_type = "binary",
  cal_data = cal_input,
  dgm_description = paste(
    "Binary DGM, 4 confounders (bm1, bm2, age, ecog),",
    "p(event)=0.30, 30% harm prevalence"),
  n_sims_per_N = nsims,
  output_dir = output_dir
)
```


## Corrected Type-I Error Table

```{r corrected-fpr}
#| tbl-cap: "Corrected FPR vs simulation.  Sim/Corrected should be closer to 1.0 than Sim/Uncorrected."

h0_corr <- copy(h0_data)

# Apply correction using the fitted L_eff
h0_corr[, L_eff_fitted := C_fit * (n_total / n_min_val)^alpha_fit]
h0_corr[, fpr_corrected := 1 - (1 - approx)^L_eff_fitted]
h0_corr[, ratio_uncorrected := round(detect / pmax(approx, 0.001), 2)]
h0_corr[, ratio_corrected := round(detect / pmax(fpr_corrected, 0.001), 2)]

gt(h0_corr[, .(N = n_total, c1, c2,
               `P1 (uncorr)` = round(approx, 3),
               `FPR (corrected)` = round(fpr_corrected, 3),
               `Sim FPR` = round(detect, 3),
               `Sim/Uncorr` = ratio_uncorrected,
               `Sim/Corr` = ratio_corrected)]) |>
  tab_header(
    title = md("**Corrected vs Uncorrected Approximation (H0)**"),
    subtitle = sprintf("L_eff = %.3f * (N/%d)^%.3f",
                        C_fit, n_min_val, alpha_fit)
  ) |>
  tab_style(style = cell_fill(color = "#E8F5E9"),
    locations = cells_body(columns = `Sim/Corr`,
      rows = abs(`Sim/Corr` - 1) <= 0.3)) |>
  tab_style(style = cell_fill(color = "#FFF9C4"),
    locations = cells_body(columns = `Sim/Corr`,
      rows = abs(`Sim/Corr` - 1) > 0.3 & abs(`Sim/Corr` - 1) <= 0.6)) |>
  tab_row_group(label = "N = 700", rows = N == 700) |>
  tab_row_group(label = "N = 500", rows = N == 500) |>
  tab_row_group(label = "N = 200", rows = N == 200) |>
  tab_source_note(
    "Green: Sim/Corr within 30% of 1.0.  Yellow: 30-60% off."
  )
```


## Corrected Scatter Plot

```{r corrected-scatter}
#| fig-height: 7
#| fig-cap: "Corrected (triangles) vs uncorrected (circles) approximation against simulation.  Under H0, corrected points should move toward the diagonal."

# Build comparison data for H0 only
scatter_corr <- rbind(
  h0_corr[, .(N = n_total, Approx = approx, Sim = detect,
              Method = "Uncorrected P1",
              config = sprintf("c1=%.2f, c2=%.2f", c1, c2))],
  h0_corr[, .(N = n_total, Approx = fpr_corrected, Sim = detect,
              Method = "Corrected",
              config = sprintf("c1=%.2f, c2=%.2f", c1, c2))]
)

# Add H1 points (uncorrected only --- correction not needed for power)
h1_scatter <- oc_sim[has_hte == TRUE,
  .(N = n_total, Approx = approx, Sim = detect,
    Method = "H1 Power (no correction needed)",
    config = sprintf("c1=%.2f, c2=%.2f", c1, c2))]

scatter_all <- rbind(scatter_corr, h1_scatter)

ggplot(scatter_all, aes(x = Approx, y = Sim)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              colour = "grey50") +
  geom_point(aes(colour = Method, shape = Method), size = 3, alpha = 0.7) +
  scale_shape_manual(values = c("Uncorrected P1" = 16,
                                 "Corrected" = 17,
                                 "H1 Power (no correction needed)" = 4)) +
  scale_colour_manual(values = c("Uncorrected P1" = "firebrick",
                                  "Corrected" = "steelblue",
                                  "H1 Power (no correction needed)" = "grey40")) +
  facet_wrap(~ paste("N =", N)) +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(x = "Approximation", y = "Simulation",
       title = "Corrected vs Uncorrected: All Configurations",
       subtitle = "Triangles (corrected H0) should cluster near diagonal",
       colour = NULL, shape = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Corrected Power-FPR Tradeoff

```{r corrected-tradeoff}
#| fig-height: 6
#| fig-cap: "Power vs FPR using corrected (right) vs uncorrected (left) FPR estimates.  The green region is the acceptable operating region (FPR <= 15%, Power >= 50%)."

# Build tradeoff with corrected FPR
tradeoff_corr <- merge(
  h0_corr[, .(n_total, c1, c2,
              FPR_uncorr = approx,
              FPR_corr = fpr_corrected,
              FPR_sim = detect)],
  oc_sim[has_hte == TRUE, .(n_total, c1, c2, Power = detect)],
  by = c("n_total", "c1", "c2")
)

tradeoff_long <- rbind(
  tradeoff_corr[, .(N = n_total, c1, c2,
                    FPR = FPR_uncorr, Power,
                    Source = "Uncorrected Approx")],
  tradeoff_corr[, .(N = n_total, c1, c2,
                    FPR = FPR_corr, Power,
                    Source = "Corrected Approx")],
  tradeoff_corr[, .(N = n_total, c1, c2,
                    FPR = FPR_sim, Power,
                    Source = "Simulation")]
)
tradeoff_long$config <- sprintf("c1=%.2f, c2=%.2f",
                                 tradeoff_long$c1, tradeoff_long$c2)

ggplot(tradeoff_long, aes(x = FPR, y = Power)) +
  geom_rect(aes(xmin = 0, xmax = 0.15, ymin = 0.50, ymax = 1.0),
    fill = "#E8F5E9", alpha = 0.2, inherit.aes = FALSE) +
  geom_point(aes(colour = Source, shape = Source), size = 2.5, alpha = 0.7) +
  scale_shape_manual(values = c("Uncorrected Approx" = 1,
                                 "Corrected Approx" = 17,
                                 "Simulation" = 16)) +
  scale_colour_manual(values = c("Uncorrected Approx" = "firebrick",
                                  "Corrected Approx" = "steelblue",
                                  "Simulation" = "black")) +
  facet_wrap(~ paste("N =", N)) +
  labs(x = "FPR (H0)", y = "Power (H1, OR = 2.0)",
       title = "Power-FPR Tradeoff: Corrected vs Uncorrected",
       subtitle = "Corrected points (triangles) should align with simulation (filled circles)",
       colour = NULL, shape = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"))
```


## Corrected Threshold Recommendations

Using the corrected approximation, identify thresholds where
the **corrected** FPR is within the acceptable range.

```{r corrected-recommendations}
#| tbl-cap: "Threshold recommendations using corrected FPR (target: corrected FPR <= 15%, simulated power >= 50%)"

# Use corrected FPR for recommendations
rec_corr <- tradeoff_corr[, .(
  N = n_total, c1, c2,
  `Corr FPR` = round(FPR_corr, 3),
  `Sim FPR` = round(FPR_sim, 3),
  `Sim Power` = round(Power, 3)
)]

# Acceptable: corrected FPR <= 15% AND sim power >= 50%
acceptable_corr <- rec_corr[`Corr FPR` <= 0.15 & `Sim Power` >= 0.50]

if (nrow(acceptable_corr) > 0) {
  setorder(acceptable_corr, N, `Corr FPR`)
  gt(acceptable_corr) |>
    tab_header(
      title = md("**Corrected-Approximation Threshold Recommendations**"),
      subtitle = sprintf("L_eff = %.3f * (N/%d)^%.3f | pcons = %.2f",
                          C_fit, n_min_val, alpha_fit, pcons)
    ) |>
    tab_row_group(label = "N = 700", rows = N == 700) |>
    tab_row_group(label = "N = 500", rows = N == 500) |>
    tab_row_group(label = "N = 200", rows = N == 200) |>
    tab_source_note("Corr FPR = corrected approximation.  Sim FPR = validation.")
} else {
  cat("No configurations met both criteria with corrected FPR.\n")
  setorder(rec_corr, N, `Corr FPR`)
  print(rec_corr)
}
```


# Summary {#sec-summary}

## Approximation Accuracy

```{r summary-assessment}
cat("=== Assessment ===\n\n")

cat("POWER (H1): The per-subgroup approximation agrees with simulated\n")
cat("power for N >= 500 (Sim/Approx near 1.0).  For N = 200, the\n")
cat("approximation overestimates because the true subgroup (n_H ~ 60)\n")
cat("is at the n.min boundary and may not enter the candidate set.\n\n")

cat("TYPE-I ERROR (H0, UNCORRECTED): The per-subgroup P1 increasingly\n")
cat("underestimates the simulated FPR as N grows because more candidate\n")
cat("subgroups survive the n.min filter at larger N.\n\n")

cat("TYPE-I ERROR (H0, CORRECTED): The L_eff-corrected approximation\n")
cat("accounts for the effective number of candidates.  The corrected\n")
cat("Sim/Approx ratio should be closer to 1.0 across all N values.\n\n")

cat(sprintf("Fitted L_eff model: %.3f * (N/%d)^%.3f  (R-sq = %.3f)\n",
            C_fit, n_min_val, alpha_fit, r2_fit))
cat(sprintf("  N = 200:  L_eff = %.1f\n", C_fit * (200/n_min_val)^alpha_fit))
cat(sprintf("  N = 500:  L_eff = %.1f\n", C_fit * (500/n_min_val)^alpha_fit))
cat(sprintf("  N = 700:  L_eff = %.1f\n", C_fit * (700/n_min_val)^alpha_fit))
cat(sprintf("  N = 1000: L_eff = %.1f\n", C_fit * (1000/n_min_val)^alpha_fit))
```


## Recommended Thresholds (Simulation-Validated)

Based on the simulation results, identify configurations
where the procedure achieves:

- Simulated FPR $\le$ 15 percent
- Simulated Power $\ge$ 50 percent (for OR = 2.0)

```{r recommendations}
#| tbl-cap: "Acceptable operating points (Sim FPR <= 15%, Power >= 50%)"

acceptable <- tradeoff[FPR <= 0.15 & Power >= 0.50]

if (nrow(acceptable) > 0) {
  setorder(acceptable, n_total, FPR)
  gt(acceptable[, .(N = n_total, c1, c2,
                    `Sim FPR` = round(FPR, 3),
                    `Sim Power` = round(Power, 3))]) |>
    tab_header(
      title = "Simulation-Validated Threshold Configurations",
      subtitle = sprintf("pcons = %.2f | %d sims per config", pcons, nsims)
    ) |>
    tab_row_group(label = "N = 700", rows = N == 700) |>
    tab_row_group(label = "N = 500", rows = N == 500) |>
    tab_row_group(label = "N = 200", rows = N == 200)
} else {
  cat("No configurations met both criteria.\n")
  cat("Consider: wider c2 range, different pconsistency, or larger N.\n\n")
  setorder(tradeoff, n_total, FPR)
  print(tradeoff[, .(N = n_total, c1, c2,
                     FPR = round(FPR, 3), Power = round(Power, 3))])
}
```


## Note on Calibration

The $L_{\text{eff}}$ correction is specific to the confounder
configuration used in this simulation (bm1, bm2, age, ecog)
with `maxk = 2` and `n.min = 60`.  For a different confounder
set or analysis configuration, recalibrate by running a small
H0 simulation (50--100 replicates at 2--3 sample sizes) and
using `calibrate_L_eff()` from the forestsearch package.


# Timing and Session Info {#sec-session}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Stage = c("Simulation (all configs)", "Total (including rendering)"),
  Seconds = c(timings$simulation, timings$total),
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(
    title = "Computational Timing Summary",
    subtitle = sprintf(
      "%d threshold pairs x %d N values x 2 hypotheses x %d sims = %d tasks | %d workers",
      length(sim_pairs), length(n_values), nsims,
      nrow(tasks), n_workers)
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)",
             Minutes = "Time (min)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total (including rendering)")
  )
```

```{r session-info}
plan(sequential)
sessionInfo()
```
