---
title: "Count / Rate Outcomes in ForestSearch"
subtitle: "Subgroup Identification for Heterogeneous Treatment Effects on Recurrent Event Rates: A COPD Exacerbation Example"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ============================================================================
# SIMULATION PARAMETERS
# ============================================================================
sim_n          <- 1000L
sim_n_boot     <- 50L
sim_n_cv_folds <- 5L
sim_seed       <- 42L

sim_n_workers <- max(1L, parallel::detectCores(logical = FALSE) - 1L)
sim_parallel  <- list(plan = "multisession", workers = sim_n_workers,
                      show_message = FALSE)
```


# Introduction {#sec-intro}

## Clinical Motivation

Many clinical trials use **recurrent event counts** as the primary
endpoint.  The outcome is the number of events a patient experiences
over a follow-up period of variable length, analyzed as a rate
(events per unit person-time).  Examples include COPD exacerbations,
asthma attacks, MS relapses, heart failure hospitalizations,
epileptic seizures, and MRI lesion counts.

In COPD, biologic therapies targeting eosinophilic inflammation
(e.g., mepolizumab, benralizumab, dupilumab) have demonstrated a
consistent pattern: the treatment effect on exacerbation rates
**varies by baseline blood eosinophil count**.  In the MATINEE trial
of mepolizumab for COPD (Sciurba et al., *NEJM* 2025), the
annualized exacerbation rate ratio was 0.79 overall, but subgroup
analyses suggested stronger effects in patients with higher
eosinophil levels.  The GALATHEA and TERRANOVA trials of
benralizumab (Criner et al., *NEJM* 2019) showed no significant
overall effect, yet pre-specified eosinophil subgroup analyses
suggested potential benefit in the high-eosinophil subgroup.

This clinical reality---that treatment effects on event rates vary
across subpopulations---motivates extending `forestsearch` to
count/rate outcomes.  The key clinical questions are:

1. **Which subgroup experiences harm?**  A treatment that reduces
   exacerbation rates overall may *increase* rates in a subgroup.
2. **How large is the heterogeneity?**  The incidence rate ratio
   (IRR) in the harm subgroup vs the complement quantifies this.
3. **Which baseline factors define the subgroup?**  Exploratory
   methods like `forestsearch` identify candidate factors from
   the data without pre-specification.


## Statistical Framework

### Poisson GLM for Count Outcomes

The standard model for count data with variable follow-up is the
Poisson regression with a log-link and log(person-time) offset:

$$
Y_i \sim \text{Poisson}(\mu_i), \quad
\log(\mu_i) = \mathbf{x}_i'\boldsymbol{\beta} + \log(t_i)
$$

The treatment coefficient $\beta_W$ estimates the **log incidence
rate ratio** (log-IRR):

$$
\text{IRR} = \exp(\beta_W) =
\frac{\text{rate}_{\text{treatment}}}{\text{rate}_{\text{control}}}
$$

### Heterogeneous Treatment Effects on the Rate Scale

Treatment effect heterogeneity is modeled via interactions on the
log-rate scale:

$$
\log(\mu_i) = \beta_0 + \beta_W W_i + \beta_S S_i +
\beta_{WS} W_i S_i + \mathbf{x}_i'\boldsymbol{\gamma} + \log(t_i)
$$

The subgroup-specific IRRs are:

- **Harm subgroup** ($S = 1$):
  $\text{IRR}_H = \exp(\beta_W + \beta_{WS})$
- **Complement** ($S = 0$):
  $\text{IRR}_{Hc} = \exp(\beta_W)$

### The DINA Estimand

The DINA framework (DIfference in NAtural parameters; *Biometrics*,
2025) proposes the **log conditional mean ratio** as the natural
estimand for count data HTE---identical to the log-IRR used by
`forestsearch`.  For each response type, the DINA estimand is:

| Response | Natural Parameter | DINA Estimand |
|---|---|---|
| Gaussian | Mean $\mu$ | Mean difference |
| Bernoulli | Log-odds | Log-odds ratio |
| **Poisson** | **Log-mean $\log\mu$** | **Log-IRR** |
| Cox | Log-cumulative hazard | Log-HR |


### Overdispersion

The Poisson model assumes $\text{Var}(Y) = E(Y)$.  Clinical count
data almost always exhibit **overdispersion** (variance $>$ mean)
due to unmeasured patient frailty, event clustering, and residual
heterogeneity.  Three variance corrections are supported:

| Method | Variance Function | SE Adjustment |
|---|---|---|
| Poisson | $\mu_i$ | Model-based |
| Quasi-Poisson | $\phi \cdot \mu_i$ | SEs $\times \sqrt{\hat\phi}$ |
| Negative Binomial | $\mu_i + \mu_i^2/\theta$ | Full MLE |

All three produce **identical point estimates** for the log-IRR;
they differ only in standard errors and CIs.


### Threshold Translation

For log-scale measures (IRR), `forestsearch()` converts thresholds:

- `hr.threshold = 1.25` $\rightarrow$ `log(1.25) = 0.223`
- `hr.consistency = 1.0` $\rightarrow$ `log(1.0) = 0.0`

A candidate subgroup is retained if its log-IRR exceeds the
threshold, and the consistency criterion requires this in both
halves of each random split.


## Document Structure

**Part I** demonstrates the standalone count-data functions on a
synthetic dataset with a *known* harm subgroup.

**Part II** runs the integrated `forestsearch()` pipeline with
`outcome_type = "count"`.

**Part III** demonstrates downstream analyses using the best
available subgroup (forestsearch > GRF > known truth).

| Function | Section | Purpose |
|---|---|---|
| `grf.subg.harm.glm()` | @sec-grf | GRF candidate factor screening |
| `create_glm_row()` | @sec-glm-row | Per-subgroup effect estimation |
| `glm_effect_profile()` | @sec-profile | Smooth biomarker profile |
| `gg_forest()` | @sec-forest-known | Publication-quality forest plot |
| `forestsearch()` | @sec-fs | Integrated pipeline |
| `sg_tables()` | @sec-ds-tables | Auto-routed summary tables |
| `plot_sg_glm_outcomes()` | @sec-ds-outcome | Bar chart with annotations |
| `forestsearch_bootstrap_dofuture()` | @sec-bootstrap | Bias correction |
| `forestsearch_Kfold()` | @sec-cv | Replication stability |


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(ggplot2)
library(patchwork)
library(gt)
library(data.table)
```

```{r timing-init}
timings <- list()
t_vignette_start <- proc.time()
```


# Synthetic Data: COPD Exacerbation Trial {#sec-data}

## Data-Generating Mechanism

We simulate a **respiratory exacerbation trial** inspired by the
COPD biologic therapy setting.  The DGM plants a harm subgroup
where treatment *increases* the exacerbation rate, while the
complement benefits.  Parameters are calibrated to published
trial data:

| Parameter | Value | Source / Rationale |
|---|---|---|
| $n$ | 1000 | Typical Phase III COPD trial |
| Follow-up | 0.5--3.0 years | Variable as-treated exposure |
| Baseline rate (control) | ~1.2 events/person-year | GALATHEA ~0.93, REACT ~1.5, MATINEE ~1.01 |
| **Harm subgroup** (eos low, 35 percent) | log-IRR = +1.0, IRR $\approx$ 2.7 | Treatment increases exacerbations |
| **Complement** (eos high, 65 percent) | log-IRR = $-$0.5, IRR $\approx$ 0.61 | Treatment reduces exacerbations |
| Biomarker interaction | Slope 0.15 per unit | Continuous effect modification |
| Age effect | 0.02 per year above 60 | Baseline rate covariate |

The model is:

$$
\log(\mu_i) = -1.0
+ 1.0 \cdot W_i \cdot \mathbf{1}(\text{eos\_low}_i)
- 0.5 \cdot W_i \cdot \mathbf{1}(\text{eos\_high}_i)
+ 0.15 \cdot W_i \cdot (z_i - 2.5)
+ 0.02 \cdot (\text{age}_i - 60)
+ \log(t_i)
$$

$$Y_i \sim \text{Poisson}(\mu_i)$$

```{r sim-data}
#| code-fold: false

set.seed(sim_seed)
n <- sim_n

df <- data.frame(
  id          = seq_len(n),
  treat       = rbinom(n, 1, 0.5),
  # Subgroup-defining factors
  eos_low     = as.factor(rbinom(n, 1, 0.35)),   # harm subgroup
  smoking     = as.factor(rbinom(n, 1, 0.50)),    # noise
  male        = as.factor(rbinom(n, 1, 0.45)),    # noise
  age         = round(rnorm(n, 60, 12)),          # baseline covariate
  fev1_pct    = round(runif(n, 0.5, 5.0), 2),     # continuous biomarker (FEV1 pct predicted / 20)
  follow_up   = round(runif(n, 0.5, 3.0), 2)      # person-time (years)
)

# Poisson outcome with heterogeneous treatment effect
log_rate <- with(df, {
  -1.0 +
    1.0  * treat * (eos_low == 1) +
    -0.5 * treat * (eos_low == 0) +
    0.15 * treat * (fev1_pct - 2.5) +
    0.02 * (age - 60) +
    log(follow_up)
})
df$exacerbations <- rpois(n, lambda = exp(log_rate))
df$flag_harm     <- as.integer(df$eos_low == 1)
```

## Data Summary

```{r data-summary}
cat("=== COPD Exacerbation Trial (Simulated) ===\n")
cat("N:", n, "\n")
cat("Exacerbations: mean =", round(mean(df$exacerbations), 2),
    ", range = [", min(df$exacerbations), ",",
    max(df$exacerbations), "]\n")
cat("Zero counts:", sum(df$exacerbations == 0),
    sprintf("(%.1f%%)\n", 100 * mean(df$exacerbations == 0)))
cat("Follow-up: mean =", round(mean(df$follow_up), 2), "years\n\n")

cat("Per-arm event rates:\n")
for (arm in 0:1) {
  idx <- df$treat == arm
  rate <- sum(df$exacerbations[idx]) / sum(df$follow_up[idx])
  cat(sprintf("  treat = %d: n = %d, events = %d, PT = %.1f yr, rate = %.2f/yr\n",
              arm, sum(idx), sum(df$exacerbations[idx]),
              sum(df$follow_up[idx]), rate))
}

cat("\nPer-subgroup event rates:\n")
for (sg in 0:1) {
  for (arm in 0:1) {
    idx <- df$eos_low == sg & df$treat == arm
    rate <- sum(df$exacerbations[idx]) / sum(df$follow_up[idx])
    sg_lbl <- ifelse(sg == 1, "eos_low = 1 (harm)", "eos_low = 0 (benefit)")
    cat(sprintf("  %s, treat = %d: rate = %.2f/yr (n = %d)\n",
                sg_lbl, arm, rate, sum(idx)))
  }
}
```

```{r truth-table}
#| tbl-cap: "DGM treatment effects by subgroup"

fit_overall <- glm(exacerbations ~ treat + offset(log(follow_up)),
                   family = poisson, data = df)
fit_eos1 <- glm(exacerbations ~ treat + offset(log(follow_up)),
                family = poisson, data = df[df$eos_low == 1, ])
fit_eos0 <- glm(exacerbations ~ treat + offset(log(follow_up)),
                family = poisson, data = df[df$eos_low == 0, ])

truth_df <- data.frame(
  Subgroup = c("eos_low = 1 (harm)", "eos_low = 0 (benefit)",
               "Overall (ITT)"),
  `True log-IRR` = c(1.0, -0.5, NA),
  `Obs log-IRR` = c(coef(fit_eos1)["treat"],
                     coef(fit_eos0)["treat"],
                     coef(fit_overall)["treat"]),
  `Obs IRR` = c(exp(coef(fit_eos1)["treat"]),
                 exp(coef(fit_eos0)["treat"]),
                 exp(coef(fit_overall)["treat"])),
  N = c(sum(df$eos_low == 1), sum(df$eos_low == 0), n),
  check.names = FALSE
)

gt(truth_df) |>
  tab_header(
    title = "DGM Treatment Effects",
    subtitle = "COPD exacerbation trial: Poisson GLM with log(follow-up) offset"
  ) |>
  fmt_number(columns = 2:4, decimals = 3) |>
  tab_source_note("True log-IRR: harm = +1.0 (IRR = 2.72), benefit = -0.5 (IRR = 0.61)")
```


# Part I: Standalone Function Demonstrations {#sec-part1}

These sections use the **known truth** (`eos_low = 1` is harm)
directly, so all outputs render unconditionally.


## GRF Screening {#sec-grf}

`grf.subg.harm.glm()` uses `grf::causal_forest()` with a
variance-stabilising $\log(Y + 0.5)$ transform.  This compresses
the count scale so that high-count patients do not dominate
splits, and approximately aligns the difference scale with the
log-ratio scale.

### Poisson with Log Transform

```{r grf-poisson}
#| code-fold: false
t0 <- proc.time()[3]

grf_pois <- grf.subg.harm.glm(
  data                = df,
  confounders.name    = c("eos_low", "smoking", "male", "age"),
  outcome.name        = "exacerbations",
  treat.name          = "treat",
  id.name             = "id",
  outcome_type        = "count",
  effect_measure      = "log_IRR",
  offset.name         = "follow_up",
  overdispersion      = "none",
  grf_count_transform = "log",
  n.min               = 60,
  conf.level          = 0.95,
  seedit              = 8316951,
  verbose             = TRUE
)

timings$grf_standalone <- proc.time()[3] - t0

cat("\n=== GRF Result ===\n")
cat("Subgroup found:", !is.null(grf_pois$sg.harm.id), "\n")
if (!is.null(grf_pois$sg.harm.id)) {
  cat("Definition:", paste(grf_pois$sg.harm.id, collapse = " & "), "\n")
  n_harm <- sum(grf_pois$data$treat.recommend == 0)
  cat(sprintf("Size: n(H) = %d (%.1f%%)\n", n_harm, 100 * n_harm / n))
}
```

### Variable Importance: Log vs Identity Transform

```{r grf-identity}
grf_id <- grf.subg.harm.glm(
  data = df, confounders.name = c("eos_low", "smoking", "male", "age"),
  outcome.name = "exacerbations", treat.name = "treat", id.name = "id",
  outcome_type = "count", effect_measure = "log_IRR",
  offset.name = "follow_up", grf_count_transform = "identity",
  n.min = 60, seedit = 8316951, verbose = FALSE
)
```

```{r vi-table}
#| tbl-cap: "GRF variable importance by transform"

vi_log <- sort(grf_pois$grf_varimp, decreasing = TRUE)
vi_id  <- sort(grf_id$grf_varimp, decreasing = TRUE)

vi_df <- data.frame(
  Variable = names(vi_log),
  `Importance (log)` = as.numeric(vi_log),
  `Importance (identity)` = as.numeric(vi_id[names(vi_log)]),
  check.names = FALSE
)

gt(vi_df) |>
  tab_header(
    title = "GRF Variable Importance by Transform",
    subtitle = sprintf("n = %d | %d zero counts | log(Y+0.5) vs raw Y",
                        n, sum(df$exacerbations == 0))
  ) |>
  fmt_number(columns = 2:3, decimals = 4) |>
  tab_source_note(sprintf("log SG: %s | identity SG: %s",
    paste(grf_pois$sg.harm.id %||% "none", collapse = " & "),
    paste(grf_id$sg.harm.id %||% "none", collapse = " & ")))
```


## Per-Subgroup Estimation with `create_glm_row()` {#sec-glm-row}

`create_glm_row()` fits a GLM within a subgroup and returns a
one-row data frame with point estimate, CI, sample size, and event
counts.  Here we use the **known truth** (`eos_low = 1`) and
compare three variance estimators.

```{r glm-row-estimates}
#| code-fold: false

df_harm <- df[df$eos_low == 1, ]
df_comp <- df[df$eos_low == 0, ]

methods <- c("none", "quasi", "negbin")
method_labels <- c("Poisson", "Quasi-Poisson", "Neg. Binomial")

build_rows <- function(dfa, sg_label) {
  lapply(seq_along(methods), function(i) {
    r <- create_glm_row(dfa, "exacerbations", "treat", "log_IRR",
                        "follow_up", methods[i], 0.95)
    r$Method <- method_labels[i]
    r$Subgroup <- sg_label
    r
  })
}

est_all <- do.call(rbind, c(
  build_rows(df, "ITT"),
  build_rows(df_harm, "Harm (H): eos_low = 1"),
  build_rows(df_comp, "Benefit (Hc): eos_low = 0")
))
est_all$IRR <- exp(est_all$est)
est_all$IRR_lower <- exp(est_all$lower)
est_all$IRR_upper <- exp(est_all$upper)
```

```{r glm-row-table}
#| tbl-cap: "Per-subgroup IRR estimates: three variance estimators"

display <- est_all[, c("Subgroup", "Method", "N", "events_treat",
                        "events_control", "est", "se", "IRR",
                        "IRR_lower", "IRR_upper")]

gt(display) |>
  tab_header(
    title = "Subgroup Effect Estimates (log-IRR Scale)",
    subtitle = "Poisson vs quasi-Poisson vs negative binomial"
  ) |>
  fmt_number(columns = c(est, se), decimals = 3) |>
  fmt_number(columns = c(IRR, IRR_lower, IRR_upper), decimals = 2) |>
  cols_label(
    Subgroup = "Subgroup", Method = "Variance",
    N = "N", events_treat = "Ev (Rx)", events_control = "Ev (Ctrl)",
    est = "log-IRR", se = "SE",
    IRR = "IRR", IRR_lower = "Lower", IRR_upper = "Upper"
  ) |>
  tab_spanner(label = "log scale", columns = c(est, se)) |>
  tab_spanner(label = "IRR (95 percent CI)",
              columns = c(IRR, IRR_lower, IRR_upper)) |>
  tab_source_note(
    "True: H log-IRR = +1.00 (IRR = 2.72), Hc log-IRR = -0.50 (IRR = 0.61)")
```


## Forest Plot (Known Truth) {#sec-forest-known}

```{r forest-plot-known}
#| fig-height: 4
#| fig-cap: "Forest plot of subgroup IRR estimates (Poisson SEs). Subgroups defined by known truth (eos_low)."

fp_data <- est_all[est_all$Method == "Poisson", ]

p_forest <- gg_forest(
  subgroups  = fp_data$Subgroup,
  est        = fp_data$IRR,
  lo         = fp_data$IRR_lower,
  hi         = fp_data$IRR_upper,
  cat_vec    = ifelse(fp_data$Subgroup == "ITT", "ITT", "Subgroup"),
  cat_colours = c("ITT" = "grey30", "Subgroup" = "steelblue"),
  annot      = list(
    "N" = as.character(fp_data$N),
    "IRR (95% CI)" = sprintf("%.2f (%.2f, %.2f)",
                              fp_data$IRR, fp_data$IRR_lower, fp_data$IRR_upper)
  ),
  ref_line   = 1.0,
  vert_lines = 1.0,
  xlim       = c(0.2, 5.0),
  ticks_at   = c(0.25, 0.5, 0.67, 1.0, 1.5, 2.0, 3.0, 4.0),
  xlab       = "Incidence Rate Ratio",
  xlog       = TRUE,
  base_size  = 11,
  widths     = c(4, 5, 1, 2.5)
)
print(p_forest)
```


## Biomarker Treatment-Effect Profile {#sec-profile}

`glm_effect_profile()` estimates the treatment effect as a smooth
function of FEV1 (the continuous biomarker) using natural cubic
splines with delta-method CIs.

```{r profile-all}
#| fig-height: 6
#| fig-cap: "Three variance estimators: point estimates identical, CIs widen with overdispersion correction. Dashed line: approximate true DGM interaction slope."

t0 <- proc.time()[3]

prof_pois <- glm_effect_profile(
  df = df, outcome.name = "exacerbations", treat.name = "treat",
  z.name = "fev1_pct", effect_measure = "log_IRR",
  offset.name = "follow_up", overdispersion = "none",
  spline_df = 3, z_by = 0.05, z_quantile = 0.95,
  alpha = 0.05, verbose = FALSE
)

prof_quasi <- glm_effect_profile(
  df = df, outcome.name = "exacerbations", treat.name = "treat",
  z.name = "fev1_pct", effect_measure = "log_IRR",
  offset.name = "follow_up", overdispersion = "quasi",
  spline_df = 3, z_by = 0.05, z_quantile = 0.95,
  alpha = 0.05, verbose = FALSE
)

prof_nb <- glm_effect_profile(
  df = df, outcome.name = "exacerbations", treat.name = "treat",
  z.name = "fev1_pct", effect_measure = "log_IRR",
  offset.name = "follow_up", overdispersion = "negbin",
  spline_df = 3, z_by = 0.05, z_quantile = 0.95,
  alpha = 0.05, verbose = FALSE
)

timings$profiles <- proc.time()[3] - t0

df_prof <- rbind(
  data.frame(fev1_pct = prof_pois$z_profile, est = prof_pois$est,
             lower = prof_pois$lower, upper = prof_pois$upper,
             Method = "Poisson"),
  data.frame(fev1_pct = prof_quasi$z_profile, est = prof_quasi$est,
             lower = prof_quasi$lower, upper = prof_quasi$upper,
             Method = "Quasi-Poisson"),
  data.frame(fev1_pct = prof_nb$z_profile, est = prof_nb$est,
             lower = prof_nb$lower, upper = prof_nb$upper,
             Method = "Neg. Binomial")
)
df_prof$Method <- factor(df_prof$Method,
  levels = c("Poisson", "Quasi-Poisson", "Neg. Binomial"))

true_intercept <- 1.0 * 0.35 + (-0.5) * 0.65
true_slope <- 0.15

ggplot(df_prof, aes(x = fev1_pct, y = est, colour = Method, fill = Method)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.08, colour = NA) +
  geom_line(linewidth = 1.0) +
  geom_abline(intercept = true_intercept - true_slope * 2.5,
              slope = true_slope,
              linetype = "dashed", color = "grey40", linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "red",
             linewidth = 0.4) +
  geom_rug(data = df, aes(x = fev1_pct, y = NULL),
           inherit.aes = FALSE,
           alpha = 0.1, sides = "b", length = unit(0.02, "npc")) +
  scale_colour_manual(values = c("Poisson" = "steelblue",
                                  "Quasi-Poisson" = "#E07B39",
                                  "Neg. Binomial" = "#7B2D8E")) +
  scale_fill_manual(values = c("Poisson" = "steelblue",
                                "Quasi-Poisson" = "#E07B39",
                                "Neg. Binomial" = "#7B2D8E")) +
  labs(x = "FEV1 (percent predicted / 20)", y = "log(IRR)",
       title = "Treatment-Effect Profile Across FEV1",
       subtitle = sprintf(
         "Dispersion: quasi = %.3f | negbin (1/theta) = %.3f",
         prof_quasi$dispersion, prof_nb$dispersion)) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
```

```{r profile-print}
cat("=== Poisson ===\n")
print(prof_pois)
cat("\n=== Quasi-Poisson ===\n")
print(prof_quasi)
cat("\n=== Negative Binomial ===\n")
print(prof_nb)
```


## SE Inflation Factor {#sec-se-comparison}

```{r se-comparison}
#| fig-height: 5
#| fig-cap: "Standard error ratio relative to Poisson across the biomarker range. Ratios near 1.0 confirm that the DGM has minimal overdispersion (as expected from a pure Poisson DGP)."

se_ratio_df <- data.frame(
  fev1_pct = prof_pois$z_profile,
  `Quasi / Poisson` = prof_quasi$se / prof_pois$se,
  `NegBin / Poisson` = prof_nb$se / prof_pois$se,
  check.names = FALSE
)

se_long <- data.frame(
  fev1_pct = rep(se_ratio_df$fev1_pct, 2),
  SE_ratio = c(se_ratio_df$`Quasi / Poisson`,
               se_ratio_df$`NegBin / Poisson`),
  Method = rep(c("Quasi-Poisson", "Neg. Binomial"),
               each = nrow(se_ratio_df))
)

ggplot(se_long, aes(x = fev1_pct, y = SE_ratio, colour = Method)) +
  geom_line(linewidth = 1.0) +
  geom_hline(yintercept = 1.0, linetype = "dashed", colour = "grey50") +
  scale_colour_manual(values = c("Quasi-Poisson" = "#E07B39",
                                  "Neg. Binomial" = "#7B2D8E")) +
  labs(x = "FEV1 (percent predicted / 20)", y = "SE Ratio (vs Poisson)",
       title = "Overdispersion Correction: SE Inflation Factor",
       subtitle = sprintf("Dispersion: quasi = %.3f | negbin (1/theta) = %.3f",
                           prof_quasi$dispersion, prof_nb$dispersion)) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
```


# Part II: Integrated ForestSearch Pipeline {#sec-fs}

`forestsearch()` with `outcome_type = "count"` runs the full
pipeline: GRF screening, LASSO variable selection, exhaustive
subgroup search with a Poisson IRR effect estimator, and
split-sample consistency evaluation.

```{r fs-count}
#| code-fold: false
t0 <- proc.time()[3]

confounders <- c("eos_low", "smoking", "male", "age")

fs <- forestsearch(
  df.analysis      = df,
  confounders.name = confounders,
  outcome.name     = "exacerbations",
  event.name       = "exacerbations",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "count",
  effect_measure   = "IRR",
  offset.name      = "follow_up",
  overdispersion   = "none",
  grf_count_transform = "log",
  hr.threshold     = 1.10,
  hr.consistency   = 1.0,
  pconsistency.threshold = 0.40,
  fs.splits        = 400,
  n.min            = 60,
  maxk             = 2,
  use_lasso        = TRUE,
  use_grf          = TRUE,
  use_twostage     = TRUE,
  is.RCT           = TRUE,
  seedit           = 8316951,
  details          = TRUE,
  plot.sg          = FALSE,
  parallel_args    = sim_parallel
)

timings$forestsearch <- proc.time()[3] - t0
cat(sprintf("\nForestSearch completed in %.1f sec\n", timings$forestsearch))
```

```{r fs-result-check}
fs_found <- !is.null(fs$sg.harm) && length(fs$sg.harm) > 0

if (fs_found) {
  cat("=== ForestSearch identified a subgroup ===\n")
  cat("Definition:", paste(fs$sg.harm, collapse = " & "), "\n")
} else {
  cat("=== ForestSearch did not identify a subgroup ===\n")
  cat("The consistency criterion across random splits was not met.\n")
  cat("This is expected for count data where half-sample event counts\n")
  cat("are smaller, producing higher-variance log-IRR estimates.\n")
  cat("Part III will use the GRF-identified subgroup as fallback.\n")
}
```

```{r fs-print}
print(fs)
```

```{r fs-summary}
summary(fs)
```


# Part III: Downstream Analyses {#sec-part3}

This section demonstrates summary tables, outcome plots, forest
plots, bootstrap, and cross-validation using the best available
subgroup.

```{r resolve-subgroup}
#| code-fold: false

# Priority cascade: forestsearch > GRF > known truth
if (fs_found) {
  sg_source <- "forestsearch"
  sg_label  <- paste(fs$sg.harm, collapse = " & ")
  sg_flag   <- fs$df.est$treat.recommend
} else if (!is.null(grf_pois$sg.harm.id)) {
  sg_source <- "GRF (grf.subg.harm.glm)"
  sg_label  <- paste(grf_pois$sg.harm.id, collapse = " & ")
  sg_flag   <- as.integer(grf_pois$data$treat.recommend)
} else {
  sg_source <- "Known truth (eos_low)"
  sg_label  <- "eos_low = 1"
  sg_flag   <- 1L - df$flag_harm
}

cat("=== Working Subgroup ===\n")
cat("Source:", sg_source, "\n")
cat("Definition:", sg_label, "\n")
cat(sprintf("H:  n = %d (%.1f%%)\n", sum(sg_flag == 0),
            100 * mean(sg_flag == 0)))
cat(sprintf("Hc: n = %d (%.1f%%)\n", sum(sg_flag == 1),
            100 * mean(sg_flag == 1)))
```


## Summary Tables {#sec-ds-tables}

```{r sg-tables}
if (fs_found) {
  tabs <- tryCatch(sg_tables(fs), error = function(e) {
    message("sg_tables(): ", e$message); NULL
  })
  if (!is.null(tabs)) print(tabs)
} else {
  cat("sg_tables() requires a forestsearch result.\n")
  cat("Using create_glm_row() with the", sg_source, "subgroup:\n\n")

  df_H  <- df[sg_flag == 0, ]
  df_Hc <- df[sg_flag == 1, ]

  manual_rows <- do.call(rbind, list(
    { r <- create_glm_row(df,    "exacerbations", "treat", "log_IRR",
                          "follow_up", "none", 0.95); r$Subgroup <- "ITT"; r },
    { r <- create_glm_row(df_H,  "exacerbations", "treat", "log_IRR",
                          "follow_up", "none", 0.95);
      r$Subgroup <- paste("H:", sg_label); r },
    { r <- create_glm_row(df_Hc, "exacerbations", "treat", "log_IRR",
                          "follow_up", "none", 0.95);
      r$Subgroup <- "Hc (Complement)"; r }
  ))
  manual_rows$IRR    <- exp(manual_rows$est)
  manual_rows$IRR_lo <- exp(manual_rows$lower)
  manual_rows$IRR_hi <- exp(manual_rows$upper)

  disp <- manual_rows[, c("Subgroup", "N", "events_treat",
                            "events_control", "IRR", "IRR_lo", "IRR_hi")]

  gt(disp) |>
    tab_header(
      title = sprintf("Subgroup Summary (source: %s)", sg_source),
      subtitle = sprintf("H = %s", sg_label)
    ) |>
    fmt_number(columns = c(IRR, IRR_lo, IRR_hi), decimals = 2) |>
    cols_label(
      Subgroup = "Subgroup", N = "N",
      events_treat = "Ev (Rx)", events_control = "Ev (Ctrl)",
      IRR = "IRR", IRR_lo = "Lower", IRR_hi = "Upper"
    ) |>
    tab_spanner(label = "IRR (95 percent CI)",
                columns = c(IRR, IRR_lo, IRR_hi))
}
```


## Subgroup Outcome Visualization {#sec-ds-outcome}

`plot_sg_glm_outcomes()` creates a grouped bar chart showing
per-arm exacerbation rates within each subgroup.

```{r outcome-plot}
#| fig-height: 5.5
#| fig-cap: "Per-arm exacerbation rates by subgroup with IRR annotations"

if (fs_found) {
  fs_plot <- fs
} else {
  fs_plot <- list(
    df.est       = data.frame(df, treat.recommend = sg_flag),
    sg.harm      = sg_label,
    outcome_type = "count",
    effect_measure = "IRR",
    args_call_all = list(
      outcome_type   = "count",
      effect_measure = "IRR",
      offset.name    = "follow_up",
      treat.name     = "treat",
      outcome.name   = "exacerbations"
    )
  )
}

outcome_viz <- tryCatch(
  plot_sg_glm_outcomes(
    fs.est         = fs_plot,
    outcome.name   = "exacerbations",
    treat.name     = "treat",
    offset.name    = "follow_up",
    effect_measure = "IRR",
    outcome_type   = "count",
    sg0_name       = "Harm (H)",
    sg1_name       = "Benefit (Hc)",
    E.name         = "Treatment",
    C.name         = "Placebo",
    show_effect    = TRUE,
    show_itt       = TRUE,
    title          = "COPD Exacerbation Rates by Subgroup",
    subtitle       = sprintf("N = %d | Source: %s | Effect: IRR",
                              n, sg_source)
  ),
  error = function(e) {
    message("plot_sg_glm_outcomes(): ", e$message)
    NULL
  }
)
if (!is.null(outcome_viz)) print(outcome_viz)
```


## Forest Plot (Identified Subgroup) {#sec-ds-forest}

```{r forest-plot-identified}
#| fig-height: 4
#| fig-cap: "Forest plot from identified subgroup"

df_H  <- df[sg_flag == 0, ]
df_Hc <- df[sg_flag == 1, ]

fp_list <- list(
  create_glm_row(df,    "exacerbations", "treat", "log_IRR",
                 "follow_up", "none", 0.95),
  create_glm_row(df_H,  "exacerbations", "treat", "log_IRR",
                 "follow_up", "none", 0.95),
  create_glm_row(df_Hc, "exacerbations", "treat", "log_IRR",
                 "follow_up", "none", 0.95)
)
fp_df <- do.call(rbind, fp_list)
fp_df$Subgroup <- c("ITT", paste("H:", sg_label), "Benefit (Hc)")
fp_df$IRR    <- exp(fp_df$est)
fp_df$IRR_lo <- exp(fp_df$lower)
fp_df$IRR_hi <- exp(fp_df$upper)

p <- gg_forest(
  subgroups  = fp_df$Subgroup,
  est        = fp_df$IRR,
  lo         = fp_df$IRR_lo,
  hi         = fp_df$IRR_hi,
  cat_vec    = ifelse(fp_df$Subgroup == "ITT", "ITT", "Subgroup"),
  cat_colours = c("ITT" = "grey30", "Subgroup" = "firebrick"),
  annot      = list(
    "N" = as.character(fp_df$N),
    "IRR (95% CI)" = sprintf("%.2f (%.2f, %.2f)",
                              fp_df$IRR, fp_df$IRR_lo, fp_df$IRR_hi)
  ),
  ref_line = 1.0, vert_lines = 1.0,
  xlim = c(0.2, 5.0),
  ticks_at = c(0.25, 0.5, 1.0, 1.5, 2.0, 3.0),
  xlab = "Incidence Rate Ratio", xlog = TRUE,
  base_size = 11, widths = c(4, 5, 1, 2.5)
)
print(p)
```


## Bootstrap Bias Correction {#sec-bootstrap}

`forestsearch_bootstrap_dofuture()` resamples the data and re-runs
the pipeline to estimate selection bias.

```{r bootstrap}
t0 <- proc.time()[3]

if (fs_found) {
  fs_bc <- tryCatch(
    forestsearch_bootstrap_dofuture(
      fs,
      nb = sim_n_boot,
      parallel_args = sim_parallel
    ),
    error = function(e) {
      message("Bootstrap failed: ", e$message)
      NULL
    }
  )
} else {
  fs_bc <- NULL
}

timings$bootstrap <- proc.time()[3] - t0
```

```{r bootstrap-results}
if (!is.null(fs_bc)) {
  cat("=== Bootstrap Summary ===\n")
  cat("Replicates:", sim_n_boot, "\n")
  bc_table <- tryCatch(
    format_bootstrap_table(fs_bc),
    error = function(e) {
      message("format_bootstrap_table(): ", e$message); NULL
    }
  )
  if (!is.null(bc_table)) print(bc_table)
} else {
  cat("Bootstrap requires forestsearch() to identify a subgroup.\n")
  cat("Working subgroup source:", sg_source, "\n")
}
```


## Cross-Validation {#sec-cv}

```{r cv}
t0 <- proc.time()[3]

if (fs_found) {
  fs_cv <- tryCatch(
    forestsearch_Kfold(
      fs,
      K = sim_n_cv_folds,
      parallel_args = sim_parallel
    ),
    error = function(e) {
      message("Kfold CV failed: ", e$message); NULL
    }
  )
} else {
  fs_cv <- NULL
}

timings$cv <- proc.time()[3] - t0
```

```{r cv-results}
if (!is.null(fs_cv)) {
  cat("=== Cross-Validation Summary ===\n")
  print(fs_cv)
} else {
  cat("Cross-validation requires forestsearch() to identify a subgroup.\n")
  cat("Working subgroup source:", sg_source, "\n")
}
```


# Timing Summary {#sec-timing}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_stages <- c("GRF standalone", "forestsearch() pipeline",
                    "Biomarker profiles")
timing_values <- c(timings$grf_standalone, timings$forestsearch,
                    timings$profiles)

if (!is.null(timings$bootstrap) && timings$bootstrap > 0.1) {
  timing_stages <- c(timing_stages, "Bootstrap")
  timing_values <- c(timing_values, timings$bootstrap)
}
if (!is.null(timings$cv) && timings$cv > 0.1) {
  timing_stages <- c(timing_stages, "Cross-validation")
  timing_values <- c(timing_values, timings$cv)
}

timing_stages <- c(timing_stages, "Total")
timing_values <- c(timing_values, timings$total)

timing_df <- data.frame(
  Stage = timing_stages,
  Seconds = timing_values,
  stringsAsFactors = FALSE
)

gt(timing_df) |>
  tab_header(
    title = "Computational Timing",
    subtitle = sprintf("n = %d | %d bootstrap | %d CV folds | %d workers",
                        n, sim_n_boot, sim_n_cv_folds, sim_n_workers)
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total")
  )
```


# Summary {#sec-summary}

This document demonstrated the full ForestSearch v0.2.0 pipeline
for count/rate outcomes using a COPD exacerbation trial simulation
calibrated to published clinical trial parameters.

**Part I** exercised the standalone functions using the known truth
subgroup, confirming that:

- `grf.subg.harm.glm()` correctly identified `eos_low` as the top
  variable importance factor via $\log(Y + 0.5)$ variance
  stabilisation
- `create_glm_row()` produced concordant point estimates across
  Poisson, quasi-Poisson, and negative binomial models, with SEs
  widening appropriately under overdispersion correction
- `gg_forest()` rendered a publication-quality IRR forest plot
- `glm_effect_profile()` recovered the continuous biomarker
  interaction across all three variance estimators

**Part II** ran `forestsearch(outcome_type = "count")` with the
integrated pipeline.  The consistency criterion across random
splits is more demanding for count data than survival data because
half-sample event counts are smaller, producing higher-variance
log-IRR estimates.

**Part III** exercised downstream analyses (summary tables, outcome
bar charts, forest plots) using a priority cascade
(forestsearch > GRF > known truth) so that every function renders
regardless of whether the consistency criterion was met.  Bootstrap
and cross-validation are available when `forestsearch()` identifies
a subgroup.


# References {#sec-references}

## Methodological

- Leon, L.F., et al. (2024). Exploratory Subgroup Identification
  for Treatment Effect Heterogeneity in Randomized Controlled
  Trials. *Statistics in Medicine*, 43, 4253--4275.
  doi:10.1002/sim.10163

- DINA: Estimating Heterogeneous Treatment Effects for General
  Responses. *Biometrics*, 81(4), 2025. (Log conditional mean
  ratio as the natural estimand for Poisson count data)

- Athey, S., Tibshirani, J., and Wager, S. (2019). Generalized
  Random Forests. *Annals of Statistics*, 47(2), 1148--1178.

- Keene, O.N., et al. (2007). Analysis of Exacerbation Rates in
  Asthma and COPD: Example from the TRISTAN Study.
  *Pharmaceutical Statistics*, 6, 89--97.


## Clinical (COPD Exacerbation Trials)

- Sciurba, F., et al. (2025). Mepolizumab to Prevent
  Exacerbations of COPD (MATINEE). *NEJM*, 392, 1710--1720.

- Criner, G.J., et al. (2019). Benralizumab for COPD
  (GALATHEA/TERRANOVA). *NEJM*, 381, 1023--1034.

- Bhatt, S.P., et al. (2023). Dupilumab for COPD with Type 2
  Inflammation (BOREAS). *NEJM*, 389, 205--214.

- Martinez, F.J., et al. (2015). Roflumilast in Severe COPD
  (REACT). *Lancet*, 385, 857--866.


# Session Information {#sec-session}

```{r session-info}
sessionInfo()
```
