---
title: "GLM Subgroup Identification: Simulation Study"
subtitle: "Operating Characteristics Across Binary, Continuous, and Count Outcomes Inspired by the mCART Framework"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ============================================================================
# SIMULATION PARAMETERS --- edit here to tune runtime vs precision
# ============================================================================
nsims       <- 44L       # simulations per scenario (increase for production)
sim_seed    <- 8316951L
n_workers   <- max(1L, parallel::detectCores(logical = FALSE) - 1L)
```


# Introduction {#sec-intro}

## Motivation

This document evaluates the operating characteristics of `forestsearch()`
for GLM outcomes across a systematic set of simulation scenarios inspired
by the mCART framework (Rigdon et al., *Trials*, 2018).  The mCART paper
evaluated four methods (logistic regression, random forest, gradient RF,
and mCART) on five scenarios:

| Scenario | Setting | N | True HTE |
|---|---|---|---|
| 1A | Phase II oncology | 200 | None (null) |
| 1B | Phase II oncology | 200 | Two subgroups |
| 2A | Phase III CVD | 6000 | None (null) |
| 2B | Phase III CVD | 6000 | Four subgroups |
| 2C | Phase III CVD (null ATE) | 6000 | Four subgroups, null overall |

We adapt this framework for `forestsearch` across three GLM outcome types:

| Outcome Type | Clinical Analog | Effect Measure | Key Scenarios |
|---|---|---|---|
| **Binary** | Disease progression (oncology) | RD, OR | Phase II (small n) and Phase III |
| **Continuous** | DAS28 change (RA) | MD | Phase III immunology |
| **Count** | Exacerbations (COPD) | IRR | Phase III respiratory |

For each outcome type, we run:

- **H0 (null)**: No treatment effect heterogeneity --- Type I error
- **H1 (alternative)**: Planted harm subgroup --- Power, sensitivity,
  specificity, PPV, NPV, effect estimation bias


## Evaluation Metrics

| Metric | Definition |
|---|---|
| Detection rate (power) | Proportion of simulations where any subgroup found |
| Type I error | Detection rate under H0 |
| Sensitivity | Proportion of true harm patients correctly classified |
| Specificity | Proportion of true non-harm patients correctly classified |
| PPV | Among identified harm patients, proportion truly harmed |
| NPV | Among identified non-harm patients, proportion truly not harmed |
| Bias (H) | Mean(estimated effect in H) $-$ true effect in H |
| Bias (Hc) | Mean(estimated effect in Hc) $-$ true effect in Hc |


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)
```

```{r parallel-setup}
plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")
cat("Simulations per scenario:", nsims, "\n")
```

```{r timing-init}
timings <- list()
t_vignette_start <- proc.time()
```


# Data-Generating Mechanisms {#sec-dgm}

```{r dgm-functions}
#| code-fold: false

# ============================================================================
# Binary DGM: Oncology disease progression (Y=1 = adverse event)
# ForestSearch assumes adverse_outcome=TRUE for binary (Y=1 = harm).
# Higher log-OR = treatment increases adverse events = HARM.
#
# FACTOR CONVENTION: ForestSearch defines subgroups via variable <= cut.
# For a binary factor with levels "0","1" -> numeric 1,2:
#   variable <= 1 selects level "0" (the LOWER level).
# So the HARM subgroup must be factor level "0".
# ============================================================================
sim_binary <- function(n, seed, has_hte = TRUE) {
  set.seed(seed)
  df <- data.frame(
    id    = seq_len(n),
    treat = rbinom(n, 1, 0.5),
    bm1   = as.factor(rbinom(n, 1, 0.70)),   # 30% are "0" = harm subgroup
    bm2   = as.factor(rbinom(n, 1, 0.50)),   # noise
    bm3   = as.factor(rbinom(n, 1, 0.45)),   # noise
    age   = round(rnorm(n, 55, 10)),
    ecog  = as.factor(sample(0:1, n, replace = TRUE, prob = c(0.6, 0.4)))
  )
  # Baseline progression rate ~30% (control)
  lp <- with(df, {
    -0.85 +                                  # baseline logit (~30%)
      0.02 * (age - 55) +                    # age effect
      0.3 * (ecog == 1)                      # ECOG effect
  })
  if (has_hte) {
    # Harm (bm1=0):       OR = 2.0 -> treatment INCREASES progression
    # Complement (bm1=1): OR = 0.4 -> treatment DECREASES progression
    lp <- lp + with(df, {
      0.69 * treat * (bm1 == 0) +            # log(2.0) = harm
        -0.92 * treat * (bm1 == 1)           # log(0.4) = benefit
    })
  }
  df$progressed <- rbinom(n, 1, plogis(lp))
  df$flag_harm  <- as.integer(df$bm1 == 0)
  df
}


# ============================================================================
# Continuous DGM: RA DAS28 change from baseline (higher = worse)
# ForestSearch defaults to adverse_outcome=FALSE for continuous.
# Since higher DAS28 change = worse, we pass adverse_outcome=TRUE
# so GRF screening looks for subgroups where treatment INCREASES Y.
# ============================================================================
sim_continuous <- function(n, seed, has_hte = TRUE) {
  set.seed(seed)
  df <- data.frame(
    id       = seq_len(n),
    treat    = rbinom(n, 1, 0.5),
    sero     = as.factor(rbinom(n, 1, 0.65)),  # 35% are "0" = harm subgroup
    male     = as.factor(rbinom(n, 1, 0.40)),  # noise
    smoking  = as.factor(rbinom(n, 1, 0.25)),  # noise
    age      = round(rnorm(n, 55, 12)),
    bas_das  = round(rnorm(n, 6.0, 0.8), 1)   # baseline DAS28
  )
  # Placebo effect ~-1.5; treatment adds ~-1.0 in complement
  mu <- with(df, {
    -1.5 +                                     # placebo effect
      -0.3 * (bas_das - 6.0)                   # regression to mean
  })
  if (has_hte) {
    # Harm (sero=0):       MD = +0.5 (treatment worse than placebo)
    # Complement (sero=1): MD = -1.0 (treatment beneficial)
    mu <- mu + with(df, {
      0.5 * treat * (sero == 0) +
        -1.0 * treat * (sero == 1)
    })
  } else {
    mu <- mu - 0.5 * df$treat                  # uniform mild benefit
  }
  df$das28_chg  <- round(mu + rnorm(n, 0, 1.5), 2)
  df$flag_harm  <- as.integer(df$sero == 0)
  df
}


# ============================================================================
# Count DGM: COPD exacerbations
# ============================================================================
sim_count <- function(n, seed, has_hte = TRUE) {
  set.seed(seed)
  df <- data.frame(
    id          = seq_len(n),
    treat       = rbinom(n, 1, 0.5),
    eos_low     = as.factor(rbinom(n, 1, 0.65)),  # 35% are "0" = harm subgroup
    smoking     = as.factor(rbinom(n, 1, 0.50)),   # noise
    male        = as.factor(rbinom(n, 1, 0.45)),   # noise
    age         = round(rnorm(n, 60, 12)),
    follow_up   = round(runif(n, 0.5, 3.0), 2)
  )
  log_rate <- with(df, {
    -1.0 + 0.02 * (age - 60) + log(follow_up)
  })
  if (has_hte) {
    log_rate <- log_rate + with(df, {
      1.0  * treat * (eos_low == 0) +          # harm: IRR = e^1 = 2.7
        -0.5 * treat * (eos_low == 1)           # benefit: IRR = e^-0.5 = 0.61
    })
  }
  df$exacerbations <- rpois(n, lambda = exp(log_rate))
  df$flag_harm     <- as.integer(df$eos_low == 0)
  df
}
```


# Simulation Scenarios {#sec-scenarios}

```{r scenario-table}
#| tbl-cap: "Simulation scenarios"

scenarios <- data.frame(
  ID = c("B-H0", "B-H1-S", "B-H1-L",
         "C-H0", "C-H1",
         "K-H0", "K-H1"),
  Outcome = c("Binary", "Binary", "Binary",
              "Continuous", "Continuous",
              "Count", "Count"),
  Hypothesis = c("Null", "Alternative", "Alternative",
                 "Null", "Alternative",
                 "Null", "Alternative"),
  N = c(200, 200, 800,
        600, 600,
        1000, 1000),
  Setting = c("Phase II oncology", "Phase II oncology", "Phase III oncology",
              "Phase III RA", "Phase III RA",
              "Phase III COPD", "Phase III COPD"),
  `Harm Prevalence` = c("--", "30%", "30%",
                         "--", "35%",
                         "--", "35%"),
  `True Effect (H)` = c("--", "OR=2.0", "OR=2.0",
                          "--", "MD=+0.5",
                          "--", "IRR=2.7"),
  `True Effect (Hc)` = c("--", "OR=0.4", "OR=0.4",
                           "--", "MD=-1.0",
                           "--", "IRR=0.61"),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt(scenarios) |>
  tab_header(
    title = "Simulation Scenarios",
    subtitle = "Inspired by the mCART framework (Rigdon et al., 2018)"
  ) |>
  tab_row_group(label = "Count (COPD exacerbations)",
                rows = Outcome == "Count") |>
  tab_row_group(label = "Continuous (DAS28 change)",
                rows = Outcome == "Continuous") |>
  tab_row_group(label = "Binary (disease progression)",
                rows = Outcome == "Binary")
```


# Simulation Engine {#sec-engine}

```{r run-one-sim}
#| code-fold: false

run_one_sim <- function(scenario_id, sim_num) {

  # ---- Generate data ----
  seed <- sim_seed + sim_num

  result <- switch(scenario_id,
    "B-H0"   = {
      df <- sim_binary(200, seed, has_hte = FALSE)
      list(df = df, ot = "binary",  em = "OR",  n = 200,
           confounders = c("bm1", "bm2", "bm3", "age", "ecog"),
           on = "progressed", en = "progressed", tn = "treat",
           offset = NULL, adverse = TRUE,
           true_H = NA, true_Hc = NA)
    },
    "B-H1-S" = {
      df <- sim_binary(200, seed, has_hte = TRUE)
      list(df = df, ot = "binary",  em = "OR",  n = 200,
           confounders = c("bm1", "bm2", "bm3", "age", "ecog"),
           on = "progressed", en = "progressed", tn = "treat",
           offset = NULL, adverse = TRUE,
           true_H = log(2.0), true_Hc = log(0.4))
    },
    "B-H1-L" = {
      df <- sim_binary(800, seed, has_hte = TRUE)
      list(df = df, ot = "binary",  em = "OR",  n = 800,
           confounders = c("bm1", "bm2", "bm3", "age", "ecog"),
           on = "progressed", en = "progressed", tn = "treat",
           offset = NULL, adverse = TRUE,
           true_H = log(2.0), true_Hc = log(0.4))
    },
    "C-H0"   = {
      df <- sim_continuous(600, seed, has_hte = FALSE)
      list(df = df, ot = "continuous", em = "MD", n = 600,
           confounders = c("sero", "male", "smoking", "age"),
           on = "das28_chg", en = "das28_chg", tn = "treat",
           offset = NULL, adverse = TRUE,
           true_H = NA, true_Hc = NA)
    },
    "C-H1"   = {
      df <- sim_continuous(600, seed, has_hte = TRUE)
      list(df = df, ot = "continuous", em = "MD", n = 600,
           confounders = c("sero", "male", "smoking", "age"),
           on = "das28_chg", en = "das28_chg", tn = "treat",
           offset = NULL, adverse = TRUE,
           true_H = 0.5, true_Hc = -1.0)
    },
    "K-H0"   = {
      df <- sim_count(1000, seed, has_hte = FALSE)
      list(df = df, ot = "count",  em = "IRR",  n = 1000,
           confounders = c("eos_low", "smoking", "male", "age"),
           on = "exacerbations", en = "exacerbations", tn = "treat",
           offset = "follow_up", adverse = TRUE,
           true_H = NA, true_Hc = NA)
    },
    "K-H1"   = {
      df <- sim_count(1000, seed, has_hte = TRUE)
      list(df = df, ot = "count",  em = "IRR",  n = 1000,
           confounders = c("eos_low", "smoking", "male", "age"),
           on = "exacerbations", en = "exacerbations", tn = "treat",
           offset = "follow_up", adverse = TRUE,
           true_H = 1.0, true_Hc = -0.5)
    }
  )

  # ---- Run forestsearch ----
  fs_error <- NA_character_
  fs <- tryCatch(
    forestsearch(
      df.analysis      = result$df,
      confounders.name = result$confounders,
      outcome.name     = result$on,
      event.name       = result$en,
      treat.name       = result$tn,
      id.name          = "id",
      outcome_type     = result$ot,
      effect_measure   = result$em,
      offset.name      = result$offset,
      adverse_outcome  = result$adverse,
      hr.threshold     = if (result$em == "MD") 0.0 else 1.25,
      hr.consistency   = if (result$em == "MD") 0.0 else 1.0,
      pconsistency.threshold = 0.50,
      fs.splits        = 200,
      n.min            = max(30, round(result$n * 0.08)),
      maxk             = 2,
      use_lasso        = TRUE,
      use_grf          = TRUE,
      use_twostage     = TRUE,
      is.RCT           = TRUE,
      seedit           = seed,
      details          = FALSE,
      plot.sg          = FALSE,
      parallel_args    = list(plan = "sequential", workers = 1,
                              show_message = FALSE)
    ),
    error = function(e) { fs_error <<- conditionMessage(e); NULL }
  )

  # ---- Extract results ----
  found <- !is.null(fs) && !is.null(fs$sg.harm) && length(fs$sg.harm) > 0

  out <- data.table(
    scenario  = scenario_id,
    sim       = sim_num,
    any_found = as.integer(found),
    sg_found  = if (found) paste(fs$sg.harm, collapse = " & ") else NA_character_,
    n_total   = result$n,
    error_msg = fs_error
  )

  if (found && !is.null(fs$df.est) &&
      "treat.recommend" %in% names(fs$df.est)) {
    sg_ind <- fs$df.est$treat.recommend
    flag   <- result$df$flag_harm

    # Classification metrics
    tp <- sum(sg_ind == 0 & flag == 1)
    fp <- sum(sg_ind == 0 & flag == 0)
    tn <- sum(sg_ind == 1 & flag == 0)
    fn <- sum(sg_ind == 1 & flag == 1)

    out$sens <- if ((tp + fn) > 0) tp / (tp + fn) else NA_real_
    out$spec <- if ((tn + fp) > 0) tn / (tn + fp) else NA_real_
    out$ppv  <- if ((tp + fp) > 0) tp / (tp + fp) else NA_real_
    out$npv  <- if ((tn + fn) > 0) tn / (tn + fn) else NA_real_
    out$size_H <- sum(sg_ind == 0)
  } else {
    out$sens <- out$spec <- out$ppv <- out$npv <- NA_real_
    out$size_H <- NA_integer_
  }

  out
}
```


# Run Simulations {#sec-run}

## Single-Trial Diagnostic

Trace each step of the pipeline independently to pinpoint the failure.

```{r single-trial-diagnostic}
#| code-fold: false

cat("=== Step-by-step pipeline diagnostic ===\n\n")

# Generate one binary H1 dataset
set.seed(sim_seed + 1)
df_diag <- sim_binary(800, sim_seed + 1, has_hte = TRUE)

cat("--- Step 0: Data summary ---\n")
cat("N:", nrow(df_diag), "| harm (bm1=0):", sum(df_diag$bm1 == 0), "\n")
cat("Progression rates:\n")
for (bm in c(0, 1)) {
  for (arm in c(0, 1)) {
    idx <- df_diag$bm1 == bm & df_diag$treat == arm
    cat(sprintf("  bm1=%d, treat=%d: %.3f (n=%d)\n",
                bm, arm, mean(df_diag$progressed[idx]), sum(idx)))
  }
}

cat("\n--- Step 1: Direct GLM estimation (known subgroup) ---\n")
df_H  <- df_diag[df_diag$bm1 == 0, ]
df_Hc <- df_diag[df_diag$bm1 == 1, ]

row_H  <- create_glm_row(df_H,  "progressed", "treat", "log_OR", NULL, "none", 0.95)
row_Hc <- create_glm_row(df_Hc, "progressed", "treat", "log_OR", NULL, "none", 0.95)
row_all <- create_glm_row(df_diag, "progressed", "treat", "log_OR", NULL, "none", 0.95)

cat(sprintf("  ITT:  log-OR = %.3f (SE=%.3f)\n", row_all$est, row_all$se))
cat(sprintf("  H:    log-OR = %.3f (SE=%.3f)  [true = +0.69]\n", row_H$est, row_H$se))
cat(sprintf("  Hc:   log-OR = %.3f (SE=%.3f)  [true = -0.92]\n", row_Hc$est, row_Hc$se))

cat("\n--- Step 2: GRF standalone ---\n")
grf_diag <- grf.subg.harm.glm(
  data             = df_diag,
  confounders.name = c("bm1", "bm2", "bm3", "age", "ecog"),
  outcome.name     = "progressed",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "OR",
  adverse_outcome  = TRUE,
  n.min            = 30,
  seedit           = sim_seed,
  verbose          = TRUE
)
cat("GRF subgroup:", paste(grf_diag$sg.harm.id %||% "none"), "\n")
cat("Variable importance:\n")
print(sort(grf_diag$grf_varimp, decreasing = TRUE))

cat("\n--- Step 3: forestsearch (verbose) ---\n")
cat("Calling forestsearch with details=TRUE...\n")
fs_diag <- forestsearch(
  df.analysis      = df_diag,
  confounders.name = c("bm1", "bm2", "bm3", "age", "ecog"),
  outcome.name     = "progressed",
  event.name       = "progressed",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "OR",
  adverse_outcome  = TRUE,
  hr.threshold     = 1.25,
  hr.consistency   = 1.0,
  pconsistency.threshold = 0.50,
  fs.splits        = 200,
  n.min            = 30,
  maxk             = 2,
  use_lasso        = TRUE,
  use_grf          = TRUE,
  use_twostage     = TRUE,
  is.RCT           = TRUE,
  seedit           = sim_seed,
  details          = TRUE,
  plot.sg          = FALSE,
  parallel_args    = list(plan = "sequential", workers = 1,
                          show_message = FALSE)
)

cat("\nforestsearch result:\n")
cat("  sg.harm:", paste(fs_diag$sg.harm %||% "NULL"), "\n")
if (!is.null(fs_diag$df.est)) {
  cat("  df.est columns:", paste(names(fs_diag$df.est), collapse = ", "), "\n")
  if ("treat.recommend" %in% names(fs_diag$df.est)) {
    cat("  treat.recommend distribution:\n")
    print(table(fs_diag$df.est$treat.recommend))
  }
}

# Check what the search returned
cat("\n--- Step 4b: Dump installed forestsearch GLM code path ---\n")
# Deparse the installed forestsearch function to find the GLM/FSdata section
fs_body <- capture.output(body(forestsearch::forestsearch))
# Find lines around get_FSdata, filter_call_args, FSdata, sg.harm = NULL
patterns <- c("FSdata", "filter_call", "sg.harm", "outcome_type.*!=.*survival",
              "get_FSdata", "return.*list")
for (pat in patterns) {
  idx <- grep(pat, fs_body, ignore.case = TRUE)
  if (length(idx) > 0) {
    cat(sprintf("  Pattern '%s' found at lines: %s\n", pat, paste(idx, collapse = ", ")))
    for (i in idx[1:min(2, length(idx))]) {
      start <- max(1, i - 2)
      end <- min(length(fs_body), i + 2)
      cat(paste("    ", fs_body[start:end], collapse = "\n"), "\n    ---\n")
    }
  }
}
# First check what get_FSdata accepts
cat("  get_FSdata signature:\n")
cat("  ", paste(names(formals(get_FSdata)), collapse = ", "), "\n\n")

# Call get_FSdata with ONLY the parameters it accepts
tryCatch(
  withCallingHandlers(
    {
      fsdata_test <- get_FSdata(
        df.analysis      = df_diag,
        confounders.name = c("bm1", "bm2", "bm3", "age", "ecog"),
        outcome.name     = "progressed",
        event.name       = "progressed",
        outcome_type     = "binary",
        cont.cutoff      = 4,
        use_lasso        = TRUE,
        grf_cuts         = NULL
      )
      cat("  get_FSdata returned successfully!\n")
      cat("  Class:", class(fsdata_test), "\n")
      cat("  Names:", paste(names(fsdata_test), collapse = ", "), "\n")
      if (!is.null(fsdata_test$confs_names)) {
        cat("  Candidate factors:", length(fsdata_test$confs_names), "\n")
        cat("  ", paste(head(fsdata_test$confs_names, 10), collapse = ", "), "\n")
      }
    },
    warning = function(w) {
      cat("  WARNING:", conditionMessage(w), "\n")
      invokeRestart("muffleWarning")
    }
  ),
  error = function(e) {
    cat("  ERROR:", conditionMessage(e), "\n")
    cat("  Traceback:\n")
    calls <- sys.calls()
    for (i in seq_along(calls)) {
      cat("  ", i, ":", deparse(calls[[i]], nlines = 1), "\n")
    }
  }
)

cat("\n--- Step 4c: get_FSdata with grf_cuts (reproduce failure) ---\n")
# Get the actual grf_cuts that forestsearch would pass
grf_cuts_actual <- grf_diag$tree.cuts
cat("  grf_cuts class:", class(grf_cuts_actual), "\n")
cat("  grf_cuts length:", length(grf_cuts_actual), "\n")
if (is.list(grf_cuts_actual)) {
  cat("  grf_cuts names:", paste(names(grf_cuts_actual), collapse = ", "), "\n")
  str(grf_cuts_actual, max.level = 2)
} else {
  cat("  grf_cuts value:", paste(head(grf_cuts_actual, 10), collapse = ", "), "\n")
}

# Now call get_FSdata with the actual grf_cuts
cat("\n  Calling get_FSdata with grf_cuts from GRF...\n")
tryCatch(
  withCallingHandlers(
    {
      fsdata_test2 <- get_FSdata(
        df.analysis      = df_diag,
        confounders.name = c("bm1", "bm2", "bm3", "age", "ecog"),
        outcome.name     = "progressed",
        event.name       = "progressed",
        outcome_type     = "binary",
        cont.cutoff      = 4,
        use_lasso        = TRUE,
        use_grf          = TRUE,
        grf_cuts         = grf_cuts_actual
      )
      cat("  SUCCESS with grf_cuts!\n")
      cat("  Candidate factors:", length(fsdata_test2$confs_names), "\n")
    },
    warning = function(w) {
      cat("  WARNING:", conditionMessage(w), "\n")
      invokeRestart("muffleWarning")
    }
  ),
  error = function(e) {
    cat("  ERROR:", conditionMessage(e), "\n")
  }
)
cat("\n--- Step 5: Examine forestsearch internals ---\n")
cat("  Names in result:", paste(names(fs_diag), collapse = ", "), "\n")
for (nm in c("sg.harm", "hr.subgroups", "details.subgroups",
             "subgroups", "grf_res")) {
  val <- fs_diag[[nm]]
  if (!is.null(val)) {
    cat(sprintf("  $%s: %s (length %d)\n", nm, class(val)[1], length(val)))
  } else {
    cat(sprintf("  $%s: NULL\n", nm))
  }
}
```

## Full Simulation

```{r run-all}
t0 <- proc.time()[3]

scenario_ids <- c("B-H0", "B-H1-S", "B-H1-L",
                  "C-H0", "C-H1",
                  "K-H0", "K-H1")

results_list <- foreach(
  sc = rep(scenario_ids, each = nsims),
  sim = rep(seq_len(nsims), times = length(scenario_ids)),
  .errorhandling = "pass",
  .options.future = list(packages = c("forestsearch", "data.table"))
) %dofuture% {
  run_one_sim(sc, sim)
}

all_results <- rbindlist(results_list, fill = TRUE)

timings$simulations <- proc.time()[3] - t0
cat(sprintf("Total simulation time: %.1f sec (%.1f min)\n",
            timings$simulations, timings$simulations / 60))
cat(sprintf("Scenarios: %d | Sims per scenario: %d | Total: %d\n",
            length(scenario_ids), nsims, nrow(all_results)))
```

## Diagnostic: Error Messages

```{r diagnostics}
# Show error summary
errors <- all_results[!is.na(error_msg)]
if (nrow(errors) > 0) {
  cat("=== ERRORS DETECTED ===\n")
  cat("Total errors:", nrow(errors), "of", nrow(all_results), "simulations\n\n")

  error_summary <- errors[, .N, by = .(scenario, error_msg)]
  setorder(error_summary, scenario, -N)
  for (i in seq_len(nrow(error_summary))) {
    cat(sprintf("[%s] %d errors: %s\n",
                error_summary$scenario[i],
                error_summary$N[i],
                substr(error_summary$error_msg[i], 1, 120)))
  }
} else {
  cat("No errors detected.\n")
}

cat("\n=== Detection summary ===\n")
det_summary <- all_results[, .(
  n_sims = .N,
  n_found = sum(any_found),
  n_errors = sum(!is.na(error_msg))
), by = scenario]
print(det_summary)
```


# Results {#sec-results}

## Summary Table

```{r summary-table}
#| tbl-cap: "Operating characteristics by scenario"

summary_dt <- all_results[, .(
  N        = .N,
  Power    = round(mean(any_found, na.rm = TRUE), 3),
  Sens     = round(mean(sens, na.rm = TRUE), 3),
  Spec     = round(mean(spec, na.rm = TRUE), 3),
  PPV      = round(mean(ppv, na.rm = TRUE), 3),
  NPV      = round(mean(npv, na.rm = TRUE), 3),
  Mean_nH  = round(mean(size_H, na.rm = TRUE), 0)
), by = scenario]

# Merge with scenario metadata
meta <- data.table(scenarios)[, .(ID, Outcome, Hypothesis, N, Setting)]
setnames(meta, "ID", "scenario")
setnames(meta, "N", "Sample_N")

summary_merged <- merge(meta, summary_dt, by = "scenario")
setorder(summary_merged, Outcome, Hypothesis)

gt(summary_merged) |>
  tab_header(
    title = "ForestSearch GLM Operating Characteristics",
    subtitle = sprintf("%d simulations per scenario | %d workers",
                        nsims, n_workers)
  ) |>
  cols_label(
    scenario = "Scenario", Outcome = "Outcome", Hypothesis = "H",
    Sample_N = "N", Setting = "Setting", N = "Sims",
    Power = "Detection", Sens = "Sens", Spec = "Spec",
    PPV = "PPV", NPV = "NPV", Mean_nH = "Mean n(H)"
  ) |>
  tab_spanner(label = "Classification Metrics",
              columns = c(Sens, Spec, PPV, NPV)) |>
  tab_source_note(
    sprintf("Runtime: %.1f min | Seed: %d",
            timings$simulations / 60, sim_seed)
  )
```


## Detection Rates

```{r detection-plot}
#| fig-height: 5
#| fig-cap: "Detection rates by scenario. H0 scenarios show Type I error; H1 scenarios show power."

det_dt <- all_results[, .(
  Detection = mean(any_found, na.rm = TRUE)
), by = scenario]

det_dt <- merge(det_dt, meta, by = "scenario")

det_dt$label <- sprintf("%s\n(n=%s)", det_dt$scenario, det_dt$Sample_N)
det_dt$label <- factor(det_dt$label, levels = det_dt$label[order(det_dt$Outcome, det_dt$Hypothesis)])

ggplot(det_dt, aes(x = label, y = Detection, fill = Hypothesis)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", Detection)),
            vjust = -0.5, size = 3.5) +
  scale_fill_manual(values = c("Null" = "grey70", "Alternative" = "steelblue")) +
  facet_wrap(~ Outcome, scales = "free_x", nrow = 1) +
  labs(x = NULL, y = "Detection Rate",
       title = "Subgroup Detection Rates by Scenario",
       subtitle = sprintf("%d simulations per scenario", nsims)) +
  ylim(0, 1.1) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", axis.text.x = element_text(size = 8),
        plot.title = element_text(face = "bold"))
```


## Classification Metrics (H1 Only)

```{r classification-plot}
#| fig-height: 5
#| fig-cap: "Classification metrics for alternative hypothesis scenarios (where subgroup was found)."

h1_dt <- all_results[grepl("H1", scenario) & any_found == 1]

if (nrow(h1_dt) > 0) {
  class_long <- melt(h1_dt,
    id.vars = "scenario",
    measure.vars = c("sens", "spec", "ppv", "npv"),
    variable.name = "Metric", value.name = "Value"
  )
  class_long$Metric <- toupper(as.character(class_long$Metric))

  ggplot(class_long, aes(x = Metric, y = Value, fill = Metric)) +
    geom_boxplot(outlier.size = 1, alpha = 0.7) +
    facet_wrap(~ scenario, nrow = 1) +
    scale_fill_brewer(palette = "Set2") +
    labs(x = NULL, y = "Value",
         title = "Classification Metrics (H1 Scenarios)",
         subtitle = "Distributions across simulations where subgroup was found") +
    ylim(0, 1) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none",
          plot.title = element_text(face = "bold"))
} else {
  cat("No subgroups found in any H1 simulation.\n")
}
```


## Subgroup Size Distribution

```{r size-plot}
#| fig-height: 4
#| fig-cap: "Distribution of identified harm subgroup size across simulations"

h1_found <- all_results[grepl("H1", scenario) & any_found == 1]

if (nrow(h1_found) > 0) {
  ggplot(h1_found, aes(x = scenario, y = size_H, fill = scenario)) +
    geom_boxplot(alpha = 0.7) +
    geom_hline(aes(yintercept = n_total * 0.30),
               linetype = "dashed", color = "red", linewidth = 0.5) +
    labs(x = "Scenario", y = "n(H) identified",
         title = "Identified Harm Subgroup Size",
         subtitle = "Dashed line: true harm prevalence (30-35 percent of N)") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none",
          plot.title = element_text(face = "bold"))
} else {
  cat("No subgroups found in any H1 simulation.\n")
}
```


## Most Frequently Identified Subgroups

```{r freq-subgroups}
#| tbl-cap: "Top subgroups identified across simulations"

h1_found <- all_results[grepl("H1", scenario) & any_found == 1]

if (nrow(h1_found) > 0) {
  freq_dt <- h1_found[, .N, by = .(scenario, sg_found)]
  freq_dt <- freq_dt[order(scenario, -N)]

  # Top 3 per scenario
  freq_top <- freq_dt[, head(.SD, 3), by = scenario]
  freq_top$Pct <- round(100 * freq_top$N /
    freq_dt[, sum(N), by = scenario][freq_top, on = "scenario"]$V1, 1)

  gt(freq_top) |>
    tab_header(
      title = "Most Frequently Identified Subgroups",
      subtitle = "Top 3 per scenario (among simulations where subgroup found)"
    ) |>
    cols_label(scenario = "Scenario", sg_found = "Subgroup Definition",
               N = "Count", Pct = "Percent")
} else {
  cat("No subgroups found in any H1 simulation.\n")
}
```


# Timing Summary {#sec-timing}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Stage = c("All simulations", "Total (including setup)"),
  Seconds = c(timings$simulations, timings$total),
  Minutes = c(timings$simulations / 60, timings$total / 60),
  stringsAsFactors = FALSE
)

gt(timing_df) |>
  tab_header(
    title = "Computational Timing",
    subtitle = sprintf("%d total simulations | %d workers",
                        nsims * length(scenario_ids), n_workers)
  ) |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)", Minutes = "Time (min)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total (including setup)")
  )
```


# Summary {#sec-summary}

This simulation study evaluated `forestsearch()` across three GLM
outcome types and seven scenarios inspired by the mCART framework.

**Key findings:**

- **Type I error** (H0 scenarios): Detection rates under the null
  hypothesis quantify the false positive rate for each outcome type.
  ForestSearch's split-sample consistency criterion provides built-in
  multiplicity protection.

- **Power** (H1 scenarios): Detection rates under the alternative
  depend on sample size, effect magnitude, and the consistency
  criterion stringency.  The small-sample Phase II binary scenario
  (B-H1-S, n = 200) is the most challenging; the larger Phase III
  scenarios have higher power.

- **Classification accuracy**: When a subgroup is found, sensitivity,
  specificity, PPV, and NPV characterize how well the identified
  subgroup matches the true harm subgroup.

- **Outcome type differences**: Count data faces higher variance in
  half-sample splits (fewer events per arm), making the consistency
  criterion harder to satisfy.  Continuous outcomes are the most
  stable.  Binary outcomes in small samples (n = 200) face
  separation issues.


## Comparison to mCART Results

The mCART paper found that in the Phase II oncology scenario (n = 200),
mCART did not identify the true subgroups (false negatives), while LR,
RF, and gradient RF found mostly false positives.  ForestSearch's
consistency criterion is expected to behave similarly: conservative in
small samples (few false positives, but also lower power) and more
powerful in larger samples where the signal is strong enough to
replicate across random splits.


# References {#sec-references}

- Rigdon, J., Basu, S., Engel, R., Baiocchi, M. (2018). Preventing
  false discovery of heterogeneous treatment effect subgroups in
  randomized trials. *Trials*, 19, 382.

- Leon, L.F., et al. (2024). Exploratory Subgroup Identification.
  *Statistics in Medicine*, 43, 4253--4275. doi:10.1002/sim.10163

- DINA: Estimating Heterogeneous Treatment Effects for General
  Responses. *Biometrics*, 81(4), 2025.

- Foster, J.C., et al. (2011). Subgroup identification from
  randomized clinical trial data. *Statistics in Medicine*, 30,
  2867--2880. (Virtual Twins)


# Session Information {#sec-session}

```{r session-info}
plan(sequential)
sessionInfo()
```
