---
title: "GLM Subgroup Identification: Simulation Study"
subtitle: "Operating Characteristics with Calibrated Thresholds Across Binary, Continuous, and Count Outcomes"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ============================================================================
# SIMULATION PARAMETERS --- edit here to tune runtime vs precision
# ============================================================================
nsims       <- 100L       # simulations per scenario (increase for production)
sim_seed    <- 8316951L
n_workers   <- max(1L, parallel::detectCores(logical = FALSE) - 1L)

# L_eff calibration (from binary_threshold_calibration.qmd, 5000 sims).
# Used for binary threshold selection.  Count outcomes use the same
# L_eff as an approximation (same number of confounders; L_eff depends
# on the candidate generation process, not the outcome type).
# Continuous uses L_eff = 1 (high d_eff makes P1 small enough without
# correction).
L_eff_C_binary     <- 0.220
L_eff_alpha_binary <- 1.298

# FPR target for threshold selection
fpr_target <- 0.12
```


# Introduction {#sec-intro}

## Motivation

This document evaluates the operating characteristics of `forestsearch()`
for GLM outcomes across a systematic set of simulation scenarios inspired
by the mCART framework (Rigdon et al., *Trials*, 2018).

We adapt this framework for `forestsearch` across three GLM outcome types:

| Outcome Type | Clinical Analog | Effect Measure | Key Scenarios |
|---|---|---|---|
| **Binary** | Disease progression (oncology) | OR | Phase II (small n) and Phase III |
| **Continuous** | DAS28 change (RA) | MD | Phase III immunology |
| **Count** | Exacerbations (COPD) | IRR | Phase III respiratory |

**Key feature of this study**: Thresholds $(c_1, c_2)$ are selected
**per outcome type** using the corrected approximation from
`suggest_thresholds()`, rather than using the survival-calibrated paper
defaults (c1 = 1.25, c2 = 1.0) for all outcomes.


## Evaluation Metrics

| Metric | Definition |
|---|---|
| Detection rate (power) | Proportion of simulations where any subgroup found |
| Type I error | Detection rate under H0 |
| Sensitivity | Proportion of true harm patients correctly classified |
| Specificity | Proportion of true non-harm patients correctly classified |
| PPV | Among identified harm patients, proportion truly harmed |
| NPV | Among identified non-harm patients, proportion truly not harmed |


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)

# Load latest source (including suggest_thresholds.R)
#if (requireNamespace("devtools", quietly = TRUE)) {
#  try(devtools::load_all(here::here(), quiet = TRUE), silent = TRUE)
#}

```

```{r parallel-setup}
plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")
cat("Simulations per scenario:", nsims, "\n")
```

```{r timing-init}
timings <- list()
t_vignette_start <- proc.time()
```


# Data-Generating Mechanisms {#sec-dgm}

```{r dgm-functions}
#| code-fold: false

# ============================================================================
# Binary DGM: Oncology disease progression (Y=1 = adverse event)
# FACTOR CONVENTION: ForestSearch defines subgroups via variable <= cut.
# For a binary factor with levels "0","1" -> numeric 1,2:
#   variable <= 1 selects level "0" (the LOWER level).
# So the HARM subgroup must be factor level "0".
# ============================================================================
sim_binary <- function(n, seed, has_hte = TRUE) {
  set.seed(seed)
  df <- data.frame(
    id    = seq_len(n),
    treat = rbinom(n, 1, 0.5),
    bm1   = as.factor(rbinom(n, 1, 0.70)),   # 30% are "0" = harm subgroup
    bm2   = as.factor(rbinom(n, 1, 0.50)),   # noise
    bm3   = as.factor(rbinom(n, 1, 0.45)),   # noise
    age   = round(rnorm(n, 55, 10)),
    ecog  = as.factor(sample(0:1, n, replace = TRUE, prob = c(0.6, 0.4)))
  )
  lp <- with(df, {
    -0.85 + 0.02 * (age - 55) + 0.3 * (ecog == 1)
  })
  if (has_hte) {
    lp <- lp + with(df, {
      0.69 * treat * (bm1 == 0) +            # log(2.0) = harm
        -0.92 * treat * (bm1 == 1)           # log(0.4) = benefit
    })
  }
  df$progressed <- rbinom(n, 1, plogis(lp))
  df$flag_harm  <- as.integer(df$bm1 == 0)
  df
}


# ============================================================================
# Continuous DGM: RA DAS28 change from baseline (higher = worse)
# ============================================================================
sim_continuous <- function(n, seed, has_hte = TRUE) {
  set.seed(seed)
  df <- data.frame(
    id       = seq_len(n),
    treat    = rbinom(n, 1, 0.5),
    sero     = as.factor(rbinom(n, 1, 0.65)),  # 35% are "0" = harm subgroup
    male     = as.factor(rbinom(n, 1, 0.40)),  # noise
    smoking  = as.factor(rbinom(n, 1, 0.25)),  # noise
    age      = round(rnorm(n, 55, 12)),
    bas_das  = round(rnorm(n, 6.0, 0.8), 1)
  )
  mu <- with(df, {
    -1.5 + -0.3 * (bas_das - 6.0)
  })
  if (has_hte) {
    mu <- mu + with(df, {
      0.5 * treat * (sero == 0) +
        -1.0 * treat * (sero == 1)
    })
  } else {
    mu <- mu - 0.5 * df$treat
  }
  df$das28_chg  <- round(mu + rnorm(n, 0, 1.5), 2)
  df$flag_harm  <- as.integer(df$sero == 0)
  df
}


# ============================================================================
# Count DGM: COPD exacerbations
# ============================================================================
sim_count <- function(n, seed, has_hte = TRUE) {
  set.seed(seed)
  df <- data.frame(
    id          = seq_len(n),
    treat       = rbinom(n, 1, 0.5),
    eos_low     = as.factor(rbinom(n, 1, 0.65)),  # 35% are "0" = harm subgroup
    smoking     = as.factor(rbinom(n, 1, 0.50)),   # noise
    male        = as.factor(rbinom(n, 1, 0.45)),   # noise
    age         = round(rnorm(n, 60, 12)),
    follow_up   = round(runif(n, 0.5, 3.0), 2)
  )
  log_rate <- with(df, {
    -1.0 + 0.02 * (age - 60) + log(follow_up)
  })
  if (has_hte) {
    log_rate <- log_rate + with(df, {
      1.0  * treat * (eos_low == 0) +          # harm: IRR = e^1 = 2.7
        -0.5 * treat * (eos_low == 1)           # benefit: IRR = 0.61
    })
  }
  df$exacerbations <- rpois(n, lambda = exp(log_rate))
  df$flag_harm     <- as.integer(df$eos_low == 0)
  df
}
```


# Threshold Selection {#sec-thresholds}

The paper defaults (c1 = 1.25, c2 = 1.0) were calibrated for
survival endpoints.  For non-survival outcomes, we use
`suggest_thresholds()` to select $(c_1, c_2)$ per outcome type
based on the corrected approximation.

## Compute $d_{\text{eff}}$ for Each Scenario

```{r compute-deff}
#| code-fold: false

# Binary: 30% harm prevalence, p(event) ~ 0.30
# N = 200: n_H = 60;  N = 800: n_H = 240
d_eff_B200 <- d_eff_binary(n_sg = round(0.30 * 200), p_event = 0.30)
d_eff_B800 <- d_eff_binary(n_sg = round(0.30 * 800), p_event = 0.30)

# Continuous: 35% harm prevalence, sigma = 1.5, N = 600
d_eff_C600 <- d_eff_continuous(
  n_sg = round(0.35 * 600), sigma_y = 1.5)

# Count: 35% harm prevalence, N = 1000
# Mean rate ~ exp(-1.0) * mean(follow_up) ~ 0.37 * 1.75 ~ 0.65/person
# D = n_H * mu ~ 350 * 0.65 ~ 227
D_count <- round(0.35 * 1000 * 0.65)
d_eff_K1000 <- d_eff_count(total_events = D_count)

cat("=== d_eff by Scenario ===\n")
cat(sprintf("  Binary  N=200: d_eff = %.1f  (n_H=%d)\n",
            d_eff_B200, round(0.30 * 200)))
cat(sprintf("  Binary  N=800: d_eff = %.1f  (n_H=%d)\n",
            d_eff_B800, round(0.30 * 800)))
cat(sprintf("  Continuous N=600: d_eff = %.1f  (n_H=%d, sigma=1.5)\n",
            d_eff_C600, round(0.35 * 600)))
cat(sprintf("  Count   N=1000: d_eff = %.0f  (D=%d events)\n",
            d_eff_K1000, D_count))
```


## Suggest Thresholds

```{r suggest-thresholds}
#| tbl-cap: "Threshold recommendations from the corrected approximation"

# Binary (with L_eff correction)
ft_B200 <- suggest_thresholds(
  d_eff_B200, N = 200, fpr_target = fpr_target,
  L_eff_C = L_eff_C_binary, L_eff_alpha = L_eff_alpha_binary)
ft_B800 <- suggest_thresholds(
  d_eff_B800, N = 800, fpr_target = fpr_target,
  L_eff_C = L_eff_C_binary, L_eff_alpha = L_eff_alpha_binary)

# Continuous (difference scale, L_eff = 1)
ft_C600 <- suggest_thresholds(
  d_eff_C600, N = 600, fpr_target = fpr_target,
  effect_scale = "difference",
  c1_range = c(0.01, 1.0), c1_step = 0.01,
  c2_candidates = seq(0.0, 0.50, by = 0.01))

# Count (ratio scale, L_eff from binary calibration as approximation:
# same number of confounders, so candidate generation is similar)
ft_K1000 <- suggest_thresholds(
  d_eff_K1000, N = 1000, fpr_target = fpr_target,
  L_eff_C = L_eff_C_binary, L_eff_alpha = L_eff_alpha_binary)

# Build recommendation summary
rec_list <- list()
if (nrow(ft_B200) > 0) rec_list$B200 <- data.frame(
  Scenario = "Binary N=200", c1 = ft_B200$c1[1],
  c2 = ft_B200$min_c2[1], FPR_corr = ft_B200$FPR_corr[1],
  d_eff = d_eff_B200, Scale = "OR")
if (nrow(ft_B800) > 0) rec_list$B800 <- data.frame(
  Scenario = "Binary N=800", c1 = ft_B800$c1[1],
  c2 = ft_B800$min_c2[1], FPR_corr = ft_B800$FPR_corr[1],
  d_eff = d_eff_B800, Scale = "OR")
if (nrow(ft_C600) > 0) rec_list$C600 <- data.frame(
  Scenario = "Continuous N=600", c1 = ft_C600$c1[1],
  c2 = ft_C600$min_c2[1], FPR_corr = ft_C600$FPR_corr[1],
  d_eff = d_eff_C600, Scale = "MD")
if (nrow(ft_K1000) > 0) rec_list$K1000 <- data.frame(
  Scenario = "Count N=1000", c1 = ft_K1000$c1[1],
  c2 = ft_K1000$min_c2[1], FPR_corr = ft_K1000$FPR_corr[1],
  d_eff = d_eff_K1000, Scale = "IRR")

rec_df <- do.call(rbind, rec_list)

gt(rec_df) |>
  tab_header(
    title = md("**Threshold Recommendations (Smallest c1)**"),
    subtitle = sprintf("FPR target = %.0f%% | Binary and count use L_eff correction",
                        100 * fpr_target)
  ) |>
  fmt_number(columns = c(c1, c2, FPR_corr, d_eff), decimals = 2) |>
  tab_source_note(
    "Continuous thresholds on the mean-difference scale; all others on the ratio (OR/IRR) scale."
  )
```


## Threshold Lookup Table

```{r threshold-lookup}
#| code-fold: false

# Build lookup used by run_one_sim()
# Use smallest c1 from each frontier
thresh <- list(
  binary_small = list(
    c1 = if (nrow(ft_B200) > 0) ft_B200$c1[1] else 1.25,
    c2 = if (nrow(ft_B200) > 0) ft_B200$min_c2[1] else 1.0),
  binary_large = list(
    c1 = if (nrow(ft_B800) > 0) ft_B800$c1[1] else 1.25,
    c2 = if (nrow(ft_B800) > 0) ft_B800$min_c2[1] else 1.0),
  continuous = list(
    c1 = if (nrow(ft_C600) > 0) ft_C600$c1[1] else 0.0,
    c2 = if (nrow(ft_C600) > 0) ft_C600$min_c2[1] else 0.0),
  count = list(
    c1 = if (nrow(ft_K1000) > 0) ft_K1000$c1[1] else 1.25,
    c2 = if (nrow(ft_K1000) > 0) ft_K1000$min_c2[1] else 1.0)
)

cat("=== Threshold Lookup ===\n")
for (nm in names(thresh)) {
  cat(sprintf("  %-14s: c1 = %.2f, c2 = %.2f\n",
              nm, thresh[[nm]]$c1, thresh[[nm]]$c2))
}
```


# Simulation Scenarios {#sec-scenarios}

```{r scenario-table}
#| tbl-cap: "Simulation scenarios with calibrated thresholds"

scenarios <- data.frame(
  ID = c("B-H0", "B-H1-S", "B-H0-L", "B-H1-L",
         "C-H0", "C-H1",
         "K-H0", "K-H1"),
  Outcome = c("Binary", "Binary", "Binary", "Binary",
              "Continuous", "Continuous",
              "Count", "Count"),
  Hypothesis = c("Null", "Alternative", "Null", "Alternative",
                 "Null", "Alternative",
                 "Null", "Alternative"),
  N = c(200, 200, 800, 800,
        600, 600,
        1000, 1000),
  Setting = c("Phase II oncology", "Phase II oncology",
              "Phase III oncology", "Phase III oncology",
              "Phase III RA", "Phase III RA",
              "Phase III COPD", "Phase III COPD"),
  c1 = c(thresh$binary_small$c1, thresh$binary_small$c1,
         thresh$binary_large$c1, thresh$binary_large$c1,
         thresh$continuous$c1, thresh$continuous$c1,
         thresh$count$c1, thresh$count$c1),
  c2 = c(thresh$binary_small$c2, thresh$binary_small$c2,
         thresh$binary_large$c2, thresh$binary_large$c2,
         thresh$continuous$c2, thresh$continuous$c2,
         thresh$count$c2, thresh$count$c2),
  `True Effect (H)` = c("--", "OR=2.0", "--", "OR=2.0",
                          "--", "MD=+0.5",
                          "--", "IRR=2.7"),
  `True Effect (Hc)` = c("--", "OR=0.4", "--", "OR=0.4",
                           "--", "MD=-1.0",
                           "--", "IRR=0.61"),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

gt(scenarios) |>
  tab_header(
    title = "Simulation Scenarios with Calibrated Thresholds",
    subtitle = sprintf("FPR target = %.0f%%", 100 * fpr_target)
  ) |>
  tab_row_group(label = "Count (COPD exacerbations)",
                rows = Outcome == "Count") |>
  tab_row_group(label = "Continuous (DAS28 change)",
                rows = Outcome == "Continuous") |>
  tab_row_group(label = "Binary (disease progression)",
                rows = Outcome == "Binary")
```


# Simulation Engine {#sec-engine}

```{r run-one-sim}
#| code-fold: false

run_one_sim <- function(scenario_id, sim_num) {

  seed <- sim_seed + sim_num

  result <- switch(scenario_id,
    "B-H0"   = {
      df <- sim_binary(200, seed, has_hte = FALSE)
      list(df = df, ot = "binary", em = "OR", n = 200,
           confounders = c("bm1", "bm2", "bm3", "age", "ecog"),
           on = "progressed", en = "progressed", tn = "treat",
           offset = NULL, adverse = TRUE,
           c1 = thresh$binary_small$c1,
           c2 = thresh$binary_small$c2,
           true_H = NA, true_Hc = NA)
    },
    "B-H1-S" = {
      df <- sim_binary(200, seed, has_hte = TRUE)
      list(df = df, ot = "binary", em = "OR", n = 200,
           confounders = c("bm1", "bm2", "bm3", "age", "ecog"),
           on = "progressed", en = "progressed", tn = "treat",
           offset = NULL, adverse = TRUE,
           c1 = thresh$binary_small$c1,
           c2 = thresh$binary_small$c2,
           true_H = log(2.0), true_Hc = log(0.4))
    },
    "B-H0-L" = {
      df <- sim_binary(800, seed, has_hte = FALSE)
      list(df = df, ot = "binary", em = "OR", n = 800,
           confounders = c("bm1", "bm2", "bm3", "age", "ecog"),
           on = "progressed", en = "progressed", tn = "treat",
           offset = NULL, adverse = TRUE,
           c1 = thresh$binary_large$c1,
           c2 = thresh$binary_large$c2,
           true_H = NA, true_Hc = NA)
    },
    "B-H1-L" = {
      df <- sim_binary(800, seed, has_hte = TRUE)
      list(df = df, ot = "binary", em = "OR", n = 800,
           confounders = c("bm1", "bm2", "bm3", "age", "ecog"),
           on = "progressed", en = "progressed", tn = "treat",
           offset = NULL, adverse = TRUE,
           c1 = thresh$binary_large$c1,
           c2 = thresh$binary_large$c2,
           true_H = log(2.0), true_Hc = log(0.4))
    },
    "C-H0"   = {
      df <- sim_continuous(600, seed, has_hte = FALSE)
      list(df = df, ot = "continuous", em = "MD", n = 600,
           confounders = c("sero", "male", "smoking", "age"),
           on = "das28_chg", en = "das28_chg", tn = "treat",
           offset = NULL, adverse = TRUE,
           c1 = thresh$continuous$c1,
           c2 = thresh$continuous$c2,
           true_H = NA, true_Hc = NA)
    },
    "C-H1"   = {
      df <- sim_continuous(600, seed, has_hte = TRUE)
      list(df = df, ot = "continuous", em = "MD", n = 600,
           confounders = c("sero", "male", "smoking", "age"),
           on = "das28_chg", en = "das28_chg", tn = "treat",
           offset = NULL, adverse = TRUE,
           c1 = thresh$continuous$c1,
           c2 = thresh$continuous$c2,
           true_H = 0.5, true_Hc = -1.0)
    },
    "K-H0"   = {
      df <- sim_count(1000, seed, has_hte = FALSE)
      list(df = df, ot = "count", em = "IRR", n = 1000,
           confounders = c("eos_low", "smoking", "male", "age"),
           on = "exacerbations", en = "exacerbations", tn = "treat",
           offset = "follow_up", adverse = TRUE,
           c1 = thresh$count$c1,
           c2 = thresh$count$c2,
           true_H = NA, true_Hc = NA)
    },
    "K-H1"   = {
      df <- sim_count(1000, seed, has_hte = TRUE)
      list(df = df, ot = "count", em = "IRR", n = 1000,
           confounders = c("eos_low", "smoking", "male", "age"),
           on = "exacerbations", en = "exacerbations", tn = "treat",
           offset = "follow_up", adverse = TRUE,
           c1 = thresh$count$c1,
           c2 = thresh$count$c2,
           true_H = 1.0, true_Hc = -0.5)
    }
  )

  # ---- Run forestsearch ----
  fs_error <- NA_character_
  fs <- tryCatch(
    forestsearch(
      df.analysis      = result$df,
      confounders.name = result$confounders,
      outcome.name     = result$on,
      event.name       = result$en,
      treat.name       = result$tn,
      id.name          = "id",
      outcome_type     = result$ot,
      effect_measure   = result$em,
      offset.name      = result$offset,
      adverse_outcome  = result$adverse,
      hr.threshold     = result$c1,
      hr.consistency   = result$c2,
      pconsistency.threshold = 0.90,
      fs.splits        = 400,
      n.min            = 60,
      maxk             = 2,
      use_lasso        = TRUE,
      use_grf          = TRUE,
      use_twostage     = TRUE,
      is.RCT           = TRUE,
      seedit           = seed,
      details          = FALSE,
      plot.sg          = FALSE,
      parallel_args    = list(plan = "sequential", workers = 1,
                              show_message = FALSE)
    ),
    error = function(e) { fs_error <<- conditionMessage(e); NULL }
  )

  # ---- Extract results ----
  found <- !is.null(fs) && !is.null(fs$sg.harm) && length(fs$sg.harm) > 0

  out <- data.table(
    scenario  = scenario_id,
    sim       = sim_num,
    any_found = as.integer(found),
    sg_found  = if (found) paste(fs$sg.harm, collapse = " & ") else NA_character_,
    n_total   = result$n,
    c1_used   = result$c1,
    c2_used   = result$c2,
    error_msg = fs_error
  )

  if (found && !is.null(fs$df.est) &&
      "treat.recommend" %in% names(fs$df.est)) {
    sg_ind <- fs$df.est$treat.recommend
    flag   <- result$df$flag_harm
    tp <- sum(sg_ind == 0 & flag == 1)
    fp <- sum(sg_ind == 0 & flag == 0)
    tn <- sum(sg_ind == 1 & flag == 0)
    fn <- sum(sg_ind == 1 & flag == 1)
    out$sens <- if ((tp + fn) > 0) tp / (tp + fn) else NA_real_
    out$spec <- if ((tn + fp) > 0) tn / (tn + fp) else NA_real_
    out$ppv  <- if ((tp + fp) > 0) tp / (tp + fp) else NA_real_
    out$npv  <- if ((tn + fn) > 0) tn / (tn + fn) else NA_real_
    out$size_H <- sum(sg_ind == 0)
  } else {
    out$sens <- out$spec <- out$ppv <- out$npv <- NA_real_
    out$size_H <- NA_integer_
  }

  out
}
```


# Run Simulations {#sec-run}

## Single-Trial Diagnostic

```{r single-trial-diagnostic}
#| code-fold: false

cat("=== Single-trial diagnostic (Binary H1, N=800) ===\n\n")

df_diag <- sim_binary(800, sim_seed + 1, has_hte = TRUE)

cat("N:", nrow(df_diag), "| harm (bm1=0):", sum(df_diag$bm1 == 0), "\n")
cat("Progression rates:\n")
for (bm in c(0, 1)) {
  for (arm in c(0, 1)) {
    idx <- df_diag$bm1 == bm & df_diag$treat == arm
    cat(sprintf("  bm1=%d, treat=%d: %.3f (n=%d)\n",
                bm, arm, mean(df_diag$progressed[idx]), sum(idx)))
  }
}

cat(sprintf("\nUsing thresholds: c1=%.2f, c2=%.2f\n",
            thresh$binary_large$c1, thresh$binary_large$c2))

fs_diag <- tryCatch(
  forestsearch(
    df.analysis      = df_diag,
    confounders.name = c("bm1", "bm2", "bm3", "age", "ecog"),
    outcome.name     = "progressed",
    event.name       = "progressed",
    treat.name       = "treat",
    id.name          = "id",
    outcome_type     = "binary",
    effect_measure   = "OR",
    adverse_outcome  = TRUE,
    hr.threshold     = thresh$binary_large$c1,
    hr.consistency   = thresh$binary_large$c2,
    pconsistency.threshold = 0.90,
    fs.splits        = 400,
    n.min            = 60,
    maxk             = 2,
    use_lasso        = TRUE,
    use_grf          = TRUE,
    use_twostage     = TRUE,
    is.RCT           = TRUE,
    seedit           = sim_seed,
    details          = TRUE,
    plot.sg          = FALSE,
    parallel_args    = list(plan = "sequential", workers = 1,
                            show_message = FALSE)
  ),
  error = function(e) { cat("ERROR:", conditionMessage(e), "\n"); NULL }
)

if (!is.null(fs_diag)) {
  cat("  sg.harm:", paste(fs_diag$sg.harm %||% "NULL"), "\n")
  if (!is.null(fs_diag$df.est) &&
      "treat.recommend" %in% names(fs_diag$df.est)) {
    cat("  treat.recommend distribution:\n")
    print(table(fs_diag$df.est$treat.recommend))
  }
} else {
  cat("  ForestSearch returned NULL.\n")
}
```


## Full Simulation

```{r run-all}
t0 <- proc.time()[3]

scenario_ids <- c("B-H0", "B-H1-S", "B-H0-L", "B-H1-L",
                  "C-H0", "C-H1",
                  "K-H0", "K-H1")

results_list <- foreach(
  sc = rep(scenario_ids, each = nsims),
  sim = rep(seq_len(nsims), times = length(scenario_ids)),
  .errorhandling = "pass",
  .options.future = list(packages = c("forestsearch", "data.table"))
) %dofuture% {
  run_one_sim(sc, sim)
}

all_results <- rbindlist(results_list, fill = TRUE)

timings$simulations <- proc.time()[3] - t0
cat(sprintf("Total simulation time: %.1f sec (%.1f min)\n",
            timings$simulations, timings$simulations / 60))
cat(sprintf("Scenarios: %d | Sims per scenario: %d | Total: %d\n",
            length(scenario_ids), nsims, nrow(all_results)))
```

## Diagnostic: Error Messages

```{r diagnostics}
errors <- all_results[!is.na(error_msg)]
if (nrow(errors) > 0) {
  cat("=== ERRORS DETECTED ===\n")
  cat("Total errors:", nrow(errors), "of", nrow(all_results), "\n\n")
  error_summary <- errors[, .N, by = .(scenario, error_msg)]
  setorder(error_summary, scenario, -N)
  for (i in seq_len(nrow(error_summary))) {
    cat(sprintf("[%s] %d errors: %s\n",
                error_summary$scenario[i],
                error_summary$N[i],
                substr(error_summary$error_msg[i], 1, 120)))
  }
} else {
  cat("No errors detected.\n")
}

cat("\n=== Detection summary ===\n")
det_summary <- all_results[, .(
  n_sims = .N, n_found = sum(any_found),
  n_errors = sum(!is.na(error_msg))
), by = scenario]
print(det_summary)
```


# Results {#sec-results}

## Summary Table

```{r summary-table}
#| tbl-cap: "Operating characteristics by scenario with calibrated thresholds"

summary_dt <- all_results[, .(
  N        = .N,
  Power    = round(mean(any_found, na.rm = TRUE), 3),
  Sens     = round(mean(sens, na.rm = TRUE), 3),
  Spec     = round(mean(spec, na.rm = TRUE), 3),
  PPV      = round(mean(ppv, na.rm = TRUE), 3),
  NPV      = round(mean(npv, na.rm = TRUE), 3),
  Mean_nH  = round(mean(size_H, na.rm = TRUE), 0)
), by = scenario]

meta <- data.table(scenarios)[, .(ID, Outcome, Hypothesis, N, Setting, c1, c2)]
setnames(meta, "ID", "scenario")
setnames(meta, "N", "Sample_N")

summary_merged <- merge(meta, summary_dt, by = "scenario")
setorder(summary_merged, Outcome, Hypothesis)

gt(summary_merged) |>
  tab_header(
    title = "ForestSearch GLM Operating Characteristics",
    subtitle = sprintf(
      "%d sims/scenario | Calibrated thresholds (FPR target=%.0f%%)",
      nsims, 100 * fpr_target)
  ) |>
  cols_label(
    scenario = "Scenario", Outcome = "Outcome", Hypothesis = "H",
    Sample_N = "N", Setting = "Setting",
    c1 = "c1", c2 = "c2",
    N = "Sims",
    Power = "Detection", Sens = "Sens", Spec = "Spec",
    PPV = "PPV", NPV = "NPV", Mean_nH = "Mean n(H)"
  ) |>
  tab_spanner(label = "Thresholds", columns = c(c1, c2)) |>
  tab_spanner(label = "Classification Metrics",
              columns = c(Sens, Spec, PPV, NPV)) |>
  tab_source_note(
    sprintf("Runtime: %.1f min | Seed: %d",
            timings$simulations / 60, sim_seed)
  )
```


## Detection Rates

```{r detection-plot}
#| fig-height: 5
#| fig-cap: "Detection rates by scenario.  H0 scenarios show Type I error; H1 scenarios show power."

det_dt <- all_results[, .(
  Detection = mean(any_found, na.rm = TRUE)
), by = scenario]

det_dt <- merge(det_dt, meta, by = "scenario")

det_dt$label <- sprintf("%s\n(n=%s)", det_dt$scenario, det_dt$Sample_N)
det_dt$label <- factor(det_dt$label,
  levels = det_dt$label[order(det_dt$Outcome, det_dt$Hypothesis)])

ggplot(det_dt, aes(x = label, y = Detection, fill = Hypothesis)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", Detection)),
            vjust = -0.5, size = 3.5) +
  scale_fill_manual(values = c("Null" = "grey70",
                                "Alternative" = "steelblue")) +
  facet_wrap(~ Outcome, scales = "free_x", nrow = 1) +
  labs(x = NULL, y = "Detection Rate",
       title = "Subgroup Detection Rates (Calibrated Thresholds)",
       subtitle = sprintf("%d sims/scenario | FPR target=%.0f%%",
                           nsims, 100 * fpr_target)) +
  ylim(0, 1.1) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        axis.text.x = element_text(size = 8),
        plot.title = element_text(face = "bold"))
```


## Classification Metrics (H1 Only)

```{r classification-plot}
#| fig-height: 5
#| fig-cap: "Classification metrics for alternative hypothesis scenarios."

h1_dt <- all_results[grepl("H1", scenario) & any_found == 1]

if (nrow(h1_dt) > 0) {
  class_long <- melt(h1_dt,
    id.vars = "scenario",
    measure.vars = c("sens", "spec", "ppv", "npv"),
    variable.name = "Metric", value.name = "Value"
  )
  class_long$Metric <- toupper(as.character(class_long$Metric))

  ggplot(class_long, aes(x = Metric, y = Value, fill = Metric)) +
    geom_boxplot(outlier.size = 1, alpha = 0.7) +
    facet_wrap(~ scenario, nrow = 1) +
    scale_fill_brewer(palette = "Set2") +
    labs(x = NULL, y = "Value",
         title = "Classification Metrics (H1 Scenarios)",
         subtitle = "Distributions among simulations where subgroup found") +
    ylim(0, 1) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none",
          plot.title = element_text(face = "bold"))
} else {
  cat("No subgroups found in any H1 simulation.\n")
}
```


## Subgroup Size Distribution

```{r size-plot}
#| fig-height: 4
#| fig-cap: "Distribution of identified harm subgroup size"

h1_found <- all_results[grepl("H1", scenario) & any_found == 1]

if (nrow(h1_found) > 0) {
  ggplot(h1_found, aes(x = scenario, y = size_H, fill = scenario)) +
    geom_boxplot(alpha = 0.7) +
    labs(x = "Scenario", y = "n(H) identified",
         title = "Identified Harm Subgroup Size",
         subtitle = "Dashed line: true harm prevalence (30-35% of N)") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none",
          plot.title = element_text(face = "bold"))
} else {
  cat("No subgroups found in any H1 simulation.\n")
}
```


## Most Frequently Identified Subgroups

```{r freq-subgroups}
#| tbl-cap: "Top subgroups identified across simulations"

h1_found <- all_results[grepl("H1", scenario) & any_found == 1]

if (nrow(h1_found) > 0) {
  freq_dt <- h1_found[, .N, by = .(scenario, sg_found)]
  freq_dt <- freq_dt[order(scenario, -N)]
  freq_top <- freq_dt[, head(.SD, 3), by = scenario]
  freq_top$Pct <- round(100 * freq_top$N /
    freq_dt[, sum(N), by = scenario][freq_top, on = "scenario"]$V1, 1)

  gt(freq_top) |>
    tab_header(
      title = "Most Frequently Identified Subgroups",
      subtitle = "Top 3 per scenario"
    ) |>
    cols_label(scenario = "Scenario",
               sg_found = "Subgroup Definition",
               N = "Count", Pct = "Percent")
} else {
  cat("No subgroups found in any H1 simulation.\n")
}
```


# Timing Summary {#sec-timing}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

timing_df <- data.frame(
  Stage = c("All simulations", "Total (including setup)"),
  Seconds = c(timings$simulations, timings$total),
  Minutes = c(timings$simulations / 60, timings$total / 60),
  stringsAsFactors = FALSE
)

gt(timing_df) |>
  tab_header(
    title = "Computational Timing",
    subtitle = sprintf("%d total simulations | %d workers",
                        nsims * length(scenario_ids), n_workers)
  ) |>
  fmt_number(columns = c(Seconds, Minutes), decimals = 1) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)",
             Minutes = "Time (min)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total (including setup)")
  )
```


# Summary {#sec-summary}

This simulation study evaluated `forestsearch()` across three GLM
outcome types with **calibrated thresholds** selected by
`suggest_thresholds()`.

**Key design features:**

- **Binary** (OR scale): Thresholds from corrected approximation
  with $L_{\text{eff}}$ calibration ($C = `r L_eff_C_binary`$,
  $\alpha = `r L_eff_alpha_binary`$).  Phase II (N = 200) and
  Phase III (N = 800) use different thresholds because $d_{\text{eff}}$
  and $L_{\text{eff}}$ depend on both $n_H$ and $N$.

- **Continuous** (MD scale): Thresholds on the mean-difference
  scale with $L_{\text{eff}} = 1$ (no calibration available).
  High $d_{\text{eff}}$ (low variance, large $n_H$) allows
  small thresholds.

- **Count** (IRR scale): Thresholds on the ratio scale using the
  binary $L_{\text{eff}}$ calibration as an approximation (same number
  of confounders, so the candidate generation process is similar).
  $d_{\text{eff}} = D$ (total events) depends on the mean event rate.
  A count-specific $L_{\text{eff}}$ calibration would be more rigorous.


# References {#sec-references}

- Leon, L.F., et al. (2024). Exploratory Subgroup Identification.
  *Statistics in Medicine*, 43, 4253--4275. doi:10.1002/sim.10163

- Rigdon, J., Basu, S., Engel, R., Baiocchi, M. (2018). Preventing
  false discovery of heterogeneous treatment effect subgroups in
  randomized trials. *Trials*, 19, 382.

- DINA: Estimating Heterogeneous Treatment Effects for General
  Responses. *Biometrics*, 81(4), 2025.


# Session Information {#sec-session}

```{r session-info}
plan(sequential)
sessionInfo()
```
