---
title: "GRF dmin.grf Calibration for ACTG175 Binary Benefit"
subtitle: "Calibrating to the 12-Confounder ACTG175 DGM"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 6
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ── All tunable parameters in this single location ──────────────────

# Simulation budget
nsims         <- 100L           # Sims per dmin × hypothesis
sim_n         <- 1000L          # Sample size per sim (matches ACTG175 doc)
sim_seed      <- 8316951L
n_workers     <- max(1L, floor(0.90 * parallel::detectCores(logical = FALSE)))

# ACTG175 data setup (replicate from binary benefit document)
actg_arms        <- c(1L, 3L)
actg_treat_arm   <- 1L
actg_age_cut     <- 34
actg_preanti_cut <- 744.5

# DGM
dgm_factor_vars   <- paste0("z", 1:12)
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)
dgm_n_super       <- 5000L

# Calibration target
cal_target_or     <- 2.0         # OR(Q) on switched scale
cal_k_grid_range  <- c(0, 3)
cal_grid_step     <- 0.05

# dmin.grf grid — focused on the actionable range
dmin_grid <- c(0, 0.02, 0.05, 0.08, 0.10, 0.12, 0.15, 0.20, 0.25, 0.30)

# GRF parameters (fixed across sweep)
grf_maxdepth     <- 2L
grf_sg_criterion <- "mDiff"
grf_n_min        <- 40L
grf_adverse_outcome <- TRUE    # Y-flip for binary benefit search
grf_tune         <- FALSE      # cross-validated GRF hyperparameter tuning

# Targets
target_fpr <- 15   # Maximum acceptable FPR (%)
target_det <- 50   # Minimum acceptable detection (%)
```


# Purpose {#sec-purpose}

The `grf_dmin_calibration_binary.qmd` document calibrated
`dmin.grf` on **synthetic 4-confounder data** and found
`dmin.grf = 0.02` achieves strong FPR control.  However, the
ACTG175 binary benefit scenario uses **12 binary confounders**,
which creates substantially more candidate splits for the
policy tree and weakens FPR control.

This document calibrates `dmin.grf` directly on the ACTG175
DGM to find the value that balances detection power and FPR
control for this specific application.

Current results at `dmin.grf = 0.10` (2,000 sims):

| Metric | FS | GRF |
|--------|------|------|
| H1 Detection | 74% | 72% |
| H1 Sensitivity | 0.57 | 0.90 |
| H1 PPV | 0.80 | 0.75 |
| H0 FPR | 4% | 28% |
| θ̂(Q̂) bias | — | 1.8% |


# Data Preparation {#sec-data}

```{r packages}
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(doFuture)
library(foreach)

plan(multisession, workers = n_workers)
cat(sprintf("Workers: %d\n", n_workers))

reset_workers <- function(msg = "") {
  if (nzchar(msg)) cat(sprintf("[Resetting workers: %s]\n", msg))
  plan(sequential); gc(verbose = FALSE); Sys.sleep(1)
  plan(multisession, workers = n_workers)
}

t_start <- proc.time()
```

```{r prepare-actg175}
# Replicate ACTG175 data preparation from binary benefit document
actg_df <- subset(speff2trial::ACTG175, arms %in% actg_arms)
actg_df$id <- seq_len(nrow(actg_df))
actg_df$treat_orig <- ifelse(actg_df$arms == actg_treat_arm, 1L, 0L)
actg_df$treat <- 1L - actg_df$treat_orig   # switched for benefit search
actg_df$y_binary <- as.integer(actg_df$cd420 > actg_df$cd40)
actg_df <- actg_df[!is.na(actg_df$cd420), ]

# Discretise covariates
actg_df$z1  <- as.factor(ifelse(actg_df$age > actg_age_cut, 1L, 0L))
actg_df$z2  <- as.factor(ifelse(actg_df$preanti <= actg_preanti_cut, 1L, 0L))
actg_df$z3  <- as.factor(ifelse(actg_df$wtkg <= 75, 1L, 0L))
actg_df$z4  <- as.factor(ifelse(actg_df$karnof <= median(actg_df$karnof), 1L, 0L))
actg_df$z5  <- as.factor(ifelse(actg_df$cd40 <= median(actg_df$cd40), 1L, 0L))
actg_df$z6  <- as.factor(ifelse(actg_df$cd80 <= median(actg_df$cd80), 1L, 0L))
actg_df$z7  <- as.factor(actg_df$hemo)
actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs)
actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender)
actg_df$z12 <- as.factor(actg_df$symptom)

actg_df$flag_harm_ref <- as.integer(actg_df$z1 == 1 & actg_df$z2 == 1)
cat(sprintf("N=%d, confounders=%d, Q prevalence=%.1f%%\n",
    nrow(actg_df), length(dgm_factor_vars),
    100 * mean(actg_df$flag_harm_ref)))
```


# DGM Construction {#sec-dgm}

```{r build-dgms}
# Calibrate H1 DGM
dgm_calibrated <- calibrate_glm_interaction(
  data = actg_df, factor_vars = dgm_factor_vars,
  outcome_var = "y_binary", treatment_var = "treat",
  target_effect = cal_target_or,
  outcome_type = "binary", effect_measure = "OR",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  k_inter_range = cal_k_grid_range, grid_step = cal_grid_step,
  n_super = dgm_n_super, seed = sim_seed, verbose = TRUE
)

cat(sprintf("\nH1 DGM: OR(Q)=%.3f (target %.1f), k_inter=%.2f\n",
    dgm_calibrated$hazard_ratios$harm_subgroup,
    cal_target_or, dgm_calibrated$model_params$k_inter))

# H0 DGM (homogeneous treatment effect)
dgm_null <- generate_glm_dgm(
  data = actg_df, factor_vars = dgm_factor_vars,
  outcome_var = "y_binary", treatment_var = "treat",
  outcome_type = "binary", effect_measure = "OR",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  model = "null", k_treat = 1,
  n_super = dgm_n_super, seed = sim_seed
)

cat(sprintf("H0 DGM: OR(ITT)=%.3f\n",
    dgm_null$hazard_ratios$overall))
```


# dmin.grf Sweep {#sec-sweep}

For each value in the grid, run `r nsims` simulations under both
H1 and H0. Each simulation calls `grf.subg.harm.glm()` directly
with 12 confounders at N = `r sim_n`.

```{r sweep-fn}
run_dmin_eval <- function(dgm, nsims, dmin_val) {
  results_raw <- foreach(
    i = seq_len(nsims),
    .errorhandling = "pass",
    .options.future = list(
      packages = c("forestsearch", "survival", "data.table"),
      seed = TRUE
    )
  ) %dofuture% {
    sim_df <- simulate_from_glm_dgm(dgm, n = sim_n,
                                     seed = sim_seed + i * 1000L)
    res <- tryCatch(
      grf.subg.harm.glm(
        data = sim_df, confounders.name = dgm_factor_vars,
        outcome.name = "y_sim", treat.name = "treat_sim",
        id.name = "id", outcome_type = "binary",
        adverse_outcome = grf_adverse_outcome,
        n.min = grf_n_min, dmin.grf = dmin_val, maxdepth = grf_maxdepth,
        RCT = TRUE, sg.criterion = grf_sg_criterion,
        tune_grf = grf_tune,
        seedit = 42L, verbose = FALSE
      ),
      error = function(e) NULL
    )

    detected <- !is.null(res) && !is.null(res$sg.harm.id) &&
      length(res$sg.harm.id) > 0 && !all(is.na(res$sg.harm.id))

    # Classification metrics (only if detected)
    sens <- ppv <- n_H <- NA_real_
    if (detected && !is.null(res$data) &&
        "treat.recommend" %in% names(res$data)) {
      in_H <- res$data$treat.recommend == 0
      true_Q <- sim_df$flag_harm
      n_H <- sum(in_H)
      tp <- sum(in_H & true_Q == 1)
      fp <- sum(in_H & true_Q == 0)
      fn <- sum(!in_H & true_Q == 1)
      sens <- tp / max(tp + fn, 1)
      ppv  <- tp / max(tp + fp, 1)
    }

    list(detected = detected, sens = sens, ppv = ppv, n_H = n_H)
  }

  is_err <- vapply(results_raw, inherits, logical(1), "error")
  good <- results_raw[!is_err]

  det_vec  <- vapply(good, function(x) x$detected, logical(1))
  sens_vec <- vapply(good, function(x) x$sens, numeric(1))
  ppv_vec  <- vapply(good, function(x) x$ppv, numeric(1))
  nH_vec   <- vapply(good, function(x) x$n_H, numeric(1))

  list(
    det_pct    = 100 * mean(det_vec),
    mean_sens  = mean(sens_vec[det_vec & !is.na(sens_vec)]),
    mean_ppv   = mean(ppv_vec[det_vec & !is.na(ppv_vec)]),
    mean_nH    = mean(nH_vec[det_vec & !is.na(nH_vec)]),
    n_errors   = sum(is_err)
  )
}
```

```{r run-sweep}
cat(sprintf("dmin.grf grid: %s\n", paste(dmin_grid, collapse = ", ")))
cat(sprintf("Sims per point: %d alt + %d null, N=%d\n", nsims, nsims, sim_n))
cat(sprintf("Confounders: %d\n\n", length(dgm_factor_vars)))

sweep_results <- data.frame(
  dmin = numeric(0), Det_H1 = numeric(0), FPR_H0 = numeric(0),
  Sens_H1 = numeric(0), PPV_H1 = numeric(0),
  Avg_nH_H1 = numeric(0),
  Sens_H0 = numeric(0), PPV_H0 = numeric(0),
  stringsAsFactors = FALSE
)

for (dmin_val in dmin_grid) {
  cat(sprintf("--- dmin = %.2f ---\n", dmin_val))

  if (dmin_val == dmin_grid[1] || dmin_val == 0.15) {
    reset_workers(sprintf("dmin=%.2f", dmin_val))
  }

  # H1
  t0 <- proc.time()[3]
  h1 <- run_dmin_eval(dgm_calibrated, nsims, dmin_val)
  t_h1 <- proc.time()[3] - t0

  # H0
  t0 <- proc.time()[3]
  h0 <- run_dmin_eval(dgm_null, nsims, dmin_val)
  t_h0 <- proc.time()[3] - t0

  sweep_results[nrow(sweep_results) + 1, ] <- data.frame(
    dmin = dmin_val,
    Det_H1 = round(h1$det_pct, 1),
    FPR_H0 = round(h0$det_pct, 1),
    Sens_H1 = round(h1$mean_sens, 3),
    PPV_H1 = round(h1$mean_ppv, 3),
    Avg_nH_H1 = round(h1$mean_nH, 0),
    Sens_H0 = round(h0$mean_sens, 3),
    PPV_H0 = round(h0$mean_ppv, 3),
    stringsAsFactors = FALSE
  )

  flag <- if (h1$det_pct >= target_det && h0$det_pct <= target_fpr) " ***" else ""
  cat(sprintf("  H1: Det=%.0f%%, Sens=%.2f, PPV=%.2f, avg|H|=%.0f (%.0fs)\n",
      h1$det_pct, h1$mean_sens, h1$mean_ppv, h1$mean_nH, t_h1))
  cat(sprintf("  H0: FPR=%.0f%%, Sens=%.2f, PPV=%.2f (%.0fs)%s\n\n",
      h0$det_pct, h0$mean_sens, h0$mean_ppv, t_h0, flag))
}

plan(sequential)
```


# Results {#sec-results}

## Full Results Table {#sec-table}

```{r results-table}
sweep_results$Pass <- sweep_results$Det_H1 >= target_det &
                      sweep_results$FPR_H0 <= target_fpr

sweep_results |>
  gt::gt() |>
  gt::tab_header(
    title = "ACTG175 Binary Benefit: dmin.grf Calibration",
    subtitle = sprintf(
      "%d alt + %d null sims, N=%d, 12 confounders | Target: Det >= %d%%, FPR <= %d%%",
      nsims, nsims, sim_n, target_det, target_fpr)
  ) |>
  gt::cols_label(
    dmin = "dmin.grf",
    Det_H1 = "Det (%)",
    FPR_H0 = "FPR (%)",
    Sens_H1 = "Sens (H1)",
    PPV_H1 = "PPV (H1)",
    Avg_nH_H1 = "Avg |Ĥ| (H1)",
    Sens_H0 = "Sens (H0)",
    PPV_H0 = "PPV (H0)"
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#c8e6c9"),
    locations = gt::cells_body(rows = Pass == TRUE)
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#fff9c4"),
    locations = gt::cells_body(
      rows = Det_H1 >= target_det & FPR_H0 <= 25 & !Pass)
  ) |>
  gt::cols_hide(columns = "Pass") |>
  gt::tab_footnote(
    footnote = "Green = meets both targets. Yellow = near-miss (FPR <= 25%)."
  )
```

## Detection vs FPR Tradeoff {#sec-tradeoff}

```{r tradeoff-plot, fig.height=5}
#| fig-cap: "Detection and FPR as a function of dmin.grf for the ACTG175 12-confounder DGM."

if (nrow(sweep_results) > 0) {
  plot_df <- rbind(
    transform(sweep_results, Metric = "Detection (H1)", Rate = Det_H1),
    transform(sweep_results, Metric = "FPR (H0)", Rate = FPR_H0)
  )

  ggplot(plot_df, aes(x = dmin, y = Rate, colour = Metric, linetype = Metric)) +
    geom_line(linewidth = 1) +
    geom_point(size = 3) +
    geom_text(aes(label = sprintf("%.0f%%", Rate)), vjust = -1.2, size = 2.8,
              show.legend = FALSE) +
    geom_hline(yintercept = target_fpr, linetype = "dotted",
               colour = "red", linewidth = 0.5) +
    geom_hline(yintercept = target_det, linetype = "dotted",
               colour = "blue", linewidth = 0.5) +
    annotate("rect", xmin = -Inf, xmax = Inf,
             ymin = 0, ymax = target_fpr,
             alpha = 0.05, fill = "green") +
    scale_colour_manual(values = c(
      "Detection (H1)" = "#2c3e50", "FPR (H0)" = "#e74c3c")) +
    scale_linetype_manual(values = c(
      "Detection (H1)" = "solid", "FPR (H0)" = "dashed")) +
    labs(
      x = "dmin.grf",
      y = "Rate (%)",
      title = "ACTG175 Binary: Detection and FPR vs dmin.grf",
      subtitle = sprintf(
        "%d sims/point, N=%d, 12 confounders | OR(Q)=%.2f (switched)",
        nsims, sim_n, dgm_calibrated$hazard_ratios$harm_subgroup)
    ) +
    coord_cartesian(ylim = c(0, 105)) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom")
}
```

## Sensitivity and PPV vs dmin.grf {#sec-classification}

```{r classification-plot, fig.height=5}
#| fig-cap: "Classification quality (among detections) as a function of dmin.grf."

if (nrow(sweep_results) > 0) {
  cls_df <- rbind(
    data.frame(dmin = sweep_results$dmin,
               Metric = "Sensitivity", Value = sweep_results$Sens_H1),
    data.frame(dmin = sweep_results$dmin,
               Metric = "PPV", Value = sweep_results$PPV_H1)
  )

  ggplot(cls_df, aes(x = dmin, y = Value, colour = Metric)) +
    geom_line(linewidth = 1) +
    geom_point(size = 3) +
    labs(
      x = "dmin.grf",
      y = "Classification Metric",
      title = "ACTG175 Binary: Classification Quality vs dmin.grf (H1)",
      subtitle = "Among simulations where GRF detected a subgroup"
    ) +
    scale_colour_manual(values = c(
      Sensitivity = "#27ae60", PPV = "#2980b9")) +
    coord_cartesian(ylim = c(0, 1.05)) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom")
}
```

## Subgroup Size vs dmin.grf {#sec-size}

```{r size-plot, fig.height=4}
#| fig-cap: "Average identified subgroup size under H1 as a function of dmin.grf."

if (nrow(sweep_results) > 0) {
  true_Q_size <- round(sim_n * mean(actg_df$flag_harm_ref))

  ggplot(sweep_results, aes(x = dmin, y = Avg_nH_H1)) +
    geom_line(linewidth = 1, colour = "#2c3e50") +
    geom_point(size = 3, colour = "#2c3e50") +
    geom_hline(yintercept = true_Q_size, linetype = "dashed",
               colour = "firebrick") +
    annotate("text", x = max(dmin_grid) * 0.8, y = true_Q_size + 20,
             label = sprintf("True |Q| ≈ %d", true_Q_size),
             colour = "firebrick", size = 3.5) +
    labs(
      x = "dmin.grf",
      y = "Average |Ĥ|",
      title = "Identified Subgroup Size vs dmin.grf",
      subtitle = "Closer to true |Q| = better calibration"
    ) +
    coord_cartesian(ylim = c(0, sim_n)) +
    theme_minimal(base_size = 12)
}
```


## Recommendation {#sec-recommendation}

```{r recommendation}
#| results: asis

# Find the best dmin that meets targets
good <- sweep_results[sweep_results$Det_H1 >= target_det &
                        sweep_results$FPR_H0 <= target_fpr, ]

if (nrow(good) > 0) {
  best <- good[which.min(good$dmin), ]
  cat(sprintf(
    "**Recommended `dmin.grf = %.2f`** for the ACTG175 binary benefit scenario:\n\n",
    best$dmin))
  cat(sprintf("- Detection: %.0f%%\n", best$Det_H1))
  cat(sprintf("- FPR: %.0f%%\n", best$FPR_H0))
  cat(sprintf("- Sensitivity: %.2f\n", best$Sens_H1))
  cat(sprintf("- PPV: %.2f\n", best$PPV_H1))
  cat(sprintf("- Average |Ĥ|: %.0f (true |Q| ≈ %d)\n",
      best$Avg_nH_H1, round(sim_n * mean(actg_df$flag_harm_ref))))
} else {
  cat("**No `dmin.grf` value achieves both targets.**\n\n")

  # Find best tradeoff
  sweep_results$score <- sweep_results$Det_H1 - 2 * sweep_results$FPR_H0
  best_trade <- sweep_results[which.max(sweep_results$score), ]
  cat(sprintf(
    "Best tradeoff: `dmin.grf = %.2f` (Det=%.0f%%, FPR=%.0f%%, Sens=%.2f, PPV=%.2f)\n\n",
    best_trade$dmin, best_trade$Det_H1, best_trade$FPR_H0,
    best_trade$Sens_H1, best_trade$PPV_H1))

  # Also show relaxed target
  relaxed <- sweep_results[sweep_results$Det_H1 >= target_det &
                             sweep_results$FPR_H0 <= 25, ]
  if (nrow(relaxed) > 0) {
    best_relaxed <- relaxed[which.min(relaxed$dmin), ]
    cat(sprintf(
      "With relaxed FPR target (≤ 25%%): `dmin.grf = %.2f` (Det=%.0f%%, FPR=%.0f%%)\n",
      best_relaxed$dmin, best_relaxed$Det_H1, best_relaxed$FPR_H0))
  }
}
```


# Comparison with 4-Confounder Calibration {#sec-comparison}

The synthetic 4-confounder calibration
(`grf_dmin_calibration_binary.qmd`) found `dmin.grf = 0.02`
sufficient for FPR $\leq 9\%$.  The difference here reflects:

- **12 vs 4 confounders**: more candidate splits = higher
  multiplicity for the policy tree
- **Correlated confounders**: ACTG175 covariates have real
  correlations (e.g., CD4/CD8, age/Karnofsky) that create
  redundant splits
- **Imbalanced prevalence**: some z-variables have very
  asymmetric marginal distributions


# Timing {#sec-timing}

```{r timing}
elapsed <- (proc.time() - t_start)["elapsed"]
cat(sprintf("Total render time: %.1f minutes (%.0f seconds)\n",
    elapsed / 60, elapsed))
cat(sprintf("Configuration: %d sims × %d dmin values × 2 hypotheses\n",
    nsims, length(dmin_grid)))
cat(sprintf("Total GRF calls: ~%d\n",
    2 * nsims * length(dmin_grid)))
```


# Session Info {#sec-session}

```{r session-info}
sessionInfo()
```
