---
title: "Correcting the Detection Probability Under the Null"
subtitle: "Accounting for the Number of Viable Candidates"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")
```


# The Problem {#sec-problem}

The Section 2.1 approximation gives the per-subgroup detection
probability $P_1$.  Simulation shows:

- **Power (H1)**: $P_1$ agrees well with simulation for
  $N \ge 500$ (Sim/Approx $\approx$ 1.0).
- **Type-I error (H0)**: $P_1$ increasingly underestimates
  the simulated FPR as $N$ grows (Sim/Approx reaches 5--15x
  at $N = 1000$).

This is anti-intuitive only if one expects the approximation
to model the *procedure-level* FPR directly.  It does not.
$P_1$ models a *single pre-specified subgroup*.  The procedure
evaluates many candidates, and the number of viable candidates
**grows with $N$**.


# Why Power Agreement Improves but FPR Agreement Worsens {#sec-why}

## Power (H1)

Under $H_1$, there exists a true subgroup $H$ with
$\beta = \log(2) = 0.69$.  This signal is much larger than
any noise candidate's $\hat\beta \approx 0$.  The procedure
reliably identifies $H$ (when $N$ is large enough for $H$
to enter the candidate set), and the approximation for
$P_1(\beta = 0.69)$ is accurate because it models exactly
the right quantity: the detection probability for a specific
subgroup with a known effect.

For $N = 200$, the approximation *overestimates* power
because the true harm subgroup ($n_H \approx 60$) is right
at `n.min = 60` and may be excluded from the candidate set
in many simulations.  The approximation does not model the
candidate-nomination step.

## Type-I Error (H0)

Under $H_0$, $\beta = 0$ for every subgroup.  The
approximation gives $P_1(0) \approx 0.15$--$0.23$ depending
on $d_{\text{eff}}$ and thresholds.  But the procedure finds
*any* subgroup among $L$ candidates, so:

$$
P(\text{any passes}) \approx 1 - (1 - P_1)^{L_{\text{eff}}}
$$

The effective number of candidates $L_{\text{eff}}$ grows
with $N$ because:

1. **More subgroup combinations survive `n.min`**.  At
   $N = 200$ with `n.min = 60`, a subgroup must be $\ge 30\%$
   of the sample.  Two-factor combinations (e.g.,
   bm1 = 0 AND bm2 = 0, prevalence $\approx 15\%$) have only
   30 patients and are excluded.  At $N = 1000$, the same
   combination has 150 patients, well above `n.min`.

2. **GRF generates more cuts on continuous variables**.  With
   more data, GRF produces additional split points on age,
   each creating new candidate factors.

3. **LASSO retains more variables**.  Increased power detects
   spurious prognostic associations.


# Empirical $L_{\text{eff}}$ {#sec-leff}

From the simulation results (Table 5), we can solve for the
implied $L_{\text{eff}}$ at each $N$:

$$
L_{\text{eff}} =
\frac{\log(1 - \text{FPR}_{\text{sim}})}{\log(1 - P_1)}
$$

```{r load-packages}
library(forestsearch)
library(ggplot2)
library(gt)
```

```{r compute-leff}
#| code-fold: false

# Simulation results from Table 5
# Using c1=1.25, c2=1.25 (moderate thresholds) for stability
sim_data <- data.frame(
  N = c(200, 500, 700, 1000),
  n_H = round(c(200, 500, 700, 1000) * 0.30),
  P1 = c(0.152, 0.109, 0.091, 0.071),
  Sim_FPR = c(0.185, 0.220, 0.270, 0.280)
)

sim_data$L_eff <- log(1 - sim_data$Sim_FPR) / log(1 - sim_data$P1)
sim_data$N_over_nmin <- sim_data$N / 60

cat("=== Implied L_eff (c1=1.25, c2=1.25) ===\n")
print(sim_data[, c("N", "P1", "Sim_FPR", "L_eff", "N_over_nmin")])
```


## $L_{\text{eff}}$ Grows with $N / n_{\min}$

```{r leff-model}
#| fig-height: 5
#| fig-cap: "L_eff grows approximately as a power of N/n.min.  The fit suggests L_eff ~ (N/n.min)^0.8."

# Fit log-linear model: log(L_eff) = a + b * log(N/n.min)
fit <- lm(log(L_eff) ~ log(N_over_nmin), data = sim_data)
alpha <- coef(fit)[2]
C <- exp(coef(fit)[1])

cat(sprintf("Fit: L_eff = %.2f * (N/n.min)^%.2f\n", C, alpha))
cat(sprintf("R-squared: %.3f\n", summary(fit)$r.squared))

# Plot
N_seq <- seq(100, 1200, by = 10)
pred_df <- data.frame(
  N = N_seq,
  N_over_nmin = N_seq / 60,
  L_eff_pred = C * (N_seq / 60)^alpha
)

ggplot() +
  geom_point(data = sim_data, aes(x = N, y = L_eff),
             size = 4, colour = "firebrick") +
  geom_line(data = pred_df, aes(x = N, y = L_eff_pred),
            colour = "steelblue", linewidth = 1) +
  labs(x = "Total Sample Size N",
       y = expression(L[eff]),
       title = expression(L[eff] ~ "from Simulation (c1=1.25, c2=1.25)"),
       subtitle = sprintf("Fit: L_eff = %.2f x (N/n.min)^%.2f",
                           C, alpha)) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```


## Validation Across Threshold Pairs

```{r leff-by-threshold}
#| tbl-cap: "Implied L_eff across threshold configurations.  L_eff should be similar across (c1, c2) for the same N if it reflects the candidate count."

# Multiple (c1,c2) pairs from Table 5
configs <- data.frame(
  c1 = c(1.25, 1.25, 1.50, 1.50, 1.50,
         1.25, 1.25, 1.50, 1.50, 1.50,
         1.25, 1.25, 1.50, 1.50, 1.50,
         1.25, 1.25, 1.50, 1.50, 1.50),
  c2 = c(1.00, 1.25, 1.25, 1.35, 1.50,
         1.00, 1.25, 1.25, 1.35, 1.50,
         1.00, 1.25, 1.25, 1.35, 1.50,
         1.00, 1.25, 1.25, 1.35, 1.50),
  N  = c(rep(200, 5), rep(500, 5), rep(700, 5), rep(1000, 5)),
  P1 = c(0.227, 0.152, 0.139, 0.121, 0.094,
         0.195, 0.109, 0.083, 0.068, 0.045,
         0.177, 0.091, 0.061, 0.049, 0.029,
         0.153, 0.071, 0.039, 0.031, 0.016),
  Sim_FPR = c(0.325, 0.185, 0.185, 0.155, 0.095,
              0.505, 0.220, 0.220, 0.160, 0.090,
              0.590, 0.270, 0.270, 0.170, 0.115,
              0.765, 0.280, 0.280, 0.190, 0.080)
)

configs$L_eff <- log(1 - configs$Sim_FPR) / log(1 - configs$P1)
configs$label <- sprintf("c1=%.2f, c2=%.2f", configs$c1, configs$c2)

# Show L_eff by N
leff_summary <- aggregate(L_eff ~ N, data = configs,
                           FUN = function(x) {
                             sprintf("%.1f (%.1f--%.1f)",
                               median(x), min(x), max(x))
                           })
names(leff_summary)[2] <- "L_eff: median (range)"

gt(leff_summary) |>
  tab_header(
    title = "Implied L_eff by Sample Size (Across Threshold Pairs)",
    subtitle = "L_eff should be roughly constant within N"
  )
```


```{r leff-scatter}
#| fig-height: 5
#| fig-cap: "L_eff by N, coloured by threshold pair.  If L_eff reflects the candidate count rather than the thresholds, points at the same N should cluster vertically."

ggplot(configs, aes(x = N, y = L_eff, colour = label)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(x = "N", y = expression(L[eff]),
       title = "Implied L_eff Across Configurations",
       colour = "Thresholds") +
  scale_colour_brewer(palette = "Set2") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


# The Corrected Approximation {#sec-corrected}

## Formula

$$
\boxed{
P_{\text{null}}^{\text{corrected}}(N) =
1 - \bigl(1 - P_1(0)\bigr)^{L_{\text{eff}}(N)}
}
$$

where $P_1(0)$ is the per-subgroup detection probability from
the GLM approximation (`compute_detection_probability_glm()`
with $\theta = 1$), and:

$$
L_{\text{eff}}(N) = C \cdot
\left(\frac{N}{n_{\min}}\right)^{\alpha}
$$

with $C$ and $\alpha$ estimated from a reference simulation.

**Important**: $L_{\text{eff}}$ is a property of the *candidate
generation process* (GRF + LASSO + confounder set + maxk +
n.min), not of the thresholds.  It should be approximately
constant across $(c_1, c_2)$ for the same $N$ and confounder
configuration.

## Implementation

```{r corrected-function}
#| code-fold: false

#' Corrected Null Detection Probability
#'
#' Adjusts the per-subgroup approximation for the number of
#' viable candidates evaluated by ForestSearch.
#'
#' @param theta Numeric. True effect (1.0 for null).
#' @param d_eff Numeric. Effective information.
#' @param c1 Numeric. Screening threshold.
#' @param c2 Numeric. Consistency threshold.
#' @param N Integer. Total sample size.
#' @param n_min Integer. Minimum subgroup size. Default: 60.
#' @param L_eff_C Numeric. Calibration constant. Default: 0.8.
#' @param L_eff_alpha Numeric. Power-law exponent. Default: 0.8.
#'
#' @return Numeric. Corrected detection probability.
p_detect_null_corrected <- function(theta, d_eff, c1, c2,
                                     N, n_min = 60,
                                     L_eff_C = NULL,
                                     L_eff_alpha = NULL) {
  # Per-subgroup probability
  P1 <- compute_detection_probability_glm(
    theta = theta, d_eff = d_eff, c1 = c1, c2 = c2,
    effect_scale = "ratio", method = "monte_carlo"
  )

  # Use fitted values if not provided
  if (is.null(L_eff_C)) L_eff_C <- C
  if (is.null(L_eff_alpha)) L_eff_alpha <- alpha

  # Effective candidates
  L_eff <- L_eff_C * (N / n_min)^L_eff_alpha

  # Union bound
  1 - (1 - P1)^L_eff
}
```


## Validation Against Simulation

```{r validate-correction}
#| tbl-cap: "Corrected approximation vs simulation.  The corrected Sim/Approx ratio should be closer to 1.0 than the uncorrected."

configs$P1_corrected <- mapply(function(N_val, d_eff, c1_val, c2_val) {
  d <- d_eff_binary(round(0.30 * N_val), p_event = 0.30)
  p_detect_null_corrected(1.0, d, c1_val, c2_val, N_val)
}, configs$N, configs$P1, configs$c1, configs$c2)

configs$Ratio_uncorrected <- round(configs$Sim_FPR / configs$P1, 2)
configs$Ratio_corrected <- round(configs$Sim_FPR /
                                   pmax(configs$P1_corrected, 0.001), 2)

gt(configs[, c("N", "c1", "c2", "P1", "P1_corrected",
               "Sim_FPR", "Ratio_uncorrected", "Ratio_corrected")]) |>
  tab_header(
    title = "Corrected vs Uncorrected Approximation (H0)",
    subtitle = sprintf("L_eff = %.2f x (N/n.min)^%.2f", C, alpha)
  ) |>
  fmt_number(columns = c(P1, P1_corrected, Sim_FPR), decimals = 3) |>
  cols_label(
    P1 = "P1 (uncorr)", P1_corrected = "P1 (corrected)",
    Sim_FPR = "Sim FPR",
    Ratio_uncorrected = "Sim/Uncorr",
    Ratio_corrected = "Sim/Corr"
  ) |>
  tab_row_group(label = "N = 1000", rows = N == 1000) |>
  tab_row_group(label = "N = 700", rows = N == 700) |>
  tab_row_group(label = "N = 500", rows = N == 500) |>
  tab_row_group(label = "N = 200", rows = N == 200) |>
  tab_source_note(
    "Sim/Corr near 1.0 = correction works. Compare to Sim/Uncorr."
  )
```


```{r validate-scatter}
#| fig-height: 5
#| fig-cap: "Corrected (triangles) vs uncorrected (circles) approximation against simulation.  Corrected points should cluster near the diagonal."

plot_df <- data.frame(
  Approx = c(configs$P1, configs$P1_corrected),
  Sim = rep(configs$Sim_FPR, 2),
  Method = rep(c("Uncorrected P1", "Corrected"), each = nrow(configs)),
  N = rep(configs$N, 2)
)

ggplot(plot_df, aes(x = Approx, y = Sim)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              colour = "grey50") +
  geom_point(aes(colour = factor(N), shape = Method), size = 3,
             alpha = 0.7) +
  scale_shape_manual(values = c("Uncorrected P1" = 16,
                                 "Corrected" = 17)) +
  coord_equal(xlim = c(0, 0.8), ylim = c(0, 0.8)) +
  labs(x = "Approximated FPR", y = "Simulated FPR",
       title = "Null FPR: Corrected vs Uncorrected Approximation",
       colour = "N", shape = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


# Discussion {#sec-discussion}

## What $L_{\text{eff}}$ Captures

$L_{\text{eff}}$ is not the total number of candidate subgroups.
It is the effective number of *approximately independent*
candidates that survive the sample-size and events filters.
It depends on:

- **$N / n_{\min}$**: the ratio of total sample size to the
  minimum subgroup size.  This determines what fraction of
  candidate combinations are viable.
- **The confounder set**: more confounders $\to$ more
  candidates $\to$ larger $L_{\text{eff}}$.
- **`maxk`**: higher maxk allows more combinations.
- **GRF and LASSO**: data-adaptive candidate selection adds
  candidates that depend on the specific dataset.

$L_{\text{eff}}$ does *not* depend on the thresholds $(c_1, c_2)$
--- the thresholds determine $P_1$, while $L_{\text{eff}}$
determines how many times $P_1$ is "rolled."

## When to Calibrate $L_{\text{eff}}$

The power-law fit $L_{\text{eff}} = C (N/n_{\min})^\alpha$ is
specific to the binary DGM with confounders
`(bm1, bm2, age, ecog)` and `maxk = 2`.  For a different
confounder set or `maxk`, $C$ and $\alpha$ would differ.

**Practical recommendation**: Run a small reference simulation
(50--100 replicates) at 2--3 sample sizes under $H_0$ to
calibrate $L_{\text{eff}}$ for the specific analysis configuration.
Then use the corrected approximation for threshold selection.

## Comparison to Survival

For the survival simulations in Leon et al. (2024), the
approximation worked well without a multiplicity correction.
This is because:

1. The GBSG-based DGM used $N = 300$--$700$ with $n_{\min} = 60$,
   giving $N/n_{\min} = 5$--$12$ and $L_{\text{eff}} \approx 1$--$3$.
2. Higher censoring rates reduced the number of viable candidates
   (events filter is more restrictive).
3. GRF for survival (causal survival forest) is more conservative
   in cut generation than the GLM GRF.

For binary outcomes with larger $N$ (e.g., $N = 1000$),
$L_{\text{eff}}$ grows sufficiently that the correction becomes
important.


```{r session-info}
sessionInfo()
```
