---
title: "Extending the ForestSearch Power Approximation to GLM Outcomes"
subtitle: "From Cox Var(log-HR) = 4/d to a Unified Framework Across Response Types"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---


# The Section 2.1 Approximation

## Original Formulation (Cox / Survival)

Leon et al. (2024, *Statistics in Medicine*) approximate the
probability of ForestSearch identifying a subgroup H via:

$$
P(\text{detect} | \beta, d) = \int_{-\infty}^{\infty}
\int_{-\infty}^{\infty}
\mathbf{1}(w_1 + w_2 \ge 2\log(c_1)) \;
\mathbf{1}(\min(w_1, w_2) \ge \log(c_2)) \;
\varphi(w_1; \beta, 8/d) \;
\varphi(w_2; \beta, 8/d) \;
dw_1 \, dw_2,
$$

where:

- $W_1, W_2 \stackrel{iid}{\sim} N(\beta, 8/d)$ represent
  Cox log-HR estimators from 50/50 random splits
- $\beta = \log(\theta^{\dagger}(H))$ is the true log-HR in
  the subgroup
- $d \approx n_H \cdot (1 - \text{censoring rate})$ is the
  expected number of events
- $c_1 = 1.25$ (screening threshold), $c_2 = 1.0$
  (consistency threshold)

The variance $8/d$ follows from $\text{Var}(\hat\beta_{\text{Cox}})
\approx 4/d$ per treatment arm.  Each split-half has $d/2$ events,
so $\text{Var}_{\text{split}} = 4/(d/2) = 8/d$.


## Two Properties That Enable the Approximation

The approximation rests on exactly two properties:

1. **Asymptotic normality**: $\hat\beta \stackrel{d}{\to}
   N(\beta, I^{-1}(\beta))$ where $I(\beta)$ is the Fisher
   information

2. **A simple variance formula**: $I^{-1}(\beta) \approx 4/d$,
   depending only on the number of events $d$

For **any** GLM estimator satisfying both properties, the
approximation extends directly with the appropriate
variance substitution.


# Extension to GLM Outcomes

## General Framework

For a GLM with natural parameter $\eta$ (the DINA estimand), the
MLE $\hat\eta$ is asymptotically normal:

$$
\hat\eta \stackrel{d}{\to}
N\!\left(\eta, \; \sigma^2_{\text{split}}(\eta, n_H)\right)
$$

The detection probability becomes:

$$
P(\text{detect} | \eta, n_H) = \int \int
\mathbf{1}(w_1 + w_2 \ge 2k_1) \;
\mathbf{1}(\min(w_1, w_2) \ge k_2) \;
\varphi(w_1; \eta, \sigma^2_s) \;
\varphi(w_2; \eta, \sigma^2_s) \;
dw_1 \, dw_2,
$$

where $\sigma^2_s \equiv \sigma^2_{\text{split}}$ and
$k_1, k_2$ are the screening and consistency thresholds
(on the natural-parameter scale).

The **only quantity that changes** across outcome types is
$\sigma^2_s$.  The integral structure, the bivariate normal
density, and the indicator functions are identical.


## Variance Formulas by Outcome Type

### Survival (Cox PH) --- Original

$$
\sigma^2_s = \frac{8}{d}, \quad
d = n_H \cdot (1 - p_{\text{cens}})
$$

where $d$ is the expected event count in the subgroup.

### Count (Poisson / log-IRR)

For Poisson regression with log-offset, the log-IRR estimator
has:

$$
\text{Var}(\hat\beta_{\text{IRR}}) =
\frac{1}{d_1} + \frac{1}{d_0}
$$

where $d_j$ = events in arm $j$.  For balanced arms and total
events $D = d_1 + d_0$ in the subgroup:

$$
\sigma^2_s = \frac{8}{D}
$$

This is **identical** to the survival formula with $d \to D$.
The Poisson--Cox connection makes this automatic: under
proportional hazards with a constant baseline hazard, the
Poisson IRR and the Cox HR estimate the same quantity with
the same asymptotic variance.

**Implication**: The existing `compute_detection_probability()`
function works unchanged for count outcomes---simply
pass $d = \text{total events in subgroup}$.

### Binary (Logistic / log-OR)

For logistic regression, the log-OR estimator has:

$$
\text{Var}(\hat\beta_{\text{OR}}) =
\frac{1}{n_1 \bar p_1 (1 - \bar p_1)} +
\frac{1}{n_0 \bar p_0 (1 - \bar p_0)}
$$

where $\bar p_j$ is the event probability in arm $j$ and
$n_j$ is the arm size.  For balanced arms ($n_1 = n_0 = n_H/2$)
and roughly equal event rates ($\bar p_1 \approx \bar p_0
\approx \bar p$):

$$
\text{Var}(\hat\beta_{\text{OR}}) \approx
\frac{4}{n_H \cdot \bar p (1 - \bar p)}
$$

In a half-sample ($n_{\text{split}} = n_H / 2$):

$$
\sigma^2_s = \frac{8}{n_H \cdot \bar p (1 - \bar p)}
$$

The "effective number of events" is
$d_{\text{eff}} = n_H \cdot \bar p (1 - \bar p)$,
which represents the Bernoulli information content.

**Key difference from survival**: The variance depends on
the event probability $\bar p$, not just the number of events.
For rare events ($\bar p \approx 0.05$), $d_{\text{eff}}
\approx 0.05 \cdot n_H$, which is much smaller than
$n_H \cdot (1 - p_{\text{cens}})$ in survival, so the
approximation will show lower power for rare binary events.

### Continuous (Gaussian / MD)

For the mean difference estimator from Gaussian regression:

$$
\text{Var}(\hat\beta_{\text{MD}}) =
\sigma^2_Y \left(\frac{1}{n_1} + \frac{1}{n_0}\right)
= \frac{4\sigma^2_Y}{n_H}
$$

In a half-sample:

$$
\sigma^2_s = \frac{8\sigma^2_Y}{n_H}
$$

**Key difference**: The thresholds are on the identity scale
(not log), so $k_1 = c_1$ and $k_2 = c_2$ directly (no
log transformation), and the variance depends on the
residual SD $\sigma_Y$, which must be specified or estimated.


## Summary Table

| Outcome | Estimand $\eta$ | Split Variance $\sigma^2_s$ | Effective d | Thresholds |
|---|---|---|---|---|
| Survival | $\log(\text{HR})$ | $8/d$ | $n_H(1-p_c)$ | $k_j = \log(c_j)$ |
| Count | $\log(\text{IRR})$ | $8/D$ | $D$ (total events) | $k_j = \log(c_j)$ |
| Binary | $\log(\text{OR})$ | $8/(n_H\bar p(1-\bar p))$ | $n_H\bar p(1-\bar p)$ | $k_j = \log(c_j)$ |
| Continuous | MD | $8\sigma^2_Y/n_H$ | $n_H/\sigma^2_Y$ | $k_j = c_j$ (identity) |


# Implementation Approach

## Unified `compute_detection_probability_glm()`

The existing `compute_detection_probability()` can be
generalized with a single additional parameter---the
split-sample variance:

```{r}
#| eval: false
#| code-fold: false

compute_detection_probability_glm <- function(
    effect,            # true effect (log-HR, log-OR, log-IRR, or MD)
    sigma2_split,      # split-sample variance (outcome-type-specific)
    k_screen,          # screening threshold (on natural scale)
    k_consist,         # consistency threshold (on natural scale)
    method = "cubature"
) {
  # Bivariate normal integration
  # P(W1 + W2 >= 2*k_screen, min(W1, W2) >= k_consist)
  # where W1, W2 ~ N(effect, sigma2_split) independently
  ...
}
```

## Convenience Wrappers

```{r}
#| eval: false
#| code-fold: false

# Survival / Count (identical formula)
sigma2_split_events <- function(n_sg, event_rate) {
  d <- n_sg * event_rate
  8 / d
}

# Binary
sigma2_split_binary <- function(n_sg, p_event) {
  8 / (n_sg * p_event * (1 - p_event))
}

# Continuous
sigma2_split_continuous <- function(n_sg, sigma_y) {
  8 * sigma_y^2 / n_sg
}
```


# Accuracy Assessment

## When the Approximation Is Good

The approximation relies on:

1. **CLT convergence**: $\hat\eta$ is approximately normal.
   For the Cox model, this requires $d \gtrsim 20$.  For
   logistic regression, the rule of thumb is
   $n_H \cdot \min(\bar p, 1 - \bar p) \gtrsim 10$.

2. **Variance formula accuracy**: The formulas assume
   balanced arms, no covariate adjustment, and (for binary)
   homogeneous event rates.

3. **Independence of splits**: The two half-samples are
   independent conditional on the true effect $\eta$.

## When Caution Is Needed

| Scenario | Concern | Mitigation |
|---|---|---|
| Binary, rare events ($\bar p < 0.1$) | Normal approximation poor; separation risk | Use exact methods or simulation |
| Binary, small subgroup ($n_H < 80$) | Finite-sample bias in log-OR | Firth correction variance |
| Continuous, heavy tails | $\sigma^2_Y$ may be poorly estimated | Robust variance or bootstrap |
| Count, overdispersion | True variance $>$ Poisson variance | Use quasi-Poisson: $\sigma^2_s = 8\hat\phi / D$ |
| All types, covariate adjustment | Variance reduction not captured | Adjust $d_{\text{eff}}$ upward |


# Connection to the DINA Framework

The DINA paper (*Biometrics*, 2025) proposes the natural
parameter as the universal estimand for HTE:

| Response | Natural Parameter | DINA Estimand |
|---|---|---|
| Gaussian | $\mu$ | $\mu_1(x) - \mu_0(x)$ |
| Bernoulli | $\log\frac{p}{1-p}$ | $\log\text{OR}$ |
| Poisson | $\log\mu$ | $\log\text{IRR}$ |
| Cox | $\log\Lambda$ | $\log\text{HR}$ |

The ForestSearch Section 2.1 approximation is naturally a
**DINA-scale** approximation: the bivariate normal integral
operates on the natural parameter $\eta$, and the
variance $\sigma^2_s$ is the Fisher information inverse for
$\eta$ in a half-sample.

This means the extension is not ad hoc---it follows
directly from the exponential family structure that unifies
all four response types.


# Practical Implications

## Recommended Default Thresholds

The original thresholds ($c_1 = 1.25$, $c_2 = 1.0$) were
calibrated for survival endpoints to yield ~10 percent type-I
error when $\theta^{\dagger}(H) \le 1$.

For GLM outcomes on the log scale (OR, IRR), the same thresholds
on the exponentiated scale apply directly:

- **OR / IRR**: $c_1 = 1.25 \to k_1 = \log(1.25)$,
  $c_2 = 1.0 \to k_2 = 0$

For continuous outcomes (MD), the thresholds are on the
identity scale and must be calibrated to the outcome's
natural units.  A general recommendation:

- $k_1 = 0.2 \cdot \sigma_Y$ (screening: effect exceeds
  0.2 SD units)
- $k_2 = 0$ (consistency: both halves show positive effect)

These yield similar operating characteristics to the survival
defaults when expressed in standardized-effect-size units.

## Power Curve Comparison

The approximation enables rapid power comparison across
outcome types without simulation:

```{r}
#| code-fold: false
library(ggplot2)

# Unified detection probability (Monte Carlo)
detect_prob <- function(effect, sigma2_s, k_screen, k_consist,
                        n_mc = 100000, seed = 42) {
  set.seed(seed)
  w1 <- rnorm(n_mc, mean = effect, sd = sqrt(sigma2_s))
  w2 <- rnorm(n_mc, mean = effect, sd = sqrt(sigma2_s))
  mean((w1 + w2 >= 2 * k_screen) & (w1 >= k_consist) & (w2 >= k_consist))
}

# Generate power curves for n_sg = 100
n_sg <- 100
effects <- seq(-0.5, 1.5, by = 0.05)

# Survival: 45% censoring -> d = 55 events
surv_power <- sapply(effects, function(e)
  detect_prob(e, 8 / 55, log(1.25), 0))

# Count: rate = 1.2/yr, follow-up = 1.5yr -> D = 180 events
count_power <- sapply(effects, function(e)
  detect_prob(e, 8 / 180, log(1.25), 0))

# Binary: p = 0.30 -> d_eff = 100 * 0.3 * 0.7 = 21
binary_power <- sapply(effects, function(e)
  detect_prob(e, 8 / 21, log(1.25), 0))

# Continuous: sigma = 1.5, threshold = 0 on identity scale
cont_power <- sapply(effects, function(e)
  detect_prob(e, 8 * 1.5^2 / 100, 0.2 * 1.5, 0))

df_power <- data.frame(
  effect = rep(effects, 4),
  power = c(surv_power, count_power, binary_power, cont_power),
  Outcome = rep(c("Survival (d=55)", "Count (D=180)",
                  "Binary (p=0.30)", "Continuous (sigma=1.5)"),
                each = length(effects))
)

ggplot(df_power, aes(x = effect, y = power, colour = Outcome)) +
  geom_line(linewidth = 1.0) +
  geom_hline(yintercept = 0.10, linetype = "dashed", colour = "grey50") +
  geom_hline(yintercept = 0.80, linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = 0, linetype = "dotted", colour = "grey70") +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "True Effect (natural parameter scale)",
       y = "P(detect)",
       title = "ForestSearch Detection Probability: Unified GLM Approximation",
       subtitle = sprintf("n(H) = %d | Thresholds: screen = 1.25, consist = 1.0 (log scale)", n_sg)) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold")) +
  annotate("text", x = -0.4, y = 0.12, label = "10%", size = 3) +
  annotate("text", x = -0.4, y = 0.82, label = "80%", size = 3)
```


# Conclusions

The Section 2.1 approximation extends naturally to all GLM
outcome types via the exponential family structure:

1. **Count outcomes**: The extension is exact---the variance
   formula $8/D$ is identical to the survival formula, and
   the existing `compute_detection_probability()` works
   unchanged.

2. **Binary outcomes**: The variance formula
   $8/(n_H \bar p(1-\bar p))$ replaces $8/d$, introducing a
   dependence on the event probability.  The approximation
   is accurate for $n_H \cdot \min(\bar p, 1-\bar p) \gtrsim 10$.

3. **Continuous outcomes**: The variance formula
   $8\sigma^2_Y / n_H$ requires specifying the residual SD.
   Thresholds must be on the identity scale.

4. **The DINA connection**: The approximation operates on
   the natural parameter for each exponential family member,
   making it a natural extension rather than an ad hoc
   modification.

No published work was found that extends split-sample
consistency power approximations to GLM outcomes in this
manner.  This appears to be a novel contribution that could
accompany the v0.2.0 GLM extension.


# References

- Leon, L.F., et al. (2024). Exploratory Subgroup
  Identification in the Heterogeneous Cox Model.
  *Statistics in Medicine*, 43, 4253--4275.

- DINA: Estimating Heterogeneous Treatment Effects for
  General Responses. *Biometrics*, 81(4), 2025.

- Agresti, A. (2002). *Categorical Data Analysis*, 2nd ed.
  Wiley. (Variance of log-OR: Section 3.1.3)

- McCullagh, P. and Nelder, J.A. (1989). *Generalized Linear
  Models*, 2nd ed. Chapman and Hall. (Fisher information for
  exponential families)

- Firth, D. (1993). Bias Reduction of Maximum Likelihood
  Estimates. *Biometrika*, 80, 27--38. (Finite-sample
  correction for logistic regression)
