---
title: "Null Threshold Search for Survival Outcomes"
subtitle: "Approximation-Driven Selection with Simulation Validation (GBSG DGM)"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# -- User-configurable parameters --

# Validation sims per (c1, c2, N) configuration
# ~ 500 hr linux for (at most) 10 triplets {c1,c2,N}
nsims_val       <- 1000L
sim_seed        <- 8316951L
n_workers       <- ceiling(0.9*max(1L, parallel::detectCores(logical = FALSE) - 1L))

# L_eff calibration.
# Loaded from _data/calibration_survival_leff.rds
# (produced by calibration_survival_fourway.qmd, aka C2).
# If the .rds is not found, falls back to hardcoded values.
# Set recalibrate = TRUE to re-derive from scratch (Part I).
C_prior_fallback  <- 0.029
alpha_prior_fallback <- 0.882
recalibrate     <- FALSE     # TRUE = run calibration sims
n_sims_recal    <- 200L      # Sims per N when calibrating

# FPR target for threshold selection
fpr_target      <- 0.12

# Output directory for saved results
output_dir      <- "_data"
```


# Introduction {#sec-intro}

## Strategy

Rather than sweeping a large $(c_1, c_2)$ grid by simulation
(expensive), we use the **corrected approximation** to
pre-screen threshold pairs and only simulate the promising
candidates:

1. **Calibrate $L_{\text{eff}}$** (Part I): Either use prior
   calibration values or run a small H0 simulation at multiple
   sample sizes to fit $L_{\text{eff}} = C (N/n_{\min})^\alpha$.

2. **Approximation grid** (Part II, instant): Evaluate a dense
   $(c_1, c_2)$ grid using
   $\text{FPR}_{\text{corr}} = 1 - (1 - P_1)^{L_{\text{eff}}}$.

3. **Pre-select**: Retain only pairs where the corrected FPR
   is at most `r fpr_target * 100`% at the **most stringent**
   sample size (largest $N$, where $L_{\text{eff}}$ is highest).

4. **Validate** (Part III, few sims): Run ForestSearch under
   $H_0$ on the selected pairs with `r nsims_val` simulations
   per configuration to confirm the approximation.

5. **Save** (Part V): Write all results to `.rds` for later
   re-analysis with more simulations or different graphics.


## $L_{\text{eff}}$ Calibration Source

$$
\text{FPR}_{\text{corr}}(N) = 1 - (1 - P_1)^{L_{\text{eff}}(N)},
\qquad
L_{\text{eff}}(N) = C \cdot (N / n_{\min})^\alpha
$$

Prior calibration (from `calibration_survival_fourway.qmd`,
GBSG survival DGM, 7 confounders):
loaded from `_data/calibration_survival_leff.rds` if available,
otherwise falls back to $C = `r C_prior_fallback`$,
$\alpha = `r alpha_prior_fallback`$.

Note: For survival, $L_{\text{eff}} < 1$ at typical sample
sizes because the events-based filters (`d0.min`, `d1.min`)
aggressively prune candidates.  The paper defaults
(c1 = 1.25, c2 = 1.0) are already conservative for survival.


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(weightedsurv)
library(data.table)
library(survival)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)

# L_eff save/load helpers
#source(here::here("../R", "leff_calibration_io.R"))
```

```{r load-calibration}
#| code-fold: false

# Load L_eff calibration from C2 (calibration_survival_fourway.qmd)
# Falls back to hardcoded values if .rds not found
cal_survival <- tryCatch(
  load_leff_calibration("survival", output_dir = output_dir),
  error = function(e) {
    message("Calibration .rds not found: ", e$message)
    message("Using fallback values: C=", C_prior_fallback,
            ", alpha=", alpha_prior_fallback)
    list(C = C_prior_fallback, alpha = alpha_prior_fallback,
         n_min = 60L, source = "fallback (hardcoded)")
  }
)

C_prior     <- cal_survival$C
alpha_prior <- cal_survival$alpha

cat(sprintf("L_eff prior: C = %.4f, alpha = %.4f\n",
            C_prior, alpha_prior))
if (!is.null(cal_survival$dgm_description))
  cat(sprintf("Source: %s\n", cal_survival$dgm_description))
if (!is.null(cal_survival$timestamp))
  cat(sprintf("Calibrated: %s\n",
              format(cal_survival$timestamp, "%Y-%m-%d %H:%M")))
```

```{r timing-init}
timings <- list()
t_vignette_start <- proc.time()
```

```{r setup-parallel}
plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")
cat("Validation sims per config:", nsims_val, "\n")
cat("FPR target:", fpr_target, "\n")
cat("Recalibrate L_eff:", recalibrate, "\n")
```


# Null DGM {#sec-dgm}

```{r create-dgm-null}
t0 <- proc.time()[3]

dgm_null <- setup_gbsg_dgm(model = "null", k_treat = 1.0,
                             verbose = TRUE)

cat("\nNull DGM:\n")
cat("  HR(ITT):", round(dgm_null$hr_causal, 3), "\n")
cat("  AHR:", round(dgm_null$AHR, 3), "\n")

timings$dgm <- proc.time()[3] - t0
```


# Part I: $L_{\text{eff}}$ Calibration {#sec-leff}

## Background

For survival, $L_{\text{eff}}$ is typically less than 1 because
the events-based filters (`d0.min`, `d1.min`) prune most
candidate subgroups before they reach the threshold check.
This means the procedure is *more conservative* than a
single per-subgroup test, and the paper defaults (c1 = 1.25,
c2 = 1.0) provide strong FPR control.

## Shared Parameters

```{r leff-params}
#| code-fold: false

n_min_val <- 60
prop_cens <- 0.46
p_harm    <- 0.13     # Prevalence of harm subgroup in GBSG DGM
pcons     <- 0.90

# Subgroup characteristics at reference N = 700
n_H_ref <- round(p_harm * 700)
d_dgm <- round(d_eff_survival(n_H_ref, prop_cens))
cat(sprintf("Reference: N=700, n_H=%d, d_eff=%d events\n",
            n_H_ref, d_dgm))

p_detect_glm <- function(theta, d_eff, c1, c2) {
  compute_detection_probability_glm(
    theta = theta, d_eff = d_eff, c1 = c1, c2 = c2,
    effect_scale = "ratio", method = "monte_carlo"
  )
}

# Simulation parameters for calibration / validation
confounders_base <- c("z1", "z2", "z3", "z4", "z5",
                       "size", "grade3")

fs_params_base <- list(
  hr.threshold = 1.25,
  hr.consistency = 1.0,
  pconsistency.threshold = pcons,
  fs.splits = 400,
  n.min = n_min_val,
  d0.min = 12, d1.min = 12,
  maxk = 2, by.risk = 12,
  vi.grf.min = -0.2,
  use_lasso = TRUE, use_grf = TRUE,
  use_twostage = TRUE, twostage_args = list()
)

fs_params_cox <- modifyList(fs_params_base, list(
  outcome.name = "y_sim", event.name = "event_sim",
  treat.name = "treat_sim", id.name = "id"
))

sim_config <- list(
  n_sample = 700, analysis_time = 84,
  cens_adjust = log(1.5)
)
```


## Calibration Simulation

```{r run-calibration}
if (recalibrate) {
  N_cal_values <- c(300, 500, 700, 1000)

  cat(sprintf("=== L_eff Calibration ===\n"))
  cat(sprintf("Reference thresholds: c1=1.25, c2=1.0\n"))
  cat(sprintf("N values: %s\n", paste(N_cal_values, collapse=", ")))
  cat(sprintf("Sims per N: %d\n\n", n_sims_recal))

  t0 <- proc.time()[3]

  cal_results <- list()
  for (N_val in N_cal_values) {
    t_N <- proc.time()[3]

    res_N <- foreach(
      sim = seq_len(n_sims_recal),
      .combine = rbind, .errorhandling = "remove",
      .options.future = list(
        packages = c("forestsearch", "survival", "data.table"),
        seed = TRUE)
    ) %dofuture% {
      run_simulation_analysis(
        sim_id = sim, dgm = dgm_null,
        n_sample = N_val,
        analysis_time = sim_config$analysis_time,
        cens_adjust = sim_config$cens_adjust,
        confounders_base = confounders_base,
        cox_formula_adj = survival::Surv(y_sim, event_sim) ~
          treat_sim + z1 + z2 + z3,
        n_add_noise = 0L,
        run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
        fs_params = fs_params_cox, verbose = FALSE
      )
    }

    fs_rows <- res_N[res_N$analysis == "FS", ]
    fpr_N <- mean(fs_rows$any.H, na.rm = TRUE)
    cal_results[[as.character(N_val)]] <- data.frame(
      N = N_val, sim_fpr = fpr_N, n_sims = nrow(fs_rows)
    )
    cat(sprintf("  N = %4d: FPR = %.3f (%d sims) | %.1f sec\n",
        N_val, fpr_N, nrow(fs_rows), proc.time()[3] - t_N))
  }

  timings$leff_cal <- proc.time()[3] - t0
  cat(sprintf("\nCalibration: %.1f min\n", timings$leff_cal / 60))

} else {
  timings$leff_cal <- 0
  cat("Calibration skipped (recalibrate = FALSE).\n")
  cat("Using prior values from calibration_survival_fourway.\n")
}
```


## Fit $L_{\text{eff}}$ Power Law

```{r fit-leff}
#| fig-height: 5
#| fig-cap: "L_eff calibration for the GBSG survival DGM (only shown when recalibrate = TRUE)."

if (recalibrate) {
  cal_df <- do.call(rbind, cal_results)
  cal_df$n_H <- round(p_harm * cal_df$N)
  cal_df$d_eff <- round(d_eff_survival(cal_df$n_H, prop_cens))
  cal_df$P1 <- sapply(cal_df$d_eff, function(d)
    p_detect_glm(1.0, d, 1.25, 1.0))

  cal_df$sim_fpr <- pmax(cal_df$sim_fpr, 0.5 / n_sims_recal)
  cal_df$sim_fpr <- pmin(cal_df$sim_fpr, 1 - 0.5 / n_sims_recal)
  cal_df$L_eff <- log(1 - cal_df$sim_fpr) / log(1 - cal_df$P1)

  cat("=== Implied L_eff ===\n")
  print(cal_df[, c("N", "n_H", "d_eff", "P1",
                    "sim_fpr", "L_eff")])

  cal_valid <- cal_df[is.finite(cal_df$L_eff) &
                       cal_df$L_eff > 0, ]

  if (nrow(cal_valid) >= 2) {
    leff_cal <- calibrate_L_eff(
      N = cal_valid$N, P1 = cal_valid$P1,
      sim_fpr = cal_valid$sim_fpr, n_min = n_min_val
    )
    C_fit <- leff_cal$C
    alpha_fit <- leff_cal$alpha
    print(leff_cal)

    N_pred <- seq(200, 1200, by = 5)
    pred_curve <- data.frame(
      N = N_pred,
      L_eff = C_fit * (N_pred / n_min_val)^alpha_fit
    )

    print(
      ggplot() +
        geom_point(data = cal_valid, aes(x = N, y = L_eff),
                   size = 4, colour = "firebrick") +
        geom_line(data = pred_curve, aes(x = N, y = L_eff),
                  colour = "steelblue", linewidth = 1) +
        labs(x = "Total Sample Size N",
             y = expression(L[eff]),
             title = expression(
               L[eff] * " Calibration: GBSG Survival DGM"),
             subtitle = sprintf(
               "L_eff = %.3f * (N/%d)^%.3f",
               C_fit, n_min_val, alpha_fit)) +
        theme_minimal(base_size = 12) +
        theme(plot.title = element_text(face = "bold"))
    )
  } else {
    cat("Insufficient calibration points; using prior.\n")
    C_fit <- C_prior; alpha_fit <- alpha_prior
  }

} else {
  C_fit <- C_prior
  alpha_fit <- alpha_prior
  cat(sprintf("Using prior: C = %.3f, alpha = %.3f\n",
              C_fit, alpha_fit))
}
```


## $L_{\text{eff}}$ Summary

```{r leff-summary}
N_target <- c(300, 500, 700, 1000)

cat("=== L_eff Model ===\n")
cat(sprintf("  L_eff(N) = %.3f * (N/%d)^%.3f\n\n",
            C_fit, n_min_val, alpha_fit))

cat("=== L_eff at Target Sample Sizes ===\n")
for (N_val in N_target) {
  L <- C_fit * (N_val / n_min_val)^alpha_fit
  cat(sprintf("  N = %4d: L_eff = %.2f\n", N_val, L))
}

cat(sprintf("\n=== Source: %s ===\n",
  if (recalibrate) "calibrated in this run" else "prior"))
cat("\nNote: L_eff < 1 means the procedure is MORE conservative\n")
cat("than a single per-subgroup test (events filters prune candidates).\n")
```


## Adapting for Other Scenarios {.unnumbered}

::: {.callout-tip}
## Blueprint: How to Calibrate for a New Survival DGM

**Step 1.** Replace `setup_gbsg_dgm()` / `simulate_from_dgm()`
with your H0 data generator.

**Step 2.** Update `confounders_base` and `fs_params_cox` to
match your analysis configuration.

**Step 3.** Update `p_harm` (harm subgroup prevalence) and
`prop_cens` (censoring proportion) for `d_eff_survival()`.

**Step 4.** Set `recalibrate = TRUE` and `n_sims_recal` to at
least 200.

**Step 5.** After the calibration run, record the fitted $C$
and $\alpha$ as your new `C_prior` / `alpha_prior`.
:::


# Part II: Approximation-Driven Selection {#sec-select}

## Dense Approximation Grid

```{r approx-grid}
c1_grid <- seq(1.00, 1.80, by = 0.05)
c2_grid <- seq(0.90, 1.50, by = 0.05)

full_grid <- expand.grid(c1 = c1_grid, c2 = c2_grid,
                          N = N_target,
                          stringsAsFactors = FALSE)
full_grid <- full_grid[full_grid$c2 <= full_grid$c1, ]

# Compute corrected FPR for every (c1, c2, N)
full_grid$n_H <- round(p_harm * full_grid$N)
full_grid$d_eff <- round(d_eff_survival(full_grid$n_H, prop_cens))
full_grid$L_eff <- C_fit * (full_grid$N / n_min_val)^alpha_fit

full_grid$P1 <- mapply(function(d, c1v, c2v) {
  p_detect_glm(1.0, d, c1v, c2v)
}, full_grid$d_eff, full_grid$c1, full_grid$c2)

full_grid$FPR_corr <- 1 - (1 - full_grid$P1)^full_grid$L_eff

full_dt <- data.table(full_grid)

cat(sprintf("Dense grid: %d c1 x %d c2 x %d N = %d configs\n",
    length(c1_grid), length(c2_grid), length(N_target),
    nrow(full_dt)))
```


## Corrected FPR Heatmaps

```{r approx-heatmaps, fig.width=10, fig.height=14}
#| fig-cap: "Corrected approximation FPR.  For survival with L_eff < 1, the corrected FPR is LOWER than P1."

c1_heat <- seq(1.00, 1.50, by = 0.05)
c2_heat <- c(0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20, 1.25)

heat_sub <- full_dt[c1 %in% c1_heat & c2 %in% c2_heat]

heatmap_list <- lapply(N_target, function(N_val) {
  sub <- heat_sub[N == N_val]
  ggplot(sub, aes(x = factor(c1), y = factor(c2),
                  fill = FPR_corr)) +
    geom_tile(colour = "white", linewidth = 0.5) +
    geom_text(aes(label = sprintf("%.1f%%", 100 * FPR_corr)),
              size = 2.8, colour = "black") +
    scale_fill_gradient2(
      low = "#E8F5E9", mid = "#FFF9C4", high = "#E53935",
      midpoint = fpr_target,
      limits = c(0, max(heat_sub$FPR_corr, 0.25)),
      name = "Corr FPR"
    ) +
    labs(x = expression(c[1]), y = expression(c[2]),
         title = sprintf("N = %d (L_eff = %.2f)", N_val,
           C_fit * (N_val / n_min_val)^alpha_fit)) +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold"),
          axis.text = element_text(size = 9))
})

wrap_plots(heatmap_list, ncol = 1) +
  plot_annotation(
    title = "Corrected Approximation: FPR Under H0 (GBSG Survival)",
    subtitle = sprintf(
      "L_eff = %.3f * (N/%d)^%.3f | Target: FPR <= %.0f%%",
      C_fit, n_min_val, alpha_fit, 100 * fpr_target))
```


## Pre-Select: Smallest $(c_1, c_2)$ Meeting the Target

```{r select-pairs}
#| tbl-cap: "Frontier: minimum c2 for each c1 to achieve corrected FPR <= target at all N"

N_max <- max(N_target)
fpr_wide <- full_dt[, .(
  FPR_corr_worst = FPR_corr[N == N_max]
), by = .(c1, c2)]

for (N_val in N_target) {
  col_name <- sprintf("FPR_N%d", N_val)
  fpr_wide[[col_name]] <- full_dt[N == N_val,
    FPR_corr[match(paste(fpr_wide$c1, fpr_wide$c2),
                   paste(c1, c2))]]
}

selected <- fpr_wide[FPR_corr_worst <= fpr_target]
setorder(selected, c1, c2)

cat(sprintf(
  "Pairs with corrected FPR <= %.0f%% at N=%d: %d of %d\n",
  100 * fpr_target, N_max, nrow(selected), nrow(fpr_wide)))

# Frontier
frontier <- selected[, .(
  min_c2 = min(c2),
  FPR_at_min_c2 = FPR_corr_worst[which.min(c2)]
), by = c1]
setorder(frontier, c1)

if (nrow(frontier) > 0) {
  cat(sprintf("\nSmallest c1 achieving target: %.2f (with c2=%.2f)\n",
      frontier$c1[1], frontier$min_c2[1]))

  gt(frontier[, .(
    c1,
    `Min c2` = min_c2,
    `Corr FPR (worst N)` = sprintf("%.1f%%", 100 * FPR_at_min_c2)
  )]) |>
    tab_header(
      title = md(sprintf(
        "**Frontier: Minimum c2 for Each c1 (FPR <= %.0f%%)**",
        100 * fpr_target)),
      subtitle = sprintf(
        "Binding at N=%d (L_eff=%.2f) | Paper defaults: c1=1.25, c2=1.0",
        N_max, C_fit * (N_max / n_min_val)^alpha_fit)
    ) |>
    tab_style(style = cell_fill(color = "#E8F5E9"),
      locations = cells_body(rows = c1 == 1.25))
} else {
  cat("No c1 values achieved the target.\n")
}
```


## Boundary Pairs

```{r boundary-pairs}
boundary <- fpr_wide[FPR_corr_worst > fpr_target &
                      FPR_corr_worst <= fpr_target + 0.05]

sim_candidates <- rbind(
  selected[, .(c1, c2, FPR_corr_worst, region = "selected")],
  boundary[, .(c1, c2, FPR_corr_worst, region = "boundary")]
)
setorder(sim_candidates, c1, c2)

max_pairs <- 40L
if (nrow(sim_candidates) > max_pairs) {
  n_sel <- sum(sim_candidates$region == "selected")
  n_boundary_keep <- max(max_pairs - n_sel, 5L)
  sim_candidates <- rbind(
    sim_candidates[region == "selected"],
    sim_candidates[region == "boundary"][
      seq(1, .N, length.out = min(.N, n_boundary_keep))]
  )
}

cat(sprintf(
  "Simulation candidates: %d selected + %d boundary = %d total\n",
  sum(sim_candidates$region == "selected"),
  sum(sim_candidates$region == "boundary"),
  nrow(sim_candidates)))
```


# Part III: Simulation Validation {#sec-sim}

Set `nsims_val = 0` to skip validation and rely on the
corrected approximation.  Increase `nsims_val` (e.g., 500)
for production-quality validation.

```{r run-validation}
run_validation <- nsims_val > 0

if (run_validation) {
  t0 <- proc.time()[3]

  # Only validate the frontier pairs: the minimum c2 for each c1
  # that achieves FPR_corr <= target.  At most 10 pairs.
  if (exists("frontier") && nrow(frontier) > 0) {
    val_candidates <- frontier[seq_len(min(10, nrow(frontier))),
                                .(c1, c2 = min_c2, FPR_corr = FPR_at_min_c2)]
  } else {
    val_candidates <- sim_candidates[region == "selected",
                                      .(c1, c2, FPR_corr = FPR_corr_worst)]
    if (nrow(val_candidates) > 10)
      val_candidates <- val_candidates[seq_len(10)]
  }

  cat(sprintf(
    "Validating %d frontier pairs (min c2 per c1, FPR_corr <= %.0f%%)\n",
    nrow(val_candidates), 100 * fpr_target))
  for (r in seq_len(nrow(val_candidates))) {
    cat(sprintf("  c1=%.2f, c2=%.2f (corr FPR=%.1f%%)\n",
                val_candidates$c1[r], val_candidates$c2[r],
                100 * val_candidates$FPR_corr[r]))
  }

  val_tasks <- do.call(rbind, lapply(N_target, function(N_val) {
    do.call(rbind, lapply(seq_len(nrow(val_candidates)), function(g) {
      data.frame(
        N = N_val,
        sim = seq_len(nsims_val),
        c1 = val_candidates$c1[g],
        c2 = val_candidates$c2[g],
        stringsAsFactors = FALSE
      )
    }))
  }))

  cat(sprintf(
    "Running %d validation tasks (%d pairs x %d N x %d sims)...\n",
    nrow(val_tasks), nrow(val_candidates),
    length(N_target), nsims_val))

  val_results <- foreach(
    i = seq_len(nrow(val_tasks)),
    .combine = rbind, .errorhandling = "remove",
    .options.future = list(
      packages = c("forestsearch", "survival", "data.table"),
      seed = TRUE)
  ) %dofuture% {
    N_val  <- val_tasks$N[i]
    sim_id <- val_tasks$sim[i]
    c1_val <- val_tasks$c1[i]
    c2_val <- val_tasks$c2[i]

    fs_params_grid <- modifyList(fs_params_cox, list(
      hr.threshold = c1_val,
      hr.consistency = c2_val
    ))

    res <- run_simulation_analysis(
      sim_id = sim_id, dgm = dgm_null,
      n_sample = N_val,
      analysis_time = sim_config$analysis_time,
      cens_adjust = sim_config$cens_adjust,
      confounders_base = confounders_base,
      cox_formula_adj = survival::Surv(y_sim, event_sim) ~
        treat_sim + z1 + z2 + z3,
      n_add_noise = 0L,
      run_fs = TRUE, run_fs_grf = FALSE, run_grf = FALSE,
      fs_params = fs_params_grid, verbose = FALSE
    )

    fs_row <- res[res$analysis == "FS", ]
    data.table(N = N_val, sim = sim_id,
               c1 = c1_val, c2 = c2_val,
               any_found = if (nrow(fs_row) > 0) fs_row$any.H[1]
                           else 0L)
  }

  timings$validation <- proc.time()[3] - t0
  cat(sprintf("Validation: %.1f min | %d results\n",
      timings$validation / 60, nrow(val_results)))

} else {
  timings$validation <- 0
  val_results <- NULL
  cat("Validation skipped (nsims_val = 0).\n")
  cat("Threshold selection based on corrected approximation only.\n")
}
```


# Part IV: Results {#sec-results}

```{r compute-fpr}
if (run_validation) {
  val_dt <- data.table(val_results)
  val_fpr <- val_dt[, .(
    Sim_FPR = mean(any_found, na.rm = TRUE),
    n_found = sum(any_found, na.rm = TRUE),
    n_sims = .N
  ), by = .(N, c1, c2)]

  val_fpr[, n_H := round(p_harm * N)]
  val_fpr[, d_eff := round(d_eff_survival(n_H, prop_cens))]
  val_fpr[, P1 := mapply(function(d, c1v, c2v) {
    p_detect_glm(1.0, d, c1v, c2v)
  }, d_eff, c1, c2)]
  val_fpr[, L_eff_N := C_fit * (N / n_min_val)^alpha_fit]
  val_fpr[, FPR_corr := 1 - (1 - P1)^L_eff_N]

  val_fpr <- merge(val_fpr,
    sim_candidates[, .(c1, c2, region)],
    by = c("c1", "c2"), all.x = TRUE)
} else {
  val_dt <- NULL
  val_fpr <- NULL
  cat("No simulation results (nsims_val = 0).\n")
  cat("Results below based on corrected approximation only.\n")
}
```


## Validation Table

```{r val-table}
#| tbl-cap: "Simulated FPR vs corrected approximation."

if (run_validation) {
  # CI under H0: FPR_corr is the truth
  val_fpr[, CI_lower := pmax(0, FPR_corr - 1.96 * sqrt(
    FPR_corr * (1 - FPR_corr) / n_sims))]
  val_fpr[, CI_upper := pmin(1, FPR_corr + 1.96 * sqrt(
    FPR_corr * (1 - FPR_corr) / n_sims))]
  val_fpr[, in_CI := Sim_FPR >= CI_lower & Sim_FPR <= CI_upper]

  gt(val_fpr[, .(N, c1, c2,
    `Corr FPR` = sprintf("%.1f%%", 100 * FPR_corr),
    `95% CI` = sprintf("[%.1f%%, %.1f%%]",
      100 * CI_lower, 100 * CI_upper),
    `Sim FPR` = sprintf("%.1f%% (%d/%d)",
      100 * Sim_FPR, n_found, n_sims),
    `Sim/Corr` = round(Sim_FPR / pmax(FPR_corr, 0.001), 2),
    `In CI` = ifelse(in_CI, "Yes", "No")
  )]) |>
    tab_header(
      title = md("**Simulation Validation of Selected Pairs**"),
      subtitle = sprintf(
        "%d sims/config | L_eff = %.3f * (N/%d)^%.3f",
        nsims_val, C_fit, n_min_val, alpha_fit)
    ) |>
    tab_row_group(label = "N = 1000", rows = N == 1000) |>
    tab_row_group(label = "N = 700", rows = N == 700) |>
    tab_row_group(label = "N = 500", rows = N == 500) |>
    tab_row_group(label = "N = 300", rows = N == 300) |>
    tab_style(style = cell_fill(color = "#E8F5E9"),
      locations = cells_body(columns = `In CI`,
        rows = `In CI` == "Yes")) |>
    tab_style(style = cell_fill(color = "#FFEBEE"),
      locations = cells_body(columns = `In CI`,
        rows = `In CI` == "No")) |>
    tab_source_note(paste(
      "95% CI: expected range of Sim FPR if Corr FPR is the true rate.",
      sprintf("Green = Sim FPR within CI; %d sims per config.",
              nsims_val)))
} else {
  cat("Validation table skipped (nsims_val = 0).\n")
}
```


## Validation Scatter

```{r val-scatter, fig.width=9, fig.height=6}
#| fig-cap: "Corrected approximation vs simulated FPR."

if (run_validation) {
  print(
    ggplot(val_fpr, aes(x = FPR_corr, y = Sim_FPR)) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed",
                  colour = "grey50") +
      geom_point(aes(colour = region, shape = region), size = 3,
                 alpha = 0.7) +
      geom_hline(yintercept = fpr_target, linetype = "dotted",
                 colour = "firebrick") +
      geom_vline(xintercept = fpr_target, linetype = "dotted",
                 colour = "firebrick") +
      scale_colour_manual(
        values = c("selected" = "steelblue",
                   "boundary" = "orange")) +
      facet_wrap(~ paste("N =", N)) +
      coord_equal(
        xlim = c(0, max(val_fpr$FPR_corr, 0.20)),
        ylim = c(0, max(val_fpr$Sim_FPR, 0.20))) +
      labs(x = "Corrected Approximation",
           y = "Simulated FPR",
           title = "Validation: Corrected Approximation vs Simulation",
           colour = "Region", shape = "Region") +
      theme_minimal(base_size = 12) +
      theme(legend.position = "top",
            plot.title = element_text(face = "bold"))
  )
} else {
  cat("Scatter plot skipped (nsims_val = 0).\n")
}
```


## Confirmed Acceptable Pairs

```{r confirmed-pairs}
#| tbl-cap: "Confirmed acceptable pairs"

if (run_validation) {
  confirmed <- val_fpr[Sim_FPR <= fpr_target &
                        FPR_corr <= fpr_target]
  setorder(confirmed, N, -Sim_FPR)

  if (nrow(confirmed) > 0) {
    gt(confirmed[, .(N, c1, c2,
      `Corr FPR` = sprintf("%.1f%%", 100 * FPR_corr),
      `95% CI` = sprintf("[%.1f%%, %.1f%%]",
        100 * CI_lower, 100 * CI_upper),
      `Sim FPR` = sprintf("%.1f%%", 100 * Sim_FPR),
      `Sim/Corr` = round(Sim_FPR / pmax(FPR_corr, 0.001), 2),
      `In CI` = ifelse(in_CI, "Yes", "No")
    )]) |>
      tab_header(
        title = md(sprintf(
          "**Confirmed Pairs: FPR <= %.0f%% (Approx AND Sim)**",
          100 * fpr_target)),
        subtitle = sprintf("%d sims per config", nsims_val)
      ) |>
      tab_style(style = cell_fill(color = "#E8F5E9"),
        locations = cells_body(columns = `In CI`,
          rows = `In CI` == "Yes")) |>
      tab_style(style = cell_fill(color = "#FFEBEE"),
        locations = cells_body(columns = `In CI`,
          rows = `In CI` == "No")) |>
      tab_source_note(paste(
        "95% CI bounds treat Corr FPR as the true rate.",
        "'In CI' = simulated FPR falls within expected range."))
  } else {
    cat("No pairs confirmed at this sim count.\n")
  }
} else {
  confirmed <- selected
  confirmed[, N := NA_integer_]
  cat("No simulation validation (nsims_val = 0).\n")
  cat(sprintf(
    "Approximation-selected pairs (corrected FPR <= %.0f%%): %d\n",
    100 * fpr_target, nrow(confirmed)))
}
```


# Part V: Save Results {#sec-save}

```{r save-results}
# If recalibrated in this run, save the new L_eff values
# in the standardized format so downstream documents can load them
if (recalibrate && exists("leff_cal")) {
  save_leff_calibration(
    C = C_fit,
    alpha = alpha_fit,
    n_min = n_min_val,
    outcome_type = "survival",
    cal_data = if (exists("cal_df")) cal_df else NULL,
    dgm_description = paste(
      "GBSG survival DGM, 7 confounders",
      "(z1-z5, size, grade3),",
      sprintf("prop_cens=%.2f, p_harm=%.2f", prop_cens, p_harm)),
    n_sims_per_N = n_sims_recal,
    output_dir = output_dir
  )
}

# Save the full selection results bundle
results_bundle <- list(
  params = list(
    nsims_val = nsims_val, sim_seed = sim_seed,
    n_workers = n_workers,
    C_fit = C_fit, alpha_fit = alpha_fit,
    recalibrate = recalibrate,
    n_sims_recal = n_sims_recal,
    fpr_target = fpr_target,
    n_min_val = n_min_val, prop_cens = prop_cens,
    p_harm = p_harm, pcons = pcons,
    N_target = N_target,
    c1_grid = c1_grid, c2_grid = c2_grid,
    timestamp = Sys.time()
  ),
  leff_source = list(
    C = C_fit, alpha = alpha_fit, n_min = n_min_val,
    source = if (recalibrate) "recalibrated in this run"
             else if (!is.null(cal_survival$timestamp))
               sprintf("loaded from .rds (calibrated %s)",
                 format(cal_survival$timestamp, "%Y-%m-%d"))
             else "fallback (hardcoded)"
  ),
  leff_raw_data = if (exists("cal_df")) cal_df else NULL,
  approx_grid = full_dt,
  sim_candidates = sim_candidates,
  frontier = frontier,
  val_results_raw = val_dt,
  val_fpr = val_fpr,
  confirmed = if (exists("confirmed")) confirmed else NULL,
  timings = timings
)

if (!is.null(output_dir)) {
  if (!dir.exists(output_dir)) dir.create(output_dir,
                                           recursive = TRUE)
  rds_path <- file.path(output_dir,
    sprintf("selection_survival_frontier_%s.rds",
            format(Sys.Date(), "%Y%m%d")))
  saveRDS(results_bundle, rds_path)
  cat(sprintf("Selection results saved to: %s\n", rds_path))
  cat(sprintf("  Size: %.1f KB\n", file.size(rds_path) / 1024))
} else {
  cat("Saving skipped (output_dir = NULL)\n")
}
```


## How to Reload

```{r reload-example}
#| eval: false
#| code-fold: false

# --- Load L_eff calibration (for use in other documents) ---
cal <- load_leff_calibration("survival")
cal$C       # 0.029
cal$alpha   # 0.882

# --- Load selection results ---
res <- readRDS("_data/selection_survival_frontier_20260330.rds")
res$leff_source    # Where L_eff came from
res$frontier       # Smallest c1 for each target
res$approx_grid    # Full approximation grid
res$val_fpr        # Simulation results (if run)
```


# Summary {#sec-summary}

```{r summary}
cat("=== Approximation-Driven Null Threshold Search ===\n\n")
cat(sprintf("DGM: GBSG survival, 7 confounders\n"))
cat(sprintf("n.min=%d, pcons=%.2f, maxk=2\n", n_min_val, pcons))
cat(sprintf("L_eff = %.3f * (N/%d)^%.3f",
            C_fit, n_min_val, alpha_fit))
if (!recalibrate) cat("  [from prior calibration]")
cat("\n\n")

cat(sprintf("Dense grid: %d c1 x %d c2 values\n",
            length(c1_grid), length(c2_grid)))
cat(sprintf("Target: corrected FPR <= %.0f%% at all N in {%s}\n\n",
            100 * fpr_target,
            paste(N_target, collapse = ", ")))

if (exists("frontier") && nrow(frontier) > 0) {
  cat("--- Frontier (smallest c1, minimum c2) ---\n")
  cat(sprintf("  Smallest c1: %.2f (with c2 >= %.2f)\n\n",
              frontier$c1[1], frontier$min_c2[1]))
  for (r in seq_len(min(10, nrow(frontier)))) {
    cat(sprintf("  c1=%.2f: min c2=%.2f (corr FPR=%.1f%%)\n",
                frontier$c1[r], frontier$min_c2[r],
                100 * frontier$FPR_at_min_c2[r]))
  }
  cat("\n")
}

cat(sprintf("Selected: %d pairs | Boundary: %d pairs\n",
            sum(sim_candidates$region == "selected"),
            sum(sim_candidates$region == "boundary")))

if (run_validation) {
  n_confirmed <- if (!is.null(confirmed)) nrow(confirmed) else 0
  cat(sprintf("Confirmed: %d configs\n", n_confirmed))
} else {
  cat("Validation: SKIPPED (nsims_val = 0)\n")
}

cat("\n--- Paper defaults (c1=1.25, c2=1.0) ---\n")
paper <- fpr_wide[c1 == 1.25 & c2 == 1.00]
if (nrow(paper) > 0) {
  cat(sprintf("  Corrected FPR at N=%d: %.1f%%\n",
              N_max, 100 * paper$FPR_corr_worst))
  cat("  (For survival, paper defaults are typically adequate.)\n")
}
```


# Timing and Session Info {#sec-session}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

stages <- c("DGM", "L_eff Calibration")
seconds <- c(timings$dgm, timings$leff_cal)
if (run_validation) {
  stages <- c(stages, "Validation Simulation")
  seconds <- c(seconds, timings$validation)
}
stages <- c(stages, "Total")
seconds <- c(seconds, timings$total)

timing_df <- data.frame(
  Stage = stages, Seconds = seconds,
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60

gt(timing_df) |>
  tab_header(
    title = "Computational Timing Summary",
    subtitle = if (run_validation)
      sprintf("%d pairs x %d N x %d sims | %d workers",
        nrow(sim_candidates), length(N_target),
        nsims_val, n_workers)
    else
      sprintf("Approximation only | %d workers", n_workers)
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)",
             Minutes = "Time (min)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total")
  )
```

```{r session-info}
plan(sequential)
sessionInfo()
```
