---
title: "Null Threshold Search for Binary Outcomes"
subtitle: "Approximation-Driven Selection with Simulation Validation"
author: "Larry F. Leon"
date: today
execute:
  warning: true
  message: true
---

# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(data.table)
library(ggplot2)
library(patchwork)
library(gt)
library(foreach)
library(doFuture)
library(future)
```

```{r timing-init}
timings <- list()
t_vignette_start <- proc.time()
```


```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ── Simulation control ──────────────────────────────────────────
# 1000/hr mac
nsims_val       <- 500L    # validation sims per (c1, c2, N)
sim_seed        <- 8316951L
n_workers       <- ceiling(0.95 * max(1L,
  parallel::detectCores(logical = FALSE) - 1L))

# ── Output ──────────────────────────────────────────────────────
output_dir      <- "_data"

# ── L_eff calibration source ──────────────────────────────────
# Loaded from C1's .rds.  Shared DGM/FS parameters are inherited
# from the calibration bundle (p_event, confounders_name, pcons,
# etc.).  Only override below if you intentionally want to differ.
C_prior_fallback     <- 0.220
alpha_prior_fallback <- 1.298
recalibrate          <- FALSE
n_sims_recal         <- 200L

# ── FPR target ─────────────────────────────────────────────────
fpr_target      <- 0.12

# ── Shared DGM / FS parameters (fallbacks if C1 .rds missing) ─
# These are overwritten by C1's saved config when available.
p_event          <- 0.30
p_harm           <- 0.30
theta_h1         <- 2.0   # OR for harm subgroup under H1
n_min_val        <- 60L
confounders_name <- c("bm1", "bm2", "age", "ecog")
pcons            <- 0.90
fs_splits        <- 400L
maxk_val         <- 2L
N_cal_values     <- c(300, 500, 700)
N_target         <- c(300, 500, 700)

# ── Threshold grid for approximation (S1-specific) ───────────
# When use_c1_grid = TRUE (default), the grid range is set to
# C1's validated range from the .rds bundle.  The step size
# (grid_step) controls density — S1 can be denser than C1.
# When FALSE, the user-specified grids below are used as-is
# (a warning is printed if they extend beyond C1's range).
use_c1_grid <- TRUE
grid_step   <- 0.05      # step size for approximation grid
c1_grid     <- seq(1.00, 2.00, by = grid_step)  # fallback only
c2_grid     <- seq(0.90, 1.60, by = grid_step)  # fallback only

# ── Validation ────────────────────────────────────────────────
max_pairs   <- 40L
nsims_power <- 200L      # H1 sims at optimal pair (0 = approx only)

# ── Heatmap display subset ────────────────────────────────────
c1_heat     <- seq(1.20, 1.90, by = 0.10)
c2_heat     <- c(1.00, 1.10, 1.15, 1.20, 1.25, 1.30, 1.40, 1.50)
```

```{r load-calibration}
#| code-fold: false

# Load L_eff calibration from C1 (calibration_binary_leff.qmd)
cal_binary <- tryCatch(
  load_leff_calibration("binary", output_dir = output_dir),
  error = function(e) {
    message("Calibration .rds not found: ", e$message)
    message("Using fallback values from setup chunk.")
    list(C = C_prior_fallback, alpha = alpha_prior_fallback,
         n_min = 60L, source = "fallback (hardcoded)",
         extra = list())
  }
)

C_prior     <- cal_binary$C
alpha_prior <- cal_binary$alpha

# Inherit shared parameters from C1's config (if available)
cfg <- cal_binary$extra
if (length(cfg) > 0) {
  cat("=== Inheriting config from C1 calibration ===\n")
  if (!is.null(cfg$p_event))          p_event          <- cfg$p_event
  if (!is.null(cfg$p_harm))           p_harm           <- cfg$p_harm
  if (!is.null(cfg$theta_h1))         theta_h1         <- cfg$theta_h1
  if (!is.null(cfg$confounders_name)) confounders_name <- cfg$confounders_name
  if (!is.null(cfg$pcons))            pcons            <- cfg$pcons
  if (!is.null(cfg$fs_splits))        fs_splits        <- cfg$fs_splits
  if (!is.null(cfg$maxk_val))         maxk_val         <- cfg$maxk_val
  if (!is.null(cfg$N_cal_values))     N_cal_values     <- cfg$N_cal_values
  if (!is.null(cfg$fpr_target))       fpr_target       <- cfg$fpr_target
  cat(sprintf("  confounders: %s\n", paste(confounders_name, collapse = ", ")))
  cat(sprintf("  p_event=%.2f, pcons=%.2f, fs_splits=%d, maxk=%d\n",
              p_event, pcons, fs_splits, maxk_val))
  cat(sprintf("  fpr_target=%.2f, N_cal=%s\n",
              fpr_target, paste(N_cal_values, collapse = ", ")))

  # Grid range: clamp to C1's validated range or warn
  if (!is.null(cfg$c1_range) && !is.null(cfg$c2_range)) {
    if (use_c1_grid) {
      # Default: adopt C1's validated range with S1's step size
      c1_grid <- seq(cfg$c1_range[1], cfg$c1_range[2], by = grid_step)
      c2_grid <- seq(cfg$c2_range[1], cfg$c2_range[2], by = grid_step)
      cat(sprintf("  Grid clamped to C1's validated range:\n"))
      cat(sprintf("    c1_grid = seq(%.2f, %.2f, by = %.2f) [%d values]\n",
                  min(c1_grid), max(c1_grid), grid_step, length(c1_grid)))
      cat(sprintf("    c2_grid = seq(%.2f, %.2f, by = %.2f) [%d values]\n",
                  min(c2_grid), max(c2_grid), grid_step, length(c2_grid)))
    } else {
      # User override: keep setup-chunk grid, warn if out of range
      cat("  Grid: using user-specified values (use_c1_grid = FALSE)\n")
      if (min(c1_grid) < cfg$c1_range[1] ||
          max(c1_grid) > cfg$c1_range[2]) {
        warning(sprintf(paste0(
          "S1 c1_grid [%.2f, %.2f] extends beyond C1's validated ",
          "range [%.2f, %.2f].\n",
          "  L_eff threshold-independence is unverified outside ",
          "this range.\n",
          "  Consider expanding c1_grid in C1 and re-running."),
          min(c1_grid), max(c1_grid),
          cfg$c1_range[1], cfg$c1_range[2]),
          call. = FALSE)
      }
      if (min(c2_grid) < cfg$c2_range[1] ||
          max(c2_grid) > cfg$c2_range[2]) {
        warning(sprintf(paste0(
          "S1 c2_grid [%.2f, %.2f] extends beyond C1's validated ",
          "range [%.2f, %.2f].\n",
          "  Consider expanding c2_grid in C1 and re-running."),
          min(c2_grid), max(c2_grid),
          cfg$c2_range[1], cfg$c2_range[2]),
          call. = FALSE)
      }
    }
  } else {
    cat("  Grid range: no C1 range info; using setup-chunk values.\n")
  }
} else {
  cat("=== Using fallback config from setup chunk ===\n")
  cat("(C1 .rds not found or has no extra config)\n")
}

cat(sprintf("\nL_eff prior: C = %.4f, alpha = %.4f\n",
            C_prior, alpha_prior))
if (!is.null(cal_binary$dgm_description))
  cat(sprintf("Source: %s\n", cal_binary$dgm_description))
if (!is.null(cal_binary$timestamp))
  cat(sprintf("Calibrated: %s\n",
              format(cal_binary$timestamp, "%Y-%m-%d %H:%M")))
```

::: {.callout-note}
## Calibration Source

The $L_{\text{eff}}$ prior and shared ForestSearch parameters
were loaded from `_data/calibration_binary_leff.rds`, produced
by `calibration_binary_leff.qmd` (C1).

- **If the `.rds` exists:** inherits `p_event`, `confounders_name`,
  `pcons`, `fs_splits`, `maxk_val`, `fpr_target`, and `N_cal_values`
  from C1's config.
- **Grid range** (`use_c1_grid = TRUE`, default): `c1_grid` and
  `c2_grid` are set to C1's validated range at S1's `grid_step`.
  Set `use_c1_grid = FALSE` to use the setup-chunk values instead
  (a warning is printed if they extend beyond C1's range).
- **If `.rds` not found:** uses the fallback values defined in the
  setup chunk.
- **To recalibrate from scratch:** set `recalibrate = TRUE`.
:::

```{r setup-parallel}
plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")
cat("Validation sims per config:", nsims_val, "\n")
cat("Recalibrate L_eff:", recalibrate, "\n")
```


# Introduction {#sec-intro}

## Strategy

Rather than sweeping a large $(c_1, c_2)$ grid by simulation
(expensive), we use the **corrected approximation** to
pre-screen threshold pairs and only simulate the promising
candidates:

1. **Calibrate $L_{\text{eff}}$** (Part I): Either use prior
   calibration values or run a small H0 simulation at multiple
   sample sizes to fit $L_{\text{eff}} = C (N/n_{\min})^\alpha$.
   This step is the key to adapting for new scenarios.

2. **Approximation grid** (Part II, instant): Evaluate a dense
   $(c_1, c_2)$ grid using
   $\text{FPR}_{\text{corr}} = 1 - (1 - P_1)^{L_{\text{eff}}}$.

3. **Pre-select**: Retain only pairs where the corrected FPR
   is at most `r fpr_target * 100`% at the **most stringent**
   sample size (largest $N$, where $L_{\text{eff}}$ is highest).

4. **Validate** (Part III, few sims): Run ForestSearch under
   $H_0$ on the selected pairs with `r nsims_val` simulations
   per configuration to confirm the approximation.

5. **Save** (Part V): Write all results to `.rds` for later
   re-analysis with more simulations or different graphics.


## $L_{\text{eff}}$ Calibration Source

The corrected approximation uses:

$$
\text{FPR}_{\text{corr}}(N) = 1 - (1 - P_1)^{L_{\text{eff}}(N)},
\qquad
L_{\text{eff}}(N) = C \cdot (N / n_{\min})^\alpha
$$

Prior calibration (from `calibration_binary_leff.qmd`,
5000 sims per config):
$C = `r C_prior`$, $\alpha = `r alpha_prior`$.




## Calibration Simulation Flow (from C1) {.unnumbered}

::: {.callout-note collapse="true"}
## How C1's Calibration Simulations Work

**C1** (`calibration_binary_leff.qmd`) runs simulations to
calibrate $L_{\text{eff}}$ and validate threshold-independence.
S1 consumes C1's fitted $(C, \alpha)$ and config.

**Grid construction.** C1 defines `c1_grid` and `c2_grid` in
its setup chunk.  `expand.grid()` with the `c2 <= c1` filter
produces the valid pairs.  Each pair is crossed with
`N_cal_values` (sample sizes), hypotheses (H0, optionally H1),
and `nsims` replications.

**Data generation.** `sim_binary(n, seed, has_hte)` generates
4 confounders (`bm1`, `bm2`, `age`, `ecog`), 50/50 treatment,
and a binary outcome with baseline event rate `p_event`.
Under H1, the harm subgroup (`bm1 == 0`, prevalence `p_harm`)
receives OR = `theta_h1`; the complement receives
OR = 1/`theta_h1`.  Under H0, no treatment effect.

**ForestSearch call.** Each task calls `forestsearch()` with
the task's `(c1, c2)` as screening/consistency thresholds and
the shared parameters (`pcons`, `fs_splits`, `n.min`, `maxk`).
Returns whether a subgroup was found (0/1).

**Aggregation.** Results are grouped by
`(n_total, has_hte, c1, c2)`.  The detection rate
`mean(any_found)` gives simulated FPR (H0) or power (H1).

**L_eff calibration (H0 only).** The per-subgroup approximation
$P_1$ is computed at `n_H = n.min`.  The implied
$L_{\text{eff}} = \log(1 - \text{FPR}_{\text{sim}}) / \log(1 - P_1)$
is computed at each `(c1, c2, N)`.  Consistency of
$L_{\text{eff}}$ across threshold pairs at the same $N$
validates the key assumption.  The power law
$L_{\text{eff}}(N) = C (N/n_{\min})^\alpha$ is fitted to the
medians.

**What S1 uses.** S1 loads $C$ and $\alpha$ from C1's `.rds`.
It also inherits the ForestSearch config and grid ranges.
S1's approximation grid can be denser than C1's simulation grid,
but defaults to C1's validated range (see `use_c1_grid`).
:::


# Binary Null DGM {#sec-dgm}

```{r sim-null-fn}
#| code-fold: false

sim_binary_null <- function(n, seed) {
  set.seed(seed)
  df <- data.frame(
    id = seq_len(n),
    treat = rbinom(n, 1, 0.5),
    bm1 = as.factor(rbinom(n, 1, 0.70)),
    bm2 = as.factor(rbinom(n, 1, 0.50)),
    age = round(rnorm(n, 55, 10)),
    ecog = as.factor(sample(0:1, n, replace = TRUE,
                             prob = c(0.6, 0.4)))
  )
  lp <- with(df, qlogis(p_event) + 0.02 * (age - 55) +
               0.3 * (ecog == 1))
  df$progressed <- rbinom(n, 1, plogis(lp))
  df
}

# H1 DGM: adds treatment effect for power validation at optimal pair
sim_binary_h1 <- function(n, seed) {
  set.seed(seed)
  df <- data.frame(
    id = seq_len(n),
    treat = rbinom(n, 1, 0.5),
    bm1 = as.factor(rbinom(n, 1, 0.70)),
    bm2 = as.factor(rbinom(n, 1, 0.50)),
    age = round(rnorm(n, 55, 10)),
    ecog = as.factor(sample(0:1, n, replace = TRUE,
                             prob = c(0.6, 0.4)))
  )
  lp <- with(df, qlogis(p_event) + 0.02 * (age - 55) +
               0.3 * (ecog == 1))
  # Harm subgroup (bm1 == 0): OR = theta_h1
  # Complement  (bm1 == 1): OR = 1/theta_h1
  lp <- lp + with(df,
    log(theta_h1)     * treat * (bm1 == 0) +
    log(1 / theta_h1) * treat * (bm1 == 1))
  df$progressed <- rbinom(n, 1, plogis(lp))
  df$flag_harm  <- as.integer(df$bm1 == 0)
  df
}
```


# Part I: $L_{\text{eff}}$ Calibration {#sec-leff}

## Background

The per-subgroup approximation $P_1$ gives the probability
that a single pre-specified subgroup passes the screening and
consistency criteria.  The procedure-level FPR is higher
because ForestSearch evaluates multiple candidates:

$$
\text{FPR}_{\text{corr}}(N) =
1 - (1 - P_1)^{L_{\text{eff}}(N)},
\qquad
L_{\text{eff}}(N) = C \cdot (N / n_{\min})^\alpha
$$

$L_{\text{eff}}$ depends on the **candidate generation process**
(confounder set, GRF, LASSO, maxk, n.min) --- not on the
thresholds $(c_1, c_2)$.  It must be calibrated from a small
$H_0$ reference simulation for each analysis configuration.

The calibration procedure is:

1. Run ForestSearch under $H_0$ at a reference threshold pair
   (e.g., c1 = 1.25, c2 = 1.0) across $\ge 3$ sample sizes.
2. At each $N$, compute $P_1$ from the approximation and
   $\text{FPR}_{\text{sim}}$ from simulation.
3. Solve for implied $L_{\text{eff}} =
   \log(1 - \text{FPR}_{\text{sim}}) / \log(1 - P_1)$.
4. Fit the power law $L_{\text{eff}} = C (N/n_{\min})^\alpha$.


## Derived Objects

```{r leff-params}
#| code-fold: false

# Convenience wrapper for detection probability
p_detect <- function(theta, d_eff, c1, c2) {
  compute_detection_probability_glm(
    theta = theta, d_eff = d_eff, c1 = c1, c2 = c2,
    effect_scale = "ratio", method = "monte_carlo"
  )
}

cat(sprintf("n.min=%d, pcons=%.2f, maxk=%d, p_event=%.2f\n",
            n_min_val, pcons, maxk_val, p_event))
```


## Calibration Simulation

This section runs ForestSearch under $H_0$ at a reference
threshold (c1 = 1.25, c2 = 1.0) across multiple sample sizes
to estimate $L_{\text{eff}}$.

Set `run_calibration = TRUE` in the setup chunk to execute.
When `FALSE`, the prior values ($C = `r C_prior`$,
$\alpha = `r alpha_prior`$) are used.

**To adapt for a new scenario**: modify `sim_binary_null()`
(Section 3), the `confounders.name` list, and
`outcome_type` / `effect_measure` in the ForestSearch call
below.  Then set `run_calibration = TRUE`.

```{r run-calibration}
if (recalibrate) {
  # Reference thresholds for calibration
  c1_cal <- 1.25
  c2_cal <- 1.0

  cat(sprintf("=== L_eff Calibration ===\n"))
  cat(sprintf("Reference thresholds: c1=%.2f, c2=%.2f\n",
              c1_cal, c2_cal))
  cat(sprintf("N values: %s\n", paste(N_cal_values, collapse=", ")))
  cat(sprintf("Sims per N: %d\n\n", n_sims_recal))

  t0 <- proc.time()[3]

  cal_results <- list()
  for (N_val in N_cal_values) {
    t_N <- proc.time()[3]

    res_N <- foreach(
      s = seq_len(n_sims_recal),
      .combine = c, .errorhandling = "remove",
      .options.future = list(
        packages = c("forestsearch", "data.table"), seed = TRUE)
    ) %dofuture% {
      future::plan("sequential")
      seed <- sim_seed + (N_val * 1000L) + s
      df_null <- sim_binary_null(N_val, seed)

      # ---- ForestSearch call ----
      # Adapt this block for other scenarios:
      #   - confounders.name: match the DGM
      #   - outcome.name / event.name: match the DGM
      #   - outcome_type / effect_measure: match the endpoint
      fs <- tryCatch(
        suppressWarnings(
          forestsearch(
            df.analysis = df_null,
            confounders.name = confounders_name,
            outcome.name = "progressed",
            event.name = "progressed",
            treat.name = "treat", id.name = "id",
            outcome_type = "binary", effect_measure = "OR",
            adverse_outcome = TRUE,
            hr.threshold = c1_cal,
            hr.consistency = c2_cal,
            pconsistency.threshold = pcons, fs.splits = fs_splits,
            n.min = n_min_val, maxk = maxk_val,
            use_lasso = TRUE, use_grf = TRUE,
            use_twostage = TRUE,
            is.RCT = TRUE, seedit = seed,
            details = FALSE, quiet = TRUE, plot.sg = FALSE,
            parallel_args = list(plan = "sequential",
                                 workers = 1,
                                 show_message = FALSE)
          )
        ),
        error = function(e) NULL
      )
      found <- !is.null(fs) && !is.null(fs$sg.harm) &&
               length(fs$sg.harm) > 0
      as.integer(found)
    }

    res_clean <- unlist(res_N[sapply(res_N, is.numeric)])
    if (length(res_clean) < n_sims_recal) {
      warning(sprintf(
        "N=%d: %d of %d tasks silently failed.",
        N_val, n_sims_recal - length(res_clean), n_sims_recal
      ))
    }
    fpr_N <- mean(res_clean, na.rm = TRUE)
    cal_results[[as.character(N_val)]] <- data.frame(
      N = N_val, sim_fpr = fpr_N,
      n_sims = length(res_clean)
    )
    cat(sprintf("  N = %4d: FPR = %.3f (%d/%d) | %.1f sec\n",
        N_val, fpr_N, sum(res_clean), length(res_clean),
        proc.time()[3] - t_N))
  }

  timings$leff_cal <- proc.time()[3] - t0
  cat(sprintf("\nCalibration: %.1f min\n", timings$leff_cal / 60))

} else {
  timings$leff_cal <- 0
  cat("Calibration skipped (recalibrate = FALSE).\n")
  cat("Using prior values from calibration_binary_leff.qmd.\n")
}
```


## Fit $L_{\text{eff}}$ Power Law

```{r fit-leff}
#| fig-height: 5
#| fig-cap: "L_eff calibration: implied L_eff at each N with power-law fit.  Only shown when recalibrate = TRUE."

if (recalibrate) {
  cal_df <- do.call(rbind, cal_results)

  # Compute d_eff and P1 at reference thresholds
  cal_df$n_H <- n_min_val
  cal_df$d_eff <- d_eff_binary(cal_df$n_H, p_event)
  cal_df$P1 <- sapply(cal_df$d_eff, function(d)
    p_detect(1.0, d, c1_cal, c2_cal))

  # Clamp sim_fpr away from 0 and 1 for log transform
  cal_df$sim_fpr <- pmax(cal_df$sim_fpr, 0.5 / n_sims_recal)
  cal_df$sim_fpr <- pmin(cal_df$sim_fpr, 1 - 0.5 / n_sims_recal)

  # Implied L_eff
  cal_df$L_eff <- log(1 - cal_df$sim_fpr) / log(1 - cal_df$P1)

  cat("=== Implied L_eff ===\n")
  print(cal_df[, c("N", "n_H", "d_eff", "P1",
                    "sim_fpr", "L_eff")])

  # Fit power law via calibrate_L_eff()
  cal_valid <- cal_df[is.finite(cal_df$L_eff) &
                       cal_df$L_eff > 0, ]

  if (nrow(cal_valid) >= 2) {
    leff_cal <- calibrate_L_eff(
      N = cal_valid$N, P1 = cal_valid$P1,
      sim_fpr = cal_valid$sim_fpr, n_min = n_min_val
    )
    C_fit <- leff_cal$C
    alpha_fit <- leff_cal$alpha
    print(leff_cal)

    # Diagnostic plot
    N_pred <- seq(100, 1200, by = 5)
    pred_curve <- data.frame(
      N = N_pred,
      L_eff = C_fit * (N_pred / n_min_val)^alpha_fit
    )

    print(
      ggplot() +
        geom_point(data = cal_valid, aes(x = N, y = L_eff),
                   size = 4, colour = "firebrick") +
        geom_line(data = pred_curve, aes(x = N, y = L_eff),
                  colour = "steelblue", linewidth = 1) +
        labs(x = "Total Sample Size N",
             y = expression(L[eff]),
             title = expression(
               L[eff] * " Calibration (from this run)"),
             subtitle = sprintf(
               "L_eff = %.3f * (N/%d)^%.3f  (R-sq = %.3f)",
               C_fit, n_min_val, alpha_fit,
               leff_cal$r_squared)) +
        theme_minimal(base_size = 12) +
        theme(plot.title = element_text(face = "bold"))
    )
  } else {
    cat("Insufficient calibration points; using prior.\n")
    C_fit <- C_prior; alpha_fit <- alpha_prior
  }

} else {
  # Use prior calibration
  C_fit <- C_prior
  alpha_fit <- alpha_prior
  cat(sprintf("Using prior: C = %.3f, alpha = %.3f\n",
              C_fit, alpha_fit))
}
```


## $L_{\text{eff}}$ Summary

```{r leff-summary}
cat("=== L_eff Model ===\n")
cat(sprintf("  L_eff(N) = %.3f * (N/%d)^%.3f\n\n",
            C_fit, n_min_val, alpha_fit))

cat("=== L_eff at Target Sample Sizes ===\n")
for (N_val in N_target) {
  L <- C_fit * (N_val / n_min_val)^alpha_fit
  cat(sprintf("  N = %4d: L_eff = %.2f\n", N_val, L))
}

cat(sprintf("\n=== Source: %s ===\n",
  if (recalibrate) "calibrated in this run" else "prior"))
```


## Adapting for Other Scenarios {.unnumbered}

::: {.callout-tip}
## Blueprint: How to Calibrate for a New DGM

To use this document as a template for a different analysis
configuration:
 
**Step 1.** Replace `sim_binary_null()` (Section 3) with your
H0 data generator.

**Step 2.** In the `run-calibration` chunk, update the
`forestsearch()` call:

- `confounders.name`: your candidate factors
- `outcome.name`, `event.name`: your endpoint
- `outcome_type`, `effect_measure`: e.g., `"survival"` /
  `"HR"`, `"count"` / `"IRR"`, `"continuous"` / `"MD"`
- `offset.name`: if using Poisson + offset

**Step 3.** Update the `d_eff` calculation to match the
outcome type:

- Binary: `d_eff_binary(n_H, p_event)`
- Survival: `d_eff_survival(n_H, prop_cens)`
- Count: `d_eff_count(total_events = D)`
- Continuous: `d_eff_continuous(n_H, sigma_y)`

**Step 4.** Set `recalibrate = TRUE` and `n_sims_recal` to at
least 100 (more is better for stable $L_{\text{eff}}$ estimates).

**Step 5.** After the calibration run, record the fitted $C$
and $\alpha$ as your new `C_prior` / `alpha_prior` for future
runs with `recalibrate = FALSE`.

The calibration typically takes 10--30 minutes on a 12-core
Mac (100 sims x 5 N values).  Subsequent threshold searches
with `recalibrate = FALSE` use the prior values and skip
directly to the approximation grid (instant) and validation
simulation.
:::


# Part II: Approximation-Driven Selection {#sec-select}

## Dense Approximation Grid

```{r approx-grid}
# Grid defined in setup: c1_grid, c2_grid, N_target
full_grid <- expand.grid(c1 = c1_grid, c2 = c2_grid,
                          N = N_target,
                          stringsAsFactors = FALSE)
full_grid <- full_grid[full_grid$c2 <= full_grid$c1, ]

# Compute corrected FPR for every (c1, c2, N)
full_grid$n_H <- n_min_val
full_grid$d_eff <- d_eff_binary(full_grid$n_H, p_event)
full_grid$L_eff <- C_fit * (full_grid$N / n_min_val)^alpha_fit

full_grid$P1 <- mapply(function(d, c1v, c2v) {
  p_detect(1.0, d, c1v, c2v)
}, full_grid$d_eff, full_grid$c1, full_grid$c2)

full_grid$FPR_corr <- 1 - (1 - full_grid$P1)^full_grid$L_eff

full_dt <- data.table(full_grid)

cat(sprintf("Dense grid: %d c1 x %d c2 x %d N = %d configs\n",
    length(c1_grid), length(c2_grid), length(N_target),
    nrow(full_dt)))
cat(sprintf("(c1 range: %.2f--%.2f, c2 range: %.2f--%.2f)\n",
    min(c1_grid), max(c1_grid), min(c2_grid), max(c2_grid)))
```


## Corrected FPR Heatmaps

```{r approx-heatmaps, fig.width=10, fig.height=12}
#| fig-cap: "Corrected approximation FPR across the (c1,c2) grid.  Only the region near the 12% boundary is relevant for threshold selection."

# Subset to readable range (c1_heat, c2_heat from setup)
heat_sub <- full_dt[c1 %in% c1_heat & c2 %in% c2_heat]

heatmap_list <- lapply(N_target, function(N_val) {
  sub <- heat_sub[N == N_val]
  ggplot(sub, aes(x = factor(c1), y = factor(c2),
                  fill = FPR_corr)) +
    geom_tile(colour = "white", linewidth = 0.5) +
    geom_text(aes(label = sprintf("%.1f%%", 100 * FPR_corr)),
              size = 2.8, colour = "black") +
    scale_fill_gradient2(
      low = "#E8F5E9", mid = "#FFF9C4", high = "#E53935",
      midpoint = fpr_target,
      limits = c(0, max(heat_sub$FPR_corr, 0.30)),
      name = "Corr FPR"
    ) +
    labs(x = expression(c[1]), y = expression(c[2]),
         title = sprintf("N = %d (L_eff = %.1f)", N_val,
           C_fit * (N_val / n_min_val)^alpha_fit)) +
    theme_minimal(base_size = 11) +
    theme(plot.title = element_text(face = "bold"),
          axis.text = element_text(size = 9))
})

wrap_plots(heatmap_list, ncol = 1) +
  plot_annotation(
    title = "Corrected Approximation: FPR Under H0",
    subtitle = sprintf(
      "L_eff = %.3f * (N/%d)^%.3f | Target: FPR <= %.0f%%",
      C_fit, n_min_val, alpha_fit, 100 * fpr_target))
```


## Pre-Select: Smallest $(c_1, c_2)$ Meeting the Target

For each $(c_1, c_2)$, the corrected FPR must be at most
`r 100 * fpr_target`% at **every** target $N$.  Since
$L_{\text{eff}}$ grows with $N$, the binding constraint is at
the largest $N$ ($= `r max(N_target)`$).

We seek the **smallest $c_1$** that achieves the target,
and for each $c_1$, the **smallest $c_2$** that suffices.

```{r select-pairs}
#| tbl-cap: "Minimum c2 required for each c1 to achieve corrected FPR <= target at all N"

# For each (c1,c2), compute corrected FPR at the most
# stringent N (largest, where L_eff is highest)
N_max <- max(N_target)
worst_case <- full_dt[N == N_max, .(c1, c2, FPR_corr)]

# Also collect FPR at all N for display
fpr_wide <- full_dt[, .(
  FPR_corr_worst = FPR_corr[N == N_max]
), by = .(c1, c2)]

# Merge with per-N values
for (N_val in N_target) {
  col_name <- sprintf("FPR_N%d", N_val)
  fpr_wide[[col_name]] <- full_dt[N == N_val,
    FPR_corr[match(paste(fpr_wide$c1, fpr_wide$c2),
                   paste(c1, c2))]]
}

# Identify pairs meeting the target
selected <- fpr_wide[FPR_corr_worst <= fpr_target]
setorder(selected, c1, c2)

cat(sprintf(
  "Pairs with corrected FPR <= %.0f%% at N=%d: %d of %d\n",
  100 * fpr_target, N_max, nrow(selected), nrow(fpr_wide)))

# Frontier: for each c1, the minimum c2 that meets the target
frontier <- selected[, .(
  min_c2 = min(c2),
  FPR_at_min_c2 = FPR_corr_worst[which.min(c2)]
), by = c1]
setorder(frontier, c1)

if (nrow(frontier) > 0) {
  cat(sprintf("\nSmallest c1 achieving target: %.2f (with c2=%.2f)\n",
      frontier$c1[1], frontier$min_c2[1]))

  gt(frontier[, .(
    c1,
    `Min c2` = min_c2,
    `Corr FPR (worst N)` = sprintf("%.1f%%", 100 * FPR_at_min_c2)
  )]) |>
    tab_header(
      title = md(sprintf(
        "**Frontier: Minimum c2 for Each c1 (FPR <= %.0f%%)**",
        100 * fpr_target)),
      subtitle = sprintf(
        "Binding constraint at N=%d (L_eff=%.1f)",
        N_max, C_fit * (N_max / n_min_val)^alpha_fit)
    ) |>
    tab_style(style = cell_fill(color = "#E8F5E9"),
      locations = cells_body(rows = c1 == frontier$c1[1]))
} else {
  cat("No c1 values achieved the target.\n")
  cat("Consider relaxing fpr_target or increasing c1/c2 grid.\n")
}
```


## Optimal Pair: Power Curve

The frontier identifies the **least restrictive** thresholds
that still control FPR.  We extract $(c_{1,\min}, c_{2,\min})$
--- the smallest $c_1$ achieving the target, paired with its
minimum $c_2$ --- and plot the power curve.

::: {.callout-important}
## Why the $L_{\text{eff}}$ correction is asymmetric

The correction serves different roles under $H_0$ and $H_1$:

**Under $H_0$ (FPR):** every candidate subgroup is null,
each with per-subgroup probability $P_1(\theta = 1)$ of
falsely passing the screening and consistency criteria.
The procedure evaluates $\sim L_{\text{eff}}$ approximately
independent candidates, so the procedure-level FPR is

$$
\text{FPR}_{\text{corr}} = 1 - (1 - P_1)^{L_{\text{eff}}}
\;\gg\; P_1
$$

The correction is essential here --- without it, the
per-subgroup rate understates the procedure-level rate by
a factor of $L_{\text{eff}}$.

**Under $H_1$ (power):** there is one true subgroup with
effect $\theta^\dagger(H)$.  Its detection probability is
$P_1(\theta^\dagger)$.  The remaining $\sim L_{\text{eff}}$
candidates are null and each has probability
$P_1(1) \approx$ `r sprintf("%.1f%%", 100 * p_detect(1.0,
round(d_eff_binary(n_min_val, p_event)), 1.25, 1.0))`
of falsely passing --- negligible once
thresholds have been selected to control FPR.  Therefore:

$$
\text{Power} \;\approx\; P_1(\theta^\dagger)
\qquad \text{(uncorrected)}
$$

The **corrected** power
$1 - (1 - P_1(\theta))^{L_{\text{eff}}}$ slightly overstates
true power because it counts detection of *any* subgroup
(including false ones) as a success.  The **uncorrected**
$P_1(\theta)$ is the probability of detecting the *correct*
subgroup and is the primary power measure.

Both curves are shown below.  The per-subgroup $P_1$ (dashed)
is the power to detect the true subgroup; the corrected curve
(solid) is the power to detect *something*.  C1's simulations
confirm that $P_1$ is accurate under $H_1$ (Sim/Approx
$\approx 1.0$).
:::

When `nsims_power > 0`, a targeted $H_1$ simulation is run at
$\theta = `r theta_h1`$ for the optimal pair to spot-check
$P_1$ accuracy at the effect size of interest.

```{r optimal-pair-power}
#| fig-height: 7
#| fig-cap: !expr 'if (exists("frontier") && nrow(frontier) > 0) sprintf("Power curve for optimal pair (c1=%.2f, c2=%.2f). Solid blue = corrected (any detection), dashed grey = per-subgroup P1 (correct subgroup), red diamonds = simulation.", frontier$c1[1], frontier$min_c2[1]) else "No frontier found."'

if (exists("frontier") && nrow(frontier) > 0) {

  c1_opt <- frontier$c1[1]
  c2_opt <- frontier$min_c2[1]

  cat(sprintf("=== Optimal Pair: c1 = %.2f, c2 = %.2f ===\n",
              c1_opt, c2_opt))
  cat(sprintf("Corrected FPR at this pair: %.1f%% (target: %.0f%%)\n",
              100 * frontier$FPR_at_min_c2[1], 100 * fpr_target))

  # ── Approximation power curve ──────────────────────────────
  d_ref <- round(d_eff_binary(n_min_val, p_event))
  or_seq <- seq(0.5, 4.0, by = 0.05)

  power_curves <- do.call(rbind, lapply(N_target, function(N_val) {
    L_eff <- C_fit * (N_val / n_min_val)^alpha_fit
    data.frame(
      OR = rep(or_seq, 2),
      Power = c(
        sapply(or_seq, function(or) {
          P1 <- p_detect(or, d_ref, c1_opt, c2_opt)
          1 - (1 - P1)^L_eff
        }),
        sapply(or_seq, function(or) {
          p_detect(or, d_ref, c1_opt, c2_opt)
        })
      ),
      Type = rep(c("Corrected (any detection)",
                   "Per-subgroup P1 (correct subgroup)"),
                 each = length(or_seq)),
      N = sprintf("N = %d (L_eff = %.2f)", N_val, L_eff)
    )
  }))

  # ── Targeted H1 simulation at theta_h1 ─────────────────────
  sim_power_points <- NULL
  if (nsims_power > 0) {
    cat(sprintf("\n--- Running %d H1 sims at OR = %.1f ---\n",
                nsims_power, theta_h1))
    t0_power <- proc.time()[3]

    sim_power_results <- do.call(rbind, lapply(N_target, function(N_val) {
      found <- foreach(
        s = seq_len(nsims_power),
        .combine = c, .errorhandling = "remove",
        .options.future = list(
          packages = c("forestsearch", "data.table"),
          seed = TRUE)
      ) %dofuture% {
        future::plan("sequential")
        seed <- sim_seed + 100000L + (N_val * 1000L) + s
        df_h1 <- sim_binary_h1(N_val, seed)
        fs <- tryCatch(
          suppressWarnings(
            forestsearch(
              df.analysis = df_h1,
              confounders.name = confounders_name,
              outcome.name = "progressed", event.name = "progressed",
              treat.name = "treat", id.name = "id",
              outcome_type = "binary", effect_measure = "OR",
              adverse_outcome = TRUE,
              hr.threshold = c1_opt,
              hr.consistency = c2_opt,
              pconsistency.threshold = pcons, fs.splits = fs_splits,
              n.min = n_min_val, maxk = maxk_val,
              use_lasso = TRUE, use_grf = TRUE,
              use_twostage = TRUE,
              is.RCT = TRUE, seedit = seed,
              details = FALSE, quiet = TRUE, plot.sg = FALSE,
              parallel_args = list(plan = "sequential", workers = 1,
                                   show_message = FALSE)
            )
          ),
          error = function(e) NULL
        )
        found <- !is.null(fs) && !is.null(fs$sg.harm) &&
                 length(fs$sg.harm) > 0
        as.integer(found)
      }
      res_clean <- unlist(found[sapply(found, is.numeric)])
      if (length(res_clean) < nsims_power) {
        warning(sprintf(
          "H1 N=%d: %d of %d tasks silently failed.",
          N_val, nsims_power - length(res_clean), nsims_power
        ))
      }
      sim_pwr <- mean(res_clean, na.rm = TRUE)
      L_eff <- C_fit * (N_val / n_min_val)^alpha_fit
      cat(sprintf("  N = %4d: Sim power = %.1f%% (%d/%d) | L_eff = %.2f\n",
                  N_val, 100 * sim_pwr, sum(res_clean),
                  length(res_clean), L_eff))
      data.frame(
        OR = theta_h1, Power = sim_pwr,
        Type = "Simulation",
        N = sprintf("N = %d (L_eff = %.2f)", N_val, L_eff))
    }))

    sim_power_points <- sim_power_results
    cat(sprintf("Power simulation: %.1f min\n",
                (proc.time()[3] - t0_power) / 60))
  }

  # ── Plot ────────────────────────────────────────────────────
  p <- ggplot(power_curves, aes(x = OR, y = Power,
                                 linetype = Type, colour = Type)) +
    geom_line(linewidth = 0.9) +
    geom_hline(yintercept = c(fpr_target, 0.80),
               linetype = "dashed", colour = "grey50", linewidth = 0.4) +
    geom_vline(xintercept = 1.0, linetype = "dotted", colour = "grey70")

  if (!is.null(sim_power_points)) {
    p <- p +
      geom_point(data = sim_power_points,
                 aes(x = OR, y = Power),
                 colour = "red", size = 3, shape = 18,
                 inherit.aes = FALSE) +
      geom_text(data = sim_power_points,
                aes(x = OR, y = Power,
                    label = sprintf("%.0f%%", 100 * Power)),
                colour = "red", vjust = -1, size = 3,
                inherit.aes = FALSE)
  }

  p <- p +
    facet_wrap(~ N, ncol = 1) +
    scale_linetype_manual(values = c(
      "Corrected (any detection)" = "solid",
      "Per-subgroup P1 (correct subgroup)" = "dashed")) +
    scale_colour_manual(values = c(
      "Corrected (any detection)" = "steelblue",
      "Per-subgroup P1 (correct subgroup)" = "grey50")) +
    labs(x = expression(theta^dagger*(H) ~ "(true OR in subgroup)"),
         y = "Detection Probability",
         title = sprintf("Optimal Pair: c1 = %.2f, c2 = %.2f",
                          c1_opt, c2_opt),
         subtitle = sprintf(
           "d_eff = %d (n_H = %d, p = %.2f) | FPR target = %.0f%%%s",
           d_ref, n_min_val, p_event, 100 * fpr_target,
           if (nsims_power > 0)
             sprintf(" | Red diamonds: %d H1 sims at OR=%.1f",
                     nsims_power, theta_h1) else ""),
         linetype = NULL, colour = NULL) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top",
          plot.title = element_text(face = "bold"))
  print(p)

  # ── Power summary table ────────────────────────────────────
  cat("\n=== Power at Key Effect Sizes ===\n")
  header <- sprintf("%-8s", "OR")
  for (N_val in N_target) {
    header <- paste0(header, sprintf("  N=%-4d (P1)  (corr)", N_val))
    if (nsims_power > 0) header <- paste0(header, "   (sim)")
  }
  cat(header, "\n")

  for (or_val in c(1.0, 1.25, 1.50, theta_h1, 2.5, 3.0)) {
    cat(sprintf("%-8.2f", or_val))
    for (N_val in N_target) {
      L_eff <- C_fit * (N_val / n_min_val)^alpha_fit
      P1 <- p_detect(or_val, d_ref, c1_opt, c2_opt)
      P_corr <- 1 - (1 - P1)^L_eff
      cat(sprintf("  %5.1f%%  %5.1f%%", 100 * P1, 100 * P_corr))
      if (nsims_power > 0) {
        sp <- sim_power_points[sim_power_points$N ==
          sprintf("N = %d (L_eff = %.2f)", N_val, L_eff), ]
        if (nrow(sp) > 0 && abs(or_val - theta_h1) < 0.01) {
          cat(sprintf("  %5.1f%%", 100 * sp$Power[1]))
        } else {
          cat("      -")
        }
      }
    }
    cat("\n")
  }

} else {
  cat("No frontier found. Cannot compute optimal pair power curve.\n")
  cat("Consider expanding c1_grid/c2_grid or relaxing fpr_target.\n")
}
```

```{r optimal-pair-table}
#| tbl-cap: !expr 'if (exists("c1_opt")) sprintf("Power at optimal pair (c1 = %.2f, c2 = %.2f) by sample size and effect size", c1_opt, c2_opt) else "Power summary"'

if (exists("c1_opt")) {

  or_vals <- c(1.0, 1.25, 1.50, theta_h1, 2.5, 3.0)
  or_vals <- unique(sort(or_vals))

  power_tbl <- do.call(rbind, lapply(or_vals, function(or_val) {
    row <- data.frame(OR = or_val, stringsAsFactors = FALSE)
    for (N_val in N_target) {
      L_eff <- C_fit * (N_val / n_min_val)^alpha_fit
      P1 <- p_detect(or_val, d_ref, c1_opt, c2_opt)
      P_corr <- 1 - (1 - P1)^L_eff
      row[[sprintf("P1_N%d", N_val)]]   <- P1
      row[[sprintf("Corr_N%d", N_val)]] <- P_corr

      # Add simulation column if available
      if (!is.null(sim_power_points) &&
          abs(or_val - theta_h1) < 0.01) {
        sp <- sim_power_points[
          sim_power_points$N == sprintf("N = %d (L_eff = %.2f)",
                                         N_val, L_eff), ]
        row[[sprintf("Sim_N%d", N_val)]] <- if (nrow(sp) > 0)
          sp$Power[1] else NA_real_
      } else {
        row[[sprintf("Sim_N%d", N_val)]] <- NA_real_
      }
    }
    row
  }))

  # Build gt table with spanners per N
  tbl <- gt(power_tbl) |>
    fmt_percent(columns = -OR, decimals = 1) |>
    sub_missing(missing_text = "—") |>
    fmt_number(columns = OR, decimals = 2)

  for (N_val in N_target) {
    tbl <- tbl |>
      tab_spanner(
        label = sprintf("N = %d", N_val),
        columns = starts_with(c(sprintf("P1_N%d", N_val),
                                sprintf("Corr_N%d", N_val),
                                sprintf("Sim_N%d", N_val)))
      ) |>
      cols_label(
        !!sprintf("P1_N%d", N_val)   := "P1",
        !!sprintf("Corr_N%d", N_val) := "Corrected",
        !!sprintf("Sim_N%d", N_val)  := "Simulation"
      )
  }

  # Remove Simulation columns if no sims were run
  if (is.null(sim_power_points)) {
    sim_cols <- sprintf("Sim_N%d", N_target)
    tbl <- tbl |> cols_hide(columns = any_of(sim_cols))
  }

  tbl <- tbl |>
    tab_header(
      title = sprintf("Power: Optimal Pair (c1 = %.2f, c2 = %.2f)",
                      c1_opt, c2_opt),
      subtitle = sprintf(
        "d_eff = %d (n_H = %d, p = %.2f) | P1 = correct subgroup, Corrected = any detection%s",
        d_ref, n_min_val, p_event,
        if (!is.null(sim_power_points))
          sprintf(" | Sim: %d sims at OR = %.1f", nsims_power, theta_h1)
        else "")
    ) |>
    tab_style(
      style = cell_fill(color = "#E8F5E9"),
      locations = cells_body(
        columns = starts_with("P1_"),
        rows = OR == theta_h1)
    ) |>
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_body(rows = OR == theta_h1)
    ) |>
    tab_source_note(sprintf(
      "P1 = per-subgroup detection (primary power measure) | Corrected = 1-(1-P1)^L_eff (overstates; see §5.4 callout) | OR = %.1f row highlighted",
      theta_h1))

  tbl
}
```

## Full Selected Pairs Table

```{r selected-table}
#| tbl-cap: "All pairs meeting the corrected FPR target, sorted by c1 then c2 (smallest first)"

if (nrow(selected) > 0) {
  display_cols <- c("c1", "c2", "FPR_corr_worst")
  for (N_val in N_target) {
    display_cols <- c(display_cols, sprintf("FPR_N%d", N_val))
  }

  display_dt <- selected[, ..display_cols]
  names(display_dt)[3] <- "Worst Case"
  for (N_val in N_target) {
    old_name <- sprintf("FPR_N%d", N_val)
    new_name <- sprintf("N=%d", N_val)
    names(display_dt)[names(display_dt) == old_name] <- new_name
  }

  # Format as percentages
  for (col in names(display_dt)[3:ncol(display_dt)]) {
    display_dt[[col]] <- sprintf("%.1f%%", 100 * display_dt[[col]])
  }

  gt(display_dt) |>
    tab_header(
      title = md(sprintf(
        "**All Pairs with Corrected FPR <= %.0f%%**",
        100 * fpr_target)),
      subtitle = "Sorted by c1 (ascending), then c2 (ascending)"
    ) |>
    tab_spanner(
      label = "Corrected FPR by N",
      columns = 4:ncol(display_dt)
    ) |>
    tab_source_note(
      "Worst Case = max corrected FPR across N values"
    )
} else {
  cat("No pairs met the target.\n")
}
```


## Boundary Pairs (Near the Target)

Include pairs slightly above the target to map the transition
and provide simulation candidates.

```{r boundary-pairs}
boundary <- fpr_wide[FPR_corr_worst > fpr_target &
                      FPR_corr_worst <= fpr_target + 0.05]

sim_candidates <- rbind(
  selected[, .(c1, c2, FPR_corr_worst, region = "selected")],
  boundary[, .(c1, c2, FPR_corr_worst, region = "boundary")]
)
setorder(sim_candidates, c1, c2)

# Thin to manageable count (max_pairs from setup)
if (nrow(sim_candidates) > max_pairs) {
  n_sel <- sum(sim_candidates$region == "selected")
  n_boundary_keep <- max(max_pairs - n_sel, 5L)
  sim_candidates <- rbind(
    sim_candidates[region == "selected"],
    sim_candidates[region == "boundary"][
      seq(1, .N, length.out = min(.N, n_boundary_keep))]
  )
}

cat(sprintf(
  "Simulation candidates: %d selected + %d boundary = %d total\n",
  sum(sim_candidates$region == "selected"),
  sum(sim_candidates$region == "boundary"),
  nrow(sim_candidates)))
```


# Part III: Simulation Validation {#sec-sim}

Set `nsims_val = 0` in the setup chunk to skip validation
and rely entirely on the corrected approximation.  Increase
`nsims_val` (e.g., 500) for production-quality validation.

```{r run-validation}
run_validation <- nsims_val > 0

if (run_validation) {
  t0 <- proc.time()[3]

  # Only validate the frontier pairs: the minimum c2 for each c1
  # that achieves FPR_corr <= target.  At most 10 pairs.
  if (exists("frontier") && nrow(frontier) > 0) {
    val_candidates <- frontier[seq_len(min(10, nrow(frontier))),
                                .(c1, c2 = min_c2, FPR_corr = FPR_at_min_c2)]
  } else {
    val_candidates <- sim_candidates[region == "selected",
                                      .(c1, c2, FPR_corr = FPR_corr_worst)]
    if (nrow(val_candidates) > 10)
      val_candidates <- val_candidates[seq_len(10)]
  }

  cat(sprintf(
    "Validating %d frontier pairs (min c2 per c1, FPR_corr <= %.0f%%)\n",
    nrow(val_candidates), 100 * fpr_target))
  for (r in seq_len(nrow(val_candidates))) {
    cat(sprintf("  c1=%.2f, c2=%.2f (corr FPR=%.1f%%)\n",
                val_candidates$c1[r], val_candidates$c2[r],
                100 * val_candidates$FPR_corr[r]))
  }

  # Build task list
  val_tasks <- do.call(rbind, lapply(N_target, function(N_val) {
    do.call(rbind, lapply(seq_len(nrow(val_candidates)), function(g) {
      data.frame(
        N = N_val,
        sim = seq_len(nsims_val),
        c1 = val_candidates$c1[g],
        c2 = val_candidates$c2[g],
        stringsAsFactors = FALSE
      )
    }))
  }))

  cat(sprintf(
    "Running %d validation tasks (%d pairs x %d N x %d sims)...\n",
    nrow(val_tasks), nrow(val_candidates),
    length(N_target), nsims_val))

  val_results <- foreach(
    i = seq_len(nrow(val_tasks)),
    .combine = rbind, .errorhandling = "remove",
    .options.future = list(
      packages = c("forestsearch", "data.table"), seed = TRUE)
  ) %dofuture% {
    future::plan("sequential")
    N_val  <- val_tasks$N[i]
    sim_id <- val_tasks$sim[i]
    c1_val <- val_tasks$c1[i]
    c2_val <- val_tasks$c2[i]

    seed <- sim_seed + 200000L + (N_val * 1000L) + sim_id
    df_null <- sim_binary_null(N_val, seed)

    fs <- tryCatch(
      suppressWarnings(
        forestsearch(
          df.analysis = df_null,
          confounders.name = confounders_name,
          outcome.name = "progressed", event.name = "progressed",
          treat.name = "treat", id.name = "id",
          outcome_type = "binary", effect_measure = "OR",
          adverse_outcome = TRUE,
          hr.threshold = c1_val, hr.consistency = c2_val,
          pconsistency.threshold = pcons, fs.splits = fs_splits,
          n.min = n_min_val, maxk = maxk_val,
          use_lasso = TRUE, use_grf = TRUE,
          use_twostage = TRUE,
          is.RCT = TRUE, seedit = seed,
          details = FALSE, quiet = TRUE, plot.sg = FALSE,
          parallel_args = list(plan = "sequential", workers = 1,
                               show_message = FALSE)
        )
      ),
      error = function(e) NULL
    )

    found <- !is.null(fs) && !is.null(fs$sg.harm) &&
             length(fs$sg.harm) > 0
    data.table(N = N_val, sim = sim_id,
               c1 = c1_val, c2 = c2_val,
               any_found = as.integer(found))
  }

  # ── Post-hoc task-count check ──────────────────────────────────
  n_expected_val <- nrow(val_tasks)
  n_received_val <- nrow(val_results)
  if (n_received_val < n_expected_val) {
    warning(sprintf(
      "Validation: %d of %d tasks silently failed (%.1f%% loss).",
      n_expected_val - n_received_val, n_expected_val,
      100 * (1 - n_received_val / n_expected_val)
    ))
  }

  timings$validation <- proc.time()[3] - t0
  cat(sprintf("Validation: %.1f min | %d/%d results\n",
      timings$validation / 60, n_received_val, n_expected_val))

} else {
  timings$validation <- 0
  val_results <- NULL
  cat("Validation skipped (nsims_val = 0).\n")
  cat("Threshold selection based on corrected approximation only.\n")
}
```


# Part IV: Results {#sec-results}

```{r compute-fpr}
if (run_validation) {
  val_dt <- data.table(val_results)
  val_fpr <- val_dt[, .(
    Sim_FPR = mean(any_found, na.rm = TRUE),
    n_found = sum(any_found, na.rm = TRUE),
    n_sims = .N
  ), by = .(N, c1, c2)]

  # Merge with approximation
  val_fpr[, n_H := n_min_val]
  val_fpr[, d_eff := d_eff_binary(n_H, p_event)]
  val_fpr[, P1 := mapply(function(d, c1v, c2v) {
    p_detect(1.0, d, c1v, c2v)
  }, d_eff, c1, c2)]
  val_fpr[, L_eff_N := C_fit * (N / n_min_val)^alpha_fit]
  val_fpr[, FPR_corr := 1 - (1 - P1)^L_eff_N]

  # Add region label
  val_fpr <- merge(val_fpr,
    sim_candidates[, .(c1, c2, region)],
    by = c("c1", "c2"), all.x = TRUE)

} else {
  val_dt <- NULL
  val_fpr <- NULL
  cat("No simulation results (nsims_val = 0).\n")
  cat("Results below are based on the corrected approximation only.\n")
}
```


## Validation Table

```{r val-table}
#| tbl-cap: "Simulated FPR vs corrected approximation for pre-selected pairs."

if (run_validation) {
  # CI under H0: FPR_corr is the truth
  val_fpr[, CI_lower := pmax(0, FPR_corr - 1.96 * sqrt(
    FPR_corr * (1 - FPR_corr) / n_sims))]
  val_fpr[, CI_upper := pmin(1, FPR_corr + 1.96 * sqrt(
    FPR_corr * (1 - FPR_corr) / n_sims))]
  val_fpr[, in_CI := Sim_FPR >= CI_lower & Sim_FPR <= CI_upper]

  gt(val_fpr[, .(N, c1, c2,
    `Corr FPR` = sprintf("%.1f%%", 100 * FPR_corr),
    `95% CI` = sprintf("[%.1f%%, %.1f%%]",
      100 * CI_lower, 100 * CI_upper),
    `Sim FPR` = sprintf("%.1f%% (%d/%d)",
      100 * Sim_FPR, n_found, n_sims),
    `Sim/Corr` = round(Sim_FPR / pmax(FPR_corr, 0.001), 2),
    `In CI` = ifelse(in_CI, "Yes", "No")
  )]) |>
    tab_header(
      title = md("**Simulation Validation of Selected Pairs**"),
      subtitle = sprintf(
        "%d sims/config | L_eff = %.3f * (N/%d)^%.3f",
        nsims_val, C_fit, n_min_val, alpha_fit)
    ) |>
    tab_row_group(label = "N = 700", rows = N == 700) |>
    tab_row_group(label = "N = 500", rows = N == 500) |>
    tab_style(style = cell_fill(color = "#E8F5E9"),
      locations = cells_body(columns = `In CI`,
        rows = `In CI` == "Yes")) |>
    tab_style(style = cell_fill(color = "#FFEBEE"),
      locations = cells_body(columns = `In CI`,
        rows = `In CI` == "No")) |>
    tab_source_note(paste(
      "95% CI: expected range of Sim FPR if Corr FPR is the true rate.",
      sprintf("Green = Sim FPR within CI; %d sims per config.",
              nsims_val)))
} else {
  cat("Validation table skipped (nsims_val = 0).\n")
}
```


## Validation Scatter

```{r val-scatter, fig.width=9, fig.height=6}
#| fig-cap: "Corrected approximation (x) vs simulated FPR (y).  Points near the diagonal confirm the approximation.  Dotted lines mark the FPR target."

if (run_validation) {
  print(
    ggplot(val_fpr, aes(x = FPR_corr, y = Sim_FPR)) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed",
                  colour = "grey50") +
      geom_point(aes(colour = region, shape = region), size = 3,
                 alpha = 0.7) +
      geom_hline(yintercept = fpr_target, linetype = "dotted",
                 colour = "firebrick") +
      geom_vline(xintercept = fpr_target, linetype = "dotted",
                 colour = "firebrick") +
      scale_colour_manual(
        values = c("selected" = "steelblue",
                   "boundary" = "orange")) +
      facet_wrap(~ paste("N =", N)) +
      coord_equal(
        xlim = c(0, max(val_fpr$FPR_corr, 0.20)),
        ylim = c(0, max(val_fpr$Sim_FPR, 0.20))) +
      labs(x = "Corrected Approximation",
           y = "Simulated FPR",
           title = "Validation: Corrected Approximation vs Simulation",
           colour = "Region", shape = "Region") +
      theme_minimal(base_size = 12) +
      theme(legend.position = "top",
            plot.title = element_text(face = "bold"))
  )
} else {
  cat("Scatter plot skipped (nsims_val = 0).\n")
}
```


## Confirmed Acceptable Pairs

```{r confirmed-pairs}
#| tbl-cap: "Confirmed acceptable pairs"

if (run_validation) {
  # Pairs where both approximation AND simulation agree
  confirmed <- val_fpr[Sim_FPR <= fpr_target &
                        FPR_corr <= fpr_target]
  setorder(confirmed, N, -Sim_FPR)

  if (nrow(confirmed) > 0) {
    gt(confirmed[, .(N, c1, c2,
      `Corr FPR` = sprintf("%.1f%%", 100 * FPR_corr),
      `95% CI` = sprintf("[%.1f%%, %.1f%%]",
        100 * CI_lower, 100 * CI_upper),
      `Sim FPR` = sprintf("%.1f%%", 100 * Sim_FPR),
      `Sim/Corr` = round(Sim_FPR / pmax(FPR_corr, 0.001), 2),
      `In CI` = ifelse(in_CI, "Yes", "No")
    )]) |>
      tab_header(
        title = md(sprintf(
          "**Confirmed Pairs: FPR <= %.0f%% (Approx AND Sim)**",
          100 * fpr_target)),
        subtitle = sprintf("%d sims per config", nsims_val)
      ) |>
      tab_row_group(label = "N = 700", rows = N == 700) |>
      tab_row_group(label = "N = 500", rows = N == 500) |>
      tab_style(style = cell_fill(color = "#E8F5E9"),
        locations = cells_body(columns = `In CI`,
          rows = `In CI` == "Yes")) |>
      tab_style(style = cell_fill(color = "#FFEBEE"),
        locations = cells_body(columns = `In CI`,
          rows = `In CI` == "No")) |>
      tab_source_note(paste(
        "95% CI bounds treat Corr FPR as the true rate.",
        "'In CI' = simulated FPR falls within expected range."))
  } else {
    cat("No pairs confirmed at this sim count.\n")
    cat("Consider increasing nsims_val.\n")
  }

} else {
  # No simulation: "confirmed" = approximation-selected only
  confirmed <- selected
  confirmed[, N := NA_integer_]

  cat("No simulation validation (nsims_val = 0).\n")
  cat(sprintf(
    "Approximation-selected pairs (corrected FPR <= %.0f%%): %d\n",
    100 * fpr_target, nrow(confirmed)))
  cat("Re-render with nsims_val > 0 to validate by simulation.\n")
}
```


# Part V: Save Results {#sec-save}

```{r save-results}
results_bundle <- list(
  # Parameters
  params = list(
    nsims_val = nsims_val, sim_seed = sim_seed,
    n_workers = n_workers,
    C_fit = C_fit, alpha_fit = alpha_fit,
    recalibrate = recalibrate,
    n_sims_recal = n_sims_recal,
    fpr_target = fpr_target,
    n_min_val = n_min_val, p_event = p_event,
    p_harm = p_harm, pcons = pcons,
    N_target = N_target,
    c1_grid = c1_grid, c2_grid = c2_grid,
    timestamp = Sys.time()
  ),

  # L_eff calibration object and raw data
  # (If recalibrate = FALSE, stores prior values only.
  #  If TRUE, stores the full calibration data for
  #  reproducing the fit.)
  leff_calibration = if (exists("leff_cal")) leff_cal else
    list(C = C_fit, alpha = alpha_fit,
         n_min = n_min_val, source = "prior"),
  leff_raw_data = if (exists("cal_df")) cal_df else NULL,

  # Approximation grid (all pairs, all N)
  approx_grid = full_dt,

  # Pre-selected candidates
  sim_candidates = sim_candidates,

  # Raw simulation results (per-replicate)
  val_results_raw = val_dt,

  # Aggregated FPR by (N, c1, c2)
  val_fpr = val_fpr,

  # Confirmed acceptable pairs
  confirmed = if (exists("confirmed")) confirmed else NULL,

  # Timings
  timings = timings
)

if (!is.null(output_dir)) {
  if (!dir.exists(output_dir)) dir.create(output_dir,
                                           recursive = TRUE)
  rds_path <- file.path(output_dir,
    sprintf("binary_null_search_%s.rds",
            format(Sys.Date(), "%Y%m%d")))
  saveRDS(results_bundle, rds_path)
  cat(sprintf("Results saved to: %s\n", rds_path))
  cat(sprintf("  Size: %.1f KB\n", file.size(rds_path) / 1024))
} else {
  cat("Saving skipped (output_dir = NULL)\n")
}
```


## How to Reload and Extend

```{r reload-example}
#| eval: false
#| code-fold: false

# Load the most recent results bundle
res <- readRDS(latest_rds("selection_binary_frontier"))

# Access components
res$params$C_fit           # L_eff parameters
res$params$alpha_fit
res$leff_calibration       # Full calibration object (or prior)
res$leff_raw_data          # Raw calibration data (if recalibrated)
res$approx_grid            # Full approximation grid
res$val_fpr                # Simulation results by (N, c1, c2)
res$confirmed              # Confirmed pairs

# Re-run with more sims: change nsims_val in setup and re-render.
# The approximation selection is instant; only validation re-runs.

# Use saved calibration as prior for a new scenario:
# C_prior     <- res$params$C_fit
# alpha_prior <- res$params$alpha_fit
# Then modify sim_binary_null() and forestsearch() call for
# the new DGM, set recalibrate = TRUE.
```


# Summary {#sec-summary}

```{r summary}
cat("=== Approximation-Driven Null Threshold Search ===\n\n")
cat(sprintf("DGM: Binary, p=%.2f, 4 confounders\n", p_event))
cat(sprintf("n.min=%d, pcons=%.2f, maxk=2\n", n_min_val, pcons))
cat(sprintf("L_eff = %.3f * (N/%d)^%.3f",
            C_fit, n_min_val, alpha_fit))
if (!recalibrate) cat("  [from prior calibration]")
cat("\n\n")

cat(sprintf("Dense grid: %d c1 x %d c2 values\n",
            length(c1_grid), length(c2_grid)))
cat(sprintf("Target: corrected FPR <= %.0f%% at all N in {%s}\n\n",
            100 * fpr_target,
            paste(N_target, collapse = ", ")))

# Frontier summary
if (exists("frontier") && nrow(frontier) > 0) {
  c1_opt <- frontier$c1[1]
  c2_opt <- frontier$min_c2[1]
  d_ref  <- round(d_eff_binary(n_min_val, p_event))

  cat("--- Optimal Pair: (c1_min, c2_min) ---\n")
  cat(sprintf("  c1_min = %.2f, c2_min = %.2f\n",
              c1_opt, c2_opt))
  cat(sprintf("  Corrected FPR: %.1f%% (target: %.0f%%)\n",
              100 * frontier$FPR_at_min_c2[1], 100 * fpr_target))
  cat(sprintf("  Per-subgroup P1(OR=1): %.1f%%\n",
              100 * p_detect(1.0, d_ref, c1_opt, c2_opt)))
  cat(sprintf("  Approx power at OR=%.1f: %.1f%%\n\n",
              theta_h1,
              100 * p_detect(theta_h1, d_ref, c1_opt, c2_opt)))

  cat("--- Full Frontier ---\n")
  for (r in seq_len(min(10, nrow(frontier)))) {
    cat(sprintf("  c1=%.2f: min c2=%.2f (corr FPR=%.1f%%)\n",
                frontier$c1[r], frontier$min_c2[r],
                100 * frontier$FPR_at_min_c2[r]))
  }
  if (nrow(frontier) > 10)
    cat(sprintf("  ... and %d more c1 values\n", nrow(frontier) - 10))
  cat("\n")
}

cat(sprintf("Approximation selected: %d pairs\n",
            sum(sim_candidates$region == "selected")))
cat(sprintf("Boundary (within 5%%):  %d pairs\n",
            sum(sim_candidates$region == "boundary")))

if (run_validation) {
  cat(sprintf("Validated with %d sims per config\n\n",
              nsims_val))

  n_confirmed <- if (exists("confirmed") &&
    !is.null(confirmed)) nrow(confirmed) else 0
  cat(sprintf(
    "Confirmed (approx + sim <= %.0f%%): %d configs\n\n",
    100 * fpr_target, n_confirmed))

  if (n_confirmed > 0) {
    cat("--- Confirmed pairs (largest Sim FPR first) ---\n")
    for (N_val in N_target) {
      sub <- confirmed[N == N_val]
      if (nrow(sub) > 0) {
        cat(sprintf("  N = %d:\n", N_val))
        setorder(sub, -Sim_FPR)
        for (r in seq_len(min(10, nrow(sub)))) {
          cat(sprintf(
            "    c1=%.2f, c2=%.2f: corr=%.1f%%, sim=%.1f%%\n",
            sub$c1[r], sub$c2[r],
            100 * sub$FPR_corr[r],
            100 * sub$Sim_FPR[r]))
        }
        if (nrow(sub) > 10)
          cat(sprintf("    ... and %d more\n", nrow(sub) - 10))
      }
    }
  }
} else {
  cat("Validation: SKIPPED (nsims_val = 0)\n\n")

  cat("--- Approximation-selected pairs ---\n")
  cat(sprintf("  %d pairs with corrected FPR <= %.0f%%\n",
              nrow(selected), 100 * fpr_target))
  if (nrow(selected) > 0) {
    setorder(selected, c1, c2)
    for (r in seq_len(min(15, nrow(selected)))) {
      cat(sprintf("    c1=%.2f, c2=%.2f: corr FPR=%.1f%%\n",
                  selected$c1[r], selected$c2[r],
                  100 * selected$FPR_corr_worst[r]))
    }
    if (nrow(selected) > 15)
      cat(sprintf("    ... and %d more\n", nrow(selected) - 15))
  }
}

cat("\n--- Next step ---\n")
if (!run_validation) {
  cat("Set nsims_val > 0 (e.g., 30 for quick check, 500 for\n")
  cat("production) and re-render to validate by simulation.\n")
} else {
  cat("Increase nsims_val (e.g., 500) and re-render to\n")
  cat("tighten the simulation confidence intervals.\n")
}
cat("The approximation selection is instant; only\n")
cat("validation re-runs.\n")
```


# Timing and Session Info {#sec-session}

```{r timing-summary}
timings$total <- (proc.time() - t_vignette_start)["elapsed"]

stages <- c("L_eff Calibration")
seconds <- c(timings$leff_cal)
if (run_validation) {
  stages <- c(stages, "Validation Simulation")
  seconds <- c(seconds, timings$validation)
}
stages <- c(stages, "Total")
seconds <- c(seconds, timings$total)

timing_df <- data.frame(
  Stage = stages, Seconds = seconds,
  stringsAsFactors = FALSE
)
timing_df$Minutes <- timing_df$Seconds / 60

subtitle_text <- if (run_validation) {
  sprintf("%d pairs x %d N x %d sims = %d tasks | %d workers",
    nrow(sim_candidates), length(N_target),
    nsims_val,
    nrow(sim_candidates) * length(N_target) * nsims_val,
    n_workers)
} else {
  sprintf("Approximation only (no simulation) | %d workers",
          n_workers)
}

gt(timing_df) |>
  tab_header(
    title = "Computational Timing Summary",
    subtitle = subtitle_text
  ) |>
  fmt_number(columns = Seconds, decimals = 1) |>
  fmt_number(columns = Minutes, decimals = 2) |>
  cols_label(Stage = "Stage", Seconds = "Time (sec)",
             Minutes = "Time (min)") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Stage == "Total")
  )
```

```{r session-info}
plan(sequential)
sessionInfo()
```
