---
title: "Threshold Selection Guide for ForestSearch"
subtitle: "Using `suggest_thresholds()` to Choose (c1, c2) for Non-Survival Outcomes"
author: "Larry F. Leon"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: false
  message: false
---


# Background {#sec-background}

## The Problem

The paper defaults for ForestSearch (c1 = 1.25, c2 = 1.0)
were calibrated for **survival endpoints** using the GBSG DGM
(Leon et al., 2024).  For non-survival outcomes (binary,
count, continuous), these defaults may not provide adequate
FPR control because:

1. **Lower $d_{\text{eff}}$**: Binary outcomes with
   $\bar{p} = 0.30$ give $d_{\text{eff}} = 0.21 \cdot n_H$,
   roughly one-third the survival value
   ($d = 0.54 \cdot n_H$ at 46% censoring).  Lower
   $d_{\text{eff}}$ means higher per-subgroup $P_1$.

2. **Higher $L_{\text{eff}}$**: Without an events-based
   minimum (`d0.min`, `d1.min`), more candidate subgroups
   survive to the threshold check, so the procedure-level
   FPR is amplified relative to the per-subgroup $P_1$.

## The Solution

`suggest_thresholds()` uses the **corrected approximation**
to find the smallest $(c_1, c_2)$ that controls the
procedure-level FPR:

$$
\text{FPR}_{\text{corr}} = 1 - (1 - P_1)^{L_{\text{eff}}(N)}
\le \text{target}
$$

where:

- $P_1 = P(W_1 + W_2 \ge 2 \log c_1,\;
  \min(W_1, W_2) \ge \log c_2)$
  with $W_1, W_2 \sim N(0, 8/d_{\text{eff}})$ under $H_0$
- $L_{\text{eff}}(N) = C \cdot (N / n_{\min})^\alpha$ is
  the effective number of candidates

The function returns a **frontier table**: for each $c_1$
(ascending), the minimum $c_2$ that achieves the target.
The first row is the answer to "what is the smallest $c_1$?"


# Setup {#sec-setup}

```{r load-packages}
#| code-fold: false
library(ggplot2)
library(gt)

# Load forestsearch from source (picks up suggest_thresholds.R
# and other new files without needing devtools::install())
devtools::load_all(here::here(), quiet = TRUE)
```


# Installation {#sec-install}

The `suggest_thresholds()` function lives in
`R/suggest_thresholds.R`.  After placing it in the package
`R/` directory:

```{r install-note}
#| eval: false
#| code-fold: false

# From your package root:
devtools::document()   # generates man/suggest_thresholds.Rd
devtools::load_all()   # makes the function available
?suggest_thresholds    # view the help page
```


# Example 1: Survival (Paper Defaults) {#sec-survival}

For survival, the paper defaults should emerge naturally when
$L_{\text{eff}} = 1$ (the default).

```{r survival-example}
#| code-fold: false

# Subgroup: n_H = 91 (pH=13%, N=700), 46% censoring
d_surv <- d_eff_survival(n_sg = 91, prop_cens = 0.46)
cat(sprintf("d_eff (survival): %.0f events\n", d_surv))

# Without L_eff correction (default: C=1, alpha=0 -> L_eff=1)
frontier_surv <- suggest_thresholds(
  d_eff = d_surv, N = 700,
  fpr_target = 0.12
)
```

```{r survival-table}
#| tbl-cap: "Survival frontier: smallest c1 values achieving P1 <= 12% (L_eff = 1, so FPR_corr = P1)"

gt(head(frontier_surv[, c("c1", "min_c2", "P1")], 10)) |>
  tab_header(
    title = "Survival: Threshold Frontier (L_eff = 1)",
    subtitle = sprintf("d_eff = %d, N = 700", round(d_surv))
  ) |>
  fmt_number(columns = P1, decimals = 3) |>
  tab_source_note(
    "With L_eff = 1: corrected FPR = P1 (no multiplicity adjustment)."
  )
```

With $L_{\text{eff}} = 1$, the per-subgroup approximation
gives $P_1(1.25, 1.0) \approx 17\%$.  The paper found
simulated FPR of ~5.6% at N = 700 because the true
$L_{\text{eff}} < 1$ (the procedure is *more* conservative
than a single test due to the events-based filters).
The paper defaults are safe for survival.


# Example 2: Binary Outcome {#sec-binary}

## Without $L_{\text{eff}}$ Correction

```{r binary-no-leff}
#| code-fold: false

# Subgroup: 30% prevalence, N=700 -> n_H = 210, p = 0.30
d_bin <- d_eff_binary(n_sg = 210, p_event = 0.30)
cat(sprintf("d_eff (binary): %.1f\n", d_bin))

# Without correction: gives the per-subgroup P1 frontier
frontier_bin_raw <- suggest_thresholds(
  d_eff = d_bin, N = 700,
  fpr_target = 0.12
)
```

```{r binary-no-leff-table}
#| tbl-cap: "Binary frontier WITHOUT L_eff correction (P1 only)"

gt(head(frontier_bin_raw[, c("c1", "min_c2", "P1")], 8)) |>
  tab_header(
    title = "Binary: Per-Subgroup Frontier (L_eff = 1)",
    subtitle = sprintf("d_eff = %.1f, N = 700", d_bin)
  ) |>
  fmt_number(columns = P1, decimals = 3) |>
  tab_source_note(
    "With L_eff = 1: corrected FPR = P1. This UNDERESTIMATES the true procedure-level FPR."
  )
```

Without the $L_{\text{eff}}$ correction, the frontier suggests
lower thresholds than are actually needed.

## With $L_{\text{eff}}$ Correction

The $L_{\text{eff}}$ calibration from
`binary_threshold_calibration.qmd` (5000 sims, 4 confounders):
$C = 0.220$, $\alpha = 1.298$.

```{r binary-with-leff}
#| code-fold: false

# With L_eff correction
frontier_bin <- suggest_thresholds(
  d_eff = d_bin, N = 700,
  fpr_target = 0.12,
  L_eff_C = 0.220,
  L_eff_alpha = 1.298
)
```

```{r binary-leff-table}
#| tbl-cap: "Binary frontier WITH L_eff correction.  Note: P1 is 2-3% but the corrected FPR is 10-12% due to the L_eff multiplier."

gt(head(frontier_bin, 10)) |>
  tab_header(
    title = "Binary: Corrected Frontier (L_eff calibrated)",
    subtitle = sprintf("d_eff = %.1f, N = 700, L_eff = %.1f",
      d_bin,
      0.220 * (700 / 60)^1.298)
  ) |>
  fmt_number(columns = c(FPR_corr, P1), decimals = 3) |>
  fmt_number(columns = L_eff, decimals = 1) |>
  tab_source_note(
    "P1 = per-subgroup probability. FPR_corr = 1-(1-P1)^L_eff = procedure-level FPR."
  )
```

The corrected frontier shows that c1 = 1.55--1.60 with
c2 = 1.50--1.55 is needed to achieve 12% FPR control.


## How It Changes With Sample Size

```{r binary-by-N, fig.height=6}
#| fig-cap: "Binary frontier across sample sizes.  Larger N requires higher thresholds because L_eff grows."
#| code-fold: false

N_values <- c(300, 500, 700, 1000)
p_harm <- 0.30
p_event <- 0.30

frontier_list <- lapply(N_values, function(N_val) {
  n_H <- round(p_harm * N_val)
  d <- d_eff_binary(n_H, p_event)
  ft <- suggest_thresholds(
    d_eff = d, N = N_val,
    fpr_target = 0.12,
    L_eff_C = 0.220, L_eff_alpha = 1.298
  )
  if (nrow(ft) > 0) {
    ft$N <- N_val
    ft$n_H <- n_H
    ft$d_eff_val <- d
  }
  ft
})

frontier_all <- do.call(rbind, frontier_list)

ggplot(frontier_all, aes(x = c1, y = min_c2,
                          colour = factor(N))) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = expression(c[1] ~ "(screening threshold)"),
       y = expression("Min " * c[2] ~ "(consistency)"),
       title = "Binary Outcome: Threshold Frontier by N",
       subtitle = "Smallest (c1, c2) achieving corrected FPR <= 12%",
       colour = "N") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right",
        plot.title = element_text(face = "bold"))
```


# Example 3: Using the Result in ForestSearch {#sec-usage}

```{r usage-example}
#| eval: false
#| code-fold: false

# --- Step 1: Determine d_eff for your expected subgroup ---
# Binary: 30% prevalence, N = 700
n_H_expected <- round(0.30 * 700)
d_eff <- d_eff_binary(n_sg = n_H_expected, p_event = 0.30)

# --- Step 2: Get threshold recommendation ---
# Use L_eff from prior calibration
frontier <- suggest_thresholds(
  d_eff = d_eff, N = 700,
  fpr_target = 0.12,
  L_eff_C = 0.220, L_eff_alpha = 1.298
)
print(frontier[1:3, ])  # smallest c1 values

# --- Step 3: Pick thresholds and run ForestSearch ---
# Choose: c1 = 1.60, c2 = 1.50 (from frontier)
fs_result <- forestsearch(
  df.analysis = my_data,
  confounders.name = c("bm1", "bm2", "age", "ecog"),
  outcome.name = "response",
  event.name = "response",
  treat.name = "treat",
  id.name = "id",
  outcome_type = "binary",
  effect_measure = "OR",
  adverse_outcome = TRUE,
  hr.threshold = 1.60,       # from suggest_thresholds()
  hr.consistency = 1.50,     # from suggest_thresholds()
  pconsistency.threshold = 0.90,
  fs.splits = 400,
  n.min = 60,
  maxk = 2,
  use_lasso = TRUE,
  use_grf = TRUE,
  use_twostage = TRUE,
  is.RCT = TRUE,
  plot.sg = TRUE
)
```


# Example 4: Cross-Outcome Comparison {#sec-compare}

```{r compare-outcomes, fig.height=7}
#| fig-cap: "Threshold frontiers by outcome type.  Survival (high d_eff) needs lower thresholds; binary (low d_eff) needs higher thresholds."

# Common parameters
N <- 700
fpr_target <- 0.12

# Survival: n_H = 91 (pH=13%), 46% censoring
d_surv <- d_eff_survival(91, 0.46)
ft_surv <- suggest_thresholds(d_surv, N, fpr_target)
ft_surv$outcome <- sprintf("Survival (d_eff=%d)", round(d_surv))

# Binary: n_H = 210 (pH=30%), p=0.30, with L_eff
d_bin <- d_eff_binary(210, 0.30)
ft_bin <- suggest_thresholds(d_bin, N, fpr_target,
  L_eff_C = 0.220, L_eff_alpha = 1.298)
ft_bin$outcome <- sprintf("Binary (d_eff=%.0f, L_eff=%.1f)",
  d_bin, 0.220 * (700 / 60)^1.298)

# Binary without L_eff (to show the difference)
ft_bin_raw <- suggest_thresholds(d_bin, N, fpr_target)
ft_bin_raw$outcome <- sprintf("Binary (d_eff=%.0f, no L_eff)", d_bin)

compare_df <- rbind(ft_surv, ft_bin_raw, ft_bin)

ggplot(compare_df, aes(x = c1, y = min_c2,
                        colour = outcome)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  geom_point(data = data.frame(c1 = 1.25, min_c2 = 1.0),
             colour = "black", size = 4, shape = 8) +
  annotate("text", x = 1.28, y = 1.03,
           label = "Paper\ndefaults", size = 3, hjust = 0) +
  labs(x = expression(c[1] ~ "(screening threshold)"),
       y = expression("Min " * c[2] ~ "(consistency)"),
       title = "Threshold Frontier: Cross-Outcome Comparison",
       subtitle = sprintf("N=%d, FPR target=%.0f%%",
                           N, 100 * fpr_target),
       colour = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        plot.title = element_text(face = "bold"))
```


# The $L_{\text{eff}}$ Calibration Step {#sec-leff}

## Why It Matters

Without the $L_{\text{eff}}$ correction, `suggest_thresholds()`
returns the **per-subgroup** frontier, which underestimates
the procedure-level FPR.  The underestimation grows with $N$
because more candidates are evaluated.

| DGM | Confounders | $L_{\text{eff}}$ at N=700 | Source |
|---|---|---|---|
| GBSG survival | 7 | ~0.3 | cox_vs_glm_approximation.qmd |
| Binary (4 conf) | 4 | ~5.3 | binary_threshold_calibration.qmd |

## How to Calibrate

See `binary_threshold_calibration_null-search.qmd` for the
full blueprint.  The short version:

```{r calibrate-example}
#| eval: false
#| code-fold: false

# --- Quick L_eff calibration (100 sims x 5 N values) ---
# 1. Define your H0 data generator
sim_null <- function(n, seed) { ... }

# 2. Run ForestSearch under H0 at multiple N values with
#    a reference threshold (e.g., c1=1.25, c2=1.0).
#    Record the simulated FPR at each N.

# 3. Compute P1 at each N
P1_vec <- sapply(N_values, function(N) {
  n_H <- round(p_harm * N)
  d <- d_eff_binary(n_H, p_event)
  compute_detection_probability_glm(1.0, d, 1.25, 1.0,
    effect_scale = "ratio")
})

# 4. Fit L_eff
cal <- calibrate_L_eff(
  N = N_values, P1 = P1_vec,
  sim_fpr = sim_fpr_vec, n_min = 60
)
# cal$C, cal$alpha

# 5. Use in suggest_thresholds()
suggest_thresholds(d_eff = d, N = 700,
  L_eff_C = cal$C, L_eff_alpha = cal$alpha,
  fpr_target = 0.12)
```


## When You Can Skip It

If you have no calibration data, calling
`suggest_thresholds()` with the defaults ($C = 1$,
$\alpha = 0$) gives the per-subgroup frontier.  This is
**conservative for survival** (where $L_{\text{eff}} < 1$)
and **optimistic for binary** (where $L_{\text{eff}} > 1$).

For a quick rule of thumb without calibration:

- **Survival**: paper defaults (c1 = 1.25, c2 = 1.0) are safe
- **Binary**: start with c1 = 1.60, c2 = 1.50 and validate


# Quick Reference {#sec-reference}

## Function Signature

```{r signature}
#| eval: false
#| code-fold: false

suggest_thresholds(
  d_eff,                              # from d_eff_*() helpers
  N,                                  # total sample size
  fpr_target = 0.10,                  # max procedure-level FPR
  n_min = 60,                         # min subgroup size
  L_eff_C = 1.0,                      # calibration constant
  L_eff_alpha = 0.0,                  # calibration exponent
  c1_range = c(1.05, 2.50),           # search range
  c1_step = 0.05,                     # grid resolution
  c2_candidates = seq(0.90, 2.0, 0.05),
  effect_scale = "ratio"              # or "difference"
)
```

## $d_{\text{eff}}$ by Outcome Type

```{r deff-table}
#| tbl-cap: "d_eff helpers and typical values"

deff_ref <- data.frame(
  `Outcome Type` = c("Survival (Cox PH)", "Binary (logit)",
                      "Count (Poisson + offset)", "Continuous"),
  `d_eff` = c("d_eff_survival(n_H, prop_cens)",
               "d_eff_binary(n_H, p_event)",
               "d_eff_count(D)",
               "d_eff_continuous(n_H, sigma_y)"),
  `Typical range` = c("33-65 (n_H=60-120, 46% cens)",
                       "13-44 (n_H=60-210, p=0.30)",
                       "Same as survival",
                       "Depends on sigma"),
  check.names = FALSE
)

gt(deff_ref) |>
  tab_header(title = "d_eff Helpers by Outcome Type")
```


## Reference $L_{\text{eff}}$ Calibrations

```{r leff-table}
#| tbl-cap: "Published L_eff calibrations"

leff_ref <- data.frame(
  DGM = c("GBSG survival", "Binary (4 confounders)"),
  C = c(0.029, 0.220),
  alpha = c(0.882, 1.298),
  `L_eff at N=700` = c(
    round(0.029 * (700/60)^0.882, 1),
    round(0.220 * (700/60)^1.298, 1)),
  Source = c("cox_vs_glm_approximation.qmd",
             "binary_threshold_calibration.qmd"),
  check.names = FALSE
)

gt(leff_ref) |>
  tab_header(title = "Reference L_eff Calibrations")
```


## Recommended Thresholds Summary

```{r recommendations}
#| tbl-cap: "Recommended starting thresholds by outcome type"

rec <- data.frame(
  `Outcome Type` = c("Survival", "Binary (p~0.30)",
                      "Count (Poisson+offset)"),
  `c1` = c("1.25", "1.55-1.60", "1.25 (same as survival)"),
  `c2` = c("1.00", "1.50-1.55", "1.00 (same as survival)"),
  `FPR at N=700` = c("~6% (GBSG)", "~10-12% (4 conf)",
                      "~6-10% (anchored to Cox)"),
  Source = c("Paper (Table 1)", "Calibration study",
             "Poisson-Cox equivalence"),
  check.names = FALSE
)

gt(rec) |>
  tab_header(
    title = "Recommended Starting Thresholds",
    subtitle = "Validate with suggest_thresholds() for your specific configuration"
  ) |>
  tab_source_note(
    "Paper defaults (c1=1.25, c2=1.0) remain the forestsearch() signature defaults."
  )
```
