---
title: "Fixed vs Random Baseline: A Design Comparison"
subtitle: "Extreme Subgroup Effects Under a Uniform Treatment Benefit — conditional-on-X vs population-resampling simulation"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-tools: true
    theme: cosmo
    self-contained: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(gt)
library(dplyr)

# =========================================================================
# Source of the numbers below
# =========================================================================
# Both quantities were extracted from the Section 6.6.1 summary tables of
# the two rendered vignettes (10,000 simulated trials each, same DGM,
# super-population seed 99, AHR-calibrated true HR = 0.70):
#
#   random-X : extreme_subgroups.html        (baseline = "resample")
#   fixed-X  : extreme_subgroups_fixed.html  (baseline = "fixed")
#
# This document is a static comparison memo.  If the vignettes later adopt
# save_sim_results() to persist sim_hrs/sim_ubs/sim_ns as RDS, this memo
# can be upgraded to recompute everything programmatically — see the
# Reproduction appendix.
cmp <- data.frame(
  subgroup = c(
    "All Patients",
    "Post-menopausal",
    "Pre-menopausal",
    "Grade 3",
    "Grade 1/2",
    "Age (young)",
    "Age (older)",
    "Age <= 50",
    "Age > 50",
    "Tumour size (small)",
    "Tumour size (large)",
    "Node-negative",
    "Nodes 1-3",
    "High nodes (>Q3)",
    "PGR low",
    "PGR high",
    "ER-low (<20)",
    "ER-high (>=20)",
    "Pre-meno / Grade 3",
    "Post-meno / Grade 3",
    "Pre-meno / Grade 1-2",
    "Post-meno / Grade 1-2",
    "Pre-meno / Age<=50",
    "Pre-meno / Age>50",
    "Post-meno / Age (young)",
    "Post-meno / Age (older)",
    "Pre-meno / ER-low",
    "Pre-meno / ER-high",
    "Post-meno / ER-low",
    "Post-meno / ER-high",
    "Grade 3 / Node-neg",
    "Grade 3 / Node-pos",
    "Grade 1-2 / Node-neg",
    "Grade 1-2 / Node-pos",
    "Grade 3 / High nodes",
    "Grade 3 / ER-low",
    "Grade 3 / ER-high",
    "Grade 1-2 / ER-low",
    "Grade 1-2 / ER-high",
    "Grade 3 / PGR low",
    "Grade 3 / PGR high",
    "Grade 1-2 / PGR low",
    "Grade 1-2 / PGR high",
    "Pre-meno/Yng/ER-low",
    "Pre-meno/Yng/ER-high",
    "Post-meno/G3/ER-low",
    "Post-meno/G3/Node-pos",
    "G3/Node-pos/ER-low",
    "Large/Node-pos",
    "Large/Node-neg",
    "Small/Node-pos",
    "Small/Node-neg",
    "random60",
    "random40",
    "random20",
    "random15"),
  N_r = c(
    686,
    389,
    297,
    161,
    525,
    373,
    313,
    295,
    391,
    357,
    329,
    NA,
    381,
    135,
    340,
    346,
    276,
    410,
    78,
    83,
    219,
    306,
    258,
    38,
    80,
    309,
    139,
    157,
    137,
    252,
    NA,
    161,
    NA,
    525,
    44,
    110,
    51,
    167,
    359,
    127,
    33,
    213,
    312,
    120,
    138,
    57,
    83,
    110,
    329,
    NA,
    357,
    NA,
    60,
    40,
    20,
    15),
  N_f = c(
    686,
    396,
    290,
    161,
    525,
    363,
    323,
    289,
    397,
    353,
    333,
    NA,
    376,
    143,
    343,
    343,
    272,
    414,
    74,
    87,
    216,
    309,
    255,
    35,
    77,
    319,
    134,
    156,
    138,
    258,
    NA,
    161,
    NA,
    525,
    46,
    106,
    55,
    166,
    359,
    127,
    34,
    216,
    309,
    116,
    139,
    58,
    87,
    106,
    333,
    NA,
    353,
    NA,
    60,
    40,
    20,
    15),
  ub2_r = c(
    0.0,
    0.0,
    0.0,
    0.5,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    NA,
    0.0,
    0.7,
    0.0,
    0.0,
    0.0,
    0.0,
    13.1,
    8.7,
    0.5,
    0.0,
    0.1,
    60.6,
    15.1,
    0.0,
    2.2,
    4.6,
    1.5,
    0.1,
    NA,
    0.5,
    NA,
    0.0,
    25.0,
    2.7,
    33.3,
    0.9,
    0.0,
    1.3,
    56.9,
    0.1,
    0.1,
    3.9,
    6.6,
    19.4,
    8.7,
    2.7,
    0.0,
    NA,
    0.0,
    NA,
    31.6,
    50.5,
    80.5,
    90.2),
  ub2_f = c(
    0.0,
    0.0,
    0.0,
    0.3,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    NA,
    0.0,
    0.4,
    0.0,
    0.0,
    0.0,
    0.0,
    13.7,
    6.6,
    0.6,
    0.0,
    0.1,
    62.3,
    16.0,
    0.0,
    2.4,
    4.2,
    1.4,
    0.1,
    NA,
    0.3,
    NA,
    0.0,
    22.2,
    2.8,
    27.9,
    1.0,
    0.0,
    1.1,
    55.2,
    0.1,
    0.1,
    4.1,
    5.8,
    17.3,
    6.6,
    2.8,
    0.0,
    NA,
    0.0,
    NA,
    31.1,
    50.6,
    80.4,
    90.1),
  ub3_r = c(
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    NA,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    1.1,
    0.5,
    0.0,
    0.0,
    0.0,
    34.2,
    1.7,
    0.0,
    0.0,
    0.1,
    0.0,
    0.0,
    NA,
    0.0,
    NA,
    0.0,
    5.2,
    0.0,
    8.7,
    0.0,
    0.0,
    0.0,
    29.1,
    0.0,
    0.0,
    0.1,
    0.2,
    3.3,
    0.5,
    0.0,
    0.0,
    NA,
    0.0,
    NA,
    8.2,
    23.1,
    61.8,
    77.4),
  ub3_f = c(
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    NA,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    1.2,
    0.2,
    0.0,
    0.0,
    0.0,
    36.0,
    1.7,
    0.0,
    0.0,
    0.1,
    0.0,
    0.0,
    NA,
    0.0,
    NA,
    0.0,
    3.4,
    0.0,
    5.9,
    0.0,
    0.0,
    0.0,
    27.1,
    0.0,
    0.0,
    0.1,
    0.3,
    2.2,
    0.2,
    0.0,
    0.0,
    NA,
    0.0,
    NA,
    7.8,
    22.8,
    60.5,
    76.7),
  mUB_r = c(
    0.90,
    0.97,
    1.03,
    1.12,
    0.94,
    0.99,
    1.01,
    1.03,
    0.98,
    1.00,
    1.00,
    NA,
    0.99,
    1.14,
    0.97,
    1.03,
    1.01,
    0.99,
    1.36,
    1.31,
    1.11,
    1.02,
    1.06,
    2.32,
    1.40,
    1.01,
    1.18,
    1.24,
    1.16,
    1.07,
    NA,
    1.12,
    NA,
    0.94,
    1.54,
    1.20,
    1.67,
    1.13,
    1.01,
    1.15,
    2.19,
    1.05,
    1.06,
    1.22,
    1.27,
    1.46,
    1.31,
    1.20,
    1.00,
    NA,
    1.00,
    NA,
    1.64,
    2.01,
    3.81,
    6.05),
  mUB_f = c(
    0.90,
    0.97,
    1.03,
    1.11,
    0.94,
    0.99,
    1.00,
    1.03,
    0.97,
    0.99,
    0.99,
    NA,
    0.99,
    1.12,
    0.96,
    1.03,
    1.01,
    0.98,
    1.37,
    1.28,
    1.11,
    1.02,
    1.06,
    2.41,
    1.42,
    1.00,
    1.19,
    1.23,
    1.16,
    1.06,
    NA,
    1.11,
    NA,
    0.94,
    1.50,
    1.20,
    1.59,
    1.14,
    1.01,
    1.14,
    2.14,
    1.05,
    1.06,
    1.23,
    1.27,
    1.43,
    1.28,
    1.20,
    0.99,
    NA,
    0.99,
    NA,
    1.63,
    2.01,
    3.67,
    5.83),
  hr05_r = c(
    0.2,
    1.3,
    3.0,
    5.6,
    0.6,
    1.7,
    2.2,
    3.1,
    1.2,
    2.5,
    1.5,
    NA,
    2.7,
    6.3,
    1.4,
    3.0,
    2.8,
    1.4,
    14.8,
    11.9,
    6.1,
    2.9,
    4.2,
    28.7,
    15.8,
    2.3,
    9.2,
    10.8,
    8.5,
    3.9,
    NA,
    5.6,
    NA,
    0.6,
    17.7,
    9.6,
    19.0,
    7.7,
    2.2,
    7.6,
    26.9,
    4.6,
    4.0,
    10.9,
    12.3,
    16.9,
    11.9,
    9.6,
    1.5,
    NA,
    2.5,
    NA,
    21.0,
    26.5,
    34.1,
    36.8),
  hr05_f = c(
    0.1,
    1.1,
    3.2,
    5.1,
    0.7,
    1.5,
    1.8,
    3.2,
    1.0,
    2.4,
    1.3,
    NA,
    2.8,
    5.1,
    1.1,
    3.3,
    2.2,
    1.4,
    14.7,
    11.1,
    6.2,
    2.5,
    4.2,
    29.9,
    15.8,
    1.9,
    9.2,
    10.2,
    8.0,
    3.6,
    NA,
    5.1,
    NA,
    0.7,
    17.4,
    9.7,
    18.3,
    7.4,
    2.2,
    6.8,
    27.3,
    4.1,
    4.3,
    10.4,
    11.5,
    17.1,
    11.1,
    9.7,
    1.3,
    NA,
    2.4,
    NA,
    20.7,
    25.7,
    34.1,
    37.6),
  hr1_r = c(
    0.2,
    1.5,
    3.8,
    7.1,
    0.9,
    1.9,
    2.8,
    3.3,
    1.6,
    2.5,
    2.3,
    NA,
    2.4,
    7.9,
    1.2,
    3.8,
    2.5,
    2.1,
    15.7,
    14.8,
    7.0,
    3.1,
    4.6,
    28.4,
    15.9,
    2.8,
    9.6,
    11.6,
    8.8,
    5.0,
    NA,
    7.1,
    NA,
    0.9,
    20.3,
    9.9,
    22.7,
    7.4,
    3.0,
    8.4,
    28.2,
    4.0,
    4.5,
    11.2,
    12.8,
    17.8,
    14.8,
    9.9,
    2.3,
    NA,
    2.5,
    NA,
    21.0,
    26.0,
    34.5,
    36.9),
  hr1_f = c(
    0.3,
    1.6,
    3.7,
    6.4,
    1.0,
    2.2,
    2.8,
    3.6,
    1.6,
    2.3,
    2.1,
    NA,
    2.2,
    7.2,
    1.1,
    3.8,
    2.7,
    2.1,
    15.8,
    12.9,
    6.9,
    3.2,
    4.7,
    28.7,
    16.6,
    2.8,
    9.5,
    11.5,
    8.7,
    4.6,
    NA,
    6.4,
    NA,
    1.0,
    19.4,
    10.1,
    20.5,
    7.5,
    2.8,
    8.1,
    27.4,
    4.2,
    4.8,
    11.3,
    12.5,
    17.2,
    12.9,
    10.1,
    2.1,
    NA,
    2.3,
    NA,
    21.4,
    25.9,
    33.7,
    35.7),
  mHR_r = c(
    0.72,
    0.72,
    0.72,
    0.72,
    0.72,
    0.72,
    0.72,
    0.71,
    0.72,
    0.71,
    0.72,
    NA,
    0.71,
    0.72,
    0.71,
    0.71,
    0.71,
    0.72,
    0.71,
    0.73,
    0.71,
    0.71,
    0.72,
    0.71,
    0.71,
    0.72,
    0.71,
    0.71,
    0.71,
    0.72,
    NA,
    0.72,
    NA,
    0.72,
    0.72,
    0.72,
    0.73,
    0.71,
    0.72,
    0.72,
    0.72,
    0.71,
    0.71,
    0.71,
    0.71,
    0.71,
    0.73,
    0.72,
    0.72,
    NA,
    0.71,
    NA,
    0.72,
    0.71,
    0.72,
    0.71),
  mHR_f = c(
    0.72,
    0.72,
    0.71,
    0.72,
    0.72,
    0.72,
    0.72,
    0.72,
    0.72,
    0.71,
    0.72,
    NA,
    0.70,
    0.72,
    0.71,
    0.71,
    0.71,
    0.72,
    0.71,
    0.72,
    0.71,
    0.71,
    0.72,
    0.71,
    0.72,
    0.71,
    0.71,
    0.71,
    0.71,
    0.72,
    NA,
    0.72,
    NA,
    0.72,
    0.72,
    0.71,
    0.73,
    0.71,
    0.71,
    0.71,
    0.72,
    0.71,
    0.71,
    0.71,
    0.72,
    0.71,
    0.72,
    0.71,
    0.72,
    NA,
    0.71,
    NA,
    0.71,
    0.72,
    0.71,
    0.69),
  stringsAsFactors = FALSE
)

# Derived columns used throughout
cmp <- cmp |>
  mutate(
    d_ub2 = ub2_f - ub2_r,          # fixed minus random, percentage points
    d_ub3 = ub3_f - ub3_r,
    d_N   = N_f - N_r,
    empty = is.na(N_r) | is.na(N_f),
    kind  = case_when(
      subgroup == "All Patients"        ~ "ITT",
      grepl("^random", subgroup)        ~ "Random benchmark",
      TRUE                              ~ "Covariate-defined"
    )
  )
valid <- cmp |> filter(!empty)
```

# Overview and provenance

This memo compares the two companion simulation designs for the
extreme-subgroups analysis, both run at **10,000 simulated trials** on the
same AHR-calibrated uniform-effect DGM (true HR = 0.70 for every patient,
GBSG source data, super-population seed 99):

- **Random-X** (`extreme_subgroups.html`, `baseline = "resample"`): every
  trial resamples 686 subjects with replacement from the super-population,
  so covariates — and hence subgroup composition and sizes — vary across
  trials.  This answers the *population-replication* question: what if the
  trial had enrolled a different sample from the same population?
- **Fixed-X** (`extreme_subgroups_fixed.html`, `baseline = "fixed"`): every
  trial uses the observed GBSG panel — each source patient exactly once —
  so subgroup composition and sizes are identical in every replicate.  This
  answers the *conditional-on-X* question: given exactly these patients,
  how variable are the subgroup analyses across re-runs of the trial?

```{r headline-stats}
#| results: asis
mstat <- valid |> summarise(
  mean_d   = mean(d_ub2),
  med_d    = median(d_ub2),
  min_d    = min(d_ub2),
  max_d    = max(d_ub2)
)
cat(sprintf(
  paste0("**Headline:** across all %d non-empty subgroups the mean ",
         "difference in Pr(UB \u2265 2) is **%+.2f pp** (fixed minus ",
         "random), median %+.2f pp, range [%+.1f, %+.1f].  Median HR is ",
         "0.69\u20130.73 in every row of both designs.\n"),
  nrow(valid), mstat$mean_d, mstat$med_d, mstat$min_d, mstat$max_d))
```

At 10,000 independent simulations per design, the Monte Carlo 95% noise
band for a difference in two proportions is approximately
$\pm 2\sqrt{2\,p(1-p)/10{,}000}$ — about **±2.0 pp at p = 50%** and
±0.9 pp at p = 5%.  The vast majority of subgroup-level differences sit
inside that band.  **The central message of both vignettes — extreme
results are routine in small subgroups, and clinical rationale confers no
statistical protection — is design-robust.**  A skeptic cannot attribute
the findings to covariate-resampling artifacts.

# Random benchmarks: the internal-consistency check

A random draw of $k$ patients is distributionally almost unaffected by
whether the panel it draws from was itself resampled.  The four random
benchmarks are therefore predicted to be *invariant* across the two
designs — and they are:

```{r tbl-benchmarks}
valid |>
  filter(kind == "Random benchmark") |>
  transmute(
    Subgroup = subgroup,
    `N (both)` = N_r,
    `Pr(UB>=2), random-X` = sprintf("%.1f%%", ub2_r),
    `Pr(UB>=2), fixed-X`  = sprintf("%.1f%%", ub2_f),
    `Pr(UB>=3), random-X` = sprintf("%.1f%%", ub3_r),
    `Pr(UB>=3), fixed-X`  = sprintf("%.1f%%", ub3_f),
    `mUB, random-X` = sprintf("%.2f", mUB_r),
    `mUB, fixed-X`  = sprintf("%.2f", mUB_f)
  ) |>
  gt() |>
  tab_header(title = "Random benchmarks are design-invariant") |>
  tab_options(table.font.size = px(13), column_labels.font.weight = "bold")
```

Under fixed-X these benchmarks isolate *pure membership noise at exactly
fixed N* (sizes are 60/40/20/15 in every replicate by construction);
under random-X they additionally inherit panel resampling.  The
equivalence above shows that second component is negligible — which is
precisely what makes the benchmarks valid calibration anchors in both
documents.

# Where the designs differ — and why

Every covariate-defined subgroup that moved meaningfully moved in
lockstep with its realized N: the sign of the change in Pr(UB ≥ 2) is
opposite the sign of the change in N essentially everywhere in the
high-risk panel.

```{r tbl-movers}
valid |>
  filter(kind == "Covariate-defined", abs(d_ub2) >= 1.5) |>
  # Drop exact-duplicate result rows here (the Node-pos duplicates,
  # explained in the Data Quirks section below) so each mover appears once.
  distinct(N_r, N_f, ub2_r, ub2_f, ub3_r, mUB_r, .keep_all = TRUE) |>
  arrange(d_ub2) |>
  transmute(
    Subgroup = subgroup,
    `N random-X` = N_r, `N fixed-X` = N_f,
    `Pr(UB>=2) random-X` = sprintf("%.1f%%", ub2_r),
    `Pr(UB>=2) fixed-X`  = sprintf("%.1f%%", ub2_f),
    `Delta (pp)` = sprintf("%+.1f", d_ub2)
  ) |>
  gt() |>
  tab_header(
    title = "Subgroups with |Delta Pr(UB>=2)| >= 1.5 pp",
    subtitle = "Direction tracks realized N: larger fixed panel => lower tail risk, and vice versa"
  ) |>
  tab_options(table.font.size = px(13), column_labels.font.weight = "bold")
```

**The mechanism has a twist worth understanding.**  The fixed-X N is the
*true* GBSG count for each subgroup, while the random-X mean N inherits
the drift of the one super-population realization (5,000 draws at
seed 99).  Grade 3 / ER-high genuinely has 55 patients in GBSG; the
super-population happened to under-represent it, so random-X simulated it
at mean N ≈ 51 and correspondingly overstated its tail risk (33.3% vs
27.9%).  The deltas above are therefore **not** purely the textbook
"random-X adds denominator variance" effect — a substantial share is the
super-population's own finite-sampling error, which fixed-X sidesteps
entirely by conditioning on the exact source composition.  (If cleaner
random-X numbers were ever needed, raising `n_super` from 5,000 to
50,000 would shrink that component.)

Beyond composition, the pure variance effect is visible but small:
same-N rows (random40, random60, Grade 3 at N = 161) show whiskers and
tail probabilities a touch tighter under fixed-X, well within one or two
percentage points.

# Section 6.6.4 high-risk panels

The two rendered Section 6.6.4 panels are near-interchangeable: the same
**11 subgroups** pass the Pr(UB ≥ 2) ≥ 10% filter in both designs, with a
single ordering swap — Grade 3 / ER-high drops below random60 under
fixed-X, driven by its larger (true) panel N of 55 vs the random-X mean
of 51.  ECI whiskers for Grade 3 / High nodes and Post-meno/G3/ER-low are
visibly slightly shorter under fixed-X; everything else is
indistinguishable by eye.

# Data quirks surfaced by the comparison

Two artefacts of the subgroup list — present in **both** designs, and
worth fixing in both vignettes:

```{r quirks}
#| results: asis
empties <- cmp |> filter(empty) |> pull(subgroup)
cat(sprintf(
  paste0("1. **Five structurally empty subgroups.** `survival::gbsg` is an ",
         "all-node-positive cohort (minimum nodes = 1), so `nodes == 0` ",
         "never matches.  Empty in both designs: %s.\n"),
  paste0("*", empties, "*", collapse = ", ")))
```

```{r dup-pairs}
#| results: asis
# Detect exact-duplicate result rows (same stats in BOTH designs):
# a consequence of the same quirk -- "Node-pos" is every patient, so each
# Node-pos subgroup duplicates its parent definition.
key <- valid |>
  mutate(k = paste(N_r, N_f, ub2_r, ub2_f, ub3_r, ub3_f, mUB_r, mUB_f,
                   hr05_r, hr05_f, mHR_r, mHR_f))
dups <- key |> group_by(k) |> filter(n() > 1) |>
  summarise(pair = paste(subgroup, collapse = "  ==  "), .groups = "drop")
cat("2. **Duplicate subgroup pairs** (identical results in both designs, ",
    "because Node-pos is everyone):\n\n", sep = "")
for (p in dups$pair) cat("   - ", p, "\n", sep = "")
cat(sprintf("\nEffective count of distinct, populated subgroups: **%d**, not 56.\n",
            nrow(valid) - nrow(dups)))
```

**Suggested fix** (both qmds): replace the `nodes == 0` / `nodes > 0`
splits with `nodes <= 3` vs `nodes > 3` — both populated in GBSG — which
removes the empties and the duplicates in one edit.

# Practical guidance

The **fixed-X** document answers the conditional question a regulatory
reviewer implicitly asks — *given exactly this trial's enrolled patients
and subgroup sizes, how extreme can chance make the analyses look?* — with
subgroup Ns that match the actual submission table exactly.  The
**random-X** document answers the population-replication question and is
the right benchmark when generalizing beyond the specific trial.

Since the results are near-identical, the suggestion is to **lead with
fixed-X for the regulatory framing** (its exact-N property makes the
forest-plot annotations directly comparable to the real trial) and cite
the random-X version as the robustness companion, with the
random-benchmark invariance table above as the bridge showing the
calibration anchors are design-invariant.

# Appendix A — full comparison table {.unnumbered}

```{r tbl-full}
cmp |>
  arrange(desc(ifelse(is.na(ub2_r), -1, ub2_r))) |>
  transmute(
    Subgroup = subgroup,
    `N r|f` = ifelse(empty, "-", paste0(N_r, " | ", N_f)),
    `Pr(UB>=2) r|f` = ifelse(empty, "-", sprintf("%.1f | %.1f", ub2_r, ub2_f)),
    `Pr(UB>=3) r|f` = ifelse(empty, "-", sprintf("%.1f | %.1f", ub3_r, ub3_f)),
    `mUB r|f` = ifelse(empty, "-", sprintf("%.2f | %.2f", mUB_r, mUB_f)),
    `Pr(HR<0.5) r|f` = ifelse(empty, "-", sprintf("%.1f | %.1f", hr05_r, hr05_f)),
    `Pr(HR>1.0) r|f` = ifelse(empty, "-", sprintf("%.1f | %.1f", hr1_r, hr1_f)),
    `mHR r|f` = ifelse(empty, "-", sprintf("%.2f | %.2f", mHR_r, mHR_f))
  ) |>
  gt() |>
  tab_header(
    title = "All 56 subgroups, random-X | fixed-X",
    subtitle = "Sorted by random-X Pr(UB>=2); percentages in points; '-' = structurally empty"
  ) |>
  tab_options(table.font.size = px(11), data_row.padding = px(2),
              column_labels.font.weight = "bold")
```

# Appendix B — reproduction notes {.unnumbered}

The per-subgroup quantities in this memo were extracted verbatim from the
printed Section 6.6.1 summary tables of the two rendered HTML vignettes
(both at `n_sims_uniform = 10000`).  They are embedded here as a static
data frame (`cmp`, first chunk) so the memo renders with no external
dependencies.

To make this memo fully programmatic, have both vignettes persist their
raw simulation matrices, e.g.:

```{r repro-code}
#| echo: true
#| eval: false
# In each vignette, after the parallel loop:
saveRDS(list(sim_hrs = sim_hrs, sim_ubs = sim_ubs, sim_ns = sim_ns,
             n_sims = n_sims_uniform, design = sim_config$baseline),
        file = sprintf("results/extreme_sims_%s_%d.rds",
                       sim_config$baseline, n_sims_uniform))

# Then this memo can load both RDS files and recompute every table above,
# plus quantities not printed in the vignettes (full ECI endpoints,
# per-subgroup quantile curves, paired design deltas with bootstrap CIs).
```
