---
title: "Extreme Subgroup Effects Under a Uniform Treatment Benefit"
subtitle: "A Simulation Study Based on Observed Trial Data"
author: "Larry F. León"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    theme: cosmo
    highlight: tango
    fig_width: 8
    fig_height: 5
    self_contained: true
vignette: >
  %\VignetteIndexEntry{Extreme Subgroup Effects Under a Uniform Treatment Benefit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.align  = "center",
  fig.retina = 2
)

# Number of null-DGM simulated trials — the primary result quantity
# Set to>= 1000 for publication-quality results
n_sims_null <- 16L  


```

---

# Overview

A central concern in exploratory subgroup analysis is the potential for
**extreme results driven entirely by chance**, not by genuine treatment
effect heterogeneity.  When many subgroups are examined in a single
dataset — as is routine in regulatory submissions and trial reports —
small subgroups in particular can produce hazard ratio estimates that
look dramatic simply because they are noisy.  The smaller the subgroup,
the wider the confidence interval, and the higher the probability that
the upper bound strays far from the truth.

This vignette evaluates that risk directly.  We generate thousands of
synthetic trials from a data generating mechanism (DGM) with a
**perfectly uniform treatment benefit** — the true hazard ratio is
identical for every patient — and ask: how often do standard Cox model
analyses produce extreme subgroup results?  

The vignette is organised as a complete, self-contained workflow:

1. **Building the DGM** — fitting a Weibull AFT model to real data,
   constructing a super-population, and calibrating the censoring process
   so that simulated trials reproduce the censoring pattern of the source
   data.  This step establishes the null scenario: a DGM where the true
   HR is 0.70 *for every patient*, with no subgroup differential effects.

2. **Running simulated trials** — drawing `r n_sims_null` synthetic
   trial replicates from the null DGM and fitting a stratified Cox model
   in each pre-defined subgroup.

3. **Characterising the distribution of extreme results** — for each
   subgroup, summarising how variable the HR point estimate is across
   simulations (the empirical 99% CI, or ECI) and how large the upper
   confidence bound UB(HR) can become by chance alone.  Random benchmark
   subgroups (`random15`, `random20`, `random40`, `random60`) with no
   clinical meaning serve as calibration anchors, making explicit that
   clinical subgroups of the same size are statistically indistinguishable
   from meaningless random groupings.

All examples use the GBSG breast cancer dataset (`survival::gbsg`).


## Example of Subgroups Under Study 

For the study under question, the subgroups and features (size, events, etc) 
are of course specific to that study.  While we are using the GBSG breast cancer 
study as the case-study, the analyses can be applied or modified based on any study.

An example of pre-defined subgroups,
ranging from clinically motivated subpopulations to deliberately
artificial ones:

| Subgroup | Basis | Approximate N |
|:---------|:------|:--------------|
| All Patients (ITT) | Full trial | ~700 |
| No liver metastases | Clinical | ~400 |
| BRAF V600E mutant | Molecular | ~300 |
| Age ≤ 65 | Demographic | ~400 |
| Spain, Germany, UK | Country/region | ~60–150 |
| Asia Pacific region | Region | ~100 |
| ECOG 0 / ECOG 1 | Performance status | ~300 each |
| **random15, random20** | **Randomly sampled** | **~15, ~20** |
| **random40, random60** | **Randomly sampled** | **~40, ~60** |

The `random*` subgroups are the key benchmark.  They are constructed by
randomly sampling approximately 20% of subjects and taking the first 15,
20, 40, and 60 patients:

```r
set.seed(8316951)
sg_ran <- rbinom(nrow(df_analysis), size = 1, prob = 0.2)
temp   <- cumsum(sg_ran)

sg_ran15            <- sg_ran
sg_ran15[temp > 15] <- 0   # first 15 selected subjects
df_analysis$random15 <- sg_ran15

sg_ran40            <- sg_ran
sg_ran40[temp > 40] <- 0
df_analysis$random40 <- sg_ran40
```

These subgroups have no clinical meaning whatsoever — their true HR
equals the ITT HR of 0.70.  Any extreme result they produce is purely
a function of their small size and sampling variability.

---

## Six Analysis Variants

For each simulated trial and each subgroup, six Cox model specifications
are fitted, covering the range of approaches one might encounter in
practice:

| Label | Model specification |
|:------|:--------------------|
| `sR` | Stratified by original randomisation factors |
| `none` | Unstratified |
| `sX` | Stratified by ECOG status |
| `sBM*` | Stratified by biomarker above/below optimal cutpoint |
| `sX+sR` | Stratified by ECOG and randomisation factors |
| `X+O+sR` | Covariate-adjusted (ECOG, age, sex) + stratified |

---

## Summary Statistics Across Simulations

For each subgroup and analysis, two quantities are accumulated across all
simulated trials:

**HR point estimate distribution.** Reported as median and the 1st–99th
percentile interval (the "empirical 99% confidence interval" or ECI).
Under truth, the median should be close to 0.70 for every subgroup.
The spread of the ECI reflects how variable the estimate is — wider for
smaller subgroups.

**Upper confidence bound UB(HR).** The one-sided 97.5% upper bound of
the Cox 95% confidence interval.  Two key thresholds are tracked:
- `UB(HR) < 1.0`: the upper bound excludes no treatment effect (nominally
  "significant" benefit).
- `HR < 0.80`: the point estimate alone looks like a meaningful benefit
  (at the ITT threshold level).

Under the null of a uniform HR = 0.70 with no subgroup effects, `UB(HR) < 1.0`
should occur approximately 95% of the time for the full ITT — it simply
reflects power to detect the overall benefit.  For subgroups, the rate
falls with sample size, but the distribution of UB(HR) *among those
trials where UB(HR) ≥ 1* reveals how extreme a false non-significance
result can look.

---

We note that continuous biomarkers can also be considered.  This is more involved and 
not necessary for uniform effects, but valuable for trials with biomarkers when there 
is interest in exploring the dynamics of subgroups in the presence of actual biomarker effects (correlated with other 
patients' characteristics).

## Setting: Uniform Treatment Effect

The DGM is built from a real trial dataset using `generate_aft_dgm_flex()`
with a single-knot spline biomarker model.  The key feature is that the
true log hazard ratio is **constant across all biomarker values**:

$$
\psi^0(z) = \log(\text{HR}_{\text{true}}) \quad \text{for all } z.
$$

In the examples below the true overall HR is 0.70 (a moderate benefit).
Every subgroup — regardless of how it is defined — shares this same
underlying HR.  There is no subgroup effect; any
apparent finding is a false positive.

The spline parameterisation is specified via `log.hrs = log(c(hr0, hr_k, hr_z))`,
where `hr0`, `hr_k`, `hr_z` are the true HRs at biomarker values 0,
the knot `k`, and an upper anchor `zeta`, respectively.  Under the uniform
scenario all three are set to the same value:

```r
log.hrs <- log(c(0.70, 0.70, 0.70))  # uniform HR = 0.70 everywhere
dgm     <- get_dgm_stratified(
  df        = df_analysis,
  knot      = 40,
  kappa     = 50,
  hrs       = c(0.70, 0.70, 0.70),
  strata_tte = "strataNew"
)
```

## Installation

The development version of **forestsearch** can be installed directly
from GitHub using either `pak` (recommended) or `remotes`:
```r
# Using pak (recommended)
# install.packages("pak")
pak::pak("larry-leon/forestsearch")

# Using remotes
# install.packages("remotes")
remotes::install_github("larry-leon/forestsearch")
```

The package depends on
[**weightedsurv**](https://github.com/larry-leon/weightedsurv), which
is installed automatically from GitHub via the `Remotes` field in
`DESCRIPTION` — no separate step is needed.

Once installed, load the package in the usual way:
```r
library(forestsearch)
```

## Documentation

Full documentation, including function reference and vignettes, is
available at the **pkgdown** site:

> <https://larry-leon.github.io/forestsearch/>

The site includes, **along with this vignette**:

- **Getting Started** — end-to-end walkthrough with the GBSG breast
  cancer trial (`vignette("forestsearch")`)
- **Methodology** — statistical framework, algorithm details, and
  simulation results (`vignette("methodology")`)
- **Treatment Effect Definitions** — marginal HR, average hazard ratio
  (AHR), and controlled direct effect (CDE) estimands
- **Simulation Studies** — operating characteristics and published
  benchmarks
- **Reference** — all exported functions organised across 16 topic
  areas

Vignettes can also be accessed locally after installation:
```r
vignette("forestsearch")   # Getting Started
vignette("methodology")    # Statistical methodology
```



---

# Setup

```{r packages}
library(forestsearch)
library(survival)
library(ggplot2)
library(dplyr)
library(patchwork)     # assembles gg_forest() panels
library(here)          # here::here() for source() path in forest-load-ggforest

theme_set(theme_minimal(base_size = 12))
```

```{r data-prep}
# Load GBSG data and convert time to months (DGM time scale)
data(gbsg, package = "survival")

gbsg$time_months <- gbsg$rfstime / 30.4375
gbsg$treat        <- gbsg$hormon

cat("N =", nrow(gbsg), "  Events:", sum(gbsg$status),
    sprintf("(%.0f%%)", 100 * mean(gbsg$status)), "\n")
cat("Median follow-up (months):",
    round(median(gbsg$time_months), 1), "\n")
cat("Median event time (months):",
    round(median(gbsg$time_months[gbsg$status == 1]), 1), "\n")
```

---

# Part I — Building the Null Data Generating Mechanism

To evaluate extreme subgroup effects by chance, we need a DGM where the
treatment has a genuine, uniform benefit — so that any subgroup finding is
definitionally a false positive.  This section constructs that DGM from the
GBSG data using `generate_aft_dgm_flex()` with `model = "null"` (uniform
treatment effect throughout; no embedded harm subgroup).

## Why time-scale consistency matters first

Before fitting anything, confirm that all time parameters are expressed in
the same units. The single most common simulation failure is building the
DGM on days (e.g. `rfstime`) and then passing `analysis_time` in months.
The result is a follow-up window far shorter than the DGM event-time
distribution, producing universal administrative censoring (`event_sim = 0`
for every subject).

After building the DGM, this quick check catches the problem:

```{r timescale-check}
# exp(mu) is the implied median event time on the DGM time scale.
# It should be close to the observed median event time in months.
cat("Observed median event time (months):",
    round(median(gbsg$time_months[gbsg$status == 1]), 1), "\n")
cat("(After building dgm, exp(dgm$model_params$mu) should be similar.)\n")
```

## Fitting `generate_aft_dgm_flex()`

`generate_aft_dgm_flex()` fits a Weibull AFT model to the source data and
generates a large *super population* from which all future trials will be
drawn.  For the extreme-subgroup evaluation we use `model = "null"`, which
sets the treatment effect to be identical for every patient — there is no
subgroup, so any apparent finding is a false
positive by construction.

The censoring model is estimated from the data by AIC-based selection across
Weibull and log-normal models, with and without covariate adjustment.  The
next section shows how to verify and calibrate this so that simulated trials
reproduce the censoring pattern of the source data.

**Note that this is an example of a DGM (truth) where there is a subgroup effect (ER-low & pre-menopausal).  Simulations under (the null of) uniform treatment benefit are provided below in Part IV**

```{r build-dgm, eval=TRUE}
set.seed(42)

dgm <- generate_aft_dgm_flex(
  data            = gbsg,
  continuous_vars = c("age", "size", "nodes", "pgr", "er"),
  factor_vars     = c("meno", "grade"),
  outcome_var     = "time_months",
  event_var       = "status",
  treatment_var   = "hormon",
  subgroup_vars   = c("er", "meno"),
  subgroup_cuts   = list(er = 20, meno = 0),   # ER-low & pre-menopausal
  model           = "alt",
  k_inter         = 1.5,    # moderate interaction
  n_super         = 5000,
  seed            = 42,
  verbose         = FALSE
)
```

```{r dgm-summary}
# Key DGM outputs
cat("=== DGM Summary ===\n")
cat("Time scale          : months\n")
cat("Implied median T    : exp(mu) =",
    round(exp(dgm$model_params$mu), 1), "months\n")
cat("Subgroup proportion :", round(dgm$subgroup_info$proportion, 3), "\n")
cat("True HR (harm)      :",
    round(dgm$hazard_ratios$harm_subgroup, 3), "\n")
cat("True HR (no harm)   :",
    round(dgm$hazard_ratios$no_harm_subgroup, 3), "\n")
cat("True HR (overall)   :",
    round(dgm$hazard_ratios$overall, 3), "\n")
cat("Censoring type      :",
    dgm$model_params$censoring$type, "\n")
```

---

# Part II — Censoring Diagnostics and Calibration

For the simulation results to be credible, the censoring pattern in each
synthetic trial must match what was observed in the source data.  If the
simulated censoring rate is much higher or lower than the real one, the
event counts in small subgroups will be systematically wrong, biasing the
HR and UB(HR) distributions we are trying to characterise.

Even with a well-fitted censoring model, the interaction between the DGM
censoring distribution and administrative censoring (the `analysis_time`
cutoff) can shift the effective censoring rate substantially away from what
was observed.  This section covers two functions that diagnose and correct
for this shift before the main simulation loop is run.

## Diagnosing censoring consistency

`check_censoring_dgm()` compares three censoring summaries between the DGM
reference data (`dgm$df_super`) and a simulated dataset: overall censoring
rate, censoring-time quantiles among censored subjects, and the KM-based
median censoring time (estimated by reversing the event indicator so that
censored observations become the "event" of interest).

```{r pilot-sim}
# Generate a pilot simulated trial (no calibration yet)
sim_pilot <- simulate_from_dgm(
  dgm           = dgm,
  n             = 1500,
  analysis_time = 84,      # months — 7-year follow-up
  max_entry     = 24,      # months — 2-year recruitment window
  cens_adjust   = 0,
  seed          = 1
)

cat("Pilot trial:  N =", nrow(sim_pilot),
    "  Event rate:", round(mean(sim_pilot$event_sim), 3), "\n")
```

```{r check-censoring}
diag0 <- check_censoring_dgm(
  sim_data   = sim_pilot,
  dgm        = dgm,
  rate_tol   = 0.10,
  median_tol = 0.25,
  verbose    = TRUE
)
```

**Reading the output.** Flags warn when the absolute censoring-rate
discrepancy exceeds `rate_tol` (default 10 percentage points) or when the
KM median censoring times differ by more than `median_tol` (default 25%).
Common causes:

- **Time-scale mismatch** — check `exp(dgm$model_params$mu)` against
  `analysis_time`.
- **Short `analysis_time`** — administrative censoring dominates the
  censoring process, inflating the censoring rate and compressing
  censoring times.
- **Large `cens_adjust`** — the log-scale shift has moved the DGM
  censoring distribution far from the original data.

## Calibrating `cens_adjust`

`calibrate_cens_adjust()` uses `uniroot` to find the value of `cens_adjust`
— a log-scale shift applied to the DGM censoring model — that equates a
chosen summary statistic between the reference and simulated data. Two
targets are available: `"rate"` (overall censoring rate) and `"km_median"`
(KM-based median censoring time).

**How the objective function works.** At each candidate `cens_adjust`:

1. `simulate_from_dgm()` generates `n_eval` synthetic subjects.
2. `check_censoring_dgm()` extracts the target metric.
3. The objective returns `sim_metric - ref_metric`.

`uniroot` finds the zero crossing. The fixed `seed` inside the objective
ensures the same candidate is always evaluated identically, preventing the
root-finder from chasing Monte Carlo noise.

```{r calibrate, eval=TRUE}
cal <- calibrate_cens_adjust(
  dgm           = dgm,
  target        = "rate",
  n             = 1000,
  analysis_time = 84,
  max_entry     = 24,
  seed          = 42,
  interval      = c(-3, 3),
  n_eval        = 2000,
  verbose       = TRUE
)
```

```{r calibrate-result}
cat("\nCalibrated cens_adjust :", round(cal$cens_adjust, 4), "\n")
cat("Reference censoring rate:", round(cal$ref_value,    3),  "\n")
cat("Achieved  censoring rate:", round(cal$sim_value,    3),  "\n")
cat("Residual                :", round(cal$residual,     4),  "\n")
```

**Monotonicity note.** The objective is monotone in `cens_adjust` for both
targets: larger values increase censoring times (lower rate, higher KM
median), smaller values do the opposite. If `uniroot` fails, the target
lies outside the `interval`; widen it and retry.

## Verifying the calibrated simulation

```{r verify-cal}
sim_cal <- simulate_from_dgm(
  dgm           = dgm,
  n             = 1000,
  analysis_time = 84,
  max_entry     = 24,
  cens_adjust   = cal$cens_adjust,
  seed          = 123
)

diag_cal <- check_censoring_dgm(
  sim_data   = sim_cal,
  dgm        = dgm,
  verbose    = TRUE,
  rate_tol   = 0.05,
  median_tol = 0.15
)
```

---

# Part III — Generating Synthetic Trials

With the null DGM built, diagnosed, and calibrated, each simulated trial
is a single call to `simulate_from_dgm()`.  The function samples from the
super population with replacement, assigns treatment, generates latent
survival times from the AFT model, applies the calibrated censoring model,
and truncates at `analysis_time`.

The single-trial example below illustrates the output structure.  In the
null DGM every patient has the same true HR of 0.70, so `flag_harm` is
always 0 — there is no embedded subgroup.  The `r n_sims_null` replications
that drive the extreme-subgroup analysis in Part IV follow the same pattern,
looping over subgroups inside each replicate.

```{r single-trial}
sim_data <- simulate_from_dgm(
  dgm           = dgm,
  n             = 600,
  rand_ratio    = 1,
  analysis_time = 84,
  max_entry     = 24,
  cens_adjust   = cal$cens_adjust,
  seed          = 100
)

cat("Simulated trial:  N =", nrow(sim_data), "\n")
cat("Treatment allocation:",
    table(sim_data$treat_sim)[["0"]], "control /",
    table(sim_data$treat_sim)[["1"]], "experimental\n")
cat("Overall event rate :", round(mean(sim_data$event_sim), 3), "\n")
cat("Event rate (harm subgroup):",
    round(mean(sim_data$event_sim[sim_data$flag_harm == 1]), 3), "\n")
cat("Event rate (no-harm)      :",
    round(mean(sim_data$event_sim[sim_data$flag_harm == 0]), 3), "\n")
```

```{r trial-km, fig.cap="Kaplan-Meier curves: one simulated trial. Solid = overall ITT; dashed = true harm subgroup; dotted = true no-harm subgroup."}
library(survival)

# KM for overall, harm subgroup, and complementary subgroup
km_all  <- survfit(Surv(y_sim, event_sim) ~ treat_sim,
                   data = sim_data)
km_harm <- survfit(Surv(y_sim, event_sim) ~ treat_sim,
                   data = subset(sim_data, flag_harm == 1))
km_noharm <- survfit(Surv(y_sim, event_sim) ~ treat_sim,
                     data = subset(sim_data, flag_harm == 0))

# Quick summary of true vs observed HRs
cox_all   <- coxph(Surv(y_sim, event_sim) ~ treat_sim, data = sim_data)
cox_harm  <- coxph(Surv(y_sim, event_sim) ~ treat_sim,
                   data = subset(sim_data, flag_harm == 1))
cox_noharm <- coxph(Surv(y_sim, event_sim) ~ treat_sim,
                    data = subset(sim_data, flag_harm == 0))

results_df <- data.frame(
  Subgroup   = c("ITT", "Harm (true)", "No-harm (true)"),
  N          = c(nrow(sim_data),
                 sum(sim_data$flag_harm == 1),
                 sum(sim_data$flag_harm == 0)),
  True_HR    = round(c(dgm$hazard_ratios$overall,
                       dgm$hazard_ratios$harm_subgroup,
                       dgm$hazard_ratios$no_harm_subgroup), 3),
  Observed_HR = round(exp(c(coef(cox_all),
                             coef(cox_harm),
                             coef(cox_noharm))), 3)
)
print(results_df)
```


---


# Part IV — Main Results: Extreme Subgroup Effects Under a Uniform Treatment Benefit

## The Core Question

Parts I–III established how to build a null DGM and generate synthetic
trials from it.  We now use those tools of `forestsearch`: **when many subgroups are examined
in a trial with a perfectly uniform treatment benefit, how extreme can a
standard Cox analysis look by chance alone?**

We generate `r n_sims_null` synthetic trials from the null DGM and fit a
stratified Cox model in each pre-defined subgroup in every trial.  Two
output statistics are accumulated:

- The **HR point estimate distribution** across trials — summarised as
  the median and the 1st–99th percentile interval (the empirical 99%
  confidence interval, ECI).  The median should track 0.70 for every
  subgroup; the width of the ECI reveals how noisy a single-trial
  estimate is.

- The **upper confidence bound UB(HR)** — the upper end of the Cox 95%
  CI.  We track `Pr(UB < 1.0)` (power to show nominal benefit) and the
  conditional distribution of UB(HR) *given* UB ≥ 1 (how extreme a
  "failure to show benefit" looks in the worst-case trial).

The `random*` subgroups (no clinical basis, same true HR as everyone
else) serve as calibration anchors.  Any clinical subgroup of the same N
has an identical chance distribution — the clinical rationale for the
grouping provides no statistical protection.

---

## Setting: Uniform Treatment Effect

The DGM is built from the GBSG breast cancer trial using
`generate_aft_dgm_flex()`. The key feature is that the true log hazard
ratio is **constant across all patients**:

$$
\psi^0(\mathbf{L}) = \log(\mathrm{HR}_{\mathrm{true}})
  \quad \text{for all } \mathbf{L}.
$$

In the examples below the true overall HR is 0.70 (a moderate benefit).
Every subgroup — regardless of how it is defined — shares this same
underlying HR. 

```{r null-dgm}
set.seed(99)

dgm_null <- generate_aft_dgm_flex(
  data            = gbsg,
  continuous_vars = c("age", "size", "nodes", "pgr", "er"),
  factor_vars     = c("meno", "grade"),
  outcome_var     = "time_months",
  event_var       = "status",
  treatment_var   = "hormon",
  subgroup_vars   = NULL,   # no embedded subgroup — uniform effect throughout
  model           = "null",
  n_super         = 5000,
  seed            = 99,
  verbose         = FALSE
)

cat("Null DGM  |  true overall HR:",
    round(dgm_null$hazard_ratios$overall, 3),
    " |  implied median event time:",
    round(exp(dgm_null$model_params$mu), 1), "months\n")
```

---

## Subgroups Under Study

Each simulated trial analyses the same set of pre-defined subgroups,
ranging from clinically motivated subpopulations to deliberately
artificial ones.

The **`random*` subgroups** are the key benchmark. They are constructed
by randomly sampling approximately 20% of subjects and taking the first
15, 20, 40, and 60 patients from that draw:

```{r define-null-subgroups}
set.seed(8316951)
gbsg$flag_itt <- 1L

# ── Scalar cut-points for continuous variables ────────────────────────────
# Computed once from the observed data; re-exposed as columns inside each
# simulated df_s so that subset expressions like "age <= age_med" resolve.
age_med  <- median(gbsg$age)          # ~53 yrs
size_med <- median(gbsg$size)         # ~25 mm
nodes_q3 <- quantile(gbsg$nodes, 0.75)  # top-quartile node count (~5)
pgr_med  <- median(gbsg$pgr)          # median PGR
er_cut   <- 20                         # ER-low clinical threshold (fmol / mg)

subgroups <- list(

  # ── A. Full trial (ITT) ─────────────────────────────────────────────────
  list(id = "flag_itt == 1",                       name = "All Patients",              grp = "ITT"),

  # ── B. Single clinical / factor subgroups ───────────────────────────────
  list(id = "meno == 1",                           name = "Post-menopausal",           grp = "Clinical"),
  list(id = "meno == 0",                           name = "Pre-menopausal",            grp = "Clinical"),
  list(id = "grade == 3",                          name = "Grade 3",                   grp = "Clinical"),
  list(id = "grade != 3",                          name = "Grade 1/2",                 grp = "Clinical"),

  # ── C. Single continuous-variable cuts ──────────────────────────────────
  list(id = "age <= age_med",                      name = "Age (young)",               grp = "Continuous"),
  list(id = "age > age_med",                       name = "Age (older)",               grp = "Continuous"),
  list(id = "age <= 50",                           name = "Age <= 50",                 grp = "Continuous"),
  list(id = "age > 50",                            name = "Age > 50",                  grp = "Continuous"),
  list(id = "size <= size_med",                    name = "Tumour size (small)",       grp = "Continuous"),
  list(id = "size > size_med",                     name = "Tumour size (large)",       grp = "Continuous"),
  list(id = "nodes == 0",                          name = "Node-negative",             grp = "Continuous"),
  list(id = "nodes > 0 & nodes <= 3",              name = "Nodes 1-3",                grp = "Continuous"),
  list(id = "nodes > nodes_q3",                    name = "High nodes (>Q3)",          grp = "Continuous"),
  list(id = "pgr <= pgr_med",                      name = "PGR low",                  grp = "Continuous"),
  list(id = "pgr > pgr_med",                       name = "PGR high",                 grp = "Continuous"),
  list(id = "er <= er_cut",                        name = "ER-low (<20)",              grp = "Continuous"),
  list(id = "er > er_cut",                         name = "ER-high (>=20)",            grp = "Continuous"),

  # ── D. Menopausal status x Grade (2x2) ──────────────────────────────────
  list(id = "meno == 0 & grade == 3",              name = "Pre-meno / Grade 3",        grp = "Interaction: meno x grade"),
  list(id = "meno == 1 & grade == 3",              name = "Post-meno / Grade 3",       grp = "Interaction: meno x grade"),
  list(id = "meno == 0 & grade != 3",              name = "Pre-meno / Grade 1-2",      grp = "Interaction: meno x grade"),
  list(id = "meno == 1 & grade != 3",              name = "Post-meno / Grade 1-2",     grp = "Interaction: meno x grade"),

  # ── E. Menopausal status x Age ───────────────────────────────────────────
  list(id = "meno == 0 & age <= 50",               name = "Pre-meno / Age<=50",        grp = "Interaction: meno x age"),
  list(id = "meno == 0 & age > 50",                name = "Pre-meno / Age>50",         grp = "Interaction: meno x age"),
  list(id = "meno == 1 & age <= age_med",          name = "Post-meno / Age (young)",   grp = "Interaction: meno x age"),
  list(id = "meno == 1 & age > age_med",           name = "Post-meno / Age (older)",   grp = "Interaction: meno x age"),

  # ── F. Menopausal status x ER status ────────────────────────────────────
  list(id = "meno == 0 & er <= er_cut",            name = "Pre-meno / ER-low",         grp = "Interaction: meno x ER"),
  list(id = "meno == 0 & er > er_cut",             name = "Pre-meno / ER-high",        grp = "Interaction: meno x ER"),
  list(id = "meno == 1 & er <= er_cut",            name = "Post-meno / ER-low",        grp = "Interaction: meno x ER"),
  list(id = "meno == 1 & er > er_cut",             name = "Post-meno / ER-high",       grp = "Interaction: meno x ER"),

  # ── G. Grade x Node status ───────────────────────────────────────────────
  list(id = "grade == 3 & nodes == 0",             name = "Grade 3 / Node-neg",        grp = "Interaction: grade x nodes"),
  list(id = "grade == 3 & nodes > 0",              name = "Grade 3 / Node-pos",        grp = "Interaction: grade x nodes"),
  list(id = "grade != 3 & nodes == 0",             name = "Grade 1-2 / Node-neg",      grp = "Interaction: grade x nodes"),
  list(id = "grade != 3 & nodes > 0",              name = "Grade 1-2 / Node-pos",      grp = "Interaction: grade x nodes"),
  list(id = "grade == 3 & nodes > nodes_q3",       name = "Grade 3 / High nodes",      grp = "Interaction: grade x nodes"),

  # ── H. Grade x ER status ─────────────────────────────────────────────────
  list(id = "grade == 3 & er <= er_cut",           name = "Grade 3 / ER-low",          grp = "Interaction: grade x ER"),
  list(id = "grade == 3 & er > er_cut",            name = "Grade 3 / ER-high",         grp = "Interaction: grade x ER"),
  list(id = "grade != 3 & er <= er_cut",           name = "Grade 1-2 / ER-low",        grp = "Interaction: grade x ER"),
  list(id = "grade != 3 & er > er_cut",            name = "Grade 1-2 / ER-high",       grp = "Interaction: grade x ER"),

  # ── I. Grade x PGR ──────────────────────────────────────────────────────
  list(id = "grade == 3 & pgr <= pgr_med",         name = "Grade 3 / PGR low",         grp = "Interaction: grade x PGR"),
  list(id = "grade == 3 & pgr > pgr_med",          name = "Grade 3 / PGR high",        grp = "Interaction: grade x PGR"),
  list(id = "grade != 3 & pgr <= pgr_med",         name = "Grade 1-2 / PGR low",       grp = "Interaction: grade x PGR"),
  list(id = "grade != 3 & pgr > pgr_med",          name = "Grade 1-2 / PGR high",      grp = "Interaction: grade x PGR"),

  # ── J. Age x ER x Menopausal (3-way) ───────────────────────────────────
  list(id = "meno == 0 & age <= 50 & er <= er_cut",  name = "Pre-meno/Yng/ER-low",    grp = "3-way"),
  list(id = "meno == 0 & age <= 50 & er > er_cut",   name = "Pre-meno/Yng/ER-high",   grp = "3-way"),
  list(id = "meno == 1 & grade == 3 & er <= er_cut", name = "Post-meno/G3/ER-low",    grp = "3-way"),
  list(id = "meno == 1 & grade == 3 & nodes > 0",    name = "Post-meno/G3/Node-pos",  grp = "3-way"),
  list(id = "grade == 3 & nodes > 0 & er <= er_cut", name = "G3/Node-pos/ER-low",     grp = "3-way"),

  # ── K. Size x Node combinations ─────────────────────────────────────────
  list(id = "size > size_med & nodes > 0",         name = "Large/Node-pos",            grp = "Interaction: size x nodes"),
  list(id = "size > size_med & nodes == 0",        name = "Large/Node-neg",            grp = "Interaction: size x nodes"),
  list(id = "size <= size_med & nodes > 0",        name = "Small/Node-pos",            grp = "Interaction: size x nodes"),
  list(id = "size <= size_med & nodes == 0",       name = "Small/Node-neg",            grp = "Interaction: size x nodes"),

  # ── L. Random benchmark subgroups ───────────────────────────────────────
  # No clinical meaning; regenerated each simulation. Calibrate variability
  # by size against the clinical and interaction subgroups above.
  list(id = "random60 == 1",                       name = "random60",                  grp = "Random (N~60)"),
  list(id = "random40 == 1",                       name = "random40",                  grp = "Random (N~40)"),
  list(id = "random20 == 1",                       name = "random20",                  grp = "Random (N~20)"),
  list(id = "random15 == 1",                       name = "random15",                  grp = "Random (N~15)")
)

# ── Show observed sizes from GBSG ────────────────────────────────────────
# Scalar cut-points exposed as columns so eval() resolves them correctly.
gbsg_eval           <- gbsg
gbsg_eval$age_med   <- age_med
gbsg_eval$size_med  <- size_med
gbsg_eval$nodes_q3  <- nodes_q3
gbsg_eval$pgr_med   <- pgr_med
gbsg_eval$er_cut    <- er_cut
gbsg_eval$random60  <- 0L
gbsg_eval$random40  <- 0L
gbsg_eval$random20  <- 0L
gbsg_eval$random15  <- 0L

sg_sizes <- sapply(subgroups, function(s) {
  tryCatch(
    sum(eval(parse(text = s$id), envir = gbsg_eval)),
    error = function(e) NA_integer_
  )
})
is_random_sg <- grepl("^random", sapply(subgroups, `[[`, "name"))
rand_sizes   <- c(60L, 40L, 20L, 15L)

sg_overview <- data.frame(
  Subgroup = sapply(subgroups, `[[`, "name"),
  Group    = sapply(subgroups, `[[`, "grp"),
  N_source = ifelse(is_random_sg, rand_sizes[cumsum(is_random_sg)], sg_sizes),
  stringsAsFactors = FALSE
)
print(sg_overview, row.names = FALSE)
cat("
Total subgroups:", nrow(sg_overview), "
")
```

These `random*` subgroups have no clinical meaning whatsoever — their
true HR equals the ITT HR of 0.70. Any extreme result they produce is
purely a function of their small size and sampling variability. Placing
them alongside clinical subgroups of similar size makes explicit how much
of the spread in those clinical findings is attributable to noise alone.

---

## Summary Statistics Across `r n_sims_null` Simulations

For each subgroup the primary analysis — Cox model stratified by the
randomisation factor (`grade`) — is fitted in every simulated trial.
Two quantities are accumulated:

**HR point estimate distribution.** Reported as median and the 1st–99th
percentile interval (the *empirical 99% confidence interval*, ECI). Under
truth, the median should be close to 0.70 for every subgroup. The spread
of the ECI reflects how variable the estimate is — wider for smaller
subgroups.

**Upper confidence bound UB(HR).** The upper end of the Cox 95%
confidence interval. Two thresholds are tracked:

- **`Pr(UB < 1.0)`**: the proportion of trials in which the upper bound
  excludes no treatment effect, i.e., the subgroup shows a nominally
  significant benefit. For the full ITT this should be close to the
  trial power (~95%); for small subgroups it falls substantially.
- **Conditional distribution of UB(HR) given UB ≥ 1**: when the upper
  bound *fails* to exclude no effect, how large does it become? This is
  what a safety reviewer sees in the worst-case trial for that subgroup.

---

## Simulation Loop

```{r null-calibrate}
cal_null <- calibrate_cens_adjust(
  dgm           = dgm_null,
  target        = "rate",
  n             = 800,
  analysis_time = 84,
  max_entry     = 24,
  seed          = 42,
  n_eval        = 1500,
  verbose       = FALSE
)
cat("Calibrated cens_adjust:", round(cal_null$cens_adjust, 4),
    " | Target censoring rate:", round(cal_null$ref_value, 3), "\n")
```

```{r null-sims}
sg_names <- sapply(subgroups, `[[`, "name")
sg_ids   <- sapply(subgroups, `[[`, "id")
n_sg     <- length(subgroups)

# Fit stratified Cox, return c(HR point estimate, upper 95% CI bound)
cox_sR <- function(data) {
  fit <- tryCatch(
    coxph(Surv(y_sim, event_sim) ~ treat_sim + strata(grade), data = data),
    error = function(e) NULL
  )
  if (is.null(fit) || nrow(summary(fit)$conf.int) == 0)
    return(c(NA_real_, NA_real_))
  ci <- summary(fit)$conf.int[1, ]
  c(ci["exp(coef)"], ci["upper .95"])
}

# Storage: HR point estimates and upper 95% confidence bounds
sim_hrs <- matrix(NA_real_, n_sims_null, n_sg, dimnames = list(NULL, sg_names))
sim_ubs <- matrix(NA_real_, n_sims_null, n_sg, dimnames = list(NULL, sg_names))
sim_ns  <- matrix(NA_real_, n_sims_null, n_sg, dimnames = list(NULL, sg_names))

for (ss in seq_len(n_sims_null)) {

  df_s <- simulate_from_dgm(
    dgm           = dgm_null,
    n             = nrow(gbsg),
    analysis_time = 84,
    max_entry     = 24,
    cens_adjust   = cal_null$cens_adjust,
    seed          = ss
  )
  df_s$flag_itt <- 1L

  # Make quantile cut-points available as scalar variables in the eval
  # environment so expressions like "size <= size_med" resolve correctly.
  # Expose scalar cut-points as columns so subset expressions resolve.
  df_s$age_med  <- age_med
  df_s$size_med <- size_med
  df_s$nodes_q3 <- nodes_q3
  df_s$pgr_med  <- pgr_med
  df_s$er_cut   <- er_cut

  # Regenerate random benchmark subgroups for this simulated trial.
  # Random subgroups are re-sampled each iteration: their purpose is to
  # benchmark the variability of a subgroup of a given size with no true
  # treatment effect heterogeneity — exactly what each simulated trial provides.
  set.seed(ss + 1e6L)
  n_s   <- nrow(df_s)
  r_idx <- sample.int(n_s, min(60L, n_s), replace = FALSE)
  df_s$random60 <- as.integer(seq_len(n_s) %in% r_idx)
  df_s$random40 <- as.integer(seq_len(n_s) %in% r_idx[seq_len(min(40L, length(r_idx)))])
  df_s$random20 <- as.integer(seq_len(n_s) %in% r_idx[seq_len(min(20L, length(r_idx)))])
  df_s$random15 <- as.integer(seq_len(n_s) %in% r_idx[seq_len(min(15L, length(r_idx)))])

  for (gg in seq_len(n_sg)) {
    df_sg <- tryCatch(
      subset(df_s, eval(parse(text = sg_ids[[gg]]))),
      error = function(e) NULL
    )
    if (is.null(df_sg) || nrow(df_sg) < 5) next
    sim_ns[ss, gg] <- nrow(df_sg)
    r <- cox_sR(df_sg)
    sim_hrs[ss, gg] <- r[1]
    sim_ubs[ss, gg] <- r[2]
  }
}

cat("Simulations completed:", n_sims_null, "\n")
```

---

## Results

### Summary table

```{r null-results-table}
# Median HR and 1-99th percentile ECI across simulations
hr_q  <- t(apply(sim_hrs, 2, quantile, probs = c(0.01, 0.50, 0.99),
                 na.rm = TRUE))
ub_q  <- t(apply(sim_ubs, 2, quantile, probs = c(0.01, 0.50, 0.99),
                 na.rm = TRUE))

# n_valid: number of simulations where Cox estimation succeeded
n_valid     <- apply(sim_hrs, 2, function(x) sum(!is.na(x)))
pct_valid   <- round(100 * n_valid / n_sims_null, 1)

mean_n      <- round(apply(sim_ns,  2, mean, na.rm = TRUE))

# Proportions denominated by n_valid (not n_sims_null) so NAs from
# estimation failures are excluded; they are separately reported.
pr_hr_lt080 <- apply(sim_hrs, 2, function(x) mean(x < 0.80, na.rm = TRUE))
pr_ub_lt1   <- apply(sim_ubs, 2, function(x) mean(x < 1.0,  na.rm = TRUE))
pr_ub_ge1   <- 1 - pr_ub_lt1

# Conditional distribution of UB given UB >= 1
cond_ub <- t(apply(sim_ubs, 2, function(x) {
  xx <- x[!is.na(x) & x >= 1.0]
  if (length(xx) < 2) return(c(NA_real_, NA_real_, NA_real_))
  quantile(xx, probs = c(0.50, 0.75, 0.99))
}))

fmt_pct <- function(x) ifelse(is.nan(x), "-", paste0(round(100 * x, 1), "%"))
fmt_num <- function(x) ifelse(is.na(x), "-", as.character(round(x, 2)))

results_tbl <- data.frame(
  Subgroup    = sg_names,
  Mean_N      = ifelse(is.nan(mean_n), "-", as.character(mean_n)),
  N_valid     = paste0(n_valid, " (", pct_valid, "%)"),
  HR_ECI      = ifelse(is.na(hr_q[,2]), "-",
                  paste0(round(hr_q[,2],2),
                         " (", round(hr_q[,1],2), ", ",
                         round(hr_q[,3],2), ")")),
  Pr_HR_lt080 = fmt_pct(pr_hr_lt080),
  Pr_UB_lt1   = fmt_pct(pr_ub_lt1),
  Pr_UB_ge1   = fmt_pct(pr_ub_ge1),
  Med_UB_cond = fmt_num(cond_ub[, 1]),
  P99_UB_cond = fmt_num(cond_ub[, 3]),
  stringsAsFactors = FALSE
)
colnames(results_tbl) <- c(
  "Subgroup", "N", "#(% converged)",
  "Median HR (1st, 99th ECI)",
  "Pr(HR<0.80)", "Pr(UB<1.0)", "Pr(UB>=1.0)",
  "Med UB|UB>=1", "P99 UB|UB>=1"
)
print(results_tbl, row.names = FALSE)
```

**Reading the table.** Every row shares the same true HR = 0.70, so the
median of the HR ECI should be close to 0.70 throughout — confirming
the Cox model is unbiased. The *spread* of the ECI is what changes:
narrow for the ITT and large clinical subgroups, wide for `random20`
and `random15`. `Pr(UB < 1.0)` measures power for the full trial; it
falls sharply with subgroup size. `P99 UB | UB ≥ 1` answers the
question a safety reviewer faces in the worst 1% of trials: *how large
can the upper bound get when the subgroup fails to show a significant
benefit?* For `random15` the answer routinely exceeds 5.0; for
`random40` it commonly exceeds 3.0.

---

### Forest plot: HR point estimate distributions

Each row shows one subgroup. The point is the median HR across
`r n_sims_null` simulations; the line spans the 1st–99th percentile ECI.
The red dashed line marks the true HR = 0.70; the grey dotted line marks
HR = 1.0. `random*` benchmarks (orange) are shown alongside same-sized
clinical subgroups (blue) so that the comparison is immediate.

```{r forest-theme-params}
# =============================================================================
# Forest Plot Sizing Parameters — edit here to resize all four panels
# =============================================================================
#
# HOW SIZING WORKS WITH gg_forest()
# ──────────────────────────────────
# gg_forest() uses ggplot2 + patchwork.  Unlike forestploter, there is NO
# hidden scaling.  Row height is simply:
#
#     row_height_in  =  fig.height  /  n_rows
#
# So to set a target row height (e.g. 0.45 in):
#     fig.height  =  n_rows  ×  0.45  +  overhead_in
#
# where overhead_in ≈ 1.5 in covers the title, x-axis, and caption.
# ─────────────────────────────────────────────────────────────────────────────

fp_row_height   <- 0.45   # in — target height per row (increase to spread rows)
fp_overhead_in  <- 1.5    # in — fixed overhead (title + x-axis + footnote)
fp_base_size    <- 14     # pt — ggplot2 base font size (controls ALL text sizes)
fp_point_size   <- 2.2    # relative size of the median point symbol
fp_line_size    <- 0.75   # line width of CI whiskers

# ── Colour map (used by all four panels) ──────────────────────────────────
cat_cols <- c(
  "ITT"         = "black",
  "Clinical"    = "steelblue",
  "Continuous"  = "darkgreen",
  "Interaction" = "purple4",
  "3-way"       = "violetred3",
  "Random"      = "darkorange"
)

# ── Recommended fig.height for each panel ─────────────────────────────────
cat_ok  <- dplyr::case_when(
  sapply(subgroups, `[[`, "grp") == "ITT"               ~ "ITT",
  sapply(subgroups, `[[`, "grp") == "Clinical"          ~ "Clinical",
  sapply(subgroups, `[[`, "grp") == "Continuous"        ~ "Continuous",
  grepl("^Interaction", sapply(subgroups, `[[`, "grp")) ~ "Interaction",
  sapply(subgroups, `[[`, "grp") == "3-way"             ~ "3-way",
  TRUE                                                  ~ "Random"
)

ok    <- !is.na(hr_q[, 2])
ok_ub <- !is.na(ub_q[, 2])

single_cats <- c("ITT", "Clinical", "Continuous")
combo_cats  <- c("Interaction", "3-way", "Random")

idx_hr_single <- which(cat_ok[ok]  %in% single_cats)
idx_hr_combo  <- which(cat_ok[ok]  %in% combo_cats)
idx_ub_single <- which(cat_ok[ok_ub] %in% single_cats)
idx_ub_combo  <- which(cat_ok[ok_ub] %in% combo_cats)

n_single <- length(idx_hr_single)
n_combo  <- length(idx_hr_combo)

fh_single <- round(n_single * fp_row_height + fp_overhead_in, 1)
fh_combo  <- round(n_combo  * fp_row_height + fp_overhead_in, 1)

cat(sprintf(
  "Single panels: %d rows  ->  recommended fig.height = %.1f in
",
  n_single, fh_single
))
cat(sprintf(
  "Combo  panels: %d rows  ->  recommended fig.height = %.1f in
",
  n_combo,  fh_combo
))
```


**Panel A — HR distributions, single-variable subgroups.**
Point = median HR; whiskers = 1st–99th percentile ECI.
The true HR is 0.70 for every subgroup (red dashed line).

```{r hr-forest-single, fig.height = fh_single, fig.width = 14, fig.cap = "HR distribution — single-variable subgroups (log scale). Spread reflects sampling variability alone; all subgroups share the same true HR = 0.70."}
sg_single   <- sg_names[ok][idx_hr_single]
cat_single  <- cat_ok[ok][idx_hr_single]

gg_forest(
  subgroups   = sg_single,
  est         = hr_q[ok, 2][idx_hr_single],
  lo          = hr_q[ok, 1][idx_hr_single],
  hi          = hr_q[ok, 3][idx_hr_single],
  cat_vec     = cat_single,
  cat_colours = cat_cols,
  annot       = list(
    "N"     = as.character(mean_n[ok][idx_hr_single]),
    "Pr(UB<1.0)" = paste0(round(100 * pr_ub_lt1[ok][idx_hr_single], 1), "%")
  ),
  ref_line    = 0.70,
  vert_lines  = 1.00,
  ref_col     = "firebrick",
  vert_col    = "grey55",
  vert_lty    = "dotted",
  xlim        = c(0.15, 3.5),
  ticks_at    = c(0.20, 0.35, 0.50, 0.70, 1.00, 1.50, 2.50),
  xlog        = TRUE,
  xlab        = "Hazard Ratio (Cox, stratified by grade)",
  footnote    = paste0(
    "Median HR (point) + 1st-99th percentile ECI  |  ",
    n_sims_null, " simulated trials, N = ", nrow(gbsg), " per trial  |  ",
    "True HR = 0.70 (red dashed)"
  ),
  point_size  = fp_point_size,
  line_size   = fp_line_size,
  base_size   = fp_base_size,
  widths      = c(3.5, 5, 1.1, 1.2)
)
```

**Panel B — HR distributions, combination and random-benchmark subgroups.**
Note how the ECI widens sharply as subgroup N falls below ~60; the
`random*` benchmarks (orange) define the noise floor for each size class.

```{r hr-forest-combo, fig.height = fh_combo, fig.width = 14, fig.cap = "HR distribution — combination subgroups and random benchmarks (log scale). Random benchmarks of matching size (orange) are the baseline for chance variability."}
sg_combo   <- sg_names[ok][idx_hr_combo]
cat_combo  <- cat_ok[ok][idx_hr_combo]

gg_forest(
  subgroups   = sg_combo,
  est         = hr_q[ok, 2][idx_hr_combo],
  lo          = hr_q[ok, 1][idx_hr_combo],
  hi          = hr_q[ok, 3][idx_hr_combo],
  cat_vec     = cat_combo,
  cat_colours = cat_cols,
  annot       = list(
    "N"     = as.character(mean_n[ok][idx_hr_combo]),
    "Pr(UB<1.0)" = paste0(round(100 * pr_ub_lt1[ok][idx_hr_combo], 1), "%")
  ),
  ref_line    = 0.70,
  vert_lines  = 1.00,
  ref_col     = "firebrick",
  vert_col    = "grey55",
  vert_lty    = "dotted",
  xlim        = c(0.15, 3.5),
  ticks_at    = c(0.20, 0.35, 0.50, 0.70, 1.00, 1.50, 2.50),
  xlog        = TRUE,
  xlab        = "Hazard Ratio (Cox, stratified by grade)",
  footnote    = paste0(
    "Median HR (point) + 1st-99th percentile ECI  |  ",
    n_sims_null, " simulated trials  |  ",
    "True HR = 0.70 (red dashed)"
  ),
  point_size  = fp_point_size,
  line_size   = fp_line_size,
  base_size   = fp_base_size,
  widths      = c(3.5, 5, 1.1, 1.2)
)
```

---

### Forest plot: UB(HR) distributions

The upper confidence bound UB(HR) is the quantity a safety reviewer
examines when asking whether a subgroup demonstrates harm or lack of
benefit. The plot below shows the distribution of UB(HR) across
simulations in the same ECI format. Values above 1.0 (grey dotted line)
represent trials where the confidence interval includes no treatment
effect. The right-hand column shows the probability of this occurring
(`Pr(UB >= 1)`) and the conditional median UB given failure to show
benefit (`Med UB | UB>=1`).

**Panel A — UB(HR) distributions, single-variable subgroups.**

```{r ub-forest-single, fig.height = fh_single, fig.width = 14, fig.cap = "UB(HR) distribution — single-variable subgroups (log scale). Values right of HR = 1.0 (dotted line) represent trials failing to show significant benefit."}
lab_cmed_ub <- ifelse(
  is.na(cond_ub[ok_ub, 1]), "-",
  as.character(round(cond_ub[ok_ub, 1], 2))
)
sg_ub_s   <- sg_names[ok_ub][idx_ub_single]
cat_ub_s  <- cat_ok[ok_ub][idx_ub_single]

gg_forest(
  subgroups   = sg_ub_s,
  est         = ub_q[ok_ub, 2][idx_ub_single],
  lo          = ub_q[ok_ub, 1][idx_ub_single],
  hi          = ub_q[ok_ub, 3][idx_ub_single],
  cat_vec     = cat_ub_s,
  cat_colours = cat_cols,
  annot       = list(
    "N"         = as.character(mean_n[ok_ub][idx_ub_single]),
    "UB>=1"    = paste0(round(100 * pr_ub_ge1[ok_ub][idx_ub_single], 1), "%"),
    "mUB | UB>=1" = lab_cmed_ub[idx_ub_single]
  ),
  ref_line    = 1.00,
  vert_lines  = 0.70,
  ref_col     = "grey55",
  ref_lty     = "dotted",
  vert_col    = "firebrick",
  vert_lty    = "dashed",
  xlim        = c(0.30, 9.0),
  ticks_at    = c(0.40, 0.70, 1.00, 1.50, 2.50, 4.00, 8.00),
  xlog        = TRUE,
  xlab        = "Upper Bound of 95% CI  [UB(HR)]",
  footnote    = paste0(
    "Median UB(HR) (point) + 1st-99th percentile ECI  |  ",
    n_sims_null, " simulated trials  |  ",
    "Dotted line: UB = 1.0;  Dashed line: true HR = 0.70"
  ),
  point_size  = fp_point_size,
  line_size   = fp_line_size,
  base_size   = fp_base_size,
  widths      = c(3.5, 5, 1.1, 1.2, 1.5)
)
```

**Panel B — UB(HR) distributions, combination subgroups and random benchmarks.**
The `Med UB | UB>=1` column is the key safety quantity — it shows how
large the upper bound grows in those trials where the CI actually fails
to demonstrate benefit.

```{r ub-forest-combo, fig.height = fh_combo, fig.width = 14, fig.cap = "UB(HR) distribution — combination and random-benchmark subgroups. The widening ECI for small subgroups is the central message: chance alone produces UB(HR) > 3-5 for N ~ 15-40."}
sg_ub_c   <- sg_names[ok_ub][idx_ub_combo]
cat_ub_c  <- cat_ok[ok_ub][idx_ub_combo]

gg_forest(
  subgroups   = sg_ub_c,
  est         = ub_q[ok_ub, 2][idx_ub_combo],
  lo          = ub_q[ok_ub, 1][idx_ub_combo],
  hi          = ub_q[ok_ub, 3][idx_ub_combo],
  cat_vec     = cat_ub_c,
  cat_colours = cat_cols,
  annot       = list(
    "N"         = as.character(mean_n[ok_ub][idx_ub_combo]),
    "UB>=1"    = paste0(round(100 * pr_ub_ge1[ok_ub][idx_ub_combo], 1), "%"),
    "mUB | UB>=1" = lab_cmed_ub[idx_ub_combo]
  ),
  ref_line    = 1.00,
  vert_lines  = 0.70,
  ref_col     = "grey55",
  ref_lty     = "dotted",
  vert_col    = "firebrick",
  vert_lty    = "dashed",
  xlim        = c(0.30, 9.0),
  ticks_at    = c(0.40, 0.70, 1.00, 1.50, 2.50, 4.00, 8.00),
  xlog        = TRUE,
  xlab        = "Upper Bound of 95% CI  [UB(HR)]",
  footnote    = paste0(
    "Median UB(HR) (point) + 1st-99th percentile ECI  |  ",
    n_sims_null, " simulated trials  |  ",
    "Dotted line: UB = 1.0;  Dashed line: true HR = 0.70"
  ),
  point_size  = fp_point_size,
  line_size   = fp_line_size,
  base_size   = fp_base_size,
  widths      = c(3.5, 5, 1.1, 1.2, 1.5)
)
```

---

## Key Finding: Small Subgroups Produce Extreme Results by Chance

The forest plots above tell a consistent story.

**Point estimates (HR).** For the ITT and large clinical subgroups the
median HR tracks the truth (≈ 0.70) and the ECI is narrow. For
`random15` and `random20` the ECI spans from roughly 0.25 to above
1.0 — the entire plausible range. Clinical subgroups of similar size
(e.g. Grade 3, Age ≤ 50) show ECI widths that are indistinguishable
from their random-benchmark counterparts of the same N. The randomisation
basis of those clinical groupings provides no protection against this
variability.

**Upper bounds UB(HR).** The upper-bound distribution is even more
striking. For `random40` (N ≈ 40), the 99th percentile of UB(HR)
commonly exceeds 3.0–4.0, meaning that 1% of simulated trials produce
an upper confidence bound above 3.0 for a subgroup where the true HR
is 0.70. For `random15`, values above 5.0 are not rare.

**This is not a failure of the Cox model** — it is working exactly as
intended. A 95% confidence interval that spans from 0.3 to 4.5 is
statistically correct: with N = 15 there is genuinely that much
uncertainty. The problem is the *interpretation*: a practitioner seeing
UB(HR) = 3.5 in a subgroup of 15 patients in a single trial has no way
to distinguish a chance fluctuation from a genuine safety signal without
additional structure.

---

---

# Summary

| Step | Function | Purpose |
|:-----|:---------|:--------|
| 1 | — | Convert source data to consistent time scale |
| 2 | `generate_aft_dgm_flex()` | Fit AFT model; embed harm subgroup; build super population |
| 3 | `check_censoring_dgm()` | Verify simulated censoring matches source data |
| 4 | `calibrate_cens_adjust()` | Find `cens_adjust` that aligns censoring rate or KM median |
| 5 | `simulate_from_dgm()` | Draw individual synthetic trials from the DGM |
| 6 | `forestsearch()` + summaries | Apply method and accumulate operating characteristics |

The extreme-subgroup evaluation in Part IV answers a question: *under a null where every patient shares the same true HR,
how alarming can a standard Cox subgroup analysis look by chance?* The
`random*` subgroups anchor the answer — any clinical subgroup of similar
size has an identical chance distribution of HR and UB(HR) regardless of
clinical rationale. 
---

# Session Information

```{r session-info}
sessionInfo()
```
