---
title: "Acceptance: `betaHhat_truth_md.R`"
subtitle: "Six criteria from `SPEC_betaHhat_md.md`, run against a calibrated ACTG175 continuous DGM"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")
options(width = 100)
```

# What this document is

An integration-style living document, not a `testthat` suite. It runs the six
acceptance criteria in `dev/glm-continuous-sims/SPEC_betaHhat_md.md` against
the module at
`quarto/simulations/actg175/continuous/betaHhat_truth_md.R`.

The output that matters is the **numbers**, not "all assertions pass". Every
criterion below prints its measured quantities before any assertion is
evaluated, and the pass/fail gate is deferred to the last section so that a
failure anywhere still leaves the full set of measurements visible.

This is the first contact between these closed forms and `forestsearch()`.

## Two structural facts this document leans on

1. **The mean difference is collapsible.** Under randomisation, for any region
   $R$, $\beta(R) = \mathbb E[\tau(X)\mid R]$. `df_super` is a fixed finite
   population, so $\beta(\widehat H)$ is an exact finite mean over it — no
   evaluation frame, no model fit, no residual Monte-Carlo noise.
2. **The subject-level effect is two-valued.**
   $\mu_{1,i} = \mu_{0,i} + \delta + \beta_{\mathrm{inter}}\mathbb 1\{i\in Q\}$,
   with $\delta := k_{\mathrm{treat}}\beta_{\mathrm{treat}}$ constant because
   the base fit is additive (`R/generate_glm_dgm.R:310-323`). Hence
   $\beta(\widehat H) = \delta + \beta_{\mathrm{inter}}\,\mathbb P(Q\mid\widehat H)$,
   and the originating handoff's $\beta_{\mathrm{inter}}\cdot\mathbb P(Q\mid\widehat H)$
   is low by $\delta$ on every region.

Both are established in `dev/glm-continuous-sims/NOTE_target_is_collapsibility.md`.
**Do not take numbers from `HANDOFF_glm_continuous_simulations.md`.**

## A tension in the SPEC, resolved here explicitly

Criterion 3 requires a region **disjoint from $Q$**, whose target is exactly
$\delta$. Criterion 5 requires that **every** region have a target
distinguishable from $0$, from $\delta$, and from $\delta+\beta_{\mathrm{inter}}$.
Taken literally these cannot both hold: the disjoint region's target *is*
$\delta$, and a region equal to (or inside) $Q$ has target exactly
$\delta+\beta_{\mathrm{inter}}$.

Resolution used below, stated before running: criterion 5's distinguishability
requirement is applied to the **interior** regions, those with
$0 < \mathbb P(Q\mid R) < 1$. The two boundary cases are not accidents to be
excluded but the pinned endpoints that criteria 3 and 4 exist to check. Both
kinds are exercised; only the interior ones carry the distinguishability
assertion.

```{r load-packages}
library(forestsearch)
library(speff2trial)
library(ggplot2)

module_path <- "../../../quarto/simulations/actg175/continuous/betaHhat_truth_md.R"
stopifnot(file.exists(module_path))
source(module_path)

results_log <- list()
timings     <- list()
```

# The fixture

## DGM settings

Reused verbatim from
`quarto/simulations/actg175/continuous/actg175_continuous_simulations.qmd`.
This is a fixture, not a sweep: one calibrated DGM.

The vignette's calibration-curve chunk (a 61-point `k_inter` sweep, each point
a full `generate_glm_dgm()` fit) is **skipped**. It is presentational and
scores no criterion.

```{r dgm-settings}
actg_arms         <- c(1L, 3L)
actg_treat_arm    <- 1L
actg_age_cut      <- 34
actg_preanti_cut  <- 744.5

dgm_factor_vars   <- paste0("z", 1:12)
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)
dgm_n_super       <- 5000L

cal_target_md     <- 40
cal_k_grid_range  <- c(0, 120)
cal_grid_step     <- 2

sim_seed          <- 8316951L

analysis_continuous_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
analysis_binary_vars     <- c("hemo", "homo", "drugs", "race", "gender", "symptom")
```

## ACTG175 data preparation

```{r prepare-data}
actg_df <- subset(speff2trial::ACTG175, arms %in% actg_arms)
actg_df$id <- seq_len(nrow(actg_df))

actg_df$treat_orig <- ifelse(actg_df$arms == actg_treat_arm, 1L, 0L)
# SWITCHED roles: ddI = treat = 1 (benefit search via treatment switching)
actg_df$treat      <- 1L - actg_df$treat_orig
actg_df$cd4_change <- actg_df$cd420 - actg_df$cd40
actg_df <- actg_df[!is.na(actg_df$cd420), ]

actg_df$z1  <- as.factor(ifelse(actg_df$age > actg_age_cut, 1L, 0L))
actg_df$z2  <- as.factor(ifelse(actg_df$preanti <= actg_preanti_cut, 1L, 0L))
actg_df$z3  <- as.factor(ifelse(actg_df$wtkg <= 75, 1L, 0L))
actg_df$z4  <- as.factor(ifelse(actg_df$karnof <= median(actg_df$karnof), 1L, 0L))
actg_df$z5  <- as.factor(ifelse(actg_df$cd40 <= median(actg_df$cd40), 1L, 0L))
actg_df$z6  <- as.factor(ifelse(actg_df$cd80 <= median(actg_df$cd80), 1L, 0L))
actg_df$z7  <- as.factor(actg_df$hemo)
actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs)
actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender)
actg_df$z12 <- as.factor(actg_df$symptom)

for (v in analysis_binary_vars) actg_df[[v]] <- as.factor(actg_df[[v]])
confounders_analysis <- c(analysis_continuous_vars, analysis_binary_vars)

cat(sprintf("ACTG175 (arms %s): n = %d\n",
            paste(actg_arms, collapse = " vs "), nrow(actg_df)))
cat(sprintf("  cd4_change: mean = %.2f, sd = %.2f\n",
            mean(actg_df$cd4_change), sd(actg_df$cd4_change)))
```

## Calibrated DGM

Built on every render. No cache: `calibrate_glm_interaction()` uses `uniroot()`
with interval extension, not a grid search (`R/calibrate_glm_interaction.R:143-146`),
and for MD the relation $\text{effect}_Q = \delta + k_{\mathrm{inter}}$ is
affine, so it converges in a few steps. Measured build time is printed below;
it does not justify a cache path.

```{r build-dgm}
t0 <- proc.time()[3]

dgm_calibrated <- calibrate_glm_interaction(
    data           = actg_df,
    factor_vars    = dgm_factor_vars,
    outcome_var    = "cd4_change",
    treatment_var  = "treat",
    target_effect  = cal_target_md,
    outcome_type   = "continuous",
    effect_measure = "MD",
    subgroup_vars  = dgm_subgroup_vars,
    subgroup_cuts  = dgm_subgroup_cuts,
    k_inter_range  = cal_k_grid_range,
    grid_step      = cal_grid_step,
    n_super        = dgm_n_super,
    seed           = sim_seed,
    verbose        = TRUE
)

timings$dgm <- proc.time()[3] - t0
cat(sprintf("\nDGM built in %.2f sec (no cache)\n", timings$dgm))

df_super <- dgm_calibrated$df_super
cat(sprintf("df_super: n = %d,  Q prevalence = %.4f\n",
            nrow(df_super), mean(df_super$flag_harm == 1L)))
cat(sprintf("NAs in mu0: %d,  NAs in mu1: %d\n",
            sum(is.na(df_super$mu0)), sum(is.na(df_super$mu1))))
```

## Derived quantities

All three are read off the calibrated object and `df_super`. **None is
hard-coded.**

```{r derived-quantities}
inQ <- df_super$flag_harm == 1L

# delta: the complement effect, via compute_aor() on Q^c
delta <- compute_aor(df_super,
                     subset_indicator = as.integer(!inQ),
                     effect_measure   = "MD")

# beta_inter: the calibrated interaction shift
beta_inter <- dgm_calibrated$model_params$beta_inter
k_inter    <- dgm_calibrated$model_params$k_inter

# P(Q | region): counted on df_super (function; used per-region below)
p_Q_given <- function(idx) mean(inQ[idx])

effect_Q  <- dgm_calibrated$hazard_ratios$harm_subgroup
effect_Qc <- dgm_calibrated$hazard_ratios$no_harm_subgroup

cat(sprintf("delta       (compute_aor on Q^c) = %.6f\n", delta))
cat(sprintf("beta_inter  (model_params)       = %.6f\n", beta_inter))
cat(sprintf("k_inter     (model_params)       = %.6f\n", k_inter))
cat(sprintf("effect_Q    (DGM .effect_one)    = %.6f\n", effect_Q))
cat(sprintf("effect_Qc   (DGM .effect_one)    = %.6f\n", effect_Qc))
cat(sprintf("delta + beta_inter               = %.6f\n", delta + beta_inter))
cat(sprintf("  vs effect_Q, difference        = %.3e\n",
            (delta + beta_inter) - effect_Q))
```

# The regions

Six rules on `df_super`, built from the **raw** covariates (not the discretised
`z` factors) so `get_dfpred()` resolves them by comparison. $Q$ is
$\{ \texttt{age} > 34 \} \cap \{ \texttt{preanti} \le 744.5 \}$.

Shapes covered, as criterion 5 requires: a conjunction, a strict subset of $Q$,
a single cut, a **disjunction** in GRF form, a **negation**, and a region
**disjoint from $Q$**.

```{r define-regions}
rules <- list(
  exact_Q     = c("{age > 34}", "{preanti <= 744.5}"),
  narrow      = c("{age > 34}", "{preanti <= 744.5}", "{wtkg <= 75}"),
  one_cut     = c("{age > 34}"),
  disjunction = "(age > 34 & preanti <= 744.5) | (wtkg <= 60)",
  negation    = c("!{cd40 <= 400}"),
  disjoint_Q  = c("{age <= 34}")
)

# Mirrors betaHhat_one_md()'s dispatch order exactly: disjunction form FIRST,
# on the unsplit string.  See the module's comment and the discussion section.
.resolve_rule <- function(r) {
  if (length(r) == 1L && grepl("|", r, fixed = TRUE)) r
  else if (length(r) == 1L && grepl(" & ", r, fixed = TRUE))
    strsplit(r, " & ", fixed = TRUE)[[1L]]
  else r
}

region_idx <- lapply(rules, function(r)
  get_dfpred(df_super, .resolve_rule(r))$treat.recommend == 0L)

region_tab <- data.frame(
  region   = names(rules),
  n        = vapply(region_idx, sum, integer(1)),
  P_Q      = vapply(region_idx, p_Q_given, numeric(1)),
  exact    = vapply(region_idx, function(i) .beta_region_md(df_super, i), numeric(1)),
  row.names = NULL
)
region_tab$closed_form <- delta + beta_inter * region_tab$P_Q
region_tab$handoff     <- beta_inter * region_tab$P_Q
region_tab$interior    <- region_tab$P_Q > 0 & region_tab$P_Q < 1

knitr::kable(region_tab, digits = 6,
             caption = "Regions on df_super. `exact` is the module's target; `closed_form` is delta + beta_inter*P(Q|R); `handoff` is the originating handoff's form.")
```

# Criterion 1 — exactness

Repeated evaluation of `.beta_region_md()` on one region must be
**bit-identical**. Asserted as `max - min == 0`, not a tolerance.

```{r criterion-1}
reps <- vapply(1:25, function(i) .beta_region_md(df_super, region_idx$one_cut),
               numeric(1))

c1_spread <- max(reps) - min(reps)
c1_unique <- length(unique(reps))

cat(sprintf("25 evaluations of .beta_region_md() on region 'one_cut'\n"))
cat(sprintf("  value          = %.15f\n", reps[1]))
cat(sprintf("  max - min      = %.3e   (identical zero required)\n", c1_spread))
cat(sprintf("  distinct values= %d\n", c1_unique))
cat(sprintf("  identical()    = %s\n", identical(reps, rep(reps[1], 25))))

results_log$c1 <- (c1_spread == 0) && (c1_unique == 1L)
```

# Criterion 2 — agreement with a fitted MD, shrinking at $n^{-1/2}$

**20 independent draws per size**, pre-declared and all reported. Within each
draw the fitted $\bar Y_1 - \bar Y_0$ on each interior region is compared
against the module's target and the RMS taken across regions; the *mean* RMS
across draws is then regressed on $\log n$. A single comparison at one $n$
cannot distinguish Monte-Carlo error from a small systematic offset — the slope
is the discriminating measurement.

::: {.callout-important title="Why one draw per size was not enough"}
The regions overlap and, within a draw, are scored on the **same** realization,
so their per-region differences are strongly positively correlated. Averaging
across regions therefore does **not** reduce draw-level noise: at one draw per
size there were effectively six independent observations in total, not six
times the number of regions.

That design produced slope $-0.237$ with 95% CI $[-1.158, 0.684]$ and
$R^2 = 0.11$. The interval contained $-0.5$, but it also contained $0$ — so the
check could not reject "no decay at all", which makes it vacuous by this
workstream's own standard. The band was **not** widened to accommodate it; the
number of draws was raised, which changes the fixture's cost rather than the
evidentiary bar. The band remains $(-0.85, -0.25)$.
:::

```{r criterion-2}
t0 <- proc.time()[3]
n_grid <- c(2000L, 4000L, 8000L, 16000L, 32000L, 64000L)

fitted_md_region <- function(df_sim, rule) {
  i <- get_dfpred(df_sim, .resolve_rule(rule))$treat.recommend == 0L
  d  <- df_sim[i, , drop = FALSE]
  if (sum(d$treat_sim == 1L) < 5L || sum(d$treat_sim == 0L) < 5L) return(NA_real_)
  mean(d$y_sim[d$treat_sim == 1L]) - mean(d$y_sim[d$treat_sim == 0L])
}

n_draws <- 20L   # PRE-DECLARED: 20 independent draws per size, all reported

interior_names <- region_tab$region[region_tab$interior]

c2_draws <- do.call(rbind, lapply(seq_along(n_grid), function(j) {
  n <- n_grid[j]
  do.call(rbind, lapply(seq_len(n_draws), function(b) {
    df_sim <- simulate_from_glm_dgm(dgm_calibrated, n = n,
                                    seed = sim_seed + 1000L * j + b)
    d2 <- vapply(interior_names, function(rn)
      fitted_md_region(df_sim, rules[[rn]]) -
        region_tab$exact[region_tab$region == rn], numeric(1))
    data.frame(n = n, draw = b, rms = sqrt(mean(d2^2, na.rm = TRUE)))
  }))
}))

agg <- aggregate(rms ~ n, data = c2_draws,
                 FUN = function(x) c(mean = mean(x), se = sd(x) / sqrt(length(x))))
c2_rms <- data.frame(n = agg$n, mean_rms = agg$rms[, "mean"], se = agg$rms[, "se"])

c2_fit   <- lm(log(mean_rms) ~ log(n), data = c2_rms)
c2_slope <- unname(coef(c2_fit)[2])
c2_se    <- summary(c2_fit)$coefficients[2, 2]
c2_ci    <- c2_slope + c(-1, 1) * qt(0.975, df = nrow(c2_rms) - 2) * c2_se

knitr::kable(c2_rms, digits = 4,
             caption = sprintf("Mean RMS |fitted - target| across interior regions, %d draws per size, with SE.",
                               n_draws))

cat(sprintf("\ndraws per size = %d (pre-declared), sizes = %d, total draws = %d\n",
            n_draws, nrow(c2_rms), nrow(c2_draws)))
cat(sprintf("slope of log(mean_rms) on log(n) = %.4f   (predicted -0.5)\n", c2_slope))
cat(sprintf("  SE = %.4f,  95%% CI = [%.4f, %.4f],  R^2 = %.4f\n",
            c2_se, c2_ci[1], c2_ci[2], summary(c2_fit)$r.squared))
cat(sprintf("  CI excludes 0 (i.e. rejects 'no decay'): %s\n",
            !(c2_ci[1] < 0 && c2_ci[2] > 0)))
cat(sprintf("  monotone decreasing in n: %s\n", all(diff(c2_rms$mean_rms) < 0)))

timings$c2 <- proc.time()[3] - t0

results_log$c2 <- is.finite(c2_slope) && c2_slope < -0.25 && c2_slope > -0.85
```

## The pre-registered alternation check

At one draw per size the six RMS values alternated cleanly — 8.51, 1.00, 5.44,
1.11, 4.48, 1.51 — and the low subsequence (1.00, 1.11, 1.51) *increased* with
$n$. That is more structure than one-draw noise should produce, so it was
pre-registered as something to watch: **surviving 20 draws would make it a seed
or design artifact and a finding in its own right; vanishing would make it
noise.**

```{r criterion-2-alternation}
lo <- c2_rms$mean_rms[c2_rms$n %in% c(4000L, 16000L, 64000L)]
hi <- c2_rms$mean_rms[c2_rms$n %in% c(2000L, 8000L, 32000L)]

cat("Former LOW subsequence  (n = 4000, 16000, 64000):\n")
cat(sprintf("  one draw : 1.0029, 1.1123, 1.5080   <- increasing in n\n"))
cat(sprintf("  %2d draws : %.4f, %.4f, %.4f\n", n_draws, lo[1], lo[2], lo[3]))
cat(sprintf("  still increasing? %s\n", all(diff(lo) > 0)))

cat("\nFormer HIGH subsequence (n = 2000, 8000, 32000):\n")
cat(sprintf("  one draw : 8.5094, 5.4433, 4.4819\n"))
cat(sprintf("  %2d draws : %.4f, %.4f, %.4f\n", n_draws, hi[1], hi[2], hi[3]))

cat(sprintf("\nAlternation gone (whole sequence monotone decreasing)? %s\n",
            all(diff(c2_rms$mean_rms) < 0)))
```

**Verdict: it was noise.** At 20 draws per size the whole sequence is monotone
decreasing, the former low subsequence decreases rather than increases, and the
slope tightens onto $-1/2$. No seed or design artifact survives. The alternation
was an artifact of scoring correlated regions on a single shared draw per size.

# Criterion 3 — the wrong form fails

The handoff's $\beta_{\mathrm{inter}}\cdot\mathbb P(Q\mid\widehat H)$ must differ
from the exact target by $\delta$ on **every** region, including the region
disjoint from $Q$ where it gives $0$ against a truth of $\delta$.

```{r criterion-3}
c3 <- data.frame(
  region       = region_tab$region,
  P_Q          = region_tab$P_Q,
  exact        = region_tab$exact,
  handoff      = region_tab$handoff,
  gap          = region_tab$exact - region_tab$handoff,
  gap_minus_d  = (region_tab$exact - region_tab$handoff) - delta,
  stringsAsFactors = FALSE
)

knitr::kable(c3, digits = 6,
             caption = "Every `gap` should equal delta; `gap_minus_d` should be identically zero.")

cat(sprintf("\ndelta = %.6f\n", delta))
cat(sprintf("max |gap - delta| over all regions = %.3e\n",
            max(abs(c3$gap_minus_d))))

dj <- c3[c3$region == "disjoint_Q", ]
cat(sprintf("\ndisjoint-from-Q region: P(Q|R) = %.6f\n", dj$P_Q))
cat(sprintf("  handoff form gives %.6f;  truth is %.6f (= delta)\n",
            dj$handoff, dj$exact))

results_log$c3 <- max(abs(c3$gap_minus_d)) < 1e-9 &&
                  dj$P_Q == 0 && abs(dj$handoff) < 1e-12
```

# Criterion 4 — the $\theta^{\dagger}$ identity, plus an external anchor

## Pre-registered failure mode

Stated **before running**. The internal half of this criterion compares
`compute_aor()` against the DGM's own `.effect_one()`
(`R/generate_glm_dgm.R:568-594`). These are two package sites, so that half is
package-against-package and cannot detect a shared error.

The two sites differ in exactly one respect: `compute_aor()` uses
`mean(..., na.rm = TRUE)` and subsets the extracted vectors, while
`.effect_one()` uses bare `mean(df$mu1)` on a subset of the frame. Their
arithmetic is otherwise the same expression. **Therefore a nonzero difference
here points at `NA`s in `mu0`/`mu1`, not at the target logic.** If this
criterion fails, look at the NA counts printed in the fixture section first.

## External anchor

Because of the above, the internal identity is paired with an anchor outside
the package's own arithmetic: the calibrator was asked for
MD$(Q) = `r cal_target_md`$ and stops if it cannot get within `tol_rel = 2.5`
percent (`R/calibrate_glm_interaction.R:53-55,105`). So `effect_Q` must
reproduce the calibration target within that band.

::: {.callout-warning title="This anchor is weaker than it reads — stated before running"}
The calibrator uses `uniroot()` with interval extension, not a grid snap
(`R/calibrate_glm_interaction.R:143-146`). For MD the relation is **affine**,
$\text{effect}_Q(k_{\mathrm{inter}}) = \delta + k_{\mathrm{inter}}$, so
`uniroot` converges to machine precision in a few steps — and its objective
function *is* `.effect_one()`. So "within 2.5 percent" measures the solver's
convergence on the package's own arithmetic. It confirms the fixture realizes
the analyst's stated target; it does **not** independently validate the MD
computation.

A second anchor is therefore added that touches none of the package's effect
arithmetic: fit `stats::lm(y_sim ~ treat_sim)` on $Q$ in a large drawn trial
and compare the coefficient to `cal_target_md`. That path goes through the
simulator and base R least squares only. It carries Monte-Carlo error, so it is
checked against the same 2.5 percent band rather than as an identity.
:::

```{r criterion-4}
td <- betaHhat_theta_dagger_check_md(df_super)

c4 <- data.frame(
  quantity = c("theta_dagger_H", "theta_dagger_Hc"),
  module   = c(unname(td["thetaDagger_H"]), unname(td["thetaDagger_Hc"])),
  dgm      = c(effect_Q, effect_Qc),
  stringsAsFactors = FALSE
)
c4$difference <- c4$module - c4$dgm

knitr::kable(c4, digits = 10,
             caption = "Internal identity: module vs the DGM's own effect_Q / effect_Qc. Exact zero required.")

cat(sprintf("\nmax |difference| = %.3e   (exact identity required, not a tolerance)\n",
            max(abs(c4$difference))))
cat(sprintf("NA counts in mu0 / mu1: %d / %d\n",
            sum(is.na(df_super$mu0)), sum(is.na(df_super$mu1))))

# --- external anchor ---
rel_err <- 100 * abs(effect_Q - cal_target_md) / abs(cal_target_md)
cat(sprintf("\nExternal anchor:\n"))
cat(sprintf("  calibration target MD(Q) = %.4f\n", cal_target_md))
cat(sprintf("  realized      effect_Q   = %.6f\n", effect_Q))
cat(sprintf("  relative error           = %.4f%%   (calibrator band: 2.5%%)\n",
            rel_err))

results_log$c4        <- max(abs(c4$difference)) == 0
results_log$c4_anchor <- rel_err <= 2.5
```

```{r criterion-4-lm-anchor}
n_anchor <- 200000L
df_anchor <- simulate_from_glm_dgm(dgm_calibrated, n = n_anchor,
                                   seed = sim_seed + 55L)
fit_anchor <- stats::lm(y_sim ~ treat_sim,
                        data = df_anchor[df_anchor$flag_harm == 1L, ])
md_lm <- unname(stats::coef(fit_anchor)[["treat_sim"]])
rel_err_lm <- 100 * abs(md_lm - cal_target_md) / abs(cal_target_md)

cat(sprintf("External anchor #2 -- stats::lm, no package effect arithmetic:\n"))
cat(sprintf("  n                        = %d\n", n_anchor))
cat(sprintf("  lm MD(Q)                 = %.6f\n", md_lm))
cat(sprintf("  calibration target       = %.4f\n", cal_target_md))
cat(sprintf("  relative error           = %.4f%%   (band: 2.5%%)\n", rel_err_lm))
cat(sprintf("  vs effect_Q, difference  = %.6f  (Monte-Carlo error)\n",
            md_lm - effect_Q))

results_log$c4_anchor_lm <- rel_err_lm <= 2.5
```

# Criterion 5 — rule shapes

The closed form is shape-free; this demonstrates that rather than assuming it.
Interior regions ($0 < \mathbb P(Q\mid R) < 1$) must have targets
distinguishable from $0$, from $\delta$, and from $\delta+\beta_{\mathrm{inter}}$
— see the resolution stated at the top of this document for why the two
boundary regions are exempt.

::: {.callout-note title="How the negation region was chosen"}
The negation region is `!{cd40 <= 400}`, selected deliberately rather than
taken as the first thing to hand. Three conjunctive negations were evaluated
first and rejected **because their targets collided**, which would have made
the assertion vacuous in exactly the way criterion 5 exists to prevent:

| candidate | target | nearest other region |
|---|---|---|
| `!{cd40 <= 400} & {preanti <= 744.5}` | 1.87 | 1.87 from $0$ |
| `!{cd40 <= 400} & !{age <= 34}` | 21.77 | **0.34** from `one_cut` |
| `!{cd80 <= 900} & {age > 34}` | 20.65 | **0.78** from `one_cut` |
| `!{cd40 <= 400}` (**used**) | −4.24 | 25.67 from `one_cut` |

This is the caution the workstream README carries about
`check_betaHhat_md.R`, whose negation row is byte-identical to its exact row
and so discriminates nothing. A negation region whose target lands on top of
another region's tests the arithmetic, not the negation handling.
:::

```{r criterion-5}
sep <- function(x, y) abs(x - y)

c5 <- data.frame(
  region      = region_tab$region,
  shape       = c("conjunction", "subset of Q", "single cut",
                  "disjunction (GRF)", "negation", "disjoint from Q"),
  P_Q         = region_tab$P_Q,
  target      = region_tab$exact,
  interior    = region_tab$interior,
  d_from_0    = sep(region_tab$exact, 0),
  d_from_delta= sep(region_tab$exact, delta),
  d_from_full = sep(region_tab$exact, delta + beta_inter),
  stringsAsFactors = FALSE
)

knitr::kable(c5, digits = 4,
             caption = "Shape coverage and separation of each target from the three degenerate values.")

int <- c5[c5$interior, ]
min_sepn <- min(c(int$d_from_0, int$d_from_delta, int$d_from_full))
cat(sprintf("\ninterior regions: %d of %d\n", nrow(int), nrow(c5)))
cat(sprintf("minimum separation over interior regions = %.4f\n", min_sepn))
cat(sprintf("shapes present: %s\n", paste(unique(c5$shape), collapse = "; ")))

results_log$c5 <- nrow(int) >= 3 && min_sepn > 1 &&
                  all(c("disjunction (GRF)", "negation", "disjoint from Q") %in% c5$shape)
```

# Criterion 6 — reachability

At least one assertion must run on a region produced by an **actual
`forestsearch()` call**, not a hand-constructed mask.

The continuous vignette warns that at modest $n$ a single replicate may fail to
retain the true subgroup-defining variables, because the analysis sees raw
covariates and LASSO/GRF must rediscover `age`/`preanti` each time. A larger
draw is used here for that reason. **The realized outcome is reported either
way — no seed is retried until one detects.** A control that passes because the
quantity is easy is not evidence the code path is reachable.

::: {.callout-important title="`adverse_outcome` is required here, and both settings are reported"}
`forestsearch()` takes `adverse_outcome`, defaulting to `NULL`, which resolves
to `outcome_type %in% c("binary", "count")` — hence **`FALSE` for continuous**
(`R/forestsearch_main.R:1220-1221`).

The continuous vignette states at lines 138-162 that `adverse_outcome = TRUE`
is **required** for this setup and that the default `FALSE` is "WRONG for
benefit searches via treatment switching": the search maximises
$\text{score} = -\text{CATE}$, so under switched roles the default looks in the
direction *opposite* to $Q$.

An earlier version of this document omitted the argument entirely, ran under
the resolved default, and recorded the resulting inversion as an observation
about the harm/benefit fork. That was a misconfigured call, not evidence.
Both settings are now run and reported, with `TRUE` as the configured one.
:::

```{r criterion-6}
t0 <- proc.time()[3]
fs_n <- 4000L
df_fs <- simulate_from_glm_dgm(dgm_calibrated, n = fs_n, seed = sim_seed + 7L)

run_fs_once <- function(adverse) {
  tryCatch(
    suppressWarnings(forestsearch(
      df.analysis            = df_fs,
      confounders.name       = confounders_analysis,
      outcome.name           = "y_sim",
      treat.name             = "treat_sim",
      id.name                = "id",
      outcome_type           = "continuous",
      effect_measure         = "MD",
      effect.threshold       = 30,
      consistency.threshold  = 10,
      pconsistency.threshold = 0.90,
      fs.splits              = 400L,
      n.min                  = 60L,
      d0.min                 = 12L,
      d1.min                 = 12L,
      maxk                   = 2L,
      vi.grf.min             = -0.2,
      use_lasso              = TRUE,
      use_grf                = TRUE,
      use_twostage           = TRUE,
      is.RCT                 = TRUE,
      adverse_outcome        = adverse,
      details                = FALSE,
      quiet                  = TRUE,
      seedit                 = sim_seed,
      parallel_args = list(plan = "sequential", workers = 1, show_message = FALSE)
    )),
    error = function(e) { message("forestsearch() failed: ", e$message); NULL }
  )
}

# Reported for contrast: the resolved default, which the vignette calls wrong.
fs_default <- run_fs_once(FALSE)
# The configured call.
fs_run     <- run_fs_once(TRUE)

timings$c6 <- proc.time()[3] - t0
cat(sprintf("two forestsearch() calls on n = %d completed in %.1f sec\n",
            fs_n, timings$c6))

describe_run <- function(fsobj, label) {
  sg <- if (!is.null(fsobj) && !is.null(fsobj$grp.consistency))
          fsobj$grp.consistency$sg.harm else NULL
  if (is.null(sg) || length(sg) == 0 || all(is.na(sg))) {
    cat(sprintf("\n[%s] NO subgroup identified.\n", label)); return(invisible(NULL))
  }
  sd_ <- paste(sg, collapse = " & ")
  v   <- betaHhat_one_md(sd_, df_super, focus = "harm")
  idx <- get_dfpred(df_super, .resolve_rule(sd_))$treat.recommend == 0L
  d   <- df_fs[get_dfpred(df_fs, .resolve_rule(sd_))$treat.recommend == 0L, ]
  mdt <- mean(d$y_sim[d$treat_sim == 1L]) - mean(d$y_sim[d$treat_sim == 0L])
  cat(sprintf("\n[%s]\n  rule            : %s\n", label, sd_))
  cat(sprintf("  P(Q | region)   : %.6f%s\n", p_Q_given(idx),
              if (p_Q_given(idx) %in% c(0, 1)) "   <- BOUNDARY region" else "   <- interior"))
  cat(sprintf("  beta(Hhat)      : %.4f    beta(Hhat^c): %.4f\n",
              v[["betaHhat_H"]], v[["betaHhat_Hc"]]))
  cat(sprintf("  nH_eval         : %d\n", v[["nH_eval"]]))
  cat(sprintf("  in-trial MD     : %.4f\n", mdt))
  cat(sprintf("  selection effect: %.4f\n", mdt - v[["betaHhat_H"]]))
  invisible(sd_)
}

cat(sprintf("reference: MD(Q) = %.4f, delta = %.4f, delta + beta_inter = %.4f\n",
            effect_Q, delta, delta + beta_inter))
describe_run(fs_default, "adverse_outcome = FALSE  (resolved default; vignette calls this wrong)")
describe_run(fs_run,     "adverse_outcome = TRUE   (configured)")

sg_found <- NULL
if (!is.null(fs_run) && !is.null(fs_run$grp.consistency)) {
  sg_found <- fs_run$grp.consistency$sg.harm
}

if (!is.null(sg_found) && length(sg_found) > 0 && !all(is.na(sg_found))) {
  sg_def <- paste(sg_found, collapse = " & ")
  cat(sprintf("\nRealized rule: %s\n", sg_def))

  v_harm    <- betaHhat_one_md(sg_def, df_super, focus = "harm")
  v_benefit <- betaHhat_one_md(sg_def, df_super, focus = "benefit")

  c6 <- data.frame(
    focus       = c("harm", "benefit"),
    betaHhat_H  = c(v_harm[["betaHhat_H"]],  v_benefit[["betaHhat_H"]]),
    betaHhat_Hc = c(v_harm[["betaHhat_Hc"]], v_benefit[["betaHhat_Hc"]]),
    nH_eval     = c(v_harm[["nH_eval"]],     v_benefit[["nH_eval"]]),
    nHc_eval    = c(v_harm[["nHc_eval"]],    v_benefit[["nHc_eval"]]),
    stringsAsFactors = FALSE
  )
  knitr::kable(c6, digits = 6,
               caption = "Module evaluated on a rule produced by forestsearch().")

  # focus exercised both ways: complementary regions, differing targets
  n_sum_ok  <- (v_harm[["nH_eval"]] + v_benefit[["nH_eval"]]) == nrow(df_super)
  targets_differ <- isTRUE(v_harm[["betaHhat_H"]] != v_benefit[["betaHhat_H"]])

  cat(sprintf("\nnH_eval(harm) + nH_eval(benefit) = %d + %d = %d   (nrow(df_super) = %d)\n",
              v_harm[["nH_eval"]], v_benefit[["nH_eval"]],
              v_harm[["nH_eval"]] + v_benefit[["nH_eval"]], nrow(df_super)))
  cat(sprintf("targets differ: %s   (harm %.6f vs benefit %.6f)\n",
              targets_differ, v_harm[["betaHhat_H"]], v_benefit[["betaHhat_H"]]))

  # How informative is the region the configured run landed on?
  idx_fs <- get_dfpred(df_super, .resolve_rule(sg_def))$treat.recommend == 0L
  pq_fs  <- p_Q_given(idx_fs)
  cat(sprintf("\nP(Q | realized region) = %.6f\n", pq_fs))
  if (pq_fs == 0 || pq_fs == 1) {
    cat("  This is a BOUNDARY region: its target is exactly ")
    cat(if (pq_fs == 0) "delta" else "delta + beta_inter", "\n")
    cat("  The reachability check therefore ran, but landed on the least\n")
    cat("  informative possible region. Reported as-is; no seed retried.\n")
  } else {
    cat("  Interior region: target is strictly between delta and delta+beta_inter,\n")
    cat("  so the check is not passing on a degenerate value.\n")
  }

  results_log$c6            <- all(is.finite(c(v_harm[["betaHhat_H"]],
                                               v_harm[["betaHhat_Hc"]])))
  results_log$focus_complement <- n_sum_ok
  results_log$focus_differ     <- targets_differ
} else {
  cat("\nforestsearch() identified NO subgroup on this draw.\n")
  cat("Reported as the measurement, not retried with another seed.\n")
  cat("Criterion 6 is UNMET for this run.\n")
  results_log$c6               <- FALSE
  results_log$focus_complement <- NA
  results_log$focus_differ     <- NA
}
```

## `focus` on a hand-constructed rule

The complementary-region and differing-target assertions are repeated on a
hand-constructed rule so they are available regardless of whether
`forestsearch()` detected anything above.

```{r focus-hand}
vh <- betaHhat_one_md(rules$one_cut, df_super, focus = "harm")
vb <- betaHhat_one_md(rules$one_cut, df_super, focus = "benefit")

cat(sprintf("rule 'one_cut':\n"))
cat(sprintf("  harm:    betaHhat_H = %.6f,  nH_eval = %d\n",
            vh[["betaHhat_H"]], vh[["nH_eval"]]))
cat(sprintf("  benefit: betaHhat_H = %.6f,  nH_eval = %d\n",
            vb[["betaHhat_H"]], vb[["nH_eval"]]))
cat(sprintf("  nH_eval sum = %d  (nrow(df_super) = %d)\n",
            vh[["nH_eval"]] + vb[["nH_eval"]], nrow(df_super)))
cat(sprintf("  targets differ: %s\n", vh[["betaHhat_H"]] != vb[["betaHhat_H"]]))

results_log$focus_complement_hand <-
  (vh[["nH_eval"]] + vb[["nH_eval"]]) == nrow(df_super)
results_log$focus_differ_hand <- vh[["betaHhat_H"]] != vb[["betaHhat_H"]]

# harm/benefit are mirror images: H under one is H^c under the other
cat(sprintf("  harm betaHhat_H == benefit betaHhat_Hc: %s\n",
            identical(vh[["betaHhat_H"]], vb[["betaHhat_Hc"]])))
```

# Reported, not asserted

Two things the task asks to be run and shown verbatim rather than folded into a
pass/fail gate.

## A missing `focus` — verbatim error

Confirming the message is the **module's**, not R's standard
missing-argument error.

```{r missing-focus}
err <- tryCatch(
  betaHhat_one_md(rules$exact_Q, df_super),
  error = function(e) conditionMessage(e)
)
cat(err, "\n")
```

## `.beta_region_md()` against the SPEC, and a `mu0`/`mu1` grep

First: is the module's `.beta_region_md()` byte-identical to the code block in
`SPEC_betaHhat_md.md`? Second: does `mu0` or `mu1` appear anywhere else in the
module in an arithmetic context — which would mean a third site for
`mean(mu1) - mean(mu0)`?

```{r spec-fidelity}
spec_lines   <- readLines("../SPEC_betaHhat_md.md")
module_lines <- readLines(module_path)

# Brace-counting extractor: the first "}" in this function is the inner one
# closing the guard, so a naive scan for "}" would truncate the block.
extract_fn <- function(lines) {
  s <- grep("^\\.beta_region_md <- function", lines)[1]
  depth <- 0L
  for (e in seq(s, length(lines))) {
    depth <- depth + lengths(regmatches(lines[e], gregexpr("\\{", lines[e]))) -
                     lengths(regmatches(lines[e], gregexpr("\\}", lines[e])))
    if (depth == 0L && e > s) return(lines[s:e])
    if (depth == 0L && grepl("\\}", lines[e])) return(lines[s:e])
  }
  stop("could not delimit .beta_region_md()")
}

spec_fn <- extract_fn(spec_lines)
mod_fn  <- extract_fn(module_lines)

cat("SPEC block:\n"); cat(spec_fn, sep = "\n")
cat("\n\nModule block:\n"); cat(mod_fn, sep = "\n")
cat(sprintf("\n\nbyte-identical: %s\n", identical(spec_fn, mod_fn)))

results_log$spec_identical <- identical(spec_fn, mod_fn)
```

```{r mu-grep}
hits <- grep("mu0|mu1", module_lines, value = TRUE)
cat(sprintf("grep 'mu0|mu1' over %s\n", basename(module_path)))
cat(sprintf("  %d matching line(s)\n\n", length(hits)))
if (length(hits)) cat(hits, sep = "\n") else cat("(empty)\n")

# arithmetic context = a match outside a comment line
code_hits <- hits[!grepl("^\\s*#", hits)]
cat(sprintf("\n  of which outside comments: %d\n", length(code_hits)))
if (length(code_hits)) cat(code_hits, sep = "\n") else cat("(empty)\n")

results_log$no_third_site <- length(code_hits) == 0L
```

# Discussion — open items

Recorded here, not acted on in this task.

## The string-parsing defect class, and the package's own answer

Criterion 5 was initially blocked. `SPEC_betaHhat_md.md` states under *Call
surface* that "GRF disjunctions live inside a single element and survive the
split". **They do not.** A GRF disjunction carries `" & "` inside each
conjunct, so a `" & "` split shreds it:

    input : "(age > 34 & preanti <= 744.5) | (wtkg <= 60)"
    split : [1] "(age > 34"   [2] "preanti <= 744.5) | (wtkg <= 60)"
    result: Column '(age' not found  ->  treat.recommend = NA  ->  NA target

The immediate cause is dispatch order: `get_dfpred()` tests for the disjunction
form **first**, on the unsplit string (`R/forestsearch_helpers.R:101`), while
the module split first. This module now mirrors `get_dfpred()`'s order, which
unblocks criterion 5.

**That is a patch, not the fix.** The package's own answer for GRF membership
is `.grf_evaluate_subgroup()` applied to the **structured** `sg_def`
(`R/forestsearch_helpers.R:1778, 1781, 1784`), chosen explicitly because
`get_dfpred()` cannot evaluate the disjunction string — see the comment at
`R/forestsearch_helpers.R:1772`. All three `betaHhat_truth_*.R` modules
re-parse the label string instead, which is why this defect exists at all.

**Open item.** If `sg_def` structure is reachable at the `attach_betaHhat_*()`
call site, the string handling should be replaced by the structured evaluator
in all three modules. The defect class then disappears rather than being
patched once per module. Not in scope here.

**Also note:** the same split-first logic is live in the binary sibling,
`quarto/simulations/actg175/binary/betaHhat_truth_glm.R:76-77`. Any GRF rule
scored through `betaHhat_one_or()` today yields `NA` targets silently rather
than erroring. Out of scope for this task; flagged for triage.

## The SPEC and NOTE carry fixture-specific numbers

Two documentation items, for the outstanding handoff-correction commit rather
than for this one:

1. The SPEC's "survive the split" sentence is wrong, as above.
2. `NOTE_target_is_collapsibility.md` section 3 reports
   $\delta = 6.5$, $\beta_{\mathrm{inter}} = 40$, $\text{effect}_Q = 46.5$.
   This calibrated fixture gives $\delta = -26.26$,
   $\beta_{\mathrm{inter}} = 66.26$, $\text{effect}_Q = 40$. The NOTE's
   **structural** claim is untouched and is confirmed below — the gap is
   exactly $\delta$ on every region — but $\delta$ is **negative** here, so
   the handoff's form runs *high* by 26.26 rather than low. The NOTE's prose
   ("low by $\delta$", "every handoff difference is $\delta = 6.5$") describes
   a different fixture.

## A withdrawn observation about search direction

An earlier version of this document reported that `forestsearch()`, run on the
vignette's own parameters and DGM, selected a region disjoint from $Q$ and in
the opposite effect direction — and offered that as an observation bearing on
the open harm/benefit fork.

**That observation is withdrawn.** The call omitted `adverse_outcome`, so it
ran under the resolved default `FALSE`
(`R/forestsearch_main.R:1220-1221`), which the continuous vignette documents at
lines 138-162 as *required to be* `TRUE` for this setup and explicitly "WRONG
for benefit searches via treatment switching". The vignette predicts that exact
inversion as the consequence of the wrong setting: the search maximises
$\text{score} = -\text{CATE}$, so under switched roles the default looks
opposite to $Q$.

The inversion was therefore a misconfigured call reproducing a documented
failure mode, not evidence about the fork. Criterion 6 now runs both settings;
with `adverse_outcome = TRUE` the selected region overlaps $Q$ substantially
and is **interior**, so the reachability check no longer passes on a degenerate
target. The harm/benefit fork remains open, and nothing in this document speaks
to it.

The general lesson is worth keeping: a misconfigured call recorded as evidence
is worse than no observation, because it looks like a finding.

## What is still unverified

Criterion 6 establishes that the module runs on a rule `forestsearch()`
actually produced. It does **not** establish that the module is wired into any
engine: no `attach_betaHhat_md()` call exists in a sweep yet, and the
harm/benefit decision that fixes `focus` at that call site is still open.

# Summary

```{r summary}
#| error: true
summary_tab <- data.frame(
  check  = names(results_log),
  result = vapply(results_log, function(x)
             if (is.na(x)) "NOT RUN" else if (isTRUE(x)) "pass" else "FAIL",
             character(1)),
  row.names = NULL
)
knitr::kable(summary_tab, caption = "Gate. Measurements above are the output; this is bookkeeping.")

cat("\nTimings (sec):\n")
print(unlist(timings))

failed <- summary_tab$check[summary_tab$result == "FAIL"]
if (length(failed)) {
  cat(sprintf("\nFAILED: %s\n", paste(failed, collapse = ", ")))
  stop("Acceptance gate not met. Do NOT widen a tolerance -- report the numbers.")
} else {
  cat("\nAll checks met.\n")
}
```

```{r session}
sessionInfo()
```
