---
title: "GRF Subgroup Identification: dmin.grf Calibration for Binary Endpoints"
subtitle: "Systematic Evaluation of Detection Power and FPR Control"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 4
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# -- Global parameters (all tunable from this single location) ----------------
N_syn         <- 1200L          # Synthetic data sample size
# 100 ~ 20m linux
nsims         <- 1500L          # Sims per dmin × scenario × effect size
sim_n         <- 800L           # Per-simulation sample size
dgm_n_super  <- 5000L          # Super-population size for DGMs
sim_seed      <- 8316951L       # Master seed
confounders   <- c("z1", "z2", "z3", "z4")
n_workers     <- max(1L, floor(0.90 * parallel::detectCores(logical = FALSE)))

# dmin.grf grid (rate difference scale for binary outcomes)
# Clinically: 0.05 = 5pp, 0.10 = 10pp, 0.20 = 20pp
dmin_grid     <- c(0, 0.02, 0.05, 0.08, 0.10, 0.12, 0.15, 0.20, 0.25, 0.30)

# Effect size multipliers (k_inter for generate_glm_dgm)
k_inter_levels <- c(weak = 0.8, moderate = 1.5, strong = 2.5)

# FPR / Detection targets
target_fpr    <- 15   # Maximum acceptable FPR (%)
target_det    <- 50   # Minimum acceptable detection (%)
```


# Purpose {#sec-purpose}

`grf.subg.harm.glm()` uses `causal_forest()` for candidate factor
screening and policy trees for subgroup identification.  The
`dmin.grf` parameter sets the minimum absolute CATE (conditional
average treatment effect) difference required for a policy tree
node to be considered valid.

For binary outcomes, the CATE is on the **risk difference** scale:

$$\text{CATE}(x) = P(Y=1 \mid T=1, X=x) - P(Y=1 \mid T=0, X=x)$$

Setting `dmin.grf = 0` means the policy tree **always** splits,
producing near-100% FPR.  This document systematically evaluates
a grid of `dmin.grf` values to find the range that balances
detection power ($\geq$ `r target_det`%) with FPR control
($\leq$ `r target_fpr`%).

## Design

The evaluation proceeds in three phases:

1. **Configuration discovery**: Test all combinations of treatment
   switching and `adverse_outcome` to identify which GRF
   configuration correctly detects the planted subgroup Q for
   each of four binary scenario types.

2. **dmin.grf sweep**: For each working configuration, evaluate
   detection and FPR across `r length(dmin_grid)` values of
   `dmin.grf` at three effect sizes (weak, moderate, strong).

3. **Summary**: Identify optimal `dmin.grf` per scenario and
   produce detection--FPR tradeoff curves.


# Setup {#sec-setup}

```{r packages}
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(doFuture)
library(foreach)

plan(multisession, workers = n_workers)
cat(sprintf("Workers: %d\n", n_workers))

reset_workers <- function(msg = "") {
  if (nzchar(msg)) cat(sprintf("[Resetting workers: %s]\n", msg))
  plan(sequential); gc(verbose = FALSE); Sys.sleep(1)
  plan(multisession, workers = n_workers)
}

t_start <- proc.time()
```


# Synthetic Data {#sec-data}

```{r syn-data}
set.seed(42)
syn <- data.frame(
  id = seq_len(N_syn), treat = rep(0:1, each = N_syn / 2),
  z1 = as.factor(sample(0:1, N_syn, TRUE)),
  z2 = as.factor(sample(0:1, N_syn, TRUE)),
  z3 = as.factor(sample(0:1, N_syn, TRUE)),
  z4 = as.factor(sample(0:1, N_syn, TRUE))
)
in_Q <- syn$z1 == 1 & syn$z2 == 1
syn$treat_sw <- 1L - syn$treat
cat(sprintf("Synthetic data: N=%d, Q prevalence=%.1f%%\n",
    N_syn, 100 * mean(in_Q)))
```

Four binary outcome scenarios, each with a planted subgroup
Q = {z1 == 1 & z2 == 1}:

```{r outcomes}
# (A) Harm + Adverse: treatment INCREASES bad outcome for Q
syn$y_harm_adv <- rbinom(N_syn, 1,
  ifelse(syn$treat == 1 & in_Q, 0.65,
    ifelse(syn$treat == 1, 0.22, 0.35)))

# (B) Harm + Positive: treatment REDUCES good outcome for Q
syn$y_harm_pos <- rbinom(N_syn, 1,
  ifelse(syn$treat == 1 & in_Q, 0.15,
    ifelse(syn$treat == 1, 0.55, 0.40)))

# (C) Benefit + Positive: treatment INCREASES good outcome MORE for Q
syn$y_ben_pos <- rbinom(N_syn, 1,
  ifelse(syn$treat == 1 & in_Q, 0.80,
    ifelse(syn$treat == 1, 0.50, 0.40)))

# (D) Benefit + Adverse: treatment REDUCES bad outcome MORE for Q
syn$y_ben_adv <- rbinom(N_syn, 1,
  ifelse(syn$treat == 1 & in_Q, 0.10,
    ifelse(syn$treat == 1, 0.33, 0.38)))

# Summary
cat("Raw CATE (risk difference) for Q:\n")
for (nm in c("y_harm_adv", "y_harm_pos", "y_ben_pos", "y_ben_adv")) {
  rd_Q  <- mean(syn[[nm]][in_Q & syn$treat == 1]) -
           mean(syn[[nm]][in_Q & syn$treat == 0])
  rd_Qc <- mean(syn[[nm]][!in_Q & syn$treat == 1]) -
           mean(syn[[nm]][!in_Q & syn$treat == 0])
  cat(sprintf("  %s: RD(Q)=%+.3f, RD(Qc)=%+.3f, diff=%.3f\n",
      nm, rd_Q, rd_Qc, abs(rd_Q - rd_Qc)))
}
```


# Phase 1: Configuration Discovery {#sec-phase1}

GRF identifies the subgroup with the **most negative CATE** as
"harm."  For adverse outcomes (higher Y = worse), this
convention is inverted — negative CATE means treatment
*reduces* the bad outcome, which is beneficial.

Two mechanisms can fix this:

- `adverse_outcome = TRUE`: flips Y to 1−Y before GRF
- **Treatment switching**: negates the CATE directly

This phase tests all 16 combinations (4 scenarios × 2 treatment
directions × 2 adverse_outcome settings) to empirically determine
which produces correct detection.

```{r phase1}
test_config <- function(scenario, y_col, treat_col, adverse_outcome) {
  res <- tryCatch(
    grf.subg.harm.glm(
      data = syn, confounders.name = confounders,
      outcome.name = y_col, treat.name = treat_col, id.name = "id",
      outcome_type = "binary", adverse_outcome = adverse_outcome,
      n.min = 40L, dmin.grf = 0, maxdepth = 2L, RCT = TRUE,
      sg.criterion = "mDiff", seedit = 42L, verbose = FALSE
    ),
    error = function(e) NULL
  )

  if (is.null(res) || is.null(res$sg.harm.id) ||
      length(res$sg.harm.id) == 0 || all(is.na(res$sg.harm.id))) {
    return(data.frame(Scenario = scenario, Treatment = treat_col,
      Adverse_Outcome = adverse_outcome, Detected = FALSE,
      Sensitivity = NA_real_, PPV = NA_real_,
      Subgroup = "", stringsAsFactors = FALSE))
  }

  in_H <- res$data$treat.recommend == 0
  tp <- sum(in_H & in_Q); fp <- sum(in_H & !in_Q)
  fn <- sum(!in_H & in_Q)
  sens <- tp / max(tp + fn, 1)
  ppv  <- tp / max(tp + fp, 1)

  data.frame(Scenario = scenario, Treatment = treat_col,
    Adverse_Outcome = adverse_outcome, Detected = TRUE,
    Sensitivity = round(sens, 2), PPV = round(ppv, 2),
    Subgroup = paste(res$sg.harm.id, collapse = " & "),
    stringsAsFactors = FALSE)
}

# Test all 16 combinations
configs <- list()
for (scen in list(
  list(name = "Harm+Adverse", y = "y_harm_adv"),
  list(name = "Harm+Positive", y = "y_harm_pos"),
  list(name = "Benefit+Positive", y = "y_ben_pos"),
  list(name = "Benefit+Adverse", y = "y_ben_adv")
)) {
  for (tr in c("treat", "treat_sw")) {
    for (ao in c(FALSE, TRUE)) {
      configs[[length(configs) + 1]] <- test_config(
        scen$name, scen$y, tr, ao)
    }
  }
}

config_df <- do.call(rbind, configs)
```

## Full Configuration Results

```{r phase1-table}
config_df |>
  gt::gt() |>
  gt::tab_header(
    title = "GRF Configuration Discovery",
    subtitle = sprintf("N=%d, Q={z1==1 & z2==1}, dmin.grf=0", N_syn)
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#c8e6c9"),
    locations = gt::cells_body(
      rows = Detected & !is.na(Sensitivity) & Sensitivity >= 0.5)
  ) |>
  gt::tab_style(
    style = gt::cell_fill(color = "#ffcdd2"),
    locations = gt::cells_body(
      rows = !Detected | is.na(Sensitivity) | Sensitivity < 0.5)
  ) |>
  gt::tab_footnote(
    footnote = "Green = correctly identifies Q (Sens >= 0.50). Red = wrong subgroup or no detection."
  )
```

## Best Configuration per Scenario

```{r best-configs}
best_configs <- list()
best_df <- data.frame()

for (scen in unique(config_df$Scenario)) {
  sub <- config_df[config_df$Scenario == scen & config_df$Detected &
                     !is.na(config_df$Sensitivity), ]
  if (nrow(sub) == 0) next
  best <- sub[which.max(sub$Sensitivity), ]
  best_configs[[scen]] <- best
  best_df <- rbind(best_df, best)
}

best_df |>
  gt::gt() |>
  gt::tab_header(
    title = "Recommended GRF Configuration per Scenario",
    subtitle = "Highest sensitivity configuration"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(columns = Sensitivity)
  )
```


# Phase 2: dmin.grf Sweep {#sec-phase2}

For each working configuration from Phase 1, we evaluate
`r length(dmin_grid)` values of `dmin.grf` at
`r length(k_inter_levels)` effect sizes, using
`r nsims` alt + `r nsims` null simulations (N=`r sim_n`)
per grid point.

```{r phase2-helpers}
run_dmin_eval <- function(dgm, nsims, dmin_val, adverse_outcome) {
  results_raw <- foreach(
    i = seq_len(nsims),
    .errorhandling = "pass",
    .options.future = list(
      packages = c("forestsearch", "survival", "data.table"),
      seed = TRUE
    )
  ) %dofuture% {
    sim_df <- simulate_from_glm_dgm(dgm, n = sim_n,
                                     seed = sim_seed + i * 1000L)
    res <- tryCatch(
      grf.subg.harm.glm(
        data = sim_df, confounders.name = confounders,
        outcome.name = "y_sim", treat.name = "treat_sim",
        id.name = "id", outcome_type = "binary",
        adverse_outcome = adverse_outcome,
        n.min = 40L, dmin.grf = dmin_val, maxdepth = 2L,
        RCT = TRUE, sg.criterion = "mDiff", seedit = 42L,
        verbose = FALSE
      ),
      error = function(e) NULL
    )
    detected <- !is.null(res) && !is.null(res$sg.harm.id) &&
      length(res$sg.harm.id) > 0 && !all(is.na(res$sg.harm.id))
    detected
  }

  # Handle any errors from foreach
  is_err <- vapply(results_raw, inherits, logical(1), "error")
  good <- unlist(results_raw[!is_err])
  100 * sum(good) / nsims
}

# Scenario DGM specs
scenario_specs <- list(
  list(name = "Harm+Adverse", y_alt = "y_harm_adv",
       null_rates = c(treat = 0.22, ctrl = 0.35)),
  list(name = "Harm+Positive", y_alt = "y_harm_pos",
       null_rates = c(treat = 0.55, ctrl = 0.40)),
  list(name = "Benefit+Positive", y_alt = "y_ben_pos",
       null_rates = c(treat = 0.55, ctrl = 0.40)),
  list(name = "Benefit+Adverse", y_alt = "y_ben_adv",
       null_rates = c(treat = 0.22, ctrl = 0.35))
)
```

```{r phase2-sweep}
sweep_results <- data.frame(
  Scenario = character(0), Effect_Size = character(0),
  k_inter = numeric(0), OR_Q = numeric(0),
  dmin = numeric(0), Det_H1 = numeric(0), FPR_H0 = numeric(0),
  stringsAsFactors = FALSE
)

cat(sprintf("dmin.grf grid: %s\n", paste(dmin_grid, collapse = ", ")))
cat(sprintf("Effect sizes: %s\n",
    paste(sprintf("%s (k=%.1f)", names(k_inter_levels), k_inter_levels),
          collapse = ", ")))
cat(sprintf("Sims per point: %d alt + %d null, N=%d\n\n", nsims, nsims, sim_n))

for (spec in scenario_specs) {
  scen_name <- spec$name
  if (!scen_name %in% names(best_configs)) {
    cat(sprintf("--- Skipping %s (no working config) ---\n\n", scen_name))
    next
  }

  bc <- best_configs[[scen_name]]
  treat_col <- bc$Treatment
  adverse_out <- bc$Adverse_Outcome

  cat(sprintf("=== %s (treat=%s, adverse_outcome=%s) ===\n",
      scen_name, treat_col, adverse_out))

  # Reset workers at start of each scenario to prevent memory buildup
  reset_workers(scen_name)

  # Create null outcome
  y_null_col <- paste0("y_null_", gsub("\\+", "", scen_name))
  syn[[y_null_col]] <- rbinom(N_syn, 1,
    ifelse(syn[[treat_col]] == 1,
           spec$null_rates["treat"], spec$null_rates["ctrl"]))

  for (es_name in names(k_inter_levels)) {
    k_val <- k_inter_levels[es_name]

    dgm_alt <- tryCatch(
      generate_glm_dgm(
        syn, confounders, spec$y_alt, treat_col, "binary", "OR",
        subgroup_vars = c("z1", "z2"),
        subgroup_cuts = list(z1 = 1L, z2 = 1L),
        model = "alt", k_inter = k_val,
        n_super = dgm_n_super, seed = sim_seed),
      error = function(e) NULL)

    dgm_null <- tryCatch(
      generate_glm_dgm(
        syn, confounders, y_null_col, treat_col, "binary", "OR",
        subgroup_vars = c("z1", "z2"),
        subgroup_cuts = list(z1 = 1L, z2 = 1L),
        model = "null", k_treat = 1,
        n_super = dgm_n_super, seed = sim_seed),
      error = function(e) NULL)

    if (is.null(dgm_alt) || is.null(dgm_null)) next

    or_Q <- dgm_alt$hazard_ratios$harm_subgroup
    cat(sprintf("  %s (k=%.1f): OR(Q)=%.2f\n", es_name, k_val, or_Q))

    for (dmin_val in dmin_grid) {
      det <- run_dmin_eval(dgm_alt, nsims, dmin_val, adverse_out)
      fpr <- run_dmin_eval(dgm_null, nsims, dmin_val, adverse_out)

      sweep_results[nrow(sweep_results) + 1, ] <- data.frame(
        Scenario = scen_name, Effect_Size = es_name,
        k_inter = k_val, OR_Q = round(or_Q, 2),
        dmin = dmin_val, Det_H1 = det, FPR_H0 = fpr,
        stringsAsFactors = FALSE)

      flag <- if (det >= target_det && fpr <= target_fpr) " ***" else ""
      cat(sprintf("    dmin=%.2f: Det=%3.0f%%, FPR=%3.0f%%%s\n",
          dmin_val, det, fpr, flag))
    }
    cat("\n")
  }
  cat("\n")
}
```


```{r cleanup}
plan(sequential)
```


# Results {#sec-results}

## Detection--FPR Tradeoff Curves {#sec-curves}

```{r tradeoff-curves, fig.height=8}
#| fig-cap: "Detection rate vs FPR across dmin.grf values. Dashed lines mark the target zone (Det >= 50%, FPR <= 15%). Points in the upper-left quadrant achieve the target."

if (nrow(sweep_results) > 0) {
  sweep_results$label <- sprintf("%.2f", sweep_results$dmin)

  ggplot(sweep_results,
         aes(x = FPR_H0, y = Det_H1, colour = Effect_Size)) +
    geom_line(aes(group = Effect_Size), linewidth = 0.8) +
    geom_point(size = 2.5) +
    geom_text(aes(label = label), vjust = -0.8, size = 2.3,
              show.legend = FALSE) +
    geom_hline(yintercept = target_det, linetype = "dashed",
               colour = "grey50") +
    geom_vline(xintercept = target_fpr, linetype = "dashed",
               colour = "grey50") +
    annotate("rect", xmin = 0, xmax = target_fpr,
             ymin = target_det, ymax = 100,
             alpha = 0.08, fill = "green") +
    facet_wrap(~ Scenario, ncol = 2, scales = "free_x") +
    scale_colour_manual(values = c(
      weak = "#e74c3c", moderate = "#f39c12", strong = "#27ae60")) +
    labs(
      x = "False Positive Rate (%)",
      y = "Detection Rate (%)",
      colour = "Effect Size",
      title = "GRF Detection--FPR Tradeoff by dmin.grf",
      subtitle = sprintf(
        "%d sims per point, N=%d | Labels = dmin.grf values",
        nsims, sim_n)
    ) +
    coord_cartesian(xlim = c(0, 105), ylim = c(0, 105)) +
    theme_minimal(base_size = 11) +
    theme(
      strip.text = element_text(face = "bold"),
      legend.position = "bottom"
    )
}
```


## dmin.grf vs Detection / FPR {#sec-dmin-curves}

```{r dmin-vs-rates, fig.height=8}
#| fig-cap: "Detection (solid) and FPR (dashed) as a function of dmin.grf. The shaded band marks the target FPR zone."

if (nrow(sweep_results) > 0) {
  plot_df <- rbind(
    transform(sweep_results, Metric = "Detection", Rate = Det_H1),
    transform(sweep_results, Metric = "FPR", Rate = FPR_H0)
  )

  ggplot(plot_df, aes(x = dmin, y = Rate, colour = Effect_Size,
                      linetype = Metric)) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 1.5) +
    geom_hline(yintercept = target_fpr, linetype = "dotted",
               colour = "red", linewidth = 0.5) +
    geom_hline(yintercept = target_det, linetype = "dotted",
               colour = "blue", linewidth = 0.5) +
    facet_wrap(~ Scenario, ncol = 2) +
    scale_linetype_manual(values = c(Detection = "solid", FPR = "dashed")) +
    scale_colour_manual(values = c(
      weak = "#e74c3c", moderate = "#f39c12", strong = "#27ae60")) +
    labs(
      x = "dmin.grf",
      y = "Rate (%)",
      colour = "Effect Size", linetype = "Metric",
      title = "Detection and FPR vs dmin.grf",
      subtitle = "Solid = Detection, Dashed = FPR"
    ) +
    coord_cartesian(ylim = c(0, 105)) +
    theme_minimal(base_size = 11) +
    theme(
      strip.text = element_text(face = "bold"),
      legend.position = "bottom"
    )
}
```


## Full Results Table {#sec-table}

```{r results-table}
if (nrow(sweep_results) > 0) {
  sweep_results$Pass <- sweep_results$Det_H1 >= target_det &
                        sweep_results$FPR_H0 <= target_fpr

  sweep_results |>
    gt::gt() |>
    gt::tab_header(
      title = "GRF dmin.grf Calibration Results",
      subtitle = sprintf(
        "%d alt + %d null sims, N=%d | Target: Det >= %d%%, FPR <= %d%%",
        nsims, nsims, sim_n, target_det, target_fpr)
    ) |>
    gt::cols_label(
      Effect_Size = "Effect", OR_Q = "OR(Q)",
      dmin = "dmin.grf", Det_H1 = "Det %", FPR_H0 = "FPR %"
    ) |>
    gt::tab_style(
      style = gt::cell_fill(color = "#c8e6c9"),
      locations = gt::cells_body(rows = Pass == TRUE)
    ) |>
    gt::cols_hide(columns = c("Pass", "k_inter"))
}
```


## Optimal dmin.grf per Scenario {#sec-optimal}

```{r optimal}
#| results: asis

if (nrow(sweep_results) > 0) {
  optimal_df <- data.frame()

  for (scen in unique(sweep_results$Scenario)) {
    for (es in unique(sweep_results$Effect_Size)) {
      sub <- sweep_results[sweep_results$Scenario == scen &
                            sweep_results$Effect_Size == es, ]
      good <- sub[sub$Det_H1 >= target_det & sub$FPR_H0 <= target_fpr, ]

      if (nrow(good) > 0) {
        best <- good[which.min(good$dmin), ]
        optimal_df <- rbind(optimal_df, data.frame(
          Scenario = scen, Effect_Size = es,
          OR_Q = best$OR_Q,
          Optimal_dmin = best$dmin,
          Detection = best$Det_H1,
          FPR = best$FPR_H0,
          Status = "Achieved",
          stringsAsFactors = FALSE))
      } else {
        sub$score <- sub$Det_H1 - 2 * sub$FPR_H0
        best_trade <- sub[which.max(sub$score), ]
        optimal_df <- rbind(optimal_df, data.frame(
          Scenario = scen, Effect_Size = es,
          OR_Q = best_trade$OR_Q,
          Optimal_dmin = best_trade$dmin,
          Detection = best_trade$Det_H1,
          FPR = best_trade$FPR_H0,
          Status = "Best tradeoff",
          stringsAsFactors = FALSE))
      }
    }
  }

  optimal_df |>
    gt::gt() |>
    gt::tab_header(
      title = "Optimal dmin.grf by Scenario and Effect Size",
      subtitle = sprintf("Target: Det >= %d%% AND FPR <= %d%%",
        target_det, target_fpr)
    ) |>
    gt::tab_style(
      style = gt::cell_fill(color = "#c8e6c9"),
      locations = gt::cells_body(rows = Status == "Achieved")
    ) |>
    gt::tab_style(
      style = gt::cell_fill(color = "#fff9c4"),
      locations = gt::cells_body(rows = Status == "Best tradeoff")
    )
}
```


## Mean FPR by dmin.grf {#sec-mean-fpr}

```{r mean-fpr}
if (nrow(sweep_results) > 0) {
  mean_by_dmin <- do.call(rbind, lapply(dmin_grid, function(d) {
    sub <- sweep_results[sweep_results$dmin == d, ]
    data.frame(
      dmin = d,
      Mean_Det = round(mean(sub$Det_H1, na.rm = TRUE), 1),
      Mean_FPR = round(mean(sub$FPR_H0, na.rm = TRUE), 1),
      stringsAsFactors = FALSE
    )
  }))

  mean_by_dmin |>
    gt::gt() |>
    gt::tab_header(
      title = "Mean Detection and FPR by dmin.grf",
      subtitle = "Averaged across all scenarios and effect sizes"
    ) |>
    gt::tab_style(
      style = gt::cell_fill(color = "#e8f5e9"),
      locations = gt::cells_body(
        rows = Mean_FPR <= target_fpr)
    )
}
```


# Phase 1 Reference: Best Configurations {#sec-configs-ref}

```{r configs-ref}
if (nrow(best_df) > 0) {
  best_df |>
    gt::gt() |>
    gt::tab_header(
      title = "Phase 1: Recommended GRF Configuration per Scenario",
      subtitle = "From single-dataset configuration discovery"
    ) |>
    gt::tab_footnote(
      footnote = paste(
        "Treatment = 'treat' (direct) or 'treat_sw' (switched).",
        "Adverse_Outcome controls the 1-Y flip before causal_forest().")
    )
}
```


# Timing {#sec-timing}

```{r timing}
elapsed <- (proc.time() - t_start)["elapsed"]
cat(sprintf("Total document render time: %.1f minutes (%.0f seconds)\n",
    elapsed / 60, elapsed))
cat(sprintf("Configuration: %d sims × %d dmin values × %d effect sizes × %d scenarios\n",
    nsims, length(dmin_grid), length(k_inter_levels),
    length(unique(sweep_results$Scenario))))
cat(sprintf("Total GRF calls: ~%d\n",
    2 * nsims * length(dmin_grid) * length(k_inter_levels) *
    length(unique(sweep_results$Scenario))))
```


# Session Info {#sec-session}

```{r session-info}
sessionInfo()
```
