<!--
  β(Ĥ) coverage table (manuscript fragment, payload-driven).

  Reads ONLY the sidecar written by betaHhat_sweep.qmd
  (betaHhat_<run_tag>.rds: a list with `summary`, `per_replicate`,
  `distinct_rules`, `meta`) — no forestsearch, no DGM, no Cox fitting at render
  time — so it renders in milliseconds. Defines @tbl-betaHhat-coverage.

  This reports coverage and relative bias against β(Ĥ), the conditional estimand
  Theorem 2 certifies (the censored observed-data least-false Cox coefficient in
  the *realized* region), as a complement to the θ†/θ‡/oracle columns of the main
  estimation table. Keep the sidecar beside this fragment or set `payload_dir`.
-->

```{r betaHhat-setup}
#| echo: false
#| message: false
#| warning: false
library(kableExtra)

# Where the sidecar lives (written by betaHhat_sweep.qmd, next to the bundles)
# and which run to show. Both must match the scorer's `rds_dir` and `run_tag`.
payload_dir <- "../analysis/simulations/gbsg/resampling_comparisons/m1"
run_tag     <- "fs_m1_h10_knoise0_n500"
rds_file    <- sprintf("betaHhat_%s.rds", run_tag)

.pdir <- if (is.null(payload_dir))
  tryCatch(dirname(knitr::current_input(dir = TRUE)), error = function(e) getwd()) else payload_dir
.rds_path <- file.path(.pdir, rds_file)
have_sc   <- file.exists(.rds_path)

fmt2 <- function(x) ifelse(is.na(x), "\u2014", sprintf("%.2f", x))
fmtb <- function(x) ifelse(is.na(x), "\u2014", sprintf("%+.1f", x))

if (have_sc) {
  SC <- readRDS(.rds_path); S <- SC$summary; M <- SC$meta
  # Wide layout: one row per estimator, columns = {cover, relbias} × {H, Hc}.
  est_lv <- levels(S$estimator); if (is.null(est_lv)) est_lv <- unique(S$estimator)
  pick <- function(rg, col) sapply(est_lv, function(e) {
    v <- S[S$region == rg & S$estimator == e, col]; if (length(v)) v[1] else NA_real_ })
  tab <- data.frame(
    Estimator = est_lv,
    cH  = fmt2(pick("H",  "cover_betaHhat")),
    bH  = fmtb(pick("H",  "relbias_betaHhat")),
    cHc = fmt2(pick("Hc", "cover_betaHhat")),
    bHc = fmtb(pick("Hc", "relbias_betaHhat")),
    check.names = FALSE)
  n_rec <- M$n_rule_recovered; n_det <- M$n_detected; n_rules <- M$n_distinct_rules
  cap_bh <- sprintf(paste0(
    "Coverage and relative bias against the conditional estimand $\\beta(\\widehat H)$ ",
    "— the least-false within-subgroup Cox coefficient on the observed, censored data, ",
    "evaluated at the realized region — for forest search, re-scoring the stored ",
    "multiplier-resampling and full-bootstrap intervals (no bootstrap re-run). ",
    "$\\beta(\\widehat H)$ was computed once per distinct realized rule (%d rules over ",
    "%d detections) on a censored evaluation population of n = %s (event rate %.2f). ",
    "Contrast with the $\\theta^{\\dagger}$/$\\theta^{\\ddagger}$/oracle columns of the ",
    "main estimation table, which target the true subgroup $H$."),
    n_rules, n_rec, format(M$n_eval, big.mark = ","), M$eval_event_rate)
} else {
  cap_bh <- sprintf(paste0("Coverage and relative bias against $\\beta(\\widehat H)$. ",
    "**Placeholder \u2014 run `betaHhat_sweep.qmd` to write `%s`.**"), rds_file)
}
```

## Coverage against the conditional estimand β(Ĥ) {#sec-betaHhat}

On the fixed forest-search family the multiplier-resampling and full-bootstrap intervals target $\beta(\widehat H)$, the population value of the standard within-subgroup analysis at the *realized* region. @tbl-betaHhat-coverage re-scores the stored intervals against $\beta(\widehat H)$ directly, the estimand Theorem 2 certifies — distinct from the $\theta^{\dagger}(H)$/$\theta^{\ddagger}(H)$ targets of the main table, which are causal effects in the *true* subgroup and coincide with $\beta(\widehat H)$ only when $\widehat H = H$ and the working model is proportional under the operative censoring.

```{r}
#| label: tbl-betaHhat-coverage
#| tbl-cap: !expr cap_bh
#| echo: false
#| message: false
#| warning: false
if (!have_sc) {
  data.frame(Estimator = c("oracle","naive","FB","MR"),
             `Cov` = "\u2014", `Bias` = "\u2014", check.names = FALSE) |>
    kbl(booktabs = TRUE, escape = FALSE) |> kable_styling()
} else {
  colnames(tab) <- c("Estimator",
                     "Coverage", "Rel. bias (%)", "Coverage", "Rel. bias (%)")
  kbl(tab, booktabs = TRUE, escape = FALSE, align = "lcccc", linesep = "") |>
    kable_styling(font_size = if (knitr::is_latex_output()) 10 else NULL) |>
    add_header_above(c(" " = 1,
                       "Harm subgroup" = 2,
                       "Complement" = 2),
                     escape = FALSE)
}
```
