---
title: "Post-selection coverage against the conditional estimand β(Ĥ)"
subtitle: "Refit-free re-scoring of stored MR / FB intervals against the realized-subgroup, censored observed-data Cox coefficient"
author: "León & Anderson"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    code-tools: true
    df-print: paged
    embed-resources: true
execute:
  warning: false
  message: false
---

## What this template does

Theorem 2 certifies coverage of **β(Ĥ)** — the population value of the *standard
within-subgroup analysis* (`coxph(Surv(time, event) ~ treat)` on the **observed,
randomized, censored** data) evaluated at whatever region the search realized.
The simulation tables instead score against the causal targets θ†(H) / θ‡(H)
(uncensored potential-outcome quantities at the *true* H). This template computes
β(Ĥ) and re-scores the **already-stored** MR and FB interval endpoints against it,
so coverage can be reported against the estimand the theorem actually certifies.

It does **not** re-run the bootstrap. For each detected replicate it (1) recovers
the realized rule (from the bundle's `sg.def` if present, else by re-identifying
with the same `forestsearch()` call your id-sweep uses, seed-linked via
`seed_base + sim_id`), (2) computes β(Ĥ) and β(Ĥᶜ) **once per distinct rule** on a
cached large censored super-population — membership via `get_dfpred()`, the
within-region log-HR via `get_Cox_sg()`, exactly the fitting path the oracle/naive
use — and (3) scores the stored `t1_*` (FB), `t2_*` (MR), `or_*` (oracle), and
`nv_*` (naive) endpoints for containment of β(Ĥ), with relative bias of each
point estimate against β(Ĥ).

The output is a sidecar `betaHhat_<run_tag>.rds` (per-replicate β plus
coverage/bias summaries, keyed by `sim_id`) that the companion fragment
`_betaHhat_coverage_table.qmd` renders alongside the existing estimation table.

> **β(Ĥ) is the *censored observed-data* Cox coefficient** — it is computed on the
> same generative pipeline as the simulation's per-replicate data
> (`simulate_from_dgm()` with `analysis_time` + `cens_adjust`), **not** on the
> uncensored potential-outcome stack (which would reproduce `hr_*_true`) and not
> the AHR average. The two differ under the AFT (non-PH) DGM; β(Ĥ) is the one
> Theorem 2 targets.

------------------------------------------------------------------------

## Configuration (knobs) {#sec-config}

```{r config}
# ── The stored t1_t2 results bundle to re-score ─────────────────────────────
# This is the SAME bundle the estimation/synthesis fragments read.
rds_dir   <- "../analysis/simulations/gbsg/resampling_comparisons/m1"
run_tag   <- "fs_m1_h10_knoise0_n500"                         # names the sidecar
bundle    <- file.path(rds_dir, "fs_t1_t2_m1_h10_knoise0_n500_combined_1_500.rds")

# ── DGM knobs — MUST match the generation of `bundle` (copied from the
#    id-sweep config so the recovered rules and β(Ĥ) align with that run). ────
dgm_model      <- "alt"
target_hr_harm <- 1.0
n_super        <- 25000L
seed_base      <- 8316951L        # replicate i used seed_base + i
analysis_time  <- 84              # administrative censoring (months) — CENSORED data
cens_adjust    <- log(1.5)        # censoring-rate adjustment
n_sample       <- 500L            # per-replicate trial size of THIS bundle

# ── Evaluation population for β(g): one large CENSORED draw, fixed across reps ─
n_eval    <- 100000L              # larger => more stable β(g); 1e5 is ample
eval_seed <- 20260628L            # fixed, distinct from any replicate seed

# ── Column map for the stored bundle (override if your bundle differs) ───────
# Estimator key -> column stem; endpoints are <stem>_<region>_{est,lo,hi}.
est_keys   <- c(oracle = "or", naive = "nv", FB = "t1", MR = "t2")
regions    <- c(H = "H", Hc = "Hc")
detect_col <- "detected"          # 0/1 detection flag in results
# Candidate columns that may carry the realized rule (first match wins).
rule_cols  <- c("sg.def", "sg", "subgroup", "sg.harm")
# Candidate columns for the per-replicate id (first match wins; else row order).
simid_cols <- c("sim_id", "sim", "rep", "replicate", "id")

# ── Rule recovery source ────────────────────────────────────────────────────
#   "idsweep"    : read the rules the id-sweep already recovered (its `sg`
#                  column -- the same " & "-joined named rule -- linked by
#                  sim_id). No forestsearch re-run; reuses the sweep you ran.
#   "auto"       : read a rule column from the t1_t2 bundle if present, else
#                  re-identify seed-linked.
#   "reidentify" : always re-identify seed-linked. This IS the id-sweep's
#                  record_replicate (re-simulate -> forestsearch -> fs$sg.harm);
#                  deterministic in seed_base + sim_id, so it reproduces the
#                  bundle's region exactly when fs_params match the run.
rule_recovery <- "idsweep"        # "idsweep" | "auto" | "reidentify"
# The id-sweep output to read rules from. fs_subgroup_id_sweep_*.qmd now names it
#   sprintf("%s_n%d_%.1f.rds", identifier, n_sample, target_hr_harm)
# and writes it to its cache_dir (default "."). Set `identifier` to match the run
# (fs / dina / grf) and `idsweep_dir` to where the sweep wrote that file.
identifier  <- "fs"               # "fs" | "dina" | "grf" -- matches the sweep run
idsweep_dir <- "."                # the sweep's cache_dir (where results_rds landed)
idsweep_rds <- file.path(idsweep_dir,
                         sprintf("%s_n%d_%.1f.rds", identifier, n_sample, target_hr_harm))

# ── forestsearch identification spec — used ONLY by the re-identify fallback.
#    Must match the config that produced `bundle`. (Copied from the id-sweep.) ─
confounders_base <- c("er", "age", "meno", "pgr", "nodes", "size", "grade")
fs_params <- list(
  subgroup_method        = "consistency",
  consistency_method     = "resample",
  sg_focus               = "eff",
  selection_rule         = "neighborhood",
  hr.threshold           = 0.90,
  hr.consistency         = 0.80,
  pconsistency.threshold = 0.90,
  n.min                  = NULL,
  maxk                   = 2L,
  d0.min                 = 10L,
  d1.min                 = 0L,
  use_lasso              = FALSE,
  use_grf                = FALSE,
  use_twostage           = TRUE,
  use_dina               = FALSE,
  conf_force             = c("meno == 0", "er <= 0", "pgr <= 0"),
  conf.cont_jcuts        = list(er = 10),
  fs.splits              = 400L
)

# Column names emitted by simulate_from_dgm().
outcome_name <- "y_sim"; event_name <- "event_sim"; treat_name <- "treat_sim"
id_name <- "id"; harm_col <- "flag_harm"

# ── Output / caching ────────────────────────────────────────────────────────
# The sidecar is written NEXT TO THE BUNDLE (rds_dir) so the synthesis figure
# and the coverage-table fragment, which look in the bundle directory, find it.
# The DGM and eval-population caches are scratch and stay in cache_dir.
cache_dir   <- "."
sidecar_rds <- file.path(rds_dir, sprintf("betaHhat_%s.rds", run_tag))
n_workers   <- if (rule_recovery == "idsweep") 1L else
  round(0.8 * max(1L, parallel::detectCores(logical = FALSE) - 1L))

# ── Smoke test: set to e.g. 1:20 to validate end-to-end fast (NULL = all) ────
smoke_ids <- NULL
```

------------------------------------------------------------------------

## Setup {#sec-setup}

```{r setup}
library(forestsearch)   # devtools::install()ed, not load_all(): workers need it
library(survival)
library(data.table)
library(foreach)
library(doFuture)
library(future)

if (n_workers > 1L) plan(multisession, workers = n_workers) else plan(sequential)

cox_formula <- build_cox_formula(outcome_name, event_name, treat_name)

# Reconstruct the sg.harm vector get_dfpred() expects from a stored rule string.
#   FS / DINA: AND-composed braced cuts joined by " & "  -> split into a vector.
#   GRF:       a single disjunction containing "|"        -> keep as one string.
rule_to_sgharm <- function(x) {
  if (is.null(x) || length(x) != 1L || is.na(x) || !nzchar(x)) return(NULL)
  if (grepl("|", x, fixed = TRUE)) return(x)                 # GRF disjunction: as-is
  trimws(strsplit(x, "&", fixed = TRUE)[[1]])                # AND form: vector of cuts
}

# Membership in the realized region on an arbitrary df, via the package's own
# evaluator (handles braces, "!{meno}" negation, and GRF disjunctions).
#   region "H"  -> treat.recommend == 0L  (the identified harm region Ĥ)
#   region "Hc" -> treat.recommend == 1L  (its complement Ĥᶜ)
in_region <- function(df, sg_harm, region = "H") {
  tr <- get_dfpred(df, sg_harm)$treat.recommend
  if (region == "H") tr == 0L else tr == 1L
}
```

------------------------------------------------------------------------

## DGM and the censored evaluation population {#sec-dgm}

```{r build-dgm}
dgm_cache <- file.path(
  cache_dir, sprintf("dgm_%s_hr%.2f_nsuper%d_seed%d.rds",
                     dgm_model, target_hr_harm, n_super, seed_base))
if (file.exists(dgm_cache)) {
  dgm <- readRDS(dgm_cache)
  message("Loaded cached DGM: ", dgm_cache)
} else {
  k_inter <- calibrate_k_inter(target_hr_harm = target_hr_harm,
                               model = dgm_model, use_ahr = FALSE)
  dgm <- setup_gbsg_dgm(model = dgm_model, k_inter = k_inter,
                        n_super = n_super, seed = seed_base)
  saveRDS(dgm, dgm_cache)
  message(sprintf("Built DGM (k_inter = %.4f); cached.", k_inter))
}
```

```{r build-eval-pop}
# One large CENSORED draw — the population at which β(g) is evaluated. Same
# pipeline (analysis_time + cens_adjust) as the per-replicate simulation data,
# so β(g) is the least-false Cox coefficient under the OBSERVED (censored) law.
eval_cache <- file.path(
  cache_dir, sprintf("evalpop_%s_n%d_at%g_ca%.3f_seed%d.rds",
                     dgm_model, n_eval, analysis_time, cens_adjust, eval_seed))
if (file.exists(eval_cache)) {
  df_eval <- readRDS(eval_cache)
  message("Loaded cached eval population: ", eval_cache)
} else {
  df_eval <- simulate_from_dgm(dgm, n = n_eval, analysis_time = analysis_time,
                               cens_adjust = cens_adjust, seed = eval_seed)
  saveRDS(df_eval, eval_cache)
  message(sprintf("Built eval population: n = %s, event rate = %.3f.",
                  format(n_eval, big.mark = ","), mean(df_eval[[event_name]])))
}
stopifnot(all(c(outcome_name, event_name, treat_name) %in% names(df_eval)))
```

------------------------------------------------------------------------

## Load the stored bundle and preflight {#sec-load}

```{r load-bundle}
# Interactive runs often have a different working directory than `quarto render`,
# so a relative `rds_dir` can miss. If the configured path is absent, try a few
# likely locations and then a recursive search from getwd(); on a unique match,
# adopt it (and re-home the sidecar beside it). Otherwise fail with diagnostics.
if (!file.exists(bundle)) {
  quick <- file.path(c(rds_dir,
                       "analysis/simulations/gbsg/resampling_comparisons/m1",
                       "../analysis/simulations/gbsg/resampling_comparisons/m1"),
                     basename(bundle))
  hit <- quick[file.exists(quick)]
  if (!length(hit))
    hit <- list.files(".", pattern = paste0("^", basename(bundle), "$"),
                      recursive = TRUE, full.names = TRUE)
  hit <- unique(hit[file.exists(hit)])
  if (length(hit) == 1L) {
    bundle      <- normalizePath(hit[1])
    rds_dir     <- dirname(bundle)                       # keep the sidecar beside it
    sidecar_rds <- file.path(rds_dir, sprintf("betaHhat_%s.rds", run_tag))
    message("Resolved bundle to: ", bundle)
  } else {
    stop("Bundle not found.",
         "\n  getwd():  ", getwd(),
         "\n  rds_dir:  ", rds_dir, "  (exists: ", dir.exists(rds_dir), ")",
         "\n  basename: ", basename(bundle),
         if (length(hit) > 1L)
           paste0("\n  Multiple matches under getwd():\n    ",
                  paste(hit, collapse = "\n    "), "\n  Set `bundle` explicitly.")
         else "\n  No matches under getwd(). Set `rds_dir`/`bundle` (absolute path safest).")
  }
}
B   <- readRDS(bundle)
res <- as.data.frame(B$results)

# Resolve the per-replicate id and the rule column (first match wins).
simid_col <- intersect(simid_cols, names(res))[1]
if (is.na(simid_col)) { res$.row_simid <- seq_len(nrow(res)); simid_col <- ".row_simid" }
rule_col  <- intersect(rule_cols, names(res))[1]   # NA => not stored in the bundle

rule_mode <- if (rule_recovery == "idsweep") "idsweep" else
             if (rule_recovery == "reidentify") "reidentify" else
             if (!is.na(rule_col)) "bundle" else "reidentify"

cat("Bundle:", basename(bundle), "\n")
cat(sprintf("  replicates: %d | detected: %d\n", nrow(res), sum(res[[detect_col]] == 1L, na.rm = TRUE)))
cat(sprintf("  sim-id column: %s | rule column: %s | rule source: %s\n",
            simid_col, if (is.na(rule_col)) "<none>" else rule_col,
            switch(rule_mode,
                   idsweep    = paste0("id-sweep output (", basename(idsweep_rds), ")"),
                   bundle     = "bundle column",
                   reidentify = "re-identify (seed-linked forestsearch = id-sweep record_replicate)")))

# Verify the endpoint columns the re-scoring needs are present.
endpoint_cols <- unlist(lapply(est_keys, function(k)
  unlist(lapply(regions, function(r) paste0(k, "_", r, c("_est","_lo","_hi"))))))
missing <- setdiff(endpoint_cols, names(res))
if (length(missing))
  stop("Bundle is missing expected endpoint columns:\n  ",
       paste(missing, collapse = ", "),
       "\nEdit `est_keys` / `regions` in the config to match this bundle.")
```

------------------------------------------------------------------------

## Recover the realized rule per replicate {#sec-recover}

```{r recover-rules}
det_idx <- which(res[[detect_col]] == 1L)
if (!is.null(smoke_ids)) det_idx <- det_idx[res[[simid_col]][det_idx] %in% smoke_ids]
det_simid <- res[[simid_col]][det_idx]

if (rule_mode == "idsweep") {
  # Read the rules the id-sweep already recovered (its `sg` column), linked by
  # sim_id. Same forestsearch identification, computed once in the sweep -- no
  # second pass here. `sg` is the " & "-joined named rule (NA where undetected).
  if (!file.exists(idsweep_rds))
    stop("rule_recovery = 'idsweep' but the id-sweep output was not found:\n  ",
         idsweep_rds,
         "\n  Run fs_subgroup_id_sweep_*.qmd (it writes results$sg by sim_id),",
         " or set rule_recovery to 'auto'/'reidentify'.")
  IS     <- as.data.frame(readRDS(idsweep_rds))
  is_sg  <- intersect(c("sg", "sg.def", "sg.harm"), names(IS))[1]
  is_sid <- intersect(simid_cols, names(IS))[1]
  if (is.na(is_sg) || is.na(is_sid))
    stop("id-sweep output lacks a rule column ('sg') or a sim-id column. Found: ",
         paste(names(IS), collapse = ", "))
  j <- match(det_simid, IS[[is_sid]])
  rule_str <- as.character(IS[[is_sg]][j])
  if ("detected" %in% names(IS)) {                  # detection sets should agree
    mis <- sum(is.na(j) | IS[["detected"]][j] != 1L)
    if (mis) message("  note: ", mis, " bundle-detected replicate(s) not flagged ",
                     "detected in the id-sweep; their rule is NA. Check the two runs ",
                     "share fs_params and the seed convention.")
  }
} else if (rule_mode == "bundle") {
  # Rule stored in the bundle: read it directly.
  rule_str <- as.character(res[[rule_col]][det_idx])
} else {
  # Re-identify seed-linked. This IS the id-sweep's record_replicate:
  # re-simulate from seed_base + sim_id, run forestsearch, read fs$sg.harm.
  reident_one <- function(sid) {
    df <- tryCatch(
      simulate_from_dgm(dgm, n = n_sample, analysis_time = analysis_time,
                        cens_adjust = cens_adjust, seed = seed_base + sid),
      error = function(e) NULL)
    if (is.null(df)) return(NA_character_)
    df[[id_name]] <- seq_len(nrow(df))
    confs <- intersect(confounders_base, names(df))
    fixed_args <- list(
      df.analysis = df, outcome.name = outcome_name, event.name = event_name,
      treat.name = treat_name, id.name = id_name, flag_harm.name = harm_col,
      confounders.name = confs, is.RCT = TRUE, seedit = seed_base + sid,
      quiet = TRUE, parallel_args = list(plan = "sequential"), debias_gate = FALSE)
    fs <- tryCatch(do.call(forestsearch, c(fixed_args, fs_params)),
                   error = function(e) NULL)
    if (is.null(fs) || is.null(fs$sg.harm) || all(is.na(fs$sg.harm)))
      return(NA_character_)
    paste(fs$sg.harm, collapse = " & ")
  }
  rule_list <- foreach(sid = det_simid, .options.future = list(seed = TRUE)) %dofuture% {
    reident_one(sid)
  }
  rule_str <- unlist(rule_list)
  plan(sequential); gc(verbose = FALSE)
}
names(rule_str) <- det_simid
ok_rule  <- !is.na(rule_str) & nzchar(rule_str)
cat(sprintf("Recovered a rule for %d / %d detected replicates.\n",
            sum(ok_rule), length(rule_str)))
```

------------------------------------------------------------------------

## β(g) once per distinct rule {#sec-beta}

```{r beta-per-rule}
# β(g) on the HR scale for a region, computed on the fixed censored eval pop.
beta_hr <- function(sg_harm, region) {
  idx <- in_region(df_eval, sg_harm, region)
  if (sum(idx) < 20L || length(unique(df_eval[[treat_name]][idx])) < 2L)
    return(NA_real_)
  est_log <- tryCatch(
    get_Cox_sg(df_eval[idx, , drop = FALSE], cox.formula = cox_formula,
               est.loghr = TRUE)$est_obs,
    error = function(e) NA_real_)
  unname(exp(est_log))                       # HR scale, matching stored endpoints
}

distinct_rules <- unique(rule_str[ok_rule])
cat(sprintf("Distinct realized rules: %d (over %d detections).\n",
            length(distinct_rules), sum(ok_rule)))

lut <- data.frame(rule = distinct_rules, betaH_hr = NA_real_, betaHc_hr = NA_real_,
                  stringsAsFactors = FALSE)
for (i in seq_len(nrow(lut))) {
  sg <- rule_to_sgharm(lut$rule[i])
  lut$betaH_hr[i]  <- beta_hr(sg, "H")
  lut$betaHc_hr[i] <- beta_hr(sg, "Hc")
}
# Map each replicate to its β(Ĥ), β(Ĥᶜ).
m <- match(rule_str, lut$rule)
betaH_hr  <- lut$betaH_hr[m]
betaHc_hr <- lut$betaHc_hr[m]
```

------------------------------------------------------------------------

## Re-score the stored endpoints against β(Ĥ) {#sec-score}

```{r score}
.cover <- function(target, lo, hi) ifelse(is.finite(target) & is.finite(lo) & is.finite(hi),
                                          target >= lo & target <= hi, NA)
tgt <- list(H = betaH_hr, Hc = betaHc_hr)

# Per-replicate frame; accumulate a coverage column per estimator × region.
per <- data.frame(sim_id = det_simid, rule = rule_str,
                  betaH_hr = betaH_hr, betaHc_hr = betaHc_hr,
                  stringsAsFactors = FALSE)

summ_rows <- list()
for (rg in names(regions)) {
  target <- tgt[[rg]]
  for (lab in names(est_keys)) {
    k  <- est_keys[[lab]]
    e  <- res[[paste0(k, "_", regions[[rg]], "_est")]][det_idx]
    lo <- res[[paste0(k, "_", regions[[rg]], "_lo")]][det_idx]
    hi <- res[[paste0(k, "_", regions[[rg]], "_hi")]][det_idx]
    cov <- .cover(target, lo, hi)
    rb  <- (e - target) / target
    per[[paste0("cov_", k, "_", rg)]] <- as.integer(cov)
    summ_rows[[length(summ_rows) + 1L]] <- data.frame(
      region = rg, estimator = lab,
      n = sum(is.finite(target) & is.finite(lo) & is.finite(hi)),
      cover_betaHhat   = mean(cov, na.rm = TRUE),
      relbias_betaHhat = 100 * mean(rb[is.finite(rb)]),
      stringsAsFactors = FALSE)
  }
}
summ <- do.call(rbind, summ_rows)
summ$region    <- factor(summ$region, levels = c("H", "Hc"))
summ$estimator <- factor(summ$estimator, levels = names(est_keys))
summ <- summ[order(summ$region, summ$estimator), ]
rownames(summ) <- NULL
summ
```

```{r save-sidecar}
sidecar <- list(
  summary    = summ,
  per_replicate = per,
  distinct_rules = lut,
  meta = list(run_tag = run_tag, bundle = bundle, n_eval = n_eval,
              eval_seed = eval_seed, analysis_time = analysis_time,
              cens_adjust = cens_adjust, n_detected = length(det_idx),
              n_rule_recovered = sum(ok_rule), n_distinct_rules = nrow(lut),
              rule_recovery = rule_mode,
              idsweep_src = if (rule_mode == "idsweep") idsweep_rds else NA_character_,
              eval_event_rate = mean(df_eval[[event_name]]),
              built = Sys.time()))
saveRDS(sidecar, sidecar_rds)
cat("Wrote sidecar:", sidecar_rds, "\n")
```

------------------------------------------------------------------------

## Notes for reuse

- **Other scenarios** — repoint `bundle` and `run_tag` and adjust the DGM knobs
  (`target_hr_harm`, `n_sample`, `analysis_time`, `cens_adjust`) to match that
  bundle's generation; the eval population is cached per (n_eval, analysis_time,
  cens_adjust, seed). DINA/GRF bundles work unchanged provided their stored rule
  (or the re-identify `fs_params`) is set to that engine.
- **Rule recovery** — `"idsweep"` reads the rules the id-sweep already saved
  (`results$sg`, linked by `sim_id`): no forestsearch re-run, and it reuses the
  identification you already trust. `"auto"` reads a rule column from the bundle
  if one is present, else re-identifies. `"reidentify"` always re-runs the
  seed-linked `forestsearch` -- which is exactly the id-sweep's `record_replicate`
  and, being deterministic in `seed_base + sim_id`, reproduces the bundle's
  region when `fs_params` match the run that built it.
- **β(g) stability** — increase `n_eval` if distinct-rule β estimates look noisy;
  1e5 is typically ample for a stable least-false Cox coefficient.
- **Validation** — set `smoke_ids <- 1:20` (and, for re-identify, `n_workers`
  small) to exercise the full path in well under a minute before the full run.
- The sidecar is consumed by `_betaHhat_coverage_table.qmd`, which renders the
  β(Ĥ) coverage and relative-bias columns next to the existing estimation table.
```
