```{r lit-gbsg-fallbacks}
#| include: false

# When this file is rendered as a stand-alone document for review,
# the parent vignette's setup chunk has not run, so any inline R
# references like `r gbsg_estrec_cut` fall out of scope.  We supply
# fallback values matching the parent vignette's setup chunk so
# inline references resolve.  When this file is included as content
# in the parent vignette, the parent's setup chunk runs first and
# the fallbacks here become no-ops because the variables already
# exist.
if (!exists("gbsg_estrec_cut"))   gbsg_estrec_cut   <- 0L
if (!exists("gbsg_n"))            gbsg_n            <- 686L
if (!exists("gbsg_itt_hr"))       gbsg_itt_hr       <- 0.69
if (!exists("gbsg_itt_hr_lcl"))   gbsg_itt_hr_lcl   <- 0.54
if (!exists("gbsg_itt_hr_ucl"))   gbsg_itt_hr_ucl   <- 0.89
```

## Literature on Subgroup Discovery in GBSG {#sec-literature-gbsg}

The German Breast Cancer Study Group dataset [@schumacher1994; @sauerbrei1999]
— commonly distributed in R as `TH.data::GBSG2`, in Python as
`lifelines.datasets.gbsg2`, and as part of the train/test split popularised
by Katzman et al. [@katzman2018] — is among the most heavily reused public
benchmark datasets in survival analysis. The trial enrolled
*N* = `r gbsg_n` node-positive breast cancer patients in a 2 × 2 randomised
design comparing duration of CMF chemotherapy with or without 2 years of
tamoxifen [@schumacher1994]. Methods papers spanning three decades have used
GBSG to illustrate fractional-polynomial prognostic modelling, random
survival forests, conditional inference forests, deep survival neural
networks, treatment-recommendation systems, transformation models,
penalised Cox regression, and recursive partitioning.

Despite the breadth of methodological reuse, GBSG occupies a notably
different position from ACTG175 in the heterogeneous-treatment-effect
literature. The dominant uses of GBSG are *prognostic* — predicting
survival from baseline covariates — rather than *predictive* in the
HTE/CATE sense (predicting differential treatment effect). Papers that
explicitly identify GBSG-based treatment subgroups are comparatively rare,
which means the forestsearch result on this dataset is validated more
strongly by the *external clinical literature* on tamoxifen and estrogen-
receptor (ER) status than by competing methodological subgroup identifications.

```{r lit-gbsg-table}
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Selected studies that have analysed GBSG/GBSG2 or tamoxifen ER-status interaction. The literature divides into a methods column (using GBSG as a benchmark) and a clinical column (independently establishing the ER mechanism). Outcome abbreviations: RFS = recurrence-free survival; OS = overall survival; treatment recommender = personalised survival prediction with counterfactual treatment assignment."

library(gt)

lit_gbsg <- data.frame(
  citation = c(
    "Schumacher et al. (1994), J Clin Oncol",
    "Sauerbrei & Royston (1999), JRSS-A",
    "Hothorn et al. (HSAUR/ipred/party), 2003+",
    "Royston & Altman (2013), BMC MedResMethodol",
    "Ishwaran et al. (2008), Annals Appl Stat",
    "Katzman et al. (2018), BMC MedResMethodol (DeepSurv)",
    "Kvamme et al. (2019), JMLR (Cox-Time, DeepHit)",
    "Kundu & Ghosh (2021), SAM (SurvCART)",
    "Hothorn / Bühlmann boosting & tram lineage",
    "Kuo et al. (2024), arXiv (CK4Gen synthetic GBSG2)",
    "EBCTCG (2011), Lancet (patient-level meta-analysis)",
    "Cuzick et al. (2013), Lancet (prevention meta-analysis)",
    "Allison et al. (2020), J Clin Oncol (ASCO/CAP guideline)",
    "León et al. (2024), Stat Med (forestsearch)"
  ),
  method = c(
    "Original RCT; stratified Cox PH; protocol-specified subgroup analyses",
    "Fractional polynomials for multivariable prognostic modelling",
    "Conditional inference trees / forests; survival ensembles",
    "Cox model external validation methodology (Rotterdam cohort)",
    "Random survival forest (RSF)",
    "Deep neural network Cox PH; treatment recommender system",
    "Discrete-time and continuous-time deep survival networks",
    "Survival CART (recursive partitioning via parameter-instability tests)",
    "Penalised Cox / boosting / transformation models",
    "Knowledge-distillation synthetic-data generation",
    "Patient-level meta-analysis of adjuvant tamoxifen efficacy by ER status",
    "Patient-level meta-analysis of tamoxifen prevention trials by ER status",
    "Hormone-receptor testing guideline (ER cut-points and predictive use)",
    "All-possible-subgroups screen with consistency criterion + bootstrap bias correction"
  ),
  context = c(
    "Methods (clinical RCT)",
    "Methods (prognostic)",
    "Methods (prognostic / benchmark)",
    "Methods (validation)",
    "Methods (prognostic)",
    "Methods (HTE / treatment recommender)",
    "Methods (prognostic / benchmark)",
    "Methods (subgroup detection)",
    "Methods (prognostic)",
    "Methods (synthetic data)",
    "Clinical (predictive biomarker)",
    "Clinical (predictive biomarker)",
    "Clinical (guideline)",
    "Methods (subgroup identification)"
  ),
  outcome = c(
    "RFS, OS",
    "RFS",
    "RFS",
    "RFS (Rotterdam)",
    "RFS / general survival",
    "RFS (Rot. + GBSG split, also illustrative recommender)",
    "RFS (Rot. + GBSG split)",
    "Generic survival (with GBSG2 as worked example)",
    "RFS",
    "RFS (synthetic)",
    "Recurrence, breast-cancer mortality (15-yr)",
    "Breast-cancer incidence (prevention)",
    "n/a",
    "RFS"
  ),
  candidate_set = c(
    "Pre-specified: hormonal therapy, chemotherapy duration",
    "8 candidate prognostic factors (age, tsize, tgrade, pnodes, progrec, estrec, menostat, horTh)",
    "Same 8 covariates as Sauerbrei/Royston",
    "8 covariates incl. ER and PR",
    "8 GBSG2 covariates as standard input",
    "8 covariates standardised; treatment indicator as one of them",
    "8 covariates per Katzman et al. preprocessing",
    "8 GBSG2 covariates as splitting candidates",
    "8 GBSG2 covariates",
    "8 covariates retained from ground truth",
    "Patient-level data from 20 trials, n = 21 457; ER and PR central",
    "5 prevention trials; ER status central",
    "n/a",
    "8 covariates; lasso pre-selects Grade, Size, Nodes, Prog; GRF adds Estrogen≤0"
  ),
  identified_role = c(
    paste0("Clinical-ITT analysis: tamoxifen vs no tamoxifen, ",
           "RR = 0.75 (0.54, 1.04) for RFS, P = 0.08 — non-significant ",
           "overall. PNODES and PROGREC the dominant prognostic factors."),
    paste0("Multivariable prognostic model with non-linear effects of ",
           "age, tsize, pnodes, progrec via fractional polynomials. ",
           "Establishes the canonical 8-covariate framing reused downstream."),
    paste0("Cforest and ctree treat GBSG2 as a benchmark for ensemble ",
           "survival prediction; not designed to identify HTE subgroups."),
    paste0("Develops external-validation pipeline using Rotterdam tumor ",
           "bank as a held-out cohort for Cox prognostic models."),
    paste0("RSF on GBSG2 ranks PNODES, PROGREC, and TSIZE as top ",
           "prognostic variables; emphasis is risk prediction, not HTE."),
    paste0("DeepSurv trains on Rotterdam, tests on GBSG; reports ",
           "survival gains under model-recommended treatment vs ",
           "observed assignment. Treatment recommender output, not an ",
           "explicit subgroup."),
    paste0("Cox-Time / DeepHit / Logistic-Hazard benchmarks on the same ",
           "Rot. + GBSG split; concordance and IBS rather than subgroup ",
           "discovery."),
    paste0("Documentation example uses GBSG2 to illustrate SurvCART; ",
           "identifies splits primarily on PNODES and PROGREC, ",
           "consistent with the prognostic literature."),
    paste0("Boosting algorithms and transformation-model variants used ",
           "GBSG2 in original papers as illustrative survival example; ",
           "again, prognostic rather than HTE focus."),
    paste0("Generates synthetic GBSG2 versions retaining covariate ",
           "structure for benchmarking; treats Sauerbrei/Royston ",
           "specification as the ground-truth model."),
    paste0("Recurrence reductions seen across all subgroups for ER+ ",
           "disease; ER-negative disease shows essentially no benefit ",
           "(RR ≈ 1 with wide CI). Tamoxifen efficacy depends ",
           "essentially on ER status — *the* established clinical fact."),
    paste0("In tamoxifen prevention trials, no effect for ER-negative ",
           "breast cancers (HR 1.22 [0.89, 1.67]), supporting the ",
           "ER-mechanism interpretation."),
    paste0("Recommends ≥ 1% nuclear ER staining as the threshold for ",
           "ER-positive status and for offering endocrine therapy. ",
           "Substantively defines what 'ER = 0' means clinically."),
    paste0("Identifies Estrogen≤0 (estrec ≤ ", gbsg_estrec_cut,
           " fmol) as Ĥ with consistency rate 95.1%; Ĥc = 88% of ITT ",
           "with bias-corrected HR 0.64 (0.44, 0.93). Recovers the ",
           "ER-mechanism subgroup *without* prior knowledge of biology.")
  ),
  stringsAsFactors = FALSE
)

gt(lit_gbsg) |>
  cols_label(
    citation        = "Citation",
    method          = "Method",
    context         = "Context",
    outcome         = "Outcome",
    candidate_set   = "Candidate covariates",
    identified_role = md("Identified role of **Estrogen / ER** (or prognostic substitute)")
  ) |>
  tab_options(
    table.font.size           = px(11),
    column_labels.font.weight = "bold",
    data_row.padding          = px(4),
    table.border.top.style    = "none",
    table.border.bottom.style = "none"
  ) |>
  cols_width(
    citation        ~ pct(15),
    method          ~ pct(17),
    context         ~ pct(10),
    outcome         ~ pct(8),
    candidate_set   ~ pct(20),
    identified_role ~ pct(30)
  )
```

### Methodological breadth

The studies in the table above split into three distinct families,
addressing GBSG from genuinely different starting points and asking
different questions.

**Prognostic-modelling tradition.** Sauerbrei & Royston [@sauerbrei1999],
Royston & Altman [@royston2013], Ishwaran et al. [@ishwaran2008] (random
survival forests), and the Hothorn/Bühlmann transformation-model lineage
all use GBSG as a benchmark for the question "given baseline covariates,
how well can we predict survival." This is the dominant use of the dataset
and shapes its canonical 8-covariate framing. None of these methods is
designed to identify treatment-effect subgroups.

**Deep survival / treatment-recommender tradition.** DeepSurv [@katzman2018]
introduced the now-standard split — train on the Rotterdam tumor bank
[@foekens2000], test on GBSG — and recast the problem as personalised
survival prediction, with counterfactual treatment assignment giving an
implicit recommender. Many follow-ups (Cox-Time, DeepHit, DySurv,
Survival MDN, NeuralSurv) [@kvamme2019] adopt the same split as a
benchmark for concordance and integrated-Brier-score performance.
The output of these systems is a per-patient recommendation rather
than a discrete subgroup defined by interpretable cuts.

**Subgroup-identification tradition (small).** Compared with ACTG175,
remarkably few methods papers explicitly identify treatment-effect
subgroups in GBSG. SurvCART [@kundu_ghosh2021] uses GBSG2 in its package
documentation as a recursive-partitioning example, but its splits are
prognostic (PNODES, PROGREC) rather than treatment-modifying. The
forestsearch result [@leon2024_forestsearch] is one of the few methods
papers to identify a treatment-modifying subgroup on GBSG and report
bias-corrected effect estimates within it.

**Clinical / predictive-biomarker tradition.** The strongest external
evidence for the forestsearch finding is not in the methods literature
at all — it is in the clinical meta-analyses by the EBCTCG [@ebctcg2011]
and Cuzick et al. [@cuzick2013], which independently established that
tamoxifen efficacy depends essentially on ER status, with no benefit in
ER-negative disease. The ASCO/CAP guideline [@allison2020] provides the
clinical definition that anchors what "ER = 0" means in practice.

These traditions disagree on what counts as the "right" answer: a
prognostic risk score, a treatment-recommendation policy, an
interpretable subgroup, or a meta-analytic biomarker. They nevertheless
converge on the same factual claim — that ER status governs tamoxifen
efficacy — even though only the clinical and forestsearch traditions
explicitly target *that* claim.

### Convergent and divergent findings

**The strongest, most consistent signal is ER status.** The clinical
meta-analyses [@ebctcg2011; @cuzick2013] establish unambiguously that
the proportional reduction in recurrence and mortality from adjuvant
tamoxifen is concentrated in ER-positive disease, and that ER-negative
disease shows essentially no benefit. The ASCO/CAP guideline
[@allison2020] formalises this by recommending ≥ 1% nuclear ER staining
as the cut-point for offering endocrine therapy. The forestsearch
result on GBSG2 — that the maximally-consistent harmful-or-null
subgroup is exactly Estrogen ≤ `r gbsg_estrec_cut` — recovers this
mechanism *without* prior biological input.

**The prognostic literature points elsewhere.** Sauerbrei & Royston
[@sauerbrei1999] and the random-survival-forest literature
[@ishwaran2008] consistently identify PNODES (number of positive
lymph nodes) and PROGREC (progesterone receptor) as the dominant
*prognostic* variables. These factors predict who is at higher
*absolute* risk regardless of treatment, which is a different question
from who *benefits more* from tamoxifen. The two literatures agree
without conflict because they target different estimands.

**The deep-survival literature does not produce a subgroup.** DeepSurv
[@katzman2018] and its successors output a continuous treatment-
recommendation score per patient. Whether the score implicitly carves
out an ER-defined subgroup at the cohort level is rarely reported,
since these papers are evaluated on concordance and Brier score
rather than on subgroup recovery. As a result, comparing forestsearch
to DeepSurv on GBSG is not directly possible without re-running both
and post-hoc summarising the recommender output.

**Sample-size caveat.** GBSG is small (*N* = `r gbsg_n`) compared with
the trials in the EBCTCG meta-analysis (*n* = 21 457). The forestsearch
ER ≤ `r gbsg_estrec_cut` subgroup has just over 80 patients
(approximately 12% of the cohort), with the bias-corrected HR
confidence interval (0.86, 2.9) wide enough that the result is best
read as *consistent with* the established clinical mechanism, not as
independent confirmation.

### Connection to the present analysis

The forestsearch ITT estimate on GBSG2 (HR =
`r gbsg_itt_hr` [`r gbsg_itt_hr_lcl`, `r gbsg_itt_hr_ucl`]) sits
within a 2 × 2 design where everyone received CMF chemotherapy and the
randomised contrast is hormonal therapy yes vs no
[@schumacher1994]. The subgroup result — that Estrogen ≤
`r gbsg_estrec_cut` may not benefit from tamoxifen, while the
complementary subgroup of ER-positive patients shows a slightly
stronger benefit (HR 0.64 [0.44, 0.93]) than the overall ITT — is
biologically expected because tamoxifen is a selective ER modulator
and acts mechanistically through the ER pathway [@allison2020;
@ebctcg2011].

The methodological contribution is therefore not the *finding* (which
the clinical literature had already established by the late 1990s)
but the *recovery of the finding from data* using only the 8 routine
baseline covariates and a procedure that does not pre-specify ER
status as the candidate effect modifier. This is a different kind of
demonstration than the ACTG175 application, where the forestsearch
finding (Preanti × Age) competed with a substantial prior HTE
literature on the same dataset.

::: {.callout-note title="What this literature does NOT establish"}
The convergent ER signal does not, on its own, validate any single
forestsearch threshold or any specific cut on the continuous estrogen-
receptor measurement. The clinical literature reports ≥ 1% as the
practical positivity threshold [@allison2020], the EBCTCG observed
substantial benefit at 10–19 fmol/mg [@ebctcg2011], and the
forestsearch result identifies estrec = 0 as the discrete cut. These
are all consistent with the same underlying biology but are not
identical operationalisations. The GBSG sample is also small relative
to the meta-analyses, and external validation on the Rotterdam cohort
[@foekens2000] shows the same direction but with the ER-negative
subgroup point estimate trending toward favourable benefit (HR 0.55,
[0.30, 1.01]) rather than the harm direction seen in GBSG2.
:::

<!-- ─────────────────────────────────────────────────────────────
     Bibliography entries (BibTeX format) for the @key citations
     used above. Move these into the vignette's references.bib or
     paste into a YAML `bibliography:` file. The keys match the
     in-text @cite calls. If the parent vignette already has a
     references.bib, deduplicate before merging.
     ───────────────────────────────────────────────────────── -->

<!--
@article{schumacher1994,
  author  = {Schumacher, M. and Bastert, G. and Bojar, H. and
             Hu{\"e}bner, K. and Olschewski, M. and Sauerbrei, W. and
             Schmoor, C. and Beyerle, C. and Neumann, R. L. A. and
             Rauschecker, H. F.},
  title   = {Randomized 2 x 2 trial evaluating hormonal treatment
             and the duration of chemotherapy in node-positive
             breast cancer patients. {German Breast Cancer Study Group}},
  journal = {Journal of Clinical Oncology},
  volume  = {12},
  number  = {10},
  pages   = {2086--2093},
  year    = {1994},
  doi     = {10.1200/JCO.1994.12.10.2086}
}

@article{sauerbrei1999,
  author  = {Sauerbrei, W. and Royston, P. and Bojar, H. and
             Schmoor, C. and Schumacher, M.},
  title   = {Modelling the effects of standard prognostic factors
             in node-positive breast cancer},
  journal = {British Journal of Cancer},
  volume  = {79},
  number  = {11--12},
  pages   = {1752--1760},
  year    = {1999}
}

@article{royston2013,
  author  = {Royston, P. and Altman, D. G.},
  title   = {External validation of a {Cox} prognostic model:
             principles and methods},
  journal = {BMC Medical Research Methodology},
  volume  = {13},
  pages   = {33},
  year    = {2013},
  doi     = {10.1186/1471-2288-13-33}
}

@article{ishwaran2008,
  author  = {Ishwaran, H. and Kogalur, U. B. and Blackstone, E. H.
             and Lauer, M. S.},
  title   = {Random survival forests},
  journal = {Annals of Applied Statistics},
  volume  = {2},
  number  = {3},
  pages   = {841--860},
  year    = {2008},
  doi     = {10.1214/08-AOAS169}
}

@article{katzman2018,
  author  = {Katzman, J. L. and Shaham, U. and Cloninger, A. and
             Bates, J. and Jiang, T. and Kluger, Y.},
  title   = {{DeepSurv}: personalized treatment recommender system
             using a {Cox} proportional hazards deep neural network},
  journal = {BMC Medical Research Methodology},
  volume  = {18},
  number  = {1},
  pages   = {24},
  year    = {2018},
  doi     = {10.1186/s12874-018-0482-1}
}

@article{kvamme2019,
  author  = {Kvamme, H. and Borgan, {\O}. and Scheel, I.},
  title   = {Time-to-event prediction with neural networks and
             {Cox} regression},
  journal = {Journal of Machine Learning Research},
  volume  = {20},
  number  = {129},
  pages   = {1--30},
  year    = {2019}
}

@article{kundu_ghosh2021,
  author  = {Kundu, M. G. and Ghosh, S.},
  title   = {Survival trees based on heterogeneity in time-to-event
             and censoring distributions using parameter instability
             test},
  journal = {Statistical Analysis and Data Mining: The ASA Data
             Science Journal},
  volume  = {14},
  number  = {5},
  pages   = {466--483},
  year    = {2021},
  doi     = {10.1002/sam.11534}
}

@article{foekens2000,
  author  = {Foekens, J. A. and Peters, H. A. and Look, M. P. and
             others},
  title   = {The urokinase system of plasminogen activation and
             prognosis in 2780 breast cancer patients},
  journal = {Cancer Research},
  volume  = {60},
  number  = {3},
  pages   = {636--643},
  year    = {2000}
}

@article{ebctcg2011,
  author  = {{Early Breast Cancer Trialists' Collaborative Group}},
  title   = {Relevance of breast cancer hormone receptors and other
             factors to the efficacy of adjuvant tamoxifen:
             patient-level meta-analysis of randomised trials},
  journal = {Lancet},
  volume  = {378},
  number  = {9793},
  pages   = {771--784},
  year    = {2011},
  doi     = {10.1016/S0140-6736(11)60993-8}
}

@article{cuzick2013,
  author  = {Cuzick, J. and Sestak, I. and Bonanni, B. and others},
  title   = {Selective oestrogen receptor modulators in prevention
             of breast cancer: an updated meta-analysis of
             individual participant data},
  journal = {Lancet},
  volume  = {381},
  number  = {9880},
  pages   = {1827--1834},
  year    = {2013},
  doi     = {10.1016/S0140-6736(13)60140-3}
}

@article{allison2020,
  author  = {Allison, K. H. and Hammond, M. E. H. and Dowsett, M.
             and others},
  title   = {Estrogen and progesterone receptor testing in breast
             cancer: {ASCO/CAP} guideline update},
  journal = {Journal of Clinical Oncology},
  volume  = {38},
  number  = {12},
  pages   = {1346--1366},
  year    = {2020},
  doi     = {10.1200/JCO.19.02309}
}

@misc{kuo2024_ck4gen,
  author  = {Kuo, N. I. and others},
  title   = {{CK4Gen}: A Knowledge Distillation Framework for
             Generating High-Utility Synthetic Survival Datasets in
             Healthcare},
  year    = {2024},
  eprint  = {2410.16872},
  archivePrefix = {arXiv}
}

@article{leon2024_forestsearch,
  author  = {Le{\'o}n, L. F. and Jemielita, T. and Guo, Z. and
             Marceau West, R. and Anderson, K. M.},
  title   = {Exploratory subgroup identification in the heterogeneous
             {Cox} model: A relatively simple procedure},
  journal = {Statistics in Medicine},
  year    = {2024},
  doi     = {10.1002/sim.10163}
}
-->
