---
title: "ForestSearch Bootstrap De-biasing — Focused Robustness Smoke-test"
subtitle: "ACTG175 binary (OR scale); targets the Finding 1 / 2 / 4 guards"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    theme: cosmo
---

```{r}
#| label: setup
#| message: false
#| warning: false
# Run AFTER devtools::install(): the Tier-1 bootstrap spawns parallel workers,
# which see only the *installed* package (load_all() is not enough).
library(forestsearch)
library(speff2trial)
library(data.table)
library(future)

set.seed(20260613)
analysis_seed <- 8675309L

# ---- knobs (single top chunk) -------------------------------------------
# Deliberately small: this is a smoke-test, not a simulation study.
NB_SMOKE        <- 60L                                   # Tier-1 bootstrap resamples
n_cores         <- max(1L, parallel::detectCores() - 2L)
run_integration <- TRUE   # FALSE -> run only Section B (fast deterministic probes)

# FS binary (OR, adverse scale) — mirrors analysis_actg175_binary_multimethod.qmd
fs_outcome_type       <- "binary"
fs_effect_measure     <- "OR"
fs_or_threshold       <- 1.25     # harm floor on the OR scale (lower to force identification)
fs_or_consistency     <- 1.0
fs_pconsistency       <- 0.90
fs_consistency_method <- "resample"
```

This is a **focused** smoke-test for the three robustness fixes made to the
Tier-1 bootstrap de-biasing path (`forestsearch_bootstrap_dofuture()` and the
helpers it calls). It is **not** a re-run of the full ACTG175 analysis; it is
scoped to exactly the changed behaviour:

- **Finding 1** — `Ystar_mat` is realigned to the surviving `boot_id`s before the
  infinitesimal-jackknife (IJ) variance is computed, and `get_targetEst()` now
  asserts `length(x) == nrow(ystar)` rather than recycling silently.
- **Finding 2** — non-finite per-resample subgroup fits (`Inf`/`NaN` from
  (near-)separated or single-arm subgroups) are mapped to `NA` in the worker so
  they drop from the bias average and the IJ variance, instead of propagating an
  `Inf` that would void the whole estimate.
- **Finding 4** — `get_targetEst()` `match.arg`s `cov_method`, and the
  `"nocorrect"` branch carries the `na.rm = TRUE` its sibling already had.

Two layers: **Section A** runs the patched bootstrap end-to-end in the realistic
binary setting (a regression check that normal runs are unchanged); **Section B**
exercises each guard's *firing* path deterministically, with no reliance on a
particular resample outcome.

```{r}
#| label: pass-fail-harness
# Lightweight PASS/FAIL recorder (no testthat scaffolding, per project style).
.smoke <- new.env()
.smoke$log <- list()
check <- function(label, ok, detail = "") {
  ok <- isTRUE(ok)
  .smoke$log[[length(.smoke$log) + 1L]] <- data.frame(
    check = label, result = if (ok) "PASS" else "FAIL",
    detail = detail, stringsAsFactors = FALSE
  )
  cat(sprintf("[%s] %s%s\n", if (ok) "PASS" else "FAIL", label,
              if (nzchar(detail)) paste0(" \u2014 ", detail) else ""))
  invisible(ok)
}
```

# Section A — end-to-end regression (patched bootstrap in situ)

Builds the ACTG175 binary adverse outcome (`y_binary_neg`, OR > 1 = harm),
identifies a subgroup with the consistency engine, and runs the Tier-1 bootstrap.
The screening engines (lasso/GRF) are switched off for speed — the de-biasing
code under test is engine-agnostic, so consistency-only fully exercises it. The
whole block is wrapped so an environment failure is recorded rather than aborting
Section B.

```{r}
#| label: integration
#| message: false
#| warning: true
fs_bc <- NULL
if (isTRUE(run_integration)) {

  # ---- data: ACTG175, arms 1 (ZDV+ddI) vs 3 (ddI) -----------------------
  actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
  actg_df$id    <- seq_len(nrow(actg_df))
  actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)
  actg_df <- actg_df[!is.na(actg_df$cd420), ]
  actg_df$y_binary_neg <- 1L - as.integer(actg_df$cd420 > actg_df$cd40)  # 1 = no improvement

  cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
  bin_vars  <- c("hemo", "homo", "drugs", "race", "gender", "symptom")
  for (v in bin_vars) actg_df[[v]] <- as.factor(actg_df[[v]])
  confounders.name <- c(cont_vars, bin_vars)

  msg <- ""
  integ_ok <- tryCatch({
    fs <- forestsearch(
      df.analysis      = actg_df,
      confounders.name = confounders.name,
      outcome.name     = "y_binary_neg",
      event.name       = "y_binary_neg",   # binary: outcome IS the event
      treat.name       = "treat",
      id.name          = "id",
      is.RCT           = TRUE,
      seedit           = analysis_seed,
      outcome_type     = fs_outcome_type,
      effect_measure   = fs_effect_measure,
      adverse_outcome  = TRUE,
      # screening off for speed; bootstrap fixes are engine-agnostic
      use_lasso        = FALSE,
      use_grf          = FALSE,
      subgroup_method  = "consistency",
      cut_type         = "default",
      cont.cutoff      = 4L,
      n.min            = 60L,
      d0.min           = 10L,
      d1.min           = 10L,
      maxk             = 2L,
      effect.threshold       = fs_or_threshold,
      consistency.threshold  = fs_or_consistency,
      pconsistency.threshold = fs_pconsistency,
      consistency_method     = fs_consistency_method,
      fs.splits        = 200L,
      use_twostage     = TRUE,
      debias_gate      = FALSE,
      details          = FALSE,
      quiet            = TRUE,
      plot.sg          = FALSE,
      parallel_args    = list(plan = "multisession", workers = n_cores)
    )
    plan("sequential"); gc(verbose = FALSE)

    fs_bc <<- forestsearch_bootstrap_dofuture(
      fs.est        = fs,
      nb_boots      = NB_SMOKE,
      show_three    = FALSE,
      details       = FALSE,
      parallel_args = list(plan = "multisession", workers = n_cores)
    )
    plan("sequential"); gc(verbose = FALSE)
    TRUE
  }, error = function(e) { msg <<- conditionMessage(e); FALSE })

  check("A1 integration: end-to-end run completed without error", integ_ok,
        if (!integ_ok) msg else "")

  if (isTRUE(integ_ok)) {
    # Edit B guarantee: IJ weights line up row-for-row with the per-resample
    # adjustments — holds whether or not any iteration was dropped/realigned.
    check("A2 integration: IJ alignment invariant length(H_biasadj_2)==nrow(Ystar_mat)",
          length(fs_bc$results$H_biasadj_2) == nrow(fs_bc$Ystar_mat),
          sprintf("len=%d, nrow=%d",
                  length(fs_bc$results$H_biasadj_2), nrow(fs_bc$Ystar_mat)))

    if (!is.null(fs_bc$H_estimates)) {
      check("A3 integration: de-biased H2 (OR*) estimate is finite",
            is.finite(fs_bc$H_estimates$H2),
            sprintf("H2 = %.3f", fs_bc$H_estimates$H2))
      check("A4 integration: de-biased H2 (OR*) CI is finite",
            all(is.finite(c(fs_bc$H_estimates$H2_lower, fs_bc$H_estimates$H2_upper))),
            sprintf("(%.3f, %.3f)",
                    fs_bc$H_estimates$H2_lower, fs_bc$H_estimates$H2_upper))
    } else {
      cat("\nNOTE: no subgroup identified at fs_or_threshold =", fs_or_threshold,
          "\n  -> the in-situ de-biasing path was not exercised. Lower the",
          "threshold to force it.\n  Section B validates the patched code",
          "deterministically regardless.\n")
    }
  }
} else {
  cat("run_integration = FALSE; skipping Section A.\n")
}
```

# Section B — deterministic guard probes

These call the patched internal helper directly with crafted inputs, so each fix
is exercised on its firing path without depending on a random resample. A small
`B \u00d7 N` multiplicity matrix and an aligned vector of per-draw adjustments stand
in for `Ystar_mat` and `results$H_biasadj_2`.

```{r}
#| label: probe-fixtures
set.seed(101)
Bdraws <- 40L
Nsub   <- 60L
ystar  <- matrix(rpois(Bdraws * Nsub, lambda = 1), nrow = Bdraws, ncol = Nsub)  # B x N multiplicities
x_full <- rnorm(Bdraws, mean = 0.40, sd = 0.30)                                 # aligned H_biasadj_2

gte <- forestsearch:::get_targetEst   # internal helper under test
```

## Baseline: aligned inputs give a finite IJ variance

```{r}
#| label: probe-baseline
e0 <- gte(x_full, ystar)
check("B0 baseline: aligned x/ystar -> finite estimate & IJ SE",
      is.finite(e0$target_est) && is.finite(e0$sehat_new),
      sprintf("est=%.3f, se=%.4f", e0$target_est, e0$sehat_new))
```

## Finding 1 — alignment assertion and `Ystar_mat` realignment

```{r}
#| label: probe-finding1
# (a) Misaligned length must FAIL LOUDLY (the recycling bug is now an error).
err_mis <- tryCatch({ gte(x_full[-1], ystar); "no-error" },
                    error = function(e) conditionMessage(e))
check("B1a Finding 1: misaligned length(x) != nrow(ystar) triggers the assertion",
      grepl("misaligned", err_mis), err_mis)

# (b) Simulate the upstream drop that Edit B repairs: two iterations errored, so
#     only the survivors carry an adjustment; boot_id records which rows survived.
dropped  <- c(3L, 17L)
boot_id  <- setdiff(seq_len(Bdraws), dropped)   # ascending, matches results' row order
x_surv   <- x_full[boot_id]                     # length Bdraws - 2

# Without realignment (the bug): full ystar vs short x -> caught by the assertion.
caught <- tryCatch({ gte(x_surv, ystar); FALSE }, error = function(e) TRUE)
check("B1b Finding 1: dropped rows without realignment are caught (not recycled)",
      caught)

# With Edit B's realignment Ystar_mat[results$boot_id, ] -> finite IJ variance.
ystar_aln <- ystar[boot_id, , drop = FALSE]
e_aln <- gte(x_surv, ystar_aln)
check("B1c Finding 1: Ystar[boot_id,] realignment restores a finite IJ variance",
      is.finite(e_aln$target_est) && is.finite(e_aln$sehat_new),
      sprintf("se=%.4f over %d surviving draws", e_aln$sehat_new, length(boot_id)))
```

## Finding 2 — non-finite handling and the worker guard rationale

```{r}
#| label: probe-finding2
# (a) NA draws (a resample that produced no usable estimate) drop via na.rm.
x_na <- x_full; x_na[c(2L, 9L)] <- NA_real_
e_na <- gte(x_na, ystar)
check("B2a Finding 2: NA resample draws drop (na.rm); estimate stays finite",
      is.finite(e_na$target_est) && is.finite(e_na$sehat_new))

# (b) Rationale: an Inf is NOT dropped by na.rm, so it must be neutralised
#     upstream. With an Inf left in x, the IJ estimate is (correctly) non-finite
#     -- this is exactly what the worker guard prevents by mapping Inf -> NA.
x_inf <- x_full; x_inf[2L] <- Inf
e_inf <- gte(x_inf, ystar)
check("B2b Finding 2 (rationale): an Inf left in x propagates to a non-finite estimate",
      !is.finite(e_inf$target_est))

# (c) The worker guard itself: any non-finite constituent -> NA adjustment,
#     which then drops cleanly from the bias mean.
mk_adj <- function(H_obs, H_star, Hstar_star, Hstar_obs) {
  if (all(is.finite(c(H_obs, H_star, Hstar_star, Hstar_obs)))) {
    2 * H_obs - (H_star + Hstar_star - Hstar_obs)
  } else NA_real_
}
adj_ok  <- mk_adj(0.50, 0.60, 0.70, 0.40)   # all finite -> real adjustment
adj_inf <- mk_adj(0.50, Inf,  0.70, 0.40)   # one Inf     -> NA
check("B2c Finding 2: worker guard maps a non-finite fit to NA",
      is.finite(adj_ok) && is.na(adj_inf))
check("B2d Finding 2: guarded NA draw drops from the bias mean (na.rm), result finite",
      is.finite(mean(c(adj_ok, adj_inf, 0.55, 0.48), na.rm = TRUE)))
```

## Finding 4 — `match.arg(cov_method)` and `nocorrect` na.rm parity

```{r}
#| label: probe-finding4
err_typo <- tryCatch({ gte(x_full, ystar, cov_method = "nocorect"); "no-error" },
                     error = function(e) "match.arg-error")
check("B4a Finding 4: cov_method typo rejected by match.arg", err_typo == "match.arg-error")

x_na2  <- x_full; x_na2[5L] <- NA_real_
e_noc  <- gte(x_na2, ystar, cov_method = "nocorrect")
check("B4b Finding 4: cov_method='nocorrect' runs with na.rm parity (finite on NA input)",
      is.finite(e_noc$target_est))
```

# Summary

```{r}
#| label: summary
res_df <- do.call(rbind, .smoke$log)
rownames(res_df) <- NULL
knitr::kable(res_df, caption = "Focused smoke-test results")
```

```{r}
#| label: verdict
n_fail <- sum(res_df$result == "FAIL")
cat(sprintf("\n%d check(s), %d PASS, %d FAIL\n",
            nrow(res_df), sum(res_df$result == "PASS"), n_fail))
# A failing smoke-test fails the render, loudly.
stopifnot("Smoke-test: all checks must pass" = n_fail == 0L)
cat("\nALL CHECKS PASSED \u2014 the de-biasing guards behave as intended.\n")
```
