---
title: "Operating Characteristics of the Tier-2 De-biased Gate"
subtitle: "Estimation/coverage head-to-head against the Tier-1 bootstrap, on the León (2024) Table 2 design"
author: "Larry León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    html-math-method: mathjax
    theme: cosmo
    fig-cap-location: bottom
    code-fold: true
---

> **Status — smoke test.** This is the development configuration of the
> simulation: tiny counts to wire the pipeline end-to-end on a single M1 cell
> before scaling up. Every tunable lives in the setup chunk (@sec-setup); the
> counts (`n_sims`, `nb_boots`, `gate_draws`) are the only things to change to
> move from smoke test to production. Results run live — no `saveRDS`/`readRDS`
> caching.

# What this evaluates {#sec-intro}

The **Tier-2 de-biased gate** (`fs_debias_gate()`, surfaced through
`forestsearch(debias_gate = TRUE)`) is a refit-free multiplier-bootstrap
approximation of León's post-selection bias correction, with an
infinitesimal-jackknife (IJ) confidence interval. To first order its point
estimate equals León's Eq. 7 correction and its interval equals León's IJ
variance, the only approximation being the per-subgroup `dfbeta` linearization.

This document turns the single-trial validation into operating characteristics by
re-running the León (2024) estimation/coverage design and recording, **per
replicate on the same data and the same selected subgroup**:

* **oracle** $\hat\theta(H)$ — refit on the *true* harm subgroup;
* **naive** $\hat\theta(\hat H)$ — plugin on the identified subgroup;
* **Tier-1** $\hat\theta^{*}(\hat H)$ — full nonparametric bootstrap
  bias-correction (`forestsearch_bootstrap_dofuture()`, León Eq. 7, the `H2`
  variant accounting for both bias sources) with its IJ interval;
* **Tier-2** $\tilde\theta(\hat H)$ — the de-biased gate with its IJ interval.

Each is compared to three truth targets: the per-replicate **oracle** estimate,
the controlled direct effect $\theta^{\ddagger}$ (CDE), and the marginal subgroup
hazard ratio $\theta^{\dagger}$. The two outputs are (1) the León Table-2-layout
estimation/coverage table with a Tier-2 block, and (2) the paired
Tier-1$\leftrightarrow$Tier-2 agreement panel — the most direct empirical check
of the leading-order correspondence.

The search runs under `consistency_method = "resample"` and `sg_focus = "eff"`;
the gate re-selects with `reselection = "maxcons"`, which reproduces the
`sg_focus = "eff"` rule (highest consistency rate) so both tiers share the same
selection map.

# Setup — all knobs {#sec-setup}

```{r setup, message=FALSE, warning=FALSE}
# NOTE: forestsearch must be INSTALLED (devtools::install()), not load_all():
# the doFuture multisession workers spawn separate R processes that only see the
# installed package.
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(ggplot2)
library(foreach)
library(doFuture)
library(future)

# ── Smoke-test counts (the only knobs to change for production) ────────────
n_sims       <- 200L       # simulated trials                (prod: 1000 -> 20000)
nb_boots     <- 500L       # Tier-1 nonparametric bootstraps (prod: 200 -> 500)
gate_draws   <- 2000L     # Tier-2 multiplier draws         (keep ~2000)

run_smoke    <- TRUE      # master flag; FALSE short-circuits the run loop

# ── DGM: León M1 cell (GBSG AFT super-population, alt hypothesis) ───────────
dgm_model       <- "alt"
target_hr_harm  <- 2.0    # calibrate k_inter to this Cox HR in the harm subgroup
n_sample        <- 700L   # M1 sample size
analysis_time   <- 84     # administrative censoring (months)
cens_adjust     <- log(1.5)
n_super         <- 5000L

# ── Search configuration (León tuning; resample identification) ────────────
consistency_method <- "resample"   # decided: resample identification (the new default)
sg_focus           <- "eff"        # -> "hr": highest consistency rate, ties by effect
gate_reselection   <- "maxcons"    # mirrors sg_focus = "eff" for a faithful comparison
hr_threshold       <- 1.25
hr_consistency     <- 1.0
pconsistency       <- 0.90
fs_splits          <- 400L         # moot under "resample" (closed-form, not literal split)
maxk               <- 2L
n_min              <- 60L
d0_min             <- 12L
d1_min             <- 12L
use_lasso          <- FALSE
use_grf            <- TRUE         # FALSE = LASSO-only (fast, "FSl"); set TRUE for "FSlg"
use_twostage       <- TRUE

fs_conf_force      <- c("z3 == 1")
fs_conf.cont_jcuts <- list(z1 = 10)

# ── Column names emitted by simulate_from_dgm() ────────────────────────────
outcome_name   <- "y_sim"
event_name     <- "event_sim"
treat_name     <- "treat_sim"
id_name        <- "id"
harm_col       <- "flag_harm"
# GBSG covariate candidates (intersected with the simulated columns at run time)
confounders_base <- c("z1", "z2", "z3", "z4", "z5", "size", "grade3")

# ── Reproducibility / parallelism ──────────────────────────────────────────
seed_base  <- 8316951L
n_workers  <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))
fs_conf_force      <- c("z3 == 1")
fs_conf.cont_jcuts <- list(z1 = 10)
z975       <- qnorm(0.975)

cat(sprintf("Smoke test: n_sims=%d, nb_boots=%d, gate_draws=%d, workers=%d\n",
            n_sims, nb_boots, gate_draws, n_workers))
```

# Build the DGM and read off the truth targets

```{r build-dgm}
k_inter <- calibrate_k_inter(target_hr_harm = target_hr_harm,
                             model = dgm_model, use_ahr = FALSE)

dgm <- setup_gbsg_dgm(model = dgm_model, k_inter = k_inter,
                      n_super = n_super, seed = seed_base)
dgm <- compute_dgm_cde(dgm)   # attaches CDE (theta-ddagger) targets

`%||%` <- function(a, b) if (is.null(a) || length(a) == 0 || all(is.na(a))) b else a

# Fixed (population) truth targets, read defensively from the DGM object.
truth <- list(
  hr_causal   = dgm$hr_causal   %||% NA_real_,                 # overall causal HR
  marg_H      = dgm$hr_H_true   %||% NA_real_,                 # theta-dagger (H)
  marg_Hc     = dgm$hr_Hc_true  %||% NA_real_,                 # theta-dagger (Hc)
  cde_H       = dgm$cde_H       %||% (dgm$hazard_ratios$CDE_harm    %||% NA_real_),  # theta-ddagger (H)
  cde_Hc      = dgm$cde_Hc      %||% (dgm$hazard_ratios$CDE_no_harm %||% NA_real_)   # theta-ddagger (Hc)
)

cat("DGM truth targets (population):\n")
cat(sprintf("  overall causal HR      : %.3f\n", truth$hr_causal))
cat(sprintf("  theta-dagger  H / Hc   : %.3f / %.3f\n", truth$marg_H, truth$marg_Hc))
cat(sprintf("  theta-ddagger H / Hc   : %.3f / %.3f\n", truth$cde_H,  truth$cde_Hc))
```

# Per-replicate paired recorder

For each simulated trial: identify $\hat H$ and obtain the naive + Tier-2 gate in
one `forestsearch()` call, run the Tier-1 bootstrap on the *same* fit (inheriting
`consistency_method = "resample"`), and refit the oracle on the true subgroup.
Inner `forestsearch()`/bootstrap parallelism is forced sequential because the
outer loop already parallelises across trials.

```{r recorder}
# Cox HR + delta-method 95% CI on a subset (for the per-replicate oracle).
.cox_hr_ci <- function(df) {
  if (sum(df[[event_name]]) < 5L || length(unique(df[[treat_name]])) < 2L)
    return(c(est = NA_real_, lo = NA_real_, hi = NA_real_))
  fml <- stats::as.formula(
    sprintf("survival::Surv(%s, %s) ~ %s", outcome_name, event_name, treat_name))
  fit <- tryCatch(survival::coxph(fml, data = df), error = function(e) NULL)
  if (is.null(fit)) return(c(est = NA_real_, lo = NA_real_, hi = NA_real_))
  b <- unname(stats::coef(fit)[1]); se <- sqrt(diag(stats::vcov(fit))[1])
  c(est = exp(b), lo = exp(b - z975 * se), hi = exp(b + z975 * se))
}

# One all-NA record (returned on failure so the loop never aborts a cell).
.na_record <- function(sim_id) data.frame(
  sim_id = sim_id, detected = 0L, n_sel = NA_integer_,
  label = NA_character_, t1_secs = NA_real_, t2_secs = NA_real_,
  or_H_est = NA, or_H_lo = NA, or_H_hi = NA,
  or_Hc_est = NA, or_Hc_lo = NA, or_Hc_hi = NA,
  nv_H_est = NA, nv_H_lo = NA, nv_H_hi = NA,
  nv_Hc_est = NA, nv_Hc_lo = NA, nv_Hc_hi = NA,
  t1_H_est = NA, t1_H_lo = NA, t1_H_hi = NA,
  t1_Hc_est = NA, t1_Hc_lo = NA, t1_Hc_hi = NA,
  t2_H_est = NA, t2_H_lo = NA, t2_H_hi = NA, t2_H_se_ij = NA,
  t2_Hc_est = NA, t2_Hc_lo = NA, t2_Hc_hi = NA, t2_Hc_se_ij = NA,
  stringsAsFactors = FALSE)

record_replicate <- function(sim_id) {
  rec <- .na_record(sim_id)

  df <- tryCatch(
    simulate_from_dgm(dgm, n = n_sample, analysis_time = analysis_time,
                      cens_adjust = cens_adjust, seed = seed_base + sim_id),
    error = function(e) NULL)
  if (is.null(df)) return(rec)
  df[[id_name]] <- seq_len(nrow(df))
  confs <- intersect(confounders_base, names(df))

  # ---- Identification + naive + Tier-2 gate (single forestsearch call) -----
  t2_t0 <- proc.time()[3]
  fs.est <- tryCatch(
    forestsearch(
      df.analysis     = df,
      outcome.name    = outcome_name, event.name = event_name,
      treat.name      = treat_name,   id.name    = id_name,
      flag_harm.name  = harm_col,     confounders.name = confs,
      is.RCT          = TRUE,         seedit = seed_base + sim_id,
      use_lasso       = use_lasso,    use_grf = use_grf,
      sg_focus        = sg_focus,
      consistency_method = consistency_method,
      conf_force          = fs_conf_force,
     conf.cont_jcuts      = fs_conf.cont_jcuts,
      hr.threshold    = hr_threshold, hr.consistency = hr_consistency,
      pconsistency.threshold = pconsistency, fs.splits = fs_splits,
      maxk = maxk, n.min = n_min, d0.min = d0_min, d1.min = d1_min,
      use_twostage    = use_twostage,
      parallel_args   = list(plan = "sequential"),   # outer loop owns parallelism
      debias_gate     = TRUE,
      debias_gate_args = list(ci_method = "ij", draws = gate_draws,
                              reselection = gate_reselection,
                              include_complement = TRUE)
    ),
    error = function(e) NULL)
  rec$t2_secs <- proc.time()[3] - t2_t0
  if (is.null(fs.est)) return(rec)

  g <- fs.est$debias_gate
  found <- !is.null(fs.est$sg.harm) && !is.null(g)
  rec$detected <- as.integer(found)
  if (!found) return(rec)

  rec$label <- g$selected_label %||% NA_character_
  rec$n_sel <- g$n_selected      %||% NA_integer_
  rec$nv_H_est <- g$naive$est;    rec$nv_H_lo <- g$naive$lower;    rec$nv_H_hi <- g$naive$upper
  rec$t2_H_est <- g$debiased$est; rec$t2_H_lo <- g$debiased$lower; rec$t2_H_hi <- g$debiased$upper
  rec$t2_H_se_ij <- g$debiased$se_ij %||% NA_real_
  if (is.list(g$complement) && !is.null(g$complement$debiased)) {
    rec$nv_Hc_est <- g$complement$naive$est
    rec$nv_Hc_lo  <- g$complement$naive$lower
    rec$nv_Hc_hi  <- g$complement$naive$upper
    rec$t2_Hc_est <- g$complement$debiased$est
    rec$t2_Hc_lo  <- g$complement$debiased$lower
    rec$t2_Hc_hi  <- g$complement$debiased$upper
    rec$t2_Hc_se_ij <- g$complement$debiased$se_ij %||% NA_real_
  }

  # ---- Tier-1 bootstrap (inherits resample via fs.est$args_call_all) -------
  t1_t0 <- proc.time()[3]
  boot <- tryCatch(
    forestsearch_bootstrap_dofuture(
      fs.est, nb_boots = nb_boots, seed = seed_base + sim_id,
      parallel_args = list(plan = "sequential")),
    error = function(e) NULL)
  rec$t1_secs <- proc.time()[3] - t1_t0
  if (!is.null(boot) && !is.null(boot$H_estimates)) {
    he <- boot$H_estimates                     # H2 = León Eq.7 (both bias sources)
    rec$t1_H_est <- he$H2; rec$t1_H_lo <- he$H2_lower; rec$t1_H_hi <- he$H2_upper
  }
  if (!is.null(boot) && !is.null(boot$Hc_estimates)) {
    hce <- boot$Hc_estimates
    rec$t1_Hc_est <- hce$H2; rec$t1_Hc_lo <- hce$H2_lower; rec$t1_Hc_hi <- hce$H2_upper
  }

  # ---- Oracle: refit on the TRUE harm subgroup of this trial ---------------
  oH  <- .cox_hr_ci(df[df[[harm_col]] == 1, , drop = FALSE])
  oHc <- .cox_hr_ci(df[df[[harm_col]] == 0, , drop = FALSE])
  rec$or_H_est <- oH["est"];   rec$or_H_lo <- oH["lo"];   rec$or_H_hi <- oH["hi"]
  rec$or_Hc_est <- oHc["est"]; rec$or_Hc_lo <- oHc["lo"]; rec$or_Hc_hi <- oHc["hi"]

  rec
}
```

# Run the simulation

```{r run-loop, message=FALSE, warning=FALSE}
results <- NULL
if (run_smoke) {
  plan("sequential"); gc()
  plan("multisession", workers = n_workers)

  t_all <- proc.time()[3]
  results <- foreach(
    s = seq_len(n_sims), .combine = rbind, .errorhandling = "remove",
    .options.future = list(packages = c("forestsearch", "survival", "data.table"),
                           seed = TRUE)
  ) %dofuture% {
    record_replicate(s)
  }
  elapsed <- proc.time()[3] - t_all

  plan("sequential"); gc()   # release workers between phases

  cat(sprintf("Completed %d/%d replicates in %.1f s (%.1f s/replicate).\n",
              sum(results$detected, na.rm = TRUE), n_sims, elapsed,
              elapsed / max(1L, n_sims)))
  cat(sprintf("Mean wall-clock: Tier-2 gate %.3f s vs Tier-1 bootstrap %.2f s (B=%d).\n",
              mean(results$t2_secs, na.rm = TRUE),
              mean(results$t1_secs, na.rm = TRUE), nb_boots))
}
```

# Estimation / coverage table (León Table 2 layout + Tier-2)

```{r estimation-table}
# Coverage of a target by a per-replicate [lo, hi] interval.
.cover <- function(target, lo, hi) mean(target >= lo & target <= hi, na.rm = TRUE)
# Percent relative bias of an estimate vs a (scalar or per-replicate) target.
.pbias <- function(est, target) 100 * mean((est - target) / target, na.rm = TRUE)
# HR-scale SD implied by log-scale CI bounds (uniform across estimators).
.sd_hat <- function(est, lo, hi) {
  se_log <- (log(hi) - log(lo)) / (2 * z975)
  mean(est * se_log, na.rm = TRUE)
}

# Build one block (H or Hc): four estimator rows against three targets.
build_block <- function(r, suffix, tgt_cde, tgt_marg) {
  est_keys <- c(oracle = "or", naive = "nv", `Tier-1` = "t1", `Tier-2` = "t2")
  or_est <- r[[paste0("or_", suffix, "_est")]]   # per-replicate oracle target
  rows <- lapply(names(est_keys), function(lab) {
    k  <- est_keys[[lab]]
    e  <- r[[paste0(k, "_", suffix, "_est")]]
    lo <- r[[paste0(k, "_", suffix, "_lo")]]
    hi <- r[[paste0(k, "_", suffix, "_hi")]]
    data.frame(
      Estimator = lab,
      Avg     = mean(e, na.rm = TRUE),
      SD      = sd(e, na.rm = TRUE),
      SD_hat  = .sd_hat(e, lo, hi),
      Length  = mean(hi - lo, na.rm = TRUE),
      b_oracle = .pbias(e, or_est),
      b_cde    = .pbias(e, tgt_cde),
      b_marg   = .pbias(e, tgt_marg),
      Cov_oracle = .cover(or_est, lo, hi),
      Cov_cde    = .cover(tgt_cde,  lo, hi),
      Cov_marg   = .cover(tgt_marg, lo, hi),
      stringsAsFactors = FALSE)
  })
  do.call(rbind, rows)
}

if (run_smoke && !is.null(results)) {
  rH  <- results[results$detected == 1, , drop = FALSE]
  blockH  <- build_block(rH, "H",  truth$cde_H,  truth$marg_H)
  blockHc <- build_block(rH, "Hc", truth$cde_Hc, truth$marg_Hc)
  est_tab <- rbind(
    cbind(Block = "Harm  \u0124", blockH),
    cbind(Block = "Complement  \u0124\u1d9c", blockHc)
  )

  det_rate <- mean(results$detected, na.rm = TRUE)

  gt(est_tab, groupname_col = "Block") |>
    tab_header(
      title = "Tier-2 de-biased gate vs Tier-1 bootstrap: estimation & coverage",
      subtitle = sprintf("M1 smoke test \u00b7 n=%d, %d sims, B=%d, draws=%d \u00b7 resample identification",
                         n_sample, n_sims, nb_boots, gate_draws)) |>
    fmt_number(columns = c(Avg, SD, SD_hat, Length), decimals = 2) |>
    fmt_number(columns = c(b_oracle, b_cde, b_marg), decimals = 1) |>
    fmt_number(columns = c(Cov_oracle, Cov_cde, Cov_marg), decimals = 2) |>
    cols_label(
      SD_hat = md("$\\widehat{\\text{SD}}$"),
      b_oracle = md("$b^{\\text{oracle}}$"), b_cde = md("$b^{\\ddagger}$"),
      b_marg = md("$b^{\\dagger}$"),
      Cov_oracle = md("$\\hat{C}^{\\text{oracle}}$"),
      Cov_cde = md("$\\hat{C}^{\\ddagger}$"), Cov_marg = md("$C^{\\dagger}$")) |>
    tab_spanner(label = "Relative bias (%)", columns = c(b_oracle, b_cde, b_marg)) |>
    tab_spanner(label = "95% coverage", columns = c(Cov_oracle, Cov_cde, Cov_marg)) |>
    tab_footnote(footnote = md(sprintf(
      "Detection (harm-subgroup identified) rate: **%.0f%%** of %d sims. Summaries over detected replicates.",
      100 * det_rate, n_sims))) |>
    tab_footnote(footnote = md(paste(
      "Targets: $b^{\\text{oracle}}$ vs per-replicate oracle $\\hat\\theta(H)$;",
      "$b^{\\ddagger}$ vs CDE $\\theta^{\\ddagger}$;",
      "$b^{\\dagger}$ vs marginal subgroup HR $\\theta^{\\dagger}$.",
      "Tier-1 = `H2` (León Eq.7, both bias sources); Tier-2 = de-biased gate, IJ interval.")))
}
```

# Paired Tier-1 ↔ Tier-2 agreement panel {#sec-paired}

The cleanest test of the leading-order correspondence: both tiers correct the
*same* selected $\hat H$ on the *same* data, so per-replicate differences isolate
the `dfbeta` linearisation. Reported on the log-effect scale.

```{r paired-panel}
if (run_smoke && !is.null(results)) {
  paired <- results[results$detected == 1 &
                      is.finite(results$t1_H_est) & is.finite(results$t2_H_est), ,
                    drop = FALSE]

  if (nrow(paired) >= 2L) {
    d_point <- log(paired$t2_H_est) - log(paired$t1_H_est)
    d_lo    <- log(paired$t2_H_lo)  - log(paired$t1_H_lo)
    d_hi    <- log(paired$t2_H_hi)  - log(paired$t1_H_hi)

    agree <- data.frame(
      Quantity = c("log point estimate", "log CI lower", "log CI upper"),
      mean_d   = c(mean(d_point), mean(d_lo, na.rm = TRUE), mean(d_hi, na.rm = TRUE)),
      sd_d     = c(sd(d_point),   sd(d_lo, na.rm = TRUE),   sd(d_hi, na.rm = TRUE)),
      stringsAsFactors = FALSE)

    print(
      gt(agree) |>
        tab_header(title = "Tier-2 \u2212 Tier-1 (log scale)",
                   subtitle = sprintf("%d paired replicates \u00b7 point-estimate correlation = %.3f",
                                      nrow(paired), cor(log(paired$t1_H_est), log(paired$t2_H_est)))) |>
        cols_label(mean_d = "Mean \u0394", sd_d = "SD \u0394") |>
        fmt_number(columns = c(mean_d, sd_d), decimals = 4)
    )

    rng <- range(c(paired$t1_H_est, paired$t2_H_est), na.rm = TRUE)
    p <- ggplot(paired, aes(x = t1_H_est, y = t2_H_est)) +
      geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey50") +
      geom_point(size = 2, alpha = 0.8) +
      scale_x_log10(limits = rng) + scale_y_log10(limits = rng) +
      coord_equal() +
      labs(x = expression("Tier-1  "*hat(theta)^"*"*(hat(H))),
           y = expression("Tier-2  "*tilde(theta)*(hat(H))),
           title = "Per-replicate de-biased point estimates (H)") +
      theme_minimal(base_size = 12)
    print(p)
  } else {
    cat("Too few paired replicates at smoke-test counts to draw the panel.\n")
  }
}
```

# Notes for scale-up

* **Counts.** Move `n_sims` $\to$ 1000 $\to$ 20000 and `nb_boots` $\to$ 200–500
  in @sec-setup; nothing else changes. Tier-1 is the wall-clock bottleneck (the
  per-replicate timing line above quantifies the gap).
* **M2 / M3 and the regime sweep.** Add the smaller-$N$ / noise-factor cells
  (`n_add_noise`, $N \in \{500, 300\}$) and the true-HR $\times$ $n_S$ sweep once
  the M1 cell looks sane; the small-$n_S$ M3 cell is the linearisation's stress
  test.
* **FSlg.** Set `use_grf = TRUE` to match León's LASSO+GRF screening; the smoke
  test uses LASSO-only for speed.
* **Target semantics to confirm.** The three targets here are (per-replicate
  oracle, DGM `cde_H`/`cde_Hc`, DGM `hr_H_true`/`hr_Hc_true`). Before publishing
  numbers, confirm these match the $\theta^{\ddagger}$/$\theta^{\dagger}$
  definitions used by `build_estimation_table()` / León Table 5.

```{r session-info}
sessionInfo()
```
