---
title: "Operating Characteristics of the Tier-2 De-biased Gate (Standalone)"
subtitle: "Tier-2-only estimation/coverage at scale, on the León (2024) Table 2 design"
author: "Larry León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    html-math-method: mathjax
    theme: cosmo
    fig-cap-location: bottom
    code-fold: true
---

> **Status — Tier-2 standalone.** This is the **fast** configuration: it runs the
> de-biased gate only and omits the Tier-1 nonparametric bootstrap entirely. The
> Tier-1$\leftrightarrow$Tier-2 head-to-head (the correspondence check and the
> paired agreement panel) lives in the sibling document
> `tier2_debias_gate_simulation.qmd`; once that validation is in hand, this
> document characterizes the gate's operating characteristics at large `n_sims`
> without the bootstrap bottleneck. Every tunable lives in the setup chunk
> (@sec-setup); `n_sims` and `gate_draws` are the only counts. Results run live —
> no `saveRDS`/`readRDS` caching.

# What this evaluates {#sec-intro}

The **Tier-2 de-biased gate** (`fs_debias_gate()`, surfaced through
`forestsearch(debias_gate = TRUE)`) is a refit-free multiplier-bootstrap
approximation of León's post-selection bias correction, with an
infinitesimal-jackknife (IJ) confidence interval. To first order its point
estimate equals León's Eq. 7 correction and its interval equals León's IJ
variance, the only approximation being the per-subgroup `dfbeta` linearization.

Because the gate costs essentially the same per replicate regardless of design
size (one fit plus the multiplier draws, no re-search), it scales to large
`n_sims` cheaply. This document records, per replicate:

* **oracle** $\hat\theta(H)$ — refit on the *true* harm subgroup;
* **naive** $\hat\theta(\hat H)$ — plugin on the identified subgroup;
* **Tier-2** $\tilde\theta(\hat H)$ — the de-biased gate with its IJ interval.

Each is compared to three truth targets: the per-replicate **oracle** estimate,
the controlled direct effect $\theta^{\ddagger}$ (CDE), and the marginal subgroup
hazard ratio $\theta^{\dagger}$. The output is the León Table-2-layout
estimation/coverage table with a Tier-2 block (no Tier-1 row).

The **subgroup-identification engine is selectable** via `subgroup_method`
(`"consistency"`, `"dina"`, or `"grf"`, with `grf_selection` for the GRF mode):
the oracle, the truth targets, and the classification metric are all
method-agnostic, so the same harness characterizes any engine. The gate
de-biases each method against **its own** candidate family, and the re-selection
rule is derived automatically from `sg_focus` so it always matches the rule the
engine actually used: the consistency search ranks by consistency rate
(`eff` $\to$ `maxcons`), DINA/GRF rank by effect (`eff` $\to$ `maxeff`), and
`maxSG`/`minSG`/`effMaxSG`/`effMinSG` map through identically. No separate
re-selection knob is needed. **Detection** is judged on the identified subgroup itself
(`fs.est$sg.harm`), independent of whether the gate produced a CI; gate
availability is reported separately.

# Setup — all knobs {#sec-setup}

```{r setup, message=FALSE, warning=FALSE}
# NOTE: forestsearch must be INSTALLED (devtools::install()), not load_all():
# the doFuture multisession workers spawn separate R processes that only see the
# installed package.
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(ggplot2)
library(foreach)
library(doFuture)
library(future)

# ── Counts (the only knobs to change for production) ───────────────────────
n_sims       <- 1000L     # simulated trials        (Tier-2 scales cheaply: try 20000)
gate_draws   <- 2000L     # Tier-2 multiplier draws (keep ~2000)

run_smoke    <- TRUE      # master flag; FALSE short-circuits the run loop

# Scale-up scenarios for the timing projection (Tier-2 cost is ~linear in n_sims). Edit freely.
proj_scenarios <- c(1000L, 5000L, 20000L)

# ── Subgroup-identification method ─────────────────────────────────────────
# The estimation/coverage targets, oracle, and classification are all
# method-agnostic, so this document characterizes ANY selection engine:
#   "consistency" - the consistency search (resample identification);
#   "dina"        - DINA selection (de-biased over DINA's qualifying family);
#   "grf"         - GRF selection (DR-candidate frontier family; tree or frontier).
# The Tier-2 gate now runs for all three (each de-biased against its OWN
# candidate family, with re-selection matching the method's selection rule).
subgroup_method <- "dina"  # "consistency" | "dina" | "grf"
grf_selection   <- "tree"      # only if subgroup_method == "grf": "tree" | "frontier"
grf_depth       <- 2L              # GRF policy-tree depth / max frontier conjunctions
dmin.grf        <- 0.0             # GRF DR-score screening floor
dina_args       <- list()          # extra args forwarded to the DINA selector (dina_args =)

# ── DGM: León M1 cell (GBSG AFT super-population, alt hypothesis) ───────────
dgm_model       <- "alt"
target_hr_harm  <- 2.0    # calibrate k_inter to this Cox HR in the harm subgroup
n_sample        <- 700L   # M1 sample size
analysis_time   <- 84     # administrative censoring (months)
cens_adjust     <- log(1.5)
n_super         <- 5000L

# ── Search configuration (León tuning; resample identification) ────────────
consistency_method <- "resample"   # decided: resample identification (the new default)
sg_focus           <- "effMaxSG"        # -> "hr": highest consistency rate, ties by effect
# Inclusion-band rule for the effMaxSG/effMinSG foci (inert for eff/maxSG/minSG):
#   "neighborhood" = effect within (1-eps) of the max (multiplicative, natural scale);
#   "pareto"       = 2-D non-dominated set in (effect, size);
#   "both"         = neighborhood AND pareto.  The gate de-biases under the SAME rule.
selection_rule     <- "neighborhood"   # "neighborhood" | "pareto" | "both"
# Gate re-selection rule is NO LONGER specified here.  forestsearch() derives it
# from sg_focus, faithfully per engine: the consistency search ranks by
# consistency rate (eff/hr -> "maxcons"), DINA/GRF rank by effect (eff/hr ->
# "maxeff"), and maxSG/minSG/effMaxSG/effMinSG map through identically.  So
# sg_focus is the single source of truth; to override, pass `reselection =`
# inside debias_gate_args below (advanced use only).
hr_threshold       <- 0.9         # different target under DINA (not Cox based)
hr_consistency     <- 0.9
pconsistency       <- 0.90
fs_splits          <- 400L         # moot under "resample" (closed-form, not literal split)
maxk               <- 2L
n_min              <- 60L
d0_min             <- 12L
d1_min             <- 0L
use_lasso          <- TRUE
use_dina           <- TRUE
use_grf            <- TRUE         # FALSE = LASSO-only (fast, "FSl"); set TRUE for "FSlg"
use_twostage       <- TRUE

fs_conf_force      <- c("meno == 0")   # premenopausal (raw analog of z3 == 1)
fs_conf.cont_jcuts <- list(er = 10)     # J-quantile grid on raw ER (analog of z1; now meaningful, ER is continuous

# fs_conf_force <- NULL
# fs_conf.cont_jcuts <- NULL


# ── Column names emitted by simulate_from_dgm() ────────────────────────────
outcome_name   <- "y_sim"
event_name     <- "event_sim"
treat_name     <- "treat_sim"
id_name        <- "id"
harm_col       <- "flag_harm"
# Raw GBSG covariate candidates (as in a real analysis): forestsearch discovers
# its own cuts. Continuous -> er, age, pgr, nodes, size; categorical -> meno (0/1),
# grade (1/2/3). The pre-dichotomized z1..z5 are NOT used. The true harm subgroup
# (er <= 25th pct & meno == 0) must be re-discovered from these raw variables.
confounders_base <- c("er", "age", "meno", "pgr", "nodes", "size", "grade")

# ── Reproducibility / parallelism ──────────────────────────────────────────
seed_base  <- 8316951L
n_workers  <- ceiling(0.95*max(1L, min(parallel::detectCores(logical = FALSE) - 1L)))
z975       <- qnorm(0.975)

cat(sprintf("Tier-2 standalone: method=%s%s, n_sims=%d, gate_draws=%d, workers=%d\n",
            subgroup_method,
            if (identical(subgroup_method, "grf")) paste0("/", grf_selection) else "",
            n_sims, gate_draws, n_workers))
```

# Build the DGM and read off the truth targets

```{r build-dgm}
k_inter <- calibrate_k_inter(target_hr_harm = target_hr_harm,
                             model = dgm_model, use_ahr = FALSE)

dgm <- setup_gbsg_dgm(model = dgm_model, k_inter = k_inter,
                      n_super = n_super, seed = seed_base)
dgm <- compute_dgm_cde(dgm)   # attaches CDE (theta-ddagger) targets

`%||%` <- function(a, b) if (is.null(a) || length(a) == 0 || all(is.na(a))) b else a

# Fixed (population) truth targets, read defensively from the DGM object.
truth <- list(
  hr_causal   = dgm$hr_causal   %||% NA_real_,                 # overall causal HR
  marg_H      = dgm$hr_H_true   %||% NA_real_,                 # theta-dagger (H)
  marg_Hc     = dgm$hr_Hc_true  %||% NA_real_,                 # theta-dagger (Hc)
  cde_H       = dgm$cde_H       %||% (dgm$hazard_ratios$CDE_harm    %||% NA_real_),  # theta-ddagger (H)
  cde_Hc      = dgm$cde_Hc      %||% (dgm$hazard_ratios$CDE_no_harm %||% NA_real_)   # theta-ddagger (Hc)
)

cat("DGM truth targets (population):\n")
cat(sprintf("  overall causal HR      : %.3f\n", truth$hr_causal))
cat(sprintf("  theta-dagger  H / Hc   : %.3f / %.3f\n", truth$marg_H, truth$marg_Hc))
cat(sprintf("  theta-ddagger H / Hc   : %.3f / %.3f\n", truth$cde_H,  truth$cde_Hc))
```

# Per-replicate recorder

For each simulated trial: identify $\hat H$ and obtain the naive + Tier-2 gate in
one `forestsearch()` call, then refit the oracle on the true subgroup. Inner
`forestsearch()` parallelism is forced sequential because the outer loop already
parallelises across trials.

```{r recorder}
# Cox HR + delta-method 95% CI on a subset (for the per-replicate oracle).
# Guards against separation: non-finite coef/SE (the "coefficient may be
# infinite" case) returns a clean NA triple rather than Inf/NaN bounds.
.cox_hr_ci <- function(df) {
  if (sum(df[[event_name]]) < 5L || length(unique(df[[treat_name]])) < 2L)
    return(c(est = NA_real_, lo = NA_real_, hi = NA_real_))
  fml <- stats::as.formula(
    sprintf("survival::Surv(%s, %s) ~ %s", outcome_name, event_name, treat_name))
  fit <- tryCatch(survival::coxph(fml, data = df), error = function(e) NULL)
  if (is.null(fit)) return(c(est = NA_real_, lo = NA_real_, hi = NA_real_))
  b  <- unname(stats::coef(fit)[1])
  v  <- unname(suppressWarnings(diag(stats::vcov(fit))[1]))  # unname: else lo/hi inherit coef name -> oH["lo"] = NA
  se <- if (is.finite(v) && v > 0) sqrt(v) else NA_real_
  if (!is.finite(b) || !is.finite(se))
    return(c(est = NA_real_, lo = NA_real_, hi = NA_real_))
  c(est = exp(b), lo = exp(b - z975 * se), hi = exp(b + z975 * se))
}

# Classification of the identified subgroup against the true harm flag.
# sg_hat is grp.consistency$sg.harm.id: a per-subject 0/1 vector, length
# nrow(df), aligned to the input rows (clean on the consistency result; the
# GRF-object naming collision does not apply here).
.classify <- function(sg_hat, truth) {
  out <- c(sens = NA_real_, spec = NA_real_, ppv = NA_real_, npv = NA_real_)
  if (is.null(sg_hat) || length(sg_hat) != length(truth)) return(out)
  sg_hat <- as.integer(sg_hat); truth <- as.integer(truth)
  tp <- sum(sg_hat == 1 & truth == 1); fp <- sum(sg_hat == 1 & truth == 0)
  fn <- sum(sg_hat == 0 & truth == 1); tn <- sum(sg_hat == 0 & truth == 0)
  out["sens"] <- if (tp + fn > 0) tp / (tp + fn) else NA_real_
  out["spec"] <- if (tn + fp > 0) tn / (tn + fp) else NA_real_
  out["ppv"]  <- if (tp + fp > 0) tp / (tp + fp) else NA_real_
  out["npv"]  <- if (tn + fn > 0) tn / (tn + fn) else NA_real_
  out
}

# One all-NA record (returned on failure so the loop never aborts a cell).
.na_record <- function(sim_id) data.frame(
  sim_id = sim_id, detected = 0L, gate_ok = 0L, n_sel = NA_integer_,
  n_harm = NA_integer_,
  label = NA_character_, t2_secs = NA_real_,
  or_H_est = NA, or_H_lo = NA, or_H_hi = NA,
  or_Hc_est = NA, or_Hc_lo = NA, or_Hc_hi = NA,
  nv_H_est = NA, nv_H_lo = NA, nv_H_hi = NA,
  nv_Hc_est = NA, nv_Hc_lo = NA, nv_Hc_hi = NA,
  t2_H_est = NA, t2_H_lo = NA, t2_H_hi = NA, t2_H_se_ij = NA,
  t2_Hc_est = NA, t2_Hc_lo = NA, t2_Hc_hi = NA, t2_Hc_se_ij = NA,
  t2_harm_flag = NA_integer_, ij_source = NA_character_,
  sens = NA_real_, spec = NA_real_, ppv = NA_real_, npv = NA_real_,
  stringsAsFactors = FALSE)

record_replicate <- function(sim_id) {
  rec <- .na_record(sim_id)

  df <- tryCatch(
    simulate_from_dgm(dgm, n = n_sample, analysis_time = analysis_time,
                      cens_adjust = cens_adjust, seed = seed_base + sim_id),
    error = function(e) NULL)
  if (is.null(df)) return(rec)
  df[[id_name]] <- seq_len(nrow(df))
  confs <- intersect(confounders_base, names(df))

  # ---- Identification + naive + Tier-2 gate (single forestsearch call) -----
  # Args common to every method, plus a method-specific block.  Consistency-only
  # tuning (LASSO/two-stage/forced cuts/maxk) is NOT forwarded to DINA/GRF, which
  # have their own candidate enumeration; all methods search the same confs.
  base_args <- list(
    df.analysis     = df,
    outcome.name    = outcome_name, event.name = event_name,
    treat.name      = treat_name,   id.name    = id_name,
    flag_harm.name  = harm_col,     confounders.name = confs,
    is.RCT          = TRUE,         seedit = seed_base + sim_id,
    quiet           = TRUE,         # suppress the "De-biased gate: ..." cat()
    sg_focus        = sg_focus,
    subgroup_method = subgroup_method,
    hr.threshold    = hr_threshold, hr.consistency = hr_consistency,
    pconsistency.threshold = pconsistency, n.min = n_min,
    selection_rule  = selection_rule,   # governs *MaxSG/*MinSG band (search + gate)
    parallel_args   = list(plan = "sequential"),   # outer loop owns parallelism
    debias_gate     = TRUE,
    debias_gate_args = list(ci_method = "ij", draws = gate_draws,
                            include_complement = TRUE)  # reselection derived from sg_focus
  )
  method_args <- switch(
    subgroup_method,
    consistency = list(
      consistency_method = consistency_method,
      use_lasso = use_lasso, use_grf = use_grf, use_twostage = use_twostage,
      use_dina = use_dina,
      conf_force = fs_conf_force, conf.cont_jcuts = fs_conf.cont_jcuts,
      fs.splits = fs_splits, maxk = maxk, d0.min = d0_min, d1.min = d1_min),
    dina = dina_args,
    grf  = list(grf_selection = grf_selection, grf_depth = grf_depth,
                dmin.grf = dmin.grf),
    stop("unknown subgroup_method: ", subgroup_method)
  )

  t2_t0 <- proc.time()[3]
  fs.est <- tryCatch(do.call(forestsearch, c(base_args, method_args)),
                     error = function(e) NULL)
  rec$t2_secs <- proc.time()[3] - t2_t0
  if (is.null(fs.est)) return(rec)

  # Detection is METHOD-AGNOSTIC (a subgroup was identified), independent of
  # whether the Tier-2 gate produced a CI; gate availability is tracked
  # separately in gate_ok so a gate failure never masks a true detection.
  g     <- fs.est$debias_gate
  found <- !is.null(fs.est$sg.harm)
  rec$detected <- as.integer(found)
  if (!found) return(rec)
  rec$gate_ok <- as.integer(!is.null(g))

  if (!is.null(g)) {
    rec$label <- g$selected_label %||% NA_character_
    rec$n_sel <- g$n_selected      %||% NA_integer_
    rec$t2_harm_flag <- as.integer(isTRUE(g$harm_flag))
    rec$ij_source    <- g$debiased$ij_source %||% NA_character_
    rec$nv_H_est <- g$naive$est;    rec$nv_H_lo <- g$naive$lower;    rec$nv_H_hi <- g$naive$upper
    rec$t2_H_est <- g$debiased$est; rec$t2_H_lo <- g$debiased$lower; rec$t2_H_hi <- g$debiased$upper
    rec$t2_H_se_ij <- g$debiased$se_ij %||% NA_real_
    if (is.list(g$complement) && !is.null(g$complement$debiased)) {
      rec$nv_Hc_est <- g$complement$naive$est
      rec$nv_Hc_lo  <- g$complement$naive$lower
      rec$nv_Hc_hi  <- g$complement$naive$upper
      rec$t2_Hc_est <- g$complement$debiased$est
      rec$t2_Hc_lo  <- g$complement$debiased$lower
      rec$t2_Hc_hi  <- g$complement$debiased$upper
      rec$t2_Hc_se_ij <- g$complement$debiased$se_ij %||% NA_real_
    }
  }

  # ---- Classification of Ĥ against the true harm flag ----------------------
  # grp.consistency$sg.harm.id is a per-subject 0/1 vector (length nrow(df),
  # aligned to input rows) populated by ALL selectors -- consistency, DINA, and
  # GRF -- so the classification metric is identical across methods.
  cl <- .classify(fs.est$grp.consistency$sg.harm.id, df[[harm_col]])
  rec$sens <- cl["sens"]; rec$spec <- cl["spec"]
  rec$ppv  <- cl["ppv"];  rec$npv <- cl["npv"]
  # Identified-subgroup size |Ĥ| from the per-subject flag (all detected reps,
  # independent of whether the gate produced a CI) -- feeds avg(Ĥ).
  rec$n_harm <- sum(fs.est$grp.consistency$sg.harm.id == 1L, na.rm = TRUE)

  # ---- Oracle: refit on the TRUE harm subgroup of this trial ---------------
  oH  <- .cox_hr_ci(df[df[[harm_col]] == 1, , drop = FALSE])
  oHc <- .cox_hr_ci(df[df[[harm_col]] == 0, , drop = FALSE])
  rec$or_H_est <- oH["est"];   rec$or_H_lo <- oH["lo"];   rec$or_H_hi <- oH["hi"]
  rec$or_Hc_est <- oHc["est"]; rec$or_Hc_lo <- oHc["lo"]; rec$or_Hc_hi <- oHc["hi"]

  rec
}
```

# Run the simulation

```{r run-loop, message=FALSE, warning=FALSE}
results <- NULL
if (run_smoke) {
  plan("sequential"); gc()
  plan("multisession", workers = n_workers)

  t_all <- proc.time()[3]
  results <- foreach(
    s = seq_len(n_sims), .combine = rbind, .errorhandling = "remove",
    .options.future = list(packages = c("forestsearch", "survival", "data.table"),
                           seed = TRUE)
  ) %dofuture% {
    record_replicate(s)
  }
  elapsed <- proc.time()[3] - t_all

  plan("sequential"); gc()   # release workers between phases

  cat(sprintf("Completed %d/%d replicates in %.1f s (%.1f s/replicate).\n",
              sum(results$detected, na.rm = TRUE), n_sims, elapsed,
              elapsed / max(1L, n_sims)))
  cat(sprintf("Mean Tier-2 gate wall-clock: %.3f s/replicate.\n",
              mean(results$t2_secs, na.rm = TRUE)))
}
```

# Timing and scale-up projection

Tier-2 cost per replicate is essentially fixed — one fit plus the multiplier
draws, with **no** per-replicate bootstrap re-search — so total cost scales
~linearly with `n_sims`. The projection uses this run's measured per-replicate
gate time and divides by the worker count for an idealized wall-clock.

```{r timing-projection}
if (run_smoke && !is.null(results)) {
  t2_mean  <- mean(results$t2_secs, na.rm = TRUE)   # forestsearch + gate, per rep (fixed)
  det_rate <- mean(results$detected, na.rm = TRUE)

  fmt_hm <- function(sec) {
    h <- floor(sec / 3600); m <- round((sec - 3600 * h) / 60)
    if (h >= 1) sprintf("%d h %02d m", h, m) else sprintf("%d m", round(sec / 60))
  }
  proj_row <- function(ns) {
    cpu_sec <- ns * t2_mean
    wall_id <- cpu_sec / n_workers                   # idealized (perfect parallelism)
    data.frame(n_sims = ns, t2_rep_s = round(t2_mean, 1),
               cpu_hours = round(cpu_sec / 3600, 2),
               wall_ideal = fmt_hm(wall_id),
               wall_real_1.3x = fmt_hm(1.3 * wall_id),
               stringsAsFactors = FALSE)
  }
  tab <- rbind(cbind(Scenario = "current run", proj_row(n_sims)),
               cbind(Scenario = "projection",
                     do.call(rbind, lapply(proj_scenarios, proj_row))))

  cat(sprintf(
    "Current run: %d reps in %.0f s on %d workers. Tier-2 gate %.1f s/rep. Detection %.0f%%.\n",
    n_sims, elapsed, n_workers, t2_mean, 100 * det_rate))

  gt(tab) |>
    tab_header(title = "Wall-clock projection (Tier-2 standalone)",
               subtitle = sprintf("Measured on this run \u00b7 %d workers", n_workers)) |>
    cols_label(n_sims = "sims", t2_rep_s = "Tier-2/rep (s)", cpu_hours = "CPU-hours",
               wall_ideal = "wall (ideal)", wall_real_1.3x = "wall (~1.3\u00d7)") |>
    tab_footnote(footnote = md(paste(
      "Tier-2/rep is fixed (no bootstrap). CPU-hours = sims \u00d7 Tier-2/rep.",
      "`wall (ideal)` divides by workers assuming perfect parallelism;",
      "`~1.3\u00d7` adds a rule-of-thumb overhead (pool spin-up, memory contention at high worker counts).")))
}
```

# Estimation / coverage table (León Table 2 layout + Tier-2)

```{r estimation-table}
# All summaries are finite-only: non-finite estimates/bounds (e.g. from Cox
# separation or an unstable small-B IJ) are dropped so a single blow-up cannot
# dominate a mean.
.cover <- function(target, lo, hi) {
  target <- rep_len(target, length(lo))   # recycle scalar targets (CDE/marginal) to n reps
  ok <- is.finite(target) & is.finite(lo) & is.finite(hi)
  if (!any(ok)) return(NA_real_)
  mean(target[ok] >= lo[ok] & target[ok] <= hi[ok])
}
.pbias <- function(est, target) {
  d <- (est - target) / target; d <- d[is.finite(d)]
  if (!length(d)) return(NA_real_)
  100 * mean(d)
}
# HR-scale SD implied by log-scale CI bounds (uniform across estimators).
.sd_hat <- function(est, lo, hi) {
  se_log <- (log(hi) - log(lo)) / (2 * z975)
  v <- est * se_log; v <- v[is.finite(v)]
  if (!length(v)) return(NA_real_)
  mean(v)
}
.mfin <- function(x) { x <- x[is.finite(x)]; if (length(x)) mean(x) else NA_real_ }
.sdfin <- function(x) { x <- x[is.finite(x)]; if (length(x) > 1L) sd(x) else NA_real_ }

# Build one block (H or Hc): three estimator rows against three targets.
build_block <- function(r, suffix, tgt_cde, tgt_marg) {
  est_keys <- c(oracle = "or", naive = "nv", `Tier-2` = "t2")
  or_est <- r[[paste0("or_", suffix, "_est")]]   # per-replicate oracle target
  rows <- lapply(names(est_keys), function(lab) {
    k  <- est_keys[[lab]]
    e  <- r[[paste0(k, "_", suffix, "_est")]]
    lo <- r[[paste0(k, "_", suffix, "_lo")]]
    hi <- r[[paste0(k, "_", suffix, "_hi")]]
    len <- hi - lo; len <- len[is.finite(len)]
    data.frame(
      Estimator = lab,
      n       = sum(is.finite(e)),
      Avg     = .mfin(e),
      SD      = .sdfin(e),
      SD_hat  = .sd_hat(e, lo, hi),
      Len_med = if (length(len)) stats::median(len) else NA_real_,
      Length  = if (length(len)) mean(len) else NA_real_,
      b_oracle = .pbias(e, or_est),
      b_cde    = .pbias(e, tgt_cde),
      b_marg   = .pbias(e, tgt_marg),
      Cov_oracle = .cover(or_est, lo, hi),
      Cov_cde    = .cover(tgt_cde,  lo, hi),
      Cov_marg   = .cover(tgt_marg, lo, hi),
      stringsAsFactors = FALSE)
  })
  do.call(rbind, rows)
}

if (run_smoke && !is.null(results)) {
  rH  <- results[results$detected == 1, , drop = FALSE]
  blockH  <- build_block(rH, "H",  truth$cde_H,  truth$marg_H)
  blockHc <- build_block(rH, "Hc", truth$cde_Hc, truth$marg_Hc)
  est_tab <- rbind(
    cbind(Block = "Harm  \u0124", blockH),
    cbind(Block = "Complement  \u0124\u1d9c", blockHc)
  )

  det_rate <- mean(results$detected, na.rm = TRUE)
  # Classification + decision + IJ-stability diagnostics (over detected reps).
  cls <- sprintf("sens %.2f \u00b7 spec %.2f \u00b7 PPV %.2f \u00b7 NPV %.2f",
                 .mfin(rH$sens), .mfin(rH$spec), .mfin(rH$ppv), .mfin(rH$npv))
  harm_rate <- mean(rH$t2_harm_flag, na.rm = TRUE)
  ij_ok     <- mean(rH$ij_source == "ij", na.rm = TRUE)
  gate_rate <- mean(rH$gate_ok, na.rm = TRUE)      # gate produced a CI | detected
  avg_size  <- .mfin(rH$n_harm)                    # average |Ĥ| over detected reps
  # Estimable-realization note: the n column is near-constant, so summarize it
  # as the common per-cell count plus any exceptions, and drop the column.
  n_common <- as.integer(names(sort(table(est_tab$n), decreasing = TRUE))[1])
  exc <- est_tab[est_tab$n != n_common, , drop = FALSE]
  n_note <- if (!nrow(exc))
    sprintf("Estimable realizations: %d per cell.", n_common)
  else sprintf("Estimable realizations: %d per cell (except %s).", n_common,
               paste(sprintf("%s in %s: %d", exc$Estimator,
                             sub("^.*\\s", "", exc$Block), exc$n),
                     collapse = "; "))
  est_tab$n <- NULL
  method_lab <- sprintf("%s%s identification", subgroup_method,
                        if (identical(subgroup_method, "grf"))
                          paste0("/", grf_selection) else "")

  gt(est_tab, groupname_col = "Block") |>
    tab_header(
      title = "Tier-2 de-biased gate: estimation & coverage",
      subtitle = sprintf("M1 \u00b7 n=%d, %d sims, draws=%d \u00b7 %s",
                         n_sample, n_sims, gate_draws, method_lab)) |>
    fmt_number(columns = c(Avg, SD, SD_hat, Len_med, Length), decimals = 2) |>
    fmt_number(columns = c(b_oracle, b_cde, b_marg), decimals = 1) |>
    fmt_number(columns = c(Cov_oracle, Cov_cde, Cov_marg), decimals = 2) |>
    cols_label(
      Len_med = "Len(med)",
      SD_hat = md("$\\widehat{\\text{SD}}$"),
      b_oracle = md("$b^{\\text{oracle}}$"), b_cde = md("$b^{\\ddagger}$"),
      b_marg = md("$b^{\\dagger}$"),
      Cov_oracle = md("$\\hat{C}^{\\text{oracle}}$"),
      Cov_cde = md("$\\hat{C}^{\\ddagger}$"), Cov_marg = md("$C^{\\dagger}$")) |>
    tab_spanner(label = "Relative bias (%)", columns = c(b_oracle, b_cde, b_marg)) |>
    tab_spanner(label = "95% coverage", columns = c(Cov_oracle, Cov_cde, Cov_marg)) |>
    tab_footnote(footnote = md(sprintf(
      "Detection rate **%.0f%%** of %d sims (the search identified a subgroup \u0124). Among those detected reps: the gate produced a CI in **%.0f%%**, and the Tier-2 de-biased estimate still cleared the harm gate in **%.0f%%** (i.e. apparent harm survived bias correction); IJ variance held (not fallback) in **%.0f%%**. Classification of \u0124 vs true subgroup: %s; avg(\u0124) size **%.0f** (%.0f%% of n=%d). %s",
      100 * det_rate, n_sims, 100 * gate_rate, 100 * harm_rate, 100 * ij_ok, cls,
      avg_size, 100 * avg_size / n_sample, n_sample, n_note))) |>
    tab_footnote(footnote = md(paste(
      "Targets: $b^{\\text{oracle}}$ vs per-replicate oracle $\\hat\\theta(H)$;",
      "$b^{\\ddagger}$ vs CDE $\\theta^{\\ddagger}$;",
      "$b^{\\dagger}$ vs marginal subgroup HR $\\theta^{\\dagger}$.",
      "`Len(med)` = median CI length (robust to blow-ups); `Length` = mean.",
      "Tier-2 = de-biased gate, IJ interval.")))
}
```

# Notes for scale-up

* **Counts.** Move `n_sims` $\to$ 5000 $\to$ 20000 in @sec-setup; nothing else
  changes. With no bootstrap, cost is ~linear in `n_sims` and modest (see the
  projection above), so large runs are cheap.
* **Validation.** This document drops the Tier-1 head-to-head; run the sibling
  `tier2_debias_gate_simulation.qmd` at small `n_sims` first to confirm the
  Tier-1$\leftrightarrow$Tier-2 correspondence before trusting Tier-2 alone here.
* **M2 / M3 and the regime sweep.** Add the smaller-$N$ / noise-factor cells
  (`n_add_noise`, $N \in \{500, 300\}$) and the true-HR $\times$ $n_S$ sweep once
  the M1 cell looks sane; the small-$n_S$ M3 cell is the linearisation's stress
  test (where Tier-2 is loosest, so worth the large-`n_sims` precision this
  standalone run affords).
* **Target semantics to confirm.** The three targets here are (per-replicate
  oracle, DGM `cde_H`/`cde_Hc`, DGM `hr_H_true`/`hr_Hc_true`). Before publishing
  numbers, confirm these match the $\theta^{\ddagger}$/$\theta^{\dagger}$
  definitions used by `build_estimation_table()` / León Table 5.

```{r session-info}
sessionInfo()
```
