---
title: "Cross-cut Summary: sg_focus × FS Bundle (ACTG175 Binary, m1)"
format: html
---

This article compares operating characteristics across the
cross-product of **`sg_focus`** and **FS parameter bundle** for a
fixed study × outcome × DGM × objective combination:

- study : ACTG175
- outcome : binary
- DGM : m1 (calibrated GLM, target OR(Q) = 2.0)
- objective : harm
- varied : `sg_focus` (e.g., `hr`, `hrMaxSG`, `maxSG`) × FS bundle (`fs*`)

Each cell of the grid corresponds to one bundle file produced by
its companion runner script. The full intended grid is declared
below; missing bundles are skipped with a message (set
`strict_bundles <- TRUE` in setup to error instead).

```{r setup}
#| code-fold: true
#| code-summary: "Setup, configs, and library loads (click to expand)"

library(forestsearch)
library(data.table)
library(gt)

# Base R adds %||% in 4.4; define a fallback for older R.
if (!exists("%||%", mode = "function")) {
  `%||%` <- function(x, y) if (is.null(x)) y else x
}

# -- Levels to compare ------------------------------------------------------
# Edit these vectors as bundles become available.  expand.grid() builds
# the full cross-product; missing bundles are skipped (or error if
# strict_bundles = TRUE).
focus_levels   <- c("hr", "hrMaxSG", "maxSG")
fs_levels      <- c("fs1", "fs2", "fs3")
strict_bundles <- FALSE

# -- Fixed axes (must match every bundle's config) --------------------------
fixed_study     <- "actg175"
fixed_outcome   <- "binary"
fixed_dgm       <- "m1"
fixed_objective <- "harm"

# -- Build configs ----------------------------------------------------------
# Path is relative to this qmd's directory
# (i.e., quarto/simulations/actg175/_summaries/).
configs <- expand.grid(
  focus = focus_levels,
  fs    = fs_levels,
  KEEP.OUT.ATTRS   = FALSE,
  stringsAsFactors = FALSE
)
# Sort by focus first, then fs (matches the naming convention).
configs <- configs[order(
  match(configs$focus, focus_levels),
  match(configs$fs,    fs_levels)
), , drop = FALSE]
rownames(configs) <- NULL
configs$config_id <- sprintf(
  "%s_%s_%s_%s_%s_%s",
  fixed_study, fixed_outcome, fixed_dgm, fixed_objective,
  configs$focus, configs$fs
)
configs$label       <- sprintf("%s × %s", configs$focus, configs$fs)
configs$bundle_path <- sprintf("../%s.rds", configs$config_id)

# -- Filter to existing bundles --------------------------------------------
exists_flag <- vapply(configs$bundle_path, file.exists, logical(1))
if (!all(exists_flag)) {
  miss_msg <- sprintf("Missing %d of %d bundles:\n  %s",
                      sum(!exists_flag), nrow(configs),
                      paste(configs$config_id[!exists_flag],
                            collapse = "\n  "))
  if (strict_bundles) {
    stop(miss_msg, call. = FALSE)
  } else {
    message(miss_msg)
    configs <- configs[exists_flag, , drop = FALSE]
  }
}

if (nrow(configs) == 0L) {
  stop("No bundles available to compare. Build at least one bundle first.",
       call. = FALSE)
}

# -- Display options --------------------------------------------------------
sg_notation <- "harm"

cat(sprintf("Comparing %d configuration(s):\n", nrow(configs)))
print(configs[, c("label", "config_id")])
```

# Load bundles {#sec-load}

```{r load-bundles}
#| code-fold: true
#| code-summary: "Bundle loading + validation (click to expand)"

bundles <- setNames(lapply(configs$bundle_path, readRDS), configs$label)

# Soft-validate config_id matches what's declared in `configs`
for (i in seq_len(nrow(configs))) {
  meta_id <- bundles[[i]]$meta$config_id
  if (!is.null(meta_id) && meta_id != configs$config_id[i]) {
    warning(sprintf(
      "Bundle '%s' has config_id '%s' but configs declares '%s'.",
      configs$label[i], meta_id, configs$config_id[i]
    ), call. = FALSE)
  }
}
```

## Bundle metadata

```{r bundle-metadata}
#| code-fold: true
#| code-summary: "Per-bundle metadata table (click to expand)"

meta_dt <- data.table::rbindlist(lapply(seq_along(bundles), function(i) {
  m <- bundles[[i]]$meta
  data.frame(
    label           = configs$label[i],
    focus           = configs$focus[i],
    fs              = configs$fs[i],
    timestamp       = format(m$timestamp     %||% NA),
    n_alt           = m$n_alt                %||% NA_integer_,
    n_null          = m$n_null               %||% NA_integer_,
    sim_n           = m$sim_n                %||% NA_integer_,
    n_failed_alt    = m$n_failed_alt         %||% NA_integer_,
    n_failed_null   = m$n_failed_null        %||% NA_integer_,
    t_alt_min       = round((m$t_alt_sec  %||% NA) / 60, 1),
    t_null_min      = round((m$t_null_sec %||% NA) / 60, 1),
    package_version = m$package_version      %||% NA_character_,
    stringsAsFactors = FALSE
  )
}))

gt::gt(meta_dt) |>
  gt::tab_header(title = "Bundle metadata") |>
  gt::sub_missing(missing_text = "—")
```

# Stacked Operating Characteristics (per cell)

For each cell of the focus × fs grid, the bundle's H1 results are
rendered using the same `format_oc_results()` and
`build_classification_table()` calls used in the per-config article
(its Section 6.1 and Section 6.3 equivalents). This preserves the
familiar tutorial-style layout per cell, stacked vertically for
visual scan-comparison across configurations.

```{r stacked-oc}
#| results: asis
#| code-fold: true
#| code-summary: "Per-cell rendering loop (click to expand)"

for (i in seq_len(nrow(configs))) {
  b <- bundles[[i]]

  # Header: focus × fs (in TOC, follows naming-convention order)
  cat(sprintf("\n## %s × %s\n\n", configs$focus[i], configs$fs[i]))

  if (is.null(b$results$alt) || nrow(b$results$alt) == 0L) {
    cat("_No H1 results in this bundle._\n\n")
    next
  }

  # ── Section 6.1 equivalent: format_oc_results() ────────────────────────
  oc_tbl <- format_oc_results(
    b$results$alt,
    metrics           = c("detection", "classification", "subgroup_size"),
    digits            = 2,
    digits_hr         = 2,
    title             = sprintf("%s × %s — H1 OC",
                                configs$focus[i], configs$fs[i]),
    dgm               = b$dgm$calibrated,
    subgroup_notation = sg_notation
  )
  cat(gt::as_raw_html(oc_tbl), "\n\n")

  # ── Section 6.3 equivalent: build_classification_table() ───────────────
  scenario_alt <- list(
    H1 = list(
      results    = b$results$alt,
      dgm        = b$dgm$calibrated,
      hypothesis = "alt",
      n_sample   = b$meta$sim_n,
      label      = sprintf("%s × %s", configs$focus[i], configs$fs[i])
    )
  )

  for (analysis_method in c("FS", "GRF")) {
    cls_tbl <- build_classification_table(
      scenario_alt,
      analyses          = analysis_method,
      subgroup_notation = sg_notation,
      title             = sprintf("%s — Classification Under H1",
                                  analysis_method)
    )
    cat(gt::as_raw_html(cls_tbl), "\n\n")
  }
}
```

# Per-config Detail

For deep-dives into any single configuration, see the corresponding
per-config article.

```{r per-config-pointers}
#| echo: false
pointers <- data.frame(
  label   = configs$label,
  article = sprintf("../%s.qmd", configs$config_id),
  bundle  = configs$bundle_path,
  stringsAsFactors = FALSE
)
gt::gt(pointers) |>
  gt::tab_header(title = "Per-config articles and bundles") |>
  gt::cols_label(
    label   = "Config",
    article = "Per-config article",
    bundle  = "Bundle (.rds)"
  )
```

# Notes

- This article assumes all bundles share the **same DGM truth**
  (study × outcome × DGM × objective). Cross-cuts that vary the
  DGM need a different summary template (with per-bundle reference
  values for target OR, true |Q|, etc.).
- The per-cell tables here use the same code path as the per-config
  article's Section 6.1 and Section 6.3 — so each cell is
  byte-equivalent to that configuration's standalone view.
- Toggle `strict_bundles <- TRUE` in setup to fail when any cell
  in the declared grid is missing — useful as a completeness check
  once all configurations are expected to be built.
