---
title: "Simulation Studies: Poisson Rate Model (GBSG)"
subtitle: "Subgroup Identification with Incidence Rate Ratios"
author: "Larry F. León"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 9
    fig-height: 6
execute:
  warning: true
  message: true
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

# ─── Simulation control ─────────────────────────────────────────────────
nsims_alt   <- 100L
nsims_null  <- 200L
sim_seed    <- 8316951L
n_workers   <- min(parallel::detectCores(logical = FALSE) - 1L, 12L)

# ─── Dataset: GBSG ──────────────────────────────────────────────────────
# Treat the binary recurrence indicator (status) as a Poisson count
# (0 or 1) with person-time offset (time_months).  Under exponential
# survival, the Poisson IRR approximates the Cox HR.

# ─── Subgroup definition ────────────────────────────────────────────────
# ER-negative patients: consistent with the Cox analysis finding.
# Dichotomised covariates follow the GBSG vignette conventions.
dgm_factor_vars   <- c("z_age", "z_meno", "z_size", "z_grade3",
                        "z_nodes", "z_pgr", "z_er")
dgm_subgroup_vars <- c("z_er")
dgm_subgroup_cuts <- list(z_er = 1L)
dgm_n_super       <- 5000L

# ─── Calibration ────────────────────────────────────────────────────────
# Target IRR(H) = 2.0 in the harm subgroup (treatment increases event rate)
cal_target_irr    <- 2.0
cal_k_grid_range  <- c(0, 3)
cal_grid_step     <- 0.05

# ─── Simulation study ──────────────────────────────────────────────────
sim_n_sample      <- 686L    # match original GBSG sample size
sim_n_noise       <- 0L

# ─── Analysis flags ────────────────────────────────────────────────────
run_fs            <- TRUE
run_fs_grf        <- FALSE
run_grf           <- FALSE

# ─── ForestSearch parameters ───────────────────────────────────────────
fs_irr_threshold       <- 1.25
fs_irr_consistency     <- 1.0
fs_pconsistency        <- 0.90
fs_splits              <- 400L
fs_n_min               <- 60L
fs_d0_min              <- 12L
fs_d1_min              <- 12L
fs_maxk                <- 2L
fs_vi_grf_min          <- -0.2
fs_seedit              <- 8316951L

# ─── ForestSearch pipeline flags ───────────────────────────────────────
fs_use_lasso           <- TRUE    # Poisson-LASSO available
fs_use_grf             <- TRUE
fs_is_rct              <- TRUE

# ─── Output persistence ───────────────────────────────────────────────
save_results           <- TRUE
output_dir             <- "_data"
```


# Introduction {#sec-intro}

## Motivation

This document validates the ForestSearch GLM extension for **count
outcomes** (Poisson regression with offset) using the German Breast
Cancer Study Group (GBSG) dataset.  By treating each patient's binary
recurrence indicator (`status`) as a count (0 or 1) with
`log(time_months)` as the exposure offset, we obtain incidence rate
ratios (IRR) that approximate the Cox hazard ratios (HR) under
exponential survival.

This design serves two purposes:

1. **Cross-validation with the Cox pipeline**: If the Poisson rate
   model identifies the same subgroup with similar effect sizes, this
   strengthens confidence in both approaches.

2. **GLM count simulation infrastructure**: Exercises `generate_glm_dgm()`
   with `outcome_type = "count"` and `offset_var`, validating the
   entire pipeline (DGM → calibration → simulation → ForestSearch →
   operating characteristics).


# Data Preparation {#sec-data}

```{r load-packages}
library(forestsearch)
library(survival)
library(data.table)
library(gt)
library(ggplot2)
library(doFuture)
library(foreach)

# Safe foreach result collection
collect_results <- function(raw_list, label = "") {
  is_err <- vapply(raw_list, inherits, logical(1), "error")
  n_failed <- sum(is_err)
  if (n_failed > 0) {
    err_msgs <- vapply(raw_list[is_err], conditionMessage, character(1))
    warning(sprintf("%s%d of %d tasks failed.",
        if (nzchar(label)) paste0(label, ": ") else "",
        n_failed, length(raw_list)))
    message("  First errors:\n    ",
        paste(head(unique(err_msgs), 3), collapse = "\n    "))
    if (n_failed == length(raw_list)) {
      stop(sprintf("%sAll %d tasks failed. First error: %s",
        if (nzchar(label)) paste0(label, ": ") else "",
        n_failed, err_msgs[1]), call. = FALSE)
    }
  }
  good <- raw_list[!is_err]
  if (length(good) == 0) {
    result <- data.frame()
  } else {
    result <- data.table::rbindlist(good, use.names = TRUE, fill = TRUE)
    result <- as.data.frame(result)
  }
  attr(result, "n_failed") <- n_failed
  attr(result, "n_total") <- length(raw_list)
  result
}

timings <- list()
```

```{r prepare-data}
t0 <- proc.time()[3]

df.analysis <- gbsg

df.analysis <- within(df.analysis, {
  id          <- seq_len(nrow(df.analysis))
  time_months <- rfstime / 30.4375
  grade3      <- ifelse(grade == "3", 1, 0)
  treat       <- hormon
})

# Dichotomise covariates for ForestSearch
df.analysis$z_age    <- as.factor(ifelse(df.analysis$age > median(df.analysis$age), 1L, 0L))
df.analysis$z_meno   <- as.factor(df.analysis$meno)
df.analysis$z_size   <- as.factor(ifelse(df.analysis$size > median(df.analysis$size), 1L, 0L))
df.analysis$z_grade3 <- as.factor(df.analysis$grade3)
df.analysis$z_nodes  <- as.factor(ifelse(df.analysis$nodes > median(df.analysis$nodes), 1L, 0L))
df.analysis$z_pgr    <- as.factor(ifelse(df.analysis$pgr > median(df.analysis$pgr), 1L, 0L))
df.analysis$z_er     <- as.factor(ifelse(df.analysis$er <= 0, 1L, 0L))

cat(sprintf("GBSG: N = %d, events = %d (%.1f%%)\n",
    nrow(df.analysis), sum(df.analysis$status),
    100 * mean(df.analysis$status)))
cat(sprintf("Median follow-up: %.1f months\n",
    median(df.analysis$time_months)))

# Verify Poisson IRR ≈ Cox HR at ITT level
cox_itt <- coxph(Surv(time_months, status) ~ hormon, data = df.analysis)
pois_itt <- glm(status ~ hormon, family = poisson(),
                offset = log(pmax(time_months, 1e-6)),
                data = df.analysis)
cat(sprintf("\nITT Cox HR:      %.4f\n", exp(coef(cox_itt))))
cat(sprintf("ITT Poisson IRR: %.4f\n", exp(coef(pois_itt)[["hormon"]])))
cat(sprintf("|HR - IRR|:      %.4f\n",
    abs(exp(coef(cox_itt)) - exp(coef(pois_itt)[["hormon"]]))))

# Subgroup prevalence
cat(sprintf("\nER-negative (z_er == 1): %d / %d (%.1f%%)\n",
    sum(df.analysis$z_er == 1), nrow(df.analysis),
    100 * mean(df.analysis$z_er == 1)))

timings$data_prep <- proc.time()[3] - t0
```


# Creating the Poisson DGM {#sec-dgm}

The GLM DGM fits a Poisson regression with offset to the observed
GBSG data, creates a super-population with individual-level potential
outcome means ($\mu_0$, $\mu_1$), and computes true subgroup effects
on the IRR scale.

## Step 1: Initial DGM

```{r create-dgm-alt}
t0 <- proc.time()[3]

dgm_alt_init <- generate_glm_dgm(
  data          = df.analysis,
  factor_vars   = dgm_factor_vars,
  outcome_var   = "status",
  treatment_var = "hormon",
  outcome_type  = "count",
  effect_measure = "IRR",
  offset_var    = "time_months",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  model         = "alt",
  k_inter       = 0,
  n_super       = dgm_n_super,
  seed          = sim_seed,
  verbose       = TRUE
)

cat("\n=== Initial GBSG Poisson DGM ===\n")
cat("  IRR(H):  ", round(dgm_alt_init$hazard_ratios$harm_subgroup, 3), "\n")
cat("  IRR(Hc): ", round(dgm_alt_init$hazard_ratios$no_harm_subgroup, 3), "\n")
cat("  IRR(ITT):", round(dgm_alt_init$hazard_ratios$overall, 3), "\n")

timings$dgm_init <- proc.time()[3] - t0
```


## Step 2: Calibrate for Target IRR {#sec-calibrate}

The `k_inter` parameter shifts the log-rate for subgroup H members
under treatment.  We search over a grid to find the value producing
the target IRR(H).

```{r calibrate}
t0 <- proc.time()[3]

dgm_calibrated <- calibrate_glm_interaction(
  data          = df.analysis,
  factor_vars   = dgm_factor_vars,
  outcome_var   = "status",
  treatment_var = "hormon",
  target_effect = cal_target_irr,
  outcome_type  = "count",
  effect_measure = "IRR",
  offset_var    = "time_months",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  k_inter_range = cal_k_grid_range,
  grid_step     = cal_grid_step,
  n_super       = dgm_n_super,
  seed          = sim_seed,
  verbose       = TRUE
)

cat("\n=== Calibrated GBSG Poisson DGM ===\n")
cat("  IRR(H):  ", round(dgm_calibrated$hazard_ratios$harm_subgroup, 3),
    " (target:", cal_target_irr, ")\n")
cat("  IRR(Hc): ", round(dgm_calibrated$hazard_ratios$no_harm_subgroup, 3), "\n")
cat("  IRR(ITT):", round(dgm_calibrated$hazard_ratios$overall, 3), "\n")
cat("  k_inter: ", dgm_calibrated$model_params$k_inter, "\n")

timings$calibrate <- proc.time()[3] - t0
```

### Calibration Curve

```{r calibration-plot, fig.height=5}
#| fig-cap: "Calibration of k_inter: IRR(H) as a function of the interaction parameter."

cal_grid <- seq(cal_k_grid_range[1], cal_k_grid_range[2], by = cal_grid_step)
cal_irr_H <- vapply(cal_grid, function(k) {
  d <- generate_glm_dgm(
    data = df.analysis, factor_vars = dgm_factor_vars,
    outcome_var = "status", treatment_var = "hormon",
    outcome_type = "count", effect_measure = "IRR",
    offset_var = "time_months",
    subgroup_vars = dgm_subgroup_vars,
    subgroup_cuts = dgm_subgroup_cuts, model = "alt",
    k_inter = k, n_super = dgm_n_super, seed = sim_seed
  )
  d$hazard_ratios$harm_subgroup
}, numeric(1))

cal_df <- data.frame(k_inter = cal_grid, irr_H = cal_irr_H)
best_k <- dgm_calibrated$model_params$k_inter

ggplot(cal_df, aes(x = k_inter, y = irr_H)) +
  geom_line(linewidth = 1, colour = "steelblue") +
  geom_hline(yintercept = cal_target_irr, linetype = "dashed",
             colour = "grey50") +
  geom_hline(yintercept = 1.0, linetype = "dotted",
             colour = "grey70") +
  geom_point(
    data = data.frame(k_inter = best_k,
                      irr_H = dgm_calibrated$hazard_ratios$harm_subgroup),
    colour = "red", size = 4, shape = 16
  ) +
  labs(
    x = "k_inter (log-rate shift for H under treatment)",
    y = "IRR(H)",
    title = "GBSG Poisson DGM Calibration",
    subtitle = sprintf("Best: k_inter = %.2f -> IRR(H) = %.3f",
      best_k, dgm_calibrated$hazard_ratios$harm_subgroup)
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```


## Step 3: Null DGM

Under `model = "null"`, the interaction is zeroed out — no
treatment effect heterogeneity.

```{r dgm-null}
t0 <- proc.time()[3]

dgm_null <- generate_glm_dgm(
  data          = df.analysis,
  factor_vars   = dgm_factor_vars,
  outcome_var   = "status",
  treatment_var = "hormon",
  outcome_type  = "count",
  effect_measure = "IRR",
  offset_var    = "time_months",
  subgroup_vars = dgm_subgroup_vars,
  subgroup_cuts = dgm_subgroup_cuts,
  model         = "null",
  n_super       = dgm_n_super,
  seed          = sim_seed,
  verbose       = TRUE
)

cat("\n=== GBSG Null DGM ===\n")
cat("  IRR(ITT):", round(dgm_null$hazard_ratios$overall, 3), "\n")
cat("  IRR(H):  ", round(dgm_null$hazard_ratios$harm_subgroup, 3), "\n")
cat("  IRR(Hc): ", round(dgm_null$hazard_ratios$no_harm_subgroup, 3), "\n")

timings$dgm_null <- proc.time()[3] - t0
```


# Single Trial Verification {#sec-single-trial}

Verify that `simulate_from_glm_dgm()` and `run_simulation_analysis()`
work end-to-end on a single replicate before the full simulation.

```{r single-trial}
#| code-fold: false
t0 <- proc.time()[3]

df_single <- simulate_from_glm_dgm(
  dgm  = dgm_calibrated,
  n    = sim_n_sample,
  seed = sim_seed + 1L
)

cat(sprintf("Single trial: N = %d, event rate = %.1f%%\n",
    nrow(df_single), 100 * mean(df_single$y_sim)))
cat(sprintf("  H prevalence: %.1f%%\n",
    100 * mean(df_single$flag_harm)))

# Verify IRR in the simulated data
fit_single <- glm(y_sim ~ treat_sim, data = df_single, family = poisson(),
                  offset = log(pmax(df_single$time_months, 1e-6)))
cat(sprintf("  Simulated ITT IRR: %.3f (DGM: %.3f)\n",
    exp(coef(fit_single)["treat_sim"]),
    dgm_calibrated$hazard_ratios$overall))

timings$single_trial <- proc.time()[3] - t0
```


# Running Simulation Studies {#sec-sims}

## Setup

```{r sim-setup}
#| code-fold: false

plan(multisession, workers = n_workers)
cat("Using", n_workers, "parallel workers\n")

confounders_gbsg <- dgm_calibrated$factor_vars

fs_params <- list(
  outcome.name           = "y_sim",
  treat.name             = "treat_sim",
  id.name                = "id",

  # GLM pipeline
  outcome_type           = "count",
  effect_measure         = "IRR",
  offset.name            = "time_months",

  # Variable screening
  use_lasso              = fs_use_lasso,
  use_grf                = fs_use_grf,

  # Subgroup detection thresholds (GLM-era names)
  effect.threshold       = fs_irr_threshold,
  consistency.threshold  = fs_irr_consistency,
  pconsistency.threshold = fs_pconsistency,

  # Computational constraints
  fs.splits              = fs_splits,
  n.min                  = fs_n_min,
  d0.min                 = fs_d0_min,
  d1.min                 = fs_d1_min,
  maxk                   = fs_maxk,
  vi.grf.min             = fs_vi_grf_min,
  seedit                 = fs_seedit,

  # Inner parallelism: sequential within each worker
  parallel_args          = list(
    plan = "sequential", workers = 1, show_message = FALSE
  )
)
```


## Alternative Hypothesis Simulations

```{r run-alt}
cat("Running", nsims_alt, "simulations under H1...\n")
t0 <- proc.time()[3]

results_alt_raw <- foreach(
  sim = seq_len(nsims_alt),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_calibrated,
    n_sample         = sim_n_sample,
    confounders_base = confounders_gbsg,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    verbose          = FALSE
  )
}

timings$sims_alt <- proc.time()[3] - t0

results_alt <- collect_results(results_alt_raw, "H1 alt")
n_failed_alt <- attr(results_alt, "n_failed")

cat("Completed in", round(timings$sims_alt / 60, 1), "minutes\n")
cat("Results:", nrow(results_alt), "rows\n")
if (n_failed_alt > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_alt))
cat(sprintf("Detection rate: %.1f%%\n",
    100 * mean(results_alt$any.H[results_alt$analysis == "FS"])))
```


## Null Hypothesis Simulations

```{r run-null}
cat("Running", nsims_null, "simulations under H0...\n")
t0 <- proc.time()[3]

results_null_raw <- foreach(
  sim = seq_len(nsims_null),
  .errorhandling = "pass",
  .options.future = list(
    packages = c("forestsearch", "survival", "data.table", "future"),
    seed = TRUE
  )
) %dofuture% {
  future::plan("sequential")
  run_simulation_analysis(
    sim_id           = sim,
    dgm              = dgm_null,
    n_sample         = sim_n_sample,
    confounders_base = confounders_gbsg,
    n_add_noise      = sim_n_noise,
    run_fs           = run_fs,
    run_fs_grf       = run_fs_grf,
    run_grf          = run_grf,
    fs_params        = fs_params,
    verbose          = FALSE
  )
}

timings$sims_null <- proc.time()[3] - t0

results_null <- collect_results(results_null_raw, "H0 null")
n_failed_null <- attr(results_null, "n_failed")

cat("Completed in", round(timings$sims_null / 60, 1), "minutes\n")
if (n_failed_null > 0) cat(sprintf("  *** %d FAILED ***\n", n_failed_null))
cat(sprintf("False discovery rate: %.1f%%\n",
    100 * mean(results_null$any.H[results_null$analysis == "FS"])))
```

```{r cleanup-plan}
plan(sequential)
```

```{r save-results}
#| code-fold: false

if (save_results) {
  if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

  bundle <- list(
    results_alt    = results_alt,
    results_null   = results_null,
    dgm_calibrated = dgm_calibrated,
    dgm_null       = dgm_null,
    config = list(
      nsims_alt    = nsims_alt,
      nsims_null   = nsims_null,
      n_sample     = sim_n_sample,
      n_noise      = sim_n_noise,
      n_workers    = n_workers,
      fs_params    = fs_params,
      target_irr   = cal_target_irr,
      subgroup     = paste(dgm_subgroup_vars, collapse = " & ")
    ),
    true_effects = list(
      irr_H    = dgm_calibrated$hazard_ratios$harm_subgroup,
      irr_Hc   = dgm_calibrated$hazard_ratios$no_harm_subgroup,
      irr_ITT  = dgm_calibrated$hazard_ratios$overall,
      k_inter  = dgm_calibrated$model_params$k_inter
    ),
    timings = timings
  )

  fname <- sprintf("gbsg_poisson_simulations_%dalts_%dnulls.rds",
      nsims_alt, nsims_null)
  saveRDS(bundle, file.path(output_dir, fname))
  cat("Saved:", file.path(output_dir, fname), "\n")
} else {
  cat("save_results = FALSE; skipping save.\n")
}
```


# Results {#sec-results}

## Operating Characteristics Under H1

```{r oc-alt}
t0 <- proc.time()[3]

format_oc_results(
  results_alt,
  metrics   = c("detection", "classification"),
  digits    = 2,
  digits_hr = 2,
  title     = "GBSG Poisson: Subgroup Identification (Harm Search)",
  subtitle  = sprintf(
    "n = %d, %d sims, IRR(H) = %.2f, IRR(Hc) = %.2f",
    sim_n_sample, nsims_alt,
    dgm_calibrated$hazard_ratios$harm_subgroup,
    dgm_calibrated$hazard_ratios$no_harm_subgroup
  )
)

build_estimation_table(
  results         = results_alt,
  dgm             = dgm_calibrated,
  analysis_method = "FS",
  subtitle = sprintf(
    "GBSG Poisson DGM | n = %d, %d sims",
    sim_n_sample, nsims_alt
  ),
  font_size = 16
)

timings$report_alt <- proc.time()[3] - t0
```


## Operating Characteristics Under H0

```{r oc-null}
format_oc_results(
  results_null,
  metrics  = c("detection", "classification"),
  digits   = 2,
  title    = "GBSG Poisson: Type I Error (Null)",
  subtitle = sprintf(
    "n = %d, %d sims, IRR(ITT) = %.2f",
    sim_n_sample, nsims_null,
    dgm_null$hazard_ratios$overall
  )
)
```


## Classification Table

```{r classification}
build_classification_table(
  scenario_results = list(
    null = list(
      results    = results_null,
      label      = "GBSG-pois",
      n_sample   = sim_n_sample,
      dgm        = dgm_null,
      hypothesis = "null"
    ),
    alt = list(
      results    = results_alt,
      label      = "GBSG-pois",
      n_sample   = sim_n_sample,
      dgm        = dgm_calibrated,
      hypothesis = "alt"
    )
  ),
  analyses = sort(unique(c(
    unique(results_null$analysis),
    unique(results_alt$analysis)
  ))),
  digits    = 2,
  font_size = 16,
  title     = "GBSG Poisson: Subgroup Identification and Classification",
  n_sims    = nsims_alt
)
```


# Timing Summary {#sec-timing}

```{r timing}
timings$total <- Reduce(`+`, timings)

timing_df <- data.frame(
  Step = names(timings),
  Seconds = round(unlist(timings), 1),
  Minutes = round(unlist(timings) / 60, 2)
)

timing_df |>
  gt() |>
  gt::tab_header(title = "Computation Timing") |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(rows = Step == "total")
  )
```


# Discussion {#sec-discussion}

## Connection to the Cox Analysis

The Poisson rate model with a single interval per patient is the
piecewise exponential approximation to Cox regression.  Under
exponential survival, IRR = HR exactly.  This simulation validates
that ForestSearch recovers the same subgroup (ER-negative) and
similar effect sizes when using the GLM count pipeline instead of
the Cox pipeline.

## Key Differences from the Binary Simulation

1. **Offset**: The Poisson model includes `log(time_months)` as an
   offset.  This carries through `generate_glm_dgm()` (formula-based
   offset), `simulate_from_glm_dgm()` (retained in simulated data),
   and `forestsearch()` (via `offset.name`).

2. **LASSO screening**: Uses `glmnet` with `family = "poisson"` and
   `offset = log(time_months)`.

3. **GRF screening**: Uses `causal_forest()` on `log(Y + 0.5)` for
   variance stabilisation.

4. **No censoring model**: The count outcome is fully observed.
   Person-time variability is captured by the offset.

## Limitations

- The DGM assumes a constant baseline hazard (implicit in the
  single-interval Poisson model).  Departures from this — such as
  non-proportional hazards — would make IRR ≠ HR.

- With `status` ∈ {0, 1}, count data machinery is applied to binary
  data.  True recurrent-event count simulations (where Y > 1 is
  possible) would exercise the Poisson DGM more thoroughly.


# Session Info {#sec-session}

```{r session-info}
sessionInfo()
```


# References {#sec-references}

- León LF, Jemielita T, Guo Z, Marceau West R, Anderson KM (2024).
  "Exploratory subgroup identification in the heterogeneous Cox model."
  *Statistics in Medicine*, 43, 3921--3942.
