---
title: "DINA Survival Analysis (Cox and Poisson + Offset) — GBSG"
subtitle: "DINA analogs to the ForestSearch GBSG analysis, with Cox–Poisson alignment"
author: "forestsearch v0.2.0"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = "#>", fig.align = "center")

timings <- list()
t_doc_start <- proc.time()
```

# Introduction

This vignette is the DINA analog of `gbsg_survival_cox_poisson.qmd`.
It applies the **DINA** (Doubly-robust Iterative Neyman-Orthogonal)
estimator to the German Breast Cancer Study Group trial, in two
parallel branches:

1. **DINA Cox** — partial-likelihood DINA, mirroring the
   ForestSearch Cox analysis.
2. **DINA Poisson + offset** — Poisson rate-model DINA with
   $\log(\text{time})$ as offset, mirroring the ForestSearch
   IRR analysis.

The Cox model is exact under the partial-likelihood / proportional
hazards assumption; the Poisson + offset model is an approximation
that is exact under a piecewise-constant hazard. Where the two
models give nearly the same answer, the proportional-hazards
assumption looks defensible; where they disagree, that's a flag.

The full inference toolkit applied to each branch:

* Cross-fit point estimation via `dina()`
* Threshold-search harm subgroup identification via
  `dina_subgroup()` with sandwich-variance Wald CI
* Selection-adjusted bootstrap percentile CI on the fixed-subgroup
  linear functional via `dina_subgroup_bootstrap()`
* Stability diagnostics: selection-frequency table,
  `n_subgroup_ci`, `threshold_ci`

The final section assesses **Cox–Poisson alignment**: do the two
DINA branches discover the same subgroup, with concordant effect
estimates?

The intent here is to mirror the FS doc's analysis flow on the same
data with the same parameter conventions, *not* to repeat all of the
FS doc's downstream artifacts (CV, KM by subgroup, forest plots).
Those exist in the FS doc; this document focuses on the DINA
inference pipeline and the Cox–Poisson comparison.

# Setup

## Load Required Packages

```{r load-packages}
#| code-fold: false
library(forestsearch)
library(survival)
library(data.table)
library(ggplot2)
library(gt)
library(doFuture)

theme_set(theme_minimal(base_size = 12))
```

## Analysis Parameters

```{r analysis-params}
#| code-fold: false

# ─── Shared parameters ───────────────────────────────────────────────────
n_cores         <- max(1L, floor(0.95 * parallel::detectCores(logical = FALSE)))
confounders     <- c("age", "meno", "size", "grade3", "nodes", "pgr", "er")
id_col          <- "id"
treat_col       <- "hormon"

# ─── DINA outcome columns ────────────────────────────────────────────────
cox_outcome     <- "time_months"
cox_status      <- "status"

pois_outcome    <- "status"          # event count (0/1)
pois_offset     <- "time_months_log" # precomputed log offset

# ─── DINA tuning ─────────────────────────────────────────────────────────
dina_n_folds    <- 2L                # cross-fitting folds inside each DINA fit
dina_seed       <- 8316951L

# ─── Harm threshold (natural-parameter scale: log-HR for Cox,
#     log-IRR for Poisson; same value since the ratio threshold is the same)
m_diff          <- log(1.25)         # = 0.2231; matches FS hr.threshold = 1.25
n_min           <- 60L               # matches FS n.min

# ─── Bootstrap ───────────────────────────────────────────────────────────
NB              <- 500L              # bootstrap iterations (use 1000+ for production)
boot_seed       <- 8316951L

# ─── Confidence level ────────────────────────────────────────────────────
alpha           <- 0.05
```

# Data: German Breast Cancer Study Group Trial

```{r data-setup}
df.analysis <- gbsg

df.analysis <- within(df.analysis, {
  id              <- seq_len(nrow(df.analysis))
  time_months     <- rfstime / 30.4375
  time_months_log <- log(pmax(time_months, 1e-6))
  grade3          <- as.integer(ifelse(grade == "3", 1, 0))
})

cat("Sample size:    ", nrow(df.analysis), "\n")
cat("Events:         ", sum(df.analysis[[cox_status]]),
    sprintf(" (%.1f%%)\n",
            100 * mean(df.analysis[[cox_status]])))
cat("Covariates:     ", paste(confounders, collapse = ", "), "\n")
cat("Cox outcome:    ", cox_outcome, "(+ status: ", cox_status, ")\n", sep = "")
cat("Poisson outcome:", pois_outcome, "(offset: ", pois_offset, ")\n", sep = "")
```

# ITT Sanity Check: Cox vs. Poisson + Offset

Quick verification that the Poisson + offset approximation lines up
with Cox at the ITT level before running the DINA branches.

```{r itt-sanity}
cox_itt <- survival::coxph(
  survival::Surv(time_months, status) ~ hormon,
  data = df.analysis
)
pois_itt <- stats::glm(
  status ~ hormon,
  family = stats::poisson(),
  offset = df.analysis$time_months_log,
  data = df.analysis
)

itt_table <- data.table(
  Quantity = c("Cox  HR",
               "Poisson IRR",
               "Absolute difference"),
  Value    = c(
    sprintf("%.4f", exp(stats::coef(cox_itt))),
    sprintf("%.4f", exp(stats::coef(pois_itt)[["hormon"]])),
    sprintf("%.4f", abs(exp(stats::coef(cox_itt)) -
                        exp(stats::coef(pois_itt)[["hormon"]])))
  )
)

itt_table |>
  gt::gt() |>
  gt::tab_header(
    title    = "ITT-level Cox vs. Poisson + offset",
    subtitle = "Sanity check before subgroup search"
  )
```

# Part I: DINA Cox Analysis

## Fit cross-fit DINA Cox model

```{r dina-cox-fit}
t0 <- proc.time()

fit_dina_cox <- dina(
  df         = df.analysis,
  outcome    = cox_outcome,
  treatment  = treat_col,
  covariates = confounders,
  family     = "cox",
  status     = cox_status,
  n_folds    = dina_n_folds,
  seed       = dina_seed
)

timings$dina_cox_fit <- (proc.time() - t0)[["elapsed"]]
cat("DINA Cox fit completed in",
    round(timings$dina_cox_fit, 1), "seconds\n")
```

## Subgroup search (sandwich Wald CI)

`m_diff = log(1.25) ≈ 0.223` on the log-HR scale corresponds to the
HR ≥ 1.25 screening threshold used by ForestSearch.

```{r dina-cox-subgroup}
sg_cox <- dina_subgroup(
  fit        = fit_dina_cox,
  df         = df.analysis,
  covariates = confounders,
  m_diff     = m_diff,
  n_min      = n_min,
  direction  = "both",
  alpha      = alpha
)

print(sg_cox)
```

## Selection-adjusted bootstrap inference

```{r dina-cox-bootstrap}
t0 <- proc.time()

future::plan(future::multisession, workers = n_cores)

bs_cox <- dina_subgroup_bootstrap(
  df         = df.analysis,
  outcome    = cox_outcome,
  treatment  = treat_col,
  covariates = confounders,
  family     = "cox",
  status     = cox_status,
  m_diff     = m_diff,
  n_min      = n_min,
  direction  = "both",
  alpha      = alpha,
  n_boot     = NB,
  parallel   = "boots",
  n_folds    = dina_n_folds,
  seed       = boot_seed
)

future::plan("sequential")
gc(verbose = FALSE)

timings$dina_cox_boot <- (proc.time() - t0)[["elapsed"]]
cat("DINA Cox bootstrap completed in",
    round(timings$dina_cox_boot / 60, 2), "minutes\n")
```

```{r dina-cox-bootstrap-print}
print(bs_cox)
```

# Part II: DINA Poisson + Offset Analysis

The Poisson + offset model approximates the Cox model under
proportional hazards. Estimands on the log-rate scale are
interpreted analogously to log-HR for Cox.

## Fit cross-fit DINA Poisson + offset model

```{r dina-pois-fit}
t0 <- proc.time()

# NOTE on offset argument naming: if dina() expects `offset.name` rather
# than `offset` for the column-name interface, rename the argument here.
fit_dina_pois <- dina(
  df         = df.analysis,
  outcome    = pois_outcome,
  treatment  = treat_col,
  covariates = confounders,
  family     = "poisson",
  offset     = pois_offset,
  n_folds    = dina_n_folds,
  seed       = dina_seed
)

timings$dina_pois_fit <- (proc.time() - t0)[["elapsed"]]
cat("DINA Poisson fit completed in",
    round(timings$dina_pois_fit, 1), "seconds\n")
```

## Subgroup search

`m_diff = log(1.25)` on the log-IRR scale — same value as Cox, since
the ratio threshold (1.25) is conceptually the same target.

```{r dina-pois-subgroup}
sg_pois <- dina_subgroup(
  fit        = fit_dina_pois,
  df         = df.analysis,
  covariates = confounders,
  m_diff     = m_diff,
  n_min      = n_min,
  direction  = "both",
  alpha      = alpha
)

print(sg_pois)
```

## Selection-adjusted bootstrap inference

```{r dina-pois-bootstrap}
t0 <- proc.time()

future::plan(future::multisession, workers = n_cores)

bs_pois <- dina_subgroup_bootstrap(
  df         = df.analysis,
  outcome    = pois_outcome,
  treatment  = treat_col,
  covariates = confounders,
  family     = "poisson",
  offset     = pois_offset,
  m_diff     = m_diff,
  n_min      = n_min,
  direction  = "both",
  alpha      = alpha,
  n_boot     = NB,
  parallel   = "boots",
  n_folds    = dina_n_folds,
  seed       = boot_seed
)

future::plan("sequential")
gc(verbose = FALSE)

timings$dina_pois_boot <- (proc.time() - t0)[["elapsed"]]
cat("DINA Poisson bootstrap completed in",
    round(timings$dina_pois_boot / 60, 2), "minutes\n")
```

```{r dina-pois-bootstrap-print}
print(bs_pois)
```

# Part III: DINA Cox vs. DINA Poisson Alignment

## Helper functions

```{r alignment-helpers}
`%||%` <- function(a, b) if (is.null(a)) b else a

# Helper to format a (covariate, direction, threshold) signature
fmt_signature <- function(sg) {
  if (!isTRUE(sg$found)) return("—")
  cmp <- if (sg$direction == "left") "<=" else ">="
  sprintf("%s %s %.3f", sg$covariate, cmp, sg$threshold)
}

# Helper to format a CI on the ratio scale (exponentiated)
fmt_ratio_ci <- function(ci, digits = 3L) {
  if (any(!is.finite(ci))) return("—")
  sprintf("[%.*f, %.*f]",
          digits, exp(ci[["lower"]]),
          digits, exp(ci[["upper"]]))
}

# Helper to format a CI on the log scale (raw)
fmt_log_ci <- function(ci, digits = 3L) {
  if (any(!is.finite(ci))) return("—")
  sprintf("[%.*f, %.*f]",
          digits, ci[["lower"]],
          digits, ci[["upper"]])
}

# Helper to format an integer CI (n_subgroup)
fmt_int_ci <- function(ci) {
  if (any(!is.finite(ci))) return("—")
  sprintf("[%.0f, %.0f]", ci[["lower"]], ci[["upper"]])
}

# Helper: top selection frequency line
fmt_top_selections <- function(bs, k = 3L) {
  if (is.null(bs$selection_frequency) || bs$n_boot_found == 0L) return("—")
  sel_long <- as.data.frame(bs$selection_frequency,
                            stringsAsFactors = FALSE)
  sel_long <- sel_long[sel_long$Freq > 0L, , drop = FALSE]
  sel_long <- sel_long[order(-sel_long$Freq), , drop = FALSE]
  top <- head(sel_long, k)
  paste(sprintf("%s (%s): %d/%d (%.0f%%)",
                top$covariate, top$direction, top$Freq,
                bs$n_boot_found,
                100 * top$Freq / bs$n_boot_found),
        collapse = "; ")
}

# Defensive Wald-CI extractor from a dina_subgroup result.
# Tries several plausible field-name conventions before falling back
# to (NA, NA).  Adjust as needed once the actual field names are
# confirmed.
extract_wald_ci <- function(sg) {
  na_ci <- c(lower = NA_real_, upper = NA_real_)
  if (!isTRUE(sg$found)) return(na_ci)

  # Common length-2 vector field names
  vec_candidates <- c("mean_tau_ci", "wald_ci", "ci")
  for (cand in vec_candidates) {
    if (cand %in% names(sg) && length(sg[[cand]]) == 2L) {
      v <- as.numeric(sg[[cand]])
      return(c(lower = v[1L], upper = v[2L]))
    }
  }

  # Common paired scalar field names
  pair_candidates <- list(
    c("mean_tau_ci_lower", "mean_tau_ci_upper"),
    c("mean_tau_ci_lo",    "mean_tau_ci_hi"),
    c("ci_lower",          "ci_upper"),
    c("lower",             "upper")
  )
  for (cand in pair_candidates) {
    if (all(cand %in% names(sg))) {
      return(c(lower = as.numeric(sg[[cand[1L]]]),
               upper = as.numeric(sg[[cand[2L]]])))
    }
  }

  na_ci
}
```

## Side-by-side summary table

```{r alignment-table}
# Pull common pieces
pt_cox  <- bs_cox$point
pt_pois <- bs_pois$point

# Extract Wald CIs defensively (try multiple field-name conventions)
wald_cox  <- extract_wald_ci(pt_cox)
wald_pois <- extract_wald_ci(pt_pois)

# Build the comparison rows.  Log-scale rows are the natural DINA
# output; ratio-scale rows are the exponentiated interpretation (HR
# for Cox, IRR for Poisson).
align_tbl <- data.table(
  Field = c(
    "Discovered subgroup signature",
    "n_subgroup",
    "Bootstrap found / total",
    "Bootstrap failed",
    "Modal-match iters / found",
    "Point: mean tau-hat (log scale)",
    "Point: exp(mean tau-hat) (ratio scale)",
    "Sandwich Wald CI (log scale)",
    "Sandwich Wald CI (ratio scale)",
    "Bootstrap effect CI (log scale)",
    "Bootstrap effect CI (ratio scale)",
    "n_subgroup CI (modal-match)",
    "Threshold CI (modal-match)",
    "Top (covariate, direction) selections"
  ),
  `DINA Cox (HR)` = c(
    fmt_signature(pt_cox),
    sprintf("%d / %d", pt_cox$n_subgroup %||% NA, pt_cox$n_total %||% NA),
    sprintf("%d / %d", bs_cox$n_boot_found, bs_cox$n_boot),
    sprintf("%d", bs_cox$n_boot_failed),
    if (!is.na(bs_cox$n_modal_match))
      sprintf("%d / %d", bs_cox$n_modal_match, bs_cox$n_boot_found)
    else "—",
    sprintf("%.3f", pt_cox$mean_tau_hat %||% NA_real_),
    sprintf("%.3f", exp(pt_cox$mean_tau_hat %||% NA_real_)),
    fmt_log_ci(wald_cox),
    fmt_ratio_ci(wald_cox),
    fmt_log_ci(bs_cox$effect_ci),
    fmt_ratio_ci(bs_cox$effect_ci),
    fmt_int_ci(bs_cox$n_subgroup_ci),
    fmt_log_ci(bs_cox$threshold_ci),
    fmt_top_selections(bs_cox)
  ),
  `DINA Poisson + offset (IRR)` = c(
    fmt_signature(pt_pois),
    sprintf("%d / %d", pt_pois$n_subgroup %||% NA, pt_pois$n_total %||% NA),
    sprintf("%d / %d", bs_pois$n_boot_found, bs_pois$n_boot),
    sprintf("%d", bs_pois$n_boot_failed),
    if (!is.na(bs_pois$n_modal_match))
      sprintf("%d / %d", bs_pois$n_modal_match, bs_pois$n_boot_found)
    else "—",
    sprintf("%.3f", pt_pois$mean_tau_hat %||% NA_real_),
    sprintf("%.3f", exp(pt_pois$mean_tau_hat %||% NA_real_)),
    fmt_log_ci(wald_pois),
    fmt_ratio_ci(wald_pois),
    fmt_log_ci(bs_pois$effect_ci),
    fmt_ratio_ci(bs_pois$effect_ci),
    fmt_int_ci(bs_pois$n_subgroup_ci),
    fmt_log_ci(bs_pois$threshold_ci),
    fmt_top_selections(bs_pois)
  )
)

align_tbl |>
  gt::gt() |>
  gt::tab_header(
    title    = gt::md("**DINA Cox vs. DINA Poisson + offset — alignment summary**"),
    subtitle = sprintf("m_diff = log(1.25) = %.3f;  n_boot = %d;  %d cross-fitting folds",
                       m_diff, NB, dina_n_folds)
  ) |>
  gt::tab_source_note(
    source_note = gt::md(
      "*Sandwich Wald CIs are from `dina_subgroup()` on the original data
       (asymptotic-normal, sandwich variance).  Bootstrap effect CIs are
       percentile CIs on `a*'beta_b` with `a*` fixed from the original-data
       subgroup, via `dina_subgroup_bootstrap()`.  If Sandwich CI rows
       display '—', the field-name conventions in `extract_wald_ci()` need
       updating to match this DINA build.*"
    )
  )
```

## Subgroup membership concordance

How many patients are jointly classified as "in the harm subgroup"
by each branch? If the proportional-hazards assumption is defensible,
the two branches should partition patients similarly.

```{r alignment-membership}
mask_cox  <- if (isTRUE(pt_cox$found))  pt_cox$mask  else rep(FALSE, nrow(df.analysis))
mask_pois <- if (isTRUE(pt_pois$found)) pt_pois$mask else rep(FALSE, nrow(df.analysis))

concord_tab <- table(
  `DINA Cox`     = factor(mask_cox,
                          levels = c(TRUE, FALSE),
                          labels = c("In Cox H", "In Cox H^c")),
  `DINA Poisson` = factor(mask_pois,
                          levels = c(TRUE, FALSE),
                          labels = c("In Pois H", "In Pois H^c"))
)

agree_n   <- concord_tab[1, 1] + concord_tab[2, 2]
agree_pct <- 100 * agree_n / sum(concord_tab)

# Cohen's kappa (no external package: hand-compute from the 2x2 table)
n      <- sum(concord_tab)
p_obs  <- agree_n / n
p_cox  <- sum(concord_tab[1, ]) / n
p_pois <- sum(concord_tab[, 1]) / n
p_exp  <- p_cox * p_pois + (1 - p_cox) * (1 - p_pois)
kappa  <- (p_obs - p_exp) / (1 - p_exp)

concord_df <- as.data.frame.matrix(concord_tab)
concord_df$.row <- rownames(concord_df)
setDT(concord_df)
setcolorder(concord_df, c(".row", setdiff(names(concord_df), ".row")))
setnames(concord_df, ".row", " ")

concord_df |>
  gt::gt() |>
  gt::tab_header(
    title    = "Subgroup membership concordance",
    subtitle = sprintf("Agreement: %d / %d (%.1f%%);  Cohen's kappa = %.3f",
                       agree_n, n, agree_pct, kappa)
  )
```

## Discussion: Cox–Poisson alignment

The natural reading of this table:

* If the **discovered subgroup signature is identical** across
  branches (same covariate, same direction, similar threshold) and
  the **bootstrap effect CIs overlap heavily** on the ratio scale,
  the Poisson + offset approximation is tracking the Cox analysis
  closely on this dataset.
* If signatures **differ** but membership concordance is still high
  and agreement is mostly on the larger complement subgroup, the
  procedures are picking up the same underlying signal at slightly
  different boundaries.
* If signatures differ **and** membership concordance is low, that's
  a flag — either the proportional-hazards assumption is shaky for
  the harm subgroup or one branch has selected an unstable boundary.
  Read the selection-frequency rows together with the threshold CIs
  to localize which.

# Timings

```{r timings-table}
tim_df <- data.table(
  Step       = c("DINA Cox fit",
                 "DINA Cox bootstrap",
                 "DINA Poisson fit",
                 "DINA Poisson bootstrap",
                 "Total"),
  `Time (s)` = c(
    sprintf("%.1f", timings$dina_cox_fit),
    sprintf("%.1f", timings$dina_cox_boot),
    sprintf("%.1f", timings$dina_pois_fit),
    sprintf("%.1f", timings$dina_pois_boot),
    sprintf("%.1f", (proc.time() - t_doc_start)[["elapsed"]])
  )
)

tim_df |>
  gt::gt() |>
  gt::tab_header(title = "Timing summary")
```

# Session Information

```{r session-info}
sessionInfo()
```
