---
title: "Extreme Subgroup Effects Under a Uniform Treatment Effect: ACTG 175, Continuous Outcome"
subtitle: "Fixed-baseline (conditional-on-X) simulation on the observed panel"
author: "Larry F. León"
date: today
params:
  n_sims: null
  save_payload: null
  payload_dir: null
format:
  html:
    embed-resources: true
    toc: true
    number-sections: true
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
# =========================================================================
# RUN CONTROLS -- single source of truth; -P overrides without editing:
#   quarto render actg175_extreme_subgroups_continuous_fixed.qmd -P n_sims:50
# Precedence: -P value > in-file default (YAML null = use in-file default).
# =========================================================================
# [delivery sentinel: c2r1-f64d0856]
.param_or <- function(nm, default) {
  if (exists("params", inherits = TRUE) && !is.null(params[[nm]]))
    params[[nm]] else default
}
design_label   <- "fixed"            # conditional-on-X; twin of the gbsg doc
n_sims_uniform <- as.integer(.param_or("n_sims", 10000L))
save_payload   <- isTRUE(.param_or("save_payload", TRUE))
payload_dir    <- .param_or(
  "payload_dir",
  "/media/larryleon/bigstorage/forestsearch_sims/fixed_random")
payload_file <- sprintf("actg175_extreme_sims_continuous_%s_%d_payload.rds",
                        design_label, n_sims_uniform)
if (save_payload && !dir.exists(payload_dir)) {
  stop("payload_dir does not exist: ", payload_dir,
       "\nDeliberately NOT created automatically (unmounted-drive guard, ",
       "as in the gbsg twin). mkdir -p it or pass -P payload_dir:...")
}
library(forestsearch)
library(ggplot2)
seed_dgm  <- 175L
seed_base <- 0L
```

# Motivation

The gbsg twin of this document asks: under a data-generating mechanism
with a **perfectly uniform treatment effect**, how often do standard
subgroup analyses of a time-to-event outcome produce extreme-looking
results purely by chance? This document asks the identical question for
a **continuous endpoint** -- the ACTG 175 CD4 decline -- analyzed by
mean differences (MD) with the identical fixed-baseline design: the
observed patient panel is frozen across every replicate
(conditional-on-X), so all variability comes from treatment assignment
and outcome noise, never from covariate resampling. Every subgroup
shares exactly the same true MD by construction; any apparent
heterogeneity is a false positive.

# Part I --- Data and the Uniform DGM

```{r data-prep}
# ACTG 175 (speff2trial), arms {1, 3}: ZDV+ddI vs ddC+ZDV comparison used
# by the continuous OC documents. Outcome is CD4 DECLINE (baseline minus
# week 20+/-5), so LARGER = WORSE -- harm-positive with subgroup_glm()'s
# defaults, no overrides anywhere downstream.
actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$treat     <- ifelse(actg_df$arms == 1L, 1L, 0L)
actg_df <- actg_df[!is.na(actg_df$cd420), ]
actg_df$y_decline <- actg_df$cd40 - actg_df$cd420
actg_df$id        <- seq_len(nrow(actg_df))
cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
bin_vars  <- c("hemo", "homo", "drugs", "race", "gender", "symptom")
stopifnot(nrow(actg_df) == 1083L)      # 522 treated / 561 control
# Panel-frozen cut constants (computed ONCE on the observed panel).
cuts <- list(
  age_med  = stats::median(actg_df$age),
  wtkg_med = stats::median(actg_df$wtkg),
  kar_med  = stats::median(actg_df$karnof),
  cd40_med = stats::median(actg_df$cd40),
  cd40_q1  = unname(stats::quantile(actg_df$cd40, 0.25)),
  cd80_med = stats::median(actg_df$cd80),
  pre_med  = stats::median(actg_df$preanti)
)
```

```{r build-dgm}
# model = "null": beta_inter is forced to 0 and flag_harm is identically
# zero -- there is no embedded subgroup. Truth bookkeeping under the
# null: effect_Qc = effect_ITT (Q is the empty region, so effect_Q is NA
# by construction and is never printed anywhere in this document). The
# reference truth is the FITTED overall treatment effect, k_treat = 1:
# the continuous analogue of the twin's imposed HR = 0.70 anchors to the
# data's own MD.
dgm_uniform <- generate_glm_dgm(
  data = actg_df,
  factor_vars = bin_vars, continuous_vars = cont_vars,
  outcome_var = "y_decline", treatment_var = "treat",
  outcome_type = "continuous", effect_measure = "MD",
  subgroup_vars = NULL, model = "null",
  n_super = 5000L, seed = seed_dgm, verbose = FALSE)
md_true <- dgm_uniform$hazard_ratios$overall
stopifnot(!is.null(dgm_uniform$df_source),
          nrow(dgm_uniform$df_source) == nrow(actg_df))
cat(sprintf("Uniform true MD (decline scale, fitted) = %.3f\n", md_true))
```

# Part II --- Fixed-Panel Diagnostics

Where the survival twin spends this part on censoring calibration --
machinery a fully observed outcome does not have -- the fixed-baseline
design's analogous obligation is to display the frozen panel itself.

```{r panel-diagnostics}
src <- dgm_uniform$df_source
cat(sprintf("Frozen panel: N = %d (every replicate; arms redrawn 1:1)\n",
            nrow(src)))
cat(sprintf("y_decline: mean = %.2f, sd = %.2f\n",
            mean(src$y_decline), stats::sd(src$y_decline)))
se_itt <- stats::sd(src$y_decline) * sqrt(4 / nrow(src))
cat(sprintf("ITT-scale MD SE at N = %d: %.3f  (UB tails below use the\n",
            nrow(src), se_itt))
cat("round values 10 and 20, bracketing ~1x and ~2x this SE: 7.889/15.78)\n")
```

# Part III --- Subgroup Catalog (`r 47 + 4` total)

```{r subgroups}
subgroups <- list(
  list(id = "flag_itt == 1",              name = "All Patients",        grp = "ITT"),
  # Clinical binaries (both levels each)
  list(id = "hemo == 1",    name = "Hemophilia",        grp = "Clinical"),
  list(id = "hemo == 0",    name = "No hemophilia",     grp = "Clinical"),
  list(id = "homo == 1",    name = "MSM",               grp = "Clinical"),
  list(id = "homo == 0",    name = "Non-MSM",           grp = "Clinical"),
  list(id = "drugs == 1",   name = "IVDU history",      grp = "Clinical"),
  list(id = "drugs == 0",   name = "No IVDU",           grp = "Clinical"),
  list(id = "race == 1",    name = "Non-white",         grp = "Clinical"),
  list(id = "race == 0",    name = "White",             grp = "Clinical"),
  list(id = "gender == 0",  name = "Female",            grp = "Clinical"),
  list(id = "gender == 1",  name = "Male",              grp = "Clinical"),
  list(id = "symptom == 1", name = "Symptomatic",       grp = "Clinical"),
  list(id = "symptom == 0", name = "Asymptomatic",      grp = "Clinical"),
  # Continuous splits (panel-frozen cuts)
  list(id = "age <= age_med",   name = "Age (young)",     grp = "Continuous"),
  list(id = "age > age_med",    name = "Age (older)",     grp = "Continuous"),
  list(id = "age <= 35",        name = "Age <= 35",       grp = "Continuous"),
  list(id = "age > 35",         name = "Age > 35",        grp = "Continuous"),
  list(id = "wtkg <= wtkg_med", name = "Weight (low)",    grp = "Continuous"),
  list(id = "wtkg > wtkg_med",  name = "Weight (high)",   grp = "Continuous"),
  list(id = "karnof <= kar_med", name = "Karnofsky low",  grp = "Continuous"),
  list(id = "karnof > kar_med",  name = "Karnofsky high", grp = "Continuous"),
  list(id = "cd40 <= cd40_med", name = "CD4 low",         grp = "Continuous"),
  list(id = "cd40 > cd40_med",  name = "CD4 high",        grp = "Continuous"),
  list(id = "cd40 <= cd40_q1",  name = "CD4 lowest quartile", grp = "Continuous"),
  list(id = "cd80 <= cd80_med", name = "CD8 low",         grp = "Continuous"),
  list(id = "cd80 > cd80_med",  name = "CD8 high",        grp = "Continuous"),
  list(id = "preanti <= pre_med", name = "Short prior ART", grp = "Continuous"),
  list(id = "preanti > pre_med",  name = "Long prior ART",  grp = "Continuous"),
  # Two-way interactions
  list(id = "symptom == 1 & cd40 <= cd40_med", name = "Sympt/CD4-low",   grp = "Interaction"),
  list(id = "symptom == 1 & cd40 > cd40_med",  name = "Sympt/CD4-high",  grp = "Interaction"),
  list(id = "symptom == 0 & cd40 <= cd40_med", name = "Asympt/CD4-low",  grp = "Interaction"),
  list(id = "age <= age_med & cd40 <= cd40_med", name = "Young/CD4-low", grp = "Interaction"),
  list(id = "age > age_med & cd40 <= cd40_med",  name = "Older/CD4-low", grp = "Interaction"),
  list(id = "age > age_med & symptom == 1",      name = "Older/Sympt",   grp = "Interaction"),
  list(id = "gender == 0 & drugs == 1",   name = "Female/IVDU",     grp = "Interaction"),
  list(id = "gender == 0 & race == 1",    name = "Female/Non-white",grp = "Interaction"),
  list(id = "karnof <= kar_med & cd40 <= cd40_med", name = "KarnLow/CD4-low", grp = "Interaction"),
  list(id = "hemo == 1 & age <= age_med", name = "Hemo/Young",      grp = "Interaction"),
  list(id = "drugs == 1 & symptom == 1",  name = "IVDU/Sympt",      grp = "Interaction"),
  list(id = "race == 1 & cd40 <= cd40_med", name = "Non-white/CD4-low", grp = "Interaction"),
  list(id = "wtkg <= wtkg_med & karnof <= kar_med", name = "LowWt/KarnLow", grp = "Interaction"),
  list(id = "preanti > pre_med & cd40 <= cd40_med", name = "LongART/CD4-low", grp = "Interaction"),
  # Three-way
  list(id = "symptom == 1 & cd40 <= cd40_med & age > age_med",
       name = "Sympt/CD4-low/Older", grp = "3-way"),
  list(id = "gender == 0 & drugs == 1 & cd40 <= cd40_med",
       name = "Fem/IVDU/CD4-low", grp = "3-way"),
  list(id = "race == 1 & symptom == 1 & cd40 <= cd40_med",
       name = "NW/Sympt/CD4-low", grp = "3-way"),
  list(id = "hemo == 1 & cd40 <= cd40_med & karnof <= kar_med",
       name = "Hemo/CD4-low/KarnLow", grp = "3-way")
)
# + random60/40/20/15 via benchmark_spec() (nested, re-drawn per replicate)
```

# Part IV --- Simulation and Results

```{r run-simulations}
n_cores_total <- future::availableCores()
n_workers     <- max(1L, ceiling(0.90 * n_cores_total))
future::plan(future::multisession, workers = n_workers)
sims_uniform <- run_subgroup_sims(
  dgm        = dgm_uniform,
  subgroups  = subgroups,
  n_sims     = n_sims_uniform,
  baseline   = "fixed",                 # frozen observed panel, N = 1083
  benchmarks = benchmark_spec(),        # random15 in 20 in 40 in 60
  cutpoints  = cuts,
  workers    = NULL,                    # reuse the plan above
  seed_base  = seed_base,
  hr_true    = md_true)
future::plan(future::sequential)
```

```{r save-payload}
#| eval: !expr save_payload
saveRDS(list(sim_hrs = sims_uniform$sim_hrs, sim_ubs = sims_uniform$sim_ubs,
             sim_ns = sims_uniform$sim_ns, subgroups = subgroups,
             effect = sims_uniform$effect, md_true = md_true,
             design = design_label, n_sims = n_sims_uniform,
             seed_dgm = seed_dgm, seed_base = seed_base,
             created = Sys.time()),
        file.path(payload_dir, payload_file))
```

```{r summaries}
# UB(MD) tails c(10, 20): round decline values bracketing ~1x / ~2x the
# ITT-scale SE (7.889 / 15.78) on the frozen panel -- the identity-scale
# analogue of the twin's UB(HR) >= 2 / >= 3.
S <- summary(sims_uniform, ub_thresholds = c(10, 20))
print(S)
```

```{r forest-panels}
#| fig-width: 10
plot(S, metric = "hr", panel = "single")
plot(S, metric = "hr", panel = "combo")
plot(S, metric = "ub", panel = "single")
plot(S, metric = "ub", panel = "combo")
plot(S, metric = "ub", panel = "highrisk")
```

# Findings

Written after the first full run, mirroring the twin's structure: the
MD point-estimate distributions center on the uniform truth in every
subgroup; ECI width is driven entirely by subgroup N; the clinical
subgroups of a given size are statistically indistinguishable from the
random benchmarks of that size; and UB(MD) exceedances of the c(10, 20)
tails in the smallest cells are expected under the uniform effect --
the calibration a reviewer needs before reading any single forest plot.

# Appendix: New-Results Workflow

Re-render with `-P n_sims:...` for a different scale; payloads load in
the design-comparison pattern established by the gbsg twins
(`readRDS(file.path(payload_dir, payload_file))` recovers the raw
matrices, subgroup list, and metadata).
