---
title: "Extreme Subgroup Effects Under a Uniform Treatment Benefit"
subtitle: "A Simulation Study Based on Observed Trial Data"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: show
    code-tools: true
    theme: cosmo
    highlight-style: tango
    fig-width: 8
    fig-height: 5
    self-contained: true
    html-math-method: mathjax
execute:
  echo: true
  warning: false
  message: false
params:
  n_sims: null
  save_payload: null
  payload_dir: null
  payload_file: null
---

```{r setup}
#| include: false

# =========================================================================
# RUN CONTROLS
# =========================================================================
# Defaults live HERE (single source of truth).  Every control below can
# also be overridden at render time -- without editing this file -- via
# Quarto params, e.g.:
#
#   quarto render extreme_subgroups.qmd -P n_sims:50
#   quarto render extreme_subgroups.qmd -P save_payload:false
#   quarto render extreme_subgroups.qmd \
#     -P payload_dir:/media/larryleon/bigstorage/forestsearch_sims/fixed_random
#
# Precedence: -P value > in-file default (a param left at its YAML
# default of `null` means "use the in-file default").  The exists()
# guard in .param_or() keeps interactive chunk execution in RStudio
# (where no `params` object exists) identical to a default render.
.param_or <- function(name, default) {
  if (exists("params", inherits = TRUE) && !is.null(params[[name]]))
    params[[name]] else default
}

# Design label: stored in the payload metadata and used in the default
# payload file name.  "resample" = random-X (population
# resampling); the fixed-baseline companion vignette uses "fixed".
design_label <- "resample"

# Number of uniform-DGM simulated trials -- the primary result quantity.
# Set to >= 1000 for publication-quality results.
n_sims_uniform <- as.integer(.param_or("n_sims", 10000L))

# =========================================================================
# PAYLOAD (RDS) PERSISTENCE -- one knob, off-switchable, fully relocatable
# =========================================================================
# save_payload = TRUE  -> after the Part IV simulation loop, a chunk
#     writes the raw result matrices (sim_hrs / sim_ubs / sim_ns), the
#     subgroup list, and run metadata as ONE RDS file ("payload") to
#     file.path(payload_dir, payload_file).  The companion memo
#     extreme_subgroups_design_comparison.qmd loads one payload from
#     this vignette and one from its companion and recomputes the full
#     fixed-vs-random design comparison from raw results.
# save_payload = FALSE -> nothing is written anywhere; the vignette
#     renders exactly as it did before payloads existed.
#
# payload_dir : any writable directory (a relative path resolves
#     against this file's folder at render time).  It is NOT created
#     automatically: if missing, the render STOPS here -- in seconds,
#     not after the simulation loop.  This is deliberate.  Silently
#     dir.create()-ing a path like /media/<user>/<drive>/... while the
#     external drive is unmounted would write to the empty mount point
#     on the root filesystem.  Create the directory yourself
#     (mkdir -p) or point payload_dir somewhere else.
#
# payload_file: the default name encodes design + n_sims
#     (e.g. extreme_sims_resample_10000_payload.rds), so smoke
#     runs and full runs never overwrite each other.  Override freely.
save_payload <- as.logical(.param_or("save_payload", TRUE))
payload_dir  <- .param_or("payload_dir", "results")
payload_file <- .param_or(
  "payload_file",
  sprintf("extreme_sims_%s_%d_payload.rds", design_label, n_sims_uniform)
)
payload_path <- file.path(payload_dir, payload_file)

if (isTRUE(save_payload) && !dir.exists(payload_dir)) {
  stop("save_payload = TRUE but payload_dir does not exist: '", payload_dir,
       "'.\n  Fix one of:\n",
       "  (a) create it:            mkdir -p ", payload_dir, "\n",
       "  (b) relocate it:          -P payload_dir:<existing-dir>\n",
       "  (c) disable persistence:  -P save_payload:false")
}
```

# Overview

A central concern in exploratory subgroup analysis is the potential for
**extreme results driven entirely by chance**, not by genuine treatment
effect heterogeneity.  When many subgroups are examined in a single
dataset — as is routine in regulatory submissions and trial reports —
small subgroups in particular can produce hazard ratio estimates that
look dramatic simply because they are noisy.  The smaller the subgroup,
the wider the confidence interval, and the higher the probability that
the upper bound strays far from the truth.

This vignette evaluates that risk directly.  We generate thousands of
synthetic trials from a data generating mechanism (DGM) with a
**perfectly uniform treatment benefit** — the true hazard ratio is
identical for every patient — and ask: how often do standard Cox model
analyses produce extreme subgroup results?  

The vignette is organised as a complete, self-contained workflow:

1. **Building the DGM** — fitting a Weibull AFT model to real data,
   constructing a super-population, and calibrating the censoring process
   so that simulated trials reproduce the censoring pattern of the source
   data.  This step establishes the uniform-effect scenario: a DGM where the true
   HR is 0.70 *for every patient*, with no subgroup differential effects.

2. **Running simulated trials** — drawing `r n_sims_uniform` synthetic
   trial replicates from the uniform DGM and fitting a stratified Cox model
   in each pre-defined subgroup.

3. **Characterising the distribution of extreme results** — for each
   subgroup, summarising how variable the Cox HR estimate is across
   simulations (the empirical 99% interval, or "ECI" — defined in
   Section 1.3 below) and how large the upper confidence bound UB(HR)
   can become by chance alone.  Random benchmark
   subgroups (`random15`, `random20`, `random40`, `random60`) with no
   clinical meaning serve as calibration anchors, making explicit that
   clinical subgroups of the same size can exhibit distributional patterns
   similar to (meaningless) random groupings.

All examples use the GBSG breast cancer dataset (`survival::gbsg`).


## Example of Subgroups Under Study 

For the study under question, the subgroups and features (size, events, etc) 
are of course specific to that study.  While we are using the GBSG breast cancer 
study as the case-study, the analyses can be applied or modified based on any study.

An example of pre-defined subgroups,
ranging from clinically motivated subpopulations to deliberately
artificial ones:

| Subgroup | Basis | Approximate N |
|:---------|:------|:--------------|
| All Patients (ITT) | Full trial | ~700 |
| No liver metastases | Clinical | ~400 |
| BRAF V600E mutant | Molecular | ~300 |
| Age ≤ 65 | Demographic | ~400 |
| Spain, Germany, UK | Country/region | ~60–150 |
| Asia Pacific region | Region | ~100 |
| ECOG 0 / ECOG 1 | Performance status | ~300 each |
| **random15, random20** | **Randomly sampled** | **~15, ~20** |
| **random40, random60** | **Randomly sampled** | **~40, ~60** |

The `random*` subgroups are the key benchmark.  They are constructed by
randomly sampling (15, 20, 40, and 60 patients).  Such subgroups have no clinical meaning whatsoever — their true HR
equals the ITT HR of 0.70.  Any extreme result they produce is purely
a function of their small size and sampling variability.


## Summary Statistics Across Simulations

For each subgroup, each simulated trial produces two per-trial values:
the **Cox HR point estimate** (stratified by `grade`) and the
**upper bound of the 95% confidence interval, UB(HR)**.  Across the
`r n_sims_uniform` simulated trials these per-trial values form an
empirical distribution for each subgroup.  All downstream summaries —
the results table and every forest panel — operate on this empirical
distribution.

### 99% empirical confidence interval (ECI)

The **99% ECI** is the **1st–99th percentile range** of the per-trial
values accumulated across all simulated trials — computed separately for
the Cox HR and for UB(HR).  It is a description of how much these
quantities vary from one synthetic trial to the next under the same
DGM.

The 99% ECI is **not** a confidence interval in the inferential sense.
A standard confidence interval is built from a single sample and is
designed to cover the unknown true parameter with a stated probability.
The ECI here is a different object: the true parameter is fixed and
known by construction (true HR = 0.70 everywhere), and what is being
summarised is the empirical spread of *estimates* across repeated
trials drawn from the DGM.  It answers "how variable is this estimator
in repeated draws?", not "what range of parameter values is consistent
with the data?".  The terminology ("empirical confidence interval") is
chosen to evoke the visual similarity to a forest-plot CI line — point
in the middle, whiskers spanning the range — not equivalence of
meaning.

### Six summary statistics (table + forest panels)

The results table and the forest-panel annotation columns report the
same six summary statistics for each subgroup.

**For the Cox HR point-estimate distribution:**

- **`Pr(HR<0.5)`** — fraction of simulated trials whose point estimate
  is below 0.5.  Measures how often the HR alone looks like an extreme
  benefit by chance.
- **`Pr(HR>1.0)`** — fraction of simulated trials whose point estimate
  is above 1.0.  Measures how often the estimate lands on the wrong
  side of no effect.
- **`mHR`** — unconditional median of the HR point estimate across all
  simulated trials.  Under the uniform DGM (true HR = 0.70) `mHR`
  should track 0.70 for every subgroup, confirming the Cox model is
  unbiased.

**For the upper 95% confidence bound UB(HR):**

- **`Pr(UB>=2)`** — fraction of simulated trials whose upper bound is
  at or above 2.0.  Measures how often a forest-plot row looks alarming
  to a safety reviewer at a moderately-strict threshold.
- **`Pr(UB>=3)`** — same, at a stricter threshold.
- **`mUB`** — unconditional median of the upper bound across all
  simulated trials.

The forest plots in the Results section visualise the median + 99% ECI
graphically; the annotation columns and the results table give the
same six summary statistics as numbers.

---

We note that continuous biomarkers can also be considered.  This is more involved and 
not necessary for uniform effects, but valuable for trials with biomarkers when there 
is interest in exploring the dynamics of subgroups in the presence of actual biomarker effects (correlated with other 
patients' characteristics).

## Setting: Uniform Treatment Effect

The DGM is built from a real trial dataset using `generate_aft_dgm_flex()`
with `model = "null"`.  The fitted Weibull AFT treatment coefficient is
then rescaled via `k_treat`, with the multiplier calibrated by `uniroot`
so that the **average hazard ratio (AHR)** matches a chosen target.
Under `model = "null"` the AHR equals the patient-level individual HR,
so calibrating the AHR fixes the true HR for every patient.  The key
feature of the resulting DGM is that the true log hazard ratio is
**constant across all patients**:

$$
\psi^0(\mathbf{L}) = \log(\text{HR}_{\text{true}}) \quad \text{for all } \mathbf{L}.
$$

In the examples below the target AHR is 0.70 (a moderate benefit).
Every subgroup — regardless of how it is defined — shares this same
underlying HR.  There is no subgroup effect; any apparent finding is
a false positive.

The calibration is performed in Part I via the helper function
`calibrate_k_treat()`, the natural sibling of the package's
`calibrate_k_inter()` (which targets the harm-subgroup HR under
`model = "alt"`) and `calibrate_cens_adjust()` (which targets the
censoring rate).  See the **Appendix: DGM with Heterogeneous Treatment
Effects** for a contrasting setup where one subgroup has a different
treatment effect than the rest of the population.

## Installation

The **forestsearch** package can be installed directly from GitHub using
either `pak` (recommended) or `remotes`.

### Stable development version (default branch)

```r
# Using pak (recommended)
# install.packages("pak")
pak::pak("larry-leon/forestsearch")

# Using remotes
# install.packages("remotes")
remotes::install_github("larry-leon/forestsearch")
```

### `feature/glm-extension` branch (optional)

As of **May 19, 2026**, the most up-to-date functionality is available on
the **`feature/glm-extension`** branch.  This branch extends the survival
simulation framework to binary, continuous, and count outcomes (logistic,
Gaussian, and Poisson GLMs) and includes additional functions such as
`generate_glm_dgm()`, `simulate_from_glm_dgm()`,
`calibrate_glm_interaction()`, and GLM dispatch in
`run_simulation_analysis()`.  Users who want access to the latest features
can install from this branch:

```r
# Using pak (recommended)
pak::pak("larry-leon/forestsearch@feature/glm-extension")

# Using remotes
remotes::install_github("larry-leon/forestsearch",
                        ref = "feature/glm-extension")
```

To verify the branch is loaded after installation:
```r
packageVersion("forestsearch")
# Confirm that GLM functions are exported:
exists("generate_glm_dgm", where = asNamespace("forestsearch"))
```

The package depends on
[**weightedsurv**](https://github.com/larry-leon/weightedsurv), which
is installed automatically from GitHub via the `Remotes` field in
`DESCRIPTION` — no separate step is needed.

Once installed, load the package in the usual way:
```r
library(forestsearch)
```

## Documentation

Full documentation, including function reference and vignettes, is
available at the **pkgdown** site:

> <https://larry-leon.github.io/forestsearch/>

The site includes, **along with this vignette**:

- **Getting Started** — end-to-end walkthrough with the GBSG breast
  cancer trial (`vignette("forestsearch")`)
- **Methodology** — statistical framework, algorithm details, and
  simulation results (`vignette("methodology")`)
- **Treatment Effect Definitions** — marginal HR, average hazard ratio
  (AHR), and controlled direct effect (CDE) estimands
- **Simulation Studies** — operating characteristics and published
  benchmarks
- **Reference** — all exported functions organised across 16 topic
  areas

Vignettes can also be accessed locally after installation:
```r
vignette("forestsearch")   # Getting Started
vignette("methodology")    # Statistical methodology
```



---

# Setup

```{r packages}
library(forestsearch)
library(survival)
library(ggplot2)
library(dplyr)
library(patchwork)     # assembles gg_forest() panels

theme_set(theme_minimal(base_size = 12))
```

```{r data-prep}
# Load GBSG data and convert time to months (DGM time scale)
data(gbsg, package = "survival")

gbsg$time_months <- gbsg$rfstime / 30.4375
gbsg$treat        <- gbsg$hormon

# Kaplan–Meier median event time.  The naive median over events only —
# median(time[event == 1]) — is biased downward in the presence of
# censoring, since the KM curve accounts for the at-risk set at each
# event time.  If the KM curve does not cross 0.5 the median is
# returned as NA (event time "not reached").
km_overall <- survfit(Surv(time_months, status) ~ 1, data = gbsg)
km_median  <- summary(km_overall)$table[["median"]]

cat("N =", nrow(gbsg), "  Events:", sum(gbsg$status),
    sprintf("(%.0f%%)", 100 * mean(gbsg$status)), "\n")
cat("Median follow-up, (months):",
    round(median(gbsg$time_months), 1), "\n")
cat("Median event time, KM (months)  :",
    if (is.na(km_median)) "not reached" else round(km_median, 1), "\n")
```

## Baseline characteristics

The table below summarises the cohort by treatment arm for the
variables used later to define subgroups in Part IV: the continuous
covariates (`age`, `size`, `nodes`, `pgr`, `er`), tumour `grade`, and
menopausal status (`meno`).  This is the source data — the
super-population in the DGM is sampled from it, so any baseline
imbalance carries through.

```{r baseline-table}
create_summary_table(
  data             = gbsg,
  treat_var        = "treat",
  vars_continuous  = c("age", "size", "nodes", "pgr", "er"),
  vars_categorical = c("grade"),
  vars_binary      = c("meno"),
  var_labels       = list(
    age   = "Age (years)",
    size  = "Tumour size (mm)",
    nodes = "Positive lymph nodes",
    pgr   = "Progesterone receptor (fmol)",
    er    = "Estrogen receptor (fmol)",
    grade = "Tumour grade (1 / 2 / 3)",
    meno  = "Post-menopausal"
  ),
  digits         = 1,
  show_pvalue    = TRUE,
  show_smd       = TRUE,
  show_missing   = TRUE,
  table_title    = "GBSG baseline characteristics — variables used to define subgroups",
  table_subtitle = paste0("Source: survival::gbsg  |  N = ", nrow(gbsg)),
  compact_mode   = TRUE
)
```

---

# Part I — Building the Uniform-Effect Data Generating Mechanism

To evaluate extreme subgroup effects by chance, we need a DGM where the
treatment has a genuine, uniform benefit — so that any subgroup finding is
definitionally a false positive.  This section constructs that DGM from the
GBSG data using `generate_aft_dgm_flex()` with `model = "null"` (uniform
treatment effect throughout; no embedded harm subgroup).

## How `generate_aft_dgm_flex()` works

The function fits **two parametric models** to the source data — one
for the event times, one for the censoring times — and uses them
jointly to build a "super population" from which all later synthetic
trials are drawn.

### Outcome model — Weibull AFT

A Weibull accelerated failure time (AFT) model is fitted to the
observed event times via `survival::survreg()`:

$$
\log T_i = \mu + \mathbf{x}_i^\top \boldsymbol{\beta} + \tau \, \varepsilon_i,
\qquad \varepsilon_i \sim \text{Extreme Value}.
$$

(Standard *minimum* extreme-value distribution with density
$f_\varepsilon(x) = \exp(x - e^x)$ — generated in the package via
`log(rexp(n_super))`.  See the companion vignettes
**treatment_effect_definitions.Rmd** and **biomarker_effects.Rmd** for
the AFT–Cox parameterisation linkage.)

The linear predictor includes treatment, the user-specified
`continuous_vars` and `factor_vars`, and — when `model = "alt"` — a
treatment-by-subgroup interaction term.  The treatment coefficient
and the interaction coefficient can be rescaled before sampling via
the `k_treat` and `k_inter` arguments, allowing the marginal
treatment effect and HTE strength to be set independently of the
fitted values.  Under `model = "null"` the interaction is removed
entirely, producing a uniform treatment effect across all patients.

### Censoring model — AFT with AIC selection (default)

With `select_censoring = TRUE` (the default), the censoring
distribution is fitted *symmetrically* by reversing the event
indicator: censored observations become the "event" of interest in
the reversed survival problem.  The function fits **four candidate
AFT models** to these reversed data and picks the best by AIC:

| Candidate    | Distribution | Covariates                       |
|:-------------|:-------------|:---------------------------------|
| `Weibull`    | Weibull      | `treat` + censoring covariates   |
| `LogNormal`  | Log-normal   | `treat` + censoring covariates   |
| `Weibull0`   | Weibull      | intercept only                   |
| `LogNormal0` | Log-normal   | intercept only                   |

The censoring covariates default to the outcome model's covariates,
which can be overridden via `continuous_vars_cens` / `factor_vars_cens`.
The intercept-only candidates (`*0`) guard against overfitting when
the censoring process is genuinely unrelated to baseline covariates —
in that case AIC will favour them.  The selected model's parameters
are stored in `dgm$model_params$censoring` (`mu`, `tau`, `gamma`, and
`type`).

Alternative censoring-model modes (set `select_censoring = FALSE`):

- **Analytical**: supply `cens_params$mu` and `cens_params$tau`
  directly — no model is fitted.
- **Force-fit**: leave `cens_params` empty — a single AFT model is
  fitted with the user-specified family (`cens_type`) and formula
  (`cens_intercept_only` toggles between intercept-only and
  covariate-adjusted).

### Super-population construction

Once both models are in hand, the function draws `n_super` rows from
the source data with replacement, attaches per-row linear predictors
for both outcome and censoring under each treatment counterfactual,
and returns the result as `dgm$df_super`.  Downstream,
`simulate_from_dgm()` samples trials from this super population
without re-fitting anything — each call is a fast row-sampling step
followed by quantile-transform draws from the fitted distributions.

### Marginal censoring rate is not yet calibrated

Even with the AIC-selected censoring model, the interaction between
the fitted censoring distribution and the administrative cutoff
`analysis_time` shifts the *realised* censoring rate in simulated
trials away from the observed rate in the source data.  Part II
addresses this directly: `check_censoring_dgm()` diagnoses the
discrepancy, and `calibrate_cens_adjust()` finds a log-scale shift
(`cens_adjust`) that aligns the simulated censoring rate (or KM
median censoring time) with the observed value.

## Why time-scale consistency matters first

Before fitting anything, confirm that all time parameters are expressed in
the same units. The single most common simulation failure is building the
DGM on days (e.g. `rfstime`) and then passing `analysis_time` in months.
The result is a follow-up window far shorter than the DGM event-time
distribution, producing universal administrative censoring (`event_sim = 0`
for every subject).

After building the DGM, this quick check catches the problem:

```{r timescale-check}
# exp(mu) is the implied median event time on the DGM time scale.
# It should be close to the KM-based observed median event time.
# We use the KM estimator (computed in Section 2 as km_median) rather
# than the naive median over events, which is biased by censoring.
cat("Observed median event time, KM (months):",
    if (is.na(km_median)) "not reached" else round(km_median, 1), "\n")
cat("(After building dgm_uniform, exp(dgm_uniform$model_params$mu) should be similar.)\n")
```

## Fitting `generate_aft_dgm_flex()`

`generate_aft_dgm_flex()` fits a Weibull AFT model to the source data and
generates a large *super population* from which all future trials will be
drawn.  For the extreme-subgroup evaluation we use `model = "null"`, which
sets the treatment effect to be identical for every patient — there is no
subgroup, so any apparent finding is a false
positive by construction.

The censoring model is estimated from the data by AIC-based selection across
Weibull and log-normal models, with and without covariate adjustment.  The
next section shows how to verify and calibrate this so that simulated trials
reproduce the censoring pattern of the source data.

**Note.** The DGM built here has a uniform treatment effect — no embedded subgroup. For a contrasting example with a true HTE (a harm subgroup defined by ER-low & pre-menopausal status), see the **Appendix: DGM with Heterogeneous Treatment Effects** at the end of this document.

### Calibrating `k_treat` to a target AHR

`generate_aft_dgm_flex()` fits a Weibull AFT to the source data and uses
the *fitted* treatment coefficient as the simulation truth.  With
`k_treat = 1` (default), the resulting hazard ratio summaries are
whatever the GBSG fit happens to produce — not exactly-specified targets.

To set the true treatment effect to **exactly HR = 0.70**, we rescale
the fitted treatment coefficient via `k_treat`, with the multiplier
chosen by `uniroot` so that a target summary of the DGM hits 0.70.  The
package function `calibrate_k_treat()` performs this calibration; it is
the natural sibling of `calibrate_k_inter()` (which calibrates the
harm-subgroup HR under `model = "alt"`) and `calibrate_cens_adjust()`
(which calibrates the censoring rate).  All three follow the same
`uniroot`-based pattern.

**Which target?** `calibrate_k_treat()` can target either of two HR
summaries computed on the super-population of potential outcomes:

- `dgm$hazard_ratios$overall` — the **marginal Cox HR** from a
  no-covariate Cox fit on stacked potential outcomes (default,
  `use_ahr = FALSE`)
- `dgm$hazard_ratios$AHR` — the **average hazard ratio**,
  $\exp(\overline{\log\text{HR}_{\text{p.o.}}})$ across the
  super-population (`use_ahr = TRUE`)

Under `model = "null"` (uniform effect) the AHR equals the
patient-level individual HR — the same number for every subject.  The
**marginal Cox HR** is attenuated toward 1 relative to the AHR because
the Cox model is non-collapsible: a no-covariate fit on heterogeneous
covariate distributions does not recover the patient-level HR.

The simulation analyses below fit Cox models *stratified by `grade`*,
which is closer to a covariate-adjusted fit than to a marginal fit.
Under proportional hazards (which the Weibull AFT satisfies), a
sufficiently-adjusted Cox model estimates the patient-level HR — i.e.,
the AHR — not the attenuated marginal HR.  We therefore calibrate
against the AHR (`use_ahr = TRUE`) so that the median across simulations
of the stratified Cox HR estimate tracks 0.70 closely.  Calibrating
against `hazard_ratios$overall` instead would force the marginal Cox
HR to 0.70 and pull the simulation median below 0.70 by the same
non-collapsibility shift.

See `?calibrate_k_treat` for full documentation.

### Build the DGM (calibrated `k_treat`)

The full argument list to `generate_aft_dgm_flex()` is laid out as a
single named list `base_args_uniform`.  This list is then used twice —
once inside `calibrate_k_treat()` to find the right `k_treat`, and once
to build the final calibrated DGM — so every knob of the constructor
appears exactly once.

```{r build-dgm}
.t_dgm_start <- Sys.time()

# ── All arguments to generate_aft_dgm_flex() except k_treat ───────────────
# k_treat is the calibration parameter; everything else is fixed up front.
base_args_uniform <- list(
  # ── Required: source data & outcome ─────────────────────────────────────
  data                = gbsg,
  continuous_vars     = c("age", "size", "nodes", "pgr", "er"),
  factor_vars         = c("meno", "grade"),
  outcome_var         = "time_months",
  event_var           = "status",
  # ── Treatment ───────────────────────────────────────────────────────────
  treatment_var       = "hormon",
  draw_treatment      = FALSE,                                  # default
  # ── HTE specification — uniform treatment effect, no embedded subgroup ──
  model               = "null",                                 # uniform HR throughout
  subgroup_vars       = NULL,                                   # default — no embedded HTE
  subgroup_cuts       = NULL,                                   # default
  k_inter             = 1,                                      # default (irrelevant under "null")
  spline_spec         = NULL,                                   # default
  set_beta_spec       = list(set_var = NULL, beta_var = NULL),  # default
  # ── Censoring model (independent of treatment effect) ───────────────────
  continuous_vars_cens = NULL,                                  # default → inherit
  factor_vars_cens     = NULL,                                  # default → inherit
  select_censoring     = TRUE,                                  # default — AIC select
  cens_type            = "weibull",                             # default
  cens_params          = list(),                                # default
  cens_intercept_only  = FALSE,                                 # default
  # ── Super population & misc ─────────────────────────────────────────────
  n_super             = 5000,
  standardize         = FALSE,                                  # default
  seed                = 99,
  verbose             = FALSE
)

# ── Step 1: calibrate k_treat so that the AHR is exactly 0.70 ─────────────
# We calibrate to the AHR (average hazard ratio over the super-population),
# not the marginal Cox HR, because the simulation analyses below fit a Cox
# model stratified by `grade` — and under Cox non-collapsibility the
# marginal Cox HR on fully-observed potential outcomes is attenuated toward
# 1 relative to both the AHR and the stratified Cox HR.  Calibrating to the
# AHR makes the simulation median HR track 0.70 closely; calibrating to the
# marginal Cox HR (`use_ahr = FALSE`, the default) would pull the simulation
# median below 0.70.  See `?calibrate_k_treat` for both options.
k_treat_uniform <- calibrate_k_treat(
  target_hr_overall = 0.70,
  base_args         = base_args_uniform,
  k_treat_range     = c(-5, 5),
  tol               = 1e-6,
  use_ahr           = TRUE,
  verbose           = TRUE
)

# ── Step 2: build the DGM with the calibrated k_treat ─────────────────────
dgm_uniform <- do.call(
  generate_aft_dgm_flex,
  c(base_args_uniform, list(k_treat = k_treat_uniform))
)

t_dgm <- as.numeric(difftime(Sys.time(), .t_dgm_start, units = "secs"))

cat("Uniform DGM (AHR-calibrated)\n",
    "  k_treat calibrated to       : ", round(k_treat_uniform, 4), "\n",
    "  AHR (target 0.70)           : ",
    round(dgm_uniform$hazard_ratios$AHR, 4), "\n",
    "  marginal Cox HR (attenuated): ",
    round(dgm_uniform$hazard_ratios$overall, 4), "\n",
    "  implied median event T      : ",
    round(exp(dgm_uniform$model_params$mu), 1), " months\n", sep = "")
```

```{r dgm-summary}
# Key DGM outputs.
#
# Terminology note: dgm$hazard_ratios$overall is the *marginal* Cox HR
# from a no-covariate fit on stacked potential outcomes — not the
# "true HR" in the patient-level sense.  Under Cox non-collapsibility
# it is attenuated toward 1 relative to the AHR (which equals the
# patient-level individual HR under model = "null").  See Section 3.3.1
# and the companion vignette `treatment_effect_definitions.Rmd`.
cat("=== Uniform DGM Summary ===\n")
cat("Time scale          : months\n")
cat("Implied median T    : exp(mu) =",
    round(exp(dgm_uniform$model_params$mu), 1), "months\n")
cat("Marginal HR         :",
    round(dgm_uniform$hazard_ratios$overall, 3), "\n")
cat("Causal AHR          :",
    round(dgm_uniform$hazard_ratios$AHR, 3), "\n")
cat("Censoring type      :",
    dgm_uniform$model_params$censoring$type, "\n")
```

---

# Part II — Censoring Diagnostics and Calibration

For the simulation results to be credible, the censoring pattern in each
synthetic trial must match what was observed in the source data.  If the
simulated censoring rate is much higher or lower than the real one, the
event counts in small subgroups will be systematically wrong, biasing the
HR and UB(HR) distributions we are trying to characterise.

Even with a well-fitted censoring model, the interaction between the DGM
censoring distribution and administrative censoring (the `analysis_time`
cutoff) can shift the effective censoring rate substantially away from what
was observed.  This section covers two functions that diagnose and correct
for this shift before the main simulation loop is run.

## Simulation pipeline configuration

The simulation pipeline has a small number of parameters that govern every
per-trial draw: the patients per trial, the administrative analysis time,
the recruitment window, and the iteration-seed scheme. Defining them
**once**, in a single `sim_config` list, means the calibration step (below),
the single-trial demonstration in Part III, and the main parallel loop
in Part IV all use exactly the same values — no duplication, no risk of
drift. Change a field here and every downstream chunk reacts automatically.

```{r sim-config}
# ── Simulation pipeline configuration ──────────────────────────────────────
# Single source of truth for every per-trial parameter.  Every downstream
# call to simulate_from_dgm() and calibrate_cens_adjust() references this
# list — nothing is hard-coded twice.
sim_config <- list(
  n_per_trial      = nrow(gbsg),  # patients per simulated trial
  analysis_time    = 84,          # months — administrative censoring time
  max_entry        = 24,          # months — uniform recruitment window
  seed_base        = 0L,          # per-trial seed = seed_base + iter
  rand_seed_offset = 1e6L         # random-subgroup seed = seed + offset
)

cat("Per-trial simulation parameters\n",
    "  n per trial   :", sim_config$n_per_trial, "\n",
    "  analysis_time :", sim_config$analysis_time, "months\n",
    "  max_entry     :", sim_config$max_entry,     "months\n",
    "  seed scheme   : iter seed = ", sim_config$seed_base,
    " + ss\n", sep = "")
```

## Diagnosing censoring consistency

`check_censoring_dgm()` compares three censoring summaries between the DGM
reference data (`dgm$df_super`) and a simulated dataset: overall censoring
rate, censoring-time quantiles among censored subjects, and the KM-based
median censoring time (estimated by reversing the event indicator so that
censored observations become the "event" of interest).

```{r pilot-sim}
# Generate a pilot simulated trial (no calibration yet)
sim_pilot <- simulate_from_dgm(
  dgm            = dgm_uniform,
  n              = 1500,
  rand_ratio     = 1,                          # default (1:1 randomisation)
  entry_var      = NULL,                       # default (synthetic entry times)
  max_entry      = sim_config$max_entry,
  analysis_time  = sim_config$analysis_time,
  cens_adjust    = 0,                          # pilot: no calibration shift
  draw_treatment = TRUE,                       # default
  seed           = 1,
  strata_rand    = NULL,                       # default
  hrz_crit       = NULL,                       # default
  keep_rand      = FALSE,                      # default
  time_eos       = NULL                        # default
)

cat("Pilot trial:  N =", nrow(sim_pilot),
    "  Event rate:", round(mean(sim_pilot$event_sim), 3), "\n")
```

```{r check-censoring}
diag0 <- check_censoring_dgm(
  sim_data   = sim_pilot,
  dgm        = dgm_uniform,
  treat_var  = "treat_sim",   # default
  rate_tol   = 0.10,
  median_tol = 0.25,
  verbose    = TRUE
)
```

**Reading the output.** Flags warn when the absolute censoring-rate
discrepancy exceeds `rate_tol` (default 10 percentage points) or when the
KM median censoring times differ by more than `median_tol` (default 25%).
Common causes:

- **Time-scale mismatch** — check `exp(dgm$model_params$mu)` against
  `analysis_time`.
- **Short `analysis_time`** — administrative censoring dominates the
  censoring process, inflating the censoring rate and compressing
  censoring times.
- **Large `cens_adjust`** — the log-scale shift has moved the DGM
  censoring distribution far from the original data.

## Calibrating `cens_adjust`

`calibrate_cens_adjust()` uses `uniroot` to find the value of `cens_adjust`
— a log-scale shift applied to the DGM censoring model — that equates a
chosen summary statistic between the reference and simulated data. Two
targets are available: `"rate"` (overall censoring rate) and `"km_median"`
(KM-based median censoring time).

**How the objective function works.** At each candidate `cens_adjust`:

1. `simulate_from_dgm()` generates `n_eval` synthetic subjects.
2. `check_censoring_dgm()` extracts the target metric.
3. The objective returns `sim_metric - ref_metric`.

`uniroot` finds the zero crossing. The fixed `seed` inside the objective
ensures the same candidate is always evaluated identically, preventing the
root-finder from chasing Monte Carlo noise.

```{r calibrate}
.t_cal_start <- Sys.time()

cal_uniform <- calibrate_cens_adjust(
  dgm           = dgm_uniform,
  target        = "rate",                       # "rate" | "km_median"
  n             = 1000,
  rand_ratio    = 1,                            # default
  analysis_time = sim_config$analysis_time,
  max_entry     = sim_config$max_entry,
  seed          = 42,
  interval      = c(-3, 3),                     # default — uniroot search interval
  tol           = 1e-4,                         # default — uniroot tolerance
  n_eval        = 2000,
  verbose       = TRUE
)

t_cal <- as.numeric(difftime(Sys.time(), .t_cal_start, units = "secs"))
```

```{r calibrate-result}
cat("\nCalibrated cens_adjust :", round(cal_uniform$cens_adjust, 4), "\n")
cat("Reference censoring rate:", round(cal_uniform$ref_value,    3),  "\n")
cat("Achieved  censoring rate:", round(cal_uniform$sim_value,    3),  "\n")
cat("Residual                :", round(cal_uniform$residual,     4),  "\n")
```

**Monotonicity note.** The objective is monotone in `cens_adjust` for both
targets: larger values increase censoring times (lower rate, higher KM
median), smaller values do the opposite. If `uniroot` fails, the target
lies outside the `interval`; widen it and retry.

## Verifying the calibrated simulation

```{r verify-cal}
sim_cal <- simulate_from_dgm(
  dgm            = dgm_uniform,
  n              = 1000,
  rand_ratio     = 1,                          # default
  entry_var      = NULL,                       # default
  max_entry      = sim_config$max_entry,
  analysis_time  = sim_config$analysis_time,
  cens_adjust    = cal_uniform$cens_adjust,
  draw_treatment = TRUE,                       # default
  seed           = 123,
  strata_rand    = NULL,                       # default
  hrz_crit       = NULL,                       # default
  keep_rand      = FALSE,                      # default
  time_eos       = NULL                        # default
)

diag_cal <- check_censoring_dgm(
  sim_data   = sim_cal,
  dgm        = dgm_uniform,
  treat_var  = "treat_sim",   # default
  rate_tol   = 0.05,
  median_tol = 0.15,
  verbose    = TRUE
)
```

---

# Part III — Generating Synthetic Trials

With the uniform DGM built, diagnosed, and calibrated, each simulated trial
is a single call to `simulate_from_dgm()`.  The function samples from the
super population with replacement, assigns treatment, generates latent
survival times from the AFT model, applies the calibrated censoring model,
and truncates at `analysis_time`.

The single-trial example below illustrates the output structure.  In the
uniform DGM every patient has the same true HR of 0.70, so `flag_harm` is
always 0 — there is no embedded subgroup.  The `r n_sims_uniform` replications
that drive the extreme-subgroup analysis in Part IV follow the same pattern,
looping over subgroups inside each replicate.

```{r single-trial}
sim_data <- simulate_from_dgm(
  dgm            = dgm_uniform,
  n              = 600,
  rand_ratio     = 1,
  entry_var      = NULL,                       # default
  max_entry      = sim_config$max_entry,
  analysis_time  = sim_config$analysis_time,
  cens_adjust    = cal_uniform$cens_adjust,
  draw_treatment = TRUE,                       # default
  seed           = 100,
  strata_rand    = NULL,                       # default
  hrz_crit       = NULL,                       # default
  keep_rand      = FALSE,                      # default
  time_eos       = NULL                        # default
)

cat("Simulated trial:  N =", nrow(sim_data), "\n")
cat("Treatment allocation:",
    table(sim_data$treat_sim)[["0"]], "control /",
    table(sim_data$treat_sim)[["1"]], "experimental\n")
cat("Overall event rate :", round(mean(sim_data$event_sim), 3), "\n")
cat("flag_harm range    :", paste(range(sim_data$flag_harm), collapse = "–"),
    "  (always 0 under the uniform DGM)\n")
```

**Cox specification — marginal here, stratified in Part IV.**  The
single-trial sanity check below fits an *unstratified* Cox model
(`treat_sim` only).  This deliberately matches the no-covariate fit
behind `dgm_uniform$hazard_ratios$overall` (the marginal HR reported
above), so that the observed and DGM-implied HRs are directly
comparable.  The full simulation loop in Part IV uses a *stratified*
Cox model (`treat_sim + strata(grade)`), which under Cox
non-collapsibility recovers a quantity closer to the AHR / patient-level
true HR — i.e., 0.70 — and which reflects how the user's analysis plan
specifies the primary subgroup analysis.

```{r trial-cox}
# Fit overall Cox model on this single simulated trial
cox_all <- coxph(Surv(y_sim, event_sim) ~ treat_sim, data = sim_data)
hr_obs  <- exp(coef(cox_all))
ci_obs  <- exp(confint(cox_all))

cat("Single-trial Cox analysis (one realisation):\n")
cat("  Marginal HR (DGM) :",
    round(dgm_uniform$hazard_ratios$overall, 3), "\n")
cat("  Observed HR       :", round(hr_obs, 3), "\n")
cat("  95% CI            : (",
    round(ci_obs[1], 3), ", ", round(ci_obs[2], 3), ")\n", sep = "")
```


---


# Part IV — Main Results: Extreme Subgroup Effects Under a Uniform Treatment Benefit

## The Core Question

Parts I–III established how to build a uniform DGM and generate synthetic
trials from it.  We now use those tools of `forestsearch` to answer the
question this vignette is built around: **when many subgroups are
examined in a trial with a perfectly uniform treatment benefit, how
extreme can a standard Cox subgroup analysis look by chance alone?**

We generate `r n_sims_uniform` synthetic trials from `dgm_uniform` and
fit a Cox model (stratified by `grade`) in each pre-defined subgroup of
every trial.  For each subgroup we accumulate the per-trial HR estimate
and the upper 95% confidence bound.  The resulting empirical
distributions are summarised using the six statistics defined in
Section 1.3 (Summary Statistics Across Simulations) — `Pr(HR<0.5)`,
`Pr(HR>1.0)`, `mHR`, `Pr(UB>=2)`, `Pr(UB>=3)`, `mUB` — and visualised
as median + 99% ECI ranges in the forest panels.  As discussed in
Section 1.3, the 99% ECI describes empirical spread across simulated
trials; it is not a confidence interval in the inferential sense.

The `random*` subgroups (no clinical basis, same true HR as everyone
else) serve as calibration anchors: any clinical subgroup of the same N
has an identical chance distribution, so clinical rationale provides no
statistical protection against the extreme results that small samples
naturally produce.

---

## Setting: Uniform Treatment Effect (recap)

The DGM built in Part I (`dgm_uniform`) sets the true log hazard ratio
constant across all patients:

$$
\psi^0(\mathbf{L}) = \log(\mathrm{HR}_{\mathrm{true}})
  \quad \text{for all } \mathbf{L}.
$$

With true HR = 0.70, every subgroup — regardless of how it is
defined — shares this same underlying HR.  Any apparent subgroup finding
in the analysis below is, by construction, a false positive.

---

## Subgroups Under Study

Each simulated trial analyses the same set of pre-defined subgroups,
ranging from clinically motivated subpopulations to deliberately
artificial ones.

The **`random*` subgroups** are the key benchmark. They are constructed
by randomly sampling approximately 20% of subjects and taking the first
15, 20, 40, and 60 patients from that draw:

```{r define-uniform-subgroups}
set.seed(8316951)
gbsg$flag_itt <- 1L

# ── Scalar cut-points for continuous variables ────────────────────────────
# Computed once from the observed data; re-exposed as columns inside each
# simulated df_s so that subset expressions like "age <= age_med" resolve.
age_med  <- median(gbsg$age)          # ~53 yrs
size_med <- median(gbsg$size)         # ~25 mm
nodes_q3 <- quantile(gbsg$nodes, 0.75)  # top-quartile node count (~5)
pgr_med  <- median(gbsg$pgr)          # median PGR
er_cut   <- 20                         # ER-low clinical threshold (fmol / mg)

subgroups <- list(

  # ── A. Full trial (ITT) ─────────────────────────────────────────────────
  list(id = "flag_itt == 1",                       name = "All Patients",              grp = "ITT"),

  # ── B. Single clinical / factor subgroups ───────────────────────────────
  list(id = "meno == 1",                           name = "Post-menopausal",           grp = "Clinical"),
  list(id = "meno == 0",                           name = "Pre-menopausal",            grp = "Clinical"),
  list(id = "grade == 3",                          name = "Grade 3",                   grp = "Clinical"),
  list(id = "grade != 3",                          name = "Grade 1/2",                 grp = "Clinical"),

  # ── C. Single continuous-variable cuts ──────────────────────────────────
  list(id = "age <= age_med",                      name = "Age (young)",               grp = "Continuous"),
  list(id = "age > age_med",                       name = "Age (older)",               grp = "Continuous"),
  list(id = "age <= 50",                           name = "Age <= 50",                 grp = "Continuous"),
  list(id = "age > 50",                            name = "Age > 50",                  grp = "Continuous"),
  list(id = "size <= size_med",                    name = "Tumour size (small)",       grp = "Continuous"),
  list(id = "size > size_med",                     name = "Tumour size (large)",       grp = "Continuous"),
  list(id = "nodes == 0",                          name = "Node-negative",             grp = "Continuous"),
  list(id = "nodes > 0 & nodes <= 3",              name = "Nodes 1-3",                grp = "Continuous"),
  list(id = "nodes > nodes_q3",                    name = "High nodes (>Q3)",          grp = "Continuous"),
  list(id = "pgr <= pgr_med",                      name = "PGR low",                  grp = "Continuous"),
  list(id = "pgr > pgr_med",                       name = "PGR high",                 grp = "Continuous"),
  list(id = "er <= er_cut",                        name = "ER-low (<20)",              grp = "Continuous"),
  list(id = "er > er_cut",                         name = "ER-high (>=20)",            grp = "Continuous"),

  # ── D. Menopausal status x Grade (2x2) ──────────────────────────────────
  list(id = "meno == 0 & grade == 3",              name = "Pre-meno / Grade 3",        grp = "Interaction: meno x grade"),
  list(id = "meno == 1 & grade == 3",              name = "Post-meno / Grade 3",       grp = "Interaction: meno x grade"),
  list(id = "meno == 0 & grade != 3",              name = "Pre-meno / Grade 1-2",      grp = "Interaction: meno x grade"),
  list(id = "meno == 1 & grade != 3",              name = "Post-meno / Grade 1-2",     grp = "Interaction: meno x grade"),

  # ── E. Menopausal status x Age ───────────────────────────────────────────
  list(id = "meno == 0 & age <= 50",               name = "Pre-meno / Age<=50",        grp = "Interaction: meno x age"),
  list(id = "meno == 0 & age > 50",                name = "Pre-meno / Age>50",         grp = "Interaction: meno x age"),
  list(id = "meno == 1 & age <= age_med",          name = "Post-meno / Age (young)",   grp = "Interaction: meno x age"),
  list(id = "meno == 1 & age > age_med",           name = "Post-meno / Age (older)",   grp = "Interaction: meno x age"),

  # ── F. Menopausal status x ER status ────────────────────────────────────
  list(id = "meno == 0 & er <= er_cut",            name = "Pre-meno / ER-low",         grp = "Interaction: meno x ER"),
  list(id = "meno == 0 & er > er_cut",             name = "Pre-meno / ER-high",        grp = "Interaction: meno x ER"),
  list(id = "meno == 1 & er <= er_cut",            name = "Post-meno / ER-low",        grp = "Interaction: meno x ER"),
  list(id = "meno == 1 & er > er_cut",             name = "Post-meno / ER-high",       grp = "Interaction: meno x ER"),

  # ── G. Grade x Node status ───────────────────────────────────────────────
  list(id = "grade == 3 & nodes == 0",             name = "Grade 3 / Node-neg",        grp = "Interaction: grade x nodes"),
  list(id = "grade == 3 & nodes > 0",              name = "Grade 3 / Node-pos",        grp = "Interaction: grade x nodes"),
  list(id = "grade != 3 & nodes == 0",             name = "Grade 1-2 / Node-neg",      grp = "Interaction: grade x nodes"),
  list(id = "grade != 3 & nodes > 0",              name = "Grade 1-2 / Node-pos",      grp = "Interaction: grade x nodes"),
  list(id = "grade == 3 & nodes > nodes_q3",       name = "Grade 3 / High nodes",      grp = "Interaction: grade x nodes"),

  # ── H. Grade x ER status ─────────────────────────────────────────────────
  list(id = "grade == 3 & er <= er_cut",           name = "Grade 3 / ER-low",          grp = "Interaction: grade x ER"),
  list(id = "grade == 3 & er > er_cut",            name = "Grade 3 / ER-high",         grp = "Interaction: grade x ER"),
  list(id = "grade != 3 & er <= er_cut",           name = "Grade 1-2 / ER-low",        grp = "Interaction: grade x ER"),
  list(id = "grade != 3 & er > er_cut",            name = "Grade 1-2 / ER-high",       grp = "Interaction: grade x ER"),

  # ── I. Grade x PGR ──────────────────────────────────────────────────────
  list(id = "grade == 3 & pgr <= pgr_med",         name = "Grade 3 / PGR low",         grp = "Interaction: grade x PGR"),
  list(id = "grade == 3 & pgr > pgr_med",          name = "Grade 3 / PGR high",        grp = "Interaction: grade x PGR"),
  list(id = "grade != 3 & pgr <= pgr_med",         name = "Grade 1-2 / PGR low",       grp = "Interaction: grade x PGR"),
  list(id = "grade != 3 & pgr > pgr_med",          name = "Grade 1-2 / PGR high",      grp = "Interaction: grade x PGR"),

  # ── J. Age x ER x Menopausal (3-way) ───────────────────────────────────
  list(id = "meno == 0 & age <= 50 & er <= er_cut",  name = "Pre-meno/Yng/ER-low",    grp = "3-way"),
  list(id = "meno == 0 & age <= 50 & er > er_cut",   name = "Pre-meno/Yng/ER-high",   grp = "3-way"),
  list(id = "meno == 1 & grade == 3 & er <= er_cut", name = "Post-meno/G3/ER-low",    grp = "3-way"),
  list(id = "meno == 1 & grade == 3 & nodes > 0",    name = "Post-meno/G3/Node-pos",  grp = "3-way"),
  list(id = "grade == 3 & nodes > 0 & er <= er_cut", name = "G3/Node-pos/ER-low",     grp = "3-way"),

  # ── K. Size x Node combinations ─────────────────────────────────────────
  list(id = "size > size_med & nodes > 0",         name = "Large/Node-pos",            grp = "Interaction: size x nodes"),
  list(id = "size > size_med & nodes == 0",        name = "Large/Node-neg",            grp = "Interaction: size x nodes"),
  list(id = "size <= size_med & nodes > 0",        name = "Small/Node-pos",            grp = "Interaction: size x nodes"),
  list(id = "size <= size_med & nodes == 0",       name = "Small/Node-neg",            grp = "Interaction: size x nodes"),

  # ── L. Random benchmark subgroups ───────────────────────────────────────
  # No clinical meaning; regenerated each simulation. Calibrate variability
  # by size against the clinical and interaction subgroups above.
  list(id = "random60 == 1",                       name = "random60",                  grp = "Random (N~60)"),
  list(id = "random40 == 1",                       name = "random40",                  grp = "Random (N~40)"),
  list(id = "random20 == 1",                       name = "random20",                  grp = "Random (N~20)"),
  list(id = "random15 == 1",                       name = "random15",                  grp = "Random (N~15)")
)

# ── Show observed sizes from GBSG ────────────────────────────────────────
# Scalar cut-points exposed as columns so eval() resolves them correctly.
gbsg_eval           <- gbsg
gbsg_eval$age_med   <- age_med
gbsg_eval$size_med  <- size_med
gbsg_eval$nodes_q3  <- nodes_q3
gbsg_eval$pgr_med   <- pgr_med
gbsg_eval$er_cut    <- er_cut
gbsg_eval$random60  <- 0L
gbsg_eval$random40  <- 0L
gbsg_eval$random20  <- 0L
gbsg_eval$random15  <- 0L

sg_sizes <- sapply(subgroups, function(s) {
  tryCatch(
    sum(eval(parse(text = s$id), envir = gbsg_eval)),
    error = function(e) NA_integer_
  )
})
is_random_sg <- grepl("^random", sapply(subgroups, `[[`, "name"))
rand_sizes   <- c(60L, 40L, 20L, 15L)

sg_overview <- data.frame(
  Subgroup = sapply(subgroups, `[[`, "name"),
  Group    = sapply(subgroups, `[[`, "grp"),
  N_source = ifelse(is_random_sg, rand_sizes[cumsum(is_random_sg)], sg_sizes),
  stringsAsFactors = FALSE
)
print(sg_overview, row.names = FALSE)
cat("
Total subgroups:", nrow(sg_overview), "
")
```

These `random*` subgroups have no clinical meaning whatsoever — their
true HR equals the ITT HR of 0.70. Any extreme result they produce is
purely a function of their small size and sampling variability. Placing
them alongside clinical subgroups of similar size makes explicit how much
of the spread in those clinical findings is attributable to noise alone.

---

## Summary Statistics

For each subgroup the primary analysis — Cox model stratified by the
randomisation factor (`grade`) — is fitted in every simulated trial.
The accumulated per-trial HR estimates and 95% upper bounds are
summarised using the six statistics defined in Section 1.3 (Summary
Statistics Across Simulations): `Pr(HR<0.5)`, `Pr(HR>1.0)`, `mHR`,
`Pr(UB>=2)`, `Pr(UB>=3)`, and `mUB`.  The Results section below
presents these in tabular form and visualises the underlying
distributions as median + 99% ECI ranges in the forest panels.

---

## Simulation Loop

The DGM (`dgm_uniform`), calibrated censoring shift (`cal_uniform`), and
simulation pipeline configuration (`sim_config`) were all built in
Parts I–II.  The Part IV loop reuses them directly — there is no
re-build, no re-calibration, no duplication of parameters.

### Parallel backend

The simulation loop below is embarrassingly parallel — every iteration
is independent and uses iteration-local seeds derived from `sim_config`
(`seed_base + ss` for `simulate_from_dgm()`, and
`seed_base + ss + rand_seed_offset` for the random-subgroup re-draw) —
so bit-for-bit reproducibility is preserved regardless of execution
order.  We use `doFuture` with a `multisession` plan (portable across
Windows / macOS / Linux) and target **at least 90% of available cores**,
so a 5,000-sim run scales close to the number of physical cores
available.

```{r parallel-setup}
#| message: false

library(future)
library(doFuture)
library(foreach)

# Target at least 90% of available cores; ceiling() guarantees the
# 90% floor on machines where 0.9 * cores is non-integer.
.t_par_start  <- Sys.time()
n_cores_total <- future::availableCores()
n_workers     <- max(1L, ceiling(0.90 * n_cores_total))

future::plan(future::multisession, workers = n_workers)
t_par_setup   <- as.numeric(difftime(Sys.time(), .t_par_start, units = "secs"))

cat(sprintf(
  "Parallel plan: multisession with %d workers (%.0f%% of %d available cores).\n",
  n_workers, 100 * n_workers / n_cores_total, n_cores_total
))
```

```{r sim-helpers}
# ── Per-subgroup analysis function ────────────────────────────────────────
# subgroup_cox() builds the analysis function applied inside every
# subgroup of every simulated trial: a stratified Cox fit returning the
# HR point estimate and upper 95% bound from the first conf.int row,
# with c(NA, NA) on estimation failure.  The rest of the former inline
# machinery -- one synthetic trial via simulate_from_dgm(), flag_itt and
# cut-point columns, nested random-benchmark regeneration, and the
# per-subgroup subset/fit loop with its < 5 skip rule -- now lives in
# forestsearch::run_subgroup_sims(), verified bit-identical to this
# document's original loop by dev/accept_phase1_bitident.R.  Swap the
# formula, or pass any function(data) returning c(estimate, upper), for
# sensitivity analyses without touching the simulation machinery.
fit_uniform <- subgroup_cox(
  Surv(y_sim, event_sim) ~ treat_sim + strata(grade)
)
```

```{r uniform-sims}
.t_sims_start <- Sys.time()

# Bundle scalar cut-points into a single argument
cut_points <- list(age_med = age_med, size_med = size_med,
                    nodes_q3 = nodes_q3, pgr_med = pgr_med, er_cut = er_cut)

# ── Simulation study ──────────────────────────────────────────────────────
# One call replaces the former helper + %dofuture% loop.  Per-iteration
# seeding is explicit inside the runner (trial ss: seed_base + ss for the
# DGM draw, + rand_seed_offset for benchmark membership), so results are
# independent of the parallel schedule.  workers = NULL reuses the
# multisession plan established in the parallel-setup chunk above instead
# of spawning a fresh worker pool.
sims_uniform <- run_subgroup_sims(
  dgm              = dgm_uniform,
  subgroups        = subgroups,
  n_sims           = n_sims_uniform,
  fit              = fit_uniform,
  baseline         = "resample",
  n                = sim_config$n_per_trial,
  analysis_time    = sim_config$analysis_time,
  max_entry        = sim_config$max_entry,
  cens_adjust      = cal_uniform$cens_adjust,
  cutpoints        = cut_points,
  benchmarks       = benchmark_spec(),      # random15 in 20 in 40 in 60
  seed_base        = sim_config$seed_base,
  rand_seed_offset = sim_config$rand_seed_offset,
  workers          = NULL,
  hr_true          = 0.70,
  k_treat          = k_treat_uniform
)

t_sims <- as.numeric(difftime(Sys.time(), .t_sims_start, units = "secs"))

cat("Simulations completed:", n_sims_uniform,
    " |  Parallel workers used:", n_workers, "\n")
```

### Persisting the payload (optional)

When `save_payload = TRUE` (see the run controls in the setup chunk for
the on/off knob, directory, and naming conventions) the chunk below
writes the raw simulation matrices plus run metadata to a single RDS
file.  This is what the companion memo
`extreme_subgroups_design_comparison.qmd` consumes to recompute the
fixed-vs-random design comparison from raw results.  With
`save_payload = FALSE` the chunk is a no-op and nothing touches disk.

```{r save-payload}
if (isTRUE(save_payload)) {
  # The runner's return object IS the payload: field names and matrix
  # layout are byte-compatible with the former hand-built list (plus a
  # "subgroup_sims" class attribute and a richer sim_config), so the
  # design-comparison memo loads it unchanged.
  saveRDS(sims_uniform, payload_path)
  cat(sprintf("Payload written: %s  (%.1f MB, forestsearch %s)\n",
              payload_path, file.size(payload_path) / 1024^2,
              sims_uniform$forestsearch_version))
} else {
  cat("Payload persistence disabled (save_payload = FALSE); nothing written.\n")
}
```

---

## Results

### Summary table

```{r uniform-results-table}
# ── Section 6.6 statistics ────────────────────────────────────────────────
# summary() is the single package-level source of truth for every
# statistic in the results table and forest panels: ECI quantiles, tail
# probabilities, convergence counts, category mapping, panel membership,
# and the high-risk panel.  Definitions were lifted verbatim from this
# document's former inline chunk and verified character-identical in the
# Phase 1 acceptance gate.
S_uniform   <- summary(sims_uniform, hr_true = 0.70)
results_tbl <- S_uniform$results_tbl
print(results_tbl, row.names = FALSE)
```

**Reading the table.** The columns mirror the annotation columns of the
forest plots below. Every row shares the same true HR = 0.70, so `mHR`
(the unconditional median HR across simulations) should track 0.70
throughout — confirming the Cox model is unbiased. The tail-probability
columns are what move with subgroup size:

- **`Pr(HR<0.5)`** and **`Pr(HR>1.0)`** measure how often the point
  estimate alone looks extremely favorable or lands on the wrong side
  of no effect by chance — relevant when a forest-plot row is read as
  evidence of differential benefit or harm.
- **`Pr(UB>=2)`** and **`Pr(UB>=3)`** measure how often the upper 95%
  confidence bound crosses thresholds a safety reviewer would treat as
  alarming.

For `random15`, these probabilities are substantial — the small-N
subgroup routinely produces upper bounds far above 1.0 even though the
true HR is 0.70. For the ITT they are negligible. The forest plots
below show the full ECI structure visually; this table gives the
underlying numbers per subgroup.

---

### Forest plot: HR point estimate distributions

Each row shows one subgroup. The point is the median HR across
`r n_sims_uniform` simulations; the line spans the 1st–99th percentile ECI.
The red dashed line marks the true HR = 0.70; the grey dotted line marks
HR = 1.0. `random*` benchmarks (orange) are shown alongside same-sized
clinical subgroups (blue) so that the comparison is immediate.

```{r forest-theme-params}
# =============================================================================
# Forest Plot Sizing Parameters — edit here to resize all four panels
# =============================================================================
#
# HOW SIZING WORKS WITH gg_forest()
# ──────────────────────────────────
# gg_forest() uses ggplot2 + patchwork.  Unlike forestploter, there is NO
# hidden scaling.  Row height is simply:
#
#     row_height_in  =  fig.height  /  n_rows
#
# So to set a target row height (e.g. 0.45 in):
#     fig.height  =  n_rows  ×  0.45  +  overhead_in
#
# where overhead_in ≈ 1.5 in covers the title, x-axis, and caption.
# ─────────────────────────────────────────────────────────────────────────────

fp_row_height   <- 0.45   # in — target height per row (increase to spread rows)
fp_overhead_in  <- 1.5    # in — fixed overhead (title + x-axis + footnote)
fp_base_size    <- 14     # pt — ggplot2 base font size (controls ALL text sizes)
fp_point_size   <- 2.2    # relative size of the median point symbol
fp_line_size    <- 0.75   # line width of CI whiskers

# ── Panel membership, colours, and validity masks ─────────────────────────
# Category labels, the colour map, ok/ok_ub masks, and the single/combo
# index vectors now live on S_uniform (see summary()); forest_height()
# converts panel row counts to the fig.height values used below.
fh_single <- forest_height(S_uniform, "single", fp_row_height, fp_overhead_in)
fh_combo  <- forest_height(S_uniform, "combo",  fp_row_height, fp_overhead_in)

cat(sprintf(
  "Single panels: %d rows  ->  recommended fig.height = %.1f in\n",
  S_uniform$n_single, fh_single))
cat(sprintf(
  "Combo  panels: %d rows  ->  recommended fig.height = %.1f in\n",
  S_uniform$n_combo,  fh_combo))
```


**Panel A — HR distributions, single-variable subgroups.**
Point = median HR; whiskers = 1st–99th percentile ECI.
The true HR is 0.70 for every subgroup (red dashed line).

```{r hr-forest-single}
#| fig-height: !expr fh_single
#| fig-width: 14
#| fig-cap: "HR distribution — single-variable subgroups (log scale). Spread reflects sampling variability alone; all subgroups share the same true HR = 0.70."
plot(S_uniform, metric = "hr", panel = "single",
     base_size  = fp_base_size, point_size = fp_point_size,
     line_size  = fp_line_size)
```

**Panel B — HR distributions, combination and random-benchmark subgroups.**
Note how the ECI widens sharply as subgroup N falls below ~60; the
`random*` benchmarks (orange) define the noise floor for each size class.

```{r hr-forest-combo}
#| fig-height: !expr fh_combo
#| fig-width: 14
#| fig-cap: "HR distribution — combination subgroups and random benchmarks (log scale). Random benchmarks of matching size (orange) are the baseline for chance variability."
plot(S_uniform, metric = "hr", panel = "combo",
     base_size  = fp_base_size, point_size = fp_point_size,
     line_size  = fp_line_size)
```

---

### Forest plot: UB(HR) distributions

The upper confidence bound UB(HR) is the quantity a safety reviewer
examines when asking whether a subgroup demonstrates harm or lack of
benefit. The plots below show the distribution of UB(HR) across
simulations in the same ECI format. Values above 1.0 (grey dotted line)
represent trials where the confidence interval includes no treatment
effect. The right-hand annotation columns show: the proportion of trials
in which UB(HR) reaches each of two clinically meaningful thresholds
(`UB>=2`, `UB>=3`), and the unconditional median UB(HR) across all
trials (`mUB`).

**Panel A — UB(HR) distributions, single-variable subgroups.**

```{r ub-forest-single}
#| fig-height: !expr fh_single
#| fig-width: 14
#| fig-cap: "UB(HR) distribution — single-variable subgroups (log scale). Values right of HR = 1.0 (dotted line) represent trials failing to show significant benefit."
plot(S_uniform, metric = "ub", panel = "single",
     base_size  = fp_base_size, point_size = fp_point_size,
     line_size  = fp_line_size)
```

**Panel B — UB(HR) distributions, combination subgroups and random benchmarks.**
The `UB>=2` and `UB>=3` columns are the key safety quantities — they
show how often a randomly-defined subgroup's upper confidence bound
crosses thresholds that a reviewer would treat as concerning, even
though the true HR is 0.70 throughout.

```{r ub-forest-combo}
#| fig-height: !expr fh_combo
#| fig-width: 14
#| fig-cap: "UB(HR) distribution — combination and random-benchmark subgroups. The widening ECI for small subgroups is the central message: chance alone produces UB(HR) > 3-5 for N ~ 15-40."
plot(S_uniform, metric = "ub", panel = "combo",
     base_size  = fp_base_size, point_size = fp_point_size,
     line_size  = fp_line_size)
```

---

### Forest plot: high-risk subgroups (Pr(UB ≥ 2.0) ≥ 10%)

The panels above present *all* pre-defined subgroups grouped by category.
A complementary view is to filter to subgroups whose UB(HR) distribution
crosses 2.0 with non-negligible frequency — concretely, those where at
least 10% of simulated trials produce UB(HR) ≥ 2.0.  These are the
subgroups most likely to generate alarming-looking forest-plot rows by
chance alone.  **All Patients (ITT)** is always retained as a reference
anchor, regardless of whether it meets the filter.

Rows are ordered with the ITT at the top, then the remaining subgroups
in descending order of `Pr(UB>=2)` — so the most chance-prone subgroups
appear immediately below the ITT for direct comparison.

```{r ub-forest-highrisk-setup}
#| include: false

# High-risk membership (Pr(UB>=2) >= threshold, ITT anchored first) is
# computed by summary(); only the recommended figure height remains
# document-side.
fh_highrisk <- forest_height(S_uniform, "highrisk",
                              fp_row_height, fp_overhead_in)
cat(sprintf(
  "High-risk panel: %d subgroups (Pr(UB>=2) >= %.0f%%, ITT included)  ->  fig.height = %.1f in\n",
  S_uniform$highrisk$n, 100 * S_uniform$highrisk$threshold, fh_highrisk))
```

```{r ub-forest-highrisk}
#| fig-height: !expr fh_highrisk
#| fig-width: 14
#| fig-cap: "UB(HR) distribution — subgroups where at least 10% of simulated trials produce UB(HR) >= 2.0, with All Patients (ITT) for comparison.  All share the same true HR = 0.70; differences are entirely sampling variability."
plot(S_uniform, metric = "ub", panel = "highrisk",
     base_size  = fp_base_size, point_size = fp_point_size,
     line_size  = fp_line_size)
```

The contrast is the point of the plot: every row shares the same true
HR (0.70), yet the filtered subgroups produce UB(HR) values that are
substantively different from the ITT.  Because the filter is on a tail
probability rather than on any clinical attribute, small subgroups
populate this panel regardless of whether they were defined by a
biomarker, a clinical interaction, or pure random sampling.

The companion view below is the *HR point-estimate distribution* for
the same set of subgroups.  Where the UB(HR) panel above asks "how
alarming can the upper confidence bound look by chance?", the HR panel
asks "how misleading can the point estimate itself be?"  The two views
together complete the picture: small subgroups produce extreme results
in both tails.

```{r hr-forest-highrisk}
#| fig-height: !expr fh_highrisk
#| fig-width: 14
#| fig-cap: "HR distribution — same high-risk subgroups as the panel above (Pr(UB(HR) >= 2.0) >= 10%, plus ITT for comparison).  Every row shares the same true HR = 0.70 (red dashed); spread reflects sampling variability alone."

plot(S_uniform, metric = "hr", panel = "highrisk",
     base_size  = fp_base_size, point_size = fp_point_size,
     line_size  = fp_line_size)
```

Read together with the UB(HR) panel, the message is direct: the same
small subgroups that produce large upper bounds also routinely produce
point estimates that overstate (HR &lt; 0.5) or understate (HR &gt; 1.0)
the true effect.  None of this reflects genuine heterogeneity — the
DGM guarantees a uniform HR of 0.70 for every patient.

---


# Summary

| Step | Function | Purpose |
|:-----|:---------|:--------|
| 1 | — | Convert source data to consistent time scale |
| 2 | `generate_aft_dgm_flex()` | Fit AFT model; embed harm subgroup; build super population |
| 3 | `check_censoring_dgm()` | Verify simulated censoring matches source data |
| 4 | `calibrate_cens_adjust()` | Find `cens_adjust` that aligns censoring rate or KM median |
| 5 | `simulate_from_dgm()` | Draw individual synthetic trials from the DGM |
| 6 | `forestsearch()` + summaries | Apply method and accumulate operating characteristics |

The extreme-subgroup evaluation in Part IV answers a question: *under a null where every patient shares the same true HR,
how alarming can a standard Cox subgroup analysis look by chance?* The
`random*` subgroups anchor the answer — any clinical subgroup of similar
size has an identical chance distribution of HR and UB(HR) regardless of
clinical rationale. 
---

### Timing summary

Wall-clock time for each computational stage and a projection for larger
simulation budgets.  The simulation loop dominates and scales linearly
in `n_sims_uniform`; the DGM and censoring-calibration steps are one-time
costs that don't change with replicate count.

```{r timing-summary}
#| echo: false

# Per-trial breakdown and projections
t_total       <- t_dgm + t_cal + t_sims
t_per_trial   <- t_sims / n_sims_uniform
trials_per_s  <- if (t_per_trial > 0) 1 / t_per_trial else NA_real_

fmt_time <- function(s) {
  if (is.na(s) || !is.finite(s))      "-"
  else if (s < 1)                     sprintf("%.0f ms", s * 1000)
  else if (s < 60)                    sprintf("%.2f sec", s)
  else if (s < 3600)                  sprintf("%.1f min", s / 60)
  else                                sprintf("%.2f hr", s / 3600)
}

# Projected total time for a future budget = one-time costs + linear sim cost.
# Per-trial cost is the *parallel* per-trial cost (wall-clock / n_sims), so
# projections assume the same parallel plan is used.
proj_for <- function(n_target) fmt_time(t_dgm + t_cal + n_target * t_per_trial)

timing_df <- data.frame(
  Step = c(
    "DGM construction (generate_aft_dgm_flex)",
    "Censoring calibration (calibrate_cens_adjust)",
    sprintf("Simulation loop (%d trials \u00d7 %d subgroups, %d workers)",
            n_sims_uniform, length(subgroups), n_workers),
    "Total"
  ),
  Elapsed = c(fmt_time(t_dgm),
              fmt_time(t_cal),
              fmt_time(t_sims),
              fmt_time(t_total)),
  `Per trial` = c("\u2014", "\u2014",
                  fmt_time(t_per_trial), "\u2014"),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

knitr::kable(timing_df,
             align   = c("l", "r", "r"),
             caption = sprintf(
               "Timing for n_sims_uniform = %d, parallel across %d of %d cores.",
               n_sims_uniform, n_workers, n_cores_total))

cat(sprintf(
  "\nThroughput: %.1f trials/sec (wall-clock, %d workers)  |  Projected total time:  1,000 sims \u2192 %s    \u2502    5,000 sims \u2192 %s\n",
  trials_per_s, n_workers, proj_for(1000), proj_for(5000)
))
```


# Appendix: DGM with Heterogeneous Treatment Effects

The main vignette uses a **uniform-effect DGM** (`model = "null"`): every
patient shares the same true HR = 0.70, and any apparent subgroup finding
is a false positive.  This appendix illustrates the alternative case — a
DGM with **genuine heterogeneous treatment effects (HTE)**, where a
specific subgroup of patients experiences a different (smaller or
larger) treatment effect than the rest of the population.

`generate_aft_dgm_flex()` supports HTE via `model = "alt"` combined with
`subgroup_vars`, `subgroup_cuts`, and the interaction modifier `k_inter`.
The example below embeds a **harm subgroup** defined by ER-low (`er ≤ 20`)
and pre-menopausal (`meno == 0`) status, with the interaction strength
chosen so the within-harm-subgroup HR is materially attenuated relative
to the no-harm complement.

```{r hte-build-dgm}
set.seed(42)

dgm_hte <- generate_aft_dgm_flex(
  # ── Required: source data & outcome ─────────────────────────────────────
  data                = gbsg,
  continuous_vars     = c("age", "size", "nodes", "pgr", "er"),
  factor_vars         = c("meno", "grade"),
  outcome_var         = "time_months",
  event_var           = "status",
  # ── Treatment ───────────────────────────────────────────────────────────
  treatment_var       = "hormon",
  draw_treatment      = FALSE,                                  # default
  # ── HTE specification — embedded harm subgroup ──────────────────────────
  model               = "alt",                                  # alternative DGM (HTE)
  subgroup_vars       = c("er", "meno"),
  subgroup_cuts       = list(er = 20, meno = 0),                # ER-low & pre-menopausal
  k_treat             = 1,                                      # default
  k_inter             = 1.5,                                    # moderate interaction
  spline_spec         = NULL,                                   # default
  set_beta_spec       = list(set_var = NULL, beta_var = NULL),  # default
  # ── Censoring model ─────────────────────────────────────────────────────
  continuous_vars_cens = NULL,                                  # default → inherit
  factor_vars_cens     = NULL,                                  # default → inherit
  select_censoring     = TRUE,                                  # default — AIC select
  cens_type            = "weibull",                             # default
  cens_params          = list(),                                # default
  cens_intercept_only  = FALSE,                                 # default
  # ── Super population & misc ─────────────────────────────────────────────
  n_super             = 5000,
  standardize         = FALSE,                                  # default
  seed                = 42,
  verbose             = FALSE
)
```

```{r hte-summary}
# Terminology: dgm_hte$hazard_ratios$harm_subgroup,
# $no_harm_subgroup, and $overall are all *marginal* Cox HRs from
# no-covariate fits on the corresponding potential-outcomes subsets.
# Under Cox non-collapsibility these are attenuated toward 1 relative
# to the corresponding AHRs (which equal the patient-level individual
# HRs within each subgroup).  See Section 3.3.1.
cat("=== HTE DGM Summary ===\n")
cat("Time scale            : months\n")
cat("Implied median T      : exp(mu) =",
    round(exp(dgm_hte$model_params$mu), 1), "months\n")
cat("Subgroup proportion   :", round(dgm_hte$subgroup_info$proportion, 3), "\n")
cat("Marginal HR (harm)    :",
    round(dgm_hte$hazard_ratios$harm_subgroup, 3), "\n")
cat("Marginal HR (no harm) :",
    round(dgm_hte$hazard_ratios$no_harm_subgroup, 3), "\n")
cat("Marginal HR (overall) :",
    round(dgm_hte$hazard_ratios$overall, 3), "\n")
cat("Censoring type        :",
    dgm_hte$model_params$censoring$type, "\n")
```

## Calibration and a single trial

Calibration follows the same procedure as in Part II — only the DGM
object changes.  We reuse `sim_config` so the per-trial pipeline
parameters (`analysis_time`, `max_entry`) are identical to the main
analysis.

```{r hte-calibrate}
cal_hte <- calibrate_cens_adjust(
  dgm           = dgm_hte,
  target        = "rate",
  n             = 1000,
  rand_ratio    = 1,                            # default
  analysis_time = sim_config$analysis_time,
  max_entry     = sim_config$max_entry,
  seed          = 42,
  interval      = c(-3, 3),                     # default
  tol           = 1e-4,                         # default
  n_eval        = 2000,
  verbose       = FALSE
)

cat("Calibrated cens_adjust :", round(cal_hte$cens_adjust, 4), "\n")
cat("Achieved censoring rate:", round(cal_hte$sim_value,   3), "\n")
```

```{r hte-single-trial}
sim_hte <- simulate_from_dgm(
  dgm            = dgm_hte,
  n              = 600,
  rand_ratio     = 1,
  entry_var      = NULL,                       # default
  max_entry      = sim_config$max_entry,
  analysis_time  = sim_config$analysis_time,
  cens_adjust    = cal_hte$cens_adjust,
  draw_treatment = TRUE,                       # default
  seed           = 100,
  strata_rand    = NULL,                       # default
  hrz_crit       = NULL,                       # default
  keep_rand      = FALSE,                      # default
  time_eos       = NULL                        # default
)

cat("Simulated HTE trial:  N =", nrow(sim_hte), "\n")
cat("Harm-subgroup size :", sum(sim_hte$flag_harm == 1),
    " (", round(100 * mean(sim_hte$flag_harm), 1), "% of trial)\n", sep = "")
cat("Overall event rate :", round(mean(sim_hte$event_sim), 3), "\n")
cat("Event rate (harm subgroup):",
    round(mean(sim_hte$event_sim[sim_hte$flag_harm == 1]), 3), "\n")
cat("Event rate (no-harm)      :",
    round(mean(sim_hte$event_sim[sim_hte$flag_harm == 0]), 3), "\n")
```

## True vs observed HR by subgroup

Fitting unstratified Cox models within each true subgroup recovers the
embedded heterogeneity — the harm subgroup's observed HR is materially
higher than the no-harm complement's.

```{r hte-results}
cox_all_hte    <- coxph(Surv(y_sim, event_sim) ~ treat_sim, data = sim_hte)
cox_harm_hte   <- coxph(Surv(y_sim, event_sim) ~ treat_sim,
                        data = subset(sim_hte, flag_harm == 1))
cox_noharm_hte <- coxph(Surv(y_sim, event_sim) ~ treat_sim,
                        data = subset(sim_hte, flag_harm == 0))

results_hte <- data.frame(
  Subgroup    = c("ITT", "Harm (true)", "No-harm (true)"),
  N           = c(nrow(sim_hte),
                  sum(sim_hte$flag_harm == 1),
                  sum(sim_hte$flag_harm == 0)),
  Marginal_HR = round(c(dgm_hte$hazard_ratios$overall,
                        dgm_hte$hazard_ratios$harm_subgroup,
                        dgm_hte$hazard_ratios$no_harm_subgroup), 3),
  Observed_HR = round(exp(c(coef(cox_all_hte),
                            coef(cox_harm_hte),
                            coef(cox_noharm_hte))), 3)
)
print(results_hte, row.names = FALSE)
```

## Kaplan–Meier curves by subgroup

```{r hte-km}
#| fig-cap: "Kaplan–Meier curves for one simulated HTE trial. Harm-subgroup curves diverge less between arms than the no-harm complement, reflecting the embedded HTE."
#| fig-width: 9
#| fig-height: 4

op <- par(mfrow = c(1, 3), mar = c(4, 4, 3, 1))

plot(survfit(Surv(y_sim, event_sim) ~ treat_sim, data = sim_hte),
     col = c("black", "red"), lwd = 2,
     xlab = "Time (months)", ylab = "Survival",
     main = "ITT")
plot(survfit(Surv(y_sim, event_sim) ~ treat_sim,
             data = subset(sim_hte, flag_harm == 1)),
     col = c("black", "red"), lwd = 2,
     xlab = "Time (months)", ylab = "Survival",
     main = "Harm subgroup")
plot(survfit(Surv(y_sim, event_sim) ~ treat_sim,
             data = subset(sim_hte, flag_harm == 0)),
     col = c("black", "red"), lwd = 2,
     xlab = "Time (months)", ylab = "Survival",
     main = "No-harm complement")

par(op)
```

This appendix DGM is a useful target for evaluating subgroup-discovery
methods (such as `forestsearch()`): we know the true harm subgroup by
construction and can measure whether the method recovers it.

---




# Session Information

```{r session-info}
sessionInfo()
```
