---
title: "Fixed vs Random Baseline: A Design Comparison"
subtitle: "Extreme Subgroup Effects Under a Uniform Treatment Benefit — conditional-on-X vs population-resampling simulation"
author: "Larry F. León"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-tools: true
    theme: cosmo
    self-contained: true
execute:
  echo: false
  warning: false
  message: false
params:
  use_payloads: null
  payload_dir: null
  payload_random: null
  payload_fixed: null
---

```{r setup}
#| include: false
library(gt)
library(dplyr)

# =========================================================================
# SOURCE-OF-NUMBERS KNOB -- live payloads vs embedded static transcription
# =========================================================================
# use_payloads = TRUE (default) -> load the two RDS payloads written by
#     the companion vignettes with save_payload = TRUE
#     (extreme_subgroups.qmd -> design "resample";
#      extreme_subgroups_fixed.qmd -> design "fixed") and RECOMPUTE every
#     number in this memo from the raw simulation matrices.  If either
#     file is missing the render STOPS with fix-it instructions -- there
#     is no silent fallback to stale numbers.
# use_payloads = FALSE -> render entirely from the static `cmp` frame
#     embedded below (transcribed from the original 10,000-simulation
#     reference renders).  Zero external dependencies.
#
# Override at render time without editing this file:
#   quarto render extreme_subgroups_design_comparison.qmd -P use_payloads:false
#   quarto render extreme_subgroups_design_comparison.qmd \
#     -P payload_random:results/extreme_sims_resample_50_payload.rds \
#     -P payload_fixed:results/extreme_sims_fixed_50_payload.rds
#
# payload_dir feeds only the two DEFAULT file names below; explicit
# payload_random / payload_fixed paths always win.  Relative paths
# resolve against this file's folder at render time.
.param_or <- function(name, default) {
  if (exists("params", inherits = TRUE) && !is.null(params[[name]]))
    params[[name]] else default
}
use_payloads   <- as.logical(.param_or("use_payloads", TRUE))
payload_dir    <- .param_or("payload_dir", "results")
payload_random <- .param_or("payload_random",
  file.path(payload_dir, "extreme_sims_resample_10000_payload.rds"))
payload_fixed  <- .param_or("payload_fixed",
  file.path(payload_dir, "extreme_sims_fixed_10000_payload.rds"))

if (isTRUE(use_payloads)) {
  # -- LIVE MODE: load payloads, validate, recompute the comparison ------
  miss <- c(payload_random, payload_fixed)
  miss <- miss[!file.exists(miss)]
  if (length(miss) > 0) {
    stop("use_payloads = TRUE but payload file(s) not found:\n  ",
         paste(miss, collapse = "\n  "), "\n",
         "Fix one of:\n",
         "  (a) render both vignettes first (their default save_payload = TRUE\n",
         "      writes results/extreme_sims_<design>_<n_sims>_payload.rds)\n",
         "  (b) point at existing files: -P payload_random:<path> -P payload_fixed:<path>\n",
         "  (c) render the static memo:  -P use_payloads:false")
  }
  pl_r <- readRDS(payload_random)   # random-X  (design "resample")
  pl_f <- readRDS(payload_fixed)    # fixed-X   (design "fixed")

  # Guard 1: payloads must come from the intended designs.
  if (!identical(pl_r$design, "resample") || !identical(pl_f$design, "fixed")) {
    stop("Design labels are not ('resample', 'fixed'): got ('",
         pl_r$design, "', '", pl_f$design, "').  Check the payload paths.")
  }
  # Guard 2: identical subgroup panels, or the row-wise comparison below
  # would be silently misaligned.
  if (!identical(colnames(pl_r$sim_hrs), colnames(pl_f$sim_hrs))) {
    stop("Subgroup names differ between the two payloads; re-run both ",
         "vignettes from the same subgroup definitions.")
  }

  # Per-design summary statistics.  Definitions replicate the vignettes'
  # Section 6.6.1 results table exactly: tail probabilities use
  # na.rm = TRUE (denominated over converged Cox fits), medians are
  # unconditional, N is the across-trial mean subgroup size.
  # Structurally empty subgroups (all-NA columns) come out NaN and are
  # mapped to NA to match the static frame's convention.
  .col_stats <- function(pl) {
    nan_to_na <- function(x) { x[is.nan(x)] <- NA_real_; x }
    data.frame(
      subgroup = colnames(pl$sim_hrs),
      N    = nan_to_na(round(colMeans(pl$sim_ns, na.rm = TRUE))),
      ub2  = nan_to_na(100 * apply(pl$sim_ubs, 2, function(x) mean(x >= 2.0, na.rm = TRUE))),
      ub3  = nan_to_na(100 * apply(pl$sim_ubs, 2, function(x) mean(x >= 3.0, na.rm = TRUE))),
      mUB  = suppressWarnings(apply(pl$sim_ubs, 2, median, na.rm = TRUE)),
      hr05 = nan_to_na(100 * apply(pl$sim_hrs, 2, function(x) mean(x < 0.5, na.rm = TRUE))),
      hr1  = nan_to_na(100 * apply(pl$sim_hrs, 2, function(x) mean(x > 1.0, na.rm = TRUE))),
      mHR  = suppressWarnings(apply(pl$sim_hrs, 2, median, na.rm = TRUE)),
      stringsAsFactors = FALSE
    )
  }
  .st_r <- .col_stats(pl_r)
  .st_f <- .col_stats(pl_f)

  cmp <- data.frame(
    subgroup = .st_r$subgroup,
    N_r    = .st_r$N,    N_f    = .st_f$N,
    ub2_r  = .st_r$ub2,  ub2_f  = .st_f$ub2,
    ub3_r  = .st_r$ub3,  ub3_f  = .st_f$ub3,
    mUB_r  = .st_r$mUB,  mUB_f  = .st_f$mUB,
    hr05_r = .st_r$hr05, hr05_f = .st_f$hr05,
    hr1_r  = .st_r$hr1,  hr1_f  = .st_f$hr1,
    mHR_r  = .st_r$mHR,  mHR_f  = .st_f$mHR,
    stringsAsFactors = FALSE
  )

  n_sims_r <- pl_r$n_sims
  n_sims_f <- pl_f$n_sims
  provenance_mode <- "live"
} else {
  # -- STATIC MODE: the embedded transcription (reference renders) -------
  n_sims_r <- 10000L
  n_sims_f <- 10000L
  provenance_mode <- "static"

  # Both quantities were extracted from the Section 6.6.1 summary tables
  # of the two rendered reference vignettes (10,000 simulated trials
  # each, same DGM, super-population seed 99, AHR-calibrated true
  # HR = 0.70):
  #
  #   random-X : extreme_subgroups.html        (baseline = "resample")
  #   fixed-X  : extreme_subgroups_fixed.html  (baseline = "fixed")
cmp <- data.frame(
  subgroup = c(
    "All Patients",
    "Post-menopausal",
    "Pre-menopausal",
    "Grade 3",
    "Grade 1/2",
    "Age (young)",
    "Age (older)",
    "Age <= 50",
    "Age > 50",
    "Tumour size (small)",
    "Tumour size (large)",
    "Node-negative",
    "Nodes 1-3",
    "High nodes (>Q3)",
    "PGR low",
    "PGR high",
    "ER-low (<20)",
    "ER-high (>=20)",
    "Pre-meno / Grade 3",
    "Post-meno / Grade 3",
    "Pre-meno / Grade 1-2",
    "Post-meno / Grade 1-2",
    "Pre-meno / Age<=50",
    "Pre-meno / Age>50",
    "Post-meno / Age (young)",
    "Post-meno / Age (older)",
    "Pre-meno / ER-low",
    "Pre-meno / ER-high",
    "Post-meno / ER-low",
    "Post-meno / ER-high",
    "Grade 3 / Node-neg",
    "Grade 3 / Node-pos",
    "Grade 1-2 / Node-neg",
    "Grade 1-2 / Node-pos",
    "Grade 3 / High nodes",
    "Grade 3 / ER-low",
    "Grade 3 / ER-high",
    "Grade 1-2 / ER-low",
    "Grade 1-2 / ER-high",
    "Grade 3 / PGR low",
    "Grade 3 / PGR high",
    "Grade 1-2 / PGR low",
    "Grade 1-2 / PGR high",
    "Pre-meno/Yng/ER-low",
    "Pre-meno/Yng/ER-high",
    "Post-meno/G3/ER-low",
    "Post-meno/G3/Node-pos",
    "G3/Node-pos/ER-low",
    "Large/Node-pos",
    "Large/Node-neg",
    "Small/Node-pos",
    "Small/Node-neg",
    "random60",
    "random40",
    "random20",
    "random15"),
  N_r = c(
    686,
    389,
    297,
    161,
    525,
    373,
    313,
    295,
    391,
    357,
    329,
    NA,
    381,
    135,
    340,
    346,
    276,
    410,
    78,
    83,
    219,
    306,
    258,
    38,
    80,
    309,
    139,
    157,
    137,
    252,
    NA,
    161,
    NA,
    525,
    44,
    110,
    51,
    167,
    359,
    127,
    33,
    213,
    312,
    120,
    138,
    57,
    83,
    110,
    329,
    NA,
    357,
    NA,
    60,
    40,
    20,
    15),
  N_f = c(
    686,
    396,
    290,
    161,
    525,
    363,
    323,
    289,
    397,
    353,
    333,
    NA,
    376,
    143,
    343,
    343,
    272,
    414,
    74,
    87,
    216,
    309,
    255,
    35,
    77,
    319,
    134,
    156,
    138,
    258,
    NA,
    161,
    NA,
    525,
    46,
    106,
    55,
    166,
    359,
    127,
    34,
    216,
    309,
    116,
    139,
    58,
    87,
    106,
    333,
    NA,
    353,
    NA,
    60,
    40,
    20,
    15),
  ub2_r = c(
    0.0,
    0.0,
    0.0,
    0.5,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    NA,
    0.0,
    0.7,
    0.0,
    0.0,
    0.0,
    0.0,
    13.1,
    8.7,
    0.5,
    0.0,
    0.1,
    60.6,
    15.1,
    0.0,
    2.2,
    4.6,
    1.5,
    0.1,
    NA,
    0.5,
    NA,
    0.0,
    25.0,
    2.7,
    33.3,
    0.9,
    0.0,
    1.3,
    56.9,
    0.1,
    0.1,
    3.9,
    6.6,
    19.4,
    8.7,
    2.7,
    0.0,
    NA,
    0.0,
    NA,
    31.6,
    50.5,
    80.5,
    90.2),
  ub2_f = c(
    0.0,
    0.0,
    0.0,
    0.3,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    NA,
    0.0,
    0.4,
    0.0,
    0.0,
    0.0,
    0.0,
    13.7,
    6.6,
    0.6,
    0.0,
    0.1,
    62.3,
    16.0,
    0.0,
    2.4,
    4.2,
    1.4,
    0.1,
    NA,
    0.3,
    NA,
    0.0,
    22.2,
    2.8,
    27.9,
    1.0,
    0.0,
    1.1,
    55.2,
    0.1,
    0.1,
    4.1,
    5.8,
    17.3,
    6.6,
    2.8,
    0.0,
    NA,
    0.0,
    NA,
    31.1,
    50.6,
    80.4,
    90.1),
  ub3_r = c(
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    NA,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    1.1,
    0.5,
    0.0,
    0.0,
    0.0,
    34.2,
    1.7,
    0.0,
    0.0,
    0.1,
    0.0,
    0.0,
    NA,
    0.0,
    NA,
    0.0,
    5.2,
    0.0,
    8.7,
    0.0,
    0.0,
    0.0,
    29.1,
    0.0,
    0.0,
    0.1,
    0.2,
    3.3,
    0.5,
    0.0,
    0.0,
    NA,
    0.0,
    NA,
    8.2,
    23.1,
    61.8,
    77.4),
  ub3_f = c(
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    NA,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    0.0,
    1.2,
    0.2,
    0.0,
    0.0,
    0.0,
    36.0,
    1.7,
    0.0,
    0.0,
    0.1,
    0.0,
    0.0,
    NA,
    0.0,
    NA,
    0.0,
    3.4,
    0.0,
    5.9,
    0.0,
    0.0,
    0.0,
    27.1,
    0.0,
    0.0,
    0.1,
    0.3,
    2.2,
    0.2,
    0.0,
    0.0,
    NA,
    0.0,
    NA,
    7.8,
    22.8,
    60.5,
    76.7),
  mUB_r = c(
    0.90,
    0.97,
    1.03,
    1.12,
    0.94,
    0.99,
    1.01,
    1.03,
    0.98,
    1.00,
    1.00,
    NA,
    0.99,
    1.14,
    0.97,
    1.03,
    1.01,
    0.99,
    1.36,
    1.31,
    1.11,
    1.02,
    1.06,
    2.32,
    1.40,
    1.01,
    1.18,
    1.24,
    1.16,
    1.07,
    NA,
    1.12,
    NA,
    0.94,
    1.54,
    1.20,
    1.67,
    1.13,
    1.01,
    1.15,
    2.19,
    1.05,
    1.06,
    1.22,
    1.27,
    1.46,
    1.31,
    1.20,
    1.00,
    NA,
    1.00,
    NA,
    1.64,
    2.01,
    3.81,
    6.05),
  mUB_f = c(
    0.90,
    0.97,
    1.03,
    1.11,
    0.94,
    0.99,
    1.00,
    1.03,
    0.97,
    0.99,
    0.99,
    NA,
    0.99,
    1.12,
    0.96,
    1.03,
    1.01,
    0.98,
    1.37,
    1.28,
    1.11,
    1.02,
    1.06,
    2.41,
    1.42,
    1.00,
    1.19,
    1.23,
    1.16,
    1.06,
    NA,
    1.11,
    NA,
    0.94,
    1.50,
    1.20,
    1.59,
    1.14,
    1.01,
    1.14,
    2.14,
    1.05,
    1.06,
    1.23,
    1.27,
    1.43,
    1.28,
    1.20,
    0.99,
    NA,
    0.99,
    NA,
    1.63,
    2.01,
    3.67,
    5.83),
  hr05_r = c(
    0.2,
    1.3,
    3.0,
    5.6,
    0.6,
    1.7,
    2.2,
    3.1,
    1.2,
    2.5,
    1.5,
    NA,
    2.7,
    6.3,
    1.4,
    3.0,
    2.8,
    1.4,
    14.8,
    11.9,
    6.1,
    2.9,
    4.2,
    28.7,
    15.8,
    2.3,
    9.2,
    10.8,
    8.5,
    3.9,
    NA,
    5.6,
    NA,
    0.6,
    17.7,
    9.6,
    19.0,
    7.7,
    2.2,
    7.6,
    26.9,
    4.6,
    4.0,
    10.9,
    12.3,
    16.9,
    11.9,
    9.6,
    1.5,
    NA,
    2.5,
    NA,
    21.0,
    26.5,
    34.1,
    36.8),
  hr05_f = c(
    0.1,
    1.1,
    3.2,
    5.1,
    0.7,
    1.5,
    1.8,
    3.2,
    1.0,
    2.4,
    1.3,
    NA,
    2.8,
    5.1,
    1.1,
    3.3,
    2.2,
    1.4,
    14.7,
    11.1,
    6.2,
    2.5,
    4.2,
    29.9,
    15.8,
    1.9,
    9.2,
    10.2,
    8.0,
    3.6,
    NA,
    5.1,
    NA,
    0.7,
    17.4,
    9.7,
    18.3,
    7.4,
    2.2,
    6.8,
    27.3,
    4.1,
    4.3,
    10.4,
    11.5,
    17.1,
    11.1,
    9.7,
    1.3,
    NA,
    2.4,
    NA,
    20.7,
    25.7,
    34.1,
    37.6),
  hr1_r = c(
    0.2,
    1.5,
    3.8,
    7.1,
    0.9,
    1.9,
    2.8,
    3.3,
    1.6,
    2.5,
    2.3,
    NA,
    2.4,
    7.9,
    1.2,
    3.8,
    2.5,
    2.1,
    15.7,
    14.8,
    7.0,
    3.1,
    4.6,
    28.4,
    15.9,
    2.8,
    9.6,
    11.6,
    8.8,
    5.0,
    NA,
    7.1,
    NA,
    0.9,
    20.3,
    9.9,
    22.7,
    7.4,
    3.0,
    8.4,
    28.2,
    4.0,
    4.5,
    11.2,
    12.8,
    17.8,
    14.8,
    9.9,
    2.3,
    NA,
    2.5,
    NA,
    21.0,
    26.0,
    34.5,
    36.9),
  hr1_f = c(
    0.3,
    1.6,
    3.7,
    6.4,
    1.0,
    2.2,
    2.8,
    3.6,
    1.6,
    2.3,
    2.1,
    NA,
    2.2,
    7.2,
    1.1,
    3.8,
    2.7,
    2.1,
    15.8,
    12.9,
    6.9,
    3.2,
    4.7,
    28.7,
    16.6,
    2.8,
    9.5,
    11.5,
    8.7,
    4.6,
    NA,
    6.4,
    NA,
    1.0,
    19.4,
    10.1,
    20.5,
    7.5,
    2.8,
    8.1,
    27.4,
    4.2,
    4.8,
    11.3,
    12.5,
    17.2,
    12.9,
    10.1,
    2.1,
    NA,
    2.3,
    NA,
    21.4,
    25.9,
    33.7,
    35.7),
  mHR_r = c(
    0.72,
    0.72,
    0.72,
    0.72,
    0.72,
    0.72,
    0.72,
    0.71,
    0.72,
    0.71,
    0.72,
    NA,
    0.71,
    0.72,
    0.71,
    0.71,
    0.71,
    0.72,
    0.71,
    0.73,
    0.71,
    0.71,
    0.72,
    0.71,
    0.71,
    0.72,
    0.71,
    0.71,
    0.71,
    0.72,
    NA,
    0.72,
    NA,
    0.72,
    0.72,
    0.72,
    0.73,
    0.71,
    0.72,
    0.72,
    0.72,
    0.71,
    0.71,
    0.71,
    0.71,
    0.71,
    0.73,
    0.72,
    0.72,
    NA,
    0.71,
    NA,
    0.72,
    0.71,
    0.72,
    0.71),
  mHR_f = c(
    0.72,
    0.72,
    0.71,
    0.72,
    0.72,
    0.72,
    0.72,
    0.72,
    0.72,
    0.71,
    0.72,
    NA,
    0.70,
    0.72,
    0.71,
    0.71,
    0.71,
    0.72,
    0.71,
    0.72,
    0.71,
    0.71,
    0.72,
    0.71,
    0.72,
    0.71,
    0.71,
    0.71,
    0.71,
    0.72,
    NA,
    0.72,
    NA,
    0.72,
    0.72,
    0.71,
    0.73,
    0.71,
    0.71,
    0.71,
    0.72,
    0.71,
    0.71,
    0.71,
    0.72,
    0.71,
    0.72,
    0.71,
    0.72,
    NA,
    0.71,
    NA,
    0.71,
    0.72,
    0.71,
    0.69),
  stringsAsFactors = FALSE
  )
}

# -- Quantities shared by both modes --------------------------------------
# n_sims label + Monte Carlo noise band referenced in the narrative.
n_sims_label <- if (n_sims_r == n_sims_f) {
  sprintf("**%s simulated trials**", format(n_sims_r, big.mark = ","))
} else {
  sprintf("**%s (random-X) / %s (fixed-X) simulated trials**",
          format(n_sims_r, big.mark = ","), format(n_sims_f, big.mark = ","))
}
n_mc     <- min(n_sims_r, n_sims_f)
.mc_band <- function(p) 2 * sqrt(2 * p * (1 - p) / n_mc)  # ~95% band for a
                                                          # difference of two
                                                          # proportions
mc_50_pp <- 100 * .mc_band(0.50)
mc_05_pp <- 100 * .mc_band(0.05)

# Derived columns used throughout
cmp <- cmp |>
  mutate(
    d_ub2 = ub2_f - ub2_r,          # fixed minus random, percentage points
    d_ub3 = ub3_f - ub3_r,
    d_N   = N_f - N_r,
    empty = is.na(N_r) | is.na(N_f),
    kind  = case_when(
      subgroup == "All Patients"        ~ "ITT",
      grepl("^random", subgroup)        ~ "Random benchmark",
      TRUE                              ~ "Covariate-defined"
    )
  )
valid <- cmp |> filter(!empty)

# Row behind the mechanism example in "Where the designs differ".
.g3eh <- cmp[cmp$subgroup == "Grade 3 / ER-high", , drop = FALSE]
if (nrow(.g3eh) == 0) .g3eh <- cmp[NA_integer_, , drop = FALSE]
```

# Overview and provenance

This memo compares the two companion simulation designs for the
extreme-subgroups analysis, both run at `r n_sims_label` on the
same AHR-calibrated uniform-effect DGM (true HR = 0.70 for every patient,
GBSG source data, super-population seed 99):

- **Random-X** (`extreme_subgroups.html`, `baseline = "resample"`): every
  trial resamples 686 subjects with replacement from the super-population,
  so covariates — and hence subgroup composition and sizes — vary across
  trials.  This answers the *population-replication* question: what if the
  trial had enrolled a different sample from the same population?
- **Fixed-X** (`extreme_subgroups_fixed.html`, `baseline = "fixed"`): every
  trial uses the observed GBSG panel — each source patient exactly once —
  so subgroup composition and sizes are identical in every replicate.  This
  answers the *conditional-on-X* question: given exactly these patients,
  how variable are the subgroup analyses across re-runs of the trial?

```{r provenance}
#| results: asis
if (provenance_mode == "live") {
  cat(sprintf(paste0(
    "::: {.callout-note title=\"Provenance: recomputed from live payloads\"}\n",
    "Every table in this memo is computed from the raw simulation matrices:\n\n",
    "- **random-X**: `%s` -- %s sims, forestsearch %s, written %s\n",
    "- **fixed-X**: `%s` -- %s sims, forestsearch %s, written %s\n\n",
    "Narrative commentary was written against the 10,000-sim reference runs; ",
    "prose that cites specific values draws them from the loaded payloads, but ",
    "qualitative statements should be cross-checked against the recomputed ",
    "tables whenever the payloads change.\n",
    ":::\n\n"),
    payload_random, format(n_sims_r, big.mark = ","),
    pl_r$forestsearch_version, format(pl_r$created, "%Y-%m-%d %H:%M"),
    payload_fixed,  format(n_sims_f, big.mark = ","),
    pl_f$forestsearch_version, format(pl_f$created, "%Y-%m-%d %H:%M")))
} else {
  cat(paste0(
    "::: {.callout-note title=\"Provenance: embedded static transcription\"}\n",
    "`use_payloads = FALSE`: every number below is the embedded transcription ",
    "from the original 10,000-simulation reference renders; no external files ",
    "were read.  Drop `-P use_payloads:false` (or set the knob back to TRUE) ",
    "to recompute from live payloads.\n",
    ":::\n\n"))
}
```

```{r headline-stats}
#| results: asis
mstat <- valid |> summarise(
  mean_d   = mean(d_ub2),
  med_d    = median(d_ub2),
  min_d    = min(d_ub2),
  max_d    = max(d_ub2)
)
cat(sprintf(
  paste0("**Headline:** across all %d non-empty subgroups the mean ",
         "difference in Pr(UB \u2265 2) is **%+.2f pp** (fixed minus ",
         "random), median %+.2f pp, range [%+.1f, %+.1f].  Median HR is ",
         "0.69\u20130.73 in every row of both designs.\n"),
  nrow(valid), mstat$mean_d, mstat$med_d, mstat$min_d, mstat$max_d))
```

At `r format(n_mc, big.mark = ",")` independent simulations per design
(the smaller of the two runs), the Monte Carlo 95% noise band for a
difference in two proportions is approximately
$\pm 2\sqrt{2\,p(1-p)/n}$ — about **±`r sprintf("%.1f", mc_50_pp)` pp at
p = 50%** and ±`r sprintf("%.1f", mc_05_pp)` pp at p = 5%.  The vast majority of subgroup-level differences sit
inside that band.  **The central message of both vignettes — extreme
results are routine in small subgroups, and clinical rationale confers no
statistical protection — is design-robust.**  A skeptic cannot attribute
the findings to covariate-resampling artifacts.

# Random benchmarks: the internal-consistency check

A random draw of $k$ patients is distributionally almost unaffected by
whether the panel it draws from was itself resampled.  The four random
benchmarks are therefore predicted to be *invariant* across the two
designs — and they are:

```{r tbl-benchmarks}
valid |>
  filter(kind == "Random benchmark") |>
  transmute(
    Subgroup = subgroup,
    `N (both)` = N_r,
    `Pr(UB>=2), random-X` = sprintf("%.1f%%", ub2_r),
    `Pr(UB>=2), fixed-X`  = sprintf("%.1f%%", ub2_f),
    `Pr(UB>=3), random-X` = sprintf("%.1f%%", ub3_r),
    `Pr(UB>=3), fixed-X`  = sprintf("%.1f%%", ub3_f),
    `mUB, random-X` = sprintf("%.2f", mUB_r),
    `mUB, fixed-X`  = sprintf("%.2f", mUB_f)
  ) |>
  gt() |>
  tab_header(title = "Random benchmarks are design-invariant") |>
  tab_options(table.font.size = px(13), column_labels.font.weight = "bold")
```

Under fixed-X these benchmarks isolate *pure membership noise at exactly
fixed N* (sizes are 60/40/20/15 in every replicate by construction);
under random-X they additionally inherit panel resampling.  The
equivalence above shows that second component is negligible — which is
precisely what makes the benchmarks valid calibration anchors in both
documents.

# Where the designs differ — and why

Every covariate-defined subgroup that moved meaningfully moved in
lockstep with its realized N: the sign of the change in Pr(UB ≥ 2) is
opposite the sign of the change in N essentially everywhere in the
high-risk panel.

```{r tbl-movers}
valid |>
  filter(kind == "Covariate-defined", abs(d_ub2) >= 1.5) |>
  # Drop exact-duplicate result rows here (the Node-pos duplicates,
  # explained in the Data Quirks section below) so each mover appears once.
  distinct(N_r, N_f, ub2_r, ub2_f, ub3_r, mUB_r, .keep_all = TRUE) |>
  arrange(d_ub2) |>
  transmute(
    Subgroup = subgroup,
    `N random-X` = N_r, `N fixed-X` = N_f,
    `Pr(UB>=2) random-X` = sprintf("%.1f%%", ub2_r),
    `Pr(UB>=2) fixed-X`  = sprintf("%.1f%%", ub2_f),
    `Delta (pp)` = sprintf("%+.1f", d_ub2)
  ) |>
  gt() |>
  tab_header(
    title = "Subgroups with |Delta Pr(UB>=2)| >= 1.5 pp",
    subtitle = "Direction tracks realized N: larger fixed panel => lower tail risk, and vice versa"
  ) |>
  tab_options(table.font.size = px(13), column_labels.font.weight = "bold")
```

**The mechanism has a twist worth understanding.**  The fixed-X N is the
*true* GBSG count for each subgroup, while the random-X mean N inherits
the drift of the one super-population realization (5,000 draws at
seed 99).  Grade 3 / ER-high genuinely has `r .g3eh$N_f` patients in GBSG; the
super-population happened to under-represent it, so random-X simulated it
at mean N ≈ `r .g3eh$N_r` and correspondingly overstated its tail risk
(`r sprintf("%.1f%%", .g3eh$ub2_r)` vs `r sprintf("%.1f%%", .g3eh$ub2_f)`).  The deltas above are therefore **not** purely the textbook
"random-X adds denominator variance" effect — a substantial share is the
super-population's own finite-sampling error, which fixed-X sidesteps
entirely by conditioning on the exact source composition.  (If cleaner
random-X numbers were ever needed, raising `n_super` from 5,000 to
50,000 would shrink that component.)

Beyond composition, the pure variance effect is visible but small:
same-N rows (random40, random60, Grade 3 at N = 161) show whiskers and
tail probabilities a touch tighter under fixed-X, well within one or two
percentage points.

# Section 6.6.4 high-risk panels

The two rendered Section 6.6.4 panels are near-interchangeable: the same
**11 subgroups** pass the Pr(UB ≥ 2) ≥ 10% filter in both designs, with a
single ordering swap — Grade 3 / ER-high drops below random60 under
fixed-X, driven by its larger (true) panel N of 55 vs the random-X mean
of 51.  ECI whiskers for Grade 3 / High nodes and Post-meno/G3/ER-low are
visibly slightly shorter under fixed-X; everything else is
indistinguishable by eye.

# Data quirks surfaced by the comparison

Two artefacts of the subgroup list — present in **both** designs, and
worth fixing in both vignettes:

```{r quirks}
#| results: asis
empties <- cmp |> filter(empty) |> pull(subgroup)
cat(sprintf(
  paste0("1. **Five structurally empty subgroups.** `survival::gbsg` is an ",
         "all-node-positive cohort (minimum nodes = 1), so `nodes == 0` ",
         "never matches.  Empty in both designs: %s.\n"),
  paste0("*", empties, "*", collapse = ", ")))
```

```{r dup-pairs}
#| results: asis
# Detect exact-duplicate result rows (same stats in BOTH designs):
# a consequence of the same quirk -- "Node-pos" is every patient, so each
# Node-pos subgroup duplicates its parent definition.
key <- valid |>
  mutate(k = paste(N_r, N_f, ub2_r, ub2_f, ub3_r, ub3_f, mUB_r, mUB_f,
                   hr05_r, hr05_f, mHR_r, mHR_f))
dups <- key |> group_by(k) |> filter(n() > 1) |>
  summarise(pair = paste(subgroup, collapse = "  ==  "), .groups = "drop")
cat("2. **Duplicate subgroup pairs** (identical results in both designs, ",
    "because Node-pos is everyone):\n\n", sep = "")
for (p in dups$pair) cat("   - ", p, "\n", sep = "")
cat(sprintf("\nEffective count of distinct, populated subgroups: **%d**, not 56.\n",
            nrow(valid) - nrow(dups)))
```

**Suggested fix** (both qmds): replace the `nodes == 0` / `nodes > 0`
splits with `nodes <= 3` vs `nodes > 3` — both populated in GBSG — which
removes the empties and the duplicates in one edit.

# Practical guidance

The **fixed-X** document answers the conditional question a regulatory
reviewer implicitly asks — *given exactly this trial's enrolled patients
and subgroup sizes, how extreme can chance make the analyses look?* — with
subgroup Ns that match the actual submission table exactly.  The
**random-X** document answers the population-replication question and is
the right benchmark when generalizing beyond the specific trial.

Since the results are near-identical, the suggestion is to **lead with
fixed-X for the regulatory framing** (its exact-N property makes the
forest-plot annotations directly comparable to the real trial) and cite
the random-X version as the robustness companion, with the
random-benchmark invariance table above as the bridge showing the
calibration anchors are design-invariant.

# Appendix A — full comparison table {.unnumbered}

```{r tbl-full}
cmp |>
  arrange(desc(ifelse(is.na(ub2_r), -1, ub2_r))) |>
  transmute(
    Subgroup = subgroup,
    `N r|f` = ifelse(empty, "-", paste0(N_r, " | ", N_f)),
    `Pr(UB>=2) r|f` = ifelse(empty, "-", sprintf("%.1f | %.1f", ub2_r, ub2_f)),
    `Pr(UB>=3) r|f` = ifelse(empty, "-", sprintf("%.1f | %.1f", ub3_r, ub3_f)),
    `mUB r|f` = ifelse(empty, "-", sprintf("%.2f | %.2f", mUB_r, mUB_f)),
    `Pr(HR<0.5) r|f` = ifelse(empty, "-", sprintf("%.1f | %.1f", hr05_r, hr05_f)),
    `Pr(HR>1.0) r|f` = ifelse(empty, "-", sprintf("%.1f | %.1f", hr1_r, hr1_f)),
    `mHR r|f` = ifelse(empty, "-", sprintf("%.2f | %.2f", mHR_r, mHR_f))
  ) |>
  gt() |>
  tab_header(
    title = "All 56 subgroups, random-X | fixed-X",
    subtitle = "Sorted by random-X Pr(UB>=2); percentages in points; '-' = structurally empty"
  ) |>
  tab_options(table.font.size = px(11), data_row.padding = px(2),
              column_labels.font.weight = "bold")
```

# Appendix B — payload knob and reproduction notes {.unnumbered}

This memo renders in one of two modes, controlled by a single knob in the
setup chunk (`use_payloads`, default `TRUE`):

- **Live mode** (`use_payloads = TRUE`, the default).  The memo loads two
  RDS *payloads* — one per design — and recomputes every table from the
  raw `n_sims × n_subgroups` matrices.  Payloads are written by the
  companion vignettes when their own knob `save_payload = TRUE` (also the
  default): `extreme_subgroups.qmd` writes the `resample` payload and
  `extreme_subgroups_fixed.qmd` the `fixed` payload, by default to
  `results/extreme_sims_<design>_<n_sims>_payload.rds` next to the
  vignettes.  Missing files stop the render with fix-it instructions —
  there is no silent fallback to stale numbers.
- **Static mode** (`use_payloads = FALSE`).  The memo renders entirely
  from the static `cmp` data frame embedded in the setup chunk
  (transcribed from the original 10,000-simulation reference renders) and
  reads no external files.

All knobs can be overridden at render time without editing any file:

```bash
# Static memo (no payloads read):
quarto render extreme_subgroups_design_comparison.qmd -P use_payloads:false

# Custom payload locations / names (e.g. the external sims drive):
quarto render extreme_subgroups_design_comparison.qmd \
  -P payload_random:/media/larryleon/bigstorage/forestsearch_sims/fixed_random/extreme_sims_resample_10000_payload.rds \
  -P payload_fixed:/media/larryleon/bigstorage/forestsearch_sims/fixed_random/extreme_sims_fixed_10000_payload.rds

# The vignettes carry the mirror-image knobs: save_payload, payload_dir,
# payload_file (plus n_sims).  A 50-sim smoke run of the whole pipeline:
quarto render extreme_subgroups.qmd       -P n_sims:50
quarto render extreme_subgroups_fixed.qmd -P n_sims:50
quarto render extreme_subgroups_design_comparison.qmd \
  -P payload_random:results/extreme_sims_resample_50_payload.rds \
  -P payload_fixed:results/extreme_sims_fixed_50_payload.rds
```

Each payload is a plain named list:

```r
str(readRDS("results/extreme_sims_fixed_10000_payload.rds"), max.level = 1)
# $ design               : chr "fixed"            ("resample" for random-X)
# $ n_sims               : int 10000
# $ sim_hrs, sim_ubs, sim_ns : num [n_sims x n_subgroups], colnames = subgroup names
# $ subgroups            : list of list(id, name, grp)
# $ sim_config           : the vignette's simulation configuration
# $ cens_adjust, k_treat, hr_true : calibration constants
# $ created, r_version, forestsearch_version : provenance stamps
```

Beyond the tables shown here, the raw matrices support paired
design-delta summaries, full ECI endpoints, and per-subgroup quantile
curves — natural extensions now that payloads are the standard
interchange between the vignettes and this memo.
