---
title: "ACTG175 Binary Harm Search: Simulation Sanity Check"
format: html
---

## Purpose

Build a DGM from the ACTG175 data targeting OR(Q) = 2.0 in the harm
subgroup identified by the real-data analysis (cd40 > 421 & wtkg > 82.3).
Simulate one trial and run FS + GRF as a sanity check before scaling to
full simulations.

No treatment switching.  No `adverse_outcome` flags.  The default
forestsearch harm-search machinery operates natively.

```{r setup}
library(forestsearch)
library(speff2trial)
library(data.table)
library(gt)
library(ggplot2)

set.seed(8316951L)
```

# Data Preparation {#sec-data}

```{r data-prep}
# ACTG175: arm 1 = ZDV+ddI (experimental), arm 3 = ddI (control)
actg_df <- subset(speff2trial::ACTG175, arms %in% c(1L, 3L))
actg_df$id <- seq_len(nrow(actg_df))

# Treatment: ZDV+ddI = 1, ddI = 0 (NO switching)
actg_df$treat <- ifelse(actg_df$arms == 1L, 1L, 0L)

# Binary outcome: adverse (failure to improve)
actg_df$y_binary <- as.integer(actg_df$cd420 <= actg_df$cd40)

actg_df <- actg_df[!is.na(actg_df$cd420), ]

cat(sprintf("N = %d\n", nrow(actg_df)))
cat(sprintf("Adverse rate overall:  %.1f%%\n", 100 * mean(actg_df$y_binary)))
cat(sprintf("  ZDV+ddI (T=1): %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$treat == 1])))
cat(sprintf("  ddI     (T=0): %.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$treat == 0])))
```

## DGM Factor Variables

The DGM requires factor covariates.  The TRUE subgroup Q is defined by
the harm subgroup identified in the observed-data analysis:
Q = {cd40 > 421 & wtkg > 82.3}.

```{r dgm-factors}
# z1, z2 define the TRUE subgroup Q (from observed-data FS result)
actg_df$z1  <- as.factor(ifelse(actg_df$cd40 > 421, 1L, 0L))
actg_df$z2  <- as.factor(ifelse(actg_df$wtkg > 82.3, 1L, 0L))

# Remaining factors: cover the covariate space for the DGM model fit
actg_df$z3  <- as.factor(ifelse(actg_df$age > 34, 1L, 0L))
actg_df$z4  <- as.factor(ifelse(actg_df$preanti <= 745, 1L, 0L))
actg_df$z5  <- as.factor(ifelse(
  actg_df$karnof <= median(actg_df$karnof), 1L, 0L))
actg_df$z6  <- as.factor(ifelse(
  actg_df$cd80 <= median(actg_df$cd80), 1L, 0L))
actg_df$z7  <- as.factor(actg_df$hemo)
actg_df$z8  <- as.factor(actg_df$homo)
actg_df$z9  <- as.factor(actg_df$drugs)
actg_df$z10 <- as.factor(actg_df$race)
actg_df$z11 <- as.factor(actg_df$gender)
actg_df$z12 <- as.factor(actg_df$symptom)

# True subgroup: Q = {z1=1 & z2=1} = {cd40 > 421 & wtkg > 82.3}
actg_df$flag_harm_ref <- as.integer(actg_df$z1 == 1 & actg_df$z2 == 1)

Q_n   <- sum(actg_df$flag_harm_ref)
Q_N   <- nrow(actg_df)
Q_pct <- 100 * Q_n / Q_N

cat(sprintf("True Q: cd40 > 421 & wtkg > 82.3\n"))
cat(sprintf("  |Q| = %d / %d (%.1f%%)\n", Q_n, Q_N, Q_pct))
cat(sprintf("  Adverse rate in Q:  ZDV+ddI=%.1f%%, ddI=%.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 1 & actg_df$treat == 1]),
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 1 & actg_df$treat == 0])))
cat(sprintf("  Adverse rate in Qc: ZDV+ddI=%.1f%%, ddI=%.1f%%\n",
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 0 & actg_df$treat == 1]),
    100 * mean(actg_df$y_binary[actg_df$flag_harm_ref == 0 & actg_df$treat == 0])))
```

## Analysis Covariates

The analyst sees raw continuous covariates — not the z-variables.

```{r covariates}
cont_vars <- c("age", "preanti", "wtkg", "karnof", "cd40", "cd80")
bin_vars  <- c("hemo", "homo", "drugs", "race", "gender", "symptom")
for (v in bin_vars) actg_df[[v]] <- as.factor(actg_df[[v]])

confounders <- c(cont_vars, bin_vars)
```

# DGM Calibration {#sec-calibration}

Target: OR(Q) = 2.0 on the adverse scale.  Treatment (ZDV+ddI) doubles
the odds of non-improvement in Q = harm.  The DGM includes both factor
variables (for subgroup definition) and raw continuous covariates (for
a richer prognostic baseline model), consistent with the survival
`generate_aft_dgm_flex()` design.

```{r calibrate}
dgm_factor_vars   <- paste0("z", 1:12)
dgm_cont_vars    <- cont_vars   # age, preanti, wtkg, karnof, cd40, cd80
dgm_subgroup_vars <- c("z1", "z2")
dgm_subgroup_cuts <- list(z1 = 1L, z2 = 1L)

cal_target <- 2.0

dgm_calibrated <- calibrate_glm_interaction(
  data            = actg_df,
  factor_vars     = dgm_factor_vars,
  continuous_vars = dgm_cont_vars,
  outcome_var     = "y_binary",
  treatment_var   = "treat",
  target_effect   = cal_target,
  outcome_type    = "binary",
  effect_measure  = "OR",
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  k_inter_range   = c(0, 3),
  grid_step       = 0.05,
  n_super         = 5000L,
  seed            = 8316951L,
  verbose         = TRUE
)

cat(sprintf("\n=== Calibrated DGM ===\n"))
cat(sprintf("  k_inter: %.2f\n", dgm_calibrated$model_params$k_inter))
cat(sprintf("  OR(Q):   %.3f (target: %.1f)\n",
    dgm_calibrated$hazard_ratios$harm_subgroup, cal_target))
cat(sprintf("  OR(Qc):  %.3f\n",
    dgm_calibrated$hazard_ratios$no_harm_subgroup))
cat(sprintf("  OR(ITT): %.3f\n",
    dgm_calibrated$hazard_ratios$overall))
```

## Null DGM

```{r null-dgm}
dgm_null <- generate_glm_dgm(
  data            = actg_df,
  factor_vars     = dgm_factor_vars,
  continuous_vars = dgm_cont_vars,
  outcome_var     = "y_binary",
  treatment_var   = "treat",
  outcome_type    = "binary",
  subgroup_vars   = dgm_subgroup_vars,
  subgroup_cuts   = dgm_subgroup_cuts,
  model           = "null",
  n_super         = 5000L,
  seed            = 8316951L,
  verbose         = TRUE
)

cat(sprintf("\n=== Null DGM ===\n"))
cat(sprintf("  OR(Q):   %.3f\n", dgm_null$hazard_ratios$harm_subgroup))
cat(sprintf("  OR(Qc):  %.3f\n", dgm_null$hazard_ratios$no_harm_subgroup))
cat(sprintf("  OR(ITT): %.3f\n", dgm_null$hazard_ratios$overall))
```

# Single Trial Simulation {#sec-single-trial}

Simulate one trial from the calibrated DGM, then run FS and GRF
to verify the pipeline reproduces the harm signal.

```{r simulate-one}
df_sim <- simulate_from_glm_dgm(dgm_calibrated, n = 1000, seed = 42)

cat(sprintf("Simulated trial: N = %d\n", nrow(df_sim)))
cat(sprintf("  Adverse rate overall:  %.1f%%\n", 100 * mean(df_sim$y_sim)))
cat(sprintf("  Adverse rate (T=1):    %.1f%%\n",
    100 * mean(df_sim$y_sim[df_sim$treat_sim == 1])))
cat(sprintf("  Adverse rate (T=0):    %.1f%%\n",
    100 * mean(df_sim$y_sim[df_sim$treat_sim == 0])))

# True subgroup in simulated data
cat(sprintf("\n  True Q prevalence: %.1f%%\n",
    100 * mean(df_sim$flag_harm)))
cat(sprintf("  Adverse rate Q (T=1):  %.1f%%\n",
    100 * mean(df_sim$y_sim[df_sim$flag_harm == 1 & df_sim$treat_sim == 1])))
cat(sprintf("  Adverse rate Q (T=0):  %.1f%%\n",
    100 * mean(df_sim$y_sim[df_sim$flag_harm == 1 & df_sim$treat_sim == 0])))
cat(sprintf("  Adverse rate Qc (T=1): %.1f%%\n",
    100 * mean(df_sim$y_sim[df_sim$flag_harm == 0 & df_sim$treat_sim == 1])))
cat(sprintf("  Adverse rate Qc (T=0): %.1f%%\n",
    100 * mean(df_sim$y_sim[df_sim$flag_harm == 0 & df_sim$treat_sim == 0])))

# Empirical OR in Q
fit_Q <- glm(y_sim ~ treat_sim,
             data = df_sim[df_sim$flag_harm == 1, ],
             family = binomial)
cat(sprintf("\n  Empirical OR(Q): %.3f (target: %.1f)\n",
    exp(coef(fit_Q)["treat_sim"]), cal_target))
```

## FS on Simulated Trial

```{r fs-sim}
# FS thresholds
c1 <- 1.25
c2 <- 1.0

# FS needs: outcome = adverse, treat as-is
fs_sim <- forestsearch(
  df.analysis      = df_sim,
  confounders.name = confounders,
  outcome.name     = "y_sim",
  event.name       = "y_sim",
  treat.name       = "treat_sim",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "OR",
  seedit           = 8316951L,
  n.min            = 60L,
  d0.min           = 10L,
  d1.min           = 10L,
  hr.threshold     = c1,
  hr.consistency   = c2,
  pconsistency.threshold = 0.80,
  fs.splits        = 500L,
  max_subgroups_search = 15,
  maxk             = 2L,
  sg_focus         = "minSG",
  use_lasso        = FALSE,
  use_grf          = TRUE,
  is.RCT           = TRUE,
  vi.grf.min       = -0.2,
  dmin.grf         = 0,
  cont.cutoff      = 4L,
  details          = TRUE,
  quiet            = FALSE,
  parallel_args    = list(plan = "sequential", workers = 1,
                          show_message = FALSE)
)
```

### FS Results

```{r fs-sim-tables}
fs_sim_tabs <- sg_tables(fs_sim, ndecimals = 3, which_df = "est")
fs_sim_tabs$sg10_out
fs_sim_tabs$tab_estimates
```

## GRF on Simulated Trial

GRF uses the beneficial outcome (1 − y_sim) with default
`adverse_outcome = FALSE`.

```{r grf-sim}
df_sim$y_sim_pos <- 1L - df_sim$y_sim

grf_sim <- grf.subg.harm.glm(
  data             = df_sim,
  confounders.name = confounders,
  outcome.name     = "y_sim_pos",
  treat.name       = "treat_sim",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "OR",
  dmin.grf         = 0.1,
  maxdepth         = 2L,
  RCT              = TRUE,
  sg.criterion     = "mDiff",
  details          = TRUE
)

cat("\n=== GRF Result (simulated trial) ===\n")
if (!is.null(grf_sim$sg.harm.id) && length(grf_sim$sg.harm.id) > 0) {
  cat("Subgroup:", paste(grf_sim$sg.harm.id, collapse = " & "), "\n")
  grf_H <- grf_sim$data$treat.recommend == 0
  cat(sprintf("Size(H): %d,  Size(Hc): %d\n", sum(grf_H), sum(!grf_H)))
} else {
  cat("GRF: no subgroup identified.\n")
}
```

### GRF Results

```{r grf-sim-tables}
grf_sim_tabs <- sg_tables(grf_sim, ndecimals = 3, which_df = "est")
grf_sim_tabs$tab_estimates
```

# Comparison: Real Data vs Simulated

```{r compare}
cat("=== FS on REAL ACTG175 data ===\n")
fs_real <- forestsearch(
  df.analysis      = actg_df,
  confounders.name = confounders,
  outcome.name     = "y_binary",
  event.name       = "y_binary",
  treat.name       = "treat",
  id.name          = "id",
  outcome_type     = "binary",
  effect_measure   = "OR",
  seedit           = 8316951L,
  n.min            = 60L,
  d0.min           = 10L,
  d1.min           = 10L,
  hr.threshold     = c1,
  hr.consistency   = c2,
  pconsistency.threshold = 0.80,
  fs.splits        = 500L,
  max_subgroups_search = 15,
  maxk             = 2L,
  sg_focus         = "minSG",
  use_lasso        = FALSE,
  use_grf          = TRUE,
  is.RCT           = TRUE,
  vi.grf.min       = -0.2,
  dmin.grf         = 0,
  cont.cutoff      = 4L,
  details          = TRUE,
  quiet            = FALSE,
  parallel_args    = list(plan = "sequential", workers = 1,
                          show_message = FALSE)
)

fs_real_tabs <- sg_tables(fs_real, ndecimals = 3, which_df = "est")
```

### Side by Side

```{r side-by-side}
cat("=== Real data ===\n")
fs_real_tabs$tab_estimates

cat("\n=== Simulated trial ===\n")
fs_sim_tabs$tab_estimates
```
